### 2016-04-21 paper/refinements.tex
Added in outline of paper

\section{Synthesis with Refinements}
\label{sec:refinements}

\subsection{Examples are Refinements}

Until now, we have thought of examples as distinct from types.  Together, we stated,
these two components form the specification for a type-directed synthesis problem.
Here, we observe that examples can be integrated into the type system.

Consider the statement \Refines[x][2], where \HasType[x][\mathsf{nat}].  Put another way, we mean
that $x$ is any inhabitant of \textsf{nat}~so long as $x$ = 2.  Gathering the example and type
information into a single statement, we might put a predicate on the type of $x$ and say that
$\{x \in \mathsf{nat}~|~x = 2\}$.  This kind of type is hardly esoteric;
to the contrary, it is an instance of a \emph{singleton refinement type}
~\cite{pfenning-refinements}.

Refinements represent predicates on the structure of inhabitants of a type. 
In their simplest, singleton form, refinements constrain the type of an expression
to a single value as in the example above.  Refinements can express more general predicates,
such as the type for a non-empty list, $\mathsf{\HasType[\Refines[ls][Cons(natlist, nat)]][natlist]}$
(alternatively written as the predicate
$\{ls \in \mathsf{natlist}~|~ls = \mathsf{Cons(natlist, nat)}\}$).

The example grammar from Figure \ref{fig:grammar1} is already a refinement system.  Every
first-order example is a singleton refinement type.  Function examples are refinements
that partially constrain a type without necessarily restricting it to a single inhabitant.
Intersection refinements allow us to express the conjunction of predicates, which is useful for
combining function refinements.

\subsection{Refinements are more than Examples}

We can exploit the insight that examples are refinements to build a far richer synthesis
specification language than that in the original type-directed synthesis system. 
Here, we introduce two new refinements: union and base refinements
(Figure \ref{fig:grammar2}).  In Section \ref{sec:negation}, we discuss negation.

Union refinements are the dual of intersections: they allow a user to express that one of several
predicates holds on a value.  For example, where the refinement \TAnds[\TFunc{0}{1}, \TFunc{1}{2}]
describes a function that maps 0 to 1 and 1 to 2, \TOrs[\TFunc{0}{1}, \TFunc{1}{2}] refers to a
function that obeys at least one (but not necessarily both) of the two mappings.
Base refinements express the seemingly weak predicate that an expression
is some inhabitant of that base type.  They are most potent when combined with other refinements like
constructors, which enable us to express predicates on, for example, all non-empty lists.

As an example, consider a possible specification for the unary
decrement function, which subtracts one from all numbers except zero.
The function maps anything described
by the refinement \TOrs[0, 1] to 0, so a partial specification includes the refinement
\TFunc{\TOrs[0, 1]}{0}.  On any other input $\mathsf{S~(S~nat)}$, the result should be
$\mathsf{S~nat}$.  Combining these disparate parts of the specification into a single
refinement, we could describe decrement with the statement
\TAnds[\TFunc{\TOrs[0, 1]}{0}, \TFunc{\mathsf{S~(S~nat)}}{\mathsf{S~nat}}].

\subsection{Synthesis with Refinements}

Refinement types introduce an additional complication into the synthesis process: subsumption.
While the original type-directed synthesis system lacked any form of subtyping, the richer
refinement language forces us to consider the interaction between subsumption and synthesis.
For example, an expression \HasType[\Refines[\DExp][\mathsf{S~(S~nat)}]][\mathsf{nat}] can be
used in any context where we expect an expression
\HasType[\Refines[\DExp'][\mathsf{S~nat}]][\mathsf{nat}].

As a prerequisite for formalizing the subtyping relations between refinements, we present
the denotational semantics of refinement types as sets of their inhabitants in Figure
\ref{fig:denotation1}.  This set-theoretic denotation clarifies the relationships between
refinements that become concrete in the subtyping judgments in Figure \ref{fig:subsumption1}.

The denotations of each type are in the top section of Figure \ref{fig:denotation1}.  Unit type
denotes the set containing only the unit value.
Any base type denotes the set of all of its inhabitants.
Product type denotes the cartesian product of the denotations of its constituents.
Function type denotes the set of total functions over the input and output types,
where each function is a set of input-output pairs.

The middle section of Figure \ref{fig:denotation1} contains the denotations of refinements,
which only have meaning in relation to the types that they refine.  The unit and base refinement
denotations are identical to those of their synonymous types.  Intersection and union refinements
denote the set intersection and union operators.  Cartesian product and function refinements
have the same denotations as their types over the more restrictive domains of their constituent
refinements.  Finally, in the bottom section of Figure \ref{fig:denotation1}, the subtype
relation denotes the subset operator.

Using these denotational semantics, we can give meaning to the subtyping rules for the refinement
system (Figure \ref{fig:subsumption1}).  Most rules are standard, with special unfolding
rules are necessary for recursive base types.  Note that these subtyping rules apply to refinements
rather than the types themselves.
That is, every rule requires that the elements on both sides of the
subsumption relation are refinements of the same type.  In this system, no subtyping relation
exists between different types.

Finally, we are prepared to update the synthesis system (Figure \ref{fig:rules2}).
New union rules mirror those for intersection
types.  When the goal type is a union, we need only synthesize a program that satisfies one
of the union's constituents.  When an expression in the context is of union type, we must synthesize
a program that handles every constituent of the union.  Base type unfolding rules serve
a purpose similar to those in subsumption.  The \textsc{Synth-Ctx} no longer looks for
an exact match between refinements in the context and in goal position.  Instead, subtypes
are also acceptable.

### 2016-04-21 paper/related-work.tex
Added in outline of paper

\section{Related Work}
\subsection{FlashFill}
\subsection{FlashExtract}
\subsection{SemFill}

### 2016-05-04 paper/dnf-regex-lenses.tex
Current state of things

\section{DNF Regular Expressions and Lenses}
We can push the boundaries of what is expressible through syntactic lenses
through finding a language, equivalent to regular expressions, with fewer
equivalences.

\input{fig_dnf-regex-definition}
A language which removes some of those equivalences is the language of regular
expressions in disjunctive normal form.  This language removes the
equivalences corresponding to distributivity, associativity, and concatenation identity.
Figure~\ref{fig:dnf-regex-definition} defines this.
The full DNF Regex is a list of Clauses.  This corresponds to a chain of $+$s
in the normal language of regular expressions.
A Clause is a list of Atoms, with strings in between.
This corresponds to a chain of composition.  These strings
correspond to regular expression base types, and their requirement to
exist between every atom removes the $\epsilon$ equivalences.
Atoms correspond to pieces that either cannot be broken up uniquely, or we feel should not be broken up.
These are the star regular expressions, and the user defined regular expressions.
The star regular expressions can be broken up in an infinite number of ways.
We feel that a user grouping together aspects of their regular expression together
signifies that it will likely stay together in the transformation.

\input{fig_dnf-lens-syntax}
\input{fig_dnf-lens-semantics}
\input{fig_dnf-lens-typing}
As we have dnf regular expressions, it also makes sense to have dnf lenses,
whose types are dnf regular expressions.  We define dnf lenses in Figure~\ref{fig:dnf-lens-syntax}.
These dnf lenses have the comparable underlying functions to the underlying functions
of normal lenses, and are given in Figure~\ref{fig:dnf-lens-semantics}
TODO: make a nice diagram of how clause lenses work.
They can be typed according to the rules given in Figure~\ref{fig:dnf-lens-typing}

One thing to note about these lenses is that they are slightly more expressive
than the syntactic lenses given previously.
Not all permutations are possible through only using swap.
Furthermore, before we did not allow for permutations on union clauses,
which previously were dealt with through the type change rule of pure lenses.
Now, these permutations are possible.  So, for example, there are no syntactic lenses
between $a+b*$ and $z*+y$.  However, with allowing for permutations, the new
language can synthesize these types of lenses.

Just as we used the parse trees of strings within a regular expression,
to help guide the synthesis for syntactic lenses, we can do the same
for dnf lenses.

### 2016-05-04 paper/intro.tex
Current state of things

\section{Introduction}
\input{fig_regex-definition}
\input{fig_lens-syntax}
\input{fig_lens-typing}
\input{fig_lens-semantics}
We have regular expressions as normal regular expressions.
We expand it with having user defined data types as well.
We have lenses, defined in Figure~\ref{fig:lens-syntax}, typed as in Figure~\ref{fig:lens-typing}.
These lenses have underlying functions, given via Figure~\ref{fig:lens-semantics}.
We would like to be able to synthesize these lenses automatically, given a
specification as two regular expressions, an a set of values that are
mapped to each other.

### 2016-05-04 paper/syntax-lenses.tex
Current state of things

\section{Syntax Driven Lenses}
Unfortunately, with retyping rule, it is a difficult task.
There are an infinite number of possible retypings.
They are still relevant, depending on what the examples are.
Retyping is needed to express all lenses.

For example, consider $r_1=a*$, $r_2=b*$, with examples $a \leftrightarrow bb, aa \leftrightarrow b, aaa \leftrightarrow bbb$.
Without retyping, the only possible lens is $iterate(const(a,b))$, which doesn't express it correctly.
However, using retyping, we can first retype to $\epsilon + a + aa + aaaa*$ and $\epsilon + b + bb + bbbb*$.
This is able to be solved.

Because of the difficulty for lenses including retyping, lets first only worry about syntactic lenses, lenses that don't include retyping.
\input{fig_syntactic-lens-typing.tex}

For these syntactic lenses, one can prove that composition and identity can
be applied only on Userdefs without losing expressibility.

We now have the means to enumerate all possible syntactic lenses between two regular
expressions in a terminating way.  We merely enumerate all the possible
terms of a certain type, which there are a finite number of.  Then we can
run all the lenses on the inputs until we find a lens that satisfies the inputs.

However, this is inefficient, as the number of lenses are possibly exponential in the size of the regular expression.  We can prune some of the possible
lenses very quickly with a key insight, when we reduce the synthesis of
the lens into subproblems, we can also make sub-problems for the input-output examples.

For example, consider the problem of synthesizing a lens $l$ between $\Star{a}\Star{b}$ and $\Star{z}\Star{y}$ that satisfies the example $a \leftrightarrow y$.
We can apply either a swap or a concat rule to this.
If we apply a concat rule, we get the subproblems of synthesizing lenses between
$\Star{a}$ and $\Star{z}$, and synthesizing lenses betwen $\Star{b}$ and $\Star{y}$.
By the underlying function of concat, we know that for this concat lens to work on the provided example,
we must have the lens between $\Star{a}$ and $\Star{z}$ send $a$ to $\epsilon$,
and must have the lens between $\Star{b}$ and $\Star{y}$ send $\epsilon$ to $y$.
However, when trying to apply an iterate rule for $\Star{a}$ and $\Star{z}$, we immediately see there is no satisfying lens,
as parsing $a$ into $\Star{a}$, we see $a$ mathes the inner regular expression $a$ once.
However, on parsing $\epsilon$ into $\Star{z}$, we see $\epsilon$ matches the inner regular expression $z$ zero times.
Through the underlying function of iterate, we know this cannot be done,
so there is no lens that does this.
So we immediately know that there is no concat lens between
$\Star{a}\Star{b}$ and $\Star{z}\Star{y}$ that sends $a$ to $y$.
We can then continue with trying the swap lens, which will eventually provide a lens.
\input{fig_parsetree-definition.tex}
\input{fig_parsetree-typing.tex}

We can formalize this first by definining a data structure for how a regular
expression parses a string, as in Figure~\ref{fig:parsetree-definition},
and typing for that data structure as in Figure~\ref{fig:parsetree-typing}.

\input{fig_lens-parsetree-typing.tex}

With the data structure for parsing, we can now define a typing for a lens
which satisfies the parse tree of an example in Figure~\ref{fig:lens-parsetree-typing}.  Now, to generate a lens for an input output example,
between two regular expressions,
needs to generate the parse tree for the example with those expressions,
and then find a lens between those two regular expressions and between the parse trees.

We say a lens $l : \Regex \Leftrightarrow \RegexAlt$ satisfies a set of parse tree input output examples $\ParseTree_1 \leftrightarrow \ParseTreeAlt_1, \ldots, \ParseTree_n \leftrightarrow \ParseTreeAlt_n$
if we can show $l : \Regex \Leftrightarrow \RegexAlt, \ParseTree_i \leftrightarrow \ParseTreeAlt_i$ for all $i$ between $1$ and $n$.

\iffalse
We now can prove some inversion properties.
\begin{itemize}
\item $concat(\Lens_1,\Lens_2) : \Regex \Leftrightarrow \RegexAlt$ or
$swap(\Lens_1,\Lens_2) : \Regex \Leftrightarrow \RegexAlt$
if, and only if $\Regex = \Regex_1\Regex_2$, and $\RegexAlt = \RegexAlt_1\RegexAlt_2$ for some $\Regex_1$,$\Regex_2$
\end{itemize}
This provides us with a nice, syntax driven strategy for synthesis.
We now have an inversion property, where i

When does there exist a lens between two regular expressions.
More, formally, given two regular expressions $r$ and $s$, when does there exist a lens $l : r \Leftrightarrow s$.
We can say that $\SynSim{r}{s}$ if there is such an $l$.
We can prove $\SynSim{r}{s}$ if, and only if:
\begin{itemize}
\item If $r = r_1 r_2$ then either $s = s_1 s_2$ or $s = s_2 s_1$, and $\SynSim{r_1}{s_1}$ and $\SynSim{r_2}{s_2}$
\item If $r = r_1 | r_2$ then $s = s_1 | s_2$ and $\SynSim{r_1}{s_1}$ and $\SynSim{r_2}{s_2}$
\item If $r = \Star{r'}$ then $s = \Star{s'}$ and $\SynSim{r'}{s'}$.
\item If $r = x\in\Star{\Sigma}$, then $s = y\in\Star{\Sigma}$.
\item If $r = u\in\Delta$, then $s = u\in\Delta$
\end{itemize}

With this, it is easy to recursively check whether there exists a syntactic lens between two regular expressions.

How can we see if there exists a lens between two regular expressions, given a set of examples.
Intuitively, a lens matches a set of examples if it brings the parse trees of the left example to the parse trees of the right example.
So, for example, for an example $e\leftrightarrow f$ to be possible under $r_1|r_2 \Leftrightarrow s_1|s_2$, we need if $e$ matches $r_1$, $f$ must match $s_1$.
To formalize that, we can define parse trees as figure \ref{fig:parsetree-definition}.
Given two regular expressions $\Regex$, $\RegexAlt$, and a list of examples $\{e_1\leftrightarrow f_1,\ldots,e_n\leftrightarrow f_n\}$.
We define a relation between parse trees and regular expressions, $\SynSim{(\Regex_1,\ParseTrees)}{(\Regex_2,\ParseTreeAlts)}$ as follows:
\begin{itemize}
\item If $\Regex = \Regex_1 \Regex_2$,
$\RegexAlt = \RegexAlt_1 \RegexAlt_2$,
$\ParseTrees = [\ParseTree_{1,1}\ParseTree_{1,2}, \ldots, \ParseTree_{n,1}\ParseTree_{n,2}]$,
and $\ParseTreeAlts = [\ParseTreeAlt_{1,1}\ParseTreeAlt_{1,2}, \ldots, \ParseTreeAlt_{n,1}\ParseTreeAlt_{n,2}]$
and either\\
($\SynSim{(\Regex_1,[\ParseTree_{1,1}, \ldots,\ParseTree_{n,1}])}{(\RegexAlt_1,[\ParseTreeAlt_{1,1}, \ldots, \ParseTreeAlt_{n,1}])}$\\
and
$\SynSim{(\Regex_2,[\ParseTree_{1,2}, \ldots,\ParseTree_{n,2}])}{(\RegexAlt_2,[\ParseTreeAlt_{1,2}, \ldots, \ParseTreeAlt_{n,2}])}$)\\
or ($\SynSim{(\Regex_1,[\ParseTree_{1,1}, \ldots,\ParseTree_{n,1}])}{(\RegexAlt_2,[\ParseTreeAlt_{1,2}, \ldots, \ParseTreeAlt_{n,2}])}$\\
and
$\SynSim{(\Regex_2,[\ParseTree_{1,2}, \ldots,\ParseTree_{n,2}])}{(\RegexAlt_1,[\ParseTreeAlt_{1,1}, \ldots, \ParseTreeAlt_{n,1}]})$)
\item If $\Regex = \Regex_1 | \Regex_2$,
$\RegexAlt = \RegexAlt_1 \RegexAlt_2$,
$\ParseTrees = [\ParseTree_1;\ldots;\ParseTree_n]$, $\ParseTreeAlts = [\ParseTreeAlt_1;\ldots;\ParseTreeAlt_n]$,
($\ParseTree_i = \LeftChoiceParse{\ParseTree_i'}$ or $\ParseTree_i = \RightChoiceParse{\ParseTree_i'}$),
and if $\ParseTree_i = \LeftChoiceParse{\ParseTree_i'}$ then $\ParseTreeAlt_i = \LeftChoiceParse{\ParseTreeAlt_i'}$,
and if $\ParseTree_i = \RightChoiceParse{\ParseTree_i'}$ then $\ParseTreeAlt_i = \RightChoiceParse{\ParseTreeAlt_i'}$,
and (define $\ParseTrees_l$ = [$\ParseTree_i'$ for each $\ParseTree_i$ = $\LeftChoiceParse{\ParseTree_i'}$ in \ParseTrees],
define $\ParseTrees_r$ = [$\ParseTree_i'$ for each $\ParseTree_i$ = $\RightChoiceParse{\ParseTree_i'}$ in \ParseTrees],
define $\ParseTreeAlts_l$ = [$\ParseTreeAlt_i'$ for each $\ParseTreeAlt_i$ = $\LeftChoiceParse{\ParseTreeAlt_i'}$ in \ParseTreeAlts],
define $\ParseTreeAlts_r$ = [$\ParseTreeAlt_i'$ for each $\ParseTreeAlt_i = \RightChoiceParse{\ParseTreeAlt_i'}$ in \ParseTreeAlts]
)\SynSim{(\Regex_1,)}{a})
\end{itemize}
\ref{fig:syntactic-lens-typing}
\fi

### 2016-05-05 paper/dnf-regex-lenses.tex
some latex and augeus access files

@@
 \input{fig_dnf-regex-definition}
+\input{fig_dnf-regex-semantics}
 A language which removes some of those equivalences is the language of regular
@@
 equivalences corresponding to distributivity, associativity, and concatenation identity.
-Figure~\ref{fig:dnf-regex-definition} defines this.
+Figure~\ref{fig:dnf-regex-definition} defines this, with semantics as in
+~\ref{fig:dnf-regex-semantics}
 The full DNF Regex is a list of Clauses.  This corresponds to a chain of $+$s

### 2016-05-12 paper/dnf-regex-lenses.tex
initial approach

@@
 whose types are dnf regular expressions.  We define dnf lenses in Figure~\ref{fig:dnf-lens-syntax}.
-These dnf lenses have the comparable underlying functions to the underlying functions
-of normal lenses, and are given in Figure~\ref{fig:dnf-lens-semantics}
+These dnf lenses have the comparable underlying relations to the underlyingrelations 
+of normal lenses, and are given in Figure~\ref{fig:dnf-lens-semantics}.
+There is also a similar theorem to Theorem~\ref{thm:lens-bij-fcn}.
+\begin{theorem}
+\label{thm:dnf-lens-bij-fcn}
+If $\Delta \vdash \DNFLens : \DNFRegex_1 \Leftrightarrow \DNFRegex_2$,
+then for all $x \in \LanguageOf{\Delta}{\DNFRegex_1}$, there exists a unique $y \in \LanguageOf{\Delta}{\DNFRegex_2}$ such that $\denot{\DNFLens}(x,y)$.
+This defines a bijective function called $\DNFLens.putr : \LanguageOf{\Delta}{\DNFRegex_1} \rightarrow \LanguageOf{\Delta}{\DNFRegex_2}$,
+whose inverse is called $\DNFLens.putl$.
+\end{theorem}
 TODO: make a nice diagram of how clause lenses work.

### 2016-05-12 paper/intro.tex
some latex and augeus access files
initial approach

@@
 \input{fig_regex-definition}
+\input{fig_regex-semantics}
 \input{fig_lens-syntax}
@@
 We have lenses, defined in Figure~\ref{fig:lens-syntax}, typed as in Figure~\ref{fig:lens-typing}.
-These lenses have underlying functions, given via Figure~\ref{fig:lens-semantics}.
+These lenses have a semantic interpretation as a relation, given via Figure~\ref{fig:lens-semantics}.
+\begin{theorem}
+\label{thm:lens-bij-fcn}
+If $\Delta \vdash \Lens : \Regex_1 \Leftrightarrow \Regex_2$,
+then for all $x \in \LanguageOf{\Delta}{\Regex_1}$, there exists a unique $y \in \LanguageOf{\Delta}{\Regex_2}$ such that $\denot{\Lens}(x,y)$.
+This defines a bijective function called $\Lens.putr : \LanguageOf{\Delta}{\Regex_1} \rightarrow \LanguageOf{\Delta}{\Regex_2}$,
+whose inverse is called $\Lens.putl$.
+\end{theorem}
 We would like to be able to synthesize these lenses automatically, given a

### 2016-05-24 paper/dnf-regex-lenses.tex
added in some tex

@@
 
-\input{fig_dnf-regex-definition}
-\input{fig_dnf-regex-semantics}
 A language which removes some of those equivalences is the language of regular
@@
 
-\input{fig_dnf-lens-syntax}
-\input{fig_dnf-lens-semantics}
-\input{fig_dnf-lens-typing}
 As we have dnf regular expressions, it also makes sense to have dnf lenses,

### 2016-05-24 paper/put-together.tex
added in some tex

Lenses are two functions, that satisfy the get-put laws.
But it is annoying to prove, automatically or through provided proofs,
that two functions satisfy the get-put laws.

We can take a language based approach, where the language only allows for
the writing of functions
that satisfy the get-put laws.
This has been explored in previous work like Boomerang\cite{boomerang}.
A close inspection of the laws behind the lenses shows that the typing of the
lenses corresponds tightly with the actual lenses.
For example, consider the concatenation rule

\begin{mathpar}
\inferrule*
{
\Lens_1 : \Regex_1 \Leftrightarrow \RegexAlt_1\\
\Lens_2 : \Regex_2 \Leftrightarrow \RegexAlt_2\\\\
\UnambigConcat{\Language(\Regex_1)}{\Language(\Regex_2)}\\
\UnambigConcat{\Language(\Regex_2)}{\Language(\RegexAlt_2)}\\
}
{
\ConcatLens{\Lens_1}{\Lens_2} :
\RegexConcat{\Regex_1}{\Regex_2} \Leftrightarrow
\RegexConcat{\RegexAlt_1}{\RegexAlt_2}
}
\end{mathpar}

Concatenation lenses come alongside a regular expression concatenation.
Perhaps, instead of writing out lenses explicitly, we can merely
write its type, and the lenses can be inferred from the types.
Utilizing types for synthesis is not a new idea, and has been explored
in systems like Myth \cite{tds-pldi}.
However, we have found that type directed synthesis is well suited
for domain specific languages like the language of lenses.
The constraints the type systems of these languages make the search space
tighter, and allow us to synthesize very complicated
programs very quickly.

\input{fig_regex-definition}
\input{fig_regex-semantics}
However, if we are to use the types to direct, and the types are
regular expressions, we would like to allow for users to easily engineer
regular expressions.
For this we have created a language of regular expressions with a context,
with syntax defined in Figure~\ref{fig:regex-definition}, and semantics defined
in Figure~\ref{fig:regex-semantics}.  In this, we only allow the context to
have finite size.
We show that adding in a context doesn't increase expressibility.
\begin{restatable}[Completeness of the Regular Expressions Without Variables]{theorem}{ecrc}
\label{thm:empty-context-regex-completeness}
For all contexts \RegexContext{} and regular expressions \Regex{}, there exists a regular expression \Regex{}',
such that \LanguageOf{\RegexContext{}}{\Regex{}}=\LanguageOf{\emptyset}{\Regex{}'}.
\end{restatable}
\begin{proof}
By induction on the size of the context and the regular expression
\end{proof}

\input{fig_lens-syntax}
\input{fig_lens-alternate-alternate-semantics}

We choose to work on a restricted class of lenses,
bijective lenses.
While normal lenses allow for loss of data in going from
the source to the target, bijective lenses do not allow for any loss of data.
With this further restriction, we arrive at the lenses shown in Figure~\ref{fig:lens-syntax}.
These lenses have a type system, shown in Figure~\ref{fig:lens-alternate-alternate-semantics},
even more suited to synthesis than the normal language of lenses.
In these, the typing $\FullContext \DerivesLens \Lens \OfType \MapsBetweenTypeOf{\Regex}{\RegexAlt} \HasSemantics \PutLeft,\PutRight$ means
under the regular expression context \RegexContext{} and the lens context \LensContext{},
there is a bijective lens \Lens{}, with two underlying functions,
\PutRight{} and \PutLeft{}, where $\PutRight \OfType \ArrowTypeOf{\LanguageOf{\RegexContext}{\Regex}}{\LanguageOf{\RegexContext}{\RegexAlt}}$
and $\PutLeft \OfType \ArrowTypeOf{\LanguageOf{\RegexContext}{\RegexAlt}}{\LanguageOf{\RegexContext}{\Regex}}$.
We can prove that the system only expresses bijective
functions, but does not express all bijective functions.
\begin{restatable}[Soundness of the Language of Bijective Lenses]{theorem}{bls}
\label{thm:bij-lens-soundness}
If $\FullContext \vdash \Lens : \Regex \Leftrightarrow \RegexAlt \HasSemantics \PutRight,\PutLeft$,
then \PutRight{} is a bijective function from \LanguageOf{\Delta{}}{\Regex{}} to \LanguageOf{\Delta{}}{\RegexAlt{}}, with inverse $\PutLeft{}$.
\end{restatable}
\begin{proof}
Through induction on typing derivations.
\end{proof}
\begin{restatable}[Incompleteness of the Language of Bijective Lenses]{theorem}{blic}
\label{thm:bij-lens-incompleteness}
There exists regular expressions $\Regex$, $\RegexAlt$,
a regular expression context $\RegexContext$, a lens context $\LensContext$, and a bijective function $f:\LanguageOf{\Delta}{\Regex}\rightarrow\LanguageOf{\Delta}{\RegexAlt}$
such that for all lenses \Lens{}, there does not exist a derivation for
$\FullContext \vdash \Lens : \Regex \Leftrightarrow \RegexAlt \HasSemantics f,\InverseOf{f}$.
\end{restatable}
\begin{proof}
Through a specific bijection between $\LanguageOf{\emptyset}{\Star{a}}$ and $\LanguageOf{\emptyset}{\Star{a}}$.
\end{proof}
or
\begin{proof}
Through a counting argument of the possible number of bijections between $\LanguageOf{\emptyset}{\Star{a}}$ and $\LanguageOf{\emptyset}{\Star{a}}$.
\end{proof}

Interestingly, the use of lens contexts and regular expression contexts
is not needed.
This can be formalized in the following theorem.
\begin{restatable}[Completeness of Bijective Lenses Without Variables]{theorem}{eclc}
\label{thm:empty-context-lens-completeness}
If $\FullContext \vdash \Lens : \Regex \Leftrightarrow \RegexAlt \HasSemantics \PutRight,\PutLeft$,
then there exists a $\Lens'$, $\Regex'$, $\RegexAlt'$, such that
\LanguageOf{\RegexContext}{\Regex}=\LanguageOf{\emptyset}{\Regex'},
\LanguageOf{\RegexContext}{\RegexAlt}=\LanguageOf{\emptyset}{\RegexAlt'}, and
$\emptyset,\emptyset \DerivesLens \Lens' \OfType \MapsBetweenTypeOf{\Regex'}{\RegexAlt'} \HasSemantics \PutRight,\PutLeft$.
\end{restatable}
\begin{proof}
Through induction on typing derivations.
\end{proof}

In this, all of the rules except Retype Lens have a clear synthesis algorithm.
For example, given the type $\RegexConcat{\Regex_1}{\Regex_2}\Leftrightarrow\RegexConcat{\RegexAlt_1}{\RegexAlt_2}$.
We know that it can be created from a Concat Lens rule,
which would create the subproblems of $\Regex_1\Leftrightarrow\RegexAlt_1$ and $\Regex_2\Leftrightarrow\RegexAlt_2$;
it can be created from a Swap Lens rule, which would create the subproblems of $\Regex_1\Leftrightarrow\RegexAlt_2$ and $\Regex_2\Leftrightarrow\RegexAlt_1$;
and it can be created from Retype Lens, for which the subproblems are not clear.

However, the Retype Lens rule is a very important rule.
Without the Retype Lens rule, it would not be possible to have a lens of the type
$a\Concat (b\Concat (c\Concat d))\Leftrightarrow(a\Concat b)\Concat(c\Concat d)$,
despite the fact that we expect this lens to be inhabited by the lens \IdentityLens{}.

\input{fig_dnf-lens-syntax}
\input{fig_dnf-lens-alternate-alternate-semantics}

Because of this, to synthesize lenses, we instead create a new, equivalent language,
who handles many of the equivalences of regular expressions through exploiting a normal form.
We have created a language which can handle many of the equivalences of regular
expressions through its normal form, and adds additional structure to the search
space, constraining it and minimizing the traversal through equivalent elements.
We call this language, the language of DNF lenses, due to the fact these lenses are
put into a form similar to the disjunctive normal form of logic.
The syntax for these lenses is provided in Figure~\ref{fig:dnf-lens-syntax}.
The typing and semantics for these lenses is provided in Figure~\ref{fig:dnf-lens-alternate-alternate-semantics}.

DNF lenses handle the equivalences of concatenation associativity,
or associativity, or commutativity, and and-or distributivity.
Furthermore, they do not require concatenation to be complete,
though we do allow for concatenation of lenses in the lens context.
While these are not needed, this is under the assumption that if lenses are present
in the context, they are likely to be used.
Because of this, the choices needed to be done by the synthesis algorithm
are fairly minimal.  There are only
a few types of equivalences related to stars and use of regular expressions
in the context that need to be explored for completeness.
Furthermore, there is the difficulty of determining which permutations create
well typed regular expressions, but our implentation reduces the problem of
finding a valid permutation, to ordering of terms under a specific order.

### 2016-05-29 paper/fig_dnf-regex-syntax.tex
added latexes

\begin{figure}
\begin{tabular}{l@{\ }l@{\ }c@{\ }l@{\ }r}

% DNF_REGEX
(Atoms)& \Atom{} & \GEq{} & \RegexVariable{} & variable\\
& & &\GBar{} \DNFRegex{}* & iterate \\
(Conjunct)& \Conjunct{} & \GEq{} &
$[\String_0;\Atom_1;\ldots;\Atom_n;\String_n]$ & conjunct\\
(DNF Regex)& \DNFRegex{} & \GEq{} & $[\Conjunct_1;\ldots;\Conjunct_n]$ & dnf regex\\
\end{tabular}
\caption{DNF Regex Syntax}
\label{fig:dnf-regex-syntax}
\end{figure}

### 2016-05-29 paper/fig_regex-syntax.tex
added latexes

\begin{figure}
\centering
\begin{tabular}{l@{\ }l@{\ }c@{\ }l@{\ }r}

% REGEX
(Regexs)& \Regex{} & \GEq{} & \RegexVariable{} & variable \\
& & & \GBar{} s $\in \Sigma^*$ & base \\
& & & \GBar{} \Regex{}* & star \\
& & & \GBar{} $\Regex_1 \Regex_2$ & concat \\
& & & \GBar{} $\Regex_1 | \Regex_2$ & or \\
\end{tabular}
\caption{Regex Syntax}
\label{fig:regex-syntax}
\end{figure}

### 2016-05-29 paper/put-together.tex
added latexes

@@
 
-\input{fig_regex-definition}
+\input{fig_regex-syntax}
 \input{fig_regex-semantics}
@@
 For this we have created a language of regular expressions with a context,
-with syntax defined in Figure~\ref{fig:regex-definition}, and semantics defined
+with syntax defined in Figure~\ref{fig:regex-syntax}, and semantics defined
 in Figure~\ref{fig:regex-semantics}.  In this, we only allow the context to
@@
 \end{proof}
+However, despite the fact these variables are not needed, they are useful for nice
+separation of concerns.
+Lenses in the context are likely to be used, much like when programming, one
+doesn't import libraries for no reason.
 
@@
 
+\input{fig_dnf-regex-syntax}
+\input{fig_dnf-regex-semantics}
+\input{fig_dnf-regex-functions}
+\input{fig_dnf-regex-rewrites}
+Because of this, to synthesize lenses, we instead create a new, equivalent language,
+who handles many of the equivalences of regular expressions through exploiting a normal form.
+We define this normal form of regular expression DNF Regular expression, with syntax
+defined as in Figure~\ref{fig:dnf-regex-syntax}, and with semantics defined as
+in Figure~\ref{fig:dnf-regex-semantics}.  These semantics rely on some functions
+defined in Figure~\ref{fig:dnf-regex-functions},
+and some rewrite rules defined in Figure~\ref{fig:dnf-regex-rewrites}.
+This normal form is called a DNF regular expression, as it is similar to the
+sentences in logical disjunctive normal form.
+These DNF regular expression expresses exactly the same amount as standard regular
+expressions.
+\begin{restatable}[Completeness of DNF Regexs]{theorem}{dnfrc}
+\label{thm:completeness-dnf-lenses}
+There exists a function $\ToDNFRegex\OfType\ArrowTypeOf{\mathit{Regex}}{\mathit{DNF Regex}}$,
+such that, for all $\RegexContext$, $\LanguageOf{\RegexContext}{\Regex}=\LanguageOf{\RegexContext}{\ToDNFRegex(\Regex)}$.
+\end{restatable}
+
+\begin{restatable}[Soundness of DNF Regexs]{theorem}{dnfrs}
+\label{thm:soundness-dnf-lenses}
+The function \ToDNFRegex{} has a right inverse \FromDNFRegex{}.
+\end{restatable}
+
 \input{fig_dnf-lens-syntax}
 \input{fig_dnf-lens-alternate-alternate-semantics}
-
-Because of this, to synthesize lenses, we instead create a new, equivalent language,
-who handles many of the equivalences of regular expressions through exploiting a normal form.
 We have created a language which can handle many of the equivalences of regular
-expressions through its normal form, and adds additional structure to the search
+expressions through operation on dnf regular expressions,
+and adds additional structure to the search
 space, constraining it and minimizing the traversal through equivalent elements.
-We call this language, the language of DNF lenses, due to the fact these lenses are
-put into a form similar to the disjunctive normal form of logic.
+We call this language, the language of DNF lenses, due to the fact the lenses
+operate on DNF regular expressions.
 The syntax for these lenses is provided in Figure~\ref{fig:dnf-lens-syntax}.
 The typing and semantics for these lenses is provided in Figure~\ref{fig:dnf-lens-alternate-alternate-semantics}.
+DNF Lenses express exactly the same amount as standard lenses.
+\begin{restatable}[Completeness of DNF Lenses]{theorem}{dnflc}
+\label{thm:completeness-dnf-lenses}
+If there exists a derivation of $\FullContext \DerivesLens \Lens \OfType \MapsBetweenTypeOf{\Regex}{\RegexAlt} \HasSemantics \PutRight,\PutLeft$,
+then there exists a derivation of $\FullContext \DerivesDNFLens \DNFLens \OfType \MapsBetweenTypeOf{\DNFRegex}{\DNFRegexAlt} \HasSemantics \PutRight,\PutLeft$ such that
+$\LanguageOf{\RegexContext}{\DNFRegex}=\LanguageOf{\RegexContext}{\Regex}$, and
+$\LanguageOf{\RegexContext}{\DNFRegexAlt}=\LanguageOf{\RegexContext}{\RegexAlt}$.
+\end{restatable}
+
+\begin{restatable}[Soundness of DNF Lenses]{theorem}{dnfls}
+\label{thm:soundness-dnf-lenses}
+If there exists a derivation of $\FullContext \DerivesDNFLens \DNFLens \OfType \MapsBetweenTypeOf{\DNFRegex}{\DNFRegexAlt} \HasSemantics \PutRight,\PutLeft$,
+then there exists a derivation of $\FullContext \DerivesLens \Lens \OfType \MapsBetweenTypeOf{\Regex}{\RegexAlt} \HasSemantics \PutRight,\PutLeft$ such that
+$\LanguageOf{\RegexContext}{\Regex}=\LanguageOf{\RegexContext}{\DNFRegex}$, and
+$\LanguageOf{\RegexContext}{\RegexAlt}=\LanguageOf{\RegexContext}{\DNFRegexAlt}$.
+\end{restatable}
 
@@
 or associativity, or commutativity, and and-or distributivity.
-Furthermore, they do not require concatenation to be complete,
+Furthermore, they do not require composition to be complete,
 though we do allow for concatenation of lenses in the lens context.
-While these are not needed, this is under the assumption that if lenses are present
+While these are not needed for full expressibility because of Theorem~\ref{thm:empty-context-regex-completeness}
+and Theorem~\ref{thm:completeness-dnf-lenses},
+we choose to use these contextual lenses and regular expression.
+This is under the assumption that if lenses are present
 in the context, they are likely to be used.
+However, this does add in additional choices for synthesis, but we've found
+that in practice, it simplifies the algorithm (TODO: this analysis).
 Because of this, the choices needed to be done by the synthesis algorithm
@@
 well typed regular expressions, but our implentation reduces the problem of
-finding a valid permutation, to ordering of terms under a specific order.
+finding a valid permutation to ordering the terms under a specific order.
+
+\input{fig_regex-equivalence-rules}
+Two regular expressions are said to be equivalent if they are equivalent through
+any of the rules specified in Figure~\ref{fig:regex-equivalence-rules}.
+Indeed, after ordering we have put our regular expressions in a very normal form.
+If we normalize two regular expressions, and those two regular expressions are
+equivalent through any number of applications of equivalences 1 through 8, then
+their ordered DNF regular expressions will be equivalent.
+What's more, we can use the ordering on ordered DNF regular expressions to create another
+equivalence class.  The ordering on DNF regular expressions says two ordered
+dnf regular expressions are equal if, and only if, there exists a lens between
+then, where the lens typing doesn't include the DNF Rewrite Lens rule.  The purpose
+of the DNF Rewrite rule is to handle the other types of equivalences.

### 2016-05-31 paper/fig_dnf-regex-syntax.tex
added in some clenaer stuff

@@
 % DNF_REGEX
-(Atoms)& \Atom{} & \GEq{} & \RegexVariable{} & variable\\
-& & &\GBar{} \DNFRegex{}* & iterate \\
-(Conjunct)& \Conjunct{} & \GEq{} &
-$[\String_0;\Atom_1;\ldots;\Atom_n;\String_n]$ & conjunct\\
-(DNF Regex)& \DNFRegex{} & \GEq{} & $[\Conjunct_1;\ldots;\Conjunct_n]$ & dnf regex\\
+(Atoms)& \Atom{},\AtomAlt{} & \GEq{} & \DNFRegex{}* & Iterate DNF\\
+(Conjunct)& \Conjunct{},\ConjunctAlt{} & \GEq{} &
+$[\String_0;\Atom_1;\ldots;\Atom_n;\String_n]$ & Conjoin\\
+(DNF Regex)& \DNFRegex{},\DNFRegexAlt{} & \GEq{} & $[\Conjunct_1;\ldots;\Conjunct_n]$ & DNF Or\\
 \end{tabular}

### 2016-05-31 paper/fig_regex-syntax.tex
added in some clenaer stuff

@@
 % REGEX
-(Regexs)& \Regex{} & \GEq{} & \RegexVariable{} & variable \\
-& & & \GBar{} s $\in \Sigma^*$ & base \\
-& & & \GBar{} \Regex{}* & star \\
-& & & \GBar{} $\Regex_1 \Regex_2$ & concat \\
-& & & \GBar{} $\Regex_1 | \Regex_2$ & or \\
+(String)& \String{},\StringAlt{} & \GEq{} & $\String\in\Star{\Sigma}$ \\
+(Regexs)& \Regex{},\RegexAlt{} & \GEq{} & s & Base \\
+& & & \GBar{} \Regex{}* & Star \\
+& & & \GBar{} $\RegexConcat{\Regex_1}{\Regex_2}$ & Concat \\
+& & & \GBar{} $\Regex_1 | \Regex_2$ & Or \\
 \end{tabular}

### 2016-05-31 paper/reordered.tex
added in some clenaer stuff

\input{fig_regex-syntax}
\input{fig_regex-semantics}

Regular expressions are a nice way to formalize a set of strings, called a language.
They have nice computational properties.  The syntax for nonempty regular expressions
is given in Figure~\ref{fig:regex-syntax}.
These regular expressions have an underlying semantics, which expresses languages,
formalized in Figure~\ref{fig:regex-semantics}.

\input{fig_lens-syntax}
\input{fig_lens-alternate-alternate-semantics}
There are things called views, which allow one a different perspective on data.
Lenses are updatable views, which satisfy the get-put laws.
One can approach creating lenses that satisfy the laws through a language approach.
We focus our attentions currently on the language of bijective lenses, where the
functions between views are bijections, and inverses to each other.
We define the syntax for these bijective lenses in Figure~\ref{fig:lens-syntax}.
The semantics and typing for these lenses is defined in Figure~\ref{fig:lens-alternate-alternate-semantics}.
The semantics is only defined alongside the typing, as the functions are only
well defined when the term is well typed.

A close inspection of the laws behind the lenses shows that the typing of the
lenses corresponds tightly with the actual lenses.
For example, consider the concatenation rule:

\begin{mathpar}
\inferrule*
{
\Lens_1 : \Regex_1 \Leftrightarrow \RegexAlt_1\\
\Lens_2 : \Regex_2 \Leftrightarrow \RegexAlt_2\\\\
\UnambigConcat{\LanguageOf{\Regex_1}}{\LanguageOf{\Regex_2}}\\
\UnambigConcat{\LanguageOf{\Regex_2}}{\LanguageOf{\RegexAlt_2}}\\
}
{
\ConcatLens{\Lens_1}{\Lens_2} :
\RegexConcat{\Regex_1}{\Regex_2} \Leftrightarrow
\RegexConcat{\RegexAlt_1}{\RegexAlt_2}
}
\end{mathpar}

Concatenation lenses come alongside a regular expression concatenation.
Perhaps, instead of writing out lenses explicitly, we can merely
write its type, and the lenses can be inferred from the types.
Utilizing types for synthesis is not a new idea, and has been explored
in systems like Myth \cite{tds-pldi}.
However, we have found that type directed synthesis is well suited
for domain specific languages like the language of lenses.
The constraints of the type systems of these languages make the search space
tighter, and allow us to synthesize very complicated
programs very quickly.

Our approach to synthesis is one of enumeration.
We would like to enumerate the programs that satisfy the desired typing until
we find one that satisfies the examples.
We enumerate these programs by creating subproblems based on typing rules
could potentially create the desired type.

For example, given the type $\RegexConcat{\Regex_1}{\Regex_2}\Leftrightarrow\RegexConcat{\RegexAlt_1}{\RegexAlt_2}$.
We know that it can be created from a Concat Lens rule,
which would create the subproblems of finding a lens of type $\Regex_1\Leftrightarrow\RegexAlt_1$
and a lens of type $\Regex_2\Leftrightarrow\RegexAlt_2$.
However, this is made more difficult due to the Compose Lens rule,
and the Retype Lens rule.
Both of these rules can be applied to any type, and the subproblems are not immediately apparent.
For example, if the lens of type
$\RegexConcat{\Regex_1}{\Regex_2}\Leftrightarrow\RegexConcat{\RegexAlt_1}{\RegexAlt_2}$
came from a Compose Lens rule, then for some regular expression $\Regex$,
we would have the subproblems of finding a lens with type
$\RegexConcat{\Regex_1}{\Regex_2}\Leftrightarrow\Regex$,
and finding a lens with type
$\Regex\Leftrightarrow\RegexConcat{\RegexAlt_1}{\RegexAlt_2}$.
However, merely from the resulting type,
$\RegexConcat{\Regex_1}{\Regex_2} \Leftrightarrow \RegexConcat{\RegexAlt_1}{\RegexAlt_2}$,
we do not know which regular expression \Regex{} to use.

Similarly, if the lens of type
$\RegexConcat{\Regex_1}{\Regex_2}\Leftrightarrow\RegexConcat{\RegexAlt_1}{\RegexAlt_2}$
came from a Retype Lens rule, then for some regular expressions $\Regex$, and $\RegexAlt$,
where $\LanguageOf{\Regex}=\LanguageOf{\RegexConcat{\Regex_1}{\Regex_2}}$
and $\LanguageOf{\RegexAlt}=\LanguageOf{\RegexConcat{\RegexAlt_1}{\RegexAlt_2}}$,
we have the subproblem of finding a lens with type
$\Regex\Leftrightarrow\RegexAlt$.
However, merely from the resulting type,
$\RegexConcat{\Regex_1}{\Regex_2}\Leftrightarrow\RegexConcat{\RegexAlt_1}{\RegexAlt_2}$,
we do not know which regular expressions \Regex{} and \RegexAlt{} to
use, we only know that they have to meet some semantic restrictions.

However, neither of these rules are admissible and both are important.
Without the Retype Lens rule, it would not be possible to have a lens of the type
$a\Concat (b\Concat (c\Concat d))\Leftrightarrow(a\Concat b)\Concat(c\Concat d)$,
despite the fact that we expect this lens to be inhabited by the lens \IdentityLens{}.
Without the Compose Lens rule, it would not be possible to have a lens $\Lens$ of the type
$\RegexConcat{\Star{a}}{\RegexConcat{\Star{b}}{\RegexConcat{\Star{c}}{\Star{d}}}} \Leftrightarrow
\RegexConcat{\Star{c}}{\RegexConcat{\Star{a}}{\RegexConcat{\Star{d}}{\Star{b}}}}$,
which has $\PutRightOf{\Lens}(a^h\Concat b^i \Concat c^j \Concat d^k)=c^j\Concat a^h\Concat d^k\Concat b^i$, a simple permutation of the letters.

\input{fig_regex-equivalence-rules}
These rules are important, but on each application of the rules, there are an
infinite number of potential subproblems.
Making the composition admissible is not the most complicated problem,
as it's lack of admissibility comes a bit from the rigid structure of the rules,
and only arises in an edge case.
However, retyping is not so easily made admissible.
For the handling of the Retype Lens
rule, we can create a search space for these potential subproblems by
searching through the lenses of equivalent regular expressions.
For this, we use the equational theory presented by Conway \cite{conway},
and proven complete by Krob \cite{Krob},
shown in Figure~\ref{fig:regex-equivalence-rules}.

However, enumerating through all possible regular expressions for lenses is inefficient.
Beyond that, we are not merely searching for the applications of these rules to
a single regular expression, but instead the applications of these to two regular expressions.
This makes this search space incredibly broad.
Instead of searching through this search space, we would instead like to find
a normal form for regular expressions, and find a way to create lenses on
regular expressions of that form.

\input{fig_dnf-regex-syntax}
\input{fig_dnf-regex-semantics}
To this end, we have created DNF regular expressions.  Intuitively, a DNF regular
expression is a regular expression with the distributivity laws (6 and 7) always
applied, with no associativity between operations of the same type,
and with base regular expressions at fixed locations.
We formalize the syntax of this language in Figure~\ref{fig:dnf-regex-syntax}.

The outermost layer is a list of conjuncts.
This layer intuitively corresponds to the choices involved in regular expression matching, and so represents where the Ors of normal regular expressions exist.
The second layer is a list of alternating strings and stars.
After the choice has been made about what will be expressed,
the base strings and iterated portions remains to be expressed.
We keep it in a normal form, by requiring a (possibly empty) string between
each clause.
The clause corresponds to the concatenated data, and is a concatenation of the
fixed data of base strings, and the iterated data of stars.
Finally is the atom, which is a star.
This corresponds to the iteration that takes place in normal stars.
There is no fixed way to break the choices of this into the clause,
as there is an arbitrarily large number of choices made in the stars.
This intuition is formalized by the semantics, given in Figure~\ref{fig:dnf-regex-semantics}.

\input{fig_dnf-regex-functions}
We can create a dnf regular expression from a regular expression through repeated
application of the distributivity rule, and through removal of association information.  To do this, we require some functions defined on DNF regular expressions,
which we define in Figure~\ref{fig:dnf-regex-functions}.
From this we can define a function which converts a regular expression into
an equivalent DNF regular expression, \ToDNFRegex{}.
\begin{definition}
\leavevmode
\begin{itemize}
\item $\ToDNFRegex(\String)=[[\String]]$
\item $\ToDNFRegex(\Star{\Regex}) = [[\Star{\ToDNFRegex(\Regex)}]]$
\item $\ToDNFRegex(\RegexConcat{\Regex_1}{\Regex_2}) =$\\
\hspace*{1em}$\ConcatDNF(\ToDNFRegex(\Regex_1),\ToDNFRegex(\Regex_2))$
\item $\ToDNFRegex(\RegexOr{\Regex_1}{\Regex_2}) =$\\
\hspace*{1em}$\OrDNF(\ToDNFRegex(\Regex_1),\ToDNFRegex(\Regex_2))$
\end{itemize}
\end{definition}
We show that this transformation is a valid transformation.
\begin{restatable}[Completeness of DNF Regexs]{theorem}{dnfrc}
\label{thm:completeness-dnf-lenses}
For all regular expressions \Regex{},
$\LanguageOf{\Regex}=\LanguageOf{\ToDNFRegex(\Regex)}$.
\end{restatable}

Furthermore, we can use this to prove that the language of DNF regular expressions is sound.

\begin{restatable}[Soundness of DNF Regexs]{theorem}{dnfrs}
\label{thm:soundness-dnf-lenses}
The function \ToDNFRegex{} has a right inverse \FromDNFRegex{}.
\end{restatable}

We can further normalize the DNF regular expressions by providing an order on conjuncts,
and ordering the conjuncts in the DNF regular expression according to that order.
With this, we have normalized the regular expression equivalences 1-7.
In other words, if two regular expressions, $\Regex_1$ and $\Regex_2$,
are equivalent up to the transformations of regular expressions in equivalences 1-7,
then $\ToDNFRegex(\Regex_1)=\ToDNFRegex(\Regex_2)$.
A reasonable approach to further normalizing this is to attempt to minimize the
regular expression as much as possible, using the star laws.
Unfortunately, while this approach works well for normalizing, it doesn't work
well for finding lenses.  For example, there is clearly a lens of type
$\MapsBetweenTypeOf{\RegexOr{\EmptyString}{(\RegexConcat{a}{\Star{a}})}}{\RegexOr{b}{(\RegexConcat{c}{\Star{d}})}}$,
namely $\OrLens{\ConstLens{\EmptyString}{a}}{\ConcatLens{\ConstLens{a}{b}}{\IterateLens{\ConstLens{a}{d}}}}$.
However, if one simplifies as much as possible, then they will have to find a lens
of type $\MapsBetweenTypeOf{\Star{a}}{\RegexOr{b}{(\RegexConcat{c}{\Star{d}})}}$.
Unlike the type before being minified, where there was a type directed way to find
this lens, there is no longer a type directed way: should the lens be an iterate lens
or an or lens.
We interpret this as the need to gain semantic information through the expansion of
the regular expressions.
A regular expression merely of the form $\Star{a}$ only cares about how the iterated case is handled, where a regular expression of the form
$\RegexOr{\EmptyString}{(\RegexConcat{a}{\Star{a}})}$ potentially acts differently on the empty
string case than on the nonempty case.

\input{fig_dnf-regex-rewrites}
Because of this, instead of writing equivalences for DNF regular expressions like
exist for normal regular expressions,
instead we write rewrite rules as shown in Figure~\ref{fig:dnf-regex-rewrites}.
Intuitively, these rewrite rules correspond to breaking a regular expression
into a regular expression that encodes additional semantics.
We define these rules as rewrites that turn atoms into DNF Regular expressions,
and a rule that expresses how these rewrites can become rewrites on
the DNF regular expressions themselves.

Roughly, the Atom Sumstar rewrite corresponds to Regular Expression equivalence 8,
the Atom Unrollstar rewrite corresponds to Regular Expression equivalence 9,
and the Atom Powerstar rewrite corresponds to Regular Expression equivalence 11.
There is no rewrite corresponding to Regular Expression equivalence 10, as that
equivalence is an ambiguity introducing equivalence.
An application of Atom Sumstar corresponds to doing different things after the last time a certain event has occured, and before the last time that event occurs.
An application of Atom Unrollstar corresponds to making a different action for
the empty case of an iteration, the first case of an iteration, and all further cases of the iteration.
An application of Atom Powerstar corresponds to doing different things if the iteration
is occuring a different number of times, modulo n, for some fixed n.

\input{fig_dnf-lens-syntax}
\input{fig_dnf-lens-alternate-alternate-semantics}
Armed with these rewrites, we have the capabilities to define a sufficiently
strong language for lenses on these regular expressions.
The syntax is defined in Figure~\ref{fig:dnf-lens-syntax}.
Similarly to the language of lenses, we aim to have the typing of the lenses
correspond closely to the syntax for the lenses themselves.
The typing and semantics of these dnf lenses are defined in Figure~\ref{fig:dnf-lens-alternate-alternate-semantics}.
Intuitively, a DNF Regex Lens corresponds roughly to an n-ary version of an or lens,
a Conjunct Lens corresponds to const lenses for the strings, and a combination of
Concat and Swap lenses for the Atoms.
Furthermore, these lenses are sufficiently strong that Compose Lenses are not
necessary.
Indeed, these lenses are strong enough to express everything expressible in the language of lenses.
\begin{restatable}[Completeness of DNF Lenses]{theorem}{dnflc}
\label{thm:completeness-dnf-lenses}
If there exists a derivation of $\Lens \OfType \MapsBetweenTypeOf{\Regex}{\RegexAlt} \HasSemantics \PutRight,\PutLeft$,
then there exists a derivation of $\DNFLens \OfType \MapsBetweenTypeOf{\DNFRegex}{\DNFRegexAlt} \HasSemantics \PutRight,\PutLeft$ such that
$\LanguageOf{\DNFRegex}=\LanguageOf{\Regex}$, and
$\LanguageOf{\DNFRegexAlt}=\LanguageOf{\RegexAlt}$.
\end{restatable}
Furthermore, they only express things expressible in the language of lenses.
\begin{restatable}[Soundness of DNF Lenses]{theorem}{dnfls}
\label{thm:soundness-dnf-lenses}
If there exists a derivation of $\DNFLens \OfType \MapsBetweenTypeOf{\DNFRegex}{\DNFRegexAlt} \HasSemantics \PutRight,\PutLeft$,
then there exists a derivation of $\Lens \OfType \MapsBetweenTypeOf{\Regex}{\RegexAlt} \HasSemantics \PutRight,\PutLeft$ such that
$\LanguageOf{\Regex}=\LanguageOf{\DNFRegex}$, and
$\LanguageOf{\RegexAlt}=\LanguageOf{\DNFRegexAlt}$.
\end{restatable}

However, these lenses are much more suited to synthesis.
The vast majority of the rules have a syntax directed synthesis algorithm.
Furthermore, even the rewrite rules that don't have an immediate syntax directed
synthesis algorithm have an underlying semantic meaning which can be used
to direct the solution.

### 2016-05-31 paper/reordered.tex
Bunch of small comments

@@
 
-Regular expressions are a nice way to formalize a set of strings, called a language.
-They have nice computational properties.  The syntax for nonempty regular expressions
+The syntax for nonempty\bcp{why not empty ones too??} regular expressions
 is given in Figure~\ref{fig:regex-syntax}.
@@
 \input{fig_lens-alternate-alternate-semantics}
-There are things called views, which allow one a different perspective on data.
-Lenses are updatable views, which satisfy the get-put laws.
-One can approach creating lenses that satisfy the laws through a language approach.
-We focus our attentions currently on the language of bijective lenses, where the
+We focus on the language of bijective lenses, where the
 functions between views are bijections, and inverses to each other.
@@
 exist for normal regular expressions,
-instead we write rewrite rules as shown in Figure~\ref{fig:dnf-regex-rewrites}.
+instead we write rewrite rules as shown in
+Figure~\ref{fig:dnf-regex-rewrites}.\bcp{Don't understand what's going on here.}
 Intuitively, these rewrite rules correspond to breaking a regular expression
@@
 equivalence is an ambiguity introducing equivalence.
-An application of Atom Sumstar corresponds to doing different things after the last time a certain event has occured, and before the last time that event occurs.
+An application of Atom Sumstar\bcp{Wrong font} corresponds to doing different things after the last time a certain event has occured, and before the last time that event occurs.
 An application of Atom Unrollstar corresponds to making a different action for

### 2016-06-01 paper/fig_regex-syntax.tex
Bunch of small comments
fixed benjmain's suggestions, half of appendix

--- 2016-05-29 paper/fig_regex-syntax.tex
@@
 \centering
-\begin{tabular}{l@{\ }l@{\ }c@{\ }l@{\ }r}
+\begin{tabular}{l@{\hspace*{5mm}}l@{\ }c@{\ }l@{\hspace*{5mm}}l}
 
 % REGEX
-(Regexs)& \Regex{} & \GEq{} & \RegexVariable{} & variable \\
-& & & \GBar{} s $\in \Sigma^*$ & base \\
-& & & \GBar{} \Regex{}* & star \\
-& & & \GBar{} $\Regex_1 \Regex_2$ & concat \\
-& & & \GBar{} $\Regex_1 | \Regex_2$ & or \\
+(String)& \String{},\StringAlt{} & \GEq{} & $\String\in\Star{\Sigma}$ \\
+(Regexs)& \Regex{},\RegexAlt{} & \GEq{} & s & \it Base \\
+& & & \GBar{} \Regex{}* & \it Star \\
+& & & \GBar{} $\RegexConcat{\Regex_1}{\Regex_2}$ & \it Concat \\
+& & & \GBar{} $\Regex_1 | \Regex_2$ & \it Or \\
 \end{tabular}
-\caption{Regex Syntax}
+\caption{Regex Syntax \bcp{Recommend using capital letters ($R$, $S$,
+    ...) for regular expressions.  And left-justifying the RH column and
+    putting it in italic (with more space separating it from the rest) so
+    that it's clear that it's commentary.}}
 \label{fig:regex-syntax}

### 2016-06-06 paper/fig_dnf-regex-syntax.tex
Fixed some issues

@@
 % DNF_REGEX
-(Atoms)& \Atom{},\AtomAlt{} & \GEq{} & \DNFRegex{}* & Iterate DNF\\
+(Atoms)& \Atom{},\AtomAlt{} & \GEq{} & \StarOf{\DNFRegex{}} & Iterate DNF\\
 (Conjunct)& \Conjunct{},\ConjunctAlt{} & \GEq{} &
-$[\String_0;\Atom_1;\ldots;\Atom_n;\String_n]$ & Conjoin\\
-(DNF Regex)& \DNFRegex{},\DNFRegexAlt{} & \GEq{} & $[\Conjunct_1;\ldots;\Conjunct_n]$ & DNF Or\\
+$\ConjunctOf{\String_0\ConjunctSep\Atom_1\ConjunctSep\ldots\ConjunctSep\Atom_n\ConjunctSep\String_n}$ & Conjoin\\
+(DNF Regex)& \DNFRegex{},\DNFRegexAlt{} & \GEq{} & $\DNFOf{\Conjunct_1\DNFSep\ldots\DNFSep\Conjunct_n}$ & DNF Or\\
 \end{tabular}

### 2016-06-06 paper/fig_regex-syntax.tex
Fixed some issues

--- 2016-05-29 paper/fig_regex-syntax.tex
@@
 \centering
-\begin{tabular}{l@{\ }l@{\ }c@{\ }l@{\ }r}
+\begin{tabular}{l@{\hspace*{5mm}}l@{\ }c@{\ }l@{\hspace*{5mm}}>{\itshape\/}l}
 
 % REGEX
-(Regexs)& \Regex{} & \GEq{} & \RegexVariable{} & variable \\
-& & & \GBar{} s $\in \Sigma^*$ & base \\
-& & & \GBar{} \Regex{}* & star \\
-& & & \GBar{} $\Regex_1 \Regex_2$ & concat \\
-& & & \GBar{} $\Regex_1 | \Regex_2$ & or \\
+(String)& \String{},\StringAlt{} & \GEq{} & $\String\in\StarOf{\Sigma}$ \\
+(Regexs)& \Regex{},\RegexAlt{} & \GEq{} & s & Base \\
+& & & \GBar{} $\emptyset$ & Empty \\
+& & & \GBar{} \Regex{}* & Star \\
+& & & \GBar{} $\RegexConcat{\Regex_1}{\Regex_2}$ & Concat \\
+& & & \GBar{} $\RegexOr{\Regex_1}{\Regex_2}$ & Or \\
 \end{tabular}

### 2016-06-07 generate-baseline.py
added python and generated data

@@ -8,9 +8,9 @@ import subprocess
 import sys
 import time
 
-TEST_EXT = '.ml'
+TEST_EXT = '.ls'
 BASELINE_EXT = '.out'
-FLAGS = ['-nosugar', '-noincomplete-warning']
+FLAGS = []
 
 def find_tests(root):
     tests = []
@@ -35,6 +35,7 @@ def gen_baseline(prog, path, base):
     except subprocess.CalledProcessError as err:
         output_string = "An error was raised by this test: {0}\n{1}".format(err.returncode, err.output)
         print("Encountered an error running test {0}".format(test_path))
+	print(output_string)
     with open(baseline_path, 'w') as f:
         f.write(gen_timestamp() + '\n')
         f.write(output_string)

### 2016-06-07 generate-data.py

@@ -1,16 +1,24 @@
-#!/usr/bin/python
+#!/usr/local/bin/python
 
 from __future__ import print_function
+from easyprocess import EasyProcess
 
 import os
+import csv
 from os.path import splitext, join
 import subprocess
 import sys
 import time
 
-TEST_EXT = '.ml'
+TEST_EXT = '.ls'
 BASELINE_EXT = '.out'
-FLAGS = ['-nosugar', '-data', '-noincomplete-warning']
+BASE_FLAGS = []
+TIMEOUT_TIME = 8
+
+REPETITION_COUNT = 10
+
+def transpose(matrix):
+    return zip(*matrix)
 
 def find_tests(root):
     tests = []
@@ -19,28 +27,88 @@ def find_tests(root):
         tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
     return tests
 
+def gather_datum(prog, path, base, additional_flags):
+    return EasyProcess([prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)]).call(timeout=TIMEOUT_TIME).stdout
+
 def gather_data(prog, path, base):
-    print(subprocess.check_output([prog] + FLAGS + [join(path, base + TEST_EXT)], stderr=subprocess.STDOUT), end='')
+    current_data = {"Test":join(path, base + TEST_EXT).replace("_","-")[6:]}
+    run_data = []
+    timeout = False
+    for iteration in range(REPETITION_COUNT):
+    	datum = gather_datum(prog, path, base,['-data'])
+	if datum == "":
+		timeout = True
+		break
+	else:
+		run_data.append(datum.split(","))
+    if timeout:
+	current_data["ComputationTime"]="Timeout"
+	current_data["ExamplesRequired"]="Timeout"
+    else:
+        run_data_transpose = transpose(run_data)
+	computation_time_col = [float(x) for x in run_data_transpose[0]]
+	examples_required_col = [float(x) for x in run_data_transpose[1]]
+	current_data["ComputationTime"]="{:.5f}".format(sum(computation_time_col)/len(computation_time_col))
+	current_data["ExamplesRequired"]="{:.1f}".format(sum(examples_required_col)/len(examples_required_col))
+    expanded_run_data = []
+    for iteration in range(REPETITION_COUNT):
+    	datum = gather_datum(prog, path, base,['-forceexpand'])
+	if datum == "":
+		timeout = True
+		break
+	else:
[…]

### 2016-06-07 paper/evaluation.tex
added some latex bug fixes

\section{Evaluation}

\input{fig_runtimes-table}
\subsection{Speed}
We ran

\input{fig_examples-required-table}
\subsection{Importance of Examples}

\input{fig_expanded-time-table}
\subsection{Importance of User Defined Data Types}

Evaluation~\ref{fig:examples-required-table}~\ref{fig:runtimes-table}~\ref{fig:expanded-time-table}.

### 2016-06-07 paper/fig_examples-required-table.tex
added some latex bug fixes

\begin{figure}
\centering
%\begin{tabular}{|l|c|}

\csvreader[tabular=|l|l|,
    table head=\hline\textbf{Test} & \textbf{Example Count} \\\hline,
    late after line = \\\hline]%
{generated-data/data.csv}{Test=\Test,ExamplesRequired=\ExamplesRequired}%
{\Test{} & \ExamplesRequired{}}%

%\hline\textbf{Test} & \textbf{Examples Required}
%\\\hline\hline
%\csvreader[head to column names]{generated-data/data.csv}{}
%{\Test& \ExamplesRequired\\\hline}
%\end{tabular}
\caption{Examples Required Table}
\label{fig:examples-required-table}
\end{figure}

### 2016-06-07 paper/fig_runtimes-table.tex
added some latex bug fixes

--- 2016-06-07 paper/fig_examples-required-table.tex
@@
 \csvreader[tabular=|l|l|,
-    table head=\hline\textbf{Test} & \textbf{Example Count} \\\hline,
-    late after line = \\\hline]%
-{generated-data/data.csv}{Test=\Test,ExamplesRequired=\ExamplesRequired}%
-{\Test{} & \ExamplesRequired{}}%
+    table head=\hline\textbf{Test} & \textbf{Time (s)} \\\hline,
+    late after line = \\\hline]
+{generated-data/data.csv}{Test=\Test,ComputationTime=\ComputationTime}
+{\Test{} & \ComputationTime{}}
 
@@
 %\end{tabular}
-\caption{Examples Required Table}
-\label{fig:examples-required-table}
+\caption{Runtime Table}
+\label{fig:runtimes-table}
 \end{figure}

### 2016-06-09 paper/evaluation.tex
added evaluation

@@
 \section{Evaluation}
+We ran our algorithm on tests validating our ability to handle regular
+expression equivalences, tests derived from the Augeus test suite, and from
+tests derived from the FlashFill tests.
 
@@
 \subsection{Speed}
-We ran
+Figure~\ref{fig:runtimes-table} provides the runtime of our algorithm on our
+test suite.
+In this figure, we want to minimize runtime, as a smaller runtime means we were
+able to synthesize very efficiently.
+We found that, on many of our difficult examples, our algorithm was able to find
+the appropriate lens very quickly.
+
+However, we can also see where the strengths and weaknesses are in our approach.
+In particular, our synthesis strategy does quite well on problems that only
+include a small number of transformations, even if the number of clauses is
+decently large.
+This is as expected, for after the appropriate transformations have been found,
+the problem of finding the lens between the resulting DNF Regular Expressions
+is merely as difficult as ordering the clauses.
+
+For example, the address problem is the second slowest lens to synthesize.
+This is because there is a large number of required expansions of user defined
+data types, and there are multiple required star transformations.
+Furthermore, there are a large number of potential star transformations because
+of the large number of stars.
+In addition to requiring a relatively deep search, this search is made much
+more difficult because of the breadth of the search.
+This shows us where our algorithm can hit issues.
+When there are a large number of potential transformations, and there are a
+large number of required transformations, our search gets more complicated,
+and things slow down.
+
+Another place where our synthesis algorithm preforms badly is where there is not
+only a large number of clauses generated, but they also have a non-identity
+lens between them.
+For example, the double capitalization problem is another slowest test, despite
+being an intuitively easy function to figure out, looking at it as a human.
+This is because the algorithm problem has an exponential blow up in terms
+of the number of clauses, when there are many regular expressions +d together.
+If we take this problem to the logical extremes, we see that the the triple
+capitalization problem takes 6.974 seconds, and the quadruple capitalization
+problem takes 33 minutes and 54.730 seconds.
+
+\subsection{Importance of Examples}
+We wanted to see how many examples were required to synthesize the correct
+bijective lens.
+However, because of our implicit understanding of the algorithm, we fear we may
+choose examples that we know will work well with the synthesis algorithm,
+instead of choosing examples that users are likely to.
+Because of that, we test the number of examples that it takes to generate the
+correct lens, with randomly generated examples.
 
 \input{fig_examples-required-table}
-\subsection{Importance of Examples}
+As we see in Figure~\ref{fig:examples-required-table}, examples are not too
+important for synthesizing bijective lenses.
+Because bijective lenses have very few well typed programs, compared to the
+number of functions, oftentimes just a few examples is sufficient.
+As would be expected, the functions that have to make choices about the
+locations of where data goes are those that require more examples.
+In particular, the address problem requires more examples, because it must make
+choices of where information goes, like which name is the a person's first name,
+and which is their last.
 
@@
 \subsection{Importance of User Defined Data Types}
-
-Evaluation~\ref{fig:examples-required-table}~\ref{fig:runtimes-table}~\ref{fig:expanded-time-table}.

### 2016-06-09 paper/fig_regex-syntax.tex
Some small comments while reading

@@
 % REGEX
-(String)& \String{},\StringAlt{} & \GEq{} & $\String\in\StarOf{\Sigma}$ \\
+(Strings)& \String{},\StringAlt{} & \GEq{} & $\String\in\StarOf{\Sigma}$ \\
 (Regexs)& \Regex{},\RegexAlt{} & \GEq{} & s & Base \\
@@
 \end{tabular}
-\caption{Regex Syntax}
+\caption{Regex Syntax\bcp{How about spelling out ``regular expression'', if
+    it doesn't take too much space?}}
 \label{fig:regex-syntax}

### 2016-06-09 paper/introduction.tex
Some small comments while reading

\bcp{May I reiterate my request to put all the text and figures into one big
.tex file?  Makes finding things, global replace, renaming sections, etc. very much easier.  :-)}

\section{Introduction}

### 2016-06-09 paper/evaluation.tex
More comments while reading

@@
 As we see in Figure~\ref{fig:examples-required-table}, examples are not too
-important for synthesizing bijective lenses.
+important for synthesizing bijective lenses.\bcp{If that's really the
+  conclusion, then maybe we should have written a paper about synthesizing
+  bijective lenses {\em without} examples!  But I don't think we can
+  actually conclude this from the examples we have.  }
 Because bijective lenses have very few well typed programs, compared to the

### 2016-06-09 paper/evaluation.tex
One more

@@
 expression equivalences, tests derived from the Augeus test suite, and from
-tests derived from the FlashFill tests.
+tests derived from the FlashFill tests.\bcp{Need to carefully explain our
+  experimental setup and methodology (on what machine config, how many test
+  runs, what kind of statistics, ...)}
 

### 2016-06-10 paper/fig_dnf-regex-syntax.tex

@@
 \end{tabular}
-\caption{DNF Regex Syntax}
+\caption{DNF Regex Syntax\bcp{Why the dot after conjuncts?  Ah, now I see;
+    OK, but maybe it's better to choose two different kinds of brackets
+    instead of distinguishing with . and + (I was also confused by the + --
+    thought it must mean some kind of repetition).  How about
+    $\langle...\rangle$ for disjuncts and $[...]$ for conjuncts?  (This is a
+  bit suggestive of type-theoretic notation.)  Or how about ``hollow square
+  brackets'' for conjuncts (using some kind of square brackets seems good,
+  since they are ordered) and ``hollow set braces'' for disjuncts?  We
+  should try to get this right, since these widgets are our core technical
+  device.}} 
 \label{fig:dnf-regex-syntax}

### 2016-06-16 paper/fig_dnf-regex-syntax.tex
removed proofs appendix, renamed to sequence

@@
 (Atoms)& \Atom{},\AtomAlt{} & \GEq{} & \StarOf{\DNFRegex{}} & Iterate DNF\\
-(Conjunct)& \Conjunct{},\ConjunctAlt{} & \GEq{} &
-$\ConjunctOf{\String_0\ConjunctSep\Atom_1\ConjunctSep\ldots\ConjunctSep\Atom_n\ConjunctSep\String_n}$ & Conjoin\\
-(DNF Regex)& \DNFRegex{},\DNFRegexAlt{} & \GEq{} & $\DNFOf{\Conjunct_1\DNFSep\ldots\DNFSep\Conjunct_n}$ & DNF Or\\
+(Sequence)& \Sequence{},\SequenceAlt{} & \GEq{} &
+$\SequenceOf{\String_0\SequenceSep\Atom_1\SequenceSep\ldots\SequenceSep\Atom_n\SequenceSep\String_n}$ & Conjoin\\
+(DNF Regex)& \DNFRegex{},\DNFRegexAlt{} & \GEq{} & $\DNFOf{\Sequence_1\DNFSep\ldots\DNFSep\Sequence_n}$ & DNF Or\\
 \end{tabular}

### 2016-06-16 paper/fig_dnf-regex-syntax.tex
responded to some feedback

--- 2016-06-06 paper/fig_dnf-regex-syntax.tex
@@
 (Atoms)& \Atom{},\AtomAlt{} & \GEq{} & \StarOf{\DNFRegex{}} & Iterate DNF\\
-(Conjunct)& \Conjunct{},\ConjunctAlt{} & \GEq{} &
-$\ConjunctOf{\String_0\ConjunctSep\Atom_1\ConjunctSep\ldots\ConjunctSep\Atom_n\ConjunctSep\String_n}$ & Conjoin\\
-(DNF Regex)& \DNFRegex{},\DNFRegexAlt{} & \GEq{} & $\DNFOf{\Conjunct_1\DNFSep\ldots\DNFSep\Conjunct_n}$ & DNF Or\\
+(Sequence)& \Sequence{},\SequenceAlt{} & \GEq{} &
+$\SequenceOf{\String_0\SequenceSep\Atom_1\SequenceSep\ldots\SequenceSep\Atom_n\SequenceSep\String_n}$ & Sequence\\
+(DNF Regex)& \DNFRegex{},\DNFRegexAlt{} & \GEq{} & $\DNFOf{\Sequence_1\DNFSep\ldots\DNFSep\Sequence_n}$ & DNF Or\\
 \end{tabular}
-\caption{DNF Regex Syntax}
+\caption{DNF Regex Syntax\bcp{Or how about ``hollow square
+  brackets'' for conjuncts (using some kind of square brackets seems good,
+  since they are ordered) and ``hollow set braces'' for disjuncts?  We
+  should try to get this right, since these widgets are our core technical
+  device.}} 
 \label{fig:dnf-regex-syntax}

### 2016-06-17 paper/evaluation.tex
responded to some feedback
added in proof outline

--- 2016-06-09 paper/evaluation.tex
@@
 tests derived from the FlashFill tests.
+The values presented are the arithmetic means of 10 test runs.
+The tests were run on a 2.5 GHz Intel Core i7 processor, with
+16 GB of 1600 MHz DDR3, running Mac OS X Yosemite.
 
-\input{fig_runtimes-table}
+% fig:runtimes-table
+\begin{figure}
+\centering
+%\begin{tabular}{|l|c|}
+
+\csvreader[tabular=|l|l|,
+    table head=\hline\textbf{Test} & \textbf{Time (s)} \\\hline,
+    late after line = \\\hline]
+{generated-data/data.csv}{Test=\Test,ComputationTime=\ComputationTime}
+{\Test{} & \ComputationTime{}}
+
+%\hline\textbf{Test} & \textbf{Examples Required}
+%\\\hline\hline
+%\csvreader[head to column names]{generated-data/data.csv}{}
+%{\Test& \ExamplesRequired\\\hline}
+%\end{tabular}
+\caption{Runtime Table\bcp{Hard to read, with all the zeros.  Maybe better
+    to display as a (log-scale) graph?  More seriously, we'll need to
+    explain the examples someplace, say why we chose them, maybe group them
+    in some way, ...  Why ws this the right set of tests?}}
+\label{fig:runtimes-table}
+\end{figure}
+
 \subsection{Speed}
@@
 
-\input{fig_examples-required-table}
+% fig:examples-required-table
+\begin{figure}
+\centering
+%\begin{tabular}{|l|c|}
+
+\csvreader[tabular=|l|l|,
+    table head=\hline\textbf{Test} & \textbf{Example Count} \\\hline,
+    late after line = \\\hline]%
+{generated-data/data.csv}{Test=\Test,ExamplesRequired=\ExamplesRequired}%
+{\Test{} & \ExamplesRequired{}}%
+
+%\hline\textbf{Test} & \textbf{Examples Required}
+%\\\hline\hline
+%\csvreader[head to column names]{generated-data/data.csv}{}
+%{\Test& \ExamplesRequired\\\hline}
+%\end{tabular}
+\caption{Examples Required Table\bcp{Is this table showing us that the
+    examples we chose are mostly not very interesting?}}
+\label{fig:examples-required-table}
+\end{figure}
+
 As we see in Figure~\ref{fig:examples-required-table}, examples are not too
@@
 
-\input{fig_expanded-time-table}
+% fig:expanded-time-table
+\begin{figure}
+\centering
+%\begin{tabular}{|l|c|}
+
+\csvreader[tabular=|l|l|,
+    table head=\hline\textbf{Test} & \textbf{Time (s)} \\\hline,
+    late after line = \\\hline]%
+{generated-data/data.csv}{Test=\Test,ForceExpandTime=\ForceExpandTime}%
+{\Test{} & \ForceExpandTime{}}%
+
+%\hline\textbf{Test} & \textbf{Examples Required}
+%\\\hline\hline
+%\csvreader[head to column names]{generated-data/data.csv}{}
+%{\Test& \ExamplesRequired\\\hline}
+%\end{tabular}
+\caption{Expanded Regexp Runtime Table}
+\label{fig:expanded-time-table}
+\end{figure}
 \subsection{Importance of User Defined Data Types}

### 2016-06-22 paper/synthesis-algorithm.tex
add in a more coherent dnfrs

\section{Synthesis Algorithm}

\subsection{Definition}

\begin{lstlisting}[mathescape]
test $t_eu$
\end{lstlisting}

let synthrigidatom (a:atom) (b:atom) : atomlens list =\\
\hspace*{.5em}let\\
\\
let synthrigidsequence (sq:sequence) (tq:sequence) : sequencelens list =\\
\hspace*{.5em}let\\
\\
let synthrigiddnf (dr:dnfregex) (ds:dnfregex) : dnflens list =\\
\hspace*{.5em}if List.length dr <> List.length ds then None else\\
\hspace*{.5em}let permutations = retrieveallpermutations (List.length dr) in\\
\hspace*{.5em}let \\
\hspace*{.5em}let if List.exists\\
\\
let retrieveallexpansions(dr:dnfregex) : dr sequence =\\
\hspace*{.5em}let seq = List.concatmap ~f:(fun e -> applyexpansion e dr) expansions in\\
\hspace*{.5em}seq @ (List.map ~f:retrieveallexpansions seq)\\
\\
let dnfsynth (dr:dnfregex) (ds:dnfregex) (es:examples) : $dnflens_{\bot}$ =\\
\hspace*{.5em}let drqueue = retrieveallexpansion(dr) in\\
\hspace*{.5em}let dlqueue = retrieveallexpansion(dl) in\\
\\
let synth (r:regex) (s:regex) (es:examples) : $lens_{\bot}$ =\\
\hspace*{.5em}let dl=dnfsynth(dnf(r),dnf(s),e)\\
\hspace*{.5em}undnf(dl)\\

\subsection{Correctness}

\begin{theorem}[Correctness of synthrigiddnf]
If $\DNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt$, then
$\DNFLens \in$ (synthrigiddnf \DNFRegex{} \DNFRegexAlt{})
\end{theorem}
\begin{proof}
\end{proof}

\begin{lemma}[Correctness of retrieveallexpansions]
If $\DNFRegex \StarOf{\RewriteDNF} \DNFRegex'$, then
$\DNFRegex'\in$ retrieveallexpansions($\DNFRegex$).
\end{lemma}
\begin{proof}
\end{proof}

\begin{theorem}[Correctness of dnfsynth]
If there exists a dnf lens
$\DNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt \models \Examples$,
then dnfsynth(\DNFRegex{},\DNFRegexAlt{},\Examples{}) will return with some lens
$\DNFLens' \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt \models \Examples'$.
\end{theorem}

\begin{theorem}[Correctness of synth]
If there exists a lens
$\Lens \OfType \Regex \Leftrightarrow \DNFLens \models \Examples$,
then synth(\DNFRegex{},\DNFRegexAlt{},\Examples{}) will return some lens
$\Lens' \OfType \Regex \Leftrightarrow \RegexAlt \models \Examples$
\end{theorem}
\begin{proof}
\end{proof}

### 2016-06-28 paper/abstract.tex
added in abstract draft and contribution list

The language of lenses expresses programs which can convert between source and
target regular languages.
We tackle the task of synthesizing these programs through the regular expressions
of their types.
This is difficult because we cannot use basic type directed synthesis techniques,
as there are many equivalences in the types of these regular expressions.
We do this by developing a language for partially normalized regular expressions,
and searching through those regular expressions for two that have a lens between
them which can be synthesized in a type-directed way.
We develop strategies for efficiently enacting this search and synthesis,
and evaluate our strategies against a suite of microbenchmarks and linux
configuration files.

### 2016-06-28 paper/introduction.tex
added in abstract draft and contribution list

@@
-\bcp{May I reiterate my request to put all the text and figures into one big
-.tex file?  Makes finding things, global replace, renaming sections, etc. very much easier.  :-)}
+\section{Introduction}
 
-\section{Introduction}
+KEY INSIGHTS
+\begin{enumerate}
+\item Search through all possible lenses by: searching for good regular
+expressions, then searching within those regular expressions for type directed
+lenses
+\item Putting the regular expressions in a normal form reduces the number of
+regular expressions we must search through
+\item Using user defined data types allows the synthesis algorithm to keep the
+internals of regular expressions abstracted, unless the internals of those
+regular expressions are changed, and the abstraction must be broken.
+\item Using a pseudometric on regular expressions, based on the
+distribution of user defined data types and size of the regular expression,
+allows for an efficient search through equivalent regular expressions
+\item We use a data structure combining examples with regular expressions to
+synthesize only candidate programs which obey the examples
+\item We develop an efficient strategy for finding if two
+regular expressions have a lens between them requiring no rewrites which amounts
+to merely ordering the components.
+\end{enumerate}

### 2016-06-29 paper/introduction.tex
minor edit

@@
 
-KEY INSIGHTS
 \begin{enumerate}
@@
 \end{enumerate}
+
+In the era of data, big and small, tools for reliably parsing,
+printing, cleaning, and transforming data are increasing important.
+One class of such tools are \emph{lenses}~\cite{?}, bi-directional
+programs that help users solve the classic ``view update problem.''
+More specifically, a lens provides a user with two functions, a
+\emph{get} function and a \emph{put} function.  The \emph{get}
+function translates native data into a new format, or \emph{view}, for
+the user.  If the user updates the new view, the corresponding
+\emph{put} function can translate the edited data back into
+the native format.
+
+One concrete example of where lenses provide value is in managing
+linux configuration files.  Such files come in a wide variety of
+different ``ad hoc'' data formats, and even experienced system
+administrators have a hard time keeping track of all the format
+variations.  Moreover, a missing comma, forgotten keyword or 
+misplaced vertical bar can cause the systems that depend on such
+configurations to crash or misbehave.  In order to manage such configuration
+files more reliably,  RedHat developed Augeas~\cite{}, a configuration
+editing system based around lenses.  At the moment, Augeas contains more
+than 200 stock lenses to provide views of common configuration files.
+
+Lenses are popular because they typically come with strong guarantees
+about the get-put (or put-get) round-trip behavior.  Such guarantees
+help users guard against data corruption while reading and editing
+high-value data sources.

### 2016-06-29 paper/abstract.tex
added in fixes yadda ydadda
abstract

@@
-The language of lenses expresses programs which can convert between source and
-target regular languages.
-We tackle the task of synthesizing these programs through the regular expressions
-of their types.
-This is difficult because we cannot use basic type directed synthesis techniques,
-as there are many equivalences in the types of these regular expressions.
-We do this by developing a language for partially normalized regular expressions,
-and searching through those regular expressions for two that have a lens between
-them which can be synthesized in a type-directed way.
-We develop strategies for efficiently enacting this search and synthesis,
-and evaluate our strategies against a suite of microbenchmarks and linux
-configuration files.
+Tools for reliably converting back and forth between different data
+formats are increasingly important in a world full of big data.
+One class of such tools are \emph{lenses}, programs that simultaneously
+represent one function to convert data from type $A$ to type $B$ and a
+second function to convert data from type $B$ back into data of type
+$A$.  While such programs often come with important invertability properties,
+they can be quite difficult to write in practice 
+as a programmer must think about data conversions in two directions as once.
+To overcome this difficulty, we develop algorithms for synthesizing such lenses 
+automatically from their types and a small number of examples.
+Unlike past work in type-directed synthesis, synthesis of lenses is
+complicated by their rich type structure, which is based on regular
+expressions.  More specifically, rather than simply having to solve
+the problem of finding a representative term to map back and forth
+between types $A$ and $B$, it is first necessary to find the right
+representative types $A'$ and $B'$ amongst many equivalent regular expressions.
+We show how to manage this search by defining a new language of
+lenses governed by semi-canonical regular expression types and prove it
+sound with respect to a standard language of bijective lenses.  We then show
+how to search the reduced space for well-typed lenses that are consistent
+with the given examples.  We demonstrate the effectiveness of our algorithms
+by experimenting on benchmark suite containing microbenchmarks, FlashFill
+examples, and linux configuration file transformations.
+
+%% The language of regular lenses expresses programs which can convert between
+%% source and target regular languages.
+%% We tackle the task of synthesizing these programs through the regular expressions
+%% of their types.
+%% This is difficult because we cannot use basic type directed synthesis techniques,
+%% as there are many equivalences in the types of these regular expressions.
+%% We do this by developing a language for partially normalized regular expressions,
+%% and searching through those regular expressions for two that have a lens between
+%% them which can be synthesized in a type-directed way.
+%% We develop strategies for efficiently enacting this search and synthesis,
+%% and evaluate our strategies against a suite of microbenchmarks and linux
+%% configuration files.

### 2016-06-29 paper/abstract.tex
abs

@@
 second function to convert data from type $B$ back into data of type
-$A$.  While such programs often come with important invertability properties,
-they can be quite difficult to write in practice 
+$A$.  While such programs often come with important invertability guarantees,
+enhancing the reliability of data transformations defined this way,
+they can be quite difficult to write 
 as a programmer must think about data conversions in two directions as once.
@@
 with the given examples.  We demonstrate the effectiveness of our algorithms
-by experimenting on benchmark suite containing microbenchmarks, FlashFill
+by experimenting on a benchmark suite containing microbenchmarks, FlashFill
 examples, and linux configuration file transformations.

### 2016-06-29 paper/abstract.tex

--- 2016-06-29 paper/abstract.tex
@@
-Tools for reliably converting back and forth between different data
-formats are increasingly important in a world full of big data.
-One class of such tools are \emph{lenses}, programs that simultaneously
-represent one function to convert data from type $A$ to type $B$ and a
-second function to convert data from type $B$ back into data of type
-$A$.  While such programs often come with important invertability properties,
-they can be quite difficult to write in practice 
-as a programmer must think about data conversions in two directions as once.
-To overcome this difficulty, we develop algorithms for synthesizing such lenses 
-automatically from their types and a small number of examples.
-Unlike past work in type-directed synthesis, synthesis of lenses is
-complicated by their rich type structure, which is based on regular
-expressions.  More specifically, rather than simply having to solve
-the problem of finding a representative term to map back and forth
-between types $A$ and $B$, it is first necessary to find the right
-representative types $A'$ and $B'$ amongst many equivalent regular expressions.
+The growing dependence on big data makes reliably converting
+between different data formats increasingly important.
+\emph{Lenses} are a bi-directional technique for specifying such
+conversions.  Specifically, a lens is a declarative program that
+simultaneously specifies a function to convert data from type $A$ to
+type $B$ and a second function to convert data from type $B$ to type
+$A$.  Such programs usually provide invertability guarantees, avoiding
+pernicious bugs that can corrupt data and downstream processing.
+Unfortunately, lenses can be quite difficult to write because they
+require programmers to think in two directions at once.  To overcome
+this difficulty, we develop an algorithm to automatically synthesize
+lenses from format descriptions, specified as regular expressions, and
+a small number of examples.  The rich structure of regular expressions
+makes lens synthesis more complicated than previous work on other
+forms of type-directed synthesis.  Specifically, finding a term to
+bi-directionally map between regular expressions $A$ and $B$ requires
+first finding equivalent regular expressions $A'$ and $B'$ that are
+more algorithmically tractable.
 We show how to manage this search by defining a new language of
 lenses governed by semi-canonical regular expression types and prove it
-sound with respect to a standard language of bijective lenses.  We then show
+sound with respect to a standard language of bijective lenses.  
+We then show
 how to search the reduced space for well-typed lenses that are consistent
-with the given examples.  We demonstrate the effectiveness of our algorithms
-by experimenting on benchmark suite containing microbenchmarks, FlashFill
-examples, and linux configuration file transformations.
+with the given examples.  We demonstrate the effectiveness of our algorithm
+by experimenting on a benchmark suite containing XX examples
+comprising microbenchmarks, FlashFill
+examples, and Augeus transformations of Linux configuration files.
+We show we are able to infer lenses for most examples in the
+benchmark suite in an average of XXX seconds.
 

### 2016-06-30 paper/extended-example.tex
added in extended example

\section{Extended Example}

Consider building a lens which converts between between citations given
in EndNote format, and in BibTex format.  For simplicity, we are only
considering the author and title fields.  The synthesis problem for these
two regular expressions is provided below.

\begin{lstlisting}
Name = [A-Z][a-z]*
Initial = [A-Z] "."
Word = ([A-Z] | [a-z] | "-" | "," | ":")+
Title = Word (" " Word)*

LastCommaFirstName = Name", " Name
		(" " (NAME | Initial))*
Type = "@inproceedings" | "@article" | "@book"
BibTex = Type "{" Word ","
	("" | ("title={" Title "},"))
	("" | ("author={" LastCommaFirstName
		(" and " LastCommaFirstName)* "},"))

FullName = Name (" " (Name | Initial))*
	(" " Name)
Class = "Conference Paper" | "Journal Article"
	| "Book"
EndNote = "%0 " Class "\n"
	("%A " FullName "\n")*
	("" | "%T TITLE")
	"%1 " Word "\n"

BibEndConv : BibTex <=> EndNote = ?
\end{lstlisting}

We are trying to synthesize \BibEndConv{}.

### 2016-06-30 paper/evaluation.tex
added in full table figure

@@
 16 GB of 1600 MHz DDR3, running Mac OS X Yosemite.
+
+\begin{figure*}
+\centering
+\begin{tabular}{|c|c|c|c|c|}
+\hline
+\bfseries Test & \bfseries ForceExpandTime & \bfseries ForceExpandExamplesRequired
+& \bfseries Computation Time & \bfseries Examples Required
+\csvreader[head to column names]{generated-data/data.csv}{}
+{\\\hline\Test & \ForceExpandTime & \ForceExpandExamplesRequired & \ComputationTime & \ExamplesRequired}
+\\\hline
+\end{tabular}
+
+\end{figure*}
 

### 2016-07-02 paper/abstract.tex
added in abstract spellings

@@
 type $B$ and a second function to convert data from type $B$ to type
-$A$.  Such programs usually provide invertability guarantees, avoiding
+$A$.  Such programs usually provide invertibility guarantees, avoiding
 pernicious bugs that can corrupt data and downstream processing.
@@
 with the given examples.  We demonstrate the effectiveness of our algorithm
-by experimenting on a benchmark suite containing XX examples
-comprising microbenchmarks, FlashFill
-examples, and Augeus transformations of Linux configuration files.
+by experimenting on a benchmark suite containing 42 examples
+comprising microbenchmarks, Flash Fill
+examples, and Augeas transformations of Linux configuration files.
 We show we are able to infer lenses for most examples in the
-benchmark suite in an average of XXX seconds.
+benchmark suite in an average of 0.134 seconds.
 

### 2016-07-02 paper/introduction.tex
intro draft

@@
 
-In the era of data, big and small, tools for reliably parsing,
-printing, cleaning, and transforming data are increasing important.
-One class of such tools are \emph{lenses}~\cite{?}, bi-directional
+The growing dependence on big data makes tools for reliably parsing
+printing, cleaning, and transforming data increasing important.
+One such class of tools are \emph{lenses}~\cite{?}, bi-directional
 programs that help users solve the classic ``view update problem.''
@@
 
+\dpw{The following paragraph isn't so great. It might be helpful if
+we cited more places lenses had been used in practice and then
+augeas was just one of many examples.}
 One concrete example of where lenses provide value is in managing
@@
 help users guard against data corruption while reading and editing
-high-value data sources.
+high-value data sources.  These guarantees are often achieved through
+the use of domain-specific languages with strong type systems that
+guarantee invertability of lens programs~\cite{?,?,?}.  Unfortunately, such
+lens languages are naturally quite non-standard, and, moreover, the
+process of writing a lens is challenging as a user must think
+about data conversion in two directions at once --- an
+unusual task for the everyday programmer.  As a consequence, 
+for the casual user of lenses, learning the language, its type system and its
+unorthodox programming style is time-consuming and, in the end,
+a disincentive to adopting the technology, despite its advantages.
+
+In order to overcome the steep learning curve involved in programming
+with lenses, we have developed a new system for synthesis of lenses
+from their types and a collection of representative examples.  More
+specifically, we focus on algorithms for the synthesis of bijective
+string lenses.  The types for such lenses are regular expressions,
+which describe the format of the source and target strings.  Given
+such types, and a very small number of representative examples (sometimes
+no examples at all), we synthesize a well-typed lens that implements
+a bijection between data formats.
+
+Recently, researchers have developed a number of new algorithms for
+type-directed synthesis~\cite{?,?,?}.  Many of those
+algorithms~\cite{osera+:pldi15,frankle+:popl16,armando+:pldi16} are
+directed in part by the syntax of types.  For instance, to find an
+expression with type $A \rightarrow B$, such algorithms will assume
+that the expression in question is a function with the form $\lambda
+x{:}A. e$, and they will continue the search for an expression $e$
+with type $B$ under the hypothesis that the parameter $x$ has type
+$A$.  In the case of the simply-typed lambda calculus, such an
+approach reduces the complexity of the search for terms inhabiting the
+type without losing completeness.  Unfortunately, such simple
+syntax-directed algorithms are not directly applicable when searching
+the space of terms defined by richer type systems, in which the types
+exhibit complex isomorphisms, as is the case for string lenses and
+their regular expression types. \dpw{I was trying to generalize a little
+in that last sentence, but it may be awkward.}
+
+\paragraph*{The Central Challenge.}
+To illustrate the central challenge facing the design of our synthesis system
+in greater detail,
+consider the concatenation lens, $\Lens_1 \Concat \Lens_2$,
+one of the simplest string lens combinators.  
+When viewed from left to right, this lens defines a function \emph{get}, which,
+given a string 
+$\String$, this lens divides $\String$ into two substrings $\String_1$ 
+and $\String_2$, applies $\Lens_1$ to the first, yielding $t_1$,
+and $\Lens_2$ to the second, yielding $t_2$.  When viewed from
+right to left, it defines a second function \emph{put}, which, maps
+data back from strings $t$ to strings $s$ in an analogous way.
+
+The type system for lenses ensures such \emph{get} and \emph{put} functions
+satisfy strong invertibility properties.  In our case, the two functions
+form a bijection.  In general, to 
+describe the type of a lens, we write $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$, which may be read ``lens $\Lens$ converts back and forth between
+data in the language of regular expression $\Regex$ and data in the language
+of regular expression $\RegexAlt$.  More specifically, to give a type
+to the concatenation lens, we use the following typing rule.
+\dpw{I think we should delete $\LanguageOf{}$ from all the unambiguity
+constraints in the typing rules, but I didn't do that.  We should be
+careful with consistency, and I didn't know how that would mess with 
+all the proofs and other definitions.}
+%To understand how
+%this lens makes the choice about how to divide up the input string $\String$,
+%it is easiest to look at the typing rule for the lens, which follows.
+\[
+\inferrule
+{
+\Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
+\Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt_2\\\\
+\UnambigConcatOf{\LanguageOf{\Regex_1}}{\LanguageOf{\Regex_2}}\\
+\UnambigConcatOf{\LanguageOf{\RegexAlt_1}}{\LanguageOf{\RegexAlt_2}}
+}
+{
+\ConcatLens{\Lens_1}{\Lens_2} \OfType \Regex_1\Regex_2 \Leftrightarrow \RegexAlt_1\RegexAlt_2
+}
+\]
+Above, the notation \UnambigConcatOf{\LanguageOf{\Regex_1}}{\LanguageOf{\Regex_2}} asserts that $\Regex_1$ and $\Regex_2$ are
+unambiguously concatenable, meaning there is
+exactly one way to divide up any
+string $\String$ in the language of $\Regex_1\Regex_2$ into $\String_1$
+and $\String_2$ such that $\String_1$ is in the language of $\Regex_1$
+and $\String_2$ is in the language of $\Regex_2$.  These unambiguous
+concatenation constraints guarantee the \emph{get} and \emph{put}
+functions are indeed functions and that they form a bijection if
+the underlying lenses are bijective.
+
+One other useful lens to know at this point is the identity lens for a 
+type $\Regex$, written $\IdentityLensOf{\Regex}$.  This lens
+has the type $\Regex \Leftrightarrow \Regex$ and it simply copies
+its input to its output (in either direction).
+
+Now, with these typing rules in hand, consider trying to search for
+a lens with the type $(A B) C \Leftrightarrow A' (B' C)$ where
+we know of lenses with types 
+$A \Leftrightarrow A'$, $B \Leftrightarrow B'$, and 
+$C \Leftrightarrow C'$.  A straightforward,
+syntax-directed search suggests applying the rule for concatenation,
+and looking for two simpler lenses with the types $A B \Leftrightarrow A'$
+and $C \Leftrightarrow B' C'$ respectively.  Unfortunately, at this point,
+we are simply stuck---we have no lenses with those types.  Of course,
+the natural
+solution to this problem
+is to introduce a type conversion rule to the system:
+\[
+\inferrule
+{
+\Lens \OfType \Regex_1 \Leftrightarrow \Regex_2\\
+\Regex_1 \equiv \Regex_1'\\
+\Regex_2 \equiv \Regex_2'
+}
+{
+\Lens \OfType \Regex_1' \Leftrightarrow \Regex_2'
+}
+\]
+With such a type conversion rule, we could first 
+convert the type $(A B) C \Leftrightarrow A' (B' C)$
+to the better aligned $(A B) C \Leftrightarrow (A' B') C$.
+When we then proceed with structure decomposition using the
+concatenation rule, we generate sub-problems involving
+finding lenses for types
+$A \Leftrightarrow A'$, $B \Leftrightarrow B'$, and 
+$C \Leftrightarrow C'$, which is just what we had in hand.
+
+Because equivalence of regular expressions is decideable, when
+given a lens $\Lens$ with type $\Regex_1 \Leftrightarrow \Regex_2$,
+it is easy to \emph{check} whether that lens also has type 
+$\Regex_1' \Leftrightarrow \Regex_2'$, and lens programming languages
+such as Boomerang \cite{foster:thesis} do so.  However, 
+when we turn to synthesis and are given the lens type 
+$\Regex_1' \Leftrightarrow \Regex_2'$ and the option to apply this
+rule, the search space becomes unmanageably large.  In essence,
+we must solve the problem of search \emph{in two dimensions}:  In one
+dimension, we must search for a regular expression \emph{type}
+with an appropriate ``syntactic shape'' and in a second dimension,
+we must search for a lens expression that has the given type and
+is consistent with the user's examples.
+
+\dpw{Note:  I recall there is another aspect to the difficulty of
+doing the search that I'm not mentioning here.  It has to do with
+``horizontal'' composition -- I think we came up with an example of
+lens that you could not write without horizontal composition, which
+also makes the search space bigger.  Ideally, we should expand the
+introduction to illustrate that additional problem.  The fact that you
+need horizontal composition was surprising.  Extending the intro 
+with that material will
+further peak the reader's interest and show them we are doing something
+still harder.}
+
+\paragraph*{Our Solution.}  Our solution to this problem is to define
+a new language of lenses that
+cuts down the synthesis search space \emph{in both dimensions}.  \dpw{Is that
+true?  In \emph{both} dimensions?  Because we eliminate the need to search
+for compositions? But we haven't said that yet.}
+This language is designed so that its regular expression types are limited 
+to those in a pseudo-canonical form.  This pseudo-canonical form
+demands that regular expressions be defined as a flattened 
+union of concatenated atoms, where each atom is a simple string or
+the Kleene star of an expression that is recursively in normal form.
+When converted into this pseudo-canonical form, two regular expressions that 
+differed only by associativity of concatenation will be syntactically equal.
+For example, rather than searching for a lens with type 
+ $(A B) C \Leftrightarrow A' (B' C)$, our algorithm will always search
+for one with type  $(A B) C \Leftrightarrow (A' B') C$
+
+More generally, many, but not all, structural incongruities that
+prevent the existence of lenses disappear.
+Not all structural incogruities disappear because our regular expressions
+are only \emph{pseudo-canonical} (no true canonical form for regular
+expressions exists~\cite{?}).  For instance, our pseudo-normal form
+does not automatically unify lenses such as
+$A B* \Leftrightarrow A (\epsilon + BB*)$ that result from an
+unrolling of Kleene Star, for instance.
+Consequently, we structure our synthesis algorithm in two parts:  The
+first is a syntax-directed search for a lens based on the structure of the given
+regular expression types and, if that fails, the second is a search,
+using Conway's axioms~\cite{conway} for 
+equivalent regular expression types that might allow a new search
+in the first phase of the search to succeed.
+
+One other complication that appears in our setting but not elsewhere in
+the literature on type-directed synthesis, is that the types 
+describing data formats may be very large --- they are regular
+grammars describing complex data formats.  On the one hand, this is
+very useful as such types provide a wealth of information that constrains
+the search process effectively.  On the other hand, there is a lot of
+data here that must be managed carefully by the synthesis system.
+One important observation is that users naturally construct
+such types compositionally, using named abbreviations for subcomponents,
+and the structure inherent in these named abbreviations can be 
+exploited by the synthesis algorithm to speed up the search dramatically.
+
+For example, if the name $A$ appears once on both the left and the right,
+the system will hypothesize that the identity lens can be used to convert
+between occurrences of $A$.  On the other hand, if the name $A$ appears
+on the left, but not on the right, the system will hypothesize that the
+name $A$ should be replaced by its definition.  Likewise, differing
+numbers of occurrences of names can serve as a guide for unrolling
+Kleene star or applying other regular expression equivalence axioms.
+Hence, in general, to guide the transformation of regular expression
+types, we define a metric based on the occurrence of names.
+
+\paragraph*{Contributions.}  Overall, we have designed and implemented
+a new system for the synthesis of bijective string lenses.
+In addition, this paper makes the following contributions:
+
+\begin{itemize}
+\item It defines a new language of lenses designed specifically for synthesis
+with types based on pseudo-canonical regular expressions, with a term 
+language to match.
+\item We formalize the semantics of the language and prove it sound with
+respect to an existing, standard language for bijective string lenses.
+\item We designed a basic, new, type-directed, and two-dimensional 
+algorithm for synthesizing lenses in our new domain-specific language.
+\item We developed a collection of optimizations that speed up the basic
+synthesis algorithm.
+\item We have implemented the language and evaluated it on a range of
+over 40 benchmarks including micro-benchmarks, a few examples from FlashFill,
+and a number of examples from the Augeas system for transforming linux
+file formats.  Our synthesis algorithm succeeds on 36 out of 42 of our
+benchmarks.  When it succeeds, synthesis time is typically less than 1 second.
+\end{itemize}
+
+%% This rule gives the lens $\ConcatLens{\Lens_1}{\Lens_2}$ the type
+%% $\Regex_1\Regex_2 \Leftrightarrow \RegexAlt_1\RegexAlt_2$, meaning it
+%% converts between strings in the language of $\Regex_1\Regex_2$ and
+%% strings in the language of $\RegexAlt_1\RegexAlt_2$.  In order for the
+%% lens to be well-typed,  $\Regex_1$ and $\Regex_2$ must be
+%% unambiguously concatenable (written \UnambigConcatOf{\LanguageOf{\Regex_1}}{\LanguageOf{\Regex_2}}), meaning there must be exactly 1 way to divide up any
+%% string $\String$ in the language of $\Regex_1\Regex_2$ into $\String_1$
+%% and $\String_2$ such that

### 2016-07-03 paper/extended-example.tex
edits to synthesis example

@@
 
+\begin{figure}
 \begin{lstlisting}
@@
 
-LastCommaFirstName = Name", " Name
-		(" " (NAME | Initial))*
-Type = "@inproceedings" | "@article" | "@book"
-BibTex = Type "{" Word ","
+FN      = Name", " Name (" " (Name | Initial))*
+Type    = "@inproceedings" | "@article" 
+BibTex  = 
+  Type "{" 
+    Word ","
     ("" | ("title={" Title "},"))
-	("" | ("author={" LastCommaFirstName
-		(" and " LastCommaFirstName)* "},"))
+    ("" | ("author={" FN (" and " FN)* "},")) 
+  "}"
 
-FullName = Name (" " (Name | Initial))*
-	(" " Name)
+EName   = Name (" " (Name | Initial))* (" " Name)
 Class   = "Conference Paper" | "Journal Article"
-	| "Book"
 EndNote = "%0 " Class "\n"
-	("%A " FullName "\n")*
+	 ("%A " EName "\n")*
 	 ("" | "%T TITLE")
@@
 \end{lstlisting}
+\caption{Synthesis Problem: Simplified Bibtex to EndNote}
+\end{figure}
 

### 2016-07-03 paper/extended-example.tex
added in extended example stuff

--- 2016-06-30 paper/extended-example.tex
@@
 LastCommaFirstName = Name", " Name
-		(" " (NAME | Initial))*
+		(" " (Name | Initial))*
 Type = "@inproceedings" | "@article" | "@book"
@@
 	("%A " FullName "\n")*
-	("" | "%T TITLE")
+	("" | ("%T " Title "\n"))
 	"%1 " Word "\n"
@@
 
+\begin{lstlisting}
+Name = [A-Z][a-z]*
+
+LastCommaFirstName = Name", " Name
+		(" " (Name | Initial))*
+BibTexAuthor = ("" | ("author={" LastCommaFirstName
+		(" and " LastCommaFirstName)* "},"))
+
+FullName = Name (" " (Name | Initial))*
+	(" " Name)
+EndNoteAuthor = ("%A " FullName "\n")*
+
+BibEndAuthorConv : BibTexAuthor <=> EndNoteAuthor = ?
+\end{lstlisting}
+
 We are trying to synthesize \BibEndConv{}.
+
+There is no lens that goes between \BibTex{} and \EndNote{}, so one of them must
+be expanded.
+Indeed, both of them must be expanded, as \BibTex{} will never be able to be
+the source of any lens in this problem, nor will \EndNote{} be able to be the
+target.
+This can be generalized for the sub-expressions of \BibTex{} and \EndNote{},  
+If, any user defined data type does not exist in the possible types of the
+other side, it must be expanded.
+
+FIGURE:
+Trees of user defined regular expressions, with red x's over those that must be
+removed.
+
+After completing these necessary expansions, the problem becomes synthesizing a
+lens between the regular expression
+\begin{lstlisting}
+BibTex'=("@inproceedings" | "@article" | "@book") "{" Word "," ("" | ("title={" Title "},")) ("" | ("author={" (Name", " Name (" " (Name | Initial))*) (" and " (Name", " Name (" " (Name | Initial))*))* "},"))
+\end{lstlisting}
+and
+\begin{lstlisting}
+EndNote'="%0 " ("Conference Paper" | "Journal Article" | "Book") "\n" ("%A " (Name (" " (Name | Initial))* (" " Name)) "\n")* ("" | ("%T " Title "\n")) "%1 " Word "\n"
+\end{lstlisting}
+
+A basic synthesis approach from here will quickly fail.  Both \texttt{BibTex'}
+and \texttt{EndNote'} are of the form $\Regex_1\Concat\Regex_2$, but there is
+not a lens between \texttt{"\%0"} and
+\texttt{("@inproceedings" | "@article" | "@book")}.
+
+These issues can be handled by normalizing the regular expressions slightly.
+By putting them in a form that normalizes away the equivalences of
+associativity, distributivity, and \EmptyString{} identity.
+
+The normalized form of these equivalences is too long to be contained in this
+paper.  Instead, we will work through \afm{or should we just do the easier
+example shown above}.
+
+With those equivalences are handled, perhaps the lens can be synthesized.
+A lens can be synthesized, if there is some pairing of sequences such that
+there are sequence lenses between the two elements of the pair for all pairings.
+A sequence lens can be synthesized if there is some pairing of atoms, such that
+there is an atom lens between the two atoms.  An atom lens can be synthesized
+if the regular expressions the same user defined data type, or if there is a dnf
+lens between the two dnfs under the stars.
+
+FIGURE:
+
+"" | Author Author*
+
+AuthorEndNote*
+
+Looking at the number of sequences, we can easily see that these dnf regular
+expressions do not have a lens between them, without some rewrite rules applied.
+However, there are a large number of possible rewrite rules that can be applied.
+Instead of taking and testing all of them, each rewrite rule is applied, and
+then placed in a priority queue.
+The priority for each problem is a linear combination of the number of rewrites
+applied, and the distance between the two dnf regular expressions by a
+pseudometric.
+
+This pseudometric is defined as the sum of two metrics.
+The first metric is a standard distance metric over the size of the two dnf
+regular expressions.
+The second metric is defined based on the distribution of user defined regular
+expressions.
+We can create an infinite dimensional free module over the integers from the
+regular expressions, generated by the set $\SetOf{(U,i)\SuchThat U\text{is a
+user defined data type}\BooleanAnd i\in\Nats}$.
+We can provide a function $f$ that maps all regular expressions into the element
+which counts how many user defined data types there are at a given depth.
+\afm{i should probably write down this function with a mapsto style, instead of
+just through text}
+The second metric is the taxicab metric on the elements of the vector space
+mapped to by the regular expressions.
+
+FIGURE:
+
+\begin{lstlisting}
+"" | (name,name) (name,name)* -> \{ ((Author,0),2) , ((Author,1),4) \}
+(name,name)* -> \{ ((name,1),2) \}
+
+d("" | (name,name) (name,name)*,(name,name)*) =
+taxicab(\{ ((Author,0),2) , ((Author,1),4) \},other)+(8-2)
+\end{lstlisting}
+
+Empirically, good transformations, like (name,name)* -> (name,name)+
+do very well, giving a distance of zero, where bad transformations, like
+name -> [A-Z][a-z]* do not do as well.
+
+Using the priority queue, the problem that gets popped would be
+"" | (name,name) (name,name)* <=> "" | (name,name) (name,name)*.
+Unlike the previous example, this cannot be so easily determined to not be
+equivalent.
+
+A naive approach to this would be to attempt all possible pairings of sequences,
+and all pairings of atoms within those sequences.  However, the complexity of
+trying all permutations is (\# permutations)!.
+
+Instead, an ordering can be provided to atoms, clauses, and dnf regular
+expressions, where two elements are equal in the ordering if there exists a lens
+between them.  Upon applying that ordering to the elements of the dnf regular
+expression, determining whether there is a lens between two dnf regular
+expressions amounts to merely seeing if each element of the ordering is
+equal with respect to the ordering, as the one on the other side.
+
+FIGURE
+\begin{lstlisting}
+e | (n,n)(n,n)* <=> e | (n,n)(n,n)*
+reorder
+e | (n,n)(n,n)* <=> e | (n,n)(n,n)*
+matched together
+e <=> e , (n,n) <=> (n,n) , etc.
+\end{lstlisting}
+
+This still isn't complete yet, for we don't know if the generated lens will
+satisfy the examples.  The same problem arises with multiple permutations
+being possible.  If all the sequences can be mapped to each other, then there is
+a valid lens for each matching of the sequences!  And not all of these
+valid lenses satisfy the examples.
+To resolve this issue, we embed the parse trees of the examples into the
+regular expressions, and treat regular expressions with different aspects of the
+parse trees embedded as differently in the ordering.
+
+FIGURE
+e | name name(name, name)* , lalalaexample
+
+\begin{lstlisting}
+ /\
+e  |
+   /\
+ name *
+      .
+     /\
+   name name
+
+lalaexample -> or2 (lala . Iterated(ex.am,p.le)
+
+combine to
+
+(the top thing but with the parse tree embedded alongside the expressions)
+\end{lstlisting}

### 2016-07-03 paper/implementation.tex
added in extended example stuff

\section{Implementation}

\subsection{Synthesis Algorithm}
\begin{lstlisting}[mathescape]

let gen_dnf_lens (q:priority_queue) (es:examples)
    : dnf_lens$_\bot$ =
  let (dnf_r,dnf_s,q) = q.pop()
  let dnf_lens_option =
    gen_tds_dnf_lens dnf_r dnf_s es in
  begin match

let gen_lens (r:regex) (s:regex) (es:examples) : lens$_\bot$ =
  let dnf_r = $\ToDNFRegex$(r) in
  let dnf_s = $\ToDNFRegex$(s) in
  let dnf_lens = gen_dnf_lens(dnf_r, dnf_s, es) in
  to_lens(dnf_lens)
\end{lstlisting}

### 2016-07-03 paper/extended-example.tex

@@
 
-LastCommaFirstName = Name", " Name
-		(" " (Name | Initial))*
-Type = "@inproceedings" | "@article" | "@book"
-BibTex = Type "{" Word ","
+FN      = Name", " Name (" " (Name | Initial))*
+Type    = "@inproceedings" | "@article"
+BibTex  = 
+  Type "{" 
+    Word ","
     ("" | ("title={" Title "},"))
-	("" | ("author={" LastCommaFirstName
-		(" and " LastCommaFirstName)* "},"))
+    ("" | ("author={" FN (" and " FN)* "},")) 
+  "}"
 
-FullName = Name (" " (Name | Initial))*
-	(" " Name)
+EName   = Name (" " (Name | Initial))* (" " Name)
 Class   = "Conference Paper" | "Journal Article"
-	| "Book"
 EndNote = "%0 " Class "\n"
-	("%A " FullName "\n")*
-	("" | ("%T " Title "\n"))
+	 ("%A " EName "\n")*
+	 ("" | "%T TITLE")
 	  "%1 " Word "\n"

### 2016-07-03 paper/extended-example.tex
added authors

@@
 
-LastCommaFirstName = Name", " Name
-		(" " (Name | Initial))*
-BibTexAuthor = ("" | ("author={" LastCommaFirstName
-		(" and " LastCommaFirstName)* "},"))
+BNames = Name", " Name (" " (Name))*
+Bib = ("" | ("author={" BNames (" and " BNames)* "},"))
 
-FullName = Name (" " (Name | Initial))*
-	(" " Name)
-EndNoteAuthor = ("%A " FullName "\n")*
+ENames = Name (" " (Name))* (" " Name)
+End = ("%A " ENames "\n")*
 
-BibEndAuthorConv : BibTexAuthor <=> EndNoteAuthor = ?
+BibEnd : Bib <=> End = ?
 \end{lstlisting}

### 2016-07-03 paper/extended-example.tex
minified
added in space in BN def
moved
asdf
added in stuff

@@
 
-BNames = Name", " Name (" " (Name))*
-Bib = ("" | ("author={" BNames (" and " BNames)* "},"))
+BNames = Name ", " Name (" " Name)*
+Bib = ""
+    | "author={" BNames (" and " BNames)* "},"
 
-ENames = Name (" " (Name))* (" " Name)
+ENames = Name (" " (Name))* " " Name
 End = ("%A " ENames "\n")*

### 2016-07-03 generate-data.py
added in %a

@@ -12,8 +12,9 @@ import time
 
 TEST_EXT = '.ls'
 BASELINE_EXT = '.out'
-BASE_FLAGS = []
+BASE_FLAGS = ["-iterativedeepenstrategy"]
 TIMEOUT_TIME = 8
+GENERATE_EXAMPLES_TIMEOUT_TIME = 8
 
 REPETITION_COUNT = 10
 
@@ -27,15 +28,16 @@ def find_tests(root):
         tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
     return tests
 
-def gather_datum(prog, path, base, additional_flags):
-    return EasyProcess([prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)]).call(timeout=TIMEOUT_TIME).stdout
+def gather_datum(prog, path, base, additional_flags, timeout):
+    #print([prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)])
+    return EasyProcess([prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)]).call(timeout=timeout).stdout
 
 def gather_data(prog, path, base):
     current_data = {"Test":join(path, base + TEST_EXT).replace("_","-")[6:]}
     run_data = []
     timeout = False
     for iteration in range(REPETITION_COUNT):
-    	datum = gather_datum(prog, path, base,['-data'])
+    	datum = gather_datum(prog, path, base,['-time'],TIMEOUT_TIME)
 	if datum == "":
 		timeout = True
 		break
@@ -43,16 +45,29 @@ def gather_data(prog, path, base):
 		run_data.append(datum.split(","))
     if timeout:
 	current_data["ComputationTime"]="Timeout"
-	current_data["ExamplesRequired"]="Timeout"
     else:
         run_data_transpose = transpose(run_data)
 	computation_time_col = [float(x) for x in run_data_transpose[0]]
-	examples_required_col = [float(x) for x in run_data_transpose[1]]
 	current_data["ComputationTime"]="{:.5f}".format(sum(computation_time_col)/len(computation_time_col))
-	current_data["ExamplesRequired"]="{:.1f}".format(sum(examples_required_col)/len(examples_required_col))
-    expanded_run_data = []
+    generate_exs_run_data = []
+    timeout = False
     for iteration in range(REPETITION_COUNT):
-    	datum = gather_datum(prog, path, base,['-forceexpand'])
[…]

### 2016-07-03 paper/extended-example.tex
added names
cleaning
added in %a
added in bib and end

@@
 
-BibEnd : Bib <=> End = ?
+BibEnd : Bib <=> End = {
+"author={Conway, John and Kleene, Stephen Cole}"
+<-> "%A John Conway\n%A Stephen Cole Kleene" }
 \end{lstlisting}
 
-We are trying to synthesize \BibEndConv{}.
+We are trying to synthesize \BibEnd{}.
 
@@
 \begin{lstlisting}
-BibTex'=("@inproceedings" | "@article" | "@book") "{" Word "," ("" | ("title={" Title "},")) ("" | ("author={" (Name", " Name (" " (Name | Initial))*) (" and " (Name", " Name (" " (Name | Initial))*))* "},"))
+Bib: ""   | "author={" (" Name ", " Name (" " Name))* (" and " (" Name ", " Name (" " Name)))* "},"
+
+End: ("%A " (Name (" " (Name))* " " Name) "\n")*
 \end{lstlisting}
 and
-\begin{lstlisting}
-EndNote'="%0 " ("Conference Paper" | "Journal Article" | "Book") "\n" ("%A " (Name (" " (Name | Initial))* (" " Name)) "\n")* ("" | ("%T " Title "\n")) "%1 " Word "\n"
-\end{lstlisting}
 

### 2016-07-03 paper/extended-example.tex
add anything left

@@
 Name = [A-Z][a-z]*
+Names = Name (" " Name)*
 
-BNames = Name ", " Name (" " Name)*
+BNames = Name ", " Name Names
 Bib = ""
@@
 
-ENames = Name (" " (Name))* " " Name
+ENames = Name Names " " Name
 End = ("%A " ENames "\n")*
@@
 \begin{lstlisting}
-Bib: ""   | "author={" (" Name ", " Name (" " Name))* (" and " (" Name ", " Name (" " Name)))* "},"
+Bib: ""   | ("author={" Name ", " Name Names (" and " (" Name ", " Name Names))* "},")
 
-End: ("%A " (Name (" " (Name))* " " Name) "\n")*
+End: ("%A " (Name Names " " Name) "\n")*
 \end{lstlisting}

### 2016-07-03 paper/implementation.tex
added in schematic
Added explanation of synthesis overview

@@
 \section{Implementation}
+We have implemented the lens synthesis algorithm in $n$ lines of OCaml code.
+
+\subsection{Synthesis Overview}
+\begin{figure}
+\includegraphics[scale=.5]{synth-lens-schematic.png}
+\label{fig:synth-lens-schematic}
+\end{figure}
+The synthesis algorithm is shown in a schematic diagram in 
+Figure~\ref{fig:synth-lens-schematic}.  The algorithm takes in a pair of
+regular expressions, and examples, and converts the regular expressions into
+DNF regular expressions, using \ToDNFRegex{}.
+These then get enqueued, and SynthDNFLens, where the bulk of the work
+occurs.
+
+In SynthDNFLens, the highest priority DNF regular expression pair is popped.
+Then, SynthRewritelessDNF sees if a DNF lens which satisfies the examples, which
+does not include any applications of the rewrite rule, can be found.
+If none are found, then each rewrite is applied once on each side of the regular
+expressions.
+This new list of DNF regular expression doubles is enqueued, and SynthDNFLens is
+called again.
+If one is found, this DNF lens gets converted to a normal lens, and is returned
+to the user.
 

### 2016-07-03 paper/introduction.tex
some work on section 2

@@
 $\Regex_1' \Leftrightarrow \Regex_2'$ and the option to apply this
-rule, the search space becomes unmanageably large.  In essence,
+rule, the search space becomes unmanageably large --- we must guess
+some arbitrary new lenses $\Regex_1$ and $\Regex_2$ that are
+equivalent to  $\Regex_1'$ and $\Regex_2'$ respectively
+and there are infinitely many such regular expressions to choose
+from.  In essence,
 we must solve the problem of search \emph{in two dimensions}:  In one
@@
 with an appropriate ``syntactic shape'' and in a second dimension,
-we must search for a lens expression that has the given type and
+we must search for a lens \emph{expression} that has the given type and
 is consistent with the user's examples.

### 2016-07-04 paper/implementation.tex
added in changes to implementation and took out stupid section

@@
 
-\subsection{Synthesis Algorithm}
-\begin{lstlisting}[mathescape]
+\subsection{Priority Queue}
+The priority queue is implemented with a list implementation of a priority queue.
+We say that the lower the priority value, the higher the priority.
+The priority value of an element is based on the number of expansions that have been
+preformed, combined with the distance between the two regular expressions, based
+on a psuedometric $\Distance$ we have implemented.  This pseudometric is implemented
+as the sum of simpler pseudometrics,
+$\Distance = \Distance_{size}+\Distance_{dist}$.
 
-let gen_dnf_lens (q:priority_queue) (es:examples)
-    : dnf_lens$_\bot$ =
-  let (dnf_r,dnf_s,q) = q.pop()
-  let dnf_lens_option =
-    gen_tds_dnf_lens dnf_r dnf_s es in
-  begin match
+$\Distance_{size}$ is defined as merely the difference in the sizes of the DNF
+regular expressions.
+$\Distance_{size}(\DNFRegex,\DNFRegexAlt)=\AbsOf{\Size(\DNFRegex)-\Size(\DNFRegexAlt)}$
 
-let gen_lens (r:regex) (s:regex) (es:examples) : lens$_\bot$ =
-  let dnf_r = $\ToDNFRegex$(r) in
-  let dnf_s = $\ToDNFRegex$(s) in
-  let dnf_lens = gen_dnf_lens(dnf_r, dnf_s, es) in
-  to_lens(dnf_lens)
-\end{lstlisting}
+$\Distance_{dist}$ captures how far apart the distribution of
+user defined regular expressions are.
+The distribution of user defined regular expressions can be viewed as an
+infinite dimensional free module over the integers, \Module{}.
+The basis for this is $\SetOf{1*(\RegexVariable,n) | n\in\Nats, \RegexVariable
+\text{ is a user defined regular expression}}$
+DNF regular expressions with variables can map into this with a function by
+counting the number of user defined data types present at a given level.
+The formal definition for this mapping is given by the function \GetDist{}.
+
+\begin{definition}\leavevmode\\
+\label{def:getdist}
+\begin{tabular}{@{}L@{}L@{}}
+\GetDist(\DNFOf{\Sequence_1;\ldots;\Sequence_n}) &
+=\GetDist(\Sequence_1)+\ldots+\GetDist(\Sequence_n)\\
+\GetDist(\SequenceOf{\String_0;\Atom_1;\ldots;\Atom_n;\String_n}) &
+=\GetDist(\Atom_1)+\ldots+\GetDist(\Atom_n)\\
+\GetDist(\RegexVariable)&=1*(\RegexVariable,1)\\
+\GetDist(\IterateLens{\DNFLens})&=\phi(\GetDist(\DNFLens))
+\end{tabular}
+
+where $\phi$ is the linear map sending $1*(\RegexVariable,n)$ to
+$1*(\RegexVariable,n+1)$
+\end{definition}
+
+Using this, we now have a mapping from regular expressions into the distribution
+of user defined regular exprssions.
+
+$\Distance_{dist}(\DNFRegex,\DNFRegexAlt)=\LOneNorm(\phi(\DNFRegex),\phi(\DNFRegexAlt))$
+where \LOneNorm{} is the taxicab metric on \Module{}.
+The priority of a pair of DNF regular expressions, $(\DNFRegex{},\DNFRegexAlt{})$,
+is $\Distance(\DNFRegex{},\DNFRegexAlt{})+8*expansion_count$.
+Keeping track of expansion count is important as it allows for lower priority
+expansions to be eventually popped, instead of getting stuck doing incorrect,
+increasingly deep expansions whose DNF regular expressions have a low distance.
+\afm{should i do an example, or is that going into too much detail?}
+Experimentally, 8 was a good value for how much weight the expansion
+count should get in the priority.  It allows for not getting stuck in a local
+minima in the search space, while still allowing for the distance to choose the
+correct expansion the majority of the time.
+
+\subsection{SynthRWLessDNF}
+We find an efficient way to synthesize DNF lenses which do not use rewrites.

### 2016-07-13 paper/extended-example.tex

@@
-\section{Extended Example}
+\section{Lens Synthesis by Example}
 
-Consider building a lens which converts between between citations given
-in EndNote format, and in BibTex format.  For simplicity, we are only
-considering the author and title fields.  The synthesis problem for these
-two regular expressions is provided below.
+\dpw{the lstlisting environment is drawing a line underneath the code.
+I don't want that.  we need to fix it but I'm focusing on writing.
+Also, it may be good to hack it so it can be in a backslash small font.
+There is perhaps another environment we could use -- a code environment.}
+
+In order to explain the major components of our synthesis algorithm,
+we will consider a running example involving conversion between citations given
+in highly simplified BibTex and EndNote formats.  Our BibTex format
+consists entirely of specifying a list of authors on a single line as follows.
+\begin{lstlisting}
+author={Conway, John and Kleene, Stephen Cole}
+\end{lstlisting}
+Our EndNote format contains the same information, but uses different separators,
+rearranges the order of first and last names, and uses a separate line for each
+author. 
+\begin{lstlisting}
+%A John Conway
+%A Stephen Cole Kleene
+\end{lstlisting}
 
@@
 Name = [A-Z][a-z]*
-Initial = [A-Z] "."
-Word    = ([A-Z] | [a-z] | "-" | "," | ":")+
-Title   = Word (" " Word)*
-
-FN      = Name", " Name (" " (Name | Initial))*
-Type    = "@inproceedings" | "@article"
-BibTex  = 
-  Type "{" 
-    Word ","
-    ("" | ("title={" Title "},"))
-    ("" | ("author={" FN (" and " FN)* "},")) 
-  "}"
-
-EName   = Name (" " (Name | Initial))* (" " Name)
-Class   = "Conference Paper" | "Journal Article"
-EndNote = "%0 " Class "\n"
-	 ("%A " EName "\n")*
-	 ("" | "%T TITLE")
-	  "%1 " Word "\n"
-
-BibEndConv : BibTex <=> EndNote = ?
-\end{lstlisting}
-\caption{Synthesis Problem: Simplified Bibtex to EndNote}
-\end{figure}
-
-\begin{lstlisting}
-Name = [A-Z][a-z]*
-Names = Name (" " Name)*
+Names = (" " Name)*
 
@@
 \end{lstlisting}
+\caption{Synthesis Problem Definition: Idealized Bibtex to EndNote}
+\label{fig:bibend-spec}
+\end{figure}
+
+Figure~\cite{fig:bibend-spec} presents
+the user's specification of such a problem.  In general, such specifications are
+given by defining the source and target data formats~\footnote{Lenses
+operate both left-to-right and right-to-left, so both formats are sometimes
+sources and sometimes targets. Nevertheless, we will often call the
+format on the left, the ``source,'' and the format on the right, 
+the ``target.''} as well as some examples.  In this case, the lens
+we wish to synthesize is \CF{BibEnd}.  It converts between 
+the formats \CF{Bib} and \CF{End}.
+Such formats are defined using a standard syntax for regular expressions.
+Intermediate regular expressions are named in order to facilitate
+construction of complex formats.
+
+\subsection{Regular Bijective Lenses}
+
+Once we have a synthesis specification, we need to find a lens that
+will convert back and forth between formats and that satisfies the
+given examples.  In this paper, we begin our work with a simple and 
+relatively standard language of bijective string lens combinators.
+In order to guarantee the functions generated by these combinators
+form proper bijections between source and target, they are given types using regular expressions.  
+We typicallly write 
+$\Lens \OfType \Regex \Leftrightarrow \RegexAlt$ to indicate that
+a lens $\Lens$ converts back and forth between data belonging
+to the languages of regular expressions
+$\Regex$ and $\RegexAlt$.
+
+The simplest lens in the combinator language is the identity lens 
+at type $R$,
+$\IdentityLensOf{R}$, which copies data matching regular expression $R$
+from source to target, and vice versa.  The
+constant lens $\ConstLens{\String_1}{\String_2}$, when operated left-to-right, 
+replaces string $\String_1$ with $\String_2$, and when operated right-to-left,
+replaces string $\String_2$ with $\String_1$.
+
+Each of the other lenses manipulate structured data.  For instance,
+$\ConcatLens{\Lens_1}{\Lens_2}$
+
+\paragraph*{Step 1:  Inlining Definitions.}
+
+\dpw{----------------}
 

### 2016-07-27 paper/implementation.tex
refactor, add in lens context and puts

@@
 
-\subsection{SynthRWLessDNF}
-We find an efficient way to synthesize DNF lenses which do not use rewrites.
+\subsection{SynthRewritelessDNF}
+We find an efficient way to synthesize DNF lenses which do not use rewrites
+through a clever encoding of the DNF regular expressions and examples.
+A key insight into efficiently finding DNF lenses between two DNF regular
+expressions is finding an ordering of DNF regular expressions which satisfy the
+property such that, for two regular expressions $\DNFRegex$ and $\DNFRegexAlt$,
+$\DNFRegex \leq \DNFRegexAlt$ and $\DNFRegexAlt \leq \DNFRegex$ if, and only if,
+for some dnf lens $\DNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt$,
+with no use of \DNFRewriteLensRule{}.
+
+\begin{definition}
+$\DNFOf{\Sequence_1;\ldots\Sequence_n} \leq
+\DNFOf{\SequenceAlt_1;\ldots\SequenceAlt_m}$
+if $n<m$ or $n=m$ and
+$sort([\Sequence_1;\ldots;\Sequence_n])\leq
+sort([\SequenceAlt_1;\ldots;\SequenceAlt_n])$ by the dictionary ordering
+on the sequence ordering.
+
+$\SequenceOf{\String_0;\Atom_1;\ldots;\Atom_n;\String_n} \leq
+\SequenceOf{\StringAlt_0;\AtomAlt_1;\ldots;\AtomAlt_n;\String_n}$
+if $n<m$ or $n=m$ and
+$sort([\Atom_1\ldots\Atom_n])\leq
+sort([\AtomAlt_1\ldots\AtomAlt_n])$ by the dictionary ordering on the
+atom ordering.
+
+$\RegexVariable < \IterateLens{\DNFLens}$
+
+$\RegexVariable \leq \RegexVariableAlt$ if
+$\RegexVariable \leq \RegexVariableAlt$ as strings.
+
+$\StarOf{\DNFRegex} \leq \StarOf{\DNFRegexAlt}$ if $\DNFRegex \leq \DNFRegexAlt$.
+\end{definition}
+
+\begin{theorem}
+$\DNFRegex\leq\DNFRegexAlt$ and $\DNFRegexAlt\leq\DNFRegex$ if, and only
+if there exists a $\DNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt$.
+\end{theorem}
+
+Beyond that, we can use these orderings to find the lenses.
+If $\DNFOf{\Sequence_1;\ldots;\Sequence_n}\leq\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}$,
+and $\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}\leq\DNFOf{\Sequence_1;\ldots;\Sequence_n}$,
+then $\Sort([\Sequence_1;\ldots;\Sequence_n])\leq\Sort([\Sequence_1;\ldots;\Sequence_n])$
+using the dictionary ordering, which means that
+$\Sort([\Sequence_1;\ldots;\Sequence_n])[i]\leq
+\Sort([\SequenceAlt_1;\ldots;\SequenceAlt_n])[i]$ and
+$\Sort([\SequenceAlt_1;\ldots;\SequenceAlt_n])[i]\leq
+\Sort([\Sequence_1;\ldots;\Sequence_n])[i]$, which means that there is a
+sequence lens between them.
+\afm{I can keep going on about this stuff, should it go here?  Should I just make a brief claim
+about it and move on}
+
+Unfortunately, while this can find a well typed DNF regular expression, it doesn't
+necessarily find the one that matches the examples.  In the situation where
+there are multiple valid sorted orderings, there are multiple different lenses
+with potentially different semantics.  Instead of iterating through all the
+possible sortings, and finding one that matches the examples, we would like to
+be able to immediately find only a sorting which satisfies the examples.
+
+The key insight here is that certain invariants must hold for a DNF lens to have
+semantics which match the examples.
+
+\begin{enumerate}
+\item If a source example string matches one of the clauses during parsing,
+it must be sent to the clause the target example string matches during parsing.
+\item If a source example string matches a user defined regular expression during
+parsing, the target must match the exact same string during parsing.
+\item If a source example string matches an iteration, then the target must iterate
+the same number of times, and all the invariants must hold for each iterated part.
+\end{enumerate}
+
+This is done by joining additional parsing information alongside the regular
+expression components.

### 2016-07-29 generate-data.py
refactor and allow for using lens context stuff

@@ -30,20 +30,27 @@ def find_tests(root):
 
 def gather_datum(prog, path, base, additional_flags, timeout):
     #print([prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)])
-    return EasyProcess([prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)]).call(timeout=timeout).stdout
+    process_output = EasyProcess([prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)]).call(timeout=timeout)
+    return (process_output.stdout,process_output.stderr)
 
 def gather_data(prog, path, base):
     current_data = {"Test":join(path, base + TEST_EXT).replace("_","-")[6:]}
     run_data = []
     timeout = False
+    error = False
     for iteration in range(REPETITION_COUNT):
-    	datum = gather_datum(prog, path, base,['-time'],TIMEOUT_TIME)
+    	(datum,err) = gather_datum(prog, path, base,['-time'],TIMEOUT_TIME)
 	if datum == "":
+            if err == "":
 		timeout = True
-		break
+            else:
+                error = True
+	    break
 	else:
-		run_data.append(datum.split(","))
-    if timeout:
+	    run_data.append(datum.split(","))
+    if error:
+        current_data["ComputationTime"]="Error"
+    elif timeout:
 	current_data["ComputationTime"]="Timeout"
     else:
         run_data_transpose = transpose(run_data)
@@ -52,13 +59,18 @@ def gather_data(prog, path, base):
     generate_exs_run_data = []
     timeout = False
     for iteration in range(REPETITION_COUNT):
-    	datum = gather_datum(prog, path, base,['-generatedexamples'],GENERATE_EXAMPLES_TIMEOUT_TIME)
+    	(datum,err) = gather_datum(prog, path, base,['-generatedexamples'],GENERATE_EXAMPLES_TIMEOUT_TIME)
 	if datum == "":
+            if err == "":
 		timeout = True
-		break
+            else:
+                error = True
+	    break
 	else:
-		generate_exs_run_data.append(datum.split(","))
-    if timeout:
+	    generate_exs_run_data.append(datum.split(","))
+    if error:
+        current_data["ExamplesRequired"]="Error"
+    elif timeout:
 	current_data["ExamplesRequired"]="Timeout"
     else:
         generate_exs_run_data_transpose = transpose(generate_exs_run_data)
[…]

### 2016-09-12 paper/conclusion.tex
added variable information on master file to each file

--- 2016-06-29 paper/abstract.tex
@@
-Tools for reliably converting back and forth between different data
-formats are increasingly important in a world full of big data.
-One class of such tools are \emph{lenses}, programs that simultaneously
-represent one function to convert data from type $A$ to type $B$ and a
-second function to convert data from type $B$ back into data of type
-$A$.  While such programs often come with important invertability guarantees,
-enhancing the reliability of data transformations defined this way,
-they can be quite difficult to write 
-as a programmer must think about data conversions in two directions as once.
-To overcome this difficulty, we develop algorithms for synthesizing such lenses 
-automatically from their types and a small number of examples.
-Unlike past work in type-directed synthesis, synthesis of lenses is
-complicated by their rich type structure, which is based on regular
-expressions.  More specifically, rather than simply having to solve
-the problem of finding a representative term to map back and forth
-between types $A$ and $B$, it is first necessary to find the right
-representative types $A'$ and $B'$ amongst many equivalent regular expressions.
-We show how to manage this search by defining a new language of
-lenses governed by semi-canonical regular expression types and prove it
-sound with respect to a standard language of bijective lenses.  We then show
-how to search the reduced space for well-typed lenses that are consistent
-with the given examples.  We demonstrate the effectiveness of our algorithms
-by experimenting on a benchmark suite containing microbenchmarks, FlashFill
-examples, and linux configuration file transformations.
+\section{Conclusion}
 
-%% The language of regular lenses expresses programs which can convert between
-%% source and target regular languages.
-%% We tackle the task of synthesizing these programs through the regular expressions
-%% of their types.
-%% This is difficult because we cannot use basic type directed synthesis techniques,
-%% as there are many equivalences in the types of these regular expressions.
-%% We do this by developing a language for partially normalized regular expressions,
-%% and searching through those regular expressions for two that have a lens between
-%% them which can be synthesized in a type-directed way.
-%% We develop strategies for efficiently enacting this search and synthesis,
-%% and evaluate our strategies against a suite of microbenchmarks and linux
-%% configuration files.
+Conclusion
+
+%%% Local Variables:
+%%% TeX-master: "main"
+%%% End:

### 2016-09-12 paper/implementation.tex
added variable information on master file to each file

@@
 The priority of a pair of DNF regular expressions, $(\DNFRegex{},\DNFRegexAlt{})$,
-is $\Distance(\DNFRegex{},\DNFRegexAlt{})+8*expansion_count$.
+is $\Distance(\DNFRegex{},\DNFRegexAlt{})+8*expansion_{count}$.
 Keeping track of expansion count is important as it allows for lower priority
@@
 
+ATTEMPT TO WRITE IT LESS FORMALLY
+
+Unfortunately, while synthesizing a DNF regular expression, we must synthesize
+two permutations, one to deal with determining which sequence
+gets mapped to which (handling the commutativity of +), and one to deal
+with determining within a sequence, which atoms get mapped to which
+(handling swaps).
+
+We found an incredibly efficient strategy for doing this, by further normalizing
+the DNF regular expressions.  Instead of allowing any order of sequences, we
+provide a total preorder on sequences, and order order the sequences
+according to that ordering.
+
+We do the same trick for normalizing the sequences.  While reordering a sequence
+changes the semantics, we can temporarily reorder the sequence, remembering the
+permutation of the reordering, and recover the original sequence later.
+
 \begin{definition}
-$\DNFOf{\Sequence_1;\ldots\Sequence_n} \leq
-\DNFOf{\SequenceAlt_1;\ldots\SequenceAlt_m}$
-if $n<m$ or $n=m$ and
-$sort([\Sequence_1;\ldots;\Sequence_n])\leq
-sort([\SequenceAlt_1;\ldots;\SequenceAlt_n])$ by the dictionary ordering
-on the sequence ordering.
-
-$\SequenceOf{\String_0;\Atom_1;\ldots;\Atom_n;\String_n} \leq
-\SequenceOf{\StringAlt_0;\AtomAlt_1;\ldots;\AtomAlt_n;\String_n}$
-if $n<m$ or $n=m$ and
-$sort([\Atom_1\ldots\Atom_n])\leq
-sort([\AtomAlt_1\ldots\AtomAlt_n])$ by the dictionary ordering on the
-atom ordering.
-
-$\RegexVariable < \IterateLens{\DNFLens}$
-
-$\RegexVariable \leq \RegexVariableAlt$ if
-$\RegexVariable \leq \RegexVariableAlt$ as strings.
-
-$\StarOf{\DNFRegex} \leq \StarOf{\DNFRegexAlt}$ if $\DNFRegex \leq \DNFRegexAlt$.
+Given an ordering $\leq : A \rightarrow A \rightarrow \SetOf{-1,0,1}$,
+we denote the dictionary ordering on lists of $As$, $\ListOf{\leq}$.
 \end{definition}
 
+\begin{definition}
+  Given an ordering $\leq : A \rightarrow A \rightarrow \SetOf{-1,0,1}$,
+  and $\ListOf{a_1;\ldots;a_n}\in A \ListType$
+  We define $\SortingOf{\leq}{\ListOf{a_1;\ldots;a_n}}$ as
+  $\sigma \in \PermutationSetOf{n}$ where if $i\leq j$,
+  then $a_{\sigma(i)} \leq a_{\sigma(j)}$.
+
+  We define $\SortOf{\leq}{\ListOf{a_1;\ldots;a_n}}$ as
+  $\ListOf{a_{\sigma(1)};\ldots;a_{\sigma(n)}}$ where
+  $\sigma=\SortingOf{\leq}{\ListOf{a_1;\ldots;a_n}}$
+\end{definition}
+
+\begin{definition}
+  \begin{enumerate}
+  \item
+    $\DNFOf{\Sequence_1;\ldots\Sequence_n}
+    \DNFLeq
+    \DNFOf{\SequenceAlt_1;\ldots\SequenceAlt_m}$
+    if
+    $\SortOf{\SequenceLeq}{\ListOf{\Sequence_1;\ldots;\Sequence_n}}
+    \ListOf{\SequenceLeq}
+    \SortOf{\SequenceLeq}{[\SequenceAlt_1;\ldots;\SequenceAlt_n]}$.
+
+  \item
+    $\SequenceOf{\String_0;\Atom_1;\ldots;\Atom_n;\String_n}
+    \SequenceLeq
+    \SequenceOf{\StringAlt_0;\AtomAlt_1;\ldots;\AtomAlt_n;\String_n}$
+    if
+    $\SortOf{\AtomLeq}{\ListOf{\Atom_1\ldots\Atom_n}}
+    \ListOf{\AtomLeq}
+    \SortOf{\AtomLeq}{[\AtomAlt_1\ldots\AtomAlt_n]}$.
+
+  \item
+    $\StarOf{\DNFRegex} \AtomLeq \StarOf{\DNFRegexAlt}$
+    if
+    $\DNFRegex \DNFLeq \DNFRegexAlt$
+  \end{enumerate}
+\end{definition}
+
+This allows us to put our DNF regular expressions into a more normal form.
+
+\begin{definition}
+  \begin{enumerate}
+  \item A DNF regular expression $\DNFOf{Sequence_1;\ldots;\Sequence_n}$ is
+    \textit{normalized} if $i \leq j$ implies $\Sequence_1\SequenceLeq\Sequence_j$,
+    and each $\Sequence_i$ is normalized.
+  \item A sequence $\SequenceOf{\Atom_1;\ldots;\Atom_n}$ is \textit{normalized}
+    if $i \leq j$ implies $\Atom_i\AtomLeq\Atom_j$,
+    and each $\Atom_i$ is normalized.
+  \item An atom $\StarOf{\DNFRegex}$ is \textit{normalized} if
+    $\DNFRegex$ is normalized.
+  \end{enumerate}
+\end{definition}
+
+We can also easily normalize a DNF regular expression
+
+\begin{definition}
+  \begin{enumerate}
+  \item $\NormalizeOf{\DNFOf{\Sequence_1;\ldots;\Sequence_n}} =
+    \DNFOf{\NormalizeOf{\Sequence_{\sigma(1)}};\ldots;\NormalizeOf{\Sequence_{\sigma(n)}}}$,
+    where
+    $\sigma=\SortingOf{\SequenceLeq}
+    {\ListOf{\NormalizeOf{\Sequence_1};\ldots;\NormalizeOf{\Sequence_n}}}$
+  \item $\NormalizeOf{\SequenceOf{\String_0;\Atom_1;\ldots;\Atom_n;\String_n}} =
+    \SequenceOf{\String_0;\Atom_{\sigma(1)};\ldots;\Atom_{\sigma(n)};\String_n}$,
+    where
+    $\sigma=\SortingOf{\SequenceLeq}
+    {\ListOf{\NormalizeOf{\Atom_1};\ldots;\NormalizeOf{\Atom_n}}}$
+  \item $\NormalizeOf{\StarOf{\DNFRegex}}=\StarOf{\NormalizeOf{\DNFRegex}}$
+  \end{enumerate}
+\end{definition}
+
+Critically, the normalized form of the DNF regular expression, and the ordering
+on DNF regular expressions, are tied to the existance of a DNF Lens without
+rewrites.
+
 \begin{theorem}
-$\DNFRegex\leq\DNFRegexAlt$ and $\DNFRegexAlt\leq\DNFRegex$ if, and only
-if there exists a $\DNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt$.
+  $\NormalizeOf{\DNFRegex}\DNFLeq\NormalizeOf{\DNFRegexAlt}$ and
+  $\NormalizeOf{\DNFRegexAlt}\DNFLeq\NormalizeOf{\DNFRegex}$ if, and only
+  if, there exists a $\DNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt$,
+  where $\DNFLens$ has no applications of a rewrite rule.
 \end{theorem}
+
+This can be proven with mutual induction, proving similar statements for
+sequences and atoms.
+
+Furthermore, two DNF regular expressions can be zipped together to form a
+DNF lens.  Consider $\DNFOf{\Sequence_{\sigma_1(0)};\ldots;\Sequence_{\sigma_1(n)}}$, and
+$\DNFOf{\SequenceAlt_{\sigma_2(0)};\ldots;\SequenceAlt_{\sigma_2(n)}}$, two normalized
+DNF regular expressions.  $\Sequence_{\sigma(0)}$ merely needs to be mapped to
+$\SequenceAlt_{\tau(0)}$.  Furthermore, the permutation mapping the two lenses
+merely needs to be $\sigma_1 \Compose \sigma_2^{-1}$
+
+We can use these orderings to find
 
@@
 and $\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}\leq\DNFOf{\Sequence_1;\ldots;\Sequence_n}$,
-then $\Sort([\Sequence_1;\ldots;\Sequence_n])\leq\Sort([\Sequence_1;\ldots;\Sequence_n])$
+then $\Sorting([\Sequence_1;\ldots;\Sequence_n])\leq\Sorting([\Sequence_1;\ldots;\Sequence_n])$
 using the dictionary ordering, which means that
-$\Sort([\Sequence_1;\ldots;\Sequence_n])[i]\leq
-\Sort([\SequenceAlt_1;\ldots;\SequenceAlt_n])[i]$ and
-$\Sort([\SequenceAlt_1;\ldots;\SequenceAlt_n])[i]\leq
-\Sort([\Sequence_1;\ldots;\Sequence_n])[i]$, which means that there is a
+$\Sorting([\Sequence_1;\ldots;\Sequence_n])[i]\leq
+\Sorting([\SequenceAlt_1;\ldots;\SequenceAlt_n])[i]$ and
+$\Sorting([\SequenceAlt_1;\ldots;\SequenceAlt_n])[i]\leq
+\Sorting([\Sequence_1;\ldots;\Sequence_n])[i]$, which means that there is a
 sequence lens between them.
@@
 expression components.
+
+%%% Local Variables:
+%%% TeX-master: "main"
+%%% End:

### 2016-09-16 paper/conclusion.tex
added rough draft of all sections

@@
 
-Conclusion
+We have coded up an implementation of type directed synthesis on a domain
+specific language of bijective lenses.  This program allows users to input two
+regular expressions, and a few input/output examples, and synthesizes a
+bijective lens between those regular expressions, matching those input/output
+examples.  While lenses are very useful for their guarantees, they are a
+difficult domain specific language to code in, in practice.  This program
+provides an easier interaction model for the creation of lenses through using
+synthesis as the way to interact with the programmer.
+
+This synthesis algorithm is competitive with state of the art tools for
+synthesis on strings transformations.  This gain is made through requiring a
+larger amount of user interaction with the synthesis algorithm.  However, in the
+for this increased interaction, the program provides the capabilities of
+synthesizing very complicated transformations, and provides stronger guarantees
+about correctness on inputs dissimilar to those provided in the examples.
 

### 2016-09-16 paper/discussion.tex
added variable information on master file to each file
added rough draft of all sections

--- 2016-09-12 paper/conclusion.tex
@@
-\section{Conclusion}
+\section{Discussion}
 
-Conclusion
+We found that in practice, requiring a user to put in more information for the
+synthesis of a string transformation creates a more reliable algorithm.
+However, this comes at a few costs.  The user needs to know how to write out
+regular expressions.
+
+The largest difficulty in this is the restriction to bijective lenses.
+Oftentimes, we find the need to hide information away in some alternate form.
+For example, there is oftentimes whitespace in the definition of a regular
+expression.  This whitespace must be copied somehow over to the other side.
+Oftentimes this requires the use of having a portion of the string dedicated to
+holding the whitespace information of the other side.
+
+Another annoyance is the fact that if the correct lens does not exist,
+rarely does the tool terminate.  Much of the time it will just spin trying to
+perform increasingly complicated transformations to find the correct lens.
+A better solution would be to find something close to the correct lens, and have
+a user interaction model that supports iterative synthesis.  The user could give
+more examples which fix where the program is wrong.
+
+\subsection{Future Work}
+Some of the issues seen in this have already been observed.  For example, in the
+quotient lens paper, the issue of having to hack a place in the source string to
+move whitespace information from the view to the source was noted.  An extension
+we would like to work on would be able to synthesize quotient lenses.
+
+Another approach to having disparate information on each side of the lens is
+through the use of symmetric lenses.  This could be another approach to the
+whitespace problem.
 

### 2016-09-16 paper/evaluation.tex
added variable information on master file to each file
added rough draft of all sections

--- 2016-06-09 paper/evaluation.tex
@@
 expression equivalences, tests derived from the Augeus test suite, and from
-tests derived from the FlashFill tests.\bcp{Need to carefully explain our
-  experimental setup and methodology (on what machine config, how many test
-  runs, what kind of statistics, ...)}
+tests derived from the FlashFill tests.
+The values presented are the arithmetic means of 10 test runs.
+The tests were run on a 2.5 GHz Intel Core i7 processor, with
+16 GB of 1600 MHz DDR3, running Mac OS X Yosemite.
 
-\input{fig_runtimes-table}
+\begin{figure*}
+\centering
+\begin{tabular}{|c|c|c|c|c|}
+\hline
+\bfseries Test & \bfseries ForceExpandTime & \bfseries ForceExpandExamplesRequired
+& \bfseries Computation Time & \bfseries Examples Required
+\csvreader[head to column names]{generated-data/data.csv}{}
+{\\\hline\Test & \ForceExpandTime & \ForceExpandExamplesRequired & \ComputationTime & \ExamplesRequired}
+\\\hline
+\end{tabular}
+
+\end{figure*}
+
 \subsection{Speed}
@@
 
-\input{fig_examples-required-table}
 As we see in Figure~\ref{fig:examples-required-table}, examples are not too
@@
   conclusion, then maybe we should have written a paper about synthesizing
-  bijective lenses {\em without} examples!  But I don't think we can
+  bijective lenses {\em without\/} examples!  But I don't think we can
   actually conclude this from the examples we have.  }
@@
 
-\input{fig_expanded-time-table}
 \subsection{Importance of User Defined Data Types}
+
+User defined data types are very critical to the algorithm and performance of
+the synthesis algorithm.  User defined data types help direct the search of
+which expansions should be performed.  Furthermore, they help to reduce the
+number of clauses we have to deal with.  There is a potentially exponentially
+large blowup in converting regular expressions into DNF form, but keeping
+certain portions of the regular expressions abstracted lessens that blowup,
+as those regular expressions are kept atomic in converting into DNF form.
+
+User defined data types are also critical in the applications of previously
+defined lenses.  We restrict the application of previously defined lenses to
+only on user defined data types in the synthesis algorithm.  We then group the
+user defined data types based on which have previously defined lenses between
+them.  This allows for efficient recognition of when two user defined regular
+expressions can be mapped to each other.  Removing the ability to use previously
+defined lenses makes certain problems intractable, like (example1, example2,
+example3).
+
+While we can simulate the use of user defined data types automatically, by
+giving regular expressions which have the same syntax the same user defined data
+type, manually inputting them performs better in practice.  This is because the
+procedure for automatically creating user defined data types creates far more
+data types, which then creates a more complicated search problem.
+
+%%% Local Variables:
+%%% TeX-master: "main"
+%%% End:

### 2016-09-16 paper/extended-example.tex
added variable information on master file to each file
added rough draft of all sections

@@
 
-\dpw{the lstlisting environment is drawing a line underneath the code.
-I don't want that.  we need to fix it but I'm focusing on writing.
-Also, it may be good to hack it so it can be in a backslash small font.
-There is perhaps another environment we could use -- a code environment.}
+\dpw{It may be good to hack lstlisting so it can be in a backslash small font.
+There is perhaps another environment we could use --- a code environment.}
 
@@
 
-Figure~\cite{fig:bibend-spec} presents
+Figure~\ref{fig:bibend-spec} presents
 the user's specification of such a problem.  In general, such specifications are
@@
 from source to target, and vice versa.  The
-constant lens $\ConstLens{\String_1}{\String_2}$, when operated left-to-right, 
+constant lens $\ConstLensOf{\String_1}{\String_2}$, when operated left-to-right, 
 replaces string $\String_1$ with $\String_2$, and when operated right-to-left,
@@
 Each of the other lenses manipulate structured data.  For instance,
-$\ConcatLens{\Lens_1}{\Lens_2}$
+$\ConcatLensOf{\Lens_1}{\Lens_2}$
 
@@
 If, any user defined data type does not exist in the possible types of the
-other side, it must be expanded.
+other side, it must be expanded, shown in Figure~\ref{fig:expanded-defs}.
 
-FIGURE:
-Trees of user defined regular expressions, with red x's over those that must be
-removed.
+\begin{figure}
+  \Tree[.\textcolor{dkred}{\sout{\texttt{Bib}}}
+    [.\textcolor{dkred}{\sout{\texttt{BNames}}}
+      \texttt{Name}
+      [.\texttt{Names}
+        \texttt{Name} ]]]
+  \Tree[.\textcolor{dkred}{\sout{\texttt{End}}}
+    [.\textcolor{dkred}{\sout{\texttt{ENames}}}
+      \texttt{Name}
+      [.\texttt{Names}
+        \texttt{Name} ]]]
+\caption{Marking of Definitions which must be expanded}
+\label{fig:expanded-defs}
+\end{figure}
+
+Trees of user defined regular expressions,
+with those that must be replaced by their definitions crossed out.
 
@@
 \begin{lstlisting}
-Bib: ""   | ("author={" Name ", " Name Names (" and " (" Name ", " Name Names))* "},")
+Bib': ""   | ("author={" Name ", " Name Names (" and " (Name ", " Name Names))* "},")
 
-End: ("%A " (Name Names " " Name) "\n")*
+End': ("%A " (Name Names " " Name) "\n")*
 \end{lstlisting}
@@
 
-A basic synthesis approach from here will quickly fail.  Both \texttt{BibTex'}
-and \texttt{EndNote'} are of the form $\Regex_1\Concat\Regex_2$, but there is
+A basic type directed synthesis approach from here will quickly fail.
+Both \texttt{Bib'}
+and \texttt{End'} are of the form $\Regex_1\Concat\Regex_2$, but there is
 not a lens between \texttt{"\%0"} and
@@
 By putting them in a form that normalizes away the equivalences of
-associativity, distributivity, and \EmptyString{} identity.
-
-The normalized form of these equivalences is too long to be contained in this
-paper.  Instead, we will work through \afm{or should we just do the easier
-example shown above}.
-
-With those equivalences are handled, perhaps the lens can be synthesized.
-A lens can be synthesized, if there is some pairing of sequences such that
-there are sequence lenses between the two elements of the pair for all pairings.
-A sequence lens can be synthesized if there is some pairing of atoms, such that
-there is an atom lens between the two atoms.  An atom lens can be synthesized
-if the regular expressions the same user defined data type, or if there is a dnf
-lens between the two dnfs under the stars.
-
-FIGURE:
-
-"" | Author Author*
-
-AuthorEndNote*
-
-Looking at the number of sequences, we can easily see that these dnf regular
-expressions do not have a lens between them, without some rewrite rules applied.
-However, there are a large number of possible rewrite rules that can be applied.
-Instead of taking and testing all of them, each rewrite rule is applied, and
-then placed in a priority queue.
-The priority for each problem is a linear combination of the number of rewrites
-applied, and the distance between the two dnf regular expressions by a
-pseudometric.
-
-This pseudometric is defined as the sum of two metrics.
-The first metric is a standard distance metric over the size of the two dnf
-regular expressions.
-The second metric is defined based on the distribution of user defined regular
-expressions.
-We can create an infinite dimensional free module over the integers from the
-regular expressions, generated by the set $\SetOf{(U,i)\SuchThat U\text{is a
-user defined data type}\BooleanAnd i\in\Nats}$.
-We can provide a function $f$ that maps all regular expressions into the element
-which counts how many user defined data types there are at a given depth.
-\afm{i should probably write down this function with a mapsto style, instead of
-just through text}
-The second metric is the taxicab metric on the elements of the vector space
-mapped to by the regular expressions.
-
-FIGURE:
+associativity, distributivity, and \EmptyString{}.
 
 \begin{lstlisting}
-"" | (name,name) (name,name)* -> \{ ((Author,0),2) , ((Author,1),4) \}
-(name,name)* -> \{ ((name,1),2) \}
+DNFBib':
+  ""
+  | "author=" Name ", " Name "" Names ""
+    (" and " Name ", " Name "" Names "")* ""
 
-d("" | (name,name) (name,name)*,(name,name)*) =
-taxicab(\{ ((Author,0),2) , ((Author,1),4) \},other)+(8-2)
+DNFEnd':
+  "" ("%A " Name "" Names " " Name "\n")* ""
 \end{lstlisting}
 
-Empirically, good transformations, like (name,name)* -> (name,name)+
-do very well, giving a distance of zero, where bad transformations, like
-name -> [A-Z][a-z]* do not do as well.
+With those equivalences handled, one might think it is now possible to
+synthesize the lens.
+Unfortunately, we are still unable to preform any synthesis because the types
+on the left and right are quite different, and the lens connectives require
+similar types of the left and right.
+For example, as \texttt{DNFBib'} has an outer |, we know an or lens must be
+synthesized.
+However, an or lens requires an outer | on both sides, and \texttt{DNFEnd'} does
+not have any outer |.
 
-Using the priority queue, the problem that gets popped would be
-"" | (name,name) (name,name)* <=> "" | (name,name) (name,name)*.
-Unlike the previous example, this cannot be so easily determined to not be
-equivalent.
+While, with normalization, certain equivalences have been handled, there are
+still many equivalences that haven't been handled.
+For example, the star in \texttt{End'} can be unrolled, turning \texttt{End'},
+and \texttt{DNFEnd'} into
 
-A naive approach to this would be to attempt all possible pairings of sequences,
-and all pairings of atoms within those sequences.  However, the complexity of
-trying all permutations is (\# permutations)!.
+\begin{lstlisting}
+End'': "" | ("%A " (Name Names " " Name) "\n"
+("%A " (Name Names " " Name) "\n")*)
 
-Instead, an ordering can be provided to atoms, clauses, and dnf regular
-expressions, where two elements are equal in the ordering if there exists a lens
-between them.  Upon applying that ordering to the elements of the dnf regular
-expression, determining whether there is a lens between two dnf regular
-expressions amounts to merely seeing if each element of the ordering is
-equal with respect to the ordering, as the one on the other side.
-
-FIGURE
-\begin{lstlisting}
-e | (n,n)(n,n)* <=> e | (n,n)(n,n)*
-reorder
-e | (n,n)(n,n)* <=> e | (n,n)(n,n)*
-matched together
-e <=> e , (n,n) <=> (n,n) , etc.
+DNFEnd'':
+  ""
+  | "%A " Name "" Names " " Name "\n"
+    ("%A " Name "" Names " " Name "\n")* ""
 \end{lstlisting}
 
-This still isn't complete yet, for we don't know if the generated lens will
-satisfy the examples.  The same problem arises with multiple permutations
-being possible.  If all the sequences can be mapped to each other, then there is
-a valid lens for each matching of the sequences!  And not all of these
-valid lenses satisfy the examples.
-To resolve this issue, we embed the parse trees of the examples into the
-regular expressions, and treat regular expressions with different aspects of the
-parse trees embedded as differently in the ordering.
+\texttt{DNFBib'} and \texttt{DNFEnd'} now look very similar.
+They both have an outer |, which has on one side a string, and on the other
+side, two instances of \texttt{Name}, and an instance of \texttt{Names} outside
+of any stars,
+with two instances of \texttt{Name}, and an instance of \texttt{Name} underneath
+a star.
+From this, a lens can easily be made.
 
-FIGURE
-e | name name(name, name)* , lalalaexample
+However, there is some element of choice in how the lens can be made.
+The outer \texttt{Name}s can each be sent to each other.
+and the inner \texttt{Name}s can each be sent to each other, so which
+should get sent to which.
+This can be determined from the examples.
+If the first \texttt{Name} in \texttt{DNFBib'} were sent to the first
+\texttt{Name} in \texttt{DNFEnd'}, then \texttt{Conway, John} would
+get mapped to \texttt{\%a Conway John}, which is inconsistent with the
+examples.
+The examples can be used, when this situation arises, to specify which parts
+of a regular expression can be mapped to which, when the types alone leave
+this ambiguous.
 
-\begin{lstlisting}
- /\
-e  |
-   /\
- name *
-      .
-     /\
-   name name
-
-lalaexample -> or2 (lala . Iterated(ex.am,p.le)
-
-combine to
-
-(the top thing but with the parse tree embedded alongside the expressions)
-\end{lstlisting}
+\afm{I am not sure going into the metric and priority queue are worth going
+  into,
+  and least not here?}
+%%% Local Variables:
+%%% TeX-master: "main"
+%%% End:

### 2016-09-16 paper/implementation.tex
added rough draft of all sections

@@
 regular expressions, and examples, and converts the regular expressions into
-DNF regular expressions, using \ToDNFRegex{}.
+DNF regular expressions, using $\ToDNFRegex{}$.
 These then get enqueued, and SynthDNFLens, where the bulk of the work
@@
 The distribution of user defined regular expressions can be viewed as an
-infinite dimensional free module over the integers, \Module{}.
+infinite dimensional vector space over the reals, $\VectorSpace{}$.
 The basis for this is $\SetOf{1*(\RegexVariable,n) | n\in\Nats, \RegexVariable
@@
 \GetDist(\RegexVariable)&=1*(\RegexVariable,1)\\
-\GetDist(\IterateLens{\DNFLens})&=\phi(\GetDist(\DNFLens))
+\GetDist(\IterateLensOf{\DNFLens})&=\phi(\GetDist(\DNFLens))
 \end{tabular}
@@
 
-$\Distance_{dist}(\DNFRegex,\DNFRegexAlt)=\LOneNorm(\phi(\DNFRegex),\phi(\DNFRegexAlt))$
-where \LOneNorm{} is the taxicab metric on \Module{}.
+$\Distance(\DNFRegex,\DNFRegexAlt)=\LOneNorm(\GetDist(\DNFRegex),\GetDist(\DNFRegexAlt))$
+where $\LOneNorm$ is the taxicab metric on $\VectorSpace{}$.
 The priority of a pair of DNF regular expressions, $(\DNFRegex{},\DNFRegexAlt{})$,
@@
   Given an ordering $\leq : A \rightarrow A \rightarrow \SetOf{-1,0,1}$,
-  and $\ListOf{a_1;\ldots;a_n}\in A \ListType$
+  and $\ListOf{a_1;\ldots;a_n}\in \ListTypeOf{A}$
   We define $\SortingOf{\leq}{\ListOf{a_1;\ldots;a_n}}$ as
@@
     $\DNFRegex \DNFLeq \DNFRegexAlt$
+
+  \item 
+    $\DNFRegex \DNFEq \DNFRegexAlt$
+    if
+    $\DNFRegex \DNFLeq \DNFRegexAlt$ and $\DNFRegexAlt \DNFLeq \DNFRegex$.
+
+  \item
+    $\Sequence \SequenceEq \SequenceAlt$
+    if
+    $\Sequence \SequenceLeq \SequenceAlt$ and $\SequenceAlt \SequenceLeq
+    \Sequence$.
+
+  \item
+    $\Atom \AtomEq \AtomAlt$
+    if
+    $\Atom \AtomLeq \AtomAlt$ and $\AtomAlt \AtomLeq \Atom$.
   \end{enumerate}
@@
   \begin{enumerate}
-  \item A DNF regular expression $\DNFOf{Sequence_1;\ldots;\Sequence_n}$ is
+  \item A DNF regular expression $\DNFOf{\Sequence_1;\ldots;\Sequence_n}$ is
     \textit{normalized} if $i \leq j$ implies $\Sequence_1\SequenceLeq\Sequence_j$,
@@
     where
-    $\sigma=\SortingOf{\SequenceLeq}
+    $\sigma=\SortingOf{\AtomLeq}
     {\ListOf{\NormalizeOf{\Atom_1};\ldots;\NormalizeOf{\Atom_n}}}$
@@
 
-Critically, the normalized form of the DNF regular expression, and the ordering
-on DNF regular expressions, are tied to the existance of a DNF Lens without
-rewrites.
+\begin{theorem}
+  \begin{enumerate}
+  \item
+    If $\DNFOf{\Sequence_1;\ldots;\Sequence_n}$ and
+    $\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_m}$ are normalized, and
+    $\DNFOf{\Sequence_1;\ldots;\Sequence_n}\DNFEq\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}$
+    then
+    $n=m$ and
+    $\Sequence_i = \SequenceAlt_i$ for all $i\in\RangeIncInc{1}{n}$
+  \item
+    If $\NormalizeOf{\SequenceOf{\Atom_1;\ldots;\Atom_n}} \DNFEq
+    \NormalizeOf{\SequenceOf{\AtomAlt_1;\ldots;\AtomAlt_n}}$ then
+    $\NormalizeOf{\Atom_{\sigma_1(i)}} =
+    \NormalizeOf{\AtomAlt_{\sigma_2(i)}}$ for all $i$
+    where $\sigma_1 =
+    \SortingOf{\SequenceLeq}{\ListOf{\NormalizeOf{\Atom_1};\ldots;\NormalizeOf{\Atom_n}}}$
+    and $\sigma_2 =
+    \SortingOf{\DNFLeq}{\ListOf{\NormalizeOf{\AtomAlt_1};\ldots;\NormalizeOf{\AtomAlt_n}}}$
+  \item
+    tester
+   % If $\NormalizeOf{\StarOf{\DNFRegex}}\DNFEq\NormalizeOf{\StarOf{\DNFRegexAlt}}$
+  \end{enumerate}
+\end{theorem}
+
+\begin{definition}
+  \begin{enumerate}
+  \item
+    Let
+    $\DNFOf{\Sequence_1;\ldots;\Sequence_n}$
+    and
+    $\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}$
+    be two DNF regular expressions, where
+    $\NormalizeOf{\DNFOf{\Sequence_1;\ldots;\Sequence_n}} \DNFEq
+    \NormalizeOf{\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}}$.
+    $\DNFLensSynthOf{\DNFOf{\Sequence_1;\ldots;\Sequence_n}}
+    {\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}}=
+    (\DNFLensOf{\SequenceLens_1;\ldots;\SequenceLens_n},\sigma)$
+    where
+    $\SequenceLens_i=\SequenceLensSynthOf{\Sequence_1}{\SequenceAlt_{\sigma(i)}}$
+    and
+    $\sigma=\SortingOf{\SequenceLeq}{\ListOf{\NormalizeOf{\Sequence_1};\ldots;\NormalizeOf{\Sequence_n}}}
+    \Compose
+    \SortingOf{\SequenceLeq}{\ListOf{\NormalizeOf{\SequenceAlt_1};\ldots;\NormalizeOf{\SequenceAlt_n}}}$
+  \item
+    Let
+    $\SequenceOf{\String_0;\Atom_1;\ldots;\Atom_n;\String_n}$
+    and
+    $\SequenceOf{\StringAlt_0;\AtomAlt_1;\ldots;\AtomAlt_n;\StringAlt_n}$
+    be two sequences, where
+    $\NormalizeOf{\SequenceOf{\String_0;\Atom_1;\ldots;\Atom_n;\String_n}} \SequenceEq
+    \NormalizeOf{\SequenceOf{\StringAlt_0;\AtomAlt_1;\ldots;\AtomAlt_n;\StringAlt_n}}$.
+    $\SequenceLensSynthOf{\SequenceOf{\String_0;\Atom_1;\ldots;\Atom_n;\String_n}}
+    {\SequenceOf{\StringAlt_0;\AtomAlt_1;\ldots;\AtomAlt_n;\StringAlt_n}}=
+    (\SequenceLensOf{(\String_0,\StringAlt_0);\AtomLens_1;\ldots;\AtomLens_n;(\String_n,\StringAlt_n)},\sigma)$
+    where
+    $\AtomLens_i=\AtomLensSynthOf{\Atom_1}{\AtomAlt_{\sigma(i)}}$
+    and
+    $\sigma=\SortingOf{\AtomLeq}{\ListOf{\NormalizeOf{\Atom_1};\ldots;\NormalizeOf{\Atom_n}}}
+    \Compose
+    \SortingOf{\SequenceLeq}{\ListOf{\NormalizeOf{\AtomAlt_1};\ldots;\NormalizeOf{\AtomAlt_n}}}$
+  \item
+    Let
+    $\StarOf{\DNFRegex}$
+    and
+    $\StarOf{\DNFRegexAlt}$
+    be two atoms, where
+    $\NormalizeOf{\StarOf{\DNFRegex}} \SequenceEq
+    \NormalizeOf{\StarOf{\DNFRegexAlt}}$.
+    $\AtomLensSynthOf{\StarOf{\DNFRegex}}
+    {\StarOf{\DNFRegexAlt}}=
+    \IterateLensOf{\DNFLensSynthOf{\DNFRegex}{\DNFRegexAlt}}$
+  \end{enumerate}
+\end{definition}
 
 \begin{theorem}
-  $\NormalizeOf{\DNFRegex}\DNFLeq\NormalizeOf{\DNFRegexAlt}$ and
-  $\NormalizeOf{\DNFRegexAlt}\DNFLeq\NormalizeOf{\DNFRegex}$ if, and only
+  $\NormalizeOf{\DNFRegex}\DNFEq\NormalizeOf{\DNFRegexAlt}$
+  if, and only
   if, there exists a $\DNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt$,
-  where $\DNFLens$ has no applications of a rewrite rule.
+  where the typing of $\DNFLens\OfType\DNFRegex\Leftrightarrow\DNFRegexAlt$
+  has no applications of a rewrite rule.
 \end{theorem}
 
-This can be proven with mutual induction, proving similar statements for
-sequences and atoms.
-
-Furthermore, two DNF regular expressions can be zipped together to form a
-DNF lens.  Consider $\DNFOf{\Sequence_{\sigma_1(0)};\ldots;\Sequence_{\sigma_1(n)}}$, and
-$\DNFOf{\SequenceAlt_{\sigma_2(0)};\ldots;\SequenceAlt_{\sigma_2(n)}}$, two normalized
-DNF regular expressions.  $\Sequence_{\sigma(0)}$ merely needs to be mapped to
-$\SequenceAlt_{\tau(0)}$.  Furthermore, the permutation mapping the two lenses
-merely needs to be $\sigma_1 \Compose \sigma_2^{-1}$
-
-We can use these orderings to find
-
-Beyond that, we can use these orderings to find the lenses.
-If $\DNFOf{\Sequence_1;\ldots;\Sequence_n}\leq\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}$,
-and $\DNFOf{\SequenceAlt_1;\ldots;\SequenceAlt_n}\leq\DNFOf{\Sequence_1;\ldots;\Sequence_n}$,
-then $\Sorting([\Sequence_1;\ldots;\Sequence_n])\leq\Sorting([\Sequence_1;\ldots;\Sequence_n])$
-using the dictionary ordering, which means that
-$\Sorting([\Sequence_1;\ldots;\Sequence_n])[i]\leq
-\Sorting([\SequenceAlt_1;\ldots;\SequenceAlt_n])[i]$ and
-$\Sorting([\SequenceAlt_1;\ldots;\SequenceAlt_n])[i]\leq
-\Sorting([\Sequence_1;\ldots;\Sequence_n])[i]$, which means that there is a
-sequence lens between them.
-\afm{I can keep going on about this stuff, should it go here?  Should I just make a brief claim
-about it and move on}
-
-Unfortunately, while this can find a well typed DNF regular expression, it doesn't
+Unfortunately, while this can find a well typed DNF lens, it doesn't
 necessarily find the one that matches the examples.  In the situation where
@@
 This is done by joining additional parsing information alongside the regular
-expression components.
+expression components, and creating an ordering on DNF Regular expressions
+with parsing information.\afm{This would be a really in depth thing to describe,
+worth going into?}
 

### 2016-09-16 paper/related-work.tex
added variable information on master file to each file
added rough draft of all sections

@@
 \subsection{FlashFill}
+FlashFill is another string transformation program.  FlashFill takes input and
+output examples as its only form of specification.  For FlashFill's primary use
+case, providing easy transformations of Excel data, input and output examples as
+the only specification makes a lot of sense.  Our tool is intended for use for
+programmers to use, who already know regular expressions.  Our tool is oriented
+towards a situation where a programmer would check the generated code into a
+repository.  From requiring this extra information, our tool is able to handle
+synthesize complicated functions that go beyond what FlashFill can synthesize.
+For example, given 21 examples of extracting the first author, with the names
+reversed, FlashFill still gave incorrect outputs for a number of inputs, even
+when those inputs are a similar form to the provided examples.
+
+Furthermore, even if FlashFill were to synthesize the correct program, it would
+be foolhardy to check the generated code into a repository.  FlashFill does not
+fail on inputs it doesn't know how to handle, rather it tries to provide an
+output.  This is different from the generated lenses, which will quickly
+fail on inputs that do not match it's intended specification.  Furthermore, the
+specification of what inputs should be handled are provided by the user, not
+inferred from the examples, giving it more reliability. \afm{talk about how it
+  did on bibtex example, more specific on what it did wrong, maybe?}
+
 \subsection{FlashExtract}
-\subsection{SemFill}
+FlashExtract is a program that takes an input of a string, and a partial
+labeling of substrings within that string, to try to label the rest of the
+substrings.  FlashExtract was able to label some amounts of the strings, but was
+unable to handle reorderings of the data well.  While it would do well in
+extracting data when the data came in the same form, for example, when given a
+Bibtex file where the fields come in a specific order, the data would do well,
+however, FlashExtract was unable to handle when fields were given in different
+orders, even if previous examples in that order were given.
+
+%%% Local Variables:
+%%% TeX-master: "main"
+%%% End:

### 2016-09-22 paper/extended-example.tex
revised sections 1-5

@@
 \section{Lens Synthesis by Example}
-
-\dpw{It may be good to hack lstlisting so it can be in a backslash small font.
-There is perhaps another environment we could use --- a code environment.}
 
@@
 \subsection{Regular Bijective Lenses}
-
+\afm{I'm not sure how necessary this section is.  While I want to give users a
+  bit of idea about how things map between one another, I think this section
+  might go into too much detail.  How would I be able to give the intuition of
+  finding parts that map to each other, without just describing lenses (assuming
+  that's the right sort of thing to do)?}
 Once we have a synthesis specification, we need to find a lens that
@@
 In order to guarantee the functions generated by these combinators
-form proper bijections between source and target, they are given types using regular expressions.  
-We typicallly write 
-$\Lens \OfType \Regex \Leftrightarrow \RegexAlt$ to indicate that
-a lens $\Lens$ converts back and forth between data belonging
+form proper bijections between source and target, they are given types using
+regular expressions.  We typically write
+$\Lens \OfType \Regex \Leftrightarrow \RegexAlt$ to indicate that a lens $\Lens$
+converts back and forth between data belonging
 to the languages of regular expressions
@@
 Each of the other lenses manipulate structured data.  For instance,
-$\ConcatLensOf{\Lens_1}{\Lens_2}$
+$\ConcatLensOf{\Lens_1}{\Lens_2}$ operates by applying $\Lens_1$ to the left
+portion of a string, and $\Lens_2$ to the right, and concatenating the results.
+$\SwapLensOf{\Lens_1}{\Lens_2}$ does the same as $\ConcatLensOf{\Lens_1}{\Lens_2}$
+but it swaps the results before concatenating.
+$\OrLensOf{\Lens_1}{\Lens_2}$ operates by applying either $\Lens_1$ or $\Lens_2$
+to the string.  $\IterateLensOf{\Lens}$ operates by repeatedly applying $\Lens$
+to subparts of a string.
 
@@
 
-\dpw{----------------}
-
-We are trying to synthesize \BibEnd{}.
+\begin{figure}
+  \Tree[.\textcolor{dkred}{\sout{\CF{Bib}}}
+    [.\textcolor{dkred}{\sout{\CF{BNames}}}
+      \CF{Name}
+      [.\CF{Names}
+        \CF{Name} ]]]
+  \Tree[.\textcolor{dkred}{\sout{\CF{End}}}
+    [.\textcolor{dkred}{\sout{\CF{ENames}}}
+      \CF{Name}
+      [.\CF{Names}
+        \CF{Name} ]]]
+\caption{
+Trees of user defined regular expressions,
+with those that must be replaced by their definitions crossed out.
+}
+\label{fig:expanded-defs}
+\end{figure}
 
@@
 \begin{figure}
-  \Tree[.\textcolor{dkred}{\sout{\texttt{Bib}}}
-    [.\textcolor{dkred}{\sout{\texttt{BNames}}}
-      \texttt{Name}
-      [.\texttt{Names}
-        \texttt{Name} ]]]
-  \Tree[.\textcolor{dkred}{\sout{\texttt{End}}}
-    [.\textcolor{dkred}{\sout{\texttt{ENames}}}
-      \texttt{Name}
-      [.\texttt{Names}
-        \texttt{Name} ]]]
-\caption{Marking of Definitions which must be expanded}
-\label{fig:expanded-defs}
+\begin{lstlisting}
+Bib':
+  "" 
+  | ("author={" Name ", " Name Names
+    (" and " (Name ", " Name Names))* "},")
+
+End':
+  ("%A " (Name Names " " Name) "\n")*
+\end{lstlisting}
+\label{fig:examples-expanded}
+\caption{
+  \CF{Bib} and \CF{End} with necessary name replacements preformed.
+}
 \end{figure}
-
-Trees of user defined regular expressions,
-with those that must be replaced by their definitions crossed out.
-
 After completing these necessary expansions, the problem becomes synthesizing a
-lens between the regular expression
-\begin{lstlisting}
-Bib': ""   | ("author={" Name ", " Name Names (" and " (Name ", " Name Names))* "},")
-
-End': ("%A " (Name Names " " Name) "\n")*
-\end{lstlisting}
-and
+lens between the regular expressions \CF{Bib'} and \CF{End'} shown in
+Figure~\ref{fig:examples-expanded}.
 
 A basic type directed synthesis approach from here will quickly fail.
-Both \texttt{Bib'}
-and \texttt{End'} are of the form $\Regex_1\Concat\Regex_2$, but there is
-not a lens between \texttt{"\%0"} and
-\texttt{("@inproceedings" | "@article" | "@book")}.
+Both \CF{Bib'}
+and \CF{End'} are of the form $\Regex_1\Concat\Regex_2$, but there is
+not a lens between \CF{"\%0"} and
+\CF{("@inproceedings" | "@article" | "@book")}.
 
-These issues can be handled by normalizing the regular expressions slightly.
-By putting them in a form that normalizes away the equivalences of
-associativity, distributivity, and \EmptyString{}.
-
+\begin{figure}
 \begin{lstlisting}
@@
   ""
-  | "author=" Name ", " Name "" Names ""
-    (" and " Name ", " Name "" Names "")* ""
+  | "author={" Name ", " Name "" Names ""
+    (" and " Name ", " Name "" Names "")* "},"
 
@@
 \end{lstlisting}
+\label{fig:examples-expanded-normalized}
+\caption{
+  \CF{Bib'} and \CF{End'} put into a semi-normal form.
+}
+\end{figure}
+
+These issues can be handled by normalizing the regular expressions slightly.
+We put them in a form that normalizes away the equivalences involving
+associativity, distributivity, and \EmptyString{}, shown in
+Figure~\ref{fig:examples-expanded-normalized}
 
@@
 similar types of the left and right.
-For example, as \texttt{DNFBib'} has an outer |, we know an or lens must be
+For example, as \CF{DNFBib'} has an outer $\Or$, we know an or lens must be
 synthesized.
-However, an or lens requires an outer | on both sides, and \texttt{DNFEnd'} does
-not have any outer |.
+However, an or lens requires an outer $\Or$ on both sides, and \CF{DNFEnd'} does
+not have any outer $\Or$.
 
-While, with normalization, certain equivalences have been handled, there are
-still many equivalences that haven't been handled.
-For example, the star in \texttt{End'} can be unrolled, turning \texttt{End'},
-and \texttt{DNFEnd'} into
-
+\begin{figure}
 \begin{lstlisting}
-End'': "" | ("%A " (Name Names " " Name) "\n"
-("%A " (Name Names " " Name) "\n")*)
-
-DNFEnd'':
+DNFEndUnroll:
   ""
@@
     ("%A " Name "" Names " " Name "\n")* ""
+
+DNFEndMod2:
+  ""
+      ("%A " Name "" Names " " Name
+        "\n%A " Name "" Names " " Name "\n")* ""
+  | "%A " Name "" Names " " Name "\n"
+      ("%A " Name "" Names " " Name
+        "\n%A " Name "" Names " " Name "\n")* ""
+
 \end{lstlisting}
+\label{fig:examples-star-equivalenced}
+\caption{
+  \CF{DNFEnd'} with some star equivalences preformed.
+}
+\end{figure}
 
-\texttt{DNFBib'} and \texttt{DNFEnd'} now look very similar.
-They both have an outer |, which has on one side a string, and on the other
-side, two instances of \texttt{Name}, and an instance of \texttt{Names} outside
-of any stars,
-with two instances of \texttt{Name}, and an instance of \texttt{Name} underneath
-a star.
-From this, a lens can easily be made.
+\paragraph*{Step 2:  Traversing Equivalences.}
+While, with normalization, certain equivalences have been handled, there are
+still many equivalences that haven't been handled.
+For example, the equivalences $\StarOf{A}=\EmptyString \Or A\StarOf{A}$ and
+$\StarOf{A}=(\EmptyString \Or A)\StarOf{(AA)}$ can be applied to the star in
+\CF{End'}, resulting in \CF{DNFEndUnroll} and \CF{DNFEndMod2}, shown in
+Figure~\ref{fig:examples-star-equivalenced}.
 
-However, there is some element of choice in how the lens can be made.
-The outer \texttt{Name}s can each be sent to each other.
-and the inner \texttt{Name}s can each be sent to each other, so which
-should get sent to which.
-This can be determined from the examples.
-If the first \texttt{Name} in \texttt{DNFBib'} were sent to the first
-\texttt{Name} in \texttt{DNFEnd'}, then \texttt{Conway, John} would
-get mapped to \texttt{\%a Conway John}, which is inconsistent with the
-examples.
-The examples can be used, when this situation arises, to specify which parts
-of a regular expression can be mapped to which, when the types alone leave
-this ambiguous.
+However, now a single problem of finding a lens between regular expression has
+become two problems!  This can be solved by using a priority queue, where the
+elements in the queue are regular expression specifications.  The priority of
+each specification is a combination of how many expansions have been preformed,
+and how different the two regular expressions in the specification are.
+\CF{DNFEndUnroll} and \CF{DNFBib'} are very similar, in that they have the same
+number of instances of \CF{Name} and \CF{Names}.  However, \CF{DNFEndMod2} is
+very different, with a different number of \CF{Name}, \CF{Names}, and a
+different number of $\Star$s.  Because of this large difference, the
+specification of \CF{DNFBib' <=> DNFEndUnroll} would get popped, instead of the
+specification of \CF{DNFBib' <=> DNFEndMod2}.
 
-\afm{I am not sure going into the metric and priority queue are worth going
-  into,
-  and least not here?}
+\paragraph*{Step 3: Creating Lenses.}
+Because the regular expressions of  \CF{DNFBib'} and \CF{DNFEndUnroll} are so
+similar, type directed synthesis can now happen between them.
+Following the types, the outer \CF{""}s
+would get mapped to each other, and the sides of the disjunct with $\Star$s
+would get mapped to each other.  Within that side, \CF{Name} would map to
+\CF{Name}, \CF{Names} would map to \CF{Names} and the $\Star$s would map to the
+$\Star$s.
+
+In doing this though, some ambiguity is introduced.  There are two
+instances of \CF{Name} present outside a star, so which one should get mapped to
+which?  The basic solution would be to map the first to the first, and the
+second to the second.  This is not correct, however, as the generated lens for
+this will not match the examples.  We can determine which maps to which, as they
+must each match the same substrings.  The first \CF{Name} in \CF{DNFBib'}
+matches \CF{"Conway"} and the second matches \CF{"John"}.  The first \CF{Name} in
+\CF{DNFEndUnroll} matches \CF{"John"} and the second matches \CF{"Conway"}.
+This means that the first \CF{Name} in \CF{DNFBib'} must map to the second
+\CF{Name} in \CF{DNFEndUnroll}, and vice-versa.  Doing this will generate a lens
+that satisfies the type and the examples.
+
 %%% Local Variables:

### 2016-09-22 paper/introduction.tex
added variable information on master file to each file
added rough draft of all sections
revised sections 1-5

--- 2016-06-29 paper/introduction.tex
@@
 
-\begin{enumerate}
-\item Search through all possible lenses by: searching for good regular
-expressions, then searching within those regular expressions for type directed
-lenses
-\item Putting the regular expressions in a normal form reduces the number of
-regular expressions we must search through
-\item Using user defined data types allows the synthesis algorithm to keep the
-internals of regular expressions abstracted, unless the internals of those
-regular expressions are changed, and the abstraction must be broken.
-\item Using a pseudometric on regular expressions, based on the
-distribution of user defined data types and size of the regular expression,
-allows for an efficient search through equivalent regular expressions
-\item We use a data structure combining examples with regular expressions to
-synthesize only candidate programs which obey the examples
-\item We develop an efficient strategy for finding if two
-regular expressions have a lens between them requiring no rewrites which amounts
-to merely ordering the components.
-\end{enumerate}
-
-In the era of data, big and small, tools for reliably parsing,
-printing, cleaning, and transforming data are increasing important.
-One class of such tools are \emph{lenses}~\cite{?}, bi-directional
+The growing dependence on big data makes tools for reliably parsing,
+printing, cleaning, and transforming data increasing important.
+One such class of tools are \emph{lenses}~\cite{Focal2005-long}, bi-directional
 programs that help users solve the classic ``view update problem.''
@@
 
-One concrete example of where lenses provide value is in managing
-linux configuration files.  Such files come in a wide variety of
-different ``ad hoc'' data formats, and even experienced system
-administrators have a hard time keeping track of all the format
-variations.  Moreover, a missing comma, forgotten keyword or 
-misplaced vertical bar can cause the systems that depend on such
-configurations to crash or misbehave.  In order to manage such configuration
-files more reliably,  RedHat developed Augeas~\cite{}, a configuration
-editing system based around lenses.  At the moment, Augeas contains more
-than 200 stock lenses to provide views of common configuration files.
+Bidirectional programming is used heavily in practice.
+Augeas~\cite{?}, a configuration editing system for linux, uses lens combinators
+to provide editable views of configuration files.
+Bixid~\cite{?} is a language developed for expressing bidirectional XML
+transformations.
+XSugar~\cite{?} is used to convert bidirectionally between XML and non-XML data,
+and is used to convert between EpiDoc~\cite{?} and Leiden+~\cite{?} formats for
+encoding ancient documents~\cite{?}.
+Serializers and deserializers combine to make a bidirectional program, though
+they are rarely written in a domain specific language for bidirectional
+programming.
 
 Lenses are popular because they typically come with strong guarantees
-about the get-put (or put-get) round-trip behavior.  Such guarantees
+about the get-put (and put-get) round-trip behavior.  Such guarantees
 help users guard against data corruption while reading and editing
-high-value data sources.
+high-value data sources.  These guarantees are often achieved through
+the use of domain-specific languages with strong type systems that
+guarantee invertability of lens programs~\cite{?,?,?}.  Unfortunately, such
+lens languages are currently quite non-standard.  The
+process of writing a lens is challenging, as a user must think
+about data conversion in two directions at once --- an
+unusual task for the everyday programmer.  As a consequence, 
+for the casual user of lenses, learning the language, its type system and its
+unorthodox programming style is time-consuming and, in the end,
+a disincentive to adopting the technology, despite its advantages.
+
+In order to overcome the steep learning curve involved in programming
+with lenses, we have developed a new system for synthesis of lenses
+from their types and a collection of representative examples.  More
+specifically, we focus on algorithms for the synthesis of bijective
+string lenses.  The types for such lenses are regular expressions,
+which describe the format of the source and target strings.  Given
+such types, and a very small number of representative examples (sometimes
+no examples at all), we synthesize a well-typed lens that implements
+a bijection between data formats.
+
+Recently, researchers have developed a number of new algorithms for
+type-directed synthesis~\cite{?,?,?}.  Many of those
+algorithms~\cite{osera+:pldi15,frankle+:popl16,armando+:pldi16} are
+directed in part by the syntax of types.  For instance, to find an
+expression with type $A \rightarrow B$, such algorithms will assume
+that the expression in question is a function with the form $\lambda
+x{:}A. e$, and they will continue the search for an expression $e$
+with type $B$ under the hypothesis that the parameter $x$ has type
+$A$.  In the case of the simply-typed lambda calculus, such an
+approach reduces the complexity of the search for terms inhabiting the
+type without losing completeness.  Lenses are defined by a richer type system
+than the simply typed lambda calculus,
+with their regular expression types, as
+there are many equivalent ways to express the same regular expression.
+Unfortunately, this makes simple
+syntax-directed algorithms not directly applicable.
+
+\paragraph*{The Central Challenge.}
+To illustrate the central challenge facing the design of our synthesis system
+in greater detail,
+consider the concatenation lens, $\Lens_1 \Concat \Lens_2$,
+one of the simplest string lens combinators.  
+When viewed from left to right, this lens defines a function \emph{get}, which,
+given a string 
+$\String$, divides $\String$ into two substrings $\String_1$ 
+and $\String_2$, applies $\Lens_1$ to the first, yielding $t_1$,
+and $\Lens_2$ to the second, yielding $t_2$.  When viewed from
+right to left, it defines a second function \emph{put}, which maps
+data back from strings $t$ to strings $s$ in an analogous way.
+
+The type system for lenses ensures such \emph{get} and \emph{put} functions
+satisfy strong invertibility properties.  In our case, the two functions
+form a bijection.  In general, to 
+describe the type of a lens, we write $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$, which may be read ``lens $\Lens$ converts back and forth between
+data in the language of regular expression $\Regex$ and data in the language
+of regular expression $\RegexAlt$''.  More specifically, to give a type
+to the concatenation lens, we use the following typing rule.
+\[
+\inferrule
+{
+\Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
+\Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt_2\\\\
+\UnambigConcatOf{\Regex_1}{\Regex_2}\\
+\UnambigConcatOf{\RegexAlt_1}{\RegexAlt_2}
+}
+{
+\ConcatLensOf{\Lens_1}{\Lens_2} \OfType \Regex_1\Regex_2 \Leftrightarrow \RegexAlt_1\RegexAlt_2
+}
+\]
+Above, the notation $\UnambigConcatOf{\Regex_1}{\Regex_2}$ asserts that $\Regex_1$ and $\Regex_2$ are
+unambiguously concatenable, meaning there is
+exactly one way to divide up any
+string $\String$ in the language of $\Regex_1\Regex_2$ into $\String_1$
+and $\String_2$ such that $\String_1$ is in the language of $\Regex_1$
+and $\String_2$ is in the language of $\Regex_2$.  These unambiguous
+concatenation constraints guarantee the \emph{get} and \emph{put}
+functions are indeed functions and that they form a bijection if
+the underlying lenses are bijective.
+
+Now, with these typing rules in hand, consider trying to search for
+a lens with the type $(A B) C \Leftrightarrow A' (B' C')$ where
+we know of lenses with types 
+$A \Leftrightarrow A'$, $B \Leftrightarrow B'$, and 
+$C \Leftrightarrow C'$.  A straightforward,
+syntax-directed search suggests applying the rule for concatenation,
+and looking for two simpler lenses with the types $A B \Leftrightarrow A'$
+and $C \Leftrightarrow B' C'$ respectively.  Unfortunately, at this point,
+we are simply stuck---we have no lenses with those types.  Of course,
+the natural
+solution to this problem
+is to introduce a type conversion rule to the system:
+\[
+\inferrule
+{
+\Lens \OfType \Regex_1 \Leftrightarrow \Regex_2\\
+\Regex_1 \equiv \Regex_1'\\
+\Regex_2 \equiv \Regex_2'
+}
+{
+\Lens \OfType \Regex_1' \Leftrightarrow \Regex_2'
+}
+\]
+With such a type conversion rule, we could first 
+convert the type $(A B) C \Leftrightarrow A' (B' C')$
+to the better aligned $(A B) C \Leftrightarrow (A' B') C'$.
+When we then proceed with structure decomposition using the
+concatenation rule, we generate sub-problems involving
+finding lenses for types
+$A \Leftrightarrow A'$, $B \Leftrightarrow B'$, and 
+$C \Leftrightarrow C'$, which is just what we had in hand.
+
+Because equivalence of regular expressions is decideable, when
+given a lens $\Lens$ with type $\Regex_1 \Leftrightarrow \Regex_2$,
+it is easy to \emph{check} whether that lens also has type 
+$\Regex_1' \Leftrightarrow \Regex_2'$, and lens programming languages
+such as Boomerang \cite{foster:thesis} do so.  However, 
+when we turn to synthesis and are given the lens type 
+$\Regex_1' \Leftrightarrow \Regex_2'$ and the option to apply this
+rule, the search space becomes unmanageably large --- we must guess
+some arbitrary new regular expressions $\Regex_1$ and $\Regex_2$ that are
+equivalent to  $\Regex_1'$ and $\Regex_2'$ respectively
+and there are infinitely many such regular expressions to choose
+from.
+
+Furthermore, the language of lenses without a composition operator is not closed
+under composition.  There exists lenses $\Lens_1 \OfType \Regex_1 \Leftrightarrow
+\Regex_2$, $\Lens_2 \OfType \Regex_2 \Leftrightarrow \Regex_3$ where there is no
+lens $\Lens \OfType \Regex_1 \Leftrightarrow \Regex_3$ such that
+$\SemanticsOf{\ComposeLensOf{\Lens_2}{\Lens_1}}=\SemanticsOf{\Lens}$.  So, if we
+would like to be able to synthesize $\ComposeLensOf{\Lens_2}{\Lens_1} \OfType
+\Regex_1 \Leftrightarrow \Regex_3$ then we
+must also search through possible $\Regex_2$ for creating compositions.
+
+In essence,
+we must solve the problem of search \emph{in three dimensions}:  In one
+dimension, we must search for a regular expression \emph{type}
+with an appropriate ``syntactic shape''.  In a second dimension,
+we must search for a lens \emph{expression} that has the given type and
+is consistent with the user's examples.  In a final dimension, we must search
+for intermediary regular expressions for composing lenses.
+
+\paragraph*{Our Solution.}  Our solution to this problem is to define
+a new language of lenses that
+cuts down the synthesis search space \emph{in all dimensions}.
+This language is designed so that its regular expression types are limited 
+to those in a pseudo-canonical form.  This pseudo-canonical form
+demands that regular expressions be defined as a flattened 
+union of concatenated strings and atoms, where each atom is a user defined data
+type or the Kleene star of an expression that is recursively in normal form.
+When converted into this pseudo-canonical form, two regular expressions that 
+differed only by associativity of concatenation will be syntactically equal.
+For example, rather than searching for a lens with type 
+ $(A B) C \Leftrightarrow A' (B' C')$, our algorithm will always search
+for one with type  $(A B) C \Leftrightarrow (A' B') C'$
+
+More generally, many, but not all, structural incongruities that
+prevent the existence of lenses disappear.
+Not all structural incogruities disappear because our regular expressions
+are only \emph{pseudo-canonical}.  For instance, our pseudo-normal form
+does not automatically unify lenses such as
+$A B* \Leftrightarrow A (\epsilon + BB*)$ that result from an
+unrolling of Kleene Star, for instance.
+Consequently, we structure our synthesis algorithm in two parts:  The
+first is a syntax-directed search for a lens based on the structure of the given
+regular expression types and, if that fails, the second is a search,
+using Conway's axioms~\cite{conway} for 
+equivalent regular expression types that might allow a new search
+in the first phase of the search to succeed.
+
+Furthermore, in this new language, we can express transformations which were
+previously inexpressible without composition.  It is still an open problem
+whether this new language is closed under composition.  However, with the
+increased expressibility, we have found searching for a central regular
+expression to compose with to be unnecessary in practice.
+
+The types
+describing data formats may be very large --- they are regular
+grammars describing complex data formats.  This is another complication that
+arises in this synthesis setting, but not elsewhere in the literature on type
+directed synthesis.  On the one hand, this is
+very useful as such types provide a wealth of information that constrains
+the search process effectively.  On the other hand, there is a lot of
+data that must be managed carefully by the synthesis system.
+One important observation is that users naturally construct
+such types compositionally, using named abbreviations for subcomponents.
+The structure inherent in these named abbreviations can be 
+exploited by the synthesis algorithm to speed up the search dramatically.
+
+For example, if the name $A$ appears once on both the left and the right,
+the system will hypothesize that the identity lens can be used to convert
+between occurrences of $A$.  On the other hand, if the name $A$ appears
+on the left, but not on the right, the system will determine that the
+name $A$ should be replaced by its definition.  Likewise, differing
+numbers of occurrences of names can serve as a guide for unrolling
+Kleene star or applying other regular expression equivalence axioms.
+Hence, in general, to guide the transformation of regular expression
+types, we define a pseudometric based on the occurrence of names.
+
+\paragraph*{Contributions.}  We have designed and implemented
+a new system for the synthesis of bijective string lenses.
+In addition, this paper makes the following contributions:
+
+\begin{itemize}
+\item It defines a new language of lenses designed specifically for synthesis
+with types based on pseudo-canonical regular expressions, with a term 
+language to match.
+\item We formalize the semantics of the language and prove it sound with
+respect to an existing, standard language for bijective string lenses.
+\item We designed a type-directed, two-dimensional 
+algorithm for synthesizing lenses in our new domain-specific language.
+\item We developed a collection of optimizations that speed up the basic
+synthesis algorithm.
+\item We have implemented the language and evaluated it on a range of
+over 40 benchmarks including micro-benchmarks, a few examples from FlashFill,
+and a number of examples from the Augeas system for transforming linux
+file formats.  Our synthesis algorithm succeeds on all of our
+benchmarks.  When it succeeds, synthesis time is typically less than 1 second.
+\end{itemize}
+
+%%% Local Variables:
+%%% TeX-master: "main"
+%%% End:

### 2016-09-23 paper/abstract.tex
added variable information on master file to each file
revised sections 1-5
Some thoughts on the paper

@@
 The growing dependence on big data makes reliably converting
-between different data formats increasingly important.
+between different data formats increasingly important\bcp{This seems both
+  contentious and not terribly relevant}.
 \emph{Lenses} are a bi-directional technique for specifying such
@@
 this difficulty, we develop an algorithm to automatically synthesize
-lenses from format descriptions, specified as regular expressions, and
+lenses\bcp{just bijective ones!} from format descriptions, specified as regular expressions, and
 a small number of examples.  The rich structure of regular expressions
@@
 examples, and Augeas transformations of Linux configuration files.
-We show we are able to infer lenses for most examples in the
-benchmark suite in an average of 0.134 seconds.
+We show we are able to infer lenses for most\bcp{!} examples in the
+benchmark suite in an average of 0.134 seconds \afm{TODO: automatically calculate this from CSV}.
 
@@
 %% configuration files.
+
+%%% Local Variables:
+%%% TeX-master: "main"
+%%% End:

### 2016-09-23 paper/introduction.tex
Some thoughts on the paper

@@
 \section{Introduction}
+
+\bcp{May I make another plea for putting the whole paper into one big .tex
+  file?}
 
@@
 
-Bidirectional programming is used heavily in practice.
+Bidirectional programming is used heavily in practice.\bcp{This may be
+  overstating the case, for lenses per se.  Augeas is the only real-world
+  example AFAIK.}
 Augeas~\cite{?}, a configuration editing system for linux, uses lens combinators

### 2016-09-25 paper/introduction.tex
added some citations

--- 2016-09-22 paper/introduction.tex
@@
 Bidirectional programming is used heavily in practice.
-Augeas~\cite{?}, a configuration editing system for linux, uses lens combinators
+Augeas~\cite{augeas}, a configuration editing system for linux, uses lens combinators
 to provide editable views of configuration files.
-Bixid~\cite{?} is a language developed for expressing bidirectional XML
+The language biXid~\cite{bixid} is developed for expressing bidirectional XML
 transformations.
-XSugar~\cite{?} is used to convert bidirectionally between XML and non-XML data,
-and is used to convert between EpiDoc~\cite{?} and Leiden+~\cite{?} formats for
-encoding ancient documents~\cite{?}.
+XSugar~\cite{xsugar} is used to convert bidirectionally between XML and non-XML data,
+and is used to convert between EpiDoc~\cite{epidoc} and Leiden+~\cite{leidenplus} formats for
+encoding ancient documents~\cite{epidocleidenplus}.
 Serializers and deserializers combine to make a bidirectional program, though
@@
 the use of domain-specific languages with strong type systems that
-guarantee invertability of lens programs~\cite{?,?,?}.  Unfortunately, such
+guarantee invertability of lens programs~\cite{Focal2005-long,boomerang,symmetric-lenses}.  Unfortunately, such
 lens languages are currently quite non-standard.  The
@@
 Recently, researchers have developed a number of new algorithms for
-type-directed synthesis~\cite{?,?,?}.  Many of those
+synthesis~\cite{yag+:pldi16,?,?}.  Many of those
 algorithms~\cite{osera+:pldi15,frankle+:popl16,armando+:pldi16} are

### 2016-09-25 paper/extended-example.tex

@@
   that's the right sort of thing to do)?}
+\bcp{I suggest rearranging things a bit.  The problem I see here is that
+  this section ends up duplicating material from section 3.  So why not put
+  it after 3?  I.e., take the example in figure 1 (and the text up to this
+  point) and put it in the intro.  Take the rest of this section, remove all
+the explanations of what lenses are (or move them top the current section 3)
+and move the rest to a new section after the current section 3 that just
+focuses on synthesis.}
 Once we have a synthesis specification, we need to find a lens that

### 2016-09-25 paper/introduction.tex

@@
 
-Bidirectional programming is used heavily in practice.
-Augeas~\cite{augeas}, a configuration editing system for linux, uses lens combinators
+Bidirectional programming is used heavily in practice.\bcp{This may be
+  overstating the case, for lenses per se.  Augeas is the only real-world
+  example AFAIK.}
+Augeas~\cite{?}, a configuration editing system for linux, uses lens combinators
 to provide editable views of configuration files.

### 2016-11-01 program/check-with-boomerang.py
moved files, added test

@@ -21,7 +21,7 @@ REPETITION_COUNT = 2
 
 def find_tests(root):
     tests = []
-    for path, dirs, files in os.walk(root):
+    for path, dirs, files in os.walk(root,followlinks=True):
         files = [(f[0], f[1]) for f in [splitext(f) for f in files]]
         tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
     return tests

### 2016-11-01 program/generate-data.py

@@ -13,7 +13,7 @@ import time
 TEST_EXT = '.ls'
 BASELINE_EXT = '.out'
 BASE_FLAGS = ["-iterativedeepenstrategy"]
-TIMEOUT_TIME = 8
+TIMEOUT_TIME = 60
 GENERATE_EXAMPLES_TIMEOUT_TIME = 60
 
 REPETITION_COUNT = 2
@@ -38,8 +38,8 @@ def gather_datum(prog, path, base, additional_flags, timeout):
     process_output = EasyProcess([prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)]).call(timeout=timeout)
     return (process_output.stdout,process_output.stderr)
 
-def gather_data(prog, path, base):
-    current_data = {"Test":join(path, base + TEST_EXT).replace("_","-")[6:]}
+def gather_data(rootlength, prog, path, base):
+    current_data = {"Test":join(path, base + TEST_EXT).replace("_","-")[rootlength:]}
     run_data = []
     timeout = False
     error = False
@@ -104,28 +104,25 @@ def gather_data(prog, path, base):
         expanded_run_data_transpose = transpose(expanded_run_data)
 	expanded_computation_time_col = [float(x) for x in expanded_run_data_transpose[0]]
 	current_data["ForceExpandTime"]="{:.5f}".format(sum(expanded_computation_time_col)/len(expanded_computation_time_col))
-    gen_exs_expanded_run_data = []
+    max_ex_run_data = []
     timeout = False
     error = False
-    for iteration in range(REPETITION_COUNT):
-    	(datum,err) = gather_datum(prog, path, base,['-forceexpand','-generatedexamples'],GENERATE_EXAMPLES_TIMEOUT_TIME)
-	if datum == "":
-            if err == "":
-		timeout = True
-            else:
-                print("error:"+err+"end")
-                error = True
-	    break
-	else:
-		gen_exs_expanded_run_data.append(datum.split(","))
-    if error:
-        current_data["ForceExpandExamplesRequired"]="Error"
-    elif timeout:
-	current_data["ForceExpandExamplesRequired"]="Timeout"
+    (datum,err) = gather_datum(prog, path, base,['-max_to_specify'],TIMEOUT_TIME)
+    if datum == "":
+        if err == "":
+	    timeout = True
+        else:
+            print(err)
+            print("there was an error")
+            error = True
     else:
-        gen_exs_expanded_run_data_transpose = transpose(gen_exs_expanded_run_data)
[…]

### 2016-11-11 program/generate-data.py
shortened paper

@@ -39,7 +39,7 @@ def gather_datum(prog, path, base, additional_flags, timeout):
     return (process_output.stdout,process_output.stderr)
 
 def gather_data(rootlength, prog, path, base):
-    current_data = {"Test":join(path, base + TEST_EXT).replace("_","-")[rootlength:]}
+    current_data = {"Test":join(path, base).replace("_","-")[rootlength:]}
     run_data = []
     timeout = False
     error = False
@@ -60,7 +60,10 @@ def gather_data(rootlength, prog, path, base):
     else:
         run_data_transpose = transpose(run_data)
 	computation_time_col = [float(x) for x in run_data_transpose[0]]
-	current_data["ComputationTime"]="{:.5f}".format(sum(computation_time_col)/len(computation_time_col))
+        ans = sum(computation_time_col)/len(computation_time_col)
+        ans = (int)(ans * 100)
+        ans = ans * 10
+	current_data["ComputationTime"]=ans if ans != 0 else "$<10$"
     generate_exs_run_data = []
     timeout = False
     error = False

### 2016-11-14 program/generate-data.py
merged master

@@ -14,7 +14,7 @@ TEST_EXT = '.ls'
 BASELINE_EXT = '.out'
 BASE_FLAGS = ["-iterativedeepenstrategy"]
 TIMEOUT_TIME = 60
-GENERATE_EXAMPLES_TIMEOUT_TIME = 60
+GENERATE_EXAMPLES_TIMEOUT_TIME = 6000
 
 REPETITION_COUNT = 2
 
@@ -126,7 +126,32 @@ def gather_data(rootlength, prog, path, base):
 	current_data["MaxExampleCount"]="Timeout"
     else:
 	current_data["MaxExampleCount"]=max_ex_run_data
+    spec_size_run_data = []
+    timeout = False
+    error = False
+    (datum,err) = gather_datum(prog, path, base,['-spec_size'],TIMEOUT_TIME)
+    if datum == "":
+        if err == "":
+	    timeout = True
+        else:
+            print(err)
+            print("there was an error")
+            error = True
+    else:
+	spec_size_run_data=int(datum)
+    if error:
+        current_data["SpecSize"]="Error"
+    elif timeout:
+	current_data["SpecSize"]="Timeout"
+    else:
+	current_data["SpecSize"]=spec_size_run_data
     return current_data
+
+def specsize_compare(x,y):
+    return x["SpecSize"]-y["SpecSize"]
+
+def sort_data(data):
+    return sorted(data,cmp=specsize_compare)
 
 def print_data(data):
     ensure_dir("generated_data/")
@@ -164,6 +189,7 @@ def main(args):
 		print(join(path, base + TEST_EXT).replace("_","-")[rootlength:])
 		current_data = gather_data(rootlength,prog, path, base)
                 data.append(current_data)
+            data = sort_data(data)
 	    print_data(data)
         else:
             path, filename = os.path.split(path)
@@ -172,6 +198,7 @@ def main(args):
                 print_usage(args)
             else:
                 data = gather_data(prog, path, base)
+                sort_data(data)
 		print_data([data])
     else:
         print_usage(args)

### 2017-01-28 paper/rebuttal.txt
add rebuttal rough draft

Thank you for your feedback.

Several reviewers asked who benefits from this technology or what problem
exactly what solved.  We will rewrite the introduction to our paper to clarify
the goals of our system.  Our primary response is that this technology is aimed
at programmers.  If one were to ask a random programmer on the street whether
they were familiar with and knew how to use regular expressions, roughly
95% (?) would say yes.  If one were to a random programmer on the street
whether they were familiar with lenses roughly 0% would say yes.  The central
idea is to exploit something that most programmers know very well (lenses) to
allow them to write complex programs in a language that most programmers do not
know well.  Second, lenses *are* tricky to program.  Here is an example of a
lens for translating between idealized bibtex and endnote author formats:

let bibtex_to_endnote_au
  : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
  del "author={"
    . ins "au - "
    . lens_swap (NAME . del ",") (lens_swap WSP NAME)* 
    . (del " and "
         . ins "\n au - "
         . lens_swap (NAME . del ",") (lens_swap WSP NAME)* )* 
    . del "}"

Note the existence of complex recursive swaps.  Programming this manually would
be complex, and programming this while conforming to the specifications of the
types even more so.

Other concerns stemmed from our lack of a formal treatment of our
implementation, in particular the function RigidSynth.  In previous drafts of
our paper, we had given a formal description of this algorithm.  Regrettably, we
had to find a way to shorten our paper, and chose to remove this formalization.
Even worse, we accidentally removed it both from our paper and the extended
report.  We will certainly be adding this formal description to the extended
report, and will look into ways to fit the formal description into the final
paper.

Concerns were also raised about our evaluation.  In particular, while we showed
averages and runtimes, these were not easily interpreted.  Some reviewers were
interested in a comparison to other, similar tools, and others were interested
in why our approach is necessary, with a comparison given to naive approaches.

Comparison to similar tools is difficult.  There are no existing tools that
fill the same bidirectional program synthesis.  Furthermore, an extended
comparison to similar tools is difficult because other tools are unable to
even express the type of complicated permutations we tackle!  For example,
FlashFill and FlashExtract are unable to express a large number of our
permutations.  We mention the fact that similar tools are unable to perform the
types of transformations we target, and find a more extended evaluation of this
would not serve to further this point much.

Why have we done these approaches, and not the simpler, naive approaches?
These concerns are valid, and we do not address them in the paper.
Our optimizations were derived from the inability to synthesize our lenses in
any reasonable amount of time with the more naive approaches.  For example,
a search procedure which traversed arbitrary equivalences was unable to
synthesize anything beyond simple microbenchmarks, even with the significant
simplification of not searching for potential compositions!  Without using our
priority queue, we were able to synthesize complicated examples when the data
formats were similar, but vastly different format specifications for the source
and target languages were too difficult caused the synthesis problem to became
intractable.

We fully agree that an evaluation highlighting these importances would
strengthen our paper significantly.  We plan to begin adding flags to disable
certain optimizations upon completion of this rebuttal.

We respond to individual concerns below

==========================
Review A:
==========================

- The implementation section, which describes choices and heuristics
  of the search, is not as clear as earlier parts of the paper.

Addressed above

- The synthesized lenses are not shown or discussed in much detail,
  nor are they compared to hand-written ones (when possible).

The synthesized lenses are fairly comprehensible.  The transformation from DNF
lenses to Boomerang lenses alone creates fairly unintelligible syntax.  This is
an issue, we would like for programmers to be able to validate the synthesized
programs through a manual inspection of them.  As such we apply a number of
optimizations, including, but not limited to:

1. Trivial sublenses are removed
2. Sublenses expressing identity transformations are turned into identities
3. Identity transformations are combined when possible
4. Distributed transformations are factored when possible

The previously shown lens, bibtex_to_endnote_au, is an autogenerated lens.
The primary difference we see between the autogenerated code, and what we would
write ourselves, is the repeated code of

ins "au - "
. lens_swap (NAME . del ",") (lens_swap WSP NAME)*

We personally do not find this to make the code significantly more difficult to
read.  However, adding this optimization is merely a rote application of common
subexpression elimination.

- The comparison and relationship to Flash Fill, clarified in the
  evaluation, is too broad in the motivating sections.

Addressed above

- The prevalence and limitations of bijective lenses (compared to
  less well-behaved ones) is not discussed.

There certainly are limitations to restricting ourselves to bijective lenses.
However, we chose the simplification of bijective lenses for this new research
area, but think there is a wealth of future work that can be done through an
investigation of less well-behaved lenses.

- How many of surveyed FlashFill examples were close to bijective? Were there
  only the 3 included in the evaluation?

Few of the surveyed FlashFill examples were close to bijective.  The tricks we
used to make our chosen 3 examples bijective - adding in extra fields to hold
bits of data - could be extended to non-bijective examples.  Our primary goal
with the use of FlashFill was to show that, when made bijective and given data
format descriptors, the problems FlashFill addresses become simple.

==========================
Review B:
==========================

- paper unclear in various (important) places

We unfortunately were had to be brief in our description of confluence,
bisimilarity, and definitional equivalence.  We go into much greater detail in
the appendix, and will make sure to refer to the appendix for a more detailed
description of these terms, and for a more formal understanding of their
importance.

- algorithm only partially presented

Addressed above

- I am not sure the remark about the implementation spanning '5515 lines' is as
positive as you make it sound at the beginning of Section 6. There is no
pseudocode for the main part of the algorithm, a function called RigidSynth.
This makes one wonder how complex it really is and, more worryingly, as a
result, there is also no correctness proof and no complexity analysis.

Addressed in part above.  Many of the lines go not to the core synthesis
algorithm, but for language transformations, pretty printing, lens
simplification, and DFA operations.  The portion of our code corresponding to
RigidSynth is just over 400 lines.

- some proofs are missing

Please refer to our appendix, where we go into a detailed proof of these
theorems.

- It would be nicer if runtimes were compared with other programs. For example,
  the related work section mentions FlashFill as 'one of the most prominent
  recent string transformation systems', so since part of the benchmarks are
  adapted from FlashFill benchmarks, it would have been interesting to compare
  the performance here. The authors also say that 'there are many other recent
  results showing how to synthesize functions from type-based specifications'
  and that they 'found we needed new techniques to manage the multidimensional
  search space generated by regular expression equivalences and other aspects of
  lens combinator languages'. I am missing either an experimental comparison
  with these other results or a more detailed explanation of why they would not
  be applicable to their problem.

Addressed above

==========================
Review C:
==========================

- The experimental section could be stronger. I understand that comparisons with
  other tools is not feasible in this setting, as there aren't any other tools
  that are solving this specific variety of synthesis. However, you could do
  some internal comparisons, for various values of the design decisions.

Discussed above

- The discussion of related work can be improved. It is true that there aren't
  other approaches that are solving the same exact problem. However, some
  aspects of your synthesis are shared with other approaches.

Agreed, while we use similar techniques, they are used in different ways,
identifying the similarities and noting the differences would strengthen our
related work.

==========================
Review D:
==========================

- The algorithmic contributions (particularly, type-directedness) are often
  unclear.

Discussed in part above.
RigidSynth is entirely type directed.  We use an ordering on the subcomponents
of the types to find which subcomponents must be mapped to each other.  After
these lenses are found, they are then combined into the final lens in a way
that keeps the lens well typed.

- Converting a regular expression to DNF may incur an exponential explosion (as
it is akin to determinizing an NFM). Have you encountered examples where that is
the case? Please state this shortcoming.

We have not encountered any examples in practice where this is a problem.
However, we include in our benchmarks cust/cap-prob and cust/2-cap-prob to
highlight this shortcoming.  Furthermore, we discuss the shortcoming in Speed of
Synthesis in the evaluation section.

- Have you considered using a SAT/SMT solver to search the space of
  permutations?

Because of our use of orderings, we are able to find the permutations in
O(n*log(n)) time.

- The experimental evaluation is limited and does not back the claimed
  contributions.  Similar to (a), what is the effect of type-directedness on
  the experimental results?

Discussed in part above.
We did not ever try an approach that was not type-directed, as we do not think
enumerative search followed by a pruning of ill typed terms would scale well.

- The fact that the avg # of examples is often 0 seems to indicate that the
  benchmarks are simple. This is perhaps because the language is very
  restricted; i.e., there are no user-defined lenses, only the core ones. Could
  you comment on that?

We discuss why the average number of examples is so low in Importance of
Examples in the evaluation section.  To highlight how complicated the input
specifications can be, the specification for the BibTex to
EndNote conversion problem - one of our 10 least complex input formats - and the
specification for cron - our most complex format - are given
below.  We are happy to provide all our specifications.

cust/bib_prob
      #use "base.decls"
      typedef NAME = UPPERCASE (LOWERCASE)*;;

      typedef LASTCOMMASTART = NAME "," (WSP NAME)*;;

      typedef STARTTOEND = (NAME WSP)* NAME;;

      typedef BIBTEXAUTHORLIST = LASTCOMMASTART (" and " LASTCOMMASTART)*;;
      typedef BIBTEXAUTHORINFO = "author={" BIBTEXAUTHORLIST "}";;

      typedef AUTAG = "au - ";;
      typedef TITAG = "ti - ";;
      typedef JOTAG = "jo - ";;

      typedef TAGGEDAUTHORDEFNS = AUTAG STARTTOEND ("\n " AUTAG STARTTOEND)*;;

      typedef TITLE = NAME (WSP NAME)*;;
      typedef BIBTEXTITLE = "title={" TITLE "}";;
      typedef TAGGEDTITLE = TITAG TITLE;;

      typedef JOURNAL = NAME (WSP NAME)*;;
      typedef BIBTEXJOURNAL = "journal={" JOURNAL "}";;
      typedef TAGGEDJOURNAL = JOTAG JOURNAL;;

      typedef FULLBIBTEX = "{" ((BIBTEXJOURNAL | BIBTEXAUTHORINFO | BIBTEXTITLE)",")* "}";;
      typedef FULLTAGS = . | ((TAGGEDAUTHORDEFNS | TAGGEDTITLE | TAGGEDJOURNAL)
      (("\n" (TAGGEDAUTHORDEFNS | TAGGEDTITLE | TAGGEDJOURNAL))*)) ;;

      bibtex_to_readable_au = [BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS
      {"author={Foster, Nathan and Pierce, Benjamin and Bohannon, Aaron}" <->
      "au - Nathan Foster
       au - Benjamin Pierce
       au - Aaron Bohannon"}]

      bibtext_to_readable_title = [BIBTEXTITLE <=> TAGGEDTITLE
      {"title={Boomerang Resourceful Lenses For String Data}" <->
       "ti - Boomerang Resourceful Lenses For String Data"}]

      journal_to_readable_journal = [BIBTEXJOURNAL <=> TAGGEDJOURNAL
      {"journal={Principals Of Programming Languages}" <->
       "jo - Principals Of Programming Languages"}]

      bibtext_to_tagged_tester= [FULLBIBTEX <=> FULLTAGS {
      "{author={Foster, Nathan and Pierce, Benjamin and Bohannon, Aaron},title={Boomerang Resourceful Lenses For String Data},journal={Principals Of Programming Languages},}"
      <->
      "au - Nathan Foster
       au - Benjamin Pierce
       au - Aaron Bohannon
      ti - Boomerang Resourceful Lenses For String Data
      jo - Principals Of Programming Languages"
      }]

      test bibtext_to_tagged_tester
      {"{author={Foster, Nathan and Pierce, Benjamin and Bohannon, Aaron},}"
      <->
      "au - Nathan Foster
       au - Benjamin Pierce
       au - Aaron Bohannon",

      "{title={Boomerang Resourceful Lenses For String Data},}"
      <->
      "ti - Boomerang Resourceful Lenses For String Data",

      "{journal={Principals Of Programming Languages},}"
      <->
      "jo - Principals Of Programming Languages"
      };;

aug/cron
      #use "base.decls"
      #use "util.decls"

      typedef INDENT = (" " | "\t")*;;
      typedef INDENT_REQ = (" " | "\t")+;;
      typedef ALPHANUM = (UPPERCASE | LOWERCASE | DIGIT)+;;
      typedef RANGE = (ALPHANUM "-" ALPHANUM | ALPHANUM );;
      typedef PREFIX = "-";;

      typedef SCHEDULE_VALUE = "reboot" | "yearly" | "annually" | "monthly"
                             | "weekly" | "daily" | "midnight" | "hourly";;
      typedef SCHEDULE = "@" SCHEDULE_VALUE;;

      typedef USER = (UPPERCASE | LOWERCASE | DIGIT)+;;

      typedef TIME = NUMBER INDENT_REQ NUMBER INDENT_REQ NUMBER INDENT_REQ RANGE INDENT_REQ RANGE;;

      typedef SHELLCOMMAND_CHAR = LOWERCASE | UPPERCASE | DIGIT | "_" | "/" | "|"  | "." ;;
      typedef SC_CHAR_OR_SPACE = LOWERCASE | UPPERCASE | DIGIT | "_" | "/" | "|" | "." | " " ;;
      typedef SHELLCOMMAND = (SHELLCOMMAND_CHAR (SC_CHAR_OR_SPACE)* SHELLCOMMAND_CHAR) | SHELLCOMMAND_CHAR;;

      typedef SHELLVAR_CHAR = LOWERCASE | UPPERCASE | DIGIT | "_";; 
      typedef SHELLVAR_NAME = SHELLVAR_CHAR+;;
      typedef SHELLVALUE_CHAR = LOWERCASE | UPPERCASE | DIGIT | "_" | "/" | "|" | ".";;
      typedef SHELLVALUE_NAME = SHELLVALUE_CHAR+;;

      typedef SHELLVAR = SHELLVAR_NAME "=" SHELLVALUE_NAME "\n";;
      typedef COMMENTLINE = COMMENT "\n";;
      typedef ENTRY = INDENT (PREFIX | . ) (TIME | SCHEDULE) INDENT_REQ USER INDENT_REQ SHELLCOMMAND "\n";;
      typedef CRON = ( "\n" | SHELLVAR | COMMENTLINE | ENTRY)*;;

      typedef PREFIX_DICT = "{\"prefix\"=" ("true" | "false") "}";;
      typedef TIME_DICT = "{\"minute\"=" NUMBER ",\"ws1\"=" INDENT_REQ ",\"hour\"=" NUMBER 
        ",\"ws2\"=" INDENT_REQ ",\"dayofmonth\"=" NUMBER ",\"ws3\"=" INDENT_REQ 
        ",\"month\"=" RANGE ",\"ws4\"=" INDENT_REQ ",\"dayofweek\"=" RANGE "}";;
      typedef SCHEDULE_DICT = "{\"schedule\"=\"" SCHEDULE_VALUE "\"}";;
      typedef ENTRY_DICT = "{\"indent\"=\"" INDENT "\"," PREFIX_DICT "," (TIME_DICT | SCHEDULE_DICT)
        ",\"indent2\"=\"" INDENT_REQ "\",\"user\"=\"" USER "\",\"indent3\"=\""
        INDENT_REQ "\",\"command\"=\"" SHELLCOMMAND "\"}";;
      typedef SHELL_DICT = "{\"varname\"=\"" SHELLVAR_NAME "\",\"value\"=\"" SHELLVALUE_NAME "\"}";;
      typedef CRON_DICT = ((EMPTYDICT | SHELL_DICT | COMMENT_DICT | ENTRY_DICT) "\n")*;;

      cron_lens = [CRON_DICT <=> CRON {}]

==========================
Review E:
==========================

- The current presentation doesn’t explain what are the difficult challenges for
  this problem. Given the source and target regular expressions, and especially
  with the strongly unambiguous constraint, it seems a simple enumerative search
  should work well. I was wondering if the authors tried simpler enumerative
  solvers baselines for this problem?

Addressed above

- The second big concern is that it was not obvious who is the target audience
  for this system. If we are expecting a user to provide the regular expressions
  for source and target formats, why can’t they also write the transformation as
  well? Can the authors provide the Boomerang program for the Bibtex example and
  explain why it would be challenging for a user to write this Boomerang program
  if they can write regular expressions for the two formats.

Addressed above

- For the Bibtex to EndNote transformation example, the paper presents an
  algorithmic way to align the regular expressions to find corresponding
  permutations of sub-expressions to match different fields. Why can’t the user
  provide regular expressions in first place that are aligned? For example,
  there is an epsilon in the Bibtex regular expression, but there is no epsilon
  in the EndNote format. Wouldn’t it be a more clear and a better specification
  on user’s part to write a regular expression where EndNode format also has an
  epsilon with an Or in front. For finding permutations as well, it isn’t clear
  why can’t the user simply annotate or label different fields using some user
  interface, rather than the algorithm having to search over different matching
  of sub-expressions.

We feel that requiring well aligned regular expressions is too large a burden on
the user.  The user themselves needs to think through various equivalence rules,
which we think is better suited to an automated tool.

The idea of adding in the ability to annotate fields is an interesting one.
The way that examples are embedded in the regular expressions is similar to an
annotation.  In certain scenarios, it may also be easier to give annotations
than to give a covering example.  Likely the best solution would be one which
uses both annotations and examples to help differentiate similar fields.

- Is there a restriction that the user-provided regular expressions need to
  parse the source and target formats unambiguously? For example, if a user
  provides the regular expression for a format as Name* Name *, and the input is
  John Conway, would the system disallow such user inputs?

There is such a restriction (though we allow for ambiguity when hidden under
user defined regular expressions), but the tool does not currently check that
this restriction is held.  Methods for preforming these checks are well studied,
and described in Boomerang as well as in Nate Foster's thesis.  An end goal for
this tool is to be integrated with the Boomerang codebase, and used as part of
the development for Boomerang programs.  Such a production quality version of
this tool would certainly provide these capabilities.

- The synthesis algorithm section presents several heuristics to assign scores
  and pseudometric scores to a pair of DNF regular expressions. How were these
  heuristics chosen? What is the effect of these heuristics on the benchmark
  problems?

Discussed in part above.  These heuristics were chosen in a way to leverage
the abstractions that programmers use for their specifications, while also
allowing enough breadth to the search space to not get stuck in local minima.

==========================
Review F:
==========================

- First, who are the intended users of the proposed tool? Are they end users
  who might not have technical knowledge about data transformation programs? Or
  are they technical users who are unfamiliar with lens programs? I initially
  thought the tool is intended for the first type of users, but as I read
  through the paper it seems apparent that the authors intend to target the
  second group, especially with the requirement that users need to provide a
  regular expression description of the input/output data format. The
  requirement to do so seems to be the weakest point of the paper, especially
  given that existing tools (such as FlashFill or Dataplay [1]) do not have such
  requirement. If the tool is indeed intended for users who are not familiar
  with lens programming, then can the authors justify why lens is the
  appropriate programming model for data transformation? Wouldn't it be easier
  to dump the input/output data into a common format (say CSV, spreadsheet, or
  JSON), and simply use FlashFill or TriFacta [2] to learn the desired
  transformation program rather than providing the ? I understand that it might
  be difficult for tools like FlashFill to learn nested constructs (given its
  relational data model), but to me that doesn't quite justify the need for
  providing regular expressions as part of the input.

Discussed above

- Second, the evaluations seem weak.  The examples from FlashFill and are those
  that can already be synthesized using FlashFill by just providing input /
  output examples. Also, can the authors describe how complex are the regular
  expressions that users need to provide? Given that the same data format can be
  described using different semantically equivalent expressions, how do they
  affect the effectiveness of the synthesis process?

Discussed in part above.  Semantically equivalent expressions can have an impact
on the effectiveness of the synthesis process.  As we discuss in Speed of
Synthesis, in the Evaluation section, the difference in the presentation of the
regular expressions being mapped between is a stronger indicator of complexity
for this system.

- Finally, the real world contains many erroneous / dirty data. What happens if
  the user provide the incorrect regular expression or inconsistent input /
  output examples?

If there is possibly dirty data, then we expect this dirty data to not be
input to the transformer through the typing.  This would allow for the
unexpected data to fail, instead of silently succeed with unexpected results.
If this odd data is found to be an expected occurence, changes to the typing can
be made to address it, and a new lens can be synthesized.

- Can the authors describe what kind of examples were needed to
  achieve the efficient synthesis results? In light of the "Importance of
  Examples" subsection on p.11, it seems that choosing "clever" examples is one
  of the key requirement for the tool to succeed in synthesizing a valid
  program. In addition, is there some way for the tool to identify the error and
  perform iterative synthesis? Also, what happens if multiple lens program are
  found for a given input?

Clever examples do not need to be provided to the system.  The "Avg # Exmpl"
column gets its value from randomly generated examples, as we did not want our
implicit understanding of the synthesis algorithm to make our system look like
it needed fewer examples than it does when used by a non-expert.  When
multiple programs are possible, the one that requires the least significant
permutation is chosen.  The "Max # Exmpl" column corresponds to a permutation
generated which chooses a wrong permutation, if such a permutation is
permutation is possible.

### 2017-01-28 paper/rebuttal.txt
revised shared portion of rebuttal

@@
-Thank you for your feedback.
+Thank you for your detailed feedback.  We will incorporate your many
+useful suggestions when we revise our paper.
 
-Several reviewers asked who benefits from this technology or what problem
-exactly what solved.  We will rewrite the introduction to our paper to clarify
-the goals of our system.  Our primary response is that this technology is aimed
-at programmers.  If one were to ask a random programmer on the street whether
-they were familiar with and knew how to use regular expressions, roughly
-95% (?) would say yes.  If one were to a random programmer on the street
-whether they were familiar with lenses roughly 0% would say yes.  The central
-idea is to exploit something that most programmers know very well (lenses) to
-allow them to write complex programs in a language that most programmers do not
-know well.  Second, lenses *are* tricky to program.  Here is an example of a
-lens for translating between idealized bibtex and endnote author formats:
+Several reviewers asked who benefits from this technology.   The
+brief answer is that this technology is aimed at programmers.  If one
+were to ask random programmers whether they know how to use regular
+expressions, most would say yes.  In contrast, almost none of them
+would know how to use lenses. The central idea of this paper is to
+exploit something that most programmers know well (regular exprssions)
+to help them write complex programs in a language that they do not
+know well (lenses).  This help is particularly important because
+lenses are *tricky* to write.  For example, consider the following
+lens for translating between idealized bibtex and endnote author
+formats:
 
@@
 
-Note the existence of complex recursive swaps.  Programming this manually would
-be complex, and programming this while conforming to the specifications of the
-types even more so.
+Writing this lens from scratch is difficult, and it is a relatively
+simple lens! We will rewrite the introduction to clarify the goals of
+our work.
 
-Other concerns stemmed from our lack of a formal treatment of our
-implementation, in particular the function RigidSynth.  In previous drafts of
-our paper, we had given a formal description of this algorithm.  Regrettably, we
-had to find a way to shorten our paper, and chose to remove this formalization.
-Even worse, we accidentally removed it both from our paper and the extended
-report.  We will certainly be adding this formal description to the extended
-report, and will look into ways to fit the formal description into the final
-paper.
+A second concern stemmed from the lack of a formal treatment of our
+implementation, in particular the function RigidSynth. In earlier
+drafts, we had a formal description of this algorithm.  Regrettably,
+we removed it to shorten the paper to fit the page limit.  We
+inadvertently removed it from the extended report at the same time.
+We will reinstate the full formal description in the extended report
+and find ways to summarize it precisely in the final paper.
 
-Concerns were also raised about our evaluation.  In particular, while we showed
-averages and runtimes, these were not easily interpreted.  Some reviewers were
-interested in a comparison to other, similar tools, and others were interested
-in why our approach is necessary, with a comparison given to naive approaches.
+A third area of concern involved aspects of our evaluation,
+specifically comparisons with other tools and measurements to evaluate
+the heuristics we chose in searching for lenses.
 
-Comparison to similar tools is difficult.  There are no existing tools that
-fill the same bidirectional program synthesis.  Furthermore, an extended
-comparison to similar tools is difficult because other tools are unable to
-even express the type of complicated permutations we tackle!  For example,
-FlashFill and FlashExtract are unable to express a large number of our
-permutations.  We mention the fact that similar tools are unable to perform the
-types of transformations we target, and find a more extended evaluation of this
-would not serve to further this point much.
+A direct comparison with related tools is not possible because there
+are no other tools that do bidirectional program synthesis.
+Performance comparisons to similar tools such as FlashFill and
+FlashExtract  are difficult because they cannot express the rich set
+of permutations we tackle.
 
-Why have we done these approaches, and not the simpler, naive approaches?
-These concerns are valid, and we do not address them in the paper.
-Our optimizations were derived from the inability to synthesize our lenses in
-any reasonable amount of time with the more naive approaches.  For example,
-a search procedure which traversed arbitrary equivalences was unable to
-synthesize anything beyond simple microbenchmarks, even with the significant
-simplification of not searching for potential compositions!  Without using our
-priority queue, we were able to synthesize complicated examples when the data
-formats were similar, but vastly different format specifications for the source
-and target languages were too difficult caused the synthesis problem to became
-intractable.
+As for our choice of heuristics, we arrived at them when more naive
+approaches failed.  For example, our initial search procedure, which
+traversed arbitrary equivalences, could not synthesize anything beyond
+simple microbenchmarks, even with the significant simplification of
+not searching for potential compositions!  Without a priority queue,
+we were able to synthesize complicated lenses when the source and
+target formats were similar, but not when the formats differed
+significantly. We agree that performance measurements showing the
+importance of these aspects of our algorithm would improve the paper,
+and we will work to add them.
 
-We fully agree that an evaluation highlighting these importances would
-strengthen our paper significantly.  We plan to begin adding flags to disable
-certain optimizations upon completion of this rebuttal.
-
-We respond to individual concerns below
+In the remainder, we respond to individual concerns not already addressed.
 
@@
 ==========================
-
-- The implementation section, which describes choices and heuristics
-  of the search, is not as clear as earlier parts of the paper.
-
-Addressed above
 

### 2017-01-28 paper/rebuttal.txt
Checkpoint work on response

@@
-Thank you for your detailed feedback.  We will incorporate your many
-useful suggestions when we revise our paper.
+TARGET AUDIENCE
 
-Several reviewers asked who benefits from this technology.   The
-brief answer is that this technology is aimed at programmers.  If one
-were to ask random programmers whether they know how to use regular
-expressions, most would say yes.  In contrast, almost none of them
-would know how to use lenses. The central idea of this paper is to
-exploit something that most programmers know well (regular exprssions)
-to help them write complex programs in a language that they do not
-know well (lenses).  This help is particularly important because
-lenses are *tricky* to write.  For example, consider the following
-lens for translating between idealized bibtex and endnote author
-formats:
+Several reviewers asked who benefits from this technology.  Briefly, the
+technology is aimed at programmers [BCP: Do we really think the reviewers
+might not have got this??  Or do we mean something more focused?].  The
+central idea is to exploit something that most programmers know well
+(regular expressions) to help them write significant programs in a language
+that they do not know well (lenses).  This help is particularly important
+because past experience has shown that lenses are often quite tricky to
+write.  For example, consider the following lens for translating between
+idealized bibtex and endnote author formats:
 
@@
 
-Writing this lens from scratch is difficult, and it is a relatively
-simple lens! We will rewrite the introduction to clarify the goals of
-our work.
+Writing this lens from scratch is nontrivial (mainly because of all the
+swaps), and it is a relatively simple lens -- Augeas users regularly work
+with lenses XXX orders of magnitude this size.
+
+FORMALIZATION
 
 A second concern stemmed from the lack of a formal treatment of our
-implementation, in particular the function RigidSynth. In earlier
-drafts, we had a formal description of this algorithm.  Regrettably,
-we removed it to shorten the paper to fit the page limit.  We
-inadvertently removed it from the extended report at the same time.
-We will reinstate the full formal description in the extended report
-and find ways to summarize it precisely in the final paper.
+implementation, in particular the function RigidSynth. In earlier drafts, we
+actually had a formal description of this algorithm, but (in retrospect
+regrettably) we removed it to get under the page limit and inadvertently
+removed it from the extended report at the same time.  We will reinstate the
+full formal description in the extended report and find a good way to
+summarize it precisely in the final paper.
+
+[BCP: Is there anything we can say about the formal description?  E.g., a
+few words about how large or complex it is?  Any theorems we've proved about
+it?]
+
+EVALUATION
 
@@
 
-In the remainder, we respond to individual concerns not already addressed.
+We will clarify all these points in the next rewrite.
+
+===========================================================================
+===========================================================================
+The remainder of this document -- to be read only as needed -- responds in
+detail to individual concerns not already addressed.
 

### 2017-01-28 paper/rebuttal.txt
Checkpoint again

@@
 Writing this lens from scratch is nontrivial (mainly because of all the
-swaps), and it is a relatively simple lens -- Augeas users regularly work
-with lenses XXX orders of magnitude this size.
+swaps), and it is a relatively simple lens -- Augeas and Boomerang users
+regularly work with lenses XXX orders of magnitude [BCP: quantify, maybe
+with a couple examples.] this size.
 
@@
 
-A second concern stemmed from the lack of a formal treatment of our
-implementation, in particular the function RigidSynth. In earlier drafts, we
-actually had a formal description of this algorithm, but (in retrospect
-regrettably) we removed it to get under the page limit and inadvertently
-removed it from the extended report at the same time.  We will reinstate the
-full formal description in the extended report and find a good way to
-summarize it precisely in the final paper.
+A second major concern was the lack of a formal treatment of our
+implementation, in particular the function RigidSynth.
+
+In earlier drafts, we actually had a formal description of this algorithm,
+but (in retrospect regrettably) we removed it to get under the page limit
+and inadvertently removed it from the extended version at the same time.  We
+will reinstate the full treatment in the extended report and find a good way
+to summarize it in the final paper.
 
@@
 
-A third area of concern involved aspects of our evaluation,
-specifically comparisons with other tools and measurements to evaluate
-the heuristics we chose in searching for lenses.
+A third area of concern was our evaluation, specifically the lack of
+comparisons with other tools or a quantitative evaluation of the heuristics
+we chose in searching for lenses.
 
-A direct comparison with related tools is not possible because there
-are no other tools that do bidirectional program synthesis.
-Performance comparisons to similar tools such as FlashFill and
-FlashExtract  are difficult because they cannot express the rich set
-of permutations we tackle.
+Direct comparison with related tools is difficult because there aren't any
+tools that do the same task: this is the first attempt at bidirectional
+program synthesis.  Even performance comparisons to similar tools such as
+FlashFill and FlashExtract are difficult because they cannot express the
+rich set of permutations we tackle.  [BCP: This paragraph feels a bit
+argumentative.  Aren't there ANY quantitative comparisons that we could do,
+even if they didn't teach us too much?]
 
-As for our choice of heuristics, we arrived at them when more naive
-approaches failed.  For example, our initial search procedure, which
+On the other hand, we agree that performance measurements quantifying the
+importance of these aspects of the various heuristics used by the algorithm
+would improve the paper.  It will not be hard to instrument the
+implementation so that we can disable heuristics individually and gather
+these numbers.  (Anecdotally, we invented each of these heuristics when more
+naive approaches failed.  For example, our initial search procedure, which
 traversed arbitrary equivalences, could not synthesize anything beyond
-simple microbenchmarks, even with the significant simplification of
-not searching for potential compositions!  Without a priority queue,
-we were able to synthesize complicated lenses when the source and
-target formats were similar, but not when the formats differed
-significantly. We agree that performance measurements showing the
-importance of these aspects of our algorithm would improve the paper,
-and we will work to add them.
+simple microbenchmarks, even with the significant simplification of not
+searching for potential compositions!  Without a priority queue, we were
+able to synthesize complicated lenses when the source and target formats
+were similar, but not when the formats differed significantly.)
 
@@
 
-The synthesized lenses are fairly comprehensible.  The transformation from DNF
-lenses to Boomerang lenses alone creates fairly unintelligible syntax.  This is
-an issue, we would like for programmers to be able to validate the synthesized
-programs through a manual inspection of them.  As such we apply a number of
-optimizations, including, but not limited to:
+Our implementation puts some work into making the synthesized lenses
+readable.  The transformation from DNF lenses to Boomerang lenses alone
+creates fairly unintelligible syntax.  This is an issue, we would like for
+programmers to be able to validate the synthesized programs through a manual
+inspection of them.  As such we apply a number of optimizations, including,
+but not limited to:
 

### 2017-01-28 paper/rebuttal.txt
Release lock

@@
 technology is aimed at programmers [BCP: Do we really think the reviewers
-might not have got this??  Or do we mean something more focused?].  The
-central idea is to exploit something that most programmers know well
-(regular expressions) to help them write significant programs in a language
-that they do not know well (lenses).  This help is particularly important
-because past experience has shown that lenses are often quite tricky to
-write.  For example, consider the following lens for translating between
-idealized bibtex and endnote author formats:
+might not have got this??  Or do we mean something more focused?  E.g.,
+Review E is not really asking who the target audience is.  Nor is review F.
+Moreover, we don't really address review F's concern -- that providing
+regular expressions is already onerous.  Indeed, we agree with this, and
+ultimately we hope to synthesize those too; what we're reporting here is
+addressing what seemed like the hardest and most novel piece of the problem;
+inferring regularly expressions is a separable, and previously studied,
+issue.].  The central idea is to exploit something that most programmers
+know well (regular expressions) to help them write significant programs in a
+language that they do not know well (lenses).  This help is particularly
+important because past experience has shown that lenses are often quite
+tricky to write.  For example, consider the following lens for translating
+between idealized bibtex and endnote author formats:
 
@@
 were similar, but not when the formats differed significantly.)
+
+MORE THINGS TO DISCUSS
+
+[BCP: Here are some other points in the reviews that I thought were
+important enough to deserve discussion in the "above the fold" part of the
+response...
+
+* the current approach works for the subset of lenses that are bijective. It
+  would be good to give a sense about how much this limits its use among,
+  say, existing Augeas programs. And is there reason to believe that the
+  approach will or will not be amenable for larger subsets of Boomerang?
 
@@
 Discussed in part above.
+
+[BCP: Do you mean "above" as in the very top section?  Or somewhere else in
+the specific review comments?  I doubt if anybody is going to read all our
+responses to all the reviews, so best to make the pointers more precise --
+"See response to review A..."]
+
 RigidSynth is entirely type directed.  We use an ordering on the subcomponents
@@
 
+[BCP: Give an example!]
+
 The idea of adding in the ability to annotate fields is an interesting one.
@@
   parse the source and target formats unambiguously? For example, if a user
-  provides the regular expression for a format as Name* Name *, and the input is
+  provides the regular expression for a format as Name* Name*, and the input is
   John Conway, would the system disallow such user inputs?
@@
 user defined regular expressions), but the tool does not currently check that
-this restriction is held.  Methods for preforming these checks are well studied,
+this restriction is satisfied.
+
+[BCP: Really???  It should, but in any case I think we don't need to mention
+that it doesn't here!]
+
+Methods for performing these checks are well studied,
 and described in Boomerang as well as in Nate Foster's thesis.  An end goal for
@@
 this tool would certainly provide these capabilities.
+
+[BCP: Doesn't answer the question very clearly.]
 

### 2017-01-28 paper/rebuttal.txt
react to feedback on rebuttal

--- 2017-01-28 paper/rebuttal.txt
@@
 Several reviewers asked who benefits from this technology.  Briefly, the
-technology is aimed at programmers [BCP: Do we really think the reviewers
-might not have got this??  Or do we mean something more focused?].  The
-central idea is to exploit something that most programmers know well
-(regular expressions) to help them write significant programs in a language
-that they do not know well (lenses).  This help is particularly important
-because past experience has shown that lenses are often quite tricky to
-write.  For example, consider the following lens for translating between
-idealized bibtex and endnote author formats:
+technology is aimed at programmers, to be used as a code writing tool.
+Previous work on data wranglers works on easily transforming data, which
+can then be analyzed manually.  This work aims to provide a tool which can
+assist in writing product quality data transformers.
+
+For this work, the central idea is to exploit something that most programmers
+know well (regular expressions) to help them write significant programs in a
+language that they do not know well (lenses).  This help is particularly
+important because past experience has shown that lenses are often quite
+tricky to write.  For example, consider the following lens for translating
+between idealized BibTeX and EndNote author formats:
 
@@
 swaps), and it is a relatively simple lens -- Augeas and Boomerang users
-regularly work with lenses XXX orders of magnitude [BCP: quantify, maybe
-with a couple examples.] this size.
+regularly work with lenses significantly larger than this.  For example,
+Boomerang's full version of an idealized BibTeX converter is over 300 lines of
+code, and all but 4 of the studied Augeas programs were over 50 lines of code,
+with 4 over 100 lines of code.
+
+Furthermore, we agree that providing regular expressions does require the
+programmers to do additional work.  While we would eventually like to find ways
+to encorporate our work with existing research on inferring regular
+expressions, we wanted to initially address the novel portion of this eventual
+goal.  We think that even without the inference of regular expressions, this
+work already is useful as a computer assisted bidirectional programming tool.
 
@@
 
-[BCP: Is there anything we can say about the formal description?  E.g., a
-few words about how large or complex it is?  Any theorems we've proved about
-it?]
+The formal description is not very complex, but takes a good amount of space to
+complete.  The formal description of RigidSynth first describes an ordering on
+atoms, sequences, and DNF regular expressions.  Next, this ordering is used to
+define a normalization procedure for DNF regular expressions.  Finally, a
+procedure for "zipping" two DNF regular expressions together is provided.
+If these two DNF regular expressions can be zipped, then there is a lens between
+them.  A transformation from zipped DNF regular expressions and the
+permutations that normalize them is then provided. These formalizations are not
+very complex, but they take up a lot of space.
 
@@
 FlashFill and FlashExtract are difficult because they cannot express the
-rich set of permutations we tackle.  [BCP: This paragraph feels a bit
-argumentative.  Aren't there ANY quantitative comparisons that we could do,
-even if they didn't teach us too much?]
+rich set of permutations we tackle.  We recognize that a quantification of this
+increase in expressibility would be important for readers of our paper.  We
+think that we should provide an analysis detailing how many of our programs are
+inexpressible using other string synthesis tools.
 
-On the other hand, we agree that performance measurements quantifying the
+Furthermore, we strongly agree that performance measurements quantifying the
 importance of these aspects of the various heuristics used by the algorithm
-would improve the paper.  It will not be hard to instrument the
+would greatly improve the paper.  It will not be hard to instrument the
 implementation so that we can disable heuristics individually and gather
@@
 were similar, but not when the formats differed significantly.)
+
+BIJECTIVE LENS RESTRICTION
+
+Our program is unable to synthesize general Boomerang lenses, but rather only
+those which express bijections.
+There certainly are limitations to restricting ourselves to bijective lenses.
+Indeed, while all existing Boomerang and Augeas lenses can be transformed into
+bijective lenses, few are bijective by default, often because of a lack of
+bijectivity guarantees on whitespace.
+We chose to restrict our domain to bijective programs as a simplification
+technique for this new research, but think there are many directions this work
+can be expanded to, and we feel this research is amenable to be expanded.
+
+In particular, bijective quotient lenses, lenses which are bijective modulo an
+equivalence on the source and target languages, are an expansion to this work
+that already has shown promise.  Merely adding quotienting makes this tool more
+practical.  XSugar - an existing, powerful language for XML and string
+bidirectional transformations - expresses bijections modulo equivalences.
+While we have only seen preliminary results so far, it seems that only minor
+modifications were needed for our tool to generalize to bijective quotient
+lenses.
 
@@
 
-Addressed above
+Discussed in the general section.
 
@@
 
-Addressed in part above.  Many of the lines go not to the core synthesis
-algorithm, but for language transformations, pretty printing, lens
+Discussed in the general section.  Many of the lines go not to the core
+synthesis algorithm, but for language transformations, pretty printing, lens
 simplification, and DFA operations.  The portion of our code corresponding to
@@
 
-Discussed above
+Discussed in the general section.
 
@@
 
-Discussed in part above.
+Discussed in part in the general section.
+
 RigidSynth is entirely type directed.  We use an ordering on the subcomponents
@@
 
-Discussed in part above.
+Discussed in part in the general section.
 We did not ever try an approach that was not type-directed, as we do not think
@@
 Examples in the evaluation section.  To highlight how complicated the input
-specifications can be, the specification for the BibTex to
+specifications can be, the specification for the BibTeX to
 EndNote conversion problem - one of our 10 least complex input formats - and the
@@
   for source and target formats, why can’t they also write the transformation as
-  well? Can the authors provide the Boomerang program for the Bibtex example and
+  well? Can the authors provide the Boomerang program for the BibTeX example and
   explain why it would be challenging for a user to write this Boomerang program
@@
 
-Addressed above
+Discussed in the general section.
 
-- For the Bibtex to EndNote transformation example, the paper presents an
+- For the BibTeX to EndNote transformation example, the paper presents an
   algorithmic way to align the regular expressions to find corresponding
@@
   provide regular expressions in first place that are aligned? For example,
-  there is an epsilon in the Bibtex regular expression, but there is no epsilon
+  there is an epsilon in the BibTeX regular expression, but there is no epsilon
   in the EndNote format. Wouldn’t it be a more clear and a better specification
@@
 
+This is very difficult when the equivalences that need to be exploited are
+nonobvious or hidden away.  Consider the two regular expressions (A|B)C*
+and (A|C(C*)A)|(B|B(C*)C).  It is nontrivial for a user to identify that the
+equivalences that must be applied to reach two well aligned lenses.
+Furthermore, the aligned regular expressions could become difficult to read and
+maintain: (A|B)(e+C+CC*C) is less understandable than (A|B)C*.
+
 The idea of adding in the ability to annotate fields is an interesting one.
@@
   parse the source and target formats unambiguously? For example, if a user
-  provides the regular expression for a format as Name* Name *, and the input is
+  provides the regular expression for a format as Name* Name*, and the input is
   John Conway, would the system disallow such user inputs?
 
-There is such a restriction (though we allow for ambiguity when hidden under
-user defined regular expressions), but the tool does not currently check that
-this restriction is held.  Methods for preforming these checks are well studied,
-and described in Boomerang as well as in Nate Foster's thesis.  An end goal for
-this tool is to be integrated with the Boomerang codebase, and used as part of
-the development for Boomerang programs.  Such a production quality version of
-this tool would certainly provide these capabilities.
+The regular expressions provided to the synthesis algorithm must be unambiguous.
+However, we provide machinery to allow ambiguous user-defined regular
+expressions to safely be subexpressions of the input format descriptors.
+Using the keyword "abstract" on the definition of a user-defined regular
+expression allows these abstract regular expressions to never be expanded into
+their subcomponents in the search procedure.  With this opacity, ambiguous
+regular expressions can safely be subexpressions of an input format descriptor,
+though only the identity transformation can act on them.
 
@@
 
-Discussed in part above.  These heuristics were chosen in a way to leverage
-the abstractions that programmers use for their specifications, while also
-allowing enough breadth to the search space to not get stuck in local minima.
+Discussed in part in the general section. These heuristics were chosen in a way
+to leverage the abstractions that programmers use for their specifications,
+while also allowing enough breadth to the search space to not get stuck in
+local minima.
 
@@
 
-Discussed above
+Discussed in part in the general section.
 
@@
 
-Discussed in part above.  Semantically equivalent expressions can have an impact
-on the effectiveness of the synthesis process.  As we discuss in Speed of
-Synthesis, in the Evaluation section, the difference in the presentation of the
-regular expressions being mapped between is a stronger indicator of complexity
-for this system.
+Discussed in part in the general section. Semantically equivalent expressions
+can have an impact on the effectiveness of the synthesis process.  As we discuss
+in Speed of Synthesis, in the Evaluation section, the difference in the
+presentation of the regular expressions being mapped between is a stronger
+indicator of complexity for this system.
 

### 2017-01-28 paper/rebuttal.txt
removed useless blocks

@@
 
-- The comparison and relationship to Flash Fill, clarified in the
-  evaluation, is too broad in the motivating sections.
-
-Discussed in the general section.
-
 - The prevalence and limitations of bijective lenses (compared to
@@
 
-- algorithm only partially presented
-
-Discussed in the general section.
-
 - I am not sure the remark about the implementation spanning '5515 lines' is as
@@
 
-Discussed in the general section.  Many of the lines go not to the core
-synthesis algorithm, but for language transformations, pretty printing, lens
-simplification, and DFA operations.  The portion of our code corresponding to
-RigidSynth is just over 400 lines.
+Discussed in the part in the general section.  Many of the lines go not to the
+core synthesis algorithm, but for language transformations, pretty printing,
+lens simplification, and DFA operations.  The portion of our code corresponding
+to RigidSynth is just over 400 lines.
 
@@
 
-- It would be nicer if runtimes were compared with other programs. For example,
-  the related work section mentions FlashFill as 'one of the most prominent
-  recent string transformation systems', so since part of the benchmarks are
-  adapted from FlashFill benchmarks, it would have been interesting to compare
-  the performance here. The authors also say that 'there are many other recent
-  results showing how to synthesize functions from type-based specifications'
-  and that they 'found we needed new techniques to manage the multidimensional
-  search space generated by regular expression equivalences and other aspects of
-  lens combinator languages'. I am missing either an experimental comparison
-  with these other results or a more detailed explanation of why they would not
-  be applicable to their problem.
-
-Discussed in the general section.
-
 ==========================
@@
 ==========================
-
-- The experimental section could be stronger. I understand that comparisons with
-  other tools is not feasible in this setting, as there aren't any other tools
-  that are solving this specific variety of synthesis. However, you could do
-  some internal comparisons, for various values of the design decisions.
-
-Discussed in the general section.
 
@@
 
-Discussed in part in the general section.
-We did not ever try an approach that was not type-directed, as we do not think
-enumerative search followed by a pruning of ill typed terms would scale well.
+Discussed in part in the general section. We did not ever try an approach that
+was not type-directed, as we do not think enumerative search followed by a
+pruning of ill typed terms would scale well, as there are many ill typed terms.
 
@@
 
-- The current presentation doesn’t explain what are the difficult challenges for
-  this problem. Given the source and target regular expressions, and especially
-  with the strongly unambiguous constraint, it seems a simple enumerative search
-  should work well. I was wondering if the authors tried simpler enumerative
-  solvers baselines for this problem?
-
-Discussed in the general section.
-
-- The second big concern is that it was not obvious who is the target audience
-  for this system. If we are expecting a user to provide the regular expressions
-  for source and target formats, why can’t they also write the transformation as
-  well? Can the authors provide the Boomerang program for the BibTeX example and
-  explain why it would be challenging for a user to write this Boomerang program
-  if they can write regular expressions for the two formats.
-
-Discussed in the general section.
-
 - For the BibTeX to EndNote transformation example, the paper presents an
@@
 
-- First, who are the intended users of the proposed tool? Are they end users
-  who might not have technical knowledge about data transformation programs? Or
-  are they technical users who are unfamiliar with lens programs? I initially
-  thought the tool is intended for the first type of users, but as I read
-  through the paper it seems apparent that the authors intend to target the
-  second group, especially with the requirement that users need to provide a
-  regular expression description of the input/output data format. The
-  requirement to do so seems to be the weakest point of the paper, especially
-  given that existing tools (such as FlashFill or Dataplay [1]) do not have such
-  requirement. If the tool is indeed intended for users who are not familiar
-  with lens programming, then can the authors justify why lens is the
-  appropriate programming model for data transformation? Wouldn't it be easier
-  to dump the input/output data into a common format (say CSV, spreadsheet, or
-  JSON), and simply use FlashFill or TriFacta [2] to learn the desired
-  transformation program rather than providing the ? I understand that it might
-  be difficult for tools like FlashFill to learn nested constructs (given its
-  relational data model), but to me that doesn't quite justify the need for
-  providing regular expressions as part of the input.
-
-Discussed in part in the general section.
-
 - Second, the evaluations seem weak.  The examples from FlashFill and are those

### 2017-01-29 paper/rebuttal-old.txt
rename rebuttal names

--- 2017-01-28 paper/rebuttal.txt
@@
 Several reviewers asked who benefits from this technology.  Briefly, the
-technology is aimed at programmers, to be used as a code writing tool.
-Previous work on data wranglers works on easily transforming data, which
-can then be analyzed manually.  This work aims to provide a tool which can
-assist in writing product quality data transformers.
+technology is aimed at programmers, as an interactive tool for creating complex
+bidirection programs. Previous automated data wranglers transform and extract
+data, which is then manually analyzed and used.  Unexpected data is handled
+through manual inspection of the outputs, with the program potentially raising
+flags when it encounters new and unexpected cases.  For some environments this
+is sufficient, but this interaction model doesn't work well for published,
+production quality code.  The data wrangler can't simply be resynthesized to
+handle the new cases, instead the generated program must be explicit about what
+types of data it can handle, and must be guaranteed to work on this data.  This
+work aims to provide a tool which can assist in writing product quality data
+transformers.
 
@@
 
-Furthermore, we agree that providing regular expressions does require the
-programmers to do additional work.  While we would eventually like to find ways
-to encorporate our work with existing research on inferring regular
+Furthermore, we agree that inputting regular expressions and examples is a
+harder task than merely inputting examples.  While we would eventually like to
+find ways to encorporate our work with existing research on inferring regular
 expressions, we wanted to initially address the novel portion of this eventual

### 2017-01-29 paper/rebuttal.txt
some spelling and grammar on rebuttal

@@
 
-Several reviewers asked who benefits from this technology.  Briefly, the
-technology is aimed at programmers, to be used as a code writing tool.
-Previous work on data wranglers works on easily transforming data, which
-can then be analyzed manually.  This work aims to provide a tool which can
-assist in writing product quality data transformers.
+Several reviewers asked who the intended users of this tool are and
+when the use case was compelling.  Briefly, the technology is aimed at
+programmers.  If one were to ask random programmers whether they were
+familiar with and knew how to use regular expressions, most would say
+yes.  If one were to random programmers whether they were familiar
+with lenses 100% would say no.  The central idea is to exploit
+something that most programmers know well (regular expressions) to
+help them write significant programs in a language that they do not
+know well (lenses).  Programmers are often reluctant to learn new
+things.  Our system helps them avoid that.
 
-For this work, the central idea is to exploit something that most programmers
-know well (regular expressions) to help them write significant programs in a
-language that they do not know well (lenses).  This help is particularly
-important because past experience has shown that lenses are often quite
-tricky to write.  For example, consider the following lens for translating
-between idealized BibTeX and EndNote author formats:
+Several reviewers wondered whether regular expressions were really
+necessary at all and whether we could get away without requiring them.
+One of the reasons we feel our work is valuable is that it illustrates
+an important practical tradeoff: by providing data formats, we can
+infer transformations over significantly more sophisticated string
+data types than related systems which do not use such information.
+For instance, FlashFill cannot infer nested iterations.  17 of our
+benchmarks have nested iterations and the deepest nesting structure
+involved 4 levels of nesting.  As another example, FlashExtract does
+not have unions, but it can represent them indirectly by using
+options.  Still, in our tests (e.g. on the BibTeX data source),
+FlashExtract could not infer transformations between string data types
+that contain unions.  By the way, both FlashExtract and FlashFill are
+wonderful systems, but they are optimized for other problems.  Our
+work demonstrates the tradeoff between designs.  We hope to make this
+point more clearly in a revised version of the paper and to quantify
+the number of benchmarks that we handle that cannot be handled by
+other string synthesis systems.
+
+Several reviewers wondered whether lenses really were tricky to write
+and suggested we show a concrete example (thank you -- we should
+have).  The following lens translates between idealized BibTeX and
+EndNote author formats (it was auto-generated by our tool):
 
@@
 
-Writing this lens from scratch is nontrivial (mainly because of all the
-swaps), and it is a relatively simple lens -- Augeas and Boomerang users
-regularly work with lenses significantly larger than this.  For example,
-Boomerang's full version of an idealized BibTeX converter is over 300 lines of
-code, and all but 4 of the studied Augeas programs were over 50 lines of code,
-with 4 over 100 lines of code.
-
-Furthermore, we agree that providing regular expressions does require the
-programmers to do additional work.  While we would eventually like to find ways
-to encorporate our work with existing research on inferring regular
-expressions, we wanted to initially address the novel portion of this eventual
-goal.  We think that even without the inference of regular expressions, this
-work already is useful as a computer assisted bidirectional programming tool.
+Writing this lens from scratch is quite tricky and fiddly.  The
+trickiest part is the nested swaps, which are fiddly to think about
+and get right.  We would also argue that the cognitive model overall
+is tricky and quite different from what many programmers are used to
+--- one has to think in two directions at once.  Importantly, this is
+an incredibly simple example -- Augeas and Boomerang users regularly
+work with lenses significantly larger than this.  For example,
+Boomerang's full version of an idealized BibTeX converter is over 300
+lines of code, and all but 4 of the Augeas programs we studied were
+over 50 lines of code, with 4 over 100 lines of code.
 
@@
 
-In earlier drafts, we actually had a formal description of this algorithm,
-but (in retrospect regrettably) we removed it to get under the page limit
-and inadvertently removed it from the extended version at the same time.  We
-will reinstate the full treatment in the extended report and find a good way
-to summarize it in the final paper.
+In earlier drafts, we actually had a formal description of this
+algorithm, but we (regrettably) removed it to get under the page limit
+and inadvertently removed it from the extended version at the same
+time.  We will reinstate the full treatment in the extended report and
+find a good way to summarize it in the final paper.
 
-The formal description is not very complex, but takes a good amount of space to
-complete.  The formal description of RigidSynth first describes an ordering on
-atoms, sequences, and DNF regular expressions.  Next, this ordering is used to
-define a normalization procedure for DNF regular expressions.  Finally, a
-procedure for "zipping" two DNF regular expressions together is provided.
-If these two DNF regular expressions can be zipped, then there is a lens between
-them.  A transformation from zipped DNF regular expressions and the
-permutations that normalize them is then provided. These formalizations are not
-very complex, but they take up a lot of space.
+The formal description is not very complex, but takes a good amount of
+space to complete.  The formal description of RigidSynth first
+describes an ordering on atoms, sequences, and DNF regular
+expressions.  Next, this ordering is used to define a normalization
+procedure for DNF regular expressions.  Finally, a procedure for
+"zipping" two DNF regular expressions together is provided.  If these
+two DNF regular expressions can be zipped, then there is a lens
+between them.  A transformation from zipped DNF regular expressions
+and the permutations that normalize them is then provided. These
+formalizations are not very complex, but they take up a lot of space.
 
@@
 
-Direct comparison with related tools is difficult because there aren't any
-tools that do the same task: this is the first attempt at bidirectional
-program synthesis.  Even performance comparisons to similar tools such as
-FlashFill and FlashExtract are difficult because they cannot express the
-rich set of permutations we tackle.  We recognize that a quantification of this
-increase in expressibility would be important for readers of our paper.  We
-think that we should provide an analysis detailing how many of our programs are
-inexpressible using other string synthesis tools.
+Above, we discussed the fact that there are fundamental differences between
+the kinds of transformations our tool can perform and the kinds of
+transformations tools such as FlashFill and FlashExtract can perform.
+We can quantify the set of benchmarks that each tool can process (at least
+in one direction).
 
-Furthermore, we strongly agree that performance measurements quantifying the
-importance of these aspects of the various heuristics used by the algorithm
-would greatly improve the paper.  It will not be hard to instrument the
-implementation so that we can disable heuristics individually and gather
-these numbers.  (Anecdotally, we invented each of these heuristics when more
-naive approaches failed.  For example, our initial search procedure, which
-traversed arbitrary equivalences, could not synthesize anything beyond
-simple microbenchmarks, even with the significant simplification of not
-searching for potential compositions!  Without a priority queue, we were
-able to synthesize complicated lenses when the source and target formats
-were similar, but not when the formats differed significantly.)
+We agree that quantifying the impact of each heuristic we use would
+strengthen our paper.  We will instrument our implementation to gather
+data on the effectiveness of each heuristic individually.
+Anecdotally, we invented each of these heuristics when more naive
+approaches failed.  For example, reviewer D asks whether we "guess an
+arbitrary regular expression to deal with composition."  Actually, our
+first attempt at this problem tried to do that -- it did not use DNF
+lenses and it could not synthesize anything beyond simple
+microbenchmarks.  Moreover, this initial implementation did not even
+try to search for programs that might use composition (a second issue
+our revised system overcomes)!  Our priority queue and ranking system
+were also motivated by failures endured by simpler system: those
+simple systems were able to synthesize complicated lenses when the
+source and target formats were similar, but not when the formats
+differed significantly.
 
@@
 
-Our program is unable to synthesize general Boomerang lenses, but rather only
-those which express bijections.
-There certainly are limitations to restricting ourselves to bijective lenses.
-Indeed, while all existing Boomerang and Augeas lenses can be transformed into
-bijective lenses, few are bijective by default, often because of a lack of
-bijectivity guarantees on whitespace.
-We chose to restrict our domain to bijective programs as a simplification
-technique for this new research, but think there are many directions this work
-can be expanded to, and we feel this research is amenable to be expanded.
-
-In particular, bijective quotient lenses, lenses which are bijective modulo an
-equivalence on the source and target languages, are an expansion to this work
-that already has shown promise.  Merely adding quotienting makes this tool more
-practical.  XSugar - an existing, powerful language for XML and string
-bidirectional transformations - expresses bijections modulo equivalences.
-While we have only seen preliminary results so far, it seems that only minor
-modifications were needed for our tool to generalize to bijective quotient
-lenses.
-
-We will clarify all these points in the next rewrite.
+Bijective lenses are a subset of Boomerang's language of lenses, but
+they are an important subset that both admit many useful
+transformations and that also present some of the crucial technical
+challenges.  In particular, overcoming the difficulties introduced by
+regular expression typing is a fundamental challenge that is useful
+for synthesizing larger subsets of Boomerang and any language related
+to Boomerang.  (We also hope many of the principles introduced here
+will apply to languages such as XDuce and other DTD or XML translation
+languages, whose schema depend upon regular expressions.)
 
@@
 
+- The search algorithm, particularly RigidSynth, is described almost
+completely in prose, rather than with inductively defined rules.
+
+As mentioned above, we have formalized this as a collection of
+inference rules.  We regrettably omitted it for space reasons from
+the conference submission.  We can include it in a future version.
+
 - The synthesized lenses are not shown or discussed in much detail,
-  nor are they compared to hand-written ones (when possible).
+  nor are they compared to hand-written ones (when possible).  How do
+  these compare to hand-written types/terms in Augeas? How do the
+  inferred lenses compare to hand-written terms in Augeas (or Flash
+  Fill, when appropriate)?
 
 Our implementation puts some work into making the synthesized lenses
-readable.  The transformation from DNF lenses to Boomerang lenses alone
-creates fairly unintelligible syntax.  This is an issue, we would like for
-programmers to be able to validate the synthesized programs through a manual
-inspection of them.  As such we apply a number of optimizations, including,
-but not limited to:
+readable.  We perform the following transformations (among others):
 
@@
 
-The previously shown lens, bibtex_to_endnote_au, is an autogenerated lens.
-The primary difference we see between the autogenerated code, and what we would
-write ourselves, is the repeated code of
+The lens we presented above bibtex_to_endnote_au, is an autogenerated
+lens.  Overall, we believe bibtex_to_endnote_au is quite similar to
+what would be written in Augeas by hand.  One thing a human might
+choose to do is to notice that:
 
-ins "au - "
-. lens_swap (NAME . del ",") (lens_swap WSP NAME)*
+lens_swap (NAME . del ",") (lens_swap WSP NAME)*
 
-We personally do not find this to make the code significantly more difficult to
-read.  However, adding this optimization is merely a rote application of common
-subexpression elimination.
-
-- The prevalence and limitations of bijective lenses (compared to
-  less well-behaved ones) is not discussed.
-
-There certainly are limitations to restricting ourselves to bijective lenses.
-However, we chose the simplification of bijective lenses for this new research
-area, but think there is a wealth of future work that can be done through an
-investigation of less well-behaved lenses.
+appears twice.  One could extract this common subexpression and bind it to a
+name.  (We could implement a common sub-expression elimination pass but did 
+not do so in the submission.)
 
@@
 
-Few of the surveyed FlashFill examples were close to bijective.  The tricks we
-used to make our chosen 3 examples bijective - adding in extra fields to hold
-bits of data - could be extended to non-bijective examples.  Our primary goal
-with the use of FlashFill was to show that, when made bijective and given data
-format descriptors, the problems FlashFill addresses become simple.
+Few of the surveyed FlashFill examples were "close to" bijective.  Most of
+them are projections.  Bijections are typically used to synchronize two
+data sources in different formats.  Hence the domains are different.  Our
+purpose in considering the FlashFill examples was to illustrate that our
+tool could handle the bijective data that FlashFill could handle.  And
+as we mentioned earlier, there are many formats that we can handle that
+FlashFill cannot.
+
+- What do the type specifications look like?
+
+Please scroll down to reviewer D (or grep for "cust/bib_prob" or
+"aug/cron" below).  You will see that the formats we are describing
+are really quite complicated.  Again, we would like to emphasize the
+tradeoff we are studying vs a tool like FlashFill: we ask users to
+supply information of kind they are familiar with (regular
+expressions), which does clearly take work.  However, the payback is
+that we now have a lot of leverage unavailable to other tools, and as
+result, makes it possible for us to synthesize transformations over
+very complex formats (with many nested iterations and unions that
+other tools cannot handle).
+
+- Is there reason to believe that the approach will or will not be
+amenable for larger subsets of Boomerang?"
+
+All of Boomerang is oriented around regular expression typing.
+Showing how to manage the core problem of regular expression typing
+and equivalence appears to be a crucial first step that any subsequent
+tool will need to build on.  We are optimistic this foundation will
+support larger subsets of Boomerang.
 
@@
 ==========================
-
-- paper unclear in various (important) places
-
-We unfortunately were had to be brief in our description of confluence,
-bisimilarity, and definitional equivalence.  We go into much greater detail in
-the appendix, and will make sure to refer to the appendix for a more detailed
-description of these terms, and for a more formal understanding of their
-importance.
 
@@
 
-Discussed in the part in the general section.  Many of the lines go not to the
-core synthesis algorithm, but for language transformations, pretty printing,
-lens simplification, and DFA operations.  The portion of our code corresponding
-to RigidSynth is just over 400 lines.
+We regrettably omitted a formalization from the paper for space
+reasons.  If authors are given additional pages in a final version of
+the paper, we can include a version.  Much of the code is not for the
+core synthesis algorithm, but for language transformations, pretty
+printing, lens simplification, and DFA operations.  The portion of our
+code corresponding to RigidSynth is just over 400 lines.
 
@@
 Please refer to our appendix, where we go into a detailed proof of these
-theorems.
-
-==========================
-Review C:
-==========================
-
-- The discussion of related work can be improved. It is true that there aren't
-  other approaches that are solving the same exact problem. However, some
-  aspects of your synthesis are shared with other approaches.
-
-Agreed, while we use similar techniques, they are used in different ways,
-identifying the similarities and noting the differences would strengthen our
-related work.
+theorems.  (The appendix was submitted with the paper as auxiliary materials ---
+please ask other reviewers or the program chair if you have difficulty
+locating the auxiliary files on hotcrp.)
 
@@
 
-Discussed in part in the general section.
+The key algorithmic contribution is our technique for reducing the
+language of general lenses to the language of DNF lenses.  We give
+a formal completeness argument for this reduction.
 
-RigidSynth is entirely type directed.  We use an ordering on the subcomponents
-of the types to find which subcomponents must be mapped to each other.  After
-these lenses are found, they are then combined into the final lens in a way
-that keeps the lens well typed.
+After this reduction, our algorithm is has two phases, which alternate.
+One phase rewrites regular expressions according to our rules.  It is
+not type directed.  The second phase is type directed:  it is a
+divide-and-conquer algorithm that makes a different decision based on
+whether the top level constructor is concatenation, union or kleene star.
 
@@
 Because of our use of orderings, we are able to find the permutations in
-O(n*log(n)) time.
+O(n*log(n)) time.  It does not seem likely that a SAT solver would help
+here.
 
@@
 
-Discussed in part in the general section. We did not ever try an approach that
-was not type-directed, as we do not think enumerative search followed by a
-pruning of ill typed terms would scale well, as there are many ill typed terms.
+Please see the introductory part of this rebuttal.
 
@@
 
-- For the BibTeX to EndNote transformation example, the paper presents an
-  algorithmic way to align the regular expressions to find corresponding
-  permutations of sub-expressions to match different fields. Why can’t the user
-  provide regular expressions in first place that are aligned? For example,
-  there is an epsilon in the BibTeX regular expression, but there is no epsilon
-  in the EndNote format. Wouldn’t it be a more clear and a better specification
-  on user’s part to write a regular expression where EndNode format also has an
-  epsilon with an Or in front. For finding permutations as well, it isn’t clear
-  why can’t the user simply annotate or label different fields using some user
-  interface, rather than the algorithm having to search over different matching
-  of sub-expressions.
+- Why can’t the user provide regular expressions in first place that 
+are aligned?
 
-We feel that requiring well aligned regular expressions is too large a burden on
-the user.  The user themselves needs to think through various equivalence rules,
-which we think is better suited to an automated tool.
-
-This is very difficult when the equivalences that need to be exploited are
-nonobvious or hidden away.  Consider the two regular expressions (A|B)C*
-and (A|C(C*)A)|(B|B(C*)C).  It is nontrivial for a user to identify that the
-equivalences that must be applied to reach two well aligned lenses.
-Furthermore, the aligned regular expressions could become difficult to read and
-maintain: (A|B)(e+C+CC*C) is less understandable than (A|B)C*.
-
-The idea of adding in the ability to annotate fields is an interesting one.
-The way that examples are embedded in the regular expressions is similar to an
-annotation.  In certain scenarios, it may also be easier to give annotations
-than to give a covering example.  Likely the best solution would be one which
-uses both annotations and examples to help differentiate similar fields.
+Describing a single format using a regular expression is a standard
+task that programmers understand.  Aligning two regular expressions
+is a level more difficult.  Because there are translations going on,
+figuring out what needs to be aligned with what does not seem so easy.
+Consider the two regular expressions (A|B)C* and
+(A|C(C*)A)|(B|B(C*)C).  We believe it is quite 
+nontrivial for a user to identify that
+the equivalences that must be applied to reach two well aligned
+lenses.  Furthermore, the aligned regular expressions could become
+difficult to read and maintain: (A|B)(e+C+CC*C) is less understandable
+than (A|B)C*.  Surely, avoiding this complexity is a win for the user.
+Fortunately, we are successful in our efforts to do alignment
+automatically so we can relieve programmers of this burden.
 
@@
 
-The regular expressions provided to the synthesis algorithm must be unambiguous.
-However, we provide machinery to allow ambiguous user-defined regular
-expressions to safely be subexpressions of the input format descriptors.
-Using the keyword "abstract" on the definition of a user-defined regular
-expression allows these abstract regular expressions to never be expanded into
-their subcomponents in the search procedure.  With this opacity, ambiguous
-regular expressions can safely be subexpressions of an input format descriptor,
-though only the identity transformation can act on them.
+In general, the regular expressions need to be unambiguous.  If they
+aren't then the kind of "round tripping" laws one would want simply
+won't hold -- i.e. when you translate back-and-forth, you may wind up
+corrupting the data.  (And making standard assumptions, like longest
+prefix match does not work either).  Nate Foster covers this in depth
+in his thesis.  Having said that, there are some places that ambiguity
+is allowed.  In our system, we allow users to write ambiguous regular
+expressions if they only want to perform the identity transformation
+on that part of the string.  (If they want to carve up and rearrange
+the data, the disambiguity constraints are important.)
 
@@
 
-Discussed in part in the general section. These heuristics were chosen in a way
-to leverage the abstractions that programmers use for their specifications,
-while also allowing enough breadth to the search space to not get stuck in
-local minima.
+Discussed above in the general section.
 
@@
 
-- Second, the evaluations seem weak.  The examples from FlashFill and are those
-  that can already be synthesized using FlashFill by just providing input /
-  output examples. Also, can the authors describe how complex are the regular
-  expressions that users need to provide? Given that the same data format can be
+- The examples from FlashFill and are those that can already be synthesized 
+  using FlashFill by just providing input / output examples. Also, can the 
+  authors describe how complex are the regular expressions that users need 
+  to provide?
+
+Please grep above for "aug/cron" and "cust/bib_prob" to see some example 
+specs.
+
+The purpose of the FlashFill examples was just to illustrate that our
+system, with the extra information we supply it, certainly has no
+problem with the FlashFill examples that are bijective.  As mentioned
+above, most of the other benchmarks cannot be handled by FlashFill
+because they contain nested iterations.  In total, 17 of the examples
+contain nested iterations.  The most complex have 4 levels of nested
+iterations.  These examples cannot be handled by FlashFill.
+
+- Given that the same data format can be
   described using different semantically equivalent expressions, how do they
@@
 
-Discussed in part in the general section. Semantically equivalent expressions
-can have an impact on the effectiveness of the synthesis process.  As we discuss
-in Speed of Synthesis, in the Evaluation section, the difference in the
-presentation of the regular expressions being mapped between is a stronger
-indicator of complexity for this system.
+As we discuss in Speed of Synthesis, in the Evaluation section, the
+difference in the presentation of the regular expressions can have an
+impact on the performance of synthesis.  The "farther away" the two
+pairs of expressions are, the more search is needed to find
+sufficiently compatible pairs of regular expressions.  Much of the
+focus of our heuristics is on identifying compatible regular
+expression pairs during search.  Our experiments suggest they are
+successful.
 
@@
 
-If there is possibly dirty data, then we expect this dirty data to not be
-input to the transformer through the typing.  This would allow for the
-unexpected data to fail, instead of silently succeed with unexpected results.
-If this odd data is found to be an expected occurence, changes to the typing can
-be made to address it, and a new lens can be synthesized.
+Briefly, one can test the regular expression against available data to
+see if it parses the available data, but we offer no special support
+here --- doing so is beyond the scope of this paper. Our experiments
+suggest we need very few examples -- often zero or one.  If in the one
+example the user puts the data in the wrong place, the user will
+simply have specified the wrong transformation.  Developing some more
+general machinery for analyzing many examples, some of which may be
+noisy/buggy, is beyond the scope of this paper.
 
@@
   of the key requirement for the tool to succeed in synthesizing a valid
-  program. In addition, is there some way for the tool to identify the error and
-  perform iterative synthesis? Also, what happens if multiple lens program are
-  found for a given input?
+  program.
 
-Clever examples do not need to be provided to the system.  The "Avg # Exmpl"
-column gets its value from randomly generated examples, as we did not want our
-implicit understanding of the synthesis algorithm to make our system look like
-it needed fewer examples than it does when used by a non-expert.  When
-multiple programs are possible, the one that requires the least significant
-permutation is chosen.  The "Max # Exmpl" column corresponds to a permutation
-generated which chooses a wrong permutation, if such a permutation is
-permutation is possible.
+"Clever" examples do not need to be provided to the system.  The "Avg # Exmpl"
+column gets its value from randomly generated examples.
+
+- What happens if multiple lens program are found for a given input?
+
+The system returns the "simplest" program (we have a heuristic notion of
+"simplest").

### 2017-01-29 paper/rebuttal.txt
Checkpoint

@@
 
-Several reviewers asked who the intended users of this tool are and
-when the use case was compelling.  Briefly, the technology is aimed at
-programmers.  If one were to ask random programmers whether they were
-familiar with and knew how to use regular expressions, most would say
-yes.  If one were to random programmers whether they were familiar
-with lenses 100% would say no.  The central idea is to exploit
-something that most programmers know well (regular expressions) to
-help them write significant programs in a language that they do not
-know well (lenses).  Programmers are often reluctant to learn new
-things.  Our system helps them avoid that.
+Several reviewers asked who the intended users of this tool are and when the
+use case is compelling.  The technology is aimed at programmers, and the
+central idea is to exploit something that most programmers know well
+(regular expressions) to help them write significant programs in a language
+that they do not know well (lenses).
 
@@
 
-We agree that quantifying the impact of each heuristic we use would
-strengthen our paper.  We will instrument our implementation to gather
-data on the effectiveness of each heuristic individually.
-Anecdotally, we invented each of these heuristics when more naive
-approaches failed.  For example, reviewer D asks whether we "guess an
-arbitrary regular expression to deal with composition."  Actually, our
-first attempt at this problem tried to do that -- it did not use DNF
-lenses and it could not synthesize anything beyond simple
-microbenchmarks.  Moreover, this initial implementation did not even
-try to search for programs that might use composition (a second issue
-our revised system overcomes)!  Our priority queue and ranking system
-were also motivated by failures endured by simpler system: those
-simple systems were able to synthesize complicated lenses when the
-source and target formats were similar, but not when the formats
-differed significantly.
+We agree that quantifying the impact of each heuristic would strengthen the
+paper.  We plan to instrument our implementation to gather data on the
+effectiveness of each heuristic individually.
+
+Anecdotally, we invented each of these heuristics when more naive approaches
+failed.  For example, reviewer D asks whether we "guess an arbitrary regular
+expression to deal with composition."  Actually, our first attempt at this
+problem tried to do that -- it did not use DNF lenses and it could not
+synthesize anything beyond simple microbenchmarks.  Moreover, this initial
+implementation did not even try to search for programs that might use
+composition (a second issue our revised system overcomes)!  The priority
+queue and ranking system were also motivated by failures endured by earlier
+implementations, which were able to synthesize complicated lenses when the
+source and target formats were similar, but not when the formats differed
+significantly.
 
@@
 
-Bijective lenses are a subset of Boomerang's language of lenses, but
-they are an important subset that both admit many useful
-transformations and that also present some of the crucial technical
-challenges.  In particular, overcoming the difficulties introduced by
-regular expression typing is a fundamental challenge that is useful
-for synthesizing larger subsets of Boomerang and any language related
-to Boomerang.  (We also hope many of the principles introduced here
-will apply to languages such as XDuce and other DTD or XML translation
-languages, whose schema depend upon regular expressions.)
+Bijective lenses are a subset of Boomerang's language of lenses, but they
+are an important subset that both admits many useful transformations and
+also presents some of the crucial technical challenges.  In particular,
+overcoming the difficulties introduced by regular expression typing is a
+fundamental challenge that is useful for synthesizing larger subsets of
+Boomerang and any language related to Boomerang.  We also hope many of the
+principles introduced here will apply to languages such as XDuce and other
+DTD or XML translation languages, whose schemas are based on upon regular
+expressions.
 
@@
 
-- Converting a regular expression to DNF may incur an exponential explosion (as
-it is akin to determinizing an NFM). Have you encountered examples where that is
-the case? Please state this shortcoming.
+- Converting a regular expression to DNF may incur an exponential explosion
+(as it is akin to determinizing an NFM). Have you encountered examples where
+that is the case? Please state this shortcoming.
 
@@
 However, we include in our benchmarks cust/cap-prob and cust/2-cap-prob to
-highlight this shortcoming.  Furthermore, we discuss the shortcoming in Speed of
-Synthesis in the evaluation section.
+highlight this shortcoming.  Furthermore, we discuss the shortcoming in
+Speed of Synthesis in the evaluation section.
 
@@
   benchmarks are simple. This is perhaps because the language is very
-  restricted; i.e., there are no user-defined lenses, only the core ones. Could
-  you comment on that?
+  restricted; i.e., there are no user-defined lenses, only the core
+  ones. Could you comment on that?
 
@@
 Examples in the evaluation section.  To highlight how complicated the input
-specifications can be, the specification for the BibTeX to
-EndNote conversion problem - one of our 10 least complex input formats - and the
-specification for cron - our most complex format - are given
-below.  We are happy to provide all our specifications.
+specifications can be, the specification for the BibTeX to EndNote
+conversion problem - one of our 10 least complex input formats - and the
+specification for cron - our most complex format - are given below.  We are
+happy to provide all our specifications.
 
@@
 
-The system returns the "simplest" program (we have a heuristic notion of
+The system returns the "simplest" program (using a heuristic notion of
 "simplest").

### 2017-01-29 paper/rebuttal.txt
Checkpoint and release lock

@@
 
-Several reviewers wondered whether regular expressions were really
-necessary at all and whether we could get away without requiring them.
-One of the reasons we feel our work is valuable is that it illustrates
-an important practical tradeoff: by providing data formats, we can
-infer transformations over significantly more sophisticated string
-data types than related systems which do not use such information.
-For instance, FlashFill cannot infer nested iterations.  17 of our
-benchmarks have nested iterations and the deepest nesting structure
-involved 4 levels of nesting.  As another example, FlashExtract does
-not have unions, but it can represent them indirectly by using
-options.  Still, in our tests (e.g. on the BibTeX data source),
-FlashExtract could not infer transformations between string data types
-that contain unions.  By the way, both FlashExtract and FlashFill are
-wonderful systems, but they are optimized for other problems.  Our
-work demonstrates the tradeoff between designs.  We hope to make this
-point more clearly in a revised version of the paper and to quantify
-the number of benchmarks that we handle that cannot be handled by
-other string synthesis systems.
+Several reviewers wondered whether regular expressions were really necessary
+at all and whether we could get away without requiring them.  One of the
+reasons we feel our work is valuable is that it illustrates an important
+practical tradeoff: by providing data formats, we can infer transformations
+over significantly more sophisticated string data types than related systems
+which do not use such information.  For instance, FlashFill cannot infer
+nested iterations.  17 of our benchmarks have nested iterations and the
+deepest nesting structure involved 4 levels of nesting.  As another example,
+FlashExtract does not have unions, but it can represent them indirectly by
+using options.  Still, in our tests (e.g. on the BibTeX data source),
+FlashExtract could not infer transformations between string data types that
+contain unions.  By the way, both FlashExtract and FlashFill are wonderful
+systems, but they are optimized for other problems.  Our work demonstrates
+the tradeoff between designs.  We will make this point more clearly in a
+revised version of the paper, and in particular we will quantify the number
+of benchmarks that we handle that cannot be handled by other string
+synthesis systems.
 

### 2017-01-29 paper/rebuttal.txt
Polishing the rebuttal.

--- 2017-01-29 paper/rebuttal.txt
@@
 familiar with and knew how to use regular expressions, most would say
-yes.  If one were to random programmers whether they were familiar
-with lenses 100% would say no.  The central idea is to exploit
+yes.  If one were to ask random programmers whether they were familiar
+with lenses, 100% would say no.  The central idea is to exploit
 something that most programmers know well (regular expressions) to
@@
 FlashExtract could not infer transformations between string data types
-that contain unions.  By the way, both FlashExtract and FlashFill are
-wonderful systems, but they are optimized for other problems.  Our
-work demonstrates the tradeoff between designs.  We hope to make this
-point more clearly in a revised version of the paper and to quantify
-the number of benchmarks that we handle that cannot be handled by
-other string synthesis systems.
+that contain unions. FlashExtract and FlashFill are wonderful systems,
+but they are optimized for other problems.  Our work demonstrates the
+tradeoff between designs.  We hope to make this point more clearly in
+a revised version of the paper and to quantify the number of
+benchmarks that we handle that cannot be handled by other string
+synthesis systems.
 
@@
 is tricky and quite different from what many programmers are used to
---- one has to think in two directions at once.  Importantly, this is
+-- one has to think in two directions at once.  Importantly, this is
 an incredibly simple example -- Augeas and Boomerang users regularly
@@
 
-The formal description is not very complex, but takes a good amount of
+The formal description is not very complex, but takes significant
 space to complete.  The formal description of RigidSynth first
@@
 
-Above, we discussed the fact that there are fundamental differences between
-the kinds of transformations our tool can perform and the kinds of
+Above, we discussed that there are fundamental differences between the
+kinds of transformations our tool can perform and the kinds of
 transformations tools such as FlashFill and FlashExtract can perform.
-We can quantify the set of benchmarks that each tool can process (at least
-in one direction).
+We can quantify the set of benchmarks that each tool can process (at
+least in one direction).
 
@@
 Describing a single format using a regular expression is a standard
-task that programmers understand.  Aligning two regular expressions
-is a level more difficult.  Because there are translations going on,
-figuring out what needs to be aligned with what does not seem so easy.
+task that programmers understand.  Aligning two regular expressions is
+considerably more difficult.  Due to the presence of translations,
+figuring out what needs to be aligned with is no so easy.
 Consider the two regular expressions (A|B)C* and
-(A|C(C*)A)|(B|B(C*)C).  We believe it is quite 
-nontrivial for a user to identify that
-the equivalences that must be applied to reach two well aligned
-lenses.  Furthermore, the aligned regular expressions could become
-difficult to read and maintain: (A|B)(e+C+CC*C) is less understandable
-than (A|B)C*.  Surely, avoiding this complexity is a win for the user.
-Fortunately, we are successful in our efforts to do alignment
-automatically so we can relieve programmers of this burden.
+(A|C(C*)A)|(B|B(C*)C).  We believe it is quite nontrivial for a user
+to identify which equivalences must be applied to reach two
+well aligned lenses.  Furthermore, the aligned regular expressions
+could become difficult to read and maintain: (A|B)(e+C+CC*C) is less
+understandable than (A|B)C*.  Surely, avoiding this complexity is a
+win for the user.  Fortunately, we are successful in our efforts to do
+alignment automatically so we can relieve programmers of this burden.
 

### 2017-01-29 paper/rebuttal.txt

--- 2017-01-29 paper/rebuttal.txt
@@
-TARGET AUDIENCE
+INTENDED USE CASE
 

### 2017-01-29 paper/rebuttal.txt
another polishing pass

@@
 
-Several reviewers asked who the intended users of this tool are and when the
-use case is compelling.  The technology is aimed at programmers, and the
-central idea is to exploit something that most programmers know well
-(regular expressions) to help them write significant programs in a language
-that they do not know well (lenses).
+Several reviewers asked who the intended users of this tool are and
+when the use case is compelling.  The technology is aimed at
+programmers, and the central idea is to exploit something that most
+programmers know well (regular expressions) to help them write
+significant programs in a language that they do not know well
+(lenses).
 
-Several reviewers wondered whether regular expressions were really necessary
-at all and whether we could get away without requiring them.  One of the
-reasons we feel our work is valuable is that it illustrates an important
-practical tradeoff: by providing data formats, we can infer transformations
-over significantly more sophisticated string data types than related systems
-which do not use such information.  For instance, FlashFill cannot infer
-nested iterations.  17 of our benchmarks have nested iterations and the
-deepest nesting structure involved 4 levels of nesting.  As another example,
-FlashExtract does not have unions, but it can represent them indirectly by
-using options.  Still, in our tests (e.g. on the BibTeX data source),
-FlashExtract could not infer transformations between string data types that
-contain unions.  By the way, both FlashExtract and FlashFill are wonderful
-systems, but they are optimized for other problems.  Our work demonstrates
-the tradeoff between designs.  We will make this point more clearly in a
-revised version of the paper, and in particular we will quantify the number
-of benchmarks that we handle that cannot be handled by other string
-synthesis systems.
+Several reviewers wondered whether regular expressions were really
+necessary at all and whether we could get away without requiring them.
+One of the reasons we feel our work is valuable is that it illustrates
+an important practical tradeoff: by providing data formats, we can
+infer transformations over significantly more sophisticated string
+data types than related systems which do not use such information.
+For instance, FlashFill cannot infer nested iterations.  17 of our
+benchmarks have nested iterations and the deepest nesting structure
+involved 4 levels of nesting.  As another example, FlashExtract does
+not have unions, but it can represent them indirectly by using
+options.  Still, in our tests (e.g. on the BibTeX data source),
+FlashExtract could not infer transformations between string data types
+that contain unions. FlashExtract and FlashFill are wonderful systems,
+but they are optimized for other problems.  Our work demonstrates the
+tradeoff between designs.  We will make this point more clearly in a
+revised version of the paper, and in particular we will quantify the
+number of benchmarks that we handle that cannot be handled by other
+string synthesis systems.
 
@@
 and get right.  We would also argue that the cognitive model overall
-is tricky and quite different from what many programmers are used to
--- one has to think in two directions at once.  Importantly, this is
-an incredibly simple example -- Augeas and Boomerang users regularly
-work with lenses significantly larger than this.  For example,
-Boomerang's full version of an idealized BibTeX converter is over 300
-lines of code, and all but 4 of the Augeas programs we studied were
-over 50 lines of code, with 4 over 100 lines of code.
+is quite different from what many programmers are used to -- one has
+to think in two directions at once.  Importantly, this is an
+incredibly simple example -- Augeas and Boomerang users regularly work
+with lenses significantly larger than this.  For example, Boomerang's
+full version of an idealized BibTeX converter is over 300 lines of
+code, and all but 4 of the Augeas programs we studied were over 50
+lines of code, with 4 over 100 lines of code.
 
@@
 The formal description is not very complex, but takes significant
-space to complete.  The formal description of RigidSynth first
+space to explain.  The formal description of RigidSynth first
 describes an ordering on atoms, sequences, and DNF regular
@@
 procedure for DNF regular expressions.  Finally, a procedure for
-"zipping" two DNF regular expressions together is provided.  If these
+"zipping" two DNF regular expressions together is provided.  If 
 two DNF regular expressions can be zipped, then there is a lens
 between them.  A transformation from zipped DNF regular expressions
-and the permutations that normalize them is then provided. These
-formalizations are not very complex, but they take up a lot of space.
+and the permutations that normalize them is then provided.
 
@@
 
-We agree that quantifying the impact of each heuristic would strengthen the
-paper.  We plan to instrument our implementation to gather data on the
-effectiveness of each heuristic individually.
+We agree that quantifying the impact of each heuristic would
+strengthen the paper.  We plan to instrument our implementation to
+gather data on the effectiveness of each heuristic individually.
 
-Anecdotally, we invented each of these heuristics when more naive approaches
-failed.  For example, reviewer D asks whether we "guess an arbitrary regular
-expression to deal with composition."  Actually, our first attempt at this
-problem tried to do that -- it did not use DNF lenses and it could not
-synthesize anything beyond simple microbenchmarks.  Moreover, this initial
-implementation did not even try to search for programs that might use
-composition (a second issue our revised system overcomes)!  The priority
-queue and ranking system were also motivated by failures endured by earlier
-implementations, which were able to synthesize complicated lenses when the
-source and target formats were similar, but not when the formats differed
-significantly.
+Anecdotally, we invented each of these heuristics when more naive
+approaches failed.  For example, reviewer D asks whether we "guess an
+arbitrary regular expression to deal with composition."  Actually, our
+first attempt at this problem tried to do that -- it did not use DNF
+lenses and it could not synthesize anything beyond simple
+microbenchmarks.  Moreover, this initial implementation did not even
+try to search for programs that might use composition (a second issue
+our revised system overcomes)!  The priority queue and ranking system
+were also motivated by failures encountered with earlier
+implementations, which were able to synthesize complicated lenses when
+the source and target formats were similar, but not when the formats
+differed significantly.
 
@@
 
-Bijective lenses are a subset of Boomerang's language of lenses, but they
-are an important subset that both admits many useful transformations and
-also presents some of the crucial technical challenges.  In particular,
-overcoming the difficulties introduced by regular expression typing is a
-fundamental challenge that is useful for synthesizing larger subsets of
-Boomerang and any language related to Boomerang.  We also hope many of the
-principles introduced here will apply to languages such as XDuce and other
-DTD or XML translation languages, whose schemas are based on upon regular
-expressions.
+Bijective lenses are a subset of Boomerang's language of lenses, but
+they are an important subset that both admits many useful
+transformations and also presents some of the crucial technical
+challenges.  In particular, overcoming the difficulties introduced by
+regular expression typing is a fundamental challenge that is useful
+for synthesizing larger subsets of Boomerang and any language related
+to Boomerang.  We also hope many of the principles introduced here
+will apply to languages such as XDuce and other DTD or XML translation
+languages, whose schemas are based on upon regular expressions.
 

### 2017-05-14 program/generate-data.py
added test for forced expansions

@@ -12,9 +12,9 @@ import time
 
 TEST_EXT = '.ls'
 BASELINE_EXT = '.out'
-BASE_FLAGS = ['-iterativedeepenstrategy']
-TIMEOUT_TIME = 8
-GENERATE_EXAMPLES_TIMEOUT_TIME = 6000
+BASE_FLAGS = []
+TIMEOUT_TIME = 120
+GENERATE_EXAMPLES_TIMEOUT_TIME = 600000
 
 REPETITION_COUNT = 2
 
@@ -41,11 +41,11 @@ def gather_datum(prog, path, base, additional_flags, timeout):
 def gather_data(rootlength, prog, path, base):
     current_data = {"Test":join(path, base).replace("_","-")[rootlength:]}
 
-    def gather_col(flags, run_combiner, col_name, timeout_time):
+    def gather_col(flags, run_combiner, col_name, timeout_time, repetition_count):
         run_data = []
         timeout = False
         error = False
-        for iteration in range(REPETITION_COUNT):
+        for iteration in range(repetition_count):
     	    (datum,err) = gather_datum(prog, path, base,flags,timeout_time)
             if datum == "":
                 if err == "":
@@ -56,9 +56,9 @@ def gather_data(rootlength, prog, path, base):
             else:
 	            run_data.append(datum.split(","))
             if error:
-                current_data[col_name]="Error"
+                current_data[col_name]=-1
             elif timeout:
-	            current_data[col_name]="Timeout"
+	            current_data[col_name]=-1
             else:
                 run_data_transpose = transpose(run_data)
                 current_data[col_name]=run_combiner(run_data_transpose)
@@ -66,9 +66,8 @@ def gather_data(rootlength, prog, path, base):
     def ctime_combiner(run_data_transpose):
         computation_time_col = [float(x) for x in run_data_transpose[0]]
         ans = sum(computation_time_col)/len(computation_time_col)
-        ans = (int)(ans * 100)
-        ans = ans * 10
-        return ans if ans != 0 else "$<10$"
+        ans = (int)(ans * 1000)
[…]

### 2017-05-23 program/generate-data.py
add in the changes i did before refactoring

@@ -90,17 +90,22 @@ def gather_data(rootlength, prog, path, base):
     gather_col(['-no_lens_context','-time'],ctime_combiner,"NoLensContext",TIMEOUT_TIME,REPETITION_COUNT)
     #gather_col(['-no_short_circuit','-no_inferred_expansions','-no_lens_context','-time'],ctime_combiner,"NoInferenceNoLCNoSC",TIMEOUT_TIME,REPETITION_COUNT)
     #gather_col(['-no_short_circuit','-no_lens_context','-time'],ctime_combiner,"NoLCNoSC",TIMEOUT_TIME,REPETITION_COUNT)
-    gather_col(['-no_inferred_expansions','-no_lens_context','-time'],ctime_combiner,"NoInferenceNoLC",TIMEOUT_TIME,REPETITION_COUNT)
-    gather_col(['-no_inferred_expansions','-time'],ctime_combiner,"NoInference",TIMEOUT_TIME,REPETITION_COUNT)
+    gather_col(['-naive_expansion_search','-no_lens_context','-time'],ctime_combiner,"NaiveExpansionNoLC",TIMEOUT_TIME,REPETITION_COUNT)
+    gather_col(['-use_only_forced_expansions','-no_lens_context','-time'],ctime_combiner,"OnlyForcedExpansionsNoLC",TIMEOUT_TIME,REPETITION_COUNT)
+    gather_col(['-naive_expansion_search','-time'],ctime_combiner,"NaiveExpansion",TIMEOUT_TIME,REPETITION_COUNT)
+    gather_col(['-use_only_forced_expansions','-time'],ctime_combiner,"OnlyForcedExpansions",TIMEOUT_TIME,REPETITION_COUNT)
     gather_col(['-generatedexamples'],exs_reqd_combiner,"ExamplesRequired",TIMEOUT_TIME,REPETITION_COUNT)
     gather_col(['-max_to_specify'],max_exs_reqd_combiner,"MaxExampleCount",TIMEOUT_TIME,1)
     gather_col(['-spec_size'],max_exs_reqd_combiner,"SpecSize",TIMEOUT_TIME,REPETITION_COUNT)
+    gather_col(['-lens_size'],max_exs_reqd_combiner,"LensSize",TIMEOUT_TIME,REPETITION_COUNT)
[…]

### 2017-06-29 program/check-with-boomerang.py
add in new tests because i have nightmares about my computer breaking

@@ -14,7 +14,7 @@ BOOMERANG_COMMAND = "boomerang"
 TEMP_BOOM_FILENAME = "generated.boom"
 TEST_EXT = '.ls'
 BASELINE_EXT = '.out'
-BASE_FLAGS = ["-iterativedeepenstrategy"]
+BASE_FLAGS = []
 TIMEOUT = 60
 
 REPETITION_COUNT = 2

### 2017-06-29 program/generate-data.py

@@ -13,10 +13,10 @@ import time
 TEST_EXT = '.ls'
 BASELINE_EXT = '.out'
 BASE_FLAGS = []
-TIMEOUT_TIME = 120
+TIMEOUT_TIME = 600
 GENERATE_EXAMPLES_TIMEOUT_TIME = 600000
 
-REPETITION_COUNT = 2
+REPETITION_COUNT = 10
 
 def ensure_dir(f):
     d = os.path.dirname(f)
@@ -52,16 +52,16 @@ def gather_data(rootlength, prog, path, base):
                     timeout = True
                 else:
                     error = True
-                    break
+                break
             else:
 	            run_data.append(datum.split(","))
-            if error:
-                current_data[col_name]=-1
-            elif timeout:
-	            current_data[col_name]=-1
-            else:
-                run_data_transpose = transpose(run_data)
-                current_data[col_name]=run_combiner(run_data_transpose)
+        if error:
+            current_data[col_name]=-1
+        elif timeout:
+	        current_data[col_name]=-1
+        else:
+            run_data_transpose = transpose(run_data)
+            current_data[col_name]=run_combiner(run_data_transpose)
 
     def ctime_combiner(run_data_transpose):
         computation_time_col = [float(x) for x in run_data_transpose[0]]
@@ -82,37 +82,38 @@ def gather_data(rootlength, prog, path, base):
 	    return int(sum(example_number_col)/len(example_number_col))
 
 
-    gather_col(['-time'],ctime_combiner,"ComputationTime",TIMEOUT_TIME,REPETITION_COUNT)
-    #gather_col(['-forceexpand','-time'],ctime_combiner,"ForceExpandTime",TIMEOUT_TIME,REPETITION_COUNT)
-    gather_col(['-naive_strategy','-time'],ctime_combiner,"NaiveStrategy",TIMEOUT_TIME,REPETITION_COUNT)
-    #gather_col(['-naive_pqueue','-time'],ctime_combiner,"NaivePQueue",TIMEOUT_TIME,REPETITION_COUNT)
[…]

### 2017-07-05 paper/transform-data.py
small changes

@@ -1,0 +1,241 @@
+#!/usr/local/bin/python
+
+from __future__ import print_function
+from easyprocess import EasyProcess
+
+import os
+import csv
+from os.path import splitext, join
+import subprocess
+import sys
+import time
+
+TEST_EXT = '.ls'
+BASELINE_EXT = '.out'
+BASE_FLAGS = []
+TIMEOUT_TIME = 600
+GENERATE_EXAMPLES_TIMEOUT_TIME = 600000
+
+REPETITION_COUNT = 10
+
+def ensure_dir(f):
+    d = os.path.dirname(f)
+    if not os.path.exists(d):
+        os.makedirs(d)
+
+def transpose(matrix):
+    return zip(*matrix)
+
+def find_tests(root):
+    tests = []
+    for path, dirs, files in os.walk(root):
+        files = [(f[0], f[1]) for f in [splitext(f) for f in files]]
+        tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
+    return tests
+
+def gather_datum(prog, path, base, additional_flags, timeout):
+    process_output = EasyProcess([prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)]).call(timeout=timeout)
+    return (process_output.stdout,process_output.stderr)
+
+
+def gather_data(rootlength, prog, path, base):
+    current_data = {"Test":join(path, base).replace("_","-")[rootlength:]}
+
+    def gather_col(flags, run_combiner, col_name, timeout_time, repetition_count):
+        print(col_name)
+        run_data = []
+        timeout = False
+        error = False
+        for iteration in range(repetition_count):
+    	    (datum,err) = gather_datum(prog, path, base,flags,timeout_time)
+            if datum == "":
+                if err == "":
+                    timeout = True
+                else:
+                    error = True
+                break
+            else:
+	            run_data.append(datum.split(","))
+        if error:
+            current_data[col_name]=-1
+        elif timeout:
+	        current_data[col_name]=-1
+        else:
+            run_data_transpose = transpose(run_data)
+            current_data[col_name]=run_combiner(run_data_transpose)
+
+    def ctime_combiner(run_data_transpose):
[…]

### 2017-07-05 combine-data.py
random changes
removed pcf from prog names, removed use of boomerang when bijective lenses sufficient

removed extraction specs to let git accept

@@ -9,12 +9,6 @@ from os.path import splitext, join
 import subprocess
 import sys
 import time
-import matplotlib
-import numpy as np
-import matplotlib.pyplot as plt
-
-matplotlib.rcParams['mathtext.fontset'] = 'stix'
-matplotlib.rcParams['font.family'] = 'STIXGeneral'
 
 TEST_EXT = '.ls'
 BASELINE_EXT = '.out'
@@ -138,7 +132,7 @@ def print_data(data):
 	datawriter.writerows(data)
 
 def print_usage(args):
-    print("Usage: {0} <file1>".format(args[0]))
+    print("Usage: {0} <file1> <file2> <file3>".format(args[0]))
 
 def transform_data(path, base, run_data):
     current_data = {"Test":join(path, base + TEST_EXT).replace("_","-")[6:]}
@@ -161,159 +155,31 @@ def retrieve_csv(filename):
             csv_rows.append(row)
     return csv_rows
 
-def project_column_from_csv(csv_obj, col_name):
-    return [r[col_name] for r in csv_obj]
-
-def write_to_filename(filename, s):
-    with open(filename, "wb") as f:
-        f.write(s)
-
-def generate_examples_required_data(input_csv):
-    zero_count_ind = 0
-    one_to_five_count_ind = 1
-    six_to_ten_count_ind = 2
-    eleven_to_fifteen_count_ind = 3
-    sixteen_to_twenty_count_ind = 4
-    over_twenty_count_ind = 5
-
-    zero_count_text = "0"
-    one_to_five_count_text = "1-5"
-    six_to_ten_count_text = "6-10"
-    eleven_to_fifteen_count_text = "11-15"
-    sixteen_to_twenty_count_text = "16-20"
-    over_twenty_count_text = ">20"
-    ind_to_text = [zero_count_text,
-                   one_to_five_count_text,
-                   six_to_ten_count_text,
-                   eleven_to_fifteen_count_text,
-                   sixteen_to_twenty_count_text,
-                   over_twenty_count_text]
-
-    experimental_values = [0,0,0,0,0,0,]
-    determinizing_values = [0,0,0,0,0,0,]
-
-    def add_to_correct_group(count_values, n):
-        if n < 0.0:
-            raise Exception("SOMETHING WENT WRONG")
-        if n == 0.0:
-            count_values[zero_count_ind] = count_values[zero_count_ind]+1
-        elif n <= 5.0:
[…]

### 2017-07-05 program/generate-data.py

@@ -42,6 +42,7 @@ def gather_data(rootlength, prog, path, base):
     current_data = {"Test":join(path, base).replace("_","-")[rootlength:]}
 
     def gather_col(flags, run_combiner, col_name, timeout_time, repetition_count):
+        print(col_name)
         run_data = []
         timeout = False
         error = False
@@ -82,38 +83,38 @@ def gather_data(rootlength, prog, path, base):
 	    return int(sum(example_number_col)/len(example_number_col))
 
 
-    #gather_col(['-time'],ctime_combiner,"ComputationTime",TIMEOUT_TIME,REPETITION_COUNT)
-    ##gather_col(['-forceexpand','-time'],ctime_combiner,"ForceExpandTime",TIMEOUT_TIME,REPETITION_COUNT)
-    #gather_col(['-naive_strategy','-time'],ctime_combiner,"NaiveStrategy",TIMEOUT_TIME,REPETITION_COUNT)
-    ##gather_col(['-naive_pqueue','-time'],ctime_combiner,"NaivePQueue",TIMEOUT_TIME,REPETITION_COUNT)
-    ##gather_col(['-no_short_circuit','-time'],ctime_combiner,"NoShortCircuit",TIMEOUT_TIME,REPETITION_COUNT)
-    #gather_col(['-no_lens_context','-time'],ctime_combiner,"NoLensContext",TIMEOUT_TIME,REPETITION_COUNT)
-    ##gather_col(['-no_short_circuit','-no_inferred_expansions','-no_lens_context','-time'],ctime_combiner,"NoInferenceNoLCNoSC",TIMEOUT_TIME,REPETITION_COUNT)
-    ##gather_col(['-no_short_circuit','-no_lens_context','-time'],ctime_combiner,"NoLCNoSC",TIMEOUT_TIME,REPETITION_COUNT)
-    #gather_col(['-naive_expansion_search','-no_lens_context','-time'],ctime_combiner,"NaiveExpansionNoLC",TIMEOUT_TIME,REPETITION_COUNT)
-    #gather_col(['-use_only_forced_expansions','-no_lens_context','-time'],ctime_combiner,"OnlyForcedExpansionsNoLC",TIMEOUT_TIME,REPETITION_COUNT)
[…]

### 2017-07-05 paper/transform-data.py
add minor changes

@@ -298,11 +298,14 @@ def generate_uninferred_expansions_data(input_csv):
 
 def generate_time_vs_tasks_data(input_csv):
     def create_step_plot(colname, outputname):
-        col_vals = project_column_from_csv(input_csv, colname)
-        col_vals_and_endpoints = [x for x in col_vals if x != "-1"] + [0,600000]
-        x_vals = sorted([float(x) for x in set(col_vals_and_endpoints)])
-
+        col_vals = [float(x) for x in project_column_from_csv(input_csv, colname) if x != "-1"]
+        col_vals_and_endpoints = col_vals + [0,600000]
+        x_vals = sorted([x for x in set(col_vals_and_endpoints)])
         print(x_vals)
+        x_count_dict = {key: 0 for key in x_vals}
+        for val in col_vals:
+            x_count_dict[val] = x_count_dict[val]+1
+        print(x_count_dict)
         return 0
     create_step_plot("FlashExtract","FlashExtract")
 

### 2017-07-06 paper/transform-data.py
add in new graphs woo

@@ -224,14 +224,15 @@ def generate_examples_required_data(input_csv):
     rects1 = ax.bar(ind, experimental_values, width, color='#ffffb3', align='center')
     rects2 = ax.bar(ind+width, determinizing_values, width, color='#998ec3', align='center')
 
-    ax.set_ylabel('Count')
+    ax.set_ylabel('Benchmark Count')
+    ax.set_xlabel('Example Count')
     ax.set_title("Examples Required for Benchmarks")
     ax.set_xticks(ind + width / 2)
     ax.set_xticklabels(ind_to_text)
 
     ax.legend((rects1[0],rects2[0]),("Experimental Average", "Determinize Permutations"))
 
-    plt.savefig("examples.eps")
+    plt.savefig("generated-graphs/examples.eps")
 
 def generate_uninferred_expansions_data(input_csv):
     zero_count_ind = 0
@@ -287,14 +288,15 @@ def generate_uninferred_expansions_data(input_csv):
     rects1 = ax.bar(ind, uninferred_values, width, color='#ffffb3', align='center')
     rects2 = ax.bar(ind+width, unforced_values, width, color='#998ec3', align='center')
 
-    ax.set_ylabel('Count')
-    ax.set_title("Examples Required for Benchmarks")
+    ax.set_ylabel('Benchmark Count')
+    ax.set_xlabel('Expansions Not Inferred')
+    ax.set_title("Number of Benchmarks with Uninferred Expansions")
     ax.set_xticks(ind + width / 2)
     ax.set_xticklabels(ind_to_text)
 
     ax.legend((rects1[0],rects2[0]),("Full Inference","Forced Only"))
 
-    plt.savefig("uninferred.eps")
+    plt.savefig("generated-graphs/uninferred.eps")
 
 def generate_time_vs_tasks_data(input_csv):
     def create_step_plot(colname, outputname):
@@ -307,7 +309,7 @@ def generate_time_vs_tasks_data(input_csv):
             x_count_dict[val] = x_count_dict[val]+1
         print(x_count_dict)
         return 0
-    create_step_plot("FlashExtract","FlashExtract")
+    create_step_plot("ComputationTime","Full")
 
 def main(args):
     if len(args) == 2:

### 2017-07-06 paper/transform-data.py
add better graphics

@@ -15,6 +15,8 @@ import matplotlib.pyplot as plt
 
 matplotlib.rcParams['mathtext.fontset'] = 'stix'
 matplotlib.rcParams['font.family'] = 'STIXGeneral'
+plt.rc('font', size=10)
+plt.rc('legend', fontsize=10)
 
 TEST_EXT = '.ls'
 BASELINE_EXT = '.out'
@@ -232,7 +234,11 @@ def generate_examples_required_data(input_csv):
 
     ax.legend((rects1[0],rects2[0]),("Experimental Average", "Determinize Permutations"))
 
-    plt.savefig("generated-graphs/examples.eps")
+    fig = plt.figure(1,tight_layout=True)
+    fig.set_figheight(1.8)
+    fig.set_figwidth(5)
+
+    fig.savefig("generated-graphs/examples.eps", bbox_inches='tight')
 
 def generate_uninferred_expansions_data(input_csv):
     zero_count_ind = 0
@@ -296,7 +302,11 @@ def generate_uninferred_expansions_data(input_csv):
 
     ax.legend((rects1[0],rects2[0]),("Full Inference","Forced Only"))
 
-    plt.savefig("generated-graphs/uninferred.eps")
+    fig = plt.figure(1,tight_layout=True)
+    fig.set_figheight(1.8)
+    fig.set_figwidth(5)
+
+    fig.savefig("generated-graphs/uninferred.eps", bbox_inches='tight')
 
 def generate_time_vs_tasks_data(input_csv):
     def create_step_plot(colname, outputname):
@@ -307,7 +317,12 @@ def generate_time_vs_tasks_data(input_csv):
         x_count_dict = {key: 0 for key in x_vals}
         for val in col_vals:
             x_count_dict[val] = x_count_dict[val]+1
-        print(x_count_dict)
+        x_completed_counts = []
+        acc = 0
+        for val in x_vals:
+            acc = acc + x_count_dict[val]
+            x_completed_counts.append(acc)
+        print(x_completed_counts)
         return 0
     create_step_plot("ComputationTime","Full")
 

### 2017-07-06 paper/transform-data.py
added new rbg etc

@@ -302,13 +302,15 @@ def generate_uninferred_expansions_data(input_csv):
 
     ax.legend((rects1[0],rects2[0]),("Full Inference","Forced Only"))
 
-    fig = plt.figure(1,tight_layout=True)
+    fig = plt.figure(2,tight_layout=True)
     fig.set_figheight(1.8)
     fig.set_figwidth(5)
 
     fig.savefig("generated-graphs/uninferred.eps", bbox_inches='tight')
 
 def generate_time_vs_tasks_data(input_csv):
+    fig, ax = plt.subplots()
+
     def create_step_plot(colname, outputname):
         col_vals = [float(x) for x in project_column_from_csv(input_csv, colname) if x != "-1"]
         col_vals_and_endpoints = col_vals + [0,600000]
@@ -323,8 +325,23 @@ def generate_time_vs_tasks_data(input_csv):
             acc = acc + x_count_dict[val]
             x_completed_counts.append(acc)
         print(x_completed_counts)
-        return 0
-    create_step_plot("ComputationTime","Full")
+
+
+        ax.step(x_vals,x_completed_counts,label=outputname)
+
+    create_step_plot("ComputationTime","DNF+UD+FE+IE+PD")
+    create_step_plot("NoLensContext","DNF+UD+FE+IE")
+    create_step_plot("OnlyForcedExpansionsNoLC","DNF+UD+FE")
+    create_step_plot("OnlyForcedExpansionsNoLC","DNF+UD")
+    create_step_plot("NaiveExpansionNoLC","DNF+UD")
+    create_step_plot("NoUDTypes","DNF")
+    create_step_plot("NaiveStrategy","Naive")
+
+    ax.legend()
+
+    fig = plt.figure(3,tight_layout=True)
+       
+    fig.savefig("generated-graphs/times.eps", bbox_inches='tight')
 
 def main(args):
     if len(args) == 2:

### 2017-07-06 paper/transform-data.py
turn modes not into additive optimizations

@@ -332,12 +332,15 @@ def generate_time_vs_tasks_data(input_csv):
     create_step_plot("ComputationTime","DNF+UD+FE+IE+PD")
     create_step_plot("NoLensContext","DNF+UD+FE+IE")
     create_step_plot("OnlyForcedExpansionsNoLC","DNF+UD+FE")
-    create_step_plot("OnlyForcedExpansionsNoLC","DNF+UD")
     create_step_plot("NaiveExpansionNoLC","DNF+UD")
     create_step_plot("NoUDTypes","DNF")
     create_step_plot("NaiveStrategy","Naive")
 
-    ax.legend()
+    ax.set_ylabel('Benchmarks Completed')
+    ax.set_xlabel('Time (ms)')
+    ax.set_title("Time vs Benchmarks Completed")
+
+    ax.legend(bbox_to_anchor=(1.4,1),borderaxespad=0)
 
     fig = plt.figure(3,tight_layout=True)
        

### 2017-07-06 paper/transform-data.py
add in new stuff

@@ -18,142 +18,17 @@ matplotlib.rcParams['font.family'] = 'STIXGeneral'
 plt.rc('font', size=10)
 plt.rc('legend', fontsize=10)
 
-TEST_EXT = '.ls'
-BASELINE_EXT = '.out'
-BASE_FLAGS = []
-TIMEOUT_TIME = 600
-GENERATE_EXAMPLES_TIMEOUT_TIME = 600000
-
-REPETITION_COUNT = 10
+generated_graphs_base = "generated-graphs/"
+transformed_data_base = "transformed-data/"
 
 def ensure_dir(f):
     d = os.path.dirname(f)
     if not os.path.exists(d):
         os.makedirs(d)
 
-def transpose(matrix):
-    return zip(*matrix)
-
-def find_tests(root):
-    tests = []
-    for path, dirs, files in os.walk(root):
-        files = [(f[0], f[1]) for f in [splitext(f) for f in files]]
-        tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
-    return tests
-
-def gather_datum(prog, path, base, additional_flags, timeout):
-    process_output = EasyProcess([prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)]).call(timeout=timeout)
-    return (process_output.stdout,process_output.stderr)
-
-
-def gather_data(rootlength, prog, path, base):
-    current_data = {"Test":join(path, base).replace("_","-")[rootlength:]}
-
-    def gather_col(flags, run_combiner, col_name, timeout_time, repetition_count):
-        print(col_name)
-        run_data = []
-        timeout = False
-        error = False
-        for iteration in range(repetition_count):
-    	    (datum,err) = gather_datum(prog, path, base,flags,timeout_time)
-            if datum == "":
-                if err == "":
-                    timeout = True
-                else:
-                    error = True
-                break
-            else:
-	            run_data.append(datum.split(","))
-        if error:
-            current_data[col_name]=-1
-        elif timeout:
-	        current_data[col_name]=-1
-        else:
-            run_data_transpose = transpose(run_data)
-            current_data[col_name]=run_combiner(run_data_transpose)
-
[…]

### 2017-07-07 paper/transform-data.py
added in new figure and data

@@ -209,8 +209,8 @@ def generate_time_vs_tasks_graph(input_csv):
     create_step_plot("OnlyForcedExpansionsNoLC","NoFPE",'-',normal_size)
     create_step_plot("NoUDTypes","NoUD",':',normal_size)
     create_step_plot("NaiveExpansionNoLC","NoER",'-',normal_size)
-    create_step_plot("NaiveStrategy","FlashExtract",':',normal_size)
-    create_step_plot("NaiveStrategy","Flash Fill",'-',normal_size)
+    create_step_plot("FlashExtract","FlashExtract",':',normal_size)
+    create_step_plot("FlashFill","Flash Fill",'-',normal_size)
     create_step_plot("NaiveStrategy",u"Na\u00EFve",':',normal_size)
 
     ax.set_ylabel('Benchmarks Completed')

### 2017-07-07 paper/transform-data.py
add benchmarks

@@ -199,7 +199,10 @@ def generate_time_vs_tasks_graph(input_csv):
             acc = acc + x_count_dict[val]
             x_completed_counts.append(acc)
 
-        ax.step(x_vals,x_completed_counts,label=outputname,linestyle=style,linewidth=width)
+        if (style != '-'):
+            ax.step(x_vals,x_completed_counts,label=outputname,linestyle=style,linewidth=width, dashes=(5,1))
+        else:
+            ax.step(x_vals,x_completed_counts,label=outputname,linestyle=style,linewidth=width)
 
     normal_size = 2
     full_size = 3

### 2017-07-07 paper/transform-data.py
replaced numbers with real numbers

@@ -175,7 +175,8 @@ def generate_uninferred_expansions_graph(input_csv):
     ax.set_xticks(ind + width / 2)
     ax.set_xticklabels(ind_to_text)
 
-    ax.legend((rects1[0],rects2[0]),("Full Inference","Forced Only"))
+    l = ax.legend((rects1[0],rects2[0]),("NoPD","NoFPE"))
+    plt.setp(l.texts, weight='bold')
 
     fig = plt.figure(2,tight_layout=True)
     fig.set_figheight(1.8)
@@ -240,6 +241,31 @@ def generate_multiple_of_five_number_of_seconds_synthesized_under(input_csv):
         num = num+5.0
     write_to_filename(transformed_data_base + "multiple-of-five-number-of-seconds-synthesized-under.txt", str(int(num)))
 
+def generate_number_augeas(input_csv):
+    names = project_column_from_csv(input_csv,"Test")
+    augeasnames = [x for x in names if x.startswith("aug")]
+    write_to_filename(transformed_data_base + "augeas-count.txt", str(len(augeasnames)))
+
+def generate_min_lens_size(input_csv):
+    sizes_strings = project_column_from_csv(input_csv,"LensSize")
+    sizes = [int(x) for x in sizes_strings]
+    write_to_filename(transformed_data_base + "min_lens_size.txt", str(min(sizes)))
+
+def generate_min_lens_size(input_csv):
+    sizes_strings = project_column_from_csv(input_csv,"LensSize")
+    sizes = [int(x) for x in sizes_strings]
+    write_to_filename(transformed_data_base + "max_lens_size.txt", str(max(sizes)))
+
+def generate_median_expands_forced(input_csv):
+    sizes_strings = project_column_from_csv(input_csv,"ExpansionsForcedNoLensContext")
+    exps = [int(x) for x in sizes_strings]
+    write_to_filename(transformed_data_base + "median_expansions_forced.txt", str(int(np.median(exps))))
+
+def generate_maximum_expands_forced(input_csv):
+    sizes_strings = project_column_from_csv(input_csv,"ExpansionsForcedNoLensContext")
+    exps = [int(x) for x in sizes_strings]
+    write_to_filename(transformed_data_base + "maximum_expansions_forced.txt", str(int(max(exps))))
+
 def main(args):
     if len(args) == 2:
         input_filepath = args[1]
@@ -251,6 +277,10 @@ def main(args):
         generate_time_vs_tasks_graph(input_csv)
[…]

### 2017-07-07 anonymized_material/comparisons/prose/Extraction.Text/LearningSamples.cs
add everything in

@@ -1,0 +1,363 @@
+﻿using System;
+using System.Collections.Generic;
+using System.Linq;
+using System.Text.RegularExpressions;
+using Microsoft.ProgramSynthesis.AST;
+using Microsoft.ProgramSynthesis.Extraction.Text;
+using Microsoft.ProgramSynthesis.Extraction.Text.Constraints;
+using Microsoft.ProgramSynthesis.Extraction.Text.Semantics;
+using Microsoft.ProgramSynthesis.Utils;
+using Microsoft.ProgramSynthesis.Transformation.Text;
+using Microsoft.ProgramSynthesis.Transformation.Text.Semantics;
+using Microsoft.ProgramSynthesis.Wrangling.Constraints;
+using Microsoft.ProgramSynthesis.VersionSpace;
+using Microsoft.ProgramSynthesis.Wrangling;
+using System.Diagnostics;
+using VBFIO = Microsoft.VisualBasic.FileIO;
+
+namespace Extraction.Text
+{
+    /// <summary>
+    ///     Extraction.Text learns programs to extract a single string region or a sequence of string regions from text files.
+    ///     This class demonstrates some common usage of Extraction.Text APIs.
+    /// </summary>
+    internal static class LearningSamples
+	{
+		private class Result
+		{
+			public ResultType ResultType { get; set; }
+			public long Milliseconds { get; set; }
+		}
+
+		private enum ResultType { Success, Exception, IncorrectResponse, NothingLearned }
+
+		private const int NUM_TESTS = 100;
+
+        private static void Main(string[] args)
+		{
+			if (args.Length != 2)
+			{
+				throw new Exception(args.Length.ToString());
+			}
+			var numExamples = int.Parse(args[0]);
+			var fileName = args[1];
+			var results = GetResults(fileName, numExamples);
+			var failedResults = results.Where(result => result.ResultType != ResultType.Success).ToList();
+			if (failedResults.IsEmpty())
+			{
+				Console.Out.WriteLine(results.Select(result => result.Milliseconds).Sum());
+			}
+			else
+			{
+				Console.Out.WriteLine(failedResults.First().ResultType);
+			}
+
+            // Learning sequence is similar to learning region. 
+            // We only illustrate some API usages. Other sequence learning APIs are similar to their region APIs counterpart.
+            // Note: we need to give positive examples continuously. 
[…]

### 2017-07-07 anonymized_material/comparisons/prose/Transformation.Text/SampleProgram.cs

@@ -1,0 +1,129 @@
+﻿using System;
+using System.IO;
+using System.Collections.Generic;
+using System.Linq;
+using System.Threading;
+using System.Security.Permissions;
+using Microsoft.ProgramSynthesis.Transformation.Text;
+using Microsoft.ProgramSynthesis.Transformation.Text.Semantics;
+using Microsoft.ProgramSynthesis.Wrangling.Constraints;
+using VBFIO = Microsoft.VisualBasic.FileIO;
+using System.Diagnostics;
+
+namespace Transformation.Text
+{
+	/// <summary>
+	///     Sample of how to use the Transformation.Text API. Transformation.Text generates string programs from input/output examples.
+	/// </summary>
+	internal static class SampleProgram
+	{
+		private enum ResultType { Success, Exception, IncorrectResponse, NothingLearned }
+
+		private const int NUM_TESTS = 100;
+
+		private static void Main(string[] args)
+		{
+			var t = new Stopwatch();
+			if (args.Length != 2) {
+				throw new Exception(args.Length.ToString());
+			}
+			var numExamples = int.Parse(args[0]);
+			var fileName = args[1];
+			t.Start();
+			var result = GetResult(fileName, numExamples);
+			if (result == ResultType.Success)
+			{
+				t.Stop();
+				Console.Out.WriteLine(t.ElapsedMilliseconds);
+			}
+			else
+			{
+				Console.Out.WriteLine(GetResult(fileName, numExamples).ToString());
+			}
+		}
+
+		private static List<KeyValuePair<string, string>> ParseOutIOList(string fileName)
+		{
+			var ioList = new List<KeyValuePair<string, string>>();
+/*			var reader = new StreamReader(fileName);
+			var csv = new CsvReader(reader);
+			csv.Configuration.Delimiter = "\t";
+			csv.Configuration.HasHeaderRecord = false;
+			while (csv.Read())
+			{
+				ioList.Add(new KeyValuePair<string,string>(csve.GetField<string>(0),csv.GetField<string>(1)));
+				Console.Out.WriteLine("in");
+				Console.Out.WriteLine(csv.GetField<string>(0));
+				Console.Out.WriteLine("out");
+				Console.Out.WriteLine(csv.GetField<string>(1));
+			}
+			Console.Out.WriteLine(ioList.Count());
+			Console.Out.WriteLine(fileName);
+			return ioList;*/
+
[…]

### 2017-07-07 anonymized_material/program/check-with-boomerang.py

@@ -1,0 +1,88 @@
+#!/usr/local/bin/python
+
+from __future__ import print_function
+from easyprocess import EasyProcess
+
+import os
+import csv
+from os.path import splitext, join
+import subprocess
+import sys
+import time
+
+BOOMERANG_COMMAND = "boomerang"
+TEMP_BOOM_FILENAME = "generated.boom"
+TEST_EXT = '.ls'
+BASELINE_EXT = '.out'
+BASE_FLAGS = []
+TIMEOUT = 60
+
+REPETITION_COUNT = 2
+
+def find_tests(root):
+    tests = []
+    for path, dirs, files in os.walk(root,followlinks=True):
+        files = [(f[0], f[1]) for f in [splitext(f) for f in files]]
+        tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
+    return tests
+
+def gather_datum(prog, path, base):
+    process_output = EasyProcess([prog] + BASE_FLAGS + [join(path, base + TEST_EXT)]).call(timeout=TIMEOUT)
+    return (process_output.stdout,process_output.stderr,process_output.return_code)
+
+def run_boomerang():
+    process_output = EasyProcess([BOOMERANG_COMMAND,TEMP_BOOM_FILENAME]).call(timeout=TIMEOUT)
+    return (process_output.stdout,process_output.stderr,process_output.return_code)
+
+def check_files(prog, path, base):
+    test_name = join(path, base + TEST_EXT)
+    (datum,err,return_code) = gather_datum(prog, path, base)
+    if return_code != 0:
+        raise Exception(test_name + " failed with stdout:" + datum + "and stderr:" + err)
+    elif datum == "":
+        if err == "":
+	    raise Exception("synthesis of " + test_name + " timed out")
+        else:
+            raise Exception("synthesis of " + test_name +
+                            " errored with error:" + err)
+    with open(TEMP_BOOM_FILENAME, 'w') as f:
+        f.write(datum)
+    (datum,err,return_code) = run_boomerang()
+    if return_code != 0:
+        raise Exception("boomerang failed on the generated files of " + test_name +
+                        "with stdout:" + datum + "stderr:" + err)
+
+def raise_usage(args):
+    raise Exception("Usage: {0} <prog> <test|testdir>".format(args[0]))
+
+def main(args):
+    if len(args) == 3:
+        prog = args[1]
+        path = args[2]
+        data = []
+        if not os.path.exists(prog):
[…]

### 2017-07-07 anonymized_material/program/generate-baseline.py

@@ -1,0 +1,69 @@
+#!/usr/bin/python
+
+from __future__ import print_function
+
+import os
+from os.path import splitext, join
+import subprocess
+import sys
+import time
+
+TEST_EXT = '.ls'
+BASELINE_EXT = '.out'
+FLAGS = []
+
+def find_tests(root):
+    tests = []
+    for path, dirs, files in os.walk(root):
+        files = [(f[0], f[1]) for f in [splitext(f) for f in files]]
+        tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
+    return tests
+
+def run_test(prog, path, base):
+    return subprocess.check_output([prog] + FLAGS + [join(path, base + TEST_EXT)], stderr=subprocess.STDOUT)
+
+def gen_timestamp():
+    return "(* Generated {0} *)".format(time.strftime("%c"))
+
+def gen_baseline(prog, path, base):
+    test          = base + TEST_EXT
+    test_path     = join(path, test)
+    baseline_path = join(path, base + BASELINE_EXT)
+    print("Generating baseline for {0}...".format(test_path))
+    try:
+        output_string = run_test(prog, path, base)
+    except subprocess.CalledProcessError as err:
+        output_string = "An error was raised by this test: {0}\n{1}".format(err.returncode, err.output)
+        print("Encountered an error running test {0}".format(test_path))
+	print(output_string)
+    with open(baseline_path, 'w') as f:
+        f.write(gen_timestamp() + '\n')
+        f.write(output_string)
+
+def gen_baselines(prog, root):
+    for path, base in find_tests(root):
+        gen_baseline(prog, path, base)
+
+def print_usage(args):
+    print("Usage: {0} <prog> <test|testdir>".format(args[0]))
+
+def main(args):
+    if len(args) == 3:
+        prog = args[1]
+        path = args[2]
+        if not os.path.exists(prog):
+            print_usage(args)
+        elif os.path.exists(path) and os.path.isdir(path):
+            gen_baselines(prog, path)
+        else:
+            path, filename = os.path.split(path)
+            base, ext = splitext(filename)
+            if ext != TEST_EXT:
+                print_usage(args)
+            else:
+                gen_baseline(prog, path, base)
+    else:
+        print_usage(args)
+
+if __name__ == '__main__':
+    main(sys.argv)

### 2017-07-07 anonymized_material/program/generate_extraction_specs.py

@@ -1,0 +1,79 @@
+#!/usr/local/bin/python
+
+from __future__ import print_function
+from easyprocess import EasyProcess
+
+import os
+from os.path import splitext, join
+import subprocess
+import sys
+import time
+
+TEST_EXT = '.ls'
+BASELINE_EXT = '.out'
+TSV_EXT = '.tsv'
+BASE_FLAGS = ['-generate_extraction_spec']
+TIMEOUT_TIME = 6000
+REPETITION_COUNT = 2
+BASE_TSV_DIR = 'generated_extraction_specs/'
+
+def ensure_dir(f):
+    if not os.path.exists(f):
+        os.makedirs(f)
+
+def find_tests_and_ensure_output_dirs(root):
+    tests = []
+    for path, dirs, files in os.walk(root):
+        files = [(f[0], f[1]) for f in [splitext(f) for f in files]]
+        tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
+        ensure_dir(BASE_TSV_DIR + path)
+    return tests
+
+def gather_tsv(rootlength, prog, path, base, num_examples):
+    process_output = EasyProcess([prog] + BASE_FLAGS + [str(num_examples)] + [join(path, base + TEST_EXT)]).call(timeout=TIMEOUT_TIME)
+    if (process_output.stderr != "" or process_output.return_code != 0):
+        raise Exception("failure")
+    return process_output.stdout
+
+def write_tsv(tsv, path, base):
+    with open(join(BASE_TSV_DIR+path, base + TSV_EXT),'w') as tsvfile:
+        tsvfile.write(tsv)
+
+def print_usage(args):
+    print("Usage: {0} <prog> <testdir>".format(args[0]))
+
+def transform_data(path, base, run_data):
+    current_data = {"Test":join(path, base + TEST_EXT).replace("_","-")[6:]}
+    run_data_transpose = transpose(run_data)
+    for index in range(len(run_data_transpose)/2):
+	col_name = run_data_transpose[index][0]
+	col_data = run_data_transpose[index+1]
+        if "" in col_data:
+	    current_data[col_name]="Timeout"
+        else:
+            col = [float(x) for x in col_data]
+            current_data[col_name] = str(sum(col)/len(col))
+    return current_data
+
+def main(args):
+    if len(args) == 4:
+        prog = args[1]
+        path = args[2]
+        num_examples = args[3]
+        rootlength = len(path)
+        data = []
+        base_tsv_dir = ensure_dir('generated_extraction_specs/')
[…]

### 2017-07-07 paper/transform-data.py

@@ -115,6 +115,66 @@ def generate_examples_required_graph(input_csv):
 
     fig.savefig(generated_graphs_base + "examples.eps", bbox_inches='tight')
 
+def generate_compositional_lenses_graph(input_csv):
+    zero_count_ind = 0
+    one_to_five_count_ind = 1
+    six_to_ten_count_ind = 2
+    eleven_to_fifteen_count_ind = 3
+    sixteen_to_twenty_count_ind = 4
+
+    zero_count_text = "0"
+    one_to_five_count_text = "1"
+    six_to_ten_count_text = "2"
+    eleven_to_fifteen_count_text = "3"
+    sixteen_to_twenty_count_text = "4"
+    ind_to_text = [zero_count_text,
+                   one_to_five_count_text,
+                   six_to_ten_count_text,
+                   eleven_to_fifteen_count_text,
+                   sixteen_to_twenty_count_text]
+
+    experimental_values = [0,0,0,0,0,]
+    determinizing_values = [0,0,0,0,0,]
+
+    def add_to_correct_group(count_values, n):
+        if n < 0.0:
+            raise Exception("SOMETHING WENT WRONG")
+        if n == 0.0:
+            count_values[zero_count_ind] = count_values[zero_count_ind]+1
+        elif n == 1.0:
+            count_values[one_to_five_count_ind] = count_values[one_to_five_count_ind]+1
+        elif n == 2.0:
+            count_values[six_to_ten_count_ind] = count_values[six_to_ten_count_ind]+1
+        elif n == 3.0:
+            count_values[eleven_to_fifteen_count_ind] = count_values[eleven_to_fifteen_count_ind]+1
+        elif n == 4.0:
+            count_values[sixteen_to_twenty_count_ind] = count_values[sixteen_to_twenty_count_ind]+1
+        else:
+            raise Exception("SOMETHING WENT WRONG")
+
+    vals = project_column_from_csv(input_csv, "CompositionalLensesUsed")
+    for example_num in vals:
+        add_to_correct_group(experimental_values, float(example_num))
+
+    ind = np.arange(5)
+    width = 0.35
+
+    fig, ax = plt.subplots()
+
+    rects1 = ax.bar(ind, experimental_values, width, color='#ffffb3', align='center')
+
+    ax.set_ylabel('Benchmark Count')
[…]

### 2017-09-11 papers/BijectiveLensSynth/popl_18_rebuttal.txt
add in some popl rebuttal notes

We thank the reviewers for their feedback.

...

Detailed Responses to individual questions by the reviewers follow below.

Review #18A
===========================================================================

Strengths/weaknesses

+ Interesting and potentially useful application of program synthesis to bidirectional programming
+ Nontrivial design/technical development
+ Approach implemented and evaluated on real(ish) examples
- Restriction of focus to bijective "lenses" not clearly justified
- Important parts of algorithm not explained (regarding how examples are used)
- Not clear how much work it is to adapt examples to be suitable for bijective lens synthesis, or how useful the results are w.r.t. the original task

But the main part of the paper is not self-contained - for example the
algorithm description relies on (example-dependent) orderings of sequences and atoms that
are not described; this is a key part of the algorithm since it is
where the provided examples are actually used.

I also have a concern: sicne the approach is
limited to bijective lenses, but many of the test examples were not
bijective, how much work was it to modify the examples
and how easy is it to manually hack the resulting bijective
lenses to make them work on the real data?  Is there any chance of
automating this or generalizing to handle non-bijective lenses
(perhaps with this algorithm as a subroutine)?

I could see future work building on this to handle the non-bijective case 
being a separate (and also interesting) paper.  But I would want the algorithm
to be described more clearly and more caveats about the examples.

p12 fig 6. I'm not sure how ule "DNF structural rewrite" could be
applied as the LHS uses the operators defined in fig. 5, rather than syntax

p17.  The desription of RigidSynth is incomplete.  It mentions orders
$\leq_{Seq}$ and $\leq_{Atom}$ (and variants that use the examples) that do
not appear to be defined.  What happens if there is ambiguity?  for
example what if we are trying to infer a lens of type (a | b | c) <->
(d|e|f) and we only have one example that says a <-> c?  If these
orderings are standard things then please give a reference.  If not,
they really need to be described in the paper.  The way examples are
used to align parts of regular expressions seems really critical to
this approach and is hardly explained.

p18: AllSome (alg. 3): I assume this is the obvious function of type
'a option list -> 'a list option (that is None if there is a None in
the list or Some [a1..an] if the list is [Some a1...Some an]) but
please clarify

p19: "easy to understand" - can you present any evidence for this?
Easy for whom?

p20: "We adapted these examples" please say mor about how many of the
examples needed to be adapted (all? a handful?) and how much work it
was, and how useful the resulting isomorphic lenses were w.r.t the
original data / how much additional work was needed to adapt them.

p20: "We developed a series of optimizations" - clarify that the
algorithm already described includes all of these, and what you are
going to do is disable them, rather than add more optimizations that
weren't discussed earlier.  Might be helpful to signpost what parts of
the algorithm are the "basic" algorithm and what parts are optional optimizations

p21: Fig 9 "Number of expansions" --> "Number of solutions"??

p21.  Also interesting to evaluate would be how time / number of
solutions varies (e.g. for the full algorithm) as the number of
examples provided varies.  Is the number of examples small? An active
approach where the algorithm asks the user for examples might also be
worth exploring (future work).

p24: "because our types do not have full canonical forms" - this is
the case for the equations presented in the paper, but what about
e.g. translating to automata and minimizing?

Review #18B
===========================================================================
Paper summary
-------------

However, I’m not sure what exact advantage the proposed approach brings to the
table. In particular, the user is still required to write exact input and output
types of the transformations via regular expressions. How hard is to write
lenses after one has such expressions? I’ll further detail my comparison now.
Also, the language has big limitations in terms of expressiveness.

### CONS:
- In the reviewer’s opinion, the specification mechanism is not very natural and
it’s unclear whether it actually simplifies writing bijective transformations.
This aspect is not evaluated and it would be nice to at least see the difference
in sizes between the domain/range regexes vs the Boomerang lenses. In general,
there is no report of the sizes of benchmarks.

- Very limited DSL for transformations. The proposed DSL can only describe
transformations that use identities and constant functions over input subparts.
For example, one cannot synthesize the function toUpperCase, which I imagine
could be useful for the kind of transformations appearing in the motivation.
Moreover, the functions can use each word in the input only once: one cannot
express the transformation abc->abcabc. Most bijective functions I can think of
(string encoders, compression algorithms, encryption algorithms) require complex
transformations over the input that go beyond identities. On this note, there
was a paper this year at PLDI on inverting string encoders (Automatic program
inversion using symbolic transducers, by Hu et al.), a topic which seems very
related, but for which I couldn’t find a citation.

- Somewhat limited evaluation. The evaluation is performed on two families of
benchmarks. The first family is an adaptation of format translation problems
from Augeas a tool for configuration editing and the second family is from
FlashFill a tool for synthesis of string transformations. All the benchmarks
presented are “adaptation” and it’s unclear whether the authors only picked
benchmarks for which they knew a priori a program in their DSL existed. Based
on the presented data, I cannot assess how often the presented technique could
be applied. Next, the comparison with the tool FlashFill/FlashExtract is somewhat
confusing. I suspect the authors had to provide some domain and range regular
expressions for the flash fill benchmarks, but this aspect, which is crucial to
help the search, is not mentioned. What happens if someone provides Sigma* and
Sigma* as domains and ranges to Optician. Also, a better explanation of the
difference between the DSLs of Lenses and Flash fill is required. For example,
FlashFill can reverse a string, but the proposed language can’t.

### QUESTIONS FOR REBUTTAL
- What is the difference in size between the regexes for domain/range spec and
the full lenses?
- What is the difference in size between manually written lenses and synthesized
ones?
- Clarify how the benchmarks were selected and whether any were not selected a
priori. Especially the flash* ones. Did you omit those that used character
transformations such as upper-case?
- Can the proposed DSL express any transformation that FlashFill can’t and
vice-versa?
- P19: maximally factors the concats and ors. What does this mean? Regex
minimization is a hard problem, what is that one factors exactly? I think it
would help to show one example of fully synthesized lens in the paper.

Minor:
- Eval the text says Full synthesized 49 benchmarks, but the plot in fig 8 only
shows 38/39?

Review #18C
===========================================================================

Overall merit
-------------
3. Weak reject - will not argue against

Reviewer expertise
------------------
Z. Outsider

Paper summary
-------------
Overall, this paper looks like a solid and useful contribution to the lens
literature. However, this reviewer has never used, let alone designed, a lens
library. As such, I found the paper very hard to read and assess. I would have
liked to see a clear problem statement (such as an example where it's annoying
to write the Boomerang code but easy to synthesize the lens using your approach
- why is there no Boomerang code for your example in Sec. 3?). It was also not
very clear to me what subset of Boomerang your approach supports. The
introduction says it is a "useful subset of Boomerang" - but what does that
mean?

I was also a bit confused about the preliminaries in Sec. 2. In particular, it
is not clear whether all or most of those preliminaries are from related work
(but there are no references) or whether they are contributions of this paper.

Minor comments:

- p.4, "naively using it in the context of lens synthesis presents several
problems"? Which problems? At this point I have no clue what "using it" may
even mean. The "big picture" is not clear at all.

- p. 3, "We have ommitted these theorems for space..." - maybe you mean
"proofs"?

-p. 7, "instead of maintaining server code for each endpoint, we envision..."
- but surely the changes to an API involve more than just data format changes.
For instance, single calls may have to be split into multiple calls. Hence it
does not seem to be realistic to maintain the old API using _only_ synthesized
lenses.

### 2017-09-11 papers/BijectiveLensSynth/popl_rebuttal_strategy_notes.txt
add in some popl rebuttal notes

Summarize points they all raise, and questions for rebuttal:

Quotient lenses allow for removing whitespace

Reviewer 1:
Specify and talk about QREs, transformations can be factored into 3 parts,
and we have a way of specifying 2 of those parts.  1 part that transforms
data into a canonical form, transformations into canonical forms and then
same transforms as usual in the middle.
We've experimented with this and our current inference algorithm works
off the shelf with it.

Rebuttal to reviewer B issues is that:
Are there any xforms that we can synth and flashfill can't
Start with that..., talk about nested iterations etc.
One of largest strengths is that it can synthesize really complex ones that
can't be synthesized by FF

Point to thing thing in Excel, as stated it can't synthesize nested stars

The DSL allows for nested stars, but it doesn't seem to work

Copy & paste text of their questions,
Differences between FF and Optician
How do we transform the examples, describe changes
Any chance of automating this or handling, cut & paste question

	2 major classes of changes we had to make: 1 is whitespace, other is
 	  projection of data
	Bijections have to preserve the amount of whitespace, this can be
	  automated by putting a wrapper around the xform that normalizes
	  the whitespace
	Our current system does that (this is called a quotient lens)
	Second transformation is forgetting information, and then restoring
	  after a round-trip.  We again believe the core of our algorithm
	  can be reused for this, and we are working on this still, but it's
	  future work.  Say how we modified the examples - added in the
	  extraneous data.  Again, a wrapper that deleted that and restored
	  it on the reverse transformation could handle it and is not hard
	  to write - believe can be automated, but need to do the research.

Reviewers are worried we're cheating about experiments

	Try and avoid subjective wording
	Concrete facts about what's going on
	In whitespace, we're making sure that each side has same amount of
	  whitespace

### 2017-09-11 papers/BijectiveLensSynth/popl_18_rebuttal.txt
add in initial rebuttal

@@
 We thank the reviewers for their feedback.
+
+*Talk about where we and FlashFill are different in expressivity of DSL*
+First, we detail the differences between the DSL of FlashFill and our DSL.
+FlashFill allows for non-bijective transformations, which lets it express
+transformations which project information.
+
+Our DSL is less expressive than Flash Fill, but encodes reverse transformations
+as well.  Furthermore, while our DSL is less expressive, we are able to
+synthesize significantly more transformations than Flash Fill.  In particular,
+Flash Fill has difficulties synthesizing transformations with nested iterations.
+In the original Flash Fill paper, the tool does not even attempt to synthesize
+terms with nested iterations for performance reasons (the last sentence of
+Section 4.4).  We could not find documentation detailing whether the same
+restriction of non-nested iterations is present in the version of Flash Fill
+bundled in the Prose API.  While the language detailed in the Flash Fill can
+express nested iterations we found it could not synthesize any, significantly
+limiting its capabilities for synthesizing the types of transformations we
+encountered in Augeas.
+
+*Talk about why bijective restriction is not an issue*
+We simplified our task to handling bijective lenses as this is an initial
+approach into this area.  This restriction did require some alterations of the
+benchmarks.  These alterations fall in 2 major categories:
+1) Whitespace was present in one format but not the other.
+2) Information was projected away in the original example.
+
+Category 1 was by far the most prevalent, and encountered in many of the Augeas
+examples.  To handle this, we added additional whitespace in the other format.
+However, this process can be automated using quotient lenses.  Quotient lenses
+canonize strings into representative elements: a large amount of whitespace
+would be canonized into a single whitespace character.  Putting these canonizers
+at the edges would allow us to not need to put the additional whitespace
+information in the target language.  We have since developed a language to
+express canonizers within regular expressions, providing the ability to write a
+"whitespace token" that allows for whitespace to be input, but also canonizes
+that whitespace intended format.
+
+Category 2 was less prevalent, but still present, typically in the FlashFill
+examples.  To address this category, we added additional information to the end
+of the target format containing the projected information.  Composing this
+generated lens with one that deletes that information at the tail of the file,
+when going left to right, and restores that information going back would provide
+the full lens.  This wrapper is not very difficult to write, and we think that
+discovering this projection automatically is possible, but we still need to do
+the research on it.
+
+This research we believe can be used in future work to solve these problems.
+This was our primary motivation to solving this first, if we can solve this
+problem we can use the generalize to more complex problems, but this step
+is an important one in making those more general problems more tractable.
+
+*Talk about how writing down the regular expressions is easier than writing down
+the lenses (for B and C)*
+
+Thinking about the ambiguity of concatenations and unions of the languages is
+hard to think about.  
+One example of this is highlighted in the BibTeX to EndNote author fields.
+
+      let bibtex_to_endnote_au
+        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
+        del "author={"
+          . ins "au - "
+          . lens_swap (NAME . del ",") (lens_swap WSP NAME)* 
+          . (del " and "
+               . ins "\n au - "
+               . lens_swap (NAME . del ",") (lens_swap WSP NAME)* )* 
+          . del "}"
 
@@
 
-+ Interesting and potentially useful application of program synthesis to bidirectional programming
-+ Nontrivial design/technical development
-+ Approach implemented and evaluated on real(ish) examples
 - Restriction of focus to bijective "lenses" not clearly justified
@@
 But the main part of the paper is not self-contained - for example the
-algorithm description relies on (example-dependent) orderings of sequences and atoms that
-are not described; this is a key part of the algorithm since it is
+algorithm description relies on (example-dependent) orderings of sequences and
+atoms that are not described; this is a key part of the algorithm since it is
 where the provided examples are actually used.
+
+    We omitted the formal description of these orderings for space.  A full
+    description of the algorithm is present in algorithm 4 in the appendix, on
+    page 119, with the orderings defined in Definition 33 on page 118 (the
+    definitions of related structures begin at page 115).  We embed the parse
+    trees of the examples into the DNF regular expressions, and make an ordering
+    on the DNF regular expression with embedded parse trees.
 
@@
 
+    We merely included that syntax as a means to make the derivation more
+    readable.  One can evaluate those operators, and it will be an arbitrary
+    DNF regular expression, we just use the operators to show the relationship
+    between the atom involved in the structural rewrite, and what it is
+    rewritten to.
+
 p17.  The desription of RigidSynth is incomplete.  It mentions orders
@@
 
+    This is correct.
+
 p19: "easy to understand" - can you present any evidence for this?
 Easy for whom?
+
+    We feel that they are easy to understand because they look similar to
+    handwritten lenses, but we have not done any user studies to justify that
+    they are easy to understand.  Evidence to the fact that they are easier to
+    understand than non-pretty printed versions is given by the sizes of the
+    generated lenses.  Without minimization, the lenses sizes averaged at (SIZE)
+    AST nodes, whereas minimization made the sizes average at (SIZE) ast nodes.
 
@@
 
+    Our DSL contains transformations that cannot be expressed by Extended Symbolic
+    Finite Transducers.  The finite lookahead does not allow for transformations
+    which swaps the locations of data, a common pattern in Boomerang lenses.
+    We will add a citation to this work and a comment detailing the restrictions
+    extended symbolic finite transducers put in place.
+
 - Somewhat limited evaluation. The evaluation is performed on two families of
@@
 FlashFill can reverse a string, but the proposed language can’t.
+
+    Boomerang doesn't currently provide such capabilities.  However, there is no
+    technical reason this can't be done by adding in a reverse-and-iterate
+    operator.  Furthermore, we can
 
@@
 
+    The text is wrong, there are only 39 benchmarks.
+
 Review #18C
@@
 to write the Boomerang code but easy to synthesize the lens using your approach
+
 - why is there no Boomerang code for your example in Sec. 3?). It was also not
@@
 
+    Correct
+
 -p. 7, "instead of maintaining server code for each endpoint, we envision..."
@@
 lenses.
+
+    Yes, there are certainly times where more than lenses needs to be added.
+    However, the specific instance of single calls being split into multiple
+    calls can be handled by this paradigm if the server allows for batch
+    requests.  In particular, APIs of the running example we use, the modern
+    format does have batching capabilities because of this type of issue,
+    https://www.visualstudio.com/en-us/docs/integrate/api/wit/batch .

### 2017-09-11 papers/BijectiveLensSynth/popl_18_rebuttal.txt
add rebuttal update

@@
 *Talk about where we and FlashFill are different in expressivity of DSL*
+*Talk about how we are /BETTER/ than them*
+
 First, we detail the differences between the DSL of FlashFill and our DSL.
@@
 
-*Talk about why bijective restriction is not an issue*
+*Talk about why bijective restriction is not an issue slash did we scam
+reviewers*
 We simplified our task to handling bijective lenses as this is an initial
@@
 
+*Discuss future work, how do we disambiguate, that makes it easier*
+
 Thinking about the ambiguity of concatenations and unions of the languages is
@@
 - Restriction of focus to bijective "lenses" not clearly justified
+
+The focus on bijective lenses is twofold.  Firstly, there are a large number of
+formats which are in bijective correspondence: for example https://www.data.gov/
+provides data in a variety of formats in bijective correspondence with each
+other.  Secondly, bijective lenses can be viewed as the core of many other types
+of lenses.  For example, standard lenses can be viewed in two parts: a bijective
+part that moves the location of data around and adds in boilerplate, and a
+projecting part that removes data in one direction, and adds it back in when
+transforming in the reverse direction.  This approach can further be extended to
+symmetric lenses by having that projection component on each side of the
+transformation.  Quotient lenses can also be viewed in two parts: a
+transformation component and a canonizing component.  This transformation
+component could be a bijective core that is transformed, and the canonizing
+component appears surrounding it.  In this way, finding a synthesis algorithm
+for bijective lenses provides great progress in synthesizing more complex
+lenses, it can be plugged into other systems in a principled way to extend the
+algorithm to more general lens forms.
+
 - Important parts of algorithm not explained (regarding how examples are used)
-- Not clear how much work it is to adapt examples to be suitable for bijective lens synthesis, or how useful the results are w.r.t. the original task
 
-But the main part of the paper is not self-contained - for example the
+We omitted the formal description of these orderings for space.  A full
+description of the algorithm is present in algorithm 4 in the appendix, on page
+119, with the orderings defined in Definition 33 on page 118 (the definitions of
+related structures begin at page 115).  We embed the parse trees of the examples
+into the DNF regular expressions, and make an ordering on the DNF regular
+expression with embedded parse trees.
+
+- Not clear how much work it is to adapt examples to be suitable for bijective
+  lens synthesis, or how useful the results are w.r.t. the original task
+
+We simplified our task to handling bijective lenses as this is an initial
+approach into this area.  This restriction did require some alterations of the
+benchmarks.  The alterations were adding data into the target format that wasn't
+there originally.  These alterations fall in 2 major categories:
+1) Whitespace was present in one format but not the other.
+2) Information was projected away in the original example.
+
+Category 1 was by far the most prevalent, and encountered in many of the Augeas
+examples.  To handle this, we added additional whitespace in the other format.
+However, this process can be automated using quotient lenses.  Quotient lenses
+canonize strings into representative elements: a large amount of whitespace
+would be canonized into a single whitespace character.  Putting these canonizers
+at the edges would allow us to not need to put the additional whitespace
+information in the target language.  We have since developed a language to
+express canonizers within regular expressions, providing the ability to write a
+"whitespace token" that allows for whitespace to be input, but also canonizes
+that whitespace intended format.
+
+Category 2 was less prevalent, but still present, typically in the FlashFill
+examples.  To address this category, we added additional information to the end
+of the target format containing the projected information.  Composing this
+generated lens with one that deletes that information at the tail of the file,
+when going left to right, and restores that information going back would provide
+the full lens.  This wrapper is not very difficult to write, and we think that
+discovering this projection automatically is possible, but we still need to do
+the research on it.
+
+- But the main part of the paper is not self-contained - for example the
   algorithm description relies on (example-dependent) orderings of sequences and
@@
 
-    We omitted the formal description of these orderings for space.  A full
-    description of the algorithm is present in algorithm 4 in the appendix, on
-    page 119, with the orderings defined in Definition 33 on page 118 (the
-    definitions of related structures begin at page 115).  We embed the parse
-    trees of the examples into the DNF regular expressions, and make an ordering
-    on the DNF regular expression with embedded parse trees.
+See above
 
-I also have a concern: sicne the approach is
+- I also have a concern: sicne the approach is
   limited to bijective lenses, but many of the test examples were not
@@
 
-I could see future work building on this to handle the non-bijective case 
+See above
+
+- I could see future work building on this to handle the non-bijective case 
   being a separate (and also interesting) paper.  But I would want the algorithm
@@
 
-p12 fig 6. I'm not sure how ule "DNF structural rewrite" could be
+It is difficult to go into more formal detail about the orderings within the
+page limits.  We will provide a more detailed informal description, and include
+an explicit reference to the extended version in the final paper.
+
+- p12 fig 6. I'm not sure how ule "DNF structural rewrite" could be
   applied as the LHS uses the operators defined in fig. 5, rather than syntax
@@
 
-p17.  The desription of RigidSynth is incomplete.  It mentions orders
+- p17.  The desription of RigidSynth is incomplete.  It mentions orders
   $\leq_{Seq}$ and $\leq_{Atom}$ (and variants that use the examples) that do
@@
 
-p18: AllSome (alg. 3): I assume this is the obvious function of type
+We can go into more detail in the paper on this.  If two subcomponents can each
+map to two subcomponents on the opposite side, a permutation which does not swap
+the order of those subcomponents is best.
+
+- p18: AllSome (alg. 3): I assume this is the obvious function of type
   'a option list -> 'a list option (that is None if there is a None in
@@
 
-p19: "easy to understand" - can you present any evidence for this?
+- p19: "easy to understand" - can you present any evidence for this?
   Easy for whom?
 
-    We feel that they are easy to understand because they look similar to
-    handwritten lenses, but we have not done any user studies to justify that
+We feel that they are easy to understand because they look similar to how we
+would handwrite the lenses,
+but we have not done any user studies to justify that
 they are easy to understand.  Evidence to the fact that they are easier to
@@
 
-p20: "We adapted these examples" please say mor about how many of the
+- p20: "We adapted these examples" please say mor about how many of the
   examples needed to be adapted (all? a handful?) and how much work it
@@
 
-p20: "We developed a series of optimizations" - clarify that the
+All of the examples taken from existing programs needed to be adapted in some
+way.  In particular, while there were many examples that didn't have projected
+information that would have necessitated projection aspects of lens
+capabilities, those examples all used whitespace.
+
+- p20: "We developed a series of optimizations" - clarify that the
   algorithm already described includes all of these, and what you are
@@
   weren't discussed earlier.  Might be helpful to signpost what parts of
-the algorithm are the "basic" algorithm and what parts are optional optimizations
+  the algorithm are the "basic" algorithm and what parts are optional
+  optimizations.
 
-p21: Fig 9 "Number of expansions" --> "Number of solutions"??
+The algorithm described includes all but compositional synthesis.  We don't
+include in the formal algorithm the means to which we use previously defined
+lenses to synthesize new ones.  The way we do it is we develop equivalence
+classes between user defined data types that have lenses, choose a
+representative, and transform the user defined data types into that
+representative, and convert the strings into elements of that representative
+data type.
 
-p21.  Also interesting to evaluate would be how time / number of
+- p21: Fig 9 "Number of expansions" --> "Number of solutions"??
+
+Number of expansions is the correct phrase.  Figure 9 shows how much random
+search through expansions is required before RigidSynth terminates.  In
+particular, this shows that full expansion inference performs significantly
+better than the partial one that only performs the expansions it knows it must
+perform.
+
+- p21.  Also interesting to evaluate would be how time / number of
   solutions varies (e.g. for the full algorithm) as the number of
@@
 
-p24: "because our types do not have full canonical forms" - this is
+The number of examples provided doesn't typically have an impact on runspeed,
+only correctness.  The number of valid solutions does vary with the number of
+examples provided.  Typically, a small number of examples can greatly reduce the
+number of possible solutions (TODO data).  The number of examples is indeed
+small, as is shown in Figure 11.  An active approach we agree is an
+interesting one, and one we too have considered.
+
+- p24: "because our types do not have full canonical forms" - this is
   the case for the equations presented in the paper, but what about
   e.g. translating to automata and minimizing?
+
+You are correct, regular expressions do have canonical forms in that manner.
+We will change the phrasing to more accurately represent our meaning.
 
@@
 
-However, I’m not sure what exact advantage the proposed approach brings to the
+- However, I’m not sure what exact advantage the proposed approach brings to the
   table. In particular, the user is still required to write exact input and output
   types of the transformations via regular expressions. How hard is to write
-lenses after one has such expressions? I’ll further detail my comparison now.
-Also, the language has big limitations in terms of expressiveness.
-
-### CONS:
-- In the reviewer’s opinion, the specification mechanism is not very natural and
+  lenses after one has such expressions? ...
+  In the reviewer’s opinion, the specification mechanism is not very natural and
   it’s unclear whether it actually simplifies writing bijective transformations.
@@
   there is no report of the sizes of benchmarks.
+
+We provide an example below on mapping the author field of BibTeX to EndNote:
+
+      typedef NAME = UPPERCASE (LOWERCASE)*;;
+
+      typedef LASTCOMMASTART = NAME "," (WSP NAME)*;;
+      typedef BIBTEXAUTHORLIST = LASTCOMMASTART (" and " LASTCOMMASTART)*;;
+      typedef BIBTEXAUTHORINFO = "author={" BIBTEXAUTHORLIST "}";;
+
+      typedef AUTAG = "au - ";;
+      typedef STARTTOEND = (NAME WSP)* NAME;;
+      typedef TAGGEDAUTHORDEFNS = AUTAG STARTTOEND ("\n " AUTAG STARTTOEND)*;;
+
+      let bibtex_to_endnote_au_synthed =
+        synth
+          BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS
+
+      let bibtex_to_endnote_au
+        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
+        del "author={"
+          . ins "au - "
+          . lens_swap (NAME . del ",") (lens_swap WSP NAME)* 
+          . (del " and "
+               . ins "\n au - "
+               . lens_swap (NAME . del ",") (lens_swap WSP NAME)* )* 
+          . del "}"
+
+      let bibtex_to_endnote_au_human_written
+        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
+        let single_author_convert =
+          ins " au - "
+            . lens_swap
+                (NAME . del ",")
+                (lens_swap WSP NAME)*
+        in
+        del "author={"
+          . single_author_convert
+          . (del " and "
+               . ins "\n"
+               . single_author_convert)* 
+          . del "}"
+
+We feel the specification is easier to read and write than the explicit lens.
+Furthermore, even the relatively simple permutation of only 3 elements needed
+for single_author_convert is fairly complex.  Beyond this, we find it is best
+practices to include the type explicitly when writing the lens, both as
+documentation and as additional assurance to correctness.  As we already spend
+the effort writing these types manually for these benefits, there is no
+additional work needed to use our synthesis algorithm, there is instead only
+saved effort.
+
+We provide the size of specifications and lenses below:
+
+--------------------------------------------
+|Test                | SpecSize | LensSize |
+|------------------------------------------|
+|cust/date-probs     | 85       | 79       |
+|cust/cap-prob       | 112      | 108      |
+|cust/2-cap-prob     | 122      | 66       |
+|ff/extr-fname-err   | 143      | 151      |
+|ff/extr-fname       | 149      | 168      |
+|ff/extr-quant       | 150      | 145      |
+|ff/extr-num         | 171      | 166      |
+|aug/activemq        | 194      | 184      |
+|cust/bib-prob       | 202      | 181      |
+|cust/workitem-probs | 221      | 252      |
+|aug/xml-l1          | 235      | 214      |
+|cust/addr-probs     | 244      | 239      |
+|aug/aptcrngsec      | 265      | 204      |
+|aug/alias           | 271      | 210      |
+|aug/backuppchosts   | 272      | 249      |
+|aug/approx          | 277      | 264      |
+|aug/apt-upd-mgr     | 278      | 210      |
+|aug/aptsources      | 278      | 238      |
+|aug/cachefilesd     | 294      | 233      |
+|aug/aptprefs        | 297      | 280      |
+|aug/afs-cellalias   | 298      | 246      |
+|aug/avahi           | 301      | 234      |
+|aug/chrony          | 308      | 289      |
+|aug/xml             | 311      | 383      |
+|aug/carbon          | 325      | 377      |
+|aug/access          | 335      | 253      |
+|aug/anacron         | 342      | 315      |
+|aug/hosts           | 343      | 328      |
+|aug/bootconf        | 350      | 271      |
+|aug/auth-keys       | 354      | 424      |
+|aug/cgrules         | 378      | 295      |
+|aug/cgconfig        | 415      | 347      |
+|aug/automaster      | 475      | 377      |
+|aug/cron            | 475      | 414      |
+|aug/channels        | 512      | 367      |
+|aug/automounter     | 535      | 432      |
+|aug/aptconf         | 540      | 450      |
+|aug/aptconf-l1      | 564      | 503      |
+|aug/bbhosts         | 670      | 651      |
+--------------------------------------------
+
+We can make sure we can provide at least some description of the sizes of specs
+and lenses in the paper.
 
@@
   on the presented data, I cannot assess how often the presented technique could
-be applied. Next, the comparison with the tool FlashFill/FlashExtract is somewhat
+  be applied.
+
+- Next, the comparison with the tool FlashFill/FlashExtract is somewhat
   confusing. I suspect the authors had to provide some domain and range regular
   expressions for the flash fill benchmarks, but this aspect, which is crucial to
-help the search, is not mentioned. What happens if someone provides Sigma* and
-Sigma* as domains and ranges to Optician. Also, a better explanation of the
+  help the search, is not mentioned.
+
+  We do require regular expressions describing the input of FlashFill specs, we
+  will make sure that is more explicitly stated in the paper.
+
+- What happens if someone provides Sigma* and Sigma* as domains and ranges to
+  Optician.
+
+If someone provides Sigma* and Sigma* as domains and ranges to optician it
+will merely return the identity between those two.  If a user presents Sigma*
+and Sigma* with some examples, we expect it will unroll the iterations until
+the number of characters in the largest example has been enumerated, and
+return something close to identity, but performing differently on those
+examples, and some others to handle the readjustment.
+
+Optician is meant to work on finding mappings between the regular expressions
+specified.  It is not intended to find subsets of the provided regular
+expressions that can be mapped.
+
+- Also, a better explanation of the
   difference between the DSLs of Lenses and Flash fill is required. For example,
@@
 technical reason this can't be done by adding in a reverse-and-iterate
-    operator.  Furthermore, we can
+operator to Boomerang.  Only a slight adjustment to the synthesis algorithm
+would be needed to incorporate this into the algorithm.
 
-### QUESTIONS FOR REBUTTAL
 - What is the difference in size between the regexes for domain/range spec and
   the full lenses?
+
+On average, the domain and range specs are slightly larger than the lens itself.
+
+We do not think that this means the work is useless, but rather it is used to
+alleviate the pain of reasoning about nested swaps and concatenations, allow for
+more simple maintenance of the code.
+
+We provide an example below on mapping the author field of BibTeX to EndNote:
+
+      typedef NAME = UPPERCASE (LOWERCASE)*;;
+
+      typedef LASTCOMMASTART = NAME "," (WSP NAME)*;;
+      typedef BIBTEXAUTHORLIST = LASTCOMMASTART (" and " LASTCOMMASTART)*;;
+      typedef BIBTEXAUTHORINFO = "author={" BIBTEXAUTHORLIST "}";;
+
+      typedef AUTAG = "au - ";;
+      typedef STARTTOEND = (NAME WSP)* NAME;;
+      typedef TAGGEDAUTHORDEFNS = AUTAG STARTTOEND ("\n " AUTAG STARTTOEND)*;;
+
+      let bibtex_to_endnote_au_synthed =
+        synth
+          BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS
+
+      let bibtex_to_endnote_au
+        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
+        del "author={"
+          . ins "au - "
+          . lens_swap (NAME . del ",") (lens_swap WSP NAME)* 
+          . (del " and "
+               . ins "\n au - "
+               . lens_swap (NAME . del ",") (lens_swap WSP NAME)* )* 
+          . del "}"
+
+      let bibtex_to_endnote_au_human_written
+        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
+        let single_author_convert =
+          ins " au - "
+            . lens_swap
+                (NAME . del ",")
+                (lens_swap WSP NAME)*
+        in
+        del "author={"
+          . single_author_convert
+          . (del " and "
+               . ins "\n"
+               . single_author_convert)* 
+          . del "}"
+
+We feel the specification is easier to read and write than the explicit lens.
+Furthermore, even the relatively simple permutation of only 3 elements needed
+for single_author_convert is fairly complex.  Beyond this, we find it is best
+practices to include the type explicitly when writing the lens, both as
+documentation and as additional assurance to correctness.  As we already spend
+the effort writing these types manually for these benefits, there is no
+additional work needed to use our synthesis algorithm, there is instead only
+saved effort.
+
 - What is the difference in size between manually written lenses and synthesized
   ones?
+
+We have not manually written lenses for all our benchmarks.  We find the sizes
+are fairly comparable, but the hand-written lenses are more modular: with common
+portions taken out and given good names.
+
 - Clarify how the benchmarks were selected and whether any were not selected a
@@
   transformations such as upper-case?
+
+The first three (extr-fname, extr-fname-err and extr-quant) we chose because
+they corresponded to the first two examples in the paper.
+The problem extr-fname only allowed input paths with a file at the end, where
+extr-fname-err provided an error message if no file was present.
+The final example, extr-num was used because extracting a phone number was used
+as an extended example.  We could not find any examples in the paper which
+applied a to_upper conversion.
+
 - Can the proposed DSL express any transformation that FlashFill can’t and
   vice-versa?
+
+While Flash Fill's DSL permits nested iterations, for performance reasons Flash
+Fill only synthesizes a single nesting of loops.  While this is sufficient for
+the types of applications present in an Excel file, the examples we found in
+Augeas were significantly more complex, requiring complex transformations with 
+many nested iterations.
+
+Flash Fill can express many transformation our DSL cannot.  For example, we
+currently cannot handle losing information, which prevents us from projecting
+information.
+
 - P19: maximally factors the concats and ors. What does this mean? Regex
@@
   would help to show one example of fully synthesized lens in the paper.
+
+Agreed, we will finish our extended example to include a fully synthesized
+title field lens.
+
+In the title field, without minimization, the title field transformation would
+appear as:
+
+const("<Field Id=2></Field>","")
+  | (const("<Field Id=2>","Title: ")
+      . Id(text_char)
+      . const("","")
+      . iterate(const("","") . Id(text_char) . const("",""))
+      . const("","")
+      . const("</Field>",", "))
+
+With minimization, the title field transformation would appear as:
+
+const("<Field Id=2>","")
+  . (const("","")
+       | (Id(text_char) . iterate(Id(text_char)) . const("", ",")))
+  . const("</Field>",""))
+
+In particular, the useless concatenations of const("","") are removed and
+the field boilerplate is removed from being on both sides of the disjunct.
 
@@
 
-Overall merit
--------------
-3. Weak reject - will not argue against
-
-Reviewer expertise
-------------------
-Z. Outsider
-
-Paper summary
--------------
-Overall, this paper looks like a solid and useful contribution to the lens
+- Overall, this paper looks like a solid and useful contribution to the lens
   literature. However, this reviewer has never used, let alone designed, a lens
@@
 
-- why is there no Boomerang code for your example in Sec. 3?). It was also not
-very clear to me what subset of Boomerang your approach supports. The
-introduction says it is a "useful subset of Boomerang" - but what does that
-mean?
+We provide an example below on mapping the author field of BibTeX to EndNote:
 
-I was also a bit confused about the preliminaries in Sec. 2. In particular, it
+      typedef NAME = UPPERCASE (LOWERCASE)*;;
+
+      typedef LASTCOMMASTART = NAME "," (WSP NAME)*;;
+      typedef BIBTEXAUTHORLIST = LASTCOMMASTART (" and " LASTCOMMASTART)*;;
+      typedef BIBTEXAUTHORINFO = "author={" BIBTEXAUTHORLIST "}";;
+
+      typedef AUTAG = "au - ";;
+      typedef STARTTOEND = (NAME WSP)* NAME;;
+      typedef TAGGEDAUTHORDEFNS = AUTAG STARTTOEND ("\n " AUTAG STARTTOEND)*;;
+
+      let bibtex_to_endnote_au_synthed =
+        synth
+          BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS
+
+      let bibtex_to_endnote_au
+        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
+        del "author={"
+          . ins "au - "
+          . lens_swap (NAME . del ",") (lens_swap WSP NAME)* 
+          . (del " and "
+               . ins "\n au - "
+               . lens_swap (NAME . del ",") (lens_swap WSP NAME)* )* 
+          . del "}"
+
+      let bibtex_to_endnote_au_human_written
+        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
+        let single_author_convert =
+          ins " au - "
+            . lens_swap
+                (NAME . del ",")
+                (lens_swap WSP NAME)*
+        in
+        del "author={"
+          . single_author_convert
+          . (del " and "
+               . ins "\n"
+               . single_author_convert)* 
+          . del "}"
+
+We feel the specification is easier to read and write than the explicit lens.
+Furthermore, even the relatively simple permutation of only 3 elements needed
+for single_author_convert is fairly complex.  Beyond this, we find it is best
+practices to include the type explicitly when writing the lens, both as
+documentation and as additional assurance to correctness.  As we already spend
+the effort writing these types manually for these benefits, there is no
+additional work needed to use our synthesis algorithm, there is instead only
+saved effort.
+
+- why is there no Boomerang code for your example in Sec. 3?).
+
+We can include example code like this to highlight the frustrations a lens
+programmer must encounter while writing their programs.
+
+- It was also not very clear to me what subset of Boomerang your approach
+  supports. The introduction says it is a "useful subset of Boomerang" - but
+  what does that mean?
+
+The subset our approach supports are the bijective lenses where all regular
+expression equivalences used in the typing derivation can be proven with
+the star-semiring equivalence rules.  We find this useful in that all lenses can
+easily be converted into a bijective lens, and we could not find a lens that
+requires a courser notion of equivalence than star-semiring equivalence.
+
+- I was also a bit confused about the preliminaries in Sec. 2. In particular, it
   is not clear whether all or most of those preliminaries are from related work
@@
 
-Minor comments:
+The preliminaries are us defining a specific subset of Boomerang that we focus
+on.  Boomerang supports a variety of lenses - asymmetric lenses, quotient
+lenses, and alignment lenses.  We define which subset of Boomerang we aim to
+support.
 
@@
   even mean. The "big picture" is not clear at all.
+
+In this context, "using it" refers to allowing retyping on the specified
+equivalences.  We are trying to carve out a specific subset of Boomerang that
+we want to synthesize, and are not allowing that subset to be included.
+The problems with including full language equivalence is detailed on pages 8
+through 10.
 
@@
 
--p. 7, "instead of maintaining server code for each endpoint, we envision..."
-- but surely the changes to an API involve more than just data format changes.
+- p. 7, "instead of maintaining server code for each endpoint, we envision..."
+  but surely the changes to an API involve more than just data format changes.
   For instance, single calls may have to be split into multiple calls. Hence it
@@
     Yes, there are certainly times where more than lenses needs to be added.
-    However, the specific instance of single calls being split into multiple
+    The specific instance of single calls being split into multiple
     calls can be handled by this paradigm if the server allows for batch
     requests.  In particular, APIs of the running example we use, the modern
-    format does have batching capabilities because of this type of issue,
+    format does have batching capabilities,
     https://www.visualstudio.com/en-us/docs/integrate/api/wit/batch .

### 2017-09-12 papers/BijectiveLensSynth/popl_18_rebuttal.txt
add the intro

@@
-We thank the reviewers for their feedback.
+We thank the reviewers for their feedback.  We will provide a common rebuttal
+to address concerns common to many reviewers, and will then have an appendix
+addressing specific concerns.
 
-*Talk about where we and FlashFill are different in expressivity of DSL*
-*Talk about how we are /BETTER/ than them*
+- What is the difference between our work and existing work, like Flash Fill,
+  in terms of expressivity?
 
-First, we detail the differences between the DSL of FlashFill and our DSL.
-FlashFill allows for non-bijective transformations, which lets it express
-transformations which project information.
+We are able to synthesize some transformations that Flash Fill cannot, and
+Flash Fill is able to synthesize some transformations that we cannot.
+In particular, by restricting ourselves to bijections, we are unable to express
+a large class of functions.  In alternative string synthesis systems, this
+bijective restriction is not in place, and additional strings can be
+synthesized.  However, we do not see this as a failure, but as a design choice.
+Our system is not targeting single direction transformations: we feel
+programmers already have the tools they need to write these by hand.  Instead,
+we are aiming to alleviate the difficulties in programming bidirectional
+transformations, which have languages that are very difficult to write in.
 
-Our DSL is less expressive than Flash Fill, but encodes reverse transformations
-as well.  Furthermore, while our DSL is less expressive, we are able to
-synthesize significantly more transformations than Flash Fill.  In particular,
-Flash Fill has difficulties synthesizing transformations with nested iterations.
-In the original Flash Fill paper, the tool does not even attempt to synthesize
-terms with nested iterations for performance reasons (the last sentence of
-Section 4.4).  We could not find documentation detailing whether the same
-restriction of non-nested iterations is present in the version of Flash Fill
-bundled in the Prose API.  While the language detailed in the Flash Fill can
-express nested iterations we found it could not synthesize any, significantly
-limiting its capabilities for synthesizing the types of transformations we
-encountered in Augeas.
+However, Optician can still solve certain classes of problems other string
+synthesis systems cannot.  In particular, existing techniques that only use
+input-output examples are unable to synthesize programs with nested loops.
+In particular, while Flash Fill's DSL does permit nested loops, for efficiency
+reasons, the synthesis algorithm only synthesizes programs with no nesting (this
+is stated at the end of section 4.4).  While a single nesting of loops is
+sufficient for the types of problems encountered in transforming Excel
+spreadsheet data, our tool is aimed at building large transformations between
+highly complex formats.  The types of formats Augeas is built to handle have
+many deep layers of nesting, with very complex transformations at each layer.
+Nested iteration is necessary for these types of transformations, and existing
+techniques are insufficient for synthesizing them.
 
-*Talk about why bijective restriction is not an issue slash did we scam
-reviewers*
+- Why did we make the bijective restriction, and what transformations did we
+  make to our benchmark suite to make them bijective.
+
 We simplified our task to handling bijective lenses as this is an initial
-approach into this area.  This restriction did require some alterations of the
-benchmarks.  These alterations fall in 2 major categories:
+approach into this research.  There are a large number of
+formats which are in bijective correspondence: for example https://www.data.gov/
+provides data in a variety of formats in bijective correspondence with each
+other.  However, we began with synthesizing bijective
+lenses primarily because we believe this bijective core can be extended in a
+principled way to handle more complex lenses.
+
+Many of our transformations were nearly bijective, but had unnecessary
+information, like whitespace or casing, that prevented them from being bijective.
+Quotient lenses provide the capabilities to ignore unnecessary information like
+whitespace.  While the combinators for quotient lenses, canonizers, can be
+intermixed with basic lenses in the original paper, we believe that a
+"canonizers-at-the-edge" approach is just as expressive.  In particular,
+we can express a transformation that is bijective up to a quotient by having a
+bijective lens core, with a canonizer at each side that normalizes the input.
+Indeed, we already have a prototype that does this, allowing specialized
+tokens that encodes both the full description of viable inpout as well as the
+information on how to canonize the data.
+
+The remaining transformations that needed to be altered were also nearly
+bijective, but projected away useful information, and this projection prevented
+them from being bijective.  Unlike with quotient lenses, we do not want to
+remove this information - we want the information to be preserved through
+round-trips.  Furthermore, there are formats where each format contains
+information not present in the other format.  Symmetric lenses provide the
+capabilities for writing such transformations.  Similarly to quotient lenses,
+projection combinators can be intermixed with bijection capabilities,
+but we can view symmetric lenses as having a bijective core with a projection
+combinator on each side that removes data in one direction, and restores it in
+the other.
+
+While we simplified our task to bijective lenses, we did this because we felt
+a synthesis algorithm for bijective lenses could form the core of a synthesis
+algorithm for other types of lenses.  We agree this motivation looking towards
+future work is currently lacking in the paper, and we can add this in for the
+final version.
+
+The changes we put on our benchmark suite fall in 2 major categories that mirror
+the intended extensions to our system.
 1) Whitespace was present in one format but not the other.
-2) Information was projected away in the original example.
+2) Useful information was projected away going from one format to the other.
 
 Category 1 was by far the most prevalent, and encountered in many of the Augeas
-examples.  To handle this, we added additional whitespace in the other format.
-However, this process can be automated using quotient lenses.  Quotient lenses
-canonize strings into representative elements: a large amount of whitespace
-would be canonized into a single whitespace character.  Putting these canonizers
-at the edges would allow us to not need to put the additional whitespace
-information in the target language.  We have since developed a language to
-express canonizers within regular expressions, providing the ability to write a
-"whitespace token" that allows for whitespace to be input, but also canonizes
-that whitespace intended format.
+examples.  To handle this, we added additional whitespace to the other format.
+This can be solved by using quotients on each side.  We think the strategy of
+using specialized tokens to represent both a language and a canonizer can solve
+this problem.
 
@@
 
-This research we believe can be used in future work to solve these problems.
-This was our primary motivation to solving this first, if we can solve this
-problem we can use the generalize to more complex problems, but this step
-is an important one in making those more general problems more tractable.
+- After having written the regular expressions, is it very difficult for users
+  to then write the lens.  Are we saving anybody any significant effort.
 
-*Talk about how writing down the regular expressions is easier than writing down
-the lenses (for B and C)*
+Engineering best practices dictate that we annotate the lenses with their types.
+While Boomerang does not need these annotations - being able to infer the types
+from the terms - the types of the term serve as documentation for
+future programmers to understand what formats the lens maps between.
+Furthermore, it provides an additional resilience to future modifications of the
+lens, as it validates that the lens maps exactly between strings of the provided
+formats - invalid inputs are not accepted nor are the outputs ever invalid.
+In this way, we are not requiring additional work by the programmer to use our
+system; we are allowing the programmer to save the difficulties of
+writing the term by using work they should already be performing.  To highlight
+these difficulties, we show an example program for mapping the author field of
+a BibTeX entry.
 
-*Discuss future work, how do we disambiguate, that makes it easier*
+      typedef NAME = UPPERCASE (LOWERCASE)*;;
 
-Thinking about the ambiguity of concatenations and unions of the languages is
-hard to think about.  
-One example of this is highlighted in the BibTeX to EndNote author fields.
+      typedef LASTCOMMASTART = NAME "," (WSP NAME)*;;
+      typedef BIBTEXAUTHORLIST = LASTCOMMASTART (" and " LASTCOMMASTART)*;;
+      typedef BIBTEXAUTHORINFO = "author={" BIBTEXAUTHORLIST "}";;
 
-      let bibtex_to_endnote_au
+      typedef AUTAG = "au - ";;
+      typedef STARTTOEND = (NAME WSP)* NAME;;
+      typedef TAGGEDAUTHORDEFNS = AUTAG STARTTOEND ("\n " AUTAG STARTTOEND)*;;
+
+      let bibtex_to_endnote_au =
+        synth
+          BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS
+
+      let bibtex_to_endnote_au_synthesized_result
         : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
@@
 
-...
+      let bibtex_to_endnote_au_human_written
+        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
+        let single_author_convert =
+          ins " au - "
+            . lens_swap
+                (NAME . del ",")
+                (lens_swap WSP NAME)*
+        in
+        del "author={"
+          . single_author_convert
+          . (del " and "
+               . ins "\n"
+               . single_author_convert)* 
+          . del "}"
+
+Even the relatively simple permutation of only 3 elements needed for
+single_author_convert is fairly complex.  These becomes even more apparent when
+working with large lenses comprised of many sub-lenses being permuted.  Lenses
+provide great power with their invertibility guarantees, but they come at the
+cost of thinking about many fiddly details when writing the terms.  Worrying
+about these fiddly details while ALSO thinking about unambiguity restrictions
+is a very difficult task!
+
+This tool is aimed at different audiences than existing string transformation
+tools.  While existing tools allow the average end-user to generalize their
+examples to programs, Optician allows programmers to get the benefits of a
+powerful language while avoiding much of the complexity of programming in that
+language.  This paradigm allows for more complex programs to be able to be
+synthesized, the types of programs that necessitate programming in such a DSL,
+but also provide a burden on the user.  However, this burden is already one
+incured by programmers using best practices, so we are merely saving them the
+time spent on tedious work.
 

### 2017-09-12 papers/BijectiveLensSynth/popl_18_rebuttal.txt
add references to intro

@@
 
-The focus on bijective lenses is twofold.  Firstly, there are a large number of
-formats which are in bijective correspondence: for example https://www.data.gov/
-provides data in a variety of formats in bijective correspondence with each
-other.  Secondly, bijective lenses can be viewed as the core of many other types
-of lenses.  For example, standard lenses can be viewed in two parts: a bijective
-part that moves the location of data around and adds in boilerplate, and a
-projecting part that removes data in one direction, and adds it back in when
-transforming in the reverse direction.  This approach can further be extended to
-symmetric lenses by having that projection component on each side of the
-transformation.  Quotient lenses can also be viewed in two parts: a
-transformation component and a canonizing component.  This transformation
-component could be a bijective core that is transformed, and the canonizing
-component appears surrounding it.  In this way, finding a synthesis algorithm
-for bijective lenses provides great progress in synthesizing more complex
-lenses, it can be plugged into other systems in a principled way to extend the
-algorithm to more general lens forms.
+See general responses.
 
@@
 into the DNF regular expressions, and make an ordering on the DNF regular
-expression with embedded parse trees.
+expression with embedded parse trees.  While we think the formal description of
+these orderings is too long to be included in the paper, we can provide a more
+involved informal description in the text, and provide a reference to an
+extended version with the formal descriptions of the orderings.
 
@@
 
-We simplified our task to handling bijective lenses as this is an initial
-approach into this area.  This restriction did require some alterations of the
-benchmarks.  The alterations were adding data into the target format that wasn't
-there originally.  These alterations fall in 2 major categories:
-1) Whitespace was present in one format but not the other.
-2) Information was projected away in the original example.
-
-Category 1 was by far the most prevalent, and encountered in many of the Augeas
-examples.  To handle this, we added additional whitespace in the other format.
-However, this process can be automated using quotient lenses.  Quotient lenses
-canonize strings into representative elements: a large amount of whitespace
-would be canonized into a single whitespace character.  Putting these canonizers
-at the edges would allow us to not need to put the additional whitespace
-information in the target language.  We have since developed a language to
-express canonizers within regular expressions, providing the ability to write a
-"whitespace token" that allows for whitespace to be input, but also canonizes
-that whitespace intended format.
-
-Category 2 was less prevalent, but still present, typically in the FlashFill
-examples.  To address this category, we added additional information to the end
-of the target format containing the projected information.  Composing this
-generated lens with one that deletes that information at the tail of the file,
-when going left to right, and restores that information going back would provide
-the full lens.  This wrapper is not very difficult to write, and we think that
-discovering this projection automatically is possible, but we still need to do
-the research on it.
+See general responses.
 
@@
 
-See above
+See above.
 
@@
 
-It is difficult to go into more formal detail about the orderings within the
-page limits.  We will provide a more detailed informal description, and include
-an explicit reference to the extended version in the final paper.
+We agree the non-bijective case would also be an interesting paper.  We can
+provide additional caveats about the examples and a better description of the
+algorithm in the final version of this paper.
 
@@
 
-We can go into more detail in the paper on this.  If two subcomponents can each
-map to two subcomponents on the opposite side, a permutation which does not swap
-the order of those subcomponents is best.
+We can go into more detail in the paper on this.  Intuitively,
+we group the subcomponents of a DNF regular expression into equivalence classes
+based on if they can map to the same subcomponents of the DNF regular expression
+on the other side.  Then, no swaps are performed within subcomponents of the
+same equivalence class, they are merely swapped with subcomponents of different
+equivalence classes, if one removed all other subcomponents, the orderings would
+be merely the identity.
+
+For example, in the provided scenario (caveat, I think the reviewer intended to
+say an example that says a <-> e) the example makes a map to e, so the
+equivalence classes are:
+
+{a}, {b,c} on the left
+{d,f}, {e} on the right
+
+Then, the algorithm will sort these equivalence classes
+
+{a}, {b,c}
+{e}, {d,f}
+
+and the equivalence classes will be mapped to each other through zipping
+
+a <-> e, b <-> d, c <-> f
 
@@
 understand than non-pretty printed versions is given by the sizes of the
-generated lenses.  Without minimization, the lenses sizes averaged at (SIZE)
+generated lenses.  Without minimization, the lenses sizes average at (SIZE)
 AST nodes, whereas minimization made the sizes average at (SIZE) ast nodes.
+In the general response, we show what a minimized vs handwritten lens look like.
+Below we the difference between a non-minimized title field and a minimized
+title field for a work item.
+
+In the title field, without minimization, the title field transformation would
+appear as:
+
+const("<Field Id=2></Field>","")
+  | (const("<Field Id=2>","Title: ")
+      . Id(text_char)
+      . const("","")
+      . iterate(const("","") . Id(text_char) . const("",""))
+      . const("","")
+      . const("</Field>",", "))
+
+With minimization, the title field transformation would appear as:
+
+const("<Field Id=2>","")
+  . (Id("")
+       | (Id(text_char) . iterate(Id(text_char)) . const("", ",")))
+  . const("</Field>",""))
+
+In particular, the useless concatenations of const("","") are removed and
+the field boilerplate is removed from being on both sides of the disjunct.
+Furthermore, when const(s,s) is present, it is converted into Id(s).
 
@@
 
-All of the examples taken from existing programs needed to be adapted in some
-way.  In particular, while there were many examples that didn't have projected
-information that would have necessitated projection aspects of lens
-capabilities, those examples all used whitespace.
+See general response.
 
@@
 
-We provide an example below on mapping the author field of BibTeX to EndNote:
+See general response.  We also provide data on the sizes below.
 
-      typedef NAME = UPPERCASE (LOWERCASE)*;;
-
-      typedef LASTCOMMASTART = NAME "," (WSP NAME)*;;
-      typedef BIBTEXAUTHORLIST = LASTCOMMASTART (" and " LASTCOMMASTART)*;;
-      typedef BIBTEXAUTHORINFO = "author={" BIBTEXAUTHORLIST "}";;
-
-      typedef AUTAG = "au - ";;
-      typedef STARTTOEND = (NAME WSP)* NAME;;
-      typedef TAGGEDAUTHORDEFNS = AUTAG STARTTOEND ("\n " AUTAG STARTTOEND)*;;
-
-      let bibtex_to_endnote_au_synthed =
-        synth
-          BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS
-
-      let bibtex_to_endnote_au
-        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
-        del "author={"
-          . ins "au - "
-          . lens_swap (NAME . del ",") (lens_swap WSP NAME)* 
-          . (del " and "
-               . ins "\n au - "
-               . lens_swap (NAME . del ",") (lens_swap WSP NAME)* )* 
-          . del "}"
-
-      let bibtex_to_endnote_au_human_written
-        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
-        let single_author_convert =
-          ins " au - "
-            . lens_swap
-                (NAME . del ",")
-                (lens_swap WSP NAME)*
-        in
-        del "author={"
-          . single_author_convert
-          . (del " and "
-               . ins "\n"
-               . single_author_convert)* 
-          . del "}"
-
-We feel the specification is easier to read and write than the explicit lens.
-Furthermore, even the relatively simple permutation of only 3 elements needed
-for single_author_convert is fairly complex.  Beyond this, we find it is best
-practices to include the type explicitly when writing the lens, both as
-documentation and as additional assurance to correctness.  As we already spend
-the effort writing these types manually for these benefits, there is no
-additional work needed to use our synthesis algorithm, there is instead only
-saved effort.
-
-We provide the size of specifications and lenses below:
-
---------------------------------------------
-|Test                | SpecSize | LensSize |
-|------------------------------------------|
+---------------------------------------------------
+|Test                | SpecSize | LensSize | TODO |
+|------------------------------------------|------|
 |cust/date-probs     | 85       | 79       |
@@
 We can make sure we can provide at least some description of the sizes of specs
-and lenses in the paper.
+and lenses in the final version of the paper.
 
@@
   could be useful for the kind of transformations appearing in the motivation.
-  Moreover, the functions can use each word in the input only once: one cannot
+
+We can convert between all uppercase and all lowercase.  Quotient lenses can be
+used to normalize the data into fully uppercase and fully lowercase data.
+
+- Moreover, the functions can use each word in the input only once: one cannot
   express the transformation abc->abcabc. Most bijective functions I can think of
   (string encoders, compression algorithms, encryption algorithms) require complex
-  transformations over the input that go beyond identities. On this note, there
-  was a paper this year at PLDI on inverting string encoders (Automatic program
-  inversion using symbolic transducers, by Hu et al.), a topic which seems very
-  related, but for which I couldn’t find a citation.
+  transformations over the input that go beyond identities.
+
+See general response for more detailed information about the bijective
+restriction.
+- On this note, there was a paper this year at PLDI on inverting string encoders
+  (Automatic program inversion using symbolic transducers, by Hu et al.), a
+  topic which seems very related, but for which I couldn’t find a citation.
 
@@
 Finite Transducers.  The finite lookahead does not allow for transformations
-which swaps the locations of data, a common pattern in Boomerang lenses.
-We will add a citation to this work and a comment detailing the restrictions
-extended symbolic finite transducers put in place.
+which swaps data, a common pattern and important pattern in Boomerang lenses. We
+will add a citation to this work and a comment detailing the restrictions of
+symbolic finite transducers.
 
@@
   be applied.
+
+See general response.
 
@@
 specified.  It is not intended to find subsets of the provided regular
-expressions that can be mapped.
+expressions that can be mapped, as those would be ill-typed programs.
 
@@
 On average, the domain and range specs are slightly larger than the lens itself.
-
-We do not think that this means the work is useless, but rather it is used to
-alleviate the pain of reasoning about nested swaps and concatenations, allow for
-more simple maintenance of the code.
-
-We provide an example below on mapping the author field of BibTeX to EndNote:
-
-      typedef NAME = UPPERCASE (LOWERCASE)*;;
-
-      typedef LASTCOMMASTART = NAME "," (WSP NAME)*;;
-      typedef BIBTEXAUTHORLIST = LASTCOMMASTART (" and " LASTCOMMASTART)*;;
-      typedef BIBTEXAUTHORINFO = "author={" BIBTEXAUTHORLIST "}";;
-
-      typedef AUTAG = "au - ";;
-      typedef STARTTOEND = (NAME WSP)* NAME;;
-      typedef TAGGEDAUTHORDEFNS = AUTAG STARTTOEND ("\n " AUTAG STARTTOEND)*;;
-
-      let bibtex_to_endnote_au_synthed =
-        synth
-          BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS
-
-      let bibtex_to_endnote_au
-        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
-        del "author={"
-          . ins "au - "
-          . lens_swap (NAME . del ",") (lens_swap WSP NAME)* 
-          . (del " and "
-               . ins "\n au - "
-               . lens_swap (NAME . del ",") (lens_swap WSP NAME)* )* 
-          . del "}"
-
-      let bibtex_to_endnote_au_human_written
-        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
-        let single_author_convert =
-          ins " au - "
-            . lens_swap
-                (NAME . del ",")
-                (lens_swap WSP NAME)*
-        in
-        del "author={"
-          . single_author_convert
-          . (del " and "
-               . ins "\n"
-               . single_author_convert)* 
-          . del "}"
-
-We feel the specification is easier to read and write than the explicit lens.
-Furthermore, even the relatively simple permutation of only 3 elements needed
-for single_author_convert is fairly complex.  Beyond this, we find it is best
-practices to include the type explicitly when writing the lens, both as
-documentation and as additional assurance to correctness.  As we already spend
-the effort writing these types manually for these benefits, there is no
-additional work needed to use our synthesis algorithm, there is instead only
-saved effort.
+See the above chart.  See the general response for a detailed commentary on the
+intent of our work.
 
@@
 The final example, extr-num was used because extracting a phone number was used
-as an extended example.  We could not find any examples in the paper which
-applied a to_upper conversion.
+as an extended example, so we used it.  We could not find any examples in the
+paper which applied a to_upper conversion.
 
@@
 
-While Flash Fill's DSL permits nested iterations, for performance reasons Flash
-Fill only synthesizes a single nesting of loops.  While this is sufficient for
-the types of applications present in an Excel file, the examples we found in
-Augeas were significantly more complex, requiring complex transformations with 
-many nested iterations.
-
-Flash Fill can express many transformation our DSL cannot.  For example, we
-currently cannot handle losing information, which prevents us from projecting
-information.
+See the general response.
 
@@
 const("<Field Id=2>","")
-  . (const("","")
+  . (Id("")
        | (Id(text_char) . iterate(Id(text_char)) . const("", ",")))
@@
 the field boilerplate is removed from being on both sides of the disjunct.
+Furthermore, when const(s,s) is present, it is converted into Id(s).
 
@@
 
-We provide an example below on mapping the author field of BibTeX to EndNote:
-
-      typedef NAME = UPPERCASE (LOWERCASE)*;;
-
-      typedef LASTCOMMASTART = NAME "," (WSP NAME)*;;
-      typedef BIBTEXAUTHORLIST = LASTCOMMASTART (" and " LASTCOMMASTART)*;;
-      typedef BIBTEXAUTHORINFO = "author={" BIBTEXAUTHORLIST "}";;
-
-      typedef AUTAG = "au - ";;
-      typedef STARTTOEND = (NAME WSP)* NAME;;
-      typedef TAGGEDAUTHORDEFNS = AUTAG STARTTOEND ("\n " AUTAG STARTTOEND)*;;
-
-      let bibtex_to_endnote_au_synthed =
-        synth
-          BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS
-
-      let bibtex_to_endnote_au
-        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
-        del "author={"
-          . ins "au - "
-          . lens_swap (NAME . del ",") (lens_swap WSP NAME)* 
-          . (del " and "
-               . ins "\n au - "
-               . lens_swap (NAME . del ",") (lens_swap WSP NAME)* )* 
-          . del "}"
-
-      let bibtex_to_endnote_au_human_written
-        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
-        let single_author_convert =
-          ins " au - "
-            . lens_swap
-                (NAME . del ",")
-                (lens_swap WSP NAME)*
-        in
-        del "author={"
-          . single_author_convert
-          . (del " and "
-               . ins "\n"
-               . single_author_convert)* 
-          . del "}"
-
-We feel the specification is easier to read and write than the explicit lens.
-Furthermore, even the relatively simple permutation of only 3 elements needed
-for single_author_convert is fairly complex.  Beyond this, we find it is best
-practices to include the type explicitly when writing the lens, both as
-documentation and as additional assurance to correctness.  As we already spend
-the effort writing these types manually for these benefits, there is no
-additional work needed to use our synthesis algorithm, there is instead only
-saved effort.
+We provide an extended example in the general response.  We will make the
+problem statement more clear in the final version.
 
@@
 the star-semiring equivalence rules.  We find this useful in that all lenses can
-easily be converted into a bijective lens, and we could not find a lens that
-requires a courser notion of equivalence than star-semiring equivalence.
+easily be converted into a bijective lens, and we could not find a lens in
+Boomerang or Augeas that requires a courser notion of equivalence than
+star-semiring equivalence.
 
@@
 equivalences.  We are trying to carve out a specific subset of Boomerang that
-we want to synthesize, and are not allowing that subset to be included.
-The problems with including full language equivalence is detailed on pages 8
-through 10.
+we want to synthesize, and are not allowing the subset requiring a courser
+notion of equivalence to be included. The problems with including full language
+equivalence is detailed on pages 8 through 10.
 

### 2017-09-12 papers/BijectiveLensSynth/popl_18_rebuttal.txt
add in a line

@@
 Category 1 was by far the most prevalent, and encountered in many of the Augeas
-examples.  To handle this, we added additional whitespace to the other format.
-This can be solved by using quotients on each side.  We think the strategy of
-using specialized tokens to represent both a language and a canonizer can solve
-this problem.
+examples.  To handle this, we added additional whitespace to the other format,
+where typically whitespace is unnecessary.
+This can be solved by using quotients on the side with whitespace.  We think the
+strategy of using specialized tokens to represent both a language and a
+canonizer can solve this problem.
 
@@
 single_author_convert is fairly complex.  These becomes even more apparent when
-working with large lenses comprised of many sub-lenses being permuted.  Lenses
-provide great power with their invertibility guarantees, but they come at the
-cost of thinking about many fiddly details when writing the terms.  Worrying
-about these fiddly details while ALSO thinking about unambiguity restrictions
-is a very difficult task!
+working with large lenses comprised of many sub-lenses being permuted (though
+we prefer to use a more minimal example).  Lenses provide great power with their
+invertibility guarantees, but they come at the cost of thinking about many
+fiddly details when writing the terms.  Worrying about these fiddly details
+while ALSO thinking about unambiguity restrictions is a very difficult task!
 

### 2017-09-12 papers/BijectiveLensSynth/popl_18_rebuttal.txt
add in changes

@@
 Flash Fill is able to synthesize some transformations that we cannot.
-In particular, by restricting ourselves to bijections, we are unable to express
-a large class of functions.  In alternative string synthesis systems, this
-bijective restriction is not in place, and additional strings can be
-synthesized.  However, we do not see this as a failure, but as a design choice.
-Our system is not targeting single direction transformations: we feel
-programmers already have the tools they need to write these by hand.  Instead,
-we are aiming to alleviate the difficulties in programming bidirectional
-transformations, which have languages that are very difficult to write in.
+
+By restricting ourselves to bijections, we are unable to express functions that
+don't preserve data, such as: mapping from a format with whitespace to one
+without whitespace, turning a mix of uppercase and lowercase characters into
+purely uppercase characters, and extracting a file name from a file path.
 
 However, Optician can still solve certain classes of problems other string
-synthesis systems cannot.  In particular, existing techniques that only use
+synthesis systems cannot - in particular, existing techniques that only use
 input-output examples are unable to synthesize programs with nested loops.
-In particular, while Flash Fill's DSL does permit nested loops, for efficiency
+While Flash Fill's DSL does permit nested loops, for efficiency
 reasons, the synthesis algorithm only synthesizes programs with no nesting (this
-is stated at the end of section 4.4).  While a single nesting of loops is
-sufficient for the types of problems encountered in transforming Excel
+is stated at the end of section 4.4 of "Automating String Processing in
+Spreadsheets Using Input-Output Examples", POPL '11).  While a single nesting of
+loops is sufficient for the types of problems encountered in transforming Excel
 spreadsheet data, our tool is aimed at building large transformations between
-highly complex formats.  The types of formats Augeas is built to handle have
-many deep layers of nesting, with very complex transformations at each layer.
-Nested iteration is necessary for these types of transformations, and existing
-techniques are insufficient for synthesizing them.
+more complex formats.  The types of formats Augeas is built to handle have
+deep layers of nested transformations, with very complex transformations at each
+layer, making existing techniques insufficient.
 
-- Why did we make the bijective restriction, and what transformations did we
-  make to our benchmark suite to make them bijective.
-
-We simplified our task to handling bijective lenses as this is an initial
-approach into this research.  There are a large number of
-formats which are in bijective correspondence: for example https://www.data.gov/
-provides data in a variety of formats in bijective correspondence with each
-other.  However, we began with synthesizing bijective
-lenses primarily because we believe this bijective core can be extended in a
-principled way to handle more complex lenses.
-
-Many of our transformations were nearly bijective, but had unnecessary
-information, like whitespace or casing, that prevented them from being bijective.
-Quotient lenses provide the capabilities to ignore unnecessary information like
-whitespace.  While the combinators for quotient lenses, canonizers, can be
-intermixed with basic lenses in the original paper, we believe that a
-"canonizers-at-the-edge" approach is just as expressive.  In particular,
-we can express a transformation that is bijective up to a quotient by having a
-bijective lens core, with a canonizer at each side that normalizes the input.
-Indeed, we already have a prototype that does this, allowing specialized
-tokens that encodes both the full description of viable inpout as well as the
-information on how to canonize the data.
-
-The remaining transformations that needed to be altered were also nearly
-bijective, but projected away useful information, and this projection prevented
-them from being bijective.  Unlike with quotient lenses, we do not want to
-remove this information - we want the information to be preserved through
-round-trips.  Furthermore, there are formats where each format contains
-information not present in the other format.  Symmetric lenses provide the
-capabilities for writing such transformations.  Similarly to quotient lenses,
-projection combinators can be intermixed with bijection capabilities,
-but we can view symmetric lenses as having a bijective core with a projection
-combinator on each side that removes data in one direction, and restores it in
-the other.
-
-While we simplified our task to bijective lenses, we did this because we felt
-a synthesis algorithm for bijective lenses could form the core of a synthesis
-algorithm for other types of lenses.  We agree this motivation looking towards
-future work is currently lacking in the paper, and we can add this in for the
-final version.
+- What transformations did we make to our benchmark suite to make them
+  bijective, and does this restriction give rise to useful systems.
 
@@
 where typically whitespace is unnecessary.
-This can be solved by using quotients on the side with whitespace.  We think the
-strategy of using specialized tokens to represent both a language and a
-canonizer can solve this problem.
+
+Instead of changing the formats, we could have extended the system to involve
+canonizers.  Canonizers can be placed on the ends of lenses to
+normalize the input data.  Canonizers can normalize an arbitrarily long sequence
+of whitespace to a single space.  Canonizers also can be used to handle casing
+issues: normalizing a sequence of uppercase and lowercase characters into purely
+uppercase characters.  We have already developed a framework to annotate a
+regular expression with additional canonizing information, allowing for a single
+term to represent an arbitrary sequence of whitespace and a means to normalize
+it to a single space.  Using this framework, non-bijective lenses can be
+generated using our framework, as long as the canonized
+forms of their source and target types are in bijective correspondence.
 
@@
 the research on it.
+
+There are a large number of formats that are in bijective correspondence: for
+example https://www.data.gov/ provides data in a variety of formats in bijective
+correspondence with each other.  Much of the encountered formats present in
+existing lens libraries uses either canonizers or projections.  The theories
+developed to handle synchronization tasks for real world data (quotient lenses
+for Category 1, and symmetric lenses for Category 2), have a bijective core,
+surrounded by either canonizers or projections.  We feel that our system can be
+used as a drop-in component for synthesis algorithms on these more complex lens
+structures.
 
@@
 We omitted the formal description of these orderings for space.  A full
-description of the algorithm is present in algorithm 4 in the appendix, on page
+description of the algorithm is present in algorithm 4 in the appendix (see the
+supplemental pdf submitted with the paper), on page
 119, with the orderings defined in Definition 33 on page 118 (the definitions of
@@
 into the DNF regular expressions, and make an ordering on the DNF regular
-expression with embedded parse trees.  While we think the formal description of
-these orderings is too long to be included in the paper, we can provide a more
-involved informal description in the text, and provide a reference to an
-extended version with the formal descriptions of the orderings.
+expression with embedded parse trees.  We would be happy to rewrite the paper to
+include whatever information the reviewer believes is most important.
 
@@
 
-We merely included that syntax as a means to make the derivation more
+We merely wrote the rule with operators as a means to make the derivation more
 readable.  One can evaluate those operators, and it will be an arbitrary
@@
 
-We can go into more detail in the paper on this.  Intuitively,
-we group the subcomponents of a DNF regular expression into equivalence classes
-based on if they can map to the same subcomponents of the DNF regular expression
-on the other side.  Then, no swaps are performed within subcomponents of the
-same equivalence class, they are merely swapped with subcomponents of different
-equivalence classes, if one removed all other subcomponents, the orderings would
-be merely the identity.
+First, we think the reviewer intended to write "we have only one example that
+says a <-> d".  We see two cases for this scenario.
+Case 1: The structure of the other two determines where they must map
+Case 2: The structure does not determine where they must mapped
 
-For example, in the provided scenario (caveat, I think the reviewer intended to
-say an example that says a <-> e) the example makes a map to e, so the
-equivalence classes are:
+For Case 1, consider the scenario where
+b = Name
+c = Name*
+e = Name*
+f = Name
+The structures of b, c, e, and f determine where they will be mapped: b <-> f
+and c <-> e.  In particular, RigidSynth will order them as:
+[a;b;c]
+[d;f;e]
+that then creates the maps previously provided.
 
-{a}, {b,c} on the left
-{d,f}, {e} on the right
-
-Then, the algorithm will sort these equivalence classes
-
-{a}, {b,c}
-{e}, {d,f}
-
-and the equivalence classes will be mapped to each other through zipping
-
-a <-> e, b <-> d, c <-> f
+For Case 2, consider the scenario where
+b = Name
+c = Name
+e = Name
+f = Name
+In this case, the structures do not determine where they will be mapped.
+The maps (a <-> d , b <-> e , c <-> f)
+     and (a <-> d , b <-> f , c <-> e) are both valid.  In this situation, the
+one we choose is based on a heuristic that minimizes swaps.  In particular,
+Rigidsynth will order them as:
+[a;b;c]
+[d;e;f]
+that creates the maps (a <-> d , b <-> e , c <-> f).
 
@@
 The algorithm described includes all but compositional synthesis.  We don't
-include in the formal algorithm the means to which we use previously defined
-lenses to synthesize new ones.  The way we do it is we develop equivalence
-classes between user defined data types that have lenses, choose a
+include in the formal algorithm a description of how we use previously defined
+lenses to synthesize new ones.  For compositional synthesis, we develop
+equivalence classes between user defined data types that have lenses, choose a
 representative, and transform the user defined data types into that
 representative, and convert the strings into elements of that representative
-data type.
+data type.  Then we proceed with synthesis as described by the formal algorithm.
 
@@
 
----------------------------------------------------
-|Test                | SpecSize | LensSize | TODO |
-|------------------------------------------|------|
-|cust/date-probs     | 85       | 79       |
-|cust/cap-prob       | 112      | 108      |
-|cust/2-cap-prob     | 122      | 66       |
-|ff/extr-fname-err   | 143      | 151      |
-|ff/extr-fname       | 149      | 168      |
-|ff/extr-quant       | 150      | 145      |
-|ff/extr-num         | 171      | 166      |
-|aug/activemq        | 194      | 184      |
-|cust/bib-prob       | 202      | 181      |
-|cust/workitem-probs | 221      | 252      |
-|aug/xml-l1          | 235      | 214      |
-|cust/addr-probs     | 244      | 239      |
-|aug/aptcrngsec      | 265      | 204      |
-|aug/alias           | 271      | 210      |
-|aug/backuppchosts   | 272      | 249      |
-|aug/approx          | 277      | 264      |
-|aug/apt-upd-mgr     | 278      | 210      |
-|aug/aptsources      | 278      | 238      |
-|aug/cachefilesd     | 294      | 233      |
-|aug/aptprefs        | 297      | 280      |
-|aug/afs-cellalias   | 298      | 246      |
-|aug/avahi           | 301      | 234      |
-|aug/chrony          | 308      | 289      |
-|aug/xml             | 311      | 383      |
-|aug/carbon          | 325      | 377      |
-|aug/access          | 335      | 253      |
-|aug/anacron         | 342      | 315      |
-|aug/hosts           | 343      | 328      |
-|aug/bootconf        | 350      | 271      |
-|aug/auth-keys       | 354      | 424      |
-|aug/cgrules         | 378      | 295      |
-|aug/cgconfig        | 415      | 347      |
-|aug/automaster      | 475      | 377      |
-|aug/cron            | 475      | 414      |
-|aug/channels        | 512      | 367      |
-|aug/automounter     | 535      | 432      |
-|aug/aptconf         | 540      | 450      |
-|aug/aptconf-l1      | 564      | 503      |
-|aug/bbhosts         | 670      | 651      |
---------------------------------------------
+|-------------------------------------------------------------|
+| Test                | SpecSize | LensSize | LensAndSpecSize |
+|-------------------------------------------------------------|
+| cust/date-probs     | 85       | 79       | 97              |
+| cust/cap-prob       | 112      | 108      | 169             |
+| cust/2-cap-prob     | 122      | 66       | 188             |
+| ff/extr-fname-err   | 143      | 151      | 172             |
+| ff/extr-fname       | 149      | 168      | 195             |
+| ff/extr-quant       | 150      | 145      | 161             |
+| ff/extr-num         | 171      | 166      | 186             |
+| aug/activemq        | 194      | 184      | 204             |
+| cust/bib-prob       | 202      | 181      | 265             |
+| cust/workitem-probs | 221      | 252      | 330             |
+| aug/xml-l1          | 235      | 214      | 301             |
+| cust/addr-probs     | 244      | 239      | 308             |
+| aug/aptcrngsec      | 265      | 204      | 336             |
+| aug/alias           | 271      | 210      | 348             |
+| aug/backuppchosts   | 272      | 249      | 318             |
+| aug/approx          | 277      | 264      | 345             |
+| aug/apt-upd-mgr     | 278      | 210      | 355             |
+| aug/aptsources      | 278      | 238      | 351             |
+| aug/cachefilesd     | 294      | 233      | 368             |
+| aug/aptprefs        | 297      | 280      | 409             |
+| aug/afs-cellalias   | 298      | 246      | 370             |
+| aug/avahi           | 301      | 234      | 370             |
+| aug/chrony          | 308      | 289      | 331             |
+| aug/xml             | 311      | 383      | 546             |
+| aug/carbon          | 325      | 377      | 507             |
+| aug/access          | 335      | 253      | 436             |
+| aug/anacron         | 342      | 315      | 447             |
+| aug/hosts           | 343      | 328      | 449             |
+| aug/bootconf        | 350      | 271      | 457             |
+| aug/auth-keys       | 354      | 424      | 553             |
+| aug/cgrules         | 378      | 295      | 487             |
+| aug/cgconfig        | 415      | 347      | 612             |
+| aug/automaster      | 475      | 377      | 621             |
+| aug/cron            | 475      | 414      | 667             |
+| aug/channels        | 512      | 367      | 682             |
+| aug/automounter     | 535      | 432      | 710             |
+| aug/aptconf         | 540      | 450      | 700             |
+| aug/aptconf-l1      | 564      | 503      | 777             |
+| aug/bbhosts         | 670      | 651      | 1041            |
+|-------------------------------------------------------------|
+| average             | 310      | 277      | 414             |
+|-------------------------------------------------------------|
 

### 2017-09-12 papers/BijectiveLensSynth/popl_18_rebuttal.txt
whiches to thats

@@
 Finite Transducers.  The finite lookahead does not allow for transformations
-which swaps data, a common pattern and important pattern in Boomerang lenses. We
+that swap data, a common pattern and important pattern in Boomerang lenses. We
 will add a citation to this work and a comment detailing the restrictions of
@@
 as an extended example, so we used it.  We could not find any examples in the
-paper which applied a to_upper conversion.
+paper that applied a to_upper conversion.
 

### 2017-09-12 papers/BijectiveLensSynth/popl_18_rebuttal.txt
Begin polishing

@@
-We thank the reviewers for their feedback.  We will provide a common rebuttal
+Thank you for the thoughtful reviews.  We'll start with brief responses to
+the most important concerns, followed by an (optional!) appendix addressing
+every nontrivial point from the reviews, in case you want more.
+
+ will provide a common rebuttal
 to address concerns common to many reviewers, and will then have an appendix

### 2017-09-12 papers/BijectiveLensSynth/popl_18_rebuttal.txt
Bunch of polishing -- the opening part is still WAY too long...

@@
 
- will provide a common rebuttal
-to address concerns common to many reviewers, and will then have an appendix
-addressing specific concerns.
+                                -------------
 
-- What is the difference between our work and existing work, like Flash Fill,
-  in terms of expressivity?
+- What is the difference in expressiveness between our work and existing
+  algorithms like FlashFill?
 
-We are able to synthesize some transformations that Flash Fill cannot, and
-Flash Fill is able to synthesize some transformations that we cannot.
+Formally, they are incomparable: each can synthesize some transformations
+that the other cannot.
 
-By restricting ourselves to bijections, we are unable to express functions that
-don't preserve data, such as: mapping from a format with whitespace to one
-without whitespace, turning a mix of uppercase and lowercase characters into
-purely uppercase characters, and extracting a file name from a file path.
+Since we restrict to bijections, we are unable to express functions that
+don't preserve data, such as
+  - mapping from a format with whitespace to one without whitespace,
+  - turning a mix of uppercase and lowercase characters into purely
+    uppercase characters, and
+  - extracting a file name from a file path.
+[BCP: Can FlashFill express all these?]
 
-However, Optician can still solve certain classes of problems other string
-synthesis systems cannot - in particular, existing techniques that only use
-input-output examples are unable to synthesize programs with nested loops.
-While Flash Fill's DSL does permit nested loops, for efficiency
-reasons, the synthesis algorithm only synthesizes programs with no nesting (this
-is stated at the end of section 4.4 of "Automating String Processing in
-Spreadsheets Using Input-Output Examples", POPL '11).  While a single nesting of
-loops is sufficient for the types of problems encountered in transforming Excel
-spreadsheet data, our tool is aimed at building large transformations between
-more complex formats.  The types of formats Augeas is built to handle have
-deep layers of nested transformations, with very complex transformations at each
-layer, making existing techniques insufficient.
+Conversely, our Optician algorithm can solve some classes of problems other
+string synthesis systems cannot.  In particular, existing techniques that
+only use input-output examples are unable to synthesize programs with nested
+loops.
 
-- What transformations did we make to our benchmark suite to make them
-  bijective, and does this restriction give rise to useful systems.
+While FlashFill's DSL does permit nested loops, for efficiency reasons the
+synthesis algorithm only synthesizes programs with no nesting (this is
+stated at the end of section 4.4 of "Automating String Processing in
+Spreadsheets Using Input-Output Examples", POPL '11).  While a single
+nesting of loops seems to be sufficient for the types of problems
+encountered in transforming Excel data, our tool is aimed at building large
+transformations between more complex formats.  In particular, the types of
+formats that Augeas handles tend to have deep layers of nested
+transformations, with complex transformations at each layer.  FlashFill
+cannot handle such examples.
 
-The changes we put on our benchmark suite fall in 2 major categories that mirror
-the intended extensions to our system.
+                                -------------
+
+- What modifications did we need to make to programs in our Augeas
+  "benchmark suite" to make them bijective?
+
+The changes we made to the benchmark suite fall in two major categories that
+mirror the intended extensions to our system.
+
 1) Whitespace was present in one format but not the other.
@@
 
+                                -------------
+
 - After having written the regular expressions, is it very difficult for users
-  to then write the lens.  Are we saving anybody any significant effort.
+  to then write the lens?  Are we saving anybody any significant effort?
 
-Engineering best practices dictate that we annotate the lenses with their types.
-While Boomerang does not need these annotations - being able to infer the types
-from the terms - the types of the term serve as documentation for
-future programmers to understand what formats the lens maps between.
-Furthermore, it provides an additional resilience to future modifications of the
-lens, as it validates that the lens maps exactly between strings of the provided
-formats - invalid inputs are not accepted nor are the outputs ever invalid.
-In this way, we are not requiring additional work by the programmer to use our
-system; we are allowing the programmer to save the difficulties of
-writing the term by using work they should already be performing.  To highlight
-these difficulties, we show an example program for mapping the author field of
-a BibTeX entry.
+Engineering best practices dictate that we annotate the lenses with their
+types.  While Boomerang does not need these annotations - being able to
+infer the types from the terms - the types of the term serve as
+documentation for future programmers to understand what formats the lens
+maps between.  Furthermore, it provides an additional resilience to future
+modifications of the lens, as it validates that the lens maps exactly
+between strings of the provided formats - invalid inputs are not accepted
+nor are the outputs ever invalid.  In this way, we are not requiring
+additional work by the programmer to use our system; we are allowing the
+programmer to save the difficulties of writing the term by using work they
+should already be performing.  To highlight these difficulties, we show an
+example program for mapping the author field of a BibTeX entry.
 
@@
 
-Detailed Responses to individual questions by the reviewers follow below.
+===========================================================================
+===========================================================================
+Detailed responses to individual questions in the reviews follow...
+
+[BCP: In the "appendix" I think it's important to quote the reviews
+verbatim, so that reviewers can see what they said rather than our
+paraphrase of it.]
 
 Review #18A
-===========================================================================
 
@@
 
+===========================================================================
 Review #18B
-===========================================================================
+
 Paper summary
@@
 
+===========================================================================
 Review #18C
-===========================================================================
 

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
add in responses to BCP

@@
   - extracting a file name from a file path.
-[BCP: Can FlashFill express all these?]
+[BCP: Can FlashFill express all these?] [AFM: Yes]
 
@@
 verbatim, so that reviewers can see what they said rather than our
-paraphrase of it.]
+paraphrase of it.] [AFM: Quoted but will add in quotes to make it explicit]
 

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
Update popl_18_rebuttal.txt

@@
 Thank you for the thoughtful reviews.  We'll start with brief responses to
-the most important concerns, followed by an (optional!) appendix addressing
-every nontrivial point from the reviews, in case you want more.
+the most significant concerns, followed by an (optional!) appendix addressing
+the more detailed questions.
 

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
Update popl_18_rebuttal.txt

@@
 nesting of loops seems to be sufficient for the types of problems
-encountered in transforming Excel data, our tool is aimed at building large
+encountered in transforming Excel data, our tool is aimed at building intricate
 transformations between more complex formats.  In particular, the types of
@@
 issues: normalizing a sequence of uppercase and lowercase characters into purely
-uppercase characters.  We have already developed a framework to annotate a
-regular expression with additional canonizing information, allowing for a single
+uppercase characters.  We have recently developed a framework to annotate a
+regular expression with canonizing information, allowing for a single
 term to represent an arbitrary sequence of whitespace and a means to normalize
-it to a single space.  Using this framework, non-bijective lenses can be
-generated using our framework, as long as the canonized
-forms of their source and target types are in bijective correspondence.
+it to a single space.  Using this framework, we can generate non-bijective lenses 
+using our framework, as long as the canonized forms of their source and target 
+types are in bijective correspondence.
 

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
Update popl_18_rebuttal.txt

@@
 when going left to right, and restores that information going back would provide
-the full lens.  This wrapper is not very difficult to write, and we think that
-discovering this projection automatically is possible, but we still need to do
-the research on it.
+the full lens.  This wrapper is not difficult to write, and we think that
+discovering this projection automatically is possible, but we have not yet done it.
 
-There are a large number of formats that are in bijective correspondence: for
+There are a large number of formats that are in bijective correspondence. For
 example https://www.data.gov/ provides data in a variety of formats in bijective
-correspondence with each other.  Much of the encountered formats present in
-existing lens libraries uses either canonizers or projections.  The theories
-developed to handle synchronization tasks for real world data (quotient lenses
-for Category 1, and symmetric lenses for Category 2), have a bijective core,
-surrounded by either canonizers or projections.  We feel that our system can be
-used as a drop-in component for synthesis algorithms on these more complex lens
-structures.
+correspondence with each other.  Many of the formats present in existing lens libraries
+use either canonizers or projections.  The theories developed to handle synchronization 
+tasks for real world data (quotient lenses for Category 1, and symmetric lenses 
+for Category 2), have a bijective core, surrounded by either canonizers or projections.  
+We believe that our system can be used as a drop-in component for synthesis algorithms 
+on these more complex lens structures.
 
@@
 should already be performing.  To highlight these difficulties, we show an
-example program for mapping the author field of a BibTeX entry.
+example program for mapping the author field of a BibTeX entry:
 
@@
 
-Even the relatively simple permutation of only 3 elements needed for
-single_author_convert is fairly complex.  These becomes even more apparent when
-working with large lenses comprised of many sub-lenses being permuted (though
-we prefer to use a more minimal example).  Lenses provide great power with their
+Even the relatively simple lens that permuts the 3 elements needed for
+single_author_convert is fairly complex.  The complexity becomes even more apparent when
+working with large lenses comprised of many sub-lenses (though for brevity
+we give here the more minimal example above).  Lenses provide great power with their
 invertibility guarantees, but they come at the cost of thinking about many
@@
 
-This tool is aimed at different audiences than existing string transformation
+Optician is aimed at different audiences than existing string transformation
 tools.  While existing tools allow the average end-user to generalize their
@@
 powerful language while avoiding much of the complexity of programming in that
-language.  This paradigm allows for more complex programs to be able to be
+language.  This paradigm allows for more complex programs to be
 synthesized, the types of programs that necessitate programming in such a DSL,

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
add in some responses to individual comments

--- 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
@@
 Thank you for the thoughtful reviews.  We'll start with brief responses to
-the most significant concerns, followed by an (optional!) appendix addressing
-the more detailed questions.
+the most important concerns, followed by an (optional!) appendix addressing
+every nontrivial point from the reviews, in case you want more.
 
@@
 In the general response, we show what a minimized vs handwritten lens look like.
-Below we the difference between a non-minimized title field and a minimized
+Below we show the difference between a non-minimized title field and a minimized
 title field for a work item.
@@
 
-The algorithm described includes all but compositional synthesis.  We don't
-include in the formal algorithm a description of how we use previously defined
-lenses to synthesize new ones.  For compositional synthesis, we develop
+The algorithm described includes all but compositional synthesis, which uses
+predefined lenses.  To use a predefined lens, l : A <-> B, the implementation
+develops equivalence classes between user defined data types with lenses between
+them (in the above example, A and B would be placed in the same equivalence
+class).  Next, a representative element of each equivalence class is chosen.
+All user defined data types are then transformed into their representative, and
+the string examples are converted into examples in the language of the
+representative.
+
 equivalence classes between user defined data types that have lenses, choose a
@@
   solutions varies (e.g. for the full algorithm) as the number of
-  examples provided varies.  Is the number of examples small? An active
-  approach where the algorithm asks the user for examples might also be
-  worth exploring (future work).
+  examples provided varies.  Is the number of examples small?
 
@@
 number of possible solutions (TODO data).  The number of examples is indeed
-small, as is shown in Figure 11.  An active approach we agree is an
-interesting one, and one we too have considered.
+small, as is shown in Figure 11.
 
@@
 
-You are correct, regular expressions do have canonical forms in that manner.
-We will change the phrasing to more accurately represent our meaning.
+This is an interesting idea.  It is unclear how an automataton for a type
+can be used to direct synthesis.  Nevertheless, this may lead to interesting
+research.
 
@@
 
-See general response.  We also provide data on the sizes below.
+See general response.  We also provide data on the sizes below.  However, we
+do not think that purely lines of code is the best metric for our data, as our
+system provides benefits beyond saving keystrokes for the programmer, detailed
+below:
+
+* Regular expressions are common - almost every programmer has a firm grip on
+regular expressions.  In contrast, almost _no_ programmers know anything about
+programming with lenses.
+
+* A regular expression describes one piece of data.  When writing a lens, one
+must think about both pieces of data at the same time, and simultaneously
+translate back and forth between them.
+
+* One must sometimes compute in their own head how to unroll or transform the
+regular expression into the right form.
 
 |-------------------------------------------------------------|
-| Test                | SpecSize | LensSize | LensAndSpecSize |
+| Test                | SpecSize | LensAndTypeSize | LensSize |
 |-------------------------------------------------------------|
-| cust/date-probs     | 85       | 79       | 97              |
-| cust/cap-prob       | 112      | 108      | 169             |
-| cust/2-cap-prob     | 122      | 66       | 188             |
-| ff/extr-fname-err   | 143      | 151      | 172             |
-| ff/extr-fname       | 149      | 168      | 195             |
-| ff/extr-quant       | 150      | 145      | 161             |
-| ff/extr-num         | 171      | 166      | 186             |
-| aug/activemq        | 194      | 184      | 204             |
-| cust/bib-prob       | 202      | 181      | 265             |
-| cust/workitem-probs | 221      | 252      | 330             |
-| aug/xml-l1          | 235      | 214      | 301             |
-| cust/addr-probs     | 244      | 239      | 308             |
-| aug/aptcrngsec      | 265      | 204      | 336             |
-| aug/alias           | 271      | 210      | 348             |
-| aug/backuppchosts   | 272      | 249      | 318             |
-| aug/approx          | 277      | 264      | 345             |
-| aug/apt-upd-mgr     | 278      | 210      | 355             |
-| aug/aptsources      | 278      | 238      | 351             |
-| aug/cachefilesd     | 294      | 233      | 368             |
-| aug/aptprefs        | 297      | 280      | 409             |
-| aug/afs-cellalias   | 298      | 246      | 370             |
-| aug/avahi           | 301      | 234      | 370             |
-| aug/chrony          | 308      | 289      | 331             |
-| aug/xml             | 311      | 383      | 546             |
-| aug/carbon          | 325      | 377      | 507             |
-| aug/access          | 335      | 253      | 436             |
-| aug/anacron         | 342      | 315      | 447             |
-| aug/hosts           | 343      | 328      | 449             |
-| aug/bootconf        | 350      | 271      | 457             |
-| aug/auth-keys       | 354      | 424      | 553             |
-| aug/cgrules         | 378      | 295      | 487             |
-| aug/cgconfig        | 415      | 347      | 612             |
-| aug/automaster      | 475      | 377      | 621             |
-| aug/cron            | 475      | 414      | 667             |
-| aug/channels        | 512      | 367      | 682             |
-| aug/automounter     | 535      | 432      | 710             |
-| aug/aptconf         | 540      | 450      | 700             |
-| aug/aptconf-l1      | 564      | 503      | 777             |
-| aug/bbhosts         | 670      | 651      | 1041            |
+| cust/date-probs     | 85       | 97              | 79       |
+| cust/cap-prob       | 112      | 169             | 108      |
+| cust/2-cap-prob     | 122      | 188             | 66       |
+| ff/extr-fname-err   | 143      | 172             | 151      |
+| ff/extr-fname       | 149      | 195             | 168      |
+| ff/extr-quant       | 150      | 161             | 145      |
+| ff/extr-num         | 171      | 186             | 166      |
+| aug/activemq        | 194      | 204             | 184      |
+| cust/bib-prob       | 202      | 265             | 181      |
+| cust/workitem-probs | 221      | 330             | 252      |
+| aug/xml-l1          | 235      | 301             | 214      |
+| cust/addr-probs     | 244      | 308             | 239      |
+| aug/aptcrngsec      | 265      | 336             | 204      |
+| aug/alias           | 271      | 348             | 210      |
+| aug/backuppchosts   | 272      | 318             | 249      |
+| aug/approx          | 277      | 345             | 264      |
+| aug/apt-upd-mgr     | 278      | 355             | 210      |
+| aug/aptsources      | 278      | 351             | 238      |
+| aug/cachefilesd     | 294      | 368             | 233      |
+| aug/aptprefs        | 297      | 409             | 280      |
+| aug/afs-cellalias   | 298      | 370             | 246      |
+| aug/avahi           | 301      | 370             | 234      |
+| aug/chrony          | 308      | 331             | 289      |
+| aug/xml             | 311      | 546             | 383      |
+| aug/carbon          | 325      | 507             | 377      |
+| aug/access          | 335      | 436             | 253      |
+| aug/anacron         | 342      | 447             | 315      |
+| aug/hosts           | 343      | 449             | 328      |
+| aug/bootconf        | 350      | 457             | 271      |
+| aug/auth-keys       | 354      | 553             | 424      |
+| aug/cgrules         | 378      | 487             | 295      |
+| aug/cgconfig        | 415      | 612             | 347      |
+| aug/automaster      | 475      | 621             | 377      |
+| aug/cron            | 475      | 667             | 414      |
+| aug/channels        | 512      | 682             | 367      |
+| aug/automounter     | 535      | 710             | 432      |
+| aug/aptconf         | 540      | 700             | 450      |
+| aug/aptconf-l1      | 564      | 777             | 503      |
+| aug/bbhosts         | 670      | 1041            | 651      |
 |-------------------------------------------------------------|
-| average             | 310      | 277      | 414             |
+| average             | 310      | 414             | 277      |
 |-------------------------------------------------------------|
@@
 We can convert between all uppercase and all lowercase.  Quotient lenses can be
-used to normalize the data into fully uppercase and fully lowercase data.
+used to normalize a format (or portions of a format) into fully uppercase or
+fully lowercase data.
 
@@
 
-We do require regular expressions describing the input of FlashFill specs, we
-will make sure that is more explicitly stated in the paper.
+We do require regular expressions describing the input of FlashFill specs (we
+require them for all), we will make sure that is more explicitly stated in the
+paper.
 
@@
 examples, and some others to handle the readjustment.
-
-Optician is meant to work on finding mappings between the regular expressions
-specified.  It is not intended to find subsets of the provided regular
-expressions that can be mapped, as those would be ill-typed programs.
 
@@
 On average, the domain and range specs are slightly larger than the lens itself.
-See the above chart.  See the general response for a detailed commentary on the
-intent of our work.
+See the above chart.  See the general response for a detailed commentary
+explaining why we believe writing lenses is harder than writing regular
+expressions, despite the fact that lenses alone are sometimes shorter than
+regular expression specifications.
 

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
Polishing

@@
-Thank you for the thoughtful reviews.  We'll start with brief responses to
-the most significant concerns, followed by an (optional!) appendix addressing
-the more detailed questions.
+Thank you for your comments!
+
+We begin with brief responses to the most significant concerns, followed by
+a (long and very optional) appendix addressing all questions in detail.
 
@@
 
-- What is the difference in expressiveness between our work and existing
-  algorithms like FlashFill?
+      - What is the difference in expressiveness between our work and
+        existing algorithms like FlashFill?
 
@@
   - extracting a file name from a file path.
-[BCP: Can FlashFill express all these?] [AFM: Yes]
+FlashFill can express all these.
 
-Conversely, our Optician algorithm can solve some classes of problems other
-string synthesis systems cannot.  In particular, existing techniques that
-only use input-output examples are unable to synthesize programs with nested
-loops.
+Conversely, our "Optician" algorithm can solve some classes of problems that
+existing synthesis systems cannot.  In particular, techniques that only use
+input-output examples are unable to synthesize programs with nested loops.
 
-While FlashFill's DSL does permit nested loops, for efficiency reasons the
-synthesis algorithm only synthesizes programs with no nesting (this is
-stated at the end of section 4.4 of "Automating String Processing in
-Spreadsheets Using Input-Output Examples", POPL '11).  While a single
-nesting of loops seems to be sufficient for the types of problems
-encountered in transforming Excel data, our tool is aimed at building intricate
-transformations between more complex formats.  In particular, the types of
-formats that Augeas handles tend to have deep layers of nested
-transformations, with complex transformations at each layer.  FlashFill
-cannot handle such examples.
+One more technical note about point: FlashFill's DSL does permit nested
+loops, but for efficiency reasons the synthesis algorithm only synthesizes
+programs with no nesting (this is discussed at the end of section 4.4 of
+"Automating String Processing in Spreadsheets Using Input-Output
+Examples", POPL '11).  While a single nesting of loops seems to be
+sufficient FlashFIll's intended uses, our benchmark suite (based on
+real-world uses of the Augeas tool) includes many more complex formats and
+transformations.  In particular, these formats tend to have deep layers of
+nested transformations, with complex transformations at each layer.
+FlashFill cannot handle such examples.
+
+[BCP: Deep layers of transformations?  Or of stars (which is the issue we
+started with)?  The last sentence or two are not convincing yet!
 
@@
 
-- What modifications did we need to make to programs in our Augeas
-  "benchmark suite" to make them bijective?
+     - What modifications did we need to make to examples in the Augeas
+       benchmark suite to make them bijective?
 
-The changes we made to the benchmark suite fall in two major categories that
-mirror the intended extensions to our system.
+[BCP: Open by reminding the reader (who has probably forgotten, or who may
+not even have read the paper yet!) what the question/issue in the reviews
+was...]
 
+The changes we made to the benchmark suite were of two forms:
 1) Whitespace was present in one format but not the other.
@@
 
-Category 1 was by far the most prevalent, and encountered in many of the Augeas
+Category 1 was by far the most common, applying to many of the Augeas
 examples.  To handle this, we added additional whitespace to the other format,
@@
 
-Instead of changing the formats, we could have extended the system to involve
-canonizers.  Canonizers can be placed on the ends of lenses to
-normalize the input data.  Canonizers can normalize an arbitrarily long sequence
-of whitespace to a single space.  Canonizers also can be used to handle casing
-issues: normalizing a sequence of uppercase and lowercase characters into purely
-uppercase characters.  We have recently developed a framework to annotate a
-regular expression with canonizing information, allowing for a single
-term to represent an arbitrary sequence of whitespace and a means to normalize
-it to a single space.  Using this framework, we can generate non-bijective lenses 
-using our framework, as long as the canonized forms of their source and target 
-types are in bijective correspondence.
+Instead of changing the formats, we could have extended the system to
+involve canonizers.  Canonizers can be placed on the ends of lenses to
+normalize the input data.  Canonizers can normalize an arbitrarily long
+sequence of whitespace to a single space.  Canonizers also can be used to
+handle casing issues: normalizing a sequence of uppercase and lowercase
+characters into purely uppercase characters.  We have recently developed a
+framework to annotate a regular expression with canonizing information,
+allowing for a single term to represent an arbitrary sequence of whitespace
+and a means to normalize it to a single space.  Using this framework, we can
+generate non-bijective lenses using our framework, as long as the canonized
+forms of their source and target types are in bijective correspondence.
 
@@
 of the target format containing the projected information.  Composing this
-generated lens with one that deletes that information at the tail of the file,
-when going left to right, and restores that information going back would provide
-the full lens.  This wrapper is not difficult to write, and we think that
-discovering this projection automatically is possible, but we have not yet done it.
+generated lens with one that deletes that information at the tail of the
+file, when going left to right, and restores that information going back
+would provide the full lens.  This wrapper is not difficult to write, and we
+think that discovering this projection automatically is possible, but we
+have not yet done it.
 
-There are a large number of formats that are in bijective correspondence. For
-example https://www.data.gov/ provides data in a variety of formats in bijective
-correspondence with each other.  Many of the formats present in existing lens libraries
-use either canonizers or projections.  The theories developed to handle synchronization 
-tasks for real world data (quotient lenses for Category 1, and symmetric lenses 
-for Category 2), have a bijective core, surrounded by either canonizers or projections.  
-We believe that our system can be used as a drop-in component for synthesis algorithms 
-on these more complex lens structures.
+There are a large number of formats that are in bijective
+correspondence. For example https://www.data.gov/ provides data in a variety
+of formats in bijective correspondence with each other.  Many of the formats
+present in existing lens libraries use either canonizers or projections.
+The theories developed to handle synchronization tasks for real world
+data (quotient lenses for Category 1, and symmetric lenses for Category 2),
+have a bijective core, surrounded by either canonizers or projections.  We
+believe that our system can be used as a drop-in component for synthesis
+algorithms on these more complex lens structures.
 
@@
 
-- After having written the regular expressions, is it very difficult for users
-  to then write the lens?  Are we saving anybody any significant effort?
+     - After having written the regular expressions, is it very difficult
+       for users to then write the lens?  Are we saving anybody any
+       significant effort?
 
@@
 Even the relatively simple lens that permuts the 3 elements needed for
-single_author_convert is fairly complex.  The complexity becomes even more apparent when
-working with large lenses comprised of many sub-lenses (though for brevity
-we give here the more minimal example above).  Lenses provide great power with their
-invertibility guarantees, but they come at the cost of thinking about many
-fiddly details when writing the terms.  Worrying about these fiddly details
-while ALSO thinking about unambiguity restrictions is a very difficult task!
+single_author_convert is fairly complex.  The complexity becomes even more
+apparent when working with large lenses comprised of many sub-lenses (though
+for brevity we give here the more minimal example above).  Lenses provide
+great power with their invertibility guarantees, but they come at the cost
+of thinking about many fiddly details when writing the terms.  Worrying
+about these fiddly details while ALSO thinking about unambiguity
+restrictions is a very difficult task!
 
@@
 examples to programs, Optician allows programmers to get the benefits of a
-powerful language while avoiding much of the complexity of programming in that
-language.  This paradigm allows for more complex programs to be
-synthesized, the types of programs that necessitate programming in such a DSL,
-but also provide a burden on the user.  However, this burden is already one
-incured by programmers using best practices, so we are merely saving them the
-time spent on tedious work.
+powerful language while avoiding much of the complexity of programming in
+that language.  This paradigm allows for more complex programs to be
+synthesized, the types of programs that necessitate programming in such a
+DSL, but also provide a burden on the user.  However, this burden is already
+one incured by programmers using best practices, so we are merely saving
+them the time spent on tedious work.
 
@@
 paraphrase of it.] [AFM: Quoted but will add in quotes to make it explicit]
+[BCP: I see -- I was misreading (in part because of "Strengths/weaknesses",
+which comes from the review but is not "quoted").  No need to add in quotes.
+But maybe indenting the comments more will separate them better, visually.
+Let's also clean up the spacing.]
 
@@
 
-- I also have a concern: sicne the approach is
+- I also have a concern: since the approach is
   limited to bijective lenses, but many of the test examples were not

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
appendix

--- 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
@@
 
-The first three (extr-fname, extr-fname-err and extr-quant) we chose because
-they corresponded to the first two examples in the paper.
-The problem extr-fname only allowed input paths with a file at the end, where
-extr-fname-err provided an error message if no file was present.
+We used benchmarks from three sources: Augeas benchmarks, FlashFill benchmarks,
+and custom benchmarks.  We used Augeas benchmarks because it is a lens-based
+tool that has seen industrial use.  We used FlashFill benchmarks because it is
+a well-known and popular string transformation synthesis system.
+We used custom benchmarks to find good uses for the tool in areas not previously
+explored (like converting user address information from a CSV format into a
+letter format).
+
+For our Augeas benchmarks, we began synthesizing lenses for arbitrary problems
+in the directory.  After a few of these, to show we weren't cherry-picking
+lenses that our system could handle, we began turning the lenses into synthesis
+problems alphabetically - first the As, then the Bs.
+
+We chose the first three FlashFill benchmarks (extr-fname, extr-fname-err and
+extr-quant) because they corresponded to the first two examples in the
+paper. The problem extr-fname only allowed input paths with a file at the end,
+where extr-fname-err provided an error message if no file was present.
 The final example, extr-num was used because extracting a phone number was used
@@
 
-The text is wrong, there are only 39 benchmarks.
+Thank you for pointing out this typo, there are only 39 benchmarks.
 
@@
 
+We support the bijective subset given by the combinators in Section 2.  See the
+general responses for a discussion on the usefulness of these combinators.
+
 The subset our approach supports are the bijective lenses where all regular
@@
 
-The preliminaries are us defining a specific subset of Boomerang that we focus
-on.  Boomerang supports a variety of lenses - asymmetric lenses, quotient
-lenses, and alignment lenses.  We define which subset of Boomerang we aim to
-support.
+The preliminaries define a specific subset of Boomerang that we aim to
+synthesize.  Boomerang supports a variety of lenses - asymmetric lenses,
+quotient lenses, and alignment lenses defined over a series of papers
+[Combinators for bi-directional tree transformations: a linguistic approach to
+the view update problem, POPL 2005; Boomerang: Resourceful Lenses for String
+Data, POPL 2008; Quotient Lenses, ICFP 2008].  In the preliminaries, we define
+which subset of Boomerang we aim to synthesize. While this language does not
+appear verbatim in a prior paper, it is a natural core language for string
+lenses.
 
@@
 
-    Yes, there are certainly times where more than lenses needs to be added.
-    The specific instance of single calls being split into multiple
-    calls can be handled by this paradigm if the server allows for batch
-    requests.  In particular, APIs of the running example we use, the modern
-    format does have batching capabilities,
+    Yes, there are certainly times where more than lenses need to be added,
+    though lenses still solve part of this problem. The specific instance of
+    single calls being split into multiple calls can be handled by this paradigm
+    if the server allows for batch requests.  In particular, APIs of the running
+    example we use, the modern format does have batching capabilities,
     https://www.visualstudio.com/en-us/docs/integrate/api/wit/batch .

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
appendix fixes

@@
 
-In this context, "using it" refers to allowing retyping on the specified
-equivalences.  We are trying to carve out a specific subset of Boomerang that
-we want to synthesize, and are not allowing the subset requiring a courser
-notion of equivalence to be included. The problems with including full language
-equivalence is detailed on pages 8 through 10.
+The big picture is: we are trying to carve out a core, bijective subset of
+Boomerang, to synthesize.  We want to find a subset that is
+sufficiently expressive, but also permits an efficient synthesis algorithm.
+For this subset to be sufficiently expressive, it is important to have a type
+equivalence rule, where a regular expression type can be substituted for an
+equivalent one.  However, we choose to use a notion of equivalence that is finer
+than language equivalence: star semiring equivalence.  The reasons we choose to
+use star semiring equivalence instead of language equivalence are detailed on
+pages 8 through 10.
 
@@
     single calls being split into multiple calls can be handled by this paradigm
-    if the server allows for batch requests.  In particular, APIs of the running
-    example we use, the modern format does have batching capabilities,
+    if the server allows for batch requests.  In particular, in the APIs of the
+    running example we use, the modern format does have batching capabilities,
     https://www.visualstudio.com/en-us/docs/integrate/api/wit/batch .

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
spacing

@@
 transformations.  In particular, these formats tend to have deep layers of
-nested transformations, with complex transformations at each layer.
+nested loops, with complex transformations at each layer.
 FlashFill cannot handle such examples.
 
-[BCP: Deep layers of transformations?  Or of stars (which is the issue we
-started with)?  The last sentence or two are not convincing yet!
+[BCP: The last sentence or two are not convincing yet!
 
@@
 
-[BCP: Open by reminding the reader (who has probably forgotten, or who may
-not even have read the paper yet!) what the question/issue in the reviews
-was...]
-
-The changes we made to the benchmark suite were of two forms:
+Many reviewers expressed concern that the changes made to the formats in our
+benchmark suite to make them bijective were major and time consuming. The
+changes we made to the benchmark suite were of two forms:
 1) Whitespace was present in one format but not the other.
@@
 
-[BCP: In the "appendix" I think it's important to quote the reviews
-verbatim, so that reviewers can see what they said rather than our
-paraphrase of it.] [AFM: Quoted but will add in quotes to make it explicit]
-[BCP: I see -- I was misreading (in part because of "Strengths/weaknesses",
-which comes from the review but is not "quoted").  No need to add in quotes.
-But maybe indenting the comments more will separate them better, visually.
-Let's also clean up the spacing.]
-
+===========================================================================
 Review #18A
 
-Strengths/weaknesses
+                                -------------
 
@@
 
-- Important parts of algorithm not explained (regarding how examples are used)
+                                -------------
+
+    - Important parts of algorithm not explained (regarding how examples are
+      used)
 
@@
 
-- Not clear how much work it is to adapt examples to be suitable for bijective
-  lens synthesis, or how useful the results are w.r.t. the original task
+                                -------------
+
+    - Not clear how much work it is to adapt examples to be suitable for
+      bijective lens synthesis, or how useful the results are w.r.t. the
+      original task
 
@@
 
+                                -------------
+
     - But the main part of the paper is not self-contained - for example the
-  algorithm description relies on (example-dependent) orderings of sequences and
-  atoms that are not described; this is a key part of the algorithm since it is
-  where the provided examples are actually used.
+      algorithm description relies on (example-dependent) orderings of
+      sequences and atoms that are not described; this is a key part of the
+      algorithm since it is where the provided examples are actually used.
 
 See above.
+
+                                -------------
 
@@
 
-- I could see future work building on this to handle the non-bijective case 
-  being a separate (and also interesting) paper.  But I would want the algorithm
-  to be described more clearly and more caveats about the examples.
+                                -------------
+
+    - I could see future work building on this to handle the non-bijective
+      case being a separate (and also interesting) paper.  But I would want
+      the algorithm to be described more clearly and more caveats about the
+      examples.
 
@@
 
+                                -------------
+
     - p12 fig 6. I'm not sure how ule "DNF structural rewrite" could be
-  applied as the LHS uses the operators defined in fig. 5, rather than syntax
+      applied as the LHS uses the operators defined in fig. 5, rather than
+      syntax
 
@@
 
+                                -------------
+
     - p17.  The desription of RigidSynth is incomplete.  It mentions orders
-  $\leq_{Seq}$ and $\leq_{Atom}$ (and variants that use the examples) that do
-  not appear to be defined.  What happens if there is ambiguity?  for
+      $\leq_{Seq}$ and $\leq_{Atom}$ (and variants that use the examples) that
+      do not appear to be defined.  What happens if there is ambiguity?  for
       example what if we are trying to infer a lens of type (a | b | c) <->
@@
 
+                                -------------
+
     - p18: AllSome (alg. 3): I assume this is the obvious function of type
@@
 This is correct.
+
+                                -------------
 
@@
 
+                                -------------
+
     - p20: "We adapted these examples" please say mor about how many of the
@@
 See general response.
+
+                                -------------
 
@@
 
+                                -------------
+
     - p21: Fig 9 "Number of expansions" --> "Number of solutions"??
@@
 perform.
+
+                                -------------
 
@@
 small, as is shown in Figure 11.
+
+                                -------------
 
@@
 
-Paper summary
                                 -------------
 
-- However, I’m not sure what exact advantage the proposed approach brings to the
-  table. In particular, the user is still required to write exact input and output
-  types of the transformations via regular expressions. How hard is to write
-  lenses after one has such expressions? ...
-  In the reviewer’s opinion, the specification mechanism is not very natural and
-  it’s unclear whether it actually simplifies writing bijective transformations.
-  This aspect is not evaluated and it would be nice to at least see the difference
-  in sizes between the domain/range regexes vs the Boomerang lenses. In general,
-  there is no report of the sizes of benchmarks.
+    - However, I’m not sure what exact advantage the proposed approach brings
+      to the table. In particular, the user is still required to write exact
+      input and output types of the transformations via regular expressions.
+      How hard is to write lenses after one has such expressions? ...
+      In the reviewer’s opinion, the specification mechanism is not very
+      natural and it’s unclear whether it actually simplifies writing
+      bijective transformations. This aspect is not evaluated and it would be
+      nice to at least see the difference in sizes between the domain/range
+      regexes vs the Boomerang lenses. In general, there is no report of the
+      sizes of benchmarks.
 
@@
 
-We can make sure we can provide at least some description of the sizes of specs
-and lenses in the final version of the paper.
+We can make sure we can provide at a description of the sizes of specs and
+lenses in the final version of the paper.
+
+                                -------------
 
     - Very limited DSL for transformations. The proposed DSL can only describe
-  transformations that use identities and constant functions over input subparts.
-  For example, one cannot synthesize the function toUpperCase, which I imagine
-  could be useful for the kind of transformations appearing in the motivation.
+      transformations that use identities and constant functions over input
+      subparts. For example, one cannot synthesize the function toUpperCase,
+      which I imagine could be useful for the kind of transformations
+      appearing in the motivation.
 
@@
 
-- Moreover, the functions can use each word in the input only once: one cannot
-  express the transformation abc->abcabc. Most bijective functions I can think of
-  (string encoders, compression algorithms, encryption algorithms) require complex
-  transformations over the input that go beyond identities.
+                                -------------
+
+    - Moreover, the functions can use each word in the input only once: one
+      cannot
+      express the transformation abc->abcabc. Most bijective functions I can
+      think of (string encoders, compression algorithms, encryption
+      algorithms) require complex transformations over the input that go
+      beyond identities.
 
@@
 restriction.
+
+                                -------------
+
     - On this note, there was a paper this year at PLDI on inverting string encoders
@@
 
-- Somewhat limited evaluation. The evaluation is performed on two families of
-  benchmarks. The first family is an adaptation of format translation problems
-  from Augeas a tool for configuration editing and the second family is from
-  FlashFill a tool for synthesis of string transformations. All the benchmarks
-  presented are “adaptation” and it’s unclear whether the authors only picked
-  benchmarks for which they knew a priori a program in their DSL existed. Based
-  on the presented data, I cannot assess how often the presented technique could
-  be applied.
+                                -------------
+
+    - Somewhat limited evaluation. The evaluation is performed on two families
+      of benchmarks. The first family is an adaptation of format translation
+      problems from Augeas a tool for configuration editing and the second
+      family is from FlashFill a tool for synthesis of string transformations.
+      All the benchmarks presented are “adaptation” and it’s unclear whether
+      the authors only picked benchmarks for which they knew a priori a
+      program in their DSL existed. Based on the presented data, I cannot
+      assess how often the presented technique could be applied.
 
@@
 
+                                -------------
+
     - Next, the comparison with the tool FlashFill/FlashExtract is somewhat
-  confusing. I suspect the authors had to provide some domain and range regular
-  expressions for the flash fill benchmarks, but this aspect, which is crucial to
-  help the search, is not mentioned.
+      confusing. I suspect the authors had to provide some domain and range
+      regular expressions for the flash fill benchmarks, but this aspect,
+      which is crucial to help the search, is not mentioned.
 
@@
 paper.
+
+                                -------------
 
@@
 
+                                -------------
+
     - Also, a better explanation of the
-  difference between the DSLs of Lenses and Flash fill is required. For example,
-  FlashFill can reverse a string, but the proposed language can’t.
+      difference between the DSLs of Lenses and Flash fill is required. For
+      example, FlashFill can reverse a string, but the proposed language
+      can’t.
 
@@
 
-- What is the difference in size between the regexes for domain/range spec and
-  the full lenses?
+                                -------------
+
+    - What is the difference in size between the regexes for domain/range spec
+      and the full lenses?
 
@@
 
-- What is the difference in size between manually written lenses and synthesized
-  ones?
+                                -------------
+
+    - What is the difference in size between manually written lenses and
+      synthesized ones?
 
@@
 
-- Clarify how the benchmarks were selected and whether any were not selected a
-  priori. Especially the flash* ones. Did you omit those that used character
-  transformations such as upper-case?
+                                -------------
+
+    - Clarify how the benchmarks were selected and whether any were not
+      selected a priori. Especially the flash* ones. Did you omit those that
+      used character transformations such as upper-case?
 
@@
 
+                                -------------
+
     - Can the proposed DSL express any transformation that FlashFill can’t and
@@
 
+                                -------------
+
     - P19: maximally factors the concats and ors. What does this mean? Regex
-  minimization is a hard problem, what is that one factors exactly? I think it
-  would help to show one example of fully synthesized lens in the paper.
+      minimization is a hard problem, what is that one factors exactly? I
+      think it would help to show one example of fully synthesized lens in the
+      paper.
 
@@
 
-Minor:
-- Eval the text says Full synthesized 49 benchmarks, but the plot in fig 8 only
-  shows 38/39?
+                                -------------
+
+    - Eval the text says Full synthesized 49 benchmarks, but the plot in fig 8
+      only shows 38/39?
 
@@
 
-- Overall, this paper looks like a solid and useful contribution to the lens
-  literature. However, this reviewer has never used, let alone designed, a lens
-  library. As such, I found the paper very hard to read and assess. I would have
-  liked to see a clear problem statement (such as an example where it's annoying
-  to write the Boomerang code but easy to synthesize the lens using your approach
+                                -------------
+
+    - Overall, this paper looks like a solid and useful contribution to the
+      lens literature. However, this reviewer has never used, let alone
+      designed, a lens library. As such, I found the paper very hard to read
+      and assess. I would have liked to see a clear problem statement (such as
+      an example where it's annoying to write the Boomerang code but easy to
+      synthesize the lens using your approach
 
@@
 problem statement more clear in the final version.
+
+                                -------------
 
@@
 
+                                -------------
+
     - It was also not very clear to me what subset of Boomerang your approach
-  supports. The introduction says it is a "useful subset of Boomerang" - but
-  what does that mean?
+      supports. The introduction says it is a "useful subset of Boomerang" -
+      but what does that mean?
 
@@
 
-- I was also a bit confused about the preliminaries in Sec. 2. In particular, it
-  is not clear whether all or most of those preliminaries are from related work
-  (but there are no references) or whether they are contributions of this paper.
+                                -------------
+
+    - I was also a bit confused about the preliminaries in Sec. 2. In
+      particular, it is not clear whether all or most of those preliminaries
+      are from related work (but there are no references) or whether they are
+      contributions of this paper.
 
@@
 
+                                -------------
+
     - p.4, "naively using it in the context of lens synthesis presents several
-  problems"? Which problems? At this point I have no clue what "using it" may
-  even mean. The "big picture" is not clear at all.
+      problems"? Which problems? At this point I have no clue what "using it"
+      may even mean. The "big picture" is not clear at all.
 
@@
 
+                                -------------
+
     - p. 3, "We have ommitted these theorems for space..." - maybe you mean
@@
 
-- p. 7, "instead of maintaining server code for each endpoint, we envision..."
-  but surely the changes to an API involve more than just data format changes.
-  For instance, single calls may have to be split into multiple calls. Hence it
-  does not seem to be realistic to maintain the old API using _only_ synthesized
-  lenses.
+                                -------------
+
+    - p. 7, "instead of maintaining server code for each endpoint, we
+      envision..." but surely the changes to an API involve more than just
+      data format changes. For instance, single calls may have to be split
+      into multiple calls. Hence it does not seem to be realistic to maintain
+      the old API using _only_ synthesized lenses.
 

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
new updates to rebuttal

@@
 One more technical note about point: FlashFill's DSL does permit nested
-loops, but for efficiency reasons the synthesis algorithm only synthesizes
+loops, but for efficiency reasons the algorithm only synthesizes
 programs with no nesting (this is discussed at the end of section 4.4 of
@@
 Examples", POPL '11).  While a single nesting of loops seems to be
-sufficient FlashFIll's intended uses, our benchmark suite (based on
+sufficient for FlashFill's intended uses, our benchmark suite (based on
 real-world uses of the Augeas tool) includes many more complex formats and
 transformations.  In particular, these formats tend to have deep layers of
-nested loops, with complex transformations at each layer.
-FlashFill cannot handle such examples.
-
-[BCP: The last sentence or two are not convincing yet!
+iterated data, with complex transformations at each layer, requiring the
+synthesis of many layers of nested loops.  FlashFill cannot handle such
+examples.
 
@@
 
-     - What modifications did we need to make to examples in the Augeas
-       benchmark suite to make them bijective?
+     - What modifications did we need to make to examples in the Augeas and
+       FlashFill benchmark suite to make them bijective?
 
-Many reviewers expressed concern that the changes made to the formats in our
-benchmark suite to make them bijective were major and time consuming. The
-changes we made to the benchmark suite were of two forms:
+Many reviewers expressed concern that major changes were added to the formats in
+our benchmark suite to make them bijective. We made changes to the benchmark
+suite to handle two forms of non-bijectivity:
 1) Whitespace was present in one format but not the other.
@@
 
-Instead of changing the formats, we could have extended the system to
-involve canonizers.  Canonizers can be placed on the ends of lenses to
-normalize the input data.  Canonizers can normalize an arbitrarily long
-sequence of whitespace to a single space.  Canonizers also can be used to
-handle casing issues: normalizing a sequence of uppercase and lowercase
-characters into purely uppercase characters.  We have recently developed a
-framework to annotate a regular expression with canonizing information,
-allowing for a single term to represent an arbitrary sequence of whitespace
-and a means to normalize it to a single space.  Using this framework, we can
-generate non-bijective lenses using our framework, as long as the canonized
-forms of their source and target types are in bijective correspondence.
-
 Category 2 was less prevalent, but still present, typically in the FlashFill
 examples.  To address this category, we added additional information to the end
-of the target format containing the projected information.  Composing this
-generated lens with one that deletes that information at the tail of the
-file, when going left to right, and restores that information going back
-would provide the full lens.  This wrapper is not difficult to write, and we
-think that discovering this projection automatically is possible, but we
-have not yet done it.
+of the target format containing the projected information.
 
-There are a large number of formats that are in bijective
-correspondence. For example https://www.data.gov/ provides data in a variety
-of formats in bijective correspondence with each other.  Many of the formats
-present in existing lens libraries use either canonizers or projections.
-The theories developed to handle synchronization tasks for real world
-data (quotient lenses for Category 1, and symmetric lenses for Category 2),
-have a bijective core, surrounded by either canonizers or projections.  We
-believe that our system can be used as a drop-in component for synthesis
-algorithms on these more complex lens structures.
+                                -------------
+
+     - The restriction to bijective lenses is not fully justified
+
+There are still many formats kept in bijective correspondence with each other,
+like those present at https://www.data.gov/ .
+
+There are lens theories developed to handle synchronization tasks for data
+formats not kept in bijective correspondence (quotient lenses to handle Category
+1 non-bijectivity, and symmetric lenses to handle Category 2 non-bijectivity).
+These theories have a bijective core, surrounded by either canonizers (for
+quotient lenses) or projections (for symmetric lenses).  We believe that our
+system can be used as a drop-in component for synthesis algorithms on these more
+complex lens structures.
+
+These justifications can be added to the paper.
 
@@
 
-Even the relatively simple lens that permuts the 3 elements needed for
+Even the relatively simple lens that permutes the 3 elements needed for
 single_author_convert is fairly complex.  The complexity becomes even more
@@
 
-See above.
+See general responses.
+
+For Category 1 non-bijectivity, instead of changing the formats, we could have
+placed canonizers on the edges of formats with large amounts of whitespace.
+Canonizers can normalize an arbitrarily long sequence of whitespace to a single
+space.  Canonizers also can be used to handle casing issues: normalizing a
+sequence of uppercase and lowercase characters into purely uppercase characters.
+In this manner, whitespace (or casing) information can be normalized, and no
+longer would need to be propagated to the target format. We have recently
+developed a framework to annotate a regular expression with canonizing
+information, allowing for a single term to represent an arbitrary sequence of
+whitespace and a means to normalize it to a single space.  Using this framework,
+we can generate non-bijective lenses using our framework, as long as the
+canonized forms of their source and target types are in bijective
+correspondence.
+
+For Category 2 non-bijectivity, we could compose the generated lens (that has
+projected information at the end) with one that deletes that information at the
+tail of the file, when going left to right, and restores that information when
+going back. This wrapper is not difficult to write, and we think that
+discovering this projection automatically is possible, but we have not yet done
+it.
 
@@
 
-equivalence classes between user defined data types that have lenses, choose a
-representative, and transform the user defined data types into that
-representative, and convert the strings into elements of that representative
-data type.  Then we proceed with synthesis as described by the formal algorithm.
-
                                 -------------
@@
      - Moreover, the functions can use each word in the input only once: one
-      cannot
-      express the transformation abc->abcabc. Most bijective functions I can
-      think of (string encoders, compression algorithms, encryption
+       cannot express the transformation abc->abcabc. Most bijective functions I
+       can think of (string encoders, compression algorithms, encryption
        algorithms) require complex transformations over the input that go
@@
 
-See general response for more detailed information about the bijective
-restriction.
+See general response.
 
@@
 
-The subset our approach supports are the bijective lenses where all regular
-expression equivalences used in the typing derivation can be proven with
-the star-semiring equivalence rules.  We find this useful in that all lenses can
-easily be converted into a bijective lens, and we could not find a lens in
-Boomerang or Augeas that requires a courser notion of equivalence than
-star-semiring equivalence.
-
                                 -------------
@@
 The big picture is: we are trying to carve out a core, bijective subset of
-Boomerang, to synthesize.  We want to find a subset that is
+Boomerang to synthesize.  We want to find a subset that is
 sufficiently expressive, but also permits an efficient synthesis algorithm.

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
minor nits

@@
 our benchmark suite to make them bijective. We made changes to the benchmark
-suite to handle two forms of non-bijectivity:
+formats to handle two forms of non-bijectivity:
 1) Whitespace was present in one format but not the other.
@@
 
-There are still many formats kept in bijective correspondence with each other,
-like those present at https://www.data.gov/ .
+There are many formats kept in bijective correspondence with each other, like
+those present at https://www.data.gov/ .
 
@@
 description of the algorithm is present in algorithm 4 in the appendix (see the
-supplemental pdf submitted with the paper), on page
-119, with the orderings defined in Definition 33 on page 118 (the definitions of
-related structures begin at page 115).  We embed the parse trees of the examples
-into the DNF regular expressions, and make an ordering on the DNF regular
-expression with embedded parse trees.  We would be happy to rewrite the paper to
-include whatever information the reviewer believes is most important.
+supplemental pdf submitted with the paper) on page 119, with the orderings
+defined in Definition 33 on page 118 (the definitions of related structures
+begin at page 115).  We embed the parse trees of the examples into the DNF
+regular expressions, and make an ordering on the DNF regular expression with
+embedded parse trees.  We would be happy to rewrite the paper to include
+whatever information the reviewer believes is most important.
 
@@
 sequence of uppercase and lowercase characters into purely uppercase characters.
-In this manner, whitespace (or casing) information can be normalized, and no
-longer would need to be propagated to the target format. We have recently
-developed a framework to annotate a regular expression with canonizing
-information, allowing for a single term to represent an arbitrary sequence of
-whitespace and a means to normalize it to a single space.  Using this framework,
-we can generate non-bijective lenses using our framework, as long as the
-canonized forms of their source and target types are in bijective
+In this manner, whitespace (or casing) information can be normalized before
+synthesis, and no longer would need to be propagated to the target format. We
+have recently developed a framework to annotate a regular expression with
+canonizing information, allowing for a single term to represent an arbitrary
+sequence of whitespace and a means to normalize it to a single space.  Using
+this framework, we easily can generate non-bijective lenses using our framework,
+as long as the canonized forms of their source and target types are in bijective
 correspondence.
@@
 We feel that they are easy to understand because they look similar to how we
-would handwrite the lenses,
-but we have not done any user studies to justify that
-they are easy to understand.  Evidence to the fact that they are easier to
+would handwrite the lenses, but we have not done any user studies to justify
+that they are easy to understand.  Evidence to the fact that they are easier to
 understand than non-pretty printed versions is given by the sizes of the
@@
 
-See general response.
+See above.
 
@@
 examples provided.  Typically, a small number of examples can greatly reduce the
-number of possible solutions (TODO data).  The number of examples is indeed
-small, as is shown in Figure 11.
+number of possible solutions.  The number of examples is indeed small, as is
+shown in Figure 11.
+
+For example, there are an average of 2709 solutions with 0 randomly generated
+examples, 2016 with 2 randomly generated examples, and 755 with 5 randomly
+generated examples.  We are omitting the data from aug/bbhosts, whose number of
+possible solutions (in the 10^17 range) overwhelmed all other data.
 
@@
 used to normalize a format (or portions of a format) into fully uppercase or
-fully lowercase data.
+fully lowercase data.  We believe quotient lenses have a bijective core, so our
+synthesis engine for bijective lenses can be used as a plug-in component for
+tasks (like to_upper) requiring quotient lenses.
 
@@
 
-     - On this note, there was a paper this year at PLDI on inverting string encoders
-       (Automatic program inversion using symbolic transducers, by Hu et al.), a
-       topic which seems very related, but for which I couldn’t find a citation.
+     - On this note, there was a paper this year at PLDI on inverting string
+       encoders (Automatic program inversion using symbolic transducers, by Hu
+       et al.), a topic which seems very related, but for which I couldn’t find
+       a citation.
 
@@
 Finite Transducers.  The finite lookahead does not allow for transformations
-that swap data, a common pattern and important pattern in Boomerang lenses. We
-will add a citation to this work and a comment detailing the restrictions of
-symbolic finite transducers.
+that swap data of indeterminite size, a common pattern and important pattern in
+Boomerang lenses. We will add a citation to this work and a comment detailing
+the restrictions of extended symbolic finite transducers.
 
@@
 
-     - Also, a better explanation of the
-       difference between the DSLs of Lenses and Flash fill is required. For
-       example, FlashFill can reverse a string, but the proposed language
-       can’t.
+     - Also, a better explanation of the difference between the DSLs of Lenses
+       and Flash fill is required. For example, FlashFill can reverse a string,
+       but the proposed language can’t.
 
@@
 On average, the domain and range specs are slightly larger than the lens itself.
-See the above chart.  See the general response for a detailed commentary
-explaining why we believe writing lenses is harder than writing regular
-expressions, despite the fact that lenses alone are sometimes shorter than
-regular expression specifications.
+See above for the chart and a commentary explaining why we believe writing
+lenses is harder than writing regular expressions, despite the fact that lenses
+alone are sometimes shorter than regular expression specifications.
 
@@
 sufficiently expressive, but also permits an efficient synthesis algorithm.
+
 For this subset to be sufficiently expressive, it is important to have a type
@@
 
-Correct
+Correct, thank you.
 

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
add in data
add in changes to intro with Dave's help

@@
 
-Since we restrict to bijections, we are unable to express functions that
+Since we restrict ourselves to bijections, we are unable to express functions
+that
 don't preserve data, such as
@@
 synthesis of many layers of nested loops.  FlashFill cannot handle such
-examples.
+examples.  Our experiments demonstrate this fact by showing that FlashFill
+only succeeds on [x] out of our [y] benchmarks, whereas our system succeeds on
+all our benchmarks.
 
@@
 
-These justifications can be added to the paper.
+Reviewer 1 asked us directly about the potential for future work.
+Along those lines, we can say we have already implemented a system that
+synthesizes quotient lenses from regular expressions augmented with equivalence
+relations that, for instance, say that whitespace is irrelevant.  This system
+uses the algorithm provided in this paper as a drop-in component without change.
+We conjecture the system could be augmented in a similar way with projections as
+well.  We are not suggesting adding more results of this form to the paper
+because the paper is already very dense, but transformations of the form
+(f o bijection o g) are very powerful, and can be synthesized using our engine
+as a drop-in component when f and g are provided.
 
@@
 programmer to save the difficulties of writing the term by using work they
-should already be performing.  To highlight these difficulties, we show an
+should already be performing.
+
+There are a few more reasons why programming with regular expressions are easier
+than programming with lenses.
+
+* Regular expressions are common - almost every programmer has a firm
+grip on regular expressions.  In contrast, only a handful of programmers know
+anything about programming with lenses.
+
+* A regular expression describes one piece of data.  When writing a lens, one
+must think about both pieces of data at the same time, and simultaneously
+translate back and forth between them.  This is a non-standard programming task.
+
+* One must sometimes compute in their own head how to unroll or transform the
+regular expression into the right form, before writing the lens.
+
+TODO: move to appropriate places in other users, and remove *s from 18B
+To highlight these difficulties, we show an
 example program for mapping the author field of a BibTeX entry:
@@
 restrictions is a very difficult task!
-
-Optician is aimed at different audiences than existing string transformation
-tools.  While existing tools allow the average end-user to generalize their
-examples to programs, Optician allows programmers to get the benefits of a
-powerful language while avoiding much of the complexity of programming in
-that language.  This paradigm allows for more complex programs to be
-synthesized, the types of programs that necessitate programming in such a
-DSL, but also provide a burden on the user.  However, this burden is already
-one incured by programmers using best practices, so we are merely saving
-them the time spent on tedious work.
 
@@
 understand than non-pretty printed versions is given by the sizes of the
-generated lenses.  Without minimization, the lenses sizes average at (SIZE)
-AST nodes, whereas minimization made the sizes average at (SIZE) ast nodes.
+generated lenses.  Without minimization, the lenses sizes average at 424 AST
+nodes, whereas minimization made the sizes average at 276 AST nodes.
 In the general response, we show what a minimized vs handwritten lens look like.

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
update intro more and move around the program

@@
 examples.  Our experiments demonstrate this fact by showing that FlashFill
-only succeeds on [x] out of our [y] benchmarks, whereas our system succeeds on
+only succeeds on 3 out of our 39 benchmarks, whereas our system succeeds on
 all our benchmarks.
@@
 regular expression into the right form, before writing the lens.
-
-TODO: move to appropriate places in other users, and remove *s from 18B
-To highlight these difficulties, we show an
-example program for mapping the author field of a BibTeX entry:
-
-      typedef NAME = UPPERCASE (LOWERCASE)*;;
-
-      typedef LASTCOMMASTART = NAME "," (WSP NAME)*;;
-      typedef BIBTEXAUTHORLIST = LASTCOMMASTART (" and " LASTCOMMASTART)*;;
-      typedef BIBTEXAUTHORINFO = "author={" BIBTEXAUTHORLIST "}";;
-
-      typedef AUTAG = "au - ";;
-      typedef STARTTOEND = (NAME WSP)* NAME;;
-      typedef TAGGEDAUTHORDEFNS = AUTAG STARTTOEND ("\n " AUTAG STARTTOEND)*;;
-
-      let bibtex_to_endnote_au =
-        synth
-          BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS
-
-      let bibtex_to_endnote_au_synthesized_result
-        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
-        del "author={"
-          . ins "au - "
-          . lens_swap (NAME . del ",") (lens_swap WSP NAME)* 
-          . (del " and "
-               . ins "\n au - "
-               . lens_swap (NAME . del ",") (lens_swap WSP NAME)* )* 
-          . del "}"
-
-      let bibtex_to_endnote_au_human_written
-        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
-        let single_author_convert =
-          ins " au - "
-            . lens_swap
-                (NAME . del ",")
-                (lens_swap WSP NAME)*
-        in
-        del "author={"
-          . single_author_convert
-          . (del " and "
-               . ins "\n"
-               . single_author_convert)* 
-          . del "}"
-
-Even the relatively simple lens that permutes the 3 elements needed for
-single_author_convert is fairly complex.  The complexity becomes even more
-apparent when working with large lenses comprised of many sub-lenses (though
-for brevity we give here the more minimal example above).  Lenses provide
-great power with their invertibility guarantees, but they come at the cost
-of thinking about many fiddly details when writing the terms.  Worrying
-about these fiddly details while ALSO thinking about unambiguity
-restrictions is a very difficult task!
 
@@
 
-See general response.  We also provide data on the sizes below.  However, we
-do not think that purely lines of code is the best metric for our data, as our
-system provides benefits beyond saving keystrokes for the programmer, detailed
-below:
+See general response.  We provide an example program (and specification) below:
 
-* Regular expressions are common - almost every programmer has a firm grip on
-regular expressions.  In contrast, almost _no_ programmers know anything about
-programming with lenses.
+      typedef WSP = (" " | "\t" | "\n")+;;
+      typedef NAME = UPPERCASE (LOWERCASE)*;;
 
-* A regular expression describes one piece of data.  When writing a lens, one
-must think about both pieces of data at the same time, and simultaneously
-translate back and forth between them.
+      typedef LASTCOMMASTART = NAME "," (WSP NAME)*;;
+      typedef BIBTEXAUTHORLIST = LASTCOMMASTART (" and " LASTCOMMASTART)*;;
+      typedef BIBTEXAUTHORINFO = "author={" BIBTEXAUTHORLIST "}";;
 
-* One must sometimes compute in their own head how to unroll or transform the
-regular expression into the right form.
+      typedef AUTAG = "au - ";;
+      typedef STARTTOEND = (NAME WSP)* NAME;;
+      typedef TAGGEDAUTHORDEFNS = AUTAG STARTTOEND ("\n " AUTAG STARTTOEND)*;;
+
+      let bibtex_to_endnote_au =
+        synth
+          BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS
+
+      let bibtex_to_endnote_au_synthesized_result
+        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
+        del "author={"
+          . ins "au - "
+          . lens_swap (NAME . del ",") (lens_swap WSP NAME)* 
+          . (del " and "
+               . ins "\n au - "
+               . lens_swap (NAME . del ",") (lens_swap WSP NAME)* )* 
+          . del "}"
+
+      let bibtex_to_endnote_au_human_written
+        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
+        let single_author_convert =
+          ins " au - "
+            . lens_swap
+                (NAME . del ",")
+                (lens_swap WSP NAME)*
+        in
+        del "author={"
+          . single_author_convert
+          . (del " and "
+               . ins "\n"
+               . single_author_convert)* 
+          . del "}"
+
+Even the relatively simple lens that permutes the 3 elements needed for
+single_author_convert is fairly complex.  The complexity becomes even more
+apparent when working with large lenses comprised of many sub-lenses (though
+for brevity we give here the more minimal example above).  Lenses provide
+great power with their invertibility guarantees, but they come at the cost
+of thinking about many fiddly details when writing the terms.  Worrying
+about these fiddly details while ALSO thinking about unambiguity
+restrictions is a very difficult task!
+
+We also provide data on the sizes below:
 
@@
 
-We provide an extended example in the general response.  We will make the
-problem statement more clear in the final version.
+We provide an example program below.
+
+      typedef WSP = (" " | "\t" | "\n")+;;
+      typedef NAME = UPPERCASE (LOWERCASE)*;;
+
+      typedef LASTCOMMASTART = NAME "," (WSP NAME)*;;
+      typedef BIBTEXAUTHORLIST = LASTCOMMASTART (" and " LASTCOMMASTART)*;;
+      typedef BIBTEXAUTHORINFO = "author={" BIBTEXAUTHORLIST "}";;
+
+      typedef AUTAG = "au - ";;
+      typedef STARTTOEND = (NAME WSP)* NAME;;
+      typedef TAGGEDAUTHORDEFNS = AUTAG STARTTOEND ("\n " AUTAG STARTTOEND)*;;
+
+      let bibtex_to_endnote_au =
+        synth
+          BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS
+
+      let bibtex_to_endnote_au_synthesized_result
+        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
+        del "author={"
+          . ins "au - "
+          . lens_swap (NAME . del ",") (lens_swap WSP NAME)* 
+          . (del " and "
+               . ins "\n au - "
+               . lens_swap (NAME . del ",") (lens_swap WSP NAME)* )* 
+          . del "}"
+
+      let bibtex_to_endnote_au_human_written
+        : (lens in BIBTEXAUTHORINFO <=> TAGGEDAUTHORDEFNS) =
+        let single_author_convert =
+          ins " au - "
+            . lens_swap
+                (NAME . del ",")
+                (lens_swap WSP NAME)*
+        in
+        del "author={"
+          . single_author_convert
+          . (del " and "
+               . ins "\n"
+               . single_author_convert)* 
+          . del "}"
+
+Even the relatively simple lens that permutes the 3 elements needed for
+single_author_convert is fairly complex.  The complexity becomes even more
+apparent when working with large lenses comprised of many sub-lenses (though
+for brevity we give here the more minimal example above).  Lenses provide
+great power with their invertibility guarantees, but they come at the cost
+of thinking about many fiddly details when writing the terms.  Worrying
+about these fiddly details while ALSO thinking about unambiguity
+restrictions is a very difficult task!
+
+We will make the problem statement more clear in the final version.
 
@@
 
-We can include example code like this to highlight the frustrations a lens
-programmer must encounter while writing their programs.
+We can include example code like we did above to highlight the frustrations a
+lens programmer must encounter while writing their programs.
 

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
Polishing

@@
 
-                                -------------
-
      - What modifications did we need to make to examples in the Augeas and
@@
 of the target format containing the projected information.
-
-                                -------------
 
@@
 
-                                -------------
-
      - After having written the regular expressions, is it very difficult
@@
 
-There are a few more reasons why programming with regular expressions are easier
-than programming with lenses.
+A few more reasons why programming with regular expressions is easier
+than programming with lenses:
 
-* Regular expressions are common - almost every programmer has a firm
-grip on regular expressions.  In contrast, only a handful of programmers know
-anything about programming with lenses.
+* Regular expressions are widely understood.  Few programmers know about
+  programming with lenses.
 
-* A regular expression describes one piece of data.  When writing a lens, one
-must think about both pieces of data at the same time, and simultaneously
-translate back and forth between them.  This is a non-standard programming task.
+* A regular expression describes one data format, while lenses deal with two
+  at the same time.
 
-* One must sometimes compute in their own head how to unroll or transform the
-regular expression into the right form, before writing the lens.
+* In the (common) case where the two formats have kleene stars in different
+  places, manual lens programming requires mentally transforming both to a
+  common "aligned" form.  Our synthesis algorithm does this alignment
+  automatically.
 

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
fixed typo

@@
 
-One more technical note about point: FlashFill's DSL does permit nested
+One more technical note on this point: FlashFill's DSL does permit nested
 loops, but for efficiency reasons the algorithm only synthesizes

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
Polishing

--- 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
@@
 
-One more technical note about point: FlashFill's DSL does permit nested
+We note that FlashFill's DSL does permit nested
 loops, but for efficiency reasons the algorithm only synthesizes

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
polishing.
polishing

@@
 
-Many reviewers expressed concern that major changes were added to the formats in
+Reviewers expressed concern that major changes were added to the formats in
 our benchmark suite to make them bijective. We made changes to the benchmark
@@
 Category 1 was by far the most common, applying to many of the Augeas
-examples.  To handle this, we added additional whitespace to the other format,
+examples.  To handle it, we added additional whitespace to the other format,
 where typically whitespace is unnecessary.
 
-Category 2 was less prevalent, but still present, typically in the FlashFill
+Category 2 was less prevalent, but present, typically in the FlashFill
 examples.  To address this category, we added additional information to the end

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
polishing

@@
 
-Reviewers expressed concern that major changes were added to the formats in
-our benchmark suite to make them bijective. We made changes to the benchmark
-formats to handle two forms of non-bijectivity:
+Reviewers expressed concern that changes were made to the formats in
+our benchmark suite to make them bijective. We made changes to handle two forms of non-bijectivity:
 1) Whitespace was present in one format but not the other.

### 2017-09-13 papers/BijectiveLensSynth/popl_18_rebuttal.txt
polishing

@@
 input-output examples are unable to synthesize programs with nested loops.
-
 We note that FlashFill's DSL does permit nested
@@
 "Automating String Processing in Spreadsheets Using Input-Output
-Examples", POPL '11).  While a single nesting of loops seems to be
-sufficient for FlashFill's intended uses, our benchmark suite (based on
+Examples", POPL '11).
+
+While a single nesting of loops seems to be sufficient for FlashFill's intended 
+uses, our benchmark suite (based on
 real-world uses of the Augeas tool) includes many more complex formats and
@@
 
-Reviewers expressed concern that changes were made to the formats in
-our benchmark suite to make them bijective. We made changes to handle two forms of non-bijectivity:
-1) Whitespace was present in one format but not the other.
-2) Useful information was projected away going from one format to the other.
+We changed some formats in our benchmark suite to handle two forms of non-bijectivity:
+1) Whitespace present in one format but not the other.
+2) Useful information projected away when going from one format to the other.
 
@@
 
-Category 2 was less prevalent, but present, typically in the FlashFill
-examples.  To address this category, we added additional information to the end
-of the target format containing the projected information.
+Category 2 was less prevalent, but occurred, typically in the FlashFill
+examples.  To address this category, we added the projected information to the end
+of the format missing the information.
 

### 2017-09-14 papers/BijectiveLensSynth/popl_18_rebuttal.txt
polishing and shortening

@@
 Reviewer 1 asked us directly about the potential for future work.
-Along those lines, we can say we have already implemented a system that
+Along those lines, we have already implemented a system that
 synthesizes quotient lenses from regular expressions augmented with equivalence
-relations that, for instance, say that whitespace is irrelevant.  This system
+relations that solve the whitespace bijection problem.  This system
 uses the algorithm provided in this paper as a drop-in component without change.
 We conjecture the system could be augmented in a similar way with projections as
-well.  We are not suggesting adding more results of this form to the paper
-because the paper is already very dense, but transformations of the form
-(f o bijection o g) are very powerful, and can be synthesized using our engine
-as a drop-in component when f and g are provided.
+well.  We are not suggesting adding more results of this form to the current paper
+because it is already very dense.
 

### 2017-09-14 papers/BijectiveLensSynth/popl_18_rebuttal.txt
polishing

@@
 documentation for future programmers to understand what formats the lens
-maps between.  Furthermore, it provides an additional resilience to future
-modifications of the lens, as it validates that the lens maps exactly
-between strings of the provided formats - invalid inputs are not accepted
+maps between.  Furthermore, the types provide resilience in the face of future
+lens modifications, as they validate that the lens maps exactly
+between strings of the provided formats. Invalid inputs are not accepted
 nor are the outputs ever invalid.  In this way, we are not requiring
-additional work by the programmer to use our system; we are allowing the
-programmer to save the difficulties of writing the term by using work they
-should already be performing.
+additional work; we are saving programmers from the difficulties of 
+writing the lens by leveraging the work they should already be performing.
 

### 2017-09-14 papers/BijectiveLensSynth/popl_18_rebuttal.txt
polishing and shortening

@@
 Engineering best practices dictate that we annotate the lenses with their
-types.  While Boomerang does not need these annotations - being able to
+types.  While Boomerang does not require these annotations - being able to
 infer the types from the terms - the types of the term serve as
@@
 maps between.  Furthermore, the types provide resilience in the face of future
-lens modifications, as they validate that the lens maps exactly
+lens modifications, as the types ensure that the lens maps exactly
 between strings of the provided formats. Invalid inputs are not accepted
-nor are the outputs ever invalid.  In this way, we are not requiring
-additional work; we are saving programmers from the difficulties of 
-writing the lens by leveraging the work they should already be performing.
+nor are the outputs ever invalid.  Consequently, we are not requiring
+additional work. Instead, we are saving programmers from having to 
+writing the lens by leveraging work they should already be performing.
 
@@
   common "aligned" form.  Our synthesis algorithm does this alignment
-  automatically.
+  automatically.  This alignment can be quite challenging for programmers.
 

### 2017-09-14 papers/BijectiveLensSynth/popl_18_rebuttal.txt
some small changes and wording updates

@@
 Since we restrict ourselves to bijections, we are unable to express functions
-that
-don't preserve data, such as
+that don't preserve data, such as
   - mapping from a format with whitespace to one without whitespace,
@@
 While a single nesting of loops seems to be sufficient for FlashFill's intended
-uses, our benchmark suite (based on
-real-world uses of the Augeas tool) includes many more complex formats and
-transformations.  In particular, these formats tend to have deep layers of
-iterated data, with complex transformations at each layer, requiring the
-synthesis of many layers of nested loops.  FlashFill cannot handle such
-examples.  Our experiments demonstrate this fact by showing that FlashFill
-only succeeds on 3 out of our 39 benchmarks, whereas our system succeeds on
-all our benchmarks.
+uses, our benchmark suite (based on real-world uses of the Augeas tool) includes
+many more complex formats and transformations.  In particular, these formats
+tend to have deep layers of iterated data, with complex transformations at each
+layer, requiring the synthesis of many layers of nested loops.  FlashFill cannot
+handle such examples.  Our experiments demonstrate this fact by showing that
+FlashFill only succeeds on 3 out of our 39 benchmarks, whereas our system
+succeeds on all our benchmarks.
 
@@
 
-Number of expansions is the correct phrase.  Figure 9 shows how much random
-search through expansions is required before RigidSynth terminates.  In
-particular, this shows that full expansion inference performs significantly
-better than the partial one that only performs the expansions it knows it must
-perform.
+Number of expansions is the correct phrase.  Figure 9 shows how much
+nondeterministic search through expansions is required before RigidSynth
+terminates.  In particular, this shows that full expansion inference performs
+significantly better than the partial one that only performs the expansions it
+knows it must perform.
 
@@
 
-The number of examples provided doesn't typically have an impact on runspeed,
-only correctness.  The number of valid solutions does vary with the number of
-examples provided.  Typically, a small number of examples can greatly reduce the
-number of possible solutions.  The number of examples is indeed small, as is
-shown in Figure 11.
+The number of examples provided doesn't typically have an impact on runtime
+speed, only correctness.  The number of valid solutions does vary with the
+number of examples provided.  Typically, a small number of examples can greatly
+reduce the number of possible solutions.  The number of examples is indeed
+small, as is shown in Figure 11.
 
@@
 single_author_convert is fairly complex.  The complexity becomes even more
-apparent when working with large lenses comprised of many sub-lenses (though
-for brevity we give here the more minimal example above).  Lenses provide
-great power with their invertibility guarantees, but they come at the cost
-of thinking about many fiddly details when writing the terms.  Worrying
-about these fiddly details while ALSO thinking about unambiguity
-restrictions is a very difficult task!
+apparent when working with large lenses comprised of many sub-lenses (though for
+brevity we give here the small example above).  Lenses provide great power with
+their invertibility guarantees, but they come at the cost of thinking about many
+fiddly details when writing the terms.  Worrying about these fiddly details
+while ALSO thinking about unambiguity restrictions is a very difficult task!
 
@@
 
-Boomerang doesn't currently provide such capabilities.  However, there is no
-technical reason this can't be done by adding in a reverse-and-iterate
-operator to Boomerang.  Only a slight adjustment to the synthesis algorithm
-would be needed to incorporate this into the algorithm.
+The Boomerang lenses that we synthesize don't currently provide such
+capabilities.  However, there is no technical reason this can't be done by
+adding in a reverse-and-iterate operator to Boomerang.  Only slight adjustments
+to the synthesis algorithm would be needed.
 
@@
 
-We have not manually written lenses for all our benchmarks.  We find the sizes
-are fairly comparable, but the hand-written lenses are more modular: with common
-portions taken out and given good names.
+We have not manually written lenses for all our benchmarks.  For the ones we
+have, we find that the sizes are fairly comparable, but the hand-written lenses
+are more modular: with common portions factored out and given good names.
 
@@
 lenses that our system could handle, we began turning the lenses into synthesis
-problems alphabetically - first the As, then the Bs.
+problems alphabetically - first the As, then the Bs, etc.
 
@@
 
-See the general response.
+Yes, see the general response.
 
@@
 
-In particular, the useless concatenations of const("","") are removed and
-the field boilerplate is removed from being on both sides of the disjunct.
-Furthermore, when const(s,s) is present, it is converted into Id(s).
+In particular, the useless concatenations of const("","") are removed and the
+field boilerplate is shared rather than being duplicated in the two cases of the
+disjunct.  Furthermore, when const(s,s) is present, it is converted into Id(s).
 
@@
 the view update problem, POPL 2005; Boomerang: Resourceful Lenses for String
-Data, POPL 2008; Quotient Lenses, ICFP 2008].  In the preliminaries, we define
-which subset of Boomerang we aim to synthesize. While this language does not
-appear verbatim in a prior paper, it is a natural core language for string
-lenses.
+Data, POPL 2008; Quotient Lenses, ICFP 2008]. While this specific sublanguage
+does not appear verbatim in a prior paper, it is a natural core language for
+string lenses.
 

### 2017-09-14 papers/BijectiveLensSynth/popl_18_rebuttal.txt
another couple tweaks
move engineering best practices into individual comments

@@
-Thank you for your comments!
+Thanks for your comments!
 
-We begin with brief responses to the most significant concerns, followed by
-a (long and very optional) appendix addressing all questions in detail.
+We begin with brief responses to the most significant concerns, followed by a
+(long and optional) appendix addressing all questions in detail.
 
@@
 existing synthesis systems cannot.  In particular, techniques that only use
-input-output examples are unable to synthesize programs with nested loops.
-We note that FlashFill's DSL does permit nested
-loops, but for efficiency reasons the algorithm only synthesizes
-programs with no nesting (this is discussed at the end of section 4.4 of
-"Automating String Processing in Spreadsheets Using Input-Output
-Examples", POPL '11).
+input-output examples are unable to synthesize programs with nested loops.  We
+note that FlashFill's DSL does permit nested loops, but for efficiency reasons
+the algorithm only synthesizes programs with no nesting (this is discussed at
+the end of section 4.4 of "Automating String Processing in Spreadsheets Using
+Input-Output Examples", POPL '11).
 
@@
 
-Engineering best practices dictate that we annotate the lenses with their
-types.  While Boomerang does not require these annotations - being able to
-infer the types from the terms - the types of the term serve as
-documentation for future programmers to understand what formats the lens
-maps between.  Furthermore, the types provide resilience in the face of future
-lens modifications, as the types ensure that the lens maps exactly
-between strings of the provided formats. Invalid inputs are not accepted
-nor are the outputs ever invalid.  Consequently, we are not requiring
-additional work. Instead, we are saving programmers from having to 
-writing the lens by leveraging work they should already be performing.
-
-A few more reasons why programming with regular expressions is easier
-than programming with lenses:
+While we still require users to write regular expressions, programming
+regular expressions is easier than programming lenses:
 
@@
 
-* In the (common) case where the two formats have kleene stars in different
+* In the (common) case where the two formats have Kleene stars in different
   places, manual lens programming requires mentally transforming both to a
@@
 
+Engineering best practices dictate that we annotate the lenses with their
+types.  While Boomerang does not need these annotations - being able to
+infer the types from the terms - the types of the term serve as
+documentation for future programmers to understand what formats the lens
+maps between.  Furthermore, it provides an additional resilience to future
+modifications of the lens, as it validates that the lens maps exactly
+between strings of the provided formats - invalid inputs are not accepted
+nor are the outputs ever invalid.  In this way, we are not requiring
+additional work by the programmer to use our system; we are allowing the
+programmer to save the difficulties of writing the term by using work they
+should already be performing.
+
 We can make sure we can provide at a description of the sizes of specs and
@@
 
+Engineering best practices dictate that we annotate the lenses with their
+types.  While Boomerang does not need these annotations - being able to
+infer the types from the terms - the types of the term serve as
+documentation for future programmers to understand what formats the lens
+maps between.  Furthermore, it provides an additional resilience to future
+modifications of the lens, as it validates that the lens maps exactly
+between strings of the provided formats - invalid inputs are not accepted
+nor are the outputs ever invalid.  In this way, we are not requiring
+additional work by the programmer to use our system; we are allowing the
+programmer to save the difficulties of writing the term by using work they
+should already be performing.
+
 We will make the problem statement more clear in the final version.

### 2017-09-14 program/generate-data.py

@@ -83,39 +83,44 @@ def gather_data(rootlength, prog, path, base):
 	    return int(sum(example_number_col)/len(example_number_col))
 
 
-    gather_col(['-time'],ctime_combiner,"ComputationTime",TIMEOUT_TIME,REPETITION_COUNT)
-    #gather_col(['-forceexpand','-time'],ctime_combiner,"ForceExpandTime",TIMEOUT_TIME,REPETITION_COUNT)
-    gather_col(['-naive_strategy','-time'],ctime_combiner,"NaiveStrategy",TIMEOUT_TIME,REPETITION_COUNT)
-    #gather_col(['-naive_pqueue','-time'],ctime_combiner,"NaivePQueue",TIMEOUT_TIME,REPETITION_COUNT)
-    #gather_col(['-no_short_circuit','-time'],ctime_combiner,"NoShortCircuit",TIMEOUT_TIME,REPETITION_COUNT)
-    gather_col(['-no_lens_context','-time'],ctime_combiner,"NoLensContext",TIMEOUT_TIME,REPETITION_COUNT)
-    #gather_col(['-no_short_circuit','-no_inferred_expansions','-no_lens_context','-time'],ctime_combiner,"NoInferenceNoLCNoSC",TIMEOUT_TIME,REPETITION_COUNT)
-    #gather_col(['-no_short_circuit','-no_lens_context','-time'],ctime_combiner,"NoLCNoSC",TIMEOUT_TIME,REPETITION_COUNT)
-    gather_col(['-naive_expansion_search','-no_lens_context','-time'],ctime_combiner,"NaiveExpansionNoLC",TIMEOUT_TIME,REPETITION_COUNT)
-    gather_col(['-use_only_forced_expansions','-no_lens_context','-time'],ctime_combiner,"OnlyForcedExpansionsNoLC",TIMEOUT_TIME,REPETITION_COUNT)
-    gather_col(['-naive_expansion_search','-time'],ctime_combiner,"NaiveExpansion",TIMEOUT_TIME,REPETITION_COUNT)
-    gather_col(['-use_only_forced_expansions','-time'],ctime_combiner,"OnlyForcedExpansions",TIMEOUT_TIME,REPETITION_COUNT)
-    gather_col(['-forceexpand','-naive_expansion_search','-time'],ctime_combiner,"NoUDTypes",TIMEOUT_TIME,REPETITION_COUNT)
[…]

### 2017-09-14 papers/BijectiveLensSynth/popl_18_rebuttal.txt
polishing and shortening

--- 2017-09-14 papers/BijectiveLensSynth/popl_18_rebuttal.txt
@@
 many more complex formats and transformations.  In particular, these formats
-tend to have deep layers of iterated data, with complex transformations at each
-layer, requiring the synthesis of many layers of nested loops.  FlashFill cannot
-handle such examples.  Our experiments demonstrate this fact by showing that
-FlashFill only succeeds on 3 out of our 39 benchmarks, whereas our system
-succeeds on all our benchmarks.
+tend to have deep layers of iterated data, requiring the synthesis of many
+layers of nested loops.  FlashFill cannot handle such examples.  Our experiments
+demonstrate that FlashFill only succeeds on 3 out of our 39 benchmarks, whereas
+our system succeeds on all our benchmarks.
 
@@
 
-We changed some formats in our benchmark suite to handle two forms of non-bijectivity:
+We changed some formats in our benchmark suite to handle two forms of
+non-bijectivity:
 1) Whitespace present in one format but not the other.
@@
 
-Category 1 was by far the most common, applying to many of the Augeas
-examples.  To handle it, we added additional whitespace to the other format,
-where typically whitespace is unnecessary.
+Category 1 was by far the most common, applying to many Augeas examples.  To
+address this category, we added additional whitespace to the other format, where
+typically whitespace is unnecessary.
 
 Category 2 was less prevalent, but occurred, typically in the FlashFill
-examples.  To address this category, we added the projected information to the end
-of the format missing the information.
+examples.  To address this category, we added the projected information to the
+end of the format missing the information.
 
@@
 
-Reviewer 1 asked us directly about the potential for future work.
-Along those lines, we have already implemented a system that
-synthesizes quotient lenses from regular expressions augmented with equivalence
-relations that solve the whitespace bijection problem.  This system
-uses the algorithm provided in this paper as a drop-in component without change.
-We conjecture the system could be augmented in a similar way with projections as
-well.  We are not suggesting adding more results of this form to the current paper
-because it is already very dense.
+Reviewer 1 asked us directly about the potential for future work. Along those
+lines, we have already implemented a system that synthesizes quotient lenses
+from regular expressions augmented with equivalence relations to solve Category
+1 non-bijectivity.  This system uses the algorithm provided in this paper as a
+drop-in component without change. We conjecture the system could be augmented in
+a similar way with projections as well.  We are not suggesting adding more
+results of this form to the current paper because it is already very dense.
 
@@
 
+Engineering best practices dictate that we annotate the lenses with their
+types.  While Boomerang does not require these annotations - being able to
+infer the types from the terms - the types of the term serve as
+documentation for future programmers to understand what formats the lens
+maps between.  Furthermore, the types provide resilience in the face of future
+lens modifications, as the types ensure that the lens maps exactly
+between strings of the provided formats. Invalid inputs are not accepted
+nor are the outputs ever invalid.  Consequently, we are not requiring
+additional work. Instead, we are saving programmers from having to 
+writing the lens by leveraging work they should already be performing.
+
 We can make sure we can provide at a description of the sizes of specs and
@@
 
+Engineering best practices dictate that we annotate the lenses with their
+types.  While Boomerang does not require these annotations - being able to
+infer the types from the terms - the types of the term serve as
+documentation for future programmers to understand what formats the lens
+maps between.  Furthermore, the types provide resilience in the face of future
+lens modifications, as the types ensure that the lens maps exactly
+between strings of the provided formats. Invalid inputs are not accepted
+nor are the outputs ever invalid.  Consequently, we are not requiring
+additional work. Instead, we are saving programmers from having to 
+writing the lens by leveraging work they should already be performing.
+
 We will make the problem statement more clear in the final version.

### 2017-09-14 papers/BijectiveLensSynth/popl_18_rebuttal.txt
really minimize

@@
 
-Conversely, our "Optician" algorithm can solve some classes of problems that
-existing synthesis systems cannot.  In particular, techniques that only use
-input-output examples are unable to synthesize programs with nested loops.  We
-note that FlashFill's DSL does permit nested loops, but for efficiency reasons
-the algorithm only synthesizes programs with no nesting (this is discussed at
-the end of section 4.4 of "Automating String Processing in Spreadsheets Using
-Input-Output Examples", POPL '11).
+Conversely, Optician can solve classes of problems that existing synthesis
+systems cannot.  In particular, techniques that only use input-output examples
+are currently unable to synthesize programs with nested loops. FlashFill's DSL
+does permit nested loops, but for efficiency reasons the algorithm only
+synthesizes programs with no nesting (see section 4.4 of "Automating String
+Processing in Spreadsheets Using Input-Output Examples", POPL '11).
 
-While a single nesting of loops seems to be sufficient for FlashFill's intended
-uses, our benchmark suite (based on real-world uses of the Augeas tool) includes
-many more complex formats and transformations.  In particular, these formats
-tend to have deep layers of iterated data, requiring the synthesis of many
-layers of nested loops.  FlashFill cannot handle such examples.  Our experiments
-demonstrate that FlashFill only succeeds on 3 out of our 39 benchmarks, whereas
-our system succeeds on all our benchmarks.
+While a single nesting of loops is sufficient for FlashFill's intended uses, our
+benchmark suite (based on real-world uses of the Augeas tool) includes many
+complex formats and transformations.  In particular, these formats have many
+layers of iterated data, requiring the synthesis of deeply nested loops.
+FlashFill cannot handle such examples; FlashFill only succeeds on 3 of our
+39 benchmarks, whereas our system succeeds on all 39 benchmarks.
 
@@
 
-We changed some formats in our benchmark suite to handle two forms of
+We modified some formats in our benchmark suite to handle two forms of
 non-bijectivity:
@@
 Category 1 was by far the most common, applying to many Augeas examples.  To
-address this category, we added additional whitespace to the other format, where
-typically whitespace is unnecessary.
+address this category, we added whitespace to the other format where typically
+whitespace is unnecessary.
 
@@
 examples.  To address this category, we added the projected information to the
-end of the format missing the information.
+end of the format missing that information.
 
@@
 
-There are lens theories developed to handle synchronization tasks for data
+Lens theories were developed to handle synchronization tasks for data
 formats not kept in bijective correspondence (quotient lenses to handle Category
@@
 
-Reviewer 1 asked us directly about the potential for future work. Along those
-lines, we have already implemented a system that synthesizes quotient lenses
-from regular expressions augmented with equivalence relations to solve Category
-1 non-bijectivity.  This system uses the algorithm provided in this paper as a
-drop-in component without change. We conjecture the system could be augmented in
-a similar way with projections as well.  We are not suggesting adding more
-results of this form to the current paper because it is already very dense.
+Reviewer 1 asked us about the potential for future work. We have already
+implemented a system that synthesizes quotient lenses from regular expressions
+augmented with equivalence relations to solve Category 1 non-bijectivity.  This
+system uses the algorithm provided in this paper as a drop-in component without
+change. We conjecture Optician could be augmented in a similar way with
+projections as well.  We are not suggesting adding more results of this form to
+the current paper because it is already very dense.
 
-     - After having written the regular expressions, is it very difficult
-       for users to then write the lens?  Are we saving anybody any
-       significant effort?
+     - After writing the regular expressions, is it very difficult for users to
+       then write the lens?  Are we saving anybody significant effort?
 
@@
 
-* A regular expression describes one data format, while lenses deal with two
+* A regular expression describes one data format, while a lens deals with two
   at the same time.

### 2017-09-14 papers/BijectiveLensSynth/popl_18_rebuttal.txt
600 words woo

@@
 We begin with brief responses to the most significant concerns, followed by a
-(long and optional) appendix addressing all questions in detail.
+(long and optional) appendix addressing individual questions.
 
@@
 Conversely, Optician can solve classes of problems that existing synthesis
-systems cannot.  In particular, techniques that only use input-output examples
-are currently unable to synthesize programs with nested loops. FlashFill's DSL
-does permit nested loops, but for efficiency reasons the algorithm only
-synthesizes programs with no nesting (see section 4.4 of "Automating String
-Processing in Spreadsheets Using Input-Output Examples", POPL '11).
+systems cannot.  Techniques that only use input-output examples are currently
+unable to synthesize programs with nested loops. FlashFill's DSL does permit
+nested loops, but for efficiency reasons the algorithm only synthesizes programs
+with no nesting ("Automating String Processing in Spreadsheets Using
+Input-Output Examples", section 4.4, POPL '11).
 
 While a single nesting of loops is sufficient for FlashFill's intended uses, our
-benchmark suite (based on real-world uses of the Augeas tool) includes many
+benchmark suite (based on real-world uses of the Augeas tool) includes highly
 complex formats and transformations.  In particular, these formats have many
@@
 Category 2 was less prevalent, but occurred, typically in the FlashFill
-examples.  To address this category, we added the projected information to the
+examples.  To address this category, we added projected information to the
 end of the format missing that information.
@@
 system can be used as a drop-in component for synthesis algorithms on these more
-complex lens structures.
-
-Reviewer 1 asked us about the potential for future work. We have already
-implemented a system that synthesizes quotient lenses from regular expressions
-augmented with equivalence relations to solve Category 1 non-bijectivity.  This
-system uses the algorithm provided in this paper as a drop-in component without
-change. We conjecture Optician could be augmented in a similar way with
-projections as well.  We are not suggesting adding more results of this form to
-the current paper because it is already very dense.
+complex lens structures.  Indeed, we have implemented a system that synthesizes
+quotient lenses from annotated regular expressions to solve Category 1
+non-bijectivity,  using the algorithm provided in this paper as a drop-in
+component without change.  We are not suggesting adding more results of this
+form to the current paper because it is already very dense.
 
@@
 
-While we still require users to write regular expressions, programming
-regular expressions is easier than programming lenses:
+While we require users to write regular expressions, programming regular
+expressions is easier than programming lenses:
 

### 2017-09-26 papers/QuotientLensSynth/listofitems.tex
Added quotient lens paper folder

% Ce fichier contient le code de l'extension "listofitems"
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%                                                                    %
\def\loiname                 {listofitems}                           %
\def\loiver                      {1.3}                              %
%                                                                    %
\def\loidate                  {2016/11/18}                           %
%                                                                    %
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% --------------------------------------------------------------------
% This work may be distributed and/or modified under the
% conditions of the LaTeX Project Public License, either version 1.3
% of this license or (at your option) any later version.
% The latest version of this license is in
%
% %     http://www.latex-project.org/lppl.txt
%
% and version 1.3 or later is part of all distributions of LaTeX
% version 2005/12/01 or later.
% --------------------------------------------------------------------
% This work has the LPPL maintenance status `maintained'.
%
% The Current Maintainer of this work is Christian Tellechea
% email: unbonpetit@openmailbox.org
%        Commentaires, suggestions et signalement de bugs bienvenus !
%        Comments, bug reports and suggestions are welcome.
% --------------------------------------------------------------------
% L'extension listofitems est compos\'ee des 7 fichiers suivants :
%   - code               : listofitems            (.tex et .sty)
%   - manuel en francais : listofitems-fr         (.tex et .pdf)
%   - manuel en anglais  : listofitems-en         (.tex et .pdf)
%   - fichier lisezmoi   : README
% --------------------------------------------------------------------
\expandafter\edef\csname loi_restorecatcode\endcsname{\catcode\number`\_=\number\catcode`\_\relax}
\catcode`\_11

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%% gestion des erreurs %%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\ifdefined\PackageError
	\def\loi_error#1{\PackageError\loiname{#1}{Read the manual}}% pour LaTeX
\else
	\def\loi_error#1{\errmessage{Package \loiname\space Error: #1^^J}}% pour TeX
\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%% v\'erification de la pr`'esence de etex %%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begingroup
\edef\__tempa{\meaning\eTeXversion}\edef\__tempb{\string\eTeXversion}%
\ifx\__tempa\__tempb
	\endgroup
\else
	\endgroup
	\loi_error{You are not using an eTeX engine, listofitems cannot work.}%
	\expandafter\loi_restorecatcode\expandafter\endinput
\fi

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%% macros auxiliaires %%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%---- macros de d\'eveloppement et de manipulation d'arguments
\chardef\loi_stop=0
\def\loi_quark{\loi_quark}
\long\def\loi_identity#1{#1}
\def\loi_gobarg#1{}
\long\def\loi_first#1#2{#1}
\long\def\loi_second#1#2{#2}
\long\def\loi_firsttonil#1#2\_nil{#1}
\long\def\loi_antefi#1#2\fi{#2\fi#1}
\long\def\loi_swaparg#1#2{#2{#1}}
\long\def\loi_exparg#1#2{\expandafter\loi_swaparg\expandafter{#2}{#1}}
\long\def\loi_expafternil#1\_nil{\begingroup\def\loi_temp{\endgroup#1}\expandafter\loi_temp}
\def\loi_macroname{\loi_ifinrange\escapechar[[0:255]]{\expandafter\loi_gobarg}{}\string}
\def\loi_argcsname#1#{\begingroup\def\loi_temp{\endgroup#1}\loi_argcsname_i}
\def\loi_argcsname_i#1{\expandafter\loi_temp\csname#1\endcsname}

%--- macros de test
\def\loi_ifnum#1{\ifnum#1\expandafter\loi_first\else\expandafter\loi_second\fi}
\long\def\loi_ifx#1{\ifx#1\expandafter\loi_first\else\expandafter\loi_second\fi}
\long\def\loi_ifempty#1{\loi_exparg\loi_ifx{\expandafter\relax\detokenize{#1}\relax}}
\def\loi_ifstar#1#2{\def\loi_ifstar_i{\loi_ifx{*\loi_nxttok}{\loi_first{#1}}{#2}}\futurelet\loi_nxttok\loi_ifstar_i}
\long\def\loi_ifcs#1{% #1 est-il une sc ?
	\loi_ifempty{#1}
		\loi_second% si #1 est vide, faux
		{\loi_ifspacefirst{#1}
			\loi_second% si espace en 1er, faux
			{\loi_exparg\loi_ifempty{\loi_gobarg#1}% 1 seul token ?
				{\begingroup \escapechar`\_
					\if\expandafter\loi_firsttonil\string#1\_nil\expandafter\loi_firsttonil\string\relax\_nil
						\endgroup\expandafter\loi_first
					\else
						\endgroup\expandafter\loi_second
					\fi
				}
				\loi_second% si plusieurs tokens, faux
			}%
		}%
}
\def\loi_ifinrange#1[[#2:#3]]{\loi_ifnum{\numexpr(#1-#2)*(#1-#3)>0 }\loi_second\loi_first}

%--- macro de type for
%      Voir codes 150 \`a 155 ici --> http://progtex.fr/wp-content/uploads/2014/09/code.txt
%      et pages 175 \`a 184 du livre "Apprendre \`a programmer en TeX"
\def\loi_fornum#1=#2to#3\do{%
	\edef#1{\number\numexpr#2}\edef\loi_sgncmp{\ifnum#1<\numexpr#3\relax>+\else<-\fi}%
	\loi_argcsname\loi_fornum_i{loi_fornum_\string#1\expandafter}\expandafter{\number\numexpr#3\expandafter}\loi_sgncmp#1%
}
\long\def\loi_fornum_i#1#2#3#4#5#6{\def#1{\unless\ifnum#5#3#2\relax\loi_antefi{#6\edef#5{\number\numexpr#5#41\relax}#1}\fi}#1}

%--- macros retirant les espaces extr\^emes
%      Voir codes 320 \`a 324 ici --> http://progtex.fr/wp-content/uploads/2014/09/code.txt
%      et pages 339 \`a 343 de "Apprendre \`a programmer en TeX"
\long\def\loi_ifspacefirst#1{\expandafter\loi_ifspacefirst_i\detokenize{#10} \_nil}
\long\def\loi_ifspacefirst_i#1 #2\_nil{\loi_ifempty{#1}}
\expandafter\def\expandafter\loi_gobspace\space{}
\def\loi_removefirstspaces{\romannumeral\loi_removefirstspaces_i}
\long\def\loi_removefirstspaces_i#1{\loi_ifspacefirst{#1}{\expandafter\loi_removefirstspaces_i\expandafter{\loi_gobspace#1}}{\loi_stop#1}}
\edef\loi_restorezerocatcode{\catcode0=\number\catcode0 \relax}
\catcode0 12
\long\def\loi_removelastspaces#1{\romannumeral\loi_removelastspaces_i#1^^00 ^^00\_nil}
\long\def\loi_removelastspaces_i#1 ^^00{\loi_removelastspaces_ii#1^^00}
\long\def\loi_removelastspaces_ii#1^^00#2\_nil{\loi_ifspacefirst{#2}{\loi_removelastspaces_i#1^^00 ^^00\_nil}{\loi_stop#1}}
\loi_restorezerocatcode
\long\def\loi_removeextremespaces#1{% #1=texte o\`u les espaces extr\^emes sont retir\'es
	\romannumeral\expandafter\expandafter\expandafter\loi_removelastspaces\expandafter\expandafter\expandafter
	{\expandafter\expandafter\expandafter\loi_stop\loi_removefirstspaces{#1}}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%% macro publique \setsepchar %%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\setsepchar{\futurelet\loi_nxttok\setsepchar_i}
\def\setsepchar_i{\loi_ifx{[\loi_nxttok}\setsepchar_ii{\setsepchar_ii[/]}}
\long\def\setsepchar_ii[#1]#2{% #1=sepcar de <liste des sepcar>  #2=<liste des sepcar>
	\loi_ifempty{#1}
		{\loi_error{Empty separator not allowed, separator "/" used}%
		\setsepchar_ii[/]{#2}%
		}
		{\def\loi_currentsep{#1}%
		\_removeextremespacesfalse
		\loi_nestcnt1 % r\'einitaliser niveau initial \`a 1
		\def\nestdepth{1}%
		\loi_argcsname\let{loi_previndex[\number\loi_nestcnt]}\empty
		\def\loi_listname{loi_listofsep}%
		\let\loi_def\def \let\loi_edef\edef \let\loi_let\let
		\loi_ifempty{#2}
			{\loi_error{Empty list of separators not allowed, "," used}%
			\readlist_iv1{,}%
			}
			{\readlist_iv1{#2}}%
		\loi_argcsname\let\nestdepth{loi_listofseplen[0]}%
		\loi_argcsname\let\loi_currentsep{loi_listofsep[1]}% 1er car de s\'eparation
		}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%% macro normalisant l'index %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\loi_normalizeindex#1#2#3{% #1=macroname  #2=liste d'index  #3=profondeur max --> renvoie {err}{indx norm}
	\loi_ifempty{#2}
		{\loi_stop{}{}}
		{\loi_normalizeindex_i1{}{#3}{#1}#2,\loi_quark,}%
}%
\def\loi_normalizeindex_i#1#2#3#4#5,{% #1=compteur de profondeur #2=index pr\'ec\'edents  #3=profondeur max #4=macroname  #5=index courant
	\loi_ifx{\loi_quark#5}
		{\loi_normalizeindex_iii#2\loi_quark}% supprimer la derni\`ere virgule
		{\loi_ifnum{#1>#3 }
			{\loi_invalidindex{Too deeply nested index, index [.] retained}{#2}}% si profondeur trop grande
			{\loi_ifinrange\ifnum\numexpr#5<0 -1*\fi(#5)[[1:\csname #4len[#20]\endcsname]]% si abs(#5) hors de [1,len]
				{\loi_exparg\loi_normalizeindex_ii{\number\numexpr#5\ifnum\numexpr#5<0 +\csname #4len[#20]\endcsname+1\fi}{#1}{#2}{#3}{#4}}
				{\loi_invalidindex{#5 is an invalid index, index [.] retained}{#2}}%
			}%
		}%
}
\def\loi_normalizeindex_ii#1#2#3{\loi_exparg\loi_normalizeindex_i{\number\numexpr#2+1}{#3#1,}}% #1=index \`a rajouter  #2=compteur de profondeur #3=index pr\'ec\'edents
\def\loi_normalizeindex_iii#1,\loi_quark{\loi_stop{}{#1}}
\def\loi_invalidindex#1#2{\loi_ifempty{#2}{\loi_invalidindex_i{#1},}\loi_invalidindex_i{#1}{#2}}
\def\loi_invalidindex_i#1#2{\loi_invalidindex_ii#1\loi_quark#2\loi_quark}
\def\loi_invalidindex_ii#1[.]#2\loi_quark#3,\loi_quark#4\loi_quark,{\loi_stop{#1[#3]#2}{#3}}% #4= index ignor\'es

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%% macro publique \readlist %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newcount\loi_nestcnt
\def\greadlist{%
	\let\loi_def\gdef \let\loi_edef\xdef \def\loi_let{\global\let}%
	\loi_nestcnt1 % niveau initial = 1
	\loi_argcsname\loi_let{loi_previndex[\number\loi_nestcnt]}\empty
	\loi_ifstar{\_removeextremespacestrue\readlist_i}{\_removeextremespacesfalse\readlist_i}%
}
\def\readlist{%
	\let\loi_def\def \let\loi_edef\edef \let\loi_let\let
	\loi_nestcnt1 % niveau initial = 1
	\loi_argcsname\let{loi_previndex[\number\loi_nestcnt]}\empty
	\loi_ifstar{\_removeextremespacestrue\readlist_i}{\_removeextremespacesfalse\readlist_i}%
}
\long\def\readlist_i#1#2{% #1=macro stockant les \'el\'ements  #2=liste des \'el\'ements
	\loi_ifcs{#2}
		{\loi_exparg{\readlist_i#1}{#2}}
		{\loi_edef\loi_listname{\loi_macroname#1}%
		\loi_argcsname\loi_let{\loi_listname nest}\nestdepth
		\loi_argcsname\loi_def{\loi_listname[]}{#2}% la liste enti\`ere
		\loi_argcsname\loi_def{\loi_listname sep[]}{}% s\'eparateur vide
		\loi_ifempty{#2}
			{\loi_def#1[##1]{}%
			\loi_argcsname\loi_def{\loi_listname len}{0}\loi_argcsname\loi_def{\loi_listname len[0]}{0}%
			\loi_error{Empty list ignored, nothing to do}%
			}
			{\loi_edef#1[##1]{\unexpanded{\romannumeral\expandafter\loi_auxmacrolistitem\romannumeral\loi_normalizeindex}{\loi_listname}{##1}{\csname\loi_listname nest\endcsname}{\loi_listname}}%
			\loi_argcsname\loi_edef{\loi_listname sep}[##1]{\unexpanded{\romannumeral\expandafter\loi_auxmacrolistitem\romannumeral\loi_normalizeindex}{\loi_listname}{##1}{\csname\loi_listname nest\endcsname}{\loi_listname sep}}%
			\readlist_ii{#2}%
			\loi_argcsname\loi_argcsname\loi_let{\loi_listname len}{\loi_listname len[0]}% longueur du niveau 0
			}%
		}%
}
\def\loi_auxmacrolistitem#1#2#3{%
	\expandafter\expandafter\expandafter\loi_stop\csname#3[#2]\expandafter\endcsname
	\romannumeral\loi_ifempty{#1}{\loi_stop}{\loi_stop\loi_error{#1}}%
}
\def\readlist_ii{%
	\loi_argcsname\loi_let\loi_currentsep{loi_listofsep[\number\loi_nestcnt]}%
	\expandafter\readlist_iii\loi_currentsep||\_nil
}
\long\def\readlist_iii#1||#2\_nil#3{\readlist_iv1{#3#1}}% #1=<sep courant simple>  #3=liste -> rajoute un \'el\'ement vide pour le test ifempty ci dessous
\long\def\readlist_iv#1#2{% #1=compteur d'index  #2=liste d'\'el\'ements \`a examiner
	\loi_ifempty{#2}
		{\loi_argcsname\loi_edef{\loi_listname len[\csname loi_previndex[\number\loi_nestcnt]\endcsname0]}{\number\numexpr#1-1\relax}%
		\advance\loi_nestcnt-1
		\loi_argcsname\loi_let\loi_currentsep{loi_listofsep[\number\loi_nestcnt]}%
		}
		{\loi_expafternil\readlist_vi{#2}{}\_nil\loi_currentsep||\loi_quark||#2\_nil{#1}}% aller isoler le 1er item
}
\long\def\readlist_v#1#2{\readlist_vi{#2}{}#1||\loi_quark||#2\_nil}% #1=liste s\'eparateurs (s\'ep=||)  #2=chaine de tokens
\long\def\readlist_vi#1#2#3||{% #1=liste restante   #2=dernier <sep utile>  #3=<sep courant>
	\loi_ifx{\loi_quark#3}
		{\loi_ifempty{#2}% si #1 vide, aucun <sep utile> n'a \'et\'e trouv\'e, il reste \`a lire "<liste compl\`ete>\_nil"
			{\long\def\readlist_vii##1\_nil##2{\loi_exparg{\readlist_ix{##2}{}}{\loi_gobarg##1}{#2}}}% ##2=compteur d'index
			{\long\def\readlist_vii##1#2{\loi_exparg\readlist_viii{\loi_gobarg##1}\relax}%
			\long\def\readlist_viii##1##2\_nil##3{\loi_exparg{\readlist_ix{##3}}{\loi_gobarg##2}{##1}{#2}}% ##3=compteur d'index
			}%
		\readlist_vii\relax% le \relax meuble l'argument d\'elimit\'e
		}
		{\long\def\readlist_vii##1#3##2\_nil{%
			\loi_ifempty{##2}% si <liste restante> ne contient pas le <sep courant>
				{\readlist_vi{#1}{#2}}% recommencer avec le m\^eme <sep utile>
				{\readlist_vi{##1#3}{#3}}% sinon raccourcir <liste restante> et <sep courant>:=<sep utile>
			}%
		\readlist_vii#1#3\_nil
		}%
}
\long\def\readlist_ix#1#2#3{% #1=compteur d'index  #2=liste restante  #3=\'el\'ement courant
	\loi_ifnum{0\loi_exparg\loi_ifspacefirst{\loi_currentsep}{}1\if_removeextremespaces1\fi=11 }% s'il faur retirer les espaces extr\^emes
		{\loi_exparg{\loi_exparg{\readlist_x{#1}{#2}}}{\loi_removeextremespaces{#3}}}% red\'efinir l'\'el\'ement courant
		{\readlist_x{#1}{#2}{#3}}%
}
\long\def\readlist_x#1#2#3#4{% #1=compteur d'index  #2=liste restante  #3=\'el\'ement courant   #4=sep utilis\'e
	\loi_ifnum{0\if_ignoreemptyitems1\fi\loi_ifempty{#3}1{}=11 }
		{\readlist_iv{#1}{#2}}% si l'on n'ignore pas les \'el\'ements vides :
		{\loi_argcsname\loi_def{\loi_listname[\csname loi_previndex[\number\loi_nestcnt]\endcsname#1]}{#3}% assignation de l'item ctuel \`a la macro
		\loi_argcsname\loi_def{\loi_listname sep[\csname loi_previndex[\number\loi_nestcnt]\endcsname#1]}{#4}% assignation du <sep> actuel \`a la macro \<macrolist>sep
		\loi_ifnum{\loi_nestcnt<\nestdepth\relax}% si imbrication max non atteinte
			{\advance\loi_nestcnt1
			\loi_argcsname\edef{loi_previndex[\number\loi_nestcnt]}{\csname loi_previndex[\number\numexpr\loi_nestcnt-1]\endcsname#1,}%
			\readlist_ii{#3}% recommencer avec l'\'el\'ement courant
			}
			{}%
		\loi_exparg\readlist_iv{\number\numexpr#1+1}{#2}% puis chercher l'\'el\'ement suivant dans la liste restante
		}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%% macro \listlen %%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\listlen#1[#2]{%
	\romannumeral\loi_ifempty{#2}
		{\expandafter\expandafter\expandafter\loi_stop\csname\loi_macroname#1len[0]\endcsname}
		{\loi_exparg\listlen_i{\romannumeral-`\.\loi_macroname#1}{#2}}%
}
\def\listlen_i#1#2{% #1=macro name   #2=index non normalis\'e    prendre <profondeur max-1>
	\loi_exparg{\expandafter\listlen_ii\romannumeral\loi_normalizeindex{#1}{#2}}{\number\numexpr\csname#1nest\endcsname-1}{#1}%
}
\def\listlen_ii#1#2#3{% #1=err   #2=index normalis\'e  #3=macroname
	\expandafter\expandafter\expandafter\loi_stop\csname#3len[#2,0]\expandafter\endcsname
	\romannumeral\loi_ifempty{#1}{\loi_stop}{\loi_stop\loi_error{#1}}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%% macro \foreachitem %%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\foreachitem#1\in#2{%
	\edef\foreachitem_i{\noexpand\foreachitem_iii\noexpand#1{\expandafter\noexpand\csname\loi_macroname#1cnt\endcsname}{\loi_macroname#2}}%
	\futurelet\loi_nxttok\foreachitem_ii
}
\def\foreachitem_ii{\loi_ifx{\loi_nxttok[}\foreachitem_i{\foreachitem_i[]}}
\def\foreachitem_iii#1#2#3[#4]{% prendre <profondeur max-1>
	\loi_exparg{\expandafter\foreachitem_iv\romannumeral\loi_normalizeindex{#3}{#4}}{\number\numexpr\csname#3nest\endcsname-1}#1{#2}{#3}%
}
\def\foreachitem_iv#1#2{\loi_ifempty{#2}{\foreachitem_v{#1}{}}{\foreachitem_v{#1}{#2,}}}% #1=err  #2=index norm
\long\def\foreachitem_v#1#2#3#4#5#6{% #1=err  #2=index norm  #3=macroiter  #4=compteur associ\'e  #5=nom de macrolist  #6=code
	\loi_ifnum{\csname#5len[#20]\endcsname>0 }
		{\loi_ifempty{#1}{}{\loi_error{#1}}%
		\loi_fornum#4=1to\csname#5len[#20]\endcsname\do{\loi_argcsname\let#3{#5[#2#4]}#6}%
		}
		{}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%% macro \showitem %%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\showitems{\loi_ifstar{\let\showitems_cmd\detokenize\showitems_i}{\let\showitems_cmd\loi_identity\showitems_i}}
\def\showitems_i#1{\def\showitems_ii{\showitems_iv#1}\futurelet\loi_nxttok\showitems_iii}
\def\showitems_iii{\loi_ifx{\loi_nxttok[}\showitems_ii{\showitems_ii[]}}
\def\showitems_iv#1[#2]{\foreachitem\showitems_iter\in#1[#2]{\showitemsmacro{\expandafter\showitems_cmd\expandafter{\showitems_iter}}}}
\unless\ifdefined\fbox
	\newdimen\fboxrule \newdimen\fboxsep \fboxrule=.4pt \fboxsep=3pt % r\'eglages identiques \`a LaTeX
	\def\fbox#1{% imitation de la macro \fbox de LaTeX, voir codes 251 \`a 254 ici --> http://progtex.fr/wp-content/uploads/2014/09/code.txt
		\hbox{% et pages 271 \`a 274 de "Apprendre \`a programmer en TeX"
			\vrule width\fboxrule
			\vtop{%
				\vbox{\hrule height\fboxrule \kern\fboxsep \hbox{\kern\fboxsep#1\kern\fboxsep}}%
				\kern\fboxsep \hrule height\fboxrule
			}\vrule width\fboxrule
		}%
	}
\fi
\def\showitemsmacro#1{% encadrement par d\'efaut
	\begingroup\fboxsep=0.25pt \fboxrule=0.5pt \fbox{\strut#1}\endgroup
	\hskip0.25em\relax
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%% macro \itemtomacro %%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\def\itemtomacro#1[#2]{% #1[#2]=item  non encore lu: #3=macro
	\edef\loi_listname{\loi_macroname#1}%
	\loi_exparg{\expandafter\itemtomacro_i\romannumeral\expandafter\loi_normalizeindex\expandafter{\loi_listname}{#2}}{\csname\loi_listname nest\endcsname}\let
}
\def\gitemtomacro#1[#2]{% #1[#2]=item
	\xdef\loi_listname{\loi_macroname#1}%
	\loi_exparg{\expandafter\itemtomacro_i\romannumeral\expandafter\loi_normalizeindex\expandafter{\loi_listname}{#2}}{\csname\loi_listname nest\endcsname}{\global\let}%
}
\def\itemtomacro_i#1#2#3#4{%
	\loi_ifempty{#1}{}{\loi_error{#1}}%
	\loi_argcsname#3#4{\loi_listname[#2]}%
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%% r\'eglages par d\'efaut %%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\if_removeextremespaces
\newif\if_ignoreemptyitems
\let\ignoreemptyitems\_ignoreemptyitemstrue
\let\reademptyitems\_ignoreemptyitemsfalse
\setsepchar{,}
\reademptyitems

\loi_restorecatcode
\endinput

######################################################################
#                             Historique                             #
######################################################################

v1.0  19/8/2016
      - Premi\`ere version publique

v1.1  01/09/2016
      - Stockage des s\'eparateurs dans <macrolist>sep
      - bug corrig\'e dans \loi_restorecatcode

v1.2  22/10/2016
      - macros \greadlist et \gitemtomacro pour la globalit\'e

v1.3  18/11/2016
      - bugs corrig\'es dans la gestion de la globalit\'e

### 2017-09-26 papers/QuotientLensSynth/main.tex
Added quotient lens paper folder

\documentclass{svproc}

\usepackage{amsmath, amssymb, verbatim, enumerate, 
graphicx, centernot, tikz, array, tikz-cd, extarrows, cleveref,
mathrsfs, mathtools, bussproofs, stmaryrd, enumitem, stackengine}

\begin{document}
\mainmatter              % start of a contribution
%
\title{A Synthesis Ready Langauge for Expressing Quotient Lenses}
%
\titlerunning{}  % abbreviated title (for running head)
%                                     also used for the TOC unless
%                                     \toctitle is used
%
\author{}
%
\authorrunning{} % abbreviated author list (for running head)
%
%%%% list of authors for the TOC (use if author list has to be modified)
\tocauthor{}
%
\institute{}

\maketitle              % typeset the title of the contribution

\begin{abstract}

\keywords{quotient lenses, quotient regular expressions, synthesis}
\end{abstract}

\section{Introduction}
TODO Introduce BibTex to Endnote example as motivation.
\subsection{Contributions}
\begin{enumerate}
  \item
  We introduce the language of Quotient Regular Expressions (QREs) which enables
  us to express a set of equivalence relations on regular languages using a
  compact, easy-to-use syntax.
  \item
  We describe a class $\mathcal{Q}$ of quotient lenses whose types are given by
  QREs. We derive a normal form for quotient lenses in $\mathcal{Q}$, and
  demonstrate that elements of $\mathcal{Q}$ that are in this normal form are
  closed under left quotienting, right quotienting, composition and regular
  operators.
  \item
  We give an algorithm for synthesizing quotient lenses in
  $\mathcal{Q}$ from their QRE types and a set of examples.
  \item
  We describe an implementation of QREs, the language $\mathcal{Q}$ of quotient
  lenses, and the algorithm for synthesizing quotient lenses in $\mathcal{Q}$ in
  the Boomerang language.
\end{enumerate}

\section{Quotient Regular Expressions}
The language of Quotient Regular Expressions (QREs) is given by the following
grammar:
$$q := id(R) \; | \; R \mapsto s \; | \; squash (R, R', f) \; | \; perm \;
(R_1, \ldots, R_n) \; with \; R \; | \; q' \circ q \; | \; q \cdot q' \; | \;
(q \; | \; q') \; | \; q^*,$$
where $R$ ranges over regular expressions, $f$ ranges over functions whose
domain and codomain are regular languages, and $s$ ranges over character
strings.\\
Each QRE $q$ enables us to express
\begin{enumerate}
  \item a regular expression $W(q)$ (the ``whole'' of $q$),
  \item an equivalence relation $EqRel(q)$ on $\mathcal{L}(W(q))$,
  \item a regular expression $K(q)$ (the ``kernel'' of $q$)
  such that $\mathcal{L}(K(q))$ forms a complete set of representatives for
  $EqRel(q)$, and
  \item a ``canonizing'' function $Canonizer(q):\mathcal{L}(W(q))
  \longrightarrow \mathcal{L}(K(q))$ which given any $w \in \mathcal{L}(W(q))$,
  computes $Canonizer(q)(w)$ as the unique $k$ in $\mathcal{L}(K(q))$ such that
  $k$ is equivalent to $w$ mod $EqRel(q)$
\end{enumerate}
Observe that since $Canonizer(q)$ computes
$Canonizer(q)(w)$ as the unique $k$ in $\mathcal{L}(K(q))$ such that $w$ is
equivalent to $k$ mod $EqRel(q)$, then the equivalence classes of $EqRel(q)$
are the same as the fibres of $Canonizer(q)$.
\subsection{Semantics of QREs}
We now give the semantics of QREs as well as inference rules for deriving them:
\begin{enumerate}
  \item $q = id(R)$
\begin{center}
\begin{enumerate}
  \item $W(id(R)) = R$,
  \item $K(id(R)) = R$, 
  \item
For all $w, w' \in W(id(R))$, $w \; EqRel(q) \; w'$ if and only if $w =
w'$.
\item
$Canonizer(id(R))(r) = r$
\end{enumerate}
\begin{prooftree}
\AxiomC{}
\UnaryInfC{$id(R)$ is well formed}
\end{prooftree}
\end{center}
\item $q=R \mapsto s$
\begin{center}
\begin{enumerate}
  \item $W(R \mapsto S) = R$,
  \item $K(R \mapsto s) = s$, 
  \item
For all $w, w' \in W(R \mapsto S)$, $w \; EqRel(R \mapsto S) \; w'$,
\item
$Canonizer(R \mapsto S)(r) = s$
\end{enumerate}
\begin{prooftree}
\AxiomC{$s \in \mathcal{L}(R)$}
\UnaryInfC{$R \mapsto s$ is well formed}
\end{prooftree}
\end{center}
\item $q=squash(R, R', f)$
\begin{center}
\begin{enumerate}
  \item $W(squash(R, R', f)) = R \; | \; R'$,
  \item $K(squash(R, R', f)) = R'$, 
  \item
For all $w, w' \in W(squash(R, R', f))$, $w \; EqRel(q) \; w'$ if and only
if $f(w) = w'$ or $w = w'$
\item
$Canonizer(squash(R, R', f))(w) = 
\begin{cases}
f(w) & \text{if } w \in \mathcal{L}(R)\\
w & \text{otherwise}
\end{cases}$
\end{enumerate}
\begin{prooftree}
\AxiomC{$\mathcal{L}(R) \cap \mathcal{L}(R') = \varnothing$}
\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(R')$}
\BinaryInfC{$squash(R, R', f)$ is well formed}
\end{prooftree}
\end{center}
\item $q=perm \; (R_1, \ldots, R_n) \; with \; S$
\begin{center}
\begin{enumerate}
  \item $W(perm \; (R_1, \ldots, R_n) \; with \; S) = \bigcup \limits_{\sigma
  \in S_n} R_{\sigma(1)} \cdot S \cdot \ldots \cdot \cdot S \cdot R_{\sigma(n)}$
,
  \item $K(perm \; (R_1, \ldots, R_n) \; with \; S) = R_{1} \cdot S
  \cdot \ldots \cdot \cdot S \cdot R_{n}$, 
  \item
For all $w, w' \in W(q)$,  $w \; EqRel(q) \; w'$ if and only if
   $$w = r_{\sigma(1)} \cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot
    r_{\sigma(n)} \text{ and } r' = {r}_{\theta(1)} \cdot s_1 \cdot \ldots
    \cdot s_{n-1} \cdot {r}_{\theta(n)}$$ for some $\sigma, \theta \in
    S_n$, where $r_i \in R_i$ and $s_j \in S$.
\item
$Canonize(q)(r_{\sigma(1)}
\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)})
= r_1 \cdot s_1 \cdot \ldots \cdot \cdot s_{n-1}\cdot r_n$
\end{enumerate}
\begin{prooftree}
\AxiomC{$\forall \sigma \neq \theta, \; R_{\sigma(1)} \cdot S \cdot \ldots \cdot
S \cdot R_{\sigma(n)} \cap R_{\theta(1)} \cdot S \cdot \ldots \cdot S \cdot
R_{\theta(n)} = \varnothing$} \AxiomC{$\forall \sigma, \; R_{\sigma(1)} \cdot^!
S \cdot^! \ldots \cdot^! S \cdot^! R_{\sigma(n)}$}
\BinaryInfC{$permute \; (R_1, \ldots, R_n) \; with \;  S$ is well formed}
\end{prooftree}
\end{center}
\item $q=q' \circ q$
\begin{center}
\begin{enumerate}
  \item $W(q' \circ q) = W(q)$,
  \item $K(q' \circ q) = K(q')$, 
  \item
For all $w, w' \in W(squash(R, R', f))$, $w \; EqRel(q' \circ q) \; w'$ if and
only $$\text{there exist }k, k' \in
  \mathcal{L}(K(q)) \text{ such that } w \; EqRel(q) \; k, \; w' \;
  EqRel(q) \; k', \text{ and } k \; EqRel(q') \; k' $$
\item
$Canonizer(q' \circ q) = Canonizer(q') \circ Canonizer(q)
$
\end{enumerate}
\begin{prooftree}
\AxiomC{$q$ is well formed}
\AxiomC{$q'$ is well formed}
\AxiomC{$K(q) = W(q')$}
\TrinaryInfC{$q' \circ q$ is well formed}
\end{prooftree}
\end{center}
\item $q=q \cdot q'$
\begin{center}
\begin{enumerate}
  \item $W(q \cdot q') = W(q) \cdot W(q')$,
  \item $K(q \cdot q') = K(q) \cdot K(q')$, 
  \item
For all $w, w' \in W(q \cdot q')$, $w \; EqRel(q' \cdot q) \; w'$ if and
only $$w = r_1
      \cdot r_2, \; w' = {r'}_1 \cdot {r'}_2 \text{ with } r_1 \; EqRel(q) \;
      {r'}_1, \; r_2 \; EqRel(q') \; {r'}_2$$
\item
$Canonizer(q \cdot q') = Canonizer(q) \cdot Canonizer(q')
$
\end{enumerate}
\begin{prooftree}
\AxiomC{$q$ is well formed}
\AxiomC{$q'$ is well formed}
\AxiomC{$W(q) \cdot^! W(q')$}
\AxiomC{$K(q) \cdot^! K(q')$}
\QuaternaryInfC{$q \cdot q$ is well formed}
\end{prooftree}
\end{center}
\item $q=q \; | \;  q'$
\begin{center}
\begin{enumerate}
  \item $W(q \cdot q') = W(q) \; | \; W(q')$,
  \item $K(q \cdot q') = K(q) \; | \; K(q')$, 
  \item
For all $w, w' \in W(q \; | \; q')$, $w \; EqRel(q' \; | \; q) \; w'$ if and
only if 
$$w \; EqRel(q) \; w' \text{ or } \; w \; EqRel(q') \; w'$$
\item
$Canonizer(q \; | \; q') = Canonizer(q) \; | \; Canonizer(q')
$
\end{enumerate}
\begin{prooftree}
\AxiomC{$q$ is well formed}
\AxiomC{$q'$ is well formed}
\AxiomC{$W(q) \cap W(q') = \varnothing$}
\TrinaryInfC{$q \cdot q$ is well formed}
\end{prooftree}
\end{center}
\item $q=q'^*$
\begin{center}
\begin{enumerate}
  \item $W(q'^*) = W(q')^*$,
  \item $K(q'^*) = K(q')^*$, 
  \item
For all $w, w' \in W(q'^*)$, $w \; EqRel(q'^*) \; w'$ if and
only if 
$$w = r_1 \cdot \ldots \cdot r_n, \; w' = {r'}_1 \cdot \ldots
      \cdot {r'}_n \text{ and } r_i \; EqRel(q) \; {r'}_i
      \text{ for all } 1 \leq i \leq n$$
      \item
$Canonizer(q'^*) = Canonizer(q')^* $
\end{enumerate}
\begin{prooftree}
\AxiomC{$q$ is well formed}
\AxiomC{$W(q)^{*!}$}
\AxiomC{$K(q)^{*!}$}
\TrinaryInfC{$q^*$ is well formed}
\end{prooftree}
\end{center}
\end{enumerate}
\begin{claim}
If $q$ is a well formed QRE, then
\begin{enumerate}
  \item  $EqRel(q)$ is an equivalence relation on $\mathcal{L}(W(q))$,
  \item  $\mathcal{L}(K(q))$ forms a complete set of representatives for
  $EqRel(q)$,
  \item $Canonizer(q):\mathcal{L}(W(q)) \longrightarrow \mathcal{L}(K(q))$ is a
  well-defined function, and
  \item  given any $w \in \mathcal{L}(W(q))$, $Canonizer(q)(w)$ is the unique
  $k$ in $\mathcal{L}(K(q))$ such that $k$ is equivalent to $w$ modulo
  $EqRel(q)$
  \end{enumerate}
\end{claim}
\section{Using QREs to Express Quotient Lenses}
Given regular expressions $R, S$ and equivalence relations $\sim_R, \sim_S$ are
defined on $\mathcal{L}(R)$ and $\mathcal{L}(S)$ respectively, a
\textit{quotient lens} $q :
R/{\sim_R}{\Longleftrightarrow} S/{\sim_S}$ from $R$ to $S$ is
a pair of functions $q.get:
\mathcal{L}(R) \longrightarrow \mathcal{L}(S)$ and $q.put : \mathcal{L}(S)
\longrightarrow \mathcal{L}(R)$ such that
\begin{align*}
q.put \; (q.get \; r) &\sim_R r\\
q.get \; (q.put \; s) &\sim_S s
\end{align*}
Additionally the components of $q$ must respect $\sim_R$ and $\sim_S$ i.e.
if $r \sim_R r'$ and $s \sim_S s'$ then $q.get \; r \; \sim_S q.get \; r'$
and $q.put \; s \sim_R q.put \; s' \; s'$.\\
Now let $c, c'$ be QREs. Since $K(c)$ and $K(c')$ form a complete set of
representatives for $W(c)$ modulo $EqRel(c)$ and $W(c')$ modulo $EqRel(c')$
respectively, then given a bijection $\ell : K(c) \Leftrightarrow K(c')$, we
may define a quotient lens $q : W(c)/EqRel(c) \Leftrightarrow
W(c')/EqRel(c')$ by 
\begin{align*}
q.get &= \ell \circ Canonizer(c)\\
q.put &= \ell^{-1} \circ Canonizer(c')
\end{align*}
[Reference needed] Define the class of \textit{bijective lenses} to be the set
of bijections between regular languages created from Boomerang lens combinators.
The syntax for the language of bijective lenses is given by
$$\ell := id(R) \; | \; const(s, s') \; | \;  swap(\ell,
\ell) \; | \; \ell \cdot \ell' \; |  \; (\ell \; | \; \ell') \; | \; \ell^* \;
| \; \ell \circ \ell',$$ where $R$ ranges over regular expressions and $s$
ranges over character strings.\\
The denotation of a bijective lens $\ell$ is a function from 
$\llbracket e \rrbracket$ of a bijective lens $\ell$ is defined
inductively by
\begin{align*}
\llbracket id(R) \rrbracket &= \{(r, r) \; | \; r \in \mathcal{L}(R)\}\\
\llbracket swap(\ell, \ell') \rrbracket &= \{(s \cdot t, t' \cdot s') \; | \;
(s, s') \in \llbracket \ell \rrbracket \text{ and } (t, t') \in \llbracket
\ell' \rrbracket\}\\
\llbracket \ell \cdot \ell' \rrbracket &= \{(s \cdot t, s' \cdot t) \; | \;
(s, s') \in \llbracket \ell \rrbracket \text{ and } (t, t') \in \llbracket
\ell' \rrbracket\}\\
\llbracket \ell \; | \; \ell' \rrbracket &= \{(s \cdot t) \; | \;
(s, t) \in \llbracket \ell \rrbracket \text{ or } (s, t) \in \llbracket
\ell' \rrbracket\}\\
\llbracket \ell^* \rrbracket &= \{(s_1 \cdot \ldots \cdot s_n, t_1 \cdot \ldots
\cdot t_n) \; | \; (s_i, t_i) \in \llbracket \ell \rrbracket \text{ for } 1
\leq i \leq n\}
\end{align*}
Having described the class of bijective lenses, we now define the class
$\mathcal{Q}$ of quotient lenses, whose language is defined inductively by the
following grammar:
$$ q := id(\ell) \; | \; lquot(q, \ell) \; | \; rquot(\ell, q) \; | \; q
\cdot q_2 \; | \; (q \; | \; q') \; | \; q^* \; | \; q \circ q',$$
where $\ell$ ranges over bijective lenses 
\subsection{Semantics of $\mathcal{Q}$}
Every element in $\mathcal{Q}$ has a type $q :
c \Longleftrightarrow c'$ where $c$ and $c'$ are QREs. The denotation
$\llbracket q \rrbracket$ of an element $q : c \Longleftrightarrow c'$ is a
quotient lens $\llbracket q \rrbracket :
W(c)/{EqRel(c)} \Longleftrightarrow W(c')/{EqRel(c')}$. We now give the
semantics of $\mathcal{Q}$, as well as inference rules that may be used to
derive elements of $\mathcal{Q}$:
\begin{enumerate}
  \item
  if $q = id(\ell)$, then
  \begin{prooftree}
\AxiomC{$\ell : R \Leftrightarrow S$}
\UnaryInfC{$id(\ell): id(R) \Leftrightarrow id(S)$}
\end{prooftree}
  \begin{align*}
  \llbracket q \rrbracket.get &=  \llbracket \ell \rrbracket, \text{ and }\\
  \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1}
  \end{align*}

  \item if $q = lquot(c, q')$, then
\begin{prooftree}
\AxiomC{$q' : c'  \Leftrightarrow c''$}
\AxiomC{$c$ is well formed}
\AxiomC{$K(c) = W(c')$}
\TrinaryInfC{$lquot(c, q'): c' \circ c \Leftrightarrow c''$}
\end{prooftree}
  \begin{align*}
  \llbracket q \rrbracket.get  &= \llbracket q'
  \rrbracket.get \circ Canonizer(c)\\
  \llbracket q \rrbracket.put &= \llbracket q' \rrbracket.put
  \end{align*}
  \item
  if $q = rquot(q', c'')$, then
  \begin{prooftree}
  \AxiomC{$q' : c \Leftrightarrow c'$}
  \AxiomC{$c''$ is well formed}

\AxiomC{$K(c'') = W(c')$}
\TrinaryInfC{$rquot(q', c):c \Leftrightarrow c'' \circ c'$}
\end{prooftree}
  \begin{align*}
  \llbracket q \rrbracket.get &= \llbracket q'
  \rrbracket.get\\
  \llbracket q \rrbracket.put &= \llbracket q'
  \rrbracket.put \circ Canonizer(c'')
  \end{align*}

  \item
  $q = q_2 \circ q_1$, then
  \begin{prooftree}
\AxiomC{$q_1 : c \Leftrightarrow c'$}
\AxiomC{$q_2 : c' \Leftrightarrow c''$}
\BinaryInfC{$q_2 \circ q_1: c \Leftrightarrow c''$}
\end{prooftree}
  \begin{align*}
  \llbracket q \rrbracket.get &= \llbracket q_2 \rrbracket.get\circ \llbracket
  q_1 \rrbracket.get, \text{ and }\\
  \llbracket q \rrbracket.put &= \llbracket q_1 \rrbracket.put \circ \llbracket
  q_2 \rrbracket.put
  \end{align*}
  \item
  $q = {q'}^*$, then
    \begin{prooftree}
\AxiomC{$q' : c \Leftrightarrow c'$}
\AxiomC{$W(c)^{*!}$ and $W(c')^{*!}$}
\AxiomC{$K(c)^{*!}$ and $K(c')^{*!}$}
\TrinaryInfC{${q'}^* : c^* \Leftrightarrow {c'}^*$}
\end{prooftree}
  \begin{align*}
  \llbracket q \rrbracket.get &= (\llbracket q' \rrbracket.get)^*, \text{
  and }\\
  \llbracket q \rrbracket.put &= (\llbracket q' \rrbracket.put)^*
  \end{align*}

  \item
  $q = q_1 \cdot q_2$, then
    \begin{prooftree}
\AxiomC{$q_1 : c_1 \Leftrightarrow d_1 $}
\AxiomC{$q_2 : c_2 \Leftrightarrow d_2$}
\AxiomC{${\substack{W(c_1) \cdot^! W(c_2)\\ K(c_1) \cdot^! K(c_2)}}$}
\AxiomC{${\substack{W(d_1) \cdot^! W(d_2)\\ K(d_1) \cdot^! K(d_2)}}$}
\QuaternaryInfC{$q_1 \cdot q_2: c_1 \cdot c_2
\Leftrightarrow d_1 \cdot d_2$}
\end{prooftree}
  \begin{align*}
  \llbracket q \rrbracket.get &= \llbracket q_1 \rrbracket.get \cdot \llbracket
  q_2 \rrbracket.get, \text{ and }\\
  \llbracket q \rrbracket.put &= \llbracket q_1 \rrbracket.put \cdot \llbracket
  q_2 \rrbracket.put
  \end{align*}

  \item
  $q = q_1 \; | \; q_2$, then
      \begin{prooftree}
\AxiomC{$q_1 : c_1 \Leftrightarrow d_1 $}
\AxiomC{$q_2 : c_2 \Leftrightarrow d_2$}
\AxiomC{$\mathcal{L}(W(c_1)) \cap \mathcal{L}(W(c_2)) = \varnothing$}
\AxiomC{$\mathcal{L}(W(d_1)) \cap \mathcal{L}(W(d_2)) = \varnothing$}
\QuaternaryInfC{$q_1 \; | \; q_2: (c_1 \; | \; c_2)
\Leftrightarrow (d_1 \; | \; d_2)$}
\end{prooftree}
  $$
  \llbracket q_1 \; | \; q_2 \rrbracket.get(s) = 
  \begin{cases}
  \llbracket q_1 \rrbracket.get (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
  \llbracket q_2 \rrbracket.get (s) & \text{if } s \in \mathcal{L}(W(c_2))\\
  \end{cases}$$
  $$\llbracket q_1 \; | \; q_2 \rrbracket.put(s) = 
  \begin{cases}
  \llbracket q_1 \rrbracket.put (s) & \text{if } s \in \mathcal{L}(W(d_1))\\
  \llbracket q_2 \rrbracket.put (s) & \text{if } s \in \mathcal{L}(W(d_2))\\
  \end{cases}
  $$
\end{enumerate}

\begin{claim}
If there exists a derivation $q:c \Leftrightarrow c'$ then
$\llbracket q \rrbracket$ is a quotient lens from $R$ to $S$.
\end{claim}

\subsection{Normal Forms for Quotient Lenses in $\mathcal{Q}$}
\begin{claim}
If there is a derivation $q : c \Leftrightarrow c'$ in
$\mathcal{Q}$ then there exists a bijective lens $\ell : K(c) \Leftrightarrow
K(c')$ such that
\begin{align*}
\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ Canonizer(c)\\
\llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
Canonizer(c')
\end{align*}
\end{claim}
\begin{proof}
Assume that $q : c \Leftrightarrow c'$. We proceed by induction over the
derivation $q : c \Leftrightarrow c'$.
\begin{enumerate}
  \item
  $id(\ell): id(R) \Leftrightarrow id(S)$ where $\ell : R \Leftrightarrow
S$. Then
  \begin{align*}
  \llbracket id(\ell) \rrbracket.get &=  \llbracket \ell \rrbracket = \llbracket \ell \rrbracket \circ
  id_{\mathcal{L}(R)} = \llbracket \ell \rrbracket \circ Canonizer(id(R)), \text{ and }\\
  \llbracket id(\ell) \rrbracket.put &= \llbracket \ell \rrbracket^{-1} = \llbracket \ell \rrbracket^{-1} \circ
  id_{\mathcal{L}(S)} = \llbracket \ell \rrbracket^{-1} \circ Canonizer(id(S))
  \end{align*}
  \item
  $lquot(c, q'): c' \circ c \Leftrightarrow c''$ where $q' : c' 
  \Leftrightarrow c''$, $c$ is well formed and $K(c) = W(c')$. Then
\begin{align*}
  \llbracket q \rrbracket.get  &= \llbracket q'
  \rrbracket.get \circ Canonizer(c)\\
  \llbracket q \rrbracket.put &= \llbracket q' \rrbracket.put
  \end{align*}
  By the induction hypothesis, there exists a bijective lens $\ell :
  K(c') \Leftrightarrow K(c'')$ such that 
  \begin{align*}
\llbracket q' \rrbracket.get &= \llbracket \ell \rrbracket \circ Canonizer(c')\\
\llbracket q' \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
Canonize(c'')
\end{align*}
Consequently
\begin{align*}
  \llbracket q \rrbracket.get  &= (\llbracket \ell \rrbracket \circ
  Canonizer(c')) \circ Canonizer(c) = \llbracket \ell \rrbracket \circ
  (Canonizer(c' \circ c))\\
  \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
  Canonize(c'')
  \end{align*}

  \item
  $rquot(q', c''):c \Leftrightarrow c'' \circ c'$ where $q' : c \Leftrightarrow
  c'$, $c''$ is well formed and $K(c'') = W(c')$. Proceed as in the previous
  case.
\item
$q_2 \circ q_1: c \Leftrightarrow c''$ where $q_1 : c \Leftrightarrow c'$ and
$q_2 : c' \Leftrightarrow c''$. Then
  \begin{align*}
  \llbracket q \rrbracket.get &= \llbracket q_2 \rrbracket.get\circ \llbracket
  q_1 \rrbracket.get, \text{ and }\\
  \llbracket q \rrbracket.put &= \llbracket q_1 \rrbracket.put \circ \llbracket
  q_2 \rrbracket.put
  \end{align*}
  By the induction hypothesis, there exist bijective lenses
  $\ell_1 :
  K(c) \Leftrightarrow K(c')$ and $\ell_2 : K(c') \Leftrightarrow K(c'')$ such
  that
  \begin{align*}
\llbracket q_1 \rrbracket.get &= \llbracket \ell_1 \rrbracket \circ
Canonizer(c)\\
\llbracket q_1 \rrbracket.put &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
Canonizer(c')
\end{align*}
and
\begin{align*}
\llbracket q_2 \rrbracket.get &= \llbracket \ell_2 \rrbracket \circ
Canonize(c')\\
\llbracket q_2 \rrbracket.put &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
Canonize(c'')
\end{align*}
Consequently,
\begin{align*}
\llbracket q_2 \rrbracket.get \circ \llbracket q_1 \rrbracket.get &=
(\llbracket \ell_2 \rrbracket \circ Canonizer(c')) \circ (\llbracket \ell_1
\rrbracket \circ Canonizer(c))\\
&= \llbracket \ell_2 \rrbracket \circ (Canonizer(c') \circ \llbracket \ell_1
\rrbracket) \circ Canonizer(c)\\
&= (\llbracket \ell_2 \rrbracket \circ \llbracket \ell_1 \rrbracket) \circ
Canonizer(c)\\
&= \llbracket \ell_2  \circ  \ell_1 \rrbracket \circ
Canonizer(c)
\end{align*} 
A similar argument shows that 
$$\llbracket q_1 \rrbracket.put \circ \llbracket q_2 \rrbracket.put =
\llbracket \ell_2  \circ  \ell_1 \rrbracket^{-1} \circ
Canonizer(c)$$
\item  
${q'}^* : c^* \Leftrightarrow {c'}^*$ where $q' : c \Leftrightarrow c'$,
$W(c)^{*!}$ and $W(c')^{*!}$ and $K(c)^{*!}$ and $K(c')^{*!}$. Then
  \begin{align*}
  \llbracket {q'}^* \rrbracket.get &= (\llbracket q' \rrbracket.get)^*, \text{
  and }\\
  \llbracket {q'}^* \rrbracket.put &= (\llbracket q' \rrbracket.put)^*
  \end{align*}
  By the induction hypothesis there exists a bijective lens $\ell : K(c)
  \Leftrightarrow K(c')$ such that 
   that
  \begin{align*}
\llbracket q' \rrbracket.get &= \llbracket \ell \rrbracket \circ
Canonizer(c)\\
\llbracket q' \rrbracket.put &= {\llbracket \ell \rrbracket}^{-1} \circ
Canonizer(c')
\end{align*}
Consequentlty
\begin{align*}
\llbracket {q'}^* \rrbracket.get &= (\llbracket \ell \rrbracket \circ
Canonizer(c))^* = \llbracket \ell \rrbracket^* \circ
Canonizer(c)^* = \llbracket \ell^* \rrbracket \circ
Canonizer(c^*)\\
\llbracket {q'}^* \rrbracket.put &= (\llbracket \ell \rrbracket^{-1} \circ
Canonizer(c'))^* = (\llbracket \ell \rrbracket^{-1})^* \circ
Canonizer(c')^* = \llbracket \ell^* \rrbracket^{-1} \circ
Canonizer(c'^*)\\
\end{align*}
\item
  $q_1 \cdot q_2: c_1 \cdot c_2 \Leftrightarrow d_1 \cdot d_2$, where $q_1 : c_1
  \Leftrightarrow d_1 $,  $q_2 : c_2 \Leftrightarrow d_2$, $W(c_1)
  \cdot^! W(c_2)$, $K(c_1) \cdot^! K(c_2)$, $W(d_1) \cdot^! W(d_2)$ and $
  K(d_1) \cdot^! K(d_2)$. Then
  \begin{align*}
  \llbracket q \rrbracket.get &= \llbracket q_1 \rrbracket.get \cdot \llbracket
  q_2 \rrbracket.get, \text{ and }\\
  \llbracket q \rrbracket.put &= \llbracket q_1 \rrbracket.put \cdot \llbracket
  q_2 \rrbracket.put
  \end{align*}
By the induction hypothesis, there exist bijective lenses $\ell_1 : K(c_1)
\Leftrightarrow K(d_1)$ and $\ell_2 : K(c_2) \Leftrightarrow K(d_2)$ such that
\begin{align*}
\llbracket q_1 \rrbracket.get &= \llbracket \ell_1 \rrbracket \circ
Canonizer(c_1)\\
\llbracket q_1 \rrbracket.put &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
Canonizer(d_1)
\end{align*}
and
\begin{align*}
\llbracket q_2 \rrbracket.get &= \llbracket \ell_2 \rrbracket \circ
Canonizer(c_2)\\
\llbracket q_2 \rrbracket.put &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
Canonizer(d_2)
\end{align*}
Consequently,
\begin{align*}
  \llbracket q \rrbracket.get &= (\llbracket \ell_1 \rrbracket \circ
Canonizer(c_1)) \cdot  (\llbracket \ell_2 \rrbracket \circ
Canonizer(c_2))\\
&= (\llbracket \ell_1 \rrbracket \cdot \llbracket \ell_2
\rrbracket) \circ (Canonizer(c_1) \cdot Canonizer(c_2))\\
&= \llbracket \ell_1 \cdot  \ell_2 \rrbracket \circ Canonizer(c_1 \cdot c_2)
\end{align*}
Similarly
$$
  \llbracket q \rrbracket.put = \llbracket \ell_1 \cdot  \ell_2 \rrbracket^{-1}
  \circ Canonizer(d_1 \cdot d_2) $$
  \item
  $q = q_1 \; | \; q_2$ where $q_1 : c_1 \Leftrightarrow d_1 $, $q_2 : c_2
  \Leftrightarrow d_2$, $\mathcal{L}(W(c_1)) \cap \mathcal{L}(W(c_2)) =
  \varnothing$ and $\mathcal{L}(W(d_1)) \cap \mathcal{L}(W(d_2)) = \varnothing$.
  Then
  $$
  \llbracket q_1 \; | \; q_2 \rrbracket.get(s) = 
  \begin{cases}
  \llbracket q_1 \rrbracket.get (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
  \llbracket q_2 \rrbracket.get (s) & \text{if } s \in \mathcal{L}(W(c_2))\\
  \end{cases}$$
  $$\llbracket q_1 \; | \; q_2 \rrbracket.put(s) = 
  \begin{cases}
  \llbracket q_1 \rrbracket.put (s) & \text{if } s \in \mathcal{L}(W(d_1))\\
  \llbracket q_2 \rrbracket.put (s) & \text{if } s \in \mathcal{L}(W(d_2))\\
  \end{cases}
  $$
By the induction hypothesis, there exist bijective lenses $\ell_1 : K(c_1)
\Leftrightarrow K(d_1)$ and $\ell_2 : K(c_2) \Leftrightarrow K(d_2)$ such that
\begin{align*}
\llbracket q_1 \rrbracket.get &= \llbracket \ell_1 \rrbracket \circ
Canonizer(c_1)\\
\llbracket q_1 \rrbracket.put &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
Canonizer(d_1)
\end{align*}
and
\begin{align*}
\llbracket q_2 \rrbracket.get &= \llbracket \ell_2 \rrbracket \circ
Canonizer(c_2)\\
\llbracket q_2 \rrbracket.put &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
Canonizer(d_2)
\end{align*}
Consequently,
$$
  \llbracket q_1 \; | \; q_2 \rrbracket.get(s) = 
  \begin{cases}
  \llbracket \ell_1 \rrbracket \circ
Canonizer(c_1) (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
  \llbracket \ell_2 \rrbracket \circ
Canonizer(c_2) (s) & \text{if } s \in \mathcal{L}(W(c_2)),\\
  \end{cases}$$
  so $\llbracket q_1 \; | \; q_2 \rrbracket.get = \llbracket \ell_1 \; | \;
  \ell_2 \rrbracket \circ Canonizer(c_1 \; | \; c_2)$. A similar argument shows
  that $\llbracket q_1 \; | \; q_2 \rrbracket.put = \llbracket \ell_1 \; | \;
  \ell_2 \rrbracket^{-1} \circ Canonizer(d_1 \; | \; d_2)$.\\
  This completes the proof.
\end{enumerate}
\end{proof}
\section{Synthesizing Quotient Lenses}
\section{Related Work}

\section{Conclusion}

\end{document}

### 2017-09-27 anonymized_material/comparisons/prose/Extraction.Text/LearningSamples.cs
remove irrelevant stuff

@@ -1,363 +1,0 @@
-﻿using System;
-using System.Collections.Generic;
-using System.Linq;
-using System.Text.RegularExpressions;
-using Microsoft.ProgramSynthesis.AST;
-using Microsoft.ProgramSynthesis.Extraction.Text;
-using Microsoft.ProgramSynthesis.Extraction.Text.Constraints;
-using Microsoft.ProgramSynthesis.Extraction.Text.Semantics;
-using Microsoft.ProgramSynthesis.Utils;
-using Microsoft.ProgramSynthesis.Transformation.Text;
-using Microsoft.ProgramSynthesis.Transformation.Text.Semantics;
-using Microsoft.ProgramSynthesis.Wrangling.Constraints;
-using Microsoft.ProgramSynthesis.VersionSpace;
-using Microsoft.ProgramSynthesis.Wrangling;
-using System.Diagnostics;
-using VBFIO = Microsoft.VisualBasic.FileIO;
-
-namespace Extraction.Text
-{
-    /// <summary>
-    ///     Extraction.Text learns programs to extract a single string region or a sequence of string regions from text files.
-    ///     This class demonstrates some common usage of Extraction.Text APIs.
-    /// </summary>
-    internal static class LearningSamples
-	{
-		private class Result
-		{
-			public ResultType ResultType { get; set; }
-			public long Milliseconds { get; set; }
-		}
-
-		private enum ResultType { Success, Exception, IncorrectResponse, NothingLearned }
-
-		private const int NUM_TESTS = 100;
-
-        private static void Main(string[] args)
-		{
-			if (args.Length != 2)
-			{
-				throw new Exception(args.Length.ToString());
-			}
-			var numExamples = int.Parse(args[0]);
-			var fileName = args[1];
-			var results = GetResults(fileName, numExamples);
-			var failedResults = results.Where(result => result.ResultType != ResultType.Success).ToList();
-			if (failedResults.IsEmpty())
-			{
-				Console.Out.WriteLine(results.Select(result => result.Milliseconds).Sum());
-			}
-			else
-			{
-				Console.Out.WriteLine(failedResults.First().ResultType);
-			}
-
-            // Learning sequence is similar to learning region. 
-            // We only illustrate some API usages. Other sequence learning APIs are similar to their region APIs counterpart.
-            // Note: we need to give positive examples continuously. 
[…]

### 2017-09-27 anonymized_material/program/check-baseline.py

@@ -1,83 +1,0 @@
-#!/usr/bin/python
-
-from __future__ import print_function
-
-import difflib
-import os
-from os.path import splitext, join
-import subprocess
-import sys
-
-TEST_EXT = '.ml'
-BASELINE_EXT = '.out'
-FLAGS = ['-nosugar', '-noincomplete-warning']
-
-def find_tests(root):
-    tests = []
-    for path, dirs, files in os.walk(root):
-        files = [(f[0], f[1]) for f in [splitext(f) for f in files]]
-        tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
-    return tests
-
-def run_test(prog, path):
-    output_string = subprocess.check_output([prog] + FLAGS + [path], stderr=subprocess.STDOUT)
-    return output_string.decode('utf-8').splitlines(1)
-
-def extract_baseline_output(path):
-    with open(path) as f:
-        lines = f.readlines()
-        # Strip the timestamp at the top of the output file
-        lines = lines[1:]
-        return lines
-
-def create_diff(output, baseline, test_path, baseline_path):
-    diff_gen = difflib.unified_diff(output, baseline, fromfile=test_path, tofile=baseline_path)
-    diff = []
-    for line in diff_gen:
-        diff.append(line)
-    return diff
-
-def check_test(prog, path, base):
-    test_path     = join(path, base + TEST_EXT)
-    baseline_path = join(path, base + BASELINE_EXT)
-    print("Checking {0}...".format(test_path))
-    if not os.path.isfile(baseline_path):
-        print("Baseline file not found: {0}".format(baseline_path))
-        return
-    try:
-        output = run_test(prog, test_path)
-    except subprocess.CalledProcessError as err:
-        print("Encountered an error running test {0}".format(test_path))
-        return
-    baseline = extract_baseline_output(baseline_path)
-    diff = create_diff(output, baseline, test_path, baseline_path)
-    if (len(diff) > 0):
-        print(''.join(diff))
-
-def check_tests(prog, root):
-    for path, base in find_tests(root):
-        check_test(prog, path, base)
-
-def print_usage(args):
-    print("Usage: {0} <prog> <test|testdir>".format(args[0]))
-
-def main(args):
-    if len(args) == 3:
-        prog = args[1]
-        path = args[2]
-        if not os.path.exists(prog):
-            print_usage(args)
-        elif os.path.exists(path) and os.path.isdir(path):
-            check_tests(prog, path)
-        else:
[…]

### 2017-09-27 anonymized_material/program/generate-data.py

@@ -1,179 +1,0 @@
-#!/usr/local/bin/python
-
-from __future__ import print_function
-from easyprocess import EasyProcess
-
-import os
-import csv
-from os.path import splitext, join
-import subprocess
-import sys
-import time
-
-TEST_EXT = '.ls'
-BASELINE_EXT = '.out'
-BASE_FLAGS = []
-TIMEOUT_TIME = 600
-GENERATE_EXAMPLES_TIMEOUT_TIME = 600000
-
-REPETITION_COUNT = 10
-
-def ensure_dir(f):
-    d = os.path.dirname(f)
-    if not os.path.exists(d):
-        os.makedirs(d)
-
-def transpose(matrix):
-    return zip(*matrix)
-
-def find_tests(root):
-    tests = []
-    for path, dirs, files in os.walk(root):
-        files = [(f[0], f[1]) for f in [splitext(f) for f in files]]
-        tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
-    return tests
-
-def gather_datum(prog, path, base, additional_flags, timeout):
-    process_output = EasyProcess([prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)]).call(timeout=timeout)
-    return (process_output.stdout,process_output.stderr)
-
-
-def gather_data(rootlength, prog, path, base):
-    current_data = {"Test":join(path, base).replace("_","-")[rootlength:]}
-
-    def gather_col(flags, run_combiner, col_name, timeout_time, repetition_count):
-        print(col_name)
-        run_data = []
-        timeout = False
-        error = False
-        for iteration in range(repetition_count):
-    	    (datum,err) = gather_datum(prog, path, base,flags,timeout_time)
-            if datum == "":
-                if err == "":
-                    timeout = True
-                else:
-                    error = True
-                break
-            else:
-	            run_data.append(datum.split(","))
-        if error:
-            current_data[col_name]=-1
-        elif timeout:
-	        current_data[col_name]=-1
-        else:
-            run_data_transpose = transpose(run_data)
-            current_data[col_name]=run_combiner(run_data_transpose)
-
[…]

### 2017-09-27 anonymized_material/program/generate_IO_specs.py

@@ -1,79 +1,0 @@
-#!/usr/local/bin/python
-
-from __future__ import print_function
-from easyprocess import EasyProcess
-
-import os
-from os.path import splitext, join
-import subprocess
-import sys
-import time
-
-TEST_EXT = '.ls'
-BASELINE_EXT = '.out'
-TSV_EXT = '.tsv'
-BASE_FLAGS = ['-generate_io_spec']
-TIMEOUT_TIME = 6000
-REPETITION_COUNT = 2
-BASE_TSV_DIR = 'generated_io_specs/'
-
-def ensure_dir(f):
-    if not os.path.exists(f):
-        os.makedirs(f)
-
-def find_tests_and_ensure_output_dirs(root):
-    tests = []
-    for path, dirs, files in os.walk(root):
-        files = [(f[0], f[1]) for f in [splitext(f) for f in files]]
-        tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
-        ensure_dir(BASE_TSV_DIR + path)
-    return tests
-
-def gather_tsv(rootlength, prog, path, base, num_examples):
-    process_output = EasyProcess([prog] + BASE_FLAGS + [str(num_examples)] + [join(path, base + TEST_EXT)]).call(timeout=TIMEOUT_TIME)
-    if (process_output.stdout == "" or process_output.stderr != "" or process_output.return_code != 0):
-        raise Exception("emptystdout")
-    return process_output.stdout
-
-def write_tsv(tsv, path, base):
-    with open(join(BASE_TSV_DIR+path, base + TSV_EXT),'w') as tsvfile:
-        tsvfile.write(tsv)
-
-def print_usage(args):
-    print("Usage: {0} <prog> <testdir>".format(args[0]))
-
-def transform_data(path, base, run_data):
-    current_data = {"Test":join(path, base + TEST_EXT).replace("_","-")[6:]}
-    run_data_transpose = transpose(run_data)
-    for index in range(len(run_data_transpose)/2):
-	col_name = run_data_transpose[index][0]
-	col_data = run_data_transpose[index+1]
-        if "" in col_data:
-	    current_data[col_name]="Timeout"
-        else:
-            col = [float(x) for x in col_data]
-            current_data[col_name] = str(sum(col)/len(col))
-    return current_data
-
-def main(args):
-    if len(args) == 4:
-        prog = args[1]
-        path = args[2]
-        num_examples = args[3]
-        rootlength = len(path)
-        data = []
[…]

### 2017-09-27 anonymized_material/program/generate_extraction_specs.py

@@ -1,79 +1,0 @@
-#!/usr/local/bin/python
-
-from __future__ import print_function
-from easyprocess import EasyProcess
-
-import os
-from os.path import splitext, join
-import subprocess
-import sys
-import time
-
-TEST_EXT = '.ls'
-BASELINE_EXT = '.out'
-TSV_EXT = '.tsv'
-BASE_FLAGS = ['-generate_extraction_spec']
-TIMEOUT_TIME = 6000
-REPETITION_COUNT = 2
-BASE_TSV_DIR = 'generated_extraction_specs/'
-
-def ensure_dir(f):
-    if not os.path.exists(f):
-        os.makedirs(f)
-
-def find_tests_and_ensure_output_dirs(root):
-    tests = []
-    for path, dirs, files in os.walk(root):
-        files = [(f[0], f[1]) for f in [splitext(f) for f in files]]
-        tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
-        ensure_dir(BASE_TSV_DIR + path)
-    return tests
-
-def gather_tsv(rootlength, prog, path, base, num_examples):
-    process_output = EasyProcess([prog] + BASE_FLAGS + [str(num_examples)] + [join(path, base + TEST_EXT)]).call(timeout=TIMEOUT_TIME)
-    if (process_output.stderr != "" or process_output.return_code != 0):
-        raise Exception("failure")
-    return process_output.stdout
-
-def write_tsv(tsv, path, base):
-    with open(join(BASE_TSV_DIR+path, base + TSV_EXT),'w') as tsvfile:
-        tsvfile.write(tsv)
-
-def print_usage(args):
-    print("Usage: {0} <prog> <testdir>".format(args[0]))
-
-def transform_data(path, base, run_data):
-    current_data = {"Test":join(path, base + TEST_EXT).replace("_","-")[6:]}
-    run_data_transpose = transpose(run_data)
-    for index in range(len(run_data_transpose)/2):
-	col_name = run_data_transpose[index][0]
-	col_data = run_data_transpose[index+1]
-        if "" in col_data:
-	    current_data[col_name]="Timeout"
-        else:
-            col = [float(x) for x in col_data]
-            current_data[col_name] = str(sum(col)/len(col))
-    return current_data
-
-def main(args):
-    if len(args) == 4:
-        prog = args[1]
-        path = args[2]
-        num_examples = args[3]
-        rootlength = len(path)
-        data = []
-        base_tsv_dir = ensure_dir('generated_extraction_specs/')
[…]

### 2017-09-28 program/generate_extraction_specs.py
add generated extraction specs

@@ -1,7 +1,6 @@
 #!/usr/local/bin/python
 
 from __future__ import print_function
-from easyprocess import EasyProcess
 
 import os
 from os.path import splitext, join
@@ -30,14 +29,14 @@ def find_tests_and_ensure_output_dirs(root):
     return tests
 
 def gather_tsv(rootlength, prog, path, base, num_examples):
-    process_output = EasyProcess([prog] + BASE_FLAGS + [str(num_examples)] + [join(path, base + TEST_EXT)]).call(timeout=TIMEOUT_TIME)
-    if (process_output.stderr != "" or process_output.return_code != 0):
-        raise Exception("failure")
-    return process_output.stdout
+    f = open("outputfile","w")
+    process_output = subprocess.Popen([prog] + BASE_FLAGS + [str(num_examples)] + [join(path, base + TEST_EXT)], stdout=f)
+    while (process_output.poll() == None):
+        time.sleep(5)
+    f.close()
 
 def write_tsv(tsv, path, base):
-    with open(join(BASE_TSV_DIR+path, base + TSV_EXT),'w') as tsvfile:
-        tsvfile.write(tsv)
+    os.rename("outputfile",join(BASE_TSV_DIR+path, base+TSV_EXT))
 
 def print_usage(args):
     print("Usage: {0} <prog> <testdir>".format(args[0]))

### 2017-09-28 comparisons/prose/Extraction.Text/generate-data.py
add prose

@@ -1,0 +1,107 @@
+#!/usr/local/bin/python
+
+from __future__ import print_function
+from easyprocess import EasyProcess
+
+import os
+import csv
+from os.path import splitext, join
+import subprocess
+import sys
+import time
+
+TEST_EXT = '.tsv'
+BASELINE_EXT = '.out'
+BASE_FLAGS = []
+TIMEOUT_TIME = 600
+
+def ensure_dir(f):
+    d = os.path.dirname(f)
+    if not os.path.exists(d):
+        os.makedirs(d)
+
+def transpose(matrix):
+    return zip(*matrix)
+
+def find_tests(root):
+    tests = []
+    for path, dirs, files in os.walk(root):
+        files = [(f[0], f[1]) for f in [splitext(f) for f in files]]
+        tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
+    return tests
+
+def gather_datum(prog, path, base, additional_flags, timeout):
+    print(["mono", prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)])
+    process_output = EasyProcess(["mono", prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)]).call(timeout=timeout)
+    return (process_output.stdout,process_output.stderr)
+
+
+def gather_data(rootlength, prog, path, base):
+    current_data = {"Test":join(path, base).replace("_","-")[rootlength:]}
+    timeout = False
+    error = False
+    num_inputs = 0
+    result = "IncorrectResponse"
+    while (result == "IncorrectResponse"):
+        num_inputs=int(float(num_inputs)*1.5+1.0)
+        print(num_inputs)
+        (datum,err) = gather_datum(prog, path, base, [str(num_inputs)],TIMEOUT_TIME)
+        if datum == "":
+            if err == "":
+	            result="-1"
+            else:
+                print(err)
+                result="-1"
+        elif (num_inputs > 100000):
+            result="TooLargeSpec"
+        else:
+	        result=datum
+    print(result)
+    current_data["Result"]=result
+    current_data["ExsTaken"]=num_inputs
+    return current_data
+
+def print_data(data):
+    ensure_dir("generated_data/")
+    with open("generated_data/data.csv", "wb") as csvfile:
+	datawriter = csv.DictWriter(csvfile,fieldnames=data[0].keys())
+	datawriter.writeheader()
+	datawriter.writerows(data)
+
[…]

### 2017-09-28 comparisons/prose/Transformation.Text/generate-data.py

@@ -1,0 +1,113 @@
+#!/usr/local/bin/python
+
+from __future__ import print_function
+from easyprocess import EasyProcess
+
+import os
+import csv
+from os.path import splitext, join
+import subprocess
+import sys
+import time
+
+TEST_EXT = '.tsv'
+BASELINE_EXT = '.out'
+BASE_FLAGS = []
+TIMEOUT_TIME = 600
+
+counter = 0
+
+def ensure_dir(f):
+    d = os.path.dirname(f)
+    if not os.path.exists(d):
+        os.makedirs(d)
+
+def transpose(matrix):
+    return zip(*matrix)
+
+def find_tests(root):
+    tests = []
+    for path, dirs, files in os.walk(root):
+        files = [(f[0], f[1]) for f in [splitext(f) for f in files]]
+        tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
+    return tests
+
+def gather_datum(prog, path, base, additional_flags, timeout):
+    print(["mono", prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)])
+    process_output = EasyProcess(["mono", prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)]).call(timeout=timeout)
+    return (process_output.stdout,process_output.stderr)
+
+
+def gather_data(rootlength, prog, path, base):
+    global counter
+    current_data = {"Test":join(path, base).replace("_","-")[rootlength:]}
+    timeout = False
+    error = False
+    num_inputs = 0
+    result = "IncorrectResponse"
+    while (result == "IncorrectResponse"):
+        num_inputs=int(num_inputs*1.5+1.0)
+        print(num_inputs)
+        (datum,err) = gather_datum(prog, path, base, [str(num_inputs)],TIMEOUT_TIME)
+        if datum == "" or err != "":
+            if err == "":
+	            result="-1"
+            else:
+                print("what")
+                print(err)
+                result="-1"
+        elif (num_inputs > 100000):
+            result="-1"
+        else:
+	        result=datum
+    print(result)
+    current_data["Result"]=result
+    current_data["ExsTaken"]=num_inputs
+    counter = counter+1
+    print_data([current_data],str(counter));
+    return current_data
+
+def print_data(data,file_addition):
+    ensure_dir("generated_data/")
+    with open("generated_data/data" + file_addition + ".csv", "wb") as csvfile:
[…]

### 2017-09-28 comparisons/prose/Extraction.Text/Properties/AssemblyInfo.cs
add in extraction and transformation program

@@ -1,0 +1,37 @@
+﻿using System.Reflection;
+using System.Runtime.CompilerServices;
+using System.Runtime.InteropServices;
+
+// General Information about an assembly is controlled through the following 
+// set of attributes. Change these attribute values to modify the information
+// associated with an assembly.
+[assembly: AssemblyTitle("Microsoft.ProgramSynthesis.Extraction.Text.Sample")]
+[assembly: AssemblyDescription("")]
+[assembly: AssemblyConfiguration("")]
+[assembly: AssemblyCompany("Microsoft")]
+[assembly: AssemblyProduct("Microsoft.ProgramSynthesis.Extraction.Text.Sample")]
+[assembly: AssemblyCopyright("Copyright ©  2015")]
+[assembly: AssemblyTrademark("")]
+[assembly: AssemblyCulture("")]
+
+// Setting ComVisible to false makes the types in this assembly not visible 
+// to COM components.  If you need to access a type in this assembly from 
+// COM, set the ComVisible attribute to true on that type.
+[assembly: ComVisible(false)]
+
+// The following GUID is for the ID of the typelib if this project is exposed to COM
+[assembly: Guid("80fa46f0-36e2-4a1a-b4ec-414676d2a38d")]
+
+// Version information for an assembly consists of the following four values:
+//
+//      Major Version
+//      Minor Version 
+//      Build Number
+//      Revision
+//
+// You can specify all the values or you can default the Build and Revision Numbers 
+// by using the '*' as shown below:
+// [assembly: AssemblyVersion("1.0.*")]
+[assembly: AssemblyVersion("2.0.1")]
+[assembly: AssemblyInformationalVersion("2.0.1")]
+[assembly: AssemblyFileVersion("2.0.1.0")]

### 2017-09-28 comparisons/prose/Transformation.Text/Properties/AssemblyInfo.cs

@@ -1,0 +1,36 @@
+﻿using System.Reflection;
+using System.Runtime.InteropServices;
+
+// General Information about an assembly is controlled through the following 
+// set of attributes. Change these attribute values to modify the information
+// associated with an assembly.
+[assembly: AssemblyTitle("Microsoft.ProgramSynthesis.Transformation.Text.Sample")]
+[assembly: AssemblyDescription("")]
+[assembly: AssemblyConfiguration("")]
+[assembly: AssemblyCompany("Microsoft")]
+[assembly: AssemblyProduct("Microsoft.ProgramSynthesis.Transformation.Text.Sample")]
+[assembly: AssemblyCopyright("Copyright ©  2015")]
+[assembly: AssemblyTrademark("")]
+[assembly: AssemblyCulture("")]
+
+// Setting ComVisible to false makes the types in this assembly not visible 
+// to COM components.  If you need to access a type in this assembly from 
+// COM, set the ComVisible attribute to true on that type.
+[assembly: ComVisible(false)]
+
+// The following GUID is for the ID of the typelib if this project is exposed to COM
+[assembly: Guid("2698a7f0-26b7-42e7-a9c6-11bb4aeaf23d")]
+
+// Version information for an assembly consists of the following four values:
+//
+//      Major Version
+//      Minor Version 
+//      Build Number
+//      Revision
+//
+// You can specify all the values or you can default the Build and Revision Numbers 
+// by using the '*' as shown below:
+// [assembly: AssemblyVersion("1.0.*")]
+[assembly: AssemblyVersion("2.0.1")]
+[assembly: AssemblyInformationalVersion("2.0.1")]
+[assembly: AssemblyFileVersion("2.0.1.0")]

### 2017-09-28 comparisons/prose/Transformation.Text/SimpleAPI.cs

@@ -1,0 +1,101 @@
+﻿using System;
+using System.Collections.Generic;
+using System.Diagnostics.CodeAnalysis;
+using System.Linq;
+using Microsoft.ProgramSynthesis;
+using Microsoft.ProgramSynthesis.AST;
+using Microsoft.ProgramSynthesis.Learning;
+using Microsoft.ProgramSynthesis.Learning.Strategies;
+using Microsoft.ProgramSynthesis.Specifications;
+using Microsoft.ProgramSynthesis.Transformation.Text;
+using Microsoft.ProgramSynthesis.Transformation.Text.Semantics;
+
+namespace Transformation.Text
+{
+    /// <summary>
+    ///     Simplified version of <see cref="Program" /> to demonstrate lower level API usage.
+    /// </summary>
+    public class TextTransformationProgram
+    {
+        /// <summary>
+        ///     Constructor for a Transformation.Text Program.
+        /// </summary>
+        /// <param name="program">The learnt program.</param>
+        private TextTransformationProgram(ProgramNode program)
+        {
+            ProgramNode = program;
+        }
+
+        public ProgramNode ProgramNode { get; set; }
+
+        /// <summary>
+        ///     Learn <paramref name="k" /> top-ranked Transformation.Text programs for a given set of input-output examples.
+        /// </summary>
+        /// <param name="trainingExamples">
+        ///     The set of input-output examples as a Tuple of the input and the output.
+        /// </param>
+        /// <param name="additionalInputs">
+        ///     The set of additional inputs that do not have output examples, which helps rank learnt programs.
+        /// </param>
+        /// <param name="k">the number of top programs</param>
+        /// <returns>The top-k ranked programs as <see cref="TextTransformationProgram" />s</returns>
+        public static IEnumerable<TextTransformationProgram> LearnTopK(IDictionary<string, string> trainingExamples,
+                                                                       IEnumerable<string> additionalInputs = null,
+                                                                       int k = 1)
+        {
+            if (trainingExamples == null) throw new ArgumentNullException(nameof(trainingExamples));
+            // Load Transformation.Text grammar
+            Grammar grammar = Language.Grammar;
+            DomainLearningLogic learningLogic = new Witnesses(grammar);
+
+            // Setup configuration of synthesis process.
+            var engine = new SynthesisEngine(grammar, new SynthesisEngine.Config
+            {
+                // Strategies perform the actual logic of the synthesis process.
+                Strategies = new[] { new DeductiveSynthesis(learningLogic) },
+                UseThreads = false,
[…]

### 2017-09-29 combine-data.py
makefile improvements

@@ -160,7 +160,7 @@ def main(args):
         prog1_path = args[1]
         prog2_path = args[2]
         prog3_path = args[3]
-        ensure_dir("paper/generated-data/")
+        ensure_dir("papers/BijectiveLensSynth/generated-data/")
         csv1 = retrieve_csv(prog1_path)
         csv2 = retrieve_csv(prog2_path)
         csv3 = retrieve_csv(prog3_path)
@@ -174,7 +174,7 @@ def main(args):
             for key, value in row.items():
                 relevant_data[key] = value
         print(data_keyed_by_test)
-        with open("paper/generated-data/data.csv", "wb") as csvfile:
+        with open("papers/BijectiveLensSynth/generated-data/data.csv", "wb") as csvfile:
             fieldnames = data_keyed_by_test[data_keyed_by_test.keys()[0]].keys()
             writer = csv.DictWriter(csvfile, fieldnames=fieldnames)
 

### 2017-09-29 comparisons/prose/Extraction.Text/generate-data.py

@@ -57,7 +57,9 @@ def gather_data(rootlength, prog, path, base):
         else:
 	        result=datum
     print(result)
-    current_data["Result"]=result
+    if result == "NothingLearned" or result == "TooLargeSpec":
+        result = -1
+    current_data["FlashExtract"]=result
     current_data["ExsTaken"]=num_inputs
     return current_data
 

### 2017-09-29 comparisons/prose/Transformation.Text/generate-data.py

@@ -58,10 +58,14 @@ def gather_data(rootlength, prog, path, base):
                 result="-1"
         elif (num_inputs > 100000):
             result="-1"
+        elif result == "NothingLearned":
+            result="-1"
         else:
 	        result=datum
+    if result == "NothingLearned":
+        result="-1"
     print(result)
-    current_data["Result"]=result
+    current_data["FlashFill"]=result
     current_data["ExsTaken"]=num_inputs
     counter = counter+1
     print_data([current_data],str(counter));

### 2017-09-29 program/generate-data.py

@@ -83,44 +83,44 @@ def gather_data(rootlength, prog, path, base):
 	    return int(sum(example_number_col)/len(example_number_col))
 
 
-    #gather_col(['-time'],ctime_combiner,"ComputationTime",TIMEOUT_TIME,REPETITION_COUNT)
-    ##gather_col(['-forceexpand','-time'],ctime_combiner,"ForceExpandTime",TIMEOUT_TIME,REPETITION_COUNT)
-    #gather_col(['-naive_strategy','-time'],ctime_combiner,"NaiveStrategy",TIMEOUT_TIME,REPETITION_COUNT)
-    ##gather_col(['-naive_pqueue','-time'],ctime_combiner,"NaivePQueue",TIMEOUT_TIME,REPETITION_COUNT)
-    ##gather_col(['-no_short_circuit','-time'],ctime_combiner,"NoShortCircuit",TIMEOUT_TIME,REPETITION_COUNT)
-    #gather_col(['-no_lens_context','-time'],ctime_combiner,"NoLensContext",TIMEOUT_TIME,REPETITION_COUNT)
-    ##gather_col(['-no_short_circuit','-no_inferred_expansions','-no_lens_context','-time'],ctime_combiner,"NoInferenceNoLCNoSC",TIMEOUT_TIME,REPETITION_COUNT)
-    ##gather_col(['-no_short_circuit','-no_lens_context','-time'],ctime_combiner,"NoLCNoSC",TIMEOUT_TIME,REPETITION_COUNT)
-    #gather_col(['-naive_expansion_search','-no_lens_context','-time'],ctime_combiner,"NaiveExpansionNoLC",TIMEOUT_TIME,REPETITION_COUNT)
-    #gather_col(['-use_only_forced_expansions','-no_lens_context','-time'],ctime_combiner,"OnlyForcedExpansionsNoLC",TIMEOUT_TIME,REPETITION_COUNT)
-    #gather_col(['-naive_expansion_search','-time'],ctime_combiner,"NaiveExpansion",TIMEOUT_TIME,REPETITION_COUNT)
-    #gather_col(['-use_only_forced_expansions','-time'],ctime_combiner,"OnlyForcedExpansions",TIMEOUT_TIME,REPETITION_COUNT)
[…]

### 2017-09-29 program/generate-data.py
add generate data stuff

@@ -127,9 +127,6 @@ def gather_data(rootlength, prog, path, base):
 def specsize_compare(x,y):
     return int(x["SpecSize"])-int(y["SpecSize"])
 
-def sort_data(data):
-    return sorted(data,cmp=specsize_compare)
-
 def print_data(data):
     ensure_dir("generated_data/")
     with open("generated_data/data.csv", "wb") as csvfile:
@@ -153,21 +150,32 @@ def transform_data(path, base, run_data):
             current_data[col_name] = str(sum(col)/len(col))
     return current_data
 
+def load_data():
+    try:
+        with open("generated_data/data.csv", "r") as csvfile:
+            datareader = csv.DictReader(csvfile)
+            return [row for row in datareader]
+    except:
+        return []
+
 def main(args):
     if len(args) == 3:
         prog = args[1]
         path = args[2]
         rootlength = len(path)
-        data = []
+        data = load_data()
         if not os.path.exists(prog):
             print_usage(args)
         elif os.path.exists(path) and os.path.isdir(path):
             for path, base in find_tests(path):
-                print(join(path, base + TEST_EXT).replace("_","-")[rootlength:])
-                current_data = gather_data(rootlength,prog, path, base)
-                data.append(current_data)
-            data = sort_data(data)
-	    print_data(data)
+                test_name = join(path, base).replace("_","-")[rootlength:]
+                print(test_name)
+                if (not (any(row["Test"] == test_name for row in data))):
+                    current_data = gather_data(rootlength,prog, path, base)
+                    data.append(current_data)
+                else:
+                    print("data already retrieved")
+                print_data(data)
         else:
             path, filename = os.path.split(path)
             base, ext = splitext(filename)
@@ -175,8 +183,7 @@ def main(args):
                 print_usage(args)
             else:
                 data = gather_data(prog, path, base)
-                sort_data(data)
-		print_data([data])
+                print_data([data])
     else:
         print_usage(args)
 

### 2017-09-29 comparisons/prose/Extraction.Text/generate-data.py
make comparisons rerunnable

@@ -86,20 +86,32 @@ def transform_data(path, base, run_data):
             current_data[col_name] = str(sum(col)/len(col))
     return current_data
 
+def load_data():
+    try:
+        with open("generated_data/data.csv", "r") as csvfile:
+            datareader = csv.DictReader(csvfile)
+            return [row for row in datareader]
+    except:
+        return []
+
 def main(args):
     if len(args) == 3:
         prog = args[1]
         path = args[2]
         rootlength = len(path)
-        data = []
+        data = load_data()
         if not os.path.exists(prog):
             print_usage(args)
         elif os.path.exists(path) and os.path.isdir(path):
             for path, base in find_tests(path):
-                print(join(path, base + TEST_EXT).replace("_","-")[rootlength:])
-                current_data = gather_data(rootlength,prog, path, base)
-                data.append(current_data)
-            print_data(data)
+                test_name = join(path, base).replace("_","-")[rootlength:]
+                print(test_name)
+                if (not (any(row["Test"] == test_name for row in data))):
+                    current_data = gather_data(rootlength,prog, path, base)
+                    data.append(current_data)
+                else:
+                    print("data already retrieved")
+                print_data(data)
         else:
             print_usage(args)
     else:

### 2017-09-29 comparisons/prose/Transformation.Text/generate-data.py

@@ -68,12 +68,11 @@ def gather_data(rootlength, prog, path, base):
     current_data["FlashFill"]=result
     current_data["ExsTaken"]=num_inputs
     counter = counter+1
-    print_data([current_data],str(counter));
     return current_data
 
-def print_data(data,file_addition):
+def print_data(data):
     ensure_dir("generated_data/")
-    with open("generated_data/data" + file_addition + ".csv", "wb") as csvfile:
+    with open("generated_data/data" + ".csv", "wb") as csvfile:
 	datawriter = csv.DictWriter(csvfile,fieldnames=data[0].keys())
 	datawriter.writeheader()
 	datawriter.writerows(data)
@@ -94,20 +93,32 @@ def transform_data(path, base, run_data):
             current_data[col_name] = str(sum(col)/len(col))
     return current_data
 
+def load_data():
+    try:
+        with open("generated_data/data.csv", "r") as csvfile:
+            datareader = csv.DictReader(csvfile)
+            return [row for row in datareader]
+    except:
+        return []
+
 def main(args):
     if len(args) == 3:
         prog = args[1]
         path = args[2]
         rootlength = len(path)
-        data = []
+        data = load_data()
         if not os.path.exists(prog):
             print_usage(args)
         elif os.path.exists(path) and os.path.isdir(path):
             for path, base in find_tests(path):
-                print(join(path, base + TEST_EXT).replace("_","-")[rootlength:])
-                current_data = gather_data(rootlength,prog, path, base)
-                data.append(current_data)
-            print_data(data,"")
+                test_name = join(path, base).replace("_","-")[rootlength:]
+                print(test_name)
+                if (not (any(row["Test"] == test_name for row in data))):
+                    current_data = gather_data(rootlength,prog, path, base)
+                    data.append(current_data)
+                else:
+                    print("data already retrieved")
+                print_data(data)
         else:
             print_usage(args)
     else:

### 2017-10-03 papers/QuotientLensSynth/main.tex
Added intro to draft

@@
 %
-\title{A Synthesis Ready Langauge for Expressing Quotient Lenses}
+\title{Synthesizing Quotient Lenses}
 %
-\titlerunning{}  % abbreviated title (for running head)
-%                                     also used for the TOC unless
+\titlerunning{Synthesizing Quotient Lenses}  % abbreviated title (for running
+% head)                                     also used for the TOC unless
 %                                     \toctitle is used
@@
 \section{Introduction}
-TODO Introduce BibTex to Endnote example as motivation.
+String quotient lenses enable one to express bidirectional string
+transformations that obey the lens laws up to equivalence relations. However,
+the various approaches that have been proposed to express quotient lenses
+make it difficult for the programmer to express complicated equivalence
+relations over the data.\\
+For example, when converting \textsc{Bib}\TeX files to EndNote files, the data
+may contain arbitrary amounts of whitespace characters in between the actual
+data. In addition, the fields may be ordered in an arbitrary order. For
+instance, we may wish to convert this \textsc{Bib}\TeX entry 
+\\
+\newline
+@Book\{conway,\\ 
+  Author = \{Conway, J. H.\},\\
+  Title = \{Regular Algebra and Finite Machines\},\\
+  Publisher = \{Printed in GB by William Clowes \& Sons Ltd\},\\
+  Year = \{1971\}\\
+\}\\
+\newline
+to this EndNote entry,\\
+\newline
+\%0 Book\\
+\%T Regular Algebra and Finite Machines\\
+\%A Conway, J. H.\\
+\%D 1971\\
+\%I Printed in GB by William Clowes \& Sons Ltd\\
+\%F conway\\
+\newline
+while allowing for this equivalent representation of the \textsc{Bib}\TeX
+\; entry :\\
+\newline
+@Book\{conway,\\ 
+    Publisher = \{Printed in GB by William Clowes \& Sons Ltd\},\\
+    \hspace*{10mm} Title = \{Regular Algebra and Finite Machines\},\\
+    \newline
+    Year = \{1971\},\\
+    \hspace*{20mm} Author = \{Conway, J. H.\},\\
+    \newline
+\}\\
+\newline
+With the existing tools, it is cumbersome to express this, and other complicated
+equivalence relations.\\
+In this paper, we introduce the formalism of Quotient Regular Expressions (or
+QREs) to solve this problem. QREs enable the programmer to express an
+equivalence relation at internal terms of a regular expression, thus giving
+the programmer more control over where quotienting occurs in the data.\\
+For example, the fields of a \textsc{Bib}\TeX\; article like in the example
+above might be expressed by the following QRE:
+$$perm(REF, AUTHOR, TITLE, PUBLISHER, YEAR) \; with \; (WSP^* \mapsto \string
+\n),$$ where $REF, AUTHOR, TITLE, PUBLISHER, YEAR$ are regular expressions
+describing the respective \textsc{Bib}\TeX\; fields of the same name. The QRE of
+an EndNote article is similarly expressed, but with the regular expressions for
+the respective fields substituted with the appropriate EndNote equivalent.\\
+\textcolor{red}{[Mention and give examples of quotient lenses here?]}
 \subsection{Contributions}
@@
   \item
-  We give an algorithm for synthesizing quotient lenses in
-  $\mathcal{Q}$ from their QRE types and a set of examples.
+  Using the normal form for quotient lenses in $\mathcal{Q}$, we reduce the
+  synthesis of quotient lenses in $\mathcal{Q}$ to the synthesis of bijective
+  lenses between regular languages. 
   \item
   We describe an implementation of QREs, the language $\mathcal{Q}$ of quotient
-  lenses, and the algorithm for synthesizing quotient lenses in $\mathcal{Q}$ in
+  lenses, and an algorithm for synthesizing quotient lenses in $\mathcal{Q}$ in
   the Boomerang language.
+  \item
+  We assess the usabibility of the QRE synthesis via a case study in which we
+  selected several data formats used by data.gov and synthesized quotient lenses
+  between them.
 \end{enumerate}
@@
 grammar:
-$$q := id(R) \; | \; R \mapsto s \; | \; squash (R, R', f) \; | \; perm \;
-(R_1, \ldots, R_n) \; with \; R \; | \; q' \circ q \; | \; q \cdot q' \; | \;
-(q \; | \; q') \; | \; q^*,$$
+\begin{align*}
+q := \; &id(R) \; | \; q \mapsto s \; | \; squash (R, R', f) \; | \;
+perm \; (q_1, \ldots, q_n) \; with \; q \; | \; normalize(R, R', f)\\
+&q' \circ q \; | \; q \cdot q' \; | \; (q \; | \; q') \; | \; q^*,
+\end{align*}
 where $R$ ranges over regular expressions, $f$ ranges over functions whose
@@
 \subsection{Semantics of QREs}
+\textcolor{red}{[Informal Meanings/Example Usages Needed]}\\
 We now give the semantics of QREs as well as inference rules for deriving them:
@@
 \end{center}
-\item $q=R \mapsto s$
+\item $q=q \mapsto s$
 \begin{center}
 \begin{enumerate}
-  \item $W(R \mapsto S) = R$,
-  \item $K(R \mapsto s) = s$, 
+  \item $W(q \mapsto s) = W(q)$,
+  \item $K(q \mapsto s) = s$, 
   \item
-For all $w, w' \in W(R \mapsto S)$, $w \; EqRel(R \mapsto S) \; w'$,
+For all $w, w' \in W(q \mapsto S)$, $w \; EqRel(q \mapsto S) \; w'$,
 \item
-$Canonizer(R \mapsto S)(r) = s$
+$Canonizer(q \mapsto S)(r) = s$
 \end{enumerate}
 \begin{prooftree}
-\AxiomC{$s \in \mathcal{L}(R)$}
-\UnaryInfC{$R \mapsto s$ is well formed}
+\AxiomC{$s \in W(q)$}
+\UnaryInfC{$q \mapsto s$ is well formed}
 \end{prooftree}
@@
 \end{center}
-\item $q=perm \; (R_1, \ldots, R_n) \; with \; S$
+\item $q=perm \; (q_1, \ldots, q_n) \; with \; q$
 \begin{center}
 \begin{enumerate}
-  \item $W(perm \; (R_1, \ldots, R_n) \; with \; S) = \bigcup \limits_{\sigma
-  \in S_n} R_{\sigma(1)} \cdot S \cdot \ldots \cdot \cdot S \cdot R_{\sigma(n)}$
-,
-  \item $K(perm \; (R_1, \ldots, R_n) \; with \; S) = R_{1} \cdot S
-  \cdot \ldots \cdot \cdot S \cdot R_{n}$, 
+  \item $W(perm \; (q_1, \ldots, q_n) \; with \; q) = \bigcup \limits_{\sigma
+  \in S_n} W(q_{\sigma(1)}) \cdot W(q) \cdot \ldots \cdot \cdot W(q) \cdot
+  W(q_{\sigma(n)}),$
+  \item $K(perm \; (q_1, \ldots, q_n) \; with \; q) = K(q_1) \cdot K(q_1)
+  \cdot \ldots \cdot \cdot K(q) \cdot K(q_n)$, 
   \item
@@
    $$w = r_{\sigma(1)} \cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot
-    r_{\sigma(n)} \text{ and } r' = {r}_{\theta(1)} \cdot s_1 \cdot \ldots
-    \cdot s_{n-1} \cdot {r}_{\theta(n)}$$ for some $\sigma, \theta \in
-    S_n$, where $r_i \in R_i$ and $s_j \in S$.
+    r_{\sigma(n)} \text{ and } r' = {r'}_{\theta(1)} \cdot s'_1 \cdot \ldots
+    \cdot s'_{n-1} \cdot {r'}_{\theta(n)}$$ for some $\sigma, \theta \in
+    S_n$, with $r_i \; EqRel(q_i) \; r'_i$ and $s_k \; EqRel(q) \; s'_{k}$.
 \item
 $Canonize(q)(r_{\sigma(1)}
-\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)})
-= r_1 \cdot s_1 \cdot \ldots \cdot \cdot s_{n-1}\cdot r_n$
+\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \newline
+= (Canonize(q_1)(r_1)) \cdot (Canonize(q)(s_1)) \cdot \ldots \cdot \cdot \cdot
+(Canonize(q_n)(r_n))$
 \end{enumerate}
@@
 \end{prooftree}
+\end{center}
+
+\item $q = normalize(R, R', f)$
+\begin{center}
+\begin{enumerate}
+  \item
+$W(normalize(f)) = R$
+\item
+$K(normalize(f)) = R'$
+\item For all $w, w' \in W(q)$, $w \; EqRel(q) \; w'$ if and only if
+$f(w)=f(w')$
+\item
+$Canonize(normalize(f)) = f$
+\begin{prooftree}
+\AxiomC{$\mathcal{L}(R') \subseteq \mathcal{L}(R)$}
+\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(R')$}
+\AxiomC{$f$ is surjective}
+\AxiomC{$f = f^2$}
+\QuaternaryInfC{$normalize \; (R,R', f)$ is well formed}
+\end{prooftree}
+\end{enumerate}
 \end{center}
@@
   $k$ in $\mathcal{L}(K(q))$ such that $k$ is equivalent to $w$ modulo
-  $EqRel(q)$
+  $EqRel(q)$.
   \end{enumerate}
@@
 \end{align*}
-[Reference needed] Define the class of \textit{bijective lenses} to be the set
-of bijections between regular languages created from Boomerang lens combinators.
+Define the class of \textit{bijective
+lenses} to be the set of bijections between regular languages created from Boomerang lens combinators.
 The syntax for the language of bijective lenses is given by
@@
 \end{align*}
+\textcolor{red}{[More explanation needed esp. inference rules for bijective
+lenses]}\\
 Having described the class of bijective lenses, we now define the class
@@
 quotient lens $\llbracket q \rrbracket :
-W(c)/{EqRel(c)} \Longleftrightarrow W(c')/{EqRel(c')}$. We now give the
+W(c)/{EqRel(c)} \Longleftrightarrow W(c')/{EqRel(c')}$.\\
+\textcolor{red}{[Informal Meaning/Usages Needed]}\\We now give the
 semantics of $\mathcal{Q}$, as well as inference rules that may be used to
@@
 \subsection{Normal Forms for Quotient Lenses in $\mathcal{Q}$}
-\begin{claim}
+\begin{theorem}\label{normal form}
 If there is a derivation $q : c \Leftrightarrow c'$ in
@@
 \end{align*}
-\end{claim}
+\end{theorem}
 \begin{proof}
@@
 \section{Synthesizing Quotient Lenses}
+By \cref{normal form}, if there is a derivation $q : c \Leftrightarrow c'$ in
+$\mathcal{Q}$ then there exists a bijective lens $\ell : K(c) \Leftrightarrow
+K(c')$ such that
+\begin{align*}
+\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ Canonizer(c)\\
+\llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
+Canonizer(c')
+\end{align*}
+In other words, every quotient lens in $\mathcal{Q}$ is the same as an ordinary
+bijective lens with canonizers at the ends.\\
+Therefore, in order to synthesize a quotient lens $q: c \Leftrightarrow c'$
+where $c, c'$ are QREs, it suffices to synthesize a bijective lens $\ell : K(c)
+\Leftrightarrow K(c')$.\\
+\textcolor{red}{[Citation Needed!]} has already shown that if there
+exists a derivation $\ell : R \Leftrightarrow S$, then there exist
+regular expressions $R', S'$ and a synthesizable \\
+\textcolor{red}{[Explain and make explicit the meaning
+of synthesizable]} \\
+bijective lens $\ell' :
+R' \Leftrightarrow S'$ such that $\mathcal{L}(R) = \mathcal{L}(R')$,
+$\mathcal{L}(S) = \mathcal{L}(S')$ and $\llbracket \ell \rrbracket =
+\llbracket \ell' \rrbracket$. This immediately implies that if there is a
+derivation $q : c \Leftrightarrow c'$ as well as a derivation $\ell : K(c)
+\Leftrightarrow K(c')$, then $q$ is synthesizable.
+\textcolor{red}{[More Needed?]}
+\section{Implementation and Evaluation}
 \section{Related Work}

### 2017-10-03 papers/QuotientLensSynth/main.tex
Added explanation for QREs and also notion of unambiguity with related syntax

@@
 \section{Quotient Regular Expressions}
-The language of Quotient Regular Expressions (QREs) is given by the following
-grammar:
-\begin{align*}
-q := \; &id(R) \; | \; q \mapsto s \; | \; squash (R, R', f) \; | \;
-perm \; (q_1, \ldots, q_n) \; with \; q \; | \; normalize(R, R', f)\\
-&q' \circ q \; | \; q \cdot q' \; | \; (q \; | \; q') \; | \; q^*,
-\end{align*}
-where $R$ ranges over regular expressions, $f$ ranges over functions whose
-domain and codomain are regular languages, and $s$ ranges over character
-strings.\\
-Each QRE $q$ enables us to express
+Quotient Regular Expressions (or QREs) enable us to give richer specifications
+for data formats than ordinary regular expressions. More concretely, each
+Quotient Regular Expression (or QRE) $q$ enables us to express
 \begin{enumerate}
@@
   \end{enumerate}
-Observe that since $Canonizer(q)$ computes
-$Canonizer(q)(w)$ as the unique $k$ in $\mathcal{L}(K(q))$ such that $w$ is
-equivalent to $k$ mod $EqRel(q)$, then the equivalence classes of $EqRel(q)$
-are the same as the fibres of $Canonizer(q)$.
+  All data to be transformed must match $W(q)$, the ``whole'' regular
+  expression of $q$. However, before transforming the using a lens, the data
+  will be quotiented out by the equivalence relation $EqRel(q)$ which has $K(q)$
+  as a complete set of representatives. That is, for each string $w \in
+  W(q)$, there is a unique $k \in K(q)$ such that $w \; EqRel(q) \; k$). This
+  unique representative $k$ is chosen using the ``canonizing'' function
+  $Canonizer(q)$.\\
+  Observe that since $Canonizer(q)$ computes $Canonizer(q)(w)$ as the unique $k$
+  in $\mathcal{L}(K(q))$ such that $w$ is equivalent to $k$ mod $EqRel(q)$,
+  then the equivalence classes of $EqRel(q)$ are the same as the fibres of
+  $Canonizer(q)$.
+  \subsection{Syntax of QREs}
+The language of Quotient Regular Expressions (QREs) is given by the following
+grammar:
+\begin{align*}
+q := \; &R \; | \; q \mapsto s \; | \; squash (R, R', f) \; | \;
+perm \; (q_1, \ldots, q_n) \; with \; q \; | \; normalize(R, R', f)\\
+&q' \circ q \; | \; q \cdot q' \; | \; (q \; | \; q') \; | \; q^*,
+\end{align*}
+where $R$ ranges over regular expressions, $f$ ranges over functions whose
+domain and codomain are regular languages, and $s$ ranges over character
+strings.
 \subsection{Semantics of QREs}
+\subsubsection{Preliminaries}
+The transformations that will be performed on the data will require
+that the data is matched against regular expressions. Consequently, we require
+that regular expressions used by QREs be \textit{strongly unambiguous}, a
+condition which ensures that if a string $s$ matches a regular expression $R$,
+then $s$ matches $R$ uniquely.\\
+To this end, we say that regular expressions $R$ and $S$ are
+\textit{unambiguosly concatenable}, written $R \cdot^! S$ if for all strings
+$r, r' \in \mathcal{L}(R)$ and $s, s' \in \mathcal{L}(S)$, if $r \cdot s = r'
+\cdot s'$, then $r = r'$ and $s = s'$. We say that a regular expression $R$ is
+\textit{unambiguosly iterable}, written $R^{*!}$ if for all strings $r_1,
+\ldots, r_m$ and $r'_1, \ldots, r'_n \in \mathcal{L}(R)$, if $r_1 \cdot \ldots
+\cdot r_m = r'_1 \cdot \ldots \cdot r'_n$, then $m = n$ and $r_i = r'_i$.\\
+We say that a regular expression $R$ is \textit{strongly unambiguous} if and
+only if (1) $R = \varnothing$, or (2) $R = S_1 \cdot S_2$ with $S_1, S_2$
+strongly unambiguous and $S_1 \cdot^! S_n$, or (3) $R = S_1 \; | \; S_2$ with
+$S_1, S_2$ strongly unambiguous and $\mathcal{L}(S_1) \cap \mathcal{L}(S_2) =
+\varnothing$, or (4) $R = S^*$ with $S$ strongly unambiguous and $S^{*!}$.\\
+We now give the semantics of QREs as well as inference rules for deriving them:
 \textcolor{red}{[Informal Meanings/Example Usages Needed]}\\
-We now give the semantics of QREs as well as inference rules for deriving them:
 \begin{enumerate}
-  \item $q = id(R)$
+  \item $q = R$
 \begin{center}
 \begin{enumerate}
-  \item $W(id(R)) = R$,
-  \item $K(id(R)) = R$, 
+  \item $W(R) = R$,
+  \item $K(R) = R$, 
   \item
-For all $w, w' \in W(id(R))$, $w \; EqRel(q) \; w'$ if and only if $w =
+For all $w, w' \in W(R)$, $w \; EqRel(q) \; w'$ if and only if $w =
 w'$.
 \item
-$Canonizer(id(R))(r) = r$
+$Canonizer(R)(r) = r$
 \end{enumerate}
@@
 \AxiomC{}
-\UnaryInfC{$id(R)$ is well formed}
+\UnaryInfC{$R$ is well formed}
 \end{prooftree}
@@
 \end{center}
-\item $q=perm \; (q_1, \ldots, q_n) \; with \; q$
+\item $q=perm \; (q_1, \ldots, q_n) \; with \; q$. Given regular expressions
+$R_1, \ldots, R_n$ and $S$, let $\prod({R_i}, S) = R_1 \cdot S \cdot \ldots
+\cdot S \cdot R_n$.
 \begin{center}
@@
 \begin{prooftree}
-\AxiomC{$\forall \sigma \neq \theta, \; R_{\sigma(1)} \cdot S \cdot \ldots \cdot
-S \cdot R_{\sigma(n)} \cap R_{\theta(1)} \cdot S \cdot \ldots \cdot S \cdot
-R_{\theta(n)} = \varnothing$} \AxiomC{$\forall \sigma, \; R_{\sigma(1)} \cdot^!
-S \cdot^! \ldots \cdot^! S \cdot^! R_{\sigma(n)}$}
-\BinaryInfC{$permute \; (R_1, \ldots, R_n) \; with \;  S$ is well formed}
+\AxiomC{$\forall \sigma \neq \theta, \; \prod(\{W(q_{\sigma(i)})\}, W(q))\cap
+ \prod(\{W(q_{\theta(i)})\}, W(q)) = \varnothing$}
+\AxiomC{$\forall \sigma, \; K(q_{\sigma(1)}) \cdot^! K(q) \cdot^! \ldots 
+\cdot^! K(q_{\sigma(n)})$}
+\BinaryInfC{$permute \; (q_1, \ldots, q_n) \; with \;  q$ is well formed}
 \end{prooftree}
@@
 The syntax for the language of bijective lenses is given by
-$$\ell := id(R) \; | \; const(s, s') \; | \;  swap(\ell,
+$$\ell := R \; | \; const(s, s') \; | \;  swap(\ell,
 \ell) \; | \; \ell \cdot \ell' \; |  \; (\ell \; | \; \ell') \; | \; \ell^* \;
@@
 \begin{align*}
-\llbracket id(R) \rrbracket &= \{(r, r) \; | \; r \in \mathcal{L}(R)\}\\
+\llbracket R \rrbracket &= \{(r, r) \; | \; r \in \mathcal{L}(R)\}\\
 \llbracket swap(\ell, \ell') \rrbracket &= \{(s \cdot t, t' \cdot s') \; | \;
@@
 \AxiomC{$\ell : R \Leftrightarrow S$}
-\UnaryInfC{$id(\ell): id(R) \Leftrightarrow id(S)$}
+\UnaryInfC{$id(\ell): R \Leftrightarrow id(S)$}
 \end{prooftree}
@@
   \item
-  $id(\ell): id(R) \Leftrightarrow id(S)$ where $\ell : R \Leftrightarrow
+  $id(\ell): R \Leftrightarrow id(S)$ where $\ell : R \Leftrightarrow
 S$. Then
@@
   \llbracket id(\ell) \rrbracket.get &=  \llbracket \ell \rrbracket = \llbracket \ell \rrbracket \circ
-  id_{\mathcal{L}(R)} = \llbracket \ell \rrbracket \circ Canonizer(id(R)), \text{ and }\\
+  id_{\mathcal{L}(R)} = \llbracket \ell \rrbracket \circ Canonizer(R), \text{ and }\\
   \llbracket id(\ell) \rrbracket.put &= \llbracket \ell \rrbracket^{-1} = \llbracket \ell \rrbracket^{-1} \circ

### 2017-10-03 papers/QuotientLensSynth/main.tex
some comments for Solomon

@@
+\newif\ifdraft\drafttrue  % set true to show comments
+
 \documentclass{svproc}
@@
 mathrsfs, mathtools, bussproofs, stmaryrd, enumitem, stackengine}
+
+%%%%%
+% Macros
+% Colors
+\definecolor{dkblue}{rgb}{0,0.1,0.5}
+\definecolor{dkgreen}{rgb}{0,0.6,0}
+\definecolor{dkred}{rgb}{0.6,0,0}
+\definecolor{dkpurple}{rgb}{0.7,0,0.4}
+\definecolor{olive}{rgb}{0.4, 0.4, 0.0}
+\definecolor{teal}{rgb}{0.0,0.5,0.5}
+\definecolor{orange}{rgb}{0.9,0.6,0.2}
+\definecolor{lightyellow}{RGB}{255, 255, 179}
+\definecolor{lightgreen}{RGB}{170, 255, 220}
+\definecolor{teal}{RGB}{141,211,199}
+\definecolor{darkbrown}{RGB}{121,37,0}
+
+\newcommand{\FINISH}[3]{\ifdraft\textcolor{#1}{[#2: #3]}\fi}
+\newcommand{\bcp}[1]{\FINISH{dkred}{B}{#1}}
+\newcommand{\BCP}[1]{\FINISH{dkred}{B}{\bf #1}}
+\newcommand{\afm}[1]{\FINISH{dkgreen}{A}{#1}}
+\newcommand{\dpw}[1]{\FINISH{dkblue}{D}{#1}}
+\newcommand{\saz}[1]{\FINISH{orange}{SZ}{#1}}
+\newcommand{\ksf}[1]{\FINISH{teal}{K}{#1}}
+\newcommand{\sm}[1]{\FINISH{dkpurple}{SM}{#1}}
+
+% FOR Regular Expression names
+\newcommand{\re}[1]{\ensuremath{\mathit{#1}}}
 
@@
 \section{Introduction}
+\saz{We need a more gentle ramp-up to the example here.}
+
+\begin{itemize}
+\item Lens programming languages [e.g. Boomerang] are designed to make it possible to express
+  bidirectional data transformations using a single program.
+
+\item Lenses have found practical application in Augeas \saz{and ...}
+
+\item The source and target types of a lens express constraints on the format of
+  the data.
+
+\item Lens programming is still quite tricky, because lens composition operators
+  are subject to side conditions (e.g. unique decomposition) that are not so
+  easy to reason about.
+
+\item Our prior work~\cite{popl18} has shown how to synthesize lenses for the
+  \textit{bijective} case.
+
+\item This still isn't as expressive as one would want in practice, because many
+  lens transformations aren't truly bijective.
+
+\item There are at least two challenges with programming quotient lenses: (1)
+  defining the quotient languages (which serve as the types of the lenses)
+  themselves is not so easy, and (2) the lenses themselves are still not easy to
+  write by hand.
+
+\end{itemize}
+
 String quotient lenses enable one to express bidirectional string
@@
 make it difficult for the programmer to express complicated equivalence
-relations over the data.\\
-For example, when converting \textsc{Bib}\TeX files to EndNote files, the data
+relations over the data.
+
+For example, when converting \textsc{Bib}\TeX{} files to EndNote files, the data
 may contain arbitrary amounts of whitespace characters in between the actual
 data. In addition, the fields may be ordered in an arbitrary order. For
-instance, we may wish to convert this \textsc{Bib}\TeX entry 
-\\
-\newline
-@Book\{conway,\\ 
-  Author = \{Conway, J. H.\},\\
-  Title = \{Regular Algebra and Finite Machines\},\\
-  Publisher = \{Printed in GB by William Clowes \& Sons Ltd\},\\
-  Year = \{1971\}\\
-\}\\
-\newline
-to this EndNote entry,\\
-\newline
-\%0 Book\\
-\%T Regular Algebra and Finite Machines\\
-\%A Conway, J. H.\\
-\%D 1971\\
-\%I Printed in GB by William Clowes \& Sons Ltd\\
-\%F conway\\
-\newline
-while allowing for this equivalent representation of the \textsc{Bib}\TeX
-\; entry :\\
-\newline
-@Book\{conway,\\ 
-    Publisher = \{Printed in GB by William Clowes \& Sons Ltd\},\\
-    \hspace*{10mm} Title = \{Regular Algebra and Finite Machines\},\\
-    \newline
-    Year = \{1971\},\\
-    \hspace*{20mm} Author = \{Conway, J. H.\},\\
-    \newline
-\}\\
-\newline
+instance, we may wish to convert this \textsc{Bib}\TeX{} entry: \saz{I propose
+  we display such examples in verbatim:}
+
+\begin{verbatim}
+@Book {conway,
+  Author = {Conway, J. H.},
+  Title = {Regular Algebra and Finite Machines},
+  Publisher = {Printed in GB by William Clowes & Sons Ltd},
+  Year = {1971}
+}
+\end{verbatim}
+
+to this EndNote entry:
+
+\begin{verbatim}
+%0 Book
+%T Regular Algebra and Finite Machines
+%A Conway, J. H.
+%D 1971
+%I Printed in GB by William Clowes & Sons Ltd
+%F conway
+\end{verbatim}
+
+\noindent while allowing for this equivalent representation of the \textsc{Bib}\TeX{}
+entry, which rearranges the fields and uses different  whitespace conventions:
+
+\begin{verbatim}
+@Book{conway, 
+    Publisher = {Printed in GB by William Clowes & Sons Ltd},
+        Title = {Regular Algebra and Finite Machines},
+
+         Year = {1971},
+       Author = {Conway, J. H.},
+}
+\end{verbatim}
+
 With the existing tools, it is cumbersome to express this, and other complicated
-equivalence relations.\\
+equivalence relations.  \saz{It might be a good idea to show what this would
+  look like in all its gory detail, just to make the problem very
+  concrete. Perhaps in a figure?}
+
 In this paper, we introduce the formalism of Quotient Regular Expressions (or
@@
 the programmer more control over where quotienting occurs in the data.\\
-For example, the fields of a \textsc{Bib}\TeX\; article like in the example
+For example, the fields of a \textsc{Bib}\TeX{} article like in the example
 above might be expressed by the following QRE:
-$$perm(REF, AUTHOR, TITLE, PUBLISHER, YEAR) \; with \; (WSP^* \mapsto \string
-\n),$$ where $REF, AUTHOR, TITLE, PUBLISHER, YEAR$ are regular expressions
-describing the respective \textsc{Bib}\TeX\; fields of the same name. The QRE of
-an EndNote article is similarly expressed, but with the regular expressions for
-the respective fields substituted with the appropriate EndNote equivalent.\\
-\textcolor{red}{[Mention and give examples of quotient lenses here?]}
+
+$$perm(\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER}, \re{YEAR}) \; with \; (\re{WSP}^* \mapsto \string
+\n),$$
+
+\noindent where $\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},$ and $\re{YEAR}$ are
+regular expressions describing the respective \textsc{Bib}\TeX\; fields of the
+same name. The QRE of an EndNote article is similarly expressed, but with the
+regular expressions for the respective fields substituted with the appropriate
+EndNote equivalent.  \saz{Do we really mean all \re{WSP} maps to newline here?
+  What about the spaces in the title, etc.?  Would this be confusing without
+  more explanation?}
+
+\sm{[Mention and give examples of quotient lenses here?]}
+
 \subsection{Contributions}
@@
 \section{Quotient Regular Expressions}
+
+\saz{Need a transition here:  This section introduces QREs.  They are regular
+  expressions augmented with syntax that lets them simultaneously denote
+  quotient languages.  
+   }
+
+  \paragraph{Syntax of QREs}
+The language of Quotient Regular Expressions (QREs) is given by the following
+grammar:
+\begin{align*}
+q := \; &R \; | \; q \mapsto s \; | \; squash (R, R', f) \; | \;
+perm \; (q_1, \ldots, q_n) \; with \; q \; | \; normalize(R, R', f)\\
+&q' \circ q \; | \; q \cdot q' \; | \; (q \; | \; q') \; | \; q^*,
+\end{align*}
+where $R$ ranges over regular expressions, $f$ ranges over functions whose
+domain and codomain are regular languages, and $s$ ranges over character
+strings.
+
+\saz{  Every regular expression $R$ is a QRE.  The main new syntactic forms are: $q \mapsto s$, which
+}
+
+\saz{Would it be simpler to include atoms and $\epsilon$ in $q$ rather than $R$?  That is, $R$ is
+  just a subset of $QRE$ built from atoms, $\cdot$ $|$ and ${}^*$.  We could
+  reserve $R$ as a metavariable that ranges over $q$'s such that $q$ is a
+  regular expression.}
+
+\saz{Need to introduce $\mathcal{L}(R)$ notation.}
+
 Quotient Regular Expressions (or QREs) enable us to give richer specifications
@@
   $Canonizer(q)$.
-  \subsection{Syntax of QREs}
-The language of Quotient Regular Expressions (QREs) is given by the following
-grammar:
-\begin{align*}
-q := \; &R \; | \; q \mapsto s \; | \; squash (R, R', f) \; | \;
-perm \; (q_1, \ldots, q_n) \; with \; q \; | \; normalize(R, R', f)\\
-&q' \circ q \; | \; q \cdot q' \; | \; (q \; | \; q') \; | \; q^*,
-\end{align*}
-where $R$ ranges over regular expressions, $f$ ranges over functions whose
-domain and codomain are regular languages, and $s$ ranges over character
-strings.
 \subsection{Semantics of QREs}
@@
 condition which ensures that if a string $s$ matches a regular expression $R$,
-then $s$ matches $R$ uniquely.\\
+then $s$ matches $R$ uniquely.
+
+\saz{Solomon: you should use blank lines instead of backslashes to separate
+  paragraphs so latex can format them properly.}
+
 To this end, we say that regular expressions $R$ and $S$ are
@@
 \ldots, r_m$ and $r'_1, \ldots, r'_n \in \mathcal{L}(R)$, if $r_1 \cdot \ldots
-\cdot r_m = r'_1 \cdot \ldots \cdot r'_n$, then $m = n$ and $r_i = r'_i$.\\
+\cdot r_m = r'_1 \cdot \ldots \cdot r'_n$, then $m = n$ and $r_i = r'_i$.
+
 We say that a regular expression $R$ is \textit{strongly unambiguous} if and
@@
 $S_1, S_2$ strongly unambiguous and $\mathcal{L}(S_1) \cap \mathcal{L}(S_2) =
-\varnothing$, or (4) $R = S^*$ with $S$ strongly unambiguous and $S^{*!}$.\\
+\varnothing$, or (4) $R = S^*$ with $S$ strongly unambiguous and $S^{*!}$.
+
 We now give the semantics of QREs as well as inference rules for deriving them:

### 2017-10-03 papers/QuotientLensSynth/main.tex
added a table for the K and W functions

@@
 \newcommand{\ksf}[1]{\FINISH{teal}{K}{#1}}
-\newcommand{\sm}[1]{\FINISH{dkpurple}{SM}{#1}}
+\newcommand{\sam}[1]{\FINISH{dkpurple}{SM}{#1}}
 
@@
 \newcommand{\re}[1]{\ensuremath{\mathit{#1}}}
+\newcommand{\squash}[3]{\ensuremath{\mathit{squash}(#1, #2, #3)}}
+\newcommand{\perm}[2]{\ensuremath{\mathit{perm}(#1)\; \mathit{with}\; #2}}
+\newcommand{\normalize}[3]{\ensuremath{\mathit{normalize}(#1, #2, #3)}}
+\newcommand{\sep}{\ensuremath{\; | \;}}
 
@@
 
-$$perm(\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER}, \re{YEAR}) \; with \; (\re{WSP}^* \mapsto \string
-\n),$$
+$$\perm{\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER}, \re{YEAR}}{(\re{WSP}^* \mapsto \string \n)},$$
 
@@
 
-\sm{[Mention and give examples of quotient lenses here?]}
+\sam{[Mention and give examples of quotient lenses here?]}
 
@@
 \begin{align*}
-q := \; &R \; | \; q \mapsto s \; | \; squash (R, R', f) \; | \;
-perm \; (q_1, \ldots, q_n) \; with \; q \; | \; normalize(R, R', f)\\
-&q' \circ q \; | \; q \cdot q' \; | \; (q \; | \; q') \; | \; q^*,
+q := \; &R \sep q \mapsto s \sep \squash{R}{R'}{f} \sep
+\perm{(q_1, \ldots, q_n)}{q} \;  | \; \normalize{R}{R'}{f}\\
+&q' \circ q \sep q \cdot q' \sep (q \sep q') \sep q^*,
 \end{align*}
@@
 only if (1) $R = \varnothing$, or (2) $R = S_1 \cdot S_2$ with $S_1, S_2$
-strongly unambiguous and $S_1 \cdot^! S_n$, or (3) $R = S_1 \; | \; S_2$ with
+strongly unambiguous and $S_1 \cdot^! S_n$, or (3) $R = S_1 \sep S_2$ with
 $S_1, S_2$ strongly unambiguous and $\mathcal{L}(S_1) \cap \mathcal{L}(S_2) =
@@
 We now give the semantics of QREs as well as inference rules for deriving them:
-\textcolor{red}{[Informal Meanings/Example Usages Needed]}\\
+\textcolor{red}{[Informal Meanings/Example Usages Needed]}
+
+\begin{figure}[t]
+  \centering
+  \[
+    \begin{array}{r@{\quad}c@{\quad}cc}
+
+      q & W(q) & K(q) \\ \hline
+      R & R & R \\
+      q \mapsto s & W(q) & s \\
+      \squash{R}{R'}{f} & R \sep R' & R' \\
+      \normalize{R}{R'}{f} & R & R' \\
+      q_1 \circ  q_2 & W(q_2) & K(q_1) \\
+      q_1 \cdot q_2 & W(q_1) \cdot W(q_2) & K(q_1) \cdot K(q_2) \\
+      q_1 \sep q_2 & W(q_1) \sep W(q_2) & K(q_1) \sep K(q_2) \\
+      q_1^* & W(q_1)^* & K(q_1)^* \\
+    \end{array}
+  \]
+\[
+\begin{array}{r@{\quad}l}
+W( \perm{(q_1, \ldots, q_n)}{q} ) = &
+\bigcup \limits_{\sigma \in S_n} W(q_{\sigma(1)}) \cdot W(q) \cdot \ldots \cdot \cdot W(q) \cdot W(q_{\sigma(n)})
+\\
+K( \perm{(q_1, \ldots, q_n)}{q} ) = &
+ K(q_1) \cdot K(q_1) \cdot \ldots \cdot \cdot K(q) \cdot K(q_n) 
+\end{array}
+\]
+  \caption{Whole and Kernel Regular Expressions.}
+  \label{fig:wk}
+\end{figure}
+
 \begin{enumerate}
@@
 \begin{enumerate}
-  \item $W(squash(R, R', f)) = R \; | \; R'$,
+  \item $W(squash(R, R', f)) = R \sep R'$,
   \item $K(squash(R, R', f)) = R'$, 
@@
 \end{center}
-\item $q=q \; | \;  q'$
+\item $q=q \sep  q'$
 \begin{center}
 \begin{enumerate}
-  \item $W(q \cdot q') = W(q) \; | \; W(q')$,
-  \item $K(q \cdot q') = K(q) \; | \; K(q')$, 
+  \item $W(q \cdot q') = W(q) \sep W(q')$,
+  \item $K(q \cdot q') = K(q) \sep K(q')$, 
   \item
-For all $w, w' \in W(q \; | \; q')$, $w \; EqRel(q' \; | \; q) \; w'$ if and
+For all $w, w' \in W(q \sep q')$, $w \; EqRel(q' \sep q) \; w'$ if and
 only if 
@@
 \item
-$Canonizer(q \; | \; q') = Canonizer(q) \; | \; Canonizer(q')
+$Canonizer(q \sep q') = Canonizer(q) \sep Canonizer(q')
 $
@@
 The syntax for the language of bijective lenses is given by
-$$\ell := R \; | \; const(s, s') \; | \;  swap(\ell,
-\ell) \; | \; \ell \cdot \ell' \; |  \; (\ell \; | \; \ell') \; | \; \ell^* \;
+$$\ell := R \sep const(s, s') \sep  swap(\ell,
+\ell) \sep \ell \cdot \ell' \; |  \; (\ell \sep \ell') \sep \ell^* \;
 | \; \ell \circ \ell',$$ where $R$ ranges over regular expressions and $s$
@@
 \begin{align*}
-\llbracket R \rrbracket &= \{(r, r) \; | \; r \in \mathcal{L}(R)\}\\
-\llbracket swap(\ell, \ell') \rrbracket &= \{(s \cdot t, t' \cdot s') \; | \;
+\llbracket R \rrbracket &= \{(r, r) \sep r \in \mathcal{L}(R)\}\\
+\llbracket swap(\ell, \ell') \rrbracket &= \{(s \cdot t, t' \cdot s') \sep
 (s, s') \in \llbracket \ell \rrbracket \text{ and } (t, t') \in \llbracket
 \ell' \rrbracket\}\\
-\llbracket \ell \cdot \ell' \rrbracket &= \{(s \cdot t, s' \cdot t) \; | \;
+\llbracket \ell \cdot \ell' \rrbracket &= \{(s \cdot t, s' \cdot t) \sep
 (s, s') \in \llbracket \ell \rrbracket \text{ and } (t, t') \in \llbracket
 \ell' \rrbracket\}\\
-\llbracket \ell \; | \; \ell' \rrbracket &= \{(s \cdot t) \; | \;
+\llbracket \ell \sep \ell' \rrbracket &= \{(s \cdot t) \sep
 (s, t) \in \llbracket \ell \rrbracket \text{ or } (s, t) \in \llbracket
@@
 \llbracket \ell^* \rrbracket &= \{(s_1 \cdot \ldots \cdot s_n, t_1 \cdot \ldots
-\cdot t_n) \; | \; (s_i, t_i) \in \llbracket \ell \rrbracket \text{ for } 1
+\cdot t_n) \sep (s_i, t_i) \in \llbracket \ell \rrbracket \text{ for } 1
 \leq i \leq n\}
@@
 following grammar:
-$$ q := id(\ell) \; | \; lquot(q, \ell) \; | \; rquot(\ell, q) \; | \; q
-\cdot q_2 \; | \; (q \; | \; q') \; | \; q^* \; | \; q \circ q',$$
+$$ q := id(\ell) \sep lquot(q, \ell) \sep rquot(\ell, q) \sep q
+\cdot q_2 \sep (q \sep q') \sep q^* \sep q \circ q',$$
 where $\ell$ ranges over bijective lenses 
@@
   \item
-  $q = q_1 \; | \; q_2$, then
+  $q = q_1 \sep q_2$, then
       \begin{prooftree}
@@
 \AxiomC{$\mathcal{L}(W(d_1)) \cap \mathcal{L}(W(d_2)) = \varnothing$}
-\QuaternaryInfC{$q_1 \; | \; q_2: (c_1 \; | \; c_2)
-\Leftrightarrow (d_1 \; | \; d_2)$}
+\QuaternaryInfC{$q_1 \sep q_2: (c_1 \sep c_2)
+\Leftrightarrow (d_1 \sep d_2)$}
 \end{prooftree}
   $$
-  \llbracket q_1 \; | \; q_2 \rrbracket.get(s) = 
+  \llbracket q_1 \sep q_2 \rrbracket.get(s) = 
   \begin{cases}
@@
   \end{cases}$$
-  $$\llbracket q_1 \; | \; q_2 \rrbracket.put(s) = 
+  $$\llbracket q_1 \sep q_2 \rrbracket.put(s) = 
   \begin{cases}
@@
   \item
-  $q = q_1 \; | \; q_2$ where $q_1 : c_1 \Leftrightarrow d_1 $, $q_2 : c_2
+  $q = q_1 \sep q_2$ where $q_1 : c_1 \Leftrightarrow d_1 $, $q_2 : c_2
   \Leftrightarrow d_2$, $\mathcal{L}(W(c_1)) \cap \mathcal{L}(W(c_2)) =
@@
   $$
-  \llbracket q_1 \; | \; q_2 \rrbracket.get(s) = 
+  \llbracket q_1 \sep q_2 \rrbracket.get(s) = 
   \begin{cases}
@@
 $$
-  \llbracket q_1 \; | \; q_2 \rrbracket.get(s) = 
+  \llbracket q_1 \sep q_2 \rrbracket.get(s) = 
   \begin{cases}
@@
   \end{cases}$$
-  so $\llbracket q_1 \; | \; q_2 \rrbracket.get = \llbracket \ell_1 \; | \;
-  \ell_2 \rrbracket \circ Canonizer(c_1 \; | \; c_2)$. A similar argument shows
-  that $\llbracket q_1 \; | \; q_2 \rrbracket.put = \llbracket \ell_1 \; | \;
-  \ell_2 \rrbracket^{-1} \circ Canonizer(d_1 \; | \; d_2)$.\\
+  so $\llbracket q_1 \sep q_2 \rrbracket.get = \llbracket \ell_1 \sep
+  \ell_2 \rrbracket \circ Canonizer(c_1 \sep c_2)$. A similar argument shows
+  that $\llbracket q_1 \sep q_2 \rrbracket.put = \llbracket \ell_1 \sep
+  \ell_2 \rrbracket^{-1} \circ Canonizer(d_1 \sep d_2)$.\\
   This completes the proof.

### 2017-10-04 combine-data.py
make nearly done

@@ -160,7 +160,7 @@ def main(args):
         prog1_path = args[1]
         prog2_path = args[2]
         prog3_path = args[3]
-        ensure_dir("papers/BijectiveLensSynth/generated-data/")
+        ensure_dir("data/generated-data/")
         csv1 = retrieve_csv(prog1_path)
         csv2 = retrieve_csv(prog2_path)
         csv3 = retrieve_csv(prog3_path)
@@ -174,7 +174,7 @@ def main(args):
             for key, value in row.items():
                 relevant_data[key] = value
         print(data_keyed_by_test)
-        with open("papers/BijectiveLensSynth/generated-data/data.csv", "wb") as csvfile:
+        with open("data/generated-data/data.csv", "wb") as csvfile:
             fieldnames = data_keyed_by_test[data_keyed_by_test.keys()[0]].keys()
             writer = csv.DictWriter(csvfile, fieldnames=fieldnames)
 

### 2017-10-05 papers/QuotientLensSynth/main.tex
Did more explaining for QREs. Added figures for QRE semantics.

@@
 \section{Introduction}
-\saz{We need a more gentle ramp-up to the example here.}
+Lense programming languages such as Boomerang enable a programmer to express a
+transformation that maps some input data to some output data, as well as a
+transformation that maps the output data back into the input data in a single
+program. For instance, lenses enable a programmer to write a program which
+enables a doctor to acquire a summary of a patient's data, and also allows the
+doctor to reflect updates made to the patient's record back into the main
+database.
 
-\begin{itemize}
-\item Lens programming languages [e.g. Boomerang] are designed to make it possible to express
-  bidirectional data transformations using a single program.
+Lenses therefore enable the  programmer to kill two birds in one
+stone by allowing the programmer to express two different data transformations
+in one program. Also, when lenses are appropriately constrained, they offer
+guarantees about how the two transformations interact with each other. Lenses
+have therefore been used effectively as a solution to the \textit{view update
+problem}, a classical problem in the database community in which the
+programmer wishes to write a program which provides a \textit{view} of data
+derived from a \textit{source} to a client, as well as a program which enables
+the client to reflect changes made in the view back into the source without
+corrupting the source.
 
-\item Lenses have found practical application in Augeas \saz{and ...}
+However, the constraints that are imposed on lenses are often too strict.
+For instance, the programmer may wish to ignore updates to the data that only
+change the order of fields, or add whitespace characters that in reality are not
+part of the data. Foster et al therefore proposed \textit{quotient lenses} as a
+solution to this problem. Quotient lenses enable a programmer to write lenses
+which transform data that is quotiented out by an equivalence relation, so that
+records which are equivalent to each other modulo this relation are considered
+the same by the underlying lens.
 
-\item The source and target types of a lens express constraints on the format of
-  the data.
+This paper addresses two main challenges that arise when programming quotient
+lenses. The first challenge is tied directly to the programming of quotient
+lenses in Boomerang, a string lens programming language where the source and
+view types are given by regular expressions. While Boomerang enables the
+programmer to express a rich set of quotient lenses, the syntax for writing
+these lenses is often too verbose and cumbersome. This makes writing quotient
+lenses in Boomerang rather difficult.
 
-\item Lens programming is still quite tricky, because lens composition operators
-  are subject to side conditions (e.g. unique decomposition) that are not so
-  easy to reason about.
-
-\item Our prior work~\cite{popl18} has shown how to synthesize lenses for the
-  \textit{bijective} case.
-
-\item This still isn't as expressive as one would want in practice, because many
-  lens transformations aren't truly bijective.
-
-\item There are at least two challenges with programming quotient lenses: (1)
-  defining the quotient languages (which serve as the types of the lenses)
-  themselves is not so easy, and (2) the lenses themselves are still not easy to
-  write by hand.
-
-\end{itemize}
-
-String quotient lenses enable one to express bidirectional string
-transformations that obey the lens laws up to equivalence relations. However,
-the various approaches that have been proposed to express quotient lenses
-make it difficult for the programmer to express complicated equivalence
-relations over the data.
-
-For example, when converting \textsc{Bib}\TeX{} files to EndNote files, the data
-may contain arbitrary amounts of whitespace characters in between the actual
-data. In addition, the fields may be ordered in an arbitrary order. For
-instance, we may wish to convert this \textsc{Bib}\TeX{} entry: \saz{I propose
-  we display such examples in verbatim:}
+For example, suppose that the programmer wishes to convert citation records
+stored in \textsc{Bib}\TeX{} format to a representation of the same data in the
+EndNote format. The programmer wishes to ``quotient'' away unnecessary
+whitespace characters in between the various fields in each record, while also
+ignoring the order in which these fields occur in the recore. For instance, the
+quotient lens which the programmer wishes to derive should convert this
+\textsc{Bib}\TeX{} record:
 
@@
 
-to this EndNote entry:
+to this EndNote record
 
@@
 
-With the existing tools, it is cumbersome to express this, and other complicated
-equivalence relations.  \saz{It might be a good idea to show what this would
-  look like in all its gory detail, just to make the problem very
-  concrete. Perhaps in a figure?}
+While Boomerang enables the programmer to write a lens which exhibits this
+behaviour, the syntax tree for the program is large and complicated.
+
+\saz{It might be a good idea to show what this would look like in all its gory
+detail, just to make the problem very concrete. Perhaps in a figure?}
 
 In this paper, we introduce the formalism of Quotient Regular Expressions (or
-QREs) to solve this problem. QREs enable the programmer to express an
-equivalence relation at internal terms of a regular expression, thus giving
-the programmer more control over where quotienting occurs in the data.\\
-For example, the fields of a \textsc{Bib}\TeX{} article like in the example
-above might be expressed by the following QRE:
+QREs) to address this issue. QREs enable the programmer to express
+a rich set of equivalence relations on regular expressions at the level of
+syntax. QREs therefore enable the programmer to express more equivalence
+relations with less effort. Also, since QREs are defined at the level of
+syntax on regular expressions, QREs give the programmer more control over where
+the quotienting occurs.
 
-$$\perm{\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER}, \re{YEAR}}{(\re{WSP}^* \mapsto \string \n)},$$
+The second challenge that we address in this paper is the synthesis of
+\textit{bijective} quotient lenses. These are quotient lenses which match the
+equivalance classes of the source data to the equivalence classes of the view
+data bijectively as in the \textsc{Bib}\TeX{} to EndNote example above. 
+Building on our prior work~\cite{popl18} where we showed how to synthesize
+(ordinary?) bijective lenses, we show how to synthesize bijective quotient
+lenses given a synthesis task described by a pair of QREs and a set of
+examples. To our knowledge this is the first time that this has been done.
 
-\noindent where $\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},$ and $\re{YEAR}$ are
-regular expressions describing the respective \textsc{Bib}\TeX\; fields of the
-same name. The QRE of an EndNote article is similarly expressed, but with the
-regular expressions for the respective fields substituted with the appropriate
-EndNote equivalent.  \saz{Do we really mean all \re{WSP} maps to newline here?
-  What about the spaces in the title, etc.?  Would this be confusing without
-  more explanation?}
-
-\sam{[Mention and give examples of quotient lenses here?]}
-
-\subsection{Contributions}
+Our plan for this paper is as follows:
 \begin{enumerate}
@@
 \section{Quotient Regular Expressions}
+This section introduces Quotient Regular Expressions or QREs. QREs are regular
+expressions augmented with syntax that lets them simultaneously express an
+equivalence relation on the language described by the regular expression.
 
-\saz{Need a transition here:  This section introduces QREs.  They are regular
-  expressions augmented with syntax that lets them simultaneously denote
-  quotient languages.  
-   }
+\sam{More explanation here?}
 
-  \paragraph{Syntax of QREs}
+  \subsection{Syntax of QREs}
 The language of Quotient Regular Expressions (QREs) is given by the following
@@
 \begin{align*}
-q := \; &R \sep q \mapsto s \sep \squash{R}{R'}{f} \sep
-\perm{(q_1, \ldots, q_n)}{q} \;  | \; \normalize{R}{R'}{f}\\
+q := \; &R \sep R \mapsto s \sep \squash{R}{R'}{f} \sep
+\perm{q_1, \ldots, q_n}{q} \;  | \; \normalize{R}{R'}{f}\\
 &q' \circ q \sep q \cdot q' \sep (q \sep q') \sep q^*,
 \end{align*}
-where $R$ ranges over regular expressions, $f$ ranges over functions whose
-domain and codomain are regular languages, and $s$ ranges over character
-strings.
+where $R$ ranges over regular expressions, $f$ ranges over functions between
+regular languages, and $s$ ranges over character strings.
 
-\saz{  Every regular expression $R$ is a QRE.  The main new syntactic forms are: $q \mapsto s$, which
-}
-
-\saz{Would it be simpler to include atoms and $\epsilon$ in $q$ rather than $R$?  That is, $R$ is
-  just a subset of $QRE$ built from atoms, $\cdot$ $|$ and ${}^*$.  We could
-  reserve $R$ as a metavariable that ranges over $q$'s such that $q$ is a
-  regular expression.}
-
-\saz{Need to introduce $\mathcal{L}(R)$ notation.}
-
-Quotient Regular Expressions (or QREs) enable us to give richer specifications
-for data formats than ordinary regular expressions. More concretely, each
-Quotient Regular Expression (or QRE) $q$ enables us to express
+Each QRE $q$ enables us to express
 \begin{enumerate}
@@
   \end{enumerate}
+  Observe that since $Canonizer(q)$ computes $Canonizer(q)(w)$ as the unique $k$
+  in $\mathcal{L}(K(q))$ such that $w$ is equivalent to $k$ mod $EqRel(q)$,
+  then the equivalence classes of $EqRel(q)$ are the same as the fibres of
+  $Canonizer(q)$.
+  \subsection{Meaning and Example Usages for QREs}
+  \begin{enumerate}
+    \item 
+    Every regular expression is a QRE. When a regular expression $R$ is used as
+    a QRE, then $\mathcal{L}(R)$ is assumed to be quotiented by the equality
+    equivalence relation on $\mathcal{L}(R)$.
+    \item
+    For the QRE $R \mapsto s$, the data is described by $\mathcal{L}(R)$, but
+    all members of $R$ are considered equivalent to each other, with $s \in
+    \mathcal{L}(R)$ as the choice of representative. This construct is useful
+    when the programmer simply wishes to ignore unnecessary characters such as
+    excess whitespace e.g. 
+    $$FIRST\_NAME \cdot (WHITESPACE^* \mapsto `` \quad ") \cdot LAST\_NAME$$
+    \item
+    $\squash{R}{R'}{f}$ enables the programmer to use one of two different
+    representations of the data if they can provide a way of mapping the first
+    representation to the second. For instance, the first representation may be
+    $$FIRST\_NAME \cdot `` \quad " \cdot LAST\_NAME$$
+    while the second representation may be
+    $$FIRST\_NAME \cdot ``," \cdot LAST\_NAME$$
+    The mapping that maps the first representation to the second simply converts
+    the space between the first and last name to a comma.
+    \item
+    The $\perm{q_1, \ldots, q_n}{q}$ construct enables the programmer to
+    consider all permutations of some data that are interspersed with a
+    separator the same. For instance, in the \textsc{Bib}\TeX{} to EndNote
+    example, the programmer may use this construct as follows:
+    $$\perm{\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},
+    \re{YEAR}}{(\re{WSP}^* \mapsto \string \n)}$$
+ where $\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},$ and $\re{YEAR}$ are
+regular expressions describing the respective \textsc{Bib}\TeX\; fields of the
+same name, and $WSP$ is a regular expression describing whitespace characters.
+\item
+The constructs $q \cdot q'$, $q \sep q'$ and $q^*$ are similar to the respective
+regular expression constructs. However, the regular constructs for QREs are more
+restrictive for QREs since the data should match the whole language $W(q)$ and
+the kernel language $K(q)$ unambiguosly so as to make transformations
+deterministic.
+\item
+The $\normalize{R}{R'}{f}$ construct allows the programmer to choose how to
+canonize $\mathcal{L}(R)$ into $\mathcal{L}(R')$ be providing the canonizing
+function $f$ themselves. This constuct is useful when the desired equivalence
+relation cannot easily be expressed by any of the other constructs.
+  \end{enumerate}
+
+\saz{Every regular expression $R$ is a QRE.  The main new syntactic
+forms are:
+$R \mapsto s$, which }
+
+\saz{Would it be simpler to include atoms and $\epsilon$ in $q$ rather than $R$?  That is, $R$ is
+  just a subset of $QRE$ built from atoms, $\cdot$ $|$ and ${}^*$.  We could
+  reserve $R$ as a metavariable that ranges over $q$'s such that $q$ is a
+  regular expression.}
+
+\saz{Need to introduce $\mathcal{L}(R)$ notation.}
+
+\iffalse
+Quotient Regular Expressions (or QREs) enable us to give richer specifications
+for data formats than ordinary regular expressions. More concretely,
   All data to be transformed must match $W(q)$, the ``whole'' regular
@@
   unique representative $k$ is chosen using the ``canonizing'' function
-  $Canonizer(q)$.\\
-  Observe that since $Canonizer(q)$ computes $Canonizer(q)(w)$ as the unique $k$
-  in $\mathcal{L}(K(q))$ such that $w$ is equivalent to $k$ mod $EqRel(q)$,
-  then the equivalence classes of $EqRel(q)$ are the same as the fibres of
   $Canonizer(q)$.
+  \fi
+
 \subsection{Semantics of QREs}
 \subsubsection{Preliminaries}
-The transformations that will be performed on the data will require
-that the data is matched against regular expressions. Consequently, we require
-that regular expressions used by QREs be \textit{strongly unambiguous}, a
-condition which ensures that if a string $s$ matches a regular expression $R$,
-then $s$ matches $R$ uniquely.
-
-\saz{Solomon: you should use blank lines instead of backslashes to separate
-  paragraphs so latex can format them properly.}
+As we mentioned earlier, when applying the regular constructs $\cdot
+\; | \;  *$ to QREs, we require that the regular languages
+described by QREs match the data in only one way so as to make transformations
+deterministic. Consequently, we require that regular expressions used
+by QREs be \textit{strongly unambiguous}, a condition which ensures that if a
+string $s$ matches a regular expression $R$, then $s$ matches $R$ uniquely.
 
@@
 
-We now give the semantics of QREs as well as inference rules for deriving them:
-\textcolor{red}{[Informal Meanings/Example Usages Needed]}
-
+The semantics for QREs that formally define the whole language $W(q)$, the
+kernel language $K(q)$, the equivalence relation $EqRel(q)$ and the
+canonizing function $Canonizer(q)$ are given in Figures ~\ref{fig:wk},
+~\ref{fig:relations} and ~\ref{fig:canonizers}.
 \begin{figure}[t]
@@
       R & R & R \\
-      q \mapsto s & W(q) & s \\
+      R \mapsto s & W(q) & s \\
       \squash{R}{R'}{f} & R \sep R' & R' \\
@@
       q_1 \sep q_2 & W(q_1) \sep W(q_2) & K(q_1) \sep K(q_2) \\
-      q_1^* & W(q_1)^* & K(q_1)^* \\
+      q^* & W(q)^* & K(q)^* \\
     \end{array}
@@
 
-\begin{enumerate}
-  \item $q = R$
+\begin{figure}[t]
   \begin{center}
-\begin{enumerate}
-  \item $W(R) = R$,
-  \item $K(R) = R$, 
-  \item
-For all $w, w' \in W(R)$, $w \; EqRel(q) \; w'$ if and only if $w =
-w'$.
-\item
-$Canonizer(R)(r) = r$
-\end{enumerate}
+\[
+    \begin{array}{l@{\quad}l} 
+      q & EqRel(q)  \\ \hline
+      R & w \; \equiv_R \; w' \iff w = w \\
+      R \mapsto s & w \; \equiv_{R \mapsto s} \; w' \text{ for all }w, w'\\
+      \squash{R}{R'}{f} & w \; \equiv_{\squash{R}{R'}{f}} \; w' \iff f(w) = w'
+      \text{ or } w = w' \\
+      \normalize{R}{R'}{f} & w \; \equiv_{\normalize{R}{R'}{f}} \; w' \iff
+      f(w)=f(w') \text{ or }w = w'\\
+      q_2 \circ  q_1 &  w \; \equiv_{q_2 \circ q_1} \; w' \iff \exists k, k' \in
+  \mathcal{L}(K(q_2)) \text{ such that } w \; \equiv_{q_1} \; k, \; w' \;
+  \equiv_{q_1} \; k', \text{ and } k \; \equiv_{q_2} \; k'\\
+      q_1 \cdot q_2 &  w \; \equiv_{q_1 \cdot q_2} \; w'  \iff w = r_1
+      \cdot r_2, \; w' = {r'}_1 \cdot {r'}_2 \text{ with } r_1 \; \equiv_{q_1}
+      \; {r'}_1, \; r_2 \; \equiv{q_n} \; {r'}_2\\
+      q_1 \sep q_2 &  w \; \equiv_{q' \sep q} \; w' \iff w \; \equiv_{q_1} \; w'
+      \text{ or } \; w \; \equiv_{q_2} \; w'\\
+      q^* &  w \; \equiv_{q*} \; w' \iff w = r_1 \cdot \ldots \cdot r_n, \; w'
+      = {r'}_1 \cdot \ldots \cdot {r'}_n \text{ and } r_i \equiv_{q} \; {r'}_i
+      \text{ for all } 1 \leq i \leq n\\
+    \end{array}
+    \]
+    \end{center}
+    $w \; \equiv_{\perm{(q_1, \ldots, q_n)}{q}} \; w' $ if and only if
+
+    $w = r_{\sigma(1)} \cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}
+    \text{ and } w' = {r'}_{\theta(1)} \cdot s'_1 \cdot \ldots \cdot s'_{n-1}
+    \cdot {r'}_{\theta(n)}$ for some $\sigma, \theta \in S_n$, with $r_i \;
+    \equiv_{q_i} \; r'_i$ and $s_k \; \equiv_{q} \; s'_{k}$.
+  \caption{QRE Equivalence Relations}
+  \label{fig:relations}
+\end{figure}
+\begin{figure}[t]
+  \begin{center}
+\[
+    \begin{array}{l@{\quad}l} 
+      q & Canonizer(q)  \\ \hline
+      R & Canonizer(R) = id_{\mathcal{L}(R)} \\
+      R \mapsto s & Canonizer(R \mapsto s)(w) = s\\
+      \squash{R}{R'}{f} & Canonizer(\squash{R}{R'}{f})(w) = 
+\begin{cases}
+f(w) & \text{if } w \in \mathcal{L}(R)\\
+w & \text{otherwise}
+\end{cases}\\
+      \normalize{R}{R'}{f} & Canonizer(\normalize{R}{R'}{f}) = f\\
+      \perm{(q_1, \ldots, q_n)}{q} & \\
+      q_2 \circ  q_1 & Canonizer(q' \circ q) = Canonizer(q') \circ Canonizer(q)\\
+      q_1 \cdot q_2 & Canonizer(q' \cdot q) = Canonizer(q') \cdot Canonizer(q)\\
+      q_1 \sep q_2 &  Canonizer(q_1 \sep q_2)(w) = 
+\begin{cases}
+Canonizer(q_1)(w) & \text{if } w \in \mathcal{L}(W(q_1))\\
+Canonizer(q_2)(w) & \text{if } w \in \mathcal{L}(W(q_2))\\
+\end{cases}\\
+      q^* &  Canonizer(q^*) = Canonizer(q)^* \\
+    \end{array}
+    \]
+    \end{center}
+    $Canonizer(\perm{(q_1, \ldots, q_n)}{q})(r_{\sigma(1)}
+\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \newline
+= (Canonizer(q_1)(r_1)) \cdot (Canonizer(q)(s_1)) \cdot \ldots \cdot
+(Canonizer(q_n)(r_n))$
+  \caption{QRE Canonizers}
+  \label{fig:canonizers}
+\end{figure}
+\begin{figure}[t]
+\centering
 \begin{prooftree}
@@
 \end{prooftree}
-\end{center}
-\item $q=q \mapsto s$
-\begin{center}
-\begin{enumerate}
-  \item $W(q \mapsto s) = W(q)$,
-  \item $K(q \mapsto s) = s$, 
-  \item
-For all $w, w' \in W(q \mapsto S)$, $w \; EqRel(q \mapsto S) \; w'$,
-\item
-$Canonizer(q \mapsto S)(r) = s$
-\end{enumerate}
+
 \begin{prooftree}
 \AxiomC{$s \in W(q)$}
-\UnaryInfC{$q \mapsto s$ is well formed}
+\UnaryInfC{$R \mapsto s$ is well formed}
 \end{prooftree}
-\end{center}
-\item $q=squash(R, R', f)$
-\begin{center}
-\begin{enumerate}
-  \item $W(squash(R, R', f)) = R \sep R'$,
-  \item $K(squash(R, R', f)) = R'$, 
-  \item
-For all $w, w' \in W(squash(R, R', f))$, $w \; EqRel(q) \; w'$ if and only
-if $f(w) = w'$ or $w = w'$
-\item
-$Canonizer(squash(R, R', f))(w) = 
-\begin{cases}
-f(w) & \text{if } w \in \mathcal{L}(R)\\
-w & \text{otherwise}
-\end{cases}$
-\end{enumerate}
+
 \begin{prooftree}
@@
 \end{prooftree}
-\end{center}
-\item $q=perm \; (q_1, \ldots, q_n) \; with \; q$. Given regular expressions
-$R_1, \ldots, R_n$ and $S$, let $\prod({R_i}, S) = R_1 \cdot S \cdot \ldots
-\cdot S \cdot R_n$.
-\begin{center}
-\begin{enumerate}
-  \item $W(perm \; (q_1, \ldots, q_n) \; with \; q) = \bigcup \limits_{\sigma
-  \in S_n} W(q_{\sigma(1)}) \cdot W(q) \cdot \ldots \cdot \cdot W(q) \cdot
-  W(q_{\sigma(n)}),$
-  \item $K(perm \; (q_1, \ldots, q_n) \; with \; q) = K(q_1) \cdot K(q_1)
-  \cdot \ldots \cdot \cdot K(q) \cdot K(q_n)$, 
-  \item
-For all $w, w' \in W(q)$,  $w \; EqRel(q) \; w'$ if and only if
-   $$w = r_{\sigma(1)} \cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot
-    r_{\sigma(n)} \text{ and } r' = {r'}_{\theta(1)} \cdot s'_1 \cdot \ldots
-    \cdot s'_{n-1} \cdot {r'}_{\theta(n)}$$ for some $\sigma, \theta \in
-    S_n$, with $r_i \; EqRel(q_i) \; r'_i$ and $s_k \; EqRel(q) \; s'_{k}$.
-\item
-$Canonize(q)(r_{\sigma(1)}
-\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \newline
-= (Canonize(q_1)(r_1)) \cdot (Canonize(q)(s_1)) \cdot \ldots \cdot \cdot \cdot
-(Canonize(q_n)(r_n))$
-\end{enumerate}
+
 \begin{prooftree}
-\AxiomC{$\forall \sigma \neq \theta, \; \prod(\{W(q_{\sigma(i)})\}, W(q))\cap
- \prod(\{W(q_{\theta(i)})\}, W(q)) = \varnothing$}
+\AxiomC{${\substack{\forall \sigma \neq \theta, \; W(q_{\sigma(1)}) \cdot W(q)
+\cdot \ldots \cdot W(q_{\sigma(n)})\\ \cap W(q_{\theta(1)}) \cdot W(q)
+\cdot \ldots \cdot W(q_{\theta(n)}) =\varnothing}}$}
+\AxiomC{$q_i, q$ are well formed}
 \AxiomC{$\forall \sigma, \; K(q_{\sigma(1)}) \cdot^! K(q) \cdot^! \ldots 
 \cdot^! K(q_{\sigma(n)})$}
-\BinaryInfC{$permute \; (q_1, \ldots, q_n) \; with \;  q$ is well formed}
+\TrinaryInfC{$\perm{q_1, \ldots, q_n}{q}$ is well formed}
 \end{prooftree}
-\end{center}
 
-\item $q = normalize(R, R', f)$
-\begin{center}
-\begin{enumerate}
-  \item
-$W(normalize(f)) = R$
-\item
-$K(normalize(f)) = R'$
-\item For all $w, w' \in W(q)$, $w \; EqRel(q) \; w'$ if and only if
-$f(w)=f(w')$
-\item
-$Canonize(normalize(f)) = f$
 \begin{prooftree}
@@
 \end{prooftree}
-\end{enumerate}
-\end{center}
-\item $q=q' \circ q$
-\begin{center}
-\begin{enumerate}
-  \item $W(q' \circ q) = W(q)$,
-  \item $K(q' \circ q) = K(q')$, 
-  \item
-For all $w, w' \in W(squash(R, R', f))$, $w \; EqRel(q' \circ q) \; w'$ if and
-only $$\text{there exist }k, k' \in
-  \mathcal{L}(K(q)) \text{ such that } w \; EqRel(q) \; k, \; w' \;
-  EqRel(q) \; k', \text{ and } k \; EqRel(q') \; k' $$
-\item
-$Canonizer(q' \circ q) = Canonizer(q') \circ Canonizer(q)
-$
-\end{enumerate}
+
 \begin{prooftree}
@@
 \end{prooftree}
-\end{center}
-\item $q=q \cdot q'$
-\begin{center}
-\begin{enumerate}
-  \item $W(q \cdot q') = W(q) \cdot W(q')$,
-  \item $K(q \cdot q') = K(q) \cdot K(q')$, 
-  \item
-For all $w, w' \in W(q \cdot q')$, $w \; EqRel(q' \cdot q) \; w'$ if and
-only $$w = r_1
-      \cdot r_2, \; w' = {r'}_1 \cdot {r'}_2 \text{ with } r_1 \; EqRel(q) \;
-      {r'}_1, \; r_2 \; EqRel(q') \; {r'}_2$$
-\item
-$Canonizer(q \cdot q') = Canonizer(q) \cdot Canonizer(q')
-$
-\end{enumerate}
+
 \begin{prooftree}
@@
 \end{prooftree}
-\end{center}
-\item $q=q \sep  q'$
-\begin{center}
-\begin{enumerate}
-  \item $W(q \cdot q') = W(q) \sep W(q')$,
-  \item $K(q \cdot q') = K(q) \sep K(q')$, 
-  \item
-For all $w, w' \in W(q \sep q')$, $w \; EqRel(q' \sep q) \; w'$ if and
-only if 
-$$w \; EqRel(q) \; w' \text{ or } \; w \; EqRel(q') \; w'$$
-\item
-$Canonizer(q \sep q') = Canonizer(q) \sep Canonizer(q')
-$
-\end{enumerate}
+
 \begin{prooftree}
@@
 \end{prooftree}
-\end{center}
-\item $q=q'^*$
-\begin{center}
-\begin{enumerate}
-  \item $W(q'^*) = W(q')^*$,
-  \item $K(q'^*) = K(q')^*$, 
-  \item
-For all $w, w' \in W(q'^*)$, $w \; EqRel(q'^*) \; w'$ if and
-only if 
-$$w = r_1 \cdot \ldots \cdot r_n, \; w' = {r'}_1 \cdot \ldots
-      \cdot {r'}_n \text{ and } r_i \; EqRel(q) \; {r'}_i
-      \text{ for all } 1 \leq i \leq n$$
-      \item
-$Canonizer(q'^*) = Canonizer(q')^* $
-\end{enumerate}
+
 \begin{prooftree}
@@
 \end{prooftree}
-\end{center}
-\end{enumerate}
-\begin{claim}
+  \caption{QRE Inference Rules}
+  \label{fig:qrerules}
+\end{figure}
+The theorem below confirms that the semantics of QREs are consistent with their
+intended meaning:
+\begin{theorem}
 If $q$ is a well formed QRE, then
 \begin{enumerate}
+  \item $W(q)$ and $K(q)$ are well defined regular expressions,
   \item  $EqRel(q)$ is an equivalence relation on $\mathcal{L}(W(q))$,
@@
   \end{enumerate}
-\end{claim}
+\end{theorem}
 \section{Using QREs to Express Quotient Lenses}

### 2017-10-05 papers/QuotientLensSynth/main.tex
Bunch of comments here and there

@@
 \saz{It might be a good idea to show what this would look like in all its gory
-detail, just to make the problem very concrete. Perhaps in a figure?}
+detail, just to make the problem very concrete. Perhaps in a figure?}\bcp{+1}
 
@@
 
-Our plan for this paper is as follows:
+Our plan for this paper is as follows:\bcp{This should not just explain the
+  plan, but rather state the contributions {\em as contributions},
+  not only summarizing what they are but suggesting why they are novel,
+  challenging, and important.}
 \begin{enumerate}
@@
 
-\sam{More explanation here?}
+\sam{More explanation here?}\bcp{Probably not needed---we're going straight
+  on to the detailed explanation anyway.}
 
@@
   $EqRel(q)$, and
-  \item a ``canonizing'' function $Canonizer(q):\mathcal{L}(W(q))
+  \item a ``canonizing'' function\bcp{Words in math mode need $\backslash${\tt
+      mathit} around them, always.} $Canonizer(q):\mathcal{L}(W(q))
   \longrightarrow \mathcal{L}(K(q))$ which given any $w \in \mathcal{L}(W(q))$,
@@
     when the programmer simply wishes to ignore unnecessary characters such as
-    excess whitespace e.g. 
+    excess whitespace e.g. \bcp{Suggest using $\backlash${\tt mathtt} for 
+      concrete program identifiers and ascii characters in examples.}
     $$FIRST\_NAME \cdot (WHITESPACE^* \mapsto `` \quad ") \cdot LAST\_NAME$$
@@
 \]
-  \caption{Whole and Kernel Regular Expressions.}
+  \caption{Whole and Kernel Regular Expressions\bcp{left-justify columns}}
   \label{fig:wk}
@@
     \equiv_{q_i} \; r'_i$ and $s_k \; \equiv_{q} \; s'_{k}$.
-  \caption{QRE Equivalence Relations}
+  \caption{QRE Equivalence Relations\bcp{too wide}}
   \label{fig:relations}
@@
 (Canonizer(q_n)(r_n))$
-  \caption{QRE Canonizers}
+  \caption{QRE Canonizers\bcp{First column not needed (and, once it's gone,
+      maybe the = signs could be vertically aligned?)}}
   \label{fig:canonizers}
@@
 \end{prooftree}
-  \caption{QRE Inference Rules}
+  \caption{QRE Inference Rules\bcp{Maybe some or all of these can be
+      suppressed in the ``short version'' of the paper}}
   \label{fig:qrerules}
@@
 semantics of $\mathcal{Q}$, as well as inference rules that may be used to
-derive elements of $\mathcal{Q}$:
+derive elements of $\mathcal{Q}$:\bcp{I think the numbering and the ``if $q
+  = ..." stuff is not needed.  And maybe we can save space by typesetting the
+  get/put parts on the same line as the typing rule, at least in some cases.}
 \begin{enumerate}
@@
 If there exists a derivation $q:c \Leftrightarrow c'$ then
-$\llbracket q \rrbracket$ is a quotient lens from $R$ to $S$.
+$\llbracket q \rrbracket$ is a quotient lens from $R$ to $S$. \bcp{??}
 \end{claim}

### 2017-10-05 papers/QuotientLensSynth/main.tex
Progress with Solomon on the QRE paper

--- 2017-10-05 papers/QuotientLensSynth/main.tex
@@
 \section{Introduction}
-Lense programming languages such as Boomerang enable a programmer to express a
-transformation that maps some input data to some output data, as well as a
-transformation that maps the output data back into the input data in a single
-program. For instance, lenses enable a programmer to write a program which
-enables a doctor to acquire a summary of a patient's data, and also allows the
-doctor to reflect updates made to the patient's record back into the main
-database.
 
-Lenses therefore enable the  programmer to kill two birds in one
-stone by allowing the programmer to express two different data transformations
-in one program. Also, when lenses are appropriately constrained, they offer
-guarantees about how the two transformations interact with each other. Lenses
-have therefore been used effectively as a solution to the \textit{view update
-problem}, a classical problem in the database community in which the
-programmer wishes to write a program which provides a \textit{view} of data
-derived from a \textit{source} to a client, as well as a program which enables
-the client to reflect changes made in the view back into the source without
-corrupting the source.
+{\em Lens programming languages} such as
+Boomerang~\cite{Boomerang}\bcp{and others!} enable programmers to kill two
+birds in one stone by expressing both directions of a bidirectional
+transformation with a single program.
 
-However, the constraints that are imposed on lenses are often too strict.
+\bcp{
+  \begin{itemize}
+  \item Small example of a boomerang program (maybe in our syntax)
+  \item Discussion of lens laws.
+  \end{itemize}
+}
+
+However, the constraints that are imposed by the lens laws are sometimes too
+strict. 
 For instance, the programmer may wish to ignore updates to the data that only
@@
 quotient lens which the programmer wishes to derive should convert this
-\textsc{Bib}\TeX{} record:
+\textsc{Bib}\TeX{} record
 
@@
 
-to this EndNote record
+\noindent to this EndNote record
 
@@
 
-\noindent while allowing for this equivalent representation of the \textsc{Bib}\TeX{}
+\noindent and vice versa.
+
+\bcp{Moreover...}
+allowing for this equivalent representation of the \textsc{Bib}\TeX{}
 entry, which rearranges the fields and uses different whitespace conventions:
@@
 @Book{conway, 
-    Publisher = {Printed in GB by William Clowes & Sons Ltd},
+    Publisher = {Printed in GB by 
+                 William Clowes & Sons Ltd},
         Title = {Regular Algebra and Finite Machines},
-
          Year = {1971},
@@
 \end{verbatim}
+\bcp{Can we write (regular and QRE) lenses that handle this version of the
+  example?  Write this in Boomerang and see how it looks!!}
 
@@
 
-The second challenge that we address in this paper is the synthesis of
-\textit{bijective} quotient lenses. These are quotient lenses which match the
+\bcp{Give a little example (perhaps for Bibtex and/or EndNote?)}
+
+The second challenge that we address in this paper is the {\em synthesis} of
+(bijective) quotient lenses. These are quotient lenses which match the
 equivalance classes of the source data to the equivalence classes of the view
@@
 lenses given a synthesis task described by a pair of QREs and a set of
-examples. To our knowledge this is the first time that this has been done.
+examples.
 
-Our plan for this paper is as follows:
+Our main contributions are:
 \begin{enumerate}
-  \item
-  We introduce the language of Quotient Regular Expressions (QREs) which enables
-  us to express a set of equivalence relations on regular languages using a
-  compact, easy-to-use syntax.
-  \item
-  We describe a class $\mathcal{Q}$ of quotient lenses whose types are given by
-  QREs. We derive a normal form for quotient lenses in $\mathcal{Q}$, and
-  demonstrate that elements of $\mathcal{Q}$ that are in this normal form are
-  closed under left quotienting, right quotienting, composition and regular
-  operators.
-  \item
-  Using the normal form for quotient lenses in $\mathcal{Q}$, we reduce the
-  synthesis of quotient lenses in $\mathcal{Q}$ to the synthesis of bijective
-  lenses between regular languages. 
-  \item
-  We describe an implementation of QREs, the language $\mathcal{Q}$ of quotient
-  lenses, and an algorithm for synthesizing quotient lenses in $\mathcal{Q}$ in
-  the Boomerang language.
-  \item
-  We assess the usabibility of the QRE synthesis via a case study in which we
-  selected several data formats used by data.gov and synthesized quotient lenses
-  between them.
+\item We introduce a novel language of {\em Quotient Regular Expressions}
+(QREs) that provide compact, convenient notation for a broad class of
+equivalence relations on regular languages (Section~\ref{QRE}).
+\item We define a language of {\em QRE lenses}, a class of quotient lenses
+whose types are given by QREs.  Our main technical contribution is a normal
+form for QRE lenses and a proof that QRE lenses in normal form are closed
+under left quotienting, right quotienting, composition, and regular
+operators (Section~\ref{QRE-lenses}).
+\item Using this normal form, we reduce the problem of {\em synthesizing}
+QRE lenses from types and examples to the previously studied problem of
+synthesizing bijective lenses between regular languages
+(Section~\ref{synth}).
+\item We extend the Boomerang {\em implementation} with QREs, QRE lens operators,
+and QRE lens synthesis and demonstrate its applicability by using it to
+synthesize QRE lenses between several real-world data formats from the
+{\tt data.go} database (Section~\ref{impl}).
 \end{enumerate}
+Sections~\ref{relwork} and~\ref{concl} discuss related and future work.
 
 \section{Quotient Regular Expressions}
+\label{QRE}
+
 This section introduces Quotient Regular Expressions or QREs. QREs are regular
@@
 equivalence relation on the language described by the regular expression.
-
-\sam{More explanation here?}
 
@@
   $Canonizer(q)$.
-  \subsection{Meaning and Example Usages for QREs}
+
+  \subsection{Examples}
+
+\bcp{Give some counterexamples too!  And can we take away the enumerate?}
+
   \begin{enumerate}
@@
     when the programmer simply wishes to ignore unnecessary characters such as
-    excess whitespace e.g. 
+    excess whitespace e.g. \bcp{Suggest using $\backslash${\tt mathtt} for 
+      concrete program identifiers and ascii characters in examples.}
     $$FIRST\_NAME \cdot (WHITESPACE^* \mapsto `` \quad ") \cdot LAST\_NAME$$
@@
 W( \perm{(q_1, \ldots, q_n)}{q} ) = &
-\bigcup \limits_{\sigma \in S_n} W(q_{\sigma(1)}) \cdot W(q) \cdot \ldots \cdot \cdot W(q) \cdot W(q_{\sigma(n)})
+\bigcup \limits_{\sigma \in S_n} W(q_{\sigma(1)}) \cdot W(q) \cdot \ldots \cdot W(q) \cdot W(q_{\sigma(n)})
 \\
 K( \perm{(q_1, \ldots, q_n)}{q} ) = &
- K(q_1) \cdot K(q_1) \cdot \ldots \cdot \cdot K(q) \cdot K(q_n) 
+ K(q_1) \cdot K(q) \cdot \ldots \cdot K(q) \cdot K(q_n) 
 \end{array}
 \]
-  \caption{Whole and Kernel Regular Expressions.}
+  \caption{Whole and Kernel Regular Expressions\bcp{left-justify columns,
+      and try putting the last two clauses into the same format as the
+      others (using $\mathit{qq} = ...$ and maybe reducing spacing around
+      cdots, etc.)}}
   \label{fig:wk}
@@
     \equiv_{q_i} \; r'_i$ and $s_k \; \equiv_{q} \; s'_{k}$.
-  \caption{QRE Equivalence Relations}
+  \caption{QRE Equivalence Relations\bcp{too wide -- remove LH column and
+      try to get the last clause into the same table}}
   \label{fig:relations}
@@
 \end{theorem}
-\section{Using QREs to Express Quotient Lenses}
+
+\section{QRE Lenses}
+\label{QRE-lenses}
+
 Given regular expressions $R, S$ and equivalence relations $\sim_R, \sim_S$ are
@@
 semantics of $\mathcal{Q}$, as well as inference rules that may be used to
-derive elements of $\mathcal{Q}$:
+derive elements of $\mathcal{Q}$:\bcp{I think the numbering and the ``if $q$
+  = ..." stuff is not needed.  And maybe we can save space by typesetting the
+  get/put parts on the same line as the typing rule, at least in some cases.}
 \begin{enumerate}
@@
 \end{proof}
-\section{Synthesizing Quotient Lenses}
+
+\section{Synthesizing QRE Lenses}
+\label{synth}
+
 By \cref{normal form}, if there is a derivation $q : c \Leftrightarrow c'$ in
@@
 \section{Implementation and Evaluation}
+\label{impl}
+
 \section{Related Work}
+\label{relwork}
 
-\section{Conclusion}
+\bcp{Make sure to list/discuss:
+  \begin{itemize}
+  \item Seminar papers on lenses (to which we can refer readers for further
+  reading about connections to view updates, etc.) and quotient lenses.
+  \item Discuss what other lens-like programming languages (besides
+  boomerang) do about quotienting.  E.g., Augeas, XSugar, ...
+  \item How quotient lenses are used in Boomerang.
+  \item Search for papers that cite the quotient lens paper and see if they
+  extend the ideas there.
+  \end{itemize}
+}
+
+\section{Conclusion and Future Work}
+\label{concl}
 

### 2017-10-06 papers/QuotientLensSynth/main.tex
More explanation added in lenses section

@@
 % FOR Regular Expression names
-\newcommand{\re}[1]{\ensuremath{\mathit{#1}}}
-\newcommand{\squash}[3]{\ensuremath{\mathit{squash}(#1, #2, #3)}}
-\newcommand{\perm}[2]{\ensuremath{\mathit{perm}(#1)\; \mathit{with}\; #2}}
-\newcommand{\normalize}[3]{\ensuremath{\mathit{normalize}(#1, #2, #3)}}
+\newcommand{\re}[1]{\ensuremath{\mathtt{#1}}}
+\newcommand{\squash}[3]{\ensuremath{\mathit{squash} \; (#1, #2, #3)}}
+\newcommand{\perm}[2]{\ensuremath{\mathit{perm}\; (#1)\; \mathit{with}\; #2}}
+\newcommand{\normalize}[3]{\ensuremath{\mathit{normalize} \; (#1, #2, #3)}}
 \newcommand{\sep}{\ensuremath{\; | \;}}
+\newcommand{\canonizer}{\ensuremath{\mathit{Canonizer}}}
 
@@
                  William Clowes & Sons Ltd},
+
     Title = {Regular Algebra and Finite Machines},
+
     Year = {1971},
+
     Author = {Conway, J. H.},
@@
 \item Using this normal form, we reduce the problem of {\em synthesizing}
-QRE lenses from types and examples to the previously studied problem of
-synthesizing bijective lenses between regular languages
+QRE lenses from their QRE types and a set of examples to the previously studied
+problem of synthesizing bijective lenses between regular languages 
 (Section~\ref{synth}).
@@
 synthesize QRE lenses between several real-world data formats from the
-{\tt data.go} database (Section~\ref{impl}).
+{\tt data.gov} database (Section~\ref{impl}).
 \end{enumerate}
@@
   $EqRel(q)$, and
-  \item a ``canonizing'' function\bcp{Words in math mode need $\backslash${\tt
-      mathit} around them, always.} $Canonizer(q):\mathcal{L}(W(q))
+  \item a ``canonizing'' function $\canonizer(q):\mathcal{L}(W(q))
   \longrightarrow \mathcal{L}(K(q))$ which given any $w \in \mathcal{L}(W(q))$,
-  computes $Canonizer(q)(w)$ as the unique $k$ in $\mathcal{L}(K(q))$ such that
+  computes $\canonizer(q)(w)$ as the unique $k$ in $\mathcal{L}(K(q))$ such that
   $k$ is equivalent to $w$ mod $EqRel(q)$
   \end{enumerate}
-  Observe that since $Canonizer(q)$ computes $Canonizer(q)(w)$ as the unique $k$
-  in $\mathcal{L}(K(q))$ such that $w$ is equivalent to $k$ mod $EqRel(q)$,
+  Observe that since $\canonizer(q)$ computes $\canonizer(q)(w)$ as the unique
+  $k$ in $\mathcal{L}(K(q))$ such that $w$ is equivalent to $k$ mod $EqRel(q)$,
   then the equivalence classes of $EqRel(q)$ are the same as the fibres of
-  $Canonizer(q)$.
+  $\canonizer(q)$.
 
@@
     \item 
-    Every regular expression is a QRE. When a regular expression $R$ is used as
-    a QRE, then $\mathcal{L}(R)$ is assumed to be quotiented by the equality
-    equivalence relation on $\mathcal{L}(R)$.
+    Every regular expression can be lifted to a QRE. When a regular expression
+    $R$ is used as a QRE, then $\mathcal{L}(R)$ is assumed to be quotiented by
+    the equality equivalence relation on $\mathcal{L}(R)$.
     \item
@@
     when the programmer simply wishes to ignore unnecessary characters such as
-    excess whitespace e.g. \bcp{Suggest using $\backslash${\tt mathtt} for 
-      concrete program identifiers and ascii characters in examples.}
-    $$FIRST\_NAME \cdot (WHITESPACE^* \mapsto `` \quad ") \cdot LAST\_NAME$$
+    excess whitespace e.g. 
+    $$\mathtt{FIRST\_NAME} \cdot (\mathtt{WHITESPACE}^* \mapsto `` \quad ")
+    \cdot \mathtt{LAST\_NAME}$$
     \item
@@
     representation to the second. For instance, the first representation may be
-    $$FIRST\_NAME \cdot `` \quad " \cdot LAST\_NAME$$
+    $$\mathtt{FIRST\_NAME} \cdot `` \quad " \cdot \mathtt{LAST\_NAME}$$
     while the second representation may be
-    $$FIRST\_NAME \cdot ``," \cdot LAST\_NAME$$
+    $$\mathtt{FIRST\_NAME} \cdot ``," \cdot \mathtt{LAST\_NAME}$$
     The mapping that maps the first representation to the second simply converts
@@
     $$\perm{\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},
-    \re{YEAR}}{(\re{WSP}^* \mapsto \string \n)}$$
+    \re{YEAR}}{(\re{WHITESPACE}^* \mapsto \string \n)}$$
  where $\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},$ and $\re{YEAR}$ are
 regular expressions describing the respective \textsc{Bib}\TeX\; fields of the
-same name, and $WSP$ is a regular expression describing whitespace characters.
+same name, and $\re{WHITESPACE}$ is a regular expression describing whitespace
+characters.
 \item
@@
 regular expression constructs. However, the regular constructs for QREs are more
-restrictive for QREs since the data should match the whole language $W(q)$ and
-the kernel language $K(q)$ unambiguosly so as to make transformations
-deterministic.
+restrictive for QREs since the transformations made to the data by
+canonizers and lenses involve matching the data against a regular expression
+deterministic. Therefore to make these transformations deterministic, the
+data should be matched uniquely against the regular expressions, and this
+constrains the regular expressions that may be used by QREs.
 \item
@@
   unique representative $k$ is chosen using the ``canonizing'' function
-  $Canonizer(q)$.
+  $\canonizer(q)$.
   \fi
@@
 \; | \;  *$ to QREs, we require that the regular languages
-described by QREs match the data in only one way so as to make transformations
-deterministic. Consequently, we require that regular expressions used
-by QREs be \textit{strongly unambiguous}, a condition which ensures that if a
-string $s$ matches a regular expression $R$, then $s$ matches $R$ uniquely.
+described by QREs match the data in only one way so as to make
+various transformations made to the data that involve matching the data against
+a regular expression deterministic. Consequently, we require that regular
+expressions used by QREs be \textit{strongly unambiguous}, a condition which
+ensures that if a string $s$ matches a regular expression $R$, then $s$ matches
+$R$ uniquely.
 
@@
 kernel language $K(q)$, the equivalence relation $EqRel(q)$ and the
-canonizing function $Canonizer(q)$ are given in Figures ~\ref{fig:wk},
+canonizing function $\canonizer(q)$ are given in Figures ~\ref{fig:wk},
 ~\ref{fig:relations} and ~\ref{fig:canonizers}.
@@
   \[
-    \begin{array}{r@{\quad}c@{\quad}cc}
+    \begin{array}{l@{\quad}l@{\quad}l}
 
@@
 \begin{array}{r@{\quad}l}
-W( \perm{(q_1, \ldots, q_n)}{q} ) = &
-\bigcup \limits_{\sigma \in S_n} W(q_{\sigma(1)}) \cdot W(q) \cdot \ldots \cdot W(q) \cdot W(q_{\sigma(n)})
+W( \perm{q_1, \ldots, q_n}{q} ) = &
+\bigcup \limits_{\sigma \in S_n} W(q_{\sigma(1)}) \cdot W(q) \cdot \ldots \cdot
+W(q) \cdot W(q_{\sigma(n)})
 \\
-K( \perm{(q_1, \ldots, q_n)}{q} ) = &
+K( \perm{q_1, \ldots, q_n}{q} ) = &
  K(q_1) \cdot K(q) \cdot \ldots \cdot K(q) \cdot K(q_n) 
@@
 \begin{figure}[t]
-  \begin{center}
+  \centering
 \[
-    \begin{array}{l@{\quad}l} 
-      q & EqRel(q)  \\ \hline
-      R & w \; \equiv_R \; w' \iff w = w \\
-      R \mapsto s & w \; \equiv_{R \mapsto s} \; w' \text{ for all }w, w'\\
-      \squash{R}{R'}{f} & w \; \equiv_{\squash{R}{R'}{f}} \; w' \iff f(w) = w'
+    \begin{array}{l@{\quad}l@{\quad}l} 
+      w \; \equiv_R \; w' &\iff& w = w \\
+      w \; \equiv_{R \mapsto s} \; w' \text{ for all }w, w'\\
+      w \; \equiv_{\squash{R}{R'}{f}} \; w' &\iff& f(w) = w'
       \text{ or } w = w' \\
-      \normalize{R}{R'}{f} & w \; \equiv_{\normalize{R}{R'}{f}} \; w' \iff
+      w \; \equiv_{\normalize{R}{R'}{f}} \; w' &\iff&
       f(w)=f(w') \text{ or }w = w'\\
-      q_2 \circ  q_1 &  w \; \equiv_{q_2 \circ q_1} \; w' \iff \exists k, k' \in
+      w \; \equiv_{q_2 \circ q_1} \; w' &\iff& \exists k, k' \in
   \mathcal{L}(K(q_2)) \text{ such that } w \; \equiv_{q_1} \; k, \; w' \;
   \equiv_{q_1} \; k', \text{ and } k \; \equiv_{q_2} \; k'\\
-      q_1 \cdot q_2 &  w \; \equiv_{q_1 \cdot q_2} \; w'  \iff w = r_1
+      w \; \equiv_{q_1 \cdot q_2} \; w'  &\iff& w = r_1
       \cdot r_2, \; w' = {r'}_1 \cdot {r'}_2 \text{ with } r_1 \; \equiv_{q_1}
       \; {r'}_1, \; r_2 \; \equiv{q_n} \; {r'}_2\\
-      q_1 \sep q_2 &  w \; \equiv_{q' \sep q} \; w' \iff w \; \equiv_{q_1} \; w'
+      w \; \equiv_{q' \sep q} \; w' &\iff& w \; \equiv_{q_1} \; w'
       \text{ or } \; w \; \equiv_{q_2} \; w'\\
-      q^* &  w \; \equiv_{q*} \; w' \iff w = r_1 \cdot \ldots \cdot r_n, \; w'
+      w \; \equiv_{q*} \; w' &\iff& w = r_1 \cdot \ldots \cdot r_n, \; w'
       = {r'}_1 \cdot \ldots \cdot {r'}_n \text{ and } r_i \equiv_{q} \; {r'}_i
-      \text{ for all } 1 \leq i \leq n\\
+      \\
+      w \; \equiv_{\perm{q_1, \ldots, q_n}{q}} \; w' &\iff& w = r_{\sigma(1)}
+      \cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}, \;
+    w' = {r'}_{\theta(1)} \cdot s'_1 \cdot \ldots \cdot s'_{n-1}
+    \cdot {r'}_{\theta(n)} \\
+    & & \text{ for some } \sigma, \theta \in S_n, \text{ with } r_i \;
+    \equiv_{q_i} \; r'_i \text{ and } s_k \; \equiv_{q} \; s'_{k}
     \end{array}
     \]
-    \end{center}
-    $w \; \equiv_{\perm{(q_1, \ldots, q_n)}{q}} \; w' $ if and only if
-
-    $w = r_{\sigma(1)} \cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}
-    \text{ and } w' = {r'}_{\theta(1)} \cdot s'_1 \cdot \ldots \cdot s'_{n-1}
-    \cdot {r'}_{\theta(n)}$ for some $\sigma, \theta \in S_n$, with $r_i \;
-    \equiv_{q_i} \; r'_i$ and $s_k \; \equiv_{q} \; s'_{k}$.
-  \caption{QRE Equivalence Relations\bcp{too wide -- remove LH column and
-      try to get the last clause into the same table}}
+  \caption{QRE Equivalence Relations}
   \label{fig:relations}
@@
 \[
-    \begin{array}{l@{\quad}l} 
-      q & Canonizer(q)  \\ \hline
-      R & Canonizer(R) = id_{\mathcal{L}(R)} \\
-      R \mapsto s & Canonizer(R \mapsto s)(w) = s\\
-      \squash{R}{R'}{f} & Canonizer(\squash{R}{R'}{f})(w) = 
+    \begin{array}{l@\quad l @\quad l} 
+      \canonizer(R) &=& id_{\mathcal{L}(R)} \\
+      \canonizer(R \mapsto s)(w) &=& s\\
+      \canonizer(\squash{R}{R'}{f})(w) &=& 
 \begin{cases}
@@
 \end{cases}\\
-      \normalize{R}{R'}{f} & Canonizer(\normalize{R}{R'}{f}) = f\\
-      \perm{(q_1, \ldots, q_n)}{q} & \\
-      q_2 \circ  q_1 & Canonizer(q' \circ q) = Canonizer(q') \circ Canonizer(q)\\
-      q_1 \cdot q_2 & Canonizer(q' \cdot q) = Canonizer(q') \cdot Canonizer(q)\\
-      q_1 \sep q_2 &  Canonizer(q_1 \sep q_2)(w) = 
+      \canonizer(\normalize{R}{R'}{f}) &=& f\\
+      \canonizer(q' \circ q) &=& \canonizer(q') \circ \canonizer(q)\\
+      \canonizer(q' \cdot q) &=& \canonizer(q') \cdot \canonizer(q)\\
+      \canonizer(q_1 \sep q_2)(w) &=& 
 \begin{cases}
-Canonizer(q_1)(w) & \text{if } w \in \mathcal{L}(W(q_1))\\
-Canonizer(q_2)(w) & \text{if } w \in \mathcal{L}(W(q_2))\\
+\canonizer(q_1)(w) & \text{if } w \in \mathcal{L}(W(q_1))\\
+\canonizer(q_2)(w) & \text{if } w \in \mathcal{L}(W(q_2))\\
 \end{cases}\\
-      q^* &  Canonizer(q^*) = Canonizer(q)^* \\
+      \canonizer(q^*) &=& \canonizer(q)^* \\
     \end{array}
@@
     \end{center}
-    $Canonizer(\perm{(q_1, \ldots, q_n)}{q})(r_{\sigma(1)}
+    $\canonizer(\perm{q_1, \ldots, q_n}{q})(r_{\sigma(1)}
 \cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \newline
-= (Canonizer(q_1)(r_1)) \cdot (Canonizer(q)(s_1)) \cdot \ldots \cdot
-(Canonizer(q_n)(r_n))$
-  \caption{QRE Canonizers\bcp{First column not needed (and, once it's gone,
-      maybe the = signs could be vertically aligned?)}}
+= (\canonizer(q_1)(r_1)) \cdot (\canonizer(q)(s_1)) \cdot \ldots \cdot
+(\canonizer(q_n)(r_n))$
+  \caption{QRE canonizers}
   \label{fig:canonizers}
@@
 \begin{prooftree}
-\AxiomC{}
-\UnaryInfC{$R$ is well formed}
+\AxiomC{$R$ is strongly unambiguous}
+\UnaryInfC{$\mathit{id}(R)$ is well formed}
 \end{prooftree}
@@
 The theorem below confirms that the semantics of QREs are consistent with their
-intended meaning:
+intended behaviour:
 \begin{theorem}
@@
   $EqRel(q)$,
-  \item $Canonizer(q):\mathcal{L}(W(q)) \longrightarrow \mathcal{L}(K(q))$ is a
+  \item $\canonizer(q):\mathcal{L}(W(q)) \longrightarrow \mathcal{L}(K(q))$ is a
   well-defined function, and
-  \item  given any $w \in \mathcal{L}(W(q))$, $Canonizer(q)(w)$ is the unique
+  \item  given any $w \in \mathcal{L}(W(q))$, $\canonizer(q)(w)$ is the unique
   $k$ in $\mathcal{L}(K(q))$ such that $k$ is equivalent to $w$ modulo
@@
 \label{QRE-lenses} 
-
-Given regular expressions $R, S$ and equivalence relations $\sim_R, \sim_S$ are
-defined on $\mathcal{L}(R)$ and $\mathcal{L}(S)$ respectively, a
-\textit{quotient lens} $q :
+As we have seen, QREs enable the programmer to express a regular expression $R$
+as well as an equivalence relation on $\mathcal{L}(R)$. QREs are therefore a
+good \textit{specification} language for quotient lenses. In this section of the
+paper, we introduce \textit{QRE Lenses}. QRE lenses are quotient lenses based on
+the Boomerang quotient lens combinators, but which use QREs to describe their
+source and view types.
+
+\subsection{Syntax of QRE Lenses}
+\subsubsection{Preliminaries}
+Given regular expressions $R, S$ and equivalence
+relations $\sim_R, \sim_S$ defined on $\mathcal{L}(R)$ and $\mathcal{L}(S)$
+respectively, a \textit{bijective quotient lens} $q :
 R/{\sim_R}{\Longleftrightarrow} S/{\sim_S}$ from $R$ to $S$ is
@@
 if $r \sim_R r'$ and $s \sim_S s'$ then $q.get \; r \; \sim_S q.get \; r'$
-and $q.put \; s \sim_R q.put \; s' \; s'$.\\
-Now let $c, c'$ be QREs. Since $K(c)$ and $K(c')$ form a complete set of
-representatives for $W(c)$ modulo $EqRel(c)$ and $W(c')$ modulo $EqRel(c')$
-respectively, then given a bijection $\ell : K(c) \Leftrightarrow K(c')$, we
-may define a quotient lens $q : W(c)/EqRel(c) \Leftrightarrow
-W(c')/EqRel(c')$ by 
-\begin{align*}
-q.get &= \ell \circ Canonizer(c)\\
-q.put &= \ell^{-1} \circ Canonizer(c')
-\end{align*}
-Define the class of \textit{bijective
-lenses} to be the set of bijections between regular languages created from Boomerang lens combinators.
+and $q.put \; s \sim_R q.put \; s' \; s'$.
+
+Our approach in defining QRE lenses is as follows. Let $c, c'$ be QREs. Given a
+bijection $\ell : K(c) \Leftrightarrow K(c')$, we may define a quotient lens $q
+: W(c)/EqRel(c) \Leftrightarrow W(c')/EqRel(c')$ by
+\begin{equation}\label{normalform}
+q.get = \ell \circ \canonizer(c) \text{ and } q.put = \ell^{-1} \circ
+\canonizer(c'),
+\end{equation}
+
+A potential problem with this approach is that quotient lenses of this form may
+not necessarily be closed under the action of the lens combinators. For
+example, these lenses may not be closed under composition, which is one of the
+combinators we would want to have quotient lens language.
+
+Consequently, in order to define quotient lenses, we need to choose a class of
+bijections that has just the right properties with respect to the quotient
+lens combinators. Our choice is to use the class of \textit{bijective
+lenses} which we identified and studied in our previous work~\cite{popl18}.
+Bijective lenses, which we briefly discuss in the next subsection, possess
+the properties needed to ensure that all QRE lenses have the normal form
+suggested in \cref{normalform}.
+
+\subsection{Bijective Lenses}
+We define the set of \textit{bijective lenses} to be the set of bijections
+between regular languages created from Boomerang lens combinators.
 The syntax for the language of bijective lenses is given by
-$$\ell := R \sep const(s, s') \sep  swap(\ell,
+$$\ell := \mathit{id} \; (R) \sep const(s, s') \sep  swap(\ell,
 \ell) \sep \ell \cdot \ell' \; |  \; (\ell \sep \ell') \sep \ell^* \;
 | \; \ell \circ \ell',$$ where $R$ ranges over regular expressions and $s$
-ranges over character strings.\\
-The denotation of a bijective lens $\ell$ is a function from 
-$\llbracket e \rrbracket$ of a bijective lens $\ell$ is defined
-inductively by
+ranges over character strings.
+
+The denotation of a lens $\ell$ is $\llbracket \ell \rrbracket \subseteq
+\mathit{String} \times \mathit{String}$. If $(s_1, s_2) \in \llbracket \ell
+\rrbracket$, then $\ell$ maps between $s_1$ and $s_2$.
+
 \begin{align*}
@@
 \end{align*}
-\textcolor{red}{[More explanation needed esp. inference rules for bijective
-lenses]}\\
-Having described the class of bijective lenses, we now define the class
-$\mathcal{Q}$ of quotient lenses, whose language is defined inductively by the
-following grammar:
-$$ q := id(\ell) \sep lquot(q, \ell) \sep rquot(\ell, q) \sep q
-\cdot q_2 \sep (q \sep q') \sep q^* \sep q \circ q',$$
-where $\ell$ ranges over bijective lenses 
-\subsection{Semantics of $\mathcal{Q}$}
-Every element in $\mathcal{Q}$ has a type $q :
+
+The $\mathit{const}(s, t)$ lens replaces the string $s$ with $t$ in the source
+data in the forward direction, and $t$ with $s$ in the view data in the backward
+direction. $\mathit{id}(R)$ applies the identity function to the source and view
+in $\mathcal{L}(R)$ in both directions. The composition lens $\ell' \circ \ell$
+applies $\ell$ followed by then $\ell'$ to the source in the forward direction,
+and applies $\ell'$ followed by $\ell$ to the view in the backward direction.
+The lens $\ell \cdot \ell'$ first splits the string $s$ into $s_1$ and $s_2$,
+applies $\ell$ and $\ell'$ to $s_1$ and $s_2$ to get $t_1$ and $t_2$
+respectively, then concatenates $t_1$ and $t_2$ and returns $t_1 \cdot t_2$ as
+the final result in the forward direction. $\ell \cdot \ell'$ operates
+similarly in the backward direction, but with $s, s_1$ and $s_2$ substituted
+for $t, t_1$ and $t_2$. The $\mathit{swap} \; (\ell, \ell')$ lens operates
+like $\ell \cdot \ell'$, except that it swaps $t_1$ and $t_2$ before
+concatenating the two for a final result of $t_2 \cdot t_1$ in the forward
+direction. In the backward direction, $\mathit{swap}(\ell, \ell')$ first undoes
+the swap, then proceeds as expected. The $\ell \; | \; \ell'$ lens
+chooses to apply $\ell$ or $\ell'$ depending on whether the source
+(resp. view) data is matched by $\ell$ or $\ell'$ in the forward (resp.
+backward direction. The $\ell^*$ lens splits the string $s$ into strings $s_1,
+\ldots, s_n$, applies $\ell$ to each $s_i$ to get $t_i$, and then concatenates
+each of the $t_i$'s for a final result of $t_1 \cdot \ldots \cdot t_n$ in the
+forward direction. $\ell^*$ operates similarly in the backward direction, but
+with $s, s_i$ substituted for $t, t_i$.
+
+Each bijective lens $\ell$ has a type $\ell : R \Leftrightarrow S$ where $R$ and
+$S$ are regular expressions. If $\ell : R \Leftrightarrow S$, then the source
+type of $\ell$ is $\mathcal{L}(R)$ and the view type of $\ell$ is
+$\mathcal{L}(S)$. Interestingly, with the bijective lens type system, a lens
+$\ell : R \Leftrightarrow S$ can also be considered to be of type $\ell : R'
+\Leftrightarrow S'$ provided that $R$ (resp. $S$) can be proven to be
+equivalent to $R'$ (resp. $S$) from the star-semiring axioms. The typing rules
+for bijective lenses are given in Figure~\ref{fig:lensrules}.
+
+ \begin{figure}[t]
+  \begin{prooftree}
+\AxiomC{$s_1 \in \Sigma^*$}
+\AxiomC{$s_2 \in \Sigma^*$}
+\BinaryInfC{$\mathit{const} \; (s_1, s_t): s_1 \Leftrightarrow s_2$}
+\end{prooftree}
+
+\begin{prooftree}
+\AxiomC{$R$ is strongly unambiguous}
+\UnaryInfC{$\mathit{id} \; (R): R \Leftrightarrow R$}
+\end{prooftree}
+
+\begin{prooftree}
+\AxiomC{$\ell_1 : R_1 \Leftrightarrow S_1$}
+\AxiomC{$\ell_2 : R_2 \Leftrightarrow S_2$}
+\AxiomC{$R_1 \cdot^! R_2$}
+\AxiomC{$S_1 \cdot^! S_2$}
+\QuaternaryInfC{$\ell_1 \cdot \ell_2: R_1 \cdot R_2 \Leftrightarrow S_1 \cdot
+S_2$}
+\end{prooftree}
+
+\begin{prooftree}
+\AxiomC{$\ell_1 : R_1 \Leftrightarrow R_2$}
+\AxiomC{$\ell_2 : R_2 \Leftrightarrow R_3$}
+\BinaryInfC{$\ell_2 \circ \ell_1: R_1 \Leftrightarrow R_3$}
+\end{prooftree}
+
+\begin{prooftree}
+\AxiomC{$\ell_1 : R_1 \Leftrightarrow S_1$}
+\AxiomC{$\ell_2 : R_2 \Leftrightarrow S_2$}
+\AxiomC{$R_1 \cdot^! R_2$}
+\AxiomC{$S_2 \cdot^! S_1$}
+\QuaternaryInfC{$\mathit{swap} \; (\ell_1, \ell_2): R_1 \cdot R_2
+\Leftrightarrow S_2 \cdot S_1$}
+\end{prooftree}
+
+\begin{prooftree}
+\AxiomC{$\ell_1 : R_1 \Leftrightarrow S_1$}
+\AxiomC{$\ell_2 : R_2 \Leftrightarrow S_2$}
+\AxiomC{$\mathcal{L}(R_1) \cap \mathcal{L}(R_2) = \varnothing$}
+\AxiomC{$\mathcal{L}(S_1) \cap \mathcal{L}(S_2) = \varnothing$}
+\QuaternaryInfC{$\ell_1 \sep \ell_2: R_1 \sep R_2 \Leftrightarrow S_1 \sep S_2$}
+\end{prooftree}
+
+\begin{prooftree}
+\AxiomC{$\ell : R \Leftrightarrow S$}
+\AxiomC{$R^{*!}$}
+\AxiomC{$S^{*!}$}
+\TrinaryInfC{$\ell^*: R \Leftrightarrow S$}
+\end{prooftree}
+
+\begin{prooftree}
+\AxiomC{$\ell : R \Leftrightarrow S$}
+\AxiomC{$R \equiv R'$}
+\AxiomC{$S \equiv S'$}
+\TrinaryInfC{$\ell : R' \Leftrightarrow S'$}
+\end{prooftree}
+  \caption{Bijective Lens Typing Rules}
+  \label{fig:lensrules}
+\end{figure}
+
+\subsection{Syntax of QRE Lenses}
+Having described the class of bijective lenses, we now introduce the class
+of QRE lenses, whose language is given by following grammar:
+$$ q := \mathit{lift}(\ell) \sep \mathit{lquot}(q, \ell) \sep
+\mathit{rquot}(\ell, q) \sep q \cdot q' \sep (q \sep q') \sep q^* \sep q \circ q',$$
+where $\ell$ ranges over bijective lenses.
+
+The QRE lens combinators are inspired by the Boomerang lens combinators of the
+same name. The $\mathit{lift}(\ell)$ quotient lens enables a bijective lens
+$\ell$ to be considered a quotient lens where the equivalence relation applied
+to the source and view types is the equality relation. The $\mathit{lquot}(c,
+q)$ combinator takes a quotient lens $q$ and quotients the source data using
+$c$, assuming that the source data forms a complete set of representatives for
+the equivalence relation $EqRel(c)$. The $\mathit{lquot}(q, c)$ does the same
+but on the view data. The regular combinators $\cdot \; * \; |$ exhibit the
+expected behaviour. \sam{Examples needed?}
+
+The difference between QRE lenses and Boomerang quotient lenses is the typing
+judgements. More concretely, each QRE lens $q$ has a type $q : c \Leftrightarrow
+c'$ where $c, c'$ are QREs. In contrast, the approach used to type quotient
+lenses in Boomerang is to classify equivalences according to whether they are
+or are not the equality relation. This type system is based on two
+observations: first, that most quotient lenses originate as lifted basic
+lenses, and therefore have types whose equivalence relations are both equality;
+and second, that equality is preserved by many of the quotient lens
+combinators. Foster et al also discuss a second possible approach to typing
+quotient lenses, where equivalence relations are represented by rational
+functions that induce them. While this second approach is more refined than the
+first, Boomerang favours the first approach since the second appears to be too
+expensive to be useful in practice~\cite{quotientlenses}.
+
+\subsection{Semantics of QRE Lenses}
+As we mentioned earlier, every quotient lens $q$ has a type $q :
 c \Longleftrightarrow c'$ where $c$ and $c'$ are QREs. The denotation
-$\llbracket q \rrbracket$ of an element $q : c \Longleftrightarrow c'$ is a
-quotient lens $\llbracket q \rrbracket :
-W(c)/{EqRel(c)} \Longleftrightarrow W(c')/{EqRel(c')}$.\\
-\textcolor{red}{[Informal Meaning/Usages Needed]}\\We now give the
-semantics of $\mathcal{Q}$, as well as inference rules that may be used to
-derive elements of $\mathcal{Q}$:\bcp{I think the numbering and the ``if $q$
-  = ..." stuff is not needed.  And maybe we can save space by typesetting the
-  get/put parts on the same line as the typing rule, at least in some cases.}
-\begin{enumerate}
-  \item
-  if $q = id(\ell)$, then
+$\llbracket q \rrbracket$ of a QRE lens $q:c \Leftrightarrow c'$ is a quotient
+lens $\llbracket q \rrbracket : W(c)/{EqRel(c)} \Longleftrightarrow
+W(c')/{EqRel(c')}$. The typing rules and denotation of QRE lenses are given in
+Figure~\ref{fig:qlenssemantics}. The following theorem confirms that QRE lenses
+exhibit the intended behaviour.
+
+\begin{theorem}
+If there is a derivation $q:c \Leftrightarrow c'$ then
+$\llbracket q \rrbracket : W(c)/EqRel(c) \Leftrightarrow W(c')/EqRel(c')$ is a
+well-defined quotient lens.
+\end{theorem}
+
+ \begin{figure}[t]
+  \centering
   \begin{prooftree}
 \AxiomC{$\ell : R \Leftrightarrow S$}
-\UnaryInfC{$id(\ell): R \Leftrightarrow id(S)$}
+\UnaryInfC{$\mathit{lift}(\ell): R \Leftrightarrow id(S)$}
 \end{prooftree}
@@
 
-  \item if $q = lquot(c, q')$, then
 \begin{prooftree}
@@
 \AxiomC{$K(c) = W(c')$}
-\TrinaryInfC{$lquot(c, q'): c' \circ c \Leftrightarrow c''$}
+\TrinaryInfC{$\mathit{lquot}(c, q'): c' \circ c \Leftrightarrow c''$}
 \end{prooftree}
@@
   \llbracket q \rrbracket.get  &= \llbracket q'
-  \rrbracket.get \circ Canonizer(c)\\
+  \rrbracket.get \circ \canonizer(c)\\
   \llbracket q \rrbracket.put &= \llbracket q' \rrbracket.put
   \end{align*}
-  \item
-  if $q = rquot(q', c'')$, then
+
   \begin{prooftree}
@@
 \AxiomC{$K(c'') = W(c')$}
-\TrinaryInfC{$rquot(q', c):c \Leftrightarrow c'' \circ c'$}
+\TrinaryInfC{$\mathit{rquot}(q', c):c \Leftrightarrow c'' \circ c'$}
 \end{prooftree}
@@
   \llbracket q \rrbracket.put &= \llbracket q'
-  \rrbracket.put \circ Canonizer(c'')
-  \end{align*}
-
-  \item
-  $q = q_2 \circ q_1$, then
+  \rrbracket.put \circ \canonizer(c'')
+  \end{align*}
+
   \begin{prooftree}
@@
   \end{align*}
-  \item
-  $q = {q'}^*$, then
+
     \begin{prooftree}
@@
 
-  \item
-  $q = q_1 \cdot q_2$, then
     \begin{prooftree}
@@
 
-  \item
-  $q = q_1 \sep q_2$, then
       \begin{prooftree}
@@
   $$
-\end{enumerate}
-
-\begin{claim}
-If there exists a derivation $q:c \Leftrightarrow c'$ then
-$\llbracket q \rrbracket$ is a quotient lens from $R$ to $S$. \bcp{??}
-\end{claim}
-
-\subsection{Normal Forms for Quotient Lenses in $\mathcal{Q}$}
+  \caption{Denotation and Typing Rules for QRE Lenses}
+  \label{fig:qlenssemantics}
+\end{figure}
+
+\section{Normal Forms QRE Lenses}
+\sam{Transition needed?}
 \begin{theorem}\label{normal form}
-If there is a derivation $q : c \Leftrightarrow c'$ in
-$\mathcal{Q}$ then there exists a bijective lens $\ell : K(c) \Leftrightarrow
-K(c')$ such that
-\begin{align*}
-\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ Canonizer(c)\\
+If there is a derivation $q : c \Leftrightarrow c'$, then there exists a
+bijective lens $\ell : K(c) \Leftrightarrow K(c')$ such that
+\begin{align*}
+\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ \canonizer(c)\\
 \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
-Canonizer(c')
+\canonizer(c')
 \end{align*}
@@
   \item
-  $id(\ell): R \Leftrightarrow id(S)$ where $\ell : R \Leftrightarrow
-S$. Then
-  \begin{align*}
-  \llbracket id(\ell) \rrbracket.get &=  \llbracket \ell \rrbracket = \llbracket \ell \rrbracket \circ
-  id_{\mathcal{L}(R)} = \llbracket \ell \rrbracket \circ Canonizer(R), \text{ and }\\
-  \llbracket id(\ell) \rrbracket.put &= \llbracket \ell \rrbracket^{-1} = \llbracket \ell \rrbracket^{-1} \circ
-  id_{\mathcal{L}(S)} = \llbracket \ell \rrbracket^{-1} \circ Canonizer(id(S))
-  \end{align*}
-  \item
-  $lquot(c, q'): c' \circ c \Leftrightarrow c''$ where $q' : c' 
+  $\mathit{lift}(\ell): R/\mathit{id}(R) \Leftrightarrow S/\mathit{id}(S)$ where
+  $\ell :
+  R \Leftrightarrow S$. Then
+  \begin{align*}
+  \llbracket \mathit{lift}(\ell) \rrbracket.get &=  \llbracket \ell \rrbracket
+  = \llbracket \ell \rrbracket \circ id_{\mathcal{L}(R)} =
+  \llbracket \ell \rrbracket \circ \canonizer(\mathit{id}(R)), \text{ and }\\
+  \llbracket \mathit{lift}(\ell) \rrbracket.put &= \llbracket \ell
+  \rrbracket^{-1} = \llbracket \ell \rrbracket^{-1} \circ id_{\mathcal{L}(S)} =
+  \llbracket \ell \rrbracket^{-1} \circ \canonizer(id(S))
+  \end{align*}
+  \item
+  $\mathit{lquot}(c, q'): c' \circ c \Leftrightarrow c''$ where $q' : c' 
   \Leftrightarrow c''$, $c$ is well formed and $K(c) = W(c')$. Then
@@
   \llbracket q \rrbracket.get  &= \llbracket q'
-  \rrbracket.get \circ Canonizer(c)\\
+  \rrbracket.get \circ \canonizer(c)\\
   \llbracket q \rrbracket.put &= \llbracket q' \rrbracket.put
@@
   \begin{align*}
-\llbracket q' \rrbracket.get &= \llbracket \ell \rrbracket \circ Canonizer(c')\\
+\llbracket q' \rrbracket.get &= \llbracket \ell \rrbracket \circ \canonizer(c')\\
 \llbracket q' \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
@@
   \llbracket q \rrbracket.get  &= (\llbracket \ell \rrbracket \circ
-  Canonizer(c')) \circ Canonizer(c) = \llbracket \ell \rrbracket \circ
-  (Canonizer(c' \circ c))\\
+  \canonizer(c')) \circ \canonizer(c) = \llbracket \ell \rrbracket \circ
+  (\canonizer(c' \circ c))\\
   \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
@@
   \item
-  $rquot(q', c''):c \Leftrightarrow c'' \circ c'$ where $q' : c \Leftrightarrow
+  $\mathit{rquot}(q', c''):c \Leftrightarrow c'' \circ c'$ where $q' : c \Leftrightarrow
   c'$, $c''$ is well formed and $K(c'') = W(c')$. Proceed as in the previous
@@
 \llbracket q_1 \rrbracket.get &= \llbracket \ell_1 \rrbracket \circ
-Canonizer(c)\\
+\canonizer(c)\\
 \llbracket q_1 \rrbracket.put &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
-Canonizer(c')
+\canonizer(c')
 \end{align*}
@@
 \llbracket q_2 \rrbracket.get \circ \llbracket q_1 \rrbracket.get &=
-(\llbracket \ell_2 \rrbracket \circ Canonizer(c')) \circ (\llbracket \ell_1
-\rrbracket \circ Canonizer(c))\\
-&= \llbracket \ell_2 \rrbracket \circ (Canonizer(c') \circ \llbracket \ell_1
-\rrbracket) \circ Canonizer(c)\\
+(\llbracket \ell_2 \rrbracket \circ \canonizer(c')) \circ (\llbracket \ell_1
+\rrbracket \circ \canonizer(c))\\
+&= \llbracket \ell_2 \rrbracket \circ (\canonizer(c') \circ \llbracket \ell_1
+\rrbracket) \circ \canonizer(c)\\
 &= (\llbracket \ell_2 \rrbracket \circ \llbracket \ell_1 \rrbracket) \circ
-Canonizer(c)\\
+\canonizer(c)\\
 &= \llbracket \ell_2  \circ  \ell_1 \rrbracket \circ
-Canonizer(c)
+\canonizer(c)
 \end{align*} 
@@
 \llbracket \ell_2  \circ  \ell_1 \rrbracket^{-1} \circ
-Canonizer(c)$$
+\canonizer(c)$$
 \item  
@@
 \llbracket q' \rrbracket.get &= \llbracket \ell \rrbracket \circ
-Canonizer(c)\\
+\canonizer(c)\\
 \llbracket q' \rrbracket.put &= {\llbracket \ell \rrbracket}^{-1} \circ
-Canonizer(c')
+\canonizer(c')
 \end{align*}
@@
 \llbracket {q'}^* \rrbracket.get &= (\llbracket \ell \rrbracket \circ
-Canonizer(c))^* = \llbracket \ell \rrbracket^* \circ
-Canonizer(c)^* = \llbracket \ell^* \rrbracket \circ
-Canonizer(c^*)\\
+\canonizer(c))^* = \llbracket \ell \rrbracket^* \circ
+\canonizer(c)^* = \llbracket \ell^* \rrbracket \circ
+\canonizer(c^*)\\
 \llbracket {q'}^* \rrbracket.put &= (\llbracket \ell \rrbracket^{-1} \circ
-Canonizer(c'))^* = (\llbracket \ell \rrbracket^{-1})^* \circ
-Canonizer(c')^* = \llbracket \ell^* \rrbracket^{-1} \circ
-Canonizer(c'^*)\\
+\canonizer(c'))^* = (\llbracket \ell \rrbracket^{-1})^* \circ
+\canonizer(c')^* = \llbracket \ell^* \rrbracket^{-1} \circ
+\canonizer(c'^*)\\
 \end{align*}
@@
 \llbracket q_1 \rrbracket.get &= \llbracket \ell_1 \rrbracket \circ
-Canonizer(c_1)\\
+\canonizer(c_1)\\
 \llbracket q_1 \rrbracket.put &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
-Canonizer(d_1)
+\canonizer(d_1)
 \end{align*}
@@
 \llbracket q_2 \rrbracket.get &= \llbracket \ell_2 \rrbracket \circ
-Canonizer(c_2)\\
+\canonizer(c_2)\\
 \llbracket q_2 \rrbracket.put &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
-Canonizer(d_2)
+\canonizer(d_2)
 \end{align*}
@@
   \llbracket q \rrbracket.get &= (\llbracket \ell_1 \rrbracket \circ
-Canonizer(c_1)) \cdot  (\llbracket \ell_2 \rrbracket \circ
-Canonizer(c_2))\\
+\canonizer(c_1)) \cdot  (\llbracket \ell_2 \rrbracket \circ
+\canonizer(c_2))\\
 &= (\llbracket \ell_1 \rrbracket \cdot \llbracket \ell_2
-\rrbracket) \circ (Canonizer(c_1) \cdot Canonizer(c_2))\\
-&= \llbracket \ell_1 \cdot  \ell_2 \rrbracket \circ Canonizer(c_1 \cdot c_2)
+\rrbracket) \circ (\canonizer(c_1) \cdot \canonizer(c_2))\\
+&= \llbracket \ell_1 \cdot  \ell_2 \rrbracket \circ \canonizer(c_1 \cdot c_2)
 \end{align*}
@@
   \llbracket q \rrbracket.put = \llbracket \ell_1 \cdot  \ell_2 \rrbracket^{-1}
-  \circ Canonizer(d_1 \cdot d_2) $$
+  \circ \canonizer(d_1 \cdot d_2) $$
   \item
@@
   \llbracket \ell_1 \rrbracket \circ
-Canonizer(c_1) (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
+\canonizer(c_1) (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
   \llbracket \ell_2 \rrbracket \circ
-Canonizer(c_2) (s) & \text{if } s \in \mathcal{L}(W(c_2)),\\
+\canonizer(c_2) (s) & \text{if } s \in \mathcal{L}(W(c_2)),\\
   \end{cases}$$
   so $\llbracket q_1 \sep q_2 \rrbracket.get = \llbracket \ell_1 \sep
-  \ell_2 \rrbracket \circ Canonizer(c_1 \sep c_2)$. A similar argument shows
+  \ell_2 \rrbracket \circ \canonizer(c_1 \sep c_2)$. A similar argument shows
   that $\llbracket q_1 \sep q_2 \rrbracket.put = \llbracket \ell_1 \sep
-  \ell_2 \rrbracket^{-1} \circ Canonizer(d_1 \sep d_2)$.\\
+  \ell_2 \rrbracket^{-1} \circ \canonizer(d_1 \sep d_2)$.\\
   This completes the proof.
@@
 \begin{align*}
-\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ Canonizer(c)\\
+\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ \canonizer(c)\\
 \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
-Canonizer(c')
+\canonizer(c')
 \end{align*}

### 2017-10-06 papers/QuotientLensSynth/main.tex
Started working on synthesis section

@@
 
-\section{Normal Forms QRE Lenses}
-\sam{Transition needed?}
+\subsection{Normal Forms QRE Lenses}
+Recall that our approach in defining QRE lenses is to have each QRE lens $q: c
+\Leftrightarrow c'$ be such that 
+\begin{align*}
+\llbracket q \rrbracket.get &= \ell \circ \canonizer(c)\\
+\llbracket q \rrbracket.put &= \ell^{-1} \circ
+\canonizer(c')
+\end{align*}
+for some bijective lens $\ell$. We now provide a proof that all QRE lenses are
+of this form.
 \begin{theorem}\label{normal form}
@@
 \label{synth}
-
-By \cref{normal form}, if there is a derivation $q : c \Leftrightarrow c'$ in
-$\mathcal{Q}$ then there exists a bijective lens $\ell : K(c) \Leftrightarrow
-K(c')$ such that
+By Theorem~\ref{normal form}, if there is a derivation $q : c \Leftrightarrow
+c'$ of a QRE lens, then there exists a bijective lens $\ell : K(c)
+\Leftrightarrow K(c')$ such that
 \begin{align*}
@@
 \end{align*}
-In other words, every quotient lens in $\mathcal{Q}$ is the same as an ordinary
-bijective lens with canonizers at the ends.\\
-Therefore, in order to synthesize a quotient lens $q: c \Leftrightarrow c'$
-where $c, c'$ are QREs, it suffices to synthesize a bijective lens $\ell : K(c)
-\Leftrightarrow K(c')$.\\
-\textcolor{red}{[Citation Needed!]} has already shown that if there
-exists a derivation $\ell : R \Leftrightarrow S$, then there exist
-regular expressions $R', S'$ and a synthesizable \\
-\textcolor{red}{[Explain and make explicit the meaning
-of synthesizable]} \\
-bijective lens $\ell' :
-R' \Leftrightarrow S'$ such that $\mathcal{L}(R) = \mathcal{L}(R')$,
-$\mathcal{L}(S) = \mathcal{L}(S')$ and $\llbracket \ell \rrbracket =
-\llbracket \ell' \rrbracket$. This immediately implies that if there is a
-derivation $q : c \Leftrightarrow c'$ as well as a derivation $\ell : K(c)
-\Leftrightarrow K(c')$, then $q$ is synthesizable.
-\textcolor{red}{[More Needed?]}
+
+Now suppose that we wish to synthesize a quotient lens $q$ of type $q: c
+\Leftrightarrow c'$ that maps a specified set of input strings to a
+specified set of output strings and vice versa.
+
+For example, in the \textsc{Bib}\TeX\ to EndNote problem, the programmer wishes
+to synthesize a QRE lens $q$ which maps the QRE
+$$\perm{\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},
+\re{YEAR}}{(\re{WHITESPACE}^* \mapsto \string \n)}$$ 
+to the QRE
+$$\perm{\re{REF'}, \re{AUTHOR'}, \re{TITLE'}, \re{PUBLISHER'},
+\re{YEAR'}}{(\re{WHITESPACE}^* \mapsto \string \n)}$$
+where $\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},$ and $\re{YEAR}$
+(resp. $\re{REF'}, \re{AUTHOR'}, \re{TITLE'}, \re{PUBLISHER'},$ and
+$\re{YEAR'}$) are regular expressions describing the respective
+\textsc{Bib}\TeX\; (resp. EndNote) fields of the same name, and
+$\re{WHITESPACE}$ is a regular expression describing whitespace
+characters. Morever, the synthesized lens should map the equivalence
+class of this \textsc{Bib}\TeX\; record
+\begin{verbatim}
+@Book {conway,
+  Author = {Conway, J. H.},
+  Title = {Regular Algebra and Finite Machines},
+  Publisher = {Printed in GB by William Clowes & Sons Ltd},
+  Year = {1971}
+}
+\end{verbatim}
+\noindent to the equivalence class of this EndNote entry:
+\begin{verbatim}
+%0 Book
+%T Regular Algebra and Finite Machines
+%A Conway, J. H.
+%D 1971
+%I Printed in GB by William Clowes & Sons Ltd
+%F conway
+\end{verbatim}
+
+In our previous work, we showed that given regular expressions $R, S$ and a set
+of input-output pairs $\{(r_1, s_1), \ldots, (r_n, s_n)\}$, if there is a
+derivation $\ell :, R \Leftrightarrow S$ of a bijective lens that maps $r_i$ to
+$s_i$ in the forward direction and $s_i$ to $r_i$ in the backward direction,
+then there is an algorithm that computes a bijective lens $\ell' : R'
+\Leftrightarrow S'$ such that $R \equiv R'$, $S \equiv S'$, $\ell'$ maps $r_i$
+to $s_i$ in the forward direction and $s_i$ to $r_i$ in the backward
+direction, and $\llbracket \ell \rrbracket = \llbracket \ell'
+\rrbracket$~\cite{popl18}.
+
 \section{Implementation and Evaluation}

### 2017-10-06 papers/QuotientLensSynth/main.tex
a few comments

@@
 \newcommand{\afm}[1]{\FINISH{dkgreen}{A}{#1}}
-\newcommand{\dpw}[1]{\FINISH{dkblue}{D}{#1}}
+\newcommand{\dpw}[1]{\FINISH{dkblue}{D}{#1}} % Toronto Maple Leafs Blue :-)
 \newcommand{\saz}[1]{\FINISH{orange}{SZ}{#1}}
@@
 
+\dpw{The next paragraph gets at the problem, but takes a while to do
+so.  Why not make the first sentence go directly to the problem:  ``Unfortunately,
+quotient lens programs are unnecessarily verbose and challenging to construct.'' ... or
+something like that.}
+
 This paper addresses two main challenges that arise when programming quotient
@@
 \saz{It might be a good idea to show what this would look like in all its gory
-detail, just to make the problem very concrete. Perhaps in a figure?}\bcp{+1}
+detail, just to make the problem very concrete. Perhaps in a figure?}\bcp{+1}\dpw{+1}
 
 In this paper, we introduce the formalism of Quotient Regular Expressions (or
-QREs) to address this issue. QREs enable the programmer to express
+QREs) to address this issue. \dpw{cut unnecessary phrases that add little information
+content.  eg, cut ``the formalism of.''  Doing so makes your text ``punchier.''}  QREs enable the programmer to express
 a rich set of equivalence relations on regular expressions at the level of
@@
 
+\dpw{I think it would be a little more effective to introduce QREs with a running,
+but informal, concrete example. Explain these concepts informally as you go.
+Then summarize the syntax at the end.}
+
   \subsection{Examples}
@@
 data should be matched uniquely against the regular expressions, and this
-constrains the regular expressions that may be used by QREs.
+constrains the regular expressions that may be used by QREs. \dpw{This is not an
+``example'' -- it is an informal description of how these operators work.}
 \item

### 2017-10-06 papers/QuotientLensSynth/main.tex
Notes from meeting

@@
 \newcommand{\canonizer}{\ensuremath{\mathit{Canonizer}}}
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
-\noindent and vice versa.
+\noindent and vice versa.  \bcp{Let's make the example simpler (leave out
+  year and publisher?)}
 
@@
 \item We introduce a novel language of {\em Quotient Regular Expressions}
-(QREs) that provide compact, convenient notation for a broad class of
-equivalence relations on regular languages (Section~\ref{QRE}).
-\item We define a language of {\em QRE lenses}, a class of quotient lenses
+(QREs) that provide compact, convenient notation for a natural and useful
+class of equivalence relations on regular languages (Section~\ref{QRE}).
+\item We define a language of {\em QRE lenses}, a form of quotient lenses
 whose types are given by QREs.  Our main technical contribution is a normal
@@
   \caption{QRE Inference Rules\bcp{Maybe some or all of these can be
-      suppressed in the ``short version'' of the paper}}
+      suppressed in the ``short version'' of the paper}\bcp{No, cancel
+      that.  These are important.  And we need to comment about the decidability
+    issues involving normalize.  (And a note that we don't really use
+    normalize, but rather introduce more specialized instances.)}}
   \label{fig:qrerules}
@@
 example, these lenses may not be closed under composition, which is one of the
-combinators we would want to have quotient lens language.
+combinators we would want to have in the quotient lens language.\bcp{Can you
+say the explanation without using the phrase ``closed under''?}
 
@@
 
-\subsection{Normal Forms QRE Lenses}
+\subsection{Normal Forms of QRE Lenses}
 Recall that our approach in defining QRE lenses is to have each QRE lens $q: c
@@
 $\re{YEAR'}$) are regular expressions describing the respective
-\textsc{Bib}\TeX\; (resp. EndNote) fields of the same name, and
+\textsc{Bib}\TeX\; (resp. EndNote\bcp{Don't like this ``resp.''
+  construction---saves space, but not reader energy}) fields of the same
+name, and 
 $\re{WHITESPACE}$ is a regular expression describing whitespace
 characters. Morever, the synthesized lens should map the equivalence
-class of this \textsc{Bib}\TeX\; record
+class of this \textsc{Bib}\TeX{} record
 \begin{verbatim}

### 2017-10-08 program/generate_boomerang_specs.py
output specs in boom format

@@ -1,0 +1,78 @@
+#!/usr/local/bin/python
+
+from __future__ import print_function
+from easyprocess import EasyProcess
+
+import os
+from os.path import splitext, join
+import subprocess
+import sys
+import time
+
+TEST_EXT = '.ls'
+BASELINE_EXT = '.out'
+BOOM_EXT = '.boom'
+BASE_FLAGS = ['-generate_boomerang_spec']
+TIMEOUT_TIME = 6000
+REPETITION_COUNT = 2
+BASE_BOOM_DIR = 'generated_boomerang_specs/'
+
+def ensure_dir(f):
+    if not os.path.exists(f):
+        os.makedirs(f)
+
+def find_tests_and_ensure_output_dirs(root):
+    tests = []
+    for path, dirs, files in os.walk(root):
+        files = [(f[0], f[1]) for f in [splitext(f) for f in files]]
+        tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
+        ensure_dir(BASE_BOOM_DIR + path)
+    return tests
+
+def gather_boom(rootlength, prog, path, base):
+    process_output = EasyProcess([prog] + BASE_FLAGS + [join(path, base + TEST_EXT)]).call(timeout=TIMEOUT_TIME)
+    if (process_output.stderr != "" or process_output.return_code != 0):
+        raise Exception("failure")
+    return process_output.stdout
+
+def write_boom(boom, path, base):
+    with open(join(BASE_BOOM_DIR+path, base + BOOM_EXT),'w') as boomfile:
+        boomfile.write(boom)
+
+def print_usage(args):
+    print("Usage: {0} <prog> <testdir>".format(args[0]))
+
+def transform_data(path, base, run_data):
+    current_data = {"Test":join(path, base + TEST_EXT).replace("_","-")[6:]}
+    run_data_transpose = transpose(run_data)
+    for index in range(len(run_data_transpose)/2):
+	col_name = run_data_transpose[index][0]
+	col_data = run_data_transpose[index+1]
+        if "" in col_data:
+	    current_data[col_name]="Timeout"
+        else:
+            col = [float(x) for x in col_data]
+            current_data[col_name] = str(sum(col)/len(col))
+    return current_data
+
+def main(args):
+    if len(args) == 3:
+        prog = args[1]
+        path = args[2]
+        rootlength = len(path)
+        data = []
+        base_boom_dir = ensure_dir('generated_boomerang_specs/')
+        if not os.path.exists(prog):
+            print_usage(args)
[…]

### 2017-10-09 papers/QuotientLensSynth/main.tex
Made changes suggested from Friday. Need to add programs and diagrams!

@@
 \newcommand{\canonizer}{\ensuremath{\mathit{Canonizer}}}
+\newcommand{\bibtex}{\textsc{Bib}\TeX{}}
+\newcommand{\get}{\ensuremath{\mathit{get}}}
+\newcommand{\lput}{\ensuremath{\mathit{put}}}
+\newcommand{\create}{\ensuremath{\mathit{create}}}
+\newcommand{\eqrel}[1]{\ensuremath{\equiv_{#1}}}
 
@@
 \begin{abstract}
-
-\keywords{quotient lenses, quotient regular expressions, synthesis}
+{\em A Quotient lens} is an expression which encapsulates a bidirectional
+transformation whereby each of the two transformations that are part of the lens
+have the same behaviour on data that are equivalent to each other modulo some
+equivalence relation. Quotient lenses save the programmer the effort of writing
+two expressions instead of one, give guarantess on how the forward and backward
+transformations interact with each other, and also allow the programmer to
+ignore differences in the data which are not intended to affect the
+behaviour of the transformations.
+
+However, quotient lenses are sometimes difficult to program because the
+equivalence relations defined on the data, as well as the underlying lens which
+operates on the equivalence classes are complicated. These two challenges make
+quotient languages impractical to use in many cases spite of their advantages.
+
+In this paper, we address both of these challenges. First, we propose the
+language of {\em Quotient Regular Expressions} or QREs, which enables the
+programmer to explicitly define a broad class of equivalence relations on
+regular languages using a compact and convenient notation. We then define a
+class of {\em QRE lenses}, a class of quotient lenses which are typed using
+QREs, and show how to {\em synthesize} QRE lenses from their QRE types and a
+set of input-output examples. Finally, we demonstrate the applicability of this
+approach by synthesizing QRE lenses between several real-world data formats
+from the {\tt data.gov} database.
 \end{abstract}
@@
 \section{Introduction}
-
-{\em Lens programming languages} such as
-Boomerang~\cite{Boomerang}\bcp{and others!} enable programmers to kill two
-birds in one stone by expressing both directions of a bidirectional
-transformation with a single program.
-
-\bcp{
-  \begin{itemize}
-  \item Small example of a boomerang program (maybe in our syntax)
-  \item Discussion of lens laws.
-  \end{itemize}
-}
-
-However, the constraints that are imposed by the lens laws are sometimes too
-strict. 
-For instance, the programmer may wish to ignore updates to the data that only
-change the order of fields, or add whitespace characters that in reality are not
-part of the data. Foster et al therefore proposed \textit{quotient lenses} as a
-solution to this problem. Quotient lenses enable a programmer to write lenses
-which transform data that is quotiented out by an equivalence relation, so that
-records which are equivalent to each other modulo this relation are considered
-the same by the underlying lens.
-
-\dpw{The next paragraph gets at the problem, but takes a while to do
-so.  Why not make the first sentence go directly to the problem:  ``Unfortunately,
-quotient lens programs are unnecessarily verbose and challenging to construct.'' ... or
-something like that.}
-
-This paper addresses two main challenges that arise when programming quotient
-lenses. The first challenge is tied directly to the programming of quotient
-lenses in Boomerang, a string lens programming language where the source and
-view types are given by regular expressions. While Boomerang enables the
-programmer to express a rich set of quotient lenses, the syntax for writing
-these lenses is often too verbose and cumbersome. This makes writing quotient
-lenses in Boomerang rather difficult.
-
-For example, suppose that the programmer wishes to convert citation records
-stored in \textsc{Bib}\TeX{} format to a representation of the same data in the
-EndNote format. The programmer wishes to ``quotient'' away unnecessary
-whitespace characters in between the various fields in each record, while also
-ignoring the order in which these fields occur in the recore. For instance, the
-quotient lens which the programmer wishes to derive should convert this
-\textsc{Bib}\TeX{} record
-
+{\em Bidirectional programming languages} such as
+Boomerang~\cite{boomerang}, Augeas~\cite{augeas}, XSugar~\cite{xsugar},
+biXid~\cite{bixid} and X/Inv~\cite{xinv}enable a programmer to express a
+transformation that maps some input data to some output data, as well as a
+transformation that maps the output data back into the input data in a single
+expression. An example of a {\em bidirectional transformation} which we shall
+revisit severally in this paper is a program which converts citation records
+such as
 \begin{verbatim}
@@
   Title = {Regular Algebra and Finite Machines},
-  Publisher = {Printed in GB by William Clowes & Sons Ltd},
-  Year = {1971}
 }
@@
 
-\noindent to this EndNote record
-
+\noindent stored in the \bibtex format to their equivalent representation 
 \begin{verbatim}
@@
 %A Conway, J. H.
-%D 1971
-%I Printed in GB by William Clowes & Sons Ltd
 %F conway
@@
 
-\noindent and vice versa.  \bcp{Let's make the example simpler (leave out
-  year and publisher?)}
-
-\bcp{Moreover...}
-allowing for this equivalent representation of the \textsc{Bib}\TeX{}
-entry, which rearranges the fields and uses different whitespace conventions:
+in the EndNote format, and vice versa.
+
+Bidirectional programs have also been called {\em lenses}. Formally, Foster et
+al~\cite{foster} defined a lens $\ell$ that maps between an ``abstract" {\em
+source} $S$ and a ``concrete" {\em view} $V$ to be a triple of functions
+\begin{align*}
+\ell.get &: S \longrightarrow V\\
+\ell.put &: V \longrightarrow S \longrightarrow S\\
+\ell.create &: S \longrightarrow V
+\end{align*}
+
+\noindent satisfying the equations
+\begin{align*}
+\ell.\lput \; (\ell.\get \; s) \; s &= s\\
+\ell.\get \; (\ell.\lput \; v \; s) &= v\\
+\ell.\lput \; (\ell.\create \; v) &= v\\
+\end{align*}
+
+The $\get$ function presents a view of the data from the source, the
+$\lput$ function enables an updated view to be folded back into the source, and
+the $\create$ component produces a copy of the source from a view, using
+defaults to fill in information that may have been discarded by the $\get$
+function in producing the view. The first law states that the $\lput$ function
+must restore all the information discarded by $\get$ when its arguments are an
+abstract structure and a concrete structure that generates the very same
+abstract structure. The second and third laws state that $\lput$ and $\create$
+must propagate all of the information contained in their abstract arguments to
+the concrete structure they produce. These laws express fundamental expectations
+about how the components of a lens should work together.
+
+However, the constraints that are imposed by the lens laws are sometimes too
+strict. For instance, the programmer may wish to ignore updates to the data
+that only change the order of fields, or add whitespace characters that in reality are not
+part of the data. Foster et al therefore proposed \textit{quotient lenses}
+~\cite{quotient} as a solution to this problem. Quotient lenses enable a
+programmer to write lenses which transform data that is quotiented out by an
+equivalence relation, so that records which are equivalent to each other modulo
+this relation are considered the same by the underlying lens.
+
+Unfortunately, quotient lenses are sometimes difficult to program because the
+equivalence relations defined on the data, as well as the underlying lens which
+operates on the equivalence classes are complicated. For example, suppose that
+the programmer wants to program the \bibtex to EndNote transformation in such a
+way that the program allows for any of two representations of the authors
+name (e.g. ``Conway, J. H." and ``Conway J. H"), ignores the order in which the
+author and title fields occur in each record, and allows for arbitrary amounts
+of whitespace in between the author and title fields. These rules allow for
+these two \bibtex citation records to be considered the same:
 
 \begin{verbatim}
+@Book {conway,
+  Author = {Conway, J. H.},
+  Title = {Regular Algebra and Finite Machines},
+}
+
 @Book{conway,
-    Publisher = {Printed in GB by 
-                 William Clowes & Sons Ltd},
-
     Title = {Regular Algebra and Finite Machines},
 
-    Year = {1971},
-
-    Author = {Conway, J. H.},
+    Author = {Conway J. H.},
 }
 \end{verbatim}
-\bcp{Can we write (regular and QRE) lenses that handle this version of the
-  example?  Write this in Boomerang and see how it looks!!}
-
-While Boomerang enables the programmer to write a lens which exhibits this
-behaviour, the syntax tree for the program is large and complicated.
-
-\saz{It might be a good idea to show what this would look like in all its gory
-detail, just to make the problem very concrete. Perhaps in a figure?}\bcp{+1}\dpw{+1}
-
-In this paper, we introduce the formalism of Quotient Regular Expressions (or
-QREs) to address this issue. \dpw{cut unnecessary phrases that add little information
-content.  eg, cut ``the formalism of.''  Doing so makes your text ``punchier.''}  QREs enable the programmer to express
-a rich set of equivalence relations on regular expressions at the level of
-syntax. QREs therefore enable the programmer to express more equivalence
-relations with less effort. Also, since QREs are defined at the level of
-syntax on regular expressions, QREs give the programmer more control over where
-the quotienting occurs.
-
-\bcp{Give a little example (perhaps for Bibtex and/or EndNote?)}
-
-The second challenge that we address in this paper is the {\em synthesis} of
-(bijective) quotient lenses. These are quotient lenses which match the
-equivalance classes of the source data to the equivalence classes of the view
-data bijectively as in the \textsc{Bib}\TeX{} to EndNote example above. 
-Building on our prior work~\cite{popl18} where we showed how to synthesize
-(ordinary?) bijective lenses, we show how to synthesize bijective quotient
-lenses given a synthesis task described by a pair of QREs and a set of
-examples.
-
-Our main contributions are:
+
+The Boomerang quotient lens which does the \bibtex to EndNote conversion is
+
+\sam{TODO: PUT BOOMERANG EXPRESSION HERE!},
+
+an expression which is difficult to write and understand.
+
+We address the challenge of programming quotient lenses in two ways. First, we
+propose the language of {\em Quotient Regular Expressions} or QREs, which
+enables the programmer to explicitly define a broad class of equivalence
+relations on regular languages using a compact and convenient notation. For
+example, the QRE which describes the equivalence relation on the \bibtex
+citation records is as follows:
+
+\sam{TODO: PUT QRE FOR BIBTEX HERE!}
+
+This expression better describes the equivalence relation since it is explicit
+and more precise.
+
+However, even though QREs enable the programmer to describe complicated
+equivalence relations on data more simply, the underlying lens which maps the
+equivalence classes of the \bibtex records to the equivalence classes of the
+EndNote records is also challenging to program, mostly because of the verbosity
+of the expression:
+
+\sam{TODO: PUT QRE LENS FOR BIBTEX HERE!}
+
+Therefore, to address this challenge we demonstrate how to {\em synthesize}
+quotient lenses from a pair of QREs and a set of input-output example pairs. Our
+key insight is to reduce the problem of synthesizing QRE lenses to the problem
+of synthesizing {\em bijective lenses}, a problem which we solved in out prior
+work ~\cite{popl18}.
+
+Our approach is to define a class of {\em QRE lenses} which are the same as
+bijective lenses with canonizers at the ends. That is, a QRE lens uses the
+source QRE to compute a unique representative for a datum modulo the
+equivalence relation defined by the QRE before applying a bijective lens to
+this representative in the forward direction, and operates similarly same in the
+backward direction but using the view QRE.
+
+\sam{TODO: PUT LENS WITH CANONIZERS ON THE EDGE DIAGRAM HERE!}
+
+One potential pitfall with QRE lenses is that the composition of QRE lenses may
+not have the normal form of a bijective lens with canonizers on the ends:
+
+\sam{TODO: PUT COMPOSITION DIAGRAM HERE!}
+
+Our main technical contribution in this paper is a proof that QRE lenses are
+closed under left quotienting, right quotienting, composition, and regular
+combinators. This proof will allow justify our approach for synthesizing
+quotient lenses so that the programmer can simply type
+
+$$\mathit{synth} \; \mathtt{BIBTEX} \; \mathtt{ENDNOTE}$$
+
+to derive a quotient lens that does the \bibtex to EndNote.
+
+In summary, our main contributions are:
 \begin{enumerate}
@@
 
+For example, suppose that the programmer wishes to write a regular expression
+which describes \bibtex citation records such as:
+\begin{verbatim}
+@Book {conway,
+  Author = {Conway, J. H.},
+  Title = {Regular Algebra and Finite Machines},
+}
+\end{verbatim}
+
+Each record may be represented in many different ways. Firstly, the author's
+name can have the last name followed by a comma, and then the
+first and second names separated by a space, or the last name, first name and
+second name separated by spaces in between them. For example, each of these
+\bibtex records is equivalent to the other:
+\begin{verbatim}
+@Book {conway,
+  Author = {Conway, J. H.},
+  Title = {Regular Algebra and Finite Machines},
+}
+\end{verbatim}
+\begin{verbatim}
+@Book {conway,
+  Author = {Conway J. H.},
+  Title = {Regular Algebra and Finite Machines},
+}
+\end{verbatim}
+
+Secondly, the author and title fields may occur in either order:
+\begin{verbatim}
+@Book {conway,
+  Author = {Conway, J. H.},
+  Title = {Regular Algebra and Finite Machines},
+}
+\end{verbatim}
+\begin{verbatim}
+@Book {conway,
+  Title = {Regular Algebra and Finite Machines},
+  Author = {Conway, J. H.},
+}
+\end{verbatim}
+Thirdly, the title and author records can be separated by arbitrary amounts of
+whitespace characters. For example, if the title field occurs first, then the
+author field may occur not on the line immediately below the title field, but
+instead on the second line below the title field:
+\begin{verbatim}
+@Book {conway,
+  Title = {Regular Algebra and Finite Machines},
+
+  Author = {Conway, J. H.},
+}
+\end{verbatim}
+In the example above, a \bibtex record may have different
+representations for three different reasons. The first reason
+is that the author's name can have the last name followed by a comma and then
+the first and second names separated by a space, or the last name, first name and
+second name separated by spaces in between them. The $\squash{R}{R'}{f}$ QRE
+enables the programmer to choose between the representations described by $R$
+and $R$', provided that they can provide a function $f$ that maps data stored in
+the first format to data stored in the second format. In this example we have
+the two regular expressions
+
+\begin{align*}
+\mathtt{COMMA\_NAME} &= \mathtt{NAME} \cdot ``," \cdot \mathtt{NAME} \cdot ``
+\quad " \cdot \mathtt{NAME}\\
+\mathtt{SPACE\_NAME} &= \mathtt{NAME} \cdot ``," \cdot \mathtt{NAME} \cdot ``
+\quad " \cdot \mathtt{NAME}\\
+\end{align*}
+
+and the bijective lens 
+$$ \ell = (\mathit{copy} \; \mathtt{COMMA\_NAME}) \cdot (\mathit{ins} \; ``,")
+\cdot (\mathit{del} \; `` \quad") \cdot (\mathit{copy} \; \mathtt{COMMA\_NAME})
+\cdot (\mathit{copy} \; \mathtt{COMMA\_NAME})$$
+therefore the QRE
+$$\mathtt{AUTHOR}
+=
+``\text{Author
+= \{}" \cdot \squash{\mathtt{COMMA\_NAME}}{\mathtt{SPACE\_NAME}}{\ell.get}
+\cdot ``\}"$$ enables the programmer to express the fact that the two different
+representations of the author's name are equivalent, with the
+$\mathtt{COMMA\_NAME}$ representation chosen as the canonical representation,
+and the function $\ell.get$ giving the ``canonizing'' function which converts
+data in the $\mathtt{SPACE\_NAME}$ format to its $\mathtt{COMMA\_NAME}$
+representation. The $\squash{R}{R'}{f}$ has the added advantage that if the
+transformation from $R$ to $R'$ can be expressed by a bijective lens,
+then the programmer may attempt to \textit{synthesize} the lens from $R$ to $R'$
+because of our previous work where we demonstrated how to synthesize bijective
+lenses from a pair of regular expressions and a set of input-output example
+pairs:
+$$\mathtt{AUTHOR}
+=
+``\text{Author
+= \{}" \cdot \squash{\mathtt{COMMA\_NAME}}{\mathtt{SPACE\_NAME}}{\mathit{synth}
+\; \mathtt{COMMA\_NAME} \; \mathtt{SPACE\_NAME}} \cdot ``\}"$$
+
+\sam{Should we have a special construct like squash-synth(R, R') in addition to
+squash(R, R', l) or maybe even just have only squash-synth(R, R')? Perhaps
+having to explicitly write the lens weakens our case that QREs are compact and
+convenient}
+
+The second reason that a record may have two different representations is that
+the author and title fields may occur in a different order, and the third
+reason is that the author and title fields may have arbitrary occurences of
+whitespace characters in between them. The $R \mapsto s$ and $\perm{q_1,
+\ldots, q_n}{q}$ QREs enable the programmer to handle each of these cases:
+$$\perm{\re{AUTHOR}, \re{TITLE}}{((``," \cdot \re{WHITESPACE}^*) \mapsto \string
+``, \backslash n")}$$
+This QRE enables the programmer to consider any permutation of the author and
+title fields interspersed with a comma and arbitrary whitespace as equivalent.
+The $((``," \cdot \re{WHITESPACE}^*) \mapsto \string ``, \backslash n")$
+subexpression is itself a QRE that denotes that all string $s$ which consists
+of a comma $``,"$ followed by any number of whitespace should be considered
+equivalent, with the canonical representative being the string $\string ``,
+\backslash n"$. The canonical representation of \bibtex records is that in
+which the author field occurs the title field with a comma then a newline
+character separating the two.
+
+The regular combinators for QREs i.e. $\cdot$ (concatenation), $|$ (union) and
+$*$ (iteration) behave like the regular combinators for regular languages. For
+instance, if the QRE for matching a \bibtex \; record is named
+$\mathtt{RECORD}$, then the QRE for matching two records is $\mathtt{RECORD}
+\cdot \mathtt{RECORD}$, the QRE for matching one or two records is $\mathtt{RECORD}
+\sep (\mathtt{RECORD} \cdot \mathtt{RECORD})$ and the QRE for matching any
+number of QRE records in $\mathtt{RECORD}^*$.
+
+The $\normalize{R}{R'}{f}$ QRE is the most general QRE in that for each of the
+the other QREs $q$, there exist regular expressions $R, R'$ and a surjective,
+idempotent function $f:\mathcal{L}(R) \longrightarrow \mathcal{L}(R')$ such that
+$q$ $\normalize{R}{R'}{f}$ each denote the same equivalence relation on the
+same regular language. $\normalize{R}{R'}{f}$ enables the programmer to
+provide a canonizing function from $\mathcal{L}(R)$ to $\mathcal{L}(R')$,
+with the equivalence class on $\mathcal{L}(R)$ being the equivalence relation
+defined by the {\em fibres} of $f$ (i.e. $r \sim r'$ if and only if $f(r) =
+f(r'))$. The canonizing function is required to be surjective and idempotent.
+The $\normalize{R}{R'}{f}$ combinator therefore enables the programmer to express
+equivalence relations which are difficult to express, or which cannot be
+expressed using the other combinators.
+
+In summary each QRE $q$ enables us to express
+\begin{enumerate}
+  \item a regular expression $W(q)$ (the ``whole'' of $q$),
+  \item an equivalence relation $\eqrel{q}$ on $\mathcal{L}(W(q))$,
+  \item a regular expression $K(q)$ (the ``kernel'' of $q$)
+  such that $\mathcal{L}(K(q))$ forms a complete set of representatives for
+  $\eqrel{q}$, and
+  \item a ``canonizing'' function $\canonizer(q):\mathcal{L}(W(q))
+  \longrightarrow \mathcal{L}(K(q))$ which given any $w \in \mathcal{L}(W(q))$,
+  computes $\canonizer(q)(w)$ as the unique $k$ in $\mathcal{L}(K(q))$ such that
+  $k$ is equivalent to $w$ mod $\eqrel{q}$
+  \end{enumerate}
+  Observe that since $\canonizer(q)$ computes $\canonizer(q)(w)$ as the unique
+  $k$ in $\mathcal{L}(K(q))$ such that $w$ is equivalent to $k$ mod $\eqrel{q}$,
+  then the equivalence classes of $\eqrel{q}$ are the same as the fibres of
+  $\canonizer(q)$.
 \subsection{Syntax of QREs}
@@
 
-Each QRE $q$ enables us to express
-\begin{enumerate}
-  \item a regular expression $W(q)$ (the ``whole'' of $q$),
-  \item an equivalence relation $EqRel(q)$ on $\mathcal{L}(W(q))$,
-  \item a regular expression $K(q)$ (the ``kernel'' of $q$)
-  such that $\mathcal{L}(K(q))$ forms a complete set of representatives for
-  $EqRel(q)$, and
-  \item a ``canonizing'' function $\canonizer(q):\mathcal{L}(W(q))
-  \longrightarrow \mathcal{L}(K(q))$ which given any $w \in \mathcal{L}(W(q))$,
-  computes $\canonizer(q)(w)$ as the unique $k$ in $\mathcal{L}(K(q))$ such that
-  $k$ is equivalent to $w$ mod $EqRel(q)$
-  \end{enumerate}
-  Observe that since $\canonizer(q)$ computes $\canonizer(q)(w)$ as the unique
-  $k$ in $\mathcal{L}(K(q))$ such that $w$ is equivalent to $k$ mod $EqRel(q)$,
-  then the equivalence classes of $EqRel(q)$ are the same as the fibres of
-  $\canonizer(q)$.
-
-\dpw{I think it would be a little more effective to introduce QREs with a running,
-but informal, concrete example. Explain these concepts informally as you go.
-Then summarize the syntax at the end.}
-
-  \subsection{Examples}
-
-\bcp{Give some counterexamples too!  And can we take away the enumerate?}
-
-  \begin{enumerate}
-    \item 
-    Every regular expression can be lifted to a QRE. When a regular expression
-    $R$ is used as a QRE, then $\mathcal{L}(R)$ is assumed to be quotiented by
-    the equality equivalence relation on $\mathcal{L}(R)$.
-    \item
-    For the QRE $R \mapsto s$, the data is described by $\mathcal{L}(R)$, but
-    all members of $R$ are considered equivalent to each other, with $s \in
-    \mathcal{L}(R)$ as the choice of representative. This construct is useful
-    when the programmer simply wishes to ignore unnecessary characters such as
-    excess whitespace e.g. 
-    $$\mathtt{FIRST\_NAME} \cdot (\mathtt{WHITESPACE}^* \mapsto `` \quad ")
-    \cdot \mathtt{LAST\_NAME}$$
-    \item
-    $\squash{R}{R'}{f}$ enables the programmer to use one of two different
-    representations of the data if they can provide a way of mapping the first
-    representation to the second. For instance, the first representation may be
-    $$\mathtt{FIRST\_NAME} \cdot `` \quad " \cdot \mathtt{LAST\_NAME}$$
-    while the second representation may be
-    $$\mathtt{FIRST\_NAME} \cdot ``," \cdot \mathtt{LAST\_NAME}$$
-    The mapping that maps the first representation to the second simply converts
-    the space between the first and last name to a comma.
-    \item
-    The $\perm{q_1, \ldots, q_n}{q}$ construct enables the programmer to
-    consider all permutations of some data that are interspersed with a
-    separator the same. For instance, in the \textsc{Bib}\TeX{} to EndNote
-    example, the programmer may use this construct as follows:
-    $$\perm{\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},
-    \re{YEAR}}{(\re{WHITESPACE}^* \mapsto \string \n)}$$
- where $\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},$ and $\re{YEAR}$ are
-regular expressions describing the respective \textsc{Bib}\TeX\; fields of the
-same name, and $\re{WHITESPACE}$ is a regular expression describing whitespace
-characters.
-\item
-The constructs $q \cdot q'$, $q \sep q'$ and $q^*$ are similar to the respective
-regular expression constructs. However, the regular constructs for QREs are more
-restrictive for QREs since the transformations made to the data by
-canonizers and lenses involve matching the data against a regular expression
-deterministic. Therefore to make these transformations deterministic, the
-data should be matched uniquely against the regular expressions, and this
-constrains the regular expressions that may be used by QREs. \dpw{This is not an
-``example'' -- it is an informal description of how these operators work.}
-\item
-The $\normalize{R}{R'}{f}$ construct allows the programmer to choose how to
-canonize $\mathcal{L}(R)$ into $\mathcal{L}(R')$ be providing the canonizing
-function $f$ themselves. This constuct is useful when the desired equivalence
-relation cannot easily be expressed by any of the other constructs.
-  \end{enumerate}
-
-\saz{Every regular expression $R$ is a QRE.  The main new syntactic
-forms are:
-$R \mapsto s$, which }
-
-\saz{Would it be simpler to include atoms and $\epsilon$ in $q$ rather than $R$?  That is, $R$ is
-  just a subset of $QRE$ built from atoms, $\cdot$ $|$ and ${}^*$.  We could
-  reserve $R$ as a metavariable that ranges over $q$'s such that $q$ is a
-  regular expression.}
-
-\saz{Need to introduce $\mathcal{L}(R)$ notation.}
-
-\iffalse
-Quotient Regular Expressions (or QREs) enable us to give richer specifications
-for data formats than ordinary regular expressions. More concretely,
-  All data to be transformed must match $W(q)$, the ``whole'' regular
-  expression of $q$. However, before transforming the using a lens, the data
-  will be quotiented out by the equivalence relation $EqRel(q)$ which has $K(q)$
-  as a complete set of representatives. That is, for each string $w \in
-  W(q)$, there is a unique $k \in K(q)$ such that $w \; EqRel(q) \; k$). This
-  unique representative $k$ is chosen using the ``canonizing'' function
-  $\canonizer(q)$.
-  \fi
-
 \subsection{Semantics of QREs}
 \subsubsection{Preliminaries}
-As we mentioned earlier, when applying the regular constructs $\cdot
-\; | \;  *$ to QREs, we require that the regular languages
-described by QREs match the data in only one way so as to make
-various transformations made to the data that involve matching the data against
-a regular expression deterministic. Consequently, we require that regular
-expressions used by QREs be \textit{strongly unambiguous}, a condition which
-ensures that if a string $s$ matches a regular expression $R$, then $s$ matches
-$R$ uniquely.
+Let $q, q'$ be QREs. When applying the regular combinators to QREs, we require
+that if a string $s$ matches any of the regular expressions $W(q) \cdot W(q')$,
+$W(q) \sep W(q')$, $W(q)^*$, $K(q) \cdot K(q')$,
+$K(q) \sep K(q')$, $K(q)^*$, then $s$ matches that regular expression in
+only one way. This unambiguity condition is called {\em strong unambiguity},
+and is necessary, firstly because it ensures that the canonizing function of a
+QRE is well-defined. For example if $\mathtt{RECORD}$ is the QRE for a \bibtex
+record in our \bibtex to and $W(\mathtt{ENDNOTE} \cdot \mathtt{ENDNOTE})$ is not
+unambiguous, that is there is a string $s$ such that $s = s_1 \cdot s_2 =
+{s_1}' \cdot {s_2}'$ with $s_1, s_2, {s_1}', {s_2}' \in W(\mathtt{ENDNOTE} \cdot
+\mathtt{ENDNOTE})$, then the canonizer $f$ for $\mathtt{ENDNOTE} \cdot
+\mathtt{ENDNOTE}$ may not be well defined since we are not guaranteed that
+$f(s_1) \cdot f(s_2) = f({s_1}') \cdot f({s_2}')$. We also require that the
+regular combinators applied to the kernels are unambiguous since the underlying
+lenses will end up operating on the kernels, and these lenses impose the same
+restrictions for similar reasons.
 
@@
 The semantics for QREs that formally define the whole language $W(q)$, the
-kernel language $K(q)$, the equivalence relation $EqRel(q)$ and the
-canonizing function $\canonizer(q)$ are given in Figures ~\ref{fig:wk},
-~\ref{fig:relations} and ~\ref{fig:canonizers}.
+kernel language $K(q)$, the equivalence relation $\eqrel{q}$ and the
+canonizing function $\canonizer(q)$ are given in
+Figures~\ref{fig:wk},~\ref{fig:relations} and~\ref{fig:canonizers}.
 \begin{figure}[t]
@@
   \item $W(q)$ and $K(q)$ are well defined regular expressions,
-  \item  $EqRel(q)$ is an equivalence relation on $\mathcal{L}(W(q))$,
+  \item  $\eqrel{q}$ is an equivalence relation on $\mathcal{L}(W(q))$,
   \item  $\mathcal{L}(K(q))$ forms a complete set of representatives for
-  $EqRel(q)$,
+  $\eqrel{q}$,
   \item $\canonizer(q):\mathcal{L}(W(q)) \longrightarrow \mathcal{L}(K(q))$ is a
@@
   $k$ in $\mathcal{L}(K(q))$ such that $k$ is equivalent to $w$ modulo
-  $EqRel(q)$.
+  $\eqrel{q}$.
   \end{enumerate}
@@
 
-\subsection{Syntax of QRE Lenses}
-\subsubsection{Preliminaries}
+Our approach in defining QRE lenses is as follows. Let $c, c'$ be QREs. Given a
+bijection $\ell : K(c) \Leftrightarrow K(c')$, we may define a quotient lens $q
+: W(c)/\eqrel{c} \Leftrightarrow W(c')/\eqrel{c'}$ by
+\begin{equation}\label{normalform}
+q.get = \ell \circ \canonizer(c) \text{ and } q.put = \ell^{-1} \circ
+\canonizer(c'),
+\end{equation}
+
+A potential problem with this approach is that quotient lenses of this form may
+not necessarily be closed under the action of the lens combinators. For
+example, given two lenses $q$ and $q'$, the composition of $q' \circ q$ may not
+be of this form.
+
+Consequently, in order to define quotient lenses, we need to choose a class of
+bijections that has just the right properties with respect to the quotient
+lens combinators. Our choice is to use the class of \textit{bijective
+lenses} which we identified and studied in our previous work~\cite{popl18}.
+Bijective lenses, which we briefly discuss in the next subsection, possess
+the properties needed to ensure that all QRE lenses have the normal form
+suggested in \cref{normalform}.
+
+\subsection{Bijective Lenses}
 Given regular expressions $R, S$ and equivalence
@@
 
-Our approach in defining QRE lenses is as follows. Let $c, c'$ be QREs. Given a
-bijection $\ell : K(c) \Leftrightarrow K(c')$, we may define a quotient lens $q
-: W(c)/EqRel(c) \Leftrightarrow W(c')/EqRel(c')$ by
-\begin{equation}\label{normalform}
-q.get = \ell \circ \canonizer(c) \text{ and } q.put = \ell^{-1} \circ
-\canonizer(c'),
-\end{equation}
-
-A potential problem with this approach is that quotient lenses of this form may
-not necessarily be closed under the action of the lens combinators. For
-example, these lenses may not be closed under composition, which is one of the
-combinators we would want to have in the quotient lens language.\bcp{Can you
-say the explanation without using the phrase ``closed under''?}
-
-Consequently, in order to define quotient lenses, we need to choose a class of
-bijections that has just the right properties with respect to the quotient
-lens combinators. Our choice is to use the class of \textit{bijective
-lenses} which we identified and studied in our previous work~\cite{popl18}.
-Bijective lenses, which we briefly discuss in the next subsection, possess
-the properties needed to ensure that all QRE lenses have the normal form
-suggested in \cref{normalform}.
-
-\subsection{Bijective Lenses}
 We define the set of \textit{bijective lenses} to be the set of bijections
@@
 \subsection{Syntax of QRE Lenses}
-Having described the class of bijective lenses, we now introduce the class
-of QRE lenses, whose language is given by following grammar:
+Having given a brief overview of the class of bijective lenses, we now introduce
+the class of QRE lenses, whose language is given by following grammar:
 $$ q := \mathit{lift}(\ell) \sep \mathit{lquot}(q, \ell) \sep
@@
 $c$, assuming that the source data forms a complete set of representatives for
-the equivalence relation $EqRel(c)$. The $\mathit{lquot}(q, c)$ does the same
+the equivalence relation $\eqrel{c}$. The $\mathit{lquot}(q, c)$ does the same
 but on the view data. The regular combinators $\cdot \; * \; |$ exhibit the
@@
 $\llbracket q \rrbracket$ of a QRE lens $q:c \Leftrightarrow c'$ is a quotient
-lens $\llbracket q \rrbracket : W(c)/{EqRel(c)} \Longleftrightarrow
-W(c')/{EqRel(c')}$. The typing rules and denotation of QRE lenses are given in
+lens $\llbracket q \rrbracket : W(c)/{\eqrel{c}} \Longleftrightarrow
+W(c')/{\eqrel{c'}}$. The typing rules and denotation of QRE lenses are given in
 Figure~\ref{fig:qlenssemantics}. The following theorem confirms that QRE lenses
@@
 If there is a derivation $q:c \Leftrightarrow c'$ then
-$\llbracket q \rrbracket : W(c)/EqRel(c) \Leftrightarrow W(c')/EqRel(c')$ is a
+$\llbracket q \rrbracket : W(c)/\eqrel{c} \Leftrightarrow W(c')/\eqrel{c'}$ is a
 well-defined quotient lens.

### 2017-10-11 papers/QuotientLensSynth/main.tex
Made more changes

@@
 
-However, quotient lenses are sometimes difficult to program because the
-equivalence relations defined on the data, as well as the underlying lens which
-operates on the equivalence classes are complicated. These two challenges make
-quotient languages impractical to use in many cases spite of their advantages.
+However, quotient lenses are sometimes difficult to program, firstly because the
+programmer has to give a specification for the data as well as give a separate
+specification for the equivalence relations defined on the data. Secondly, even
+after giving both of these specifications, the programmer also has to manually
+write the underlying lens which maps between the respective equivalence
+classes, a task which is itself challenging.
 
@@
 language of {\em Quotient Regular Expressions} or QREs, which enables the
-programmer to explicitly define a broad class of equivalence relations on
-regular languages using a compact and convenient notation. We then define a
-class of {\em QRE lenses}, a class of quotient lenses which are typed using
-QREs, and show how to {\em synthesize} QRE lenses from their QRE types and a
-set of input-output examples. Finally, we demonstrate the applicability of this
-approach by synthesizing QRE lenses between several real-world data formats
-from the {\tt data.gov} database.
+programmer to explicitly define a broad class of equivalence relations
+directly on regular languages using a compact and convenient notation, and hence
+negates the need to specify the data and the equivalence relation on the data
+separately. We then define a class of {\em QRE lenses}, a class of quotient
+lenses which are typed using QREs, and show how to {\em synthesize} QRE lenses
+from their QRE types and a set of input-output examples. Finally, we
+demonstrate the applicability of this approach by synthesizing QRE lenses
+between several real-world data formats from the {\tt data.gov} database.
 \end{abstract}
@@
 
-in the EndNote format, and vice versa.
+\noindent in the EndNote format, and vice versa.
 
@@
 strict. For instance, the programmer may wish to ignore updates to the data
-that only change the order of fields, or add whitespace characters that in reality are not
-part of the data. Foster et al therefore proposed \textit{quotient lenses}
-~\cite{quotient} as a solution to this problem. Quotient lenses enable a
-programmer to write lenses which transform data that is quotiented out by an
-equivalence relation, so that records which are equivalent to each other modulo
-this relation are considered the same by the underlying lens.
-
-Unfortunately, quotient lenses are sometimes difficult to program because the
-equivalence relations defined on the data, as well as the underlying lens which
-operates on the equivalence classes are complicated. For example, suppose that
-the programmer wants to program the \bibtex to EndNote transformation in such a
-way that the program allows for any of two representations of the authors
-name (e.g. ``Conway, J. H." and ``Conway J. H"), ignores the order in which the
-author and title fields occur in each record, and allows for arbitrary amounts
-of whitespace in between the author and title fields. These rules allow for
-these two \bibtex citation records to be considered the same:
+that only change the order of fields, or add whitespace characters that in
+reality are not part of the data. Foster et al therefore proposed
+\textit{quotient lenses} ~\cite{quotient} as a solution to this problem
+Quotient lenses, which are implemented as a refinement of the bidirectional
+string processing language Boomerang, enable a programmer to write lenses which
+transform data that is quotiented out by an equivalence relation, so that
+records which are equivalent to each other modulo this relation are considered
+the same by the underlying lens.
+
+\sam{Go over quotient lens laws?}
+
+Unfortunately, quotient lenses are sometimes difficult to program in Boomerang,
+firstly because the programmer has to give a specification for the data as well
+as give a separate specification for the equivalence relations defined on the
+data. For example, suppose that the programmer wants to program the \bibtex to
+EndNote transformation in such a way that the program allows for any of two
+representations of the authors name (e.g. ``Conway, J. H." and ``Conway J. H"),
+ignores the order in which the author and title fields occur in each record,
+and allows for arbitrary amounts of whitespace in between the author and title
+fields. These rules allow for these two \bibtex citation records to be
+considered the same,
 
@@
 
-The Boomerang quotient lens which does the \bibtex to EndNote conversion is
-
-\sam{TODO: PUT BOOMERANG EXPRESSION HERE!},
-
-an expression which is difficult to write and understand.
-
-We address the challenge of programming quotient lenses in two ways. First, we
-propose the language of {\em Quotient Regular Expressions} or QREs, which
-enables the programmer to explicitly define a broad class of equivalence
-relations on regular languages using a compact and convenient notation. For
-example, the QRE which describes the equivalence relation on the \bibtex
-citation records is as follows:
-
-\sam{TODO: PUT QRE FOR BIBTEX HERE!}
-
-This expression better describes the equivalence relation since it is explicit
-and more precise.
-
-However, even though QREs enable the programmer to describe complicated
-equivalence relations on data more simply, the underlying lens which maps the
-equivalence classes of the \bibtex records to the equivalence classes of the
-EndNote records is also challenging to program, mostly because of the verbosity
-of the expression:
-
-\sam{TODO: PUT QRE LENS FOR BIBTEX HERE!}
+with the EndNote records having a similar equivalence relation defined on them.
+
+In Boomerang, equivalence relations are expressed using {\em canonizers}
+which are functions that map each element of a regular language to a
+canonical representation of that element. Therefore, in order for the programmer
+to give a specification for the \bibtex records, then the programmer
+has to write a regular expression which matches all possible representations of
+\bibtex records, and then program a set of canonizers which does the job of
+quotienting the data. This often makes the code verbose and error-prone.
+
+\sam{TODO: PUT BOOMERANG CANONIZERS HERE!}
+
+To address this issue, we propose the language of {\em Quotient Regular
+Expressions} or QREs, which enables the programmer to explicitly define a broad
+class of equivalence relations directly on regular languages using a compact
+and convenient notation, and hence negates the need to specify the data and the
+equivalence relation on the data separately. For example, a QRE which
+specifies the equivalence relation on the \bibtex citation records is as
+follows:
+
+\begin{verbatim}
+let bib_author  = 
+"author = {" . squash bib_space_names -> bib_comma_names using 
+(get (synth bib_space_names <=> bib_comma_names)) . "}"
+
+let bib_ignore_wsp =  project ("," . [ \n\t\r]* ) -> ",\n"
+
+let bib_perm : canonizer = perm (label, bib_author, bib_title) with
+bib_ignore_wsp
+
+let bibtex  = "@book{" . bib_perm . "}"
+\end{verbatim}
+
+\sam{Should we have a special construct like squash-synth(R, R') in addition to
+squash(R, R', l), or maybe even just have only squash-synth(R, R')?}
+
+The first expression specifies the author field as
+having one of two representations (\verb!bib_space_names! and
+\verb!bib_comma_names!), with the \verb!bib_comma_names! being the
+canonical representation and the expression
+
+\noindent \verb!(get (synth bib_space_names <=> bib_comma_names))! giving the
+function which converts names described by \verb!bib_space_names! to their canonical
+form described  by \verb!bib_comma_names!.
+
+The second expression specifies the regular language which is a comma followed
+by arbitrary occurences of whitespace characters quotiented by the equivalence
+relation which makes all members of this language equivalent, with the canonical
+representation being a comma followed by a newline character.
+
+The third expression specifies that the label, title and author fields
+are can occur in any order, with each of these fields separated by a comma and
+arbitrary amounts of whitespace. The canonical representaion is that in which
+the label, author and title fields occur in that order, separated by a comma and
+a newline.
+
+However, even though QREs enable the programmer to specify the data as well as
+the equivalence relation on the data in the same expression, the programmer
+still has to manually program the underlying lens which maps the equivalence
+classes of the \bibtex records to the equivalence classes of the EndNote
+records
+
+\sam{TODO: PUT BIJECTIVE LENS HERE!},
+
+\noindent task which is itself challenging and erro-prone.
 
@@
 bijective lenses with canonizers at the ends. That is, a QRE lens uses the
-source QRE to compute a unique representative for a datum modulo the
+source QRE to compute a unique representative for the data modulo the
 equivalence relation defined by the QRE before applying a bijective lens to
 this representative in the forward direction, and operates similarly same in the
-backward direction but using the view QRE.
+backward direction, but using the view QRE.
 
@@
 
-$$\mathit{synth} \; \mathtt{BIBTEX} \; \mathtt{ENDNOTE}$$
-
-to derive a quotient lens that does the \bibtex to EndNote.
+$$\mathit{synth} \; \verb!bibtex! \; \verb!endnote!$$
+
+to derive a quotient lens that does the \bibtex to EndNote transformation.
 
@@
 \; \mathtt{COMMA\_NAME} \; \mathtt{SPACE\_NAME}} \cdot ``\}"$$
-
-\sam{Should we have a special construct like squash-synth(R, R') in addition to
-squash(R, R', l) or maybe even just have only squash-synth(R, R')? Perhaps
-having to explicitly write the lens weakens our case that QREs are compact and
-convenient}
 

### 2017-10-12 papers/QuotientLensSynth/main.tex

@@
 \begin{abstract}
-{\em A Quotient lens} is an expression which encapsulates a bidirectional
-transformation whereby each of the two transformations that are part of the lens
-have the same behaviour on data that are equivalent to each other modulo some
-equivalence relation. Quotient lenses save the programmer the effort of writing
-two expressions instead of one, give guarantess on how the forward and backward
-transformations interact with each other, and also allow the programmer to
-ignore differences in the data which are not intended to affect the
-behaviour of the transformations.
-
-However, quotient lenses are sometimes difficult to program, firstly because the
-programmer has to give a specification for the data as well as give a separate
-specification for the equivalence relations defined on the data. Secondly, even
-after giving both of these specifications, the programmer also has to manually
-write the underlying lens which maps between the respective equivalence
-classes, a task which is itself challenging.
-
-In this paper, we address both of these challenges. First, we propose the
-language of {\em Quotient Regular Expressions} or QREs, which enables the
-programmer to explicitly define a broad class of equivalence relations
-directly on regular languages using a compact and convenient notation, and hence
-negates the need to specify the data and the equivalence relation on the data
-separately. We then define a class of {\em QRE lenses}, a class of quotient
-lenses which are typed using QREs, and show how to {\em synthesize} QRE lenses
+
+  {\em Quotient lenses} express bidirectional transformations between data
+  formats, where the inputs to the transformation are considered modulo some
+  equivalence relation.  Such lenses save the programmer from having to write
+  two expressions and satisfy strong ``lens laws'' that ensure that the forward
+  and backward transformations interact well.  The use of quotients allows the
+  programmer to ignore ``inessential'' differences in the data, such as the
+  amount whitespace or the ordering of fields.
+
+  However, quotient lenses are often difficult to program.  The programmer has
+  to give a specification for the data on each side of the lens and a separate
+  specification for the equivalence relations defined on the data. The
+  programmer must also manually implement a lens program that maps between the
+  specifications while respecting the equivalence relations---this is itself a
+  challenging task thanks to the restrictions on how lenses can be combined
+  while still satisfying the lens laws.
+
+This paper addresses both of these challenges. First, we propose the
+language of {\em Quotient Regular Expressions}, or QREs, which 
+define a broad class of equivalence relations
+directly on regular languages using a compact and convenient notation. This
+eliminates the need to specify the data and the equivalence relation on the data
+separately. We then define {\em QRE lenses}, a class of quotient
+lenses that are typed using QREs, and show how to {\em synthesize} QRE lenses
 from their QRE types and a set of input-output examples. Finally, we
@@
 
-\noindent stored in the \bibtex format to their equivalent representation 
+\noindent stored in the \bibtex{} format to their equivalent representation 
 \begin{verbatim}
@@
 as give a separate specification for the equivalence relations defined on the
-data. For example, suppose that the programmer wants to program the \bibtex to
+data. For example, suppose that the programmer wants to program the \bibtex{} to
 EndNote transformation in such a way that the program allows for any of two
@@
 and allows for arbitrary amounts of whitespace in between the author and title
-fields. These rules allow for these two \bibtex citation records to be
+fields. These rules allow for these two \bibtex{} citation records to be
 considered the same,
@@
 canonical representation of that element. Therefore, in order for the programmer
-to give a specification for the \bibtex records, then the programmer
+to give a specification for the \bibtex{} records, then the programmer
 has to write a regular expression which matches all possible representations of
-\bibtex records, and then program a set of canonizers which does the job of
+\bibtex{} records, and then program a set of canonizers which does the job of
 quotienting the data. This often makes the code verbose and error-prone.
@@
 equivalence relation on the data separately. For example, a QRE which
-specifies the equivalence relation on the \bibtex citation records is as
+specifies the equivalence relation on the \bibtex{} citation records is as
 follows:
@@
 still has to manually program the underlying lens which maps the equivalence
-classes of the \bibtex records to the equivalence classes of the EndNote
+classes of the \bibtex{} records to the equivalence classes of the EndNote
 records
@@
 
-to derive a quotient lens that does the \bibtex to EndNote transformation.
+to derive a quotient lens that does the \bibtex{} to EndNote transformation.
 
@@
 For example, suppose that the programmer wishes to write a regular expression
-which describes \bibtex citation records such as:
+which describes \bibtex{} citation records such as:
 \begin{verbatim}
@@
 second name separated by spaces in between them. For example, each of these
-\bibtex records is equivalent to the other:
+\bibtex{} records is equivalent to the other:
 \begin{verbatim}
@@
 \end{verbatim}
-In the example above, a \bibtex record may have different
+In the example above, a \bibtex{} record may have different
 representations for three different reasons. The first reason
@@
 equivalent, with the canonical representative being the string $\string ``,
-\backslash n"$. The canonical representation of \bibtex records is that in
+\backslash n"$. The canonical representation of \bibtex{} records is that in
 which the author field occurs the title field with a comma then a newline
@@
 $*$ (iteration) behave like the regular combinators for regular languages. For
-instance, if the QRE for matching a \bibtex \; record is named
+instance, if the QRE for matching a \bibtex{} \; record is named
 $\mathtt{RECORD}$, then the QRE for matching two records is $\mathtt{RECORD}
@@
 QRE is well-defined. For example if $\mathtt{RECORD}$ is the QRE for a \bibtex
-record in our \bibtex to and $W(\mathtt{ENDNOTE} \cdot \mathtt{ENDNOTE})$ is not
+record in our \bibtex{} to and $W(\mathtt{ENDNOTE} \cdot \mathtt{ENDNOTE})$ is not
 unambiguous, that is there is a string $s$ such that $s = s_1 \cdot s_2 =

### 2017-10-13 papers/QuotientLensSynth/main.tex
Comments

@@
 that only change the order of fields, or add whitespace characters that in
-reality are not part of the data. Foster et al therefore proposed
+reality are not part of the data. Foster {\em et al.\relax} therefore proposed
 \textit{quotient lenses} ~\cite{quotient} as a solution to this problem
@@
 EndNote transformation in such a way that the program allows for any of two
-representations of the authors name (e.g. ``Conway, J. H." and ``Conway J. H"),
+representations of the authors name (e.g. ``Conway, J. H.'' and ``Conway
+J. H''\bcp{``J. H. Conway''??}), 
 ignores the order in which the author and title fields occur in each record,
@@
 fields. These rules allow for these two \bibtex{} citation records to be
-considered the same,
+considered the same,\bcp{I thought
+  we were going to allow newlines within titles and authors...?}
 
@@
 to derive a quotient lens that does the \bibtex{} to EndNote transformation.
+\bcp{What about examples??}
 
@@
 for some bijective lens $\ell$. We now provide a proof that all QRE lenses are
-of this form.
+of this form.\bcp{Maybe some cases of the proof can be elided in the short
+  version?} 
 \begin{theorem}\label{normal form}
@@
   extend the ideas there.
+  \item Search for theoretical papers involving equivalence relations over
+  regular languages. 
   \end{itemize}

### 2017-10-15 papers/QuotientLensSynth/main.tex
add eval rough draft

--- 2017-10-12 papers/QuotientLensSynth/main.tex
@@
 \newcommand{\eqrel}[1]{\ensuremath{\equiv_{#1}}}
+
+% Evaluation AST Metrics
+\newcommand{\QRESize}{\textbf{QS}}
+\newcommand{\CanonizerAndSpecSize}{\textbf{TCS}}
+\newcommand{\LensAndSpecSize}{\textbf{TLS}}
+
+% Evaluation Runtime Metrics
+\newcommand{\SystemOnBenchmarks}{\textbf{MM}}
+\newcommand{\SystemOnOptician}{\textbf{MO}}
+\newcommand{\OpticianRuntime}{\textbf{OO}}
+
+\newcommand{\Name}{Optometrist}
 
@@
 \begin{verbatim}
-%0 Book
-%T Regular Algebra and Finite Machines
-%A Conway, J. H.
-%D 1971
-%I Printed in GB by William Clowes & Sons Ltd
-%F conway
+% 0 Book
+% T Regular Algebra and Finite Machines
+% A Conway, J. H.
+% D 1971
+% I Printed in GB by William Clowes & Sons Ltd
+% F conway
 \end{verbatim}
@@
 \label{impl}
+
+We have integrated QREs and quotient lens synthesis into Boomerang with an
+additional \afm{loc} lines of OCaml code.
+
+All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
+of 1600 MHz DDR3 running macOS Sierra.
+
+Our evaluations serve to answer the following questions:
+\begin{enumerate}
+\item Does synthesizing quotient lenses from QREs permit for an easier
+  development process than existing approaches.
+
+\item Are QREs sufficiently expressive that more lenses can be synthesized
+  without modifications to the format.
+
+\item Is our synthesis algorithm competitive with existing lens synthesis
+  tools.
+\end{enumerate}
+
+\subsection{Benchmark Suite Construction} 
+Our benchmarks were constructed from two sources:
+\begin{enumerate}
+\item \afm{39} of our benchmarks were derived from the Optician system.  These
+  benchmarks were a combination of custom benchmarks, benchmarks derived from
+  FlashFill, and benchmarks derived from Augeas.  Many of the formats mapped
+  between had to be modified to work with the bijectivity constraints
+  Optician required.  Commonly, one format had whitespace where the other did
+  not allow for whitespace.  To address the bijectivity issues surrounding this,
+  we added whitespace to the format that typically did not have whitespace.  We
+  were able to remove these alterations in adopting these benchmarks to our
+  system.
+
+\item \afm{10} of our benchmarks were derived from data.gov.  Data.gov contains
+  many transformations between formats.  In addition to the formats previously
+  adopted for Optician, we synthesized transformers between these formats.
+\end{enumerate}
+
+\subsection{Easing the Development Process}
+
+To test if our tool eases the development process, we compare the following
+number of AST nodes.
+%
+\begin{itemize}
+\item[\QRESize{}] \QRESize{} is the number of AST nodes in the QRE
+  specifications involved in QRE lens synthesis task.
+\item[\CanonizerAndSpecSize{}] \CanonizerAndSpecSize{} is the number of
+  AST nodes in $W(q)$ summed with the number of AST nodes in $\canonizer(q)$,
+  for each QRE used in QRE lens synthesis task.
+\item[\LensAndSpecSize{}] \LensAndSpecSize{} is the number of AST nodes
+  in $W(q)$ summed with the number of AST nodes in the synthesized QRE lens.
+\end{itemize}
+
+The value of \QRESize{} represents the difficulty in writing our benchmarks.
+
+The value of \CanonizerAndSpecSize{} represents the difficulty in writing our
+benchmarks without QREs.  This is only an approximation, as both $W(q)$ and
+$\canonizer(q)$ are automatically generated from the QRE.  In our
+experience, these automatically generated regular expressions and canonizers are
+the ones we would naturally write.
+
+The value of \LensAndSpecSize{} represents the difficulty in writing our benchmarks
+with neither QREs nor synthesis.  This is also an approximation, as $W(q)$ and
+$\canonizer(q)$ and the bijective lens are all automatically generated.
+We have found that the generated bijective lens is fairly similar in size to the
+one we would manually write, though it is typically less modular.
+
+Figure~\ref{?} shows the number of benchmarks written with at most a given
+number of AST nodes.  We found that \QRESize{} was consistently smaller than
+\CanonizerAndSpecSize{} and \LensAndSpecSize{}, demonstrating that introducing
+QREs saves programmers programming effort compared to both Optician and basic
+Boomerang.
+
+\subsection{Increase of Expressivity}
+
+Optician was only able to synthesize fully bijective data transformations.
+\Name loosens that restriction, and is able to synthesize data which is
+bijective when put in canonical form.
+To determine the increase of expressivity, we analyzed:
+\begin{enumerate}
+\item How many data formats in Optician's benchmark suite required modifications
+  to be bijective.
+\item How many data formats in Optician's benchmark suite required modifications
+  to be bijective when put in canonical form.
+\end{enumerate}
+
+We found that \afm{35} of Opticians 39 benchmarks required modifications to be
+bijective when put in canonical form. We found that \afm{25} of Optician's 39
+benchmarks required modifications to be in bijective correspondence when in
+canonical form.  Furthermore, we found that \afm{all} of the data.gov examples
+required no modifications to be in canonical form.  While we did not put the
+data.gov examples into a form suitable for Optician, we found only \afm{1} of
+the \afm{5} data.gov examples did not require canonization, so making suitable
+for Optician would have required a modification of \afm{4} of the data.gov
+examples.
+
+\subsection{Competitive Performance}
+
+While we expanded on the Optician system, we found it important to validate
+performance did not degrade in this expansion.  To validate this, we tested:
+\begin{itemize}
+\item[\OpticianRuntime{}] Optician runs on its own benchmarks.
+\item[\SystemOnOptician{}] \Name{} runs on Optician's benchmarks.
+\item[\SystemOnBenchmarks{}] \Name{} runs on its own benchmarks.
+\end{itemize}
+
+We summarize the results of these experiments in Figure~\ref{?}.  \Name{} was
+able to synthesize all of Optician's benchmarks at a competitive speed with
+Optician.  Furthermore, adopting these tasks to use QREs actually sped up
+tasks!  After canonization, the regular expressions sent to synthesis are
+\emph{smaller} than before canonization - making the bijective lens synthesis
+portion of these tasks easier.
 

### 2017-10-15 papers/QuotientLensSynth/main.tex

@@
   extend the ideas there.
+  \item Search for theoretical papers involving equivalence relations over
+  regular languages. 
+    extend the ideas there.
   \end{itemize}

### 2017-10-17 papers/QuotientLensSynth/main.tex
another draft

--- 2017-10-13 papers/QuotientLensSynth/main.tex
@@
 
-\usepackage{amsmath, amssymb, verbatim, enumerate, 
-graphicx, centernot, tikz, array, tikz-cd, extarrows, cleveref,
-mathrsfs, mathtools, bussproofs, stmaryrd, enumitem, stackengine}
+\usepackage{amsmath, amssymb, verbatim, enumerate, graphicx, centernot, tikz,
+array, mathtools, bussproofs, stmaryrd, enumitem, stackengine}
 
@@
 
-  However, quotient lenses are often difficult to program.  The programmer has
-  to give a specification for the data on each side of the lens and a separate
-  specification for the equivalence relations defined on the data. The
+  However, quotient lenses are often difficult to program.  The programmer
+  often has to give a specification for the data on each side of the lens and a
+  separate specification for the equivalence relations defined on the data. The
   programmer must also manually implement a lens program that maps between the
@@
 @Book {conway,
-  Author = {Conway, J. H.},
-  Title = {Regular Algebra and Finite Machines},
+  Author = "J. H. Conway",
+  Title = {Regular Algebra and Finite Machines}
 }
@@
 %T Regular Algebra and Finite Machines
-%A Conway, J. H.
+%A J. H. Conway
 %F conway
@@
 Bidirectional programs have also been called {\em lenses}. Formally, Foster et
-al~\cite{foster} defined a lens $\ell$ that maps between an ``abstract" {\em
-source} $S$ and a ``concrete" {\em view} $V$ to be a triple of functions
-\begin{align*}
-\ell.get &: S \longrightarrow V\\
-\ell.put &: V \longrightarrow S \longrightarrow S\\
-\ell.create &: S \longrightarrow V
-\end{align*}
-
-\noindent satisfying the equations
-\begin{align*}
-\ell.\lput \; (\ell.\get \; s) \; s &= s\\
-\ell.\get \; (\ell.\lput \; v \; s) &= v\\
-\ell.\lput \; (\ell.\create \; v) &= v\\
+al~\cite{foster} defined a lens $\ell : S \Leftrightarrow V$ that maps between
+an ``abstract" {\em source} $S$ and a ``concrete" {\em view} $V$ to be a triple
+of functions $(\ell.\get, \; \ell.\lput, \; \ell.\create)$ satisfying the
+following equations:
+\begin{align*}
+\ell.get &: S \longrightarrow V & \ell.\lput \; (\ell.\get \; s) \; s &= s\\
+\ell.put &: V \longrightarrow S \longrightarrow S & \ell.\get \; (\ell.\lput \;
+v \; s) &= v\\
+\ell.create &: S \longrightarrow V & \ell.\lput \; (\ell.\create \; v) &= v
 \end{align*}
@@
 reality are not part of the data. Foster {\em et al.\relax} therefore proposed
-\textit{quotient lenses} ~\cite{quotient} as a solution to this problem
+\textit{quotient lenses} ~\cite{quotient} as a solution to this problem.
 Quotient lenses, which are implemented as a refinement of the bidirectional
@@
 
-\sam{Go over quotient lens laws?}
-
-Unfortunately, quotient lenses are sometimes difficult to program in Boomerang,
-firstly because the programmer has to give a specification for the data as well
-as give a separate specification for the equivalence relations defined on the
-data. For example, suppose that the programmer wants to program the \bibtex{} to
-EndNote transformation in such a way that the program allows for any of two
-representations of the authors name (e.g. ``Conway, J. H.'' and ``Conway
-J. H''\bcp{``J. H. Conway''??}), 
-ignores the order in which the author and title fields occur in each record,
-and allows for arbitrary amounts of whitespace in between the author and title
-fields. These rules allow for these two \bibtex{} citation records to be
-considered the same,\bcp{I thought
-  we were going to allow newlines within titles and authors...?}
-
+Formally, given equivalence relations $\sim_S$ and $\sim_V$ defined on the
+source $S$ and the view $V$ respectively, a quotient lens $q : S/{\sim_S}
+\Leftrightarrow V/{\sim_V}$ is a triple of functions $(\ell.\get, \; \ell.\lput,
+\; \ell.\create)$ which satisfy the lens laws up to equivalence:
+\begin{align*}
+\ell.get &: S \longrightarrow V & \ell.\lput \; (\ell.\get \; s) \; s &\sim_S
+s\\
+\ell.put &: V \longrightarrow S \longrightarrow S & \ell.\get \; (\ell.\lput \;
+v \; s) &\sim_V v\\
+\ell.create &: S \longrightarrow V & \ell.\lput \; (\ell.\create \; v) &\sim_V v
+\end{align*}
+Moreover, each of the $\get$, $\lput$ and $\create$ functions must respect
+$\sim_S$ and $\sim_V$. That is, if $s \sim_S s'$ then $\get \; s \sim_V
+\get \; s'$, if $s \sim_S s'$ and $v \sim_V v'$ then $\lput \; v \; s \sim_S
+\lput \; v' \; s'$ and if $v \sim_V v'$, then $\create \; v \sim_S \create \;
+v'$.
+
+In Boomerang, equivalence relations are expressed using {\em canonizers}
+which are functions that map members of a regular language to their canonical
+representation modulo an equivalence relation. Therefore, in order to give a
+specification for the data, the programmer often has to write a
+regular expression which matches all possible representations of the data, and
+then provide a set of canonizers which does the job of quotienting the data.
+
+For example, suppose that the programmer wants to
+derive the \bibtex{} to EndNote transformation in such a way that the program
+allows for whitespace to occur arbitrarily between the words in the author and
+title fields, ignores the order in which the author and title fields occur in
+each record, and allows for arbitrary amounts of whitespace in between the
+author and title fields. These rules allow for these two \bibtex{} citation
+records to be considered the same,
+
+\bcp{I thought we were going to allow newlines within titles and authors...?}
+
+\sam{I wanted to illustrate how to use the squash QRE which is why I changed my
+mind on that, but on second thought since the problem of quotienting away
+whitespace at will is going to be more familiar to the reader I changed it back
+to the original plan. I will explain squash in the QRE section.}
 \begin{verbatim}
 @Book {conway,
-  Author = {Conway, J. H.},
+  Author = "J. H. Conway",
   Title = {Regular Algebra and Finite Machines},
@@
 @Book{conway,
-    Title = {Regular Algebra and Finite Machines},
-
-    Author = {Conway J. H.},
+
+    Title = {Regular Algebra 
+            and Finite Machines},
+
+    Author = "J. H. Conway",
 }
@@
 
-with the EndNote records having a similar equivalence relation defined on them.
-
-In Boomerang, equivalence relations are expressed using {\em canonizers}
-which are functions that map each element of a regular language to a
-canonical representation of that element. Therefore, in order for the programmer
-to give a specification for the \bibtex{} records, then the programmer
-has to write a regular expression which matches all possible representations of
-\bibtex{} records, and then program a set of canonizers which does the job of
-quotienting the data. This often makes the code verbose and error-prone.
-
-\sam{TODO: PUT BOOMERANG CANONIZERS HERE!}
+\noindent with the EndNote records having a similar equivalence relation defined
+on them.
+
+The code sample below shows how the programmer might specify these equivalence
+relations on the \bibtex{} records in Boomerang using canonizers:
+
+\begin{verbatim}
+let project (r : regexp) (s : string) = canonizer_of_lens ((del r) . (ins s))
+
+let wsp : regexp = [ \n\t\r]+
+
+let wsp_to_space : canonizer = project wsp " "
+
+let valid_name : canonizer = 
+name | (name . wsp_to_space . (name .wsp_to_space)* .name)
+
+let wsp_to_and : canonizer = project (wsp . "and" . wsp) " and "
+
+let bib_names : canonizer =
+
+valid_name | (valid_name . wsp_to_and . (valid_name . wsp_to_and)* . valid_name)
+
+let bib_author : canonizer = "author = \"" . bib_names . "\""
+
+let label : regexp = [a-zA-Z0-9]+
+
+let title : canonizer = 
+word | (word . wsp_to_space . (word . wsp_to_space)* . word)
+
+let bib_title : canonizer = "title = {" . title . "}"
+
+let wsp_to_comma : canonizer = project ("," . wsp) ",\n"
+
+let first_perm : canonizer = bib_author . wsp_to_comma . bib_title
+
+let second_perm : canonizer = bib_title . wsp_to_comma . bib_author
+
+let bibtex : canonizer = 
+"@book{" . label. wsp_to_comma . (first_perm | second_perm) . "}"
+\end{verbatim}
+
+Using the in-built \verb!canonizer_of_lens! function which extracts a canonizer
+from a lens, it is easy to define a an equivalence relation that sends all
+members of a regular language to a canonical representative. For example
+the expression 
+\begin{verbatim}
+let wsp_to_space : canonizer = canonizer_of_lens ((del wsp) . (ins " "))
+\end{verbatim}
+
+\noindent defines a canonizer which sends arbitrary occurences of whitespace
+characters to a single space. However, a problem arises when trying to specify
+the equivalence relation which nests this projection relation with the relation
+that considers all permutations of the \bibtex{} fields equivalent.
+
+Specifically, it becomes necessary to specify each of the permutations
+seperately, since Boomerang's in-built permutation combinator does not account
+for separators in between the regular expressions being permuted:
+
+\begin{verbatim}
+let first_perm : canonizer = bib_author . wsp_to_comma . bib_title
+
+let second_perm : canonizer = bib_title . wsp_to_comma . bib_author
+\end{verbatim}
+
+One possible solution is to permute the alternative canonizers
+
+\noindent \verb!bib_author. wsp_to_comma! and \verb!bib_title . wsp_to_comma!,
+but Boomerang's type-checker will reject this expression since the permutation
+combinator accepts regular expressions rather than canonizers. In any case, the
+regular expression
+\begin{verbatim}
+bib_author . wsp_to_comma . bib_title . wsp_to_comma | 
+bib_title . wsp_to_comma . bib_author . wsp_to_comma
+\end{verbatim}
+is not equivalent to,
+\begin{verbatim}
+bib_author . wsp_to_comma . bib_title  | 
+bib_title . wsp_to_comma . bib_author
+\end{verbatim}
+thus the programmer would have to write more lenses and canonizers to correct
+this. Therefore, in order to utilize this trick, the programmer might choose to
+specify all the regular expressions first (including the whole regular
+expressions canonized by the \verb!bib_author! and \verb!bib_title!
+canonizers), and then the canonizers separately. Alternatively, he/she may stick
+to using canonizers and then liberally make use of the built-in
+functions \verb!uncanonized_type! and \verb!canonized_type! which return the
+regular expression matching all the data and the regular expression matching
+the canonical representatives respectively.
+
+In any case, all of these alternatives lead to verbose code that is
+difficult to understand. Moreover, if the \bibtex{} records had 3 or more
+fields or even nested permutations, then the programmer would quickly find
+themselves in a bind with these workarounds.
 
 To address this issue, we propose the language of {\em Quotient Regular
-Expressions} or QREs, which enables the programmer to explicitly define a broad
-class of equivalence relations directly on regular languages using a compact
-and convenient notation, and hence negates the need to specify the data and the
-equivalence relation on the data separately. For example, a QRE which
-specifies the equivalence relation on the \bibtex{} citation records is as
-follows:
-
-\begin{verbatim}
-let bib_author  = 
-"author = {" . squash bib_space_names -> bib_comma_names using 
-(get (synth bib_space_names <=> bib_comma_names)) . "}"
-
-let bib_ignore_wsp =  project ("," . [ \n\t\r]* ) -> ",\n"
+Expressions} or QREs, which express a broad class of equivalence relations
+directly on regular languages using a compact and convenient notation.
+
+With QREs we can specify \bibtex{} records as follows:
+
+\begin{verbatim}
+let wsp_to_space : canonizer = project [ \n\t\r]+ -> " "
+
+let valid_name : canonizer = 
+name | (name . wsp_to_space . (name . wsp_to_space)* . name)
+
+let wsp_to_and : canonizer = 
+project ([ \n\t\r]+ . "and" . [ \n\t\r]+) -> " and "
+
+let bib_names : canonizer =
+
+valid_name | (valid_name . wsp_to_and . (valid_name . wsp_to_and)* . valid_name)
+
+let bib_author : canonizer = "author = \"" . bib_names . "\""
+
+let label : regexp = [a-zA-Z0-9]+
+
+let title : canonizer = word | (word . wsp_to_space . (word . wsp_to_space)* .
+word)
+
+let bib_title : canonizer = "title = {" . title . "}"
 
 let bib_perm  : canonizer = perm (label, bib_author, bib_title) with
-bib_ignore_wsp
-
-let bibtex  = "@book{" . bib_perm . "}"
-\end{verbatim}
-
-\sam{Should we have a special construct like squash-synth(R, R') in addition to
-squash(R, R', l), or maybe even just have only squash-synth(R, R')?}
-
-The first expression specifies the author field as
-having one of two representations (\verb!bib_space_names! and
-\verb!bib_comma_names!), with the \verb!bib_comma_names! being the
-canonical representation and the expression
-
-\noindent \verb!(get (synth bib_space_names <=> bib_comma_names))! giving the
-function which converts names described by \verb!bib_space_names! to their canonical
-form described  by \verb!bib_comma_names!.
-
-The second expression specifies the regular language which is a comma followed
-by arbitrary occurences of whitespace characters quotiented by the equivalence
-relation which makes all members of this language equivalent, with the canonical
-representation being a comma followed by a newline character.
-
-The third expression specifies that the label, title and author fields
-are can occur in any order, with each of these fields separated by a comma and
-arbitrary amounts of whitespace. The canonical representaion is that in which
-the label, author and title fields occur in that order, separated by a comma and
-a newline.
-
-However, even though QREs enable the programmer to specify the data as well as
-the equivalence relation on the data in the same expression, the programmer
+(project ("," . [ \n\t\r]* ) -> ",\n")
+
+let bibtex : canonizer = "@book{" . bib_perm . "}"
+\end{verbatim}
+
+The expression \verb!project [ \n\t\r]+ -> " "! is a QRE which matches any
+number of whitespace characters, and then considers all strings of whitespace
+characters as equivalent with a single space as the canonical representative. 
+Another QRE in the example above is the
+
+\begin{center}
+\begin{verbatim}
+perm (label, bib_author, bib_title) with (project ("," . [\n\t\r]* ) -> ",\n")
+\end{verbatim}
+\end{center}
+
+\noindent QRE, which specifies that the label, title and author fields can occur
+in any order, with the fields separated by a comma followed by arbitrary amounts
+of whitespace. The canonical representation is that in which the label, author
+and title fields occur in that order, separated by a comma followed a newline.
+The \verb!perm! QRE is defined at the level of canonizers which allows for
+easy nesting of permutations. Similarly, it is easy to specify any number of
+fields within the \verb!perm! QRE.
+
+While QREs kill two birds with one stone by specifying the
+data and the equivalence relation on the data in one expression, he/she
 still has to manually program the underlying lens which maps the equivalence
-classes of the \bibtex{} records to the equivalence classes of the EndNote
-records
-
-\sam{TODO: PUT BIJECTIVE LENS HERE!},
-
-\noindent task which is itself challenging and erro-prone.
-
-Therefore, to address this challenge we demonstrate how to {\em synthesize}
-quotient lenses from a pair of QREs and a set of input-output example pairs. Our
-key insight is to reduce the problem of synthesizing QRE lenses to the problem
-of synthesizing {\em bijective lenses}, a problem which we solved in out prior
-work ~\cite{popl18}.
+classes of the source data to the equivalence classes of the view data, a task
+which is itself tedious and error-prone as the code snippet below illustrates:
+\begin{verbatim}
+let preamble : (lens in "@book{" <=> "%0 Book\n%F ") = 
+    del "@book{" . ins "%0 Book\n%F "
+
+let author_lens : (lens in (",\n" . bib_author) <=> ("\n" . end_author)) = 
+    del ",\nauthor = \""
+    . ins "\n%A "
+    . valid_name
+    . (del " and "
+         . ins "\n"
+         . (ins "%A " . valid_name . del " and " . ins "\n")*  . ins "%A "
+         . valid_name
+           || "")
+
+let title_lens : 
+(lens in ("\",\n" . bib_title . "},\n}") <=> ("\n" . end_title)) =
+del "\",\ntitle = {" . ins "\n%T " . title . del "},\n}"
+
+let bib_to_end : (lens in bib_tex <=> end_note) =
+preamble . label . author_lens . title_lens
+\end{verbatim}
+
+This lens transforms the \bibtex{} data to the EndNote (and vice versa) data by
+matching the delimiters and separators in a bijective manner. For example,
+the \verb!preamble! lens replaces the \bibtex{} preamble \verb!@book! with the
+EndNote preamble \verb!%0 Book\n%F ! in the \bibtex{record}. The programmer
+can ill afford to make a mistake with any one of the small but vital details
+of the transformations since even a one character mistake might render the lens
+incorrect. This makes lens programming tedious and error-prone.
+
+Therefore, to address this challenge, we demonstrate how to {\em synthesize}
+quotient lenses from a pair of QREs and a set of input-output example pairs.
+Our key insight is to reduce the problem of synthesizing QRE lenses to the
+problem of synthesizing {\em bijective lenses}, a particular subset of the
+lenses that can be expressed in Boomerang. This is a problem which we solved in
+our prior work ~\cite{popl18}.
 
 Our approach is to define a class of {\em QRE lenses} which are the same as
-bijective lenses with canonizers at the ends. That is, a QRE lens uses the
-source QRE to compute a unique representative for the data modulo the
-equivalence relation defined by the QRE before applying a bijective lens to
-this representative in the forward direction, and operates similarly same in the
-backward direction, but using the view QRE.
-
-\sam{TODO: PUT LENS WITH CANONIZERS ON THE EDGE DIAGRAM HERE!}
-
-One potential pitfall with QRE lenses is that the composition of QRE lenses may
-not have the normal form of a bijective lens with canonizers on the ends:
-
-\sam{TODO: PUT COMPOSITION DIAGRAM HERE!}
-
-Our main technical contribution in this paper is a proof that QRE lenses are
-closed under left quotienting, right quotienting, composition, and regular
-combinators. This proof will allow justify our approach for synthesizing
-quotient lenses so that the programmer can simply type
-
-$$\mathit{synth} \; \verb!bibtex! \; \verb!endnote!$$
-
-to derive a quotient lens that does the \bibtex{} to EndNote transformation.
-\bcp{What about examples??}
+bijective lenses with ``canonizers at the edges'' as in
+Figure~\ref{fig:attheedges}. In the forward direction, a QRE lens $q$ uses the
+source QRE $c$ to compute a unique representative for the data modulo the
+equivalence relation defined by $c$ and then applies the $\get$ function of a
+bijective lens $\ell$ to this representative. In the backward direction, $q$
+operates similarly, but using the view QRE $c'$ and the $\lput$ function of
+$\ell$.
+
+\begin{figure}[t]
+  \centering
+  \includegraphics[width=\textwidth]{attheedges}
+  \caption{Quotient Lens with ``Canonizers at the Edges''\sam{This is a
+  screenshot from Nate's thesis. Are we allowed to use it?}}
+  \label{fig:attheedges}
+\end{figure}
+
+One potential pitfall with QRE lenses is that the combinators which
+we define on them may not preserve this normal form. For example, the
+composition $q' \circ q$ of QRE lenses $q$ and $q'$ may not be equivalent to
+any quotient lens which has the canonical normal form of a bijective lens with
+canonizers a the edges. Our main technical contribution in this paper is a
+proof that QRE lenses are closed under left quotienting, right quotienting,
+composition, and regular combinators. This proof will allow justify our
+approach for synthesizing quotient lenses so that the programmer can simply
+type
+
+\begin{verbatim}
+let bib_to_end : (lens in bibtex <=> endnot) 
+= synth bibtex <=> endnote  using  {(bib_example, end_example)}
+\end{verbatim}
+\noindent to synthesize a quotient lens that maps between \verb!bibtex! and
+\verb!endnote! QREs, and sends the equivalence class of \verb!bib_bexample! to
+the equivalence class of \verb!end_example!.
 
@@
 
-For example, suppose that the programmer wishes to write a regular expression
-which describes \bibtex{} citation records such as:
+We have already encountered the \verb!project R -> s! QRE which sends all
+elements of the language described by the regular expression $R$, and the
+
+\noindent \verb!perm (q_1, ..., q_n) with q! QRE which sends all permutations of
+the QREs $q_1, \ldots, q_n$ separated by the QRE $q$ to the canonical
+QRE $q_1 \cdot q \cdot \ldots \cdot q \cdot q_n$.
+
+Now consider a variant of the \bibtex{} to EndNote example transformation,
+where, the only quotienting occurs in the author field, where the author's name
+can have the last name followed by a comma and then the rest of the name, or
+remain as before where the names are separated by a space.
+
+For example, these two records should be considered equivalent:
 \begin{verbatim}
 @Book {conway,
-  Author = {Conway, J. H.},
+  Author = "Conway, J. H.",
   Title = {Regular Algebra and Finite Machines},
@@
 \end{verbatim}
-
-Each record may be represented in many different ways. Firstly, the author's
-name can have the last name followed by a comma, and then the
-first and second names separated by a space, or the last name, first name and
-second name separated by spaces in between them. For example, each of these
-\bibtex{} records is equivalent to the other:
 \begin{verbatim}
 @Book {conway,
-  Author = {Conway, J. H.},
+  Author = "J. H. Conway",
   Title = {Regular Algebra and Finite Machines},
@@
 \end{verbatim}
-\begin{verbatim}
-@Book {conway,
-  Author = {Conway J. H.},
-  Title = {Regular Algebra and Finite Machines},
-}
-\end{verbatim}
-
-Secondly, the author and title fields may occur in either order:
-\begin{verbatim}
-@Book {conway,
-  Author = {Conway, J. H.},
-  Title = {Regular Algebra and Finite Machines},
-}
-\end{verbatim}
-\begin{verbatim}
-@Book {conway,
-  Title = {Regular Algebra and Finite Machines},
-  Author = {Conway, J. H.},
-}
-\end{verbatim}
-Thirdly, the title and author records can be separated by arbitrary amounts of
-whitespace characters. For example, if the title field occurs first, then the
-author field may occur not on the line immediately below the title field, but
-instead on the second line below the title field:
-\begin{verbatim}
-@Book {conway,
-  Title = {Regular Algebra and Finite Machines},
-
-  Author = {Conway, J. H.},
-}
-\end{verbatim}
-In the example above, a \bibtex{} record may have different
-representations for three different reasons. The first reason
-is that the author's name can have the last name followed by a comma and then
-the first and second names separated by a space, or the last name, first name and
-second name separated by spaces in between them. The $\squash{R}{R'}{f}$ QRE
-enables the programmer to choose between the representations described by $R$
-and $R$', provided that they can provide a function $f$ that maps data stored in
-the first format to data stored in the second format. In this example we have
-the two regular expressions
-
-\begin{align*}
-\mathtt{COMMA\_NAME} &= \mathtt{NAME} \cdot ``," \cdot \mathtt{NAME} \cdot ``
-\quad " \cdot \mathtt{NAME}\\
-\mathtt{SPACE\_NAME} &= \mathtt{NAME} \cdot ``," \cdot \mathtt{NAME} \cdot ``
-\quad " \cdot \mathtt{NAME}\\
-\end{align*}
-
-and the bijective lens 
-$$ \ell = (\mathit{copy} \; \mathtt{COMMA\_NAME}) \cdot (\mathit{ins} \; ``,")
-\cdot (\mathit{del} \; `` \quad") \cdot (\mathit{copy} \; \mathtt{COMMA\_NAME})
-\cdot (\mathit{copy} \; \mathtt{COMMA\_NAME})$$
-therefore the QRE
-$$\mathtt{AUTHOR}
-=
-``\text{Author
-= \{}" \cdot \squash{\mathtt{COMMA\_NAME}}{\mathtt{SPACE\_NAME}}{\ell.get}
-\cdot ``\}"$$ enables the programmer to express the fact that the two different
-representations of the author's name are equivalent, with the
-$\mathtt{COMMA\_NAME}$ representation chosen as the canonical representation,
-and the function $\ell.get$ giving the ``canonizing'' function which converts
-data in the $\mathtt{SPACE\_NAME}$ format to its $\mathtt{COMMA\_NAME}$
-representation. The $\squash{R}{R'}{f}$ has the added advantage that if the
-transformation from $R$ to $R'$ can be expressed by a bijective lens,
-then the programmer may attempt to \textit{synthesize} the lens from $R$ to $R'$
+
+The equivalence described is one in which there is a bijection between 
+the regular expressions \verb!comma_name! and \verb!space_name! given by,
+\begin{verbatim}
+let comma_name : regexp = name | (name . ", " . name . (" " . name)*)
+
+let space_name : regexp = name | (name . " " . (name . wsp_to_space)* . name)
+
+\end{verbatim}
+
+\noindent with the lens \verb!comma_to_space! mapping between them given by 
+\begin{verbatim}
+let comma_to_space : (lens in space_name <=> comma_name) =
+lens_swap 
+(name . del " " . (ins " " . name . del " ")* ) (name . ins ", ") || name 
+\end{verbatim}
+
+The $\squash{R}{R'}{f}$ QRE is designed to make it easy to express such
+equivalences easily. Specifically, given regular expressions $R$ and $R'$, and
+a function $f : \mathcal{L}(R) \longrightarrow \mathcal{L}(R')$, the
+$\squash{R}{R'}{f}$ QRE expresses the equivalence relation on $\mathcal{L}(R \;
+R')$ which sends each element $r \in \mathcal{L}(R)$ to the canonical element
+$f(r) \in \mathcal{L}(R')$, and sends each element $r' \in \mathcal{L}(R')$ to
+itself:
+
+\begin{verbatim}
+let squash_author : canonizer =  
+squash space_name -> comma_name using (get comma_to_space)
+
+let author : canonizer = "author = \"" . squash_author . "\""
+\end{verbatim}
+
+In this example, the \verb!squash_author! QRE expresses the fact that the
+\verb!space_name! and \verb!comma_name! representations of the authors' name
+are equivalent, with the \verb!comma_name! representation chosen as the
+canonical representation, and the function \verb!comma_to_space.get! giving
+the canonizing function which converts data in the \verb!space_name! format
+to its \verb!comma_name! representation.
+
+The $\squash{R}{R'}{f}$ has the added advantage that if the transformation from
+$R$ to $R'$ can be expressed by a bijective lens that is consistent with a set
+of examples, then the programmer can synthesize the lens from $R$ to $R'$
 because of our previous work where we demonstrated how to synthesize bijective
@@
 pairs:
-$$\mathtt{AUTHOR}
-=
-``\text{Author
-= \{}" \cdot \squash{\mathtt{COMMA\_NAME}}{\mathtt{SPACE\_NAME}}{\mathit{synth}
-\; \mathtt{COMMA\_NAME} \; \mathtt{SPACE\_NAME}} \cdot ``\}"$$
-
-The second reason that a record may have two different representations is that
-the author and title fields may occur in a different order, and the third
-reason is that the author and title fields may have arbitrary occurences of
-whitespace characters in between them. The $R \mapsto s$ and $\perm{q_1,
-\ldots, q_n}{q}$ QREs enable the programmer to handle each of these cases:
-$$\perm{\re{AUTHOR}, \re{TITLE}}{((``," \cdot \re{WHITESPACE}^*) \mapsto \string
-``, \backslash n")}$$
-This QRE enables the programmer to consider any permutation of the author and
-title fields interspersed with a comma and arbitrary whitespace as equivalent.
-The $((``," \cdot \re{WHITESPACE}^*) \mapsto \string ``, \backslash n")$
-subexpression is itself a QRE that denotes that all string $s$ which consists
-of a comma $``,"$ followed by any number of whitespace should be considered
-equivalent, with the canonical representative being the string $\string ``,
-\backslash n"$. The canonical representation of \bibtex{} records is that in
-which the author field occurs the title field with a comma then a newline
-character separating the two.
+
+\begin{verbatim}
+let comma_to_space : (lens in space_name <=> comma_name) = 
+synth space_name <=> comma_name using {("J. H. Conway", "Conway, J. H.")}
+\end{verbatim}
 
@@
 $*$ (iteration) behave like the regular combinators for regular languages. For
-instance, if the QRE for matching a \bibtex{} \; record is named
-$\mathtt{RECORD}$, then the QRE for matching two records is $\mathtt{RECORD}
-\cdot \mathtt{RECORD}$, the QRE for matching one or two records is $\mathtt{RECORD}
-\sep (\mathtt{RECORD} \cdot \mathtt{RECORD})$ and the QRE for matching any
-number of QRE records in $\mathtt{RECORD}^*$.
-
-The $\normalize{R}{R'}{f}$ QRE is the most general QRE in that for each of the
-the other QREs $q$, there exist regular expressions $R, R'$ and a surjective,
-idempotent function $f:\mathcal{L}(R) \longrightarrow \mathcal{L}(R')$ such that
-$q$ $\normalize{R}{R'}{f}$ each denote the same equivalence relation on the
-same regular language. $\normalize{R}{R'}{f}$ enables the programmer to
-provide a canonizing function from $\mathcal{L}(R)$ to $\mathcal{L}(R')$,
-with the equivalence class on $\mathcal{L}(R)$ being the equivalence relation
-defined by the {\em fibres} of $f$ (i.e. $r \sim r'$ if and only if $f(r) =
-f(r'))$. The canonizing function is required to be surjective and idempotent.
-The $\normalize{R}{R'}{f}$ combinator therefore enables the programmer to express
-equivalence relations which are difficult to express, or which cannot be
-expressed using the other combinators.
+instance, if the QRE for matching a \bibtex{} record is named
+\verb!bibtex!, then the QRE for matching two records is 
+\verb!bibtex . bibtex!, the QRE for matching one or two records is
+\verb! bibtex | (bibtex . bibtex)! and the QRE for matching any number of QRE
+records \verb!bibtex*!
+
+The $\normalize{R}{R'}{f}$ expresses the equivalence relation which sends each
+element in $\mathcal{L}(R)$ to its canonical representation $f(r) \in
+\mathcal{L}(R')$. The equivalence relation defined by
+$\normalize{R}{R'}{f}$ is thus the equivalence relation on
+$\mathcal{L}(R)$ defined by the {\em fibres} of $f$ (i.e. $r \sim r'$ if and
+only if $f(r) = f(r'))$. The canonizing function $f$ is required to be
+surjective and idempotent.
+
+The $\normalize{R}{R'}{f}$ QRE is the most general QRE in that for each QREs
+$q$, there exist regular expressions $R, R'$ and a surjective, idempotent
+function $f:\mathcal{L}(R) \longrightarrow \mathcal{L}(R')$ such that $q$ and
+$\normalize{R}{R'}{f}$ each define the same equivalence relation on
+$\mathcal{L}(R)$, with $\mathcal{L}(R')$ forming a complete set of
+representatives for this equivalence relation. The $\normalize{R}{R'}{f}$
+combinator is included so as to express equivalence relations which are
+difficult to express, or which cannot be expressed, using the other
+combinators. However, checking that the canonizing function $f$ is
+surjective and idempotent is undecidable, thus we do not use the
+$\normalize{R}{R'}{f}$ in practice. On the other hand, the
+$\normalize{R}{R'}{f}$ combinator gives a sufficient condition for potential
+QREs to be ``well-behaved'', particularly with respect to our strategy for
+synthesizing quotient lenses which we shall introduce in Section~\ref{synth}.
+
+While QREs express a broad class of equivalence relations on regular languages,
+they do not express all the equivalence relations that can be expressed in
+Boomerang, nor of course all the equivalence relations which cannot be
+expressed using Boomerang canonizers. For instance, Boomerang canonizers can
+map a regular language described by $R$ to a regular language described by
+$S$, with $\mathcal{L}(S) \not \subseteq \mathcal{L}(R)$, whereas each QRE $q$
+canonizes a regular language described by $R$ to regular subset
+$\mathcal{L}(R') \subseteq \mathcal{L}(R)$ described by $R'$.
 
@@
 and is necessary, firstly because it ensures that the canonizing function of a
-QRE is well-defined. For example if $\mathtt{RECORD}$ is the QRE for a \bibtex
-record in our \bibtex{} to and $W(\mathtt{ENDNOTE} \cdot \mathtt{ENDNOTE})$ is not
-unambiguous, that is there is a string $s$ such that $s = s_1 \cdot s_2 =
-{s_1}' \cdot {s_2}'$ with $s_1, s_2, {s_1}', {s_2}' \in W(\mathtt{ENDNOTE} \cdot
-\mathtt{ENDNOTE})$, then the canonizer $f$ for $\mathtt{ENDNOTE} \cdot
-\mathtt{ENDNOTE}$ may not be well defined since we are not guaranteed that
-$f(s_1) \cdot f(s_2) = f({s_1}') \cdot f({s_2}')$. We also require that the
-regular combinators applied to the kernels are unambiguous since the underlying
-lenses will end up operating on the kernels, and these lenses impose the same
-restrictions for similar reasons.
+QRE is well-defined.
+
+For example consider the QRE
+
+\begin{verbatim}
+let ambiguous : canonizer = "a"* . (project "a"* -> "a")
+\end{verbatim}
+
+\noindent The behaviour of \verb!ambiguous! is not-well defined since the string
+\verb!"aaa"! can be canonized to any of \verb!"a", "aa", "aaa"! depending on how
+\verb!"aaa"! is parsed. We also require that the regular combinators applied to
+the kernels are unambiguous since the underlying lenses will end up operating
+on the kernels, and these lenses impose the same restrictions for similar
+reasons.
 
@@
 \label{QRE-lenses} 
-As we have seen, QREs enable the programmer to express a regular expression $R$
-as well as an equivalence relation on $\mathcal{L}(R)$. QREs are therefore a
-good \textit{specification} language for quotient lenses. In this section of the
-paper, we introduce \textit{QRE Lenses}. QRE lenses are quotient lenses based on
-the Boomerang quotient lens combinators, but which use QREs to describe their
-source and view types.
+As we have seen, QREs express a broad class of equivalence relations
+directly on regular languages.  QREs are therefore a good \textit{specification}
+language for quotient lenses. In this section of the paper, we introduce
+\textit{QRE Lenses}. QRE lenses are quotient lenses based on the Boomerang
+quotient lens combinators, but which use QREs to describe their source and view
+types.
 
@@
 
-A potential problem with this approach is that quotient lenses of this form may
-not necessarily be closed under the action of the lens combinators. For
-example, given two lenses $q$ and $q'$, the composition of $q' \circ q$ may not
-be of this form.
-
-Consequently, in order to define quotient lenses, we need to choose a class of
-bijections that has just the right properties with respect to the quotient
-lens combinators. Our choice is to use the class of \textit{bijective
-lenses} which we identified and studied in our previous work~\cite{popl18}.
-Bijective lenses, which we briefly discuss in the next subsection, possess
-the properties needed to ensure that all QRE lenses have the normal form
-suggested in \cref{normalform}.
+\noindent One potential pitfall with QRE lenses is that the combinators which
+we define on them may not preserve this normal form. For example, the
+composition $q' \circ q$ of QRE lenses $q$ and $q'$ may not be equivalent to
+any quotient lens which has the canonical normal form of a bijective lens with
+canonizers a the edges. Consequently, in order to define quotient lenses, we
+need to choose a class of bijections that has just the right properties with
+respect to the quotient lens combinators. Our choice is to use the class of
+\textit{bijective lenses}, a subset of the lenses expressible in Boomerang which
+we identified and studied in our previous work~\cite{popl18}. Bijective lenses,
+which we briefly discuss in the next subsection, possess the properties needed
+to ensure that all QRE lenses have the normal form suggested in
+\ref{normalform}.
 
 \subsection{Bijective Lenses}
-Given regular expressions $R, S$ and equivalence
-relations $\sim_R, \sim_S$ defined on $\mathcal{L}(R)$ and $\mathcal{L}(S)$
-respectively, a \textit{bijective quotient lens} $q :
-R/{\sim_R}{\Longleftrightarrow} S/{\sim_S}$ from $R$ to $S$ is
-a pair of functions $q.get:
-\mathcal{L}(R) \longrightarrow \mathcal{L}(S)$ and $q.put : \mathcal{L}(S)
-\longrightarrow \mathcal{L}(R)$ such that
-\begin{align*}
-q.put \; (q.get \; r) &\sim_R r\\
-q.get \; (q.put \; s) &\sim_S s
-\end{align*}
-Additionally the components of $q$ must respect $\sim_R$ and $\sim_S$ i.e.
-if $r \sim_R r'$ and $s \sim_S s'$ then $q.get \; r \; \sim_S q.get \; r'$
-and $q.put \; s \sim_R q.put \; s' \; s'$.
-
-We define the set of \textit{bijective lenses} to be the set of bijections
+Define the set of \textit{bijective lenses} to be the set of bijections
 between regular languages created from Boomerang lens combinators.
@@
 $\ell$ to be considered a quotient lens where the equivalence relation applied
-to the source and view types is the equality relation. The $\mathit{lquot}(c,
-q)$ combinator takes a quotient lens $q$ and quotients the source data using
-$c$, assuming that the source data forms a complete set of representatives for
-the equivalence relation $\eqrel{c}$. The $\mathit{lquot}(q, c)$ does the same
-but on the view data. The regular combinators $\cdot \; * \; |$ exhibit the
-expected behaviour. \sam{Examples needed?}
-
-The difference between QRE lenses and Boomerang quotient lenses is the typing
-judgements. More concretely, each QRE lens $q$ has a type $q : c \Leftrightarrow
-c'$ where $c, c'$ are QREs. In contrast, the approach used to type quotient
-lenses in Boomerang is to classify equivalences according to whether they are
-or are not the equality relation. This type system is based on two
-observations: first, that most quotient lenses originate as lifted basic
+to the source and view types is the equality relation, and the regular
+combinators $\cdot \; * \; |$ on QRE lenses exhibit the expected behaviour,
+with the caveat that the regular expressions used by these lenses are strongly
+unambiguous.
+
+The more interesting combinators are the $\mathit{lquot}$ and $\mathit{rquot}$
+combinators. The $\mathit{lquot}(c, q)$ combinator takes a quotient lens $q$ and
+quotients the source data using $c$, assuming that the source data forms a
+complete set of representatives for the equivalence relation $\eqrel{c}$. The
+$\mathit{lquot}(q, c)$ combinator does the same but on the view data. For
+example, recall that in the \bibtex{} to EndNote transformation, we had the QREs
+\verb!bibtex! and \verb!endnote! describing the \bibtex{} and EndNote records
+respectively, and the bijective lens \verb!bib_to_end! that maps between
+\verb!bibtex! and \verb!endnote!. The quotient lens which maps between
+\verb!bibtex! and \verb!endnote! is given by the QRE lens
+
+\begin{verbatim}
+let bib_to_end_q : (lens in bibtex <=> endnote) = 
+rquot (lquot (bibtex, bib_to_end), endnote)
+ \end{verbatim}
+
+The difference between QRE lenses and Boomerang quotient lenses lies in the
+typing judgements. More concretely, each QRE lens $q$ has a type $q : c
+\Leftrightarrow c'$ where $c, c'$ are QREs. In contrast, the approach used to
+type quotient lenses in Boomerang is to classify equivalences according to
+whether they are or are not the equality relation. This type system is based on
+two observations: first, that most quotient lenses originate as lifted basic
 lenses, and therefore have types whose equivalence relations are both equality;
@@
 \subsection{Semantics of QRE Lenses}
-As we mentioned earlier, every quotient lens $q$ has a type $q :
-c \Longleftrightarrow c'$ where $c$ and $c'$ are QREs. The denotation
-$\llbracket q \rrbracket$ of a QRE lens $q:c \Leftrightarrow c'$ is a quotient
-lens $\llbracket q \rrbracket : W(c)/{\eqrel{c}} \Longleftrightarrow
-W(c')/{\eqrel{c'}}$. The typing rules and denotation of QRE lenses are given in
-Figure~\ref{fig:qlenssemantics}. The following theorem confirms that QRE lenses
-exhibit the intended behaviour.
+The denotation $\llbracket q \rrbracket$ of a QRE lens $q:c \Leftrightarrow c'$
+is a quotient lens $\llbracket q \rrbracket : W(c)/{\eqrel{c}}
+\Longleftrightarrow W(c')/{\eqrel{c'}}$. The typing rules and denotation of QRE
+lenses are given in Figure~\ref{fig:qlenssemantics}. The following theorem
+confirms that QRE lenses exhibit the intended behaviour.
 
@@
 If there is a derivation $q:c \Leftrightarrow c'$ then
-$\llbracket q \rrbracket : W(c)/\eqrel{c} \Leftrightarrow W(c')/\eqrel{c'}$ is a
-well-defined quotient lens.
+$\llbracket q \rrbracket : W(c)/{\eqrel{c}} \Leftrightarrow
+W(c')/{\eqrel{c'}}$ is a well-defined quotient lens.
 \end{theorem}
@@
 \label{synth}
-By Theorem~\ref{normal form}, if there is a derivation $q : c \Leftrightarrow
-c'$ of a QRE lens, then there exists a bijective lens $\ell : K(c)
-\Leftrightarrow K(c')$ such that
+Our goal in this section is to demonstrate how to synthesize a quotient lens
+$q: c \Leftrightarrow c'$ from the QREs $c, c'$ and a set of input-output pairs
+$\{(x_1, y_1), \ldots, (x_n, y_n)\}$ where the $x_i$'s are in $W(c)$ and the
+$y_i$'s are in $W(c')$.
+
+We have shown in Theorem~\ref{normal form} that if there is a derivation $q : c
+\Leftrightarrow c'$ of a QRE lens, then there exists a bijective lens $\ell :
+K(c) \Leftrightarrow K(c')$ such that
 \begin{align*}
@@
 
-Now suppose that we wish to synthesize a quotient lens $q$ of type $q: c
-\Leftrightarrow c'$ that maps a specified set of input strings to a
-specified set of output strings and vice versa.
-
-For example, in the \textsc{Bib}\TeX\ to EndNote problem, the programmer wishes
-to synthesize a QRE lens $q$ which maps the QRE
-$$\perm{\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},
-\re{YEAR}}{(\re{WHITESPACE}^* \mapsto \string \n)}$$ 
-to the QRE
-$$\perm{\re{REF'}, \re{AUTHOR'}, \re{TITLE'}, \re{PUBLISHER'},
-\re{YEAR'}}{(\re{WHITESPACE}^* \mapsto \string \n)}$$
-where $\re{REF}, \re{AUTHOR}, \re{TITLE}, \re{PUBLISHER},$ and $\re{YEAR}$
-(resp. $\re{REF'}, \re{AUTHOR'}, \re{TITLE'}, \re{PUBLISHER'},$ and
-$\re{YEAR'}$) are regular expressions describing the respective
-\textsc{Bib}\TeX\; (resp. EndNote\bcp{Don't like this ``resp.''
-  construction---saves space, but not reader energy}) fields of the same
-name, and 
-$\re{WHITESPACE}$ is a regular expression describing whitespace
-characters. Morever, the synthesized lens should map the equivalence
-class of this \textsc{Bib}\TeX{} record
-\begin{verbatim}
-@Book {conway,
-  Author = {Conway, J. H.},
-  Title = {Regular Algebra and Finite Machines},
-  Publisher = {Printed in GB by William Clowes & Sons Ltd},
-  Year = {1971}
-}
-\end{verbatim}
-\noindent to the equivalence class of this EndNote entry:
-\begin{verbatim}
-%0 Book
-%T Regular Algebra and Finite Machines
-%A Conway, J. H.
-%D 1971
-%I Printed in GB by William Clowes & Sons Ltd
-%F conway
-\end{verbatim}
+Now since the $x_i$'s are in $W(c)$ and the $y_i$'s are in $W(c')$, then
+
+\noindent ${x_i}' = \canonizer(c)(x_i)$ is in $K(c)$ and ${y_i}' =
+\canonizer(c')(y_i)$ is in $K(c')$, hence in order to synthesize the desired
+quotient lens $q$ it suffices to synthesize a bijective lens $l : K(c)
+\Leftrightarrow K(c')$ that is consistent with the canonized examples
+$\{({x_1}', {y_1}'), \ldots, ({x_n}', {y_n}')\}$
 
@@
 of input-output pairs $\{(r_1, s_1), \ldots, (r_n, s_n)\}$, if there is a
-derivation $\ell :, R \Leftrightarrow S$ of a bijective lens that maps $r_i$ to
+derivation $\ell : R \Leftrightarrow S$ of a bijective lens that maps $r_i$ to
 $s_i$ in the forward direction and $s_i$ to $r_i$ in the backward direction,
@@
 \Leftrightarrow S'$ such that $R \equiv R'$, $S \equiv S'$, $\ell'$ maps $r_i$
-to $s_i$ in the forward direction and $s_i$ to $r_i$ in the backward
-direction, and $\llbracket \ell \rrbracket = \llbracket \ell'
-\rrbracket$~\cite{popl18}.
+to $s_i$ in the forward direction and $s_i$ to $r_i$ in the backward direction,
+and $\llbracket \ell \rrbracket = \llbracket \ell' \rrbracket$~\cite{popl18}.
+
+This implies that if there is a derivation $q : c \Leftrightarrow
+c'$ of a QRE lens as well as a derivation $\ell : K(c) \Leftrightarrow K(c')$
+that is consistent with the canonized examples $\{({x_1}', {y_1}'),
+\ldots, ({x_n}', {y_n}')\}$, then we can synthesize a quotient lens $q$ of
+type $c \Leftrightarrow c'$ that is consistent with $\{({x_1}', {y_1}'),
+\ldots, ({x_n}', {y_n}')\}$.
+
+More specifically, in order to synthesize a quotient lens $q: c \Leftrightarrow
+c'$, that is consistent with the examples $\{(x_1, y_1), \ldots, (x_n, y_n)\}$
+ it suffices to perform the following procedure:
+\begin{enumerate}
+  \item
+  Compute $K(c)$, $K(c')$,
+  \item
+  compute $\{({x_1}', {y_1}'), \ldots, ({x_n}', {y_n}')\}$ where ${x_i}' =
+  \canonizer(c)(x_i)$ and ${y_i}' = \canonizer(c')(y_i)$,
+  \item
+    synthesize a bijective lens $\ell : K(c) \Leftrightarrow K(c')$ that is
+    consistent with the examples $\{({x_1}', {y_1}'), \ldots, ({x_n}',
+    {y_n}')\}$
+  \item 
+  return $q = \mathit{rquot}(\mathit{lquot}(c, \ell), c') : c \Leftrightarrow
+  c'$
+\end{enumerate}
+As an example, recall that
+in the \bibtex{} to EndNote transformation we had the QREs \verb!bibtex! and \verb!endnote! describing a \bibtex{} record and an
+Endnote record respectively. We also had the example pair 
+\verb!(bib_example , end_example)! where
+
+\begin{verbatim}
+bib_example = 
+"@Book {conway,
+  Author = \"J. H. Conway\",
+  Title = {Regular Algebra and Finite Machines},
+}"
+\end{verbatim}
+
+\noindent and
+
+\begin{verbatim}
+end_example = 
+"%0 Book
+%T Regular Algebra and Finite Machines
+%A J. H. Conway
+%F conway"
+\end{verbatim}
+
+Since the transformation between the kernel of \verb!bibtex! and the kernel of
+\verb!endnote! that maps \verb!bib_example! to \verb!end_example! can be
+expressed by a bijective lens, it suffices to type the expression
+
+\begin{verbatim}
+let bib_to_end : (lens in bibtex <=> endnot) 
+= synth bibtex <=> endnote  using  {(bib_example, end_example)}
+\end{verbatim}
 

### 2017-10-17 papers/QuotientLensSynth/main.tex
add in some hard data

--- 2017-10-15 papers/QuotientLensSynth/main.tex
@@
 
-\item \afm{10} of our benchmarks were derived from data.gov.  Data.gov contains
+\item \afm{8} of our benchmarks were derived from data.gov.  Data.gov contains
   many transformations between formats.  In addition to the formats previously
@@
 
-We found that \afm{35} of Opticians 39 benchmarks required modifications to be
-bijective when put in canonical form. We found that \afm{25} of Optician's 39
+We found that 10 of Opticians 39 benchmarks required modifications to be
+bijective. We found that 4 of Optician's 39
 benchmarks required modifications to be in bijective correspondence when in
-canonical form.  Furthermore, we found that \afm{all} of the data.gov examples
+canonical form.  Furthermore, we found that all of the data.gov examples
 required no modifications to be in canonical form.  While we did not put the
-data.gov examples into a form suitable for Optician, we found only \afm{1} of
-the \afm{5} data.gov examples did not require canonization, so making suitable
-for Optician would have required a modification of \afm{4} of the data.gov
-examples.
+data.gov examples into a form suitable for Optician, we found only 4 of
+the 8 data.gov examples did not require canonization, so making suitable
+for Optician would have required a modification of 4 of the data.gov
+formats.
 
@@
 We summarize the results of these experiments in Figure~\ref{?}.  \Name{} was
-able to synthesize all of Optician's benchmarks at a competitive speed with
+able to synthesize all of Optician's benchmarks at a speed competitive with
 Optician.  Furthermore, adopting these tasks to use QREs actually sped up

### 2017-10-17 papers/QuotientLensSynth/main.tex

--- 2017-10-17 papers/QuotientLensSynth/main.tex
@@
 
+We have integrated QREs and quotient lens synthesis into Boomerang with an
+additional \afm{loc} lines of OCaml code.
+
+All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
+of 1600 MHz DDR3 running macOS Sierra.
+
+Our evaluations serve to answer the following questions:
+\begin{enumerate}
+\item Does synthesizing quotient lenses from QREs permit for an easier
+  development process than existing approaches.
+
+\item Are QREs sufficiently expressive that more lenses can be synthesized
+  without modifications to the format.
+
+\item Is our synthesis algorithm competitive with existing lens synthesis
+  tools.
+\end{enumerate}
+
+\subsection{Benchmark Suite Construction} 
+Our benchmarks were constructed from two sources:
+\begin{enumerate}
+\item \afm{39} of our benchmarks were derived from the Optician system.  These
+  benchmarks were a combination of custom benchmarks, benchmarks derived from
+  FlashFill, and benchmarks derived from Augeas.  Many of the formats mapped
+  between had to be modified to work with the bijectivity constraints
+  Optician required.  Commonly, one format had whitespace where the other did
+  not allow for whitespace.  To address the bijectivity issues surrounding this,
+  we added whitespace to the format that typically did not have whitespace.  We
+  were able to remove these alterations in adopting these benchmarks to our
+  system.
+
+\item \afm{8} of our benchmarks were derived from data.gov.  Data.gov contains
+  many transformations between formats.  In addition to the formats previously
+  adopted for Optician, we synthesized transformers between these formats.
+\end{enumerate}
+
+\subsection{Easing the Development Process}
+
+To test if our tool eases the development process, we compare the following
+number of AST nodes.
+%
+\begin{itemize}
+\item[\QRESize{}] \QRESize{} is the number of AST nodes in the QRE
+  specifications involved in QRE lens synthesis task.
+\item[\CanonizerAndSpecSize{}] \CanonizerAndSpecSize{} is the number of
+  AST nodes in $W(q)$ summed with the number of AST nodes in $\canonizer(q)$,
+  for each QRE used in QRE lens synthesis task.
+\item[\LensAndSpecSize{}] \LensAndSpecSize{} is the number of AST nodes
+  in $W(q)$ summed with the number of AST nodes in the synthesized QRE lens.
+\end{itemize}
+
+The value of \QRESize{} represents the difficulty in writing our benchmarks.
+
+The value of \CanonizerAndSpecSize{} represents the difficulty in writing our
+benchmarks without QREs.  This is only an approximation, as both $W(q)$ and
+$\canonizer(q)$ are automatically generated from the QRE.  In our
+experience, these automatically generated regular expressions and canonizers are
+the ones we would naturally write.
+
+The value of \LensAndSpecSize{} represents the difficulty in writing our benchmarks
+with neither QREs nor synthesis.  This is also an approximation, as $W(q)$ and
+$\canonizer(q)$ and the bijective lens are all automatically generated.
+We have found that the generated bijective lens is fairly similar in size to the
+one we would manually write, though it is typically less modular.
+
+Figure~\ref{?} shows the number of benchmarks written with at most a given
+number of AST nodes.  We found that \QRESize{} was consistently smaller than
+\CanonizerAndSpecSize{} and \LensAndSpecSize{}, demonstrating that introducing
+QREs saves programmers programming effort compared to both Optician and basic
+Boomerang.
+
+\subsection{Increase of Expressivity}
+
+Optician was only able to synthesize fully bijective data transformations.
+\Name loosens that restriction, and is able to synthesize data which is
+bijective when put in canonical form.
+To determine the increase of expressivity, we analyzed:
+\begin{enumerate}
+\item How many data formats in Optician's benchmark suite required modifications
+  to be bijective.
+\item How many data formats in Optician's benchmark suite required modifications
+  to be bijective when put in canonical form.
+\end{enumerate}
+
+We found that 10 of Opticians 39 benchmarks required modifications to be
+bijective. We found that 4 of Optician's 39
+benchmarks required modifications to be in bijective correspondence when in
+canonical form.  Furthermore, we found that all of the data.gov examples
+required no modifications to be in canonical form.  While we did not put the
+data.gov examples into a form suitable for Optician, we found only 4 of
+the 8 data.gov examples did not require canonization, so making suitable
+for Optician would have required a modification of 4 of the data.gov
+formats.
+
+\subsection{Competitive Performance}
+
+While we expanded on the Optician system, we found it important to validate
+performance did not degrade in this expansion.  To validate this, we tested:
+\begin{itemize}
+\item[\OpticianRuntime{}] Optician runs on its own benchmarks.
+\item[\SystemOnOptician{}] \Name{} runs on Optician's benchmarks.
+\item[\SystemOnBenchmarks{}] \Name{} runs on its own benchmarks.
+\end{itemize}
+
+We summarize the results of these experiments in Figure~\ref{?}.  \Name{} was
+able to synthesize all of Optician's benchmarks at a speed competitive with
+Optician.  Furthermore, adopting these tasks to use QREs actually sped up
+tasks!  After canonization, the regular expressions sent to synthesis are
+\emph{smaller} than before canonization - making the bijective lens synthesis
+portion of these tasks easier.
+
 \section{Related Work}

### 2017-10-17 papers/QuotientLensSynth/main.tex
fix build breaks

@@
 \newcommand{\eqrel}[1]{\ensuremath{\equiv_{#1}}}
+
+\newcommand{\Name}{Optometrist}
+
+\newcommand{\OpticianRuntime}{OO}
+\newcommand{\SystemOnOptician}{SO}
+\newcommand{\SystemOnBenchmarks}{SS}
 
@@
 \begin{itemize}
-\item[\QRESize{}] \QRESize{} is the number of AST nodes in the QRE
+\item[QRESize{}] QRESize{} is the number of AST nodes in the QRE
   specifications involved in QRE lens synthesis task.
-\item[\CanonizerAndSpecSize{}] \CanonizerAndSpecSize{} is the number of
+\item[CanonizerAndSpecSize{}] CanonizerAndSpecSize{} is the number of
   AST nodes in $W(q)$ summed with the number of AST nodes in $\canonizer(q)$,
   for each QRE used in QRE lens synthesis task.
-\item[\LensAndSpecSize{}] \LensAndSpecSize{} is the number of AST nodes
+\item[LensAndSpecSize{}] LensAndSpecSize{} is the number of AST nodes
   in $W(q)$ summed with the number of AST nodes in the synthesized QRE lens.
@@
 
-The value of \QRESize{} represents the difficulty in writing our benchmarks.
-
-The value of \CanonizerAndSpecSize{} represents the difficulty in writing our
+The value of QRESize{} represents the difficulty in writing our benchmarks.
+
+The value of CanonizerAndSpecSize{} represents the difficulty in writing our
 benchmarks without QREs.  This is only an approximation, as both $W(q)$ and
@@
 
-The value of \LensAndSpecSize{} represents the difficulty in writing our benchmarks
+The value of LensAndSpecSize{} represents the difficulty in writing our benchmarks
 with neither QREs nor synthesis.  This is also an approximation, as $W(q)$ and
@@
 Figure~\ref{?} shows the number of benchmarks written with at most a given
-number of AST nodes.  We found that \QRESize{} was consistently smaller than
-\CanonizerAndSpecSize{} and \LensAndSpecSize{}, demonstrating that introducing
+number of AST nodes.  We found that QRESize{} was consistently smaller than
+CanonizerAndSpecSize{} and LensAndSpecSize{}, demonstrating that introducing
 QREs saves programmers programming effort compared to both Optician and basic

### 2017-10-17 papers/QuotientLensSynth/main.tex
Comments and a bit of working on the abstract (needs more)

@@
 
-  {\em Quotient lenses} express bidirectional transformations between data
-  formats, where the inputs to the transformation are considered modulo some
-  equivalence relation.  Such lenses save the programmer from having to write
-  two expressions and satisfy strong ``lens laws'' that ensure that the forward
-  and backward transformations interact well.  The use of quotients allows the
-  programmer to ignore ``inessential'' differences in the data, such as the
-  amount whitespace or the ordering of fields.
-
-  However, quotient lenses are often difficult to program.  The programmer
-  often has to give a specification for the data on each side of the lens and a
-  separate specification for the equivalence relations defined on the data. The
+{\em Lenses}~\cite{Lenses} are programs that specify bidirectional
+transformations between different data formats.  Writing a single lens
+rather than two separate uni-directional transformations saves both
+keystrokes and bugs, by reducing redundancy and by automatically
+guaranteeing strong ``lens laws'' ensuring that the forward and backward
+transformations interact well.
+%
+{\em Quotient lenses}~\cite{quotientlenses} are a generalization of lenses
+where the two data formats are considered modulo some equivalence relation.
+The use of quotients allows the programmer to ignore inessential differences
+in the data, such as the amount of whitespace or the ordering of fields.
+
+Quotient lenses can be difficult to program, requiring the programmer to
+give both a specification for the data on each side of the lens and a
+separate specification for the equivalence relations on the data.  The
 programmer must also manually implement a lens program that maps between the
-  specifications while respecting the equivalence relations---this is itself a
+formats while respecting the equivalence relations---this is itself a
 challenging task thanks to the restrictions on how lenses can be combined
-  while still satisfying the lens laws.
-
-This paper addresses both of these challenges. First, we propose the
-language of {\em Quotient Regular Expressions}, or QREs, which 
-define a broad class of equivalence relations
+while still satisfying the lens laws.  \bcp{Don't really understand that
+  paragraph!}
+
+This paper addresses both of these challenges\bcp{what are the two
+  challenges?}. First, we propose the language of {\em Quotient Regular
+  Expressions}, or QREs, which define a broad class of equivalence relations
 directly on regular languages using a compact and convenient notation. This
-eliminates the need to specify the data and the equivalence relation on the data
-separately. We then define {\em QRE lenses}, a class of quotient
-lenses that are typed using QREs, and show how to {\em synthesize} QRE lenses
-from their QRE types and a set of input-output examples. Finally, we
-demonstrate the applicability of this approach by synthesizing QRE lenses
-between several real-world data formats from the {\tt data.gov} database.
+eliminates the need to specify the data and the equivalence relation on the
+data separately. We then define {\em QRE lenses}, a class of quotient lenses
+that are typed using QREs, and show how to {\em synthesize} QRE lenses from
+their QRE types and a set of input-output examples. Finally, we demonstrate
+the applicability of this approach by synthesizing QRE lenses between
+several real-world data formats from the {\tt data.gov} database.
 \end{abstract}
@@
 of functions $(\ell.\get, \; \ell.\lput, \; \ell.\create)$ satisfying the
-following equations:
+following equations:\bcp{Why are we showing all the complexity of asymmetric
+lenses?  We should just say that we're considering only the bijective case
+for now and then just give the details of that.}
 \begin{align*}
@@
 
-\bcp{I thought we were going to allow newlines within titles and authors...?}
-
-\sam{I wanted to illustrate how to use the squash QRE which is why I changed my
-mind on that, but on second thought since the problem of quotienting away
-whitespace at will is going to be more familiar to the reader I changed it back
-to the original plan. I will explain squash in the QRE section.}
 \begin{verbatim}
@@
 The code sample below shows how the programmer might specify these equivalence
-relations on the \bibtex{} records in Boomerang using canonizers:
+relations on the \bibtex{} records in Boomerang using canonizers:\bcp{Too
+  big to display in-line.  Better to put it in a figure.}
 
@@
 
-With QREs we can specify \bibtex{} records as follows:
+With QREs we can specify \bibtex{} records as follows:\bcp{Again, belongs in
+a figure.  And, since both figures are long, maybe there is a visual way to
+highlight the differences...?}
 
@@
   \caption{Quotient Lens with ``Canonizers at the Edges''\sam{This is a
-  screenshot from Nate's thesis. Are we allowed to use it?}}
+      screenshot from Nate's thesis. Are we allowed to use it?}\bcp{We can
+      (if we put ``from [Nate's thesis]'' in the caption), but I don't think
+      we want the full complexity of asymmetric lenses here.  We should draw
+      the figure again.  (I might have the OmniGraffle file for it lying
+      around someplace...)}}
   \label{fig:attheedges}
@@
 
+\bcp{No bibliography??}
+
 \end{document}

### 2017-10-17 papers/QuotientLensSynth/main.tex
More comments

@@
 \section{Introduction}
+\bcp{Not sure there is time to fix this, but the intro seems very long and
+  technical---it takes a long time to get to saying what our actual
+  contributions are!  Wonder if we should move most of it to a second
+  ``Example'' section.} 
 {\em Bidirectional programming languages} such as

### 2017-10-17 papers/QuotientLensSynth/main.tex
add commands

--- 2017-10-17 papers/QuotientLensSynth/main.tex
@@
 \newcommand{\Name}{Optometrist}
+
+\newcommand{\QRESize}{QS}
+\newcommand{\CanonizerAndSpecSize}{CS}
+\newcommand{\LensAndSpecSize}{LS}
 
@@
 \begin{itemize}
-\item[QRESize{}] QRESize{} is the number of AST nodes in the QRE
+\item[\QRESize{}] \QRESize{} is the number of AST nodes in the QRE
   specifications involved in QRE lens synthesis task.
-\item[CanonizerAndSpecSize{}] CanonizerAndSpecSize{} is the number of
+\item[\CanonizerAndSpecSize{}] \CanonizerAndSpecSize{} is the number of
   AST nodes in $W(q)$ summed with the number of AST nodes in $\canonizer(q)$,
   for each QRE used in QRE lens synthesis task.
-\item[LensAndSpecSize{}] LensAndSpecSize{} is the number of AST nodes
+\item[\LensAndSpecSize{}] \LensAndSpecSize{} is the number of AST nodes
   in $W(q)$ summed with the number of AST nodes in the synthesized QRE lens.
@@
 
-The value of QRESize{} represents the difficulty in writing our benchmarks.
-
-The value of CanonizerAndSpecSize{} represents the difficulty in writing our
+The value of \QRESize{} represents the difficulty in writing our benchmarks.
+
+The value of \CanonizerAndSpecSize{} represents the difficulty in writing our
 benchmarks without QREs.  This is only an approximation, as both $W(q)$ and
@@
 
-The value of LensAndSpecSize{} represents the difficulty in writing our benchmarks
+The value of \LensAndSpecSize{} represents the difficulty in writing our benchmarks
 with neither QREs nor synthesis.  This is also an approximation, as $W(q)$ and
@@
 Figure~\ref{?} shows the number of benchmarks written with at most a given
-number of AST nodes.  We found that QRESize{} was consistently smaller than
-CanonizerAndSpecSize{} and LensAndSpecSize{}, demonstrating that introducing
+number of AST nodes.  We found that \QRESize{} was consistently smaller than
+\CanonizerAndSpecSize{} and \LensAndSpecSize{}, demonstrating that introducing
 QREs saves programmers programming effort compared to both Optician and basic

### 2017-10-18 papers/QuotientLensSynth/main.tex
some work on the intro to move the code into figures

@@
 array, mathtools, bussproofs, stmaryrd, enumitem, stackengine}
+\usepackage{relsize}
+\usepackage{listings}
+\lstset{
+  language=Caml,
+  basicstyle=\upshape\sffamily,
+  keywordstyle=\upshape\sffamily\color{dkblue},
+  keepspaces=true,
+  framexleftmargin=1ex,
+  framexrightmargin=1ex,
+  showstringspaces=false,
+  commentstyle=\itshape\rmfamily,
+  columns=flexible,
+  stringstyle=\sffamily\color{dkred},
+  mathescape
+}
+\newcommand{\cd}[1]{\lstinline[backgroundcolor=\color{white}]$#1$}
 
@@
 
-The code sample below shows how the programmer might specify these equivalence
-relations on the \bibtex{} records in Boomerang using canonizers:\bcp{Too
-  big to display in-line.  Better to put it in a figure.}
-
-\begin{verbatim}
+The code sample in Figure~\ref{fig:example-canonizers} shows how the programmer
+might specify these equivalence relations on the \bibtex{} records in Boomerang
+using canonizers.
+
+\begin{figure}[t]
+
+\begin{lstlisting}
 let project (r : regexp) (s : string) = canonizer_of_lens ((del r) . (ins s))
-
-let wsp : regexp = [ \n\t\r]+
-
+let wsp : regexp = [ \n\t\r]$^+$
 let wsp_to_space : canonizer = project wsp " "
-
 let valid_name : canonizer = 
-name | (name . wsp_to_space . (name .wsp_to_space)* .name)
-
+    name | (name . wsp_to_space . (name .wsp_to_space)$^*$ .name)
 let wsp_to_and : canonizer = project (wsp . "and" . wsp) " and "
-
 let bib_names : canonizer =  
-
-valid_name | (valid_name . wsp_to_and . (valid_name . wsp_to_and)* . valid_name)
-
+    valid_name | (valid_name . wsp_to_and . (valid_name . wsp_to_and)$^*$ . valid_name)
 let bib_author : canonizer = "author = \"" . bib_names . "\""
-
-let label : regexp = [a-zA-Z0-9]+
-
+let label : regexp = [a-zA-Z0-9]$^+$
 let title : canonizer = 
-word | (word . wsp_to_space . (word . wsp_to_space)* . word)
-
+    word | (word . wsp_to_space . (word . wsp_to_space)$^*$ . word)
 let bib_title : canonizer = "title = {" . title . "}"
-
 let wsp_to_comma : canonizer = project ("," . wsp) ",\n"
-
 let first_perm : canonizer = bib_author . wsp_to_comma . bib_title
-
 let second_perm : canonizer = bib_title . wsp_to_comma . bib_author
-
 let bibtex : canonizer = 
     "@book{" . label. wsp_to_comma . (first_perm | second_perm) . "}"
-\end{verbatim}
-
-Using the in-built \verb!canonizer_of_lens! function which extracts a canonizer
-from a lens, it is easy to define a an equivalence relation that sends all
+\end{lstlisting}
+
+  \caption{Boomerang canonizers for the \bibtex{}--EndnNote equivalence
+    relations.}
+  \label{fig:example-canonizers}
+\end{figure}
+
+Using the in-built \cd{canonizer_of_lens} function which extracts a canonizer
+from a lens, it is easy to define an equivalence relation that sends all
 members of a regular language to a canonical representative. For example
 the expression 
-\begin{verbatim}
+\begin{lstlisting}
 let wsp_to_space : canonizer = canonizer_of_lens ((del wsp) . (ins " "))
-\end{verbatim}
-
+\end{lstlisting}
 \noindent defines a canonizer which sends arbitrary occurences of whitespace
@@
 
-\begin{verbatim}
+\begin{lstlisting}
 let first_perm : canonizer = bib_author . wsp_to_comma . bib_title
-
 let second_perm : canonizer = bib_title . wsp_to_comma . bib_author
-\end{verbatim}
+\end{lstlisting}
 
 One possible solution is to permute the alternative canonizers
-
-\noindent \verb!bib_author. wsp_to_comma! and \verb!bib_title . wsp_to_comma!,
+\cd{bib_author.wsp_to_comma} and \cd{bib_title.wsp_to_comma},
 but Boomerang's type-checker will reject this expression since the permutation
@@
 regular expression
-\begin{verbatim}
+\begin{lstlisting}
 bib_author . wsp_to_comma . bib_title . wsp_to_comma | 
 bib_title . wsp_to_comma . bib_author . wsp_to_comma
-\end{verbatim}
+\end{lstlisting}
 is not equivalent to,
-\begin{verbatim}
+\begin{lstlisting}
 bib_author . wsp_to_comma . bib_title  | 
 bib_title . wsp_to_comma . bib_author
-\end{verbatim}
+\end{lstlisting}
 thus the programmer would have to write more lenses and canonizers to correct
@@
 specify all the regular expressions first (including the whole regular
-expressions canonized by the \verb!bib_author! and \verb!bib_title!
+expressions canonized by the \cd{bib_author} and \cd{bib_title}
 canonizers), and then the canonizers separately. Alternatively, he/she may stick
 to using canonizers and then liberally make use of the built-in
-functions \verb!uncanonized_type! and \verb!canonized_type! which return the
+functions \cd{uncanonized_type} and \cd{canonized_type} which return the
 regular expression matching all the data and the regular expression matching
@@
 
-With QREs we can specify \bibtex{} records as follows:\bcp{Again, belongs in
-a figure.  And, since both figures are long, maybe there is a visual way to
-highlight the differences...?}
-
-\begin{verbatim}
-let wsp_to_space : canonizer = project [ \n\t\r]+ -> " "
-
+With QREs we can specify \bibtex{} records as shown in Figure~\ref{fig:example-qre}.
+
+\begin{figure}[t]
+\begin{lstlisting}
+let wsp_to_space : canonizer = project [ \n\t\r]$^+$ $\to$ " "
 let valid_name : canonizer = 
-name | (name . wsp_to_space . (name . wsp_to_space)* . name)
-
+   name | (name . wsp_to_space . (name . wsp_to_space)$^*$ . name)
 let wsp_to_and : canonizer = 
-project ([ \n\t\r]+ . "and" . [ \n\t\r]+) -> " and "
-
+   project ([ \n\t\r]$^+$ . "and" . [ \n\t\r]$^+$) $\to$ " and "
 let bib_names : canonizer =  
-
-valid_name | (valid_name . wsp_to_and . (valid_name . wsp_to_and)* . valid_name)
-
+   valid_name | (valid_name . wsp_to_and . (valid_name . wsp_to_and)$^*$ . valid_name)
 let bib_author : canonizer = "author = \"" . bib_names . "\""
-
-let label : regexp = [a-zA-Z0-9]+
-
-let title : canonizer = word | (word . wsp_to_space . (word . wsp_to_space)* .
-word)
-
+let label : regexp = [a-zA-Z0-9]$^+$
+let title : canonizer = word | (word . wsp_to_space . (word . wsp_to_space)$^*$ . word)
 let bib_title : canonizer = "title = {" . title . "}"
-
 let bib_perm  : canonizer = perm (label, bib_author, bib_title) with
-(project ("," . [ \n\t\r]* ) -> ",\n")
-
+   (project ("," . [ \n\t\r]$^*$ ) $\to$ ",\n")
 let bibtex : canonizer = "@book{" . bib_perm . "}"
-\end{verbatim}
-
-The expression \verb!project [ \n\t\r]+ -> " "! is a QRE which matches any
+\end{lstlisting}
+
+  \caption{QRE specification of \bibtex{} records.}
+  \label{fig:example-qre}
+\end{figure}
+
+The expression \cd{project [ \n\t\r]+ -> " "} is a QRE that matches any
 number of whitespace characters, and then considers all strings of whitespace
@@
 \begin{center}
-\begin{verbatim}
-perm (label, bib_author, bib_title) with (project ("," . [\n\t\r]* ) -> ",\n")
-\end{verbatim}
+\begin{lstlisting}
+perm (label, bib_author, bib_title) with (project ("," . [\n\t\r]$^*$ ) $\to$ ",\n")
+\end{lstlisting}
 \end{center}
@@
 and title fields occur in that order, separated by a comma followed a newline.
-The \verb!perm! QRE is defined at the level of canonizers which allows for
+The \cd{perm} QRE is defined at the level of canonizers which allows for
 easy nesting of permutations. Similarly, it is easy to specify any number of
-fields within the \verb!perm! QRE.
+fields within the \cd{perm} QRE.
 
@@
 which is itself tedious and error-prone as the code snippet below illustrates:
-\begin{verbatim}
-let preamble : (lens in "@book{" <=> "%0 Book\n%F ") = 
+
+\begin{lstlisting}
+let preamble : (lens in "@book{" $\Leftrightarrow$ "%0 Book\n%F ") = 
     del "@book{" . ins "%0 Book\n%F "
 
-let author_lens : (lens in (",\n" . bib_author) <=> ("\n" . end_author)) = 
+let author_lens : (lens in (",\n" . bib_author) $\Leftrightarrow$ ("\n" . end_author)) = 
     del ",\nauthor = \""
@@
          . ins "\n"
-         . (ins "%A " . valid_name . del " and " . ins "\n")*  . ins "%A "
+         . (ins "%A " . valid_name . del " and " . ins "\n")$^*$ . ins "%A "
          . valid_name
@@
 
-let title_lens : 
-(lens in ("\",\n" . bib_title . "},\n}") <=> ("\n" . end_title)) =
+let title_lens : (lens in ("\",\n" . bib_title . "},\n}") $\Leftrightarrow$ ("\n" . end_title)) =
    del "\",\ntitle = {" . ins "\n%T " . title . del "},\n}"
 
-let bib_to_end : (lens in bib_tex <=> end_note) =
+let bib_to_end : (lens in bib_tex $\Leftrightarrow$ end_note) =
    preamble . label . author_lens . title_lens
-\end{verbatim}
+\end{lstlisting}
 
@@
 matching the delimiters and separators in a bijective manner. For example,
-the \verb!preamble! lens replaces the \bibtex{} preamble \verb!@book! with the
-EndNote preamble \verb!%0 Book\n%F ! in the \bibtex{record}. The programmer
+the \cd{preamble} lens replaces the \bibtex{} preamble \cd{@book} with the
+EndNote preamble \cd{\%0 Book\n\%F } in the \bibtex{record}. The programmer
 can ill afford to make a mistake with any one of the small but vital details
@@
 
-We have already encountered the \verb!project R -> s! QRE which sends all
+We have already encountered the \verb!project R $\to$ s! QRE which sends all
 elements of the language described by the regular expression $R$, and the
 
-\noindent \verb!perm (q_1, ..., q_n) with q! QRE which sends all permutations of
+\noindent \cd{perm (q_1, ..., q_n) with q} QRE which sends all permutations of
 the QREs $q_1, \ldots, q_n$ separated by the QRE $q$ to the canonical
@@
 let squash_author : canonizer =  
-squash space_name -> comma_name using (get comma_to_space)
+squash space_name $\to$ comma_name using (get comma_to_space)
 
@@
 \begin{verbatim}
-let ambiguous : canonizer = "a"* . (project "a"* -> "a")
+let ambiguous : canonizer = "a"* . (project "a"* $\to$ "a")
 \end{verbatim}

### 2017-10-18 papers/QuotientLensSynth/main.tex
update performance, todo add graph

--- 2017-10-17 papers/QuotientLensSynth/main.tex
@@
 able to synthesize all of Optician's benchmarks at a speed competitive with
-Optician.  Furthermore, adopting these tasks to use QREs actually sped up
-tasks!  After canonization, the regular expressions sent to synthesis are
-\emph{smaller} than before canonization - making the bijective lens synthesis
-portion of these tasks easier.
+Optician.  There was additional overhead in calculating $W(q)$ and $K(q)$,
+resulting in a very slight decrease in performance.
+
+Also notable was one of the data.gov examples, which converted demographic
+statistics by zip code from xml form into rdf form.  This merely took longer because it was
+an exceptionally large format, converting over 40 fields.  After adding in additional
+fields to make this format bijective, it still took over 15 seconds to complete.
 

### 2017-10-18 papers/QuotientLensSynth/main.tex
Note
Little edits and comments here and there

@@
   columns=flexible,
-  stringstyle=\sffamily\color{dkred},
+  % BCP: I find this distracting:
+  % stringstyle=\sffamily\color{dkred},
   mathescape
@@
       we want the full complexity of asymmetric lenses here.  We should draw
-      the figure again.  (I might have the OmniGraffle file for it lying
-      around someplace...)}}
+      the figure again.  (I have the OmniGraffle file for it, so this is
+      easy if I know how it should look.)}}
   \label{fig:attheedges}
@@
 class of equivalence relations on regular languages (Section~\ref{QRE}).
-\item We define a language of {\em QRE lenses}, a form of quotient lenses
-whose types are given by QREs.  Our main technical contribution is a normal
-form for QRE lenses and a proof that QRE lenses in normal form are closed
-under left quotienting, right quotienting, composition, and regular
+\item We define a language of {\em QRE lenses}, a subset of bijective quotient
+lenses whose types are given by QREs.  Our main technical contribution is a
+normal form for QRE lenses and a proof that normal forms are
+closed under left quotienting, right quotienting, composition, and regular
 operators (Section~\ref{QRE-lenses}).
@@
 equivalence relation on the language described by the regular expression.
+
+\subsection{Overview}
 
@@
 \]
-  \caption{Whole and Kernel Regular Expressions\bcp{left-justify columns,
-      and try putting the last two clauses into the same format as the
+  \caption{Whole and Kernel Regular Expressions\bcp{try putting the 
+      last two clauses into the same format as the
       others (using $\mathit{qq} = ...$ and maybe reducing spacing around
@@
   $$
-  \caption{Denotation and Typing Rules for QRE Lenses}
+  \caption{Denotation and Typing Rules for QRE Lenses\bcp{I think we can
+      display this more compactly if we make the inference rules narrower by
+    stacking the premises vertically and then put each inference rule {\em
+      beside} its semantics definition.}}
   \label{fig:qlenssemantics}

### 2017-10-18 papers/QuotientLensSynth/main.tex
Another rewrite of the abstract

@@
 \newif\ifdraft\drafttrue  % set true to show comments
+\newif\ifplentyoftime\plentyoftimefalse  % :-)
 
@@
 
-{\em Lenses}~\cite{Lenses} are programs that specify bidirectional
+{\em Lenses}~\cite{Lenses} specify bidirectional 
 transformations between different data formats.  Writing a single lens
@@
 {\em Quotient lenses}~\cite{quotientlenses} are a generalization of lenses
-where the two data formats are considered modulo some equivalence relation.
-The use of quotients allows the programmer to ignore inessential differences
-in the data, such as the amount of whitespace or the ordering of fields.
-
-Quotient lenses can be difficult to program, requiring the programmer to
-give both a specification for the data on each side of the lens and a
-separate specification for the equivalence relations on the data.  The
-programmer must also manually implement a lens program that maps between the
-formats while respecting the equivalence relations---this is itself a
-challenging task thanks to the restrictions on how lenses can be combined
-while still satisfying the lens laws.  \bcp{Don't really understand that
-  paragraph!}
-
-This paper addresses both of these challenges\bcp{what are the two
-  challenges?}. First, we propose the language of {\em Quotient Regular
-  Expressions}, or QREs, which define a broad class of equivalence relations
-directly on regular languages using a compact and convenient notation. This
-eliminates the need to specify the data and the equivalence relation on the
-data separately. We then define {\em QRE lenses}, a class of quotient lenses
-that are typed using QREs, and show how to {\em synthesize} QRE lenses from
-their QRE types and a set of input-output examples. Finally, we demonstrate
-the applicability of this approach by synthesizing QRE lenses between
-several real-world data formats from the {\tt data.gov} database.
+where each of the data formats is quotiented by some equivalence relation
+and the lens laws are taken modulo equivalence.  This allows lenses to be
+both simpler and more flexible, since whole equivalence classes of data
+values can be handled all at once and inessential details like whitespace
+and ordering of fields can be ignored.
+
+\smallskip
+
+However, while plain lenses are widely used in both standalone systems and
+popular libraries, Boomerang~\cite{Boomerang} is the only implementation
+currently offering quotient lenses.  One reason for this is that there is
+some annoying redundancy in the quotient lens notation: The programmer must
+separately specify (1) the data formats on each side of the lens, (2) the
+equivalence relations over them, and (3) a lens between the canonical
+representatives of the equivalence classes, even though, typically, these
+three share a great deal of structure.  Moreover, writing the lens between
+the equivalence classes can be challenging due to the rather fiddly, mainly
+due to ``non-ambiguity'' conditions in the lens well-formedness rules.
+
+\smallskip
+
+In this paper, we address both of these issues.  First, we propose a novel
+language of {\em Quotient Regular Expressions}, or QREs, which allows
+programmers to specify a broad class of equivalence relations over regular
+languages using a compact and convenient notation in which the regular
+language and its equivalence classes are captured by a single expression.
+We then define {\em QRE lenses}, a class of bijective quotient lenses that
+are typed using QREs, and show how to {\em synthesize} QRE lenses from their
+QRE types and a set of input-output examples.  We demonstrate the
+practicality of this approach by synthesizing QRE lenses between several
+real-world data formats from the {\tt data.gov} database.
 \end{abstract}
@@
 \section{Introduction}
+\ifplentyoftime
 \bcp{Not sure there is time to fix this, but the intro seems very long and
@@
   ``Example'' section.} 
+\fi
 {\em Bidirectional programming languages} such as
 Boomerang~\cite{boomerang}, Augeas~\cite{augeas}, XSugar~\cite{xsugar},
-biXid~\cite{bixid} and X/Inv~\cite{xinv}enable a programmer to express a
+biXid~\cite{bixid} and X/Inv~\cite{xinv}\bcp{there are many others; we
+  should mention all of them} enable a programmer to express a
 transformation that maps some input data to some output data, as well as a

### 2017-10-18 papers/QuotientLensSynth/main.tex
Checkpoint

@@
 We then define {\em QRE lenses}, a class of bijective quotient lenses that
-are typed using QREs, and show how to {\em synthesize} QRE lenses from their
+are typed using QREs, and we show how to {\em synthesize} QRE lenses from their
 QRE types and a set of input-output examples.  We demonstrate the
@@
 \fi
-{\em Bidirectional programming languages} such as
+\bcp{This sentence could be clearer!}{\em Bidirectional programming languages} such as
 Boomerang~\cite{boomerang}, Augeas~\cite{augeas}, XSugar~\cite{xsugar},
@@
 lenses?  We should just say that we're considering only the bijective case
-for now and then just give the details of that.}
+for now and then just give the details of that.  (Related: The
+``abstract/concrete'' terminology doesn't make much sense in the bijective
+setting.)} 
 \begin{align*}

### 2017-10-18 papers/QuotientLensSynth/main.tex
Pass through the intro

@@
 
+\bcp{I'm nervous about making the concept of canonizer so central, since it
+  is not going to be much of a focus for the rest of the paper.  In general,
+it feels like the intro is explaining too much.  We need to find a more
+streamlined path through the ideas!}
 In Boomerang, equivalence relations are expressed using {\em canonizers}
@@
 regular expression which matches all possible representations of the data, and
-then provide a set of canonizers which does the job of quotienting the data.
+then provide a canonizer that does the job of quotienting the data.
 
@@
 each record, and allows for arbitrary amounts of whitespace in between the
-author and title fields. These rules allow for these two \bibtex{} citation
-records to be considered the same,
-
+author and title fields. These rules allow for the \bibtex{} citations
 \begin{verbatim}
@@
     }
-
-@Book{conway,
-
+\end{verbatim}
+and
+\begin{verbatim}
+    @Book{     conway,
         Title = {Regular Algebra 
                  and Finite Machines},
-
-    Author = "J. H. Conway",
-}
+           Author = "J. H. Conway",     }
 \end{verbatim}
-
-\noindent with the EndNote records having a similar equivalence relation defined
-on them.
-
-The code sample in Figure~\ref{fig:example-canonizers} shows how the programmer
-might specify these equivalence relations on the \bibtex{} records in Boomerang
-using canonizers.
+to be considered the same (with the EndNote records having a similar
+equivalence relation defined on them).
 
 \begin{figure}[t]
-
 \begin{lstlisting}
@@
 
-  \caption{Boomerang canonizers for the \bibtex{}--EndnNote equivalence
+  \caption{Boomerang canonizers for the \bibtex{}--EndNote equivalence
     relations.}
@@
 \end{figure}
-
+The code sample in Figure~\ref{fig:example-canonizers} shows how the programmer
+might specify these equivalence relations on the \bibtex{} records in Boomerang
+using canonizers.
+%
 Using the in-built \cd{canonizer_of_lens} function which extracts a canonizer
 from a lens, it is easy to define an equivalence relation that sends all
-members of a regular language to a canonical representative. For example
+members of a regular language to a canonical representative.\bcp{Oh dear.
+  The \cd{canonizer_of_lens} thing is really kind of a hack---we're using
+  the fancy/specialized language of lenses to write what could really be a
+  simple program in an arbitrary language (and then throwing away half of the
+  lens, when the whole reason we give for having lenses in the first place
+  is because we want both directions).  As an implementation trick, it was a
+good idea.  But trying to get readers to understand it here does not seem
+like a good idea.} For example
 the expression 
@@
 \noindent defines a canonizer which sends arbitrary occurences of whitespace
-characters to a single space. However, a problem arises when trying to specify
+characters to a single space\bcp{Going into this level of detail encourages
+readers to actually try to understand the details, but we don't actually
+give enough details to allow them to do that!}. However, a problem arises
+when trying to specify 
 the equivalence relation which nests this projection relation with the relation
@@
 seperately, since Boomerang's in-built permutation combinator does not account
-for separators in between the regular expressions being permuted:
+for separators in between the regular expressions being permuted:\bcp{Why
+  couldn't we just enrich Boomerang's permutation combinator to do this?}
 
@@
 but Boomerang's type-checker will reject this expression since the permutation
-combinator accepts regular expressions rather than canonizers. In any case, the
+combinator accepts regular expressions rather than canonizers\bcp{It feels
+  like we're lost, here, in details of exactly what Boomerang does and
+  doesn't do.  This doesn't seem especially relevant to quotient lenses in
+  general, or to what we're doing in this paper.}. In any case, the
 regular expression
@@
 
-  \caption{QRE specification of \bibtex{} records.}
+\caption{QRE specification of \bibtex{} records.  \bcp{I'm afraid this pair
+    of examples is not making the point we intended.  I thought we were
+    hoping to just say ``Here's the old Boomerang version; here's the QRE
+    version; look how much smaller and simpler the second one is.''  But the
+    second one {\em isn't} actually much smaller and, at least on first
+    impression, doesn't look simpler.}}
   \label{fig:example-qre}
@@
 
-The expression \cd{project [ \n\t\r]+ -> " "} is a QRE that matches any
+The expression \bcp{mangled---did you proofread this?}\cd{project [
+  \n\t\r]+ -> " "} is a QRE that matches any 
 number of whitespace characters, and then considers all strings of whitespace
@@
 While QREs kill two birds with one stone by specifying the
-data and the equivalence relation on the data in one expression, he/she
+data and the equivalence relation on the data in one expression,
+he/she\bcp{Awkward: just use ``she'' (or find a way to alternate pronouns,
+  if that doesn't create confusion)}
 still has to manually program the underlying lens which maps the equivalence
 classes of the source data to the equivalence classes of the view data, a task
-which is itself tedious and error-prone as the code snippet below illustrates:
+which is itself tedious and error-prone as the code snippet below
+illustrates:\bcp{If we added this to the old-Boomerang figure it might make
+  a better case for size and simplicity!}
 
@@
 problem of synthesizing {\em bijective lenses}, a particular subset of the
-lenses that can be expressed in Boomerang. This is a problem which we solved in
-our prior work ~\cite{popl18}.
+lenses that can be expressed in Boomerang.\bcp{As stated, this is
+  misleading: the reader might have the impression that we somehow reduce
+  the problem of synthesizing {\em asymmetric} quotient lenses to the
+  problem of synthesizing bijective ordinary lenses!} This is a problem
+which we solved in our prior work ~\cite{popl18}.
 
@@
 proof that QRE lenses are closed under left quotienting, right quotienting,
-composition, and regular combinators. This proof will allow justify our
+composition, and regular combinators.\bcp{If this is our main technical
+  contribution, then it deserves more than a sentence in the intro!  E.g.,
+  can we give an example?} This proof will allow justify our
 approach for synthesizing quotient lenses so that the programmer can simply
@@
 
-We have already encountered the \verb!project R $\to$ s! QRE which sends all
-elements of the language described by the regular expression $R$, and the
+We have already encountered the \bcp{Again, formatting is mangled.  Also,
+  are we setting lens expressions in typewriter font, or in snas-serif?  I
+  don't care, but we need to be consistent.}\verb!project R $\to$ s! QRE
+which sends all elements of the language described by the regular expression
+$R$, and the
 

### 2017-10-18 papers/QuotientLensSynth/main.tex
add new graphs and fix some text

--- 2017-10-18 papers/QuotientLensSynth/main.tex
@@
 \newif\ifdraft\drafttrue  % set true to show comments
-\newif\ifplentyoftime\plentyoftimefalse  % :-)
 
@@
 \usepackage{amsmath, amssymb, verbatim, enumerate, graphicx, centernot, tikz,
-array, mathtools, bussproofs, stmaryrd, enumitem, stackengine}
+  array, mathtools, bussproofs, stmaryrd, enumitem, stackengine, subcaption}
+\captionsetup{compatibility=false}
 \usepackage{relsize}
@@
   columns=flexible,
-  % BCP: I find this distracting:
-  % stringstyle=\sffamily\color{dkred},
+  stringstyle=\sffamily\color{dkred},
   mathescape
@@
 
-\newcommand{\OpticianRuntime}{OO}
-\newcommand{\SystemOnOptician}{SO}
-\newcommand{\SystemOnBenchmarks}{SS}
+\newcommand{\OpticianRuntime}{\textbf{OO}}
+\newcommand{\SystemOnOptician}{\textbf{SO}}
+\newcommand{\SystemOnBenchmarks}{\textbf{SS}}
 
@@
 
-{\em Lenses}~\cite{Lenses} specify bidirectional 
+{\em Lenses}~\cite{Lenses} are programs that specify bidirectional
 transformations between different data formats.  Writing a single lens
@@
 {\em Quotient lenses}~\cite{quotientlenses} are a generalization of lenses
-where each of the data formats is quotiented by some equivalence relation
-and the lens laws are taken modulo equivalence.  This allows lenses to be
-both simpler and more flexible, since whole equivalence classes of data
-values can be handled all at once and inessential details like whitespace
-and ordering of fields can be ignored.
-
-\smallskip
-
-However, while plain lenses are widely used in both standalone systems and
-popular libraries, Boomerang~\cite{Boomerang} is the only implementation
-currently offering quotient lenses.  One reason for this is that there is
-some annoying redundancy in the quotient lens notation: The programmer must
-separately specify (1) the data formats on each side of the lens, (2) the
-equivalence relations over them, and (3) a lens between the canonical
-representatives of the equivalence classes, even though, typically, these
-three share a great deal of structure.  Moreover, writing the lens between
-the equivalence classes can be challenging due to the rather fiddly, mainly
-due to ``non-ambiguity'' conditions in the lens well-formedness rules.
-
-\smallskip
-
-In this paper, we address both of these issues.  First, we propose a novel
-language of {\em Quotient Regular Expressions}, or QREs, which allows
-programmers to specify a broad class of equivalence relations over regular
-languages using a compact and convenient notation in which the regular
-language and its equivalence classes are captured by a single expression.
-We then define {\em QRE lenses}, a class of bijective quotient lenses that
-are typed using QREs, and show how to {\em synthesize} QRE lenses from their
-QRE types and a set of input-output examples.  We demonstrate the
-practicality of this approach by synthesizing QRE lenses between several
-real-world data formats from the {\tt data.gov} database.
+where the two data formats are considered modulo some equivalence relation.
+The use of quotients allows the programmer to ignore inessential differences
+in the data, such as the amount of whitespace or the ordering of fields.
+
+Quotient lenses can be difficult to program, requiring the programmer to
+give both a specification for the data on each side of the lens and a
+separate specification for the equivalence relations on the data.  The
+programmer must also manually implement a lens program that maps between the
+formats while respecting the equivalence relations---this is itself a
+challenging task thanks to the restrictions on how lenses can be combined
+while still satisfying the lens laws.  \bcp{Don't really understand that
+  paragraph!}
+
+This paper addresses both of these challenges\bcp{what are the two
+  challenges?}. First, we propose the language of {\em Quotient Regular
+  Expressions}, or QREs, which define a broad class of equivalence relations
+directly on regular languages using a compact and convenient notation. This
+eliminates the need to specify the data and the equivalence relation on the
+data separately. We then define {\em QRE lenses}, a class of quotient lenses
+that are typed using QREs, and show how to {\em synthesize} QRE lenses from
+their QRE types and a set of input-output examples. Finally, we demonstrate
+the applicability of this approach by synthesizing QRE lenses between
+several real-world data formats from the {\tt data.gov} database.
 \end{abstract}
@@
 \section{Introduction}
-\ifplentyoftime
 \bcp{Not sure there is time to fix this, but the intro seems very long and
@@
   ``Example'' section.} 
-\fi
 {\em Bidirectional programming languages} such as
 Boomerang~\cite{boomerang}, Augeas~\cite{augeas}, XSugar~\cite{xsugar},
-biXid~\cite{bixid} and X/Inv~\cite{xinv}\bcp{there are many others; we
-  should mention all of them} enable a programmer to express a
+biXid~\cite{bixid} and X/Inv~\cite{xinv}enable a programmer to express a
 transformation that maps some input data to some output data, as well as a
@@
       we want the full complexity of asymmetric lenses here.  We should draw
-      the figure again.  (I have the OmniGraffle file for it, so this is
-      easy if I know how it should look.)}}
+      the figure again.  (I might have the OmniGraffle file for it lying
+      around someplace...)}}
   \label{fig:attheedges}
@@
 class of equivalence relations on regular languages (Section~\ref{QRE}).
-\item We define a language of {\em QRE lenses}, a subset of bijective quotient
-lenses whose types are given by QREs.  Our main technical contribution is a
-normal form for QRE lenses and a proof that normal forms are
-closed under left quotienting, right quotienting, composition, and regular
+\item We define a language of {\em QRE lenses}, a form of quotient lenses
+whose types are given by QREs.  Our main technical contribution is a normal
+form for QRE lenses and a proof that QRE lenses in normal form are closed
+under left quotienting, right quotienting, composition, and regular
 operators (Section~\ref{QRE-lenses}).
@@
 equivalence relation on the language described by the regular expression.
-
-\subsection{Overview}
 
@@
 \]
-  \caption{Whole and Kernel Regular Expressions\bcp{try putting the 
-      last two clauses into the same format as the
+  \caption{Whole and Kernel Regular Expressions\bcp{left-justify columns,
+      and try putting the last two clauses into the same format as the
       others (using $\mathit{qq} = ...$ and maybe reducing spacing around
@@
   $$
-  \caption{Denotation and Typing Rules for QRE Lenses\bcp{I think we can
-      display this more compactly if we make the inference rules narrower by
-    stacking the premises vertically and then put each inference rule {\em
-      beside} its semantics definition.}}
+  \caption{Denotation and Typing Rules for QRE Lenses}
   \label{fig:qlenssemantics}
@@
 
-We have integrated QREs and quotient lens synthesis into Boomerang with an
-additional \afm{loc} lines of OCaml code.
+We have integrated QREs and quotient lens synthesis into Boomerang allowing QREs
+and synthesis tasks to be expressed alongside existing Boomerang expressions.
 
 All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
-of 1600 MHz DDR3 running macOS Sierra.
+of 1600 MHz DDR3 running macOS High Sierra.
 
@@
 \begin{enumerate}
-\item \afm{39} of our benchmarks were derived from the Optician system.  These
+\item 39 of our benchmarks were derived from the Optician system.  These
   benchmarks were a combination of custom benchmarks, benchmarks derived from
@@
 
-\item \afm{8} of our benchmarks were derived from data.gov.  Data.gov contains
+\item 8 of our benchmarks were derived from data.gov.  Data.gov contains
   many transformations between formats.  In addition to the formats previously
@@
 
+\begin{figure}
+  \centering
+  \begin{subfigure}[b]{.49\textwidth}
+    \centering
+    \includegraphics{generated-graphs/times_opt}
+    \caption{}
+    \label{subfig:lenssize}
+  \end{subfigure}
+  \begin{subfigure}[b]{.49\textwidth}
+    \includegraphics{generated-graphs/times_new}
+    \caption{}
+    \label{subfig:examplesused}
+  \end{subfigure}
+  \caption{Runtimes of Optician and \Name{}.
+    In (a), we run Optician and \Name{} on Optician's benchmarks.  We find that
+    there is only a negligible difference incurred by Optometrist's overhead.
+    In (b), we run Optometrist on its benchmark suite.  We find that it is able
+    to synthesize all quotient lenses in under 20 seconds, and typically
+    finished in under 5 seconds. }
+  \label{fig:times}
+\end{figure}
+
 While we expanded on the Optician system, we found it important to validate
@@
 
-We summarize the results of these experiments in Figure~\ref{?}.  \Name{} was
+We summarize the results of these experiments in Figure~\ref{fig:times}.
+\Name{} was
 able to synthesize all of Optician's benchmarks at a speed competitive with

### 2017-10-18 papers/QuotientLensSynth/main.tex
Fix figure 3.
fixed the formatting of lens expressions

@@
 
-The expression \bcp{mangled---did you proofread this?}\cd{project [
-  \n\t\r]+ -> " "} is a QRE that matches any 
+The expression \cd{project [\n\t\r]}$^+$ $\to$ \cd{" "} is a QRE that matches any 
 number of whitespace characters, and then considers all strings of whitespace
@@
   \centering
-  \includegraphics[width=\textwidth]{attheedges}
+  \includegraphics[width=\textwidth]{canonizers-outside}
   \caption{Quotient Lens with ``Canonizers at the Edges''\sam{This is a
@@
       the figure again.  (I have the OmniGraffle file for it, so this is
-      easy if I know how it should look.)}}
+      easy if I know how it should look.)}\bcp{Done---what do you think?}}
   \label{fig:attheedges}
@@
 
-\begin{verbatim}
+\begin{lstlisting}
 let bib_to_end : (lens in bibtex <=> endnot) 
 = synth bibtex <=> endnote  using  {(bib_example, end_example)}
-\end{verbatim}
-\noindent to synthesize a quotient lens that maps between \verb!bibtex! and
-\verb!endnote! QREs, and sends the equivalence class of \verb!bib_bexample! to
-the equivalence class of \verb!end_example!.
+\end{lstlisting}
+\noindent to synthesize a quotient lens that maps between \cd{bibtex} and
+\cd{endnote} QREs, and sends the equivalence class of \cd{bib_bexample} to
+the equivalence class of \cd{end_example}.
 
@@
 
-We have already encountered the \bcp{Again, formatting is mangled.  Also,
-  are we setting lens expressions in typewriter font, or in snas-serif?  I
-  don't care, but we need to be consistent.}\verb!project R $\to$ s! QRE
+We have already encountered the \cd{project R} $\to$ \cd{s} QRE
 which sends all elements of the language described by the regular expression
@@
 The equivalence described is one in which there is a bijection between 
-the regular expressions \verb!comma_name! and \verb!space_name! given by,
-\begin{verbatim}
+the regular expressions \cd{comma_name} and \cd{space_name} given by,
+\begin{lstlisting}
 let comma_name : regexp = name | (name . ", " . name . (" " . name)*)
@@
 
-\end{verbatim}
-
-\noindent with the lens \verb!comma_to_space! mapping between them given by 
-\begin{verbatim}
+\end{lstlisting}
+
+\noindent with the lens \cd{comma_to_space} mapping between them given by 
+\begin{lstlisting}
 let comma_to_space : (lens in space_name <=> comma_name) =
@@
 (name . del " " . (ins " " . name . del " ")* ) (name . ins ", ") || name 
-\end{verbatim}
+\end{lstlisting}
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}
 let squash_author : canonizer =  
@@
 let author : canonizer = "author = \"" . squash_author . "\""
-\end{verbatim}
-
-In this example, the \verb!squash_author! QRE expresses the fact that the
-\verb!space_name! and \verb!comma_name! representations of the authors' name
-are equivalent, with the \verb!comma_name! representation chosen as the
-canonical representation, and the function \verb!comma_to_space.get! giving
-the canonizing function which converts data in the \verb!space_name! format
-to its \verb!comma_name! representation.
+\end{lstlisting}
+
+In this example, the \cd{squash_author} QRE expresses the fact that the
+\cd{space_name} and \cd{comma_name} representations of the authors' name
+are equivalent, with the \cd{comma_name} representation chosen as the
+canonical representation, and the function \cd{comma_to_space.get} giving
+the canonizing function which converts data in the \cd{space_name} format
+to its \cd{comma_name} representation.
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}
 let comma_to_space : (lens in space_name <=> comma_name) = 
 synth space_name <=> comma_name using {("J. H. Conway", "Conway, J. H.")}
-\end{verbatim}
+\end{lstlisting}
 
@@
 instance, if the QRE for matching a \bibtex{} record is named
-\verb!bibtex!, then the QRE for matching two records is 
-\verb!bibtex . bibtex!, the QRE for matching one or two records is
-\verb! bibtex | (bibtex . bibtex)! and the QRE for matching any number of QRE
-records \verb!bibtex*!
+\cd{bibtex}, then the QRE for matching two records is 
+\cd{bibtex . bibtex}, the QRE for matching one or two records is
+\cd{ bibtex | (bibtex . bibtex)} and the QRE for matching any number of QRE
+records \cd{bibtex*}
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}
 let ambiguous : canonizer = "a"* . (project "a"* $\to$ "a")
-\end{verbatim}
-
-\noindent The behaviour of \verb!ambiguous! is not-well defined since the string
-\verb!"aaa"! can be canonized to any of \verb!"a", "aa", "aaa"! depending on how
-\verb!"aaa"! is parsed. We also require that the regular combinators applied to
+\end{lstlisting}
+
+\noindent The behaviour of \cd{ambiguous} is not-well defined since the string
+\cd{"aaa"} can be canonized to any of \cd{"a", "aa", "aaa"} depending on how
+\cd{"aaa"} is parsed. We also require that the regular combinators applied to
 the kernels are unambiguous since the underlying lenses will end up operating
@@
 example, recall that in the \bibtex{} to EndNote transformation, we had the QREs
-\verb!bibtex! and \verb!endnote! describing the \bibtex{} and EndNote records
-respectively, and the bijective lens \verb!bib_to_end! that maps between
-\verb!bibtex! and \verb!endnote!. The quotient lens which maps between
-\verb!bibtex! and \verb!endnote! is given by the QRE lens
-
-\begin{verbatim}
+\cd{bibtex} and \cd{endnote} describing the \bibtex{} and EndNote records
+respectively, and the bijective lens \cd{bib_to_end} that maps between
+\cd{bibtex} and \cd{endnote}. The quotient lens which maps between
+\cd{bibtex} and \cd{endnote} is given by the QRE lens
+
+\begin{lstlisting}
 let bib_to_end_q : (lens in bibtex <=> endnote) = 
 rquot (lquot (bibtex, bib_to_end), endnote)
- \end{verbatim}
+ \end{lstlisting}
 
@@
 As an example, recall that
-in the \bibtex{} to EndNote transformation we had the QREs \verb!bibtex! and \verb!endnote! describing a \bibtex{} record and an
+in the \bibtex{} to EndNote transformation we had the QREs \cd{bibtex} and \cd{endnote} describing a \bibtex{} record and an
 Endnote record respectively. We also had the example pair 
-\verb!(bib_example , end_example)! where
-
-\begin{verbatim}
+\cd{(bib_example , end_example)} where
+
+\begin{lstlisting}
 bib_example = 
@@
 
-Since the transformation between the kernel of \verb!bibtex! and the kernel of
-\verb!endnote! that maps \verb!bib_example! to \verb!end_example! can be
+Since the transformation between the kernel of \cd{bibtex} and the kernel of
+\cd{endnote! that maps \cd{bib_example! to \cd{end_example} can be
 expressed by a bijective lens, it suffices to type the expression
 
-\begin{verbatim}
+\begin{lstlisting}
 let bib_to_end : (lens in bibtex <=> endnot) 
   = synth bibtex <=> endnote  using  {(bib_example, end_example)}
-\end{verbatim}
+\end{lstlisting}
 

### 2017-10-18 papers/QuotientLensSynth/main.tex
Massive changes to intro

--- 2017-10-18 papers/QuotientLensSynth/main.tex
@@
 \section{Introduction}
-\ifplentyoftime
-\bcp{Not sure there is time to fix this, but the intro seems very long and
-  technical---it takes a long time to get to saying what our actual
-  contributions are!  Wonder if we should move most of it to a second
-  ``Example'' section.} 
-\fi
-\bcp{This sentence could be clearer!}{\em Bidirectional programming languages} such as
-Boomerang~\cite{boomerang}, Augeas~\cite{augeas}, XSugar~\cite{xsugar},
-biXid~\cite{bixid} and X/Inv~\cite{xinv}\bcp{there are many others; we
-  should mention all of them} enable a programmer to express a
-transformation that maps some input data to some output data, as well as a
-transformation that maps the output data back into the input data in a single
-expression.  An example of a {\em bidirectional transformation} which we shall
-revisit severally in this paper is a program which converts citation records
-such as
+A {\em lens} is a {\em bidirectional transformation} between a source
+type $S$ and a target type $T$ in which the forward transformation and the
+backward transformation satisfy a certain set of round-tripping laws (e.g.
+invertibility).
+
+An example of a bidirectional transformation which we shall revisit
+severally in this paper is a lens which converts citation records such as
 \begin{verbatim}
@@
 
-Bidirectional programs have also been called {\em lenses}. Formally, Foster et
-al~\cite{foster} defined a lens $\ell : S \Leftrightarrow V$ that maps between
-an ``abstract" {\em source} $S$ and a ``concrete" {\em view} $V$ to be a triple
-of functions $(\ell.\get, \; \ell.\lput, \; \ell.\create)$ satisfying the
-following equations:\bcp{Why are we showing all the complexity of asymmetric
-lenses?  We should just say that we're considering only the bijective case
-for now and then just give the details of that.  (Related: The
-``abstract/concrete'' terminology doesn't make much sense in the bijective
-setting.)} 
-\begin{align*}
-\ell.get &: S \longrightarrow V & \ell.\lput \; (\ell.\get \; s) \; s &= s\\
-\ell.put &: V \longrightarrow S \longrightarrow S & \ell.\get \; (\ell.\lput \;
-v \; s) &= v\\
-\ell.create &: S \longrightarrow V & \ell.\lput \; (\ell.\create \; v) &= v
-\end{align*}
-
-The $\get$ function presents a view of the data from the source, the
-$\lput$ function enables an updated view to be folded back into the source, and
-the $\create$ component produces a copy of the source from a view, using
-defaults to fill in information that may have been discarded by the $\get$
-function in producing the view. The first law states that the $\lput$ function
-must restore all the information discarded by $\get$ when its arguments are an
-abstract structure and a concrete structure that generates the very same
-abstract structure. The second and third laws state that $\lput$ and $\create$
-must propagate all of the information contained in their abstract arguments to
-the concrete structure they produce. These laws express fundamental expectations
-about how the components of a lens should work together.
-
-However, the constraints that are imposed by the lens laws are sometimes too
-strict. For instance, the programmer may wish to ignore updates to the data
-that only change the order of fields, or add whitespace characters that in
-reality are not part of the data. Foster {\em et al.\relax} therefore proposed
-\textit{quotient lenses} ~\cite{quotient} as a solution to this problem.
-Quotient lenses, which are implemented as a refinement of the bidirectional
-string processing language Boomerang, enable a programmer to write lenses which
-transform data that is quotiented out by an equivalence relation, so that
-records which are equivalent to each other modulo this relation are considered
-the same by the underlying lens.
-
-Formally, given equivalence relations $\sim_S$ and $\sim_V$ defined on the
-source $S$ and the view $V$ respectively, a quotient lens $q : S/{\sim_S}
-\Leftrightarrow V/{\sim_V}$ is a triple of functions $(\ell.\get, \; \ell.\lput,
-\; \ell.\create)$ which satisfy the lens laws up to equivalence:
-\begin{align*}
-\ell.get &: S \longrightarrow V & \ell.\lput \; (\ell.\get \; s) \; s &\sim_S
-s\\
-\ell.put &: V \longrightarrow S \longrightarrow S & \ell.\get \; (\ell.\lput \;
-v \; s) &\sim_V v\\
-\ell.create &: S \longrightarrow V & \ell.\lput \; (\ell.\create \; v) &\sim_V v
-\end{align*}
-Moreover, each of the $\get$, $\lput$ and $\create$ functions must respect
-$\sim_S$ and $\sim_V$. That is, if $s \sim_S s'$ then $\get \; s \sim_V
-\get \; s'$, if $s \sim_S s'$ and $v \sim_V v'$ then $\lput \; v \; s \sim_S
-\lput \; v' \; s'$ and if $v \sim_V v'$, then $\create \; v \sim_S \create \;
-v'$.
-
-\bcp{I'm nervous about making the concept of canonizer so central, since it
-  is not going to be much of a focus for the rest of the paper.  In general,
-it feels like the intro is explaining too much.  We need to find a more
-streamlined path through the ideas!}
-In Boomerang, equivalence relations are expressed using {\em canonizers}
-which are functions that map members of a regular language to their canonical
-representation modulo an equivalence relation. Therefore, in order to give a
-specification for the data, the programmer often has to write a
-regular expression which matches all possible representations of the data, and
-then provide a canonizer that does the job of quotienting the data.
-
-For example, suppose that the programmer wants to
-derive the \bibtex{} to EndNote transformation in such a way that the program
-allows for whitespace to occur arbitrarily between the words in the author and
-title fields, ignores the order in which the author and title fields occur in
-each record, and allows for arbitrary amounts of whitespace in between the
-author and title fields. These rules allow for the \bibtex{} citations
+\sam{I need a better introduction\ldots}
+
+Lenses are desirable since they save the programmer from having to write two
+seperate transformations rather than one. Also, lenses offer strong
+guarantees about how the two transformations interact since lenses are
+constrained by lens laws.
+
+However, these constraints are sometimes too strict. For example,  the
+programmer may want the \bibtex{} to EndNote transformation to allow for
+whitespace to occur arbitrarily between the words in the label, author and
+title fields as well as in between them, and ignore the order in which the
+author and title fields occur in each record. These rules allow for the
+\bibtex{} citations
 \begin{verbatim}
@@
 \end{verbatim}
-to be considered the same (with the EndNote records having a similar
-equivalence relation defined on them).
+to be considered the same, with the EndNote records having similar rules
+defined on them.
+
+As a solution to this problem, Foster et al introduced {\em quotient lenses},
+which are a generalization of lenses in which the source and target data are
+considered modulo an equivalence relation. Quotient lenses are
+both simpler and more flexible, since whole equivalence classes of data
+values can be handled all at once and inessential details like whitespace
+and ordering of fields can be ignored.
+
+However many complicated equivalence relations are difficult to express in
+Boomerang, the lens programming language introduced by Foster et al which
+has quotient lens capability. For instance, while it is possible to express the
+equivalence on the \bibtex{} data above in Boomerang, actually doing so is
+difficult. The difficulty arises from the fact that the entire equivalence is
+composed of three separate equivalences which nest and interact with each other
+in a way that makes them difficult to express with the standard tools
+available. Moreover, many of the equivalences expressible in Boomerang require
+that the programmer first specify the data itself in form of a regular
+expression before specifying the equivalence relation on the data.
+
+To address this issue, we propose the language of {\em Quotient Regular
+Expressions} or QREs, which express a broad class of equivalence relations
+directly on regular languages using a compact and convenient notation. For
+example, with QREs we can specify \bibtex{} records as shown in
+Figure~\ref{fig:example-qre}.
 
@@
 \begin{lstlisting}
-let project (r : regexp) (s : string) = canonizer_of_lens ((del r) . (ins s))
-let wsp : regexp = [ \n\t\r]$^+$
-let wsp_to_space : canonizer = project wsp " "
-let valid_name : canonizer = 
-    name | (name . wsp_to_space . (name .wsp_to_space)$^*$ .name)
-let wsp_to_and : canonizer = project (wsp . "and" . wsp) " and "
-let bib_names : canonizer =  
-    valid_name | (valid_name . wsp_to_and . (valid_name . wsp_to_and)$^*$ . valid_name)
-let bib_author : canonizer = "author = \"" . bib_names . "\""
-let label : regexp = [a-zA-Z0-9]$^+$
-let title : canonizer = 
-    word | (word . wsp_to_space . (word . wsp_to_space)$^*$ . word)
-let bib_title : canonizer = "title = {" . title . "}"
-let wsp_to_comma : canonizer = project ("," . wsp) ",\n"
-let first_perm : canonizer = bib_author . wsp_to_comma . bib_title
-let second_perm : canonizer = bib_title . wsp_to_comma . bib_author
-let bibtex : canonizer = 
-    "@book{" . label. wsp_to_comma . (first_perm | second_perm) . "}"
+let wsp_to_space = project [ \n\t\r]$^+$ $\to$ " "
+let valid_name = name | (name . wsp_to_space . (name . wsp_to_space)$^*$ . name)
+let wsp_to_and = project ([ \n\t\r]$^+$ . "and" . [ \n\t\r]$^+$) $\to$ " and "
+let bib_names = valid_name | (valid_name . wsp_to_and . (valid_name . wsp_to_and)$^*$ . valid_name)
+let bib_author = "author = \"" . bib_names . "\""
+let title = word | (word . wsp_to_space . (word . wsp_to_space)$^*$ . word)
+let bib_title = "title = {" . title . "}"
+let bib_perm = perm (label, bib_author, bib_title) with(project ("," . [ \n\t\r]$^*$ ) $\to$ ",\n")
+let bibtex = "@book{" . bib_perm . "}"
 \end{lstlisting}
-
-  \caption{Boomerang canonizers for the \bibtex{}--EndNote equivalence
-    relations.}
-  \label{fig:example-canonizers}
-\end{figure}
-The code sample in Figure~\ref{fig:example-canonizers} shows how the programmer
-might specify these equivalence relations on the \bibtex{} records in Boomerang
-using canonizers.
-%
-Using the in-built \cd{canonizer_of_lens} function which extracts a canonizer
-from a lens, it is easy to define an equivalence relation that sends all
-members of a regular language to a canonical representative.\bcp{Oh dear.
-  The \cd{canonizer_of_lens} thing is really kind of a hack---we're using
-  the fancy/specialized language of lenses to write what could really be a
-  simple program in an arbitrary language (and then throwing away half of the
-  lens, when the whole reason we give for having lenses in the first place
-  is because we want both directions).  As an implementation trick, it was a
-good idea.  But trying to get readers to understand it here does not seem
-like a good idea.} For example
-the expression 
-\begin{lstlisting}
-let wsp_to_space : canonizer = canonizer_of_lens ((del wsp) . (ins " "))
-\end{lstlisting}
-\noindent defines a canonizer which sends arbitrary occurences of whitespace
-characters to a single space\bcp{Going into this level of detail encourages
-readers to actually try to understand the details, but we don't actually
-give enough details to allow them to do that!}. However, a problem arises
-when trying to specify 
-the equivalence relation which nests this projection relation with the relation
-that considers all permutations of the \bibtex{} fields equivalent.
-
-Specifically, it becomes necessary to specify each of the permutations
-seperately, since Boomerang's in-built permutation combinator does not account
-for separators in between the regular expressions being permuted:\bcp{Why
-  couldn't we just enrich Boomerang's permutation combinator to do this?}
-
-\begin{lstlisting}
-let first_perm : canonizer = bib_author . wsp_to_comma . bib_title
-let second_perm : canonizer = bib_title . wsp_to_comma . bib_author
-\end{lstlisting}
-
-One possible solution is to permute the alternative canonizers
-\cd{bib_author.wsp_to_comma} and \cd{bib_title.wsp_to_comma},
-but Boomerang's type-checker will reject this expression since the permutation
-combinator accepts regular expressions rather than canonizers\bcp{It feels
-  like we're lost, here, in details of exactly what Boomerang does and
-  doesn't do.  This doesn't seem especially relevant to quotient lenses in
-  general, or to what we're doing in this paper.}. In any case, the
-regular expression
-\begin{lstlisting}
-bib_author . wsp_to_comma . bib_title . wsp_to_comma | 
-bib_title . wsp_to_comma . bib_author . wsp_to_comma
-\end{lstlisting}
-is not equivalent to,
-\begin{lstlisting}
-bib_author . wsp_to_comma . bib_title  | 
-bib_title . wsp_to_comma . bib_author
-\end{lstlisting}
-thus the programmer would have to write more lenses and canonizers to correct
-this. Therefore, in order to utilize this trick, the programmer might choose to
-specify all the regular expressions first (including the whole regular
-expressions canonized by the \cd{bib_author} and \cd{bib_title}
-canonizers), and then the canonizers separately. Alternatively, he/she may stick
-to using canonizers and then liberally make use of the built-in
-functions \cd{uncanonized_type} and \cd{canonized_type} which return the
-regular expression matching all the data and the regular expression matching
-the canonical representatives respectively.
-
-In any case, all of these alternatives lead to verbose code that is
-difficult to understand. Moreover, if the \bibtex{} records had 3 or more
-fields or even nested permutations, then the programmer would quickly find
-themselves in a bind with these workarounds.
-
-To address this issue, we propose the language of {\em Quotient Regular
-Expressions} or QREs, which express a broad class of equivalence relations
-directly on regular languages using a compact and convenient notation.
-
-With QREs we can specify \bibtex{} records as shown in Figure~\ref{fig:example-qre}.
-
-\begin{figure}[t]
-\begin{lstlisting}
-let wsp_to_space : canonizer = project [ \n\t\r]$^+$ $\to$ " "
-let valid_name : canonizer = 
-   name | (name . wsp_to_space . (name . wsp_to_space)$^*$ . name)
-let wsp_to_and : canonizer = 
-   project ([ \n\t\r]$^+$ . "and" . [ \n\t\r]$^+$) $\to$ " and "
-let bib_names : canonizer =  
-   valid_name | (valid_name . wsp_to_and . (valid_name . wsp_to_and)$^*$ . valid_name)
-let bib_author : canonizer = "author = \"" . bib_names . "\""
-let label : regexp = [a-zA-Z0-9]$^+$
-let title : canonizer = word | (word . wsp_to_space . (word . wsp_to_space)$^*$ . word)
-let bib_title : canonizer = "title = {" . title . "}"
-let bib_perm  : canonizer = perm (label, bib_author, bib_title) with
-   (project ("," . [ \n\t\r]$^*$ ) $\to$ ",\n")
-let bibtex : canonizer = "@book{" . bib_perm . "}"
-\end{lstlisting}
-
-\caption{QRE specification of \bibtex{} records.  \bcp{I'm afraid this pair
-    of examples is not making the point we intended.  I thought we were
-    hoping to just say ``Here's the old Boomerang version; here's the QRE
-    version; look how much smaller and simpler the second one is.''  But the
-    second one {\em isn't} actually much smaller and, at least on first
-    impression, doesn't look simpler.}}
+\caption{QRE specification of \bibtex{} records. }
   \label{fig:example-qre}
@@
 
-The expression \bcp{mangled---did you proofread this?}\cd{project [
-  \n\t\r]+ -> " "} is a QRE that matches any 
+The expression \cd{project [ \\n\\t\\r]+ -> " "} is a QRE that matches any 
 number of whitespace characters, and then considers all strings of whitespace
@@
 and title fields occur in that order, separated by a comma followed a newline.
-The \cd{perm} QRE is defined at the level of canonizers which allows for
-easy nesting of permutations. Similarly, it is easy to specify any number of
-fields within the \cd{perm} QRE.
+The \cd{perm} QRE allows for easy nesting of permutations. Similarly, it is easy
+to specify any number of fields within the \cd{perm} QRE.
 
 While QREs kill two birds with one stone by specifying the
-data and the equivalence relation on the data in one expression,
-he/she\bcp{Awkward: just use ``she'' (or find a way to alternate pronouns,
-  if that doesn't create confusion)}
-still has to manually program the underlying lens which maps the equivalence
+data and the equivalence relation on the data in one expression, the programmer
+still has to manually derive the underlying lens which maps the equivalence
 classes of the source data to the equivalence classes of the view data, a task
 which is itself tedious and error-prone as the code snippet below
-illustrates:\bcp{If we added this to the old-Boomerang figure it might make
-  a better case for size and simplicity!}
+illustrates:
 
@@
 
-Our approach is to define a class of {\em QRE lenses} which are the same as
-bijective lenses with ``canonizers at the edges'' as in
-Figure~\ref{fig:attheedges}. In the forward direction, a QRE lens $q$ uses the
-source QRE $c$ to compute a unique representative for the data modulo the
-equivalence relation defined by $c$ and then applies the $\get$ function of a
-bijective lens $\ell$ to this representative. In the backward direction, $q$
-operates similarly, but using the view QRE $c'$ and the $\lput$ function of
-$\ell$.
+Our approach, which follows the design of quotient lenses in Boomerang, is to
+define a class of {\em QRE lenses} which are the same as bijective lenses with
+``canonizers at the edges'' as in Figure~\ref{fig:attheedges}. In the forward
+direction, a QRE lens $q$ uses the source QRE $c$ to compute a unique
+representative for the data modulo the equivalence relation defined by $c$ and
+then applies the $\get$ function of a bijective lens $\ell$ to this
+representative. In the backward direction, $q$ operates similarly, but using
+the view QRE $c'$ and the $\lput$ function of $\ell$.
 

### 2017-10-18 papers/QuotientLensSynth/main.tex

@@
 
-The expression \cd{project [ \\n\\t\\r]+ -> " "} is a QRE that matches any 
+The expression \cd{project [\n\t\r]}$^+$ $\to$ \cd{" "} is a QRE that matches any 
 number of whitespace characters, and then considers all strings of whitespace

### 2017-10-18 papers/QuotientLensSynth/main.tex
More rewriting. Started on related work

@@
 guarantees about how the two transformations interact since lenses are
-constrained by lens laws.
-
-However, these constraints are sometimes too strict. For example,  the
-programmer may want the \bibtex{} to EndNote transformation to allow for
-whitespace to occur arbitrarily between the words in the label, author and
-title fields as well as in between them, and ignore the order in which the
-author and title fields occur in each record. These rules allow for the
-\bibtex{} citations
+constrained by lens laws. However, these constraints are sometimes too strict.
+For example, the data may be such that a record can be represented in many ways
+because of the presence of spurious whitespace characters or reordering of
+fields. In such cases, deriving a lens which obeys the lens laws becomes a
+daunting task.
+
+As a solution to this problem, Foster et al proposed {\em quotient lenses},
+which are a generalization of lenses in which the source and target data are
+considered modulo an equivalence relation. Quotient lenses are
+both simpler and more flexible, since whole equivalence classes of data
+values can be handled all at once and inessential details like whitespace
+and ordering of fields can be ignored.
+
+However many complicated equivalence relations are difficult to express in
+Boomerang, the lens programming language in which Foster et al implemented
+quotient lenses. For example, suppose that the \bibtex{} to EndNote
+transformation is required to allow for whitespace to occur arbitrarily between
+the words in the label, author and title fields as well as in between them, and
+ignore the order in which the author and title fields occur in each record.
+These rules allow for the \bibtex{} citations
 \begin{verbatim}
@@
 
-As a solution to this problem, Foster et al introduced {\em quotient lenses},
-which are a generalization of lenses in which the source and target data are
-considered modulo an equivalence relation. Quotient lenses are
-both simpler and more flexible, since whole equivalence classes of data
-values can be handled all at once and inessential details like whitespace
-and ordering of fields can be ignored.
-
-However many complicated equivalence relations are difficult to express in
-Boomerang, the lens programming language introduced by Foster et al which
-has quotient lens capability. For instance, while it is possible to express the
-equivalence on the \bibtex{} data above in Boomerang, actually doing so is
-difficult. The difficulty arises from the fact that the entire equivalence is
-composed of three separate equivalences which nest and interact with each other
-in a way that makes them difficult to express with the standard tools
-available. Moreover, many of the equivalences expressible in Boomerang require
-that the programmer first specify the data itself in form of a regular
-expression before specifying the equivalence relation on the data.
+While it is possible to express the equivalence on the \bibtex{} data above in
+Boomerang, actually doing so is difficult. The difficulty arises from the fact
+that the entire equivalence is composed of three separate equivalences which
+nest and interact with each other in a way that makes them difficult to express
+with the standard tools available. Moreover, many of the equivalences
+expressible in Boomerang require that the programmer first specify the data
+itself in form of a regular expression before specifying the equivalence
+relation on the data.
 
@@
 \begin{lstlisting}
-let wsp_to_space = project [ \n\t\r]$^+$ $\to$ " "
+let wsp = [ \n\t\r]$^+$
+let wsp_to_space = project wsp $\to$ " "
 let valid_name = name | (name . wsp_to_space . (name . wsp_to_space)$^*$ . name)
-let wsp_to_and = project ([ \n\t\r]$^+$ . "and" . [ \n\t\r]$^+$) $\to$ " and "
+let wsp_to_and = project (wsp . "and" . wsp) $\to$ " and "
 let bib_names = valid_name | (valid_name . wsp_to_and . (valid_name . wsp_to_and)$^*$ . valid_name)
@@
 let bib_title = "title = {" . title . "}"
-let bib_perm = perm (label, bib_author, bib_title) with(project ("," . [ \n\t\r]$^*$ ) $\to$ ",\n")
+let bib_perm = perm (label, bib_author, bib_title) with (project ("," .wsp) $\to$ ",\n")
 let bibtex = "@book{" . bib_perm . "}"
@@
 \begin{lstlisting}
-let bib_to_end : (lens in bibtex <=> endnot) 
+let bib_to_end : (lens in bibtex <=> endnote) 
 = synth bibtex <=> endnote  using  {(bib_example, end_example)}
@@
 
+Bidirectional transformations have been used in diverse areas of computing
+where they arise as parsers and pretty printers, marshallers
+and unmarshallers, serializers and deserializers, database views and view
+updaters and many others\sam{TODO put citations}. Such transformations have
+been extensively studied since they were proposed as a solution to the classical
+{\em view-update problem} in the database community, where the challenge is to
+derive a program which extracts a view of data from a source, as well as a
+program which folds view updates back into the source safely and correctly.
+
+This paper builds on the work of Foster et al~\cite{quotientlenses} who
+introduced the theory of quotient lenses and implemented quotient lenses as a
+refinement of the bidirectional string processing language Boomerang.
+In Boomerang, the source and target types are specified using regular
+expressions, and the equivalence relations are expressed using canonizers, which
+are functions that map elements of a regular language to their canonical
+representative. Boomerang canonizers can express a very broad class of
+equivalence relations between regular languages, but actually doing so is often
+difficult with complicated equivalence relations. For instance, Boomerang's
+in-build permutation combinator does not account for separators in between the
+regular expressions. Also, this combinator permutes regular expressions rather
+than canonizers, thus making it difficult to express nested permutations which
+occur in many data formats, especially XML and XML-like formats.
+
+In their paper, Foster et al also discuss other bidirectional programming
+languages which support quotienting of data including XSugar~\cite{xsugar},
+biXid~\cite{bixid} and X/Inv~\cite{xinv}. XSugar programs bidirectionally
+convert data stored in XML and ASCII formats respectively, with the
+transformations specified by a pair of unambiguous grammars. The quotienting
+occurs on the XML side by use of a generic canonizer which standardizes the
+representation of trees. Well-formed XSugar programs are guaranteed to be
+bijective modulo an equivalence relation that captures XML normalization. biXid
+programs convert between pairs of XML documents, with the XML formats specified
+using a pair of grammars as in XSugar. However, biXid grammars can be ambiguous.
+This ambiguity is what allows biXid to express equivalences on the data.
+Finally, Foster et al discuss a possible connection with the language languages
+X and Inv which support a primitive duplication combinator that does not work
+well with the lens laws, but can be expressed using Boomerang quotient lenses.
+
  \bcp{Make sure to list/discuss:
@@
   reading about connections to view updates, etc.) and quotient lenses.
-  \item Discuss what other lens-like programming languages (besides
-  boomerang) do about quotienting.  E.g., Augeas, XSugar, ...
-  \item How quotient lenses are used in Boomerang.
   \item Search for papers that cite the quotient lens paper and see if they

### 2017-10-19 papers/QuotientLensSynth/main.tex
more changes

@@
 \section{Introduction}
-A {\em lens} is a {\em bidirectional transformation} between a source
-type $S$ and a target type $T$ in which the forward transformation and the
-backward transformation satisfy a certain set of round-tripping laws (e.g.
-invertibility).
+A {\em lens} $\ell$ is a {\em bidirectional transformation} between a source
+type $S$ and a target type $T$ in which the forward transformation
+$\ell.\get$ and the backward transformation $\ell.\lput$ satisfy certain
+round-tripping laws such as invertibility.
 
@@
 
-\sam{I need a better introduction\ldots}
-
-Lenses are desirable since they save the programmer from having to write two
-seperate transformations rather than one. Also, lenses offer strong
+Lenses have the advantage of saving the programmer from having to
+write two seperate transformations rather than one. Also, lenses offer strong
 guarantees about how the two transformations interact since lenses are
-constrained by lens laws. However, these constraints are sometimes too strict.
-For example, the data may be such that a record can be represented in many ways
-because of the presence of spurious whitespace characters or reordering of
-fields. In such cases, deriving a lens which obeys the lens laws becomes a
-daunting task.
+constrained by lens laws.
+
+However, these constraints are sometimes too strict. For example, a record in
+the data may represented in many ways because of the presence of spurious
+whitespace characters or because of the reordering of fields. Deriving a lens
+which obeys the lens laws in such a case can be daunting and frustrating.
 
@@
 nest and interact with each other in a way that makes them difficult to express
-with the standard tools available. Moreover, many of the equivalences
+with the standard Boomerang combinators. Moreover, many of the equivalences
 expressible in Boomerang require that the programmer first specify the data
@@
 \begin{lstlisting}
-perm (label, bib_author, bib_title) with (project ("," . [\n\t\r]$^*$ ) $\to$ ",\n")
+perm (label, bib_author, bib_title) with (project ("," . wsp ) $\to$ ",\n")
 \end{lstlisting}
@@
 
-While QREs kill two birds with one stone by specifying the
-data and the equivalence relation on the data in one expression, the programmer
-still has to manually derive the underlying lens which maps the equivalence
-classes of the source data to the equivalence classes of the view data, a task
-which is itself tedious and error-prone as the code snippet below
-illustrates:
+While QREs make it easier to write quotient lenses by specifying the data and
+the equivalence relation on the data in one expression, the programmer still
+has to manually derive the underlying lens which maps the equivalence classes
+of the source data to the equivalence classes of the target data. This is
+itself tedious and error-prone task, as the code snippet below illustrates:
 
@@
 the \cd{preamble} lens replaces the \bibtex{} preamble \cd{@book} with the
-EndNote preamble \cd{\%0 Book\n\%F } in the \bibtex{record}. The programmer
+EndNote preamble \cd{\%0 Book\n\%F } in the \bibtex{} record. The programmer
 can ill afford to make a mistake with any one of the small but vital details
@@
 
-Therefore, to address this challenge, we demonstrate how to {\em synthesize}
-quotient lenses from a pair of QREs and a set of input-output example pairs.
-Our key insight is to reduce the problem of synthesizing QRE lenses to the
-problem of synthesizing {\em bijective lenses}, a particular subset of the
-lenses that can be expressed in Boomerang.\bcp{As stated, this is
-  misleading: the reader might have the impression that we somehow reduce
-  the problem of synthesizing {\em asymmetric} quotient lenses to the
-  problem of synthesizing bijective ordinary lenses!} This is a problem
+Therefore, to address this challenge, we identify a special class of quotient
+lenses which we call {\em QRE lenses}, and show how to {\em
+synthesize} QRE lenses from a pair of QREs and a set of input-output
+example pairs. Our key insight is to reduce the problem of synthesizing QRE
+lenses to the problem of synthesizing {\em bijective lenses}, which are the
+bijections created from the Boomerang lens combinators. This is a problem
 which we solved in our prior work ~\cite{popl18}.
 
-Our approach, which follows the design of quotient lenses in Boomerang, is to
-define a class of {\em QRE lenses} which are the same as bijective lenses with
-``canonizers at the edges'' as in Figure~\ref{fig:attheedges}. In the forward
-direction, a QRE lens $q$ uses the source QRE $c$ to compute a unique
+QRE lenses lenses are the same as bijective lenses with ``canonizers at the
+edges'' as in Figure~\ref{fig:attheedges}, and are typed using QREs. In the
+forward direction, a QRE lens $q$ uses the source QRE $c$ to compute a unique
 representative for the data modulo the equivalence relation defined by $c$ and
@@
 which sends all elements of the language described by the regular expression
-$R$, and the
+$R$ to the canonical representative $s$, and the
 
@@
 \begin{lstlisting}
-let squash_author : canonizer =  
-squash space_name $\to$ comma_name using (get comma_to_space)
+let squash_author = squash space_name $\to$ comma_name using comma_to_space
 
@@
 are equivalent, with the \cd{comma_name} representation chosen as the
-canonical representation, and the function \cd{comma_to_space.get} giving
+canonical representation, and the lens \cd{comma_to_space} giving
 the canonizing function which converts data in the \cd{space_name} format
@@
 
-The $\squash{R}{R'}{f}$ has the added advantage that if the transformation from
-$R$ to $R'$ can be expressed by a bijective lens that is consistent with a set
+The $\squash{R}{R'}{\ell}$ has the added advantage that if the transformation
+from $R$ to $R'$ can be expressed by a bijective lens that is consistent with a set
 of examples, then the programmer can synthesize the lens from $R$ to $R'$
@@
 }"
-\end{verbatim}
+\end{lstlisting}
 
@@
 
-\begin{verbatim}
+\begin{lstlisting}
 end_example = 
@@
 %F conway"
-\end{verbatim}
+\end{lstlisting}
 
 Since the transformation between the kernel of \cd{bibtex} and the kernel of
-\cd{endnote! that maps \cd{bib_example! to \cd{end_example} can be
+\cd{endnote!} that maps \cd{bib_example!} to \cd{end_example} can be
 expressed by a bijective lens, it suffices to type the expression
@@
   \begin{subfigure}[b]{.49\textwidth}
-    \includegraphics{generated-graphs/times_new}
+    \includegraphics{times_new.eps}
     \caption{}
@@
 
- \bcp{Make sure to list/discuss:
-  \begin{itemize}
-  \item Seminar papers on lenses (to which we can refer readers for further
-  reading about connections to view updates, etc.) and quotient lenses.
-  \item Search for papers that cite the quotient lens paper and see if they
-  extend the ideas there.
-  \item Search for theoretical papers involving equivalence relations over
-  regular languages. 
-  \end{itemize}
-}
+This paper also builds on our previous work where we showed how to synthesize
+bijective lenses from a pair of regular expressions describing the source and
+target types and a set of input-output examples. We refer the reader to the
+Related Work section of \cite{popl18} for a survey of the area of
+type-and-example driven synthesis.
 

### 2017-10-20 papers/QuotientLensSynth/main.tex
added related work section

@@
     \centering
-    \includegraphics{generated-graphs/times_opt}
+    %\includegraphics{generated-graphs/times_opt}
     \caption{}
@@
   \begin{subfigure}[b]{.49\textwidth}
-    \includegraphics{times_new.eps}
+    %\includegraphics{times_new.eps}
     \caption{}
@@
 
-This paper also builds on our previous work where we showed how to synthesize
-bijective lenses from a pair of regular expressions describing the source and
-target types and a set of input-output examples. We refer the reader to the
-Related Work section of \cite{popl18} for a survey of the area of
-type-and-example driven synthesis.
+As far as synthesis goes, though there is a good deal of recent research on
+synthesizing unidirectional string
+transformations~\cite{singh2012learning,le-pldi-2014,gulwani-popl-2014,
+perelman2014test,Singh:blinkfill}, our system Optician, which we introduced in
+~\cite{popl18}, is the first attempt to synthesize bidirectional
+transformations that we are aware of. We compared Optician to two of these
+unidirectional string transformers, Flash Fill~\cite{gulwani-popl-2014} and
+FlashExtract~\cite{le-pldi-2014}, and found that these tools were unsuccessful
+in synthesizing the complex transformations that can be expressed in Optician.
+Furthermore, neither of these tools were able to infer transformations 
+which occurred under two iterations.
+
+Much of the research in synthesis assumes that the synthesizer is provided with
+a collection of examples. Optician differs in that it requires that the
+programmer supplies both examples {\em and} format descriptions in the form of
+regular expressions. There are many other recent results showing how to
+synthesize functions from type-based
+specifications~\cite{augustsson-2004,osera+:pldi15,
+feser-pldi-2015,scherer-icfp-2015,frankle+:popl16,armando+:pldi16}.
+These systems enumerate programs of their target language, orienting their
+search procedures to process only terms that are well-typed.
+Optician is distinctive in that it synthesizes terms in a language with many
+type equivalences. 
+Perhaps the system most similar to Optician is InSynth~\cite{gvero-pldi-2013}, a
+system for synthesizing terms in the simply-typed lambda calculus that addresses
+equivalences on types.
+
+Morpheus~\cite{morpheus} is another synthesis system that uses two
+communicating synthesizers to generate programs.  In both Morpheus and
+Optician, one synthesizer provides an 
+outline for the program, and the other fills in that outline with program
+details that satisfy the user's specifications.
+This approach works well in large search spaces, which
+require some enumerative search.
+One important way in which Optician differs from Morpheus is that in
+Morpheus, an outline is a sketch---an
+\emph{expression}
+containing holes---whereas
+an outline in Optician is a pair of regular
+expressions, i.e., a 
+\emph{type}.  Moreover, in order to implement an efficient
+search procedure, we had to create both a new type language and a new
+term language for lenses.  Once we did so, we proved our new, more
+constrained language
+designed for synthesis was just as expressive as the original, more
+flexible and compositional language designed for human programmers.
 

### 2017-10-20 papers/QuotientLensSynth/main.tex
Made changes to the intro

@@
 
-However, these constraints are sometimes too strict. For example, a record in
-the data may represented in many ways because of the presence of spurious
-whitespace characters or because of the reordering of fields. Deriving a lens
-which obeys the lens laws in such a case can be daunting and frustrating.
+However, the constraints imposed by the lens laws are often too strict. For
+example, when a record may be represented in many ways, perhaps because
+whitespace characters are allowed to occur arbitrarily in certain parts of the
+data, deriving a lens which obeys the lens laws in such cases can be daunting
+and frustrating.
 
@@
 which are a generalization of lenses in which the source and target data are
-considered modulo an equivalence relation. Quotient lenses are
-both simpler and more flexible, since whole equivalence classes of data
-values can be handled all at once and inessential details like whitespace
-and ordering of fields can be ignored.
+considered modulo an equivalence relation. Quotient lenses, which are
+implemented as a refinement of the bidirectional string processing language
+Boomerang, are both simpler and more flexible, since whole equivalence classes
+of data values can be handled all at once and inessential details like
+whitespace and ordering of fields can be ignored.
 
 However many complicated equivalence relations are difficult to express in
-Boomerang, the lens programming language in which Foster et al implemented
-quotient lenses. For example, suppose that the \bibtex{} to EndNote
-transformation is required to allow for whitespace to occur arbitrarily between
-the words in the label, author and title fields as well as in between them, and
-ignore the order in which the author and title fields occur in each record.
+Boomerang. For example, suppose that the \bibtex{} to EndNote
+transformation which we introduced is required to allow for whitespace to occur
+arbitrarily between the words in the label, author and title fields as well as
+in between them, and ignore the order in which the author and title fields
+occur in each record.
 These rules allow for the \bibtex{} citations
@@
 with the standard Boomerang combinators. Moreover, many of the equivalences
-expressible in Boomerang require that the programmer first specify the data
-itself in form of a regular expression before specifying the equivalence
-relation on the data.
+expressible in Boomerang, such as Boomerang's permutation combinator, require
+that the programmer first specify the data itself in form of a regular
+expression before specifying the equivalence relation on the data. This leads to
+verbose code that is difficult to understand.
 

### 2017-10-24 papers/QuotientLensSynth/main.tex

@@
       the figure again.  (I have the OmniGraffle file for it, so this is
-      easy if I know how it should look.)}\bcp{Done---what do you think?}}
+      easy if I know how it should look.)}\bcp{Done---what do you think?}
+      \sam{Yes this is more accurate, except that we don't need a general choose
+      function. With QREs the kernel language is always contained in the whole
+      language so we just use the inclusion as the choose function. Also,
+      perhaps it would be clearer if we said source structures and target
+      structres instead of all structures?}}
   \label{fig:attheedges}
@@
 \AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(R')$}
-\BinaryInfC{$squash(R, R', f)$ is well formed}
+\BinaryInfC{$squash(R, R', \ell)$ is well formed}
 \end{prooftree}
@@
 language for quotient lenses. In this section of the paper, we introduce
-\textit{QRE Lenses}. QRE lenses are quotient lenses based on the Boomerang
-quotient lens combinators, but which use QREs to describe their source and view
-types.
+\textit{QRE Lenses}, a class of quotient lenses specified using QREs.
 
@@
 
-\noindent One potential pitfall with QRE lenses is that the combinators which
+\noindent In other words, the get function of the QRE first canonizes the source
+data using $\mathit{Canonizer}(c)$ and then applies $f$ to the result, while the
+put function of the QRE first canonizes the target data using
+$\mathit{Canonizer}(c')$ and the applies $f^{-1}$ to the result.
+
+One potential pitfall with QRE lenses is that the combinators which
 we define on them may not preserve this normal form. For example, the
@@
 canonizers a the edges. Consequently, in order to define quotient lenses, we
-need to choose a class of bijections that has just the right properties with
-respect to the quotient lens combinators. Our choice is to use the class of
-\textit{bijective lenses}, a subset of the lenses expressible in Boomerang which
-we identified and studied in our previous work~\cite{popl18}. Bijective lenses,
-which we briefly discuss in the next subsection, possess the properties needed
-to ensure that all QRE lenses have the normal form suggested in
-\ref{normalform}.
+need to choose a set $S$ of bijections that has the right properties in the
+sense that QRE lenses built from the bijections in $S$ are closed under the
+action of QRE lens combinators. Our choice is to use the class of
+\textit{bijective lenses}, which are the bijections that can be expressed in
+Boomerang. In the next subsection we briefly introduce bijective lenses and
+discuss the properties of bijective lenses which we will use to prove that QRE
+lenses have the normal form of bijective lens with canonizers at the edges.
 
 \subsection{Bijective Lenses}
-Define the set of \textit{bijective lenses} to be the set of bijections
-between regular languages created from Boomerang lens combinators.
+We define the set of \textit{bijective lenses} to be the set of bijections
+between regular languages created using Boomerang lens combinators.
 The syntax for the language of bijective lenses is given by
@@
 
-The denotation of a lens $\ell$ is $\llbracket \ell \rrbracket \subseteq
+The denotation of a lens $\ell$ is $\llbracket \ell \rrbracket \subseteq
 \mathit{String} \times \mathit{String}$. If $(s_1, s_2) \in \llbracket \ell
@@
 The $\mathit{const}(s, t)$ lens replaces the string $s$ with $t$ in the source
-data in the forward direction, and $t$ with $s$ in the view data in the backward
-direction. $\mathit{id}(R)$ applies the identity function to the source and view
+data in the forward direction, and $t$ with $s$ in the target data in the backward
+direction. $\mathit{id}(R)$ applies the identity function to the source and target
 in $\mathcal{L}(R)$ in both directions. The composition lens $\ell' \circ \ell$
 applies $\ell$ followed by then $\ell'$ to the source in the forward direction,
-and applies $\ell'$ followed by $\ell$ to the view in the backward direction.
+and applies $\ell'$ followed by $\ell$ to the target in the backward direction.
 The lens $\ell \cdot \ell'$ first splits the string $s$ into $s_1$ and $s_2$,
@@
 chooses to apply $\ell$ or $\ell'$ depending on whether the source
-(resp. view) data is matched by $\ell$ or $\ell'$ in the forward (resp.
-backward direction. The $\ell^*$ lens splits the string $s$ into strings $s_1,
+(resp. target) data is matched by $\ell$ or $\ell'$ in the forward (resp.
+backward direction). The $\ell^*$ lens splits the string $s$ into strings $s_1,
 \ldots, s_n$, applies $\ell$ to each $s_i$ to get $t_i$, and then concatenates
@@
 $S$ are regular expressions. If $\ell : R \Leftrightarrow S$, then the source
-type of $\ell$ is $\mathcal{L}(R)$ and the view type of $\ell$ is
+type of $\ell$ is $\mathcal{L}(R)$ and the target type of $\ell$ is
 $\mathcal{L}(S)$. Interestingly, with the bijective lens type system, a lens
@@
 \Leftrightarrow S'$ provided that $R$ (resp. $S$) can be proven to be
-equivalent to $R'$ (resp. $S$) from the star-semiring axioms. The typing rules
-for bijective lenses are given in Figure~\ref{fig:lensrules}.
+equivalent to $R'$ (resp. $S$) from the star-semiring axioms:
+
+\begin{prooftree}
+\AxiomC{$\ell : R \Leftrightarrow S$}
+\AxiomC{$R \equiv^d R'$}
+\AxiomC{$S \equiv^d S'$}
+\TrinaryInfC{$\ell : R' \Leftrightarrow S'$}
+\end{prooftree}
+
+The other typing rules for bijective lenses are given in
+Figure~\ref{fig:lensrules}.
 
@@
 \AxiomC{$\ell : R \Leftrightarrow S$}
-\AxiomC{$R \equiv R'$}
-\AxiomC{$S \equiv S'$}
+\AxiomC{$R \equiv^d R'$}
+\AxiomC{$S \equiv^d S'$}
 \TrinaryInfC{$\ell : R' \Leftrightarrow S'$}
@@
 Having given a brief overview of the class of bijective lenses, we now introduce
-the class of QRE lenses, whose language is given by following grammar:
+the class of QRE lenses. The language of QRE lenses is given by following
+grammar:
 $$ q := \mathit{lift}(\ell) \sep \mathit{lquot}(q, \ell) \sep
@@
 same name. The $\mathit{lift}(\ell)$ quotient lens enables a bijective lens
-$\ell$ to be considered a quotient lens where the equivalence relation applied
-to the source and view types is the equality relation, and the regular
-combinators $\cdot \; * \; |$ on QRE lenses exhibit the expected behaviour,
-with the caveat that the regular expressions used by these lenses are strongly
-unambiguous.
+$\ell$ to be considered a quotient lens, with the equivalence relation applied
+to both the source and target formats being the equality relation. The
+regular combinators $\cdot \; * \; |$ on QRE lenses exhibit the expected
+behaviour, with the caveat that the regular expressions used by these lenses
+are strongly unambiguous. The composition operator $\circ$ also exhibits the
+expected behaviout, though the conditions that must be checked in order for the
+composition to be well defined are perhaps unintuitive. The reasons why
+these conditions are necesaary will become more apparent when proving normal
+forms for QRE lenses.
 
@@
 complete set of representatives for the equivalence relation $\eqrel{c}$. The
-$\mathit{lquot}(q, c)$ combinator does the same but on the view data. For
-example, recall that in the \bibtex{} to EndNote transformation, we had the QREs
-\cd{bibtex} and \cd{endnote} describing the \bibtex{} and EndNote records
+$\mathit{lquot}(q, c)$ combinator does the same but on the view data.
+
+For example, recall that in the \bibtex{} to EndNote transformation, we had the
+QREs \cd{bibtex} and \cd{endnote} describing the \bibtex{} and EndNote records
 respectively, and the bijective lens \cd{bib_to_end} that maps between
@@
 
+\iffalse
 The difference between QRE lenses and Boomerang quotient lenses lies in the
@@
 expensive to be useful in practice~\cite{quotientlenses}.
+\fi
 
 \subsection{Semantics of QRE Lenses}
+Each QRE lens $q$ has a type $q : c \Leftrightarrow c'$ where $c, c'$ are QREs.
+If $q : c \Leftrightarrow c'$, then the source format is described by $W(c)$ and
+the canonical set of representative for the source data is described by $K(c)$.
+Similarly, the target format is described by $W(c')$ and the canonical set of
+representatives for the target data is described by $K(c')$. The underlying lens
+of $q$ is a bijective lens $\ell : K(c) \Leftrightarrow K(c')$.
+
 The denotation $\llbracket q \rrbracket$ of a QRE lens $q:c \Leftrightarrow c'$
@@
 \Longleftrightarrow W(c')/{\eqrel{c'}}$. The typing rules and denotation of QRE
-lenses are given in Figure~\ref{fig:qlenssemantics}. The following theorem
-confirms that QRE lenses exhibit the intended behaviour.
-
+lenses are given in Figure~\ref{fig:qlenssemantics}.
+
+The following theorem confirms that the semantics for QRE lenses are correct:
 \begin{theorem}
-If there is a derivation $q:c \Leftrightarrow c'$ then
-$\llbracket q \rrbracket : W(c)/{\eqrel{c}} \Leftrightarrow
-W(c')/{\eqrel{c'}}$ is a well-defined quotient lens.
+If there is a derivation $q : c \Leftrightarrow c'$, then $\llbracket q
+\rrbracket : W(c)/{\eqrel{c}} \Leftrightarrow W(c')/{\eqrel{c'}}$ is a
+well-defined quotient lens.
 \end{theorem}
@@
 \AxiomC{$\ell : R \Leftrightarrow S$}
-\UnaryInfC{$\mathit{lift}(\ell): R \Leftrightarrow id(S)$}
+\UnaryInfC{$\mathit{lift}(\ell): id(R) \Leftrightarrow id(S)$}
 \end{prooftree}
@@
 \AxiomC{$q_1 : c \Leftrightarrow c'$}
-\AxiomC{$q_2 : c' \Leftrightarrow c''$}
-\BinaryInfC{$q_2 \circ q_1: c \Leftrightarrow c''$}
+\AxiomC{$q_2 : c'' \Leftrightarrow c'''$}
+\AxiomC{$\substack{\mathcal{L}(W(c'))= \mathcal{L}(W(c'')) \\ K(c') \equiv^d
+K(c'')}$}
+\AxiomC{$\mathit{Canonizer}(c') = \mathit{Canonizer}(c'')$}
+\QuaternaryInfC{$q_2 \circ q_1: c \Leftrightarrow c'''$}
 \end{prooftree}
@@
 \end{align*}
-for some bijective lens $\ell$. We now provide a proof that all QRE lenses are
-of this form.\bcp{Maybe some cases of the proof can be elided in the short
-  version?} 
+for some bijective lens $\ell$. In other words, each QRE lens is the same as a
+bijective lens with canonizers at the edges. The following theorem confirms that
+this indeed is the case:
+
 \begin{theorem}\label{normal form}
@@
 \end{theorem}
+We elide a full proof of this statement, since the proof is attained via a
+straightforward induction on the derivation $q : c \Leftrightarrow c$. Since
+bijective lenses are closed under strongly unambiguous regular operators, the
+tricky case is composition. The composition case is more involved since we need
+to show that we can get rid of the canonizers in the middle:
+
+\sam{TODO: Put a diagram here to show what you mean.}
+
+The derivation rule and denotation for composition are as follows:
+  \begin{prooftree}
+\AxiomC{$q_1 : c \Leftrightarrow c'$}
+\AxiomC{$q_2 : c'' \Leftrightarrow c'''$}
+\AxiomC{$\substack{\mathcal{L}(W(c'))= \mathcal{L}(W(c'')) \\ K(c') \equiv^d
+K(c'')}$}
+\AxiomC{$\mathit{Canonizer}(c') = \mathit{Canonizer}(c'')$}
+\QuaternaryInfC{$q_2 \circ q_1: c \Leftrightarrow c'''$}
+\end{prooftree}
+
+\begin{align*}
+  \llbracket q \rrbracket.get &= \llbracket q_2 \rrbracket.get\circ \llbracket
+  q_1 \rrbracket.get, \text{ and }\\
+  \llbracket q \rrbracket.put &= \llbracket q_1 \rrbracket.put \circ \llbracket
+  q_2 \rrbracket.put
+  \end{align*}
+
+This derivation rule says that the composition $q_2 \circ q_1$ is well defined
+if and only if the intermediary QREs $c''$ and $c'''$ define the same
+equivalence relation on the same regular language. The condition
+$\mathcal{L}(W(c')) = \mathcal{L}(W(c''))$ says that the intermediary language
+is the same on both sides, while the condition $K(c') \equiv^d K(c'')$ says that
+the kernel regular expressions are equivalent via the star-semiring axioms.
+Finally, the condition $\mathit{Canonizer}(c') = \mathit{Canonizer}(c'')$ says
+that $c'$ and $c''$ define the same equivalence relation on $\mathcal{L}(W(c'))=
+\mathcal{L}(W(c''))$.
+
+  By the induction hypothesis, there exist bijective lenses
+  $\ell_1 :
+  K(c) \Leftrightarrow K(c')$ and $\ell_2 : K(c'') \Leftrightarrow K(c''')$ such
+  that
+  \begin{align*}
+\llbracket q_1 \rrbracket.get &= \llbracket \ell_1 \rrbracket \circ
+\canonizer(c)\\
+\llbracket q_1 \rrbracket.put &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
+\canonizer(c')
+\end{align*}
+and
+\begin{align*}
+\llbracket q_2 \rrbracket.get &= \llbracket \ell_2 \rrbracket \circ
+Canonize(c')\\
+\llbracket q_2 \rrbracket.put &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
+Canonize(c'')
+\end{align*}
+Consequently,
+\begin{align*}
+\llbracket q_2 \rrbracket.get \circ \llbracket q_1 \rrbracket.get &=
+(\llbracket \ell_2 \rrbracket \circ \canonizer(c')) \circ (\llbracket \ell_1
+\rrbracket \circ \canonizer(c))\\
+&= \llbracket \ell_2 \rrbracket \circ (\canonizer(c') \circ \llbracket \ell_1
+\rrbracket) \circ \canonizer(c)\\
+&= (\llbracket \ell_2 \rrbracket \circ \llbracket \ell_1 \rrbracket) \circ
+\canonizer(c)\\
+&= \llbracket \ell_2  \circ  \ell_1 \rrbracket \circ
+\canonizer(c)
+\end{align*} 
+We are permitted to claim the third step from the second since
+$\mathit{Canonizer(c')}$ is the identity function on $K(c'')$ which is
+equivalent to $K(c')$ via the star semiring axioms by assumption.
+
+A similar argument shows that 
+$$\llbracket q_1 \rrbracket.put \circ \llbracket q_2 \rrbracket.put =
+\llbracket \ell_2  \circ  \ell_1 \rrbracket^{-1} \circ
+\canonizer(c)$$
+
+The other cases of the proof are similar in that they follow from a
+straightforward application of the induction hypothesis followed by unrolling
+the equations that give the denotation for QRE lenses.
+\iffalse
 \begin{proof}
@@
 
+\fi
 \section{Synthesizing QRE Lenses}
@@
     \centering
-    %\includegraphics{generated-graphs/times_opt}
+    \includegraphics{generated-graphs/times_opt}
     \caption{}
@@
   \begin{subfigure}[b]{.49\textwidth}
-    %\includegraphics{times_new.eps}
+    \includegraphics{generated-graphs/times_new.eps}
     \caption{}

### 2017-10-24 papers/QuotientLensSynth/main.tex
starting to restructure the intro

@@
 \begin{abstract} 
-
 {\em Lenses}~\cite{Lenses} specify bidirectional 
-transformations between different data formats.  Writing a single lens
+transformations between different data formats.
+\saz{The old abstract read more like an introduction, so we modified it to be
+  the ``new'' introduction.}
+\end{abstract}
+
+\section{Introduction}
+A {\em lens} $\ell$ is a {\em bidirectional transformation} between a source
+type $S$ and a target type $T$ in which the forward transformation
+$\ell.\get$ and the backward transformation $\ell.\lput$ satisfy certain
+round-tripping laws such as invertibility.
+ Writing a single lens
 rather than two separate uni-directional transformations saves both
@@
 transformations interact well.
-%
+
 {\em Quotient lenses}~\cite{quotientlenses} are a generalization of lenses
@@
 and ordering of fields can be ignored.
-
-\smallskip
 
@@
 
-\smallskip
-
-In this paper, we address both of these issues.  First, we propose a novel
-language of {\em Quotient Regular Expressions}, or QREs, which allows
-programmers to specify a broad class of equivalence relations over regular
-languages using a compact and convenient notation in which the regular
-language and its equivalence classes are captured by a single expression.
-We then define {\em QRE lenses}, a class of bijective quotient lenses that
-are typed using QREs, and we show how to {\em synthesize} QRE lenses from their
-QRE types and a set of input-output examples.  We demonstrate the
-practicality of this approach by synthesizing QRE lenses between several
-real-world data formats from the {\tt data.gov} database.
-\end{abstract}
-
-\section{Introduction}
-A {\em lens} $\ell$ is a {\em bidirectional transformation} between a source
-type $S$ and a target type $T$ in which the forward transformation
-$\ell.\get$ and the backward transformation $\ell.\lput$ satisfy certain
-round-tripping laws such as invertibility.
+This paper addresses both of these issues.  First, in Section~\ref{example} we
+explain the difficulties of programming with quotient lenses and show, by way of
+example, how our novel language of {\em Quotient Regular Expressions}, or QREs
+addresses those difficulties.  Quotient regular expressions allow programmers to
+specify a broad class of equivalence relations over regular languages using a
+compact and convenient notation in which the regular language and its
+equivalence classes are captured by a single expression.  We then define {\em
+  QRE lenses}, a class of bijective quotient lenses that are typed using QREs,
+and we show how to {\em synthesize} QRE lenses from their QRE types and a set of
+input-output examples.  We demonstrate the practicality of this approach by
+synthesizing QRE lenses between several real-world data formats from the {\tt
+  data.gov} database.
+
+\saz{There is still some redundancy between the following contributions list and
+  the paragraph above.}
+
+In summary, our main contributions are:
+\begin{enumerate}
+\item We introduce a novel language of {\em Quotient Regular Expressions}
+(QREs) that provide compact, convenient notation for a natural and useful
+class of equivalence relations on regular languages (Section~\ref{QRE}).
+\item We define a language of {\em QRE lenses}, a subset of bijective quotient
+lenses whose types are given by QREs.  Our main technical contribution is a
+normal form for QRE lenses and a proof that normal forms are
+closed under left quotienting, right quotienting, composition, and regular
+operators (Section~\ref{QRE-lenses}).
+\item Using this normal form, we reduce the problem of {\em synthesizing}
+QRE lenses from their QRE types and a set of examples to the previously studied
+problem of synthesizing bijective lenses between regular languages 
+(Section~\ref{synth}).
+\item We extend the Boomerang {\em implementation} with QREs, QRE lens operators,
+and QRE lens synthesis and demonstrate its applicability by using it to
+synthesize QRE lenses between several real-world data formats from the
+{\tt data.gov} database (Section~\ref{impl}).
+\end{enumerate}
+Sections~\ref{relwork} and~\ref{concl} discuss related and future work.
+
+\section{Lenses and Quotient Regular Expressions by Example}
 
@@
 the equivalence class of \cd{end_example}.
-
-In summary, our main contributions are:
-\begin{enumerate}
-\item We introduce a novel language of {\em Quotient Regular Expressions}
-(QREs) that provide compact, convenient notation for a natural and useful
-class of equivalence relations on regular languages (Section~\ref{QRE}).
-\item We define a language of {\em QRE lenses}, a subset of bijective quotient
-lenses whose types are given by QREs.  Our main technical contribution is a
-normal form for QRE lenses and a proof that normal forms are
-closed under left quotienting, right quotienting, composition, and regular
-operators (Section~\ref{QRE-lenses}).
-\item Using this normal form, we reduce the problem of {\em synthesizing}
-QRE lenses from their QRE types and a set of examples to the previously studied
-problem of synthesizing bijective lenses between regular languages 
-(Section~\ref{synth}).
-\item We extend the Boomerang {\em implementation} with QREs, QRE lens operators,
-and QRE lens synthesis and demonstrate its applicability by using it to
-synthesize QRE lenses between several real-world data formats from the
-{\tt data.gov} database (Section~\ref{impl}).
-\end{enumerate}
-Sections~\ref{relwork} and~\ref{concl} discuss related and future work.
 

### 2017-10-26 papers/QuotientLensSynth/main.tex
Added squash to example. Need to make QRE section better.

@@
 \begin{abstract} 
-{\em Lenses}~\cite{Lenses} specify bidirectional 
-transformations between different data formats.
 \saz{The old abstract read more like an introduction, so we modified it to be
@@
 $\ell.\get$ and the backward transformation $\ell.\lput$ satisfy certain
-round-tripping laws such as invertibility.
- Writing a single lens
-rather than two separate uni-directional transformations saves both
-keystrokes and bugs, by reducing redundancy and by automatically
-guaranteeing strong ``lens laws'' ensuring that the forward and backward
-transformations interact well.
+round-tripping laws such as invertibility. Lenses are advantageous
+because writing a single lens rather than two separate uni-directional
+transformations saves both keystrokes and bugs, by reducing redundancy and by
+automatically guaranteeing strong ``lens laws'' ensuring that the forward and
+backward transformations interact well.
 
@@
 currently offering quotient lenses.  One reason for this is that there is
-some annoying redundancy in the quotient lens notation: The programmer must
-separately specify (1) the data formats on each side of the lens, (2) the
+some annoying redundancy in the quotient lens notation: The programmer often
+has to separately specify (1) the data formats on each side of the lens, (2) the
 equivalence relations over them, and (3) a lens between the canonical
-representatives of the equivalence classes, even though, typically, these
-three share a great deal of structure.  Moreover, writing the lens between
-the equivalence classes can be challenging due to the rather fiddly, mainly
-due to ``non-ambiguity'' conditions in the lens well-formedness rules.
+representatives of the equivalence classes, even though, typically, these three
+share a great deal of structure.  Moreover, writing the lens between the
+equivalence classes can be challenging due to the rather fiddly, mainly due to
+``non-ambiguity'' conditions in the lens well-formedness rules.
 
@@
 example, how our novel language of {\em Quotient Regular Expressions}, or QREs
-addresses those difficulties.  Quotient regular expressions allow programmers to
-specify a broad class of equivalence relations over regular languages using a
-compact and convenient notation in which the regular language and its
-equivalence classes are captured by a single expression.  We then define {\em
-  QRE lenses}, a class of bijective quotient lenses that are typed using QREs,
-and we show how to {\em synthesize} QRE lenses from their QRE types and a set of
-input-output examples.  We demonstrate the practicality of this approach by
-synthesizing QRE lenses between several real-world data formats from the {\tt
-  data.gov} database.
+and QRE lenses address these difficulties.  Quotient regular expressions allow
+programmers to specify a broad class of equivalence relations over regular
+languages using a compact and convenient notation in which the regular language
+and its equivalence classes are captured by a single expression.  We then
+define {\em QRE lenses}, a class of bijective quotient lenses that are typed
+using QREs, and we show how to {\em synthesize} QRE lenses from their QRE types
+and a set of input-output examples.  We demonstrate the practicality of this
+approach by synthesizing QRE lenses between several real-world data formats
+from the {\tt data.gov} database.
 
@@
 \item We define a language of {\em QRE lenses}, a subset of bijective quotient
-lenses whose types are given by QREs.  Our main technical contribution is a
-normal form for QRE lenses and a proof that normal forms are
-closed under left quotienting, right quotienting, composition, and regular
-operators (Section~\ref{QRE-lenses}).
+lenses whose types are given by QREs. Our main technical contribution in this
+paper is a proof that each QRE lenses $q$ defines the same transformation as a
+QRE lens $q'$ which is in normal form. (Section~\ref{QRE-lenses}).
 \item Using this normal form, we reduce the problem of {\em synthesizing}
@@
 
-\section{Lenses and Quotient Regular Expressions by Example}
-
-An example of a bidirectional transformation which we shall revisit
-severally in this paper is a lens which converts citation records such as
+\section{Quotient Regular Expressions and Lenses by Example}
+Consider a bidirectional transformation which converts citation records such as
 \begin{verbatim}
@@
 
-Lenses have the advantage of saving the programmer from having to
-write two seperate transformations rather than one. Also, lenses offer strong
-guarantees about how the two transformations interact since lenses are
-constrained by lens laws.
-
-However, the constraints imposed by the lens laws are often too strict. For
-example, when a record may be represented in many ways, perhaps because
-whitespace characters are allowed to occur arbitrarily in certain parts of the
-data, deriving a lens which obeys the lens laws in such cases can be daunting
-and frustrating.
-
-As a solution to this problem, Foster et al proposed {\em quotient lenses},
-which are a generalization of lenses in which the source and target data are
-considered modulo an equivalence relation. Quotient lenses, which are
-implemented as a refinement of the bidirectional string processing language
-Boomerang, are both simpler and more flexible, since whole equivalence classes
-of data values can be handled all at once and inessential details like
-whitespace and ordering of fields can be ignored.
-
-However many complicated equivalence relations are difficult to express in
-Boomerang. For example, suppose that the \bibtex{} to EndNote
-transformation which we introduced is required to allow for whitespace to occur
-arbitrarily between the words in the label, author and title fields as well as
-in between them, and ignore the order in which the author and title fields
-occur in each record.
-These rules allow for the \bibtex{} citations
+We wish for this transformation to have the same behaviour on records which
+differ only in ``inessential'' details. First, the transformation
+should allow whitespace characters to occur arbitrarily between the words in the
+label, author and title fields as well as in between them. Second, an
+author's name can have the surname followed by a comma and then the other
+names separated by arbitrary whitespace, or just all the names separated by
+arbitrary whitespace (for example, the names ``Conway, J. H.'' and ``J. H.
+Conway'' are equivalent). Lastly, the transformation should also ignore the
+order in which the author and title fields occur in each citation record.
+
+For example, with these rules, the following \bibtex{} citations should be
+considered the same by the lens,
 \begin{verbatim}
@@
     }
-\end{verbatim}
-and
-\begin{verbatim}
+
     @Book{     conway,
@@
                  and Finite Machines},
-           Author = "J. H. Conway",     }
+           Author = "Conway, J. H.",     }
 \end{verbatim}
-to be considered the same, with the EndNote records having similar rules
-defined on them.
+with the EndNote records having similar rules defined on them.
 
 While it is possible to express the equivalence on the \bibtex{} data above in
-Boomerang, actually doing so is difficult. The difficulty arises from the fact
-that the entire equivalence is composed of three separate equivalences which
-nest and interact with each other in a way that makes them difficult to express
-with the standard Boomerang combinators. Moreover, many of the equivalences
-expressible in Boomerang, such as Boomerang's permutation combinator, require
-that the programmer first specify the data itself in form of a regular
-expression before specifying the equivalence relation on the data. This leads to
-verbose code that is difficult to understand.
-
-To address this issue, we propose the language of {\em Quotient Regular
-Expressions} or QREs, which express a broad class of equivalence relations
-directly on regular languages using a compact and convenient notation. For
-example, with QREs we can specify \bibtex{} records as shown in
-Figure~\ref{fig:example-qre}.
+Boomerang, actually doing so is a challenge. This is because the entire
+equivalence is composed of three separate equivalences which nest and interact
+with each other in a way that makes them difficult to express with the standard
+Boomerang combinators. Moreover, many of the equivalences expressible in
+Boomerang require that the programmer first specify the data itself in form of
+a regular expression before specifying the equivalence relation on the data.
+This leads to verbose, repetitive code.
+
+To address this issue, we propose {\em Quotient Regular Expressions} or QREs,
+which express a broad class of equivalence relations directly on regular
+languages using a compact and convenient notation. For example, with QREs we
+can specify \bibtex{} records as shown in Figure~\ref{fig:example-qre}.
 
@@
 \begin{lstlisting}
-let wsp = [ \n\t\r]$^+$
-let wsp_to_space = project wsp $\to$ " "
-let valid_name = name | (name . wsp_to_space . (name . wsp_to_space)$^*$ . name)
-let wsp_to_and = project (wsp . "and" . wsp) $\to$ " and "
+let wsp = [ \n\t\r]
+let wsp_to_space = project wsp+ $\to$ " "
+let wsp_to_space_epsilon = project wsp* $\to$ ""
+let space_name  = name | (name . wsp_to_space . (name . wsp_to_space)* . name)
+let wsp_to_comma = wsp_to_space_epsilon . "," . wsp_to_space
+let comma_name = name | (name . wsp_to_comma . (name . wsp_to_space)* . name)
+let space_to_comma = synth comma_name <=> space_name using {("Conway, J. H.", "J. H. Conway")}
+let valid_name = squash comma_name -> space_name using (get space_to_comma)
+let wsp_to_and = project (wsp+ . "and" . wsp+) $\to$ " and "
 let bib_names = valid_name | (valid_name . wsp_to_and . (valid_name . wsp_to_and)$^*$ . valid_name)
@@
 let bib_title = "title = {" . title . "}"
-let bib_perm = perm (label, bib_author, bib_title) with (project ("," .wsp) $\to$ ",\n")
-let bibtex = "@book{" . bib_perm . "}"
+let bib_perm = perm (label, bib_author, bib_title) with (project ("," .wsp+)
+$\to$ ",\n") let bibtex = "@book{" . bib_perm . "}"
 \end{lstlisting}
@@
 
-The expression \cd{project [\n\t\r]}$^+$ $\to$ \cd{" "} is a QRE that matches any 
-number of whitespace characters, and then considers all strings of whitespace
-characters as equivalent with a single space as the canonical representative. 
-Another QRE in the example above is the
-
+The expression \cd{project [ \\n\\t\\r]}$^+$ $\to$ \cd{" "} is a QRE that
+matches any number of whitespace characters, and then considers all strings of
+whitespace characters as equivalent with a single space as the canonical
+representative. Another QRE in the example above is the expression
+
+\begin{center}
+\begin{lstlisting}
+squash comma_name -> space_name using (get space_to_comma)
+\end{lstlisting}
+\end{center}
+
+\noindent which considers the \cd{comma_name} representation of the author's
+name to be equivalent to the \cd{space_name} representation, with \cd{get
+space_to_comma} giving the function that transforms names in the \cd{comma_name}
+representation to names in the \cd{space_name} representation.
+
+The third example of a QRE in the example is the expression
 \begin{center}
@@
 
-While QREs make it easier to write quotient lenses by specifying the data and
-the equivalence relation on the data in one expression, the programmer still
+Now QREs make it easier to write quotient lenses by specifying the data and
+the equivalence relation on the data in one expression, but the programmer still
 has to manually derive the underlying lens which maps the equivalence classes
 of the source data to the equivalence classes of the target data. This is
-itself tedious and error-prone task, as the code snippet below illustrates:
+itself a tedious and error-prone task, as the code snippet below illustrates:
 
@@
 QRE lenses lenses are the same as bijective lenses with ``canonizers at the
-edges'' as in Figure~\ref{fig:attheedges}, and are typed using QREs. In the
-forward direction, a QRE lens $q$ uses the source QRE $c$ to compute a unique
-representative for the data modulo the equivalence relation defined by $c$ and
-then applies the $\get$ function of a bijective lens $\ell$ to this
-representative. In the backward direction, $q$ operates similarly,  but using
-the view QRE $c'$ and the $\lput$ function of $\ell$.
+edges'' as in Figure~\ref{fig:attheedges}, and are typed using QREs - every QRE
+lens $q$ has a type $q: c \Leftrightarrow c'$ where $c$ and $c'$ are QREs.
+In the forward direction, a QRE lens $q: c \Leftrightarrow c'$ uses the source
+QRE $c$ to compute a unique representative for the data modulo the equivalence
+relation defined by $c$ and then applies the $\get$ function of a bijective
+lens $\ell$ to this representative. In the backward direction, $q$ operates
+similarly,  but using the view QRE $c'$ and the $\lput$ function of $\ell$.
 
@@
       function. With QREs the kernel language is always contained in the whole
-      language so we just use the inclusion as the choose function. Also,
-      perhaps it would be clearer if we said source structures and target
-      structres instead of all structures?}}
+      language so we just use the inclusion as the choose function.}}
   \label{fig:attheedges}
@@
 canonizers a the edges. Our main technical contribution in this paper is a
-proof that QRE lenses are closed under left quotienting, right quotienting,
-composition, and regular combinators.\bcp{If this is our main technical
-  contribution, then it deserves more than a sentence in the intro!  E.g.,
-  can we give an example?} This proof will allow justify our
-approach for synthesizing quotient lenses so that the programmer can simply
-type
+proof that each QRE lenses $q$ defines the same transformation as a QRE lens
+$q'$ which is in normal form.
+
+For example, suppose \cd{q_1 : bibtex <=> endnote} is a QRE lens that maps
+between \bibtex{} and EndNote formats, and that \cd{q_2 : endnote <=> xml}
+is a QRE lens that maps between EndNote and XML formats. The composition 
+\cd{(q_1 ; q2) : bibtex <=> xml} of \cd{q_1} and \cd{q_2} is not in normal form
+due to the presence of canonizers on each side of the intermediary EndNote
+format:
+
+\sam{TODO: Put diagram of composition here}
+
+\noindent However, while \cd{q_1 ; q_2} is not in normal form, there is another
+QRE lens \cd{q : bibtex <=> xml} of the same type as \cd{q_1 ; q_2} such that
+\cd{q_1 ; q_2} and \cd{q} define the same transformation. This proof will allow
+justify our approach for synthesizing quotient lenses so that the programmer
+can simply type
 
@@
 \end{lstlisting}
-\noindent to synthesize a quotient lens that maps between \cd{bibtex} and
+\noindent to synthesize a quotient lens that maps between the \cd{bibtex} and
 \cd{endnote} QREs, and sends the equivalence class of \cd{bib_bexample} to

### 2017-10-27 papers/QuotientLensSynth/main.tex
Made changes to QRE section

@@
 \newcommand{\re}[1]{\ensuremath{\mathtt{#1}}}
-\newcommand{\squash}[3]{\ensuremath{\mathit{squash} \; (#1, #2, #3)}}
+\newcommand{\project}[2]{\ensuremath{\mathit{project} \; #1 \mapsto #2}}
+\newcommand{\squash}[3]{\ensuremath{\mathit{squash} \; #1 \rightarrow #2
+\mathit{using} \; #3}}
 \newcommand{\perm}[2]{\ensuremath{\mathit{perm}\; (#1)\; \mathit{with}\; #2}}
@@
 \newcommand{\get}{\ensuremath{\mathit{get}}}
+\newcommand{\semicolon}{\ensuremath{\; ; \;}}
 \newcommand{\lput}{\ensuremath{\mathit{put}}}
@@
 let wsp_to_space = project wsp+ $\to$ " "
-let wsp_to_space_epsilon = project wsp* $\to$ ""
 let space_name  = name | (name . wsp_to_space . (name . wsp_to_space)* . name)
-let wsp_to_comma = wsp_to_space_epsilon . "," . wsp_to_space
-let comma_name = name | (name . wsp_to_comma . (name . wsp_to_space)* . name)
+let comma_name = name | (name . "," . wsp_to_space . (name . wsp_to_space)* . name)
 let space_to_comma = synth comma_name <=> space_name using {("Conway, J. H.", "J. H. Conway")} 
@@
 let bib_title = "title = {" . title . "}" 
-let bib_perm = perm (label, bib_author, bib_title) with (project ("," .wsp+)
-$\to$ ",\n") let bibtex = "@book{" . bib_perm . "}"
+let bib_perm = perm (label, bib_author, bib_title) with (project ("," .wsp+) $\to$ ",\n")
+let bibtex = "@book{" . bib_perm . "}"
 \end{lstlisting}
@@
 
-\subsection{Overview}
-
-We have already encountered the \cd{project R} $\to$ \cd{s} QRE
-which sends all elements of the language described by the regular expression
-$R$ to the canonical representative $s$, and the
-
-\noindent \cd{perm (q_1, ..., q_n) with q} QRE which sends all permutations of
-the QREs $q_1, \ldots, q_n$ separated by the QRE $q$ to the canonical
-QRE $q_1 \cdot q \cdot \ldots \cdot q \cdot q_n$.
-
-Now consider a variant of the \bibtex{} to EndNote example transformation,
-where, the only quotienting occurs in the author field, where the author's name
-can have the last name followed by a comma and then the rest of the name, or
-remain as before where the names are separated by a space.
-
-For example, these two records should be considered equivalent:
-\begin{verbatim}
-@Book {conway,
-  Author = "Conway, J. H.",
-  Title = {Regular Algebra and Finite Machines},
-}
-\end{verbatim}
-\begin{verbatim}
-@Book {conway,
-  Author = "J. H. Conway",
-  Title = {Regular Algebra and Finite Machines},
-}
-\end{verbatim}
-
-The equivalence described is one in which there is a bijection between 
-the regular expressions \cd{comma_name} and \cd{space_name} given by,
+\subsection{Syntax of Quotient Regular Expressions}
+The language of Quotient Regular Expressions (QREs) is given by the following
+grammar:
+\begin{align*}
+q := \; &R \sep \project{R}{s} \sep \squash{q}{q'}{f} \sep
+\perm{q_1, \ldots, q_n}{q} \;  | \; \normalize{R}{R'}{f}\\
+&q \semicolon q' \sep q \cdot q' \sep (q \sep q') \sep q^*,
+\end{align*}
+where $R$ ranges over regular expressions, $f$ ranges over functions between
+regular languages, and $s$ ranges over character strings.
+
+\subsection{Semantics of Quotient Regular Expression}
+In the \bibtex{} to EndNote example, we encountered the $\project{R}{s}$,
+
+\noindent $\squash{q}{q'}{f}$ and $\perm{q_1, \ldots, q_n}{q}$ QREs, as well as
+the regular operators applied to these, as in
+
 \begin{lstlisting}
-let comma_name : regexp = name | (name . ", " . name . (" " . name)*)
-
-let space_name : regexp = name | (name . " " . (name . wsp_to_space)* . name)
-
+let space_name  = name | (name . wsp_to_space . (name . wsp_to_space)* . name)
 \end{lstlisting}
 
-\noindent with the lens \cd{comma_to_space} mapping between them given by 
+The $q \semicolon q'$ QRE is a compostion combinator used to successively apply
+two equivalence relations to a regular language. For example, in the \bibtex{}
+to EndNote example, when defining the \cd{valid_name} QRE which maps
+\cd{comma_name} into \cd{space_name}, we first defined the equivalence
+relations on the \cd{comma_name} and \cd{space_name} representations before
+defining the \cd{valid_name} QRE. Using the composition combinator, we could
+instead have defined \cd{valid_name} as follows:
+
 \begin{lstlisting}
-let comma_to_space : (lens in space_name <=> comma_name) =
-lens_swap 
-(name . del " " . (ins " " . name . del " ")* ) (name . ins ", ") || name 
+(* Define a QRE which matches sends all comma names to a space name equivalent *)
+let wsp = [ \n\t\r] 
+let whole_space_name  = name | (name . wsp+ . (name . wsp+)* . name) 
+let whole_comma_name = name | (name . "," . wsp+ . (name . wsp+)* . name) 
+let whole_space_to_comma = 
+synth whole_comma_name <=> whole_space_name using {("Conway, J. H.", "J. H. Conway")} 
+let whole_valid_name = 
+squash whole_comma_name -> whole_space_name using (get whole_space_to_comma)
+
+(* Define QRE that sends space names to their canonical representation *)
+let wsp_to_space = project wsp+ $\to$ " "
+let space_name  = name | (name . wsp_to_space . (name . wsp_to_space)* . name)
+
+(* Compose the two QREs *)
+let valid_name = whole_valid_name ; space_name
 \end{lstlisting}
 
-The $\squash{R}{R'}{f}$ QRE is designed to make it easy to express such
-equivalences easily. Specifically, given regular expressions $R$ and $R'$, and
-a function $f : \mathcal{L}(R) \longrightarrow \mathcal{L}(R')$, the
-$\squash{R}{R'}{f}$ QRE expresses the equivalence relation on $\mathcal{L}(R \;
-R')$ which sends each element $r \in \mathcal{L}(R)$ to the canonical element
-$f(r) \in \mathcal{L}(R')$, and sends each element $r' \in \mathcal{L}(R')$ to
-itself:
-
-\begin{lstlisting}
-let squash_author = squash space_name $\to$ comma_name using comma_to_space
-
-let author : canonizer = "author = \"" . squash_author . "\""
-\end{lstlisting}
-
-In this example, the \cd{squash_author} QRE expresses the fact that the
-\cd{space_name} and \cd{comma_name} representations of the authors' name
-are equivalent, with the \cd{comma_name} representation chosen as the
-canonical representation, and the lens \cd{comma_to_space} giving
-the canonizing function which converts data in the \cd{space_name} format
-to its \cd{comma_name} representation.
-
-The $\squash{R}{R'}{\ell}$ has the added advantage that if the transformation
-from $R$ to $R'$ can be expressed by a bijective lens that is consistent with a set
-of examples, then the programmer can synthesize the lens from $R$ to $R'$
-because of our previous work where we demonstrated how to synthesize bijective
-lenses from a pair of regular expressions and a set of input-output example
-pairs:
-
-\begin{lstlisting}
-let comma_to_space : (lens in space_name <=> comma_name) = 
-synth space_name <=> comma_name using {("J. H. Conway", "Conway, J. H.")}
-\end{lstlisting}
-
-The regular combinators for QREs i.e. $\cdot$ (concatenation), $|$ (union) and
-$*$ (iteration) behave like the regular combinators for regular languages. For
-instance, if the QRE for matching a \bibtex{} record is named
-\cd{bibtex}, then the QRE for matching two records is 
-\cd{bibtex . bibtex}, the QRE for matching one or two records is
-\cd{ bibtex | (bibtex . bibtex)} and the QRE for matching any number of QRE
-records \cd{bibtex*}
-
-The $\normalize{R}{R'}{f}$ expresses the equivalence relation which sends each
-element in $\mathcal{L}(R)$ to its canonical representation $f(r) \in
-\mathcal{L}(R')$. The equivalence relation defined by
-$\normalize{R}{R'}{f}$ is thus the equivalence relation on
+The final combinator is the  $\normalize{R}{R'}{f}$ combinator, which expresses
+the equivalence relation which sends each element in $\mathcal{L}(R)$ to its
+canonical representation $f(r) \in \mathcal{L}(R')$. The equivalence relation
+defined by $\normalize{R}{R'}{f}$ is thus the equivalence relation on
 $\mathcal{L}(R)$ defined by the {\em fibres} of $f$ (i.e. $r \sim r'$ if and
@@
 difficult to express, or which cannot be expressed, using the other
-combinators. However, checking that the canonizing function $f$ is
+combinators.
+
+However, checking that the canonizing function $f$ is
 surjective and idempotent is undecidable, thus we do not use the
@@
   $\canonizer(q)$.
-\subsection{Syntax of QREs}
-The language of Quotient Regular Expressions (QREs) is given by the following
-grammar:
-\begin{align*}
-q := \; &R \sep R \mapsto s \sep \squash{R}{R'}{f} \sep
-\perm{q_1, \ldots, q_n}{q} \;  | \; \normalize{R}{R'}{f}\\
-&q' \circ q \sep q \cdot q' \sep (q \sep q') \sep q^*,
-\end{align*}
-where $R$ ranges over regular expressions, $f$ ranges over functions between
-regular languages, and $s$ ranges over character strings.
-
-\subsection{Semantics of QREs}
-\subsubsection{Preliminaries}
-Let $q, q'$ be QREs. When applying the regular combinators to QREs, we require
-that if a string $s$ matches any of the regular expressions $W(q) \cdot W(q')$,
-$W(q) \sep W(q')$, $W(q)^*$, $K(q) \cdot K(q')$,
-$K(q) \sep K(q')$, $K(q)^*$, then $s$ matches that regular expression in
-only one way. This unambiguity condition is called {\em strong unambiguity},
-and is necessary, firstly because it ensures that the canonizing function of a
-QRE is well-defined.
+
+Before giving the formal semantics for QREs, we address the issue of {\em
+ambiguity} which comes up when defining QREs. If $q, q'$ are QREs, then
+when applying the regular combinators to QREs, we require that if a string $s$
+matches any of the regular expressions $W(q) \cdot W(q')$, $W(q) \sep W(q')$,
+$W(q)^*$, $K(q) \cdot K(q')$, $K(q) \sep K(q')$, $K(q)^*$, then $s$ matches
+that regular expression in only one way. This unambiguity condition is called
+{\em strong unambiguity}, and is necessary, firstly because it ensures that the
+canonizing function of a QRE is well-defined.
 

### 2017-11-02 papers/QuotientLensSynth/main.tex
some edits and comments in sections 2 and 3

@@
 \newcommand{\re}[1]{\ensuremath{\mathtt{#1}}}
-\newcommand{\project}[2]{\ensuremath{\mathit{project} \; #1 \mapsto #2}}
-\newcommand{\squash}[3]{\ensuremath{\mathit{squash} \; #1 \rightarrow #2
-\mathit{using} \; #3}}
-\newcommand{\perm}[2]{\ensuremath{\mathit{perm}\; (#1)\; \mathit{with}\; #2}}
-\newcommand{\normalize}[3]{\ensuremath{\mathit{normalize} \; (#1, #2, #3)}}
-\newcommand{\sep}{\ensuremath{\; | \;}}
-\newcommand{\canonizer}{\ensuremath{\mathit{Canonizer}}}
+\newcommand{\kw}[1]{\ensuremath{\mathsf{#1}}}
+\newcommand{\project}[2]{\ensuremath{\kw{project} \; #1 \mapsto #2}}
+\newcommand{\squash}[3]{\ensuremath{\kw{squash} \; #1 \rightarrow #2
+\kw{using} \; #3}}
+\newcommand{\perm}[2]{\ensuremath{\kw{perm}\; (#1)\; \kw{with}\; #2}}
+\newcommand{\normalize}[3]{\ensuremath{\kw{normalize} \; (#1, #2, #3)}}
+\newcommand{\sep}{\ensuremath{\ | \ }}
+\newcommand{\canonizer}{\ensuremath{\kw{Canonizer}}}
 \newcommand{\bibtex}{\textsc{Bib}\TeX{}}
-\newcommand{\get}{\ensuremath{\mathit{get}}}
+\newcommand{\get}{\ensuremath{\kw{get}}}
 \newcommand{\semicolon}{\ensuremath{\; ; \;}}
-\newcommand{\lput}{\ensuremath{\mathit{put}}}
-\newcommand{\create}{\ensuremath{\mathit{create}}}
+\newcommand{\lput}{\ensuremath{\kw{put}}}
+\newcommand{\create}{\ensuremath{\kw{create}}}
 \newcommand{\eqrel}[1]{\ensuremath{\equiv_{#1}}}
@@
 share a great deal of structure.  Moreover, writing the lens between the
-equivalence classes can be challenging due to the rather fiddly, mainly due to
+equivalence classes can be challenging due to the rather fiddly typechecking, mainly due to
 ``non-ambiguity'' conditions in the lens well-formedness rules.
 
-This paper addresses both of these issues.  First, in Section~\ref{example} we
+This paper addresses both of these issues.  First, in Section~\ref{sec:example} we
 explain the difficulties of programming with quotient lenses and show, by way of
@@
 \section{Quotient Regular Expressions and Lenses by Example}
+\label{sec:example}
+
 Consider a bidirectional transformation which converts citation records such as
@@
 For example, with these rules, the following \bibtex{} citations should be
-considered the same by the lens,
+considered the same by the lens \saz{``the lens'' hasn't been introduced to the
+  reader yet},
 \begin{verbatim}
@@
 \end{verbatim}
-with the EndNote records having similar rules defined on them.
+with the EndNote records having similar rules \saz{similarly, ``rules'' haven't been
+  mentioned yet, do you mean similar ``equivalence''} defined on them.
 
@@
 Boomerang, actually doing so is a challenge. This is because the entire
-equivalence is composed of three separate equivalences which nest and interact
+equivalence is composed of three separate equivalences \saz{explicitly mention
+  whitespace and permutation?} which nest and interact
 with each other in a way that makes them difficult to express with the standard
@@
 a regular expression before specifying the equivalence relation on the data.
+\saz{Something like this instead?:  Moreover, Boomerang requires that the
+  programmer first specify the data itself in the form of a regular expression
+  before she can define the equivalence relation on that data.}
 This leads to verbose, repetitive code.
 
-To address this issue, we propose {\em Quotient Regular Expressions} or QREs,
+To address this issue\saz{these issues? these problems with Boomerang?}, we propose {\em Quotient Regular Expressions} or QREs,
 which express a broad class of equivalence relations directly on regular
@@
 let wsp = [ \n\t\r]
-let wsp_to_space = project wsp+ $\to$ " "
-let space_name  = name | (name . wsp_to_space . (name . wsp_to_space)* . name)
-let comma_name = name | (name . "," . wsp_to_space . (name . wsp_to_space)* . name)
-let space_to_comma = synth comma_name <=> space_name using {("Conway, J. H.", "J. H. Conway")} 
-let valid_name = squash comma_name -> space_name using (get space_to_comma)
-let wsp_to_and = project (wsp+ . "and" . wsp+) $\to$ " and " 
+let wsp_to_space   = project wsp$^+$ $\to$ " "
+let space_name     = name | (name . wsp_to_space . (name . wsp_to_space)$^*$ . name)
+let comma_name   = name | (name . "," . wsp_to_space . (name . wsp_to_space)$^*$ . name)
+let space_to_comma = synth comma_name $\Leftrightarrow$ space_name using {("Conway, J. H.", "J. H. Conway")} 
+let valid_name = squash comma_name $\to$ space_name using (get space_to_comma)
+let wsp_to_and = project (wsp$^+$ . "and" . wsp$^+$) $\to$ " and " 
 let bib_names  = valid_name | (valid_name . wsp_to_and . (valid_name . wsp_to_and)$^*$ . valid_name) 
@@
 let bib_title = "title = {" . title . "}" 
-let bib_perm = perm (label, bib_author, bib_title) with (project ("," .wsp+) $\to$ ",\n")
+let bib_perm  = perm (label, bib_author, bib_title) with (project ("," .wsp$^+$) $\to$ ",\n")
 let bibtex    = "@book{" . bib_perm . "}"
@@
 
-The expression \cd{project [ \\n\\t\\r]}$^+$ $\to$ \cd{" "} is a QRE that
-matches any number of whitespace characters, and then considers all strings of
-whitespace characters as equivalent with a single space as the canonical
-representative. Another QRE in the example above is the expression
+\saz{Throughout these examples, refer to the definitions in Figure 1.}  In the
+figure, the QRE \cd{wsp_to_space}, which is defined to be \cd{project wsp}$^+$
+$\to$ \cd{" "}, matches any non-zero number of whitespace characters, but
+considers the matched string of whitespace characters as equivalent with a
+single space as the canonical representative.  \saz{We call the type of QRE a
+\textit{projection} because it projects an entire regular language down to a single
+representative string in the language.}
+
+\saz{
+Another form of QRE is the \textit{squash} QRE, an example of which is given by
+\cd{valid_name}, which is defined to be:
+}
 
@@
 \begin{lstlisting}
-squash comma_name -> space_name using (get space_to_comma)
+squash comma_name $\to$ space_name using (get space_to_comma)
 \end{lstlisting}
@@
 
-\noindent which considers the \cd{comma_name} representation of the author's
-name to be equivalent to the \cd{space_name} representation, with \cd{get
-space_to_comma} giving the function that transforms names in the \cd{comma_name}
-representation to names in the \cd{space_name} representation.
-
-The third example of a QRE in the example is the expression
+\noindent This QRE specifies that the \cd{comma_name} representation of the
+author's name is equivalent to the \cd{space_name} representation.  Here \cd{get
+  space_to_comma} is the function that transforms names in the \cd{comma_name}
+representation to names in the \cd{space_name} representation.  \saz{In this
+  example, \cd{space_to_comma} is itself an automatically \textit{synthesized}
+  lens between the \cd{space_name} and \cd{comma_name} QREs, and we need to use
+  the ``get'' direction of the lens to specify the squash transformation. (We'll
+  say more about synthesis below.)}
+
+The third example of a QRE in the Figure is \cd{bib_perm}, defined as:
 \begin{center}
 \begin{lstlisting}
-perm (label, bib_author, bib_title) with (project ("," . wsp ) $\to$ ",\n")
+perm (label, bib_author, bib_title) with (project ("," . wsp$^+$ ) $\to$ ",\n")
 \end{lstlisting}
@@
 
-\noindent QRE, which specifies that the label, title and author fields can occur
+\noindent This is a \textit{permutation} QRE, which specifies that the label, title, and author fields can occur
 in any order, with the fields separated by a comma followed by arbitrary amounts
@@
 and title fields occur in that order, separated by a comma followed a newline.
-The \cd{perm} QRE allows for easy nesting of permutations. Similarly, it is easy
+Such \cd{perm} QREs allow for easy nesting of permutations. Similarly, it is easy
 to specify any number of fields within the \cd{perm} QRE.
 
-Now QREs make it easier to write quotient lenses by specifying the data and
-the equivalence relation on the data in one expression, but the programmer still
-has to manually derive the underlying lens which maps the equivalence classes
-of the source data to the equivalence classes of the target data. This is
-itself a tedious and error-prone task, as the code snippet below illustrates:
-
+\saz{Given the \textit{projection}, \textit{squash}, and \textit{permutation}
+  QRE primitives, we can build more complex languages using the usual regular
+  expression combinators of alternative choice `\cd{|}', Kleene star `$^*$', and
+  sequential composition ``.''.  As shown in the rest of
+  Figure~\ref{fig:example}, it is straightforward to define both a regular
+  language that accepts \bibtex{} records \textit{and}, simultaneously, an
+  equivalence relation that quotients out irrelevant details.}
+
+\subsection{QRE Lenses}
+QREs make it easier to write quotient lenses by specifying the data and the
+equivalence relation on the data in one expression, but to use them to define a
+lens, the programmer still has to manually derive the underlying lens that maps
+the equivalence classes of the source data to the equivalence classes of the
+target data.
+
+\begin{figure}[t]
 \begin{lstlisting}
@@
 
-let bib_to_end : (lens in bib_tex $\Leftrightarrow$ end_note) =
+let bib_to_end : (lens in bibtex $\Leftrightarrow$ end_note) =
    preamble . label . author_lens . title_lens
 \end{lstlisting}
-
-This lens transforms the \bibtex{} data to the EndNote (and vice versa) data by
-matching the delimiters and separators in a bijective manner. For example,
-the \cd{preamble} lens replaces the \bibtex{} preamble \cd{@book} with the
-EndNote preamble \cd{\%0 Book\n\%F } in the \bibtex{} record. The programmer
-can ill afford to make a mistake with any one of the small but vital details
-of the transformations since even a one character mistake might render the lens
+\caption{An example lens between the \cd{bibtex} and \cd{end_note} QRE languages.}
+\label{fig:example-lens}
+\end{figure}
+
+This is itself a tedious and potentially error-prone task, as the code snippet
+in Figure~\ref{fig:example-lens} illustrates.  The lens shown in the Figure
+transforms \bibtex{} data to EndNote data  (and vice versa) by matching
+the delimiters and separators in a bijective manner. In more detail, the
+\cd{preamble} lens replaces the \bibtex{} preamble \verb|@Book| with the EndNote
+preamble \verb|%0 Book\n%F|. \saz{This next sentence isn't very well-supported
+  by the previous sentence's example, since that doesn't seem too hard.  Maybe
+  explain some of the harder parts too?}  The programmer can ill
+afford to make a mistake with any one of the small but vital details of the
+transformations since even a one character mistake might render the lens
 incorrect. This makes lens programming tedious and error-prone.
 
-Therefore, to address this challenge, we identify a special class of quotient
-lenses which we call {\em QRE lenses}, and show how to {\em
+To address this challenge, we identify a \saz{avoid ``special''} class of quotient
+lenses that \saz{instead of ``which''} we call {\em QRE lenses}, and show how to {\em
 synthesize} QRE lenses from a pair of QREs and a set of input-output
@@
 QRE lenses lenses are the same as bijective lenses with ``canonizers at the
-edges'' as in Figure~\ref{fig:attheedges}, and are typed using QREs - every QRE
+edges'' as in Figure~\ref{fig:attheedges}, and are typed using QREs---\saz{in
+  latex use three dashes --- for this (not one -).  Put no spaces around ``---''}every QRE
 lens $q$ has a type $q: c \Leftrightarrow c'$ where $c$ and $c'$ are QREs.
@@
 
-One potential pitfall with QRE lenses is that the combinators which
+One potential pitfall with QRE lenses is that the combinators that \saz{instead
+  of ``which''}
 we define on them may not preserve this normal form. For example, the
-composition $q' \circ q$ of QRE lenses $q$ and $q'$ may not be equivalent to
-any quotient lens which has the canonical normal form of a bijective lens with
-canonizers a the edges. Our main technical contribution in this paper is a
+composition $q' \circ q$ of QRE lenses $q$ and $q'$ might not be equivalent to
+any quotient lens that \saz{instead of ``which''} has the canonical normal form of a bijective lens with
+canonizers at the edges. Our main technical contribution in this paper is a
 proof that each QRE lenses $q$ defines the same transformation as a QRE lens
@@
 
-For example, suppose \cd{q_1 : bibtex <=> endnote} is a QRE lens that maps
-between \bibtex{} and EndNote formats, and that \cd{q_2 : endnote <=> xml}
-is a QRE lens that maps between EndNote and XML formats. The composition 
-\cd{(q_1 ; q2) : bibtex <=> xml} of \cd{q_1} and \cd{q_2} is not in normal form
-due to the presence of canonizers on each side of the intermediary EndNote
-format:
+For example, suppose \cd{q_1 : bibtex} $\Leftrightarrow$ \cd{endnote} is a QRE
+lens that maps between \bibtex{} and EndNote formats, and that \cd{q_2 : endnote
+  <=> xml} is a QRE lens that maps between EndNote and XML formats. The
+composition \cd{(q_1 ; q2) : bibtex} $\Leftrightarrow$ \cd{xml} of \cd{q_1} and
+\cd{q_2} is not (necessarily) in normal form due to the presence of canonizers
+on each side of the intermediary EndNote format:
 
@@
 
-\noindent However, while \cd{q_1 ; q_2} is not in normal form, there is another
-QRE lens \cd{q : bibtex <=> xml} of the same type as \cd{q_1 ; q_2} such that
-\cd{q_1 ; q_2} and \cd{q} define the same transformation. This proof will allow
-justify our approach for synthesizing quotient lenses so that the programmer
-can simply type
+\noindent However, while \cd{q_1 ; q_2} is not in normal form, our main theorem
+shows that there is another QRE lens \cd{q : bibtex} $\Leftrightarrow$ \cd{xml} of the same type as
+\cd{q_1 ; q_2} such that \cd{q_1 ; q_2} and \cd{q} define the same
+transformation. This proof will allow justify our approach for synthesizing
+quotient lenses so that the programmer can \saz{replace all of the code in
+  Figure~\ref{fig:example-lens} with a call to the synthesis prodedure:}
 
 \begin{lstlisting}
-let bib_to_end : (lens in bibtex <=> endnote) 
-= synth bibtex <=> endnote  using  {(bib_example, end_example)}
+let bib_to_end : (lens in bibtex $\Leftrightarrow$ endnote) 
+= synth bibtex $\Leftrightarrow$ endnote  using  {(bib_example, end_example)}
 \end{lstlisting}
-\noindent to synthesize a quotient lens that maps between the \cd{bibtex} and
-\cd{endnote} QREs, and sends the equivalence class of \cd{bib_bexample} to
-the equivalence class of \cd{end_example}.
+\noindent Here, the resulting  quotient lens maps between the \cd{bibtex} and
+\cd{endnote} QREs, and sends the equivalence class of \cd{bib_example} to
+the equivalence class of \cd{end_example}, which are the two concrete example
+strings given at the beginning of this Section.
+
+\saz{
+As we saw with the definition of \cd{space_to_comma} in
+Figure~\ref{fig:example-qre}, the synthesis procedure itself can be used to
+creates lenses that are in turn used to define other QREs.  This ability to mix
+QRE specification with QRE lens synthesis yields a powerful and flexible way of
+creating bidirectional transformations.}
 
@@
 The language of Quotient Regular Expressions (QREs) is given by the following
-grammar:
+grammar:  \saz{Use the same font that we use above---I added a \textbackslash kw
+macro}
 \begin{align*}
 q := \; &R \sep \project{R}{s} \sep \squash{q}{q'}{f} \sep
-\perm{q_1, \ldots, q_n}{q} \;  | \; \normalize{R}{R'}{f}\\
-&q \semicolon q' \sep q \cdot q' \sep (q \sep q') \sep q^*,
+\perm{q_1, \ldots, q_n}{q} \\
+  & | \; \normalize{R}{R'}{f} \sep q \semicolon q' \sep q \cdot q' \sep (q \sep q') \sep q^*,
 \end{align*}
@@
 
+\saz{We use the notation $\mathcal{L}(R)$ to mean the language accepted by the
+  regular expression $R$.}
+
 \subsection{Semantics of Quotient Regular Expression}
@@
 \begin{lstlisting}
-let space_name  = name | (name . wsp_to_space . (name . wsp_to_space)* . name)
+let space_name  = name | (name . wsp_to_space . (name . wsp_to_space)$^*$ . name)
 \end{lstlisting}
 
-The $q \semicolon q'$ QRE is a compostion combinator used to successively apply
+\saz{One combinator that we haven't yet introduced is sequential composition.}
+The $q \semicolon q'$ QRE successively applis
 two equivalence relations to a regular language. For example, in the \bibtex{}
@@
 instead have defined \cd{valid_name} as follows:
+\saz{Instead of replaying the \cd{valid_name} example, perhaps use a simpler
+  one---even an artificial example might be better.}
 
@@
 
-The final combinator is the  $\normalize{R}{R'}{f}$ combinator, which expresses
-the equivalence relation which sends each element in $\mathcal{L}(R)$ to its
+The final combinator is  $\normalize{R}{R'}{f}$, which expresses
+the equivalence relation that \saz{instead of ``which''} sends each element in $\mathcal{L}(R)$ to its
 canonical representation $f(r) \in \mathcal{L}(R')$. The equivalence relation
@@
 
-The $\normalize{R}{R'}{f}$ QRE is the most general QRE in that for each QREs
+The $\normalize{R}{R'}{f}$ QRE is the most general QRE in that for each QRE
 $q$, there exist regular expressions $R, R'$ and a surjective, idempotent
@@
 representatives for this equivalence relation. The $\normalize{R}{R'}{f}$
-combinator is included so as to express equivalence relations which are
-difficult to express, or which cannot be expressed, using the other
+combinator is included so as to express equivalence relations that are
+difficult to express, or that cannot be expressed, using the other
 combinators.
@@
 However, checking that the canonizing function $f$ is
-surjective and idempotent is undecidable, thus we do not use the
+surjective and idempotent is undecidable in general, thus we do not use the
 $\normalize{R}{R'}{f}$ in practice. On the other hand, the
@@
 
+\saz{This is a bit confusing.} \saz{The equivalence relations definable by
+  QREs are a subset of those that can be expressed in Boomerang.} 
 While QREs express a broad class of equivalence relations on regular languages,
@@
 
-In summary each QRE $q$ enables us to express
+The net effect of these design decisions means that each QRE $q$ enables us to
+simultaneously express
 \begin{enumerate}
@@
 
+\subsection{QREs and ambiguity}
 Before giving the formal semantics for QREs, we address the issue of {\em
-ambiguity} which comes up when defining QREs. If $q, q'$ are QREs, then
+ambiguity} which comes up when defining QREs. If $q$ and  $q'$ are QREs, then
 when applying the regular combinators to QREs, we require that if a string $s$
@@
 
+\subsection{Semantics of QREs}
+
+\saz{This needs a bit more setting up / explanation. Explain what the reader
+  should expect to see in the figures.}
 The semantics for QREs that formally define the whole language $W(q)$, the
@@
 \begin{figure}[t]
+\saz{composition notation}
   \centering
@@
   \centering
+\saz{Fix the projection case}
 \[
     \begin{array}{l@{\quad}l@{\quad}l} 
-      w \; \equiv_R \; w' &\iff& w = w \\
+      w \; \equiv_R \; w' &\iff& w = w' \\
       w \; \equiv_{R \mapsto s} \; w' \text{ for all }w, w'\\

### 2017-11-02 papers/QuotientLensSynth/main.tex
finished a pass through section 4

@@
 \varnothing$, or (4) $R = S^*$ with $S$ strongly unambiguous and $S^{*!}$.
+
+\saz{Is this where we should explain the rules of  Figure~\ref{fig:qrerules}?}
 
@@
 \end{figure}
-\begin{figure}[t]
+
+\begin{figure}[p!]
 \centering
@@
 \end{prooftree}
-  \caption{QRE Inference Rules\bcp{Maybe some or all of these can be
+  \caption{Well-formed QREs \bcp{Maybe some or all of these can be
       suppressed in the ``short version'' of the paper}\bcp{No, cancel
@@
     issues involving normalize.  (And a note that we don't really use
-    normalize, but rather introduce more specialized instances.)}}
+    normalize, but rather introduce more specialized instances.)}
+  \saz{maybe shorten the judgment from ``is well formed'' to ``wf''?}
+}
   \label{fig:qrerules}
 \end{figure}
+
+\saz{We need to explain the rules in Figure~\ref{fig:qrerules}---otherwise they
+  aren't mentioned anywhere in the text!}
 The theorem below confirms that the semantics of QREs are consistent with their
@@
 \label{QRE-lenses}
+
+\saz{Throughout this section: do a which hunt to look for ``which''es that
+  should be ``that''s.}
+
 As we have seen, QREs express a broad class of equivalence relations
@@
 
-Our approach in defining QRE lenses is as follows. Let $c, c'$ be QREs. Given a
+\saz{We need to be careful with our metavariable conventions.  Up until now $q$
+  has meant ``a QRE'', but this sections says that $q$ and $q'$ are
+  \textit{lenses} and switches to using $c$ and $c'$ for QREs. We need to
+  pick a metavariable that ranges over QRE lenses and use it consistently and
+  switch these $c$'s to $q$'s.
+}
+
+Our approach in defining QRE lenses is as follows. Let $c$ and $c'$ be QREs. Given a
 bijection $\ell : K(c) \Leftrightarrow K(c')$, we may define a quotient lens $q
@@
 
-One potential pitfall with QRE lenses is that the combinators which
+One potential pitfall with QRE lenses is that the combinators that
 we define on them may not preserve this normal form. For example, the
@@
 
+\saz{include the denotations of const and $\circ$}
 \begin{align*}
@@
 $S$ are regular expressions. If $\ell : R \Leftrightarrow S$, then the source
-type of $\ell$ is $\mathcal{L}(R)$ and the target type of $\ell$ is
+language of $\ell$ is $\mathcal{L}(R)$ and the target language of $\ell$ is
 $\mathcal{L}(S)$. Interestingly, with the bijective lens type system, a lens
@@
 
+\saz{Watch metavariables}
 The more interesting combinators are the $\mathit{lquot}$ and $\mathit{rquot}$
@@
 \begin{lstlisting}
-let bib_to_end_q : (lens in bibtex <=> endnote) = 
+let bib_to_end_q : (lens in bibtex $\Leftrightarrow$ endnote) = 
 rquot (lquot (bibtex, bib_to_end), endnote)
@@
 \subsection{Semantics of QRE Lenses}
+
+\saz{metavariables\ldots}
 Each QRE lens $q$ has a type $q : c \Leftrightarrow c'$ where $c, c'$ are QREs.
@@
 
- \begin{figure}[t]
+ \begin{figure}[ht]
   \centering
@@
 \end{theorem}
+\begin{proof} 
 We elide a full proof of this statement, since the proof is attained via a
-straightforward induction on the derivation $q : c \Leftrightarrow c$. Since
+straightforward induction on the derivation $q : c \Leftrightarrow c$. \saz{(The
+  full details can be found in the appendix?)}Since
 bijective lenses are closed under strongly unambiguous regular operators, the
@@
 straightforward application of the induction hypothesis followed by unrolling
-the equations that give the denotation for QRE lenses.
+the equations that give the denotation for QRE lenses.\qed
+\end{proof}
 \iffalse

### 2017-11-03 papers/QuotientLensSynth/main.tex

@@
 
-For example, with these rules, the following \bibtex{} citations should be
-considered the same by the lens \saz{``the lens'' hasn't been introduced to the
-  reader yet},
+For example, the desired transformation should map the following
+\bibtex{} citations to the same EndNote record in the forward direction:
 \begin{verbatim}
@@
 \end{verbatim}
-with the EndNote records having similar rules \saz{similarly, ``rules'' haven't been
-  mentioned yet, do you mean similar ``equivalence''} defined on them.
+\noindent EndNote records should also have a similar equivalence relation
+defined on them.
 
@@
 with each other in a way that makes them difficult to express with the standard
-Boomerang combinators. Moreover, many of the equivalences expressible in
-Boomerang require that the programmer first specify the data itself in form of
-a regular expression before specifying the equivalence relation on the data.
-\saz{Something like this instead?:  Moreover, Boomerang requires that the
+Boomerang combinators. Moreover, in many cases, Boomerang requires that the
 programmer first specify the data itself in the form of a regular expression
-  before she can define the equivalence relation on that data.}
-This leads to verbose, repetitive code.
-
-To address this issue\saz{these issues? these problems with Boomerang?}, we propose {\em Quotient Regular Expressions} or QREs,
-which express a broad class of equivalence relations directly on regular
-languages using a compact and convenient notation. For example, with QREs we
-can specify \bibtex{} records as shown in Figure~\ref{fig:example-qre}.
+before she can define the equivalence relation on that data. This leads to
+verbose, repetitive code.
+
+Instead, we propose {\em Quotient Regular Expressions} or QREs which enable
+the programmer to define a regular expression $R$ as well as an equivalence
+relation on $R$ in the same expression.For example, with QREs we can specify
+\bibtex{} records as shown in Figure~\ref{fig:example-qre}.
 
@@
 
-\saz{Throughout these examples, refer to the definitions in Figure 1.}  In the
-figure, the QRE \cd{wsp_to_space}, which is defined to be \cd{project wsp}$^+$
-$\to$ \cd{" "}, matches any non-zero number of whitespace characters, but
-considers the matched string of whitespace characters as equivalent with a
-single space as the canonical representative.  \saz{We call the type of QRE a
-\textit{projection} because it projects an entire regular language down to a single
-representative string in the language.}
-
-\saz{
+In Figure~\ref{fig:example-qre}, the QRE \cd{wsp_to_space}, which is defined to
+be
+
+\noindent \cd{project wsp}+ $\to$ \cd{" "}, matches any non-zero number of
+whitespace characters, but considers the matched string of whitespace
+characters as equivalent with a single space as the canonical representative. 
+We call this QRE a \textit{projection} because it projects an
+entire regular language down to a single representative string in the
+language.
+
 Another form of QRE is the \textit{squash} QRE, an example of which is given by
 \cd{valid_name}, which is defined to be:
-}
 
@@
   space_to_comma} is the function that transforms names in the \cd{comma_name}
-representation to names in the \cd{space_name} representation.  \saz{In this
+representation to names in the \cd{space_name} representation.  In this
   example, \cd{space_to_comma} is itself an automatically \textit{synthesized}
@@
   the ``get'' direction of the lens to specify the squash transformation. (We'll
-  say more about synthesis below.)}
+  say more about synthesis below.)
 
@@
 
-\saz{Given the \textit{projection}, \textit{squash}, and \textit{permutation}
+Given the \textit{projection}, \textit{squash}, and \textit{permutation}
   QRE primitives, we can build more complex languages using the usual regular
-  expression combinators of alternative choice `\cd{|}', Kleene star `$^*$', and
-  sequential composition ``.''.  As shown in the rest of
-  Figure~\ref{fig:example}, it is straightforward to define both a regular
-  language that accepts \bibtex{} records \textit{and}, simultaneously, an
-  equivalence relation that quotients out irrelevant details.}
+  expression combinators of alternative choice `\cd{|}', concatenation `\cd{.}'
+  and Kleene star `$^*$', and also sequential composition ``;''.  As shown in
+  the rest of Figure~\ref{fig:example-qre}, it is straightforward to define both a
+  regular language that accepts \bibtex{} records \textit{and}, simultaneously,
+  an equivalence relation that quotients out irrelevant details.
 
@@
 \end{lstlisting}
-\caption{An example lens between the \cd{bibtex} and \cd{end_note} QRE languages.}
+\caption{An example lens between \cd{bibtex} and \cd{end_note} regular
+expressions with no equivalences.}
 \label{fig:example-lens}
@@
 transforms \bibtex{} data to EndNote data  (and vice versa) by matching
-the delimiters and separators in a bijective manner. In more detail, the
+the delimiters and separators in a bijective manner. For example, the
 \cd{preamble} lens replaces the \bibtex{} preamble \verb|@Book| with the EndNote
-preamble \verb|%0 Book\n%F|. \saz{This next sentence isn't very well-supported
-  by the previous sentence's example, since that doesn't seem too hard.  Maybe
-  explain some of the harder parts too?}  The programmer can ill
-afford to make a mistake with any one of the small but vital details of the
-transformations since even a one character mistake might render the lens
-incorrect. This makes lens programming tedious and error-prone.
-
-To address this challenge, we identify a \saz{avoid ``special''} class of quotient
-lenses that \saz{instead of ``which''} we call {\em QRE lenses}, and show how to {\em
-synthesize} QRE lenses from a pair of QREs and a set of input-output
-example pairs. Our key insight is to reduce the problem of synthesizing QRE
-lenses to the problem of synthesizing {\em bijective lenses}, which are the
-bijections created from the Boomerang lens combinators. This is a problem
-which we solved in our prior work ~\cite{popl18}.
+preamble \verb|%0 Book\n%F|. When the programmer is writing the lens, she has
+to make sure that all of these small deletions and insertions are correct,
+which can be a challenging task if the lens is large enough. This makes lens
+programming tedious and error-prone.
+
+To address this challenge, we identify a  class of quotient lenses that we call
+{\em QRE lenses}, and show how to {\em synthesize} QRE lenses from a pair of
+QREs and a set of input-output example pairs. Our key insight is to reduce the
+problem of synthesizing QRE lenses to the problem of synthesizing {\em
+bijective lenses}, which are the bijections created from the Boomerang lens
+combinators. This is a problem which we solved in our prior work ~\cite{popl18}.
 
 QRE lenses lenses are the same as bijective lenses with ``canonizers at the
-edges'' as in Figure~\ref{fig:attheedges}, and are typed using QREs---\saz{in
-  latex use three dashes --- for this (not one -).  Put no spaces around ``---''}every QRE
+edges'' as in Figure~\ref{fig:attheedges}, and are typed using QREs---every QRE
 lens $q$ has a type $q: c \Leftrightarrow c'$ where $c$ and $c'$ are QREs.
@@
 
-One potential pitfall with QRE lenses is that the combinators that \saz{instead
-  of ``which''}
-we define on them may not preserve this normal form. For example, the
-composition $q' \circ q$ of QRE lenses $q$ and $q'$ might not be equivalent to
-any quotient lens that \saz{instead of ``which''} has the canonical normal form of a bijective lens with
-canonizers at the edges. Our main technical contribution in this paper is a
-proof that each QRE lenses $q$ defines the same transformation as a QRE lens
-$q'$ which is in normal form.
+One potential pitfall with QRE lenses is that the combinators that we define on
+them may not preserve this normal form. For example, the composition $q'; q$ of
+QRE lenses $q$ and $q'$ might not be equivalent to any quotient lens that has
+the canonical normal form of a bijective lens with canonizers at the edges.
+Our main technical contribution in this paper is a proof that each QRE lenses
+$q$ defines the same transformation as a QRE lens $q'$ which is in normal form.
 
@@
 
-\saz{
 As we saw with the definition of \cd{space_to_comma} in
@@
 QRE specification with QRE lens synthesis yields a powerful and flexible way of
-creating bidirectional transformations.}
+creating bidirectional transformations.
 
@@
 
-\saz{One combinator that we haven't yet introduced is sequential composition.}
+One combinator that we haven't yet introduced is sequential composition.
 The $q \semicolon q'$ QRE successively applis
@@
 The final combinator is  $\normalize{R}{R'}{f}$, which expresses
-the equivalence relation that \saz{instead of ``which''} sends each element in $\mathcal{L}(R)$ to its
+the equivalence relation that sends each element in $\mathcal{L}(R)$ to its
 canonical representation $f(r) \in \mathcal{L}(R')$. The equivalence relation
@@
 
-\saz{This is a bit confusing.} \saz{The equivalence relations definable by
-  QREs are a subset of those that can be expressed in Boomerang.} 
-While QREs express a broad class of equivalence relations on regular languages,
-they do not express all the equivalence relations that can be expressed in
-Boomerang, nor of course all the equivalence relations which cannot be
-expressed using Boomerang canonizers. For instance, Boomerang canonizers can
-map a regular language described by $R$ to a regular language described by
-$S$, with $\mathcal{L}(S) \not \subseteq \mathcal{L}(R)$, whereas each QRE $q$
-canonizes a regular language described by $R$ to regular subset
-$\mathcal{L}(R') \subseteq \mathcal{L}(R)$ described by $R'$.
+The equivalence relations definable by QREs are a strict subset of those that
+can be expressed in Boomerang. For instance, in Boomerang, one can map a
+regular language described by $R$ to a regular language described by $S$, with
+$\mathcal{L}(S) \not \subseteq \mathcal{L}(R)$, whereas each QRE $q$ canonizes
+a regular language described by $R$ to regular subset $\mathcal{L}(R')
+\subseteq \mathcal{L}(R)$ described by $R'$.
 
@@
 \end{enumerate}
-As an example, recall that
-in the \bibtex{} to EndNote transformation we had the QREs \cd{bibtex} and \cd{endnote} describing a \bibtex{} record and an
-Endnote record respectively. We also had the example pair 
-\cd{(bib_example , end_example)} where
+For example, recall that in the \bibtex{} to EndNote transformation we had the
+QREs \cd{bibtex} and \cd{endnote} describing a \bibtex{} record and an Endnote
+record respectively. We also had the example pair \cd{(bib_example ,
+end_example)} where
 
@@
 Since the transformation between the kernel of \cd{bibtex} and the kernel of
-\cd{endnote!} that maps \cd{bib_example!} to \cd{end_example} can be
+\cd{endnote} that maps \cd{bib_example} to \cd{end_example} can be
 expressed by a bijective lens, it suffices to type the expression
@@
 
-\bcp{No bibliography??}
+\bibliography{local}
 

### 2017-11-03 papers/BijectiveLensSynth/ConferencePaper/old/rebuttal.txt
move paper into conferencepaper directory

--- 2017-01-29 paper/rebuttal.txt
@@
-INTENDED USE CASE
+# INTENDED USE CASE
 
@@
 
-FORMALIZATION
+# FORMALIZATION
 
@@
 
-EVALUATION
+# EVALUATION
 
@@
 
-BIJECTIVE LENS RESTRICTION
+# BIJECTIVE LENS RESTRICTION
 
@@
 Because of our use of orderings, we are able to find the permutations in
-O(n*log(n)) time.  It does not seem likely that a SAT solver would help
+O(n * log(n)) time.  It does not seem likely that a SAT solver would help
 here.
@@
 well aligned lenses.  Furthermore, the aligned regular expressions
-could become difficult to read and maintain: (A|B)(e+C+CC*C) is less
-understandable than (A|B)C*.  Surely, avoiding this complexity is a
+could become difficult to read and maintain: (A|B)(e+C+CC* C) is less
+understandable than (A|B)C* .  Surely, avoiding this complexity is a
 win for the user.  Fortunately, we are successful in our efforts to do

### 2017-11-08 papers/QuotientLensSynth/main.tex

@@
 lenses whose types are given by QREs. Our main technical contribution in this
-paper is a proof that each QRE lenses $q$ defines the same transformation as a
-QRE lens $q'$ which is in normal form. (Section~\ref{QRE-lenses}).
+paper is a proof that each QRE lens $q$ defines the same transformation as a
+QRE lens $q'$ that is in normal form. (Section~\ref{QRE-lenses}).
 \item Using this normal form, we reduce the problem of {\em synthesizing}
@@
 
-Consider a bidirectional transformation which converts citation records such as
+Consider a bidirectional transformation that converts citation records such as
 \begin{verbatim}
@@
 
-We wish for this transformation to have the same behaviour on records which
+We wish for this transformation to have the same behaviour on records that
 differ only in ``inessential'' details. First, the transformation
@@
 let wsp = [ \n\t\r]
+
+(* project QRE maps arbitrary occurrences of whitespace to a single space *)
 let wsp_to_space   = project wsp$^+$ $\to$ " "
+
+(* define name representations with a space and with a comma *)
 let space_name     = name | (name . wsp_to_space . (name . wsp_to_space)$^*$ . name)
 let comma_name   = name | (name . "," . wsp_to_space . (name . wsp_to_space)$^*$ . name)
+
+(* synthesize a lens that maps the comma representation to the space representation *)
 let space_to_comma = synth comma_name $\Leftrightarrow$ space_name using {("Conway, J. H.", "J. H. Conway")}
+
+(* squash QRE maps comma_name to space_name *)
 let valid_name = squash comma_name $\to$ space_name using (get space_to_comma)
+
+(* define rest of bibtex fields *)
 let wsp_to_and = project (wsp$^+$ . "and" . wsp$^+$) $\to$ " and " 
@@
 let bib_title = "title = {" . title . "}"
+
+(* perm QRE allows permutations of fields with arbitrary whitespace in between *)
 let bib_perm  = perm (label, bib_author, bib_title) with (project ("," .wsp$^+$) $\to$ ",\n")
@@
 
-\noindent This is a \textit{permutation} QRE, which specifies that the label, title, and author fields can occur
-in any order, with the fields separated by a comma followed by arbitrary amounts
-of whitespace. The canonical representation is that in which the label, author
-and title fields occur in that order, separated by a comma followed a newline.
+\noindent This is a \textit{permutation} QRE, which specifies that the label,
+title, and author fields can occur in any order, with the fields separated by a
+comma followed by arbitrary amounts of whitespace. The canonical representation
+is that in which the label, author and title fields occur in that order,
+separated by a comma followed a newline.
 Such \cd{perm} QREs allow for easy nesting of permutations. Similarly, it is easy
@@
 Our main technical contribution in this paper is a proof that each QRE lenses
-$q$ defines the same transformation as a QRE lens $q'$ which is in normal form.
+$q$ defines the same transformation as a QRE lens $q'$ that is in normal form.
 
@@
 \noindent However, while \cd{q_1 ; q_2} is not in normal form, our main theorem
-shows that there is another QRE lens \cd{q : bibtex} $\Leftrightarrow$ \cd{xml} of the same type as
-\cd{q_1 ; q_2} such that \cd{q_1 ; q_2} and \cd{q} define the same
-transformation. This proof will allow justify our approach for synthesizing
-quotient lenses so that the programmer can \saz{replace all of the code in
-  Figure~\ref{fig:example-lens} with a call to the synthesis prodedure:}
+shows that there is another QRE lens \cd{q : bibtex} $\Leftrightarrow$ \cd{xml}
+of the same type as \cd{q_1 ; q_2} such that \cd{q_1 ; q_2} and \cd{q} define
+the same transformation. This proof will allow justify our approach for
+synthesizing quotient lenses so that the programmer can replace all of the
+code in Figure~\ref{fig:example-lens} with a call to the synthesis prodedure:
 
@@
 The language of Quotient Regular Expressions (QREs) is given by the following
-grammar:  \saz{Use the same font that we use above---I added a \textbackslash kw
-macro}
-\begin{align*}
-q := \; &R \sep \project{R}{s} \sep \squash{q}{q'}{f} \sep
-\perm{q_1, \ldots, q_n}{q} \\
-  & | \; \normalize{R}{R'}{f} \sep q \semicolon q' \sep q \cdot q' \sep (q \sep q') \sep q^*,
-\end{align*}
-where $R$ ranges over regular expressions, $f$ ranges over functions between
-regular languages, and $s$ ranges over character strings.
-
-\saz{We use the notation $\mathcal{L}(R)$ to mean the language accepted by the
-  regular expression $R$.}
-
-\subsection{Semantics of Quotient Regular Expression}
+grammar: 
+\begin{align*}
+c := \; &R \sep \project{R}{s} \sep \squash{c}{c'}{f} \sep
+\perm{c_1, \ldots, c_n}{c} \\
+  & | \; \normalize{R}{R'}{f} \sep c \semicolon c' \sep c \cdot c' \sep (c \sep
+  c') \sep c^*,
+\end{align*}
+where $c$ ranges over QREs, $R$ ranges over regular expressions, $f$ ranges over
+functions between regular languages, and $s$ ranges over character strings. In
+subsequent sections, we will use the the notation $\mathcal{L}(R)$ to mean the
+language accepted by the regular expression $R$.
+
+\subsection{Overview of Quotient Regular Expression}
 In the \bibtex{} to EndNote example, we encountered the $\project{R}{s}$,
 
-\noindent $\squash{q}{q'}{f}$ and $\perm{q_1, \ldots, q_n}{q}$ QREs, as well as
+\noindent $\squash{c}{c'}{f}$ and $\perm{c_1, \ldots, c_n}{c}$ QREs, as well as
 the regular operators applied to these, as in
@@
 One combinator that we haven't yet introduced is sequential composition.
-The $q \semicolon q'$ QRE successively applis
-two equivalence relations to a regular language. For example, in the \bibtex{}
-to EndNote example, when defining the \cd{valid_name} QRE which maps
-\cd{comma_name} into \cd{space_name}, we first defined the equivalence
-relations on the \cd{comma_name} and \cd{space_name} representations before
-defining the \cd{valid_name} QRE. Using the composition combinator, we could
-instead have defined \cd{valid_name} as follows:
-\saz{Instead of replaying the \cd{valid_name} example, perhaps use a simpler
-  one---even an artificial example might be better.}
+The $c \semicolon c'$ QRE successively applies
+two equivalence relations to a regular language. For example, suppose that
+we wish to express a name as
+
+\noindent $\kw{last\_name \cdot ``,\text{''} \cdot wsp^+ \cdot first\_name}$,
+or as $\kw{first\_name\cdot wsp^+ \cdot last\_name}$, with
+
+\noindent $\kw{first\_name\cdot wsp^+ \cdot last\_name}$ as the canonical
+representation. We also wish to project away the extra whitespace in between
+\kw{first\_name} and \kw{last\_name} once this identification has been made.
+Then we can express this equivalence relation by a composition of a
+\textit{squash} QRE and a QRE that projects away the extra whitespace in between
+\kw{first\_name} and \kw{last\_name}:
 
 \begin{lstlisting}
-(* Define a QRE which matches sends all comma names to a space name equivalent *)
-let wsp = [ \n\t\r] 
-let whole_space_name  = name | (name . wsp+ . (name . wsp+)* . name) 
-let whole_comma_name = name | (name . "," . wsp+ . (name . wsp+)* . name) 
-let whole_space_to_comma = 
-synth whole_comma_name <=> whole_space_name using {("Conway, J. H.", "J. H. Conway")} 
-let whole_valid_name = 
-squash whole_comma_name -> whole_space_name using (get whole_space_to_comma)
-
-(* Define QRE that sends space names to their canonical representation *)
-let wsp_to_space = project wsp+ $\to$ " "
-let space_name  = name | (name . wsp_to_space . (name . wsp_to_space)* . name)
+(* Regular expressions matching different representations *)
+let comma_name = last_name . "," . wsp+ . first_name
+let space_name = first_name . wsp+ . last_name
+
+(* Squash comma representation to space representation *)
+let comma_to_space = 
+synth comma_name <=> space_name using {("Conway, J. H.", "J. H. Conway")} 
+let comma_to_space = squash comma_name $\to$ space_name using (get comma_to_space)
+
+(* project whitespace in space_name representation *)
+let project_in_space = first_name . (project wsp+ $\to$ " ") . last_name
 
 (* Compose the two QREs *)
-let valid_name = whole_valid_name ; space_name
+let valid_name = project_in_space ; comma_to_space
 \end{lstlisting}
@@
 regular language described by $R$ to a regular language described by $S$, with
-$\mathcal{L}(S) \not \subseteq \mathcal{L}(R)$, whereas each QRE $q$ canonizes
+$\mathcal{L}(S) \not \subseteq \mathcal{L}(R)$, whereas each QRE $c$ canonizes
 a regular language described by $R$ to regular subset $\mathcal{L}(R')
@@
 
-The net effect of these design decisions means that each QRE $q$ enables us to
+The net effect of these design decisions means that each QRE $c$ enables us to
 simultaneously express
 \begin{enumerate}
-  \item a regular expression $W(q)$ (the ``whole'' of $q$),
-  \item an equivalence relation $\eqrel{q}$ on $\mathcal{L}(W(q))$,
-  \item a regular expression $K(q)$ (the ``kernel'' of $q$)
-  such that $\mathcal{L}(K(q))$ forms a complete set of representatives for
-  $\eqrel{q}$, and
-  \item a ``canonizing'' function $\canonizer(q):\mathcal{L}(W(q))
-  \longrightarrow \mathcal{L}(K(q))$ which given any $w \in \mathcal{L}(W(q))$,
-  computes $\canonizer(q)(w)$ as the unique $k$ in $\mathcal{L}(K(q))$ such that
-  $k$ is equivalent to $w$ mod $\eqrel{q}$
+  \item a regular expression $W(c)$ (the ``whole'' of $c$),
+  \item an equivalence relation $\eqrel{c}$ on $\mathcal{L}(W(c))$,
+  \item a regular expression $K(c)$ (the ``kernel'' of $c$)
+  such that $\mathcal{L}(K(c))$ forms a complete set of representatives for
+  $\eqrel{c}$, and
+  \item a ``canonizing'' function $\canonizer(c):\mathcal{L}(W(c))
+  \longrightarrow \mathcal{L}(K(c))$ which given any $w \in \mathcal{L}(W(c))$,
+  computes $\canonizer(c)(w)$ as the unique $k$ in $\mathcal{L}(K(c))$ such that
+  $k$ is equivalent to $w$ mod $\eqrel{c}$
   \end{enumerate}
-  Observe that since $\canonizer(q)$ computes $\canonizer(q)(w)$ as the unique
-  $k$ in $\mathcal{L}(K(q))$ such that $w$ is equivalent to $k$ mod $\eqrel{q}$,
-  then the equivalence classes of $\eqrel{q}$ are the same as the fibres of
-  $\canonizer(q)$.
-
-\subsection{QREs and ambiguity}
+  Observe that since $\canonizer(c)$ computes $\canonizer(c)(w)$ as the unique
+  $k$ in $\mathcal{L}(K(c))$ such that $w$ is equivalent to $k$ mod $\eqrel{c}$,
+  then the equivalence classes of $\eqrel{c}$ are the same as the fibres of
+  $\canonizer(c)$.
+
+\subsection{Well Formed QREs and ambiguity}
 Before giving the formal semantics for QREs, we address the issue of {\em
-ambiguity} which comes up when defining QREs. If $q$ and  $q'$ are QREs, then
+ambiguity} which comes up when defining QREs. If $c$ and  $c'$ are QREs, then
 when applying the regular combinators to QREs, we require that if a string $s$
-matches any of the regular expressions $W(q) \cdot W(q')$, $W(q) \sep W(q')$,
-$W(q)^*$, $K(q) \cdot K(q')$, $K(q) \sep K(q')$, $K(q)^*$, then $s$ matches
+matches any of the regular expressions $W(c) \cdot W(c')$, $W(c) \sep W(c')$,
+$W(c)^*$, $K(c) \cdot K(c')$, $K(c) \sep K(c')$, $K(c)^*$, then $s$ matches
 that regular expression in only one way. This unambiguity condition is called
@@
 
-\saz{Is this where we should explain the rules of  Figure~\ref{fig:qrerules}?}
-
+\begin{figure}[p!]
+\begin{center}
+\AxiomC{$R$ is strongly unambiguous}
+\UnaryInfC{$\mathit{id}(R)$ \textbf{WF}}
+\DisplayProof
+\hskip 1.5em
+\AxiomC{$s \in \mathcal{L}(R)$}
+\UnaryInfC{$\project{R}{s}$ \textbf{WF}}
+\DisplayProof
+\hskip 1.5em 
+\end{center}
+
+\begin{prooftree}
+\AxiomC{$c_i, c$ are well formed}
+\AxiomC{${\substack{\forall \sigma \neq \theta, \; W(c_{\sigma(1)}) \cdot W(c)
+\cdot \ldots \cdot W(c_{\sigma(n)})\\ \cap W(c_{\theta(1)}) \cdot W(c)
+\cdot \ldots \cdot W(c_{\theta(n)}) =\varnothing}}$}
+\AxiomC{$K(c_1) \cdot^! K(c) \cdot^! \ldots \cdot^! K(c) \cdot^! K(c_n)$}
+\TrinaryInfC{$\perm{c_1, \ldots, c_n}{c}$ \textbf{WF}}
+\end{prooftree}
+
+\begin{center}
+\AxiomC{$c$ \textbf{WF}}
+\AxiomC{$c'$ \textbf{WF}}
+\AxiomC{$\mathcal{L}(W(c)) \cap \mathcal{L}(W(c')) = \varnothing$}
+\AxiomC{$f : \mathcal{L}(W(c)) \longrightarrow \mathcal{L}(W(c'))$}
+\QuaternaryInfC{$\squash{c}{c'}{f}$ \textbf{WF}}
+\DisplayProof
+\end{center}
+\begin{center}
+
+\begin{prooftree}
+\AxiomC{$\mathcal{L}(R') \subseteq \mathcal{L}(R)$}
+\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(R')$}
+\AxiomC{$f$ is surjective}
+\AxiomC{$f = f^2$}
+\QuaternaryInfC{$\normalize{R}{R'}{f}$ \textbf{WF}}
+\end{prooftree}
+
+\end{center}
+\begin{center}
+\AxiomC{$c$ \textbf{WF}}
+\AxiomC{$c'$ \textbf{WF}}
+\AxiomC{$K(c) = W(c')$}
+\TrinaryInfC{$c' \; ; \; c$ \textbf{WF}}
+\DisplayProof
+\hskip 1.5em
+\AxiomC{$c$ \textbf{WF}}
+\AxiomC{$W(c)^{*!}$}
+\AxiomC{$K(c)^{*!}$}
+\TrinaryInfC{$c^*$ \textbf{WF}}
+\DisplayProof
+\end{center}
+\begin{center}
+
+\begin{center}
+\AxiomC{$c$ \textbf{WF}}
+\AxiomC{$c'$ \textbf{WF}}
+\AxiomC{$W(c) \cdot^! W(c')$}
+\AxiomC{$K(c) \cdot^! K(c')$}
+\QuaternaryInfC{$c \cdot c'$ \textbf{WF}}
+\DisplayProof
+\end{center}
+
+\begin{center}
+\AxiomC{$c$ \textbf{WF}}
+\AxiomC{$c'$ \textbf{WF}}
+\AxiomC{$W(c) \cap W(c') = \varnothing$}
+\TrinaryInfC{$c \cdot c$ \textbf{WF}}
+\DisplayProof
+\hskip 1.5em
+\AxiomC{$c$ \textbf{WF}}
+\AxiomC{$W(c)^{*!}$}
+\AxiomC{$K(c)^{*!}$}
+\TrinaryInfC{$c^*$ \textbf{WF}}
+\DisplayProof
+\end{center}
+\end{center}
+  \caption{Well-formed QREs 
+}
+  \label{fig:qrerules}
+\end{figure}
+
+Figure~\ref{fig:qrerules} gives the inferrence rules for derving well-formed
+QREs. As expected, the unambiguity conditions come into play when defining QREs
+using the regular combinators. For example, the inferrence rule for
+concatenation is as follows:
+\begin{prooftree}
+\AxiomC{$c$ \textbf{WF}}
+\AxiomC{$c'$ \textbf{WF}}
+\AxiomC{$W(c) \cdot^! W(c')$}
+\AxiomC{$K(c) \cdot^! K(c')$}
+\QuaternaryInfC{$c \cdot c'$ \textbf{WF}}
+\end{prooftree}
+
+\noindent This rule says that we can concatenate two QREs $c$ and $c'$ only when
+the concatentions of $W(c)$ and $W(c')$, and $K(c)$ and $K(c')$ are ambiguous.
+
+The most complicated inferrence rule is the rule for the \textit{permutation}
+combinator:
+
+\begin{prooftree}
+\AxiomC{$c_i, c$ are well formed}
+\AxiomC{${\substack{\forall \sigma \neq \theta, \; W(c_{\sigma(1)}) \cdot W(c)
+\cdot \ldots \cdot W(c_{\sigma(n)})\\ \cap W(c_{\theta(1)}) \cdot W(c)
+\cdot \ldots \cdot W(c_{\theta(n)}) =\varnothing}}$}
+\AxiomC{$K(c_1) \cdot^! K(c) \cdot^! \ldots \cdot^! K(c) \cdot^! K(c_n)$}
+\TrinaryInfC{$\perm{c_1, \ldots, c_n}{c}$ \textbf{WF}}
+\end{prooftree}
+
+\noindent The second hypothesis for the \textit{permutation} rule says that for
+any two different permutations $\sigma$ and $\theta$, the languages
+$W(c_{\sigma(1)}) \cdot W(c) \cdot \ldots \cdot W(c_{\sigma(n)})$ and
+$W(c_{\theta(1)}) \cdot W(c) \cdot \ldots \cdot W(c_{\theta(n)})$ must be
+disjoint. This is because any string that matches the \textit{permutation}
+combinator matches the language $W(c_{\theta(1)}) \cdot W(c) \cdot \ldots \cdot
+W(c_{\theta(n)})$ for some permutation $\theta$, so we require that these
+regular expressions be disjoint.
+
+The third hypothesis says that the regular expression $K(c_1) \cdot K(c)
+\cdot \ldots \cdot K(c) \cdot K(c_n)$ must be unambiguous. This is because the
+underlying lens that maps to or from a \textit{permutation} QRE operates on
+the language $K(c_1) \cdot K(c) \cdot \ldots \cdot K(c) \cdot K(c_n)$ which is
+the kernel of the \textit{permutation} QRE, and as we mentioned, the underlying
+lens will require that the source and target regular expressions of the lens be
+unambiguous so that the lens can match strings uniquely.
+
+The rest of the inferrence rules for deriving QREs are given in
+Figure~\ref{fig:qrerules}.
 \subsection{Semantics of QREs}
 
-\saz{This needs a bit more setting up / explanation. Explain what the reader
-  should expect to see in the figures.}
-The semantics for QREs that formally define the whole language $W(q)$, the
-kernel language $K(q)$, the equivalence relation $\eqrel{q}$ and the
-canonizing function $\canonizer(q)$ are given in
-Figures~\ref{fig:wk},~\ref{fig:relations} and~\ref{fig:canonizers}.
+As we mentioned before, each QRE $c$ enables us to express:
+\begin{enumerate}
+  \item a regular expression $W(c)$ (the ``whole'' of $c$),
+  \item an equivalence relation $\eqrel{c}$ on $\mathcal{L}(W(c))$,
+  \item a regular expression $K(c)$ (the ``kernel'' of $c$)
+  such that $\mathcal{L}(K(c))$ forms a complete set of representatives for
+  $\eqrel{c}$, and
+  \item a ``canonizing'' function $\canonizer(c):\mathcal{L}(W(c))
+  \longrightarrow \mathcal{L}(K(c))$ which given any $w \in \mathcal{L}(W(c))$,
+  computes $\canonizer(c)(w)$ as the unique $k$ in $\mathcal{L}(K(c))$ such that
+  $k$ is equivalent to $w$ mod $\eqrel{c}$
+  \end{enumerate}
+Figures~\ref{fig:wk} gives the inductive rules that define the whole language
+$W(c)$ of a QRE. The two interesting rules are the rules for the
+\textit{squash}, \textit{normalize} and \textit{permutation} combinators. If $c
+= \squash{c}{c'}{f}$, then the whole language of $c$ is $W(c) \sep W(c')$. This
+is because the \textit{squash} combinator takes the whole language $W(c)$ of
+$c$, the whole language $W(c')$ of $c'$ and a function $f : \mathcal{L}(W(c))
+\longrightarrow \mathcal{L}(W(c'))$, maps $\mathcal{L}(W(c))$ into
+$\mathcal{L}(W(c'))$ using $f$ and then canonizes $W(c')$ into $K(c')$ using
+$\canonizer(c')$.
+
+For the $\perm{(c_1, \ldots, c_n)}{c}$ combinator, the whole language is the
+union of languages of the form $W(c_{\sigma(1)}) \cdot W(c) \cdot \ldots \cdot
+W(c) \cdot W(c_{\sigma(n)})$. This is because the \textit{permutation}
+combinator allows for the string to match any permutation the $c_i$'s while also
+taking into account the separator $c$ in between each of the $c_i$'s. The kernel
+of the \textit{permutation} combinator is the language $K(c_1) \cdot K(c) \cdot
+\ldots \cdot K(c) \cdot K(c_n)$, because the canonical permutation is the
+identity permuation, and then each of the parts of the string that match $c_i$
+and $c$ are canonized into $K(c_i)$ and $K(c)$ respectively.
+
+Figure~\ref{fig:canonizers} gives the inductive definitions of the
+$\canonizer{}$ function for each QRE. Again, the the least obvious definition
+is
+
+\begin{align*}\canonizer(\perm{c_1, \ldots, c_n}{c})(r_{\sigma(1)}
+\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \\
+= (\canonizer(c_1)(r_1)) \cdot (\canonizer(c)(s_1)) \cdot \ldots \cdot
+(\canonizer(c_n)(r_n))
+\end{align*}
+
+\noindent which as we have explained is defined so that the strings that match
+$c_i, c$ are placed in the canonical permutation which is the identity
+permutation before $\canonizer(c_i)$ and $\canonizer(c)$ are applied
+respectively.
+
+Finally, Figure ~\ref{fig:relations} gives the induction defintion of the
+equivalence relation $\eqrel{c}$. These are the set-theoretic semantics of QREs
+as an equivalence relation on a regular language.
+
  \begin{figure}[t]
-\saz{composition notation}
   \centering
@@
 
-      q & W(q) & K(q) \\ \hline
+      c & W(c) & K(c) \\ \hline
       R & R & R \\
-      R \mapsto s & W(q) & s \\
-      \squash{R}{R'}{f} & R \sep R' & R' \\
+      \project{R}{s} & R & s \\
+      \squash{c}{c'}{f} & W(c) \sep W(c') & K(c') \\
       \normalize{R}{R'}{f} & R & R' \\
-      q_1 \circ  q_2 & W(q_2) & K(q_1) \\
-      q_1 \cdot q_2 & W(q_1) \cdot W(q_2) & K(q_1) \cdot K(q_2) \\
-      q_1 \sep q_2 & W(q_1) \sep W(q_2) & K(q_1) \sep K(q_2) \\
-      q^* & W(q)^* & K(q)^* \\
+      c_2 \; ; \; c_1 & W(c_1) & K(c_2) \\
+      c_1 \cdot c_2 & W(c_1) \cdot W(c_2) & K(c_1) \cdot K(c_2) \\
+      c_1 \sep c_2 & W(c_1) \sep W(c_2) & K(c_1) \sep K(c_2) \\
+      c^* & W(c)^* & K(c)^* \\
     \end{array}
@@
 \begin{array}{r@{\quad}l}
-W( \perm{q_1, \ldots, q_n}{q} ) = &
-\bigcup \limits_{\sigma \in S_n} W(q_{\sigma(1)}) \cdot W(q) \cdot \ldots \cdot
-W(q) \cdot W(q_{\sigma(n)})
-\\
-K( \perm{q_1, \ldots, q_n}{q} ) = &
- K(q_1) \cdot K(q) \cdot \ldots \cdot K(q) \cdot K(q_n) 
+W( \perm{c_1, \ldots, c_n}{c} ) = &
+\bigcup \limits_{\sigma \in S_n} W(c_{\sigma(1)}) \cdot W(c) \cdot \ldots \cdot
+W(c) \cdot W(c_{\sigma(n)})\\
+K( \perm{c_1, \ldots, c_n}{c} ) = & K(c_1) \cdot K(c) \cdot \ldots \cdot K(c)
+\cdot K(c_n)
 \end{array}
@@
 \begin{figure}[t]
+  \begin{center}
+\[
+    \begin{array}{l@\quad l @\quad l} 
+      \canonizer(R) &=& id_{\mathcal{L}(R)} \\
+      \canonizer(\project{R}{s})(w) &=& s\\
+      \canonizer(\squash{c}{c'}{f})(w) &=& 
+\begin{cases}
+\canonizer(c')(f(w)) & \text{if } w \in \mathcal{L}(W(c))\\
+\canonizer(c')(w) & \text{otherwise}
+\end{cases}\\
+      \canonizer(\normalize{R}{R'}{f}) &=& f\\
+      \canonizer(c' \; ; \; c) &=& \canonizer(c') \circ \canonizer(c)\\
+      \canonizer(c' \cdot c) &=& \canonizer(c') \cdot \canonizer(c)\\
+      \canonizer(c_1 \sep c_2)(w) &=& 
+\begin{cases}
+\canonizer(c_1)(w) & \text{if } w \in \mathcal{L}(W(c_1))\\
+\canonizer(c_2)(w) & \text{if } w \in \mathcal{L}(W(c_2))\\
+\end{cases}\\
+      \canonizer(c^*) &=& \canonizer(c)^* \\
+    \end{array}
+    \]
+    \end{center}
+    $\canonizer(\perm{c_1, \ldots, c_n}{c})(r_{\sigma(1)}
+\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \newline
+= (\canonizer(c_1)(r_1)) \cdot (\canonizer(c)(s_1)) \cdot \ldots \cdot
+(\canonizer(c_n)(r_n))$
+  \caption{QRE canonizers}
+  \label{fig:canonizers}
+\end{figure}
+
+\begin{figure}[t]
   \centering
-\saz{Fix the projection case}
 \[
@@
       w \; \equiv_R \; w' &\iff& w = w' \\
-      w \; \equiv_{R \mapsto s} \; w' \text{ for all }w, w'\\
-      w \; \equiv_{\squash{R}{R'}{f}} \; w' &\iff& f(w) = w'
-      \text{ or } w = w' \\
+      w \; \equiv_{\project{R}{s}} \; w' \text{ for all }w, w'\\
+      w \; \equiv_{\squash{c}{c'}{f}} \; w' &\iff& f(w) \equiv_{c'} w'
+      \text{ or } w \equiv_{c'} w' \\
       w \; \equiv_{\normalize{R}{R'}{f}} \; w' &\iff&
       f(w)=f(w') \text{ or }w = w'\\
-      w \; \equiv_{q_2 \circ q_1} \; w' &\iff& \exists k, k' \in
-  \mathcal{L}(K(q_2)) \text{ such that } w \; \equiv_{q_1} \; k, \; w' \;
-  \equiv_{q_1} \; k', \text{ and } k \; \equiv_{q_2} \; k'\\
-      w \; \equiv_{q_1 \cdot q_2} \; w'  &\iff& w = r_1
-      \cdot r_2, \; w' = {r'}_1 \cdot {r'}_2 \text{ with } r_1 \; \equiv_{q_1}
-      \; {r'}_1, \; r_2 \; \equiv{q_n} \; {r'}_2\\
-      w \; \equiv_{q' \sep q} \; w' &\iff& w \; \equiv_{q_1} \; w'
-      \text{ or } \; w \; \equiv_{q_2} \; w'\\
-      w \; \equiv_{q*} \; w' &\iff& w = r_1 \cdot \ldots \cdot r_n, \; w'
-      = {r'}_1 \cdot \ldots \cdot {r'}_n \text{ and } r_i \equiv_{q} \; {r'}_i
+      w \; \equiv_{c_2 \; ; \; c_1} \; w' &\iff& \exists k, k' \in
+  \mathcal{L}(K(c_2)) \text{ such that } w \; \equiv_{c_1} \; k, \; w' \;
+  \equiv_{c_1} \; k', \text{ and } k \; \equiv_{c_2} \; k'\\
+      w \; \equiv_{c_1 \cdot c_2} \; w'  &\iff& w = r_1
+      \cdot r_2, \; w' = {r'}_1 \cdot {r'}_2 \text{ with } r_1 \; \equiv_{c_1}
+      \; {r'}_1, \; r_2 \; \equiv{c_n} \; {r'}_2\\
+      w \; \equiv_{c' \sep c} \; w' &\iff& w \; \equiv_{c_1} \; w'
+      \text{ or } \; w \; \equiv_{c_2} \; w'\\
+      w \; \equiv_{c^*} \; w' &\iff& w = r_1 \cdot \ldots \cdot r_n, \; w'
+      = {r'}_1 \cdot \ldots \cdot {r'}_n \text{ and } r_i \equiv_{c} \; {r'}_i
       \\
-      w \; \equiv_{\perm{q_1, \ldots, q_n}{q}} \; w' &\iff& w = r_{\sigma(1)}
+      w \; \equiv_{\perm{c_1, \ldots, c_n}{c}} \; w' &\iff& w = r_{\sigma(1)}
       \cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}, \;
@@
     & & \text{ for some } \sigma, \theta \in S_n, \text{ with } r_i \;
-    \equiv_{q_i} \; r'_i \text{ and } s_k \; \equiv_{q} \; s'_{k}
+    \equiv_{q_i} \; r'_i \text{ and } s_k \; \equiv_{c} \; s'_{k}
     \end{array}
@@
 \end{figure}
-\begin{figure}[t]
-  \begin{center}
-\[
-    \begin{array}{l@\quad l @\quad l} 
-      \canonizer(R) &=& id_{\mathcal{L}(R)} \\
-      \canonizer(R \mapsto s)(w) &=& s\\
-      \canonizer(\squash{R}{R'}{f})(w) &=& 
-\begin{cases}
-f(w) & \text{if } w \in \mathcal{L}(R)\\
-w & \text{otherwise}
-\end{cases}\\
-      \canonizer(\normalize{R}{R'}{f}) &=& f\\
-      \canonizer(q' \circ q) &=& \canonizer(q') \circ \canonizer(q)\\
-      \canonizer(q' \cdot q) &=& \canonizer(q') \cdot \canonizer(q)\\
-      \canonizer(q_1 \sep q_2)(w) &=& 
-\begin{cases}
-\canonizer(q_1)(w) & \text{if } w \in \mathcal{L}(W(q_1))\\
-\canonizer(q_2)(w) & \text{if } w \in \mathcal{L}(W(q_2))\\
-\end{cases}\\
-      \canonizer(q^*) &=& \canonizer(q)^* \\
-    \end{array}
-    \]
-    \end{center}
-    $\canonizer(\perm{q_1, \ldots, q_n}{q})(r_{\sigma(1)}
-\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \newline
-= (\canonizer(q_1)(r_1)) \cdot (\canonizer(q)(s_1)) \cdot \ldots \cdot
-(\canonizer(q_n)(r_n))$
-  \caption{QRE canonizers}
-  \label{fig:canonizers}
-\end{figure}
-
-\begin{figure}[p!]
-\centering
-\begin{prooftree}
-\AxiomC{$R$ is strongly unambiguous}
-\UnaryInfC{$\mathit{id}(R)$ is well formed}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$s \in W(q)$}
-\UnaryInfC{$R \mapsto s$ is well formed}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$\mathcal{L}(R) \cap \mathcal{L}(R') = \varnothing$}
-\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(R')$}
-\BinaryInfC{$squash(R, R', \ell)$ is well formed}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{${\substack{\forall \sigma \neq \theta, \; W(q_{\sigma(1)}) \cdot W(q)
-\cdot \ldots \cdot W(q_{\sigma(n)})\\ \cap W(q_{\theta(1)}) \cdot W(q)
-\cdot \ldots \cdot W(q_{\theta(n)}) =\varnothing}}$}
-\AxiomC{$q_i, q$ are well formed}
-\AxiomC{$\forall \sigma, \; K(q_{\sigma(1)}) \cdot^! K(q) \cdot^! \ldots 
-\cdot^! K(q_{\sigma(n)})$}
-\TrinaryInfC{$\perm{q_1, \ldots, q_n}{q}$ is well formed}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$\mathcal{L}(R') \subseteq \mathcal{L}(R)$}
-\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(R')$}
-\AxiomC{$f$ is surjective}
-\AxiomC{$f = f^2$}
-\QuaternaryInfC{$normalize \; (R,R', f)$ is well formed}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$q$ is well formed}
-\AxiomC{$q'$ is well formed}
-\AxiomC{$K(q) = W(q')$}
-\TrinaryInfC{$q' \circ q$ is well formed}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$q$ is well formed}
-\AxiomC{$q'$ is well formed}
-\AxiomC{$W(q) \cdot^! W(q')$}
-\AxiomC{$K(q) \cdot^! K(q')$}
-\QuaternaryInfC{$q \cdot q$ is well formed}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$q$ is well formed}
-\AxiomC{$q'$ is well formed}
-\AxiomC{$W(q) \cap W(q') = \varnothing$}
-\TrinaryInfC{$q \cdot q$ is well formed}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$q$ is well formed}
-\AxiomC{$W(q)^{*!}$}
-\AxiomC{$K(q)^{*!}$}
-\TrinaryInfC{$q^*$ is well formed}
-\end{prooftree}
-  \caption{Well-formed QREs \bcp{Maybe some or all of these can be
-      suppressed in the ``short version'' of the paper}\bcp{No, cancel
-      that.  These are important.  And we need to comment about the decidability
-    issues involving normalize.  (And a note that we don't really use
-    normalize, but rather introduce more specialized instances.)}
-  \saz{maybe shorten the judgment from ``is well formed'' to ``wf''?}
-}
-  \label{fig:qrerules}
-\end{figure}
-
-\saz{We need to explain the rules in Figure~\ref{fig:qrerules}---otherwise they
-  aren't mentioned anywhere in the text!}
-The theorem below confirms that the semantics of QREs are consistent with their
-intended behaviour:
+
+The formal semantics of QREs satisfy the following theorem:
 \begin{theorem}
-If $q$ is a well formed QRE, then
+If $c$ is a well formed QRE, then
 \begin{enumerate}
-  \item $W(q)$ and $K(q)$ are well defined regular expressions,
-  \item  $\eqrel{q}$ is an equivalence relation on $\mathcal{L}(W(q))$,
-  \item  $\mathcal{L}(K(q))$ forms a complete set of representatives for
-  $\eqrel{q}$,
-  \item $\canonizer(q):\mathcal{L}(W(q)) \longrightarrow \mathcal{L}(K(q))$ is a
-  well-defined function, and
-  \item  given any $w \in \mathcal{L}(W(q))$, $\canonizer(q)(w)$ is the unique
-  $k$ in $\mathcal{L}(K(q))$ such that $k$ is equivalent to $w$ modulo
-  $\eqrel{q}$.
+  \item $W(c)$ and $K(c)$ are well defined regular expressions,
+  \item  $\eqrel{c}$ is an equivalence relation on $\mathcal{L}(W(c))$,
+  \item  $\mathcal{L}(K(c))$ forms a complete set of representatives for
+  $\eqrel{c}$,
+  \item $\canonizer(c):\mathcal{L}(W(c)) \longrightarrow \mathcal{L}(K(c))$ is
+  a well-defined function, and
+  \item  given any $w \in \mathcal{L}(W(c))$, $\canonizer(c)(w)$ is the unique
+  $k$ in $\mathcal{L}(K(c))$ such that $k$ is equivalent to $w$ modulo
+  $\eqrel{c}$.
   \end{enumerate}
@@
 \label{QRE-lenses} 
-
-\saz{Throughout this section: do a which hunt to look for ``which''es that
-  should be ``that''s.}
-
 As we have seen, QREs express a broad class of equivalence relations
@@
 \textit{QRE Lenses}, a class of quotient lenses specified using QREs.
-
-\saz{We need to be careful with our metavariable conventions.  Up until now $q$
-  has meant ``a QRE'', but this sections says that $q$ and $q'$ are
-  \textit{lenses} and switches to using $c$ and $c'$ for QREs. We need to
-  pick a metavariable that ranges over QRE lenses and use it consistently and
-  switch these $c$'s to $q$'s.
-}
 
@@
 \noindent In other words, the get function of the QRE first canonizes the source
-data using $\mathit{Canonizer}(c)$ and then applies $f$ to the result, while the
+data using $\canonizer(c)$ and then applies $f$ to the result, while the
 put function of the QRE first canonizes the target data using
-$\mathit{Canonizer}(c')$ and the applies $f^{-1}$ to the result.
+$\canonizer(c')$ and the applies $f^{-1}$ to the result.
 
@@
 we define on them may not preserve this normal form. For example, the
-composition $q' \circ q$ of QRE lenses $q$ and $q'$ may not be equivalent to
-any quotient lens which has the canonical normal form of a bijective lens with
+composition $q' \; ; \; q$ of QRE lenses $q$ and $q'$ may not be equivalent to
+any quotient lens that has the canonical normal form of a bijective lens with
 canonizers a the edges. Consequently, in order to define quotient lenses, we
@@
 Boomerang. In the next subsection we briefly introduce bijective lenses and
-discuss the properties of bijective lenses which we will use to prove that QRE
+discuss the properties of bijective lenses that we will use to prove that QRE
 lenses have the normal form of bijective lens with canonizers at the edges.
@@
 
-\saz{include the denotations of const and $\circ$}
 \begin{align*}
 \llbracket R \rrbracket &= \{(r, r) \sep r \in \mathcal{L}(R)\}\\
+\llbracket const(r, s) \rrbracket &= \{(r, s)\}\\
 \llbracket swap(\ell, \ell') \rrbracket &= \{(s \cdot t, t' \cdot s') \sep
@@
 \cdot t_n) \sep (s_i, t_i) \in \llbracket \ell \rrbracket \text{ for } 1
-\leq i \leq n\}
+\leq i \leq n\}\\
+\llbracket \ell_2 \circ \ell_1 \rrbracket &= \{(r, t) \sep \text{there exists }s
+\text{ such that } (r, s) \in \llbracket \ell_1 \rrbracket \text{ and } (s, t)
+\in \llbracket \ell_2 \rrbracket\}
 \end{align*}
@@
  \begin{figure}[t]
-  \begin{prooftree}
+\begin{center}
 \AxiomC{$s_1 \in \Sigma^*$}
@@
 \BinaryInfC{$\mathit{const} \; (s_1, s_t): s_1 \Leftrightarrow s_2$}
-\end{prooftree}
-
-\begin{prooftree}
+\DisplayProof
+\hskip 1.5em
 \AxiomC{$R$ is strongly unambiguous}
 \UnaryInfC{$\mathit{id} \; (R): R \Leftrightarrow R$}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$\ell_1 : R_1 \Leftrightarrow S_1$}
-\AxiomC{$\ell_2 : R_2 \Leftrightarrow S_2$}
-\AxiomC{$R_1 \cdot^! R_2$}
-\AxiomC{$S_1 \cdot^! S_2$}
-\QuaternaryInfC{$\ell_1 \cdot \ell_2: R_1 \cdot R_2 \Leftrightarrow S_1 \cdot
+\DisplayProof
+\end{center}
+
+\begin{center}
+\AxiomC{$\substack{\ell_1 : R_1 \Leftrightarrow S_1 \\ \ell_2 : R_2
+\Leftrightarrow S_2}$}
+\AxiomC{$\substack{R_1 \cdot^! R_2 \\S_1 \cdot^! S_2 }$}
+\BinaryInfC{$\ell_1 \cdot \ell_2: R_1 \cdot R_2 \Leftrightarrow S_1 \cdot
 S_2$}
-\end{prooftree}
-
-\begin{prooftree}
+\DisplayProof
+\hskip 1.5em
+\AxiomC{$\substack{\ell_1 : R_1 \Leftrightarrow S_1 \\ \ell_2 : R_2
+\Leftrightarrow S_2}$}
+\AxiomC{$\substack{R_1 \cdot^! R_2 \\ S_2 \cdot^! S_1}$}
+\BinaryInfC{$\mathit{swap} \; (\ell_1, \ell_2): R_1 \cdot R_2
+\Leftrightarrow S_2 \cdot S_1$}
+\DisplayProof
+\end{center}
+
+\begin{center}
 \AxiomC{$\ell_1 : R_1 \Leftrightarrow R_2$}
@@
 \BinaryInfC{$\ell_2 \circ \ell_1: R_1 \Leftrightarrow R_3$}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$\ell_1 : R_1 \Leftrightarrow S_1$}
-\AxiomC{$\ell_2 : R_2 \Leftrightarrow S_2$}
-\AxiomC{$R_1 \cdot^! R_2$}
-\AxiomC{$S_2 \cdot^! S_1$}
-\QuaternaryInfC{$\mathit{swap} \; (\ell_1, \ell_2): R_1 \cdot R_2
-\Leftrightarrow S_2 \cdot S_1$}
-\end{prooftree}
-
-\begin{prooftree}
-\AxiomC{$\ell_1 : R_1 \Leftrightarrow S_1$}
-\AxiomC{$\ell_2 : R_2 \Leftrightarrow S_2$}
-\AxiomC{$\mathcal{L}(R_1) \cap \mathcal{L}(R_2) = \varnothing$}
-\AxiomC{$\mathcal{L}(S_1) \cap \mathcal{L}(S_2) = \varnothing$}
-\QuaternaryInfC{$\ell_1 \sep \ell_2: R_1 \sep R_2 \Leftrightarrow S_1 \sep S_2$}
-\end{prooftree}
-
-\begin{prooftree}
+\DisplayProof
+\hskip 1.5em
+\AxiomC{$\substack{\ell_1 : R_1 \Leftrightarrow S_1 \\ \ell_2 : R_2
+\Leftrightarrow S_2}$}
+\AxiomC{$\substack{\mathcal{L}(R_1) \cap \mathcal{L}(R_2) =
+\varnothing \\ \mathcal{L}(S_1) \cap \mathcal{L}(S_2) = \varnothing}$}
+\BinaryInfC{$\ell_1 \sep \ell_2: R_1 \sep R_2 \Leftrightarrow S_1 \sep S_2$}
+\DisplayProof
+\end{center}
+
+\begin{center}
 \AxiomC{$\ell : R \Leftrightarrow S$}
@@
 \TrinaryInfC{$\ell^*: R \Leftrightarrow S$}
-\end{prooftree}
-
-\begin{prooftree}
+\DisplayProof
+\hskip 1.5em
 \AxiomC{$\ell : R \Leftrightarrow S$}
@@
 \TrinaryInfC{$\ell : R' \Leftrightarrow S'$}
-\end{prooftree}
+\DisplayProof
+\end{center}
   \caption{Bijective Lens Typing Rules}
@@
 grammar:
-$$ q := \mathit{lift}(\ell) \sep \mathit{lquot}(q, \ell) \sep
-\mathit{rquot}(\ell, q) \sep q \cdot q' \sep (q \sep q') \sep q^* \sep q \circ q',$$
+$$ q := \kw{lift}(\ell) \sep \kw{lquot}(q, \ell) \sep
+\kw{rquot}(\ell, q) \sep q \cdot q' \sep (q \sep q') \sep q^* \sep q \circ q',$$
 where $\ell$ ranges over bijective lenses.
@@
 The QRE lens combinators are inspired by the Boomerang lens combinators of the
-same name. The $\mathit{lift}(\ell)$ quotient lens enables a bijective lens
+same name. The $\kw{lift}(\ell)$ quotient lens enables a bijective lens
 $\ell$ to be considered a quotient lens, with the equivalence relation applied
 to both the source and target formats being the equality relation. The
-regular combinators $\cdot \; * \; |$ on QRE lenses exhibit the expected
-behaviour, with the caveat that the regular expressions used by these lenses
-are strongly unambiguous. The composition operator $\circ$ also exhibits the
-expected behaviout, though the conditions that must be checked in order for the
-composition to be well defined are perhaps unintuitive. The reasons why
-these conditions are necesaary will become more apparent when proving normal
-forms for QRE lenses.
-
-\saz{Watch metavariables}
-The more interesting combinators are the $\mathit{lquot}$ and $\mathit{rquot}$
-combinators. The $\mathit{lquot}(c, q)$ combinator takes a quotient lens $q$ and
+regular combinators of concatenation `$\cdot$', Kleene star `$*$' and 
+`$|$' on QRE lenses exhibit the expected behaviour, with the caveat that these
+combinators are strongly unambiguous. The composition operator `$;$' also
+exhibits the expected behaviout, though the conditions that must be checked in
+order for the composition to be well defined are perhaps unintuitive. The
+reasons why these conditions are necesary will become more apparent when
+proving normal forms for QRE lenses.
+
+The more interesting combinators are the $\kw{lquot}$ and $\kw{rquot}$
+combinators. The $\kw{lquot}(c, q)$ combinator takes a quotient lens $q$ and
 quotients the source data using $c$, assuming that the source data forms a
 complete set of representatives for the equivalence relation $\eqrel{c}$. The
-$\mathit{lquot}(q, c)$ combinator does the same but on the view data.
+$\kw{lquot}(q, c)$ combinator does the same but on the view data.
 
@@
 respectively, and the bijective lens \cd{bib_to_end} that maps between
-\cd{bibtex} and \cd{endnote}. The quotient lens which maps between
+\cd{bibtex} and \cd{endnote}. The quotient lens that maps between
 \cd{bibtex} and \cd{endnote} is given by the QRE lens
@@
 
-\saz{metavariables\ldots}
 Each QRE lens $q$ has a type $q : c \Leftrightarrow c'$ where $c, c'$ are QREs.
@@
 \AxiomC{$\ell : R \Leftrightarrow S$}
-\UnaryInfC{$\mathit{lift}(\ell): id(R) \Leftrightarrow id(S)$}
+\UnaryInfC{$\kw{lift}(\ell): id(R) \Leftrightarrow id(S)$}
 \end{prooftree}
   \begin{align*}
-  \llbracket q \rrbracket.get &=  \llbracket \ell \rrbracket, \text{ and }\\
+  \llbracket q \rrbracket.get &=  \llbracket \ell \rrbracket\\
   \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1}
@@
 \AxiomC{$q' : c'  \Leftrightarrow c''$}
-\AxiomC{$c$ is well formed}
+\AxiomC{$c$ \textbf{WF}}
 \AxiomC{$K(c) = W(c')$}
-\TrinaryInfC{$\mathit{lquot}(c, q'): c' \circ c \Leftrightarrow c''$}
+\TrinaryInfC{$\kw{lquot}(c, q'): c' \circ c \Leftrightarrow c''$}
 \end{prooftree}
@@
   \AxiomC{$q' : c \Leftrightarrow c'$}
-  \AxiomC{$c''$ is well formed}
+  \AxiomC{$c''$ \textbf{WF}}
 
 \AxiomC{$K(c'') = W(c')$}
-\TrinaryInfC{$\mathit{rquot}(q', c):c \Leftrightarrow c'' \circ c'$}
+\TrinaryInfC{$\kw{rquot}(q', c):c \Leftrightarrow c'' \circ c'$}
 \end{prooftree}
@@
   \begin{prooftree}
-\AxiomC{$q_1 : c \Leftrightarrow c'$}
-\AxiomC{$q_2 : c'' \Leftrightarrow c'''$}
+\AxiomC{$\substack{q_1 : c \Leftrightarrow c' \\ q_2 : c'' \Leftrightarrow
+c'''}$}
 \AxiomC{$\substack{\mathcal{L}(W(c'))= \mathcal{L}(W(c'')) \\ K(c') \equiv^d
 K(c'')}$}
-\AxiomC{$\mathit{Canonizer}(c') = \mathit{Canonizer}(c'')$}
-\QuaternaryInfC{$q_2 \circ q_1: c \Leftrightarrow c'''$}
+\AxiomC{$\canonizer(c') = \canonizer(c'')$}
+\TrinaryInfC{$q_2 \circ q_1: c \Leftrightarrow c'''$}
 \end{prooftree}
@@
     \begin{prooftree}
-\AxiomC{$q' : c \Leftrightarrow c'$}
-\AxiomC{$W(c)^{*!}$ and $W(c')^{*!}$}
-\AxiomC{$K(c)^{*!}$ and $K(c')^{*!}$}
-\TrinaryInfC{${q'}^* : c^* \Leftrightarrow {c'}^*$}
+\AxiomC{$q : c \Leftrightarrow c$}
+\AxiomC{$\substack{W(c)^{*!} \text{ and } W(c')^{*!} \\ K(c)^{*!} \text{ and }
+K(c')^{*!}}$}
+\BinaryInfC{${q}^* : c^* \Leftrightarrow {c'}^*$}
 \end{prooftree}
   \begin{align*}
-  \llbracket q \rrbracket.get &= (\llbracket q' \rrbracket.get)^*, \text{
-  and }\\
-  \llbracket q \rrbracket.put &= (\llbracket q' \rrbracket.put)^*
+  \llbracket q \rrbracket.get &= (\llbracket q \rrbracket.get)^* \\
+  \llbracket q \rrbracket.put &= (\llbracket q \rrbracket.put)^*
   \end{align*}
@@
     \begin{prooftree}
-\AxiomC{$q_1 : c_1 \Leftrightarrow d_1 $}
-\AxiomC{$q_2 : c_2 \Leftrightarrow d_2$}
+\AxiomC{$\substack{q_1 : c_1 \Leftrightarrow d_1 \\ q_2 : c_2 \Leftrightarrow
+d_2}$}
 \AxiomC{${\substack{W(c_1) \cdot^! W(c_2)\\ K(c_1) \cdot^! K(c_2)}}$}
 \AxiomC{${\substack{W(d_1) \cdot^! W(d_2)\\ K(d_1) \cdot^! K(d_2)}}$}
-\QuaternaryInfC{$q_1 \cdot q_2: c_1 \cdot c_2
+\TrinaryInfC{$q_1 \cdot q_2: c_1 \cdot c_2
 \Leftrightarrow d_1 \cdot d_2$}
@@
   \llbracket q \rrbracket.get &= \llbracket q_1 \rrbracket.get \cdot \llbracket
-  q_2 \rrbracket.get, \text{ and }\\
+  q_2 \rrbracket.get \\
   \llbracket q \rrbracket.put &= \llbracket q_1 \rrbracket.put \cdot \llbracket
@@
       \begin{prooftree}
-\AxiomC{$q_1 : c_1 \Leftrightarrow d_1 $}
-\AxiomC{$q_2 : c_2 \Leftrightarrow d_2$}
-\AxiomC{$\mathcal{L}(W(c_1)) \cap \mathcal{L}(W(c_2)) = \varnothing$}
-\AxiomC{$\mathcal{L}(W(d_1)) \cap \mathcal{L}(W(d_2)) = \varnothing$}
-\QuaternaryInfC{$q_1 \sep q_2: (c_1 \sep c_2)
+\AxiomC{$\substack{q_1 : c_1 \Leftrightarrow d_1 \\ q_2 : c_2 \Leftrightarrow
+d_2}$}
+\AxiomC{$\substack{\mathcal{L}(W(c_1)) \cap \mathcal{L}(W(c_2)) = \varnothing
+\\ \mathcal{L}(W(d_1)) \cap \mathcal{L}(W(d_2)) = \varnothing}$}
+\BinaryInfC{$q_1 \sep q_2: (c_1 \sep c_2)
 \Leftrightarrow (d_1 \sep d_2)$}
@@
 the kernel regular expressions are equivalent via the star-semiring axioms.
-Finally, the condition $\mathit{Canonizer}(c') = \mathit{Canonizer}(c'')$ says
+Finally, the condition $\canonizer(c') = \canonizer(c'')$ says
 that $c'$ and $c''$ define the same equivalence relation on $\mathcal{L}(W(c'))=
@@
 We are permitted to claim the third step from the second since
-$\mathit{Canonizer(c')}$ is the identity function on $K(c'')$ which is
+$\canonizer(c')$ is the identity function on $K(c'')$ which is
 equivalent to $K(c')$ via the star semiring axioms by assumption.
@@
   \item
-  $\mathit{lift}(\ell): R/\mathit{id}(R) \Leftrightarrow S/\mathit{id}(S)$ where
+  $\kw{lift}(\ell): R/\mathit{id}(R) \Leftrightarrow S/\mathit{id}(S)$ where
   $\ell :
@@
   \begin{align*}
-  \llbracket \mathit{lift}(\ell) \rrbracket.get &=  \llbracket \ell \rrbracket
+  \llbracket \kw{lift}(\ell) \rrbracket.get &=  \llbracket \ell \rrbracket
   = \llbracket \ell \rrbracket \circ id_{\mathcal{L}(R)} =
   \llbracket \ell \rrbracket \circ \canonizer(\mathit{id}(R)), \text{ and }\\
-  \llbracket \mathit{lift}(\ell) \rrbracket.put &= \llbracket \ell
+  \llbracket \kw{lift}(\ell) \rrbracket.put &= \llbracket \ell
   \rrbracket^{-1} = \llbracket \ell \rrbracket^{-1} \circ id_{\mathcal{L}(S)} =
@@
   \item
-  $\mathit{lquot}(c, q'): c' \circ c \Leftrightarrow c''$ where $q' : c' 
+  $\kw{lquot}(c, q'): c' \circ c \Leftrightarrow c''$ where $q' : c' 
   \Leftrightarrow c''$, $c$ is well formed and $K(c) = W(c')$. Then
@@
   \item
-  $\mathit{rquot}(q', c''):c \Leftrightarrow c'' \circ c'$ where $q' : c \Leftrightarrow
+  $\kw{rquot}(q', c''):c \Leftrightarrow c'' \circ c'$ where $q' : c \Leftrightarrow
   c'$, $c''$ is well formed and $K(c'') = W(c')$. Proceed as in the previous
@@
   \item 
-  return $q = \mathit{rquot}(\mathit{lquot}(c, \ell), c') : c \Leftrightarrow
+  return $q = \kw{rquot}(\kw{lquot}(c, \ell), c') : c \Leftrightarrow
   c'$
@@
 \cd{endnote} that maps \cd{bib_example} to \cd{end_example} can be
-expressed by a bijective lens, it suffices to type the expression
+expressed by a bijective lens, we can synthesize a lens that maps between the
+QREs \cd{bibtex} and \cd{endnote}. We do this using the expression
 
 \begin{lstlisting}
-let bib_to_end : (lens in bibtex <=> endnot) 
+let bib_to_end : (lens in bibtex <=> endnote) 
   = synth bibtex <=> endnote  using  {(bib_example, end_example)}
@@
 {\em view-update problem} in the database community, where the challenge is to
-derive a program which extracts a view of data from a source, as well as a
-program which folds view updates back into the source safely and correctly.
+derive a program that extracts a view of data from a source, as well as a
+program that folds view updates back into the source safely and correctly.
 
@@
 regular expressions. Also, this combinator permutes regular expressions rather
-than canonizers, thus making it difficult to express nested permutations which
+than canonizers, thus making it difficult to express nested permutations that
 occur in many data formats, especially XML and XML-like formats.
@@
 In their paper, Foster et al also discuss other bidirectional programming
-languages which support quotienting of data including XSugar~\cite{xsugar},
+languages that support quotienting of data including XSugar~\cite{xsugar},
 biXid~\cite{bixid} and X/Inv~\cite{xinv}. XSugar programs bidirectionally
@@
 transformations specified by a pair of unambiguous grammars. The quotienting
-occurs on the XML side by use of a generic canonizer which standardizes the
+occurs on the XML side by use of a generic canonizer that standardizes the
 representation of trees. Well-formed XSugar programs are guaranteed to be
@@
 This ambiguity is what allows biXid to express equivalences on the data.
-Finally, Foster et al discuss a possible connection with the language languages
+Finally, Foster et al discuss a possible connection with the languages
 X and Inv which support a primitive duplication combinator that does not work
@@
 Furthermore, neither of these tools were able to infer transformations 
-which occurred under two iterations.
+that occurred under two iterations.
 
@@
 details that satisfy the user's specifications.
-This approach works well in large search spaces, which
-require some enumerative search.
-One important way in which Optician differs from Morpheus is that in
+This approach works well in large search spaces that require some enumerative
+search.
+One important way that Optician differs from Morpheus is that in
 Morpheus, an outline is a sketch---an

### 2017-11-10 papers/QuotientLensSynth/main.tex
Made changes

@@
 \section{Introduction}
-A {\em lens} $\ell$ is a {\em bidirectional transformation} between a source
-type $S$ and a target type $T$ in which the forward transformation
-$\ell.\get$ and the backward transformation $\ell.\lput$ satisfy certain
-round-tripping laws such as invertibility. Lenses are advantageous
-because writing a single lens rather than two separate uni-directional
-transformations saves both keystrokes and bugs, by reducing redundancy and by
-automatically guaranteeing strong ``lens laws'' ensuring that the forward and
-backward transformations interact well.
+{\em Lenses}~\cite{Lenses} specify bidirectional 
+transformations between different data formats.  Writing a single lens
+rather than two separate uni-directional transformations saves both
+keystrokes and bugs, by reducing redundancy and by automatically
+guaranteeing strong ``lens laws'' ensuring that the forward and backward
+transformations interact well.
 
@@
 share a great deal of structure.  Moreover, writing the lens between the
-equivalence classes can be challenging due to the rather fiddly typechecking, mainly due to
-``non-ambiguity'' conditions in the lens well-formedness rules.
-
-This paper addresses both of these issues.  First, in Section~\ref{sec:example} we
-explain the difficulties of programming with quotient lenses and show, by way of
-example, how our novel language of {\em Quotient Regular Expressions}, or QREs
-and QRE lenses address these difficulties.  Quotient regular expressions allow
-programmers to specify a broad class of equivalence relations over regular
-languages using a compact and convenient notation in which the regular language
-and its equivalence classes are captured by a single expression.  We then
-define {\em QRE lenses}, a class of bijective quotient lenses that are typed
-using QREs, and we show how to {\em synthesize} QRE lenses from their QRE types
-and a set of input-output examples.  We demonstrate the practicality of this
-approach by synthesizing QRE lenses between several real-world data formats
-from the {\tt data.gov} database.
-
-\saz{There is still some redundancy between the following contributions list and
-  the paragraph above.}
+equivalence classes can be challenging due to the rather fiddly typechecking,
+mainly due to ``non-ambiguity'' conditions in the lens well-formedness rules.
+
+In this paper, we make programming quotient lenses easier in two ways. First, we
+introduce a novel language of {\em Quotient Regular Expressions}, or QREs, which
+enable a programmer to simultaneously specify a regular expression as well as an
+equivalence relation on that regular expression. QREs make programming quotient
+lenses easire by saving the programmer the effort of defining the data and the
+equivalence relation on the data separately.
+
+Secondly, we use QREs to define a class of quotient lenses called {\em QRE
+lenses} which map betweeen QREs, and show how to synthesize QRE lenses given
+an specification described by pair of QREs and a set of input-output examples.
+This way, the programmer need not manually derive the lens herself. The net
+effect of these two contributions, that is QREs and QRE lens synthesis, is that
+in order to derive a quotient lens, the programmer will only need to specify the
+data formats on each side of the desired quotient lens and possibly a set of
+example input-output pairs that guide synthesis of the lens, a marked
+improvement over the current situation.
 
@@
 \item We introduce a novel language of {\em Quotient Regular Expressions}
-(QREs) that provide compact, convenient notation for a natural and useful
+(QREs) that a provide compact, convenient notation for a natural and useful
 class of equivalence relations on regular languages (Section~\ref{QRE}).
@@
 \label{sec:example}
+In this section, we explain the difficulties of programming with quotient lenses
+and show, by way of example, how our novel language of {\em Quotient Regular
+Expressions}, or QREs and QRE lenses address these difficulties.
 
@@
 should allow whitespace characters to occur arbitrarily between the words in the
-label, author and title fields as well as in between them. Second, an
+label, author and title fields as well, as in between them. Second, an
 author's name can have the surname followed by a comma and then the other
@@
 Boomerang, actually doing so is a challenge. This is because the entire
-equivalence is composed of three separate equivalences \saz{explicitly mention
-  whitespace and permutation?} which nest and interact
-with each other in a way that makes them difficult to express with the standard
-Boomerang combinators. Moreover, in many cases, Boomerang requires that the
-programmer first specify the data itself in the form of a regular expression
-before she can define the equivalence relation on that data. This leads to
-verbose, repetitive code.
-
-Instead, we propose {\em Quotient Regular Expressions} or QREs which enable
-the programmer to define a regular expression $R$ as well as an equivalence
-relation on $R$ in the same expression.For example, with QREs we can specify
-\bibtex{} records as shown in Figure~\ref{fig:example-qre}.
+equivalence is composed of three separate equivalences---whitespace quotienting
+inside fields, whitespace quotenting in between fields, and permutation
+quotenting of fields---nest with each other in a way that makes them difficult
+to express with the standard Boomerang combinators.
+
+In general, there is also another issue in describing equivalences in
+Boomerang, which is that that the programmer often has to first specify the data
+itself in the form of a regular expression before she can define the
+equivalence relation on that data. This leads to repetitive code since the
+regular expression that defines the data and the equivalence relation defined on
+the regular expression follow the same syntactic patter---one can do both things
+with the same code with only a few changes.
+
+\sam{TODO: PUT EXAMPLE HERE!}
+
+To address these issues, we propose {\em Quotient Regular Expressions} or QREs
+which enable the programmer to define a regular expression $R$ as well as an
+equivalence relation on $R$ in the same expression. For example, with QREs we
+can specify \bibtex{} records as shown in Figure~\ref{fig:example-qre}.
 
@@
 
-(* project QRE maps arbitrary occurrences of whitespace to a single space *)
+(* map arbitrary occurrences of whitespace to a single space *)
 let wsp_to_space   = project wsp$^+$ $\to$ " "
@@
 
-(* perm QRE allows permutations of fields with arbitrary whitespace in between *)
+(* allow any permutation of fields interspersed with arbitrary whitespace *)
 let bib_perm  = perm (label, bib_author, bib_title) with (project ("," .wsp$^+$) $\to$ ",\n")
@@
 
-In Figure~\ref{fig:example-qre}, the QRE \cd{wsp_to_space}, which is defined to
-be
-
-\noindent \cd{project wsp}+ $\to$ \cd{" "}, matches any non-zero number of
-whitespace characters, but considers the matched string of whitespace
-characters as equivalent with a single space as the canonical representative. 
-We call this QRE a \textit{projection} because it projects an
-entire regular language down to a single representative string in the
-language.
-
-Another form of QRE is the \textit{squash} QRE, an example of which is given by
-\cd{valid_name}, which is defined to be:
+In Figure~\ref{fig:example-qre}, the QRE \cd{project wsp}+ $\to$ \cd{" "}
+matches any non-zero number of whitespace characters, but considers the matched
+string of whitespace characters as equivalent with a single space as the
+canonical representative. We call this QRE a \textit{projection} because it
+projects an entire regular language down to a single representative string in
+the language.
+
+Another QRE in Figure~\ref{fig:example-qre} is given by
 
@@
 
-\noindent This QRE specifies that the \cd{comma_name} representation of the
-author's name is equivalent to the \cd{space_name} representation.  Here \cd{get
-  space_to_comma} is the function that transforms names in the \cd{comma_name}
-representation to names in the \cd{space_name} representation.  In this
-  example, \cd{space_to_comma} is itself an automatically \textit{synthesized}
-  lens between the \cd{space_name} and \cd{comma_name} QREs, and we need to use
-  the ``get'' direction of the lens to specify the squash transformation. (We'll
-  say more about synthesis below.)
-
-The third example of a QRE in the Figure is \cd{bib_perm}, defined as:
+\noindent This QRE is a \textit{squash} QRE. A \textit{squash} QRE ``glues''
+two data formats using a user-defined function. For example,
+the QRE in the code above specifies that the \cd{comma_name} representation of
+the author's name is equivalent to the \cd{space_name} representation, with
+\cd{space_name} taken to be the canonical representation. The
+function that identifies the two formats is given by \cd{get space_to_comma},
+which transforms names in the \cd{comma_name} representation to names in the
+\cd{space_name} representation. In this example, \cd{space_to_comma} is itself
+an automatically \textit{synthesized} lens between the \cd{space_name} and
+\cd{comma_name} QREs, and we need to use the ``get'' direction of the lens to
+specify the squash transformation. (We'll say more about synthesis below.)
+
+The third example of a QRE in the Figure is \cd{bib_perm}, defined by:
 \begin{center}
@@
 
-\noindent This is a \textit{permutation} QRE, which specifies that the label,
+\noindent This is a \textit{permutation} QRE that specifies that the label,
 title, and author fields can occur in any order, with the fields separated by a
@@
 is that in which the label, author and title fields occur in that order,
-separated by a comma followed a newline.
-Such \cd{perm} QREs allow for easy nesting of permutations. Similarly, it is easy
-to specify any number of fields within the \cd{perm} QRE.
+separated by a comma followed a newline. The \textit{permutation} QREs is
+designed for equivalences such as this, where different permutations of
+equivalences are interspersed with a separator which itself has an
+equivalence defined on it.
 
@@
   expression combinators of alternative choice `\cd{|}', concatenation `\cd{.}'
-  and Kleene star `$^*$', and also sequential composition ``;''.  As shown in
-  the rest of Figure~\ref{fig:example-qre}, it is straightforward to define both a
-  regular language that accepts \bibtex{} records \textit{and}, simultaneously,
-  an equivalence relation that quotients out irrelevant details.
+  and Kleene star `$^*$', and also sequential composition ``;''.
 
 \subsection{QRE Lenses}
-QREs make it easier to write quotient lenses by specifying the data and the
-equivalence relation on the data in one expression, but to use them to define a
-lens, the programmer still has to manually derive the underlying lens that maps
-the equivalence classes of the source data to the equivalence classes of the
-target data.
+While QREs make it easier to write equivalence relations on regular expressions
+by specifying the data and the equivalence relation on the data in one
+expression, in order to define a quotient lens, the programmer still has to
+manually derive the underlying lens that maps the equivalence classes of the
+source data to the equivalence classes of the target data. This is itself a
+tricky task as Figure~\ref{fig:example-lens} illustrates.
 
@@
 
-This is itself a tedious and potentially error-prone task, as the code snippet
-in Figure~\ref{fig:example-lens} illustrates.  The lens shown in the Figure
-transforms \bibtex{} data to EndNote data  (and vice versa) by matching
-the delimiters and separators in a bijective manner. For example, the
-\cd{preamble} lens replaces the \bibtex{} preamble \verb|@Book| with the EndNote
-preamble \verb|%0 Book\n%F|. When the programmer is writing the lens, she has
-to make sure that all of these small deletions and insertions are correct,
-which can be a challenging task if the lens is large enough. This makes lens
-programming tedious and error-prone.
-
-To address this challenge, we identify a  class of quotient lenses that we call
-{\em QRE lenses}, and show how to {\em synthesize} QRE lenses from a pair of
-QREs and a set of input-output example pairs. Our key insight is to reduce the
-problem of synthesizing QRE lenses to the problem of synthesizing {\em
-bijective lenses}, which are the bijections created from the Boomerang lens
-combinators. This is a problem which we solved in our prior work ~\cite{popl18}.
+The lens shown in Figure~\ref{fig:example-lens} transforms \bibtex{} data to
+EndNote data  (and vice versa) by matching the delimiters and separators in a
+bijective manner. For example, the \cd{preamble} lens replaces the \bibtex{}
+preamble \verb|@Book| with the EndNote preamble \verb|%0 Book\n%F|. When the
+programmer is writing the lens, she has to make sure that all of these small
+deletions and insertions are correct, which can be a challenging task if the
+lens is large enough, as in the definition of the lens \verb!author_lens! in
+Figure~\ref{fig:example-lens}. This makes lens programming tedious and
+error-prone.
+
+Consequently, we offer the programmer some relief by identifying a class of
+quotient lenses that we call {\em QRE lenses}, and show how to {\em synthesize}
+QRE lenses from a pair of QREs and a set of input-output example pairs. Our key
+insight is to reduce the problem of synthesizing QRE lenses to the problem of
+synthesizing {\em bijective lenses}, which are the bijections created from the
+Boomerang lens combinators. This is a problem which we solved in our prior work
+~\cite{popl18}.
 
@@
 composition \cd{(q_1 ; q2) : bibtex} $\Leftrightarrow$ \cd{xml} of \cd{q_1} and
-\cd{q_2} is not (necessarily) in normal form due to the presence of canonizers
+\cd{q_2} is not necessarily in normal form due to the presence of canonizers
 on each side of the intermediary EndNote format:
@@
 shows that there is another QRE lens \cd{q : bibtex} $\Leftrightarrow$ \cd{xml}
-of the same type as \cd{q_1 ; q_2} such that \cd{q_1 ; q_2} and \cd{q} define
-the same transformation. This proof will allow justify our approach for
-synthesizing quotient lenses so that the programmer can replace all of the
-code in Figure~\ref{fig:example-lens} with a call to the synthesis prodedure:
+of the same type as \cd{(q_1 ; q2) : bibtex} $\Leftrightarrow$ \cd{xml} such
+that \cd{q_1 ; q_2} and \cd{q} define the same transformation. This proof
+will allow justify our approach for synthesizing quotient lenses so that the
+programmer can replace all of the code in Figure~\ref{fig:example-lens} with a
+call to the synthesis prodedure:
 
@@
 functions between regular languages, and $s$ ranges over character strings. In
-subsequent sections, we will use the the notation $\mathcal{L}(R)$ to mean the
-language accepted by the regular expression $R$.
+the following sections, we will use the the notation $\mathcal{L}(R)$ to mean
+the language accepted by the regular expression $R$.
 
@@
 \begin{lstlisting}
-(* Regular expressions matching different representations *)
+(* Regular expressions matching different comma and space representations *)
 let comma_name = last_name . "," . wsp+ . first_name
@@
 
-(* Squash comma representation to space representation *)
+(* Squash comma representation into space representation *)
 let comma_to_space = 
@@
 
-(* project whitespace in space_name representation *)
+(* Project away whitespace in space_name representation *)
 let project_in_space = first_name . (project wsp+ $\to$ " ") . last_name
@@
 The equivalence relations definable by QREs are a strict subset of those that
-can be expressed in Boomerang. For instance, in Boomerang, one can map a
-regular language described by $R$ to a regular language described by $S$, with
-$\mathcal{L}(S) \not \subseteq \mathcal{L}(R)$, whereas each QRE $c$ canonizes
-a regular language described by $R$ to regular subset $\mathcal{L}(R')
-\subseteq \mathcal{L}(R)$ described by $R'$.
-
-The net effect of these design decisions means that each QRE $c$ enables us to
-simultaneously express
+can be expressed in Boomerang. For instance, in Boomerang, one can use any lens
+to canonize a regular language described by $R$ to a regular language described
+by $S$, with $\mathcal{L}(S) \not \subseteq \mathcal{L}(R)$, whereas each QRE
+$c$ canonizes a regular language described by $R$ to regular subset
+$\mathcal{L}(R') \subseteq \mathcal{L}(R)$ described by $R'$. QREs therefore are
+not as expressive as Boomerang canonizers. On the other hand, QREs enable us to
+express many equivalences which occur in real world data as we shall demonstrate
+in the evaluation section (Section~\ref{impl}).
+
+The net effect of these design decisions means that each QRE $c$ enables
+us to simultaneously express
 \begin{enumerate}
@@
 
-Figure~\ref{fig:qrerules} gives the inferrence rules for derving well-formed
+Figure~\ref{fig:qrerules} gives the inferrence rules for deriving well-formed
 QREs. As expected, the unambiguity conditions come into play when defining QREs
@@
 
-\noindent This rule says that we can concatenate two QREs $c$ and $c'$ only when
-the concatentions of $W(c)$ and $W(c')$, and $K(c)$ and $K(c')$ are ambiguous.
+\noindent This rule says that the concatenation $c \cdot c'$ of QREs $c$ and
+$c'$ is well formed only if the concatentions of $W(c)$ and $W(c')$, and
+$K(c)$ and $K(c')$ are ambiguous.
 
@@
 the language $K(c_1) \cdot K(c) \cdot \ldots \cdot K(c) \cdot K(c_n)$ which is
-the kernel of the \textit{permutation} QRE, and as we mentioned, the underlying
+the kernel of the \textit{permutation} QRE. As we mentioned, the underlying
 lens will require that the source and target regular expressions of the lens be
@@
   \end{enumerate}
-Figures~\ref{fig:wk} gives the inductive rules that define the whole language
-$W(c)$ of a QRE. The two interesting rules are the rules for the
-\textit{squash}, \textit{normalize} and \textit{permutation} combinators. If $c
-= \squash{c}{c'}{f}$, then the whole language of $c$ is $W(c) \sep W(c')$. This
-is because the \textit{squash} combinator takes the whole language $W(c)$ of
-$c$, the whole language $W(c')$ of $c'$ and a function $f : \mathcal{L}(W(c))
-\longrightarrow \mathcal{L}(W(c'))$, maps $\mathcal{L}(W(c))$ into
-$\mathcal{L}(W(c'))$ using $f$ and then canonizes $W(c')$ into $K(c')$ using
-$\canonizer(c')$.
+Figures~\ref{fig:wk} gives the inductive definitions for the whole language
+$W(c)$ of a QRE. The \textit{squash} and \textit{permutation} combinators
+give the two interesting definitions . If $c = \squash{c}{c'}{f}$, then
+the whole language of $c$ is $W(c) \sep W(c')$. This is because the
+\textit{squash} combinator takes the whole language $W(c)$ of $c$, the whole
+language $W(c')$ of $c'$ and a function $f : \mathcal{L}(W(c)) \longrightarrow
+\mathcal{L}(W(c'))$, maps $\mathcal{L}(W(c))$ into $\mathcal{L}(W(c'))$ using
+$f$ and then canonizes $W(c')$ into $K(c')$ using $\canonizer(c')$.
 
@@
 \ldots \cdot K(c) \cdot K(c_n)$, because the canonical permutation is the
-identity permuation, and then each of the parts of the string that match $c_i$
-and $c$ are canonized into $K(c_i)$ and $K(c)$ respectively.
+identity permuation, with each of the parts of the string that match $c_i$
+and $c$ canonized into $K(c_i)$ and $K(c)$ respectively.
 
 Figure~\ref{fig:canonizers} gives the inductive definitions of the
-$\canonizer{}$ function for each QRE. Again, the the least obvious definition
-is
+$\canonizer{}$ function for each QRE. Again, the \textit{permutation} combinator
+gives the most interesting definition,
 
@@
 
-Our approach in defining QRE lenses is as follows. Let $c$ and $c'$ be QREs. Given a
-bijection $\ell : K(c) \Leftrightarrow K(c')$, we may define a quotient lens $q
-: W(c)/\eqrel{c} \Leftrightarrow W(c')/\eqrel{c'}$ by
+Given regular expressions $R, S$ and equivalence relations
+$\sim_R$ and $\sim_S$ defined on $\mathcal{L}(R)$ and $\mathcal{L}(S)$
+respectively, we define a {\em bijective quotient lens} $q : R /{\sim_R}
+\Leftrightarrow S/{\sim_S}$ to be a pair of functions $q.get :
+\mathcal{L}(R) \longrightarrow \mathcal{L}(S)$ and $q.put : \mathcal{L}(S)
+\longrightarrow \mathcal{L}(R)$ such that for all $r \in \mathcal{L}(R)$ and $s
+\in \mathcal{L}(S)$,
+\begin{align*}
+q.put(q.get(r)) &\sim_R r \text{, and}\\
+q.get(q.put(s)) &\sim_R s
+\end{align*}
+Moreover, if $r \sim_R r'$, then $q.get(r) = q.get(r')$, and if $s \sim_S s'$,
+then $q.put(s) = q.put(s')$.
+
+In words, a bijective quotient lens $q$ from $R$ to $S$ modulo $\sim_R$
+and $\sim_S$ is a pair of functions $q.get$ and $q.put$ such that $q.get$
+respects $\sim_R$ and $q.put$ respects $\sim_S$, and such that the respective
+functions defined on $\mathcal{L}(R)/{\sim_R}$ and $\mathcal{L}(S)/{\sim_S}$ are
+mutual inverses of each other.
+
+Now let $c$ and $c'$ be QREs. Given a bijection $\ell : K(c) \Leftrightarrow
+K(c')$, we may define a quotient lens $q : W(c)/\eqrel{c} \Leftrightarrow
+W(c')/\eqrel{c'}$ by
 \begin{equation}\label{normalform}
@@
 
-\noindent In other words, the get function of the QRE first canonizes the source
+\noindent In other words, the get function of the $q$ first canonizes the source
 data using $\canonizer(c)$ and then applies $f$ to the result, while the
-put function of the QRE first canonizes the target data using
-$\canonizer(c')$ and the applies $f^{-1}$ to the result.
+put function of the $q$ first canonizes the target data using
+$\canonizer(c')$ and the applies $f^{-1}$ to the result. This is our approach in
+defining QRE lenses.
 
@@
 any quotient lens that has the canonical normal form of a bijective lens with
-canonizers a the edges. Consequently, in order to define quotient lenses, we
-need to choose a set $S$ of bijections that has the right properties in the
-sense that QRE lenses built from the bijections in $S$ are closed under the
-action of QRE lens combinators. Our choice is to use the class of
-\textit{bijective lenses}, which are the bijections that can be expressed in
-Boomerang. In the next subsection we briefly introduce bijective lenses and
-discuss the properties of bijective lenses that we will use to prove that QRE
-lenses have the normal form of bijective lens with canonizers at the edges.
+canonizers a the edges. Consequently, in order to guarantee that every QRE
+lens $q$ is equivalent to a QRE lens $q'$ which has the canonical form of
+a bijection with canonizers at the edges, we need to identify a set $S$ of
+bijections that will ensure that QRE lenses are closed under the action of QRE
+lens combinators. Our choice is to use the class of \textit{bijective lenses},
+which we briefly discuss in the next section.
 
@@
 
-The more interesting combinators are the $\kw{lquot}$ and $\kw{rquot}$
+The interesting combinators are the $\kw{lquot}$ and $\kw{rquot}$
 combinators. The $\kw{lquot}(c, q)$ combinator takes a quotient lens $q$ and
@@
 complete set of representatives for the equivalence relation $\eqrel{c}$. The
-$\kw{lquot}(q, c)$ combinator does the same but on the view data.
+$\kw{lquot}(q, c)$ combinator does the same but on the target data.
 
@@
 \Longleftrightarrow W(c')/{\eqrel{c'}}$. The typing rules and denotation of QRE
-lenses are given in Figure~\ref{fig:qlenssemantics}.
-
-The following theorem confirms that the semantics for QRE lenses are correct:
+lenses are given in Figure~\ref{fig:qlenssemantics}. The trickiest typing rule
+is the typing rule for composition:
+
+  \begin{prooftree}
+\AxiomC{$\substack{q_1 : c \Leftrightarrow c' \\ q_2 : c'' \Leftrightarrow
+c'''}$}
+\AxiomC{$\substack{\mathcal{L}(W(c'))= \mathcal{L}(W(c'')) \\ K(c') \equiv
+K(c'')}$}
+\AxiomC{$\canonizer(c') = \canonizer(c'')$}
+\TrinaryInfC{$q_2 \circ q_1: c \Leftrightarrow c'''$}
+\end{prooftree}
+
+This derivation rule essentially says that the composition $q_2 \circ q_1$ is
+well defined if and only if the intermediary QREs $c''$ and $c'''$ define the
+same equivalence relation on the same regular language. The condition
+$\mathcal{L}(W(c')) = \mathcal{L}(W(c''))$ says that the intermediary language
+is the same on both sides, while the condition $K(c') \equiv K(c'')$ says that
+the kernel regular expressions are syntactically the same. Finally, the
+condition $\canonizer(c') = \canonizer(c'')$ says that $c'$ and $c''$ define
+the same equivalence relation on $\mathcal{L}(W(c'))= \mathcal{L}(W(c''))$.
+
+In general, checking that $\canonizer(c') = \canonizer(c'')$ is an undecidable
+task, especially because of the $\normalize{R}{R'}{f}$ QRE. Therefore, our
+actual implementation of the composition rule for QRE lenses follows Boomerang's
+approach, which is to define the composition on QRE lenses if and only if
+$c'$ and $c''$ are syntactially equal to $id(R)$ and $id(R')$ for some $R$and
+$R'$ satisfying $\mathcal{L}(R) = \mathcal{L}(R')$.
+
+Foster et al based this approach on two observations: first, that most quotient
+lenses originate as lifted basic lenses, and therefore have types whose
+equivalence relations are both equality; and second, that equality is preserved
+by many of the quotient lens combinators. Foster et al also discuss a second
+possible approach, where equivalence relations are represented by rational
+functions that induce them. While this second approach is more refined than the
+first, Boomerang favours the first approach since the second appears to be too
+expensive to be useful in practice~\cite{quotientlenses}.
+
+The semantics defined on QRE lenses imply the following theorem:
 \begin{theorem}
@@
 c'''}$}
-\AxiomC{$\substack{\mathcal{L}(W(c'))= \mathcal{L}(W(c'')) \\ K(c') \equiv^d
+\AxiomC{$\substack{\mathcal{L}(W(c'))= \mathcal{L}(W(c'')) \\ K(c') \equiv
 K(c'')}$}
@@
 for some bijective lens $\ell$. In other words, each QRE lens is the same as a
-bijective lens with canonizers at the edges. The following theorem confirms that
-this indeed is the case:
+bijective lens with canonizers at the edges. The following theorem,
+which is the main technical contribution of this paper, confirms that this
+indeed is the case:
 
@@
   \begin{prooftree}
-\AxiomC{$q_1 : c \Leftrightarrow c'$}
-\AxiomC{$q_2 : c'' \Leftrightarrow c'''$}
-\AxiomC{$\substack{\mathcal{L}(W(c'))= \mathcal{L}(W(c'')) \\ K(c') \equiv^d
+\AxiomC{$\substack{q_1 : c \Leftrightarrow c' \\ q_2 : c'' \Leftrightarrow
+c'''}$}
+\AxiomC{$\substack{\mathcal{L}(W(c'))= \mathcal{L}(W(c'')) \\ K(c') \equiv
 K(c'')}$}
-\AxiomC{$\mathit{Canonizer}(c') = \mathit{Canonizer}(c'')$}
-\QuaternaryInfC{$q_2 \circ q_1: c \Leftrightarrow c'''$}
+\AxiomC{$\canonizer(c') = \canonizer(c'')$}
+\TrinaryInfC{$q_2 \circ q_1: c \Leftrightarrow c'''$}
 \end{prooftree}
@@
   \end{align*}
-
-This derivation rule says that the composition $q_2 \circ q_1$ is well defined
-if and only if the intermediary QREs $c''$ and $c'''$ define the same
-equivalence relation on the same regular language. The condition
-$\mathcal{L}(W(c')) = \mathcal{L}(W(c''))$ says that the intermediary language
-is the same on both sides, while the condition $K(c') \equiv^d K(c'')$ says that
-the kernel regular expressions are equivalent via the star-semiring axioms.
-Finally, the condition $\canonizer(c') = \canonizer(c'')$ says
-that $c'$ and $c''$ define the same equivalence relation on $\mathcal{L}(W(c'))=
-\mathcal{L}(W(c''))$.
 
@@
 $\canonizer(c')$ is the identity function on $K(c'')$ which is
-equivalent to $K(c')$ via the star semiring axioms by assumption.
-
-A similar argument shows that 
+syntactically equal to $K(c')$ by assumption. A similar argument shows that 
 $$\llbracket q_1 \rrbracket.put \circ \llbracket q_2 \rrbracket.put =
@@
 $\{(x_1, y_1), \ldots, (x_n, y_n)\}$ where the $x_i$'s are in $W(c)$ and the
-$y_i$'s are in $W(c')$.
-
-We have shown in Theorem~\ref{normal form} that if there is a derivation $q : c
-\Leftrightarrow c'$ of a QRE lens, then there exists a bijective lens $\ell :
+$y_i$'s are in $W(c')$. We wish $q$ to map the equivalence class of $x_i$ to the
+equivalence class of $y_i$ and vice versa:
+\begin{align*}
+q.get(x_i) \equiv_{c'} y_i \text{, and}\\
+q.put(y_i) \equiv_c x_i
+\end{align*}
+
+Our approach in synthesizing QRE lenses is guided by Theorem~\ref{normal
+form}, which says that if there is a derivation $q :
+c \Leftrightarrow c'$ of a QRE lens, then there exists a bijective lens $\ell :
 K(c) \Leftrightarrow K(c')$ such that
@@
 
-Now since the $x_i$'s are in $W(c)$ and the $y_i$'s are in $W(c')$, then
-
-\noindent ${x_i}' = \canonizer(c)(x_i)$ is in $K(c)$ and ${y_i}' =
-\canonizer(c')(y_i)$ is in $K(c')$, hence in order to synthesize the desired
-quotient lens $q$ it suffices to synthesize a bijective lens $l : K(c)
-\Leftrightarrow K(c')$ that is consistent with the canonized examples
+For our synthesis task, since the $x_i$'s are in $W(c)$ and the $y_i$'s are
+in $W(c')$, then ${x_i}' = \canonizer(c)(x_i)$ is in $K(c)$ and ${y_i}' =
+\canonizer(c')(y_i)$ is in $K(c')$. Consequently in order to synthesize the
+desired quotient lens $q: c \Leftrightarrow c'$ that is consistent with the
+input-output examples $\{(x_1, y_1), \ldots, (x_n, y_n)\}$ it suffices to
+synthesize a bijective lens $l :
+K(c) \Leftrightarrow K(c')$ that is consistent with the canonized examples
 $\{({x_1}', {y_1}'), \ldots, ({x_n}', {y_n}')\}$
@@
 type $c \Leftrightarrow c'$ that is consistent with $\{({x_1}', {y_1}'),
-\ldots, ({x_n}', {y_n}')\}$.
-
-More specifically, in order to synthesize a quotient lens $q: c \Leftrightarrow
-c'$, that is consistent with the examples $\{(x_1, y_1), \ldots, (x_n, y_n)\}$
- it suffices to perform the following procedure:
+\ldots, ({x_n}', {y_n}')\}$. More concretely, in order to synthesize a
+quotient lens $q: c \Leftrightarrow c'$, that is consistent with the examples
+$\{(x_1, y_1), \ldots, (x_n, y_n)\}$ it suffices to perform the following
+procedure:
 \begin{enumerate}

### 2017-11-10 papers/BijectiveLenses/ArtifactEvaluationPaper/main.tex

\newif\ifdraft\drafttrue  % set true to show comments
%\newif\ifdraft\draftfalse  % set true to show comments
\newif\ifanon\anonfalse    % set true to suppress names, etc.
\newif\iffull\fullfalse   % set true for long version
\newif\ifappendices\appendicesfalse

\PassOptionsToPackage{usenames,dvipsnames,svgnames,table}{xcolor}
\documentclass[sigplan,acmsmall]{acmart}

\listfiles

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{bussproofs}
\usepackage{amsthm}
\usepackage{csvsimple}
\usepackage{thmtools,thm-restate}
\usepackage{changepage}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage[inline]{enumitem}
\usepackage{multirow,bigdelim}
\usepackage{multicol}
\usepackage{siunitx}
\usepackage{listings}
\usepackage{sansmath}
\usepackage{url}
\usepackage{flushend}
\usepackage{microtype}
\usepackage[utf8]{inputenc}
\usepackage{mathpartir}
\usepackage{empheq}
\usepackage{array}
\usepackage{pgfplots}
\usepackage{stmaryrd}
\usepackage{courier}
\usepackage{qtree}
\usepackage[normalem]{ulem}
\usepackage{relsize}
\usepackage{tikz}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{tabularx}
\usepackage{stackengine}
\usepackage{caption}
\usepackage{wrapfig}
\usepackage{remreset}

\settopmatter{printacmref=false}
\renewcommand\footnotetextcopyrightpermission[1]{}

\usetikzlibrary{
  er,
  matrix,
  shapes,
  arrows,
  positioning,
  fit,
  calc,
  pgfplots.groupplots,
  arrows.meta
}
\tikzset{>={Latex}}

%%%% Hyperlinks – must come late!
%\usepackage[pdftex,%
%            pdfpagelabels,%
%            linkcolor=blue,%
%            citecolor=blue,%
%            filecolor=blue,%
%            urlcolor=blue]
%           {hyperref}

\clubpenalty = 10000
\widowpenalty = 10000
\displaywidowpenalty = 10000

%\setlength{\belowcaptionskip}{-5pt}
%\setlength{\textfloatsep}{15pt}

% Creates a display mode for code in sans serif font
\lstnewenvironment{sflisting}[1][]
  {\lstset{%
    mathescape,
    basicstyle=\small\sffamily,
    aboveskip=5pt,
    belowskip=5pt,
    columns=flexible,
    frame=,
    xleftmargin=1em,#1}\sansmath}
  {}
% end

% Macros

  \acmYear{2018}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}
%\toappear{}

%\conferenceinfo{POPL '16}{January 20--22, 2016, St. Petersburg, FL, USA} 
%\copyrightyear{2016} 
%\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}

% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

%\titlebanner{DRAFT---do not distribute}        % These are ignored unless
%\preprintfooter{DRAFT---do not distribute}   % 'preprint' option specified.

\title{Synthesizing Bijective Lenses: Artifact Evaluation}

\author{Anders Miltner}
\affiliation{Princeton University, USA}
\email{amiltner@cs.princeton.edu}

\author{Kathleen Fisher}
\affiliation{Tufts University, USA}
\email{kfisher@eecs.tufts.edu}

\author{Benjamin Pierce}
\affiliation{University of Pennsylvania, USA}
\email{bcpierce@cis.upenn.edu}

\author{David Walker}
\affiliation{Princeton University, USA}
\email{dpw@cs.princeton.edu}

\author{Steve Zdancewic}
\affiliation{University of Pennsylvania, USA}
\email{stevez@cis.upenn.edu}
\maketitle

% \category{D.3.1}
% {Programming Languages}
% {Formal Definitions and Theory}
% [Semantics]

% begin introduction
\section{Setup Environment}
\subsection{Virtual Machine Environment (recommended)}
To make validation easier, we provide the Artifact Evaluation Committee with a
virtual machine that has all components installed.  Setup for the virtual
machine environment detailed below:
\begin{enumerate}
\item Install a Virtual Machine Manager (we tested with VirtualBox)
\item Download our virtual machine image (BijectiveLensSynthesis.ova in\\
  {\color{blue}
    \url{https://drive.google.com/drive/folders/0B5OYxhNvX675cTN4ZzVEMmJ0Mlk})}
\item Load our virtual machine from your virtual machine manager
\end{enumerate}

We suggest the committe dedicate over 8GB RAM to this environment.
While we get similar results with less RAM, some of the existing tools we
compare our artifact to use up to 8GB RAM.  Less RAM may cause these tools to
perform worse than presented in our data.

The user for the virtual machine is bls, with password bls.

\subsection{Custom Environment}
If you want to install on a custom environment, certain programs and libraries
must be installed.  We provide the commands for installation in Ubuntu 16.04.3
in parenthesis next to the program or library name.

\begin{enumerate}
\item Install opam (sudo apt install opam; opam init; eval `opam config env`;
  opam update)
\item Switch opam to OCaml 4.03.0 (opam switch 4.03.0; eval `opam config env`)
\item Install necessary system packages (opam depext conf-m4.1)
\item Install OCaml's Core (opam install core)
\item Install OCaml's Menhir (opam install menhir)
\item Install OCaml's OUnit (opam install ounit)
\item Install OCaml's ppx\_deriving (opam install ppx\_deriving)

\item Install python-pip (sudo apt install python-pip)
\item Install Python's EasyProcess (pip install EasyProcess)
\item Install Python's matplotlib (pip install matplotlib)
\item Install Python's tk (sudo apt install python-tk)

\item Install DotNet runtime environment, and C\# and Visual Basic libraries and
  compilers (sudo apt install mono-complete; sudo apt install mono-vbnc)

\item Install the codebase\\
  (git clone https://github.com/Miltnoid/BijectiveLensSynthArtifactEvaluation.git)
\end{enumerate}

\section{Tool Validation}

\subsection{Simple Validation}

To validate: navigate to the directory the codebase is installed in
(/home/bls/LensSynth in the virtual machine).
Then run the following commands:
\begin{enumerate}
\item make regenerate-specs
\item make regenerate-data
\end{enumerate}

After these commands are run, the figures will be present at the following
locations: 
\begin{itemize}
\item Figure 8 will be present at \$/generated-graphs/times.eps
\item Figure 9 will be present at \$/generated-graphs/uninferred.eps
\item Figure 10 will be present at \$/generated-graphs/compositional.eps
\item Figure 11 will be present at \$/generated-graphs/examples.eps
\end{itemize}

The command ''make regenerate-specs`` will take about an hour to complete.  It
must run to completion for the data to be saved.

The full data from the runs will be aggregated at
\$/generated-data/data.csv

The command, ``make regenerate-data'' can be stopped when partially completed.
After a benchmark has had all tests run on it, that data is saved, so future
runs will not rerun tests on that benchmark.  Each benchmark takes between 0 and
3 hours to complete, with information being printed to stdout about what
test and benchmark is being run.  The total time to complete all tests takes
between 2 and 3 days.

\subsection{Directory Information}

If the artifact committee want to do more in-depth testing, we provide
information on the individual project directories.  The means to run individual
programs are described in README.txt files in the directories.

\paragraph*{\$/program}  Our tool is present in the program directory.  The
source code is provided in \$/program/src .  The directory \$/program/tests
contains a test suite, consisting of microbenchmarks and the tested benchmarks.
The directory \$/program/example-inputs contains the tested benchmarks.
The directory \$/program/generated\_data is the output location for the csv created from
program runs.  The directories \$/program/generated\_extraction\_specs and
\$/program/generated\_io\_specs are the output locations for the specifications
given to Flash Fill and FlashExtract.

\paragraph*{\$/comparisons}  When created by ``make regenerate-specs'', the
directories \$/comparisons/extraction\_specs and \$/comparisons/io\_specs
contain the specifications given to Flash Fill and FlashExtract.  The directory
\$/comparisons/prose/Transformation.Text contains the generated data for its
runs under generated\_data, and the source code for our tests in
SampleProgram.cs.  The directory
\$/comparisons/prose/Extraction.Text contains the generated data for its
runs under generated\_data, and the source code for our tests in
LearningSamples.cs.

\paragraph*{\$/generated-data}  After data is created by the individual tests,
it is then combined and placed into \$/generated-data/data.csv.

\paragraph*{\$/generated-graphs}  After data is created by the individual tests
and combined into \$/generated-data/data.csv, the data is converted into graphs
and placed into \$/generated-graphs.

\end{document}

%%% Local Variables:
%%% TeX-master: "main"
%%% End:

### 2017-11-13 papers/BijectiveLenses/ICFPSRCPaper/main.tex
add src paper

\PassOptionsToPackage{usenames,dvipsnames,svgnames,table}{xcolor}
\documentclass[a4paper,twoside]{article}

\listfiles

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{bussproofs}
\usepackage{amsthm}
\usepackage{csvsimple}
\usepackage{thmtools,thm-restate}
\usepackage{changepage}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage[inline]{enumitem}
\usepackage{multirow,bigdelim}
\usepackage{multicol}
\usepackage{siunitx}
\usepackage{listings}
\usepackage{sansmath}
\usepackage{url}
\usepackage{flushend}
\usepackage{microtype}
\usepackage[utf8]{inputenc}
\usepackage{mathpartir}
\usepackage{empheq}
\usepackage{array}
\usepackage{pgfplots}
\usepackage{stmaryrd}
\usepackage{courier}
\usepackage{qtree}
\usepackage[normalem]{ulem}
\usepackage{relsize}
\usepackage{tikz}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{tabularx}
\usepackage{stackengine}
\usepackage{caption}
\usepackage{wrapfig}
\usepackage{remreset}

\usepackage{natbib}
\bibliographystyle{abbrvnat}
\setcitestyle{numbers,square}

\usetikzlibrary{
  er,
  matrix,
  shapes,
  arrows,
  positioning,
  fit,
  calc,
  pgfplots.groupplots,
  arrows.meta
}
\tikzset{>={Latex}}

%%%% Hyperlinks – must come late!
%\usepackage[pdftex,%
%            pdfpagelabels,%
%            linkcolor=blue,%
%            citecolor=blue,%
%            filecolor=blue,%
%            urlcolor=blue]
%           {hyperref}

\input{macros}

\clubpenalty = 10000
\widowpenalty = 10000
\displaywidowpenalty = 10000

%\setlength{\belowcaptionskip}{-5pt}
%\setlength{\textfloatsep}{15pt}

% Creates a display mode for code in sans serif font
\lstnewenvironment{sflisting}[1][]
  {\lstset{%
    mathescape,
    basicstyle=\small\sffamily,
    aboveskip=5pt,
    belowskip=5pt,
    columns=flexible,
    frame=,
    xleftmargin=1em,#1}\sansmath}
  {}
% end

% Macros
\newcommand{\NameOf}[1]{\CF{#1}}

\begin{document}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\author{
  \normalsize Name: Anders Miltner\\
  \normalsize Affiliation: Princeton University\\
  \normalsize Email: amiltner@cs.princeton.edu\\
  \normalsize Address: 35 Olden Street, Princeton NJ 08540\\
  \normalsize Advisor: David Walker\\
  \normalsize Category: Graduate\\
  \normalsize ACM Student ID: 7351631
}

\title{\Large Synthesizing Bijective Lenses}

\maketitle

\paragraph*{Motivation}
Companies are increasingly processing data in ad-hoc formats for customer
analytics, for understanding error logs, and more.
Special programs are often needed to extract the
underlying data contained in these ad-hoc formats oftentimes need special
programs to extract the underlying data.  When data needs to be synchronized
accross multiple ad-hoc formats, programs need to be able to both read data from
the first format, and add that data into the second, and read data in the second
format, and add that data to the first.  Boomerang~\cite{boomerang} is an approach
to this problem:  Boomerang programs use a single expression to express
the conversions from the first source to the second and from the second source
back to the first.  Furthermore, Boomerang provides guarantees about the
round-trip behavior of the data sources.  Unfortunately, providing these
guarantees causes difficulties: Boomerang programs are typically quite difficult
to program in!  Boomerang has a complex type system which requires the user to
think about unambiguity of data formats while also thinking about the fiddly
details of data conversions.  To reduce the difficulty in programming in
Boomerang, we create a tool which synthesizes Boomerang programs, \Optician{}.
\Optician{} takes in data format
descriptors and pairs of input/output examples, and outputs a Boomerang program
mapping between those data formats which respects the examples.

\paragraph*{Background and Related Work}
Boomerang is a certain instance of a class of programming languages:
bidirectional programming languages.  A survey by Czarnecki et al. provides a (slightly dated)
summary of these languages~\cite{DBLP:conf/icmt/CzarneckiFHLST09}.  Furthermore, there have recently been
theoretical approaches to bidirectional languages in general~\cite{DBLP:journals/chinaf/FischerHP15}.

In Boomerang,
if $\Lens$ is a lens, and $\Regex$ and $\RegexAlt$ are regular expressions, then
the typing derivation $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$ means
that the lens $\Lens$ maps between the languages defined by regular expressions
$\Regex$ and $\RegexAlt$.  We focus on a subset of lenses, \emph{bijective
  lenses}.  This subset requires stricter round trip properties than normal
lenses; in particular the two functions defined by the lens must be inverses.

A large amount of work has been put into synthesizing programs by
examples~\cite{tds-pldi,le-pldi-2014,Singh:blinkfill,gulwani-popl-2014,morpheus}. 
Some of this work focuses on data transformation and processing.  However, this
previous data processing work has focused on unidirectional transformations.
Furthermore, most work in synthesizing data transformers rely only on examples
to guide synthesis.  This gives these synthesis tools the complex job of both
understanding the shape of the data as well as synthesizing the program that
transforms the data.  We, however, ask users to provide format descriptions to
our algorithm, and apply type-directed synthesis to these format descriptions.
This approach allows our algorithm to synthesize much more complex
transformations than existing tools.

\paragraph*{Approach and Uniqueness}
From two regular expressions, $\Regex$ and $\RegexAlt$ and a set of examples, we
desire to synthesize a 
lens $\Lens$ such that $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$.
Furthermore, the two functions defined by $\Lens$ should map between the two
strings in each pair of examples.
Because we have the pair of regular expressions forming the type, we preform
type-directed synthesis: synthesis which enumerates only well-typed terms.
Unfortunately, this is difficult to do in Boomerang, as Boomerang has many
equivalences on its types (as every regular expression is equivalent to an
infinite number of other regular expressions), and a composition operator.
Using a na\"ive type-directed synthesis strategy on lenses is very inefficient,
as the above aspects of the language creates a large amount of nondeterminism in
the search through the typing derivation.

To resolve these issues, we have introduced alternative languages, DNF lenses
and DNF regular expressions.  Like regular expressions, DNF regular expressions
describe languages.  Like lenses, DNF lenses encode two functions.  Furthermore,
like lenses, DNF lenses are typed by pairs of expressions.  In the DNF lens case
a derivation $\DNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt$ means
that a DNF lens provides functions between the languages of $\DNFRegex$ and $\DNFRegexAlt$.
We also prove the soundness and completeness of DNF lenses to lenses.

DNF lenses are better suited to synthesis than lenses for a number of reasons.
DNF lenses contain no composition operator, there are fewer equivalences between
the types (each DNF regular expression is equivalent to fewer other
expressions than regular expressions), and equivalences between types can only
be processed at certain areas of the typing derivation.  This reduces the
nondeterminism from the search for a well-typed lens, making the problem permit
tractable solutions.

\begin{figure}
  \centering
  \begin{tikzpicture}[auto,node distance=1.5cm]
    \node[text width=1.5cm,minimum height=.6cm,align=center,draw,rectangle] (todnfregex) {\ToDNFRegex{}};

    \node[align=right, anchor=east] (regex1) [left = .6cm of todnfregex.north west]{\Regex{}};
    \node[align=right, anchor=east] (regex2) [left = .6cm of todnfregex.south west]{\RegexAlt{}};
    \node[align=right, anchor=east] (exs) [below = .2cm of regex2]{ \Examples{} };

    \node[align=center] (dnfregex1) [right = .4cm of todnfregex.north east]{\DNFRegex{}};
    \node[align=center] (dnfregex2) [right = .4cm of todnfregex.south east]{\DNFRegexAlt{}};

    \node[text width=3.0cm,minimum height=.6cm,align=center,draw,rectangle] [right = 1.45cm of todnfregex.east] (synthdnflens) {\SynthDNFLens{}};
    \node[align=center] [above = .7cm of synthdnflens] (optician) {\Optician{}};

    \node[align=center] [right = .4cm of synthdnflens] (dnflens) {\DNFLens{}};

    \node[text width=1.5cm,minimum height=.6cm,align=center,draw,rectangle] [right = .4cm of dnflens] (tolens) {\ToLens{}};

    \node[align=center] [right = .6cm of tolens] (lens) {\Lens{}};

    \path[->] (regex1.east) edge (todnfregex.north west);
    \path[->] (regex2.east) edge (todnfregex.south west);

    \path[->] (todnfregex.north east) edge (dnfregex1.west);
    \path[->] (todnfregex.south east) edge (dnfregex2.west);

    \path[->] (dnfregex1.east) edge (synthdnflens.north west);
    \path[->] (dnfregex2.east) edge (synthdnflens.south west);

    \path[->] (synthdnflens) edge (dnflens);

    \path[->] (dnflens) edge (tolens);

    \path[->] (tolens) edge (lens);
    \draw[->] ($(exs.east)+(-3pt,0)$) -| node(exsedge) {} (synthdnflens);
    \node[fit={($(todnfregex.west)+(-4pt,0)$) ($(tolens.east)+(4pt,0)$) (exsedge) (dnfregex1) (optician) (dnfregex2)},draw] (surrounding) {};
    % Now place a relation (ID=rel1)
    %\node[text width=2cm,align=center,draw, rectangle] (sketch-gen) [right = .75cm of spec] {\TypeProp{}};
    %\node (below-gen) [below=.5cm of sketch-gen] {};
    %\node[text width=2cm,align=center,draw, rectangle] (sketch-compl)
    %     [right = .25cm of sketch-gen] {\RigidSynth{}};
    %\node (below-compl) [below=.5cm of sketch-compl] {};
    %\node[align=center] (lens) [right = .75cm of sketch-compl] {Lens}; 
    %% Draw an edge between rel1 and node1; rel1 and node2
    %\path[->] (spec) edge node (start-alg) {} (sketch-gen);
    %\path[->] (sketch-gen) edge node(middle) {} (sketch-compl);
    %\path[->] (sketch-compl) edge node[near start](success) {\Success{}} (lens);
    %\draw[<-] (sketch-gen.south) -- +(0,-.5) -| node[above left](failure){\Failure{}} (sketch-compl.south);

    %\node (synth-name) [above=.5cm of middle] {\Optician{}};
    %
    %\node[fit=(sketch-gen) (sketch-compl) (start-alg) (synth-name) (failure) (success) ,draw] (surrounding) {};
  \end{tikzpicture}
  \caption{Schematic Diagram for \Optician{}.  Regular expressions, \Regex{} and
    \RegexAlt{}, and examples, \Examples{}, are given as input.
    First, the function \ToDNFRegex{} converts \Regex{} and \RegexAlt{} into
    their respective DNF forms, \DNFRegex{} and \DNFRegexAlt{}.
    Next, \SynthDNFLens{} synthesizes a DNF lens, \DNFLens{}, from \Regex{},
    \RegexAlt{}, and \Examples{}.
    Finally, \ToLens{} converts \DNFLens{} into \Lens{}, a lens in Boomerang
    that is equivalent to \DNFLens{}.}
  \label{fig:schematic-diagram-synthesis}
\end{figure}

Figure~\ref{fig:schematic-diagram-synthesis} shows the schematic diagram for
\Optician{}.  Two regular expressions, \Regex{} and \RegexAlt{}, and a set of
examples, \Examples{}, are given as input.  The regular expressions are
converted into DNF regular expressions with $\ToDNFRegex$.  Then, type-directed
synthesis is performed on the input DNF regular expressions and the examples, in
\SynthDNFLens{}.  After \SynthDNFLens{} generates a satisfying DNF lens, it is
converted back into a lens with $\ToRegex$.  We prove the correctness of our
synthesis algorithm using DNF lenses: if there is a lens satisfying the
specification, then the synthesis algorithm will return a lens which satisfies
the specification (though not necessarily the same one).

Our work is unique in that we are the first synthesis algorithm for a
bidirectional language.  While it is common to develop domain specific languages
which permit efficient synthesis algorithms, and also it is common to develop
ways of compiling these languages to existing ones, it is less common to provide
completeness results stating exactly which elements of the existing language are
able to be synthesized.  Furthermore, \Optician{} synthesizes string processing
programs from types as well as examples.

\paragraph*{Results and Contributions}
As we are the first synthesis algorithm on bidirectional programs, we have no
direct comparisons, but by changing the specifications, we were able to compare
our tool with 2 tools which synthesize string transformers (Flash Fill~\cite{gulwani-popl-2014} and
FlashExtract~\cite{le-pldi-2014}), and found our tool 
was able to outperform them.
Our synthesis algorithm was able to synthesize all 39
of the examples in less than 5 seconds, where the other tools synthesized 3 and
4 within the time limit of 10 minutes.
Furthermore, we found that the optimizations we used were critical in making
this synthesis tractible.

Our contributions are as follows:
\begin{enumerate}
\item We developed a domain specific language custom built for synthesis.  We
  proved this language equivalent to a higher-level, declarative, bijective
  subset of Boomerang.

\item We develop a synthesis algorithm for this domain specific language, and
  leverage it for the synthesis of Boomerang programs.  We prove this algorithm
  correct.

\item We evaluate our algorithm on a set of 39 benchmarks.  We find that the
  optimizations are critical for synthesis of all benchmarks, and find that we
  outperform existing tools on the types of complex tasks that necessitate the
  use of bidirectional languages.
\end{enumerate}

\paragraph*{My Contributions}
This work was done with the collaboration of 4 other researchers: David Walker,
Kathleen Fisher, Benjamin Pierce, and Steve Zdancewic.  Their insights
were critical to the entire process.  I did all the proofs and coded
\Optician{}, however.

% \category{D.3.1}
% {Programming Languages} 
% {Formal Definitions and Theory}
% [Semantics]
%\ifanon\else
%\category{F.4.1}
%{Mathematical Logic and Formal Languages}
%{Mathematical Logic}
%[Proof Theory]
%\category{I.2.2}
%{Artificial Intelligence}
%{Automatic Pro\-gramming}
%[Program Synthesis]
%
%\terms Languages, Theory
%
%\keywords Functional Programming, Proof Search, Program Synthesis, Type Theory
%\fi

% The bibliography should be embedded for final submission.

\bibliography{local,bcp}

\end{document}

%%% Local Variables:
%%% TeX-master: "main"
%%% End:

### 2017-11-13 anonymized_material/comparisons/prose/Extraction.Text/generate-data.py
clean up data

@@ -1,107 +1,0 @@
-#!/usr/local/bin/python
-
-from __future__ import print_function
-from easyprocess import EasyProcess
-
-import os
-import csv
-from os.path import splitext, join
-import subprocess
-import sys
-import time
-
-TEST_EXT = '.tsv'
-BASELINE_EXT = '.out'
-BASE_FLAGS = []
-TIMEOUT_TIME = 600
-
-def ensure_dir(f):
-    d = os.path.dirname(f)
-    if not os.path.exists(d):
-        os.makedirs(d)
-
-def transpose(matrix):
-    return zip(*matrix)
-
-def find_tests(root):
-    tests = []
-    for path, dirs, files in os.walk(root):
-        files = [(f[0], f[1]) for f in [splitext(f) for f in files]]
-        tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
-    return tests
-
-def gather_datum(prog, path, base, additional_flags, timeout):
-    print(["mono", prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)])
-    process_output = EasyProcess(["mono", prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)]).call(timeout=timeout)
-    return (process_output.stdout,process_output.stderr)
-
-
-def gather_data(rootlength, prog, path, base):
-    current_data = {"Test":join(path, base).replace("_","-")[rootlength:]}
-    timeout = False
-    error = False
-    num_inputs = 0
-    result = "IncorrectResponse"
-    while (result == "IncorrectResponse"):
-        num_inputs=int(float(num_inputs)*1.5+1.0)
-        print(num_inputs)
-        (datum,err) = gather_datum(prog, path, base, [str(num_inputs)],TIMEOUT_TIME)
-        if datum == "":
-            if err == "":
-	            result="-1"
-            else:
-                print(err)
-                result="-1"
-        elif (num_inputs > 100000):
-            result="TooLargeSpec"
-        else:
-	        result=datum
-    print(result)
-    current_data["Result"]=result
-    current_data["ExsTaken"]=num_inputs
-    return current_data
-
-def print_data(data):
-    ensure_dir("generated_data/")
-    with open("generated_data/data.csv", "wb") as csvfile:
-	datawriter = csv.DictWriter(csvfile,fieldnames=data[0].keys())
-	datawriter.writeheader()
-	datawriter.writerows(data)
-
[…]

### 2017-11-13 anonymized_material/comparisons/prose/Transformation.Text/SampleProgram.cs

@@ -1,129 +1,0 @@
-﻿using System;
-using System.IO;
-using System.Collections.Generic;
-using System.Linq;
-using System.Threading;
-using System.Security.Permissions;
-using Microsoft.ProgramSynthesis.Transformation.Text;
-using Microsoft.ProgramSynthesis.Transformation.Text.Semantics;
-using Microsoft.ProgramSynthesis.Wrangling.Constraints;
-using VBFIO = Microsoft.VisualBasic.FileIO;
-using System.Diagnostics;
-
-namespace Transformation.Text
-{
-	/// <summary>
-	///     Sample of how to use the Transformation.Text API. Transformation.Text generates string programs from input/output examples.
-	/// </summary>
-	internal static class SampleProgram
-	{
-		private enum ResultType { Success, Exception, IncorrectResponse, NothingLearned }
-
-		private const int NUM_TESTS = 100;
-
-		private static void Main(string[] args)
-		{
-			var t = new Stopwatch();
-			if (args.Length != 2) {
-				throw new Exception(args.Length.ToString());
-			}
-			var numExamples = int.Parse(args[0]);
-			var fileName = args[1];
-			t.Start();
-			var result = GetResult(fileName, numExamples);
-			if (result == ResultType.Success)
-			{
-				t.Stop();
-				Console.Out.WriteLine(t.ElapsedMilliseconds);
-			}
-			else
-			{
-				Console.Out.WriteLine(GetResult(fileName, numExamples).ToString());
-			}
-		}
-
-		private static List<KeyValuePair<string, string>> ParseOutIOList(string fileName)
-		{
-			var ioList = new List<KeyValuePair<string, string>>();
-/*			var reader = new StreamReader(fileName);
-			var csv = new CsvReader(reader);
-			csv.Configuration.Delimiter = "\t";
-			csv.Configuration.HasHeaderRecord = false;
-			while (csv.Read())
-			{
-				ioList.Add(new KeyValuePair<string,string>(csve.GetField<string>(0),csv.GetField<string>(1)));
-				Console.Out.WriteLine("in");
-				Console.Out.WriteLine(csv.GetField<string>(0));
-				Console.Out.WriteLine("out");
-				Console.Out.WriteLine(csv.GetField<string>(1));
-			}
-			Console.Out.WriteLine(ioList.Count());
-			Console.Out.WriteLine(fileName);
-			return ioList;*/
-
[…]

### 2017-11-13 anonymized_material/comparisons/prose/Transformation.Text/generate-data.py

@@ -1,113 +1,0 @@
-#!/usr/local/bin/python
-
-from __future__ import print_function
-from easyprocess import EasyProcess
-
-import os
-import csv
-from os.path import splitext, join
-import subprocess
-import sys
-import time
-
-TEST_EXT = '.tsv'
-BASELINE_EXT = '.out'
-BASE_FLAGS = []
-TIMEOUT_TIME = 600
-
-counter = 0
-
-def ensure_dir(f):
-    d = os.path.dirname(f)
-    if not os.path.exists(d):
-        os.makedirs(d)
-
-def transpose(matrix):
-    return zip(*matrix)
-
-def find_tests(root):
-    tests = []
-    for path, dirs, files in os.walk(root):
-        files = [(f[0], f[1]) for f in [splitext(f) for f in files]]
-        tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
-    return tests
-
-def gather_datum(prog, path, base, additional_flags, timeout):
-    print(["mono", prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)])
-    process_output = EasyProcess(["mono", prog] + BASE_FLAGS + additional_flags + [join(path, base + TEST_EXT)]).call(timeout=timeout)
-    return (process_output.stdout,process_output.stderr)
-
-
-def gather_data(rootlength, prog, path, base):
-    global counter
-    current_data = {"Test":join(path, base).replace("_","-")[rootlength:]}
-    timeout = False
-    error = False
-    num_inputs = 0
-    result = "IncorrectResponse"
-    while (result == "IncorrectResponse"):
-        num_inputs=int(num_inputs*1.5+1.0)
-        print(num_inputs)
-        (datum,err) = gather_datum(prog, path, base, [str(num_inputs)],TIMEOUT_TIME)
-        if datum == "" or err != "":
-            if err == "":
-	            result="-1"
-            else:
-                print("what")
-                print(err)
-                result="-1"
-        elif (num_inputs > 100000):
-            result="-1"
-        else:
-	        result=datum
-    print(result)
-    current_data["Result"]=result
-    current_data["ExsTaken"]=num_inputs
-    counter = counter+1
-    print_data([current_data],str(counter));
-    return current_data
-
-def print_data(data,file_addition):
-    ensure_dir("generated_data/")
-    with open("generated_data/data" + file_addition + ".csv", "wb") as csvfile:
[…]

### 2017-11-13 anonymized_material/program/check-with-boomerang.py

@@ -1,88 +1,0 @@
-#!/usr/local/bin/python
-
-from __future__ import print_function
-from easyprocess import EasyProcess
-
-import os
-import csv
-from os.path import splitext, join
-import subprocess
-import sys
-import time
-
-BOOMERANG_COMMAND = "boomerang"
-TEMP_BOOM_FILENAME = "generated.boom"
-TEST_EXT = '.ls'
-BASELINE_EXT = '.out'
-BASE_FLAGS = []
-TIMEOUT = 60
-
-REPETITION_COUNT = 2
-
-def find_tests(root):
-    tests = []
-    for path, dirs, files in os.walk(root,followlinks=True):
-        files = [(f[0], f[1]) for f in [splitext(f) for f in files]]
-        tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
-    return tests
-
-def gather_datum(prog, path, base):
-    process_output = EasyProcess([prog] + BASE_FLAGS + [join(path, base + TEST_EXT)]).call(timeout=TIMEOUT)
-    return (process_output.stdout,process_output.stderr,process_output.return_code)
-
-def run_boomerang():
-    process_output = EasyProcess([BOOMERANG_COMMAND,TEMP_BOOM_FILENAME]).call(timeout=TIMEOUT)
-    return (process_output.stdout,process_output.stderr,process_output.return_code)
-
-def check_files(prog, path, base):
-    test_name = join(path, base + TEST_EXT)
-    (datum,err,return_code) = gather_datum(prog, path, base)
-    if return_code != 0:
-        raise Exception(test_name + " failed with stdout:" + datum + "and stderr:" + err)
-    elif datum == "":
-        if err == "":
-	    raise Exception("synthesis of " + test_name + " timed out")
-        else:
-            raise Exception("synthesis of " + test_name +
-                            " errored with error:" + err)
-    with open(TEMP_BOOM_FILENAME, 'w') as f:
-        f.write(datum)
-    (datum,err,return_code) = run_boomerang()
-    if return_code != 0:
-        raise Exception("boomerang failed on the generated files of " + test_name +
-                        "with stdout:" + datum + "stderr:" + err)
-
-def raise_usage(args):
-    raise Exception("Usage: {0} <prog> <test|testdir>".format(args[0]))
-
-def main(args):
-    if len(args) == 3:
-        prog = args[1]
-        path = args[2]
-        data = []
-        if not os.path.exists(prog):
[…]

### 2017-11-13 anonymized_material/program/generate-baseline.py

@@ -1,69 +1,0 @@
-#!/usr/bin/python
-
-from __future__ import print_function
-
-import os
-from os.path import splitext, join
-import subprocess
-import sys
-import time
-
-TEST_EXT = '.ls'
-BASELINE_EXT = '.out'
-FLAGS = []
-
-def find_tests(root):
-    tests = []
-    for path, dirs, files in os.walk(root):
-        files = [(f[0], f[1]) for f in [splitext(f) for f in files]]
-        tests.extend([(path, f[0]) for f in files if f[1] == TEST_EXT])
-    return tests
-
-def run_test(prog, path, base):
-    return subprocess.check_output([prog] + FLAGS + [join(path, base + TEST_EXT)], stderr=subprocess.STDOUT)
-
-def gen_timestamp():
-    return "(* Generated {0} *)".format(time.strftime("%c"))
-
-def gen_baseline(prog, path, base):
-    test          = base + TEST_EXT
-    test_path     = join(path, test)
-    baseline_path = join(path, base + BASELINE_EXT)
-    print("Generating baseline for {0}...".format(test_path))
-    try:
-        output_string = run_test(prog, path, base)
-    except subprocess.CalledProcessError as err:
-        output_string = "An error was raised by this test: {0}\n{1}".format(err.returncode, err.output)
-        print("Encountered an error running test {0}".format(test_path))
-	print(output_string)
-    with open(baseline_path, 'w') as f:
-        f.write(gen_timestamp() + '\n')
-        f.write(output_string)
-
-def gen_baselines(prog, root):
-    for path, base in find_tests(root):
-        gen_baseline(prog, path, base)
-
-def print_usage(args):
-    print("Usage: {0} <prog> <test|testdir>".format(args[0]))
-
-def main(args):
-    if len(args) == 3:
-        prog = args[1]
-        path = args[2]
-        if not os.path.exists(prog):
-            print_usage(args)
-        elif os.path.exists(path) and os.path.isdir(path):
-            gen_baselines(prog, path)
-        else:
-            path, filename = os.path.split(path)
-            base, ext = splitext(filename)
-            if ext != TEST_EXT:
-                print_usage(args)
-            else:
-                gen_baseline(prog, path, base)
-    else:
-        print_usage(args)
-
-if __name__ == '__main__':
-    main(sys.argv)

### 2017-11-13 papers/QuotientLenses/main.tex
change quotientlensssynth folder to quotientlenses folder

--- 2017-11-10 papers/QuotientLensSynth/main.tex
@@
 itself in the form of a regular expression before she can define the
-equivalence relation on that data. This leads to repetitive code since the
-regular expression that defines the data and the equivalence relation defined on
-the regular expression follow the same syntactic patter---one can do both things
-with the same code with only a few changes.
-
-\sam{TODO: PUT EXAMPLE HERE!}
+equivalence relation on that data. This leads to verbose code.
 
@@
 synthesizing unidirectional string
-transformations~\cite{singh2012learning,le-pldi-2014,gulwani-popl-2014,
-perelman2014test,Singh:blinkfill}, our system Optician, which we introduced in
+transformations~\cite{singh2012learning,le-pldi-2014,gulwani-popl-2014,perelman2014test,Singh:blinkfill}, our system Optician, which we introduced in
 ~\cite{popl18}, is the first attempt to synthesize bidirectional
@@
 synthesize functions from type-based
-specifications~\cite{augustsson-2004,osera+:pldi15,
-feser-pldi-2015,scherer-icfp-2015,frankle+:popl16,armando+:pldi16}.
+specifications~\cite{augustsson-2004,osera+:pldi15,feser-pldi-2015,scherer-icfp-2015,frankle+:popl16,armando+:pldi16}.
 These systems enumerate programs of their target language, orienting their

### 2017-11-14 papers/QuotientLenses/main.tex
make build succeed on first building
added proof.sty and some comments for Solomon

@@
 \usepackage{listings}
+\usepackage{proof}
 \lstset{
@@
 \newcommand{\sam}[1]{\FINISH{dkpurple}{SM}{#1}}
+
+% For Inference Rules
+\newcommand{\Rule}[2]{\infer{#2}{#1}}
+\newcommand{\RuleSide}[3]{\infer[#3]{#2}{#1}}
+%\newcommand{\Axiom}[1]{\Rule{}{#1}}
+
+\newcommand{\wf}[1]{\vdash \ensuremath{#1\;\mathsf{wf}}}
 
@@
 \section{Introduction}
+
 {\em Lenses}~\cite{Lenses} specify bidirectional 
@@
 
+\saz{Fix the notation in the following code.}
+
 \begin{lstlisting}
@@
 $\mathcal{L}(R)$, with $\mathcal{L}(R')$ forming a complete set of
-representatives for this equivalence relation. The $\normalize{R}{R'}{f}$
+representatives for this equivalence relation. \saz{rewrite the next sentence to
+  avoid so many uses of ``express''?} The $\normalize{R}{R'}{f}$
 combinator is included so as to express equivalence relations that are
@@
 
+\saz{Add something about being ``complete'' here?  Say more about how normalize
+  wouldn't be used in ``ordinary'' QRE programming?}
 However, checking that the canonizing function $f$ is
@@
 not as expressive as Boomerang canonizers. On the other hand, QREs enable us to
-express many equivalences which occur in real world data as we shall demonstrate
+express many equivalences which\saz{``which'' $\to$ ``that''} occur in real world data as we shall demonstrate
 in the evaluation section (Section~\ref{impl}).
@@
 Before giving the formal semantics for QREs, we address the issue of {\em
-ambiguity} which comes up when defining QREs. If $c$ and  $c'$ are QREs, then
+ambiguity}, which comes up\saz{``comes up'' is a little colloquial; perhaps ``arises''?} when defining QREs. If $c$ and  $c'$ are QREs, then
 when applying the regular combinators to QREs, we require that if a string $s$
@@
 \noindent The behaviour of \cd{ambiguous} is not-well defined since the string
-\cd{"aaa"} can be canonized to any of \cd{"a", "aa", "aaa"} depending on how
-\cd{"aaa"} is parsed. We also require that the regular combinators applied to
+\cd{"aaa"} can be canonized to any of \cd{"a"}, \cd{"aa"}, or \cd{"aaa"} depending on how
+\cd{"aaa"} is parsed.
+
+\saz{Add citations to prior work that uses these definitions?}
+We also require that the regular combinators applied to
 the kernels are unambiguous since the underlying lenses will end up operating
@@
 reasons.
-
+%
 To this end, we say that regular expressions $R$ and $S$ are
@@
 \begin{figure}[p!]
+\saz{This well-formedness judgment is better, but the \textbf{WF} is a bit
+  jarring -- it detracts from the content of the judgment.  Maybe something
+  lighter, like just $\mathit{wf}$?}
 \begin{center}
@@
 The most complicated inferrence rule is the rule for the \textit{permutation}
-combinator:
+combinator: \saz{fix the overflow by stacking ?  Also, it should use the WF
+  predicate in the premise, right?}
 
@@
 
-As we mentioned before, each QRE $c$ enables us to express:
+As we mentioned before, each QRE $c$ enables us to express: \saz{I'm not sure we
+  need to repeat the whole list here.}
 \begin{enumerate}
@@
 $W(c)$ of a QRE. The \textit{squash} and \textit{permutation} combinators
-give the two interesting definitions . If $c = \squash{c}{c'}{f}$, then
+give the two interesting definitions. If $c = \squash{c}{c'}{f}$, then
 the whole language of $c$ is $W(c) \sep W(c')$. This is because the
@@
   \centering
-  \begin{prooftree}
-\AxiomC{$\ell : R \Leftrightarrow S$}
-\UnaryInfC{$\kw{lift}(\ell): id(R) \Leftrightarrow id(S)$}
-\end{prooftree}
-  \begin{align*}
-  \llbracket q \rrbracket.get &=  \llbracket \ell \rrbracket\\
-  \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1}
-  \end{align*}
+
+\[
+\begin{array}[b]{c@{\qquad}c}
+\Rule{\ell : R \Leftrightarrow S}{\kw{lift}(\ell): id(R) \Leftrightarrow id(S)}
+&
+  \begin{array}[b]{rcl}
+  \llbracket q \rrbracket.get &=&  \llbracket \ell \rrbracket\\
+  \llbracket q \rrbracket.put &=& \llbracket \ell \rrbracket^{-1}
+  \end{array}
+\end{array}
+\]
+
+% \begin{tabular}[t]{cc}
+
+% \begin{minipage}{\textwidth/2}
+%   \begin{prooftree}
+% \AxiomC{$\ell : R \Leftrightarrow S$}
+% \UnaryInfC{$\kw{lift}(\ell): id(R) \Leftrightarrow id(S)$}
+% \end{prooftree}
+% \end{minipage}
+% &
+% \(
+% \)
+% \end{tabular}
+
+\[
+\Rule{q' : c'  \Leftrightarrow c'' &
+ \wf{c} &
+ K(c) = W(c')}
+{\kw{lquot}(c, q'): c' \circ c \Leftrightarrow c''}
+\]
 
@@
 
-In our previous work, we showed that given regular expressions $R, S$ and a set
+In our previous work \saz{citation}, we showed that given regular expressions $R, S$ and a set
 of input-output pairs $\{(r_1, s_1), \ldots, (r_n, s_n)\}$, if there is a
@@
 \item Does synthesizing quotient lenses from QREs permit for an easier
-  development process than existing approaches.
+  development process than existing approaches?
 
 \item Are QREs sufficiently expressive that more lenses can be synthesized
-  without modifications to the format.
+  without modifications to the format?
 
 \item Is our synthesis algorithm competitive with existing lens synthesis
-  tools.
+  tools?
 \end{enumerate}
@@
 Optician was only able to synthesize fully bijective data transformations.
-\Name loosens that restriction, and is able to synthesize data which is
+\Name{} loosens that restriction, and is able to synthesize data which is
 bijective when put in canonical form.
@@
 
+\bibliographystyle{plain}
 \bibliography{local}

### 2017-11-14 papers/QuotientLenses/main.tex

@@
 
+\saz{Maybe say again here how \Name{} improves on Optician?   Is it the case
+  that if you use the ``regular language'' subset of QRE's that our new algorithm
+  degenerates to the Optician one?  Is it worth trying that out experimentally?}
+
 Much of the research in synthesis assumes that the synthesizer is provided with
@@
 programmer supplies both examples {\em and} format descriptions in the form of
-regular expressions. There are many other recent results showing how to
+regular expressions.  \saz{Maybe elaborate more on the tradeoffs here?}
+
+There are many other recent results showing how to
 synthesize functions from type-based
@@
 system for synthesizing terms in the simply-typed lambda calculus that addresses
-equivalences on types.
+equivalences on types.  \saz{This kind of comparison is OK, but a bit
+  superficial.  Is there a crisper, technical description of the differences?}
 

### 2017-12-15 papers/QuotientLenses/main.tex

--- 2017-11-13 papers/QuotientLenses/main.tex
@@
 \usepackage{listings}
+\usepackage{proof}
+
 \lstset{
@@
 }
-\newcommand{\cd}[1]{\lstinline[backgroundcolor=\color{white}]$#1$}
 
@@
 \newcommand{\sam}[1]{\FINISH{dkpurple}{SM}{#1}}
+
+% For Inference Rules
+\newcommand{\Rule}[2]{\infer{#2}{#1}}
+\newcommand{\RuleSide}[3]{\infer[#3]{#2}{#1}}
+%\newcommand{\Axiom}[1]{\Rule{}{#1}}
+
+\newcommand{\wf}[1]{\ensuremath{#1\;\mathsf{wf}}}
 
@@
 \newcommand{\SystemOnBenchmarks}{\textbf{SS}}
+\newcommand{\cd}[1]{\lstinline[backgroundcolor=\color{white}]$#1$}
 
@@
 However, while plain lenses are widely used in both standalone systems and
-popular libraries, Boomerang~\cite{Boomerang} is the only implementation
+popular libraries, Boomerang~\cite{boomerang} is the only implementation
 currently offering quotient lenses.  One reason for this is that there is
@@
 representatives for this equivalence relation. The $\normalize{R}{R'}{f}$
-combinator is included so as to express equivalence relations that are
-difficult to express, or that cannot be expressed, using the other
-combinators.
-
-However, checking that the canonizing function $f$ is
+combinator is included so as to express equivalence relations that are either 
+difficult to define, or that cannot be defined using the other combinators. 
+\saz{Add something about being ``complete'' here?}\sam{I'm not sure how to do
+this here. The precise statement about how normalize is strong enough to
+express all QREs requires that we introduce W(c), K(c) and Canonizer(c) which
+we haven't introduced at this point in the paper}
+
+\saz{Say more about how normalize
+wouldn't be used in ``ordinary'' QRE programming?}\sam{I'm not sure what you
+mean here exactly\ldots} However, checking that the canonizing function $f$ is
 surjective and idempotent is undecidable in general, thus we do not use the
@@
 not as expressive as Boomerang canonizers. On the other hand, QREs enable us to
-express many equivalences which occur in real world data as we shall demonstrate
+express many equivalences that occur in real world data as we shall demonstrate
 in the evaluation section (Section~\ref{impl}).
@@
 Before giving the formal semantics for QREs, we address the issue of {\em
-ambiguity} which comes up when defining QREs. If $c$ and  $c'$ are QREs, then
+ambiguity}, which arises when defining QREs. If $c$ and  $c'$ are QREs, then
 when applying the regular combinators to QREs, we require that if a string $s$
@@
 that regular expression in only one way. This unambiguity condition is called
-{\em strong unambiguity}, and is necessary, firstly because it ensures that the
-canonizing function of a QRE is well-defined.
+{\em strong unambiguity}~\cite{Sippu1988}, and is necessary, firstly because it
+ensures that the canonizing function of a QRE is well-defined.
 
@@
 \noindent The behaviour of \cd{ambiguous} is not-well defined since the string
-\cd{"aaa"} can be canonized to any of \cd{"a", "aa", "aaa"} depending on how
-\cd{"aaa"} is parsed. We also require that the regular combinators applied to
+\cd{"aaa"} can be canonized to any of \cd{"a"}, \cd{"aa"}, or \cd{"aaa"} depending on how
+\cd{"aaa"} is parsed.
+
+We also require that the regular combinators applied to
 the kernels are unambiguous since the underlying lenses will end up operating
@@
 \AxiomC{$R$ is strongly unambiguous}
-\UnaryInfC{$\mathit{id}(R)$ \textbf{WF}}
+\UnaryInfC{$\wf{\mathit{id}(R)}$}
 \DisplayProof
@@
 \AxiomC{$s \in \mathcal{L}(R)$}
-\UnaryInfC{$\project{R}{s}$ \textbf{WF}}
+\UnaryInfC{$\wf{\project{R}{s}}$}
 \DisplayProof
@@
 \begin{prooftree}
-\AxiomC{$c_i, c$ are well formed}
+\AxiomC{$\wf{c_i, c}$}
 \AxiomC{${\substack{\forall \sigma \neq \theta, \; W(c_{\sigma(1)}) \cdot W(c)
@@
 \AxiomC{$K(c_1) \cdot^! K(c) \cdot^! \ldots \cdot^! K(c) \cdot^! K(c_n)$}
-\TrinaryInfC{$\perm{c_1, \ldots, c_n}{c}$ \textbf{WF}}
+\TrinaryInfC{$\wf{\perm{c_1, \ldots, c_n}{c}}$}
 \end{prooftree}
@@
 \begin{center}
-\AxiomC{$c$ \textbf{WF}}
-\AxiomC{$c'$ \textbf{WF}}
+\AxiomC{$\wf{c, c'}$}
 \AxiomC{$\mathcal{L}(W(c)) \cap \mathcal{L}(W(c')) = \varnothing$}
 \AxiomC{$f : \mathcal{L}(W(c)) \longrightarrow \mathcal{L}(W(c'))$}
-\QuaternaryInfC{$\squash{c}{c'}{f}$ \textbf{WF}}
+\TrinaryInfC{$\wf{\squash{c}{c'}{f}}$}
 \DisplayProof
@@
 \AxiomC{$f = f^2$}
-\QuaternaryInfC{$\normalize{R}{R'}{f}$ \textbf{WF}}
+\QuaternaryInfC{$\wf{\normalize{R}{R'}{f}}$}
 \end{prooftree}
@@
 \begin{center}
-\AxiomC{$c$ \textbf{WF}}
-\AxiomC{$c'$ \textbf{WF}}
+\AxiomC{$\wf{c, c'}$}
 \AxiomC{$K(c) = W(c')$}
-\TrinaryInfC{$c' \; ; \; c$ \textbf{WF}}
+\BinaryInfC{$\wf{c' \; ; \; c}$}
 \DisplayProof
 \hskip 1.5em
-\AxiomC{$c$ \textbf{WF}}
+\AxiomC{$\wf{c}$}
 \AxiomC{$W(c)^{*!}$}
 \AxiomC{$K(c)^{*!}$}
-\TrinaryInfC{$c^*$ \textbf{WF}}
+\TrinaryInfC{$\wf{c^*}$}
 \DisplayProof
@@
 \begin{center}
-\AxiomC{$c$ \textbf{WF}}
-\AxiomC{$c'$ \textbf{WF}}
+\AxiomC{$\wf{c, c'}$}
 \AxiomC{$W(c) \cdot^! W(c')$}
 \AxiomC{$K(c) \cdot^! K(c')$}
-\QuaternaryInfC{$c \cdot c'$ \textbf{WF}}
+\TrinaryInfC{$\wf{c \cdot c'}$}
 \DisplayProof
@@
 \begin{center}
-\AxiomC{$c$ \textbf{WF}}
-\AxiomC{$c'$ \textbf{WF}}
+\AxiomC{$\wf{c, c'}$}
 \AxiomC{$W(c) \cap W(c') = \varnothing$}
-\TrinaryInfC{$c \cdot c$ \textbf{WF}}
+\BinaryInfC{$\wf{c \cdot c}$}
 \DisplayProof
 \hskip 1.5em
-\AxiomC{$c$ \textbf{WF}}
+\AxiomC{$\wf{c}$}
 \AxiomC{$W(c)^{*!}$}
 \AxiomC{$K(c)^{*!}$}
-\TrinaryInfC{$c^*$ \textbf{WF}}
+\TrinaryInfC{$\wf{c^*}$}
 \DisplayProof
@@
 \begin{prooftree}
-\AxiomC{$c$ \textbf{WF}}
-\AxiomC{$c'$ \textbf{WF}}
+\AxiomC{$\wf{c, c'}$}
 \AxiomC{$W(c) \cdot^! W(c')$}
 \AxiomC{$K(c) \cdot^! K(c')$}
-\QuaternaryInfC{$c \cdot c'$ \textbf{WF}}
+\TrinaryInfC{$\wf{c \cdot c'}$}
 \end{prooftree}
@@
 \subsection{Semantics of QREs}
-
-As we mentioned before, each QRE $c$ enables us to express:
-\begin{enumerate}
-  \item a regular expression $W(c)$ (the ``whole'' of $c$),
-  \item an equivalence relation $\eqrel{c}$ on $\mathcal{L}(W(c))$,
-  \item a regular expression $K(c)$ (the ``kernel'' of $c$)
-  such that $\mathcal{L}(K(c))$ forms a complete set of representatives for
-  $\eqrel{c}$, and
-  \item a ``canonizing'' function $\canonizer(c):\mathcal{L}(W(c))
-  \longrightarrow \mathcal{L}(K(c))$ which given any $w \in \mathcal{L}(W(c))$,
-  computes $\canonizer(c)(w)$ as the unique $k$ in $\mathcal{L}(K(c))$ such that
-  $k$ is equivalent to $w$ mod $\eqrel{c}$
-  \end{enumerate}
-Figures~\ref{fig:wk} gives the inductive definitions for the whole language
-$W(c)$ of a QRE. The \textit{squash} and \textit{permutation} combinators
-give the two interesting definitions . If $c = \squash{c}{c'}{f}$, then
-the whole language of $c$ is $W(c) \sep W(c')$. This is because the
-\textit{squash} combinator takes the whole language $W(c)$ of $c$, the whole
-language $W(c')$ of $c'$ and a function $f : \mathcal{L}(W(c)) \longrightarrow
-\mathcal{L}(W(c'))$, maps $\mathcal{L}(W(c))$ into $\mathcal{L}(W(c'))$ using
-$f$ and then canonizes $W(c')$ into $K(c')$ using $\canonizer(c')$.
+As we mentioned before, each QRE $c$ enables us to express the whole and kernel
+languages $W(c)$ and $K(c)$, an equivalence relation $\eqrel{c}$ on
+$\mathcal{L}(W(c))$ that has $\mathcal{L}(K(c))$ as a complete set of
+representatives and a canonizing function $\canonizer(c):\mathcal{L}(W(c))
+  \longrightarrow \mathcal{L}(K(c))$ whose fibres are the same as $\eqrel{c}$.
+
+Figures~\ref{fig:wk} gives the inductive definitions for the whole and
+kernel languages $W(c)$ and $K(c)$ of a QRE. The \textit{squash} and
+\textit{permutation} combinators give the two interesting definitions. If $c =
+\squash{c}{c'}{f}$, then the whole language of $c$ is $W(c) \sep W(c')$. This
+is because the \textit{squash} combinator takes the whole language $W(c)$ of
+$c$, the whole language $W(c')$ of $c'$ and a function $f : \mathcal{L}(W(c))
+\longrightarrow \mathcal{L}(W(c'))$, maps $\mathcal{L}(W(c))$ into
+$\mathcal{L}(W(c'))$ using $f$ and then canonizes $W(c')$ into $K(c')$ using
+$\canonizer(c')$.
 
@@
 \]
-  \caption{Whole and Kernel Regular Expressions\bcp{try putting the 
-      last two clauses into the same format as the
-      others (using $\mathit{qq} = ...$ and maybe reducing spacing around
-      cdots, etc.)}}
+  \caption{Whole and Kernel Regular Expressions}
   \label{fig:wk}
@@
   \centering
-  \begin{prooftree}
-\AxiomC{$\ell : R \Leftrightarrow S$}
-\UnaryInfC{$\kw{lift}(\ell): id(R) \Leftrightarrow id(S)$}
-\end{prooftree}
-  \begin{align*}
-  \llbracket q \rrbracket.get &=  \llbracket \ell \rrbracket\\
-  \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1}
-  \end{align*}
-
-\begin{prooftree}
-\AxiomC{$q' : c'  \Leftrightarrow c''$}
-\AxiomC{$c$ \textbf{WF}}
-\AxiomC{$K(c) = W(c')$}
-\TrinaryInfC{$\kw{lquot}(c, q'): c' \circ c \Leftrightarrow c''$}
-\end{prooftree}
-  \begin{align*}
-  \llbracket q \rrbracket.get  &= \llbracket q'
+\[
+\begin{array}[b]{l@{\qquad}l}
+\Rule{\ell : R \Leftrightarrow S}{\kw{lift}(\ell): id(R) \Leftrightarrow id(S)}
+&
+  \begin{array}[b]{rcl}
+  \llbracket \kw{lift}(\ell) \rrbracket.get &=&  \llbracket \ell \rrbracket\\
+  \llbracket \kw{lift}(\ell) \rrbracket.put &=& \llbracket \ell \rrbracket^{-1}
+  \end{array}
+  \end{array}
+  \]
+  \[
+\begin{array}[b]{c@{\qquad}l}
+  \Rule{q : c'  \Leftrightarrow c'' &
+ \wf{c} &
+ K(c) = W(c')}
+{\kw{lquot}(c, q): c' \circ c \Leftrightarrow c''} &
+  \begin{array}[b]{rcl}
+  \llbracket \kw{lquot}(c, q) \rrbracket.get  &=& \llbracket q
   \rrbracket.get \circ \canonizer(c)\\
-  \llbracket q \rrbracket.put &= \llbracket q' \rrbracket.put
-  \end{align*}
-
-  \begin{prooftree}
-  \AxiomC{$q' : c \Leftrightarrow c'$}
-  \AxiomC{$c''$ \textbf{WF}}
-
-\AxiomC{$K(c'') = W(c')$}
-\TrinaryInfC{$\kw{rquot}(q', c):c \Leftrightarrow c'' \circ c'$}
-\end{prooftree}
-  \begin{align*}
-  \llbracket q \rrbracket.get &= \llbracket q'
+  \llbracket \kw{lquot}(c, q) \rrbracket.put &=& \llbracket q \rrbracket.put
+  \end{array}
+\end{array}
+\]
+
+\[
+\begin{array}[b]{c@{\qquad}c}
+\Rule{q : c'  \Leftrightarrow c'' &
+ \wf{c} &
+ K(c) = W(c')}
+{\kw{lquot}(c, q): c' \circ c \Leftrightarrow c''} &
+  \begin{array}[b]{rcl}
+  \llbracket \kw{lquot}(c, q) \rrbracket.get  &=& \llbracket q
+  \rrbracket.get \circ \canonizer(c)\\
+  \llbracket \kw{lquot}(c, q) \rrbracket.put &=& \llbracket q \rrbracket.put
+  \end{array}
+\end{array}
+\]
+
+\[
+\begin{array}[b]{c@{\qquad}c}
+\Rule{q : c \Leftrightarrow c'' &
+ \wf{c''} &
+ W(c'') = K(c')}
+{\kw{rquot}(q, c):c \Leftrightarrow c'' \circ c'} &
+  \begin{array}[b]{rcl}
+  \llbracket \kw{rquot}(q, c) \rrbracket.get  &=& \llbracket q
   \rrbracket.get\\
-  \llbracket q \rrbracket.put &= \llbracket q'
+  \llbracket \kw{rquot}(q, c) \rrbracket.put &=& \llbracket q
   \rrbracket.put \circ \canonizer(c'')
-  \end{align*}
+  \end{array}
+\end{array}
+\]
+
+\[
+\begin{array}[b]{l@{\qquad}c}
+\Rule{q : c \Leftrightarrow c' &
+ W(c)^{*!},W(c')^{*!} & K(c)^{*!}, K(c')^{*!}}
+{{q}^* : c^* \Leftrightarrow {c'}^*} &
+  \begin{array}[b]{rcl}
+  \llbracket {q}^* \rrbracket.get  &=& (\llbracket q \rrbracket.get)^*\\
+  \llbracket {q}^* \rrbracket.put &=& (\llbracket q \rrbracket.get)^*
+  \end{array}
+\end{array}
+\]
+
+\[
+\begin{array}[b]{l@{\qquad}c}
+\Rule{\substack{q_1 : c_1 \Leftrightarrow d_1 \\ q_2 : c_2 \Leftrightarrow
+d_2} & {\substack{W(c_1) \cdot^! W(c_2)\\ K(c_1) \cdot^! K(c_2)}} 
+& {\substack{W(d_1) \cdot^! W(d_2)\\ K(d_1) \cdot^! K(d_2)}}}
+{q_1 \cdot q_2: c_1 \cdot c_2 \Leftrightarrow d_1 \cdot d_2} &
+  \begin{array}[b]{rcl}
+  \llbracket q_1 \cdot q_2 \rrbracket.get  &=& \llbracket q_1 \rrbracket.get
+  \cdot \llbracket q_2 \rrbracket.get\\
+  \llbracket q_1 \cdot q_2\rrbracket.put &=& \llbracket q_1 \rrbracket.put
+  \cdot \llbracket q_2 \rrbracket.put
+  \end{array}
+\end{array}
+\]
+
+\[
+\begin{array}[b]{l@{\qquad}c}
+\Rule{\substack{q_1 : c_1 \Leftrightarrow d_1 \\ q_2 : c_2 \Leftrightarrow
+d_2} & \substack{\mathcal{L}(W(c_1)) \cap \mathcal{L}(W(c_2)) = \varnothing
+\\ \mathcal{L}(W(d_1)) \cap \mathcal{L}(W(d_2)) = \varnothing }}
+{q_1 \sep q_2: (c_1 \sep c_2)\Leftrightarrow (d_1 \sep d_2)} &
+  \begin{array}[b]{rcl}
+  \llbracket q_1 \sep q_2 \rrbracket.get(s) = 
+  \begin{cases}
+  \llbracket q_1 \rrbracket.get (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
+  \llbracket q_2 \rrbracket.get (s) & \text{if } s \in \mathcal{L}(W(c_2))\\
+  \end{cases}\\
+  \llbracket q_1 \sep q_2 \rrbracket.put(s) = 
+  \begin{cases}
+  \llbracket q_1 \rrbracket.put (s) & \text{if } s \in \mathcal{L}(W(d_1))\\
+  \llbracket q_2 \rrbracket.put (s) & \text{if } s \in \mathcal{L}(W(d_2))\\
+  \end{cases}
+  \end{array}
+\end{array}
+\]
 
@@
 
-    \begin{prooftree}
-\AxiomC{$q : c \Leftrightarrow c$}
-\AxiomC{$\substack{W(c)^{*!} \text{ and } W(c')^{*!} \\ K(c)^{*!} \text{ and }
-K(c')^{*!}}$}
-\BinaryInfC{${q}^* : c^* \Leftrightarrow {c'}^*$}
-\end{prooftree}
-  \begin{align*}
-  \llbracket q \rrbracket.get &= (\llbracket q \rrbracket.get)^* \\
-  \llbracket q \rrbracket.put &= (\llbracket q \rrbracket.put)^*
-  \end{align*}
-
-    \begin{prooftree}
-\AxiomC{$\substack{q_1 : c_1 \Leftrightarrow d_1 \\ q_2 : c_2 \Leftrightarrow
-d_2}$}
-\AxiomC{${\substack{W(c_1) \cdot^! W(c_2)\\ K(c_1) \cdot^! K(c_2)}}$}
-\AxiomC{${\substack{W(d_1) \cdot^! W(d_2)\\ K(d_1) \cdot^! K(d_2)}}$}
-\TrinaryInfC{$q_1 \cdot q_2: c_1 \cdot c_2
-\Leftrightarrow d_1 \cdot d_2$}
-\end{prooftree}
-  \begin{align*}
-  \llbracket q \rrbracket.get &= \llbracket q_1 \rrbracket.get \cdot \llbracket
-  q_2 \rrbracket.get \\
-  \llbracket q \rrbracket.put &= \llbracket q_1 \rrbracket.put \cdot \llbracket
-  q_2 \rrbracket.put
-  \end{align*}
-
-      \begin{prooftree}
-\AxiomC{$\substack{q_1 : c_1 \Leftrightarrow d_1 \\ q_2 : c_2 \Leftrightarrow
-d_2}$}
-\AxiomC{$\substack{\mathcal{L}(W(c_1)) \cap \mathcal{L}(W(c_2)) = \varnothing
-\\ \mathcal{L}(W(d_1)) \cap \mathcal{L}(W(d_2)) = \varnothing}$}
-\BinaryInfC{$q_1 \sep q_2: (c_1 \sep c_2)
-\Leftrightarrow (d_1 \sep d_2)$}
-\end{prooftree}
-  $$
-  \llbracket q_1 \sep q_2 \rrbracket.get(s) = 
-  \begin{cases}
-  \llbracket q_1 \rrbracket.get (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
-  \llbracket q_2 \rrbracket.get (s) & \text{if } s \in \mathcal{L}(W(c_2))\\
-  \end{cases}$$
-  $$\llbracket q_1 \sep q_2 \rrbracket.put(s) = 
-  \begin{cases}
-  \llbracket q_1 \rrbracket.put (s) & \text{if } s \in \mathcal{L}(W(d_1))\\
-  \llbracket q_2 \rrbracket.put (s) & \text{if } s \in \mathcal{L}(W(d_2))\\
-  \end{cases}
-  $$
-  \caption{Denotation and Typing Rules for QRE Lenses\bcp{I think we can
-      display this more compactly if we make the inference rules narrower by
-    stacking the premises vertically and then put each inference rule {\em
-      beside} its semantics definition.}}
+  \caption{Denotation and Typing Rules for QRE Lenses}
   \label{fig:qlenssemantics}
@@
 
-In our previous work, we showed that given regular expressions $R, S$ and a set
-of input-output pairs $\{(r_1, s_1), \ldots, (r_n, s_n)\}$, if there is a
-derivation $\ell : R \Leftrightarrow S$ of a bijective lens that maps $r_i$ to
-$s_i$ in the forward direction and $s_i$ to $r_i$ in the backward direction,
-then there is an algorithm that computes a bijective lens $\ell' : R'
-\Leftrightarrow S'$ such that $R \equiv R'$, $S \equiv S'$, $\ell'$ maps $r_i$
-to $s_i$ in the forward direction and $s_i$ to $r_i$ in the backward direction,
-and $\llbracket \ell \rrbracket = \llbracket \ell' \rrbracket$~\cite{popl18}.
+In our previous work~\cite{popl18}, we showed that given regular expressions
+$R, S$ and a set of input-output pairs $\{(r_1, s_1), \ldots, (r_n, s_n)\}$, if
+there is a derivation $\ell : R \Leftrightarrow S$ of a bijective lens that
+maps $r_i$ to $s_i$ in the forward direction and $s_i$ to $r_i$ in the backward
+direction, then there is an algorithm that computes a bijective lens $\ell' :
+R' \Leftrightarrow S'$ such that $R \equiv R'$, $S \equiv S'$, $\ell'$ maps
+$r_i$ to $s_i$ in the forward direction and $s_i$ to $r_i$ in the backward
+direction, and $\llbracket \ell \rrbracket = \llbracket \ell'
+\rrbracket$.
 
@@
 
+\iffalse
 Bidirectional transformations have been used in diverse areas of computing
@@
 program that folds view updates back into the source safely and correctly.
-
+\fi
 This paper builds on the work of Foster et al~\cite{quotientlenses} who
@@
 languages that support quotienting of data including XSugar~\cite{xsugar},
-biXid~\cite{bixid} and X/Inv~\cite{xinv}. XSugar programs bidirectionally
-convert data stored in XML and ASCII formats respectively, with the
+biXid~\cite{bixid} and X/Inv~\cite{Hu2004,Mu2004,Mu2006}. XSugar programs
+bidirectionally convert data stored in XML and ASCII formats respectively, with the
 transformations specified by a pair of unambiguous grammars. The quotienting
@@
 synthesizing unidirectional string
-transformations~\cite{singh2012learning,le-pldi-2014,gulwani-popl-2014,perelman2014test,Singh:blinkfill}, our system Optician, which we introduced in
-~\cite{popl18}, is the first attempt to synthesize bidirectional
-transformations that we are aware of. We compared Optician to two of these
-unidirectional string transformers, Flash Fill~\cite{gulwani-popl-2014} and
+transformations~\cite{singh2012learning,le-pldi-2014,gulwani-popl-2014,perelman2014test,Singh:blinkfill},
+the system Optician, which we introduced in ~\cite{popl18}, is the first attempt
+to synthesize bidirectional transformations that we are aware of. In that
+publication, we compared Optician to two of these unidirectional string
+transformers, Flash Fill~\cite{gulwani-popl-2014} and
 FlashExtract~\cite{le-pldi-2014}, and found that these tools were unsuccessful
-in synthesizing the complex transformations that can be expressed in Optician.
-Furthermore, neither of these tools were able to infer transformations 
-that occurred under two iterations.
+in synthesizing the complex transformatiunons that can be expressed in Optician.
+
+While Optician improved on prior effors at synthesizing unidirectional string
+transformers and introduced synthesis for bidirectional string transformations,
+Optician was only able to synthesize fully bijective data transformations. Our
+new system \Name{} improves on Optician in that \Name{} loosens this
+restriction, and is able to synthesize data which is bijective when put in
+canonical form. \Name{} is therefore a refinement of Optician since the
+new synthesis algorithm (i.e. the quotient lens synthesis algorithm) restricted
+to ``pure'' regular expressions degenerates to the synthesis algorithm that
+we introduced in Optician. Indeed, as we mentioned in the evaluation section,
+\Name{} was able to synthesize all of Optician's benchmarks at a speed
+competitive with Optician, though there was additional overhead in calculating
+the whole language $W(c)$ and the kernel language $K(c)$ of QREs, which
+resulted in a very slight decrease in performance.
 
 Much of the research in synthesis assumes that the synthesizer is provided with
-a collection of examples. Optician differs in that it requires that the
-programmer supplies both examples {\em and} format descriptions in the form of
-regular expressions. There are many other recent results showing how to
-synthesize functions from type-based
+a collection of examples. Optician and Optometrist differ in that they requires
+that the programmer supplies both examples {\em and} format descriptions in the
+form of regular expressions or QREs.  There is a trade-off here.  On the one
+hand, a user must have some programming expertise to write regular expression
+(or QRE) specifications and it requires some work. On the other hand, such
+specifications provide a great deal of information to the synthesis system,
+which decreases the number of examples needed (often to zero), makes the system
+scale well, and allows it to handle large, complex formats.  By providing these
+format specifications, the synthesis engine does not have to both infer the
+format of the data as well as the transformations on it, obviating the need to
+infer tricky formats like those involving nested iterations. Furthermore, by
+focusing on bidirectional transformations, we limit the space of synthesized
+functions to bijective ones, reducing the search space, and the expressiveness
+of the search space.
+
+There are many other recent results showing how to synthesize functions from
+type-based
 specifications~\cite{augustsson-2004,osera+:pldi15,feser-pldi-2015,scherer-icfp-2015,frankle+:popl16,armando+:pldi16}.
@@
 system for synthesizing terms in the simply-typed lambda calculus that addresses
-equivalences on types.
+equivalences on types. Instead of trying to directly synthesize terms of the
+simply-typed lambda calculus, InSynth synthesizes a well-typed term
+in the succinct calculus, a language with types
+that are equivalent ``modulo isomorphisms of products and
+currying''~\cite{gvero-pldi-2013}. The type structure that we used in Optician
+is significantly more complex.  In particular, because Optician types do not
+have full canonical forms, we used a pseudo-canonical form, which captures part
+of the equivalence relation over types. To preserve completeness, we pushed
+some of the remaining parts of the type equivalence relation into a set of
+rewriting rules and other parts into the synthesis algorithm itself.
 

### 2018-01-31 papers/DifferenceLenses/document.tex
Added difference lenses paper

\newif\ifdraft\drafttrue  % set true to show comments
\newif\ifplentyoftime\plentyoftimefalse  % :-)

% % For double-blind review submission, w/o CCS and ACM Reference (max
% submission space)
\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
% % For double-blind review submission, w/ CCS and ACM Reference
% \documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
% % For single-blind review submission, w/o CCS and ACM Reference (max
% submission space)
% \documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
% % For single-blind review submission, w/ CCS and ACM Reference
% \documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true} % For
% final camera-ready submission, w/ required CCS and ACM Reference
% \documentclass[acmsmall]{acmart}\settopmatter{}

% % Journal information % Supplied to authors by publisher for camera-ready
% submission; % use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{ICDP} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

% % Copyright information % Supplied to authors (based on authors' rights
% management selection; % see authors.acm.org) by publisher for camera-ready
% submission; % use 'none' for review submission.
\setcopyright{none}
% \setcopyright{acmcopyright} \setcopyright{acmlicensed}
% \setcopyright{rightsretained} \ccopyrightyear{2018}           %% If different
% from \acmYear

% % Bibliography style %\bibliographystyle{ACM-Reference-Format} % Citation
% style % Note: author/year citations are required for papers published as an %
% issue of PACMPL.
% %\citestyle{acmauthoryear}   %% For author/year citations

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% % Note:
% Authors migrating a paper from PACMPL format to traditional % SIGPLAN
% proceedings format must update the '\documentclass' and % topmatter commands
% above; see 'acmart-sigplanproc-template.tex'.
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{amsmath, amssymb, amsthm, enumerate, graphicx, centernot, tikz,
array, tikz-cd, extarrows, cleveref, mathrsfs, mathtools, stmaryrd}
\usepackage{listings}

\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{conjecture}{Conjecture}
\newtheorem{question}{Question}
\newtheorem{fact}{Fact}
\newtheorem{claim}{Claim}

% Macros
% Colors
\definecolor{dkblue}{rgb}{0,0.1,0.5}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{dkred}{rgb}{0.6,0,0}
\definecolor{dkpurple}{rgb}{0.7,0,0.4}
\definecolor{olive}{rgb}{0.4, 0.4, 0.0}
\definecolor{teal}{rgb}{0.0,0.5,0.5}
\definecolor{orange}{rgb}{0.9,0.6,0.2}
\definecolor{lightyellow}{RGB}{255, 255, 179}
\definecolor{lightgreen}{RGB}{170, 255, 220}
\definecolor{teal}{RGB}{141,211,199}
\definecolor{darkbrown}{RGB}{121,37,0}

\newcommand{\FINISH}[3]{\ifdraft\textcolor{#1}{[#2: #3]}\fi}
\newcommand{\bcp}[1]{\FINISH{dkred}{B}{#1}}
\newcommand{\BCP}[1]{\FINISH{dkred}{B}{\bf #1}}
\newcommand{\afm}[1]{\FINISH{dkgreen}{A}{#1}}
\newcommand{\dpw}[1]{\FINISH{dkblue}{D}{#1}} % Toronto Maple Leafs Blue :-)
\newcommand{\saz}[1]{\FINISH{orange}{SZ}{#1}}
\newcommand{\ksf}[1]{\FINISH{teal}{K}{#1}}
\newcommand{\sam}[1]{\FINISH{dkpurple}{SM}{#1}}

\newcommand{\kw}[1]{\ensuremath{\mathsf{#1}}}
\newcommand{\get}{\ensuremath{\kw{get}}}
\newcommand{\pput}{\ensuremath{\kw{put}}}
\newcommand{\create}{\ensuremath{\kw{create}}}
\newcommand{\ccopy}{\ensuremath{\kw{copy}}}
\newcommand{\const}{\ensuremath{\kw{const}}}
\newcommand{\default}{\ensuremath{\kw{default}}}

\lstset{ basicstyle=\ttfamily, escapeinside=|| }
\begin{document}

% % Title information
\title{Lenses With a Difference}         %% [Short Title] is optional;
% % when present, will be used in % header instead of Full Title.
% %\titlenote{with title note}             %% \titlenote is optional; % can be
% repeated if necessary; % contents suppressed with 'anonymous'
% %\subtitle{Subtitle}                     %% \subtitle is optional
% %\subtitlenote{with subtitle note}       %% \subtitlenote is optional; % can
% be repeated if necessary; % contents suppressed with 'anonymous'

% % Author information % Contents and number of authors suppressed with
% 'anonymous'.
% % Each author should be introduced by \author, followed by % \authornote
% (optional), \orcid (optional), \affiliation, and % \email.
% % An author may have multiple affiliations and/or emails; repeat the %
% appropriate command.
% % Many elements are not rendered, but should be provided for metadata %
% extraction tools.

% % Author with single affiliation.
%\author{First1 Last1}
%\authornote{with author1 note}          %% \authornote is optional;
% % can be repeated if necessary
%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional \affiliation{
%\position{Position1}
%\department{Department1}              %% \department is recommended
%\institution{Institution1}            %% \institution is required
%\streetaddress{Street1 Address1}
%\city{City1}
%\state{State1}
%\postcode{Post-Code1}
%\country{Country1}                    %% \country is recommended }
%\email{first1.last1@inst1.edu}          %% \email is recommended

\author{Solomon Maina}
\position{PhD Student}
\department{Computer Science}              %% \department is recommended
\institution{University of Pennsylvania}            %% \institution is required
\country{USA}                    %% \country is recommended }
\email{smaina@seas.upenn.edu}          %% \email is recommended

% % Abstract % Note: \begin{abstract}...\end{abstract} environment must come %
% before \maketitle command
\begin{abstract}
A {\em lens} is a program that can be run in two directions. In the
forward direction, the \get component of a lens produces a {\em view}
from a {\em source}, while in the backward direction, the \pput component folds
an updated view back into the original source. One of the reasons that
lenses are appealing is that they satisfy laws which give guarantees on how the
two transformations interact.

One of these laws requires that the \pput component if a lens returns the
original source unmodified whenever the update to the view is a no-op.
Unfortunately, this (and other) laws do not specify the behaviour of the \pput
component when the update to the view is non-trivial. Consequently, a small
update to the view can result in a large update to the source. For example, if
the source and view are order agnostic, then a lens which updates records
positionally may be valid, even though such a lens may update all the records
in the source when only a single record is added at the beginning of the view.

This paper strengthens the traditional lens laws by adding a law that
constrains the behaviour of the \pput component of a lens even when the update
made to the view is non-trivial. To do this, we will first interprete sources
and views as {\em difference posets} which are posets where we can compute the
difference between any two elements of the poset. With this interpretation, we
will then express the new law which will require that the \pput component of a
lens preserve differences. For example, we will be able to invalidate
positional lenses defined on order agnostic data on the grounds that the \pput
component of positional lenses do not respect order agnostic differences.

\sam{Also we actually need two new laws: one for put and one for create! But
create is often elided from lens papers. Do we need to talk about create in the
abstract?}

\sam{What should an evaluation section for this paper entail?}

\end{abstract}

%% Keywords
%% comma separated list
\keywords{lenses, difference posets}  %% \keywords
% are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}
TODO
\iffalse
A {\em lens} $\ell$ is a program that can be run in two directions. In the
forward direction, the \get component of $\ell$ produces a {\em view} $v$
from a {\em source} $s$ such as a database. In the backward direction,
the \pput component of $\ell$ folds an updated view $v'$ back into the
original source $s$.

For example suppose that the source is an XML document representing the
names, dates and nationalities of a collection of classical music
composers such as,

\begin{lstlisting}
<composers>
<composer>
<name>Jean Sibelius</name>
<dates>1865-1956</dates>
<nationality>Finnish</nationality>
</composer>
<composer>
<name>Aaron Copland</name>
<dates>1910-1990</dates>
<nationality>American</nationality>
</composer>
<composer>
<name>Benjamin Briten</name>
<dates>1913-1976</dates>
<nationality>English</nationality>
</composer>
</composers>
\end{lstlisting}

while the view is a list if ASCII text representing the names and dates of each
composer:
\begin{lstlisting}
Jean Sibelius, 1865-1956
Aaron Copland, 1910-1990
Benjamin Briten, 1913-1976
\end{lstlisting}

After computing the initial view, we might want to edit it in some way---e.g.,
correcting the error in Sibelius's death date and the misspelling in Britten's
name
\begin{lstlisting}
Jean Sibelius, 1865-|\colorbox{orange}{1957}|
Aaron Copland, 1910-1990
Benjamin |\colorbox{orange}{Britten}|, 1913-1976
\end{lstlisting}
and push the changes back into the original XML format:
\begin{lstlisting}
<composers>
<composer>
<name>Jean Sibelius</name>
<dates>1865-|\colorbox{orange}{1957}|</dates>
<nationality>Finnish</nationality>
</composer>
<composer>
<name>Aaron Copland</name>
<dates>1910-1990</dates>
<nationality>American</nationality>
</composer>
<composer>
<name>Benjamin |\colorbox{orange}{Britten}|</name>
<dates>1913-1976</dates>
<nationality>English</nationality>
</composer>
</composers>
\end{lstlisting}

For traditional lenses, these two transformations---i.e. computing the initial
view from the source and pushing the updated view back into the source---are
required to satisfy two important properties. First, if we \pput a view $v$
back into a source $s$ and the \get a view $v'$ from $s$, then $v' = v$. Second,
if we \get a view $v$ from a source $s$ and then \pput $v$ back into $s$ to \get a
source $s'$, then $s = s'$. These two laws are encapsulated in the equations
\begin{align}
\ell.\get \; (\ell.\pput \; s \; v) &= v \tag{PUTGET}\\
\ell.\pput \; s \; (\ell.\get \; s) &= s \tag{GETPUT}
\end{align}

While these two laws give some guarantees on how the \get and \pput components
of a lens interact, they are not strong enough to guarantee that the
updates made to a view are indeed the updates reflected back into the source.
This is because the GETPUT law only says that the \pput component must return
the original source unmodified whenever the update to the view is a no-op; it
does not say what the \pput component should do with an unmodified view.
Consequently, any non-trivial update to a view can lead to arbitrary update to
the source, or to be more precise, the lens laws do not prohibit such a
situation from occuring.

For example consider a lens $\ell$ whose \get component maps source data in the
form of integer pairs to view data in the form of an integer by simply
discarding the first element of the pair: that is, $\ell.\get (a, b) = b$. Given
a source $(a, b)$ and a view $c$, the \pput component of $\ell$ evaluates to
$(a, c)$ if $b=c$, and otherwise evaluates to $(0, c)$.

It is easy to verify that the PUTGET law and the GETPUT law hold for $\ell$.
However, observe that $\ell.\pput (1, 2) \; 3 = (0,3)$ despite the fact that
intuitively, $\ell$ should never update the first factor of a pair, since the
view does not contain any information about the first factor. In particular,
one would expect that $\ell.\pput (1, 2) \; 3 = (1,3)$, and in general,
$\ell.\pput (a, b) \; c = (a,c)$.
\iffalse
We would like a lens to guarantee that if we produce the view $\ell.\get \; s$
from $s$ and then modify $\ell.\get \; s$ to \get an updated view $v$, then the
changes reflected back into $s$ via $\ell.\pput \; s \; v$ are at most the changes
made to $\ell.\get \; s$ to \get $v$.

This implies that if $v \neq \ell.\get \; s$ and $v' \neq \ell.\get \; s$ are
arbitrary views with $v$ being very similar to $\ell.\get \; s$ and $v'$ being
very different from $\ell.\get \; s$, then $\ell.\pput \; s \; v$ should be very
similar to $s$, whereas $\ell.\pput \; s \; v'$ should be very different from $s$.

However, the GETPUT law only says that
$$\ell.\pput \; s \; (\ell.\get \; s) = s$$
This law nothing about the behaviour of $\ell.\pput$ when it is invoked with a
view different from $\ell.\get \; s$. So we could have $\ell.\pput \; s \; v =
\ell.\pput \; s \; v'$, even though they should be very different.

However, the GETPUT law only says that
$$\ell.\pput \; s \; (\ell.\get \; s) = s$$
This law nothing about the behaviour of $\ell.\pput$ when it is invoked with a
view different from $\ell.\get \; s$. Thus we could have $\ell.\pput \; s \; v$
reflecting changes that are different from the updates made to $\ell.\get \; s$
to acquire $v$, thus corrupting the source $s$.

In the rest of this article, we strengthen the lens laws so that if we produce
an updated view $v$ from $\ell.\get \; s$, then the changes reflected in
$\ell.\pput \; s \; v$ are at most the changes made to $\ell.\get \; s$ that
resulted in the updated view $v$.
\fi
\fi

\section{Difference Posets}
\begin{definition}
A {\em difference poset (DP)} is a triple $(P, \rho, \delta)$ where $P$ is
a poset and $\rho : P \longrightarrow \mathbb{R}^{\geq 0}$ and $\delta : P
\times P \longrightarrow P$ are functions satisfying :
\begin{align}
&\delta(a, b) \leq a\tag{A1}\\
&\text{if } a < a',\text{ then } \rho(a) < \rho(a') \tag{A2}\\
&\text{if } a \leq a' \text{ then for all }b , \; \rho(\delta(b, a)) \geq
\rho(\delta(b, a'))\tag{A3}
\end{align}
We call $\rho$ the weight function of $(P, \rho, \delta)$ and $\delta$ its
difference function.
\end{definition}
\iffalse
Intuitively, $\pi_1(\delta(a, b))$ tells us how much $a$ differs from $a \wedge
b$ while $\pi_2(\delta(a, b))$ tells us how much $b$ differs from $a \wedge b$.
The advantage of the definition of difference functions is that $\delta(a, b)$
is defined even if $a \wedge b$ does not exist. Still it is instructive to
think of $a \wedge b$ to understand what the axioms are saying.

The first law says that the difference between $a$ and $a \wedge b$ is less
than or equal to $a$, and the difference between $b$ and $a \wedge b$
is less than or equal to $b$.

The second law says that the difference between $a$ and $a \wedge b$ is equal
to the difference between $b$ and $a \wedge b$ if and only if $a = b$.

The third law says that switching $a$ and $b$ and then taking the difference
is the same as taking the difference and then switching the components.

The fourth law says that if $a \leq a'$ then the difference between $b$
and $a \wedge b$ is greater than or equal to the difference between $b$ and $a'
\wedge b$.
\fi
\begin{example}\label{embedding}
Let $P$ be any set with equality giving the poset structure. Given any
function $\rho: P \longrightarrow \mathbb{R}^{\geq 0}$ on $P$, and function
$\delta$ defined by $\delta(a, b) = a$, then $(P, \rho, \delta)$ is a difference
poset.
\end{example}
\begin{example}\label{naturalnumbers}
Let $P$ be the poset given by non-negative integers with the order $\leq$ and
weight function $\rho(a) = a$. Define $\delta$ by $\delta(a, b) = \max\{0,
a-b\}$. Then $(P, \rho, \delta)$ is a difference poset.
\end{example}
\begin{example}\label{divisibility}
Let $P$ be the poset given by non-negative integers with the division
relation and weight function $\rho(p_1^{a_1} \cdot \ldots \cdot p_n^{a_n}) =
a_1 + \ldots + a_n$. Define $\delta$ by $\delta(a, b) = a/gcd(a, b)$. Then $(P,
\rho, \delta)$ is a difference poset.
\end{example}
\begin{example}\label{boolean}
Let $X$ be a set and $P$ the poset given by the finite subsets of $2^X$ with the
containment relation and weight function $\rho(A) = |A|$. Define $\delta$
by $\delta(A, B) = A \setminus B$. Then $(P, \rho, \delta)$ is a difference
poset.
\end{example}
\begin{example}\label{general}
Generalizing \cref{divisibility} and \cref{boolean}, let $P$ be a complete
lattice and $\rho$ a strict order preserving map from $P$ to the non-negative
real numbers. Define $$\delta(a, b) = \rho\left(\bigwedge \{x \; | \; (a \wedge
b) \vee x = a\}\right)$$
Then $(P, \rho, \delta)$ is a difference poset.
\end{example}
\begin{proof}
Let $a, a', b \in P$.  Let $X = \{x \; | \; (a \wedge b) \vee x =  a\}$, $Y =
\{y \; | \; (b \wedge a) \vee y = b\}$ and $Y' = \{y' \; | \; (b \wedge a')
\vee y' = b\}$.
\begin{enumerate}
  \item[(A1)]
  Since $a \in X$ and $\delta(a, b) = \bigwedge X$, it follows that
  $\bigwedge X \leq a$.
  \item[(A2)]
  Follows from the fact that $\rho$ is a strict order preserving map from $P$ to
  the non-negative real numbers.
  \item[(A3)]
  Assume that $a \leq a'$. First we show that $Y \subseteq Y'$. Assume that
  $y \in Y$. Since $a \wedge b \leq a' \wedge b \leq b = (a \wedge b) \vee y$,
  it follows that, $(a' \wedge b) \vee y = (a \wedge b) \vee y = b$, hence $y
  \in Y'$. Consequently, $\bigwedge Y \geq \bigwedge Y'$ from which it follows
  that $$\rho(\delta(b, a)) =\rho\left(\bigwedge Y\right) \geq
  \rho\left(\bigwedge Y'\right) = \rho(\delta(b, a'))$$
  This completes the proof.
\end{enumerate}
\end{proof}
\iffalse
Now define a
function $d:
P \longrightarrow \mathbb{R}$ on $P$ by $$d(a, b) =
\begin{cases}
0 & \text{if } a = b\\
\rho(\pi_1(\delta(a, b))) + \rho(\pi_2(\delta(a, b))) & \text{otherwise}
\end{cases}$$
\begin{fact}
\begin{enumerate}
  \item
  $d(a,b) \geq 0$,
  \item
  $d(a, b) = 0$ if and only if $a = b$, and
  \item
  $d(a, b) = d(b, a)$
\end{enumerate}
\end{fact}
\begin{fact}
If $b \leq c$, then for all $b$, $d(a, b) \geq d(a, c)$. Thus if $b \leq c$,
$d(a, c) \leq d(a, b) + d(b, c)$
\end{fact}
\begin{question}
Under what conditions is $d$ a metric?
\end{question}
\fi
\section{New Difference Posets from Old}
Let $(P, \rho, \delta)$ and $(P', \rho', \delta)$ be difference posets.

\subsection{Product of Posets}
Define the product $P \times P'$ of $P$ and $P'$ to be the set $P \times P'$
with $\leq_{P \times P'} = \leq_P \times \leq_{P'}$. Define the weight function
$(\rho \times \rho')$ on $P \times P'$ by $(\rho \times \rho')(a, b) = \rho(a)
+ \rho(b)$. Define the difference function $\delta_P \times \delta_{P'}$ on $P
\times P$ by $$(\delta \times \delta')((a_1, b_1), (a_2, b_2)) = (\delta(a_1,
a_2), \delta(b_1, b_2))$$
\begin{claim}
$(P \times P', \rho \times \rho', \delta \times \delta)$ is a difference poset.
\end{claim}
\begin{proof}
We check that the definition that each of the axioms for a difference poset
holds.
\begin{enumerate}
  \item[(A1)]
  Since $\delta$ and $\delta'$ are difference functions, then $\delta(a_1, a_2)
  \leq a_1$ and $\delta'(b_1, b_2) \leq b_1$, hence $$
  (\delta \times \delta')((a_1, b_1), (a_2, b_2)) = (\delta(a_1, a_2),
  \delta(b_1, b_2)) \leq (a_1, b_1)$$
  \item[(A2)]
  Assume that $(a_1, b_1) < (a_2, b_2)$. Then since $\rho$ and $\rho'$ are
  weight functions, it follows that $\rho(a_1) < \rho(a_2)$ and
  $\rho'(b_1) < \rho'(b_2)$. Consequently,
  $$(\rho \times \rho')(a_1, b_1) = \rho(a_1) + \rho(b_1) < \rho(a_2) +
  \rho(b_2) = (\rho \times \rho)(a_2, b_2)$$
  \item[(A3)]
  Assume that $(a_1, b_1) \leq (a_2, b_2)$. Then since $\delta$ and $\delta'$ are
  difference functions, it follows that $\rho(\delta(a_3, b_1))
  \geq \rho(\delta(a_3, b_2))$ and $\rho'(\delta'(b_3, b_1)) \geq
  \rho'(\delta'(b_3, b_2))$.
  Consequently,
  \begin{align*}
((\rho \times \rho') \circ (\delta \times \delta'))((a_3, b_3), (a_1, b_1)))
&= (\rho \times \rho')(\delta(a_3, a_1), \delta'(b_3, b_1))\\
&= \rho(\delta(a_3, a_1)) + \rho'(\delta(b_3, b_1))\\
&\geq \rho(\delta(a_3, a_2)) + \rho'(\delta(b_3, b_2))\\
&= (\rho \times \rho')(\delta(a_3, a_2), \delta'(b_3, b_2))\\
&= ((\rho \times \rho') \circ (\delta \times \delta'))((a_3, b_3), (a_2,
b_2))\end{align*}

This completes the proof.
\end{enumerate}
\end{proof}
\subsection{Coproduct of Posets}
Define the coproduct $P \sqcup P'$ of $P$ and $P'$ to be the set $P \sqcup P'$
(disjoint union of $P$ and $P'$) with $x \leq_{P \sqcup P'} y$ if and only if
$x \leq_P y$ or $x \leq_{P'} y$. Define the weight function
$(\rho \sqcup \rho')$ on $P \sqcup P'$ by
$$(\rho \sqcup \rho')(a) =
\begin{cases}
\rho(a) &\text{if }a \in P\\
\rho'(a) &\text{if }b \in P
\end{cases}$$
Define the difference function $\delta_P \sqcup \delta_{P'}$ on $P \sqcup P'$ by
$$(\delta \sqcup \delta')(a, b) =
\begin{cases}
\delta(a, b) &\text{if }a, b \in P\\
\delta'(a, b) &\text{if }a, b \in P'\\
a & \text{otherwise}
\end{cases}$$
\begin{claim}\label{coproductdifference}
$(P \sqcup P', \rho \sqcup \rho', \delta \sqcup \delta')$ is a difference poset.
\end{claim}
\begin{proof}
We check that $\delta_P \sqcup \delta_{P'}$ satisfies the axioms for difference
functions.
\begin{enumerate}
  \item[(A1)]
  If $x, y \in P$ then $(\delta \sqcup \delta')(x, y) = \delta(x, y) \leq x$.
  A similar argument shows the same result if $x, y \in P'$. Otherwise $\delta
  \sqcup \delta'(a, b) = a$.
  \item[(A2)]
  Assume that $a \leq_{P \sqcup P'} b$. If $a \leq_P b$,
  $$(\rho \sqcup \rho')(a) = \rho(a) < \rho(b) = (\rho \sqcup \rho')(b)$$
  If $a \leq_{P'} b$, then
  $$(\rho \sqcup \rho')(a) = \rho'(a) < \rho'(b) = (\rho \sqcup \rho')(b)$$
  \item[(A3)]
  Assume that $a \leq_{P \sqcup P'} b$. Assume first that $a \leq_P b$. If $c
  \in P$, then $\rho(\delta(c, a)) \geq \rho(\delta(c, a'))$ since $\delta$ is a
  difference function. If $c \in P'$, then $\delta(c, a) = \delta(c, b) = c$,
  hence $\rho(\delta(c, a)) \geq \rho(\delta(c, a))$. A similar argument shows
  that if $a \leq_{P'} b$, then for all $c \in P \sqcup P'$, $\rho(\delta(c, a))
  \geq \rho(\delta(c, a'))$. This completes the proof.
\end{enumerate}
\end{proof}
\subsection{List Posets}
Assume that $\rho > 0$. Define the list poset $P^*$ of $P$ to be the set
$$P^0 \sqcup P \sqcup (P \times P) \sqcup \ldots \sqcup P^i \sqcup \ldots$$
This is just the set of all finite sequences of elements on $P$, with $P^0 =
\epsilon$, the empty sequence. The poset structure on $P^*$ is the poset
structure inherited from the product and coproduct constructions. Define the
weight function $\rho^*$ on $P^*$ by $\rho((x_1, \ldots, x_m)) = \rho(x_1) +
\ldots + \rho(x_m)$. Define the difference function $\delta^*$ on $P^*$ by
$$\delta^*(a, b) =
\begin{cases}
\delta^i(a, b) &\text{if }a, b \in P^i\\
a & \text{otherwise}
\end{cases}$$
\begin{claim}
$(P^*, \rho^*, \delta^*)$ is difference poset.
\end{claim}
\begin{proof}
Proceed as in the proof of \cref{coproductdifference}.
\end{proof}
\subsubsection{A Finer Poset Structure on $P^*$}
Define a finer poset structure on $P^*$ by $x \; \leq \; y$
if and only if $x = (x_1, \ldots, x_m), \; y = (y_1, \ldots, y_n)$, $m \leq n$
and $x_i \leq y_i$ for $1 \leq i \leq m$. Now let $m, n$ be any non-negative
integers and let $k = \min\{m, n\}$. We define the difference function
$\Delta^*$ on $P^*$ by 
$$\Delta^*((x_1, \ldots, x_m), (y_1, \ldots, y_n)) = ((\delta(x_1, y_1), \ldots,
\delta(x_k, y_k), x_{k+1}, \ldots, x_m)$$
\begin{claim}
$(P^*, \rho^*, \Delta^*)$ is a difference poset.
\end{claim}
\begin{proof}
We check that each of the axioms for a difference poset holds.
\begin{enumerate}
  \item[(A1)]
  Since $\delta$ is a difference function, $\delta(x_i, y_i)
  \leq x_i$ for $1 \leq i \leq \min\{m, n\}$, from which it follows that
  $\Delta^*((x_1, \ldots, x_m), (y_1, \ldots, y_n)) \leq (x_1, \ldots, x_m)$
  \item[(A2)]
  Assume that $(x_1, \ldots, x_m) < (y_1, \ldots, y_n)$. Then $m \leq
  n$ and $\rho(x_k) < \rho(y_k)$ for some $1 \leq k \leq m$ and $\rho(x_i)
  \leq \rho(y_i)$ for all $1 \leq i \leq m$, from which it follows that
  $$\rho^*(x_1, \ldots, x_m) = \sum_{i=1}^m \rho(x_i) < \sum_{i=1}^{n}\rho(y_i)
  = \rho^*(y_1, \ldots, y_n)$$
  \item[(A3)]
  Assume that $(x_1, \ldots, x_m) \leq (y_1, \ldots, y_{m'})$. Then $m \leq
  m'$ and $x_i \leq y_i$ for $1 \leq i \leq m$. Let $k = \min\{m, n\}$ and $k' =
  \min\{m', n\}$. By the definition of $\Delta^*$,
  \begin{align*}
\Delta^*((z_1, \ldots, z_n), (x_1, \ldots, x_m)) &= (\delta(z_1, x_1),
\delta(z_k, x_k), z_{k+1}, \ldots, z_n)\\
\Delta^*((z_1, \ldots, z_n), (y_1, \ldots, y_{m'})) &= (\delta(z_1, y_1),
\delta(z_k, y_{k'}), z_{k'+1}, \ldots, z_n)
\end{align*}
As $\delta$ is a difference function, it follows that $\delta(z_i, x_i) \geq
\delta(z_i, y_i)$ for $1 \leq i \leq k$, hence
$$\Delta^*((z_1, \ldots, z_n), (x_1, \ldots, x_m)) \geq \Delta^*((z_1,
\ldots, z_n), (y_1, \ldots, y_{m'}))$$
This completes the proof.
\end{enumerate}
\end{proof}

\subsubsection{An Even Finer Poset Structure on $P^*$}\label{nocrossing}
Define a finer poset structure on $P^*$ by $x \leq y$ if
and only if $x = (x_1, \ldots, x_m), \; y = (y_1, \ldots, y_n)$, $m \leq n$ and
there exists an increasing injective function $\varphi : [m] \longrightarrow
[n]$ such that $x_i \leq y_{\varphi(i)}$ for $1 \leq i \leq m$.

Now let $x = (x_1, \ldots, x_m), \; y = (y_1, \ldots, y_n)$. We define the
function $\Gamma^*$ on $P^*$ by taking the difference of $x_i$ and
$y_{\varphi(i)}$ for some ``maximal'' increasing injective function $\varphi :
A \longrightarrow [n]$ where $A$ is a subset of $[m]$, except that we eliminate
all $x_i$ where $x_i = y_{\varphi(i)}$.

For example, assume that $P$ is the set consisting of the letters of the
alphabet where the poset structure is given by equality, with $\rho(a) = 1$
and $\delta(a, b) = a$ for all $a, b \in P$. Then $\Gamma^*((a, c, d, f, b, e),
(a, b, d, e)) = (c, f, b)$ where we eliminate $a$, $d$ and $e$ from the first
sequence. However, even though $b$ occurs in both sequences, we are not allowed
to eliminate both $b$ and $d$ since these two characters occur in the order $d,
b$ in the first sequence and in the order $b, d$ in the second sequence. If we
were to eliminate both of them, then the matching would have to have a
crossing, thus contradicting the monotonicity of $\varphi$.

Given any increasing injective $\varphi : A \longrightarrow [n]$ for some
subset $A$ of $[m]$, define $\Gamma^*_{\varphi}(x, y)$ to be the difference
between $x$ and $y$ computed using $\varphi$. Then the maximal $\varphi$ is the
$\varphi$ that minimizes $\rho^*(\Gamma^*_{\varphi}(x, y))$. If there are
multiple such $\varphi$, we choose the one that maximizes $|A|$. If there are
still multiple such $\varphi$, we choose the one for which $A$ is
lexicographically minimal.
\begin{claim}\label{nocrossingproof}
$(P^*, \rho^*, \Gamma^*)$ is a difference poset.
\end{claim}
\begin{proof}
Let $x = (x_1, \ldots, x_m)$, $y = (y_1, \ldots, y_n)$ and $z = (z_1, \ldots,
z_k)$. We check that each of the axioms for a difference poset holds.
\begin{enumerate}
  \item[(A1)]
  Since each element of $\Gamma^*(x, y)$ is either equal to $\Gamma(x_i,
  y_{\varphi(i)})$ for some $\varphi$, or to $x_j$ for some $j$, and since these
  elements occur in increasing order, it follows that $\Gamma^*(x, y) \leq x$.
  \item[(A2)]
  Assume that $(x_1, \ldots, x_m) < (y_1, \ldots, y_n)$. Then $m \leq
  n$ and there exists an increasing injective function $\varphi:[m]
  \longrightarrow [n]$ such that $\rho(x_k) < \rho(y_{\varphi(k)})$ for some $1
  \leq k \leq m$ and $\rho(x_i) \leq \rho(y_{\varphi(i)})$ for all $1 \leq i
  \leq m$. It thus follows that $$\rho^*(x_1, \ldots, x_m) = \sum_{i=1}^m
  \rho(x_i) < \sum_{i=1}^{n}\rho(y_i) = \rho^*(y_1, \ldots, y_n)$$
  \item[(A3)]
  Assume that $(x_1, \ldots, x_m) \leq (y_1, \ldots, y_n)$. Assume that
  $\Gamma^*(z, x) = \Gamma^*_{\varphi}(z, x)$ for some $\varphi$. Then
  $\rho^*(\Gamma^*_{\varphi} (z, x)) \geq \rho^*(\Gamma^*_{\varphi}(z, y))$
  since $\Gamma$ is a difference function by assumption. Consequently
  $\rho^*(\Gamma^*_{\varphi} (z, x)) \geq \rho^*(\Gamma^*_{\varphi}(z, y)) \geq
  \rho^*(\Gamma^*(z, y))$ since $\Gamma^*(z, y)$ is computed using a $\psi$ that
  minimizes the quantity $\rho^*(\Gamma^*_{\psi}(z, y))$. This completes the
  proof.
\end{enumerate}
\end{proof}
\subsubsection{Quotienting $P^*$ to Ignore Order}
We may try to define an even finer poset structure on $P^*$ by $x \leq y$ if and
only if $x = (x_1, \ldots, x_m), \; y = (y_1, \ldots, y_n)$, $m \leq n$ and
there exists an injective function $\varphi : [m] \longrightarrow [n]$ such
that $x_i \leq y_{\varphi(i)}$ for $1 \leq i \leq m$: we drop the requirement
that $\varphi$ be increasing.

With this poset structure, $(x_{\varphi(1)}, \ldots, x_{\varphi(n)}) \leq
(x_{\psi(1)}, \ldots, x_{\psi(n)})$ for all permutations
$\varphi, \psi$ of $[n]$. The poset axioms thus force us to identify
$(x_{\varphi(1)}, \ldots, x_{\varphi(n)})$ and $(x_{\psi(1)}, \ldots,
x_{\psi(n)})$ for all $\varphi, \psi$. Thus we need to formalize the notion of
{\em quotient posets}.

\begin{definition}
Let $P$ be a poset and let $\sim$ be an equivalence relation on $P$. Define the
quotient $P/{\sim}$ to be the set of equivalence classes with the binary
relation $\leq$ defined by $[x] \leq [y]$ if and only if $x' \leq y'$ for some
$x' \in [x]$ and $y' \in y$
\end{definition}
The quotient $P/{\sim}$ in general is not a poset. For example, take a
3-element chain and identify the bottom and top elements. The relation you
obtain is reflexive and transitive, but not antisymmetric. However, the
following claim gives a condition that guarantees that $P/{\sim}$ is indeed a
poset.

\begin{claim}\label{quotientposet}
Assume that $P$ is a poset and $\rho$ is a strict order preserving map from $P$
to the non-negative real numbers. Let $\sim$ be an equivalence relation on $P$.
Assume that if $x \sim x'$ then $\rho(x) = \rho(x')$. Further, assume that if
$[x] \leq [y]$, then for all $x' \in [x]$ there is a $y' \in y$ such that $x'
\leq y'$. Then $P/{\sim}$ is a poset, and $\rho$ induced on $P/{\sim}$ is a
a strict order preserving map from $P/{\sim}$ to the non-negative real numbers.
\end{claim}
\begin{proof}
Certainly $\leq$ is reflexive in $P/{\sim}$.

To prove transitivity, assume that $[x] \leq [y]$ and $[y] \leq [z]$. Then
there exist $x' \in [x]$, $y' \in [y]$ such that $x' \leq y'$. Since $[y] \leq
[z]$, then by assumption, there exists a $z' \in [z]$ such that $y' \leq z'$.
Consequently, $x' \leq y' \leq z'$, thus $[x] \leq [z]$.

Finally, to prove antisymmetry, assume that $[x] \leq [y]$ and $[y] \leq [x]$.
Then there exist $x', x'' \in [x]$ and $y', y'' \in [y]$ such that $x' \leq y'$
and $y'' \leq x''$. Assume for a contradiction that $[x] \neq [y]$. Since $\sim$
is an equivalence relation, $[x]$ and $[y]$ are disjoint. Consequently $x' <
y'$ and $y'' < x''$. By assumption, $\rho(x') = \rho(x)$ for all $x' \in [x]$
and $\rho(y') = \rho(y)$ for all $y' \in [y]$. As $\rho$ is strict, then
$\rho(x) = \rho(x') < \rho(y') = \rho(y)$ on the one hand, but on the other
hand, $\rho(y) = \rho(y'') < \rho(x'') = \rho(x)$, a contradiction. Thus $[x] =
[y]$.

Thus $P/{\sim}$ is a poset, and the assumption that $\rho(x) = \rho(x')$
if $x \sim x'$ assures us that $\rho$ induced on $P/{\sim}$ is a
a strict order preserving map from $P/{\sim}$ to the non-negative real numbers.
\end{proof}

Recall that we wished to define an even finer poset structure on $P^*$ by $x
\leq y$ if and only if $x = (x_1, \ldots, x_m), \; y = (y_1, \ldots, y_n)$, $m
\leq n$ and there exists an injective function $\varphi : [m] \longrightarrow
[n]$ such that $x_i \leq y_{\varphi(i)}$ for $1 \leq i \leq m$. We observed
that with this finer poset structure on $P^*$, we need to identify
$(x_{\varphi(1)}, \ldots, x_{\varphi(n)})$ and $(x_{\psi(1)}, \ldots,
x_{\psi(n)})$ for all permutations $\varphi$ and $\psi$. In other words we take
the poset structure $\leq_{NC}$ (NC for non-crossing) described in
\cref{nocrossing} and quotient $P^*$ by the equivalence relation which
identifies $(x_{\varphi(1)}, \ldots, x_{\varphi(n)})$ and $(x_{\psi(1)},
\ldots, x_{\psi(n)})$ for all permutations $\varphi$ and $\psi$

In order to confirm that $P^*/{\sim}$ is indeed a poset, then we show that the
hypotheses of \cref{quotientposet} apply to $P^*/{\sim}$. Let $x=(x_1, \ldots,
x_m)$ and $y = (y_1, \ldots, y_n)$. First, we observe that we only identify
$(x_{\varphi(1)}, \ldots, x_{\varphi(n)})$ and $(x_{\psi(1)}, \ldots,
x_{\psi(n)})$ for all permutations $\varphi$ and $\psi$, hence if $x \sim x'$,
then $\rho^*(x) = \rho^*(x')$. Next, we need to show that if $[x] \leq_{NC}
[y]$, then for all $x'' \in [x]$ there is a $y'' \in y$ such that $x''
\leq_{NC} y''$.

Assume that $[x] \leq_{NC} [y]$. Then by the definition of $\leq_{NC}$ and
$\sim$, there exists a permutation $\varphi$ of $[m]$ and a permutation
$\psi$ of $[n]$ such that $(x_{\varphi(1)}, \ldots, x_{\varphi(m)}) \leq_{NC}
(y_{\psi(1)}, \ldots, y_{\psi(n)})$.  Let $x'' = (x_{\theta(1)}, \ldots,
x_{\theta(m)})$ for some permutation of $[m]$. We need to show that there
exists a permutation $\pi$ of $[n]$ such that $(x_{\theta(1)}, \ldots,
x_{\theta(m)}) \leq_{NC} (y_{\pi(1)}, \ldots, y_{\pi(n)})$, but this follows
from a graphical argument; just permute the $y_i$'s to remove crossings.
Therefore, $P^*/{\sim}$ is a poset, and $\rho$ induced on $P/{\sim}$ is a
a strict order preserving map from $P/{\sim}$ to the non-negative real numbers.

We define the difference function $\Gamma^*/{\sim}$ on $P/{\sim}$ by taking
the matching which minimizes $\rho^*$ on the difference as in the non-crossing
case. If there are multiple matchings we just pick one! As in the non-crossing
case, we define $\Gamma^*/{\sim}$ on $P^*/{\sim}$ by eliminating all $x_i$ from
$x$ where $x_i = y_{\varphi(i)}$ and $\varphi$ is the canonical maximal
matching. Going back to our previous example, then $ \Gamma^*([(a, c, d, f, b,
e)], [(a, b, d, e)]) = [(c, f)]$ where we now can eliminate $a, b, d$ and $e$
from the first sequence.
\begin{claim}
$\Gamma^*/{\sim}$ is a difference function.
\end{claim}
\begin{proof}
Proceed as in the proof of \cref{nocrossingproof}.
\end{proof}
\section{Lenses}
\begin{definition}
Let $S, V$ be difference posets. A lens $\ell : S \Leftrightarrow V$ is a
3-tuple of functions
\begin{align*}
\ell.\get &: S \longrightarrow V\\
\ell.\pput &: S \longrightarrow V \longrightarrow S\\
\ell.\create &: V \longrightarrow S
\end{align*}
obeying the following laws for every $s \in S$ and $v \in V$:
\begin{align*}
\ell.\get \; (\ell.\pput \; s \; v) &= v \tag{PUTGET}\\
\ell.\get \; (\ell.\create \; v) &= v \tag{CREATEGET}\\
\ell.\pput \; s \; (\ell.\get \; s) &= s \tag{GETPUT}\\
\delta_S(\ell.\pput \; s \; v, s) &= \ell.\pput \; s \; (\delta_V(v, \ell.\get \; s))
\tag{PUTEXACT}\\
\delta_S(\ell.\create \; v, s) &= \ell.\create \; (\delta_V(v,
\ell.\get \; s))
\tag{CREATEEXACT}
\end{align*}
\end{definition}
These laws are the same as the classical lens laws except for new EXACT rule,
which says that $\ell.\pput$ and $\ell.\create$ preserve differences.

Every classical lens can be made into a lens by taking the poset structure on
$S$ and $V$ to be given by equality with arbitrary weight functions, and
difference functions $\delta$ given by $\delta(a, b) = a$ as in
\cref{embedding}.
\subsection{Boomerang Lens Combinators}
\subsubsection{Copy}
Let $(P, \rho, \delta)$ be any difference poset. Define the $\ccopy : P
\Leftrightarrow P$ by
\begin{align}
\ccopy.\get \; s &= s\\
\ccopy.\pput \; s \; v &= v\\
\ccopy.\create \; v &= v
\end{align}
\begin{claim}
\ccopy is a valid lens.
\end{claim}
\begin{proof}
First, we need to show that $\delta(\ccopy.\pput \; s \;v, s) = \ccopy.\pput \;
s \; (\delta(v, \ccopy.\get \; s))$, and this follows from the fact that
$$\delta(\ccopy.\pput \; s \;v, s) = \delta(v, s) = \delta(v, \ccopy.\get \;s)
= \ccopy.\pput \; s \; (\delta(v, \ccopy.\get \; s))
$$
Next, we need to show that $\delta(\ccopy.\create \;v, s) = \ccopy.\create \;
(\delta(v, \ccopy.\get \; s))$
\end{proof}

\subsubsection{Constant}
Let $P$ be any weighted poset with difference function $\delta(a, b) = a$.
Let $\{*\}$ be a one element weighted poset. Let $p \in P$. Define the \const
lens $\const : P \Leftrightarrow \{*\}$ by
\begin{align*}
\const.\get \; s &= *\\
\const.\pput \; s \; v &= s\\
\const.\create \; v &= p
\end{align*}
\begin{claim}
\const is a valid lens.
\end{claim}
\begin{proof}
First, we need to show that $\delta(\const.\pput \; s \;v, s) = \delta(v,
\const.\get \; s)$, and this follows from the fact that
$$\delta(\const.\pput \; s \;v, s) =
\delta(s, s) = s = \const.\pput \; s \; (\delta(v, v)) = \const.\pput \; s
\; (\delta(v, \const.\get \;s)) $$

Next, we need to show that $\delta(\const.\create \;v, s) =
\const.\create \; \delta(v, \const.\get \; s)$, and this follows from the fact
that
$$\delta(\const.\create \;v, s) = \delta(p, s) = p = \const.\create \; \delta(v,
\const.\get \; s)$$
\end{proof}
\subsubsection{Default}
Let $\ell : S \longrightarrow V$ be a lens. Let $f : V \longrightarrow S$ be any
function. Define the lens $\default(\ell) : V \Leftrightarrow S$ by
\begin{align*}
\default(\ell).\get \; s &= \ell.\get \; s\\
\default(\ell).\pput \; s \; v &= \ell.\pput \; s \; v\\
\default(\ell).\create \; v &= \ell.\pput \; (f \; v) \; v
\end{align*}
\begin{proof}
We need to show that
$$\delta(\default.\create \; v, s) = \default.\create \; s \; (\delta(v,
\ell.\get \; s))$$
and this follows from the fact that 
$$\delta(\default.\create \; v, s) = \delta(\ell.\pput \; (f \; v), s) =
\ell.\pput \; (f \; v) \; (\delta(v, \ell.\get \; s)) = \default.\create \; s
\; (\delta(v, \ell.\get \; s))$$
\end{proof}
\subsubsection{Product}
Let $\ell_1 : S_1 \Leftrightarrow V_1$ and $\ell_2 : S_2 \Leftrightarrow V_2$ be
lenses. Define the lens $\ell_1 \times \ell_2 : (S_1 \times S_2) \Leftrightarrow
(V_1 \times V_2)$ by
\begin{align*}
(\ell_1 \times \ell_2).\get \; (s_1, s_2) &= (\ell_1.\get \; s_1, \ell_2.\get
\; s_1)\\
(\ell_1 \times \ell_2).\pput \; (s_1, s_2) \; (v_1, v_2) &= (\ell_1.\pput \; s_1
\; v_1, \ell_2.\pput \; s_2 \; v_2)\\
(\ell_1 \times \ell_2).\create \; (v_1, v_2) &= (\ell_1.\create \; v_1,
\ell_2.\create \; v_2)
\end{align*}
\begin{claim}
$\ell_1 \times \ell_2$ is a valid lens.
\end{claim}
\begin{proof}
First, we need to show that
$$\delta((\ell_1 \times \ell_2).\pput \; (s_1, s_2) \; (v_1, v_2), (s_1, s_2)) =
(\ell_1 \times \ell_2).\pput \; (s_1, s_2) \; \delta((v_1, v_2), (\ell_1 \times
\ell_2).\get (s_1, s_2))$$ This follows from the fact that
\begin{align*}
\delta((\ell_1 \times \ell_2).\pput \; (s_1, s_2) \; (v_1, v_2), (s_1, s_2)) &=
\delta((\ell_1.\pput \; s_1\; v_1, \ell_2.\pput \; s_2 \; v_2), (s_1, s_2))\\
&= (\delta (\ell_1.\pput \; s_1\; v_1, s_1), \delta (\ell_2.\pput \; s_2 \;
v_2, s_1))\\
&= (\ell_1.\pput \; s_1\; (\delta(v_1, \ell_1.\get \; s_1)), \ell_2.\pput
\; s_2\; (\delta(v_2, \ell_1.\get \; s_2)))\\
&= (\ell_1 \times \ell_2).\pput \; (s_1, s_2) \; (\delta(v_1, \ell_1.\get \;
s_1), \delta(v_2, \ell_1.\get \; s_2))\\
&= (\ell_1 \times \ell_2).\pput \; (s_1, s_2) \; \delta((v_1, v_2), (\ell_1.\get
\; s_1, \ell_2.\get s_2))\\
&= (\ell_1 \times \ell_2).\pput \; (s_1, s_2) \; \delta((v_1, v_2), (\ell_1
\times \ell_2).\get (s_1, s_2))
\end{align*}
Next, we need to show that
$$\delta((\ell_1 \times \ell_2).\create \; (v_1, v_2), (s_1, s_2))
= (\ell_1 \times \ell_2).\create \; \delta((v_1, v_2), (\ell_1
\times \ell_2).\get (s_1, s_2))$$
which follows from the fact that
\begin{align*}
\delta((\ell_1 \times \ell_2).\create \; (v_1, v_2), (s_1, s_2)) &=
\delta((\ell_1.\create \; v_1, \ell_2.\create \; v_2), (s_1, s_2))\\
&= (\delta(\ell_1.\create \; v_1, s_1), \delta(\ell_2.\create \; v_2, s_2))\\
&= (\ell_1.\create \; (\delta(v_1, \ell_1.\get \; s_1)), \ell_2.\create \;
(\delta(v_2, \ell_2.\get \; s_2)))\\
&= (\ell_1 \times \ell_2).\create \; ((\delta(v_1, \ell_1.\get \; s_1)),
(\delta(v_2, \ell_2.\get \; s_2)))\\
&= (\ell_1 \times \ell_2).\create \; \delta((v_1, v_2), (\ell_1.\get \;
s_1, \ell_2.\get \; s_2))\\
&= (\ell_1 \times \ell_2).\create \; \delta((v_1, v_2), (\ell_1
\times \ell_2).\get (s_1, s_2))
\end{align*}
\end{proof}
\subsubsection{Sum}
Let $\ell_1 : S_1 \Leftrightarrow V_1$ and $\ell_2 : S_2 \Leftrightarrow V_2$ be
lenses. Define the lens $\ell_1 \sqcup \ell_2 : (S_1 \sqcup S_2) \Leftrightarrow
(V_1 \sqcup V_2)$ by
\begin{align*}
(\ell_1 \sqcup \ell_2).\get \; s &= 
\begin{cases}
\ell_1.\get \; s & \text{if } s \in S_1\\
\ell_2.\get \; s & \text{if } s \in S_2\\
\end{cases}\\
(\ell_1 \sqcup \ell_2).\pput \; s \; v&= 
\begin{cases}
\ell_1.\pput \; s \; v& \text{if } s \in S_1, \; v \in V_1\\
\ell_2.\pput \; s \; v& \text{if } s \in S_2, \; v \in V_2\\
\ell_1.\create \; v & \text{if } s \in S_2, \; v \in V_1\\
\ell_2.\create \; v & \text{if } s \in S_1, \; v \in V_2\\
\end{cases}\\
(\ell_1 \sqcup \ell_2).\create \; v &= 
\begin{cases}
\ell_1.\create \; v & \text{if } v \in V_1\\
\ell_2.\create \; v & \text{if } v \in V_2\\
\end{cases}
\end{align*}
\begin{claim}
$\ell_1 \sqcup \ell_2$ is a valid lens.
\end{claim}
\begin{proof}
We need to show that
$$\delta((\ell_1 \sqcup \ell_2).\pput \; s \; v, s) = (\ell_1 \sqcup
\ell_2).\pput \; s \; \delta(v, (\ell_1 \sqcup \ell_2).\get \; s)$$

In the first case, assume that $s \in S_1$, $v \in V_1$. Then we need to show
that $$\delta(\ell_1.\pput \; s \; v, s) = \ell_1.\pput \; s \; \delta(v,
\ell_1.\get \; s)$$
which follows from the fact that $\ell_1$ is a lens. A similar proof gives the
same result if $s \in S_2$ and $v \in v_2$.

Now assume that $s \in S_1$ and $v \in V_2$. Then we need to show
that 
$$\delta(\ell_2.\create \; v, s) = \ell_2.\create \; (\delta(v, \ell_2.\get \;
s))$$ which follows from the fact that $\ell_2$ is a lens. A similar proof gives
the same result if $s \in S_2$ and $v \in V_1$. This completes the proof.
\end{proof}
\subsubsection{Iteration}
Let $\ell: S \Leftrightarrow V$ be a lens. Define the lens $\ell^*: S
\Leftrightarrow V$ by
\begin{align*}
\ell^* .\get \; (s_1, \ldots, s_n) &= (\ell.\get \; s_1) \cdot \ldots \cdot
(\ell.\get \; s_n)\\
\ell^* .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n) &= s'_1 \cdot \ldots
\cdot s'_n\\
\text{ where } s'_i &= \begin{cases}
\ell.\pput \; s_i \; v_i & \text{if } 1 \leq i \leq \min\{m, n\}\\
\ell.\create \; v_i & \text{if } \min\{m, n\} < i \leq n
\end{cases}\\
\ell^*.\create \; (v_1, \ldots, v_n) &= (\ell.\create \; v_1) \cdot \ldots \cdot
(\ell.\create \; v_n)
\end{align*}
\begin{claim}
$\ell_1 \times \ell_2$ is a valid lens.
\end{claim}
\begin{proof}
First, we need to show that
\begin{multline}\label{positionalputexact}
\delta(\ell^*.\pput \; (s_1, \ldots, s_n) \; (v_1, \ldots, v_m),
(s_1, \ldots, s_n)) = \\ \ell^*.\pput \; (s_1, \ldots, s_n) \; \delta((v_1,
\ldots, v_m), \ell^*.\get \; (s_1, \ldots, s_n))
\end{multline}
Now
\begin{align*}
\ell^* .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n) &= s'_1 \cdot \ldots
\cdot s'_n\\
\text{ where } s'_i &= \begin{cases}
\ell.\pput \; s_i \; v_i & \text{if } 1 \leq i \leq \min\{m, n\}\\
\ell.\create \; v_i & \text{if } \min\{m, n\} < i \leq n
\end{cases}\\
\end{align*}
hence
\begin{align*}
\delta(\ell^* .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n)) &= s'_1 \cdot
\ldots \cdot s'_n\\
\text{ where } s'_i &= \begin{cases}
\delta(\ell.\pput \; s_i \; v_i) & \text{if } 1 \leq i \leq \min\{m, n\}\\
\delta(\ell.\create \; v_i) & \text{if } \min\{m, n\} < i \leq n
\end{cases}\\
\end{align*}
Since $\delta$ is a difference function, then
\begin{align*}
\delta(\ell^* .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n)) &= s'_1 \cdot
\ldots \cdot s'_n\\
\text{ where } s'_i &= \begin{cases}
\ell.\pput \; s_i \; \delta(v_i, \ell.\get \; s_i) & \text{if } 1 \leq i \leq
\min\{m, n\}\\
\ell.\create \; \delta(v_i, \ell.\get \; s_i) & \text{if } \min\{m, n\} < i \leq
n \end{cases}\\
\end{align*}
Now since
\begin{equation}\label{expanddelta^n}
\delta((v_1, \ldots, v_m), \ell^*.\get \; (s_1, \ldots, s_n)) = 
\delta(v_1, \ell.\get \; s_1) \cdot \ldots \cdot \delta((v_n, \ell.\get \; s_n))
\end{equation}
then
\begin{multline}
\ell^*.\pput \; (s_1, \ldots, s_n) \; \delta((v_1,
\ldots, v_m), \ell^*.\get \; (s_1, \ldots, s_n)) = \\
\ell^*.\pput \; (s_1, \ldots, s_n) \; \delta(v_1, \ell.\get \; s_1) \cdot
\ldots \cdot \delta((v_n, \ell.\get \; s_n))
\end{multline}

\end{proof}
\subsubsection{Composition}
Let $S, U, V$ be weighted posets. Let $\ell_1 : S \Leftrightarrow U$ and $\ell_2
: U \Leftrightarrow V$ be lenses. To show that $\ell_2 \circ \ell_1$ is also a
lens, we need to show that $\ell_2 \circ \ell_1$ satisfies the EXACT law:
\begin{multline}\label{composeresult}
\delta_S(\ell_1.\pput \; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; v), s) =
\ell_1.\pput \; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; (\delta_V(v, \ell_2.\get
\; (\ell_1.\get \; s))))
\end{multline}
Since $\ell_1$ and $\ell_2$ are lenses, then for all $s \in S$, $u \in U$ and $v
\in V$, we have
\begin{equation}\label{compose1}
\delta_S(\ell_1.\pput \; s \; u, s) = \ell_1.\pput \; s \; (\delta_U(u, \ell_1.\get
\; s))
\end{equation}
and
\begin{equation}\label{compose2}
\delta_U(\ell_2.\pput \; u \; v, u) = \ell_2.\pput \; u \; (\delta_V(v, \ell_2.\get
\; u))
\end{equation}

\noindent Substituting $\ell_2.\pput \; (\ell_1.\get \; s) \; v$ for $u$ in
\cref{compose1} and $\ell_1.\get \; s$ for $u$ in \cref{compose2} we \get
\begin{equation}\label{compose1subst}
\delta_S(\ell_1.\pput \; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; v), s) =
\ell_1.\pput \; s \; (\delta_U(\ell_2.\pput \; (\ell_1.\get \; s) \; v, \ell_1.\get
\; s))
\end{equation}
and
\begin{equation}\label{compose2subst}
\delta_U(\ell_2.\pput \; (\ell_1.\get \; s) \; v, (\ell_1.\get \; s)) = \ell_2.\pput
\; (\ell_1.\get \; s) \; (\delta_V(v, \ell_2.\get \; (\ell_1.\get \; s)))
\end{equation}
Applying $(\ell_1.\pput \; s)$ to \cref{compose2subst} we \get
\begin{equation}\label{compose3subst}
\ell_1. \pput \; s \; (\delta_U(\ell_2.\pput \; (\ell_1.\get \; s) \; v, \ell_1.\get
\; s)) = \ell_1.\pput \; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; (\delta_V(v,
\ell_2.\get \; (\ell_1.\get \; s))))
\end{equation}
Equating \cref{compose1subst} and \cref{compose3subst} hence gives
\cref{composeresult}, which is what we wanted to show.

\iffalse
\section{Weighted Posets and Difference Functions}
\begin{definition}
Let $P$ be a poset. A weight function $\rho$ on $P$ is simply an order
preserving map from $P$ to $\mathbb{R}^{\geq 0}$, the non-negative real
numbers.
\end{definition}

\begin{definition}
Let $(P, \rho)$ be a weighted poset. A function $\delta : P
\times P \longrightarrow \mathbb{R}^{\geq 0}$ is a difference function if and
only if
\begin{align*}
&\delta(a, b) \leq \rho(a) \tag{A1}\\
&\text{if } a \leq b \text{ then } \delta(a, b) = 0 \tag{A2}\\
&\text{if } a \leq a' \text{ then for all }b , \; \delta(b, a) \geq \delta(b,
a')\tag{A3}
\end{align*}
\end{definition}
\iffalse
Intuitively, $\pi_1(\delta(a, b))$ tells us how much $a$ differs from $a \wedge
b$ while $\pi_2(\delta(a, b))$ tells us how much $b$ differs from $a \wedge b$.
The advantage of the definition of difference functions is that $\delta(a, b)$
is defined even if $a \wedge b$ does not exist. Still it is instructive to
think of $a \wedge b$ to understand what the axioms are saying.

The first law says that the difference between $a$ and $a \wedge b$ is less
than or equal to $a$, and the difference between $b$ and $a \wedge b$
is less than or equal to $b$.

The second law says that the difference between $a$ and $a \wedge b$ is equal
to the difference between $b$ and $a \wedge b$ if and only if $a = b$.

The third law says that switching $a$ and $b$ and then taking the difference
is the same as taking the difference and then switching the components.

The fourth law says that if $a \leq a'$ then the difference between $b$
and $a \wedge b$ is greater than or equal to the difference between $b$ and $a'
\wedge b$.
\fi
\begin{example}\label{embedding}
Let $P$ be any set with equality giving the poset structure. Define
a weight function $\rho$ on $P$ by $\rho(a) = 1$ for all $a$. Then $\delta$
defined by $$ \delta(a, b) =
\begin{cases}
0 & \text{if } a = b\\
1 & \text{otherwise}
\end{cases}
$$
is a difference function on $P$.
\end{example}
\begin{example}\label{naturalnumbers}
Let $P$ be the weighed poset on the non-negative integers with the order $\leq$,
and weight function $\rho$ given by $Id_P$. Define $\delta(a, b) = \max\{0,
a-b\}$
\end{example}
\begin{example}\label{divisibility}
Let $P$ be the weighted poset on the non-negative integers with the division
relation and weight function $\rho$ given by $\rho(p_1^{a_1} \cdot \ldots \cdot
p_n^{a_n}) = a_1 + \ldots + a_n$. Define $\delta(a, b) = \rho(a/gcd(a, b))$.
\end{example}
\begin{example}\label{boolean}
Let $X$ be a set and $P$ the weighted poset on the finite subsets of $2^X$ with
the containment relation, and weight function $\rho$ given by $\rho(A) = |A|$.
Define $\delta(A, B) = A \setminus B$
\end{example}
\begin{example}\label{general}
Generalizing \cref{divisibility} and
\cref{boolean}, let $P$ be a complete weighted lattice with weight function
$\rho$ satisfying $\rho(x) = 0$ if and only if $x = \bigwedge P$. Define
$$\delta(a, b) = \rho\left(\bigwedge \{x \; | \; (a \wedge b) \vee x =
a\}\right)$$
\end{example}
\begin{proof}
Let $a, a', b \in P$.  Let $X = \{x \; | \; (a \wedge b) \vee x =  a\}$, $Y =
\{y \; | \; (b \wedge a) \vee y = b\}$ and $Y' = \{y' \; | \; (b \wedge a')
\vee y' = b\}$.
\begin{enumerate}
  \item[(A1)]
  Since $a \in X$ and $\delta(a, b) = \bigwedge X$, it follows that
  $\bigwedge X \leq a$, hence $\delta(a, b) = \rho\left(\bigwedge X \right) \leq
  \rho(a)$.
  \item[(A2)]
  Assume that $a \leq b$. Then $(a \wedge b) = a$, hence
  $$(a \wedge b) \vee \left(\bigwedge P\right) = a \vee\left(\bigwedge P\right)
  = a$$
  Hence $\bigwedge P \in X$, from which it follows that $\delta(a, b) =
  \rho\left(\bigwedge P\right) = 0$
  \item[(A4)]
  Assume that $a \leq a'$. First we show that $Y \subseteq Y'$. Assume that
  $y \in Y$. Since $a \wedge b \leq a' \wedge b \leq b = (a \wedge b) \vee y$,
  it follows that, $(a' \wedge b) \vee y = (a \wedge b) \vee y = b$, hence $y
  \in Y'$. Consequently, $\bigwedge Y \geq \bigwedge Y'$ from which it follows
  that $$\delta(b, a) =\rho\left(\bigwedge Y\right) \geq \rho\left(\bigwedge
  Y'\right) = \delta(b, a')$$
  This completes the proof.
\end{enumerate}
\end{proof}
\iffalse
Now define a
function $d:
P \longrightarrow \mathbb{R}$ on $P$ by $$d(a, b) =
\begin{cases}
0 & \text{if } a = b\\
\rho(\pi_1(\delta(a, b))) + \rho(\pi_2(\delta(a, b))) & \text{otherwise}
\end{cases}$$
\begin{fact}
\begin{enumerate}
  \item
  $d(a,b) \geq 0$,
  \item
  $d(a, b) = 0$ if and only if $a = b$, and
  \item
  $d(a, b) = d(b, a)$
\end{enumerate}
\end{fact}
\begin{fact}
If $b \leq c$, then for all $b$, $d(a, b) \geq d(a, c)$. Thus if $b \leq c$,
$d(a, c) \leq d(a, b) + d(b, c)$
\end{fact}
\begin{question}
Under what conditions is $d$ a metric?
\end{question}
\fi
\begin{definition}
Let $S, V$ be weighted posets. A lens $\ell : S \Leftrightarrow V$ is a
5-tuple of functions
\begin{align*}
\ell.\get &: S \longrightarrow V\\
\ell.\pput &: S \longrightarrow V \longrightarrow S\\
\ell.\create &: V \longrightarrow S\\
\delta_S &: S \times S \longrightarrow \mathbb{R}^{\geq 0}\\
\delta_V &: V \times V \longrightarrow \mathbb{R}^{\geq 0}
\end{align*}
such that $\delta_S$ and $\delta_V$ are difference functions on $S$ and $V$
respectively, and for all $s \in S, v \in V$
\begin{align*}
\ell.\get \; (\ell.\pput \; s \; v) &= v \tag{PUTGET}\\
\ell.\get \; (\ell.\create \; v) &= v \tag{CREATEGET}\\
\ell.\pput \; s \; (\ell.\get \; s) &= s \tag{GETPUT}\\
\delta_S(\ell.\pput \; s \; v, s) &= \delta_V(v, \ell.\get \; s) \tag{ISOMETRY}
\end{align*}
\end{definition}
These laws are the same as the classical lens laws except for new ISOMETRY rule,
which says that $\ell$ preserves difference functions.

Every classical lens can be made into a lens in this new sense in various
ways: for example, we can take the poset structure on $S$
and $V$ to be given by equality, with weight function 0 and $\delta$ function 0.

\section{New Posets and Difference Functions from Old}
Let $(P, \rho)$ and $(P', \rho')$ be weighted posets.

\subsection{Product of Posets}
Define the product $P \times P'$ of $P$ and $P'$ to be the set $P \times P'$
with $\leq_{P \times P'} = \leq_P \times \leq_{P'}$. Define a weight function
$(\rho \times \rho')$ on $P \times P'$ by $(\rho \times \rho')(a, b) = \rho(a)
+ \rho(b)$.
\begin{claim}
$(P \times P', \rho \times \rho')$ is a weighted poset.
\end{claim}
\begin{claim}
If $\delta_P$ and $\delta_{P'}$ are difference functions on $P$ and $P'$
respectively, then the product $\delta_P \times \delta_{P'}$ defined by
$$(\delta \times \delta')((a_1, b_1), (a_2, b_2)) = \delta(a_1, a_2) +
\delta(b_1, b_2)$$
is a difference function on $P \times P'$.
\end{claim}
\begin{proof}
Let $P, P'$ be posets with difference functions $\delta$ and $\delta'$
respectively. Then for all $a_1, a_2 \in A$ and $b_1, b_2 \in B$,
$$
(\delta \times \delta')((a_1, b_1), (a_2, b_2)) = (\delta(a_1, a_2),
\delta(b_1, b_2))
$$
We check that this definition satisfies the axioms for difference functions.
\begin{enumerate}
  \item[(A1)]
  Since $\delta$ and $\delta'$ are difference functions, then $\delta(a_1, a_2)
  \leq \rho(a_1)$ and $\delta'(b_1, b_2) \leq \rho'(b_1)$, hence $$
  (\delta \times \delta')((a_1, b_1), (a_2, b_2)) = \delta(a_1, a_2) +
  \delta(b_1, b_2) \leq \rho(a_1) + \rho'(b_1) = (\rho \times
  \rho')(a_1, b_1)$$
  \item[(A2)]
  Assume that $(a_1, b_1) \leq (a_2, b_2)$. Since $\delta$ and
  $\delta'$ are difference functions. It follows that that $\delta(a_1, a_2) =
  0$ and $\delta'(b_1, b_2) = 0$, hence $(\delta \times \delta')((a_1, b_1),
  (a_2, b_2)) = 0$.
  \item[(A3)]
  $\text{if } a \leq a' \text{ then for all }b , \; \delta(b, a) \geq \delta(b,
  a')$

  Assume that $(a_1, b_1) \leq (a_2, b_2)$. Then since $\delta$ and $\delta'$ are
  difference functions, it follows that $\delta(a_3, b_1) \geq \delta(a_3, b_2)$
  and $\delta'(b_3, b_1) \geq \delta'(b_3, b_2)$. Consequently,
  $$(\delta \times \delta')((a_3, b_3), (a_1, b_1)) = \delta(a_3, a_1) +
  \delta(b_3, b_1) \geq \delta(a_3, a_2) + \delta(b_3, b_2) = (\delta \times
  \delta')((a_3, b_3), (a_2, b_2))$$
  This completes the proof.
\end{enumerate}
\end{proof}
Assume that $R$ and $R'$ are regular expressions with some poset structure
defined on $\mathcal{L}(R)$ and $\mathcal{L}(R')$, and let $\delta_R$ and
$\delta_{R'}$ be difference functions defined on $\mathcal{L}(R)$ and
$\mathcal{L}(R')$ respectively. Assume that $R \cdot^! R'$. Then the product
construction gives us a way of defining a difference function $\delta_{R \cdot
R'}$ on $\mathcal{L}(R \cdot R')$ since the unambiguity condition implies that
$R \cdot R'$ is (categorically) isomorphic to $R \times R'$, the categorical
product of $R$ and $R'$.

\subsection{Coproduct of Posets}
Define the coproduct $P \sqcup P'$ of $P$ and $P'$ to be the set $P \sqcup P'$
(disjoint union of $P$ and $P'$) with $x \leq_{P \sqcup P'} y$ if and only if
$x \leq_P y$ or $x \leq_{P'} y$. Define a weight function
$(\rho \sqcup \rho')$ on $P \sqcup P'$ by
$$(\rho \sqcup \rho')(a) =
\begin{cases}
\rho(a) &\text{if }a \in P\\
\rho'(a) &\text{if }b \in P
\end{cases}$$
\begin{claim}
$(P \sqcup P', \rho \sqcup \rho')$ is a weighted poset.
\end{claim}
\begin{claim}
If $\delta_P$ and $\delta_{P'}$ are difference
functions on $P$ and $P'$ respectively, then the coproduct $\delta_P \sqcup
\delta_{P'}$ of $\delta_P$ and $\delta_{P'}$ defined by $$(\delta \sqcup
\delta')(a, b) =
\begin{cases}
\delta(a, b) &\text{if }a, b \in P\\
\delta'(a, b) &\text{if }a, b \in P'\\
(\rho \sqcup \rho')(a) & \text{otherwise}
\end{cases}$$
is a difference function on $P \sqcup P'$.
\end{claim}
\begin{proof}
We check that $\delta_P \sqcup \delta_{P'}$ satisfies the axioms for difference
functions.
\begin{enumerate}
  \item[(A1)]
  If $x, y \in P$ then
  $$
  (\delta \sqcup \delta')(x, y) = \delta(x, y) \leq \rho(x) = (\rho \sqcup
  \rho')(x)
  $$
  A similar argument shows that if $x, y \in P'$, then
  $$
  (\delta \sqcup \delta')(x, y) = \delta'(x, y) \leq \rho'(x) = (\rho \sqcup
  \rho')(x)
  $$
  Otherwise $\delta \sqcup \delta'(a, b) = (\rho \sqcup \rho')(a)$.
  \item[(A2)]
  Assume that $a \leq_{P \sqcup P'} a'$. Then either $a \leq_P a'$ or $a
  \leq_{P'} a'$, and since $\delta$ and $\delta'$ are difference functions, it
  follows that $\delta(a, a') = 0$ in either case.
  \item[(A3)]
  Assume that $a \leq_{P \sqcup P'} b$. Assume first that $a \leq_P b$. If $c
  \in P$, then $\delta(c, a) \geq \delta(c, a')$ since $\delta$ is a difference
  function. If $c \in P'$, then $\delta(c, a) = \delta(c, b) = \rho(c)$. A
  similar argument shows that if $a \leq_{P'} b$, then for all $c \in P \sqcup
  P'$, $\delta(c, a) \geq \delta(c, b)$. This completes the proof.
\end{enumerate}
\end{proof}
Assume that $R$ and $R'$ are regular expressions with some poset structure
defined on $\mathcal{L}(R)$ and $\mathcal{L}(R')$, and let $\delta_R$ and
$\delta_{R'}$ be difference functions defined on $\mathcal{L}(R)$ and
$\mathcal{L}(R')$ respectively. Assume that $\mathcal{L}(R) \cap
\mathcal{L}(R') = \varnothing$. Then the coproduct construction gives us a way
of defining a difference function $\delta_{R \; | \; R'}$ on $\mathcal{L}(R \;
| \; R')$ since the unambiguity condition implies that $R \; | \; R'$ is
(categorically) isomorphic to $R \sqcup R'$, the categorical coproduct of $R$
and $R'$.
\subsection{List Posets}
Let $P$ be a poset. Define the list poset $P^*$ of $P$ to be the set
$$P^0 \sqcup P \sqcup (P \times P) \sqcup \ldots \sqcup P^i \sqcup \ldots$$
This is just the set of all finite sequences of elements on $P$, with $P^0 =
\epsilon$, the empty sequence. The poset structure on $P^*$ is the poset
structure inherited from the product and coproduct constructions.

Assume that $R$ is a regular expression with some poset structure with $R^{*!}$.
Assume that $\delta_R$ is a difference function on $\mathcal{L}(R)$. Then
the free poset construction gives us a way of defining a difference function
on $\delta_{R^*}$ on $\mathcal{L}(R^*)$ since the unambiguity condition implies
that $R^*$ is (categorically) isomorphic to $F(R)$, the free monoid on $R$,
with $R^0 = \epsilon$, the literal empty string (i.e. the literal empty string
is sent to the categorical empty string by this isomorphism).

\subsection{A Finer Poset Structure on $P^*$}
Define a finer poset structure on $P^*$ by $x \; \leq \; y$
if and only if $x = (x_1, \ldots, x_m), \; y = (y_1, \ldots, y_n)$, $m \leq n$
and $x_i \leq y_i$ for $1 \leq i \leq m$.

Let $\delta_P$ be a difference function on $P$. We define the function
$\Delta^*$ on $P^*$ by
$$\Delta^*((x_1, \ldots, x_m), (y_1, \ldots, y_n)) = \sum_{i=1}^{\min\{m,
n\}}\delta(x_i, y_i) + \sum_{j = \min\{m, n\} + 1}^{m}\rho(x_i)$$
\begin{claim}
$\Delta^*$ is a difference function.
\end{claim}
\begin{proof}
We check that each of the axioms for difference functions holds.
\begin{enumerate}
  \item[(A1)]
  Since $\delta$ is a difference function, then $\delta(x_i, y_i)
  \leq \rho(x_i)$ for $1 \leq i \leq \min\{m, n\}$ and for $\min\{m, n\} < j <
  m$ from which it follows that
  $$\Delta^*((x_1, \ldots, x_m), (y_1,
  \ldots, y_n)) \leq \sum_{i=1}^m\rho(x_i) = \rho^*((x_1, \ldots, x_n))$$
  \item[(A2)]
  Assume that $(x_1, \ldots, x_m) \leq (y_1, \ldots, y_n)$. Then $m \leq n$,
  hence
  $$\Gamma^*((x_1, \ldots, x_m), (y_1, \ldots, y_n)) =
  \sum_{i=1}^{m}\delta(x_i, y_i) = 0,$$
  since $x_i \leq y_i$ for each $i$ and $\delta$ is a difference function.
  \item[(A3)]
  Assume that $(x_1, \ldots, x_m) \leq (y_1, \ldots, y_{m'})$. Then $m \leq
  m'$ and $x_i \leq y_i$. Also, $$\Delta^*((z_1, \ldots, z_n), (x_1,
  \ldots, x_m)) = \sum_{i=1}^{\min\{m, n\}}\delta(z_i, x_i) + \sum_{j =
  \min\{m, n\} + 1}^{n}\rho(z_i)$$ and
  $$\Delta^*((z_1, \ldots, z_n), (y_1, \ldots, y_{m'})) =
  \sum_{i=1}^{\min\{m', n\}}\delta(z_i, y_i) + \sum_{j = \min\{m', n\} +
  1}^{n}\rho(z_i)$$
  As $\delta$ is a difference function, it follows that $\delta(z_i, x_i) \geq
  \delta(z_i, y_i)$ for $1 \leq i \leq m$, hence
  $$\Delta^*((z_1, \ldots, z_n), (x_1, \ldots, x_m)) \geq \Delta^*((z_1,
  \ldots, z_n), (y_1, \ldots, y_{m'}))$$ This completes the proof.
\end{enumerate}
\end{proof}
\fi
%% Acknowledgments
\begin{acks}                            %% acks environment is optional
%% contents suppressed with 'anonymous'
%% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
%% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
%% acknowledge financial support and will be used by metadata
%% extraction tools.
This material is based upon work supported by the
\grantsponsor{GS100000001}{National Science
Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
No.~\grantnum{GS100000001}{nnnnnnn} and Grant
No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
conclusions or recommendations expressed in this material are those
of the author and do not necessarily reflect the views of the
National Science Foundation.
\end{acks}

\bibliographystyle{plain}
\bibliography{local}

%% Appendix
%%\appendix
%%\section{Appendix}

\end{document}

### 2018-02-02 papers/QuotientLenses/main.tex
edit quotient lens paper

@@
 
-{\em Lenses}~\cite{Lenses} specify bidirectional 
-transformations between different data formats.  Writing a single lens
-rather than two separate uni-directional transformations saves both
-keystrokes and bugs, by reducing redundancy and by automatically
-guaranteeing strong ``lens laws'' ensuring that the forward and backward
-transformations interact well.
-
-{\em Quotient lenses}~\cite{quotientlenses} are a generalization of lenses
-where each of the data formats is quotiented by some equivalence relation
-and the lens laws are taken modulo equivalence.  This allows lenses to be
-both simpler and more flexible, since whole equivalence classes of data
-values can be handled all at once and inessential details like whitespace
-and ordering of fields can be ignored.
+Oftentimes, different applications store the same data in custom formats. For
+example, \bibtex{} and EndNote both store bibliography data; spreadsheet data
+can be represented as a CSV and as a TSV; web apis can return data as a JSON or
+XML object. Synchronizing these applications amounts to synchronizing the data
+formats. \emph{Lenses} comprise such synchronization functions, and the ``lens
+laws'' provide restrictions on how these synchronization functions interoperate.
+Bidirectional languages like Boomerang~\cite{boomerang} and
+BiGul~\cite{DBLP:conf/pepm/KoZH16} use a single expression to encode all of the
+lens's synchronization functions at the same time.
+
+However, these lens laws are too restrictive for many important transformations.
+In particular, these lens laws do not permit transformations that put the data
+in canonical forms. In \bibtex{}, an author can be specified as \cd{lastname,
+  firstname} and \cd{firstname lastname}. While it is important to allow authors
+to be input in both ways, we would like these authors to only be output as
+\cd{firstname lastname}, but the lens laws prevent such functions.
+{\em Quotient lenses}~\cite{quotientlenses} are a generalization of lenses where
+each of the data formats is quotiented by some equivalence relation and the lens
+laws are taken modulo equivalence. This both permits canonization functions and
+allows lenses to be simpler and more flexible. Inessential details like
+whitespace and orderings of fields can be ignored when writing the
+transformations between the formats.
 
@@
 popular libraries, Boomerang~\cite{boomerang} is the only implementation
-currently offering quotient lenses.  One reason for this is that there is
-some annoying redundancy in the quotient lens notation: The programmer often
-has to separately specify (1) the data formats on each side of the lens, (2) the
-equivalence relations over them, and (3) a lens between the canonical
-representatives of the equivalence classes, even though, typically, these three
-share a great deal of structure.  Moreover, writing the lens between the
-equivalence classes can be challenging due to the rather fiddly typechecking,
-mainly due to ``non-ambiguity'' conditions in the lens well-formedness rules.
+currently offering quotient lenses. Even though quotient lenses permit simpler
+transformations between the formats, they merely push complexity elsewhere. To
+write a quotient lens, programmers must write with both standard lens
+combinators and \emph{canonizers}. Canonizers encode two functions, one that
+maps a format to a canonical version of that format, and another that chooses
+which elements of the original format correspond to values in the canonical
+format. Furthermore, if a user wants to enforce that a quotient lens maps
+between two different formats, they must write regular expressions specifying
+such formats. In this way, writing quotient lenses involves writing 3 terms: (1)
+the data formats on each side of the lens, (2) canonizers to normalize the data
+and choose representatives, and (3) a lens between the data formats when in
+canonical form, even though, typically, these three share a great deal of
+structure. Moreover, writing the lens between the equivalence classes can be
+challenging due to the rather fiddly typechecking, mainly due to
+``non-ambiguity'' conditions in the lens well-formedness rules.
 
@@
 introduce a novel language of {\em Quotient Regular Expressions}, or QREs, which
-enable a programmer to simultaneously specify a regular expression as well as an
-equivalence relation on that regular expression. QREs make programming quotient
-lenses easire by saving the programmer the effort of defining the data and the
-equivalence relation on the data separately.
+enable a programmer to simultaneously specify a regular expression while
+simultaneously specifying the canonical representatives on the regular
+expression's language. With QREs, programmers needn't define the data and
+canonization functions separately -- by adding simple annotations to the data,
+the canonization functions can be inferred.
 
 Secondly, we use QREs to define a class of quotient lenses called {\em QRE
-lenses} which map betweeen QREs, and show how to synthesize QRE lenses given
-an specification described by pair of QREs and a set of input-output examples.
-This way, the programmer need not manually derive the lens herself. The net
-effect of these two contributions, that is QREs and QRE lens synthesis, is that
-in order to derive a quotient lens, the programmer will only need to specify the
-data formats on each side of the desired quotient lens and possibly a set of
-example input-output pairs that guide synthesis of the lens, a marked
-improvement over the current situation.
+  lenses} which map betweeen QREs, and show how to synthesize QRE lenses. Using
+this tool, a lens programmer need not manually derive the lens herself, she
+merely must provide a pair of QREs (describing the formats the lens should map
+between), and a suite of examples (demonstrating how the lens should act on
+specific instances).  The resulting lens is guaranteed to map between the
+languages of the QREs and satisfy the provided examples.
 
@@
 \begin{enumerate}
-\item We introduce a novel language of {\em Quotient Regular Expressions}
-(QREs) that a provide compact, convenient notation for a natural and useful
-class of equivalence relations on regular languages (Section~\ref{QRE}).
+\item We introduce a novel language of {\em Quotient Regular Expressions} (QREs)
+  that provide a compact, convenient notation for regular languages and their
+  canonical representatives (Section~\ref{QRE}).
 \item We define a language of {\em QRE lenses}, a subset of bijective quotient
@@
 \item Using this normal form, we reduce the problem of {\em synthesizing}
-QRE lenses from their QRE types and a set of examples to the previously studied
+QRE lenses to the previously studied
 problem of synthesizing bijective lenses between regular languages 
 (Section~\ref{synth}).
-\item We extend the Boomerang {\em implementation} with QREs, QRE lens operators,
-and QRE lens synthesis and demonstrate its applicability by using it to
+\item We extend the Boomerang {\em implementation} with QREs
+and QRE lens synthesis.  We demonstrate its applicability by using it to
 synthesize QRE lenses between several real-world data formats from the
@@
 and show, by way of example, how our novel language of {\em Quotient Regular
-Expressions}, or QREs and QRE lenses address these difficulties.
+Expressions}, or QREs, and QRE lenses address these difficulties.
 
@@
 
-We wish for this transformation to have the same behaviour on records that
-differ only in ``inessential'' details. First, the transformation
-should allow whitespace characters to occur arbitrarily between the words in the
-label, author and title fields as well, as in between them. Second, an
-author's name can have the surname followed by a comma and then the other
-names separated by arbitrary whitespace, or just all the names separated by
-arbitrary whitespace (for example, the names ``Conway, J. H.'' and ``J. H.
-Conway'' are equivalent). Lastly, the transformation should also ignore the
-order in which the author and title fields occur in each citation record.
+If two \bibtex{} files differ in only ``inessential'' details, we would like
+these files to map to the same EndNote representations.
+First, the transformation should allow whitespace characters to occur
+arbitrarily between the words in the label, author, and title fields, as well as
+in between them. Second, an author's name can have the surname followed by a
+comma and then the other names separated by arbitrary whitespace, or just all
+the names separated by arbitrary whitespace (for example, the names ``Conway, J.
+H.'' and ``J. H. Conway'' are equivalent). Lastly, the transformation should
+also ignore the order in which the author and title fields occur in each
+citation record.
 
@@
 While it is possible to express the equivalence on the \bibtex{} data above in
-Boomerang, actually doing so is a challenge. This is because the entire
-equivalence is composed of three separate equivalences---whitespace quotienting
-inside fields, whitespace quotenting in between fields, and permutation
-quotenting of fields---nest with each other in a way that makes them difficult
-to express with the standard Boomerang combinators.
-
-In general, there is also another issue in describing equivalences in
-Boomerang, which is that that the programmer often has to first specify the data
-itself in the form of a regular expression before she can define the
-equivalence relation on that data. This leads to verbose code.
-
-To address these issues, we propose {\em Quotient Regular Expressions} or QREs
-which enable the programmer to define a regular expression $R$ as well as an
-equivalence relation on $R$ in the same expression. For example, with QREs we
-can specify \bibtex{} records as shown in Figure~\ref{fig:example-qre}.
+Boomerang, actually doing so is a challenge.  Three of the inessential details
+we wish to ignore---whitespace quotienting
+inside fields, whitespace quotienting in between fields, and permutation
+quotenting of fields---nest within each other in a way that makes them difficult
+to express with the existing Boomerang combinators.
+
+Furthermore, programming best practices require programmers provide type
+annotations to their lenses to make explicit the formats the quotient lenses map
+between. This practice prevents nefarious bugs that can result from the lenses
+mapping between incorrect formats, but it is also verbose and time consuming.
+
+To address these issues, we propose {\em Quotient Regular Expressions}, or QREs,
+which enable the programmer to define a regular expression $R$ and its canonical
+elements in the same expression. For example, with QREs we can specify \bibtex{}
+records and their canonical elements, as shown in Figure~\ref{fig:example-qre}.
 
@@
 \begin{lstlisting}
-let wsp = [ \n\t\r]
-
-(* map arbitrary occurrences of whitespace to a single space *)
-let wsp_to_space   = project wsp$^+$ $\to$ " "
+(* whitespace with " " as a canonical representative *)
+let wsp = project [ \n\t\r]$^+$ $\to$ " "
 
 (* define name representations with a space and with a comma *)
-let space_name     = name | (name . wsp_to_space . (name . wsp_to_space)$^*$ . name)
-let comma_name   = name | (name . "," . wsp_to_space . (name . wsp_to_space)$^*$ . name)
+let space_name     = name | (name . wsp . (name . wsp)$^*$ . name)
+let comma_name   = name | (name . "," . wsp . (name . wsp)$^*$ . name)
 
@@
 (* squash QRE maps comma_name to space_name *)
-let valid_name = squash comma_name $\to$ space_name using (get space_to_comma)
+let name = squash comma_name $\to$ space_name using (get space_to_comma)
 
 (* define rest of bibtex fields *)
-let wsp_to_and = project (wsp$^+$ . "and" . wsp$^+$) $\to$ " and " 
-let bib_names  = valid_name | (valid_name . wsp_to_and . (valid_name . wsp_to_and)$^*$ . valid_name) 
+let name_sep = wsp . "and" . wsp
+let bib_names  = name | (name . name_sep . (name . name_sep)$^*$ . name) 
 let bib_author = "author = \"" . bib_names . "\""
-let title     = word | (word . wsp_to_space . (word . wsp_to_space)$^*$ . word) 
+let title     = word | (word . wsp . (word . wsp)$^*$ . word) 
 let bib_title = "title = {" . title . "}"
@@
 (* allow any permutation of fields interspersed with arbitrary whitespace *)
-let bib_perm  = perm (label, bib_author, bib_title) with (project ("," .wsp$^+$) $\to$ ",\n")
-let bibtex    = "@book{" . bib_perm . "}"
+let bib_fields  = perm (label, bib_author, bib_title) with (project ("," .wsp$^+$) $\to$ ",\n")
+let bibtex    = "@book{" . bib_fields . "}"
 \end{lstlisting}
@@
 
-In Figure~\ref{fig:example-qre}, the QRE \cd{project wsp}+ $\to$ \cd{" "}
-matches any non-zero number of whitespace characters, but considers the matched
-string of whitespace characters as equivalent with a single space as the
-canonical representative. We call this QRE a \textit{projection} because it
-projects an entire regular language down to a single representative string in
-the language.
+In Figure~\ref{fig:example-qre}, the QRE \verb+project [ \n\t\r]+$^{+} \to$
+\verb+" "+ matches any non-zero number of whitespace characters, with a single
+space as the canonical representative. We call this QRE a \textit{projection}
+because an entire regular language is projected to a single representative
+string during canonization.
 
@@
 
-The third example of a QRE in the Figure is \cd{bib_perm}, defined by:
+The third example of a QRE in the Figure is \cd{bib_fields}, defined by:
 \begin{center}
@@
 is that in which the label, author and title fields occur in that order,
-separated by a comma followed a newline. The \textit{permutation} QREs is
+separated by a comma followed by a newline. The \textit{permutation} QRE is
 designed for equivalences such as this, where different permutations of
@@
 insight is to reduce the problem of synthesizing QRE lenses to the problem of
-synthesizing {\em bijective lenses}, which are the bijections created from the
-Boomerang lens combinators. This is a problem which we solved in our prior work
-~\cite{popl18}.
+synthesizing {\em bijective lenses}~\cite{optician}, which are bijections created
+from the Boomerang lens combinators.
 
@@
 This section introduces Quotient Regular Expressions or QREs. QREs are regular
-expressions augmented with syntax that lets them simultaneously express an
-equivalence relation on the language described by the regular expression.
+expressions augmented with syntax that lets them simultaneously express a
+language and how to convert from that language into its canonical elements.
 
@@
 \perm{c_1, \ldots, c_n}{c} \\
-  & | \; \normalize{R}{R'}{f} \sep c \semicolon c' \sep c \cdot c' \sep (c \sep
+  & | \; \normalize{R}{R'}{f} \sep c \semicolon c' \sep c \cdot c''' \sep (c \sep
   c') \sep c^*,
@@
 (* Regular expressions matching different comma and space representations *)
-let comma_name = last_name . "," . wsp+ . first_name
-let space_name = first_name . wsp+ . last_name
+let comma_name = last_name . "," . wsp . first_name
+let space_name = first_name . wsp . last_name
 
@@
 \AxiomC{$\ell : R \Leftrightarrow S$}
-\AxiomC{$R \equiv^d R'$}
-\AxiomC{$S \equiv^d S'$}
+\AxiomC{$R \equiv^s R'$}
+\AxiomC{$S \equiv^s S'$}
 \TrinaryInfC{$\ell : R' \Leftrightarrow S'$}
@@
 
-In our previous work~\cite{popl18}, we showed that given regular expressions
+Optician~\cite{optician} showed that given regular expressions
 $R, S$ and a set of input-output pairs $\{(r_1, s_1), \ldots, (r_n, s_n)\}$, if
@@
   \item
-  compute $\{({x_1}', {y_1}'), \ldots, ({x_n}', {y_n}')\}$ where ${x_i}' =
+  Compute $\{({x_1}', {y_1}'), \ldots, ({x_n}', {y_n}')\}$ where ${x_i}' =
   \canonizer(c)(x_i)$ and ${y_i}' = \canonizer(c')(y_i)$,
   \item
-    synthesize a bijective lens $\ell : K(c) \Leftrightarrow K(c')$ that is
+    Synthesize a bijective lens $\ell : K(c) \Leftrightarrow K(c')$ that is
     consistent with the examples $\{({x_1}', {y_1}'), \ldots, ({x_n}',
@@
   \item 
-  return $q = \kw{rquot}(\kw{lquot}(c, \ell), c') : c \Leftrightarrow
+  Return $q = \kw{rquot}(\kw{lquot}(c, \ell), c') : c \Leftrightarrow
   c'$
@@
 transformations~\cite{singh2012learning,le-pldi-2014,gulwani-popl-2014,perelman2014test,Singh:blinkfill},
-the system Optician, which we introduced in ~\cite{popl18}, is the first attempt
+the system Optician, which we introduced in ~\cite{optician}, is the first attempt
 to synthesize bidirectional transformations that we are aware of. In that
@@
 \bibliographystyle{plain}
-\bibliography{local}
+\bibliography{local,bcp}
 

### 2018-02-03 papers/DifferenceLenses/differencelenses.tex
Finished difference and lens proofs

--- 2018-01-31 papers/DifferenceLenses/document.tex
@@
 
-% Macros
-% Colors
+% Macros Colors
 \definecolor{dkblue}{rgb}{0,0.1,0.5}
@@
 % % Author with single affiliation.
-%\author{First1 Last1}
-%\authornote{with author1 note}          %% \authornote is optional;
-% % can be repeated if necessary
-%\orcid{nnnn-nnnn-nnnn-nnnn}             %% \orcid is optional \affiliation{
-%\position{Position1}
-%\department{Department1}              %% \department is recommended
-%\institution{Institution1}            %% \institution is required
-%\streetaddress{Street1 Address1}
-%\city{City1}
-%\state{State1}
-%\postcode{Post-Code1}
-%\country{Country1}                    %% \country is recommended }
-%\email{first1.last1@inst1.edu}          %% \email is recommended
+% \author{First1 Last1} \authornote{with author1 note}          %% \authornote
+% is optional; % can be repeated if necessary \orcid{nnnn-nnnn-nnnn-nnnn}       
+%      %% \orcid is optional \affiliation{ \position{Position1}
+% \department{Department1}              %% \department is recommended
+% \institution{Institution1}            %% \institution is required
+% \streetaddress{Street1 Address1} \city{City1} \state{State1}
+% \postcode{Post-Code1} \country{Country1}                    %% \country is
+% recommended } \email{first1.last1@inst1.edu}          %% \email is recommended
 
@@
 
-One of these laws requires that the \pput component if a lens returns the
+One of these laws requires that the \pput component of a lens returns the
 original source unmodified whenever the update to the view is a no-op.
-Unfortunately, this (and other) laws do not specify the behaviour of the \pput
+Unfortunately, this law does not specify the behaviour of the \pput
 component when the update to the view is non-trivial. Consequently, a small
@@
 made to the view is non-trivial. To do this, we will first interprete sources
-and views as {\em difference posets} which are posets where we can compute the
-difference between any two elements of the poset. With this interpretation, we
-will then express the new law which will require that the \pput component of a
-lens preserve differences. For example, we will be able to invalidate
-positional lenses defined on order agnostic data on the grounds that the \pput
-component of positional lenses do not respect order agnostic differences.
+and views as {\em difference posets} which are posets where we can measure the
+difference between any two elements of the poset. We will then use this
+difference structure on sources and views to express the new lens law which will
+require that the \pput component of a lens respect differences. For example, we
+will be able to invalidate positional lenses defined on order agnostic data on
+the grounds that the \pput component of a positional lenses does not respect
+order agnostic differences.
 
@@
 \begin{definition}
-A {\em difference poset (DP)} is a triple $(P, \rho, \delta)$ where $P$ is
-a poset and $\rho : P \longrightarrow \mathbb{R}^{\geq 0}$ and $\delta : P
-\times P \longrightarrow P$ are functions satisfying :
+A {\em difference poset (DP)} is a triple $(P, \rho, \delta)$ where $(P, \rho)$
+is a graded poset $\delta : P \times P \longrightarrow \mathbb{N}$ is a function
+satisfying :
 \begin{align}
-&\delta(a, b) \leq a\tag{A1}\\
-&\text{if } a < a',\text{ then } \rho(a) < \rho(a') \tag{A2}\\
-&\text{if } a \leq a' \text{ then for all }b , \; \rho(\delta(b, a)) \geq
-\rho(\delta(b, a'))\tag{A3}
+&\delta(a, b) \leq \rho(a)\tag{A1}\\
+&\text{if } a \leq a',\text{ then } \delta(a, b) = 0\tag{A2}\\
+&\text{if } a \leq a' \text{ then for all }b , \; \delta(b, a) \geq
+\delta(b, a')\tag{A3}
 \end{align}
@@
 \begin{example}\label{embedding}
-Let $P$ be any set with equality giving the poset structure. Given any
-function $\rho: P \longrightarrow \mathbb{R}^{\geq 0}$ on $P$, and function
-$\delta$ defined by $\delta(a, b) = a$, then $(P, \rho, \delta)$ is a difference
-poset.
+Let $P$ be any set with equality giving the poset structure. Define
+$\rho(a) = 1$ for all $a$ and $\delta(a, b) = \delta_{a, b}$ where
+$$\delta_{a, b} =
+\begin{cases}
+0 & \text{if } a = b\\
+1 & \text{otherwise}
+\end{cases}$$
 \end{example}
+Then $(P, \rho, \delta)$ is a difference poset.
 \begin{example}\label{naturalnumbers}
@@
 relation and weight function $\rho(p_1^{a_1} \cdot \ldots \cdot p_n^{a_n}) =
-a_1 + \ldots + a_n$. Define $\delta$ by $\delta(a, b) = a/gcd(a, b)$. Then $(P,
-\rho, \delta)$ is a difference poset.
+a_1 + \ldots + a_n$. Define $\delta$ by $\delta(a, b) = \rho(a/gcd(a, b))$. Then
+$(P, \rho, \delta)$ is a difference poset.
 \end{example}
@@
 containment relation and weight function $\rho(A) = |A|$. Define $\delta$
-by $\delta(A, B) = A \setminus B$. Then $(P, \rho, \delta)$ is a difference
-poset.
+by $\delta(A, B) = \rho(A \setminus B)$. Then $(P, \rho, \delta)$ is a
+difference poset.
 \end{example}
@@
 lattice and $\rho$ a strict order preserving map from $P$ to the non-negative
-real numbers. Define $$\delta(a, b) = \rho\left(\bigwedge \{x \; | \; (a \wedge
-b) \vee x = a\}\right)$$
-Then $(P, \rho, \delta)$ is a difference poset.
+real numbers such that $\rho(a) = 0$ if and only if $a = \bigwedge P$. Define
+$$\delta(a, b) = \rho\left(\bigwedge \{x \; | \; (a \wedge b) \vee x =
+a\}\right)$$ Then $(P, \rho, \delta)$ is a difference poset.
 \end{example}
@@
   \item[(A2)]
-  Follows from the fact that $\rho$ is a strict order preserving map from $P$ to
-  the non-negative real numbers.
+  If $a \leq b$, then $(a \wedge b) = a$ hence $\bigwedge X = \bigwedge P$.
+  Consequently, $\delta(a, a') = \left(\bigwedge X\right) =
+  \left(\bigwedge P \right) = 0$
   \item[(A3)]
@@
   \in Y'$. Consequently, $\bigwedge Y \geq \bigwedge Y'$ from which it follows
-  that $$\rho(\delta(b, a)) =\rho\left(\bigwedge Y\right) \geq
-  \rho\left(\bigwedge Y'\right) = \rho(\delta(b, a'))$$
+  that $$\delta(b, a) =\rho\left(\bigwedge Y\right) \geq
+  \rho\left(\bigwedge Y'\right) = \delta(b, a')$$
   This completes the proof.
@@
 \end{proof}
-\iffalse
-Now define a
-function $d:
-P \longrightarrow \mathbb{R}$ on $P$ by $$d(a, b) =
-\begin{cases}
-0 & \text{if } a = b\\
-\rho(\pi_1(\delta(a, b))) + \rho(\pi_2(\delta(a, b))) & \text{otherwise}
-\end{cases}$$
-\begin{fact}
-\begin{enumerate}
-  \item
-  $d(a,b) \geq 0$,
-  \item
-  $d(a, b) = 0$ if and only if $a = b$, and
-  \item
-  $d(a, b) = d(b, a)$
-\end{enumerate}
-\end{fact}
-\begin{fact}
-If $b \leq c$, then for all $b$, $d(a, b) \geq d(a, c)$. Thus if $b \leq c$,
-$d(a, c) \leq d(a, b) + d(b, c)$
-\end{fact}
-\begin{question}
-Under what conditions is $d$ a metric?
-\end{question}
-\fi
+
 \section{New Difference Posets from Old}
@@
 + \rho(b)$. Define the difference function $\delta_P \times \delta_{P'}$ on $P
-\times P$ by $$(\delta \times \delta')((a_1, b_1), (a_2, b_2)) = (\delta(a_1,
-a_2), \delta(b_1, b_2))$$
+\times P$ by $$(\delta \times \delta')((a_1, b_1), (a_2, b_2)) = \delta(a_1,
+a_2) + \delta(b_1, b_2)$$
 \begin{claim}
@@
   Since $\delta$ and $\delta'$ are difference functions, then $\delta(a_1, a_2)
-  \leq a_1$ and $\delta'(b_1, b_2) \leq b_1$, hence $$
-  (\delta \times \delta')((a_1, b_1), (a_2, b_2)) = (\delta(a_1, a_2),
-  \delta(b_1, b_2)) \leq (a_1, b_1)$$
+  \leq \rho(a_1)$ and $\delta'(b_1, b_2) \leq \rho(b_1)$, hence $$
+  (\delta \times \delta')((a_1, b_1), (a_2, b_2)) = \delta(a_1, a_2) +
+  \delta(b_1, b_2) \leq \rho(a_1) + \rho(b_1) = \rho(a_1, b_1)$$
   \item[(A2)]
-  Assume that $(a_1, b_1) < (a_2, b_2)$. Then since $\rho$ and $\rho'$ are
-  weight functions, it follows that $\rho(a_1) < \rho(a_2)$ and
-  $\rho'(b_1) < \rho'(b_2)$. Consequently,
-  $$(\rho \times \rho')(a_1, b_1) = \rho(a_1) + \rho(b_1) < \rho(a_2) +
-  \rho(b_2) = (\rho \times \rho)(a_2, b_2)$$
+  Assume that $(a_1, b_1) \leq (a_2, b_2)$. Then $a_1 \leq a_2$ and $b_1 \leq
+  b_2$. Since $\delta$ is a difference function, $\delta(a_1, a_2) = 0$ and
+  $\delta(b_1, b_2) = 0$, hence
+  $$\delta((a_1, b_1), (a_2, b_2)) = \delta(a_1, a_2) + \delta(b_1, b_2) = 0$$
   \item[(A3)]
-  Assume that $(a_1, b_1) \leq (a_2, b_2)$. Then since $\delta$ and $\delta'$ are
-  difference functions, it follows that $\rho(\delta(a_3, b_1))
-  \geq \rho(\delta(a_3, b_2))$ and $\rho'(\delta'(b_3, b_1)) \geq
-  \rho'(\delta'(b_3, b_2))$.
+  Assume that $(a_1, b_1) \leq (a_2, b_2)$. Then since $\delta$ and $\delta'$
+  are difference functions, it follows that $\delta(a_3, b_1)
+  \geq \delta(a_3, b_2)$ and $\delta'(b_3, b_1) \geq \delta'(b_3, b_2)$.
   Consequently,
   \begin{align*}
-((\rho \times \rho') \circ (\delta \times \delta'))((a_3, b_3), (a_1, b_1)))
-&= (\rho \times \rho')(\delta(a_3, a_1), \delta'(b_3, b_1))\\
-&= \rho(\delta(a_3, a_1)) + \rho'(\delta(b_3, b_1))\\
-&\geq \rho(\delta(a_3, a_2)) + \rho'(\delta(b_3, b_2))\\
-&= (\rho \times \rho')(\delta(a_3, a_2), \delta'(b_3, b_2))\\
-&= ((\rho \times \rho') \circ (\delta \times \delta'))((a_3, b_3), (a_2,
-b_2))\end{align*}
+(\delta \times \delta')((a_3, b_3), (a_1, b_1)))
+&=\delta(a_3, a_1) + \delta'(b_3, b_1)\\
+&\geq \delta(a_3, a_2) + \delta(b_3, b_2)\\
+&= \delta(a_3, a_2) + \delta'(b_3, b_2)\\
+&= (\delta \times \delta')((a_3, b_3), (a_2, b_2))
+\end{align*}
 
@@
 \delta'(a, b) &\text{if }a, b \in P'\\
-a & \text{otherwise}
+\rho(a) & \text{otherwise}
 \end{cases}$$
@@
   \item[(A1)]
-  If $x, y \in P$ then $(\delta \sqcup \delta')(x, y) = \delta(x, y) \leq x$.
-  A similar argument shows the same result if $x, y \in P'$. Otherwise $\delta
-  \sqcup \delta'(a, b) = a$.
+  If $x, y \in P$ then $(\delta \sqcup \delta')(x, y) = \delta(x, y) \leq
+  \rho(x) = (\rho \sqcup \rho')(y)$. A similar argument shows the same result if
+  $x, y \in P'$. Otherwise $\delta \sqcup \delta'(a, b) = \rho(a) = (\rho \sqcup
+  \rho')(y)$.
   \item[(A2)]
-  Assume that $a \leq_{P \sqcup P'} b$. If $a \leq_P b$,
-  $$(\rho \sqcup \rho')(a) = \rho(a) < \rho(b) = (\rho \sqcup \rho')(b)$$
-  If $a \leq_{P'} b$, then
-  $$(\rho \sqcup \rho')(a) = \rho'(a) < \rho'(b) = (\rho \sqcup \rho')(b)$$
+  Assume that $a \leq_{P \sqcup P'} b$. If $a \leq_P b$, then $\delta(a, b) =
+  \delta_P(a, b) = 0$ and if $a \leq_{P'} b$, then $\delta(a, b) =
+  \delta_{P'}(a, b) = 0$ since $\delta_P$ and $\delta_{P'}$ are difference
+  functions.
   \item[(A3)]
   Assume that $a \leq_{P \sqcup P'} b$. Assume first that $a \leq_P b$. If $c
-  \in P$, then $\rho(\delta(c, a)) \geq \rho(\delta(c, a'))$ since $\delta$ is a
-  difference function. If $c \in P'$, then $\delta(c, a) = \delta(c, b) = c$,
-  hence $\rho(\delta(c, a)) \geq \rho(\delta(c, a))$. A similar argument shows
-  that if $a \leq_{P'} b$, then for all $c \in P \sqcup P'$, $\rho(\delta(c, a))
-  \geq \rho(\delta(c, a'))$. This completes the proof.
+  \in P$, then $\delta(c, a) \geq \delta(c, b)$ since $\delta_P$ is a
+  difference function. If $c \in P'$, then $\delta(c, a) = \delta(c, b) =
+  \rho(c)$. A similar argument shows that if $a \leq_{P'} b$, then for all $c
+  \in P \sqcup P'$, $\delta(c, a) \geq \delta(c, a')$. This completes the
+  proof.
 \end{enumerate}
@@
 \delta^i(a, b) &\text{if }a, b \in P^i\\
-a & \text{otherwise}
+\rho(a) & \text{otherwise}
 \end{cases}$$
@@
 integers and let $k = \min\{m, n\}$. We define the difference function
-$\Delta^*$ on $P^*$ by 
-$$\Delta^*((x_1, \ldots, x_m), (y_1, \ldots, y_n)) = ((\delta(x_1, y_1), \ldots,
-\delta(x_k, y_k), x_{k+1}, \ldots, x_m)$$
-\begin{claim}
-$(P^*, \rho^*, \Delta^*)$ is a difference poset.
+$\delta^*$ on $P^*$ by
+$$\delta^*((x_1, \ldots, x_m), (y_1, \ldots, y_n)) =
+\sum_{i=1}^{\min\{m, n\}} \delta(x_i, y_i) + \sum_{j > \min\{m, n\}}^n \rho(x_j)
+$$
+\begin{claim}
+$(P^*, \rho^*, \delta^*)$ is a difference poset.
 \end{claim}
@@
   Since $\delta$ is a difference function, $\delta(x_i, y_i)
-  \leq x_i$ for $1 \leq i \leq \min\{m, n\}$, from which it follows that
-  $\Delta^*((x_1, \ldots, x_m), (y_1, \ldots, y_n)) \leq (x_1, \ldots, x_m)$
+  \leq \rho(x_i)$ for $1 \leq i \leq \min\{m, n\}$, from which it follows that
+  \begin{align*}
+\delta^*((x_1, \ldots, x_m), (y_1, \ldots, y_n)) &=
+\sum_{i=1}^{\min\{m, n\}} \delta(x_i, y_i) + \sum_{j > \min\{m, n\}}^n
+\rho(x_j)\\
+&\leq  \sum_{i=1}^{\min\{m, n\}} \rho(x_i) + \sum_{j > \min\{m, n\}}^n
+\rho(x_j)\\
+&= \rho(x_1, \ldots, x_n)
+\end{align*}
 \item[(A2)]
-  Assume that $(x_1, \ldots, x_m) < (y_1, \ldots, y_n)$. Then $m \leq
-  n$ and $\rho(x_k) < \rho(y_k)$ for some $1 \leq k \leq m$ and $\rho(x_i)
-  \leq \rho(y_i)$ for all $1 \leq i \leq m$, from which it follows that
-  $$\rho^*(x_1, \ldots, x_m) = \sum_{i=1}^m \rho(x_i) < \sum_{i=1}^{n}\rho(y_i)
-  = \rho^*(y_1, \ldots, y_n)$$
+Assume that $(x_1, \ldots, x_m) \leq (y_1, \ldots, y_n)$. Then $m \leq
+n$ and since $\delta$ is a difference function, then $\delta(x_i, y_i) = 0$
+for $1 \leq i \leq m$, from which it follows that
+\begin{align*}
+\delta^*((x_1, \ldots, x_m), (y_1, \ldots, y_n)) = \sum_{i=1}^m \delta(x_i,
+y_i) = \sum_{i=1}^m 0 = 0
+\end{align*}
 \item[(A3)]
@@
 m'$ and $x_i \leq y_i$ for $1 \leq i \leq m$. Let $k = \min\{m, n\}$ and $k' =
-  \min\{m', n\}$. By the definition of $\Delta^*$,
-  \begin{align*}
-\Delta^*((z_1, \ldots, z_n), (x_1, \ldots, x_m)) &= (\delta(z_1, x_1),
-\delta(z_k, x_k), z_{k+1}, \ldots, z_n)\\
-\Delta^*((z_1, \ldots, z_n), (y_1, \ldots, y_{m'})) &= (\delta(z_1, y_1),
-\delta(z_k, y_{k'}), z_{k'+1}, \ldots, z_n)
+\min\{m', n\}$. By the definition of $\delta_{NC}^*$,
+\begin{align*}
+\delta^*((z_1, \ldots, z_n), (x_1, \ldots, x_m)) &= \sum_{i=1}^{k} \delta(x_i,
+y_i) + \sum_{j > \min\{m, n\}}^n \rho(x_j)\\
+\delta^*((z_1, \ldots, z_n), (y_1, \ldots, y_{m'})) &= \sum_{i=1}^{k'}
+\delta(x_i, y_i) + \sum_{j > \min\{m, n\}}^n \rho(x_j)
 \end{align*}
@@
 \delta(z_i, y_i)$ for $1 \leq i \leq k$, hence
-$$\Delta^*((z_1, \ldots, z_n), (x_1, \ldots, x_m)) \geq \Delta^*((z_1,
+$$\delta^*((z_1, \ldots, z_n), (x_1, \ldots, x_m)) \geq \delta^*((z_1,
 \ldots, z_n), (y_1, \ldots, y_{m'}))$$
@@
 Now let $x = (x_1, \ldots, x_m), \; y = (y_1, \ldots, y_n)$. We define the
-function $\Gamma^*$ on $P^*$ by taking the difference of $x_i$ and
+function $\delta_{NC}^*$ on $P^*$ by taking the difference of $x_i$ and
 $y_{\varphi(i)}$ for some ``maximal'' increasing injective function $\varphi :
-A \longrightarrow [n]$ where $A$ is a subset of $[m]$, except that we eliminate
-all $x_i$ where $x_i = y_{\varphi(i)}$.
-
+A \longrightarrow [n]$ where $A$ is a subset of $[m]$ that minimizes $\sum_{i \in
+A }\delta(x_i, y_{\varphi(i)}) + \sum_{j \not \in A}\rho(x_j)$. If there are
+multiple such $\varphi$, we choose a canonical one.
+
+For a maximal $\varphi$, we define $\delta^*_{NC}$ by
+$$\delta^*_{NC}((x_1, \ldots, x_m), (y_1, \ldots, y_n)) = \sum_{i \in
+A }\delta(x_i, y_{\varphi(i)}) + \sum_{j \not \in A}\rho(x_j)$$
 For example, assume that $P$ is the set consisting of the letters of the
 alphabet where the poset structure is given by equality, with $\rho(a) = 1$
-and $\delta(a, b) = a$ for all $a, b \in P$. Then $\Gamma^*((a, c, d, f, b, e),
-(a, b, d, e)) = (c, f, b)$ where we eliminate $a$, $d$ and $e$ from the first
-sequence. However, even though $b$ occurs in both sequences, we are not allowed
-to eliminate both $b$ and $d$ since these two characters occur in the order $d,
-b$ in the first sequence and in the order $b, d$ in the second sequence. If we
-were to eliminate both of them, then the matching would have to have a
-crossing, thus contradicting the monotonicity of $\varphi$.
-
-Given any increasing injective $\varphi : A \longrightarrow [n]$ for some
-subset $A$ of $[m]$, define $\Gamma^*_{\varphi}(x, y)$ to be the difference
-between $x$ and $y$ computed using $\varphi$. Then the maximal $\varphi$ is the
-$\varphi$ that minimizes $\rho^*(\Gamma^*_{\varphi}(x, y))$. If there are
-multiple such $\varphi$, we choose the one that maximizes $|A|$. If there are
-still multiple such $\varphi$, we choose the one for which $A$ is
-lexicographically minimal.
+and $\delta(a, b) = \begin{cases}0 & \text{if } a = b\\ 1 & 
+\text{otherwise}\end{cases}$ for all $a, b \in P$. Then $\delta_{NC}^*((a, c,
+d, f, b, e), (a, b, d, e)) = 3$ where we match $a$, $d$ and $e$ between the two
+sequences thus leave $c, b$ and $f$ intact. However, even though $b$ occurs in
+both sequences, we are not allowed to match $b$ and $d$ since these two
+characters occur in the order $d, b$ in the first sequence and in the order $b,
+d$ in the second sequence. If we were to eliminate both of them, then the
+matching would have to have a crossing, thus contradicting the monotonicity of
+$\varphi$.
 \begin{claim}\label{nocrossingproof}
-$(P^*, \rho^*, \Gamma^*)$ is a difference poset.
+$(P^*, \rho^*, \delta_{NC}^*)$ is a difference poset.
 \end{claim}
@@
   \item[(A1)]
-  Since each element of $\Gamma^*(x, y)$ is either equal to $\Gamma(x_i,
-  y_{\varphi(i)})$ for some $\varphi$, or to $x_j$ for some $j$, and since these
-  elements occur in increasing order, it follows that $\Gamma^*(x, y) \leq x$.
+  Since each term in $\delta_{NC}^*(x, y)$ is either equal to $\delta(x_i,
+  y_{\varphi(i)})$ for some $\varphi$, or to $\rho(x_j)$ for some $j$, it
+  follows that $\delta_{NC}^*(x, y) \leq \rho(x)$.
   \item[(A2)]
-  Assume that $(x_1, \ldots, x_m) < (y_1, \ldots, y_n)$. Then $m \leq
+  Assume that $(x_1, \ldots, x_m) \leq (y_1, \ldots, y_n)$. Then $m \leq
   n$ and there exists an increasing injective function $\varphi:[m]
-  \longrightarrow [n]$ such that $\rho(x_k) < \rho(y_{\varphi(k)})$ for some $1
-  \leq k \leq m$ and $\rho(x_i) \leq \rho(y_{\varphi(i)})$ for all $1 \leq i
-  \leq m$. It thus follows that $$\rho^*(x_1, \ldots, x_m) = \sum_{i=1}^m
-  \rho(x_i) < \sum_{i=1}^{n}\rho(y_i) = \rho^*(y_1, \ldots, y_n)$$
+  \longrightarrow [n]$ such that $x_k \leq y_{\varphi(k)}$ for all $1
+  \leq k \leq m$. Consequently, $\delta^*_{NC, \varphi}((x_1, \ldots, x_m),
+  (y_1, \ldots, y_m)) = 0$, hence $\delta^*_{NC}((x_1, \ldots, x_m), (y_1,
+  \ldots, y_m)) = 0$
   \item[(A3)]
   Assume that $(x_1, \ldots, x_m) \leq (y_1, \ldots, y_n)$. Assume that
-  $\Gamma^*(z, x) = \Gamma^*_{\varphi}(z, x)$ for some $\varphi$. Then
-  $\rho^*(\Gamma^*_{\varphi} (z, x)) \geq \rho^*(\Gamma^*_{\varphi}(z, y))$
-  since $\Gamma$ is a difference function by assumption. Consequently
-  $\rho^*(\Gamma^*_{\varphi} (z, x)) \geq \rho^*(\Gamma^*_{\varphi}(z, y)) \geq
-  \rho^*(\Gamma^*(z, y))$ since $\Gamma^*(z, y)$ is computed using a $\psi$ that
-  minimizes the quantity $\rho^*(\Gamma^*_{\psi}(z, y))$. This completes the
-  proof.
+  $\delta_{NC}^*(z, x) = \delta_{NC, \varphi}^*(z, x)$ for some
+  $\varphi$.
+  Then $\delta_{NC, \varphi}^* (z, x) \geq \delta_{NC, \varphi}^*(z, y)$ since
+  $\delta$ is a differene function. Consequently $\delta_{NC, \varphi}^* (z, x)
+  \geq \delta_{NC, \varphi}^*(z, y) \geq \delta_{NC, \varphi}^*(z, y)$ since
+  $\delta_{NC}^*(z, y)$ is computed using a $\psi$ that minimizes the quantity
+  $\delta_{NC, \psi}^*(z, y)$. This completes the proof.
 \end{enumerate}
@@
 x_{\psi(n)})$ for all permutations $\varphi$ and $\psi$. In other words we take
-the poset structure $\leq_{NC}$ (NC for non-crossing) described in
-\cref{nocrossing} and quotient $P^*$ by the equivalence relation which
-identifies $(x_{\varphi(1)}, \ldots, x_{\varphi(n)})$ and $(x_{\psi(1)},
-\ldots, x_{\psi(n)})$ for all permutations $\varphi$ and $\psi$
+the poset structure $\leq_{NC}$ described in \cref{nocrossing} and take
+the quotient of $P^*$ by the equivalence relation which identifies
+$(x_{\varphi(1)}, \ldots, x_{\varphi(n)})$ and $(x_{\psi(1)}, \ldots,
+x_{\psi(n)})$ for all permutations $\varphi$ and $\psi$
 
@@
 
-We define the difference function $\Gamma^*/{\sim}$ on $P/{\sim}$ by taking
-the matching which minimizes $\rho^*$ on the difference as in the non-crossing
-case. If there are multiple matchings we just pick one! As in the non-crossing
-case, we define $\Gamma^*/{\sim}$ on $P^*/{\sim}$ by eliminating all $x_i$ from
-$x$ where $x_i = y_{\varphi(i)}$ and $\varphi$ is the canonical maximal
-matching. Going back to our previous example, then $ \Gamma^*([(a, c, d, f, b,
-e)], [(a, b, d, e)]) = [(c, f)]$ where we now can eliminate $a, b, d$ and $e$
-from the first sequence.
-\begin{claim}
-$\Gamma^*/{\sim}$ is a difference function.
+We define the difference function $\delta_{NC}^*/{\sim}([a], [b])$ on $P/{\sim}$
+similar to $\delta^*_{NC}$ except that we remove the restriction that matchings
+must be increasing. Going back to our previous example, then
+$\delta_{NC}^*/{\sim}([(a, c, d, f, b, e)], [(a, b, d, e)]) = 2$ where we now
+can match $a, b, d$ and $e$ in both sequences thus leaving only $c$ and $f$ in
+tact.
+\begin{claim}
+$\delta_{NC}^*/{\sim}$ is a difference function.
 \end{claim}
@@
 \ell.\pput \; s \; (\ell.\get \; s) &= s \tag{GETPUT}\\
-\delta_S(\ell.\pput \; s \; v, s) &= \ell.\pput \; s \; (\delta_V(v, \ell.\get \; s))
-\tag{PUTEXACT}\\
-\delta_S(\ell.\create \; v, s) &= \ell.\create \; (\delta_V(v,
-\ell.\get \; s))
-\tag{CREATEEXACT}
+\delta_S(\ell.\pput \; s \; v, s) &\leq \delta_V(v, \ell.\get \; s)
+\tag{PUTRESPECT}\\
+\delta_S(\ell.\create \; v, s) &\leq \delta_V(v, \ell.\get \; s)
+\tag{CREATERESPECT}
 \end{align*}
@@
 These laws are the same as the classical lens laws except for new EXACT rule,
-which says that $\ell.\pput$ and $\ell.\create$ preserve differences.
+which says that $\ell.\pput$ and $\ell.\create$ respect differences.
 
 Every classical lens can be made into a lens by taking the poset structure on
-$S$ and $V$ to be given by equality with arbitrary weight functions, and
-difference functions $\delta$ given by $\delta(a, b) = a$ as in
+$S$ and $V$ to be given by equality with weight functions $\rho$ given by
+$\rho(a) = 1$ and difference functions $\delta$ given by $\delta_{a, b}$ as in
 \cref{embedding}.
@@
 \begin{proof}
-First, we need to show that $\delta(\ccopy.\pput \; s \;v, s) = \ccopy.\pput \;
-s \; (\delta(v, \ccopy.\get \; s))$, and this follows from the fact that
-$$\delta(\ccopy.\pput \; s \;v, s) = \delta(v, s) = \delta(v, \ccopy.\get \;s)
-= \ccopy.\pput \; s \; (\delta(v, \ccopy.\get \; s))
-$$
-Next, we need to show that $\delta(\ccopy.\create \;v, s) = \ccopy.\create \;
-(\delta(v, \ccopy.\get \; s))$
+First, we need to show that $\delta(\ccopy.\pput \; s \;v, s) \leq \delta(v,
+\ccopy.\get \; s) $, and this follows from the fact that $\delta(\ccopy.\pput
+\; s \;v, s) = \delta(v, s) = \delta(v, \ccopy.\get \;s)$. Next, we need to show
+that $\delta(\ccopy.\create \;v, s) \leq \delta(v, \ccopy.\get \; s)$, and this
+follows from the fact that $\delta(\ccopy.\create \;v, s) = \delta(v, s) =
+\delta(v, \ccopy.\get \;s)$. This completes the proof.
 \end{proof}
@@
 \subsubsection{Constant}
-Let $P$ be any weighted poset with difference function $\delta(a, b) = a$.
-Let $\{*\}$ be a one element weighted poset. Let $p \in P$. Define the \const
+Let $P$ be a graded poset and define $\delta(a, b) = 0$ for all $a, b \in P$.
+Let $\{*\}$ be a one element difference poset. Let $p \in P$. Define the \const
 lens $\const : P \Leftrightarrow \{*\}$ by
@@
 \begin{proof}
-First, we need to show that $\delta(\const.\pput \; s \;v, s) = \delta(v,
-\const.\get \; s)$, and this follows from the fact that
-$$\delta(\const.\pput \; s \;v, s) =
-\delta(s, s) = s = \const.\pput \; s \; (\delta(v, v)) = \const.\pput \; s
-\; (\delta(v, \const.\get \;s)) $$
-
-Next, we need to show that $\delta(\const.\create \;v, s) =
-\const.\create \; \delta(v, \const.\get \; s)$, and this follows from the fact
-that
-$$\delta(\const.\create \;v, s) = \delta(p, s) = p = \const.\create \; \delta(v,
-\const.\get \; s)$$
+Follows from the fact that $\delta(a, b) = 0$ for all $a, b \in P$.
 \end{proof}
@@
 Let $\ell : S \longrightarrow V$ be a lens. Let $f : V \longrightarrow S$ be any
-function. Define the lens $\default(\ell) : V \Leftrightarrow S$ by
-\begin{align*}
-\default(\ell).\get \; s &= \ell.\get \; s\\
-\default(\ell).\pput \; s \; v &= \ell.\pput \; s \; v\\
-\default(\ell).\create \; v &= \ell.\pput \; (f \; v) \; v
+function satisfying $\delta(\ell.\pput \; (f \; v) \; v, s) \leq \delta(v,
+\ell.\get \; s)$ for all $s, v$. Define the lens $\default(\ell, f) :
+V \Leftrightarrow S$ by
+\begin{align*}
+\default(\ell, f).\get \; s &= \ell.\get \; s\\
+\default(\ell, f).\pput \; s \; v &= \ell.\pput \; s \; v\\
+\default(\ell, f).\create \; v &= \ell.\pput \; (f \; v) \; v
 \end{align*}
@@
 We need to show that
-$$\delta(\default.\create \; v, s) = \default.\create \; s \; (\delta(v,
-\ell.\get \; s))$$
-and this follows from the fact that 
-$$\delta(\default.\create \; v, s) = \delta(\ell.\pput \; (f \; v), s) =
-\ell.\pput \; (f \; v) \; (\delta(v, \ell.\get \; s)) = \default.\create \; s
-\; (\delta(v, \ell.\get \; s))$$
+$\delta(\default.\create \; v, s) \leq \delta(v,
+\ell.\get \; s)$ and this follows from the fact that
+$\delta(\default.\create \; v, s) = \delta(\ell.\pput \; (f \; v) \; v, s) \leq
+\delta(v, \ell.\get \; s)$
 \end{proof}
@@
 First, we need to show that
-$$\delta((\ell_1 \times \ell_2).\pput \; (s_1, s_2) \; (v_1, v_2), (s_1, s_2)) =
-(\ell_1 \times \ell_2).\pput \; (s_1, s_2) \; \delta((v_1, v_2), (\ell_1 \times
-\ell_2).\get (s_1, s_2))$$ This follows from the fact that
+$$\delta((\ell_1 \times \ell_2).\pput \; (s_1, s_2) \; (v_1, v_2), (s_1, s_2))
+\leq \delta((v_1, v_2), (\ell_1 \times \ell_2).\get (s_1, s_2))$$ This follows
+from the fact that
 \begin{align*}
@@
 \delta((\ell_1.\pput \; s_1\; v_1, \ell_2.\pput \; s_2 \; v_2), (s_1, s_2))\\
-&= (\delta (\ell_1.\pput \; s_1\; v_1, s_1), \delta (\ell_2.\pput \; s_2 \;
-v_2, s_1))\\
-&= (\ell_1.\pput \; s_1\; (\delta(v_1, \ell_1.\get \; s_1)), \ell_2.\pput
-\; s_2\; (\delta(v_2, \ell_1.\get \; s_2)))\\
-&= (\ell_1 \times \ell_2).\pput \; (s_1, s_2) \; (\delta(v_1, \ell_1.\get \;
-s_1), \delta(v_2, \ell_1.\get \; s_2))\\
-&= (\ell_1 \times \ell_2).\pput \; (s_1, s_2) \; \delta((v_1, v_2), (\ell_1.\get
-\; s_1, \ell_2.\get s_2))\\
-&= (\ell_1 \times \ell_2).\pput \; (s_1, s_2) \; \delta((v_1, v_2), (\ell_1
-\times \ell_2).\get (s_1, s_2))
-\end{align*}
-Next, we need to show that
-$$\delta((\ell_1 \times \ell_2).\create \; (v_1, v_2), (s_1, s_2))
-= (\ell_1 \times \ell_2).\create \; \delta((v_1, v_2), (\ell_1
-\times \ell_2).\get (s_1, s_2))$$
-which follows from the fact that
-\begin{align*}
-\delta((\ell_1 \times \ell_2).\create \; (v_1, v_2), (s_1, s_2)) &=
-\delta((\ell_1.\create \; v_1, \ell_2.\create \; v_2), (s_1, s_2))\\
-&= (\delta(\ell_1.\create \; v_1, s_1), \delta(\ell_2.\create \; v_2, s_2))\\
-&= (\ell_1.\create \; (\delta(v_1, \ell_1.\get \; s_1)), \ell_2.\create \;
-(\delta(v_2, \ell_2.\get \; s_2)))\\
-&= (\ell_1 \times \ell_2).\create \; ((\delta(v_1, \ell_1.\get \; s_1)),
-(\delta(v_2, \ell_2.\get \; s_2)))\\
-&= (\ell_1 \times \ell_2).\create \; \delta((v_1, v_2), (\ell_1.\get \;
-s_1, \ell_2.\get \; s_2))\\
-&= (\ell_1 \times \ell_2).\create \; \delta((v_1, v_2), (\ell_1
-\times \ell_2).\get (s_1, s_2))
-\end{align*}
+&= \delta (\ell_1.\pput \; s_1\; v_1, s_1) + \delta (\ell_2.\pput \; s_2 \;
+v_2, s_1)\\
+&\leq \delta(v_1, \ell_1.\get \; s_1) + \delta(v_2, \ell_2.\get \; s_2)\\
+&=\delta((v_1, v_2), (\ell_1.\get \; s_1, \ell_2.\get s_2))\\
+&= \delta((v_1, v_2), (\ell_1 \times \ell_2).\get (s_1, s_2))
+\end{align*}
+A similar argument shows that $\delta((\ell_1 \times \ell_2).\create \; (v_1,
+v_2), (s_1, s_2)) = \delta((v_1, v_2), (\ell_1 \times \ell_2).\get (s_1, s_2))$.
+This completes the proof.
 \end{proof}
@@
 Let $\ell_1 : S_1 \Leftrightarrow V_1$ and $\ell_2 : S_2 \Leftrightarrow V_2$ be
-lenses. Define the lens $\ell_1 \sqcup \ell_2 : (S_1 \sqcup S_2) \Leftrightarrow
-(V_1 \sqcup V_2)$ by
+lenses. Assume that $\rho(\ell_1.\create \; v_1) \leq
+\rho(v_1)$ and $\rho(\ell_2.\create \; v_2) \leq
+\rho(v_2)$ for all $v_1 \in V_1$ and $v_2 \in V_2$. Define the lens $\ell_1
+\sqcup \ell_2 : (S_1 \sqcup S_2) \Leftrightarrow (V_1 \sqcup V_2)$ by
 \begin{align*}
@@
 \begin{proof}
-We need to show that
-$$\delta((\ell_1 \sqcup \ell_2).\pput \; s \; v, s) = (\ell_1 \sqcup
-\ell_2).\pput \; s \; \delta(v, (\ell_1 \sqcup \ell_2).\get \; s)$$
+First we need to show that $\delta((\ell_1 \sqcup \ell_2).\pput \; s \; v, s)
+\leq \delta(v, (\ell_1 \sqcup \ell_2).\get \; s)$
 
 In the first case, assume that $s \in S_1$, $v \in V_1$. Then we need to show
-that $$\delta(\ell_1.\pput \; s \; v, s) = \ell_1.\pput \; s \; \delta(v,
-\ell_1.\get \; s)$$
+that $\delta(\ell_1.\pput \; s \; v, s) \leq \delta(v, \ell_1.\get \; s)$
 which follows from the fact that $\ell_1$ is a lens. A similar proof gives the
-same result if $s \in S_2$ and $v \in v_2$.
-
-Now assume that $s \in S_1$ and $v \in V_2$. Then we need to show
-that 
-$$\delta(\ell_2.\create \; v, s) = \ell_2.\create \; (\delta(v, \ell_2.\get \;
-s))$$ which follows from the fact that $\ell_2$ is a lens. A similar proof gives
-the same result if $s \in S_2$ and $v \in V_1$. This completes the proof.
-\end{proof}
-\subsubsection{Iteration}
-Let $\ell: S \Leftrightarrow V$ be a lens. Define the lens $\ell^*: S
-\Leftrightarrow V$ by
+same result if $s \in S_2$ and $v \in v_2$. Now assume that $s \in S_1$ and $v
+\in V_2$. Then we need to show that
+$\delta(\ell_2.\create \; v, s) \leq \delta(v, \ell_2.\get \; s)$ which follows
+from the fact that 
+$$\delta(\ell_2.\create \; v, s) = \rho(\ell_2.\create \; v) \leq
+\rho(v) = \delta(v, \ell_2.\get \; s)$$
+A similar argument shows that $\delta(\ell_1.\create \; v, s) \leq \delta(v,
+\ell_1.\get \; s)$ if $s \in S_2$ and $v \in V_1$. This completes the proof.
+\end{proof}
+\subsubsection{Positional Iteration}
+Let $\ell: S \Leftrightarrow V$ be a lens. Assume that $\rho(\ell.\create \;
+v) \leq \rho(v)$ for all $v \in V$. Define the lens $\ell^*:
+S \Leftrightarrow V$ by
 \begin{align*}
@@
 \begin{claim}
-$\ell_1 \times \ell_2$ is a valid lens.
+$\ell^*$ is a valid lens.
 \end{claim}
@@
 First, we need to show that
-\begin{multline}\label{positionalputexact}
-\delta(\ell^*.\pput \; (s_1, \ldots, s_n) \; (v_1, \ldots, v_m),
-(s_1, \ldots, s_n)) = \\ \ell^*.\pput \; (s_1, \ldots, s_n) \; \delta((v_1,
-\ldots, v_m), \ell^*.\get \; (s_1, \ldots, s_n))
-\end{multline}
+\begin{equation}\label{positionalputexact}
+\delta(\ell^*.\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n),
+(s_1, \ldots, s_m)) = \delta((v_1, \ldots, v_n), \ell^*.\get \; (s_1, \ldots,
+s_m))
+\end{equation}
 Now
@@
 \ell.\create \; v_i & \text{if } \min\{m, n\} < i \leq n
+\end{cases}
+\end{align*}
+hence
+\begin{multline}\label{positionalleft}
+\delta(\ell^* .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n), (s_1,
+\ldots, s_m))\\ = \sum_{i=1}^{\min\{m, n\}} \delta(\ell.\pput \; s_i \; v_i,
+s_i) + \sum_{j > \min\{m, n\}} \rho(\ell.\create \; v_j)
+\end{multline}
+Since $\delta$ is a difference function, then $\delta(\ell.\pput \; s_i \;
+v_i, s_i) \leq \delta(v_i, \ell.\get \; s_i)$ for $1 \leq i \leq \min\{m,n\}$
+and $\rho(\ell.\create \; v) \leq \rho(v)$ for all $v \in V$, then
+\begin{align*}
+&\delta(\ell^* .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n), (s_1,
+\ldots, s_m))\\ 
+&=\sum_{i=1}^{\min\{m, n\}} \delta(\ell.\pput \; s_i \; v_i, s_i) + \sum_{j >
+\min\{m, n\}} \rho(\ell.\create \; v_j)\\
+&\leq \sum_{i=1}^{\min\{m, n\}} \delta(v_i,
+\ell.\get \; s_i) + \sum_{j > \min\{m, n\}} \rho(v_j)\\
+&= \delta((v_1, \ldots, v_n), \ell^*.\get \; (s_1, \ldots, s_m))
+\end{align*}
+A similar argument shows that 
+$$\delta(\ell^*.\create \; (v_1, \ldots, v_n), (s_1, \ldots, s_m))
+=  \delta((v_1, \ldots, v_n), \ell^*.\get \; (s_1, \ldots, s_m)))
+$$
+This completes the proof.
+\end{proof}
+\subsubsection{Non-Crossing Iteration}
+Let $\ell: S \Leftrightarrow V$ be a lens. $\rho(\ell.\create \;
+v) \leq \rho(v)$ for all $v \in V$. Define the lens $\ell^*_{NC}: S
+\Leftrightarrow V$ by
+\begin{align*}
+\ell^*_{NC} .\get \; (s_1, \ldots, s_n) &= (\ell.\get \; s_1) \cdot \ldots \cdot
+(\ell.\get \; s_n)\\
+\ell^*_{NC} .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n) &= (s'_1 , \ldots
+, s'_n)\\
+\text{ where } s'_i &= \begin{cases}
+\ell.\pput \; s_{\varphi(i)} \; v_i & \text{if } i \in Dom(\varphi)\\
+\ell.\create \; v_i & \text{if } i \not \in Dom(\varphi)
 \end{cases}\\
-\end{align*}
-hence
-\begin{align*}
-\delta(\ell^* .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n)) &= s'_1 \cdot
-\ldots \cdot s'_n\\
+\text{ and } \varphi \text{ is the minimal matching for }
+&\delta^*_{NC, \psi}((v_1, \ldots, v_n), (\ell.\get \; s_1, \ldots, \ell.\get
+\; s_m)))\text{ over all }\psi\\
+\ell^*_{NC}.\create \; (v_1, \ldots, v_n) &= (\ell.\create
+\; v_1) \cdot \ldots \cdot (\ell.\create \; v_n)
+\end{align*}
+\begin{claim}\label{nocrossingvalid}
+$\ell^*_{NC}$ is a valid lens.
+\end{claim}
+\begin{proof}
+First, we need to show that
+\begin{multline}\label{noncrossingputexact}
+\delta_{NC}(\ell^*_{NC}.\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n),
+(s_1, \ldots, s_m)) \\ \leq  \delta_{NC}((v_1, \ldots, v_n), (\ell.\get \; s_1)
+\cdot \ldots \cdot (\ell.\get \; s_m))
+\end{multline}
+Let $\varphi$ be the minimal matching for $\delta^*_{NC, \psi}((v_1, \ldots,
+v_n), (\ell.\get \; s_1, \ldots, \ell.\get \; s_m)))$ over all matchings. Then
+\begin{align*}
+&\delta_{NC}(\ell^*_{NC} .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots,
+v_n), (s_1, \ldots, s_m))\\ 
+&\leq \sum_{i \in Dom(\varphi)} \delta(\ell.\pput \;
+s_{\varphi(i)} \; v_i, s_{\varphi(i)}) + \sum_{j \not \in Dom(\varphi)}
+\rho(\ell.\create \; v_j)\\
+&\leq \sum_{i \in Dom(\varphi)} \delta(v_i, s_{\varphi(i)}) + \sum_{j \not
+\in Dom(\varphi)} \rho(v_j)\\
+&= \delta_{NC}((v_1, \ldots, v_n), (\ell.\get \; s_1) \cdot \ldots \cdot
+(\ell.\get \; s_m))
+\end{align*}
+A similar argument shows that 
+$$\delta_{NC}(\ell^*_{NC}.\create \; (v_1, \ldots, v_n),(s_1, \ldots, s_m)) \leq
+\delta_{NC}((v_1, \ldots, v_n), (\ell.\get \; s_1) \cdot \ldots \cdot
+(\ell.\get \; s_m))$$
+\end{proof}
+\subsubsection{Min-Edit Distance Iteration}
+Let $\ell: S \Leftrightarrow V$ be a lens. $\rho(\ell.\create \;
+v) \leq \rho(v)$ for all $v \in V$. Let $\sim$ be the equivalence relation
+given by $(x_{\sigma(1)}, \ldots, x_{\sigma(n)}) \sim (x_{\theta(1)},
+\ldots, x_{\theta(n)})$ for all permutations $\sigma$ and $\theta$ of
+$[n]$. Define the lens $\ell^*_{MED}: S^*/{\sim} \Leftrightarrow V^*/{\sim}$ by
+\begin{align*}
+\ell^*_{MED} .\get \; [(s_1, \ldots, s_n)] &= [(\ell.\get \; s_1) \cdot \ldots
+\cdot (\ell.\get \; s_n)]\\
+\ell^*_{MED} .\pput \; [(s_1, \ldots, s_m)] \; [(v_1, \ldots, v_n)] &= [(s'_1,
+\ldots , s'_n)]\\
 \text{ where } s'_i &= \begin{cases}
-\delta(\ell.\pput \; s_i \; v_i) & \text{if } 1 \leq i \leq \min\{m, n\}\\
-\delta(\ell.\create \; v_i) & \text{if } \min\{m, n\} < i \leq n
+\ell.\pput \; s_{\varphi(i)} \; v_i & \text{if } i \in Dom(\varphi)\\
+\ell.\create \; v_i & \text{if } i \not \in Dom(\varphi)
 \end{cases}\\
-\end{align*}
-Since $\delta$ is a difference function, then
-\begin{align*}
-\delta(\ell^* .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n)) &= s'_1 \cdot
-\ldots \cdot s'_n\\
-\text{ where } s'_i &= \begin{cases}
-\ell.\pput \; s_i \; \delta(v_i, \ell.\get \; s_i) & \text{if } 1 \leq i \leq
-\min\{m, n\}\\
-\ell.\create \; \delta(v_i, \ell.\get \; s_i) & \text{if } \min\{m, n\} < i \leq
-n \end{cases}\\
-\end{align*}
-Now since
-\begin{equation}\label{expanddelta^n}
-\delta((v_1, \ldots, v_m), \ell^*.\get \; (s_1, \ldots, s_n)) = 
-\delta(v_1, \ell.\get \; s_1) \cdot \ldots \cdot \delta((v_n, \ell.\get \; s_n))
+\text{ and } \varphi \text{ is the minimal matching for }
+&\delta^*_{MED, \psi}([(v_1, \ldots, v_n)], [(\ell.\get \; s_1, \ldots,
+\ell.\get \; s_m)])\text{ over all }\psi\\
+\ell^*_{MED}.\create \; [(v_1, \ldots, v_n)] &= [(\ell.\create
+\; v_1) \cdot \ldots \cdot (\ell.\create \; v_n)]
+\end{align*}
+\begin{claim}
+$\ell^*_{MED}$ is a valid lens.
+\end{claim}
+\begin{proof}
+Proceed as in the proof of \cref{nocrossingvalid}.
+\end{proof}
+\subsubsection{Composition}
+
+Let $S, U, V$ be difference posets. Let $\ell_1 : S \Leftrightarrow U$
+and $\ell_2 : U \Leftrightarrow V$ be lenses. Define the lens $\ell_2
+\circ \ell_1 : S \Leftrightarrow V$ by
+\begin{align*}
+(\ell_2 \circ \ell_1).\get \; s &= \ell_2.\get \; (\ell_1.\get \; s)\\
+(\ell_2 \circ \ell_1).\pput \; s \; v &= \ell_1.\pput \; (\ell_2.\pput \;
+(\ell_1.\get \; s)) \; v\\
+(\ell_2 \circ \ell_1).\create \; v &= \ell_1.\create \; (\ell_2.\create \; v)
+\end{align*}
+
+First, we show that
+\begin{equation}\label{composeresult}
+\delta_S(\ell_1.\pput \; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; v), s) \leq
+\delta(\ell_2.\pput \; (\ell_1.\get \; s) \; v, \ell_1.\get \;(\ell_2.\get
+\; s))
 \end{equation}
-then
-\begin{multline}
-\ell^*.\pput \; (s_1, \ldots, s_n) \; \delta((v_1,
-\ldots, v_m), \ell^*.\get \; (s_1, \ldots, s_n)) = \\
-\ell^*.\pput \; (s_1, \ldots, s_n) \; \delta(v_1, \ell.\get \; s_1) \cdot
-\ldots \cdot \delta((v_n, \ell.\get \; s_n))
-\end{multline}
-
-\end{proof}
-\subsubsection{Composition}
-Let $S, U, V$ be weighted posets. Let $\ell_1 : S \Leftrightarrow U$ and $\ell_2
-: U \Leftrightarrow V$ be lenses. To show that $\ell_2 \circ \ell_1$ is also a
-lens, we need to show that $\ell_2 \circ \ell_1$ satisfies the EXACT law:
-\begin{multline}\label{composeresult}
-\delta_S(\ell_1.\pput \; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; v), s) =
-\ell_1.\pput \; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; (\delta_V(v, \ell_2.\get
-\; (\ell_1.\get \; s))))
-\end{multline}
 Since $\ell_1$ and $\ell_2$ are lenses, then for all $s \in S$, $u \in U$ and $v
@@
 \begin{equation}\label{compose1}
-\delta_S(\ell_1.\pput \; s \; u, s) = \ell_1.\pput \; s \; (\delta_U(u, \ell_1.\get
-\; s))
+\delta_S(\ell_1.\pput \; s \; u, s) \leq \delta_U(u, \ell_1.\get \; s)
 \end{equation}
@@
 \begin{equation}\label{compose2}
-\delta_U(\ell_2.\pput \; u \; v, u) = \ell_2.\pput \; u \; (\delta_V(v, \ell_2.\get
-\; u))
+\delta_U(\ell_2.\pput \; u \; v, u) \leq \delta_V(v, \ell_2.\get \; u)
 \end{equation}
@@
 \begin{equation}\label{compose1subst}
-\delta_S(\ell_1.\pput \; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; v), s) =
-\ell_1.\pput \; s \; (\delta_U(\ell_2.\pput \; (\ell_1.\get \; s) \; v, \ell_1.\get
-\; s))
+\delta_S(\ell_1.\pput \; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; v), s) \leq
+\delta_U(\ell_2.\pput \; (\ell_1.\get \; s) \; v, \ell_1.\get
+\; s)
 \end{equation}
@@
 \begin{equation}\label{compose2subst}
-\delta_U(\ell_2.\pput \; (\ell_1.\get \; s) \; v, (\ell_1.\get \; s)) = \ell_2.\pput
-\; (\ell_1.\get \; s) \; (\delta_V(v, \ell_2.\get \; (\ell_1.\get \; s)))
+\delta_U(\ell_2.\pput \; (\ell_1.\get \; s) \; v, \ell_1.\get \; s) \leq
+\delta_V(v, \ell_2.\get \; (\ell_1.\get \; s))
 \end{equation}
-Applying $(\ell_1.\pput \; s)$ to \cref{compose2subst} we \get
-\begin{equation}\label{compose3subst}
-\ell_1. \pput \; s \; (\delta_U(\ell_2.\pput \; (\ell_1.\get \; s) \; v, \ell_1.\get
-\; s)) = \ell_1.\pput \; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; (\delta_V(v,
-\ell_2.\get \; (\ell_1.\get \; s))))
-\end{equation}
-Equating \cref{compose1subst} and \cref{compose3subst} hence gives
+Comparing \cref{compose1subst} and \cref{compose2subst} hence gives
 \cref{composeresult}, which is what we wanted to show.
@@
 \iffalse
-\section{Weighted Posets and Difference Functions}
-\begin{definition}
-Let $P$ be a poset. A weight function $\rho$ on $P$ is simply an order
-preserving map from $P$ to $\mathbb{R}^{\geq 0}$, the non-negative real
-numbers.
-\end{definition}
 
@@
 Let $(P, \rho)$ be a weighted poset. A function $\delta : P
-\times P \longrightarrow \mathbb{R}^{\geq 0}$ is a difference function if and
+\times P \longrightarrow \mathbb{N}$ is a difference function if and
 only if
@@
 \end{proof}
-\iffalse
-Now define a
-function $d:
-P \longrightarrow \mathbb{R}$ on $P$ by $$d(a, b) =
-\begin{cases}
-0 & \text{if } a = b\\
-\rho(\pi_1(\delta(a, b))) + \rho(\pi_2(\delta(a, b))) & \text{otherwise}
-\end{cases}$$
-\begin{fact}
-\begin{enumerate}
-  \item
-  $d(a,b) \geq 0$,
-  \item
-  $d(a, b) = 0$ if and only if $a = b$, and
-  \item
-  $d(a, b) = d(b, a)$
-\end{enumerate}
-\end{fact}
-\begin{fact}
-If $b \leq c$, then for all $b$, $d(a, b) \geq d(a, c)$. Thus if $b \leq c$,
-$d(a, c) \leq d(a, b) + d(b, c)$
-\end{fact}
-\begin{question}
-Under what conditions is $d$ a metric?
-\end{question}
-\fi
+
 \begin{definition}
@@
 \ell.\create &: V \longrightarrow S\\
-\delta_S &: S \times S \longrightarrow \mathbb{R}^{\geq 0}\\
-\delta_V &: V \times V \longrightarrow \mathbb{R}^{\geq 0}
+\delta_S &: S \times S \longrightarrow \mathbb{N}\\
+\delta_V &: V \times V \longrightarrow \mathbb{N}
 \end{align*}
@@
 Let $\delta_P$ be a difference function on $P$. We define the function
-$\Delta^*$ on $P^*$ by
-$$\Delta^*((x_1, \ldots, x_m), (y_1, \ldots, y_n)) = \sum_{i=1}^{\min\{m,
+$\delta_{NC}^*$ on $P^*$ by
+$$\delta_{NC}^*((x_1, \ldots, x_m), (y_1, \ldots, y_n)) = \sum_{i=1}^{\min\{m,
 n\}}\delta(x_i, y_i) + \sum_{j = \min\{m, n\} + 1}^{m}\rho(x_i)$$
 \begin{claim}
-$\Delta^*$ is a difference function.
+$\delta_{NC}^*$ is a difference function.
 \end{claim}
@@
   m$ from which it follows that
-  $$\Delta^*((x_1, \ldots, x_m), (y_1,
+  $$\delta_{NC}^*((x_1, \ldots, x_m), (y_1,
   \ldots, y_n)) \leq \sum_{i=1}^m\rho(x_i) = \rho^*((x_1, \ldots, x_n))$$
@@
   hence
-  $$\Gamma^*((x_1, \ldots, x_m), (y_1, \ldots, y_n)) =
+  $$\delta_{NC}^*((x_1, \ldots, x_m), (y_1, \ldots, y_n)) =
   \sum_{i=1}^{m}\delta(x_i, y_i) = 0,$$
@@
   Assume that $(x_1, \ldots, x_m) \leq (y_1, \ldots, y_{m'})$. Then $m \leq
-  m'$ and $x_i \leq y_i$. Also, $$\Delta^*((z_1, \ldots, z_n), (x_1,
+  m'$ and $x_i \leq y_i$. Also, $$\delta_{NC}^*((z_1, \ldots, z_n), (x_1,
   \ldots, x_m)) = \sum_{i=1}^{\min\{m, n\}}\delta(z_i, x_i) + \sum_{j =
   \min\{m, n\} + 1}^{n}\rho(z_i)$$ and
-  $$\Delta^*((z_1, \ldots, z_n), (y_1, \ldots, y_{m'})) =
+  $$\delta_{NC}^*((z_1, \ldots, z_n), (y_1, \ldots, y_{m'})) =
   \sum_{i=1}^{\min\{m', n\}}\delta(z_i, y_i) + \sum_{j = \min\{m', n\} +
@@
   \delta(z_i, y_i)$ for $1 \leq i \leq m$, hence
-  $$\Delta^*((z_1, \ldots, z_n), (x_1, \ldots, x_m)) \geq \Delta^*((z_1,
+  $$\delta_{NC}^*((z_1, \ldots, z_n), (x_1, \ldots, x_m)) \geq \delta_{NC}^*((z_1,
   \ldots, z_n), (y_1, \ldots, y_{m'}))$$ This completes the proof.

### 2018-02-05 papers/QuotientLenses/main.tex
Put Quotient Lenses file in ICFP format

@@
 
-\documentclass{svproc}
-
+\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+% % For double-blind review submission, w/ CCS and ACM Reference
+% \documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
+% % For single-blind review submission, w/o CCS and ACM Reference (max
+% submission space)
+% \documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+% % For single-blind review submission, w/ CCS and ACM Reference
+% \documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true} % For
+% final camera-ready submission, w/ required CCS and ACM Reference
+% \documentclass[acmsmall]{acmart}\settopmatter{}
+
+% % Journal information % Supplied to authors by publisher for camera-ready
+% submission; % use defaults for review submission.
+\acmJournal{PACMPL}
+\acmVolume{1}
+\acmNumber{ICDP} % CONF = POPL or ICFP or OOPSLA
+\acmArticle{1}
+\acmYear{2018}
+\acmMonth{1}
+\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
+\startPage{1}
+
+% % Copyright information % Supplied to authors (based on authors' rights
+% management selection; % see authors.acm.org) by publisher for camera-ready
+% submission; % use 'none' for review submission.
+\setcopyright{none}
 \usepackage{amsmath, amssymb, verbatim, enumerate, graphicx, centernot, tikz,
@@
 \lstset{
-  language=Caml,
-  basicstyle=\upshape\sffamily,
-  keywordstyle=\upshape\sffamily\color{dkblue},
-  keepspaces=true,
-  framexleftmargin=1ex,
-  framexrightmargin=1ex,
-  showstringspaces=false,
-  commentstyle=\itshape\rmfamily,
-  columns=flexible,
+language=Caml, basicstyle=\upshape\sffamily,
+keywordstyle=\upshape\sffamily\color{dkblue}, keepspaces=true,
+framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=false,
+commentstyle=\itshape\rmfamily, columns=flexible,
 % BCP: I find this distracting:
@@
 
-%%%%%
-% Macros
-% Colors
+% %%%% Macros Colors
 \definecolor{dkblue}{rgb}{0,0.1,0.5}
@@
 \newcommand{\RuleSide}[3]{\infer[#3]{#2}{#1}}
-%\newcommand{\Axiom}[1]{\Rule{}{#1}}
+% \newcommand{\Axiom}[1]{\Rule{}{#1}}
 
@@
 
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
 \begin{document}
-\mainmatter              % start of a contribution
-%
 \title{Synthesizing Quotient Lenses}
-%
-\titlerunning{Synthesizing Quotient Lenses}  % abbreviated title (for running
-% head)                                     also used for the TOC unless
-%                                     \toctitle is used
-%
-\author{}
-%
-\authorrunning{} % abbreviated author list (for running head)
-%
-%%%% list of authors for the TOC (use if author list has to be modified)
-\tocauthor{}
-%
-\institute{}
-
-\maketitle              % typeset the title of the contribution
-
 \begin{abstract}
@@
 \end{abstract}
+
+\keywords{quotient lenses, synthesis}
+\maketitle              % typeset the title of the contribution
 
@@
 each of the data formats is quotiented by some equivalence relation and the lens
-laws are taken modulo equivalence. This both permits canonization functions and
-allows lenses to be simpler and more flexible. Inessential details like
-whitespace and orderings of fields can be ignored when writing the
-transformations between the formats.
+laws are taken modulo equivalence. This allows lenses to be simpler and more
+flexible. For example, inessential details like whitespace and orderings of
+fields can be ignored when writing the transformations between the formats.
 
@@
 transformations between the formats, they merely push complexity elsewhere. To
-write a quotient lens, programmers must write with both standard lens
+write a quotient lens, programmers must write with standard lens
 combinators and \emph{canonizers}. Canonizers encode two functions, one that
@@
 introduce a novel language of {\em Quotient Regular Expressions}, or QREs, which
-enable a programmer to simultaneously specify a regular expression while
-simultaneously specifying the canonical representatives on the regular
-expression's language. With QREs, programmers needn't define the data and
-canonization functions separately -- by adding simple annotations to the data,
-the canonization functions can be inferred.
-
-Secondly, we use QREs to define a class of quotient lenses called {\em QRE
-  lenses} which map betweeen QREs, and show how to synthesize QRE lenses. Using
-this tool, a lens programmer need not manually derive the lens herself, she
-merely must provide a pair of QREs (describing the formats the lens should map
-between), and a suite of examples (demonstrating how the lens should act on
-specific instances).  The resulting lens is guaranteed to map between the
-languages of the QREs and satisfy the provided examples.
+enable a programmer to simultaneously specify a regular expression as well as
+a set of canonical representatives on the regular expression's language. With
+QREs, programmers do not need to define the data and canonization functions
+separately -- by adding simple annotations to the data, the canonization
+functions can be inferred.
+
+Second, we use QREs to define a class of quotient lenses called {\em QRE
+lenses} which map betweeen QREs, and show how to synthesize QRE lenses. In
+order to derive a lens, a programmer need not manually derive the lens
+herself, she only needs to provide a pair of QREs describing the formats the
+lens should map between, and a suite of examples that demonstrates how the
+lens should act on specific examples.  The resulting lens is guaranteed to
+map between the languages of the QREs and satisfy the provided examples.
 
@@
 \begin{enumerate}
-\item We introduce a novel language of {\em Quotient Regular Expressions} (QREs)
-  that provide a compact, convenient notation for regular languages and their
-  canonical representatives (Section~\ref{QRE}).
+  \item We introduce {\em Quotient Regular Expressions} (QREs)
+  which provide a compact, convenient notation that expresses an equivalence
+  relation on a regular languages as well as as a canonical set representatives
+  for this equavalence relation (Section~\ref{QRE}).
   \item We define a language of {\em QRE lenses}, a subset of bijective quotient
@@
   \item Using this normal form, we reduce the problem of {\em synthesizing}
-QRE lenses to the previously studied
-problem of synthesizing bijective lenses between regular languages 
-(Section~\ref{synth}).
+  QRE lenses to the previously studied problem of synthesizing bijective lenses
+  between regular languages (Section~\ref{synth}).
   \item We extend the Boomerang {\em implementation} with QREs
@@
 \bibliographystyle{plain}
-\bibliography{local,bcp}
+\bibliography{local}
 

### 2018-02-07 papers/QuotientLenses/main.tex
Made a pass through paper

@@
 In this section, we explain the difficulties of programming with quotient lenses
-and show, by way of example, how our novel language of {\em Quotient Regular
-Expressions}, or QREs, and QRE lenses address these difficulties.
+in Boomerang and show, by way of example, how our novel language of {\em
+Quotient Regular Expressions}, or QREs, and QRE lenses address these
+difficulties.
 
@@
 quotenting of fields---nest within each other in a way that makes them difficult
-to express with the existing Boomerang combinators.
-
-Furthermore, programming best practices require programmers provide type
-annotations to their lenses to make explicit the formats the quotient lenses map
-between. This practice prevents nefarious bugs that can result from the lenses
-mapping between incorrect formats, but it is also verbose and time consuming.
+to express with the existing Boomerang combinators. Furthermore, if the
+programmer wishes to specify the type of the lens, then she can only specify
+data formats on each side of the lens with a regular expression that does not
+keep track of the equivalence relation defined on the language of the regular
+expression. Consequently, the programmer has to write a regular expression as
+well as a canonizer that expresses the equivalence relation on that regular
+expression, even though a regular expression and the canonizer defined on it
+typically share much of the syntactic structure. Having to write both of these
+expressions makes programming with quotient lenses verbose and time consuming.
 
 To address these issues, we propose {\em Quotient Regular Expressions}, or QREs,
-which enable the programmer to define a regular expression $R$ and its canonical
-elements in the same expression. For example, with QREs we can specify \bibtex{}
-records and their canonical elements, as shown in Figure~\ref{fig:example-qre}.
+which enable the programmer to define a regular expression $R$, an equivalence
+relation $\sim$ on the language described by $R$ as well as a canonical set
+of representatives of $\sim$, all in one syntactic term. For example, with QREs
+we can specify \bibtex{} records and their canonical elements, as shown in
+Figure~\ref{fig:example-qre}.
 
@@
 In Figure~\ref{fig:example-qre}, the QRE \verb+project [ \n\t\r]+$^{+} \to$
-\verb+" "+ matches any non-zero number of whitespace characters, with a single
-space as the canonical representative. We call this QRE a \textit{projection}
-because an entire regular language is projected to a single representative
-string during canonization.
+\verb+" "+ matches any non-zero number of whitespace characters and quotients
+them to a single equivalence class with a single space as the canonical
+representative. We call this QRE a \textit{projection} because an entire regular
+language is projected to a single representative string.
 
@@
 \cd{comma_name} QREs, and we need to use the ``get'' direction of the lens to
-specify the squash transformation. (We'll say more about synthesis below.)
+specify the squash transformation.
 
@@
 
-\sam{TODO: Put diagram of composition here}
+\sam{TODO: We need to put a diagram of composition here.}
 
@@
 Figure~\ref{fig:example-qre}, the synthesis procedure itself can be used to
-creates lenses that are in turn used to define other QREs.  This ability to mix
+create lenses that are in turn used to define other QREs.  This ability to mix
 QRE specification with QRE lens synthesis yields a powerful and flexible way of
@@
 \kw{first\_name} and \kw{last\_name}:
-
-\saz{Fix the notation in the following code.}
 
@@
 
-\saz{Say more about how normalize
-wouldn't be used in ``ordinary'' QRE programming?}\sam{I'm not sure what you
-mean here exactly\ldots} However, checking that the canonizing function $f$ is
-surjective and idempotent is undecidable in general, thus we do not use the
-$\normalize{R}{R'}{f}$ in practice. On the other hand, the
+Now in order for the programmer to ensure that the {\em normalize} QRE is well
+formed, she must define a canonizing function $f$ between two regular
+languages and prove that $f$ is surjective and idempotent: the compiler cannot
+verify this since this problem is undecidable in general. Consequently, we do
+not use the $\normalize{R}{R'}{f}$ QRE in practice. On the other hand, the
 $\normalize{R}{R'}{f}$ combinator gives a sufficient condition for potential
@@
 gives the most interesting definition,
-
 \begin{align*}\canonizer(\perm{c_1, \ldots, c_n}{c})(r_{\sigma(1)}
@@
 \end{align*}
-
 \noindent which as we have explained is defined so that the strings that match
@@
 \rrbracket$, then $\ell$ maps between $s_1$ and $s_2$.
-
 \begin{align*}
@@
 \end{align*}
-
 The $\mathit{const}(s, t)$ lens replaces the string $s$ with $t$ in the source
@@
 \begin{lstlisting}
-let bib_to_end_q : (lens in bibtex $\Leftrightarrow$ endnote) =
-rquot (lquot (bibtex, bib_to_end), endnote)
+let bib_to_end_q : (lens in bibtex $\Leftrightarrow$ endnote)  = rquot (lquot(bibtex, bib_to_end), endnote)
 \end{lstlisting}
@@
 approach, which is to define the composition on QRE lenses if and only if
-$c'$ and $c''$ are syntactially equal to $id(R)$ and $id(R')$ for some $R$and
+$c'$ and $c''$ are syntactially equal to $id(R)$ and $id(R')$ for some $R$ and
 $R'$ satisfying $\mathcal{L}(R) = \mathcal{L}(R')$.
@@
 We elide a full proof of this statement, since the proof is attained via a
-straightforward induction on the derivation $q : c \Leftrightarrow c$. \saz{(The
-full details can be found in the appendix?)}Since
+straightforward induction on the derivation $q : c \Leftrightarrow c$ (the
+full details can be found in the appendix {\sam{which I need to write}}). Since
 bijective lenses are closed under strongly unambiguous regular operators, the
@@
 \begin{align*}
-q.get(x_i) \equiv_{c'} y_i \text{, and}\\
-q.put(y_i) \equiv_c x_i
+q.get(x_i) &\equiv_{c'} y_i \text{, and}\\
+q.put(y_i) &\equiv_c x_i
 \end{align*}

### 2018-02-09 papers/QuotientLenses/main.tex
added asts count graph

@@
 
-\newcommand{\QRESize}{QS}
-\newcommand{\CanonizerAndSpecSize}{CS}
-\newcommand{\LensAndSpecSize}{LS}
+\newcommand{\QRESize}{\textbf{QS}}
+\newcommand{\CanonizerAndSpecSize}{\textbf{CS}}
+\newcommand{\LensAndSpecSize}{\textbf{LS}}
 
@@
 
+\begin{figure}
+\includegraphics{generated-graphs/asts.eps}
+\caption{Count of benchmark programs definable using a given AST count. We find
+  that it takes far fewer AST nodes to define benchmark lenses in \Name{} than
+  with Optician or without synthesis. }
+\label{fig:asts}
+\end{figure}
+
 To test if our tool eases the development process, we compare the following
@@
 
-Figure~\ref{?} shows the number of benchmarks written with at most a given
-number of AST nodes.  We found that \QRESize{} was consistently smaller than
-\CanonizerAndSpecSize{} and \LensAndSpecSize{}, demonstrating that introducing
-QREs saves programmers programming effort compared to both Optician and basic
-Boomerang.
+Figure~\ref{fig:asts} shows the number of benchmarks written with at most a
+given number of AST nodes. For a given benchmark, we found \QRESize{} to be
+about half as large as \CanonizerAndSpecSize, and about a third as large as
+\LensAndSpecSize{}. This demonstrates that introducing QREs saves programmers
+significant effort compared to both Optician and basic Boomerang.
 

### 2018-02-09 papers/QuotientLenses/transform-data.py

@@ -1,0 +1,107 @@
+#!/usr/local/bin/python
+
+from __future__ import print_function
+from easyprocess import EasyProcess
+
+import os
+import csv
+from os.path import splitext, join
+import subprocess
+import sys
+import time
+import matplotlib
+import matplotlib as mpl
+mpl.use('pgf')
+import numpy as np
+import matplotlib.pyplot as plt
+
+plt.rc('font', size=10)
+plt.rc('legend', fontsize=10)
+plt.rcParams['text.usetex'] = True
+plt.rcParams['text.latex.preamble'] = '\usepackage{libertine},\usepackage[libertine]{newtxmath},\usepackage[T1]{fontenc}'
+
+generated_graphs_base = "generated-graphs/"
+transformed_data_base = "transformed-data/"
+
+def ensure_dir(f):
+    d = os.path.dirname(f)
+    if not os.path.exists(d):
+        os.makedirs(d)
+
+
+def print_usage(args):
+    print("Usage: {0} <file1>".format(args[0]))
+
+def retrieve_csv(filename):
+    csv_rows = []
+    with open(filename, 'rb') as csvfile:
+        reader = csv.DictReader(csvfile)
+        for row in reader:
+            csv_rows.append(row)
+    return csv_rows
+
+def project_column_from_csv(csv_obj, col_name):
+    return [r[col_name] for r in csv_obj]
+
+def write_to_filename(filename, s):
+    with open(filename, "wb") as f:
+        f.write(s)
+
+def generate_time_vs_tasks_graph(input_csv):
+    fig, ax = plt.subplots()
+
+    def create_step_plot(colname, outputname,style,width):
+        col_vals = [float(x) for x in project_column_from_csv(input_csv, colname) if int(x) < 7500]
+        print(col_vals)
+        col_vals_and_endpoints = col_vals + [0,8000]
+        x_vals = sorted([x for x in set(col_vals_and_endpoints)])
+        x_count_dict = {key: 0 for key in x_vals}
+        for val in col_vals:
+            x_count_dict[val] = x_count_dict[val]+1
+        x_completed_counts = []
+        acc = 0
+        for val in x_vals:
+            acc = acc + x_count_dict[val]
+            x_completed_counts.append(acc)
+        x_completed_counts = [0] + x_completed_counts[:len(x_completed_counts)-1]
[…]

### 2018-02-14 papers/SDLenses/main.tex
Added first draft

\newif\ifdraft\drafttrue  % set true to show comments
\newif\ifplentyoftime\plentyoftimefalse  % :-)

% % For double-blind review submission, w/o CCS and ACM Reference (max
% submission space)
\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
% % For double-blind review submission, w/ CCS and ACM Reference
% \documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
% % For single-blind review submission, w/o CCS and ACM Reference (max
% submission space)
% \documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
% % For single-blind review submission, w/ CCS and ACM Reference
% \documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true} % For
% final camera-ready submission, w/ required CCS and ACM Reference
% \documentclass[acmsmall]{acmart}\settopmatter{}

% % Journal information % Supplied to authors by publisher for camera-ready
% submission; % use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{ICDP} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

% % Copyright information % Supplied to authors (based on authors' rights
% management selection; % see authors.acm.org) by publisher for camera-ready
% submission; % use 'none' for review submission.
\setcopyright{none}
% \setcopyright{acmcopyright} \setcopyright{acmlicensed}
% \setcopyright{rightsretained} \ccopyrightyear{2018}           %% If different
% from \acmYear

% % Bibliography style %\bibliographystyle{ACM-Reference-Format} % Citation
% style % Note: author/year citations are required for papers published as an %
% issue of PACMPL.
% %\citestyle{acmauthoryear}   %% For author/year citations

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% % Note:
% Authors migrating a paper from PACMPL format to traditional % SIGPLAN
% proceedings format must update the '\documentclass' and % topmatter commands
% above; see 'acmart-sigplanproc-template.tex'.
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{amsmath, amssymb, amsthm, enumerate, graphicx, centernot, tikz,
array, tikz-cd, extarrows, cleveref, mathrsfs, mathtools, stmaryrd, listings}

\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{conjecture}{Conjecture}
\newtheorem{question}{Question}
\newtheorem{fact}{Fact}
\newtheorem{claim}{Claim}

% Macros Colors
\definecolor{dkblue}{rgb}{0,0.1,0.5}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{dkred}{rgb}{0.6,0,0}
\definecolor{dkpurple}{rgb}{0.7,0,0.4}
\definecolor{olive}{rgb}{0.4, 0.4, 0.0}
\definecolor{teal}{rgb}{0.0,0.5,0.5}
\definecolor{orange}{rgb}{0.9,0.6,0.2}
\definecolor{lightyellow}{RGB}{255, 255, 179}
\definecolor{lightgreen}{RGB}{170, 255, 220}
\definecolor{teal}{RGB}{141,211,199}
\definecolor{darkbrown}{RGB}{121,37,0}

\newcommand{\FINISH}[3]{\ifdraft\textcolor{#1}{[#2: #3]}\fi}
\newcommand{\bcp}[1]{\FINISH{dkred}{B}{#1}}
\newcommand{\BCP}[1]{\FINISH{dkred}{B}{\bf #1}}
\newcommand{\afm}[1]{\FINISH{dkgreen}{A}{#1}}
\newcommand{\dpw}[1]{\FINISH{dkblue}{D}{#1}} % Toronto Maple Leafs Blue :-)
\newcommand{\saz}[1]{\FINISH{orange}{SZ}{#1}}
\newcommand{\ksf}[1]{\FINISH{teal}{K}{#1}}
\newcommand{\sam}[1]{\FINISH{dkpurple}{SM}{#1}}

\newcommand{\kw}[1]{\ensuremath{\mathsf{#1}}}
\newcommand{\get}{\ensuremath{\kw{get}}}
\newcommand{\pput}{\ensuremath{\kw{put}}}
\newcommand{\create}{\ensuremath{\kw{create}}}
\newcommand{\ccopy}{\ensuremath{\kw{copy}}}
\newcommand{\const}{\ensuremath{\kw{const}}}
\newcommand{\default}{\ensuremath{\kw{default}}}
\newcommand{\pos}{\ensuremath{\kw{posit}}}
\newcommand{\length}{\ensuremath{\kw{length}}}
\newcommand{\match}{\ensuremath{\kw{match}}}

\lstset{ basicstyle=\ttfamily, escapeinside=|| }
\begin{document}

% % Title information
\title{Stateful Delta Lenses}         %% [Short Title] is optional;
% % when present, will be used in % header instead of Full Title.
% %\titlenote{with title note}             %% \titlenote is optional; % can be
% repeated if necessary; % contents suppressed with 'anonymous'
% %\subtitle{Subtitle}                     %% \subtitle is optional
% %\subtitlenote{with subtitle note}       %% \subtitlenote is optional; % can
% be repeated if necessary; % contents suppressed with 'anonymous'

% % Author information % Contents and number of authors suppressed with
% 'anonymous'.
% % Each author should be introduced by \author, followed by % \authornote
% (optional), \orcid (optional), \affiliation, and % \email.
% % An author may have multiple affiliations and/or emails; repeat the %
% appropriate command.
% % Many elements are not rendered, but should be provided for metadata %
% extraction tools.

% % Author with single affiliation.
% \author{First1 Last1} \authornote{with author1 note}          %% \authornote
% is optional; % can be repeated if necessary \orcid{nnnn-nnnn-nnnn-nnnn} %%
% \orcid is optional \affiliation{ \position{Position1} \department{Department1}
%              %% \department is recommended \institution{Institution1}
%   %% \institution is required \streetaddress{Street1 Address1} \city{City1}
% \state{State1} \postcode{Post-Code1} \country{Country1}                    %%
% \country is recommended } \email{first1.last1@inst1.edu}          %% \email is
% recommended

\author{Solomon Maina}
\position{PhD Student}
\department{Computer Science}              %% \department is recommended
\institution{University of Pennsylvania}            %% \institution is required
\country{USA}                    %% \country is recommended }
\email{smaina@seas.upenn.edu}          %% \email is recommended

% % Abstract % Note: \begin{abstract}...\end{abstract} environment must come %
% before \maketitle command
\begin{abstract}
A {\em lens} is a program that can be run in two directions. In the
forward direction, the \get component of a lens produces a {\em view}
from a {\em source}, while in the backward direction, the \pput component folds
an updated view back into the original source. One of the reasons that
lenses are appealing is that they satisfy laws which give guarantees on how the
two transformations interact.

Unfortunately, the lens laws only require that the \pput component of a lens
return the original source unmodified whenever the update to the view is a
no-op; they do not specify the behaviour of the \pput component when the update
to the view is non-trivial. Consequently, a lens may be valid even though it
does not propagate non-trivial updates made to the view back into the source in
the manner intended. For example, a small update to the view can result
in a large update to the source, such as when a lens updates records
positionally even though the source and view are order agnostic. This paper
introduces {\em stateful delta lenses} or SD-lenses as a solution to this
shortcoming of traditional lenses. SD-lenses are lenses augmented with extra
information that specifies what kind of updates are valid for a given lens, and
how these updates should be propagated by the lens, and hence enable the
programmer to specify more exactly the desired behaviour of a lens.
\end{abstract}

%% Keywords
%% comma separated list
\keywords{difference structure, stateful delta lens}  %% \keywords
% are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}
A {\em lens} is a program that can be run in two directions. In the
forward direction, the \get component of a lens produces a {\em view}
from a {\em source}, while in the backward direction, the \pput component folds
an updated view back into the original source.

For example suppose that the source is an XML document representing the
names, dates and nationalities of a collection of classical music
composers such as,

\begin{lstlisting}
<composers>
    <composer>
        <name>Jean Sibelius</name>
        <dates>1865-1956</dates>
        <nationality>Finnish</nationality>
    </composer>
    <composer>
        <name>Aaron Copland</name>
        <dates>1910-1990</dates>
        <nationality>American</nationality>
    </composer>
    <composer>
        <name>Benjamin Briten</name>
        <dates>1913-1976</dates>
        <nationality>English</nationality>
    </composer>
</composers>
\end{lstlisting}

while the view is a list if ASCII text representing the names and dates of each
composer:
\begin{lstlisting}
Jean Sibelius, 1865-1956
Aaron Copland, 1910-1990
Benjamin Briten, 1913-1976
\end{lstlisting}

After computing the initial view, we might want to edit it in some way---e.g.,
correcting the error in Sibelius's death date and the misspelling in Britten's
name
\begin{lstlisting}
Jean Sibelius, 1865-|\colorbox{orange}{1957}|
Aaron Copland, 1910-1990
Benjamin |\colorbox{orange}{Britten}|, 1913-1976
\end{lstlisting}
and push the changes back into the original XML format:
\begin{lstlisting}
<composers>
    <composer>
        <name>Jean Sibelius</name>
        <dates>1865-|\colorbox{orange}{1957}|</dates>
        <nationality>Finnish</nationality>
    </composer>
        <composer>
        <name>Aaron Copland</name>
        <dates>1910-1990</dates>
        <nationality>American</nationality>
    </composer>
    <composer>
        <name>Benjamin |\colorbox{orange}{Britten}|</name>
        <dates>1913-1976</dates>
        <nationality>English</nationality>
    </composer>
</composers>
\end{lstlisting}

Lenses are designed to perfom these kinds of computations. If $\ell$ is
a suitably defined lens for the bidirectional tranfsormation that we
have just described, then the \get component of the lens performs the
XML to ASCII transformation that produces an ASCII view from an XML source,
while the \pput component of the lens performs the ASCII to XML transformation
that folds back an updated ASCII view back into the original XML source.

Formally, Foster \cite{foster2007combinators,foster2009bidirectional} defined a lens $\ell :
S \Leftrightarrow V$ from a source $S$ to a view $V$ to be a triple of
functions $\ell.\get : S \longrightarrow V$, $\ell.\pput : S \longrightarrow S
\longrightarrow V$ and $\ell.\create : V \longrightarrow S$ satisfying the
following properties:
\begin{align*}
\ell.\get \; (\ell.\pput \; s \; v) &= v \tag{PUTGET}\\
\ell.\get \; (\ell.\create \; v) &= v \tag{CREATEGET}\\
\ell.\pput \; s \; (\ell.\get \; s) &= s \tag{GETPUT}
\end{align*}
The original definition of lenses in \cite{foster2007combinators} involved only the \get
and \pput functions, but a \create function was later added in
\cite{foster2009bidirectional} to handle the special case where we need to construct a
source from a view, but we have no source to use as the original. \sam{Perhaps
we should include this in the footnotes?}

The PUTGET and CREATEGET laws ensure that updates to the view are translated
``exactly''---i.e. that, given a view, the \pput and \create functions produce
a source that \get maps back to the very same view, while the GETPUT law ensures
a ``stability'' property for the source---i.e. it requires that the \pput
function returns the original source unmodified whenever the update to the view
is a no-op.

While these two laws give some guarantees on how the \get and \pput components
of a lens interact, they are not strong enough to guarantee that the
updates made to a view are indeed the updates reflected back into the source.
This is because the GETPUT law only says that the \pput component must return
the original source unmodified whenever the update to the view is a no-op; it
does not say what the \pput component should do with a modified view.
Consequently, a non-trivial update to a view could lead to an undesired update
to the source.

For example, suppose that a source $S$ consists of a pair of integers, and
that the view $V$ for $S$ consists of a single integer. Further, suppose that we
wish for the \get of any lens from $S$ to $V$ to simply project away
the first factor:
$\ell.\get \; (a, b) = b$. Suppose further that we wish to force the \pput
function of any lens $\ell$ defined on the data to restore the factor projected
away by the \get function: that is, we wish to enforce that for
all integers $a, b, c$, $\pi_1(\ell.\pput \; (a, b) \; c) = a$, where $\pi_1$ is
the projection on the first factor.

Now consider the lens $\ell_{bad}$ defined by $\ell_{bad}.\get \; (a, b) = b$
and
\begin{align*}
\ell_{bad}.\get \; (a, b) &= b\\
\ell_{bad}.\pput \; (a, b) \; c &= \begin{cases}
(a, c) & \text{if }b = c\\
(0, c) & \text{otherwise}\\
\end{cases}\\
\ell.\create \; b &= (0,b)
\end{align*}
Then $\ell_{bad}$ is a valid lens even though $\ell_{bad}$ does not
always restore the first factor. For example, observe that $\ell.\pput (1, 2) \;
3 = (0,3)$. The lens laws do not allow us to specify how non-trivial updates are
propagated, hence the \pput component can cause any update to the source, so
long as the \get component maps the updated source back to the updated view.

In this paper, we introduce {\em stateful delta lenses} or SD-lenses, which are
lenses augmented with extra information that specifies what kind of updates are
valid for a given lens, and how these updates should be propagated.

In order to specify the intended behaviour of a lens, we will first need to
define what updates can be made to the source and view. We do this by embedding
source and views in {\em difference structures}:
\begin{definition}
A difference structure $(X, \partial X, \delta_X, i_X)$ is a 4-tuple such that
$X$ is a set, $\partial X$ is a set denoting the set of updates that can be
made in $X$, $\delta_X : X \times X \longrightarrow \partial X$, the {\em
difference function} of the structure, is a map that takes two elements of $x,
x' \in X$ and computes their difference $\delta_X$ as an element of $\partial
X$, and $i_X : X \longrightarrow \partial X$, the {\em inclusion function} of
the structure, is a map that given an $x \in X$ creates an update $i_X(x)$ in
$\partial X$.
\end{definition}

Going back to our previous example, let $\epsilon$ denote an ``empty''
symbol. On the source side, we define the ordered difference $\delta_S((a, b),
(c, d))$ between $(a, b)$ and $(c, d)$ by
$$\delta_S((a, b), (c,
d)) = \begin{cases}
(\epsilon, \epsilon) & \text{if }a = c, b = d\\
(\epsilon, b) & \text{if }a=c, b \neq d\\
(a, \epsilon) & \text{if }a \neq c, b = d\\
(a, b) & \text{if }a \neq c, b \neq d
\end{cases}$$
On the view side, we define the ordered difference $\delta_V(a, b)$
between $a$ and $b$ by, $$\delta_V(a, b) = \begin{cases}
(\epsilon, \epsilon) & \text{if } a = b\\
(\epsilon, a) & \text{if } a \neq b
\end{cases}$$
where we include an $\epsilon$ in the first factor to enforce the property that
the first factor should always be left unchanged. Thus the update space
$\partial S$ of $S$ is given by $(\{\epsilon\} \sqcup \mathbb{Z})
\times(\{\epsilon\} \sqcup \mathbb{Z})$ and the update space $\partial V$ of
$V$ is also given by $(\{\epsilon\} \sqcup \mathbb{Z}) \times(\{\epsilon\}
\sqcup \mathbb{Z})$. For inclusion functions, we define $i_S : S \longrightarrow
\partial S$ by $i_S(a, b) = (a, b)$ and $i_V : V \longrightarrow \partial V$ by
$i_V(a) = (\epsilon, a)$.

Next we define a lens $\partial(S, V) : \partial S \Leftrightarrow \partial V$
which describes how updates should be propagated. For this example, we let
define $\partial(S,V)$ to simply be the \ccopy lens from $\partial S$ to
$\partial V$, keeping in mind that $\partial S = \partial V$:
\begin{align*}
(\partial(S, V)).\get \; (a, b) &= (a, b)\\
(\partial(S, V)).\pput \; (a, b) \; (c, d) &= (c, d)\\
(\partial(S, V)).\create \; (a, b) &= (a, b)
\end{align*}
To specify the intended behaviour of a lens $\ell$ from $S$ to $V$ we
require that
\begin{equation}\label{partialpreservesdelta}
(\partial (S, V)).\get \; (\delta_S(\ell.\pput \; s \; v, s)) = \delta_V(v,
\ell.\get \; s)
\end{equation}
This condition says that if we start with a source $s$, map $s$ to the view
to get $\ell.\get \; s$ and then update $\ell.\get \; s$ to get a new view $v$,
then the \get component of $\partial(S, V)$ applied to $\delta_S(\ell.\pput \; s
\; v, s)$ is equal to $\delta_V(v, \ell.\get \; s)$.

For example, observe that $\ell_{bad}$ does not satisfy this condition since
\begin{align}
(\partial (S, V)).\get \; (\delta_S(\ell_{bad}.\pput \; (1,2) \; 3, (1,2))) &=
(\partial (S, V)).\get \; (\delta_S((0,3), (1,2)))\\
&= (\partial (S, V)).\get \; (0,3)\\
&= (0,3)\\
&\neq (\epsilon, 3)\\
&= \delta_V(3, 2)\\
&= \delta_V(3, \ell_{bad}.\get \; (1, 2))
\end{align}
Consequently $\ell_{bad}$ is not a valid SD-lens. On the
other hand, the lens $\ell_{good} : S \Leftrightarrow V$ defined by
\begin{align*}
\ell_{good}.\get \; (a, b) &= b\\
\ell_{good}.\pput \; (a, b) \; c &= (a, c)\\
\ell_{good}.\create \; b &= (0, b)
\end{align*}
is indeed a SD-lens since the its \pput component always restores the first
factor.

Up to this point we have not yet explained the role of inclusion functions. Now
assume that we have valid lenses $\ell_1$ and $\ell_2$ from source $S_i$ to
view $V_i$ for $i \in  \{1, 2\}$. One of the standard lens combinators is the
conditional combinator $\ell_1 \sqcup \ell_2$ which uses $\ell_1$ for $s \in
S_1$ and $v \in V_1$ and $\ell_2$ for $s \in S_2$ and $v \in V_2$. Now assume
that $s_1 \in S_1$ and $s_2 \in S_2$. When computing the difference between
$s_1$ and $s_2$, we wish to say that $\delta_{S_1 \sqcup S_2}(s_1, s_2) = s_1$
since $s_1$ and $s_2$ are in disjoing sets and are hence incomparable, but $s_1$
is in $S_1$ rather than in $\partial S_1$! Hence for any source $S$ we need to
embed $s$ as an update inside $\partial S$ and similarly for views, hence the
need for inclusion functions $i_S : S \longrightarrow \partial S$ and $i_V : V
\longrightarrow \partial V$. For inclusion functions $i_S$ and $i_V$, we
require a law similar to \cref{partialpreservesdelta} to hold:
$$(\partial (S, V)).\get \; (i_{S}(s)) = i_{V}(\ell.\get \; s)$$
\section{Stateful Delta Lenses}
\begin{definition}
Let $(S, \partial S, \delta_S, i_S)$ and $(V, \partial V, \delta_V, i_V)$ be
difference structures. A stateful delta-lens (SD-Lens) $\ell : (S, \partial S,
\delta_S, i_S) \Leftrightarrow (V, \partial V, \delta_V, i_V)$ is a pair
$(\ell, \partial (S, V))$ such that for all $s \in S$ and $v \in V$,
\begin{enumerate}
  \item
  $\ell : S \Leftrightarrow V$ is a lens,
  \item
  $\partial(S, V) : \partial S \Leftrightarrow \partial V$ is a lens,
  \item
  $(\partial (S, V)).\get \; (\delta_S(\ell.\pput \; s \; v, s)) = \delta_V(v,
  \ell.\get \; s)$
  \item
  $(\partial (S, V)).\get \; (i_{S}(s)) = i_{V}(\ell.\get \; s)$
\end{enumerate}

\end{definition}
\subsection{Boomerang Lens Combinators}
\subsubsection{Copy}
Let $P$ be any set. Define the $\ccopy(P) : P \Leftrightarrow P$ by
\begin{align}
\ccopy(P).\get \; s &= s\\
\ccopy(P).\pput \; s \; v &= v\\
\ccopy(P).\create \; v &= v
\end{align}
Then $\ccopy(P)$ is a lens. Now let $(S, \partial S, \delta_S, i_S)$ be a
difference structure.
\begin{claim}
$(\ccopy(S),  \ccopy(\partial S)) : (S, \partial S, \delta, i_S)
\Leftrightarrow (S, \partial S, \delta, i_S)$ is an SD-lens.
\end{claim}
\begin{proof}
First we need to show that $(\ccopy(\partial S)).\get \; (\delta(\ccopy(S).\pput
\; s \; v, s)) = \delta(v, \ccopy(S).\get \; s)$ which follows from the fact that
$$ (\ccopy(\partial S)).\get \; (\delta(\ccopy(S).\pput \; s
\; v, s)) = \delta(\ccopy(S).\pput \; s \; v, s) = \delta(v, \ccopy(S).\get
\; s) $$
Next we need to show that $(\ccopy(\partial S)).\get \; (i_{S}(s)) =
i_{S}(\ell.\get \; s)$, which follows from the fact that $(\ccopy(\partial
S)).\get \; (i_{S}(s)) = (\ccopy(\partial S)).\get \; (i_{S}(\ccopy(S).\get \;
s))$
\end{proof}
\subsubsection{Constant}
Let $P$ be any set. Define the \const lens $\const(P) : P \Leftrightarrow \{*\}$
by
\begin{align*}
\const.\get \; s &= *\\
\const.\pput \; s \; v &= s\\
\const.\create \; v &= p
\end{align*}
Now let $(S, \partial S, \delta, i_S)$ be a difference structure.
\begin{claim}
$(\const(S),  \const(\partial S)) : (S, \partial S, \delta_S, i_S)
\Leftrightarrow (\{*\}, \{*\}, \delta_{\{*\}}, i_{\{*\}})$ is an SD-Lens
\end{claim}
\begin{proof}
First we need to show that $(\const(\partial S)).\get \;
(\delta_S(\const(S).\pput \; s \; v, s)) = \delta_{\{*\}}(v, \const(S).\get \;
s)$ which follows from the fact that $ (\const(\partial S)).\get \;
(\delta_S(\const(S).\pput \; s \; v, s)) = * = \delta_{\{*\}}(v, \const(S).\get
\; s) $
Next we need to show that $(\const(\partial S)).\get \; (i_{S}(s)) =
i_{\{*\}}(\const(S).\get \; s)$ which follows from the fact that
$(\const(\partial S)).\get \; (i_{S}(s)) = * = i_{\{*\}}(s)$
\end{proof}

\subsubsection{Default}
Let $\ell : S \Leftrightarrow V$ be a lens. Let $f :
V \longrightarrow S$ be any function. Define the lens $\default(\ell, f) :
S \Leftrightarrow V$ by
\begin{align*}
\default(\ell, f).\get \; s &= \ell.\get \; s\\
\default(\ell, f).\pput \; s \; v &= \ell.\pput \; s \; v\\
\default(\ell, f).\create \; v &= \ell.\pput \; (f \; v) \; v
\end{align*}
Now let $(\ell, \partial (S, V)) : (S, \partial S, \delta_S, i_S) \Leftrightarrow
(V, \partial V, \delta_V, i_V)$ be an SD-Lens and $f: S \longrightarrow V$ be
any function.
\begin{claim}
$(\default(\ell, f), \partial (S, V))$ is an SD-lens.
\end{claim}
\begin{proof}
Follows from the fact that $(\ell, \partial (S, V))$ is an SD-lens.
\end{proof}
\subsubsection{Product}
Let $\ell_1 : S_1 \Leftrightarrow V_1$ and $\ell_2 : S_2 \Leftrightarrow V_2$ be
lenses. Define the lens $\ell_1 \times \ell_2 : (S_1 \times S_2) \Leftrightarrow
(V_1 \times V_2)$ by
\begin{align*}
(\ell_1 \times \ell_2).\get \; (s_1, s_2) &= (\ell_1.\get \; s_1, \ell_2.\get
\; s_1)\\
(\ell_1 \times \ell_2).\pput \; (s_1, s_2) \; (v_1, v_2) &= (\ell_1.\pput \; s_1
\; v_1, \ell_2.\pput \; s_2 \; v_2)\\
(\ell_1 \times \ell_2).\create \; (v_1, v_2) &= (\ell_1.\create \; v_1,
\ell_2.\create \; v_2)
\end{align*}
Given difference functions $\delta_S$ and $\delta_{S'}$ define $\delta_{S}
\times \delta_{S'}$ by $$(\delta_{S} \times \delta_{S'})((s_1, s_2), (s'_1,
s'_2)) = (\delta_S(s_1, s'_1), \delta_{S'}(s'_1, s'_2))$$
Given inclusion functions $i_S$ and $i_{S'}$ define $i_S \times i_{S'}$ by
$$(i_S \times i_{S'})(s, s') = (i_S(s), i_{S'}(s'))$$

Now let $(\ell_1, \partial (S_1, V_1)) : (S_1, \partial S_1, \delta_{S_1},
i_{S_1}) \Leftrightarrow (V_1, \partial V_1, \delta_{V_1}, i_{V_1})$ and $(\ell_2,
\partial (S_2, V_2)) : (S_2, \partial S_2, \delta_{S_2}, i_{S_2})
\Leftrightarrow (V_2, \partial V_2, \delta_{V_2}, i_{V_2})$ be SD-lenses. Define the lens
$$(\ell_1, \partial (S_1, V_1)) \times (\ell_2, \partial (S_2, V_2)) : (S_1 \times S_2,
\partial S_1 \times \partial S_2, \delta_{S_1} \times \delta_{S_2}, i_{S_1}
\times i_{S_2}) \Leftrightarrow (V_1 \times V_2, \partial V_1 \times \partial
V_2, \delta_{V_1} \times \delta_{V_2}, i_{V_1} \times i_{V_2})$$ by $(\ell_1,
\partial (S_1, V_1)) \times (\ell_2, \partial (S_2, V_2)) = (\ell_1 \times \ell_2,
\partial (S_1, V_1) \times \partial (S_2, V_2))$

\begin{claim}\label{productislens}
$(\ell_1 \times \ell_2, \partial (S_1, V_1) \times \partial (S_2, V_2))$ is an SD-lens.
\end{claim}
\begin{proof}
First need to show that
$$(\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \; (\delta((\ell_1 \times
\ell_2).\pput \; (s_1, s_2) \; (v_1, v_2), (s_1, s_2))) = \delta((v_1, v_2),
(\ell_1 \times \ell_2).\get \; (s_1, s_2))$$ which follows from the fact that
\begin{align*}
&(\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \; (\delta((\ell_1 \times
\ell_2).\pput \; (s_1, s_2) \; (v_1, v_2), (s_1, s_2)))\\
&= (\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \; (\delta((\ell_1.\pput \;
s_1 \; v_1, \ell_2.\pput \; s_2 \; v_2), (s_1, s_2))\\
&= (\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \; (\delta(\ell_1.\pput \;
s_1 \; v_1, s_1), \delta(\ell_2.\pput \; s_2 \; v_2, s_2))\\
&= ((\partial (S_1, V_1)).\get \; (\delta(\ell_1.\pput \;
s_1 \; v_1, s_1)), (\partial (S_2, V_2)).\get \; (\delta(\ell_2.\pput \; s_2 \;
v_2, s_2)))\\
&= (\delta(v_1, \ell_1.\get \; s_1), \delta(v_2, \ell_2.\get \; s_2))\\
&= \delta((v_1, v_2), (\ell_1.\get \; s_1, \ell_2.\get \; s_2))\\
&= \delta((v_1, v_2), (\ell_1 \times \ell_2).\get \; (s_1, s_2))
\end{align*}
Next, we need to show that $(\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \;
(i_{S_1 \times S_2}(s_1, s_2)) = i_{V_1 \times V_2}((\ell_1 \times
\ell_2).\get \; (s_1, s_2))$ which follows from the fact that
\begin{align*}
(\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \;
(i_{S_1 \times S_2}(s_1, s_2)) &= (\partial (S_1, V_1) \times \partial
\ell_2).\get \; (i_{S_1}(s_1), i_{S_2}(s_2))\\
&= ((\partial (S_1, V_1)).\get \; (i_{S_1}(s_1)), (\partial (S_2, V_2)).\get \;
(i_{S_2}(s_2)))\\
&= (i_{V_1}(\ell_1.\get \; s_1), i_{V_2}(\ell_2.\get \; s_2))\\
&= i_{V_1 \times V_2}(\ell_1.\get \; s_1, \ell_2.\get \; s_2)\\
&= i_{V_1 \times V_2}((\ell_1 \times \ell_2).\get \; (s_1, s_2))
\end{align*}
\end{proof}
\subsubsection{Sum}
Let $\ell_1 : S_1 \Leftrightarrow V_1$ and $\ell_2 : S_2 \Leftrightarrow V_2$ be
lenses. Define the lens
$\ell_1 \sqcup \ell_2 :
(S_1 \sqcup S_2) \Leftrightarrow (V_1 \sqcup V_2)$ by
\begin{align*}
(\ell_1 \sqcup \ell_2).\get \; s &=
\begin{cases}
\ell_1.\get \; s & \text{if } s \in S_1\\
\ell_2.\get \; s & \text{if } s \in S_2\\
\end{cases}\\
(\ell_1 \sqcup \ell_2).\pput \; s \; v&=
\begin{cases}
\ell_1.\pput \; s \; v& \text{if } s \in S_1, \; v \in V_1\\
\ell_2.\pput \; s \; v& \text{if } s \in S_2, \; v \in V_2\\
\ell_1.\create \; v & \text{if } s \in S_2, \; v \in V_1\\
\ell_2.\create \; v & \text{if } s \in S_1, \; v \in V_2\\
\end{cases}\\
(\ell_1 \sqcup \ell_2).\create \; v &=
\begin{cases}
\ell_1.\create \; v & \text{if } v \in V_1\\
\ell_2.\create \; v & \text{if } v \in V_2\\
\end{cases}
\end{align*}
Given difference functions $\delta_S, \delta_{S'}$ and inclusion functions $i_S,
i_{S'}$ define $\delta_{S} \sqcup \delta_{S'}$ and $i_{S} \sqcup i_{S'}$ by
$$(\delta_{S} \sqcup \delta_{S'})(x, y) =
\begin{cases}
\delta_S(x, y) & \text{if } x, y \in S\\
\delta_{S'}(x, y) & \text{if } x, y \in S'\\
i_S(x) & \text{if } x \in S, y \in S'\\
i_{S'}(x) & \text{if } x \in S', y \in S
\end{cases}$$
and
$$(i_S \sqcup i_{S'})(x) =
\begin{cases}
i_S(x) & \text{if } x \in S\\
i_{S'}(x) & \text{if } x \in S'
\end{cases}$$
Now let $(\ell_1, \partial (S_1, V_1)) : (S_1, \partial S_1, \delta_{S_1}, i_{S_1})
\Leftrightarrow (V_1, \partial V_1, \delta_{V_1}, i_{V_1})$ and $(\ell_2,
\partial (S_2, V_2)) : (S_2, \partial S_2, \delta_{S_2}, i_{S_2}) \Leftrightarrow
(V_2, \partial V_2, \delta_{V_2}, i_{V_2})$ be SD-lenses. Define the lens
$$(\ell_1, \partial (S_1, V_1)) \sqcup (\ell_2, \partial (S_2, V_2)) : (S_1 \sqcup S_2,
\partial S_1 \sqcup \partial S_2, \delta_{S_1} \sqcup \delta_{S_2}, i_{S_1}
\sqcup i_{S_2}) \Leftrightarrow (V_1 \sqcup V_2, \partial V_1 \sqcup \partial
V_2, \delta_{V_1} \sqcup \delta_{V_2}, i_{V_1} \sqcup i_{V_2})$$ by $(\ell_1,
\partial (S_1, V_1)) \sqcup (\ell_2, \partial (S_2, V_2)) = (\ell_1 \sqcup \ell_2,
\partial (S_1, V_1) \sqcup \partial (S_2, V_2))$
\begin{claim}
$\ell_1 \sqcup \ell_2$ is an SD-lens.
\end{claim}
\begin{proof}
We need to show that $(\partial (S_1, V_1) \sqcup \partial (S_2, V_2)).\get \;
(\delta((\ell_1 \sqcup \ell_2).\pput \; s \; v, s)) = \delta(v, (\ell_1 \sqcup
\ell_2).\get \; s)$. If $s \in S_1$ and $v \in V_1$ or $s \in S_2$ and $v \in
V_2$, then the result follows from the fact that $(\ell_1, \partial (S_1, V_1))$ and
$(\ell_2, \partial (S_2, V_2))$ are lenses. Otherwise, assume without loss of
generality that $s \in S_1$ and $v \in V_2$. Then
\begin{align*}
(\partial (S_1, V_1) \sqcup \partial (S_2, V_2)).\get \;
(\delta((\ell_1 \sqcup \ell_2).\pput \; s \; v, s)) &=
(\partial (S_1, V_1) \sqcup \partial (S_2, V_2)).\get \; (\delta(\ell_2.\create \; v,
s))\\
&= (\partial (S_1, V_1) \sqcup \partial (S_2, V_2)).\get \; (i_2(\ell_2.\create \; v))\\
&= (\partial (S_2, V_2)).\get \; (i_{S_2}(\ell_2.\create \; v))\\
&= i_{V_2}(\ell_2.\get \; (\ell_2.\create \; v))\\
&= i_{V_2}(v)\\
&= \delta(v, (\ell_1 \sqcup \ell_2).\get \; s)
\end{align*}
\end{proof}
\subsubsection{Positional Iteration}
Let $\ell: S \Leftrightarrow V$ be a lens. Define the lens $\pos(\ell):
S^* \Leftrightarrow V^*$ by
\begin{align*}
\pos(\ell) .\get \; (s_1, \ldots, s_n) &= (\ell.\get \; s_1) \cdot \ldots \cdot
(\ell.\get \; s_n)\\
\pos(\ell) .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n) &= s'_1 \cdot
\ldots \cdot s'_n\\
\text{ where } s'_i &= \begin{cases}
\ell.\pput \; s_i \; v_i & \text{if } 1 \leq i \leq \min\{m, n\}\\
\ell.\create \; v_i & \text{if } \min\{m, n\} < i \leq n
\end{cases}\\
\pos(\ell).\create \; (v_1, \ldots, v_n) &= (\ell.\create \; v_1) \cdot \ldots
\cdot (\ell.\create \; v_n)
\end{align*}
Now let $\delta_S$ be a difference function and $i_S$ and inclusion function.
We define the positional inclusion function $\pos(i_S)(s_1, \ldots, s_m) =
(i_S(s_1), \ldots, i_S(s_m))$. Let $(x_1, \ldots, x_m), (y_1, \ldots, y_n) \in
S^*$. We define the positional difference function $\pos(\delta)((x_1, \ldots,
x_m), (y_1, \ldots, y_n))$ to be the list given by $\delta(x_i, y_i)$ for $1
\leq i \leq \min\{m, n\}$, except that we omit $\delta(x_i, y_i)$ in position
$i$ if $x_i = y_i$. For $\min\{m, n\} < i \leq m$ we add $i_S(x_i)$ at position
$i$. Finally we collapse the list to get rid of the empty spaces left after the
omissions.

For instance suppose that $S$ is the set of all letters in the alphabet and
$\partial S = S^*$ with $$\delta(x, y) =
\begin{cases}
\epsilon & \text{if } x = y\\
x & \text{if } x \neq y
\end{cases}$$
and $i_S(x) = x$. Then $\pos(\delta_S)((a, b, c, d, e), (a, c, b, d)) = (b, c,
e)$ where we match and remove the first $a$ and the fourth $d$ in $(a, b, d, c,
e)$. Since $b \neq c$ we retain the $b$ and $c$ in the second and third
positions, and since $e$ is unmatched, we include $e$ via the inclusion
function $i_S(e) = e$. Finally we collapse the list to get rid of the two empty
spaces created.

Now let $(\ell, \partial (S, V)) : (S, \partial S, \delta_{S}, i_{S})
\Leftrightarrow (V, \partial V, \delta_{V}, i_{V})$ be an SD-lens.
\begin{claim}\label{positionalisvalid}
$(\pos(\ell), \pos(\partial (S, V))) : (S^*, \partial S^*, \pos(\delta_{S}),
\pos(i_{S})) \Leftrightarrow (V^*, \partial V^*, \pos(\delta_{V}), \pos(i_{V}))$
is an SD-lens.
\end{claim}
\begin{proof}
First, we need to show that
\begin{multline}
(\pos(\partial (S, V)).\get \;
((\pos(\delta_S))(\pos(\ell).\pput \; (s_1, \ldots, s_n) \; (v_1, \ldots,
v_m), (s_1, \ldots, s_n))) \\
= (\pos(\delta_V))((v_1, \ldots, v_m), (\pos(\ell)).\get \; (s_1, \ldots, s_n))
\end{multline}
Let $1 \leq i \leq \min\{m, n\}$. Then $v_i = \ell.\get \; s_i$ if and only if
$\ell.\pput \; s_i \; v_i = s_i$, so $\pos(\delta_V)$ omits $\delta_V(v_i,
\ell.\get \; s_i)$ in position $i$ if and only if $\pos(\delta_S)$ also omits
$\delta_S(\ell.\pput \; s_i \; v_i, s_i)$ in position $i$. For $\min\{m,n\}
< i \leq m$, then $\pos(\delta_V)$ adds $i_V(v_i)$ in position $i$, and
$\pos(\delta_S)$ adds $i_S(\ell.\create \; v_i)$ in position $i$, which gives
the desired result since
$$(\partial (S, V)).\get \; (i_S(\ell.\create \; v_i)) = i_V(\ell.\get \;
(\ell.\create \; v_i)) = i_V(v_i)$$
Next we need to show that
$$(\pos(\partial (S, V))).\get \; ({\pos(i_S)}(s_1, \ldots, s_m)) =
\pos(i_V)(\pos(\ell).\get \; (s_1, \ldots, s_m))$$
which follows from the fact that
\begin{align*}
(\pos(\partial (S, V))).\get \; ({\pos(i_S)}(s_1, \ldots, s_m)) &=
(\pos(\partial (S, V))).\get \; ({i_S}(s_1), \ldots, {i_S}(s_m))\\
&= (\ell.\get \; (i_S(s_1)), \ldots, \ell.\get \; (i_S(s_m)))\\
&= (i_V(\ell.\get \; s_1), \ldots, i_V(\ell.\get \; s_m))\\
&= \pos(i_V)(\ell.\get \; s_1, \ldots, \ell.\get \; s_m)\\
&= \pos(i_V)(\pos(\ell).\get \; (s_1, \ldots, s_m))
\end{align*}
\end{proof}
\subsubsection{Matching Lenses}
Let $X$ be a set. Let $x = (x_1, \ldots, x_m) \in X^*$. Define $\length(x) = m$
Now $(y_1, \ldots, y_n) \in X^*$. Let $A_{x, y} \subseteq [m]$. We call any
injective function $\varphi_{x, y} : A_{x, y} \longrightarrow [n]$ a {\em
matching} from $x$ to $y$.

Again let $x = (x_1, \ldots, x_m) \in X^*, (y_1, \ldots, y_n) \in X^*$ and
$\varphi$ a matching from $x$ to $y$. Let $\delta$ be a difference function on
$X$. Define $\delta^*_{\varphi}(x, y)$ to be the list given by $\delta(x_i,
y_{\varphi(i)})$ for $i \in A$, except that we omit $\delta(x_i,
y_{\varphi(i)})$ in position $i$ if $x_i = y_{\varphi(i)}$. If $i \not \in A$
we add $i_S(x_i)$ at position $i$. Finally we collapse the list to get rid of
the empty spaces.

Going back to our previous example, let $x = (a, b, c, d, e)$ and $y = (a, c, b,
d)$, and assume that $\varphi : \{1, 2, 3, 4\} \longrightarrow [4]$ is
given by $$\varphi(1) = 1, \; \varphi(2) = 3, \; \varphi(3) = 2, \text{ and }
\varphi(4) = 4$$ Then $\delta_{\varphi}^*((a, b, c, d, e), (a, f, b, d)) = (c,
e)$ Since $x_1 = y_{\varphi(1)} = a$, $x_2 = y_{\varphi(2)} = b$ and $x_{4} =
y_{\varphi(4)} = d$, $a, b$ and $d$ will be omitted from the list at
positions 1, 2 and 4. Also since $\delta(c, f) = c$ then $c$ will be included in
the list at position 3 and since $e$ is unmatched, $i(e) = e$ will be
included in the list at position 5. Finally we collapse the list to get rid of
the three empty spaces created.

Now let $M = \{\varphi_{x, y} \; | \; x, y \in X\}$ be a set of matchings
between $x$ and $y$ for each $x, y \in X$. We define the difference function
$\delta^*_M$ on $X^*$ by $\delta^*_{M}(x, y) = \delta^*_{\varphi_{x, y}}$. Let
$(\ell, \partial (S, V)) : (S, \partial S, \delta_{S}, i_{S}) \Leftrightarrow
(V, \partial V, \delta_{V}, i_{V})$ be an SD-lens. Let $M_V = \{\varphi_{x, y}
\; | \; x, y \in V\}$ be a set of matchings between $x$ and $y$ for each $x, y
\in V$. For $s = (s_1, \ldots, s_m), s' = (s'_1, \ldots, s'_n) \in S$, let
$\varphi_{s, s'} = \varphi_{((\ell.\get \; s_1), \ldots, (\ell.\get \; s_m)),
((\ell.\get \; s'_1), \ldots, (\ell.\get \; s'_n))}$. Let $M_S = \{\varphi_{s,
s'} \; | \; s, s' \in S\}$.

Define the lens $\match(\ell, M_V) : S^* \Leftrightarrow V^*$ by
\begin{align*}
\match(\ell, M_V) .\get \; (s_1, \ldots, s_n) &= (\ell.\get \; s_1) \cdot
\ldots \cdot (\ell.\get \; s_n)\\
\match(\ell, M_V) .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n) &= (s'_1
, \ldots , s'_n)\\
\text{ where } s'_i &= \begin{cases}
\ell.\pput \; s_{\psi(i)} \; v_i & \text{if } i \in Dom(\psi)\\
\ell.\create \; v_i & \text{if } i \not \in Dom(\psi)
\end{cases}\\
\text{ and } \psi &= \varphi_{(v_1, \ldots, v_n), ((\ell.\get \; s_1), \ldots,
(\ell.\get \; s_m))}\\
\match(\ell, M_V).\create \; (v_1, \ldots, v_n) &= (\ell.\create \; v_1) \cdot
\ldots \cdot (\ell.\create \; v_n)
\end{align*}
\begin{claim}
$(\match(\ell, M_V), \pos(\partial (S, V))) : (S^*, \partial S^*, \delta^*_{M_S},
\pos(i_{S})) \Leftrightarrow (V^*, \partial V^*, \delta^*_{M_V}, \pos(i_{V}))$
is an SD-lens.
\end{claim}
\begin{proof}
Observe that the matching used to compute the difference on the source side is
the same as the matching used to compute the matching on the view side, hence we
may take the update lens to be position and proceed with the rest of the proof
as in \cref{positionalisvalid}.
\end{proof}
The \match combinator takes an SD-lens $(\ell, \partial (S, V)) : (S, \partial
S, \delta_{S}, i_{S}) \Leftrightarrow (V, \partial V, \delta_{V}, i_{V})$ and a
set $M_V$ of matchings $V$. If $S, \partial S$ and $V, \partial V$ have an
appropriate notion of size defined on them, then we can define canonical
matchings which minimize the size of the update according to some alignment
strategy. We thus defined a {\em normed difference structure} as follows:
\begin{definition}
A normed set is a pair $(X, |\cdot|)$ where $X$ is a set and $|\cdot| : X
\longrightarrow \mathbb{R}^{\geq 0}$
\end{definition}
\begin{definition}
A difference structure $(X, \partial X, \delta_X, i_X)$ is a normed differene
structure if $X$ and $\partial X$ are normed sets satisfing
\begin{enumerate}
  \item
  $|\delta_X(x, x)| = 0$,
  \item
  $|\delta_X(x, x')| \leq |x|$, and
  \item
  $|i_X(x)| = |x|$
\end{enumerate}
\end{definition}
The first axiom says that the size of the difference between $x$ and itself is
0, the second axiom says that the size of the ordered difference between $x$
and $x'$ is at most the size of $x$, and third axiom says that the size of $x$
when it is included as an update in $\partial X$ is equal to the size of $x$.

Now given a view $V$ we may define the set of matchings $M_{POS}$ by 
$$M_{POS} = \{\varphi_{v, v'} \; | \; v = (v_1, \ldots, v_m), v' = (v'_1,
\ldots, v'_n), \varphi_{v, v'}(i) = i \text{ for }1 \leq i \leq \min\{m, n\}\}$$
Then $M_{POS}$ is an optimal set of matchings for a positional alignment
strategy.

For the non-crossing strategy, we want to our matchings to be increasing rather
than positional. Hence we define the set of matchings $M_{NC}$ by
$$M_{NC} = \{\varphi_{v, v'} \; | \; v, v' \in V, \varphi_{v, v'} \text{ is
increasing}, |\delta^*_{\varphi_{v, v}}(v, v')| \text{ is minimal among all
matchings from }v \text{ to }v'\}$$
For any $v, v'$, there may be multiple matchings which for which the size of the
difference is minimal, in which case we pick a canonical one. Then $M_{NC}$ is
an optimal set of matchings for a non-crossing alignment strategy.

Finally, for a min-edit distance strategy, we simply remove the condition that
the matchings should be increasing. Hence we define the set of matchings
$M_{MED}$ by
$$M_{MED} = \{\varphi_{v, v'} \; | \; v, v' \in V, |\delta^*_{\varphi_{v,
v'}}(v, v)| \text{ is minimal among all matchings from }v \text{ to }v'\}$$
Again, for any $v, v'$, there may be multiple matchings which for which the size
of the difference is minimal, in which case we pick a canonical one. Then
$M_{MED}$ is an optimal set of matchings for a min-edit distance alignment
strategy.
\subsubsection{Composition}
Let $\ell_1 : S \Leftrightarrow U$ and $\ell_2 : U \Leftrightarrow V$ be lenses.
Define the lens $\ell_2 \circ \ell_1 : S \Leftrightarrow V$ by
\begin{align*}
(\ell_2 \circ \ell_1).\get \; s &= \ell_2.\get \; (\ell_1.\get \; s)\\
(\ell_2 \circ \ell_1).\pput \; s \; v &= \ell_1.\pput \; s \; (\ell_2.\pput \;
(\ell_1.\get \; s) \; v) \\
(\ell_2 \circ \ell_1).\create \; v &= \ell_1.\create \; (\ell_2.\create \; v)
\end{align*}
Now let $(\ell_1, \partial (S_1, V_1)) : (S,
\partial S, \delta_S, i_{S}) \Leftrightarrow (U,
\partial U, \delta_U, i_{U})$ and $\ell_2 : (U,
\partial U, \delta_U, i_{U}) \Leftrightarrow (V,
\partial V, \delta_V, i_{V})$ be SD-lenses.
\begin{claim}
$(\ell_2 \circ \ell_1, \partial (S_2, V_2) \circ \partial (S_1, V_1))$ is an SD-lens.
\end{claim}
\begin{proof}
We need to show that
\begin{align}
\partial (S_2, V_2) \circ \partial (S_1, V_1)).\get \; (\delta(\ell_2 \circ
\ell_1).\pput \; s \; v, s)&= \delta(v, (\ell_2 \circ \ell_1).\get \; s)
\text{, and}\\
(\partial (S_2, V_2) \circ \partial (S_1, V_1)).\get \; (i_{S}(s)) &=
i_{V}((\ell_2 \circ \ell_1).\get \; s)
\end{align}

For the first equation, observe that
\begin{align*}
&(\partial (S_2, V_2) \circ \partial (S_1, V_1)).\get \; (\delta((\ell_2 \circ
\ell_1).\pput \; s \; v, s))\\
&=(\partial (S_2, V_2) \circ \partial (S_1, V_1)).\get \; (\delta(\ell_1.\pput
\; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; v), s))\\
&= (\partial (S_2, V_2)).\get \; ((\partial (S_1, V_1)).\get \; (\delta(\ell_1.\pput \;
s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; v), s)))\\
&= (\partial (S_2, V_2)).\get \; (\delta(\ell_2.\pput \; (\ell_1.\get \; s) \; v,
\ell_1.\get \; s))\\
&= \delta(v, \ell_2.\get \; (\ell_1.\get \; s))\\
&= \delta(v, (\ell_2 \circ \ell_1).\get \; s)\\
\end{align*}
For the second equation, observe that
\begin{align*}
(\partial (S_2, V_2) \circ \partial (S_1, V_1)).\get \; (i_{S}(s)) &= 
(\partial (S_2, V_2)).\get \; ((\partial (S_1, V_1)).\get \; (i_{S}(s)))\\
&= (\partial (S_2, V_2)).\get \; (i_{U}(\ell_1.\get \; s))\\
&= i_{V}(\ell_2.\get \; (\ell_1.\get \; s))\\
&= i_{V}((\ell_2
\circ \ell_1).\get \; s)
\end{align*}
This completes the proof.
\end{proof}

%% Acknowledgments
\begin{acks}                            %% acks environment is optional
%% contents suppressed with 'anonymous'
%% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
%% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
%% acknowledge financial support and will be used by metadata
%% extraction tools.
This material is based upon work supported by the
\grantsponsor{GS100000001}{National Science
Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
No.~\grantnum{GS100000001}{nnnnnnn} and Grant
No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
conclusions or recommendations expressed in this material are those
of the author and do not necessarily reflect the views of the
National Science Foundation.
\end{acks}

\section{Related Work}
In his thesis \cite{foster2009bidirectional}, Foster observed that while it would be
ideal for the lens laws to force the \pput component of a lens to always have
``minimal'' side effects, even stating this condition seems to require
building a notion of what constitutes an update into the semantics---we need to
be able to compare two updates to determine which one has a ``smaller'' effect.

This idea of lenses carrying a notion of updates and update propagation has been
explored in various ways. First, various categorical approaches have been
explored under the name of {\em delta lenses}
\cite{diskin2011state,johnson2017symmetric,pacheco2012delta,johnson2012lenses}
Delta lenses are lenses defined in an appropriate category, where sources and
views are the objects of the category, and the arrows are deltas (or updates)
between the objects of the category. In this setting, lens transformations are
functors which map objects to objects as well as deltas to deltas. Categorical
delta lenses are thus similar to the SD-lenses described in this paper in that
they map states to states as well as updates to updates. The main difference
between SD-lenses and categorical delta lenses is that SD-lenses are
set-based rather than categorical and arguably simpler to understand. Moreover,
only \cite{pacheco2012delta} defines a fairly rich set of combinators for delta
lenses over inductive types in the same vane as SD-lenses.

{\em Alignment lenses} have also been explored in tackling the issue of update
propagation in \cite{bohannon2008boomerang,barbosa2010matching}. Prior to this
work, lenses could only perform updates positionally on data lists, but
alignment lenses can match chunks using an optimal matching with
respect to a certain alignment strategy such as a non-crossing strategy or
min-edit distance strategy. While alignment lenses solved the problem of
matching chunks in a way that causes minimal change with respect to some
alignment strategy, they do not address the general problem of specifying how
updates should be propagated for an arbitrary lens. The SD-lenses introduced in
this paper do give a fairly general method for specifying and updates and update
propagation.

{\em Edit lenses} \cite{hofmann2012edit} are another approach to tackling the
issue of update propagation. Edit lenses work with descriptions of changes to
structures, rather than with the structures themselves. These lenses separate
the description of edits from the action of applying an edit to a state, with
the action of applying an edit to a state modelled by a monoid action, a design
choice which guarantees compositionality of edits. SD-lenses differ from edit
lenses in that edit lenses are not stateful---the only information passed to the
lens is the intended update, hence edit lenses pass around less information. On
the other hand, SD-lenses are again arguably simpler since they are stateful
and are also built on top of traditional lenses which are ubiquitous in the
bidirectional programming literature.
\bibliographystyle{plain}
\bibliography{local}

%% Appendix
%%\appendix
%%\section{Appendix}

\end{document}

### 2018-02-14 papers/SDLenses/main.tex
Made some edits

@@
 to the view is non-trivial. Consequently, a lens may be valid even though it
-does not propagate non-trivial updates made to the view back into the source in
-the manner intended. For example, a small update to the view can result
-in a large update to the source, such as when a lens updates records
-positionally even though the source and view are order agnostic. This paper
-introduces {\em stateful delta lenses} or SD-lenses as a solution to this
-shortcoming of traditional lenses. SD-lenses are lenses augmented with extra
-information that specifies what kind of updates are valid for a given lens, and
-how these updates should be propagated by the lens, and hence enable the
-programmer to specify more exactly the desired behaviour of a lens.
+does not propagate non-trivial updates in the manner intended. For example, a
+small update to the view can result in a large update to the source, such as
+when a lens updates records positionally even though the source and view are
+order agnostic. This paper introduces {\em stateful delta lenses} or SD-lenses
+as a solution to this shortcoming of traditional lenses. SD-lenses are lenses
+augmented with extra information that explicitly specifies the updates that can
+be made for a given lens and how these updates should be propagated. SD-lenses
+explicitly specify the intended behaviour of a lens hence enable programmers to
+give much stronger guarantees even when the update to a view is non-trivial.
 \end{abstract}
@@
 
-Lenses are designed to perfom these kinds of computations. If $\ell$ is
-a suitably defined lens for the bidirectional tranfsormation that we
+Lenses are designed to perfom these kinds of bidirectinoal transformations. If
+$\ell$ is a suitably defined lens for the tranfsormation that we
 have just described, then the \get component of the lens performs the
@@
 \end{align*}
-The original definition of lenses in \cite{foster2007combinators} involved only the \get
-and \pput functions, but a \create function was later added in
-\cite{foster2009bidirectional} to handle the special case where we need to construct a
-source from a view, but we have no source to use as the original. \sam{Perhaps
-we should include this in the footnotes?}
+The original definition of lenses in \cite{foster2007combinators} involved only
+the \get and \pput functions, but a \create function was later added in
+\cite{foster2009bidirectional} to handle the special case where we need to
+construct a source from a view, but we have no source to use as the original.
 
@@
 
-While these two laws give some guarantees on how the \get and \pput components
-of a lens interact, they are not strong enough to guarantee that the
-updates made to a view are indeed the updates reflected back into the source.
-This is because the GETPUT law only says that the \pput component must return
-the original source unmodified whenever the update to the view is a no-op; it
-does not say what the \pput component should do with a modified view.
-Consequently, a non-trivial update to a view could lead to an undesired update
-to the source.
+While these two laws give some guarantees on how the \get, \pput and \create
+components of a lens interact, they do not specify the behaviour of the \pput
+component when the update to the view is non-trivial, because the GETPUT law
+only says that the \pput component must return the original source unmodified
+whenever the update to the view is a no-op Consequently, a non-trivial update
+to a view could lead to an undesired update to the source.
 
@@
 that the view $V$ for $S$ consists of a single integer. Further, suppose that we
-wish for the \get of any lens from $S$ to $V$ to simply project away
-the first factor:
-$\ell.\get \; (a, b) = b$. Suppose further that we wish to force the \pput
-function of any lens $\ell$ defined on the data to restore the factor projected
-away by the \get function: that is, we wish to enforce that for
-all integers $a, b, c$, $\pi_1(\ell.\pput \; (a, b) \; c) = a$, where $\pi_1$ is
-the projection on the first factor.
+wish for the \get component of any lens from $S$ to $V$ to simply project away
+the first factor:that is for any $\ell : \mathbb{Z} \times \mathbb{Z}
+\Leftrightarrow \mathbb{Z}$, $\ell.\get \; (a, b) = b$ for all integers $a, b$.
+Suppose further that we wish to force the \pput function of any lens $\ell$
+defined on the data to restore the factor projected away by the \get function:
+that is, we wish to enforce that for all integers $a, b, c$, $\pi_1(\ell.\pput
+\; (a, b) \; c) = a$, where $\pi_1$ is the projection on the first factor.
 
-Now consider the lens $\ell_{bad}$ defined by $\ell_{bad}.\get \; (a, b) = b$
-and
+Now consider the lens $\ell_{bad}$ defined by
 \begin{align*}
@@
 
-In this paper, we introduce {\em stateful delta lenses} or SD-lenses, which are
-lenses augmented with extra information that specifies what kind of updates are
-valid for a given lens, and how these updates should be propagated.
+SD-lenses are lenses augmented with extra information that explicitly specifies
+the updates that can be made for a given lens and how these updates should be
+propagated.
 
@@
 Next we define a lens $\partial(S, V) : \partial S \Leftrightarrow \partial V$
-which describes how updates should be propagated. For this example, we let
+which describes how updates should be propagated. For this example, we
 define $\partial(S,V)$ to simply be the \ccopy lens from $\partial S$ to
@@
 
-For example, observe that $\ell_{bad}$ does not satisfy this condition since
-\begin{align}
+This condition enable us to rule out lenses which do not propagate updates
+correctly. For example, observe that $\ell_{bad}$ does not satisfy
+\cref{partialpreservesdelta} since
+\begin{align*}
 (\partial (S, V)).\get \; (\delta_S(\ell_{bad}.\pput \; (1,2) \; 3, (1,2))) &=
@@
 &= \delta_V(3, \ell_{bad}.\get \; (1, 2))
-\end{align}
+\end{align*}
 Consequently $\ell_{bad}$ is not a valid SD-lens. On the
@@
 \end{align*}
-is indeed a SD-lens since the its \pput component always restores the first
-factor.
+is indeed a valid SD-lens since the its \pput component always restores the
+first factor.
 
@@
 $s_1$ and $s_2$, we wish to say that $\delta_{S_1 \sqcup S_2}(s_1, s_2) = s_1$
-since $s_1$ and $s_2$ are in disjoing sets and are hence incomparable, but $s_1$
-is in $S_1$ rather than in $\partial S_1$! Hence for any source $S$ we need to
-embed $s$ as an update inside $\partial S$ and similarly for views, hence the
-need for inclusion functions $i_S : S \longrightarrow \partial S$ and $i_V : V
-\longrightarrow \partial V$. For inclusion functions $i_S$ and $i_V$, we
-require a law similar to \cref{partialpreservesdelta} to hold:
-$$(\partial (S, V)).\get \; (i_{S}(s)) = i_{V}(\ell.\get \; s)$$
+since $s_1$ and $s_2$ are in disjoint sets and are hence incomparable, but $s_1$
+is in $S_1$ rather than in $\partial S_1$ and hence is not an update! Hence for
+any source $S$ we need to embed $s$ as an update inside $\partial S$ and
+similarly for views, hence the need for inclusion functions $i_S : S
+\longrightarrow \partial S$ and $i_V : V \longrightarrow \partial V$. For
+inclusion functions $i_S$ and $i_V$, we require a law similar to
+\cref{partialpreservesdelta} to hold: $$(\partial (S, V)).\get \; (i_{S}(s)) =
+i_{V}(\ell.\get \; s)$$
 \section{Stateful Delta Lenses}

### 2018-02-15 papers/SDLenses/main.tex
Made changes

@@
 \usepackage{amsmath, amssymb, amsthm, enumerate, graphicx, centernot, tikz,
-array, tikz-cd, extarrows, cleveref, mathrsfs, mathtools, stmaryrd, listings}
+array, tikz-cd, extarrows, cleveref, mathrsfs, mathtools, stmaryrd, listings,
+xspace}
 
@@
 
-\newcommand{\kw}[1]{\ensuremath{\mathsf{#1}}}
-\newcommand{\get}{\ensuremath{\kw{get}}}
-\newcommand{\pput}{\ensuremath{\kw{put}}}
-\newcommand{\create}{\ensuremath{\kw{create}}}
-\newcommand{\ccopy}{\ensuremath{\kw{copy}}}
-\newcommand{\const}{\ensuremath{\kw{const}}}
-\newcommand{\default}{\ensuremath{\kw{default}}}
-\newcommand{\pos}{\ensuremath{\kw{posit}}}
-\newcommand{\length}{\ensuremath{\kw{length}}}
-\newcommand{\match}{\ensuremath{\kw{match}}}
+\newcommand{\kw}[1]{\ensuremath{\mathsf{#1}}\xspace}
+\newcommand{\get}{\ensuremath{\kw{get}}\xspace}
+\newcommand{\pput}{\ensuremath{\kw{put}}\xspace}
+\newcommand{\create}{\ensuremath{\kw{create}}\xspace}
+\newcommand{\ccopy}{\ensuremath{\kw{copy}}\xspace}
+\newcommand{\const}{\ensuremath{\kw{const}}\xspace}
+\newcommand{\default}{\ensuremath{\kw{default}}\xspace}
+\newcommand{\pos}{\ensuremath{\kw{posit}}\xspace}
+\newcommand{\length}{\ensuremath{\kw{length}}\xspace}
+\newcommand{\match}{\ensuremath{\kw{match}}\xspace}
 
@@
 explored under the name of {\em delta lenses}
-\cite{diskin2011state,johnson2017symmetric,pacheco2012delta,johnson2012lenses}
-Delta lenses are lenses defined in an appropriate category, where sources and
+\cite{diskin2011state,diskin2011asymmetric,johnson2017symmetric,pacheco2012delta,johnson2012lenses} Delta lenses are lenses defined in an appropriate category, where sources and
 views are the objects of the category, and the arrows are deltas (or updates)
@@
 delta lenses are thus similar to the SD-lenses described in this paper in that
-they map states to states as well as updates to updates. The main difference
-between SD-lenses and categorical delta lenses is that SD-lenses are
-set-based rather than categorical and arguably simpler to understand. Moreover,
-only \cite{pacheco2012delta} defines a fairly rich set of combinators for delta
-lenses over inductive types in the same vane as SD-lenses.
+they map states to states as well as updates to updates.
+
+Our SD-lenses are set-based version of the delta lenses introduced by Pacheco et
+al in \cite{pacheco2012delta}. Pacheco et al. define a delta lens $\ell$ (d-lens
+for short), denote by $\ell: SA \trianglerighteq_{\blacktriangle} VB$ to be a
+bidirectional transformations that comprises four total functions:
+\begin{align*}
+\get &: SA \longrightarrow SB\\
+\pput &: \forall(v, s) : VB \times SA. (v \triangle \get \;s) \longrightarrow
+SA\\
+\get_{\blacktriangle} &: \forall\{s':SA, s:SA\}. (s' \triangle s)
+\longrightarrow (\get \; s \triangle \get \; s)\\
+\pput_{\blacktriangle} &:\forall\{(v, s):VB \times SA\},d:(v \triangle \get
+\;s).(\pput \; (v, s) \; d) \triangle s
+\end{align*}
+The d-lens is called {\em well-behaved} iff it satisfies the following
+propertes:
+\begin{align*}
+\get \; (\pput \; (v, s) \; d) = v \tag{PUTGET}\\
+\pput \; (\get \; s, s) \; id = s \tag{GETPUT}\\
+\get_{\blacktriangle} \; (\pput_{\blacktriangle} \; d) = d
+\tag{$\text{PUTGET}_{\blacktriangle}$}\\
+\pput_{\blacktriangle} \; id = id \tag{$\text{PUTGET}_{\blacktriangle}$}
+\end{align*}
+In the above definition, the state-based component of the d-lens is given by
+the functions get, that computes a view of a source value, and put, that takes
+a pair containing a modified view and an original source, together with a delta
+from the modified view to the original view, and returns a new modified value.
+The delta-based function $\get_{\blacktriangle}$ translates a source delta into
+a delta between views produced by $\get$, and $\pput_{\blacktriangle}$ receives
+a view delta and computes a delta from the new source produced by put to the
+original source. Properties PUTGET and GETPUT are the traditional state-based
+ones: view-to-view roundtrips preserve view modifications; and $\pput$ must
+preserve the original source for identity updates.
+$\text{PUTGET}_{\blacktriangle}$ and $\text{PUTID}_{\blacktriangle}$ denote
+similar laws on deltas: view-to-view roundtrips preserve view updates; and
+$\text{PUTID}_{\blacktriangle}$ must preserve identity updates.
+
+SD-lenses are essentially equivalent to well-behaved d-lenses in the sense of
+Pacheco et al.
+by mapping the state-based and delta based components of an SD-lens to
+their respective state and delta based components in a well-behaved d-lens.
+The axiom $\text{PUTID}_{\blacktriangle}$ combined with the dependent type of
+$\pput_{\blacktriangle}$ hence implies the axiom
+$$(\partial (S, V)).\get \; (\delta_S(\ell.\pput \; s \; v, s)) = \delta_V(v,
+\ell.\get \; s)$$ 
+for SD-lenses. Pacheco et al. also observe that their formulation is equivalent
+to the well-behaved d-lenses defined by Diskin et al. in
+\cite{diskin2011asymmetric}, hence our SD-lenses are essentially equivalent to
+these lenses as well.
 

### 2018-02-16 papers/QuotientLenses/main.tex
a few intro edits; checkpoint

@@
 
-Oftentimes, different applications store the same data in custom formats. For
+Oftentimes, the same data is stored in a variety different formats. For
 example, \bibtex{} and EndNote both store bibliography data; spreadsheet data
-can be represented as a CSV and as a TSV; web apis can return data as a JSON or
-XML object. Synchronizing these applications amounts to synchronizing the data
-formats. \emph{Lenses} comprise such synchronization functions, and the ``lens
-laws'' provide restrictions on how these synchronization functions interoperate.
-Bidirectional languages like Boomerang~\cite{boomerang} and
-BiGul~\cite{DBLP:conf/pepm/KoZH16} use a single expression to encode all of the
-lens's synchronization functions at the same time.
-
-However, these lens laws are too restrictive for many important transformations.
+can be represented as a CSV file or a TSV file; web APIs can return data as a
+JSON or
+XML object. When different applications manipulate data in different formats,
+synchronizing the applications amounts to synchronizing their respective data
+formats. One way to synchronize data formats is to use \emph{lenses}, \emph{i.e.,} programs that represent pairs of functions for translating data from
+format A to B (the \emph{get} direction) and back again (the \emph{put}
+direction).  These lenses, when defined using domain-specific
+languages like Boomerang~\cite{boomerang} or
+BiGul~\cite{DBLP:conf/pepm/KoZH16}, typically come with a variety of guarantees,
+the \emph{lens laws}, that help ensure put and get functions are effective
+inverses of one another.
+%comprise such synchronization functions, and the ``lens
+%laws'' provide restrictions on how these synchronization functions interoperate%.
+%Bidirectional languages like Boomerang~\cite{boomerang} and
+%BiGul~\cite{DBLP:conf/pepm/KoZH16} use a single expression to encode all of the
+%lens's synchronization functions at the same time.
+
+One way to classify lenses is by the laws they obey.  Amongst the most
+restrictive lenses are the \emph{bijective lenses}.  These lenses demand
+that get and put functions for a bijection, meaning no information
+is lost when one translates from data format to another.  While this
+strong guarantee can be useful in some contexts, it is more often the case
+that some of the information in one or both of the data formats is superfluous.
+For instance, one typically does not care to preserve the amount of
+whitespace between characters.  In other circumstances, capitalization
+or the order of fields in a record or items in a list may be unimportant.
+%
+%bijective lenses have
+%strong correctness guarantees, they are often
+%However, these lens laws are too restrictive for many important transformations.
 In particular, these lens laws do not permit transformations that put the data
-in canonical forms. In \bibtex{}, an author can be specified as \cd{lastname,
-firstname} and \cd{firstname lastname}. While it is important to allow authors
-to be input in both ways, we would like these authors to only be output as
-\cd{firstname lastname}, but the lens laws prevent such functions.
-{\em Quotient lenses}~\cite{quotientlenses} are a generalization of lenses where
-each of the data formats is quotiented by some equivalence relation and the lens
-laws are taken modulo equivalence. This allows lenses to be simpler and more
-flexible. For example, inessential details like whitespace and orderings of
-fields can be ignored when writing the transformations between the formats.
+in canonical forms.
+For instance, to return to our bibliography example, in \bibtex{},
+an author can be specified as
+``\cd{lastname, firstname}'' or as ``\cd{firstname lastname}''---it does not matter which of the
+two formats is used.
+%While it is important to allow authors
+%to be input in both ways, we would like these authors to only be output as
+%\cd{firstname lastname}, but the lens laws prevent such functions.
+{\em Quotient lenses}~\cite{quotientlenses} loosen many of the restrictions
+of bijective lenses by introducing an equivalence relation and considering
+bijections modulo that relation.  Such quotient lenses make it possible
+to align many more data sources than would otherwise be possible by
+ignoring details that a programmer views as inessential.
+%each of the data formats is quotiented by some equivalence relation and the lens
+%laws are taken modulo equivalence.
+%This allows lenses to be simpler and more
+%flexible. For example, inessential details like whitespace and orderings of
+%fields can be ignored when writing the transformations between the formats.
 

### 2018-02-16 papers/QuotientLenses/main.tex
intro edits

@@
 
-However, while plain lenses are widely used in both standalone systems and
-popular libraries, Boomerang~\cite{boomerang} is the only implementation
-currently offering quotient lenses. Even though quotient lenses permit simpler
-transformations between the formats, they merely push complexity elsewhere. To
-write a quotient lens, programmers must write with standard lens
-combinators and \emph{canonizers}. Canonizers encode two functions, one that
-maps a format to a canonical version of that format, and another that chooses
+%However, while plain lenses are widely used in both standalone systems and
+%popular libraries, Boomerang~\cite{boomerang} is the only implementation
+%currently offering quotient lenses. Even though quotient lenses permit simpler
+%transformations between the formats, they merely push complexity elsewhere.
+Ordinary bijective lenses are fairly tricky to program, especially for
+beginners, as they require programmers to think in two directions at once
+and learn a new type system, which comes with some tricky ambiguity constraints.
+Unfortunately, quotient lenses do not improve the situation:  To
+write a quotient lens in Boomerang, programmers must work with standard lens
+combinators as well as a system of \emph{canonizers}.
+These canonizers encode two functions, one that
+maps data represented in the original external format
+to a canonical element, and another, in the opposite direction, that chooses
 which elements of the original format correspond to values in the canonical
-format. Furthermore, if a user wants to enforce that a quotient lens maps
-between two different formats, they must write regular expressions specifying
-such formats. In this way, writing quotient lenses involves writing 3 terms: (1)
-the data formats on each side of the lens, (2) canonizers to normalize the data
-and choose representatives, and (3) a lens between the data formats when in
-canonical form, even though, typically, these three share a great deal of
-structure. Moreover, writing the lens between the equivalence classes can be
-challenging due to the rather fiddly typechecking, mainly due to
-``non-ambiguity'' conditions in the lens well-formedness rules.
-
-In this paper, we make programming quotient lenses easier in two ways. First, we
-introduce a novel language of {\em Quotient Regular Expressions}, or QREs, which
+format. Furthermore, if a user wants to define the domain and range of
+a lens explicitly (a programming practice that is encouraged), she must
+provide regular expressions defining input and output formats.
+Consequently, to define a quotient lens, one must often three different sorts
+of terms: (1) regular expressions that define the two data formats
+to be processed, (2) canonizers to normalize the data
+and choose representatives, and (3) a bijective lens maps
+one canonical form into another.  Somewhat frustratingly,
+these three terms share a fair amount of structure.
+%Moreover, writing the lens between the equivalence classes can be
+%challenging due to the rather fiddly typechecking, mainly due to
+%``non-ambiguity'' conditions in the lens well-formedness rules.
+
+In this paper, we improve the process of programming quotient lenses
+in two distinct, but connected ways. First, we
+introduce a new language of {\em Quotient Regular Expressions}, or QREs, which
 enable a programmer to simultaneously specify a regular expression as well as
-a set of canonical representatives on the regular expression's language. With
+a set of canonical representatives for strings that match the regular
+expression. With
 QREs, programmers do not need to define the data and canonization functions
-separately -- by adding simple annotations to the data, the canonization
-functions can be inferred.
+separately -- by adding simple annotations to ordinary regular expressions,
+canonization functions can be inferred.
 
 Second, we use QREs to define a class of quotient lenses called {\em QRE
-lenses} which map betweeen QREs, and show how to synthesize QRE lenses. In
-order to derive a lens, a programmer need not manually derive the lens
-herself, she only needs to provide a pair of QREs describing the formats the
-lens should map between, and a suite of examples that demonstrates how the
-lens should act on specific examples.  The resulting lens is guaranteed to
-map between the languages of the QREs and satisfy the provided examples.
-
-In summary, our main contributions are:
+lenses}, and, we show how to synthesize such QRE
+lenses automatically.  
+More specifically, in
+order to generate a lens, a programmer provides a pair of QREs
+describing the source and target formats, and, optionally,
+a suite of examples that demonstrates how the
+lens should act on specific examples.  Given this information,
+we are able to exploit and extend recent work on algorithms for synthesis of
+bijective lenses~\cite{optician} to generate quotient lenses.
+Importantly, the lenses generated by
+our synthesis algorithm are guaranteed to translate data back and forth
+between source and target QREs and to satisfy the provided examples.
+Moreover, while our synthesized quotient lenses have a relatively
+restricted form (a fact that cuts down the amount of non-determinism in
+a specification and thus helps speed up synthesis), we prove that
+this rigidity does not compromise expressiveness.
+
+The following list summarizes the central contributions of our work.
 \begin{enumerate}
@@
   which provide a compact, convenient notation that expresses an equivalence
-  relation on a regular languages as well as as a canonical set representatives
+  relation on a regular languages as well as a canonical set representatives
   for this equavalence relation (Section~\ref{QRE}).
-  \item We define a language of {\em QRE lenses}, a subset of bijective quotient
-  lenses whose types are given by QREs. Our main technical contribution in this
-  paper is a proof that each QRE lens $q$ defines the same transformation as a
+  \item We define a language of {\em QRE lenses}.  This language is
+  a subset of the bijective quotient
+  lenses whose types are given by QREs. Our main technical contribution
+  is a proof that each QRE lens $q$ defines the same transformation as a
   QRE lens $q'$ that is in normal form. (Section~\ref{QRE-lenses}).
@@
   \item We extend the Boomerang {\em implementation} with QREs
-  and QRE lens synthesis.  We demonstrate its applicability by using it to
+  and QRE lens synthesis.  We demonstrate the utility of our
+  implementation by using it to
   synthesize QRE lenses between several real-world data formats from the

### 2018-02-17 papers/SDLenses/main.tex

@@
 \grantsponsor{GS100000001}{National Science
-Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
+Foundation}{https://doi.org/10.13039/100000001} under Grant
 No.~\grantnum{GS100000001}{nnnnnnn} and Grant

### 2018-02-20 papers/QuotientLenses/main.tex
edits to sec 2

@@
 \label{sec:example}
-In this section, we explain the difficulties of programming with quotient lenses
-in Boomerang and show, by way of example, how our novel language of {\em
-Quotient Regular Expressions}, or QREs, and QRE lenses address these
-difficulties.
-
-Consider a bidirectional transformation that converts citation records such as
+
+%In this section, we explain the difficulties of programming with quotient
+%lenses in Boomerang and show, by way of example, how our 
+%{\em Quotient Regular Expressions} (QREs) help alleviate these difficulties.
+
+Consider a bidirectional transformation that converts
+\bibtex{} citation records such as
 \begin{verbatim}
@@
 \end{verbatim}
-
-\noindent stored in the \bibtex{} format to their equivalent representation
+\noindent
+to an equivalent EndNote format like the following
 \begin{verbatim}
@@
 \end{verbatim}
-
-\noindent in the EndNote format, and vice versa.
-
-If two \bibtex{} files differ in only ``inessential'' details, we would like
-these files to map to the same EndNote representations.
-First, the transformation should allow whitespace characters to occur
-arbitrarily between the words in the label, author, and title fields, as well as
-in between them. Second, an author's name can have the surname followed by a
-comma and then the other names separated by arbitrary whitespace, or just all
-the names separated by arbitrary whitespace (for example, the names ``Conway, J.
-H.'' and ``J. H. Conway'' are equivalent). Lastly, the transformation should
-also ignore the order in which the author and title fields occur in each
-citation record.
-
-For example, the desired transformation should map the following
-\bibtex{} citations to the same EndNote record in the forward direction:
+\noindent
+and vice versa.
+%
+Transformations from one such representation to the other will preserve
+most of the information content of the data:  the title, the author,
+and the key, for instance.  However, intuitively,
+some of the information is inessential and may be discarded---indeed, this
+inessential information will likely appear so trivial that a casual
+reader may not even realize it is there to begin with.  For instance,
+there is no need to preserve the amount of whitespace between words, at the ends of lines, or
+between fields.  In addition, the order of author and title fields is irrelevant.
+There may also be additional equivalent ways of writing the same
+name:  ``Conway, J. H.'' and ``J. H. Conway'' should probably be treated
+the same.
+
+%% Moreover,
+%% if two \bibtex{} files differ only in ``inessential'' details, we would like
+%% to treat them the same by mapping them to identical EndNote representations.
+%% In this , the transformation should allow whitespace characters to occur
+%% arbitrarily between the words in the label, author, and title fields, as well as
+%% in between them. Second, an author's name can have the surname followed by a
+%% comma and then the other names separated by arbitrary whitespace, or just all
+%% the names separated by arbitrary whitespace (for example, the names ``Conway, J.
+%% H.'' and ``J. H. Conway'' are equivalent). Lastly, the transformation should
+%% also ignore the order in which the author and title fields occur in each
+%% citation record.
+
+When two instances of the same format differ only terms of such inessential
+information, it is often useful to treat them as identical.
+For example, the following two
+\bibtex{} citations represent the same logical object and as a result, should
+be translated to the same EndNote record.
 \begin{verbatim}

### 2018-02-21 papers/QuotientLenses/main.tex
section 2 edits checkpoint

@@
 
-\section{Quotient Regular Expressions and Lenses by Example}
+\section{Quotients by Example}
 \label{sec:example}
@@
 
-Consider a bidirectional transformation that converts
+\emph{Lenses} define computations that can be read in two directions, forwards and
+backwards.  The simplest lenses define bijections between source and
+data, but we are interested in \emph{quotient lenses}, which define
+bijections modulo some equivalence relation.  The remains of this section
+describe and motivate a new language and system for defining such lenses.
+
+To begin, consider a bidirectional transformation that converts
 \bibtex{} citation records such as
@@
 \noindent
-to an equivalent EndNote format like the following
+into an equivalent EndNote format like the following
 \begin{verbatim}
@@
 %
-Transformations from one such representation to the other will preserve
-most of the information content of the data:  the title, the author,
-and the key, for instance.  However, intuitively,
-some of the information is inessential and may be discarded---indeed, this
-inessential information will likely appear so trivial that a casual
-reader may not even realize it is there to begin with.  For instance,
-there is no need to preserve the amount of whitespace between words, at the ends of lines, or
-between fields.  In addition, the order of author and title fields is irrelevant.
-There may also be additional equivalent ways of writing the same
-name:  ``Conway, J. H.'' and ``J. H. Conway'' should probably be treated
-the same.
+Notice that transformations from one format to the other preserve
+most of the information present in the data:  the title, the author,
+and the key, for instance.  However, some information in the syntax
+may be lost.  For instance, users are typically uninterested in preserving
+the amount of the whitespace between words or
+between fields.  The order of author and title fields is also likely
+irrelevant and there may be equivalent ways of writing the same
+name:  ``Conway, J. H.'' vs ``J. H. Conway.''
 
@@
 When two instances of the same format differ only terms of such inessential
-information, it is often useful to treat them as identical.
-For example, the following two
+information, it is often useful to treat them as identical and to map them
+into canonical representations.
+Continuing our example, the following two
 \bibtex{} citations represent the same logical object and as a result, should
@@
 \end{verbatim}
-\noindent EndNote records should also have a similar equivalence relation
-defined on them.
-
-While it is possible to express the equivalence on the \bibtex{} data above in
-Boomerang, actually doing so is a challenge.  Three of the inessential details
-we wish to ignore---whitespace quotienting
-inside fields, whitespace quotienting in between fields, and permutation
-quotenting of fields---nest within each other in a way that makes them difficult
-to express with the existing Boomerang combinators. Furthermore, if the
-programmer wishes to specify the type of the lens, then she can only specify
-data formats on each side of the lens with a regular expression that does not
-keep track of the equivalence relation defined on the language of the regular
-expression. Consequently, the programmer has to write a regular expression as
-well as a canonizer that expresses the equivalence relation on that regular
-expression, even though a regular expression and the canonizer defined on it
-typically share much of the syntactic structure. Having to write both of these
-expressions makes programming with quotient lenses verbose and time consuming.
-
-To address these issues, we propose {\em Quotient Regular Expressions}, or QREs,
-which enable the programmer to define a regular expression $R$, an equivalence
-relation $\sim$ on the language described by $R$ as well as a canonical set
-of representatives of $\sim$, all in one syntactic term. For example, with QREs
-we can specify \bibtex{} records and their canonical elements, as shown in
-Figure~\ref{fig:example-qre}.
+
+One natural way to define lenses that will synchronize data in these formats
+is to provide the following items:
+
+\begin{enumerate}
+\item A pair of descriptions of the external source and target formats.
+Such descriptions may be given by regular expressions, for instance.
+They serve as documentation to describe the domains over which the lens
+is to operate.
+\item A pair of \emph{canonizers} that map external source (and target)
+objects into internal, canonical forms.  Such canonizers will take
+care of stripping away irrelevant information and otherwise normalizing
+differences between equivalent representations of the same logical data.
+When given an internal representation, such canonizers must also operate
+in the reverse direction, \emph{choosing} an element in the external format
+to represent the canonical internal element.
+\item A pair of descriptions of the internal source and
+target formats, again, to document the structure of the transformation.
+\item A bijective lens to map back and forth between canonical elements
+of the internal source and target formats.
+\end{enumerate}
+
+\noindent
+Indeed, Boomerang's quotient lenses are structured this way~\ref{boomerang}.
+Of course, some of these elements are optional.  For instance, it is
+not necessary to document internal and external formats.  However, we do
+view it as good engineering practice to do so, as such documentation makes
+lenses easier to read, understand, use and maintain.  In this case, there is
+significant redundancy here:  External and internal formats
+typically have some overlap; external formats and canonizers have related
+structure; and the structure of bijective lens computations is largely driven
+by the structure of the internal formats.  Hence, overall, a more compact
+notation would simplify description and legibility of such computations.
+In addition,
+the lens and canonizer languages themselves can be somewhat fiddly and
+difficult to
+master.  For instance, to quotient out whitespace inside, and in between,
+fields while also ignoring field order one must nest canonizing combinators
+within one another in a somewhat tricky and fiddly way.
+
+%% %Foster's Boomerang~\ref{boomerang} language presents one design
+%% While it is possible to express the equivalence on the \bibtex{} data above in
+%% Boomerang, actually doing so is a challenge.  Three of the inessential details
+%% we wish to ignore---whitespace quotienting
+%% inside fields, whitespace quotienting in between fields, and permutation
+%% quotenting of fields---nest within each other in a way that makes them difficult
+%% to express with the existing Boomerang combinators. Furthermore, if the
+%% programmer wishes to specify the type of the lens, then she can only specify
+%% data formats on each side of the lens with a regular expression that does not
+%% keep track of the equivalence relation defined on the language of the regular
+%% expression. Consequently, the programmer has to write a regular expression as
+%% well as a canonizer that expresses the equivalence relation on that regular
+%% expression, even though a regular expression and the canonizer defined on it
+%% typically share much of the syntactic structure. Having to write both of these
+%% expressions makes programming with quotient lenses verbose and time consuming.
+
+We propose to do away with much of this overhead and the burden of a
+fiddly programming process by defining a single new compact notation,
+the \emph{Quotient Regular Expression}, which simultaneously describes
+external format, equivalence relation and canonical forms, and then to
+synthesize lenses that map between these objects automatically.
+
+\subsection{Quotient Regular Expressions}
+
+A {\em Quotient Regular Expression} (QRE) defines an external regular
+expression $R$,
+an equivalence relation $\sim$ on the language described by $R$, and a canonical set
+of representatives of $\sim$, all in one syntactic term.
+Figure~\ref{fig:example-qre} illustrates how to 
+specify \bibtex{} records and their canonical elements using QREs.
 

### 2018-02-22 papers/QuotientLenses/main.tex
qres

@@
 \newcommand{\kw}[1]{\ensuremath{\mathsf{#1}}}
+\newcommand{\codefont}[1]{\ensuremath{\mathsf{#1}}}
 \newcommand{\project}[2]{\ensuremath{\kw{project} \; #1 \mapsto #2}}
@@
 
-We propose to do away with much of this overhead and the burden of a
-fiddly programming process by defining a single new compact notation,
-the \emph{Quotient Regular Expression}, which simultaneously describes
-external format, equivalence relation and canonical forms, and then to
-synthesize lenses that map between these objects automatically.
+We propose to do away with much of this overhead, and the burden of a
+fiddly programming process, by defining a single new compact notation
+to specify data formats modulo an equivalence relation and then by
+synthesizing the lenses that map between these formats automatically.
 
@@
 A {\em Quotient Regular Expression} (QRE) defines an external regular
-expression $R$,
-an equivalence relation $\sim$ on the language described by $R$, and a canonical set
-of representatives of $\sim$, all in one syntactic term.
-Figure~\ref{fig:example-qre} illustrates how to 
-specify \bibtex{} records and their canonical elements using QREs.
+expression $R$, an equivalence relation $\sim$ on the language
+described by $R$, and a canonical set of representatives of $\sim$,
+all at once.  QREs are constructed from a simple set of primitives
+and composed using familiar regular operators
+(choice `\cd{|}', concatenation `\cd{.}', sequential composition ``;''
+and Kleene star `$^*$').
+
+For example, to begin construction of our \bibtex{} format, we might
+first define a whitespace format, which, externally, matches
+any non-zero number of whitespace characters, and converts
+any such whitespace into a single space character (its canonical form).
+Such a definition uses the \kw{project} primitive to implement the
+specification:
+
+\begin{lstlisting}
+let wsp = project [ \n\t\r]$^+$ $\to$ " "
+\end{lstlisting}
+
+While a projection chooses a single canonical representative for all elements
+of a regular language, the \kw{squash} combinator maps data from one
+regular language to another via a user-defined
+function.  For instance,
+if \codefont{comma_name} describes the format for ``Conway, J. H.''
+and \codefont{space_name} describes the format for ``J. H. Conway''
+and \codefont{s_to_c} is a function from one to the other then
+the following instance of squash creates the desired canonizer.
+
+%Figure~\ref{fig:example-qre} illustrates
+%how to specify \bibtex{} records and their canonical elements using
+%QREs.
+
+\begin{lstlisting}
+let name = squash comma_name $\to$ space_name using s_to_c
+\end{lstlisting}
+
+One way to define the function used by squash is simply to write it
+from scratch in OCaml.  However, our system will also synthesize
+such functions automatically from pairs of regular languages and
+examples.
+%
+\begin{lstlisting}
+let l = synth comma_name $\Leftrightarrow$ space_name using {("Conway, J. H.", "J. H. Conway")}
+let s_to_c = get l
+\end{lstlisting}
+%
+The first line above synthesizes a \emph{bidirectional} lens between
+\codefont{comma_name} and \codefont{space_name} using the listed example
+transformation as a guide.  The second line simply
+extracts the left-to-right transformation from the lens, which is what we need for squash.
+
+A third useful QRE primitive is the permutation combinator.  For example,
+the following instance of \kw{perm} allows label, author and title fields
+(which we assume have been defined earlier) to appear in any order.
+%
+\begin{lstlisting}
+let bib_fields = perm (label, bib_author, bib_title) 
+\end{lstlisting}
+%
+To normalize any repeated separator between fields, one can extend a permutation
+with any other QRE.  For instance, below, we normalize whitespace between fields,
+leaving only a single newline.
+%
+\begin{lstlisting}
+let bib_fields = perm (label, bib_author, bib_title) with (project ("," .wsp$^+$) $\to$ ",\n")
+\end{lstlisting}
+%
+\noindent
+The QRE for the bibtex format is presented in its entirety in Figure~\ref{fig:example-qre}.
 
@@
 
-In Figure~\ref{fig:example-qre}, the QRE \verb+project [ \n\t\r]+$^{+} \to$
-\verb+" "+ matches any non-zero number of whitespace characters and quotients
-them to a single equivalence class with a single space as the canonical
-representative. We call this QRE a \textit{projection} because an entire regular
-language is projected to a single representative string.
-
-Another QRE in Figure~\ref{fig:example-qre} is given by
-
-\begin{center}
-\begin{lstlisting}
-squash comma_name $\to$ space_name using (get space_to_comma)
-\end{lstlisting}
-\end{center}
-
-\noindent This QRE is a \textit{squash} QRE. A \textit{squash} QRE ``glues''
-two data formats using a user-defined function. For example,
-the QRE in the code above specifies that the \cd{comma_name} representation of
-the author's name is equivalent to the \cd{space_name} representation, with
-\cd{space_name} taken to be the canonical representation. The
-function that identifies the two formats is given by \cd{get space_to_comma},
-which transforms names in the \cd{comma_name} representation to names in the
-\cd{space_name} representation. In this example, \cd{space_to_comma} is itself
-an automatically \textit{synthesized} lens between the \cd{space_name} and
-\cd{comma_name} QREs, and we need to use the ``get'' direction of the lens to
-specify the squash transformation.
-
-The third example of a QRE in the Figure is \cd{bib_fields}, defined by:
-\begin{center}
-\begin{lstlisting}
-perm (label, bib_author, bib_title) with (project ("," . wsp$^+$ ) $\to$ ",\n")
-\end{lstlisting}
-\end{center}
-
-\noindent This is a \textit{permutation} QRE that specifies that the label,
-title, and author fields can occur in any order, with the fields separated by a
-comma followed by arbitrary amounts of whitespace. The canonical representation
-is that in which the label, author and title fields occur in that order,
-separated by a comma followed by a newline. The \textit{permutation} QRE is
-designed for equivalences such as this, where different permutations of
-equivalences are interspersed with a separator which itself has an
-equivalence defined on it.
-
-Given the \textit{projection}, \textit{squash}, and \textit{permutation}
-QRE primitives, we can build more complex languages using the usual regular
-expression combinators of alternative choice `\cd{|}', concatenation `\cd{.}'
-and Kleene star `$^*$', and also sequential composition ``;''.
+%% In Figure~\ref{fig:example-qre}, the QRE \verb+project [ \n\t\r]+$^{+} \to$
+%% \verb+" "+ matches any non-zero number of whitespace characters and quotients
+%% them to a single equivalence class with a single space as the canonical
+%% representative. We call this QRE a \textit{projection} because an entire regular
+%% language is projected to a single representative string.
+
+%% Another QRE in Figure~\ref{fig:example-qre} is given by
+
+%% \begin{center}
+%% \begin{lstlisting}
+%% squash comma_name $\to$ space_name using (get space_to_comma)
+%% \end{lstlisting}
+%% \end{center}
+
+%% \noindent This QRE is a \textit{squash} QRE. A \textit{squash} QRE ``glues''
+%% two data formats using a user-defined function. For example,
+%% the QRE in the code above specifies that the \cd{comma_name} representation of
+%% the author's name is equivalent to the \cd{space_name} representation, with
+%% \cd{space_name} taken to be the canonical representation. The
+%% function that identifies the two formats is given by \cd{get space_to_comma},
+%% which transforms names in the \cd{comma_name} representation to names in the
+%% \cd{space_name} representation. In this example, \cd{space_to_comma} is itself
+%% an automatically \textit{synthesized} lens between the \cd{space_name} and
+%% \cd{comma_name} QREs, and we need to use the ``get'' direction of the lens to
+%% specify the squash transformation.
+
+%% The third example of a QRE in the Figure is \cd{bib_fields}, defined by:
+%% \begin{center}
+%% \begin{lstlisting}
+%% perm (label, bib_author, bib_title) with (project ("," . wsp$^+$ ) $\to$ ",\n")
+%% \end{lstlisting}
+%% \end{center}
+
+%% \noindent This is a \textit{permutation} QRE that specifies that the label,
+%% title, and author fields can occur in any order, with the fields separated by a
+%% comma followed by arbitrary amounts of whitespace. The canonical representation
+%% is that in which the label, author and title fields occur in that order,
+%% separated by a comma followed by a newline. The \textit{permutation} QRE is
+%% designed for equivalences such as this, where different permutations of
+%% equivalences are interspersed with a separator which itself has an
+%% equivalence defined on it.
+
+%% Given the \textit{projection}, \textit{squash}, and \textit{permutation}
+%% QRE primitives, we can build more complex languages using the usual regular
+%% expression combinators of alternative choice `\cd{|}', concatenation `\cd{.}'
+%% and Kleene star `$^*$', and also sequential composition ``;''.
 

### 2018-02-23 papers/QuotientLenses/main.tex
section 2 editing

@@
 One natural way to define lenses that will synchronize data in these formats
-is to provide the following items:
+is to provide the following items: \dpw{this is slightly redundant from the intro.  It
+adds a little more info, but not much.  Should this be cut?  Should the intro be expanded
+slightly?  Should it be left as is (more or less)?}
 
@@
 (* synthesize a lens that maps the comma representation to the space representation *)
-let space_to_comma = synth comma_name $\Leftrightarrow$ space_name using {("Conway, J. H.", "J. H. Conway")}
+let l = synth comma_name $\Leftrightarrow$ space_name using {("Conway, J. H.", "J. H. Conway")}
+let s_to_c = get l
 
 (* squash QRE maps comma_name to space_name *)
-let name = squash comma_name $\to$ space_name using (get space_to_comma)
+let name = squash comma_name $\to$ space_name using s_to_c
 
@@
 \subsection{QRE Lenses}
-While QREs make it easier to write equivalence relations on regular expressions
-by specifying the data and the equivalence relation on the data in one
-expression, in order to define a quotient lens, the programmer still has to
-manually derive the underlying lens that maps the equivalence classes of the
-source data to the equivalence classes of the target data. This is itself a
-tricky task as Figure~\ref{fig:example-lens} illustrates.
+QREs allow users to specify regular languages modulo equivalence relations,
+a first step in writing many data transformations.  However, writing the
+quotient lens itself still takes a fair amount of work.  Each such lens
+is constructed from simple primitive combinators including:
+%
+\begin{itemize}
+\item \kw{del} $s$: delete the constant string $s$ in the left-to-right direction; produce the constant string $s$ in the right-to-left direction.
+\item \kw{ins} $s$: produce the string $s$ in the left-to-right direction;
+delete the string $s$ in the right-to-left direction.
+\item \kw{id} $r$: copy the text matching $r$ from left-to-right (also from
+right-to-left) \dpw{this was not the exact syntax.  fix this.}
+\end{itemize}
+%
+In addition, such primitives may be combined using the regular operators
+Kleene star, alternation and concatenation, or functional composition of
+lenses.  Figure~\ref{fig:example-lens} illustrates the use of these
+combinators to transform \bibtex{} to EndNote (and vice versa).
+A fair amount of trickiness in writing these transformations as users
+must keep mental track of what they have parsed so far \emph{in both
+directions}, carefully inserting and deleting the relevant syntax on both
+sides.
+
+%% While QREs make it easier to write equivalence relations on regular expressions
+%% by specifying the data and the equivalence relation on the data in one
+%% expression, in order to define a quotient lens, the programmer still has to
+%% manually derive the underlying lens that maps the equivalence classes of the
+%% source data to the equivalence classes of the target data. This is itself a
+%% tricky task as Figure~\ref{fig:example-lens} illustrates.
 
@@
 
-The lens shown in Figure~\ref{fig:example-lens} transforms \bibtex{} data to
-EndNote data  (and vice versa) by matching the delimiters and separators in a
-bijective manner. For example, the \cd{preamble} lens replaces the \bibtex{}
-preamble \verb|@Book| with the EndNote preamble \verb|%0 Book\n%F|. When the
-programmer is writing the lens, she has to make sure that all of these small
-deletions and insertions are correct, which can be a challenging task if the
-lens is large enough, as in the definition of the lens \verb!author_lens! in
-Figure~\ref{fig:example-lens}. This makes lens programming tedious and
-error-prone.
+%% The lens shown in Figure~\ref{fig:example-lens} transforms \bibtex{} data to
+%% EndNote data  (and vice versa) by matching the delimiters and separators in a
+%% bijective manner. For example, the \cd{preamble} lens replaces the \bibtex{}
+%% preamble \verb|@Book| with the EndNote preamble \verb|%0 Book\n%F|. When the
+%% programmer is writing the lens, she has to make sure that all of these small
+%% deletions and insertions are correct, which can be a challenging task if the
+%% lens is large enough, as in the definition of the lens \verb!author_lens! in
+%% Figure~\ref{fig:example-lens}. This makes lens programming tedious and
+%% error-prone.
 
 Consequently, we offer the programmer some relief by identifying a class of
-quotient lenses that we call {\em QRE lenses}, and show how to {\em synthesize}
+quotient lenses that we call {\em QRE lenses}, and by showing how to {\em synthesize}
 QRE lenses from a pair of QREs and a set of input-output example pairs. Our key
 insight is to reduce the problem of synthesizing QRE lenses to the problem of
-synthesizing {\em bijective lenses}~\cite{optician}, which are bijections created
-from the Boomerang lens combinators.
-
-QRE lenses lenses are the same as bijective lenses with ``canonizers at the
-edges'' as in Figure~\ref{fig:attheedges}, and are typed using QREs---every QRE
-lens $q$ has a type $q: c \Leftrightarrow c'$ where $c$ and $c'$ are QREs.
+synthesizing {\em bijective lenses}~\cite{optician}, a simpler subset of
+lenses that support strict bijections, rather than bijections modulo
+equivalence relations.
+
+Intuitively, our QRE lenses are bijective lenses with ``canonizers at the
+edges''---Figure~\ref{fig:attheedges} presents the architecture of each lens.
+Every QRE lens $q$ has a type $q: c \Leftrightarrow c'$ where $c$ and $c'$ are QREs.
 In the forward direction, a QRE lens $q: c \Leftrightarrow c'$ uses the source
@@
 similarly,  but using the view QRE $c'$ and the $\lput$ function of $\ell$.
+
+Because the QREs $c$ and $c'$ determine the internal formats for data after
+canonization, and because the algorithm for synthesizing bijective lenses is
+directed by these formats, $c$ and $c'$ are all that is required to synthesize
+QRE lenses end-to-end.  Indeed, we structured QRE lenses as bijective lenses with
+canonizers at the edges, precisely to facilitate the use of this synthesis algorithm.
 
@@
 
-One potential pitfall with QRE lenses is that the combinators that we define on
-them may not preserve this normal form. For example, the composition $q'; q$ of
-QRE lenses $q$ and $q'$ might not be equivalent to any quotient lens that has
-the canonical normal form of a bijective lens with canonizers at the edges.
-Our main technical contribution in this paper is a proof that each QRE lenses
-$q$ defines the same transformation as a QRE lens $q'$ that is in normal form.
-
-For example, suppose \cd{q_1 : bibtex} $\Leftrightarrow$ \cd{endnote} is a QRE
-lens that maps between \bibtex{} and EndNote formats, and that \cd{q_2 : endnote
-<=> xml} is a QRE lens that maps between EndNote and XML formats. The
-composition \cd{(q_1 ; q2) : bibtex} $\Leftrightarrow$ \cd{xml} of \cd{q_1} and
-\cd{q_2} is not necessarily in normal form due to the presence of canonizers
-on each side of the intermediary EndNote format:
-
-\sam{TODO: We need to put a diagram of composition here.}
-
-\noindent However, while \cd{q_1 ; q_2} is not in normal form, our main theorem
-shows that there is another QRE lens \cd{q : bibtex} $\Leftrightarrow$ \cd{xml}
-of the same type as \cd{(q_1 ; q2) : bibtex} $\Leftrightarrow$ \cd{xml} such
-that \cd{q_1 ; q_2} and \cd{q} define the same transformation. This proof
-will allow justify our approach for synthesizing quotient lenses so that the
-programmer can replace all of the code in Figure~\ref{fig:example-lens} with a
-call to the synthesis prodedure:
+%% One potential pitfall with QRE lenses is that the combinators that we define on
+%% them may not preserve this normal form. For example, the composition $q'; q$ of
+%% QRE lenses $q$ and $q'$ might not be equivalent to any quotient lens that has
+%% the canonical normal form of a bijective lens with canonizers at the edges.
+
+On the other hand, our restriction that canonizers appear only at the edges
+raises a key technical question:  Are we limiting the expressiveness of our transformations
+by demanding all programs fit into this normal form?  It turns out that we are not:
+any lens that uses canonizers internally can be transformed into a lens that uses canonizers
+at the edges.  The main technical contribution of this paper is a proof of this fact.  The
+crucial step in the proof is a demonstration that functional composition of two lenses
+with ``canonizers at the edge'' can be represented as some single lens with ``canonizers at the
+edge.''
+
+%% For example, suppose \cd{q_1 : bibtex} $\Leftrightarrow$ \cd{endnote} is a QRE
+%% lens that maps between \bibtex{} and EndNote formats, and that \cd{q_2 : endnote
+%% <=> xml} is a QRE lens that maps between EndNote and XML formats. The
+%% composition \cd{(q_1 ; q2) : bibtex} $\Leftrightarrow$ \cd{xml} of \cd{q_1} and
+%% \cd{q_2} is not necessarily in normal form due to the presence of canonizers
+%% on each side of the intermediary EndNote format:
+
+%% \sam{TODO: We need to put a diagram of composition here.}
+
+%% \noindent However, while \cd{q_1 ; q_2} is not in normal form, our main theorem
+%% shows that there is another QRE lens \cd{q : bibtex} $\Leftrightarrow$ \cd{xml}
+%% of the same type as \cd{(q_1 ; q2) : bibtex} $\Leftrightarrow$ \cd{xml} such
+%% that \cd{q_1 ; q_2} and \cd{q} define the same transformation.
+
+This technical investigation justifies replacing manually constructed lenses that may use
+functional composition and nested canonizers with synthesized lenses in normal form.
+The end result is that users may synthesize complex transformations with a single line
+of code.  For instance, all the code in Figure~\ref{fig:example-lens} may be replaced by
+a single call to the synthesis prodedure:
 
@@
 \end{lstlisting}
-\noindent Here, the resulting  quotient lens maps between the \cd{bibtex} and
-\cd{endnote} QREs, and sends the equivalence class of \cd{bib_example} to
-the equivalence class of \cd{end_example}, which are the two concrete example
-strings given at the beginning of this Section.
-
-As we saw with the definition of \cd{space_to_comma} in
-Figure~\ref{fig:example-qre}, the synthesis procedure itself can be used to
-create lenses that are in turn used to define other QREs.  This ability to mix
+\noindent Here, the generated  quotient lens synchronizes \cd{bibtex} and
+\cd{endnote} formats.  It uses \cd{bib_example} and \cd{end_example} (the two concrete example
+strings given at the beginning of this section) to disambiguate.
+In additiona, and as we saw earlier with the definition of \cd{s_to_c}, the synthesis procedure itself
+can be used to create lenses that are in turn used to define other QREs.  The ability to intermix
 QRE specification with QRE lens synthesis yields a powerful and flexible way of

### 2018-02-23 papers/QuotientLenses/main.tex
many intro typos fixed

@@
 JSON or
-XML object. When different applications manipulate data in different formats,
+XML object. When different applications manipulate the same data in different formats,
 synchronizing the applications amounts to synchronizing their respective data
@@
 restrictive lenses are the \emph{bijective lenses}.  These lenses demand
-that get and put functions for a bijection, meaning no information
-is lost when one translates from data format to another.  While this
+that get and put functions form a bijection, meaning no information
+is lost when one translates from one data format to another.  While this
 strong guarantee can be useful in some contexts, it is more often the case
@@
 %However, these lens laws are too restrictive for many important transformations.
-In particular, these lens laws do not permit transformations that put the data
-in canonical forms.
-For instance, to return to our bibliography example, in \bibtex{},
+%The bijective lens laws do not permit transformations that ignore this
+%unimportant information---.
+Many data formats have other ad hoc rules as well.  
+Returning to our bibliography example, recall that in \bibtex{},
 an author can be specified as
@@
 %\cd{firstname lastname}, but the lens laws prevent such functions.
+%
 {\em Quotient lenses}~\cite{quotientlenses} loosen many of the restrictions
@@
 provide regular expressions defining input and output formats.
-Consequently, to define a quotient lens, one must often three different sorts
+Consequently, to define a quotient lens, one must often
+construct three different sorts
 of terms: (1) regular expressions that define the two data formats
 to be processed, (2) canonizers to normalize the data
-and choose representatives, and (3) a bijective lens maps
+and choose representatives, and (3) a bijective lens that maps
 one canonical form into another.  Somewhat frustratingly,
@@
 QREs, programmers do not need to define the data and canonization functions
-separately -- by adding simple annotations to ordinary regular expressions,
+separately---by adding simple annotations to ordinary regular expressions,
 canonization functions can be inferred.
@@
 \begin{enumerate}
-  \item We introduce {\em Quotient Regular Expressions} (QREs)
-  which provide a compact, convenient notation that expresses an equivalence
-  relation on a regular languages as well as a canonical set representatives
-  for this equavalence relation (Section~\ref{QRE}).
+  \item We introduce {\em Quotient Regular Expressions} (QREs),
+  a compact, convenient notation for defining (1) a
+  regular language modulo some equivalence relation, and (2) a canonizer
+  for that relation (Section~\ref{QRE}).
   \item We define a language of {\em QRE lenses}.  This language is
-  a subset of the bijective quotient
+  a subset of the full set of bijective quotient
   lenses whose types are given by QREs. Our main technical contribution
-  is a proof that each QRE lens $q$ defines the same transformation as a
-  QRE lens $q'$ that is in normal form. (Section~\ref{QRE-lenses}).
-  \item Using this normal form, we reduce the problem of {\em synthesizing}
+  is a proof that our QREs are just as expressive as the full set of bijective
+  quotient lenses.
+  % In other words, if there eeach QRE lens $q$ defines the same transformation as a
+  %QRE lens $q'$ that is in normal form.
+  (Section~\ref{QRE-lenses}).
+  \item We reduce the problem of {\em synthesizing}
   QRE lenses to the previously studied problem of synthesizing bijective lenses
@@
   implementation by using it to
-  synthesize QRE lenses between several real-world data formats from the
+  synthesize QRE lenses between data formats drawn from the
   {\tt data.gov} database (Section~\ref{impl}).

### 2018-02-23 papers/QuotientLenses/main.tex
edits on 2

@@
 most of the information present in the data:  the title, the author,
-and the key, for instance.  However, some information in the syntax
+and the label, for instance.  However, some of the syntactic information
 may be lost.  For instance, users are typically uninterested in preserving
-the amount of the whitespace between words or
+whitespace between words or
 between fields.  The order of author and title fields is also likely
@@
 master.  For instance, to quotient out whitespace inside, and in between,
-fields while also ignoring field order one must nest canonizing combinators
-within one another in a somewhat tricky and fiddly way.
+fields while also ignoring field order, one must nest canonizing combinators
+within one another in a somewhat tricky way.
 

### 2018-02-23 papers/QuotientLenses/main.tex
Tidying

@@
 \caption{An example lens between \cd{bibtex} and \cd{end_note} regular
-expressions with no equivalences.}
+expressions with no equivalences.\bcp{Indentation needs work.  And there was
+a suggestion to colorize the new keywords that we're introducing here
+(including symbolic ones, I guess)}} 
 \label{fig:example-lens}
@@
 \centering
-\includegraphics[width=\textwidth]{canonizers-outside}
+\includegraphics[width=0.7\textwidth]{canonizers-outside}
 \caption{Quotient Lens with ``Canonizers at the Edges''\sam{This is a
@@
 function. With QREs the kernel language is always contained in the whole
-language so we just use the inclusion as the choose function.}}
+language so we just use the inclusion as the choose function.}\bcp{I can
+fix.}} 
 \label{fig:attheedges}
@@
 
-\begin{figure}[p!]
+\begin{figure}[tp!]
 \begin{center}

### 2018-02-23 papers/QuotientLenses/main.tex
minor edits

@@
 
-\emph{Lenses} define computations that can be read in two directions, forwards and
-backwards.  The simplest lenses define bijections between source and
+\emph{Lenses} define computations that can be read in two directions,
+the get direction (forwards) and the put direction (backwards).  The simplest lenses define bijections between source and
 data, but we are interested in \emph{quotient lenses}, which define
@@
 function.  For instance,
-if \codefont{comma_name} describes the format for ``Conway, J. H.''
-and \codefont{space_name} describes the format for ``J. H. Conway''
-and \codefont{s_to_c} is a function from one to the other then
+if \codefont{comma\_name} describes the format for ``Conway, J. H.''
+and \codefont{space\_name} describes the format for ``J. H. Conway''
+and \codefont{s\_to\_c} is a function from one to the other then
 the following instance of squash creates the desired canonizer.
@@
 transformation as a guide.  The second line simply
-extracts the left-to-right transformation from the lens, which is what we need for squash.
+extracts the get direction transformation from the lens,
+which is what we need for squash.
 
@@
 \begin{itemize}
-\item \kw{del} $s$: delete the constant string $s$ in the left-to-right direction; produce the constant string $s$ in the right-to-left direction.
-\item \kw{ins} $s$: produce the string $s$ in the left-to-right direction;
-delete the string $s$ in the right-to-left direction.
-\item \kw{id} $r$: copy the text matching $r$ from left-to-right (also from
-right-to-left) \dpw{this was not the exact syntax.  fix this.}
+\item \kw{del} $s$: delete the constant string $s$ in the get direction;
+produce the constant string $s$ in the put direction.
+\item \kw{ins} $s$: produce the string $s$ in the get direction;
+delete the string $s$ in the put direction.
+\item \kw{id} $r$: copy the text matching $r$ in both directions \dpw{this was not the exact syntax.  fix this.}
 \end{itemize}
@@
 Every QRE lens $q$ has a type $q: c \Leftrightarrow c'$ where $c$ and $c'$ are QREs.
-In the forward direction, a QRE lens $q: c \Leftrightarrow c'$ uses the source
+In the get direction, a QRE lens $q: c \Leftrightarrow c'$ uses the source
 QRE $c$ to compute a unique representative for the data modulo the equivalence

### 2018-02-23 papers/QuotientLenses/main.tex
edits

@@
 and composed using familiar regular operators
-(choice `\cd{|}', concatenation `\cd{.}', sequential composition ``;''
-and Kleene star `$^*$').
+(choice `\cd{|}', concatenation `\cd{.}',
+and Kleene star `$^*$').  Users may also treat QREs as pairs of
+get and put functions and compose such pairs (the functional
+composition operator `;').
 
@@
 any such whitespace into a single space character (its canonical form).
-Such a definition uses the \kw{project} primitive to implement the
-specification:
+Such a definition uses the \kw{project} primitive.
 
@@
 
-While a projection chooses a single canonical representative for all elements
-of a regular language, the \kw{squash} combinator maps data from one
-regular language to another via a user-defined
-function.  For instance,
-if \codefont{comma\_name} describes the format for ``Conway, J. H.''
-and \codefont{space\_name} describes the format for ``J. H. Conway''
-and \codefont{s\_to\_c} is a function from one to the other then
-the following instance of squash creates the desired canonizer.
+Sometimes there are multiple disjoint representations of the same data.
+In such situations, the \kw{squash} combinator creates a QRE that
+allows external data to be in either format, and converts any
+data in one format to the other.
+For instance, assume
+the \codefont{comma\_name} format describes ``Conway, J. H.''
+and the \codefont{space\_name} format describes ``J. H. Conway''
+and \codefont{s\_to\_c} is a function from the first to the second.  In this case,
+the following instance of \kw{squash} creates the desired canonizer.
 

### 2018-02-25 papers/QuotientLenses/main.tex
Made syntax highlighting changes

@@
 
-\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
+\documentclass[acmsmall,review,anonymous]{acmart}
+\settopmatter{printfolios=true,printccs=false,printacmref=false}
 % % For double-blind review submission, w/ CCS and ACM Reference
@@
 \lstset{
-language=Caml, basicstyle=\upshape\sffamily,
-keywordstyle=\upshape\sffamily\color{dkblue}, keepspaces=true,
-framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=false,
-commentstyle=\itshape\rmfamily, columns=flexible,
+language=Caml, 
+basicstyle=\upshape\sffamily,
+keywordstyle=\upshape\sffamily\color{dkblue},
+keepspaces=true,
+framexleftmargin=1ex,
+framexrightmargin=1ex,
+showstringspaces=false,
+commentstyle=\itshape\rmfamily,
+emph={synth,project,perm,squash,normalize,using,ins,del,lens},
+emphstyle=\upshape\sffamily\color{dkblue},
+columns=flexible,
 % BCP: I find this distracting:
 % stringstyle=\sffamily\color{dkred},
-mathescape
+mathescape,
  }
@@
 languages like Boomerang~\cite{boomerang} or
-BiGul~\cite{DBLP:conf/pepm/KoZH16}, typically come with a variety of guarantees,
+BiGul~\cite{ko2016bigul}, typically come with a variety of guarantees,
 the \emph{lens laws}, that help ensure put and get functions are effective
@@
 \noindent
-Indeed, Boomerang's quotient lenses are structured this way~\ref{boomerang}.
+Indeed, Boomerang's quotient lenses are structured this
+way~\cite{quotientlenses}.
 Of course, some of these elements are optional.  For instance, it is
@@
 The first line above synthesizes a \emph{bidirectional} lens between
-\codefont{comma_name} and \codefont{space_name} using the listed example
+\codefont{comma\_name} and \codefont{space\_name} using the listed example
 transformation as a guide.  The second line simply
@@
 a first step in writing many data transformations.  However, writing the
-quotient lens itself still takes a fair amount of work.  Each such lens
+inner bijective lens itself still takes a fair amount of work.  Each such lens
 is constructed from simple primitive combinators including:
@@
   delete the string $s$ in the put direction.
-\item \kw{id} $r$: copy the text matching $r$ in both directions \dpw{this was not the exact syntax.  fix this.}
+  \item \kw{copy} $R$: copy the text matching the regular expression $R$ in both
+  directions.
 \end{itemize}
-%
 In addition, such primitives may be combined using the regular operators
@@
         . ins "\n"
-. (ins "%A " . valid_name . del " and " . ins "\n")$^*$ . ins "%A "
+        . (ins "%A " . valid_name . del " and " . ins "\n")* 
+        . ins "%A "
         . valid_name
@@
 \includegraphics[width=0.7\textwidth]{canonizers-outside}
-\caption{Quotient Lens with ``Canonizers at the Edges''\sam{This is a
-screenshot from Nate's thesis. Are we allowed to use it?}\bcp{We can
-(if we put ``from [Nate's thesis]'' in the caption), but I don't think
-we want the full complexity of asymmetric lenses here.  We should draw
-the figure again.  (I have the OmniGraffle file for it, so this is
-easy if I know how it should look.)}\bcp{Done---what do you think?}
-\sam{Yes this is more accurate, except that we don't need a general choose
-function. With QREs the kernel language is always contained in the whole
-language so we just use the inclusion as the choose function.}\bcp{I can
-fix.}} 
+\caption{Quotient Lens with ``Canonizers at the Edges''
+\sam{We don't need a general choose function. With QREs the kernel language is
+always contained in the whole language so we just use the inclusion as the
+choose function.}\bcp{I can fix.}}
 \label{fig:attheedges}
@@
 two equivalence relations to a regular language. For example, suppose that
-we wish to express a name as
-
-\noindent $\kw{last\_name \cdot ``,\text{''} \cdot wsp^+ \cdot first\_name}$,
-or as $\kw{first\_name\cdot wsp^+ \cdot last\_name}$, with
-
-\noindent $\kw{first\_name\cdot wsp^+ \cdot last\_name}$ as the canonical
-representation. We also wish to project away the extra whitespace in between
-\kw{first\_name} and \kw{last\_name} once this identification has been made.
-Then we can express this equivalence relation by a composition of a
+we wish to express a name as $\kw{last\_name \cdot ``,\text{''} \cdot
+\text{wsp+} \cdot first\_name}$, or as $\kw{first\_name\cdot \text{wsp+} \cdot
+last\_name}$, with $\kw{first\_name\cdot \text{wsp+} \cdot last\_name}$ as the
+canonical representation. We also wish to project away the extra whitespace in
+between \kw{first\_name} and \kw{last\_name} once this identification has been
+made. Then we can express this equivalence relation by a composition of a
 \textit{squash} QRE and a QRE that projects away the extra whitespace in between
@@
 $c'$ is well formed only if the concatentions of $W(c)$ and $W(c')$, and
-$K(c)$ and $K(c')$ are ambiguous.
+$K(c)$ and $K(c')$ are unambiguous.
 
@@
 \end{center}
-$\canonizer(\perm{c_1, \ldots, c_n}{c})(r_{\sigma(1)}
-\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \newline
-= (\canonizer(c_1)(r_1)) \cdot (\canonizer(c)(s_1)) \cdot \ldots \cdot
-(\canonizer(c_n)(r_n))$
+\begin{align*}
+&\canonizer(\perm{c_1, \ldots, c_n}{c})(r_{\sigma(1)}
+\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)})\\
+&= (\canonizer(c_1)(r_1)) \cdot (\canonizer(c)(s_1)) \cdot \ldots \cdot
+(\canonizer(c_n)(r_n))
+\end{align*}
 \caption{QRE canonizers}
@@
 
-\section{Conclusion and Future Work}
+\section{Conclusion}
 \label{concl}

### 2018-02-25 papers/QuotientLenses/main.tex
Added abstract

@@
 \begin{abstract}
-\saz{The old abstract read more like an introduction, so we modified it to be
-the ``new'' introduction.}
+While there currently exist many bidirectional programming languages, many of
+them do not have built-in support for transformations that operate on data
+modulo equivalence relations such as whitespace or permutation quotienting.
+Moreover, langauges such as Boomerang which do allow for such transformations
+are often difficult to program \sam{why?}. To address these shortcomings, this
+paper (1) introduces a language of {\em quotient regular expressions} or QREs
+for defining equivalences on regular expressions as well as a language of {\em
+QRE lenses}, which are lenses that map between QREs, and (2) shows how to {\em
+synthesize} QRE lenses from a pair of QREs and a set of example input-output
+pairs. We also describe an implementation of QREs and QRE lens synthesis in
+Boomerang, and demonstrate the utility of our implementation by using it to
+synthesize QRE lenses between data formats drawn from {\tt data.gov} database.
 \end{abstract}
 
-\keywords{quotient lenses, synthesis}
-\maketitle              % typeset the title of the contribution
+\keywords{quotient regular expressions, QRE lens, synthesis}
+\maketitle
 
@@
 \end{enumerate}
-Sections~\ref{relwork} and~\ref{concl} discuss related and future work.
+Sections~\ref{relwork} discusses related work and~\ref{concl} summarizes the
+results of the paper.
 
@@
 \subsection{Quotient Regular Expressions}
+\label{subsec:qre-expressions}
 
@@
 
-This section introduces Quotient Regular Expressions or QREs. QREs are regular
-expressions augmented with syntax that lets them simultaneously express a
-language and how to convert from that language into its canonical elements.
-
-\subsection{Syntax of Quotient Regular Expressions}
+Quotient Regular Expressions allow programs to specify a regular language,
+an equivalence relation over elements of the language and a function to convert
+elements of the language into normal form.
+
+\subsection{Syntax}
+
 The language of Quotient Regular Expressions (QREs) is given by the following
@@
 where $c$ ranges over QREs, $R$ ranges over regular expressions, $f$ ranges over
-functions between regular languages, and $s$ ranges over character strings. In
-the following sections, we will use the the notation $\mathcal{L}(R)$ to mean
+functions between regular languages, and $s$ ranges over character strings.
+$\mathcal{L}(R)$ is
 the language accepted by the regular expression $R$.
+
+%Section~\ref{subsec:qre-expressions} presented a number of examples
 

### 2018-02-25 papers/QuotientLenses/main.tex
performed suggested surgery

@@
 \begin{verbatim}
-@Book {conway,
-Author = "J. H. Conway",
-Title = {Regular Algebra and Finite Machines}
+@Book {Doe,
+Author = "John Doe",
+Title = {Generic Title}
 }
@@
 %0 Book
-%T Regular Algebra and Finite Machines
-%A J. H. Conway
-%F conway
+%T Generic Title
+%A John Doe
+%F Doe
 \end{verbatim}
@@
 irrelevant and there may be equivalent ways of writing the same
-name:  ``Conway, J. H.'' vs ``J. H. Conway.''
+name:  ``Doe, John'' vs ``John Doe.''
 
@@
 %% comma and then the other names separated by arbitrary whitespace, or just all
-%% the names separated by arbitrary whitespace (for example, the names ``Conway, J.
-%% H.'' and ``J. H. Conway'' are equivalent). Lastly, the transformation should
+%% the names separated by arbitrary whitespace (for example, the names ``Doe, J.
+%% H.'' and ``John Doe'' are equivalent). Lastly, the transformation should
 %% also ignore the order in which the author and title fields occur in each
@@
 \begin{verbatim}
-@Book {conway,
-Author = "J. H. Conway",
-Title = {Regular Algebra and Finite Machines},
+@Book {Doe,
+Author = "John Doe",
+Title = {Generic Title},
 }
 
-@Book{     conway,
-Title = {Regular Algebra
-and Finite Machines},
-Author = "Conway, J. H.",     }
+@Book{     Doe,
+Title = {Generic Title},
+Author = "Doe, John",     }
 \end{verbatim}
@@
 
-%% %Foster's Boomerang~\ref{boomerang} language presents one design
-%% While it is possible to express the equivalence on the \bibtex{} data above in
-%% Boomerang, actually doing so is a challenge.  Three of the inessential details
-%% we wish to ignore---whitespace quotienting
-%% inside fields, whitespace quotienting in between fields, and permutation
-%% quotenting of fields---nest within each other in a way that makes them difficult
-%% to express with the existing Boomerang combinators. Furthermore, if the
-%% programmer wishes to specify the type of the lens, then she can only specify
-%% data formats on each side of the lens with a regular expression that does not
-%% keep track of the equivalence relation defined on the language of the regular
-%% expression. Consequently, the programmer has to write a regular expression as
-%% well as a canonizer that expresses the equivalence relation on that regular
-%% expression, even though a regular expression and the canonizer defined on it
-%% typically share much of the syntactic structure. Having to write both of these
-%% expressions makes programming with quotient lenses verbose and time consuming.
-
 We propose to do away with much of this overhead, and the burden of a
@@
 \begin{lstlisting}
-let wsp = project [ \n\t\r]$^+$ $\to$ " "
+let p_wsp = project wsp+ $\to$ " "
 \end{lstlisting}
@@
 For instance, assume
-the \codefont{comma\_name} format describes ``Conway, J. H.''
-and the \codefont{space\_name} format describes ``J. H. Conway''
+the \codefont{comma\_name} format describes ``Doe, John''
+and the \codefont{space\_name} format describes ``John Doe''
 and \codefont{s\_to\_c} is a function from the first to the second.  In this case,
@@
 \begin{lstlisting}
-let l = synth comma_name $\Leftrightarrow$ space_name using {("Conway, J. H.", "J. H. Conway")}
+let l = synth comma_name $\Leftrightarrow$ space_name using {("Doe, John", "John Doe")}
 let s_to_c = get l
@@
 which is what we need for squash.
+
+Another QRE primitive is the composition combinator ``;''. This combinator
+applies two equivalence relations to a regular language in sequence. For
+example, suppose that we wish to express a name as $\kw{last\_name \cdot
+``,\text{''} \cdot \text{wsp+} \cdot first\_name}$, or as $\kw{first\_name\cdot
+\text{wsp+} \cdot last\_name}$, with $\kw{first\_name\cdot \text{wsp+} \cdot
+last\_name}$ as the canonical representation. We also wish to project away the
+extra whitespace in between \kw{first\_name} and \kw{last\_name} once this
+identification has been made. Then we can express this compound equivalence
+relation by a composition of a \textit{squash} QRE and a QRE that projects away
+the extra whitespace in between \kw{first\_name} and \kw{last\_name}
 
@@
 \begin{lstlisting}
-let bib_fields = perm (label, bib_author, bib_title) with (project ("," .wsp$^+$) $\to$ ",\n")
+let bib_fields = perm (label, bib_author, bib_title) with (project (wsp+ . "," . wsp+) $\to$ ",\n")
 \end{lstlisting}
 %
-\noindent
-The QRE for the bibtex format is presented in its entirety in Figure~\ref{fig:example-qre}.
-
+The final QRE combinator is the normalize combinator.
+This combinator allows a programmer to manually define a function
+$f$ which sends each string that matches a regular expression $R$ to some
+canonical representative in another regular expression $R'$. The equivalence
+relation defined by the normalize combinator is hence the equivalence relation
+defined by the {\em fibres} of $f$; that is, for all strings $s$ and $s'$ that
+match $R$, $s$ is equivalent to $s'$ if and only if $f(s) = f(s')$.
+
+For instance, assume that $f(s)=$ `` '' for all whitespace strings $s$. Then the
+project QRE \kw{p_wsp} which matches any non-zero number of whitespace
+characters and converts any such whitespace into a single space character can
+be expressed using the normalize combinator as
+\begin{lstlisting}
+let p_wsp = normalize (wsp+, " ", f)
+\end{lstlisting}
+The QRE for the bibtex format is presented in its entirety in
+Figure~\ref{fig:example-qre}.
 \begin{figure}[t]
 \begin{lstlisting}
-(* whitespace with " " as a canonical representative *)
-let wsp = project [ \n\t\r]$^+$ $\to$ " "
+let wsp = [ \n\t\r]
 
 (* define name representations with a space and with a comma *)
-let space_name = name | (name . wsp . (name . wsp)$^*$ . name)
-let comma_name = name | (name . "," . wsp . (name . wsp)$^*$ . name)
-
-(* synthesize a lens that maps the comma representation to the space representation *)
-let l = synth comma_name $\Leftrightarrow$ space_name using {("Conway, J. H.", "J. H. Conway")}
+let space_name = first_name . wsp+ . last_name
+let comma_name = last_name . "," . wsp+ . first_name
+
+(* synthesize a lens that maps comma representation to space representation *)
+let l = synth comma_name $\Leftrightarrow$ space_name using {("Doe, John", "John Doe")}
 let s_to_c = get l
@@
 (* squash QRE maps comma_name to space_name *)
-let name = squash comma_name $\to$ space_name using s_to_c
+let comma_to_space = squash comma_name $\to$ space_name using (get comma_to_space)
+
+(* map arbitrary whitespace into " " as a canonical representative *)
+let p_wsp = project wsp+ $\to$ " "
+
+(* Project away whitespace in space_name representation *)
+let project_in_space = first_name . p_wsp . last_name
+
+(* Compose the two QREs to get a canonical representation for a name *)
+let name = project_in_space ; comma_to_space
 
 (* define rest of bibtex fields *)
-let name_sep = wsp . "and" . wsp
+let name_sep = p_wsp . "and" . p_wsp
 let bib_names = name | (name . name_sep . (name . name_sep)$^*$ . name)
 let bib_author = "author = \"" . bib_names . "\""
-let title = word | (word . wsp . (word . wsp)$^*$ . word)
+let title = word | (word . p_wsp . (word . p_wsp)$^*$ . word)
 let bib_title = "title = {" . title . "}"
@@
 (* allow any permutation of fields interspersed with arbitrary whitespace *)
-let bib_fields = perm (label, bib_author, bib_title) with (project ("," .wsp$^+$) $\to$ ",\n")
+let bib_fields = perm (label, bib_author, bib_title) with (project (wsp+ . "," . wsp+) $\to$ ",\n")
 let bibtex = "@book{" . bib_fields . "}"
@@
 \end{figure}
-
-%% In Figure~\ref{fig:example-qre}, the QRE \verb+project [ \n\t\r]+$^{+} \to$
-%% \verb+" "+ matches any non-zero number of whitespace characters and quotients
-%% them to a single equivalence class with a single space as the canonical
-%% representative. We call this QRE a \textit{projection} because an entire regular
-%% language is projected to a single representative string.
-
-%% Another QRE in Figure~\ref{fig:example-qre} is given by
-
-%% \begin{center}
-%% \begin{lstlisting}
-%% squash comma_name $\to$ space_name using (get space_to_comma)
-%% \end{lstlisting}
-%% \end{center}
-
-%% \noindent This QRE is a \textit{squash} QRE. A \textit{squash} QRE ``glues''
-%% two data formats using a user-defined function. For example,
-%% the QRE in the code above specifies that the \cd{comma_name} representation of
-%% the author's name is equivalent to the \cd{space_name} representation, with
-%% \cd{space_name} taken to be the canonical representation. The
-%% function that identifies the two formats is given by \cd{get space_to_comma},
-%% which transforms names in the \cd{comma_name} representation to names in the
-%% \cd{space_name} representation. In this example, \cd{space_to_comma} is itself
-%% an automatically \textit{synthesized} lens between the \cd{space_name} and
-%% \cd{comma_name} QREs, and we need to use the ``get'' direction of the lens to
-%% specify the squash transformation.
-
-%% The third example of a QRE in the Figure is \cd{bib_fields}, defined by:
-%% \begin{center}
-%% \begin{lstlisting}
-%% perm (label, bib_author, bib_title) with (project ("," . wsp$^+$ ) $\to$ ",\n")
-%% \end{lstlisting}
-%% \end{center}
-
-%% \noindent This is a \textit{permutation} QRE that specifies that the label,
-%% title, and author fields can occur in any order, with the fields separated by a
-%% comma followed by arbitrary amounts of whitespace. The canonical representation
-%% is that in which the label, author and title fields occur in that order,
-%% separated by a comma followed by a newline. The \textit{permutation} QRE is
-%% designed for equivalences such as this, where different permutations of
-%% equivalences are interspersed with a separator which itself has an
-%% equivalence defined on it.
-
-%% Given the \textit{projection}, \textit{squash}, and \textit{permutation}
-%% QRE primitives, we can build more complex languages using the usual regular
-%% expression combinators of alternative choice `\cd{|}', concatenation `\cd{.}'
-%% and Kleene star `$^*$', and also sequential composition ``;''.
-
 \subsection{QRE Lenses}
@@
         . ins "\n%A "
-        . valid_name
+        . name
         . (del " and "
         . ins "\n"
-        . (ins "%A " . valid_name . del " and " . ins "\n")* 
+        . (ins "%A " . name . del " and " . ins "\n")* 
         . ins "%A "
-        . valid_name
+        . name
         || "")
@@
 \caption{An example lens between \cd{bibtex} and \cd{end_note} regular
-expressions with no equivalences.\bcp{Indentation needs work.  And there was
-a suggestion to colorize the new keywords that we're introducing here
-(including symbolic ones, I guess)}}
+expressions with no equivalences.}
 \label{fig:example-lens}
@@
 %% error-prone.
-
 Consequently, we offer the programmer some relief by identifying a class of
@@
 a single call to the synthesis prodedure:
-
 \begin{lstlisting}
@@
 \noindent Here, the generated  quotient lens synchronizes \cd{bibtex} and
-\cd{endnote} formats.  It uses \cd{bib_example} and \cd{end_example} (the two concrete example
-strings given at the beginning of this section) to disambiguate.
-In additiona, and as we saw earlier with the definition of \cd{s_to_c}, the synthesis procedure itself
-can be used to create lenses that are in turn used to define other QREs.  The ability to intermix
-QRE specification with QRE lens synthesis yields a powerful and flexible way of
-creating bidirectional transformations.
+\cd{endnote} formats.  It uses \cd{bib_example} and \cd{end_example} (the two
+concrete example strings givn at the beginning of this section) to disambiguate.
+In addition. as we saw earlier with the definition of \cd{s_to_c}, the synthesis
+procedure itself can be used to create lenses that are in turn used to define
+other QREs. The ability to mix QRE specification with QRE lens synthesis
+yields a powerful and flexible way of creating bidirectional transformations.
 
@@
 \label{QRE}
-
-Quotient Regular Expressions allow programs to specify a regular language,
-an equivalence relation over elements of the language and a function to convert
-elements of the language into normal form.
-
-\subsection{Syntax}
-
-The language of Quotient Regular Expressions (QREs) is given by the following
-grammar:
+\subsection{Syntax and Semantics of QREs}
+Formally, the language of Quotient Regular Expressions (QREs) is given by the
+following grammar:
 \begin{align*}
@@
 functions between regular languages, and $s$ ranges over character strings.
-$\mathcal{L}(R)$ is
-the language accepted by the regular expression $R$.
-
-%Section~\ref{subsec:qre-expressions} presented a number of examples
-
-\subsection{Overview of Quotient Regular Expression}
-In the \bibtex{} to EndNote example, we encountered the $\project{R}{s}$,
-
-\noindent $\squash{c}{c'}{f}$ and $\perm{c_1, \ldots, c_n}{c}$ QREs, as well as
-the regular operators applied to these, as in
-
-\begin{lstlisting}
-let space_name  = name | (name . wsp_to_space . (name . wsp_to_space)$^*$ . name)
-\end{lstlisting}
-
-One combinator that we haven't yet introduced is sequential composition.
-The $c \semicolon c'$ QRE successively applies
-two equivalence relations to a regular language. For example, suppose that
-we wish to express a name as $\kw{last\_name \cdot ``,\text{''} \cdot
-\text{wsp+} \cdot first\_name}$, or as $\kw{first\_name\cdot \text{wsp+} \cdot
-last\_name}$, with $\kw{first\_name\cdot \text{wsp+} \cdot last\_name}$ as the
-canonical representation. We also wish to project away the extra whitespace in
-between \kw{first\_name} and \kw{last\_name} once this identification has been
-made. Then we can express this equivalence relation by a composition of a
-\textit{squash} QRE and a QRE that projects away the extra whitespace in between
-\kw{first\_name} and \kw{last\_name}:
-
-\begin{lstlisting}
-(* Regular expressions matching different comma and space representations *)
-let comma_name = last_name . "," . wsp . first_name
-let space_name = first_name . wsp . last_name
-
-(* Squash comma representation into space representation *)
-let comma_to_space =
-synth comma_name <=> space_name using {("Conway, J. H.", "J. H. Conway")}
-let comma_to_space = squash comma_name $\to$ space_name using (get comma_to_space)
-
-(* Project away whitespace in space_name representation *)
-let project_in_space = first_name . (project wsp+ $\to$ " ") . last_name
-
-(* Compose the two QREs *)
-let valid_name = project_in_space ; comma_to_space
-\end{lstlisting}
-
-The final combinator is $\normalize{R}{R'}{f}$, which expresses
-the equivalence relation that sends each element in $\mathcal{L}(R)$ to its
-canonical representation $f(r) \in \mathcal{L}(R')$. The equivalence relation
-defined by $\normalize{R}{R'}{f}$ is thus the equivalence relation on
-$\mathcal{L}(R)$ defined by the {\em fibres} of $f$ (i.e. $r \sim r'$ if and
-only if $f(r) = f(r'))$. The canonizing function $f$ is required to be
-surjective and idempotent.
-
-The $\normalize{R}{R'}{f}$ QRE is the most general QRE in that for each QRE
-$q$, there exist regular expressions $R, R'$ and a surjective, idempotent
-function $f:\mathcal{L}(R) \longrightarrow \mathcal{L}(R')$ such that $q$ and
-$\normalize{R}{R'}{f}$ each define the same equivalence relation on
-$\mathcal{L}(R)$, with $\mathcal{L}(R')$ forming a complete set of
-representatives for this equivalence relation. The $\normalize{R}{R'}{f}$
-combinator is included so as to express equivalence relations that are either
-difficult to define, or that cannot be defined using the other combinators.
-\saz{Add something about being ``complete'' here?}\sam{I'm not sure how to do
-this here. The precise statement about how normalize is strong enough to
-express all QREs requires that we introduce W(c), K(c) and Canonizer(c) which
-we haven't introduced at this point in the paper}
-
-Now in order for the programmer to ensure that the {\em normalize} QRE is well
-formed, she must define a canonizing function $f$ between two regular
-languages and prove that $f$ is surjective and idempotent: the compiler cannot
-verify this since this problem is undecidable in general. Consequently, we do
-not use the $\normalize{R}{R'}{f}$ QRE in practice. On the other hand, the
-$\normalize{R}{R'}{f}$ combinator gives a sufficient condition for potential
-QREs to be ``well-behaved'', particularly with respect to our strategy for
-synthesizing quotient lenses which we shall introduce in Section~\ref{synth}.
-
-The equivalence relations definable by QREs are a strict subset of those that
-can be expressed in Boomerang. For instance, in Boomerang, one can use any lens
-to canonize a regular language described by $R$ to a regular language described
-by $S$, with $\mathcal{L}(S) \not \subseteq \mathcal{L}(R)$, whereas each QRE
-$c$ canonizes a regular language described by $R$ to regular subset
-$\mathcal{L}(R') \subseteq \mathcal{L}(R)$ described by $R'$. QREs therefore are
-not as expressive as Boomerang canonizers. On the other hand, QREs enable us to
-express many equivalences that occur in real world data as we shall demonstrate
-in the evaluation section (Section~\ref{impl}).
-
-The net effect of these design decisions means that each QRE $c$ enables
-us to simultaneously express
+
+Using the conventional notation that $\mathcal{L}(R)$ is the language accepted
+by the regular expression $R$, then each QRE $c$ enables us to express
 \begin{enumerate}
@@
 \end{enumerate}
-Observe that since $\canonizer(c)$ computes $\canonizer(c)(w)$ as the unique
-$k$ in $\mathcal{L}(K(c))$ such that $w$ is equivalent to $k$ mod $\eqrel{c}$,
-then the equivalence classes of $\eqrel{c}$ are the same as the fibres of
-$\canonizer(c)$.
-
-\subsection{Well Formed QREs and ambiguity}
-Before giving the formal semantics for QREs, we address the issue of {\em
-ambiguity}, which arises when defining QREs. If $c$ and  $c'$ are QREs, then
-when applying the regular combinators to QREs, we require that if a string $s$
-matches any of the regular expressions $W(c) \cdot W(c')$, $W(c) \sep W(c')$,
-$W(c)^*$, $K(c) \cdot K(c')$, $K(c) \sep K(c')$, $K(c)^*$, then $s$ matches
-that regular expression in only one way. This unambiguity condition is called
-{\em strong unambiguity}~\cite{Sippu1988}, and is necessary, firstly because it
-ensures that the canonizing function of a QRE is well-defined.
-
-For example consider the QRE
-
-\begin{lstlisting}
-let ambiguous : canonizer = "a"* . (project "a"* $\to$ "a")
-\end{lstlisting}
-
-\noindent The behaviour of \cd{ambiguous} is not-well defined since the string
-\cd{"aaa"} can be canonized to any of \cd{"a"}, \cd{"aa"}, or \cd{"aaa"} depending on how
-\cd{"aaa"} is parsed.
-
-We also require that the regular combinators applied to
-the kernels are unambiguous since the underlying lenses will end up operating
-on the kernels, and these lenses impose the same restrictions for similar
-reasons.
-%
-To this end, we say that regular expressions $R$ and $S$ are
-\textit{unambiguosly concatenable}, written $R \cdot^! S$ if for all strings
-$r, r' \in \mathcal{L}(R)$ and $s, s' \in \mathcal{L}(S)$, if $r \cdot s = r'
-\cdot s'$, then $r = r'$ and $s = s'$. We say that a regular expression $R$ is
-\textit{unambiguosly iterable}, written $R^{*!}$ if for all strings $r_1,
-\ldots, r_m$ and $r'_1, \ldots, r'_n \in \mathcal{L}(R)$, if $r_1 \cdot \ldots
-\cdot r_m = r'_1 \cdot \ldots \cdot r'_n$, then $m = n$ and $r_i = r'_i$.
-
-We say that a regular expression $R$ is \textit{strongly unambiguous} if and
-only if (1) $R = \varnothing$, or (2) $R = S_1 \cdot S_2$ with $S_1, S_2$
-strongly unambiguous and $S_1 \cdot^! S_n$, or (3) $R = S_1 \sep S_2$ with
-$S_1, S_2$ strongly unambiguous and $\mathcal{L}(S_1) \cap \mathcal{L}(S_2) =
-\varnothing$, or (4) $R = S^*$ with $S$ strongly unambiguous and $S^{*!}$.
-
-\begin{figure}[tp!]
-\begin{center}
-\AxiomC{$R$ is strongly unambiguous}
-\UnaryInfC{$\wf{\mathit{id}(R)}$}
-\DisplayProof
-\hskip 1.5em
-\AxiomC{$s \in \mathcal{L}(R)$}
-\UnaryInfC{$\wf{\project{R}{s}}$}
-\DisplayProof
-\hskip 1.5em
-\end{center}
-
-\begin{prooftree}
-\AxiomC{$\wf{c_i, c}$}
-\AxiomC{${\substack{\forall \sigma \neq \theta, \; W(c_{\sigma(1)}) \cdot W(c)
-\cdot \ldots \cdot W(c_{\sigma(n)})\\ \cap W(c_{\theta(1)}) \cdot W(c)
-\cdot \ldots \cdot W(c_{\theta(n)}) =\varnothing}}$}
-\AxiomC{$K(c_1) \cdot^! K(c) \cdot^! \ldots \cdot^! K(c) \cdot^! K(c_n)$}
-\TrinaryInfC{$\wf{\perm{c_1, \ldots, c_n}{c}}$}
-\end{prooftree}
-
-\begin{center}
-\AxiomC{$\wf{c, c'}$}
-\AxiomC{$\mathcal{L}(W(c)) \cap \mathcal{L}(W(c')) = \varnothing$}
-\AxiomC{$f : \mathcal{L}(W(c)) \longrightarrow \mathcal{L}(W(c'))$}
-\TrinaryInfC{$\wf{\squash{c}{c'}{f}}$}
-\DisplayProof
-\end{center}
-\begin{center}
-
-\begin{prooftree}
-\AxiomC{$\mathcal{L}(R') \subseteq \mathcal{L}(R)$}
-\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(R')$}
-\AxiomC{$f$ is surjective}
-\AxiomC{$f = f^2$}
-\QuaternaryInfC{$\wf{\normalize{R}{R'}{f}}$}
-\end{prooftree}
-
-\end{center}
-\begin{center}
-\AxiomC{$\wf{c, c'}$}
-\AxiomC{$K(c) = W(c')$}
-\BinaryInfC{$\wf{c' \; ; \; c}$}
-\DisplayProof
-\hskip 1.5em
-\AxiomC{$\wf{c}$}
-\AxiomC{$W(c)^{*!}$}
-\AxiomC{$K(c)^{*!}$}
-\TrinaryInfC{$\wf{c^*}$}
-\DisplayProof
-\end{center}
-\begin{center}
-
-\begin{center}
-\AxiomC{$\wf{c, c'}$}
-\AxiomC{$W(c) \cdot^! W(c')$}
-\AxiomC{$K(c) \cdot^! K(c')$}
-\TrinaryInfC{$\wf{c \cdot c'}$}
-\DisplayProof
-\end{center}
-
-\begin{center}
-\AxiomC{$\wf{c, c'}$}
-\AxiomC{$W(c) \cap W(c') = \varnothing$}
-\BinaryInfC{$\wf{c \cdot c}$}
-\DisplayProof
-\hskip 1.5em
-\AxiomC{$\wf{c}$}
-\AxiomC{$W(c)^{*!}$}
-\AxiomC{$K(c)^{*!}$}
-\TrinaryInfC{$\wf{c^*}$}
-\DisplayProof
-\end{center}
-\end{center}
-\caption{Well-formed QREs
-}
-\label{fig:qrerules}
-\end{figure}
-
-Figure~\ref{fig:qrerules} gives the inferrence rules for deriving well-formed
-QREs. As expected, the unambiguity conditions come into play when defining QREs
-using the regular combinators. For example, the inferrence rule for
-concatenation is as follows:
-\begin{prooftree}
-\AxiomC{$\wf{c, c'}$}
-\AxiomC{$W(c) \cdot^! W(c')$}
-\AxiomC{$K(c) \cdot^! K(c')$}
-\TrinaryInfC{$\wf{c \cdot c'}$}
-\end{prooftree}
-
-\noindent This rule says that the concatenation $c \cdot c'$ of QREs $c$ and
-$c'$ is well formed only if the concatentions of $W(c)$ and $W(c')$, and
-$K(c)$ and $K(c')$ are unambiguous.
-
-The most complicated inferrence rule is the rule for the \textit{permutation}
-combinator:
-
-\begin{prooftree}
-\AxiomC{$\wf{c_i, c}$}
-\AxiomC{${\substack{\forall \sigma \neq \theta, \; W(c_{\sigma(1)}) \cdot W(c)
-\cdot \ldots \cdot W(c_{\sigma(n)})\\ \cap W(c_{\theta(1)}) \cdot W(c)
-\cdot \ldots \cdot W(c_{\theta(n)}) =\varnothing}}$}
-\AxiomC{$K(c_1) \cdot^! K(c) \cdot^! \ldots \cdot^! K(c) \cdot^! K(c_n)$}
-\TrinaryInfC{$\wf{\perm{c_1, \ldots, c_n}{c}}$}
-\end{prooftree}
-
-\noindent The second hypothesis for the \textit{permutation} rule says that for
-any two different permutations $\sigma$ and $\theta$, the languages
-$W(c_{\sigma(1)}) \cdot W(c) \cdot \ldots \cdot W(c_{\sigma(n)})$ and
-$W(c_{\theta(1)}) \cdot W(c) \cdot \ldots \cdot W(c_{\theta(n)})$ must be
-disjoint. This is because any string that matches the \textit{permutation}
-combinator matches the language $W(c_{\theta(1)}) \cdot W(c) \cdot \ldots \cdot
-W(c_{\theta(n)})$ for some permutation $\theta$, so we require that these
-regular expressions be disjoint.
-
-The third hypothesis says that the regular expression $K(c_1) \cdot K(c)
-\cdot \ldots \cdot K(c) \cdot K(c_n)$ must be unambiguous. This is because the
-underlying lens that maps to or from a \textit{permutation} QRE operates on
-the language $K(c_1) \cdot K(c) \cdot \ldots \cdot K(c) \cdot K(c_n)$ which is
-the kernel of the \textit{permutation} QRE. As we mentioned, the underlying
-lens will require that the source and target regular expressions of the lens be
-unambiguous so that the lens can match strings uniquely.
-
-The rest of the inferrence rules for deriving QREs are given in
-Figure~\ref{fig:qrerules}.
-\subsection{Semantics of QREs}
-As we mentioned before, each QRE $c$ enables us to express the whole and kernel
-languages $W(c)$ and $K(c)$, an equivalence relation $\eqrel{c}$ on
-$\mathcal{L}(W(c))$ that has $\mathcal{L}(K(c))$ as a complete set of
-representatives and a canonizing function $\canonizer(c):\mathcal{L}(W(c))
-\longrightarrow \mathcal{L}(K(c))$ whose fibres are the same as $\eqrel{c}$.
-
 Figures~\ref{fig:wk} gives the inductive definitions for the whole and
-kernel languages $W(c)$ and $K(c)$ of a QRE. The \textit{squash} and
-\textit{permutation} combinators give the two interesting definitions. If $c =
-\squash{c}{c'}{f}$, then the whole language of $c$ is $W(c) \sep W(c')$. This
-is because the \textit{squash} combinator takes the whole language $W(c)$ of
-$c$, the whole language $W(c')$ of $c'$ and a function $f : \mathcal{L}(W(c))
-\longrightarrow \mathcal{L}(W(c'))$, maps $\mathcal{L}(W(c))$ into
-$\mathcal{L}(W(c'))$ using $f$ and then canonizes $W(c')$ into $K(c')$ using
-$\canonizer(c')$.
-
-For the $\perm{(c_1, \ldots, c_n)}{c}$ combinator, the whole language is the
-union of languages of the form $W(c_{\sigma(1)}) \cdot W(c) \cdot \ldots \cdot
-W(c) \cdot W(c_{\sigma(n)})$. This is because the \textit{permutation}
-combinator allows for the string to match any permutation the $c_i$'s while also
-taking into account the separator $c$ in between each of the $c_i$'s. The kernel
-of the \textit{permutation} combinator is the language $K(c_1) \cdot K(c) \cdot
-\ldots \cdot K(c) \cdot K(c_n)$, because the canonical permutation is the
-identity permuation, with each of the parts of the string that match $c_i$
-and $c$ canonized into $K(c_i)$ and $K(c)$ respectively.
-
-Figure~\ref{fig:canonizers} gives the inductive definitions of the
-$\canonizer{}$ function for each QRE. Again, the \textit{permutation} combinator
-gives the most interesting definition,
-\begin{align*}\canonizer(\perm{c_1, \ldots, c_n}{c})(r_{\sigma(1)}
-\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \\
-= (\canonizer(c_1)(r_1)) \cdot (\canonizer(c)(s_1)) \cdot \ldots \cdot
-(\canonizer(c_n)(r_n))
-\end{align*}
-\noindent which as we have explained is defined so that the strings that match
-$c_i, c$ are placed in the canonical permutation which is the identity
-permutation before $\canonizer(c_i)$ and $\canonizer(c)$ are applied
-respectively.
-
-Finally, Figure ~\ref{fig:relations} gives the induction defintion of the
-equivalence relation $\eqrel{c}$. These are the set-theoretic semantics of QREs
-as an equivalence relation on a regular language.
-
+kernel languages $W(c)$ and $K(c)$ of a QRE.
 \begin{figure}[t]
@@
 \end{figure}
-
+The \textit{squash} and \textit{permutation} combinators give the two
+interesting definitions. If $c = \squash{c}{c'}{f}$, then the whole language of
+$c$ is $W(c) \sep W(c')$. This is because the \textit{squash} combinator takes
+the whole language $W(c)$ of $c$, the whole language $W(c')$ of $c'$ and a
+function $f : \mathcal{L}(W(c)) \longrightarrow \mathcal{L}(W(c'))$, maps
+$\mathcal{L}(W(c))$ into $\mathcal{L}(W(c'))$ using $f$ and then canonizes
+$W(c')$ into $K(c')$ using $\canonizer(c')$.
+
+For the $\perm{c_1, \ldots, c_n}{c}$ combinator, the whole language is the
+union of languages of the form $W(c_{\sigma(1)}) \cdot W(c) \cdot \ldots \cdot
+W(c) \cdot W(c_{\sigma(n)})$. This is because the \textit{permutation}
+combinator allows for the string to match any permutation the $c_i$'s while also
+taking into account the separator $c$ in between each of the $c_i$'s. The kernel
+of the \textit{permutation} combinator is the language $K(c_1) \cdot K(c) \cdot
+\ldots \cdot K(c) \cdot K(c_n)$, because the canonical permutation is the
+identity permuation, with each of the parts of the string that match $c_i$
+and $c$ canonized into $K(c_i)$ and $K(c)$ respectively.
+
+Figure~\ref{fig:canonizers} gives the inductive definitions of the
+$\canonizer{}$ function for each QRE.
 \begin{figure}[t]
@@
 \begin{cases}
-\canonizer(c')(f(w)) & \text{if } w \in \mathcal{L}(W(c))\\
+\canonizer(c)(f(w)) & \text{if } w \in \mathcal{L}(W(c))\\
 \canonizer(c')(w) & \text{otherwise}
@@
 \end{figure}
+Again, the \textit{permutation} combinator gives the most interesting
+definition,
+\begin{align*}\canonizer(\perm{c_1, \ldots, c_n}{c})(r_{\sigma(1)}
+\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \\
+= (\canonizer(c_1)(r_1)) \cdot (\canonizer(c)(s_1)) \cdot \ldots \cdot
+(\canonizer(c_n)(r_n))
+\end{align*}
+\noindent which as we have explained is defined so that the strings that match
+$c_i, c$ are placed in the canonical permutation which is the identity
+permutation before $\canonizer(c_i)$ and $\canonizer(c)$ are applied
+respectively.
+
+Finally, Figure ~\ref{fig:relations} gives the induction defintion of the
+equivalence relation $\eqrel{c}$. These are the set-theoretic semantics of QREs
+as an equivalence relation on a regular language.
 
@@
 \end{figure}
+\subsection{Ambiguity and Well Formed QREs}
+Before giving the inferrence rules for deriving well formed QREs, we address
+the issue of {\em ambiguity}, which appears in the premises of the inferrence
+rules for the regular combinators. If $c$ and $c'$ are QREs, then when applying
+the regular combinators to QREs, we require that if a string $s$ matches any of
+the regular expressions $W(c) \cdot W(c')$, $W(c) \sep W(c')$, $W(c)^*$, $K(c)
+\cdot K(c')$, $K(c) \sep K(c')$, $K(c)^*$, then $s$ matches that regular
+expression in only one way. This unambiguity condition is called {\em strong
+unambiguity}~\cite{Sippu1988}, and is necessary because it ensures
+that the canonizing function of a QRE is well-defined.
+
+For example consider the QRE
+
+\begin{lstlisting}
+let ambiguous : canonizer = "a"* . (project "a"* $\to$ "a")
+\end{lstlisting}
+
+\noindent The behaviour of \cd{ambiguous} is not-well defined since the string
+\cd{"aaa"} can be canonized to any of \cd{"a"}, \cd{"aa"}, or \cd{"aaa"} depending on how
+\cd{"aaa"} is parsed. We also require that the regular combinators applied to
+the kernels be unambiguous since the underlying lenses will end up operating
+on the kernels, and these lenses impose the same restrictions for similar
+reasons.
+
+To this end, we say that regular expressions $R$ and $S$ are
+\textit{unambiguosly concatenable}, written $R \cdot^! S$ if for all strings
+$r, r' \in \mathcal{L}(R)$ and $s, s' \in \mathcal{L}(S)$, if $r \cdot s = r'
+\cdot s'$, then $r = r'$ and $s = s'$. We say that a regular expression $R$ is
+\textit{unambiguosly iterable}, written $R^{*!}$ if for all strings $r_1,
+\ldots, r_m$ and $r'_1, \ldots, r'_n \in \mathcal{L}(R)$, if $r_1 \cdot \ldots
+\cdot r_m = r'_1 \cdot \ldots \cdot r'_n$, then $m = n$ and $r_i = r'_i$.
+
+We say that a regular expression $R$ is \textit{strongly unambiguous} if and
+only if (1) $R = \varnothing$, or (2) $R = S_1 \cdot S_2$ with $S_1, S_2$
+strongly unambiguous and $S_1 \cdot^! S_n$, or (3) $R = S_1 \sep S_2$ with
+$S_1, S_2$ strongly unambiguous and $\mathcal{L}(S_1) \cap \mathcal{L}(S_2) =
+\varnothing$, or (4) $R = S^*$ with $S$ strongly unambiguous and $S^{*!}$.
+
+Figure~\ref{fig:qrerules} gives the inferrence rules for deriving well-formed
+QREs.
+\begin{figure}[tp!]
+\begin{center}
+\AxiomC{$R$ is strongly unambiguous}
+\UnaryInfC{$\wf{\mathit{id}(R)}$}
+\DisplayProof
+\hskip 1.5em
+\AxiomC{$s \in \mathcal{L}(R)$}
+\UnaryInfC{$\wf{\project{R}{s}}$}
+\DisplayProof
+\hskip 1.5em
+\end{center}
+
+\begin{prooftree}
+\AxiomC{$\wf{c_i, c}$}
+\AxiomC{${\substack{\forall \sigma \neq \theta, \; W(c_{\sigma(1)}) \cdot W(c)
+\cdot \ldots \cdot W(c_{\sigma(n)})\\ \cap W(c_{\theta(1)}) \cdot W(c)
+\cdot \ldots \cdot W(c_{\theta(n)}) =\varnothing}}$}
+\AxiomC{$K(c_1) \cdot^! K(c) \cdot^! \ldots \cdot^! K(c) \cdot^! K(c_n)$}
+\TrinaryInfC{$\wf{\perm{c_1, \ldots, c_n}{c}}$}
+\end{prooftree}
+
+\begin{center}
+\AxiomC{$\wf{c, c'}$}
+\AxiomC{$\mathcal{L}(W(c)) \cap \mathcal{L}(W(c')) = \varnothing$}
+\AxiomC{$f : \mathcal{L}(W(c)) \longrightarrow \mathcal{L}(W(c'))$}
+\TrinaryInfC{$\wf{\squash{c}{c'}{f}}$}
+\DisplayProof
+\end{center}
+\begin{center}
+
+\begin{prooftree}
+\AxiomC{$\mathcal{L}(R') \subseteq \mathcal{L}(R)$}
+\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(R')$}
+\AxiomC{$f$ is surjective}
+\AxiomC{$f = f^2$}
+\QuaternaryInfC{$\wf{\normalize{R}{R'}{f}}$}
+\end{prooftree}
+
+\end{center}
+\begin{center}
+\AxiomC{$\wf{c, c'}$}
+\AxiomC{$K(c) = W(c')$}
+\BinaryInfC{$\wf{c' \; ; \; c}$}
+\DisplayProof
+\hskip 1.5em
+\AxiomC{$\wf{c}$}
+\AxiomC{$W(c)^{*!}$}
+\AxiomC{$K(c)^{*!}$}
+\TrinaryInfC{$\wf{c^*}$}
+\DisplayProof
+\end{center}
+\begin{center}
+
+\begin{center}
+\AxiomC{$\wf{c, c'}$}
+\AxiomC{$W(c) \cdot^! W(c')$}
+\AxiomC{$K(c) \cdot^! K(c')$}
+\TrinaryInfC{$\wf{c \cdot c'}$}
+\DisplayProof
+\end{center}
+
+\begin{center}
+\AxiomC{$\wf{c, c'}$}
+\AxiomC{$W(c) \cap W(c') = \varnothing$}
+\BinaryInfC{$\wf{c \cdot c}$}
+\DisplayProof
+\hskip 1.5em
+\AxiomC{$\wf{c}$}
+\AxiomC{$W(c)^{*!}$}
+\AxiomC{$K(c)^{*!}$}
+\TrinaryInfC{$\wf{c^*}$}
+\DisplayProof
+\end{center}
+\end{center}
+\caption{Well-formed QREs
+}
+\label{fig:qrerules}
+\end{figure}
+As we already mentioned, the unambiguity conditions come into play when defining
+QREs using the regular combinators. For example, the inferrence rule for
+concatenation is as follows:
+\begin{prooftree}
+\AxiomC{$\wf{c, c'}$}
+\AxiomC{$W(c) \cdot^! W(c')$}
+\AxiomC{$K(c) \cdot^! K(c')$}
+\TrinaryInfC{$\wf{c \cdot c'}$}
+\end{prooftree}
+
+\noindent This rule says that the concatenation $c \cdot c'$ of QREs $c$ and
+$c'$ is well formed only if the concatentions of $W(c)$ and $W(c')$, and
+$K(c)$ and $K(c')$ are unambiguous.
+
+The most complicated inferrence rule is the rule for the \textit{permutation}
+combinator:
+
+\begin{prooftree}
+\AxiomC{$\wf{c_i, c}$}
+\AxiomC{${\substack{\forall \sigma \neq \theta, \; W(c_{\sigma(1)}) \cdot W(c)
+\cdot \ldots \cdot W(c_{\sigma(n)})\\ \cap W(c_{\theta(1)}) \cdot W(c)
+\cdot \ldots \cdot W(c_{\theta(n)}) =\varnothing}}$}
+\AxiomC{$K(c_1) \cdot^! K(c) \cdot^! \ldots \cdot^! K(c) \cdot^! K(c_n)$}
+\TrinaryInfC{$\wf{\perm{c_1, \ldots, c_n}{c}}$}
+\end{prooftree}
+
+\noindent The second hypothesis for the \textit{permutation} rule says that for
+any two different permutations $\sigma$ and $\theta$, the languages
+$W(c_{\sigma(1)}) \cdot W(c) \cdot \ldots \cdot W(c_{\sigma(n)})$ and
+$W(c_{\theta(1)}) \cdot W(c) \cdot \ldots \cdot W(c_{\theta(n)})$ must be
+disjoint. This is because any string that matches the \textit{permutation}
+combinator matches the language $W(c_{\theta(1)}) \cdot W(c) \cdot \ldots \cdot
+W(c_{\theta(n)})$ for some permutation $\theta$, so we require that these
+regular expressions be disjoint.
+
+The third hypothesis says that the regular expression $K(c_1) \cdot K(c)
+\cdot \ldots \cdot K(c) \cdot K(c_n)$ must be unambiguous. This is because the
+underlying lens that maps to or from a \textit{permutation} QRE operates on
+the language $K(c_1) \cdot K(c) \cdot \ldots \cdot K(c) \cdot K(c_n)$ which is
+the kernel of the \textit{permutation} QRE. As we mentioned, the underlying
+lens will require that the source and target regular expressions of the lens be
+unambiguous so that the lens can match strings uniquely.
+
+At this point we also bring to attention the fact that in order for the {\em
+normalize} QRE to be well-formed, then canonizing function $f$ used by this
+combinator must be surjective and idempotent. Unfortunately, verifying
+that a function satisfies these properties is in general undecidable.
+Nonetheless, we allow a programmer to use normalize at their own risk without
+checking these side conditions.
+
+We do this firstly because the {\em normalize} combinator allows one to express
+equivalence relations that are either difficult to define, or that cannot be
+defined using the other combinators. Secondly, the {\em normalize} canonizer
+enjoys the property that each of the other combinators $c$ is semantically
+equivalent to $\normalize{W(c)}{K(c)}{f}$ for some surjective, idempotent
+function $f : \mathcal{L}(W(c)) \longrightarrow \mathcal{L}(K(c))$. The {\em
+normalize} combinator hence gives a sufficient condition for potential QREs to
+be ``well-behaved'', particularly with respect to our strategy for synthesizing
+quotient lenses which we shall introduce in Section~\ref{synth}.
+
+In spite of this ``completeness'' property of the {\em normalize} combinator,
+our experience with QREs shows that using the {\em project}, {\em squash} and
+{\em perm} combinators as well as the regular operators applied to these is
+much easier than using just {\em normalize} because of the compositional nature
+of the other combinators.
 
@@
 \end{theorem}
-
 \section{QRE Lenses}
@@
 
-\subsection{Normal Forms of QRE Lenses}
+\subsection{Normal Forms for QRE Lenses}
 Recall that our approach in defining QRE lenses is to have each QRE lens $q: c
@@
 bib_example =
-"@Book {conway,
-Author = \"J. H. Conway\",
-Title = {Regular Algebra and Finite Machines},
+"@Book {Doe,
+Author = \"John Doe\",
+Title = {Generic Title},
 }"
@@
 "%0 Book
-%T Regular Algebra and Finite Machines
-%A J. H. Conway
-%F conway"
+%T Generic Title
+%A John Doe
+%F Doe"
 \end{lstlisting}

### 2018-02-26 papers/QuotientLenses/main.tex
checkpoint

--- 2018-02-25 papers/QuotientLenses/main.tex
@@
 where $c$ ranges over QREs, $R$ ranges over regular expressions, $f$ ranges over
-functions between regular languages, and $s$ ranges over character strings. In
-the following sections, we will use the the notation $\mathcal{L}(R)$ to mean
+functions between regular languages, and $s$ ranges over character strings.
+$\mathcal{L}(R)$ is
 the language accepted by the regular expression $R$.
+
+Section~\ref{subsec:qre-expressions} presented a number of examples
 

### 2018-03-04 papers/QuotientLenses/main.tex
edit
minor fixes

--- 2018-02-25 papers/QuotientLenses/main.tex
@@
 \usepackage{proof}
+\usepackage{xspace}
 
@@
 
-\newcommand{\Name}{Optometrist}
+\newcommand{\Name}{Optometrist\xspace}
 
@@
 \newcommand{\OpticianRuntime}{\textbf{OO}}
-\newcommand{\SystemOnOptician}{\textbf{SO}}
-\newcommand{\SystemOnBenchmarks}{\textbf{SS}}
+\newcommand{\SystemOnOptician}{\textbf{QO}}
+\newcommand{\SystemOnBenchmarks}{\textbf{QQ}}
 \newcommand{\cd}[1]{\lstinline[backgroundcolor=\color{white}]$#1$}
@@
 
-We have integrated QREs and quotient lens synthesis into Boomerang allowing QREs
-and synthesis tasks to be expressed alongside existing Boomerang expressions.
+We have integrated QREs and quotient lens synthesis into Boomerang. With this
+integration, QREs and quotient lens synthesis tasks can be expressed alongside
+existing Boomerang terms.
 
@@
 \begin{enumerate}
-  \item Does synthesizing quotient lenses from QREs permit for an easier
-  development process than existing approaches?
-
-  \item Are QREs sufficiently expressive that more lenses can be synthesized
-  without modifications to the format?
-
-  \item Is our synthesis algorithm competitive with existing lens synthesis
-  tools?
+\item Does synthesizing quotient lenses from QREs permit an easier development
+  process than existing approaches?
+
+\item Does the increase in expressivity gained from synthesizing from QREs
+  permit synthesis of more useful lenses than existing approaches.
+
+\item Does our tool synthesize lenses fast enough to be used as part of a
+  standard development process.
 \end{enumerate}
@@
 
-\begin{figure}
+\begin{figure}[t]
 \includegraphics{generated-graphs/asts.eps}
 \caption{Count of benchmark programs definable using a given AST count. We find
-that it takes far fewer AST nodes to define benchmark lenses in \Name{} than
-with Optician or without synthesis. }
+  that it takes far fewer AST nodes to define benchmark lenses using QRE
+  synthesis than with Optician without QRE synthesis or without synthesis. }
 \label{fig:asts}
@@
 \begin{itemize}
-  \item[\QRESize{}] \QRESize{} is the number of AST nodes in the QRE
+  \item[\QRESize{}] is the number of AST nodes in the QRE
   specifications involved in QRE lens synthesis task.
-  \item[\CanonizerAndSpecSize{}] \CanonizerAndSpecSize{} is the number of
+  \item[\CanonizerAndSpecSize{}] is the number of
   AST nodes in $W(q)$ summed with the number of AST nodes in $\canonizer(q)$,
   for each QRE used in QRE lens synthesis task.
-  \item[\LensAndSpecSize{}] \LensAndSpecSize{} is the number of AST nodes
+  \item[\LensAndSpecSize{}] is the number of AST nodes
   in $W(q)$ summed with the number of AST nodes in the synthesized QRE lens.
@@
 The value of \CanonizerAndSpecSize{} represents the difficulty in writing our
-benchmarks without QREs.  This is only an approximation, as both $W(q)$ and
-$\canonizer(q)$ are automatically generated from the QRE.  In our
-experience, these automatically generated regular expressions and canonizers are
-the ones we would naturally write.
-
-The value of \LensAndSpecSize{} represents the difficulty in writing our benchmarks
-with neither QREs nor synthesis.  This is also an approximation, as $W(q)$ and
-$\canonizer(q)$ and the bijective lens are all automatically generated.
-We have found that the generated bijective lens is fairly similar in size to the
-one we would manually write, though it is typically less modular.
+benchmarks within the Optician system. This is only an approximation, as both
+$W(q)$ and $\canonizer(q)$ are automatically generated from the QRE, instead of
+written by a human.
+
+The value of \LensAndSpecSize{} represents the difficulty in writing our
+benchmarks in Boomerang with neither QREs nor synthesis. This is also an
+approximation, as $W(q)$ and $\canonizer(q)$ and the bijective lens are all
+automatically generated.
 
@@
 
-\subsection{Increase of Expressivity}
+\subsection{Increasing Expressivity}
 
 Optician was only able to synthesize fully bijective data transformations.
-\Name{} loosens that restriction, and is able to synthesize data which is
-bijective when put in canonical form.
-To determine the increase of expressivity, we analyzed:
+Synthesizing from QREs loosens that restriction, making it possible to
+synthesize data which is bijective when put in canonical form. To determine the
+increase of expressivity, we analyzed:
 \begin{enumerate}
@@
 We found that 10 of Opticians 39 benchmarks required modifications to be
-bijective. We found that 4 of Optician's 39
-benchmarks required modifications to be in bijective correspondence when in
-canonical form.  Furthermore, we found that all of the data.gov examples
-required no modifications to be in canonical form.  While we did not put the
-data.gov examples into a form suitable for Optician, we found only 4 of
-the 8 data.gov examples did not require canonization, so making suitable
-for Optician would have required a modification of 4 of the data.gov
-formats.
-
-\subsection{Competitive Performance}
-
-\begin{figure}
+bijective. We found that 4 of Optician's 39 benchmarks required modifications to
+be in bijective correspondence when in canonical form. Furthermore, we found
+that all of the data.gov examples required no modifications to be in bijective
+correspondence when in canonical form. While we did not put the data.gov
+examples into a form suitable for Optician, we found only 4 of the 8 data.gov
+examples did not require canonization, so making suitable for Optician would
+have required a modification of 4 of the data.gov formats.
+
+\subsection{Maintaining Competitive Performance}
+
+\begin{figure}[t]
 \centering
@@
 \end{subfigure}
-\caption{Runtimes of Optician and \Name{}.
-In (a), we run Optician and \Name{} on Optician's benchmarks.  We find that
+\caption{Runtimes of Optician and our system.
+In (a), we run Optician and our system on Optician's benchmarks.  We find that
 there is only a negligible difference incurred by Optometrist's overhead.
@@
 
-While we expanded on the Optician system, we found it important to validate
-performance did not degrade in this expansion.  To validate this, we tested:
+To be used in a development environment, we found it important to validate
+performance did not significantly degrade from Optician. To validate this, we
+tested the runtimes when:
 \begin{itemize}
   \item[\OpticianRuntime{}] Optician runs on its own benchmarks.
-  \item[\SystemOnOptician{}] \Name{} runs on Optician's benchmarks.
-  \item[\SystemOnBenchmarks{}] \Name{} runs on its own benchmarks.
+  \item[\SystemOnOptician{}] Our system runs on Optician's benchmarks.
+  \item[\SystemOnBenchmarks{}] Our system runs on its own benchmarks.
 \end{itemize}
@@
 We summarize the results of these experiments in Figure~\ref{fig:times}.
-\Name{} was
+Our system was
 able to synthesize all of Optician's benchmarks at a speed competitive with
@@
 Also notable was one of the data.gov examples, which converted demographic
-statistics by zip code from xml form into rdf form.  This merely took longer because it was
-an exceptionally large format, converting over 40 fields.  After adding in additional
-fields to make this format bijective, it still took over 15 seconds to complete.
+statistics by zip code from xml form into rdf form. This merely took longer
+because it was an exceptionally large format, converting over 40 fields. We
+added in additional fields to make this format suitable for Optician, which
+took over 15 seconds synthesize the modified lens.
 

### 2018-03-05 papers/QuotientLenses/main.tex
pass up to 3.1

--- 2018-02-25 papers/QuotientLenses/main.tex
@@
 language=Caml, 
-basicstyle=\upshape\sffamily,
+basicstyle=\linespread{0.8}\upshape\sffamily,
 keywordstyle=\upshape\sffamily\color{dkblue},
@@
 \newcommand{\project}[2]{\ensuremath{\kw{project} \; #1 \mapsto #2}}
-\newcommand{\squash}[3]{\ensuremath{\kw{squash} \; #1 \rightarrow #2
-\kw{using} \; #3}}
+\newcommand{\squash}[3]{\ensuremath{\kw{squash} \; #1 \rightarrow #2\; \kw{using} \; #3}}
 \newcommand{\perm}[2]{\ensuremath{\kw{perm}\; (#1)\; \kw{with}\; #2}}
@@
 BiGul~\cite{ko2016bigul}, typically come with a variety of guarantees,
-the \emph{lens laws}, that help ensure put and get functions are effective
-inverses of one another.
-%comprise such synchronization functions, and the ``lens
+\emph{lens laws}, that help ensure put and get functions are effective
+inverses of one another.%comprise such synchronization functions, and the ``lens
 %laws'' provide restrictions on how these synchronization functions interoperate%.
@@
 Many data formats have other ad hoc rules as well.
-Returning to our bibliography example, recall that in \bibtex{},
+In bibliographies, for example, recall that in \bibtex{},
 an author can be specified as
@@
 beginners, as they require programmers to think in two directions at once
-and learn a new type system, which comes with some tricky ambiguity constraints.
+and to learn a new type system, which comes with some tricky ambiguity constraints.
 Unfortunately, quotient lenses do not improve the situation:  To
@@
 a suite of examples that demonstrates how the
-lens should act on specific examples.  Given this information,
+lens should act on certain inputs.  Given this information,
 we are able to exploit and extend recent work on algorithms for synthesis of
@@
   \item We introduce {\em Quotient Regular Expressions} (QREs),
-  a compact, convenient notation for defining (1) a
-  regular language modulo some equivalence relation, and (2) a canonizer
+  a compact, convenient notation for simultaneously defining a
+  regular language modulo some equivalence relation, and a canonizer
   for that relation (Section~\ref{QRE}).
@@
 data, but we are interested in \emph{quotient lenses}, which define
-bijections modulo some equivalence relation.  The remains of this section
-describe and motivate a new language and system for defining such lenses.
+bijections modulo some equivalence relation.  The remainder of this section
+describes and motivates a new language and system for defining such lenses.
 
@@
 adds a little more info, but not much.  Should this be cut?  Should the intro be expanded
-slightly?  Should it be left as is (more or less)?}
+slightly?  Should it be left as is (more or less)?} \saz{I think that this is
+not too redundant and that it can be left as-is, modulo space crunching.}
 
 \begin{enumerate}
-  \item A pair of descriptions of the external source and target formats.
+  \item A pair of descriptions of the external source and target formats, here
+    \bibtex{} and EndNote.
   Such descriptions may be given by regular expressions, for instance.
@@
 difficult to
-master.  For instance, to quotient out whitespace inside, and in between,
+master.  For instance, to quotient out whitespace inside, and between,
 fields while also ignoring field order, one must nest canonizing combinators
-within one another in a somewhat tricky way.
+in a somewhat tricky way.
 
@@
 (choice `\cd{|}', concatenation `\cd{.}',
-and Kleene star `$^*$').  Users may also treat QREs as pairs of
-get and put functions and compose such pairs (the functional
-composition operator `;').
+and Kleene star `$^*$').  % Users may also treat QREs as pairs of
+% get and put functions and compose such pairs using the functional
+% composition operator `;'.
 
@@
 any such whitespace into a single space character (its canonical form).
-Such a definition uses the \kw{project} primitive.
+This whitespace-normalizing QRE is easily defined by using the \kw{project} primitive:
 
@@
 allows external data to be in either format, and converts any
-data in one format to the other.
-For instance, assume
+data in the first format to the second.
+For instance, assume that
 the \codefont{comma\_name} format describes ``Doe, John''
-and the \codefont{space\_name} format describes ``John Doe''
+and that the \codefont{space\_name} format describes ``John Doe''
 and \codefont{s\_to\_c} is a function from the first to the second.  In this case,
@@
 
-One way to define the function used by squash is simply to write it
-from scratch in OCaml.  However, our system will also synthesize
-such functions automatically from pairs of regular languages and
-examples.
+One way to define the \codefont{s\_to\_c} canonization function used by squash
+is simply to write it from scratch. (Boomerang canonizers are written in OCaml.)
+However, our system will also synthesize such functions automatically from pairs
+of regular languages and examples---here, \codefont{s\_to\_c} is the
+\codefont{get} direction of a lens that can be synthesized by:
 %
@@
 extracts the get direction transformation from the lens,
-which is what we need for squash.
-
-Another QRE primitive is the composition combinator ``;''. This combinator
-applies two equivalence relations to a regular language in sequence. For
+which is what we need for \kw{squash}.
+
+Another QRE primitive is the sequential composition combinator ``;''.  For
 example, suppose that we wish to express a name as $\kw{last\_name \cdot
@@
 identification has been made. Then we can express this compound equivalence
-relation by a composition of a \textit{squash} QRE and a QRE that projects away
-the extra whitespace in between \kw{first\_name} and \kw{last\_name}
-
-A third useful QRE primitive is the permutation combinator.  For example,
-the following instance of \kw{perm} allows label, author and title fields
-(which we assume have been defined earlier) to appear in any order.
+relation by a composition of a \kw{squash} QRE and a QRE that projects away
+the extra whitespace in between \codefont{first\_name} and \codefont{last\_name}
+
+\begin{figure}[t]
+
+\begin{lstlisting}
+let wsp = [ \n\t\r]
+
+(* define name representations with a space and with a comma *)
+let space_name = first_name . wsp+ . last_name
+let comma_name = last_name . "," . wsp+ . first_name
+
+(* synthesize a lens that maps comma representation to space representation *)
+let l = synth comma_name $\Leftrightarrow$ space_name using {("Doe, John", "John Doe")}
+let s_to_c = get l
+
+(* squash QRE maps comma_name to space_name *)
+let comma_to_space = squash comma_name $\to$ space_name using (get comma_to_space)
+
+(* map arbitrary whitespace into " " as a canonical representative *)
+let p_wsp = project wsp+ $\to$ " "
+
+(* Project away whitespace in space_name representation *)
+let project_in_space = first_name . p_wsp . last_name
+
+(* Compose the two QREs to get a canonical representation for a name *)
+let name = project_in_space ; comma_to_space
+
+(* define rest of bibtex fields *)
+let name_sep = p_wsp . "and" . p_wsp
+let bib_names = name | (name . name_sep . (name . name_sep)$^*$ . name)
+let bib_author = "author = \"" . bib_names . "\""
+let title = word | (word . p_wsp . (word . p_wsp)$^*$ . word)
+let bib_title = "title = {" . title . "}"
+
+(* allow any permutation of fields interspersed with arbitrary whitespace *)
+let bib_fields = perm (label, bib_author, bib_title) with (project (wsp+ . "," . wsp+) $\to$ ",\n")
+let bibtex = "@book{" . bib_fields . "}"
+\end{lstlisting}
+
+\caption{QRE specification of \bibtex{} records. }
+\label{fig:example-qre}
+\end{figure}
+
+The permutation QRE combinator, \kw{perm}, allows data to be unordered. For
+example, the following instance of \kw{perm} allows the label, author and title
+fields (which we assume have been defined earlier) to appear in any order among
+the fields of a bibliography entry.
 %
@@
 with any other QRE.  For instance, below, we normalize whitespace between fields,
-leaving only a single newline.
+leaving only comma followed by a single newline.
 %
 \begin{lstlisting}
-let bib_fields = perm (label, bib_author, bib_title) with (project (wsp+ . "," . wsp+) $\to$ ",\n")
+let bib_fields = perm (label, bib_author, bib_title) 
+                 with (project (wsp+ . "," . wsp+) $\to$ ",\n")
 \end{lstlisting}
 %
-The final QRE combinator is the normalize combinator.
-This combinator allows a programmer to manually define a function
-$f$ which sends each string that matches a regular expression $R$ to some
-canonical representative in another regular expression $R'$. The equivalence
-relation defined by the normalize combinator is hence the equivalence relation
-defined by the {\em fibres} of $f$; that is, for all strings $s$ and $s'$ that
-match $R$, $s$ is equivalent to $s'$ if and only if $f(s) = f(s')$.
+The final QRE combinator is the \kw{normalize} combinator.  This combinator is an
+``escape hatch'' that allows a programmer to manually define a function $f$
+that sends each string matching a regular expression $R$ to some canonical
+representative in another regular expression $R'$. The equivalence relation
+defined by the normalize combinator is hence the equivalence relation defined by
+the {\em fibres} of $f$; that is, for all strings $s$ and $s'$ that match $R$,
+$s$ is equivalent to $s'$ if and only if $f(s) = f(s')$.
 
 For instance, assume that $f(s)=$ `` '' for all whitespace strings $s$. Then the
-project QRE \kw{p_wsp} which matches any non-zero number of whitespace
-characters and converts any such whitespace into a single space character can
-be expressed using the normalize combinator as
+project QRE \kw{p\_wsp} which matches any non-zero number of whitespace
+characters and converts any such whitespace into a single space character could
+be expressed using the \kw{normalize} combinator as
 \begin{lstlisting}
@@
 \end{lstlisting}
-The QRE for the bibtex format is presented in its entirety in
-Figure~\ref{fig:example-qre}.
-\begin{figure}[t]
-\begin{lstlisting}
-let wsp = [ \n\t\r]
-
-(* define name representations with a space and with a comma *)
-let space_name = first_name . wsp+ . last_name
-let comma_name = last_name . "," . wsp+ . first_name
-
-(* synthesize a lens that maps comma representation to space representation *)
-let l = synth comma_name $\Leftrightarrow$ space_name using {("Doe, John", "John Doe")}
-let s_to_c = get l
-
-(* squash QRE maps comma_name to space_name *)
-let comma_to_space = squash comma_name $\to$ space_name using (get comma_to_space)
-
-(* map arbitrary whitespace into " " as a canonical representative *)
-let p_wsp = project wsp+ $\to$ " "
-
-(* Project away whitespace in space_name representation *)
-let project_in_space = first_name . p_wsp . last_name
-
-(* Compose the two QREs to get a canonical representation for a name *)
-let name = project_in_space ; comma_to_space
-
-(* define rest of bibtex fields *)
-let name_sep = p_wsp . "and" . p_wsp
-let bib_names = name | (name . name_sep . (name . name_sep)$^*$ . name)
-let bib_author = "author = \"" . bib_names . "\""
-let title = word | (word . p_wsp . (word . p_wsp)$^*$ . word)
-let bib_title = "title = {" . title . "}"
-
-(* allow any permutation of fields interspersed with arbitrary whitespace *)
-let bib_fields = perm (label, bib_author, bib_title) with (project (wsp+ . "," . wsp+) $\to$ ",\n")
-let bibtex = "@book{" . bib_fields . "}"
-\end{lstlisting}
-\caption{QRE specification of \bibtex{} records. }
-\label{fig:example-qre}
-\end{figure}
+However, although it is in some sense ``universal'' (see the discussion below
+\saz{TODO}))the \kw{normalize} QRE imposes significant requirements on the
+canonization function $f$, so it is usually simpler to use the other QRE
+constructs where possible.
+
+A QRE for the \bibtex{} format is presented in its
+entirety in Figure~\ref{fig:example-qre}.
+
 \subsection{QRE Lenses}
@@
 \begin{itemize}
-  \item \kw{del} $s$: delete the constant string $s$ in the get direction;
-  produce the constant string $s$ in the put direction.
-  \item \kw{ins} $s$: produce the string $s$ in the get direction;
-  delete the string $s$ in the put direction.
+  \item \kw{del} $s$: \textit{delete} the constant string $s$ in the get direction;
+  insert it in the put direction.
+  \item \kw{ins} $s$: \textit{insert} the constant string $s$ in the get direction;
+  delete it in the put direction.
   \item \kw{copy} $R$: copy the text matching the regular expression $R$ in both
@@
 combinators to transform \bibtex{} to EndNote (and vice versa).
-A fair amount of trickiness in writing these transformations as users
-must keep mental track of what they have parsed so far \emph{in both
+There is a fair amount of trickiness in writing these transformations, because
+the programmer must keep mental track of what they have parsed so far \emph{in both
 directions}, carefully inserting and deleting the relevant syntax on both
@@
 %% error-prone.
-Consequently, we offer the programmer some relief by identifying a class of
-quotient lenses that we call {\em QRE lenses}, and by showing how to {\em synthesize}
-QRE lenses from a pair of QREs and a set of input-output example pairs. Our key
+To make this unsavory programming task more palatable, we identify a class of
+quotient lenses that we call {\em QRE lenses} and show how to {\em synthesize}
+them from a pair of QREs and a set of input-output examples. Our key
 insight is to reduce the problem of synthesizing QRE lenses to the problem of
@@
 \begin{lstlisting}
-let bib_to_end : (lens in bibtex $\Leftrightarrow$ endnote)
-= synth bibtex $\Leftrightarrow$ endnote  using  {(bib_example, end_example)}
+let bib_to_end : (lens in bibtex $\Leftrightarrow$ endnote) =
+    synth bibtex $\Leftrightarrow$ endnote  using  {(bib_example, end_example)}
 \end{lstlisting}
@@
 concrete example strings givn at the beginning of this section) to disambiguate.
-In addition. as we saw earlier with the definition of \cd{s_to_c}, the synthesis
+In addition, as we saw earlier with the definition of \cd{s_to_c}, the synthesis
 procedure itself can be used to create lenses that are in turn used to define
@@
 \begin{align*}
-c := \; &R \sep \project{R}{s} \sep \squash{c}{c'}{f} \sep
+c := \; &R \sep \project{R}{s} \sep \squash{c_1}{c_2}{f} \sep
 \perm{c_1, \ldots, c_n}{c} \\
-& | \; \normalize{R}{R'}{f} \sep c \semicolon c' \sep c \cdot c''' \sep (c \sep
-c') \sep c^*,
+& | \; \normalize{R_1}{R_2}{f} \sep c_1 \semicolon c_2 \sep c_1 \cdot c_2 \sep (c_1 \sep
+c_2) \sep c^*,
 \end{align*}
@@
 \project{R}{s} & R & s \\
-\squash{c}{c'}{f} & W(c) \sep W(c') & K(c') \\
-\normalize{R}{R'}{f} & R & R' \\
-c_2 \; ; \; c_1 & W(c_1) & K(c_2) \\
+\squash{c_1}{c_2}{f} & W(c_1) \sep W(c_2) & K(c_2) \\
+\normalize{R_1}{R_2}{f} & R_1 & R_2 \\
+c_1 \; ; \; c_2 & W(c_1) & K(c_2) \\
 c_1 \cdot c_2 & W(c_1) \cdot W(c_2) & K(c_1) \cdot K(c_2) \\
@@
 \]
-\caption{Whole and Kernel Regular Expressions}
+\caption{Whole and Kernel Regular Expressions \saz{Mentions somewhere that $S_n$
+    is the set of all permutations.} \saz{Check that the semi-colon
+    case is still correct}}
 \label{fig:wk}
 \end{figure}
-The \textit{squash} and \textit{permutation} combinators give the two
-interesting definitions. If $c = \squash{c}{c'}{f}$, then the whole language of
-$c$ is $W(c) \sep W(c')$. This is because the \textit{squash} combinator takes
-the whole language $W(c)$ of $c$, the whole language $W(c')$ of $c'$ and a
-function $f : \mathcal{L}(W(c)) \longrightarrow \mathcal{L}(W(c'))$, maps
-$\mathcal{L}(W(c))$ into $\mathcal{L}(W(c'))$ using $f$ and then canonizes
-$W(c')$ into $K(c')$ using $\canonizer(c')$.
+The \kw{squash} and \kw{permutation} combinators give the two interesting
+definitions. If $c = \squash{c_1}{c_2}{f}$, then the whole language of $c$ is
+$W(c_1) \sep W(c_2)$. This is because the \kw{squash} combinator uses the
+function $f : \mathcal{L}(W(c_1)) \longrightarrow \mathcal{L}(W(c_2))$ to maps
+any strings from $\mathcal{L}(W(c_1))$ into $\mathcal{L}(W(c_2))$ (leaving the
+$c_2$ strings alone), and then canonizes $W(c_2)$ into $K(c_2)$ using
+$\canonizer(c_2)$.
 
@@
 \end{align*}
-\caption{QRE canonizers}
+\caption{QRE canonizers \saz{TODO: use $c_1$ and $c_2$ everywhere.}}
 \label{fig:canonizers}
@@
 \]
-\caption{QRE Equivalence Relations}
+\caption{QRE Equivalence Relations \saz{TODO: use $c_1$ and $c_2$ everywhere.}}
 \label{fig:relations}

### 2018-03-05 papers/QuotientLenses/main.tex
Polishing and comments on early sections

--- 2018-03-04 papers/QuotientLenses/main.tex
@@
 \acmVolume{1}
-\acmNumber{ICDP} % CONF = POPL or ICFP or OOPSLA
+\acmNumber{ICFP} % CONF = POPL or ICFP or OOPSLA
 \acmArticle{1}
@@
 them do not have built-in support for transformations that operate on data
-modulo equivalence relations such as whitespace or permutation quotienting.
+modulo equivalence relations such as whitespace or permutation
+quotienting.\bcp{I'm worried that most readers will not be able to follow
+  that.  Moreover, this way of introducing the paper feels kind of
+  negative---most languages are insufficient, and Boomerang is hard to
+  program.  Could we say it in a more accessible, more positive way?}
 Moreover, langauges such as Boomerang which do allow for such transformations
@@
 Boomerang, and demonstrate the utility of our implementation by using it to
-synthesize QRE lenses between data formats drawn from {\tt data.gov} database.
+synthesize QRE lenses between a variety of real-world data formats used by
+the {\tt data.gov} database. 
 \end{abstract}
@@
 JSON or
-XML object. When different applications manipulate the same data in different formats,
-synchronizing the applications amounts to synchronizing their respective data
+XML object.\bcp{These examples show that the same data {\em could} be
+  represented in different formats, not that it needs to be.} When different applications manipulate the same data in different formats,
+synchronizing the applications\bcp{Why are we talking about synchronizing
+  applications?  Comes out of the blue.} amounts to synchronizing their
+respective data
 formats. One way to synchronize data formats is to use \emph{lenses}, \emph{i.e.,} programs that represent pairs of functions for translating data from
@@
 languages like Boomerang~\cite{boomerang} or
-BiGul~\cite{ko2016bigul}, typically come with a variety of guarantees,
-the \emph{lens laws}, that help ensure put and get functions are effective
-inverses of one another.
-%comprise such synchronization functions, and the ``lens
-%laws'' provide restrictions on how these synchronization functions interoperate%.
-%Bidirectional languages like Boomerang~\cite{boomerang} and
-%BiGul~\cite{DBLP:conf/pepm/KoZH16} use a single expression to encode all of the
-%lens's synchronization functions at the same time.
-
-One way to classify lenses is by the laws they obey.  Amongst the most
+BiGul~\cite{ko2016bigul}\bcp{why mention exactly these two, out of all the
+  bidirectional languages out there?}, typically come with a variety of
+guarantees\bcp{grammatically, that says that each lens comes with a variety
+  of guarantees---is that what you meant?},
+the \emph{lens laws}, that help\bcp{why ``help''?} ensure put and get
+functions are effective 
+inverses\bcp{what is an effective inverse?} of one another.
+
+One way\bcp{Do you mean to highlight the suggestion that there are other
+  ways too?} to classify lenses is by the laws they obey.  Amongst the most
 restrictive lenses are the \emph{bijective lenses}.  These lenses demand
@@
 {\em Quotient lenses}~\cite{quotientlenses} loosen many of the restrictions
-of bijective lenses by introducing an equivalence relation and considering
+of bijective lenses by introducing an equivalence relation\bcp{actually,
+  two: one on each side; but more importantly ``introducing an equivalence
+  relation'' is rather vague.} and considering
 bijections modulo that relation.  Such quotient lenses make it possible
@@
 ignoring details that a programmer views as inessential.
-%each of the data formats is quotiented by some equivalence relation and the lens
-%laws are taken modulo equivalence.
-%This allows lenses to be simpler and more
-%flexible. For example, inessential details like whitespace and orderings of
-%fields can be ignored when writing the transformations between the formats.
-
-%However, while plain lenses are widely used in both standalone systems and
-%popular libraries, Boomerang~\cite{boomerang} is the only implementation
-%currently offering quotient lenses. Even though quotient lenses permit simpler
-%transformations between the formats, they merely push complexity elsewhere.
-Ordinary bijective lenses are fairly tricky to program, especially for
-beginners, as they require programmers to think in two directions at once
-and learn a new type system, which comes with some tricky ambiguity constraints.
-Unfortunately, quotient lenses do not improve the situation:  To
+
+Ordinary bijective lenses are fairly tricky to program\bcp{do we have
+  concrete evidence for this?}, especially for
+beginners\bcp{or this?}, as they require programmers to think in two
+directions at once\bcp{my own experience suggests that this is the less
+  important issue...}
+and learn a new type system, which comes with some tricky ambiguity
+constraints\bcp{...and this is the bigger one.  But we don't explain what
+  these ambiguity constraints are or why they must be imposed.}.
+Unfortunately, quotient lenses do not improve the situation\bcp{Again, it
+  feels rather negative.}:  To
 write a quotient lens in Boomerang, programmers must work with standard lens
@@
 in two distinct, but connected ways. First, we
-introduce a new language of {\em Quotient Regular Expressions}, or QREs, which
+introduce a new language of {\em Quotient Regular Expressions} (QREs), which
 enable a programmer to simultaneously specify a regular expression as well as
@@
 order to generate a lens, a programmer provides a pair of QREs
-describing the source and target formats, and, optionally,
-a suite of examples that demonstrates how the
-lens should act on specific examples.  Given this information,
-we are able to exploit and extend recent work on algorithms for synthesis of
+describing the source and target formats, plus 
+a (possibly empty) suite of examples that demonstrates how the
+lens should act in specific cases.  Given this information,
+we can exploit and extend recent work on algorithms for synthesis of
 bijective lenses~\cite{optician} to generate quotient lenses.
-Importantly, the lenses generated by
+The lenses generated by
 our synthesis algorithm are guaranteed to translate data back and forth
@@
 Moreover, while our synthesized quotient lenses have a relatively
-restricted form (a fact that cuts down the amount of non-determinism in
-a specification and thus helps speed up synthesis), we prove that
-this rigidity does not compromise expressiveness.
-
-The following list summarizes the central contributions of our work.
+restricted form (which cuts down the amount of non-determinism in
+a specification and greatly speeds up synthesis), we prove that
+this rigidity does not compromise their expressiveness.
+
+The following list summarizes the central contributions of our
+work.\bcp{Ugh.} 
 \begin{enumerate}
   \item We introduce {\em Quotient Regular Expressions} (QREs),
-  a compact, convenient notation for defining (1) a
-  regular language modulo some equivalence relation, and (2) a canonizer
+  a compact, convenient notation for simultaneously defining a
+  regular language modulo some equivalence relation and a canonizer
   for that relation (Section~\ref{QRE}).
@@
   quotient lenses.
-  % In other words, if there eeach QRE lens $q$ defines the same transformation as a
-  %QRE lens $q'$ that is in normal form.
   (Section~\ref{QRE-lenses}).
@@
 \end{enumerate}
-Sections~\ref{relwork} discusses related work and~\ref{concl} summarizes the
-results of the paper.
-
-\section{Quotients by Example}
+Sections~\ref{relwork} and~\ref{concl} discuss related and future work.
+
+\bcp{We've said that one of the main contributions of the paper is
+  synthesizing lenses, but we don't actually show this happening in the
+  introduction.  Should we?}
+
+\section{QRE Lenses by Example}
 \label{sec:example}
@@
 
-\emph{Lenses} define computations that can be read in two directions,
-the get direction (forwards) and the put direction (backwards).  The simplest lenses define bijections between source and
-data, but we are interested in \emph{quotient lenses}, which define
-bijections modulo some equivalence relation.  The remains of this section
-describe and motivate a new language and system for defining such lenses.
+\emph{Lenses} define computations that can be read in two directions, the
+get direction (forwards) and the put direction (backwards).  The simplest
+lenses define bijections between source and data, but we are interested in
+\emph{quotient lenses}, which define bijections modulo some equivalence
+relation\bcp{two!}.  The core ideas of quotient lenses were developed by
+Foster et al~\cite{quotientlenses} and are embodied in the Boomerang
+language.  This section describes and motivates a 
+new notation for defining such lenses.
+
+\bcp{We need two subsections here --- one on plain old bijective lenses, and
+one on bijective lens synthesis.}
+
+\subsection{Quotient Lenses}
 
@@
 When two instances of the same format differ only terms of such inessential
-information, it is often useful to treat them as identical and to map them
+information, it is useful\bcp{Would it be worth spelling out the alternative
+(i.e., what lenses would look like if they had to preserve all this stuff?)}
+to treat them as identical and to map them 
 into canonical representations.
@@
 adds a little more info, but not much.  Should this be cut?  Should the intro be expanded
-slightly?  Should it be left as is (more or less)?}
+slightly?  Should it be left as is (more or less)?} \saz{I think that this is
+not too redundant and that it can be left as-is, modulo space crunching.}
 
 \begin{enumerate}
-  \item A pair of descriptions of the external source and target formats.
+  \item A pair of descriptions of the external source and target formats, here
+    \bibtex{} and EndNote.
   Such descriptions may be given by regular expressions, for instance.
   They serve as documentation to describe the domains over which the lens
-  is to operate.
+  is to operate.\bcp{This way of saying it implies that they serve {\em
+      only} as documentation.}
   \item A pair of \emph{canonizers} that map external source (and target)
@@
   \item A pair of descriptions of the internal source and
-  target formats, again, to document the structure of the transformation.
+  target formats, again, to document the structure of the
+  transformation.\bcp{Same question.}
   \item A bijective lens to map back and forth between canonical elements
@@
 view it as good engineering practice to do so, as such documentation makes
-lenses easier to read, understand, use and maintain.  In this case, there is
+lenses easier to read, understand, use and maintain.\bcp{Is it really true
+  that these types are optional in Boomerang?}  In this case, there is 
 significant redundancy here:  External and internal formats
@@
 structure; and the structure of bijective lens computations is largely driven
-by the structure of the internal formats.  Hence, overall, a more compact
+by the structure of the internal formats.\bcp{Would it be worth
+  demonstrating this concretely?}  Hence, overall, a more compact
 notation would simplify description and legibility of such computations.
@@
 synthesizing the lenses that map between these formats automatically.
+\bcp{If our main contribution is ``eliminating fiddliness,'' we should be
+  more explicit about what the fiddliness is.  (But, again, I would be even
+  happier if we could phrase everything more positively... ``Quotient lenses
+  make lens programming more realistic.  But here are some opportunities for
+  eliminating redundancy...'')}
 
@@
 A {\em Quotient Regular Expression} (QRE) defines an external regular
-expression $R$, an equivalence relation $\sim$ on the language
-described by $R$, and a canonical set of representatives of $\sim$,
-all at once.  QREs are constructed from a simple set of primitives
-and composed using familiar regular operators
-(choice `\cd{|}', concatenation `\cd{.}',
-and Kleene star `$^*$').  Users may also treat QREs as pairs of
-get and put functions and compose such pairs (the functional
-composition operator `;').
+expression $R$, an equivalence relation $\sim$ on the language described by
+$R$, and a canonical set of representatives of $\sim$, all at once.  QREs
+are constructed from a simple set of primitives and composed using familiar
+regular operators (`\cd{|}' for choice, `\cd{.}' for concatenation, and
+`$^*$' for repetition).  Users\bcp{This sentence feels clunky because it
+  ``uses up'' its head/subject position, and needlessly focuses the reader's
+  attention, on something (``users'') that is not especially important or
+  germane to the conversation.  (There are lots of other sentences elsewhere
+  in the paper that do the same thing. :-) Please rewrite (all of them).}
+may also treat QREs as pairs of get and put functions and compose such pairs
+(the functional composition operator `;').\saz{I agree with Benjamin about
+  useless subjects, but here the sentence could be deleted entirely.}
 
@@
 any such whitespace into a single space character (its canonical form).
-Such a definition uses the \kw{project} primitive.
+This whitespace-normalizing QRE is easily defined by using the \kw{project} primitive: \bcp{Why is it called ``project''?}
 
@@
 \end{lstlisting}
+\bcp{Can we display blanks as a distinct ``visible blank'' symbol
+  (throughout)?  That will make examples easier to read.}
 
@@
 allows external data to be in either format, and converts any
-data in one format to the other.
-For instance, assume
+data in the first format to the second. 
+For instance, assume that
 the \codefont{comma\_name} format describes ``Conway, J. H.''
-and the \codefont{space\_name} format describes ``J. H. Conway''
+and that the \codefont{space\_name} format describes ``J. H. Conway''
 and \codefont{s\_to\_c} is a function from the first to the second.  In this case,
@@
 
-One way to define the function used by squash is simply to write it
-from scratch in OCaml.  However, our system will also synthesize
-such functions automatically from pairs of regular languages and
-examples.
+One way to define the \codefont{s\_to\_c} canonization function used by squash
+is simply to write it from scratch in some ordinary programming language. (Boomerang canonizers are written in OCaml.)
+However, our system will also synthesize such functions automatically from pairs
+of regular languages and examples---here, \codefont{s\_to\_c} is the
+\codefont{get} direction of a lens that can be synthesized by:
 %
@@
 %
-The first line above synthesizes a \emph{bidirectional} lens between
+\bcp{Four problems here.  The first is that the notion of synthesizing lenses
+comes out of nowhere (of course, it is in the background, but we're using it
+here as if it had already been introduced concretely---this will be helped
+by adding the new subsection I've suggested at the top of this section).
+The second is that the idea of using half of a lens 
+as a canonizer comes out of nowhere.  The third is that this idea is a hack (and
+always was).  Why bother defining / synthesizing a lens and then using only
+half of it?  And the fourth is that, in the original quotient lens paper,
+we used this trick on lenses that were not even bijective, which may confuse
+people familiar with the earlier paper.}The first line above synthesizes a
+\emph{bidirectional}\bcp{Why is this italicized?  Are some lenses not
+  bidirectional?} lens between 
 \codefont{comma\_name} and \codefont{space\_name} using the listed example
@@
 extracts the get direction transformation from the lens,
-which is what we need for squash.
-
-A third useful QRE primitive is the permutation combinator.  For example,
+which is what we need for \kw{squash}.
+
+Another QRE primitive is the sequential composition combinator ``;''.  For
+example, suppose that we wish to express a name as $\kw{last\_name \cdot
+``,\text{''} \cdot \text{wsp+} \cdot first\_name}$, or as $\kw{first\_name\cdot
+\text{wsp+} \cdot last\_name}$, with $\kw{first\_name\cdot \text{wsp+} \cdot
+last\_name}$ as the canonical representation. We also wish to project away the
+extra whitespace in between \kw{first\_name} and \kw{last\_name} once this
+identification has been made. Then we can express this compound equivalence
+relation by a composition of a \kw{squash} QRE and a QRE that projects away
+the extra whitespace in between \codefont{first\_name} and \codefont{last\_name}
+
+\begin{figure}[t]
+\begin{lstlisting}
+(* whitespace with " " as a canonical representative *)
+let wsp = project [ \n\t\r]$^+$ $\to$ " "
+
+(* define name representations with a space and with a comma *)
+let space_name = name | (name . wsp . (name . wsp)$^*$ . name)
+let comma_name = name | (name . "," . wsp . (name . wsp)$^*$ . name)
+
+(* synthesize a lens that maps the comma representation to the space representation *)
+let l = synth comma_name $\Leftrightarrow$ space_name using {("Conway, J. H.", "J. H. Conway")}
+let s_to_c = get l
+
+(* squash QRE maps comma_name to space_name *)
+let name = squash comma_name $\to$ space_name using s_to_c
+
+(* define rest of bibtex fields *)
+let name_sep = wsp . "and" . wsp
+let bib_names = name | (name . name_sep . (name . name_sep)$^*$ . name)
+let bib_author = "author = \"" . bib_names . "\""
+let title = word | (word . wsp . (word . wsp)$^*$ . word)
+let bib_title = "title = {" . title . "}"
+
+(* allow any permutation of fields interspersed with arbitrary whitespace *)
+let bib_fields = perm (label, bib_author, bib_title) with (project ("," .wsp$^+$) $\to$ ",\n")
+let bibtex = "@book{" . bib_fields . "}"
+\end{lstlisting}
+
+\caption{QRE specification of \bibtex{} records. \bcp{The spacing of the
+    comments is weird...}}
+\label{fig:example-qre}
+\end{figure}
+
+The permutation QRE combinator, \kw{perm}, allows data to be unordered. For example,
 the following instance of \kw{perm} allows label, author and title fields
@@
 %
-To normalize any repeated separator between fields, one can extend a permutation
-with any other QRE.  For instance, below, we normalize whitespace between fields,
-leaving only a single newline.
+To normalize the field separators, one can specify that the
+components of the permutation are conjoined by another QRE. For instance, below, we normalize whitespace
+between fields, leaving only a single newline.
 %
@@
 \end{lstlisting}
-%
-\noindent
-The QRE for the bibtex format is presented in its entirety in Figure~\ref{fig:example-qre}.
-
-\begin{figure}[t]
-\begin{lstlisting}
-(* whitespace with " " as a canonical representative *)
-let wsp = project [ \n\t\r]$^+$ $\to$ " "
-
-(* define name representations with a space and with a comma *)
-let space_name = name | (name . wsp . (name . wsp)$^*$ . name)
-let comma_name = name | (name . "," . wsp . (name . wsp)$^*$ . name)
-
-(* synthesize a lens that maps the comma representation to the space representation *)
-let l = synth comma_name $\Leftrightarrow$ space_name using {("Conway, J. H.", "J. H. Conway")}
-let s_to_c = get l
-
-(* squash QRE maps comma_name to space_name *)
-let name = squash comma_name $\to$ space_name using s_to_c
-
-(* define rest of bibtex fields *)
-let name_sep = wsp . "and" . wsp
-let bib_names = name | (name . name_sep . (name . name_sep)$^*$ . name)
-let bib_author = "author = \"" . bib_names . "\""
-let title = word | (word . wsp . (word . wsp)$^*$ . word)
-let bib_title = "title = {" . title . "}"
-
-(* allow any permutation of fields interspersed with arbitrary whitespace *)
-let bib_fields = perm (label, bib_author, bib_title) with (project ("," .wsp$^+$) $\to$ ",\n")
-let bibtex = "@book{" . bib_fields . "}"
-\end{lstlisting}
-\caption{QRE specification of \bibtex{} records. }
-\label{fig:example-qre}
-\end{figure}
-
-%% In Figure~\ref{fig:example-qre}, the QRE \verb+project [ \n\t\r]+$^{+} \to$
-%% \verb+" "+ matches any non-zero number of whitespace characters and quotients
-%% them to a single equivalence class with a single space as the canonical
-%% representative. We call this QRE a \textit{projection} because an entire regular
-%% language is projected to a single representative string.
-
-%% Another QRE in Figure~\ref{fig:example-qre} is given by
-
-%% \begin{center}
-%% \begin{lstlisting}
-%% squash comma_name $\to$ space_name using (get space_to_comma)
-%% \end{lstlisting}
-%% \end{center}
-
-%% \noindent This QRE is a \textit{squash} QRE. A \textit{squash} QRE ``glues''
-%% two data formats using a user-defined function. For example,
-%% the QRE in the code above specifies that the \cd{comma_name} representation of
-%% the author's name is equivalent to the \cd{space_name} representation, with
-%% \cd{space_name} taken to be the canonical representation. The
-%% function that identifies the two formats is given by \cd{get space_to_comma},
-%% which transforms names in the \cd{comma_name} representation to names in the
-%% \cd{space_name} representation. In this example, \cd{space_to_comma} is itself
-%% an automatically \textit{synthesized} lens between the \cd{space_name} and
-%% \cd{comma_name} QREs, and we need to use the ``get'' direction of the lens to
-%% specify the squash transformation.
-
-%% The third example of a QRE in the Figure is \cd{bib_fields}, defined by:
-%% \begin{center}
-%% \begin{lstlisting}
-%% perm (label, bib_author, bib_title) with (project ("," . wsp$^+$ ) $\to$ ",\n")
-%% \end{lstlisting}
-%% \end{center}
-
-%% \noindent This is a \textit{permutation} QRE that specifies that the label,
-%% title, and author fields can occur in any order, with the fields separated by a
-%% comma followed by arbitrary amounts of whitespace. The canonical representation
-%% is that in which the label, author and title fields occur in that order,
-%% separated by a comma followed by a newline. The \textit{permutation} QRE is
-%% designed for equivalences such as this, where different permutations of
-%% equivalences are interspersed with a separator which itself has an
-%% equivalence defined on it.
-
-%% Given the \textit{projection}, \textit{squash}, and \textit{permutation}
-%% QRE primitives, we can build more complex languages using the usual regular
-%% expression combinators of alternative choice `\cd{|}', concatenation `\cd{.}'
-%% and Kleene star `$^*$', and also sequential composition ``;''.
+
+% The final QRE combinator is the \kw{normalize} combinator.  This combinator is an
+% ``escape hatch'' that allows a programmer to manually define a function $f$
+% that sends each string matching a regular expression $R$ to some canonical
+% representative in another regular expression $R'$. The equivalence relation
+% defined by the normalize combinator is hence the equivalence relation defined by
+% the {\em fibres} of $f$; that is, for all strings $s$ and $s'$ that match $R$,
+% $s$ is equivalent to $s'$ if and only if $f(s) = f(s')$.
+
+% For instance, assume that $f(s)=$ `` '' for all whitespace strings $s$. Then the
+% project QRE \kw{p\_wsp} which matches any non-zero number of whitespace
+% characters and converts any such whitespace into a single space character could
+% be expressed using the \kw{normalize} combinator as
+% \begin{lstlisting}
+% let p_wsp = normalize (wsp+, " ", f)
+% \end{lstlisting}
+% However, although it is in some sense ``universal'' (see the discussion below
+% \saz{TODO}))the \kw{normalize} QRE imposes significant requirements on the
+% canonization function $f$, so it is usually simpler to use the other QRE
+% constructs where possible.
+
+A QRE for the \bibtex{} format is presented in its
+entirety in Figure~\ref{fig:example-qre}. \bcp{Walk through it a bit?}\saz{All of the
+  examples in this section are walking through the figure, we just have to say so.}
 
@@
 a first step in writing many data transformations.  However, writing the
-inner bijective lens itself still takes a fair amount of work.  Each such lens
+inner bijective lens itself still takes a fair amount of work\bcp{Why do we
+  want to emphasize this here?}.  Each such lens
 is constructed from simple primitive combinators including:
@@
 \caption{An example lens between \cd{bibtex} and \cd{end_note} regular
-expressions with no equivalences.\bcp{Indentation needs work.  And there was
+expressions with no equivalences.\bcp{There was
 a suggestion to colorize the new keywords that we're introducing here
-(including symbolic ones, I guess)}}
+({\em including} symbolic ones, I guess)}\bcp{And a readability suggestion:
+define a named constant for a quote character.}}
 \label{fig:example-lens}
@@
 
-Consequently, we offer the programmer some relief by identifying a class of
-quotient lenses that we call {\em QRE lenses}, and by showing how to {\em synthesize}
-QRE lenses from a pair of QREs and a set of input-output example pairs. Our key
+\bcp{How about a subsection break here, to introduce the idea of QRE
+  lens synthesis?}
+
+To make this unsavory programming task more palatable, we identify a class of
+quotient lenses that we call {\em QRE lenses} and show how to {\em synthesize}
+them from a pair of QREs and a set of input-output examples.\bcp{that's a pretty weak
+  way of describing our main contribution!} Our key
 insight is to reduce the problem of synthesizing QRE lenses to the problem of
@@
 raises a key technical question:  Are we limiting the expressiveness of our transformations
-by demanding all programs fit into this normal form?  It turns out that we are not:
+by demanding all programs fit into this normal form?\bcp{I wish we could
+  have previewed this issue a bit more in the introduction, since it is our
+  main contribution.}  It turns out that we are not:
 any lens that uses canonizers internally can be transformed into a lens that uses canonizers
 at the edges.  The main technical contribution of this paper is a proof of this fact.  The
-crucial step in the proof is a demonstration that functional composition of two lenses
-with ``canonizers at the edge'' can be represented as some single lens with ``canonizers at the
-edge.''
+crucial step in the proof is a demonstration that the functional composition
+of two QRE lenses with ``all canonizers at the edge'' can be represented as some
+single QRE lens with ``canonizers at the edge.''  \bcp{Saying ``all
+  canonizers at the edge'' seems confusing: there is no other place for them
+to go, synmtactically.  And: Could we draw a picture of this transformation?}
 
@@
 This technical investigation justifies replacing manually constructed lenses that may use
-functional composition and nested canonizers with synthesized lenses in normal form.
-The end result is that users may synthesize complex transformations with a single line
-of code.  For instance, all the code in Figure~\ref{fig:example-lens} may be replaced by
+functional composition and nested canonizers\bcp{nested canonizers have not
+  been mentioned!} with synthesized lenses in normal form.
+The end result is that users can synthesize complex transformations with a
+single line of code.\bcp{Well, the QRE definitions and examples may well use
+multiple lines all by themselves!}  For instance, all the code in Figure~\ref{fig:example-lens} may be replaced by
 a single call to the synthesis prodedure:
@@
 \begin{lstlisting}
-let bib_to_end : (lens in bibtex $\Leftrightarrow$ endnote)
-= synth bibtex $\Leftrightarrow$ endnote  using  {(bib_example, end_example)}
+let bib_to_end : (lens in bibtex $\Leftrightarrow$ endnote) =
+  synth bibtex $\Leftrightarrow$ endnote
+  using {(bib_example, end_example)}
 \end{lstlisting}
+\bcp{Again, the spacing is weird}
 \noindent Here, the generated  quotient lens synchronizes \cd{bibtex} and
@@
 strings given at the beginning of this section) to disambiguate.
-In additiona, and as we saw earlier with the definition of \cd{s_to_c}, the synthesis procedure itself
+In addition, and as we saw earlier with the definition of \cd{s_to_c}, the synthesis procedure itself
 can be used to create lenses that are in turn used to define other QREs.  The ability to intermix
@@
 
-Quotient Regular Expressions allow programs to specify a regular language,
+Quotient Regular Expressions allow programs\bcp{Why bring ``programs'' into
+  it?} to specify a regular language,
 an equivalence relation over elements of the language and a function to convert
@@
 
-\subsection{Overview of Quotient Regular Expression}
+\subsection{Intuitions}
 In the \bibtex{} to EndNote example, we encountered the $\project{R}{s}$,
@@
 \noindent $\squash{c}{c'}{f}$ and $\perm{c_1, \ldots, c_n}{c}$ QREs, as well as
-the regular operators applied to these, as in
+the regular operators applied to these, as in\bcp{linebreak}
 
@@
 
-One combinator that we haven't yet introduced is sequential composition.
+One combinator that we haven't yet introduced\bcp{haven't we? we've
+  mentioned it, at least.} is sequential composition.
 The $c \semicolon c'$ QRE successively applies
@@
 made. Then we can express this equivalence relation by a composition of a
-\textit{squash} QRE and a QRE that projects away the extra whitespace in between
+\kw{squash} QRE and a QRE that projects away the extra whitespace in between
 \kw{first\_name} and \kw{last\_name}:
@@
 
-Figures~\ref{fig:wk} gives the inductive definitions for the whole and
-kernel languages $W(c)$ and $K(c)$ of a QRE. The \textit{squash} and
-\textit{permutation} combinators give the two interesting definitions. If $c =
-\squash{c}{c'}{f}$, then the whole language of $c$ is $W(c) \sep W(c')$. This
-is because the \textit{squash} combinator takes the whole language $W(c)$ of
-$c$, the whole language $W(c')$ of $c'$ and a function $f : \mathcal{L}(W(c))
-\longrightarrow \mathcal{L}(W(c'))$, maps $\mathcal{L}(W(c))$ into
-$\mathcal{L}(W(c'))$ using $f$ and then canonizes $W(c')$ into $K(c')$ using
-$\canonizer(c')$.
-
-For the $\perm{(c_1, \ldots, c_n)}{c}$ combinator, the whole language is the
-union of languages of the form $W(c_{\sigma(1)}) \cdot W(c) \cdot \ldots \cdot
-W(c) \cdot W(c_{\sigma(n)})$. This is because the \textit{permutation}
-combinator allows for the string to match any permutation the $c_i$'s while also
-taking into account the separator $c$ in between each of the $c_i$'s. The kernel
-of the \textit{permutation} combinator is the language $K(c_1) \cdot K(c) \cdot
-\ldots \cdot K(c) \cdot K(c_n)$, because the canonical permutation is the
-identity permuation, with each of the parts of the string that match $c_i$
-and $c$ canonized into $K(c_i)$ and $K(c)$ respectively.
-
-Figure~\ref{fig:canonizers} gives the inductive definitions of the
-$\canonizer{}$ function for each QRE. Again, the \textit{permutation} combinator
-gives the most interesting definition,
-\begin{align*}\canonizer(\perm{c_1, \ldots, c_n}{c})(r_{\sigma(1)}
-\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \\
-= (\canonizer(c_1)(r_1)) \cdot (\canonizer(c)(s_1)) \cdot \ldots \cdot
-(\canonizer(c_n)(r_n))
-\end{align*}
-\noindent which as we have explained is defined so that the strings that match
-$c_i, c$ are placed in the canonical permutation which is the identity
-permutation before $\canonizer(c_i)$ and $\canonizer(c)$ are applied
-respectively.
-
-Finally, Figure ~\ref{fig:relations} gives the induction defintion of the
-equivalence relation $\eqrel{c}$. These are the set-theoretic semantics of QREs
-as an equivalence relation on a regular language.
+Figure~\ref{fig:wk} gives the inductive definitions for the whole and
+kernel languages $W(c)$ and $K(c)$ of a QRE.
+The \kw{squash} and \kw{permutation} combinators give the two interesting
+definitions. If $c = \squash{c_1}{c_2}{f}$, then the whole language of $c$ is
+$W(c_1) \sep W(c_2)$. This is because the \kw{squash} combinator uses the
+function $f : \mathcal{L}(W(c_1)) \longrightarrow \mathcal{L}(W(c_2))$ to maps
+any strings from $\mathcal{L}(W(c_1))$ into $\mathcal{L}(W(c_2))$ (leaving the
+$c_2$ strings alone), and then canonizes $W(c_2)$ into $K(c_2)$ using
+$\canonizer(c_2)$.
 
@@
 \]
-\caption{Whole and Kernel Regular Expressions}
+\caption{Whole and Kernel Regular Expressions \saz{Mention somewhere that $S_n$
+    is the set of all permutations.} \saz{Check that the semi-colon
+    case is still correct}}
 \label{fig:wk}
@@
 
+For the $\perm{c_1, \ldots, c_n}{c}$ combinator, the whole language is the
+union of languages of the form $W(c_{\sigma(1)}) \cdot W(c) \cdot \ldots \cdot
+W(c) \cdot W(c_{\sigma(n)})$. This is because the \kw{perm}
+combinator allows for the string to match any permutation the $c_i$'s while also
+taking into account the separator $c$ in between each of the $c_i$'s. The kernel
+of the \kw{perm} combinator is the language $K(c_1) \cdot K(c) \cdot
+\ldots \cdot K(c) \cdot K(c_n)$, because the canonical permutation is the
+identity permuation, with each of the parts of the string that match $c_i$
+and $c$ canonized into $K(c_i)$ and $K(c)$ respectively.
+
+Figure~\ref{fig:canonizers} gives the inductive definitions of the
+$\canonizer{}$ function for each QRE.
 \begin{figure}[t]
@@
 \end{align*}
-\caption{QRE canonizers}
+\caption{QRE canonizers \saz{TODO: use $c_1$ and $c_2$ everywhere.}}
 \label{fig:canonizers}
 \end{figure}
+Again, the \kw{perm} combinator gives the most interesting
+definition,
+\begin{align*}\canonizer(\perm{c_1, \ldots, c_n}{c})(r_{\sigma(1)}
+\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \\
+= (\canonizer(c_1)(r_1)) \cdot (\canonizer(c)(s_1)) \cdot \ldots \cdot
+(\canonizer(c_n)(r_n))
+\end{align*}
+\noindent which as we have explained is defined so that the strings that match
+$c_i, c$ are placed in the canonical permutation which is the identity
+permutation before $\canonizer(c_i)$ and $\canonizer(c)$ are applied
+respectively.
+
+Finally, Figure ~\ref{fig:relations} gives the inductive defintion of the
+equivalence relation $\eqrel{c}$. These are the set-theoretic semantics of QREs
+as an equivalence relation on a regular language.
 
@@
 \section{Conclusion}
+\bcp{And future work!}  \saz{I'm lukewarm about including a description of
+  future work.  Unless there are specific conjectures or open problems that
+  are worth mentioning, I'm not sure that it's really that important.}
+\bcp{Agreed. But don't we have any ideas for how to extend or generalize or
+  things to do next??  E.g., is it completely trivial to extend all this to
+  asymmetric and symmetric lenses?  (If so, we should say so!)  Do we have
+  more ideas for QRE primitives?  Etc.}
 \label{concl}

### 2018-03-06 papers/QuotientLenses/main.tex
merged Benjamins changes into mine

@@
 ignoring details that a programmer views as inessential.
-
 Ordinary bijective lenses are fairly tricky to program\bcp{do we have
@@
 
-\section{Conclusion}
-\bcp{And future work!}  \saz{I'm lukewarm about including a description of
-  future work.  Unless there are specific conjectures or open problems that
-  are worth mentioning, I'm not sure that it's really that important.}
-\bcp{Agreed. But don't we have any ideas for how to extend or generalize or
-  things to do next??  E.g., is it completely trivial to extend all this to
-  asymmetric and symmetric lenses?  (If so, we should say so!)  Do we have
-  more ideas for QRE primitives?  Etc.}
+\section{Conclusion}\bcp{And future work!!!}  \saz{I'm luke warm about including
+  a description of future work.  Unless there are specific conjectures or open
+  problems that are worth mentioning, I'm not sure that it's really that important.}
 \label{concl}

### 2018-03-06 papers/QuotientLenses/main.tex
A bit more tidying

--- 2018-03-05 papers/QuotientLenses/main.tex
@@
 always contained in the whole language so we just use the inclusion as the
-choose function.}\bcp{I can fix.}}
+choose function.}\bcp{Done.  But I wonder if the picture should mention QREs
+explicitly, somehow... any ideas?}}
 \label{fig:attheedges}
@@
 $q: c \Leftrightarrow c'$ from the QREs $c, c'$ and a set of input-output pairs
-$\{(x_1, y_1), \ldots, (x_n, y_n)\}$ where the $x_i$'s are in $W(c)$ and the
+$\{(x_1, y_1),\allowbreak \ldots,\allowbreak (x_n, y_n)\}$ where the $x_i$'s are in $W(c)$ and the
 $y_i$'s are in $W(c')$. We wish $q$ to map the equivalence class of $x_i$ to the
@@
 Optician~\cite{optician} showed that given regular expressions
-$R, S$ and a set of input-output pairs $\{(r_1, s_1), \ldots, (r_n, s_n)\}$, if
+$R, S$ and a set of input-output pairs $\{(r_1, s_1),\allowbreak \ldots,\allowbreak (r_n, s_n)\}$, if
 there is a derivation $\ell : R \Leftrightarrow S$ of a bijective lens that
@@
 c'$ of a QRE lens as well as a derivation $\ell : K(c) \Leftrightarrow K(c')$
-that is consistent with the canonized examples $\{({x_1}', {y_1}'),
-\ldots, ({x_n}', {y_n}')\}$, then we can synthesize a quotient lens $q$ of
+that is consistent with the canonized examples $\{({x_1}', {y_1}'),\allowbreak
+\ldots,\allowbreak ({x_n}', {y_n}')\}$, then we can synthesize a quotient lens $q$ of
 type $c \Leftrightarrow c'$ that is consistent with $\{({x_1}', {y_1}'),
@@
 quotient lens $q: c \Leftrightarrow c'$, that is consistent with the examples
-$\{(x_1, y_1), \ldots, (x_n, y_n)\}$ it suffices to perform the following
+$\{(x_1, y_1),\allowbreak \ldots, (x_n, y_n)\}$ it suffices to perform the following
 procedure:
@@
   Synthesize a bijective lens $\ell : K(c) \Leftrightarrow K(c')$ that is
-  consistent with the examples $\{({x_1}', {y_1}'), \ldots, ({x_n}',
+  consistent with the examples $\{({x_1}', {y_1}'),\allowbreak
+  \ldots,\allowbreak ({x_n}', 
   {y_n}')\}$

### 2018-03-07 papers/QuotientLenses/main.tex
Polishing in 3.1 and 3.2 -- release locks

@@
 
-\bcp{We need two subsections here --- one on plain old bijective lenses, and
-one on bijective lens synthesis.}
+\bcp{We need to write two new subsections here --- one on plain old
+  bijective lenses, and one on bijective lens synthesis.}
 
@@
 
-The language of Quotient Regular Expressions (QREs) is given by the following
-grammar:
+The language of Quotient Regular Expressions (QREs) is given by the grammar
 \begin{align*}
@@
 & | \; \normalize{R_1}{R_2}{f} \sep c_1 \semicolon c_2 \sep c_1 \cdot c_2 \sep (c_1 \sep
-c_2) \sep c^*,
+c_2) \sep c^*
 \end{align*}
@@
 \subsection{Intuitions}
-In the \bibtex{} to EndNote example, we encountered the $\project{R}{s}$,
+In the \bibtex{} to EndNote example, we encountered lenses of the form $\project{R}{s}$,
 
 \noindent $\squash{c}{c'}{f}$ and $\perm{c_1, \ldots, c_n}{c}$ QREs, as well as
-the regular operators applied to these, as in\bcp{linebreak}
+the regular operators applied to these, as in\bcp{bad linebreak: rewrite}
 
@@
 \kw{squash} QRE and a QRE that projects away the extra whitespace in between
-\kw{first\_name} and \kw{last\_name}:
+\kw{first\_name} and \kw{last\_name}: \bcp{It would be easier to distinguish
+code examples from text if they were indented a bit.}
 
@@
 defined by $\normalize{R}{R'}{f}$ is thus the equivalence relation on
-$\mathcal{L}(R)$ defined by the {\em fibres} of $f$ (i.e. $r \sim r'$ if and
-only if $f(r) = f(r'))$. The canonizing function $f$ is required to be
+$\mathcal{L}(R)$ defined by the ``{fibres}'' of $f$---i.e., $r \sim r'$ if and
+only if $f(r) = f(r')$. The canonizing function $f$ is required to be
 surjective and idempotent.
 
-The $\normalize{R}{R'}{f}$ QRE is the most general QRE in that for each QRE
-$q$, there exist regular expressions $R, R'$ and a surjective, idempotent
+The $\normalize{R}{R'}{f}$ QRE is the most general QRE in that, for any QRE
+$q$, there exist regular expressions $R$ and $R'$ and a surjective, idempotent
 function $f:\mathcal{L}(R) \longrightarrow \mathcal{L}(R')$ such that $q$ and
@@
 express all QREs requires that we introduce W(c), K(c) and Canonizer(c) which
-we haven't introduced at this point in the paper}
+we haven't introduced at this point in the paper}\bcp{But we can at least
+state an informal claim and give a forward pointer to where it is made more
+precise later.  Or, since these definitions appear quite soon, perhaps we
+can just reorganize the material a bit so that the definitions come before
+the intuition plus a precise claim.}
 
@@
 $\normalize{R}{R'}{f}$ combinator gives a sufficient condition for potential
-QREs to be ``well-behaved'', particularly with respect to our strategy for
-synthesizing quotient lenses which we shall introduce in Section~\ref{synth}.
+QREs to be ``well behaved,'' particularly with respect to the strategy for
+synthesizing quotient lenses that we will introduce in Section~\ref{synth}.
 
@@
 $\mathcal{L}(R') \subseteq \mathcal{L}(R)$ described by $R'$. QREs therefore are
-not as expressive as Boomerang canonizers. On the other hand, QREs enable us to
+not as expressive as Boomerang canonizers.\bcp{Not sure I understand what
+  this means, practically.  First, is it true that QREs can express any
+  boomerang canonizer that maps from a language to a subset of the same
+  language?  And second, can't at least some Boomerang canonizers be expressed as the
+  composition of a QRE canonizer from R to R' plus a lens from R' to S?} On
+the other hand, QREs enable us to 
 express many equivalences that occur in real world data as we shall demonstrate
@@
 $k$ in $\mathcal{L}(K(c))$ such that $w$ is equivalent to $k$ mod $\eqrel{c}$,
-then the equivalence classes of $\eqrel{c}$ are the same as the fibres of
+then the equivalence classes of $\eqrel{c}$ are the same as the
+fibres\bcp{this is not a very familiar term for lots of people; if we're
+  going to use it a lot, we should define it more carefully} of
 $\canonizer(c)$.

### 2018-03-07 papers/QuotientLenses/main.tex
Polishing in 3.3 -- still holding no locks

@@
 \end{enumerate}
-Observe that since $\canonizer(c)$ computes $\canonizer(c)(w)$ as the unique
+Observe that, since $\canonizer(c)$ computes $\canonizer(c)(w)$ as the unique
 $k$ in $\mathcal{L}(K(c))$ such that $w$ is equivalent to $k$ mod $\eqrel{c}$,
-then the equivalence classes of $\eqrel{c}$ are the same as the
+it follows that the equivalence classes of $\eqrel{c}$ are the same as the
 fibres\bcp{this is not a very familiar term for lots of people; if we're
@@
 when applying the regular combinators to QREs, we require that if a string $s$
-matches any of the regular expressions $W(c) \cdot W(c')$, $W(c) \sep W(c')$,
+matches any of the regular expressions\bcp{The following list is hard to
+  parse.  A general rule is: Always separate adjacent math expressions by
+  something more than just punctuation.} $W(c) \cdot W(c')$, $W(c) \sep W(c')$,
 $W(c)^*$, $K(c) \cdot K(c')$, $K(c) \sep K(c')$, $K(c)^*$, then $s$ matches
-that regular expression in only one way. This unambiguity condition is called
-{\em strong unambiguity}~\cite{Sippu1988}, and is necessary, firstly because it
+that regular expression in only one way.
+
+This  condition is called
+{\em strong unambiguity}~\cite{Sippu1988}, and it is necessary,
+firstly,\bcp{is there a ``secondly'' coming?  I don't see it} because it
 ensures that the canonizing function of a QRE is well-defined.
-
-For example consider the QRE
+%
+For example consider the QRE\bcp{Displayed math should be punctuated either
+  (1) exactly as if it were inline (i.e., this would need a period at the
+  end) or (2) with a sentence terminator (a colon or period) beforehand.  In
+this case, (1) would be confusing (since . is also part of QRE syntax), so
+(2) is preferable.}
 
@@
 We also require that the regular combinators applied to
-the kernels are unambiguous since the underlying lenses will end up operating
-on the kernels, and these lenses impose the same restrictions for similar
-reasons.
+the kernels of QREs be unambiguous, since the underlying\bcp{?} lenses will
+end up operating on these kernels, and these lenses impose the same
+restrictions for similar reasons.
 %
-To this end, we say that regular expressions $R$ and $S$ are
-\textit{unambiguosly concatenable}, written $R \cdot^! S$ if for all strings
+To this end\bcp{which end?}, we say that regular expressions $R$ and $S$ are
+\textit{unambiguosly concatenable}, written $R \cdot^! S$, if, for all strings
 $r, r' \in \mathcal{L}(R)$ and $s, s' \in \mathcal{L}(S)$, if $r \cdot s = r'
 \cdot s'$, then $r = r'$ and $s = s'$. We say that a regular expression $R$ is
-\textit{unambiguosly iterable}, written $R^{*!}$ if for all strings $r_1,
+\textit{unambiguosly iterable}, written $R^{*!}$ if, for all strings $r_1,
 \ldots, r_m$ and $r'_1, \ldots, r'_n \in \mathcal{L}(R)$, if $r_1 \cdot \ldots
@@
 
-We say that a regular expression $R$ is \textit{strongly unambiguous} if and
-only if (1) $R = \varnothing$, or (2) $R = S_1 \cdot S_2$ with $S_1, S_2$
+We say that a regular expression $R$ is \textit{strongly unambiguous} when
+(1) $R = \varnothing$, or (2) $R = S_1 \cdot S_2$ with $S_1, S_2$
 strongly unambiguous and $S_1 \cdot^! S_n$, or (3) $R = S_1 \sep S_2$ with
@@
 
-Figure~\ref{fig:qrerules} gives the inferrence rules for deriving well-formed
+Figure~\ref{fig:qrerules} gives the inference rules for deriving well-formed
 QREs. As expected, the unambiguity conditions come into play when defining QREs
-using the regular combinators. For example, the inferrence rule for
-concatenation is as follows:
+using the regular combinators. For example, the inferrence\bcp{check spelling} rule for
+concatenation is as follows:\bcp{Rather than repeating rules inline, give
+  them names in the figure.}
 \begin{prooftree}
@@
 
-\noindent The second hypothesis for the \textit{permutation} rule says that for
+\noindent The second hypothesis says that, for
 any two different permutations $\sigma$ and $\theta$, the languages

### 2018-03-07 papers/QuotientLenses/main.tex
Bit of polishing in 3.4

@@
 \subsection{Semantics of QREs}
+
+\bcp{General typesetting comment: I frequently find the mixture of lens
+  expressions and ordinary text hard to read because the fonts used for them
+  are rather similar -- I have to look closely to notice that some
+  characters have serifs and others don't.  We should not change the regular
+  font -- that's mandated by the SIGPLAN style -- but could we make the lens
+  font be something a bit more different?}
+
 As we mentioned before, each QRE $c$ enables us to express the whole and kernel
@@
 $\mathcal{L}(W(c))$ that has $\mathcal{L}(K(c))$ as a complete set of
-representatives and a canonizing function $\canonizer(c):\mathcal{L}(W(c))
+representatives, and a canonizing function $\canonizer(c):\mathcal{L}(W(c))
 \longrightarrow \mathcal{L}(K(c))$ whose fibres are the same as $\eqrel{c}$.
@@
 $W(c_1) \sep W(c_2)$. This is because the \kw{squash} combinator uses the
-function $f : \mathcal{L}(W(c_1)) \longrightarrow \mathcal{L}(W(c_2))$ to maps
+function $f : \mathcal{L}(W(c_1)) \longrightarrow \mathcal{L}(W(c_2))$ to map
 any strings from $\mathcal{L}(W(c_1))$ into $\mathcal{L}(W(c_2))$ (leaving the

### 2018-03-08 papers/QuotientLenses/main.tex
sec 4 edits

@@
 \label{QRE-lenses}
+
+\dpw{I believe the purpose of the preamble was to explain the structure of the subsections and why the
+material presented there appears.  I tried to rewrite this to communicate that a bit more clearly.
+Please edit to improve!}
 As we have seen, QREs express a broad class of equivalence relations
@@
 
-Now let $c$ and $c'$ be QREs. Given a bijection $\ell : K(c) \Leftrightarrow
-K(c')$, we may define a quotient lens $q : W(c)/\eqrel{c} \Leftrightarrow
-W(c')/\eqrel{c'}$ by
-\begin{equation}\label{normalform}
-q.get = \ell \circ \canonizer(c) \text{ and } q.put = \ell^{-1} \circ
-\canonizer(c'),
-\end{equation}
-
-\noindent In other words, the get function of the $q$ first canonizes the source
-data using $\canonizer(c)$ and then applies $f$ to the result, while the
-put function of the $q$ first canonizes the target data using
-$\canonizer(c')$ and the applies $f^{-1}$ to the result. This is our approach in
-defining QRE lenses.
-
-One potential pitfall with QRE lenses is that the combinators that
-we define on them may not preserve this normal form. For example, the
-composition $q' \; ; \; q$ of QRE lenses $q$ and $q'$ may not be equivalent to
-any quotient lens that has the canonical normal form of a bijective lens with
-canonizers a the edges. Consequently, in order to guarantee that every QRE
-lens $q$ is equivalent to a QRE lens $q'$ which has the canonical form of
-a bijection with canonizers at the edges, we need to identify a set $S$ of
-bijections that will ensure that QRE lenses are closed under the action of QRE
-lens combinators. Our choice is to use the class of \textit{bijective lenses},
-which we briefly discuss in the next section.
+One natural way to provide a language for defining bijective quotient lenses is to
+(1) define a language of bijections and then (2) add quotients by allowing canonizers
+to be prepended or postpended to bijections and (3) allowing composition of such quotient
+lenses via the regular operators and functional composition.  Indeed, this is the program
+that we follow, with Section \ref{subsec:bijective-lenses}
+defining bijections and Sections\ref{subsec:qre-lenses-syntax} and
+\ref{subsec:qre-lenses-semantics}
+defining the quotients and their composition operators.
+
+However, we also have a secondary objective, which is to support synthesis of lenses: When
+provided with QREs describing the source and target languages, we would like to be able to
+generate quotient lenses automatically.  One way to do that is to use the QREs to generate
+canonizers $c$ and $c'$ and then to synthesize bijective lenses between the kernel languages
+for $c$ and $c'$.  Doing so provides us with quotient lenses defined by a particular normal
+form, one in which lenses are defined by a canonizer to transform source (and target) into a normal
+form and a bijection to transform between canonical forms for the source and target.
+However, this normal form does not allow for sequential composition (or composition via
+the regular operators) of quotients. Hence, an important technical question is whether we give
+up expressiveness if we restrict ourselves to this normal form.  Fortunately, as shown in
+section~\ref{subsec:qre-lens-normal-forms}, we do not.
+
+%% Now let $c$ and $c'$ be QREs. Given a bijection $\ell : K(c) \Leftrightarrow
+%% K(c')$, we may define a quotient lens $q : W(c)/\eqrel{c} \Leftrightarrow
+%% W(c')/\eqrel{c'}$ as follows.
+%% \begin{equation}\label{normalform}
+%% q.get = \ell \circ \canonizer(c) \text{ and } q.put = \ell^{-1} \circ
+%% \canonizer(c'),
+%% \end{equation}
+
+%% \noindent In other words, the get function of the $q$ first canonizes the source
+%% data using $\canonizer(c)$ and then applies $f$ to the result, while the
+%% put function of the $q$ first canonizes the target data using
+%% $\canonizer(c')$ and the applies $f^{-1}$ to the result.
+%% We define QRE lenses with canonizers on the edge and bijective lenses
+%% in the middle because QREs define the canonizers ($c$ and $c'$) and the kernel
+%% languages $K(c)$ and $K(c')$ directly.  The kernel languages can then
+%% serve as inputs to a synthesis procedure for bijections~\cite{optician}.
+%% Hence, from QREs alone, we are able to generate full bijective quotient
+%% lenses.
+
+%% \dpw{Check that I'm getting this right -- I replaced the previous paragraph.}
+%% One potential drawback of our approach is that it does not admit
+%% composition of quotient lenses ($q' \; ; \; q$) that might interleave canonizers and bijections.
+%% Our key technical result is that the lack of
+%% composition does not reduce the expressiveness of the language:  Any such composition can be
+%% rewritten as a single quotient lens with canonizers on the edges.
+
+%% One potential pitfall with QRE lenses is that the combinators that we define
+%% on them may not preserve this normal form.
+%% For example, the composition of
+%% two quotient lenses $q' \; ; \; q$ of QRE lenses $q$ and $q'$ may
+%% not be equivalent to
+%% any quotient lens that has the canonical normal form of a bijective lens with
+%% canonizers a the edges. Consequently, in order to guarantee that every QRE
+%% lens $q$ is equivalent to a QRE lens $q'$ which has the canonical form of
+%% a bijection with canonizers at the edges, we need to identify a set $S$ of
+%% bijections that will ensure that QRE lenses are closed under the action of QRE
+%% lens combinators. Our choice is to use the class of \textit{bijective lenses},
+%% which we briefly discuss in the next section.
 
 \subsection{Bijective Lenses}
+\label{subsec:bijective-lenses}
+
 We define the set of \textit{bijective lenses} to be the set of bijections
-between regular languages created using Boomerang lens combinators.
-The syntax for the language of bijective lenses is given by
+between regular languages constructed using the following combinators. \dpw{I said ``lens combinators''
+instead of ``Boomerang lens
+combinators'' because Boomerang has many other combinators... wasn't sure what the perfect thing to say was.}
+%The syntax for the language of bijective lenses is given by
 $$\ell := \mathit{id} \; (R) \sep const(s, s') \sep  swap(\ell,
@@
 
-The denotation of a lens $\ell$ is $\llbracket \ell \rrbracket \subseteq
-\mathit{String} \times \mathit{String}$. If $(s_1, s_2) \in \llbracket \ell
-\rrbracket$, then $\ell$ maps between $s_1$ and $s_2$.
-\begin{align*}
-\llbracket R \rrbracket &= \{(r, r) \sep r \in \mathcal{L}(R)\}\\
-\llbracket const(r, s) \rrbracket &= \{(r, s)\}\\
-\llbracket swap(\ell, \ell') \rrbracket &= \{(s \cdot t, t' \cdot s') \sep
-(s, s') \in \llbracket \ell \rrbracket \text{ and } (t, t') \in \llbracket
-\ell' \rrbracket\}\\
-\llbracket \ell \cdot \ell' \rrbracket &= \{(s \cdot t, s' \cdot t) \sep
-(s, s') \in \llbracket \ell \rrbracket \text{ and } (t, t') \in \llbracket
-\ell' \rrbracket\}\\
-\llbracket \ell \sep \ell' \rrbracket &= \{(s \cdot t) \sep
-(s, t) \in \llbracket \ell \rrbracket \text{ or } (s, t) \in \llbracket
-\ell' \rrbracket\}\\
-\llbracket \ell^* \rrbracket &= \{(s_1 \cdot \ldots \cdot s_n, t_1 \cdot \ldots
-\cdot t_n) \sep (s_i, t_i) \in \llbracket \ell \rrbracket \text{ for } 1
-\leq i \leq n\}\\
-\llbracket \ell_2 \circ \ell_1 \rrbracket &= \{(r, t) \sep \text{there exists }s
-\text{ such that } (r, s) \in \llbracket \ell_1 \rrbracket \text{ and } (s, t)
-\in \llbracket \ell_2 \rrbracket\}
-\end{align*}
 The $\mathit{const}(s, t)$ lens replaces the string $s$ with $t$ in the source
 data in the forward direction, and $t$ with $s$ in the target data in the backward
-direction. $\mathit{id}(R)$ applies the identity function to the source and target
+direction. The $\mathit{id}(R)$ lens applies the identity function to the source and target
 in $\mathcal{L}(R)$ in both directions. The composition lens $\ell' \circ \ell$
@@
 
+The denotation of a lens $\ell$ is $\llbracket \ell \rrbracket \subseteq
+\mathit{String} \times \mathit{String}$. If $(s_1, s_2) \in \llbracket \ell
+\rrbracket$, then $\ell$ maps between $s_1$ and $s_2$.
+\begin{align*}
+\llbracket R \rrbracket &= \{(r, r) \sep r \in \mathcal{L}(R)\}\\
+\llbracket const(r, s) \rrbracket &= \{(r, s)\}\\
+\llbracket swap(\ell, \ell') \rrbracket &= \{(s \cdot t, t' \cdot s') \sep
+(s, s') \in \llbracket \ell \rrbracket \text{ and } (t, t') \in \llbracket
+\ell' \rrbracket\}\\
+\llbracket \ell \cdot \ell' \rrbracket &= \{(s \cdot t, s' \cdot t) \sep
+(s, s') \in \llbracket \ell \rrbracket \text{ and } (t, t') \in \llbracket
+\ell' \rrbracket\}\\
+\llbracket \ell \sep \ell' \rrbracket &= \{(s \cdot t) \sep
+(s, t) \in \llbracket \ell \rrbracket \text{ or } (s, t) \in \llbracket
+\ell' \rrbracket\}\\
+\llbracket \ell^* \rrbracket &= \{(s_1 \cdot \ldots \cdot s_n, t_1 \cdot \ldots
+\cdot t_n) \sep (s_i, t_i) \in \llbracket \ell \rrbracket \text{ for } 1
+\leq i \leq n\}\\
+\llbracket \ell_2 \circ \ell_1 \rrbracket &= \{(r, t) \sep \text{there exists }s
+\text{ such that } (r, s) \in \llbracket \ell_1 \rrbracket \text{ and } (s, t)
+\in \llbracket \ell_2 \rrbracket\}
+\end{align*}
+
 Each bijective lens $\ell$ has a type $\ell : R \Leftrightarrow S$ where $R$ and
@@
 language of $\ell$ is $\mathcal{L}(R)$ and the target language of $\ell$ is
-$\mathcal{L}(S)$. Interestingly, with the bijective lens type system, a lens
+$\mathcal{L}(S)$.  The typing rules for bijective lenses are given in
+Figure~\ref{fig:lensrules}.  The final rule in the figure deserves special attention.
+It states that a lens
 $\ell : R \Leftrightarrow S$ can also be considered to be of type $\ell : R'
@@
 
-The other typing rules for bijective lenses are given in
-Figure~\ref{fig:lensrules}.
+\noindent
+This rule makes it significantly more difficult to synthesize a bijective lens from its type.
 
@@
 \subsection{Syntax of QRE Lenses}
+\label{subsec:qre-lenses-syntax}
 Having given a brief overview of the class of bijective lenses, we now introduce
@@
 \subsection{Semantics of QRE Lenses}
+\label{subsec:qre-lenses-semantics}
 
@@
 \]
-\[
-\begin{array}[b]{c@{\qquad}l}
-\Rule{q : c'  \Leftrightarrow c'' &
-\wf{c} &
-K(c) = W(c')}
-{\kw{lquot}(c, q): c' \circ c \Leftrightarrow c''} &
-\begin{array}[b]{rcl}
-\llbracket \kw{lquot}(c, q) \rrbracket.get  &=& \llbracket q
-\rrbracket.get \circ \canonizer(c)\\
-\llbracket \kw{lquot}(c, q) \rrbracket.put &=& \llbracket q \rrbracket.put
-\end{array}
-\end{array}
-\]
+%% \[
+%% \begin{array}[b]{c@{\qquad}l}
+%% \Rule{q : c'  \Leftrightarrow c'' &
+%% \wf{c} &
+%% K(c) = W(c')}
+%% {\kw{lquot}(c, q): c' \circ c \Leftrightarrow c''} &
+%% \begin{array}[b]{rcl}
+%% \llbracket \kw{lquot}(c, q) \rrbracket.get  &=& \llbracket q
+%% \rrbracket.get \circ \canonizer(c)\\
+%% \llbracket \kw{lquot}(c, q) \rrbracket.put &=& \llbracket q \rrbracket.put
+%% \end{array}
+%% \end{array}
+%% \]
 
@@
 \subsection{Normal Forms of QRE Lenses}
+\label{subsec:qre-lens-normal-forms}
 Recall that our approach in defining QRE lenses is to have each QRE lens $q: c
@@
 \begin{proof}
-We elide a full proof of this statement, since the proof is attained via a
-straightforward induction on the derivation $q : c \Leftrightarrow c$ (the
-full details can be found in the appendix {\sam{which I need to write}}). Since
-bijective lenses are closed under strongly unambiguous regular operators, the
-tricky case is composition. The composition case is more involved since we need
-to show that we can get rid of the canonizers in the middle:
+Sketch:
+% We elide a full proof of this statement, since the proof is attained via a
+% straightforward induction on the derivation $q : c \Leftrightarrow c$ (the
+% full details can be found in the appendix {\sam{which I need to write}}). Since
+% bijective lenses are closed under strongly unambiguous regular operators, the
+% tricky case is composition.
+The proof follows by induction on the derivation of $q : c \Leftrightarrow c$.
+The most interesting part of the proof is the case for functional composition
+as we must demonstrate that it is possible to eliminate the canonizers in the
+middle of the term.
 
@@
 \fi
+
 \section{Synthesizing QRE Lenses}

### 2018-03-08 papers/QuotientLenses/main.tex
Made changes

--- 2018-03-07 papers/QuotientLenses/main.tex
@@
 
-\lstset{
-language=Caml, 
-basicstyle=\linespread{0.8}\upshape\sffamily,
-keywordstyle=\upshape\sffamily\color{dkblue},
-keepspaces=true,
-framexleftmargin=1ex,
-framexrightmargin=1ex,
-showstringspaces=false,
+\lstset{ language=Caml, basicstyle=\linespread{0.8}\upshape\sffamily,
+keywordstyle=\upshape\sffamily\color{dkblue}, keepspaces=true,
+framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=false,
 commentstyle=\itshape\rmfamily,
 emph={synth,project,perm,squash,normalize,using,ins,del,lens},
-emphstyle=\upshape\sffamily\color{dkblue},
-columns=flexible,
+emphstyle=\upshape\sffamily\color{dkblue}, columns=flexible,
 % BCP: I find this distracting:
 % stringstyle=\sffamily\color{dkred},
-mathescape,
- }
+mathescape, }
 
@@
 \begin{abstract}
-While there currently exist many bidirectional programming languages, many of
-them do not have built-in support for transformations that operate on data
-modulo equivalence relations such as whitespace or permutation
-quotienting.\bcp{I'm worried that most readers will not be able to follow
-  that.  Moreover, this way of introducing the paper feels kind of
-  negative---most languages are insufficient, and Boomerang is hard to
-  program.  Could we say it in a more accessible, more positive way?}
-Moreover, langauges such as Boomerang which do allow for such transformations
-are often difficult to program \sam{why?}. To address these shortcomings, this
-paper (1) introduces a language of {\em quotient regular expressions} or QREs
-for defining equivalences on regular expressions as well as a language of {\em
-QRE lenses}, which are lenses that map between QREs, and (2) shows how to {\em
-synthesize} QRE lenses from a pair of QREs and a set of example input-output
-pairs. We also describe an implementation of QREs and QRE lens synthesis in
-Boomerang, and demonstrate the utility of our implementation by using it to
-synthesize QRE lenses between a variety of real-world data formats used by
-the {\tt data.gov} database. 
+There currently exist many bidirectional programming languages, some of which
+allow a programmer to define a transformation that respects an equivalence
+relation. For example, a language with this feature may allow a programmer to
+define a transformation which ignores the order of fields in an XML file, so
+that two files that have the same fields with the same data but with the fields
+occuring in a different order are considered the same by the
+transformation. This paper simplifies the task of programming such {\em
+quotient lenses} in two main ways. First, we introduce {\em
+quotient regular expressions} or QREs which are regular expressions
+augmented with extra syntax that makes it easier to specify equivalence
+relations on regular langauges than currently existing tools. Secondly, we
+introduce {\em QRE lenses}, which are lenses that map between QREs, and show
+how to {\em synthesize} QRE lenses from a pair of QREs and a set of example
+input-output pairs. We also describe an implementation of QREs and QRE lens
+synthesis in the bidirectional programming language Boomerang, and demonstrate
+the utility of our implementation by using it to synthesize QRE lenses between
+various real-world data formats used by the {\tt data.gov} database.
 \end{abstract}
@@
 \section{Introduction}
-
 Oftentimes, the same data is stored in a variety different formats. For
-example, \bibtex{} and EndNote both store bibliography data; spreadsheet data
-can be represented as a CSV file or a TSV file; web APIs can return data as a
-JSON or
-XML object.\bcp{These examples show that the same data {\em could} be
-  represented in different formats, not that it needs to be.} When different applications manipulate the same data in different formats,
-synchronizing the applications\bcp{Why are we talking about synchronizing
-  applications?  Comes out of the blue.} amounts to synchronizing their
-respective data
-formats. One way to synchronize data formats is to use \emph{lenses}, \emph{i.e.,} programs that represent pairs of functions for translating data from
-format A to B (the \emph{get} direction) and back again (the \emph{put}
-direction).  These lenses, when defined using domain-specific
-languages like Boomerang~\cite{boomerang} or
-BiGul~\cite{ko2016bigul}\bcp{why mention exactly these two, out of all the
-  bidirectional languages out there?}, typically come with a variety of
-guarantees\bcp{grammatically, that says that each lens comes with a variety
-  of guarantees---is that what you meant?},
-the \emph{lens laws}, that help\bcp{why ``help''?} ensure put and get
-functions are effective 
-inverses\bcp{what is an effective inverse?} of one another.
-
-One way\bcp{Do you mean to highlight the suggestion that there are other
-  ways too?} to classify lenses is by the laws they obey.  Amongst the most
-restrictive lenses are the \emph{bijective lenses}.  These lenses demand
-that get and put functions form a bijection, meaning no information
-is lost when one translates from one data format to another.  While this
-strong guarantee can be useful in some contexts, it is more often the case
+example, bibliography data may be stored in \bibtex{} and
+EndNote formats; spreadsheet data can be represented as a CSV file
+or a TSV file; web APIs can return data as a JSON or XML object. In such cases,
+it is necessary to synchronize the two or more copies of data so that they
+store the same information.
+
+One way to synchronize data stored in different formats is to use \emph{lenses},
+\emph{i.e.,} programs that represent pairs of functions for translating data
+from format A to B (the \emph{get} direction) and back again (the \emph{put}
+direction). Lenses are usually derived using domain-specific
+bidirectional languages which guarantee that the lenses defined in that language
+satisfy a certain set of \emph{lens laws} that give guarantees on how the two
+transformations interact.
+
+Amongst the most restrictive lenses are the \emph{bijective lenses}.  These
+lenses demand that get and put functions form a bijection, meaning no
+information is lost when one translates from one data format to another.  While
+this strong guarantee can be useful in some contexts, it is more often the case
 that some of the information in one or both of the data formats is superfluous.
@@
 whitespace between characters.  In other circumstances, capitalization
-or the order of fields in a record or items in a list may be unimportant.
-%
-%bijective lenses have
-%strong correctness guarantees, they are often
-%However, these lens laws are too restrictive for many important transformations.
-%The bijective lens laws do not permit transformations that ignore this
-%unimportant information---.
-Many data formats have other ad hoc rules as well.
-In bibliographies, for example, recall that in \bibtex{},
-an author can be specified as
-``\cd{lastname, firstname}'' or as ``\cd{firstname lastname}''---it does not matter which of the
-two formats is used.
-%While it is important to allow authors
-%to be input in both ways, we would like these authors to only be output as
-%\cd{firstname lastname}, but the lens laws prevent such functions.
-%
-{\em Quotient lenses}~\cite{quotientlenses} loosen many of the restrictions
-of bijective lenses by introducing an equivalence relation\bcp{actually,
-  two: one on each side; but more importantly ``introducing an equivalence
-  relation'' is rather vague.} and considering
-bijections modulo that relation.  Such quotient lenses make it possible
-to align many more data sources than would otherwise be possible by
-ignoring details that a programmer views as inessential.
-
-Ordinary bijective lenses are fairly tricky to program\bcp{do we have
-  concrete evidence for this?}, especially for
-beginners\bcp{or this?}, as they require programmers to think in two
-directions at once\bcp{my own experience suggests that this is the less
-  important issue...}
-and learn a new type system, which comes with some tricky ambiguity
-constraints\bcp{...and this is the bigger one.  But we don't explain what
-  these ambiguity constraints are or why they must be imposed.}.
-Unfortunately, quotient lenses do not improve the situation\bcp{Again, it
-  feels rather negative.}:  To
-write a quotient lens in Boomerang, programmers must work with standard lens
-combinators as well as a system of \emph{canonizers}.
-These canonizers encode two functions, one that
-maps data represented in the original external format
-to a canonical element, and another, in the opposite direction, that chooses
-which elements of the original format correspond to values in the canonical
-format. Furthermore, if a user wants to define the domain and range of
-a lens explicitly (a programming practice that is encouraged), she must
-provide regular expressions defining input and output formats.
-Consequently, to define a quotient lens, one must often
-construct three different sorts
-of terms: (1) regular expressions that define the two data formats
-to be processed, (2) canonizers to normalize the data
-and choose representatives, and (3) a bijective lens that maps
-one canonical form into another.  Somewhat frustratingly,
-these three terms share a fair amount of structure.
-%Moreover, writing the lens between the equivalence classes can be
-%challenging due to the rather fiddly typechecking, mainly due to
-%``non-ambiguity'' conditions in the lens well-formedness rules.
-
-In this paper, we improve the process of programming quotient lenses
-in two distinct, but connected ways. First, we
-introduce a new language of {\em Quotient Regular Expressions} (QREs), which
-enable a programmer to simultaneously specify a regular expression as well as
-a set of canonical representatives for strings that match the regular
-expression. With
-QREs, programmers do not need to define the data and canonization functions
-separately---by adding simple annotations to ordinary regular expressions,
-canonization functions can be inferred.
-
-Second, we use QREs to define a class of quotient lenses called {\em QRE
-lenses}, and, we show how to synthesize such QRE
-lenses automatically.
-More specifically, in
-order to generate a lens, a programmer provides a pair of QREs
-describing the source and target formats, plus 
+or the order of fields in a record or items in a list may be unimportant. Many
+data formats have other ad hoc rules as well. In bibliographies, for example,
+recall that in \bibtex{}, an author can be specified as
+``\cd{lastname, firstname}'' or as ``\cd{firstname lastname}''---it does not
+matter which of the two formats is used.
+
+{\em Quotient lenses} loosen many of the restrictions of bijective lenses by
+allowing a programmer to define an equivalence relation on the source and
+target formats so that the round-tripping lens laws hold modulo these
+equivalence rather than modulo equality. Quotient lenses thus make it possible
+to align many more data sources than would otherwise be possible, by ignoring
+details that a programmer views as inessential. The only language we are aware
+of that implements a general class of quotient lenses is {\em
+Boomerang}~\cite{quotientlenses}. While Boomerang's quotient lenses enable a
+programmer to define a very general class of quotient lenses, these lenses to
+have two main drawbacks.
+
+First, in order to define an equivalence relation in Boomerang, a programmer
+must define a {\em canonizer}; pair of functions $f$ and $g$ where $f$ maps
+data represented in the original external format to a canonical element and $g$
+chooses which elements of the original format correspond to values in the
+canonical format. The disadvantage of this approach is that in order for a
+programmer to {\em specify} a quotient lens in Boomerang, she must define four
+different data formats: two ``outer'' formats which specify the data that is to
+be transformed by the lens, and two ``inner'' formats which specify the
+canonical formats for the source and target data.
+
+Our solution to this shortcoming is to introduce {\em Quotient Regular
+Expressions} or QREs, which enable a programmer to simultaneously specify a
+regular expression as well as a set of canonical representatives for strings
+that match the regular expression. QREs have the added advantage that once the
+programmer has defined a well formed QRE, then she need not define a
+canonization function seperately; the canonization function can automatically be
+inferred from the QRE.
+
+The second challenge with programming lenses in Boomerang is that even after the
+programmer has specified the data as well as a canonizer that maps each member
+in the source and target sets to their respective canonical representative, she
+must still define the underlying lens that maps between the respective sets of
+canonical representatives. This can be a slow and tedious task due to the
+constraints that Boomerang imposes on its terms. For example, Boomerang
+programmers often must rearrange the order of data items by recursively using
+operators that swap adjacent fields. Furthermore, the Boomerang type checker is
+very strict, disallowing many programs because they contain ambiguity about how
+certain data is transformed.
+
+Our solution to this problem is that we use QREs to define a class of quotient
+lenses called {\em QRE lenses}, and, we show how to synthesize such QRE lenses
+automatically. More specifically, in order to generate a lens, a programmer
+provides a pair of QREs describing the source and target formats, plus
 a (possibly empty) suite of examples that demonstrates how the
-lens should act in specific cases.  Given this information,
-we can exploit and extend recent work on algorithms for synthesis of
-bijective lenses~\cite{optician} to generate quotient lenses.
-The lenses generated by
+lens should act in specific cases.  Given this information, we can exploit and
+extend recent work on algorithms for synthesis of bijective
+lenses~\cite{optician} to generate quotient lenses. The lenses generated by
 our synthesis algorithm are guaranteed to translate data back and forth
@@
 
-The following list summarizes the central contributions of our
-work.\bcp{Ugh.} 
+The following list summarizes the central contributions of our work.
 \begin{enumerate}
@@
   lenses whose types are given by QREs. Our main technical contribution
-  is a proof that our QREs are just as expressive as the full set of bijective
-  quotient lenses.
+  is a proof that each QRE lens is functionally equivalent to a QRE lens in the
+  canonical form of a bijective lens with ``canonizers at the edges''.
    (Section~\ref{QRE-lenses}).
@@
 \label{sec:example}
-
-%In this section, we explain the difficulties of programming with quotient
-%lenses in Boomerang and show, by way of example, how our
-%{\em Quotient Regular Expressions} (QREs) help alleviate these difficulties.
-
-\emph{Lenses} define computations that can be read in two directions, the
-get direction (forwards) and the put direction (backwards).  The simplest
-lenses define bijections between source and data, but we are interested in
-\emph{quotient lenses}, which define bijections modulo some equivalence
-relation\bcp{two!}.  The core ideas of quotient lenses were developed by
-Foster et al~\cite{quotientlenses} and are embodied in the Boomerang
-language.  This section describes and motivates a 
-new notation for defining such lenses.
-
-\bcp{We need to write two new subsections here --- one on plain old
-  bijective lenses, and one on bijective lens synthesis.}
-
-\subsection{Quotient Lenses}
-
-To begin, consider a bidirectional transformation that converts
-\bibtex{} citation records such as
+\subsection{Bijective Lenses}
+Consider a bidirectional transformation that converts \bibtex{} citation records
+such as
 \begin{verbatim}
-@Book {conway,
-Author = "J. H. Conway",
-Title = {Regular Algebra and Finite Machines}
+@Book {Doe,
+Author = "John Doe",
+Title = {Generic Title}
 }
@@
 %0 Book
-%T Regular Algebra and Finite Machines
-%A J. H. Conway
-%F conway
+%F Doe
+%A John Doe
+%T Generic Title
 \end{verbatim}
@@
 and vice versa.
-%
-Notice that transformations from one format to the other preserve
-most of the information present in the data:  the title, the author,
-and the label, for instance.  However, some of the syntactic information
-may be lost.  For instance, users are typically uninterested in preserving
-whitespace between words or
-between fields.  The order of author and title fields is also likely
-irrelevant and there may be equivalent ways of writing the same
-name:  ``Conway, J. H.'' vs ``J. H. Conway.''
-
-%% Moreover,
-%% if two \bibtex{} files differ only in ``inessential'' details, we would like
-%% to treat them the same by mapping them to identical EndNote representations.
-%% In this , the transformation should allow whitespace characters to occur
-%% arbitrarily between the words in the label, author, and title fields, as well as
-%% in between them. Second, an author's name can have the surname followed by a
-%% comma and then the other names separated by arbitrary whitespace, or just all
-%% the names separated by arbitrary whitespace (for example, the names ``Conway, J.
-%% H.'' and ``J. H. Conway'' are equivalent). Lastly, the transformation should
-%% also ignore the order in which the author and title fields occur in each
-%% citation record.
+
+{\em Bijective lenses} are designed to define a bidirectional transformation
+such as this, where the data formats can be matched in a one-to-one manner, in
+this case by matching the label, author and title fields. Boomerang allows
+programmers to define such bijecive transformations using combinators which
+favour a compositional approach in which the programmer defines small lenses
+which are then put together using the combinators to define larger and larger
+lenses until the final lens is defined.
+
+Each bijective lens is constructed from simple primitive combinators including:
+\begin{itemize}
+  \item \kw{del} $s$: \textit{delete} the constant string $s$ in the get
+  direction; insert it in the put direction.
+  \item \kw{ins} $s$: \textit{insert} the constant string $s$ in the get
+  direction; delete it in the put direction.
+  \item \kw{copy} $R$: copy the text matching the regular expression $R$ in
+  both directions.
+\end{itemize}
+In addition, such primitives may be combined using the regular operators
+Kleene star, alternation and concatenation, or functional composition of
+lenses.  Figure~\ref{fig:example-lens} illustrates the use of these
+combinators to transform \bibtex{} to EndNote (and vice versa).
+\begin{figure}[t]
+\begin{lstlisting}
+let preamble : (lens in "@book{" $\Leftrightarrow$ "%0 Book\n%F ") =
+del "@book{" . ins "%0 Book\n%F "
+
+let author_lens : (lens in (",\n" . bib_author) $\Leftrightarrow$ ("\n" . end_author)) =
+del ",\nauthor = \""
+. ins "\n%A "
+. name
+. (del " and "
+. ins "\n"
+. (ins "%A " . name . del " and " . ins "\n")*
+. ins "%A "
+. name
+|| "")
+
+let title_lens : (lens in ("\",\n" . bib_title . "},\n}") $\Leftrightarrow$ ("\n" . end_title)) =
+del "\",\ntitle = {" . ins "\n%T " . title . del "},\n}"
+
+let bib_to_end : (lens in bibtex $\Leftrightarrow$ end_note) = preamble . label . author_lens . title_lens
+\end{lstlisting}
+\caption{An example lens between \cd{bibtex} and \cd{end_note} regular
+expressions with no equivalences.\bcp{And a readability suggestion:
+define a named constant for a quote character.}}
+\label{fig:example-lens}
+\end{figure}
+
+Unfortunately, programming bijective lenses in Boomerang is often slow and
+tediuos because of the fiddly constraints that Boomerang imposes on its
+terms. making lens programming slow and tedious. For example, Boomerang
+programmers often must rearrange the order of data items by recursively using
+operators that swap adjacent fields. Furthermore, the Boomerang type checker is
+very strict, disallowing many programs because they contain ambiguity about how
+certain data is transformed.
+
+In our previous work in~\cite{optician}, we presented a solution to these
+shortcomings of bijective lenses by developing a tool called {\em Optician}
+which enables a programmer to synthesize a bijective lens from a pair of regular
+expressions $(S, T)$ and a set of input-output example \cd{exs} pairs that
+determines how the synthesized lens should behave on those examples. With {\em
+Optician}, the programmer need only spend the effort of specifying the data
+formats and the behaviour of the lens on a few inputs in order to define a
+bijective lens. As a result defining bijective lenses becomes faster and less
+error prone, since the programmer need only invoke the synthesis procedure
+\lstinline{synth S <=> T using exs} in order to define a bijective lens rather
+than manually derive the lens herself.
+
+\subsection{Quotient Lenses}
+Unfortunately, while our synthesis tool greatly simplifies the task of
+programming bijective lenses, not all bidirectional tranformations are bijective
+in nature. For instance, users are typically uninterested in preserving
+whitespace between words or between fields.  The order of author and title
+fields is also likely irrelevant and there may be equivalent ways of writing
+the same name:  ``Doe, John'' vs ``John Doe.''
 
@@
 (i.e., what lenses would look like if they had to preserve all this stuff?)}
-to treat them as identical and to map them 
-into canonical representations.
-Continuing our example, the following two
-\bibtex{} citations represent the same logical object and as a result, should
-be translated to the same EndNote record.
+to treat them as identical and to map them into canonical representations.
+Continuing our example, the following two \bibtex{} citations represent the same
+logical object and as a result, should be translated to the same EndNote record.
 \begin{verbatim}
-@Book {conway,
-Author = "J. H. Conway",
-Title = {Regular Algebra and Finite Machines},
+@Book {Doe,
+Author = "John Doe",
+Title = {Generic Title},
 }
 
-@Book{     conway,
-Title = {Regular Algebra
-and Finite Machines},
-Author = "Conway, J. H.",     }
+@Book{     Doe,
+Title = {Generic
+Title},
+Author = "Doe, John",     }
 \end{verbatim}
@@
 One natural way to define lenses that will synchronize data in these formats
-is to provide the following items: \dpw{this is slightly redundant from the intro.  It
-adds a little more info, but not much.  Should this be cut?  Should the intro be expanded
-slightly?  Should it be left as is (more or less)?} \saz{I think that this is
-not too redundant and that it can be left as-is, modulo space crunching.}
-
+is to provide the following items:
 \begin{enumerate}
   \item A pair of descriptions of the external source and target formats, here
-    \bibtex{} and EndNote.
-  Such descriptions may be given by regular expressions, for instance.
-  They serve as documentation to describe the domains over which the lens
-  is to operate.\bcp{This way of saying it implies that they serve {\em
-      only} as documentation.}
+  \bibtex{} and EndNote. Such descriptions may be given by regular
+  expressions, for instance.
   \item A pair of \emph{canonizers} that map external source (and target)
@@
   \item A pair of descriptions of the internal source and
-  target formats, again, to document the structure of the
-  transformation.\bcp{Same question.}
+  target formats.
   \item A bijective lens to map back and forth between canonical elements
@@
 Indeed, Boomerang's quotient lenses are structured this
-way~\cite{quotientlenses}.
-Of course, some of these elements are optional.  For instance, it is
-not necessary to document internal and external formats.  However, we do
-view it as good engineering practice to do so, as such documentation makes
-lenses easier to read, understand, use and maintain.\bcp{Is it really true
-  that these types are optional in Boomerang?}  In this case, there is 
-significant redundancy here:  External and internal formats
-typically have some overlap; external formats and canonizers have related
-structure; and the structure of bijective lens computations is largely driven
-by the structure of the internal formats.\bcp{Would it be worth
-  demonstrating this concretely?}  Hence, overall, a more compact
-notation would simplify description and legibility of such computations.
-In addition,
-the lens and canonizer languages themselves can be somewhat fiddly and
-difficult to
-master.  For instance, to quotient out whitespace inside, and between,
-fields while also ignoring field order, one must nest canonizing combinators
-in a somewhat tricky way.
-
-%% %Foster's Boomerang~\ref{boomerang} language presents one design
-%% While it is possible to express the equivalence on the \bibtex{} data above in
-%% Boomerang, actually doing so is a challenge.  Three of the inessential details
-%% we wish to ignore---whitespace quotienting
-%% inside fields, whitespace quotienting in between fields, and permutation
-%% quotenting of fields---nest within each other in a way that makes them difficult
-%% to express with the existing Boomerang combinators. Furthermore, if the
-%% programmer wishes to specify the type of the lens, then she can only specify
-%% data formats on each side of the lens with a regular expression that does not
-%% keep track of the equivalence relation defined on the language of the regular
-%% expression. Consequently, the programmer has to write a regular expression as
-%% well as a canonizer that expresses the equivalence relation on that regular
-%% expression, even though a regular expression and the canonizer defined on it
-%% typically share much of the syntactic structure. Having to write both of these
-%% expressions makes programming with quotient lenses verbose and time consuming.
-
-We propose to do away with much of this overhead, and the burden of a
-fiddly programming process, by defining a single new compact notation
-to specify data formats modulo an equivalence relation and then by
-synthesizing the lenses that map between these formats automatically.
-\bcp{If our main contribution is ``eliminating fiddliness,'' we should be
-  more explicit about what the fiddliness is.  (But, again, I would be even
-  happier if we could phrase everything more positively... ``Quotient lenses
-  make lens programming more realistic.  But here are some opportunities for
-  eliminating redundancy...'')}
-
-\subsection{Quotient Regular Expressions}
+way~\cite{quotientlenses}. Of course, some of these elements are optional.  For
+instance, it is not necessary to document internal and external formats.
+However, we do view it as good engineering practice to do so, as such
+documentation makes lenses easier to read, understand, use and maintain.
+
+The first way in which our work improves upon Boomerang's quotient lenses is
+that we merge the first three pieces of quotient lenses that we just described
+into a single concept, {\em Quotient Regular Expression}, or QREs. A single QRE
+enables a programmer to add annotations to the outer regular expression; these
+annotations can then be used to infer an inner regular expression, as well as a
+canonizer from the outer regular expression to the inner one. In the following
+subsection, we describe the QRE combinators and use give an example of how a
+programmer might define a QRE specification of \bibtex.
+
+\subsection{Specifying \bibtex{} Using QREs}
 \label{subsec:qre-expressions}
-
-A {\em Quotient Regular Expression} (QRE) defines an external regular
-expression $R$, an equivalence relation $\sim$ on the language described by
-$R$, and a canonical set of representatives of $\sim$, all at once.  QREs
-are constructed from a simple set of primitives and composed using familiar
-regular operators (`\cd{|}' for choice, `\cd{.}' for concatenation, and
-`$^*$' for repetition).  Users\bcp{This sentence feels clunky because it
-  ``uses up'' its head/subject position, and needlessly focuses the reader's
-  attention, on something (``users'') that is not especially important or
-  germane to the conversation.  (There are lots of other sentences elsewhere
-  in the paper that do the same thing. :-) Please rewrite (all of them).}
-may also treat QREs as pairs of get and put functions and compose such pairs
-(the functional composition operator `;').\saz{I agree with Benjamin about
-  useless subjects, but here the sentence could be deleted entirely.}
-
-For example, to begin construction of our \bibtex{} format, we might
-first define a whitespace format, which, externally, matches
-any non-zero number of whitespace characters, and converts
-any such whitespace into a single space character (its canonical form).
-This whitespace-normalizing QRE is easily defined by using the \kw{project} primitive: \bcp{Why is it called ``project''?}
+To construct a QRE specification of a \bibtex{} format, we might first define a
+whitespace format, which, externally, matches any non-zero number of whitespace
+characters, and converts any such whitespace into a single space character (its
+canonical form). This whitespace-normalizing QRE is easily defined by using the
+\kw{project} primitive: \bcp{Why is it called ``project''?}
 
@@
 For instance, assume that
-the \codefont{comma\_name} format describes ``Conway, J. H.''
-and that the \codefont{space\_name} format describes ``J. H. Conway''
+the \codefont{comma\_name} format describes ``Doe, John''
+and that the \codefont{space\_name} format describes ``John Doe''
 and \codefont{s\_to\_c} is a function from the first to the second.  In this case,
@@
 %
-\bcp{Four problems here.  The first is that the notion of synthesizing lenses
-comes out of nowhere (of course, it is in the background, but we're using it
-here as if it had already been introduced concretely---this will be helped
-by adding the new subsection I've suggested at the top of this section).
-The second is that the idea of using half of a lens 
-as a canonizer comes out of nowhere.  The third is that this idea is a hack (and
-always was).  Why bother defining / synthesizing a lens and then using only
-half of it?  And the fourth is that, in the original quotient lens paper,
+\bcp{Three problems here. First, the idea of using half of a lens
+as a canonizer comes out of nowhere.  The second is that this idea is a hack
+(and always was).  Why bother defining / synthesizing a lens and then using only
+half of it?  And the third is that, in the original quotient lens paper,
 we used this trick on lenses that were not even bijective, which may confuse
-people familiar with the earlier paper.}The first line above synthesizes a
-\emph{bidirectional}\bcp{Why is this italicized?  Are some lenses not
-  bidirectional?} lens between 
-\codefont{comma\_name} and \codefont{space\_name} using the listed example
-transformation as a guide.  The second line simply
+people familiar with the earlier paper.}
+
+\sam{Here is my response to this. You observe that the idea of using half a lens
+as a canonizer comes out of nowhere, and that this ideas is a hack and always
+was. I don't think that's quite right. I think that it is useful to have a
+combinator which chooses between two or more formats. Indeed, the permutation
+combinator is just a special case of this. The bibtex name field is a good
+example of where such a combinator would be useful.  But if we choose to have
+such a combinator, then we need a way of mapping one format to another, and
+really the only interesting string-to-string transformations in Boomerang are
+given by lenses, which is why we use the get function of a lens in a canonizer.
+We could have used any other function between the two formats but Boomerang
+doesn't give us too many options for writing unidirectional transformations. So
+I don't think that using a lens as a canonizer is a trick. I think that
+Boomerang just isn't designed for unidirectional tranformations. But I also
+agree that it doesn't make much sense to synthesize a lens only to get rid of
+one half of it.
+I'm not sure what the solution is to be honest.}
+
+The first line above synthesizes a
+lens between \codefont{comma\_name} and \codefont{space\_name} using the listed
+example transformation as a guide.  The second line simply
 extracts the get direction transformation from the lens,
@@
 relation by a composition of a \kw{squash} QRE and a QRE that projects away
-the extra whitespace in between \codefont{first\_name} and \codefont{last\_name}
-
-\begin{figure}[t]
-\begin{lstlisting}
-(* whitespace with " " as a canonical representative *)
-let wsp = project [ \n\t\r]$^+$ $\to$ " "
-
-(* define name representations with a space and with a comma *)
-let space_name = name | (name . wsp . (name . wsp)$^*$ . name)
-let comma_name = name | (name . "," . wsp . (name . wsp)$^*$ . name)
-
-(* synthesize a lens that maps the comma representation to the space representation *)
-let l = synth comma_name $\Leftrightarrow$ space_name using {("Conway, J. H.", "J. H. Conway")}
-let s_to_c = get l
-
-(* squash QRE maps comma_name to space_name *)
-let name = squash comma_name $\to$ space_name using s_to_c
-
-(* define rest of bibtex fields *)
-let name_sep = wsp . "and" . wsp
-let bib_names = name | (name . name_sep . (name . name_sep)$^*$ . name)
-let bib_author = "author = \"" . bib_names . "\""
-let title = word | (word . wsp . (word . wsp)$^*$ . word)
-let bib_title = "title = {" . title . "}"
-
-(* allow any permutation of fields interspersed with arbitrary whitespace *)
-let bib_fields = perm (label, bib_author, bib_title) with (project ("," .wsp$^+$) $\to$ ",\n")
-let bibtex = "@book{" . bib_fields . "}"
-\end{lstlisting}
-
-\caption{QRE specification of \bibtex{} records. \bcp{The spacing of the
-    comments is weird...}}
-\label{fig:example-qre}
-\end{figure}
+the extra whitespace in between \codefont{first\_name} and
+\codefont{last\_name}.
 
@@
 %
-To normalize the field separators, one can specify that the
-components of the permutation are conjoined by another QRE. For instance, below, we normalize whitespace
-between fields, leaving only a single newline.
+To normalize the field separators, one can specify that the components of the
+permutation are conjoined by another QRE. For instance, below, we normalize
+whitespace between fields, leaving only a single newline.
 %
@@
 
-% The final QRE combinator is the \kw{normalize} combinator.  This combinator is an
-% ``escape hatch'' that allows a programmer to manually define a function $f$
-% that sends each string matching a regular expression $R$ to some canonical
-% representative in another regular expression $R'$. The equivalence relation
-% defined by the normalize combinator is hence the equivalence relation defined by
-% the {\em fibres} of $f$; that is, for all strings $s$ and $s'$ that match $R$,
-% $s$ is equivalent to $s'$ if and only if $f(s) = f(s')$.
-
-% For instance, assume that $f(s)=$ `` '' for all whitespace strings $s$. Then the
-% project QRE \kw{p\_wsp} which matches any non-zero number of whitespace
-% characters and converts any such whitespace into a single space character could
-% be expressed using the \kw{normalize} combinator as
-% \begin{lstlisting}
-% let p_wsp = normalize (wsp+, " ", f)
-% \end{lstlisting}
-% However, although it is in some sense ``universal'' (see the discussion below
-% \saz{TODO}))the \kw{normalize} QRE imposes significant requirements on the
-% canonization function $f$, so it is usually simpler to use the other QRE
-% constructs where possible.
-
-A QRE for the \bibtex{} format is presented in its
-entirety in Figure~\ref{fig:example-qre}. \bcp{Walk through it a bit?}\saz{All of the
-  examples in this section are walking through the figure, we just have to say so.}
-
-\subsection{QRE Lenses}
-QREs allow users to specify regular languages modulo equivalence relations,
-a first step in writing many data transformations.  However, writing the
-inner bijective lens itself still takes a fair amount of work\bcp{Why do we
-  want to emphasize this here?}.  Each such lens
-is constructed from simple primitive combinators including:
-%
-\begin{itemize}
-  \item \kw{del} $s$: \textit{delete} the constant string $s$ in the get direction;
-  insert it in the put direction.
-  \item \kw{ins} $s$: \textit{insert} the constant string $s$ in the get direction;
-  delete it in the put direction.
-  \item \kw{copy} $R$: copy the text matching the regular expression $R$ in both
-  directions.
-\end{itemize}
-In addition, such primitives may be combined using the regular operators
-Kleene star, alternation and concatenation, or functional composition of
-lenses.  Figure~\ref{fig:example-lens} illustrates the use of these
-combinators to transform \bibtex{} to EndNote (and vice versa).
-There is a fair amount of trickiness in writing these transformations, because
-the programmer must keep mental track of what they have parsed so far \emph{in both
-directions}, carefully inserting and deleting the relevant syntax on both
-sides.
-
-%% While QREs make it easier to write equivalence relations on regular expressions
-%% by specifying the data and the equivalence relation on the data in one
-%% expression, in order to define a quotient lens, the programmer still has to
-%% manually derive the underlying lens that maps the equivalence classes of the
-%% source data to the equivalence classes of the target data. This is itself a
-%% tricky task as Figure~\ref{fig:example-lens} illustrates.
-
-\begin{figure}[t]
+The final QRE combinator is the normalize combinator.
+This combinator allows a programmer to manually define a function
+$f$ which sends each string that matches a regular expression $R$ to some
+canonical representative in another regular expression $R'$. The equivalence
+relation defined by the normalize combinator is hence the equivalence relation
+defined by the {\em fibres} of $f$; that is, for all strings $s$ and $s'$ that
+match $R$, $s$ is equivalent to $s'$ if and only if $f(s) = f(s')$.
+
+For instance, assume that $f(s)=$ `` '' for all whitespace strings $s$. Then the
+project QRE \kw{p\_wsp} which matches any non-zero number of whitespace
+characters and converts any such whitespace into a single space character can
+be expressed using the normalize combinator as 
+\lstinline{normalize (wsp+, " ", f)}.
+
+Figure~\ref{fig:example-qre} gives the complete QRE specification for the
+\bibtex{}.
+\begin{figure}
 \begin{lstlisting}
-let preamble : (lens in "@book{" $\Leftrightarrow$ "%0 Book\n%F ") =
-    del "@book{" . ins "%0 Book\n%F "
-
-let author_lens : (lens in (",\n" . bib_author) $\Leftrightarrow$ ("\n" . end_author)) =
-    del ",\nauthor = \""
-        . ins "\n%A "
-        . valid_name
-        . (del " and "
-        . ins "\n"
-        . (ins "%A " . valid_name . del " and " . ins "\n")* 
-        . ins "%A "
-        . valid_name
-        || "")
-
-let title_lens : (lens in ("\",\n" . bib_title . "},\n}") $\Leftrightarrow$ ("\n" . end_title)) =
-    del "\",\ntitle = {" . ins "\n%T " . title . del "},\n}"
-
-let bib_to_end : (lens in bibtex $\Leftrightarrow$ end_note) =
-    preamble . label . author_lens . title_lens
+let wsp = [ \n\t\r]
+
+(* define name representations with a space and with a comma *)
+let space_name = first_name . wsp+ . last_name
+let comma_name = last_name . "," . wsp+ . first_name
+
+(* synthesize a lens that maps comma representation to space representation *)
+let l = synth comma_name $\Leftrightarrow$ space_name using {("Doe, John", "John Doe")}
+let s_to_c = get l
+
+(* squash QRE maps comma_name to space_name *)
+let comma_to_space = squash comma_name $\to$ space_name using (get comma_to_space)
+
+(* map arbitrary whitespace into " " as a canonical representative *)
+let p_wsp = project wsp+ $\to$ " "
+
+(* Project away whitespace in space_name representation *)
+let project_in_space = first_name . p_wsp . last_name
+
+(* Compose the two QREs to get a canonical representation for a name *)
+let name = project_in_space ; comma_to_space
+
+(* define rest of bibtex fields *)
+let name_sep = p_wsp . "and" . p_wsp
+let bib_names = name | (name . name_sep . (name . name_sep)$^*$ . name)
+let bib_author = "author = \"" . bib_names . "\""
+let title = word | (word . p_wsp . (word . p_wsp)$^*$ . word)
+let bib_title = "title = {" . title . "}"
+
+(* allow any permutation of fields interspersed with arbitrary whitespace *)
+let bib_fields = perm (label, bib_author, bib_title) with (project (wsp+ . "," . wsp+) $\to$ ",\n")
+let bibtex = "@book{" . bib_fields . "}"
 \end{lstlisting}
-\caption{An example lens between \cd{bibtex} and \cd{end_note} regular
-expressions with no equivalences.\bcp{There was
-a suggestion to colorize the new keywords that we're introducing here
-({\em including} symbolic ones, I guess)}\bcp{And a readability suggestion:
-define a named constant for a quote character.}}
-\label{fig:example-lens}
+\caption{QRE specification of \bibtex{} records. }
+\label{fig:example-qre}
 \end{figure}
-
-%% The lens shown in Figure~\ref{fig:example-lens} transforms \bibtex{} data to
-%% EndNote data  (and vice versa) by matching the delimiters and separators in a
-%% bijective manner. For example, the \cd{preamble} lens replaces the \bibtex{}
-%% preamble \verb|@Book| with the EndNote preamble \verb|%0 Book\n%F|. When the
-%% programmer is writing the lens, she has to make sure that all of these small
-%% deletions and insertions are correct, which can be a challenging task if the
-%% lens is large enough, as in the definition of the lens \verb!author_lens! in
-%% Figure~\ref{fig:example-lens}. This makes lens programming tedious and
-%% error-prone.
-
-\bcp{How about a subsection break here, to introduce the idea of QRE
-  lens synthesis?}
-
-To make this unsavory programming task more palatable, we identify a class of
-quotient lenses that we call {\em QRE lenses} and show how to {\em synthesize}
-them from a pair of QREs and a set of input-output examples.\bcp{that's a pretty weak
-  way of describing our main contribution!} Our key
-insight is to reduce the problem of synthesizing QRE lenses to the problem of
-synthesizing {\em bijective lenses}~\cite{optician}, a simpler subset of
-lenses that support strict bijections, rather than bijections modulo
-equivalence relations.
-
-Intuitively, our QRE lenses are bijective lenses with ``canonizers at the
-edges''---Figure~\ref{fig:attheedges} presents the architecture of each lens.
-Every QRE lens $q$ has a type $q: c \Leftrightarrow c'$ where $c$ and $c'$ are QREs.
-In the get direction, a QRE lens $q: c \Leftrightarrow c'$ uses the source
-QRE $c$ to compute a unique representative for the data modulo the equivalence
-relation defined by $c$ and then applies the $\get$ function of a bijective
-lens $\ell$ to this representative. In the backward direction, $q$ operates
-similarly,  but using the view QRE $c'$ and the $\lput$ function of $\ell$.
-
-Because the QREs $c$ and $c'$ determine the internal formats for data after
-canonization, and because the algorithm for synthesizing bijective lenses is
-directed by these formats, $c$ and $c'$ are all that is required to synthesize
-QRE lenses end-to-end.  Indeed, we structured QRE lenses as bijective lenses with
-canonizers at the edges, precisely to facilitate the use of this synthesis algorithm.
-
+\subsection{QRE Lenses and QRE Lens Synthesis}
+At this point we already have a tool for synthesizing bijective lenses from a
+pair of regular expressions and a set of example input-output pairs, as well as
+the language of QREs for defining equivalence relations on regular expressions.
+An obvious next step in simplifying the quotient lens programming is thus to use
+the bijective lens synthesis procedure as a subroutine for a quotient lens
+synthesis procedure. This new procedure should take two QREs and a set
+of example input-output pairs as input, compute the canonical data formats on
+both sides of the lens from the pair of QREs, map the example input-output pairs
+to their canonical representatives to get a set of examples in canonical form,
+and then invoke the bijective lens synthesis procedure on the canonical data
+formats and the new set of examples that are in canonical form.
+
+Indeed this basic idea is what motivates our definition of {\em QRE Lenses}.
+Intuitively, just like Boomerang's quotient lenses, our QRE lenses are bijective
+lenses with ``canonizers at the edges''---Figure~\ref{fig:attheedges} presents
+the architecture of each lens.
 \begin{figure}[t]
@@
 choose function.}\bcp{Done.  But I wonder if the picture should mention QREs
-explicitly, somehow... any ideas?}}
+explicitly, somehow... any ideas?}}\sam{Maybe we can put all structures, the
+canonizer and the canonical strcutre in a box that says QRE?}
 \label{fig:attheedges}
 \end{figure}
-
-%% One potential pitfall with QRE lenses is that the combinators that we define on
-%% them may not preserve this normal form. For example, the composition $q'; q$ of
-%% QRE lenses $q$ and $q'$ might not be equivalent to any quotient lens that has
-%% the canonical normal form of a bijective lens with canonizers at the edges.
+Every QRE lens $q$ has a type $q: c_1 \Leftrightarrow c_2$ where $c_1$ and $c_2$
+are QREs. In the get direction, a QRE lens $q: c_1 \Leftrightarrow c_2$ uses the
+source QRE $c_1$ to compute a unique representative for the data modulo the
+equivalence relation defined by $c_1$ and then applies the $\get$ function of a
+bijective lens $\ell$ to this representative. In the backward direction, $q$
+operates similarly, but using the view QRE $c_2$ and the $\lput$ function of
+$\ell$.
+
+Because the QREs $c_1$ and $c_2$ determine the internal formats for data after
+canonization, and because the algorithm for synthesizing bijective lenses is
+directed by these formats, $c_1$ and $c_1$ are all that is required to
+synthesize QRE lenses end-to-end. Indeed, we structured QRE lenses as bijective
+lenses with canonizers at the edges, precisely to facilitate the use of this
+synthesis algorithm.
 
 On the other hand, our restriction that canonizers appear only at the edges
-raises a key technical question:  Are we limiting the expressiveness of our transformations
-by demanding all programs fit into this normal form?\bcp{I wish we could
-  have previewed this issue a bit more in the introduction, since it is our
-  main contribution.}  It turns out that we are not:
-any lens that uses canonizers internally can be transformed into a lens that uses canonizers
-at the edges.  The main technical contribution of this paper is a proof of this fact.  The
-crucial step in the proof is a demonstration that the functional composition
-of two QRE lenses with ``all canonizers at the edge'' can be represented as some
-single QRE lens with ``canonizers at the edge.''  \bcp{Saying ``all
-  canonizers at the edge'' seems confusing: there is no other place for them
-to go, synmtactically.  And: Could we draw a picture of this transformation?}
-
-%% For example, suppose \cd{q_1 : bibtex} $\Leftrightarrow$ \cd{endnote} is a QRE
-%% lens that maps between \bibtex{} and EndNote formats, and that \cd{q_2 : endnote
-%% <=> xml} is a QRE lens that maps between EndNote and XML formats. The
-%% composition \cd{(q_1 ; q2) : bibtex} $\Leftrightarrow$ \cd{xml} of \cd{q_1} and
-%% \cd{q_2} is not necessarily in normal form due to the presence of canonizers
-%% on each side of the intermediary EndNote format:
-
-%% \sam{TODO: We need to put a diagram of composition here.}
-
-%% \noindent However, while \cd{q_1 ; q_2} is not in normal form, our main theorem
-%% shows that there is another QRE lens \cd{q : bibtex} $\Leftrightarrow$ \cd{xml}
-%% of the same type as \cd{(q_1 ; q2) : bibtex} $\Leftrightarrow$ \cd{xml} such
-%% that \cd{q_1 ; q_2} and \cd{q} define the same transformation.
-
-This technical investigation justifies replacing manually constructed lenses that may use
-functional composition and nested canonizers\bcp{nested canonizers have not
-  been mentioned!} with synthesized lenses in normal form.
-The end result is that users can synthesize complex transformations with a
-single line of code.\bcp{Well, the QRE definitions and examples may well use
-multiple lines all by themselves!}  For instance, all the code in Figure~\ref{fig:example-lens} may be replaced by
-a single call to the synthesis prodedure:
+raises a key technical question:  Are we limiting the expressiveness of our
+transformations by demanding all programs fit into this normal form?\bcp{I wish
+we could have previewed this issue a bit more in the introduction, since it is
+our main contribution.}  It turns out that we are not: any lens that uses
+canonizers internally can be transformed into a lens that uses canonizers at
+the edges. The main technical contribution of this paper is a proof of this
+fact.
+
+This technical investigation justifies replacing manually constructed lenses
+using the QRE lens combinators with synthesized QRE lenses in normal form, thus
+saving the programmer. For instance, after defining the \bibtex{} and EndNote
+QREs and storing them in the variables \cd{bibtex} and \cd{endnote}
+respectively, then all the code in Figure~\ref{fig:example-lens} may be
+replaced by a single call to the synthesis prodedure:
 
@@
 let bib_to_end : (lens in bibtex $\Leftrightarrow$ endnote) =
-  synth bibtex $\Leftrightarrow$ endnote
-  using {(bib_example, end_example)}
+synth bibtex $\Leftrightarrow$ endnote using {(bib_example, end_example)}
 \end{lstlisting}
@@
 \noindent Here, the generated  quotient lens synchronizes \cd{bibtex} and
-\cd{endnote} formats.  It uses \cd{bib_example} and \cd{end_example} (the two concrete example
-strings given at the beginning of this section) to disambiguate.
-In addition, and as we saw earlier with the definition of \cd{s_to_c}, the synthesis procedure itself
-can be used to create lenses that are in turn used to define other QREs.  The ability to intermix
-QRE specification with QRE lens synthesis yields a powerful and flexible way of
+\cd{endnote} formats, using \cd{bib_example} and \cd{end_example} (the two
+concrete example strings given at the beginning of this section) to
+disambiguate. In addition, and as we saw earlier with the definition of
+\cd{s_to_c}, the synthesis procedure itself can be used to create lenses that
+are in turn used to define other QREs.  The ability to intermix QRE
+specification with QRE lens synthesis yields a powerful and flexible way of
 creating bidirectional transformations.
@@
 \label{QRE}
-
-Quotient Regular Expressions allow programs\bcp{Why bring ``programs'' into
-  it?} to specify a regular language,
-an equivalence relation over elements of the language and a function to convert
-elements of the language into normal form.
-
-\subsection{Syntax}
-
-The language of Quotient Regular Expressions (QREs) is given by the grammar
+\subsection{Syntax and Semantics of QREs}
+Formally, the language of Quotient Regular Expressions (QREs) is given by the
+following grammar:
 \begin{align*}
@@
 \perm{c_1, \ldots, c_n}{c} \\
-& | \; \normalize{R_1}{R_2}{f} \sep c_1 \semicolon c_2 \sep c_1 \cdot c_2 \sep (c_1 \sep
-c_2) \sep c^*
+& | \; \normalize{R}{R'}{f} \sep c_1 \semicolon c_2 \sep c_1 \cdot c_2 \sep (c_1
+\sep c_2) \sep c^*,
 \end{align*}
@@
 functions between regular languages, and $s$ ranges over character strings.
-$\mathcal{L}(R)$ is
-the language accepted by the regular expression $R$.
-
-%Section~\ref{subsec:qre-expressions} presented a number of examples
-
-\subsection{Intuitions}
-In the \bibtex{} to EndNote example, we encountered lenses of the form $\project{R}{s}$,
-
-\noindent $\squash{c}{c'}{f}$ and $\perm{c_1, \ldots, c_n}{c}$ QREs, as well as
-the regular operators applied to these, as in\bcp{bad linebreak: rewrite}
-
-\begin{lstlisting}
-let space_name  = name | (name . wsp_to_space . (name . wsp_to_space)$^*$ . name)
-\end{lstlisting}
-
-One combinator that we haven't yet introduced\bcp{haven't we? we've
-  mentioned it, at least.} is sequential composition.
-The $c \semicolon c'$ QRE successively applies
-two equivalence relations to a regular language. For example, suppose that
-we wish to express a name as $\kw{last\_name \cdot ``,\text{''} \cdot
-\text{wsp+} \cdot first\_name}$, or as $\kw{first\_name\cdot \text{wsp+} \cdot
-last\_name}$, with $\kw{first\_name\cdot \text{wsp+} \cdot last\_name}$ as the
-canonical representation. We also wish to project away the extra whitespace in
-between \kw{first\_name} and \kw{last\_name} once this identification has been
-made. Then we can express this equivalence relation by a composition of a
-\kw{squash} QRE and a QRE that projects away the extra whitespace in between
-\kw{first\_name} and \kw{last\_name}: \bcp{It would be easier to distinguish
-code examples from text if they were indented a bit.}
-
-\begin{lstlisting}
-(* Regular expressions matching different comma and space representations *)
-let comma_name = last_name . "," . wsp . first_name
-let space_name = first_name . wsp . last_name
-
-(* Squash comma representation into space representation *)
-let comma_to_space =
-synth comma_name <=> space_name using {("Conway, J. H.", "J. H. Conway")}
-let comma_to_space = squash comma_name $\to$ space_name using (get comma_to_space)
-
-(* Project away whitespace in space_name representation *)
-let project_in_space = first_name . (project wsp+ $\to$ " ") . last_name
-
-(* Compose the two QREs *)
-let valid_name = project_in_space ; comma_to_space
-\end{lstlisting}
-
-The final combinator is $\normalize{R}{R'}{f}$, which expresses
-the equivalence relation that sends each element in $\mathcal{L}(R)$ to its
-canonical representation $f(r) \in \mathcal{L}(R')$. The equivalence relation
-defined by $\normalize{R}{R'}{f}$ is thus the equivalence relation on
-$\mathcal{L}(R)$ defined by the ``{fibres}'' of $f$---i.e., $r \sim r'$ if and
-only if $f(r) = f(r')$. The canonizing function $f$ is required to be
-surjective and idempotent.
-
-The $\normalize{R}{R'}{f}$ QRE is the most general QRE in that, for any QRE
-$q$, there exist regular expressions $R$ and $R'$ and a surjective, idempotent
-function $f:\mathcal{L}(R) \longrightarrow \mathcal{L}(R')$ such that $q$ and
-$\normalize{R}{R'}{f}$ each define the same equivalence relation on
-$\mathcal{L}(R)$, with $\mathcal{L}(R')$ forming a complete set of
-representatives for this equivalence relation. The $\normalize{R}{R'}{f}$
-combinator is included so as to express equivalence relations that are either
-difficult to define, or that cannot be defined using the other combinators.
-\saz{Add something about being ``complete'' here?}\sam{I'm not sure how to do
-this here. The precise statement about how normalize is strong enough to
-express all QREs requires that we introduce W(c), K(c) and Canonizer(c) which
-we haven't introduced at this point in the paper}\bcp{But we can at least
-state an informal claim and give a forward pointer to where it is made more
-precise later.  Or, since these definitions appear quite soon, perhaps we
-can just reorganize the material a bit so that the definitions come before
-the intuition plus a precise claim.}
-
-Now in order for the programmer to ensure that the {\em normalize} QRE is well
-formed, she must define a canonizing function $f$ between two regular
-languages and prove that $f$ is surjective and idempotent: the compiler cannot
-verify this since this problem is undecidable in general. Consequently, we do
-not use the $\normalize{R}{R'}{f}$ QRE in practice. On the other hand, the
-$\normalize{R}{R'}{f}$ combinator gives a sufficient condition for potential
-QREs to be ``well behaved,'' particularly with respect to the strategy for
-synthesizing quotient lenses that we will introduce in Section~\ref{synth}.
-
-The equivalence relations definable by QREs are a strict subset of those that
-can be expressed in Boomerang. For instance, in Boomerang, one can use any lens
-to canonize a regular language described by $R$ to a regular language described
-by $S$, with $\mathcal{L}(S) \not \subseteq \mathcal{L}(R)$, whereas each QRE
-$c$ canonizes a regular language described by $R$ to regular subset
-$\mathcal{L}(R') \subseteq \mathcal{L}(R)$ described by $R'$. QREs therefore are
-not as expressive as Boomerang canonizers.\bcp{Not sure I understand what
-  this means, practically.  First, is it true that QREs can express any
-  boomerang canonizer that maps from a language to a subset of the same
-  language?  And second, can't at least some Boomerang canonizers be expressed as the
-  composition of a QRE canonizer from R to R' plus a lens from R' to S?} On
-the other hand, QREs enable us to 
-express many equivalences that occur in real world data as we shall demonstrate
-in the evaluation section (Section~\ref{impl}).
-
-The net effect of these design decisions means that each QRE $c$ enables
-us to simultaneously express
+
+Using the conventional notation that $\mathcal{L}(R)$ is the language accepted
+by the regular expression $R$, then each QRE $c$ enables us to express
 \begin{enumerate}
@@
 \end{enumerate}
-Observe that, since $\canonizer(c)$ computes $\canonizer(c)(w)$ as the unique
-$k$ in $\mathcal{L}(K(c))$ such that $w$ is equivalent to $k$ mod $\eqrel{c}$,
-it follows that the equivalence classes of $\eqrel{c}$ are the same as the
-fibres\bcp{this is not a very familiar term for lots of people; if we're
-  going to use it a lot, we should define it more carefully} of
-$\canonizer(c)$.
-
-\subsection{Well Formed QREs and ambiguity}
-Before giving the formal semantics for QREs, we address the issue of {\em
-ambiguity}, which arises when defining QREs. If $c$ and  $c'$ are QREs, then
-when applying the regular combinators to QREs, we require that if a string $s$
-matches any of the regular expressions\bcp{The following list is hard to
-  parse.  A general rule is: Always separate adjacent math expressions by
-  something more than just punctuation.} $W(c) \cdot W(c')$, $W(c) \sep W(c')$,
-$W(c)^*$, $K(c) \cdot K(c')$, $K(c) \sep K(c')$, $K(c)^*$, then $s$ matches
-that regular expression in only one way.
-
-This  condition is called
-{\em strong unambiguity}~\cite{Sippu1988}, and it is necessary,
-firstly,\bcp{is there a ``secondly'' coming?  I don't see it} because it
-ensures that the canonizing function of a QRE is well-defined.
-%
-For example consider the QRE\bcp{Displayed math should be punctuated either
-  (1) exactly as if it were inline (i.e., this would need a period at the
-  end) or (2) with a sentence terminator (a colon or period) beforehand.  In
-this case, (1) would be confusing (since . is also part of QRE syntax), so
-(2) is preferable.}
-
-\begin{lstlisting}
-let ambiguous : canonizer = "a"* . (project "a"* $\to$ "a")
-\end{lstlisting}
-
-\noindent The behaviour of \cd{ambiguous} is not-well defined since the string
-\cd{"aaa"} can be canonized to any of \cd{"a"}, \cd{"aa"}, or \cd{"aaa"} depending on how
-\cd{"aaa"} is parsed.
-
-We also require that the regular combinators applied to
-the kernels of QREs be unambiguous, since the underlying\bcp{?} lenses will
-end up operating on these kernels, and these lenses impose the same
-restrictions for similar reasons.
-%
-To this end\bcp{which end?}, we say that regular expressions $R$ and $S$ are
-\textit{unambiguosly concatenable}, written $R \cdot^! S$, if, for all strings
-$r, r' \in \mathcal{L}(R)$ and $s, s' \in \mathcal{L}(S)$, if $r \cdot s = r'
-\cdot s'$, then $r = r'$ and $s = s'$. We say that a regular expression $R$ is
-\textit{unambiguosly iterable}, written $R^{*!}$ if, for all strings $r_1,
-\ldots, r_m$ and $r'_1, \ldots, r'_n \in \mathcal{L}(R)$, if $r_1 \cdot \ldots
-\cdot r_m = r'_1 \cdot \ldots \cdot r'_n$, then $m = n$ and $r_i = r'_i$.
-
-We say that a regular expression $R$ is \textit{strongly unambiguous} when
-(1) $R = \varnothing$, or (2) $R = S_1 \cdot S_2$ with $S_1, S_2$
-strongly unambiguous and $S_1 \cdot^! S_n$, or (3) $R = S_1 \sep S_2$ with
-$S_1, S_2$ strongly unambiguous and $\mathcal{L}(S_1) \cap \mathcal{L}(S_2) =
-\varnothing$, or (4) $R = S^*$ with $S$ strongly unambiguous and $S^{*!}$.
-
-\begin{figure}[tp!]
-\begin{center}
-\AxiomC{$R$ is strongly unambiguous}
-\UnaryInfC{$\wf{\mathit{id}(R)}$}
-\DisplayProof
-\hskip 1.5em
-\AxiomC{$s \in \mathcal{L}(R)$}
-\UnaryInfC{$\wf{\project{R}{s}}$}
-\DisplayProof
-\hskip 1.5em
-\end{center}
-
-\begin{prooftree}
-\AxiomC{$\wf{c_i, c}$}
-\AxiomC{${\substack{\forall \sigma \neq \theta, \; W(c_{\sigma(1)}) \cdot W(c)
-\cdot \ldots \cdot W(c_{\sigma(n)})\\ \cap W(c_{\theta(1)}) \cdot W(c)
-\cdot \ldots \cdot W(c_{\theta(n)}) =\varnothing}}$}
-\AxiomC{$K(c_1) \cdot^! K(c) \cdot^! \ldots \cdot^! K(c) \cdot^! K(c_n)$}
-\TrinaryInfC{$\wf{\perm{c_1, \ldots, c_n}{c}}$}
-\end{prooftree}
-
-\begin{center}
-\AxiomC{$\wf{c, c'}$}
-\AxiomC{$\mathcal{L}(W(c)) \cap \mathcal{L}(W(c')) = \varnothing$}
-\AxiomC{$f : \mathcal{L}(W(c)) \longrightarrow \mathcal{L}(W(c'))$}
-\TrinaryInfC{$\wf{\squash{c}{c'}{f}}$}
-\DisplayProof
-\end{center}
-\begin{center}
-
-\begin{prooftree}
-\AxiomC{$\mathcal{L}(R') \subseteq \mathcal{L}(R)$}
-\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(R')$}
-\AxiomC{$f$ is surjective}
-\AxiomC{$f = f^2$}
-\QuaternaryInfC{$\wf{\normalize{R}{R'}{f}}$}
-\end{prooftree}
-
-\end{center}
-\begin{center}
-\AxiomC{$\wf{c, c'}$}
-\AxiomC{$K(c) = W(c')$}
-\BinaryInfC{$\wf{c' \; ; \; c}$}
-\DisplayProof
-\hskip 1.5em
-\AxiomC{$\wf{c}$}
-\AxiomC{$W(c)^{*!}$}
-\AxiomC{$K(c)^{*!}$}
-\TrinaryInfC{$\wf{c^*}$}
-\DisplayProof
-\end{center}
-\begin{center}
-
-\begin{center}
-\AxiomC{$\wf{c, c'}$}
-\AxiomC{$W(c) \cdot^! W(c')$}
-\AxiomC{$K(c) \cdot^! K(c')$}
-\TrinaryInfC{$\wf{c \cdot c'}$}
-\DisplayProof
-\end{center}
-
-\begin{center}
-\AxiomC{$\wf{c, c'}$}
-\AxiomC{$W(c) \cap W(c') = \varnothing$}
-\BinaryInfC{$\wf{c \cdot c}$}
-\DisplayProof
-\hskip 1.5em
-\AxiomC{$\wf{c}$}
-\AxiomC{$W(c)^{*!}$}
-\AxiomC{$K(c)^{*!}$}
-\TrinaryInfC{$\wf{c^*}$}
-\DisplayProof
-\end{center}
-\end{center}
-\caption{Well-formed QREs
-}
-\label{fig:qrerules}
-\end{figure}
-
-Figure~\ref{fig:qrerules} gives the inference rules for deriving well-formed
-QREs. As expected, the unambiguity conditions come into play when defining QREs
-using the regular combinators. For example, the inferrence\bcp{check spelling} rule for
-concatenation is as follows:\bcp{Rather than repeating rules inline, give
-  them names in the figure.}
-\begin{prooftree}
-\AxiomC{$\wf{c, c'}$}
-\AxiomC{$W(c) \cdot^! W(c')$}
-\AxiomC{$K(c) \cdot^! K(c')$}
-\TrinaryInfC{$\wf{c \cdot c'}$}
-\end{prooftree}
-
-\noindent This rule says that the concatenation $c \cdot c'$ of QREs $c$ and
-$c'$ is well formed only if the concatentions of $W(c)$ and $W(c')$, and
-$K(c)$ and $K(c')$ are unambiguous.
-
-The most complicated inferrence rule is the rule for the \textit{permutation}
-combinator:
-
-\begin{prooftree}
-\AxiomC{$\wf{c_i, c}$}
-\AxiomC{${\substack{\forall \sigma \neq \theta, \; W(c_{\sigma(1)}) \cdot W(c)
-\cdot \ldots \cdot W(c_{\sigma(n)})\\ \cap W(c_{\theta(1)}) \cdot W(c)
-\cdot \ldots \cdot W(c_{\theta(n)}) =\varnothing}}$}
-\AxiomC{$K(c_1) \cdot^! K(c) \cdot^! \ldots \cdot^! K(c) \cdot^! K(c_n)$}
-\TrinaryInfC{$\wf{\perm{c_1, \ldots, c_n}{c}}$}
-\end{prooftree}
-
-\noindent The second hypothesis says that, for
-any two different permutations $\sigma$ and $\theta$, the languages
-$W(c_{\sigma(1)}) \cdot W(c) \cdot \ldots \cdot W(c_{\sigma(n)})$ and
-$W(c_{\theta(1)}) \cdot W(c) \cdot \ldots \cdot W(c_{\theta(n)})$ must be
-disjoint. This is because any string that matches the \textit{permutation}
-combinator matches the language $W(c_{\theta(1)}) \cdot W(c) \cdot \ldots \cdot
-W(c_{\theta(n)})$ for some permutation $\theta$, so we require that these
-regular expressions be disjoint.
-
-The third hypothesis says that the regular expression $K(c_1) \cdot K(c)
-\cdot \ldots \cdot K(c) \cdot K(c_n)$ must be unambiguous. This is because the
-underlying lens that maps to or from a \textit{permutation} QRE operates on
-the language $K(c_1) \cdot K(c) \cdot \ldots \cdot K(c) \cdot K(c_n)$ which is
-the kernel of the \textit{permutation} QRE. As we mentioned, the underlying
-lens will require that the source and target regular expressions of the lens be
-unambiguous so that the lens can match strings uniquely.
-
-The rest of the inferrence rules for deriving QREs are given in
-Figure~\ref{fig:qrerules}.
-\subsection{Semantics of QREs}
-As we mentioned before, each QRE $c$ enables us to express the whole and kernel
-languages $W(c)$ and $K(c)$, an equivalence relation $\eqrel{c}$ on
-$\mathcal{L}(W(c))$ that has $\mathcal{L}(K(c))$ as a complete set of
-representatives and a canonizing function $\canonizer(c):\mathcal{L}(W(c))
-\longrightarrow \mathcal{L}(K(c))$ whose fibres are the same as $\eqrel{c}$.
-
-Figure~\ref{fig:wk} gives the inductive definitions for the whole and
+Figures~\ref{fig:wk} gives the inductive definitions for the whole and
 kernel languages $W(c)$ and $K(c)$ of a QRE.
-The \kw{squash} and \kw{permutation} combinators give the two interesting
-definitions. If $c = \squash{c_1}{c_2}{f}$, then the whole language of $c$ is
-$W(c_1) \sep W(c_2)$. This is because the \kw{squash} combinator uses the
-function $f : \mathcal{L}(W(c_1)) \longrightarrow \mathcal{L}(W(c_2))$ to maps
-any strings from $\mathcal{L}(W(c_1))$ into $\mathcal{L}(W(c_2))$ (leaving the
-$c_2$ strings alone), and then canonizes $W(c_2)$ into $K(c_2)$ using
-$\canonizer(c_2)$.
-
 \begin{figure}[t]
@@
 \project{R}{s} & R & s \\
-\squash{c_1}{c_2}{f} & W(c_1) \sep W(c_2) & K(c_2) \\
-\normalize{R_1}{R_2}{f} & R_1 & R_2 \\
+\squash{c}{c_1}{f} & W(c) \sep W(c_1) & K(c_1) \\
+\normalize{R}{R'}{f} & R & R' \\
 c_1 \; ; \; c_2 & W(c_1) & K(c_2) \\
@@
 \]
-\caption{Whole and Kernel Regular Expressions \saz{Mention somewhere that $S_n$
-    is the set of all permutations.} \saz{Check that the semi-colon
-    case is still correct}}
+\caption{Whole and Kernel Regular Expressions}
 \label{fig:wk}
 \end{figure}
+The \textit{squash} and \textit{permutation} combinators give the two
+interesting definitions. If $c = \squash{c}{c_1}{f}$, then the whole language of
+$c$ is $W(c) \sep W(c_1)$. This is because the \textit{squash} combinator takes
+the whole language $W(c)$ of $c$, the whole language $W(c_1)$ of $c_1$ and a
+function $f : \mathcal{L}(W(c)) \longrightarrow \mathcal{L}(W(c_1))$, maps
+$\mathcal{L}(W(c))$ into $\mathcal{L}(W(c_1))$ using $f$ and then canonizes
+$W(c_1)$ into $K(c_1)$ using $\canonizer(c_1)$.
 
@@
 union of languages of the form $W(c_{\sigma(1)}) \cdot W(c) \cdot \ldots \cdot
-W(c) \cdot W(c_{\sigma(n)})$. This is because the \kw{perm}
-combinator allows for the string to match any permutation the $c_i$'s while also
-taking into account the separator $c$ in between each of the $c_i$'s. The kernel
-of the \kw{perm} combinator is the language $K(c_1) \cdot K(c) \cdot
-\ldots \cdot K(c) \cdot K(c_n)$, because the canonical permutation is the
-identity permuation, with each of the parts of the string that match $c_i$
-and $c$ canonized into $K(c_i)$ and $K(c)$ respectively.
+W(c) \cdot W(c_{\sigma(n)})$ for any permutation $\sigma$ in $S_n$, where
+$S_n$ is the set of all permutations of the numbers $1$ to $n$. This is because
+the \textit{permutation} combinator allows for the string to match any
+permutation the $c_i$'s while also taking into account the separator $c$ in
+between each of the $c_i$'s. The kernel of the \textit{permutation} combinator
+is the language $K(c_1) \cdot K(c) \cdot \ldots \cdot K(c) \cdot K(c_n)$,
+because the canonical permutation is the identity permuation, with each of the
+parts of the string that match $c_i$ and $c$ canonized into $K(c_i)$ and $K(c)$
+respectively.
 
@@
 \canonizer(\project{R}{s})(w) &=& s\\
-\canonizer(\squash{c}{c'}{f})(w) &=&
+\canonizer(\squash{c_1}{c_2}{f})(w) &=&
 \begin{cases}
-\canonizer(c)(f(w)) & \text{if } w \in \mathcal{L}(W(c))\\
-\canonizer(c')(w) & \text{otherwise}
+\canonizer(c_1)(f(w)) & \text{if } w \in \mathcal{L}(W(c_1))\\
+\canonizer(c_2)(w) & \text{otherwise}
 \end{cases}\\
 \canonizer(\normalize{R}{R'}{f}) &=& f\\
-\canonizer(c' \; ; \; c) &=& \canonizer(c') \circ \canonizer(c)\\
-\canonizer(c' \cdot c) &=& \canonizer(c') \cdot \canonizer(c)\\
+\canonizer(c_1 \; ; \; c_2) &=& \canonizer(c_2) \circ \canonizer(c_1)\\
+\canonizer(c_1 \cdot c_2) &=& \canonizer(c_1) \cdot \canonizer(c_2)\\
 \canonizer(c_1 \sep c_2)(w) &=&
@@
 \end{align*}
-\caption{QRE canonizers \saz{TODO: use $c_1$ and $c_2$ everywhere.}}
+\caption{QRE canonizers}
 \label{fig:canonizers}
 \end{figure}
-Again, the \kw{perm} combinator gives the most interesting
+Again, the \textit{permutation} combinator gives the most interesting
 definition,
@@
 
-Finally, Figure ~\ref{fig:relations} gives the inductive defintion of the
+Finally, Figure ~\ref{fig:relations} gives the induction defintion of the
 equivalence relation $\eqrel{c}$. These are the set-theoretic semantics of QREs
@@
 w \; \equiv_{\project{R}{s}} \; w' \text{ for all }w, w'\\
-w \; \equiv_{\squash{c}{c'}{f}} \; w' &\iff& f(w) \equiv_{c'} w'
-\text{ or } w \equiv_{c'} w' \\
+w \; \equiv_{\squash{c_1}{c_2}{f}} \; w' &\iff& f(w) \equiv_{c_2} w'
+\text{ or } w \equiv_{c_2} w' \\
 w \; \equiv_{\normalize{R}{R'}{f}} \; w' &\iff&
 f(w)=f(w') \text{ or }w = w'\\
-w \; \equiv_{c_2 \; ; \; c_1} \; w' &\iff& \exists k, k' \in
+w \; \equiv_{c_1 \; ; \; c_2} \; w' &\iff& \exists k, k' \in
 \mathcal{L}(K(c_2)) \text{ such that } w \; \equiv_{c_1} \; k, \; w' \;
@@
 \; {r'}_1, \; r_2 \; \equiv{c_n} \; {r'}_2\\
-w \; \equiv_{c' \sep c} \; w' &\iff& w \; \equiv_{c_1} \; w'
+w \; \equiv_{c_1 \sep c_2} \; w' &\iff& w \; \equiv_{c_1} \; w'
 \text{ or } \; w \; \equiv_{c_2} \; w'\\
@@
 \]
-\caption{QRE Equivalence Relations \saz{TODO: use $c_1$ and $c_2$ everywhere.}}
+\caption{QRE Equivalence Relations}
 \label{fig:relations}
 \end{figure}
+\subsection{Ambiguity and Well Formed QREs}
+Before giving the inference rules for deriving well formed QREs, we address
+the issue of {\em ambiguity}, which appears in the premises of the inference
+rules for the regular combinators. If $c$ and $c_1$ are QREs, then when applying
+the regular combinators to QREs, we require that if a string $s$ matches any of
+the regular expressions $W(c) \cdot W(c_1)$, \quad $W(c) \sep W(c_1)$,
+\quad $W(c)^*$, \quad $K(c) \cdot K(c_1)$, \quad $K(c) \sep K(c_1)$,
+\quad $K(c)^*$, then $s$ matches that regular expression in only one way.
+
+More formally, we say that regular expressions $R$ and $S$ are
+\textit{unambiguosly concatenable}, written $R \cdot^! S$ if for all strings
+$r, r' \in \mathcal{L}(R)$ and $s, s' \in \mathcal{L}(S)$, if $r \cdot s = r'
+\cdot s'$, then $r = r'$ and $s = s'$. We say that a regular expression $R$ is
+\textit{unambiguosly iterable}, written $R^{*!}$ if for all strings $r_1,
+\ldots, r_m$ and $r'_1, \ldots, r'_n \in \mathcal{L}(R)$, if $r_1 \cdot \ldots
+\cdot r_m = r'_1 \cdot \ldots \cdot r'_n$, then $m = n$ and $r_i = r'_i$. We say
+that a regular expression $R$ is \textit{strongly unambiguous}\cite{Sippu1988}
+if and only if (1) $R = \varnothing$, or (2) $R = S_1 \cdot S_2$ with $S_1,
+S_2$ strongly unambiguous and $S_1 \cdot^! S_n$, or (3) $R = S_1 \sep S_2$ with
+$S_1, S_2$ strongly unambiguous and $\mathcal{L}(S_1) \cap \mathcal{L}(S_2) =
+\varnothing$, or (4) $R = S^*$ with $S$ strongly unambiguous and $S^{*!}$.
+
+Strong unambiguity is necessary in the premises of the regular QRE combinators
+because it ensures that the canonizing function of a QRE is well-defined.
+For example consider the QRE \lstinline{"a"* . (project "a"* $\to$ "a")}. The
+behaviour of this QRE is not-well defined since the string
+\cd{"aaa"} can be canonized to any of \cd{"a"}, \cd{"aa"}, or \cd{"aaa"}
+depending on how \cd{"aaa"} is parsed. We also require that the regular
+combinators applied to the kernels of QREs be unambiguous since the underlying
+bijective lens of a QRE lens operates on kernels, and bijective lenses
+impose the same unambiguity restrictions so that they too are well defined as
+functions.
+
+Figure~\ref{fig:qrerules} gives the inferrence rules for deriving well-formed
+QREs.
+\begin{figure}[tp!]
+\begin{center}
+\AxiomC{$R$ is strongly unambiguous}
+\UnaryInfC{$\wf{\mathit{id}(R)}$}
+\DisplayProof
+\hskip 1.5em
+\AxiomC{$s \in \mathcal{L}(R)$}
+\UnaryInfC{$\wf{\project{R}{s}}$}
+\DisplayProof
+\hskip 1.5em
+\end{center}
+
+\begin{prooftree}
+\AxiomC{$\wf{c_i, c}$}
+\AxiomC{${\substack{\forall \sigma \neq \theta, \; W(c_{\sigma(1)}) \cdot W(c)
+\cdot \ldots \cdot W(c_{\sigma(n)})\\ \cap W(c_{\theta(1)}) \cdot W(c)
+\cdot \ldots \cdot W(c_{\theta(n)}) =\varnothing}}$}
+\AxiomC{$K(c_1) \cdot^! K(c) \cdot^! \ldots \cdot^! K(c) \cdot^! K(c_n)$}
+\TrinaryInfC{$\wf{\perm{c_1, \ldots, c_n}{c}}$}
+\end{prooftree}
+
+\begin{center}
+\AxiomC{$\wf{c_1, c_2}$}
+\AxiomC{$\mathcal{L}(W(c_1)) \cap \mathcal{L}(W(c_2)) = \varnothing$}
+\AxiomC{$f : \mathcal{L}(W(c_1)) \longrightarrow \mathcal{L}(W(c_2))$}
+\TrinaryInfC{$\wf{\squash{c_1}{c_2}{f}}$}
+\DisplayProof
+\end{center}
+\begin{center}
+
+\begin{prooftree}
+\AxiomC{$\mathcal{L}(R') \subseteq \mathcal{L}(R)$}
+\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(R')$}
+\AxiomC{$f$ is surjective}
+\AxiomC{$f = f^2$}
+\QuaternaryInfC{$\wf{\normalize{R}{R'}{f}}$}
+\end{prooftree}
+
+\end{center}
+\begin{center}
+\AxiomC{$\wf{c_1, c_2}$}
+\AxiomC{$K(c_1) = W(c_2)$}
+\BinaryInfC{$\wf{c_2 \; ; \; c_1}$}
+\DisplayProof
+\hskip 1.5em
+\AxiomC{$\wf{c}$}
+\AxiomC{$W(c)^{*!}$}
+\AxiomC{$K(c)^{*!}$}
+\TrinaryInfC{$\wf{c^*}$}
+\DisplayProof
+\end{center}
+\begin{center}
+
+\begin{center}
+\AxiomC{$\wf{c_1, c_2}$}
+\AxiomC{$W(c_1) \cdot^! W(c_2)$}
+\AxiomC{$K(c_1) \cdot^! K(c_2)$}
+\TrinaryInfC{$\wf{c_1 \cdot c_2}$}
+\DisplayProof
+\end{center}
+
+\begin{center}
+\AxiomC{$\wf{c_1, c_2}$}
+\AxiomC{$W(c_1) \cap W(c_2) = \varnothing$}
+\BinaryInfC{$\wf{c_1 \cdot c_2}$}
+\DisplayProof
+\hskip 1.5em
+\AxiomC{$\wf{c}$}
+\AxiomC{$W(c)^{*!}$}
+\AxiomC{$K(c)^{*!}$}
+\TrinaryInfC{$\wf{c^*}$}
+\DisplayProof
+\end{center}
+\end{center}
+\caption{Well-formed QREs
+}
+\label{fig:qrerules}
+\end{figure}
+As we already mentioned, the unambiguity conditions come into play when defining
+QREs using the regular combinators. For example, the inferrence rule for
+concatenation is as follows:\bcp{Rather than repeating rules inline, give
+  them names in the figure.}
+\begin{prooftree}
+\AxiomC{$\wf{c, c_1}$}
+\AxiomC{$W(c) \cdot^! W(c_1)$}
+\AxiomC{$K(c) \cdot^! K(c_1)$}
+\TrinaryInfC{$\wf{c \cdot c_1}$}
+\end{prooftree}
+
+\noindent This rule says that the concatenation $c \cdot c_1$ of QREs $c$ and
+$c_1$ is well formed only if the concatentions of $W(c)$ and $W(c_1)$, and
+$K(c)$ and $K(c_1)$ are unambiguous.
+
+The most complicated inferrence rule is the rule for the \textit{permutation}
+combinator:
+
+\begin{prooftree}
+\AxiomC{$\wf{c_i, c}$}
+\AxiomC{${\substack{\forall \sigma \neq \theta, \; W(c_{\sigma(1)}) \cdot W(c)
+\cdot \ldots \cdot W(c_{\sigma(n)})\\ \cap W(c_{\theta(1)}) \cdot W(c)
+\cdot \ldots \cdot W(c_{\theta(n)}) =\varnothing}}$}
+\AxiomC{$K(c_1) \cdot^! K(c) \cdot^! \ldots \cdot^! K(c) \cdot^! K(c_n)$}
+\TrinaryInfC{$\wf{\perm{c_1, \ldots, c_n}{c}}$}
+\end{prooftree}
+
+\noindent The second hypothesis for the \textit{permutation} rule says that for
+any two different permutations $\sigma$ and $\theta$, the languages
+$W(c_{\sigma(1)}) \cdot W(c) \cdot \ldots \cdot W(c_{\sigma(n)})$ and
+$W(c_{\theta(1)}) \cdot W(c) \cdot \ldots \cdot W(c_{\theta(n)})$ must be
+disjoint. This is because any string that matches the \textit{permutation}
+combinator matches the language $W(c_{\theta(1)}) \cdot W(c) \cdot \ldots \cdot
+W(c_{\theta(n)})$ for some permutation $\theta$, so we require that these
+regular expressions be disjoint.
+
+The third hypothesis says that the regular expression $K(c_1) \cdot K(c)
+\cdot \ldots \cdot K(c) \cdot K(c_n)$ must be unambiguous. This is because the
+underlying lens that maps to or from a \textit{permutation} QRE operates on
+the language $K(c_1) \cdot K(c) \cdot \ldots \cdot K(c) \cdot K(c_n)$ which is
+the kernel of the \textit{permutation} QRE. As we mentioned, the underlying
+lens will require that the source and target regular expressions of the lens be
+unambiguous so that the lens can match strings uniquely.
+
+At this point we also bring to attention the fact that in order for the {\em
+normalize} QRE to be well-formed, then canonizing function $f$ used by this
+combinator must be surjective and idempotent. Unfortunately, verifying
+that a function satisfies these properties is in general undecidable.
+Nonetheless, we allow a programmer to use normalize at their own risk without
+checking these side conditions.
+
+We do this firstly because the {\em normalize} combinator allows one to express
+equivalence relations that are either difficult to define, or that cannot be
+defined using the other combinators. Secondly, the {\em normalize} canonizer
+enjoys the property that each of the other combinators $c$ is semantically
+equivalent to $\normalize{W(c)}{K(c)}{f}$ for some surjective, idempotent
+function $f : \mathcal{L}(W(c)) \longrightarrow \mathcal{L}(K(c))$. The {\em
+normalize} combinator hence gives a sufficient condition for potential QREs to
+be ``well-behaved'', particularly with respect to our strategy for synthesizing
+quotient lenses which we shall introduce in Section~\ref{synth}.
+
+In spite of this ``completeness'' property of the {\em normalize} combinator,
+our experience with QREs shows that using the {\em project}, {\em squash} and
+{\em perm} combinators as well as the regular operators applied to these is
+much easier than using just {\em normalize} because of the compositional nature
+of the other combinators.
 
@@
 
-Now let $c$ and $c'$ be QREs. Given a bijection $\ell : K(c) \Leftrightarrow
-K(c')$, we may define a quotient lens $q : W(c)/\eqrel{c} \Leftrightarrow
-W(c')/\eqrel{c'}$ by
+Now let $c$ and $c_1$ be QREs. Given a bijection $\ell : K(c) \Leftrightarrow
+K(c_1)$, we may define a quotient lens $q : W(c)/\eqrel{c} \Leftrightarrow
+W(c_1)/\eqrel{c_1}$ by
 \begin{equation}\label{normalform}
 q.get = \ell \circ \canonizer(c) \text{ and } q.put = \ell^{-1} \circ
-\canonizer(c'),
+\canonizer(c_1),
 \end{equation}
@@
 put function of the $q$ first canonizes the target data using
-$\canonizer(c')$ and the applies $f^{-1}$ to the result. This is our approach in
+$\canonizer(c_1)$ and the applies $f^{-1}$ to the result. This is our approach in
 defining QRE lenses.
@@
 typing judgements. More concretely, each QRE lens $q$ has a type $q : c
-\Leftrightarrow c'$ where $c, c'$ are QREs. In contrast, the approach used to
+\Leftrightarrow c_1$ where $c, c_1$ are QREs. In contrast, the approach used to
 type quotient lenses in Boomerang is to classify equivalences according to
@@
 
-Each QRE lens $q$ has a type $q : c \Leftrightarrow c'$ where $c, c'$ are QREs.
-If $q : c \Leftrightarrow c'$, then the source format is described by $W(c)$ and
+Each QRE lens $q$ has a type $q : c \Leftrightarrow c_1$ where $c, c_1$ are QREs.
+If $q : c \Leftrightarrow c_1$, then the source format is described by $W(c)$ and
 the canonical set of representative for the source data is described by $K(c)$.
-Similarly, the target format is described by $W(c')$ and the canonical set of
-representatives for the target data is described by $K(c')$. The underlying lens
-of $q$ is a bijective lens $\ell : K(c) \Leftrightarrow K(c')$.
-
-The denotation $\llbracket q \rrbracket$ of a QRE lens $q:c \Leftrightarrow c'$
+Similarly, the target format is described by $W(c_1)$ and the canonical set of
+representatives for the target data is described by $K(c_1)$. The underlying lens
+of $q$ is a bijective lens $\ell : K(c) \Leftrightarrow K(c_1)$.
+
+The denotation $\llbracket q \rrbracket$ of a QRE lens $q:c \Leftrightarrow c_1$
 is a quotient lens $\llbracket q \rrbracket : W(c)/{\eqrel{c}}
-\Longleftrightarrow W(c')/{\eqrel{c'}}$. The typing rules and denotation of QRE
+\Longleftrightarrow W(c_1)/{\eqrel{c_1}}$. The typing rules and denotation of QRE
 lenses are given in Figure~\ref{fig:qlenssemantics}. The trickiest typing rule
@@
 \begin{prooftree}
-\AxiomC{$\substack{q_1 : c \Leftrightarrow c' \\ q_2 : c'' \Leftrightarrow
-c'''}$}
-\AxiomC{$\substack{\mathcal{L}(W(c'))= \mathcal{L}(W(c'')) \\ K(c') \equiv
-K(c'')}$}
-\AxiomC{$\canonizer(c') = \canonizer(c'')$}
-\TrinaryInfC{$q_2 \circ q_1: c \Leftrightarrow c'''$}
+\AxiomC{$\substack{q_1 : c \Leftrightarrow c_1 \\ q_2 : c_2 \Leftrightarrow
+c_3}$}
+\AxiomC{$\substack{\mathcal{L}(W(c_1))= \mathcal{L}(W(c_2)) \\ K(c_1) \equiv
+K(c_2)}$}
+\AxiomC{$\canonizer(c_1) = \canonizer(c_2)$}
+\TrinaryInfC{$q_2 \circ q_1: c \Leftrightarrow c_3$}
 \end{prooftree}
@@
 This derivation rule essentially says that the composition $q_2 \circ q_1$ is
-well defined if and only if the intermediary QREs $c''$ and $c'''$ define the
+well defined if and only if the intermediary QREs $c_2$ and $c_3$ define the
 same equivalence relation on the same regular language. The condition
-$\mathcal{L}(W(c')) = \mathcal{L}(W(c''))$ says that the intermediary language
-is the same on both sides, while the condition $K(c') \equiv K(c'')$ says that
+$\mathcal{L}(W(c_1)) = \mathcal{L}(W(c_2))$ says that the intermediary language
+is the same on both sides, while the condition $K(c_1) \equiv K(c_2)$ says that
 the kernel regular expressions are syntactically the same. Finally, the
-condition $\canonizer(c') = \canonizer(c'')$ says that $c'$ and $c''$ define
-the same equivalence relation on $\mathcal{L}(W(c'))= \mathcal{L}(W(c''))$.
-
-In general, checking that $\canonizer(c') = \canonizer(c'')$ is an undecidable
+condition $\canonizer(c_1) = \canonizer(c_2)$ says that $c_1$ and $c_2$ define
+the same equivalence relation on $\mathcal{L}(W(c_1))= \mathcal{L}(W(c_2))$.
+
+In general, checking that $\canonizer(c_1) = \canonizer(c_2)$ is an undecidable
 task, especially because of the $\normalize{R}{R'}{f}$ QRE. Therefore, our
@@
 approach, which is to define the composition on QRE lenses if and only if
-$c'$ and $c''$ are syntactially equal to $id(R)$ and $id(R')$ for some $R$ and
+$c_1$ and $c_2$ are syntactially equal to $id(R)$ and $id(R')$ for some $R$ and
 $R'$ satisfying $\mathcal{L}(R) = \mathcal{L}(R')$.
@@
 \begin{theorem}
-If there is a derivation $q : c \Leftrightarrow c'$, then $\llbracket q
-\rrbracket : W(c)/{\eqrel{c}} \Leftrightarrow W(c')/{\eqrel{c'}}$ is a
+If there is a derivation $q : c \Leftrightarrow c_1$, then $\llbracket q
+\rrbracket : W(c)/{\eqrel{c}} \Leftrightarrow W(c_1)/{\eqrel{c_1}}$ is a
 well-defined quotient lens.
@@
 \[
-\begin{array}[b]{c@{\qquad}l}
-\Rule{q : c'  \Leftrightarrow c'' &
+\begin{array}[b]{c@{\qquad}c}
+\Rule{q : c_1  \Leftrightarrow c_2 &
 \wf{c} &
-K(c) = W(c')}
-{\kw{lquot}(c, q): c' \circ c \Leftrightarrow c''} &
+K(c) = W(c_1)}
+{\kw{lquot}(c, q): c_1 \circ c \Leftrightarrow c_2} &
 \begin{array}[b]{rcl}
@@
 \begin{array}[b]{c@{\qquad}c}
-\Rule{q : c'  \Leftrightarrow c'' &
-\wf{c} &
-K(c) = W(c')}
-{\kw{lquot}(c, q): c' \circ c \Leftrightarrow c''} &
-\begin{array}[b]{rcl}
-\llbracket \kw{lquot}(c, q) \rrbracket.get  &=& \llbracket q
-\rrbracket.get \circ \canonizer(c)\\
-\llbracket \kw{lquot}(c, q) \rrbracket.put &=& \llbracket q \rrbracket.put
-\end{array}
-\end{array}
-\]
-
-\[
-\begin{array}[b]{c@{\qquad}c}
-\Rule{q : c \Leftrightarrow c'' &
-\wf{c''} &
-W(c'') = K(c')}
-{\kw{rquot}(q, c):c \Leftrightarrow c'' \circ c'} &
+\Rule{q : c \Leftrightarrow c_2 &
+\wf{c_2} &
+W(c_2) = K(c_1)}
+{\kw{rquot}(q, c):c \Leftrightarrow c_2 \circ c_1} &
 \begin{array}[b]{rcl}
@@
 \llbracket \kw{rquot}(q, c) \rrbracket.put &=& \llbracket q
-\rrbracket.put \circ \canonizer(c'')
+\rrbracket.put \circ \canonizer(c_2)
 \end{array}
@@
 \begin{array}[b]{l@{\qquad}c}
-\Rule{q : c \Leftrightarrow c' &
-W(c)^{*!},W(c')^{*!} & K(c)^{*!}, K(c')^{*!}}
-{{q}^* : c^* \Leftrightarrow {c'}^*} &
+\Rule{q : c \Leftrightarrow c_1 &
+W(c)^{*!},W(c_1)^{*!} & K(c)^{*!}, K(c_1)^{*!}}
+{{q}^* : c^* \Leftrightarrow {c_1}^*} &
 \begin{array}[b]{rcl}
@@
 Recall that our approach in defining QRE lenses is to have each QRE lens $q: c
-\Leftrightarrow c'$ be such that
+\Leftrightarrow c_1$ be such that
 \begin{align*}
@@
 \llbracket q \rrbracket.put &= \ell^{-1} \circ
-\canonizer(c')
+\canonizer(c_1)
 \end{align*}
@@
 \begin{theorem}\label{normal form}
-If there is a derivation $q : c \Leftrightarrow c'$, then there exists a
-bijective lens $\ell : K(c) \Leftrightarrow K(c')$ such that
+If there is a derivation $q : c \Leftrightarrow c_1$, then there exists a
+bijective lens $\ell : K(c) \Leftrightarrow K(c_1)$ such that
 \begin{align*}
@@
 \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
-\canonizer(c')
+\canonizer(c_1)
 \end{align*}
@@
 $\ell_1 :
-K(c) \Leftrightarrow K(c')$ and $\ell_2 : K(c'') \Leftrightarrow K(c''')$ such
+K(c) \Leftrightarrow K(c_1)$ and $\ell_2 : K(c_2) \Leftrightarrow K(c_3)$ such
 that
@@
 \llbracket q_1 \rrbracket.put &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
-\canonizer(c')
+\canonizer(c_1)
 \end{align*}
@@
 \llbracket q_2 \rrbracket.get &= \llbracket \ell_2 \rrbracket \circ
-Canonize(c')\\
+Canonize(c_1)\\
 \llbracket q_2 \rrbracket.put &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
-Canonize(c'')
+Canonize(c_2)
 \end{align*}
@@
 \llbracket q_2 \rrbracket.get \circ \llbracket q_1 \rrbracket.get &=
-(\llbracket \ell_2 \rrbracket \circ \canonizer(c')) \circ (\llbracket \ell_1
+(\llbracket \ell_2 \rrbracket \circ \canonizer(c_1)) \circ (\llbracket \ell_1
 \rrbracket \circ \canonizer(c))\\
-&= \llbracket \ell_2 \rrbracket \circ (\canonizer(c') \circ \llbracket \ell_1
+&= \llbracket \ell_2 \rrbracket \circ (\canonizer(c_1) \circ \llbracket \ell_1
 \rrbracket) \circ \canonizer(c)\\
@@
 We are permitted to claim the third step from the second since
-$\canonizer(c')$ is the identity function on $K(c'')$ which is
-syntactically equal to $K(c')$ by assumption. A similar argument shows that
+$\canonizer(c_1)$ is the identity function on $K(c_2)$ which is
+syntactically equal to $K(c_1)$ by assumption. A similar argument shows that
 $$\llbracket q_1 \rrbracket.put \circ \llbracket q_2 \rrbracket.put =
@@
 \begin{proof}
-Assume that $q : c \Leftrightarrow c'$. We proceed by induction over the
-derivation $q : c \Leftrightarrow c'$.
+Assume that $q : c \Leftrightarrow c_1$. We proceed by induction over the
+derivation $q : c \Leftrightarrow c_1$.
 \begin{enumerate}
@@
 \item
-$\kw{lquot}(c, q'): c' \circ c \Leftrightarrow c''$ where $q' : c'
-\Leftrightarrow c''$, $c$ is well formed and $K(c) = W(c')$. Then
+$\kw{lquot}(c, q'): c_1 \circ c \Leftrightarrow c_2$ where $q' : c_1
+\Leftrightarrow c_2$, $c$ is well formed and $K(c) = W(c_1)$. Then
 \begin{align*}
@@
 By the induction hypothesis, there exists a bijective lens $\ell :
-K(c') \Leftrightarrow K(c'')$ such that
-\begin{align*}
-\llbracket q' \rrbracket.get &= \llbracket \ell \rrbracket \circ \canonizer(c')\\
+K(c_1) \Leftrightarrow K(c_2)$ such that
+\begin{align*}
+\llbracket q' \rrbracket.get &= \llbracket \ell \rrbracket \circ \canonizer(c_1)\\
 \llbracket q' \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
-Canonize(c'')
+Canonize(c_2)
 \end{align*}
@@
 \llbracket q \rrbracket.get  &= (\llbracket \ell \rrbracket \circ
-\canonizer(c')) \circ \canonizer(c) = \llbracket \ell \rrbracket \circ
-(\canonizer(c' \circ c))\\
+\canonizer(c_1)) \circ \canonizer(c) = \llbracket \ell \rrbracket \circ
+(\canonizer(c_1 \circ c))\\
 \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
-Canonize(c'')
+Canonize(c_2)
 \end{align*}
@@
 \item
-$\kw{rquot}(q', c''):c \Leftrightarrow c'' \circ c'$ where $q' : c \Leftrightarrow
-c'$, $c''$ is well formed and $K(c'') = W(c')$. Proceed as in the previous
+$\kw{rquot}(q', c_2):c \Leftrightarrow c_2 \circ c_1$ where $q' : c \Leftrightarrow
+c_1$, $c_2$ is well formed and $K(c_2) = W(c_1)$. Proceed as in the previous
 case.
 \item
-$q_2 \circ q_1: c \Leftrightarrow c''$ where $q_1 : c \Leftrightarrow c'$ and
-$q_2 : c' \Leftrightarrow c''$. Then
+$q_2 \circ q_1: c \Leftrightarrow c_2$ where $q_1 : c \Leftrightarrow c_1$ and
+$q_2 : c_1 \Leftrightarrow c_2$. Then
 \begin{align*}
@@
 $\ell_1 :
-K(c) \Leftrightarrow K(c')$ and $\ell_2 : K(c') \Leftrightarrow K(c'')$ such
+K(c) \Leftrightarrow K(c_1)$ and $\ell_2 : K(c_1) \Leftrightarrow K(c_2)$ such
 that
@@
 \item
-${q'}^* : c^* \Leftrightarrow {c'}^*$ where $q' : c \Leftrightarrow c'$,
-$W(c)^{*!}$ and $W(c')^{*!}$ and $K(c)^{*!}$ and $K(c')^{*!}$. Then
+${q'}^* : c^* \Leftrightarrow {c_1}^*$ where $q' : c \Leftrightarrow c_1$,
+$W(c)^{*!}$ and $W(c_1)^{*!}$ and $K(c)^{*!}$ and $K(c_1)^{*!}$. Then
 \begin{align*}
@@
 By the induction hypothesis there exists a bijective lens $\ell : K(c)
-\Leftrightarrow K(c')$ such that
+\Leftrightarrow K(c_1)$ such that
 that
@@
 \llbracket q' \rrbracket.put &= {\llbracket \ell \rrbracket}^{-1} \circ
-\canonizer(c')
+\canonizer(c_1)
 \end{align*}
@@
 \llbracket {q'}^* \rrbracket.put &= (\llbracket \ell \rrbracket^{-1} \circ
-\canonizer(c'))^* = (\llbracket \ell \rrbracket^{-1})^* \circ
-\canonizer(c')^* = \llbracket \ell^* \rrbracket^{-1} \circ
-\canonizer(c'^*)\\
+\canonizer(c_1))^* = (\llbracket \ell \rrbracket^{-1})^* \circ
+\canonizer(c_1)^* = \llbracket \ell^* \rrbracket^{-1} \circ
+\canonizer(c_1^*)\\
 \end{align*}
@@
 Our goal in this section is to demonstrate how to synthesize a quotient lens
-$q: c \Leftrightarrow c'$ from the QREs $c, c'$ and a set of input-output pairs
+$q: c \Leftrightarrow c_1$ from the QREs $c, c_1$ and a set of input-output pairs
 $\{(x_1, y_1),\allowbreak \ldots,\allowbreak (x_n, y_n)\}$ where the $x_i$'s are in $W(c)$ and the
-$y_i$'s are in $W(c')$. We wish $q$ to map the equivalence class of $x_i$ to the
+$y_i$'s are in $W(c_1)$. We wish $q$ to map the equivalence class of $x_i$ to the
 equivalence class of $y_i$ and vice versa:
 \begin{align*}
-q.get(x_i) &\equiv_{c'} y_i \text{, and}\\
+q.get(x_i) &\equiv_{c_1} y_i \text{, and}\\
 q.put(y_i) &\equiv_c x_i
@@
 form}, which says that if there is a derivation $q :
-c \Leftrightarrow c'$ of a QRE lens, then there exists a bijective lens $\ell :
-K(c) \Leftrightarrow K(c')$ such that
+c \Leftrightarrow c_1$ of a QRE lens, then there exists a bijective lens $\ell :
+K(c) \Leftrightarrow K(c_1)$ such that
 \begin{align*}
@@
 For our synthesis task, since the $x_i$'s are in $W(c)$ and the $y_i$'s are
-in $W(c')$, then ${x_i}' = \canonizer(c)(x_i)$ is in $K(c)$ and ${y_i}' =
-\canonizer(c')(y_i)$ is in $K(c')$. Consequently in order to synthesize the
-desired quotient lens $q: c \Leftrightarrow c'$ that is consistent with the
+in $W(c_1)$, then ${x_i}' = \canonizer(c)(x_i)$ is in $K(c)$ and ${y_i}' =
+\canonizer(c_1)(y_i)$ is in $K(c_1)$. Consequently in order to synthesize the
+desired quotient lens $q: c \Leftrightarrow c_1$ that is consistent with the
 input-output examples $\{(x_1, y_1), \ldots, (x_n, y_n)\}$ it suffices to
 synthesize a bijective lens $l :
-K(c) \Leftrightarrow K(c')$ that is consistent with the canonized examples
+K(c) \Leftrightarrow K(c_1)$ that is consistent with the canonized examples
 $\{({x_1}', {y_1}'), \ldots, ({x_n}', {y_n}')\}$
@@
 This implies that if there is a derivation $q : c \Leftrightarrow
-c'$ of a QRE lens as well as a derivation $\ell : K(c) \Leftrightarrow K(c')$
+c_1$ of a QRE lens as well as a derivation $\ell : K(c) \Leftrightarrow K(c_1)$
 that is consistent with the canonized examples $\{({x_1}', {y_1}'),\allowbreak
 \ldots,\allowbreak ({x_n}', {y_n}')\}$, then we can synthesize a quotient lens $q$ of
-type $c \Leftrightarrow c'$ that is consistent with $\{({x_1}', {y_1}'),
+type $c \Leftrightarrow c_1$ that is consistent with $\{({x_1}', {y_1}'),
 \ldots, ({x_n}', {y_n}')\}$. More concretely, in order to synthesize a
-quotient lens $q: c \Leftrightarrow c'$, that is consistent with the examples
+quotient lens $q: c \Leftrightarrow c_1$, that is consistent with the examples
 $\{(x_1, y_1),\allowbreak \ldots, (x_n, y_n)\}$ it suffices to perform the following
@@
   \item
-  Compute $K(c)$, $K(c')$,
+  Compute $K(c)$, $K(c_1)$,
   \item
   Compute $\{({x_1}', {y_1}'), \ldots, ({x_n}', {y_n}')\}$ where ${x_i}' =
-  \canonizer(c)(x_i)$ and ${y_i}' = \canonizer(c')(y_i)$,
+  \canonizer(c)(x_i)$ and ${y_i}' = \canonizer(c_1)(y_i)$,
   \item
-  Synthesize a bijective lens $\ell : K(c) \Leftrightarrow K(c')$ that is
+  Synthesize a bijective lens $\ell : K(c) \Leftrightarrow K(c_1)$ that is
   consistent with the examples $\{({x_1}', {y_1}'),\allowbreak
@@
   \item
-  Return $q = \kw{rquot}(\kw{lquot}(c, \ell), c') : c \Leftrightarrow
-  c'$
+  Return $q = \kw{rquot}(\kw{lquot}(c, \ell), c_1) : c \Leftrightarrow
+  c_1$
 \end{enumerate}

### 2018-03-08 papers/QuotientLenses/main.tex
Added more edits

@@
 \newif\ifdraft\drafttrue  % set true to show comments
-\newif\ifplentyoftime\plentyoftimefalse  % :-)
+\newif\iflastminute\lastminutefalse  % for things to check at the end
 
@@
 \usepackage{amsmath, amssymb, verbatim, enumerate, graphicx, centernot, tikz,
-array, mathtools, bussproofs, stmaryrd, enumitem, stackengine, subcaption}
+array, mathtools, bussproofs, stmaryrd, enumitem, stackengine, subcaption,
+cleveref}
 \captionsetup{compatibility=false}
@@
 \newcommand{\re}[1]{\ensuremath{\mathtt{#1}}}
-\newcommand{\kw}[1]{\ensuremath{\mathsf{#1}}}
+\newcommand{\kw}[1]{\ensuremath{\mathit{#1}}}
 \newcommand{\codefont}[1]{\ensuremath{\mathsf{#1}}}
@@
 \newcommand{\sep}{\ensuremath{\ | \ }}
-\newcommand{\canonizer}{\ensuremath{\kw{Canonizer}}}
+\newcommand{\canonizer}{\ensuremath{\kw{canonizer}}}
 \newcommand{\bibtex}{\textsc{Bib}\TeX{}}
@@
 \newcommand{\create}{\ensuremath{\kw{create}}}
+\newcommand{\const}{\ensuremath{\kw{const}}}
+\newcommand{\swap}{\ensuremath{\kw{swap}}}
+
 \newcommand{\eqrel}[1]{\ensuremath{\equiv_{#1}}}
@@
 \emph{i.e.,} programs that represent pairs of functions for translating data
-from format A to B (the \emph{get} direction) and back again (the \emph{put}
-direction). Lenses are usually derived using domain-specific
-bidirectional languages which guarantee that the lenses defined in that language
-satisfy a certain set of \emph{lens laws} that give guarantees on how the two
-transformations interact.
+from format the source format to the target formant (the \emph{get}
+direction) and back again (the \emph{put} direction). Lenses are usually
+derived using domain-specific bidirectional languages which guarantee that the
+lenses defined in that language satisfy a certain set of \emph{lens laws} that
+give guarantees on how the two transformations interact.
 
@@
 programmers often must rearrange the order of data items by recursively using
-operators that swap adjacent fields. Furthermore, the Boomerang type checker is
+operators that \swap adjacent fields. Furthermore, the Boomerang type checker is
 very strict, disallowing many programs because they contain ambiguity about how
@@
 
-The first way in which our work improves upon Boomerang's quotient lenses is
+The first way in which our work simplifies Boomerang's quotient lenses is
 that we merge the first three pieces of quotient lenses that we just described
@@
 \end{lstlisting}
+\iflastminute
 \bcp{Can we display blanks as a distinct ``visible blank'' symbol
-(throughout)?  That will make examples easier to read.}
+(throughout)?  That will make examples easier to read.}\bcp{I tried turning
+on the showstringspaces option at the top, but that doesn't do everything we
+want--it only works inside quotes.  But the latex ``textvisiblespace'' command
+produces a single one, and the lstlisting ``escapechar'' option will allow
+us to use it inside listings.  But we should save this for the final
+tweaking stage.}
+% Here is an example: if you set escapechar=\%, you could use it inside the
+% listings as follows: \begin{lstlisting} t %\leftarrow% 0 \end{lstlisting} 
+\fi
 
@@
 
-\sam{Here is my response to this. You observe that the idea of using half a lens
-as a canonizer comes out of nowhere, and that this ideas is a hack and always
-was. I don't think that's quite right. I think that it is useful to have a
-combinator which chooses between two or more formats. Indeed, the permutation
-combinator is just a special case of this. The bibtex name field is a good
-example of where such a combinator would be useful.  But if we choose to have
-such a combinator, then we need a way of mapping one format to another, and
-really the only interesting string-to-string transformations in Boomerang are
-given by lenses, which is why we use the get function of a lens in a canonizer.
-We could have used any other function between the two formats but Boomerang
-doesn't give us too many options for writing unidirectional transformations. So
-I don't think that using a lens as a canonizer is a trick. I think that
-Boomerang just isn't designed for unidirectional tranformations. But I also
-agree that it doesn't make much sense to synthesize a lens only to get rid of
-one half of it.
-I'm not sure what the solution is to be honest.}
-
-The first line above synthesizes a
-lens between \codefont{comma\_name} and \codefont{space\_name} using the listed
-example transformation as a guide.  The second line simply
-extracts the get direction transformation from the lens,
+\sam{OK I suppose it's a trick though I don't see an obvious solution. }
+
+The first line above synthesizes a lens between \codefont{comma\_name} and
+\codefont{space\_name} using the listed example transformation as a guide.  The
+second line simply extracts the get direction transformation from the lens,
 which is what we need for \kw{squash}.
@@
 \begin{lstlisting}
-let bib_fields = perm (label, bib_author, bib_title) with (project ("," .wsp$^+$) $\to$ ",\n")
+let bib_fields = perm (label, bib_author, bib_title) with (project ("," .wsp+) $\to$ ",\n")
 \end{lstlisting}
 
-The final QRE combinator is the normalize combinator.
-This combinator allows a programmer to manually define a function
+The final QRE combinator is the normalize combinator, which is also implemented
+in Boomerang. This combinator allows a programmer to manually define a function
 $f$ which sends each string that matches a regular expression $R$ to some
@@
 (* allow any permutation of fields interspersed with arbitrary whitespace *)
-let bib_fields = perm (label, bib_author, bib_title) with (project (wsp+ . "," . wsp+) $\to$ ",\n")
+let bib_fields = perm (label, bib_author, bib_title) with (project ("," . wsp+) $\to$ ",\n")
 let bibtex = "@book{" . bib_fields . "}"
@@
 \subsection{QRE Lenses and QRE Lens Synthesis}
+\label{sec:examplesynth}
 At this point we already have a tool for synthesizing bijective lenses from a
@@
 \caption{Quotient Lens with ``Canonizers at the Edges''
-\sam{We don't need a general choose function. With QREs the kernel language is
-always contained in the whole language so we just use the inclusion as the
-choose function.}\bcp{Done.  But I wonder if the picture should mention QREs
-explicitly, somehow... any ideas?}}\sam{Maybe we can put all structures, the
-canonizer and the canonical strcutre in a box that says QRE?}
+\bcp{Maybe we should go a step further and label things like $K(c)$,
+ $W(c)$, Canonizer(c), etc.  And probably we should display K as a subset of
+ W.  I can make these changes later.)}}
 \label{fig:attheedges}
 \end{figure}
-Every QRE lens $q$ has a type $q: c_1 \Leftrightarrow c_2$ where $c_1$ and $c_2$
-are QREs. In the get direction, a QRE lens $q: c_1 \Leftrightarrow c_2$ uses the
+Every QRE lens $q_1$ has a type $q_1: c_1 \Leftrightarrow c_2$ where $c_1$ and $c_2$
+are QREs. In the get direction, a QRE lens $q_1: c_1 \Leftrightarrow c_2$ uses the
 source QRE $c_1$ to compute a unique representative for the data modulo the
 equivalence relation defined by $c_1$ and then applies the $\get$ function of a
-bijective lens $\ell$ to this representative. In the backward direction, $q$
+bijective lens $\ell_1$ to this representative. In the backward direction, $q_1$
 operates similarly, but using the view QRE $c_2$ and the $\lput$ function of
-$\ell$.
+$\ell_1$.
 
@@
 canonization, and because the algorithm for synthesizing bijective lenses is
-directed by these formats, $c_1$ and $c_1$ are all that is required to
+directed by these formats, $c_1$ and $c_2$ are all that is required to
 synthesize QRE lenses end-to-end. Indeed, we structured QRE lenses as bijective
@@
 The \textit{squash} and \textit{permutation} combinators give the two
-interesting definitions. If $c = \squash{c}{c_1}{f}$, then the whole language of
-$c$ is $W(c) \sep W(c_1)$. This is because the \textit{squash} combinator takes
-the whole language $W(c)$ of $c$, the whole language $W(c_1)$ of $c_1$ and a
-function $f : \mathcal{L}(W(c)) \longrightarrow \mathcal{L}(W(c_1))$, maps
-$\mathcal{L}(W(c))$ into $\mathcal{L}(W(c_1))$ using $f$ and then canonizes
-$W(c_1)$ into $K(c_1)$ using $\canonizer(c_1)$.
+interesting definitions. If $c = \squash{c_1}{c_2}{f}$, then the whole language
+of $c$ is $W(c_1) \sep W(c_2)$. This is because the \textit{squash} combinator
+takes the whole language $W(c_1)$ of $c_1$, the whole language $W(c_2)$ of $c_2$
+and a function $f : \mathcal{L}(W(c_1)) \longrightarrow \mathcal{L}(W(c_2))$,
+maps $\mathcal{L}(W(c_1))$ into $\mathcal{L}(W(c_2))$ using $f$ and then
+canonizes $W(c_2)$ into $K(c_2)$ using $\canonizer(c_2)$.
 
@@
 \end{figure}
-\subsection{Ambiguity and Well Formed QREs}
-Before giving the inference rules for deriving well formed QREs, we address
-the issue of {\em ambiguity}, which appears in the premises of the inference
-rules for the regular combinators. If $c$ and $c_1$ are QREs, then when applying
-the regular combinators to QREs, we require that if a string $s$ matches any of
-the regular expressions $W(c) \cdot W(c_1)$, \quad $W(c) \sep W(c_1)$,
-\quad $W(c)^*$, \quad $K(c) \cdot K(c_1)$, \quad $K(c) \sep K(c_1)$,
-\quad $K(c)^*$, then $s$ matches that regular expression in only one way.
+\subsection{Well Formed QREs}
+\subsubsection{Completeness of {\em normalize}}
+Before we present the full inference rules, for QREs, we discuss the {\em
+normalize} QRE which is special in two main ways, the first of which has to do
+with the premises in its inference rule:
+\begin{prooftree}
+\AxiomC{$\mathcal{L}(R') \subseteq \mathcal{L}(R)$}
+\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(R')$}
+\AxiomC{$f$ is surjective}
+\AxiomC{$f = f^2$}
+\RightLabel{(Normalize)}
+\QuaternaryInfC{$\wf{\normalize{R}{R'}{f}}$}
+\end{prooftree}
+In order for the {\em normalize} QRE to be well-formed, then canonizing function
+$f$ used by this combinator must be surjective and idempotent. Unfortunately,
+verifying that a function satisfies these properties is in general undecidable.
+Nonetheless, we allow a programmer to use normalize at their own risk without
+checking these side conditions.
+
+We do this firstly because the {\em normalize} combinator allows one to express
+equivalence relations that are either difficult to define, or that cannot be
+defined using the other combinators. Secondly, the {\em normalize} canonizer
+enjoys the property that each of the other combinators $c$ is semantically
+equivalent to $\normalize{W(c)}{K(c)}{f}$ for some surjective, idempotent
+function $f : \mathcal{L}(W(c)) \longrightarrow \mathcal{L}(K(c))$; this is the
+second property that distinguishes the {\em normalize} combinator from the other
+QREs. The {\em normalize} combinator hence gives a sufficient condition for
+potential QREs to be ``well-behaved'', particularly with respect to the
+strategy for synthesizing quotient lenses which we shall introduce in
+Section~\ref{synth}. This is the
+\subsubsection{Ambiguity}
+Another issue which we address before presenting the full set of inference
+rules for QREs, we discuss the issue of {\em ambiguity}, which appears in the
+premises of the inference rules for the regular combinators. If $c_1$ and $c_2$
+are QREs, then when applying the regular combinators to QREs, we require that
+if a string $s$ matches any of the regular expressions $W(c_1) \cdot W(c_2)$,
+\quad $W(c_1) \sep W(c_2)$, \quad $W(c_1)^*$, \quad $K(c_1) \cdot K(c_2)$,
+\quad $K(c_1) \sep K(c_2)$, \quad $K(c_1)^*$, then $s$ matches that regular
+expression in only one way.
 
@@
 \AxiomC{$R$ is strongly unambiguous}
+\RightLabel{(Id)}
 \UnaryInfC{$\wf{\mathit{id}(R)}$}
@@
 \AxiomC{$s \in \mathcal{L}(R)$}
+\RightLabel{(Project)}
 \UnaryInfC{$\wf{\project{R}{s}}$}
@@
 \AxiomC{$\wf{c_i, c}$}
+\RightLabel{(Perm)}
 \AxiomC{${\substack{\forall \sigma \neq \theta, \; W(c_{\sigma(1)}) \cdot W(c)
@@
 \AxiomC{$f : \mathcal{L}(W(c_1)) \longrightarrow \mathcal{L}(W(c_2))$}
+\RightLabel{(Squash)}
 \TrinaryInfC{$\wf{\squash{c_1}{c_2}{f}}$}
@@
 \AxiomC{$f = f^2$}
+\RightLabel{(Normalize)}
 \QuaternaryInfC{$\wf{\normalize{R}{R'}{f}}$}
@@
 \AxiomC{$K(c_1) = W(c_2)$}
-\BinaryInfC{$\wf{c_2 \; ; \; c_1}$}
+\RightLabel{(Compose)}
+\BinaryInfC{$\wf{c_1 \; ; \; c_2}$}
 \DisplayProof
@@
 \AxiomC{$K(c)^{*!}$}
+\RightLabel{(Star)}
 \TrinaryInfC{$\wf{c^*}$}
@@
 \AxiomC{$K(c_1) \cdot^! K(c_2)$}
+\RightLabel{(Concat)}
 \TrinaryInfC{$\wf{c_1 \cdot c_2}$}
@@
 \AxiomC{$W(c_1) \cap W(c_2) = \varnothing$}
-\BinaryInfC{$\wf{c_1 \cdot c_2}$}
+\RightLabel{(Union)}
+\BinaryInfC{$\wf{c_1 \sep c_2}$}
 \DisplayProof
-\hskip 1.5em
-\AxiomC{$\wf{c}$}
-\AxiomC{$W(c)^{*!}$}
-\AxiomC{$K(c)^{*!}$}
-\TrinaryInfC{$\wf{c^*}$}
-\DisplayProof
+\
 \end{center}
 \end{center}
-\caption{Well-formed QREs
-}
+\caption{Well-formed QREs}
 \label{fig:qrerules}
@@
 As we already mentioned, the unambiguity conditions come into play when defining
-QREs using the regular combinators. For example, the inferrence rule for
-concatenation is as follows:\bcp{Rather than repeating rules inline, give
-  them names in the figure.}
-\begin{prooftree}
-\AxiomC{$\wf{c, c_1}$}
-\AxiomC{$W(c) \cdot^! W(c_1)$}
-\AxiomC{$K(c) \cdot^! K(c_1)$}
-\TrinaryInfC{$\wf{c \cdot c_1}$}
-\end{prooftree}
-
-\noindent This rule says that the concatenation $c \cdot c_1$ of QREs $c$ and
-$c_1$ is well formed only if the concatentions of $W(c)$ and $W(c_1)$, and
-$K(c)$ and $K(c_1)$ are unambiguous.
-
-The most complicated inferrence rule is the rule for the \textit{permutation}
-combinator:
-
-\begin{prooftree}
-\AxiomC{$\wf{c_i, c}$}
-\AxiomC{${\substack{\forall \sigma \neq \theta, \; W(c_{\sigma(1)}) \cdot W(c)
-\cdot \ldots \cdot W(c_{\sigma(n)})\\ \cap W(c_{\theta(1)}) \cdot W(c)
-\cdot \ldots \cdot W(c_{\theta(n)}) =\varnothing}}$}
-\AxiomC{$K(c_1) \cdot^! K(c) \cdot^! \ldots \cdot^! K(c) \cdot^! K(c_n)$}
-\TrinaryInfC{$\wf{\perm{c_1, \ldots, c_n}{c}}$}
-\end{prooftree}
-
-\noindent The second hypothesis for the \textit{permutation} rule says that for
-any two different permutations $\sigma$ and $\theta$, the languages
-$W(c_{\sigma(1)}) \cdot W(c) \cdot \ldots \cdot W(c_{\sigma(n)})$ and
-$W(c_{\theta(1)}) \cdot W(c) \cdot \ldots \cdot W(c_{\theta(n)})$ must be
-disjoint. This is because any string that matches the \textit{permutation}
-combinator matches the language $W(c_{\theta(1)}) \cdot W(c) \cdot \ldots \cdot
-W(c_{\theta(n)})$ for some permutation $\theta$, so we require that these
-regular expressions be disjoint.
+QREs using the regular combinators. For example, the (Concat) inferrence rule
+for concatenation says that the concatenation $c_1 \cdot c_2$ of QREs $c_1$ and
+$c_2$ is well formed only if the concatentions of $W(c_1)$ and $W(c_2)$, and
+$K(c_1)$ and $K(c_2)$ are unambiguous.
+
+The most complicated inference rule is the (Perm) rule for the
+\textit{permutation} combinator. The second hypothesis for the
+\textit{permutation} rule says that for any two different permutations $\sigma$
+and $\theta$, the languages $W(c_{\sigma(1)}) \cdot W(c) \cdot \ldots \cdot
+W(c_{\sigma(n)})$ and $W(c_{\theta(1)}) \cdot W(c) \cdot \ldots \cdot
+W(c_{\theta(n)})$ must be disjoint. This is because any string that matches the
+\textit{permutation} combinator matches the language $W(c_{\theta(1)}) \cdot
+W(c) \cdot \ldots \cdot W(c_{\theta(n)})$ for some permutation $\theta$, so we
+require that these regular expressions be disjoint.
 
@@
 
-At this point we also bring to attention the fact that in order for the {\em
-normalize} QRE to be well-formed, then canonizing function $f$ used by this
-combinator must be surjective and idempotent. Unfortunately, verifying
-that a function satisfies these properties is in general undecidable.
-Nonetheless, we allow a programmer to use normalize at their own risk without
-checking these side conditions.
-
-We do this firstly because the {\em normalize} combinator allows one to express
-equivalence relations that are either difficult to define, or that cannot be
-defined using the other combinators. Secondly, the {\em normalize} canonizer
-enjoys the property that each of the other combinators $c$ is semantically
-equivalent to $\normalize{W(c)}{K(c)}{f}$ for some surjective, idempotent
-function $f : \mathcal{L}(W(c)) \longrightarrow \mathcal{L}(K(c))$. The {\em
-normalize} combinator hence gives a sufficient condition for potential QREs to
-be ``well-behaved'', particularly with respect to our strategy for synthesizing
-quotient lenses which we shall introduce in Section~\ref{synth}.
-
-In spite of this ``completeness'' property of the {\em normalize} combinator,
-our experience with QREs shows that using the {\em project}, {\em squash} and
-{\em perm} combinators as well as the regular operators applied to these is
-much easier than using just {\em normalize} because of the compositional nature
-of the other combinators.
-
-The formal semantics of QREs satisfy the following theorem:
-\begin{theorem}
-If $c$ is a well formed QRE, then
-\begin{enumerate}
-  \item $W(c)$ and $K(c)$ are well defined regular expressions,
-  \item  $\eqrel{c}$ is an equivalence relation on $\mathcal{L}(W(c))$,
-  \item  $\mathcal{L}(K(c))$ forms a complete set of representatives for
-  $\eqrel{c}$,
-  \item $\canonizer(c):\mathcal{L}(W(c)) \longrightarrow \mathcal{L}(K(c))$ is
-  a well-defined function, and
-  \item  given any $w \in \mathcal{L}(W(c))$, $\canonizer(c)(w)$ is the unique
-  $k$ in $\mathcal{L}(K(c))$ such that $k$ is equivalent to $w$ modulo
-  $\eqrel{c}$.
-\end{enumerate}
-\end{theorem}
-
 \section{QRE Lenses}
@@
 directly on regular languages.  QREs are therefore a good \textit{specification}
-language for quotient lenses. In this section of the paper, we introduce
+language for quotient lenses. We next introduce
 \textit{QRE Lenses}, a class of quotient lenses specified using QREs.
+
+\bcp{A question about variable names: If we use $R$ and $S$ for plain
+  regular expressions, it seems inconsistent to use $c$ for QREs.  (And even
+  more inconsistent to use $c$ and $c'$ or $c$ and $c_1$!) I suggest $P$ and
+  $Q$ for QREs (with $Q$ as the default when we are only talking about
+  one).}\sam{I made the notation consistent by just using $c_1$ and $c_2$. I
+  personally feel that it is better not to use P and Q for QREs since we use p
+  and q for QRE lenses, but if you disagree I will make the change.}
 
@@
 $\sim_R$ and $\sim_S$ defined on $\mathcal{L}(R)$ and $\mathcal{L}(S)$
-respectively, we define a {\em bijective quotient lens} $q : R /{\sim_R}
-\Leftrightarrow S/{\sim_S}$ to be a pair of functions $q.get :
-\mathcal{L}(R) \longrightarrow \mathcal{L}(S)$ and $q.put : \mathcal{L}(S)
+respectively, we\bcp{can we clarify that this is exactly Foster et al's
+definition?}\sam{this is not Foster et al's definition. It's just a
+loosening of the definition for bijective lens used in the Optician paper}define
+a {\em bijective quotient lens} $q :
+R /{\sim_R} \Leftrightarrow S/{\sim_S}$ to be a pair of functions $q.\get :
+\mathcal{L}(R) \longrightarrow \mathcal{L}(S)$ and $q.\lput : \mathcal{L}(S)
 \longrightarrow \mathcal{L}(R)$ such that for all $r \in \mathcal{L}(R)$ and $s
-\in \mathcal{L}(S)$,
-\begin{align*}
-q.put(q.get(r)) &\sim_R r \text{, and}\\
-q.get(q.put(s)) &\sim_R s
-\end{align*}
-Moreover, if $r \sim_R r'$, then $q.get(r) = q.get(r')$, and if $s \sim_S s'$,
-then $q.put(s) = q.put(s')$.
+\in \mathcal{L}(S)$, then $q.\lput(q_1.\get(r)) \sim_R r$ and
+$q.\get(q_1.\lput(s)) \sim_R s$. Moreover, if $r \sim_R r'$, then $q.\get(r) =
+q.\get(r')$, and if $s \sim_S s'$, then $q.\lput(s) = q.\lput(s')$.
 
 In words, a bijective quotient lens $q$ from $R$ to $S$ modulo $\sim_R$
-and $\sim_S$ is a pair of functions $q.get$ and $q.put$ such that $q.get$
-respects $\sim_R$ and $q.put$ respects $\sim_S$, and such that the respective
-functions defined on $\mathcal{L}(R)/{\sim_R}$ and $\mathcal{L}(S)/{\sim_S}$ are
-mutual inverses of each other.
-
-Now let $c$ and $c_1$ be QREs. Given a bijection $\ell : K(c) \Leftrightarrow
-K(c_1)$, we may define a quotient lens $q : W(c)/\eqrel{c} \Leftrightarrow
-W(c_1)/\eqrel{c_1}$ by
-\begin{equation}\label{normalform}
-q.get = \ell \circ \canonizer(c) \text{ and } q.put = \ell^{-1} \circ
-\canonizer(c_1),
-\end{equation}
-
-\noindent In other words, the get function of the $q$ first canonizes the source
-data using $\canonizer(c)$ and then applies $f$ to the result, while the
-put function of the $q$ first canonizes the target data using
-$\canonizer(c_1)$ and the applies $f^{-1}$ to the result. This is our approach in
-defining QRE lenses.
+and $\sim_S$ is a pair of functions $q.\get$ and $q.\lput$ such that
+$q.\get$ respects $\sim_R$ and $q.\lput$ respects $\sim_S$, and such that
+the {\em get} and {\em put} functions lifted to $\mathcal{L}(R)/{\sim_R}$ and
+$\mathcal{L}(S)/{\sim_S}$ are mutual inverses.
+
+These laws are similar to the laws for bijective lenses, except that the
+equality restriction in the bijective lens laws are loosened to allow for
+equivalence relations. Also, the condition that $r \sim_R r'$ implies
+$q.\get(r) = q.\get(r')$ ensures that the {\em get} function induced on
+the equivalence classes if $\equiv_c$ is well-defined, and similarly for
+the {\em put} function.
+
+Now let $c_1$ and $c_2$ be QREs. Given a bijection $\ell : K(c_1)
+\Leftrightarrow K(c_2)$, we may define a quotient lens $q : W(c_1)/\eqrel{c_1}
+\Leftrightarrow W(c_2)/\eqrel{c_2}$ by $q.\get = \ell \circ \canonizer(c_1)
+$ and $q.\lput = \ell_1^{-1} \circ \canonizer(c_2)$. In other words, the get
+function of the $q$ first canonizes the source data using $\canonizer(c_1)$
+and then applies $f$ to the result, while the put function of the $q$ first
+canonizes the target data using $\canonizer(c_2)$ and the applies $f^{-1}$ to
+the result. This is our approach in defining QRE lenses.
 
@@
 we define on them may not preserve this normal form. For example, the
-composition $q' \; ; \; q$ of QRE lenses $q$ and $q'$ may not be equivalent to
-any quotient lens that has the canonical normal form of a bijective lens with
-canonizers a the edges. Consequently, in order to guarantee that every QRE
-lens $q$ is equivalent to a QRE lens $q'$ which has the canonical form of
-a bijection with canonizers at the edges, we need to identify a set $S$ of
-bijections that will ensure that QRE lenses are closed under the action of QRE
-lens combinators. Our choice is to use the class of \textit{bijective lenses},
-which we briefly discuss in the next section.
+composition $q_1 \; ; \; q_2$ of QRE lenses $q_1$ and $q_2$ may not be
+equivalent to any quotient lens that has the canonical normal form of a
+bijective lens with canonizers a the edges. Consequently, in order to guarantee
+that every QRE lens $q_1$ is equivalent to a QRE lens $q_2$ which has this
+canonical form, we need to identify a set $S$ of bijections that will ensure
+that all QRE lenses is functionally equivalent to a QRE lens in canonical form.
+Our choice is to use the class of \textit{bijective lenses}, which we studied in
+\cite{optician} and has this desired property. The following part of the paper
+briefly discusses bijective lenses.
 
@@
 The syntax for the language of bijective lenses is given by
-$$\ell := \mathit{id} \; (R) \sep const(s, s') \sep  swap(\ell,
-\ell) \sep \ell \cdot \ell' \; |  \; (\ell \sep \ell') \sep \ell^* \;
-| \; \ell \circ \ell',$$ where $R$ ranges over regular expressions and $s$
+$$\ell_1 := \mathit{id} \; (R) \sep \const(s_1, s_2) \sep  \swap(\ell_1,
+\ell_2) \sep \ell_1 \cdot \ell_2 \; |  \; (\ell_1 \sep \ell_2) \sep \ell^* \;
+| \; \ell_1 \; ; \;  \ell_2,$$ where $R$ ranges over regular expressions and $s$
 ranges over character strings.
 
-The denotation of a lens $\ell$ is $\llbracket \ell \rrbracket \subseteq
-\mathit{String} \times \mathit{String}$. If $(s_1, s_2) \in \llbracket \ell
-\rrbracket$, then $\ell$ maps between $s_1$ and $s_2$.
+The denotation of a lens $\ell_1$ is $\llbracket \ell_1 \rrbracket \subseteq
+\mathit{String} \times \mathit{String}$. If $(s_1, s_2) \in \llbracket \ell_1
+\rrbracket$, then $\ell_1$ maps between $s_1$ and $s_2$.
 \begin{align*}
 \llbracket R \rrbracket &= \{(r, r) \sep r \in \mathcal{L}(R)\}\\
-\llbracket const(r, s) \rrbracket &= \{(r, s)\}\\
-\llbracket swap(\ell, \ell') \rrbracket &= \{(s \cdot t, t' \cdot s') \sep
-(s, s') \in \llbracket \ell \rrbracket \text{ and } (t, t') \in \llbracket
-\ell' \rrbracket\}\\
-\llbracket \ell \cdot \ell' \rrbracket &= \{(s \cdot t, s' \cdot t) \sep
-(s, s') \in \llbracket \ell \rrbracket \text{ and } (t, t') \in \llbracket
-\ell' \rrbracket\}\\
-\llbracket \ell \sep \ell' \rrbracket &= \{(s \cdot t) \sep
-(s, t) \in \llbracket \ell \rrbracket \text{ or } (s, t) \in \llbracket
-\ell' \rrbracket\}\\
+\llbracket \const(r, s) \rrbracket &= \{(r, s)\}\\
+\llbracket \swap(\ell_1, \ell_2) \rrbracket &= \{(s \cdot t, t' \cdot s') \sep
+(s, s') \in \llbracket \ell_1 \rrbracket \text{ and } (t, t') \in \llbracket
+\ell_2 \rrbracket\}\\
+\llbracket \ell_1 \cdot \ell_2 \rrbracket &= \{(s \cdot t, s' \cdot t) \sep
+(s, s') \in \llbracket \ell_1 \rrbracket \text{ and } (t, t') \in \llbracket
+\ell_2 \rrbracket\}\\
+\llbracket \ell_1 \sep \ell_2 \rrbracket &= \{(s \cdot t) \sep
+(s, t) \in \llbracket \ell_1 \rrbracket \text{ or } (s, t) \in \llbracket
+\ell_2 \rrbracket\}\\
 \llbracket \ell^* \rrbracket &= \{(s_1 \cdot \ldots \cdot s_n, t_1 \cdot \ldots
@@
 \leq i \leq n\}\\
-\llbracket \ell_2 \circ \ell_1 \rrbracket &= \{(r, t) \sep \text{there exists }s
+\llbracket \ell_1 \; ; \; \ell_2 \rrbracket &= \{(r, t) \sep \text{there exists }s
 \text{ such that } (r, s) \in \llbracket \ell_1 \rrbracket \text{ and } (s, t)
@@
 \end{align*}
-The $\mathit{const}(s, t)$ lens replaces the string $s$ with $t$ in the source
+The $\mathit{\const}(s, t)$ lens replaces the string $s$ with $t$ in the source
 data in the forward direction, and $t$ with $s$ in the target data in the backward
 direction. $\mathit{id}(R)$ applies the identity function to the source and target
-in $\mathcal{L}(R)$ in both directions. The composition lens $\ell' \circ \ell$
-applies $\ell$ followed by then $\ell'$ to the source in the forward direction,
-and applies $\ell'$ followed by $\ell$ to the target in the backward direction.
-The lens $\ell \cdot \ell'$ first splits the string $s$ into $s_1$ and $s_2$,
-applies $\ell$ and $\ell'$ to $s_1$ and $s_2$ to get $t_1$ and $t_2$
+in $\mathcal{L}(R)$ in both directions. The composition lens $\ell_1 \; ; \; \ell_2$
+applies $\ell_1$ followed by then $\ell_2$ to the source in the forward direction,
+and applies $\ell_2$ followed by $\ell_1$ to the target in the backward direction.
+The lens $\ell_1 \cdot \ell_2$ first splits the string $s$ into $s_1$ and $s_2$,
+applies $\ell_1$ and $\ell_2$ to $s_1$ and $s_2$ to get $t_1$ and $t_2$
 respectively, then concatenates $t_1$ and $t_2$ and returns $t_1 \cdot t_2$ as
-the final result in the forward direction. $\ell \cdot \ell'$ operates
+the final result in the forward direction. $\ell_1 \cdot \ell_2$ operates
 similarly in the backward direction, but with $s, s_1$ and $s_2$ substituted
-for $t, t_1$ and $t_2$. The $\mathit{swap} \; (\ell, \ell')$ lens operates
-like $\ell \cdot \ell'$, except that it swaps $t_1$ and $t_2$ before
+for $t, t_1$ and $t_2$. The $\mathit{\swap} \; (\ell_1, \ell_2)$ lens operates
+like $\ell_1 \cdot \ell_2$, except that it swaps $t_1$ and $t_2$ before
 concatenating the two for a final result of $t_2 \cdot t_1$ in the forward
-direction. In the backward direction, $\mathit{swap}(\ell, \ell')$ first undoes
-the swap, then proceeds as expected. The $\ell \; | \; \ell'$ lens
-chooses to apply $\ell$ or $\ell'$ depending on whether the source
-(resp. target) data is matched by $\ell$ or $\ell'$ in the forward (resp.
+direction. In the backward direction, $\mathit{\swap}(\ell_1, \ell_2)$ first undoes
+the \swap, then proceeds as expected. The $\ell_1 \; | \; \ell_2$ lens
+chooses to apply $\ell_1$ or $\ell_2$ depending on whether the source
+(resp. target) data is matched by $\ell_1$ or $\ell_2$ in the forward (resp.
 backward direction). The $\ell^*$ lens splits the string $s$ into strings $s_1,
-\ldots, s_n$, applies $\ell$ to each $s_i$ to get $t_i$, and then concatenates
+\ldots, s_n$, applies $\ell_1$ to each $s_i$ to get $t_i$, and then concatenates
 each of the $t_i$'s for a final result of $t_1 \cdot \ldots \cdot t_n$ in the
@@
 \begin{prooftree}
-\AxiomC{$\ell : R \Leftrightarrow S$}
+\AxiomC{$\ell_1 : R \Leftrightarrow S$}
 \AxiomC{$R \equiv^s R'$}
 \AxiomC{$S \equiv^s S'$}
-\TrinaryInfC{$\ell : R' \Leftrightarrow S'$}
+\TrinaryInfC{$\ell_1 : R' \Leftrightarrow S'$}
 \end{prooftree}
@@
 \AxiomC{$s_2 \in \Sigma^*$}
-\BinaryInfC{$\mathit{const} \; (s_1, s_t): s_1 \Leftrightarrow s_2$}
+\BinaryInfC{$\mathit{\const} \; (s_1, s_t): s_1 \Leftrightarrow s_2$}
 \DisplayProof
@@
 \AxiomC{$\substack{R_1 \cdot^! R_2 \\ S_2 \cdot^! S_1}$}
-\BinaryInfC{$\mathit{swap} \; (\ell_1, \ell_2): R_1 \cdot R_2
+\BinaryInfC{$\mathit{\swap} \; (\ell_1, \ell_2): R_1 \cdot R_2
 \Leftrightarrow S_2 \cdot S_1$}
@@
 \AxiomC{$\ell_2 : R_2 \Leftrightarrow R_3$}
-\BinaryInfC{$\ell_2 \circ \ell_1: R_1 \Leftrightarrow R_3$}
+\BinaryInfC{$\ell_1 \; ; \; \ell_2: R_1 \Leftrightarrow R_3$}
 \DisplayProof
@@
 $$ q := \kw{lift}(\ell) \sep \kw{lquot}(q, \ell) \sep
-\kw{rquot}(\ell, q) \sep q \cdot q' \sep (q \sep q') \sep q^* \sep q \circ q',$$
-where $\ell$ ranges over bijective lenses.
+\kw{rquot}(\ell, q) \sep q_1 \cdot q_2 \sep (q_1 \sep q_2) \sep q^* \sep
+q_1 \; ; \; q_2,$$ where $\ell$ ranges over bijective lenses.
 
@@
 The difference between QRE lenses and Boomerang quotient lenses lies in the
-typing judgements. More concretely, each QRE lens $q$ has a type $q : c
+typing judgements. More concretely, each QRE lens $q_1$ has a type $q_1 : c
 \Leftrightarrow c_1$ where $c, c_1$ are QREs. In contrast, the approach used to
@@
 
-Each QRE lens $q$ has a type $q : c \Leftrightarrow c_1$ where $c, c_1$ are QREs.
-If $q : c \Leftrightarrow c_1$, then the source format is described by $W(c)$ and
-the canonical set of representative for the source data is described by $K(c)$.
-Similarly, the target format is described by $W(c_1)$ and the canonical set of
-representatives for the target data is described by $K(c_1)$. The underlying lens
-of $q$ is a bijective lens $\ell : K(c) \Leftrightarrow K(c_1)$.
-
-The denotation $\llbracket q \rrbracket$ of a QRE lens $q:c \Leftrightarrow c_1$
-is a quotient lens $\llbracket q \rrbracket : W(c)/{\eqrel{c}}
-\Longleftrightarrow W(c_1)/{\eqrel{c_1}}$. The typing rules and denotation of QRE
-lenses are given in Figure~\ref{fig:qlenssemantics}. The trickiest typing rule
-is the typing rule for composition:
+Each QRE lens $q$ has a type $q : c_1 \Leftrightarrow c_2$ where $c_1, c_2$ are
+QREs. If $q : c_1 \Leftrightarrow c_2$, then the source format is described by
+$W(c_1)$ and the canonical set of representative for the source data is
+described by $K(c_1)$.
+Similarly, the target format is described by $W(c_2)$ and the canonical set of
+representatives for the target data is described by $K(c_2)$. The underlying
+lens of $q$ is a bijective lens $\ell : K(c_1) \Leftrightarrow K(c_2)$.
+
+The denotation $\llbracket q \rrbracket$ of a QRE lens $q : c_1 \Leftrightarrow
+c_2$ is a quotient lens $\llbracket q \rrbracket : W(c_1)/{\eqrel{c_1}}
+\Longleftrightarrow W(c_2)/{\eqrel{c_2}}$. The typing rules and denotation of
+QRE lenses are given in Figure~\ref{fig:qlenssemantics}.
+\begin{figure}[ht]
+\centering
+\[
+\begin{array}[b]{l@{\qquad}l}
+\RuleSide{\ell : R \Leftrightarrow S}{\kw{lift}(\ell): id(R) \Leftrightarrow
+id(S)}{\mathit{(Lift)}} &
+\begin{array}[b]{rcl}
+\llbracket \kw{lift}(\ell) \rrbracket.\get &=&  \llbracket \ell \rrbracket\\
+\llbracket \kw{lift}(\ell) \rrbracket.\lput &=& \llbracket \ell \rrbracket^{-1}
+\end{array}
+\end{array}
+\]
+\[
+\begin{array}[b]{c@{\qquad}c}
+\RuleSide{q : c_2  \Leftrightarrow c_3 &
+\wf{c_1} &
+K(c_1) = W(c_2)}
+{\kw{lquot}(c_1, q): c_1 \; ; \; c_2 \Leftrightarrow c_3}{\mathit(Lquot)} &
+\begin{array}[b]{rcl}
+\llbracket \kw{lquot}(c_1, q) \rrbracket.\get  &=& \llbracket q
+\rrbracket.\get \circ \canonizer(c_1)\\
+\llbracket \kw{lquot}(c_1, q) \rrbracket.\lput &=& \llbracket q
+\rrbracket.\lput
+\end{array}
+\end{array}
+\]
+
+\[
+\begin{array}[b]{c@{\qquad}c}
+\RuleSide{q : c_1 \Leftrightarrow c_3 & \wf{c_3} & W(c_3) = K(c_2)}
+{\kw{rquot}(q, c_1):c_1 \Leftrightarrow c_2 \; ; \; c_3}{\mathit{(Rquot)}} &
+\begin{array}[b]{rcl}
+\llbracket \kw{rquot}(q, c_1) \rrbracket.\get  &=& \llbracket q
+\rrbracket.\get\\
+\llbracket \kw{rquot}(q, c_1) \rrbracket.\lput &=& \llbracket q
+\rrbracket.\lput \circ \canonizer(c_3)
+\end{array}
+\end{array}
+\]
+
+\[
+\begin{array}[b]{l@{\qquad}c}
+\RuleSide{q : c_1 \Leftrightarrow c_2 &
+W(c_1)^{*!},W(c_2)^{*!} & K(c_1)^{*!}, K(c_2)^{*!}}
+{{q}^* : {c_1}^* \Leftrightarrow {c_2}^*}{\mathit{(Star)}} &
+\begin{array}[b]{rcl}
+\llbracket {q}^* \rrbracket.\get  &=& (\llbracket q \rrbracket.\get)^*\\
+\llbracket {q}^* \rrbracket.\lput &=& (\llbracket q \rrbracket.\get)^*
+\end{array}
+\end{array}
+\]
+
+\[
+\begin{array}[b]{l@{\qquad}c}
+\RuleSide{\substack{q_1 : c_1 \Leftrightarrow d_1 \\ q_2 : c_2 \Leftrightarrow
+d_2} & {\substack{W(c_1) \cdot^! W(c_2)\\ K(c_1) \cdot^! K(c_2)}}
+& {\substack{W(d_1) \cdot^! W(d_2)\\ K(d_1) \cdot^! K(d_2)}}}
+{q_1 \cdot q_2: c_1 \cdot c_2 \Leftrightarrow d_1 \cdot d_2}{\mathit{(Concat)}}
+&
+\begin{array}[b]{rcl}
+\llbracket q_1 \cdot q_2 \rrbracket.\get  &=& \llbracket q_1 \rrbracket.\get
+\cdot \llbracket q_2 \rrbracket.\get\\
+\llbracket q_1 \cdot q_2\rrbracket.\lput &=& \llbracket q_1 \rrbracket.\lput
+\cdot \llbracket q_2 \rrbracket.\lput
+\end{array}
+\end{array}
+\]
+
+\[
+\begin{array}[b]{l@{\qquad}c}
+\RuleSide{\substack{q_1 : c_1 \Leftrightarrow d_1 \\ q_2 : c_2 \Leftrightarrow
+d_2} & \substack{\mathcal{L}(W(c_1)) \cap \mathcal{L}(W(c_2)) = \varnothing
+\\ \mathcal{L}(W(d_1)) \cap \mathcal{L}(W(d_2)) = \varnothing }}
+{q_1 \sep q_2: (c_1 \sep c_2)\Leftrightarrow (d_1 \sep d_2)}{\mathit{(Or)}} &
+\begin{array}[b]{rcl}
+\llbracket q_1 \sep q_2 \rrbracket.\get(s) =
+\begin{cases}
+\llbracket q_1 \rrbracket.\get (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
+\llbracket q_2 \rrbracket.\get (s) & \text{if } s \in \mathcal{L}(W(c_2))\\
+\end{cases}\\
+\llbracket q_1 \sep q_2 \rrbracket.\lput(s) =
+\begin{cases}
+\llbracket q_1 \rrbracket.\lput (s) & \text{if } s \in \mathcal{L}(W(d_1))\\
+\llbracket q_2 \rrbracket.\lput (s) & \text{if } s \in \mathcal{L}(W(d_2))\\
+\end{cases}
+\end{array}
+\end{array}
+\]
 
 \begin{prooftree}
-\AxiomC{$\substack{q_1 : c \Leftrightarrow c_1 \\ q_2 : c_2 \Leftrightarrow
-c_3}$}
-\AxiomC{$\substack{\mathcal{L}(W(c_1))= \mathcal{L}(W(c_2)) \\ K(c_1) \equiv
-K(c_2)}$}
-\AxiomC{$\canonizer(c_1) = \canonizer(c_2)$}
-\TrinaryInfC{$q_2 \circ q_1: c \Leftrightarrow c_3$}
+\AxiomC{$\substack{q_1 : c_1 \Leftrightarrow c_2 \\ q_2 : c_3 \Leftrightarrow
+c_4}$}
+\RightLabel{(Compose)}
+\AxiomC{$\substack{\mathcal{L}(W(c_2))= \mathcal{L}(W(c_3)) \\ K(c_2) \equiv^s
+K(c_3)}$}
+\AxiomC{$\canonizer(c_2) = \canonizer(c_3)$}
+\TrinaryInfC{$q_1 \; ; \; q_2: c_1 \Leftrightarrow c_4$}
 \end{prooftree}
-
-This derivation rule essentially says that the composition $q_2 \circ q_1$ is
-well defined if and only if the intermediary QREs $c_2$ and $c_3$ define the
-same equivalence relation on the same regular language. The condition
-$\mathcal{L}(W(c_1)) = \mathcal{L}(W(c_2))$ says that the intermediary language
-is the same on both sides, while the condition $K(c_1) \equiv K(c_2)$ says that
-the kernel regular expressions are syntactically the same. Finally, the
-condition $\canonizer(c_1) = \canonizer(c_2)$ says that $c_1$ and $c_2$ define
-the same equivalence relation on $\mathcal{L}(W(c_1))= \mathcal{L}(W(c_2))$.
-
-In general, checking that $\canonizer(c_1) = \canonizer(c_2)$ is an undecidable
-task, especially because of the $\normalize{R}{R'}{f}$ QRE. Therefore, our
-actual implementation of the composition rule for QRE lenses follows Boomerang's
-approach, which is to define the composition on QRE lenses if and only if
-$c_1$ and $c_2$ are syntactially equal to $id(R)$ and $id(R')$ for some $R$ and
-$R'$ satisfying $\mathcal{L}(R) = \mathcal{L}(R')$.
+\begin{align*}
+\llbracket q_1 \; ; \; q_2 \rrbracket.\get &= \llbracket q_2
+\rrbracket.\get\circ \llbracket q_1 \rrbracket.\get\\
+\llbracket q_1 \; ; \; q_2 \rrbracket.\lput &= \llbracket q_1 \rrbracket.\lput
+\circ \llbracket q_2 \rrbracket.\lput
+\end{align*}
+\caption{Denotation and Typing Rules for QRE Lenses}
+\label{fig:qlenssemantics}
+\end{figure}
+The trickiest typing rule is the typing rule for composition:
+
+\begin{prooftree}
+\AxiomC{$\substack{q_1 : c_1 \Leftrightarrow c_2 \\ q_2 : c_3 \Leftrightarrow
+c_4}$}
+\AxiomC{$\substack{\mathcal{L}(W(c_2))= \mathcal{L}(W(c_3)) \\ K(c_2) \equiv^s
+K(c_3)}$}
+\AxiomC{$\canonizer(c_2) = \canonizer(c_3)$}
+\TrinaryInfC{$q_1 \; ; \; q_2: c_1 \Leftrightarrow c_4$}
+\end{prooftree}
+
+This derivation rule essentially says that the composition $q_1 \; ; \; q_2$ is
+well defined if and only if the intermediary QREs $c_3$ and $c_4$ define the
+same equivalence relation on regular languages that are equivalent modulo the
+star-semiring axioms. The condition $\mathcal{L}(W(c_2)) = \mathcal{L}(W(c_3))$
+says that the intermediary language is the same on both sides, while the
+condition $K(c_2) \equiv^s K(c_3)$ says that the kernel regular expressions are
+equivalent modulo the star-semiring axioms. Finally, the condition
+$\canonizer(c_2) = \canonizer(c_3)$ says that $c_2$ and $c_3$ define the same
+equivalence relation on $\mathcal{L}(W(c_2))= \mathcal{L}(W(c_3))$.
+
+In general, checking that $\canonizer(c_2) = \canonizer(c_3)$ is undecidable
+because of the {\em normalize} QRE which allows uses an arbitrary
+idempotent function as the canonizer. Therefore, our actual implementation of
+the composition rule for QRE lenses follows Boomerang's approach, which is to
+define the composition on QRE lenses if and only if $c_2$ and $c_3$ are
+syntactially equal to $id(R)$ and $id(R')$ for some $R$ and $R'$ satisfying
+$\mathcal{L}(R) = \mathcal{L}(R')$.
 
@@
 \begin{theorem}
-If there is a derivation $q : c \Leftrightarrow c_1$, then $\llbracket q
-\rrbracket : W(c)/{\eqrel{c}} \Leftrightarrow W(c_1)/{\eqrel{c_1}}$ is a
+If there is a derivation $q : c_1 \Leftrightarrow c_2$, then $\llbracket q
+\rrbracket : W(c_1)/{\eqrel{c_1}} \Leftrightarrow W(c_2)/{\eqrel{c_2}}$ is a
 well-defined quotient lens.
 \end{theorem}
-
-\begin{figure}[ht]
-\centering
-\[
-\begin{array}[b]{l@{\qquad}l}
-\Rule{\ell : R \Leftrightarrow S}{\kw{lift}(\ell): id(R) \Leftrightarrow id(S)}
-&
-\begin{array}[b]{rcl}
-\llbracket \kw{lift}(\ell) \rrbracket.get &=&  \llbracket \ell \rrbracket\\
-\llbracket \kw{lift}(\ell) \rrbracket.put &=& \llbracket \ell \rrbracket^{-1}
-\end{array}
-\end{array}
-\]
-\[
-\begin{array}[b]{c@{\qquad}c}
-\Rule{q : c_1  \Leftrightarrow c_2 &
-\wf{c} &
-K(c) = W(c_1)}
-{\kw{lquot}(c, q): c_1 \circ c \Leftrightarrow c_2} &
-\begin{array}[b]{rcl}
-\llbracket \kw{lquot}(c, q) \rrbracket.get  &=& \llbracket q
-\rrbracket.get \circ \canonizer(c)\\
-\llbracket \kw{lquot}(c, q) \rrbracket.put &=& \llbracket q \rrbracket.put
-\end{array}
-\end{array}
-\]
-
-\[
-\begin{array}[b]{c@{\qquad}c}
-\Rule{q : c \Leftrightarrow c_2 &
-\wf{c_2} &
-W(c_2) = K(c_1)}
-{\kw{rquot}(q, c):c \Leftrightarrow c_2 \circ c_1} &
-\begin{array}[b]{rcl}
-\llbracket \kw{rquot}(q, c) \rrbracket.get  &=& \llbracket q
-\rrbracket.get\\
-\llbracket \kw{rquot}(q, c) \rrbracket.put &=& \llbracket q
-\rrbracket.put \circ \canonizer(c_2)
-\end{array}
-\end{array}
-\]
-
-\[
-\begin{array}[b]{l@{\qquad}c}
-\Rule{q : c \Leftrightarrow c_1 &
-W(c)^{*!},W(c_1)^{*!} & K(c)^{*!}, K(c_1)^{*!}}
-{{q}^* : c^* \Leftrightarrow {c_1}^*} &
-\begin{array}[b]{rcl}
-\llbracket {q}^* \rrbracket.get  &=& (\llbracket q \rrbracket.get)^*\\
-\llbracket {q}^* \rrbracket.put &=& (\llbracket q \rrbracket.get)^*
-\end{array}
-\end{array}
-\]
-
-\[
-\begin{array}[b]{l@{\qquad}c}
-\Rule{\substack{q_1 : c_1 \Leftrightarrow d_1 \\ q_2 : c_2 \Leftrightarrow
-d_2} & {\substack{W(c_1) \cdot^! W(c_2)\\ K(c_1) \cdot^! K(c_2)}}
-& {\substack{W(d_1) \cdot^! W(d_2)\\ K(d_1) \cdot^! K(d_2)}}}
-{q_1 \cdot q_2: c_1 \cdot c_2 \Leftrightarrow d_1 \cdot d_2} &
-\begin{array}[b]{rcl}
-\llbracket q_1 \cdot q_2 \rrbracket.get  &=& \llbracket q_1 \rrbracket.get
-\cdot \llbracket q_2 \rrbracket.get\\
-\llbracket q_1 \cdot q_2\rrbracket.put &=& \llbracket q_1 \rrbracket.put
-\cdot \llbracket q_2 \rrbracket.put
-\end{array}
-\end{array}
-\]
-
-\[
-\begin{array}[b]{l@{\qquad}c}
-\Rule{\substack{q_1 : c_1 \Leftrightarrow d_1 \\ q_2 : c_2 \Leftrightarrow
-d_2} & \substack{\mathcal{L}(W(c_1)) \cap \mathcal{L}(W(c_2)) = \varnothing
-\\ \mathcal{L}(W(d_1)) \cap \mathcal{L}(W(d_2)) = \varnothing }}
-{q_1 \sep q_2: (c_1 \sep c_2)\Leftrightarrow (d_1 \sep d_2)} &
-\begin{array}[b]{rcl}
-\llbracket q_1 \sep q_2 \rrbracket.get(s) =
-\begin{cases}
-\llbracket q_1 \rrbracket.get (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
-\llbracket q_2 \rrbracket.get (s) & \text{if } s \in \mathcal{L}(W(c_2))\\
-\end{cases}\\
-\llbracket q_1 \sep q_2 \rrbracket.put(s) =
-\begin{cases}
-\llbracket q_1 \rrbracket.put (s) & \text{if } s \in \mathcal{L}(W(d_1))\\
-\llbracket q_2 \rrbracket.put (s) & \text{if } s \in \mathcal{L}(W(d_2))\\
-\end{cases}
-\end{array}
-\end{array}
-\]
-
-\begin{prooftree}
-\AxiomC{$\substack{q_1 : c \Leftrightarrow c_1 \\ q_2 : c_2 \Leftrightarrow
-c_3}$}
-\AxiomC{$\substack{\mathcal{L}(W(c_1))= \mathcal{L}(W(c_2)) \\ K(c_1) \equiv
-K(c_2)}$}
-\AxiomC{$\canonizer(c_1) = \canonizer(c_2)$}
-\TrinaryInfC{$q_2 \circ q_1: c \Leftrightarrow c_3$}
-\end{prooftree}
-\begin{align*}
-\llbracket q \rrbracket.get &= \llbracket q_2 \rrbracket.get\circ \llbracket
-q_1 \rrbracket.get, \text{ and }\\
-\llbracket q \rrbracket.put &= \llbracket q_1 \rrbracket.put \circ \llbracket
-q_2 \rrbracket.put
-\end{align*}
-
-\caption{Denotation and Typing Rules for QRE Lenses}
-\label{fig:qlenssemantics}
-\end{figure}
-
 \subsection{Normal Forms of QRE Lenses}
-Recall that our approach in defining QRE lenses is to have each QRE lens $q: c
-\Leftrightarrow c_1$ be such that
-\begin{align*}
-\llbracket q \rrbracket.get &= \ell \circ \canonizer(c)\\
-\llbracket q \rrbracket.put &= \ell^{-1} \circ
-\canonizer(c_1)
+Recall that our approach in defining QRE lenses is to have each QRE lens $q: c_1
+\Leftrightarrow c_2$ be such that
+\begin{align*}
+\llbracket q \rrbracket.\get &= \ell \circ \canonizer(c_1)\\
+\llbracket q \rrbracket.\lput &= \ell^{-1} \circ \canonizer(c_2)
 \end{align*}
@@
 \begin{theorem}\label{normal form}
-If there is a derivation $q : c \Leftrightarrow c_1$, then there exists a
-bijective lens $\ell : K(c) \Leftrightarrow K(c_1)$ such that
-\begin{align*}
-\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ \canonizer(c)\\
-\llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
-\canonizer(c_1)
+If there is a derivation $q : c_1 \Leftrightarrow c_2$, then there exists a
+bijective lens $\ell : K(c_1) \Leftrightarrow K(c_2)$ such that:
+\begin{align*}
+\llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket\circ
+\canonizer(c_1)\\
+\llbracket q \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
+\canonizer(c_2)
 \end{align*}
@@
 We elide a full proof of this statement, since the proof is attained via a
-straightforward induction on the derivation $q : c \Leftrightarrow c$ (the
-full details can be found in the appendix {\sam{which I need to write}}). Since
-bijective lenses are closed under strongly unambiguous regular operators, the
-tricky case is composition. The composition case is more involved since we need
-to show that we can get rid of the canonizers in the middle:
+straightforward induction on the derivation $q_1 : c_1 \Leftrightarrow c_2$ (the
+full details can be found in the appendix). Since bijective lenses are closed
+under strongly unambiguous regular operators, the tricky case is composition.
+The composition case is more involved since we need to show that we can get rid
+of the canonizers in the middle:
 
@@
 \begin{prooftree}
-\AxiomC{$\substack{q_1 : c \Leftrightarrow c_1 \\ q_2 : c_2 \Leftrightarrow
-c_3}$}
-\AxiomC{$\substack{\mathcal{L}(W(c_1))= \mathcal{L}(W(c_2)) \\ K(c_1) \equiv
-K(c_2)}$}
-\AxiomC{$\canonizer(c_1) = \canonizer(c_2)$}
-\TrinaryInfC{$q_2 \circ q_1: c \Leftrightarrow c_3$}
+\AxiomC{$\substack{q_1 : c_1 \Leftrightarrow c_2 \\ q_2 : c_3 \Leftrightarrow
+c_4}$}
+\AxiomC{$\substack{\mathcal{L}(W(c_2))= \mathcal{L}(W(c_3)) \\ K(c_2) \equiv
+K(c_3)}$}
+\AxiomC{$\canonizer(c_2) = \canonizer(c_3)$}
+\TrinaryInfC{$q_1 \; ; \; q_2: c_1 \Leftrightarrow c_4$}
 \end{prooftree}
@@
 \begin{align*}
-\llbracket q \rrbracket.get &= \llbracket q_2 \rrbracket.get\circ \llbracket
-q_1 \rrbracket.get, \text{ and }\\
-\llbracket q \rrbracket.put &= \llbracket q_1 \rrbracket.put \circ \llbracket
-q_2 \rrbracket.put
+\llbracket q_1 \; ; \; q_2 \rrbracket.\get &= \llbracket q_2 \rrbracket.\get\circ
+\llbracket q_1 \rrbracket.\get\\
+\llbracket q_1 \; ; \; q_2 \rrbracket.\lput &= \llbracket q_1 \rrbracket.\lput
+\circ \llbracket q_2 \rrbracket.\lput
 \end{align*}
@@
 By the induction hypothesis, there exist bijective lenses
-$\ell_1 :
-K(c) \Leftrightarrow K(c_1)$ and $\ell_2 : K(c_2) \Leftrightarrow K(c_3)$ such
-that
-\begin{align*}
-\llbracket q_1 \rrbracket.get &= \llbracket \ell_1 \rrbracket \circ
-\canonizer(c)\\
-\llbracket q_1 \rrbracket.put &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
-\canonizer(c_1)
-\end{align*}
-and
-\begin{align*}
-\llbracket q_2 \rrbracket.get &= \llbracket \ell_2 \rrbracket \circ
-Canonize(c_1)\\
-\llbracket q_2 \rrbracket.put &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
-Canonize(c_2)
-\end{align*}
+$\ell_1 : K(c_1) \Leftrightarrow K(c_2)$ and $\ell_2 : K(c_3) \Leftrightarrow
+K(c_4)$ such that
+\begin{center}
+
+\begin{tabular}{l l}
+$\llbracket q_1 \rrbracket.\get = \llbracket \ell_1 \rrbracket \circ
+\canonizer(c_1)$ & $\llbracket q_2 \rrbracket.\get = \llbracket \ell_2
+\rrbracket \circ \canonizer(c_3)$ \\
+$\llbracket q_1 \rrbracket.\lput = {\llbracket \ell_1 \rrbracket}^{-1} \circ
+\canonizer(c_2)$ & $\llbracket q_2 \rrbracket.\lput = {\llbracket \ell_2
+\rrbracket}^{-1} \circ \canonizer(c_4)$
+\end{tabular}
+
+\end{center}
+
 Consequently,
 \begin{align*}
-\llbracket q_2 \rrbracket.get \circ \llbracket q_1 \rrbracket.get &=
-(\llbracket \ell_2 \rrbracket \circ \canonizer(c_1)) \circ (\llbracket \ell_1
-\rrbracket \circ \canonizer(c))\\
-&= \llbracket \ell_2 \rrbracket \circ (\canonizer(c_1) \circ \llbracket \ell_1
-\rrbracket) \circ \canonizer(c)\\
+\llbracket q_2 \rrbracket.\get \circ \llbracket q_1 \rrbracket.\get &=
+(\llbracket \ell_2 \rrbracket \circ \canonizer(c_3)) \circ (\llbracket \ell_1
+\rrbracket \circ \canonizer(c_1))\\
+&= \llbracket \ell_2 \rrbracket \circ (\canonizer(c_3) \circ \llbracket \ell_1
+\rrbracket) \circ \canonizer(c_1)\\
 &= (\llbracket \ell_2 \rrbracket \circ \llbracket \ell_1 \rrbracket) \circ
-\canonizer(c)\\
-&= \llbracket \ell_2  \circ  \ell_1 \rrbracket \circ
-\canonizer(c)
+\canonizer(c_1)\\
+&= \llbracket \ell_1 \; ; \; \ell_2 \rrbracket \circ \canonizer(c_1)
 \end{align*}
 We are permitted to claim the third step from the second since
-$\canonizer(c_1)$ is the identity function on $K(c_2)$ which is
-syntactically equal to $K(c_1)$ by assumption. A similar argument shows that
-$$\llbracket q_1 \rrbracket.put \circ \llbracket q_2 \rrbracket.put =
-\llbracket \ell_2  \circ  \ell_1 \rrbracket^{-1} \circ
-\canonizer(c)$$
+$\canonizer(c_3)$ is the identity function on $K(c_2)$ which is
+syntactically equal to $K(c_3)$ by assumption. A similar argument shows that
+$$\llbracket q_1 \rrbracket.\lput \circ \llbracket q_2 \rrbracket.\lput =
+\llbracket \ell_1 \; ; \; \ell_2 \rrbracket^{-1} \circ
+\canonizer(c_4)$$
 
@@
 straightforward application of the induction hypothesis followed by unrolling
-the equations that give the denotation for QRE lenses.\qed
+the equations that give the denotation for QRE lenses.
 \end{proof}
-\iffalse
-\begin{proof}
-Assume that $q : c \Leftrightarrow c_1$. We proceed by induction over the
-derivation $q : c \Leftrightarrow c_1$.
+
+\section{Synthesizing QRE Lenses}
+\label{synth}
+
+QRE lenses address some of the limitations of bijective lenses because a single
+lens program expresses both the canonizers and the transformation between kernel
+languages simultaneously, which reduces programmer effort.  But we can go even
+further by recognizing that the type structure of QRE lenses contains
+information that can be exploited to automatically synthesize lenses from their
+types.  Rather than writing the QRE lens manually, the programmer can instead
+specify the desired behavior of a lens by giving its interface types and providing
+examples, if necessary, to disambiguate among possible implementations.  This
+way of constructing lenses can often be much simpler than building them by
+hand---see the example of lens synthesis in
+Section~\ref{sec:examplesynth}.
+
+The Optician framework~\cite{optician} showed how to do such lens synthesis in
+the case for bijective lenses. Here we show how to reduce QRE lens synthesis to
+that case, so that we can re-use the Optician algorithm but in the more
+expressive context of QRE lenses.  The basic idea is straightforward: we run the
+Optician algorithm to synthesize a lens between the kernels of two QREs and then
+apply the canonizers at the edges to recover a lens between the whole
+languages.  This simple strategy turns out to be remarkably effective, and the
+idea of using Optician in this way is part the impetus for the design of our QRE
+lenses.
+
+The Optician synthesis algorithm works as follows. Given regular expressions $R$
+and $S$ and a set of example input--output pairs
+$\{(r_1, s_1),\allowbreak \ldots,\allowbreak (r_n, s_n)\}$, Optician will try to
+find a bijective lens $\ell : R \Leftrightarrow S$ that that agrees on the
+examples, \textit{i.e.}, it maps $r_i$ to $s_i$ in the forward direction and
+$s_i$ to $r_i$ in the backward direction.  The Optician algorithm is guaranteed
+to succeed if such a lens exists.    \saz{We should probably mention what
+happens if the examples aren't sufficient to uniquely determine the lens.}
+
+In our setting, we want to synthesize a quotient lens $q: c_1 \Leftrightarrow
+c_2$ from the QREs $c_1$ and $c_2$ and a set of example input-output pairs
+$\{(x_1, y_1),\allowbreak \ldots,\allowbreak (x_n, y_n)\}$ where the $x_i$'s are
+in $W(c_1)$ and the $y_i$'s are in $W(c_2)$. We furthermore wish $q$ to map the
+equivalence class of $x_i$ to the equivalence class of $y_i$ and vice versa:
+\begin{align*}
+q.\get(x_i) &\equiv_{c_2} y_i \text{, and}\\
+q.\lput(y_i) &\equiv_{c_1} x_i
+\end{align*}
+Our approach to synthesizing QRE lenses is guided by Theorem~\ref{normal form},
+which says that, if there is a derivation $q : c_1 \Leftrightarrow c_2$ of a
+QRE lens, then there exists a bijective lens $\ell : K(c_1) \Leftrightarrow
+K(c_2)$ such that
+\begin{align*}
+\llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket\circ \canonizer(c_1)\\
+\llbracket q \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
+\canonizer(c_2).
+\end{align*}
+
+For the QRE examples, the $x_i$'s are in $W(c_1)$ and the $y_i$'s are
+in $W(c_2)$, so we can construct ${x'_i} = \canonizer(c_1)(x_i)$ in $K(c_1)$
+and ${y'_i} = \canonizer(c_2)(y_i)$ is in $K(c_2)$. To synthesize the
+desired quotient lens $q: c_1 \Leftrightarrow c_2$ that is consistent with the
+input-output examples $\{(x_1, y_1), \ldots, (x_n, y_n)\}$ it suffices to
+synthesize a bijective lens $\ell : K(c_1) \Leftrightarrow K(c_2)$ that is
+consistent with the canonized examples $\{({x'_1}, {y'_1}), \ldots, ({x'_n},
+{y'_n})\}$ and then apply the canonizers at the outside. Our procedure for QRE
+lens synthesis is thus
 \begin{enumerate}
   \item
-  $\kw{lift}(\ell): R/\mathit{id}(R) \Leftrightarrow S/\mathit{id}(S)$ where
-  $\ell :
-  R \Leftrightarrow S$. Then
-  \begin{align*}
-\llbracket \kw{lift}(\ell) \rrbracket.get &=  \llbracket \ell \rrbracket
-= \llbracket \ell \rrbracket \circ id_{\mathcal{L}(R)} =
-\llbracket \ell \rrbracket \circ \canonizer(\mathit{id}(R)), \text{ and }\\
-\llbracket \kw{lift}(\ell) \rrbracket.put &= \llbracket \ell
-\rrbracket^{-1} = \llbracket \ell \rrbracket^{-1} \circ id_{\mathcal{L}(S)} =
-\llbracket \ell \rrbracket^{-1} \circ \canonizer(id(S))
-\end{align*}
+  Compute $K(c_1)$, $K(c_2)$,
   \item
-$\kw{lquot}(c, q'): c_1 \circ c \Leftrightarrow c_2$ where $q' : c_1
-\Leftrightarrow c_2$, $c$ is well formed and $K(c) = W(c_1)$. Then
-\begin{align*}
-\llbracket q \rrbracket.get  &= \llbracket q'
-\rrbracket.get \circ \canonizer(c)\\
-\llbracket q \rrbracket.put &= \llbracket q' \rrbracket.put
-\end{align*}
-By the induction hypothesis, there exists a bijective lens $\ell :
-K(c_1) \Leftrightarrow K(c_2)$ such that
-\begin{align*}
-\llbracket q' \rrbracket.get &= \llbracket \ell \rrbracket \circ \canonizer(c_1)\\
-\llbracket q' \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
-Canonize(c_2)
-\end{align*}
-Consequently
-\begin{align*}
-\llbracket q \rrbracket.get  &= (\llbracket \ell \rrbracket \circ
-\canonizer(c_1)) \circ \canonizer(c) = \llbracket \ell \rrbracket \circ
-(\canonizer(c_1 \circ c))\\
-\llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
-Canonize(c_2)
-\end{align*}
-
+  Compute $\{({x'_1}, {y'_1}), \ldots, ({x'_n}, {y'_n})\}$ where ${x'_i} =
+  \canonizer(c_1)(x_i)$ and ${y'_i} = \canonizer(c_2)(y_i)$,
+\item Synthesize, using Optician's algorithm, a bijective lens
+  $\ell : K(c_1) \Leftrightarrow K(c_2)$ that is consistent with the examples
+  $\{({x'_1}, {y'_1}),\allowbreak \ldots,\allowbreak ({x'_n}, {y'_n})\}$
   \item
-$\kw{rquot}(q', c_2):c \Leftrightarrow c_2 \circ c_1$ where $q' : c \Leftrightarrow
-c_1$, $c_2$ is well formed and $K(c_2) = W(c_1)$. Proceed as in the previous
-case.
-\item
-$q_2 \circ q_1: c \Leftrightarrow c_2$ where $q_1 : c \Leftrightarrow c_1$ and
-$q_2 : c_1 \Leftrightarrow c_2$. Then
-\begin{align*}
-\llbracket q \rrbracket.get &= \llbracket q_2 \rrbracket.get\circ \llbracket
-q_1 \rrbracket.get, \text{ and }\\
-\llbracket q \rrbracket.put &= \llbracket q_1 \rrbracket.put \circ \llbracket
-q_2 \rrbracket.put
-\end{align*}
-By the induction hypothesis, there exist bijective lenses
-$\ell_1 :
-K(c) \Leftrightarrow K(c_1)$ and $\ell_2 : K(c_1) \Leftrightarrow K(c_2)$ such
-that
-\begin{align*}
-\llbracket q_1 \rrbracket.get &= \llbracket \ell_1 \rrbracket \circ
-\canonizer(c)\\
-\llbracket q_1 \rrbracket.put &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
-\canonizer(c_1)
-\end{align*}
-and
-\begin{align*}
-\llbracket q_2 \rrbracket.get &= \llbracket \ell_2 \rrbracket \circ
-Canonize(c_1)\\
-\llbracket q_2 \rrbracket.put &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
-Canonize(c_2)
-\end{align*}
-Consequently,
-\begin{align*}
-\llbracket q_2 \rrbracket.get \circ \llbracket q_1 \rrbracket.get &=
-(\llbracket \ell_2 \rrbracket \circ \canonizer(c_1)) \circ (\llbracket \ell_1
-\rrbracket \circ \canonizer(c))\\
-&= \llbracket \ell_2 \rrbracket \circ (\canonizer(c_1) \circ \llbracket \ell_1
-\rrbracket) \circ \canonizer(c)\\
-&= (\llbracket \ell_2 \rrbracket \circ \llbracket \ell_1 \rrbracket) \circ
-\canonizer(c)\\
-&= \llbracket \ell_2  \circ  \ell_1 \rrbracket \circ
-\canonizer(c)
-\end{align*}
-A similar argument shows that
-$$\llbracket q_1 \rrbracket.put \circ \llbracket q_2 \rrbracket.put =
-\llbracket \ell_2  \circ  \ell_1 \rrbracket^{-1} \circ
-\canonizer(c)$$
-\item
-${q'}^* : c^* \Leftrightarrow {c_1}^*$ where $q' : c \Leftrightarrow c_1$,
-$W(c)^{*!}$ and $W(c_1)^{*!}$ and $K(c)^{*!}$ and $K(c_1)^{*!}$. Then
-\begin{align*}
-\llbracket {q'}^* \rrbracket.get &= (\llbracket q' \rrbracket.get)^*, \text{
-and }\\
-\llbracket {q'}^* \rrbracket.put &= (\llbracket q' \rrbracket.put)^*
-\end{align*}
-By the induction hypothesis there exists a bijective lens $\ell : K(c)
-\Leftrightarrow K(c_1)$ such that
-that
-\begin{align*}
-\llbracket q' \rrbracket.get &= \llbracket \ell \rrbracket \circ
-\canonizer(c)\\
-\llbracket q' \rrbracket.put &= {\llbracket \ell \rrbracket}^{-1} \circ
-\canonizer(c_1)
-\end{align*}
-Consequentlty
-\begin{align*}
-\llbracket {q'}^* \rrbracket.get &= (\llbracket \ell \rrbracket \circ
-\canonizer(c))^* = \llbracket \ell \rrbracket^* \circ
-\canonizer(c)^* = \llbracket \ell^* \rrbracket \circ
-\canonizer(c^*)\\
-\llbracket {q'}^* \rrbracket.put &= (\llbracket \ell \rrbracket^{-1} \circ
-\canonizer(c_1))^* = (\llbracket \ell \rrbracket^{-1})^* \circ
-\canonizer(c_1)^* = \llbracket \ell^* \rrbracket^{-1} \circ
-\canonizer(c_1^*)\\
-\end{align*}
-\item
-$q_1 \cdot q_2: c_1 \cdot c_2 \Leftrightarrow d_1 \cdot d_2$, where $q_1 : c_1
-\Leftrightarrow d_1 $,  $q_2 : c_2 \Leftrightarrow d_2$, $W(c_1)
-\cdot^! W(c_2)$, $K(c_1) \cdot^! K(c_2)$, $W(d_1) \cdot^! W(d_2)$ and $
-K(d_1) \cdot^! K(d_2)$. Then
-\begin{align*}
-\llbracket q \rrbracket.get &= \llbracket q_1 \rrbracket.get \cdot \llbracket
-q_2 \rrbracket.get, \text{ and }\\
-\llbracket q \rrbracket.put &= \llbracket q_1 \rrbracket.put \cdot \llbracket
-q_2 \rrbracket.put
-\end{align*}
-By the induction hypothesis, there exist bijective lenses $\ell_1 : K(c_1)
-\Leftrightarrow K(d_1)$ and $\ell_2 : K(c_2) \Leftrightarrow K(d_2)$ such that
-\begin{align*}
-\llbracket q_1 \rrbracket.get &= \llbracket \ell_1 \rrbracket \circ
-\canonizer(c_1)\\
-\llbracket q_1 \rrbracket.put &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
-\canonizer(d_1)
-\end{align*}
-and
-\begin{align*}
-\llbracket q_2 \rrbracket.get &= \llbracket \ell_2 \rrbracket \circ
-\canonizer(c_2)\\
-\llbracket q_2 \rrbracket.put &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
-\canonizer(d_2)
-\end{align*}
-Consequently,
-\begin{align*}
-\llbracket q \rrbracket.get &= (\llbracket \ell_1 \rrbracket \circ
-\canonizer(c_1)) \cdot  (\llbracket \ell_2 \rrbracket \circ
-\canonizer(c_2))\\
-&= (\llbracket \ell_1 \rrbracket \cdot \llbracket \ell_2
-\rrbracket) \circ (\canonizer(c_1) \cdot \canonizer(c_2))\\
-&= \llbracket \ell_1 \cdot  \ell_2 \rrbracket \circ \canonizer(c_1 \cdot c_2)
-\end{align*}
-Similarly
-$$
-\llbracket q \rrbracket.put = \llbracket \ell_1 \cdot  \ell_2 \rrbracket^{-1}
-\circ \canonizer(d_1 \cdot d_2) $$
-\item
-$q = q_1 \sep q_2$ where $q_1 : c_1 \Leftrightarrow d_1 $, $q_2 : c_2
-\Leftrightarrow d_2$, $\mathcal{L}(W(c_1)) \cap \mathcal{L}(W(c_2)) =
-\varnothing$ and $\mathcal{L}(W(d_1)) \cap \mathcal{L}(W(d_2)) = \varnothing$.
-Then
-$$
-\llbracket q_1 \sep q_2 \rrbracket.get(s) =
-\begin{cases}
-\llbracket q_1 \rrbracket.get (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
-\llbracket q_2 \rrbracket.get (s) & \text{if } s \in \mathcal{L}(W(c_2))\\
-\end{cases}$$
-$$\llbracket q_1 \sep q_2 \rrbracket.put(s) =
-\begin{cases}
-\llbracket q_1 \rrbracket.put (s) & \text{if } s \in \mathcal{L}(W(d_1))\\
-\llbracket q_2 \rrbracket.put (s) & \text{if } s \in \mathcal{L}(W(d_2))\\
-\end{cases}
-$$
-By the induction hypothesis, there exist bijective lenses $\ell_1 : K(c_1)
-\Leftrightarrow K(d_1)$ and $\ell_2 : K(c_2) \Leftrightarrow K(d_2)$ such that
-\begin{align*}
-\llbracket q_1 \rrbracket.get &= \llbracket \ell_1 \rrbracket \circ
-\canonizer(c_1)\\
-\llbracket q_1 \rrbracket.put &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
-\canonizer(d_1)
-\end{align*}
-and
-\begin{align*}
-\llbracket q_2 \rrbracket.get &= \llbracket \ell_2 \rrbracket \circ
-\canonizer(c_2)\\
-\llbracket q_2 \rrbracket.put &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
-\canonizer(d_2)
-\end{align*}
-Consequently,
-$$
-\llbracket q_1 \sep q_2 \rrbracket.get(s) =
-\begin{cases}
-\llbracket \ell_1 \rrbracket \circ
-\canonizer(c_1) (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
-\llbracket \ell_2 \rrbracket \circ
-\canonizer(c_2) (s) & \text{if } s \in \mathcal{L}(W(c_2)),\\
-\end{cases}$$
-so $\llbracket q_1 \sep q_2 \rrbracket.get = \llbracket \ell_1 \sep
-\ell_2 \rrbracket \circ \canonizer(c_1 \sep c_2)$. A similar argument shows
-that $\llbracket q_1 \sep q_2 \rrbracket.put = \llbracket \ell_1 \sep
-\ell_2 \rrbracket^{-1} \circ \canonizer(d_1 \sep d_2)$.\\
-This completes the proof.
+  Return $q = \kw{rquot}(\kw{lquot}(c_1, \ell), c_2) : c_1 \Leftrightarrow
+  c_2$
 \end{enumerate}
-\end{proof}
-
-\fi
-\section{Synthesizing QRE Lenses}
-\label{synth}
-Our goal in this section is to demonstrate how to synthesize a quotient lens
-$q: c \Leftrightarrow c_1$ from the QREs $c, c_1$ and a set of input-output pairs
-$\{(x_1, y_1),\allowbreak \ldots,\allowbreak (x_n, y_n)\}$ where the $x_i$'s are in $W(c)$ and the
-$y_i$'s are in $W(c_1)$. We wish $q$ to map the equivalence class of $x_i$ to the
-equivalence class of $y_i$ and vice versa:
-\begin{align*}
-q.get(x_i) &\equiv_{c_1} y_i \text{, and}\\
-q.put(y_i) &\equiv_c x_i
-\end{align*}
-
-Our approach in synthesizing QRE lenses is guided by Theorem~\ref{normal
-form}, which says that if there is a derivation $q :
-c \Leftrightarrow c_1$ of a QRE lens, then there exists a bijective lens $\ell :
-K(c) \Leftrightarrow K(c_1)$ such that
-\begin{align*}
-\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ \canonizer(c)\\
-\llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
-\canonizer(c_1)
-\end{align*}
-
-For our synthesis task, since the $x_i$'s are in $W(c)$ and the $y_i$'s are
-in $W(c_1)$, then ${x_i}' = \canonizer(c)(x_i)$ is in $K(c)$ and ${y_i}' =
-\canonizer(c_1)(y_i)$ is in $K(c_1)$. Consequently in order to synthesize the
-desired quotient lens $q: c \Leftrightarrow c_1$ that is consistent with the
-input-output examples $\{(x_1, y_1), \ldots, (x_n, y_n)\}$ it suffices to
-synthesize a bijective lens $l :
-K(c) \Leftrightarrow K(c_1)$ that is consistent with the canonized examples
-$\{({x_1}', {y_1}'), \ldots, ({x_n}', {y_n}')\}$
-
-Optician~\cite{optician} showed that given regular expressions
-$R, S$ and a set of input-output pairs $\{(r_1, s_1),\allowbreak \ldots,\allowbreak (r_n, s_n)\}$, if
-there is a derivation $\ell : R \Leftrightarrow S$ of a bijective lens that
-maps $r_i$ to $s_i$ in the forward direction and $s_i$ to $r_i$ in the backward
-direction, then there is an algorithm that computes a bijective lens $\ell' :
-R' \Leftrightarrow S'$ such that $R \equiv R'$, $S \equiv S'$, $\ell'$ maps
-$r_i$ to $s_i$ in the forward direction and $s_i$ to $r_i$ in the backward
-direction, and $\llbracket \ell \rrbracket = \llbracket \ell'
-\rrbracket$.
-
-This implies that if there is a derivation $q : c \Leftrightarrow
-c_1$ of a QRE lens as well as a derivation $\ell : K(c) \Leftrightarrow K(c_1)$
-that is consistent with the canonized examples $\{({x_1}', {y_1}'),\allowbreak
-\ldots,\allowbreak ({x_n}', {y_n}')\}$, then we can synthesize a quotient lens $q$ of
-type $c \Leftrightarrow c_1$ that is consistent with $\{({x_1}', {y_1}'),
-\ldots, ({x_n}', {y_n}')\}$. More concretely, in order to synthesize a
-quotient lens $q: c \Leftrightarrow c_1$, that is consistent with the examples
-$\{(x_1, y_1),\allowbreak \ldots, (x_n, y_n)\}$ it suffices to perform the following
-procedure:
-\begin{enumerate}
-  \item
-  Compute $K(c)$, $K(c_1)$,
-  \item
-  Compute $\{({x_1}', {y_1}'), \ldots, ({x_n}', {y_n}')\}$ where ${x_i}' =
-  \canonizer(c)(x_i)$ and ${y_i}' = \canonizer(c_1)(y_i)$,
-  \item
-  Synthesize a bijective lens $\ell : K(c) \Leftrightarrow K(c_1)$ that is
-  consistent with the examples $\{({x_1}', {y_1}'),\allowbreak
-  \ldots,\allowbreak ({x_n}',
-  {y_n}')\}$
-  \item
-  Return $q = \kw{rquot}(\kw{lquot}(c, \ell), c_1) : c \Leftrightarrow
-  c_1$
-\end{enumerate}
-For example, recall that in the \bibtex{} to EndNote transformation we had the
-QREs \cd{bibtex} and \cd{endnote} describing a \bibtex{} record and an Endnote
-record respectively. We also had the example pair \cd{(bib_example ,
-end_example)} where
-
+
+\saz{Do we want to state formally the completeness of this algorithm?}
+
+Returning to the \bibtex{} to EndNote transformation of
+Section~\ref{sec:examplesynth}, we had the QREs \cd{bibtex} and \cd{endnote}
+describing a \bibtex{} record and an Endnote record respectively. We also had
+the example pair \cd{(bib_example , end_example)} where
+
+\begin{minipage}{3in}
 \begin{lstlisting}
@@
 \end{lstlisting}
-
-\noindent and
-
+\end{minipage}
+\begin{minipage}{3in}
 \begin{lstlisting}
@@
 \end{lstlisting}
+\end{minipage}
 
@@
 \cd{endnote} that maps \cd{bib_example} to \cd{end_example} can be
-expressed by a bijective lens, we can synthesize a lens that maps between the
-QREs \cd{bibtex} and \cd{endnote}. We do this using the expression
+expressed as a bijective lens, we can synthesize a lens that maps between the
+QREs \cd{bibtex} and \cd{endnote}, using the code below:
 
@@
 
+In this case, one example is sufficient to uniquely determine the desired
+transformation.\saz{Is that true?}
+
+\saz{We might want to say something more precise about the running time or
+  expected performance of this algorithm, perhaps to set up the evaluation
+  section.  Something along the lines of: ``Although the algorithm is complete
+  in theory, in the worst case, it may take an unbounded [???IS THIS CORRECT?]
+  amount time to search for a satisfactory lens.  Nevertheless, this algorithm
+  performs well in practice on numerous examples, as we show in the next
+  section.}
+
 \section{Implementation and Evaluation}
@@
 
-We have integrated QREs and quotient lens synthesis into Boomerang. With this
-integration, QREs and quotient lens synthesis tasks can be expressed alongside
-existing Boomerang terms.
-
-All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
-of 1600 MHz DDR3 running macOS High Sierra.
-
-Our evaluations serve to answer the following questions:
+We have implemented QREs and the quotient lens synthesis algorithm described
+above as part of Boomerang. Thanks to this integration, QREs and quotient lens
+synthesis tasks can now be expressed alongside existing Boomerang terms, and we
+can combine the two kinds of lenses seamlessly in the same application.
+Intuition suggests that the increased expressiveness of QREs should make many
+lens programming tasks easier, but we would like more quantitative understanding
+of that claim.  We thus conducted several experiments that aim to provide
+answers to the following questions:
+
 \begin{enumerate}
@@
 
+All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
+of 1600 MHz DDR3 running macOS High Sierra.
+
 \subsection{Benchmark Suite Construction}
-Our benchmarks were constructed from two sources:
+
+To evaluate our QRE implementation, we constructed a set of lens synthesis tasks
+drawn from two sources:
 \begin{enumerate}
-  \item 39 of our benchmarks were derived from the Optician system.  These
-  benchmarks were a combination of custom benchmarks, benchmarks derived from
-  FlashFill, and benchmarks derived from Augeas.  Many of the formats mapped
-  between had to be modified to work with the bijectivity constraints
-  Optician required.  Commonly, one format had whitespace where the other did
-  not allow for whitespace.  To address the bijectivity issues surrounding this,
-  we added whitespace to the format that typically did not have whitespace.  We
-  were able to remove these alterations in adopting these benchmarks to our
-  system.
-
-  \item 8 of our benchmarks were derived from data.gov.  Data.gov contains
-  many transformations between formats.  In addition to the formats previously
-  adopted for Optician, we synthesized transformers between these formats.
+\item 39 of our benchmarks were adapted from the evaluation of the original
+  Optician system.  These benchmarks are a combination of custom benchmarks,
+  benchmarks derived from FlashFill~\cite{flashfill}, and benchmarks
+  derived from Augeas~\cite{augeas2}. In the pre-QRE version of
+  Optician, many of these data formats had to be modified to work with the
+  bijectivity constraints that Optician required.  For instance, when one format
+  permits whitespace where the other does not allow it, the original versions of
+  benchmark formats were modified to allow more whitespace, thereby restoring
+  bijectivity.  With the new QRE support, we were able to remove these
+  alterations when porting these benchmarks.  (This experience alone suggests
+  that QREs make the lens development process easier.)
+
+\item The remaining eight benchmarks were derived from
+  data.gov.\footnote{\url{http://data.gov}} These datasets include census
+  population data, wage comparison data, demographics statistics data and
+  crime index data. We chose these datasets since they were stored in XML, CSV
+  and RDF formats which are more easily specified using regular expressions than
+  plain text or PDF documents.
 \end{enumerate}
@@
 that it takes far fewer AST nodes to define benchmark lenses using QRE
-synthesis than with Optician without QRE synthesis or without synthesis. }
+synthesis than with Optician without QRE synthesis or without synthesis.\saz{Why
+is the bar for ``All QRE Benchmarks'' at 14? Shouldn't it be at 47 = 39+8? Are
+we not testing all QRE benchmarks?}}
 \label{fig:asts}
@@
 
-To test if our tool eases the development process, we compare the following
-number of AST nodes.
+To get a sense of how much the new expressiveness that combining QRE lenses with
+synthesis saves in terms of programmer effort, we compared the amount of code
+that would need to be written for each of the lens benchmarks.  We use number of
+nodes in the abstract syntax tree as a measure of code size, and calculated
+three quantities for each lens synthesis benchmark task.  \saz{Say something
+  about why there are only 14 benchmarks?}
 %
 \begin{itemize}
-  \item[\QRESize{}] is the number of AST nodes in the QRE
-  specifications involved in QRE lens synthesis task.
-  \item[\CanonizerAndSpecSize{}] is the number of
-  AST nodes in $W(q)$ summed with the number of AST nodes in $\canonizer(q)$,
-  for each QRE used in QRE lens synthesis task.
-  \item[\LensAndSpecSize{}] is the number of AST nodes
-  in $W(q)$ summed with the number of AST nodes in the synthesized QRE lens.
+  \item[\QRESize{}] the number of AST nodes in the QRE
+  specifications (including examples)\saz{CHECK THIS}
+  \item[\CanonizerAndSpecSize{}]  the number of
+  AST nodes in $W(q_1)$ summed with the number of AST nodes in $\canonizer(q_1)$
+  \item[\LensAndSpecSize{}] the number of AST nodes
+  in $W(q_1)$ summed with the number of AST nodes in the synthesized QRE lens.
 \end{itemize}
 
-The value of \QRESize{} represents the difficulty in writing our benchmarks.
-
-The value of \CanonizerAndSpecSize{} represents the difficulty in writing our
-benchmarks within the Optician system. This is only an approximation, as both
-$W(q)$ and $\canonizer(q)$ are automatically generated from the QRE, instead of
-written by a human.
-
-The value of \LensAndSpecSize{} represents the difficulty in writing our
-benchmarks in Boomerang with neither QREs nor synthesis. This is also an
-approximation, as $W(q)$ and $\canonizer(q)$ and the bijective lens are all
+The value of \QRESize{} is our measure of the programmer burden for
+writing our benchmarks---it measures the number of AST nodes the programmer has
+to write using the full expressiveness of QREs and synthesis.
+
+The value of \CanonizerAndSpecSize{} estimates the burden for writing our
+benchmarks within the Optician system, without QREs but still using
+synthesis. This is only an approximation, as both $W(q_1)$ and $\canonizer(q_1)$ are
+automatically generated from the QRE, and it is possible that a human-written
+version might be smaller.
+
+The value of \LensAndSpecSize{} estimates the burden in writing our benchmarks
+in Boomerang, using neither QREs nor synthesis. This value is also an
+approximation, as $W(q_1)$ and $\canonizer(q_1)$ and the bijective lens are all
 automatically generated.
 
-Figure~\ref{fig:asts} shows the number of benchmarks written with at most a
-given number of AST nodes. For a given benchmark, we found \QRESize{} to be
-about half as large as \CanonizerAndSpecSize, and about a third as large as
-\LensAndSpecSize{}. This demonstrates that introducing QREs saves programmers
+Figure~\ref{fig:asts} shows the number of benchmarks that can be expressed using
+at most a given number of AST nodes. For a given benchmark, we found \QRESize{}
+to be about half as large as \CanonizerAndSpecSize, and about a third as large
+as \LensAndSpecSize{}. This demonstrates that introducing QREs saves programmers
 significant effort compared to both Optician and basic Boomerang.
@@
 
-Optician was only able to synthesize fully bijective data transformations.
-Synthesizing from QREs loosens that restriction, making it possible to
-synthesize data which is bijective when put in canonical form. To determine the
-increase of expressivity, we analyzed:
+The pre-QRE version of Optician was only able to synthesize fully bijective data
+transformations.  Synthesizing from QREs loosens that restriction, making it
+possible to synthesize data which is bijective when put in canonical form. To
+determine the increase of expressivity, we analyzed:
 \begin{enumerate}
@@
 
-We found that 10 of Opticians 39 benchmarks required modifications to be
+We found that 10 of Optician's 39 benchmarks required modifications to be
 bijective. We found that 4 of Optician's 39 benchmarks required modifications to
-be in bijective correspondence when in canonical form. Furthermore, we found
-that all of the data.gov examples required no modifications to be in bijective
-correspondence when in canonical form. While we did not put the data.gov
-examples into a form suitable for Optician, we found only 4 of the 8 data.gov
-examples did not require canonization, so making suitable for Optician would
-have required a modification of 4 of the data.gov formats.
+be in bijective correspondence when in canonical form.\saz{Is this somehow where
+  the 14 = 10+4 points in Fig. 10 come from?}\saz{Do these 10 and 4 intersect?}
+Furthermore, we found that all eight of the data.gov examples required no
+modifications to be in bijective correspondence when in canonical form. While we
+did not put the data.gov examples into a form suitable for pre-QRE Optician, we
+found that only four of the eight data.gov examples did not require
+canonization, so making them suitable for pre-QRE Optician would have required 
+changes to of four of the data.gov formats.  Here again, we see that QREs add
+significant expressive power.
 
@@
 
-To be used in a development environment, we found it important to validate
-performance did not significantly degrade from Optician. To validate this, we
-tested the runtimes when:
+To assess the performance impact of adding QRE support to Optician, we measured
+the running time of the following configurations:
 \begin{itemize}
-  \item[\OpticianRuntime{}] Optician runs on its own benchmarks.
-  \item[\SystemOnOptician{}] Our system runs on Optician's benchmarks.
-  \item[\SystemOnBenchmarks{}] Our system runs on its own benchmarks.
+  \item[\OpticianRuntime{}] pre-QRE Optician running on its own benchmarks.
+  \item[\SystemOnOptician{}] Optician with QREs running on the pre-QRE benchmarks.
+  \item[\SystemOnBenchmarks{}] Optician with QREs running on its own benchmarks.
 \end{itemize}
 
-We summarize the results of these experiments in Figure~\ref{fig:times}.
-Our system was
-able to synthesize all of Optician's benchmarks at a speed competitive with
-Optician.  There was additional overhead in calculating $W(q)$ and $K(q)$,
+We summarize the results of these experiments in Figure~\ref{fig:times}.  The
+new version of Optician was able to synthesize all of pre-QRE Optician's
+benchmarks at a speed competitive with the old version.  There is a small amount
+of additional overhead introduced by QREs when calculating $W(q_1)$ and $K(q_1)$,
 resulting in a very slight decrease in performance.
 
-Also notable was one of the data.gov examples, which converted demographic
-statistics by zip code from xml form into rdf form. This merely took longer
-because it was an exceptionally large format, converting over 40 fields. We
-added in additional fields to make this format suitable for Optician, which
-took over 15 seconds synthesize the modified lens.
+Also notable is one of the data.gov examples, which converts demographic
+statistics by zip code from xml form into rdf form. This took longer because it
+was an exceptionally large format, converting over 40 fields.\saz{We need to
+  explain more detail here.  I guess this is about the last benchmark in Fig. 11
+  (b), which takes a long time?}  We added in additional fields to make this
+format suitable for Optician, which took over 15 seconds synthesize the modified
+lens.\saz{This needs more context, explanation.}
 
@@
 
+\section{Appendix}
+We give a full proof of \cref{normal form}: If there is a derivation $q : c_1
+\Leftrightarrow c_2$, then there exists a bijective lens $\ell : K(c_1)
+\Leftrightarrow K(c_2)$ such that:
+\begin{align*}
+\llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket\circ
+\canonizer(c_1)\\
+\llbracket q \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
+\canonizer(c_2)
+\end{align*}
+\begin{proof}
+We proceed by induction over the derivation $q : c_1 \Leftrightarrow c_2$.
+\begin{enumerate}
+  \item
+  $\kw{lift}(\ell): R/\mathit{id}(R) \Leftrightarrow S/\mathit{id}(S)$ where
+  $\ell : R \Leftrightarrow S$. Then
+  \begin{align*}
+\llbracket \kw{lift}(\ell) \rrbracket.\get &=  \llbracket \ell \rrbracket
+= \llbracket \ell \rrbracket \circ id_{\mathcal{L}(R)} =
+\llbracket \ell \rrbracket \circ \canonizer(\mathit{id}(R)), \text{ and }\\
+\llbracket \kw{lift}(\ell) \rrbracket.\lput &= \llbracket \ell
+\rrbracket^{-1} = \llbracket \ell \rrbracket^{-1} \circ id_{\mathcal{L}(S)} =
+\llbracket \ell \rrbracket^{-1} \circ \canonizer(id(S))
+\end{align*}
+\item
+$\kw{lquot}(c_1, q): c_1 \; ; \; c_2 \Leftrightarrow c_3$ where $q : c_2
+\Leftrightarrow c_3$, $c_1$ is well formed and $K(c_1) = W(c_2)$. Then
+\begin{align*}
+\llbracket \kw{lquot}(c_1, q) \rrbracket.\get  &= \llbracket q
+\rrbracket.\get \circ \canonizer(c_1)\\
+\llbracket \kw{lquot}(c_1, q) \rrbracket.\lput &= \llbracket q \rrbracket.\lput
+\end{align*}
+By the induction hypothesis, there exists a bijective lens $\ell :
+K(c_2) \Leftrightarrow K(c_3)$ such that
+\begin{align*}
+\llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket \circ
+\canonizer(c_2)\\
+\llbracket q \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
+\canonizer(c_3)
+\end{align*}
+Consequently
+\begin{align*}
+\llbracket \kw{lquot}(c_1, q)\rrbracket.\get  &= (\llbracket \ell \rrbracket \circ
+\canonizer(c_2)) \circ \canonizer(c_1) = \llbracket \ell \rrbracket \circ
+(\canonizer(c_1 \; ; \; c_2))\\
+\llbracket \kw{lquot}(c_1, q) \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
+\canonizer(c_3)
+\end{align*}
+
+\item
+$\kw{rquot}(q, c_3):c_1 \Leftrightarrow c_2 \; ; \; c_3$ where $q : c_1
+\Leftrightarrow c_2$, $c_3$ is well formed and $K(c_3) = W(c_2)$. Proceed as in
+the previous case.
+\item
+$q_1 \; ; \; q_2: c \Leftrightarrow c_2$ where $q_1 : c \Leftrightarrow c_1$ and
+$q_2 : c_1 \Leftrightarrow c_2$. Then
+\begin{align*}
+\llbracket q_1 \rrbracket.\get &= \llbracket q_2 \rrbracket.\get\circ \llbracket
+q_1 \rrbracket.\get, \text{ and }\\
+\llbracket q_1 \rrbracket.\lput &= \llbracket q_1 \rrbracket.\lput \circ \llbracket
+q_2 \rrbracket.\lput
+\end{align*}
+By the induction hypothesis, there exist bijective lenses
+$\ell_1 :
+K(c) \Leftrightarrow K(c_1)$ and $\ell_2 : K(c_1) \Leftrightarrow K(c_2)$ such
+that
+\begin{align*}
+\llbracket q_1 \rrbracket.\get &= \llbracket \ell_1 \rrbracket \circ
+\canonizer(c)\\
+\llbracket q_1 \rrbracket.\lput &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
+\canonizer(c_1)
+\end{align*}
+and
+\begin{align*}
+\llbracket q_2 \rrbracket.\get &= \llbracket \ell_2 \rrbracket \circ
+\canonizer(c_1)\\
+\llbracket q_2 \rrbracket.\lput &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
+\canonizer(c_2)
+\end{align*}
+Consequently,
+\begin{align*}
+\llbracket q_2 \rrbracket.\get \circ \llbracket q_1 \rrbracket.\get &=
+(\llbracket \ell_2 \rrbracket \circ \canonizer(c_1)) \circ (\llbracket \ell_1
+\rrbracket \circ \canonizer(c))\\
+&= \llbracket \ell_2 \rrbracket \circ (\canonizer(c_1) \circ \llbracket \ell_1
+\rrbracket) \circ \canonizer(c)\\
+&= (\llbracket \ell_2 \rrbracket \circ \llbracket \ell_1 \rrbracket) \circ
+\canonizer(c)\\
+&= \llbracket \ell_1 \; ; \; \ell_2 \rrbracket \circ
+\canonizer(c)
+\end{align*}
+A similar argument shows that
+$$\llbracket q_1 \rrbracket.\lput \circ \llbracket q_2 \rrbracket.\lput =
+\llbracket \ell_1 \; ; \; \ell_2 \rrbracket^{-1} \circ
+\canonizer(c)$$
+\item
+$q^* : {c_1}^* \Leftrightarrow {c_2}^*$ where $q : c_1 \Leftrightarrow c_2$,
+$W(c_1)^{*!}$ and $W(c_2)^{*!}$ and $K(c_1)^{*!}$ and $K(c_2)^{*!}$. Then
+\begin{align*}
+\llbracket q^* \rrbracket.\get &= (\llbracket q \rrbracket.\get)^*, \text{
+and }\\
+\llbracket q^* \rrbracket.\lput &= (\llbracket q \rrbracket.\lput)^*
+\end{align*}
+By the induction hypothesis there exists a bijective lens $\ell : K(c_1)
+\Leftrightarrow K(c_2)$ such that
+that
+\begin{align*}
+\llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket \circ
+\canonizer(c_1)\\
+\llbracket q \rrbracket.\lput &= {\llbracket \ell \rrbracket}^{-1} \circ
+\canonizer(c_2)
+\end{align*}
+Consequentlty
+\begin{align*}
+\llbracket q^* \rrbracket.\get &= (\llbracket \ell \rrbracket \circ
+\canonizer(c_1))^* = \llbracket \ell \rrbracket^* \circ
+\canonizer(c_1)^* = \llbracket \ell^* \rrbracket \circ
+\canonizer(c_1^*)\\
+\llbracket q^* \rrbracket.\lput &= (\llbracket \ell \rrbracket^{-1} \circ
+\canonizer(c_2))^* = (\llbracket \ell \rrbracket^{-1})^* \circ
+\canonizer(c_2)^* = \llbracket \ell^* \rrbracket^{-1} \circ
+\canonizer(c_2^*)\\
+\end{align*}
+\item
+$q_1 \cdot q_2: c_1 \cdot c_2 \Leftrightarrow d_1 \cdot d_2$, where $q_1 : c_1
+\Leftrightarrow d_1 $,  $q_2 : c_2 \Leftrightarrow d_2$, $W(c_1)
+\cdot^! W(c_2)$, $K(c_1) \cdot^! K(c_2)$, $W(d_1) \cdot^! W(d_2)$ and $
+K(d_1) \cdot^! K(d_2)$. Then
+\begin{align*}
+\llbracket q_1 \rrbracket.\get &= \llbracket q_1 \rrbracket.\get \cdot \llbracket
+q_2 \rrbracket.\get, \text{ and }\\
+\llbracket q_1 \rrbracket.\lput &= \llbracket q_1 \rrbracket.\lput \cdot \llbracket
+q_2 \rrbracket.\lput
+\end{align*}
+By the induction hypothesis, there exist bijective lenses $\ell_1 : K(c_1)
+\Leftrightarrow K(d_1)$ and $\ell_2 : K(c_2) \Leftrightarrow K(d_2)$ such that
+\begin{align*}
+\llbracket q_1 \rrbracket.\get &= \llbracket \ell_1 \rrbracket \circ
+\canonizer(c_1)\\
+\llbracket q_1 \rrbracket.\lput &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
+\canonizer(d_1)
+\end{align*}
+and
+\begin{align*}
+\llbracket q_2 \rrbracket.\get &= \llbracket \ell_2 \rrbracket \circ
+\canonizer(c_2)\\
+\llbracket q_2 \rrbracket.\lput &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
+\canonizer(d_2)
+\end{align*}
+Consequently,
+\begin{align*}
+\llbracket q_1 \rrbracket.\get &= (\llbracket \ell_1 \rrbracket \circ
+\canonizer(c_1)) \cdot  (\llbracket \ell_2 \rrbracket \circ
+\canonizer(c_2))\\
+&= (\llbracket \ell_1 \rrbracket \cdot \llbracket \ell_2
+\rrbracket) \circ (\canonizer(c_1) \cdot \canonizer(c_2))\\
+&= \llbracket \ell_1 \cdot  \ell_2 \rrbracket \circ \canonizer(c_1 \cdot c_2)
+\end{align*}
+Similarly
+$$
+\llbracket q_1 \rrbracket.\lput = \llbracket \ell_1 \cdot  \ell_2 \rrbracket^{-1}
+\circ \canonizer(d_1 \cdot d_2) $$
+\item
+$q_1 = q_1 \sep q_2$ where $q_1 : c_1 \Leftrightarrow d_1 $, $q_2 : c_2
+\Leftrightarrow d_2$, $\mathcal{L}(W(c_1)) \cap \mathcal{L}(W(c_2)) =
+\varnothing$ and $\mathcal{L}(W(d_1)) \cap \mathcal{L}(W(d_2)) = \varnothing$.
+Then
+$$
+\llbracket q_1 \sep q_2 \rrbracket.\get(s) =
+\begin{cases}
+\llbracket q_1 \rrbracket.\get (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
+\llbracket q_2 \rrbracket.\get (s) & \text{if } s \in \mathcal{L}(W(c_2))\\
+\end{cases}$$
+$$\llbracket q_1 \sep q_2 \rrbracket.\lput(s) =
+\begin{cases}
+\llbracket q_1 \rrbracket.\lput (s) & \text{if } s \in \mathcal{L}(W(d_1))\\
+\llbracket q_2 \rrbracket.\lput (s) & \text{if } s \in \mathcal{L}(W(d_2))\\
+\end{cases}
+$$
+By the induction hypothesis, there exist bijective lenses $\ell_1 : K(c_1)
+\Leftrightarrow K(d_1)$ and $\ell_2 : K(c_2) \Leftrightarrow K(d_2)$ such that
+\begin{align*}
+\llbracket q_1 \rrbracket.\get &= \llbracket \ell_1 \rrbracket \circ
+\canonizer(c_1)\\
+\llbracket q_1 \rrbracket.\lput &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
+\canonizer(d_1)
+\end{align*}
+and
+\begin{align*}
+\llbracket q_2 \rrbracket.\get &= \llbracket \ell_2 \rrbracket \circ
+\canonizer(c_2)\\
+\llbracket q_2 \rrbracket.\lput &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
+\canonizer(d_2)
+\end{align*}
+Consequently,
+$$
+\llbracket q_1 \sep q_2 \rrbracket.\get(s) =
+\begin{cases}
+\llbracket \ell_1 \rrbracket \circ
+\canonizer(c_1) (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
+\llbracket \ell_2 \rrbracket \circ
+\canonizer(c_2) (s) & \text{if } s \in \mathcal{L}(W(c_2)),\\
+\end{cases}$$
+so $\llbracket q_1 \sep q_2 \rrbracket.\get = \llbracket \ell_1 \sep
+\ell_2 \rrbracket \circ \canonizer(c_1 \sep c_2)$. A similar argument shows
+that $\llbracket q_1 \sep q_2 \rrbracket.\lput = \llbracket \ell_1 \sep
+\ell_2 \rrbracket^{-1} \circ \canonizer(d_1 \sep d_2)$.\\
+This completes the proof.
+\end{enumerate}
+\end{proof}
+
 \bibliographystyle{plain}

### 2018-03-08 papers/QuotientLenses/main.tex
Responses to comments and updates to figure -- holding no locks

--- 2018-03-08 papers/QuotientLenses/main.tex
@@
 I'm not sure what the solution is to be honest.}
+\bcp{Maybe just admit a bit more frankly that this is a convenient trick
+  to avoid introducing a whole new language for writing (or synthesizing)
+  unidirectional functions.}
 
@@
 explicitly, somehow... any ideas?}}\sam{Maybe we can put all structures, the
-canonizer and the canonical strcutre in a box that says QRE?}
+canonizer and the canonical strcutre in a box that says QRE?}\bcp{Like this?
+ (But maybe we should go a step further and label things like $K(c)$,
+ $W(c)$, Canonizer(c), etc.  And probably we should display K as a subset of
+ W.  I can make these changes later.)}
 \label{fig:attheedges}
@@
 
-The formal semantics of QREs satisfy the following theorem:
+The formal semantics of QREs satisfy the following theorem:\bcp{That bit is
+  entirely redundant with the theorem itself.  Either simply delete it or
+  replace with something more interesting. :-)}
 \begin{theorem}

### 2018-03-08 papers/QuotientLenses/main.tex
More comments -- still not holding locks

@@
 normalize} combinator hence gives a sufficient condition for potential QREs to
-be ``well-behaved'', particularly with respect to our strategy for synthesizing
-quotient lenses which we shall introduce in Section~\ref{synth}.
+be ``well-behaved'', particularly with respect to the strategy for synthesizing
+quotient lenses which we shall introduce in Section~\ref{synth}.  \bcp{I
+  think here we're mixing concerns of language design with the concerns of
+  the underlying theory.  Indeed, I think it would be clearer to reorganize
+  the whole section as follows: first present just normalize and discuss its
+  properties; then observe that the side condition is annoying because it
+  cannot be checked automatically; then introduce project, squash, and perm
+  as a pragmatically motivated, extensible set of safe combinators that
+  we've found useful for a broad range of examples.  This will also help
+  address the complaint that extracting half of a lens is a hack.}
 
@@
 much easier than using just {\em normalize} because of the compositional nature
-of the other combinators.
+of the other combinators.  \bcp{Maybe this comment is not needed?  It seems
+  obvious, besides the fact that these other combinators are safe while
+  normalize requires a proof by hand.}
 
@@
 directly on regular languages.  QREs are therefore a good \textit{specification}
-language for quotient lenses. In this section of the paper, we introduce
+language for quotient lenses. We next introduce
 \textit{QRE Lenses}, a class of quotient lenses specified using QREs.
+
+\bcp{A question about variable names: If we use $R$ and $S$ for plain
+  regular expressions, it seems inconsistent to use $c$ for QREs.  (And even
+  more inconsistent to use $c$ and $c'$ or $c$ and $c_1$!) I suggest $P$ and
+  $Q$ for QREs (with $Q$ as the default when we are only talking about
+  one).}
 
@@
 $\sim_R$ and $\sim_S$ defined on $\mathcal{L}(R)$ and $\mathcal{L}(S)$
-respectively, we define a {\em bijective quotient lens} $q : R /{\sim_R}
+respectively, we\bcp{can we clarify that this is exactly Foster et al's definition?} define a {\em bijective quotient lens} $q : R /{\sim_R}
 \Leftrightarrow S/{\sim_S}$ to be a pair of functions $q.get :
@@
 Moreover, if $r \sim_R r'$, then $q.get(r) = q.get(r')$, and if $s \sim_S s'$,
-then $q.put(s) = q.put(s')$.
+then $q.put(s) = q.put(s')$.  \bcp{We should emphasize that this is
+  precisely the desired loosening of the too-strict plain lens laws.}
 
@@
 and $\sim_S$ is a pair of functions $q.get$ and $q.put$ such that $q.get$
-respects $\sim_R$ and $q.put$ respects $\sim_S$, and such that the respective
-functions defined on $\mathcal{L}(R)/{\sim_R}$ and $\mathcal{L}(S)/{\sim_S}$ are
-mutual inverses of each other.
+respects $\sim_R$ and $q.put$ respects $\sim_S$, and such that the $get$ and
+$put$ functions lifted to $\mathcal{L}(R)/{\sim_R}$ and
+$\mathcal{L}(S)/{\sim_S}$ are mutual inverses.
+
+\bcp{Typesetting: $q.get$ and $q.\get$ are subtly different, as are
+  $q.put$ and $q.\lput$, and the latter forms are the correct ones.
+  In general, you should never put a whole word in math mode without
+  wrapping it in mathit: the inter-letter spacing will be wrong.}
 
@@
 If there is a derivation $q : c \Leftrightarrow c_1$, then there exists a
-bijective lens $\ell : K(c) \Leftrightarrow K(c_1)$ such that
+bijective lens $\ell : K(c) \Leftrightarrow K(c_1)$ such that:
 \begin{align*}
@@
 \label{synth}
-Our goal in this section is to demonstrate how to synthesize a quotient lens
+We next demonstrate how to use previous results from~\cite{optician} to
+synthesize a quotient lens 
 $q: c \Leftrightarrow c_1$ from the QREs $c, c_1$ and a set of input-output pairs
@@
 
+\bcp{Another question about choice of variables: if we rename $c$ and $c_1$
+  to $P$ and $Q$, could we also choose names like $p$ and $q$ for their
+  elements?  (I guess $q$ was used to range over QRE-lenses, so we can't use
+  that suggestion without further renaming...)}
+
 Our approach in synthesizing QRE lenses is guided by Theorem~\ref{normal
-form}, which says that if there is a derivation $q :
+form}, which says that, if there is a derivation $q :
 c \Leftrightarrow c_1$ of a QRE lens, then there exists a bijective lens $\ell :
@@
 \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
-\canonizer(c_1)
+\canonizer(c_1).
 \end{align*}

### 2018-03-08 papers/QuotientLenses/main.tex
pass through sections 5 and 6

@@
 \label{synth}
-We next demonstrate how to use previous results from~\cite{optician} to
-synthesize a quotient lens 
-$q: c \Leftrightarrow c_1$ from the QREs $c, c_1$ and a set of input-output pairs
-$\{(x_1, y_1),\allowbreak \ldots,\allowbreak (x_n, y_n)\}$ where the $x_i$'s are in $W(c)$ and the
-$y_i$'s are in $W(c_1)$. We wish $q$ to map the equivalence class of $x_i$ to the
-equivalence class of $y_i$ and vice versa:
-\begin{align*}
-q.get(x_i) &\equiv_{c_1} y_i \text{, and}\\
-q.put(y_i) &\equiv_c x_i
+
+QRE lenses address some of the limitations of bijective lenses because a single
+lens program expresses both the canonizers and the transformation between kernel
+languages simultaneously, which reduces programmer effort.  But we can go even
+further by recognizing that the type structure of QRE lenses contains
+information that can be exploited to automatically synthesize lenses from their
+types.  Rather than writing the QRE lens manually, the programmer can instead
+specify the desired behavior of a lens by giving its interface types and providing
+examples, if necessary, to disambiguate among possible implementations.  This
+way of constructing lenses can often be much simpler than building them by
+hand---see the example of lens synthesis in
+Section~\ref{sec:examplesynth}.
+
+The Optician framework~\cite{optician} showed how to do such lens synthesis in
+the case for bijective lenses. Here we show how to reduce QRE lens synthesis to
+that case, so that we can re-use the Optician algorithm but in the more
+expressive context of QRE lenses.  The basic idea is straightforward: we run the
+Optician algorithm to synthesize a lens between the kernels of two QREs and then
+apply the canonizers at the edges to recover a lens between the whole
+languages.  This simple strategy turns out to be remarkably effective, and the
+idea of using Optician in this way is part the impetus for the design of our QRE
+lenses.
+
+The Optician synthesis algorithm works as follows. Given regular expressions $R$
+and $S$ and a set of example input--output pairs
+$\{(r_1, s_1),\allowbreak \ldots,\allowbreak (r_n, s_n)\}$, Optician will try to
+find a bijective lens $\ell : R \Leftrightarrow S$ that that agrees on the
+examples, \textit{i.e.}, it maps $r_i$ to $s_i$ in the forward direction and
+$s_i$ to $r_i$ in the backward direction.  The Optician algorithm is guaranteed
+to succeed if such a lens exists.    \saz{We should probably mention what happens if the examples aren't sufficient to uniquely
+  determine the lens.}
+
+% equivalent regular expressions $R' \equiv R$ and $S' \equiv S$
+
+% if there is a , $\ell'$
+% derivation 
+% then there is an algorithm that computes a bijective lens
+% $\ell' : R' \Leftrightarrow S'$ such that 
+% maps $r_i$ to $s_i$ in the forward direction and $s_i$ to $r_i$ in the backward
+% direction, and $\llbracket \ell \rrbracket = \llbracket \ell' \rrbracket$.
+
+In our setting, we want to synthesize a quotient lens $q: c \Leftrightarrow c_1$
+from the QREs $c_1$ and $c_2$ and a set of example input-output pairs
+$\{(x_1, y_1),\allowbreak \ldots,\allowbreak (x_n, y_n)\}$ where the $x_i$'s are
+in $W(c_1)$ and the $y_i$'s are in $W(c_2)$. We furthermore wish $q$ to map the
+equivalence class of $x_i$ to the equivalence class of $y_i$ and vice versa:
+\begin{align*}
+q.get(x_i) &\equiv_{c_2} y_i \text{, and}\\
+q.put(y_i) &\equiv_{c_1} x_i
 \end{align*}
@@
   elements?  (I guess $q$ was used to range over QRE-lenses, so we can't use
-  that suggestion without further renaming...)}
-
-Our approach in synthesizing QRE lenses is guided by Theorem~\ref{normal
-form}, which says that, if there is a derivation $q :
-c \Leftrightarrow c_1$ of a QRE lens, then there exists a bijective lens $\ell :
-K(c) \Leftrightarrow K(c_1)$ such that
-\begin{align*}
-\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ \canonizer(c)\\
+  that suggestion without further renaming...)}\saz{I like the idea of doing a
+  consistent renaming of the variables in the paper following this suggestion.
+  I would also suggest not using so many primes. }
+
+Our approach to synthesizing QRE lenses is guided by Theorem~\ref{normal form},
+which says that, if there is a derivation $q : c_1 \Leftrightarrow c_2$ of a QRE
+lens, then there exists a bijective lens $\ell : K(c_1) \Leftrightarrow K(c_2)$
+such that
+\begin{align*}
+\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ \canonizer(c_1)\\
 \llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
-\canonizer(c_1).
-\end{align*}
-
-For our synthesis task, since the $x_i$'s are in $W(c)$ and the $y_i$'s are
-in $W(c_1)$, then ${x_i}' = \canonizer(c)(x_i)$ is in $K(c)$ and ${y_i}' =
-\canonizer(c_1)(y_i)$ is in $K(c_1)$. Consequently in order to synthesize the
-desired quotient lens $q: c \Leftrightarrow c_1$ that is consistent with the
+\canonizer(c_2).
+\end{align*}
+
+For the QRE examples, the $x_i$'s are in $W(c_1)$ and the $y_i$'s are
+in $W(c_2)$, so we can construct ${x'_i} = \canonizer(c_1)(x_i)$ in $K(c_1)$ and ${y'_i} =
+\canonizer(c_2)(y_i)$ is in $K(c_2)$. To synthesize the
+desired quotient lens $q: c_1 \Leftrightarrow c_2$ that is consistent with the
 input-output examples $\{(x_1, y_1), \ldots, (x_n, y_n)\}$ it suffices to
 synthesize a bijective lens $l :
-K(c) \Leftrightarrow K(c_1)$ that is consistent with the canonized examples
-$\{({x_1}', {y_1}'), \ldots, ({x_n}', {y_n}')\}$
-
-Optician~\cite{optician} showed that given regular expressions
-$R, S$ and a set of input-output pairs $\{(r_1, s_1),\allowbreak \ldots,\allowbreak (r_n, s_n)\}$, if
-there is a derivation $\ell : R \Leftrightarrow S$ of a bijective lens that
-maps $r_i$ to $s_i$ in the forward direction and $s_i$ to $r_i$ in the backward
-direction, then there is an algorithm that computes a bijective lens $\ell' :
-R' \Leftrightarrow S'$ such that $R \equiv R'$, $S \equiv S'$, $\ell'$ maps
-$r_i$ to $s_i$ in the forward direction and $s_i$ to $r_i$ in the backward
-direction, and $\llbracket \ell \rrbracket = \llbracket \ell'
-\rrbracket$.
-
-This implies that if there is a derivation $q : c \Leftrightarrow
-c_1$ of a QRE lens as well as a derivation $\ell : K(c) \Leftrightarrow K(c_1)$
-that is consistent with the canonized examples $\{({x_1}', {y_1}'),\allowbreak
-\ldots,\allowbreak ({x_n}', {y_n}')\}$, then we can synthesize a quotient lens $q$ of
-type $c \Leftrightarrow c_1$ that is consistent with $\{({x_1}', {y_1}'),
-\ldots, ({x_n}', {y_n}')\}$. More concretely, in order to synthesize a
-quotient lens $q: c \Leftrightarrow c_1$, that is consistent with the examples
-$\{(x_1, y_1),\allowbreak \ldots, (x_n, y_n)\}$ it suffices to perform the following
-procedure:
+K(c_1) \Leftrightarrow K(c_2)$ that is consistent with the canonized examples
+$\{({x'_1}, {y'_1}), \ldots, ({x'_n}, {y'_n})\}$ and then apply the canonizers
+at the outside.  QRE lens synthesis is
+
+% This implies that if there is a derivation $q : c \Leftrightarrow
+% c_1$ of a QRE lens as well as a derivation $\ell : K(c) \Leftrightarrow K(c_1)$
+% that is consistent with the canonized examples $\{({x_1}', {y_1}'),\allowbreak
+% \ldots,\allowbreak ({x_n}', {y_n}')\}$, then we can synthesize a quotient lens $q$ of
+% type $c \Leftrightarrow c_1$ that is consistent with $\{({x_1}', {y_1}'),
+% \ldots, ({x_n}', {y_n}')\}$. More concretely, in order to synthesize a
+% quotient lens $q: c \Leftrightarrow c_1$, that is consistent with the examples
+% $\{(x_1, y_1),\allowbreak \ldots, (x_n, y_n)\}$ it suffices to perform the following
+% procedure:
 \begin{enumerate}
   \item
-  Compute $K(c)$, $K(c_1)$,
+  Compute $K(c_1)$, $K(c_2)$,
   \item
-  Compute $\{({x_1}', {y_1}'), \ldots, ({x_n}', {y_n}')\}$ where ${x_i}' =
-  \canonizer(c)(x_i)$ and ${y_i}' = \canonizer(c_1)(y_i)$,
+  Compute $\{({x'_1}, {y'_1}), \ldots, ({x'_n}, {y'_n})\}$ where ${x'_i} =
+  \canonizer(c_1)(x_i)$ and ${y'_i} = \canonizer(c_2)(y_i)$,
+\item Synthesize, using Optician's algorithm, a bijective lens
+  $\ell : K(c_1) \Leftrightarrow K(c_2)$ that is consistent with the examples
+  $\{({x'_1}, {y'_1}),\allowbreak \ldots,\allowbreak ({x'_n}, {y'_n})\}$
   \item
-  Synthesize a bijective lens $\ell : K(c) \Leftrightarrow K(c_1)$ that is
-  consistent with the examples $\{({x_1}', {y_1}'),\allowbreak
-  \ldots,\allowbreak ({x_n}',
-  {y_n}')\}$
-  \item
-  Return $q = \kw{rquot}(\kw{lquot}(c, \ell), c_1) : c \Leftrightarrow
-  c_1$
+  Return $q = \kw{rquot}(\kw{lquot}(c_1, \ell), c_2) : c_1 \Leftrightarrow
+  c_2$
 \end{enumerate}
-For example, recall that in the \bibtex{} to EndNote transformation we had the
-QREs \cd{bibtex} and \cd{endnote} describing a \bibtex{} record and an Endnote
-record respectively. We also had the example pair \cd{(bib_example ,
-end_example)} where
-
+
+\saz{Do we want to state formally the completeness of this algorithm?}
+
+Returning to the \bibtex{} to EndNote transformation of
+Section~\ref{sec:examplesynth}, we had the QREs \cd{bibtex} and \cd{endnote}
+describing a \bibtex{} record and an Endnote record respectively. We also had
+the example pair \cd{(bib_example , end_example)} where
+
+\begin{minipage}{3in}
 \begin{lstlisting}
@@
 
+All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
+of 1600 MHz DDR3 running macOS High Sierra.
+
 \subsection{Benchmark Suite Construction}
-Our benchmarks were constructed from two sources:
+
+To evaluate our QRE implementation, we constructed a set of lens synthesis tasks
+drawn from two sources:
 \begin{enumerate}
-  \item 39 of our benchmarks were derived from the Optician system.  These
-  benchmarks were a combination of custom benchmarks, benchmarks derived from
-  FlashFill, and benchmarks derived from Augeas.  Many of the formats mapped
-  between had to be modified to work with the bijectivity constraints
-  Optician required.  Commonly, one format had whitespace where the other did
-  not allow for whitespace.  To address the bijectivity issues surrounding this,
-  we added whitespace to the format that typically did not have whitespace.  We
-  were able to remove these alterations in adopting these benchmarks to our
-  system.
-
-  \item 8 of our benchmarks were derived from data.gov.  Data.gov contains
-  many transformations between formats.  In addition to the formats previously
-  adopted for Optician, we synthesized transformers between these formats.
+\item 39 of our benchmarks were adapted from the evaluation of the original
+  Optician system.  These benchmarks are a combination of custom benchmarks,
+  benchmarks derived from FlashFill~\cite{}\saz{TODO:cite}, and benchmarks
+  derived from Augeas~\cite{}\saz{TODO:cite}.  In the pre-QRE version of
+  Optician, many of these data formats had to be modified to work with the
+  bijectivity constraints that Optician required.  For instance, when one format
+  permits whitespace where the other does not allow it, the original versions of
+  benchmark formats were modified to allow more whitespace, thereby restoring
+  bijectivity.  With the new QRE support, we were able to remove these
+  alterations when porting these benchmarks.  (This experience alone suggests
+  that QREs make the lens development process easier.)
+
+\item The remaining eight benchmarks were derived from
+  data.gov.\footnote{\url{http://data.gov}} Data.gov contains many
+  transformations between formats.  In addition to the formats previously
+  used for Optician, we synthesized transformers between these formats.\saz{This
+    description is not specific enough.  What are the data formats for?  How did
+    we pick which ones to try?}
 \end{enumerate}
@@
 
-To test if our tool eases the development process, we compare the following
-number of AST nodes.
+To get a sense of how much the new expressiveness that combining QRE lenses with
+synthesis saves in terms of programmer effort, we compared the amount of code
+that would need to be written for each of the lens benchmarks.  We use number of
+nodes in the abstract syntax tree as a measure of code size, and calculated
+three quantities for each lens synthesis benchmark task.  \saz{Say something
+  about why there are only 14 benchmarks?}
 %
 \begin{itemize}
-  \item[\QRESize{}] is the number of AST nodes in the QRE
-  specifications involved in QRE lens synthesis task.
-  \item[\CanonizerAndSpecSize{}] is the number of
-  AST nodes in $W(q)$ summed with the number of AST nodes in $\canonizer(q)$,
-  for each QRE used in QRE lens synthesis task.
-  \item[\LensAndSpecSize{}] is the number of AST nodes
+  \item[\QRESize{}] the number of AST nodes in the QRE
+  specifications (including examples)\saz{CHECK THIS}
+  \item[\CanonizerAndSpecSize{}]  the number of
+  AST nodes in $W(q)$ summed with the number of AST nodes in $\canonizer(q)$
+  \item[\LensAndSpecSize{}] the number of AST nodes
   in $W(q)$ summed with the number of AST nodes in the synthesized QRE lens.
@@
 
-The value of \QRESize{} represents the difficulty in writing our benchmarks.
-
-The value of \CanonizerAndSpecSize{} represents the difficulty in writing our
-benchmarks within the Optician system. This is only an approximation, as both
-$W(q)$ and $\canonizer(q)$ are automatically generated from the QRE, instead of
-written by a human.
-
-The value of \LensAndSpecSize{} represents the difficulty in writing our
-benchmarks in Boomerang with neither QREs nor synthesis. This is also an
+The value of \QRESize{} is our measure of the programmer burden for
+writing our benchmarks---it measures the number of AST nodes the programmer has
+to write using the full expressiveness of QREs and synthesis.
+
+The value of \CanonizerAndSpecSize{} estimates the burden for writing our
+benchmarks within the Optician system, without QREs but still using
+synthesis. This is only an approximation, as both $W(q)$ and $\canonizer(q)$ are
+automatically generated from the QRE, and it is possible that a human-written
+version might be smaller.
+
+The value of \LensAndSpecSize{} estimates the burden in writing our benchmarks
+in Boomerang, using neither QREs nor synthesis. This value is also an
 approximation, as $W(q)$ and $\canonizer(q)$ and the bijective lens are all
@@
 
-Figure~\ref{fig:asts} shows the number of benchmarks written with at most a
-given number of AST nodes. For a given benchmark, we found \QRESize{} to be
-about half as large as \CanonizerAndSpecSize, and about a third as large as
-\LensAndSpecSize{}. This demonstrates that introducing QREs saves programmers
+Figure~\ref{fig:asts} shows the number of benchmarks that can be expressed using
+at most a given number of AST nodes. For a given benchmark, we found \QRESize{}
+to be about half as large as \CanonizerAndSpecSize, and about a third as large
+as \LensAndSpecSize{}. This demonstrates that introducing QREs saves programmers
 significant effort compared to both Optician and basic Boomerang.
@@
 
-To be used in a development environment, we found it important to validate
-performance did not significantly degrade from Optician. To validate this, we
-tested the runtimes when:
+To assess the performance impact of adding QRE support to Optician, we measured
+the running time of the following configurations:
 \begin{itemize}
-  \item[\OpticianRuntime{}] Optician runs on its own benchmarks.
-  \item[\SystemOnOptician{}] Our system runs on Optician's benchmarks.
-  \item[\SystemOnBenchmarks{}] Our system runs on its own benchmarks.
+  \item[\OpticianRuntime{}] pre-QRE Optician running on its own benchmarks.
+  \item[\SystemOnOptician{}] Optician with QREs running on the pre-QRE benchmarks.
+  \item[\SystemOnBenchmarks{}] Optician with QREs running on its own benchmarks.
 \end{itemize}
 
-We summarize the results of these experiments in Figure~\ref{fig:times}.
-Our system was
-able to synthesize all of Optician's benchmarks at a speed competitive with
-Optician.  There was additional overhead in calculating $W(q)$ and $K(q)$,
+We summarize the results of these experiments in Figure~\ref{fig:times}.  The
+new version of Optician was able to synthesize all of pre-QRE Optician's
+benchmarks at a speed competitive with the old version.  There is a small amount
+of additional overhead introduced by QREs when calculating $W(q)$ and $K(q)$,
 resulting in a very slight decrease in performance.
 
-Also notable was one of the data.gov examples, which converted demographic
-statistics by zip code from xml form into rdf form. This merely took longer
-because it was an exceptionally large format, converting over 40 fields. We
-added in additional fields to make this format suitable for Optician, which
-took over 15 seconds synthesize the modified lens.
+Also notable is one of the data.gov examples, which converts demographic
+statistics by zip code from xml form into rdf form. This took longer because it
+was an exceptionally large format, converting over 40 fields.\saz{We need to
+  explain more detail here.  I guess this is about the last benchmark in Fig. 11
+  (b), which takes a long time?}  We added in additional fields to make this
+format suitable for Optician, which took over 15 seconds synthesize the modified
+lens.\saz{This needs more context, explanation.}
 

### 2018-03-09 papers/QuotientLenses/main.tex
A few comments on comments

@@
 
-\sam{Here is my response to this. You observe that the idea of using half a lens
-as a canonizer comes out of nowhere, and that this ideas is a hack and always
-was. I don't think that's quite right. I think that it is useful to have a
-combinator which chooses between two or more formats. Indeed, the permutation
-combinator is just a special case of this. The bibtex name field is a good
-example of where such a combinator would be useful.  But if we choose to have
-such a combinator, then we need a way of mapping one format to another, and
-really the only interesting string-to-string transformations in Boomerang are
-given by lenses, which is why we use the get function of a lens in a canonizer.
-We could have used any other function between the two formats but Boomerang
-doesn't give us too many options for writing unidirectional transformations. So
-I don't think that using a lens as a canonizer is a trick. I think that
-Boomerang just isn't designed for unidirectional tranformations. But I also
-agree that it doesn't make much sense to synthesize a lens only to get rid of
-one half of it.
-I'm not sure what the solution is to be honest.}
-\bcp{Maybe just admit a bit more frankly that this is a convenient trick
-  to avoid introducing a whole new language for writing (or synthesizing)
-  unidirectional functions.}
-
-The first line above synthesizes a
-lens between \codefont{comma\_name} and \codefont{space\_name} using the listed
-example transformation as a guide.  The second line simply
-extracts the get direction transformation from the lens,
+\sam{OK I suppose it's a trick though I don't see an obvious solution. }
+
+The first line above synthesizes a lens between \codefont{comma\_name} and
+\codefont{space\_name} using the listed example transformation as a guide.  The
+second line simply extracts the get direction transformation from the lens,
 which is what we need for \kw{squash}.
@@
 \caption{Quotient Lens with ``Canonizers at the Edges''
-\sam{We don't need a general choose function. With QREs the kernel language is
-always contained in the whole language so we just use the inclusion as the
-choose function.}\bcp{Done.  But I wonder if the picture should mention QREs
-explicitly, somehow... any ideas?}}\sam{Maybe we can put all structures, the
-canonizer and the canonical strcutre in a box that says QRE?}\bcp{Like this?
- (But maybe we should go a step further and label things like $K(c)$,
+\bcp{Maybe we should go a step further and label things like $K(c)$,
  $W(c)$, Canonizer(c), etc.  And probably we should display K as a subset of
- W.  I can make these changes later.)}
+ W.  I can make these changes later.)}}
 \label{fig:attheedges}
 \end{figure}
-Every QRE lens $q$ has a type $q: c_1 \Leftrightarrow c_2$ where $c_1$ and $c_2$
-are QREs. In the get direction, a QRE lens $q: c_1 \Leftrightarrow c_2$ uses the
+Every QRE lens $q_1$ has a type $q_1: c_1 \Leftrightarrow c_2$ where $c_1$ and $c_2$
+are QREs. In the get direction, a QRE lens $q_1: c_1 \Leftrightarrow c_2$ uses the
 source QRE $c_1$ to compute a unique representative for the data modulo the
 equivalence relation defined by $c_1$ and then applies the $\get$ function of a
-bijective lens $\ell$ to this representative. In the backward direction, $q$
+bijective lens $\ell_1$ to this representative. In the backward direction, $q_1$
 operates similarly, but using the view QRE $c_2$ and the $\lput$ function of
-$\ell$.
+$\ell_1$.
 
@@
 
-At this point we also bring to attention the fact that in order for the {\em
-normalize} QRE to be well-formed, then canonizing function $f$ used by this
-combinator must be surjective and idempotent. Unfortunately, verifying
-that a function satisfies these properties is in general undecidable.
-Nonetheless, we allow a programmer to use normalize at their own risk without
-checking these side conditions.
-
-We do this firstly because the {\em normalize} combinator allows one to express
-equivalence relations that are either difficult to define, or that cannot be
-defined using the other combinators. Secondly, the {\em normalize} canonizer
-enjoys the property that each of the other combinators $c$ is semantically
-equivalent to $\normalize{W(c)}{K(c)}{f}$ for some surjective, idempotent
-function $f : \mathcal{L}(W(c)) \longrightarrow \mathcal{L}(K(c))$. The {\em
-normalize} combinator hence gives a sufficient condition for potential QREs to
-be ``well-behaved'', particularly with respect to the strategy for synthesizing
-quotient lenses which we shall introduce in Section~\ref{synth}.  \bcp{I
-  think here we're mixing concerns of language design with the concerns of
-  the underlying theory.  Indeed, I think it would be clearer to reorganize
-  the whole section as follows: first present just normalize and discuss its
-  properties; then observe that the side condition is annoying because it
-  cannot be checked automatically; then introduce project, squash, and perm
-  as a pragmatically motivated, extensible set of safe combinators that
-  we've found useful for a broad range of examples.  This will also help
-  address the complaint that extracting half of a lens is a hack.}
-
-In spite of this ``completeness'' property of the {\em normalize} combinator,
-our experience with QREs shows that using the {\em project}, {\em squash} and
-{\em perm} combinators as well as the regular operators applied to these is
-much easier than using just {\em normalize} because of the compositional nature
-of the other combinators.  \bcp{Maybe this comment is not needed?  It seems
-  obvious, besides the fact that these other combinators are safe while
-  normalize requires a proof by hand.}
-
-The formal semantics of QREs satisfy the following theorem:\bcp{That bit is
-  entirely redundant with the theorem itself.  Either simply delete it or
-  replace with something more interesting. :-)}
-\begin{theorem}
-If $c$ is a well formed QRE, then
-\begin{enumerate}
-  \item $W(c)$ and $K(c)$ are well defined regular expressions,
-  \item  $\eqrel{c}$ is an equivalence relation on $\mathcal{L}(W(c))$,
-  \item  $\mathcal{L}(K(c))$ forms a complete set of representatives for
-  $\eqrel{c}$,
-  \item $\canonizer(c):\mathcal{L}(W(c)) \longrightarrow \mathcal{L}(K(c))$ is
-  a well-defined function, and
-  \item  given any $w \in \mathcal{L}(W(c))$, $\canonizer(c)(w)$ is the unique
-  $k$ in $\mathcal{L}(K(c))$ such that $k$ is equivalent to $w$ modulo
-  $\eqrel{c}$.
-\end{enumerate}
-\end{theorem}
-
 \section{QRE Lenses}
@@
   $Q$ for QREs (with $Q$ as the default when we are only talking about
-  one).}
+  one).}\sam{I made the notation consistent by just using $c_1$ and $c_2$. I
+  personally feel that it is better not to use P and Q for QREs since we use p
+  and q for QRE lenses, but if you disagree I will make the change.}\bcp{I
+  agree that this is a problem.  But maybe the right fix is to choose
+  different metavariables for QRE lenses.  Or to use capital letters other
+  than $P$ and $Q$ for QREs.  I don't like $c_1$ and $c_2$, both for the
+  lower-case letter and for the subscripts---it's not at all consistent with
+our conventions elsewhere.}
 
@@
 $\sim_R$ and $\sim_S$ defined on $\mathcal{L}(R)$ and $\mathcal{L}(S)$
-respectively, we\bcp{can we clarify that this is exactly Foster et al's definition?} define a {\em bijective quotient lens} $q : R /{\sim_R}
-\Leftrightarrow S/{\sim_S}$ to be a pair of functions $q.get :
-\mathcal{L}(R) \longrightarrow \mathcal{L}(S)$ and $q.put : \mathcal{L}(S)
+respectively, we\bcp{can we clarify that this is exactly Foster et al's
+definition?}\sam{this is not Foster et al's definition. It's just a
+loosening of the definition for bijective lens used in the Optician
+paper}\bcp{OK, then we should say this}define
+a {\em bijective quotient lens} $q :
+R /{\sim_R} \Leftrightarrow S/{\sim_S}$ to be a pair of functions $q.\get :
+\mathcal{L}(R) \longrightarrow \mathcal{L}(S)$ and $q.\lput : \mathcal{L}(S)
 \longrightarrow \mathcal{L}(R)$ such that for all $r \in \mathcal{L}(R)$ and $s
-\in \mathcal{L}(S)$,
-\begin{align*}
-q.put(q.get(r)) &\sim_R r \text{, and}\\
-q.get(q.put(s)) &\sim_R s
-\end{align*}
-Moreover, if $r \sim_R r'$, then $q.get(r) = q.get(r')$, and if $s \sim_S s'$,
-then $q.put(s) = q.put(s')$.  \bcp{We should emphasize that this is
-  precisely the desired loosening of the too-strict plain lens laws.}
+\in \mathcal{L}(S)$, then $q.\lput(q_1.\get(r)) \sim_R r$ and
+$q.\get(q_1.\lput(s)) \sim_R s$. Moreover, if $r \sim_R r'$, then $q.\get(r) =
+q.\get(r')$, and if $s \sim_S s'$, then $q.\lput(s) = q.\lput(s')$.
 
 In words, a bijective quotient lens $q$ from $R$ to $S$ modulo $\sim_R$
-and $\sim_S$ is a pair of functions $q.get$ and $q.put$ such that $q.get$
-respects $\sim_R$ and $q.put$ respects $\sim_S$, and such that the $get$ and
-$put$ functions lifted to $\mathcal{L}(R)/{\sim_R}$ and
+and $\sim_S$ is a pair of functions $q.\get$ and $q.\lput$ such that
+$q.\get$ respects $\sim_R$ and $q.\lput$ respects $\sim_S$, and such that
+the {\em get} and {\em put} functions lifted to $\mathcal{L}(R)/{\sim_R}$ and
 $\mathcal{L}(S)/{\sim_S}$ are mutual inverses.
 
-\bcp{Typesetting: $q.get$ and $q.\get$ are subtly different, as are
-  $q.put$ and $q.\lput$, and the latter forms are the correct ones.
-  In general, you should never put a whole word in math mode without
-  wrapping it in mathit: the inter-letter spacing will be wrong.}
-
-Now let $c$ and $c_1$ be QREs. Given a bijection $\ell : K(c) \Leftrightarrow
-K(c_1)$, we may define a quotient lens $q : W(c)/\eqrel{c} \Leftrightarrow
-W(c_1)/\eqrel{c_1}$ by
-\begin{equation}\label{normalform}
-q.get = \ell \circ \canonizer(c) \text{ and } q.put = \ell^{-1} \circ
-\canonizer(c_1),
-\end{equation}
-
-\noindent In other words, the get function of the $q$ first canonizes the source
-data using $\canonizer(c)$ and then applies $f$ to the result, while the
-put function of the $q$ first canonizes the target data using
-$\canonizer(c_1)$ and the applies $f^{-1}$ to the result. This is our approach in
-defining QRE lenses.
+These laws are similar to the laws for bijective lenses, except that the
+equality restriction in the bijective lens laws are loosened to allow for
+equivalence relations. Also, the condition that $r \sim_R r'$ implies
+$q.\get(r) = q.\get(r')$ ensures that the {\em get} function induced on
+the equivalence classes if $\equiv_c$ is well-defined, and similarly for
+the {\em put} function.
+
+Now let $c_1$ and $c_2$ be QREs. Given a bijection $\ell : K(c_1)
+\Leftrightarrow K(c_2)$, we may define a quotient lens $q : W(c_1)/\eqrel{c_1}
+\Leftrightarrow W(c_2)/\eqrel{c_2}$ by $q.\get = \ell \circ \canonizer(c_1)
+$ and $q.\lput = \ell_1^{-1} \circ \canonizer(c_2)$. In other words, the get
+function of the $q$ first canonizes the source data using $\canonizer(c_1)$
+and then applies $f$ to the result, while the put function of the $q$ first
+canonizes the target data using $\canonizer(c_2)$ and the applies $f^{-1}$ to
+the result. This is our approach in defining QRE lenses.
 
@@
 \begin{theorem}\label{normal form}
-If there is a derivation $q : c \Leftrightarrow c_1$, then there exists a
-bijective lens $\ell : K(c) \Leftrightarrow K(c_1)$ such that:
-\begin{align*}
-\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ \canonizer(c)\\
-\llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
-\canonizer(c_1)
+If there is a derivation $q : c_1 \Leftrightarrow c_2$, then there exists a
+bijective lens $\ell : K(c_1) \Leftrightarrow K(c_2)$ such that:
+\begin{align*}
+\llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket\circ
+\canonizer(c_1)\\
+\llbracket q \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
+\canonizer(c_2)
 \end{align*}
@@
 straightforward application of the induction hypothesis followed by unrolling
-the equations that give the denotation for QRE lenses.\qed
+the equations that give the denotation for QRE lenses.
 \end{proof}
-\iffalse
-\begin{proof}
-Assume that $q : c \Leftrightarrow c_1$. We proceed by induction over the
-derivation $q : c \Leftrightarrow c_1$.
-\begin{enumerate}
-  \item
-  $\kw{lift}(\ell): R/\mathit{id}(R) \Leftrightarrow S/\mathit{id}(S)$ where
-  $\ell :
-  R \Leftrightarrow S$. Then
-  \begin{align*}
-\llbracket \kw{lift}(\ell) \rrbracket.get &=  \llbracket \ell \rrbracket
-= \llbracket \ell \rrbracket \circ id_{\mathcal{L}(R)} =
-\llbracket \ell \rrbracket \circ \canonizer(\mathit{id}(R)), \text{ and }\\
-\llbracket \kw{lift}(\ell) \rrbracket.put &= \llbracket \ell
-\rrbracket^{-1} = \llbracket \ell \rrbracket^{-1} \circ id_{\mathcal{L}(S)} =
-\llbracket \ell \rrbracket^{-1} \circ \canonizer(id(S))
-\end{align*}
-\item
-$\kw{lquot}(c, q'): c_1 \circ c \Leftrightarrow c_2$ where $q' : c_1
-\Leftrightarrow c_2$, $c$ is well formed and $K(c) = W(c_1)$. Then
-\begin{align*}
-\llbracket q \rrbracket.get  &= \llbracket q'
-\rrbracket.get \circ \canonizer(c)\\
-\llbracket q \rrbracket.put &= \llbracket q' \rrbracket.put
-\end{align*}
-By the induction hypothesis, there exists a bijective lens $\ell :
-K(c_1) \Leftrightarrow K(c_2)$ such that
-\begin{align*}
-\llbracket q' \rrbracket.get &= \llbracket \ell \rrbracket \circ \canonizer(c_1)\\
-\llbracket q' \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
-Canonize(c_2)
-\end{align*}
-Consequently
-\begin{align*}
-\llbracket q \rrbracket.get  &= (\llbracket \ell \rrbracket \circ
-\canonizer(c_1)) \circ \canonizer(c) = \llbracket \ell \rrbracket \circ
-(\canonizer(c_1 \circ c))\\
-\llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
-Canonize(c_2)
-\end{align*}
-
-\item
-$\kw{rquot}(q', c_2):c \Leftrightarrow c_2 \circ c_1$ where $q' : c \Leftrightarrow
-c_1$, $c_2$ is well formed and $K(c_2) = W(c_1)$. Proceed as in the previous
-case.
-\item
-$q_2 \circ q_1: c \Leftrightarrow c_2$ where $q_1 : c \Leftrightarrow c_1$ and
-$q_2 : c_1 \Leftrightarrow c_2$. Then
-\begin{align*}
-\llbracket q \rrbracket.get &= \llbracket q_2 \rrbracket.get\circ \llbracket
-q_1 \rrbracket.get, \text{ and }\\
-\llbracket q \rrbracket.put &= \llbracket q_1 \rrbracket.put \circ \llbracket
-q_2 \rrbracket.put
-\end{align*}
-By the induction hypothesis, there exist bijective lenses
-$\ell_1 :
-K(c) \Leftrightarrow K(c_1)$ and $\ell_2 : K(c_1) \Leftrightarrow K(c_2)$ such
-that
-\begin{align*}
-\llbracket q_1 \rrbracket.get &= \llbracket \ell_1 \rrbracket \circ
-\canonizer(c)\\
-\llbracket q_1 \rrbracket.put &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
-\canonizer(c_1)
-\end{align*}
-and
-\begin{align*}
-\llbracket q_2 \rrbracket.get &= \llbracket \ell_2 \rrbracket \circ
-Canonize(c_1)\\
-\llbracket q_2 \rrbracket.put &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
-Canonize(c_2)
-\end{align*}
-Consequently,
-\begin{align*}
-\llbracket q_2 \rrbracket.get \circ \llbracket q_1 \rrbracket.get &=
-(\llbracket \ell_2 \rrbracket \circ \canonizer(c_1)) \circ (\llbracket \ell_1
-\rrbracket \circ \canonizer(c))\\
-&= \llbracket \ell_2 \rrbracket \circ (\canonizer(c_1) \circ \llbracket \ell_1
-\rrbracket) \circ \canonizer(c)\\
-&= (\llbracket \ell_2 \rrbracket \circ \llbracket \ell_1 \rrbracket) \circ
-\canonizer(c)\\
-&= \llbracket \ell_2  \circ  \ell_1 \rrbracket \circ
-\canonizer(c)
-\end{align*}
-A similar argument shows that
-$$\llbracket q_1 \rrbracket.put \circ \llbracket q_2 \rrbracket.put =
-\llbracket \ell_2  \circ  \ell_1 \rrbracket^{-1} \circ
-\canonizer(c)$$
-\item
-${q'}^* : c^* \Leftrightarrow {c_1}^*$ where $q' : c \Leftrightarrow c_1$,
-$W(c)^{*!}$ and $W(c_1)^{*!}$ and $K(c)^{*!}$ and $K(c_1)^{*!}$. Then
-\begin{align*}
-\llbracket {q'}^* \rrbracket.get &= (\llbracket q' \rrbracket.get)^*, \text{
-and }\\
-\llbracket {q'}^* \rrbracket.put &= (\llbracket q' \rrbracket.put)^*
-\end{align*}
-By the induction hypothesis there exists a bijective lens $\ell : K(c)
-\Leftrightarrow K(c_1)$ such that
-that
-\begin{align*}
-\llbracket q' \rrbracket.get &= \llbracket \ell \rrbracket \circ
-\canonizer(c)\\
-\llbracket q' \rrbracket.put &= {\llbracket \ell \rrbracket}^{-1} \circ
-\canonizer(c_1)
-\end{align*}
-Consequentlty
-\begin{align*}
-\llbracket {q'}^* \rrbracket.get &= (\llbracket \ell \rrbracket \circ
-\canonizer(c))^* = \llbracket \ell \rrbracket^* \circ
-\canonizer(c)^* = \llbracket \ell^* \rrbracket \circ
-\canonizer(c^*)\\
-\llbracket {q'}^* \rrbracket.put &= (\llbracket \ell \rrbracket^{-1} \circ
-\canonizer(c_1))^* = (\llbracket \ell \rrbracket^{-1})^* \circ
-\canonizer(c_1)^* = \llbracket \ell^* \rrbracket^{-1} \circ
-\canonizer(c_1^*)\\
-\end{align*}
-\item
-$q_1 \cdot q_2: c_1 \cdot c_2 \Leftrightarrow d_1 \cdot d_2$, where $q_1 : c_1
-\Leftrightarrow d_1 $,  $q_2 : c_2 \Leftrightarrow d_2$, $W(c_1)
-\cdot^! W(c_2)$, $K(c_1) \cdot^! K(c_2)$, $W(d_1) \cdot^! W(d_2)$ and $
-K(d_1) \cdot^! K(d_2)$. Then
-\begin{align*}
-\llbracket q \rrbracket.get &= \llbracket q_1 \rrbracket.get \cdot \llbracket
-q_2 \rrbracket.get, \text{ and }\\
-\llbracket q \rrbracket.put &= \llbracket q_1 \rrbracket.put \cdot \llbracket
-q_2 \rrbracket.put
-\end{align*}
-By the induction hypothesis, there exist bijective lenses $\ell_1 : K(c_1)
-\Leftrightarrow K(d_1)$ and $\ell_2 : K(c_2) \Leftrightarrow K(d_2)$ such that
-\begin{align*}
-\llbracket q_1 \rrbracket.get &= \llbracket \ell_1 \rrbracket \circ
-\canonizer(c_1)\\
-\llbracket q_1 \rrbracket.put &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
-\canonizer(d_1)
-\end{align*}
-and
-\begin{align*}
-\llbracket q_2 \rrbracket.get &= \llbracket \ell_2 \rrbracket \circ
-\canonizer(c_2)\\
-\llbracket q_2 \rrbracket.put &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
-\canonizer(d_2)
-\end{align*}
-Consequently,
-\begin{align*}
-\llbracket q \rrbracket.get &= (\llbracket \ell_1 \rrbracket \circ
-\canonizer(c_1)) \cdot  (\llbracket \ell_2 \rrbracket \circ
-\canonizer(c_2))\\
-&= (\llbracket \ell_1 \rrbracket \cdot \llbracket \ell_2
-\rrbracket) \circ (\canonizer(c_1) \cdot \canonizer(c_2))\\
-&= \llbracket \ell_1 \cdot  \ell_2 \rrbracket \circ \canonizer(c_1 \cdot c_2)
-\end{align*}
-Similarly
-$$
-\llbracket q \rrbracket.put = \llbracket \ell_1 \cdot  \ell_2 \rrbracket^{-1}
-\circ \canonizer(d_1 \cdot d_2) $$
-\item
-$q = q_1 \sep q_2$ where $q_1 : c_1 \Leftrightarrow d_1 $, $q_2 : c_2
-\Leftrightarrow d_2$, $\mathcal{L}(W(c_1)) \cap \mathcal{L}(W(c_2)) =
-\varnothing$ and $\mathcal{L}(W(d_1)) \cap \mathcal{L}(W(d_2)) = \varnothing$.
-Then
-$$
-\llbracket q_1 \sep q_2 \rrbracket.get(s) =
-\begin{cases}
-\llbracket q_1 \rrbracket.get (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
-\llbracket q_2 \rrbracket.get (s) & \text{if } s \in \mathcal{L}(W(c_2))\\
-\end{cases}$$
-$$\llbracket q_1 \sep q_2 \rrbracket.put(s) =
-\begin{cases}
-\llbracket q_1 \rrbracket.put (s) & \text{if } s \in \mathcal{L}(W(d_1))\\
-\llbracket q_2 \rrbracket.put (s) & \text{if } s \in \mathcal{L}(W(d_2))\\
-\end{cases}
-$$
-By the induction hypothesis, there exist bijective lenses $\ell_1 : K(c_1)
-\Leftrightarrow K(d_1)$ and $\ell_2 : K(c_2) \Leftrightarrow K(d_2)$ such that
-\begin{align*}
-\llbracket q_1 \rrbracket.get &= \llbracket \ell_1 \rrbracket \circ
-\canonizer(c_1)\\
-\llbracket q_1 \rrbracket.put &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
-\canonizer(d_1)
-\end{align*}
-and
-\begin{align*}
-\llbracket q_2 \rrbracket.get &= \llbracket \ell_2 \rrbracket \circ
-\canonizer(c_2)\\
-\llbracket q_2 \rrbracket.put &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
-\canonizer(d_2)
-\end{align*}
-Consequently,
-$$
-\llbracket q_1 \sep q_2 \rrbracket.get(s) =
-\begin{cases}
-\llbracket \ell_1 \rrbracket \circ
-\canonizer(c_1) (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
-\llbracket \ell_2 \rrbracket \circ
-\canonizer(c_2) (s) & \text{if } s \in \mathcal{L}(W(c_2)),\\
-\end{cases}$$
-so $\llbracket q_1 \sep q_2 \rrbracket.get = \llbracket \ell_1 \sep
-\ell_2 \rrbracket \circ \canonizer(c_1 \sep c_2)$. A similar argument shows
-that $\llbracket q_1 \sep q_2 \rrbracket.put = \llbracket \ell_1 \sep
-\ell_2 \rrbracket^{-1} \circ \canonizer(d_1 \sep d_2)$.\\
-This completes the proof.
-\end{enumerate}
-\end{proof}
-
-\fi
+
 \section{Synthesizing QRE Lenses}
@@
 $s_i$ to $r_i$ in the backward direction.  The Optician algorithm is guaranteed
-to succeed if such a lens exists.    \saz{We should probably mention what happens if the examples aren't sufficient to uniquely
-  determine the lens.}
-
-% equivalent regular expressions $R' \equiv R$ and $S' \equiv S$
-
-% if there is a , $\ell'$
-% derivation 
-% then there is an algorithm that computes a bijective lens
-% $\ell' : R' \Leftrightarrow S'$ such that 
-% maps $r_i$ to $s_i$ in the forward direction and $s_i$ to $r_i$ in the backward
-% direction, and $\llbracket \ell \rrbracket = \llbracket \ell' \rrbracket$.
-
-In our setting, we want to synthesize a quotient lens $q: c \Leftrightarrow c_1$
-from the QREs $c_1$ and $c_2$ and a set of example input-output pairs
+to succeed if such a lens exists.    \saz{We should probably mention what
+happens if the examples aren't sufficient to uniquely determine the lens.}
+
+In our setting, we want to synthesize a quotient lens $q: c_1 \Leftrightarrow
+c_2$ from the QREs $c_1$ and $c_2$ and a set of example input-output pairs
 $\{(x_1, y_1),\allowbreak \ldots,\allowbreak (x_n, y_n)\}$ where the $x_i$'s are
@@
 \begin{align*}
-q.get(x_i) &\equiv_{c_2} y_i \text{, and}\\
-q.put(y_i) &\equiv_{c_1} x_i
-\end{align*}
-
-\bcp{Another question about choice of variables: if we rename $c$ and $c_1$
-  to $P$ and $Q$, could we also choose names like $p$ and $q$ for their
-  elements?  (I guess $q$ was used to range over QRE-lenses, so we can't use
-  that suggestion without further renaming...)}\saz{I like the idea of doing a
-  consistent renaming of the variables in the paper following this suggestion.
-  I would also suggest not using so many primes. }
-
+q.\get(x_i) &\equiv_{c_2} y_i \text{, and}\\
+q.\lput(y_i) &\equiv_{c_1} x_i
+\end{align*}
 Our approach to synthesizing QRE lenses is guided by Theorem~\ref{normal form},
-which says that, if there is a derivation $q : c_1 \Leftrightarrow c_2$ of a QRE
-lens, then there exists a bijective lens $\ell : K(c_1) \Leftrightarrow K(c_2)$
-such that
-\begin{align*}
-\llbracket q \rrbracket.get &= \llbracket \ell \rrbracket\circ \canonizer(c_1)\\
-\llbracket q \rrbracket.put &= \llbracket \ell \rrbracket^{-1} \circ
+which says that, if there is a derivation $q : c_1 \Leftrightarrow c_2$ of a
+QRE lens, then there exists a bijective lens $\ell : K(c_1) \Leftrightarrow
+K(c_2)$ such that
+\begin{align*}
+\llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket\circ \canonizer(c_1)\\
+\llbracket q \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
 \canonizer(c_2).
@@
 For the QRE examples, the $x_i$'s are in $W(c_1)$ and the $y_i$'s are
-in $W(c_2)$, so we can construct ${x'_i} = \canonizer(c_1)(x_i)$ in $K(c_1)$ and ${y'_i} =
-\canonizer(c_2)(y_i)$ is in $K(c_2)$. To synthesize the
+in $W(c_2)$, so we can construct ${x'_i} = \canonizer(c_1)(x_i)$ in $K(c_1)$
+and ${y'_i} = \canonizer(c_2)(y_i)$ is in $K(c_2)$. To synthesize the
 desired quotient lens $q: c_1 \Leftrightarrow c_2$ that is consistent with the
 input-output examples $\{(x_1, y_1), \ldots, (x_n, y_n)\}$ it suffices to
-synthesize a bijective lens $l :
-K(c_1) \Leftrightarrow K(c_2)$ that is consistent with the canonized examples
-$\{({x'_1}, {y'_1}), \ldots, ({x'_n}, {y'_n})\}$ and then apply the canonizers
-at the outside.  QRE lens synthesis is
-
-% This implies that if there is a derivation $q : c \Leftrightarrow
-% c_1$ of a QRE lens as well as a derivation $\ell : K(c) \Leftrightarrow K(c_1)$
-% that is consistent with the canonized examples $\{({x_1}', {y_1}'),\allowbreak
-% \ldots,\allowbreak ({x_n}', {y_n}')\}$, then we can synthesize a quotient lens $q$ of
-% type $c \Leftrightarrow c_1$ that is consistent with $\{({x_1}', {y_1}'),
-% \ldots, ({x_n}', {y_n}')\}$. More concretely, in order to synthesize a
-% quotient lens $q: c \Leftrightarrow c_1$, that is consistent with the examples
-% $\{(x_1, y_1),\allowbreak \ldots, (x_n, y_n)\}$ it suffices to perform the following
-% procedure:
+synthesize a bijective lens $\ell : K(c_1) \Leftrightarrow K(c_2)$ that is
+consistent with the canonized examples $\{({x'_1}, {y'_1}), \ldots, ({x'_n},
+{y'_n})\}$ and then apply the canonizers at the outside. Our procedure for QRE
+lens synthesis is thus
 \begin{enumerate}
@@
   Optician system.  These benchmarks are a combination of custom benchmarks,
-  benchmarks derived from FlashFill~\cite{}\saz{TODO:cite}, and benchmarks
-  derived from Augeas~\cite{}\saz{TODO:cite}.  In the pre-QRE version of
+  benchmarks derived from FlashFill~\cite{flashfill}, and benchmarks
+  derived from Augeas~\cite{augeas2}. In the pre-QRE version of
   Optician, many of these data formats had to be modified to work with the
@@
 \item The remaining eight benchmarks were derived from
-  data.gov.\footnote{\url{http://data.gov}} Data.gov contains many
-  transformations between formats.  In addition to the formats previously
-  used for Optician, we synthesized transformers between these formats.\saz{This
-    description is not specific enough.  What are the data formats for?  How did
-    we pick which ones to try?}
+  data.gov.\footnote{\url{http://data.gov}} These datasets include census
+  population data, wage comparison data, demographics statistics data and
+  crime index data. We chose these datasets since they were stored in XML, CSV
+  and RDF formats which are more easily specified using regular expressions than
+  plain text or PDF documents.
 \end{enumerate}
@@
   \item[\CanonizerAndSpecSize{}]  the number of
-  AST nodes in $W(q)$ summed with the number of AST nodes in $\canonizer(q)$
+  AST nodes in $W(q_1)$ summed with the number of AST nodes in $\canonizer(q_1)$
   \item[\LensAndSpecSize{}] the number of AST nodes
-  in $W(q)$ summed with the number of AST nodes in the synthesized QRE lens.
+  in $W(q_1)$ summed with the number of AST nodes in the synthesized QRE lens.
 \end{itemize}
@@
 benchmarks within the Optician system, without QREs but still using
-synthesis. This is only an approximation, as both $W(q)$ and $\canonizer(q)$ are
+synthesis. This is only an approximation, as both $W(q_1)$ and $\canonizer(q_1)$ are
 automatically generated from the QRE, and it is possible that a human-written
@@
 in Boomerang, using neither QREs nor synthesis. This value is also an
-approximation, as $W(q)$ and $\canonizer(q)$ and the bijective lens are all
+approximation, as $W(q_1)$ and $\canonizer(q_1)$ and the bijective lens are all
 automatically generated.
@@
 to synthesize all quotient lenses in under 20 seconds, and typically
-finished in under 5 seconds. }
+finished in under 5 seconds. \bcp{The left-hand graph, especially, is hard
+  to make anything out of, because the curves are squished against the
+  axes.  Not sure what to do about this.  Also, I'm finding the terms
+  ``Optician'' and ``Optometrist'' really hard to keep straight---in part
+  because, in English, they are almost synonyms.}}
 \label{fig:times}
@@
 benchmarks at a speed competitive with the old version.  There is a small amount
-of additional overhead introduced by QREs when calculating $W(q)$ and $K(q)$,
+of additional overhead introduced by QREs when calculating $W(q_1)$ and $K(q_1)$,
 resulting in a very slight decrease in performance.
@@
 
+\section{Appendix}
+\bcp{Appendices should be numbered differently (with letters, probably) and
+  belong after the references.}
+%
+We give a full proof of \cref{normal form}: If there is a derivation $q : c_1
+\Leftrightarrow c_2$, then there exists a bijective lens $\ell : K(c_1)
+\Leftrightarrow K(c_2)$ such that:
+\begin{align*}
+\llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket\circ
+\canonizer(c_1)\\
+\llbracket q \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
+\canonizer(c_2)
+\end{align*}
+\begin{proof}
+We proceed by induction over the derivation $q : c_1 \Leftrightarrow c_2$.
+\begin{enumerate}
+  \item
+  $\kw{lift}(\ell): R/\mathit{id}(R) \Leftrightarrow S/\mathit{id}(S)$ where
+  $\ell : R \Leftrightarrow S$. Then
+  \begin{align*}
+\llbracket \kw{lift}(\ell) \rrbracket.\get &=  \llbracket \ell \rrbracket
+= \llbracket \ell \rrbracket \circ id_{\mathcal{L}(R)} =
+\llbracket \ell \rrbracket \circ \canonizer(\mathit{id}(R)), \text{ and }\\
+\llbracket \kw{lift}(\ell) \rrbracket.\lput &= \llbracket \ell
+\rrbracket^{-1} = \llbracket \ell \rrbracket^{-1} \circ id_{\mathcal{L}(S)} =
+\llbracket \ell \rrbracket^{-1} \circ \canonizer(id(S))
+\end{align*}
+\item
+$\kw{lquot}(c_1, q): c_1 \; ; \; c_2 \Leftrightarrow c_3$ where $q : c_2
+\Leftrightarrow c_3$, $c_1$ is well formed and $K(c_1) = W(c_2)$. Then
+\begin{align*}
+\llbracket \kw{lquot}(c_1, q) \rrbracket.\get  &= \llbracket q
+\rrbracket.\get \circ \canonizer(c_1)\\
+\llbracket \kw{lquot}(c_1, q) \rrbracket.\lput &= \llbracket q \rrbracket.\lput
+\end{align*}
+By the induction hypothesis, there exists a bijective lens $\ell :
+K(c_2) \Leftrightarrow K(c_3)$ such that
+\begin{align*}
+\llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket \circ
+\canonizer(c_2)\\
+\llbracket q \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
+\canonizer(c_3)
+\end{align*}
+Consequently
+\begin{align*}
+\llbracket \kw{lquot}(c_1, q)\rrbracket.\get  &= (\llbracket \ell \rrbracket \circ
+\canonizer(c_2)) \circ \canonizer(c_1) = \llbracket \ell \rrbracket \circ
+(\canonizer(c_1 \; ; \; c_2))\\
+\llbracket \kw{lquot}(c_1, q) \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
+\canonizer(c_3)
+\end{align*}
+
+\item
+$\kw{rquot}(q, c_3):c_1 \Leftrightarrow c_2 \; ; \; c_3$ where $q : c_1
+\Leftrightarrow c_2$, $c_3$ is well formed and $K(c_3) = W(c_2)$. Proceed as in
+the previous case.
+\item
+$q_1 \; ; \; q_2: c \Leftrightarrow c_2$ where $q_1 : c \Leftrightarrow c_1$ and
+$q_2 : c_1 \Leftrightarrow c_2$. Then
+\begin{align*}
+\llbracket q_1 \rrbracket.\get &= \llbracket q_2 \rrbracket.\get\circ \llbracket
+q_1 \rrbracket.\get, \text{ and }\\
+\llbracket q_1 \rrbracket.\lput &= \llbracket q_1 \rrbracket.\lput \circ \llbracket
+q_2 \rrbracket.\lput
+\end{align*}
+By the induction hypothesis, there exist bijective lenses
+$\ell_1 :
+K(c) \Leftrightarrow K(c_1)$ and $\ell_2 : K(c_1) \Leftrightarrow K(c_2)$ such
+that
+\begin{align*}
+\llbracket q_1 \rrbracket.\get &= \llbracket \ell_1 \rrbracket \circ
+\canonizer(c)\\
+\llbracket q_1 \rrbracket.\lput &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
+\canonizer(c_1)
+\end{align*}
+and
+\begin{align*}
+\llbracket q_2 \rrbracket.\get &= \llbracket \ell_2 \rrbracket \circ
+\canonizer(c_1)\\
+\llbracket q_2 \rrbracket.\lput &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
+\canonizer(c_2)
+\end{align*}
+Consequently,
+\begin{align*}
+\llbracket q_2 \rrbracket.\get \circ \llbracket q_1 \rrbracket.\get &=
+(\llbracket \ell_2 \rrbracket \circ \canonizer(c_1)) \circ (\llbracket \ell_1
+\rrbracket \circ \canonizer(c))\\
+&= \llbracket \ell_2 \rrbracket \circ (\canonizer(c_1) \circ \llbracket \ell_1
+\rrbracket) \circ \canonizer(c)\\
+&= (\llbracket \ell_2 \rrbracket \circ \llbracket \ell_1 \rrbracket) \circ
+\canonizer(c)\\
+&= \llbracket \ell_1 \; ; \; \ell_2 \rrbracket \circ
+\canonizer(c)
+\end{align*}
+A similar argument shows that
+$$\llbracket q_1 \rrbracket.\lput \circ \llbracket q_2 \rrbracket.\lput =
+\llbracket \ell_1 \; ; \; \ell_2 \rrbracket^{-1} \circ
+\canonizer(c)$$
+\item
+$q^* : {c_1}^* \Leftrightarrow {c_2}^*$ where $q : c_1 \Leftrightarrow c_2$,
+$W(c_1)^{*!}$ and $W(c_2)^{*!}$ and $K(c_1)^{*!}$ and $K(c_2)^{*!}$. Then
+\begin{align*}
+\llbracket q^* \rrbracket.\get &= (\llbracket q \rrbracket.\get)^*, \text{
+and }\\
+\llbracket q^* \rrbracket.\lput &= (\llbracket q \rrbracket.\lput)^*
+\end{align*}
+By the induction hypothesis there exists a bijective lens $\ell : K(c_1)
+\Leftrightarrow K(c_2)$ such that
+that
+\begin{align*}
+\llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket \circ
+\canonizer(c_1)\\
+\llbracket q \rrbracket.\lput &= {\llbracket \ell \rrbracket}^{-1} \circ
+\canonizer(c_2)
+\end{align*}
+Consequentlty
+\begin{align*}
+\llbracket q^* \rrbracket.\get &= (\llbracket \ell \rrbracket \circ
+\canonizer(c_1))^* = \llbracket \ell \rrbracket^* \circ
+\canonizer(c_1)^* = \llbracket \ell^* \rrbracket \circ
+\canonizer(c_1^*)\\
+\llbracket q^* \rrbracket.\lput &= (\llbracket \ell \rrbracket^{-1} \circ
+\canonizer(c_2))^* = (\llbracket \ell \rrbracket^{-1})^* \circ
+\canonizer(c_2)^* = \llbracket \ell^* \rrbracket^{-1} \circ
+\canonizer(c_2^*)\\
+\end{align*}
+\item
+$q_1 \cdot q_2: c_1 \cdot c_2 \Leftrightarrow d_1 \cdot d_2$, where $q_1 : c_1
+\Leftrightarrow d_1 $,  $q_2 : c_2 \Leftrightarrow d_2$, $W(c_1)
+\cdot^! W(c_2)$, $K(c_1) \cdot^! K(c_2)$, $W(d_1) \cdot^! W(d_2)$ and $
+K(d_1) \cdot^! K(d_2)$. Then
+\begin{align*}
+\llbracket q_1 \rrbracket.\get &= \llbracket q_1 \rrbracket.\get \cdot \llbracket
+q_2 \rrbracket.\get, \text{ and }\\
+\llbracket q_1 \rrbracket.\lput &= \llbracket q_1 \rrbracket.\lput \cdot \llbracket
+q_2 \rrbracket.\lput
+\end{align*}
+By the induction hypothesis, there exist bijective lenses $\ell_1 : K(c_1)
+\Leftrightarrow K(d_1)$ and $\ell_2 : K(c_2) \Leftrightarrow K(d_2)$ such that
+\begin{align*}
+\llbracket q_1 \rrbracket.\get &= \llbracket \ell_1 \rrbracket \circ
+\canonizer(c_1)\\
+\llbracket q_1 \rrbracket.\lput &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
+\canonizer(d_1)
+\end{align*}
+and
+\begin{align*}
+\llbracket q_2 \rrbracket.\get &= \llbracket \ell_2 \rrbracket \circ
+\canonizer(c_2)\\
+\llbracket q_2 \rrbracket.\lput &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
+\canonizer(d_2)
+\end{align*}
+Consequently,
+\begin{align*}
+\llbracket q_1 \rrbracket.\get &= (\llbracket \ell_1 \rrbracket \circ
+\canonizer(c_1)) \cdot  (\llbracket \ell_2 \rrbracket \circ
+\canonizer(c_2))\\
+&= (\llbracket \ell_1 \rrbracket \cdot \llbracket \ell_2
+\rrbracket) \circ (\canonizer(c_1) \cdot \canonizer(c_2))\\
+&= \llbracket \ell_1 \cdot  \ell_2 \rrbracket \circ \canonizer(c_1 \cdot c_2)
+\end{align*}
+Similarly
+$$
+\llbracket q_1 \rrbracket.\lput = \llbracket \ell_1 \cdot  \ell_2 \rrbracket^{-1}
+\circ \canonizer(d_1 \cdot d_2) $$
+\item
+$q_1 = q_1 \sep q_2$ where $q_1 : c_1 \Leftrightarrow d_1 $, $q_2 : c_2
+\Leftrightarrow d_2$, $\mathcal{L}(W(c_1)) \cap \mathcal{L}(W(c_2)) =
+\varnothing$ and $\mathcal{L}(W(d_1)) \cap \mathcal{L}(W(d_2)) = \varnothing$.
+Then
+$$
+\llbracket q_1 \sep q_2 \rrbracket.\get(s) =
+\begin{cases}
+\llbracket q_1 \rrbracket.\get (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
+\llbracket q_2 \rrbracket.\get (s) & \text{if } s \in \mathcal{L}(W(c_2))\\
+\end{cases}$$
+$$\llbracket q_1 \sep q_2 \rrbracket.\lput(s) =
+\begin{cases}
+\llbracket q_1 \rrbracket.\lput (s) & \text{if } s \in \mathcal{L}(W(d_1))\\
+\llbracket q_2 \rrbracket.\lput (s) & \text{if } s \in \mathcal{L}(W(d_2))\\
+\end{cases}
+$$
+By the induction hypothesis, there exist bijective lenses $\ell_1 : K(c_1)
+\Leftrightarrow K(d_1)$ and $\ell_2 : K(c_2) \Leftrightarrow K(d_2)$ such that
+\begin{align*}
+\llbracket q_1 \rrbracket.\get &= \llbracket \ell_1 \rrbracket \circ
+\canonizer(c_1)\\
+\llbracket q_1 \rrbracket.\lput &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
+\canonizer(d_1)
+\end{align*}
+and
+\begin{align*}
+\llbracket q_2 \rrbracket.\get &= \llbracket \ell_2 \rrbracket \circ
+\canonizer(c_2)\\
+\llbracket q_2 \rrbracket.\lput &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
+\canonizer(d_2)
+\end{align*}
+Consequently,
+$$
+\llbracket q_1 \sep q_2 \rrbracket.\get(s) =
+\begin{cases}
+\llbracket \ell_1 \rrbracket \circ
+\canonizer(c_1) (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
+\llbracket \ell_2 \rrbracket \circ
+\canonizer(c_2) (s) & \text{if } s \in \mathcal{L}(W(c_2)),\\
+\end{cases}$$
+so $\llbracket q_1 \sep q_2 \rrbracket.\get = \llbracket \ell_1 \sep
+\ell_2 \rrbracket \circ \canonizer(c_1 \sep c_2)$. A similar argument shows
+that $\llbracket q_1 \sep q_2 \rrbracket.\lput = \llbracket \ell_1 \sep
+\ell_2 \rrbracket^{-1} \circ \canonizer(d_1 \sep d_2)$.\\
+This completes the proof.
+\end{enumerate}
+\end{proof}
+
 \bibliographystyle{plain}

### 2018-03-09 papers/QuotientLenses/main.tex
Checkpoint polishing and comments

@@
 Optician}, the programmer need only spend the effort of specifying the data
-formats and the behaviour of the lens on a few inputs in order to define a
+formats and the behaviour\bcp{Let's use American spellings: behavior} of the
+lens on a few inputs in order to define a 
 bijective lens. As a result defining bijective lenses becomes faster and less
@@
   $\eqrel{c}$, and
-  \item a ``canonizing'' function $\canonizer(c):\mathcal{L}(W(c))
+  \item a ``canonizing'' function\bcp{Why is it called ``canonizer'' rather
+    than ``canonize''??} $\canonizer(c):\mathcal{L}(W(c))
   \longrightarrow \mathcal{L}(K(c))$ which given any $w \in \mathcal{L}(W(c))$,
@@
 \end{align*}
-\noindent which as we have explained is defined so that the strings that match
+\noindent which, as we have explained, is defined so that the strings that match
 $c_i, c$ are placed in the canonical permutation which is the identity
@@
 
-Finally, Figure ~\ref{fig:relations} gives the induction defintion of the
-equivalence relation $\eqrel{c}$. These are the set-theoretic semantics of QREs
+Finally, Figure ~\ref{fig:relations} gives the inductive definition of the
+equivalence relation $\eqrel{c}$. This is the set-theoretic semantics of QREs
 as an equivalence relation on a regular language.
@@
 \[
-\begin{array}{l@{\quad}l@{\quad}l}
+\begin{array}{l@{}l@{\ \ }l}
 w \; \equiv_R \; w' &\iff& w = w' \\
-w \; \equiv_{\project{R}{s}} \; w' \text{ for all }w, w'\\
+w \; \equiv_{\project{R}{s}} \; w' && \text{for all }w, w'\\
 w \; \equiv_{\squash{c_1}{c_2}{f}} \; w' &\iff& f(w) \equiv_{c_2} w'
@@
 \cdot {r'}_{\theta(n)} \\
-& & \text{ for some } \sigma, \theta \in S_n, \text{ with } r_i \;
+& & \text{for some } \sigma, \theta \in S_n, \text{ with } r_i \;
 \equiv_{q_i} \; r'_i \text{ and } s_k \; \equiv_{c} \; s'_{k}
@@
 
-More formally, we say that regular expressions $R$ and $S$ are
+Formally, we say that regular expressions $R$ and $S$ are
 \textit{unambiguosly concatenable}, written $R \cdot^! S$ if for all strings
@@
 
-Figure~\ref{fig:qrerules} gives the inferrence rules for deriving well-formed
+Figure~\ref{fig:qrerules} gives the inference rules for deriving well-formed
 QREs.
@@
 As we already mentioned, the unambiguity conditions come into play when defining
-QREs using the regular combinators. For example, the (Concat) inferrence rule
+QREs using the regular combinators. For example, the (Concat) inference rule
 for concatenation says that the concatenation $c_1 \cdot c_2$ of QREs $c_1$ and
@@
 require that these regular expressions be disjoint.
-
+%
 The third hypothesis says that the regular expression $K(c_1) \cdot K(c)
@@
   one).}\sam{I made the notation consistent by just using $c_1$ and $c_2$. I
-  personally feel that it is better not to use P and Q for QREs since we use p
-  and q for QRE lenses, but if you disagree I will make the change.}\bcp{I
-  agree that this is a problem.  But maybe the right fix is to choose
-  different metavariables for QRE lenses.  Or to use capital letters other
-  than $P$ and $Q$ for QREs.  I don't like $c_1$ and $c_2$, both for the
-  lower-case letter and for the subscripts---it's not at all consistent with
-our conventions elsewhere.}
+  personally feel that it is better not to use P and Q for QREs since we use
+  p and q for QRE lenses, but if you disagree I will make the change.}\bcp{I
+  agree that this is suboptimal, but I don't think it's worse than what we
+  have now.  Or maybe the right fix is to choose different metavariables for
+  QRE lenses.  Or to use capital letters other than $P$ and $Q$ for QREs.
+  In any case, I don't like $c_1$ and $c_2$, both for the lower-case letter
+  and for the subscripts---it's not at all consistent with our conventions
+  elsewhere.}
 
 Given regular expressions $R, S$ and equivalence relations
-$\sim_R$ and $\sim_S$ defined on $\mathcal{L}(R)$ and $\mathcal{L}(S)$
-respectively, we\bcp{can we clarify that this is exactly Foster et al's
+$\sim_R$ and $\sim_S$ defined on $\mathcal{L}(R)$ and $\mathcal{L}(S)$,
+we\bcp{can we clarify that this is exactly Foster et al's 
 definition?}\sam{this is not Foster et al's definition. It's just a
@@
 \longrightarrow \mathcal{L}(R)$ such that for all $r \in \mathcal{L}(R)$ and $s
-\in \mathcal{L}(S)$, then $q.\lput(q_1.\get(r)) \sim_R r$ and
+\in \mathcal{L}(S)$, we have $q.\lput(q_1.\get(r)) \sim_R r$ and
 $q.\get(q_1.\lput(s)) \sim_R s$. Moreover, if $r \sim_R r'$, then $q.\get(r) =
@@
 
-These laws are similar to the laws for bijective lenses, except that the
-equality restriction in the bijective lens laws are loosened to allow for
+These laws are similar to the laws for bijective lenses\bcp{we should state
+  them explicitly somewhere in this paper and refer to them here by number},
+except that the 
+equality restrictions in the bijective lens laws are loosened to allow for
 equivalence relations. Also, the condition that $r \sim_R r'$ implies
@@
 $ and $q.\lput = \ell_1^{-1} \circ \canonizer(c_2)$. In other words, the get
-function of the $q$ first canonizes the source data using $\canonizer(c_1)$
-and then applies $f$ to the result, while the put function of the $q$ first
+function of $q$ first canonizes the source data using $\canonizer(c_1)$
+and then applies $f$ to the result, while the put function of $q$ first
 canonizes the target data using $\canonizer(c_2)$ and the applies $f^{-1}$ to
@@
 
-One potential pitfall with QRE lenses is that the combinators that
-we define on them may not preserve this normal form. For example, the
+One potential pitfall\bcp{Why is this a pitfall?  I.e., why do we care about
+this norman form?  (The answer is synthesis, of course, but we haven't said
+so here.)} with QRE lenses is that the combinators that
+we define on them may not preserve this normal form\bcp{In what sense is
+  this a ``normal form''?  What you've given above is just a definition:
+  quotient lenses can be defined in such and such way.  You haven't said we
+  want to consider it a normal form, or specified what process it is the
+  normal form {\em of}.}. For example, the
 composition $q_1 \; ; \; q_2$ of QRE lenses $q_1$ and $q_2$ may not be
@@
 bijective lens with canonizers a the edges. Consequently, in order to guarantee
-that every QRE lens $q_1$ is equivalent to a QRE lens $q_2$ which has this
-canonical form, we need to identify a set $S$ of bijections that will ensure
-that all QRE lenses is functionally equivalent to a QRE lens in canonical form.
+that every QRE lens $q_1$ is equivalent to a QRE lens $q_2$ in this
+canonical form\bcp{This paragraph uses variables $q_1$ and $q_2$ twice, with
+two different meanings.}, we need to identify a set $S$ of bijections that
+will ensure 
+that every QRE lens is functionally equivalent to a QRE lens in canonical
+form.\bcp{This is not very clear, in part because it is not very precise:
+  It is assuming, implicitly, that the class of QRE lenses is 
+  parameterized on a particular set of bijections on kernels, and then
+  asking what sets of bijections can be used.  This question makes sense,
+  but the reader may not realize that this is what's being asked. } 
 Our choice is to use the class of \textit{bijective lenses}, which we studied in
@@
 \subsection{Bijective Lenses}
+\bcp{There's a persistent opportunity for confusion between the semantic
+  space of {\em all} bijective lenses (perhaps between arbitrary structures,
+  not just strings) and the rather specific bijections that are picked out
+  by our combinators.  One way to help, I think, would be to use the phrase
+  ``bijective string lenses'' instead of just ``bijective lenses.''}
 We define the set of \textit{bijective lenses} to be the set of bijections
-between regular languages created using Boomerang lens combinators.
+between regular languages created using Boomerang lens combinators.\bcp{Why
+  do we choose Boomerang's set of combinators?  Does it have some formal
+  property that we find attractive?  Or did it just happen to be lying around?}
 The syntax for the language of bijective lenses is given by
@@
 data in the forward direction, and $t$ with $s$ in the target data in the backward
-direction. $\mathit{id}(R)$ applies the identity function to the source and target
+direction. \bcp{Don't begin a sentence with a lower-case letter, even if it
+  is a math symbol.  Find a way to rewrite so that it doesn't need to be at
+  the beginning.}$\mathit{id}(R)$ applies the identity function to the
+source and target 
 in $\mathcal{L}(R)$ in both directions. The composition lens $\ell_1 \; ; \; \ell_2$
@@
 direction. In the backward direction, $\mathit{\swap}(\ell_1, \ell_2)$ first undoes
-the \swap, then proceeds as expected. The $\ell_1 \; | \; \ell_2$ lens
+the \swap, then behaves like $cdot$. The $\ell_1 \; | \; \ell_2$ lens
 chooses to apply $\ell_1$ or $\ell_2$ depending on whether the source
@@
 each of the $t_i$'s for a final result of $t_1 \cdot \ldots \cdot t_n$ in the
-forward direction. $\ell^*$ operates similarly in the backward direction, but
+forward direction. \bcp{ditto}$\ell^*$ operates similarly in the backward direction, but
 with $s, s_i$ substituted for $t, t_i$.
@@
 \Leftrightarrow S'$ provided that $R$ (resp. $S$) can be proven to be
-equivalent to $R'$ (resp. $S$) from the star-semiring axioms:
+equivalent to $R'$ (resp. $S$) from the star-semiring axioms\bcp{people may
+  not know what these are}:
 
@@
 q_1 \; ; \; q_2,$$ where $\ell$ ranges over bijective lenses.
+\BCP{Is this really what we want?  E.g., isn't it reasonable for the 
+  programmer to write $\mathit{\swap}(q_1, q_2)$?}
 
@@
 combinators are strongly unambiguous. The composition operator `$;$' also
-exhibits the expected behaviout, though the conditions that must be checked in
+exhibits the expected behaviour, though the conditions that must be checked in
 order for the composition to be well defined are perhaps unintuitive. The
-reasons why these conditions are necesary will become more apparent when
+reasons why these conditions are necessary will become more apparent when
 proving normal forms for QRE lenses.
@@
 \bcp{Agreed. But don't we have any ideas for how to extend or generalize or
-things to do next??  E.g., is it completely trivial to extend all this to
-asymmetric and symmetric lenses?  (If so, we should say so!)  Do we have
-more ideas for QRE primitives?  Etc.}
+things to do next or additional questions to consider??  E.g.: 
+\begin{itemize}
+\item is it easy to extend all this to
+asymmetric and symmetric lenses?  (If so, we should say so!)
+\item What properties do the bijective lens combinators have to have in
+order to validate the normalization property?  (I.e., if we tried to extend
+Boomerang's bijective lens combinators with another one, what would happen?)
+\item Do we have more ideas for QRE primitives?
+\item Etc.
+\end{itemize}
+}
 \label{concl}

### 2018-03-09 papers/QuotientLenses/main.tex
Checkpoint polishing and new try at dfiagram

@@
 framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=false,
+keepspaces=true,
 commentstyle=\itshape\rmfamily,
@@
 regular combinators of concatenation `$\cdot$', Kleene star `$*$' and
-`$|$' on QRE lenses exhibit the expected behaviour, with the caveat that these
+`$|$' on QRE lenses exhibit the expected behaviour\bcp{Is the behavior really
+  ``expected''---i.e., will it be obvious to all readers?}, with the caveat
+that these 
 combinators are strongly unambiguous. The composition operator `$;$' also
@@
 For example, recall that in the \bibtex{} to EndNote transformation, we had the
-QREs \cd{bibtex} and \cd{endnote} describing the \bibtex{} and EndNote records
-respectively, and the bijective lens \cd{bib_to_end} that maps between
+QREs \cd{bibtex} and \cd{endnote} and the bijective lens \cd{bib_to_end}
+that maps between 
 \cd{bibtex} and \cd{endnote}. The quotient lens that maps between
-\cd{bibtex} and \cd{endnote} is given by the QRE lens
+\cd{bibtex} and \cd{endnote} is then given by the following QRE lens:
 
 \begin{lstlisting}
-let bib_to_end_q : (lens in bibtex $\Leftrightarrow$ endnote)  = rquot (lquot(bibtex, bib_to_end), endnote)
+let bib_to_end_q : (lens in bibtex $\Leftrightarrow$ endnote) = 
+  rquot (lquot(bibtex, bib_to_end), endnote)
 \end{lstlisting}

### 2018-03-09 papers/QuotientLenses/main.tex
Checkpoint through section 4

@@
 syntactially equal to $id(R)$ and $id(R')$ for some $R$ and $R'$ satisfying
-$\mathcal{L}(R) = \mathcal{L}(R')$.
+$\mathcal{L}(R) = \mathcal{L}(R')$.  \BCP{Here, once again, I worry that we
+  are mixing fundamental theory with pragmatic language design /
+  implementation, to the detriment of both.  At a theoretical level, this
+  restriction is nonsensical, and making it here raises questions like
+  ``Does the proof that composition can be normalized away depend on this
+  restriction?''}
 

### 2018-03-09 papers/QuotientLenses/main.tex
Further polishing

@@
 \begin{proof}
-We elide a full proof of this statement, since the proof is attained via a
-straightforward induction on the derivation $q_1 : c_1 \Leftrightarrow c_2$ (the
-full details can be found in the appendix). Since bijective lenses are closed
-under strongly unambiguous regular operators, the tricky case is composition.
-The composition case is more involved since we need to show that we can get rid
-of the canonizers in the middle:
+By straightforward induction on the derivation
+$q_1 : c_1 \Leftrightarrow c_2$. Since bijective lenses are closed under strongly unambiguous
+regular operators, the tricky case is composition.  The composition case is
+more involved since we need to show that we can get rid of the canonizers in
+the middle:
 
@@
 \end{center}
-
+%
 Consequently,
@@
 
-The other cases of the proof are similar in that they follow from a
+The other cases of the proof are similar, proceeding by a
 straightforward application of the induction hypothesis followed by unrolling
-the equations that give the denotation for QRE lenses.
+the equations that give the denotation for QRE lenses.  Full details can be
+found in the appendix.
 \end{proof}

### 2018-03-09 papers/QuotientLenses/main.tex
Checkpoint

@@
 way of constructing lenses can often be much simpler than building them by
-hand---see the example of lens synthesis in
-Section~\ref{sec:examplesynth}.
+hand, as we saw in Section~\ref{sec:examplesynth}.
 
@@
 languages.  This simple strategy turns out to be remarkably effective, and the
-idea of using Optician in this way is part the impetus for the design of our QRE
-lenses.
+idea of using Optician in this way is part of the impetus\bcp{What was the
+  other part??} for the design of our QRE lenses.
 
@@
 $s_i$ to $r_i$ in the backward direction.  The Optician algorithm is guaranteed
-to succeed if such a lens exists.    \saz{We should probably mention what
-happens if the examples aren't sufficient to uniquely determine the lens.}
+to succeed (eventually!) if such a lens exists.    \saz{We should probably
+  mention what happens if the examples aren't sufficient to uniquely
+  determine the lens.}
 
@@
 \end{align*}
-
-For the QRE examples, the $x_i$'s are in $W(c_1)$ and the $y_i$'s are
+\bcp{This math display and the previous one are punctuated perfedtly: In the
+first, the colon replaces a period at the end of the sentence (so no need
+for another period at the end of the display).  In the second, the display
+is punctuated as if it were inline.  Please go through and make sure all the
+other displayed material follows one or the other of these rules.}
+
+For the\bcp{why ``the''?  Are we talking about some set of QRE examples
+  already?} QRE examples, the $x_i$'s are in $W(c_1)$ and the $y_i$'s are 
 in $W(c_2)$, so we can construct ${x'_i} = \canonizer(c_1)(x_i)$ in $K(c_1)$
@@
   Return $q = \kw{rquot}(\kw{lquot}(c_1, \ell), c_2) : c_1 \Leftrightarrow
-  c_2$
+  c_2$ \bcp{punctuate this enumeration as if it were running text: ``, and''
+  at the end of (3) and period here.}
 \end{enumerate}
@@
 
-\begin{minipage}{3in}
+\begin{center}
+\begin{minipage}{2.5in}
 \begin{lstlisting}
@@
 \end{minipage}
-\begin{minipage}{3in}
+\begin{minipage}{1.5in}
 \begin{lstlisting}
@@
 \end{minipage}
+\end{center}
 
@@
 \begin{lstlisting}
-let bib_to_end : (lens in bibtex <=> endnote)
-= synth bibtex <=> endnote  using  {(bib_example, end_example)}
+let bib_to_end : (lens in bibtex <=> endnote) = 
+  synth bibtex <=> endnote
+  using {(bib_example, end_example)}
 \end{lstlisting}

### 2018-03-09 papers/QuotientLenses/main.tex
Fixed spacing weirdness in listings

@@
 emph={synth,project,perm,squash,normalize,using,ins,del,lens},
-emphstyle=\upshape\sffamily\color{dkblue}, columns=flexible,
+emphstyle=\upshape\sffamily\color{dkblue}, columns=fullflexible,
 % BCP: I find this distracting:
@@
 \end{lstlisting}
-\bcp{Again, the spacing is weird}
 \noindent Here, the generated  quotient lens synchronizes \cd{bibtex} and

### 2018-03-09 papers/QuotientLenses/main.tex
Section 5 polishing / comments

@@
 \newcommand{\saz}[1]{\FINISH{orange}{SZ}{#1}}
+\newcommand{\SAZ}[1]{\FINISH{orange}{SZ}{\bf #1}}
 \newcommand{\ksf}[1]{\FINISH{teal}{K}{#1}}
@@
   c_2$ \bcp{punctuate this enumeration as if it were running text: ``, and''
-  at the end of (3) and period here.}
+  at the end of (3) and period here.  And lowercase letters at the
+  beginnings of the items.}
 \end{enumerate}
 
-\saz{Do we want to state formally the completeness of this algorithm?}
+\saz{Do we want to state formally the completeness of this algorithm?}\BCP{+1}
 
@@
 In this case, one example is sufficient to uniquely determine the desired
-transformation.\saz{Is that true?}
-
-\saz{We might want to say something more precise about the running time or
+transformation.\SAZ{Is that true?}\BCP{Seems more likely that the desired
+  transformation happened to be the first one that the synthesis algorithm
+  reached.  Indeed, my intuition is that {\em every} synthesis problem with
+  a finite set of examples has infinitely many solutions.  Is that right?}
+
+\SAZ{We might want to say something more precise about the running time or
   expected performance of this algorithm, perhaps to set up the evaluation
   section.  Something along the lines of: ``Although the algorithm is complete
-  in theory, in the worst case, it may take an unbounded [???IS THIS CORRECT?]
+  in theory, in the worst case, it may take an unbounded [???IS THIS
+  CORRECT?]\bcp{Surely it's bounded by some exponential formula involving the
+    size of the inputs...?} 
   amount time to search for a satisfactory lens.  Nevertheless, this algorithm

### 2018-03-09 papers/QuotientLenses/main.tex
Checkpoint

@@
 of that implements a general class of quotient lenses is {\em
-Boomerang}~\cite{quotientlenses}. While Boomerang's quotient lenses enable a
-programmer to define a very general class of quotient lenses, these lenses to
-have two main drawbacks.
+Boomerang}~\cite{quotientlenses}\bcp{Is this the right citation for
+Boomerang?  Or should it be \cite{boomerang}?}. While Boomerang's quotient
+lenses enable a 
+programmer to define a very general class of quotient lenses\bcp{read
+  literally, that part of the sentence is nonsense: boomerang's lenses
+  (themselves) do not enable the programmer to define lenses.  Of course,
+  the reader can probably figure out what is meant, but that's a needless
+  tax on their finite mental energy.}, these lenses to have two main
+drawbacks.
 
@@
 We have implemented QREs and the quotient lens synthesis algorithm described
-above as part of Boomerang. Thanks to this integration, QREs and quotient lens
+above as an extension to the Boomerang interpreter~\cite{Boomerang}. Thanks to this integration, QREs and quotient lens
 synthesis tasks can now be expressed alongside existing Boomerang terms, and we

### 2018-03-09 papers/QuotientLenses/main.tex
Checkpoint

@@
 We have implemented QREs and the quotient lens synthesis algorithm described
-above as an extension to the Boomerang interpreter~\cite{Boomerang}. Thanks to this integration, QREs and quotient lens
-synthesis tasks can now be expressed alongside existing Boomerang terms, and we
+above as an extension to the Boomerang interpreter~\cite{boomerang,quotientlenses}. Thanks to this integration, QREs and quotient lens
+synthesis tasks can now be expressed alongside existing Boomerang
+terms\bcp{what's the difference between a task and a term (and how do each
+  of them relate to ``lenses'' just below)?}, and we
 can combine the two kinds of lenses seamlessly in the same application.
@@
   \item Does synthesizing quotient lenses from QREs permit an easier development
-  process than existing approaches?
+  process than existing approaches?  \bcp{Needlessly vague: What existing approaches?  Writing
+  them by hand?}
 
   \item Does the increase in expressivity gained from synthesizing from QREs
-  permit synthesis of more useful lenses than existing approaches.
+  permit synthesis of more useful lenses than existing approaches.  \bcp{Is
+    this really an interesting empirical question?  Seems obvious, from the
+    arguments we made in Section 2 and the intro.}
 
   \item Does our tool synthesize lenses fast enough to be used as part of a
-  standard development process.
+  standard development process.  \bcp{Should there be some question marks
+    around here.}
 \end{enumerate}
@@
 
-\item The remaining eight benchmarks were derived from
-  data.gov.\footnote{\url{http://data.gov}} These datasets include census
+\item The remaining eight benchmarks were derived from datasets found at 
+  data.gov. These datasets include census
   population data, wage comparison data, demographics statistics data and
@@
   and RDF formats which are more easily specified using regular expressions than
-  plain text or PDF documents.
+  plain text or PDF documents.  \bcp{Why did we include these eight new
+    ones?  Are they significantly different (more complex, exercising
+    different features of the synthesis algorithm or QRE syntax, ...)?}
 \end{enumerate}
@@
 benchmarks within the Optician system, without QREs but still using
-synthesis. This is only an approximation, as both $W(q_1)$ and $\canonizer(q_1)$ are
+synthesis. (This is only an approximation, as both $W(q_1)$ and $\canonizer(q_1)$ are
 automatically generated from the QRE, and it is possible that a human-written
-version might be smaller.
+version might be smaller.)
 
 The value of \LensAndSpecSize{} estimates the burden in writing our benchmarks
-in Boomerang, using neither QREs nor synthesis. This value is also an
+in Boomerang, using neither QREs nor synthesis. (This is also an
 approximation, as $W(q_1)$ and $\canonizer(q_1)$ and the bijective lens are all
-automatically generated.
+automatically generated.)
 

### 2018-03-09 papers/QuotientLenses/main.tex
Checkpoint comment

@@
 to be about half as large as \CanonizerAndSpecSize, and about a third as large
-as \LensAndSpecSize{}. This demonstrates that introducing QREs saves programmers
+as \LensAndSpecSize{}.\BCP{These are really the critical conclusions, and
+  they are hard to see, the way the graph is drawn.  I wonder whether we
+  could instead present with the individual examples on the X axis (sorted
+  by number of AST nodes), with a bar above each one containing three
+  regions, colored differently, for LS, CS-LS, and QC-CS...} This suggests
+that introducing QREs saves programmers 
 significant effort compared to both Optician and basic Boomerang.

### 2018-03-09 papers/QuotientLenses/main.tex
Checkpoint comment

@@
   and RDF formats which are more easily specified using regular expressions than
-  plain text or PDF documents.  \bcp{Why did we include these eight new
+  plain text or PDF documents.  \BCP{Besides differing formats, why did we
+    include these eight new 
     ones?  Are they significantly different (more complex, exercising
@@
 transformations.  Synthesizing from QREs loosens that restriction, making it
-possible to synthesize data which is bijective when put in canonical form. To
+possible to synthesize data which is bijective when put in canonical
+form\bcp{No: We do not synthesize data, nor is the data itself bijective.}. To
 determine the increase of expressivity, we analyzed:
@@
   \item How many data formats in Optician's benchmark suite required modifications
-  to be bijective when put in canonical form.
+  to be bijective when put in canonical form.  \bcp{I don't understand the
+    difference between these.}
 \end{enumerate}
@@
 Furthermore, we found that all eight of the data.gov examples required no
-modifications to be in bijective correspondence when in canonical form. While we
-did not put the data.gov examples into a form suitable for pre-QRE Optician, we
+modifications to be in bijective correspondence when in canonical
+form.\BCP{Is this surprising?  We could have chosen the examples so that
+  this is true.} While we
+did not put the data.gov examples into a form suitable for pre-QRE
+Optician\bcp{What does that mean?}, we
 found that only four of the eight data.gov examples did not require
@@
   to make anything out of, because the curves are squished against the
-  axes.  Not sure what to do about this.  Also, I'm finding the terms
+  axes and have almost identical shapes.  Not sure what to do about this.
+  Also, I'm finding the terms 
   ``Optician'' and ``Optometrist'' really hard to keep straight---in part
@@
   \item[\SystemOnOptician{}] Optician with QREs running on the pre-QRE benchmarks.
-  \item[\SystemOnBenchmarks{}] Optician with QREs running on its own benchmarks.
+  \item[\SystemOnBenchmarks{}] Optician with QREs running on its own
+  benchmarks. \BCP{What is the difference between QO and QQ??  The eight
+    additional benchmarks?  Why not measure just these?  Seems like this
+    graphj repeats some information from the one to the left, obscuring the
+    new information.}
 \end{itemize}
@@
   explain more detail here.  I guess this is about the last benchmark in Fig. 11
-  (b), which takes a long time?}  We added in additional fields to make this
+  (b), which takes a long time?\bcp{It's a bit hidden in the figure...}}  We
+added in additional fields to make this 
 format suitable for Optician, which took over 15 seconds synthesize the modified

### 2018-03-09 papers/QuotientLenses/main.tex
Comments on related work

@@
 introduced the theory of quotient lenses and implemented quotient lenses as a
-refinement of the bidirectional string processing language Boomerang.
+refinement of the bidirectional string processing language
+Boomerang~\cite{boomerang}.  
 In Boomerang, the source and target types are specified using regular
@@
 are functions that map elements of a regular language to their canonical
-representative. Boomerang canonizers can express a very broad class of
+representative\bcp{If their use of the word is exactly what we've been using
+throughout, then why define it?  Or, if we mean their exact set of
+expressible canonizers, rather than the core concept of canonizer, can we be
+precise about that distinction?}. Boomerang canonizers can express a very
+broad class of 
 equivalence relations between regular languages, but actually doing so is often
@@
 in-build permutation combinator does not account for separators in between the
-regular expressions. Also, this combinator permutes regular expressions rather
+regular expressions\bcp{This sounds like a kind of trivial distinction.}. Also, this combinator permutes regular expressions rather
 than canonizers, thus making it difficult to express nested permutations that
@@
 
-In their paper, Foster et al also discuss other bidirectional programming
+Foster et al.\relax also discuss other bidirectional programming
 languages that support quotienting of data including XSugar~\cite{xsugar},
-biXid~\cite{bixid} and X/Inv~\cite{Hu2004,Mu2004,Mu2006}. XSugar programs
+biXid~\cite{bixid} and X/Inv~\cite{Hu2004,Mu2004,Mu2006}.\bcp{The fact that
+  Fosater et al discuss these things is not germane to the present
+  discussion.  We need to make our own comparisons (unless they are
+  exactly the same, in which case we can just refer to Foster et al.).}
+XSugar programs 
 bidirectionally convert data stored in XML and ASCII formats respectively, with the
@@
 X and Inv which support a primitive duplication combinator that does not work
-well with the lens laws, but can be expressed using Boomerang quotient lenses.
+well with the lens laws, but can be expressed using Boomerang quotient
+lenses.\bcp{Same comment.}
+
+\BCP{Are there really no bidirectional languages since 2008 that support any
+form of quotienting??}
 
@@
 the system Optician, which we introduced in ~\cite{optician}, is the first attempt
-to synthesize bidirectional transformations that we are aware of. In that
-publication, we compared Optician to two of these unidirectional string
+to synthesize bidirectional transformations. In that
+publication, we\bcp{confusing: Not the same ``we''!} compared Optician to
+two of these unidirectional string 
 transformers, Flash Fill~\cite{gulwani-popl-2014} and
@@
 
-While Optician improved on prior effors at synthesizing unidirectional string
+\bcp{This paragraph seems redundant.}While Optician improved on prior effors
+at synthesizing unidirectional string 
 transformers and introduced synthesis for bidirectional string transformations,
@@
 that the programmer supplies both examples {\em and} format descriptions in the
-form of regular expressions or QREs.  There is a trade-off here.  On the one
+form of regular expressions or QREs.\bcp{But we are far from the only ones
+  to consider type-based synthesis!!}  There is a trade-off here.  On the one
 hand, a user must have some programming expertise to write regular expression
-(or QRE) specifications and it requires some work. On the other hand, such
+(or QRE) specifications, and it requires some work. On the other hand, such
 specifications provide a great deal of information to the synthesis system,

### 2018-03-09 papers/QuotientLenses/main.tex
Final few comments for now
add some changes to eval

--- 2018-03-09 papers/QuotientLenses/main.tex
@@
 \newcommand{\saz}[1]{\FINISH{orange}{SZ}{#1}}
-\newcommand{\SAZ}[1]{\FINISH{orange}{SZ}{\bf #1}}
 \newcommand{\ksf}[1]{\FINISH{teal}{K}{#1}}
@@
 of that implements a general class of quotient lenses is {\em
-Boomerang}~\cite{quotientlenses}\bcp{Is this the right citation for
-Boomerang?  Or should it be \cite{boomerang}?}. While Boomerang's quotient
-lenses enable a 
-programmer to define a very general class of quotient lenses\bcp{read
-  literally, that part of the sentence is nonsense: boomerang's lenses
-  (themselves) do not enable the programmer to define lenses.  Of course,
-  the reader can probably figure out what is meant, but that's a needless
-  tax on their finite mental energy.}, these lenses to have two main
-drawbacks.
+Boomerang}~\cite{quotientlenses}. While Boomerang's quotient lenses enable a
+programmer to define a very general class of quotient lenses, these lenses to
+have two main drawbacks.
 
@@
   c_2$ \bcp{punctuate this enumeration as if it were running text: ``, and''
-  at the end of (3) and period here.  And lowercase letters at the
-  beginnings of the items.}
+  at the end of (3) and period here.}
 \end{enumerate}
 
-\saz{Do we want to state formally the completeness of this algorithm?}\BCP{+1}
+\saz{Do we want to state formally the completeness of this algorithm?}
 
@@
 In this case, one example is sufficient to uniquely determine the desired
-transformation.\SAZ{Is that true?}\BCP{Seems more likely that the desired
-  transformation happened to be the first one that the synthesis algorithm
-  reached.  Indeed, my intuition is that {\em every} synthesis problem with
-  a finite set of examples has infinitely many solutions.  Is that right?}
-
-\SAZ{We might want to say something more precise about the running time or
+transformation.\saz{Is that true?}
+
+\saz{We might want to say something more precise about the running time or
   expected performance of this algorithm, perhaps to set up the evaluation
   section.  Something along the lines of: ``Although the algorithm is complete
-  in theory, in the worst case, it may take an unbounded [???IS THIS
-  CORRECT?]\bcp{Surely it's bounded by some exponential formula involving the
-    size of the inputs...?} 
+  in theory, in the worst case, it may take an unbounded [???IS THIS CORRECT?]
   amount time to search for a satisfactory lens.  Nevertheless, this algorithm
@@
 We have implemented QREs and the quotient lens synthesis algorithm described
-above as an extension to the Boomerang interpreter~\cite{boomerang,quotientlenses}. Thanks to this integration, QREs and quotient lens
-synthesis tasks can now be expressed alongside existing Boomerang
-terms\bcp{what's the difference between a task and a term (and how do each
-  of them relate to ``lenses'' just below)?}, and we
+above as part of Boomerang. Thanks to this integration, QREs and quotient lens
+synthesis tasks can now be expressed alongside existing Boomerang terms, and we
 can combine the two kinds of lenses seamlessly in the same application.
@@
   \item Does synthesizing quotient lenses from QREs permit an easier development
-  process than existing approaches?  \bcp{Needlessly vague: What existing approaches?  Writing
-  them by hand?}
+  process than existing approaches?
 
   \item Does the increase in expressivity gained from synthesizing from QREs
-  permit synthesis of more useful lenses than existing approaches.  \bcp{Is
-    this really an interesting empirical question?  Seems obvious, from the
-    arguments we made in Section 2 and the intro.}
+  permit synthesis of more useful lenses than existing approaches.
 
   \item Does our tool synthesize lenses fast enough to be used as part of a
-  standard development process.  \bcp{Should there be some question marks
-    around here.}
+  standard development process.
 \end{enumerate}
@@
 
-\item The remaining eight benchmarks were derived from datasets found at 
-  data.gov. These datasets include census
+\item The remaining eight benchmarks were derived from
+  data.gov.\footnote{\url{http://data.gov}} These datasets include census
   population data, wage comparison data, demographics statistics data and
@@
   and RDF formats which are more easily specified using regular expressions than
-  plain text or PDF documents.  \bcp{Why did we include these eight new
-    ones?  Are they significantly different (more complex, exercising
-    different features of the synthesis algorithm or QRE syntax, ...)?}
+  plain text or PDF documents.
 \end{enumerate}
@@
 that it takes far fewer AST nodes to define benchmark lenses using QRE
-synthesis than with Optician without QRE synthesis or without synthesis.\saz{Why
-is the bar for ``All QRE Benchmarks'' at 14? Shouldn't it be at 47 = 39+8? Are
-we not testing all QRE benchmarks?}}
+synthesis than with Optician without QRE synthesis or without synthesis.}
 \label{fig:asts}
@@
 nodes in the abstract syntax tree as a measure of code size, and calculated
-three quantities for each lens synthesis benchmark task.  \saz{Say something
-  about why there are only 14 benchmarks?}
+three quantities for each lens synthesis benchmark task. We only compared the
+sizes on benchmarks where the canonization functions are nontrivial.
 %
@@
 benchmarks within the Optician system, without QREs but still using
-synthesis. (This is only an approximation, as both $W(q_1)$ and $\canonizer(q_1)$ are
+synthesis. This is only an approximation, as both $W(q_1)$ and $\canonizer(q_1)$ are
 automatically generated from the QRE, and it is possible that a human-written
-version might be smaller.)
+version might be smaller.
 
 The value of \LensAndSpecSize{} estimates the burden in writing our benchmarks
-in Boomerang, using neither QREs nor synthesis. (This is also an
+in Boomerang, using neither QREs nor synthesis. This value is also an
 approximation, as $W(q_1)$ and $\canonizer(q_1)$ and the bijective lens are all
-automatically generated.)
+automatically generated.
 
@@
 to be about half as large as \CanonizerAndSpecSize, and about a third as large
-as \LensAndSpecSize{}.\BCP{These are really the critical conclusions, and
-  they are hard to see, the way the graph is drawn.  I wonder whether we
-  could instead present with the individual examples on the X axis (sorted
-  by number of AST nodes), with a bar above each one containing three
-  regions, colored differently, for LS, CS-LS, and QC-CS...} This suggests
-that introducing QREs saves programmers 
+as \LensAndSpecSize{}. This demonstrates that introducing QREs saves programmers
 significant effort compared to both Optician and basic Boomerang.
@@
 We found that 10 of Optician's 39 benchmarks required modifications to be
-bijective. We found that 4 of Optician's 39 benchmarks required modifications to
-be in bijective correspondence when in canonical form.\saz{Is this somehow where
-  the 14 = 10+4 points in Fig. 10 come from?}\saz{Do these 10 and 4 intersect?}
+bijective. Of these We found that 4 of Optician's 39 benchmarks required
+modifications to be in bijective correspondence when in canonical form.
 Furthermore, we found that all eight of the data.gov examples required no
@@
 Also notable is one of the data.gov examples, which converts demographic
-statistics by zip code from xml form into rdf form. This took longer because it
-was an exceptionally large format, converting over 40 fields.\saz{We need to
-  explain more detail here.  I guess this is about the last benchmark in Fig. 11
-  (b), which takes a long time?}  We added in additional fields to make this
-format suitable for Optician, which took over 15 seconds synthesize the modified
-lens.\saz{This needs more context, explanation.}
+statistics by zip code from xml form into rdf form. To validate this performance
+was not due to our overhead, we modified this format to be bijective, so it
+could be input into Optician. Optician took over 15 seconds synthesize this
+modified lens. We believe this conversion took longer because the input formats
+were exceptionally large, requiring conversions between over 40 fields.
 

### 2018-03-09 papers/QuotientLenses/transform-data.py

@@ -72,7 +72,7 @@ def generate_time_vs_tasks_graph(input_csv):
     normal_size = 2
     full_size = 3
 
-    ax.step([0,8000],[14,14],label="All QRE\nBenchmarks",linestyle=":",
+    ax.step([0,8000],[14,14],label="Benchmarks With\nNontrivial Quotients",linestyle=":",
             linewidth=1, dashes=(1,1))
 
     create_step_plot("OO","\\textbf{QS}",'-',normal_size)
@@ -83,8 +83,8 @@ def generate_time_vs_tasks_graph(input_csv):
     ax.set_xlabel('AST Count')
     ax.set_title("AST Count vs Benchmarks Definable")
 
-    l = ax.legend(bbox_to_anchor=(1.6,1),borderaxespad=0)
-    plt.setp(l.texts,weight='bold') 
+    l = ax.legend(bbox_to_anchor=(1.56,1),borderaxespad=0)
+    plt.setp(l.texts,weight='bold')
 
     plt.xticks(np.arange(0, 7501, 2500))
     plt.yticks(np.arange(0, 15.1, 5))

### 2018-03-09 papers/QuotientLenses/main.tex

--- 2018-03-09 papers/QuotientLenses/main.tex
@@
 Also notable is one of the data.gov examples, which converts demographic
-statistics by zip code from xml form into rdf form. This took longer because it
-was an exceptionally large format, converting over 40 fields.\saz{We need to
-  explain more detail here.  I guess this is about the last benchmark in Fig. 11
-  (b), which takes a long time?\bcp{It's a bit hidden in the figure...}}  We
-added in additional fields to make this 
-format suitable for Optician, which took over 15 seconds synthesize the modified
-lens.\saz{This needs more context, explanation.}
+statistics by zip code from xml form into rdf form.\bcp{It's a bit hidden in the figure...} To validate this performance
+was not due to our overhead, we modified this format to be bijective, so it
+could be input into Optician. Optician took over 15 seconds synthesize this
+modified lens. We believe this conversion took longer because the input formats
+were exceptionally large, requiring conversions between over 40 fields.
 
@@
 
+\bcp{Some of the citations need some love---e.g., the first one.}
 \bibliographystyle{plain}

### 2018-03-10 papers/QuotientLenses/main.tex
More edits

@@
 framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=false,
-keepspaces=true,
 commentstyle=\itshape\rmfamily,
@@
 \newcommand{\saz}[1]{\FINISH{orange}{SZ}{#1}}
-\newcommand{\SAZ}[1]{\FINISH{orange}{SZ}{\bf #1}}
+\newcommand{\SAZ}[1]{\FINISH{orange}{SZ}{#1}}
 \newcommand{\ksf}[1]{\FINISH{teal}{K}{#1}}
@@
 
-Amongst the most restrictive lenses are the \emph{bijective lenses}.  These
-lenses demand that get and put functions form a bijection, meaning no
-information is lost when one translates from one data format to another.  While
-this strong guarantee can be useful in some contexts, it is more often the case
-that some of the information in one or both of the data formats is superfluous.
-For instance, one typically does not care to preserve the amount of
+Amongst the most restrictive lenses are the \emph{bijective lenses}.
+A bijective lens $\ell : S \Leftrightarrow T$ is simply a bijection from
+a source set $S$ to a target set $T$:
+\begin{equation}\label{bijectivelenslaws}
+\ell.\get \; (\ell.\lput \; t) = t \text{, and } \ell.\lput \; (\ell.\get \; s)
+= s
+\end{equation}
+Since these lenses demand that get and put functions form a bijection, then
+no information is lost when one translates from one data format to another. 
+While this strong guarantee can be useful in some contexts, it is more often
+the case that some of the information in one or both of the data formats is
+superfluous. For instance, one typically does not care to preserve the amount of
 whitespace between characters.  In other circumstances, capitalization
@@
 
-{\em Quotient lenses} loosen many of the restrictions of bijective lenses by
-allowing a programmer to define an equivalence relation on the source and
-target formats so that the round-tripping lens laws hold modulo these
-equivalence rather than modulo equality. Quotient lenses thus make it possible
-to align many more data sources than would otherwise be possible, by ignoring
-details that a programmer views as inessential. The only language we are aware
-of that implements a general class of quotient lenses is {\em
-Boomerang}~\cite{quotientlenses}\bcp{Is this the right citation for
-Boomerang?  Or should it be \cite{boomerang}?}. While Boomerang's quotient
-lenses enable a 
-programmer to define a very general class of quotient lenses\bcp{read
-  literally, that part of the sentence is nonsense: boomerang's lenses
-  (themselves) do not enable the programmer to define lenses.  Of course,
-  the reader can probably figure out what is meant, but that's a needless
-  tax on their finite mental energy.}, these lenses to have two main
+{\em Quotient lenses}~\cite{quotientlenses} loosen many of the restrictions of
+bijective string lenses by allowing a programmer to define an equivalence
+relation on the source and target formats so that the round-tripping lens laws
+hold modulo these equivalence rather than modulo equality. Quotient lenses thus
+make it possible to align many more data sources than would otherwise be
+possible, by ignoring details that a programmer views as inessential. The only
+language we are aware of that implements a general class of quotient lenses is
+{\em Boomerang}~\cite{boomerang}. While Boomerang enables a programmer to
+define a very general class of quotient lenses, these lenses to have two main
 drawbacks.
@@
   for that relation (Section~\ref{QRE}).
-  \item We define a language of {\em QRE lenses}.  This language is
-  a subset of the full set of bijective quotient
-  lenses whose types are given by QREs. Our main technical contribution
-  is a proof that each QRE lens is functionally equivalent to a QRE lens in the
-  canonical form of a bijective lens with ``canonizers at the edges''.
+  \item We introduce {\em QRE lenses} which are lenses that map data specified
+  using QREs.
+  In addition, we will descibe a particular set of QRE lens combinators and
+  prove that QRE lenses defined using these combinators can be transformed to
+  a QRE lens that us defined using just two of these combinators; this is the
+  main technical contribution of this paper.
   (Section~\ref{QRE-lenses}).
-  \item We reduce the problem of {\em synthesizing}
-  QRE lenses to the previously studied problem of synthesizing bijective lenses
-  between regular languages (Section~\ref{synth}).
+  \item We use this result to reduce the problem of {\em synthesizing} QRE
+  lenses to the problem of synthesizing bijective string lenses which we studied
+  in~\cite{optician}(Section~\ref{synth}).
   \item We extend the Boomerang {\em implementation} with QREs
-  and QRE lens synthesis.  We demonstrate the utility of our
-  implementation by using it to
-  synthesize QRE lenses between data formats drawn from the
+  and QRE lens synthesis.  We demonstrate the utility of our implementation by
+  using it to synthesize QRE lenses between data formats drawn from the
   {\tt data.gov} database (Section~\ref{impl}).
@@
 \label{sec:example}
-\subsection{Bijective Lenses}
+\subsection{Bijective String Lenses}
 Consider a bidirectional transformation that converts \bibtex{} citation records
@@
 
-{\em Bijective lenses} are designed to define a bidirectional transformation
-such as this, where the data formats can be matched in a one-to-one manner, in
-this case by matching the label, author and title fields. Boomerang allows
-programmers to define such bijecive transformations using combinators which
-favour a compositional approach in which the programmer defines small lenses
-which are then put together using the combinators to define larger and larger
-lenses until the final lens is defined.
-
-Each bijective lens is constructed from simple primitive combinators including:
+Boomerang's bijective string lenses are designed to define a bidirectional
+transformation such as this, where the data formats can be matched in a
+one-to-one manner, in this case by matching the label, author and title fields.
+Boomerang allows programmers to define such bijecive transformations using
+combinators which favour a compositional approach in which the programmer
+defines small lenses which are then put together using the combinators to
+define larger and larger lenses until the final lens is defined.
+
+Each bijective string lens is constructed from simple primitive combinators including:
 \begin{itemize}
@@
 
-Unfortunately, programming bijective lenses in Boomerang is often slow and
+Unfortunately, programming bijective string lenses in Boomerang is often slow and
 tediuos because of the fiddly constraints that Boomerang imposes on its
@@
 In our previous work in~\cite{optician}, we presented a solution to these
-shortcomings of bijective lenses by developing a tool called {\em Optician}
-which enables a programmer to synthesize a bijective lens from a pair of regular
+shortcomings of bijective string lenses by developing a tool called {\em Optician}
+which enables a programmer to synthesize a bijective string lens from a pair of regular
 expressions $(S, T)$ and a set of input-output example \cd{exs} pairs that
@@
 Optician}, the programmer need only spend the effort of specifying the data
-formats and the behaviour\bcp{Let's use American spellings: behavior} of the
-lens on a few inputs in order to define a 
-bijective lens. As a result defining bijective lenses becomes faster and less
+formats and the behavior of the lens on a few inputs in order to define a
+bijective string lens. As a result defining bijective string lenses becomes faster and less
 error prone, since the programmer need only invoke the synthesis procedure
-\lstinline{synth S <=> T using exs} in order to define a bijective lens rather
+\lstinline{synth S <=> T using exs} in order to define a bijective string lens rather
 than manually derive the lens herself.
@@
 Unfortunately, while our synthesis tool greatly simplifies the task of
-programming bijective lenses, not all bidirectional tranformations are bijective
+programming bijective string lenses, not all bidirectional tranformations are bijective
 in nature. For instance, users are typically uninterested in preserving
@@
   target formats.
-  \item A bijective lens to map back and forth between canonical elements
+  \item A bijective string lens to map back and forth between canonical elements
   of the internal source and target formats.
@@
 canonizer from the outer regular expression to the inner one. In the following
-subsection, we describe the QRE combinators and use give an example of how a
-programmer might define a QRE specification of \bibtex.
+subsection, we describe a set of QRE combinators that we have integrated into
+Boomerang and give an example of how a programmer might define a QRE
+specification of \bibtex{} using these combinators.
 
@@
 and that the \codefont{space\_name} format describes ``John Doe''
-and \codefont{s\_to\_c} is a function from the first to the second.  In this case,
-the following instance of \kw{squash} creates the desired canonizer.
+and \codefont{c\_to\_s} is a function from the first to the second.  In this
+case, the following instance of \kw{squash} creates the desired canonizer.
 
@@
 \begin{lstlisting}
-let name = squash comma_name $\to$ space_name using s_to_c
+let name = squash comma_name $\to$ space_name using c_to_s
 \end{lstlisting}
 
-One way to define the \codefont{s\_to\_c} canonization function used by squash
-is simply to write it from scratch in some ordinary programming language. (Boomerang canonizers are written in OCaml.)
+One way to define the \codefont{c\_to\_s} canonization function used by squash
+is simply to write it from scratch in some ordinary programming language.
 However, our system will also synthesize such functions automatically from pairs
-of regular languages and examples---here, \codefont{s\_to\_c} is the
+of regular languages and examples---here, \codefont{c\_to\_s} is the
 \codefont{get} direction of a lens that can be synthesized by:
@@
 let l = synth comma_name $\Leftrightarrow$ space_name using {("Doe, John", "John Doe")}
-let s_to_c = get l
+let c_to_s = get l
 \end{lstlisting}
-%
-\bcp{Three problems here. First, the idea of using half of a lens
-as a canonizer comes out of nowhere.  The second is that this idea is a hack
-(and always was).  Why bother defining / synthesizing a lens and then using only
-half of it?  And the third is that, in the original quotient lens paper,
-we used this trick on lenses that were not even bijective, which may confuse
-people familiar with the earlier paper.}
-
-\sam{OK I suppose it's a trick though I don't see an obvious solution. }
 
@@
 let l = synth comma_name $\Leftrightarrow$ space_name using {("Doe, John", "John Doe")}
-let s_to_c = get l
+let c_to_s = get l
 
@@
 \label{sec:examplesynth}
-At this point we already have a tool for synthesizing bijective lenses from a
+At this point we already have a tool for synthesizing bijective string lenses from a
 pair of regular expressions and a set of example input-output pairs, as well as
@@
 An obvious next step in simplifying the quotient lens programming is thus to use
-the bijective lens synthesis procedure as a subroutine for a quotient lens
+the bijective string lens synthesis procedure as a subroutine for a quotient lens
 synthesis procedure. This new procedure should take two QREs and a set
@@
 to their canonical representatives to get a set of examples in canonical form,
-and then invoke the bijective lens synthesis procedure on the canonical data
+and then invoke the bijective string lens synthesis procedure on the canonical data
 formats and the new set of examples that are in canonical form.
@@
 \includegraphics[width=0.7\textwidth]{canonizers-outside}
-\caption{Quotient Lens with ``Canonizers at the Edges''
-\bcp{Maybe we should go a step further and label things like $K(c)$,
- $W(c)$, Canonizer(c), etc.  And probably we should display K as a subset of
- W.  I can make these changes later.)}}
+\caption{QRE Lens with ``Canonizers at the Edges''}
 \label{fig:attheedges}
 \end{figure}
-Every QRE lens $q_1$ has a type $q_1: c_1 \Leftrightarrow c_2$ where $c_1$ and $c_2$
-are QREs. In the get direction, a QRE lens $q_1: c_1 \Leftrightarrow c_2$ uses the
-source QRE $c_1$ to compute a unique representative for the data modulo the
-equivalence relation defined by $c_1$ and then applies the $\get$ function of a
-bijective lens $\ell_1$ to this representative. In the backward direction, $q_1$
-operates similarly, but using the view QRE $c_2$ and the $\lput$ function of
+Every QRE lens $q_1$ has a type $q_1: Q_1 \Leftrightarrow Q_2$ where $Q_1$ and $Q_2$
+are QREs. In the get direction, a QRE lens $q_1: Q_1 \Leftrightarrow Q_2$ uses the
+source QRE $Q_1$ to compute a unique representative for the data modulo the
+equivalence relation defined by $Q_1$ and then applies the $\get$ function of a
+bijective string lens $\ell_1$ to this representative. In the backward direction, $q_1$
+operates similarly, but using the view QRE $Q_2$ and the $\lput$ function of
 $\ell_1$.
 
-Because the QREs $c_1$ and $c_2$ determine the internal formats for data after
-canonization, and because the algorithm for synthesizing bijective lenses is
-directed by these formats, $c_1$ and $c_2$ are all that is required to
+Because the QREs $Q_1$ and $Q_2$ determine the internal formats for data after
+canonization, and because the algorithm for synthesizing bijective string lenses is
+directed by these formats, $Q_1$ and $Q_2$ are all that is required to
 synthesize QRE lenses end-to-end. Indeed, we structured QRE lenses as bijective
@@
 This technical investigation justifies replacing manually constructed lenses
-using the QRE lens combinators with synthesized QRE lenses in normal form, thus
-saving the programmer. For instance, after defining the \bibtex{} and EndNote
-QREs and storing them in the variables \cd{bibtex} and \cd{endnote}
-respectively, then all the code in Figure~\ref{fig:example-lens} may be
-replaced by a single call to the synthesis prodedure:
+using the QRE lens combinators with synthesized QRE lenses with canonizers at
+the edges. For instance, after defining the \bibtex{} and EndNote QREs and
+storing them in the variables \cd{bibtex} and \cd{endnote} respectively, then
+all the code in Figure~\ref{fig:example-lens} may be replaced by a single call
+to the synthesis prodedure:
 
@@
 disambiguate. In addition, and as we saw earlier with the definition of
-\cd{s_to_c}, the synthesis procedure itself can be used to create lenses that
+\cd{c_to_s}, the synthesis procedure itself can be used to create lenses that
 are in turn used to define other QREs.  The ability to intermix QRE
@@
 \section{Quotient Regular Expressions}
+A Quotient Regular Expression (or QRE) is a regular expression $R$ augmented
+with syntax that expresses an equivalence relation on the language of $R$. This
+section describes in further detail the set of QRE combinators that we described
+in \cref{sec:example}.
 \label{QRE}
@@
 \begin{align*}
-c := \; &R \sep \project{R}{s} \sep \squash{c_1}{c_2}{f} \sep
-\perm{c_1, \ldots, c_n}{c} \\
-& | \; \normalize{R}{R'}{f} \sep c_1 \semicolon c_2 \sep c_1 \cdot c_2 \sep (c_1
-\sep c_2) \sep c^*,
-\end{align*}
-where $c$ ranges over QREs, $R$ ranges over regular expressions, $f$ ranges over
+Q := \; &R \sep \project{R}{s} \sep \squash{Q_1}{Q_2}{f} \sep
+\perm{Q_1, \ldots, Q_n}{Q} \\
+& | \; \normalize{R}{R'}{f} \sep Q_1 \semicolon Q_2 \sep Q_1 \cdot Q_2 \sep (Q_1
+\sep Q_2) \sep Q^*,
+\end{align*}
+where $Q$ ranges over QREs, $R$ ranges over regular expressions, $f$ ranges over
 functions between regular languages, and $s$ ranges over character strings.
@@
 Using the conventional notation that $\mathcal{L}(R)$ is the language accepted
-by the regular expression $R$, then each QRE $c$ enables us to express
+by the regular expression $R$, then each QRE $Q$ enables us to express
 \begin{enumerate}
-  \item a regular expression $W(c)$ (the ``whole'' of $c$),
-  \item an equivalence relation $\eqrel{c}$ on $\mathcal{L}(W(c))$,
-  \item a regular expression $K(c)$ (the ``kernel'' of $c$)
-  such that $\mathcal{L}(K(c))$ forms a complete set of representatives for
-  $\eqrel{c}$, and
+  \item a regular expression $W(Q)$ (the ``whole'' of $Q$),
+  \item an equivalence relation $\eqrel{Q}$ on $\mathcal{L}(W(Q))$,
+  \item a regular expression $K(Q)$ (the ``kernel'' of $Q$)
+  such that $\mathcal{L}(K(Q))$ forms a complete set of representatives for
+  $\eqrel{Q}$, and
   \item a ``canonizing'' function\bcp{Why is it called ``canonizer'' rather
-    than ``canonize''??} $\canonizer(c):\mathcal{L}(W(c))
-  \longrightarrow \mathcal{L}(K(c))$ which given any $w \in \mathcal{L}(W(c))$,
-  computes $\canonizer(c)(w)$ as the unique $k$ in $\mathcal{L}(K(c))$ such that
-  $k$ is equivalent to $w$ mod $\eqrel{c}$
+  than ``canonize''??}\sam{because \canonizer(Q) is a function}
+  $\canonizer(Q):\mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$ which
+  given any $w \in \mathcal{L}(W(Q))$, computes $\canonizer(Q)(w)$ as the
+  unique $k$ in $\mathcal{L}(K(Q))$ such that $k$ is equivalent to $w$ mod
+  $\eqrel{Q}$
 \end{enumerate}
 Figures~\ref{fig:wk} gives the inductive definitions for the whole and
-kernel languages $W(c)$ and $K(c)$ of a QRE.
+kernel languages $W(Q)$ and $K(Q)$ of a QRE.
 \begin{figure}[t]
@@
 
-c & W(c) & K(c) \\ \hline
+Q & W(Q) & K(Q) \\ \hline
 R & R & R \\
 \project{R}{s} & R & s \\
-\squash{c}{c_1}{f} & W(c) \sep W(c_1) & K(c_1) \\
+\squash{Q}{Q_1}{f} & W(Q) \sep W(Q_1) & K(Q_1) \\
 \normalize{R}{R'}{f} & R & R' \\
-c_1 \; ; \; c_2 & W(c_1) & K(c_2) \\
-c_1 \cdot c_2 & W(c_1) \cdot W(c_2) & K(c_1) \cdot K(c_2) \\
-c_1 \sep c_2 & W(c_1) \sep W(c_2) & K(c_1) \sep K(c_2) \\
-c^* & W(c)^* & K(c)^* \\
+Q_1 \; ; \; Q_2 & W(Q_1) & K(Q_2) \\
+Q_1 \cdot Q_2 & W(Q_1) \cdot W(Q_2) & K(Q_1) \cdot K(Q_2) \\
+Q_1 \sep Q_2 & W(Q_1) \sep W(Q_2) & K(Q_1) \sep K(Q_2) \\
+Q^* & W(Q)^* & K(Q)^* \\
 \end{array}
@@
 \begin{array}{r@{\quad}l}
-W( \perm{c_1, \ldots, c_n}{c} ) = &
-\bigcup \limits_{\sigma \in S_n} W(c_{\sigma(1)}) \cdot W(c) \cdot \ldots \cdot
-W(c) \cdot W(c_{\sigma(n)})\\
-K( \perm{c_1, \ldots, c_n}{c} ) = & K(c_1) \cdot K(c) \cdot \ldots \cdot K(c)
-\cdot K(c_n)
+W( \perm{Q_1, \ldots, Q_n}{Q} ) = &
+\bigcup \limits_{\sigma \in S_n} W(c_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
+W(Q) \cdot W(c_{\sigma(n)})\\
+K( \perm{Q_1, \ldots, Q_n}{Q} ) = & K(Q_1) \cdot K(Q) \cdot \ldots \cdot K(Q)
+\cdot K(Q_n)
 \end{array}
@@
 The \textit{squash} and \textit{permutation} combinators give the two
-interesting definitions. If $c = \squash{c_1}{c_2}{f}$, then the whole language
-of $c$ is $W(c_1) \sep W(c_2)$. This is because the \textit{squash} combinator
-takes the whole language $W(c_1)$ of $c_1$, the whole language $W(c_2)$ of $c_2$
-and a function $f : \mathcal{L}(W(c_1)) \longrightarrow \mathcal{L}(W(c_2))$,
-maps $\mathcal{L}(W(c_1))$ into $\mathcal{L}(W(c_2))$ using $f$ and then
-canonizes $W(c_2)$ into $K(c_2)$ using $\canonizer(c_2)$.
-
-For the $\perm{c_1, \ldots, c_n}{c}$ combinator, the whole language is the
-union of languages of the form $W(c_{\sigma(1)}) \cdot W(c) \cdot \ldots \cdot
-W(c) \cdot W(c_{\sigma(n)})$ for any permutation $\sigma$ in $S_n$, where
+interesting definitions. If $Q = \squash{Q_1}{Q_2}{f}$, then the whole language
+of $Q$ is $W(Q_1) \sep W(Q_2)$. This is because the \textit{squash} combinator
+takes the whole language $W(Q_1)$ of $Q_1$, the whole language $W(Q_2)$ of $Q_2$
+and a function $f : \mathcal{L}(W(Q_1)) \longrightarrow \mathcal{L}(W(Q_2))$,
+maps $\mathcal{L}(W(Q_1))$ into $\mathcal{L}(W(Q_2))$ using $f$ and then
+canonizes $W(Q_2)$ into $K(Q_2)$ using $\canonizer(Q_2)$.
+
+For the $\perm{Q_1, \ldots, Q_n}{Q}$ combinator, the whole language is the
+union of languages of the form $W(c_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
+W(Q) \cdot W(c_{\sigma(n)})$ for any permutation $\sigma$ in $S_n$, where
 $S_n$ is the set of all permutations of the numbers $1$ to $n$. This is because
 the \textit{permutation} combinator allows for the string to match any
-permutation the $c_i$'s while also taking into account the separator $c$ in
-between each of the $c_i$'s. The kernel of the \textit{permutation} combinator
-is the language $K(c_1) \cdot K(c) \cdot \ldots \cdot K(c) \cdot K(c_n)$,
+permutation the $Q_i$'s while also taking into account the separator $Q$ in
+between each of the $Q_i$'s. The kernel of the \textit{permutation} combinator
+is the language $K(Q_1) \cdot K(Q) \cdot \ldots \cdot K(Q) \cdot K(Q_n)$,
 because the canonical permutation is the identity permuation, with each of the
-parts of the string that match $c_i$ and $c$ canonized into $K(c_i)$ and $K(c)$
+parts of the string that match $Q_i$ and $Q$ canonized into $K(Q_i)$ and $K(Q)$
 respectively.
@@
 \canonizer(\project{R}{s})(w) &=& s\\
-\canonizer(\squash{c_1}{c_2}{f})(w) &=&
+\canonizer(\squash{Q_1}{Q_2}{f})(w) &=&
 \begin{cases}
-\canonizer(c_1)(f(w)) & \text{if } w \in \mathcal{L}(W(c_1))\\
-\canonizer(c_2)(w) & \text{otherwise}
+\canonizer(Q_1)(f(w)) & \text{if } w \in \mathcal{L}(W(Q_1))\\
+\canonizer(Q_2)(w) & \text{otherwise}
 \end{cases}\\
 \canonizer(\normalize{R}{R'}{f}) &=& f\\
-\canonizer(c_1 \; ; \; c_2) &=& \canonizer(c_2) \circ \canonizer(c_1)\\
-\canonizer(c_1 \cdot c_2) &=& \canonizer(c_1) \cdot \canonizer(c_2)\\
-\canonizer(c_1 \sep c_2)(w) &=&
+\canonizer(Q_1 \; ; \; Q_2) &=& \canonizer(Q_2) \circ \canonizer(Q_1)\\
+\canonizer(Q_1 \cdot Q_2) &=& \canonizer(Q_1) \cdot \canonizer(Q_2)\\
+\canonizer(Q_1 \sep Q_2)(w) &=&
 \begin{cases}
-\canonizer(c_1)(w) & \text{if } w \in \mathcal{L}(W(c_1))\\
-\canonizer(c_2)(w) & \text{if } w \in \mathcal{L}(W(c_2))\\
+\canonizer(Q_1)(w) & \text{if } w \in \mathcal{L}(W(Q_1))\\
+\canonizer(Q_2)(w) & \text{if } w \in \mathcal{L}(W(Q_2))\\
 \end{cases}\\
-\canonizer(c^*) &=& \canonizer(c)^* \\
+\canonizer(Q^*) &=& \canonizer(Q)^* \\
 \end{array}
@@
 \begin{align*}
-&\canonizer(\perm{c_1, \ldots, c_n}{c})(r_{\sigma(1)}
+&\canonizer(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
 \cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)})\\
-&= (\canonizer(c_1)(r_1)) \cdot (\canonizer(c)(s_1)) \cdot \ldots \cdot
-(\canonizer(c_n)(r_n))
+&= (\canonizer(Q_1)(r_1)) \cdot (\canonizer(Q)(s_1)) \cdot \ldots \cdot
+(\canonizer(Q_n)(r_n))
 \end{align*}
@@
 definition,
-\begin{align*}\canonizer(\perm{c_1, \ldots, c_n}{c})(r_{\sigma(1)}
+\begin{align*}\canonizer(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
 \cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \\
-= (\canonizer(c_1)(r_1)) \cdot (\canonizer(c)(s_1)) \cdot \ldots \cdot
-(\canonizer(c_n)(r_n))
+= (\canonizer(Q_1)(r_1)) \cdot (\canonizer(Q)(s_1)) \cdot \ldots \cdot
+(\canonizer(Q_n)(r_n))
 \end{align*}
 \noindent which, as we have explained, is defined so that the strings that match
-$c_i, c$ are placed in the canonical permutation which is the identity
-permutation before $\canonizer(c_i)$ and $\canonizer(c)$ are applied
+$Q_i, Q$ are placed in the canonical permutation which is the identity
+permutation before $\canonizer(Q_i)$ and $\canonizer(Q)$ are applied
 respectively.
@@
 Finally, Figure ~\ref{fig:relations} gives the inductive definition of the
-equivalence relation $\eqrel{c}$. This is the set-theoretic semantics of QREs
+equivalence relation $\eqrel{Q}$. This is the set-theoretic semantics of QREs
 as an equivalence relation on a regular language.
@@
 w \; \equiv_{\project{R}{s}} \; w' && \text{for all }w, w'\\
-w \; \equiv_{\squash{c_1}{c_2}{f}} \; w' &\iff& f(w) \equiv_{c_2} w'
-\text{ or } w \equiv_{c_2} w' \\
+w \; \equiv_{\squash{Q_1}{Q_2}{f}} \; w' &\iff& f(w) \equiv_{Q_2} w'
+\text{ or } w \equiv_{Q_2} w' \\
 w \; \equiv_{\normalize{R}{R'}{f}} \; w' &\iff&
 f(w)=f(w') \text{ or }w = w'\\
-w \; \equiv_{c_1 \; ; \; c_2} \; w' &\iff& \exists k, k' \in
-\mathcal{L}(K(c_2)) \text{ such that } w \; \equiv_{c_1} \; k, \; w' \;
-\equiv_{c_1} \; k', \text{ and } k \; \equiv_{c_2} \; k'\\
-w \; \equiv_{c_1 \cdot c_2} \; w'  &\iff& w = r_1
-\cdot r_2, \; w' = {r'}_1 \cdot {r'}_2 \text{ with } r_1 \; \equiv_{c_1}
-\; {r'}_1, \; r_2 \; \equiv{c_n} \; {r'}_2\\
-w \; \equiv_{c_1 \sep c_2} \; w' &\iff& w \; \equiv_{c_1} \; w'
-\text{ or } \; w \; \equiv_{c_2} \; w'\\
-w \; \equiv_{c^*} \; w' &\iff& w = r_1 \cdot \ldots \cdot r_n, \; w'
-= {r'}_1 \cdot \ldots \cdot {r'}_n \text{ and } r_i \equiv_{c} \; {r'}_i
+w \; \equiv_{Q_1 \; ; \; Q_2} \; w' &\iff& \exists k, k' \in
+\mathcal{L}(K(Q_2)) \text{ such that } w \; \equiv_{Q_1} \; k, \; w' \;
+\equiv_{Q_1} \; k', \text{ and } k \; \equiv_{Q_2} \; k'\\
+w \; \equiv_{Q_1 \cdot Q_2} \; w'  &\iff& w = r_1
+\cdot r_2, \; w' = {r'}_1 \cdot {r'}_2 \text{ with } r_1 \; \equiv_{Q_1}
+\; {r'}_1, \; r_2 \; \equiv{Q_n} \; {r'}_2\\
+w \; \equiv_{Q_1 \sep Q_2} \; w' &\iff& w \; \equiv_{Q_1} \; w'
+\text{ or } \; w \; \equiv_{Q_2} \; w'\\
+w \; \equiv_{Q^*} \; w' &\iff& w = r_1 \cdot \ldots \cdot r_n, \; w'
+= {r'}_1 \cdot \ldots \cdot {r'}_n \text{ and } r_i \equiv_{Q} \; {r'}_i
 \\
-w \; \equiv_{\perm{c_1, \ldots, c_n}{c}} \; w' &\iff& w = r_{\sigma(1)}
+w \; \equiv_{\perm{Q_1, \ldots, Q_n}{Q}} \; w' &\iff& w = r_{\sigma(1)}
 \cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}, \;
@@
 & & \text{for some } \sigma, \theta \in S_n, \text{ with } r_i \;
-\equiv_{q_i} \; r'_i \text{ and } s_k \; \equiv_{c} \; s'_{k}
+\equiv_{q_i} \; r'_i \text{ and } s_k \; \equiv_{Q} \; s'_{k}
 \end{array}
@@
 defined using the other combinators. Secondly, the {\em normalize} canonizer
-enjoys the property that each of the other combinators $c$ is semantically
-equivalent to $\normalize{W(c)}{K(c)}{f}$ for some surjective, idempotent
-function $f : \mathcal{L}(W(c)) \longrightarrow \mathcal{L}(K(c))$; this is the
+enjoys the property that each of the other combinators $Q$ is semantically
+equivalent to $\normalize{W(Q)}{K(Q)}{f}$ for some surjective, idempotent
+function $f : \mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$; this is the
 second property that distinguishes the {\em normalize} combinator from the other
@@
 rules for QREs, we discuss the issue of {\em ambiguity}, which appears in the
-premises of the inference rules for the regular combinators. If $c_1$ and $c_2$
+premises of the inference rules for the regular combinators. If $Q_1$ and $Q_2$
 are QREs, then when applying the regular combinators to QREs, we require that
-if a string $s$ matches any of the regular expressions $W(c_1) \cdot W(c_2)$,
-\quad $W(c_1) \sep W(c_2)$, \quad $W(c_1)^*$, \quad $K(c_1) \cdot K(c_2)$,
-\quad $K(c_1) \sep K(c_2)$, \quad $K(c_1)^*$, then $s$ matches that regular
+if a string $s$ matches any of the regular expressions $W(Q_1) \cdot W(Q_2)$,
+\quad $W(Q_1) \sep W(Q_2)$, \quad $W(Q_1)^*$, \quad $K(Q_1) \cdot K(Q_2)$,
+\quad $K(Q_1) \sep K(Q_2)$, \quad $K(Q_1)^*$, then $s$ matches that regular
 expression in only one way.
@@
 combinators applied to the kernels of QREs be unambiguous since the underlying
-bijective lens of a QRE lens operates on kernels, and bijective lenses
+bijective string lens of a QRE lens operates on kernels, and bijective string lenses
 impose the same unambiguity restrictions so that they too are well defined as
@@
 \begin{prooftree}
-\AxiomC{$\wf{c_i, c}$}
+\AxiomC{$\wf{Q_i, Q}$}
 \RightLabel{(Perm)}
-\AxiomC{${\substack{\forall \sigma \neq \theta, \; W(c_{\sigma(1)}) \cdot W(c)
-\cdot \ldots \cdot W(c_{\sigma(n)})\\ \cap W(c_{\theta(1)}) \cdot W(c)
+\AxiomC{${\substack{\forall \sigma \neq \theta, \; W(c_{\sigma(1)}) \cdot W(Q)
+\cdot \ldots \cdot W(c_{\sigma(n)})\\ \cap W(c_{\theta(1)}) \cdot W(Q)
 \cdot \ldots \cdot W(c_{\theta(n)}) =\varnothing}}$}
-\AxiomC{$K(c_1) \cdot^! K(c) \cdot^! \ldots \cdot^! K(c) \cdot^! K(c_n)$}
-\TrinaryInfC{$\wf{\perm{c_1, \ldots, c_n}{c}}$}
+\AxiomC{$K(Q_1) \cdot^! K(Q) \cdot^! \ldots \cdot^! K(Q) \cdot^! K(Q_n)$}
+\TrinaryInfC{$\wf{\perm{Q_1, \ldots, Q_n}{Q}}$}
 \end{prooftree}
@@
 \begin{center}
-\AxiomC{$\wf{c_1, c_2}$}
-\AxiomC{$\mathcal{L}(W(c_1)) \cap \mathcal{L}(W(c_2)) = \varnothing$}
-\AxiomC{$f : \mathcal{L}(W(c_1)) \longrightarrow \mathcal{L}(W(c_2))$}
+\AxiomC{$\wf{Q_1, Q_2}$}
+\AxiomC{$\mathcal{L}(W(Q_1)) \cap \mathcal{L}(W(Q_2)) = \varnothing$}
+\AxiomC{$f : \mathcal{L}(W(Q_1)) \longrightarrow \mathcal{L}(W(Q_2))$}
 \RightLabel{(Squash)}
-\TrinaryInfC{$\wf{\squash{c_1}{c_2}{f}}$}
+\TrinaryInfC{$\wf{\squash{Q_1}{Q_2}{f}}$}
 \DisplayProof
@@
 \begin{center}
-\AxiomC{$\wf{c_1, c_2}$}
-\AxiomC{$K(c_1) = W(c_2)$}
+\AxiomC{$\wf{Q_1, Q_2}$}
+\AxiomC{$K(Q_1) = W(Q_2)$}
 \RightLabel{(Compose)}
-\BinaryInfC{$\wf{c_1 \; ; \; c_2}$}
+\BinaryInfC{$\wf{Q_1 \; ; \; Q_2}$}
 \DisplayProof
 \hskip 1.5em
-\AxiomC{$\wf{c}$}
-\AxiomC{$W(c)^{*!}$}
-\AxiomC{$K(c)^{*!}$}
+\AxiomC{$\wf{Q}$}
+\AxiomC{$W(Q)^{*!}$}
+\AxiomC{$K(Q)^{*!}$}
 \RightLabel{(Star)}
-\TrinaryInfC{$\wf{c^*}$}
+\TrinaryInfC{$\wf{Q^*}$}
 \DisplayProof
@@
 \begin{center}
-\AxiomC{$\wf{c_1, c_2}$}
-\AxiomC{$W(c_1) \cdot^! W(c_2)$}
-\AxiomC{$K(c_1) \cdot^! K(c_2)$}
+\AxiomC{$\wf{Q_1, Q_2}$}
+\AxiomC{$W(Q_1) \cdot^! W(Q_2)$}
+\AxiomC{$K(Q_1) \cdot^! K(Q_2)$}
 \RightLabel{(Concat)}
-\TrinaryInfC{$\wf{c_1 \cdot c_2}$}
+\TrinaryInfC{$\wf{Q_1 \cdot Q_2}$}
 \DisplayProof
@@
 \begin{center}
-\AxiomC{$\wf{c_1, c_2}$}
-\AxiomC{$W(c_1) \cap W(c_2) = \varnothing$}
+\AxiomC{$\wf{Q_1, Q_2}$}
+\AxiomC{$W(Q_1) \cap W(Q_2) = \varnothing$}
 \RightLabel{(Union)}
-\BinaryInfC{$\wf{c_1 \sep c_2}$}
+\BinaryInfC{$\wf{Q_1 \sep Q_2}$}
 \DisplayProof
@@
 QREs using the regular combinators. For example, the (Concat) inference rule
-for concatenation says that the concatenation $c_1 \cdot c_2$ of QREs $c_1$ and
-$c_2$ is well formed only if the concatentions of $W(c_1)$ and $W(c_2)$, and
-$K(c_1)$ and $K(c_2)$ are unambiguous.
+for concatenation says that the concatenation $Q_1 \cdot Q_2$ of QREs $Q_1$ and
+$Q_2$ is well formed only if the concatentions of $W(Q_1)$ and $W(Q_2)$, and
+$K(Q_1)$ and $K(Q_2)$ are unambiguous.
 
@@
 \textit{permutation} rule says that for any two different permutations $\sigma$
-and $\theta$, the languages $W(c_{\sigma(1)}) \cdot W(c) \cdot \ldots \cdot
-W(c_{\sigma(n)})$ and $W(c_{\theta(1)}) \cdot W(c) \cdot \ldots \cdot
+and $\theta$, the languages $W(c_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
+W(c_{\sigma(n)})$ and $W(c_{\theta(1)}) \cdot W(Q) \cdot \ldots \cdot
 W(c_{\theta(n)})$ must be disjoint. This is because any string that matches the
 \textit{permutation} combinator matches the language $W(c_{\theta(1)}) \cdot
-W(c) \cdot \ldots \cdot W(c_{\theta(n)})$ for some permutation $\theta$, so we
+W(Q) \cdot \ldots \cdot W(c_{\theta(n)})$ for some permutation $\theta$, so we
 require that these regular expressions be disjoint.
 %
-The third hypothesis says that the regular expression $K(c_1) \cdot K(c)
-\cdot \ldots \cdot K(c) \cdot K(c_n)$ must be unambiguous. This is because the
+The third hypothesis says that the regular expression $K(Q_1) \cdot K(Q)
+\cdot \ldots \cdot K(Q) \cdot K(Q_n)$ must be unambiguous. This is because the
 underlying lens that maps to or from a \textit{permutation} QRE operates on
-the language $K(c_1) \cdot K(c) \cdot \ldots \cdot K(c) \cdot K(c_n)$ which is
+the language $K(Q_1) \cdot K(Q) \cdot \ldots \cdot K(Q) \cdot K(Q_n)$ which is
 the kernel of the \textit{permutation} QRE. As we mentioned, the underlying
@@
 directly on regular languages. QREs are therefore a good \textit{specification}
-language for quotient lenses. We next introduce
-\textit{QRE Lenses}, a class of quotient lenses specified using QREs.
-
-\bcp{A question about variable names: If we use $R$ and $S$ for plain
-  regular expressions, it seems inconsistent to use $c$ for QREs.  (And even
-  more inconsistent to use $c$ and $c'$ or $c$ and $c_1$!) I suggest $P$ and
-  $Q$ for QREs (with $Q$ as the default when we are only talking about
-  one).}\sam{I made the notation consistent by just using $c_1$ and $c_2$. I
-  personally feel that it is better not to use P and Q for QREs since we use
-  p and q for QRE lenses, but if you disagree I will make the change.}\bcp{I
-  agree that this is suboptimal, but I don't think it's worse than what we
-  have now.  Or maybe the right fix is to choose different metavariables for
-  QRE lenses.  Or to use capital letters other than $P$ and $Q$ for QREs.
-  In any case, I don't like $c_1$ and $c_2$, both for the lower-case letter
-  and for the subscripts---it's not at all consistent with our conventions
-  elsewhere.}
-
-Given regular expressions $R, S$ and equivalence relations
-$\sim_R$ and $\sim_S$ defined on $\mathcal{L}(R)$ and $\mathcal{L}(S)$,
-we\bcp{can we clarify that this is exactly Foster et al's 
-definition?}\sam{this is not Foster et al's definition. It's just a
-loosening of the definition for bijective lens used in the Optician
-paper}\bcp{OK, then we should say this}define
-a {\em bijective quotient lens} $q :
-R /{\sim_R} \Leftrightarrow S/{\sim_S}$ to be a pair of functions $q.\get :
+language for quotient lenses. In this section we introduce \textit{QRE Lenses},
+a class of quotient lenses that map between data that is specified using QREs.
+
+First we give the formal definition of a {\em bijective quotient lens}; QRE
+lenses are a specific class of bijective quotient lenses that map between data
+that is specified by QREs. Given regular expressions $R, S$ and equivalence
+relations $\sim_R$ and $\sim_S$ defined on $\mathcal{L}(R)$ and $\mathcal{L}(S)$,
+we define a {\em bijective quotient lens} $q : R /{\sim_R} \Leftrightarrow
+S/{\sim_S}$ to be a pair of functions $q.\get :
 \mathcal{L}(R) \longrightarrow \mathcal{L}(S)$ and $q.\lput : \mathcal{L}(S)
@@
 
-These laws are similar to the laws for bijective lenses\bcp{we should state
-  them explicitly somewhere in this paper and refer to them here by number},
-except that the 
-equality restrictions in the bijective lens laws are loosened to allow for
-equivalence relations. Also, the condition that $r \sim_R r'$ implies
-$q.\get(r) = q.\get(r')$ ensures that the {\em get} function induced on
-the equivalence classes if $\equiv_c$ is well-defined, and similarly for
-the {\em put} function.
-
-Now let $c_1$ and $c_2$ be QREs. Given a bijection $\ell : K(c_1)
-\Leftrightarrow K(c_2)$, we may define a quotient lens $q : W(c_1)/\eqrel{c_1}
-\Leftrightarrow W(c_2)/\eqrel{c_2}$ by $q.\get = \ell \circ \canonizer(c_1)
-$ and $q.\lput = \ell_1^{-1} \circ \canonizer(c_2)$. In other words, the get
-function of $q$ first canonizes the source data using $\canonizer(c_1)$
+These laws are similar to the laws for bijective lenses
+\cref{bijectivelenslaws}, except that the  equality restrictions in the
+bijective lens laws are loosened to allow for equivalence relations. Also, the
+condition that $r \sim_R r'$ implies $q.\get(r) = q.\get(r')$ ensures that the
+{\em get} function induced on the equivalence classes if $\equiv_c$ is
+well-defined, and similarly for the {\em put} function.
+
+Now let $Q_1$ and $Q_2$ be QREs. Given a bijection $\ell : K(Q_1)
+\Leftrightarrow K(Q_2)$, we may define a quotient lens $q : W(Q_1)/\eqrel{Q_1}
+\Leftrightarrow W(Q_2)/\eqrel{Q_2}$ by $q.\get = \ell \circ \canonizer(Q_1)
+$ and $q.\lput = \ell_1^{-1} \circ \canonizer(Q_2)$. In other words, the get
+function of $q$ first canonizes the source data using $\canonizer(Q_1)$
 and then applies $f$ to the result, while the put function of $q$ first
-canonizes the target data using $\canonizer(c_2)$ and the applies $f^{-1}$ to
+canonizes the target data using $\canonizer(Q_2)$ and the applies $f^{-1}$ to
 the result. This is our approach in defining QRE lenses.
 
-One potential pitfall\bcp{Why is this a pitfall?  I.e., why do we care about
-this norman form?  (The answer is synthesis, of course, but we haven't said
-so here.)} with QRE lenses is that the combinators that
-we define on them may not preserve this normal form\bcp{In what sense is
-  this a ``normal form''?  What you've given above is just a definition:
-  quotient lenses can be defined in such and such way.  You haven't said we
-  want to consider it a normal form, or specified what process it is the
-  normal form {\em of}.}. For example, the
-composition $q_1 \; ; \; q_2$ of QRE lenses $q_1$ and $q_2$ may not be
-equivalent to any quotient lens that has the canonical normal form of a
-bijective lens with canonizers a the edges. Consequently, in order to guarantee
-that every QRE lens $q_1$ is equivalent to a QRE lens $q_2$ in this
-canonical form\bcp{This paragraph uses variables $q_1$ and $q_2$ twice, with
-two different meanings.}, we need to identify a set $S$ of bijections that
-will ensure 
-that every QRE lens is functionally equivalent to a QRE lens in canonical
-form.\bcp{This is not very clear, in part because it is not very precise:
-  It is assuming, implicitly, that the class of QRE lenses is 
-  parameterized on a particular set of bijections on kernels, and then
-  asking what sets of bijections can be used.  This question makes sense,
-  but the reader may not realize that this is what's being asked. } 
-Our choice is to use the class of \textit{bijective lenses}, which we studied in
-\cite{optician} and has this desired property. The following part of the paper
-briefly discusses bijective lenses.
-
-\subsection{Bijective Lenses}
-\bcp{There's a persistent opportunity for confusion between the semantic
-  space of {\em all} bijective lenses (perhaps between arbitrary structures,
-  not just strings) and the rather specific bijections that are picked out
-  by our combinators.  One way to help, I think, would be to use the phrase
-  ``bijective string lenses'' instead of just ``bijective lenses.''}
-We define the set of \textit{bijective lenses} to be the set of bijections
-between regular languages created using Boomerang lens combinators.\bcp{Why
-  do we choose Boomerang's set of combinators?  Does it have some formal
-  property that we find attractive?  Or did it just happen to be lying around?}
-The syntax for the language of bijective lenses is given by
-$$\ell_1 := \mathit{id} \; (R) \sep \const(s_1, s_2) \sep  \swap(\ell_1,
-\ell_2) \sep \ell_1 \cdot \ell_2 \; |  \; (\ell_1 \sep \ell_2) \sep \ell^* \;
-| \; \ell_1 \; ; \;  \ell_2,$$ where $R$ ranges over regular expressions and $s$
-ranges over character strings.
+One potential pitfall with QRE lenses is that a lens derived using the QRE lens
+combinators may not have the form of a bijection with ``canonizers at the
+edges''. Such a situation would be undesirable since our strategy for
+synthesizing QRE lenses is to synthesize a bijective string lens which plugs
+into the two QREs provided as input to the synthesis procedure; if there were
+some QREs which could not be transformed to the desired form, then our synthesis
+algorithm would necessarily be incomplete.
+
+Thankfully, this turns out not to be an issue if we use the bijections that
+can be expressed in Boomerang; following ~\cite{optician} we call these
+lenses {\em bijective string lenses}. Once we restrict the bijections that we
+use to derive QRE lenses to include just bijective string lenses, we will be
+able show that every QRE lens can be transformed to a QRE lens which has the
+form of a bijective string lens with canonizers at the edges.
+
+\subsection{Bijective String Lenses}
+The syntax for the language of bijective string lenses is given by $$\ell_1 :=
+\mathit{id} \; (R) \sep \const(s_1, s_2) \sep  \swap(\ell_1, \ell_2) \sep
+\ell_1 \cdot \ell_2 \; |  \; (\ell_1 \sep \ell_2) \sep \ell^* \; | \; \ell_1 \;
+; \;  \ell_2,$$ where $R$ ranges over regular expressions and $s$ ranges over
+character strings.
 
@@
 data in the forward direction, and $t$ with $s$ in the target data in the backward
-direction. \bcp{Don't begin a sentence with a lower-case letter, even if it
-  is a math symbol.  Find a way to rewrite so that it doesn't need to be at
-  the beginning.}$\mathit{id}(R)$ applies the identity function to the
-source and target 
-in $\mathcal{L}(R)$ in both directions. The composition lens $\ell_1 \; ; \; \ell_2$
-applies $\ell_1$ followed by then $\ell_2$ to the source in the forward direction,
-and applies $\ell_2$ followed by $\ell_1$ to the target in the backward direction.
-The lens $\ell_1 \cdot \ell_2$ first splits the string $s$ into $s_1$ and $s_2$,
-applies $\ell_1$ and $\ell_2$ to $s_1$ and $s_2$ to get $t_1$ and $t_2$
-respectively, then concatenates $t_1$ and $t_2$ and returns $t_1 \cdot t_2$ as
-the final result in the forward direction. $\ell_1 \cdot \ell_2$ operates
-similarly in the backward direction, but with $s, s_1$ and $s_2$ substituted
-for $t, t_1$ and $t_2$. The $\mathit{\swap} \; (\ell_1, \ell_2)$ lens operates
-like $\ell_1 \cdot \ell_2$, except that it swaps $t_1$ and $t_2$ before
-concatenating the two for a final result of $t_2 \cdot t_1$ in the forward
-direction. In the backward direction, $\mathit{\swap}(\ell_1, \ell_2)$ first undoes
-the \swap, then behaves like $cdot$. The $\ell_1 \; | \; \ell_2$ lens
-chooses to apply $\ell_1$ or $\ell_2$ depending on whether the source
-(resp. target) data is matched by $\ell_1$ or $\ell_2$ in the forward (resp.
-backward direction). The $\ell^*$ lens splits the string $s$ into strings $s_1,
-\ldots, s_n$, applies $\ell_1$ to each $s_i$ to get $t_i$, and then concatenates
-each of the $t_i$'s for a final result of $t_1 \cdot \ldots \cdot t_n$ in the
-forward direction. \bcp{ditto}$\ell^*$ operates similarly in the backward direction, but
-with $s, s_i$ substituted for $t, t_i$.
-
-Each bijective lens $\ell$ has a type $\ell : R \Leftrightarrow S$ where $R$ and
-$S$ are regular expressions. If $\ell : R \Leftrightarrow S$, then the source
-language of $\ell$ is $\mathcal{L}(R)$ and the target language of $\ell$ is
-$\mathcal{L}(S)$. Interestingly, with the bijective lens type system, a lens
-$\ell : R \Leftrightarrow S$ can also be considered to be of type $\ell : R'
-\Leftrightarrow S'$ provided that $R$ (resp. $S$) can be proven to be
-equivalent to $R'$ (resp. $S$) from the star-semiring axioms\bcp{people may
-  not know what these are}:
+direction, while $\mathit{id}(R)$ applies the identity function to the
+source and target  in $\mathcal{L}(R)$ in both directions. The composition lens
+$\ell_1 \; ; \; \ell_2$ applies $\ell_1$ followed by then $\ell_2$ to the
+source in the forward direction, and applies $\ell_2$ followed by $\ell_1$ to
+the target in the backward direction. The lens $\ell_1 \cdot \ell_2$ first
+splits the string $s$ into $s_1$ and $s_2$, applies $\ell_1$ and $\ell_2$ to
+$s_1$ and $s_2$ to get $t_1$ and $t_2$ respectively, then concatenates $t_1$
+and $t_2$ and returns $t_1 \cdot t_2$ as the final result in the forward
+direction. $\ell_1 \cdot \ell_2$ operates similarly in the backward direction,
+but with $s, s_1$ and $s_2$ substituted for $t, t_1$ and $t_2$. The
+$\mathit{\swap} \; (\ell_1, \ell_2)$ lens operates like $\ell_1 \cdot \ell_2$,
+except that it swaps $t_1$ and $t_2$ before concatenating the two for a final
+result of $t_2 \cdot t_1$ in the forward direction. In the backward direction,
+$\mathit{\swap}(\ell_1, \ell_2)$ first undoes the \swap, then behaves like
+$cdot$. The $\ell_1 \; | \; \ell_2$ lens chooses to apply $\ell_1$ or $\ell_2$
+depending on whether the source (resp. target) data is matched by $\ell_1$ or
+$\ell_2$ in the forward (resp. backward direction). The $\ell^*$ lens splits the
+string $s$ into strings $s_1, \ldots, s_n$, applies $\ell_1$ to each $s_i$ to
+get $t_i$, and then concatenates each of the $t_i$'s for a final result of $t_1
+\cdot \ldots \cdot t_n$ in the forward direction. The iteration lens $\ell^*$
+operates similarly in the backward direction, but with $s, s_i$ substituted for
+$t, t_i$.
+
+Each bijective string lens $\ell$ has a type $\ell : R \Leftrightarrow S$ where
+$R$ and $S$ are regular expressions. If $\ell : R \Leftrightarrow S$, then the
+source language of $\ell$ is $\mathcal{L}(R)$ and the target language of $\ell$
+is $\mathcal{L}(S)$. Interestingly, with the bijective string lens type system,
+a lens $\ell : R \Leftrightarrow S$ can also be considered to be of type $\ell
+: R' \Leftrightarrow S'$ provided that $R$ (resp. $S$) can be proven to be
+equivalent to $R'$ (resp. $S$) from the star-semiring axioms (associativity and
+commutativity of | and $\cdot$, identity of the empty string $\epsilon$,
+distibutivity of $\cdot$ over |, annihilative property of $\varnothing$ with
+respect to $\cdot$, and that $R^* = \epsilon \; | \; R \cdot R* = \epsilon \; |
+\; R^* \cdot R$ for all $R$):
 
@@
 
-The other typing rules for bijective lenses are given in
+The other typing rules for bijective string lenses are given in
 Figure~\ref{fig:lensrules}.
@@
 \end{center}
-\caption{Bijective Lens Typing Rules}
+\caption{Bijective String Lens Typing Rules}
 \label{fig:lensrules}
@@
 grammar:
-$$ q := \kw{lift}(\ell) \sep \kw{lquot}(q, \ell) \sep
-\kw{rquot}(\ell, q) \sep q_1 \cdot q_2 \sep (q_1 \sep q_2) \sep q^* \sep
-q_1 \; ; \; q_2,$$ where $\ell$ ranges over bijective lenses.
-\BCP{Is this really what we want?  E.g., isn't it reasonable for the 
-  programmer to write $\mathit{\swap}(q_1, q_2)$?}
-
-The QRE lens combinators are inspired by the Boomerang lens combinators of the
-same name. The $\kw{lift}(\ell)$ quotient lens enables a bijective lens
-$\ell$ to be considered a quotient lens, with the equivalence relation applied
-to both the source and target formats being the equality relation. The
-regular combinators of concatenation `$\cdot$', Kleene star `$*$' and
-`$|$' on QRE lenses exhibit the expected behaviour\bcp{Is the behavior really
-  ``expected''---i.e., will it be obvious to all readers?}, with the caveat
-that these 
-combinators are strongly unambiguous. The composition operator `$;$' also
-exhibits the expected behaviour, though the conditions that must be checked in
-order for the composition to be well defined are perhaps unintuitive. The
-reasons why these conditions are necessary will become more apparent when
-proving normal forms for QRE lenses.
-
-The interesting combinators are the $\kw{lquot}$ and $\kw{rquot}$
-combinators. The $\kw{lquot}(c, q)$ combinator takes a quotient lens $q$ and
-quotients the source data using $c$, assuming that the source data forms a
-complete set of representatives for the equivalence relation $\eqrel{c}$. The
-$\kw{lquot}(q, c)$ combinator does the same but on the target data.
+$$ q := \kw {id}(R) \sep  \kw{const}(r, s) \sep q_1 \cdot q_2 \sep
+\kw{swap}(q_1, q_2) \sep (q_1 \sep q_2) \sep q^* \sep q_1 \; ; \; q_2 \sep
+\kw{lquot}(q, \ell) \sep \kw{rquot}(\ell, q)$$ where $R$ ranges over
+regular expressions and $\ell$ ranges over bijective lenses.
+
+TThe QRE lens combinators are the same as the bijective string lens
+combinators but with two extra combinators where quotienting actually
+occurs: tthe $\kw{lquot}$ and $\kw{rquot}$ combinators. The $\kw{lquot}(Q, q)$
+combinator takes a quotient lens $q$ and quotients the source data using $Q$,
+assuming that the source data forms a complete set of representatives for the
+equivalence relation $\eqrel{Q}$. The $\kw{lquot}(q, Q)$ combinator does the
+same but on the target data.
 
@@
 \begin{lstlisting}
-let bib_to_end_q : (lens in bibtex $\Leftrightarrow$ endnote) = 
-  rquot (lquot(bibtex, bib_to_end), endnote)
+let bib_to_end_q : (lens in bibtex $\Leftrightarrow$ endnote) =  rquot (lquot(bibtex, bib_to_end), endnote)
 \end{lstlisting}
@@
 typing judgements. More concretely, each QRE lens $q_1$ has a type $q_1 : c
-\Leftrightarrow c_1$ where $c, c_1$ are QREs. In contrast, the approach used to
+\Leftrightarrow Q_1$ where $c, Q_1$ are QREs. In contrast, the approach used to
 type quotient lenses in Boomerang is to classify equivalences according to
@@
 
-Each QRE lens $q$ has a type $q : c_1 \Leftrightarrow c_2$ where $c_1, c_2$ are
-QREs. If $q : c_1 \Leftrightarrow c_2$, then the source format is described by
-$W(c_1)$ and the canonical set of representative for the source data is
-described by $K(c_1)$.
-Similarly, the target format is described by $W(c_2)$ and the canonical set of
-representatives for the target data is described by $K(c_2)$. The underlying
-lens of $q$ is a bijective lens $\ell : K(c_1) \Leftrightarrow K(c_2)$.
-
-The denotation $\llbracket q \rrbracket$ of a QRE lens $q : c_1 \Leftrightarrow
-c_2$ is a quotient lens $\llbracket q \rrbracket : W(c_1)/{\eqrel{c_1}}
-\Longleftrightarrow W(c_2)/{\eqrel{c_2}}$. The typing rules and denotation of
+Each QRE lens $q$ has a type $q : Q_1 \Leftrightarrow Q_2$ where $Q_1, Q_2$ are
+QREs. If $q : Q_1 \Leftrightarrow Q_2$, then the source format is described by
+$W(Q_1)$ and the canonical set of representative for the source data is
+described by $K(Q_1)$.
+Similarly, the target format is described by $W(Q_2)$ and the canonical set of
+representatives for the target data is described by $K(Q_2)$. The underlying
+lens of $q$ is a bijective lens $\ell : K(Q_1) \Leftrightarrow K(Q_2)$.
+
+The denotation $\llbracket q \rrbracket$ of a QRE lens $q : Q_1 \Leftrightarrow
+Q_2$ is a quotient lens $\llbracket q \rrbracket : W(Q_1)/{\eqrel{Q_1}}
+\Longleftrightarrow W(Q_2)/{\eqrel{Q_2}}$. The typing rules and denotation of
 QRE lenses are given in Figure~\ref{fig:qlenssemantics}.
 \begin{figure}[ht]
+TODO: Need to add some inferrence rules here.
 \centering
@@
 \begin{array}[b]{c@{\qquad}c}
-\RuleSide{q : c_2  \Leftrightarrow c_3 &
-\wf{c_1} &
-K(c_1) = W(c_2)}
-{\kw{lquot}(c_1, q): c_1 \; ; \; c_2 \Leftrightarrow c_3}{\mathit(Lquot)} &
+\RuleSide{q : Q_2  \Leftrightarrow Q_3 &
+\wf{Q_1} &
+K(Q_1) = W(Q_2)}
+{\kw{lquot}(Q_1, q): Q_1 \; ; \; Q_2 \Leftrightarrow Q_3}{\mathit(Lquot)} &
 \begin{array}[b]{rcl}
-\llbracket \kw{lquot}(c_1, q) \rrbracket.\get  &=& \llbracket q
-\rrbracket.\get \circ \canonizer(c_1)\\
-\llbracket \kw{lquot}(c_1, q) \rrbracket.\lput &=& \llbracket q
+\llbracket \kw{lquot}(Q_1, q) \rrbracket.\get  &=& \llbracket q
+\rrbracket.\get \circ \canonizer(Q_1)\\
+\llbracket \kw{lquot}(Q_1, q) \rrbracket.\lput &=& \llbracket q
 \rrbracket.\lput
@@
 \begin{array}[b]{c@{\qquad}c}
-\RuleSide{q : c_1 \Leftrightarrow c_3 & \wf{c_3} & W(c_3) = K(c_2)}
-{\kw{rquot}(q, c_1):c_1 \Leftrightarrow c_2 \; ; \; c_3}{\mathit{(Rquot)}} &
+\RuleSide{q : Q_1 \Leftrightarrow Q_3 & \wf{Q_3} & W(Q_3) = K(Q_2)}
+{\kw{rquot}(q, Q_1):Q_1 \Leftrightarrow Q_2 \; ; \; Q_3}{\mathit{(Rquot)}} &
 \begin{array}[b]{rcl}
-\llbracket \kw{rquot}(q, c_1) \rrbracket.\get  &=& \llbracket q
+\llbracket \kw{rquot}(q, Q_1) \rrbracket.\get  &=& \llbracket q
 \rrbracket.\get\\
-\llbracket \kw{rquot}(q, c_1) \rrbracket.\lput &=& \llbracket q
-\rrbracket.\lput \circ \canonizer(c_3)
+\llbracket \kw{rquot}(q, Q_1) \rrbracket.\lput &=& \llbracket q
+\rrbracket.\lput \circ \canonizer(Q_3)
 \end{array}
@@
 \begin{array}[b]{l@{\qquad}c}
-\RuleSide{q : c_1 \Leftrightarrow c_2 &
-W(c_1)^{*!},W(c_2)^{*!} & K(c_1)^{*!}, K(c_2)^{*!}}
-{{q}^* : {c_1}^* \Leftrightarrow {c_2}^*}{\mathit{(Star)}} &
+\RuleSide{q : Q_1 \Leftrightarrow Q_2 &
+W(Q_1)^{*!},W(Q_2)^{*!} & K(Q_1)^{*!}, K(Q_2)^{*!}}
+{{q}^* : {Q_1}^* \Leftrightarrow {Q_2}^*}{\mathit{(Star)}} &
 \begin{array}[b]{rcl}
@@
 \begin{array}[b]{l@{\qquad}c}
-\RuleSide{\substack{q_1 : c_1 \Leftrightarrow d_1 \\ q_2 : c_2 \Leftrightarrow
-d_2} & {\substack{W(c_1) \cdot^! W(c_2)\\ K(c_1) \cdot^! K(c_2)}}
+\RuleSide{\substack{q_1 : Q_1 \Leftrightarrow d_1 \\ q_2 : Q_2 \Leftrightarrow
+d_2} & {\substack{W(Q_1) \cdot^! W(Q_2)\\ K(Q_1) \cdot^! K(Q_2)}}
 & {\substack{W(d_1) \cdot^! W(d_2)\\ K(d_1) \cdot^! K(d_2)}}}
-{q_1 \cdot q_2: c_1 \cdot c_2 \Leftrightarrow d_1 \cdot d_2}{\mathit{(Concat)}}
+{q_1 \cdot q_2: Q_1 \cdot Q_2 \Leftrightarrow d_1 \cdot d_2}{\mathit{(Concat)}}
 &
@@
 \begin{array}[b]{l@{\qquad}c}
-\RuleSide{\substack{q_1 : c_1 \Leftrightarrow d_1 \\ q_2 : c_2 \Leftrightarrow
-d_2} & \substack{\mathcal{L}(W(c_1)) \cap \mathcal{L}(W(c_2)) = \varnothing
+\RuleSide{\substack{q_1 : Q_1 \Leftrightarrow d_1 \\ q_2 : Q_2 \Leftrightarrow
+d_2} & \substack{\mathcal{L}(W(Q_1)) \cap \mathcal{L}(W(Q_2)) = \varnothing
 \\ \mathcal{L}(W(d_1)) \cap \mathcal{L}(W(d_2)) = \varnothing }}
-{q_1 \sep q_2: (c_1 \sep c_2)\Leftrightarrow (d_1 \sep d_2)}{\mathit{(Or)}} &
+{q_1 \sep q_2: (Q_1 \sep Q_2)\Leftrightarrow (d_1 \sep d_2)}{\mathit{(Or)}} &
 \begin{array}[b]{rcl}
@@
 \begin{cases}
-\llbracket q_1 \rrbracket.\get (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
-\llbracket q_2 \rrbracket.\get (s) & \text{if } s \in \mathcal{L}(W(c_2))\\
+\llbracket q_1 \rrbracket.\get (s) & \text{if } s \in \mathcal{L}(W(Q_1))\\
+\llbracket q_2 \rrbracket.\get (s) & \text{if } s \in \mathcal{L}(W(Q_2))\\
 \end{cases}\\
@@
 \begin{prooftree}
-\AxiomC{$\substack{q_1 : c_1 \Leftrightarrow c_2 \\ q_2 : c_3 \Leftrightarrow
-c_4}$}
+\AxiomC{$\substack{q_1 : Q_1 \Leftrightarrow Q_2 \\ q_2 : Q_3 \Leftrightarrow
+Q_4}$}
 \RightLabel{(Compose)}
-\AxiomC{$\substack{\mathcal{L}(W(c_2))= \mathcal{L}(W(c_3)) \\ K(c_2) \equiv^s
-K(c_3)}$}
-\AxiomC{$\canonizer(c_2) = \canonizer(c_3)$}
-\TrinaryInfC{$q_1 \; ; \; q_2: c_1 \Leftrightarrow c_4$}
+\AxiomC{$\substack{\mathcal{L}(W(Q_2))= \mathcal{L}(W(Q_3)) \\ K(Q_2) \equiv^s
+K(Q_3)}$}
+\AxiomC{$\canonizer(Q_2) = \canonizer(Q_3)$}
+\TrinaryInfC{$q_1 \; ; \; q_2: Q_1 \Leftrightarrow Q_4$}
 \end{prooftree}
@@
 \begin{prooftree}
-\AxiomC{$\substack{q_1 : c_1 \Leftrightarrow c_2 \\ q_2 : c_3 \Leftrightarrow
-c_4}$}
-\AxiomC{$\substack{\mathcal{L}(W(c_2))= \mathcal{L}(W(c_3)) \\ K(c_2) \equiv^s
-K(c_3)}$}
-\AxiomC{$\canonizer(c_2) = \canonizer(c_3)$}
-\TrinaryInfC{$q_1 \; ; \; q_2: c_1 \Leftrightarrow c_4$}
+\AxiomC{$\substack{q_1 : Q_1 \Leftrightarrow Q_2 \\ q_2 : Q_3 \Leftrightarrow
+Q_4}$}
+\AxiomC{$\substack{\mathcal{L}(W(Q_2))= \mathcal{L}(W(Q_3)) \\ K(Q_2) \equiv^s
+K(Q_3)}$}
+\AxiomC{$\canonizer(Q_2) = \canonizer(Q_3)$}
+\TrinaryInfC{$q_1 \; ; \; q_2: Q_1 \Leftrightarrow Q_4$}
 \end{prooftree}
@@
 This derivation rule essentially says that the composition $q_1 \; ; \; q_2$ is
-well defined if and only if the intermediary QREs $c_3$ and $c_4$ define the
+well defined if and only if the intermediary QREs $Q_3$ and $Q_4$ define the
 same equivalence relation on regular languages that are equivalent modulo the
-star-semiring axioms. The condition $\mathcal{L}(W(c_2)) = \mathcal{L}(W(c_3))$
+star-semiring axioms. The condition $\mathcal{L}(W(Q_2)) = \mathcal{L}(W(Q_3))$
 says that the intermediary language is the same on both sides, while the
-condition $K(c_2) \equiv^s K(c_3)$ says that the kernel regular expressions are
+condition $K(Q_2) \equiv^s K(Q_3)$ says that the kernel regular expressions are
 equivalent modulo the star-semiring axioms. Finally, the condition
-$\canonizer(c_2) = \canonizer(c_3)$ says that $c_2$ and $c_3$ define the same
-equivalence relation on $\mathcal{L}(W(c_2))= \mathcal{L}(W(c_3))$.
-
-In general, checking that $\canonizer(c_2) = \canonizer(c_3)$ is undecidable
-because of the {\em normalize} QRE which allows uses an arbitrary
-idempotent function as the canonizer. Therefore, our actual implementation of
-the composition rule for QRE lenses follows Boomerang's approach, which is to
-define the composition on QRE lenses if and only if $c_2$ and $c_3$ are
-syntactially equal to $id(R)$ and $id(R')$ for some $R$ and $R'$ satisfying
-$\mathcal{L}(R) = \mathcal{L}(R')$.  \BCP{Here, once again, I worry that we
-  are mixing fundamental theory with pragmatic language design /
-  implementation, to the detriment of both.  At a theoretical level, this
-  restriction is nonsensical, and making it here raises questions like
-  ``Does the proof that composition can be normalized away depend on this
-  restriction?''}
-
-Foster et al based this approach on two observations: first, that most quotient
-lenses originate as lifted basic lenses, and therefore have types whose
-equivalence relations are both equality; and second, that equality is preserved
-by many of the quotient lens combinators. Foster et al also discuss a second
-possible approach, where equivalence relations are represented by rational
-functions that induce them. While this second approach is more refined than the
-first, Boomerang favours the first approach since the second appears to be too
-expensive to be useful in practice~\cite{quotientlenses}.
+$\canonizer(Q_2) = \canonizer(Q_3)$ says that $Q_2$ and $Q_3$ define the same
+equivalence relation on $\mathcal{L}(W(Q_2))= \mathcal{L}(W(Q_3))$.
 
@@
 \begin{theorem}
-If there is a derivation $q : c_1 \Leftrightarrow c_2$, then $\llbracket q
-\rrbracket : W(c_1)/{\eqrel{c_1}} \Leftrightarrow W(c_2)/{\eqrel{c_2}}$ is a
+If there is a derivation $q : Q_1 \Leftrightarrow Q_2$, then $\llbracket q
+\rrbracket : W(Q_1)/{\eqrel{Q_1}} \Leftrightarrow W(Q_2)/{\eqrel{Q_2}}$ is a
 well-defined quotient lens.
@@
 \subsection{Normal Forms of QRE Lenses}
-Recall that our approach in defining QRE lenses is to have each QRE lens $q: c_1
-\Leftrightarrow c_2$ be such that
-\begin{align*}
-\llbracket q \rrbracket.\get &= \ell \circ \canonizer(c_1)\\
-\llbracket q \rrbracket.\lput &= \ell^{-1} \circ \canonizer(c_2)
+Recall that our approach in defining QRE lenses is to have each QRE lens $q: Q_1
+\Leftrightarrow Q_2$ be such that
+\begin{align*}
+\llbracket q \rrbracket.\get &= \ell \circ \canonizer(Q_1)\\
+\llbracket q \rrbracket.\lput &= \ell^{-1} \circ \canonizer(Q_2)
 \end{align*}
@@
 \begin{theorem}\label{normal form}
-If there is a derivation $q : c_1 \Leftrightarrow c_2$, then there exists a
-bijective lens $\ell : K(c_1) \Leftrightarrow K(c_2)$ such that:
+If there is a derivation $q : Q_1 \Leftrightarrow Q_2$, then there exists a
+bijective lens $\ell : K(Q_1) \Leftrightarrow K(Q_2)$ such that:
 \begin{align*}
 \llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket\circ
-\canonizer(c_1)\\
+\canonizer(Q_1)\\
 \llbracket q \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
-\canonizer(c_2)
+\canonizer(Q_2)
 \end{align*}
@@
 By straightforward induction on the derivation
-$q_1 : c_1 \Leftrightarrow c_2$. Since bijective lenses are closed under strongly unambiguous
+$q_1 : Q_1 \Leftrightarrow Q_2$. Since bijective lenses are closed under strongly unambiguous
 regular operators, the tricky case is composition.  The composition case is
@@
 \begin{prooftree}
-\AxiomC{$\substack{q_1 : c_1 \Leftrightarrow c_2 \\ q_2 : c_3 \Leftrightarrow
-c_4}$}
-\AxiomC{$\substack{\mathcal{L}(W(c_2))= \mathcal{L}(W(c_3)) \\ K(c_2) \equiv
-K(c_3)}$}
-\AxiomC{$\canonizer(c_2) = \canonizer(c_3)$}
-\TrinaryInfC{$q_1 \; ; \; q_2: c_1 \Leftrightarrow c_4$}
+\AxiomC{$\substack{q_1 : Q_1 \Leftrightarrow Q_2 \\ q_2 : Q_3 \Leftrightarrow
+Q_4}$}
+\AxiomC{$\substack{\mathcal{L}(W(Q_2))= \mathcal{L}(W(Q_3)) \\ K(Q_2) \equiv
+K(Q_3)}$}
+\AxiomC{$\canonizer(Q_2) = \canonizer(Q_3)$}
+\TrinaryInfC{$q_1 \; ; \; q_2: Q_1 \Leftrightarrow Q_4$}
 \end{prooftree}
@@
 By the induction hypothesis, there exist bijective lenses
-$\ell_1 : K(c_1) \Leftrightarrow K(c_2)$ and $\ell_2 : K(c_3) \Leftrightarrow
-K(c_4)$ such that
+$\ell_1 : K(Q_1) \Leftrightarrow K(Q_2)$ and $\ell_2 : K(Q_3) \Leftrightarrow
+K(Q_4)$ such that
 \begin{center}
@@
 $\llbracket q_1 \rrbracket.\get = \llbracket \ell_1 \rrbracket \circ
-\canonizer(c_1)$ & $\llbracket q_2 \rrbracket.\get = \llbracket \ell_2
-\rrbracket \circ \canonizer(c_3)$ \\
+\canonizer(Q_1)$ & $\llbracket q_2 \rrbracket.\get = \llbracket \ell_2
+\rrbracket \circ \canonizer(Q_3)$ \\
 $\llbracket q_1 \rrbracket.\lput = {\llbracket \ell_1 \rrbracket}^{-1} \circ
-\canonizer(c_2)$ & $\llbracket q_2 \rrbracket.\lput = {\llbracket \ell_2
-\rrbracket}^{-1} \circ \canonizer(c_4)$
+\canonizer(Q_2)$ & $\llbracket q_2 \rrbracket.\lput = {\llbracket \ell_2
+\rrbracket}^{-1} \circ \canonizer(Q_4)$
 \end{tabular}
@@
 \llbracket q_2 \rrbracket.\get \circ \llbracket q_1 \rrbracket.\get &=
-(\llbracket \ell_2 \rrbracket \circ \canonizer(c_3)) \circ (\llbracket \ell_1
-\rrbracket \circ \canonizer(c_1))\\
-&= \llbracket \ell_2 \rrbracket \circ (\canonizer(c_3) \circ \llbracket \ell_1
-\rrbracket) \circ \canonizer(c_1)\\
+(\llbracket \ell_2 \rrbracket \circ \canonizer(Q_3)) \circ (\llbracket \ell_1
+\rrbracket \circ \canonizer(Q_1))\\
+&= \llbracket \ell_2 \rrbracket \circ (\canonizer(Q_3) \circ \llbracket \ell_1
+\rrbracket) \circ \canonizer(Q_1)\\
 &= (\llbracket \ell_2 \rrbracket \circ \llbracket \ell_1 \rrbracket) \circ
-\canonizer(c_1)\\
-&= \llbracket \ell_1 \; ; \; \ell_2 \rrbracket \circ \canonizer(c_1)
+\canonizer(Q_1)\\
+&= \llbracket \ell_1 \; ; \; \ell_2 \rrbracket \circ \canonizer(Q_1)
 \end{align*}
 We are permitted to claim the third step from the second since
-$\canonizer(c_3)$ is the identity function on $K(c_2)$ which is
-syntactically equal to $K(c_3)$ by assumption. A similar argument shows that
+$\canonizer(Q_3)$ is the identity function on $K(Q_2)$ which is
+syntactically equal to $K(Q_3)$ by assumption. A similar argument shows that
 $$\llbracket q_1 \rrbracket.\lput \circ \llbracket q_2 \rrbracket.\lput =
 \llbracket \ell_1 \; ; \; \ell_2 \rrbracket^{-1} \circ
-\canonizer(c_4)$$
+\canonizer(Q_4)$$
 
@@
 
-In our setting, we want to synthesize a quotient lens $q: c_1 \Leftrightarrow
-c_2$ from the QREs $c_1$ and $c_2$ and a set of example input-output pairs
+In our setting, we want to synthesize a quotient lens $q: Q_1 \Leftrightarrow
+Q_2$ from the QREs $Q_1$ and $Q_2$ and a set of example input-output pairs
 $\{(x_1, y_1),\allowbreak \ldots,\allowbreak (x_n, y_n)\}$ where the $x_i$'s are
-in $W(c_1)$ and the $y_i$'s are in $W(c_2)$. We furthermore wish $q$ to map the
+in $W(Q_1)$ and the $y_i$'s are in $W(Q_2)$. We furthermore wish $q$ to map the
 equivalence class of $x_i$ to the equivalence class of $y_i$ and vice versa:
 \begin{align*}
-q.\get(x_i) &\equiv_{c_2} y_i \text{, and}\\
-q.\lput(y_i) &\equiv_{c_1} x_i
+q.\get(x_i) &\equiv_{Q_2} y_i \text{, and}\\
+q.\lput(y_i) &\equiv_{Q_1} x_i
 \end{align*}
 Our approach to synthesizing QRE lenses is guided by Theorem~\ref{normal form},
-which says that, if there is a derivation $q : c_1 \Leftrightarrow c_2$ of a
-QRE lens, then there exists a bijective lens $\ell : K(c_1) \Leftrightarrow
-K(c_2)$ such that
-\begin{align*}
-\llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket\circ \canonizer(c_1)\\
+which says that, if there is a derivation $q : Q_1 \Leftrightarrow Q_2$ of a
+QRE lens, then there exists a bijective lens $\ell : K(Q_1) \Leftrightarrow
+K(Q_2)$ such that
+\begin{align*}
+\llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket\circ \canonizer(Q_1)\\
 \llbracket q \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
-\canonizer(c_2).
-\end{align*}
-\bcp{This math display and the previous one are punctuated perfedtly: In the
+\canonizer(Q_2).
+\end{align*}
+\bcp{This math display and the previous one are punctuated perfectly: In the
 first, the colon replaces a period at the end of the sentence (so no need
@@
 For the\bcp{why ``the''?  Are we talking about some set of QRE examples
-  already?} QRE examples, the $x_i$'s are in $W(c_1)$ and the $y_i$'s are 
-in $W(c_2)$, so we can construct ${x'_i} = \canonizer(c_1)(x_i)$ in $K(c_1)$
-and ${y'_i} = \canonizer(c_2)(y_i)$ is in $K(c_2)$. To synthesize the
-desired quotient lens $q: c_1 \Leftrightarrow c_2$ that is consistent with the
+  already?} QRE examples, the $x_i$'s are in $W(Q_1)$ and the $y_i$'s are 
+in $W(Q_2)$, so we can construct ${x'_i} = \canonizer(Q_1)(x_i)$ in $K(Q_1)$
+and ${y'_i} = \canonizer(Q_2)(y_i)$ is in $K(Q_2)$. To synthesize the
+desired quotient lens $q: Q_1 \Leftrightarrow Q_2$ that is consistent with the
 input-output examples $\{(x_1, y_1), \ldots, (x_n, y_n)\}$ it suffices to
-synthesize a bijective lens $\ell : K(c_1) \Leftrightarrow K(c_2)$ that is
+synthesize a bijective lens $\ell : K(Q_1) \Leftrightarrow K(Q_2)$ that is
 consistent with the canonized examples $\{({x'_1}, {y'_1}), \ldots, ({x'_n},
@@
   \item
-  Compute $K(c_1)$, $K(c_2)$,
+  Compute $K(Q_1)$, $K(Q_2)$,
   \item
   Compute $\{({x'_1}, {y'_1}), \ldots, ({x'_n}, {y'_n})\}$ where ${x'_i} =
-  \canonizer(c_1)(x_i)$ and ${y'_i} = \canonizer(c_2)(y_i)$,
+  \canonizer(Q_1)(x_i)$ and ${y'_i} = \canonizer(Q_2)(y_i)$,
 \item Synthesize, using Optician's algorithm, a bijective lens
-  $\ell : K(c_1) \Leftrightarrow K(c_2)$ that is consistent with the examples
+  $\ell : K(Q_1) \Leftrightarrow K(Q_2)$ that is consistent with the examples
   $\{({x'_1}, {y'_1}),\allowbreak \ldots,\allowbreak ({x'_n}, {y'_n})\}$
   \item
-  Return $q = \kw{rquot}(\kw{lquot}(c_1, \ell), c_2) : c_1 \Leftrightarrow
-  c_2$ \bcp{punctuate this enumeration as if it were running text: ``, and''
+  Return $q = \kw{rquot}(\kw{lquot}(Q_1, \ell), Q_2) : Q_1 \Leftrightarrow
+  Q_2$ \bcp{punctuate this enumeration as if it were running text: ``, and''
   at the end of (3) and period here.  And lowercase letters at the
@@
 We have implemented QREs and the quotient lens synthesis algorithm described
-above as an extension to the Boomerang interpreter~\cite{boomerang,quotientlenses}. Thanks to this integration, QREs and quotient lens
-synthesis tasks can now be expressed alongside existing Boomerang
-terms\bcp{what's the difference between a task and a term (and how do each
-  of them relate to ``lenses'' just below)?}, and we
-can combine the two kinds of lenses seamlessly in the same application.
-Intuition suggests that the increased expressiveness of QREs should make many
-lens programming tasks easier, but we would like more quantitative understanding
-of that claim.  We thus conducted several experiments that aim to provide
-answers to the following questions:
+above as an extension to the Boomerang
+interpreter~\cite{boomerang,quotientlenses}. We have added QREs and QRE lens
+synthesis tasks to Boomerang. Furthermore, these synthesis tasks evaluate to
+Boomerang lens values, so Boomerang treats synthesized lenses the same as
+hand-written ones. Intuition suggests that the increased expressiveness
+of QREs should make many lens programming tasks easier, but we would like more
+quantitative understanding of that claim. We thus conducted several experiments
+that aim to provide answers to the following questions:
 
 \begin{enumerate}
-  \item Does synthesizing quotient lenses from QREs permit an easier development
-  process than existing approaches?  \bcp{Needlessly vague: What existing approaches?  Writing
-  them by hand?}
+  \item Does synthesizing QRE lenses from QREs permit an easier development
+  process than manually writing canonizers, and synthesizing the lenses between
+  their canonized forms?  Does synthesizing QRE lenses from QREs permit an
+  easier development process than writing the lenses by hand?
 
@@
     this really an interesting empirical question?  Seems obvious, from the
-    arguments we made in Section 2 and the intro.}
+    arguments we made in Section 2 and the intro.}\afm{I'm happy to remove it}
 
   \item Does our tool synthesize lenses fast enough to be used as part of a
-  standard development process.  \bcp{Should there be some question marks
-    around here.}
+  standard development process?
 \end{enumerate}
@@
     ones?  Are they significantly different (more complex, exercising
-    different features of the synthesis algorithm or QRE syntax, ...)?}
+    different features of the synthesis algorithm or QRE syntax,
+    ...)?}\afm{Solomon, could you fill this out?}
+    \sam{Well these data formats were the ones I was playing with before we got
+    the Augeus and flashfill examples. I chose them simply because they had
+    formats which could be expressed using regular expressions. They basically
+    use the same QREs as the flashfill and augeus examples, mostly the project
+    QRE. The fields in these files are occur in the same order in each record;
+    we tried using the perm combinator anyways but the synthesis algorithm took
+    too long. I do not think that they are significantly different from the
+    flashfill and augues examples.}
 \end{enumerate}
@@
   \item[\QRESize{}] the number of AST nodes in the QRE
-  specifications (including examples)\saz{CHECK THIS}
+  specifications (including examples)\saz{CHECK THIS}\afm{no but i'll add it in}
   \item[\CanonizerAndSpecSize{}]  the number of
@@
   by number of AST nodes), with a bar above each one containing three
-  regions, colored differently, for LS, CS-LS, and QC-CS...} This suggests
+  regions, colored differently, for LS, CS-LS, and QC-CS...}\afm{will do} This suggests
 that introducing QREs saves programmers 
@@
 transformations. Synthesizing from QREs loosens that restriction, making it
-possible to synthesize data which is bijective when put in canonical
-form\bcp{No: We do not synthesize data, nor is the data itself bijective.}. To
-determine the increase of expressivity, we analyzed:
-\begin{enumerate}
-  \item How many data formats in Optician's benchmark suite required modifications
-  to be bijective.
-  \item How many data formats in Optician's benchmark suite required modifications
-  to be bijective when put in canonical form.  \bcp{I don't understand the
-    difference between these.}
-\end{enumerate}
-
-We found that 10 of Optician's 39 benchmarks required modifications to be
-bijective. Of these We found that 4 of Optician's 39 benchmarks required
-modifications to be in bijective correspondence when in canonical form.
-Furthermore, we found that all eight of the data.gov examples required no
-modifications to be in bijective correspondence when in canonical
-form.\BCP{Is this surprising?  We could have chosen the examples so that
-  this is true.} While we
-did not put the data.gov examples into a form suitable for pre-QRE
-Optician\bcp{What does that mean?}, we
-found that only four of the eight data.gov examples did not require
-canonization, so making them suitable for pre-QRE Optician would have required
-changes to of four of the data.gov formats. Here again, we see that QREs add
-significant expressive power.
+possible to synthesize data transformations involving bijections with
+canonization functions. Because Optician could previously only synthesize
+bijective lenses, some of the data formats in the example suite were modified to
+include additional fields containing information should should be canonized from
+the other format. In total, 12 of Optician's original benchmarks and 3 of the
+data.gov benchmarks would require such additional fields when input to Optician
+without QREs.
 
@@
 \end{subfigure}
-\caption{Runtimes of Optician and our system.
-In (a), we run Optician and our system on Optician's benchmarks.  We find that
-there is only a negligible difference incurred by Optometrist's overhead.
-In (b), we run Optometrist on its benchmark suite.  We find that it is able
-to synthesize all quotient lenses in under 20 seconds, and typically
-finished in under 5 seconds. \bcp{The left-hand graph, especially, is hard
-  to make anything out of, because the curves are squished against the
-  axes and have almost identical shapes.  Not sure what to do about this.
-  Also, I'm finding the terms 
-  ``Optician'' and ``Optometrist'' really hard to keep straight---in part
-  because, in English, they are almost synonyms.}}
+\caption{Runtimes of Optician and our system. In (a), we run Optician and our
+  system on Optician's benchmarks. We find that there is only a negligible
+  performance overhead incurred by using QREs. In (b), we run Optician with QREs
+  on its benchmark suite. We find that it is able to synthesize all quotient
+  lenses in under 20 seconds, and typically finished in under 5 seconds.
+  \bcp{The left-hand graph, especially, is hard to make anything out of, because
+    the curves are squished against the axes and have almost identical shapes.
+    }\afm{will fix}}
 \label{fig:times}
@@
     graphj repeats some information from the one to the left, obscuring the
-    new information.}
+    new information.}\afm{will do}
 \end{itemize}
@@
 Also notable is one of the data.gov examples, which converts demographic
-statistics by zip code from xml form into rdf form.\bcp{It's a bit hidden in the figure...} To validate this performance
-was not due to our overhead, we modified this format to be bijective, so it
+statistics by zip code from xml form into rdf form. Synthesizing this
+transformation certainly took the longest, nearly fifteen seconds, where the
+next slowest took hardly over five. To validate this performance was not due to
+our overhead, we modified these format to be bijective, so the synthesis task
 could be input into Optician. Optician took over 15 seconds synthesize this
 modified lens. We believe this conversion took longer because the input formats
-were exceptionally large, requiring conversions between over 40 fields.
-
+were exceptionally large, requiring conversions between over 40 fields, not due
+to an overhead from synthesizing from QREs.
 \section{Related Work}
@@
   belong after the references.}
-%
-We give a full proof of \cref{normal form}: If there is a derivation $q : c_1
-\Leftrightarrow c_2$, then there exists a bijective lens $\ell : K(c_1)
-\Leftrightarrow K(c_2)$ such that:
+\sam{Also need to add a few cases here.}
+We give a full proof of \cref{normal form}: If there is a derivation $q : Q_1
+\Leftrightarrow Q_2$, then there exists a bijective lens $\ell : K(Q_1)
+\Leftrightarrow K(Q_2)$ such that:
 \begin{align*}
 \llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket\circ
-\canonizer(c_1)\\
+\canonizer(Q_1)\\
 \llbracket q \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
-\canonizer(c_2)
+\canonizer(Q_2)
 \end{align*}
 \begin{proof}
-We proceed by induction over the derivation $q : c_1 \Leftrightarrow c_2$.
+We proceed by induction over the derivation $q : Q_1 \Leftrightarrow Q_2$.
 \begin{enumerate}
@@
 \item
-$\kw{lquot}(c_1, q): c_1 \; ; \; c_2 \Leftrightarrow c_3$ where $q : c_2
-\Leftrightarrow c_3$, $c_1$ is well formed and $K(c_1) = W(c_2)$. Then
-\begin{align*}
-\llbracket \kw{lquot}(c_1, q) \rrbracket.\get  &= \llbracket q
-\rrbracket.\get \circ \canonizer(c_1)\\
-\llbracket \kw{lquot}(c_1, q) \rrbracket.\lput &= \llbracket q \rrbracket.\lput
+$\kw{lquot}(Q_1, q): Q_1 \; ; \; Q_2 \Leftrightarrow Q_3$ where $q : Q_2
+\Leftrightarrow Q_3$, $Q_1$ is well formed and $K(Q_1) = W(Q_2)$. Then
+\begin{align*}
+\llbracket \kw{lquot}(Q_1, q) \rrbracket.\get  &= \llbracket q
+\rrbracket.\get \circ \canonizer(Q_1)\\
+\llbracket \kw{lquot}(Q_1, q) \rrbracket.\lput &= \llbracket q \rrbracket.\lput
 \end{align*}
 By the induction hypothesis, there exists a bijective lens $\ell :
-K(c_2) \Leftrightarrow K(c_3)$ such that
+K(Q_2) \Leftrightarrow K(Q_3)$ such that
 \begin{align*}
 \llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket \circ
-\canonizer(c_2)\\
+\canonizer(Q_2)\\
 \llbracket q \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
-\canonizer(c_3)
+\canonizer(Q_3)
 \end{align*}
@@
 \begin{align*}
-\llbracket \kw{lquot}(c_1, q)\rrbracket.\get  &= (\llbracket \ell \rrbracket \circ
-\canonizer(c_2)) \circ \canonizer(c_1) = \llbracket \ell \rrbracket \circ
-(\canonizer(c_1 \; ; \; c_2))\\
-\llbracket \kw{lquot}(c_1, q) \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
-\canonizer(c_3)
+\llbracket \kw{lquot}(Q_1, q)\rrbracket.\get  &= (\llbracket \ell \rrbracket \circ
+\canonizer(Q_2)) \circ \canonizer(Q_1) = \llbracket \ell \rrbracket \circ
+(\canonizer(Q_1 \; ; \; Q_2))\\
+\llbracket \kw{lquot}(Q_1, q) \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
+\canonizer(Q_3)
 \end{align*}
@@
 \item
-$\kw{rquot}(q, c_3):c_1 \Leftrightarrow c_2 \; ; \; c_3$ where $q : c_1
-\Leftrightarrow c_2$, $c_3$ is well formed and $K(c_3) = W(c_2)$. Proceed as in
+$\kw{rquot}(q, Q_3):Q_1 \Leftrightarrow Q_2 \; ; \; Q_3$ where $q : Q_1
+\Leftrightarrow Q_2$, $Q_3$ is well formed and $K(Q_3) = W(Q_2)$. Proceed as in
 the previous case.
 \item
-$q_1 \; ; \; q_2: c \Leftrightarrow c_2$ where $q_1 : c \Leftrightarrow c_1$ and
-$q_2 : c_1 \Leftrightarrow c_2$. Then
+$q_1 \; ; \; q_2: c \Leftrightarrow Q_2$ where $q_1 : c \Leftrightarrow Q_1$ and
+$q_2 : Q_1 \Leftrightarrow Q_2$. Then
 \begin{align*}
@@
 $\ell_1 :
-K(c) \Leftrightarrow K(c_1)$ and $\ell_2 : K(c_1) \Leftrightarrow K(c_2)$ such
+K(c) \Leftrightarrow K(Q_1)$ and $\ell_2 : K(Q_1) \Leftrightarrow K(Q_2)$ such
 that
@@
 \llbracket q_1 \rrbracket.\lput &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
-\canonizer(c_1)
+\canonizer(Q_1)
 \end{align*}
@@
 \llbracket q_2 \rrbracket.\get &= \llbracket \ell_2 \rrbracket \circ
-\canonizer(c_1)\\
+\canonizer(Q_1)\\
 \llbracket q_2 \rrbracket.\lput &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
-\canonizer(c_2)
+\canonizer(Q_2)
 \end{align*}
@@
 \llbracket q_2 \rrbracket.\get \circ \llbracket q_1 \rrbracket.\get &=
-(\llbracket \ell_2 \rrbracket \circ \canonizer(c_1)) \circ (\llbracket \ell_1
+(\llbracket \ell_2 \rrbracket \circ \canonizer(Q_1)) \circ (\llbracket \ell_1
 \rrbracket \circ \canonizer(c))\\
-&= \llbracket \ell_2 \rrbracket \circ (\canonizer(c_1) \circ \llbracket \ell_1
+&= \llbracket \ell_2 \rrbracket \circ (\canonizer(Q_1) \circ \llbracket \ell_1
 \rrbracket) \circ \canonizer(c)\\
@@
 \item
-$q^* : {c_1}^* \Leftrightarrow {c_2}^*$ where $q : c_1 \Leftrightarrow c_2$,
-$W(c_1)^{*!}$ and $W(c_2)^{*!}$ and $K(c_1)^{*!}$ and $K(c_2)^{*!}$. Then
+$q^* : {Q_1}^* \Leftrightarrow {Q_2}^*$ where $q : Q_1 \Leftrightarrow Q_2$,
+$W(Q_1)^{*!}$ and $W(Q_2)^{*!}$ and $K(Q_1)^{*!}$ and $K(Q_2)^{*!}$. Then
 \begin{align*}
@@
 \end{align*}
-By the induction hypothesis there exists a bijective lens $\ell : K(c_1)
-\Leftrightarrow K(c_2)$ such that
+By the induction hypothesis there exists a bijective lens $\ell : K(Q_1)
+\Leftrightarrow K(Q_2)$ such that
 that
@@
 \llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket \circ
-\canonizer(c_1)\\
+\canonizer(Q_1)\\
 \llbracket q \rrbracket.\lput &= {\llbracket \ell \rrbracket}^{-1} \circ
-\canonizer(c_2)
+\canonizer(Q_2)
 \end{align*}
@@
 \llbracket q^* \rrbracket.\get &= (\llbracket \ell \rrbracket \circ
-\canonizer(c_1))^* = \llbracket \ell \rrbracket^* \circ
-\canonizer(c_1)^* = \llbracket \ell^* \rrbracket \circ
-\canonizer(c_1^*)\\
+\canonizer(Q_1))^* = \llbracket \ell \rrbracket^* \circ
+\canonizer(Q_1)^* = \llbracket \ell^* \rrbracket \circ
+\canonizer(Q_1^*)\\
 \llbracket q^* \rrbracket.\lput &= (\llbracket \ell \rrbracket^{-1} \circ
-\canonizer(c_2))^* = (\llbracket \ell \rrbracket^{-1})^* \circ
-\canonizer(c_2)^* = \llbracket \ell^* \rrbracket^{-1} \circ
-\canonizer(c_2^*)\\
+\canonizer(Q_2))^* = (\llbracket \ell \rrbracket^{-1})^* \circ
+\canonizer(Q_2)^* = \llbracket \ell^* \rrbracket^{-1} \circ
+\canonizer(Q_2^*)\\
 \end{align*}
 \item
-$q_1 \cdot q_2: c_1 \cdot c_2 \Leftrightarrow d_1 \cdot d_2$, where $q_1 : c_1
-\Leftrightarrow d_1 $,  $q_2 : c_2 \Leftrightarrow d_2$, $W(c_1)
-\cdot^! W(c_2)$, $K(c_1) \cdot^! K(c_2)$, $W(d_1) \cdot^! W(d_2)$ and $
+$q_1 \cdot q_2: Q_1 \cdot Q_2 \Leftrightarrow d_1 \cdot d_2$, where $q_1 : Q_1
+\Leftrightarrow d_1 $,  $q_2 : Q_2 \Leftrightarrow d_2$, $W(Q_1)
+\cdot^! W(Q_2)$, $K(Q_1) \cdot^! K(Q_2)$, $W(d_1) \cdot^! W(d_2)$ and $
 K(d_1) \cdot^! K(d_2)$. Then
@@
 \end{align*}
-By the induction hypothesis, there exist bijective lenses $\ell_1 : K(c_1)
-\Leftrightarrow K(d_1)$ and $\ell_2 : K(c_2) \Leftrightarrow K(d_2)$ such that
+By the induction hypothesis, there exist bijective lenses $\ell_1 : K(Q_1)
+\Leftrightarrow K(d_1)$ and $\ell_2 : K(Q_2) \Leftrightarrow K(d_2)$ such that
 \begin{align*}
 \llbracket q_1 \rrbracket.\get &= \llbracket \ell_1 \rrbracket \circ
-\canonizer(c_1)\\
+\canonizer(Q_1)\\
 \llbracket q_1 \rrbracket.\lput &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
@@
 \llbracket q_2 \rrbracket.\get &= \llbracket \ell_2 \rrbracket \circ
-\canonizer(c_2)\\
+\canonizer(Q_2)\\
 \llbracket q_2 \rrbracket.\lput &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
@@
 \llbracket q_1 \rrbracket.\get &= (\llbracket \ell_1 \rrbracket \circ
-\canonizer(c_1)) \cdot  (\llbracket \ell_2 \rrbracket \circ
-\canonizer(c_2))\\
+\canonizer(Q_1)) \cdot  (\llbracket \ell_2 \rrbracket \circ
+\canonizer(Q_2))\\
 &= (\llbracket \ell_1 \rrbracket \cdot \llbracket \ell_2
-\rrbracket) \circ (\canonizer(c_1) \cdot \canonizer(c_2))\\
-&= \llbracket \ell_1 \cdot  \ell_2 \rrbracket \circ \canonizer(c_1 \cdot c_2)
+\rrbracket) \circ (\canonizer(Q_1) \cdot \canonizer(Q_2))\\
+&= \llbracket \ell_1 \cdot  \ell_2 \rrbracket \circ \canonizer(Q_1 \cdot Q_2)
 \end{align*}
@@
 \item
-$q_1 = q_1 \sep q_2$ where $q_1 : c_1 \Leftrightarrow d_1 $, $q_2 : c_2
-\Leftrightarrow d_2$, $\mathcal{L}(W(c_1)) \cap \mathcal{L}(W(c_2)) =
+$q_1 = q_1 \sep q_2$ where $q_1 : Q_1 \Leftrightarrow d_1 $, $q_2 : Q_2
+\Leftrightarrow d_2$, $\mathcal{L}(W(Q_1)) \cap \mathcal{L}(W(Q_2)) =
 \varnothing$ and $\mathcal{L}(W(d_1)) \cap \mathcal{L}(W(d_2)) = \varnothing$.
@@
 \begin{cases}
-\llbracket q_1 \rrbracket.\get (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
-\llbracket q_2 \rrbracket.\get (s) & \text{if } s \in \mathcal{L}(W(c_2))\\
+\llbracket q_1 \rrbracket.\get (s) & \text{if } s \in \mathcal{L}(W(Q_1))\\
+\llbracket q_2 \rrbracket.\get (s) & \text{if } s \in \mathcal{L}(W(Q_2))\\
 \end{cases}$$
@@
 $$
-By the induction hypothesis, there exist bijective lenses $\ell_1 : K(c_1)
-\Leftrightarrow K(d_1)$ and $\ell_2 : K(c_2) \Leftrightarrow K(d_2)$ such that
+By the induction hypothesis, there exist bijective lenses $\ell_1 : K(Q_1)
+\Leftrightarrow K(d_1)$ and $\ell_2 : K(Q_2) \Leftrightarrow K(d_2)$ such that
 \begin{align*}
 \llbracket q_1 \rrbracket.\get &= \llbracket \ell_1 \rrbracket \circ
-\canonizer(c_1)\\
+\canonizer(Q_1)\\
 \llbracket q_1 \rrbracket.\lput &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
@@
 \llbracket \ell_1 \rrbracket \circ
-\canonizer(c_1) (s) & \text{if } s \in \mathcal{L}(W(c_1))\\
+\canonizer(Q_1) (s) & \text{if } s \in \mathcal{L}(W(Q_1))\\
 \llbracket \ell_2 \rrbracket \circ
-\canonizer(c_2) (s) & \text{if } s \in \mathcal{L}(W(c_2)),\\
+\canonizer(Q_2) (s) & \text{if } s \in \mathcal{L}(W(Q_2)),\\
 \end{cases}$$
 so $\llbracket q_1 \sep q_2 \rrbracket.\get = \llbracket \ell_1 \sep
-\ell_2 \rrbracket \circ \canonizer(c_1 \sep c_2)$. A similar argument shows
+\ell_2 \rrbracket \circ \canonizer(Q_1 \sep Q_2)$. A similar argument shows
 that $\llbracket q_1 \sep q_2 \rrbracket.\lput = \llbracket \ell_1 \sep

### 2018-03-10 papers/QuotientLenses/main.tex
Merged Dave's work back in

@@
 \begin{abstract}
-There currently exist many bidirectional programming languages, some of which
-allow a programmer to define a transformation that respects an equivalence
-relation. For example, a language with this feature may allow a programmer to
-define a transformation which ignores the order of fields in an XML file, so
-that two files that have the same fields with the same data but with the fields
-occuring in a different order are considered the same by the
-transformation. This paper simplifies the task of programming such {\em
-quotient lenses} in two main ways. First, we introduce {\em
-quotient regular expressions} or QREs which are regular expressions
-augmented with extra syntax that makes it easier to specify equivalence
-relations on regular langauges than currently existing tools. Secondly, we
-introduce {\em QRE lenses}, which are lenses that map between QREs, and show
-how to {\em synthesize} QRE lenses from a pair of QREs and a set of example
-input-output pairs. We also describe an implementation of QREs and QRE lens
-synthesis in the bidirectional programming language Boomerang, and demonstrate
-the utility of our implementation by using it to synthesize QRE lenses between
-various real-world data formats used by the {\tt data.gov} database.
+{\em Quotient lenses} are bidirectional transformations that respect
+specified equivalence relations with the goal of 
+appropriately handling inessential details in data formats. 
+For example, a programmer could use a quotient lens to define 
+a transformation that ignores the order of fields in XML data, so
+that two XML files with the same fields but in different orders would be
+considered the same. 
+This paper simplifies the task of programming quotient lenses in three
+steps.
+First, we introduce {\em quotient regular expressions} (QREs) to
+make it easy to express which portions of a data format are
+inessential.  A QRE is an augmented regular expression that
+simultaneously defines a regular expression and an equivalence
+relation over matching strings.  Variations within the same
+equivalence class are deemed inessential. 
+Next, we introduce {\em QRE lenses}, which are lenses that map
+between QREs.
+Our key technical result is a proof that every QRE lens can be
+transformed into a QRE lens that has the form $c_1 \cdot \ell \cdot
+c_2$, where $c_1$ and $c_2$ are {\em canonizers} 
+and $\ell$ is a bijective transformation.  Intuitively,
+the canonizers normalize the data  by picking a
+canonical representative of the relevant equivalence relations. 
+This theorem means we can push all the normalization to the edges of a
+bidirectional transformation, leaving in the middle a pure bijective
+transformation.  
+Finally,
+we levarage this insight to {\em synthesize} QRE lenses from a pair
+of QREs and example input-output pairs by reusing earlier work on
+synthesizing bijective lenses.
+We have implemented QREs and QRE lens synthesis 
+in the bidirectional programming language Boomerang.
+We evaluate the effectiveness of our approach by using it to
+synthesize QRE lenses between various real-world data formats in
+the {\tt data.gov} database.
+
 \end{abstract}
@@
 \section{Introduction}
-Oftentimes, the same data is stored in a variety different formats. For
-example, bibliography data may be stored in \bibtex{} and
-EndNote formats; spreadsheet data can be represented as a CSV file
-or a TSV file; web APIs can return data as a JSON or XML object. In such cases,
-it is necessary to synchronize the two or more copies of data so that they
-store the same information.
-
-One way to synchronize data stored in different formats is to use \emph{lenses},
-\emph{i.e.,} programs that represent pairs of functions for translating data
-from format the source format to the target formant (the \emph{get}
+Often, the same data is stored in more than one format. For example,
+bibliographic data may be stored in \bibtex{} or EndNote formats,
+spreadsheet data can be represented as either CSV and TSV files, and web
+APIs can return data as either JSON or XML objects.  When changes are
+made to data in one format, data in the other must be updated so the
+two representations continue to store the same information.
+One way to synchronize such data is to use \emph{lenses}\ksf{include citations?}, which
+are programs that simultaneously define pairs of functions for translating
+the data 
+from the source format to the target formant (the \emph{get}
 direction) and back again (the \emph{put} direction). Lenses are usually
-derived using domain-specific bidirectional languages which guarantee that the
-lenses defined in that language satisfy a certain set of \emph{lens laws} that
-give guarantees on how the two transformations interact.
-
-Amongst the most restrictive lenses are the \emph{bijective lenses}.
-A bijective lens $\ell : S \Leftrightarrow T$ is simply a bijection from
+defined using domain-specific bidirectional languages that 
+ensure the specified lenses satisfy various \emph{lens laws} that
+guarantee how the \emph{put} and \emph{get} functions interact.
+
+\emph{Bijective lenses} are amongst the most restrictive. 
+A bijective lens $\ell : S \Leftrightarrow T$ specifies a bijection from
 a source set $S$ to a target set $T$:
@@
 \end{equation}
-Since these lenses demand that get and put functions form a bijection, then
-no information is lost when one translates from one data format to another. 
-While this strong guarantee can be useful in some contexts, it is more often
-the case that some of the information in one or both of the data formats is
-superfluous. For instance, one typically does not care to preserve the amount of
-whitespace between characters.  In other circumstances, capitalization
-or the order of fields in a record or items in a list may be unimportant. Many
-data formats have other ad hoc rules as well. In bibliographies, for example,
-recall that in \bibtex{}, an author can be specified as
-``\cd{lastname, firstname}'' or as ``\cd{firstname lastname}''---it does not
-matter which of the two formats is used.
-
-{\em Quotient lenses}~\cite{quotientlenses} loosen many of the restrictions of
-bijective string lenses by allowing a programmer to define an equivalence
-relation on the source and target formats so that the round-tripping lens laws
+Since these lenses demand that the \emph{get} and \emph{put} functions form a bijection, 
+no information is lost during translation from one data format to another. 
+While this very strong guarantee can be useful, more often
+some of the information in one or both of the data formats is
+superfluous. For instance, the number of white space characters
+between data items, the capitalization of various strings, the
+sequence of fields in record, or the order of items in a list are
+often irrelevant.  Many
+data formats have their own ad hoc rules as well. In \bibtex{}, for example,
+an author can be specified as
+``\cd{lastname, firstname}'' or as ``\cd{firstname lastname}''---
+the specific choice does not matter.
+
+{\em Quotient lenses}~\cite{quotientlenses} loosen the restrictions of
+bijective string lenses by including equivalence
+relations on the source and target formats so that the round-tripping lens laws
 hold modulo these equivalence rather than modulo equality. Quotient lenses thus
-make it possible to align many more data sources than would otherwise be
-possible, by ignoring details that a programmer views as inessential. The only
-language we are aware of that implements a general class of quotient lenses is
-{\em Boomerang}~\cite{boomerang}. While Boomerang enables a programmer to
-define a very general class of quotient lenses, these lenses to have two main
-drawbacks.
-
-First, in order to define an equivalence relation in Boomerang, a programmer
-must define a {\em canonizer}; pair of functions $f$ and $g$ where $f$ maps
-data represented in the original external format to a canonical element and $g$
-chooses which elements of the original format correspond to values in the
-canonical format. The disadvantage of this approach is that in order for a
-programmer to {\em specify} a quotient lens in Boomerang, she must define four
-different data formats: two ``outer'' formats which specify the data that is to
-be transformed by the lens, and two ``inner'' formats which specify the
-canonical formats for the source and target data.
+make it possible to synchronize many more data formats than would otherwise be
+possible by ignoring details deemed inessential.
+
+The bidirectional programming language {\em
+Boomerang}~\cite{boomerang} is the only language we know of that
+implements a general class of quotient lenses.  Unfortunately, these
+lenses to have two main drawbacks.
+
+The first drawback is that defining an equivalence
+relation in Boomerang requires specifying a {\em canonizer}, which is
+a pair of functions $f$ and $g$ where $f$
+normalizes data from the
+external format by choosing canonical representations
+and $g$ provides an
+inverse by specifying the external form of each canonical
+representation. 
+The disadvantage of this approach is that writing
+a quotient lens in Boomerang requires defining four
+different data formats: two ``external'' formats which specify the data to
+be transformed by the lens, and two ``internal'' formats which specify the
+canonical representations for the source and target data.
 
 Our solution to this shortcoming is to introduce {\em Quotient Regular
-Expressions} or QREs, which enable a programmer to simultaneously specify a
-regular expression as well as a set of canonical representatives for strings
-that match the regular expression. QREs have the added advantage that once the
-programmer has defined a well formed QRE, then she need not define a
-canonization function seperately; the canonization function can automatically be
-inferred from the QRE.
-
-The second challenge with programming lenses in Boomerang is that even after the
-programmer has specified the data as well as a canonizer that maps each member
-in the source and target sets to their respective canonical representative, she
-must still define the underlying lens that maps between the respective sets of
-canonical representatives. This can be a slow and tedious task due to the
-constraints that Boomerang imposes on its terms. For example, Boomerang
-programmers often must rearrange the order of data items by recursively using
-operators that \swap adjacent fields. Furthermore, the Boomerang type checker is
-very strict, disallowing many programs because they contain ambiguity about how
-certain data is transformed.
-
-Our solution to this problem is that we use QREs to define a class of quotient
-lenses called {\em QRE lenses}, and, we show how to synthesize such QRE lenses
-automatically. More specifically, in order to generate a lens, a programmer
-provides a pair of QREs describing the source and target formats, plus
-a (possibly empty) suite of examples that demonstrates how the
-lens should act in specific cases.  Given this information, we can exploit and
-extend recent work on algorithms for synthesis of bijective
-lenses~\cite{optician} to generate quotient lenses. The lenses generated by
+Expressions} (QREs), which enable programmers to simultaneously specify a
+regular expression and a set of canonical representatives for matching
+strings.  Given an QRE, we can automatically infer the corresponding
+canonization functions.  
+\ksf{Define comma\_name here as a really simple example? }
+
+The second drawback with writing quotient lenses in Boomerang is that
+even after specifying the four required data formats, the programmer must
+still define a lens for converting between the canonical
+representations.   Experience shows that writing such lenses can be
+tedious 
+(consider having to rearrange the order of data items by recursively
+using operators to \swap adjacent fields) 
+and that satisfying the type checker can be difficult because 
+many seemingly useful lenses contain ambiguities disallowed by the 
+Boomerang type checker~\cite{optician}.
+
+Our solution to this problem is to introduce {\em QRE lenses}, which
+are a new class of quotient lens that can be automatically
+synthesized, obviating the need to write lenses by hand.  A QRE lens
+is like a normal lens except it maps between formats described by
+QREs. We prove a normal form theorem (Theorem~\ref{normal form}) that says that every
+composition of QRE lenses can be rewritten 
+to be the composition of a source canonizer, a bijective lens, and a
+target canonizer. This normalization property means we can synthesize
+QRE lenses by extending an existing algorithm for synthesizing
+bijective lenses~\cite{optician}.
+
+Given this framework, generating a QRE lens requires only a pair of QREs to describe the
+source and target formats and a (possibly empty) suite of examples
+demonstrating the mapping.  For example, the following code
+
+\begin{lstlisting}
+let l = synth comma_name $\Leftrightarrow$ space_name using {("Lovelace, Ada", "Ada Lovelace")}
+\end{lstlisting}
+binds \codefont{l} to a synthesized QRE lens mapping between names in
+the comma-separated form describd by the QRE \codefont{comma\_name}
+and the space-separated form described by the
+QRE \codefont{space\_name}. 
+The lenses generated by
 our synthesis algorithm are guaranteed to translate data back and forth
 between source and target QREs and to satisfy the provided examples.
-Moreover, while our synthesized quotient lenses have a relatively
-restricted form (which cuts down the amount of non-determinism in
-a specification and greatly speeds up synthesis), we prove that
-this rigidity does not compromise their expressiveness.
-
-The following list summarizes the central contributions of our work.
+Our synthesized quotient lenses have a relatively
+restricted form (which greatly speeds up synthesis), but the 
+normalization theorem means 
+this restriction does not compromise expressiveness.
+
+The following list summarizes the contributions of our work.
 \begin{enumerate}
-  \item We introduce {\em Quotient Regular Expressions} (QREs),
+  \item We introduce {\em Quotient Regular Expressions} (QREs), which are
   a compact, convenient notation for simultaneously defining a
@@
   for that relation (Section~\ref{QRE}).
-  \item We introduce {\em QRE lenses} which are lenses that map data specified
-  using QREs.
-  In addition, we will descibe a particular set of QRE lens combinators and
-  prove that QRE lenses defined using these combinators can be transformed to
-  a QRE lens that us defined using just two of these combinators; this is the
-  main technical contribution of this paper.
+  \item We introduce {\em QRE lenses}, which are lenses that translate
+  between data formats specified using QREs. (Section~\ref{QRE-lenses}).
+  \item We descibe a particular set of QRE lens combinators and prove
+  that QRE lenses defined using these combinators can be normalized
+  (Theorem~\ref{normal form}), which is the main technical
+  contribution of this paper.
   (Section~\ref{QRE-lenses}).
   \item We use this result to reduce the problem of {\em synthesizing} QRE
-  lenses to the problem of synthesizing bijective string lenses which we studied
-  in~\cite{optician}(Section~\ref{synth}).
-  \item We extend the Boomerang {\em implementation} with QREs
-  and QRE lens synthesis.  We demonstrate the utility of our implementation by
-  using it to synthesize QRE lenses between data formats drawn from the
-  {\tt data.gov} database (Section~\ref{impl}).
+  lenses to the problem of synthesizing bijective string lenses
+  previously studied in~\cite{optician}(Section~\ref{synth}).
+  \item We extend Boomerang with QREs and QRE lens synthesis. 
+  We use the resulting implementation to demonstrate the utility of
+  our approach by synthesizing QRE lenses between
+  data formats drawn from the {\tt data.gov} database (Section~\ref{impl}).
 \end{enumerate}
 Sections~\ref{relwork} and~\ref{concl} discuss related and future work.
-
-\bcp{We've said that one of the main contributions of the paper is
-synthesizing lenses, but we don't actually show this happening in the
-introduction.  Should we?}
 
@@
 \begin{verbatim}
-@Book {Doe,
-Author = "John Doe",
+@Book {Lovelace,
+Author = "Ada Lovelace",
 Title = {Generic Title}
@@
 %0 Book
-%F Doe
-%A John Doe
+%F Lovelace
+%A Ada Lovelace
 %T Generic Title
@@
 fields is also likely irrelevant and there may be equivalent ways of writing
-the same name:  ``Doe, John'' vs ``John Doe.''
+the same name:  ``Lovelace, Adan'' vs ``Ada Lovelace.''
 
@@
 \begin{verbatim}
-@Book {Doe,
-Author = "John Doe",
+@Book {Lovelace,
+Author = "Ada Lovelace",
 Title = {Generic Title},
@@
 
-@Book{     Doe,
+@Book{     Lovelace,
 Title = {Generic
 Title},
-Author = "Doe, John",     }
+Author = "Lovelace, Ada",     }
 \end{verbatim}
@@
 For instance, assume that
-the \codefont{comma\_name} format describes ``Doe, John''
-and that the \codefont{space\_name} format describes ``John Doe''
+the \codefont{comma\_name} format describes ``Lovelace, Ada''
+and that the \codefont{space\_name} format describes ``Ada Lovelace''
 and \codefont{c\_to\_s} is a function from the first to the second.  In this
@@
 \begin{lstlisting}
-let l = synth comma_name $\Leftrightarrow$ space_name using {("Doe, John", "John Doe")}
+let l = synth comma_name $\Leftrightarrow$ space_name using {("Lovelace, Ada", "Ada Lovelace")}
 let c_to_s = get l
@@
 (* synthesize a lens that maps comma representation to space representation *)
-let l = synth comma_name $\Leftrightarrow$ space_name using {("Doe, John", "John Doe")}
+let l = synth comma_name $\Leftrightarrow$ space_name using {("Lovelace, Ada", "Ada Lovelace")}
 let c_to_s = get l
@@
 
-Now let $Q_1$ and $Q_2$ be QREs. Given a bijection $\ell : K(Q_1)
-\Leftrightarrow K(Q_2)$, we may define a quotient lens $q : W(Q_1)/\eqrel{Q_1}
-\Leftrightarrow W(Q_2)/\eqrel{Q_2}$ by $q.\get = \ell \circ \canonizer(Q_1)
-$ and $q.\lput = \ell_1^{-1} \circ \canonizer(Q_2)$. In other words, the get
-function of $q$ first canonizes the source data using $\canonizer(Q_1)$
-and then applies $f$ to the result, while the put function of $q$ first
-canonizes the target data using $\canonizer(Q_2)$ and the applies $f^{-1}$ to
-the result. This is our approach in defining QRE lenses.
-
-One potential pitfall with QRE lenses is that a lens derived using the QRE lens
-combinators may not have the form of a bijection with ``canonizers at the
-edges''. Such a situation would be undesirable since our strategy for
-synthesizing QRE lenses is to synthesize a bijective string lens which plugs
-into the two QREs provided as input to the synthesis procedure; if there were
-some QREs which could not be transformed to the desired form, then our synthesis
-algorithm would necessarily be incomplete.
-
-Thankfully, this turns out not to be an issue if we use the bijections that
-can be expressed in Boomerang; following ~\cite{optician} we call these
-lenses {\em bijective string lenses}. Once we restrict the bijections that we
-use to derive QRE lenses to include just bijective string lenses, we will be
-able show that every QRE lens can be transformed to a QRE lens which has the
-form of a bijective string lens with canonizers at the edges.
+Having identified QREs as a natural way of specifing equivalence relations on
+regular expressions, a natural next step in defining quotient lenses is to map
+between two QREs via a bijection between their; this is our
+approach in defining {\em QRE lenses}.
+
+More concretely, in order to define a language of QRE lenses, our approach is
+to (1) define a language of bijections (Section \ref{subsec:bijective-lenses}),
+and then (2) add quotients by allowing canonizers to be prepended or postpended
+to bijections and (3) allowing composition of such quotient lenses via the
+regular operators and functional composition
+(Sections \ref{subsec:qre-lenses-syntax} and \ref{subsec:qre-lenses-semantics}).
+
+However, we also have a secondary objective, which is to support synthesis of
+lenses: When provided with QREs describing the source and target languages, we
+would like to be able to generate quotient lenses automatically. One way to do
+that is to generate canonizers $\canonizer(Q_1)$ and $\canonizer(Q_2)$ from QREs
+$Q_1$ and $Q_2$ and then to synthesize bijective lenses between the kernel
+languages for $Q_1$ and $Q_2$.
+
+Unfortunately, the composition of two such quotient lenses does not have have
+the form of a bijective lens with canonizers at the edges. Hence, an important
+technical question is whether we give up expressiveness if we restrict
+ourselves to this form. Thankfully, this turns out not to be an issue if
+the bijections that we use to define QRE lenses are derived using a particular
+set of combinators that we describe in the next section.
 
 \subsection{Bijective String Lenses}
-The syntax for the language of bijective string lenses is given by $$\ell_1 :=
-\mathit{id} \; (R) \sep \const(s_1, s_2) \sep  \swap(\ell_1, \ell_2) \sep
-\ell_1 \cdot \ell_2 \; |  \; (\ell_1 \sep \ell_2) \sep \ell^* \; | \; \ell_1 \;
-; \;  \ell_2,$$ where $R$ ranges over regular expressions and $s$ ranges over
-character strings.
-
-The denotation of a lens $\ell_1$ is $\llbracket \ell_1 \rrbracket \subseteq
-\mathit{String} \times \mathit{String}$. If $(s_1, s_2) \in \llbracket \ell_1
-\rrbracket$, then $\ell_1$ maps between $s_1$ and $s_2$.
-\begin{align*}
-\llbracket R \rrbracket &= \{(r, r) \sep r \in \mathcal{L}(R)\}\\
-\llbracket \const(r, s) \rrbracket &= \{(r, s)\}\\
-\llbracket \swap(\ell_1, \ell_2) \rrbracket &= \{(s \cdot t, t' \cdot s') \sep
-(s, s') \in \llbracket \ell_1 \rrbracket \text{ and } (t, t') \in \llbracket
-\ell_2 \rrbracket\}\\
-\llbracket \ell_1 \cdot \ell_2 \rrbracket &= \{(s \cdot t, s' \cdot t) \sep
-(s, s') \in \llbracket \ell_1 \rrbracket \text{ and } (t, t') \in \llbracket
-\ell_2 \rrbracket\}\\
-\llbracket \ell_1 \sep \ell_2 \rrbracket &= \{(s \cdot t) \sep
-(s, t) \in \llbracket \ell_1 \rrbracket \text{ or } (s, t) \in \llbracket
-\ell_2 \rrbracket\}\\
-\llbracket \ell^* \rrbracket &= \{(s_1 \cdot \ldots \cdot s_n, t_1 \cdot \ldots
-\cdot t_n) \sep (s_i, t_i) \in \llbracket \ell \rrbracket \text{ for } 1
-\leq i \leq n\}\\
-\llbracket \ell_1 \; ; \; \ell_2 \rrbracket &= \{(r, t) \sep \text{there exists }s
-\text{ such that } (r, s) \in \llbracket \ell_1 \rrbracket \text{ and } (s, t)
-\in \llbracket \ell_2 \rrbracket\}
-\end{align*}
+\label{subsec:bijective-lenses}
+We define the set of \textit{bijective string lenses} to be the set of
+bijections between regular languages constructed using the following combinators.
+
+$$\ell := \mathit{id} \; (R) \sep \const(s_1, s_2) \sep  \swap(\ell_1, \ell_2)
+\sep \ell_1 \cdot \ell_2 \; |  \; (\ell_1 \sep \ell_2) \sep \ell^* \; | \;
+\ell_1 \; ; \;  \ell_2,$$ where $R$ ranges over regular expressions and $s$
+ranges over character strings.
+
 The $\mathit{\const}(s, t)$ lens replaces the string $s$ with $t$ in the source
-data in the forward direction, and $t$ with $s$ in the target data in the backward
-direction, while $\mathit{id}(R)$ applies the identity function to the
+data in the forward direction, and $t$ with $s$ in the target data in the
+backward direction, while $\mathit{id}(R)$ applies the identity function to the
 source and target  in $\mathcal{L}(R)$ in both directions. The composition lens
@@
 
+The denotation of a lens $\ell_1$ is $\llbracket \ell_1 \rrbracket \subseteq
+\mathit{String} \times \mathit{String}$. If $(s_1, s_2) \in \llbracket \ell_1
+\rrbracket$, then $\ell_1$ maps between $s_1$ and $s_2$.
+\begin{align*}
+\llbracket R \rrbracket &= \{(r, r) \sep r \in \mathcal{L}(R)\}\\
+\llbracket \const(r, s) \rrbracket &= \{(r, s)\}\\
+\llbracket \swap(\ell_1, \ell_2) \rrbracket &= \{(s \cdot t, t' \cdot s') \sep
+(s, s') \in \llbracket \ell_1 \rrbracket \text{ and } (t, t') \in \llbracket
+\ell_2 \rrbracket\}\\
+\llbracket \ell_1 \cdot \ell_2 \rrbracket &= \{(s \cdot t, s' \cdot t) \sep
+(s, s') \in \llbracket \ell_1 \rrbracket \text{ and } (t, t') \in \llbracket
+\ell_2 \rrbracket\}\\
+\llbracket \ell_1 \sep \ell_2 \rrbracket &= \{(s \cdot t) \sep
+(s, t) \in \llbracket \ell_1 \rrbracket \text{ or } (s, t) \in \llbracket
+\ell_2 \rrbracket\}\\
+\llbracket \ell^* \rrbracket &= \{(s_1 \cdot \ldots \cdot s_n, t_1 \cdot \ldots
+\cdot t_n) \sep (s_i, t_i) \in \llbracket \ell \rrbracket \text{ for } 1
+\leq i \leq n\}\\
+\llbracket \ell_1 \; ; \; \ell_2 \rrbracket &= \{(r, t) \sep \text{there exists }s
+\text{ such that } (r, s) \in \llbracket \ell_1 \rrbracket \text{ and } (s, t)
+\in \llbracket \ell_2 \rrbracket\}
+\end{align*}
+
 Each bijective string lens $\ell$ has a type $\ell : R \Leftrightarrow S$ where
@@
 source language of $\ell$ is $\mathcal{L}(R)$ and the target language of $\ell$
-is $\mathcal{L}(S)$. Interestingly, with the bijective string lens type system,
-a lens $\ell : R \Leftrightarrow S$ can also be considered to be of type $\ell
-: R' \Leftrightarrow S'$ provided that $R$ (resp. $S$) can be proven to be
-equivalent to $R'$ (resp. $S$) from the star-semiring axioms (associativity and
-commutativity of | and $\cdot$, identity of the empty string $\epsilon$,
-distibutivity of $\cdot$ over |, annihilative property of $\varnothing$ with
-respect to $\cdot$, and that $R^* = \epsilon \; | \; R \cdot R* = \epsilon \; |
-\; R^* \cdot R$ for all $R$):
+is $\mathcal{L}(S)$. IThe final rule in the figure deserves special attention.
+It states that a lens $\ell : R \Leftrightarrow S$ can also be considered to be
+of type $\ell : R' \Leftrightarrow S'$ provided that $R$ (resp. $S$) can be
+proven to be equivalent to $R'$ (resp. $S$) from the star-semiring axioms
+(associativity and commutativity of | and $\cdot$, identity of the empty string
+$\epsilon$, distibutivity of $\cdot$ over |, annihilative property of
+$\varnothing$ with respect to $\cdot$, and that $R^* = \epsilon \; | \; R \cdot
+R* = \epsilon \; | \; R^* \cdot R$ for all $R$):
 
@@
 
-The other typing rules for bijective string lenses are given in
-Figure~\ref{fig:lensrules}.
+\noindent
+This rule makes it significantly more difficult to synthesize a bijective lens
+from its type.\sam{Should we explain why?}
 
@@
 \subsection{Syntax of QRE Lenses}
-Having given a brief overview of the class of bijective lenses, we now introduce
-the class of QRE lenses. The language of QRE lenses is given by following
+\label{subsec:qre-lenses-syntax}
+Having given a brief overview of the class of bijective string lenses, we now
+introduce the class of QRE lenses. The language of QRE lenses is given by following
 grammar:
@@
 
-\iffalse
-The difference between QRE lenses and Boomerang quotient lenses lies in the
-typing judgements. More concretely, each QRE lens $q_1$ has a type $q_1 : c
-\Leftrightarrow Q_1$ where $c, Q_1$ are QREs. In contrast, the approach used to
-type quotient lenses in Boomerang is to classify equivalences according to
-whether they are or are not the equality relation. This type system is based on
-two observations: first, that most quotient lenses originate as lifted basic
-lenses, and therefore have types whose equivalence relations are both equality;
-and second, that equality is preserved by many of the quotient lens
-combinators. Foster et al also discuss a second possible approach to typing
-quotient lenses, where equivalence relations are represented by rational
-functions that induce them. While this second approach is more refined than the
-first, Boomerang favours the first approach since the second appears to be too
-expensive to be useful in practice~\cite{quotientlenses}.
-\fi
-
 \subsection{Semantics of QRE Lenses}
-
+\label{subsec:qre-lenses-semantics}
 Each QRE lens $q$ has a type $q : Q_1 \Leftrightarrow Q_2$ where $Q_1, Q_2$ are
@@
 \begin{proof}
-By straightforward induction on the derivation
-$q_1 : Q_1 \Leftrightarrow Q_2$. Since bijective lenses are closed under strongly unambiguous
-regular operators, the tricky case is composition.  The composition case is
-more involved since we need to show that we can get rid of the canonizers in
-the middle:
-
-\sam{TODO: Put a diagram here to show what you mean.}
+The proof follows by induction on the derivation of $q : c \Leftrightarrow c$.
+The most interesting part of the proof is the case for functional composition
+as we must demonstrate that it is possible to eliminate the canonizers in the
+middle of the term.
 
@@
 bib_example =
-"@Book {Doe,
- Author = \"John Doe\",
+"@Book {Lovelace,
+ Author = \"Ada Lovelace\",
  Title = {Generic Title},
@@
  %T Generic Title
- %A John Doe
- %F Doe"
+ %A Ada Lovelace
+ %F Lovelace"
 \end{lstlisting}

### 2018-03-10 papers/QuotientLenses/main.tex
add eval fixes

--- 2018-03-09 papers/QuotientLenses/main.tex
@@
     ones?  Are they significantly different (more complex, exercising
-    different features of the synthesis algorithm or QRE syntax, ...)?}
+    different features of the synthesis algorithm or QRE syntax,
+    ...)?}\afm{Solomon, could you fill this out?}
 \end{enumerate}
@@
 Also notable is one of the data.gov examples, which converts demographic
-statistics by zip code from xml form into rdf form.\bcp{It's a bit hidden in the figure...} To validate this performance
-was not due to our overhead, we modified this format to be bijective, so it
+statistics by zip code from xml form into rdf form. Synthesizing this
+transformation certainly took the longest, nearly fifteen seconds, where the
+next slowest took hardly over five. To validate this performance was not due to
+our overhead, we modified these format to be bijective, so the synthesis task
 could be input into Optician. Optician took over 15 seconds synthesize this
 modified lens. We believe this conversion took longer because the input formats
-were exceptionally large, requiring conversions between over 40 fields.
+were exceptionally large, requiring conversions between over 40 fields, not due
+to an overhead from synthesizing from QREs.
 

### 2018-03-10 papers/QuotientLenses/main.tex
in progress edits to intro; changed John Doe to Ada Lovelace

@@
 \end{equation}
-Since these lenses demand that get and put functions form a bijection, then
-no information is lost when one translates from one data format to another. 
-While this strong guarantee can be useful in some contexts, it is more often
-the case that some of the information in one or both of the data formats is
-superfluous. For instance, one typically does not care to preserve the amount of
-whitespace between characters.  In other circumstances, capitalization
-or the order of fields in a record or items in a list may be unimportant. Many
-data formats have other ad hoc rules as well. In bibliographies, for example,
-recall that in \bibtex{}, an author can be specified as
-``\cd{lastname, firstname}'' or as ``\cd{firstname lastname}''---it does not
-matter which of the two formats is used.
-
-{\em Quotient lenses}~\cite{quotientlenses} loosen many of the restrictions of
-bijective string lenses by allowing a programmer to define an equivalence
-relation on the source and target formats so that the round-tripping lens laws
+Since these lenses demand that the \emph{get} and \emph{put} functions form a bijection, 
+no information is lost during translation from one data format to another. 
+While this very strong guarantee can be useful, more often
+some of the information in one or both of the data formats is
+superfluous. For instance, the number of white space characters
+between data items, the capitalization of various strings, the
+sequence of fields in record, or the order of items in a list are
+often irrelevant.  Many
+data formats have their own ad hoc rules as well. In \bibtex{}, for example,
+an author can be specified as
+``\cd{lastname, firstname}'' or as ``\cd{firstname lastname}''---
+the specific choice does not matter.
+
+{\em Quotient lenses}~\cite{quotientlenses} loosen the restrictions of
+bijective string lenses by including equivalence
+relations on the source and target formats so that the round-tripping lens laws
 hold modulo these equivalence rather than modulo equality. Quotient lenses thus
-make it possible to align many more data sources than would otherwise be
-possible, by ignoring details that a programmer views as inessential. The only
-language we are aware of that implements a general class of quotient lenses is
-{\em Boomerang}~\cite{boomerang}. While Boomerang enables a programmer to
-define a very general class of quotient lenses, these lenses to have two main
-drawbacks.
-
-First, in order to define an equivalence relation in Boomerang, a programmer
-must define a {\em canonizer}; pair of functions $f$ and $g$ where $f$ maps
-data represented in the original external format to a canonical element and $g$
-chooses which elements of the original format correspond to values in the
-canonical format. The disadvantage of this approach is that in order for a
-programmer to {\em specify} a quotient lens in Boomerang, she must define four
-different data formats: two ``outer'' formats which specify the data that is to
-be transformed by the lens, and two ``inner'' formats which specify the
-canonical formats for the source and target data.
+make it possible to synchronize many more data formats than would otherwise be
+possible by ignoring details deemed inessential.
+
+The bidirectional programming language {\em
+Boomerang}~\cite{boomerang} is the only language we know of that
+implements a general class of quotient lenses.  Unfortunately, these
+lenses to have two main drawbacks.
+
+The first drawback is that defining an equivalence
+relation in Boomerang requires specifying a {\em canonizer}, which is
+a pair of functions $f$ and $g$ where $f$
+normalizes data from the
+external format by choosing canonical representations
+and $g$ provides an
+inverse by specifying the external form of each canonical
+representation. 
+The disadvantage of this approach is that writing
+a quotient lens in Boomerang requires defining four
+different data formats: two ``external'' formats which specify the data to
+be transformed by the lens, and two ``internal'' formats which specify the
+canonical representations for the source and target data.
 
 Our solution to this shortcoming is to introduce {\em Quotient Regular
-Expressions} or QREs, which enable a programmer to simultaneously specify a
-regular expression as well as a set of canonical representatives for strings
-that match the regular expression. QREs have the added advantage that once the
-programmer has defined a well formed QRE, then she need not define a
-canonization function seperately; the canonization function can automatically be
-inferred from the QRE.
-
-The second challenge with programming lenses in Boomerang is that even after the
-programmer has specified the data as well as a canonizer that maps each member
-in the source and target sets to their respective canonical representative, she
-must still define the underlying lens that maps between the respective sets of
-canonical representatives. This can be a slow and tedious task due to the
-constraints that Boomerang imposes on its terms. For example, Boomerang
-programmers often must rearrange the order of data items by recursively using
-operators that \swap adjacent fields. Furthermore, the Boomerang type checker is
-very strict, disallowing many programs because they contain ambiguity about how
-certain data is transformed.
-
-Our solution to this problem is that we use QREs to define a class of quotient
-lenses called {\em QRE lenses}, and, we show how to synthesize such QRE lenses
-automatically. More specifically, in order to generate a lens, a programmer
-provides a pair of QREs describing the source and target formats, plus
-a (possibly empty) suite of examples that demonstrates how the
-lens should act in specific cases.  Given this information, we can exploit and
+Expressions} (QREs), which enable programmers to simultaneously specify a
+regular expression and a set of canonical representatives for matching
+strings.  Given an QRE, we can automatically infer the corresponding
+canonization functions.
+
+The second drawback with writing quotient lenses in Boomerang is that
+even after specifying the four required data formats, the programmer must
+still define a lens for converting between the canonical
+representations.   Experience shows that writing such lenses can be
+tedious 
+(consider having to rearrange the order of data items by recursively
+using operators to \swap adjacent fields) 
+and that satisfying the type checker can be difficult because 
+many seemingly useful lenses contain ambiguities disallowed by the 
+Boomerang type checker~\cite{optician}.
+
+Our solution to this problem is to introduce {\em QRE lenses}, which
+are a new class of quotient lens that can be automatically
+synthesized, obviating the need to write lenses by hand.  A QRE lens
+is like a normal lens except it maps between formats described by
+QREs. We prove that every composition of QRE lenses can be rewritten
+to be the composition of a source canonizer, a bijective lens, and a
+target canonizer. This normalization property means we can synthesize
+QRE lenses by extending an existing algorithm for synthesizing
+bijective lenses~\cite{optician}.
+
+Generating a QRE lens requires only a pair of QREs to describe the
+source and target formats and a (possibly empty) suite of examples
+demonstrating the mapping.  For example, the following code
+
+\begin{lstlisting}
+let l = synth comma_name $\Leftrightarrow$ space_name using {("Lovelace, Ada", "Ada Lovelace")}
+\end{lstlisting}
+
+binds \codefont{l} to a synthesized QRE lens mapping between names in
+the comma-separated form describd by the QRE \codefont{comma_name}
+and the space-separated form described by the
+QRE \codefont{space_name}.
+
+Given this information, we can exploit and
 extend recent work on algorithms for synthesis of bijective

### 2018-03-10 papers/QuotientLenses/main.tex
finished pass over intro

@@
 canonization functions.  
+\ksf{Define comma\_name here as a really simple example? }
 
@@
 is like a normal lens except it maps between formats described by
-QREs. We prove that every composition of QRE lenses can be rewritten
+QREs. We prove a normal form theorem (Theorem~\ref{normal form}) that says that every
+composition of QRE lenses can be rewritten 
 to be the composition of a source canonizer, a bijective lens, and a
@@
 
-Generating a QRE lens requires only a pair of QREs to describe the
+Given this framework, generating a QRE lens requires only a pair of QREs to describe the
 source and target formats and a (possibly empty) suite of examples
@@
 \end{lstlisting}
-
 binds \codefont{l} to a synthesized QRE lens mapping between names in
-the comma-separated form describd by the QRE \codefont{comma_name}
+the comma-separated form describd by the QRE \codefont{comma\_name}
 and the space-separated form described by the
-QRE \codefont{space_name}.
-
-Given this information, we can exploit and
-extend recent work on algorithms for synthesis of bijective
-lenses~\cite{optician} to generate quotient lenses. The lenses generated by
+QRE \codefont{space\_name}. 
+The lenses generated by
 our synthesis algorithm are guaranteed to translate data back and forth
 between source and target QREs and to satisfy the provided examples.
-Moreover, while our synthesized quotient lenses have a relatively
-restricted form (which cuts down the amount of non-determinism in
-a specification and greatly speeds up synthesis), we prove that
-this rigidity does not compromise their expressiveness.
-
-The following list summarizes the central contributions of our work.
+Our synthesized quotient lenses have a relatively
+restricted form (which greatly speeds up synthesis), but the 
+normalization theorem means 
+this restriction does not compromise expressiveness.
+
+The following list summarizes the contributions of our work.
 \begin{enumerate}
-  \item We introduce {\em Quotient Regular Expressions} (QREs),
+  \item We introduce {\em Quotient Regular Expressions} (QREs), which are
   a compact, convenient notation for simultaneously defining a

### 2018-03-10 papers/QuotientLenses/main.tex
In progress edits to background/example section

@@
 
-\section{QRE Lenses by Example}
-\label{sec:example}
-\subsection{Bijective String Lenses}
-Consider a bidirectional transformation that converts \bibtex{} citation records
-such as
+\section{Background: Bijective String Languages}
+\label{sec:background}
+Before describing QREs and QRE lenses, we briefly review bijective
+string lenses. Consider a bidirectional transformation that converts
+between \bibtex{} citation records such as
 \begin{verbatim}
@@
   Author = "Ada Lovelace",
-Title = {Generic Title}
+  Title = {Notes}
 }
@@
 \noindent
-into an equivalent EndNote format like the following
+and equivalent EndNote records like the following
 \begin{verbatim}
@@
   %A Ada Lovelace
-%T Generic Title
+  %T Notes
 \end{verbatim}
 \noindent
-and vice versa.
-
-Boomerang's bijective string lenses are designed to define a bidirectional
-transformation such as this, where the data formats can be matched in a
+Boomerang's bijective string lenses are designed to define bidirectional
+transformations such as this one, where the data formats can be matched in a
 one-to-one manner, in this case by matching the label, author and title fields.
-Boomerang allows programmers to define such bijecive transformations using
-combinators which favour a compositional approach in which the programmer
-defines small lenses which are then put together using the combinators to
-define larger and larger lenses until the final lens is defined.
-
-Each bijective string lens is constructed from simple primitive combinators including:
+Boomerang encourages a compositional approach in which programmers
+define simple lenses and then compose them using a variety of
+combinators.
+
+Primitive lenses include:
 \begin{itemize}
-  \item \kw{del} $s$: \textit{delete} the constant string $s$ in the get
+  \item \kw{del} $s$: delete the constant string $s$ in the get
   direction; insert it in the put direction.
-  \item \kw{ins} $s$: \textit{insert} the constant string $s$ in the get
+  \item \kw{ins} $s$: insert the constant string $s$ in the get
   direction; delete it in the put direction.
@@
 \end{itemize}
-In addition, such primitives may be combined using the regular operators
-Kleene star, alternation and concatenation, or functional composition of
-lenses.  Figure~\ref{fig:example-lens} illustrates the use of these
-combinators to transform \bibtex{} to EndNote (and vice versa).
+Such primitives may be combined using the regular operators
+Kleene star, alternation and concatenation, as well as lens composition.
+Figure~\ref{fig:example-lens} illustrates the use of these
+combinators to define a lens \cd{bib_to_end} that transforms data
+in \bibtex{} to EndNote (and vice versa). 
 \begin{figure}[t]
@@
 \end{lstlisting}
-\caption{An example lens between \cd{bibtex} and \cd{end_note} regular
-expressions with no equivalences.\bcp{And a readability suggestion:
+\caption{An example bijective lens \cd{bib_to_end} between data
+matching \cd{bibtex} and \cd{end_note} regular 
+expressions, which are omitted for brevity.  
+\bcp{And a readability suggestion:
 define a named constant for a quote character.}}
@@
 
-Unfortunately, programming bijective string lenses in Boomerang is often slow and
-tediuos because of the fiddly constraints that Boomerang imposes on its
-terms. making lens programming slow and tedious. For example, Boomerang
-programmers often must rearrange the order of data items by recursively using
-operators that \swap adjacent fields. Furthermore, the Boomerang type checker is
-very strict, disallowing many programs because they contain ambiguity about how
-certain data is transformed.
-
-In our previous work in~\cite{optician}, we presented a solution to these
-shortcomings of bijective string lenses by developing a tool called {\em Optician}
-which enables a programmer to synthesize a bijective string lens from a pair of regular
-expressions $(S, T)$ and a set of input-output example \cd{exs} pairs that
-determines how the synthesized lens should behave on those examples. With {\em
-Optician}, the programmer need only spend the effort of specifying the data
-formats and the behavior of the lens on a few inputs in order to define a
-bijective string lens. As a result defining bijective string lenses becomes faster and less
-error prone, since the programmer need only invoke the synthesis procedure
-\lstinline{synth S <=> T using exs} in order to define a bijective string lens rather
-than manually derive the lens herself.
+Recent work described the {\em Optician}
+algorithm/tool~\cite{optician}, which can synthesize bijective lenses
+such as \cd{bib_to_end}, obviating the sometimes tedious tasks
+involved in writing such lenses by hand.  Given the directive
+\begin{lstlisting}
+synth S <=> T using exs
+\end{lstlisting}
+\noindent
+{\em Optician} will synthesize a bijective lens between 
+source and target formats described by regular expressions
+\cd{S} and \cd{T}, respectively, and a set of input-output example pairs \cd{exs} specifying how
+the synthesized lens should behave on those examples.
+
+\section{QRE Lenses by Example}
+\label{sec:example}
 
 \subsection{Quotient Lenses}
-Unfortunately, while our synthesis tool greatly simplifies the task of
+Unfortunately, while {\em Optician} greatly simplifies the task of
 programming bijective string lenses, not all bidirectional tranformations are bijective
@@
 fields is also likely irrelevant and there may be equivalent ways of writing
-the same name:  ``Lovelace, Adan'' vs ``Ada Lovelace.''
+the same name:  ``Lovelace, Ada'' vs ``Ada Lovelace.''
 
 When two instances of the same format differ only terms of such inessential
-information, it is useful\bcp{Would it be worth spelling out the alternative
-(i.e., what lenses would look like if they had to preserve all this stuff?)}
-to treat them as identical and to map them into canonical representations.
+information, it is desirable
+to treat them as identical and to map them into canonical
+representations.
+The alternative is to require extra machinery to preserve the
+inessential details, which leads to complex and brittle lenses.
 Continuing our example, the following two \bibtex{} citations represent the same

### 2018-03-10 papers/QuotientLenses/main.tex
partial pass over background/example section; all buyt section on QRE Lenses subsection

@@
 
-\subsection{Quotient Lenses}
-Unfortunately, while {\em Optician} greatly simplifies the task of
-programming bijective string lenses, not all bidirectional tranformations are bijective
-in nature. For instance, users are typically uninterested in preserving
-whitespace between words or between fields.  The order of author and title
-fields is also likely irrelevant and there may be equivalent ways of writing
-the same name:  ``Lovelace, Ada'' vs ``Ada Lovelace.''
-
-When two instances of the same format differ only terms of such inessential
-information, it is desirable
-to treat them as identical and to map them into canonical
-representations.
-The alternative is to require extra machinery to preserve the
-inessential details, which leads to complex and brittle lenses.
-Continuing our example, the following two \bibtex{} citations represent the same
-logical object and as a result, should be translated to the same EndNote record.
+{\em Optician} greatly simplifies the task of programming bijective
+string lenses, but not all bidirectional transformations are
+bijective.  For instance, \bibtex{} users are not typically interested
+in preserving whitespace between words.  The order of author and title
+fields is also likely irrelevant, and there may be equivalent ways of
+writing the same name: ``Lovelace, Ada'' vs ``Ada Lovelace.''
+Consequently, the following two \bibtex{} citations represent the same
+logical object even though they differ in nonessential details.
 \begin{verbatim}
@@
 Author = "Ada Lovelace",
-Title = {Generic Title},
+Title = {Notes},
 }
@@
 @Book{     Lovelace,
-Title = {Generic
-Title},
+Title = {Notes},
 Author = "Lovelace, Ada",     }
@@
 
-One natural way to define lenses that will synchronize data in these formats
-is to provide the following items:
-\begin{enumerate}
-  \item A pair of descriptions of the external source and target formats, here
-  \bibtex{} and EndNote. Such descriptions may be given by regular
-  expressions, for instance.
-  \item A pair of \emph{canonizers} that map external source (and target)
-  objects into internal, canonical forms.  Such canonizers will take
-  care of stripping away irrelevant information and otherwise normalizing
-  differences between equivalent representations of the same logical data.
-  When given an internal representation, such canonizers must also operate
-  in the reverse direction, \emph{choosing} an element in the external format
-  to represent the canonical internal element.
-  \item A pair of descriptions of the internal source and
-  target formats.
-  \item A bijective string lens to map back and forth between canonical elements
-  of the internal source and target formats.
-\end{enumerate}
-
-\noindent
-Indeed, Boomerang's quotient lenses are structured this
-way~\cite{quotientlenses}. Of course, some of these elements are optional.  For
-instance, it is not necessary to document internal and external formats.
-However, we do view it as good engineering practice to do so, as such
-documentation makes lenses easier to read, understand, use and maintain.
-
-The first way in which our work simplifies Boomerang's quotient lenses is
-that we merge the first three pieces of quotient lenses that we just described
-into a single concept, {\em Quotient Regular Expression}, or QREs. A single QRE
-enables a programmer to add annotations to the outer regular expression; these
-annotations can then be used to infer an inner regular expression, as well as a
-canonizer from the outer regular expression to the inner one. In the following
-subsection, we describe a set of QRE combinators that we have integrated into
-Boomerang and give an example of how a programmer might define a QRE
-specification of \bibtex{} using these combinators.
+When mapping these records into another format, such as EndNote, we
+must decide what to do with the nonessential information.  A bijective
+mapping must preserve all the information, including the extraneous
+details, which leads to complex and brittle lenses complicated by the
+need to preserve unimportant details.  
+A better approach is to identify records that differ only in the
+nonessential information, mapping them into a canonical representation.
+This canonical representation is then mapped into the target format.
+With this approach, both of the above \bibtex{} records would be
+mapped to the same EndNote record.
+
+We use {\em Quotient Regular Expresions} (QREs) to specify the
+external format in full detail and to mark which pieces of it are
+inessential.  From a QRE, we can infer a regular expression that
+describes only the essential information, which we call the internal
+format, and we can derive a canonizer that maps between the external
+and internal formats.
 
@@
 \label{subsec:qre-expressions}
-To construct a QRE specification of a \bibtex{} format, we might first define a
-whitespace format, which, externally, matches any non-zero number of whitespace
-characters, and converts any such whitespace into a single space character (its
-canonical form). This whitespace-normalizing QRE is easily defined by using the
-\kw{project} primitive: \bcp{Why is it called ``project''?}
+
+In this subsection, we develop a QRE specification of \bibtex{}
+records, introducing various QRE combinators along the way. 
+Our first step in this process is to define a
+whitespace format, which externally matches any non-zero number of whitespace
+characters. It converts any such whitespace into a single space
+character, its canonical form. 
+We use the QRE \kw{project} primitive to define this
+whitespace-normalizing QRE. \bcp{Why is it called
+``project''?} \ksf{select seems like it might be a better name as we
+are choosing one representative element from a set of possibilities.}
 
 \begin{lstlisting}
-let wsp = project [ \n\t\r]$^+$ $\to$ " "
+let p_wsp = project [ \n\t\r]$^+$ $\to$ " "
 \end{lstlisting}
@@
 Sometimes there are multiple disjoint representations of the same data.
-In such situations, the \kw{squash} combinator creates a QRE that
+In such situations, the ARE \kw{squash} combinator creates a QRE that
 allows external data to be in either format, and converts any
@@
 
-One way to define the \codefont{c\_to\_s} canonization function used by squash
+One way to define the \codefont{c\_to\_s} function
 is simply to write it from scratch in some ordinary programming language.
-However, our system will also synthesize such functions automatically from pairs
-of regular languages and examples---here, \codefont{c\_to\_s} is the
-\codefont{get} direction of a lens that can be synthesized by:
+However, we can synthesize such functions automatically---here, \codefont{c\_to\_s} is the
+\codefont{get} direction of a lens that can be synthesized using the 
+\kw{synth} combinator:
 %
@@
 \end{lstlisting}
-
+\noindent
 The first line above synthesizes a lens between \codefont{comma\_name} and
 \codefont{space\_name} using the listed example transformation as a guide.  The
-second line simply extracts the get direction transformation from the lens,
+second line extracts the \codefont{get} direction transformation from the lens,
 which is what we need for \kw{squash}.
 
-Another QRE primitive is the sequential composition combinator ``;''.  For
-example, suppose that we wish to express a name as $\kw{last\_name \cdot
-``,\text{''} \cdot \text{wsp+} \cdot first\_name}$, or as $\kw{first\_name\cdot
-\text{wsp+} \cdot last\_name}$, with $\kw{first\_name\cdot \text{wsp+} \cdot
-last\_name}$ as the canonical representation. We also wish to project away the
-extra whitespace in between \kw{first\_name} and \kw{last\_name} once this
-identification has been made. Then we can express this compound equivalence
-relation by a composition of a \kw{squash} QRE and a QRE that projects away
-the extra whitespace in between \codefont{first\_name} and
-\codefont{last\_name}.
+Another QRE primitive is the sequential composition combinator,
+written ``;''.
+For an example of its use, suppose we have already defined a QRE
+\codefont{comma_to_space} that accepts names in either space-separated
+or name-separated format and chooses the space-separated format as canonical.
+Suppose we have also already defined a
+QRE \codefont{project_in_space}, which projects away extra white space
+between the first and last name in the space-separated name format. 
+We can use the sequential composition combinator to combine these two QREs:
+\begin{lstlisting}
+let name = project_in_space ; comma_to_space
+\end{lstlisting}
 
@@
 %
-To normalize the field separators, one can specify that the components of the
+To normalize the field separators, one can specify 
+in an optional \kw{with} clause that the components of the
 permutation are conjoined by another QRE. For instance, below, we normalize
@@
 
-The final QRE combinator is the normalize combinator, which is also implemented
-in Boomerang. This combinator allows a programmer to manually define a function
-$f$ which sends each string that matches a regular expression $R$ to some
-canonical representative in another regular expression $R'$. The equivalence
-relation defined by the normalize combinator is hence the equivalence relation
-defined by the {\em fibres} of $f$; that is, for all strings $s$ and $s'$ that
-match $R$, $s$ is equivalent to $s'$ if and only if $f(s) = f(s')$.
+The final QRE combinator is the \kw{normalize} combinator. This combinator
+allows a programmer to manually define a function $f$ which sends each
+string that matches a regular expression $R$ to some canonical
+representative in another regular expression $R'$. The equivalence
+relation defined by the normalize combinator is hence the equivalence
+relation defined by the {\em fibres} of $f$; that is, for all strings
+$s$ and $s'$ that match $R$, $s$ is equivalent to $s'$ if and only if
+$f(s) = f(s')$.
 
 For instance, assume that $f(s)=$ `` '' for all whitespace strings $s$. Then the
-project QRE \kw{p\_wsp} which matches any non-zero number of whitespace
-characters and converts any such whitespace into a single space character can
+project QRE \kw{p\_wsp} defined above can
 be expressed using the normalize combinator as
@@
 \end{lstlisting}
-\caption{QRE specification of \bibtex{} records. }
+\caption{QRE specification of \bibtex{} records. \ksf{Is all this code
+correct?  Does it run through the system correctly?  I'm suspicious of
+the comma\_to\_space definition which seems to be recurisive.  If it's
+correct, we should comment on it.}}
 \label{fig:example-qre}
 \end{figure}
+
 \subsection{QRE Lenses and QRE Lens Synthesis}

### 2018-03-11 papers/QuotientLenses/main.tex
finished pass over background/QREs by example

@@
 \label{sec:background}
+\ksf{We are inconsistent in how we typeset get/put.  Sometimes we use
+codefont, sometimes we use italics of various flavors, and sometimes
+we don't do anything distinctive.  It would be good to be consistent!}
 Before describing QREs and QRE lenses, we briefly review bijective
@@
 \label{subsec:qre-expressions}
+\ksf{Are these QRE combinators or QRE lens combinators?  
+We are currently mushy about this distinction.  
+The section needs a pass to clarify this point.
+The headers of this section and the next imply that this section is
+about QREs and the next section is about QRE lenses, but this section
+is where we introduce the QRE lens combinators, right?  If so, we need
+to do something to make the structure clearer.}
 
@@
 Sometimes there are multiple disjoint representations of the same data.
-In such situations, the ARE \kw{squash} combinator creates a QRE that
+In such situations, the QRE \kw{squash} combinator creates a QRE that
 allows external data to be in either format, and converts any
@@
 the comma\_to\_space definition which seems to be recurisive.  If it's
-correct, we should comment on it.}}
+correct, we should comment on it.}\ksf{In what sense is this a QRE
+spec vs a QRE lens spec?}}
 \label{fig:example-qre}
@@
 \label{sec:examplesynth}
-At this point we already have a tool for synthesizing bijective string lenses from a
-pair of regular expressions and a set of example input-output pairs, as well as
-the language of QREs for defining equivalence relations on regular expressions.
-An obvious next step in simplifying the quotient lens programming is thus to use
-the bijective string lens synthesis procedure as a subroutine for a quotient lens
-synthesis procedure. This new procedure should take two QREs and a set
-of example input-output pairs as input, compute the canonical data formats on
-both sides of the lens from the pair of QREs, map the example input-output pairs
-to their canonical representatives to get a set of examples in canonical form,
+At this point we have a tool for synthesizing bijective string lenses from a
+pair of regular expressions and a set of example input-output pairs
+({\em Optician}), 
+and we have a way of defining regular expressions with equivalence
+relations indicating the essential information (QREs).
+A tantalyzing possibility would be to use the bijective string lens
+synthesis procedure as a subroutine for a quotient lens synthesis procedure. 
+This new synthesizer would take as input source and target QREs and 
+example input-output pairs, 
+compute the canonical source/target formats from the QREs, 
+map the example input-output pairs to their canonical representations,
 and then invoke the bijective string lens synthesis procedure on the canonical data
-formats and the new set of examples that are in canonical form.
-
-Indeed this basic idea is what motivates our definition of {\em QRE Lenses}.
-Intuitively, just like Boomerang's quotient lenses, our QRE lenses are bijective
-lenses with ``canonizers at the edges''---Figure~\ref{fig:attheedges} presents
-the architecture of each lens.
+formats and the canonical examples.
+
+Indeed this idea is what motivates our definition of {\em QRE lenses}.
+Intuitively, our QRE lenses are bijective lenses with ``canonizers at the edges,''
+just like Boomerang's quotient lenses. 
+Figure~\ref{fig:attheedges} depicts the architecture of QRE lenses.
 \begin{figure}[t]
@@
 Every QRE lens $q_1$ has a type $q_1: Q_1 \Leftrightarrow Q_2$ where $Q_1$ and $Q_2$
-are QREs. In the get direction, a QRE lens $q_1: Q_1 \Leftrightarrow Q_2$ uses the
-source QRE $Q_1$ to compute a unique representative for the data modulo the
+are QREs. In the {\em get} direction, a QRE lens $q_1: Q_1 \Leftrightarrow Q_2$ uses the
+source QRE $Q_1$ to compute a canonical representative for the data modulo the
 equivalence relation defined by $Q_1$ and then applies the $\get$ function of a
-bijective string lens $\ell_1$ to this representative. In the backward direction, $q_1$
-operates similarly, but using the view QRE $Q_2$ and the $\lput$ function of
+bijective string lens $\ell_1$ to this representative. In the {\em put} direction, $q_1$
+operates similarly, but using the QRE $Q_2$ and the $\lput$ function of
 $\ell_1$.
@@
 directed by these formats, $Q_1$ and $Q_2$ are all that is required to
-synthesize QRE lenses end-to-end. Indeed, we structured QRE lenses as bijective
-lenses with canonizers at the edges, precisely to facilitate the use of this
-synthesis algorithm.
-
-On the other hand, our restriction that canonizers appear only at the edges
-raises a key technical question:  Are we limiting the expressiveness of our
-transformations by demanding all programs fit into this normal form?\bcp{I wish
-we could have previewed this issue a bit more in the introduction, since it is
-our main contribution.}  It turns out that we are not: any lens that uses
-canonizers internally can be transformed into a lens that uses canonizers at
-the edges. The main technical contribution of this paper is a proof of this
-fact.
-
-This technical investigation justifies replacing manually constructed lenses
-using the QRE lens combinators with synthesized QRE lenses with canonizers at
-the edges. For instance, after defining the \bibtex{} and EndNote QREs and
-storing them in the variables \cd{bibtex} and \cd{endnote} respectively, then
+synthesize QRE lenses end-to-end.
+
+However, our restriction that canonizers appear only at the edges
+raises a key technical question: Are we limiting the expressiveness of
+our transformations by demanding all programs fit into this normal
+form?  It turns out that we are not: any lens that uses canonizers
+internally can be transformed into a lens that uses canonizers only at
+the edges. The main technical contribution of this paper
+(Theorem~\ref{normal form})
+is a proof of this fact.
+This technical result justifies 
+using synthesis to produce QRE lenses instead of manually writing
+them,
+which can lead to substantial savings in program complexity.
+For instance, after defining the \bibtex{} and EndNote QREs and
+binding them to the variables \cd{bibtex} and \cd{endnote} respectively, then
 all the code in Figure~\ref{fig:example-lens} may be replaced by a single call
 to the synthesis prodedure:
-
+\ksf{What is the notation ``lens in'' below?  We should explain it
+intuitively here.}
 \begin{lstlisting}
@@
 \end{lstlisting}
-\noindent Here, the generated  quotient lens synchronizes \cd{bibtex} and
+\noindent 
+Here, the generated quotient lens synchronizes \cd{bibtex} and
 \cd{endnote} formats, using \cd{bib_example} and \cd{end_example} (the two
@@
 \cd{c_to_s}, the synthesis procedure itself can be used to create lenses that
-are in turn used to define other QREs.  The ability to intermix QRE
+are in turn used to define other QREs.  The ability to interleave QRE
 specification with QRE lens synthesis yields a powerful and flexible way of

### 2018-03-11 papers/QuotientLenses/main.tex
minor fixes and proposed alternate qre example

@@
 between data items, the capitalization of various strings, the
-sequence of fields in record, or the order of items in a list are
+sequence of fields in a record, or the order of items in a list are
 often irrelevant.  Many
@@
 relations on the source and target formats so that the round-tripping lens laws
-hold modulo these equivalence rather than modulo equality. Quotient lenses thus
+hold modulo these equivalences rather than modulo equality. Quotient lenses thus
 make it possible to synchronize many more data formats than would otherwise be
@@
 \ksf{Define comma\_name here as a really simple example? }
+\afm{Right now comma\_name is just a regular expression.  Further details in Fig
+2}
 
@@
 \begin{lstlisting}
-let name = project_in_space ; comma_to_space
+let name = comma_to_space ; space_name_p
 \end{lstlisting}
@@
 (* squash QRE maps comma_name to space_name *)
-let comma_to_space = squash comma_name $\to$ space_name using (get comma_to_space)
+let comma_to_space = squash comma_name $\to$ space_name using c_to_s
 
@@
 (* Project away whitespace in space_name representation *)
-let project_in_space = first_name . p_wsp . last_name
+let space_name_p = first_name . p_wsp . last_name
 
 (* Compose the two QREs to get a canonical representation for a name *)
-let name = project_in_space ; comma_to_space
+let name = comma_to_space ; space_name_p
 
@@
 let name_sep = p_wsp . "and" . p_wsp
-let bib_names = name | (name . name_sep . (name . name_sep)$^*$ . name)
+let bib_names = name . (name_sep . name)$^*$)
 let bib_author = "author = \"" . bib_names . "\""
-let title = word | (word . p_wsp . (word . p_wsp)$^*$ . word)
+let title = word . (p_wsp . word)$^*$
 let bib_title = "title = {" . title . "}"
@@
 \end{lstlisting}
-\caption{QRE specification of \bibtex{} records. \ksf{Is all this code
-correct?  Does it run through the system correctly?  I'm suspicious of
-the comma\_to\_space definition which seems to be recurisive.  If it's
-correct, we should comment on it.}\ksf{In what sense is this a QRE
+\caption{QRE specification of \bibtex{} records.\ksf{In what sense is this a QRE
+spec vs a QRE lens spec?}\afm{I don't really like the way this is set up.  The
+use of the sequential composition operator makes the QREs appear really
+operational, and in my opinion makes them seem less cool, and distinct from
+canonizers, than they actually are.  I include a proposed alternative in the
+next figure.  However, this would mean that the previous section which describes
+the sequential composition operator would require it be talked about without the
+concrete example.}}
+\label{fig:example-qre}
+\end{figure}
+
+\begin{figure}
+\begin{lstlisting}
+let wsp = [ \n\t\r]+
+let wsp_s = project wsp $\to$ " "
+
+(* define name representations with a space and with a comma *)
+let space_name = first_name . wsp_s . last_name
+let comma_name = last_name . "," . wsp_s . first_name
+
+(* synthesize a lens that maps comma representation to space representation *)
+let l = synth comma_name $\Leftrightarrow$ space_name using {("Lovelace, Ada", "Ada Lovelace")}
+let c_to_s = get l
+
+(* squash QRE maps comma_name to space_name *)
+let name = squash comma_name $\to$ space_name using c_to_s
+
+(* define rest of bibtex fields *)
+let bib_names = name . (wsp_s . "and" . wsp_s . name)$^*$
+let bib_author = "author = \"" . bib_names . "\""
+let title = word . (wsp_s . word)$^*$
+let bib_title = "title = {" . title . "}"
+
+(* allow any permutation of fields interspersed with arbitrary whitespace *)
+let bib_fields = perm (label, bib_author, bib_title) with (project ("," . wsp) $\to$ ",\n")
+let bibtex = "@book{" . bib_fields . "}"
+\end{lstlisting}
+\caption{QRE specification of \bibtex{} records.\ksf{In what sense is this a QRE
 spec vs a QRE lens spec?}}
-\label{fig:example-qre}
+\label{fig:example-qre-anders-proposal}
 \end{figure}

### 2018-03-11 papers/QuotientLenses/main.tex
add additional question

@@
     too long. I do not think that they are significantly different from the
-    flashfill and augues examples.}
+    flashfill and augues examples.}\afm{So would it be appropriate to say it was
+    these were merely the first examples in the data.gov website whose languages
+    are regular?}
 \end{enumerate}

### 2018-03-12 papers/QuotientLenses/main.tex
Edited related work section

@@
 \subsection{Well Formed QREs}
+\label{subsec:well-formed-qres}
 \subsubsection{Completeness of {\em normalize}}
@@
 \label{relwork}
-
-\iffalse
-Bidirectional transformations have been used in diverse areas of computing
-where they arise as parsers and pretty printers, marshallers
-and unmarshallers, serializers and deserializers, database views and view
-updaters and many others\sam{TODO put citations}. Such transformations have
-been extensively studied since they were proposed as a solution to the classical
-{\em view-update problem} in the database community, where the challenge is to
-derive a program that extracts a view of data from a source, as well as a
-program that folds view updates back into the source safely and correctly.
-\fi
 This paper builds on the work of Foster et al~\cite{quotientlenses} who
@@
 refinement of the bidirectional string processing language
-Boomerang~\cite{boomerang}.  
-In Boomerang, the source and target types are specified using regular
-expressions, and the equivalence relations are expressed using canonizers, which
-are functions that map elements of a regular language to their canonical
-representative\bcp{If their use of the word is exactly what we've been using
-throughout, then why define it?  Or, if we mean their exact set of
-expressible canonizers, rather than the core concept of canonizer, can we be
-precise about that distinction?}. Boomerang canonizers can express a very
-broad class of 
-equivalence relations between regular languages, but actually doing so is often
-difficult with complicated equivalence relations. For instance, Boomerang's
-in-build permutation combinator does not account for separators in between the
-regular expressions\bcp{This sounds like a kind of trivial distinction.}. Also, this combinator permutes regular expressions rather
-than canonizers, thus making it difficult to express nested permutations that
-occur in many data formats, especially XML and XML-like formats.
-
-Foster et al.\relax also discuss other bidirectional programming
-languages that support quotienting of data including XSugar~\cite{xsugar},
-biXid~\cite{bixid} and X/Inv~\cite{Hu2004,Mu2004,Mu2006}.\bcp{The fact that
-  Fosater et al discuss these things is not germane to the present
-  discussion.  We need to make our own comparisons (unless they are
-  exactly the same, in which case we can just refer to Foster et al.).}
-XSugar programs 
-bidirectionally convert data stored in XML and ASCII formats respectively, with the
-transformations specified by a pair of unambiguous grammars. The quotienting
-occurs on the XML side by use of a generic canonizer that standardizes the
-representation of trees. Well-formed XSugar programs are guaranteed to be
-bijective modulo an equivalence relation that captures XML normalization. biXid
-programs convert between pairs of XML documents, with the XML formats specified
-using a pair of grammars as in XSugar. However, biXid grammars can be ambiguous.
-This ambiguity is what allows biXid to express equivalences on the data.
-Finally, Foster et al discuss a possible connection with the languages
-X and Inv which support a primitive duplication combinator that does not work
-well with the lens laws, but can be expressed using Boomerang quotient
-lenses.\bcp{Same comment.}
-
-\BCP{Are there really no bidirectional languages since 2008 that support any
-form of quotienting??}
-
-As far as synthesis goes, though there is a good deal of recent research on
-synthesizing unidirectional string
-transformations~\cite{singh2012learning,le-pldi-2014,gulwani-popl-2014,perelman2014test,Singh:blinkfill},
-the system Optician, which we introduced in ~\cite{optician}, is the first attempt
-to synthesize bidirectional transformations. In that
-publication, we\bcp{confusing: Not the same ``we''!} compared Optician to
-two of these unidirectional string 
-transformers, Flash Fill~\cite{gulwani-popl-2014} and
-FlashExtract~\cite{le-pldi-2014}, and found that these tools were unsuccessful
-in synthesizing the complex transformatiunons that can be expressed in Optician.
-
-\bcp{This paragraph seems redundant.}While Optician improved on prior effors
-at synthesizing unidirectional string 
-transformers and introduced synthesis for bidirectional string transformations,
-Optician was only able to synthesize fully bijective data transformations. Our
-new system \Name{} improves on Optician in that \Name{} loosens this
-restriction, and is able to synthesize data which is bijective when put in
-canonical form. \Name{} is therefore a refinement of Optician since the
-new synthesis algorithm (i.e. the quotient lens synthesis algorithm) restricted
-to ``pure'' regular expressions degenerates to the synthesis algorithm that
-we introduced in Optician. Indeed, as we mentioned in the evaluation section,
-\Name{} was able to synthesize all of Optician's benchmarks at a speed
-competitive with Optician, though there was additional overhead in calculating
-the whole language $W(c)$ and the kernel language $K(c)$ of QREs, which
-resulted in a very slight decrease in performance.
-
-Much of the research in synthesis assumes that the synthesizer is provided with
-a collection of examples. Optician and Optometrist differ in that they requires
-that the programmer supplies both examples {\em and} format descriptions in the
-form of regular expressions or QREs.\bcp{But we are far from the only ones
-  to consider type-based synthesis!!}  There is a trade-off here.  On the one
-hand, a user must have some programming expertise to write regular expression
-(or QRE) specifications, and it requires some work. On the other hand, such
-specifications provide a great deal of information to the synthesis system,
-which decreases the number of examples needed (often to zero), makes the system
-scale well, and allows it to handle large, complex formats.  By providing these
-format specifications, the synthesis engine does not have to both infer the
-format of the data as well as the transformations on it, obviating the need to
-infer tricky formats like those involving nested iterations. Furthermore, by
-focusing on bidirectional transformations, we limit the space of synthesized
-functions to bijective ones, reducing the search space, and the expressiveness
-of the search space.
+Boomerang~\cite{boomerang}. As we mentioned in
+Section\cref{subsec:well-formed-qres}, all our QRE combinators can be expressed
+using just the {\em normalize} combinator, which is one of the canonizer
+primitives that Boomerang already supports. Also, all our QRE lens combinators
+are already supported in Boomerang. Consequently Boomerang quotient lenses are
+at least as expressive as our language of QRE lenses.
+
+Boomerang's canonizers allow one to canonize a regular language $R$ to by
+mapping it to another regular language $S$ which may not be contained in $S$.
+Formally, given sets $C$ and $B$ and an equivalence relation on $B$, Foster et
+al defined a {\em canonizer} $q$ from $C$ to $B/{\sim_B}$ to be a pair of
+functions $q.\kw{canonize} : C \longrightarrow B$ and $q.\kw{choose} : B
+\longrightarrow C$ such that for every $b \in B$,
+$$q.\kw{canonize} \; (q.\kw{choose} \; b) \sim_B \; b$$
+This definition gives allows much more latitude for defining canonizers than
+QREs. For example, if $\sim_B = \kw{Tot}(B)$, the equivalence relation that
+relates every element in $B$ to every other element in $B$, then every function
+from $C$ to $B$ is a canonizer.
+
+Because of this extra elbow-room, Boomerang has two primitive {\em duplication}
+quotient lenses, the first of which can be derived using the following inference
+rule,
+\begin{prooftree}
+\AxiomC{$\ell : C/{\sim_C} \Leftrightarrow A_1/{\sim_{A_1}}$}
+\AxiomC{$f : C \longrightarrow A_2$}
+\AxiomC{$A_1 \cdot^! A_2$}
+\AxiomC{$\sim_A = \sim_{A_1} \cdot \kw{Tot}(A_2)$}
+\QuaternaryInfC{$\kw{dup}_1 \ell \; f \; : \; C/{\sim_C} \Longrightarrow A_1
+\cdot A_2/{\sim_A}$}
+\end{prooftree}
+\begin{align*}
+(\kw{dup}_1 \; \ell \; f).\get \; c &= (\ell.\get \; c) \cdot (f \; c)\\
+(\kw{dup}_1 \; \ell \; f).\lput \; (a_1 \cdot a_2) \; c &= \ell.\lput \; a_1 \; c\\
+(\kw{dup}_1 \; \ell \; f).\create \; (a_1 \cdot a_2) &= \ell.\create \; a_1
+\end{align*}
+wiht the symmetric $\kw{dup}_2$ combinator discarding the first copy instead of
+the second in the \lput/\create direction.
+
+Boomerang's more general definition for canonizers also allows quotient lenses
+to be used as canonizers by using the taking the \kw{canonize} function to be
+the \get component of a lens and the \kw{choose} function to be its \create
+component. Naturally, QREs also take advantage of this ability to use lenses as
+canonizers by allowing for user-defined functions to be used by the \kw{squash}
+and \kw{normalize} combinators. Moreover, the added functionality of
+synthesizing lenses further makes it easier to define canonizers, as well as
+lenses.
+
+Foster et al also discuss other bidirectional programming languages that
+support quotienting of data including XSugar~\cite{xsugar}, biXid~\cite{bixid}
+and X/Inv~\cite{Hu2004,Mu2004,Mu2006}; we refer the reader to the related work
+section in \cite{quotientlenses} for this discussion.
+
+Another language which supports quotenting is BiFluX~\cite{pacheco2014biflux}, a
+bidirectional functional update language for XML. BiFluX is inspired by
+the FLUX XML update language~\cite{cheney2008flux}, and adopts a bidirectional
+programming by update paradigm, where a program succintly and precisely
+describes {\em how} to update a source document with a target document, in an
+intuitive way, such that there is a unique ``inverse'' source query for each
+update program. The source and view types of BiFluX programs are given by
+regular expressions.
+
+In BiFluX, a regular expression $R$ is said to be unambiguous if there is only
+one way to parse a {\em flat} value of type $R$ to a stuctured value of type
+$R$; here a flat value of $R$ is a value of $R$ with all left/right tags,
+parantheses and list brackets removed. The bidirectional part of BiFluX involves
+transforming source regular expressions to view regular expressions so that
+data sequences described by the two types can be matched. This process can
+result in intermediate types that are unambiguous, and this ambiguity can cause
+unintended updates to be made the source, especially when information is
+discared in this process. The typechecking phase of a BiFluX program therefore
+includes a {\em type normalization} phase that tries to normalize a source types
+using automata reduction techniques, and derive a lens between ambiguous and
+unambiguous types. On the view side, this phase only tries to normalize view
+types into isomorphic types.
+
+{\em Formlenses}~\cite{rajkumar2014lenses} are a variant of lenses that perform
+transformations up to equivalence. Formlenses are a bidirectional generalization
+of {\em formlets}~\cite{cooper2008essence}. Formlets are a high-level
+abstraction for building Web forms, and can be represented in Haskell as
+functions that take a source of names---concretely, an \kw{Int}---as an
+argument and produce a triple consisting of an HTML document, a {\em collect}
+function, and a modified namesource:
+\begin{center}
+\begin{lstlisting}
+type Formlet a = Int -> (Html, Env -> a, Int)
+\end{lstlisting}
+\end{center}
+The names of any generated form fields are drawn from the namesource, and the
+{\em collect} function looks up precisely these names.
+Rajkumar et. al.~\cite{rajkumar2014lenses} observe that formlets are a useful
+abstraction, but they only address one half of the problem---they make it easy
+to construct a function that {\em produces} a value of type \kw{a}, but they do
+not provide a way to describe a function that {\em consumes} a value of type
+\kw{a} and embeds it in a form. A different way to think about this problem is
+to observe that forms are often used to present an {\em updateable view} of the
+data source. Therefore, Rajkumar et. al. leverage the power of bidirectional
+transformations to define formlets that behave like updatable views, yielding
+the abstraction of formlenses.
+
+They \kw{Formless} type is given by,
+\begin{center}
+\begin{lstlisting}
+type Formlens a = Maybe a -> [Int] -> (Html,Env -> Maybe a,[Int])
+\end{lstlisting}
+\end{center}
+with the differences between this definition and the standard definition for
+formlets being an extra \kw{Maybe a} parameter that provides an optional initial
+value of the form, the namesource being changed to a list of integers, and the
+{\em collect} function being allowed to return an optional value.
+
+Rajkumar et. al. consider only formlenses that are well-formed in that a
+well-formed formlens should only draw names from the namesource provided as an
+argument, and the {\em collect} function should only look up corresponding
+names.
+Under this assumption, Rajkumar et. al. define what it means for a formlens to
+be well behaved by defining a version of the classical GETPUT and PUTGET laws
+for formlenses. These laws hold modulo an equivalence relation on environments,
+since a formlens may be invoked with two separate namespaces, but this should
+not affect the value stored in the information encapsulated in the type $a$.
+Consequently, the GETPUT and PUTGET laws defined for formlenses are in spirit
+analogous to the GETPUT and PUTGET laws for quotient lenses, since they hold up
+to renaming of environment variables.
+
+\iffalse
+Well-behaved formlenses satisfy properties that are analogous to the classical
+GETPUT and PUTGET law. If {\em extract} is a function from \kw{Html} to \kw{Env} that
+extracts an environment containing the names and values of all the form fields
+from an HTML document, then a formlens $f$ satisfies the Acceptability law if
+for any $v$ of type $a$, $n$ of type \kw{Int} and environment $e$ of type
+\kw{Env} if $f$ applied to \kw{Just v} and $n$ yields an HTML file $h$ and a
+{\em collect} function $c$ where \kw{extract h} is contained in $e$, then
+collect $e$
+\fi
+
+Finally, in Cunha~\cite{cunha2010relational} uses relational algebra to
+encompass many of the existing approaches to bidirectional programming. For
+example, using $\cdot$ to express relational composition, the GETPUT law states
+that $\get \cdot \lput \subseteq \sim_V \cdot \pi_1$, where $\pi_1$ is the
+projection onto the first factor. In addition to its generality, this
+``point-free'' approach to defining bidirectional transformations has the
+advantage of creating equational proofs of lens laws that proceed by folding
+and unfolding combinator laws.
+
+Now we turn to related work in synthesis. Much of the research in synthesis
+assumes that the synthesizer is provided with a collection of examples. Optician
+and Optometrist differ in that they require that the programmer supplies both
+examples {\em and} format descriptions in the form of regular expressions or
+QREs, though we are far from the only ones to consider type-based synthesis.
+There is a trade-off here.  On the one hand, a user must have some programming
+expertise to write regular expression (or QRE) specifications, and it requires
+some work. On the other hand, such specifications provide a great deal of
+information to the synthesis system, which decreases the number of examples
+needed (often to zero), makes the system scale well, and allows it to handle
+large, complex formats.  By providing these format specifications, the 
+synthesis engine does not have to both infer the format of the data as well as
+the transformations on it, obviating the need to infer tricky formats like
+those involving nested iterations. Furthermore, by focusing on bidirectional
+transformations, we limit the space of synthesized functions to bijective ones,
+reducing the search space, and the expressiveness of the search space.
 

### 2018-03-12 papers/QuotientLenses/main.tex
A few responses and further comments

@@
 expressions, which are omitted for brevity.  
+\iflater
 \bcp{And a readability suggestion:
-define a named constant for a quote character.}}
+  define a named constant for a quote character.}\fi}
 \label{fig:example-lens}
@@
 ``project''?} \ksf{select seems like it might be a better name as we
-are choosing one representative element from a set of possibilities.}
+are choosing one representative element from a set of possibilities.}\bcp{Or
+``collapse''?}
 
@@
 \section{Quotient Regular Expressions}
+
+\BCP{What became of my suggestion that we start with just ``normalize'' and
+  then, later, introduce the other, more specific (and pragmatic, and
+  incomplete) ones?}
+
 A Quotient Regular Expression (or QRE) is a regular expression $R$ augmented
@@
   \item a ``canonizing'' function\bcp{Why is it called ``canonizer'' rather
-  than ``canonize''??}\sam{because \canonizer(Q) is a function}
+  than ``canonize''??}\sam{because \canonizer(Q) is a function}\bcp{My ear
+  consistently finds it quite awkward.  I'd be happier with
+  ``canonize'' (which describes an action).}
   $\canonizer(Q):\mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$ which
@@
     these were merely the first examples in the data.gov website whose languages
-    are regular?}
+    are regular?}\BCP{Or would anything be lost by just deleting them?  It
+    seems like we're tying ourselves in knots at bit trying to explain how
+    they relate to the others and what is interesting about them, when the
+    answer seems to be they don't and nothing.}
 \end{enumerate}
@@
 assumes that the synthesizer is provided with a collection of examples. Optician
-and Optometrist differ in that they require that the programmer supplies both
+and Optometrist\bcp{Sole use?!} differ in that they require that the
+programmer supplies both 
 examples {\em and} format descriptions in the form of regular expressions or

### 2018-03-12 papers/QuotientLenses/main.tex
change the bibtex example

@@
 strings.  Given an QRE, we can automatically infer the corresponding
-canonization functions.  
-\ksf{Define comma\_name here as a really simple example? }
-\afm{Right now comma\_name is just a regular expression.  Further details in Fig
-2}
+canonization functions.  As an example, consider the following QRE for writing
+author names:
+%
+\begin{lstlisting}
+let wsp_sp = project wsp $\to$ " "
+let comma_name = first_name . "," . wsp_sp . last_name
+\end{lstlisting}
+%
+In this example, \lstinline{wsp} is an existing regular expression for a
+nonempty sequence of whitespace characters, and \lstinline{first_name} and
+\lstinline{last_name} are existing regular expressions for first and last names.
+The QRE \lstinline{wsp_sp} is a QRE with the same underlying language as
+\lstinline{wsp}, but with a single canonical representative: \lstinline{" "}.
+This is then used to build up \lstinline{comma_name}, a QRE with an
+underlying language of two names, separated by a command and whitespace, where
+the names with a single space between them are canonical.
 
@@
 expressions, which are omitted for brevity.  
-\iflater
+\ifdraft
 \bcp{And a readability suggestion:
-  define a named constant for a quote character.}\fi}
+  define a named constant for a quote character.}
+\fi}
 \label{fig:example-lens}
@@
 is where we introduce the QRE lens combinators, right?  If so, we need
-to do something to make the structure clearer.}
+to do something to make the structure clearer.}\afm{I believe this are QRE
+combinators, so we should be good.}
 
@@
 \begin{lstlisting}
-let p_wsp = project [ \n\t\r]$^+$ $\to$ " "
+let wsp_sp = project wsp $\to$ " "
 \end{lstlisting}
@@
 
-Another QRE primitive is the sequential composition combinator,
-written ``;''.
-For an example of its use, suppose we have already defined a QRE
-\codefont{comma_to_space} that accepts names in either space-separated
-or name-separated format and chooses the space-separated format as canonical.
-Suppose we have also already defined a
-QRE \codefont{project_in_space}, which projects away extra white space
-between the first and last name in the space-separated name format. 
-We can use the sequential composition combinator to combine these two QREs:
-\begin{lstlisting}
-let name = comma_to_space ; space_name_p
-\end{lstlisting}
-
 The permutation QRE combinator, \kw{perm}, allows data to be unordered. For example,
@@
 \begin{lstlisting}
-let bib_fields = perm (label, bib_author, bib_title) with (project ("," .wsp+) $\to$ ",\n")
+let bib_fields = perm (label, bib_author, bib_title) with (project ("," . wsp) $\to$ ",\n")
 \end{lstlisting}
+
+Another QRE primitive is the sequential composition combinator, written ``;''.
+For an example of its use, suppose we have already defined a QRE,
+\lstinline{canonized_whitespace}, that accepts XML documents and chooses
+documents with no whitespace as canonical. Suppose that we also have defined a
+QRE, \lstinline{canonized_order}, which accepts whitespace-normalized XML
+documents, and chooses a specific ordering of XML elements as canonical. We can
+use the sequential composition combinator to combine these two QREs into
+\lstinline{canonized_whitespace ; canonized_order}, a QRE that accepts all XML
+documents, and chooses ordered XML documents without whitespace as canonical.
 
@@
 For instance, assume that $f(s)=$ `` '' for all whitespace strings $s$. Then the
-project QRE \kw{p\_wsp} defined above can
+project QRE \lstinline{wsp_sp} defined above can
 be expressed using the normalize combinator as
-\lstinline{normalize (wsp+, " ", f)}.
+\lstinline{normalize (wsp," ", f)}.
 
@@
 \begin{lstlisting}
-let wsp = [ \n\t\r]
+let wsp = [ \n\t\r]+
+let wsp_sp = project wsp $\to$ " "
+let last_name = [A-Z][a-z]
+let first_name = [A-Z][a-z]
 
 (* define name representations with a space and with a comma *)
-let space_name = first_name . wsp+ . last_name
-let comma_name = last_name . "," . wsp+ . first_name
+let space_name = first_name . wsp_sp . last_name
+let comma_name = last_name . "," . wsp_sp . first_name
 
@@
 (* squash QRE maps comma_name to space_name *)
-let comma_to_space = squash comma_name $\to$ space_name using c_to_s
-
-(* map arbitrary whitespace into " " as a canonical representative *)
-let p_wsp = project wsp+ $\to$ " "
-
-(* Project away whitespace in space_name representation *)
-let space_name_p = first_name . p_wsp . last_name
-
-(* Compose the two QREs to get a canonical representation for a name *)
-let name = comma_to_space ; space_name_p
+let name = squash comma_name $\to$ space_name using c_to_s
 
 (* define rest of bibtex fields *)
-let name_sep = p_wsp . "and" . p_wsp
-let bib_names = name . (name_sep . name)$^*$)
+let bib_names = name . (wsp_sp . "and" . wsp_sp . name)$^*$
 let bib_author = "author = \"" . bib_names . "\""
-let title = word . (p_wsp . word)$^*$
-let bib_title = "title = {" . title . "}"
-
-(* allow any permutation of fields interspersed with arbitrary whitespace *)
-let bib_fields = perm (label, bib_author, bib_title) with (project ("," . wsp+) $\to$ ",\n")
-let bibtex = "@book{" . bib_fields . "}"
-\end{lstlisting}
-\caption{QRE specification of \bibtex{} records.\ksf{In what sense is this a QRE
-spec vs a QRE lens spec?}\afm{I don't really like the way this is set up.  The
-use of the sequential composition operator makes the QREs appear really
-operational, and in my opinion makes them seem less cool, and distinct from
-canonizers, than they actually are.  I include a proposed alternative in the
-next figure.  However, this would mean that the previous section which describes
-the sequential composition operator would require it be talked about without the
-concrete example.}}
-\label{fig:example-qre}
-\end{figure}
-
-\begin{figure}
-\begin{lstlisting}
-let wsp = [ \n\t\r]+
-let wsp_s = project wsp $\to$ " "
-
-(* define name representations with a space and with a comma *)
-let space_name = first_name . wsp_s . last_name
-let comma_name = last_name . "," . wsp_s . first_name
-
-(* synthesize a lens that maps comma representation to space representation *)
-let l = synth comma_name $\Leftrightarrow$ space_name using {("Lovelace, Ada", "Ada Lovelace")}
-let c_to_s = get l
-
-(* squash QRE maps comma_name to space_name *)
-let name = squash comma_name $\to$ space_name using c_to_s
-
-(* define rest of bibtex fields *)
-let bib_names = name . (wsp_s . "and" . wsp_s . name)$^*$
-let bib_author = "author = \"" . bib_names . "\""
-let title = word . (wsp_s . word)$^*$
+let title = word . (wsp_sp . word)$^*$
 let bib_title = "title = {" . title . "}"
@@
 spec vs a QRE lens spec?}}
-\label{fig:example-qre-anders-proposal}
+\label{fig:example-qre}
 \end{figure}

### 2018-03-12 papers/QuotientLenses/main.tex
evaluation text changes

@@
 
-  \item Does the increase in expressivity gained from synthesizing from QREs
-  permit synthesis of more useful lenses than existing approaches.  \bcp{Is
-    this really an interesting empirical question?  Seems obvious, from the
-    arguments we made in Section 2 and the intro.}\afm{I'm happy to remove it}
-
   \item Does our tool synthesize lenses fast enough to be used as part of a
@@
 
-To evaluate our QRE implementation, we constructed a set of lens synthesis tasks
-drawn from two sources:
-\begin{enumerate}
-\item 39 of our benchmarks were adapted from the evaluation of the original
-  Optician system.  These benchmarks are a combination of custom benchmarks,
-  benchmarks derived from FlashFill~\cite{flashfill}, and benchmarks
-  derived from Augeas~\cite{augeas2}. In the pre-QRE version of
-  Optician, many of these data formats had to be modified to work with the
-  bijectivity constraints that Optician required.  For instance, when one format
-  permits whitespace where the other does not allow it, the original versions of
-  benchmark formats were modified to allow more whitespace, thereby restoring
-  bijectivity.  With the new QRE support, we were able to remove these
-  alterations when porting these benchmarks.  (This experience alone suggests
-  that QREs make the lens development process easier.)
-
-\item The remaining eight benchmarks were derived from datasets found at 
-  data.gov. These datasets include census
-  population data, wage comparison data, demographics statistics data and
-  crime index data. We chose these datasets since they were stored in XML, CSV
-  and RDF formats which are more easily specified using regular expressions than
-  plain text or PDF documents.  \BCP{Besides differing formats, why did we
-    include these eight new 
-    ones?  Are they significantly different (more complex, exercising
-    different features of the synthesis algorithm or QRE syntax,
-    ...)?}\afm{Solomon, could you fill this out?}
-    \sam{Well these data formats were the ones I was playing with before we got
-    the Augeus and flashfill examples. I chose them simply because they had
-    formats which could be expressed using regular expressions. They basically
-    use the same QREs as the flashfill and augeus examples, mostly the project
-    QRE. The fields in these files are occur in the same order in each record;
-    we tried using the perm combinator anyways but the synthesis algorithm took
-    too long. I do not think that they are significantly different from the
-    flashfill and augues examples.}\afm{So would it be appropriate to say it was
-    these were merely the first examples in the data.gov website whose languages
-    are regular?}\BCP{Or would anything be lost by just deleting them?  It
-    seems like we're tying ourselves in knots at bit trying to explain how
-    they relate to the others and what is interesting about them, when the
-    answer seems to be they don't and nothing.}
-\end{enumerate}
+To evaluate our QRE implementation, we adapted 39 lens synthesis tasks from the
+benchmark suite of the original Optician system. These benchmarks are a
+combination of custom benchmarks, benchmarks derived from
+FlashFill~\cite{flashfill}, and benchmarks derived from Augeas~\cite{augeas2}.
+In the pre-QRE version of Optician, many of these data formats had to be
+modified to work with the bijectivity constraints that Optician required. For
+instance, when one format permits whitespace where the other does not allow it,
+the original versions of benchmark formats were modified to allow more
+whitespace, thereby restoring bijectivity (but altering the data format). With
+the new QRE support, we were able to remove these alterations when porting these
+benchmarks. (This experience alone suggests that QREs make the lens development
+process easier.)
 
@@
 to be about half as large as \CanonizerAndSpecSize, and about a third as large
-as \LensAndSpecSize{}.\BCP{These are really the critical conclusions, and
-  they are hard to see, the way the graph is drawn.  I wonder whether we
-  could instead present with the individual examples on the X axis (sorted
-  by number of AST nodes), with a bar above each one containing three
-  regions, colored differently, for LS, CS-LS, and QC-CS...}\afm{will do} This suggests
-that introducing QREs saves programmers 
-significant effort compared to both Optician and basic Boomerang.
-
-\subsection{Increasing Expressivity}
-
-The pre-QRE version of Optician was only able to synthesize fully bijective data
-transformations. Synthesizing from QREs loosens that restriction, making it
-possible to synthesize data transformations involving bijections with
-canonization functions. Because Optician could previously only synthesize
-bijective lenses, some of the data formats in the example suite were modified to
-include additional fields containing information should should be canonized from
-the other format. In total, 12 of Optician's original benchmarks and 3 of the
-data.gov benchmarks would require such additional fields when input to Optician
-without QREs.
+as \LensAndSpecSize{}.\BCP{These are really the critical conclusions, and they
+  are hard to see, the way the graph is drawn. I wonder whether we could instead
+  present with the individual examples on the X axis (sorted by number of AST
+  nodes), with a bar above each one containing three regions, colored
+  differently, for LS, CS-LS, and QC-CS...}\afm{will do} This suggests that
+introducing QREs saves programmers significant effort compared to both Optician
+and basic Boomerang.
 

### 2018-03-12 papers/QuotientLenses/main.tex
More edits

@@
 \newcommand{\normalize}[3]{\ensuremath{\kw{normalize} \; (#1, #2, #3)}}
+\newcommand{\eqrel}[1]{\ensuremath{\equiv_{#1}}}
 \newcommand{\sep}{\ensuremath{\ | \ }}
-\newcommand{\canonizer}{\ensuremath{\kw{canonizer}}}
+\newcommand{\canonize}{\ensuremath{\kw{canonize}}}
 \newcommand{\bibtex}{\textsc{Bib}\TeX{}}
@@
 \newcommand{\swap}{\ensuremath{\kw{swap}}}
-
-\newcommand{\eqrel}[1]{\ensuremath{\equiv_{#1}}}
+\newcommand{\id}{\ensuremath{\kw{id}}}
+\newcommand{\lquot}{\ensuremath{\kw{lquot}}}
+\newcommand{\rquot}{\ensuremath{\kw{rquot}}}
+\newcommand{\lift}{\ensuremath{\kw{lift}}}
 
@@
 \newcommand{\QRESize}{\textbf{QS}}
-\newcommand{\CanonizerAndSpecSize}{\textbf{CS}}
+\newcommand{\canonizeAndSpecSize}{\textbf{CS}}
 \newcommand{\LensAndSpecSize}{\textbf{LS}}
@@
 two representations continue to store the same information.
-One way to synchronize such data is to use \emph{lenses}\ksf{include citations?}, which
+One way to synchronize such data is to use \emph{lenses}~\cite{Lenses}, which
 are programs that simultaneously define pairs of functions for translating
-the data 
-from the source format to the target formant (the \emph{get}
+the data from the source format to the target formant (the \emph{get}
 direction) and back again (the \emph{put} direction). Lenses are usually
@@
 representations.   Experience shows that writing such lenses can be
-tedious 
-(consider having to rearrange the order of data items by recursively
-using operators to \swap adjacent fields) 
-and that satisfying the type checker can be difficult because 
-many seemingly useful lenses contain ambiguities disallowed by the 
-Boomerang type checker~\cite{optician}.
+tedious (consider having to rearrange the order of data items by recursively
+using operators to \kw{swap} adjacent fields)  and that satisfying the type
+checker can be difficult because many seemingly useful lenses contain
+ambiguities disallowed by the Boomerang type checker~\cite{optician}.
 
@@
 is like a normal lens except it maps between formats described by
-QREs. We prove a normal form theorem (Theorem~\ref{normal form}) that says that every
-composition of QRE lenses can be rewritten 
+QREs. We prove a normal form theorem (Theorem~\ref{normal form}) that says that
+every composition of QRE lenses can be rewritten 
 to be the composition of a source canonizer, a bijective lens, and a
@@
 \section{Quotient Regular Expressions}
-
-\BCP{What became of my suggestion that we start with just ``normalize'' and
-  then, later, introduce the other, more specific (and pragmatic, and
-  incomplete) ones?}
-
+\label{QRE}
 A Quotient Regular Expression (or QRE) is a regular expression $R$ augmented
@@
 section describes in further detail the set of QRE combinators that we described
-in \cref{sec:example}.
-\label{QRE}
+in \cref{sec:example}, but first we discuss the \kw{normalize} combinator which
+enjoys some special properties.
+
+\subsection{The \kw{normalize} Combinator}
+The \kw{normalize} combinator is special in two main ways, the first of
+which has to do with the premises in its inference rule:
+\begin{prooftree}
+\AxiomC{$\mathcal{L}(R') \subseteq \mathcal{L}(R)$}
+\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(R')$}
+\AxiomC{$f$ is surjective}
+\AxiomC{$f = f^2$}
+\RightLabel{(Normalize)}
+\QuaternaryInfC{$\wf{\normalize{R}{R'}{f}}$}
+\end{prooftree}
+In order for the {\em normalize} QRE to be well-formed, then canonizing function
+$f$ used by this combinator must be surjective and idempotent. Unfortunately,
+verifying that a function satisfies these properties is in general undecidable.
+Nonetheless, we allow a programmer to use normalize at their own risk without
+checking these side conditions.
+
+We do this firstly because the {\em normalize} combinator allows one to express
+equivalence relations that are either difficult to define, or that cannot be
+defined using the other combinators. Secondly, the {\em normalize} canonizer
+enjoys the property that each of the other combinators $Q$ is semantically
+equivalent to $\normalize{W(Q)}{K(Q)}{f}$ for some surjective, idempotent
+function $f : \mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$; this is the
+second property that distinguishes the {\em normalize} combinator from the other
+QREs. The {\em normalize} combinator hence gives a sufficient condition for
+potential QREs to be ``well-behaved'', particularly with respect to the
+strategy for synthesizing quotient lenses which we shall introduce in
+Section~\ref{synth}.
+
+Having discussed the special properties of the \kw{normalize} combinator, we now
+give the syntax and semantics of the other QRE combinators.
 \subsection{Syntax and Semantics of QREs}
@@
 \begin{align*}
-Q := \; &R \sep \project{R}{s} \sep \squash{Q_1}{Q_2}{f} \sep
+Q := \; &\id(R) \sep \project{R}{s} \sep \squash{Q_1}{Q_2}{f} \sep
 \perm{Q_1, \ldots, Q_n}{Q} \\
@@
   $\eqrel{Q}$, and
-  \item a ``canonizing'' function\bcp{Why is it called ``canonizer'' rather
-  than ``canonize''??}\sam{because \canonizer(Q) is a function}\bcp{My ear
-  consistently finds it quite awkward.  I'd be happier with
-  ``canonize'' (which describes an action).}
-  $\canonizer(Q):\mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$ which
-  given any $w \in \mathcal{L}(W(Q))$, computes $\canonizer(Q)(w)$ as the
-  unique $k$ in $\mathcal{L}(K(Q))$ such that $k$ is equivalent to $w$ mod
-  $\eqrel{Q}$
+  \item a ``canonizing'' function $\canonize(Q):\mathcal{L}(W(Q))
+  \longrightarrow \mathcal{L}(K(Q))$ which given any $w \in \mathcal{L}(W(Q))$,
+  computes $\canonize(Q)(w)$ as the unique $k$ in $\mathcal{L}(K(Q))$ such
+  that $k$ is equivalent to $w$ mod $\eqrel{Q}$
 \end{enumerate}
@@
 maps $\mathcal{L}(W(Q_1))$ into $\mathcal{L}(W(Q_2))$ using $f$ and then
-canonizes $W(Q_2)$ into $K(Q_2)$ using $\canonizer(Q_2)$.
+canonizes $W(Q_2)$ into $K(Q_2)$ using $\canonize(Q_2)$.
 
@@
 Figure~\ref{fig:canonizers} gives the inductive definitions of the
-$\canonizer{}$ function for each QRE.
+$\canonize{}$ function for each QRE.
 \begin{figure}[t]
@@
 \begin{array}{l@\quad l @\quad l}
-\canonizer(R) &=& id_{\mathcal{L}(R)} \\
-\canonizer(\project{R}{s})(w) &=& s\\
-\canonizer(\squash{Q_1}{Q_2}{f})(w) &=&
+\canonize(\id(R)) &=& id_{\mathcal{L}(R)} \\
+\canonize(\project{R}{s})(w) &=& s\\
+\canonize(\squash{Q_1}{Q_2}{f})(w) &=&
 \begin{cases}
-\canonizer(Q_1)(f(w)) & \text{if } w \in \mathcal{L}(W(Q_1))\\
-\canonizer(Q_2)(w) & \text{otherwise}
+\canonize(Q_1)(f(w)) & \text{if } w \in \mathcal{L}(W(Q_1))\\
+\canonize(Q_2)(w) & \text{otherwise}
 \end{cases}\\
-\canonizer(\normalize{R}{R'}{f}) &=& f\\
-\canonizer(Q_1 \; ; \; Q_2) &=& \canonizer(Q_2) \circ \canonizer(Q_1)\\
-\canonizer(Q_1 \cdot Q_2) &=& \canonizer(Q_1) \cdot \canonizer(Q_2)\\
-\canonizer(Q_1 \sep Q_2)(w) &=&
+\canonize(\normalize{R}{R'}{f}) &=& f\\
+\canonize(Q_1 \; ; \; Q_2) &=& \canonize(Q_2) \circ \canonize(Q_1)\\
+\canonize(Q_1 \cdot Q_2) &=& \canonize(Q_1) \cdot \canonize(Q_2)\\
+\canonize(Q_1 \sep Q_2)(w) &=&
 \begin{cases}
-\canonizer(Q_1)(w) & \text{if } w \in \mathcal{L}(W(Q_1))\\
-\canonizer(Q_2)(w) & \text{if } w \in \mathcal{L}(W(Q_2))\\
+\canonize(Q_1)(w) & \text{if } w \in \mathcal{L}(W(Q_1))\\
+\canonize(Q_2)(w) & \text{if } w \in \mathcal{L}(W(Q_2))\\
 \end{cases}\\
-\canonizer(Q^*) &=& \canonizer(Q)^* \\
+\canonize(Q^*) &=& \canonize(Q)^* \\
 \end{array}
@@
 \begin{align*}
-&\canonizer(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
+&\canonize(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
 \cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)})\\
-&= (\canonizer(Q_1)(r_1)) \cdot (\canonizer(Q)(s_1)) \cdot \ldots \cdot
-(\canonizer(Q_n)(r_n))
+&= (\canonize(Q_1)(r_1)) \cdot (\canonize(Q)(s_1)) \cdot \ldots \cdot
+(\canonize(Q_n)(r_n))
 \end{align*}
@@
 definition,
-\begin{align*}\canonizer(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
+\begin{align*}\canonize(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
 \cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \\
-= (\canonizer(Q_1)(r_1)) \cdot (\canonizer(Q)(s_1)) \cdot \ldots \cdot
-(\canonizer(Q_n)(r_n))
+= (\canonize(Q_1)(r_1)) \cdot (\canonize(Q)(s_1)) \cdot \ldots \cdot
+(\canonize(Q_n)(r_n))
 \end{align*}
@@
 $Q_i, Q$ are placed in the canonical permutation which is the identity
-permutation before $\canonizer(Q_i)$ and $\canonizer(Q)$ are applied
+permutation before $\canonize(Q_i)$ and $\canonize(Q)$ are applied
 respectively.
@@
 \begin{array}{l@{}l@{\ \ }l}
-w \; \equiv_R \; w' &\iff& w = w' \\
+w \; \equiv_{\id(R)} \; w' &\iff& w = w' \\
 w \; \equiv_{\project{R}{s}} \; w' && \text{for all }w, w'\\
@@
 \end{figure}
-\subsection{Well Formed QREs}
+\subsection{Ambiguity and Well Formed QREs}
 \label{subsec:well-formed-qres}
-\subsubsection{Completeness of {\em normalize}}
-Before we present the full inference rules, for QREs, we discuss the {\em
-normalize} QRE which is special in two main ways, the first of which has to do
-with the premises in its inference rule:
-\begin{prooftree}
-\AxiomC{$\mathcal{L}(R') \subseteq \mathcal{L}(R)$}
-\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(R')$}
-\AxiomC{$f$ is surjective}
-\AxiomC{$f = f^2$}
-\RightLabel{(Normalize)}
-\QuaternaryInfC{$\wf{\normalize{R}{R'}{f}}$}
-\end{prooftree}
-In order for the {\em normalize} QRE to be well-formed, then canonizing function
-$f$ used by this combinator must be surjective and idempotent. Unfortunately,
-verifying that a function satisfies these properties is in general undecidable.
-Nonetheless, we allow a programmer to use normalize at their own risk without
-checking these side conditions.
-
-We do this firstly because the {\em normalize} combinator allows one to express
-equivalence relations that are either difficult to define, or that cannot be
-defined using the other combinators. Secondly, the {\em normalize} canonizer
-enjoys the property that each of the other combinators $Q$ is semantically
-equivalent to $\normalize{W(Q)}{K(Q)}{f}$ for some surjective, idempotent
-function $f : \mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$; this is the
-second property that distinguishes the {\em normalize} combinator from the other
-QREs. The {\em normalize} combinator hence gives a sufficient condition for
-potential QREs to be ``well-behaved'', particularly with respect to the
-strategy for synthesizing quotient lenses which we shall introduce in
-Section~\ref{synth}. This is the
-\subsubsection{Ambiguity}
-Another issue which we address before presenting the full set of inference
-rules for QREs, we discuss the issue of {\em ambiguity}, which appears in the
+An issue that we address before presenting the full set of inference
+rules for QREs is the issue of {\em ambiguity}, which appears in the
 premises of the inference rules for the regular combinators. If $Q_1$ and $Q_2$
@@
 would like to be able to generate quotient lenses automatically. One way to do
-that is to generate canonizers $\canonizer(Q_1)$ and $\canonizer(Q_2)$ from QREs
+that is to generate canonizers $\canonize(Q_1)$ and $\canonize(Q_2)$ from QREs
 $Q_1$ and $Q_2$ and then to synthesize bijective lenses between the kernel
@@
 
-$$\ell := \mathit{id} \; (R) \sep \const(s_1, s_2) \sep  \swap(\ell_1, \ell_2)
+$$\ell := \id(R) \sep \const(s_1, s_2) \sep  \swap(\ell_1, \ell_2)
 \sep \ell_1 \cdot \ell_2 \; |  \; (\ell_1 \sep \ell_2) \sep \ell^* \; | \;
@@
 $\mathit{\swap}(\ell_1, \ell_2)$ first undoes the \swap, then behaves like
-$cdot$. The $\ell_1 \; | \; \ell_2$ lens chooses to apply $\ell_1$ or $\ell_2$
-depending on whether the source (resp. target) data is matched by $\ell_1$ or
-$\ell_2$ in the forward (resp. backward direction). The $\ell^*$ lens splits the
-string $s$ into strings $s_1, \ldots, s_n$, applies $\ell_1$ to each $s_i$ to
-get $t_i$, and then concatenates each of the $t_i$'s for a final result of $t_1
-\cdot \ldots \cdot t_n$ in the forward direction. The iteration lens $\ell^*$
-operates similarly in the backward direction, but with $s, s_i$ substituted for
-$t, t_i$.
+the concatenation lens. The $\ell_1 \; | \; \ell_2$ lens chooses to apply
+$\ell_1$ or $\ell_2$ depending on whether the source (resp. target) data is
+matched by $\ell_1$ or $\ell_2$ in the forward (resp. backward direction). The
+$\ell^*$ lens splits the string $s$ into strings $s_1, \ldots, s_n$, applies
+$\ell_1$ to each $s_i$ to get $t_i$, and then concatenates each of the $t_i$'s
+for a final result of $t_1 \cdot \ldots \cdot t_n$ in the forward direction.
+The iteration lens $\ell^*$ operates similarly in the backward direction, but
+with $s, s_i$ substituted for $t, t_i$.
 
@@
 \begin{align*}
-\llbracket R \rrbracket &= \{(r, r) \sep r \in \mathcal{L}(R)\}\\
+\llbracket \id(R) \rrbracket &= \{(r, r) \sep r \in \mathcal{L}(R)\}\\
 \llbracket \const(r, s) \rrbracket &= \{(r, s)\}\\
@@
 $\varnothing$ with respect to $\cdot$, and that $R^* = \epsilon \; | \; R \cdot
-R* = \epsilon \; | \; R^* \cdot R$ for all $R$):
+R^* = \epsilon \; | \; R^* \cdot R$ for all $R$):
 
@@
 This rule makes it significantly more difficult to synthesize a bijective lens
-from its type.\sam{Should we explain why?}
+from its type.
 
@@
 grammar:
-$$ q := \kw {id}(R) \sep  \kw{const}(r, s) \sep q_1 \cdot q_2 \sep
-\kw{swap}(q_1, q_2) \sep (q_1 \sep q_2) \sep q^* \sep q_1 \; ; \; q_2 \sep
-\kw{lquot}(q, \ell) \sep \kw{rquot}(\ell, q)$$ where $R$ ranges over
-regular expressions and $\ell$ ranges over bijective lenses.
-
-TThe QRE lens combinators are the same as the bijective string lens
-combinators but with two extra combinators where quotienting actually
-occurs: tthe $\kw{lquot}$ and $\kw{rquot}$ combinators. The $\kw{lquot}(Q, q)$
-combinator takes a quotient lens $q$ and quotients the source data using $Q$,
-assuming that the source data forms a complete set of representatives for the
-equivalence relation $\eqrel{Q}$. The $\kw{lquot}(q, Q)$ combinator does the
-same but on the target data.
+$$ q := \lift(\ell) \sep  q_1 \cdot q_2 \sep \swap(q_1, q_2) \sep (q_1 \sep q_2)
+\sep q^* \sep q_1 \; ; \; q_2 \sep \lquot(q, \ell) \sep \rquot(\ell, q),$$ where
+$\ell$ ranges over bijective lenses.
+
+Other than the \kw{lift} combinator which allows a bijective lens to be
+treated as a quotient lens, the QRE lens combinators are the same as the
+bijective string lens combinators but with two extra combinators where
+quotienting actually occurs: the $\lquot$ and $\rquot$ combinators. The
+$\lquot(Q, q)$ combinator takes a quotient lens $q$ and quotients the source
+data using $Q$, assuming that the source data forms a complete set of
+representatives for the equivalence relation $\eqrel{Q}$. The $\lquot(q, Q)$
+combinator does the same but on the target data.
 
@@
 \begin{figure}[ht]
-TODO: Need to add some inferrence rules here.
 \centering
@@
 K(Q_1) = W(Q_2)}
-{\kw{lquot}(Q_1, q): Q_1 \; ; \; Q_2 \Leftrightarrow Q_3}{\mathit(Lquot)} &
+{\lquot(Q_1, q): Q_1 \; ; \; Q_2 \Leftrightarrow Q_3}{\mathit(Lquot)} &
 \begin{array}[b]{rcl}
-\llbracket \kw{lquot}(Q_1, q) \rrbracket.\get  &=& \llbracket q
-\rrbracket.\get \circ \canonizer(Q_1)\\
-\llbracket \kw{lquot}(Q_1, q) \rrbracket.\lput &=& \llbracket q
+\llbracket \lquot(Q_1, q) \rrbracket.\get  &=& \llbracket q
+\rrbracket.\get \circ \canonize(Q_1)\\
+\llbracket \lquot(Q_1, q) \rrbracket.\lput &=& \llbracket q
 \rrbracket.\lput
@@
 \RuleSide{q : Q_1 \Leftrightarrow Q_3 & \wf{Q_3} & W(Q_3) = K(Q_2)}
-{\kw{rquot}(q, Q_1):Q_1 \Leftrightarrow Q_2 \; ; \; Q_3}{\mathit{(Rquot)}} &
+{\rquot(q, Q_1):Q_1 \Leftrightarrow Q_2 \; ; \; Q_3}{\mathit{(Rquot)}} &
 \begin{array}[b]{rcl}
-\llbracket \kw{rquot}(q, Q_1) \rrbracket.\get  &=& \llbracket q
+\llbracket \rquot(q, Q_1) \rrbracket.\get  &=& \llbracket q
 \rrbracket.\get\\
-\llbracket \kw{rquot}(q, Q_1) \rrbracket.\lput &=& \llbracket q
-\rrbracket.\lput \circ \canonizer(Q_3)
+\llbracket \rquot(q, Q_1) \rrbracket.\lput &=& \llbracket q
+\rrbracket.\lput \circ \canonize(Q_3)
 \end{array}
@@
 \begin{array}[b]{l@{\qquad}c}
-\RuleSide{\substack{q_1 : Q_1 \Leftrightarrow d_1 \\ q_2 : Q_2 \Leftrightarrow
-d_2} & {\substack{W(Q_1) \cdot^! W(Q_2)\\ K(Q_1) \cdot^! K(Q_2)}}
-& {\substack{W(d_1) \cdot^! W(d_2)\\ K(d_1) \cdot^! K(d_2)}}}
-{q_1 \cdot q_2: Q_1 \cdot Q_2 \Leftrightarrow d_1 \cdot d_2}{\mathit{(Concat)}}
+\RuleSide{\substack{q_1 : Q_1 \Leftrightarrow Q_3 \\ q_2 : Q_2 \Leftrightarrow
+Q_4} & {\substack{W(Q_1) \cdot^! W(Q_2)\\ K(Q_1) \cdot^! K(Q_2)}}
+& {\substack{W(Q_3) \cdot^! W(Q_4)\\ K(Q_3) \cdot^! K(Q_4)}}}
+{q_1 \cdot q_2: Q_1 \cdot Q_2 \Leftrightarrow Q_3 \cdot Q_4}{\mathit{(Concat)}}
 &
@@
 \begin{array}[b]{l@{\qquad}c}
-\RuleSide{\substack{q_1 : Q_1 \Leftrightarrow d_1 \\ q_2 : Q_2 \Leftrightarrow
-d_2} & \substack{\mathcal{L}(W(Q_1)) \cap \mathcal{L}(W(Q_2)) = \varnothing
-\\ \mathcal{L}(W(d_1)) \cap \mathcal{L}(W(d_2)) = \varnothing }}
-{q_1 \sep q_2: (Q_1 \sep Q_2)\Leftrightarrow (d_1 \sep d_2)}{\mathit{(Or)}} &
+\RuleSide{\substack{q_1 : Q_1 \Leftrightarrow Q_3 \\ q_2 : Q_2 \Leftrightarrow
+Q_4} & {\substack{W(Q_1) \cdot^! W(Q_2)\\ K(Q_1) \cdot^! K(Q_2)}}
+& {\substack{W(Q_4) \cdot^! W(Q_3)\\ K(Q_4) \cdot^! K(Q_3)}}}
+{q= q_1 \cdot q_2: Q_1 \cdot Q_2 \Leftrightarrow Q_4 \cdot
+Q_3}{\mathit{(Swap)}} &
+\begin{array}[b]{rcl}
+\llbracket q \rrbracket.\get(s_1 \cdot s_2)  &=& \llbracket q_2
+\rrbracket.\get(s_2) \cdot \llbracket q_1 \rrbracket.\get(s_1)\\
+\llbracket q \rrbracket.\lput(t_2 \cdot t_1) &=&
+\llbracket q_1 \rrbracket.\lput(t_1) \cdot \llbracket q_2 \rrbracket.\lput(t_2)
+\end{array}
+\end{array}
+\]
+
+\[
+\begin{array}[b]{l@{\qquad}c}
+\RuleSide{\substack{q_1 : Q_1 \Leftrightarrow Q_3 \\ q_2 : Q_2 \Leftrightarrow
+Q_4} & \substack{\mathcal{L}(W(Q_1)) \cap \mathcal{L}(W(Q_2)) = \varnothing
+\\ \mathcal{L}(W(Q_3)) \cap \mathcal{L}(W(Q_4)) = \varnothing }}
+{q_1 \sep q_2: (Q_1 \sep Q_2)\Leftrightarrow (Q_3 \sep Q_4)}{\mathit{(Or)}} &
 \begin{array}[b]{rcl}
@@
 \begin{cases}
-\llbracket q_1 \rrbracket.\lput (s) & \text{if } s \in \mathcal{L}(W(d_1))\\
-\llbracket q_2 \rrbracket.\lput (s) & \text{if } s \in \mathcal{L}(W(d_2))\\
+\llbracket q_1 \rrbracket.\lput (s) & \text{if } s \in \mathcal{L}(W(Q_3))\\
+\llbracket q_2 \rrbracket.\lput (s) & \text{if } s \in \mathcal{L}(W(Q_4))\\
 \end{cases}
@@
 K(Q_3)}$}
-\AxiomC{$\canonizer(Q_2) = \canonizer(Q_3)$}
+\AxiomC{$\canonize(Q_2) = \canonize(Q_3)$}
 \TrinaryInfC{$q_1 \; ; \; q_2: Q_1 \Leftrightarrow Q_4$}
@@
 equivalent modulo the star-semiring axioms. Finally, the condition
-$\canonizer(Q_2) = \canonizer(Q_3)$ says that $Q_2$ and $Q_3$ define the same
+$\canonize(Q_2) = \canonize(Q_3)$ says that $Q_2$ and $Q_3$ define the same
 equivalence relation on $\mathcal{L}(W(Q_2))= \mathcal{L}(W(Q_3))$.
@@
 \begin{align*}
-\llbracket q \rrbracket.\get &= \ell \circ \canonizer(Q_1)\\
-\llbracket q \rrbracket.\lput &= \ell^{-1} \circ \canonizer(Q_2)
+\llbracket q \rrbracket.\get &= \ell \circ \canonize(Q_1)\\
+\llbracket q \rrbracket.\lput &= \ell^{-1} \circ \canonize(Q_2)
 \end{align*}
@@
 \llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket\circ
-\canonizer(Q_1)\\
+\canonize(Q_1)\\
 \llbracket q \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
-\canonizer(Q_2)
+\canonize(Q_2)
 \end{align*}
@@
 $\llbracket q_1 \rrbracket.\get = \llbracket \ell_1 \rrbracket \circ
-\canonizer(Q_1)$ & $\llbracket q_2 \rrbracket.\get = \llbracket \ell_2
-\rrbracket \circ \canonizer(Q_3)$ \\
+\canonize(Q_1)$ & $\llbracket q_2 \rrbracket.\get = \llbracket \ell_2
+\rrbracket \circ \canonize(Q_3)$ \\
 $\llbracket q_1 \rrbracket.\lput = {\llbracket \ell_1 \rrbracket}^{-1} \circ
-\canonizer(Q_2)$ & $\llbracket q_2 \rrbracket.\lput = {\llbracket \ell_2
-\rrbracket}^{-1} \circ \canonizer(Q_4)$
+\canonize(Q_2)$ & $\llbracket q_2 \rrbracket.\lput = {\llbracket \ell_2
+\rrbracket}^{-1} \circ \canonize(Q_4)$
 \end{tabular}
@@
 \llbracket q_2 \rrbracket.\get \circ \llbracket q_1 \rrbracket.\get &=
-(\llbracket \ell_2 \rrbracket \circ \canonizer(Q_3)) \circ (\llbracket \ell_1
-\rrbracket \circ \canonizer(Q_1))\\
-&= \llbracket \ell_2 \rrbracket \circ (\canonizer(Q_3) \circ \llbracket \ell_1
-\rrbracket) \circ \canonizer(Q_1)\\
+(\llbracket \ell_2 \rrbracket \circ \canonize(Q_3)) \circ (\llbracket \ell_1
+\rrbracket \circ \canonize(Q_1))\\
+&= \llbracket \ell_2 \rrbracket \circ (\canonize(Q_3) \circ \llbracket \ell_1
+\rrbracket) \circ \canonize(Q_1)\\
 &= (\llbracket \ell_2 \rrbracket \circ \llbracket \ell_1 \rrbracket) \circ
-\canonizer(Q_1)\\
-&= \llbracket \ell_1 \; ; \; \ell_2 \rrbracket \circ \canonizer(Q_1)
+\canonize(Q_1)\\
+&= \llbracket \ell_1 \; ; \; \ell_2 \rrbracket \circ \canonize(Q_1)
 \end{align*}
 We are permitted to claim the third step from the second since
-$\canonizer(Q_3)$ is the identity function on $K(Q_2)$ which is
+$\canonize(Q_3)$ is the identity function on $K(Q_2)$ which is
 syntactically equal to $K(Q_3)$ by assumption. A similar argument shows that
@@
 \llbracket \ell_1 \; ; \; \ell_2 \rrbracket^{-1} \circ
-\canonizer(Q_4)$$
+\canonize(Q_4)$$
 
@@
 languages.  This simple strategy turns out to be remarkably effective, and the
-idea of using Optician in this way is part of the impetus\bcp{What was the
-  other part??} for the design of our QRE lenses.
+idea of using Optician in this way inspired the design of our QRE lenses.
 
@@
 QRE lens, then there exists a bijective lens $\ell : K(Q_1) \Leftrightarrow
-K(Q_2)$ such that
-\begin{align*}
-\llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket\circ \canonizer(Q_1)\\
+K(Q_2)$ such that:
+\begin{align*}
+\llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket\circ \canonize(Q_1)\\
 \llbracket q \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
-\canonizer(Q_2).
-\end{align*}
-\bcp{This math display and the previous one are punctuated perfectly: In the
-first, the colon replaces a period at the end of the sentence (so no need
-for another period at the end of the display).  In the second, the display
-is punctuated as if it were inline.  Please go through and make sure all the
-other displayed material follows one or the other of these rules.}
-
-For the\bcp{why ``the''?  Are we talking about some set of QRE examples
-  already?} QRE examples, the $x_i$'s are in $W(Q_1)$ and the $y_i$'s are 
-in $W(Q_2)$, so we can construct ${x'_i} = \canonizer(Q_1)(x_i)$ in $K(Q_1)$
-and ${y'_i} = \canonizer(Q_2)(y_i)$ is in $K(Q_2)$. To synthesize the
-desired quotient lens $q: Q_1 \Leftrightarrow Q_2$ that is consistent with the
+\canonize(Q_2)
+\end{align*}
+For the examples, the $x_i$'s are in $W(Q_1)$ and the $y_i$'s are 
+in $W(Q_2)$, so we can construct ${x'_i} = \canonize(Q_1)(x_i)$ in $K(Q_1)$
+and ${y'_i} = \canonize(Q_2)(y_i)$ is in $K(Q_2)$. To synthesize the desired
+quotient lens $q: Q_1 \Leftrightarrow Q_2$ that is consistent with the
 input-output examples $\{(x_1, y_1), \ldots, (x_n, y_n)\}$ it suffices to
@@
   \item
-  Compute $K(Q_1)$, $K(Q_2)$,
+  compute $K(Q_1)$, $K(Q_2)$,
   \item
-  Compute $\{({x'_1}, {y'_1}), \ldots, ({x'_n}, {y'_n})\}$ where ${x'_i} =
-  \canonizer(Q_1)(x_i)$ and ${y'_i} = \canonizer(Q_2)(y_i)$,
-\item Synthesize, using Optician's algorithm, a bijective lens
+  compute $\{({x'_1}, {y'_1}), \ldots, ({x'_n}, {y'_n})\}$ where ${x'_i} =
+  \canonize(Q_1)(x_i)$ and ${y'_i} = \canonize(Q_2)(y_i)$,
+\item synthesize, using Optician's algorithm, a bijective lens
   $\ell : K(Q_1) \Leftrightarrow K(Q_2)$ that is consistent with the examples
-  $\{({x'_1}, {y'_1}),\allowbreak \ldots,\allowbreak ({x'_n}, {y'_n})\}$
+  $\{({x'_1}, {y'_1}),\allowbreak \ldots,\allowbreak ({x'_n}, {y'_n})\}$,
   \item
-  Return $q = \kw{rquot}(\kw{lquot}(Q_1, \ell), Q_2) : Q_1 \Leftrightarrow
-  Q_2$ \bcp{punctuate this enumeration as if it were running text: ``, and''
-  at the end of (3) and period here.  And lowercase letters at the
-  beginnings of the items.}
+  return $q = \rquot(\lquot(Q_1, \ell), Q_2) : Q_1 \Leftrightarrow
+  Q_2$.
 \end{enumerate}
-
-\saz{Do we want to state formally the completeness of this algorithm?}\BCP{+1}
+This algorithm is complete in the following sense: given any QREs $Q_1$ and
+$Q_2$ and set of examples $\{(x_1, y_2), \ldots, (x_n, y_n)\}$ with $x_i \in
+W(Q_1)$ and $y_i \in W(Q_2)$, then if there exists a derivation of a QRE lens
+$q : Q_1 \Leftrightarrow Q_2$ such that 
+\begin{equation}\label{qrelensequation}
+q.\get(x_i) \equiv_{Q_2} y_i \text{ and } q.\lput(y_i) \equiv_{Q_1} x_i \text{
+for all } i,
+\end{equation}
+then the algorithm will return a QRE lens $q': Q_1 \Leftrightarrow Q_2$ that
+also satisfies \cref{qrelensequation}. The justification for this fact follows
+from the completeness of the Optician algorithm and \cref{normal form}.
 
@@
   specifications (including examples)\saz{CHECK THIS}\afm{no but i'll add it in}
-  \item[\CanonizerAndSpecSize{}]  the number of
-  AST nodes in $W(q_1)$ summed with the number of AST nodes in $\canonizer(q_1)$
+  \item[\canonizeAndSpecSize{}]  the number of
+  AST nodes in $W(q_1)$ summed with the number of AST nodes in $\canonize(q_1)$
   \item[\LensAndSpecSize{}] the number of AST nodes
@@
 
-The value of \CanonizerAndSpecSize{} estimates the burden for writing our
+The value of \canonizeAndSpecSize{} estimates the burden for writing our
 benchmarks within the Optician system, without QREs but still using
-synthesis. (This is only an approximation, as both $W(q_1)$ and $\canonizer(q_1)$ are
+synthesis. (This is only an approximation, as both $W(q_1)$ and $\canonize(q_1)$ are
 automatically generated from the QRE, and it is possible that a human-written
@@
 in Boomerang, using neither QREs nor synthesis. (This is also an
-approximation, as $W(q_1)$ and $\canonizer(q_1)$ and the bijective lens are all
+approximation, as $W(q_1)$ and $\canonize(q_1)$ and the bijective lens are all
 automatically generated.)
@@
 at most a given number of AST nodes. For a given benchmark, we found \QRESize{}
-to be about half as large as \CanonizerAndSpecSize, and about a third as large
+to be about half as large as \canonizeAndSpecSize, and about a third as large
 as \LensAndSpecSize{}.\BCP{These are really the critical conclusions, and they
@@
 wiht the symmetric $\kw{dup}_2$ combinator discarding the first copy instead of
-the second in the \lput/\create direction.
+the second in the \lput/\create{} direction.
 
@@
 to be used as canonizers by using the taking the \kw{canonize} function to be
-the \get component of a lens and the \kw{choose} function to be its \create
+the \get{} component of a lens and the \kw{choose} function to be its \create{}
 component. Naturally, QREs also take advantage of this ability to use lenses as
@@
 {\em collect} function looks up precisely these names.
-Rajkumar et. al.~\cite{rajkumar2014lenses} observe that formlets are a useful
+Rajkumar et. al. observe that formlets are a useful
 abstraction, but they only address one half of the problem---they make it easy
@@
 Now we turn to related work in synthesis. Much of the research in synthesis
-assumes that the synthesizer is provided with a collection of examples. Optician
-and Optometrist\bcp{Sole use?!} differ in that they require that the
-programmer supplies both 
-examples {\em and} format descriptions in the form of regular expressions or
-QREs, though we are far from the only ones to consider type-based synthesis.
+assumes that the synthesizer is provided with a collection of examples.
+Optician, as well as Optician enhanced with QREs for QRE lens synthesis, differ
+in that they require that the programmer supplies both examples {\em and}
+format descriptions in the form of regular expressions or QREs, though we are
+far from the only ones to consider type-based synthesis.
 There is a trade-off here.  On the one hand, a user must have some programming
@@
 
-\section{Appendix}
-\bcp{Appendices should be numbered differently (with letters, probably) and
-  belong after the references.}
-\sam{Also need to add a few cases here.}
-We give a full proof of \cref{normal form}: If there is a derivation $q : Q_1
+\appendix
+\addcontentsline{toc}{section}{Appendices}
+\section*{Appendices}
+\section{Proof of Normal Form Theorem for QRE lenses}
+\cref{normal form} claims that if there is a derivation $q : Q_1
 \Leftrightarrow Q_2$, then there exists a bijective lens $\ell : K(Q_1)
@@
 \llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket\circ
-\canonizer(Q_1)\\
+\canonize(Q_1)\\
 \llbracket q \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
-\canonizer(Q_2)
-\end{align*}
+\canonize(Q_2)
+\end{align*}
+We now prove this theorem.
 \begin{proof}
@@
 = \llbracket \ell \rrbracket \circ id_{\mathcal{L}(R)} =
-\llbracket \ell \rrbracket \circ \canonizer(\mathit{id}(R)), \text{ and }\\
+\llbracket \ell \rrbracket \circ \canonize(\mathit{id}(R)), \text{ and }\\
 \llbracket \kw{lift}(\ell) \rrbracket.\lput &= \llbracket \ell
 \rrbracket^{-1} = \llbracket \ell \rrbracket^{-1} \circ id_{\mathcal{L}(S)} =
-\llbracket \ell \rrbracket^{-1} \circ \canonizer(id(S))
+\llbracket \ell \rrbracket^{-1} \circ \canonize(id(S))
 \end{align*}
 \item
-$\kw{lquot}(Q_1, q): Q_1 \; ; \; Q_2 \Leftrightarrow Q_3$ where $q : Q_2
+$\lquot(Q_1, q): Q_1 \; ; \; Q_2 \Leftrightarrow Q_3$ where $q : Q_2
 \Leftrightarrow Q_3$, $Q_1$ is well formed and $K(Q_1) = W(Q_2)$. Then
 \begin{align*}
-\llbracket \kw{lquot}(Q_1, q) \rrbracket.\get  &= \llbracket q
-\rrbracket.\get \circ \canonizer(Q_1)\\
-\llbracket \kw{lquot}(Q_1, q) \rrbracket.\lput &= \llbracket q \rrbracket.\lput
+\llbracket \lquot(Q_1, q) \rrbracket.\get  &= \llbracket q
+\rrbracket.\get \circ \canonize(Q_1)\\
+\llbracket \lquot(Q_1, q) \rrbracket.\lput &= \llbracket q \rrbracket.\lput
 \end{align*}
@@
 \llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket \circ
-\canonizer(Q_2)\\
+\canonize(Q_2)\\
 \llbracket q \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
-\canonizer(Q_3)
+\canonize(Q_3)
 \end{align*}
@@
 \begin{align*}
-\llbracket \kw{lquot}(Q_1, q)\rrbracket.\get  &= (\llbracket \ell \rrbracket \circ
-\canonizer(Q_2)) \circ \canonizer(Q_1) = \llbracket \ell \rrbracket \circ
-(\canonizer(Q_1 \; ; \; Q_2))\\
-\llbracket \kw{lquot}(Q_1, q) \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
-\canonizer(Q_3)
+\llbracket \lquot(Q_1, q)\rrbracket.\get  &= (\llbracket \ell \rrbracket \circ
+\canonize(Q_2)) \circ \canonize(Q_1) = \llbracket \ell \rrbracket \circ
+(\canonize(Q_1 \; ; \; Q_2))\\
+\llbracket \lquot(Q_1, q) \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
+\canonize(Q_3)
 \end{align*}
@@
 \item
-$\kw{rquot}(q, Q_3):Q_1 \Leftrightarrow Q_2 \; ; \; Q_3$ where $q : Q_1
+$\rquot(q, Q_3):Q_1 \Leftrightarrow Q_2 \; ; \; Q_3$ where $q : Q_1
 \Leftrightarrow Q_2$, $Q_3$ is well formed and $K(Q_3) = W(Q_2)$. Proceed as in
@@
 \begin{align*}
-\llbracket q_1 \rrbracket.\get &= \llbracket q_2 \rrbracket.\get\circ \llbracket
-q_1 \rrbracket.\get, \text{ and }\\
-\llbracket q_1 \rrbracket.\lput &= \llbracket q_1 \rrbracket.\lput \circ \llbracket
-q_2 \rrbracket.\lput
+\llbracket q_1 \; ; \; q_2 \rrbracket.\get &= \llbracket q_2
+\rrbracket.\get\circ \llbracket q_1 \rrbracket.\get, \text{ and }\\
+\llbracket q_1 \; ; \; q_2 \rrbracket.\lput &= \llbracket q_1 \rrbracket.\lput
+\circ \llbracket q_2 \rrbracket.\lput
 \end{align*}
@@
 \llbracket q_1 \rrbracket.\get &= \llbracket \ell_1 \rrbracket \circ
-\canonizer(c)\\
+\canonize(c)\\
 \llbracket q_1 \rrbracket.\lput &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
-\canonizer(Q_1)
+\canonize(Q_1)
 \end{align*}
@@
 \llbracket q_2 \rrbracket.\get &= \llbracket \ell_2 \rrbracket \circ
-\canonizer(Q_1)\\
+\canonize(Q_1)\\
 \llbracket q_2 \rrbracket.\lput &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
-\canonizer(Q_2)
+\canonize(Q_2)
 \end{align*}
@@
 \begin{align*}
-\llbracket q_2 \rrbracket.\get \circ \llbracket q_1 \rrbracket.\get &=
-(\llbracket \ell_2 \rrbracket \circ \canonizer(Q_1)) \circ (\llbracket \ell_1
-\rrbracket \circ \canonizer(c))\\
-&= \llbracket \ell_2 \rrbracket \circ (\canonizer(Q_1) \circ \llbracket \ell_1
-\rrbracket) \circ \canonizer(c)\\
+\llbracket q_1 \; ; \; q_2 \rrbracket.\get &=
+\llbracket q_2 \rrbracket.\get \circ \llbracket q_1 \rrbracket.\get \\
+&=(\llbracket \ell_2 \rrbracket \circ \canonize(Q_1)) \circ (\llbracket \ell_1
+\rrbracket \circ \canonize(c))\\
+&= \llbracket \ell_2 \rrbracket \circ (\canonize(Q_1) \circ \llbracket \ell_1
+\rrbracket) \circ \canonize(c)\\
 &= (\llbracket \ell_2 \rrbracket \circ \llbracket \ell_1 \rrbracket) \circ
-\canonizer(c)\\
+\canonize(c)\\
 &= \llbracket \ell_1 \; ; \; \ell_2 \rrbracket \circ
-\canonizer(c)
+\canonize(c)
 \end{align*}
 A similar argument shows that
-$$\llbracket q_1 \rrbracket.\lput \circ \llbracket q_2 \rrbracket.\lput =
+$$\llbracket q_1 \; ; \; q_2 \rrbracket.\lput =
 \llbracket \ell_1 \; ; \; \ell_2 \rrbracket^{-1} \circ
-\canonizer(c)$$
+\canonize(c)$$
 \item
@@
 \llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket \circ
-\canonizer(Q_1)\\
+\canonize(Q_1)\\
 \llbracket q \rrbracket.\lput &= {\llbracket \ell \rrbracket}^{-1} \circ
-\canonizer(Q_2)
+\canonize(Q_2)
 \end{align*}
@@
 \llbracket q^* \rrbracket.\get &= (\llbracket \ell \rrbracket \circ
-\canonizer(Q_1))^* = \llbracket \ell \rrbracket^* \circ
-\canonizer(Q_1)^* = \llbracket \ell^* \rrbracket \circ
-\canonizer(Q_1^*)\\
+\canonize(Q_1))^* = \llbracket \ell \rrbracket^* \circ
+\canonize(Q_1)^* = \llbracket \ell^* \rrbracket \circ
+\canonize(Q_1^*)\\
 \llbracket q^* \rrbracket.\lput &= (\llbracket \ell \rrbracket^{-1} \circ
-\canonizer(Q_2))^* = (\llbracket \ell \rrbracket^{-1})^* \circ
-\canonizer(Q_2)^* = \llbracket \ell^* \rrbracket^{-1} \circ
-\canonizer(Q_2^*)\\
+\canonize(Q_2))^* = (\llbracket \ell \rrbracket^{-1})^* \circ
+\canonize(Q_2)^* = \llbracket \ell^* \rrbracket^{-1} \circ
+\canonize(Q_2^*)\\
 \end{align*}
 \item
-$q_1 \cdot q_2: Q_1 \cdot Q_2 \Leftrightarrow d_1 \cdot d_2$, where $q_1 : Q_1
-\Leftrightarrow d_1 $,  $q_2 : Q_2 \Leftrightarrow d_2$, $W(Q_1)
-\cdot^! W(Q_2)$, $K(Q_1) \cdot^! K(Q_2)$, $W(d_1) \cdot^! W(d_2)$ and $
-K(d_1) \cdot^! K(d_2)$. Then
-\begin{align*}
-\llbracket q_1 \rrbracket.\get &= \llbracket q_1 \rrbracket.\get \cdot \llbracket
-q_2 \rrbracket.\get, \text{ and }\\
-\llbracket q_1 \rrbracket.\lput &= \llbracket q_1 \rrbracket.\lput \cdot \llbracket
-q_2 \rrbracket.\lput
+$q_1 \cdot q_2: Q_1 \cdot Q_2 \Leftrightarrow Q_3 \cdot Q_4$, where $q_1 : Q_1
+\Leftrightarrow Q_3 $,  $q_2 : Q_2 \Leftrightarrow Q_4$, $W(Q_1)
+\cdot^! W(Q_2)$, $K(Q_1) \cdot^! K(Q_2)$, $W(Q_3) \cdot^! W(Q_4)$ and $
+K(Q_3) \cdot^! K(Q_4)$. Then
+\begin{align*}
+\llbracket q_1 \cdot q_2 \rrbracket.\get &= \llbracket q_1 \rrbracket.\get \cdot
+\llbracket q_2 \rrbracket.\get, \text{ and }\\
+\llbracket q_1 \cdot q_2 \rrbracket.\lput &= \llbracket q_1 \rrbracket.\lput
+\cdot \llbracket q_2 \rrbracket.\lput
 \end{align*}
 By the induction hypothesis, there exist bijective lenses $\ell_1 : K(Q_1)
-\Leftrightarrow K(d_1)$ and $\ell_2 : K(Q_2) \Leftrightarrow K(d_2)$ such that
-\begin{align*}
-\llbracket q_1 \rrbracket.\get &= \llbracket \ell_1 \rrbracket \circ
-\canonizer(Q_1)\\
-\llbracket q_1 \rrbracket.\lput &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
-\canonizer(d_1)
+\Leftrightarrow K(Q_3)$ and $\ell_2 : K(Q_2) \Leftrightarrow K(Q_4)$ such that
+\begin{align*}
+\llbracket q \rrbracket.\get &= \llbracket \ell_1 \rrbracket \circ
+\canonize(Q_1)\\
+\llbracket q \rrbracket.\lput &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
+\canonize(Q_3)
 \end{align*}
@@
 \llbracket q_2 \rrbracket.\get &= \llbracket \ell_2 \rrbracket \circ
-\canonizer(Q_2)\\
+\canonize(Q_2)\\
 \llbracket q_2 \rrbracket.\lput &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
-\canonizer(d_2)
+\canonize(Q_4)
 \end{align*}
@@
 \begin{align*}
-\llbracket q_1 \rrbracket.\get &= (\llbracket \ell_1 \rrbracket \circ
-\canonizer(Q_1)) \cdot  (\llbracket \ell_2 \rrbracket \circ
-\canonizer(Q_2))\\
+\llbracket q_1 \cdot q_2 \rrbracket.\get &= (\llbracket \ell_1 \rrbracket \circ
+\canonize(Q_1)) \cdot  (\llbracket \ell_2 \rrbracket \circ
+\canonize(Q_2))\\
 &= (\llbracket \ell_1 \rrbracket \cdot \llbracket \ell_2
-\rrbracket) \circ (\canonizer(Q_1) \cdot \canonizer(Q_2))\\
-&= \llbracket \ell_1 \cdot  \ell_2 \rrbracket \circ \canonizer(Q_1 \cdot Q_2)
+\rrbracket) \circ (\canonize(Q_1) \cdot \canonize(Q_2))\\
+&= \llbracket \ell_1 \cdot  \ell_2 \rrbracket \circ \canonize(Q_1 \cdot Q_2)
 \end{align*}
@@
 $$
-\llbracket q_1 \rrbracket.\lput = \llbracket \ell_1 \cdot  \ell_2 \rrbracket^{-1}
-\circ \canonizer(d_1 \cdot d_2) $$
+\llbracket q_1 \cdot q_2\rrbracket.\lput = \llbracket \ell_1 \cdot  \ell_2
+\rrbracket^{-1} \circ \canonize(Q_3 \cdot Q_4) $$
 \item
-$q_1 = q_1 \sep q_2$ where $q_1 : Q_1 \Leftrightarrow d_1 $, $q_2 : Q_2
-\Leftrightarrow d_2$, $\mathcal{L}(W(Q_1)) \cap \mathcal{L}(W(Q_2)) =
-\varnothing$ and $\mathcal{L}(W(d_1)) \cap \mathcal{L}(W(d_2)) = \varnothing$.
+$\swap(q_1,q_2): Q_1 \cdot Q_2 \Leftrightarrow Q_4 \cdot Q_3$, where $q_1 : Q_1
+\Leftrightarrow Q_3 $,  $q_2 : Q_2 \Leftrightarrow Q_4$, $W(Q_1)
+\cdot^! W(Q_2)$, $K(Q_1) \cdot^! K(Q_2)$, $W(Q_4) \cdot^! W(Q_3)$ and $
+K(Q_4) \cdot^! K(Q_3)$. Then
+\begin{align*}
+\llbracket \swap(q_1, q_2) \rrbracket.\get(s_1 \cdot s_2) &= \llbracket q_2
+\rrbracket.\get(s_2) \cdot \llbracket q_1 \rrbracket.\get(s_1), \text{ and }\\
+\llbracket \swap(q_1, q_2) \rrbracket.\lput(t_1, t_2) &= \llbracket q_1
+\rrbracket.\lput(t_1) \cdot \llbracket q_2 \rrbracket.\lput(t_2)
+\end{align*}
+By the induction hypothesis, there exist bijective lenses $\ell_1 : K(Q_1)
+\Leftrightarrow K(Q_3)$ and $\ell_2 : K(Q_2) \Leftrightarrow K(Q_4)$ such that
+\begin{align*}
+\llbracket q_1 \rrbracket.\get &= \llbracket \ell_1 \rrbracket \circ
+\canonize(Q_1)\\
+\llbracket q_1 \rrbracket.\lput &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
+\canonize(Q_3)
+\end{align*}
+and
+\begin{align*}
+\llbracket q_2 \rrbracket.\get &= \llbracket \ell_2 \rrbracket \circ
+\canonize(Q_2)\\
+\llbracket q_2 \rrbracket.\lput &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
+\canonize(Q_4)
+\end{align*}
+Consequently,
+\begin{align*}
+\llbracket \swap(q_1, q_2) \rrbracket.\get(s_1 \cdot s_2) &= \llbracket q_2
+\rrbracket.\get(s_2) \cdot \llbracket q_1 \rrbracket.\get(s_1)\\
+&= (\llbracket \ell_2 \rrbracket \circ
+\canonize(Q_2))(s_2) \cdot  (\llbracket \ell_1 \rrbracket \circ
+\canonize(Q_1))(s_1)\\
+&= (\llbracket \swap(\ell_1, \ell_2) \rrbracket) \circ (\canonize(Q_1) \cdot
+\canonize(Q_2)) (s_1, s_2)
+\end{align*}
+Similarly
+$$
+\llbracket \swap(q_1, q_2) \rrbracket.\lput = (\llbracket \swap(\ell_1, \ell_2)
+\rrbracket)^{-1} \circ (\canonize(Q_4) \cdot \canonize(Q_3))$$
+\item
+$q_1 = q_1 \sep q_2$ where $q_1 : Q_1 \Leftrightarrow Q_3 $, $q_2 : Q_2
+\Leftrightarrow Q_4$, $\mathcal{L}(W(Q_1)) \cap \mathcal{L}(W(Q_2)) =
+\varnothing$ and $\mathcal{L}(W(Q_3)) \cap \mathcal{L}(W(Q_4)) = \varnothing$.
 Then
@@
 By the induction hypothesis, there exist bijective lenses $\ell_1 : K(Q_1)
-\Leftrightarrow K(d_1)$ and $\ell_2 : K(Q_2) \Leftrightarrow K(d_2)$ such that
+\Leftrightarrow K(Q_3)$ and $\ell_2 : K(Q_2) \Leftrightarrow K(Q_4)$ such that
 \begin{align*}
 \llbracket q_1 \rrbracket.\get &= \llbracket \ell_1 \rrbracket \circ
-\canonizer(Q_1)\\
+\canonize(Q_1)\\
 \llbracket q_1 \rrbracket.\lput &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
-\canonizer(d_1)
+\canonize(Q_3)
 \end{align*}
@@
 \llbracket \ell_1 \rrbracket \circ
-\canonizer(Q_1) (s) & \text{if } s \in \mathcal{L}(W(Q_1))\\
+\canonize(Q_1) (s) & \text{if } s \in \mathcal{L}(W(Q_1))\\
 \llbracket \ell_2 \rrbracket \circ
-\canonizer(Q_2) (s) & \text{if } s \in \mathcal{L}(W(Q_2)),\\
+\canonize(Q_2) (s) & \text{if } s \in \mathcal{L}(W(Q_2)),\\
 \end{cases}$$
 so $\llbracket q_1 \sep q_2 \rrbracket.\get = \llbracket \ell_1 \sep
-\ell_2 \rrbracket \circ \canonizer(Q_1 \sep Q_2)$. A similar argument shows
+\ell_2 \rrbracket \circ \canonize(Q_1 \sep Q_2)$. A similar argument shows
 that $\llbracket q_1 \sep q_2 \rrbracket.\lput = \llbracket \ell_1 \sep
-\ell_2 \rrbracket^{-1} \circ \canonizer(d_1 \sep d_2)$.\\
+\ell_2 \rrbracket^{-1} \circ \canonize(Q_3 \sep Q_4)$.\\
 This completes the proof. 

### 2018-03-13 papers/QuotientLenses/main.tex
fixed eval issues

--- 2018-03-12 papers/QuotientLenses/main.tex
@@
 synthesize QRE lenses between various real-world data formats in
-the {\tt data.gov} database.
+the Optician benchmark suite.
 
@@
   our approach by synthesizing QRE lenses between
-  data formats drawn from the {\tt data.gov} database (Section~\ref{impl}).
+  data formats drawn from the Optician benchmark suite (Section~\ref{impl}).
 \end{enumerate}
@@
 the new QRE support, we were able to remove these alterations when porting these
-benchmarks. (This experience alone suggests that QREs make the lens development
-process easier.)
+benchmarks. 10 of Optician's 39 benchmarks had formats modified in this way ---
+introducing QREs removed these format alterations. (This experience alone
+suggests that QREs make the lens development process easier.)
 
@@
   \item[\QRESize{}] the number of AST nodes in the QRE
-  specifications (including examples)\saz{CHECK THIS}\afm{no but i'll add it in}
+  specifications (including examples).
   \item[\CanonizerAndSpecSize{}]  the number of
   AST nodes in $W(q_1)$ summed with the number of AST nodes in $\canonizer(q_1)$
+  and the number of examples.
   \item[\LensAndSpecSize{}] the number of AST nodes
@@
 Figure~\ref{fig:asts} shows the number of benchmarks that can be expressed using
-at most a given number of AST nodes. For a given benchmark, we found \QRESize{}
-to be about half as large as \CanonizerAndSpecSize, and about a third as large
-as \LensAndSpecSize{}.\BCP{These are really the critical conclusions, and they
-  are hard to see, the way the graph is drawn. I wonder whether we could instead
-  present with the individual examples on the X axis (sorted by number of AST
-  nodes), with a bar above each one containing three regions, colored
-  differently, for LS, CS-LS, and QC-CS...}\afm{will do} This suggests that
-introducing QREs saves programmers significant effort compared to both Optician
-and basic Boomerang.
+at most a given number of AST nodes. On average, \CanonizerAndSpecSize{} used
+44\% more AST nodes than \QRESize{}, using an average of 237 more AST nodes. On
+average, \LensAndSpecSize{} used 193\% more AST nodes than \QRESize{}, using an
+average of 1267 more AST nodes. This suggests that introducing QREs saves
+programmers significant effort compared to both pre-QRE Optician and basic
+Boomerang.
 
@@
   performance overhead incurred by using QREs. In (b), we run Optician with QREs
-  on its benchmark suite. We find that it is able to synthesize all quotient
-  lenses in under 20 seconds, and typically finished in under 5 seconds.
-  \bcp{The left-hand graph, especially, is hard to make anything out of, because
-    the curves are squished against the axes and have almost identical shapes.
-    }\afm{will fix}}
+  on the Optician benchmarks we modified to use QREs. We find that it is able to
+  synthesize all quotient lenses in under 20 seconds, and typically finished in
+  under 5 seconds.
 \label{fig:times}
@@
   \item[\OpticianRuntime{}] pre-QRE Optician running on its own benchmarks.
-  \item[\SystemOnOptician{}] Optician with QREs running on the pre-QRE benchmarks.
-  \item[\SystemOnBenchmarks{}] Optician with QREs running on its own
-  benchmarks. \BCP{What is the difference between QO and QQ??  The eight
-    additional benchmarks?  Why not measure just these?  Seems like this
-    graphj repeats some information from the one to the left, obscuring the
-    new information.}\afm{will do}
+  \item[\SystemOnOptician{}] Optician with QREs running on the pre-QRE
+    benchmarks.
+  \item[\SystemOnBenchmarks{}] Optician with QREs running on the benchmarks that
+    were altered from Optician to include quotients.
 \end{itemize}
@@
 
-Also notable is one of the data.gov examples, which converts demographic
-statistics by zip code from xml form into rdf form. Synthesizing this
-transformation certainly took the longest, nearly fifteen seconds, where the
-next slowest took hardly over five. To validate this performance was not due to
-our overhead, we modified these format to be bijective, so the synthesis task
-could be input into Optician. Optician took over 15 seconds synthesize this
-modified lens. We believe this conversion took longer because the input formats
-were exceptionally large, requiring conversions between over 40 fields, not due
-to an overhead from synthesizing from QREs.
 \section{Related Work}

### 2018-03-13 papers/QuotientLenses/transform-data.py

@@ -47,6 +47,67 @@ def write_to_filename(filename, s):
     with open(filename, "wb") as f:
         f.write(s)
 
+def generate_ast_sizes_graph_confused(input_csv):
+    N = 10
+    opt_quot_ast = project_column_from_csv(input_csv, "OO")
+
+    ind = np.arange(N)  # the x locations for the groups
+    width = 0.35       # the width of the bars
+
+    fig, ax = plt.subplots()
+    rects1 = ax.bar(ind - width, opt_quot_ast, width=width, color='r',align='center')
+
+    opt_noquot_ast = project_column_from_csv(input_csv, "SO")
+    rects2 = ax.bar(ind, opt_noquot_ast, width=width,align='center')
+
+    pure_boom_ast = project_column_from_csv(input_csv, "SS")
+    rects3 = ax.bar(ind + width, pure_boom_ast, width=width, color='y',align='center')
+
+    # add some text for labels, title and axes ticks
+    ax.set_ylabel('Scores')
+    ax.set_title('Scores by group and gender')
+    ax.set_xticks(ind + width*(1.5))
+    ax.set_xticklabels(project_column_from_csv(input_csv, "Test"))
+
+    ax.legend((rects1[0], rects2[0], rects3[0]), ('OO', 'SS', 'SO'))
+
+    fig = plt.figure(1,tight_layout=True)
+    fig.set_figheight(2)
+    fig.set_figwidth(6)
+
+    fig.savefig(generated_graphs_base + "asts.eps", bbox_inches='tight')
+
+def generate_ast_sizes_graph(input_csv):
+    N = 10
+    ind = np.arange(N)  # the x locations for the groups
+
+    width = 0.25       # the width of the bars
+
+    fig, ax = plt.subplots()
+
+    qre_optician_means = project_column_from_csv(input_csv, "OO")
+    rects1 = ax.bar(ind, qre_optician_means, width, color='r')
+
+    basic_boomerang_means = project_column_from_csv(input_csv, "SO")
+    rects2 = ax.bar(ind + width, basic_boomerang_means, width, color='y')
+
+    optician_means = project_column_from_csv(input_csv, "SS")
+    rects3 = ax.bar(ind+width*2, optician_means, width, color='b')
+
+    # add some text for labels, title and axes ticks
+    ax.set_ylabel('AST Count')
+    ax.set_title('AST Counts by Benchmark')
[…]

### 2018-03-13 papers/QuotientLenses/main.tex
Tweaking in abstract

--- 2018-03-12 papers/QuotientLenses/main.tex
@@
 \begin{abstract}
-{\em Quotient lenses} are bidirectional transformations that respect
-specified equivalence relations with the goal of 
-appropriately handling inessential details in data formats. 
+{\em Quotient lenses} are bidirectional transformations whose correctness
+laws are ``loosened'' by specified equivalence relations, allowing
+inessential details in concrete data formats to be suppressed. 
 For example, a programmer could use a quotient lens to define 
@@
 that two XML files with the same fields but in different orders would be
-considered the same. 
-This paper simplifies the task of programming quotient lenses in three
-steps.
-First, we introduce {\em quotient regular expressions} (QREs) to
-make it easy to express which portions of a data format are
+considered the same, allowing a single, simple lens to handle them both.
+
+This paper simplifies the task of programming quotient lenses with
+three connected innovations.
+First, we introduce {\em quotient regular expressions} (QREs), to
+easily mark aspects of data formats that are
 inessential.  A QRE is an augmented regular expression that
-simultaneously defines a regular expression and an equivalence
-relation over matching strings.  Variations within the same
+simultaneously defines a regular language and an equivalence
+relation on this language.  Variations within the same
 equivalence class are deemed inessential. 
-Next, we introduce {\em QRE lenses}, which are lenses that map
+Next, we introduce {\em QRE lenses}---lenses mapping
 between QREs.
@@
 and $\ell$ is a bijective transformation.  Intuitively,
-the canonizers normalize the data  by picking a
+the canonizers normalize data by picking a
 canonical representative of the relevant equivalence relations. 
-This theorem means we can push all the normalization to the edges of a
+This theorem means we can push all normalization to the outer edges of a
 bidirectional transformation, leaving in the middle a pure bijective
@@
 Finally,
-we levarage this insight to {\em synthesize} QRE lenses from a pair
-of QREs and example input-output pairs by reusing earlier work on
-synthesizing bijective lenses.
+we leverage this normalization theorem to {\em synthesize} QRE lenses from a pair
+of QREs and example input-output pairs, reusing earlier work on
+synthesizing plain bijective lenses.
 We have implemented QREs and QRE lens synthesis 
@@
 synthesize QRE lenses between various real-world data formats in
-the {\tt data.gov} database.
-
+the {\tt data.gov} database.\bcp{The very last bit is inaccurate now.}
 \end{abstract}

### 2018-03-13 papers/QuotientLenses/main.tex

--- 2018-03-12 papers/QuotientLenses/main.tex
@@
   \item[\QRESize{}] the number of AST nodes in the QRE
-  specifications (including examples)\saz{CHECK THIS}\afm{no but i'll add it in}
+  specifications (including examples).
   \item[\canonizeAndSpecSize{}]  the number of
   AST nodes in $W(q_1)$ summed with the number of AST nodes in $\canonize(q_1)$
+  and the number of examples.
   \item[\LensAndSpecSize{}] the number of AST nodes
@@
 Figure~\ref{fig:asts} shows the number of benchmarks that can be expressed using
-at most a given number of AST nodes. For a given benchmark, we found \QRESize{}
-to be about half as large as \canonizeAndSpecSize, and about a third as large
-as \LensAndSpecSize{}.\BCP{These are really the critical conclusions, and they
-  are hard to see, the way the graph is drawn. I wonder whether we could instead
-  present with the individual examples on the X axis (sorted by number of AST
-  nodes), with a bar above each one containing three regions, colored
-  differently, for LS, CS-LS, and QC-CS...}\afm{will do} This suggests that
-introducing QREs saves programmers significant effort compared to both Optician
-and basic Boomerang.
+at most a given number of AST nodes. On average, \canonizeAndSpecSize{} used
+44\% more AST nodes than \QRESize{}, using an average of 237 more AST nodes. On
+average, \LensAndSpecSize{} used 193\% more AST nodes than \QRESize{}, using an
+average of 1267 more AST nodes. This suggests that introducing QREs saves
+programmers significant effort compared to both pre-QRE Optician and basic
+Boomerang.
 

### 2018-03-13 papers/QuotientLenses/main.tex
merge

@@
   synthesize all quotient lenses in under 20 seconds, and typically finished in
-  under 5 seconds.
+  under 5 seconds.}
 \label{fig:times}

### 2018-03-13 papers/QuotientLenses/main.tex
pass through section 4

--- 2018-03-13 papers/QuotientLenses/main.tex
@@
 \newcommand{\re}[1]{\ensuremath{\mathtt{#1}}}
-\newcommand{\kw}[1]{\ensuremath{\mathit{#1}}}
 \newcommand{\codefont}[1]{\ensuremath{\mathsf{#1}}}
+\newcommand{\kw}[1]{\codefont{#1}}
 \newcommand{\project}[2]{\ensuremath{\kw{project} \; #1 \mapsto #2}}
@@
 \newcommand{\perm}[2]{\ensuremath{\kw{perm}\; (#1)\; \kw{with}\; #2}}
-\newcommand{\normalize}[3]{\ensuremath{\kw{normalize} \; (#1, #2, #3)}}
+\newcommand{\normalize}[3]{\ensuremath{\kw{normalize}(#1, #2, #3)}}
 \newcommand{\eqrel}[1]{\ensuremath{\equiv_{#1}}}
@@
 \begin{abstract}
-{\em Quotient lenses} are bidirectional transformations whose correctness
-laws are ``loosened'' by specified equivalence relations, allowing
-inessential details in concrete data formats to be suppressed. 
+{\em Quotient lenses} are bidirectional transformations that respect
+specified equivalence relations with the goal of 
+appropriately handling inessential details in data formats. 
 For example, a programmer could use a quotient lens to define 
@@
 that two XML files with the same fields but in different orders would be
-considered the same, allowing a single, simple lens to handle them both.
-
-This paper simplifies the task of programming quotient lenses with
-three connected innovations.
-First, we introduce {\em quotient regular expressions} (QREs), to
-easily mark aspects of data formats that are
+considered the same. 
+This paper simplifies the task of programming quotient lenses in three
+steps.
+First, we introduce {\em quotient regular expressions} (QREs) to
+make it easy to express which portions of a data format are
 inessential.  A QRE is an augmented regular expression that
-simultaneously defines a regular language and an equivalence
-relation on this language.  Variations within the same
+simultaneously defines a regular expression and an equivalence
+relation over matching strings.  Variations within the same
 equivalence class are deemed inessential. 
-Next, we introduce {\em QRE lenses}---lenses mapping
+Next, we introduce {\em QRE lenses}, which are lenses that map
 between QREs.
@@
 and $\ell$ is a bijective transformation.  Intuitively,
-the canonizers normalize data by picking a
+the canonizers normalize the data  by picking a
 canonical representative of the relevant equivalence relations. 
-This theorem means we can push all normalization to the outer edges of a
+This theorem means we can push all the normalization to the edges of a
 bidirectional transformation, leaving in the middle a pure bijective
@@
 Finally,
-we leverage this normalization theorem to {\em synthesize} QRE lenses from a pair
-of QREs and example input-output pairs, reusing earlier work on
-synthesizing plain bijective lenses.
+we levarage this insight to {\em synthesize} QRE lenses from a pair
+of QREs and example input-output pairs by reusing earlier work on
+synthesizing bijective lenses.
 We have implemented QREs and QRE lens synthesis 
@@
 synthesize QRE lenses between various real-world data formats in
-the {\tt data.gov} database.\bcp{The very last bit is inaccurate now.}
+the {\tt data.gov} database.
+
 \end{abstract}
@@
 section describes in further detail the set of QRE combinators that we described
-in \cref{sec:example}, but first we discuss the \kw{normalize} combinator which
-enjoys some special properties.
+in \cref{sec:example}.\saz{Why does the ``cref'' macro not capitalize Section?}
+
+\subsection{Syntax of Semantics of QREs}
+Formally, the language of Quotient Regular Expressions (QREs) is given by the
+following grammar:
+\newcommand{\bsep}{\ \ \sep{} \ \ }
+\begin{align*}
+Q := \; & \normalize{R}{R'}{f} \bsep{} \id(R) \bsep \project{R}{s} \bsep \squash{Q_1}{Q_2}{f} \\
+ | \; & \perm{Q_1, \ldots, Q_n}{Q} \bsep Q_1 \semicolon Q_2 \bsep Q_1 \cdot Q_2 \bsep (Q_1
+\sep Q_2) \bsep Q^*,
+\end{align*}
+where $Q$ ranges over QREs, $R$ ranges over regular expressions, $f$ ranges over
+functions between regular languages, and $s$ ranges over character strings.
+
+Using the conventional notation that $\mathcal{L}(R)$ is the language accepted
+by the regular expression $R$, each QRE $Q$ enables us to express four semantic objects
+
+\begin{tabular}{rl}
+  $W(Q)$  & a regular expression (the ``whole'' of $Q$),\\
+  $\eqrel{Q}$  & an equivalence relation on $\mathcal{L}(W(Q))$,\\
+  $K(Q)$ & a regular expression (the ``kernel'' of $Q$) such that
+           $\mathcal{L}(K(Q))$ forms a complete \\
+             & set of representatives for $\eqrel{Q}$, and \\
+  $\canonize(Q)$ & 
+  a ``canonizing'' function that, given any $w \in \mathcal{L}(W(Q))$,
+  yields $\canonize(Q)(w)$ \\ & as the unique $k$ in $\mathcal{L}(K(Q))$ such
+  that $k$ is equivalent to $w$ mod $\eqrel{Q}$
+\end{tabular}
+  % of type $\mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$  which
+
+The well-formedness constraints for QREs ensure that these four semantic
+components fit together to form a coherent quotient $W(Q)/\eqrel{Q}$ whose
+equivalence classes are determined by \canonize(Q).
 
 \subsection{The \kw{normalize} Combinator}
-The \kw{normalize} combinator is special in two main ways, the first of
-which has to do with the premises in its inference rule:
+
+The relationship among the semantic objects of a QRE can be understood in terms
+of the \normalize{R}{S}{f} combinator, which exposes each of these pieces
+explicitly.  Its whole language is just $R$, its kernel language is just
+$S$, and $\canonize(\normalize{R}{S}{f}) = f$.  Its equivalence relation is
+determined by the fibres of $f$, so we have
+$s_1 \equiv s_2 \Leftrightarrow f(s_1) = f(s_2)$.
+
+These components form a coherent quotient language when the canonization
+function $f$ is surjective and idempotent (intuitively, $f$ picks out a unique
+representative for each equivalence class).  We also require that the kernel
+language be a subset of the whole language, which enables QRE composition.
+These considerations lead to the following well-formedness rule.
+
 \begin{prooftree}
-\AxiomC{$\mathcal{L}(R') \subseteq \mathcal{L}(R)$}
-\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(R')$}
+\AxiomC{$\mathcal{L}(S) \subseteq \mathcal{L}(R)$}
+\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(S)$}
 \AxiomC{$f$ is surjective}
@@
 \end{prooftree}
-In order for the {\em normalize} QRE to be well-formed, then canonizing function
-$f$ used by this combinator must be surjective and idempotent. Unfortunately,
-verifying that a function satisfies these properties is in general undecidable.
-Nonetheless, we allow a programmer to use normalize at their own risk without
-checking these side conditions.
-
-We do this firstly because the {\em normalize} combinator allows one to express
-equivalence relations that are either difficult to define, or that cannot be
-defined using the other combinators. Secondly, the {\em normalize} canonizer
-enjoys the property that each of the other combinators $Q$ is semantically
-equivalent to $\normalize{W(Q)}{K(Q)}{f}$ for some surjective, idempotent
-function $f : \mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$; this is the
-second property that distinguishes the {\em normalize} combinator from the other
-QREs. The {\em normalize} combinator hence gives a sufficient condition for
-potential QREs to be ``well-behaved'', particularly with respect to the
-strategy for synthesizing quotient lenses which we shall introduce in
-Section~\ref{synth}.
-
-Having discussed the special properties of the \kw{normalize} combinator, we now
-give the syntax and semantics of the other QRE combinators.
-\subsection{Syntax and Semantics of QREs}
-Formally, the language of Quotient Regular Expressions (QREs) is given by the
-following grammar:
-\begin{align*}
-Q := \; &\id(R) \sep \project{R}{s} \sep \squash{Q_1}{Q_2}{f} \sep
-\perm{Q_1, \ldots, Q_n}{Q} \\
-& | \; \normalize{R}{R'}{f} \sep Q_1 \semicolon Q_2 \sep Q_1 \cdot Q_2 \sep (Q_1
-\sep Q_2) \sep Q^*,
-\end{align*}
-where $Q$ ranges over QREs, $R$ ranges over regular expressions, $f$ ranges over
-functions between regular languages, and $s$ ranges over character strings.
-
-Using the conventional notation that $\mathcal{L}(R)$ is the language accepted
-by the regular expression $R$, then each QRE $Q$ enables us to express
-\begin{enumerate}
-  \item a regular expression $W(Q)$ (the ``whole'' of $Q$),
-  \item an equivalence relation $\eqrel{Q}$ on $\mathcal{L}(W(Q))$,
-  \item a regular expression $K(Q)$ (the ``kernel'' of $Q$)
-  such that $\mathcal{L}(K(Q))$ forms a complete set of representatives for
-  $\eqrel{Q}$, and
-  \item a ``canonizing'' function $\canonize(Q):\mathcal{L}(W(Q))
-  \longrightarrow \mathcal{L}(K(Q))$ which given any $w \in \mathcal{L}(W(Q))$,
-  computes $\canonize(Q)(w)$ as the unique $k$ in $\mathcal{L}(K(Q))$ such
-  that $k$ is equivalent to $w$ mod $\eqrel{Q}$
-\end{enumerate}
-Figures~\ref{fig:wk} gives the inductive definitions for the whole and
-kernel languages $W(Q)$ and $K(Q)$ of a QRE.
+
+Semantically, the \kw{normalize} QRE is universal---each of the other
+combinators $Q$ is equivalent to $\normalize{W(Q)}{K(Q)}{f}$ for some
+surjective, idempotent function
+$f : \mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$.   However,
+although \kw{normalize} is powerful, verifying that a canonization function $f$
+is surjective and idempotent is in general undecidable.  This means that a
+programmer wishing to use \kw{normalize} must discharge strong proof
+obligations, which is cumbersome in practice.\footnote{In our implementation we allow a programmer to use \kw{normalize} at their
+own risk without checking these side conditions as an ``escape hatch'' for the
+case when other QRE combinators are insufficient.}
+
+The remaining QRE combinators provide simpler, more compositional ways of
+building canonizers.  Nevertheless, the \kw{normalize} combinator gives a
+sufficient condition for potential QREs to be well-behaved, particularly with
+respect to the strategy for synthesizing quotient lenses which we shall
+introduce in Section~\ref{synth}, so we use it to guide the design of our QRE
+combinators, as we see next.
+
+% ; this is the second
+% property that distinguishes the \kw{normalize} combinator from the other QREs.
+% The \kw{normalize} combinator is special in two main ways, the first of 
+% which has to do with the premises in its inference rule:
+
+\subsection{QRE Combinator Semantics}
+
+Figure~\ref{fig:wk} gives the inductive definitions of the whole and
+kernel languages $W(Q)$ and $K(Q)$ for all of the QRE forms.
 \begin{figure}[t]
@@
 \end{figure}
-The \textit{squash} and \textit{permutation} combinators give the two
+The \kw{squash} and \kw{permutation} combinators have the two most
 interesting definitions. If $Q = \squash{Q_1}{Q_2}{f}$, then the whole language
-of $Q$ is $W(Q_1) \sep W(Q_2)$. This is because the \textit{squash} combinator
+of $Q$ is $W(Q_1) \sep W(Q_2)$. This is because the \kw{squash} combinator
 takes the whole language $W(Q_1)$ of $Q_1$, the whole language $W(Q_2)$ of $Q_2$
@@
 $S_n$ is the set of all permutations of the numbers $1$ to $n$. This is because
-the \textit{permutation} combinator allows for the string to match any
+the \kw{permutation} combinator allows for the string to match any
 permutation the $Q_i$'s while also taking into account the separator $Q$ in
-between each of the $Q_i$'s. The kernel of the \textit{permutation} combinator
+between each of the $Q_i$'s. The kernel of the \kw{permutation} combinator
 is the language $K(Q_1) \cdot K(Q) \cdot \ldots \cdot K(Q) \cdot K(Q_n)$,
@@
 \end{figure}
-Again, the \textit{permutation} combinator gives the most interesting
+Again, the \kw{permutation} combinator gives the most interesting
 definition,
@@
 The most complicated inference rule is the (Perm) rule for the
-\textit{permutation} combinator. The second hypothesis for the
-\textit{permutation} rule says that for any two different permutations $\sigma$
+\kw{permutation} combinator. The second hypothesis for the
+Perm rule says that for any two different permutations $\sigma$
 and $\theta$, the languages $W(c_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
@@
 W(c_{\theta(n)})$ must be disjoint. This is because any string that matches the
-\textit{permutation} combinator matches the language $W(c_{\theta(1)}) \cdot
+\kw{permutation} combinator matches the language $W(c_{\theta(1)}) \cdot
 W(Q) \cdot \ldots \cdot W(c_{\theta(n)})$ for some permutation $\theta$, so we
@@
 \cdot \ldots \cdot K(Q) \cdot K(Q_n)$ must be unambiguous. This is because the
-underlying lens that maps to or from a \textit{permutation} QRE operates on
+underlying lens that maps to or from a \kw{permutation} QRE operates on
 the language $K(Q_1) \cdot K(Q) \cdot \ldots \cdot K(Q) \cdot K(Q_n)$ which is
-the kernel of the \textit{permutation} QRE. As we mentioned, the underlying
+the kernel of the \kw{permutation} QRE. As we mentioned, the underlying
 lens will require that the source and target regular expressions of the lens be

### 2018-03-13 papers/QuotientLenses/main.tex

@@
   performance overhead incurred by using QREs. In (b), we run Optician with QREs
-  on its benchmark suite. We find that it is able to synthesize all quotient
-  lenses in under 20 seconds, and typically finished in under 5 seconds.
-  \bcp{The left-hand graph, especially, is hard to make anything out of, because
-    the curves are squished against the axes and have almost identical shapes.
-    }\afm{will fix}}
+  on the Optician benchmarks we modified to use QREs. We find that it is able to
+  synthesize all quotient lenses in under 20 seconds, and typically finished in
+  under 5 seconds.}
 \label{fig:times}

### 2018-03-13 papers/QuotientLenses/main.tex
minor tweak
Checkpoint some tweaking and comments, mostly toward the beginning

@@
 keywordstyle=\upshape\sffamily\color{dkblue}, keepspaces=true,
-framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=false,
+framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=true,
 commentstyle=\itshape\rmfamily,
-emph={synth,project,perm,squash,normalize,using,ins,del,lens},
-emphstyle=\upshape\sffamily\color{dkblue}, columns=fullflexible,
+emph={synth,project,perm,squash,normalize,using,ins,del,lens,let},
+emphstyle=\upshape\sffamily\color{dkred}, 
+columns=fullflexible,
+escapechar=\#,
 % BCP: I find this distracting:
@@
 
-This paper simplifies the task of programming quotient lenses with
-three connected innovations.
+We offer three interconnected innovations to simplify the task of programming
+quotient lenses.  
 First, we introduce {\em quotient regular expressions} (QREs), to
@@
 equivalence class are deemed inessential. 
-Next, we introduce {\em QRE lenses}---lenses mapping
+Next, we introduce {\em QRE lenses}, i.e., lenses mapping
 between QREs.
 Our key technical result is a proof that every QRE lens can be
-transformed into a QRE lens that has the form $c_1 \cdot \ell \cdot
-c_2$, where $c_1$ and $c_2$ are {\em canonizers} 
-and $\ell$ is a bijective transformation.  Intuitively,
+transformed into one of the form $c_1 ; \ell ;
+c_2^{-1}$, where $c_1$ and $c_2^{-1}$ are {\em canonizers}, $\ell$ is a bijective
+transformation, and $;$ is composition\iflastminute\bcp{$c_2^{-1}$ is ugly, but I think
+  we need the superscript for accuracy.  Perhaps calling them $c$ and $d$
+  instead of $c_1$ and $c_2$ would be better?}\fi.  Intuitively,
 the canonizers normalize data by picking a
@@
 APIs can return data as either JSON or XML objects.  When changes are
-made to data in one format, data in the other must be updated so the
-two representations continue to store the same information.
-One way to synchronize such data is to use \emph{lenses}~\cite{Lenses}, which
-are programs that simultaneously define pairs of functions for translating
+made one version, the data in the other must be updated so the
+two continue to represent the same information.
+One way to synchronize such data is to use \emph{lenses}~\cite{Lenses}, 
+programs that simultaneously define pairs of functions for translating
 the data from the source format to the target formant (the \emph{get}
-direction) and back again (the \emph{put} direction). Lenses are usually
-defined using domain-specific bidirectional languages that 
-ensure the specified lenses satisfy various \emph{lens laws} that
-guarantee how the \emph{put} and \emph{get} functions interact.
+direction) and back again (the \emph{put} direction). Lenses can be
+written in one of several domain-specific bidirectional languages that 
+guarantee various \emph{lens laws} constraining how the \emph{put} and
+\emph{get} functions must fit together.
 
@@
 \end{equation}
-Since these lenses demand that the \emph{get} and \emph{put} functions form a bijection, 
+Since the \emph{get} and \emph{put} functions form a bijection, 
 no information is lost during translation from one data format to another. 
-While this very strong guarantee can be useful, more often
-some of the information in one or both of the data formats is
+While this very strong guarantee can sometimes be useful, it is more common for 
+some of the details in one or both of the data formats to be
 superfluous. For instance, the number of white space characters
@@
 sequence of fields in a record, or the order of items in a list are
-often irrelevant.  Many
-data formats have their own ad hoc rules as well. In \bibtex{}, for example,
-an author can be specified as
-``\cd{lastname, firstname}'' or as ``\cd{firstname lastname}''---
-the specific choice does not matter.
-
-{\em Quotient lenses}~\cite{quotientlenses} loosen the restrictions of
+often irrelevant.  Also, many
+data formats introduce their own ad hoc conventions about formatting details
+that are taken to be inessential. In \bibtex{}, for example,
+an author can be specified either as 
+``\cd{lastname, firstname}'' or as ``\cd{firstname lastname}''.
+
+{\em Quotient lenses}~\cite{quotientlenses} loosen the strong restrictions of
 bijective string lenses by including equivalence
-relations on the source and target formats so that the round-tripping lens laws
-hold modulo these equivalences rather than modulo equality. Quotient lenses thus
-make it possible to synchronize many more data formats than would otherwise be
-possible by ignoring details deemed inessential.
+relations on the source and target formats and requiring that the
+round-tripping lens laws 
+hold  modulo these equivalences, rather than ``on the nose.'' Quotient lenses 
+make it possible to synchronize many more data formats than is possible with
+plain bijective lenses\bcp{Do we justify this claim later?}.
 
@@
 author names:
+\iflastminute
+\bcp{I'd love to use a visible space symbol for important space characters
+  -- this would make the code examples very much clearer!  But it will
+  require a bit of hacking: We can use showstringspaces=false}
+\fi
 %
@@
 \label{sec:background}
+\iflastminute
 \ksf{We are inconsistent in how we typeset get/put.  Sometimes we use
@@
 we don't do anything distinctive.  It would be good to be consistent!}
+\fi
 Before describing QREs and QRE lenses, we briefly review bijective
@@
 expressions, which are omitted for brevity.  
-\ifdraft
-\bcp{And a readability suggestion:
-  define a named constant for a quote character.}
+\iflastminute
+\bcp{a possible readability suggestion: define a named constant for a quote
+  character.  (Not sure if this makes sense, or if it improves things...)}
 \fi}
@@
 us to use it inside listings.  But we should save this for the final
-tweaking stage.}
+tweaking stage.}\bcp{Done, I think.}
 % Here is an example: if you set escapechar=\%, you could use it inside the
@@
 
-For instance, assume that $f(s)=$ `` '' for all whitespace strings $s$. Then the
+For instance, assume that $f(s)=$ ``\textvisiblespace'' (a space
+character) for all whitespace strings $s$. Then the 
 project QRE \lstinline{wsp_sp} defined above can
@@
 \begin{lstlisting}
-let wsp = [ \n\t\r]+
+let wsp = [#\,\textvisiblespace\,#\n\t\r]+
 let wsp_sp = project wsp $\to$ " "
@@
 section describes in further detail the set of QRE combinators that we described
-in \cref{sec:example}.\saz{Why does the ``cref'' macro not capitalize Section?}
+in \cref{sec:example}.
+\iflastminute 
+\saz{Why does the ``cref'' macro not capitalize Section?}
+\bcp{Dunno.  Because SIGPLAN does not want it to??  But maybe we should be
+  using autoref instead anyway...}
+\fi
 
@@
 own risk without checking these side conditions as an ``escape hatch'' for the
-case when other QRE combinators are insufficient.}
+case when other QRE combinators are insufficient.\bcp{Do we use this in any
+  of the programs we've written?}}
 
@@
 \begin{array}{l@{}l@{\ \ }l}
+w \; \equiv_{\normalize{R}{R'}{f}} \; w' &\iff&
+f(w)=f(w') \\
 w \; \equiv_{\id(R)} \; w' &\iff& w = w' \\
@@
 \text{ or } w \equiv_{Q_2} w' \\
-w \; \equiv_{\normalize{R}{R'}{f}} \; w' &\iff&
-f(w)=f(w') \text{ or }w = w'\\
 w \; \equiv_{Q_1 \; ; \; Q_2} \; w' &\iff& \exists k, k' \in

### 2018-03-13 papers/QuotientLenses/main.tex
synth qre lenses

--- 2018-03-13 papers/QuotientLenses/main.tex
@@
 \setcopyright{none}
-\usepackage{amsmath, amssymb, verbatim, enumerate, graphicx, centernot, tikz,
-array, mathtools, bussproofs, stmaryrd, enumitem, stackengine, subcaption,
-cleveref}
+\usepackage{algorithm, amsmath, amssymb, verbatim, enumerate, graphicx,
+  centernot, tikz, array, mathtools, bussproofs, stmaryrd, enumitem,
+  stackengine, subcaption, cleveref}
 \captionsetup{compatibility=false}
@@
 \usepackage{xspace}
+\usepackage[noend]{algpseudocode}
 
@@
 \newcommand{\re}[1]{\ensuremath{\mathtt{#1}}}
+\newcommand{\kw}[1]{\ensuremath{\mathit{#1}}}
 \newcommand{\codefont}[1]{\ensuremath{\mathsf{#1}}}
-\newcommand{\kw}[1]{\codefont{#1}}
 \newcommand{\project}[2]{\ensuremath{\kw{project} \; #1 \mapsto #2}}
@@
 \newcommand{\perm}[2]{\ensuremath{\kw{perm}\; (#1)\; \kw{with}\; #2}}
-\newcommand{\normalize}[3]{\ensuremath{\kw{normalize}(#1, #2, #3)}}
+\newcommand{\normalize}[3]{\ensuremath{\kw{normalize} \; (#1, #2, #3)}}
 \newcommand{\eqrel}[1]{\ensuremath{\equiv_{#1}}}
@@
 \begin{lstlisting}
-let preamble : (lens in "@book{" $\Leftrightarrow$ "%0 Book\n%F ") =
+let preamble : "@book{" $\Leftrightarrow$ "%0 Book\n%F " =
 del "@book{" . ins "%0 Book\n%F "
 
-let author_lens : (lens in (",\n" . bib_author) $\Leftrightarrow$ ("\n" . end_author)) =
+let author_lens : (",\n" . bib_author) $\Leftrightarrow$ ("\n" . end_author) =
 del ",\nauthor = \""
@@
 
-let title_lens : (lens in ("\",\n" . bib_title . "},\n}") $\Leftrightarrow$ ("\n" . end_title)) =
+let title_lens : ("\",\n" . bib_title . "},\n}") $\Leftrightarrow$ ("\n" . end_title) =
 del "\",\ntitle = {" . ins "\n%T " . title . del "},\n}"
 
-let bib_to_end : (lens in bibtex $\Leftrightarrow$ end_note) = preamble . label . author_lens . title_lens
+let bib_to_end : bibtex $\Leftrightarrow$ end_note = preamble . label . author_lens . title_lens
 \end{lstlisting}
@@
 to the synthesis prodedure:
-\ksf{What is the notation ``lens in'' below?  We should explain it
-intuitively here.}
 \begin{lstlisting}
-let bib_to_end : (lens in bibtex $\Leftrightarrow$ endnote) =
+let bib_to_end : bibtex $\Leftrightarrow$ endnote =
 synth bibtex $\Leftrightarrow$ endnote using {(bib_example, end_example)}
@@
 section describes in further detail the set of QRE combinators that we described
-in \cref{sec:example}.\saz{Why does the ``cref'' macro not capitalize Section?}
-
-\subsection{Syntax of Semantics of QREs}
-Formally, the language of Quotient Regular Expressions (QREs) is given by the
-following grammar:
-\newcommand{\bsep}{\ \ \sep{} \ \ }
-\begin{align*}
-Q := \; & \normalize{R}{R'}{f} \bsep{} \id(R) \bsep \project{R}{s} \bsep \squash{Q_1}{Q_2}{f} \\
- | \; & \perm{Q_1, \ldots, Q_n}{Q} \bsep Q_1 \semicolon Q_2 \bsep Q_1 \cdot Q_2 \bsep (Q_1
-\sep Q_2) \bsep Q^*,
-\end{align*}
-where $Q$ ranges over QREs, $R$ ranges over regular expressions, $f$ ranges over
-functions between regular languages, and $s$ ranges over character strings.
-
-Using the conventional notation that $\mathcal{L}(R)$ is the language accepted
-by the regular expression $R$, each QRE $Q$ enables us to express four semantic objects
-
-\begin{tabular}{rl}
-  $W(Q)$  & a regular expression (the ``whole'' of $Q$),\\
-  $\eqrel{Q}$  & an equivalence relation on $\mathcal{L}(W(Q))$,\\
-  $K(Q)$ & a regular expression (the ``kernel'' of $Q$) such that
-           $\mathcal{L}(K(Q))$ forms a complete \\
-             & set of representatives for $\eqrel{Q}$, and \\
-  $\canonize(Q)$ & 
-  a ``canonizing'' function that, given any $w \in \mathcal{L}(W(Q))$,
-  yields $\canonize(Q)(w)$ \\ & as the unique $k$ in $\mathcal{L}(K(Q))$ such
-  that $k$ is equivalent to $w$ mod $\eqrel{Q}$
-\end{tabular}
-  % of type $\mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$  which
-
-The well-formedness constraints for QREs ensure that these four semantic
-components fit together to form a coherent quotient $W(Q)/\eqrel{Q}$ whose
-equivalence classes are determined by \canonize(Q).
+in \cref{sec:example}, but first we discuss the \kw{normalize} combinator which
+enjoys some special properties.
 
 \subsection{The \kw{normalize} Combinator}
-
-The relationship among the semantic objects of a QRE can be understood in terms
-of the \normalize{R}{S}{f} combinator, which exposes each of these pieces
-explicitly.  Its whole language is just $R$, its kernel language is just
-$S$, and $\canonize(\normalize{R}{S}{f}) = f$.  Its equivalence relation is
-determined by the fibres of $f$, so we have
-$s_1 \equiv s_2 \Leftrightarrow f(s_1) = f(s_2)$.
-
-These components form a coherent quotient language when the canonization
-function $f$ is surjective and idempotent (intuitively, $f$ picks out a unique
-representative for each equivalence class).  We also require that the kernel
-language be a subset of the whole language, which enables QRE composition.
-These considerations lead to the following well-formedness rule.
-
+The \kw{normalize} combinator is special in two main ways, the first of
+which has to do with the premises in its inference rule:
 \begin{prooftree}
-\AxiomC{$\mathcal{L}(S) \subseteq \mathcal{L}(R)$}
-\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(S)$}
+\AxiomC{$\mathcal{L}(R') \subseteq \mathcal{L}(R)$}
+\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(R')$}
 \AxiomC{$f$ is surjective}
@@
 \end{prooftree}
-
-Semantically, the \kw{normalize} QRE is universal---each of the other
-combinators $Q$ is equivalent to $\normalize{W(Q)}{K(Q)}{f}$ for some
-surjective, idempotent function
-$f : \mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$.   However,
-although \kw{normalize} is powerful, verifying that a canonization function $f$
-is surjective and idempotent is in general undecidable.  This means that a
-programmer wishing to use \kw{normalize} must discharge strong proof
-obligations, which is cumbersome in practice.\footnote{In our implementation we allow a programmer to use \kw{normalize} at their
-own risk without checking these side conditions as an ``escape hatch'' for the
-case when other QRE combinators are insufficient.}
-
-The remaining QRE combinators provide simpler, more compositional ways of
-building canonizers.  Nevertheless, the \kw{normalize} combinator gives a
-sufficient condition for potential QREs to be well-behaved, particularly with
-respect to the strategy for synthesizing quotient lenses which we shall
-introduce in Section~\ref{synth}, so we use it to guide the design of our QRE
-combinators, as we see next.
-
-% ; this is the second
-% property that distinguishes the \kw{normalize} combinator from the other QREs.
-% The \kw{normalize} combinator is special in two main ways, the first of 
-% which has to do with the premises in its inference rule:
-
-\subsection{QRE Combinator Semantics}
-
-Figure~\ref{fig:wk} gives the inductive definitions of the whole and
-kernel languages $W(Q)$ and $K(Q)$ for all of the QRE forms.
+In order for the {\em normalize} QRE to be well-formed, then canonizing function
+$f$ used by this combinator must be surjective and idempotent. Unfortunately,
+verifying that a function satisfies these properties is in general undecidable.
+Nonetheless, we allow a programmer to use normalize at their own risk without
+checking these side conditions.
+
+We do this firstly because the {\em normalize} combinator allows one to express
+equivalence relations that are either difficult to define, or that cannot be
+defined using the other combinators. Secondly, the {\em normalize} canonizer
+enjoys the property that each of the other combinators $Q$ is semantically
+equivalent to $\normalize{W(Q)}{K(Q)}{f}$ for some surjective, idempotent
+function $f : \mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$; this is the
+second property that distinguishes the {\em normalize} combinator from the other
+QREs. The {\em normalize} combinator hence gives a sufficient condition for
+potential QREs to be ``well-behaved'', particularly with respect to the
+strategy for synthesizing quotient lenses which we shall introduce in
+Section~\ref{synth}.
+
+Having discussed the special properties of the \kw{normalize} combinator, we now
+give the syntax and semantics of the other QRE combinators.
+\subsection{Syntax and Semantics of QREs}
+Formally, the language of Quotient Regular Expressions (QREs) is given by the
+following grammar:
+\begin{align*}
+Q := \; &\id(R) \sep \project{R}{s} \sep \squash{Q_1}{Q_2}{f} \sep
+\perm{Q_1, \ldots, Q_n}{Q} \\
+& | \; \normalize{R}{R'}{f} \sep Q_1 \semicolon Q_2 \sep Q_1 \cdot Q_2 \sep (Q_1
+\sep Q_2) \sep Q^*,
+\end{align*}
+where $Q$ ranges over QREs, $R$ ranges over regular expressions, $f$ ranges over
+functions between regular languages, and $s$ ranges over character strings.
+
+Using the conventional notation that $\mathcal{L}(R)$ is the language accepted
+by the regular expression $R$, then each QRE $Q$ enables us to express
+\begin{enumerate}
+  \item a regular expression $W(Q)$ (the ``whole'' of $Q$),
+  \item an equivalence relation $\eqrel{Q}$ on $\mathcal{L}(W(Q))$,
+  \item a regular expression $K(Q)$ (the ``kernel'' of $Q$)
+  such that $\mathcal{L}(K(Q))$ forms a complete set of representatives for
+  $\eqrel{Q}$, and
+  \item a ``canonizing'' function $\canonize(Q):\mathcal{L}(W(Q))
+  \longrightarrow \mathcal{L}(K(Q))$ which given any $w \in \mathcal{L}(W(Q))$,
+  computes $\canonize(Q)(w)$ as the unique $k$ in $\mathcal{L}(K(Q))$ such
+  that $k$ is equivalent to $w$ mod $\eqrel{Q}$
+\end{enumerate}
+Figures~\ref{fig:wk} gives the inductive definitions for the whole and
+kernel languages $W(Q)$ and $K(Q)$ of a QRE.
 \begin{figure}[t]
@@
 \end{figure}
-The \kw{squash} and \kw{permutation} combinators have the two most
+The \textit{squash} and \textit{permutation} combinators give the two
 interesting definitions. If $Q = \squash{Q_1}{Q_2}{f}$, then the whole language
-of $Q$ is $W(Q_1) \sep W(Q_2)$. This is because the \kw{squash} combinator
+of $Q$ is $W(Q_1) \sep W(Q_2)$. This is because the \textit{squash} combinator
 takes the whole language $W(Q_1)$ of $Q_1$, the whole language $W(Q_2)$ of $Q_2$
@@
 $S_n$ is the set of all permutations of the numbers $1$ to $n$. This is because
-the \kw{permutation} combinator allows for the string to match any
+the \textit{permutation} combinator allows for the string to match any
 permutation the $Q_i$'s while also taking into account the separator $Q$ in
-between each of the $Q_i$'s. The kernel of the \kw{permutation} combinator
+between each of the $Q_i$'s. The kernel of the \textit{permutation} combinator
 is the language $K(Q_1) \cdot K(Q) \cdot \ldots \cdot K(Q) \cdot K(Q_n)$,
@@
 \end{figure}
-Again, the \kw{permutation} combinator gives the most interesting
+Again, the \textit{permutation} combinator gives the most interesting
 definition,
@@
 The most complicated inference rule is the (Perm) rule for the
-\kw{permutation} combinator. The second hypothesis for the
-Perm rule says that for any two different permutations $\sigma$
+\textit{permutation} combinator. The second hypothesis for the
+\textit{permutation} rule says that for any two different permutations $\sigma$
 and $\theta$, the languages $W(c_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
@@
 W(c_{\theta(n)})$ must be disjoint. This is because any string that matches the
-\kw{permutation} combinator matches the language $W(c_{\theta(1)}) \cdot
+\textit{permutation} combinator matches the language $W(c_{\theta(1)}) \cdot
 W(Q) \cdot \ldots \cdot W(c_{\theta(n)})$ for some permutation $\theta$, so we
@@
 \cdot \ldots \cdot K(Q) \cdot K(Q_n)$ must be unambiguous. This is because the
-underlying lens that maps to or from a \kw{permutation} QRE operates on
+underlying lens that maps to or from a \textit{permutation} QRE operates on
 the language $K(Q_1) \cdot K(Q) \cdot \ldots \cdot K(Q) \cdot K(Q_n)$ which is
-the kernel of the \kw{permutation} QRE. As we mentioned, the underlying
+the kernel of the \textit{permutation} QRE. As we mentioned, the underlying
 lens will require that the source and target regular expressions of the lens be
@@
 \begin{lstlisting}
-let bib_to_end_q : (lens in bibtex $\Leftrightarrow$ endnote) =  rquot (lquot(bibtex, bib_to_end), endnote)
+let bib_to_end_q : bibtex $\Leftrightarrow$ endnote =  rquot (lquot(bibtex, bib_to_end), endnote)
 \end{lstlisting}
@@
 
-The Optician synthesis algorithm works as follows. Given regular expressions $R$
-and $S$ and a set of example input--output pairs
-$\{(r_1, s_1),\allowbreak \ldots,\allowbreak (r_n, s_n)\}$, Optician will try to
-find a bijective lens $\ell : R \Leftrightarrow S$ that that agrees on the
-examples, \textit{i.e.}, it maps $r_i$ to $s_i$ in the forward direction and
-$s_i$ to $r_i$ in the backward direction.  The Optician algorithm is guaranteed
-to succeed (eventually!) if such a lens exists.    \saz{We should probably
-  mention what happens if the examples aren't sufficient to uniquely
-  determine the lens.}
+The Optician bijective lens synthesis algorithm works as follows. Given regular
+expressions $R$ and $S$ and a set of example input--output pairs $\{(r_1,
+s_1),\allowbreak \ldots,\allowbreak (r_n, s_n)\}$, Optician will try to find a
+bijective lens $\ell : R \Leftrightarrow S$ that agrees on the examples,
+\textit{i.e.}, it maps $r_i$ to $s_i$ in the forward direction and $s_i$ to
+$r_i$ in the backward direction. The bijective lens synthesis algorithm is
+guaranteed to succeed (eventually!) if such a bijective lens exists. There are
+typically a large number of such lenses, so this algorithm chooses the one that
+corresponds to a minimal \emph{alignment} of the regular expressions. Additional
+details on this algorithm can be found in the original paper on synthesizing
+bijective lenses~\cite{optician}.
 
@@
 {y'_n})\}$ and then apply the canonizers at the outside. Our procedure for QRE
-lens synthesis is thus
-\begin{enumerate}
-  \item
-  compute $K(Q_1)$, $K(Q_2)$,
-  \item
-  compute $\{({x'_1}, {y'_1}), \ldots, ({x'_n}, {y'_n})\}$ where ${x'_i} =
-  \canonize(Q_1)(x_i)$ and ${y'_i} = \canonize(Q_2)(y_i)$,
-\item synthesize, using Optician's algorithm, a bijective lens
-  $\ell : K(Q_1) \Leftrightarrow K(Q_2)$ that is consistent with the examples
-  $\{({x'_1}, {y'_1}),\allowbreak \ldots,\allowbreak ({x'_n}, {y'_n})\}$,
-  \item
-  return $q = \rquot(\lquot(Q_1, \ell), Q_2) : Q_1 \Leftrightarrow
-  Q_2$.
-\end{enumerate}
-This algorithm is complete in the following sense: given any QREs $Q_1$ and
-$Q_2$ and set of examples $\{(x_1, y_2), \ldots, (x_n, y_n)\}$ with $x_i \in
-W(Q_1)$ and $y_i \in W(Q_2)$, then if there exists a derivation of a QRE lens
-$q : Q_1 \Leftrightarrow Q_2$ such that 
-\begin{equation}\label{qrelensequation}
-q.\get(x_i) \equiv_{Q_2} y_i \text{ and } q.\lput(y_i) \equiv_{Q_1} x_i \text{
-for all } i,
-\end{equation}
-then the algorithm will return a QRE lens $q': Q_1 \Leftrightarrow Q_2$ that
-also satisfies \cref{qrelensequation}. The justification for this fact follows
-from the completeness of the Optician algorithm and \cref{normal form}.
+lens synthesis is given formally in Algorithm~\ref{alg:synthqrelens}.
+
+\begin{algorithm}
+  \caption{\textproc{SynthQRELens}}
+  \label{alg:synthqrelens}
+  \begin{algorithmic}[1]
+    \Function{\textproc{SynthQRELens}}{$Q_1,Q_2,exs$}
+    \State $R_1 \gets K(Q_1)$
+    \State $R_2 \gets K(Q_2)$
+    \State $c_1 \gets \canonize(Q_1)$
+    \State $c_2 \gets \canonize(Q_2)$
+    \State $exs' \gets \Call{Map}{exs,fun \, (ex_l,ex_r) \to
+      (c_1(ex_l),c_2(ex_r))}$
+    \State $l \gets \Call{SynthBijectiveLens}{R_1,R_2,exs'}$
+    \State $\Return \, \rquot(\lquot(Q_1, \ell), Q_2)$
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
+\noindent
+We have also proven the correctness of this algorithm.
+\begin{theorem}\label{thm:alg-correct}
+  Given QREs $Q_1$ and $Q_2$, and a set of examples
+  $\{(x_1,y_1),\ldots,(x_n,y_n)\}$, if there is a QRE lens $q : Q_1
+  \Leftrightarrow Q_2$ such that $q.\get(x_i) \equiv_{Q_2} y_i \text{ and }
+q.\lput(y_i) \equiv_{Q_1} x_i$, then $\textproc{SynthQRELens}(Q_1,Q_2,exs)$ will
+return such a lens.
+\end{theorem}
+
+\noindent
+This theorem follows from the correctness of the Optician algorithm and
+Theorem~\ref{normal form}.
 
 Returning to the \bibtex{} to EndNote transformation of
-Section~\ref{sec:examplesynth}, we had the QREs \cd{bibtex} and \cd{endnote}
-describing a \bibtex{} record and an Endnote record respectively. We also had
+Section~\ref{sec:examplesynth}, the QREs \cd{bibtex} and \cd{endnote}
+describe a \bibtex{} record and an Endnote record respectively. We also had
 the example pair \cd{(bib_example , end_example)} where
@@
 
-Since the transformation between the kernel of \cd{bibtex} and the kernel of
-\cd{endnote} that maps \cd{bib_example} to \cd{end_example} can be
-expressed as a bijective lens, we can synthesize a lens that maps between the
-QREs \cd{bibtex} and \cd{endnote}, using the code below:
-
-\begin{lstlisting}
-let bib_to_end : (lens in bibtex <=> endnote) = 
-  synth bibtex <=> endnote
-  using {(bib_example, end_example)}
-\end{lstlisting}
-
-In this case, one example is sufficient to uniquely determine the desired
-transformation.\SAZ{Is that true?}\BCP{Seems more likely that the desired
-  transformation happened to be the first one that the synthesis algorithm
-  reached.  Indeed, my intuition is that {\em every} synthesis problem with
-  a finite set of examples has infinitely many solutions.  Is that right?}
-
-\SAZ{We might want to say something more precise about the running time or
-  expected performance of this algorithm, perhaps to set up the evaluation
-  section.  Something along the lines of: ``Although the algorithm is complete
-  in theory, in the worst case, it may take an unbounded [???IS THIS
-  CORRECT?]\bcp{Surely it's bounded by some exponential formula involving the
-    size of the inputs...?} 
-  amount time to search for a satisfactory lens.  Nevertheless, this algorithm
-  performs well in practice on numerous examples, as we show in the next
-  section.}
+There exists a bijective lens between the kernel of \cd{bibtex} and the kernel
+of \cd{endnote} that maps the normalized form of \cd{bib_example} to the
+normalized form of \cd{end_example}, so calling \textproc{SynthQRELens} on the
+QREs \cd{bibtex} and \cd{endnote}, with the example set of
+$\{(\text{\cd{bib\_example}},\text{\cd{end\_example}})\}$ will return a
+satisfying lens. In this instance, the example set guides the algorithm to also
+find the desired lens.
 

### 2018-03-13 papers/QuotientLenses/main.tex
More polishing in intro

@@
 escapechar=\#,
+mathescape, 
+xleftmargin=1.5em,
 % BCP: I find this distracting:
 % stringstyle=\sffamily\color{dkred},
-mathescape, }
+}
 
@@
 plain bijective lenses\bcp{Do we justify this claim later?}. 
-
+%
 The bidirectional programming language {\em
-Boomerang}~\cite{boomerang} is the only language we know of that
+Boomerang}~\cite{boomerang} 
 implements a general class of quotient lenses.  Unfortunately, these
-lenses to have two main drawbacks.
-
-The first drawback is that defining an equivalence
-relation in Boomerang requires specifying a {\em canonizer}, which is
-a pair of functions $f$ and $g$ where $f$
+lenses to have two main drawbacks.  \bcp{This wording may raise a question
+  in the reader's mind about whether this paper is really about quotient
+  lenses in general or whether it is specifically a proposal to tweak
+  Boomerang.  I.e., ``if I am not a Boomerang user, do I want to bother
+  reading this paper?''  My answer would be yes, because quotient
+  lenses and QREs are interesting in themselves.  But I don't think we've
+  made this point very convincingly in the way we've framed the paper.}
+
+The first shortcoming is that defining an equivalence
+relation in Boomerang requires explicitly providing a {\em canonizer}, 
+a pair of functions $(f,g)$ where $f$
 normalizes data from the
@@
 The disadvantage of this approach is that writing
-a quotient lens in Boomerang requires defining four
+a quotient lens in Boomerang requires separately defining four
 different data formats: two ``external'' formats which specify the data to
 be transformed by the lens, and two ``internal'' formats which specify the
-canonical representations for the source and target data.
+canonical representations for the source and target data\bcp{Plus writing
+  two canonizers!}.
 
@@
 author names:
-\iflastminute
-\bcp{I'd love to use a visible space symbol for important space characters
-  -- this would make the code examples very much clearer!  But it will
-  require a bit of hacking: We can use showstringspaces=false}
-\fi
 %
@@
 
-The second drawback with writing quotient lenses in Boomerang is that
-even after specifying the four required data formats, the programmer must
-still define a lens for converting between the canonical
-representations.   Experience shows that writing such lenses can be
-tedious (consider having to rearrange the order of data items by recursively
-using operators to \kw{swap} adjacent fields)  and that satisfying the type
-checker can be difficult because many seemingly useful lenses contain
-ambiguities disallowed by the Boomerang type checker~\cite{optician}.
+The second drawback with writing quotient lenses in Boomerang is that even
+after specifying the four required data formats, the programmer must still
+define a lens for converting between the canonical representations.
+Experience shows that writing such lenses can be tedious (e.g., consider
+having to rearrange the order of data items by recursively using operators
+to \kw{swap} adjacent fields) and that satisfying the type checker can be
+difficult because the most obvious way to express a lens runs afoul of one
+or more ``ambiguity checks'' that the type checker performs in order to make
+sure that the lens laws hold~\cite{optician}.
 
@@
 \end{lstlisting}
-binds \codefont{l} to a synthesized QRE lens mapping between names in
-the comma-separated form describd by the QRE \codefont{comma\_name}
-and the space-separated form described by the
-QRE \codefont{space\_name}. 
-The lenses generated by
-our synthesis algorithm are guaranteed to translate data back and forth
-between source and target QREs and to satisfy the provided examples.
-Our synthesized quotient lenses have a relatively
-restricted form (which greatly speeds up synthesis), but the 
-normalization theorem means 
-this restriction does not compromise expressiveness.
-
-The following list summarizes the contributions of our work.
+binds \codefont{l} to a synthesized QRE lens mapping between names in the
+comma-separated form describd by the QRE \codefont{comma\_name} and the
+space-separated form described by the QRE \codefont{space\_name}.  The
+lenses generated by our synthesis algorithm are guaranteed to translate data
+back and forth between source and target QREs and to satisfy the provided
+examples.  Our synthesized quotient lenses have a relatively restricted
+form, where the sequential composition operator is not used and all
+canonization occurs at the very edges (which greatly speeds up synthesis),
+but the normalization theorem means this restriction does not compromise
+expressiveness.
+
+In summary, our main contributions are:
 \begin{enumerate}
-  \item We introduce {\em Quotient Regular Expressions} (QREs), which are
+  \item We introduce {\em Quotient Regular Expressions} (QREs), 
   a compact, convenient notation for simultaneously defining a
@@
   for that relation (Section~\ref{QRE}).
-  \item We introduce {\em QRE lenses}, which are lenses that translate
-  between data formats specified using QREs. (Section~\ref{QRE-lenses}).
-  \item We descibe a particular set of QRE lens combinators and prove
-  that QRE lenses defined using these combinators can be normalized
-  (Theorem~\ref{normal form}), which is the main technical
-  contribution of this paper.
-  (Section~\ref{QRE-lenses}).
-  \item We use this result to reduce the problem of {\em synthesizing} QRE
-  lenses to the problem of synthesizing bijective string lenses
-  previously studied in~\cite{optician}(Section~\ref{synth}).
-  \item We extend Boomerang with QREs and QRE lens synthesis. 
-  We use the resulting implementation to demonstrate the utility of
-  our approach by synthesizing QRE lenses between
-  data formats drawn from the Optician benchmark suite (Section~\ref{impl}).
+  \item We introduce {\em QRE lenses}, which translate between data formats
+  specified using QREs (Section~\ref{QRE-lenses}).
+  \item We design a specific set of useful QRE lens combinators and prove
+  that QRE lenses defined using these combinators can be put into a
+  restricted normal form (Section~\ref{QRE-lenses}).  This is
+  the main technical contribution of this paper.
+  \item We leverage this result to reduce the problem of {\em synthesizing} QRE
+  lenses to the problem of synthesizing bijective string lenses, which was
+  previously studied in~\cite{optician} (Section~\ref{synth}).
+  \item We extend Boomerang with QREs and QRE lens synthesis and demonstrate
+  the utility and practicality of our approach by synthesizing QRE lenses
+  between a variety of data formats drawn from the Optician benchmark suite
+  (Section~\ref{impl}).
 \end{enumerate}
-Sections~\ref{relwork} and~\ref{concl} discuss related and future work.
+Sections~\ref{relwork} and~\ref{concl} present related and future work.
 
@@
 \end{lstlisting}
-\caption{An example bijective lens \cd{bib_to_end} between data
+\caption{A plain bijective lens \cd{bib_to_end} between data
 matching \cd{bibtex} and \cd{end_note} regular 
-expressions, which are omitted for brevity.  
+expressions (which are omitted for brevity).  
 \iflastminute

### 2018-03-13 papers/QuotientLenses/main.tex
add correctness proof

--- 2018-03-13 papers/QuotientLenses/main.tex
@@
 
+\bcp{Some of the citations need some love---e.g., the first one.}
+
+\bibliographystyle{plain}
+\bibliography{local}
+
 \appendix
@@
 
-\bcp{Some of the citations need some love---e.g., the first one.}
-\bibliographystyle{plain}
-\bibliography{local}
+\section{Proof of Correctness of \textproc{SynthQRELens}}
+Theorem~\ref{thm:alg-correct} states that
+  Given QREs $Q_1$ and $Q_2$, and a set of examples
+  $\{(x_1,y_1),\ldots,(x_n,y_n)\}$, if there is a QRE lens $q : Q_1
+  \Leftrightarrow Q_2$ such that $q.\get(x_i) \equiv_{Q_2} y_i \text{ and }
+q.\lput(y_i) \equiv_{Q_1} x_i$, then $\textproc{SynthQRELens}(Q_1,Q_2,exs)$ will
+return such a lens.  We prove that here.
+
+\begin{lemma}[Algorithm Soundness]
+  If $q = \textproc{SynthQRELens}(Q_1,Q_2,exs)$, then $q : Q_1 \Leftrightarrow
+  Q_2$ and $q.\get(x_i) \equiv_{Q_2} y_i$  and 
+$q.\lput(y_i) \equiv_{Q_1} x_i$ for all $(x_i,y_i)\in exs$.
+\end{lemma}
+\begin{proof}
+  As $q = \textproc{SynthQRELens}(Q_1,Q_2,exs)$, we know that $q =
+  \rquot(\lquot(Q_1, \ell), Q_2)$. Furthermore, we know that $\ell =
+  \Call{SynthBijectiveLens}{R_1,R_2,exs'}$, where $exs' = \Call{Map}{exs,fun \,
+    (ex_l,ex_r) \to (\canonize(Q_1)(ex_l),\canonize(Q_2)(ex_r))}$. By the correctness of
+  $\textproc{SynthBijectiveLens}$, we know that $\ell.\get(x_i') = y_i'$ and
+  $\ell.\lput(y_i') = x_i'$ for all $(x_i',y_i')\in exs'$. By definitions of
+  $\rquot$ and $\lquot$, this means that:\\
+$\rquot(\lquot(Q_1, \ell),
+  Q_2).\get(x_i)$\\
+ = $\ell.get(\canonize(Q_1)(x_i))$\\ $= \ell.get(x_i') = y_i' =
+  \canonize(Q_2)(y_i) \equiv_{Q_2} y_i$ and\\
+\\
+$\rquot(\lquot(Q_1, \ell),
+  Q_2).\lput(y_i)$\\
+ = $\ell.\lput(\canonize(Q_2)(y_i))$\\ $= \ell.\lput(y_i') = x_i' =
+  \canonize(Q_1)(x_i) \equiv_{Q_1} x_i$ as desired.
+\end{proof}
+
+\begin{lemma}[Algorithm Completeness]
+  If there exists a QRE lens $q : Q_1 \Leftrightarrow Q_2$ such that
+  $q.\get(x_i) \equiv_{Q_2} y_i$ and $q.\lput(y_i) \equiv_{Q_1} x_i$ for all
+  $(x_i,y_i)\in exs$, then $\textproc{SynthQRELens}(Q_1,Q_2,exs)$ terminates.
+\end{lemma}
+\begin{proof}
+  By Theorem~\ref{normal form}, there exists a bijective lens $\ell : K(Q_1)
+  \Leftrightarrow K(Q_2)$, such that $q' = \rquot(\lquot(Q_1, \ell), Q_2)$ is
+  semantically equal to $q$.
+
+  This means that $q'.\get(x_i) \equiv_{Q_2} y_i$ and $q'.\lput(y_i)
+  \equiv_{Q_1} x_i$ for all $(x_i,y_i)\in exs$. Unfolding the definitions of
+  $\rquot$ and $\lquot$, we get:\\
+  $q'.\get(x_i) = \ell.\get(\canonize(Q_1)(x_i)) = \canonize(Q_2)(y_i)$ and\\
+  $q'.\lput(y_i) = \ell.\lput(\canonize(Q_2)(y_i)) = \canonize(Q_1)(x_i)$ for all
+  $(x_i,y_i)\in exs$.
+
+  By correctness of $\textproc{SynthBijectiveLens}$, we know that
+  $\Call{SynthBijectiveLens}{R_1,R_2,exs'}$ terminates, so the entire algorithm
+  terminates. 
+\end{proof}
+
+With soundness and completeness, we trivially get the correctness theorem.
 

### 2018-03-13 papers/QuotientLenses/main.tex
edits in progress of evaluation section; changes to macros for three different measures.

@@
 \newcommand{\QRESize}{\textbf{QS}}
-\newcommand{\canonizeAndSpecSize}{\textbf{CS}}
-\newcommand{\LensAndSpecSize}{\textbf{LS}}
+\newcommand{\canonizeAndSpecSize}{\textbf{BS}}
+\newcommand{\LensAndSpecSize}{\textbf{NS}}
 
@@
 above as an extension to the Boomerang
-interpreter~\cite{boomerang,quotientlenses}. We have added QREs and QRE lens
-synthesis tasks to Boomerang. Furthermore, these synthesis tasks evaluate to
-Boomerang lens values, so Boomerang treats synthesized lenses the same as
-hand-written ones. Intuition suggests that the increased expressiveness
-of QREs should make many lens programming tasks easier, but we would like more
-quantitative understanding of that claim. We thus conducted several experiments
-that aim to provide answers to the following questions:
-
-\begin{enumerate}
-  \item Does synthesizing QRE lenses from QREs permit an easier development
-  process than manually writing canonizers, and synthesizing the lenses between
-  their canonized forms?  Does synthesizing QRE lenses from QREs permit an
-  easier development process than writing the lenses by hand?
-
-  \item Does our tool synthesize lenses fast enough to be used as part of a
-  standard development process?
-\end{enumerate}
-
+interpreter~\cite{boomerang,quotientlenses}. 
+The synthesis algorithm produces Boomerang lens values, so Boomerang
+treats synthesized lenses the same as hand-written ones.
+To evaluate the effectiveness of QREs, QRE lenses, and QRE lens
+synthesis, we conducted experiments to answer the following questions:
+%
+\begin{itemize}
+  \item {\bf Ease of use.} Does synthesizing QRE lenses from QREs permit an
+  easier development process than writing lenses by hand?
+  Does synthesizing QRE lenses from QREs permit an easier development
+  process than manually writing canonizers and then synthesizing lenses between
+  their canonized forms?  
+  \item {\bf Performance.} Is the synthesis algorithm/implementation fast enough to be
+  used as part of a standard development process?
+\end{itemize}
+%
 All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
@@
 
-To evaluate our QRE implementation, we adapted 39 lens synthesis tasks from the
-benchmark suite of the original Optician system. These benchmarks are a
-combination of custom benchmarks, benchmarks derived from
-FlashFill~\cite{flashfill}, and benchmarks derived from Augeas~\cite{augeas2}.
-In the pre-QRE version of Optician, many of these data formats had to be
-modified to work with the bijectivity constraints that Optician required. For
-instance, when one format permits whitespace where the other does not allow it,
-the original versions of benchmark formats were modified to allow more
-whitespace, thereby restoring bijectivity (but altering the data format). With
-the new QRE support, we were able to remove these alterations when porting these
-benchmarks. 10 of Optician's 39 benchmarks had formats modified in this way ---
-introducing QREs removed these format alterations. (This experience alone
-suggests that QREs make the lens development process easier.)
-
-\subsection{Easing the Development Process}
+To evaluate our QRE implementation, we adapted 39 lens synthesis tasks
+from the benchmark suite of the original Optician system. These
+benchmarks are a combination of custom benchmarks, benchmarks derived
+from FlashFill~\cite{flashfill}, and benchmarks derived from
+Augeas~\cite{augeas2}.  In the pre-QRE version of Optician, 10 of
+these data formats had to be modified to work with the bijectivity
+constraints that Optician required. For instance, when one
+representation permits whitespace where the other does not, we
+modified the original version of the benchmark to allow more whitespace,
+thereby restoring bijectivity (but altering the data format). With the
+new QRE support, we were able to remove these alterations.  This
+experience alone suggests that QREs make the lens development process
+more flexible.
+
+\subsection{Ease of Use}
 
@@
 
-To get a sense of how much the new expressiveness that combining QRE lenses with
-synthesis saves in terms of programmer effort, we compared the amount of code
-that would need to be written for each of the lens benchmarks. We use number of
-nodes in the abstract syntax tree as a measure of code size, and calculated
-three quantities for each lens synthesis benchmark task. We only compared the
-sizes on benchmarks where the canonization functions are nontrivial.
+To evaluate the impact of QRE lens synthsis on programmer effort,
+we focus our attention on the 10 problems in the benchmark suite that
+are not bijective and hence require non-trivial canonizers. 
+(Optician already handles the other problems with minimal programmer
+effort.)
+
+We are interested in comparing three different approaches, which vary
+in the amount of synthesis used. 
+In the first approach, which we call \QRESize{} for QRE Synthesis, the programmer uses
+QRE synthesis.  She must write QRE specifications of the source and
+target formats and she may give examples.
+In the second approach, which we call \canonizeAndSpecSize{} for
+Bijective Synthesis, the
+programmer uses bijective lens synthesis \`a la the original version
+of Optician.  She must write canonizers by hand, along with the
+regular expressions describing the external
+representations of the source and target formats. (The internal
+regular expressions can be inferred from the canonizers.) She may also
+provide examples to help in the synthesis of the bijective lens.
+In the third approach, which we call \LensAndSpecSize{} for No Synthesis, the
+programmer writes the lens between the source and target lenses
+entirely by hand.
+
+To measure the programmer effort using each of these approaches, we
+calculate three different quantities for each problem in the benchmark suite:
 %
 \begin{itemize}
-  \item[\QRESize{}] the number of AST nodes in the QRE
-  specifications (including examples).
-  \item[\canonizeAndSpecSize{}]  the number of
-  AST nodes in $W(q_1)$ summed with the number of AST nodes in $\canonize(q_1)$
-  and the number of examples.
-  \item[\LensAndSpecSize{}] the number of AST nodes
-  in $W(q_1)$ summed with the number of AST nodes in the synthesized QRE lens.
+  \item[\QRESize{}:] 
+  The number of AST nodes in the QRE specifications for the source and
+  target formats, including examples. 
+  \item[\canonizeAndSpecSize{}:] 
+  The sum of (1) the number of AST nodes in $W(q)$ for each QRE $q$ in the source and target
+  formats, (2) the number of AST nodes $\canonize(q)$ for each such QRE $q$ with a
+  non-trivial canonizer, and (3) the number of AST nodes in the
+  examples.  We use (1) and (2) to estimate the burden of describing
+  the external source and target formats and of writing the requisite canonizers
+  by hand.  We count the nodes in the examples because they would be
+  fed to the bijective synthesizer.  
+  These counts are an approximation, as both $W(q)$ and $\canonize(q)$ are
+  automatically generated from the QRE, and it is possible that a human-written
+  version might be smaller.
+  \item[\LensAndSpecSize{}:] The sum of (1) the number of AST nodes in
+  $W(q)$ for each QRE $q$ in the source and target formats and (2) the
+  number of AST nodes in the synthesized QRE lens.  We use (1) to
+  estimate the burdern for describing the source and target formats
+  and (2) to estimate the burdern for writing the appropriate lens by
+  hand. These counts are also approximations, as
+  $W(q)$ and the synthesized lens may be larger than one written by hand.
 \end{itemize}
-
-The value of \QRESize{} is our measure of the programmer burden for
-writing our benchmarks---it measures the number of AST nodes the programmer has
-to write using the full expressiveness of QREs and synthesis.
-
-The value of \canonizeAndSpecSize{} estimates the burden for writing our
-benchmarks within the Optician system, without QREs but still using
-synthesis. (This is only an approximation, as both $W(q_1)$ and $\canonize(q_1)$ are
-automatically generated from the QRE, and it is possible that a human-written
-version might be smaller.)
-
-The value of \LensAndSpecSize{} estimates the burden in writing our benchmarks
-in Boomerang, using neither QREs nor synthesis. (This is also an
-approximation, as $W(q_1)$ and $\canonize(q_1)$ and the bijective lens are all
-automatically generated.)
 

### 2018-03-13 papers/QuotientLenses/transform-data.py
update labels on graph

@@ -100,7 +100,7 @@ def generate_ast_sizes_graph(input_csv):
     ax.set_xticks(ind + width*1.5)
     ax.set_xticklabels(project_column_from_csv(input_csv, "Test"),rotation=45)
 
-    ax.legend((rects1[0], rects2[0], rects3[0]), ('\\textbf{QS}', '\\textbf{CS}', '\\textbf{LS}'))
+    ax.legend((rects1[0], rects2[0], rects3[0]), ('\\textbf{QS}', '\\textbf{BS}', '\\textbf{NS}'))
 
     fig = plt.figure(1,tight_layout=True)
     fig.set_figheight(2)

### 2018-03-13 papers/QuotientLenses/main.tex

@@
 
+%\newcommand{\QOpt}{Optician_Q}
+\newcommand{\QOpt}{QRE-enhanced Optician}
 \newcommand{\OpticianRuntime}{\textbf{OO}}
+\newcommand{\QREOptician}{\textbf{OQ}}
 \newcommand{\SystemOnOptician}{\textbf{QO}}
@@
 \includegraphics{generated-graphs/asts.eps}
-\caption{Count of benchmark programs definable using a given AST count. We find
-that it takes far fewer AST nodes to define benchmark lenses using QRE
-synthesis than with Optician without QRE synthesis or without synthesis.}
+\caption{AST node measurements for each of the three approaches
+on each of the 10 non-bijective benchmark problems.
+QRE Synthesis requires far fewer AST nodes than the other
+two approaches. \ksf{Anders, is there an order to the benchmarks in
+this list?  It might be useful to list them in order of increasing
+complexity of the format, perhaps measured by the sum of the number of
+AST nodes in the source and target format descriptions.  It seems
+better to have some order than just listing them randomly, even if it
+doesn't show much interesting.}}
+%%\caption{Count of benchmark programs definable using a given AST count. We find
+%%that it takes far fewer AST nodes to define benchmark lenses using QRE
+%%synthesis than with Optician without QRE synthesis or without synthesis.}
 \label{fig:asts}
@@
 In the first approach, which we call \QRESize{} for QRE Synthesis, the programmer uses
-QRE synthesis.  She must write QRE specifications of the source and
+QRE lens synthesis.  She must write QRE specifications of the source and
 target formats and she may give examples.
@@
 programmer uses bijective lens synthesis \`a la the original version
-of Optician.  She must write canonizers by hand, along with the
-regular expressions describing the external
+of Optician.  She must write canonizers by hand, along with 
+regular expressions to describe the external
 representations of the source and target formats. (The internal
-regular expressions can be inferred from the canonizers.) She may also
+formats can be inferred from the canonizers.) She may also
 provide examples to help in the synthesis of the bijective lens.
 In the third approach, which we call \LensAndSpecSize{} for No Synthesis, the
-programmer writes the lens between the source and target lenses
-entirely by hand.
-
-To measure the programmer effort using each of these approaches, we
-calculate three different quantities for each problem in the benchmark suite:
+programmer writes the lens between the source and target formats
+entirely by hand, including the descriptions of the source and target
+formats.
+
+For each problem in the benchmark suite, we calculate the following
+measures as proxies for the level of programmer effort when using each
+the three approaches:
+
 %
@@
   The sum of (1) the number of AST nodes in $W(q)$ for each QRE $q$ in the source and target
-  formats, (2) the number of AST nodes $\canonize(q)$ for each such QRE $q$ with a
+  formats, (2) the number of AST nodes in $\canonize(q)$ for each QRE $q$ with a
   non-trivial canonizer, and (3) the number of AST nodes in the
-  examples.  We use (1) and (2) to estimate the burden of describing
-  the external source and target formats and of writing the requisite canonizers
+  examples.  We use (1) to estimate the burden of describing
+  the external source and target formats and (2) to estimate the
+  burdern of writing the requisite canonizers
   by hand.  We count the nodes in the examples because they would be
@@
   These counts are an approximation, as both $W(q)$ and $\canonize(q)$ are
-  automatically generated from the QRE, and it is possible that a human-written
-  version might be smaller.
+  automatically generated from the corresponding QRE $q$, and it is
+  possible that a human-written version might be smaller.
   \item[\LensAndSpecSize{}:] The sum of (1) the number of AST nodes in
@@
   number of AST nodes in the synthesized QRE lens.  We use (1) to
-  estimate the burdern for describing the source and target formats
-  and (2) to estimate the burdern for writing the appropriate lens by
+  estimate the burdern of describing the source and target formats
+  and (2) to estimate the burdern of writing the appropriate lens by
   hand. These counts are also approximations, as
@@
 
-Figure~\ref{fig:asts} shows the number of benchmarks that can be expressed using
-at most a given number of AST nodes. On average, \canonizeAndSpecSize{} used
-44\% more AST nodes than \QRESize{}, using an average of 237 more AST nodes. On
-average, \LensAndSpecSize{} used 193\% more AST nodes than \QRESize{}, using an
-average of 1267 more AST nodes. This suggests that introducing QREs saves
+\ksf{Anders: are these figures up to date?}
+Figure~\ref{fig:asts} shows each of these measures for the 10
+non-bijective problems in the benchmark suite.  On
+average, \canonizeAndSpecSize{} used~44\% more AST nodes
+than \QRESize{}, requiring an average of~237 more AST nodes. On 
+average, \LensAndSpecSize{} used~193\% more AST nodes than \QRESize{}, requiring an
+average of~1267 more AST nodes. These figures suggest that introducing QREs saves
 programmers significant effort compared to both pre-QRE Optician and basic
@@
 \end{subfigure}
-\caption{Runtimes of Optician and our system. In (a), we run Optician and our
-  system on Optician's benchmarks. We find that there is only a negligible
-  performance overhead incurred by using QREs. In (b), we run Optician with QREs
-  on the Optician benchmarks we modified to use QREs. We find that it is able to
-  synthesize all quotient lenses in under 20 seconds, and typically finished in
+\caption{Runtimes measurements. In (a), we run the
+  original version of Optician (\OpticianRuntime{}) and \QOpt{}
+  (\QREOptician{}) on the Optician benchmarks.  We find that there is only a negligible
+  performance overhead incurred by using QREs. 
+  In (b), we run \QOpt{} on the 10 Optician benchmarks
+  previously edited to make them bijective, after removing those edits
+  (In other words, we restored them to their original state and then
+  ran \QOpt{}). 
+  We find that \QOpt{} is able to
+  synthesize all quotient lenses in under 20 seconds, and typically finishes in
   under 5 seconds.}
@@
 
-To assess the performance impact of adding QRE support to Optician, we measured
-the running time of the following configurations:
-\begin{itemize}
-  \item[\OpticianRuntime{}] pre-QRE Optician running on its own benchmarks.
-  \item[\SystemOnOptician{}] Optician with QREs running on the pre-QRE
-    benchmarks.
-  \item[\SystemOnBenchmarks{}] Optician with QREs running on the benchmarks that
-    were altered from Optician to include quotients.
-\end{itemize}
-
-We summarize the results of these experiments in Figure~\ref{fig:times}.  The
-new version of Optician was able to synthesize all of pre-QRE Optician's
-benchmarks at a speed competitive with the old version.  There is a small amount
-of additional overhead introduced by QREs when calculating $W(q_1)$ and $K(q_1)$,
-resulting in a very slight decrease in performance.
+To assess the performance of QRE synthesis, we are interested in two
+different questions.  First, how does the performance of \QOpt{} 
+compare to the performance of the original
+version of Optician on benchmarks that do not require QREs?  The
+answer to this question tells us how much overhead we have introduced
+by adopting the more general mechanism.  
+Figure~\ref{fig:times}(a) shows that \QOpt{} was able to synthesize
+all of the Optician benchmarks at a speed competative with the old
+version. 
+There is a small amount
+of additional overhead introduced by QREs in calculating the $W$ and
+$K$ functions, resulting in a slight decrease in performance.
+
+Second, how much time does it take for \QOpt{} to synthesize a QRE
+lens when running on a non-bijective benchmark problem?  
+Figure~\ref{fig:times}(b) shows the amount of time required to infer a
+lens for each of the 10 benchmark programs.  
+We find that \QOpt{} is able to synthesize all quotient lenses in
+under 20~seconds, and typically finishes in under 5~seconds.
 

### 2018-03-13 papers/QuotientLenses/main.tex
minor textual changes:

@@
 \newcommand{\QOpt}{QRE-enhanced Optician}
-\newcommand{\OpticianRuntime}{\textbf{OO}}
-\newcommand{\QREOptician}{\textbf{OQ}}
+\newcommand{\OpticianRuntime}{\textbf{Optician}}
+\newcommand{\QREOptician}{\textbf{Optician\textsubscript{Q}}}
 \newcommand{\SystemOnOptician}{\textbf{QO}}
@@
 QRE Synthesis requires far fewer AST nodes than the other
-two approaches. \ksf{Anders, is there an order to the benchmarks in
-this list?  It might be useful to list them in order of increasing
-complexity of the format, perhaps measured by the sum of the number of
-AST nodes in the source and target format descriptions.  It seems
-better to have some order than just listing them randomly, even if it
-doesn't show much interesting.}}
+two approaches.}
 %%\caption{Count of benchmark programs definable using a given AST count. We find
@@
 
-\ksf{Anders: are these figures up to date?}
 Figure~\ref{fig:asts} shows each of these measures for the 10
@@
   previously edited to make them bijective, after removing those edits
-  (In other words, we restored them to their original state and then
-  ran \QOpt{}). 
+  (In other words, we restored them to their original state with QREs
+  and then ran \QOpt{}). 
   We find that \QOpt{} is able to
-  synthesize all quotient lenses in under 20 seconds, and typically finishes in
+  synthesize all quotient lenses in under 10 seconds, and typically finishes in
   under 5 seconds.}
@@
 Figure~\ref{fig:times}(b) shows the amount of time required to infer a
-lens for each of the 10 benchmark programs.  
+lens for each of the 10 benchmark programs with nontrivial quotients.  
 We find that \QOpt{} is able to synthesize all quotient lenses in
-under 20~seconds, and typically finishes in under 5~seconds.
+under 10~seconds, and typically finishes in under 5~seconds.
 

### 2018-03-13 papers/QuotientLenses/transform-data.py

@@ -84,6 +84,7 @@ def generate_ast_sizes_graph(input_csv):
     width = 0.25       # the width of the bars
 
     fig, ax = plt.subplots()
+    input_csv.sort(key=lambda x : int(x["OO"]))
 
     qre_optician_means = project_column_from_csv(input_csv, "OO")
     rects1 = ax.bar(ind, qre_optician_means, width, color='r')
@@ -100,7 +101,7 @@ def generate_ast_sizes_graph(input_csv):
     ax.set_xticks(ind + width*1.5)
     ax.set_xticklabels(project_column_from_csv(input_csv, "Test"),rotation=45)
 
-    ax.legend((rects1[0], rects2[0], rects3[0]), ('\\textbf{QS}', '\\textbf{BS}', '\\textbf{NS}'))
+    ax.legend((rects1[0], rects2[0], rects3[0]), ('\\textbf{QS}', '\\textbf{BS}', '\\textbf{NS}'),bbox_to_anchor=(0.92,.922),borderaxespad=0)
 
     fig = plt.figure(1,tight_layout=True)
     fig.set_figheight(2)

### 2018-03-13 papers/QuotientLenses/main.tex
extended caption on ast graph
fixed caption in performance graph

@@
 on each of the 10 non-bijective benchmark problems.
+Benchmarks are sorted in order of increasing complexity as measured by
+the number of AST nodes in the source and target format descriptions. 
 QRE Synthesis requires far fewer AST nodes than the other
@@
   previously edited to make them bijective, after removing those edits
-  (In other words, we restored them to their original state with QREs
+  and then extending the synthesis specification to include QREs.
+  (In other words, we restored them to their original state, added QREs,
   and then ran \QOpt{}). 

### 2018-03-13 papers/QuotientLenses/main.tex
in progress edits to section 4

@@
 with syntax that expresses an equivalence relation on the language of $R$. This
-section describes in further detail the set of QRE combinators that we described
+section formalizes the set of QRE combinators that we introduced informally
 in \cref{sec:example}.
@@
 
-\subsection{Syntax of Semantics of QREs}
+\subsection{Syntax and Semantics of QREs}
 Formally, the language of Quotient Regular Expressions (QREs) is given by the
@@
 \begin{align*}
-Q := \; & \normalize{R}{R'}{f} \bsep{} \id(R) \bsep \project{R}{s} \bsep \squash{Q_1}{Q_2}{f} \\
+Q := \; & \normalize{R_1}{R_2}{f} \bsep{} \id(R) \bsep \project{R}{s} \bsep \squash{Q_1}{Q_2}{f} \\
  | \; & \perm{Q_1, \ldots, Q_n}{Q} \bsep Q_1 \semicolon Q_2 \bsep Q_1 \cdot Q_2 \bsep (Q_1
-\sep Q_2) \bsep Q^*,
-\end{align*}
-where $Q$ ranges over QREs, $R$ ranges over regular expressions, $f$ ranges over
-functions between regular languages, and $s$ ranges over character strings.
+\sep Q_2) \bsep Q^*
+\end{align*}
+where $Q$ ranges over QREs, 
+$R$ ranges over regular expressions, 
+$f$ ranges over functions between regular languages, and 
+$s$ ranges over character strings.
 
 Using the conventional notation that $\mathcal{L}(R)$ is the language accepted
-by the regular expression $R$, each QRE $Q$ enables us to express four semantic objects
+by the regular expression $R$, each QRE $Q$ yields four semantic
+objects:
 
 \begin{tabular}{rl}
-  $W(Q)$  & a regular expression (the ``whole'' of $Q$),\\
-  $\eqrel{Q}$  & an equivalence relation on $\mathcal{L}(W(Q))$,\\
-  $K(Q)$ & a regular expression (the ``kernel'' of $Q$) such that
-           $\mathcal{L}(K(Q))$ forms a complete \\
-             & set of representatives for $\eqrel{Q}$, and \\
+  $W(Q)$       & A regular expression, denoting the ``whole'' of $Q$\\
+  $\eqrel{Q}$  & An equivalence relation on $\mathcal{L}(W(Q))$\\
+  $K(Q)$       & A regular expression, denoting the ``kernel'' of $Q$, such that
+                  $\mathcal{L}(K(Q))$ forms a \\
+               & complete set of representatives for $\eqrel{Q}$ \\
   $\canonize(Q)$ & 
-  a ``canonizing'' function that, given any $w \in \mathcal{L}(W(Q))$,
-  yields $\canonize(Q)(w)$ \\ & as the unique $k$ in $\mathcal{L}(K(Q))$ such
+  A ``canonizing'' function. Given any $w \in \mathcal{L}(W(Q))$,
+  $\canonize(Q)(w)$ \\ & is the unique $k$ in $\mathcal{L}(K(Q))$ such
   that $k$ is equivalent to $w$ mod $\eqrel{Q}$
@@
 
+\noindent
+Intuitivley, $W(Q)$ is the regular expression representing the
+external format, while $K(Q)$ is the regular expression representing
+the internal format.  The equivalence relation $\eqrel{Q}$ groups
+together elements in the language of $W(Q)$ that contain the same
+essential information.  The function $\canonize(Q)$ picks the
+representative element from each of the resulting equivalence classes.
+
 The well-formedness constraints for QREs ensure that these four semantic
-components fit together to form a coherent quotient $W(Q)/\eqrel{Q}$ whose
+objects fit together to form a coherent quotient $W(Q)/\eqrel{Q}$ whose
 equivalence classes are determined by \canonize(Q).
@@
 The relationship among the semantic objects of a QRE can be understood in terms
-of the \normalize{R}{S}{f} combinator, which exposes each of these pieces
-explicitly.  Its whole language is just $R$, its kernel language is just
-$S$, and $\canonize(\normalize{R}{S}{f}) = f$.  Its equivalence relation is
+of the combinator \normalize{R_1}{R_2}{f}, which expresses each of these pieces
+explicitly.  Its whole language is just $R_1$, its kernel language is just
+$R_2$, and its canonizer is just $f$; 
+its equivalence relation $\equiv$ is
 determined by the fibres of $f$, so we have
-$s_1 \equiv s_2 \Leftrightarrow f(s_1) = f(s_2)$.
+$s_1 \equiv s_2 \Leftrightarrow f(s_1) = f(s_2)$ for $s_1$ and $s_2$
+in $\mathcal{L}(R_1)$.
 
@@
 \begin{prooftree}
-\AxiomC{$\mathcal{L}(S) \subseteq \mathcal{L}(R)$}
-\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(S)$}
+\AxiomC{$\mathcal{L}(R_2) \subseteq \mathcal{L}(R_1)$}
+\AxiomC{$f : \mathcal{L}(R_1) \longrightarrow \mathcal{L}(R_2)$}
 \AxiomC{$f$ is surjective}
@@
 \RightLabel{(Normalize)}
-\QuaternaryInfC{$\wf{\normalize{R}{R'}{f}}$}
+\QuaternaryInfC{$\wf{\normalize{R_1}{R_2}{f}}$}
 \end{prooftree}
@@
 surjective, idempotent function 
-$f : \mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$.   However,
-although \kw{normalize} is powerful, verifying that a canonization function $f$
-is surjective and idempotent is in general undecidable.  This means that a
-programmer wishing to use \kw{normalize} must discharge strong proof
-obligations, which is cumbersome in practice.\footnote{In our implementation we allow a programmer to use \kw{normalize} at their
-own risk without checking these side conditions as an ``escape hatch'' for the
-case when other QRE combinators are insufficient.\bcp{Do we use this in any
-  of the programs we've written?}}
-
-The remaining QRE combinators provide simpler, more compositional ways of
-building canonizers.  Nevertheless, the \kw{normalize} combinator gives a
-sufficient condition for potential QREs to be well-behaved, particularly with
-respect to the strategy for synthesizing quotient lenses which we shall
-introduce in Section~\ref{synth}, so we use it to guide the design of our QRE
-combinators, as we see next.
+$f : \mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$.  
+However,
+verifying that a canonization
+function $f$ is surjective and idempotent is in general undecidable.
+Consequently, a programmer wishing to use \kw{normalize} must
+discharge strong proof obligations, which is cumbersome in
+practice.\footnote{In our implementation we allow a programmer to
+use \kw{normalize} at their own risk without checking these side
+conditions as an ``escape hatch'' for the case when other QRE
+combinators are insufficient.\bcp{Do we use this in any of the
+programs we've written?}}
+
+The remaining QRE combinators, which we discuss next, provide simpler,
+more compositional ways of building canonizers.  Nevertheless,
+the \kw{normalize} combinator provides a useful guide in the design of
+these combinators because it gives a sufficient condition for the
+well-formedness of any potential QREs.
+
+% particularly with respect to the strategy for synthesizing quotient
+% lenses (See Section~\ref{synth}), so we use it to guide the design of
+% our QRE remaining combinators, which we discuss next.
 
@@
 Figure~\ref{fig:wk} gives the inductive definitions of the whole and
-kernel languages $W(Q)$ and $K(Q)$ for all of the QRE forms.
+kernel languages $W(Q)$ and $K(Q)$ for all of the QRE combinators.
 \begin{figure}[t]
@@
 \[
-\begin{array}{l@{\quad}l@{\quad}l}
+\begin{array}{l@{\quad\quad}l@{\quad\quad}l}
 
@@
 \project{R}{s} & R & s \\
-\squash{Q}{Q_1}{f} & W(Q) \sep W(Q_1) & K(Q_1) \\
-\normalize{R}{R'}{f} & R & R' \\
+\squash{Q_1}{Q_2}{f} & W(Q_1) \sep W(Q_2) & K(Q_2) \\
+\normalize{R_1}{R_2}{f} & R_1 & R_2 \\
 Q_1 \; ; \; Q_2 & W(Q_1) & K(Q_2) \\
-Q_1 \cdot Q_2 & W(Q_1) \cdot W(Q_2) & K(Q_1) \cdot K(Q_2) \\
-Q_1 \sep Q_2 & W(Q_1) \sep W(Q_2) & K(Q_1) \sep K(Q_2) \\
+Q_1 \cdot\; Q_2 & W(Q_1) \cdot W(Q_2) & K(Q_1) \cdot K(Q_2) \\
+Q_1 \sep \; Q_2 & W(Q_1) \sep W(Q_2) & K(Q_1) \sep K(Q_2) \\
 Q^* & W(Q)^* & K(Q)^* \\
@@
 W( \perm{Q_1, \ldots, Q_n}{Q} ) = &
-\bigcup \limits_{\sigma \in S_n} W(c_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
-W(Q) \cdot W(c_{\sigma(n)})\\
+\bigcup \limits_{\sigma \in S_n} W(Q_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
+W(Q) \cdot W(Q_{\sigma(n)})\\
 K( \perm{Q_1, \ldots, Q_n}{Q} ) = & K(Q_1) \cdot K(Q) \cdot \ldots \cdot K(Q)
@@
 \]
-\caption{Whole and Kernel Regular Expressions}
+\caption{Whole and Kernel Regular Expressions for QRE combinators. 
+In describing regular expressions, we use the notations $\sep$ and $\bigcup$ for
+alternation, the notation $\cdot$ for concatenation, and the notation
+$*$ for Kleene closure.  
+We use the notation $S_n$ to denote the set of all permutations of the
+numbers $1$ to $n$.
+}
 \label{fig:wk}
@@
 interesting definitions. If $Q = \squash{Q_1}{Q_2}{f}$, then the whole language
-of $Q$ is $W(Q_1) \sep W(Q_2)$. This is because the \kw{squash} combinator
-takes the whole language $W(Q_1)$ of $Q_1$, the whole language $W(Q_2)$ of $Q_2$
-and a function $f : \mathcal{L}(W(Q_1)) \longrightarrow \mathcal{L}(W(Q_2))$,
+of $Q$ is $W(Q_1) \sep W(Q_2)$ because the \kw{squash} combinator
+merges the whole language $W(Q_1)$ of $Q_1$ with the whole language $W(Q_2)$ of $Q_2$.
+The function $f : \mathcal{L}(W(Q_1)) \longrightarrow \mathcal{L}(W(Q_2))$,
 maps $\mathcal{L}(W(Q_1))$ into $\mathcal{L}(W(Q_2))$ using $f$ and then
@@
 For the $\perm{Q_1, \ldots, Q_n}{Q}$ combinator, the whole language is the
-union of languages of the form $W(c_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
-W(Q) \cdot W(c_{\sigma(n)})$ for any permutation $\sigma$ in $S_n$, where
-$S_n$ is the set of all permutations of the numbers $1$ to $n$. This is because
+union of languages of the form $W(Q_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
+W(Q) \cdot W(Q_{\sigma(n)})$ for any permutation $\sigma$ in $S_n$, where
+$S_n$ is the set of all permutations of the numbers $1$ to $n$. Intuitively,
 the \kw{permutation} combinator allows for the string to match any
-permutation the $Q_i$'s while also taking into account the separator $Q$ in
+permutation of the $Q_i$'s while preserving the separator $Q$ in
 between each of the $Q_i$'s. The kernel of the \kw{permutation} combinator
@@
 because the canonical permutation is the identity permuation, with each of the
-parts of the string that match $Q_i$ and $Q$ canonized into $K(Q_i)$ and $K(Q)$
+parts of the input that match $Q_i$ and $Q$ canonized into $K(Q_i)$ and $K(Q)$
 respectively.
@@
 \canonize(\id(R)) &=& id_{\mathcal{L}(R)} \\
-\canonize(\project{R}{s})(w) &=& s\\
+\canonize(\project{R}{s})(w) &=& s\\[1ex]
 \canonize(\squash{Q_1}{Q_2}{f})(w) &=&
@@
 \canonize(Q_2)(w) & \text{otherwise}
-\end{cases}\\
-\canonize(\normalize{R}{R'}{f}) &=& f\\
+\end{cases}\\[3ex]
+\canonize(\normalize{R_1}{R_2}{f}) &=& f\\
 \canonize(Q_1 \; ; \; Q_2) &=& \canonize(Q_2) \circ \canonize(Q_1)\\
-\canonize(Q_1 \cdot Q_2) &=& \canonize(Q_1) \cdot \canonize(Q_2)\\
+\canonize(Q_1 \cdot Q_2) &=& \canonize(Q_1) \cdot \canonize(Q_2)\\[1ex]
 \canonize(Q_1 \sep Q_2)(w) &=&
@@
 \canonize(Q_2)(w) & \text{if } w \in \mathcal{L}(W(Q_2))\\
-\end{cases}\\
+\end{cases}\\[3ex]
 \canonize(Q^*) &=& \canonize(Q)^* \\
@@
 &= (\canonize(Q_1)(r_1)) \cdot (\canonize(Q)(s_1)) \cdot \ldots \cdot
+(\canonize(Q)(s_{n-1}))\cdot
 (\canonize(Q_n)(r_n))
 \end{align*}
-\caption{QRE canonizers}
+\caption{QRE canonizers. \ksf{Solomon: please define/explain the . and *
+operators on functions.}}
 \label{fig:canonizers}
 \end{figure}
-Again, the \kw{permutation} combinator gives the most interesting
-definition,
+The \kw{permutation} combinator gives rise to the most interesting
+definition:
 \begin{align*}\canonize(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
@@
 = (\canonize(Q_1)(r_1)) \cdot (\canonize(Q)(s_1)) \cdot \ldots \cdot
+(\canonize(Q)(s_{n-1}))\cdot
 (\canonize(Q_n)(r_n))
 \end{align*}
-\noindent which, as we have explained, is defined so that the strings that match
-$Q_i, Q$ are placed in the canonical permutation which is the identity
-permutation before $\canonize(Q_i)$ and $\canonize(Q)$ are applied
-respectively.
+\noindent 
+which places the strings that match $Q_i$ and $Q$ according to the
+canonical permutation (i.e., the identity permutation) before applying
+the canonizers $\canonize(Q_i)$ and $\canonize(Q)$, respectively.
 
 Finally, Figure ~\ref{fig:relations} gives the inductive definition of the
-equivalence relation $\eqrel{Q}$. This is the set-theoretic semantics of QREs
-as an equivalence relation on a regular language.
+equivalence relation $\eqrel{Q}$, which is the set-theoretic semantics
+of a QRE $Q$ as an equivalence relation on the regular language $W(Q)$.
 
@@
 \begin{array}{l@{}l@{\ \ }l}
-w \; \equiv_{\normalize{R}{R'}{f}} \; w' &\iff&
+w \; \equiv_{\normalize{R_1}{R_2}{f}} \; w' &\iff&
 f(w)=f(w') \\

### 2018-03-13 papers/QuotientLenses/main.tex
finished pass over section 4

@@
 w \; \equiv_{\id(R)} \; w' &\iff& w = w' \\
-w \; \equiv_{\project{R}{s}} \; w' && \text{for all }w, w'\\
+w \; \equiv_{\project{R}{s}} \; w' &\iff& True \\ % \text{for all }w, w'\\
 w \; \equiv_{\squash{Q_1}{Q_2}{f}} \; w' &\iff& f(w) \equiv_{Q_2} w'
@@
 \end{figure}
-\subsection{Ambiguity and Well Formed QREs}
+\subsection{Ambiguity and Well-Formed QREs}
 \label{subsec:well-formed-qres}
-An issue that we address before presenting the full set of inference
-rules for QREs is the issue of {\em ambiguity}, which appears in the
-premises of the inference rules for the regular combinators. If $Q_1$ and $Q_2$
-are QREs, then when applying the regular combinators to QREs, we require that
+
+To ensure that regular combinations of QREs are well-formed, we need
+to enforce a variety of {\em unambiguity} constraints.  
+Specifically, when applying regular combinators to QREs $Q_1$ and $Q_2$, we require that
 if a string $s$ matches any of the regular expressions $W(Q_1) \cdot W(Q_2)$,
@@
 expression in only one way.
-
 Formally, we say that regular expressions $R$ and $S$ are
@@
 because it ensures that the canonizing function of a QRE is well-defined.
-For example consider the QRE \lstinline{"a"* . (project "a"* $\to$ "a")}. The
-behaviour of this QRE is not-well defined since the string
+For example, consider the QRE \lstinline{"a"* . (project "a"* $\to$ "a")}. The
+behaviour of this QRE is not well-defined since the string
 \cd{"aaa"} can be canonized to any of \cd{"a"}, \cd{"aa"}, or \cd{"aaa"}
@@
 functions.
+
+\subsection{Well-formedness of QREs}
 
@@
 \RightLabel{(Perm)}
-\AxiomC{${\substack{\forall \sigma \neq \theta, \; W(c_{\sigma(1)}) \cdot W(Q)
-\cdot \ldots \cdot W(c_{\sigma(n)})\\ \cap W(c_{\theta(1)}) \cdot W(Q)
-\cdot \ldots \cdot W(c_{\theta(n)}) =\varnothing}}$}
+\AxiomC{${\substack{\forall \sigma \neq \theta, \; W(Q_{\sigma(1)}) \cdot W(Q)
+\cdot \ldots \cdot W(Q_{\sigma(n)})\\ \cap W(Q_{\theta(1)}) \cdot W(Q)
+\cdot \ldots \cdot W(Q_{\theta(n)}) =\varnothing}}$}
 \AxiomC{$K(Q_1) \cdot^! K(Q) \cdot^! \ldots \cdot^! K(Q) \cdot^! K(Q_n)$}
@@
 \end{figure}
-As we already mentioned, the unambiguity conditions come into play when defining
-QREs using the regular combinators. For example, the (Concat) inference rule
-for concatenation says that the concatenation $Q_1 \cdot Q_2$ of QREs $Q_1$ and
-$Q_2$ is well formed only if the concatentions of $W(Q_1)$ and $W(Q_2)$, and
-$K(Q_1)$ and $K(Q_2)$ are unambiguous.
+The unambiguity conditions are pertinent when defining QREs using the
+regular combinators. For example, the (Concat) inference rule 
+says that the concatenation $Q_1 \cdot Q_2$ of QREs
+$Q_1$ and $Q_2$ is well formed only if the concatentions of $W(Q_1)$
+and $W(Q_2)$, and $K(Q_1)$ and $K(Q_2)$ are unambiguous.
 
@@
 Perm rule says that for any two different permutations $\sigma$
-and $\theta$, the languages $W(c_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
-W(c_{\sigma(n)})$ and $W(c_{\theta(1)}) \cdot W(Q) \cdot \ldots \cdot
-W(c_{\theta(n)})$ must be disjoint. This is because any string that matches the
-\kw{permutation} combinator matches the language $W(c_{\theta(1)}) \cdot
-W(Q) \cdot \ldots \cdot W(c_{\theta(n)})$ for some permutation $\theta$, so we
-require that these regular expressions be disjoint.
+and $\theta$, the languages $W(Q_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
+W(Q_{\sigma(n)})$ and $W(Q_{\theta(1)}) \cdot W(Q) \cdot \ldots \cdot
+W(Q_{\theta(n)})$ must be disjoint. This restriction is important
+because any string that matches the 
+\kw{permutation} combinator matches the language $W(Q_{\theta(1)}) \cdot
+W(Q) \cdot \ldots \cdot W(Q_{\theta(n)})$ for some permutation $\theta$, so we
+must require that these regular expressions be disjoint.
 %
 The third hypothesis says that the regular expression $K(Q_1) \cdot K(Q)
-\cdot \ldots \cdot K(Q) \cdot K(Q_n)$ must be unambiguous. This is because the
+\cdot \ldots \cdot K(Q) \cdot K(Q_n)$ must be unambiguous. This
+restriction arises because the
 underlying lens that maps to or from a \kw{permutation} QRE operates on
-the language $K(Q_1) \cdot K(Q) \cdot \ldots \cdot K(Q) \cdot K(Q_n)$ which is
-the kernel of the \kw{permutation} QRE. As we mentioned, the underlying
-lens will require that the source and target regular expressions of the lens be
+the language $K(Q_1) \cdot K(Q) \cdot \ldots \cdot K(Q) \cdot K(Q_n)$, 
+the kernel of the \kw{permutation} QRE. The underlying
+lens requires that the source and target regular expressions be
 unambiguous so that the lens can match strings uniquely.

### 2018-03-13 papers/QuotientLenses/main.tex
partial pass over section 5

@@
 regular expressions, a natural next step in defining quotient lenses is to map
-between two QREs via a bijection between their; this is our
-approach in defining {\em QRE lenses}.
-
-More concretely, in order to define a language of QRE lenses, our approach is
-to (1) define a language of bijections (Section \ref{subsec:bijective-lenses}),
-and then (2) add quotients by allowing canonizers to be prepended or postpended
-to bijections and (3) allowing composition of such quotient lenses via the
+between two QREs via a bijection between their kernels; indeed, this approach
+is the one we adopt in defining {\em QRE lenses}.
+More concretely, to define a language of QRE lenses, we
+(1) define a language of bijections (Section \ref{subsec:bijective-lenses}),
+(2) add quotients by allowing canonizers to be prepended or postpended
+to bijections, and 
+(3) allow composition of such quotient lenses via the
 regular operators and functional composition
@@
 
-However, we also have a secondary objective, which is to support synthesis of
-lenses: When provided with QREs describing the source and target languages, we
-would like to be able to generate quotient lenses automatically. One way to do
-that is to generate canonizers $\canonize(Q_1)$ and $\canonize(Q_2)$ from QREs
-$Q_1$ and $Q_2$ and then to synthesize bijective lenses between the kernel
-languages for $Q_1$ and $Q_2$.
-
-Unfortunately, the composition of two such quotient lenses does not have have
+However, we also have a secondary objective, which is to support lens
+synthesis. When provided with QREs describing the source and target
+languages, we would like to be able to generate quotient lenses
+automatically. One way to achieve that goal is to generate canonizers
+$\canonize(Q_1)$ and $\canonize(Q_2)$ from QREs $Q_1$ and $Q_2$ and
+then to synthesize bijective lenses between the kernel languages for
+$Q_1$ and $Q_2$.
+Unfortunately, the composition of two such quotient lenses does not have 
 the form of a bijective lens with canonizers at the edges. Hence, an important
 technical question is whether we give up expressiveness if we restrict
-ourselves to this form. Thankfully, this turns out not to be an issue if
-the bijections that we use to define QRE lenses are derived using a particular
-set of combinators that we describe in the next section.
+ourselves to this form. 
+Thankfully, we can show that there is no loss of expressiveness if
+the bijections used to define QRE lenses are derived with a particular
+set of combinators.  We describe these combinators next.
 

### 2018-03-14 papers/QuotientLenses/main.tex
introduct Optician_Q as a shorthand for optician enhanced with QREs

--- 2018-03-13 papers/QuotientLenses/main.tex
@@
 %\newcommand{\QOpt}{Optician_Q}
-\newcommand{\QOpt}{QRE-enhanced Optician}
+\newcommand{\QOpt}{Optician\textsubscript{Q}}
 \newcommand{\OpticianRuntime}{\textbf{Optician}}
@@
 with syntax that expresses an equivalence relation on the language of $R$. This
-section formalizes the set of QRE combinators that we introduced informally
+section describes in further detail the set of QRE combinators that we described
 in \cref{sec:example}.
@@
 
-\subsection{Syntax and Semantics of QREs}
+\subsection{Syntax of Semantics of QREs}
 Formally, the language of Quotient Regular Expressions (QREs) is given by the
@@
 \begin{align*}
-Q := \; & \normalize{R_1}{R_2}{f} \bsep{} \id(R) \bsep \project{R}{s} \bsep \squash{Q_1}{Q_2}{f} \\
+Q := \; & \normalize{R}{R'}{f} \bsep{} \id(R) \bsep \project{R}{s} \bsep \squash{Q_1}{Q_2}{f} \\
  | \; & \perm{Q_1, \ldots, Q_n}{Q} \bsep Q_1 \semicolon Q_2 \bsep Q_1 \cdot Q_2 \bsep (Q_1
-\sep Q_2) \bsep Q^*
-\end{align*}
-where $Q$ ranges over QREs, 
-$R$ ranges over regular expressions, 
-$f$ ranges over functions between regular languages, and 
-$s$ ranges over character strings.
+\sep Q_2) \bsep Q^*,
+\end{align*}
+where $Q$ ranges over QREs, $R$ ranges over regular expressions, $f$ ranges over
+functions between regular languages, and $s$ ranges over character strings.
 
 Using the conventional notation that $\mathcal{L}(R)$ is the language accepted
-by the regular expression $R$, each QRE $Q$ yields four semantic
-objects:
+by the regular expression $R$, each QRE $Q$ enables us to express four semantic objects
 
 \begin{tabular}{rl}
-  $W(Q)$       & A regular expression, denoting the ``whole'' of $Q$\\
-  $\eqrel{Q}$  & An equivalence relation on $\mathcal{L}(W(Q))$\\
-  $K(Q)$       & A regular expression, denoting the ``kernel'' of $Q$, such that
-                  $\mathcal{L}(K(Q))$ forms a \\
-               & complete set of representatives for $\eqrel{Q}$ \\
+  $W(Q)$  & a regular expression (the ``whole'' of $Q$),\\
+  $\eqrel{Q}$  & an equivalence relation on $\mathcal{L}(W(Q))$,\\
+  $K(Q)$ & a regular expression (the ``kernel'' of $Q$) such that
+           $\mathcal{L}(K(Q))$ forms a complete \\
+             & set of representatives for $\eqrel{Q}$, and \\
   $\canonize(Q)$ & 
-  A ``canonizing'' function. Given any $w \in \mathcal{L}(W(Q))$,
-  $\canonize(Q)(w)$ \\ & is the unique $k$ in $\mathcal{L}(K(Q))$ such
+  a ``canonizing'' function that, given any $w \in \mathcal{L}(W(Q))$,
+  yields $\canonize(Q)(w)$ \\ & as the unique $k$ in $\mathcal{L}(K(Q))$ such
   that $k$ is equivalent to $w$ mod $\eqrel{Q}$
@@
 
-\noindent
-Intuitivley, $W(Q)$ is the regular expression representing the
-external format, while $K(Q)$ is the regular expression representing
-the internal format.  The equivalence relation $\eqrel{Q}$ groups
-together elements in the language of $W(Q)$ that contain the same
-essential information.  The function $\canonize(Q)$ picks the
-representative element from each of the resulting equivalence classes.
-
 The well-formedness constraints for QREs ensure that these four semantic
-objects fit together to form a coherent quotient $W(Q)/\eqrel{Q}$ whose
+components fit together to form a coherent quotient $W(Q)/\eqrel{Q}$ whose
 equivalence classes are determined by \canonize(Q).
@@
 The relationship among the semantic objects of a QRE can be understood in terms
-of the combinator \normalize{R_1}{R_2}{f}, which expresses each of these pieces
-explicitly.  Its whole language is just $R_1$, its kernel language is just
-$R_2$, and its canonizer is just $f$; 
-its equivalence relation $\equiv$ is
+of the \normalize{R}{S}{f} combinator, which exposes each of these pieces
+explicitly.  Its whole language is just $R$, its kernel language is just
+$S$, and $\canonize(\normalize{R}{S}{f}) = f$.  Its equivalence relation is
 determined by the fibres of $f$, so we have
-$s_1 \equiv s_2 \Leftrightarrow f(s_1) = f(s_2)$ for $s_1$ and $s_2$
-in $\mathcal{L}(R_1)$.
+$s_1 \equiv s_2 \Leftrightarrow f(s_1) = f(s_2)$.
 
@@
 \begin{prooftree}
-\AxiomC{$\mathcal{L}(R_2) \subseteq \mathcal{L}(R_1)$}
-\AxiomC{$f : \mathcal{L}(R_1) \longrightarrow \mathcal{L}(R_2)$}
+\AxiomC{$\mathcal{L}(S) \subseteq \mathcal{L}(R)$}
+\AxiomC{$f : \mathcal{L}(R) \longrightarrow \mathcal{L}(S)$}
 \AxiomC{$f$ is surjective}
@@
 \RightLabel{(Normalize)}
-\QuaternaryInfC{$\wf{\normalize{R_1}{R_2}{f}}$}
+\QuaternaryInfC{$\wf{\normalize{R}{R'}{f}}$}
 \end{prooftree}
@@
 surjective, idempotent function
-$f : \mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$.  
-However,
-verifying that a canonization
-function $f$ is surjective and idempotent is in general undecidable.
-Consequently, a programmer wishing to use \kw{normalize} must
-discharge strong proof obligations, which is cumbersome in
-practice.\footnote{In our implementation we allow a programmer to
-use \kw{normalize} at their own risk without checking these side
-conditions as an ``escape hatch'' for the case when other QRE
-combinators are insufficient.\bcp{Do we use this in any of the
-programs we've written?}}
-
-The remaining QRE combinators, which we discuss next, provide simpler,
-more compositional ways of building canonizers.  Nevertheless,
-the \kw{normalize} combinator provides a useful guide in the design of
-these combinators because it gives a sufficient condition for the
-well-formedness of any potential QREs.
-
-% particularly with respect to the strategy for synthesizing quotient
-% lenses (See Section~\ref{synth}), so we use it to guide the design of
-% our QRE remaining combinators, which we discuss next.
+$f : \mathcal{L}(W(Q)) \longrightarrow \mathcal{L}(K(Q))$.   However,
+although \kw{normalize} is powerful, verifying that a canonization function $f$
+is surjective and idempotent is in general undecidable.  This means that a
+programmer wishing to use \kw{normalize} must discharge strong proof
+obligations, which is cumbersome in practice.\footnote{In our implementation we allow a programmer to use \kw{normalize} at their
+own risk without checking these side conditions as an ``escape hatch'' for the
+case when other QRE combinators are insufficient.\bcp{Do we use this in any
+  of the programs we've written?}}
+
+The remaining QRE combinators provide simpler, more compositional ways of
+building canonizers.  Nevertheless, the \kw{normalize} combinator gives a
+sufficient condition for potential QREs to be well-behaved, particularly with
+respect to the strategy for synthesizing quotient lenses which we shall
+introduce in Section~\ref{synth}, so we use it to guide the design of our QRE
+combinators, as we see next.
 
@@
 Figure~\ref{fig:wk} gives the inductive definitions of the whole and
-kernel languages $W(Q)$ and $K(Q)$ for all of the QRE combinators.
+kernel languages $W(Q)$ and $K(Q)$ for all of the QRE forms.
 \begin{figure}[t]
@@
 \[
-\begin{array}{l@{\quad\quad}l@{\quad\quad}l}
+\begin{array}{l@{\quad}l@{\quad}l}
 
@@
 \project{R}{s} & R & s \\
-\squash{Q_1}{Q_2}{f} & W(Q_1) \sep W(Q_2) & K(Q_2) \\
-\normalize{R_1}{R_2}{f} & R_1 & R_2 \\
+\squash{Q}{Q_1}{f} & W(Q) \sep W(Q_1) & K(Q_1) \\
+\normalize{R}{R'}{f} & R & R' \\
 Q_1 \; ; \; Q_2 & W(Q_1) & K(Q_2) \\
-Q_1 \cdot\; Q_2 & W(Q_1) \cdot W(Q_2) & K(Q_1) \cdot K(Q_2) \\
-Q_1 \sep \; Q_2 & W(Q_1) \sep W(Q_2) & K(Q_1) \sep K(Q_2) \\
+Q_1 \cdot Q_2 & W(Q_1) \cdot W(Q_2) & K(Q_1) \cdot K(Q_2) \\
+Q_1 \sep Q_2 & W(Q_1) \sep W(Q_2) & K(Q_1) \sep K(Q_2) \\
 Q^* & W(Q)^* & K(Q)^* \\
@@
 W( \perm{Q_1, \ldots, Q_n}{Q} ) = &
-\bigcup \limits_{\sigma \in S_n} W(Q_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
-W(Q) \cdot W(Q_{\sigma(n)})\\
+\bigcup \limits_{\sigma \in S_n} W(c_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
+W(Q) \cdot W(c_{\sigma(n)})\\
 K( \perm{Q_1, \ldots, Q_n}{Q} ) = & K(Q_1) \cdot K(Q) \cdot \ldots \cdot K(Q)
@@
 \]
-\caption{Whole and Kernel Regular Expressions for QRE combinators. 
-In describing regular expressions, we use the notations $\sep$ and $\bigcup$ for
-alternation, the notation $\cdot$ for concatenation, and the notation
-$*$ for Kleene closure.  
-We use the notation $S_n$ to denote the set of all permutations of the
-numbers $1$ to $n$.
-}
+\caption{Whole and Kernel Regular Expressions}
 \label{fig:wk}
@@
 interesting definitions. If $Q = \squash{Q_1}{Q_2}{f}$, then the whole language
-of $Q$ is $W(Q_1) \sep W(Q_2)$ because the \kw{squash} combinator
-merges the whole language $W(Q_1)$ of $Q_1$ with the whole language $W(Q_2)$ of $Q_2$.
-The function $f : \mathcal{L}(W(Q_1)) \longrightarrow \mathcal{L}(W(Q_2))$,
+of $Q$ is $W(Q_1) \sep W(Q_2)$. This is because the \kw{squash} combinator
+takes the whole language $W(Q_1)$ of $Q_1$, the whole language $W(Q_2)$ of $Q_2$
+and a function $f : \mathcal{L}(W(Q_1)) \longrightarrow \mathcal{L}(W(Q_2))$,
 maps $\mathcal{L}(W(Q_1))$ into $\mathcal{L}(W(Q_2))$ using $f$ and then
@@
 For the $\perm{Q_1, \ldots, Q_n}{Q}$ combinator, the whole language is the
-union of languages of the form $W(Q_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
-W(Q) \cdot W(Q_{\sigma(n)})$ for any permutation $\sigma$ in $S_n$, where
-$S_n$ is the set of all permutations of the numbers $1$ to $n$. Intuitively,
+union of languages of the form $W(c_{\sigma(1)}) \cdot W(Q) \cdot \ldots \cdot
+W(Q) \cdot W(c_{\sigma(n)})$ for any permutation $\sigma$ in $S_n$, where
+$S_n$ is the set of all permutations of the numbers $1$ to $n$. This is because
 the \kw{permutation} combinator allows for the string to match any
-permutation of the $Q_i$'s while preserving the separator $Q$ in
+permutation the $Q_i$'s while also taking into account the separator $Q$ in
 between each of the $Q_i$'s. The kernel of the \kw{permutation} combinator
@@
 because the canonical permutation is the identity permuation, with each of the
-parts of the input that match $Q_i$ and $Q$ canonized into $K(Q_i)$ and $K(Q)$
+parts of the string that match $Q_i$ and $Q$ canonized into $K(Q_i)$ and $K(Q)$
 respectively.
@@
 \canonize(\id(R)) &=& id_{\mathcal{L}(R)} \\
-\canonize(\project{R}{s})(w) &=& s\\[1ex]
+\canonize(\project{R}{s})(w) &=& s\\
 \canonize(\squash{Q_1}{Q_2}{f})(w) &=&
@@
 \canonize(Q_2)(w) & \text{otherwise}
-\end{cases}\\[3ex]
-\canonize(\normalize{R_1}{R_2}{f}) &=& f\\
+\end{cases}\\
+\canonize(\normalize{R}{R'}{f}) &=& f\\
 \canonize(Q_1 \; ; \; Q_2) &=& \canonize(Q_2) \circ \canonize(Q_1)\\
-\canonize(Q_1 \cdot Q_2) &=& \canonize(Q_1) \cdot \canonize(Q_2)\\[1ex]
+\canonize(Q_1 \cdot Q_2) &=& \canonize(Q_1) \cdot \canonize(Q_2)\\
 \canonize(Q_1 \sep Q_2)(w) &=&
@@
 \canonize(Q_2)(w) & \text{if } w \in \mathcal{L}(W(Q_2))\\
-\end{cases}\\[3ex]
+\end{cases}\\
 \canonize(Q^*) &=& \canonize(Q)^* \\
@@
 &= (\canonize(Q_1)(r_1)) \cdot (\canonize(Q)(s_1)) \cdot \ldots \cdot
-(\canonize(Q)(s_{n-1}))\cdot
 (\canonize(Q_n)(r_n))
 \end{align*}
-\caption{QRE canonizers. \ksf{Solomon: please define/explain the . and *
-operators on functions.}}
+\caption{QRE canonizers}
 \label{fig:canonizers}
 \end{figure}
-The \kw{permutation} combinator gives rise to the most interesting
-definition:
+Again, the \kw{permutation} combinator gives the most interesting
+definition,
 \begin{align*}\canonize(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
@@
 = (\canonize(Q_1)(r_1)) \cdot (\canonize(Q)(s_1)) \cdot \ldots \cdot
-(\canonize(Q)(s_{n-1}))\cdot
 (\canonize(Q_n)(r_n))
 \end{align*}
-\noindent 
-which places the strings that match $Q_i$ and $Q$ according to the
-canonical permutation (i.e., the identity permutation) before applying
-the canonizers $\canonize(Q_i)$ and $\canonize(Q)$, respectively.
+\noindent which, as we have explained, is defined so that the strings that match
+$Q_i, Q$ are placed in the canonical permutation which is the identity
+permutation before $\canonize(Q_i)$ and $\canonize(Q)$ are applied
+respectively.
 
 Finally, Figure ~\ref{fig:relations} gives the inductive definition of the
-equivalence relation $\eqrel{Q}$, which is the set-theoretic semantics
-of a QRE $Q$ as an equivalence relation on the regular language $W(Q)$.
+equivalence relation $\eqrel{Q}$. This is the set-theoretic semantics of QREs
+as an equivalence relation on a regular language.
 
@@
 \begin{array}{l@{}l@{\ \ }l}
-w \; \equiv_{\normalize{R_1}{R_2}{f}} \; w' &\iff&
+w \; \equiv_{\normalize{R}{R'}{f}} \; w' &\iff&
 f(w)=f(w') \\
@@
 above as an extension to the Boomerang
-interpreter~\cite{boomerang,quotientlenses}. 
-The synthesis algorithm produces Boomerang lens values, so Boomerang
-treats synthesized lenses the same as hand-written ones.
-To evaluate the effectiveness of QREs, QRE lenses, and QRE lens
-synthesis, we conducted experiments to answer the following questions:
+interpreter~\cite{boomerang,quotientlenses}. We have extended the existing
+Optician tool to synthesize QRE lenses from QREs. We will use word
+\emph{\QOpt} to denote our extended version of Optician, and Optician to denote
+the pre-QRE version of Optician. The synthesis algorithm produces Boomerang lens
+values, so Boomerang treats synthesized lenses the same as hand-written ones. To
+evaluate the effectiveness of QREs, QRE lenses, and QRE lens synthesis, we
+conducted experiments to answer the following questions:
 %
@@
 from FlashFill~\cite{flashfill}, and benchmarks derived from
-Augeas~\cite{augeas2}.  In the pre-QRE version of Optician, 10 of
-these data formats had to be modified to work with the bijectivity
+Augeas~\cite{augeas2}.  In these 39 benchmarks, 10 of
+the data formats had to be modified to work with the bijectivity
 constraints that Optician required. For instance, when one
@@
 on each of the 10 non-bijective benchmark problems.
-Benchmarks are sorted in order of increasing complexity as measured by
-the number of AST nodes in the source and target format descriptions. 
 QRE Synthesis requires far fewer AST nodes than the other
@@
 Bijective Synthesis, the
-programmer uses bijective lens synthesis \`a la the original version
-of Optician.  She must write canonizers by hand, along with 
+programmer uses bijective lens synthesis \`a la Optician.
+She must write canonizers by hand, along with 
 regular expressions to describe the external
@@
 average of~1267 more AST nodes. These figures suggest that introducing QREs saves
-programmers significant effort compared to both pre-QRE Optician and basic
+programmers significant effort compared to both Optician and basic
 Boomerang.
@@
 \end{subfigure}
-\caption{Runtimes measurements. In (a), we run the
-  original version of Optician (\OpticianRuntime{}) and \QOpt{}
-  (\QREOptician{}) on the Optician benchmarks.  We find that there is only a negligible
+\caption{Runtimes measurements. In (a), we run Optician and \QOpt{}
+  on the Optician benchmarks.  We find that there is only a negligible
   performance overhead incurred by using QREs. 
@@
   previously edited to make them bijective, after removing those edits
-  and then extending the synthesis specification to include QREs.
-  (In other words, we restored them to their original state, added QREs,
+  (In other words, we restored them to their original state with QREs
   and then ran \QOpt{}). 
@@
 
-To assess the performance of QRE synthesis, we are interested in two
-different questions.  First, how does the performance of \QOpt{} 
-compare to the performance of the original
-version of Optician on benchmarks that do not require QREs?  The
-answer to this question tells us how much overhead we have introduced
-by adopting the more general mechanism.  
-Figure~\ref{fig:times}(a) shows that \QOpt{} was able to synthesize
-all of the Optician benchmarks at a speed competative with the old
-version. 
-There is a small amount
-of additional overhead introduced by QREs in calculating the $W$ and
-$K$ functions, resulting in a slight decrease in performance.
+To assess the performance of QRE synthesis, we are interested in two different
+questions. First, how does the performance of \QOpt{} compare to the performance
+of Optician on benchmarks that do not require QREs? The answer to this question
+tells us how much overhead we have introduced by adopting the more general
+mechanism. Figure~\ref{fig:times}(a) shows that \QOpt{} was able to synthesize
+all of the Optician benchmarks at a speed competative with the old version.
+There is a small amount of additional overhead introduced by QREs in calculating
+the $W$ and $K$ functions, resulting in a slight decrease in performance.
 
@@
 Boomerang~\cite{boomerang}. As we mentioned in
-Section\cref{subsec:well-formed-qres}, all our QRE combinators can be expressed
+Section~\ref{subsec:well-formed-qres}, all our QRE combinators can be expressed
 using just the {\em normalize} combinator, which is one of the canonizer

### 2018-03-14 papers/QuotientLenses/main.tex
Tweak spacing of visible space char

@@
 }
+\makeatletter
+     \let\lst@oldvisiblespace\lst@visiblespace
+     \def\lst@visiblespace{\,\lst@oldvisiblespace\,}
+\makeatother
 
@@
 %\newcommand{\QOpt}{Optician_Q}
-\newcommand{\QOpt}{Optician\textsubscript{Q}}
+\newcommand{\QOpt}{QRE-enhanced Optician}
 \newcommand{\OpticianRuntime}{\textbf{Optician}}
@@
 
-We offer three interconnected innovations to simplify the task of programming
-quotient lenses.  
-First, we introduce {\em quotient regular expressions} (QREs), to
-easily mark aspects of data formats that are
-inessential.  A QRE is an augmented regular expression that
-simultaneously defines a regular language and an equivalence
-relation on this language.  Variations within the same
-equivalence class are deemed inessential. 
-Next, we introduce {\em QRE lenses}, i.e., lenses mapping
-between QREs.
-Our key technical result is a proof that every QRE lens can be
-transformed into one of the form $c_1 ; \ell ;
-c_2^{-1}$, where $c_1$ and $c_2^{-1}$ are {\em canonizers}, $\ell$ is a bijective
-transformation, and $;$ is composition\iflastminute\bcp{$c_2^{-1}$ is ugly, but I think
-  we need the superscript for accuracy.  Perhaps calling them $c$ and $d$
-  instead of $c_1$ and $c_2$ would be better?}\fi.  Intuitively,
-the canonizers normalize data by picking a
-canonical representative of the relevant equivalence relations. 
-This theorem means we can push all normalization to the outer edges of a
+Building on recent work that showed how to synthesize bijective lenses from high
+level specifications, we offer three innovations to synthesize quotient lenses.
+First, we introduce {\em quotient regular expressions} (QREs), that easily mark
+aspects of data formats that are inessential and hence give a high-level
+specification of regular languages with an equivalence relation defined on
+them. Next, we introduce {\em QRE lenses}, i.e., lenses mapping between QREs.
+Our key technical result is a proof that every QRE lens can be transformed into
+one of the form $c_1 ; \ell ; c_2^{-1}$, where $c_1$ and $c_2^{-1}$ are {\em
+canonizers}, $\ell$ is a bijective transformation, and $;$ is
+composition\iflastminute\bcp{$c_2^{-1}$ is ugly, but I think we need the
+superscript for accuracy.  Perhaps calling them $c$ and $d$ instead of $c_1$
+and $c_2$ would be better?}\fi.  Intuitively, the canonizers normalize data by
+picking a canonical representative of the relevant equivalence relations. This
+theorem means we can push all normalization to the outer edges of a
 bidirectional transformation, leaving in the middle a pure bijective
-transformation.  
-Finally,
-we leverage this normalization theorem to {\em synthesize} QRE lenses from a pair
-of QREs and example input-output pairs, reusing earlier work on
-synthesizing plain bijective lenses.
-We have implemented QREs and QRE lens synthesis 
-in the bidirectional programming language Boomerang.
-We evaluate the effectiveness of our approach by using it to
-synthesize QRE lenses between various real-world data formats in
-the Optician benchmark suite.
+transformation. Finally, we leverage this normalization theorem to {\em
+synthesize} QRE lenses from a pair of QREs and example input-output pairs,
+reusing earlier work on synthesizing plain bijective lenses. We have
+implemented QREs and QRE lens synthesis in the bidirectional programming
+language Boomerang. We evaluate the effectiveness of our approach by using it
+to synthesize QRE lenses between various real-world data formats in the
+Optician benchmark suite.
 
@@
 
-\emph{Bijective lenses} are amongst the most restrictive. 
-A bijective lens $\ell : S \Leftrightarrow T$ specifies a bijection from
-a source set $S$ to a target set $T$:
-\begin{equation}\label{bijectivelenslaws}
-\ell.\get \; (\ell.\lput \; t) = t \text{, and } \ell.\lput \; (\ell.\get \; s)
-= s
+In ~\cite{optician}, we developed a tool called {\em Optician} to synthesize the
+class of {\em bijective lenses} that can be derived in the lens programming
+language {\em Boomerang}\cite{boomerang}. Each lens $\ell : S \Leftrightarrow T$
+that Optician can synthesized specifies a bijection from the language of a
+source regular expression $S$ to a target language of a target regular
+expression $T$: \begin{equation}\label{bijectivelenslaws} \ell.\get \;
+(\ell.\lput \; t) = t \text{, and } \ell.\lput \; (\ell.\get \; s) = s
 \end{equation}
-Since the \emph{get} and \emph{put} functions form a bijection, 
-no information is lost during translation from one data format to another. 
-While this very strong guarantee can sometimes be useful, it is more common for 
-some of the details in one or both of the data formats to be
-superfluous. For instance, the number of white space characters
-between data items, the capitalization of various strings, the
-sequence of fields in a record, or the order of items in a list are
-often irrelevant.  Also, many
-data formats introduce their own ad hoc conventions about formatting details
-that are taken to be inessential. In \bibtex{}, for example,
-an author can be specified either as 
-``\cd{lastname, firstname}'' or as ``\cd{firstname lastname}''.
-
-{\em Quotient lenses}~\cite{quotientlenses} loosen the strong restrictions of
-bijective string lenses by including equivalence
-relations on the source and target formats and requiring that the
-round-tripping lens laws 
-hold  modulo these equivalences, rather than ``on the nose.'' Quotient lenses 
-make it possible to synchronize many more data formats than is possible with
-plain bijective lenses\bcp{Do we justify this claim later?}. 
-%
-The bidirectional programming language {\em
-Boomerang}~\cite{boomerang} 
-implements a general class of quotient lenses.  Unfortunately, these
-lenses to have two main drawbacks.  \bcp{This wording may raise a question
-  in the reader's mind about whether this paper is really about quotient
-  lenses in general or whether it is specifically a proposal to tweak
-  Boomerang.  I.e., ``if I am not a Boomerang user, do I want to bother
-  reading this paper?''  My answer would be yes, because quotient
-  lenses and QREs are interesting in themselves.  But I don't think we've
-  made this point very convincingly in the way we've framed the paper.}
-
-The first shortcoming is that defining an equivalence
-relation in Boomerang requires explicitly providing a {\em canonizer}, 
-a pair of functions $(f,g)$ where $f$
-normalizes data from the
-external format by choosing canonical representations
-and $g$ provides an
-inverse by specifying the external form of each canonical
-representation. 
-The disadvantage of this approach is that writing
-a quotient lens in Boomerang requires separately defining four
-different data formats: two ``external'' formats which specify the data to
-be transformed by the lens, and two ``internal'' formats which specify the
-canonical representations for the source and target data\bcp{Plus writing
-  two canonizers!}.
-
-Our solution to this shortcoming is to introduce {\em Quotient Regular
-Expressions} (QREs), which enable programmers to simultaneously specify a
-regular expression and a set of canonical representatives for matching
-strings.  Given an QRE, we can automatically infer the corresponding
-canonization functions.  As an example, consider the following QRE for writing
-author names:
+Optician makes programming bijective lenses faster and easier, and though
+many useful lenses are not bijective in nature, they do have a bijective core
+surrounded by structured transformations, hence Optician can be used as a
+drop-in component for synthesizing more complex lenses.
+
+This paper applies this idea to synthesize {\em quotient lenses}
+~\cite{quotientlenses}. Quotient lenses are lenses in which the lens laws are
+loosened so that they hold modulo an equivalence relation on the source and
+target data respectively; in this paper we are concerned with {\em bijective
+quotient lenses} which are lenses for which Equation \ref{bijectivelenslaws}
+holds modulo equivalence relations $\sim_S$ and $\sim_T$ defined on the source
+and target data respectively:
+\begin{equation}\label{quotientlenslaws}
+\ell.\get \; (\ell.\lput \; t) \sim_T t \text{, and } \ell.\lput \; (\ell.\get
+\; s) \sim_S s
+\end{equation}
+Quotient lenses are useful in situations where a programmer wishes for the
+transformation defined by a lens to have the same behavior on data that differ
+only in inessential details. For instance, a programmer may wish to ``quotient
+away'' the number of white space characters between data items, the
+capitalization of various strings, the sequence of fields in a record, or the
+order of items in a list.
+
+Optician synthesizes bijective lenses from a pair $(S, T)$ of regular expression
+specifying the source and target types and a set of example input-output pairs
+that guide the synthesis algorithm. This presents a challenge for synthesizing
+quotient lenses since a specification of the source and target formats for a
+quotient lens needs to account for the equivalence relations defined on the
+respective formats. Our solution is to introduce {\em Quotient Regular
+Expressions} (QREs), which are regular expressions augmented with extra
+syntax that enables programmers to simultaneously specify a regular expression
+as well as an equivalence relation on the language of that regular expression.
+Further, given a QRE, we can automatically infer a set of canonical
+representatives the equivalence relation defined by the QRE as well as a
+canonizing function to the canonical set of representatives.
+
+For example, consider the following QRE for writing author names:
 %
@@
 
-The second drawback with writing quotient lenses in Boomerang is that even
-after specifying the four required data formats, the programmer must still
-define a lens for converting between the canonical representations.
-Experience shows that writing such lenses can be tedious (e.g., consider
-having to rearrange the order of data items by recursively using operators
-to \kw{swap} adjacent fields) and that satisfying the type checker can be
-difficult because the most obvious way to express a lens runs afoul of one
-or more ``ambiguity checks'' that the type checker performs in order to make
-sure that the lens laws hold~\cite{optician}.
-
-Our solution to this problem is to introduce {\em QRE lenses}, which
-are a new class of quotient lens that can be automatically
-synthesized, obviating the need to write lenses by hand.  A QRE lens
-is like a normal lens except it maps between formats described by
-QREs. We prove a normal form theorem (Theorem~\ref{normal form}) that says that
-every composition of QRE lenses can be rewritten 
-to be the composition of a source canonizer, a bijective lens, and a
-target canonizer. This normalization property means we can synthesize
-QRE lenses by extending an existing algorithm for synthesizing
-bijective lenses~\cite{optician}.
-
-Given this framework, generating a QRE lens requires only a pair of QREs to describe the
-source and target formats and a (possibly empty) suite of examples
+Our second main contribution in this paper is to introduce {\em QRE lenses},
+which are bijective quotient lenses that map between QREs via bijective lenses
+that map between the respective canonical formats defined by the source and
+target QREs. This idea is simple and natural, but begs the following question:
+do we give up expressiveness if we restrict ourselves to this form? Our main
+technical contribution to this question asserts that we do not; more
+specifically, we will prove a normal form theorem (Theorem~\ref{normal form})
+that says that every QRE lens can be rewritten to be the composition of a
+source canonizer, a bijective lens, and a target canonizer. This normalization
+property will enable us to (1) synthesize QRE lenses by extending the synthesis
+algorithm used by Optician, and (2) prove that if there is a QRE lens that
+satisfies the input specification, then this extended algorithm will return
+such a lens.
+
+Given this framework, generating a QRE lens requires only a pair of QREs to
+describe the source and target formats and a (possibly empty) suite of examples
 demonstrating the mapping.  For example, the following code
@@
 comma-separated form describd by the QRE \codefont{comma\_name} and the
-space-separated form described by the QRE \codefont{space\_name}.  The
-lenses generated by our synthesis algorithm are guaranteed to translate data
-back and forth between source and target QREs and to satisfy the provided
-examples.  Our synthesized quotient lenses have a relatively restricted
-form, where the sequential composition operator is not used and all
-canonization occurs at the very edges (which greatly speeds up synthesis),
-but the normalization theorem means this restriction does not compromise
-expressiveness.
+space-separated form described by the QRE \codefont{space\_name}.
 

### 2018-03-14 papers/QuotientLenses/main.tex
Redid Intro

@@
 }
-\makeatletter
-     \let\lst@oldvisiblespace\lst@visiblespace
-     \def\lst@visiblespace{\,\lst@oldvisiblespace\,}
-\makeatother
 
@@
 
-Building on recent work that showed how to synthesize bijective lenses from high
-level specifications, we offer three innovations to synthesize quotient lenses.
-First, we introduce {\em quotient regular expressions} (QREs), that easily mark
-aspects of data formats that are inessential and hence give a high-level
-specification of regular languages with an equivalence relation defined on
-them. Next, we introduce {\em QRE lenses}, i.e., lenses mapping between QREs.
-Our key technical result is a proof that every QRE lens can be transformed into
-one of the form $c_1 ; \ell ; c_2^{-1}$, where $c_1$ and $c_2^{-1}$ are {\em
-canonizers}, $\ell$ is a bijective transformation, and $;$ is
-composition\iflastminute\bcp{$c_2^{-1}$ is ugly, but I think we need the
-superscript for accuracy.  Perhaps calling them $c$ and $d$ instead of $c_1$
-and $c_2$ would be better?}\fi.  Intuitively, the canonizers normalize data by
-picking a canonical representative of the relevant equivalence relations. This
-theorem means we can push all normalization to the outer edges of a
+We offer three interconnected innovations to simplify the task of programming
+quotient lenses.  
+First, we introduce {\em quotient regular expressions} (QREs), to
+easily mark aspects of data formats that are
+inessential.  A QRE is an augmented regular expression that
+simultaneously defines a regular language and an equivalence
+relation on this language.  Variations within the same
+equivalence class are deemed inessential. 
+Next, we introduce {\em QRE lenses}, i.e., lenses mapping
+between QREs.
+Our key technical result is a proof that every QRE lens can be
+transformed into one of the form $c_1 ; \ell ;
+c_2^{-1}$, where $c_1$ and $c_2^{-1}$ are {\em canonizers}, $\ell$ is a bijective
+transformation, and $;$ is composition\iflastminute\bcp{$c_2^{-1}$ is ugly, but I think
+  we need the superscript for accuracy.  Perhaps calling them $c$ and $d$
+  instead of $c_1$ and $c_2$ would be better?}\fi.  Intuitively,
+the canonizers normalize data by picking a
+canonical representative of the relevant equivalence relations. 
+This theorem means we can push all normalization to the outer edges of a
 bidirectional transformation, leaving in the middle a pure bijective
-transformation. Finally, we leverage this normalization theorem to {\em
-synthesize} QRE lenses from a pair of QREs and example input-output pairs,
-reusing earlier work on synthesizing plain bijective lenses. We have
-implemented QREs and QRE lens synthesis in the bidirectional programming
-language Boomerang. We evaluate the effectiveness of our approach by using it
-to synthesize QRE lenses between various real-world data formats in the
-Optician benchmark suite.
+transformation.  
+Finally,
+we leverage this normalization theorem to {\em synthesize} QRE lenses from a pair
+of QREs and example input-output pairs, reusing earlier work on
+synthesizing plain bijective lenses.
+We have implemented QREs and QRE lens synthesis 
+in the bidirectional programming language Boomerang.
+We evaluate the effectiveness of our approach by using it to
+synthesize QRE lenses between various real-world data formats in
+the Optician benchmark suite.
 

### 2018-03-14 papers/QuotientLenses/main.tex
Edited abstract
New figure

--- 2018-03-14 papers/QuotientLenses/main.tex
@@
 \centering
-\includegraphics[width=0.7\textwidth]{canonizers-outside}
-\caption{QRE Lens with ``Canonizers at the Edges''}
+\includegraphics[width=0.7\textwidth]{QREs}
+\caption{QRE lens from source $P$ to target $Q$.  $P$
+  and $Q$ are QREs, each consisting of a ``whole'' set ($W(P)$ and $W(Q)$)
+  and a ``kernel'' part ($K(P)$ and $K(Q)$)
+  Between the two kernels is a plain bijective lens.  The {\emph get}
+  function of the whole lens takes an argument from $W(P)$, applies the
+  canonizer for $P$ to obtain a canonical representative in $K(P)$, then
+  applies the {\em get} of the plain lens, yielding an element of $K(Q)$,
+  which is a subset of $W(Q)$.  The {\em put} function does the reverse,
+  mapping from $W(Q)$ to $K(P)$ (hence $W(P)$.\BCP{Both the figure and the
+    caption are new---what do you think??  Too detailed?  Unreadable?}}
 \label{fig:attheedges}

### 2018-03-14 papers/QuotientLenses/main.tex
intro edits

@@
 
-In ~\cite{optician}, we developed a tool called {\em Optician} to synthesize the
+In recent work~\cite{optician}, we developed a tool called {\em Optician} to synthesize the
 class of {\em bijective lenses} that can be derived in the lens programming
 language {\em Boomerang}\cite{boomerang}. Each lens $\ell : S \Leftrightarrow T$
-that Optician can synthesized specifies a bijection from the language of a
+that Optician can synthesize specifies a bijection from the language of a
 source regular expression $S$ to a target language of a target regular
@@
 many useful lenses are not bijective in nature, they do have a bijective core
-surrounded by structured transformations, hence Optician can be used as a
-drop-in component for synthesizing more complex lenses.
-
-This paper applies this idea to synthesize {\em quotient lenses}
-~\cite{quotientlenses}. Quotient lenses are lenses in which the lens laws are
+surrounded by structured transformations. Hence, Optician can be used as a
+component in a system that synthesizes more complex lenses.
+
+This paper applies this idea to the problem of
+synthesizing {\em quotient lenses}~\cite{quotientlenses}.
+Quotient lenses are lenses in which the lens laws are
 loosened so that they hold modulo an equivalence relation on the source and
@@
 
-Optician synthesizes bijective lenses from a pair $(S, T)$ of regular expression
+Optician synthesizes bijective lenses from a pair $(S, T)$ of regular
+expressions
 specifying the source and target types and a set of example input-output pairs
@@
 as well as an equivalence relation on the language of that regular expression.
-Further, given a QRE, we can automatically infer a set of canonical
-representatives the equivalence relation defined by the QRE as well as a
-canonizing function to the canonical set of representatives.
+Further, given a QRE, we can automatically infer a \emph{canonizer}---a
+function that converts string in the language of regular expression
+to a canonical form.
 
@@
   lenses to the problem of synthesizing bijective string lenses, which was
-  previously studied in~\cite{optician} (Section~\ref{synth}).
+  previously studied in past work~\cite{optician} (Section~\ref{synth}).
   \item We extend Boomerang with QREs and QRE lens synthesis and demonstrate

### 2018-03-14 papers/QuotientLenses/main.tex
edit
Bunch more polishing and comments

@@
 \newif\ifdraft\drafttrue  % set true to show comments
-\newif\iflastminute\lastminutefalse  % for things to check at the end
+\newif\iflastminute\lastminutetrue  % for things to check at the end
 
@@
 
-Recent work described the {\em Optician}
+Recent work described the {\em Optician}\bcp{Why is this sometimes
+  italicized and sometimes not?  (My preference would be not---we have too
+  much italics.)}
 algorithm/tool~\cite{optician}, which can synthesize bijective lenses
@@
 
-Another QRE primitive is the sequential composition combinator, written ``;''.
+Another QRE primitive is the functional composition combinator, written ``;''.
 For an example of its use, suppose we have already defined a QRE,
@@
 documents, and chooses a specific ordering of XML elements as canonical. We can
-use the sequential composition combinator to combine these two QREs into
+use the functional composition combinator to combine these two QREs into
 \lstinline{canonized_whitespace ; canonized_order}, a QRE that accepts all XML
@@
 
-Figure~\ref{fig:example-qre} gives the complete QRE specification for the
-\bibtex{}.
+Figure~\ref{fig:example-qre} gives a QRE specification for the simple
+\bibtex{} records we consider here.
 \begin{figure}
@@
 \includegraphics[width=0.7\textwidth]{QREs}
-\caption{QRE lens from source $P$ to target $Q$.  $P$
-  and $Q$ are QREs, each consisting of a ``whole'' set ($W(P)$ and $W(Q)$)
-  and a ``kernel'' part ($K(P)$ and $K(Q)$)
-  Between the two kernels is a plain bijective lens.  The {\emph get}
-  function of the whole lens takes an argument from $W(P)$, applies the
-  canonizer for $P$ to obtain a canonical representative in $K(P)$, then
-  applies the {\em get} of the plain lens, yielding an element of $K(Q)$,
-  which is a subset of $W(Q)$.  The {\em put} function does the reverse,
-  mapping from $W(Q)$ to $K(P)$ (hence $W(P)$.\BCP{Both the figure and the
-    caption are new---what do you think??  Too detailed?  Unreadable?}}
+\caption{QRE lens from source $P$ to target $Q$.  $P$ and $Q$ are QREs, each
+  consisting of a ``whole'' set ($W(P)$ and $W(Q)$) and a ``kernel'' part
+  ($K(P)$ and $K(Q)$) Between the two kernels is a plain bijective lens.
+  The {\emph get} function of the whole lens takes an argument from $W(P)$,
+  applies the canonizer for $P$ to obtain a canonical representative in
+  $K(P)$, then applies the {\em get} of the plain lens, yielding an element
+  of $K(Q)$, which is a subset of $W(Q)$.  The {\em put} function does the
+  reverse, mapping from $W(Q)$ to $K(P)$ (hence $W(P)$.  This lens is in
+  normal form, with canonizers (specified by QREs) at the outer edges and an
+  ordinary bijective lens in the center. \BCP{Both the figure and the
+    caption are new---what do you think??}}
 \label{fig:attheedges}
@@
 The \kw{permutation} combinator gives rise to the most interesting
-definition:
+definition:\bcp{Why is this right-justified?}
 \begin{align*}\canonize(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
@@
 \]
-\caption{QRE Equivalence Relations}
+\caption{QRE Equivalence Relations\bcp{The inter-letter spacing of ``True''
+    is wrong, and anybody who has worked with tex for a while will be
+    bothered by this rookie mistake.  A general rule in tex is that english
+    words in math mode {\em must always} be enclosed in {\tt
+      $\backslash$mathit}.}}
 \label{fig:relations}
@@
 ourselves to this form. 
-Thankfully, we can show that there is no loss of expressiveness if
+Fortunately, we can show that there is no loss of expressiveness if
 the bijections used to define QRE lenses are derived with a particular
@@
 
-This derivation rule essentially says that the composition $q_1 \; ; \; q_2$ is
+This rule essentially says that the composition $q_1 \; ; \; q_2$ is
 well defined if and only if the intermediary QREs $Q_3$ and $Q_4$ define the
 same equivalence relation on regular languages that are equivalent modulo the
-star-semiring axioms. The condition $\mathcal{L}(W(Q_2)) = \mathcal{L}(W(Q_3))$
+star-semiring axioms\bcp{No: the regular expressions are equivalent modulo
+  these axioms; the regular {\em languages} are
+  identical}. (See~\cite[\S 4]{quotientlenses} for an example of what goes
+wrong if this premise is dropped.\bcp{Didn't we used to have a discussion
+  about how implementing this rule is hard because it involves checking
+  language equivalence so we follow Boomerang's rule?  I think this is
+  important to mention because this was a significant restriction on the
+  power of composition in Boomerang.  It doesn't apply here (right?) because
+we intend to synthesize these lenses, not write them by hand, so we will
+never actually need to typecheck a functional composition expression.}) The
+condition $\mathcal{L}(W(Q_2)) = \mathcal{L}(W(Q_3))$ 
 says that the intermediary language is the same on both sides, while the
@@
 \noindent
-We have also proven the correctness of this algorithm.
 \begin{theorem}\label{thm:alg-correct}
@@
 \noindent
-This theorem follows from the correctness of the Optician algorithm and
-Theorem~\ref{normal form}.
+(This follows from the correctness of the Optician algorithm and
+Theorem~\ref{normal form}.)
 
@@
 
-Because of this extra elbow-room, Boomerang has two primitive {\em duplication}
-quotient lenses, the first of which can be derived using the following inference
-rule,
+Because of this extra elbow-room, Boomerang is able to offer two primitive
+{\em duplication} quotient lenses, the first of which can be derived using
+the following inference rule,
 \begin{prooftree}
@@
 \end{align*}
-wiht the symmetric $\kw{dup}_2$ combinator discarding the first copy instead of
+with the symmetric $\kw{dup}_2$ combinator discarding the first copy instead of
 the second in the \lput/\create{} direction.
@@
 
-Another language which supports quotenting is BiFluX~\cite{pacheco2014biflux}, a
+A newer language which supports quotenting is BiFluX~\cite{pacheco2014biflux}, a
 bidirectional functional update language for XML. BiFluX is inspired by
@@
 data sequences described by the two types can be matched. This process can
-result in intermediate types that are unambiguous, and this ambiguity can cause
+result in intermediate types that are unambiguous\BCP{???}, and this
+ambiguity can cause 
 unintended updates to be made the source, especially when information is
@@
 
-{\em Formlenses}~\cite{rajkumar2014lenses} are a variant of lenses that perform
-transformations up to equivalence. Formlenses are a bidirectional generalization
-of {\em formlets}~\cite{cooper2008essence}. Formlets are a high-level
-abstraction for building Web forms, and can be represented in Haskell as
-functions that take a source of names---concretely, an \kw{Int}---as an
-argument and produce a triple consisting of an HTML document, a {\em collect}
-function, and a modified namesource:
+{\em Formlenses}~\cite{rajkumar2014lenses} are a variant of lenses that
+perform transformations up to equivalence. Formlenses are a bidirectional
+generalization of {\em formlets}~\cite{cooper2008essence}. \bcp{The rest of
+  this paragraph is hard to understand.  Can we be a little more
+  explicit?}Formlets are a high-level abstraction for building Web forms,
+and can be represented in Haskell as functions that take a source of
+names\bcp{It was puzzling to me, for a while, what an Int had to do with a
+  source of names: spell it out}---concretely, an \kw{Int}---as an argument
+and produce a triple consisting of an HTML document, a {\em collect}
+function, and a modified namesource:\bcp{typesetting of arrows is funny}
 \begin{center}
@@
 
-They \kw{Formless} type is given by,
+The \kw{Formless} type is given by,\bcp{I notice that several displays now
+  have commas before them.  If you imagine them typeset inline instead of
+  displayed, would this be good grammar?}
 \begin{center}
@@
 Under this assumption, Rajkumar et. al. define what it means for a formlens to
-be well behaved by defining a version of the classical GETPUT and PUTGET laws
+be well behaved by defining a version of the classical \bcp{typeset in
+  small caps}GETPUT and PUTGET laws
 for formlenses. These laws hold modulo an equivalence relation on environments,
@@
 
-Finally, in Cunha~\cite{cunha2010relational} uses relational algebra to
+Finally, Cunha~\cite{cunha2010relational} uses relational algebra to
 encompass many of the existing approaches to bidirectional programming. For
@@
 
-Now we turn to related work in synthesis. Much of the research in synthesis
+Next we turn to related work in program synthesis. Much of the research in
+synthesis 
 assumes that the synthesizer is provided with a collection of examples.
-Optician, as well as Optician enhanced with QREs for QRE lens synthesis, differ
-in that they require that the programmer supplies both examples {\em and}
+Optician and Optician$_Q$  differ
+in that they require the programmer to supply both examples {\em and}
 format descriptions in the form of regular expressions or QREs, though we are

### 2018-03-14 papers/QuotientLenses/main.tex
regularize notation for get and put

@@
 \lstset{ language=Caml, basicstyle=\linespread{0.8}\upshape\sffamily,
-keywordstyle=\upshape\sffamily\color{dkblue}, keepspaces=true,
+keywordstyle=\upshape\sffamily\color{dkred}, keepspaces=true,
 framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=true,
 commentstyle=\itshape\rmfamily,
-emph={synth,project,perm,squash,normalize,using,ins,del,lens,let},
+emph={synth,project,perm,squash,normalize,using,ins,del,lens,let,get,put,rquot,lquot},
 emphstyle=\upshape\sffamily\color{dkred}, 
@@
 \newcommand{\codefont}[1]{\ensuremath{\mathsf{#1}}}
-\newcommand{\kw}[1]{\codefont{#1}}
+\newcommand{\kw}[1]{\textcolor{dkred}{\ensuremath{\mathsf{#1}}}}
 \newcommand{\project}[2]{\ensuremath{\kw{project} \; #1 \mapsto #2}}
@@
 Our key technical result is a proof that every QRE lens can be transformed into
-one of the form $c_1 ; \ell ; c_2^{-1}$, where $c_1$ and $c_2^{-1}$ are {\em
+one of the form $c ; \ell ; d^{-1}$, where $c$ and $d^{-1}$ are {\em
 canonizers}, $\ell$ is a bijective transformation, and $;$ is
@@
 written in one of several domain-specific bidirectional languages that 
-guarantee various \emph{lens laws} constraining how the \emph{put} and
-\emph{get} functions must fit together.
-
-In recent work~\cite{optician}, we developed a tool called {\em Optician} to synthesize the
+guarantee various \emph{lens laws} constraining how the \kw{put} and
+\kw{get} functions must fit together.
+
+Recent work~\cite{optician}\saz{removed ``we'' here to maintain the double-blind
+  review standard.} developed a tool called Optician to synthesize the
 class of {\em bijective lenses} that can be derived in the lens programming
-language {\em Boomerang}\cite{boomerang}. Each lens $\ell : S \Leftrightarrow T$
+language {\em Boomerang}~\cite{boomerang}. Each lens $\ell : S \Leftrightarrow T$
 that Optician can synthesize specifies a bijection from the language of a
@@
 quotient lenses} which are lenses for which Equation \ref{bijectivelenslaws}
-holds modulo equivalence relations $\sim_S$ and $\sim_T$ defined on the source
+holds modulo equivalence relations $\equiv_S$ and $\equiv_T$ defined on the source
 and target data respectively:
 \begin{equation}\label{quotientlenslaws}
-\ell.\get \; (\ell.\lput \; t) \sim_T t \text{, and } \ell.\lput \; (\ell.\get
-\; s) \sim_S s
+\ell.\get \; (\ell.\lput \; t) \equiv_T t \text{, and } \ell.\lput \; (\ell.\get
+\; s) \equiv_S s
 \end{equation}
@@
 that guide the synthesis algorithm. This presents a challenge for synthesizing
-quotient lenses since a specification of the source and target formats for a
-quotient lens needs to account for the equivalence relations defined on the
+quotient lenses since a specification of its source and target formats
+needs to account for the equivalence relations defined on the
 respective formats. Our solution is to introduce {\em Quotient Regular
@@
 syntax that enables programmers to simultaneously specify a regular expression
-as well as an equivalence relation on the language of that regular expression.
+and an equivalence relation on the language of that regular expression.
 Further, given a QRE, we can automatically infer a \emph{canonizer}---a
-function that converts string in the language of regular expression
+function that converts strings in the language of the regular expression
 to a canonical form.
@@
 do we give up expressiveness if we restrict ourselves to this form? Our main
-technical contribution to this question asserts that we do not; more
-specifically, we will prove a normal form theorem (Theorem~\ref{normal form})
+technical contribution to this question asserts that we do not---more
+specifically, we prove a normal form theorem (Theorem~\ref{normal form})
 that says that every QRE lens can be rewritten to be the composition of a
@@
 \begin{lstlisting}
-let l = synth comma_name $\Leftrightarrow$ space_name using {("Lovelace, Ada", "Ada Lovelace")}
+let $\ell$ = synth comma_name $\Leftrightarrow$ space_name using {("Lovelace, Ada", "Ada Lovelace")}
 \end{lstlisting}
-binds \codefont{l} to a synthesized QRE lens mapping between names in the
-comma-separated form describd by the QRE \codefont{comma\_name} and the
+binds $\ell$ to a synthesized QRE lens mapping between names in the
+comma-separated form described by the QRE \codefont{comma\_name} and the
 space-separated form described by the QRE \codefont{space\_name}.
@@
 \label{sec:background}
-\iflastminute
-\ksf{We are inconsistent in how we typeset get/put.  Sometimes we use
-codefont, sometimes we use italics of various flavors, and sometimes
-we don't do anything distinctive.  It would be good to be consistent!}
-\fi
 Before describing QREs and QRE lenses, we briefly review bijective
@@
 
-Recent work described the {\em Optician}\bcp{Why is this sometimes
-  italicized and sometimes not?  (My preference would be not---we have too
-  much italics.)}
+Recent work described the Optician
 algorithm/tool~\cite{optician}, which can synthesize bijective lenses
@@
 \begin{lstlisting}
-synth S <=> T using exs
+synth S $\Leftrightarrow$ T using exs
 \end{lstlisting}
 \noindent
-{\em Optician} will synthesize a bijective lens between 
+Optician will synthesize a bijective lens between 
 source and target formats described by regular expressions
-\cd{S} and \cd{T}, respectively, and a set of input-output example pairs \cd{exs} specifying how
+\cd{S} and \cd{T}, respectively, and constrained by a set of input--output example pairs \cd{exs} specifying how
 the synthesized lens should behave on those examples.
@@
 
-{\em Optician} greatly simplifies the task of programming bijective
+Optician greatly simplifies the task of programming bijective
 string lenses, but not all bidirectional transformations are
@@
 Consequently, the following two \bibtex{} citations represent the same
-logical object even though they differ in nonessential details.
+logical object even though they differ in nonessential details.\saz{should we
+  place such examples side-by-side?}
 \begin{verbatim}
@@
 However, we can synthesize such functions automatically---here, \codefont{c\_to\_s} is the
-\codefont{get} direction of a lens that can be synthesized using the 
+get direction of a lens that can be synthesized using the 
 \kw{synth} combinator:
@@
 \begin{lstlisting}
-let l = synth comma_name $\Leftrightarrow$ space_name using {("Lovelace, Ada", "Ada Lovelace")}
-let c_to_s = get l
+let $\ell$ = synth comma_name $\Leftrightarrow$ space_name using {("Lovelace, Ada", "Ada Lovelace")}
+let c_to_s = $\ell$.get 
 \end{lstlisting}
-\noindent
+\noindent \saz{I changed \codefont{get} $\ell$ to $\ell$.\kw{get} to be
+  consistent with the earlier text.}
 The first line above synthesizes a lens between \codefont{comma\_name} and
 \codefont{space\_name} using the listed example transformation as a guide.  The
-second line extracts the \codefont{get} direction transformation from the lens,
+second line extracts the get direction transformation from the lens,
 which is what we need for \kw{squash}.
@@
 (* synthesize a lens that maps comma representation to space representation *)
-let l = synth comma_name $\Leftrightarrow$ space_name using {("Lovelace, Ada", "Ada Lovelace")}
-let c_to_s = get l
+let $\ell$ = synth comma_name $\Leftrightarrow$ space_name using {("Lovelace, Ada", "Ada Lovelace")}
+let c_to_s = $\ell$.get
 
@@
 \caption{QRE specification of \bibtex{} records.\ksf{In what sense is this a QRE
-spec vs a QRE lens spec?}}
+spec vs a QRE lens spec?}}\saz{This defines only one endpoint of the eventual
+\bibtex{}--EndNote lens.  Maybe we should say ``definition'' instead of ``specification''?}
 \label{fig:example-qre}
@@
 pair of regular expressions and a set of example input-output pairs
-({\em Optician}), 
+(Optician), 
 and we have a way of defining regular expressions with equivalence
@@
   ($K(P)$ and $K(Q)$) Between the two kernels is a plain bijective lens.
-  The {\emph get} function of the whole lens takes an argument from $W(P)$,
+  The \kw{get} function of the whole lens takes an argument from $W(P)$,
   applies the canonizer for $P$ to obtain a canonical representative in
-  $K(P)$, then applies the {\em get} of the plain lens, yielding an element
-  of $K(Q)$, which is a subset of $W(Q)$.  The {\em put} function does the
+  $K(P)$, then applies the \kw{get} of the plain lens, yielding an element
+  of $K(Q)$, which is a subset of $W(Q)$.  The \kw{put} function does the
   reverse, mapping from $W(Q)$ to $K(P)$ (hence $W(P)$.  This lens is in
@@
   ordinary bijective lens in the center. \BCP{Both the figure and the
-    caption are new---what do you think??}}
+    caption are new---what do you think??}}\saz{I think that this figure is a
+  big improvement.  One tiny possible tweak: color the \kw{get} and \kw{put}
+  keywords in the same dark red as in the text.}
 \label{fig:attheedges}
@@
 source QRE $Q_1$ to compute a canonical representative for the data modulo the
-equivalence relation defined by $Q_1$ and then applies the $\get$ function of a
+equivalence relation defined by $Q_1$ and then applies the \kw{get} function of a
 bijective string lens $\ell_1$ to this representative. In the {\em put} direction, $q_1$
@@
 binding them to the variables \cd{bibtex} and \cd{endnote} respectively, then
-all the code in Figure~\ref{fig:example-lens} may be replaced by a single call
+all of the code in Figure~\ref{fig:example-lens} may be replaced by a single call
 to the synthesis prodedure:
@@
 The \kw{permutation} combinator gives rise to the most interesting
-definition:\bcp{Why is this right-justified?}
-\begin{align*}\canonize(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
+definition:
+\[
+\begin{array}{cl}
+ & \canonize(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
 \cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \\
-= (\canonize(Q_1)(r_1)) \cdot (\canonize(Q)(s_1)) \cdot \ldots \cdot
+= & (\canonize(Q_1)(r_1)) \cdot (\canonize(Q)(s_1)) \cdot \ldots \cdot
 (\canonize(Q)(s_{n-1}))\cdot
 (\canonize(Q_n)(r_n))
-\end{align*}
+\end{array}
+\]
 \noindent 
@@
 w \; \equiv_{\id(R)} \; w' &\iff& w = w' \\
-w \; \equiv_{\project{R}{s}} \; w' &\iff& True \\ % \text{for all }w, w'\\
+w \; \equiv_{\project{R}{s}} \; w' &\iff& \mathit{True} \\ % \text{for all }w, w'\\
 w \; \equiv_{\squash{Q_1}{Q_2}{f}} \; w' &\iff& f(w) \equiv_{Q_2} w'
@@
 \]
-\caption{QRE Equivalence Relations\bcp{The inter-letter spacing of ``True''
-    is wrong, and anybody who has worked with tex for a while will be
-    bothered by this rookie mistake.  A general rule in tex is that english
-    words in math mode {\em must always} be enclosed in {\tt
-      $\backslash$mathit}.}}
+\caption{QRE Equivalence Relations}
 \label{fig:relations}
@@
 \RightLabel{(Id)}
-\UnaryInfC{$\wf{\mathit{id}(R)}$}
+\UnaryInfC{$\wf{\kw{id}(R)}$}
 \DisplayProof
@@
 The most complicated inference rule is the (Perm) rule for the
-\kw{permutation} combinator. The second hypothesis for the
+permutation combinator. The second hypothesis for the
 Perm rule says that for any two different permutations $\sigma$
@@
 W(Q_{\theta(n)})$ must be disjoint. This restriction is important
-because any string that matches the 
-\kw{permutation} combinator matches the language $W(Q_{\theta(1)}) \cdot
+because an input string could match any
+of the regular expressions $W(Q_{\theta(1)}) \cdot
 W(Q) \cdot \ldots \cdot W(Q_{\theta(n)})$ for some permutation $\theta$, so we
-must require that these regular expressions be disjoint.
+must require that all of them be disjoint.
 %
@@
 restriction arises because the
-underlying lens that maps to or from a \kw{permutation} QRE operates on
+underlying lens that maps to or from a \kw{perm} QRE operates on
 the language $K(Q_1) \cdot K(Q) \cdot \ldots \cdot K(Q) \cdot K(Q_n)$, 
-the kernel of the \kw{permutation} QRE. The underlying
+the kernel of the \kw{perm} QRE. The underlying
 lens requires that the source and target regular expressions be
@@
 that is specified by QREs. Given regular expressions $R, S$ and equivalence
-relations $\sim_R$ and $\sim_S$ defined on $\mathcal{L}(R)$ and $\mathcal{L}(S)$,
-we define a {\em bijective quotient lens} $q : R /{\sim_R} \Leftrightarrow
-S/{\sim_S}$ to be a pair of functions $q.\get :
+relations $\equiv_R$ and $\equiv_S$ defined on $\mathcal{L}(R)$ and $\mathcal{L}(S)$,
+we define a bijective quotient lens $q : R /{\equiv_R} \Leftrightarrow
+S/{\equiv_S}$ to be a pair of functions $q.\get :
 \mathcal{L}(R) \longrightarrow \mathcal{L}(S)$ and $q.\lput : \mathcal{L}(S)
 \longrightarrow \mathcal{L}(R)$ such that for all $r \in \mathcal{L}(R)$ and $s
-\in \mathcal{L}(S)$, we have $q.\lput(q_1.\get(r)) \sim_R r$ and
-$q.\get(q_1.\lput(s)) \sim_R s$. Moreover, if $r \sim_R r'$, then $q.\get(r) =
-q.\get(r')$, and if $s \sim_S s'$, then $q.\lput(s) = q.\lput(s')$.
-
-In words, a bijective quotient lens $q$ from $R$ to $S$ modulo $\sim_R$
-and $\sim_S$ is a pair of functions $q.\get$ and $q.\lput$ such that
-$q.\get$ respects $\sim_R$ and $q.\lput$ respects $\sim_S$, and such that
-the {\em get} and {\em put} functions lifted to $\mathcal{L}(R)/{\sim_R}$ and
-$\mathcal{L}(S)/{\sim_S}$ are mutual inverses.
+\in \mathcal{L}(S)$, we have $q.\lput(q_1.\get(r)) \equiv_R r$ and
+$q.\get(q_1.\lput(s)) \equiv_R s$. Moreover, if $r \equiv_R r'$, then $q.\get(r) =
+q.\get(r')$, and if $s \equiv_S s'$, then $q.\lput(s) = q.\lput(s')$.
+
+In words, a bijective quotient lens $q$ from $R$ to $S$ modulo $\equiv_R$
+and $\equiv_S$ is a pair of functions $q.\get$ and $q.\lput$ such that
+$q.\get$ respects $\equiv_R$ and $q.\lput$ respects $\equiv_S$, and such that
+the {\em get} and {\em put} functions lifted to $\mathcal{L}(R)/{\equiv_R}$ and
+$\mathcal{L}(S)/{\equiv_S}$ are mutual inverses.
 
 These laws are similar to the laws for bijective lenses
-\cref{bijectivelenslaws}, except that the  equality restrictions in the
+\cref{bijectivelenslaws}\saz{again, cref uses a silly format}, except that the  equality restrictions in the
 bijective lens laws are loosened to allow for equivalence relations. Also, the
-condition that $r \sim_R r'$ implies $q.\get(r) = q.\get(r')$ ensures that the
+condition that $r \equiv_R r'$ implies $q.\get(r) = q.\get(r')$ ensures that the
 {\em get} function induced on the equivalence classes if $\equiv_c$ is
@@
 
-$$\ell := \id(R) \sep \const(s_1, s_2) \sep  \swap(\ell_1, \ell_2)
+$$\ell := \id(R) \sep \const(s, t) \sep  \swap(\ell_1, \ell_2)
 \sep \ell_1 \cdot \ell_2 \; |  \; (\ell_1 \sep \ell_2) \sep \ell^* \; | \;
@@
 data in the forward direction, and $t$ with $s$ in the target data in the
-backward direction, while $\mathit{id}(R)$ applies the identity function to the
+backward direction, while $\kw{id}(R)$ applies the identity function to the
 source and target  in $\mathcal{L}(R)$ in both directions. The composition lens
@@
 but with $s, s_1$ and $s_2$ substituted for $t, t_1$ and $t_2$. The
-$\mathit{\swap} \; (\ell_1, \ell_2)$ lens operates like $\ell_1 \cdot \ell_2$,
+$\swap(\ell_1, \ell_2)$ lens operates like $\ell_1 \cdot \ell_2$,
 except that it swaps $t_1$ and $t_2$ before concatenating the two for a final
 result of $t_2 \cdot t_1$ in the forward direction. In the backward direction,
-$\mathit{\swap}(\ell_1, \ell_2)$ first undoes the \swap, then behaves like
+$\swap(\ell_1, \ell_2)$ first undoes the \swap, then behaves like
 the concatenation lens. The $\ell_1 \; | \; \ell_2$ lens chooses to apply
@@
 $\ell^*$ lens splits the string $s$ into strings $s_1, \ldots, s_n$, applies
-$\ell_1$ to each $s_i$ to get $t_i$, and then concatenates each of the $t_i$'s
+$\ell_1$ to each $s_i$ to obtain $t_i$, and then concatenates each of the $t_i$'s
 for a final result of $t_1 \cdot \ldots \cdot t_n$ in the forward direction.
 The iteration lens $\ell^*$ operates similarly in the backward direction, but
-with $s, s_i$ substituted for $t, t_i$.
+with $s$ and $s_i$ substituted for $t$ and $t_i$.
 
@@
 \AxiomC{$R$ is strongly unambiguous}
-\UnaryInfC{$\mathit{id} \; (R): R \Leftrightarrow R$}
+\UnaryInfC{$\kw{id}(R): R \Leftrightarrow R$}
 \DisplayProof
@@
 Formally, given sets $C$ and $B$ and an equivalence relation on $B$, Foster et
-al defined a {\em canonizer} $q$ from $C$ to $B/{\sim_B}$ to be a pair of
+al defined a {\em canonizer} $q$ from $C$ to $B/{\equiv_B}$ to be a pair of
 functions $q.\kw{canonize} : C \longrightarrow B$ and $q.\kw{choose} : B
 \longrightarrow C$ such that for every $b \in B$,
-$$q.\kw{canonize} \; (q.\kw{choose} \; b) \sim_B \; b$$
+$$q.\kw{canonize} \; (q.\kw{choose} \; b) \equiv_B \; b$$
 This definition gives allows much more latitude for defining canonizers than
-QREs. For example, if $\sim_B = \kw{Tot}(B)$, the equivalence relation that
+QREs. For example, if $\equiv_B = \kw{Tot}(B)$, the equivalence relation that
 relates every element in $B$ to every other element in $B$, then every function
@@
 \begin{prooftree}
-\AxiomC{$\ell : C/{\sim_C} \Leftrightarrow A_1/{\sim_{A_1}}$}
+\AxiomC{$\ell : C/{\equiv_C} \Leftrightarrow A_1/{\equiv_{A_1}}$}
 \AxiomC{$f : C \longrightarrow A_2$}
 \AxiomC{$A_1 \cdot^! A_2$}
-\AxiomC{$\sim_A = \sim_{A_1} \cdot \kw{Tot}(A_2)$}
-\QuaternaryInfC{$\kw{dup}_1 \ell \; f \; : \; C/{\sim_C} \Longrightarrow A_1
-\cdot A_2/{\sim_A}$}
+\AxiomC{$\equiv_A = \equiv_{A_1} \cdot \kw{Tot}(A_2)$}
+\QuaternaryInfC{$\kw{dup}_1 \ell \; f \; : \; C/{\equiv_C} \Longrightarrow A_1
+\cdot A_2/{\equiv_A}$}
 \end{prooftree}
@@
 and produce a triple consisting of an HTML document, a {\em collect}
-function, and a modified namesource:\bcp{typesetting of arrows is funny}
+function, and a modified namesource:
 \begin{center}
 \begin{lstlisting}
-type Formlet a = Int -> (Html, Env -> a, Int)
+type Formlet a = Int $\to$ (Html, Env $\to$ a, Int)
 \end{lstlisting}
@@
 abstraction, but they only address one half of the problem---they make it easy
-to construct a function that {\em produces} a value of type \kw{a}, but they do
+to construct a function that {\em produces} a value of type \codefont{a}, but they do
 not provide a way to describe a function that {\em consumes} a value of type
-\kw{a} and embeds it in a form. A different way to think about this problem is
+\codefont{a} and embeds it in a form. A different way to think about this problem is
 to observe that forms are often used to present an {\em updateable view} of the
@@
 \begin{lstlisting}
-type Formlens a = Maybe a -> [Int] -> (Html,Env -> Maybe a,[Int])
+type Formlens a = Maybe a $\to$ [Int] $\to$ (Html,Env $\to$ Maybe a,[Int])
 \end{lstlisting}
@@
 with the differences between this definition and the standard definition for
-formlets being an extra \kw{Maybe a} parameter that provides an optional initial
+formlets being an extra \codefont{Maybe\; a} parameter that provides an optional initial
 value of the form, the namesource being changed to a list of integers, and the
@@
 Under this assumption, Rajkumar et. al. define what it means for a formlens to
-be well behaved by defining a version of the classical \bcp{typeset in
-  small caps}GETPUT and PUTGET laws
+be well behaved by defining a version of the classical \textsc{GetPut} and \textsc{PutGet} laws
 for formlenses. These laws hold modulo an equivalence relation on environments,
@@
 not affect the value stored in the information encapsulated in the type $a$.
-Consequently, the GETPUT and PUTGET laws defined for formlenses are in spirit
-analogous to the GETPUT and PUTGET laws for quotient lenses, since they hold up
+Consequently, the \textsc{GetPut} and \textsc{PutGet} laws defined for formlenses are in spirit
+analogous to the \textsc{GetPut} and \textsc{PutGet} laws for quotient lenses, since they hold up
 to renaming of environment variables.
@@
 Well-behaved formlenses satisfy properties that are analogous to the classical
-GETPUT and PUTGET law. If {\em extract} is a function from \kw{Html} to \kw{Env} that
+\textsc{GetPut} and \textsc{PutGet} law. If {\em extract} is a function from \kw{Html} to \kw{Env} that
 extracts an environment containing the names and values of all the form fields
@@
 encompass many of the existing approaches to bidirectional programming. For
-example, using $\cdot$ to express relational composition, the GETPUT law states
-that $\get \cdot \lput \subseteq \sim_V \cdot \pi_1$, where $\pi_1$ is the
+example, using $\cdot$\saz{Why not use $\circ$ or ``;''} to express relational composition, the \textsc{GetPut} law states
+that $\get \cdot \lput \subseteq \equiv_V \cdot \pi_1$, where $\pi_1$ is the
 projection onto the first factor. In addition to its generality, this

### 2018-03-14 papers/QuotientLenses/main.tex
project to collapse

@@
 commentstyle=\itshape\rmfamily,
-emph={synth,project,perm,squash,normalize,using,ins,del,lens,let,get,put,rquot,lquot},
+emph={synth,collapse,perm,squash,normalize,using,ins,del,lens,let,get,put,rquot,lquot},
 emphstyle=\upshape\sffamily\color{dkred}, 
@@
 \newcommand{\kw}[1]{\textcolor{dkred}{\ensuremath{\mathsf{#1}}}}
-\newcommand{\project}[2]{\ensuremath{\kw{project} \; #1 \mapsto #2}}
+\newcommand{\collapse}[2]{\ensuremath{\kw{collapse} \; #1 \mapsto #2}}
 \newcommand{\squash}[3]{\ensuremath{\kw{squash} \; #1 \rightarrow #2\; \kw{using} \; #3}}
@@
 \begin{lstlisting}
-let wsp_sp = project wsp $\to$ " "
+let wsp_sp = collapse wsp $\to$ " "
 let comma_name = first_name . "," . wsp_sp . last_name
@@
 character, its canonical form. 
-We use the QRE \kw{project} primitive to define this
-whitespace-normalizing QRE. \bcp{Why is it called
-``project''?} \ksf{select seems like it might be a better name as we
-are choosing one representative element from a set of possibilities.}\bcp{Or
-``collapse''?}
+We use the QRE \kw{collapse} primitive to define this
+whitespace-normalizing QRE.
 
 \begin{lstlisting}
-let wsp_sp = project wsp $\to$ " "
+let wsp_sp = collapse wsp $\to$ " "
 \end{lstlisting}
@@
 \begin{lstlisting}
-let bib_fields = perm (label, bib_author, bib_title) with (project ("," . wsp) $\to$ ",\n")
+let bib_fields = perm (label, bib_author, bib_title) with (collapse ("," . wsp) $\to$ ",\n")
 \end{lstlisting}
@@
 character) for all whitespace strings $s$. Then the 
-project QRE \lstinline{wsp_sp} defined above can
+collapse QRE \lstinline{wsp_sp} defined above can
 be expressed using the normalize combinator as
@@
 let wsp = [#\,\textvisiblespace\,#\n\t\r]+
-let wsp_sp = project wsp $\to$ " "
+let wsp_sp = collapse wsp $\to$ " "
 let last_name = [A-Z][a-z]
@@
 (* allow any permutation of fields interspersed with arbitrary whitespace *)
-let bib_fields = perm (label, bib_author, bib_title) with (project ("," . wsp) $\to$ ",\n")
+let bib_fields = perm (label, bib_author, bib_title) with (collapse ("," . wsp) $\to$ ",\n")
 let bibtex = "@book{" . bib_fields . "}"
@@
 \begin{align*}
-Q := \; & \normalize{R_1}{R_2}{f} \bsep{} \id(R) \bsep \project{R}{s} \bsep \squash{Q_1}{Q_2}{f} \\
+Q := \; & \normalize{R_1}{R_2}{f} \bsep{} \id(R) \bsep \collapse{R}{s} \bsep \squash{Q_1}{Q_2}{f} \\
  | \; & \perm{Q_1, \ldots, Q_n}{Q} \bsep Q_1 \semicolon Q_2 \bsep Q_1 \cdot Q_2 \bsep (Q_1
@@
 R & R & R \\
-\project{R}{s} & R & s \\
+\collapse{R}{s} & R & s \\
 \squash{Q_1}{Q_2}{f} & W(Q_1) \sep W(Q_2) & K(Q_2) \\
@@
 \canonize(\id(R)) &=& id_{\mathcal{L}(R)} \\
-\canonize(\project{R}{s})(w) &=& s\\[1ex]
+\canonize(\collapse{R}{s})(w) &=& s\\[1ex]
 \canonize(\squash{Q_1}{Q_2}{f})(w) &=&
@@
 w \; \equiv_{\id(R)} \; w' &\iff& w = w' \\
-w \; \equiv_{\project{R}{s}} \; w' &\iff& \mathit{True} \\ % \text{for all }w, w'\\
+w \; \equiv_{\collapse{R}{s}} \; w' &\iff& \mathit{True} \\ % \text{for all }w, w'\\
 w \; \equiv_{\squash{Q_1}{Q_2}{f}} \; w' &\iff& f(w) \equiv_{Q_2} w'
@@
 because it ensures that the canonizing function of a QRE is well-defined.
-For example, consider the QRE \lstinline{"a"* . (project "a"* $\to$ "a")}. The
+For example, consider the QRE \lstinline{"a"* . (collapse "a"* $\to$ "a")}. The
 behaviour of this QRE is not well-defined since the string
@@
 \AxiomC{$s \in \mathcal{L}(R)$}
-\RightLabel{(Project)}
-\UnaryInfC{$\wf{\project{R}{s}}$}
+\RightLabel{(Collapse)}
+\UnaryInfC{$\wf{\collapse{R}{s}}$}
 \DisplayProof

### 2018-03-14 papers/QuotientLenses/main.tex
pass over section 3

@@
 Consequently, the following two \bibtex{} citations represent the same
-logical object even though they differ in nonessential details.\saz{should we
-  place such examples side-by-side?}
-\begin{verbatim}
+logical object even though they differ in nonessential details.
+
+\begin{center}
+  \begin{minipage}{2.2in}
+    \centering
+\begin{lstlisting}
 @Book {Lovelace,
@@
 }
-
+\end{lstlisting}
+  \end{minipage}
+  \begin{minipage}{2in}
+    \centering
+\begin{lstlisting}
 @Book{     Lovelace,
@@
 Author = "Lovelace, Ada",     }
-\end{verbatim}
+\end{lstlisting}
+  \end{minipage}
+\end{center}
 
@@
 \label{subsec:qre-expressions}
-\ksf{Are these QRE combinators or QRE lens combinators?  
-We are currently mushy about this distinction.  
-The section needs a pass to clarify this point.
-The headers of this section and the next imply that this section is
-about QREs and the next section is about QRE lenses, but this section
-is where we introduce the QRE lens combinators, right?  If so, we need
-to do something to make the structure clearer.}\afm{I believe this are QRE
-combinators, so we should be good.}
-
 In this subsection, we develop a QRE specification of \bibtex{}
@@
 \iflastminute
-\bcp{Can we display blanks as a distinct ``visible blank'' symbol
-(throughout)?  That will make examples easier to read.}\bcp{I tried turning
-on the showstringspaces option at the top, but that doesn't do everything we
-want--it only works inside quotes.  But the latex ``textvisiblespace'' command
-produces a single one, and the lstlisting ``escapechar'' option will allow
-us to use it inside listings.  But we should save this for the final
-tweaking stage.}\bcp{Done, I think.}
 % Here is an example: if you set escapechar=\%, you could use it inside the
@@
 \end{lstlisting}
-\noindent \saz{I changed \codefont{get} $\ell$ to $\ell$.\kw{get} to be
-  consistent with the earlier text.}
+\noindent
 The first line above synthesizes a lens between \codefont{comma\_name} and
@@
 
-Figure~\ref{fig:example-qre} gives a QRE specification for the simple
+Figure~\ref{fig:example-qre} gives a QRE definition for the simple
 \bibtex{} records we consider here.
@@
 \end{lstlisting}
-\caption{QRE specification of \bibtex{} records.\ksf{In what sense is this a QRE
-spec vs a QRE lens spec?}}\saz{This defines only one endpoint of the eventual
-\bibtex{}--EndNote lens.  Maybe we should say ``definition'' instead of ``specification''?}
+\caption{QRE definition of \bibtex{} records.}
 \label{fig:example-qre}

### 2018-03-14 papers/QuotientLenses/main.tex
update cref options

@@
   centernot, tikz, array, mathtools, bussproofs, stmaryrd, enumitem,
-  stackengine, subcaption, cleveref}
+  stackengine, subcaption}
 \captionsetup{compatibility=false}
@@
 \usepackage[noend]{algpseudocode}
+\usepackage[capitalize]{cleveref}
 

### 2018-03-14 papers/QuotientLenses/main.tex
minor nit fixes in section 4

@@
 in \cref{sec:example}.
-\iflastminute 
-\saz{Why does the ``cref'' macro not capitalize Section?}
-\bcp{Dunno.  Because SIGPLAN does not want it to??  But maybe we should be
-  using autoref instead anyway...}
-\fi
 
@@
   $\canonize(Q)(w)$ \\ & is the unique $k$ in $\mathcal{L}(K(Q))$ such
-  that $k$ is equivalent to $w$ mod $\eqrel{Q}$
+  that $k \eqrel{Q} w$.
 \end{tabular}
@@
 combinators are insufficient.\bcp{Do we use this in any of the
-programs we've written?}}
+programs we've written?}\afm{Nope.}}
 
@@
 \canonize(Q_1 \; ; \; Q_2) &=& \canonize(Q_2) \circ \canonize(Q_1)\\
-\canonize(Q_1 \cdot Q_2) &=& \canonize(Q_1) \cdot \canonize(Q_2)\\[1ex]
+\canonize(Q_1 \cdot Q_2)(w_1 \cdot w_2) &=& \canonize(Q_1)(w_1) \cdot \canonize(Q_2)(w_2)\\[1ex]
 \canonize(Q_1 \sep Q_2)(w) &=&
@@
 \end{cases}\\[3ex]
-\canonize(Q^*) &=& \canonize(Q)^* \\
+\canonize(Q^*)(\epsilon) &=& \epsilon \\
+\canonize(Q^*)(w_1 \cdot w_2) &=& \canonize(Q)(w_1)\cdot\canonize(Q^*)(w_2) \\
 \end{array}
@@
 \caption{QRE canonizers. \ksf{Solomon: please define/explain the . and *
-operators on functions.}}
+operators on functions.}\afm{I tried rewriting with explicit arguments,
+thoughts?}}
 \label{fig:canonizers}
@@
 because it ensures that the canonizing function of a QRE is well-defined.
-For example, consider the QRE \lstinline{"a"* . (collapse "a"* $\to$ "a")}. The
+For example, consider the QRE \lstinline{"a"* . (collapse "a"$^*$ $\to$ "a")}. The
 behaviour of this QRE is not well-defined since the string

### 2018-03-14 papers/QuotientLenses/main.tex
tweaks to figures 4 and 5

@@
 Q & W(Q) & K(Q) \\ \hline
-R & R & R \\
+\id(R) & R & R \\
 \collapse{R}{s} & R & s \\
@@
 \]
-\caption{Whole and Kernel Regular Expressions for QRE combinators. 
+\caption{Whole and Kernel regular expressions for QRE combinators. 
 In describing regular expressions, we use the notations $\sep$ and $\bigcup$ for
@@
 \begin{array}{l@\quad l @\quad l}
-\canonize(\id(R)) &=& id_{\mathcal{L}(R)} \\
+\canonize(\id(R))(w) &=& w \\
 \canonize(\collapse{R}{s})(w) &=& s\\[1ex]
@@
 operators on functions.}\afm{I tried rewriting with explicit arguments,
-thoughts?}}
+thoughts?}\saz{If so, do it consistently everywhere. I fixed \id{} but the
+others need to be fixed too.}}
 \label{fig:canonizers}

### 2018-03-14 papers/QuotientLenses/main.tex
updated Figure 5

@@
 \end{cases}\\[3ex]
-\canonize(\normalize{R_1}{R_2}{f}) &=& f\\
-\canonize(Q_1 \; ; \; Q_2) &=& \canonize(Q_2) \circ \canonize(Q_1)\\
+\canonize(\normalize{R_1}{R_2}{f})(w) &=& f(w)\\
+\canonize(Q_1 \; ; \; Q_2)(w) &=& \canonize(Q_2)(\canonize(Q_1)(w))\\
 \canonize(Q_1 \cdot Q_2)(w_1 \cdot w_2) &=& \canonize(Q_1)(w_1) \cdot \canonize(Q_2)(w_2)\\[1ex]
@@
 \canonize(Q^*)(\epsilon) &=& \epsilon \\
-\canonize(Q^*)(w_1 \cdot w_2) &=& \canonize(Q)(w_1)\cdot\canonize(Q^*)(w_2) \\
+\canonize(Q^*)(w_1 \cdot w_2) &=& \canonize(Q)(w_1)\cdot\canonize(Q^*)(w_2) \\[2ex]
+\multicolumn{3}{l}{\canonize(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
+\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)})(w_{\sigma(1)} \cdot s_1
+  \cdot \ldots \cdot s_{n-1} \cdot w_{\sigma(n)})}\\[.5ex]
+\multicolumn{3}{l}{\ =\ \ (\canonize(Q_1)(r_1))(w_1) \cdot (\canonize(Q)(s_1)) \cdot \ldots \cdot
+(\canonize(Q)(s_{n-1}))\cdot(\canonize(Q_n)(r_n))(w_n)}
 \end{array}
@@
 \end{center}
-\begin{align*}
-&\canonize(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
-\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)})\\
-&= (\canonize(Q_1)(r_1)) \cdot (\canonize(Q)(s_1)) \cdot \ldots \cdot
-(\canonize(Q)(s_{n-1}))\cdot
-(\canonize(Q_n)(r_n))
-\end{align*}
-\caption{QRE canonizers. \ksf{Solomon: please define/explain the . and *
-operators on functions.}\afm{I tried rewriting with explicit arguments,
-thoughts?}\saz{If so, do it consistently everywhere. I fixed \id{} but the
-others need to be fixed too.}}
+\caption{QRE canonizers.  Here we assume that the input $w$ to the canonizers
+  has been partitioned in the unique way guaranteed to exist by the lens unambiguity conditions.}
 \label{fig:canonizers}

### 2018-03-14 papers/QuotientLenses/main.tex
more tweaks to fig 5
fixed text of 4.3
Edited related work section

@@
 \newcommand{\squash}[3]{\ensuremath{\kw{squash} \; #1 \rightarrow #2\; \kw{using} \; #3}}
-\newcommand{\perm}[2]{\ensuremath{\kw{perm}\; (#1)\; \kw{with}\; #2}}
+\newcommand{\perm}[2]{\ensuremath{\kw{perm}(#1)\; \kw{with}\; #2}}
 \newcommand{\normalize}[3]{\ensuremath{\kw{normalize}(#1, #2, #3)}}
@@
   \cdot \ldots \cdot s_{n-1} \cdot w_{\sigma(n)})}\\[.5ex]
-\multicolumn{3}{l}{\ =\ \ (\canonize(Q_1)(r_1))(w_1) \cdot (\canonize(Q)(s_1)) \cdot \ldots \cdot
-(\canonize(Q)(s_{n-1}))\cdot(\canonize(Q_n)(r_n))(w_n)}
+\multicolumn{3}{l}{\ =\ \ \canonize(Q_1)(r_1)(w_1) \cdot \canonize(Q)(s_1) \cdot \ldots \cdot
+\canonize(Q)(s_{n-1})\cdot\canonize(Q_n)(r_n)(w_n)}
 \end{array}
@@
 \[
-\begin{array}{cl}
- & \canonize(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
-\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)}) \\
-= & (\canonize(Q_1)(r_1)) \cdot (\canonize(Q)(s_1)) \cdot \ldots \cdot
-(\canonize(Q)(s_{n-1}))\cdot
-(\canonize(Q_n)(r_n))
+\begin{array}{l}
+\canonize(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
+\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)})(w_{\sigma(1)} \cdot s_1
+  \cdot \ldots \cdot s_{n-1} \cdot w_{\sigma(n)})\\[.5ex]
+\ =\ \ \canonize(Q_1)(r_1)(w_1) \cdot \canonize(Q)(s_1) \cdot \ldots \cdot
+\canonize(Q)(s_{n-1})\cdot\canonize(Q_n)(r_n)(w_n)
 \end{array}
@@
 This definition gives allows much more latitude for defining canonizers than
-QREs. For example, if $\equiv_B = \kw{Tot}(B)$, the equivalence relation that
-relates every element in $B$ to every other element in $B$, then every function
-from $C$ to $B$ is a canonizer.
+QREs. For example, if $\equiv_B$ is equal to $\kw{Tot}(B)$, the equivalence
+relation that relates every element in $B$ to every other element in $B$, then
+every function from $C$ to $B$ is a canonizer.
 
@@
 data sequences described by the two types can be matched. This process can
-result in intermediate types that are unambiguous\BCP{???}, and this
-ambiguity can cause 
-unintended updates to be made the source, especially when information is
-discared in this process. The typechecking phase of a BiFluX program therefore
-includes a {\em type normalization} phase that tries to normalize a source types
-using automata reduction techniques, and derive a lens between ambiguous and
-unambiguous types. On the view side, this phase only tries to normalize view
-types into isomorphic types.
+result in intermediate types that are ambiguous\BCP{???}\sam{Before I had this
+as unambiguous. Was this the issue? Or does this need more explaining even if
+the word unambiguous was the issue}, and this ambiguity can cause unintended
+updates to be made the source, especially when information is discarded in this
+process.
+
+\iffalse
+For example, imagine that a book database has source type \\
+$S$ = \lstinline{books[book[title[string], author[string]+]*]}
+with view type $V$ = \lstinline{title[string]}. When the source type is
+transformed to the view type for matching, only titles are kept, the element
+labels are destroyed and all authors are replaced for the empty sequence, thus
+leaving the intermediate type $S'$ = \lstinline{(title[string],(),()*)*}. In
+BiFluX, $V$ is considered a subtype of $S'$ because the flat values of $V$ are
+also flat values of $S'$. Since the intermediate type is $S'$ ambiguous, a view
+sequence \lstinline{[title["mybook"]]} would be upcast into an intermediate
+value \lstinline{[(title["mybook"],((), [ ]))]} that would in turn lead to an
+updated source \lstinline{books[[book[(title["mybook"],(aut, [ ]))]]]}, which is
+clearly unsatisfactory as it would discard the entire book database except the
+first author \lstinline{aut} of the first book.
+\fi
+The typechecking phase of a BiFluX program therefore includes a {\em type
+normalization} phase that tries to normalize a source types using automata
+reduction techniques, and derive a lens between ambiguous and unambiguous
+types. On the view side, this phase only tries to normalize view types into
+isomorphic types. Pacheco et. al. do not claim that their normalization
+procedures are complete in the sense that they can disambiguate any ambiguous
+type, but they do show that when these procedures succed the normalized types
+are unambiguous.
 
@@
 perform transformations up to equivalence. Formlenses are a bidirectional
-generalization of {\em formlets}~\cite{cooper2008essence}. \bcp{The rest of
-  this paragraph is hard to understand.  Can we be a little more
-  explicit?}Formlets are a high-level abstraction for building Web forms,
-and can be represented in Haskell as functions that take a source of
-names\bcp{It was puzzling to me, for a while, what an Int had to do with a
-  source of names: spell it out}---concretely, an \kw{Int}---as an argument
-and produce a triple consisting of an HTML document, a {\em collect}
-function, and a modified namesource:
+generalization of {\em formlets}~\cite{cooper2008essence}, which are a
+high-level abstraction for building Web forms. Formlets encapsulate several
+low-level details including selecting field names for elements and parsing data
+from client responses. In their work, Rajkumar et. al. represent formlets as
+functions that take a source of names---concretely, an integer $n$ that should
+be used as the next name---as an argument and produce a triple consisting of an
+HTML document, a {\em collect} function, and a modified namesource:
 \begin{center}
@@
 The names of any generated form fields are drawn from the namesource, and the
-{\em collect} function looks up precisely these names.
-Rajkumar et. al. observe that formlets are a useful
-abstraction, but they only address one half of the problem---they make it easy
-to construct a function that {\em produces} a value of type \codefont{a}, but they do
-not provide a way to describe a function that {\em consumes} a value of type
-\codefont{a} and embeds it in a form. A different way to think about this problem is
-to observe that forms are often used to present an {\em updateable view} of the
-data source. Therefore, Rajkumar et. al. leverage the power of bidirectional
+{\em collect} function looks up precisely these names. For example the
+\kw{inputInt} combinator below builds a form that accepts an integer,
+\begin{center}
+\begin{lstlisting}
+inputInt :: Formlet Int
+inputInt i = let n = show i in
+             ($\kw{inputTag}$ {name = n, value = ""},$\lambda e \longrightarrow \kw{read}$ ($\kw{fromJust}$ ($\kw{lookup}$ n e)),i+1)
+\end{lstlisting}
+\end{center}
+where \kw{inputTag} constructs an input element, \kw{lookup}
+retrieves a value from an association list \kw{fromJust} extracts the
+encapsulated value from a \kw{Maybe} value, and \kw{read} parses an \kw{Int}
+from a \kw{String}.
+
+Rajkumar et. al. observe that formlets are a useful abstraction, but they only
+address one half of the problem---they make it easy to construct a function
+that {\em produces} a value of type \codefont{a}, but they do not provide a way
+to describe a function that {\em consumes} a value of type \codefont{a} and
+embeds it in a form. A different way to think about this problem is to observe
+that forms are often used to present an {\em updateable view} of the data
+source. Therefore, Rajkumar et. al. leverage the power of bidirectional
 transformations to define formlets that behave like updatable views, yielding
@@
 
-The \kw{Formless} type is given by,\bcp{I notice that several displays now
-  have commas before them.  If you imagine them typeset inline instead of
-  displayed, would this be good grammar?}
+The \kw{Formless} type is given by
 \begin{center}
@@
 with the differences between this definition and the standard definition for
-formlets being an extra \codefont{Maybe\; a} parameter that provides an optional initial
-value of the form, the namesource being changed to a list of integers, and the
-{\em collect} function being allowed to return an optional value.
+formlets being an extra \codefont{Maybe\; a} parameter that provides an
+optional initial value of the form, the namesource being changed to a list of
+integers, and the {\em collect} function being allowed to return an optional
+value.
 
@@
 argument, and the {\em collect} function should only look up corresponding
-names.
-Under this assumption, Rajkumar et. al. define what it means for a formlens to
-be well behaved by defining a version of the classical \textsc{GetPut} and \textsc{PutGet} laws
-for formlenses. These laws hold modulo an equivalence relation on environments,
-since a formlens may be invoked with two separate namespaces, but this should
-not affect the value stored in the information encapsulated in the type $a$.
-Consequently, the \textsc{GetPut} and \textsc{PutGet} laws defined for formlenses are in spirit
-analogous to the \textsc{GetPut} and \textsc{PutGet} laws for quotient lenses, since they hold up
-to renaming of environment variables.
+names. Under this assumption, Rajkumar et. al. define what it means for a
+formlens to be well behaved by defining a version of the classical
+\textsc{GetPut} and \textsc{PutGet} laws for formlenses. These laws hold modulo
+an equivalence relation on environments, since a formlens may be invoked with
+two separate namespaces, but this should not affect the value stored in the
+information encapsulated in the type $a$. Consequently, the \textsc{GetPut} and
+\textsc{PutGet} laws defined for formlenses are in spirit analogous to the
+\textsc{GetPut} and \textsc{PutGet} laws for quotient lenses, since they hold
+up to renaming of environment variables.
 
@@
 Well-behaved formlenses satisfy properties that are analogous to the classical
-\textsc{GetPut} and \textsc{PutGet} law. If {\em extract} is a function from \kw{Html} to \kw{Env} that
-extracts an environment containing the names and values of all the form fields
-from an HTML document, then a formlens $f$ satisfies the Acceptability law if
-for any $v$ of type $a$, $n$ of type \kw{Int} and environment $e$ of type
-\kw{Env} if $f$ applied to \kw{Just v} and $n$ yields an HTML file $h$ and a
-{\em collect} function $c$ where \kw{extract h} is contained in $e$, then
-collect $e$
+\textsc{GetPut} and \textsc{PutGet} law. If {\em extract} is a function from
+\kw{Html} to \kw{Env} that extracts an environment containing the names and
+values of all the form fields from an HTML document, then a formlens $f$
+satisfies the Acceptability law if for any $v$ of type $a$, $n$ of type
+\kw{Int} and environment $e$ of type \kw{Env} if $f$ applied to \kw{Just v} and
+$n$ yields an HTML file $h$ and a {\em collect} function $c$ where \kw{extract
+h} is contained in $e$, then collect $e$
 \fi
-
-Finally, Cunha~\cite{cunha2010relational} uses relational algebra to
-encompass many of the existing approaches to bidirectional programming. For
-example, using $\cdot$\saz{Why not use $\circ$ or ``;''} to express relational composition, the \textsc{GetPut} law states
-that $\get \cdot \lput \subseteq \equiv_V \cdot \pi_1$, where $\pi_1$ is the
+Another recent line of work that exploits equivalence relations on data is by
+Hilken et. al. in \cite{hilken2016testing}. This work is concerned mainly with
+testing, and uses {\em Equivalence Partitioning}, a software testing technique
+that divides the input data of a software unit into partitions of equivalent
+data from which test cases can be derived. This approach has the advantage that
+it can greatly reduce the number of test cases and hence the amount of time
+spent testing since only one representative from each class is tested.
+
+More concretely, their approach gives the developer an explicit option to
+formulat her understanding of two object models being different. The technical
+realization is as follows: The developer specifies ``classifying term'', a
+closed Object Constraint Language (OCL) query term that can be evaluated in an
+object model and returns a characteristic value. Two object models with the same
+characteristic value belong to the same equivalence class. For example under a
+configuration requiring at least 2 and at most 4 Person objects, the
+classifying term \lstinline{Person.allInstances()->size()} would yield three
+object models with 2,3 and 4 Person objects respectively.
+
+Classifying terms are therefore analogous to our QREs or to Boomerang canonizers
+in the sense that they define an equivalence relation on data. However, while
+QREs and canonizers aim at identifying or characterizing the elements that
+should be treated equally by a lens, the focus of the work by Hilken et. al. is
+on classes that represent specific patterns of particular relevance to the
+modeler who is interested in analyzing the behavior of a transformation.
+
+Finally, Cunha~\cite{cunha2010relational} uses relational algebra to encompass
+many of the existing approaches to bidirectional programming. For example,
+using $\circ$ to express relational composition, the \textsc{GetPut} law states
+that $\get \circ \lput \subseteq \equiv_V \circ \pi_1$, where $\pi_1$ is the
 projection onto the first factor. In addition to its generality, this
@@
 Next we turn to related work in program synthesis. Much of the research in
-synthesis 
-assumes that the synthesizer is provided with a collection of examples.
-Optician and Optician$_Q$  differ
-in that they require the programmer to supply both examples {\em and}
-format descriptions in the form of regular expressions or QREs, though we are
-far from the only ones to consider type-based synthesis.
-There is a trade-off here.  On the one hand, a user must have some programming
-expertise to write regular expression (or QRE) specifications, and it requires
-some work. On the other hand, such specifications provide a great deal of
-information to the synthesis system, which decreases the number of examples
-needed (often to zero), makes the system scale well, and allows it to handle
-large, complex formats.  By providing these format specifications, the 
-synthesis engine does not have to both infer the format of the data as well as
-the transformations on it, obviating the need to infer tricky formats like
-those involving nested iterations. Furthermore, by focusing on bidirectional
-transformations, we limit the space of synthesized functions to bijective ones,
-reducing the search space, and the expressiveness of the search space.
+synthesis assumes that the synthesizer is provided with a collection of examples.
+Optician and Optician$_Q$ differ in that they require the programmer to supply
+both examples {\em and} format descriptions in the form of regular expressions
+or QREs, though these two systems are far from the only ones to consider
+type-based synthesis. There is a trade-off here.  On the one hand, a user must
+have some programming expertise to write regular expression (or QRE)
+specifications, and it requires some work. On the other hand, such
+specifications provide a great deal of information to the synthesis system,
+which decreases the number of examples needed (often to zero), makes the system
+scale well, and allows it to handle large, complex formats.  By providing these
+format specifications, the synthesis engine does not have to both infer the
+format of the data as well as the transformations on it, obviating the need to
+infer tricky formats like those involving nested iterations.
 
@@
 Optician is distinctive in that it synthesizes terms in a language with many
-type equivalences.
-Perhaps the system most similar to Optician is InSynth~\cite{gvero-pldi-2013}, a
-system for synthesizing terms in the simply-typed lambda calculus that addresses
-equivalences on types. Instead of trying to directly synthesize terms of the
-simply-typed lambda calculus, InSynth synthesizes a well-typed term
-in the succinct calculus, a language with types
-that are equivalent ``modulo isomorphisms of products and
-currying''~\cite{gvero-pldi-2013}. The type structure that we used in Optician
+type equivalences. Perhaps the system most similar to Optician is
+InSynth~\cite{gvero-pldi-2013}, a system for synthesizing terms in the
+simply-typed lambda calculus that addresses equivalences on types. Instead of
+trying to directly synthesize terms of the simply-typed lambda calculus,
+InSynth synthesizes a well-typed term in the succinct calculus, a language with
+types that are equivalent ``modulo isomorphisms of products and
+currying''~\cite{gvero-pldi-2013}. The type structure used in Optician
 is significantly more complex.  In particular, because Optician types do not
-have full canonical forms, we used a pseudo-canonical form, which captures part
-of the equivalence relation over types. To preserve completeness, we pushed
-some of the remaining parts of the type equivalence relation into a set of
-rewriting rules and other parts into the synthesis algorithm itself.
+have full canonical forms, Anders et. al. used a pseudo-canonical form that
+captures part of the equivalence relation over types were used. To preserve
+completeness, they pushed some of the remaining parts of the type equivalence
+relation into a set of rewriting rules and other parts into the synthesis
+algorithm itself.
 
@@
 communicating synthesizers to generate programs.  In both Morpheus and
-Optician, one synthesizer provides an
-outline for the program, and the other fills in that outline with program
-details that satisfy the user's specifications.
-This approach works well in large search spaces that require some enumerative
-search.
-One important way that Optician differs from Morpheus is that in
-Morpheus, an outline is a sketch---an
-\emph{expression}
-containing holes---whereas
-an outline in Optician is a pair of regular
-expressions, i.e., a
-\emph{type}.  Moreover, in order to implement an efficient
-search procedure, we had to create both a new type language and a new
-term language for lenses.  Once we did so, we proved our new, more
-constrained language
-designed for synthesis was just as expressive as the original, more
-flexible and compositional language designed for human programmers.
+Optician, one synthesizer provides an outline for the program, and the other
+fills in that outline with program details that satisfy the user's
+specifications. This approach works well in large search spaces that require
+some enumerative search. One important way that Optician differs from Morpheus
+is that in Morpheus, an outline is a sketch---an \emph{expression}
+containing holes---whereas an outline in Optician is a pair of regular
+expressions, i.e., a \emph{type}.  Moreover, in order to implement an efficient
+search procedure, Anders et. al had to create both a new type language and a new
+term language for lenses.  Once they did so, they proved their new, more
+constrained language designed for synthesis was just as expressive as the
+original, more flexible and compositional language designed for human
+programmers.
 

### 2018-03-14 papers/QuotientLenses/main.tex
minor responses to things in section 5

--- 2018-03-14 papers/QuotientLenses/main.tex
@@
 
-This rule essentially says that the composition $q_1 \; ; \; q_2$ is
-well defined if and only if the intermediary QREs $Q_3$ and $Q_4$ define the
-same equivalence relation on regular languages that are equivalent modulo the
-star-semiring axioms\bcp{No: the regular expressions are equivalent modulo
-  these axioms; the regular {\em languages} are
-  identical}. (See~\cite[\S 4]{quotientlenses} for an example of what goes
-wrong if this premise is dropped.\bcp{Didn't we used to have a discussion
-  about how implementing this rule is hard because it involves checking
-  language equivalence so we follow Boomerang's rule?  I think this is
-  important to mention because this was a significant restriction on the
-  power of composition in Boomerang.  It doesn't apply here (right?) because
-we intend to synthesize these lenses, not write them by hand, so we will
-never actually need to typecheck a functional composition expression.}) The
-condition $\mathcal{L}(W(Q_2)) = \mathcal{L}(W(Q_3))$ 
-says that the intermediary language is the same on both sides, while the
-condition $K(Q_2) \equiv^s K(Q_3)$ says that the kernel regular expressions are
-equivalent modulo the star-semiring axioms. Finally, the condition
-$\canonize(Q_2) = \canonize(Q_3)$ says that $Q_2$ and $Q_3$ define the same
-equivalence relation on $\mathcal{L}(W(Q_2))= \mathcal{L}(W(Q_3))$.
+This rule essentially says that the composition $q_1 \; ; \; q_2$ is well
+defined if and only if the intermediary QREs $Q_3$ and $Q_4$ define the same
+equivalence relation on regular expressions that are equivalent modulo the
+star-semiring axioms. (See~\cite[\S 4]{quotientlenses} for an example of what
+goes wrong if this premise is dropped.\bcp{Didn't we used to have a discussion
+  about how implementing this rule is hard because it involves checking language
+  equivalence so we follow Boomerang's rule? I think this is important to
+  mention because this was a significant restriction on the power of composition
+  in Boomerang. It doesn't apply here (right?) because we intend to synthesize
+  these lenses, not write them by hand, so we will never actually need to
+  typecheck a functional composition expression.}\afm{Boomerang does check
+  language equivalence, we restrict the check to star-semiring equivalence
+  because we aren't complete when we allow arbitrary equivalence.}) The
+condition $\mathcal{L}(W(Q_2)) = \mathcal{L}(W(Q_3))$ says that the intermediary
+language is the same on both sides, while the condition $K(Q_2) \equiv^s K(Q_3)$
+says that the kernel regular expressions are equivalent modulo the star-semiring
+axioms. Finally, the condition $\canonize(Q_2) = \canonize(Q_3)$ says that $Q_2$
+and $Q_3$ define the same equivalence relation on $\mathcal{L}(W(Q_2))=
+\mathcal{L}(W(Q_3))$.
 

### 2018-03-14 papers/QuotientLenses/main.tex
Added Conclusion

@@
 \section{Conclusion}
+In this paper, we showed how to synthesize a class of bijective quotient lenses
+using the bijective lens synthesis system Optician as a plug-in for our
+synthesis algorithm. In order to achieve this, we first introduced {\em
+Quotient regular languages} to specify regular languages with an equivalence
+relation defined on them. Then, we introduced {\em QRE lenses}, which
+are bijective quotient lenses that map between QREs via bijective lenses. We
+proved a normal form theorem for QREs that enabled us to (1) extend the
+synthesis algorithm used by Optician to synthesize QRE lenses, and (2) prove
+that if there is a QRE lens that satisfies the input specification, then the
+extended algorithm returns such a lens. Finally we tested QRE-enhanced optician
+on the Optician benchmark suite and demonstrated that QREs can save programmers
+significant effort compared to both Optician and basic Boomerang, and that
+QRE-enhanced Optician maintains competitive performance over both of these
+systems.
 \bcp{And future work!}  \saz{I'm lukewarm about including a description of

### 2018-03-14 papers/QuotientLenses/main.tex
Nits

@@
 class of {\em bijective lenses} that can be derived in the lens programming
-language {\em Boomerang}~\cite{boomerang}. Each lens $\ell : S \Leftrightarrow T$
+language {\em Boomerang}~\cite{boomerang}\bcp{All of them?  I'm not certain
+  that's accurate.}. Each lens $\ell : S \Leftrightarrow T$
 that Optician can synthesize specifies a bijection from the language of a
@@
 holds modulo equivalence relations $\equiv_S$ and $\equiv_T$ defined on the source
-and target data respectively:
+and target data respectively:\bcp{Not too convinced about the colors.  The
+  fact that lens expressions mix black and red is distracting -- my eyes
+  can't easily tell which bits of a larger expression are lenses.}
 \begin{equation}\label{quotientlenslaws}
@@
 
-For example, consider the following QRE for writing author names:
+For example, consider the following QRE for writing author
+names:\bcp{Solomon: Someone noted on their copy of the intro from
+  class yesterday that this example is backwards.  Did you go over the
+  marked-up copies that I gave you?}
 %

### 2018-03-14 papers/QuotientLenses/main.tex
Bunch of comments

@@
 guarantee various \emph{lens laws} constraining how the \kw{put} and
-\kw{get} functions must fit together.
+\kw{get} functions must fit together. \bcp{See my email about deemphasizing
+  ``storing'' multiple formats.}
 
@@
   fact that lens expressions mix black and red is distracting -- my eyes
-  can't easily tell which bits of a larger expression are lenses.}
+  can't easily tell which bits of a larger expression are lenses.  I would
+  suggest either:
+  \begin{itemize}
+  \item get and put italicized, plus
+  \item either and lens expressions (completely) in some subtle color (maybe
+  brown), or simply black.
+  \end{itemize}
+}
 \begin{equation}\label{quotientlenslaws}
@@
 \noindent
-and equivalent EndNote records like the following
+and equivalent EndNote records like the following\bcp{bad punctuation: needs
+a colon}
 \begin{verbatim}
@@
 expressions (which are omitted for brevity).  
-\iflastminute
-\bcp{a possible readability suggestion: define a named constant for a quote
-  character.  (Not sure if this makes sense, or if it improves things...)}
-\fi}
+}
 \label{fig:example-lens}
@@
   big improvement.  One tiny possible tweak: color the \kw{get} and \kw{put}
-  keywords in the same dark red as in the text.}
+  keywords in the same dark red as in the text.}\bcp{Can do, if we keep the
+  color in the text :-).  How are we going to decide about that?}
 \label{fig:attheedges}
@@
 Formally, the language of Quotient Regular Expressions (QREs) is given by the
-following grammar:
+following grammar:\bcp{bad punctuation}
 \newcommand{\bsep}{\ \ \sep{} \ \ }
@@
 We define the set of \textit{bijective string lenses} to be the set of
-bijections between regular languages constructed using the following combinators.
+bijections between regular languages constructed using the following
+combinators.\bcp{bad punctuation}
 
@@
 introduce the class of QRE lenses. The language of QRE lenses is given by following
-grammar:
+grammar:\bcp{bad punctuation}
 $$ q := \lift(\ell) \sep  q_1 \cdot q_2 \sep \swap(q_1, q_2) \sep (q_1 \sep q_2)
@@
 
-By the induction hypothesis, there exist bijective lenses
+By the induction hypothesis, there exist bijective lenses\bcp{every display
+  for the rest of the section needs punctuation help}
 $\ell_1 : K(Q_1) \Leftrightarrow K(Q_2)$ and $\ell_2 : K(Q_3) \Leftrightarrow
@@
 describe a \bibtex{} record and an Endnote record respectively. We also had
-the example pair \cd{(bib_example , end_example)} where
+the example pair \cd{(bib_example , end_example)} where\bcp{punctuation}
 
@@
 functions $q.\kw{canonize} : C \longrightarrow B$ and $q.\kw{choose} : B
-\longrightarrow C$ such that for every $b \in B$,
+\longrightarrow C$ such that for every $b \in B$,\bcp{punctuation}
 $$q.\kw{canonize} \; (q.\kw{choose} \; b) \equiv_B \; b$$

### 2018-03-14 papers/QuotientLenses/main.tex
More edits

@@
 \newcommand{\lift}{\ensuremath{\kw{lift}}}
+\newcommand{\bsep}{\ \ \sep{} \ \ }
 
@@
 \section{Introduction}
-Often, the same data is stored in more than one format. For example,
-bibliographic data may be stored in \bibtex{} or EndNote formats,
-spreadsheet data can be represented as either CSV and TSV files, and web
-APIs can return data as either JSON or XML objects.  When changes are
-made one version, the data in the other must be updated so the
-two continue to represent the same information.
-One way to synchronize such data is to use \emph{lenses}~\cite{Lenses}, 
-programs that simultaneously define pairs of functions for translating
-the data from the source format to the target formant (the \emph{get}
-direction) and back again (the \emph{put} direction). Lenses can be
-written in one of several domain-specific bidirectional languages that 
-guarantee various \emph{lens laws} constraining how the \kw{put} and
-\kw{get} functions must fit together. \bcp{See my email about deemphasizing
-  ``storing'' multiple formats.}
-
-Recent work~\cite{optician}\saz{removed ``we'' here to maintain the double-blind
-  review standard.} developed a tool called Optician to synthesize the
-class of {\em bijective lenses} that can be derived in the lens programming
-language {\em Boomerang}~\cite{boomerang}\bcp{All of them?  I'm not certain
-  that's accurate.}. Each lens $\ell : S \Leftrightarrow T$
-that Optician can synthesize specifies a bijection from the language of a
-source regular expression $S$ to a target language of a target regular
-expression $T$: \begin{equation}\label{bijectivelenslaws} \ell.\get \;
+Programmers often need to write programs that bidirectionally convert data
+between two different formats. A time-tested and appealing way to do this is
+to use \emph{lenses}~\cite{Lenses}, programs that simultaneously define pairs of
+functions for translating the data from the source format to the target format
+(the \emph{get} direction) and back again (the \emph{put} direction). The
+main reason that lenses are appealing is because lens programming languages
+guarantee that the \kw{put} and \kw{get} functions of a lens satisfy
+certain \emph{lens laws} constraining how the two functions interact with each
+other.
+
+Recent work~\cite{optician} developed a tool called Optician to synthesize a
+subset of the {\em bijective lenses} that can be derived in the lens
+programming language {\em Boomerang}~\cite{boomerang}. Each lens $\ell : S
+\Leftrightarrow T$ that Optician can synthesize specifies a bijection from the
+language of a source regular expression $S$ to a target language of a target
+regular expression $T$: \begin{equation}\label{bijectivelenslaws} \ell.\get \;
 (\ell.\lput \; t) = t \text{, and } \ell.\lput \; (\ell.\get \; s) = s
@@
   class yesterday that this example is backwards.  Did you go over the
-  marked-up copies that I gave you?}
+  marked-up copies that I gave you?}\sam{I'm not sure what this comment means
+  and also I forgot those notes in the office. I'll go through them more
+  thoroughly tomorrow morning.}
 %
@@
 \noindent
-and equivalent EndNote records like the following\bcp{bad punctuation: needs
-a colon}
+and equivalent EndNote records like the following:
 \begin{verbatim}
@@
 Formally, the language of Quotient Regular Expressions (QREs) is given by the
-following grammar:\bcp{bad punctuation}
-\newcommand{\bsep}{\ \ \sep{} \ \ }
-\begin{align*}
-Q := \; & \normalize{R_1}{R_2}{f} \bsep{} \id(R) \bsep \collapse{R}{s} \bsep \squash{Q_1}{Q_2}{f} \\
- | \; & \perm{Q_1, \ldots, Q_n}{Q} \bsep Q_1 \semicolon Q_2 \bsep Q_1 \cdot Q_2 \bsep (Q_1
-\sep Q_2) \bsep Q^*
+following grammar,
+\begin{align*}
+Q := \; & \normalize{R_1}{R_2}{f} \bsep{} \id(R) \bsep \collapse{R}{s} \bsep
+\squash{Q_1}{Q_2}{f} \\ | \; & \perm{Q_1, \ldots, Q_n}{Q} \bsep Q_1 \semicolon
+Q_2 \bsep Q_1 \cdot Q_2 \bsep (Q_1 \sep Q_2) \bsep Q^*
 \end{align*}
@@
 objects fit together to form a coherent quotient $W(Q)/\eqrel{Q}$ whose
-equivalence classes are determined by \canonize(Q).
+equivalence classes are determined by $\canonize(Q)$.
 
@@
 conditions as an ``escape hatch'' for the case when other QRE
-combinators are insufficient.\bcp{Do we use this in any of the
-programs we've written?}\afm{Nope.}}
+combinators are insufficient.}
 
@@
 \end{cases}\\[3ex]
-\canonize(Q^*)(\epsilon) &=& \epsilon \\
-\canonize(Q^*)(w_1 \cdot w_2) &=& \canonize(Q)(w_1)\cdot\canonize(Q^*)(w_2) \\[2ex]
+\canonize(Q^*)(w_1 \cdot \ldots \cdot w_n) &=&
+\begin{cases}
+\epsilon & \text{if } n = 0\\
+\canonize(Q)(w_1)\cdot \ldots \cdot \canonize(Q)(w_n) & \text{if } n > 0
+\end{cases}
+\\[2ex]
 \multicolumn{3}{l}{\canonize(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
-\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)})(w_{\sigma(1)} \cdot s_1
-  \cdot \ldots \cdot s_{n-1} \cdot w_{\sigma(n)})}\\[.5ex]
-\multicolumn{3}{l}{\ =\ \ \canonize(Q_1)(r_1)(w_1) \cdot \canonize(Q)(s_1) \cdot \ldots \cdot
-\canonize(Q)(s_{n-1})\cdot\canonize(Q_n)(r_n)(w_n)}
+\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)})(w_{\sigma(1)} \cdot
+s_1 \cdot \ldots \cdot s_{n-1} \cdot w_{\sigma(n)})}\\[.5ex]
+\multicolumn{3}{l}{\ =\ \ \canonize(Q_1)(r_1)(w_1) \cdot \canonize(Q)(s_1)
+\cdot \ldots \cdot \canonize(Q)(s_{n-1})\cdot\canonize(Q_n)(r_n)(w_n)}
 \end{array}
@@
 
-First we give the formal definition of a {\em bijective quotient lens}; QRE
-lenses are a specific class of bijective quotient lenses that map between data
-that is specified by QREs. Given regular expressions $R, S$ and equivalence
-relations $\equiv_R$ and $\equiv_S$ defined on $\mathcal{L}(R)$ and $\mathcal{L}(S)$,
-we define a bijective quotient lens $q : R /{\equiv_R} \Leftrightarrow
-S/{\equiv_S}$ to be a pair of functions $q.\get :
-\mathcal{L}(R) \longrightarrow \mathcal{L}(S)$ and $q.\lput : \mathcal{L}(S)
-\longrightarrow \mathcal{L}(R)$ such that for all $r \in \mathcal{L}(R)$ and $s
-\in \mathcal{L}(S)$, we have $q.\lput(q_1.\get(r)) \equiv_R r$ and
-$q.\get(q_1.\lput(s)) \equiv_R s$. Moreover, if $r \equiv_R r'$, then $q.\get(r) =
-q.\get(r')$, and if $s \equiv_S s'$, then $q.\lput(s) = q.\lput(s')$.
-
-In words, a bijective quotient lens $q$ from $R$ to $S$ modulo $\equiv_R$
-and $\equiv_S$ is a pair of functions $q.\get$ and $q.\lput$ such that
-$q.\get$ respects $\equiv_R$ and $q.\lput$ respects $\equiv_S$, and such that
-the {\em get} and {\em put} functions lifted to $\mathcal{L}(R)/{\equiv_R}$ and
-$\mathcal{L}(S)/{\equiv_S}$ are mutual inverses.
-
-These laws are similar to the laws for bijective lenses
-\cref{bijectivelenslaws}\saz{again, cref uses a silly format}, except that the  equality restrictions in the
-bijective lens laws are loosened to allow for equivalence relations. Also, the
-condition that $r \equiv_R r'$ implies $q.\get(r) = q.\get(r')$ ensures that the
-{\em get} function induced on the equivalence classes if $\equiv_c$ is
-well-defined, and similarly for the {\em put} function.
+Recall that given regular expressions $R, S$ and equivalence relations
+$\equiv_R$ and $\equiv_S$ defined on $\mathcal{L}(R)$ and $\mathcal{L}(S)$, a
+{\em bijective quotient lens} $q : R /{\equiv_R} \Leftrightarrow S/{\equiv_S}$
+is a pair of functions $q.\get : \mathcal{L}(R) \longrightarrow \mathcal{L}(S)$
+and $q.\lput : \mathcal{L}(S) \longrightarrow \mathcal{L}(R)$ such that for all
+$r \in \mathcal{L}(R)$ and $s \in \mathcal{L}(S)$, we have $q.\lput(q_1.\get(r))
+\equiv_R r$ and $q.\get(q_1.\lput(s)) \equiv_R s$. Moreover, if $r \equiv_R
+r'$, then $q.\get(r) = q.\get(r')$, and if $s \equiv_S s'$, then $q.\lput(s) =
+q.\lput(s')$. In words, a bijective quotient lens $q$ from $R$ to $S$ modulo
+$\equiv_R$ and $\equiv_S$ is a pair of functions $q.\get$ and $q.\lput$ such
+that $q.\get$ respects $\equiv_R$ and $q.\lput$ respects $\equiv_S$, and such
+that the {\em get} and {\em put} functions lifted to $\mathcal{L}(R)/{\equiv_R}$
+and $\mathcal{L}(S)/{\equiv_S}$ are mutual inverses. These laws are similar
+bijective lens laws \cref{bijectivelenslaws}, except that the equality
+restrictions in the bijective lens laws are loosened to allow for equivalence
+relations. Also, the condition that $r \equiv_R r'$ implies $q.\get(r) =
+q.\get(r')$ ensures that the {\em get} function induced on the equivalence
+classes if $\equiv_c$ is well-defined, and similarly for the {\em put}
+function.
 
@@
 bijections between regular languages constructed using the following
-combinators.\bcp{bad punctuation}
-
+combinators,
 $$\ell := \id(R) \sep \const(s, t) \sep  \swap(\ell_1, \ell_2)
 \sep \ell_1 \cdot \ell_2 \; |  \; (\ell_1 \sep \ell_2) \sep \ell^* \; | \;
-\ell_1 \; ; \;  \ell_2,$$ where $R$ ranges over regular expressions and $s$
-ranges over character strings.
+\ell_1 \; ; \;  \ell_2$$
+where $R$ ranges over regular expressions and $s$ ranges over character strings.
 
@@
 Having given a brief overview of the class of bijective string lenses, we now
-introduce the class of QRE lenses. The language of QRE lenses is given by following
-grammar:\bcp{bad punctuation}
+introduce the class of QRE lenses. The language of QRE lenses is given by
+following grammar,
 $$ q := \lift(\ell) \sep  q_1 \cdot q_2 \sep \swap(q_1, q_2) \sep (q_1 \sep q_2)
-\sep q^* \sep q_1 \; ; \; q_2 \sep \lquot(q, \ell) \sep \rquot(\ell, q),$$ where
-$\ell$ ranges over bijective lenses.
+\sep q^* \sep q_1 \; ; \; q_2 \sep \lquot(q, \ell) \sep \rquot(\ell, q)$$
+where $\ell$ ranges over bijective lenses.
 
@@
 star-semiring axioms. (See~\cite[\S 4]{quotientlenses} for an example of what
-goes wrong if this premise is dropped.\bcp{Didn't we used to have a discussion
-  about how implementing this rule is hard because it involves checking language
-  equivalence so we follow Boomerang's rule? I think this is important to
-  mention because this was a significant restriction on the power of composition
-  in Boomerang. It doesn't apply here (right?) because we intend to synthesize
-  these lenses, not write them by hand, so we will never actually need to
-  typecheck a functional composition expression.}\afm{Boomerang does check
-  language equivalence, we restrict the check to star-semiring equivalence
-  because we aren't complete when we allow arbitrary equivalence.}) The
-condition $\mathcal{L}(W(Q_2)) = \mathcal{L}(W(Q_3))$ says that the intermediary
-language is the same on both sides, while the condition $K(Q_2) \equiv^s K(Q_3)$
-says that the kernel regular expressions are equivalent modulo the star-semiring
-axioms. Finally, the condition $\canonize(Q_2) = \canonize(Q_3)$ says that $Q_2$
-and $Q_3$ define the same equivalence relation on $\mathcal{L}(W(Q_2))=
+goes wrong if this premise is dropped). The condition $\mathcal{L}(W(Q_2)) =
+\mathcal{L}(W(Q_3))$ says that the intermediary language is the same on both
+sides, while the condition $K(Q_2) \equiv^s K(Q_3)$ says that the kernel
+regular expressions are equivalent modulo the star-semiring axioms. Finally,
+the condition $\canonize(Q_2) = \canonize(Q_3)$ says that $Q_2$ and $Q_3$
+define the same equivalence relation on $\mathcal{L}(W(Q_2))=
 \mathcal{L}(W(Q_3))$.
+
+Of course checking the condition $\canonize(Q_2) = \canonize(Q_3)$ is
+undecidable in general; indeed the approach used to type quotient lenses in
+Boomerang is to classify equivalences according to whether they are or are not
+the equality relation. While this decision seemingly restricts the
+power of composition in Boomerang significantly, the practice of writing
+quotient lenses shows that this restriction is not overly restrictive. This is 
+because most quotient lenses originate as lifted basic lenses, and
+therefore have types whose equivalence relations are both equality, and further,
+equality is preserved by many of the quotient lens combinators
+~\cite{quotientlenses}. Foster et al also discuss a second possible approach to
+typing quotient lenses, where equivalence relations are represented by rational
+functions that induce them. While this second approach is more refined than the
+first, Boomerang favours the first approach since the second appears to be too
+expensive to be useful in practice.
+
+Thankfully, we do not face the issue of checking equivalence relation equality
+in our work since our end goal is to synthesize lenses and not write them my
+hand, so the programmer will never actually need to typecheck a functional
+composition expression.
 
@@
 
-By the induction hypothesis, there exist bijective lenses\bcp{every display
-  for the rest of the section needs punctuation help}
+By the induction hypothesis, there exist bijective lenses,
 $\ell_1 : K(Q_1) \Leftrightarrow K(Q_2)$ and $\ell_2 : K(Q_3) \Leftrightarrow
 K(Q_4)$ such that
-\begin{center}
-
+\begin{center}:
 \begin{tabular}{l l}
@@
 \end{tabular}
-
 \end{center}
 %
-Consequently,
+Consequently:
 \begin{align*}
@@
 $\canonize(Q_3)$ is the identity function on $K(Q_2)$ which is
-syntactically equal to $K(Q_3)$ by assumption. A similar argument shows that
+syntactically equal to $K(Q_3)$ by assumption. A similar argument shows that:
 $$\llbracket q_1 \rrbracket.\lput \circ \llbracket q_2 \rrbracket.\lput =
@@
 
-The other cases of the proof are similar, proceeding by a
-straightforward application of the induction hypothesis followed by unrolling
-the equations that give the denotation for QRE lenses.  Full details can be
-found in the appendix.
+The other cases of the proof are similar, proceeding by a straightforward
+application of the induction hypothesis followed by unrolling the equations
+that give the denotation for QRE lenses.  Full details can be found in the
+appendix.
 \end{proof}
@@
 types.  Rather than writing the QRE lens manually, the programmer can instead
-specify the desired behavior of a lens by giving its interface types and providing
-examples, if necessary, to disambiguate among possible implementations.  This
-way of constructing lenses can often be much simpler than building them by
-hand, as we saw in Section~\ref{sec:examplesynth}.
+specify the desired behavior of a lens by giving its interface types and
+providing examples, if necessary, to disambiguate among possible
+implementations.  This way of constructing lenses can often be much simpler
+than building them by hand, as we saw in Section~\ref{sec:examplesynth}.
 
@@
 describe a \bibtex{} record and an Endnote record respectively. We also had
-the example pair \cd{(bib_example , end_example)} where\bcp{punctuation}
-
+the example pair \cd{(bib_example , end_example)}, where:
 \begin{center}
@@
 \end{center}
-
 There exists a bijective lens between the kernel of \cd{bibtex} and the kernel
@@
 functions $q.\kw{canonize} : C \longrightarrow B$ and $q.\kw{choose} : B
-\longrightarrow C$ such that for every $b \in B$,\bcp{punctuation}
+\longrightarrow C$ such that for every $b \in B$:
 $$q.\kw{canonize} \; (q.\kw{choose} \; b) \equiv_B \; b$$
@@
 data sequences described by the two types can be matched. This process can
-result in intermediate types that are ambiguous\BCP{???}\sam{Before I had this
-as unambiguous. Was this the issue? Or does this need more explaining even if
-the word unambiguous was the issue}, and this ambiguity can cause unintended
-updates to be made the source, especially when information is discarded in this
-process.
+result in intermediate types that are ambiguous, and this ambiguity can cause
+unintended updates to be made the source, especially when information is
+discarded in this process.
 
@@
 
-\section{Conclusion}
+\section{Conclusion and Future Work}
 In this paper, we showed how to synthesize a class of bijective quotient lenses
@@
 systems.
+
 \bcp{And future work!}  \saz{I'm lukewarm about including a description of
@@
 
-\bcp{Some of the citations need some love---e.g., the first one.}
-
 \bibliographystyle{plain}
@@
   $\kw{lift}(\ell): R/\mathit{id}(R) \Leftrightarrow S/\mathit{id}(S)$ where
-  $\ell : R \Leftrightarrow S$. Then
+  $\ell : R \Leftrightarrow S$. Then:
   \begin{align*}
@@
 $\lquot(Q_1, q): Q_1 \; ; \; Q_2 \Leftrightarrow Q_3$ where $q : Q_2
-\Leftrightarrow Q_3$, $Q_1$ is well formed and $K(Q_1) = W(Q_2)$. Then
+\Leftrightarrow Q_3$, $Q_1$ is well formed and $K(Q_1) = W(Q_2)$. Then:
 \begin{align*}
@@
 By the induction hypothesis, there exists a bijective lens $\ell :
-K(Q_2) \Leftrightarrow K(Q_3)$ such that
+K(Q_2) \Leftrightarrow K(Q_3)$ such that:
 \begin{align*}
@@
 (\canonize(Q_1 \; ; \; Q_2))\\
-\llbracket \lquot(Q_1, q) \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
-\canonize(Q_3)
+\llbracket \lquot(Q_1, q) \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1}
+\circ \canonize(Q_3)
 \end{align*}
@@
 $q_1 \; ; \; q_2: c \Leftrightarrow Q_2$ where $q_1 : c \Leftrightarrow Q_1$ and
-$q_2 : Q_1 \Leftrightarrow Q_2$. Then
+$q_2 : Q_1 \Leftrightarrow Q_2$. Then:
 \begin{align*}
@@
 \end{align*}
-By the induction hypothesis, there exist bijective lenses
+By the induction hypothesis, there exist bijective lenses,
 $\ell_1 :
 K(c) \Leftrightarrow K(Q_1)$ and $\ell_2 : K(Q_1) \Leftrightarrow K(Q_2)$ such
-that
+that,
 \begin{align*}
@@
 \end{align*}
-Consequently,
+Consequently:
 \begin{align*}
@@
 \end{align*}
-A similar argument shows that
+A similar argument shows that:
 $$\llbracket q_1 \; ; \; q_2 \rrbracket.\lput =
@@
 $q^* : {Q_1}^* \Leftrightarrow {Q_2}^*$ where $q : Q_1 \Leftrightarrow Q_2$,
-$W(Q_1)^{*!}$ and $W(Q_2)^{*!}$ and $K(Q_1)^{*!}$ and $K(Q_2)^{*!}$. Then
+$W(Q_1)^{*!}$ and $W(Q_2)^{*!}$ and $K(Q_1)^{*!}$ and $K(Q_2)^{*!}$. Then:
 \begin{align*}
@@
 By the induction hypothesis there exists a bijective lens $\ell : K(Q_1)
-\Leftrightarrow K(Q_2)$ such that
-that
+\Leftrightarrow K(Q_2)$ such that:
 \begin{align*}
@@
 \end{align*}
-Consequentlty
+Consequentlty:
 \begin{align*}
@@
 \cdot^! W(Q_2)$, $K(Q_1) \cdot^! K(Q_2)$, $W(Q_3) \cdot^! W(Q_4)$ and $
-K(Q_3) \cdot^! K(Q_4)$. Then
+K(Q_3) \cdot^! K(Q_4)$. Then:
 \begin{align*}
@@
 By the induction hypothesis, there exist bijective lenses $\ell_1 : K(Q_1)
-\Leftrightarrow K(Q_3)$ and $\ell_2 : K(Q_2) \Leftrightarrow K(Q_4)$ such that
+\Leftrightarrow K(Q_3)$ and $\ell_2 : K(Q_2) \Leftrightarrow K(Q_4)$ such that,
 \begin{align*}
@@
 \end{align*}
-Similarly
+Similarly:
 $$
@@
 \cdot^! W(Q_2)$, $K(Q_1) \cdot^! K(Q_2)$, $W(Q_4) \cdot^! W(Q_3)$ and $
-K(Q_4) \cdot^! K(Q_3)$. Then
+K(Q_4) \cdot^! K(Q_3)$. Then:
 \begin{align*}
@@
 \varnothing$ and $\mathcal{L}(W(Q_3)) \cap \mathcal{L}(W(Q_4)) = \varnothing$.
-Then
-$$
+Then:$$
 \llbracket q_1 \sep q_2 \rrbracket.\get(s) =
@@
 \end{align*}
-Consequently,
+Consequently:
 $$
@@
   \Call{SynthBijectiveLens}{R_1,R_2,exs'}$, where $exs' = \Call{Map}{exs,fun \,
-    (ex_l,ex_r) \to (\canonize(Q_1)(ex_l),\canonize(Q_2)(ex_r))}$. By the correctness of
-  $\textproc{SynthBijectiveLens}$, we know that $\ell.\get(x_i') = y_i'$ and
-  $\ell.\lput(y_i') = x_i'$ for all $(x_i',y_i')\in exs'$. By definitions of
-  $\rquot$ and $\lquot$, this means that:\\
+    (ex_l,ex_r) \to (\canonize(Q_1)(ex_l),\canonize(Q_2)(ex_r))}$. By the
+    correctness of $\textproc{SynthBijectiveLens}$, we know that
+    $\ell.\get(x_i') = y_i'$ and $\ell.\lput(y_i') = x_i'$ for all
+    $(x_i',y_i')\in exs'$. By definitions of $\rquot$ and $\lquot$, this means
+    that:\\
 $\rquot(\lquot(Q_1, \ell),

### 2018-03-15 papers/QuotientLenses/main.tex
Quick future work paragraphs

@@
 Programmers often need to write programs that bidirectionally convert data
-between two different formats. A time-tested and appealing way to do this is
+between two different formats.\bcp{It's a shame to lose the concrete
+  examples here.} A time-tested and appealing way to do this is
 to use \emph{lenses}~\cite{Lenses}, programs that simultaneously define pairs of
@@
   and also I forgot those notes in the office. I'll go through them more
-  thoroughly tomorrow morning.}
+  thoroughly tomorrow morning.}\bcp{What the comment means is that names are
+not usually formatted as ``Solomon, Maina'' but rather as ``Maina, Solomon''.}
 %
@@
 id(S)}{\mathit{(Lift)}} &
-\begin{array}[b]{rcl}
+\begin{array}[b]{r@{\ }c@{\ }l}
 \llbracket \kw{lift}(\ell) \rrbracket.\get &=&  \llbracket \ell \rrbracket\\
@@
 \]
+
 \[
@@
 {\lquot(Q_1, q): Q_1 \; ; \; Q_2 \Leftrightarrow Q_3}{\mathit(Lquot)} &
-\begin{array}[b]{rcl}
+\begin{array}[b]{r@{\ }c@{\ }l}
 \llbracket \lquot(Q_1, q) \rrbracket.\get  &=& \llbracket q
@@
 {\rquot(q, Q_1):Q_1 \Leftrightarrow Q_2 \; ; \; Q_3}{\mathit{(Rquot)}} &
-\begin{array}[b]{rcl}
+\begin{array}[b]{r@{\ }c@{\ }l}
 \llbracket \rquot(q, Q_1) \rrbracket.\get  &=& \llbracket q
@@
 {{q}^* : {Q_1}^* \Leftrightarrow {Q_2}^*}{\mathit{(Star)}} &
-\begin{array}[b]{rcl}
+\begin{array}[b]{r@{\ }c@{\ }l}
 \llbracket {q}^* \rrbracket.\get  &=& (\llbracket q \rrbracket.\get)^*\\
@@
 &
-\begin{array}[b]{rcl}
+\begin{array}[b]{r@{\ }c@{\ }l}
 \llbracket q_1 \cdot q_2 \rrbracket.\get  &=& \llbracket q_1 \rrbracket.\get
@@
 Q_3}{\mathit{(Swap)}} &
-\begin{array}[b]{rcl}
+\begin{array}[b]{r@{\ }c@{\ }l}
 \llbracket q \rrbracket.\get(s_1 \cdot s_2)  &=& \llbracket q_2
@@
 {q_1 \sep q_2: (Q_1 \sep Q_2)\Leftrightarrow (Q_3 \sep Q_4)}{\mathit{(Or)}} &
-\begin{array}[b]{rcl}
+\begin{array}[b]{r@{\ }c@{\ }l}
 \llbracket q_1 \sep q_2 \rrbracket.\get(s) =
@@
 $\ell_1 : K(Q_1) \Leftrightarrow K(Q_2)$ and $\ell_2 : K(Q_3) \Leftrightarrow
-K(Q_4)$ such that
-\begin{center}:
+K(Q_4)$ such that:
+\begin{center}
 \begin{tabular}{l l}
@@
 \section{Conclusion and Future Work}
+\label{concl}
+
 In this paper, we showed how to synthesize a class of bijective quotient lenses
@@
 
-\bcp{And future work!}  \saz{I'm lukewarm about including a description of
-future work.  Unless there are specific conjectures or open problems that
-are worth mentioning, I'm not sure that it's really that important.}
-\bcp{Agreed. But don't we have any ideas for how to extend or generalize or
-things to do next or additional questions to consider??  E.g.: 
-\begin{itemize}
-\item is it easy to extend all this to
-asymmetric and symmetric lenses?  (If so, we should say so!)
-\item What properties do the bijective lens combinators have to have in
-order to validate the normalization property?  (I.e., if we tried to extend
-Boomerang's bijective lens combinators with another one, what would happen?)
-\item Do we have more ideas for QRE primitives?
-\item Etc.
-\end{itemize}
-}
-\label{concl}
+\BCP{Wrote the rest---comments?}
+Looking ahead, we would like to investigate whether the techniques proposed
+here can be applied to synthesizing quotiented variants of richer classes of
+lenses, such as ``classic'' asymmetric lenses~\cite{FosterLenses} or
+symmetric lenses~\cite{symmetriclenses}.  The main technical issues are
+developing synthesis procedures for the unquotientied variants of a given
+class---we are currently attacking the case of symmetric lenses---and
+showing that quotiented lenses in this class can be normalized along the
+lines of Theorem~\ref{normal form}.
+
+Along a different dimension, we are experimenting with larger examples of
+quotiented bijective lenses.  So far, the three ``specialized'' QRE
+combinators (\kw{squash}, \kw{perm}, and \kw{collapse}) have proved
+sufficient, but we expect eventually to encounter examples that will lead us
+to consider adding further combinators to supplement these.
 

### 2018-03-15 papers/QuotientLenses/main.tex
change color to dkblue and respond to B's commments

@@
 \lstset{ language=Caml, basicstyle=\linespread{0.8}\upshape\sffamily,
-keywordstyle=\upshape\sffamily\color{dkred}, keepspaces=true,
+keywordstyle=\upshape\sffamily\color{dkblue}, keepspaces=true,
 framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=true,
@@
 emph={synth,collapse,perm,squash,normalize,using,ins,del,lens,let,get,put,rquot,lquot},
-emphstyle=\upshape\sffamily\color{dkred}, 
+emphstyle=\upshape\sffamily\color{dkblue}, 
 columns=fullflexible,
@@
 \newcommand{\codefont}[1]{\ensuremath{\mathsf{#1}}}
-\newcommand{\kw}[1]{\textcolor{dkred}{\ensuremath{\mathsf{#1}}}}
+\newcommand{\kw}[1]{\textcolor{dkblue}{\ensuremath{\mathsf{#1}}}}
 \newcommand{\collapse}[2]{\ensuremath{\kw{collapse} \; #1 \mapsto #2}}
@@
   \end{itemize}
-}
+}\saz{I agree that the coloring can be a bit distracting, but I do think that
+  highlighting the keywords in the code helps for parsing--maybe using dark blue
+  is better? Why do ``get'' and ``put'' deserve special treatment as keywords,
+  compared to, say, ``synth'' or ``del''?}
 \begin{equation}\label{quotientlenslaws}
@@
   big improvement.  One tiny possible tweak: color the \kw{get} and \kw{put}
-  keywords in the same dark red as in the text.}\bcp{Can do, if we keep the
-  color in the text :-).  How are we going to decide about that?}
+  keywords in the same color as in the text.}\bcp{Can do, if we keep the
+  color in the text :-).  How are we going to decide about that?}\saz{Vote?}
 \label{fig:attheedges}
@@
 
-\BCP{Wrote the rest---comments?}
+\BCP{Wrote the rest---comments?}\saz{This seems OK---maybe swap the two
+  paragraphs.  The second one sounds more incremental and less exciting than the
+  first, so it's maybe not as good of a note to end on?}
 Looking ahead, we would like to investigate whether the techniques proposed

### 2018-03-15 papers/QuotientLenses/main.tex
slight reworking of the first couple paragraphs

@@
 canonizers}, $\ell$ is a bijective transformation, and $;$ is
-composition\iflastminute\bcp{$c_2^{-1}$ is ugly, but I think we need the
-superscript for accuracy.  Perhaps calling them $c$ and $d$ instead of $c_1$
-and $c_2$ would be better?}\fi.  Intuitively, the canonizers normalize data by
+composition.  Intuitively, the canonizers normalize data by
 picking a canonical representative of the relevant equivalence relations. This
@@
 Programmers often need to write programs that bidirectionally convert data
-between two different formats.\bcp{It's a shame to lose the concrete
-  examples here.} A time-tested and appealing way to do this is
-to use \emph{lenses}~\cite{Lenses}, programs that simultaneously define pairs of
+between two different formats. For example, bibliographic data may be stored in
+\bibtex{} or EndNote formats, spreadsheet data can be represented as either CSV
+and TSV files, and web APIs can return data as either JSON or XML objects.  A
+time-tested and appealing way to implement such conversions is to use
+\emph{lenses}~\cite{Lenses}, programs that simultaneously define pairs of
 functions for translating the data from the source format to the target format
-(the \emph{get} direction) and back again (the \emph{put} direction). The
-main reason that lenses are appealing is because lens programming languages
-guarantee that the \kw{put} and \kw{get} functions of a lens satisfy
-certain \emph{lens laws} constraining how the two functions interact with each
-other.
+(the \emph{get} direction) and back again (the \emph{put} direction).  Lens
+programming languages guarantee that these \kw{put} and \kw{get} functions
+satisfy certain \emph{lens laws}, which imply that round trips, from source to
+target and back again, behave properly.
 
@@
 \end{equation}
-Optician makes programming bijective lenses faster and easier, and though
-many useful lenses are not bijective in nature, they do have a bijective core
-surrounded by structured transformations. Hence, Optician can be used as a
-component in a system that synthesizes more complex lenses.
+Optician makes programming bijective lenses easier.  However, many useful lenses
+are not strictly bijective in nature---the desired transformation might ignore
+whitespace or the exact ordering of data fields, for instance---and it is a
+shame that Optician's synthesis algorithm cannot be directly applied in such
+cases.  One observation, however, is that non-bijective transformations can
+often be structured as a bijective ``core'' surrounded by some kind of data
+normalization at the edges. We can therefore hope to use the Optician algorithm
+as component in a system that synthesizes more complex lenses.
 

### 2018-03-15 papers/QuotientLenses/main.tex
tweaks to the intro

@@
 For example, consider the following QRE for writing author
-names:\bcp{Solomon: Someone noted on their copy of the intro from
-  class yesterday that this example is backwards.  Did you go over the
-  marked-up copies that I gave you?}\sam{I'm not sure what this comment means
-  and also I forgot those notes in the office. I'll go through them more
-  thoroughly tomorrow morning.}\bcp{What the comment means is that names are
-not usually formatted as ``Solomon, Maina'' but rather as ``Maina, Solomon''.}
+names:
 %
@@
 let wsp_sp = collapse wsp $\to$ " "
-let comma_name = first_name . "," . wsp_sp . last_name
+let comma_name = last_name . "," . wsp_sp . first_name
 \end{lstlisting}
@@
 \lstinline{wsp}, but with a single canonical representative: \lstinline{" "}.
-This is then used to build up \lstinline{comma_name}, a QRE with an
+It is used as a component of \lstinline{comma_name}, a QRE with an
 underlying language of two names, separated by a command and whitespace, where

### 2018-03-15 papers/QuotientLenses/main.tex
sections 2 and 3 (and part of 4)

@@
     \centering
-\begin{lstlisting}
+\begin{verbatim}
 @Book {Lovelace,
@@
 }
-\end{lstlisting}
+\end{verbatim}
   \end{minipage}
@@
     \centering
-\begin{lstlisting}
+\begin{verbatim}
 @Book{     Lovelace,
@@
 Author = "Lovelace, Ada",     }
-\end{lstlisting}
+\end{verbatim}
   \end{minipage}
@@
 mapping must preserve all the information, including the extraneous
-details, which leads to complex and brittle lenses complicated by the
-need to preserve unimportant details.  
+details, which leads to complex and brittle lenses.
 A better approach is to identify records that differ only in the
@@
 The permutation QRE combinator, \kw{perm}, allows data to be unordered. For example,
-the following instance of \kw{perm} allows label, author and title fields
+the following instance of \kw{perm} allows label, author, and title fields
 (which we assume have been defined earlier) to appear in any order.
@@
 \end{figure}
-Every QRE lens $q_1$ has a type $q_1: Q_1 \Leftrightarrow Q_2$ where $Q_1$ and $Q_2$
-are QREs. In the {\em get} direction, a QRE lens $q_1: Q_1 \Leftrightarrow Q_2$ uses the
-source QRE $Q_1$ to compute a canonical representative for the data modulo the
-equivalence relation defined by $Q_1$ and then applies the \kw{get} function of a
-bijective string lens $\ell_1$ to this representative. In the {\em put} direction, $q_1$
-operates similarly, but using the QRE $Q_2$ and the $\lput$ function of
-$\ell_1$.
-
-Because the QREs $Q_1$ and $Q_2$ determine the internal formats for data after
+Every QRE lens $q$ has a type $P \Leftrightarrow Q$ where $P$ and $Q$
+are QREs. In the {\em get} direction, a QRE lens $q: P \Leftrightarrow Q$ uses the
+source QRE $P$ to compute a canonical representative for the data modulo the
+equivalence relation defined by $P$ and then applies the \kw{get} function of a
+bijective string lens $\ell$ to this representative. In the {\em put} direction, $q$
+operates similarly, but using the QRE $Q$ and the $\lput$ function of
+$\ell$.
+
+Because the QREs $P$ and $Q$ determine the internal formats for data after
 canonization, and because the algorithm for synthesizing bijective string lenses is
-directed by these formats, $Q_1$ and $Q_2$ are all that is required to
+directed by these formats, $P$ and $Q$ are all that is required to
 synthesize QRE lenses end-to-end.
 
-However, our restriction that canonizers appear only at the edges
+However, our requirement that canonizers appear only at the edges
 raises a key technical question: Are we limiting the expressiveness of
 our transformations by demanding all programs fit into this normal
-form?  It turns out that we are not: any lens that uses canonizers
+form?  It turns out that we are not---any lens that uses canonizers
 internally can be transformed into a lens that uses canonizers only at
@@
 
-The remaining QRE combinators, which we discuss next, provide simpler,
-more compositional ways of building canonizers.  Nevertheless,
-the \kw{normalize} combinator provides a useful guide in the design of
-these combinators because it gives a sufficient condition for the
-well-formedness of any potential QREs.
+The remaining QRE combinators, which we discuss next, provide simpler, more
+compositional ways of building canonizers that meet these requirements by
+construction.  Nevertheless, the \kw{normalize} combinator provides a useful
+guide in the design of these combinators because it gives a sufficient condition
+for the well-formedness of any potential QREs.
 

### 2018-03-15 papers/QuotientLenses/main.tex
more polishing

@@
 because it ensures that the canonizing function of a QRE is well-defined.
-For example, consider the QRE \lstinline{"a"* . (collapse "a"$^*$ $\to$ "a")}. The
-behaviour of this QRE is not well-defined since the string
+For example, consider the QRE \lstinline{"a"$^*$ . (collapse "a"$^*$ $\to$ "a")}. The
+behavior of this QRE is not well-defined since the string
 \cd{"aaa"} can be canonized to any of \cd{"a"}, \cd{"aa"}, or \cd{"aaa"}
@@
 relations. Also, the condition that $r \equiv_R r'$ implies $q.\get(r) =
-q.\get(r')$ ensures that the {\em get} function induced on the equivalence
-classes if $\equiv_c$ is well-defined, and similarly for the {\em put}
+q.\get(r')$ ensures that the \kw{get} function induced on the equivalence
+classes of $\equiv_R$ is well-defined, and similarly for the \kw{put}
 function.
@@
 interpreter~\cite{boomerang,quotientlenses}. We have extended the existing
-Optician tool to synthesize QRE lenses from QREs. We will use word
-\emph{\QOpt} to denote our extended version of Optician, and Optician to denote
+Optician tool to synthesize QRE lenses from QREs. We will use 
+``\QOpt'' to denote our extended version of Optician, and just plain ``Optician'' to denote
 the pre-QRE version of Optician. The synthesis algorithm produces Boomerang lens
-values, so Boomerang treats synthesized lenses the same as hand-written ones. To
+values, so Boomerang gives synthesized lenses the same first-class status as hand-written ones. To
 evaluate the effectiveness of QREs, QRE lenses, and QRE lens synthesis, we
@@
 
-\BCP{Wrote the rest---comments?}\saz{This seems OK---maybe swap the two
-  paragraphs.  The second one sounds more incremental and less exciting than the
-  first, so it's maybe not as good of a note to end on?}
-Looking ahead, we would like to investigate whether the techniques proposed
-here can be applied to synthesizing quotiented variants of richer classes of
-lenses, such as ``classic'' asymmetric lenses~\cite{FosterLenses} or
-symmetric lenses~\cite{symmetriclenses}.  The main technical issues are
-developing synthesis procedures for the unquotientied variants of a given
-class---we are currently attacking the case of symmetric lenses---and
-showing that quotiented lenses in this class can be normalized along the
-lines of Theorem~\ref{normal form}.
-
-Along a different dimension, we are experimenting with larger examples of
+\BCP{Wrote the rest---comments?}\saz{I swapped the last two paragraphs to end on
+  a more exciting note.}
+Looking ahead, we are experimenting with larger examples of
 quotiented bijective lenses.  So far, the three ``specialized'' QRE
@@
 to consider adding further combinators to supplement these.
+
+Along a different dimension, we would like to investigate whether the techniques
+proposed here can be applied to synthesizing quotiented variants of richer
+classes of lenses, such as ``classic'' asymmetric lenses~\cite{FosterLenses} or
+symmetric lenses~\cite{symmetriclenses}.  The main technical issues are
+developing synthesis procedures for the unquotientied variants of a given
+class---we are currently attacking the case of symmetric lenses---and showing
+that quotiented lenses in this class can be normalized along the lines of
+Theorem~\ref{normal form}.
 

### 2018-03-15 papers/QuotientLenses/main.tex
comment in the related work section
Polish abstract

@@
 that two XML files with the same fields but in different orders would be
-considered the same, allowing a single, simple lens to handle them both.
-
-Building on recent work that showed how to synthesize bijective lenses from high
-level specifications, we offer three innovations to synthesize quotient lenses.
-First, we introduce {\em quotient regular expressions} (QREs), that easily mark
-aspects of data formats that are inessential and hence give a high-level
-specification of regular languages with an equivalence relation defined on
-them. Next, we introduce {\em QRE lenses}, i.e., lenses mapping between QREs.
+considered the same, allowing a single, simple program to handle them both. 
+%
+Building on a recently published algorithm for synthesizing plain bijective
+lenses from high-level specifications, we show how to synthesize bijective
+{\em quotient} lenses in three steps.  First, we introduce {\em quotient regular
+  expressions} (QREs), annotated regular expressions that conveniently mark
+inessential aspects of string data formatsl; each QRE specifies,
+simulteneously, a regular language and an equivalence relation on it.
+Second, we introduce {\em QRE lenses}, i.e., lenses mapping between QREs.
 Our key technical result is a proof that every QRE lens can be transformed into
-one of the form $c ; \ell ; d^{-1}$, where $c$ and $d^{-1}$ are {\em
+one of the form $c ; \ell ; d^{-1}$, where $c$ and $d$ are {\em
 canonizers}, $\ell$ is a bijective transformation, and $;$ is
-composition.  Intuitively, the canonizers normalize data by
-picking a canonical representative of the relevant equivalence relations. This
-theorem means we can push all normalization to the outer edges of a
-bidirectional transformation, leaving in the middle a pure bijective
-transformation. Finally, we leverage this normalization theorem to {\em
+composition\bcp{This seems not quite accurate.  It makes it look like
+  canonizers are bidirectional things, but they aren't---at least, we
+don't talk about them in that way (I suppose we could define the ``other
+direction'' to be the identity, but we do not).  What we really mean are
+left quotient and right quotient (and then we don't need the $^{-1}$).
+Similarly, the meaning of 
+``;'' here is not sequential composition of either lenses or QREs, which are
+the only two meanings we give to this symbol in the paper.}.  
+% Intuitively,
+% these canonizers normalize data by 
+% picking a canonical representative from the relevant equivalence
+% relations. 
+% BCP: Don't think this adds much, and it gets between the theorem and its use.
+This theorem means we can push all normalization to the outer edges of a
+composite quotient lens, leaving in the middle a pure bijective
+transformation. Third, we leverage this normalization theorem to {\em
 synthesize} QRE lenses from a pair of QREs and example input-output pairs,
 reusing earlier work on synthesizing plain bijective lenses. We have
-implemented QREs and QRE lens synthesis in the bidirectional programming
-language Boomerang. We evaluate the effectiveness of our approach by using it
-to synthesize QRE lenses between various real-world data formats in the
+implemented QREs and QRE lens synthesis as an extension to the bidirectional programming
+language Boomerang. We evaluate the effectiveness of our approach by 
+synthesizing QRE lenses between various real-world data formats in the
 Optician benchmark suite.
@@
 
-The \kw{Formless} type is given by
+The \kw{Formless}\saz{Is this ``Formless'' ``Formlens'' or ``Formlet''? The code
+  below says ``Formlens'' but the earlier parts of the paragraph say ``Formlet''}type is given by
 \begin{center}

### 2018-03-15 papers/QuotientLenses/main.tex
Made edits to abstract and intro

@@
 lenses from high-level specifications, we show how to synthesize bijective
-{\em quotient} lenses in three steps.  First, we introduce {\em quotient regular
+quotient lenses in three steps. First, we introduce {\em quotient regular
   expressions} (QREs), annotated regular expressions that conveniently mark
-inessential aspects of string data formatsl; each QRE specifies,
+inessential aspects of string data formats; each QRE specifies,
 simulteneously, a regular language and an equivalence relation on it.
 Second, we introduce {\em QRE lenses}, i.e., lenses mapping between QREs.
-Our key technical result is a proof that every QRE lens can be transformed into
-one of the form $c ; \ell ; d^{-1}$, where $c$ and $d$ are {\em
-canonizers}, $\ell$ is a bijective transformation, and $;$ is
-composition\bcp{This seems not quite accurate.  It makes it look like
-  canonizers are bidirectional things, but they aren't---at least, we
-don't talk about them in that way (I suppose we could define the ``other
-direction'' to be the identity, but we do not).  What we really mean are
-left quotient and right quotient (and then we don't need the $^{-1}$).
-Similarly, the meaning of 
-``;'' here is not sequential composition of either lenses or QREs, which are
-the only two meanings we give to this symbol in the paper.}.  
-% Intuitively,
-% these canonizers normalize data by 
-% picking a canonical representative from the relevant equivalence
-% relations. 
-% BCP: Don't think this adds much, and it gets between the theorem and its use.
-This theorem means we can push all normalization to the outer edges of a
-composite quotient lens, leaving in the middle a pure bijective
-transformation. Third, we leverage this normalization theorem to {\em
-synthesize} QRE lenses from a pair of QREs and example input-output pairs,
-reusing earlier work on synthesizing plain bijective lenses. We have
-implemented QREs and QRE lens synthesis as an extension to the bidirectional programming
-language Boomerang. We evaluate the effectiveness of our approach by 
-synthesizing QRE lenses between various real-world data formats in the
-Optician benchmark suite.
-
+Our key technical result is a proof that every QRE lens can be transformed
+into a functionally equivalent lens that canonizes source and target data only
+once at the ``edges'' then uses a bijective lens to map between the respective
+canonical elements; no internal canonization occurs in a lens in this normal
+form. Third, we leverage this normalization theorem to {\em synthesize} QRE
+lenses from a pair of QREs and example input-output pairs, reusing earlier work
+on synthesizing plain bijective lenses. We have implemented QREs and QRE lens
+synthesis as an extension to the bidirectional programming language Boomerang.
+We evaluate the effectiveness of our approach by synthesizing QRE lenses
+between various real-world data formats in the Optician benchmark suite.
 \end{abstract}
@@
 
+% Our end goal is to synthesize bijective quotient lenses that map between QREs
+% via a bijective lens between the respective canonical formats of the QREs. This
+% idea is simple and natural, but begs the following question: do we give up
+% expressiveness if we restrict ourselves to this form? We will show that this
+% is not the case by doing the following. First, we will introduce {\em QRE
+% lenses}, which are bijective quotient lenses formed by ``lifting'' bijective
+% lenses to a bijective quotient lens type, then freely quotienting these lenses
+% by QREs or applying familiar regular and composition operators to them. Then,
+% we will prove a normal form theorem (Theorem~\ref{normal form})
+% that says that every QRE lens can be transformed into a functionally equivalent
+% lens that canonizes source and target data only once at the ``edges'' then uses
+% a bijective lens to map between the respective canonical elements; no internal
+% canonization occurs in a lens in this normal form. This normalization
+% property will enable us to (1) synthesize QRE lenses by extending the synthesis
+% algorithm used by Optician, and (2) prove that if there is a QRE lens that
+% satisfies the input specification, then this extended algorithm will return
+% such a lens.
+
 Our second main contribution in this paper is to introduce {\em QRE lenses},
-which are bijective quotient lenses that map between QREs via bijective lenses
-that map between the respective canonical formats defined by the source and
-target QREs. This idea is simple and natural, but begs the following question:
-do we give up expressiveness if we restrict ourselves to this form? Our main
-technical contribution to this question asserts that we do not---more
-specifically, we prove a normal form theorem (Theorem~\ref{normal form})
-that says that every QRE lens can be rewritten to be the composition of a
-source canonizer, a bijective lens, and a target canonizer. This normalization
-property will enable us to (1) synthesize QRE lenses by extending the synthesis
-algorithm used by Optician, and (2) prove that if there is a QRE lens that
-satisfies the input specification, then this extended algorithm will return
-such a lens.
+with the end goal of synthesizing QRE lenses that map between QREs via a
+synthesized bijective lens between the respective canonical formats. This idea
+is simple and natural but begs the following question: do we give up
+expressiveness if we restrict ourselves to this form? Our main technical
+contribution to this question asserts that we do not---more specifically, we
+prove a normal form theorem (Theorem~\ref{normal form}) that says that every
+QRE lens formed by freely composing QRE lenses, applying regular
+operators to them, and quotenting them with QREs can be rewritten to be the
+composition of a source canonizer, a bijective lens, and a target canonizer.
+This normalization property will enable us to (1) synthesize QRE lenses by
+extending the synthesis algorithm used by Optician, and (2) prove that if there
+is a QRE lens that satisfies the input specification, then this extended
+algorithm will return such a lens.
 
@@
   the main technical contribution of this paper.
-  \item We leverage this result to reduce the problem of {\em synthesizing} QRE
-  lenses to the problem of synthesizing bijective string lenses, which was
+  \item We leverage this result to reduce the problem of {\em synthesizing}
+  QRE lenses to the problem of synthesizing bijective string lenses, which was
   previously studied in past work~\cite{optician} (Section~\ref{synth}).
@@
 Indeed this idea is what motivates our definition of {\em QRE lenses}.
-Intuitively, our QRE lenses are bijective lenses with ``canonizers at the edges,''
-just like Boomerang's quotient lenses. 
+Intuitively, our QRE lenses are bijective lenses with ``canonizers at the
+edges''.
 Figure~\ref{fig:attheedges} depicts the architecture of QRE lenses.
@@
 Every QRE lens $q$ has a type $P \Leftrightarrow Q$ where $P$ and $Q$
-are QREs. In the {\em get} direction, a QRE lens $q: P \Leftrightarrow Q$ uses the
-source QRE $P$ to compute a canonical representative for the data modulo the
-equivalence relation defined by $P$ and then applies the \kw{get} function of a
-bijective string lens $\ell$ to this representative. In the {\em put} direction, $q$
-operates similarly, but using the QRE $Q$ and the $\lput$ function of
-$\ell$.
+are QREs. In the {\em get} direction, a QRE lens $q: P \Leftrightarrow Q$ uses
+the source QRE $P$ to compute a canonical representative for the data modulo
+the equivalence relation defined by $P$ and then applies the \kw{get} function
+of a bijective string lens $\ell$ to this representative. In the {\em put}
+direction, $q$ operates similarly, but using the QRE $Q$ and the $\lput$
+function of $\ell$.
 
 Because the QREs $P$ and $Q$ determine the internal formats for data after
-canonization, and because the algorithm for synthesizing bijective string lenses is
-directed by these formats, $P$ and $Q$ are all that is required to
+canonization, and because the algorithm for synthesizing bijective string
+lenses is directed by these formats, $P$ and $Q$ are all that is required to
 synthesize QRE lenses end-to-end.
@@
 
-The \kw{Formless}\saz{Is this ``Formless'' ``Formlens'' or ``Formlet''? The code
-  below says ``Formlens'' but the earlier parts of the paragraph say ``Formlet''}type is given by
+The Formlens type is given by
 \begin{center}
@@
 
-\BCP{Wrote the rest---comments?}\saz{I swapped the last two paragraphs to end on
-  a more exciting note.}
 Looking ahead, we are experimenting with larger examples of
@@
 proposed here can be applied to synthesizing quotiented variants of richer
-classes of lenses, such as ``classic'' asymmetric lenses~\cite{FosterLenses} or
-symmetric lenses~\cite{symmetriclenses}.  The main technical issues are
+classes of lenses, such as ``classic'' asymmetric lenses~\cite{Lenses} or
+symmetric lenses~\cite{hofmann2011symmetric}.  The main technical issues are
 developing synthesis procedures for the unquotientied variants of a given

### 2018-03-16 papers/QuotientLenses/main.tex
Response about colors

@@
 \newif\iflastminute\lastminutetrue  % for things to check at the end
+\newif\iflater\laterfalse  % for things to think about after submission
 
@@
 % %%%% Macros Colors
-\definecolor{dkblue}{rgb}{0,0.1,0.5}
+\definecolor{dkblue}{rgb}{0,0.1,0.7}
 \definecolor{dkgreen}{rgb}{0,0.6,0}
@@
   compared to, say, ``synth'' or ``del''?}
+%
+\bcp{This gets at the basic difficulty with code coloring: we are dealing,
+  in different parts of the paper, with (1) Boomerang code and (2)
+  mathematical expressions involving lenses. We might want to make different
+  highlighting choices in the two cases (e.g., we might want to highlight
+  just individual keywords in large pieces of code, rather than the whole
+  thing, while lens expressions within larger math formulas should perhaps
+  be colored all the same, to better distinguish them from their context.
+  My suggestion for get and put was based on the fact that they are not
+  really part of boomerang programs: they are only mathematical entities (I
+  think you {\em can} extract them as separate functions in boomerang, but
+  usually there is no reason to, and we don't do it here).  However, I think
+  it doesn't matter that much and we should use our remaining time now for
+  polishing the text.  I propose we try dark blue for today and, if it's not
+  distracting, just leave it alone.}
+%
 \begin{equation}\label{quotientlenslaws}
@@
   normal form, with canonizers (specified by QREs) at the outer edges and an
-  ordinary bijective lens in the center. \BCP{Both the figure and the
-    caption are new---what do you think??}}\saz{I think that this figure is a
+  ordinary bijective lens in the center. }
+\iflater
+\saz{I think that this figure is a
   big improvement.  One tiny possible tweak: color the \kw{get} and \kw{put}
@@
   color in the text :-).  How are we going to decide about that?}\saz{Vote?}
+\fi
 \label{fig:attheedges}

### 2018-03-16 papers/QuotientLenses/main.tex
Tweaks

@@
 Our key technical result is a proof that every QRE lens can be transformed
-into a functionally equivalent lens that canonizes source and target data only
-once at the ``edges'' then uses a bijective lens to map between the respective
+into a functionally equivalent lens that canonizes source and target data just
+at the ``edges'' and that uses a bijective lens to map between the respective
 canonical elements; no internal canonization occurs in a lens in this normal

### 2018-05-07 papers/QuotientLenses/main.tex
Commented out some suggestions

@@
 holds modulo equivalence relations $\equiv_S$ and $\equiv_T$ defined on the source
-and target data respectively:\bcp{Not too convinced about the colors.  The
+and target data respectively:
+\iflater
+\bcp{Not too convinced about the colors.  The
   fact that lens expressions mix black and red is distracting -- my eyes
@@
 %
+\fi
 \begin{equation}\label{quotientlenslaws}

### 2018-05-21 papers/SymmetricLenses/main.tex
merge

\newif\ifdraft\drafttrue  % set true to show comments
%\newif\ifdraft\draftfalse  % set true to show comments
\newif\ifanon\anonfalse    % set true to suppress names, etc.
\newif\iffull\fullfalse   % set true for long version
\newif\ifappendices\appendicestrue

\PassOptionsToPackage{usenames,dvipsnames,svgnames,table}{xcolor}
\documentclass[acmsmall,screen,anonymous]{acmart}
\settopmatter{}

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{bussproofs}
\usepackage{varwidth}
\usepackage{amsthm}
\usepackage{csvsimple}
\usepackage{thmtools,thm-restate}
\usepackage{changepage}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage[inline]{enumitem}
\usepackage{multirow,bigdelim}
\usepackage{multicol}
\usepackage{siunitx}
\usepackage{listings}
\usepackage{sansmath}
\usepackage{url}
\usepackage{flushend}
\usepackage{microtype}
\usepackage[utf8]{inputenc}
\usepackage{mathpartir}
\usepackage{empheq}
\usepackage{array}
\usepackage{pgfplots}
\usepackage{stmaryrd}
\usepackage{courier}
\usepackage{qtree}
\usepackage[normalem]{ulem}
\usepackage{relsize}
\usepackage{tikz}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{tabularx}
\usepackage{stackengine}
\usepackage{caption}
\usepackage{wrapfig}
\usepackage{remreset}
\usepackage{tabulary}
\usepackage{xspace}

\usetikzlibrary{
  er,
  matrix,
  shapes,
  arrows,
  positioning,
  fit,
  calc,
  pgfplots.groupplots,
  arrows.meta
}
\tikzset{>={Latex}}

%%%% Hyperlinks – must come late!
%\usepackage[pdftex,%
%            pdfpagelabels,%
%            linkcolor=blue,%
%            citecolor=blue,%
%            filecolor=blue,%
%            urlcolor=blue]
%           {hyperref}

\input{macros}

\clubpenalty = 10000
\widowpenalty = 10000
\displaywidowpenalty = 10000

%\setlength{\belowcaptionskip}{-5pt}
%\setlength{\textfloatsep}{15pt}

% Creates a display mode for code in sans serif font
\lstnewenvironment{sflisting}[1][]
  {\lstset{%
    mathescape,
    basicstyle=\small\sffamily,
    aboveskip=5pt,
    belowskip=5pt,
    columns=flexible,
    frame=,
    xleftmargin=1em,#1}\sansmath}
  {}
% end

% Macros
  \newcommand{\NameOf}[1]{\CF{#1}}

%%% If you see 'ACMUNKNOWN' in the 'setcopyright' statement below,
%%% please first submit your publishing-rights agreement with ACM (follow link on submission page).
%%% Then please update our instructions page and copy-and-paste the NEW commands into your article.
%%% Please contact us in case of questions; allow up to 10 min for the system to propagate the information.
%%%
%%% The following is specific to POPL'18 and the paper
%%% 'Synthesizing Bijective Lenses'
%%% by Anders Miltner, Kathleen Fisher, Benjamin C. Pierce, David Walker, and Steve Zdancewic.
%%%

\setcopyright{rightsretained}
\acmPrice{}
\acmDOI{10.1145/3158089}
\acmYear{2018}
\copyrightyear{2018}
\acmJournal{PACMPL}
\acmVolume{2}
\acmNumber{POPL}
\acmArticle{1}
\acmMonth{1}
\startPage{1}

\bibliographystyle{ACM-Reference-Format}
\citestyle{acmauthoryear}

  \begin{document}

  %%% The following is specific to POPL'18 and the paper
%%% 'Synthesizing Bijective Lenses'
%%% by Anders Miltner, Kathleen Fisher, Benjamin C. Pierce, David Walker, and Steve Zdancewic.
%%%

%\toappear{}

%\conferenceinfo{POPL '16}{January 20--22, 2016, St. Petersburg, FL, USA} 
%\copyrightyear{2016} 
%\copyrightdata{978-1-nnnn-nnnn-n/yy/mm}

% Uncomment one of the following two, if you are not going for the 
% traditional copyright transfer agreement.

%\exclusivelicense                % ACM gets exclusive license to publish, 
                                  % you retain copyright

%\permissiontopublish             % ACM gets nonexclusive license to publish
                                  % (paid open-access papers, 
                                  % short abstracts)

%\titlebanner{DRAFT---do not distribute}        % These are ignored unless
%\preprintfooter{DRAFT---do not distribute}   % 'preprint' option specified.

\title{Simple Symmetric Lenses - Foundations and Synthesis}

\begin{abstract}
  Lenses are bidirectional transformers that are often used for synchronizing
  data in different formats. A generalization of lenses, symmetric lenses,
  enable synchronizations between formats where each format contains information
  the other lacks. The theory of symmetric lenses relies heavily on complements
  for maintaining synchronization. Despite the increased expressivity of
  symmetric lenses, no major bidirectional languages or tools have adopted
  symmetric lenses. While the use of complements allows for a clean and general
  formulation of symmetric lenses, it also requires maintaining internal state
  -- in addition to maintaining synchronized files, the complement must also be
  maintained. We simplify the formulation of symmetric lenses to \emph{simple
    symmetric lenses}. These lenses require no complement, instead using
  familiar $create$ and $put$ functions to maintain synchronization.
  Furthermore, we demonstrate the relationship between simple symmetric lenses,
  by carving out a subset of symmetric lenses which are equivalent to simple
  symmetric lenses. We integrate symmetric lenses into a modern bidirectional
  language, Boomerang, while maintaining full backward compatibility.

  Additionally, to enable easy use of symmetric lens combinators, we extend the
  Optician algorithm to synthesize simple symmetric lenses from type and example
  specifications. Unlike the bijective lenses Optician originally addresses,
  there are many symmetric lenses between two data formats. Hypothesizing that
  typically users want the lens that synchronizes the most data, we use an
  information-theoretic measure based on stochastic regular expressions, to
  quantify how much data is left non-synchronized, and develop an algorithm that
  attempts to minimize this measure. Furthermore, as our synthesis algorithm
  requires rewrites to our regular expressions, we prove that these the
  probability distribution of strings within our stochastic regular expressions
  are stable with respect to these rewrites. We experimentally demonstrate that
  despite addressing a much harder problem, our extension to Optician still
  maintains synthesis times competitive to Optician on Optician's bijective
  benchmark suite, and generalizes to benchmarks synchronizing data not in
  bijective correspondence.
\end{abstract}

\begin{CCSXML}
<ccs2012>
<concept>
<concept_id>10011007.10011006.10011050.10011017</concept_id>
<concept_desc>Software and its engineering~Domain specific languages</concept_desc>
<concept_significance>500</concept_significance>
</concept>
<concept>
<concept_id>10011007.10011006.10011066.10011070</concept_id>
<concept_desc>Software and its engineering~Application specific development environments</concept_desc>
<concept_significance>300</concept_significance>
</concept>
</ccs2012>
\end{CCSXML}

\ccsdesc[500]{Software and its engineering~Domain specific languages}
\ccsdesc[300]{Software and its engineering~Application specific development environments}

\ifanon
%\authorinfo{}
%           {}
%           {}
\maketitle
% \vspace*{-6cm}
\else
\author{Anders Miltner}
\affiliation{
  \institution{Princeton University}
  \country{USA}
}
\email{amiltner@cs.princeton.edu}

\author{Solomon Maina}
\affiliation{
  \institution{Princeton University}
  \country{USA}
}
\email{smaina@cis.upenn.edu}

\author{Kathleen Fisher}
\affiliation{
  \institution{Tufts University}
  \country{USA}
}
\email{kfisher@eecs.tufts.edu}

\author{Benjamin C. Pierce}
\affiliation{
  \institution{University of Pennsylvania}
  \country{USA}
}
\email{bcpierce@cis.upenn.edu}

\author{David Walker}
\affiliation{
  \institution{Princeton University}
  \country{USA}
}
\email{dpw@cs.princeton.edu}

\author{Steve Zdancewic}
\affiliation{
  \institution{University of Pennsylvania}
  \country{USA}
}
\email{stevez@cis.upenn.edu}

\keywords{Bidirectional Programming, Program Synthesis, Type-Directed Synthesis,
Type Systems}

\maketitle
\fi

% \category{D.3.1}
% {Programming Languages}
% {Formal Definitions and Theory}
% [Semantics]
\ifanon\else
\fi

% begin introduction
\section{Introduction}
% state of the world
Lenses, groups of three functions that satisfy invertibility laws, are useful
formulations for synchronization. Given data in one format, you can view it is a
different data format using \emph{get} function. Given data in another format,
you can turn it into data in the first format using a \emph{create} function.
Given edits on the second format, you can propagate those updates to the first
format using a \emph{put} function. However, writing these three functions is
tedious and error prone. Much better would be to write a single program that
expresses all three of the functions while also guaranteeing invertibility
properties about them.  This is the approach used in many bidirectional
languages: users write code in a bidirectional DSL, and all three functions are
maintained.  These bidirectional languages are often used for data
synchronization tasks - keeping data in different forms synchronized.

While lenses are used to synchronize data formats, but they remain incapable of
synchronizing formats where each format maintains information the other lacks.
To address this, lenses have been generalized to symmetric lenses.  Symmetric
lenses address these synchronization tasks by maintaining an explicit complement
to keep track of projected data.  However, despite the benefits of symmetric
lenses, they have not been adopted into bidirectional languages, remaining instead
as a rarely used Haskell package.

We believe this is largely due to complements. While complements provide a
simple, theoretical framework to express symmetric lenses, they become tricky
when used for synchronization tasks.  Critically, instead of maintaining two
pieces of data (one for each synchronized data format), the lens must now
maintain three (one for each synchronized data format, and one for the
complement).  This can then be difficult when migrating data - I can no longer
copy my formats to a new computer and tell a bidirectional language to keep them
synchronized, I instead must copy my formats and the complement to the new
computer.

We propose an alternative method. Instead of using complements, we can instead
use familiar \emph{create} and \emph{put} methods to synchronize the data
formats. In this paper, we present \emph{simple symmetric lenses}, a
reformulation of symmetric lenses consisting of 4 functions: 2 create functions
and 2 put functions. While these are less expressive than symmetric lenses,
simple symmetric lenses are able to express all synchronization tasks tackled by
asymmetric lenses - all asymmetric lenses are simple symmetric lenses.
Furthermore, we find a predicate over symmetric lenses, where symmetric lenses
satisfy the predicate if, and only if, they can be written as a simple symmetric lens.

As part of our work, we extend Optician to handle symmetric lens combinators,
while maintaining full backwards compatibility. This means simple symmetric
lenses compose simply with both the matching lens and quotient lens combinators
Boomerang provides. Furthermore, despite providing additional capabilities, our
implementation of Boomerang uses fewer lines of code in the core lens library
than were originally present - the symmetry of our lenses enable more code
reuse than the asymmetric versions permitted.

Furthermore, we enable easy use of these combinators by developing a synthesis
algorithm for symmetric lenses within Boomerang. We do this by extending
Optician, a tool for synthesizing bijective lenses, with capabilities for
synthesizing symmetric lenses as well. However, this is not trivial. Bijective
lenses permit relatively few well-typed terms: finding a bijective lens between
two data formats is not easy. However, finding a symmetric lens between two data
formats is relatively easy, as the data synchronization function can be a
``disconnect'' synchronizer -- changing the data in one format doesn't impact
the data in the other. Therefore, we need a means to find the intended symmetric
lens, not merely a symmetric lens. While input-output examples can help, we wish
to make our tool able to synthesize many functions without requiring users to
specify their intent with large example suites.

Instead, we assume that users want to synchronize as much data as possible. To
these ends, we develop an information thereoretic metric measuring the entropy
of how much data is left non-synchronized, utilizing \emph{stochastic regular
  expressions}. We syntactically find the entropy of stochastic regular
expressions for unambiguous stochastic regular expresions. Furthermore, just as
Optician applies rewrites to regular expressions to find well-aligned ones, our
enhancement to Optician applies rewrites to stochastic regular expressions to
find well-aligned ones. We prove these rewrites maintain both the language of
the stochastic regular expressions and the probability distributions over them.

Lastly, we demonstrate the effectiveness of our algorithm by synthesizing all
the bijective transformations present in Optician's benchmark suite with the
new, symmetric algorithm.  Furthermore, we synthesize an additional
\NumBenchmarks new, non-bijective benchmarks.

In summary, our contributions are as follows:
\begin{enumerate}
\item We define simple symmetric lenses and show their relationship with
  symmetric and asymmetric lenses (\S\ref{sec:ssl} and \S\ref{sec:ssl-rels}).
\item We introduce simple symmetric lens combinators that can be used for
  synchronizing string data formats (\S\ref{sec:ssl-string}).
\item We define stochastic regular expressions, demonstrate how to find their
  entropy, and show how certain rewrites on them maintain language and
  probability distribution (\S\ref{sec:stoch-regexp}).
\item We define symmetric DNF lenses, and stochastic DNF regular expressions.
  We show the conversion of regular expression to DNF regular expression
  maintains language and probability distribution.  We show how to infer the
  entropy of a DNF regular expression (\S\ref{sec:dnf-regex-lens}).
\item We describe a synthesis algorithm that attempts to minimize unsynchronized
  information (\S\ref{sec:synthesis}).
\item We provide a fully backwards compatible implementation of symmetric lenses
  in Boomerang, and an implementation of our synthesis algorithm into Optician
  (\S\ref{sec:implementation}).
\item We evaluate our synthesis algorithm on a variety of benchmarks, and find
  that we lose very little synthesis time, while gaining a large amount of
  expressivity (\S\ref{sec:evaluation}).
\end{enumerate}

% end introduction

% begin preliminaries
\section{Preliminaries}
\label{sec:preliminaries}
A \emph{symmetric lens} between $X$ and $Y$ is an element of the type
$\exists C.\{ \mathit{missing} :
C,\,get : X \times C \to Y \times C,\,put : Y \times C \to X \times C\}$
satisfying the following laws:
\begin{equation}
  \tag{\PutRL}
  \frac{\PutR(x,c) = (y,c')}{\PutL(y,c') = (x,c')}
\end{equation}
\begin{equation}
  \tag{\PutLR}
  \frac{\PutL(y,c) = (x,c')}{\PutR(x,c') = (y,c')}
\end{equation}

Given a lens $l \in X \leftrightarrow Y$, a \emph{put object} for $l$ is a
member of $X + Y$. Consider the function $apply$, which, given a lens and an
element of that lens's complement, is a function from sequences of put objects
to sequences of put objects. If $apply(\Lens,c,es) = es'$, then that means, starting
with complement $c$, after edit $es_i$, the lens $\Lens$ synchronizes the other
data to $es'_i$.
\[
  \inferrule
  {
  }
  {
    apply(\Lens,c,[]) = []
  }
\]
\[
  \inferrule
  {
    \Lens.putr(x,c) = (y,c')\\
    apply(\Lens,c',es) = es'
  }
  {
    apply(\Lens,c,(\InLOf{x})::es) = (\InROf{y})::es'
  }
\]
\[
  \inferrule
  {
    \Lens.putl(y,c) = (x,c')\\
    apply(\Lens,c',es) = es'
  }
  {
    apply(\Lens,c,(\InROf{y})::es) = (\InLOf{x})::es'
  }
\]

Two lenses, $\Lens_1$ and $\Lens_2$, are equivalent if
$apply(\Lens_1,\Lens_1.init,es) = apply(\Lens_2,\Lens_2.init,es)$ for all edit
sequences $es$.
% end preliminaries

\section{Simple Symmetric Lenses}
\label{sec:ssl}
Classically, asymmetric lenses comprised three functions, \Put{}, \Get{}, and
\Create{}.  We extend this style of lens to symmetric lenses with
\emph{simple symmetric lenses}.  Like the classical formulation of asymmetric
lenses, simple symmetric lenses have no complement.  Simple symmetric
lenses comprise 4 functions, \CreateR, \CreateL, \PutR, and \PutL.
These functions satisfy the following round-tripping laws:
\begin{equation}
  \tag{\CreatePutRL}
  \PutLOf{(\CreateROf{x})}{x} = x
\end{equation}
\begin{equation}
  \tag{\CreatePutLR}
  \PutROf{(\CreateLOf{y})}{y} = y
\end{equation}
\begin{equation}
  \tag{\PutRL}
  \PutLOf{(\PutROf{x}{y})}{x} = x
\end{equation}
\begin{equation}
  \tag{\PutLR}
  \PutROf{(\PutLOf{y}{x})}{y} = y
\end{equation}

Two simple symmetric lenses are equivalent if the four functions they comprise
are equivalent as functions. However, to formalize how these lenses work as
synchronization tools, we define an $apply$ function on them, much like we
define an $apply$ on symmetric lenses.

The function $apply$, given a symmetric lens and an optional pair of elements of
the source and target, is a function from sequences of put objects to sequences
of put objects.  If $apply(\Lens,\None,es,es')$, then that means, starting with
no prior data, after edit $es_i$, the lens \Lens synchronizes the other data to
$es_i'$.  If $apply(\Lens,\SomeOf{(x,y)},es,es')$, then that means, starting with
data $x$ and $y$ on the left and right, respectively, after edit $es_i$, the
lens \Lens synchronizes the other data to $es_i'$.

\[
  \inferrule
  {
  }
  {
    apply(\Lens,xyo,[]) = []
  }
\]
\[
  \inferrule
  {
    \Lens.\CreateROf{x} = y\\
    apply(\Lens,\SomeOf{(x,y)},es) = es'
  }
  {
    apply(l,\None,\InLOf{x}::es) = \InROf{y}::es'
  }
\]
\[
  \inferrule
  {
    \Lens.\CreateLOf{y} = x\\
    apply(\Lens,\SomeOf{(x,y)},es) = es'
  }
  {
    apply(l,\None,\InROf{y}::es) = \InLOf{x}::es'
  }
\]
\[
  \inferrule
  {
    \Lens.\PutROf{x'}{y}  = y'\\
    apply(\Lens,\SomeOf{(x',y')},es) = es'
  }
  {
    apply(l,\SomeOf{(x,y)},\InLOf{x'}::es) = \InROf{y'}::es'
  }
\]
\[
  \inferrule
  {
    \Lens.\PutLOf{y'}{x}  = x'\\
    apply(\Lens,\SomeOf{(x',y')},es) = es'
  }
  {
    apply(l,\SomeOf{(x,y)},\InROf{y'}::es) = \InLOf{x'}::es'
  }
\]

Furthermore, equivalence of the $apply$ function corresponds exactly with
equivalence of lenses.

\begin{theorem}
  $\Lens_1 \equiv \Lens_2$ if, and only if, for all sequences $es$,
  $apply(\Lens_1,\None,es) = apply(\Lens_2,\None,es)$.
\end{theorem}

% begin relationship
\section{Relationships Between Simple Symmetric Lenses and Other Lenses}
\label{sec:ssl-rels}

We've reformulated symmetric lenses in a style similar to existing asymmetric
lens formulations. However, some symmetric lenses certainly are inexpressible in
the stateless symmetric lens formulation. We formalize which are expressible
with a restriction on symmetric lenses. We define \emph{forgetful symmetric
  lenses} to be symmetric lenses which satisfy the following additional laws:
\begin{equation}
  \tag{\ForgetfulRL}
  \begin{mathprooftree}
    \AxiomC{$\Lens.putr(x,c_1) = (\_,c_1')$}
    \def\extraVskip{.5pt}
    \noLine 
    \UnaryInfC{$\Lens.putr(x,c_2) = (\_,c_2')$}
    \AxiomC{$\Lens.putl(y,c_1') = (\_,c_1'')$}
    \def\extraVskip{.5pt}
    \noLine 
    \UnaryInfC{$\Lens.putl(y,c_2') = (\_,c_2'')$}
    \def\extraVskip{2pt}
    \singleLine
    \BinaryInfC{$c_1'' = c_2''$}
  \end{mathprooftree}
\end{equation}
\begin{equation}
  \tag{\ForgetfulLR}
  \begin{mathprooftree}
    \AxiomC{$\Lens.putl(y,c_1) = (\_,c_1')$}
    \def\extraVskip{.5pt}
    \noLine 
    \UnaryInfC{$\Lens.putl(y,c_2) = (\_,c_2')$}
    \AxiomC{$putr(x,c_1') = (\_,c_1'')$}
    \def\extraVskip{.5pt}
    \noLine 
    \UnaryInfC{$\Lens.putr(x,c_2') = (\_,c_2'')$}
    \def\extraVskip{2pt}
    \singleLine
    \BinaryInfC{$c_1'' = c_2''$}
  \end{mathprooftree}
\end{equation}

Intuitively, this says that the complements are uniquely determined by the most
recently input $x$ and $y$. This corresponds closely with simple symmetric
lenses, where all state is maintained by the current state of the $x$ and $y$
data. Forgetful symmetric lenses express exactly the same $apply$ function as
stateless symmetric lenses.

\begin{theorem}
  Let $\Lens$ be a symmetric lens. The lens $\Lens$ is equivalent to a forgetful
  lens if, and only if, there exists a simple symmetric lens $\Lens'$ where
  $apply(\Lens,\Lens.init,es) = apply(\Lens',\None,es)$, for all put sequences $es$.
\end{theorem}

Despite being less expressive than symmetric lenses, simple symmetric lenses are
strictly more expressive than classical asymmetric lenses.

\begin{theorem}
  Let $\Lens$ be an asymmetric lens. $\Lens$ is also a simple symmetric lens,
  where
  \begin{enumerate}
  \item $\Lens.creater \App x = \Lens.get \App x$
  \item $\Lens.createl \App y = \Lens.create \App y$
  \item $\Lens.putr \App x \App y = \Lens.get \App x$
  \item $\Lens.putl \App y \App x = \Lens.put \App y \App x$
  \end{enumerate}
\end{theorem}
% end relationship

% begin simple symmetric string lenses
\section{Simple Symmetric String Lenses}
\label{sec:ssl-string}
While we have a theory for simple symmetric lenses, to integrate with Boomerang,
we need combinators to express these lenses. While Boomerang has additional,
complex combinators, we focus on the core ones, which are also those we
eventually wish to synthesize.

\subsection{Identity}
The identity lens operates on a single regular expression.  It fully
synchronizes the data within that regular expression: if data gets changed in
the left, that change gets reflected in the right.

\subsection{Disconnect}
The disconnect lens performs the opposite function as the identity lens.  It
leave the two data formats provided separate.  If data gets changed on the left,
the right is left unaffected.  When the data is being newly added, default
behavior comes from one side or the other.

\subsection{Concat}
The concat lens combines two existing lenses, working on their concatenated
regular expressions.  It performs the left lens on the left portion of the
string, and the right lens on the right portion.  It requires unambiguity
constraints to break up the left and right portions.

\subsection{Swap}
The swap lens operates like the concat lens, but when going from one side to the
other, it swaps the order of the lenses.

\subsection{Compose}
The swap lens operates like the concat lens, but when going from one side to the
other, it swaps the order of the lenses.

\subsection{Iterate}
The iterate lens breaks the provided strings into chunks, and performs the
provided lens on those chunks, then concatenates them together.

\subsection{Or}
The or lens performs one lens or another, depending on the inputs to the lens.

\subsection{MergeL}
The mergel lens is used when the same piece of data on the left is synchronized
to data in potentially many different forms on the right. The first provided
lens is the one defaulted to on creates.

\subsection{MergeR}
The merger lens is symmetric to the mergel lens.

\subsection{Compose}
The compose lens applies one lens to go to an intermediary format, then applies
the other to go to the final format.
% end simple symmetric string lenses

% begin stochastic regular expressions
\section{Stochastic Regular Expressions}
\label{sec:stoch-regexp}
To define the entropy for data recovery, we must first develop a probability
model for our languages. We do this by enhancing regular expressions into
\emph{stochastic regular expressions}. With this, we can jointly
express both a language, and a probability distribution over that language.

\begin{center}
  \begin{tabular}{rcl}
    \Regex{},\RegexAlt{}
    & \GEq{} & $\String$ \\
    & \GBar{} & $\emptyset$ \\
    & \GBar{} & $\PRegexStar{\Regex}{p}$ \\
    & \GBar{} & $\RegexConcat{\Regex_1}{\Regex_2}$ \\
    & \GBar{} & $\PRegexOr{\Regex_1}{\Regex_2}{p}$
  \end{tabular}
\end{center}

The language is defined like the regular expression without probability
annotations would be.  The probabilities are defined as follows:

\begin{center}
  \begin{tabular}{rcl}
    $\ProbabilityOf{\String}{\String'}$
    & =
    & $\begin{cases*}1 & if $\String = \String'$\\ 0 & otherwise\end{cases*}$ \\

    $\ProbabilityOf{\emptyset}{\String}$
    & =
    & $0$ \\

    $\ProbabilityOf{\RegexConcat{\Regex_1}{\Regex_2}}{\Character_1 \ldots \Character_n}$

    & =
    & $\Sigma_{\String = \String_1\String_2}\ProbabilityOf{\Regex_1}{\String_1}*\ProbabilityOf{\Regex_2}{\String_2}$ \\

    $\ProbabilityOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability}}{\String}$
    & =
    & $\Probability * \ProbabilityOf{\Regex_1}{\String} +
      (1-\Probability) * \ProbabilityOf{\Regex_2}{\String}$\\

    $\ProbabilityOf{\PRegexStar{\Regex}{\Probability}}{\String}$
    & =
    & $\Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\Regex}{\String_i}$\\
  \end{tabular}
\end{center}

Much like the synthesis algorithm for Optician, we need to apply rewrites to
handle equivalences on the provided stochastic regular expressions.  However, no
such equivalences have yet been defined.  We extend the \emph{star-semiring}
equivalences to stochastic regular expressions.

\begin{center}
  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{(1-\Probability_1)*\Probability_2})}{\Probability_1*\Probability_2} & \OrAssociativityRule{}  \\
    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability} & \UnrollstarLeftRule{} \\
    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{\Probability} & \UnrollstarRightRule{} 
  \end{tabular}
\end{center}
% \caption{Regular Expression Equivalences}
% \label{fig:regex-equivalence-rules}

\begin{theorem}
  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
\end{theorem}

% end stochastic regular expressions

% begin DNF regular expressions and lenses
\section{Stochastic DNF Regular Expressions}
\label{sec:dnf-regex-lens}
We extend stochastic regular expressions to stochastic DNF regular expressions.
\begin{center}
  \begin{tabular}{l@{\ }c@{\ }l@{\ }>{\itshape\/}r}
    % DNF_REGEX
    \Atom{},\AtomAlt{} & \GEq{} & \PRegexStar{\DNFRegex{}}{\Probability}
% & \StarAtomType{}
\\
    \Sequence{},\SequenceAlt{} & \GEq{} &
                                                       $\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n}$ 
%& \MultiConcatSequenceType{} 
\\
    \DNFRegex{},\DNFRegexAlt{} & \GEq{} & $\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}$ %& \MultiOrDNFRegexType{} 
  \end{tabular}
\end{center}

As before, DNF regular expressions are essentially regular expressions with all
concatenations fully distributed over all disjunctions.  As these DNF regular
expressions are \emph{stochastic}, they are annotated with probabilities to
express a probability distribution, in addition to a language.  This probability
distribution is defined below:

\begin{center}
  \begin{tabular}{rcl}
    $\ProbabilityOf{\PRegexStar{\DNFRegex}{\Probability}}{\String}$
    & =
    & $\Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\DNFRegex}{\String_i}$\\

    $\ProbabilityOf{\SequenceOf{\String_0 \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_n \SeqSep \String_n}}{\String'}$
    & =
    & $\Sigma_{\String' = \String_0\String_1'\ldots\String_n'\String_n}\Pi_{i = 1}^n\ProbabilityOf{\Atom_i}{\String_i'}$ \\

    $\ProbabilityOf{\DNFOf{(\Sequence_1,\Probability_1) \DNFSep \ldots \DNFSep (\Sequence_n,\Probability_n)}}{\String}$
    & =
    & $\Sigma_{i=1}^n\Probability_i * \ProbabilityOf{\Sequence_i}{\String}$\\
  \end{tabular}
\end{center}

\begin{figure}
  \raggedright
  $\ConcatSequence{} \OfType{} \ArrowTypeOf{\SequenceType{}}{\ArrowTypeOf{\SequenceType{}}{\SequenceType{}}}$\\
  $\ConcatSequenceOf{[\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n]}{[\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\StringAlt_m]}=
  [\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n\Concat\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\StringAlt_m]$\\

  \medskip

  $\ConcatDNF{} \OfType{} \ArrowTypeOf{\DNFRegexType{}}{\ArrowTypeOf{\DNFRegexType{}}{\DNFRegexType{}}}$\\
  $\ConcatDNFOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}{\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}=$\\
      $\DNFLeft (\ConcatSequenceOf{\Sequence_1}{\SequenceAlt_1},\Probability_1*\ProbabilityAlt_1)\DNFSep \ldots
      \DNFSep
      (\ConcatSequenceOf{\Sequence_1}{\SequenceAlt_m},\Probability_1*\ProbabilityAlt_m)\DNFSep
      \ldots$\\
      $\DNFSep
      (\ConcatSequenceOf{\Sequence_n}{\SequenceAlt_1},\Probability_n*\ProbabilityAlt_1)\DNFSep
      \ldots \DNFSep
      (\ConcatSequenceOf{\Sequence_n}{\SequenceAlt_m},\Probability_n * \ProbabilityAlt_m) \DNFRight$

  \medskip

  $\OrDNF{}_{\Probability} \OfType{}
  \ArrowTypeOf{\DNFRegexType{}}{\ArrowTypeOf{\DNFRegexType{}}{\DNFRegexType{}}
  }$ \\
  $\OrDNFOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}{\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}{\Probability} =$\\
  $\DNFOf{(\Sequence_1,\Probability_1*\Probability)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n*\Probability)\DNFSep(\SequenceAlt_1,\ProbabilityAlt_1*(1-\Probability))\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m*(1-\Probability))}$

  \medskip

  \AtomToDNF{} \OfType
  \ArrowTypeOf{\AtomType{}}{\DNFRegexType{}}\\
  $\AtomToDNFOf{\Atom} = \DNFOf{(\SequenceOf{\EmptyString \SeqSep \Atom \SeqSep
      \EmptyString},1)}$
  \caption{DNF Regular Expression Functions}
  \label{fig:dnf-regex-functions}
\end{figure}

We have a means of turning stochastic regular expressions into stochastic DNF
regular expressions.
The conversion algorithm itself, written $\ToDNFRegexOf{\Regex}$, is defined below.
\[
  \begin{array}{rcl}
    \ToDNFRegexOf{\String} & = & \DNFOf{\SequenceOf{\String}}\\
    \ToDNFRegexOf{\emptyset} & = & \DNFOf{}\\
    \ToDNFRegexOf{(\PRegexStar{\Regex}{\Probability})} & = & \AtomToDNFOf{\PRegexStar{(\ToDNFRegexOf{\Regex})}{\Probability}}\\
    \ToDNFRegexOf{(\RegexConcat{\Regex_1}{\Regex_2})} & = & \ToDNFRegexOf{\Regex_1} \ConcatDNF \ToDNFRegexOf{\Regex_2}\\
    \ToDNFRegexOf{(\PRegexOr{\Regex_1}{\Regex_2}{\Probability})} & = & \ToDNFRegexOf{\Regex_1} \OrDNF_{\Probability} \ToDNFRegexOf{\Regex_2}\\
  \end{array}
\]
We have also proven an algorithm, demonstrating this conversion respects
probability distributions.

\begin{theorem}
  $\ProbabilityOf{\Regex}{\String} = \ProbabilityOf{\ToDNFRegex \Regex}{\String}$
\end{theorem}

Furthermore, in the situation where a DNF regular expression is unambiguous, we
have defined an efficient, syntactic means to find the entropy of the DNF
regular expression.

\begin{center}
  \begin{tabular}{rcl}
    $\EntropyOf{\PRegexStar{\DNFRegex}{\Probability}}$
    & =
    & $\frac{\Probability}{1-\Probability}(\EntropyOf{\DNFRegex} - \Log_2\Probability)
      - \Log_2(1-\Probability)
      $\\

    $\EntropyOf{\SequenceOf{\String_0 \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_n \SeqSep \String_n}}{\String'}$
    & =
    & $\Sigma_{i = 1}^n\EntropyOf{\Atom_i}$ \\

    $\EntropyOf{\DNFOf{(\Sequence_1,\Probability_1) \DNFSep \ldots \DNFSep (\Sequence_n,\Probability_n)}}$
    & =
    & $\Sigma_{i=1}^n\Probability_i(\EntropyOf{\Sequence_i}+\Log_2\Probability_i)$\\
  \end{tabular}
\end{center}

\begin{theorem}
  $\EntropyOf{\DNFRegex}$ is the entropy of $P_{\DNFRegex}$.
\end{theorem}

% end DNF regular expressions and lenses

% begin symmetric dnf lenses
\section{Symmetric DNF Lenses}
\label{symmetric-dnf-lenses}
\subsection{Syntax}
Symmetric DNF Lenses are what we synthesize, and we then turn them into
Symmetric Lenses. The nonterminals $i$, $j$, $p$, $q$, $r$, $c$, and $d$ all
represent natural numbers. The nonterminals $\String$ and $\StringAlt$ represent
strings.

\begin{center}
  \begin{tabular}{@{}r@{\ }c@{}l@{}}
    % REGEX
    \SAtomLens{} & \GEq{} & $\IterateLensOf{\SDNFLens}$ \\
    \SSQLens{} & \GEq{} & $(\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
                          \ldots\SeqLSep
                          (i_p,j_p,\SAtomLens_p)}
                          ,\ListOf{(\String_1,\Atom_1);\ldots;(\String_q,\Atom_q)}
                          ,\ListOf{\StringAlt_1;\ldots;\StringAlt_r})$ \\
    \SDNFLens{} & \GEq{} & $(\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
                           \ldots\DNFLSep
                           (i_p,j_p,\SSQLens_p)}
                           ,\ListOf{c_1;\ldots;c_q}
                           ,\ListOf{d_1;\ldots;d_r})$ \\
  \end{tabular}
\end{center}

Lets go through these from bottom to top.

A DNF Lens \SDNFLens{} consists of three components. First is the mapping
component. The mapping component is a list of triples $(i,j,\SSQLens)$. In this,
$i$ and $j$ refer to sequence $i$ and sequence $j$ in the left and right DNF
regular expressions, respectively. The third component, \SSQLens{}, is a
sequence lens that maps between the two sequences specified by $i$ and $j$. The
second component is a list of integers $c$. The integer, $c_i$ at position $i$
states that when performing a $\CreateR$ on a string matching sequence
$\Sequence_i$, it will use the sequence lens that maps it to sequence
$\SequenceAlt_{c_i}$ on the right. The third component is a list of integers
$d$, and follows the same pattern as the second component, but maps from right
to left.

A Sequence lens \SSQLens{} consists of three components. First is the mapping
component. The mapping component is a list of triples $(i,j,\SAtomLens)$. In
this, $i$ and $j$ refer to atom $i$ and atom $j$ on the left and right
sequences, respectively. The third component, \SAtomLens{}, is an atom lens that
maps between the two atoms specified by $i$ and $j$. The second component is a
list of strings, $\String$. The string, $s_i$ at position $i$ states that when
performing a $\CreateL$ on a string matching atom $\Atom_i$, it will do one of
two things. If $\Atom_i$ is mapped to in the mapping component, then performing
the atom's $\CreateL$ will actually $\PutL$ into that string. If $\Atom_i$ is
not mapped to, then $\String_i$ will be created as a representative of
$\Atom_i$.

An atom lens \AtomLens{} is merely the iteration of a symmetric DNF Lens.

\subsection{Typing and Semantics}
Because these lenses only encode functions on well-typed programs, we introduce
the semantics alongside the typing rules.  The semantics are over typing
derivations, instead of over individual lenses.

\paragraph*{DNF Lenses}
The typing judgment is a 3-ary relation over a single DNF lens, and two DNF
regular expressions. If $\SDNFLens \OfType \DNFRegex \Leftrightarrow
\DNFRegexAlt$, then the $\SDNFLens.\CreateR$, $\SDNFLens.\CreateL$,
$\SDNFLens.\PutR$, and $\SDNFLens.\PutL$ functions form a symmetric lens.

The typing judgement has 3 components. The first is the sublens components,
confirming that the sequence lenses the DNF lens is comprised of are all
well-typed. The second guarantees that if each sequence on the left has a
sequence lens that can be used for \CreateR{}s and \PutR{}s. The last guarantees
the same for sequences on the right, with \CreateL{} and \PutL{}.

\[
  \inferrule*
  {
    \SSQLens_1 \OfType \Sequence_{i_1} \Leftrightarrow \SequenceAlt_{j_1}\\
    \ldots\\
    \SSQLens_p \OfType \Sequence_{i_p} \Leftrightarrow \SequenceAlt_{j_p}\\\\
    i_{c_1} = 1\\
    \ldots\\
    i_{c_q} = q\\\\
    j_{d_1} = 1\\
    \ldots\\
    j_{d_r} = r
  }
  {
    (\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
      \ldots\DNFLSep
      (i_p,j_p,\SSQLens_p)}
    ,\ListOf{c_1;\ldots;c_q}
    ,\ListOf{d_1;\ldots;d_r})
    \OfType\\
    \DNFOf{(\Sequence_1,\Probability_1) \DNFSep \ldots \DNFSep (\Sequence_q,\Probability_q)}
    \Leftrightarrow
    \DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1) \DNFSep \ldots \DNFSep (\SequenceAlt_r,\ProbabilityAlt_q)}
  }
\]

The \CreateR{} function looks for the sequence the provided string matches. If
the string matches sequence $\Sequence_x$, then the lens will look in the create
list at position $x$ to find which sequence lens to use. Then, the sequence lens
transforms the given string using that sequence lens.

The \PutR{} function finds what pairs sequences the input source and view
strings match.  If there that pair of sequences have a sequence lens between
them, then the DNF lens merely performs that sequence lens on the provided
strings.  If there isn't a pair of sequence lenses between them, then \CreateR{}
is performed on the source, with the view forgotten.

The \CreateL{} and \PutL{} functions are defined symmetrically.

\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
  $\CreateR{} \App s$ & = & $\SSQLens_{c_x}.\CreateR{} \App s$ if $s \in \Sequence_x$\\
  $\CreateL{} \App v$ & = & $\SSQLens_{d_y}.\CreateL{} \App v$ if $v \in \SequenceAlt_y$\\
  $\PutR{} \App s \App v$ & = &
                               $\begin{cases*}
                                 \SSQLens_x.\PutR{} \App s \App v & if $s \in \Sequence_{i_x}$ and $v \in \SequenceAlt_{j_x}$\\
                                 \CreateR{} \App s & if $\nexists x.$ $s \in \Sequence_{i_x}$ and $v \in \SequenceAlt_{j_x}$
                               \end{cases*}$\\
  $\PutL{} \App v \App s$ & = &
                               $\begin{cases*}
                                 \SSQLens_y.\PutL{} \App v \App s & if $v \in \SequenceAlt_{j_y}$ and $s \in \Sequence_{i_y}$\\
                                 \CreateL{} \App v & if $\nexists y.$ $v \in \SequenceAlt_{j_y}$ and $s \in \Sequence_{i_y}$
                               \end{cases*}$
\end{tabular}

\paragraph*{Sequence Lenses}
The typing judgment is a 3-ary relation over a single Sequence lens, and two
sequences. If $\SSQLens \OfType \Sequence \Leftrightarrow \SequenceAlt$, then
the $\SSQLens.\CreateR$, $\SSQLens.\CreateL$, $\SSQLens.\PutR$, and
$\SSQLens.\PutL$ functions form a symmetric lens.

The typing judgement has 4 components. The first is the sublens components,
confirming that the atom lenses the sequence lens is comprised of all are
well-typed. The second guarantees that if each string that will be used for
\CreateR{}s are members of the correct atoms.  The third guarantees
the same for strings and atoms on the right, with \CreateL.  The last guarantees
that each atom is mapped by at most one atom lens.

\[
  \inferrule*
  {
    \SAtomLens_1 \OfType \Atom_{i_1} \Leftrightarrow \AtomAlt_{j_1}\\
    \ldots\\
    \SAtomLens_p \OfType \Atom_{i_p} \Leftrightarrow \AtomAlt_{j_p}\\\\
    \String_1 \in \Atom_1\\
    \ldots\\
    \String_q \in \Atom_q\\\\
    \StringAlt_1 \in \AtomAlt_1\\
    \ldots\\
    \StringAlt_r \in \AtomAlt_r\\\\
    i_x = i_y \BooleanImplies x = y\\
    j_x = j_y \BooleanImplies x = y
  }
  {
    (\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
      \ldots\SeqLSep
      (i_p,j_p,\SAtomLens_p)}
    ,\ListOf{\String_1;\ldots;\String_q}
    ,\ListOf{\StringAlt_1;\ldots;\StringAlt_r})\\
    \OfType
    \SequenceOf{\String_0' \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_q
      \SeqSep \String_q'}
    \Leftrightarrow
    \SequenceOf{\StringAlt_0' \SeqSep \AtomAlt_1 \SeqSep \ldots \SeqSep \AtomAlt_r \SeqSep \StringAlt_r'}
  }
\]

For each component of the string matching an atom, the \CreateR function looks
for the atom lens that maps on the atom. If there is such an atom lens,
\SAtomLens, then that the sequence lens puts the provided string into the
default string for the target atom. If there is no such atom lens, then the
sequence lens merely uses the default string.

For each component of the string matching an atom, the \PutR function looks
for the atom lens that maps on the atom. If there is such an atom lens,
\SAtomLens, then that the atom lens puts the provided string of the source atom into the
string of the target atom. If there is no such atom lens, then the
sequence lens merely recovers the target's string.

The \CreateL{} and \PutL{} functions are defined symmetrically.

\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
  $\CreateR{} \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q'$
  & = 
  & $\StringAlt_0' \Concat \StringAlt_1'' \Concat \ldots \Concat
    \StringAlt_r'' \Concat \StringAlt_r'$\\
  & & where $\StringAlt_y'' =
    \begin{cases*}
      \SAtomLens_k.\PutRight \App \String_{i_k}'' \App \StringAlt_y & if $j_k = y$\\
      \StringAlt_y & if $\nexists k. j_k = y$\\
    \end{cases*}$\\
  $\CreateL{} \App \StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q''
  \Concat \StringAlt_q'$
  & = 
  & $\String_0' \Concat \String_1'' \Concat \ldots \Concat
    \String_r'' \Concat \String_r'$\\
  & & where $\String_x'' =
    \begin{cases*}
      \SAtomLens_k.\PutRight \App \StringAlt_{j_k}'' \App \String_x & if $i_k = x$\\
      \String_x & if $\nexists k. i_k = x$\\
    \end{cases*}$\\
  $\PutR{} \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q' \App \StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q'' \Concat \StringAlt_q'$
  & =
  & $\StringAlt_0'\Concat \StringAlt_1''' \Concat \ldots \Concat \StringAlt_q''' \Concat \StringAlt_q'$\\
  & & where
      $t_y''' =
      \begin{cases*}
        \SAtomLens_k.\PutRight \App \String_{i_k}'' \App \StringAlt_{y}'' & if $j_k = y$\\
        \StringAlt_y'' & if $\nexists k. j_k = y$\\
      \end{cases*}$\\
  $\PutL{} \App \StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q'' \Concat \StringAlt_q' \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q'$
  & =
  & $\String_0'\Concat \String_1''' \Concat \ldots \Concat \String_r''' \Concat \String_r'$\\
  & & where
      $s_x''' =
      \begin{cases*}
        \SAtomLens_k.\PutLeft \App \StringAlt_{j_k}'' \App \String_{x}'' & if $i_k = x$\\
        \String_x'' & if $\nexists k. i_k = x$\\
      \end{cases*}$\\
\end{tabular}

\paragraph*{Atom Lenses}
The typing judgment is a 3-ary relation over a single atom lens, and two
atoms. If $\SAtomLens \OfType \Atom \Leftrightarrow \AtomAlt$, then
the $\SAtomLens.\CreateR$, $\SAtomLens.\CreateL$, $\SAtomLens.\PutR$, and
$\SAtomLens.\PutL$ functions form a symmetric lens.

The typing judgement just confirms that the DNF lens that comprises the
sequence lens is also well typed.

\[
  \inferrule*
  {
    \SDNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt
  }
  {
    \IterateLensOf{\SDNFLens}
    \OfType \PRegexStar{\DNFRegex}{\Probability}
    \Leftrightarrow \PRegexStar{\DNFRegexAlt}{\ProbabilityAlt}
  }
\]

For each component of the string matching an atom, the \CreateR function looks
for the atom lens that maps on the atom. If there is such an atom lens,
\SAtomLens, then that the sequence lens puts the provided string into the
default string for the target atom. If there is no such atom lens, then the
sequence lens merely uses the default string.

For each component of the string matching an atom, the \PutR function looks
for the atom lens that maps on the atom. If there is such an atom lens,
\SAtomLens, then that the atom lens puts the provided string of the source atom into the
string of the target atom. If there is no such atom lens, then the
sequence lens merely recovers the target's string.

The \CreateL{} and \PutL{} functions are defined symmetrically.

\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
  $\CreateR{} \App \String_0 \Concat \ldots \Concat \String_n$
  & = 
  & $\StringAlt_1 \Concat \ldots \Concat \StringAlt_n$
  where $\StringAlt_i = \SDNFLens.\CreateR \App \String_i$\\
  $\CreateL{} \App \StringAlt_0 \Concat \ldots \Concat \StringAlt_m$
  & = 
  & $\String_1 \Concat \ldots \Concat \String_m$
  where $\String_i = \SDNFLens.\CreateL \App \StringAlt_i$\\
  $\PutR{} \App \String_0 \Concat \ldots \Concat \String_n \App
  \StringAlt_1 \Concat \ldots \Concat \StringAlt_m$
  & = 
  & $\StringAlt_1' \Concat \ldots \Concat \StringAlt_n'$
    where $\StringAlt_i' =
    \begin{cases*}
      \SDNFLens.\PutR \App \String_i \App \StringAlt_i & if $i \leq m$\\
      \SDNFLens.\CreateR \App \String_i & otherwise
    \end{cases*}$\\
  $\PutL{} \App \StringAlt_1 \Concat \ldots \Concat \StringAlt_m \App
  \String_0 \Concat \ldots \Concat \String_n$
  & = 
  & $\String_1' \Concat \ldots \Concat \String_m'$
    where $\String_i' =
    \begin{cases*}
      \SDNFLens.\PutL \App \StringAlt_i \App \String_i & if $i \leq n$\\
      \SDNFLens.\CreateL \App \StringAlt_i & otherwise
    \end{cases*}$\\
\end{tabular}

Now that we've developed symmetric DNF lenses, it's important to develop a means
to guage how bijective the lens is.

Instead of searching for \emph{any} term that satisfies the specification, we
search for the \emph{best} term that satisfies the specification. This begs the
question, by what metric do we use to determine which term is best. Intuitively,
we aim to synthesize the lens which is \emph{closest} to being a bijection.
Somewhat more formally, we aim to synthesize a lens which minimizes the expected
number of bits which must be used to recover the source from the target, and the
target from the source. The expected number of bits that are needed to encode
informatation from a data source is a well known concept in information theory,
and is known as the \emph{entropy} of the data source.

The entropy of $\DNFRegexAlt$ given $\DNFRegex$ and $\SDNFLens$,
$\EntropyOf{\DNFRegexAlt \Given \DNFRegex,\SDNFLens}$ is defined symmetrically.
With these $\operatorname*{argmin}_\theta$ terms defined, the ``best'' DNF lens
between $\DNFRegex$ and $\DNFRegexAlt$ is $\ArgminOver{\SetOf{\SDNFLens \SuchThat
    \SDNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt}}
\EntropyOf{\DNFRegex \Given \DNFRegexAlt,\SDNFLens} + \EntropyOf{\DNFRegexAlt
  \Given \DNFRegex,\SDNFLens}$.

\begin{centering}
  \[
    \begin{array}{rl}
      & \EntropyOf{\DNFOf{\Sequence_1 \DNFSep \ldots \DNFSep \Sequence_n}\\
      & \hspace*{1em}\Given
        \DNFOf{\SequenceAlt_1 \DNFSep \ldots \DNFSep \SequenceAlt_m}\\
      & \hspace*{1.21em},~(\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
        \ldots\DNFLSep
        (i_p,j_p,\SSQLens_p)}
        ,\ListOf{c_1;\ldots;c_q}
        ,\ListOf{d_1;\ldots;d_r})}\\
      =\\
      & \frac{\Sigma_{j=1}^m(\frac{\Sigma_{k\SuchThat j_k = j}
        \EntropyOf{\Sequence_{i_k}\Given\SequenceAlt_j,\SequenceLens_k}}{\SizeOf{\SetOf{k\SuchThat
        j_k = j}}} + log_2\SizeOf{\SetOf{k\SuchThat j_k = j}})}{m}\\\\
      \\

      & \EntropyOf{
        \SequenceOf{\String_0,\Atom_1,\ldots,\Atom_n,\String_n}\\
      & \hspace*{1em}\Given
        \SequenceOf{\StringAlt_0,\AtomAlt_1,\ldots,\AtomAlt_m,\StringAlt_m}\\
      & \hspace*{1.21em},~
        (\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
        \ldots\SeqLSep
        (i_p,j_p,\SAtomLens_p)}
        ,\ListOf{(k_1,\String_1);\ldots;(k_q,\String_q)}
        ,\ListOf{(l_1,\StringAlt_1);\ldots;(l_r,\StringAlt_r)})}\\
      =\\
      & \Sigma_{x=1}^p\EntropyOf{\Atom_{i_x} \Given \AtomAlt_{i_x},\AtomLens_x} +
        \Sigma_{x=1}^q\EntropyOf{\Atom_{k_x}}\\\\
      \\

      & \EntropyOf{\StarOf{\DNFRegex} \Given \StarOf{\DNFRegexAlt},\StarOf{\SDNFLens}}\\
      =\\
      & 4*\EntropyOf{\DNFRegex\Given\DNFRegexAlt,\SDNFLens}
    \end{array}
  \]
\end{centering}

% end symmetric dnf lenses

% begin synthesis
\section{Synthesis}
\label{sec:synthesis}
While examples can be used to tighten the
specification, a large number of well-constructed examples are required to
sufficiently constrain the search space.

Unfortunately, trying to find the best symmetric lens is a daunting task!
Unless we find a bijective lens (in which case the sum is zero), we can never be
sure if we will find a better lens than the one we found.
% end synthesis

% begin implementation
\section{Implementation}
\label{sec:implementation}
% end implementation

% begin evaluation
\section{Evaluation}
\label{sec:evaluation}
% end evaluation

% begin related-work
\section{Related Work}
\label{sec:related}
% end related-work

% begin conclusion
\section{Conclusion}
\label{sec:conc}
% end conclusion

% begin acknowledgements
\begin{acks}
\end{acks}
% end acknowledgements

\ifanon\else
\fi

% We recommend abbrvnat bibliography style.

% The bibliography should be embedded for final submission.

\bibliography{local,bcp}

\appendix

\ifappendices

\onecolumn

\section{Forgetful Symmetric Lenses}

\begin{property}[Starting Forgetfulness RL]
  \label{prop:forget-rl}
  Let $\Lens$ be a forgetful symmetric lens.  If $(x_1',c_1') = \Lens.\PutL \App
  (y,\Snd \App (\Lens.\PutR \App (x,c_1)))$, and
  $(x_2',c_2') = \Lens.\PutL \App
  (y,\Snd \App (\Lens.\PutR \App (x,c_2)))$, then $x_1' = x_2'$.
\end{property}
\begin{proof}
  By \ForgetfulRL, $c_1' = c_2'$. We know $(y,c_1') = \Lens.\PutR \App
  (x_1',c_1')$ and $(y,c_1') = \Lens.\PutR \App (x_2',c_1')$ by \PutLR. By
  \PutRL, we know $(x_1' = \Lens.\PutL \App (y,c_1'))$ and $(x_2' = \Lens.\PutL
  \App (y,c_1')))$. Therefore, by transitivity of equality, $x_1' = x_2'$.
\end{proof}

\begin{property}[Starting Forgetfulness LR]
  \label{prop:forget-lr}
  Let $\Lens$ be a forgetful symmetric lens.  If $(y_1',c_1') = \Lens.\PutR \App
  (x,\Snd \App (\Lens.\PutL \App (y,c_1)))$, and
  $(x_2',c_2') = \Lens.\PutL \App
  (x,\Snd \App (\Lens.\PutR \App (y,c_2)))$,
  then $x_1' = x_2'$.
\end{property}
\begin{proof}
  Symmetric to Starting Forgetfulness RL.
\end{proof}

\begin{definition}[S]
  Let $\Lens$ be a forgetful lens.

  Consider the following four functions $S(\Lens)$, that we wish to satisfy the
  simple symmetric lens laws.

  \begin{centering}
    \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
      $\CreateR{} \App s$
      & =
      & $\Fst \App (\Lens.\PutR{} \App (s,\Lens.init))$\\

      $\CreateL{} \App v$
      & =
      & $\Fst \App (\Lens.\PutL{} \App (v,\Lens.init))$\\

      $\PutR{} \App s \App v$
      & =
      & $\LetIn{(\_,c)}{\Lens.\PutL \App (v,\Lens.init)}$\\
      &
      & $\LetIn{(s',\_)}{\Lens.\PutR \App (s,c)}$\\
      &
      & $s'$\\

      $\PutL{} \App v \App s$
      & =
      & $\LetIn{(\_,c)}{\Lens.\PutR \App (s,\Lens.init)}$\\
      &
      & $\LetIn{(y',\_)}{\Lens.\PutL \App (v,c)}$\\
      &
      & $y'$\\
    \end{tabular}
  \end{centering}
\end{definition}

\begin{mylemma}
  If $\Lens$ is a forgetful symmetric lens, then $S(\Lens)$ is a simple symmetric
  lens.
\end{mylemma}
\begin{proof}

  \CreatePutRL{}:

  \begin{centering}
    \begin{tabular}{@{}r@{\ }c@{\ }l@{\ }l}
      $S(\Lens).\PutLOf{(S(\Lens).\CreateROf{x})}{x}$
      & =
      & $S(\Lens).\PutLOf{(\Fst \App (\Lens.\PutR{(x,\Lens.init)}))}{x}$
      & By unfolding definitions
      \\

      & =
      & $\LetIn{(\_,c)}{\Lens.\PutR \App (x,\Lens.init)}$\\
      &
      & $\LetIn{(y',\_)}{\Lens.\PutL \App (\Fst \App (\Lens.\PutR{(x,\Lens.init)}),c)}$\\
      &
      & $y'$
      & By unfolding definitions\\

      & =
      & $\LetIn{(y',\_)}{\Lens.\PutL \App (\Lens.\PutR \App (x,\Lens.init))}$\\
      &
      & $y'$
      & By tuple harmony \\

      & =
      & $x$
      & By \PutRL \\
    \end{tabular}
  \end{centering}

  \CreatePutLR{}:  Symmetric to \CreatePutRL{}

  \PutRL{}:

  \begin{centering}
    \begin{tabular}{@{}r@{\ }c@{\ }l@{\ }l}
      $S(\Lens).\PutLOf{(S(\Lens).\PutROf{x}{y})}{x}$
      & =
      & $\LetIn{(\_,c)}{\Lens.\PutL \App (y,\Lens.init)}$\\
      & & $\LetIn{(y',c')}{\Lens.\PutR \App (x,c)}$\\
      & & $S(\Lens).\PutLOf{y'}{x}$
      & By unfolding definitions
      \\

      & =
      & $\LetIn{(\_,c)}{\Lens.\PutL \App (y,\Lens.init)}$\\
      & & $\LetIn{(y',c')}{\Lens.\PutR \App (x,c)}$\\
      & & $\LetIn{(\_,c'')}{\Lens.\PutR \App (x,\Lens.init)}$\\
      & & $\LetIn{(x',c''')}{\Lens.\PutL \App (y',c'')}$\\
      & & $x'$
      & By unfolding definitions
    \end{tabular}
  \end{centering}

  At this point, we know from \PutRL that $(x,c') = \Lens.\PutL \App (y',c')$.
  By Property~\ref{prop:forget-rl}, this means that $x' = x$, as desired.

  \PutLR{}:  Symmetric to \PutRL{}
\end{proof}

\begin{definition}
  Fix a symmetric lens $\Lens$ beteween $X$ and $Y$. Consider the
  function, $\SingleApp_{\Lens} \OfType (X + Y) \times \Lens.C
  \rightarrow ((X + Y) \times \Lens.C)$, defined as:

  \begin{tabular}{@{}r@{\ }c@{\ }l@{\ }l}
    $\SingleApp_{\Lens}(\InLOf{x},c)$
    & =
    & $\LetIn{(y,c')}{\Lens.\PutRight \App (x,c)}$\\
    &
    & $(\InROf{y},c')$\\

    $\SingleApp_{\Lens}(\InROf{y},c)$
    & =
    & $\LetIn{(x,c')}{\Lens.\PutLeft \App (y,c)}$\\
    &
    & $(\InLOf{x},\SomeOf{(x,y)})$
  \end{tabular}
\end{definition}

\begin{definition}
  Fix a simple symmetric lens $\Lens$ beteween $X$ and $Y$. Consider the
  function, $\SingleApp_{\Lens} \OfType ((X + Y) \times \OptionOf{(X \times Y)})
  \rightarrow ((X + Y) \times \OptionOf{(X \times Y)})$, defined as:

  \begin{tabular}{@{}r@{\ }c@{\ }l@{\ }l}
    $\SingleApp_{\Lens}(\InLOf{x},\None)$
    & =
    & $\LetIn{y}{\Lens.\CreateROf{x}}$\\
    &
    & $(\InROf{y},\SomeOf{(x,y)})$\\

    $\SingleApp_{\Lens}(\InROf{y},\None)$
    & =
    & $\LetIn{x}{\Lens.\CreateLOf{y}}$\\
    &
    & $(\InLOf{x},\SomeOf{(x,y)})$\\

    $\SingleApp_{\Lens}(\InLOf{x'},\SomeOf{(x,y)})$
    & =
    & $\LetIn{y'}{\Lens.\PutROf{x'}{y}}$\\
    &
    & $(\InROf{y'},\SomeOf{(x',y')})$\\

    $\SingleApp_{\Lens}(\InROf{y'},\SomeOf{(x,y)})$
    & =
    & $\LetIn{x'}{\Lens.\PutLOf{y'}{x}}$\\
    &
    & $(\InROf{x'},\SomeOf{(x',y')})$\\
  \end{tabular}
\end{definition}

\begin{definition}
  Fix a symmetric lens \Lens over $X$ and $Y$. We define the relation $R_\Lens$
  over $\Lens.C$ and $\OptionOf{(X \times Y)}$ as the largest relation such
  that:
  \begin{enumerate}
  \item $R_\Lens(c,None) \BooleanImplies c = \Lens.init$
  \item $R_\Lens(c,Some (x,y)) \BooleanImplies
    \Lens.\PutRight \App (x,c) = (y,c) \BooleanAnd
    \Lens.\PutLeft  \App (y,c) = (x,c)$
  \end{enumerate}
\end{definition}

\begin{mylemma}
  \label{lem:s-equiv}
  Let $\Lens$ be a symmetric lens.  Let $c \in \Lens.C$ be a complement, and
  $xyo \in \OptionOf{(X \times Y)}$.  If $R_{\Lens}(c,xyo)$, then
  $apply(\Lens,c,es) = apply(S(\Lens),xyo,es)$.
\end{mylemma}
\begin{proof}
  By induction on the derivation the application of $apply(S(\Lens),xyo,es)$
  \begin{case}[empty list]
    So by the case, $apply(S(\Lens),xyo,[]) = []$. Furthermore,
    $apply(\Lens,c,[]) = []$, as desired.
  \end{case}
  \begin{case}[first edit is a create right]
    So by the case, $apply(S(l),\None,\InLOf{x}::es) = \InROf{y}::es'$, where
    $S(\Lens).\CreateROf{x} = y$ and $apply(S(\Lens),\SomeOf{(x,y)},es) = es'$.

    As $R_{\Lens}(None,c)$, $c = \Lens.init$. So, performing $apply$ on
    $\Lens.init$, we get $apply(\Lens,\Lens.init,(\InLOf{x})::es) =
    (\InROf{y'})::es''$ where $\Lens.putr(x,c) = (y',c')$ and $apply(\Lens,c',es)
    = es'$.

    So, by definition, $S(\Lens).\CreateROf{x} = \Fst \App (\Lens.\PutRight \App
    (x,\Lens.init))$, so $y = y'$.

    Furthermore, by \PutRL, $\Lens.\PutLeft \App (y,c') = (x,c')$, and by another
    application of \PutLR, $\Lens.\PutRight \App (x,c') = (y,c')$.  This means
    that $R_{\Lens}(c',Some (x,y))$.

    So, by induction assumption, $apply(\Lens,c',es) = apply(S(\Lens),xyo,es)$, so
    $es' = es''$.  This means, $apply(S(l),\None,\InLOf{x}::es) =
    \InROf{y}::es'$ and $apply(\Lens,\Lens.init,(\InLOf{x})::es) =
    (\InROf{y})::es'$, so they are equal, as desired.
  \end{case}
  \begin{case}[first edit is a create left]
    Symmetric to previous case
  \end{case}
  \begin{case}[first edit is a put right]
    So by the case, $apply(S(\Lens),\SomeOf{(x,y)},\InLOf{x'}::es) = \InROf{y'}::es'$, where
    $S(\Lens).\PutROf{x}{y} = y'$ and $apply(S(\Lens),\SomeOf{(x',y')},es) = es'$.

    Performing $apply$ on $c$, we get $apply(\Lens,c,(\InLOf{x'})::es) =
    (\InROf{y'})::es''$ where $\Lens.\PutRight \App (x',c) = (y',c')$ and
    $apply(\Lens,c',es) = es''$.

    So, by definition, $S(\Lens).\PutROf{x}{y} = \Fst \App (\Lens.\PutRight \App
    (x,c''))$, where $c'' = \Snd \App (\Lens.\PutLeft \App (y,\Lens.init))$.

    By assumption, $R_{\Lens}(c,\SomeOf{(x,y)})$, so $\Lens.\PutLeft \App (y,c)
    = (x,c)$.  So, by Property~\ref{prop:forget-lr}, we know $y' = y''$.

    Furthermore, by \PutRL, $\Lens.\PutLeft \App (y',c') = (x',c')$, and by another
    application of \PutLR, $\Lens.\PutRight \App (x',c') = (y',c')$.  This means
    that $R_{\Lens}(c',Some (x',y'))$.

    So, by induction assumption, $apply(\Lens,c',es) = apply(S(\Lens),xyo,es)$, so
    $es' = es''$.  This means, $apply(S(l),\SomeOf{(x,y)},\InLOf{x'}::es) =
    \InROf{y'}::es'$ and $apply(\Lens,c,(\InLOf{x'})::es) =
    (\InROf{y'})::es'$, so they are equal, as desired.
  \end{case}
  \begin{case}[first edit is a put left]
    Symmetric to previous case
  \end{case}
\end{proof}

\begin{definition}[F]
  Let $\Lens$ be a simple symmetric lens between $X$ and $Y$.

  Consider the following set $C$, distinguished element of that set, $init$, and
  pair of functions $\PutRight$ and $\PutLeft$, that we wish to satisfy the
  symmetric lens laws, and that we also wish to be forgetful.

  \begin{centering}
    \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
      $C$
      & =
      & $\OptionOf{(X \times Y)}$\\

      $init$
      & =
      & $\None$\\

      $\PutRight \App (x,c)$
      & =
      & $(y,\SomeOf{(x,y)})$ where $y = \begin{cases*}
        \Lens.\CreateROf{x} & if $c = \None$\\
        \Lens.\PutROf{x}{y'} & if $c = \SomeOf{(x',y')}$\\
        \end{cases*}$\\

      $\PutLeft \App (y,c)$
      & =
      & $(x,\SomeOf{(x,y)})$ where $x = \begin{cases*}
        \Lens.\CreateLOf{y} & if $c = \None$\\
        \Lens.\PutLOf{y}{x'} & if $c = \Some{(x',y')}$\\
        \end{cases*}$\\
    \end{tabular}
  \end{centering}
\end{definition}

\begin{mylemma}
  \label{lem:f-sym}
  If $\Lens$ is a simple symmetric lens, then $F(\Lens)$ is a symmetric lens.
\end{mylemma}
\begin{proof}
  \PutRL: There are two cases, $c = \None$, and $c = \SomeOf{(x,y)}$.

  \begin{case}[c = \None]
    Let $(y',c') = F(\Lens).\PutRight \App (x',\None)$. This means that $y' =
    \Lens.\CreateROf{x'}$, and $c' = \SomeOf{(x',y')}$.

    Now, consider $(x'',c'') = F(\Lens).\PutLeft \App (y',\SomeOf{(x',y')})$. By
    unfolding definitions, $x'' = \Lens.\PutLOf{y'}{x'}$. By \CreatePutRL, $x''
    = x'$, meaning $c'' = \SomeOf{(x',y')}$. This means $(x',c') =
    F(\Lens).\PutLeft \App (y',\SomeOf{(x',y')})$, as desired.
  \end{case}

  \begin{case}[c = \SomeOf{(x,y)}]
    Let $(y',c') = F(\Lens).\PutRight \App (x',\SomeOf{(x,y)})$. This means that $y' =
    \Lens.\PutROf{x'}{y}$, and $c' = \SomeOf{(x',y')}$.

    Now, consider $(x'',c'') = F(\Lens).\PutLeft \App (y',\SomeOf{(x',y')})$. By
    unfolding definitions, $x'' = \Lens.\PutLOf{y'}{x'}$. By \PutRL, $x''
    = x'$, meaning $c'' = \SomeOf{(x',y')}$. This means $(x',c') =
    F(\Lens).\PutLeft \App (y',\SomeOf{(x',y')})$, as desired.
  \end{case}

  The second requirement, \PutLR, is symmetric.
\end{proof}

\begin{mylemma}
  If $\Lens$ is a simple symmetric lens, then $F(\Lens)$ is a forgetful symmetric lens.
\end{mylemma}
\begin{proof}
  By Lemma~\ref{lem:f-sym}, we know $F(\Lens)$ is symmetric, so we merely need
  to show it is forgetful.  We will tackle merely \ForgetfulRL, as the proof for
  \ForgetfulLR is symmetric.

  Let $c_1$ and $c_2$ be two arbitrary complements, and $x$ and $y$ two
  arbitrary values of $X$ and $Y$, respectively.

  We know that $c_1' = \Snd \App (F(\Lens).\PutRight \App (x,c_1))$ and $c_2' =
  \Snd \App (F(\Lens).\PutRight \App (x,c_2))$.  Now, by inversion on
  $F(\Lens).\PutRight$, we know that both $c_1' = Some(x,y_1')$ and $c_2' =
  Some(x,y_2')$ for some values of $y_1$ and $y_2$ (though we don't actually
  care about the values of $y_1$ and $y_2$).

  Now by unfolding definitions we know, $\Snd \App (F(\Lens).\PutLeft \App
  (y,\SomeOf{(x,y_1')})) = \SomeOf(\Lens.\PutLeftOf{x}{y},y) = c_1''$.  Similarly, we
  know $\Snd \App (F(\Lens).\PutLeft \App
  (y,\SomeOf{(x,y_2')})) = \SomeOf(\Lens.\PutLeftOf{x}{y},y) = c_2''$, so $c_1''
  = c_2''$, as intended.
\end{proof}

\begin{mylemma}
\label{lem:f-equiv}
  If $\Lens$ be a simple symmetric lens, then $apply(\Lens,xyo,es) =
  apply(F(\Lens),xyo,es)$.
\end{mylemma}
\begin{proof}
  By induction on the derivation of $apply$ on $\Lens$!

  \begin{case}[empty list]
    So by the case, $apply(S(\Lens),xyo,[]) = []$. Furthermore,
    $apply(\Lens,xyo,[]) = []$, as desired.
  \end{case}

  \begin{case}[first edit is a create right]
    So by the case, $apply(\Lens,\None,\InLOf{x}::es) = \InROf{y}::es'$, where
    $\Lens.\CreateROf{x} = y$ and $apply(\Lens,\SomeOf{(x,y)},es) = es'$.

    Performing $apply$ on
    $\None$, we get $apply(F(\Lens),\None,(\InLOf{x})::es) =
    (\InROf{y'})::es''$ where $F(\Lens).putr(x,\None) = (y',c)$ and $apply(\Lens,c,es)
    = es'$.

    Unfolding definitions, $F(\Lens).putr(x,\None) = \Lens.\CreateROf{x} =
    (y,\SomeOf{(x,y)})$, so $c = \SomeOf{(x,y)}$ and $y = y'$

    So, by induction assumption, $apply(\Lens,\SomeOf{(x,y)},es) = apply(F(\Lens),\SomeOf{(x,y)},es)$, so
    $es' = es''$.  This means, $apply(\Lens,\None,\InLOf{x}::es) =
    \InROf{y}::es'$ and $apply(F(\Lens),\None,(\InLOf{x})::es) =
    (\InROf{y})::es'$, so they are equal, as desired.
  \end{case}

  \begin{case}[first edit is a create left]
    Symmetric to previous case
  \end{case}

  \begin{case}[first edit is a put right]
    So by the case, $apply(\Lens,\SomeOf{(x,y)},\InLOf{x'}::es) = \InROf{y'}::es'$, where
    $\Lens.\PutROf{x}{y} = y'$ and $apply(\Lens,\SomeOf{(x',y')},es) = es'$.

    Performing $apply$ on $F(\Lens)$, we get
    $apply(F(\Lens),\SomeOf{(x,y)},(\InLOf{x'})::es) = (\InROf{y'})::es''$ where
    $F(\Lens).\PutRight \App (x',\SomeOf{(x,y)}) = (y'',\SomeOf{(x',y'')})$ and
    $apply(\Lens,c',es) = es''$.

    So, by definition, $\Fst \App (F(\Lens).\PutRight \App (x',\SomeOf{(x,y)}))
    = \Lens.PutROf{x'}{y'}$, so also by definition, $c' = \SomeOf{(x',y')}$.

    So, by induction assumption, $apply(\Lens,\SomeOf{(x',y')},es) =
    apply(F(\Lens),\SomeOf{(x',y')},es)$, so $es' = es''$. This means,
    $apply(\Lens,\SomeOf{(x,y)},\InLOf{x'}::es) = \InROf{y'}::es'$ and
    $apply(F(\Lens),\SomeOf{(x,y)},(\InLOf{x'})::es) = (\InROf{y'})::es'$, so
    they are equal, as desired.
  \end{case}

  \begin{case}[first edit is a put left]
    Symmetric to previous case
  \end{case}
\end{proof}

\begin{theorem}
  Let $\Lens$ be a symmetric lens. The lens $\Lens$ is equivalent to a forgetful
  lens if, and only if, there exists a simple symmetric lens $\Lens'$ where
  $apply(\Lens,\Lens.init,es) = apply(\Lens',\None,es)$, for all put sequences
  $es$.
\end{theorem}

\begin{proof}
  \begin{case}[$\Rightarrow$]
    Let $\Lens$ be equivalent to a forgetful lens $\Lens'$.  Consider the
    simple symmetric lens, $S(\Lens')$.  By Lemma~\ref{lem:s-equiv},
    $apply(\Lens',\Lens'.init,es) = apply(S(\Lens'),\None,es)$.  As $\Lens$ is
    equivalent to $\Lens'$, $apply(\Lens,\Lens.init,es) =
    apply(\Lens',\Lens'.init,es)$.  So, by transitivity,
    $apply(\Lens,\Lens.init,es) = apply(S(\Lens),\None,es)$.
  \end{case}

  \begin{case}[$\Leftarrow$]
    Let $\Lens'$ be a simple symmetric lens where $apply(\Lens,\Lens.init,es) =
    apply(\Lens',\None,es)$, for all put sequences $es$.

    Consider $F(\Lens')$, a forgetful symmetric lens where $apply(\Lens',\None,es) =
    apply(F(\Lens'),\Lens'.init,es)$, for all put sequences $es$, as
    $\Lens'.init = \None$, by Lemma~\ref{lem:f-equiv}.  By transitivity, $apply(\Lens,\Lens.init,es) =
    apply(\Lens',\Lens'.init,es)$, so $\Lens$ and $\Lens'$ are equivalent.
  \end{case}

\end{proof}

\fi

\end{document}

%%% Local Variables:
%%% TeX-master: "main"
%%% End:

### 2018-05-23 papers/SymmetricLenses/main.tex
add better intro abstract

@@
 \begin{abstract}
-  Lenses are bidirectional transformers that are often used for synchronizing
-  data in different formats. A generalization of lenses, symmetric lenses,
-  enable synchronizations between formats where each format contains information
-  the other lacks. The theory of symmetric lenses relies heavily on complements
-  for maintaining synchronization. Despite the increased expressivity of
-  symmetric lenses, no major bidirectional languages or tools have adopted
-  symmetric lenses. While the use of complements allows for a clean and general
-  formulation of symmetric lenses, it also requires maintaining internal state
-  -- in addition to maintaining synchronized files, the complement must also be
-  maintained. We simplify the formulation of symmetric lenses to \emph{simple
-    symmetric lenses}. These lenses require no complement, instead using
-  familiar $create$ and $put$ functions to maintain synchronization.
-  Furthermore, we demonstrate the relationship between simple symmetric lenses,
-  by carving out a subset of symmetric lenses which are equivalent to simple
-  symmetric lenses. We integrate symmetric lenses into a modern bidirectional
-  language, Boomerang, while maintaining full backward compatibility.
-
-  Additionally, to enable easy use of symmetric lens combinators, we extend the
-  Optician algorithm to synthesize simple symmetric lenses from type and example
-  specifications. Unlike the bijective lenses Optician originally addresses,
-  there are many symmetric lenses between two data formats. Hypothesizing that
-  typically users want the lens that synchronizes the most data, we use an
-  information-theoretic measure based on stochastic regular expressions, to
-  quantify how much data is left non-synchronized, and develop an algorithm that
-  attempts to minimize this measure. Furthermore, as our synthesis algorithm
-  requires rewrites to our regular expressions, we prove that these the
-  probability distribution of strings within our stochastic regular expressions
-  are stable with respect to these rewrites. We experimentally demonstrate that
-  despite addressing a much harder problem, our extension to Optician still
-  maintains synthesis times competitive to Optician on Optician's bijective
-  benchmark suite, and generalizes to benchmarks synchronizing data not in
-  bijective correspondence.
+  Similar data is oftentimes represented and stored in different formats. Users
+  commonly need to synchronize the data shared between formats, where edits to
+  the data in one format gets propagated to the data in the other format.
+  Manually writing these data synchronizers is tedious and error prone. To
+  alleviate these difficulties, tools have been developed that automatically
+  generate these synchronization functions by generating bidirectional lens
+  programs from data format descriptors~\cite{optician,quotient-optician}.
+  However, these tools only work when the two formats describe exactly the same
+  data; it assumes that all important information present in one format is
+  present in the other. This is often not the case.
+
+  In this work, we automatically synthesize conversion functions between formats
+  that each may contain information the other lacks. We define \emph{simple
+    symmetric lenses}, a class of bidirectional programs for expressing these
+  types of conversion functions, and we specify a language that expresses simple
+  symmetric lenses that operate on string formats. We develop a type-directed
+  synthesis algorithm for generating lenses in this language from regular
+  expressions describing the data formats. A significant technical challenge in
+  this work arises from the fact that our target language permits
+  synchronization between formats that don't have exactly the same data -- data
+  in one format need not be present in the other format. This gives the
+  synthesis algorithm much more freedom, as data can be completely
+  unsynchronized (edits in one format are never propagated), completely
+  synchronized (the formats are in bijective correspondence), or anywhere in
+  between. Under the assumption that users typically want to synchronize as much
+  data as possible, we provide an information theoretic metric for how much data
+  is left unsynchronized based on the theory of \emph{stochastic regular
+    expressions}. Furthermore, we use this metric to develop a greedy synthesis
+  algorithm that tries to minimize unsynchronized data. We experimentally
+  demonstrate that despite addressing a much harder problem, our extension to
+  Optician still maintains synthesis times competitive to Optician on Optician's
+  bijective benchmark suite, while generalizing to benchmarks synchronizing data
+  not in bijective correspondence.
 \end{abstract}
@@
 % state of the world
-Lenses, groups of three functions that satisfy invertibility laws, are useful
-formulations for synchronization. Given data in one format, you can view it is a
-different data format using \emph{get} function. Given data in another format,
-you can turn it into data in the first format using a \emph{create} function.
-Given edits on the second format, you can propagate those updates to the first
-format using a \emph{put} function. However, writing these three functions is
-tedious and error prone. Much better would be to write a single program that
-expresses all three of the functions while also guaranteeing invertibility
-properties about them.  This is the approach used in many bidirectional
-languages: users write code in a bidirectional DSL, and all three functions are
-maintained.  These bidirectional languages are often used for data
-synchronization tasks - keeping data in different forms synchronized.
-
-While lenses are used to synchronize data formats, but they remain incapable of
-synchronizing formats where each format maintains information the other lacks.
-To address this, lenses have been generalized to symmetric lenses.  Symmetric
-lenses address these synchronization tasks by maintaining an explicit complement
-to keep track of projected data.  However, despite the benefits of symmetric
-lenses, they have not been adopted into bidirectional languages, remaining instead
-as a rarely used Haskell package.
-
-We believe this is largely due to complements. While complements provide a
-simple, theoretical framework to express symmetric lenses, they become tricky
-when used for synchronization tasks.  Critically, instead of maintaining two
-pieces of data (one for each synchronized data format), the lens must now
-maintain three (one for each synchronized data format, and one for the
-complement).  This can then be difficult when migrating data - I can no longer
-copy my formats to a new computer and tell a bidirectional language to keep them
-synchronized, I instead must copy my formats and the complement to the new
-computer.
-
-We propose an alternative method. Instead of using complements, we can instead
-use familiar \emph{create} and \emph{put} methods to synchronize the data
-formats. In this paper, we present \emph{simple symmetric lenses}, a
-reformulation of symmetric lenses consisting of 4 functions: 2 create functions
-and 2 put functions. While these are less expressive than symmetric lenses,
-simple symmetric lenses are able to express all synchronization tasks tackled by
-asymmetric lenses - all asymmetric lenses are simple symmetric lenses.
-Furthermore, we find a predicate over symmetric lenses, where symmetric lenses
-satisfy the predicate if, and only if, they can be written as a simple symmetric lens.
-
-As part of our work, we extend Optician to handle symmetric lens combinators,
-while maintaining full backwards compatibility. This means simple symmetric
-lenses compose simply with both the matching lens and quotient lens combinators
-Boomerang provides. Furthermore, despite providing additional capabilities, our
-implementation of Boomerang uses fewer lines of code in the core lens library
-than were originally present - the symmetry of our lenses enable more code
-reuse than the asymmetric versions permitted.
-
-Furthermore, we enable easy use of these combinators by developing a synthesis
-algorithm for symmetric lenses within Boomerang. We do this by extending
-Optician, a tool for synthesizing bijective lenses, with capabilities for
-synthesizing symmetric lenses as well. However, this is not trivial. Bijective
-lenses permit relatively few well-typed terms: finding a bijective lens between
-two data formats is not easy. However, finding a symmetric lens between two data
-formats is relatively easy, as the data synchronization function can be a
-``disconnect'' synchronizer -- changing the data in one format doesn't impact
-the data in the other. Therefore, we need a means to find the intended symmetric
-lens, not merely a symmetric lens. While input-output examples can help, we wish
-to make our tool able to synthesize many functions without requiring users to
-specify their intent with large example suites.
-
-Instead, we assume that users want to synchronize as much data as possible. To
-these ends, we develop an information thereoretic metric measuring the entropy
-of how much data is left non-synchronized, utilizing \emph{stochastic regular
-  expressions}. We syntactically find the entropy of stochastic regular
-expressions for unambiguous stochastic regular expresions. Furthermore, just as
-Optician applies rewrites to regular expressions to find well-aligned ones, our
-enhancement to Optician applies rewrites to stochastic regular expressions to
-find well-aligned ones. We prove these rewrites maintain both the language of
-the stochastic regular expressions and the probability distributions over them.
-
-Lastly, we demonstrate the effectiveness of our algorithm by synthesizing all
-the bijective transformations present in Optician's benchmark suite with the
-new, symmetric algorithm.  Furthermore, we synthesize an additional
-\NumBenchmarks new, non-bijective benchmarks.
-
-In summary, our contributions are as follows:
+Similar data can come in varieties of formats: bibliography information can be
+stored in the BibTeX, EndNote, or ACM reference formats; web service resources
+can be provided to clients in legacy or modern formats; calendar events can be
+stored in CSV files or ICS files. Oftentimes, we want to synchronize files
+across formats: when my collaborator edits her bibliography stored in the ACM
+reference format, I want these edits reflected in my EndNote file; when I update
+data through a legacy API, I want those changes to be reflected when I retrieve
+it through a modern API; when I create a new calendar event in my ICS file, I
+want my CSV calendar files to be updated to reflect this change.
+
+Unfortunately, writing such synchronizers is tedious and error prone.
+Bidirectional languages ease the difficulty of writing such synchronizers, as a
+single, well-typed program expresses multiple conversion functions, and
+guarantee these functions satisfy round-tripping laws. However, bidirectional
+languages are difficult to program in as well -- often requiring users to
+permute data in fiddly ways while satisfying the type system's complex
+unambiguity constraints.
+
+A better approach is to synthesize such conversion functions. Recent research
+has gone into automatically synthesizing such conversion functions from
+input-output examples and formal descriptions of the data
+formats~\cite{optician,quotient-optician}. However, this work assumes that all
+data fields present in one format are present in the other, but this is not
+always true. Oftentimes, fields present in one format are not present in the
+other: updated APIs can expose new information, or omit legacy information; the
+ACM reference format includes article number information, whereas EndNote
+includes page number information; ICS files contain a URL field that Google
+Calendar CSV files do not expose. We aim to synthesize transformations where the
+data fields are not necessarily in bijective correspondence, where each data
+format can contain important information not present in the other format.
+
+We follow a similar approach to existing work on synthesizing conversion
+functions; instead of synthesizing these functions individually, we synthesize
+bidirectional programs. To this end, we develop a new formulation of
+bidirectional programs, \emph{simple symmetric lenses}, which permit
+synchronizations between formats not in bijective correspondence. We develop a
+language for expressing simple symmetric lenses that operate on string data.
+Terms in this language are typed between regular expressions, where the regular
+expressions describe the synchronized data formats.
+
+Our synthesis algorithm takes as input a suite of input-output examples and two
+regular expressions describing the data formats. Our synthesis algorithm uses
+type-directed synthesis on the two regular expressions to synthesize a lens that
+is well-typed between the provided regular expressions and satisfies the
+provided input-output examples.
+
+A primary novel difficulty is that, unlike previous work where there are
+relatively few well-typed lenses, there are a relatively many well-typed simple
+symmetric lenses. As symmetric lenses permit data to be present in only one
+format, for any data component there is now a choice: is this data only relevant
+to this format and as such should be left unsynchronized, or is this data
+present in both formats and as such should be synchronized.
+
+We operate under the assumption that users wish to synchronize as much data as
+possible, and try to synchronize as much data as the types and examples allow.
+This task gets harder when there are more complex decisions involved. For
+example, if I have two fields in one format, and one in the other, which field
+on the left should be synchronized to the one of the right? Or would it be
+better for the field on the right to be synchronized to a mish-mash of
+subcomponents from each field? To answer these questions, we develop a cost
+metric for reasoning about which lens synchronizes the most information.
+
+As we wish to synchronize as much data as we can, our cost model gives higher
+cost to those with more data is left unsynchronized. In particular, the cost of
+a lens is the expected number of bits required to recover a string in one format
+when a synchronized string in the other format is provided. In this model,
+bijective lenses have zero cost, and as more information is left unsychronized,
+the cost is increased.
+
+Our model for this is based on \emph{stochastic regular expressions}~\cite{?}.
+We represent a probability distribution for strings by a stochastic regular
+expression, and find a way to syntactically infer the expected information
+content (or \emph{entropy}) of an unambiguous regular expression. We use this
+entropy to calculate the expected information content to recover a string, when
+we are given a string it is synchronized to.
+
+We develop an algorithm that synthesizes a low-cost lens. Our algorithm consists
+of two communicating synthesizers, one that proposes candidate stochastic
+regular expression pairs that it thinks will align well, and one that greedily
+searches for a lens between those two regular expressions that minimizes cost
+and involves no rewrites.
+
+Our first algorithm, \TypeProp, proposes candidate stochastic regular
+expressions. These candidate stochastic regular expressions are generated by
+traversing \emph{star semiring} equivalences on stochastic regular expressions.
+These proposed stochastic regular expression pairs have equivalent languages and
+probability distributions as the original ones: we prove the star-semiring
+rewrites this synthesizer applies are sound rewrites on stochastic regular
+expression.
+
+Our second algorithm, \GreedySynth, utilizes alternative languages for
+stochastic regular expressions and lenses, \emph{stochastic DNF regular
+  expressions} and \emph{symmetric DNF lenses}. While these languages are hard
+to read and understand, they permit an efficient type-directed synthesis
+algorithm. We develop our cost metric on DNF lenses, and provide a greedy
+algorithm that tries to find a symmetric DNF lens of minimal cost.
+
+We evaluate our algorithm on a benchmark suite consisting of the original
+Optician benchmarks, as well as additional benchmarks consisting of
+synchronization tasks where the data is not in bijective correspondence.
+
+In summary, our contributions are:
 \begin{enumerate}
-\item We define simple symmetric lenses and show their relationship with
-  symmetric and asymmetric lenses (\S\ref{sec:ssl} and \S\ref{sec:ssl-rels}).
-\item We introduce simple symmetric lens combinators that can be used for
-  synchronizing string data formats (\S\ref{sec:ssl-string}).
-\item We define stochastic regular expressions, demonstrate how to find their
-  entropy, and show how certain rewrites on them maintain language and
-  probability distribution (\S\ref{sec:stoch-regexp}).
-\item We define symmetric DNF lenses, and stochastic DNF regular expressions.
-  We show the conversion of regular expression to DNF regular expression
-  maintains language and probability distribution.  We show how to infer the
-  entropy of a DNF regular expression (\S\ref{sec:dnf-regex-lens}).
-\item We describe a synthesis algorithm that attempts to minimize unsynchronized
-  information (\S\ref{sec:synthesis}).
-\item We provide a fully backwards compatible implementation of symmetric lenses
-  in Boomerang, and an implementation of our synthesis algorithm into Optician
-  (\S\ref{sec:implementation}).
-\item We evaluate our synthesis algorithm on a variety of benchmarks, and find
-  that we lose very little synthesis time, while gaining a large amount of
-  expressivity (\S\ref{sec:evaluation}).
+\item We develop \emph{simple symmetric lenses} a new formulation for
+  bidirectional programs that can project data from both sides~(\S\ref{sec:ssl}),
+  and describe a language that expresses such lenses on string
+  data~(\S\ref{sec:ssl-string}).
+\item We develop a means to traverse stochastic regular expression equivalences,
+  while preserving probability distributions and
+  languages~(\S\ref{sec:preliminaries}).
+\item We develop stochastic DNF regular expressions and DNF lenses. We develop a
+  sound algorithm for converting stochastic regular expressions to stochastic
+  DNF regular expressions~(\S\ref{sec:dnf-regex-lens}), and a
+  information-theoretic cost metric on DNF
+  lenses~(\S\ref{sec:symmetric-dnf-lenses}).
+\item We extended Boomerang with our simple symmetric lens combinators. We
+  updated the Optician synthesis system to synthesize symmetric lenses with our
+  new algorithm.
+\item We evaluate our synthesis algorithm on Opticians benchmark suite as well
+  as new, symmetric benchmarks, and find that we lose very little synthesis
+  time, while gaining a large amount of expressivity (\S\ref{sec:evaluation}).
 \end{enumerate}
-
 % end introduction
@@
 \label{sec:preliminaries}
-A \emph{symmetric lens} between $X$ and $Y$ is an element of the type
-$\exists C.\{ \mathit{missing} :
-C,\,get : X \times C \to Y \times C,\,put : Y \times C \to X \times C\}$
-satisfying the following laws:
-\begin{equation}
-  \tag{\PutRL}
-  \frac{\PutR(x,c) = (y,c')}{\PutL(y,c') = (x,c')}
-\end{equation}
-\begin{equation}
-  \tag{\PutLR}
-  \frac{\PutL(y,c) = (x,c')}{\PutR(x,c') = (y,c')}
-\end{equation}
-
-Given a lens $l \in X \leftrightarrow Y$, a \emph{put object} for $l$ is a
-member of $X + Y$. Consider the function $apply$, which, given a lens and an
-element of that lens's complement, is a function from sequences of put objects
-to sequences of put objects. If $apply(\Lens,c,es) = es'$, then that means, starting
-with complement $c$, after edit $es_i$, the lens $\Lens$ synchronizes the other
-data to $es'_i$.
-\[
-  \inferrule
-  {
-  }
-  {
-    apply(\Lens,c,[]) = []
-  }
-\]
-\[
-  \inferrule
-  {
-    \Lens.putr(x,c) = (y,c')\\
-    apply(\Lens,c',es) = es'
-  }
-  {
-    apply(\Lens,c,(\InLOf{x})::es) = (\InROf{y})::es'
-  }
-\]
-\[
-  \inferrule
-  {
-    \Lens.putl(y,c) = (x,c')\\
-    apply(\Lens,c',es) = es'
-  }
-  {
-    apply(\Lens,c,(\InROf{y})::es) = (\InLOf{x})::es'
-  }
-\]
-
-Two lenses, $\Lens_1$ and $\Lens_2$, are equivalent if
-$apply(\Lens_1,\Lens_1.init,es) = apply(\Lens_2,\Lens_2.init,es)$ for all edit
-sequences $es$.
+To define the entropy for data recovery, we must first develop a probability
+model for our languages. We do this by enhancing regular expressions into
+\emph{stochastic regular expressions}. With this, we can jointly
+express both a language and a probability distribution over that language.
+
+\begin{center}
+  \begin{tabular}{rcl}
+    \Regex{},\RegexAlt{}
+    & \GEq{} & $\String$ \\
+    & \GBar{} & $\emptyset$ \\
+    & \GBar{} & $\PRegexStar{\Regex}{p}$ \\
+    & \GBar{} & $\RegexConcat{\Regex_1}{\Regex_2}$ \\
+    & \GBar{} & $\PRegexOr{\Regex_1}{\Regex_2}{p}$
+  \end{tabular}
+\end{center}
+
+The language is defined like the regular expression without probability
+annotations would be.  The probabilities are defined as follows:
+
+\begin{center}
+  \begin{tabular}{rcl}
+    $\ProbabilityOf{\String}{\String'}$
+    & =
+    & $\begin{cases*}1 & if $\String = \String'$\\ 0 & otherwise\end{cases*}$ \\
+
+    $\ProbabilityOf{\emptyset}{\String}$
+    & =
+    & $0$ \\
+
+    $\ProbabilityOf{\RegexConcat{\Regex_1}{\Regex_2}}{\Character_1 \ldots \Character_n}$
+
+    & =
+    & $\Sigma_{\String = \String_1\String_2}\ProbabilityOf{\Regex_1}{\String_1}*\ProbabilityOf{\Regex_2}{\String_2}$ \\
+
+    $\ProbabilityOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability}}{\String}$
+    & =
+    & $\Probability * \ProbabilityOf{\Regex_1}{\String} +
+      (1-\Probability) * \ProbabilityOf{\Regex_2}{\String}$\\
+
+    $\ProbabilityOf{\PRegexStar{\Regex}{\Probability}}{\String}$
+    & =
+    & $\Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\Regex}{\String_i}$\\
+  \end{tabular}
+\end{center}
+
+Much like the synthesis algorithm for Optician, we need to apply rewrites to
+handle equivalences on the provided stochastic regular expressions.  However, no
+such equivalences have yet been defined.  We extend the \emph{star-semiring}
+equivalences to stochastic regular expressions.
+
+\begin{center}
+  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
+    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
+    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
+    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
+    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
+    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{(1-\Probability_1)*\Probability_2})}{\Probability_1*\Probability_2} & \OrAssociativityRule{}  \\
+    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
+    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
+    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
+    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
+    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
+    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability} & \UnrollstarLeftRule{} \\
+    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{\Probability} & \UnrollstarRightRule{} 
+  \end{tabular}
+\end{center}
+% \caption{Regular Expression Equivalences}
+% \label{fig:regex-equivalence-rules}
+
+\begin{theorem}
+  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
+  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
+\end{theorem}
+
 % end preliminaries
@@
 
+% begin simple symmetric string lenses
+\section{Simple Symmetric String Lenses}
+\label{sec:ssl-string}
+While we have a theory for simple symmetric lenses, to integrate with Boomerang,
+we need combinators to express these lenses.
+
+\subsection{Identity}
+The identity lens operates on a single regular expression.  It fully
+synchronizes the data within that regular expression: if data gets changed in
+the left, that change gets reflected in the right.
+
+\subsection{Disconnect}
+The disconnect lens performs the opposite function as the identity lens.  It
+leave the two data formats provided separate.  If data gets changed on the left,
+the right is left unaffected.  When the data is being newly added, default
+behavior comes from one side or the other.
+
+\subsection{Concat}
+The concat lens combines two existing lenses, working on their concatenated
+regular expressions.  It performs the left lens on the left portion of the
+string, and the right lens on the right portion.  It requires unambiguity
+constraints to break up the left and right portions.
+
+\subsection{Swap}
+The swap lens operates like the concat lens, but when going from one side to the
+other, it swaps the order of the lenses.
+
+\subsection{Compose}
+The swap lens operates like the concat lens, but when going from one side to the
+other, it swaps the order of the lenses.
+
+\subsection{Iterate}
+The iterate lens breaks the provided strings into chunks, and performs the
+provided lens on those chunks, then concatenates them together.
+
+\subsection{Or}
+The or lens performs one lens or another, depending on the inputs to the lens.
+
+\subsection{MergeL}
+The mergel lens is used when the same piece of data on the left is synchronized
+to data in potentially many different forms on the right. The first provided
+lens is the one defaulted to on creates.
+
+\subsection{MergeR}
+The merger lens is symmetric to the mergel lens.
+
+\subsection{Compose}
+The compose lens applies one lens to go to an intermediary format, then applies
+the other to go to the final format.
+% end simple symmetric string lenses
+
+% begin DNF regular expressions and lenses
+\section{Stochastic DNF Regular Expressions}
+\label{sec:dnf-regex-lens}
+We extend stochastic regular expressions to stochastic DNF regular expressions.
+\begin{center}
+  \begin{tabular}{l@{\ }c@{\ }l@{\ }>{\itshape\/}r}
+    % DNF_REGEX
+    \Atom{},\AtomAlt{} & \GEq{} & \PRegexStar{\DNFRegex{}}{\Probability}
+% & \StarAtomType{}
+\\
+    \Sequence{},\SequenceAlt{} & \GEq{} &
+                                                       $\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n}$ 
+%& \MultiConcatSequenceType{} 
+\\
+    \DNFRegex{},\DNFRegexAlt{} & \GEq{} & $\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}$ %& \MultiOrDNFRegexType{} 
+  \end{tabular}
+\end{center}
+
+As before, DNF regular expressions are essentially regular expressions with all
+concatenations fully distributed over all disjunctions.  As these DNF regular
+expressions are \emph{stochastic}, they are annotated with probabilities to
+express a probability distribution, in addition to a language.  This probability
+distribution is defined below:
+
+\begin{center}
+  \begin{tabular}{rcl}
+    $\ProbabilityOf{\PRegexStar{\DNFRegex}{\Probability}}{\String}$
+    & =
+    & $\Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\DNFRegex}{\String_i}$\\
+
+    $\ProbabilityOf{\SequenceOf{\String_0 \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_n \SeqSep \String_n}}{\String'}$
+    & =
+    & $\Sigma_{\String' = \String_0\String_1'\ldots\String_n'\String_n}\Pi_{i = 1}^n\ProbabilityOf{\Atom_i}{\String_i'}$ \\
+
+    $\ProbabilityOf{\DNFOf{(\Sequence_1,\Probability_1) \DNFSep \ldots \DNFSep (\Sequence_n,\Probability_n)}}{\String}$
+    & =
+    & $\Sigma_{i=1}^n\Probability_i * \ProbabilityOf{\Sequence_i}{\String}$\\
+  \end{tabular}
+\end{center}
+
+\begin{figure}
+  \raggedright
+  $\ConcatSequence{} \OfType{} \ArrowTypeOf{\SequenceType{}}{\ArrowTypeOf{\SequenceType{}}{\SequenceType{}}}$\\
+  $\ConcatSequenceOf{[\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n]}{[\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\StringAlt_m]}=
+  [\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n\Concat\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\StringAlt_m]$\\
+
+  \medskip
+
+  $\ConcatDNF{} \OfType{} \ArrowTypeOf{\DNFRegexType{}}{\ArrowTypeOf{\DNFRegexType{}}{\DNFRegexType{}}}$\\
+  $\ConcatDNFOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}{\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}=$\\
+      $\DNFLeft (\ConcatSequenceOf{\Sequence_1}{\SequenceAlt_1},\Probability_1*\ProbabilityAlt_1)\DNFSep \ldots
+      \DNFSep
+      (\ConcatSequenceOf{\Sequence_1}{\SequenceAlt_m},\Probability_1*\ProbabilityAlt_m)\DNFSep
+      \ldots$\\
+      $\DNFSep
+      (\ConcatSequenceOf{\Sequence_n}{\SequenceAlt_1},\Probability_n*\ProbabilityAlt_1)\DNFSep
+      \ldots \DNFSep
+      (\ConcatSequenceOf{\Sequence_n}{\SequenceAlt_m},\Probability_n * \ProbabilityAlt_m) \DNFRight$
+
+  \medskip
+
+  $\OrDNF{}_{\Probability} \OfType{}
+  \ArrowTypeOf{\DNFRegexType{}}{\ArrowTypeOf{\DNFRegexType{}}{\DNFRegexType{}}
+  }$ \\
+  $\OrDNFOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}{\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}{\Probability} =$\\
+  $\DNFOf{(\Sequence_1,\Probability_1*\Probability)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n*\Probability)\DNFSep(\SequenceAlt_1,\ProbabilityAlt_1*(1-\Probability))\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m*(1-\Probability))}$
+
+  \medskip
+
+  \AtomToDNF{} \OfType
+  \ArrowTypeOf{\AtomType{}}{\DNFRegexType{}}\\
+  $\AtomToDNFOf{\Atom} = \DNFOf{(\SequenceOf{\EmptyString \SeqSep \Atom \SeqSep
+      \EmptyString},1)}$
+  \caption{DNF Regular Expression Functions}
+  \label{fig:dnf-regex-functions}
+\end{figure}
+
+We have a means of turning stochastic regular expressions into stochastic DNF
+regular expressions.
+The conversion algorithm itself, written $\ToDNFRegexOf{\Regex}$, is defined below.
+\[
+  \begin{array}{rcl}
+    \ToDNFRegexOf{\String} & = & \DNFOf{(\SequenceOf{\String},1)}\\
+    \ToDNFRegexOf{\emptyset} & = & \DNFOf{}\\
+    \ToDNFRegexOf{(\PRegexStar{\Regex}{\Probability})} & = & \AtomToDNFOf{\PRegexStar{(\ToDNFRegexOf{\Regex})}{\Probability}}\\
+    \ToDNFRegexOf{(\RegexConcat{\Regex_1}{\Regex_2})} & = & \ToDNFRegexOf{\Regex_1} \ConcatDNF \ToDNFRegexOf{\Regex_2}\\
+    \ToDNFRegexOf{(\PRegexOr{\Regex_1}{\Regex_2}{\Probability})} & = & \ToDNFRegexOf{\Regex_1} \OrDNF_{\Probability} \ToDNFRegexOf{\Regex_2}\\
+  \end{array}
+\]
+We have also proven an algorithm, demonstrating this conversion respects
+probability distributions.
+
+\begin{theorem}
+  $\ProbabilityOf{\Regex}{\String} = \ProbabilityOf{\ToDNFRegex \Regex}{\String}$
+\end{theorem}
+
+Furthermore, in the situation where a DNF regular expression is unambiguous, we
+have defined an efficient, syntactic means to find the entropy of the DNF
+regular expression.
+
+\begin{center}
+  \begin{tabular}{rcl}
+    $\EntropyOf{\PRegexStar{\DNFRegex}{\Probability}}$
+    & =
+    & $\frac{\Probability}{1-\Probability}(\EntropyOf{\DNFRegex} - \Log_2\Probability)
+      - \Log_2(1-\Probability)
+      $\\
+
+    $\EntropyOf{\SequenceOf{\String_0 \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_n \SeqSep \String_n}}{\String'}$
+    & =
+    & $\Sigma_{i = 1}^n\EntropyOf{\Atom_i}$ \\
+
+    $\EntropyOf{\DNFOf{(\Sequence_1,\Probability_1) \DNFSep \ldots \DNFSep (\Sequence_n,\Probability_n)}}$
+    & =
+    & $\Sigma_{i=1}^n\Probability_i(\EntropyOf{\Sequence_i}+\Log_2\Probability_i)$\\
+  \end{tabular}
+\end{center}
+
+\begin{theorem}
+  $\EntropyOf{\DNFRegex}$ is the entropy of $P_{\DNFRegex}$.
+\end{theorem}
+
+This means we also have an algorithm for finding the entropy of an unambiguous
+stochastic regular expression, as we can merely convert the regular expression
+into DNF form, and find the entropy of it.
+% end DNF regular expressions and lenses
+
+% begin symmetric dnf lenses
+\section{Symmetric DNF Lenses}
+\label{sec:symmetric-dnf-lenses}
+\subsection{Syntax}
+Symmetric DNF Lenses are what we synthesize, and we then turn them into
+Symmetric Lenses. The nonterminals $i$, $j$, $p$, $q$, $r$, $c$, and $d$ all
+represent natural numbers. The nonterminals $\String$ and $\StringAlt$ represent
+strings.
+
+\begin{center}
+  \begin{tabular}{@{}r@{\ }c@{}l@{}}
+    % REGEX
+    \SAtomLens{} & \GEq{} & $\IterateLensOf{\SDNFLens}$ \\
+    \SSQLens{} & \GEq{} & $(\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
+                          \ldots\SeqLSep
+                          (i_p,j_p,\SAtomLens_p)}
+                          ,\ListOf{(\String_1,\Atom_1);\ldots;(\String_q,\Atom_q)}
+                          ,\ListOf{\StringAlt_1;\ldots;\StringAlt_r})$ \\
+    \SDNFLens{} & \GEq{} & $(\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
+                           \ldots\DNFLSep
+                           (i_p,j_p,\SSQLens_p)}
+                           ,\ListOf{c_1;\ldots;c_q}
+                           ,\ListOf{d_1;\ldots;d_r})$ \\
+  \end{tabular}
+\end{center}
+
+Lets go through these from bottom to top.
+
+A DNF Lens \SDNFLens{} consists of three components. First is the mapping
+component. The mapping component is a list of triples $(i,j,\SSQLens)$. In this,
+$i$ and $j$ refer to sequence $i$ and sequence $j$ in the left and right DNF
+regular expressions, respectively. The third component, \SSQLens{}, is a
+sequence lens that maps between the two sequences specified by $i$ and $j$. The
+second component is a list of integers $c$. The integer, $c_i$ at position $i$
+states that when performing a $\CreateR$ on a string matching sequence
+$\Sequence_i$, it will use the sequence lens that maps it to sequence
+$\SequenceAlt_{c_i}$ on the right. The third component is a list of integers
+$d$, and follows the same pattern as the second component, but maps from right
+to left.
+
+A Sequence lens \SSQLens{} consists of three components. First is the mapping
+component. The mapping component is a list of triples $(i,j,\SAtomLens)$. In
+this, $i$ and $j$ refer to atom $i$ and atom $j$ on the left and right
+sequences, respectively. The third component, \SAtomLens{}, is an atom lens that
+maps between the two atoms specified by $i$ and $j$. The second component is a
+list of strings, $\String$. The string, $s_i$ at position $i$ states that when
+performing a $\CreateL$ on a string matching atom $\Atom_i$, it will do one of
+two things. If $\Atom_i$ is mapped to in the mapping component, then performing
+the atom's $\CreateL$ will actually $\PutL$ into that string. If $\Atom_i$ is
+not mapped to, then $\String_i$ will be created as a representative of
+$\Atom_i$.
+
+An atom lens \AtomLens{} is merely the iteration of a symmetric DNF Lens.
+
+\subsection{Typing and Semantics}
+Because these lenses only encode functions on well-typed programs, we introduce
+the semantics alongside the typing rules.  The semantics are over typing
+derivations, instead of over individual lenses.
+
+\paragraph*{DNF Lenses}
+The typing judgment is a 3-ary relation over a single DNF lens, and two DNF
+regular expressions. If $\SDNFLens \OfType \DNFRegex \Leftrightarrow
+\DNFRegexAlt$, then the $\SDNFLens.\CreateR$, $\SDNFLens.\CreateL$,
+$\SDNFLens.\PutR$, and $\SDNFLens.\PutL$ functions form a symmetric lens.
+
+The typing judgement has 3 components. The first is the sublens components,
+confirming that the sequence lenses the DNF lens is comprised of are all
+well-typed. The second guarantees that if each sequence on the left has a
+sequence lens that can be used for \CreateR{}s and \PutR{}s. The last guarantees
+the same for sequences on the right, with \CreateL{} and \PutL{}.
+
+\[
+  \inferrule*
+  {
+    \SSQLens_1 \OfType \Sequence_{i_1} \Leftrightarrow \SequenceAlt_{j_1}\\
+    \ldots\\
+    \SSQLens_p \OfType \Sequence_{i_p} \Leftrightarrow \SequenceAlt_{j_p}\\\\
+    i_{c_1} = 1\\
+    \ldots\\
+    i_{c_q} = q\\\\
+    j_{d_1} = 1\\
+    \ldots\\
+    j_{d_r} = r
+  }
+  {
+    (\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
+      \ldots\DNFLSep
+      (i_p,j_p,\SSQLens_p)}
+    ,\ListOf{c_1;\ldots;c_q}
+    ,\ListOf{d_1;\ldots;d_r})
+    \OfType\\\\
+    \DNFOf{(\Sequence_1,\Probability_1) \DNFSep \ldots \DNFSep (\Sequence_q,\Probability_q)}
+    \Leftrightarrow
+    \DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1) \DNFSep \ldots \DNFSep (\SequenceAlt_r,\ProbabilityAlt_q)}
+  }
+\]
+
+The \CreateR{} function looks for the sequence the provided string matches. If
+the string matches sequence $\Sequence_x$, then the lens will look in the create
+list at position $x$ to find which sequence lens to use. Then, the sequence lens
+transforms the given string using that sequence lens.
+
+The \PutR{} function finds what pairs sequences the input source and view
+strings match.  If there that pair of sequences have a sequence lens between
+them, then the DNF lens merely performs that sequence lens on the provided
+strings.  If there isn't a pair of sequence lenses between them, then \CreateR{}
+is performed on the source, with the view forgotten.
+
+The \CreateL{} and \PutL{} functions are defined symmetrically.
+
+\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+  $\CreateR{} \App s$ & = & $\SSQLens_{c_x}.\CreateR{} \App s$ if $s \in \Sequence_x$\\
+  $\CreateL{} \App v$ & = & $\SSQLens_{d_y}.\CreateL{} \App v$ if $v \in \SequenceAlt_y$\\
+  $\PutR{} \App s \App v$ & = &
+                               $\begin{cases*}
+                                 \SSQLens_x.\PutR{} \App s \App v & if $s \in \Sequence_{i_x}$ and $v \in \SequenceAlt_{j_x}$\\
+                                 \CreateR{} \App s & if $\nexists x.$ $s \in \Sequence_{i_x}$ and $v \in \SequenceAlt_{j_x}$
+                               \end{cases*}$\\
+  $\PutL{} \App v \App s$ & = &
+                               $\begin{cases*}
+                                 \SSQLens_y.\PutL{} \App v \App s & if $v \in \SequenceAlt_{j_y}$ and $s \in \Sequence_{i_y}$\\
+                                 \CreateL{} \App v & if $\nexists y.$ $v \in \SequenceAlt_{j_y}$ and $s \in \Sequence_{i_y}$
+                               \end{cases*}$
+\end{tabular}
+
+\paragraph*{Sequence Lenses}
+The typing judgment is a 3-ary relation over a single Sequence lens, and two
+sequences. If $\SSQLens \OfType \Sequence \Leftrightarrow \SequenceAlt$, then
+the $\SSQLens.\CreateR$, $\SSQLens.\CreateL$, $\SSQLens.\PutR$, and
+$\SSQLens.\PutL$ functions form a symmetric lens.
+
+The typing judgement has 4 components. The first is the sublens components,
+confirming that the atom lenses the sequence lens is comprised of all are
+well-typed. The second guarantees that if each string that will be used for
+\CreateR{}s are members of the correct atoms.  The third guarantees
+the same for strings and atoms on the right, with \CreateL.  The last guarantees
+that each atom is mapped by at most one atom lens.
+
+\[
+  \inferrule*
+  {
+    \SAtomLens_1 \OfType \Atom_{i_1} \Leftrightarrow \AtomAlt_{j_1}\\
+    \ldots\\
+    \SAtomLens_p \OfType \Atom_{i_p} \Leftrightarrow \AtomAlt_{j_p}\\\\
+    \String_1 \in \Atom_1\\
+    \ldots\\
+    \String_q \in \Atom_q\\\\
+    \StringAlt_1 \in \AtomAlt_1\\
+    \ldots\\
+    \StringAlt_r \in \AtomAlt_r\\\\
+    i_x = i_y \BooleanImplies x = y\\
+    j_x = j_y \BooleanImplies x = y
+  }
+  {
+    (\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
+      \ldots\SeqLSep
+      (i_p,j_p,\SAtomLens_p)}
+    ,\ListOf{\String_1;\ldots;\String_q}
+    ,\ListOf{\StringAlt_1;\ldots;\StringAlt_r})\\
+    \OfType
+    \SequenceOf{\String_0' \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_q
+      \SeqSep \String_q'}
+    \Leftrightarrow
+    \SequenceOf{\StringAlt_0' \SeqSep \AtomAlt_1 \SeqSep \ldots \SeqSep \AtomAlt_r \SeqSep \StringAlt_r'}
+  }
+\]
+
+For each component of the string matching an atom, the \CreateR function looks
+for the atom lens that maps on the atom. If there is such an atom lens,
+\SAtomLens, then that the sequence lens puts the provided string into the
+default string for the target atom. If there is no such atom lens, then the
+sequence lens merely uses the default string.
+
+For each component of the string matching an atom, the \PutR function looks
+for the atom lens that maps on the atom. If there is such an atom lens,
+\SAtomLens, then that the atom lens puts the provided string of the source atom into the
+string of the target atom. If there is no such atom lens, then the
+sequence lens merely recovers the target's string.
+
+The \CreateL{} and \PutL{} functions are defined symmetrically.
+
+\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+  $\CreateR{} \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q'$
+  & = 
+  & $\StringAlt_0' \Concat \StringAlt_1'' \Concat \ldots \Concat
+    \StringAlt_r'' \Concat \StringAlt_r'$\\
+  & & where $\StringAlt_y'' =
+    \begin{cases*}
+      \SAtomLens_k.\PutRight \App \String_{i_k}'' \App \StringAlt_y & if $j_k = y$\\
+      \StringAlt_y & if $\nexists k. j_k = y$\\
+    \end{cases*}$\\
+  $\CreateL{} \App \StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q''
+  \Concat \StringAlt_q'$
+  & = 
+  & $\String_0' \Concat \String_1'' \Concat \ldots \Concat
+    \String_r'' \Concat \String_r'$\\
+  & & where $\String_x'' =
+    \begin{cases*}
+      \SAtomLens_k.\PutRight \App \StringAlt_{j_k}'' \App \String_x & if $i_k = x$\\
+      \String_x & if $\nexists k. i_k = x$\\
+    \end{cases*}$\\
+  $\PutR{} \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q' \App \StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q'' \Concat \StringAlt_q'$
+  & =
+  & $\StringAlt_0'\Concat \StringAlt_1''' \Concat \ldots \Concat \StringAlt_q''' \Concat \StringAlt_q'$\\
+  & & where
+      $t_y''' =
+      \begin{cases*}
+        \SAtomLens_k.\PutRight \App \String_{i_k}'' \App \StringAlt_{y}'' & if $j_k = y$\\
+        \StringAlt_y'' & if $\nexists k. j_k = y$\\
+      \end{cases*}$\\
+  $\PutL{} \App \StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q'' \Concat \StringAlt_q' \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q'$
+  & =
+  & $\String_0'\Concat \String_1''' \Concat \ldots \Concat \String_r''' \Concat \String_r'$\\
+  & & where
+      $s_x''' =
+      \begin{cases*}
+        \SAtomLens_k.\PutLeft \App \StringAlt_{j_k}'' \App \String_{x}'' & if $i_k = x$\\
+        \String_x'' & if $\nexists k. i_k = x$\\
+      \end{cases*}$\\
+\end{tabular}
+
+\paragraph*{Atom Lenses}
+The typing judgment is a 3-ary relation over a single atom lens, and two
+atoms. If $\SAtomLens \OfType \Atom \Leftrightarrow \AtomAlt$, then
+the $\SAtomLens.\CreateR$, $\SAtomLens.\CreateL$, $\SAtomLens.\PutR$, and
+$\SAtomLens.\PutL$ functions form a symmetric lens.
+
+The typing judgement just confirms that the DNF lens that comprises the
+sequence lens is also well typed.
+
+\[
+  \inferrule*
+  {
+    \SDNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt
+  }
+  {
+    \IterateLensOf{\SDNFLens}
+    \OfType \PRegexStar{\DNFRegex}{\Probability}
+    \Leftrightarrow \PRegexStar{\DNFRegexAlt}{\ProbabilityAlt}
+  }
+\]
+
+For each component of the string matching an atom, the \CreateR function looks
+for the atom lens that maps on the atom. If there is such an atom lens,
+\SAtomLens, then that the sequence lens puts the provided string into the
+default string for the target atom. If there is no such atom lens, then the
+sequence lens merely uses the default string.
+
+For each component of the string matching an atom, the \PutR function looks
+for the atom lens that maps on the atom. If there is such an atom lens,
+\SAtomLens, then that the atom lens puts the provided string of the source atom into the
+string of the target atom. If there is no such atom lens, then the
+sequence lens merely recovers the target's string.
+
+The \CreateL{} and \PutL{} functions are defined symmetrically.
+
+\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+  $\CreateR{} \App \String_0 \Concat \ldots \Concat \String_n$
+  & = 
+  & $\StringAlt_1 \Concat \ldots \Concat \StringAlt_n$
+  where $\StringAlt_i = \SDNFLens.\CreateR \App \String_i$\\
+  $\CreateL{} \App \StringAlt_0 \Concat \ldots \Concat \StringAlt_m$
+  & = 
+  & $\String_1 \Concat \ldots \Concat \String_m$
+  where $\String_i = \SDNFLens.\CreateL \App \StringAlt_i$\\
+  $\PutR{} \App \String_0 \Concat \ldots \Concat \String_n \App
+  \StringAlt_1 \Concat \ldots \Concat \StringAlt_m$
+  & = 
+  & $\StringAlt_1' \Concat \ldots \Concat \StringAlt_n'$
+    where $\StringAlt_i' =
+    \begin{cases*}
+      \SDNFLens.\PutR \App \String_i \App \StringAlt_i & if $i \leq m$\\
+      \SDNFLens.\CreateR \App \String_i & otherwise
+    \end{cases*}$\\
+  $\PutL{} \App \StringAlt_1 \Concat \ldots \Concat \StringAlt_m \App
+  \String_0 \Concat \ldots \Concat \String_n$
+  & = 
+  & $\String_1' \Concat \ldots \Concat \String_m'$
+    where $\String_i' =
+    \begin{cases*}
+      \SDNFLens.\PutL \App \StringAlt_i \App \String_i & if $i \leq n$\\
+      \SDNFLens.\CreateL \App \StringAlt_i & otherwise
+    \end{cases*}$\\
+\end{tabular}
+
+Now that we've developed symmetric DNF lenses, it's important to develop a means
+to guage how bijective the lens is.
+
+Instead of searching for \emph{any} term that satisfies the specification, we
+search for the \emph{best} term that satisfies the specification. This begs the
+question, by what metric do we use to determine which term is best. Intuitively,
+we aim to synthesize the lens which is \emph{closest} to being a bijection.
+Somewhat more formally, we aim to synthesize a lens which minimizes the expected
+number of bits which must be used to recover the source from the target, and the
+target from the source. The expected number of bits that are needed to encode
+informatation from a data source is a well known concept in information theory,
+and is known as the \emph{entropy} of the data source.
+
+The entropy of $\DNFRegexAlt$ given $\DNFRegex$ and $\SDNFLens$,
+$\EntropyOf{\DNFRegexAlt \Given \DNFRegex,\SDNFLens}$ is defined symmetrically.
+With these $\operatorname*{argmin}_\theta$ terms defined, the ``best'' DNF lens
+between $\DNFRegex$ and $\DNFRegexAlt$ is $\ArgminOver{\SetOf{\SDNFLens \SuchThat
+    \SDNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt}}
+\EntropyOf{\DNFRegex \Given \DNFRegexAlt,\SDNFLens} + \EntropyOf{\DNFRegexAlt
+  \Given \DNFRegex,\SDNFLens}$.
+
+\begin{centering}
+  \[
+    \begin{array}{rl}
+      & \EntropyOf{\DNFOf{\Sequence_1 \DNFSep \ldots \DNFSep \Sequence_n}\\
+      & \hspace*{1em}\Given
+        \DNFOf{\SequenceAlt_1 \DNFSep \ldots \DNFSep \SequenceAlt_m}\\
+      & \hspace*{1.21em},~(\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
+        \ldots\DNFLSep
+        (i_p,j_p,\SSQLens_p)}
+        ,\ListOf{c_1;\ldots;c_q}
+        ,\ListOf{d_1;\ldots;d_r})}\\
+      =\\
+      & \frac{\Sigma_{j=1}^m(\frac{\Sigma_{k\SuchThat j_k = j}
+        \EntropyOf{\Sequence_{i_k}\Given\SequenceAlt_j,\SequenceLens_k}}{\SizeOf{\SetOf{k\SuchThat
+        j_k = j}}} + log_2\SizeOf{\SetOf{k\SuchThat j_k = j}})}{m}\\\\
+      \\
+
+      & \EntropyOf{
+        \SequenceOf{\String_0,\Atom_1,\ldots,\Atom_n,\String_n}\\
+      & \hspace*{1em}\Given
+        \SequenceOf{\StringAlt_0,\AtomAlt_1,\ldots,\AtomAlt_m,\StringAlt_m}\\
+      & \hspace*{1.21em},~
+        (\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
+        \ldots\SeqLSep
+        (i_p,j_p,\SAtomLens_p)}
+        ,\ListOf{(k_1,\String_1);\ldots;(k_q,\String_q)}
+        ,\ListOf{(l_1,\StringAlt_1);\ldots;(l_r,\StringAlt_r)})}\\
+      =\\
+      & \Sigma_{x=1}^p\EntropyOf{\Atom_{i_x} \Given \AtomAlt_{i_x},\AtomLens_x} +
+        \Sigma_{x=1}^q\EntropyOf{\Atom_{k_x}}\\\\
+      \\
+
+      & \EntropyOf{\StarOf{\DNFRegex} \Given \StarOf{\DNFRegexAlt},\StarOf{\SDNFLens}}\\
+      =\\
+      & 4*\EntropyOf{\DNFRegex\Given\DNFRegexAlt,\SDNFLens}
+    \end{array}
+  \]
+\end{centering}
+
+% end symmetric dnf lenses
+
+% begin synthesis
+\section{Synthesis}
+\label{sec:synthesis}
+While examples can be used to tighten the
+specification, a large number of well-constructed examples are required to
+sufficiently constrain the search space.
+
+Unfortunately, trying to find the best symmetric lens is a daunting task!
+Unless we find a bijective lens (in which case the sum is zero), we can never be
+sure if we will find a better lens than the one we found.
+% end synthesis
+
+% begin implementation
+\section{Implementation}
+\label{sec:implementation}
+% end implementation
+
+% begin evaluation
+\section{Evaluation}
+\label{sec:evaluation}
+% end evaluation
+
+% begin related-work
+\section{Related Work}
+\label{sec:related}
+A \emph{symmetric lens} between $X$ and $Y$ is an element of the type
+$\exists C.\{ \mathit{missing} :
+C,\,get : X \times C \to Y \times C,\,put : Y \times C \to X \times C\}$
+satisfying the following laws:
+\begin{equation}
+  \tag{\PutRL}
+  \frac{\PutR(x,c) = (y,c')}{\PutL(y,c') = (x,c')}
+\end{equation}
+\begin{equation}
+  \tag{\PutLR}
+  \frac{\PutL(y,c) = (x,c')}{\PutR(x,c') = (y,c')}
+\end{equation}
+
+Given a lens $l \in X \leftrightarrow Y$, a \emph{put object} for $l$ is a
+member of $X + Y$. Consider the function $apply$, which, given a lens and an
+element of that lens's complement, is a function from sequences of put objects
+to sequences of put objects. If $apply(\Lens,c,es) = es'$, then that means, starting
+with complement $c$, after edit $es_i$, the lens $\Lens$ synchronizes the other
+data to $es'_i$.
+\[
+  \inferrule
+  {
+  }
+  {
+    apply(\Lens,c,[]) = []
+  }
+\]
+\[
+  \inferrule
+  {
+    \Lens.putr(x,c) = (y,c')\\
+    apply(\Lens,c',es) = es'
+  }
+  {
+    apply(\Lens,c,(\InLOf{x})::es) = (\InROf{y})::es'
+  }
+\]
+\[
+  \inferrule
+  {
+    \Lens.putl(y,c) = (x,c')\\
+    apply(\Lens,c',es) = es'
+  }
+  {
+    apply(\Lens,c,(\InROf{y})::es) = (\InLOf{x})::es'
+  }
+\]
+
+Two lenses, $\Lens_1$ and $\Lens_2$, are equivalent if
+$apply(\Lens_1,\Lens_1.init,es) = apply(\Lens_2,\Lens_2.init,es)$ for all edit
+sequences $es$.
+
 Two simple symmetric lenses are equivalent if the four functions they comprise
@@
 \end{theorem}
-
-% begin relationship
-\section{Relationships Between Simple Symmetric Lenses and Other Lenses}
-\label{sec:ssl-rels}
 
@@
 \end{theorem}
-% end relationship
-
-% begin simple symmetric string lenses
-\section{Simple Symmetric String Lenses}
-\label{sec:ssl-string}
-While we have a theory for simple symmetric lenses, to integrate with Boomerang,
-we need combinators to express these lenses. While Boomerang has additional,
-complex combinators, we focus on the core ones, which are also those we
-eventually wish to synthesize.
-
-\subsection{Identity}
-The identity lens operates on a single regular expression.  It fully
-synchronizes the data within that regular expression: if data gets changed in
-the left, that change gets reflected in the right.
-
-\subsection{Disconnect}
-The disconnect lens performs the opposite function as the identity lens.  It
-leave the two data formats provided separate.  If data gets changed on the left,
-the right is left unaffected.  When the data is being newly added, default
-behavior comes from one side or the other.
-
-\subsection{Concat}
-The concat lens combines two existing lenses, working on their concatenated
-regular expressions.  It performs the left lens on the left portion of the
-string, and the right lens on the right portion.  It requires unambiguity
-constraints to break up the left and right portions.
-
-\subsection{Swap}
-The swap lens operates like the concat lens, but when going from one side to the
-other, it swaps the order of the lenses.
-
-\subsection{Compose}
-The swap lens operates like the concat lens, but when going from one side to the
-other, it swaps the order of the lenses.
-
-\subsection{Iterate}
-The iterate lens breaks the provided strings into chunks, and performs the
-provided lens on those chunks, then concatenates them together.
-
-\subsection{Or}
-The or lens performs one lens or another, depending on the inputs to the lens.
-
-\subsection{MergeL}
-The mergel lens is used when the same piece of data on the left is synchronized
-to data in potentially many different forms on the right. The first provided
-lens is the one defaulted to on creates.
-
-\subsection{MergeR}
-The merger lens is symmetric to the mergel lens.
-
-\subsection{Compose}
-The compose lens applies one lens to go to an intermediary format, then applies
-the other to go to the final format.
-% end simple symmetric string lenses
-
-% begin stochastic regular expressions
-\section{Stochastic Regular Expressions}
-\label{sec:stoch-regexp}
-To define the entropy for data recovery, we must first develop a probability
-model for our languages. We do this by enhancing regular expressions into
-\emph{stochastic regular expressions}. With this, we can jointly
-express both a language, and a probability distribution over that language.
-
-\begin{center}
-  \begin{tabular}{rcl}
-    \Regex{},\RegexAlt{}
-    & \GEq{} & $\String$ \\
-    & \GBar{} & $\emptyset$ \\
-    & \GBar{} & $\PRegexStar{\Regex}{p}$ \\
-    & \GBar{} & $\RegexConcat{\Regex_1}{\Regex_2}$ \\
-    & \GBar{} & $\PRegexOr{\Regex_1}{\Regex_2}{p}$
-  \end{tabular}
-\end{center}
-
-The language is defined like the regular expression without probability
-annotations would be.  The probabilities are defined as follows:
-
-\begin{center}
-  \begin{tabular}{rcl}
-    $\ProbabilityOf{\String}{\String'}$
-    & =
-    & $\begin{cases*}1 & if $\String = \String'$\\ 0 & otherwise\end{cases*}$ \\
-
-    $\ProbabilityOf{\emptyset}{\String}$
-    & =
-    & $0$ \\
-
-    $\ProbabilityOf{\RegexConcat{\Regex_1}{\Regex_2}}{\Character_1 \ldots \Character_n}$
-
-    & =
-    & $\Sigma_{\String = \String_1\String_2}\ProbabilityOf{\Regex_1}{\String_1}*\ProbabilityOf{\Regex_2}{\String_2}$ \\
-
-    $\ProbabilityOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability}}{\String}$
-    & =
-    & $\Probability * \ProbabilityOf{\Regex_1}{\String} +
-      (1-\Probability) * \ProbabilityOf{\Regex_2}{\String}$\\
-
-    $\ProbabilityOf{\PRegexStar{\Regex}{\Probability}}{\String}$
-    & =
-    & $\Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\Regex}{\String_i}$\\
-  \end{tabular}
-\end{center}
-
-Much like the synthesis algorithm for Optician, we need to apply rewrites to
-handle equivalences on the provided stochastic regular expressions.  However, no
-such equivalences have yet been defined.  We extend the \emph{star-semiring}
-equivalences to stochastic regular expressions.
-
-\begin{center}
-  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
-    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
-    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
-    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
-    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
-    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{(1-\Probability_1)*\Probability_2})}{\Probability_1*\Probability_2} & \OrAssociativityRule{}  \\
-    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
-    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
-    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
-    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
-    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
-    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability} & \UnrollstarLeftRule{} \\
-    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{\Probability} & \UnrollstarRightRule{} 
-  \end{tabular}
-\end{center}
-% \caption{Regular Expression Equivalences}
-% \label{fig:regex-equivalence-rules}
-
-\begin{theorem}
-  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
-  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
-\end{theorem}
-
-% end stochastic regular expressions
-
-% begin DNF regular expressions and lenses
-\section{Stochastic DNF Regular Expressions}
-\label{sec:dnf-regex-lens}
-We extend stochastic regular expressions to stochastic DNF regular expressions.
-\begin{center}
-  \begin{tabular}{l@{\ }c@{\ }l@{\ }>{\itshape\/}r}
-    % DNF_REGEX
-    \Atom{},\AtomAlt{} & \GEq{} & \PRegexStar{\DNFRegex{}}{\Probability}
-% & \StarAtomType{}
-\\
-    \Sequence{},\SequenceAlt{} & \GEq{} &
-                                                       $\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n}$ 
-%& \MultiConcatSequenceType{} 
-\\
-    \DNFRegex{},\DNFRegexAlt{} & \GEq{} & $\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}$ %& \MultiOrDNFRegexType{} 
-  \end{tabular}
-\end{center}
-
-As before, DNF regular expressions are essentially regular expressions with all
-concatenations fully distributed over all disjunctions.  As these DNF regular
-expressions are \emph{stochastic}, they are annotated with probabilities to
-express a probability distribution, in addition to a language.  This probability
-distribution is defined below:
-
-\begin{center}
-  \begin{tabular}{rcl}
-    $\ProbabilityOf{\PRegexStar{\DNFRegex}{\Probability}}{\String}$
-    & =
-    & $\Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\DNFRegex}{\String_i}$\\
-
-    $\ProbabilityOf{\SequenceOf{\String_0 \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_n \SeqSep \String_n}}{\String'}$
-    & =
-    & $\Sigma_{\String' = \String_0\String_1'\ldots\String_n'\String_n}\Pi_{i = 1}^n\ProbabilityOf{\Atom_i}{\String_i'}$ \\
-
-    $\ProbabilityOf{\DNFOf{(\Sequence_1,\Probability_1) \DNFSep \ldots \DNFSep (\Sequence_n,\Probability_n)}}{\String}$
-    & =
-    & $\Sigma_{i=1}^n\Probability_i * \ProbabilityOf{\Sequence_i}{\String}$\\
-  \end{tabular}
-\end{center}
-
-\begin{figure}
-  \raggedright
-  $\ConcatSequence{} \OfType{} \ArrowTypeOf{\SequenceType{}}{\ArrowTypeOf{\SequenceType{}}{\SequenceType{}}}$\\
-  $\ConcatSequenceOf{[\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n]}{[\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\StringAlt_m]}=
-  [\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n\Concat\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\StringAlt_m]$\\
-
-  \medskip
-
-  $\ConcatDNF{} \OfType{} \ArrowTypeOf{\DNFRegexType{}}{\ArrowTypeOf{\DNFRegexType{}}{\DNFRegexType{}}}$\\
-  $\ConcatDNFOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}{\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}=$\\
-      $\DNFLeft (\ConcatSequenceOf{\Sequence_1}{\SequenceAlt_1},\Probability_1*\ProbabilityAlt_1)\DNFSep \ldots
-      \DNFSep
-      (\ConcatSequenceOf{\Sequence_1}{\SequenceAlt_m},\Probability_1*\ProbabilityAlt_m)\DNFSep
-      \ldots$\\
-      $\DNFSep
-      (\ConcatSequenceOf{\Sequence_n}{\SequenceAlt_1},\Probability_n*\ProbabilityAlt_1)\DNFSep
-      \ldots \DNFSep
-      (\ConcatSequenceOf{\Sequence_n}{\SequenceAlt_m},\Probability_n * \ProbabilityAlt_m) \DNFRight$
-
-  \medskip
-
-  $\OrDNF{}_{\Probability} \OfType{}
-  \ArrowTypeOf{\DNFRegexType{}}{\ArrowTypeOf{\DNFRegexType{}}{\DNFRegexType{}}
-  }$ \\
-  $\OrDNFOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}{\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}{\Probability} =$\\
-  $\DNFOf{(\Sequence_1,\Probability_1*\Probability)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n*\Probability)\DNFSep(\SequenceAlt_1,\ProbabilityAlt_1*(1-\Probability))\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m*(1-\Probability))}$
-
-  \medskip
-
-  \AtomToDNF{} \OfType
-  \ArrowTypeOf{\AtomType{}}{\DNFRegexType{}}\\
-  $\AtomToDNFOf{\Atom} = \DNFOf{(\SequenceOf{\EmptyString \SeqSep \Atom \SeqSep
-      \EmptyString},1)}$
-  \caption{DNF Regular Expression Functions}
-  \label{fig:dnf-regex-functions}
-\end{figure}
-
-We have a means of turning stochastic regular expressions into stochastic DNF
-regular expressions.
-The conversion algorithm itself, written $\ToDNFRegexOf{\Regex}$, is defined below.
-\[
-  \begin{array}{rcl}
-    \ToDNFRegexOf{\String} & = & \DNFOf{\SequenceOf{\String}}\\
-    \ToDNFRegexOf{\emptyset} & = & \DNFOf{}\\
-    \ToDNFRegexOf{(\PRegexStar{\Regex}{\Probability})} & = & \AtomToDNFOf{\PRegexStar{(\ToDNFRegexOf{\Regex})}{\Probability}}\\
-    \ToDNFRegexOf{(\RegexConcat{\Regex_1}{\Regex_2})} & = & \ToDNFRegexOf{\Regex_1} \ConcatDNF \ToDNFRegexOf{\Regex_2}\\
-    \ToDNFRegexOf{(\PRegexOr{\Regex_1}{\Regex_2}{\Probability})} & = & \ToDNFRegexOf{\Regex_1} \OrDNF_{\Probability} \ToDNFRegexOf{\Regex_2}\\
-  \end{array}
-\]
-We have also proven an algorithm, demonstrating this conversion respects
-probability distributions.
-
-\begin{theorem}
-  $\ProbabilityOf{\Regex}{\String} = \ProbabilityOf{\ToDNFRegex \Regex}{\String}$
-\end{theorem}
-
-Furthermore, in the situation where a DNF regular expression is unambiguous, we
-have defined an efficient, syntactic means to find the entropy of the DNF
-regular expression.
-
-\begin{center}
-  \begin{tabular}{rcl}
-    $\EntropyOf{\PRegexStar{\DNFRegex}{\Probability}}$
-    & =
-    & $\frac{\Probability}{1-\Probability}(\EntropyOf{\DNFRegex} - \Log_2\Probability)
-      - \Log_2(1-\Probability)
-      $\\
-
-    $\EntropyOf{\SequenceOf{\String_0 \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_n \SeqSep \String_n}}{\String'}$
-    & =
-    & $\Sigma_{i = 1}^n\EntropyOf{\Atom_i}$ \\
-
-    $\EntropyOf{\DNFOf{(\Sequence_1,\Probability_1) \DNFSep \ldots \DNFSep (\Sequence_n,\Probability_n)}}$
-    & =
-    & $\Sigma_{i=1}^n\Probability_i(\EntropyOf{\Sequence_i}+\Log_2\Probability_i)$\\
-  \end{tabular}
-\end{center}
-
-\begin{theorem}
-  $\EntropyOf{\DNFRegex}$ is the entropy of $P_{\DNFRegex}$.
-\end{theorem}
-
-% end DNF regular expressions and lenses
-
-% begin symmetric dnf lenses
-\section{Symmetric DNF Lenses}
-\label{symmetric-dnf-lenses}
-\subsection{Syntax}
-Symmetric DNF Lenses are what we synthesize, and we then turn them into
-Symmetric Lenses. The nonterminals $i$, $j$, $p$, $q$, $r$, $c$, and $d$ all
-represent natural numbers. The nonterminals $\String$ and $\StringAlt$ represent
-strings.
-
-\begin{center}
-  \begin{tabular}{@{}r@{\ }c@{}l@{}}
-    % REGEX
-    \SAtomLens{} & \GEq{} & $\IterateLensOf{\SDNFLens}$ \\
-    \SSQLens{} & \GEq{} & $(\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
-                          \ldots\SeqLSep
-                          (i_p,j_p,\SAtomLens_p)}
-                          ,\ListOf{(\String_1,\Atom_1);\ldots;(\String_q,\Atom_q)}
-                          ,\ListOf{\StringAlt_1;\ldots;\StringAlt_r})$ \\
-    \SDNFLens{} & \GEq{} & $(\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
-                           \ldots\DNFLSep
-                           (i_p,j_p,\SSQLens_p)}
-                           ,\ListOf{c_1;\ldots;c_q}
-                           ,\ListOf{d_1;\ldots;d_r})$ \\
-  \end{tabular}
-\end{center}
-
-Lets go through these from bottom to top.
-
-A DNF Lens \SDNFLens{} consists of three components. First is the mapping
-component. The mapping component is a list of triples $(i,j,\SSQLens)$. In this,
-$i$ and $j$ refer to sequence $i$ and sequence $j$ in the left and right DNF
-regular expressions, respectively. The third component, \SSQLens{}, is a
-sequence lens that maps between the two sequences specified by $i$ and $j$. The
-second component is a list of integers $c$. The integer, $c_i$ at position $i$
-states that when performing a $\CreateR$ on a string matching sequence
-$\Sequence_i$, it will use the sequence lens that maps it to sequence
-$\SequenceAlt_{c_i}$ on the right. The third component is a list of integers
-$d$, and follows the same pattern as the second component, but maps from right
-to left.
-
-A Sequence lens \SSQLens{} consists of three components. First is the mapping
-component. The mapping component is a list of triples $(i,j,\SAtomLens)$. In
-this, $i$ and $j$ refer to atom $i$ and atom $j$ on the left and right
-sequences, respectively. The third component, \SAtomLens{}, is an atom lens that
-maps between the two atoms specified by $i$ and $j$. The second component is a
-list of strings, $\String$. The string, $s_i$ at position $i$ states that when
-performing a $\CreateL$ on a string matching atom $\Atom_i$, it will do one of
-two things. If $\Atom_i$ is mapped to in the mapping component, then performing
-the atom's $\CreateL$ will actually $\PutL$ into that string. If $\Atom_i$ is
-not mapped to, then $\String_i$ will be created as a representative of
-$\Atom_i$.
-
-An atom lens \AtomLens{} is merely the iteration of a symmetric DNF Lens.
-
-\subsection{Typing and Semantics}
-Because these lenses only encode functions on well-typed programs, we introduce
-the semantics alongside the typing rules.  The semantics are over typing
-derivations, instead of over individual lenses.
-
-\paragraph*{DNF Lenses}
-The typing judgment is a 3-ary relation over a single DNF lens, and two DNF
-regular expressions. If $\SDNFLens \OfType \DNFRegex \Leftrightarrow
-\DNFRegexAlt$, then the $\SDNFLens.\CreateR$, $\SDNFLens.\CreateL$,
-$\SDNFLens.\PutR$, and $\SDNFLens.\PutL$ functions form a symmetric lens.
-
-The typing judgement has 3 components. The first is the sublens components,
-confirming that the sequence lenses the DNF lens is comprised of are all
-well-typed. The second guarantees that if each sequence on the left has a
-sequence lens that can be used for \CreateR{}s and \PutR{}s. The last guarantees
-the same for sequences on the right, with \CreateL{} and \PutL{}.
-
-\[
-  \inferrule*
-  {
-    \SSQLens_1 \OfType \Sequence_{i_1} \Leftrightarrow \SequenceAlt_{j_1}\\
-    \ldots\\
-    \SSQLens_p \OfType \Sequence_{i_p} \Leftrightarrow \SequenceAlt_{j_p}\\\\
-    i_{c_1} = 1\\
-    \ldots\\
-    i_{c_q} = q\\\\
-    j_{d_1} = 1\\
-    \ldots\\
-    j_{d_r} = r
-  }
-  {
-    (\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
-      \ldots\DNFLSep
-      (i_p,j_p,\SSQLens_p)}
-    ,\ListOf{c_1;\ldots;c_q}
-    ,\ListOf{d_1;\ldots;d_r})
-    \OfType\\
-    \DNFOf{(\Sequence_1,\Probability_1) \DNFSep \ldots \DNFSep (\Sequence_q,\Probability_q)}
-    \Leftrightarrow
-    \DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1) \DNFSep \ldots \DNFSep (\SequenceAlt_r,\ProbabilityAlt_q)}
-  }
-\]
-
-The \CreateR{} function looks for the sequence the provided string matches. If
-the string matches sequence $\Sequence_x$, then the lens will look in the create
-list at position $x$ to find which sequence lens to use. Then, the sequence lens
-transforms the given string using that sequence lens.
-
-The \PutR{} function finds what pairs sequences the input source and view
-strings match.  If there that pair of sequences have a sequence lens between
-them, then the DNF lens merely performs that sequence lens on the provided
-strings.  If there isn't a pair of sequence lenses between them, then \CreateR{}
-is performed on the source, with the view forgotten.
-
-The \CreateL{} and \PutL{} functions are defined symmetrically.
-
-\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
-  $\CreateR{} \App s$ & = & $\SSQLens_{c_x}.\CreateR{} \App s$ if $s \in \Sequence_x$\\
-  $\CreateL{} \App v$ & = & $\SSQLens_{d_y}.\CreateL{} \App v$ if $v \in \SequenceAlt_y$\\
-  $\PutR{} \App s \App v$ & = &
-                               $\begin{cases*}
-                                 \SSQLens_x.\PutR{} \App s \App v & if $s \in \Sequence_{i_x}$ and $v \in \SequenceAlt_{j_x}$\\
-                                 \CreateR{} \App s & if $\nexists x.$ $s \in \Sequence_{i_x}$ and $v \in \SequenceAlt_{j_x}$
-                               \end{cases*}$\\
-  $\PutL{} \App v \App s$ & = &
-                               $\begin{cases*}
-                                 \SSQLens_y.\PutL{} \App v \App s & if $v \in \SequenceAlt_{j_y}$ and $s \in \Sequence_{i_y}$\\
-                                 \CreateL{} \App v & if $\nexists y.$ $v \in \SequenceAlt_{j_y}$ and $s \in \Sequence_{i_y}$
-                               \end{cases*}$
-\end{tabular}
-
-\paragraph*{Sequence Lenses}
-The typing judgment is a 3-ary relation over a single Sequence lens, and two
-sequences. If $\SSQLens \OfType \Sequence \Leftrightarrow \SequenceAlt$, then
-the $\SSQLens.\CreateR$, $\SSQLens.\CreateL$, $\SSQLens.\PutR$, and
-$\SSQLens.\PutL$ functions form a symmetric lens.
-
-The typing judgement has 4 components. The first is the sublens components,
-confirming that the atom lenses the sequence lens is comprised of all are
-well-typed. The second guarantees that if each string that will be used for
-\CreateR{}s are members of the correct atoms.  The third guarantees
-the same for strings and atoms on the right, with \CreateL.  The last guarantees
-that each atom is mapped by at most one atom lens.
-
-\[
-  \inferrule*
-  {
-    \SAtomLens_1 \OfType \Atom_{i_1} \Leftrightarrow \AtomAlt_{j_1}\\
-    \ldots\\
-    \SAtomLens_p \OfType \Atom_{i_p} \Leftrightarrow \AtomAlt_{j_p}\\\\
-    \String_1 \in \Atom_1\\
-    \ldots\\
-    \String_q \in \Atom_q\\\\
-    \StringAlt_1 \in \AtomAlt_1\\
-    \ldots\\
-    \StringAlt_r \in \AtomAlt_r\\\\
-    i_x = i_y \BooleanImplies x = y\\
-    j_x = j_y \BooleanImplies x = y
-  }
-  {
-    (\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
-      \ldots\SeqLSep
-      (i_p,j_p,\SAtomLens_p)}
-    ,\ListOf{\String_1;\ldots;\String_q}
-    ,\ListOf{\StringAlt_1;\ldots;\StringAlt_r})\\
-    \OfType
-    \SequenceOf{\String_0' \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_q
-      \SeqSep \String_q'}
-    \Leftrightarrow
-    \SequenceOf{\StringAlt_0' \SeqSep \AtomAlt_1 \SeqSep \ldots \SeqSep \AtomAlt_r \SeqSep \StringAlt_r'}
-  }
-\]
-
-For each component of the string matching an atom, the \CreateR function looks
-for the atom lens that maps on the atom. If there is such an atom lens,
-\SAtomLens, then that the sequence lens puts the provided string into the
-default string for the target atom. If there is no such atom lens, then the
-sequence lens merely uses the default string.
-
-For each component of the string matching an atom, the \PutR function looks
-for the atom lens that maps on the atom. If there is such an atom lens,
-\SAtomLens, then that the atom lens puts the provided string of the source atom into the
-string of the target atom. If there is no such atom lens, then the
-sequence lens merely recovers the target's string.
-
-The \CreateL{} and \PutL{} functions are defined symmetrically.
-
-\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
-  $\CreateR{} \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q'$
-  & = 
-  & $\StringAlt_0' \Concat \StringAlt_1'' \Concat \ldots \Concat
-    \StringAlt_r'' \Concat \StringAlt_r'$\\
-  & & where $\StringAlt_y'' =
-    \begin{cases*}
-      \SAtomLens_k.\PutRight \App \String_{i_k}'' \App \StringAlt_y & if $j_k = y$\\
-      \StringAlt_y & if $\nexists k. j_k = y$\\
-    \end{cases*}$\\
-  $\CreateL{} \App \StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q''
-  \Concat \StringAlt_q'$
-  & = 
-  & $\String_0' \Concat \String_1'' \Concat \ldots \Concat
-    \String_r'' \Concat \String_r'$\\
-  & & where $\String_x'' =
-    \begin{cases*}
-      \SAtomLens_k.\PutRight \App \StringAlt_{j_k}'' \App \String_x & if $i_k = x$\\
-      \String_x & if $\nexists k. i_k = x$\\
-    \end{cases*}$\\
-  $\PutR{} \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q' \App \StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q'' \Concat \StringAlt_q'$
-  & =
-  & $\StringAlt_0'\Concat \StringAlt_1''' \Concat \ldots \Concat \StringAlt_q''' \Concat \StringAlt_q'$\\
-  & & where
-      $t_y''' =
-      \begin{cases*}
-        \SAtomLens_k.\PutRight \App \String_{i_k}'' \App \StringAlt_{y}'' & if $j_k = y$\\
-        \StringAlt_y'' & if $\nexists k. j_k = y$\\
-      \end{cases*}$\\
-  $\PutL{} \App \StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q'' \Concat \StringAlt_q' \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q'$
-  & =
-  & $\String_0'\Concat \String_1''' \Concat \ldots \Concat \String_r''' \Concat \String_r'$\\
-  & & where
-      $s_x''' =
-      \begin{cases*}
-        \SAtomLens_k.\PutLeft \App \StringAlt_{j_k}'' \App \String_{x}'' & if $i_k = x$\\
-        \String_x'' & if $\nexists k. i_k = x$\\
-      \end{cases*}$\\
-\end{tabular}
-
-\paragraph*{Atom Lenses}
-The typing judgment is a 3-ary relation over a single atom lens, and two
-atoms. If $\SAtomLens \OfType \Atom \Leftrightarrow \AtomAlt$, then
-the $\SAtomLens.\CreateR$, $\SAtomLens.\CreateL$, $\SAtomLens.\PutR$, and
-$\SAtomLens.\PutL$ functions form a symmetric lens.
-
-The typing judgement just confirms that the DNF lens that comprises the
-sequence lens is also well typed.
-
-\[
-  \inferrule*
-  {
-    \SDNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt
-  }
-  {
-    \IterateLensOf{\SDNFLens}
-    \OfType \PRegexStar{\DNFRegex}{\Probability}
-    \Leftrightarrow \PRegexStar{\DNFRegexAlt}{\ProbabilityAlt}
-  }
-\]
-
-For each component of the string matching an atom, the \CreateR function looks
-for the atom lens that maps on the atom. If there is such an atom lens,
-\SAtomLens, then that the sequence lens puts the provided string into the
-default string for the target atom. If there is no such atom lens, then the
-sequence lens merely uses the default string.
-
-For each component of the string matching an atom, the \PutR function looks
-for the atom lens that maps on the atom. If there is such an atom lens,
-\SAtomLens, then that the atom lens puts the provided string of the source atom into the
-string of the target atom. If there is no such atom lens, then the
-sequence lens merely recovers the target's string.
-
-The \CreateL{} and \PutL{} functions are defined symmetrically.
-
-\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
-  $\CreateR{} \App \String_0 \Concat \ldots \Concat \String_n$
-  & = 
-  & $\StringAlt_1 \Concat \ldots \Concat \StringAlt_n$
-  where $\StringAlt_i = \SDNFLens.\CreateR \App \String_i$\\
-  $\CreateL{} \App \StringAlt_0 \Concat \ldots \Concat \StringAlt_m$
-  & = 
-  & $\String_1 \Concat \ldots \Concat \String_m$
-  where $\String_i = \SDNFLens.\CreateL \App \StringAlt_i$\\
-  $\PutR{} \App \String_0 \Concat \ldots \Concat \String_n \App
-  \StringAlt_1 \Concat \ldots \Concat \StringAlt_m$
-  & = 
-  & $\StringAlt_1' \Concat \ldots \Concat \StringAlt_n'$
-    where $\StringAlt_i' =
-    \begin{cases*}
-      \SDNFLens.\PutR \App \String_i \App \StringAlt_i & if $i \leq m$\\
-      \SDNFLens.\CreateR \App \String_i & otherwise
-    \end{cases*}$\\
-  $\PutL{} \App \StringAlt_1 \Concat \ldots \Concat \StringAlt_m \App
-  \String_0 \Concat \ldots \Concat \String_n$
-  & = 
-  & $\String_1' \Concat \ldots \Concat \String_m'$
-    where $\String_i' =
-    \begin{cases*}
-      \SDNFLens.\PutL \App \StringAlt_i \App \String_i & if $i \leq n$\\
-      \SDNFLens.\CreateL \App \StringAlt_i & otherwise
-    \end{cases*}$\\
-\end{tabular}
-
-Now that we've developed symmetric DNF lenses, it's important to develop a means
-to guage how bijective the lens is.
-
-Instead of searching for \emph{any} term that satisfies the specification, we
-search for the \emph{best} term that satisfies the specification. This begs the
-question, by what metric do we use to determine which term is best. Intuitively,
-we aim to synthesize the lens which is \emph{closest} to being a bijection.
-Somewhat more formally, we aim to synthesize a lens which minimizes the expected
-number of bits which must be used to recover the source from the target, and the
-target from the source. The expected number of bits that are needed to encode
-informatation from a data source is a well known concept in information theory,
-and is known as the \emph{entropy} of the data source.
-
-The entropy of $\DNFRegexAlt$ given $\DNFRegex$ and $\SDNFLens$,
-$\EntropyOf{\DNFRegexAlt \Given \DNFRegex,\SDNFLens}$ is defined symmetrically.
-With these $\operatorname*{argmin}_\theta$ terms defined, the ``best'' DNF lens
-between $\DNFRegex$ and $\DNFRegexAlt$ is $\ArgminOver{\SetOf{\SDNFLens \SuchThat
-    \SDNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt}}
-\EntropyOf{\DNFRegex \Given \DNFRegexAlt,\SDNFLens} + \EntropyOf{\DNFRegexAlt
-  \Given \DNFRegex,\SDNFLens}$.
-
-\begin{centering}
-  \[
-    \begin{array}{rl}
-      & \EntropyOf{\DNFOf{\Sequence_1 \DNFSep \ldots \DNFSep \Sequence_n}\\
-      & \hspace*{1em}\Given
-        \DNFOf{\SequenceAlt_1 \DNFSep \ldots \DNFSep \SequenceAlt_m}\\
-      & \hspace*{1.21em},~(\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
-        \ldots\DNFLSep
-        (i_p,j_p,\SSQLens_p)}
-        ,\ListOf{c_1;\ldots;c_q}
-        ,\ListOf{d_1;\ldots;d_r})}\\
-      =\\
-      & \frac{\Sigma_{j=1}^m(\frac{\Sigma_{k\SuchThat j_k = j}
-        \EntropyOf{\Sequence_{i_k}\Given\SequenceAlt_j,\SequenceLens_k}}{\SizeOf{\SetOf{k\SuchThat
-        j_k = j}}} + log_2\SizeOf{\SetOf{k\SuchThat j_k = j}})}{m}\\\\
-      \\
-
-      & \EntropyOf{
-        \SequenceOf{\String_0,\Atom_1,\ldots,\Atom_n,\String_n}\\
-      & \hspace*{1em}\Given
-        \SequenceOf{\StringAlt_0,\AtomAlt_1,\ldots,\AtomAlt_m,\StringAlt_m}\\
-      & \hspace*{1.21em},~
-        (\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
-        \ldots\SeqLSep
-        (i_p,j_p,\SAtomLens_p)}
-        ,\ListOf{(k_1,\String_1);\ldots;(k_q,\String_q)}
-        ,\ListOf{(l_1,\StringAlt_1);\ldots;(l_r,\StringAlt_r)})}\\
-      =\\
-      & \Sigma_{x=1}^p\EntropyOf{\Atom_{i_x} \Given \AtomAlt_{i_x},\AtomLens_x} +
-        \Sigma_{x=1}^q\EntropyOf{\Atom_{k_x}}\\\\
-      \\
-
-      & \EntropyOf{\StarOf{\DNFRegex} \Given \StarOf{\DNFRegexAlt},\StarOf{\SDNFLens}}\\
-      =\\
-      & 4*\EntropyOf{\DNFRegex\Given\DNFRegexAlt,\SDNFLens}
-    \end{array}
-  \]
-\end{centering}
-
-% end symmetric dnf lenses
-
-% begin synthesis
-\section{Synthesis}
-\label{sec:synthesis}
-While examples can be used to tighten the
-specification, a large number of well-constructed examples are required to
-sufficiently constrain the search space.
-
-Unfortunately, trying to find the best symmetric lens is a daunting task!
-Unless we find a bijective lens (in which case the sum is zero), we can never be
-sure if we will find a better lens than the one we found.
-% end synthesis
-
-% begin implementation
-\section{Implementation}
-\label{sec:implementation}
-% end implementation
-
-% begin evaluation
-\section{Evaluation}
-\label{sec:evaluation}
-% end evaluation
-
-% begin related-work
-\section{Related Work}
-\label{sec:related}
+
 % end related-work

### 2018-05-24 papers/SymmetricLenses/main.tex
add better title
better examples

@@
 
-\title{Simple Symmetric Lenses - Foundations and Synthesis}
+\title{Synthesizing Symmetric Lenses}
 
@@
 % state of the world
-Similar data can come in varieties of formats: bibliography information can be
-stored in the BibTeX, EndNote, or ACM reference formats; web service resources
-can be provided to clients in legacy or modern formats; calendar events can be
+Similar data can come in varieties of formats: scheduled jobs can be stored as
+cron jobs for Linux or as Scheduled Tasks for Windows; web service resources can
+be provided to clients in legacy or modern formats; calendar events can be
 stored in CSV files or ICS files. Oftentimes, we want to synchronize files
-across formats: when my collaborator edits her bibliography stored in the ACM
-reference format, I want these edits reflected in my EndNote file; when I update
-data through a legacy API, I want those changes to be reflected when I retrieve
-it through a modern API; when I create a new calendar event in my ICS file, I
-want my CSV calendar files to be updated to reflect this change.
+across formats: when I edit a task to run more often on my Linux machine, I want
+this task to run more often on my Windows machine; when I update data through a
+legacy API, I want those changes to be reflected when I retrieve it through a
+modern API; when I create a new calendar event in my ICS file, I want my CSV
+calendar files to be updated to reflect this change.
 
@@
 always true. Oftentimes, fields present in one format are not present in the
-other: updated APIs can expose new information, or omit legacy information; the
-ACM reference format includes article number information, whereas EndNote
-includes page number information; ICS files contain a URL field that Google
-Calendar CSV files do not expose. We aim to synthesize transformations where the
-data fields are not necessarily in bijective correspondence, where each data
-format can contain important information not present in the other format.
+other: the commands scheduled to run on Linux machines are typically different
+than the commands scheduled to run on Windows machines; updated APIs can expose
+new information, or omit legacy information; ICS files contain a URL field that
+Google Calendar CSV files do not expose. We aim to synthesize transformations
+where the data fields are not necessarily in bijective correspondence, where
+each data format can contain important information not present in the other
+format.
 
@@
   while preserving probability distributions and
-  languages~(\S\ref{sec:preliminaries}).
+  languages~(\S\ref{sec:stoch-rx}).
 \item We develop stochastic DNF regular expressions and DNF lenses. We develop a
@@
 
-% begin preliminaries
-\section{Preliminaries}
-\label{sec:preliminaries}
-To define the entropy for data recovery, we must first develop a probability
-model for our languages. We do this by enhancing regular expressions into
-\emph{stochastic regular expressions}. With this, we can jointly
-express both a language and a probability distribution over that language.
+% begin extended example
+\section{Extended Example}
+We motivate our problem with an extended example.  Consider the previously
+described scenario of keeping Linux and Windows scheduled jobs synchronized.
+% end extended example
+
+% begin stoch-rx
+\section{Stochastic Regular Expressions}
+\label{sec:stoch-rx}
+Important to our algorithm is finding the expected number of bits needed to
+recover a string in one data source from a synchronized string in the other data
+source. To do this, we must first develop a probability model for our language.
+We do this with \emph{stochastic regular expressions}, regular expressions
+annotated with probability information. With this, we can jointly express a
+language and a probability distribution over that language.
 
@@
 \end{theorem}
-
-% end preliminaries
+% end stoch-rx
 
@@
 \label{sec:ssl}
-Classically, asymmetric lenses comprised three functions, \Put{}, \Get{}, and
-\Create{}.  We extend this style of lens to symmetric lenses with
-\emph{simple symmetric lenses}.  Like the classical formulation of asymmetric
-lenses, simple symmetric lenses have no complement.  Simple symmetric
-lenses comprise 4 functions, \CreateR, \CreateL, \PutR, and \PutL.
-These functions satisfy the following round-tripping laws:
+Classically, asymmetric lenses comprise three functions, \Put{}, \Get{}, and
+\Create{}. These lenses only allow information to be lost on one side. We
+enhance this style of lens to symmetric lenses with \emph{simple
+  symmetric lenses}. Like the classical formulation of asymmetric lenses, simple
+symmetric lenses have no complement. Simple symmetric lenses comprise 4
+functions, \CreateR, \CreateL, \PutR, and \PutL. These functions satisfy the
+following round-tripping laws:
 \begin{equation}

### 2018-05-25 papers/SymmetricLenses/main.tex
push a version

@@
 \usepackage{amsmath}
+\usepackage{nccmath}
 \usepackage{mathtools}
@@
 
+\setlength{\belowdisplayskip}{0pt} \setlength{\belowdisplayshortskip}{0pt}
+\setlength{\abovedisplayskip}{0pt} \setlength{\abovedisplayshortskip}{0pt}
+
 \usetikzlibrary{
@@
 \acmDOI{10.1145/3158089}
-\acmYear{2018}
-\copyrightyear{2018}
+\acmYear{2019}
+\copyrightyear{2019}
 \acmJournal{PACMPL}
-\acmVolume{2}
+\acmVolume{3}
 \acmNumber{POPL}
@@
 
-We develop an algorithm that synthesizes a low-cost lens. Our algorithm consists
-of two communicating synthesizers, one that proposes candidate stochastic
-regular expression pairs that it thinks will align well, and one that greedily
-searches for a lens between those two regular expressions that minimizes cost
-and involves no rewrites.
-
-Our first algorithm, \TypeProp, proposes candidate stochastic regular
+\begin{figure}
+  \includegraphics[width=.5\textwidth]{high-level-algorithm.pdf}
+  \caption{Schematic Diagram for \SOptician. Regular expressions \Regex and
+    \RegexAlt, and a set of examples \Examples, are provided as input. \RXSearch
+    searches through stochastic regular expression pairs, equivalent to the
+    originals, and proposes them to \GreedySynth. \GreedySynth finds a lens,
+    typed between the generated equivalent pairs. When \RXSearch determines it
+    has an optimal lens, it returns that lens.}
+  \label{fig:high-level-algorithm}
+\end{figure}
+
+We develop an algorithm that synthesizes a low-cost lens, shown in
+Figure~\ref{fig:high-level-algorithm}. Our algorithm consists of two
+communicating synthesizers, one that proposes candidate stochastic regular
+expression pairs that it thinks will align well, and one that greedily searches
+for a lens between those two regular expressions that minimizes cost and
+involves no rewrites. When the first algorithm thinks it has found the best
+possible lens, it returns that lens.
+
+Our first algorithm, \RXSearch, proposes candidate stochastic regular
 expressions. These candidate stochastic regular expressions are generated by
@@
 \section{Extended Example}
+
+\begin{figure}
+  % https://tex.stackexchange.com/questions/168741/showing-two-listings-in-a-table-side-by-side
+  \setbox0=\hbox{%
+    \begin{minipage}{3in}
+\begin{lstlisting}[
+tabsize=1,
+language={xml},
+]
+0,20,40 cp -r /Library/ "$\$$(date)"
+
+* df -h | sendmail admin@domain.com
+\end{lstlisting}
+    \end{minipage}
+  }
+  \savestack{\listingA}{\box0}
+
+  \setbox0=\hbox{%
+    \begin{minipage}{3in}
+\begin{lstlisting}[
+tabsize=1,
+]
+<Tasks>
+ <Task>
+  <Command>c</Command>
+  <Time>0</Time><Time>20</Time><Time>40</Time>
+ </Task>
+ <Task>
+  <Command>c</Command>
+  <AllTimes/>
+ </Task>
+<Tasks>
+
+{
+  Title: Return 400 on
+    bad PUT request,
+  Type: Bug,
+  AssignedTo: J Smith,
+}
+\end{lstlisting}
+    \end{minipage}
+  }
+  \savestack{\listingB}{\box0}
+  \centering
+  \resizebox{\columnwidth}{!}{%
+    \begin{tabular}{|>{\columncolor{lightyellow}}c|>{\columncolor{lightgreen}}c|}
+      \hline
+      Legacy Work Item Representation & Modern Work Item Representation \\
+      \hline
+      \listingA & \listingB \\
+      \hline
+    \end{tabular}
+  }
+  \caption{Hello there this is the thing}
+  \label{fig:minimized-representations}
+\end{figure}
+
 We motivate our problem with an extended example.  Consider the previously
-described scenario of keeping Linux and Windows scheduled jobs synchronized.
+described example of keeping Linux and Windows scheduled jobs synchronized.
+In this example, we describe keeping synchronized an minified version of
+crontab and the task scheduler.  These are each formats specifying when to run
+recurring tasks.  For simplicity, we are using an minified version of these
+configuration files where we can schedule tasks to occur at fixed intervals,
+hourly, daily, or weekly.  Each task consists of a command to be run.
+
+In this example, we wish for the jobs to be synchronized across our windows and
+linux devices. However, the commands that need to be run are typically different
+for each device. Paths are different in windows and Linux, and while there is
+typically a Windows equvialent for Linux commands, that equivalent command is
+not always named the same thing.
+
+To synchronize these files, we wish to synchronize how often the commands are
+run, but not synchronize the names of the commands. We consider 4 functions to
+be necessary to synchronize these two files: a function that makes a new cron
+file from a newly created task scheduler, a function that makes a new task
+scheduler file from a new cron file, a function that merges edits to the cron
+file to the synchronized task scheduler file, and a function that merges edits
+on the task scheduler file to the cron file. These are the only four functions
+we need to synchronize these two files, so these four functions comprise a
+simple symmetric lens between cron files and task scheduler files.  More
+formally, a simple symmetric lens between $X$ and $Y$ comprise the following
+four functions:
+\begin{enumerate}
+\item $\CreateR \OfType X \rightarrow Y$
+\item $\CreateL \OfType Y \rightarrow X$
+\item $\PutR \OfType X \rightarrow Y \rightarrow Y$
+\item $\PutL \OfType Y \rightarrow X \rightarrow X$
+\end{enumerate}
+
+Furthermore, we expect these functions to satisfy certain properties.  For
+example, if two files are synchronized by one of these functions, using another
+function to resynchronize it will not change the result.  These are viewed as
+the ``round-tripping'' laws:
+\vspace*{-1em}
+\begin{gather}
+  \tag{\CreatePutRL}
+  \PutLOf{(\CreateROf{x})}{x} = x\\
+  \tag{\CreatePutLR}
+  \PutROf{(\CreateLOf{y})}{y} = y\\
+  \tag{\PutRL}
+  \PutLOf{(\PutROf{x}{y})}{x} = x\\
+  \tag{\PutLR}
+  \PutROf{(\PutLOf{y}{x})}{y} = y
+\end{gather}
+
+\vspace*{-1em}
+We wish 
 % end extended example
@@
   \begin{tabular}{rcl}
-    $\ProbabilityOf{\String}{\String'}$
+    $\ProbabilityOf{\String}{\String''}$
     & =

### 2018-05-29 papers/SymmetricLenses/main.tex
add data

@@
 
+\lstset{ language=Caml, basicstyle=\upshape\sffamily,
+keywordstyle=\upshape\sffamily\color{dkpurple}, keepspaces=true,
+framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=true,
+commentstyle=\itshape\rmfamily,
+emph={rep,iterate,synth,collapse,perm,squash,normalize,using,ins,del,lens,let,get,put,rquot,lquot,id,swap,concat,or,disc,merge_left,merge_right,const},
+emphstyle=\upshape\sffamily\color{dkpurple}, 
+columns=fullflexible,
+mathescape, 
+xleftmargin=1.5em,
+% BCP: I find this distracting:
+stringstyle=\sffamily\color{dkblue},
+}
+\makeatletter
+     \let\lst@oldvisiblespace\lst@visiblespace
+     \def\lst@visiblespace{\,\lst@oldvisiblespace\,}
+\makeatother
+
 \setlength{\belowdisplayskip}{0pt} \setlength{\belowdisplayshortskip}{0pt}
@@
 % Creates a display mode for code in sans serif font
-\lstnewenvironment{sflisting}[1][]
-  {\lstset{%
-    mathescape,
-    basicstyle=\small\sffamily,
-    aboveskip=5pt,
-    belowskip=5pt,
-    columns=flexible,
-    frame=,
-    xleftmargin=1em,#1}\sansmath}
-  {}
 % end
@@
 \begin{abstract}
-  Similar data is oftentimes represented and stored in different formats. Users
-  commonly need to synchronize the data shared between formats, where edits to
-  the data in one format gets propagated to the data in the other format.
-  Manually writing these data synchronizers is tedious and error prone. To
-  alleviate these difficulties, tools have been developed that automatically
-  generate these synchronization functions by generating bidirectional lens
-  programs from data format descriptors~\cite{optician,quotient-optician}.
-  However, these tools only work when the two formats describe exactly the same
-  data; it assumes that all important information present in one format is
-  present in the other. This is often not the case.
-
-  In this work, we automatically synthesize conversion functions between formats
-  that each may contain information the other lacks. We define \emph{simple
-    symmetric lenses}, a class of bidirectional programs for expressing these
-  types of conversion functions, and we specify a language that expresses simple
-  symmetric lenses that operate on string formats. We develop a type-directed
-  synthesis algorithm for generating lenses in this language from regular
-  expressions describing the data formats. A significant technical challenge in
-  this work arises from the fact that our target language permits
-  synchronization between formats that don't have exactly the same data -- data
-  in one format need not be present in the other format. This gives the
+  Users commonly need to synchronize the data shared between formats, so edits
+  to the data in one format gets propagated to the data in the other format.
+  Manually writing the functions for these data synchronizers is tedious and
+  error prone. Instead of manually writing these synchronizers, bidirectional
+  languages can be used to express \emph{lenses}, groups of synchronization
+  functions with guarantees about round-trip behavior. While these bidirectional
+  languages help make programming these synthesizers less error-prone, they
+  remain difficult to program in. Synthesizing bidirectional languages allows
+  users to use lenses without manually programming them. Current
+  synthesizers work only on the simplest types of lenses, when the two formats
+  describe exactly the same data; this is often not the case.
+
+  Oftentimes, each format may contain information the other lacks. Such
+  synchronization tasks are described by one of the richest classes of lenses,
+  symmetric lenses. However, symmetric lenses in their current formulation
+  require external state in the form of a \emph{complement}. This statefulness
+  makes symmetric lenses both difficult to use and difficult to synthesize.
+
+  In this work, we reformulate symmetric lenses as \emph{simple symmetric
+    lenses}. While this formulation is less expressive than symmetric lenses, it
+  expresses many symmetric lenses while requiring \emph{no external state}. We
+  provide combinators for simple symmetric lenses on string data, and integrate
+  these combinators into Boomerang while providing full backwards compatibility.
+
+  We develop a type-directed synthesis algorithm for generating lenses in this
+  language from regular expressions describing the data formats. A significant
+  technical challenge in this work arises from the fact that our target language
+  permits synchronization between formats that don't have exactly the same data
+  -- data in one format need not be present in the other format. This gives the
   synthesis algorithm much more freedom, as data can be completely
@@
   algorithm that tries to minimize unsynchronized data. We experimentally
-  demonstrate that despite addressing a much harder problem, our extension to
-  Optician still maintains synthesis times competitive to Optician on Optician's
-  bijective benchmark suite, while generalizing to benchmarks synchronizing data
-  not in bijective correspondence.
+  demonstrate that our algorithm can efficiently synthesize simple symmetric
+  lenses on a variety of benchmark suites consisting of synchronizing file
+  formats from configuration files, application-specific data storage files, and
+  data cleaning tasks.
+
 \end{abstract}
@@
 Bidirectional languages ease the difficulty of writing such synchronizers, as a
-single, well-typed program expresses multiple conversion functions, and
-guarantee these functions satisfy round-tripping laws. However, bidirectional
-languages are difficult to program in as well -- often requiring users to
-permute data in fiddly ways while satisfying the type system's complex
-unambiguity constraints.
+single, well-typed program expresses a \emph{lens}, a group of conversion
+functions guaranteed to satisfy round-tripping laws. However, bidirectional
+languages are difficult to program in as well, often requiring users to permute
+data in fiddly ways while satisfying the type system's complex unambiguity
+constraints.
 
 A better approach is to synthesize such conversion functions. Recent research
-has gone into automatically synthesizing such conversion functions from
-input-output examples and formal descriptions of the data
+has gone into Optician, a tool for automatically synthesizing such conversion
+functions from input-output examples and formal descriptions of the data
 formats~\cite{optician,quotient-optician}. However, this work assumes that all
@@
 new information, or omit legacy information; ICS files contain a URL field that
-Google Calendar CSV files do not expose. We aim to synthesize transformations
-where the data fields are not necessarily in bijective correspondence, where
-each data format can contain important information not present in the other
-format.
-
-We follow a similar approach to existing work on synthesizing conversion
-functions; instead of synthesizing these functions individually, we synthesize
-bidirectional programs. To this end, we develop a new formulation of
-bidirectional programs, \emph{simple symmetric lenses}, which permit
-synchronizations between formats not in bijective correspondence. We develop a
-language for expressing simple symmetric lenses that operate on string data.
-Terms in this language are typed between regular expressions, where the regular
-expressions describe the synchronized data formats.
-
-Our synthesis algorithm takes as input a suite of input-output examples and two
-regular expressions describing the data formats. Our synthesis algorithm uses
-type-directed synthesis on the two regular expressions to synthesize a lens that
-is well-typed between the provided regular expressions and satisfies the
-provided input-output examples.
-
-A primary novel difficulty is that, unlike previous work where there are
-relatively few well-typed lenses, there are a relatively many well-typed simple
+Google Calendar CSV files do not expose.
+
+\emph{Symmetric lenses} formalize such conversion functions. However, these
+symmetric lenses use external state for synchronization, in the form of a
+\emph{complement}. This makes them difficult to use in real-world
+synchronization tasks, as we must maintain the files we synchronize as well as
+maintaining the complements. The complement further complicates synthesis, as
+input-output examples are no longer sufficient, instead edit sequences must be
+provided.
+
+To get around this, we reformulate symmetric lenses as \emph{simple symmetric
+  lenses}. This reformulation does not use a complement, and only uses the files
+that we aim to synchronize as state. We identify a property,
+\emph{forgetfulness}, where if a symmetric lens can be written as a forgetful
+symmetric lens, it can also be written a simple symmetric lens. We provide
+combinators for building simple symmetric lenses on string data, and integrate
+these combinators with Boomerang, a language for bidirectional programming on
+strings. In Boomerang, and our extension to it, lens expressions are typed
+between regular expressions, where the regular expressions describe the
+synchronized data formats.
+
+We enhance Optician with simple symmetric lens synthesis capabilities. This
+extension synthesizes symmetric lenses from a suite of input-output examples and
+a pair of regular expressions describing the data formats. The synthesis
+algorithm uses type-directed synthesis on the two regular expressions to
+synthesize a lens that is well-typed between the provided regular expressions
+and satisfies the provided input-output examples.
+
+A primary difficulty is that there are a relatively many well-typed simple
 symmetric lenses. As symmetric lenses permit data to be present in only one
@@
 
-As we wish to synchronize as much data as we can, our cost model gives higher
-cost to those with more data is left unsynchronized. In particular, the cost of
-a lens is the expected number of bits required to recover a string in one format
-when a synchronized string in the other format is provided. In this model,
-bijective lenses have zero cost, and as more information is left unsychronized,
-the cost is increased.
-
-Our model for this is based on \emph{stochastic regular expressions}~\cite{?}.
-We represent a probability distribution for strings by a stochastic regular
-expression, and find a way to syntactically infer the expected information
-content (or \emph{entropy}) of an unambiguous regular expression. We use this
-entropy to calculate the expected information content to recover a string, when
-we are given a string it is synchronized to.
+In particular, the cost of a lens is the expected number of bits required to
+recover a string in one format when a synchronized string in the other format is
+provided. In this model, bijective lenses have zero cost, and as more
+information is left unsychronized, the cost is increased.
+
+Our model for this is based on \emph{stochastic regular expressions}~\cite{?}. A
+probability distribution for strings is represented by a stochastic regular
+expression, and we develop a way to syntactically infer the expected information
+content (or \emph{entropy}) of an unambiguous regular expression. This entropy
+is used to calculate the expected information content to recover a string, when
+given a string it is synchronized to.
 
@@
 We develop an algorithm that synthesizes a low-cost lens, shown in
-Figure~\ref{fig:high-level-algorithm}. Our algorithm consists of two
+Figure~\ref{fig:high-level-algorithm}. This algorithm consists of two
 communicating synthesizers, one that proposes candidate stochastic regular
@@
   as new, symmetric benchmarks, and find that we lose very little synthesis
-  time, while gaining a large amount of expressivity (\S\ref{sec:evaluation}).
+  time, while being significantly more expressive (\S\ref{sec:evaluation}).
 \end{enumerate}
@@
 
-% begin extended example
-\section{Extended Example}
+% begin overview
+\section{Overview}
+\label{sec:overview}
 
@@
   \setbox0=\hbox{%
-    \begin{minipage}{3in}
+    \begin{minipage}{3.5in}
 \begin{lstlisting}[
-tabsize=1,
-language={xml},
+stringstyle=\upshape\sffamily
 ]
@@
   \setbox0=\hbox{%
-    \begin{minipage}{3in}
+    \begin{minipage}{3.5in}
 \begin{lstlisting}[
-tabsize=1,
+stringstyle=\upshape\sffamily
 ]
@@
   <Command>c</Command>
-  <Time>0</Time><Time>20</Time><Time>40</Time>
+  <Time>0</Time><Time>20</Time> <Time>40</Time>
  </Task>
@@
 <Tasks>
-
-{
-  Title: Return 400 on
-    bad PUT request,
-  Type: Bug,
-  AssignedTo: J Smith,
-}
 \end{lstlisting}
@@
   \savestack{\listingB}{\box0}
+
+  \setbox0=\hbox{%
+    \begin{minipage}{3.5in}
+\begin{lstlisting}
+let minute_list = minute . ("," . minute)$^*$ in
+let times = "*" | minute_list in
+let line = times . " " . command in
+let crontab = "" | line . ("\n\n" line)*
+\end{lstlisting}
+    \end{minipage}
+  }
+  \savestack{\listingC}{\box0}
+
+  \setbox0=\hbox{%
+    \begin{minipage}{3.5in}
+\begin{lstlisting}
+let minute_xml = "<Time>" . minute . "</Time>" in
+let minutes_xml = minute_xml . minute_xml$^*$ in
+let xml_time = "<AllTimes/>" | minutes_xml in
+let command_xml = 
+  "<Command>" . command . "</Command>"
+in
+let xml_task = command_xml . xml_time in
+let scheduled_job = "<Tasks>" . xml_task$^*$ . "</Tasks>"
+\end{lstlisting}
+    \end{minipage}
+  }
+  \savestack{\listingD}{\box0}
+
   \centering
@@
       \hline
-      Legacy Work Item Representation & Modern Work Item Representation \\
+      crontab & scheduled.job \\
       \hline
@@
       \hline
-    \end{tabular}
-  }
-  \caption{Hello there this is the thing}
+      \listingC & \listingD \\
+      \hline
+    \end{tabular}
+  }
+  \caption{An example crontab and Scheduled Tasks file, and regular expressions
+    describing the formats. For simplicity, these are modified versions of the
+    real formats.}
   \label{fig:minimized-representations}
@@
 We motivate our problem with an extended example. Consider the previously
-described example of keeping Linux and Windows scheduled jobs synchronized.
-In this example, we describe keeping synchronized an minified version of
-crontab and the task scheduler.  These are each formats specifying when to run
-recurring tasks.  For simplicity, we are using an minified version of these
-configuration files where we can schedule tasks to occur at fixed intervals,
-hourly, daily, or weekly.  Each task consists of a command to be run.
+described example of keeping Linux and Windows scheduled jobs synchronized. In
+this example, we describe keeping synchronized a crontab (the file containing
+cron jobs) and a Scheduled Tasks file, scheduled.job. These are each formats
+specifying when to run recurring tasks. For simplicity of description, we use a
+minified version of each of these, with examples and regular expressions shown
+in Figure~\ref{fig:minimized-representations}.
+
+For task is set in these files at specific times each hour. In both these
+formats, users can specify specific minutes they want to run the program during,
+or they can specify they want to run the program at every minute.
 
@@
 linux devices. However, the commands that need to be run are typically different
-for each device. Paths are different in windows and Linux, and while there is
+for each device. Paths are different in Windows and Linux, and while there is
 typically a Windows equvialent for Linux commands, that equivalent command is
-not always named the same thing.
-
-To synchronize these files, we wish to synchronize how often the commands are
-run, but not synchronize the names of the commands. We consider 4 functions to
-be necessary to synchronize these two files: a function that makes a new cron
-file from a newly created task scheduler, a function that makes a new task
-scheduler file from a new cron file, a function that merges edits to the cron
-file to the synchronized task scheduler file, and a function that merges edits
-on the task scheduler file to the cron file. These are the only four functions
-we need to synchronize these two files, so these four functions comprise a
-simple symmetric lens between cron files and task scheduler files.  More
-formally, a simple symmetric lens between $X$ and $Y$ comprise the following
-four functions:
+not always named the same thing.  So, while we want to be able to add commands
+and edit how often they are run, we don't want to synchronize which commands are
+run themselves.
+
+To synchronize these files, a few functions are needed. Two \emph{creates} are
+needed; if a crontab is initialized, a synchronized Scheduled Tasks file is also
+created, and vice versa. Furthermore, we require \emph{puts}, functions that
+take an updated crontab and an old Scheduled Tasks file, and output a new
+Scheduled Tasks file with the information synchronized to the updated crontab.
+To have a simple symmetric lens is to have these required synchronization
+functions. Formally, a simple symmetric lens between $X$ and $Y$ comprises 4
+functions:
 \begin{enumerate}
@@
 \end{enumerate}
-
-Furthermore, we expect these functions to satisfy certain properties.  For
-example, if two files are synchronized by one of these functions, using another
-function to resynchronize it will not change the result.  These are viewed as
-the ``round-tripping'' laws:
-\vspace*{-1em}
+Furthermore, if a file is not edited, but is then resynchronized using a put
+function, no data should be changed. To be a lens, one must satisfy the
+following ``round-tripping'' laws that guarantee identity edits won't change the
+file data.
 \begin{gather}
@@
 
-\vspace*{-1em}
-We wish 
-% end extended example
+For synchronization, when the original crontab or Scheduled Tasks file is
+created, a synchronized file of the other form is created from it with
+$\CreateR$ or $\CreateL$ (depending on which file was initially created). On
+future edits, changes are made with the puts, a \PutR would propagate the shared
+information from the crontab, and knit it together with the information specific
+to the Scheduled Tasks file.
+
+\subsection{Programming Symmetric Lenses} We now build up a symmetric lens by
+example. Our symmetric lens expressions are typed between regular expression
+pairs, if \lstinline{$\Lens$ : $\Regex \Leftrightarrow \RegexAlt$}, then $\Lens$
+encodes a lens between $\LanguageOf{\Regex}$ and $\LanguageOf{\RegexAlt}$. At
+the end, we want our symmetric lenses to by typed between \lstinline{crontab}
+and \lstinline{scheduled_job}.
+
+The identity lens takes, as argument, a regular expression, and performs the
+identity function on strings in that regular expression. The identity lens fully
+synchronizes the data formats: when one data format is updated, the other format
+gets the exact same value. In our example, when a job is told to run at a
+specific time in crontab, it is also expected to run at that same time in the
+Scheduled Jobs list, so the identity lens would propagate that information well.
+%
+\begin{lstlisting}
+id(minute) : minute $\Leftrightarrow$ minute
+\end{lstlisting}
+%
+
+Where the identity lens synchronizes data, the disconnect lens does the
+opposite: it leaves data unsynchronized. The disconnect lens takes 4 arguments:
+two regular expressions and two strings. On creates, because there is no
+association between data in the first format and data in the second format,
+designated default strings $\String$ and $\StringAlt$ are put into the other
+format. In our extended example, the Linux commands should not by synchronized
+with the Windows commands, so a disconnect lens would be used. We want our
+default behavior to be failures, under the assumption this failure will raise a
+warning, in case somebody forgets to find the analogous command of the other
+operating system.
+%
+\begin{lstlisting}
+disc(command,command_xml,"false","<Command>SET ERRORLEVEL=1</Command>") 
+  : command $\Leftrightarrow$ command_xml
+\end{lstlisting}
+%
+When building a disconnect lens between two strings, one of which is merely the
+empty string, the lenses \lstinline{ins} and \lstinline{del} can instead be
+used, where \lstinline{ins($t$) = disc("",$t$,"",$t$)} and
+\lstinline{del($s$) = disc($s$,"",$s$,"")}.
+
+Finally, there are a number of lenses that compositionally build up new lenses
+from old ones. For example, data can appear in one form or another, a program
+can run during a specific set of minutes, or that program can run the every
+minute. If \lstinline{minutes_map} and \lstinline{alltimes_map} are lenses that
+synchronize those settings in each format, then those lenses can be composed
+with an \OrLens lens.
+%
+\begin{lstlisting}
+or(minutes_map,alltimes_map) : times $\Leftrightarrow$ xml_time
+\end{lstlisting}
+
+Data can also be concatenated, and so lens mappings can be concatenated.
+However, data concatenations can only happen in a number of ways, as data can
+appear in the same order, and data can be reordered. For example, in the
+crontab, time data appears before command data, whereas in the Scheduled Jobs
+file, command data comes before time data. If \lstinline{command_map} provides a
+lens between commands in each format, and \lstinline{time_map} provides a lens
+between time data in each format, those lenses can be composed with a \SwapLens
+lens.
+%
+\begin{lstlisting}
+swap(command_map,time_map) : line $\Leftrightarrow$ xml_task
+\end{lstlisting}
+%
+Two lenses, $\Lens_1$ and $\Lens_2$, can be concatenated as
+\lstinline{concat($\Lens_1$,$\Lens_2$)} or as the infix style
+\lstinline{$\Lens_1$ . $\Lens_2$} for data that is not reordered.
+
+Lastly, the \IterateLens lens synchronizes repeated data, synchronizing each
+element of the iteration, and synchronizing the number of iterations. For
+example, if \lstinline{full_task_map} synchronizes crontab lines prepended by
+two newlines and xml for a single task, iterating that lens builds a lens that
+synchronizes a list of tasks.
+%
+\begin{lstlisting}
+iterate(full_task_map) : ("\n\n" . line)$^*$ $\Leftrightarrow$ xml_task$^*$
+\end{lstlisting}
+%
+
+\begin{figure}
+\begin{lstlisting}
+let full_minute_map = ins("<Time>") . id(minute) . ins("</Time>") in
+let comma_minute_map = del(",") . full_minute_map in
+let minutes_map = full_minute_map . iterate(comma_minute_map) in
+let alltimes_map = del("*") . ins("<AllTimes/>") in
+let times_map = or(minutes_map,alltimes_map) in
+
+let command_core = 
+  disc(command,command_xml,"false","<Command>SET ERRORLEVEL=1</Command>")
+in
+let command_map = 
+  ins("<Command>") . command_core . ins("</Command>")) . del(" ") 
+in
+
+let task_map = swap(command_map,time_map) in
+let full_task_map = 
+  ins("<Task>") . del("\n\n") . task_map . ins("</Task>") 
+in
+let repeated_task_map = iterate(full_task_map)
+let nonempty_task_map = ins("<Tasks>") . repeated_task_map . ins("</Tasks>")
+
+or(ins("<Tasks></Tasks>"),nonempty_task_map) : crontab $\Leftrightarrow$ scheduled_job
+\end{lstlisting}
+  \caption{A lens, written in our language for building simple symmetric lenses
+    on string data, that synchronizes crontab files and Scheduled Task files.
+    Note that we can annotate our lenses with types, to confirm the lens
+    operates on the languages of the regular expressions.}
+  \label{fig:example_lens}
+\end{figure}
+
+Using these combinators, we can build up a lens that synchronizes crontabs and
+Scheduled Task files, shown in Figure~\ref{fig:example_lens}. This lens uses one
+typing rule not previously mentioned: type equivalence.  While
+\lstinline{scheduled_job} is a equivalent to a lens with an outermost
+disjunction, it is not syntactically equivalent to one.  Without type
+equivalence, the provided lens wouldn't actually be well-typed, but type
+equivalence states that if a lens is typed between a pair of regular
+expressions, it is also typed between an equivalent pair.
+
+Note that, while the programmed lens should be quite simple, it was relatively
+difficult to write. We had to provide a lot of scaffolding to map between the
+desired formats, even though the idea was quite simple: synchronize the times
+but leave the commands unsynchronized. Furthermore, this is inherent to manually
+writing these synchronization functions, as a programmer must describe which
+components are synchronized with which, and which components are scaffolding.
+
+\subsection{Synthesizing Symmetric Lenses}
+To alleviate these difficulties, we wish to synthesize symmetric lenses from
+types and examples. Given a pair of regular expressions and a set of
+input-output examples, we want to synthesize a symmetric lens expression that is
+typed between the regular expression pair, and satisfies the input/output
+examples. For example, we wish to synthesize a lens from \lstinline{crontab} and
+\lstinline{scheduled_job}, using the strings in
+Figure~\ref{fig:minimized-representations} as input-output example for \CreateR.
+
+However, our simple symmetric lens combinators permit many well-typed programs
+between two regular expressions. For example, one possible lens between
+\lstinline{crontab} and \lstinline{scheduled_job} is the one in
+Figure~\ref{fig:example_lens}, but the lens
+%
+\begin{lstlisting}
+disc(crontab,scheduled_job,rep(crontab),rep(scheduled_job))
+\end{lstlisting}
+%
+is also a valid lens (where \lstinline{rep($\Regex$)} is a string in the
+language of $\Regex$). Furthermore, adding the synchronized files from
+Figure~\ref{fig:minimized-representations} as an input-output example helps
+little, but the default strings given to \Disconnect merely need to be those
+strings and the lens still satisfies the examples. While the disconnect lens is
+prohibited by two examples, or by a carefully chosen input-output example on
+\PutR or \PutL, this problem continues presenting itself for the sublenses, and
+is highlighted even more on larger formats. While we could merely ask
+programmers to provide large, tight specifications for their synthesis tasks, we
+would prefer the algorithm to choose the right lens automatically. Instead of
+merely finding \emph{any} satisfying lens, we wish to synthesize the \emph{best}
+satisfying lens.
+
+But what is the best satisfying lens?  We think the best satisfying lens is the
+one that synchronizes as much data as is permitted by the type system and the
+examples.  More formally, we wish to minimize the expected number of bits
+required to recover data in one format from the other.  The expected number of
+bits is a well known concept in information theory, the expected number of bits
+of a data source is the \emph{entropy} of that data source.
+
+The entropy of a data source is based on the probability distribution of that
+data source. Given a set $S$, and a probability distribution over that set $P
+\OfType S \rightarrow \Reals$, the entropy of $(S,P)$ is $-\Sigma_{s\in
+  S}P(s)\Log_2P(s)$. While we have ways of expressing sets, we do not have
+probability distributions. In this work, we use \emph{stochastic regular
+  expressions}, regular expressions annotated with probabilities that express
+languages and probability distributions over those languages. In this work, we
+guess a reasonable probability distribution from the syntax, though if users
+manually provided stochastic regular expressions, or large suite of data in the
+formats through which we could learn the annotations, we could instead use that.
+
+With sets and probability distributions, we can find the entropy of the
+stochastic regular expression, given another stochastic regular expression, and
+a lens that shows which strings are synchronized with which. We aim to find the
+lens that minimizes that entropy.  Now that we have a notion of best, we need a
+way to search for that best lens.
+
+Prior work on lens synthesis works with two communicating synthesizers, one that
+proposes regular expressions, semantically but not syntactically equivalent to
+each other, and another that finds lenses between them, if there are any lenses
+between them that don't use a type equivalence rule involving star equivalences.
+We utilize this same framework, except instead of having the second synthesizer
+find \emph{any} lens between the types, \GreedySynth instead uses a greedy
+algorithm to try finding the \emph{best} lens between the types.
+
+This poses a new problem: unless we find a lens with no cost, further search may
+find a better lens. While we may have found a great lens between two regular
+expressions, a different regular expression pair may provide an even better
+lens. There may be a great lens between \lstinline{crontab} and
+\lstinline{scheduled_job} that uses no star equivalences, but the lens shown in
+Figure~\ref{fig:example_lens} does. However, there is a tension, we want to find
+a good lens, but we also want to give quick feedback to the user. After finding
+the lens in Figure~\ref{fig:example_lens}, the algorithm shouldn't spend much
+longer searching for better lenses.
+
+We address this by keeping running track of the best lens found thus far. Before
+proposing new types, the regular expression generator, \RXSearch, evaluates
+whether it should continue proposing regular expression pairs, or whether it is
+content with the current best. For a given regular expression pair, \RXSearch
+looks at the number of rewrite rules performed on the pair, and identifies how
+many pairs with involving the same number of rewrites may be proposed. If there
+are a large number of pairs that can be proposed with that many rewrites, or the
+current lens is very well synchronized, \RXSearch does not propose that pairs.
+In this way, \RXSearch will continue searching for better lenses, but does not
+tackle huge search spaces unless it needs to. In this way, the algorithm can
+provide lenses quickly to the users, which users can refine with additional
+examples.
+% end overview
+
+% begin ssl
+\section{Simple Symmetric String Lenses}
+\label{sec:ssl}
+We now give a formal description of all the combinators in our symmetric lens
+language. These lens terms are typed between pairs of regular expressions: the
+typing judgment $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$ means that
+$\Lens$ is a well typed lens between $\LanguageOf{\Regex}$ and
+$\LanguageOf{\RegexAlt}$.
+
+\begin{center}
+  \begin{tabular}{l@{\ }l@{\ }l@{\ }l}
+    % REGEX
+    \Lens{} & \GEq{} & & $\IdentityLensOf{\Regex}$\\
+            & & \GBar{} & $\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}$ \\
+            & & \GBar{} & $\IterateLensOf{\Lens}$ \\
+            & & \GBar{} & $\ConcatLensOf{\Lens_1}{\Lens_2}$ \\
+            & & \GBar{} & $\SwapLensOf{\Lens_1}{\Lens_2}$ \\
+            & & \GBar{} & $\OrLensOf{\Lens_1}{\Lens_2}$ \\
+            & & \GBar{} & $\ComposeLensOf{\Lens_1}{\Lens_2}$ \\
+            & & \GBar{} & $\MergeROf{\Lens_1}{\Lens_2}$ \\
+            & & \GBar{} & $\MergeLOf{\Lens_1}{\Lens_2}$ \\
+            & & \GBar{} & $\InvertOf{\Lens}$ \\
+  \end{tabular}
+\end{center}
+
+We have already described over a number of these lenses in \S\ref{sec:overview}.
+The compose lens $\Lens_1 ; \Lens_2$ composes two lenses sequentially, it
+transforms from one data format to the other by first transforming to an
+intermediate format, shared as the target of $\Lens_1$ and the source of
+$\Lens_2$. The \MergeR lens combines two sublenses which share the same target.
+Which lens gets used depends on what data is in the left-hand format, if it is
+in the domain of $\Lens_1$ then $\Lens_1$ gets used, and similarly for
+$\Lens_2$.  The \MergeL lens is symmetric to \MergeR, the two sublenses must
+share the same source. Finally, $\InvertOf{\Lens}$ inverts a lens, \CreateR
+becomes \CreateL and \PutR becomes \PutL, and vice-versa.
+
+We now provide the typing derivations and semantics for these lenses. For
+simplicity, we define the typing derivations and semantics only over unambiguous
+regular expressions, even though some lenses can be defined over ambiguous
+regular expressions (for example, the identity lens doesn't need any ambiguity
+constraints).
+
+\[
+  \centering
+  \inferrule*
+  {
+  }
+  {
+    \IdentityLensOf{\Regex} \OfType \Regex \Leftrightarrow \Regex
+  }
+\]
+\begin{center}
+  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App s$ & = & $s$\\
+    $\CreateL{} \App s$ & = & $s$\\
+    $\PutR{} \App s_1 \App s_2$ & = & $s_1$\\
+    $\PutL{} \App s_1 \App s_2$ & = & $s_1$
+  \end{tabular}
+\end{center}
+
+Note that the identity lens ignores the second argument in the put functions.
+Because the two formats are fully synchronized, no knowledge of the prior data
+is needed.
+
+\[
+  \centering
+  \inferrule*
+  {
+    \String \in \LanguageOf{\Regex}\\
+    \StringAlt \in \LanguageOf{\RegexAlt}
+  }
+  {
+    \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}
+    \OfType \Regex \Leftrightarrow \RegexAlt
+  }
+\]
+\begin{center}
+  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App \String'$ & = & $\StringAlt$\\
+    $\CreateL{} \App \StringAlt'$ & = & $\String$\\
+    $\PutR{} \App \String' \App \StringAlt'$ & = & $\StringAlt'$\\
+    $\PutL{} \App \StringAlt' \App \String'$ & = & $\String'$
+  \end{tabular}
+\end{center}
+Just as the identity lens ignores the second argument in puts, disconnect lenses
+ignore the first in both puts and creates.  The data is unsynchronized in these
+two formats, information from one format doesn't impact the other.
+
+\[
+  \centering
+  \inferrule*
+  {
+    \Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
+    \Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt_2
+  }
+  {
+    \ConcatLensOf{\Lens_1}{\Lens_2} \OfType \Regex_1 \Concat \Regex_2
+    \Leftrightarrow
+    \RegexAlt_1 \Concat \RegexAlt_2
+  }
+\]
+\begin{center}
+  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App \String_1 \String_2$ & = & $(\Lens_1.\CreateROf{\String_1})(\Lens_2.\CreateROf{\String_2})$\\
+    $\CreateL{} \App \StringAlt_1 \StringAlt_2$ & = & $(\Lens_1.\CreateLOf{\StringAlt_1})(\Lens_2.\CreateLOf{\StringAlt_2})$\\
+    $\PutR{} \App (\String_1\String_2) \App (\StringAlt_1\StringAlt_2)$ & = & $(\Lens_1.\PutROf{\String_1}{\StringAlt_1})(\Lens_2.\PutROf{\String_2}{\StringAlt_2})$\\
+    $\PutL{} \App (\StringAlt_1\StringAlt_2) \App (\String_1\String_2)$ & = & $(\Lens_1.\PutLOf{\StringAlt_1}{\String_1})(\Lens_2.\PutLOf{\StringAlt_2}{\String_2})$\\
+  \end{tabular}
+\end{center}
+
+\afm{there isn't much to say here, should I put these in a figure or something?}
+
+\[
+  \centering
+  \inferrule*
+  {
+    \Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
+    \Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt_2
+  }
+  {
+    \SwapLensOf{\Lens_1}{\Lens_2} \OfType \Regex_1 \Concat \Regex_2
+    \Leftrightarrow
+    \RegexAlt_2 \Concat \RegexAlt_1
+  }
+\]
+\begin{center}
+  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App \String_1\String_2$ & = & $(\Lens_2.\CreateROf{\String_2})(\Lens_1.\CreateROf{\String_1})$\\
+    $\CreateL{} \App \StringAlt_2\StringAlt_1$ & = & $(\Lens_1.\CreateLOf{\StringAlt_1})(\Lens_2.\CreateLOf{\StringAlt_2})$\\
+    $\PutR{} \App (\String_1\String_2) \App (\StringAlt_2\StringAlt_1)$ & = & $(\Lens_2.\PutROf{\String_2}{\StringAlt_2})(\Lens_1.\PutROf{\String_1}{\StringAlt_1})$\\
+    $\PutL{} \App (\StringAlt_2\StringAlt_1) \App (\String_1\String_2)$ & = & $(\Lens_1.\PutLOf{\StringAlt_1}{\String_1})(\Lens_2.\PutLOf{\StringAlt_2}{\String_2})$\\
+  \end{tabular}
+\end{center}
+
+\[
+  \centering
+  \inferrule*
+  {
+    \Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
+    \Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt_2
+  }
+  {
+    \OrLensOf{\Lens_1}{\Lens_2} \OfType
+    \RegexOr{\Regex_1}{\Regex_2}
+    \Leftrightarrow
+    \RegexOr{\RegexAlt_1}{\RegexAlt_2}
+  }
+\]
+\begin{center}
+  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App \String$
+    & =
+    & $\begin{cases*}
+      \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\Regex_1}$\\
+      \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\Regex_2}$
+      \end{cases*}$\\
+
+    $\CreateL{} \App \StringAlt$
+    & =
+    & $\begin{cases*}
+      \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\RegexAlt_1}$\\
+      \Lens_2.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\RegexAlt_2}$
+      \end{cases*}$\\
+
+    $\PutR{} \App \String \App \StringAlt$
+    & =
+    & $\begin{cases*}
+        \Lens_1.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\Regex_1} \BooleanAnd \StringAlt\in\LanguageOf{\RegexAlt_1}$\\
+        \Lens_2.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\Regex_2} \BooleanAnd \StringAlt\in\LanguageOf{\RegexAlt_2}$\\
+        \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\Regex_1} \BooleanAnd \StringAlt\in\LanguageOf{\RegexAlt_2}$\\
+        \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\Regex_2} \BooleanAnd \StringAlt\in\LanguageOf{\RegexAlt_1}$
+      \end{cases*}$\\
+
+    $\PutL{} \App \StringAlt \App \String$
+    & =
+    & $\begin{cases*}
+        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_1} \BooleanAnd \String\in\LanguageOf{\Regex_1}$\\
+        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_2} \BooleanAnd \String\in\LanguageOf{\Regex_2}$\\
+        \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\RegexAlt_1} \BooleanAnd \String\in\LanguageOf{\Regex_2}$\\
+        \Lens_2.\CreateLOf{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_2} \BooleanAnd \String\in\LanguageOf{\Regex_1}$
+      \end{cases*}$\\
+  \end{tabular}
+\end{center}
+
+\[
+  \centering
+  \inferrule*
+  {
+    \Lens \OfType \Regex \Leftrightarrow \RegexAlt
+  }
+  {
+    \IterateLensOf{\Lens} \OfType
+    \StarOf{\Regex}
+    \Leftrightarrow
+    \StarOf{\RegexAlt}
+  }
+\]
+\begin{center}
+  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App \String_1\ldots\String_n$
+    & =
+    & $(\Lens.\CreateROf{\String_1})\ldots(\Lens.\CreateROf{\String_n})$\\
+
+    $\CreateL{} \App \StringAlt_1\ldots\StringAlt_n$
+    & =
+    & $(\Lens.\CreateLOf{\StringAlt_1})\ldots(\Lens.\CreateLOf{\StringAlt_n})$\\
+
+    $\PutR{} \App (\String_1\ldots\String_n) \App (\StringAlt_1\ldots\StringAlt_m)$
+    & =
+    & $\StringAlt_1'\ldots\StringAlt_n'$ where $\StringAlt_i' =
+      \begin{cases*}
+        \Lens.\PutROf{\String_i}{\StringAlt_i} & if $i \leq m$\\
+        \Lens.\CreateROf{\String_i} & otherwise
+      \end{cases*}$\\
+    $\PutL{} \App (\StringAlt_1\ldots\StringAlt_m) \App (\String_1\ldots\String_n)$
+    & =
+    & $\String_1'\ldots\String_n'$ where $\String_i' =
+      \begin{cases*}
+        \Lens.\PutROf{\StringAlt_i}{\String_i} & if $i \leq n$\\
+        \Lens.\CreateROf{\StringAlt_i} & otherwise
+      \end{cases*}$
+  \end{tabular}
+\end{center}
+
+\[
+  \centering
+  \inferrule*
+  {
+    \Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt\\
+    \Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt
+  }
+  {
+    \MergeROf{\Lens_1}{\Lens_2} \OfType
+    \RegexOr{\Regex_1}{\Regex_2}
+    \Leftrightarrow
+    \RegexAlt
+  }
+\]
+\begin{center}
+  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App \String$
+    & =
+    & $\begin{cases*}
+      \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\Regex_1}$\\
+      \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\Regex_2}$
+      \end{cases*}$\\
+
+    $\CreateL{} \App \StringAlt$
+    & =
+    & $\Lens_1.\CreateLOf{\StringAlt}$\\
+
+    $\PutR{} \App \String \App \StringAlt$
+    & =
+    & $\begin{cases*}
+      \Lens_1.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\Regex_1}$\\
+      \Lens_2.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\Regex_2}$
+    \end{cases*}$\\
+
+    $\PutL{} \App \StringAlt \App \String$
+    & =
+    & $\begin{cases*}
+        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\String\in\LanguageOf{\Regex_1}$\\
+        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\String\in\LanguageOf{\Regex_2}$
+      \end{cases*}$\\
+  \end{tabular}
+\end{center}
+
+\[
+  \centering
+  \inferrule*
+  {
+    \Lens_1 \OfType \Regex \Leftrightarrow \RegexAlt_1\\
+    \Lens_2 \OfType \Regex \Leftrightarrow \RegexAlt_2
+  }
+  {
+    \MergeLOf{\Lens_1}{\Lens_2} \OfType
+    \Regex
+    \Leftrightarrow
+    \RegexOr{\RegexAlt_1}{\RegexAlt_2}
+  }
+\]
+\begin{center}
+  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App \String$
+    & =
+    & $\Lens_1.\CreateROf{\String}$\\
+
+    $\CreateL{} \App \StringAlt$
+    & =
+    & $\begin{cases*}
+      \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\RegexAlt_1}$\\
+      \Lens_2.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\RegexAlt_2}$
+      \end{cases*}$\\
+
+    $\PutR{} \App \String \App \StringAlt$
+    & =
+    & $\begin{cases*}
+      \Lens_1.\PutROf{\String}{\StringAlt} & if $\StringAlt\in\LanguageOf{\RegexAlt_1}$\\
+      \Lens_2.\PutROf{\String}{\StringAlt} & if $\StringAlt\in\LanguageOf{\RegexAlt_2}$
+    \end{cases*}$\\
+
+    $\PutL{} \App \StringAlt \App \String$
+    & =
+    & $\begin{cases*}
+        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_1}$\\
+        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_2}$
+      \end{cases*}$\\
+  \end{tabular}
+\end{center}
+
+\[
+  \centering
+  \inferrule*
+  {
+    \Lens_1 \OfType \Regex \Leftrightarrow \RegexAlt\\
+    \Lens_2 \OfType \RegexAlt \Leftrightarrow \RegexAltAlt\\
+  }
+  {
+    \ComposeLensOf{\Lens_1}{\Lens_2} \OfType
+    \Regex \Leftrightarrow \RegexAltAlt
+  }
+\]
+\begin{center}
+  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App \String$ & = & $\Lens_2.\CreateROf{(\Lens_1.\CreateROf{\String})}$\\
+    $\CreateL{} \App \StringAlt$ & = & $\Lens_1.\CreateLOf{(\Lens_2.\CreateLOf{\StringAlt})}$\\
+    $\PutR{} \App \String \App \StringAltAlt$ & = & $\Lens_2.\PutROf{(\Lens_1.\PutROf{\String}{(\Lens_2.\CreateLOf{\StringAltAlt})})}{\StringAltAlt}$\\
+    $\PutL{} \App \StringAltAlt \App \String$ & = & $\Lens_1.\PutLOf{(\Lens_2.\PutLOf{\StringAltAlt}{(\Lens_2.\CreateROf{\String})})}{\String}$
+  \end{tabular}
+\end{center}
+
+\[
+  \centering
+  \inferrule*
+  {
+    \Lens \OfType \Regex \Leftrightarrow \RegexAlt
+  }
+  {
+    \InvertOf{\Lens} \OfType \RegexAlt \Leftrightarrow \Regex
+  }
+\]
+\begin{center}
+  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App \StringAlt$ & = & $\Lens.\CreateLOf{\StringAlt}$\\
+    $\CreateL{} \App \String$ & = & $\Lens.\CreateROf{\String}$\\
+    $\PutR{} \App \StringAlt \App \String$ & = & $\Lens.\PutLOf{\StringAlt}{\String}$\\
+    $\PutL{} \App \String \App \StringAlt$ & = & $\Lens.\PutROf{\String}{\StringAlt}$
+  \end{tabular}
+\end{center}
+% end ssl
+
+\section{Synthesis Algorithm}
 
@@
     & =
-    & $\begin{cases*}1 & if $\String = \String'$\\ 0 & otherwise\end{cases*}$ \\
+    & $\begin{cases*}1 & if $\String = \String''$\\ 0 & otherwise\end{cases*}$ \\
 
@@
 % end stoch-rx
-
-\section{Simple Symmetric Lenses}
-\label{sec:ssl}
-Classically, asymmetric lenses comprise three functions, \Put{}, \Get{}, and
-\Create{}. These lenses only allow information to be lost on one side. We
-enhance this style of lens to symmetric lenses with \emph{simple
-  symmetric lenses}. Like the classical formulation of asymmetric lenses, simple
-symmetric lenses have no complement. Simple symmetric lenses comprise 4
-functions, \CreateR, \CreateL, \PutR, and \PutL. These functions satisfy the
-following round-tripping laws:
-\begin{equation}
-  \tag{\CreatePutRL}
-  \PutLOf{(\CreateROf{x})}{x} = x
-\end{equation}
-\begin{equation}
-  \tag{\CreatePutLR}
-  \PutROf{(\CreateLOf{y})}{y} = y
-\end{equation}
-\begin{equation}
-  \tag{\PutRL}
-  \PutLOf{(\PutROf{x}{y})}{x} = x
-\end{equation}
-\begin{equation}
-  \tag{\PutLR}
-  \PutROf{(\PutLOf{y}{x})}{y} = y
-\end{equation}
-
-% begin simple symmetric string lenses
-\section{Simple Symmetric String Lenses}
-\label{sec:ssl-string}
-While we have a theory for simple symmetric lenses, to integrate with Boomerang,
-we need combinators to express these lenses.
-
-\subsection{Identity}
-The identity lens operates on a single regular expression.  It fully
-synchronizes the data within that regular expression: if data gets changed in
-the left, that change gets reflected in the right.
-
-\subsection{Disconnect}
-The disconnect lens performs the opposite function as the identity lens.  It
-leave the two data formats provided separate.  If data gets changed on the left,
-the right is left unaffected.  When the data is being newly added, default
-behavior comes from one side or the other.
-
-\subsection{Concat}
-The concat lens combines two existing lenses, working on their concatenated
-regular expressions.  It performs the left lens on the left portion of the
-string, and the right lens on the right portion.  It requires unambiguity
-constraints to break up the left and right portions.
-
-\subsection{Swap}
-The swap lens operates like the concat lens, but when going from one side to the
-other, it swaps the order of the lenses.
-
-\subsection{Compose}
-The swap lens operates like the concat lens, but when going from one side to the
-other, it swaps the order of the lenses.
-
-\subsection{Iterate}
-The iterate lens breaks the provided strings into chunks, and performs the
-provided lens on those chunks, then concatenates them together.
-
-\subsection{Or}
-The or lens performs one lens or another, depending on the inputs to the lens.
-
-\subsection{MergeL}
-The mergel lens is used when the same piece of data on the left is synchronized
-to data in potentially many different forms on the right. The first provided
-lens is the one defaulted to on creates.
-
-\subsection{MergeR}
-The merger lens is symmetric to the mergel lens.
-
-\subsection{Compose}
-The compose lens applies one lens to go to an intermediary format, then applies
-the other to go to the final format.
-% end simple symmetric string lenses
 
@@
 
-\paragraph*{DNF Lenses}
+\subsection{DNF Lenses}
 The typing judgment is a 3-ary relation over a single DNF lens, and two DNF
@@
     \begin{cases*}
-      \SAtomLens_k.\PutRight \App \String_{i_k}'' \App \StringAlt_y & if $j_k = y$\\
+      \SAtomLens_k.\PutROf{\String_{i_k}''}{\StringAlt_y} & if $j_k = y$\\
       \StringAlt_y & if $\nexists k. j_k = y$\\
@@
     \begin{cases*}
-      \SAtomLens_k.\PutRight \App \StringAlt_{j_k}'' \App \String_x & if $i_k = x$\\
+      \SAtomLens_k.\PutROf{\StringAlt_{j_k}''}{\String_x} & if $i_k = x$\\
       \String_x & if $\nexists k. i_k = x$\\
@@
       \begin{cases*}
-        \SAtomLens_k.\PutRight \App \String_{i_k}'' \App \StringAlt_{y}'' & if $j_k = y$\\
+        \SAtomLens_k.\PutROf{\String_{i_k}''}{\StringAlt_{y}''} & if $j_k = y$\\
         \StringAlt_y'' & if $\nexists k. j_k = y$\\
@@
       \begin{cases*}
-        \SAtomLens_k.\PutLeft \App \StringAlt_{j_k}'' \App \String_{x}'' & if $i_k = x$\\
+        \SAtomLens_k.\PutLOf{\StringAlt_{j_k}''}{\String_{x}''} & if $i_k = x$\\
         \String_x'' & if $\nexists k. i_k = x$\\
@@
 \label{sec:related}
+
+\section{Simple Symmetric Lenses}
+\label{sec:ssl}
+Intuitively, simple symmetric lenses are much like symmetric lenses, but they
+avoid complements by keeping all information required to stitch edits to a data
+source in the data source itself. Instead of having puts operate on a
+complement, they operate on both data sources, where the alterations of the
+first argument update the second. Furthermore, on initial creations of the
+files, instead of using an $init$ complement, two additional functions,
+$\CreateR$ and $\CreateL$ are used.
+
+More formally, a simple symmetric lens between $X$ and $Y$ comprise the
+following four functions:
+\begin{enumerate}
+\item $\CreateR \OfType X \rightarrow Y$
+\item $\CreateL \OfType Y \rightarrow X$
+\item $\PutR \OfType X \rightarrow Y \rightarrow Y$
+\item $\PutL \OfType Y \rightarrow X \rightarrow X$
+\end{enumerate}
+These functions are guaranteed to satisfy the following ``round-tripping'' laws.
+\begin{gather}
+  \tag{\CreatePutRL}
+  \PutLOf{(\CreateROf{x})}{x} = x\\
+  \tag{\CreatePutLR}
+  \PutROf{(\CreateLOf{y})}{y} = y\\
+  \tag{\PutRL}
+  \PutLOf{(\PutROf{x}{y})}{x} = x\\
+  \tag{\PutLR}
+  \PutROf{(\PutLOf{y}{x})}{y} = y
+\end{gather}
+
+For synchronization, when the original crontab or Scheduled Tasks file is
+created, a synchronized file of the other form is created from it with
+$\CreateR$ or $\CreateL$ (depending on which file was initially created). On
+future edits, changes are made with the puts, a \PutR would propagate the shared
+information from the crontab, and knit it together with the information specific
+to the Scheduled Tasks file.
+
+\section{Symmetric Lenses}
+\label{sec:symmetric-lenses}
+
+Classically, these types of synchronization tasks are represented as symmetric
+lenses. A symmetric lens $\Lens$ between $X$ and $Y$ consists of 4 components: a
+complement $C$, a designated element $init \in C$, and two functions for
+propagating changes in each format to the other:
+\begin{enumerate}
+\item $\PutRSym \OfType X \times C \rightarrow Y \times C$
+\item $\PutLSym \OfType Y \times C \rightarrow X \times C$
+\end{enumerate}
+In this formulation, the components of each format that aren't shared with the
+other format are stored in the complement. When one format is edited, the \PutR
+or \PutL function stitches together that edited data with the data for the other
+side, stored in the complement.  The designated element, $init$, is used as the
+initial value of $C$, and specifies default behavior when there isn't existing
+data to stitch with.  In our example, the complement would consist of two lists
+of commands, where one list contains the Linux commands and one list contains
+the Windows commands.
+
+These components only make sense for synchronization when they satisfy
+round-tripping laws.  This guarantees that if an identity edit is propagated, no
+data gets changed.
+\begin{equation}
+  \tag{\PutRL}
+  \frac{\PutRSymOf{(x,c)} = (y,c')}{\PutLSymOf{(y,c'')} = (x,c')}
+\end{equation}
+\begin{equation}
+  \tag{\PutLR}
+  \frac{\PutLSymOf{(y,c)} = (x,c')}{\PutRSymOf{(x,c')} = (y,c')}
+\end{equation}
+
+Furthermore, two symmetric lenses are equivalent if they always output the same
+formats on any sequence of edits. Formally, given a lens $\Lens \in X
+\leftrightarrow Y$, an \emph{edit} for $\Lens$ is a member of $X + Y$. Consider
+the function $apply$, which, given a lens and an element of that lens's
+complement, is a function from sequences of edits to sequences of edits. If
+$apply(\Lens,c,es) = es'$, then that means, starting with complement $c$, after
+edit $es_i$, the lens $\Lens$ synchronizes the other data to $es'_i$.
+\[
+  \inferrule
+  {
+  }
+  {
+    apply(\Lens,c,[]) = []
+  }
+\]
+\[
+  \inferrule
+  {
+    \Lens.putr(x,c) = (y,c')\\
+    apply(\Lens,c',es) = es'
+  }
+  {
+    apply(\Lens,c,(\InLOf{x})::es) = (\InROf{y})::es'
+  }
+\]
+\[
+  \inferrule
+  {
+    \Lens.putl(y,c) = (x,c')\\
+    apply(\Lens,c',es) = es'
+  }
+  {
+    apply(\Lens,c,(\InROf{y})::es) = (\InLOf{x})::es'
+  }
+\]
+
+Two lenses, $\Lens_1$ and $\Lens_2$, are equivalent if
+$apply(\Lens_1,\Lens_1.init,es) = apply(\Lens_2,\Lens_2.init,es)$ for all edit
+sequences $es$.
+
+\subsection{Difficulties Posed}
+
+By describing our lenses in this complement style, we can express an incredibly
+wide class of lenses, as these complements permit storing any amount of
+information in them.  However, they pose problems when implementing them in
+real-world systems and in synthesizing them.
+
+Using symmetric lenses for synchronization forces us to maintain not just two
+files that we aim to sycnrhonize, but three.  This is unintuitive, and creates
+many problems in practice.  For example, consider the situation where I am
+synchronizing a crontab and job manager file.  I am synchronizing across
+computers, so I must either maintain a copy of the complement on each computer,
+and keep them synchronized, or figure out a computer to hold the complement, and
+make sure all edits go through that complement.  Furthermore, if I want to
+change the synchronization to a different computer, I don't need to merely make
+sure that my files are moved to the computer, I also have to make sure that I
+correctly move the complement, and hook up that complement to be used for that
+synchronization task.  This is a huge divergence from other synchronization
+utilities, where all that needs to be done is maintain the files.  Using those
+for synchronization, edits can be sent over the wire, and the appropriate
+function can stitch together those edits, no additional state needed.
+
+For synthesis, because each lens has a custom complement, we can no longer
+specify the put functions though mere input/output examples.  Instead, we have
+to specify the behavior through the observable behavior of edit sequences.
+Furthermore, there are lenses that require arbitrarily long edit sequences to
+differentiate between: consider the lens that keeps two formats equal for the
+first $k$ non-identity edits, then keeps them completely unsynchronized for all
+future edits.  This requires an edit sequence of $k$ edits to differentiate
+between this lens and the identity lens.  Simple symmetric lenses avoid this
+problem by not allowing such lenses: we can't keep track of the number of
+non-identity edits performed unless there are explicit fields in the data
+formats keeping track of them.  While we are reducing the expressivity of our
+lenses with this restriction, we feel the benefits outweigh the costs.
+
 A \emph{symmetric lens} between $X$ and $Y$ is an element of the type
@@
 lens formulations. However, some symmetric lenses certainly are inexpressible in
-the stateless symmetric lens formulation. We formalize which are expressible
+the simple symmetric lens formulation. We formalize which are expressible
 with a restriction on symmetric lenses. We define \emph{forgetful symmetric
@@
 data. Forgetful symmetric lenses express exactly the same $apply$ function as
-stateless symmetric lenses.
+simple symmetric lenses.
 
@@
     & =
-    & $\LetIn{(y,c')}{\Lens.\PutRight \App (x,c)}$\\
+    & $\LetIn{(y,c')}{\Lens.\PutRSymOf{(x,c)}}$\\
     &
@@
     & =
-    & $\LetIn{(x,c')}{\Lens.\PutLeft \App (y,c)}$\\
+    & $\LetIn{(x,c')}{\Lens.\PutLSymOf{(y,c)}}$\\
     &
@@
   \item $R_\Lens(c,Some (x,y)) \BooleanImplies
-    \Lens.\PutRight \App (x,c) = (y,c) \BooleanAnd
-    \Lens.\PutLeft  \App (y,c) = (x,c)$
+    \Lens.\PutRSymOf{(x,c)} = (y,c) \BooleanAnd
+    \Lens.\PutLSymOf{(y,c)} = (x,c)$
   \end{enumerate}
@@
 
-    So, by definition, $S(\Lens).\CreateROf{x} = \Fst \App (\Lens.\PutRight \App
-    (x,\Lens.init))$, so $y = y'$.
-
-    Furthermore, by \PutRL, $\Lens.\PutLeft \App (y,c') = (x,c')$, and by another
-    application of \PutLR, $\Lens.\PutRight \App (x,c') = (y,c')$.  This means
+    So, by definition, $S(\Lens).\CreateROf{x} = \Fst \App (\Lens.\PutRSymOf{(x,\Lens.init)})$, so $y = y'$.
+
+    Furthermore, by \PutRL, $\Lens.\PutLSymOf{(y,c')} = (x,c')$, and by another
+    application of \PutLR, $\Lens.\PutRSymOf{(x,c')} = (y,c')$.  This means
     that $R_{\Lens}(c',Some (x,y))$.
@@
     Performing $apply$ on $c$, we get $apply(\Lens,c,(\InLOf{x'})::es) =
-    (\InROf{y'})::es''$ where $\Lens.\PutRight \App (x',c) = (y',c')$ and
+    (\InROf{y'})::es''$ where $\Lens.\PutRSymOf{(x',c)} = (y',c')$ and
     $apply(\Lens,c',es) = es''$.
 
-    So, by definition, $S(\Lens).\PutROf{x}{y} = \Fst \App (\Lens.\PutRight \App
-    (x,c''))$, where $c'' = \Snd \App (\Lens.\PutLeft \App (y,\Lens.init))$.
-
-    By assumption, $R_{\Lens}(c,\SomeOf{(x,y)})$, so $\Lens.\PutLeft \App (y,c)
+    So, by definition, $S(\Lens).\PutROf{x}{y} = \Fst \App (\Lens.\PutRSymOf{(x,c'')})$, where $c'' = \Snd \App (\Lens.\PutLSymOf{(y,\Lens.init)})$.
+
+    By assumption, $R_{\Lens}(c,\SomeOf{(x,y)})$, so $\Lens.\PutLSymOf{(y,c)}
     = (x,c)$.  So, by Property~\ref{prop:forget-lr}, we know $y' = y''$.
 
-    Furthermore, by \PutRL, $\Lens.\PutLeft \App (y',c') = (x',c')$, and by another
-    application of \PutLR, $\Lens.\PutRight \App (x',c') = (y',c')$.  This means
+    Furthermore, by \PutRL, $\Lens.\PutLSymOf{(y',c')} = (x',c')$, and by another
+    application of \PutLR, $\Lens.\PutRSymOf{(x',c')} = (y',c')$.  This means
     that $R_{\Lens}(c',Some (x',y'))$.
@@
   Consider the following set $C$, distinguished element of that set, $init$, and
-  pair of functions $\PutRight$ and $\PutLeft$, that we wish to satisfy the
+  pair of functions $\PutRSym$ and $\PutLSym$, that we wish to satisfy the
   symmetric lens laws, and that we also wish to be forgetful.
@@
 
-      $\PutRight \App (x,c)$
+      $\PutRSymOf{(x,c)}$
       & =
@@
 
-      $\PutLeft \App (y,c)$
+      $\PutLSymOf{(y,c)}$
       & =
@@
   \begin{case}[c = \None]
-    Let $(y',c') = F(\Lens).\PutRight \App (x',\None)$. This means that $y' =
+    Let $(y',c') = F(\Lens).\PutRSymOf{(x',\None)}$. This means that $y' =
     \Lens.\CreateROf{x'}$, and $c' = \SomeOf{(x',y')}$.
 
-    Now, consider $(x'',c'') = F(\Lens).\PutLeft \App (y',\SomeOf{(x',y')})$. By
+    Now, consider $(x'',c'') = F(\Lens).\PutLSymOf{(y',\SomeOf{(x',y')})}$. By
     unfolding definitions, $x'' = \Lens.\PutLOf{y'}{x'}$. By \CreatePutRL, $x''
     = x'$, meaning $c'' = \SomeOf{(x',y')}$. This means $(x',c') =
-    F(\Lens).\PutLeft \App (y',\SomeOf{(x',y')})$, as desired.
+    F(\Lens).\PutLSymOf{(y',\SomeOf{(x',y')})}$, as desired.
   \end{case}
@@
   \begin{case}[c = \SomeOf{(x,y)}]
-    Let $(y',c') = F(\Lens).\PutRight \App (x',\SomeOf{(x,y)})$. This means that $y' =
+    Let $(y',c') = F(\Lens).\PutRSymOf{(x',\SomeOf{(x,y)})}$. This means that $y' =
     \Lens.\PutROf{x'}{y}$, and $c' = \SomeOf{(x',y')}$.
 
-    Now, consider $(x'',c'') = F(\Lens).\PutLeft \App (y',\SomeOf{(x',y')})$. By
+    Now, consider $(x'',c'') = F(\Lens).\PutLSymOf{(y',\SomeOf{(x',y')})}$. By
     unfolding definitions, $x'' = \Lens.\PutLOf{y'}{x'}$. By \PutRL, $x''
     = x'$, meaning $c'' = \SomeOf{(x',y')}$. This means $(x',c') =
-    F(\Lens).\PutLeft \App (y',\SomeOf{(x',y')})$, as desired.
+    F(\Lens).\PutLSymOf{(y',\SomeOf{(x',y')})}$, as desired.
   \end{case}
@@
 
-  We know that $c_1' = \Snd \App (F(\Lens).\PutRight \App (x,c_1))$ and $c_2' =
-  \Snd \App (F(\Lens).\PutRight \App (x,c_2))$.  Now, by inversion on
-  $F(\Lens).\PutRight$, we know that both $c_1' = Some(x,y_1')$ and $c_2' =
+  We know that $c_1' = \Snd \App (F(\Lens).\PutRSymOf{(x,c_1)})$ and $c_2' =
+  \Snd \App (F(\Lens).\PutRSymOf{(x,c_2)})$.  Now, by inversion on
+  $F(\Lens).\PutRSym$, we know that both $c_1' = Some(x,y_1')$ and $c_2' =
   Some(x,y_2')$ for some values of $y_1$ and $y_2$ (though we don't actually
@@
 
-  Now by unfolding definitions we know, $\Snd \App (F(\Lens).\PutLeft \App
-  (y,\SomeOf{(x,y_1')})) = \SomeOf(\Lens.\PutLeftOf{x}{y},y) = c_1''$.  Similarly, we
-  know $\Snd \App (F(\Lens).\PutLeft \App
-  (y,\SomeOf{(x,y_2')})) = \SomeOf(\Lens.\PutLeftOf{x}{y},y) = c_2''$, so $c_1''
-  = c_2''$, as intended.
+  Now by unfolding definitions we know, $\Snd \App
+  (F(\Lens).\PutLSymOf{(y,\SomeOf{(x,y_1')})}) =
+  \SomeOf(\Lens.\PutLOf{x}{y},y) = c_1''$. Similarly, we know $\Snd \App
+  (F(\Lens).\PutLSymOf{(y,\SomeOf{(x,y_2')})}) =
+  \SomeOf(\Lens.\PutLOf{x}{y},y) = c_2''$, so $c_1'' = c_2''$, as intended.
 \end{proof}
@@
     $apply(F(\Lens),\SomeOf{(x,y)},(\InLOf{x'})::es) = (\InROf{y'})::es''$ where
-    $F(\Lens).\PutRight \App (x',\SomeOf{(x,y)}) = (y'',\SomeOf{(x',y'')})$ and
+    $F(\Lens).\PutRSymOf{(x',\SomeOf{(x,y)})} = (y'',\SomeOf{(x',y'')})$ and
     $apply(\Lens,c',es) = es''$.
 
-    So, by definition, $\Fst \App (F(\Lens).\PutRight \App (x',\SomeOf{(x,y)}))
+    So, by definition, $\Fst \App (F(\Lens).\PutRSymOf{(x',\SomeOf{(x,y)})})
     = \Lens.PutROf{x'}{y'}$, so also by definition, $c' = \SomeOf{(x',y')}$.

### 2018-05-29 papers/SymmetricLenses/main.tex
add text ahh

@@
 be provided to clients in legacy or modern formats; calendar events can be
-stored in CSV files or ICS files. Oftentimes, we want to synchronize files
-across formats: when I edit a task to run more often on my Linux machine, I want
-this task to run more often on my Windows machine; when I update data through a
-legacy API, I want those changes to be reflected when I retrieve it through a
-modern API; when I create a new calendar event in my ICS file, I want my CSV
-calendar files to be updated to reflect this change.
+stored in CSV files or ICS files. Oftentimes, files should be synchronized
+across these formats: when I edit a task to run more often on my Linux machine,
+I want this task to run more often on my Windows machine; when I update data
+through a legacy API, I want those changes to be reflected when I retrieve it
+through a modern API; when I create a new calendar event in my ICS file, I want
+my CSV calendar files to be updated to reflect this change.
 
@@
 
-A better approach is to synthesize such conversion functions. Recent research
-has gone into Optician, a tool for automatically synthesizing such conversion
-functions from input-output examples and formal descriptions of the data
-formats~\cite{optician,quotient-optician}. However, this work assumes that all
-data fields present in one format are present in the other, but this is not
-always true. Oftentimes, fields present in one format are not present in the
-other: the commands scheduled to run on Linux machines are typically different
-than the commands scheduled to run on Windows machines; updated APIs can expose
-new information, or omit legacy information; ICS files contain a URL field that
-Google Calendar CSV files do not expose.
+A better approach is to synthesize such conversion functions. Prior work on
+synthesizing these functions assumes that all data fields present in one format
+are present in the other, but this is not always true. Oftentimes, fields
+present in one format are not present in the other: the commands scheduled to
+run on Linux machines are typically different than the commands scheduled to run
+on Windows machines; updated APIs can expose new information, or omit legacy
+information; ICS files contain a URL field that Google Calendar CSV files do not
+expose.
 
@@
 \emph{complement}. This makes them difficult to use in real-world
-synchronization tasks, as we must maintain the files we synchronize as well as
-maintaining the complements. The complement further complicates synthesis, as
+synchronization tasks, as users must maintain the synchronized files synchronize
+as well as the complements. The complement further complicates synthesis, as
 input-output examples are no longer sufficient, instead edit sequences must be
@@
 
-We enhance Optician with simple symmetric lens synthesis capabilities. This
-extension synthesizes symmetric lenses from a suite of input-output examples and
-a pair of regular expressions describing the data formats. The synthesis
-algorithm uses type-directed synthesis on the two regular expressions to
-synthesize a lens that is well-typed between the provided regular expressions
-and satisfies the provided input-output examples.
+We provide a tool for synthesizing simple symmetric lenses. This tool
+synthesizes symmetric lenses from a suite of input-output examples and a pair of
+regular expressions describing the data formats. The synthesis algorithm uses
+type-directed synthesis on the two regular expressions to synthesize a lens that
+is well-typed between the provided regular expressions and satisfies the
+provided input-output examples.
 
@@
 
-We operate under the assumption that users wish to synchronize as much data as
-possible, and try to synchronize as much data as the types and examples allow.
-This task gets harder when there are more complex decisions involved. For
-example, if I have two fields in one format, and one in the other, which field
-on the left should be synchronized to the one of the right? Or would it be
-better for the field on the right to be synchronized to a mish-mash of
+Operate under the assumption that users wish to synchronize as much data as
+possible, our algorithm try to synchronize as much data as the types and
+examples allow. This task gets harder when there are more complex decisions
+involved. For example, if I have two fields in one format, and one in the other,
+which field on the left should be synchronized to the one of the right? Or would
+it be better for the field on the right to be synchronized to a mish-mash of
 subcomponents from each field? To answer these questions, we develop a cost
@@
 algorithm. We develop our cost metric on DNF lenses, and provide a greedy
-algorithm that tries to find a symmetric DNF lens of minimal cost.
+algorithm that tries to find a symmetric DNF lens of minimal cost. The
+synthesized symmetric DNF lens is then converted to a lens in the original
+simple symmetric lens language.
 
@@
 \item We develop \emph{simple symmetric lenses} a new formulation for
-  bidirectional programs that can project data from both sides~(\S\ref{sec:ssl}),
+  bidirectional programs that can project data from both sides~(\S\ref{sec:overview}),
   and describe a language that expresses such lenses on string
-  data~(\S\ref{sec:ssl-string}).
+  data~(\S\ref{sec:ssl}).
 \item We develop a means to traverse stochastic regular expression equivalences,
@@
 needed; if a crontab is initialized, a synchronized Scheduled Tasks file is also
-created, and vice versa. Furthermore, we require \emph{puts}, functions that
+created, and vice versa. Furthermore, \emph{puts} are needed, functions that
 take an updated crontab and an old Scheduled Tasks file, and output a new
@@
 between them that don't use a type equivalence rule involving star equivalences.
-We utilize this same framework, except instead of having the second synthesizer
-find \emph{any} lens between the types, \GreedySynth instead uses a greedy
-algorithm to try finding the \emph{best} lens between the types.
+We utilize this same framework, with two modifications. (1) The first
+synthesizer does not propose regular expressions, but stochastic regular
+expressions, equivalent to the originals. We prove that these proposed
+stochastic regular expressions don't merely have the same languages as the
+originals, but also the same probability distributions. (2) Instead of having
+the second synthesizer find \emph{any} lens between the types, \GreedySynth
+instead uses a greedy algorithm to try finding the \emph{best} lens between the
+types. Furthermore, the first synthesizer proposes stochastic regular
+expressions with equivalent languages and probability distributions, instead of
+just equivalent languages.
 
@@
 \section{Simple Symmetric Lenses}
-\label{sec:ssl}
+\label{sec:ssl2}
 Intuitively, simple symmetric lenses are much like symmetric lenses, but they

### 2018-05-29 papers/SymmetricLenses/main.tex
add some more impl etc.

@@
 
-\section{Synthesis Algorithm}
-
-% begin stoch-rx
-\section{Stochastic Regular Expressions}
-\label{sec:stoch-rx}
+% begin synthesis
+\section{Synthesis}
+In this work, we aim to quickly find a good lens between two regular expressions
+that satisfies the input-output examples. To this end, we develop an information
+theoretic metric for how much data is synchronized. Information theoretic
+metrics require a notion of probability distributions \emph{how are probability
+  distributions being used, are certain things weighted more etc, assign weights
+  to different subcomponents and parts of the data}. We annotate
+the regular expressions with a reasonable probability distribution, making them
+\emph{stochastic regular expressions}. Armed with probability distributions over
+the languages, we have the capabilities to define information theoretic metrics
+over these probability distributions.
+
+Our overall synthesis algorithm utilizes two synthesizers shown in
+Figure~\ref{fig:high-level-algorithm}. There are an infinite number of possible
+lenses between two regular expressions that can be searched through. Instead of
+searching through all possible lenses, the lenses are grouped into classes of
+lenses based on the syntax of the regular expressions they synchronize. Two
+lenses are in the same class if they can be typed by the same regular
+expressions, using no star equivalences. This works well for synthesis, as
+finding a good lens within these classes permits an efficient synthesis
+procedure.
+
+However, proposing these classes is complicated by stochastic regular
+expressions. While the equivalences on regular expressions are well studied, the
+equivalences on stochastic regular expressions are not. We want to propose
+equivalent stochastic regular expressions, but both the languages and
+probability distributions must be maintained. If the probability distributions
+are not maintained, then the metric by which we compare our lenses changes for
+each class of lenses, we can't compare lenses between classes. To alleviate
+this, we describe the \emph{star-semiring} equivalences on stochastic regular
+expressions, and prove they are sound with respect to language and probability
+distributions. Our \RXSearch proposes stochastic regular expressions by
+traversing a subset of these equivalences, so \RXSearch only proposes regular
+expression pairs equivalent to the originals.
+
+The second algorithm, \GreedySynth, searches within a class for the lowest cost
+lens.  It is a greedy algorithm, and not guaranteed to find the lowest cost
+lens.  However, we haven't found any instances in which \GreedySynth doesn't find
+the optimal lens, due to the partial backtracking we implemented. \GreedySynth
+works by searching, not through symmetric lenses, but through an alternative
+language of Symmetric DNF lenses. After finding such a symmetric DNF lens, that
+symmetric DNF lens is converted into a simple symmetric lens.
+
+\subsection{Stochastic Regular Expressions}
+\label{subsec:stoch-rx}
 Important to our algorithm is finding the expected number of bits needed to
@@
 \end{theorem}
-% end stoch-rx
-
-% begin DNF regular expressions and lenses
-\section{Stochastic DNF Regular Expressions}
+
+\subsection{\RXSearch}
+The first synthesizer is one that searches through regular expression pairs.
+This synthesizer, \RXSearch, proposes regular expressions, semantically but not
+syntactically equivalent to the originals.  In particular, this synthesizer
+performs the following rewrites:
+\begin{center}
+  \begin{tabular}{rcl}
+    $\PRegexStar{\Regex}{\Probability}$
+    & \Rewrite
+    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability}\\
+
+    $\PRegexStar{\Regex}{\Probability}$
+    & \Rewrite
+    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{\Regex{}}{\Probability}}{\Regex{}})}{\Probability}
+  \end{tabular}
+\end{center}
+As well as the congruence rules that allow these rewrites to be applied on
+subexpressions.
+
+\RXSearch proposes rewrites incrementally, where it proposes regular expression
+pairs with the smallest number of rewrites first, and when there are no more
+proposed regular expressions of that size, \RXSearch proposes regular
+expressions of larger size.
+
+\RXSearch terminates when the number of regular expressions of the smallest
+unproposed size exceeds the cost of the current best lens. At this point,
+\RXSearch stops proposing regular expression pairs, and instead returns to the
+user the best lens found thus far.  Note, if \RXSearch finds a bijective lens (a
+lens of cost 0), it will immediately return, as would be expected.
+
+\afm{dont forget to add in description of termination}
+%
+
+\subsection{\GreedySynth}
+The second synthesizer, \GreedySynth, actually generates the lenses. It does
+this by first converting the regular expressions into \emph{stochastic DNF
+  regular expressions}, intuitively stochastic regular expressions with all
+disjunctions distributed over all concatenations. Next, a good DNF lens is
+greedily found between these stochastic DNF regular expressions, using the
+probability annotations on the DNF regular expressions to find the cost.
+Finally, this DNF lens is converted into a simple symmetric lens.
+
+\paragraph*{Stochastic DNF Regular Expressions}
 \label{sec:dnf-regex-lens}
@@
 into DNF form, and find the entropy of it.
-% end DNF regular expressions and lenses
-
-% begin symmetric dnf lenses
-\section{Symmetric DNF Lenses}
+
+\paragraph*{Symmetric DNF, Sequence, and Atom Lenses}
 \label{sec:symmetric-dnf-lenses}
-\subsection{Syntax}
 Symmetric DNF Lenses are what we synthesize, and we then turn them into
@@
     % REGEX
-    \SAtomLens{} & \GEq{} & $\IterateLensOf{\SDNFLens}$ \\
+    \SAtomLens{} & \GEq{} & $\StarOf{\SDNFLens}$ \\
     \SSQLens{} & \GEq{} & $(\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
@@
 
-\subsection{Typing and Semantics}
 Because these lenses only encode functions on well-typed programs, we introduce
@@
 
-\subsection{DNF Lenses}
+\paragraph*{Symmetric DNF Lenses, Typing and Semantics}
 The typing judgment is a 3-ary relation over a single DNF lens, and two DNF
@@
 
-\paragraph*{Sequence Lenses}
+\paragraph*{Sequence Lenses, Typing and Semantics}
 The typing judgment is a 3-ary relation over a single Sequence lens, and two
@@
 
-\paragraph*{Atom Lenses}
+\paragraph*{Atom Lenses, Typing and Semantics}
 The typing judgment is a 3-ary relation over a single atom lens, and two
@@
   {
-    \IterateLensOf{\SDNFLens}
+    \StarOf{\SDNFLens}
     \OfType \PRegexStar{\DNFRegex}{\Probability}
@@
 
+\paragraph*{Information-Theoretic Cost Metric}
 Now that we've developed symmetric DNF lenses, it's important to develop a means
@@
 
-% end symmetric dnf lenses
-
-% begin synthesis
-\section{Synthesis}
-\label{sec:synthesis}
-While examples can be used to tighten the
-specification, a large number of well-constructed examples are required to
-sufficiently constrain the search space.
-
-Unfortunately, trying to find the best symmetric lens is a daunting task!
-Unless we find a bijective lens (in which case the sum is zero), we can never be
-sure if we will find a better lens than the one we found.
+\paragraph*{Greedy Algorithm}
+The greedy algorithm is actually two greedy algorithms, one that greedily finds
+symmetric DNF lenses and one that greedily finds symmetric sequence lenses.
+
+The algorithm for symmetric DNF lenses is relatively straightforward.  First,
+using examples, we identify which atoms in one format \emph{must} be
+synchronized to certain atoms in the other.  After that, we iteratively find the
+pairs of atoms with lowest cost, where at least one of the atoms is not mapped
+to another atom.  The cost of these pairs is not merely the cost of the lens,
+but also the cost incurred by adding that lens, if one of the atoms is already
+involved in a map, the lens has additional cost to account for the merge cost.
+This continues until there are no more atoms that aren't mapped to another atom.
+At this point, the creates for each lens is determined.  If a examples dictate
+one lens must create to another, that is the create behavior, otherwise atoms
+create along the lens of least cost.
+
+The algorithm for symmetric sequence lenses is less straightforward.  Unlike
+symmetric DNF lenses, where multiple subcomponents can map to the same
+subcomponent, each atom can map to at most one other subcomponent.  This is
+complicated by the fact that examples sometimes dictate one atom must map to
+another.
+
+To address this, we introduce the \emph{profit} of a adding a mapping between
+two subcomponents. The profit is the cost of the lens, minus the information
+content of the subcomponents it maps between, plus how much profit would be lost
+by freeing those subcomponents to make the subcomponents available. Then, lenses
+are only added if they have positive profit, or if one of the atoms requires a
+mapping. After there are no more lenses with positive profit or with
+subomponents requiring a mapping, the sequence lens is returned.
+
+\subsection{Optimizations}
+
+We included a number of optimizations in our algorithm. We added ``open'' and
+``closed'' annotations to regular expressions, an expansion inference algorithm,
+hash-consing and memoization, and compositional synthesis to our algorithm.
+
+\paragraph*{Open and Closed Regular Expressions} While GreedySynth acts
+relatively efficiently, it suffers from an exponential blowup when converting
+regular expressions to DNF form. This can be minimized by not converting
+\emph{all} of the regular expression into DNF form, but only converting
+\emph{some} of it to DNF form. We add in an additional atom, where an atom can
+be an either an iteration of a DNF regular expression or it can be a regular
+expression. When converted into a DNF regular expression, if $\Regex$ is a
+stochastic regular expression, annotated as closed, then $\ToDNFRegexOf{\Regex}
+= \AtomToDNFOf{\Regex}$.
+
+At the start of synthesis, we annotate our regular expressions as closed. Now,
+we enhance our rewrites by changing an annotation to from closed to open. This
+forces changes to \RXSearch and \GreedySynth.  \RXSearch is changed in that now
+it can perform new rewrites.  \GreedySynth is changed in that now there is an
+additional type of atom that can be mapped between.  Two regular expressions can
+map to each other only through the identity lens: two regular expressions can
+map to each other by the identity lens, if they are the same and all the
+examples say they are the same, otherwise each regular expression is left
+disconnected.
+
+\paragraph*{Expansion Inference} The prior optimization helped \GreedySynth, but
+it made \RXSearch have a much harder job.  To alleviate this difficulty, we
+added in expansion inference.  Expansion inference identifies when a regular
+expression does not have a chance to map through any means other than
+disconnected.  Expansion inference opens those regular expressions.
+
+\paragraph*{Hash-consing and Memoization} We further help \GreedySynth by
+memoizing its results.  However, our stochastic DNF regular expressions are
+quite large and so hashing them is computationally intensive.  We use less
+memory and allow for quick comparisons and hashing of these DNF regular
+expressions by hash-consing them.
+
+\paragraph*{Compositional Synthesis} Compositional synthesis allows us to use
+previously defined lenses - whether they are previously defined by users or by
+synthesis. We alter \GreedySynth such that two regular expression atoms can be
+mapped to each other either if they are the same regular expression and have
+examples agreeing with the examples, or if there is a lens which is a
+composition of predefined lenses between them, where the lens alters the
+examples in a desired way.
 % end synthesis
@@
 \section{Implementation}
-\label{sec:implementation}
+We have implemented our symmetric DNF lenses and the synthesis of them as an
+extension to Boomerang. We added symmetric lens combinators to Boomerang, and
+implemented Boomerang's previous asymmetric lens combinators within the
+symmetric lens combinators.
+
+We have extended Optician with our symmetric synthesis algorithm.
 % end implementation
@@
 \label{sec:related}
+There is related work in a number of directions.  There is related work in
+synthesizing synchronization functions, related work on symmetric lenses,
+related work on string transformation synthesis, and related work on
+type-directed synthesis.
+
+\subsection{Symmetric Lenses}
+We have reformulated the problem addressed by symmetric lenses with simple
+symmetric lenses.
 
@@
   \tag{\PutRL}
-  \frac{\PutRSymOf{(x,c)} = (y,c')}{\PutLSymOf{(y,c'')} = (x,c')}
+  \frac{\PutRSymOf{(x,c)} = (y,c')}{\PutLSymOf{(y,c''''')} = (x,c')}
 \end{equation}
@@
   \tag{\PutLR}
-  \frac{\PutLSymOf{(y,c)} = (x,c')}{\PutRSymOf{(x,c')} = (y,c')}
+  \frac{\PutLSymOf{(y,c)} = (x,c''')}{\PutRSymOf{(x,c')} = (y,c')}
 \end{equation}

### 2018-05-29 papers/SymmetricLenses/main.tex
add in related work and stuff

@@
 There is related work in a number of directions.  There is related work in
-synthesizing synchronization functions, related work on symmetric lenses,
+symmetric lenses, related work on synthesizing synchronization functions,
 related work on string transformation synthesis, and related work on
@@
 \subsection{Symmetric Lenses}
-We have reformulated the problem addressed by symmetric lenses with simple
-symmetric lenses.
-
-\section{Simple Symmetric Lenses}
-\label{sec:ssl2}
-Intuitively, simple symmetric lenses are much like symmetric lenses, but they
-avoid complements by keeping all information required to stitch edits to a data
-source in the data source itself. Instead of having puts operate on a
-complement, they operate on both data sources, where the alterations of the
-first argument update the second. Furthermore, on initial creations of the
-files, instead of using an $init$ complement, two additional functions,
-$\CreateR$ and $\CreateL$ are used.
-
-More formally, a simple symmetric lens between $X$ and $Y$ comprise the
-following four functions:
-\begin{enumerate}
-\item $\CreateR \OfType X \rightarrow Y$
-\item $\CreateL \OfType Y \rightarrow X$
-\item $\PutR \OfType X \rightarrow Y \rightarrow Y$
-\item $\PutL \OfType Y \rightarrow X \rightarrow X$
-\end{enumerate}
-These functions are guaranteed to satisfy the following ``round-tripping'' laws.
-\begin{gather}
-  \tag{\CreatePutRL}
-  \PutLOf{(\CreateROf{x})}{x} = x\\
-  \tag{\CreatePutLR}
-  \PutROf{(\CreateLOf{y})}{y} = y\\
-  \tag{\PutRL}
-  \PutLOf{(\PutROf{x}{y})}{x} = x\\
-  \tag{\PutLR}
-  \PutROf{(\PutLOf{y}{x})}{y} = y
-\end{gather}
-
-For synchronization, when the original crontab or Scheduled Tasks file is
-created, a synchronized file of the other form is created from it with
-$\CreateR$ or $\CreateL$ (depending on which file was initially created). On
-future edits, changes are made with the puts, a \PutR would propagate the shared
-information from the crontab, and knit it together with the information specific
-to the Scheduled Tasks file.
-
-\section{Symmetric Lenses}
-\label{sec:symmetric-lenses}
-
-Classically, these types of synchronization tasks are represented as symmetric
-lenses. A symmetric lens $\Lens$ between $X$ and $Y$ consists of 4 components: a
-complement $C$, a designated element $init \in C$, and two functions for
-propagating changes in each format to the other:
+\paragraph*{Formalization} We have reformulated the problem addressed by
+symmetric lenses with simple symmetric lenses. Classically, these types of
+synchronization tasks are represented as symmetric lenses. A symmetric lens
+$\Lens$ between $X$ and $Y$ consists of 4 components: a complement $C$, a
+designated element $init \in C$, and two functions for propagating changes in
+each format to the other:
 \begin{enumerate}
@@
   \tag{\PutRL}
-  \frac{\PutRSymOf{(x,c)} = (y,c')}{\PutLSymOf{(y,c''''')} = (x,c')}
+  \frac{\PutRSymOf{(x,c)} = (y,c')}{\PutLSymOf{(y,c')} = (x,c')}
 \end{equation}
@@
   \tag{\PutLR}
-  \frac{\PutLSymOf{(y,c)} = (x,c''')}{\PutRSymOf{(x,c')} = (y,c')}
+  \frac{\PutLSymOf{(y,c)} = (x,c')}{\PutRSymOf{(x,c')} = (y,c')}
 \end{equation}
@@
 
-\subsection{Difficulties Posed}
-
 By describing our lenses in this complement style, we can express an incredibly
@@
 
-Using symmetric lenses for synchronization forces us to maintain not just two
-files that we aim to sycnrhonize, but three.  This is unintuitive, and creates
-many problems in practice.  For example, consider the situation where I am
-synchronizing a crontab and job manager file.  I am synchronizing across
-computers, so I must either maintain a copy of the complement on each computer,
-and keep them synchronized, or figure out a computer to hold the complement, and
-make sure all edits go through that complement.  Furthermore, if I want to
-change the synchronization to a different computer, I don't need to merely make
-sure that my files are moved to the computer, I also have to make sure that I
-correctly move the complement, and hook up that complement to be used for that
-synchronization task.  This is a huge divergence from other synchronization
-utilities, where all that needs to be done is maintain the files.  Using those
-for synchronization, edits can be sent over the wire, and the appropriate
-function can stitch together those edits, no additional state needed.
+\paragraph*{Issues with Symmetric Lenses} Using symmetric lenses for
+synchronization forces us to maintain not just two files that we aim to
+sycnrhonize, but three. This is unintuitive, and creates many problems in
+practice. For example, consider the situation where I am synchronizing a crontab
+and job manager file. I am synchronizing across computers, so I must either
+maintain a copy of the complement on each computer, and keep them synchronized,
+or figure out a computer to hold the complement, and make sure all edits go
+through that complement. Furthermore, if I want to change the synchronization to
+a different computer, I don't need to merely make sure that my files are moved
+to the computer, I also have to make sure that I correctly move the complement,
+and hook up that complement to be used for that synchronization task. This is a
+huge divergence from other synchronization utilities, where all that needs to be
+done is maintain the files. Using those for synchronization, edits can be sent
+over the wire, and the appropriate function can stitch together those edits, no
+additional state needed.
 
@@
 
-A \emph{symmetric lens} between $X$ and $Y$ is an element of the type
-$\exists C.\{ \mathit{missing} :
-C,\,get : X \times C \to Y \times C,\,put : Y \times C \to X \times C\}$
-satisfying the following laws:
-\begin{equation}
-  \tag{\PutRL}
-  \frac{\PutR(x,c) = (y,c')}{\PutL(y,c') = (x,c')}
-\end{equation}
-\begin{equation}
-  \tag{\PutLR}
-  \frac{\PutL(y,c) = (x,c')}{\PutR(x,c') = (y,c')}
-\end{equation}
-
-Given a lens $l \in X \leftrightarrow Y$, a \emph{put object} for $l$ is a
-member of $X + Y$. Consider the function $apply$, which, given a lens and an
-element of that lens's complement, is a function from sequences of put objects
-to sequences of put objects. If $apply(\Lens,c,es) = es'$, then that means, starting
-with complement $c$, after edit $es_i$, the lens $\Lens$ synchronizes the other
-data to $es'_i$.
+\paragraph*{Relation with Simple Symmetric Lenses} To formalize how simple
+symmetric lenses lenses work as synchronization tools, we define an $apply$
+function on them, much like we define an $apply$ on symmetric lenses. The
+function $apply$, given a symmetric lens and an optional pair of elements of the
+source and target, is a function from sequences of put objects to sequences of
+put objects. If $apply(\Lens,\None,es,es')$, then that means, starting with no
+prior data, after edit $es_i$, the lens \Lens synchronizes the other data to
+$es_i'$. If $apply(\Lens,\SomeOf{(x,y)},es,es')$, then that means, starting with
+data $x$ and $y$ on the left and right, respectively, after edit $es_i$, the
+lens \Lens synchronizes the other data to $es_i'$.
+
 \[
@@
   {
-    apply(\Lens,c,[]) = []
+    apply(\Lens,xyo,[]) = []
   }
@@
   {
-    \Lens.putr(x,c) = (y,c')\\
-    apply(\Lens,c',es) = es'
-  }
-  {
-    apply(\Lens,c,(\InLOf{x})::es) = (\InROf{y})::es'
+    \Lens.\CreateROf{x} = y\\
+    apply(\Lens,\SomeOf{(x,y)},es) = es'
+  }
+  {
+    apply(l,\None,\InLOf{x}::es) = \InROf{y}::es'
   }
@@
   {
-    \Lens.putl(y,c) = (x,c')\\
-    apply(\Lens,c',es) = es'
-  }
-  {
-    apply(\Lens,c,(\InROf{y})::es) = (\InLOf{x})::es'
-  }
-\]
-
-Two lenses, $\Lens_1$ and $\Lens_2$, are equivalent if
-$apply(\Lens_1,\Lens_1.init,es) = apply(\Lens_2,\Lens_2.init,es)$ for all edit
-sequences $es$.
-
-Two simple symmetric lenses are equivalent if the four functions they comprise
-are equivalent as functions. However, to formalize how these lenses work as
-synchronization tools, we define an $apply$ function on them, much like we
-define an $apply$ on symmetric lenses.
-
-The function $apply$, given a symmetric lens and an optional pair of elements of
-the source and target, is a function from sequences of put objects to sequences
-of put objects.  If $apply(\Lens,\None,es,es')$, then that means, starting with
-no prior data, after edit $es_i$, the lens \Lens synchronizes the other data to
-$es_i'$.  If $apply(\Lens,\SomeOf{(x,y)},es,es')$, then that means, starting with
-data $x$ and $y$ on the left and right, respectively, after edit $es_i$, the
-lens \Lens synchronizes the other data to $es_i'$.
-
-\[
-  \inferrule
-  {
-  }
-  {
-    apply(\Lens,xyo,[]) = []
+    \Lens.\CreateLOf{y} = x\\
+    apply(\Lens,\SomeOf{(x,y)},es) = es'
+  }
+  {
+    apply(l,\None,\InROf{y}::es) = \InLOf{x}::es'
   }
@@
   {
-    \Lens.\CreateROf{x} = y\\
-    apply(\Lens,\SomeOf{(x,y)},es) = es'
-  }
-  {
-    apply(l,\None,\InLOf{x}::es) = \InROf{y}::es'
+    \Lens.\PutROf{x'}{y}  = y'\\
+    apply(\Lens,\SomeOf{(x',y')},es) = es'
+  }
+  {
+    apply(l,\SomeOf{(x,y)},\InLOf{x'}::es) = \InROf{y'}::es'
   }
@@
   {
-    \Lens.\CreateLOf{y} = x\\
-    apply(\Lens,\SomeOf{(x,y)},es) = es'
-  }
-  {
-    apply(l,\None,\InROf{y}::es) = \InLOf{x}::es'
-  }
-\]
-\[
-  \inferrule
-  {
-    \Lens.\PutROf{x'}{y}  = y'\\
-    apply(\Lens,\SomeOf{(x',y')},es) = es'
-  }
-  {
-    apply(l,\SomeOf{(x,y)},\InLOf{x'}::es) = \InROf{y'}::es'
-  }
-\]
-\[
-  \inferrule
-  {
     \Lens.\PutLOf{y'}{x}  = x'\\
@@
 \]
-
-Furthermore, equivalence of the $apply$ function corresponds exactly with
-equivalence of lenses.
-
-\begin{theorem}
-  $\Lens_1 \equiv \Lens_2$ if, and only if, for all sequences $es$,
-  $apply(\Lens_1,\None,es) = apply(\Lens_2,\None,es)$.
-\end{theorem}
 
@@
   where
-  \begin{enumerate}
-  \item $\Lens.creater \App x = \Lens.get \App x$
-  \item $\Lens.createl \App y = \Lens.create \App y$
-  \item $\Lens.putr \App x \App y = \Lens.get \App x$
-  \item $\Lens.putl \App y \App x = \Lens.put \App y \App x$
-  \end{enumerate}
+  \begin{center}
+    \begin{tabular}{rcl}
+      $\Lens.\CreateROf{x}$ & $=$ & $\Lens.get \App x$\\
+      $\Lens.\CreateLOf{y}$ & $=$ & $\Lens.create \App y$\\
+      $\Lens.\PutROf{x}{y}$ & $=$ & $\Lens.get \App x$\\
+      $\Lens.\PutLOf{y}{x}$ & $=$ & $\Lens.put \App y \App x$
+    \end{tabular}
+  \end{center}
 \end{theorem}
+
+\subsection{Synthesizing Synchronization Functions}
+
+\subsection{String Transformation Synthesis}
+
+\subsection{Type-Directed Synthesis}
 

### 2018-05-31 papers/SymmetricLenses/main.tex
clean up overview

@@
   -- data in one format need not be present in the other format. This gives the
-  synthesis algorithm much more freedom, as data can be completely
-  unsynchronized (edits in one format are never propagated), completely
-  synchronized (the formats are in bijective correspondence), or anywhere in
-  between. Under the assumption that users typically want to synchronize as much
-  data as possible, we provide an information theoretic metric for how much data
-  is left unsynchronized based on the theory of \emph{stochastic regular
-    expressions}. Furthermore, we use this metric to develop a greedy synthesis
-  algorithm that tries to minimize unsynchronized data. We experimentally
-  demonstrate that our algorithm can efficiently synthesize simple symmetric
-  lenses on a variety of benchmark suites consisting of synchronizing file
-  formats from configuration files, application-specific data storage files, and
-  data cleaning tasks.
+  synthesis algorithm much more freedom, as data can be completely unaligned
+  (edits in one format are never propagated), completely aligned (the formats
+  are in bijective correspondence), or anywhere in between. Under the assumption
+  that users typically want to align as much data as possible, we provide an
+  information theoretic metric for how much data is left unaligned based on the
+  theory of \emph{stochastic regular expressions}. Furthermore, we use this
+  metric to develop a greedy synthesis algorithm that tries to minimize
+  unaligned data. We experimentally demonstrate that our algorithm can
+  efficiently synthesize simple symmetric lenses on a variety of benchmark
+  suites consisting of synchronizing file formats from configuration files,
+  application-specific data storage files, and data cleaning tasks.
 
@@
 symmetric lenses use external state for synchronization, in the form of a
-\emph{complement}. This makes them difficult to use in real-world
-synchronization tasks, as users must maintain the synchronized files synchronize
+\emph{complement}. These complements makes symmetric lenses difficult to use in
+real-world synchronization tasks, as users must maintain the synchronized files
 as well as the complements. The complement further complicates synthesis, as
@@
 format, for any data component there is now a choice: is this data only relevant
-to this format and as such should be left unsynchronized, or is this data
-present in both formats and as such should be synchronized.
-
-Operate under the assumption that users wish to synchronize as much data as
-possible, our algorithm try to synchronize as much data as the types and
-examples allow. This task gets harder when there are more complex decisions
-involved. For example, if I have two fields in one format, and one in the other,
-which field on the left should be synchronized to the one of the right? Or would
-it be better for the field on the right to be synchronized to a mish-mash of
-subcomponents from each field? To answer these questions, we develop a cost
-metric for reasoning about which lens synchronizes the most information.
+to this format and as such should be left unaligned (changes to this data are
+not propagated to the other format), or is this data present in both formats and
+as such should be aligned (changes to this data change are propagated to the
+other format).
+
+Operating under the assumption that users wish to align as much data as
+possible, our algorithm tries to align as much data as the types and examples
+allow. This task gets harder when there are more complex decisions involved. For
+example, if I have two fields in one format, and one in the other, which field
+in the first format should be aligned to the sole field of the second format? Or
+would it be better for the field on the right to be aligneded to a mish-mash
+of subcomponents from each field? To answer these questions, we develop a cost
+metric for reasoning about which lens aligns the most information.
 
@@
 
-Our model for this is based on \emph{stochastic regular expressions}~\cite{?}. A
-probability distribution for strings is represented by a stochastic regular
-expression, and we develop a way to syntactically infer the expected information
-content (or \emph{entropy}) of an unambiguous regular expression. This entropy
-is used to calculate the expected information content to recover a string, when
-given a string it is synchronized to.
+Our model for this is based on \emph{stochastic regular expressions}~\cite{?}.
+Both a language and a probability distribution over that language is represented
+by a stochastic regular expression, and we develop a way to syntactically infer
+the expected information content (or \emph{entropy}) of an unambiguous regular
+expression. This entropy is used to calculate the expected information content
+to recover a string, when given a string it is synchronized to.
 
@@
 expression pairs that it thinks will align well, and one that greedily searches
-for a lens between those two regular expressions that minimizes cost and
-involves no rewrites. When the first algorithm thinks it has found the best
-possible lens, it returns that lens.
+for a lens between those two regular expressions that minimizes cost and does
+not traverse star-related equivalences. When the first algorithm is content with
+the lens it has found, it returns that lens.
 
@@
 rewrites this synthesizer applies are sound rewrites on stochastic regular
-expression.
+expressions.
 
@@
   expressions} and \emph{symmetric DNF lenses}. While these languages are hard
-to read and understand, they permit an efficient type-directed synthesis
-algorithm. We develop our cost metric on DNF lenses, and provide a greedy
-algorithm that tries to find a symmetric DNF lens of minimal cost. The
-synthesized symmetric DNF lens is then converted to a lens in the original
-simple symmetric lens language.
-
-We evaluate our algorithm on a benchmark suite consisting of the original
-Optician benchmarks, as well as additional benchmarks consisting of
-synchronization tasks where the data is not in bijective correspondence.
+to read and understand, they are highly normalized and permit an efficient
+type-directed synthesis algorithm. We develop our cost metric on DNF lenses, and
+provide a greedy algorithm that tries to find a symmetric DNF lens of minimal
+cost. The synthesized symmetric DNF lens is then converted to a lens in the
+original simple symmetric lens language.
+
+We evaluate our algorithm on a benchmark suite consisting of synchronizing file
+formats from configuration files, application-specific data storage files, and
+data cleaning tasks. We find that our algorithm is able to synthesize all these
+lenses in under 20 seconds.
 
@@
   while preserving probability distributions and
-  languages~(\S\ref{sec:stoch-rx}).
-\item We develop stochastic DNF regular expressions and DNF lenses. We develop a
-  sound algorithm for converting stochastic regular expressions to stochastic
-  DNF regular expressions~(\S\ref{sec:dnf-regex-lens}), and a
-  information-theoretic cost metric on DNF
-  lenses~(\S\ref{sec:symmetric-dnf-lenses}).
-\item We extended Boomerang with our simple symmetric lens combinators. We
-  updated the Optician synthesis system to synthesize symmetric lenses with our
-  new algorithm.
-\item We evaluate our synthesis algorithm on Opticians benchmark suite as well
-  as new, symmetric benchmarks, and find that we lose very little synthesis
-  time, while being significantly more expressive (\S\ref{sec:evaluation}).
+  languages~(\S\ref{subsec:stoch-rx}).
+\item We develop stochastic DNF regular expressions and DNF
+  lenses~(\S\ref{subsec:greedy-synth}). We develop a sound algorithm for
+  converting stochastic regular expressions to stochastic DNF regular
+  expressions and a information-theoretic cost metric on DNF lenses.
+\item We extended Boomerang with our simple symmetric lens combinators, and
+  integrated our algorithm with the Optician synthesis system within Boomerang,
+  allowing synthesis tasks as Boomerang
+  expressions~(\S\ref{sec:implementation}).
+\item We evaluate our algorithm on a benchmark suite consisting of synchronizing
+  file formats from configuration files, application-specific data storage
+  files, and data cleaning tasks (\S\ref{sec:evaluation}).
 \end{enumerate}
@@
 described example of keeping Linux and Windows scheduled jobs synchronized. In
-this example, we describe keeping synchronized a crontab (the file containing
-cron jobs) and a Scheduled Tasks file, scheduled.job. These are each formats
+this example, we describe keeping the file containing cron jobs (crontab) and a
+Scheduled Tasks file (scheduled.job) synchronized. These are each formats
 specifying when to run recurring tasks. For simplicity of description, we use a
 minified version of each of these, with examples and regular expressions shown
-in Figure~\ref{fig:minimized-representations}.
-
-For task is set in these files at specific times each hour. In both these
-formats, users can specify specific minutes they want to run the program during,
-or they can specify they want to run the program at every minute.
-
-In this example, we wish for the jobs to be synchronized across our windows and
-linux devices. However, the commands that need to be run are typically different
-for each device. Paths are different in Windows and Linux, and while there is
-typically a Windows equvialent for Linux commands, that equivalent command is
-not always named the same thing.  So, while we want to be able to add commands
-and edit how often they are run, we don't want to synchronize which commands are
-run themselves.
+in Figure~\ref{fig:minimized-representations}. Each file contains a list of
+tasks set to run at specific times every hour. In both these formats, users can
+either provide specific minutes they want to run the program during or they can
+specify they want to run the program at every minute.
+
+We wish for the jobs to be synchronized across our Windows and Linux devices.
+However, the commands that need to be run are typically different for each
+device. Paths are different in Windows and Linux, and while there is typically a
+Windows equvialent for Linux commands, that equivalent command is not always
+named the same thing. So while we want to be able to have the same number of
+commands, with the commands running at the same times, we don't want to
+align the actual commands that are run.
 
 To synchronize these files, a few functions are needed. Two \emph{creates} are
-needed; if a crontab is initialized, a synchronized Scheduled Tasks file is also
-created, and vice versa. Furthermore, \emph{puts} are needed, functions that
-take an updated crontab and an old Scheduled Tasks file, and output a new
-Scheduled Tasks file with the information synchronized to the updated crontab.
-To have a simple symmetric lens is to have these required synchronization
-functions. Formally, a simple symmetric lens between $X$ and $Y$ comprises 4
-functions:
+needed; if a crontab is initialized, a synchronized Scheduled Tasks file should
+also be created, and vice versa. Furthermore, \emph{puts} are needed; if a
+crontab is updated, we should be able to combine it with a non-updated Scheduled
+Tasks file, and output a new Scheduled Tasks file, synchronized to the updated
+crontab. To have a simple symmetric lens is to have these required
+synchronization functions. Formally, a simple symmetric lens between $X$ and $Y$
+comprises 4 functions:
 \begin{enumerate}
@@
 Furthermore, if a file is not edited, but is then resynchronized using a put
-function, no data should be changed. To be a lens, one must satisfy the
-following ``round-tripping'' laws that guarantee identity edits won't change the
-file data.
+function, no data should be changed. To be a simple symmetric lens, one must
+satisfy the following ``round-tripping'' laws, guaranteeing identity edits won't
+change the file data.
 \begin{gather}
@@
 pairs, if \lstinline{$\Lens$ : $\Regex \Leftrightarrow \RegexAlt$}, then $\Lens$
-encodes a lens between $\LanguageOf{\Regex}$ and $\LanguageOf{\RegexAlt}$. At
-the end, we want our symmetric lenses to by typed between \lstinline{crontab}
-and \lstinline{scheduled_job}.
+encodes a simple symmetric lens between $\LanguageOf{\Regex}$ and
+$\LanguageOf{\RegexAlt}$. At the end, we want our symmetric lenses to by typed
+between \lstinline{crontab} and \lstinline{scheduled_job}.
 
@@
 identity function on strings in that regular expression. The identity lens fully
-synchronizes the data formats: when one data format is updated, the other format
-gets the exact same value. In our example, when a job is told to run at a
-specific time in crontab, it is also expected to run at that same time in the
-Scheduled Jobs list, so the identity lens would propagate that information well.
+aligns the data formats: when one data format is updated, the other format gets
+the exact same value. In our example, when a job is told to run at a specific
+time in crontab, it is also expected to run at that same time in the Scheduled
+Jobs list. The identity lens propagates that information.
 %
@@
 
-Where the identity lens synchronizes data, the disconnect lens does the
-opposite: it leaves data unsynchronized. The disconnect lens takes 4 arguments:
-two regular expressions and two strings. On creates, because there is no
-association between data in the first format and data in the second format,
-designated default strings $\String$ and $\StringAlt$ are put into the other
-format. In our extended example, the Linux commands should not by synchronized
-with the Windows commands, so a disconnect lens would be used. We want our
-default behavior to be failures, under the assumption this failure will raise a
-warning, in case somebody forgets to find the analogous command of the other
-operating system.
+Where the identity lens aligns data, the disconnect lens does the opposite: it
+leaves data unaligned. The disconnect lens takes 4 arguments: two regular
+expressions and two strings. On creates, because there is no association between
+data in the first format and data in the second format, designated default
+strings are used for the other format. In our extended example, the Linux
+commands should not be synchronized with the Windows commands, so a disconnect
+lens would be used. We want our default behavior to be failures, under the
+assumption this failure will raise a warning, in case somebody forgets to find
+the analogous command of the other operating system.
 %
@@
 Finally, there are a number of lenses that compositionally build up new lenses
-from old ones. For example, data can appear in one form or another, a program
+from old ones. For example, data can appear in one form or another: a program
 can run during a specific set of minutes, or that program can run the every
@@
 Data can also be concatenated, and so lens mappings can be concatenated.
-However, data concatenations can only happen in a number of ways, as data can
-appear in the same order, and data can be reordered. For example, in the
+However, synchronizing concatenated data can happen in a number of ways, as data
+can appear in the same order, and data can be reordered. For example, in the
 crontab, time data appears before command data, whereas in the Scheduled Jobs
@@
 
-Lastly, the \IterateLens lens synchronizes repeated data, synchronizing each
-element of the iteration, and synchronizing the number of iterations. For
+Lastly, the \IterateLens lens synchronizes repeated data, aligning the number of
+times the data is repeated. The \IterateLens combinator takes a lens as an
+argument and synchronizes each element of the iteration with that lens. For
 example, if \lstinline{full_task_map} synchronizes crontab lines prepended by
@@
 in
-let repeated_task_map = iterate(full_task_map)
+let repeated_task_map = full_task_map . iterate(full_task_map)
 let nonempty_task_map = ins("<Tasks>") . repeated_task_map . ins("</Tasks>")
@@
 typing rule not previously mentioned: type equivalence. While
-\lstinline{scheduled_job} is a equivalent to a lens with an outermost
-disjunction, it is not syntactically equivalent to one.  Without type
+\lstinline{scheduled_job} is a equivalent to a regular expression with an
+outermost disjunction, it is not syntactically equivalent to one. Without type
 equivalence, the provided lens wouldn't actually be well-typed, but type
@@
 difficult to write. We had to provide a lot of scaffolding to map between the
-desired formats, even though the idea was quite simple: synchronize the times
-but leave the commands unsynchronized. Furthermore, this is inherent to manually
-writing these synchronization functions, as a programmer must describe which
-components are synchronized with which, and which components are scaffolding.
+desired formats, even though the idea was quite simple: align the times but
+leave the commands unaligned. Furthermore, this is inherent to manually writing
+these synchronization functions as a programmer must describe which components
+of the strings are aligned with which, and which components are scaffolding.
 
@@
 typed between the regular expression pair, and satisfies the input/output
-examples. For example, we wish to synthesize a lens from \lstinline{crontab} and
-\lstinline{scheduled_job}, using the strings in
-Figure~\ref{fig:minimized-representations} as input-output example for \CreateR.
+examples. In this example, we wish to synthesize a lens from \lstinline{crontab}
+and \lstinline{scheduled_job}, using the strings in
+Figure~\ref{fig:minimized-representations} as an input-output example for
+\CreateR.
 
@@
 Figure~\ref{fig:minimized-representations} as an input-output example helps
-little, but the default strings given to \Disconnect merely need to be those
-strings and the lens still satisfies the examples. While the disconnect lens is
-prohibited by two examples, or by a carefully chosen input-output example on
-\PutR or \PutL, this problem continues presenting itself for the sublenses, and
-is highlighted even more on larger formats. While we could merely ask
-programmers to provide large, tight specifications for their synthesis tasks, we
-would prefer the algorithm to choose the right lens automatically. Instead of
-merely finding \emph{any} satisfying lens, we wish to synthesize the \emph{best}
-satisfying lens.
+little. The default strings given to \Disconnect merely need to be the strings
+in the input-output examples, and the lens will still satisfy the specification.
+While the disconnect lens is prohibited by two examples, or by a carefully
+chosen input-output example on \PutR or \PutL, this problem continues presenting
+itself for the sublenses, and is highlighted even more on larger formats. While
+we could merely ask programmers to provide large, tight specifications for their
+synthesis tasks, we would prefer the algorithm to choose the right lens
+automatically. Instead of merely finding \emph{any} satisfying lens, we wish to
+synthesize the \emph{best} satisfying lens.
 
 But what is the best satisfying lens? We think the best satisfying lens is the
-one that synchronizes as much data as is permitted by the type system and the
+one that aligns as much data as is permitted by the type system and the
 examples. More formally, we wish to minimize the expected number of bits
-required to recover data in one format from the other.  The expected number of
-bits is a well known concept in information theory, the expected number of bits
-of a data source is the \emph{entropy} of that data source.
+required to recover data in one format from a synchronized string in the other.
+The expected number of bits is a well known concept in information theory: the
+expected number of bits of a data source is the \emph{entropy} of that data
+source.
 
@@
 data source. Given a set $S$, and a probability distribution over that set $P
-\OfType S \rightarrow \Reals$, the entropy of $(S,P)$ is $-\Sigma_{s\in
-  S}P(s)\Log_2P(s)$. While we have ways of expressing sets, we do not have
-probability distributions. In this work, we use \emph{stochastic regular
-  expressions}, regular expressions annotated with probabilities that express
-languages and probability distributions over those languages. In this work, we
-guess a reasonable probability distribution from the syntax, though if users
-manually provided stochastic regular expressions, or large suite of data in the
-formats through which we could learn the annotations, we could instead use that.
-
-With sets and probability distributions, we can find the entropy of the
-stochastic regular expression, given another stochastic regular expression, and
-a lens that shows which strings are synchronized with which. We aim to find the
-lens that minimizes that entropy.  Now that we have a notion of best, we need a
-way to search for that best lens.
-
-Prior work on lens synthesis works with two communicating synthesizers, one that
-proposes regular expressions, semantically but not syntactically equivalent to
-each other, and another that finds lenses between them, if there are any lenses
-between them that don't use a type equivalence rule involving star equivalences.
-We utilize this same framework, with two modifications. (1) The first
-synthesizer does not propose regular expressions, but stochastic regular
-expressions, equivalent to the originals. We prove that these proposed
-stochastic regular expressions don't merely have the same languages as the
-originals, but also the same probability distributions. (2) Instead of having
-the second synthesizer find \emph{any} lens between the types, \GreedySynth
-instead uses a greedy algorithm to try finding the \emph{best} lens between the
-types. Furthermore, the first synthesizer proposes stochastic regular
-expressions with equivalent languages and probability distributions, instead of
-just equivalent languages.
-
-This poses a new problem: unless we find a lens with no cost, further search may
-find a better lens. While we may have found a great lens between two regular
-expressions, a different regular expression pair may provide an even better
-lens. There may be a great lens between \lstinline{crontab} and
-\lstinline{scheduled_job} that uses no star equivalences, but the lens shown in
-Figure~\ref{fig:example_lens} does. However, there is a tension, we want to find
-a good lens, but we also want to give quick feedback to the user. After finding
-the lens in Figure~\ref{fig:example_lens}, the algorithm shouldn't spend much
-longer searching for better lenses.
-
-We address this by keeping running track of the best lens found thus far. Before
-proposing new types, the regular expression generator, \RXSearch, evaluates
-whether it should continue proposing regular expression pairs, or whether it is
-content with the current best. For a given regular expression pair, \RXSearch
-looks at the number of rewrite rules performed on the pair, and identifies how
-many pairs with involving the same number of rewrites may be proposed. If there
-are a large number of pairs that can be proposed with that many rewrites, or the
-current lens is very well synchronized, \RXSearch does not propose that pairs.
-In this way, \RXSearch will continue searching for better lenses, but does not
-tackle huge search spaces unless it needs to. In this way, the algorithm can
-provide lenses quickly to the users, which users can refine with additional
-examples.
+\OfType S \rightarrow \Reals$, the entropy of the $(S,P)$ pair is $-\Sigma_{s\in
+  S}P(s)\Log_2P(s)$. While we have ways of expressing sets of strings (with
+regular expressions), we do not have probability distributions. In this work, we
+express these probability distributions jointly with the language through
+\emph{stochastic regular expressions}, regular expressions annotated with
+probabilities. In this work, we guess a reasonable probability distribution from
+the syntax, though if users manually provided stochastic regular expressions, or
+large suite of data in the formats through which we could learn the annotations,
+we could instead use that.
+
+With sets and probability distributions, we can find the expected number of bits
+required to recover one string from a synchronized string. This expected value
+is the conditional entropy of the stochastic regular expression ($\Regex$) given
+another stochastic regular expression ($\RegexAlt$), and a lens that shows which
+strings are synchronized with which ($\Lens$). We aim to find the lens that
+minimizes the sum of that entropy with the conditional entropy of $\RegexAlt$
+given $\Regex$ and $\Lens$. Now that we have a notion of best, we need a way to
+search for that best lens.
+
+Our overall approach involves two communicating synthesizers: one (\RXSearch)
+that proposes regular expression pairs, semantically but not syntactically
+equivalent to the originals; one (\GreedySynth) that tries to find the best lens
+within a class of lenses, characterized by such regular expression pairs. Given
+two regular expressions, \GreedySynth can quickly search through the class of
+lenses that can be typed between them, where all instances of type equivalence
+involve no star equivalences. For example, consider the lens shown in
+Figure~\ref{fig:example_lens}. This lens would \emph{not} be in the class
+characterized by \lstinline{crontab} and \lstinline{scheduled_job}, as that
+typing derivation requires unrolling \lstinline{xml_task$^*$} within
+\lstinline{scheduled_job} into \lstinline{"" | (xml_task . xml_task$^*$)}.
+
+How does \RXSearch propose stochastic regular expression pairs? We define
+rewrite rules for stochastic regular expressions, and prove those rewrites
+preserve language and probability distributions. By applying these rewrites, we
+can generate new regular expression pairs, syntactically distinct but
+semantically equivalent to the originals.
+
+Critically for this work, \GreedySynth efficiently searches for one of the best
+lenses in a class of lenses, characterized by a stochastic regular expression
+pair. \GreedySynth works by searching through a smaller search space. First,
+\GreedySynth converts the provided stochastic regular expression pair into a
+\emph{stochastic DNF regular expression} pair. Where a stochastic regular
+expression pair characterizes a given class, a stochastic DNF regular expression
+pair is highly normalized -- it is the \emph{unique} pair that characterizes a
+given class. Then, we use the syntax of the DNF regular expression pair to
+greedily find a highly normalized version of a simple symmetric lens, a
+\emph{symmetric DNF lens}. Finally, this symmetric DNF lens is converted into a
+simple symmetric lens.
+
+However, these communicating synthesizers pose a problem: unless we find a lens
+with no cost, further search may find a better lens. While we may have found a
+great lens between two regular expressions, a different regular expression pair
+may provide an even better lens. There may be a great lens in the class
+characterized by \lstinline{crontab} and \lstinline{scheduled_job}, but it is
+not the the intended lens shown in Figure~\ref{fig:example_lens}. However, there
+is a tension, we want to find a good lens, but we also want to give quick
+feedback to the user. After finding the lens in Figure~\ref{fig:example_lens},
+we dont want the algorithm to spend much longer searching for better lenses.
+
+We address this by keeping running track of the best lens found thus far, and
+eventually cutting the proposal-synthesis loop short. Before proposing new
+types, the regular expression generator, \RXSearch, evaluates whether it should
+continue proposing regular expression pairs, or whether it is content with the
+current best. For a given regular expression pair, \RXSearch looks at the number
+of rewrite rules performed on the pair, and identifies how many pairs with
+involving the same number of rewrites may be proposed. If there are a large
+number of pairs that can be proposed with that many rewrites, or the current
+lens is very well synchronized, \RXSearch does not propose that pairs. In this
+way, \RXSearch will continue searching for better lenses, but does not tackle
+huge search spaces unless it needs to. In this way, the algorithm can provide
+lenses quickly to the users, which users can refine with additional examples.
 % end overview
@@
 \subsection{\GreedySynth}
+\label{subsec:greedy-synth}
 The second synthesizer, \GreedySynth, actually generates the lenses. It does
@@
 \paragraph*{Stochastic DNF Regular Expressions}
-\label{sec:dnf-regex-lens}
+\label{subsec:stoch-dnf-regex}
 We extend stochastic regular expressions to stochastic DNF regular expressions.
@@
 \section{Implementation}
+\label{sec:implementation}
 We have implemented our symmetric DNF lenses and the synthesis of them as an

### 2018-06-04 papers/SymmetricLenses/main.tex
add my new lens stuff

@@
 \GreedySynth converts the provided stochastic regular expression pair into a
-\emph{stochastic DNF regular expression} pair. Where a stochastic regular
-expression pair characterizes a given class, a stochastic DNF regular expression
-pair is highly normalized -- it is the \emph{unique} pair that characterizes a
-given class. Then, we use the syntax of the DNF regular expression pair to
-greedily find a highly normalized version of a simple symmetric lens, a
-\emph{symmetric DNF lens}. Finally, this symmetric DNF lens is converted into a
-simple symmetric lens.
-
-However, these communicating synthesizers pose a problem: unless we find a lens
-with no cost, further search may find a better lens. While we may have found a
-great lens between two regular expressions, a different regular expression pair
-may provide an even better lens. There may be a great lens in the class
-characterized by \lstinline{crontab} and \lstinline{scheduled_job}, but it is
-not the the intended lens shown in Figure~\ref{fig:example_lens}. However, there
-is a tension, we want to find a good lens, but we also want to give quick
-feedback to the user. After finding the lens in Figure~\ref{fig:example_lens},
-we dont want the algorithm to spend much longer searching for better lenses.
+\emph{stochastic DNF regular expression} pair. Then, the algorithm uses the
+syntax of the DNF regular expression pair to greedily find a highly normalized
+version of a simple symmetric lens, a \emph{symmetric DNF lens}. Finally, this
+symmetric DNF lens is converted into a simple symmetric lens. DNF Lenses include
+neither a composition operator nor a type equivalence rule. Unlike simple
+symmetric lenses, for which a single lens term is typed between many possible
+regular expression pairs, each DNF lens is only typed between a single regular
+expression pair. The relationship between DNF lens syntax and DNF regular
+expression syntax is incredibly tight, permitting an incredibly efficient
+synthesis algorithm for DNF lenses from DNF regular expressions.
+
+Our strategy involving communicating synthesizers permits both a way of
+enumerating classes and efficiently searching through them, but it poses a
+problem: unless we find a lens with no cost, other lens classes may contain
+better lenses. While we may have found a great lens between two regular
+expressions, a different regular expression pair may provide an even better
+lens. There may be a great lens in the class characterized by
+\lstinline{crontab} and \lstinline{scheduled_job}, but it is not the the
+intended lens shown in Figure~\ref{fig:example_lens}. However, there is a
+tension, we want to find a good lens, but we also want to give quick feedback to
+the user. After finding the lens in Figure~\ref{fig:example_lens}, we dont want
+the algorithm to spend much longer searching for better lenses.
 
@@
 \]
-\begin{center}
-  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
-    $\CreateR{} \App \String_1 \String_2$ & = & $(\Lens_1.\CreateROf{\String_1})(\Lens_2.\CreateROf{\String_2})$\\
-    $\CreateL{} \App \StringAlt_1 \StringAlt_2$ & = & $(\Lens_1.\CreateLOf{\StringAlt_1})(\Lens_2.\CreateLOf{\StringAlt_2})$\\
-    $\PutR{} \App (\String_1\String_2) \App (\StringAlt_1\StringAlt_2)$ & = & $(\Lens_1.\PutROf{\String_1}{\StringAlt_1})(\Lens_2.\PutROf{\String_2}{\StringAlt_2})$\\
-    $\PutL{} \App (\StringAlt_1\StringAlt_2) \App (\String_1\String_2)$ & = & $(\Lens_1.\PutLOf{\StringAlt_1}{\String_1})(\Lens_2.\PutLOf{\StringAlt_2}{\String_2})$\\
-  \end{tabular}
-\end{center}
-
-\afm{there isn't much to say here, should I put these in a figure or something?}
+%\begin{center}
+%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+%    $\CreateR{} \App \String_1 \String_2$ & = & $(\Lens_1.\CreateROf{\String_1})(\Lens_2.\CreateROf{\String_2})$\\
+%    $\CreateL{} \App \StringAlt_1 \StringAlt_2$ & = & $(\Lens_1.\CreateLOf{\StringAlt_1})(\Lens_2.\CreateLOf{\StringAlt_2})$\\
+%    $\PutR{} \App (\String_1\String_2) \App (\StringAlt_1\StringAlt_2)$ & = & $(\Lens_1.\PutROf{\String_1}{\StringAlt_1})(\Lens_2.\PutROf{\String_2}{\StringAlt_2})$\\
+%    $\PutL{} \App (\StringAlt_1\StringAlt_2) \App (\String_1\String_2)$ & = & $(\Lens_1.\PutLOf{\StringAlt_1}{\String_1})(\Lens_2.\PutLOf{\StringAlt_2}{\String_2})$\\
+%  \end{tabular}
+%\end{center}
+
+Concat is similar to concatenation in existing string lens languages like
+Boomerang.  For such terms, we do not provide the semantics, and merely refer
+readers to existing work.
 
@@
 \]
-\begin{center}
-  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
-    $\CreateR{} \App \String_1\String_2$ & = & $(\Lens_2.\CreateROf{\String_2})(\Lens_1.\CreateROf{\String_1})$\\
-    $\CreateL{} \App \StringAlt_2\StringAlt_1$ & = & $(\Lens_1.\CreateLOf{\StringAlt_1})(\Lens_2.\CreateLOf{\StringAlt_2})$\\
-    $\PutR{} \App (\String_1\String_2) \App (\StringAlt_2\StringAlt_1)$ & = & $(\Lens_2.\PutROf{\String_2}{\StringAlt_2})(\Lens_1.\PutROf{\String_1}{\StringAlt_1})$\\
-    $\PutL{} \App (\StringAlt_2\StringAlt_1) \App (\String_1\String_2)$ & = & $(\Lens_1.\PutLOf{\StringAlt_1}{\String_1})(\Lens_2.\PutLOf{\StringAlt_2}{\String_2})$\\
-  \end{tabular}
-\end{center}
+%\begin{center}
+%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+%    $\CreateR{} \App \String_1\String_2$ & = & $(\Lens_2.\CreateROf{\String_2})(\Lens_1.\CreateROf{\String_1})$\\
+%    $\CreateL{} \App \StringAlt_2\StringAlt_1$ & = & $(\Lens_1.\CreateLOf{\StringAlt_1})(\Lens_2.\CreateLOf{\StringAlt_2})$\\
+%    $\PutR{} \App (\String_1\String_2) \App (\StringAlt_2\StringAlt_1)$ & = & $(\Lens_2.\PutROf{\String_2}{\StringAlt_2})(\Lens_1.\PutROf{\String_1}{\StringAlt_1})$\\
+%    $\PutL{} \App (\StringAlt_2\StringAlt_1) \App (\String_1\String_2)$ & = & $(\Lens_1.\PutLOf{\StringAlt_1}{\String_1})(\Lens_2.\PutLOf{\StringAlt_2}{\String_2})$\\
+%  \end{tabular}
+%\end{center}
+The swap combinator is similar to concat, though the second regular expression
+is swapped.
 
@@
 \end{center}
+The \OrLens lens deals with data that can come in one form or another. If the
+data gets changed from one format to the other, information in the old format is
+lost.
 
@@
 \]
-\begin{center}
-  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
-    $\CreateR{} \App \String_1\ldots\String_n$
-    & =
-    & $(\Lens.\CreateROf{\String_1})\ldots(\Lens.\CreateROf{\String_n})$\\
-
-    $\CreateL{} \App \StringAlt_1\ldots\StringAlt_n$
-    & =
-    & $(\Lens.\CreateLOf{\StringAlt_1})\ldots(\Lens.\CreateLOf{\StringAlt_n})$\\
-
-    $\PutR{} \App (\String_1\ldots\String_n) \App (\StringAlt_1\ldots\StringAlt_m)$
-    & =
-    & $\StringAlt_1'\ldots\StringAlt_n'$ where $\StringAlt_i' =
-      \begin{cases*}
-        \Lens.\PutROf{\String_i}{\StringAlt_i} & if $i \leq m$\\
-        \Lens.\CreateROf{\String_i} & otherwise
-      \end{cases*}$\\
-    $\PutL{} \App (\StringAlt_1\ldots\StringAlt_m) \App (\String_1\ldots\String_n)$
-    & =
-    & $\String_1'\ldots\String_n'$ where $\String_i' =
-      \begin{cases*}
-        \Lens.\PutROf{\StringAlt_i}{\String_i} & if $i \leq n$\\
-        \Lens.\CreateROf{\StringAlt_i} & otherwise
-      \end{cases*}$
-  \end{tabular}
-\end{center}
+%\begin{center}
+%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+%    $\CreateR{} \App \String_1\ldots\String_n$
+%    & =
+%    & $(\Lens.\CreateROf{\String_1})\ldots(\Lens.\CreateROf{\String_n})$\\
+%    
+%    $\CreateL{} \App \StringAlt_1\ldots\StringAlt_n$
+%    & =
+%    & $(\Lens.\CreateLOf{\StringAlt_1})\ldots(\Lens.\CreateLOf{\StringAlt_n})$\\
+%    
+%    $\PutR{} \App (\String_1\ldots\String_n) \App (\StringAlt_1\ldots\StringAlt_m)$
+%    & =
+%    & $\StringAlt_1'\ldots\StringAlt_n'$ where $\StringAlt_i' =
+%      \begin{cases*}
+%        \Lens.\PutROf{\String_i}{\StringAlt_i} & if $i \leq m$\\
+%        \Lens.\CreateROf{\String_i} & otherwise
+%      \end{cases*}$\\
+%    $\PutL{} \App (\StringAlt_1\ldots\StringAlt_m) \App (\String_1\ldots\String_n)$
+%    & =
+%    & $\String_1'\ldots\String_n'$ where $\String_i' =
+%      \begin{cases*}
+%        \Lens.\PutROf{\StringAlt_i}{\String_i} & if $i \leq n$\\
+%        \Lens.\CreateROf{\StringAlt_i} & otherwise
+%      \end{cases*}$
+%  \end{tabular}
+%\end{center}
+The \IterateLens lens deals with iterated data.
 
@@
 \end{center}
+The \MergeR lens is interesting because it merges data where one data can be in
+two formats, and one data has only one format. In previous work, this was
+combined into \OrLens{}, where \OrLens{} could have ambiguous types, but we
+find it more clear to have explicit merge operators: it is easier to see what
+lens the synthesis algorithm is creating.
 
@@
 \end{center}
+The \MergeL lens is symmetric to \MergeR.
 
@@
 \end{center}
+Composing is interesting in the put functions. Because puts require intermediary
+data, we recreate that intermediary data with creates.
 
@@
 \]
-\begin{center}
-  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
-    $\CreateR{} \App \StringAlt$ & = & $\Lens.\CreateLOf{\StringAlt}$\\
-    $\CreateL{} \App \String$ & = & $\Lens.\CreateROf{\String}$\\
-    $\PutR{} \App \StringAlt \App \String$ & = & $\Lens.\PutLOf{\StringAlt}{\String}$\\
-    $\PutL{} \App \String \App \StringAlt$ & = & $\Lens.\PutROf{\String}{\StringAlt}$
-  \end{tabular}
-\end{center}
+%\begin{center}
+%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+%    $\CreateR{} \App \StringAlt$ & = & $\Lens.\CreateLOf{\StringAlt}$\\
+%    $\CreateL{} \App \String$ & = & $\Lens.\CreateROf{\String}$\\
+%    $\PutR{} \App \StringAlt \App \String$ & = & $\Lens.\PutLOf{\StringAlt}{\String}$\\
+%    $\PutL{} \App \String \App \StringAlt$ & = & $\Lens.\PutROf{\String}{\StringAlt}$
+%  \end{tabular}
+%\end{center}
+Inverting is particularly useful when chaining many compositions together, as it
+can be used to align the central types.
+
+\[
+  \centering
+  \inferrule*
+  {
+    \Lens \OfType \Regex \Leftrightarrow \RegexAlt\\
+    \Regex \equiv \Regex'\\
+    \RegexAlt \equiv \RegexAlt'
+  }
+  {
+    \Lens \OfType \Regex' \Leftrightarrow \RegexAlt'
+  }
+\]
+Type equivalence is useful both for addressing type annotations and for
+letting compositions go through.
+%\begin{center}
+%  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+%    $\CreateR{} \App \StringAlt$ & = & $\Lens.\CreateLOf{\StringAlt}$\\
+%    $\CreateL{} \App \String$ & = & $\Lens.\CreateROf{\String}$\\
+%    $\PutR{} \App \StringAlt \App \String$ & = & $\Lens.\PutLOf{\StringAlt}{\String}$\\
+%    $\PutL{} \App \String \App \StringAlt$ & = & $\Lens.\PutROf{\String}{\StringAlt}$
+%  \end{tabular}
+%\end{center}
+Inverting is particularly useful when chaining many compositions together, as it
+can be used to align the central types.
 % end ssl
@@
 \section{Synthesis}
-In this work, we aim to quickly find a good lens between two regular expressions
-that satisfies the input-output examples. To this end, we develop an information
-theoretic metric for how much data is synchronized. Information theoretic
-metrics require a notion of probability distributions \emph{how are probability
-  distributions being used, are certain things weighted more etc, assign weights
-  to different subcomponents and parts of the data}. We annotate
-the regular expressions with a reasonable probability distribution, making them
-\emph{stochastic regular expressions}. Armed with probability distributions over
-the languages, we have the capabilities to define information theoretic metrics
-over these probability distributions.
-
-Our overall synthesis algorithm utilizes two synthesizers shown in
-Figure~\ref{fig:high-level-algorithm}. There are an infinite number of possible
-lenses between two regular expressions that can be searched through. Instead of
-searching through all possible lenses, the lenses are grouped into classes of
-lenses based on the syntax of the regular expressions they synchronize. Two
-lenses are in the same class if they can be typed by the same regular
-expressions, using no star equivalences. This works well for synthesis, as
-finding a good lens within these classes permits an efficient synthesis
-procedure.
-
-However, proposing these classes is complicated by stochastic regular
-expressions. While the equivalences on regular expressions are well studied, the
-equivalences on stochastic regular expressions are not. We want to propose
-equivalent stochastic regular expressions, but both the languages and
-probability distributions must be maintained. If the probability distributions
-are not maintained, then the metric by which we compare our lenses changes for
-each class of lenses, we can't compare lenses between classes. To alleviate
-this, we describe the \emph{star-semiring} equivalences on stochastic regular
-expressions, and prove they are sound with respect to language and probability
-distributions. Our \RXSearch proposes stochastic regular expressions by
-traversing a subset of these equivalences, so \RXSearch only proposes regular
-expression pairs equivalent to the originals.
-
-The second algorithm, \GreedySynth, searches within a class for the lowest cost
-lens.  It is a greedy algorithm, and not guaranteed to find the lowest cost
-lens.  However, we haven't found any instances in which \GreedySynth doesn't find
-the optimal lens, due to the partial backtracking we implemented. \GreedySynth
-works by searching, not through symmetric lenses, but through an alternative
-language of Symmetric DNF lenses. After finding such a symmetric DNF lens, that
-symmetric DNF lens is converted into a simple symmetric lens.
+Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm. The
+provided regular expressions are first converted into stochastic regular
+expressions with \ToStochastic. Next, $\RXSearch$ is initialized with that pair
+of stochastic regular expressions. $\RXSearch$ then proposes candidate classes
+of lenses by proposing regular expression pairs that characterize those classes.
+$\GreedySynth$ greedily finds a low cost lens within each proposed lens class,
+if there is one that satisfies the specification. These lenses are then provided
+to $\RXSearch$. Eventually, $\RXSearch$ is content with one of the lenses, and
+returns that lens to the user.
+
+\begin{algorithm}
+  \caption{\SynthSymLens}
+  \label{alg:synth-sym-lens}
+  \begin{algorithmic}[1]
+    \Function{SynthSymLens}{$\Regex,\RegexAlt,\Examples$}
+    \State $\Regex \gets \Call{\ToStochastic}{\Regex}$
+    \State $\RegexAlt \gets \Call{\ToStochastic}{\RegexAlt}$
+    \State $\RXSearchState \gets \Call{\RXSearch.Init}{\Regex,\RegexAlt}$
+    \While{$\Call{\RXSearch.Continue}{\RXSearchState}$}
+    \State $(\Regex,\RegexAlt,\RXSearchState) \gets \Call{\RXSearch.Propose}{\RXSearchState}$
+    \State $\LCO \gets \Call{\GreedySynth}{\Examples,\Regex,\RegexAlt}$
+    \State $\RXSearchState \gets \Call{\RXSearch.Consider}{\RXSearchState,\LCO}$
+    \EndWhile
+    \ReturnVal{\Call{RXSearch.Best}{\RXSearchState}}
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
+%In this work, we aim to quickly find a good lens between two regular expressions
+%that satisfies the input-output examples. To this end, we develop an information
+%theoretic metric for how much data is synchronized. Information theoretic
+%metrics require a notion of probability distributions \emph{how are probability
+%  distributions being used, are certain things weighted more etc, assign weights
+%  to different subcomponents and parts of the data}. We annotate the regular
+%expressions with a reasonable probability distribution, making them
+%\emph{stochastic regular expressions}. Armed with probability distributions over
+%the languages, we have the capabilities to define information theoretic metrics
+%over these probability distributions.
+%
+%Our overall synthesis algorithm utilizes two synthesizers shown in
+%Figure~\ref{fig:high-level-algorithm}. There are an infinite number of possible
+%lenses between two regular expressions that can be searched through. Instead of
+%searching through all possible lenses, the lenses are grouped into classes of
+%lenses based on the syntax of the regular expressions they synchronize. Two
+%lenses are in the same class if they can be typed by the same regular
+%expressions, using no star equivalences. This works well for synthesis, as
+%finding a good lens within these classes permits an efficient synthesis
+%procedure.
+%
+%However, proposing these classes is complicated by stochastic regular
+%expressions. While the equivalences on regular expressions are well studied, the
+%equivalences on stochastic regular expressions are not. We want to propose
+%equivalent stochastic regular expressions, but both the languages and
+%probability distributions must be maintained. If the probability distributions
+%are not maintained, then the metric by which we compare our lenses changes for
+%each class of lenses, we can't compare lenses between classes. To alleviate
+%this, we describe the \emph{star-semiring} equivalences on stochastic regular
+%expressions, and prove they are sound with respect to language and probability
+%distributions. Our \RXSearch proposes stochastic regular expressions by
+%traversing a subset of these equivalences, so \RXSearch only proposes regular
+%expression pairs equivalent to the originals.
+%
+%The second algorithm, \GreedySynth, searches within a class for the lowest cost
+%lens.  It is a greedy algorithm, and not guaranteed to find the lowest cost
+%lens.  However, we haven't found any instances in which \GreedySynth doesn't find
+%the optimal lens, due to the partial backtracking we implemented. \GreedySynth
+%works by searching, not through symmetric lenses, but through an alternative
+%language of Symmetric DNF lenses. After finding such a symmetric DNF lens, that
+%symmetric DNF lens is converted into a simple symmetric lens.
 
@@
 
-Much like the synthesis algorithm for Optician, we need to apply rewrites to
-handle equivalences on the provided stochastic regular expressions.  However, no
-such equivalences have yet been defined.  We extend the \emph{star-semiring}
-equivalences to stochastic regular expressions.
+$\RXSearch$ needs to be able to find equivalent stochastic regular expressions.
+However, as of yet, there is no syntactic means to reason about stochastic
+regular expression equivalence. We extend the \emph{star-semiring} equivalences
+to stochastic regular expressions.
 
@@
 \end{center}
-% \caption{Regular Expression Equivalences}
-% \label{fig:regex-equivalence-rules}
 
@@
 \end{theorem}
+
+Finally, we wish to infer a reasonable probability distribution for stochastic
+regular expressions.  While these probability distributions could be either
+input by users or inferred from example data, we found that our inferred
+probability distributions work in practice. We defer our description of
+\ToStochastic until \S\ref{sec:dnf}, as it is easier to describe our desired
+probability distribution in the context of DNF regular expressions.
+% \caption{Regular Expression Equivalences}
+% \label{fig:regex-equivalence-rules}
 
@@
 \RXSearch stops proposing regular expression pairs, and instead returns to the
-user the best lens found thus far.  Note, if \RXSearch finds a bijective lens (a
-lens of cost 0), it will immediately return, as would be expected.
-
-\afm{dont forget to add in description of termination}
+user the best lens found thus far. Note this means that if \RXSearch finds a
+bijective lens (a lens of cost 0) it will immediately return.
 %
-
-\subsection{\GreedySynth}
-\label{subsec:greedy-synth}
-The second synthesizer, \GreedySynth, actually generates the lenses. It does
-this by first converting the regular expressions into \emph{stochastic DNF
-  regular expressions}, intuitively stochastic regular expressions with all
-disjunctions distributed over all concatenations. Next, a good DNF lens is
-greedily found between these stochastic DNF regular expressions, using the
-probability annotations on the DNF regular expressions to find the cost.
-Finally, this DNF lens is converted into a simple symmetric lens.
-
-\paragraph*{Stochastic DNF Regular Expressions}
-\label{subsec:stoch-dnf-regex}
-We extend stochastic regular expressions to stochastic DNF regular expressions.
+\subsection{Stochastic DNF Regular Expressions}
+Stochastic DNF regular expressions are used as a highly normalized form of
+stochastic regular expressions. Due to this highly normalized nature, where type
+equivalence is often needed for simple symmetric lenses, such rules aren't
+needed when typed between DNF regular expressions, as many semantically equal
+but syntactically distinct regular expressions have syntactically equal DNF
+forms.
+
+Stochastic DNF regular expressions are hierarchically built, where stochastic
+DNF regular expressions are lists of stochastic sequences. Stochastic sequences
+themselves are lists of interleaved strings and stochastic atoms.  Stochastic
+atoms are iterated stochastic DNF regular expressions.
 \begin{center}
@@
 
-As before, DNF regular expressions are essentially regular expressions with all
-concatenations fully distributed over all disjunctions.  As these DNF regular
-expressions are \emph{stochastic}, they are annotated with probabilities to
-express a probability distribution, in addition to a language.  This probability
-distribution is defined below:
-
+Intuitively, stochastic DNF regular expressions are essentially stochastic
+regular expressions with all concatenations fully distributed over all
+disjunctions. As such, the language of a stochastic DNF regular expression is a
+union of its subcomponents, the language of a stochastic sequence is the
+concatenation of its subcomponents, and the language of a stochastoc atom is the
+iteration of its subcomponent.
+
+\begin{trivlist}
+  \centering
+\item 
+  % \begin{center}
+  \begin{tabular}{@{\ }v@{\ }q}
+    \LanguageOf{\PRegexStar{\DNFRegex{}}{\Probability}} \ =\  &
+                                            \{\String_1\Concat\ldots\Concat\String_n
+                                            \SuchThat \forall i \String_i\in\LanguageOf{\DNFRegex}\}\\
+    \LanguageOf{\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n}}\ =\  & 
+    % \\
+    % \multicolumn{2}{L}{\ \ \ \ 
+    \{\String_0\Concat\StringAlt_1\cdots\StringAlt_n\Concat\String_n \SuchThat \StringAlt_i\in\LanguageOf{\Atom_i}\}
+    % }
+    \\
+    \LanguageOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}\ =\  &
+    % \\
+    % \multicolumn{2}{L}{\ \ \ \ 
+    \{\String \SuchThat \String \in \LanguageOf{\Sequence_i} \text{\ and $i\in\RangeIncInc{1}{n}$}\}
+    % }
+  \end{tabular}
+  % \end{center}
+\end{trivlist}
+
+As these DNF regular expressions are \emph{stochastic}, they are annotated with
+probabilities to express a probability distribution, in addition to a language.
 \begin{center}
@@
 We have a means of turning stochastic regular expressions into stochastic DNF
-regular expressions.
-The conversion algorithm itself, written $\ToDNFRegexOf{\Regex}$, is defined below.
+regular expressions. The conversion algorithm itself, written
+$\ToDNFRegexOf{\Regex}$, is defined below.
 \[
@@
 \]
-We have also proven an algorithm, demonstrating this conversion respects
-probability distributions.
+We have proven this conversion respects languages and probability distributions.
 
 \begin{theorem}
-  $\ProbabilityOf{\Regex}{\String} = \ProbabilityOf{\ToDNFRegex \Regex}{\String}$
+  $\ProbabilityOf{\Regex}{\String} = \ProbabilityOf{\ToDNFRegex
+    \Regex}{\String}$ and $\LanguageOf{\Regex} =
+  \LanguageOf{\ToDNFRegexOf{\Regex}}$
 \end{theorem}
 
-Furthermore, in the situation where a DNF regular expression is unambiguous, we
-have defined an efficient, syntactic means to find the entropy of the DNF
-regular expression.
+\paragraph*{\ToStochastic} With stochastic DNF regular expressions and
+\ToDNFRegex defined, it is easier to explain the function that converts regular
+expressions into stochastic regular expressions, \ToStochastic. If $\Regex$ is a
+stochastic DNF regular expression generated by \ToStochastic, then
+$\ToDNFRegexOf{\Regex} = \DNFOf{(\Sequence_1,\frac{1}{n}) \DNFSep \ldots \DNFSep
+  (\Sequence_n,\frac{1}{n})}$ for some sequences $\Sequence_1\ldots\Sequence_n$,
+and every stochastic atom generated by \ToStochastic is
+$\PRegexStar{\DNFRegex}{.8}$. In particular, $\ToDNFRegex$ generates regular
+expressions whose DNF form gives equal probability to all sequences, and gives
+iterations a .8 chance to continue iterating.
+
+\paragraph*{Entropy} In the situation where a DNF regular expression is
+unambiguous, we have defined an efficient, syntactic means to find the entropy
+of the DNF regular expression.
 
@@
 
-\paragraph*{Symmetric DNF, Sequence, and Atom Lenses}
-\label{sec:symmetric-dnf-lenses}
-Symmetric DNF Lenses are what we synthesize, and we then turn them into
-Symmetric Lenses. The nonterminals $i$, $j$, $p$, $q$, $r$, $c$, and $d$ all
+\subsection{Symmetric DNF Lenses}
+Symmetric DNF lenses are an intermediate synthesis target for \GreedySynth.
+Because of the lack of a type equivalence rule or a composition rule, a
+symmetric DNF lens is syntactically quite similar to the stochastic DNF regular
+expression pair it is typed between. This similarity is part of what permits
+such an efficient type-directed synthesis algorithm of them.
+
+\paragraph*{Syntax}
+Like stochastic DNF regular expressions, symmetric DNF Lenses are hierarchically
+structured. Symmetric DNF lenses consist of a list of symmetric sequence lenses,
+as well as some additional information. Symmetric sequence lenses consist of a
+list of symmetric atom lenses, as well as some additional information, and
+symmetric atom lenses are merely iterations of symmetric DNF lenses. In the
+following syntax, the nonterminals $i$, $j$, $p$, $q$, $r$, $c$, and $d$ all
 represent natural numbers. The nonterminals $\String$ and $\StringAlt$ represent
 strings.
-
 \begin{center}
@@
 
-Lets go through these from bottom to top.
-
-A DNF Lens \SDNFLens{} consists of three components. First is the mapping
-component. The mapping component is a list of triples $(i,j,\SSQLens)$. In this,
-$i$ and $j$ refer to sequence $i$ and sequence $j$ in the left and right DNF
-regular expressions, respectively. The third component, \SSQLens{}, is a
-sequence lens that maps between the two sequences specified by $i$ and $j$. The
-second component is a list of integers $c$. The integer, $c_i$ at position $i$
-states that when performing a $\CreateR$ on a string matching sequence
-$\Sequence_i$, it will use the sequence lens that maps it to sequence
-$\SequenceAlt_{c_i}$ on the right. The third component is a list of integers
-$d$, and follows the same pattern as the second component, but maps from right
-to left.
+In greater detail, a symmetric DNF Lens \SDNFLens{} consists of three
+components. First is the mapping component. The mapping component is a list of
+triples $(i,j,\SSQLens)$. In this, $i$ and $j$ refer to sequence $i$ and
+sequence $j$ in the left and right DNF regular expressions, respectively. The
+third component, \SSQLens{}, is a sequence lens that maps between the two
+sequences specified by $i$ and $j$. The second component is a list of integers
+$c$. The integer, $c_i$ at position $i$ states that when performing a $\CreateR$
+on a string matching sequence $\Sequence_i$, it will use the sequence lens that
+maps it to sequence $\SequenceAlt_{c_i}$ on the right. The third component is a
+list of integers $d$, and follows the same pattern as the second component, but
+maps from right to left.
 
@@
 
-An atom lens \AtomLens{} is merely the iteration of a symmetric DNF Lens.
-
-Because these lenses only encode functions on well-typed programs, we introduce
-the semantics alongside the typing rules.  The semantics are over typing
-derivations, instead of over individual lenses.
-
-\paragraph*{Symmetric DNF Lenses, Typing and Semantics}
+As before, an atom lens \AtomLens{} is merely the iteration of a symmetric DNF
+Lens.
+
+\paragraph*{Symmetric DNF Lenses: Typing and Semantics}
 The typing judgment is a 3-ary relation over a single DNF lens, and two DNF
@@
 
-\paragraph*{Sequence Lenses, Typing and Semantics}
+\paragraph*{Symmetric Sequence Lenses: Typing and Semantics}
 The typing judgment is a 3-ary relation over a single Sequence lens, and two
@@
 
-\paragraph*{Atom Lenses, Typing and Semantics}
+\paragraph*{Symmetric Atom Lenses: Typing and Semantics}
 The typing judgment is a 3-ary relation over a single atom lens, and two
@@
 
-\paragraph*{Greedy Algorithm}
+\subsection{\GreedySynth}
+\label{subsec:greedy-synth}
+
 The greedy algorithm is actually two greedy algorithms, one that greedily finds
-symmetric DNF lenses and one that greedily finds symmetric sequence lenses.
-
-The algorithm for symmetric DNF lenses is relatively straightforward.  First,
-using examples, we identify which atoms in one format \emph{must} be
-synchronized to certain atoms in the other.  After that, we iteratively find the
-pairs of atoms with lowest cost, where at least one of the atoms is not mapped
-to another atom.  The cost of these pairs is not merely the cost of the lens,
-but also the cost incurred by adding that lens, if one of the atoms is already
-involved in a map, the lens has additional cost to account for the merge cost.
-This continues until there are no more atoms that aren't mapped to another atom.
-At this point, the creates for each lens is determined.  If a examples dictate
-one lens must create to another, that is the create behavior, otherwise atoms
-create along the lens of least cost.
-
-The algorithm for symmetric sequence lenses is less straightforward.  Unlike
-symmetric DNF lenses, where multiple subcomponents can map to the same
-subcomponent, each atom can map to at most one other subcomponent.  This is
-complicated by the fact that examples sometimes dictate one atom must map to
-another.
-
-To address this, we introduce the \emph{profit} of a adding a mapping between
-two subcomponents. The profit is the cost of the lens, minus the information
-content of the subcomponents it maps between, plus how much profit would be lost
-by freeing those subcomponents to make the subcomponents available. Then, lenses
-are only added if they have positive profit, or if one of the atoms requires a
-mapping. After there are no more lenses with positive profit or with
-subomponents requiring a mapping, the sequence lens is returned.
+symmetric DNF lenses (\GreedySynth) and one that greedily finds symmetric
+sequence lenses (\GreedySeqSynth).
+
+\begin{algorithm}
+  \caption{\GreedySynth}
+  \label{alg:greedysynth}
+  \begin{algorithmic}[1]
+    \Function{GreedySynth}{$\Examples,\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)},\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}$}
+    \If{$\Call{CannotMap}{\Examples,\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)},\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}$}
+    \State $\ReturnVal{\None}$
+    \EndIf
+    \State $\SLS \gets
+    \Call{\CartesianMap}{\GreedySeqSynth(\Examples),\ListOf{\Sequence_1;\ldots;\Sequence_n},\ListOf{\SequenceAlt_1;\ldots;\SequenceAlt_m}}$
+    \State $\GreedyState \gets \Call{GreedyState.Init}{\SLS}$
+    \While{$\Call{GreedyState.Continue}{\GreedyState}$}
+    \State $(\GreedyState,\SSQLens) \gets
+    \Call{GreedyState.PopBest}{\GreedyState}$
+    \If{$\Call{GreedyState.UsefulAdd}{\GreedyState,\SSQLens}$}
+    \State $\GreedyState \gets \Call{GreedyState.AddSeq}{\GreedyState,\SSQLens}$
+    \EndIf
+    \EndWhile
+    \ReturnVal{\Call{GreedyState.ToDNFLens}{\GreedyState}}
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
+\paragraph*{Symmetric DNF Lenses} Algorithm~\ref{alg:greedysynth} presents
+\GreedySynth, a means to synthesize symmetric DNF lenses. The inputs to
+\GreedySynth are a pair of stochastic DNF regular expressions and a suite of
+input-output examples. First, \PCF{CannotMap} checks if there is possibly a lens
+between the given DNF regular expressions that is consistent with the examples.
+If there is not, \GreedySynth returns \None early. After this, \GreedySynth
+finds the best lenses between each sequence pair of the left and right DNF
+regular expressions. An internal state for the algorithm is initialized from
+these. This state consists of two things: an intermediate partial DNF Lens and a
+priority queue consisting of potential sequence lenses. On initialization,
+\GreedySynth checks which sequence lenses are \emph{required} to be included,
+from the examples. Those are added into the partial lens, and the rest are
+thrown into the priority queue.
+
+\GreedySynth then loops until there are no more sequence lenses in the priority
+queue. Within this loop, sequence lenses are popped. If a popped sequence lens
+maps between two sequences which aren't yet included in a sequence lens, then
+this sequence lens is added to the partial DNF lens. Furthermore, we update the
+priorities of our sequence lenses as the algorithm proceeds: if two sequence
+lenses have the same source, the second one to be popped gets a higher cost than
+it originally had, to address the information that needs to be stored for
+including that source of non-bijectivity. Finally, after all sequences have been
+popped, the partial DNF lens is converted into a symmetric DNF lens. This
+is only possible if all sequences are involved in some sequence lens: if they
+are not, \PCF{GreedyState.ToDNFLens} instead returns \None.
+
+\begin{algorithm}
+  \caption{\GreedySeqSynth}
+  \label{alg:greedyseqsynth}
+  \begin{algorithmic}[1]
+    \Function{AtomSynth}{$\Examples,\PRegexStar{\DNFRegex}{\Probability},\PRegexStar{\DNFRegexAlt}{\ProbabilityAlt}$}
+    \If{$\Call{CannotMap}{\Examples,\PRegexStar{\DNFRegex}{\Probability},\PRegexStar{\DNFRegexAlt}{\ProbabilityAlt}}$}
+    \State $\ReturnVal{\None}$
+    \Else
+    \State $\ReturnVal{\Call{Map}{\PCF{Iterate},\GreedySynth(\Examples,\DNFRegex,\DNFRegexAlt)}}$
+    \EndIf
+    \EndFunction
+
+    \Function{GreedySeqSynth}{$\Examples,\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n},\SequenceOf{\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\String_m}$}
+    \If{$\Call{CannotMap}{\Examples,\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n},\SequenceOf{\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\String_m}}$}
+    \State $\ReturnVal{\None}$
+    \EndIf
+    \State $\ALS \gets
+    \Call{\CartesianMap}{\AtomSynth(\Examples),\ListOf{\Atom_1;\ldots;\Atom_n},\ListOf{\AtomAlt_1;\ldots;\AtomAlt_m}}$
+    \State $\GreedyState \gets \Call{GreedySeqState.Init}{\ALS}$
+    \While{$\Call{GreedySeqState.Continue}{\GreedyState}$}
+    \State $(\GreedyState,\SAtomLens) \gets
+    \Call{GreedySeqState.PopBest}{\GreedyState}$
+    \If{$\Call{GreedySeqState.UsefulAdd}{\GreedyState,\SAtomLens}
+      \BooleanOr \Call{NecessaryAdd}{\Examples,\SAtomLens}$}
+    \State $\GreedyState \gets \Call{GreedyState.AddAtom}{\GreedyState,\SAtomLens}$
+    \EndIf
+    \EndWhile
+    \ReturnVal{\Call{GreedyState.ToDNFLens}{\GreedyState}}
+    \EndFunction
+  \end{algorithmic}
+\end{algorithm}
+
+\paragraph*{Symmetric Sequence lenses} Algorithm~\ref{alg:greedyseqsynth}
+presents \GreedySeqSynth, a means to synthesizes symmetric sequence lenses. This
+algorithm uses \PCF{AtomSynth}, a simple function that synthesizes atom lenses
+by iterating the DNF lens between its subcomponents. The inputs to
+\GreedySeqSynth are a pair of stochastic atoms and a suite of input-output
+examples. First, \PCF{CannotMap} checks if there is a possibility of a lens
+between the given sequences regular expressions that is consistent with the
+examples. If there is not, \GreedySeqSynth returns \None early. After this,
+\GreedySeqSynth finds the best lenses between each atom pair of the left and
+right sequences. An internal state for the algorithm is initialized from these.
+This state consists of two things: an intermediate partial sequence lens and a
+priority queue consisting of potential sequence lenses.  On initialization,
+\GreedySeqSynth throws all the atom lenses into the priority queue.
+
+\GreedySeqSynth then loops until there are no more atom lenses in the priority
+queue. Within this loop, atom lenses are popped. The atom lens is then added if
+it is either necessary, or useful. Within a sequence lens, each atom can either
+be projected or be synchronized with exactly one atom in the other sequence.
+Certain atoms need to be mapped, for example if a put changes the data of that
+field. In such situations, an atom lens is necessary. A useful atom lens is
+where adding that lens reduces the overall cost of the partial lens. If adding
+that lens requires removing another lens, where the added cost by removing that
+other lens is more than is removed by adding the proposed lens. Finally, after
+all atoms have been popped, the partial sequence lens is converted into a
+symmetric DNF lens. This is only possible if all atoms that need to be mapped,
+are: if they are not, \PCF{GreedyState.ToDNFLens} instead returns \None.
 

### 2018-06-04 papers/SymmetricLenses/main.tex
add stuff

@@
       \BooleanOr \Call{NecessaryAdd}{\Examples,\SAtomLens}$}
-    \State $\GreedyState \gets \Call{GreedyState.AddAtom}{\GreedyState,\SAtomLens}$
+    \State $\GreedyState \gets \Call{GreedySeqState.AddAtom}{\GreedyState,\SAtomLens}$
     \EndIf
     \EndWhile
-    \ReturnVal{\Call{GreedyState.ToDNFLens}{\GreedyState}}
+    \ReturnVal{\Call{GreedySeqState.ToDNFLens}{\GreedyState}}
     \EndFunction
@@
 right sequences. An internal state for the algorithm is initialized from these.
-This state consists of two things: an intermediate partial sequence lens and a
-priority queue consisting of potential sequence lenses.  On initialization,
-\GreedySeqSynth throws all the atom lenses into the priority queue.
+This state consists of two things: an intermediate sequence lens and a priority
+queue consisting of potential sequence lenses. On initialization,
+\GreedySeqSynth throws all the atom lenses into the priority queue, and the
+intermediate sequence lens is one that projects all the atoms.
 
@@
 it is either necessary, or useful. Within a sequence lens, each atom can either
-be projected or be synchronized with exactly one atom in the other sequence.
-Certain atoms need to be mapped, for example if a put changes the data of that
-field. In such situations, an atom lens is necessary. A useful atom lens is
-where adding that lens reduces the overall cost of the partial lens. If adding
-that lens requires removing another lens, where the added cost by removing that
-other lens is more than is removed by adding the proposed lens. Finally, after
-all atoms have been popped, the partial sequence lens is converted into a
-symmetric DNF lens. This is only possible if all atoms that need to be mapped,
-are: if they are not, \PCF{GreedyState.ToDNFLens} instead returns \None.
+be projected or be synchronized with exactly one atom in the other sequence, and
+examples can show that certain atoms \emph{need} to be mapped. In such
+situations, that atom lens is necessary. A useful atom lens is where adding that
+lens reduces the overall cost of the partial lens. If both atoms are not
+involved in an atom lens, the lens will always be useful, but sometimes adding
+that lens requires removing another lens. If the cost added by removing that
+other lens is more than is removed by adding the proposed lens, then this atom
+lens is not included in the sequence lens. Finally, after all atoms have been
+popped, the sequence lens is validated and returned. The sequence lens is
+validated if all atoms that need to be mapped, are: if they are not,
+\PCF{GreedySeqState.ToDNFLens} instead returns \None.
 

### 2018-06-05 papers/SymmetricLenses/main.tex
light editing and some comments through the end of section 2

@@
 \begin{abstract}
+\saz{This abstract too long and doesn't get to the point quickly enough. We
+  should tighten it up considerably.}
   Users commonly need to synchronize the data shared between formats, so edits
@@
 
-Unfortunately, writing such synchronizers is tedious and error prone.
-Bidirectional languages ease the difficulty of writing such synchronizers, as a
-single, well-typed program expresses a \emph{lens}, a group of conversion
+Unfortunately, writing such synchronizers is tedious and error
+prone.\saz{``tedious and error prone'' is such a PL clich\'{e}, can we do better?}
+Bidirectional languages ease the difficulty of writing such synchronizers, because a
+single, well-typed program expresses a \emph{lens}---a group of conversion
 functions guaranteed to satisfy round-tripping laws. However, bidirectional
-languages are difficult to program in as well, often requiring users to permute
+languages are themselves difficult to program in, often requiring users to permute
 data in fiddly ways while satisfying the type system's complex unambiguity
@@
 
-A better approach is to synthesize such conversion functions. Prior work on
-synthesizing these functions assumes that all data fields present in one format
-are present in the other, but this is not always true. Oftentimes, fields
-present in one format are not present in the other: the commands scheduled to
-run on Linux machines are typically different than the commands scheduled to run
-on Windows machines; updated APIs can expose new information, or omit legacy
-information; ICS files contain a URL field that Google Calendar CSV files do not
-expose.
-
-\emph{Symmetric lenses} formalize such conversion functions. However, these
-symmetric lenses use external state for synchronization, in the form of a
-\emph{complement}. These complements makes symmetric lenses difficult to use in
-real-world synchronization tasks, as users must maintain the synchronized files
-as well as the complements. The complement further complicates synthesis, as
-input-output examples are no longer sufficient, instead edit sequences must be
-provided.
-
-To get around this, we reformulate symmetric lenses as \emph{simple symmetric
-  lenses}. This reformulation does not use a complement, and only uses the files
-that we aim to synchronize as state. We identify a property,
-\emph{forgetfulness}, where if a symmetric lens can be written as a forgetful
-symmetric lens, it can also be written a simple symmetric lens. We provide
+A better approach is to \emph{synthesize} such conversion functions. However,
+prior work on synthesizing lenses assumes a bijection between the source and
+target, which is a severe limitation in practice. Oftentimes, fields present in
+one format are not present in the other: the commands scheduled to run on Linux
+machines are typically different than the commands scheduled to run on Windows
+machines; updated APIs can expose new information, or omit legacy information;
+ICS files contain a URL field that Google Calendar CSV files do not expose.
+
+\emph{Symmetric lenses} \saz{citations?} formalize this richer class of
+non-bijective conversion functions by using additional state, called the
+\emph{complement}, to help with synchronization. While increasing
+expressiveness, the existence of the complement makes symmetric lenses difficult
+to use in real-world synchronization tasks, because users\saz{lens developers?
+  lens programs?}  must maintain the synchronized files as well as the
+complements. The complement further complicates synthesis, because input-output
+examples no longer suffice; instead edit sequences must be provided.
+
+In this paper, we reformulate \saz{this sounds like they may be of the same
+  expressiveness, but perhaps we mean to ``constrain'' symmetric lenses?}
+symmetric lenses as \emph{simple symmetric lenses}. This reformulation does not
+use a complement, and only uses the files that we aim to synchronize as
+state. We identify a property, \emph{forgetfulness}, that identifies the subset
+of symmetric lenses that can be written as simple symmetric lenses. We provide
 combinators for building simple symmetric lenses on string data, and integrate
@@
 
-We provide a tool for synthesizing simple symmetric lenses. This tool
-synthesizes symmetric lenses from a suite of input-output examples and a pair of
-regular expressions describing the data formats. The synthesis algorithm uses
-type-directed synthesis on the two regular expressions to synthesize a lens that
-is well-typed between the provided regular expressions and satisfies the
-provided input-output examples.
-
+We provide a tool for synthesizing simple symmetric lenses when given a suite of
+input-output examples and a pair of regular expressions describing the data
+formats. The synthesis algorithm uses type-directed synthesis on the two regular
+expressions to synthesize a lens that is well-typed between the provided regular
+expressions and satisfies the provided input-output examples.
+
+\saz{This paragraph is hard to follow --- maybe introduce $\ell : S
+  \Leftrightarrow T$ and use the names $S$ and $T$?}
 A primary difficulty is that there are a relatively many well-typed simple
@@
 allow. This task gets harder when there are more complex decisions involved. For
-example, if I have two fields in one format, and one in the other, which field
+example, if there are two fields in one format, and one in the other, which field
 in the first format should be aligned to the sole field of the second format? Or
@@
 
-In particular, the cost of a lens is the expected number of bits required to
+In particular, the \emph{cost} of a lens is the expected number of bits required to
 recover a string in one format when a synchronized string in the other format is
@@
 
-Our model for this is based on \emph{stochastic regular expressions}~\cite{?}.
-Both a language and a probability distribution over that language is represented
-by a stochastic regular expression, and we develop a way to syntactically infer
-the expected information content (or \emph{entropy}) of an unambiguous regular
+Our model is based on \emph{stochastic regular expressions}~\cite{?}, which
+represent both a language and a probability distribution over that language
+simultaneously.  We develop a way to syntactically infer the expected
+information content (or \emph{entropy}) of an unambiguous regular
 expression. This entropy is used to calculate the expected information content
-to recover a string, when given a string it is synchronized to.
+to recover a string, when given a string it is synchronized with.
 
@@
 These proposed stochastic regular expression pairs have equivalent languages and
-probability distributions as the original ones: we prove the star-semiring
+probability distributions as the original ones: we prove that the star-semiring
 rewrites this synthesizer applies are sound rewrites on stochastic regular
@@
 formats from configuration files, application-specific data storage files, and
-data cleaning tasks. We find that our algorithm is able to synthesize all these
-lenses in under 20 seconds.
+data cleaning tasks. We find that our algorithm is able to synthesize all
+\saz{or each?}  of these lenses in under 20 seconds.
 
@@
 We motivate our problem with an extended example. Consider the previously
-described example of keeping Linux and Windows scheduled jobs synchronized. In
-this example, we describe keeping the file containing cron jobs (crontab) and a
-Scheduled Tasks file (scheduled.job) synchronized. These are each formats
-specifying when to run recurring tasks. For simplicity of description, we use a
-minified version of each of these, with examples and regular expressions shown
-in Figure~\ref{fig:minimized-representations}. Each file contains a list of
-tasks set to run at specific times every hour. In both these formats, users can
-either provide specific minutes they want to run the program during or they can
-specify they want to run the program at every minute.
+described goal of keeping Linux cron jobs (crontab) and Windows Scheduled Tasks
+(scheduled.job) synchronized. These are each formats specifying when to run
+recurring tasks. For the sake of exposition, we use simplified versions crontab
+and scheduled.job formats, with examples and corresponding regular expressions
+shown in Figure~\ref{fig:minimized-representations}. Each file contains a list
+of tasks set to run at specific times every hour. In both formats, users
+can either provide specific minutes at which they want the program to run, or they
+can specify that the program should run at every minute.
 
@@
 However, the commands that need to be run are typically different for each
-device. Paths are different in Windows and Linux, and while there is typically a
-Windows equvialent for Linux commands, that equivalent command is not always
-named the same thing. So while we want to be able to have the same number of
+device. Paths are different in Windows and Linux, and, while there is typically a
+Windows equvialent for a Linux command, that equivalent command is not always
+named the same thing. So, while we want to be able to have the same number of
 commands, with the commands running at the same times, we don't want to
@@
 
-To synchronize these files, a few functions are needed. Two \emph{creates} are
+\saz{Perhaps we should explicitly say here that we are trying to define a lens,
+  or describing the functions that a lens provides.}
+To synchronize these files, a few functions are needed. Two \emph{creates}
+\saz{we haven't introduced ``create'' as a noun yet.} are
 needed; if a crontab is initialized, a synchronized Scheduled Tasks file should
@@
     Note that we can annotate our lenses with types, to confirm the lens
-    operates on the languages of the regular expressions.}
+    operates on the languages of the regular expressions.\saz{Say something
+      about the possibility of synthesizing this lens?}}
   \label{fig:example_lens}
@@
 \subsection{Synthesizing Symmetric Lenses}
-To alleviate these difficulties, we wish to synthesize symmetric lenses from
-types and examples. Given a pair of regular expressions and a set of
-input-output examples, we want to synthesize a symmetric lens expression that is
-typed between the regular expression pair, and satisfies the input/output
-examples. In this example, we wish to synthesize a lens from \lstinline{crontab}
-and \lstinline{scheduled_job}, using the strings in
-Figure~\ref{fig:minimized-representations} as an input-output example for
-\CreateR.
+
+To alleviate these difficulties, we wish to synthesize symmetric lenses---given
+a pair of regular expressions and a set of input-output examples, we want to
+find a symmetric lens expression that is typed between the regular expression
+pair, and satisfies the input/output examples. In this example, we wish to
+synthesize a lens from \lstinline{crontab} and \lstinline{scheduled_job}, using
+the strings in Figure~\ref{fig:minimized-representations} as an input-output
+example for \CreateR.
 
@@
 
-But what is the best satisfying lens? We think the best satisfying lens is the
+But what is the best satisfying lens? Here, we take the view that the best satisfying lens is the
 one that aligns as much data as is permitted by the type system and the
@@
 The expected number of bits is a well known concept in information theory: the
-expected number of bits of a data source is the \emph{entropy} of that data
-source.
+\emph{entropy} of that data source.
 
@@
   S}P(s)\Log_2P(s)$. While we have ways of expressing sets of strings (with
-regular expressions), we do not have probability distributions. In this work, we
+regular expressions), we do not have probability distributions. Here, we
 express these probability distributions jointly with the language through
 \emph{stochastic regular expressions}, regular expressions annotated with
-probabilities. In this work, we guess a reasonable probability distribution from
+probabilities. Our algorithm guesses a reasonable probability distribution from
 the syntax, though if users manually provided stochastic regular expressions, or
-large suite of data in the formats through which we could learn the annotations,
+a large suite of data in the formats through which we could learn the annotations,
 we could instead use that.
@@
 minimizes the sum of that entropy with the conditional entropy of $\RegexAlt$
-given $\Regex$ and $\Lens$. Now that we have a notion of best, we need a way to
+given $\Regex$ and $\Lens$.
+
+Now that we have a notion of best, we need a way to
 search for that best lens.
-
 Our overall approach involves two communicating synthesizers: one (\RXSearch)
@@
 symmetric lenses, for which a single lens term is typed between many possible
-regular expression pairs, each DNF lens is only typed between a single regular
+regular expression pairs, each DNF lens is typed only between a single regular
 expression pair. The relationship between DNF lens syntax and DNF regular
-expression syntax is incredibly tight, permitting an incredibly efficient
+expression syntax is very tight, permitting an efficient
 synthesis algorithm for DNF lenses from DNF regular expressions.
@@
 enumerating classes and efficiently searching through them, but it poses a
-problem: unless we find a lens with no cost, other lens classes may contain
-better lenses. While we may have found a great lens between two regular
+problem: when do we stop? While we may have found a great lens between two regular
 expressions, a different regular expression pair may provide an even better
-lens. There may be a great lens in the class characterized by
-\lstinline{crontab} and \lstinline{scheduled_job}, but it is not the the
-intended lens shown in Figure~\ref{fig:example_lens}. However, there is a
-tension, we want to find a good lens, but we also want to give quick feedback to
+lens. Moreover, we may find a great lens characterized by
+\lstinline{crontab} and \lstinline{scheduled_job} that is not the
+desired lens shown in Figure~\ref{fig:example_lens}.  The
+tension is that we want to find a good lens, but we also want to give quick feedback to
 the user. After finding the lens in Figure~\ref{fig:example_lens}, we dont want
@@
 
-We address this by keeping running track of the best lens found thus far, and
+\saz{The following is hard to follow}
+We address this issue by keeping track of the best lens found thus far, and
 eventually cutting the proposal-synthesis loop short. Before proposing new
@@
 current best. For a given regular expression pair, \RXSearch looks at the number
-of rewrite rules performed on the pair, and identifies how many pairs with
+of rewrite rules performed on the pair, and identifies how many pairs
 involving the same number of rewrites may be proposed. If there are a large
 number of pairs that can be proposed with that many rewrites, or the current
-lens is very well synchronized, \RXSearch does not propose that pairs. In this
+lens is very well synchronized, \RXSearch stops. In this
 way, \RXSearch will continue searching for better lenses, but does not tackle

### 2018-06-07 papers/SymmetricLenses/main.tex
react to comments

@@
 
-Unfortunately, writing such synchronizers is tedious and error
-prone.\saz{``tedious and error prone'' is such a PL clich\'{e}, can we do better?}
-Bidirectional languages ease the difficulty of writing such synchronizers, because a
-single, well-typed program expresses a \emph{lens}---a group of conversion
-functions guaranteed to satisfy round-tripping laws. However, bidirectional
-languages are themselves difficult to program in, often requiring users to permute
-data in fiddly ways while satisfying the type system's complex unambiguity
-constraints.
+Unfortunately, writing such synchronizers is quite difficult; we must write
+parsers both formats, specify how the parsed data is converted into the other
+format, and specify how to propagate updates from one format to the other.
+Bidirectional languages ease the difficulty of writing such synchronizers,
+because a single, well-typed program expresses a \emph{lens}---a group of
+conversion functions guaranteed to satisfy round-tripping laws. However,
+bidirectional languages are themselves difficult to program in, often requiring
+users to permute data in fiddly ways while satisfying the type system's complex
+unambiguity constraints.
 
@@
 
-\emph{Symmetric lenses} \saz{citations?} formalize this richer class of
+\emph{Symmetric lenses}~\cite{symmetric-lenses} formalize this richer class of
 non-bijective conversion functions by using additional state, called the
@@
 expressiveness, the existence of the complement makes symmetric lenses difficult
-to use in real-world synchronization tasks, because users\saz{lens developers?
-  lens programs?}  must maintain the synchronized files as well as the
-complements. The complement further complicates synthesis, because input-output
-examples no longer suffice; instead edit sequences must be provided.
-
-In this paper, we reformulate \saz{this sounds like they may be of the same
-  expressiveness, but perhaps we mean to ``constrain'' symmetric lenses?}
-symmetric lenses as \emph{simple symmetric lenses}. This reformulation does not
-use a complement, and only uses the files that we aim to synchronize as
-state. We identify a property, \emph{forgetfulness}, that identifies the subset
-of symmetric lenses that can be written as simple symmetric lenses. We provide
-combinators for building simple symmetric lenses on string data, and integrate
-these combinators with Boomerang, a language for bidirectional programming on
-strings. In Boomerang, and our extension to it, lens expressions are typed
-between regular expressions, where the regular expressions describe the
-synchronized data formats.
+to use in real-world synchronization tasks, because lens programs must maintain
+the synchronized files as well as the complements. The complement further
+complicates synthesis, because input-output examples no longer suffice; instead
+edit sequences must be provided.
+
+In this paper, we contrain symmetric lenses to \emph{simple symmetric lenses}.
+These constrained lenses do not use a complement, and only use the files that we
+aim to synchronize as state. We identify a property, \emph{forgetfulness}, that
+identifies the subset of symmetric lenses that can be written as simple
+symmetric lenses. We provide combinators for building simple symmetric lenses on
+string data, and integrate these combinators with Boomerang, a language for
+bidirectional programming on strings. Both Boomerang and our extension type lens
+terms between pairs of regular expressions. The typing judgment $\Lens : \Regex
+\Leftrightarrow \RegexAlt$ means the lens term $\Lens$ encodes synchronizing
+functions between the format described by $\Regex$ and the format described by
+$\RegexAlt$.
 
@@
 input-output examples and a pair of regular expressions describing the data
-formats. The synthesis algorithm uses type-directed synthesis on the two regular
-expressions to synthesize a lens that is well-typed between the provided regular
-expressions and satisfies the provided input-output examples.
+formats. The synthesis algorithm uses type-directed synthesis
+on the two regular expressions to synthesize a lens that is well-typed between
+the provided regular expressions and satisfies the provided input-output
+examples.
 
 \saz{This paragraph is hard to follow --- maybe introduce $\ell : S
-  \Leftrightarrow T$ and use the names $S$ and $T$?}
-A primary difficulty is that there are a relatively many well-typed simple
-symmetric lenses. As symmetric lenses permit data to be present in only one
-format, for any data component there is now a choice: is this data only relevant
-to this format and as such should be left unaligned (changes to this data are
-not propagated to the other format), or is this data present in both formats and
-as such should be aligned (changes to this data change are propagated to the
-other format).
+  \Leftrightarrow T$ and use the names $S$ and $T$?} A primary difficulty is
+that there are a relatively many well-typed simple symmetric lenses. If we are
+trying to find a lens between $\Regex$ and $\RegexAlt$, there is now a choice:
+do $\Regex$ and $\RegexAlt$ represent the same data, and should be aligned
+(changes to data in the language of $\Regex$ are propagated to synchronized
+strings in the language of $\RegexAlt$), or should it be left unaligned (changes
+to data in $\Regex$ update synchronized strings in $\RegexAlt$). Or instead
+should certain components of $\Regex$ be aligned with certain components of
+$\RegexAlt$, with the other components left unaligned.
 
@@
 formats from configuration files, application-specific data storage files, and
-data cleaning tasks. We find that our algorithm is able to synthesize all
-\saz{or each?}  of these lenses in under 20 seconds.
+data cleaning tasks. We find that our algorithm is able to synthesize each of
+these lenses in under 20 seconds.
 
@@
 
-\saz{Perhaps we should explicitly say here that we are trying to define a lens,
-  or describing the functions that a lens provides.}
-To synchronize these files, a few functions are needed. Two \emph{creates}
-\saz{we haven't introduced ``create'' as a noun yet.} are
-needed; if a crontab is initialized, a synchronized Scheduled Tasks file should
-also be created, and vice versa. Furthermore, \emph{puts} are needed; if a
-crontab is updated, we should be able to combine it with a non-updated Scheduled
-Tasks file, and output a new Scheduled Tasks file, synchronized to the updated
+Lenses synchronize files by providing 4 functions, two \emph{create} functions,
+and two \emph{put} functions. Create functions are used to initialize data in
+one format when provided with data in the other. If a crontab is initialized, a
+synchronized Scheduled Tasks file should also be created, and vice versa. Put
+functions are used to propagate edits from one format to the other. If a crontab
+is updated, we should be able to combine it with a non-updated Scheduled Tasks
+file, and output a new Scheduled Tasks file, synchronized to the updated
 crontab. To have a simple symmetric lens is to have these required
@@
     Note that we can annotate our lenses with types, to confirm the lens
-    operates on the languages of the regular expressions.\saz{Say something
+    operates on the languages of the regular expressions. In this work, we
+    synthesize the lens itself from such regular expressions. \saz{Say something
       about the possibility of synthesizing this lens?}}
@@
 can generate new regular expression pairs, syntactically distinct but
-semantically equivalent to the originals.
+semantically equivalent to the originals. \RXSearch proposes the regular
+expression pairs which are most similar to the originals first (in that they
+require fewer rewrite rules to generate), and iteratively propose pairs that are
+increasingly dissimilar to the original pair (in that they require more rewrite
+rules to generate).
 
@@
 
-\saz{The following is hard to follow}
-We address this issue by keeping track of the best lens found thus far, and
-eventually cutting the proposal-synthesis loop short. Before proposing new
-types, the regular expression generator, \RXSearch, evaluates whether it should
-continue proposing regular expression pairs, or whether it is content with the
-current best. For a given regular expression pair, \RXSearch looks at the number
-of rewrite rules performed on the pair, and identifies how many pairs
-involving the same number of rewrites may be proposed. If there are a large
-number of pairs that can be proposed with that many rewrites, or the current
-lens is very well synchronized, \RXSearch stops. In this
-way, \RXSearch will continue searching for better lenses, but does not tackle
-huge search spaces unless it needs to. In this way, the algorithm can provide
-lenses quickly to the users, which users can refine with additional examples.
+\saz{The following is hard to follow} Eventually, we cut the proposal-synthesis
+loop short. Before proposing new types \RXSearch evaluates whether it should
+continue proposing regular expression pairs, or whether it is content with lens
+of lowest cost seen thus far. If there are a large number of regular expression
+pairs that are equally dissimilar to the originals or if the current lens is
+very well synchronized, \RXSearch stops. By doing this \RXSearch continues
+searching for better lenses, but does not tackle huge search spaces unless it
+needs to. In this way, the algorithm can provide lenses quickly to the users,
+which users can refine with additional examples.
 % end overview

### 2018-06-07 papers/SymmetricLenses/main.tex
Comments on the intro

@@
 \saz{This abstract too long and doesn't get to the point quickly enough. We
-  should tighten it up considerably.}
+  should tighten it up considerably.}\bcp{+1.  But not just tighten ---
+  the whole story doesn't feel very compelling.}
   Users commonly need to synchronize the data shared between formats, so edits
@@
   symmetric lenses. However, symmetric lenses in their current formulation
-  require external state in the form of a \emph{complement}. This statefulness
-  makes symmetric lenses both difficult to use and difficult to synthesize.
-
-  In this work, we reformulate symmetric lenses as \emph{simple symmetric
+  require external\bcp{why not ``internal''?} state in the form of a \emph{complement}. This statefulness
+  makes symmetric lenses both difficult to use and difficult to
+  synthesize. \bcp{This paragraph seems too dense.}
+
+  In this work, we reformulate\bcp{More a restriction than a reformulation.}
+  symmetric lenses as \emph{simple symmetric 
     lenses}. While this formulation is less expressive than symmetric lenses, it
-  expresses many symmetric lenses while requiring \emph{no external state}. We
+  expresses many\bcp{weak} symmetric lenses while requiring \emph{no external state}. We
   provide combinators for simple symmetric lenses on string data, and integrate
-  these combinators into Boomerang while providing full backwards compatibility.
-
+  these combinators into Boomerang\bcp{undefined} while providing full backwards compatibility.
+
+\bcp{This bit feels more detailed / technical than needed for an abstract.}
   We develop a type-directed synthesis algorithm for generating lenses in this
@@
 % state of the world
-Similar data can come in varieties of formats: scheduled jobs can be stored as
+\bcp{Not very compelling:}Similar data can come in varieties of formats: scheduled jobs can be stored as
 cron jobs for Linux or as Scheduled Tasks for Windows; web service resources can
@@
 \saz{This paragraph is hard to follow --- maybe introduce $\ell : S
-  \Leftrightarrow T$ and use the names $S$ and $T$?} A primary difficulty is
+  \Leftrightarrow T$ and use the names $S$ and $T$?}\bcp{The names make it a
+bit better, but it is all still rather abstract.  Indeed, the whole intro is
+quite abstract---could it not be driven by an example?} A primary difficulty is
 that there are a relatively many well-typed simple symmetric lenses. If we are
@@
 would it be better for the field on the right to be aligneded to a mish-mash
-of subcomponents from each field? To answer these questions, we develop a cost
+of subcomponents from each field?\bcp{Wouldn't these questions be answered
+  by examples?} To answer these questions, we develop a cost
 metric for reasoning about which lens aligns the most information.
@@
 expression. This entropy is used to calculate the expected information content
-to recover a string, when given a string it is synchronized with.
+to recover a string, when given a string it is synchronized with.\bcp{I
+know we're going to get to more technical detail later, but at this point I
+am confused: a string matching a RE can in general be arbitrarily long, so
+what kind of sense does it make to talk about its ``expected'' information content?}
 
@@
 
+\bcp{It gets rather technical and incomprehensible for the next three
+  paragraphs...}
+
 We develop an algorithm that synthesizes a low-cost lens, shown in
@@
 for a lens between those two regular expressions that minimizes cost and does
-not traverse star-related equivalences. When the first algorithm is content with
+not traverse star-related equivalences\bcp{???}. When the first algorithm is content with
 the lens it has found, it returns that lens.
 
-Our first algorithm, \RXSearch, proposes candidate stochastic regular
+\bcp{Make clear that these algorithms are based on earlier / published ones.}Our first algorithm, \RXSearch, proposes candidate stochastic regular
 expressions. These candidate stochastic regular expressions are generated by
@@
   and describe a language that expresses such lenses on string
-  data~(\S\ref{sec:ssl}).
+  data~(\S\ref{sec:ssl}).  \bcp{A new language, or a variant of
+    Hofmann/Pierce/Wagner?} 
 \item We develop a means to traverse stochastic regular expression equivalences,
@@
   file formats from configuration files, application-specific data storage
-  files, and data cleaning tasks (\S\ref{sec:evaluation}).
+  files, and data cleaning tasks (\S\ref{sec:evaluation}).  \bcp{... and
+    conclude what?}
 \end{enumerate}

### 2018-06-11 papers/SymmetricLenses/main.tex
remove rights stuff
Added comments to Sections 3 and 4

@@
 
-\setcopyright{rightsretained}
-\acmPrice{}
-\acmDOI{10.1145/3158089}
-\acmYear{2019}
-\copyrightyear{2019}
-\acmJournal{PACMPL}
-\acmVolume{3}
-\acmNumber{POPL}
-\acmArticle{1}
-\acmMonth{1}
-\startPage{1}
+  \renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
+
+%\setcopyright{rightsretained}
+%\acmPrice{}
+%\acmDOI{10.1145/3158089}
+%\acmYear{2019}
+%\copyrightyear{2019}
+%\acmJournal{PACMPL}
+%\acmVolume{3}
+%\acmNumber{POPL}
+%\acmArticle{1}
+%\acmMonth{1}
+%\startPage{1}
 
@@
 \label{sec:ssl}
+\sam{Perhaps we shoud introduce stochastic regular expressions first before introducing the lens combinators?
+This is because the type equivalence combinator mentions regexp equivalence, but this has not been introduced by this point in the paper.}
 We now give a formal description of all the combinators in our symmetric lens
@@
 %\end{center}
-Inverting is particularly useful when chaining many compositions together, as it
-can be used to align the central types.
+Inverting reverses the direction of a lens: creating on the right becomes creating on the left and vice versa, and putting on the right becomes putting on the left and vice versa.
+The invert combonator is particularly useful when chaining many compositions together, as it can be used to align the central types.
 
@@
 \]
+Type equivalence enables a lens of type $S \Leftrightarrow T$ to be used as a lens of type $S' \Leftrightarrow T'$ if $S$ equivalent to $S'$ and $T$ is equivalent to $T'$.
 Type equivalence is useful both for addressing type annotations and for
-letting compositions go through.
+letting compositions go through.\sam{I do not understand this statement.}
 %\begin{center}
@@
 %\end{center}
-Inverting is particularly useful when chaining many compositions together, as it
-can be used to align the central types.
 % end ssl
@@
 annotations would be.  The probabilities are defined as follows:
-
+\sam{Why in the definition of star do we multiply by $p^n * (1-p)$? I think this 
+should be explained}
 \begin{center}
@@
 
-    $\ProbabilityOf{\RegexConcat{\Regex_1}{\Regex_2}}{\Character_1 \ldots \Character_n}$
+    $\ProbabilityOf{\RegexConcat{\Regex_1}{\Regex_2}}{s}$
 
@@
 atoms are iterated stochastic DNF regular expressions.
+\sam{I think we should explicitly say what sets $A, B$, $SQ, TQ$ and $DS, DT$ range over.}
 \begin{center}
@@
 union of its subcomponents, the language of a stochastic sequence is the
-concatenation of its subcomponents, and the language of a stochastoc atom is the
+concatenation of its subcomponents, and the language of a stochastic atom is the
 iteration of its subcomponent.
@@
 expressions whose DNF form gives equal probability to all sequences, and gives
-iterations a .8 chance to continue iterating.
-
-\paragraph*{Entropy} In the situation where a DNF regular expression is
+iterations a .8 chance to continue iterating.\sam{I think we should mention that
+.8 is chosen because it works well in practice.}
+
+\paragraph*{Entropy}
+\sam{We should define what entropy means and why it is significant}
+In the situation where a DNF regular expression is
 unambiguous, we have defined an efficient, syntactic means to find the entropy
 of the DNF regular expression.
-
 \begin{center}
@@
 sequence $j$ in the left and right DNF regular expressions, respectively. The
-third component, \SSQLens{}, is a sequence lens that maps between the two
+third component of the triple, \SSQLens{}, is a sequence lens that maps between the two
 sequences specified by $i$ and $j$. The second component is a list of integers
@@
 this, $i$ and $j$ refer to atom $i$ and atom $j$ on the left and right
-sequences, respectively. The third component, \SAtomLens{}, is an atom lens that
+sequences, respectively. The third component of the triple, \SAtomLens{}, is an atom lens that
 maps between the two atoms specified by $i$ and $j$. The second component is a
@@
 \paragraph*{Symmetric DNF Lenses: Typing and Semantics}
+\sam{I found that I understood the semantics easier than the typing
+rules. Perhaps we should switch the order that we do them in?}
 The typing judgment is a 3-ary relation over a single DNF lens, and two DNF
@@
     \end{cases*}$\\
-  $\PutR{} \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q' \App \StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q'' \Concat \StringAlt_q'$
+  $\PutR{} \App (\String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q') \App (\StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q'' \Concat \StringAlt_q')$
   & =
@@
       \end{cases*}$\\
-  $\PutL{} \App \StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q'' \Concat \StringAlt_q' \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q'$
+  $\PutL{} \App (\StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q'') \Concat (\StringAlt_q' \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q')$
   & =
@@
   where $\String_i = \SDNFLens.\CreateL \App \StringAlt_i$\\
-  $\PutR{} \App \String_0 \Concat \ldots \Concat \String_n \App
-  \StringAlt_1 \Concat \ldots \Concat \StringAlt_m$
+  $\PutR{} \App (\String_0 \Concat \ldots \Concat \String_n) \App
+  (\StringAlt_1 \Concat \ldots \Concat \StringAlt_m)$
   & = 
@@
     \end{cases*}$\\
-  $\PutL{} \App \StringAlt_1 \Concat \ldots \Concat \StringAlt_m \App
-  \String_0 \Concat \ldots \Concat \String_n$
+  $\PutL{} \App (\StringAlt_1 \Concat \ldots \Concat \StringAlt_m) \App
+  (\String_0 \Concat \ldots \Concat \String_n)$
   & = 
@@
 search for the \emph{best} term that satisfies the specification. This begs the
-question, by what metric do we use to determine which term is best. Intuitively,
+question: what metric should we use to determine which term is best? Intuitively,
 we aim to synthesize the lens which is \emph{closest} to being a bijection.
@@
   \Given \DNFRegex,\SDNFLens}$.
-
+\sam{We should definitely explain the formulae for entropy computing entropy. Also I did not understand $argmin_{\theta}$ terms were being talked about in the 
+previous sentence.}
 \begin{centering}
@@
 sequence lenses (\GreedySeqSynth).
-
+\sam{Could it be better to describe the algorithm in English first before giving
+pseudo code? I think that putting the algorithm first might cause readers to read the 
+pseudo code and get confused whereas if the algorithm is first described in English then they have a better chance of understanding it.}
 \begin{algorithm}
@@
 between the given DNF regular expressions that is consistent with the examples.
-If there is not, \GreedySynth returns \None early. After this, \GreedySynth
+\sam{I think we should reword this to say something like CANNOTMAP checks to see if the specification cannot be satisfied.
+To me it sounds like checking if a lens possibly exists is already going into synthesis.}If there is not, \GreedySynth returns \None early.
+After this, \GreedySynth
 finds the best lenses between each sequence pair of the left and right DNF
@@
 thrown into the priority queue.
-
+\sam{Perhaps a small example on initialization would help? I was wondering what
+the partial DNF lens and potential sequence lenses could be initialized as}
 \GreedySynth then loops until there are no more sequence lenses in the priority
@@
 relatively efficiently, it suffers from an exponential blowup when converting
-regular expressions to DNF form. This can be minimized by not converting
+regular expressions to DNF form.\sam{Perhaps explain why this blowup arises?} This can be minimized by not converting
 \emph{all} of the regular expression into DNF form, but only converting
@@
 At the start of synthesis, we annotate our regular expressions as closed. Now,
-we enhance our rewrites by changing an annotation to from closed to open. This
+we enhance our rewrites by changing an annotation from closed to open. This
 forces changes to \RXSearch and \GreedySynth.  \RXSearch is changed in that now
@@
 disconnected.
-
+\sam{I did not understand this section. In particluar, I'm still not sure I
+understand a closed regexp vs an open regexp. Perhaps an example would help?}
 \paragraph*{Expansion Inference} The prior optimization helped \GreedySynth, but

### 2018-06-14 papers/DifferenceLenses/main.tex
Differernce lenses draft

\newif\ifdraft\drafttrue  % set true to show comments
\newif\ifplentyoftime\plentyoftimefalse  % :-)

% % For double-blind review submission, w/o CCS and ACM Reference (max
% submission space)
\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
% % For double-blind review submission, w/ CCS and ACM Reference
% \documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
% % For single-blind review submission, w/o CCS and ACM Reference (max
% submission space)
% \documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
% % For single-blind review submission, w/ CCS and ACM Reference
% \documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true} % For
% final camera-ready submission, w/ required CCS and ACM Reference
% \documentclass[acmsmall]{acmart}\settopmatter{}

% % Journal information % Supplied to authors by publisher for camera-ready
% submission; % use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

% % Copyright information % Supplied to authors (based on authors' rights
% management selection; % see authors.acm.org) by publisher for camera-ready
% submission; % use 'none' for review submission.
\setcopyright{none}
% \setcopyright{acmcopyright} \setcopyright{acmlicensed}
% \setcopyright{rightsretained} \ccopyrightyear{2018}           %% If different
% from \acmYear

% % Bibliography style %\bibliographystyle{ACM-Reference-Format} % Citation
% style % Note: author/year citations are required for papers published as an %
% issue of PACMPL.
% %\citestyle{acmauthoryear}   %% For author/year citations

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% % Note:
% Authors migrating a paper from PACMPL format to traditional % SIGPLAN
% proceedings format must update the '\documentclass' and % topmatter commands
% above; see 'acmart-sigplanproc-template.tex'.
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{amsmath, amssymb, amsthm, enumerate, array, extarrows, mathrsfs,
mathtools, stmaryrd, listings, xspace, bussproofs}
\theoremstyle{definition}
\usepackage[capitalize]{cleveref}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{conjecture}{Conjecture}
\newtheorem{question}{Question}
\newtheorem{fact}{Fact}
\newtheorem{claim}{Claim}
\newtheorem{remark}{Remark}

% Macros Colors
\definecolor{dkblue}{rgb}{0,0.1,0.5}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{dkred}{rgb}{0.6,0,0}
\definecolor{dkpurple}{rgb}{0.7,0,0.4}
\definecolor{olive}{rgb}{0.4, 0.4, 0.0}
\definecolor{teal}{rgb}{0.0,0.5,0.5}
\definecolor{orange}{rgb}{0.9,0.6,0.2}
\definecolor{lightyellow}{RGB}{255, 255, 179}
\definecolor{lightgreen}{RGB}{170, 255, 220}
\definecolor{teal}{RGB}{141,211,199}
\definecolor{darkbrown}{RGB}{121,37,0}

\newcommand{\FINISH}[3]{\ifdraft\textcolor{#1}{[#2: #3]}\fi}
\newcommand{\bcp}[1]{\FINISH{dkred}{B}{#1}}
\newcommand{\BCP}[1]{\FINISH{dkred}{B}{\bf #1}}
\newcommand{\afm}[1]{\FINISH{dkgreen}{A}{#1}}
\newcommand{\dpw}[1]{\FINISH{dkblue}{D}{#1}} % Toronto Maple Leafs Blue :-)
\newcommand{\saz}[1]{\FINISH{orange}{SZ}{#1}}
\newcommand{\ksf}[1]{\FINISH{teal}{K}{#1}}
\newcommand{\sam}[1]{\FINISH{dkpurple}{SM}{#1}}

\newcommand{\kw}[1]{\ensuremath{\mathsf{#1}}\xspace}
\newcommand{\get}{\ensuremath{\kw{get}}\xspace}
\newcommand{\pput}{\ensuremath{\kw{put}}\xspace}
\newcommand{\create}{\ensuremath{\kw{create}}\xspace}
\newcommand{\res}{\ensuremath{\kw{res}}\xspace}
\newcommand{\ccopy}{\ensuremath{\kw{copy}}\xspace}
\newcommand{\const}{\ensuremath{\kw{const}}\xspace}
\newcommand{\default}{\ensuremath{\kw{default}}\xspace}
\newcommand{\length}{\ensuremath{\kw{length}}\xspace}
\newcommand{\match}{\ensuremath{\kw{match}}\xspace}

\DeclareMathOperator*{\argmin}{arg\,min} % thin space, limits underneath in
\DeclareMathOperator*{\argmax}{argmax} % no space, limits underneath in

\lstset{ basicstyle=\ttfamily, escapeinside=|| }
\begin{document}

% % Title information
\title{Lenses with a Difference}         %% [Short Title] is optional;
% % when present, will be used in % header instead of Full Title.
% %\titlenote{with title note}             %% \titlenote is optional; % can be
% repeated if necessary; % contents suppressed with 'anonymous'
% %\subtitle{Subtitle}                     %% \subtitle is optional
% %\subtitlenote{with subtitle note}       %% \subtitlenote is optional; % can
% be repeated if necessary; % contents suppressed with 'anonymous'

% % Author information % Contents and number of authors suppressed with
% 'anonymous'.
% % Each author should be introduced by \author, followed by % \authornote
% (optional), \orcid (optional), \affiliation, and % \email.
% % An author may have multiple affiliations and/or emails; repeat the %
% appropriate command.
% % Many elements are not rendered, but should be provided for metadata %
% extraction tools.

% % Author with single affiliation.
% \author{First1 Last1} \authornote{with author1 note}          %% \authornote
% is optional; % can be repeated if necessary \orcid{nnnn-nnnn-nnnn-nnnn} %%
% \orcid is optional \affiliation{ \positition{Position1}
% \department{Department1} %% \department is recommended
% \institution{Institution1} %% \institution is required \streetaddress{Street1
% Address1} \city{City1} \state{State1} \posittcode{Post-Code1}
% \country{Country1}                    %% \country is recommended }
% \email{first1.last1@inst1.edu}          %% \email is recommended

\author{Solomon Maina}
\position{PhD Student}
\department{Computer Science}              %% \department is recommended
\institution{University of Pennsylvania}            %% \institution is required
\country{USA}                    %% \country is recommended }
\email{smaina@seas.upenn.edu}          %% \email is recommended

% % Abstract % Note: \begin{abstract}...\end{abstract} environment must come %
% before \maketitle command
\begin{abstract}
Many of the bidirectional (bx) programming languages that exist today focus on
restoring consistency between source and target data. Consequently, these languages often provide two guarantees: (1) that a transformation restores consistency when an
update is made to source or target data, and (2) that trivial updates to a the source result to trivial updates to the target and vice versa. Unfortunately, in situations
where there are many different ways of restoring consistency, these languages
do not constrain how a transformation chooses to do so. As a result, the transformations derived using these languages can lead to unexpected results. In this paper we present a formalization of bx transformations that enables a programmer to express non-trivial guarantees on update propagation. We will then use this formalization of bx transformations to describe a compositional language of {\em difference lenses}: each difference lens gives non-trivial guarantees on how it propagates non-trivial updates. We will then embed various lens programming languages into the language of difference lenses, and describe non-trivial guarantees that these languages provide beyond consistency restoration.
\end{abstract}

%% Keywords
%% comma separated list
\keywords{bidirectional programming, lenses}  %% \keywords
% are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}
Many of the bidirectional (bx) programming languages that exist today focus on
restoring consistency between source and target data. Consequently, these languages often provide two guarantees: (1) a {\em correctness} guarantee that requires that a transformation restores consistency when an update is made to source or target data, and (2) a {\em hippocraticness} guarantee that requires that trivial updates to source data result is trivial updates to target data and vice versa. Unfortunately, in situations
where there are many different ways of restoring consistency, these languages
do not constrain how a transformation chooses to do so. As a result, the transformations derived using these languages can lead to unexpected results.

For example, assume that a source consists of a list of pairs of names, and that the target consists of a list of single names. Assume that $\ell$ is a bx which in the forward direction (i.e. the \get direction) simply discards the first of the two names:
$$\ell.\get \; [``Ada \; Lovelace", \; ``Alan \; Turing"] = [``Lovelace", \; ``Turing"]$$
If the order of the names in the target is reversed then a user might expect that in the backward direction (i.e. the \pput direction), the lens would simply restore the original names:
$$\ell.\pput \; [``Ada \; Lovelace", \; ``Alan \; Turing"] \; [``Turing", \; ``Lovelace"] = [``Ada \; Lovelace", \; ``Alan \; Turing"]$$
However, $\ell$ may not perform this alignment so that instead, the following result occurs:
$$\ell.\pput \; [``Ada \; Lovelace", \; ``Alan \; Turing"] \; [``Turing", \; ``Lovelace"] = [``Ada \; Turing", \; ``Alan \; Lovelace"]$$
Consider another example where a source consists of a list of names and a number while the target consists of a list of names. Assume that $\ell$ is a bx which in the \get direction simply discards the number:
$$\ell.\get \; ([``Ada \; Lovelace", \; ``Alan \; Turing"], \; 2)  = [``Lovelace", \; ``Turing"]$$
A user might expect that $\ell$ should always restore the number in the \pput direction so that the following result occurs:
$$\ell.\pput \; ([``Ada \; Lovelace", \; ``Alan \; Turing"], \; 2) \; [``Lovelace"] = ([``Ada \; Lovelace"], \; 2)$$
However, perhaps the intended semantics of the source is that the number which is the second component of a source denotes the length of the list in the first component of the source so that an application of \pput leads to the following result:
$$\ell.\pput \; ([``Ada \; Lovelace", \; ``Alan \; Turing"], \; 2) \; [``Lovelace"] = ([``Ada \; Lovelace"], \; 1)$$
In both of these examples, there are multiple ways to restore consistency; that is, there are multiple ways to define $\ell.\pput \; s \; v$ in such a way that $\ell.\get \; (\ell.\pput \; s \; v) = v$. Moreover, the source that results from a \pput application may be vastly different from what a user expects, or may break an invariant that is not explicitly maintained by the source.

Various solutions have been proposed to address the issue of propagating updates in a ``minimal'' way, for example {\em matching lenses} that deal with the issue of alignment \cite{barbosa2010matching}, {\em edit lenses} \cite{hofmann2012edit} and {\em delta lenses} \cite{diskin2011asymmetric,diskin2011state,pacheco2012delta}
that explicitly propagate edits through the lens, and {\em least change
lenses} \cite{macedo2013composing} that require that the source value resulting from a \pput application is not only acceptable, but also one of closest to the original $s$ among the sources that share the same view $v$.

We take the approach that a bx should not have to explicitly propagate updates when going from the source to the target and vice versa. Instead, the programmer should be able provide guarantees on how a bx propagates updates. For instance, in the second example, let $\delta_S$ be a function which takes two source elements $(\ell_1, n_1)$ and $(\ell_2, n_2)$ and returns the number $n_1 - n_2$, and let $\delta_T$ be a function which takes two target lists $\ell_1$ and $\ell_2$ and returns the number $length(\ell_1) - \length(\ell_2)$. Then the programmer should be able to express the fact that the bx restores the length of the list in the backward direction, which is expressed using the following equation:
$$\delta_S((\ell_1, n_1), \ell.\pput \; (\ell_1, n_1) \; \ell_2) = \delta_V(\ell.\get \; (\ell_1, n_1), \ell_2) $$
In this paper, we introduce a formalization of bx transformations that enables a programmer to express guarantees such as the one above and show how our formalism can be used to characterize update propagation behaviour of bidirectional transformations.

The major technical contributions that we will make in this paper are the following:
\begin{enumerate}
  \item
  We present a formalization of bx transformations that enables a programmer to express non-trivial update propagation behaviour.
  \item
  We use our formalization to develop a compositional language of {\em difference lenses} based on the {\em basic lenses} implemented in the lens programming language Boomerang \cite{foster2009bidirectional}, and describe non-trivial update propagation guarantees that these lenses provide.
  \item
  We show how to represent matching lenses and  relational lenses \cite{bohannon2006relational} as difference lenses, and hence describe non-trivial guarantees that these lens languages provide as well.
\end{enumerate}
\iffalse
 we describe a compositional language of lenses called {\em difference lenses}, where each difference $\ell$ satisfies the following equation,
\begin{equation}\label{difflenslaw}
\delta_S(s, \ell.\pput \; s \; v) = \delta_V(\ell.\get \; s, v)
\end{equation}
for appropriate real-valued $\delta_S : S \times S \longrightarrow \mathbb{R}^{\geq 0}$ and $\delta_S : S \times S \longrightarrow \mathbb{R}^{\geq 0}$.

\cref{difflenslaw}, which already appears in Gibbons et. al. \cite{gibbons2017principles} as the principle of {\em Weak Least Surprise}, is simple and intuitive but presents a challenge when trying to develop a compositional lens langauge that satisfies it. For example, assume that a source consists of a name-and-date pairs while a view consists of a name. Assume that a lens $\ell$ simply discards dates in the \get direction:
\begin{lstlisting}[mathescape=true]
$\ell.\get$ "Ada Lovelace, 12-10-1815" = "Ada Lovelace"
\end{lstlisting}
Consider the iteration lens $\ell^*$, where the source is now a list of name-and-date pairs while a view consists of a list of names. Whenever the length of a view list is larger than the length of a source list, then a \pput application of $\ell^*$ must initialize a source from a view from scratch:
\begin{lstlisting}[mathescape=true]
$\ell.\pput$ [] ["Ada Lovelace"] = ["Ada Lovelace, 00-00-0000"]
\end{lstlisting}
Now if we are to define $\delta_S^*$ and $\delta_V^*$ inductively on source and view lists respectively in such a way that \cref{difflenslaw} holds of $\ell^*$, then the following equation would have to hold:
\begin{lstlisting}[mathescape=true]
$\delta_S^*$(["Ada Lovelace, 00-00-0000"], []) = $\delta_V^*$(["Ada Lovelace"], [])
\end{lstlisting}
In a case such as this when a source is {\em created} from a view as opposed to arising from a view being \pput into an existing source, the source delta will be defined so as to defer to the view delta by mapping the created source back into the view; in particular, for the lens $\ell^*$ in the example, then we have
\begin{lstlisting}[mathescape=true]
$\delta_S^*$(["Ada Lovelace, 00-00-0000"], [])
= $\delta_V^*$([$\ell.\get$ "Ada Lovelace, 00-00-0000"], [])
= $\delta_V^*$(["Ada Lovelace"], [])
\end{lstlisting}

By now there exist many programming languages that enable one to define {\em
bidirectional transformations}, programs that can be run in two
directions. Each of these languages provides guarantees on how the two
transformations interact e.g. that they are bijective, that a round-trip on
one or both directions does not cause non-trivial updates etc. Many of these
guarantees have to do with restoring or preserving consistency. To formalize
this statement, we describe a widely used formulation of bidirectional
transformations due to Stevens \cite{stevens2010bidirectional} but originally
articulated in a more general form by Meertens \cite{meertens1998designing}:
\begin{definition}
A bidirectional transformation (bx) $R: M \Leftrightarrow N$ comprises of a
consistency relation $R \subseteq M \times N$ and restorers $\overrightarrow{R}
: M \times N \rightarrow N$ and $\overleftarrow{R} : M \times N \longrightarrow
N$.
\end{definition}
\begin{definition}
A bx $R : M \Leftrightarrow N$ is {\em correct} if the consistency restorers do
restore consistency: that is, for all $m \in M, n \in N$, we have $R(m,
\overrightarrow{R}(m, n))$, and dually. $R$ is {\em hippocratic} if the
consistency restorers make no change to an already-consistent pair of models:
that is, for all $m \in M$, $n \in N$, if $R(m, n)$ then $\overrightarrow{R}(m,
n) = n$ and dually. $R$ is {\em well behaved} if it is correct and hippocratic.
\end{definition}

Many bidirectional programming languages thus enable one to define
well behaved bx's: consistency restoration is guaranteed by correctness of the
bx while consistency preservation is guaranteed by the hippocraticness of the
bx. This paper is primarily concerned with well behaved bx's where
$\overrightarrow{R}$, the \get function of the bx, ignores it's $N$-argument and
where the consistency relation $R$ is defined by $R(m, \overrightarrow{R}(m))$
for all $m \in M$--- these transformations are the well known and widely
understood {\em lenses} of Foster et al \cite{foster2007combinators}--- though
the ideas we develop could be used in other bidirectional settings as well.

When the consistency relation of a bx $R$ is very fine, i.e. when each $m \in M$
is consistent with only a few $n \in N$ and vice versa, then a well behaved lens
gives fairly strong guarantees on $\overleftarrow{R}$ and $\overrightarrow{R}$.
However, there are many cases that arise in practice where there are many
different ways to restore consistency; when this happens, well-behavedness may
be too weak to rule out ``bad'' behaviour or to clarify the behaviour of a bx in
subtle cases. Various solutions have been proposed to address this issue, for
example {\em matching lenses} that deal with the issue of alignment
\cite{barbosa2010matching}, {\em edit lenses} \cite{hofmann2012edit} and {\em
delta lenses} \cite{diskin2011asymmetric,diskin2011state,pacheco2012delta}
that explicitly propagate edits through the lens, and {\em least change
lenses} that require that the source value resulting from a \pput application
is not only acceptable, but also one of closest to the original $s$ among the
sources that share the same view $v$ \cite{macedo2013composing}.

This paper takes a different approach to specifying update propagation by
identifying a condition called {\em $\partial D$ preservation} which we describe
below:
\begin{definition}
Let $R : M \Leftrightarrow N$ be a bidirectional transformation and $\partial D
\subseteq (M \times M) \times (N \times N)$. Then $R$ {\em preserves $\partial
D$} if for all $m \in M$, $n, n' \in N$ , if $R(m, n)$ then $\partial D((m,
\overleftarrow{R}(m, n')), (\overrightarrow{R}(m, n), n'))$. $R$ {\em
strongly preserves $\partial D$} if the dual condition also holds.
\end{definition}
\noindent The idea behind $\partial D$ preservation is that if $m$ and $n$ are
already consistent and $\overrightarrow{R}(m,n)$ is updated to produce $n'$,
then the difference between $\overrightarrow{R}(m,n)$ and $n'$ is related
to the difference between $m$ and $\overleftarrow{R}(m,n')$. $R$
strongly preserves $\partial D$ if it preserves $\partial D$ in both directions.

For lenses, $\partial D$ preservation says that difference between $\ell.\get \;
s$ and $v$ is related to the difference between $s$ and $\ell.\pput \; s \;
v$. One way to formalize this idea is to model the difference between two sources $s, s'
\in S$ and two views $v, v' \in V$ as functions $\delta_S : S \times S
\longrightarrow \partial S$ and $\delta_V : V \times V \longrightarrow \partial
V$ for some sets $\partial S$ and $\partial V$, and to capture the
relationship between $\delta_S$ and $\delta_V$ by a lens. In
particular, given a lens $\partial(S, V) : \partial S \Leftrightarrow \partial
V$ we will ask if the following equation holds of $\ell$:
\begin{equation}\label{difflenslaw}
\partial(S, V).\get \; (\delta_S(s, \ell.\pput \; s \; v)) = \delta_V(\ell.\get
\; s, v)
\end{equation}
If $\ell : S \Leftrightarrow V$ satisfies \cref{difflenslaw} we say that 
$\ell$ is a {\em difference  lens} with respect to $(\delta_S, 
\partial(S, V), \delta_V)$, which we denote with the typing relation 
$\ell : S \xLeftrightarrow{\delta_S, \partial(S, V), \delta_V} V$.

For example, consider two versions of an iteration lens $\ell_1 : S^*
\Leftrightarrow V^*$ and $\ell_2 : S^* \Leftrightarrow V^*$ defined inductively
from a lens $\ell : S \Leftrightarrow V$. Both of these lens have the same get
behaviour: given a list $s_1, \ldots, s_n$ of sources, they both produce
$\ell.\get \; s_1, \ldots, \ell.\get \; s_n$. The \pput behaviour of the two
lenses is different though: $\ell_1$ simply aligns views to sources
positionally, while $\ell_2$ computes the optimal alignment on the view side and
uses this alignment to put views back into sources.

Let \lstinline|name = [A-Z] . [a-z]*|, let the source $S$ equal 
\lstinline|name . " " . name| and the view $V$ equal \lstinline|name|.
Let $\ell : S \Leftrightarrow V$ be the lens
\lstinline|del name . del " " . copy name| which given a pair of names produces
the second name in the \get direction, and restores the first name in the \pput
direction:
$$\ell.\get \; ``Ada \; Turing" = ``Turing" \text{ and }\; \ell.\pput \; ``Ada
\; Turing" \; ``Lovelace" = ``Ada \; Lovelace"$$
Define the
function $\delta_V : V \times V \longrightarrow \mathbb{N}^{\geq 0}$ by 
$$\delta_V(n, n') = 
\begin{cases}
0 & \text{if }n=n'\\
1 & \text{otherwise}
\end{cases}$$
Define $\delta_S : S \times S \longrightarrow \mathbb{N}^{\geq 0}$ by
$$\delta_S(m \cdot `` \quad " \cdot n, m' \cdot `` \quad " \cdot n') =
\begin{cases}
0 & \text{if } m = m', n = n'\\
1 & \text{if } m = m', n \neq n'\\
1 & \text{if } m \neq m', n = n'\\
2 & \text{if } m \neq m', n \neq n'
\end{cases}$$
Let the lens $\partial(S, V) : \mathbb{N}^{\geq 0} \Leftrightarrow
\mathbb{N}^{\geq 0}$ be given by $\partial(S, V) = \ccopy(\mathbb{N}^{\geq
0})$. Then $\ell : S \xLeftrightarrow{\delta_S, \partial(S, V), \delta_V} V$.

Now let $xs, ys$ be two lists of single names, with $length(xs) \geq length(ys)$
Define $\delta_V^* : V^* \times V^* \longrightarrow \mathbb{Z}^* \times V^*$ on $xs$ and
$ys$ by aligning elements of $xs$ maximally with elements of $ys$ and then using $\delta_V$
on these pairs to produce a list $ds$ of numbers. In this process, some elements of
$ys$ will not be aligned if $length(xs) < length(ys)$, in which case we simply collect
these in another list of elements of $V$. For instance,
$$\delta_V^*([``Lovelace", ``Kleene", ``Turing"], [``Turing", ``Lovelace"]) = ([0,0], [``Kleene"])$$
since we align \lstinline|"Lovelace"| to \lstinline|"Lovelace"| and 
\lstinline|"Turing"| to \lstinline|"Turing"| for a difference of 0 in each case, which leaves
\lstinline|"Kleene"| unaligned.

On the source side, let $ss, ts$ be two lists of two names with $length(ss) \geq length(ts)$.
Define $\delta_{S, \ell}^* : S^* \times S^* \longrightarrow \mathbb{Z}^* \times S^*$ by taking the 
alignment used to compute $\delta_V^*(\ell^*.\get \; ss, \ell^*.\get \; ts)$ and then using that
alignment in the same manner that $\delta_V^*$ does. For instance
\begin{align*}
&\delta_{S, \ell}^*([``Ada \; Lovelace", ``Stephen \; Kleene", ``Alan \; Turing"], 
[``Alan \; Turing", ``Ada \; Lovelace"])\\
&= ([0,0], [``Stephen \; Kleene"])
\end{align*}
Recall that the \pput behaviour of $\ell_1$ simply aligns views to sources
positionally, while $\ell_2$ computes the optimal alignment on the view side and
uses this alignment to put views back into sources. From this we can prove that
$$\ell_2 : S^* \xLeftrightarrow{\delta_{S, \ell}^*, \partial(S, V)^* \times \ell^*, \delta_V^*} V^*$$
On the other hand, the judgement $\ell_1 : S^* \xLeftrightarrow{\delta_{S, \ell}^*, 
\partial(S, V)^* \times \ell^*, \delta_V^*} V^*$ does not holds since $\ell_1$ does not satisfy \cref{difflenslaw} :
$$(\partial(S, V)^* \times \ell^*).\get \; (\delta_{S, \ell}^*(s, \ell.\pput \; s \; v)) = \delta_V^*(\ell.\get
\; s, v)
$$
This is witnessed by the fact that if $s = [``Ada \; Lovelace", ``Stephen \; Kleene", ``Alan \; Turing"]$ and \\
$v = [``Kleene", ``Turing", ``Lovelace"]$, then 
\begin{align*}
&\ell_1.\pput \; [``Ada \; Lovelace", ``Stephen \; Kleene", ``Alan \; Turing"] \; [``Kleene", ``Turing", ``Lovelace"]\\
&= [``Ada \; Kleene", ``Stephen \; Turing", ``Alan \; Lovelace"]
\end{align*}
Also, $\delta_{S, \ell}^*(s, \ell.\pput \; s \; v) = ([1,1,1], [])$, $\delta_V^*(\ell.\get
\; s, v) = ([0,0,0], [])$, but
\begin{align*}
(\partial(S, V)^* \times \ell^*).\get \; (\delta_{S, \ell}^*(s, \ell.\pput \; s \; v)) &=
(\partial(S, V)^* \times \ell^*).\get \; ([1,1,1], [])\\
&= ([1,1,1], [])\\
&\neq ([0,0,0], [])\\
&= \delta_V^*(\ell.\get \; s, v)
\end{align*}
Here the positional lens $\ell_1$ is not a difference lens with respect to $(\delta_{S, \ell}^*, 
\partial(S, V)^* \times \ell^*, \delta_V^*)$ since it does not align sources to views in the same way that
$\delta_V^*$ does. On the other hand, $\ell_2$ does, hence is a difference lens with respect to $(\delta_{S, \ell}^*, 
\partial(S, V)^* \times \ell^*, \delta_V^*)$.

In the rest of this paper, we give concrete examples of how to use $\partial D$ preservation to characterize 
the guarantees on update propagation provided by various lens programming languages. In particular,
the major technical contributions that we will make in this paper are the following:
\begin{enumerate}
  \item
  We formalize a condition called $\partial D$ preservation which provides a way
  of constraining the updates that can be made by a bidirectional
  transformation. We give examples of relations $\partial D$ as well as
  examples and counterexamples of lenses that are $\partial D$ preserving.
  \item
  We describe an instance of $\partial D$ preservation that is satisfied by {\em difference lenses}. We give a set of combinators for {\em difference lenses} and describe the update propagation guarantees that the lenses derived using these combinators give.
  \item
  Finally, we show how to represent matching lenses and relational lenses
  \cite{bohannon2006relational} as difference lenses, and hence
  describe non-trivial guarantees beyond correctness and hippocraticness that
  these lens languages provide as well.
\end{enumerate}
\fi
\section{Laying Down BX Laws}
\begin{definition}
A bidirectional transformation (bx) $R : M \Leftrightarrow N$ comprises of 
\begin{enumerate}
\item
a consistency relation $R \subseteq M \times N$,
\item
a relation $S \subseteq (M \times M) \times (N \times N)$ such that for all $m, m' \in M$, $n, n' \in N$,
\begin{enumerate}
\item
if $(m, m') \stackrel{S}{\sim} (n, n)$ then $m = m'$,
\item
if $(m, m) \stackrel{S}{\sim} (n, n')$ then $n = n'$, and
\end{enumerate}
\item
restorers $\overrightarrow{R} : M \times N \rightarrow N$ and $\overleftarrow{R} : M \times N \longrightarrow
N$.
\end{enumerate}
\end{definition}
Intuitively, the relation $S$ relates differences on the $M$ side to differences on the $N$ side. For instance, $S$ could be the relation $(m, m') \stackrel{S}{\sim} (n,n')$ if and only if $\delta_M(m,m') = \delta_N(n,n')$ for some metrics $\delta_M$ and $\delta_N$ on $M$ and $N$ respectively. The relation is required to send only trivial deltas to trivial deltas.
\begin{definition}
A bx $R : M \Leftrightarrow N$ is {\em $S$-preserving} for all $m \in M$, $n, n' \in N$, then if $R(m, n)$ then $(m, \overleftarrow{R}(m,n')) \stackrel{S}{\sim} (n, n')$
\end{definition}
Intuitively $S$-preservation says that if we start in a situation where $m$ is consistent with $n$, update $n$ to get $n'$, then map $n'$ back into $m$ via $\overleftarrow{R}(m,n')$, then the difference between $n$ and $n'$ is related to the difference between $m$ and $\overleftarrow{R}(m,n')$.
\begin{definition}
A bx $R : M \Leftrightarrow N$ is hippocratic if for all $m \in M$, $n \in N$, if $R(m, n)$ then $\overrightarrow{R}(m, n) = n$ and $\overleftarrow{R}(m,n) = m$.
\end{definition}
\begin{fact}
For any $R : N \Leftrightarrow N$ and any $S \subseteq (M \times M) \times (N \times N)$, if $R$ is $S$-preserving, then $R$ is hippocratic.
\end{fact}
\begin{definition}
A bx $R : M \Leftrightarrow N$ is {\em correct} if the consistency restorers do
restore consistency: that is, for all $m \in M, n \in N$, we have $R(m,
\overrightarrow{R}(m, n))$, and dually.
\end{definition}
\iffalse
A {\em lens} is a program that can be run in two directions. In the
forward direction, the \get component of a lens produces a {\em view}
from a {\em source}, while in the backward direction, the \pput component folds
an updated view back into the original source. Formally, Foster
\cite{foster2007combinators} defined a lens $\ell : S
\Leftrightarrow V$ from a source $S$ to a view $V$ to be a pair of functions
$\ell.\get : S \longrightarrow V$ and $\ell.\pput : S \longrightarrow V
\longrightarrow S$ satisfying the
following properties:
\begin{align*}
\ell.\get \; (\ell.\pput \; s \; v) &= v \tag{PUTGET}\\
\ell.\pput \; s \; (\ell.\get \; s) &= s \tag{GETPUT}
\end{align*}
The PUTGET law ensures that updates to the view are translated
``exactly''---i.e. that, given a view, the \pput function produces a source
that \get maps back to the very same view, while the GETPUT law ensures a
``stability'' property for the source---i.e. it requires that the \pput
function returns the original source unmodified whenever the update to the view
is a no-op.

Unfortunately, while these laws give some guarantees on how the \get and \pput
components of a lens interact, they do not specify the behaviour of the \pput
component when the update to the view is non-trivial. Consequently, a
non-trivial update to a view could lead to an unexpected update to the source.
Various solutions have been proposed to address this issue, for example {\em
matching lenses} to deal with the issue of alignment \cite{barbosa2010matching},
{\em edit lenses} \cite{hofmann2012edit} and {\em delta lenses}
\cite{diskin2011asymmetric,diskin2011state,pacheco2012delta} which explicitly
propagate edits through the lens, and {\em least change lenses} which
add a new LEASTPUTGET law which requires that the source value resulting from a
\pput application is not only acceptable, but also one of closest to the
original $s$ among the sources that share the same view $v$
\cite{macedo2013composing}.

This paper describes a general method for specifying what updates are allowable
for a given lens without requiring that the lens explicitly propagate updates.
Our approach is guided by the simple idea the idea that if one starts with a
source $s$, calls the \get function on $s$ to acquire the view $\ell.\get \;
s$, updates $\ell.\get \; s$ to get a new view $v$, then the ``difference''
between $v$ and $\ell.\get \; s$ and $v$ should be related to the difference
between $\ell.\pput \; s \; v$ and $v$.

To formalize this idea, we first recall a widely used definition of
bidirectional transformations
\cite{meertens1998designing,stevens2010bidirectional} definition of a
bidirectional transformations:
\begin{definition}
A bidirectional transformation (bx) $R: M \Leftrightarrow N$ comprises of a
consistency relation $R \subseteq M \times N$ and restorers $\overrightarrow{R}
: M \times N \rightarrow N$ and $\overleftarrow{R} : M \times N \longrightarrow
N$.
\end{definition}
\begin{definition}
A bx $R : M \leftrightarrow N$ is {\em correct} if the consistency restorers do
restore consistency: that is, for all $m \in M, n \in N$, we have $R(m,
\overrightarrow{R}(m, n))$, and dually. $R$ is {\em hippocratic} if the
consistency restores make no change to and already-consistent pair of models:
that is, for all $m \in M$, $n \in N$, if $R(m, n)$ then $\overrightarrow{R}(m,
n) = n$ and dually. $R$ is {\em well behaved} if it is correct and hippocratic.
\end{definition}

Observe that $\ell : S \Leftrightarrow V$ is a lens, let $R$ be the
consistency relation given by $R(s, \ell.\get \; s)$ for all $s \in
S$, $\overrightarrow{\ell}(s, v) = \ell.\get \; s$, and $\overleftarrow{\ell} =
\ell.\pput$. By the PUTGET law, $\ell.\get \; (\ell.\pput \; s \; v) = v$,
hence $\ell$ is correct, and by the GETPUT laws, $\ell.\pput \; s \; (\ell.\get
\; s) = s$ and $\ell.\get \; (\ell.\create \; v) = v$, hence $\ell$ is
hippocractic. Therefore $\ell$ is a well-behaved bx.

If $R$ is a well-behaved bx then $R$ restores consistency when a source or view
is updated (correctness) and preserves consistency when the source and view are
already consistent (hippocraticness). In order to express

We will then apply this method to define a typing judgement for lenses which
constrains how a lens propagates updates. We will describe a compositional
language of {\em difference lenses} which satisfy this new typing judgement,
and give examples of ordinary lenses that are not difference lenses. Finally,
we will show how to represent matching lenses and relational lenses
\cite{bohannon2006relational} as difference lenses, and hence describe the
guarantees that these lens languages provide.

For example suppose that the source is an XML document representing the
names, dates and nationalities of a collection of classical music
composers such as,

\begin{lstlisting}
<composers>
<composer>
<name>Jean Sibelius</name>
<dates>1865-1956</dates>
<nationality>Finnish</nationality>
</composer>
<composer>
<name>Aaron Copland</name>
<dates>1910-1990</dates>
<nationality>American</nationality>
</composer>
<composer>
<name>Benjamin Briten</name>
<dates>1913-1976</dates>
<nationality>English</nationality>
</composer>
</composers>
\end{lstlisting}

while the view is a list if ASCII text representing the names and dates of each
composer:
\begin{lstlisting}
Jean Sibelius, 1865-1956
Aaron Copland, 1910-1990
Benjamin Briten, 1913-1976
\end{lstlisting}

After computing the initial view, we might want to edit it in some way---e.g.,
correcting the error in Sibelius's death date and the misspelling in Britten's
name
\begin{lstlisting}
Jean Sibelius, 1865-|\colorbox{orange}{1957}|
Aaron Copland, 1910-1990
Benjamin |\colorbox{orange}{Britten}|, 1913-1976
\end{lstlisting}
and push the changes back into the original XML format:
\begin{lstlisting}
<composers>
<composer>
<name>Jean Sibelius</name>
<dates>1865-|\colorbox{orange}{1957}|</dates>
<nationality>Finnish</nationality>
</composer>
<composer>
<name>Aaron Copland</name>
<dates>1910-1990</dates>
<nationality>American</nationality>
</composer>
<composer>
<name>Benjamin |\colorbox{orange}{Britten}|</name>
<dates>1913-1976</dates>
<nationality>English</nationality>
</composer>
</composers>
\end{lstlisting}

Lenses are designed to perform these kinds of bidirectional transformations. If
$\ell$ is a suitably defined lens for the transformation that we
have just described, then the \get component of the lens performs the
XML to ASCII transformation that produces an ASCII view from an XML source,
while the \pput component of the lens performs the ASCII to XML transformation
that folds back an updated ASCII view back into the original XML source.

For example, suppose that a source $S$ consists of a pair consisting of a
boolean value and an integer, and that the view $V$ for $S$ consists of a
single integer. Further, suppose that we wish to define a \pput function for a
lens $\ell$ whose \get component is the function which given a pair $(b, n)$
simply produces the boolean value $b$. Let $f : \mathbb{B} \times \mathbb{Z}
\longrightarrow \mathbb{Z}$ be any function from $\mathbb{B} \times \mathbb{Z}
\longrightarrow \mathbb{Z}$. Then the lens $\ell_{f}$ defined by
\begin{align*}
\ell_{f}.\get \; (b, n) &= b\\
\ell_{f}.\pput \; (b, n) \; b' &= \begin{cases}
(b', n) & \text{if }b = b'\\
(b', f(b', n)) & \text{otherwise}\\
\end{cases}\\
\ell.\create \; b &= (b,f(b, 0))
\end{align*}
is a valid lens.

A user who applies a lens $\ell$ of type $\ell : S \Leftrightarrow V$ may expect
that $\ell.\pput \; (true, 1) \; false = (false, 0)$, reasoning that the
integer produced by the \pput function should be the number representation of
the boolean value $false$. Or perhaps she might expect that $\ell.\pput \;
(true, 1) \; false = (false, 1)$, reasoning that the two pieces of data in a
pair are independent and that the reasonable \pput behaviour in this case is to
restore the orginial integer. In either case, the user may be shocked to find
that in fact $\ell.\pput (true, 0) \; false = (false, 48)$! Given the \get
behaviour which just produces the boolean value from a pair, since $\ell_f$ is a
valid lens for any $f$, then the \pput function of a valid lens of type
$\mathbb{B} \times \mathbb{Z} \Leftrightarrow \mathbb{Z}$ can restore the
integer in a pair in any way so long as the old boolean value and the updated
boolean value are different.

In order to give a more constrained specification for the desired
lens we define the difference $\delta_S((a, b), (c, d))$ between $(a, b)$ and
$(c, d)$ by $$\delta_S((a, b), (c,
d)) = \begin{cases}
0 & \text{if }a = c, b = d\\
1 & \text{if }a=c, b \neq d\\
1 & \text{if }a \neq c, b = d\\
2 & \text{if }a \neq c, b \neq d
\end{cases}$$
On the view side, we define the difference $\delta_V(a, b)$
between $a$ and $b$ by, $$\delta_V(a, b) = \begin{cases}
0 & \text{if } a = b\\
1 & \text{if } a \neq b
\end{cases}$$

Next we define a lens $\partial(S, V) : \partial S \Leftrightarrow \partial V$
which describes how differences should be propagated. For this example, we
define $\partial(S,V)$ to simply be the \ccopy lens from $\partial S$ to
$\partial V$, keeping in mind that $\partial S = \partial V = \mathbb{Z}$.
\cref{difflenslaw} enables us to rule out $\ell_{bad}$, since
\begin{align*}
(\partial (S, V)).\get \; (\delta_S(\ell_{bad}.\pput \; (1,2) \; 3, (1,2))) &=
(\partial (S, V)).\get \; (\delta_S((0,3), (1,2)))\\
&= (\partial (S, V)).\get \; 2\\
&= 2\\
&\neq 1\\
&= \delta_V(3, 2)\\
&= \delta_V(3, \ell_{bad}.\get \; (1, 2))
\end{align*}
Consequently $\ell_{bad}$ is not a valid delta lens. On the
other hand, the lens $\ell_{good} : S \Leftrightarrow V$ defined by
\begin{align*}
\ell_{good}.\get \; (a, b) &= b\\
\ell_{good}.\pput \; (a, b) \; c &= (a, c)\\
\ell_{good}.\create \; b &= (0, b)
\end{align*}
is indeed a valid delta lens since the its \pput component always restores the
first factor.

\section{Strengthening Lens Laws}
\begin{definition}
A bidirectional transformation (bx) $R: M \Leftrightarrow N$ comprises of a
consistency relation $R \subseteq M \times N$ and restorers $\overrightarrow{R}
: M \times N \rightarrow N$ and $\overleftarrow{R} : M \times N \longrightarrow
N$.
\end{definition}
\begin{definition}
A bx is {\em correct} if the consistency restorers do restore consistency:
that is, for all $m \in M, n \in N$, we have $R(m, \overrightarrow{R}(m, n))$,
and dually.
\end{definition}
\begin{definition}
A bx is {\em hippocratic} if the consistency restores make no change
to and already-consistent pair of models: that is, for all $m \in M$, $n \in N$,
if $R(m, n)$ then $\overrightarrow{R}(m, n) = n$ and dually.
\end{definition}
\begin{definition}
A bx is {\em well behaved} if it is correct and hippocratic.
\end{definition}
For lenses, if $\ell : S \Leftrightarrow V$ is a lens, we take the
consistency relation $R$ to be given by $R(s, \ell.\get \; s)$ for all $s \in
S$, $\overrightarrow{\ell}(s, v) = \ell.\get \; s$, and $\overleftarrow{\ell} =
\ell.\pput$. By the PUTGET law, $\ell.\get \; (\ell.\pput \; s \; v) = v$,
hence $\ell$ is correct, and by the GETPUT laws, $\ell.\pput \; s \; (\ell.\get
\; s) = s$ and $\ell.\get \; (\ell.\create \; v) = v$, hence $\ell$ is
hippocractic. Therefore $\ell$ is a well-behaved bx.

Using the notation of Stevens \cite{stevens2010bidirectional},

The major shortcoming of this model of bx transformations is that it does not
take into account how changes should be propagated by transformation; a
well-behaved bx need only restore consistency when models go out of sync or
maintain consisteny if models are already in sync. This becomes an issue when
there are many different ways of restoring consistency

\section{Motivation and Related Work}
Foster et. al. \cite{foster2007combinators,foster2009bidirectional}
defined a lens $\ell : S \Leftrightarrow V$ from a source $S$ to a view $V$ to
be a triple of functions $\ell.\get : S \longrightarrow V$, $\ell.\pput : S
\longrightarrow S \longrightarrow V$ and $\ell.\create : V \longrightarrow S$
satisfying the following properties:
\begin{align*}
\ell.\get \; (\ell.\pput \; s \; v) &= v \tag{PUTGET}\\
\ell.\get \; (\ell.\create \; v) &= v \tag{CREATEGET}\\
\ell.\pput \; s \; (\ell.\get \; s) &= s \tag{GETPUT}
\end{align*}
While these two laws give some guarantees on how the \get, \pput and \create
components of a lens interact, they do not specify the behaviour of the \pput
component when the update to the view is non-trivial, because the GETPUT law
only says that the \pput component must return the original source unmodified
whenever the update to the view is a no-op. Consequently, a non-trivial update
to a view could lead to an undesired update to the source.

Therefore, several lens formalism have been proposed that explicitly have a
notion of an {\em update}. The first of these was Barbosa et. al.'s {\em
matching lenses}\cite{barbosa2010matching}. Matching lenses specifically deal
with the issue of {\em alignment}, since Foster et. al.'s original iteration
lens updates records positionally:
\begin{align*}
\ell^* .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n) &= t_1 \cdot
\ldots \cdot t_n\\
\text{ where } t_i &= \begin{cases}
\ell.\pput \; s_i \; v_i & \text{if } 1 \leq i \leq \min\{m, n\}\\
\ell.\create \; v_i & \text{if } \min\{m, n\} < i \leq n
\end{cases}\\
\end{align*}
In their work, Barbosa et. al. enrich the types of lenses with ``chunks''
identifying reorderable pieces of the source and view that should be re-aligned
after an update, and formulate behavioral laws that capture essential
constraints on the handling of chunks. They develop a core language
of matching lenses for strings, together with a set of ``alignment
combinators'' that implement a variety of alignment strategies. The obvious
shortcoming of matching lenses is that they only deal with alignment; they do
not solve the general issue of update propagation.

The next line of work is Diskin et. al.'s {\em delta
lenses}\cite{diskin2011asymmetric,diskin2011state}. Diskin et. al.'s delta
lenses work as follows: let $\ell : S \Leftrightarrow V$ be a lens from
$S$ to $V$. Assume that $s$ is a source, and that $v = \ell.\get s$. A user
performs an edits $d$ on $v$ to get $v'$: $v \xrightarrow{d} v'$. The \pput
function now takes the view edit $d$ and the original source $s$ and produces a
new source $s'$: $s \xrightarrow{\pput(d, s)} s'$. One thing that Diskin et. al.
do not do is that they do not give any concrete examples of delta lenses,
no composition, no notion of equivalence, and no combinators for constructing
delta lenses.

The next line of work is Hofmann et. al.'s {\em edit
lenses}\cite{hofmann2012edit}. Edit lenses are similar to delta lenses, but edit
lenses are more explicit. Specifically, the source and view edits are
represented as elements of a {\em monoid}, and the lens acts on sources and
views via monoid actions. That is if $v$ is a view and $d_v$ is an edit that
can be made on $v$ to produce $v'$, then $d_v \cdot v = v'$. The intended usage
of an edit lens is as follows. There are two users, one holding an element of
$S$ the other one an element of $V$ both referred to as {\em replicas}.
Initially, they hold $init_S \in S$ and $init_V \in V$ respectively. The users
then perform actions and propagate them across the lens. An action consists of
producing an edit $d_s$ (or $d_v$), applying it to one's current replica $s$
(resp. $v$), putting the edit through the lens to obtain an edit $d_v$ (resp.
$d_s$), and asking the user on the other side to apply $d_v$ (resp. $d_s$) to
their replica. One thing that Hofmann et. al. do that Diskin et. al. do not do
is that they do give concrete examples of edit lenses, define a suitable notion
of composition on them, and define combinators for constructing edit lenses.

The final line of work in this area is Pacheco et. al.'s
version of delta lenses\cite{pacheco2012delta}. Pacheco et. al. are the
first to give concrete examples of delta lenses, define a suitable notion
of composition on them, and define combinators for constructing delta
lenses. Pacheco et al. define a delta lens $\ell$ (d-lens for short), denote by
$\ell: S \trianglerighteq_{\blacktriangle} V$ to be a bidirectional
transformations that comprises four total functions:
\begin{align*}
\get &: S \longrightarrow V\\
\pput &: \forall(v, s) : V \times S. (v \triangle \get \;s) \longrightarrow
SA\\
\get_{\blacktriangle} &: \forall\{s':S, s:S\}. (s' \triangle s)
\longrightarrow (\get \; s')\triangle (\get \; s)\\
\pput_{\blacktriangle} &:\forall\{(v, s):V \times S\},d:(v \triangle \get
\;s).(\pput \; (v, s) \; d) \triangle s
\end{align*}
that satisfy the following propertes:
\begin{align*}
\get \; (\pput \; (v, s) \; d) = v \tag{PUTGET}\\
\pput \; (\get \; s, s) \; id = s \tag{GETPUT}\\
\get_{\blacktriangle} \; (\pput_{\blacktriangle} \; d) = d
\tag{$\text{PUTGET}_{\blacktriangle}$}\\
\pput_{\blacktriangle} \; id = id \tag{$\text{PUTGET}_{\blacktriangle}$}
\end{align*}
The \get function is a just like a \get function in Foster's et. al.'s original
work. The \pput function is different though. Let $s \in S$ and $v \in V$, and
let $d$ be an update from $v$ to $\ell.\get \; s$; for example, $d$ is a
sequence of insertions and deletions that transforms $v$ to $\ell.\get \; s$.
Then \pput takes in $s$ and $v$ as input, and in addition receives $d$ to
compute $\pput \; (v, s) \; d$.

The $\get_{\blacktriangle}$ and $\pput_{\blacktriangle}$ functions are functions
that transform deltas to deltas. The last two axioms for delta lenses say that
if $d$ is a delta from $v$ to $\ell.\get \; s$ and $\Delta$ is a delta from
$\pput \; (v, s) \; d$ to $s$, then $\get_{\blacktriangle} \; \Delta = d$, and
that if $d$ is the identity delta, then $\Delta$ is also the identity delta.

In addition to describing a delta lens language, Pacheco et. al. also
address the following shortcoming with edit lenses: edit lenses include
combinators for constructing lenses over {\em container types}.
In the simplest case, container types can be thought of as algebraic datatypes
such as lists and trees; each value of a container type has a certain {\em
shape}, and each shape has a set of {\em positions} associated with it. For
instance, the shape of a list of size $n$ is just $n$, while the positions of
such a list are the numbers $0, 1, \ldots, n-1$.

The shortcoming of the container edit lenses addressed by Pacheco et. al.
elements is that these lenses always preserve the number of positions; for
instance if $\ell$ is a container lens that maps lists to trees, and if $xs$ is
a list of size $n$, then the number of positions of $\ell.\get \; xs$ is also
$n$. Pacheco et. al. define a new set of combinators for container lenses that
can say map a list of size $5$ to a tree with 4 nodes.

In this article we show that it is not necessary for the \pput component of a lens
to explicitly propagate updates: instead, we can specify what the put function should
evaluate to, by imposing a more precise type judgement on lenses.

The idea of the stronger typing judgement is that we embed sources and views inside
of {\em difference structures} and then we define a lens that describes how the
differences behave as they cross in-between the source and the view. A
lens that is validly typed is then a lens that preserves this overall structure.

Here are the actual definitions:
\begin{definition}
A weighted set is a pair $(X, |\cdot|)$ where $X$ is a set and $|\cdot| : X
\longrightarrow \mathbb{R}^{\geq 0}$ is a map from $X$ to the non-negative
reals.
\end{definition}
\begin{definition}
A difference structure is a triple $(S, \partial S, \delta_S)$ where $S$ and
$\partial S$ are weighted sets, and $\delta_S:
S \times S \longrightarrow \partial S$ is a function satisfying
\begin{align*}
\delta(s, s') &= \delta(s', s)\tag{A1}\\
|\delta(s, s)| &= 0\tag{A2}\\
|\delta(s, s')| &\leq |s| + |s'|\tag{A3}
\end{align*}
\end{definition}
\begin{definition}
Let $(S, \partial S, \delta_S)$ and $(V, \partial V, \delta_V)$ be
difference structures. Let $\partial(S, V) : \partial S \Leftrightarrow \partial
V$ be a lens from $\partial S$ to $\partial V$. Then a lens $\ell : S
\Leftrightarrow V$ also has type $\ell : (S, \partial S,
\delta_S) \xLeftrightarrow{\partial(S, V)} (V, \partial V, \delta_V)$ if for
all $s \in S, v \in V$,
\begin{equation}\label{difflenslaw}
(\partial (S, V)).\get\; (\delta_S(\ell.\pput \; s \; v, s)) = \delta_V(v,
\ell.\get \; s)
\end{equation}
\end{definition}
The extra difference structure on the sources and the views, as well as the lens
$\partial(S, V)$ force the lens $\ell$ to satisfy \cref{difflenslaw}; this law is
therefore a specification that constrains what $\ell$ can do.

For example, suppose that a source $S$ consists of a pair of integers, and that the
view $V$ for $S$ consists of a single integer. Further, suppose that we wish
for the \get component of any lens from $S$ to $V$ to simply project away the
first factor:that is for any $\ell : \mathbb{Z} \times \mathbb{Z}
\Leftrightarrow \mathbb{Z}$, $\ell.\get \; (a, b) = b$ for all integers $a, b$.
Suppose further that we wish to force the \pput function of any lens $\ell$
defined on the data to restore the factor projected away by the \get function:
that is, we wish to enforce that for all integers $a, b, c$, $\pi_1(\ell.\pput
\; (a, b) \; c) = a$, where $\pi_1$ is the projection on the first factor.

Now consider the lens $\ell_{bad}$ defined by
\begin{align*}
\ell_{bad}.\get \; (a, b) &= b\\
\ell_{bad}.\pput \; (a, b) \; c &= \begin{cases}
(a, c) & \text{if }b = c\\
(0, c) & \text{otherwise}\\
\end{cases}\\
\ell.\create \; b &= (0,b)
\end{align*}
Then $\ell_{bad}$ is a valid lens even though $\ell_{bad}$ does not
always restore the first factor. For example, observe that $\ell.\pput (1, 2) \;
3 = (0,3)$. In order to give a more constrained specification for the desired
lens we define the difference $\delta_S((a, b), (c, d))$ between $(a, b)$ and
$(c, d)$ by $$\delta_S((a, b), (c,
d)) = \begin{cases}
0 & \text{if }a = c, b = d\\
1 & \text{if }a=c, b \neq d\\
1 & \text{if }a \neq c, b = d\\
2 & \text{if }a \neq c, b \neq d
\end{cases}$$
On the view side, we define the difference $\delta_V(a, b)$
between $a$ and $b$ by, $$\delta_V(a, b) = \begin{cases}
0 & \text{if } a = b\\
1 & \text{if } a \neq b
\end{cases}$$

Next we define a lens $\partial(S, V) : \partial S \Leftrightarrow \partial V$
which describes how differences should be propagated. For this example, we
define $\partial(S,V)$ to simply be the \ccopy lens from $\partial S$ to
$\partial V$, keeping in mind that $\partial S = \partial V = \mathbb{Z}$.
\cref{difflenslaw} enables us to rule out $\ell_{bad}$, since
\begin{align*}
(\partial (S, V)).\get \; (\delta_S(\ell_{bad}.\pput \; (1,2) \; 3, (1,2))) &=
(\partial (S, V)).\get \; (\delta_S((0,3), (1,2)))\\
&= (\partial (S, V)).\get \; 2\\
&= 2\\
&\neq 1\\
&= \delta_V(3, 2)\\
&= \delta_V(3, \ell_{bad}.\get \; (1, 2))
\end{align*}
Consequently $\ell_{bad}$ is not a valid delta lens. On the
other hand, the lens $\ell_{good} : S \Leftrightarrow V$ defined by
\begin{align*}
\ell_{good}.\get \; (a, b) &= b\\
\ell_{good}.\pput \; (a, b) \; c &= (a, c)\\
\ell_{good}.\create \; b &= (0, b)
\end{align*}
is indeed a valid delta lens since the its \pput component always restores the
first factor.

For a more realistic example using a difference lens combinator, let $S
= name \cdot dates$ and $V = name$. Let $\ell :
S \Leftrightarrow V = (copy \; name) \cdot (del \; dates)$, and suppose that we
wish to define a lens of type $S^* \Leftrightarrow V^*$, where the difference
function used on the view side is the min-edit distance difference function.
Assuming that we inductively use the difference function
$$ \delta(d_1, d_2) =
\begin{cases}
0 & \text{if }d_1 = d_2 \\
1 & \text{otherwise}
\end{cases}
$$
for both the name and date fields, then with the min-edit distance difference
function, the difference between the string $v$ given by
\begin{lstlisting}
John Doe
Jane Smith
\end{lstlisting}
and the string $v'$ given by
\begin{lstlisting}
Jane Smith
John Doe
\end{lstlisting}
is the empty list. Let $s$ =
\begin{lstlisting}
John Doe,1900-2000
Jane Smith,1901-2001
\end{lstlisting}Then $\ell.\get \; s = v$. Now assume that $v$ is edited to $v'$
given by
\begin{lstlisting}
Mark Moe
John Doe
Jane Smith
\end{lstlisting}
where a new record \lstinline|Mark Moe| has been added to $v$.
With the min-edit distance on the view side, $\delta(v', \ell.\get \; s)$ is
the singleton list [``Mark Moe''].

On the source side, the min-edit distance matching lens $\ell^*_{MIN-EDIT}$,
$\ell^*_{MIN-EDIT}.\pput \; s \; v'$ is the string
\begin{lstlisting}
Mark Moe,0000-0000
John Doe,1900-2000
Jane Smith,1901-2001
\end{lstlisting}
while with the positional lens $\ell^*$, then $\ell^*.\pput \; s \; v'$ is the
string
\begin{lstlisting}
Mark Moe,1900-2000
John Doe,1901-2001
Jane Smith,0000-0000
\end{lstlisting}
With the min-edit distance difference function,
$\delta(\ell^*_{MIN-EDIT}.\pput \; s \; v', s)$ is the singleton list
$[\text{``Mark Moe, 0000-0000''}]$, while with the
positional lens, $\delta(\ell^*.\pput \; s \; v', s)$ is the list
given by
\begin{lstlisting}
Mark Moe,1900-2000
(0,1)
(0,1)
\end{lstlisting}
The min-edit distance satisfies \cref{difflenslaw} since
\begin{align*}
(\partial (S, V))^*.\get\; (\delta_S(\ell^*_{MIN-EDIT}.\pput \; s \; v', s)) &=
(\partial (S, V))^*.\get\; ([\text{``Mark Moe, 1900-2000''}])\\
&= [\text{``Mark Moe''}]\\
&=
\delta_V(v', \ell.\get \; s)
\end{align*}
On the other hand, the positional lens does not satisfy \cref{difflenslaw} since
\begin{align*}
(\partial (S, V))^*.\get\; (\delta_S(\ell^*.\pput \; s \; v', s)) &=
(\partial (S, V))^*.\get\; ([\text{``Mark Moe, 1900-2000''}, (0,1), (0,1)])\\
&=[\text{``Mark Moe''}, (0,*), (0,*)]\\
& \neq \delta_V(v', \ell.\get \; s)
\end{align*}
Let $\mathcal{M}_V$ be the set of min-edit distance alignments on the view side,
and $\mathcal{M}_S$ the corresponding alignments on the source side.
Then, the typing judgement
$$\ell^*_{MIN-EDIT}:(S, (\partial S \sqcup S)^*, \match(\mathcal{M}_S, \delta_s))
\xLeftrightarrow{\partial(S, V)^*} (V, (\partial V \sqcup V)^*, \match(\mathcal{M}_V, \delta_v))$$
is valid, whereas the typing judgement
$$\ell^*:(S, (\partial S \sqcup S)^*, \match(\mathcal{M}_S, \delta_s))
\xLeftrightarrow{\partial(S, V)^*} (V, (\partial V \sqcup V)^*, \match(\mathcal{M}_V, \delta_v))$$
is not.
\iffalse
For example, if $\ell : S \Leftrightarrow V$ is a lens, Foster's initial
definition of the iteration lens $\ell^* : S^* \Leftrightarrow V^*$ was
\begin{align*}
\ell^* .\get \; (s_1, \ldots, s_n) &= (\ell.\get \; s_1) \cdot \ldots
\cdot (\ell.\get \; s_n)\\
\ell^* .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n) &= t_1 \cdot
\ldots \cdot t_n\\
\text{ where } t_i &= \begin{cases}
\ell.\pput \; s_i \; v_i & \text{if } 1 \leq i \leq \min\{m, n\}\\
\ell.\create \; v_i & \text{if } \min\{m, n\} < i \leq n
\end{cases}\\
\ell^*.\create \; (v_1, \ldots, v_n) &= (\ell.\create \; v_1) \cdot \ldots
\cdot (\ell.\create \; v_n)
\end{align*}
This lens always propagates updates positionally, which is unintended behaviour
in many cases. For example suppose that $\ell$ is a lens that transforms
source data in the form of an XML document representing the names, dates and
nationalities of a collection of classical music composers such as,

\begin{lstlisting}
<composers>
<composer>
<name>Jean Sibelius</name>
<dates>1865-1957</dates>
<nationality>Finnish</nationality>
</composer>
<composer>
<name>Aaron Copland</name>
<dates>1910-1990</dates>
<nationality>American</nationality>
</composer>
<composer>
<name>Benjamin Britten</name>
<dates>1913-1976</dates>
<nationality>English</nationality>
</composer>
</composers>
\end{lstlisting}
into view data stoored in ASCII text representing the names and dates of each
composer:
\begin{lstlisting}
Jean Sibelius, 1865-1957
Aaron Copland, 1910-1990
Benjamin Britten, 1913-1976
\end{lstlisting}
Now suppose that the elements in the view are reordered:
\begin{lstlisting}
Aaron Copland, 1910-1990
Benjamin Britten, 1913-1976
Jean Sibelius, 1865-1957
\end{lstlisting}
Then with the positional iteration lens, the updated source is
\begin{lstlisting}
<composers>
<composer>
<name>Aaron Copeland</name>
<dates>1910-1990</dates>
<nationality>Finnish</nationality>
</composer>
<composer>
<name>Benjamin Britten</name>
<dates>1913-1976</dates>
<nationality>American</nationality>
</composer>
<composer>
<name>Jean Sibelius</name>
<dates>1865-1957</dates>
<nationality>English</nationality>
</composer>
</composers>
\end{lstlisting}
so each of the composers now has the wrong nationality in the source.

\section{Difference functions}
\begin{definition}
A weighted set is a pair $(X, |\cdot|)$ where $X$ is a set and $|\cdot| : X
\longrightarrow \mathbb{R}^{\geq 0}$ is a map from $X$ to the non-negative
reals.
\end{definition}
\begin{definition}
A difference structure is a triple $(S, \partial S, \delta_S)$ where $S$ and
$\partial S$ are weighted sets, and $\delta_S:
S \times S \longrightarrow \partial S$ is a function satisfying
\begin{align*}
\delta(s, s') &= \delta(s', s)\tag{A1}\\
|\delta(s, s)| &= 0\tag{A2}\\
|\delta(s, s')| &\leq |s| + |s'|\tag{A3}
\end{align*}
We refer to the function $\delta$ as the difference function of the structure.
\end{definition}
\iffalse

\begin{example}
Let $X$ be any set. Define $x \leq y$ if and only if $x=y$, and define $|X|=1$
for all $x \in X$. Let $Q = \{0,1\}$, and define
$$\delta(a, b) = \begin{cases} 1 & \text{if } a \neq b\\
0 & \text{otherwise}
\end{cases}$$
Then $\delta$ is a difference function.
\end{example}

Various different edit distances on strings give a way of defining delta
functions on sets of strings:
\begin{example}
The {\em Levenshtein distance} between two words is the minimum number of
single-character edits (insertions, deletions or substitutions) required to
change one word into the other. For example, the Levenshtein distance between
"kitten" and "sitting" is 3, since the following three edits change one into
the other, and there is no way to do it with fewer than three edits:

\begin{enumerate}
  \item
  kitten -> sitten (substitution of "s" for "k")
  \item
  sitten -> sittin (substitution of "i" for "e")
  \item
  sittin -> sitting (insertion of "g" at the end).
\end{enumerate}
\end{example}
\begin{example}
In contrast, the Longest Common Subsequence(LCS) distance only allows for
insertions and deletions:
\begin{enumerate}
  \item
  delete k at 0
  \item
  insert s at 0
  \item
  delete e at 4
  \item
  insert i at 4
  \item
  insert g at 6
\end{enumerate}
for a total cost/distance of 5 operations.
\end{example}
Both Levenshtein and LCS distances are integer-valued distance functions since
they are metrics.
\begin{example}
If all the strings in a set are of the same length, then the {\em Hamming
distance} is a difference function since the Hamming distance is also a metric.
\end{example}
\begin{example}
Alignment algorithms also give us a way of defining integer-valued distance
functions on sets of strings by taking the difference of two strings to be the
total number of characters that are not equal according to the alignment.

When deciding which alignment to choose between two strings, we often wish to
choose an alignment that is optimal for alignment strategy. With
the positional alignment strategy, given two strings $s$ and $t$, then $s_1$ is
matched to $t_1$, $s_2$ is matched to $t_2$ and so on.
Also, $\mathcal{M}_{POS}$ is an optimal set of alignments for a positional
alignment strategy.

For example, using the positional alignment strategy on the strings "kitten" and
"sitting", then "k" is matched to "s", "i" to "i" and so on. The characters
which are included in the delta from "kitten" are "k", "e" while the
in the characters included from "sitting" are "s", "i" and "g", hence the delta
is equal to 5.

For the non-crossing strategy, we choose among alignments which are increasing
rather than positional, while for the min-edit distance strategy, we consider
all possible alignments.
\end{example}

\begin{example}\label{naturalnumbers}
Let $P$ be the set of non-negative integers with weight function $|a| = a$. Then
$\delta$ is a difference function.
Define $\delta(a, b) = |a-b|$. Then $\delta$ is a difference function.
\end{example}
\begin{example}
Let $X$ be a normed vector space and define $\delta(x, y) = |x-y|$. Then
$\delta$ is a difference function (in fact, $\delta$ is also a metric).
\end{example}
\begin{example}\label{divisibility}
Let $P$ be the set of positive integers with weight function $|\cdot|$ defined
by $|p_1^{a_1} \cdot \ldots \cdot p_n^{a_n}| = a_1 + \ldots + a_n$. Define
$\delta(a, b) = ab/(gcd(a, b))^2$.
Then $\delta$ is a difference function.
\end{example}
\begin{example}\label{boolean}
Let $X$ be a set and $P$ the set of finite subsets of $X$ with weight function
given by the size of the set. Define $\delta(A, B) = (A \setminus B) \cup (B
\setminus A)$. Then $\delta$ is a difference function.
\end{example}
\begin{example}\label{general}
Generalizing \cref{divisibility} and
\cref{boolean}, let $P$ be a graded complete semi-modular lattice satisfying
$|\bigwedge P|=0$. Define $\delta(a, b) = \bigwedge \{x \; | \; (a \wedge b)
\vee x = a\} \vee \bigwedge \{x \; | \; (a \wedge b) \vee x = b\}$. Then
$\delta$ is a difference function.
\end{example}
\begin{proof}
Let $a, b \in P$. Let $X = \{x \; | \; (a \wedge b) \vee x = a\}$ and $Y = \{y
\; | \; (a \wedge b) \vee y = b\}$
\begin{enumerate}
  \item[(A1)]
  Certainly $\delta(a, b) = \delta(b, a)$.
  \item[(A2)]
  Observe that $\bigwedge X = \bigwedge Y = \bot$, hence $\delta(a,
  a) = \bot$, and $|\delta(a, a)| = |\bot| = 0$.
  \item[(A3)]
  Since $P$ is semi-modular, then $|x \wedge y| + |x \vee y| \leq |x| + |y|$
  for all $x, y \in P$, from which it follows that $$|\delta(a, b)| =
  \left|\bigwedge X \vee \bigwedge Y\right| \leq \left|\bigwedge X \vee
  \bigwedge Y\right| + \left|\bigwedge X \wedge \bigwedge Y\right| \leq
  \left|\bigwedge X\right| + \left|\bigwedge Y\right|$$
\end{enumerate}
\end{proof}
\fi

\subsection{Constructions on difference functions}
\subsubsection{Products}
Given difference functions $\delta_i: P_i \times P_i \longrightarrow Q_i$ for
$i \in \{1,2\}$, define $\delta_1 \times \delta_2 :(P_1 \times P_2) \times
(P_1 \times P_2) \longrightarrow (Q_1 \times Q_2)$ by
$$(\delta_1 \times \delta_2)((x, y), (x', y')) = (\delta_1(x, x'), \delta_2(y,
y'))$$
\begin{claim}
$\delta_1 \times \delta_2$ is a difference function.
\end{claim}
\begin{proof}
Let $x, x' \in P_1$ and $y, y' \in P_2$.
\begin{enumerate}
  \item[(A1)]
  Observe that
  \begin{align*}
(\delta_1 \times \delta_2)((x, y), (x', y')) &= (\delta_1(x, x'),
\delta_2(y, y'))\\
&= (\delta_1(x', x), \delta_2(y', y))\\
&= (\delta_1 \times \delta_2)((x', y'), (x, y))
\end{align*}
\item[(A2)] Observe that
$$|(\delta_1 \times \delta_2)((x, y), (x, y))| = |(\delta_1(x, x),
\delta_2(y, y))| = |\delta_1(x, x)| + |\delta_2(y,y)| = 0 + 0 = 0$$
\item[(A3)]
Observe that
\begin{align*}
|(\delta_1 \times \delta_2)((x, y), (x', y'))| &= |(\delta_1(x, x'),
\delta_2(y, y'))|\\
&= |(\delta_1(x, x')| + |(\delta_2(y, y')|\\
&\leq (|x| + |x'|) + (|y| + |y'|)\\
&= (|x| + |y|) + (|x'| + |y'|)\\
&= |(x, y)| + |(x', y')|
\end{align*}
\end{enumerate}
\end{proof}
\subsubsection{Coproducts}
Given difference functions $\delta_i: P_i \times P_i \longrightarrow Q_i$ for
$i \in \{1,2\}$, define $\delta_1 \sqcup \delta_2 :P_1 \sqcup P_2
\longrightarrow Q_1 \sqcup Q_2 \sqcup (P_1 \times P_2)$ by
$$(\delta_1 \sqcup \delta_2)(x, y) =
\begin{cases}
\delta_1(x, y) & \text{if }x, y \in P_1\\
\delta_2(x, y) & \text{if }x, y \in P_2\\
(x, y) & \text{if }x \in P_1, y \in P_2\\
(y, x) & \text{if }x \in P_2, y \in P_1
\end{cases}$$
\begin{claim}
$\delta_1 \sqcup \delta_2$ is a difference function
\end{claim}
\begin{proof}
Let $x, x' \in P_1$ and $y, y' \in P_2$.
\begin{enumerate}
  \item[(A1)]
  If $x, y \in P_i$, then $(\delta_1 \sqcup \delta_2)(x, y) = \delta_i(x, y) =
  \delta_i(y, x)$. If $x \in P_1$ and $y \in P_2$, then $(\delta_1 \sqcup
  \delta_2)(x, y) = (x, y) = (\delta_1 \sqcup \delta_2)(y, x)$, and similarly
  if $x \in P_2$ and $y \in P_1$, then $(\delta_1 \sqcup \delta_2)(x, y) =
  (y,x) = (\delta_1 \sqcup \delta_2)(y, x)$
  \item[(A2)] If $x \in P_i$, then $|(\delta_1 \sqcup \delta_2)(x, x)| =
  |\delta_i(x, x)| = 0$
  \item[(A3)]
  If $x, y \in P_i$, then $|(\delta_1 \sqcup \delta_2)(x, y)| = |\delta_i(x, y)|
  \leq |x| + |y|$. If $x \in P_1$ and $y \in P_2$, then $|(\delta_1 \sqcup
  \delta_2)(x, y)| = |(x, y)| = |x| + |y|$, and if $x \in P_2$ and $y \in
  P_1$, then $|(\delta_1 \sqcup \delta_2)(x, y)| = (|(y, x)|) = |y| + |x|$.
\end{enumerate}
\end{proof}
\subsubsection{List difference functions}
Let $P$ be a weighted set. Let $x=(x_1, \ldots, x_m), y=(y_1, \ldots, y_n)$ be
in $P^*$. Given a difference function $\delta: P \times P \longrightarrow Q$, we
define the list difference function $\delta^*:P^* \longrightarrow (Q \sqcup
P)^*$ on $x$ and $y$ to be the list given by $\delta(x_i, y_i)$ for $1 \leq i
\leq m$, except that we omit $\delta(x_i, y_i)$ in position $i$ if $x_i = y_i$.
For $m < i \leq n$ we add $y_i$ at position $i$. Finally we collapse the list
to get rid of the empty spaces left after the omissions.

For instance suppose that $S$ is the set of all letters in the alphabet and
$\partial S = \{\epsilon\} \sqcup \{\{x, y\} \; | \; x, y \in S, x \neq y\}$
with $$\delta(x, y) =
\begin{cases}
\epsilon & \text{if } x = y\\
\{x,y\} & \text{if } x \neq y
\end{cases}$$
Then $\delta_S^*((a, c, b, d), (a, b, c, d, e)) = (\{b, c\}, \{b, c\},
e)$ where we match and remove the first $a$ and the fourth $d$ in. Since $b \neq
c$ we have $\{b,c\}$ in the second and third positions, and since $e$ is
unmatched, we include $e$ in the fifth position. Finally we collapse the list to
get rid of the two empty spaces created.

\begin{claim}
$\delta^*$ is a difference function
\end{claim}
\begin{proof}
Let $x = (x_1, \ldots, x_m), y = (y_1, \ldots, y_n) \in P^*$.
\begin{enumerate}
  \item[(A1)]
  By construction $\delta^*(x, y) = \delta^*(y, x)$.
  \item[(A2)] Observe that $|\delta^*(x, x)| = \sum_{i=1}^m|\delta(x_i,
  x_i)| =  \sum_{i=1}^m 0 = 0$.
  \item[(A3)]
  Assume WLOG that $m \leq n$. Then
  \begin{align*}
|\delta^*(x, y)| &=
\sum_{i=1}^m|\delta(x_i, y_i)| + \sum_{i=m+1}^n |y_i|\\
&\leq \sum_{i=1}^m|x_i| + |y_i| + \sum_{i=m+1}^n |y_i|\\
&= \sum_{i=1}^m|x_i| + \sum_{i=1}^n |y_i|\\
&= |x| + |y|
\end{align*}
\end{enumerate}
\end{proof}
\subsubsection{Matching difference functions}
Now let $\mathcal{M} = \{\varphi_{(x_1, \ldots, x_m), (y_1, \ldots, y_n)} \; |
\; x_i, y_j \in P, m \leq n\}$ be a set of injective functions $\varphi_{(x_1,
\ldots, x_m), (y_1, \ldots, y_n)} : [m] \longrightarrow [n]$. Given a difference
function $\delta:
P \times P \longrightarrow Q$, we define the difference function
$\match(\mathcal{M}, \delta):P^* \longrightarrow (Q \sqcup P)^*$ on $x$
and $y$ to be the list given by $\delta(x_i, y_{\varphi(i)})$ in position
$\varphi(i)$ for $1 \leq i \leq m$, except that we omit $\delta(x_i,
y_{\varphi(i)})$ in position $\varphi(i)$ if $x_i = y_{\varphi(i)}$. For $i \in
[n] \setminus \varphi([m])$ we add $y_i$ at position $i$. Finally we collapse
the list to get rid of the empty spaces left after the omissions.

Going back to our previous example, let $x = \{a, c, b, d\}$ and $y = \{a, b, c,
d, e\}$, and assume that $\varphi_{x, y} : [4] \longrightarrow [5]$ is given by
$$\varphi(1) = 1, \; \varphi(2) = 3, \; \varphi(3) = 2, \text{ and } \varphi(4)
= 4$$ Then $\match(\mathcal{M}, \delta)((a, c, b, d), (a, b, c, d, e)) =
(e)$.
\begin{claim}
$\match(\mathcal{M}, \delta)$ is a difference function.
\end{claim}
\begin{proof}
Let $x = \{x_1, \ldots, x_m\}, y = \{y_1, \ldots, y_n\} \in \match(P)$.
\begin{enumerate}
  \item[(A1)]
  By construction $\match(\mathcal{M}, \delta)(x, y) =
  \match(\mathcal{M}, \delta)(y, x)$.
  \item[(A2)] Observe that $|(\match(\mathcal{M}, \delta))(x, x)| =
  \sum_{i=1}^m|\delta(x_i, x_{\varphi(i)}| = |\delta(x_i, x_i)| =   \sum_{i=1}^m
  0 = 0$.
  \item[(A3)]
  Assume WLOG that $m \leq n$. Then
  \begin{align*}
|\match(\mathcal{M}, \delta)(x, y)| &=
\sum_{i=1}^m|\delta(x_i, y_{\varphi(i)})| + \sum_{i \in [n] \setminus
\varphi([m])} |y_i|\\
&\leq \sum_{i=1}^m|x_i| + |y_{\varphi(i)}| + \sum_{i \in [n] \setminus
\varphi([m])} |y_i|\\
&= \sum_{i=1}^m|x_i| + \sum_{i=1}^n |y_i|\\
&= |x| + |y|
\end{align*}
\end{enumerate}
\end{proof}
\fi
\section{Lens Combinators}
\subsection{Copy}
Let $P$ be any set. Define the $\ccopy(P) : P \Leftrightarrow P$ by
\begin{align}
\ccopy(P).\get \; s &= s\\
\ccopy(P).\pput \; s \; v &= v\\
\ccopy(P).\create \; v &= v
\end{align}
Then $\ccopy(P)$ is a lens. Now let $\partial S$ be a set, $\delta :
S \times S \longrightarrow \partial S$ a function.
\begin{claim}
$\ccopy(S) : S \xLeftrightarrow{\delta , \ccopy(\partial S), \delta} S$.
\end{claim}
\begin{proof}
We need to show that $(\ccopy(\partial S)).\get \; (\delta(\ccopy(S).\pput
\; s \; v, s)) = \delta(v, \ccopy(S).\get \; s)$ which follows from the fact
that $$ (\ccopy(\partial S)).\get \; (\delta(\ccopy(S).\pput \; s
\; v, s)) = \delta(\ccopy(S).\pput \; s \; v, s) = \delta(v, \ccopy(S).\get
\; s) $$
\end{proof}
\subsection{Constant}
Let $S$ be any set with $s'\in S$. Define the \const lens $\const(S) : S
\Leftrightarrow \{\star\}$ by
\begin{align*}
\const.\get \; s &= \star\\
\const.\pput \; s \; v &= s\\
\const.\create \; v &= s'
\end{align*}
Then $\const(S)$ is a lens. 
\begin{claim}
Let $\partial S$ be any set, and $\delta : S \times S \longrightarrow \partial
S$ any function satisfying $\delta(s, s) = \delta(s', s')$ for all $s, s' \in
S$. Define $\beta : \{\star\} \times \{\star\} \longrightarrow \partial S$ by
$\beta(\star, \star) = \delta(s, s)$ for any $s \in S$. Let $\ell : \partial S
\Leftrightarrow \partial S$ be any lens satisfying $\ell.\get \; \delta(s, s) =
\delta(s, s)$ for all $s \in S$. Then $\const(S): S \xLeftrightarrow{\delta,
\ell, \beta} \{\star\}$.
\end{claim}
\begin{proof}
We need to show that $\ell.\get \; (\delta(\const(S).\pput \; s \; \star, s)) =
\beta (\star, \const(S).\get \; s)$
which follows from the fact that
$$ \ell.\get \; (\delta(\const(S).\pput \; s \; \star, s)) = \ell.\get
\; \delta(s, s) = \delta(s, s) = \beta(\star, \star) = \beta(\star,
\const(S).\get \; s)$$
\end{proof}

\subsection{Default}
Let $\ell : S \Leftrightarrow V$ be a lens. Let $f :
V \longrightarrow S$ be any function. Define the lens $\default(\ell, f) :
S \Leftrightarrow V$ by
\begin{align*}
\default(\ell, f).\get \; s &= \ell.\get \; s\\
\default(\ell, f).\pput \; s \; v &= \ell.\pput \; s \; v\\
\default(\ell, f).\create \; v &= \ell.\pput \; (f \; v) \; v
\end{align*}

\begin{claim}
If $\ell : S \xLeftrightarrow{\delta_S, \partial (S, V), \partial V}
V$ and $f: S \longrightarrow V$, then $\default(\ell, f) : S
\xLeftrightarrow{\delta_S, \partial (S, V), \partial V} V$.
\end{claim}
\begin{proof}
Follows from the fact that $\ell : S \xLeftrightarrow{\delta_S, \partial (S, V),
\partial V} V$.
\end{proof}
\subsection{Product}
Let $\ell_1 : S_1 \Leftrightarrow V_1$ and $\ell_2 : S_2 \Leftrightarrow V_2$ be
lenses. Define the lens $\ell_1 \times \ell_2 : (S_1 \times S_2) \Leftrightarrow
(V_1 \times V_2)$ by
\begin{align*}
(\ell_1 \times \ell_2).\get \; (s_1, s_2) &= (\ell_1.\get \; s_1, \ell_2.\get
\; s_1)\\
(\ell_1 \times \ell_2).\pput \; (s_1, s_2) \; (v_1, v_2) &= (\ell_1.\pput \; s_1
\; v_1, \ell_2.\pput \; s_2 \; v_2)\\
(\ell_1 \times \ell_2).\create \; (v_1, v_2) &= (\ell_1.\create \; v_1,
\ell_2.\create \; v_2)
\end{align*}
\begin{claim}\label{productislens}
If $\ell_1 :S_1 \xLeftrightarrow{\delta_{S_1}, \partial (S_1, V_1),
\delta_{V_1}} V_1$ and $\ell_2 : S_2 \xLeftrightarrow{\delta_{S_2}, \partial
(S_2, V_2), \delta_{V_2}} V_2$, then \\
$\ell_1 \times \ell_2  : S_1 \times S_2 
\xLeftrightarrow{\delta_{S_1} \times \delta_{S_2}, \partial (S_1, V_1) \times
\partial (S_2, V_2), \delta_{V_1} \times \delta_{V_2}} V_1 \times V_2$.
\end{claim}
\begin{proof}
We need to show that
$$(\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \; (\delta((\ell_1 \times
\ell_2).\pput \; (s_1, s_2) \; (v_1, v_2), (s_1, s_2))) = \delta((v_1, v_2),
(\ell_1 \times \ell_2).\get \; (s_1, s_2))$$ which follows from the fact that
\begin{align*}
&(\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \; (\delta((\ell_1 \times
\ell_2).\pput \; (s_1, s_2) \; (v_1, v_2), (s_1, s_2)))\\
&= (\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \;
(\delta((\ell_1.\pput \; s_1 \; v_1, \ell_2.\pput \; s_2 \; v_2), (s_1, s_2))\\
&= (\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \;
(\delta(\ell_1.\pput \; s_1 \; v_1, s_1), \delta(\ell_2.\pput \; s_2 \; v_2,
s_2))\\
&= ((\partial (S_1, V_1)).\get \; (\delta(\ell_1.\pput \;
s_1 \; v_1, s_1)), (\partial (S_2, V_2)).\get \; (\delta(\ell_2.\pput \; s_2 \;
v_2, s_2)))\\
&= (\delta(v_1, \ell_1.\get \; s_1), \delta(v_2, \ell_2.\get \; s_2))\\
&= \delta((v_1, v_2), (\ell_1.\get \; s_1, \ell_2.\get \; s_2))\\
&= \delta((v_1, v_2), (\ell_1 \times \ell_2).\get \; (s_1, s_2))
\end{align*}
\end{proof}
\subsection{Sum}
Let $\ell_1 : S_1 \Leftrightarrow V_1$ and $\ell_2 : S_2 \Leftrightarrow V_2$ be
lenses. Define the lens
$\ell_1 \sqcup \ell_2 :
(S_1 \sqcup S_2) \Leftrightarrow (V_1 \sqcup V_2)$ by
\begin{align*}
(\ell_1 \sqcup \ell_2).\get \; s &=
\begin{cases}
\ell_1.\get \; s & \text{if } s \in S_1\\
\ell_2.\get \; s & \text{if } s \in S_2\\
\end{cases}\\
(\ell_1 \sqcup \ell_2).\pput \; s \; v&=
\begin{cases}
\ell_1.\pput \; s \; v& \text{if } s \in S_1, \; v \in V_1\\
\ell_2.\pput \; s \; v& \text{if } s \in S_2, \; v \in V_2\\
\ell_1.\create \; v & \text{if } s \in S_2, \; v \in V_1\\
\ell_2.\create \; v & \text{if } s \in S_1, \; v \in V_2\\
\end{cases}\\
(\ell_1 \sqcup \ell_2).\create \; v &=
\begin{cases}
\ell_1.\create \; v & \text{if } v \in V_1\\
\ell_2.\create \; v & \text{if } v \in V_2\\
\end{cases}
\end{align*}

\begin{claim}
If $\ell_1 : S_1  \xLeftrightarrow{\delta_{S_1}, \partial (S_1, V_1),
\delta_{V_1}} V_1$ and $\ell_2 : S_2 \xLeftrightarrow{\delta_{S_2}, \partial
(S_2, V_2), \delta_{V_2}} V_2$, then \\
$\ell_1 \sqcup \ell_2 : S_1 \sqcup S_2 \xLeftrightarrow{\delta_{S_1} \sqcup
\delta_{S_2}, \partial (S_1, V_1) \sqcup \partial (S_2, V_2) \sqcup (\ell_1
\times \ell_2), \delta_{V_1} \sqcup \delta_{V_2}} V_1 \sqcup V_2$
\end{claim}
\begin{proof}
We need to show that
$$(\partial (S_1, V_1) \sqcup \partial (S_2, V_2)).\get \;
(\delta((\ell_1 \sqcup \ell_2).\pput \; s \; v, s)) = \delta(v, (\ell_1 \sqcup
\ell_2).\get \; s)$$

The result holds inductively if $s \in S_1$ and $v \in V_1$ or $s \in S_2$ and
$v \in V_2$. Otherwise, assume without loss of generality
that $s \in S_1$ and $v \in V_2$. Then
\begin{align*}
(\partial (S_1, V_1) \sqcup \partial (S_2, V_2)).\get \;
(\delta((\ell_1 \sqcup \ell_2).\pput \; s \; v, s)) &=
(\partial (S_1, V_1) \sqcup \partial (S_2, V_2)).\get \; (\delta(\ell_2.\create
\; v, s))\\
&= (\ell_1 \times \ell_2).\get \; (s, \ell_2.\create \; v)\\
&= (\ell_1.\get \; s, \ell_2.\get \; (\ell_2.\create \; v))\\
&= (\ell_1.\get \; s, v)\\
&= \delta(v, \ell_1.\get \; s)\\
&= \delta(v, (\ell_1 \sqcup \ell_2).\get \; s)
\end{align*}
\end{proof}
\subsection{Positional Iteration}
Let $\ell : S \Leftrightarrow V$ be a lens and define the lens $\ell^*:
S^* \Leftrightarrow V^*$ by
\begin{align*}
\ell^* .\get \; (s_1, \ldots, s_n) &= (\ell.\get \; s_1) \cdot \ldots
\cdot (\ell.\get \; s_n)\\
\ell^*.\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n) &= s'_1 \cdot
\ldots \cdot s'_n\\
\text{ where } s'_i &= \begin{cases}
\ell.\pput \; s_i \; v_i & \text{if } 1 \leq i \leq \min\{m, n\}\\
\ell.\create \; v_i & \text{if } \min\{m, n\} < i \leq n
\end{cases}\\
\ell^*.\create \; (v_1, \ldots, v_n) &= (\ell.\create \; v_1) \cdot \ldots
\cdot (\ell.\create \; v_n)
\end{align*}

\begin{claim}\label{positionalisvalid}
If $\ell: S \xLeftrightarrow{\delta_S, \partial (S, V), \delta_V}
V$, then $\ell^*: S^* \xLeftrightarrow{\delta_{S}^*, (\partial (S, V) \sqcup
\ell)^*, \delta_{V}^*} V^*$.
\end{claim}
\begin{proof}
We need to show that
\begin{multline}
(\partial (S, V) \sqcup \ell)^*.\get \;
(\delta_S^*(\ell^*.\pput \; (s_1, \ldots, s_n) \; (v_1, \ldots,
v_m), (s_1, \ldots, s_n))) \\
= \delta_V^*((v_1, \ldots, v_m), \ell^*.\get \; (s_1, \ldots,
s_n))
\end{multline}
Let $1 \leq i \leq \min\{m, n\}$. Observe that $v_i = \ell.\get \; s_i$ if and
only if $\ell.\pput \; s_i \; v_i = s_i$, so $\delta_V^*$ omits
$\delta_V(v_i, \ell.\get \; s_i)$ in position $i$ if and only if
$\delta_S^*$ also omits $\delta_S(\ell.\pput \; s_i \; v_i, s_i)$ in
position $i$; if no omission occurs, then $\delta(S, V).\get \;
(\delta(\ell.\pput \; s_i \; v_i, s_i)) = \delta(v_i, \ell.\get \; s_i)$ since
$(\ell, \partial(S,V))$ is a delta lens. For $\min\{m,n\} < i \leq m$, then
$\delta_V^*$ adds $v_i$ in position $i$, and $\delta_S^*$ adds
$\ell.\create \; v_i$ in position $i$, which gives the desired result since
$$(\partial (S, V) \sqcup \ell).\get \; (\ell.\create \; v_i) = \ell.\get \;
(\ell.\create \; v_i)) = v_i$$
\end{proof}
\subsection{Matching Iteration}
\iffalse
When defining a $\match(\mathcal{M}, \delta)$ difference function, we often wish for
$\mathcal{M}$ to be an optimal set of matchings for some alignment strategy.

For instance, define the set of matchings $\mathcal{M}_{POS}$ by
$$\mathcal{M}_{POS} = \{\varphi_{(a_1, \ldots, a_m), (b_1, \ldots, b_n)}:[m]
\longrightarrow [n] \; | \; m \leq n, \varphi_{(a_1, \ldots, a_m), (b_1,
\ldots, b_n)}(i) = i\}$$ Then $\mathcal{M}_{POS}$ is an optimal set of
matchings for a positional alignment strategy.

For the non-crossing strategy, we want to our matchings to be increasing rather
than positional. That is, given a valid set of matchings $\mathcal{M}$, let
$\mathcal{M}_{a, b} = \varphi_{a, b} \in \mathcal{M}$, and for $a, a' \in P$
with $a=(a_1, \ldots, a_m)$ and $a'=(a'_1, \ldots, a'_n)$, define
$\mathcal{M}^{NC}_{a, b}$ by
$$\mathcal{M}^{NC}_{a, a'} = \argmin_{\substack{\mathcal{M}_{a, a'} \\
\mathcal{M}_{a, a'} \text{ is increasing}}} |\match(\mathcal{M}_{a, a'},
\delta)(a, a')|$$
There may be multiple matchings which for which the size of the delta is
minimal, in which case we pick the lexicographically smallest matching.
Then $\mathcal{M}^{NC}$ is an optimal set of matchings for a non-crossing
alignment strategy.

Finally, for a min-edit distance strategy, we simply remove the condition that
the matchings should be increasing. That is, define
$\mathcal{M}^{MED}_{a, b}$ by
$$\mathcal{M}^{MED}_{a, a'} = \argmin_{\mathcal{M}_{a, a'}}
|\match(\mathcal{M}_{a, a'}, \delta)(a, a')|$$
Then $\mathcal{M}_{MED}$ is an optimal set of matchings for a min-edit distance
alignment strategy.
\fi
Assume that $\ell: S \Leftrightarrow V$. Let $\mathcal{M}_V = \{\varphi_{x, y}
\; | \; x, y \in V^*\}$ be a set of alignments between $x$ and $y$ for each $x,
y \in V^*$. For each $s, s' \in S^*$, let $\varphi_{s, s'} = \varphi_{\ell.\get
\; s, \ell.\get \; s'}$. Define $\mathcal{M}_S$ by $\mathcal{M}_S =
\{\varphi_{s, s'} \; | \; s, s' \in S^*\}$. Define the lens
$\match(\mathcal{M}_V, \ell) : S^* \Leftrightarrow V^*$ by
\begin{align*}
\match(\mathcal{M}_V, \ell) .\get \; (s_1, \ldots, s_n) &= (\ell.\get \; s_1)
\cdot \ldots \cdot (\ell.\get \; s_n)\\
\match(\mathcal{M}_V, \ell) .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n)
&= (s'_1 , \ldots , s'_n)\\
\text{ where } s'_i &= \begin{cases}
\ell.\pput \; s_{\psi(i)} \; v_i & \text{if } i \in Dom(\psi)\\
\ell.\create \; v_i & \text{if } i \not \in Dom(\psi)
\end{cases}\\
\text{ and } \psi &= \varphi_{(v_1, \ldots, v_n), ((\ell.\get \; s_1), \ldots,
(\ell.\get \; s_m))}\\
\match(\mathcal{M}_V, \ell).\create \; (v_1, \ldots, v_n) &= (\ell.\create \;
v_1) \cdot \ldots \cdot (\ell.\create \; v_n)
\end{align*}
Assume that 
\begin{claim}
If $\ell: S \xLeftrightarrow{\delta_{S}, \partial (S, V), \delta_{V}}
V$, then $\match(\mathcal{M}_V, \ell) : S^*
\xLeftrightarrow{\match(\mathcal{M}_S, \delta_{S}), \partial (S, V)^*,
\match(\mathcal{M}_V, \delta_V)} V^*$.
\end{claim}
\begin{proof}
Observe that the matching used to compute the delta on the source side is
the same as the matching used to compute the matching on the view side, hence we
may take the update lens to be position and proceed with the rest of the proof
as in \cref{positionalisvalid}.
\end{proof}
\subsection{Composition}
Let $\ell_1 : S \Leftrightarrow U$ and $\ell_2 : U \Leftrightarrow V$ be lenses.
Define the lens $\ell_2 \circ \ell_1 : S \Leftrightarrow V$ by
\begin{align*}
(\ell_2 \circ \ell_1).\get \; s &= \ell_2.\get \; (\ell_1.\get \; s)\\
(\ell_2 \circ \ell_1).\pput \; s \; v &= \ell_1.\pput \; s \; (\ell_2.\pput \;
(\ell_1.\get \; s) \; v) \\
(\ell_2 \circ \ell_1).\create \; v &= \ell_1.\create \; (\ell_2.\create \; v)
\end{align*}

\begin{claim}
If $\ell_1 : S_1  \xLeftrightarrow{\delta_{S_1}, \partial (S_1, V_1),
\delta_{V_1}} V_1$ and $\ell_2 : S_2 \xLeftrightarrow{\delta_{S_2}, \partial
(S_2, V_2), \delta_{V_2}} V_2$, then \\
$\ell_2 \circ \ell_1 :  S \xLeftrightarrow{\delta_S, \partial
(S_2, V_2) \circ \partial (S_1, V_1), \delta_V} V$.
\end{claim}
\begin{proof}
We need to show that
$$
(\partial (S_2, V_2) \circ \partial (S_1, V_1)).\get \; (\delta(\ell_2 \circ
\ell_1).\pput \; s \; v, s) = \delta(v, (\ell_2 \circ \ell_1).\get \; s)$$

This follows from the fact that
\begin{align*}
&(\partial (S_2, V_2) \circ \partial (S_1, V_1)).\get \; (\delta_S((\ell_2 \circ
\ell_1).\pput \; s \; v, s))\\
&=(\partial (S_2, V_2) \circ \partial (S_1, V_1)).\get \; (\delta_S(\ell_1.\pput
\; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; v), s))\\
&= (\partial (S_2, V_2)).\get \; ((\partial (S_1, V_1)).\get \;
(\delta_S(\ell_1.\pput \; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; v), s)))\\
&= (\partial (S_2, V_2)).\get \; (\delta_U(\ell_2.\pput \; (\ell_1.\get \; s)
\; v, \ell_1.\get \; s))\\
&= \delta_V(v, \ell_2.\get \; (\ell_1.\get \; s))\\
&= \delta_V(v, (\ell_2 \circ \ell_1).\get \; s)
\end{align*}
\end{proof}
\iffalse
\section{Symmetric Delta Lenses}
Assume that $(X, \partial X, \delta_X)$ and $(Y, \partial Y, \delta_Y)$
are difference structures. A symmetric delta lens is a pair $(\ell, \partial
\ell)$ where $\ell : X \Leftrightarrow Y$ is a symmetric lens, and
$\partial \ell : \partial X \Leftrightarrow \partial Y$ is a symmetric lens
such that
\begin{enumerate}
  \item
  if $\ell.putr(x, c) = (y, c)$ and $\ell.putl(y', c) = (x', c')$,
  then $\partial \ell.putl \; (\delta(y, y'), \delta(c, c')) = (\delta(x,
  x'), \delta(c, c'))$
  \item
  if $\ell.putl(y, c) = (x, c)$ and $\ell.putr(x', c) = (y', c')$, then
  $\partial \ell.putr \; (\delta(x, x'), \delta(c, c')) = (\delta(y, y'),
  \delta(c, c'))$
\end{enumerate}
\begin{fact}[Identity Lens]
Given any symmetric lens $\partial X : X \Leftrightarrow X$, $(id_X, \partial
X)$ is a symmetric delta lens
\end{fact}
\begin{fact}[Bijective Lenses]
Assume that $f : X \longrightarrow Y$ and $g : \partial X \longrightarrow
\partial Y$ are bijections. Assume that $g(\delta(x, x')) = \delta(f(x),
f(x'))$. Then $(bij_f, bij_g)$ is a symmetric delta lens.
\end{fact}
\begin{fact}[Lens Composition]
If $(\ell_1, \partial \ell_1) : (X, \partial X, \delta_X) \Leftrightarrow (Y,
\partial Y, \delta_Y)$ and $(\ell_2, \partial \ell_2): (Y, \partial Y,
\delta_Y) \Leftrightarrow (Z, \partial Z, \delta_Z)$ are symmetric delta
lenses, then $(\ell_1;\ell_2, \partial \ell_1;\partial \ell_2)$ is a symmetric
delta lens.
\end{fact}
\begin{fact}[Dual of a Lens]
If $(\ell, \partial \ell)$ is symmetric delta lens, then $(\ell^{op}, \partial
\ell^{op})$ is a symmetric delta lens.
\end{fact}
\begin{fact}[Terminal Lens]
$(term_x, term_{\delta(x, x)})$ is a symmetric delta lens
\end{fact}
\begin{corollary}[Disconnect Lens]
If $x \in X$ and $y \in Y$, then $(term_x;term_y^{op},
term_{\delta(x, x)};term_{\delta(y, y)}^{op})$ is a symmetric delta lens.
\end{corollary}
\begin{fact}[Tensor Product Lens]
If $(\ell_1, \partial \ell_1)$ and $(\ell_2, \partial \ell_2)$ are symmetric
delta lenses, then $(\ell_1 \otimes \ell_2, \partial \ell_1 \otimes \partial
\ell_2)$ is a symmetric delta lens.
\end{fact}
\begin{corollary}[Projection Lens]
Projection lenses are symmetric delta lenses.
\end{corollary}
\begin{fact}[Forgetful Tensor Sum Lens]
If $(\ell_1, \partial \ell_1)$ and $(\ell_2, \partial \ell_2)$ are symmetric
delta lenses, then $(\ell_1 \oplus^f \ell_2, \partial \ell_1 \oplus^f \partial
\ell_2 \oplus (\ell_1 \otimes \ell_2))$ is a symmetric delta lens.
\end{fact}
\begin{fact}[Forgetful List Mapping Lens]
If $(\ell, \partial \ell)$ is a symmetric delta lens, then $(map^f(\ell)
\ell_2, map^f(\partial \ell))$ is a symmetric delta lens.
\end{fact}
\begin{remark}[Retentive Lenses]
The retentive tensor sum and list mapping lenses are in general not symmetric
delta lenses with the obvious definitions of difference functions.
\end{remark}
\fi
\section{Relational Lenses}
We define a {\em relation} $R$ of type $((T_1, \ldots, T_n), P, F)$ to be a
multiset $R$ where $T_i, P$ are sets satisfying $R \subseteq P \subseteq T_1
\times \ldots \times T_n$, $F \subseteq [n] \times [n]$, and if $i, j \in F$,
then for all $r, r' \in R$, if $r_i = r'_i$ then $r_j = r'_j$. We interpret the
$T_i$ to be a set of {\em attributes}, $R$ to be a set of {\em records} with
$dom(r) = (T_1, \ldots, T_n)$ for all $r \in R$, $P$ to be a {\em predicate}
such that the records in $R$ satisfy $P$, and $F$ to be a set of {\em
functional dependencies} on the attributes of the records in $R$.

Let $G(F)$ be the graph induced by $F$ on $T_1, \ldots, T_n$ i.e. there is an
edge from $T_i$ to $T_j$ if and only if $(i, j) \in F$. We say that $F$ is
in {\em tree form} if $G(F)$ is a DAG and every vertex has indegree at most 1;
following \cite{bohannon2006relational},we will always assume that $F$ is in
tree form from here on. Observe that every record $r$ is completely determined
by nodes of indegree 0 in $G(F)$: if $Roots(G(F)) = \{T_{i_1}, \ldots,
T_{i_k}\}$ is the set of nodes of indegree 0 and outdegree greater than 0 in
$G(F)$ and $Isolated(G(F)) = \{T_{j_1}, \ldots, T_{j_\ell}\}$ is the set of
nodes of indegree 0 and outdegree 0 in $G(F)$, then for any $r = (r_1, \ldots,
r_n)\in R$, then the map $((r_{i_1}, \ldots, r_{i_k}), (r_{j_1}, \ldots,
r_{j_\ell}))$ is a bijection since the $r_{i_1}, \ldots, r_{i_k}$ determine all
nodes of indegree greater than 0, which leaves just the nodes $r_{j_1}, \ldots,
r_{j_\ell}$.

In order to define lenses in relations, we first need to describe data
structures representing the relation; lenses will operate on the data
structures rather than on relations. Let $Outputs(F) = G(F) \setminus
(Roots(G(F)) \cup Isolated(G(F)))$. We choose to represent a relation with the
following data:
\begin{enumerate}
  \item
  A partial map $data : (T_{i_1} \times \ldots \times T_{i_k})
  \longrightarrow 2^{(T_{j_1} \times \ldots \times T_{j_{\ell}})}$ such that
  there exists a record $r \in R$ with $r[T_{i_1}, \ldots, T_{i_k}] =
  (t_{i_1}, \ldots, t_{i_k})$ if and only if
  $$data(t_{i_1}, \ldots, t_{i_k}) =  \{\{r[T_{j_1}, \ldots, T_{j_{\ell}}] \in
  R \; | \; r_{i_m} = t_{i_m} \text{ for }1 \leq m \leq k\}\}$$
  \item
  A map $dependencies : outputs(F) \longrightarrow \cup_{i,j} 2^{T_i \times
  T_j}$ such that $(p, q) \in dependencies(j)$ if and only if there exists a
  record $r \in R$ such that $r[T_i, T_j] = (p, q)$, where $i$ is such that
  $T_i \longrightarrow T_j \in F$ ($i$ is unique as $F$ is assumed to be in tree
  form).
\end{enumerate}
\subsection{Select}
We say that a predicate $P$ ignores a set of attributes $X$ if for all $r, r'
\in R$, if $r[dom(R) \setminus X] = r'[dom(R) \setminus X]$ implies that $r
\in P$ if and only if $r' \in P$.

The {\em select} lens on $R$ evaluates to a relation which is the same as $R$
but which satisfies a predicate $P$. The inferrence rule for deriving the select
lens is the following:
\begin{prooftree}
\AxiomC{$R : ((T_1, \ldots, T_2), Q, F)$}
\AxiomC{$S : ((T_1, \ldots, T_2), P \cap Q, F)$}
\AxiomC{$F$ is in tree form}
\AxiomC{$P$ ignores $Outputs(F)$}
\QuaternaryInfC{$(select \; from \; R\; where \; P \; as \; S) : R
\Leftrightarrow S$}
\end{prooftree}
In the get direction, the select lens simply restricts $R$ to $P$. Now consider
$put \; s \; r$. The select lens proceeds as follows:
\begin{enumerate}
  \item
  First, it computes $(data_s, dependencies_s)$ and $(data_r, dependencies_r)$.
  \item
  Next, it updates $dependencies_r$ using $dependencies_s$ to get
  $dependencies_r'$.
  \item
  Next it filters $data_r$ to contain only records not in $P$ to get $data_r'$.
  \item
  Next, it computes $(data_r' \cup data_s, dependencies_r')$.
  \item
  Finally, it deletes the records from $data_r' \cup data_s$ that would cause
  PUT-GET to fail.
\end{enumerate}
\section{Matching Lenses}
\begin{definition}
A {\em basic lens} $\ell : S \xLeftrightarrow{C} V$ is a 4-tuple of functions
$\ell.\get : S \longrightarrow V$, $\ell.\res : S \longrightarrow C$,
$\ell.\pput : V \longrightarrow C \longrightarrow S$ and $\ell.\create : V
\longrightarrow S$ satisfying the following properties:
\begin{align*}
\ell.\get \; (\ell.\pput \; v \; c) &= v \tag{PUTGET}\\
\ell.\pput \; (\ell.\get \; s) \; (\ell.\res \; s) &= s \tag{GETPUT}
\end{align*}
\end{definition}
Idea: First remove matched parts to leave unmatched skeleton. Next remove
unmatched parts to leave matched skeleton and then flatten.
%% Acknowledgments
\begin{acks}                            %% acks environment is optional
%% contents suppressed with 'anonymous'
%% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
%% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
%% acknowledge financial support and will be used by metadata
%% extraction tools.
This material is based upon work supported by the
\grantsponsor{GS100000001}{National Science
Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
No.~\grantnum{GS100000001}{nnnnnnn} and Grant
No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
conclusions or recommendations expressed in this material are those
of the author and do not necessarily reflect the views of the
National Science Foundation.
\end{acks}
\fi
\bibliographystyle{plain}
\bibliography{local}
%% Appendix
%%\appendix
%%\section{Appendix}

\end{document}
\newif\ifdraft\drafttrue  % set true to show comments
\newif\ifplentyoftime\plentyoftimefalse  % :-)

% % For double-blind review submission, w/o CCS and ACM Reference (max
% submission space)
\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
% % For double-blind review submission, w/ CCS and ACM Reference
% \documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
% % For single-blind review submission, w/o CCS and ACM Reference (max
% submission space)
% \documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
% % For single-blind review submission, w/ CCS and ACM Reference
% \documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true} % For
% final camera-ready submission, w/ required CCS and ACM Reference
% \documentclass[acmsmall]{acmart}\settopmatter{}

% % Journal information % Supplied to authors by publisher for camera-ready
% submission; % use defaults for review submission.
\acmJournal{PACMPL}
\acmVolume{1}
\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
\acmArticle{1}
\acmYear{2018}
\acmMonth{1}
\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
\startPage{1}

% % Copyright information % Supplied to authors (based on authors' rights
% management selection; % see authors.acm.org) by publisher for camera-ready
% submission; % use 'none' for review submission.
\setcopyright{none}
% \setcopyright{acmcopyright} \setcopyright{acmlicensed}
% \setcopyright{rightsretained} \ccopyrightyear{2018}           %% If different
% from \acmYear

% % Bibliography style %\bibliographystyle{ACM-Reference-Format} % Citation
% style % Note: author/year citations are required for papers published as an %
% issue of PACMPL.
% %\citestyle{acmauthoryear}   %% For author/year citations

% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% % Note:
% Authors migrating a paper from PACMPL format to traditional % SIGPLAN
% proceedings format must update the '\documentclass' and % topmatter commands
% above; see 'acmart-sigplanproc-template.tex'.
% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{amsmath, amssymb, amsthm, enumerate, array, extarrows, mathrsfs,
mathtools, stmaryrd, listings, xspace, bussproofs}
\theoremstyle{definition}
\usepackage[capitalize]{cleveref}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\newtheorem{conjecture}{Conjecture}
\newtheorem{question}{Question}
\newtheorem{fact}{Fact}
\newtheorem{claim}{Claim}
\newtheorem{remark}{Remark}

% Macros Colors
\definecolor{dkblue}{rgb}{0,0.1,0.5}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{dkred}{rgb}{0.6,0,0}
\definecolor{dkpurple}{rgb}{0.7,0,0.4}
\definecolor{olive}{rgb}{0.4, 0.4, 0.0}
\definecolor{teal}{rgb}{0.0,0.5,0.5}
\definecolor{orange}{rgb}{0.9,0.6,0.2}
\definecolor{lightyellow}{RGB}{255, 255, 179}
\definecolor{lightgreen}{RGB}{170, 255, 220}
\definecolor{teal}{RGB}{141,211,199}
\definecolor{darkbrown}{RGB}{121,37,0}

\newcommand{\FINISH}[3]{\ifdraft\textcolor{#1}{[#2: #3]}\fi}
\newcommand{\bcp}[1]{\FINISH{dkred}{B}{#1}}
\newcommand{\BCP}[1]{\FINISH{dkred}{B}{\bf #1}}
\newcommand{\afm}[1]{\FINISH{dkgreen}{A}{#1}}
\newcommand{\dpw}[1]{\FINISH{dkblue}{D}{#1}} % Toronto Maple Leafs Blue :-)
\newcommand{\saz}[1]{\FINISH{orange}{SZ}{#1}}
\newcommand{\ksf}[1]{\FINISH{teal}{K}{#1}}
\newcommand{\sam}[1]{\FINISH{dkpurple}{SM}{#1}}

\newcommand{\kw}[1]{\ensuremath{\mathsf{#1}}\xspace}
\newcommand{\get}{\ensuremath{\kw{get}}\xspace}
\newcommand{\pput}{\ensuremath{\kw{put}}\xspace}
\newcommand{\create}{\ensuremath{\kw{create}}\xspace}
\newcommand{\res}{\ensuremath{\kw{res}}\xspace}
\newcommand{\ccopy}{\ensuremath{\kw{copy}}\xspace}
\newcommand{\const}{\ensuremath{\kw{const}}\xspace}
\newcommand{\default}{\ensuremath{\kw{default}}\xspace}
\newcommand{\length}{\ensuremath{\kw{length}}\xspace}
\newcommand{\match}{\ensuremath{\kw{match}}\xspace}

\DeclareMathOperator*{\argmin}{arg\,min} % thin space, limits underneath in
\DeclareMathOperator*{\argmax}{argmax} % no space, limits underneath in

\lstset{ basicstyle=\ttfamily, escapeinside=|| }
\begin{document}

% % Title information
\title{Lenses with a Difference}         %% [Short Title] is optional;
% % when present, will be used in % header instead of Full Title.
% %\titlenote{with title note}             %% \titlenote is optional; % can be
% repeated if necessary; % contents suppressed with 'anonymous'
% %\subtitle{Subtitle}                     %% \subtitle is optional
% %\subtitlenote{with subtitle note}       %% \subtitlenote is optional; % can
% be repeated if necessary; % contents suppressed with 'anonymous'

% % Author information % Contents and number of authors suppressed with
% 'anonymous'.
% % Each author should be introduced by \author, followed by % \authornote
% (optional), \orcid (optional), \affiliation, and % \email.
% % An author may have multiple affiliations and/or emails; repeat the %
% appropriate command.
% % Many elements are not rendered, but should be provided for metadata %
% extraction tools.

% % Author with single affiliation.
% \author{First1 Last1} \authornote{with author1 note}          %% \authornote
% is optional; % can be repeated if necessary \orcid{nnnn-nnnn-nnnn-nnnn} %%
% \orcid is optional \affiliation{ \positition{Position1}
% \department{Department1} %% \department is recommended
% \institution{Institution1} %% \institution is required \streetaddress{Street1
% Address1} \city{City1} \state{State1} \posittcode{Post-Code1}
% \country{Country1}                    %% \country is recommended }
% \email{first1.last1@inst1.edu}          %% \email is recommended

\author{Solomon Maina}
\position{PhD Student}
\department{Computer Science}              %% \department is recommended
\institution{University of Pennsylvania}            %% \institution is required
\country{USA}                    %% \country is recommended }
\email{smaina@seas.upenn.edu}          %% \email is recommended

% % Abstract % Note: \begin{abstract}...\end{abstract} environment must come %
% before \maketitle command
\begin{abstract}
Many of the bidirectional (bx) programming languages that exist today focus on
restoring consistency of models. Consequently, these languages often provide two
guarantees: (1) that a transformation restores consistency when an
update is made to one of the models, and (2) that models that are already
consistent are not updated by the transformation. Unfortunately, in situations
where there are many different ways of restoring consistency, these languages
do not constrain how a lens chooses to do so. As a result, the bx
transformations derived using these languages can lead to unexpected results. In
this paper we formalize a condition called {\em $\partial D$ preservation} that
describes whether a bx transformation propagates updates correctly with respect
to the relation $\partial D$, and then use this condition to characterize the
guarantees that various lens programming languages make with regard to update
propagation.
\end{abstract}

%% Keywords
%% comma separated list
\keywords{bidirectional programming, lenses}  %% \keywords
% are mandatory in final camera-ready submission

%% \maketitle
%% Note: \maketitle command must come after title commands, author
%% commands, abstract environment, Computing Classification System
%% environment and commands, and keywords command.
\maketitle

\section{Introduction}
By now there exist many programming languages that enable one to define {\em
bidirectional transformations}, programs that can be run in two
directions. Each of these languages provides guarantees on how the two
transformations interact e.g. that they are bijective, that a round-trip on
one or both directions does not cause non-trivial updates etc. Many of these
guarantees have to do with restoring or preserving consistency. To formalize
this statement, we describe a widely used formulation of bidirectional
transformations due to Stevens \cite{stevens2010bidirectional} but originally
articulated in a more general form by Meertens \cite{meertens1998designing}:
\begin{definition}
A bidirectional transformation (bx) $R: M \Leftrightarrow N$ comprises of a
consistency relation $R \subseteq M \times N$ and restorers $\overrightarrow{R}
: M \times N \rightarrow N$ and $\overleftarrow{R} : M \times N \longrightarrow
N$.
\end{definition}
\begin{definition}
A bx $R : M \Leftrightarrow N$ is {\em correct} if the consistency restorers do
restore consistency: that is, for all $m \in M, n \in N$, we have $R(m,
\overrightarrow{R}(m, n))$, and dually. $R$ is {\em hippocratic} if the
consistency restorers make no change to an already-consistent pair of models:
that is, for all $m \in M$, $n \in N$, if $R(m, n)$ then $\overrightarrow{R}(m,
n) = n$ and dually. $R$ is {\em well behaved} if it is correct and hippocratic.
\end{definition}

Many bidirectional programming languages thus enable one to define
well behaved bx's: consistency restoration is guaranteed by correctness of the
bx while consistency preservation is guaranteed by the hippocraticness of the
bx. This paper is primarily concerned with well behaved bx's where
$\overrightarrow{R}$, the \get function of the bx, ignores it's $N$-argument and
where the consistency relation $R$ is defined by $R(m, \overrightarrow{R}(m))$
for all $m \in M$--- these transformations are the well known and widely
understood {\em lenses} of Foster et al \cite{foster2007combinators}--- though
the ideas we develop could be used in other bidirectional settings as well.

When the consistency relation of a bx $R$ is very fine, i.e. when each $m \in M$
is consistent with only a few $n \in N$ and vice versa, then a well behaved lens
gives fairly strong guarantees on $\overleftarrow{R}$ and $\overrightarrow{R}$.
However, there are many cases that arise in practice where there are many
different ways to restore consistency; when this happens, well-behavedness may
be too weak to rule out ``bad'' behaviour or to clarify the behaviour of a bx in
subtle cases. Various solutions have been proposed to address this issue, for
example {\em matching lenses} that deal with the issue of alignment
\cite{barbosa2010matching}, {\em edit lenses} \cite{hofmann2012edit} and {\em
delta lenses} \cite{diskin2011asymmetric,diskin2011state,pacheco2012delta}
that explicitly propagate edits through the lens, and {\em least change
lenses} that require that the source value resulting from a \pput application
is not only acceptable, but also one of closest to the original $s$ among the
sources that share the same view $v$ \cite{macedo2013composing}.

This paper takes a different approach to specifying update propagation by
identifying a condition called {\em $\partial D$ preservation} which we describe
below:
\begin{definition}
Let $R : M \Leftrightarrow N$ be a bidirectional transformation and $\partial D
\subseteq (M \times M) \times (N \times N)$. Then $R$ {\em preserves $\partial
D$} if for all $m \in M$, $n, n' \in N$ , if $R(m, n)$ then $\partial D((m,
\overleftarrow{R}(m, n')), (\overrightarrow{R}(m, n), n'))$. $R$ {\em
strongly preserves $\partial D$} if the dual condition also holds.
\end{definition}
\noindent The idea behind $\partial D$ preservation is that if $m$ and $n$ are
already consistent and $\overrightarrow{R}(m,n)$ is updated to produce $n'$,
then the difference between $\overrightarrow{R}(m,n)$ and $n'$ is related
to the difference between $m$ and $\overleftarrow{R}(m,n')$. $R$
strongly preserves $\partial D$ if it preserves $\partial D$ in both directions.

For lenses, $\partial D$ preservation says that difference between $\ell.\get \;
s$ and $v$ is related to the difference between $s$ and $\ell.\pput \; s \;
v$. One way to formalize this idea is to model the difference between two sources $s, s'
\in S$ and two views $v, v' \in V$ as functions $\delta_S : S \times S
\longrightarrow \partial S$ and $\delta_V : V \times V \longrightarrow \partial
V$ for some sets $\partial S$ and $\partial V$, and to capture the
relationship between $\delta_S$ and $\delta_V$ by a lens. In
particular, given a lens $\partial(S, V) : \partial S \Leftrightarrow \partial
V$ we will ask if the following equation holds of $\ell$:
\begin{equation}\label{difflenslaw}
\partial(S, V).\get \; (\delta_S(s, \ell.\pput \; s \; v)) = \delta_V(\ell.\get
\; s, v)
\end{equation}
If $\ell : S \Leftrightarrow V$ satisfies \cref{difflenslaw} we say that 
$\ell$ is a {\em difference  lens} with respect to $(\delta_S, 
\partial(S, V), \delta_V)$, which we denote with the typing relation 
$\ell : S \xLeftrightarrow{\delta_S, \partial(S, V), \delta_V} V$.

For example, consider two versions of an iteration lens $\ell_1 : S^*
\Leftrightarrow V^*$ and $\ell_2 : S^* \Leftrightarrow V^*$ defined inductively
from a lens $\ell : S \Leftrightarrow V$. Both of these lens have the same get
behaviour: given a list $s_1, \ldots, s_n$ of sources, they both produce
$\ell.\get \; s_1, \ldots, \ell.\get \; s_n$. The \pput behaviour of the two
lenses is different though: $\ell_1$ simply aligns views to sources
positionally, while $\ell_2$ computes the optimal alignment on the view side and
uses this alignment to put views back into sources.

Let \lstinline|name = [A-Z] . [a-z]*|, let the source $S$ equal 
\lstinline|name . " " . name| and the view $V$ equal \lstinline|name|.
Let $\ell : S \Leftrightarrow V$ be the lens
\lstinline|del name . del " " . copy name| which given a pair of names produces
the second name in the \get direction, and restores the first name in the \pput
direction:
$$\ell.\get \; ``Ada \; Turing" = ``Turing" \text{ and }\; \ell.\pput \; ``Ada
\; Turing" \; ``Lovelace" = ``Ada \; Lovelace"$$
Define the
function $\delta_V : V \times V \longrightarrow \mathbb{N}^{\geq 0}$ by 
$$\delta_V(n, n') = 
\begin{cases}
0 & \text{if }n=n'\\
1 & \text{otherwise}
\end{cases}$$
Define $\delta_S : S \times S \longrightarrow \mathbb{N}^{\geq 0}$ by
$$\delta_S(m \cdot `` \quad " \cdot n, m' \cdot `` \quad " \cdot n') =
\begin{cases}
0 & \text{if } m = m', n = n'\\
1 & \text{if } m = m', n \neq n'\\
1 & \text{if } m \neq m', n = n'\\
2 & \text{if } m \neq m', n \neq n'
\end{cases}$$
Let the lens $\partial(S, V) : \mathbb{N}^{\geq 0} \Leftrightarrow
\mathbb{N}^{\geq 0}$ be given by $\partial(S, V) = \ccopy(\mathbb{N}^{\geq
0})$. Then $\ell : S \xLeftrightarrow{\delta_S, \partial(S, V), \delta_V} V$.

Now let $xs, ys$ be two lists of single names, with $length(xs) \geq length(ys)$
Define $\delta_V^* : V^* \times V^* \longrightarrow \mathbb{Z}^* \times V^*$ on $xs$ and
$ys$ by aligning elements of $xs$ maximally with elements of $ys$ and then using $\delta_V$
on these pairs to produce a list $ds$ of numbers. In this process, some elements of
$ys$ will not be aligned if $length(xs) < length(ys)$, in which case we simply collect
these in another list of elements of $V$. For instance,
$$\delta_V^*([``Lovelace", ``Kleene", ``Turing"], [``Turing", ``Lovelace"]) = ([0,0], [``Kleene"])$$
since we align \lstinline|"Lovelace"| to \lstinline|"Lovelace"| and 
\lstinline|"Turing"| to \lstinline|"Turing"| for a difference of 0 in each case, which leaves
\lstinline|"Kleene"| unaligned.

On the source side, let $ss, ts$ be two lists of two names with $length(ss) \geq length(ts)$.
Define $\delta_{S, \ell}^* : S^* \times S^* \longrightarrow \mathbb{Z}^* \times S^*$ by taking the 
alignment used to compute $\delta_V^*(\ell^*.\get \; ss, \ell^*.\get \; ts)$ and then using that
alignment in the same manner that $\delta_V^*$ does. For instance
\begin{align*}
&\delta_{S, \ell}^*([``Ada \; Lovelace", ``Stephen \; Kleene", ``Alan \; Turing"], 
[``Alan \; Turing", ``Ada \; Lovelace"])\\
&= ([0,0], [``Stephen \; Kleene"])
\end{align*}
Recall that the \pput behaviour of $\ell_1$ simply aligns views to sources
positionally, while $\ell_2$ computes the optimal alignment on the view side and
uses this alignment to put views back into sources. From this we can prove that
$$\ell_2 : S^* \xLeftrightarrow{\delta_{S, \ell}^*, \partial(S, V)^* \times \ell^*, \delta_V^*} V^*$$
On the other hand, the judgement $\ell_1 : S^* \xLeftrightarrow{\delta_{S, \ell}^*, 
\partial(S, V)^* \times \ell^*, \delta_V^*} V^*$ does not holds since $\ell_1$ does not satisfy \cref{difflenslaw} :
$$(\partial(S, V)^* \times \ell^*).\get \; (\delta_{S, \ell}^*(s, \ell.\pput \; s \; v)) = \delta_V^*(\ell.\get
\; s, v)
$$
This is witnessed by the fact that if $s = [``Ada \; Lovelace", ``Stephen \; Kleene", ``Alan \; Turing"]$ and \\
$v = [``Kleene", ``Turing", ``Lovelace"]$, then 
\begin{align*}
&\ell_1.\pput \; [``Ada \; Lovelace", ``Stephen \; Kleene", ``Alan \; Turing"] \; [``Kleene", ``Turing", ``Lovelace"]\\
&= [``Ada \; Kleene", ``Stephen \; Turing", ``Alan \; Lovelace"]
\end{align*}
Also, $\delta_{S, \ell}^*(s, \ell.\pput \; s \; v) = ([1,1,1], [])$, $\delta_V^*(\ell.\get
\; s, v) = ([0,0,0], [])$, but
\begin{align*}
(\partial(S, V)^* \times \ell^*).\get \; (\delta_{S, \ell}^*(s, \ell.\pput \; s \; v)) &=
(\partial(S, V)^* \times \ell^*).\get \; ([1,1,1], [])\\
&= ([1,1,1], [])\\
&\neq ([0,0,0], [])\\
&= \delta_V^*(\ell.\get \; s, v)
\end{align*}
Here the positional lens $\ell_1$ is not a difference lens with respect to $(\delta_{S, \ell}^*, 
\partial(S, V)^* \times \ell^*, \delta_V^*)$ since it does not align sources to views in the same way that
$\delta_V^*$ does. On the other hand, $\ell_2$ does, hence is a difference lens with respect to $(\delta_{S, \ell}^*, 
\partial(S, V)^* \times \ell^*, \delta_V^*)$.

In the rest of this paper, we give concrete examples of how to use $\partial D$ preservation to characterize 
the guarantees on update propagation provided by various lens programming languages. In particular,
the major technical contributions that we will make in this paper are the following:
\begin{enumerate}
  \item
  We formalize a condition called $\partial D$ preservation which provides a way
  of constraining the updates that can be made by a bidirectional
  transformation. We give examples of relations $\partial D$ as well as
  examples and counterexamples of lenses that are $\partial D$ preserving.
  \item
  We describe an instance of $\partial D$ preservation that is satisfied by {\em difference lenses}. We give a set of combinators for {\em difference lenses} and describe the update propagation guarantees that the lenses derived using these combinators give.
  \item
  Finally, we show how to represent matching lenses and relational lenses
  \cite{bohannon2006relational} as difference lenses, and hence
  describe non-trivial guarantees beyond correctness and hippocraticness that
  these lens languages provide as well.
\end{enumerate}
\iffalse
A {\em lens} is a program that can be run in two directions. In the
forward direction, the \get component of a lens produces a {\em view}
from a {\em source}, while in the backward direction, the \pput component folds
an updated view back into the original source. Formally, Foster
\cite{foster2007combinators} defined a lens $\ell : S
\Leftrightarrow V$ from a source $S$ to a view $V$ to be a pair of functions
$\ell.\get : S \longrightarrow V$ and $\ell.\pput : S \longrightarrow V
\longrightarrow S$ satisfying the
following properties:
\begin{align*}
\ell.\get \; (\ell.\pput \; s \; v) &= v \tag{PUTGET}\\
\ell.\pput \; s \; (\ell.\get \; s) &= s \tag{GETPUT}
\end{align*}
The PUTGET law ensures that updates to the view are translated
``exactly''---i.e. that, given a view, the \pput function produces a source
that \get maps back to the very same view, while the GETPUT law ensures a
``stability'' property for the source---i.e. it requires that the \pput
function returns the original source unmodified whenever the update to the view
is a no-op.

Unfortunately, while these laws give some guarantees on how the \get and \pput
components of a lens interact, they do not specify the behaviour of the \pput
component when the update to the view is non-trivial. Consequently, a
non-trivial update to a view could lead to an unexpected update to the source.
Various solutions have been proposed to address this issue, for example {\em
matching lenses} to deal with the issue of alignment \cite{barbosa2010matching},
{\em edit lenses} \cite{hofmann2012edit} and {\em delta lenses}
\cite{diskin2011asymmetric,diskin2011state,pacheco2012delta} which explicitly
propagate edits through the lens, and {\em least change lenses} which
add a new LEASTPUTGET law which requires that the source value resulting from a
\pput application is not only acceptable, but also one of closest to the
original $s$ among the sources that share the same view $v$
\cite{macedo2013composing}.

This paper describes a general method for specifying what updates are allowable
for a given lens without requiring that the lens explicitly propagate updates.
Our approach is guided by the simple idea the idea that if one starts with a
source $s$, calls the \get function on $s$ to acquire the view $\ell.\get \;
s$, updates $\ell.\get \; s$ to get a new view $v$, then the ``difference''
between $v$ and $\ell.\get \; s$ and $v$ should be related to the difference
between $\ell.\pput \; s \; v$ and $v$.

To formalize this idea, we first recall a widely used definition of
bidirectional transformations
\cite{meertens1998designing,stevens2010bidirectional} definition of a
bidirectional transformations:
\begin{definition}
A bidirectional transformation (bx) $R: M \Leftrightarrow N$ comprises of a
consistency relation $R \subseteq M \times N$ and restorers $\overrightarrow{R}
: M \times N \rightarrow N$ and $\overleftarrow{R} : M \times N \longrightarrow
N$.
\end{definition}
\begin{definition}
A bx $R : M \leftrightarrow N$ is {\em correct} if the consistency restorers do
restore consistency: that is, for all $m \in M, n \in N$, we have $R(m,
\overrightarrow{R}(m, n))$, and dually. $R$ is {\em hippocratic} if the
consistency restores make no change to and already-consistent pair of models:
that is, for all $m \in M$, $n \in N$, if $R(m, n)$ then $\overrightarrow{R}(m,
n) = n$ and dually. $R$ is {\em well behaved} if it is correct and hippocratic.
\end{definition}

Observe that $\ell : S \Leftrightarrow V$ is a lens, let $R$ be the
consistency relation given by $R(s, \ell.\get \; s)$ for all $s \in
S$, $\overrightarrow{\ell}(s, v) = \ell.\get \; s$, and $\overleftarrow{\ell} =
\ell.\pput$. By the PUTGET law, $\ell.\get \; (\ell.\pput \; s \; v) = v$,
hence $\ell$ is correct, and by the GETPUT laws, $\ell.\pput \; s \; (\ell.\get
\; s) = s$ and $\ell.\get \; (\ell.\create \; v) = v$, hence $\ell$ is
hippocractic. Therefore $\ell$ is a well-behaved bx.

If $R$ is a well-behaved bx then $R$ restores consistency when a source or view
is updated (correctness) and preserves consistency when the source and view are
already consistent (hippocraticness). In order to express

We will then apply this method to define a typing judgement for lenses which
constrains how a lens propagates updates. We will describe a compositional
language of {\em difference lenses} which satisfy this new typing judgement,
and give examples of ordinary lenses that are not difference lenses. Finally,
we will show how to represent matching lenses and relational lenses
\cite{bohannon2006relational} as difference lenses, and hence describe the
guarantees that these lens languages provide.

For example suppose that the source is an XML document representing the
names, dates and nationalities of a collection of classical music
composers such as,

\begin{lstlisting}
<composers>
<composer>
<name>Jean Sibelius</name>
<dates>1865-1956</dates>
<nationality>Finnish</nationality>
</composer>
<composer>
<name>Aaron Copland</name>
<dates>1910-1990</dates>
<nationality>American</nationality>
</composer>
<composer>
<name>Benjamin Briten</name>
<dates>1913-1976</dates>
<nationality>English</nationality>
</composer>
</composers>
\end{lstlisting}

while the view is a list if ASCII text representing the names and dates of each
composer:
\begin{lstlisting}
Jean Sibelius, 1865-1956
Aaron Copland, 1910-1990
Benjamin Briten, 1913-1976
\end{lstlisting}

After computing the initial view, we might want to edit it in some way---e.g.,
correcting the error in Sibelius's death date and the misspelling in Britten's
name
\begin{lstlisting}
Jean Sibelius, 1865-|\colorbox{orange}{1957}|
Aaron Copland, 1910-1990
Benjamin |\colorbox{orange}{Britten}|, 1913-1976
\end{lstlisting}
and push the changes back into the original XML format:
\begin{lstlisting}
<composers>
<composer>
<name>Jean Sibelius</name>
<dates>1865-|\colorbox{orange}{1957}|</dates>
<nationality>Finnish</nationality>
</composer>
<composer>
<name>Aaron Copland</name>
<dates>1910-1990</dates>
<nationality>American</nationality>
</composer>
<composer>
<name>Benjamin |\colorbox{orange}{Britten}|</name>
<dates>1913-1976</dates>
<nationality>English</nationality>
</composer>
</composers>
\end{lstlisting}

Lenses are designed to perform these kinds of bidirectional transformations. If
$\ell$ is a suitably defined lens for the transformation that we
have just described, then the \get component of the lens performs the
XML to ASCII transformation that produces an ASCII view from an XML source,
while the \pput component of the lens performs the ASCII to XML transformation
that folds back an updated ASCII view back into the original XML source.

For example, suppose that a source $S$ consists of a pair consisting of a
boolean value and an integer, and that the view $V$ for $S$ consists of a
single integer. Further, suppose that we wish to define a \pput function for a
lens $\ell$ whose \get component is the function which given a pair $(b, n)$
simply produces the boolean value $b$. Let $f : \mathbb{B} \times \mathbb{Z}
\longrightarrow \mathbb{Z}$ be any function from $\mathbb{B} \times \mathbb{Z}
\longrightarrow \mathbb{Z}$. Then the lens $\ell_{f}$ defined by
\begin{align*}
\ell_{f}.\get \; (b, n) &= b\\
\ell_{f}.\pput \; (b, n) \; b' &= \begin{cases}
(b', n) & \text{if }b = b'\\
(b', f(b', n)) & \text{otherwise}\\
\end{cases}\\
\ell.\create \; b &= (b,f(b, 0))
\end{align*}
is a valid lens.

A user who applies a lens $\ell$ of type $\ell : S \Leftrightarrow V$ may expect
that $\ell.\pput \; (true, 1) \; false = (false, 0)$, reasoning that the
integer produced by the \pput function should be the number representation of
the boolean value $false$. Or perhaps she might expect that $\ell.\pput \;
(true, 1) \; false = (false, 1)$, reasoning that the two pieces of data in a
pair are independent and that the reasonable \pput behaviour in this case is to
restore the orginial integer. In either case, the user may be shocked to find
that in fact $\ell.\pput (true, 0) \; false = (false, 48)$! Given the \get
behaviour which just produces the boolean value from a pair, since $\ell_f$ is a
valid lens for any $f$, then the \pput function of a valid lens of type
$\mathbb{B} \times \mathbb{Z} \Leftrightarrow \mathbb{Z}$ can restore the
integer in a pair in any way so long as the old boolean value and the updated
boolean value are different.

In order to give a more constrained specification for the desired
lens we define the difference $\delta_S((a, b), (c, d))$ between $(a, b)$ and
$(c, d)$ by $$\delta_S((a, b), (c,
d)) = \begin{cases}
0 & \text{if }a = c, b = d\\
1 & \text{if }a=c, b \neq d\\
1 & \text{if }a \neq c, b = d\\
2 & \text{if }a \neq c, b \neq d
\end{cases}$$
On the view side, we define the difference $\delta_V(a, b)$
between $a$ and $b$ by, $$\delta_V(a, b) = \begin{cases}
0 & \text{if } a = b\\
1 & \text{if } a \neq b
\end{cases}$$

Next we define a lens $\partial(S, V) : \partial S \Leftrightarrow \partial V$
which describes how differences should be propagated. For this example, we
define $\partial(S,V)$ to simply be the \ccopy lens from $\partial S$ to
$\partial V$, keeping in mind that $\partial S = \partial V = \mathbb{Z}$.
\cref{difflenslaw} enables us to rule out $\ell_{bad}$, since
\begin{align*}
(\partial (S, V)).\get \; (\delta_S(\ell_{bad}.\pput \; (1,2) \; 3, (1,2))) &=
(\partial (S, V)).\get \; (\delta_S((0,3), (1,2)))\\
&= (\partial (S, V)).\get \; 2\\
&= 2\\
&\neq 1\\
&= \delta_V(3, 2)\\
&= \delta_V(3, \ell_{bad}.\get \; (1, 2))
\end{align*}
Consequently $\ell_{bad}$ is not a valid delta lens. On the
other hand, the lens $\ell_{good} : S \Leftrightarrow V$ defined by
\begin{align*}
\ell_{good}.\get \; (a, b) &= b\\
\ell_{good}.\pput \; (a, b) \; c &= (a, c)\\
\ell_{good}.\create \; b &= (0, b)
\end{align*}
is indeed a valid delta lens since the its \pput component always restores the
first factor.

\section{Strengthening Lens Laws}
\begin{definition}
A bidirectional transformation (bx) $R: M \Leftrightarrow N$ comprises of a
consistency relation $R \subseteq M \times N$ and restorers $\overrightarrow{R}
: M \times N \rightarrow N$ and $\overleftarrow{R} : M \times N \longrightarrow
N$.
\end{definition}
\begin{definition}
A bx is {\em correct} if the consistency restorers do restore consistency:
that is, for all $m \in M, n \in N$, we have $R(m, \overrightarrow{R}(m, n))$,
and dually.
\end{definition}
\begin{definition}
A bx is {\em hippocratic} if the consistency restores make no change
to and already-consistent pair of models: that is, for all $m \in M$, $n \in N$,
if $R(m, n)$ then $\overrightarrow{R}(m, n) = n$ and dually.
\end{definition}
\begin{definition}
A bx is {\em well behaved} if it is correct and hippocratic.
\end{definition}
For lenses, if $\ell : S \Leftrightarrow V$ is a lens, we take the
consistency relation $R$ to be given by $R(s, \ell.\get \; s)$ for all $s \in
S$, $\overrightarrow{\ell}(s, v) = \ell.\get \; s$, and $\overleftarrow{\ell} =
\ell.\pput$. By the PUTGET law, $\ell.\get \; (\ell.\pput \; s \; v) = v$,
hence $\ell$ is correct, and by the GETPUT laws, $\ell.\pput \; s \; (\ell.\get
\; s) = s$ and $\ell.\get \; (\ell.\create \; v) = v$, hence $\ell$ is
hippocractic. Therefore $\ell$ is a well-behaved bx.

Using the notation of Stevens \cite{stevens2010bidirectional},

The major shortcoming of this model of bx transformations is that it does not
take into account how changes should be propagated by transformation; a
well-behaved bx need only restore consistency when models go out of sync or
maintain consisteny if models are already in sync. This becomes an issue when
there are many different ways of restoring consistency

\section{Motivation and Related Work}
Foster et. al. \cite{foster2007combinators,foster2009bidirectional}
defined a lens $\ell : S \Leftrightarrow V$ from a source $S$ to a view $V$ to
be a triple of functions $\ell.\get : S \longrightarrow V$, $\ell.\pput : S
\longrightarrow S \longrightarrow V$ and $\ell.\create : V \longrightarrow S$
satisfying the following properties:
\begin{align*}
\ell.\get \; (\ell.\pput \; s \; v) &= v \tag{PUTGET}\\
\ell.\get \; (\ell.\create \; v) &= v \tag{CREATEGET}\\
\ell.\pput \; s \; (\ell.\get \; s) &= s \tag{GETPUT}
\end{align*}
While these two laws give some guarantees on how the \get, \pput and \create
components of a lens interact, they do not specify the behaviour of the \pput
component when the update to the view is non-trivial, because the GETPUT law
only says that the \pput component must return the original source unmodified
whenever the update to the view is a no-op. Consequently, a non-trivial update
to a view could lead to an undesired update to the source.

Therefore, several lens formalism have been proposed that explicitly have a
notion of an {\em update}. The first of these was Barbosa et. al.'s {\em
matching lenses}\cite{barbosa2010matching}. Matching lenses specifically deal
with the issue of {\em alignment}, since Foster et. al.'s original iteration
lens updates records positionally:
\begin{align*}
\ell^* .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n) &= t_1 \cdot
\ldots \cdot t_n\\
\text{ where } t_i &= \begin{cases}
\ell.\pput \; s_i \; v_i & \text{if } 1 \leq i \leq \min\{m, n\}\\
\ell.\create \; v_i & \text{if } \min\{m, n\} < i \leq n
\end{cases}\\
\end{align*}
In their work, Barbosa et. al. enrich the types of lenses with ``chunks''
identifying reorderable pieces of the source and view that should be re-aligned
after an update, and formulate behavioral laws that capture essential
constraints on the handling of chunks. They develop a core language
of matching lenses for strings, together with a set of ``alignment
combinators'' that implement a variety of alignment strategies. The obvious
shortcoming of matching lenses is that they only deal with alignment; they do
not solve the general issue of update propagation.

The next line of work is Diskin et. al.'s {\em delta
lenses}\cite{diskin2011asymmetric,diskin2011state}. Diskin et. al.'s delta
lenses work as follows: let $\ell : S \Leftrightarrow V$ be a lens from
$S$ to $V$. Assume that $s$ is a source, and that $v = \ell.\get s$. A user
performs an edits $d$ on $v$ to get $v'$: $v \xrightarrow{d} v'$. The \pput
function now takes the view edit $d$ and the original source $s$ and produces a
new source $s'$: $s \xrightarrow{\pput(d, s)} s'$. One thing that Diskin et. al.
do not do is that they do not give any concrete examples of delta lenses,
no composition, no notion of equivalence, and no combinators for constructing
delta lenses.

The next line of work is Hofmann et. al.'s {\em edit
lenses}\cite{hofmann2012edit}. Edit lenses are similar to delta lenses, but edit
lenses are more explicit. Specifically, the source and view edits are
represented as elements of a {\em monoid}, and the lens acts on sources and
views via monoid actions. That is if $v$ is a view and $d_v$ is an edit that
can be made on $v$ to produce $v'$, then $d_v \cdot v = v'$. The intended usage
of an edit lens is as follows. There are two users, one holding an element of
$S$ the other one an element of $V$ both referred to as {\em replicas}.
Initially, they hold $init_S \in S$ and $init_V \in V$ respectively. The users
then perform actions and propagate them across the lens. An action consists of
producing an edit $d_s$ (or $d_v$), applying it to one's current replica $s$
(resp. $v$), putting the edit through the lens to obtain an edit $d_v$ (resp.
$d_s$), and asking the user on the other side to apply $d_v$ (resp. $d_s$) to
their replica. One thing that Hofmann et. al. do that Diskin et. al. do not do
is that they do give concrete examples of edit lenses, define a suitable notion
of composition on them, and define combinators for constructing edit lenses.

The final line of work in this area is Pacheco et. al.'s
version of delta lenses\cite{pacheco2012delta}. Pacheco et. al. are the
first to give concrete examples of delta lenses, define a suitable notion
of composition on them, and define combinators for constructing delta
lenses. Pacheco et al. define a delta lens $\ell$ (d-lens for short), denote by
$\ell: S \trianglerighteq_{\blacktriangle} V$ to be a bidirectional
transformations that comprises four total functions:
\begin{align*}
\get &: S \longrightarrow V\\
\pput &: \forall(v, s) : V \times S. (v \triangle \get \;s) \longrightarrow
SA\\
\get_{\blacktriangle} &: \forall\{s':S, s:S\}. (s' \triangle s)
\longrightarrow (\get \; s')\triangle (\get \; s)\\
\pput_{\blacktriangle} &:\forall\{(v, s):V \times S\},d:(v \triangle \get
\;s).(\pput \; (v, s) \; d) \triangle s
\end{align*}
that satisfy the following propertes:
\begin{align*}
\get \; (\pput \; (v, s) \; d) = v \tag{PUTGET}\\
\pput \; (\get \; s, s) \; id = s \tag{GETPUT}\\
\get_{\blacktriangle} \; (\pput_{\blacktriangle} \; d) = d
\tag{$\text{PUTGET}_{\blacktriangle}$}\\
\pput_{\blacktriangle} \; id = id \tag{$\text{PUTGET}_{\blacktriangle}$}
\end{align*}
The \get function is a just like a \get function in Foster's et. al.'s original
work. The \pput function is different though. Let $s \in S$ and $v \in V$, and
let $d$ be an update from $v$ to $\ell.\get \; s$; for example, $d$ is a
sequence of insertions and deletions that transforms $v$ to $\ell.\get \; s$.
Then \pput takes in $s$ and $v$ as input, and in addition receives $d$ to
compute $\pput \; (v, s) \; d$.

The $\get_{\blacktriangle}$ and $\pput_{\blacktriangle}$ functions are functions
that transform deltas to deltas. The last two axioms for delta lenses say that
if $d$ is a delta from $v$ to $\ell.\get \; s$ and $\Delta$ is a delta from
$\pput \; (v, s) \; d$ to $s$, then $\get_{\blacktriangle} \; \Delta = d$, and
that if $d$ is the identity delta, then $\Delta$ is also the identity delta.

In addition to describing a delta lens language, Pacheco et. al. also
address the following shortcoming with edit lenses: edit lenses include
combinators for constructing lenses over {\em container types}.
In the simplest case, container types can be thought of as algebraic datatypes
such as lists and trees; each value of a container type has a certain {\em
shape}, and each shape has a set of {\em positions} associated with it. For
instance, the shape of a list of size $n$ is just $n$, while the positions of
such a list are the numbers $0, 1, \ldots, n-1$.

The shortcoming of the container edit lenses addressed by Pacheco et. al.
elements is that these lenses always preserve the number of positions; for
instance if $\ell$ is a container lens that maps lists to trees, and if $xs$ is
a list of size $n$, then the number of positions of $\ell.\get \; xs$ is also
$n$. Pacheco et. al. define a new set of combinators for container lenses that
can say map a list of size $5$ to a tree with 4 nodes.

In this article we show that it is not necessary for the \pput component of a lens
to explicitly propagate updates: instead, we can specify what the put function should
evaluate to, by imposing a more precise type judgement on lenses.

The idea of the stronger typing judgement is that we embed sources and views inside
of {\em difference structures} and then we define a lens that describes how the
differences behave as they cross in-between the source and the view. A
lens that is validly typed is then a lens that preserves this overall structure.

Here are the actual definitions:
\begin{definition}
A weighted set is a pair $(X, |\cdot|)$ where $X$ is a set and $|\cdot| : X
\longrightarrow \mathbb{R}^{\geq 0}$ is a map from $X$ to the non-negative
reals.
\end{definition}
\begin{definition}
A difference structure is a triple $(S, \partial S, \delta_S)$ where $S$ and
$\partial S$ are weighted sets, and $\delta_S:
S \times S \longrightarrow \partial S$ is a function satisfying
\begin{align*}
\delta(s, s') &= \delta(s', s)\tag{A1}\\
|\delta(s, s)| &= 0\tag{A2}\\
|\delta(s, s')| &\leq |s| + |s'|\tag{A3}
\end{align*}
\end{definition}
\begin{definition}
Let $(S, \partial S, \delta_S)$ and $(V, \partial V, \delta_V)$ be
difference structures. Let $\partial(S, V) : \partial S \Leftrightarrow \partial
V$ be a lens from $\partial S$ to $\partial V$. Then a lens $\ell : S
\Leftrightarrow V$ also has type $\ell : (S, \partial S,
\delta_S) \xLeftrightarrow{\partial(S, V)} (V, \partial V, \delta_V)$ if for
all $s \in S, v \in V$,
\begin{equation}\label{difflenslaw}
(\partial (S, V)).\get\; (\delta_S(\ell.\pput \; s \; v, s)) = \delta_V(v,
\ell.\get \; s)
\end{equation}
\end{definition}
The extra difference structure on the sources and the views, as well as the lens
$\partial(S, V)$ force the lens $\ell$ to satisfy \cref{difflenslaw}; this law is
therefore a specification that constrains what $\ell$ can do.

For example, suppose that a source $S$ consists of a pair of integers, and that the
view $V$ for $S$ consists of a single integer. Further, suppose that we wish
for the \get component of any lens from $S$ to $V$ to simply project away the
first factor:that is for any $\ell : \mathbb{Z} \times \mathbb{Z}
\Leftrightarrow \mathbb{Z}$, $\ell.\get \; (a, b) = b$ for all integers $a, b$.
Suppose further that we wish to force the \pput function of any lens $\ell$
defined on the data to restore the factor projected away by the \get function:
that is, we wish to enforce that for all integers $a, b, c$, $\pi_1(\ell.\pput
\; (a, b) \; c) = a$, where $\pi_1$ is the projection on the first factor.

Now consider the lens $\ell_{bad}$ defined by
\begin{align*}
\ell_{bad}.\get \; (a, b) &= b\\
\ell_{bad}.\pput \; (a, b) \; c &= \begin{cases}
(a, c) & \text{if }b = c\\
(0, c) & \text{otherwise}\\
\end{cases}\\
\ell.\create \; b &= (0,b)
\end{align*}
Then $\ell_{bad}$ is a valid lens even though $\ell_{bad}$ does not
always restore the first factor. For example, observe that $\ell.\pput (1, 2) \;
3 = (0,3)$. In order to give a more constrained specification for the desired
lens we define the difference $\delta_S((a, b), (c, d))$ between $(a, b)$ and
$(c, d)$ by $$\delta_S((a, b), (c,
d)) = \begin{cases}
0 & \text{if }a = c, b = d\\
1 & \text{if }a=c, b \neq d\\
1 & \text{if }a \neq c, b = d\\
2 & \text{if }a \neq c, b \neq d
\end{cases}$$
On the view side, we define the difference $\delta_V(a, b)$
between $a$ and $b$ by, $$\delta_V(a, b) = \begin{cases}
0 & \text{if } a = b\\
1 & \text{if } a \neq b
\end{cases}$$

Next we define a lens $\partial(S, V) : \partial S \Leftrightarrow \partial V$
which describes how differences should be propagated. For this example, we
define $\partial(S,V)$ to simply be the \ccopy lens from $\partial S$ to
$\partial V$, keeping in mind that $\partial S = \partial V = \mathbb{Z}$.
\cref{difflenslaw} enables us to rule out $\ell_{bad}$, since
\begin{align*}
(\partial (S, V)).\get \; (\delta_S(\ell_{bad}.\pput \; (1,2) \; 3, (1,2))) &=
(\partial (S, V)).\get \; (\delta_S((0,3), (1,2)))\\
&= (\partial (S, V)).\get \; 2\\
&= 2\\
&\neq 1\\
&= \delta_V(3, 2)\\
&= \delta_V(3, \ell_{bad}.\get \; (1, 2))
\end{align*}
Consequently $\ell_{bad}$ is not a valid delta lens. On the
other hand, the lens $\ell_{good} : S \Leftrightarrow V$ defined by
\begin{align*}
\ell_{good}.\get \; (a, b) &= b\\
\ell_{good}.\pput \; (a, b) \; c &= (a, c)\\
\ell_{good}.\create \; b &= (0, b)
\end{align*}
is indeed a valid delta lens since the its \pput component always restores the
first factor.

For a more realistic example using a difference lens combinator, let $S
= name \cdot dates$ and $V = name$. Let $\ell :
S \Leftrightarrow V = (copy \; name) \cdot (del \; dates)$, and suppose that we
wish to define a lens of type $S^* \Leftrightarrow V^*$, where the difference
function used on the view side is the min-edit distance difference function.
Assuming that we inductively use the difference function
$$ \delta(d_1, d_2) =
\begin{cases}
0 & \text{if }d_1 = d_2 \\
1 & \text{otherwise}
\end{cases}
$$
for both the name and date fields, then with the min-edit distance difference
function, the difference between the string $v$ given by
\begin{lstlisting}
John Doe
Jane Smith
\end{lstlisting}
and the string $v'$ given by
\begin{lstlisting}
Jane Smith
John Doe
\end{lstlisting}
is the empty list. Let $s$ =
\begin{lstlisting}
John Doe,1900-2000
Jane Smith,1901-2001
\end{lstlisting}Then $\ell.\get \; s = v$. Now assume that $v$ is edited to $v'$
given by
\begin{lstlisting}
Mark Moe
John Doe
Jane Smith
\end{lstlisting}
where a new record \lstinline|Mark Moe| has been added to $v$.
With the min-edit distance on the view side, $\delta(v', \ell.\get \; s)$ is
the singleton list [``Mark Moe''].

On the source side, the min-edit distance matching lens $\ell^*_{MIN-EDIT}$,
$\ell^*_{MIN-EDIT}.\pput \; s \; v'$ is the string
\begin{lstlisting}
Mark Moe,0000-0000
John Doe,1900-2000
Jane Smith,1901-2001
\end{lstlisting}
while with the positional lens $\ell^*$, then $\ell^*.\pput \; s \; v'$ is the
string
\begin{lstlisting}
Mark Moe,1900-2000
John Doe,1901-2001
Jane Smith,0000-0000
\end{lstlisting}
With the min-edit distance difference function,
$\delta(\ell^*_{MIN-EDIT}.\pput \; s \; v', s)$ is the singleton list
$[\text{``Mark Moe, 0000-0000''}]$, while with the
positional lens, $\delta(\ell^*.\pput \; s \; v', s)$ is the list
given by
\begin{lstlisting}
Mark Moe,1900-2000
(0,1)
(0,1)
\end{lstlisting}
The min-edit distance satisfies \cref{difflenslaw} since
\begin{align*}
(\partial (S, V))^*.\get\; (\delta_S(\ell^*_{MIN-EDIT}.\pput \; s \; v', s)) &=
(\partial (S, V))^*.\get\; ([\text{``Mark Moe, 1900-2000''}])\\
&= [\text{``Mark Moe''}]\\
&=
\delta_V(v', \ell.\get \; s)
\end{align*}
On the other hand, the positional lens does not satisfy \cref{difflenslaw} since
\begin{align*}
(\partial (S, V))^*.\get\; (\delta_S(\ell^*.\pput \; s \; v', s)) &=
(\partial (S, V))^*.\get\; ([\text{``Mark Moe, 1900-2000''}, (0,1), (0,1)])\\
&=[\text{``Mark Moe''}, (0,*), (0,*)]\\
& \neq \delta_V(v', \ell.\get \; s)
\end{align*}
Let $\mathcal{M}_V$ be the set of min-edit distance alignments on the view side,
and $\mathcal{M}_S$ the corresponding alignments on the source side.
Then, the typing judgement
$$\ell^*_{MIN-EDIT}:(S, (\partial S \sqcup S)^*, \match(\mathcal{M}_S, \delta_s))
\xLeftrightarrow{\partial(S, V)^*} (V, (\partial V \sqcup V)^*, \match(\mathcal{M}_V, \delta_v))$$
is valid, whereas the typing judgement
$$\ell^*:(S, (\partial S \sqcup S)^*, \match(\mathcal{M}_S, \delta_s))
\xLeftrightarrow{\partial(S, V)^*} (V, (\partial V \sqcup V)^*, \match(\mathcal{M}_V, \delta_v))$$
is not.
\iffalse
For example, if $\ell : S \Leftrightarrow V$ is a lens, Foster's initial
definition of the iteration lens $\ell^* : S^* \Leftrightarrow V^*$ was
\begin{align*}
\ell^* .\get \; (s_1, \ldots, s_n) &= (\ell.\get \; s_1) \cdot \ldots
\cdot (\ell.\get \; s_n)\\
\ell^* .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n) &= t_1 \cdot
\ldots \cdot t_n\\
\text{ where } t_i &= \begin{cases}
\ell.\pput \; s_i \; v_i & \text{if } 1 \leq i \leq \min\{m, n\}\\
\ell.\create \; v_i & \text{if } \min\{m, n\} < i \leq n
\end{cases}\\
\ell^*.\create \; (v_1, \ldots, v_n) &= (\ell.\create \; v_1) \cdot \ldots
\cdot (\ell.\create \; v_n)
\end{align*}
This lens always propagates updates positionally, which is unintended behaviour
in many cases. For example suppose that $\ell$ is a lens that transforms
source data in the form of an XML document representing the names, dates and
nationalities of a collection of classical music composers such as,

\begin{lstlisting}
<composers>
<composer>
<name>Jean Sibelius</name>
<dates>1865-1957</dates>
<nationality>Finnish</nationality>
</composer>
<composer>
<name>Aaron Copland</name>
<dates>1910-1990</dates>
<nationality>American</nationality>
</composer>
<composer>
<name>Benjamin Britten</name>
<dates>1913-1976</dates>
<nationality>English</nationality>
</composer>
</composers>
\end{lstlisting}
into view data stoored in ASCII text representing the names and dates of each
composer:
\begin{lstlisting}
Jean Sibelius, 1865-1957
Aaron Copland, 1910-1990
Benjamin Britten, 1913-1976
\end{lstlisting}
Now suppose that the elements in the view are reordered:
\begin{lstlisting}
Aaron Copland, 1910-1990
Benjamin Britten, 1913-1976
Jean Sibelius, 1865-1957
\end{lstlisting}
Then with the positional iteration lens, the updated source is
\begin{lstlisting}
<composers>
<composer>
<name>Aaron Copeland</name>
<dates>1910-1990</dates>
<nationality>Finnish</nationality>
</composer>
<composer>
<name>Benjamin Britten</name>
<dates>1913-1976</dates>
<nationality>American</nationality>
</composer>
<composer>
<name>Jean Sibelius</name>
<dates>1865-1957</dates>
<nationality>English</nationality>
</composer>
</composers>
\end{lstlisting}
so each of the composers now has the wrong nationality in the source.

\section{Difference functions}
\begin{definition}
A weighted set is a pair $(X, |\cdot|)$ where $X$ is a set and $|\cdot| : X
\longrightarrow \mathbb{R}^{\geq 0}$ is a map from $X$ to the non-negative
reals.
\end{definition}
\begin{definition}
A difference structure is a triple $(S, \partial S, \delta_S)$ where $S$ and
$\partial S$ are weighted sets, and $\delta_S:
S \times S \longrightarrow \partial S$ is a function satisfying
\begin{align*}
\delta(s, s') &= \delta(s', s)\tag{A1}\\
|\delta(s, s)| &= 0\tag{A2}\\
|\delta(s, s')| &\leq |s| + |s'|\tag{A3}
\end{align*}
We refer to the function $\delta$ as the difference function of the structure.
\end{definition}
\iffalse

\begin{example}
Let $X$ be any set. Define $x \leq y$ if and only if $x=y$, and define $|X|=1$
for all $x \in X$. Let $Q = \{0,1\}$, and define
$$\delta(a, b) = \begin{cases} 1 & \text{if } a \neq b\\
0 & \text{otherwise}
\end{cases}$$
Then $\delta$ is a difference function.
\end{example}

Various different edit distances on strings give a way of defining delta
functions on sets of strings:
\begin{example}
The {\em Levenshtein distance} between two words is the minimum number of
single-character edits (insertions, deletions or substitutions) required to
change one word into the other. For example, the Levenshtein distance between
"kitten" and "sitting" is 3, since the following three edits change one into
the other, and there is no way to do it with fewer than three edits:

\begin{enumerate}
  \item
  kitten -> sitten (substitution of "s" for "k")
  \item
  sitten -> sittin (substitution of "i" for "e")
  \item
  sittin -> sitting (insertion of "g" at the end).
\end{enumerate}
\end{example}
\begin{example}
In contrast, the Longest Common Subsequence(LCS) distance only allows for
insertions and deletions:
\begin{enumerate}
  \item
  delete k at 0
  \item
  insert s at 0
  \item
  delete e at 4
  \item
  insert i at 4
  \item
  insert g at 6
\end{enumerate}
for a total cost/distance of 5 operations.
\end{example}
Both Levenshtein and LCS distances are integer-valued distance functions since
they are metrics.
\begin{example}
If all the strings in a set are of the same length, then the {\em Hamming
distance} is a difference function since the Hamming distance is also a metric.
\end{example}
\begin{example}
Alignment algorithms also give us a way of defining integer-valued distance
functions on sets of strings by taking the difference of two strings to be the
total number of characters that are not equal according to the alignment.

When deciding which alignment to choose between two strings, we often wish to
choose an alignment that is optimal for alignment strategy. With
the positional alignment strategy, given two strings $s$ and $t$, then $s_1$ is
matched to $t_1$, $s_2$ is matched to $t_2$ and so on.
Also, $\mathcal{M}_{POS}$ is an optimal set of alignments for a positional
alignment strategy.

For example, using the positional alignment strategy on the strings "kitten" and
"sitting", then "k" is matched to "s", "i" to "i" and so on. The characters
which are included in the delta from "kitten" are "k", "e" while the
in the characters included from "sitting" are "s", "i" and "g", hence the delta
is equal to 5.

For the non-crossing strategy, we choose among alignments which are increasing
rather than positional, while for the min-edit distance strategy, we consider
all possible alignments.
\end{example}

\begin{example}\label{naturalnumbers}
Let $P$ be the set of non-negative integers with weight function $|a| = a$. Then
$\delta$ is a difference function.
Define $\delta(a, b) = |a-b|$. Then $\delta$ is a difference function.
\end{example}
\begin{example}
Let $X$ be a normed vector space and define $\delta(x, y) = |x-y|$. Then
$\delta$ is a difference function (in fact, $\delta$ is also a metric).
\end{example}
\begin{example}\label{divisibility}
Let $P$ be the set of positive integers with weight function $|\cdot|$ defined
by $|p_1^{a_1} \cdot \ldots \cdot p_n^{a_n}| = a_1 + \ldots + a_n$. Define
$\delta(a, b) = ab/(gcd(a, b))^2$.
Then $\delta$ is a difference function.
\end{example}
\begin{example}\label{boolean}
Let $X$ be a set and $P$ the set of finite subsets of $X$ with weight function
given by the size of the set. Define $\delta(A, B) = (A \setminus B) \cup (B
\setminus A)$. Then $\delta$ is a difference function.
\end{example}
\begin{example}\label{general}
Generalizing \cref{divisibility} and
\cref{boolean}, let $P$ be a graded complete semi-modular lattice satisfying
$|\bigwedge P|=0$. Define $\delta(a, b) = \bigwedge \{x \; | \; (a \wedge b)
\vee x = a\} \vee \bigwedge \{x \; | \; (a \wedge b) \vee x = b\}$. Then
$\delta$ is a difference function.
\end{example}
\begin{proof}
Let $a, b \in P$. Let $X = \{x \; | \; (a \wedge b) \vee x = a\}$ and $Y = \{y
\; | \; (a \wedge b) \vee y = b\}$
\begin{enumerate}
  \item[(A1)]
  Certainly $\delta(a, b) = \delta(b, a)$.
  \item[(A2)]
  Observe that $\bigwedge X = \bigwedge Y = \bot$, hence $\delta(a,
  a) = \bot$, and $|\delta(a, a)| = |\bot| = 0$.
  \item[(A3)]
  Since $P$ is semi-modular, then $|x \wedge y| + |x \vee y| \leq |x| + |y|$
  for all $x, y \in P$, from which it follows that $$|\delta(a, b)| =
  \left|\bigwedge X \vee \bigwedge Y\right| \leq \left|\bigwedge X \vee
  \bigwedge Y\right| + \left|\bigwedge X \wedge \bigwedge Y\right| \leq
  \left|\bigwedge X\right| + \left|\bigwedge Y\right|$$
\end{enumerate}
\end{proof}
\fi

\subsection{Constructions on difference functions}
\subsubsection{Products}
Given difference functions $\delta_i: P_i \times P_i \longrightarrow Q_i$ for
$i \in \{1,2\}$, define $\delta_1 \times \delta_2 :(P_1 \times P_2) \times
(P_1 \times P_2) \longrightarrow (Q_1 \times Q_2)$ by
$$(\delta_1 \times \delta_2)((x, y), (x', y')) = (\delta_1(x, x'), \delta_2(y,
y'))$$
\begin{claim}
$\delta_1 \times \delta_2$ is a difference function.
\end{claim}
\begin{proof}
Let $x, x' \in P_1$ and $y, y' \in P_2$.
\begin{enumerate}
  \item[(A1)]
  Observe that
  \begin{align*}
(\delta_1 \times \delta_2)((x, y), (x', y')) &= (\delta_1(x, x'),
\delta_2(y, y'))\\
&= (\delta_1(x', x), \delta_2(y', y))\\
&= (\delta_1 \times \delta_2)((x', y'), (x, y))
\end{align*}
\item[(A2)] Observe that
$$|(\delta_1 \times \delta_2)((x, y), (x, y))| = |(\delta_1(x, x),
\delta_2(y, y))| = |\delta_1(x, x)| + |\delta_2(y,y)| = 0 + 0 = 0$$
\item[(A3)]
Observe that
\begin{align*}
|(\delta_1 \times \delta_2)((x, y), (x', y'))| &= |(\delta_1(x, x'),
\delta_2(y, y'))|\\
&= |(\delta_1(x, x')| + |(\delta_2(y, y')|\\
&\leq (|x| + |x'|) + (|y| + |y'|)\\
&= (|x| + |y|) + (|x'| + |y'|)\\
&= |(x, y)| + |(x', y')|
\end{align*}
\end{enumerate}
\end{proof}
\subsubsection{Coproducts}
Given difference functions $\delta_i: P_i \times P_i \longrightarrow Q_i$ for
$i \in \{1,2\}$, define $\delta_1 \sqcup \delta_2 :P_1 \sqcup P_2
\longrightarrow Q_1 \sqcup Q_2 \sqcup (P_1 \times P_2)$ by
$$(\delta_1 \sqcup \delta_2)(x, y) =
\begin{cases}
\delta_1(x, y) & \text{if }x, y \in P_1\\
\delta_2(x, y) & \text{if }x, y \in P_2\\
(x, y) & \text{if }x \in P_1, y \in P_2\\
(y, x) & \text{if }x \in P_2, y \in P_1
\end{cases}$$
\begin{claim}
$\delta_1 \sqcup \delta_2$ is a difference function
\end{claim}
\begin{proof}
Let $x, x' \in P_1$ and $y, y' \in P_2$.
\begin{enumerate}
  \item[(A1)]
  If $x, y \in P_i$, then $(\delta_1 \sqcup \delta_2)(x, y) = \delta_i(x, y) =
  \delta_i(y, x)$. If $x \in P_1$ and $y \in P_2$, then $(\delta_1 \sqcup
  \delta_2)(x, y) = (x, y) = (\delta_1 \sqcup \delta_2)(y, x)$, and similarly
  if $x \in P_2$ and $y \in P_1$, then $(\delta_1 \sqcup \delta_2)(x, y) =
  (y,x) = (\delta_1 \sqcup \delta_2)(y, x)$
  \item[(A2)] If $x \in P_i$, then $|(\delta_1 \sqcup \delta_2)(x, x)| =
  |\delta_i(x, x)| = 0$
  \item[(A3)]
  If $x, y \in P_i$, then $|(\delta_1 \sqcup \delta_2)(x, y)| = |\delta_i(x, y)|
  \leq |x| + |y|$. If $x \in P_1$ and $y \in P_2$, then $|(\delta_1 \sqcup
  \delta_2)(x, y)| = |(x, y)| = |x| + |y|$, and if $x \in P_2$ and $y \in
  P_1$, then $|(\delta_1 \sqcup \delta_2)(x, y)| = (|(y, x)|) = |y| + |x|$.
\end{enumerate}
\end{proof}
\subsubsection{List difference functions}
Let $P$ be a weighted set. Let $x=(x_1, \ldots, x_m), y=(y_1, \ldots, y_n)$ be
in $P^*$. Given a difference function $\delta: P \times P \longrightarrow Q$, we
define the list difference function $\delta^*:P^* \longrightarrow (Q \sqcup
P)^*$ on $x$ and $y$ to be the list given by $\delta(x_i, y_i)$ for $1 \leq i
\leq m$, except that we omit $\delta(x_i, y_i)$ in position $i$ if $x_i = y_i$.
For $m < i \leq n$ we add $y_i$ at position $i$. Finally we collapse the list
to get rid of the empty spaces left after the omissions.

For instance suppose that $S$ is the set of all letters in the alphabet and
$\partial S = \{\epsilon\} \sqcup \{\{x, y\} \; | \; x, y \in S, x \neq y\}$
with $$\delta(x, y) =
\begin{cases}
\epsilon & \text{if } x = y\\
\{x,y\} & \text{if } x \neq y
\end{cases}$$
Then $\delta_S^*((a, c, b, d), (a, b, c, d, e)) = (\{b, c\}, \{b, c\},
e)$ where we match and remove the first $a$ and the fourth $d$ in. Since $b \neq
c$ we have $\{b,c\}$ in the second and third positions, and since $e$ is
unmatched, we include $e$ in the fifth position. Finally we collapse the list to
get rid of the two empty spaces created.

\begin{claim}
$\delta^*$ is a difference function
\end{claim}
\begin{proof}
Let $x = (x_1, \ldots, x_m), y = (y_1, \ldots, y_n) \in P^*$.
\begin{enumerate}
  \item[(A1)]
  By construction $\delta^*(x, y) = \delta^*(y, x)$.
  \item[(A2)] Observe that $|\delta^*(x, x)| = \sum_{i=1}^m|\delta(x_i,
  x_i)| =  \sum_{i=1}^m 0 = 0$.
  \item[(A3)]
  Assume WLOG that $m \leq n$. Then
  \begin{align*}
|\delta^*(x, y)| &=
\sum_{i=1}^m|\delta(x_i, y_i)| + \sum_{i=m+1}^n |y_i|\\
&\leq \sum_{i=1}^m|x_i| + |y_i| + \sum_{i=m+1}^n |y_i|\\
&= \sum_{i=1}^m|x_i| + \sum_{i=1}^n |y_i|\\
&= |x| + |y|
\end{align*}
\end{enumerate}
\end{proof}
\subsubsection{Matching difference functions}
Now let $\mathcal{M} = \{\varphi_{(x_1, \ldots, x_m), (y_1, \ldots, y_n)} \; |
\; x_i, y_j \in P, m \leq n\}$ be a set of injective functions $\varphi_{(x_1,
\ldots, x_m), (y_1, \ldots, y_n)} : [m] \longrightarrow [n]$. Given a difference
function $\delta:
P \times P \longrightarrow Q$, we define the difference function
$\match(\mathcal{M}, \delta):P^* \longrightarrow (Q \sqcup P)^*$ on $x$
and $y$ to be the list given by $\delta(x_i, y_{\varphi(i)})$ in position
$\varphi(i)$ for $1 \leq i \leq m$, except that we omit $\delta(x_i,
y_{\varphi(i)})$ in position $\varphi(i)$ if $x_i = y_{\varphi(i)}$. For $i \in
[n] \setminus \varphi([m])$ we add $y_i$ at position $i$. Finally we collapse
the list to get rid of the empty spaces left after the omissions.

Going back to our previous example, let $x = \{a, c, b, d\}$ and $y = \{a, b, c,
d, e\}$, and assume that $\varphi_{x, y} : [4] \longrightarrow [5]$ is given by
$$\varphi(1) = 1, \; \varphi(2) = 3, \; \varphi(3) = 2, \text{ and } \varphi(4)
= 4$$ Then $\match(\mathcal{M}, \delta)((a, c, b, d), (a, b, c, d, e)) =
(e)$.
\begin{claim}
$\match(\mathcal{M}, \delta)$ is a difference function.
\end{claim}
\begin{proof}
Let $x = \{x_1, \ldots, x_m\}, y = \{y_1, \ldots, y_n\} \in \match(P)$.
\begin{enumerate}
  \item[(A1)]
  By construction $\match(\mathcal{M}, \delta)(x, y) =
  \match(\mathcal{M}, \delta)(y, x)$.
  \item[(A2)] Observe that $|(\match(\mathcal{M}, \delta))(x, x)| =
  \sum_{i=1}^m|\delta(x_i, x_{\varphi(i)}| = |\delta(x_i, x_i)| =   \sum_{i=1}^m
  0 = 0$.
  \item[(A3)]
  Assume WLOG that $m \leq n$. Then
  \begin{align*}
|\match(\mathcal{M}, \delta)(x, y)| &=
\sum_{i=1}^m|\delta(x_i, y_{\varphi(i)})| + \sum_{i \in [n] \setminus
\varphi([m])} |y_i|\\
&\leq \sum_{i=1}^m|x_i| + |y_{\varphi(i)}| + \sum_{i \in [n] \setminus
\varphi([m])} |y_i|\\
&= \sum_{i=1}^m|x_i| + \sum_{i=1}^n |y_i|\\
&= |x| + |y|
\end{align*}
\end{enumerate}
\end{proof}
\fi
\section{Lens Combinators}
\subsection{Copy}
Let $P$ be any set. Define the $\ccopy(P) : P \Leftrightarrow P$ by
\begin{align}
\ccopy(P).\get \; s &= s\\
\ccopy(P).\pput \; s \; v &= v\\
\ccopy(P).\create \; v &= v
\end{align}
Then $\ccopy(P)$ is a lens. Now let $\partial S$ be a set, $\delta :
S \times S \longrightarrow \partial S$ a function.
\begin{claim}
$\ccopy(S) : S \xLeftrightarrow{\delta , \ccopy(\partial S), \delta} S$.
\end{claim}
\begin{proof}
We need to show that $(\ccopy(\partial S)).\get \; (\delta(\ccopy(S).\pput
\; s \; v, s)) = \delta(v, \ccopy(S).\get \; s)$ which follows from the fact
that $$ (\ccopy(\partial S)).\get \; (\delta(\ccopy(S).\pput \; s
\; v, s)) = \delta(\ccopy(S).\pput \; s \; v, s) = \delta(v, \ccopy(S).\get
\; s) $$
\end{proof}
\subsection{Constant}
Let $S$ be any set with $s'\in S$. Define the \const lens $\const(S) : S
\Leftrightarrow \{\star\}$ by
\begin{align*}
\const.\get \; s &= \star\\
\const.\pput \; s \; v &= s\\
\const.\create \; v &= s'
\end{align*}
Then $\const(S)$ is a lens. 
\begin{claim}
Let $\partial S$ be any set, and $\delta : S \times S \longrightarrow \partial
S$ any function satisfying $\delta(s, s) = \delta(s', s')$ for all $s, s' \in
S$. Define $\beta : \{\star\} \times \{\star\} \longrightarrow \partial S$ by
$\beta(\star, \star) = \delta(s, s)$ for any $s \in S$. Let $\ell : \partial S
\Leftrightarrow \partial S$ be any lens satisfying $\ell.\get \; \delta(s, s) =
\delta(s, s)$ for all $s \in S$. Then $\const(S): S \xLeftrightarrow{\delta,
\ell, \beta} \{\star\}$.
\end{claim}
\begin{proof}
We need to show that $\ell.\get \; (\delta(\const(S).\pput \; s \; \star, s)) =
\beta (\star, \const(S).\get \; s)$
which follows from the fact that
$$ \ell.\get \; (\delta(\const(S).\pput \; s \; \star, s)) = \ell.\get
\; \delta(s, s) = \delta(s, s) = \beta(\star, \star) = \beta(\star,
\const(S).\get \; s)$$
\end{proof}

\subsection{Default}
Let $\ell : S \Leftrightarrow V$ be a lens. Let $f :
V \longrightarrow S$ be any function. Define the lens $\default(\ell, f) :
S \Leftrightarrow V$ by
\begin{align*}
\default(\ell, f).\get \; s &= \ell.\get \; s\\
\default(\ell, f).\pput \; s \; v &= \ell.\pput \; s \; v\\
\default(\ell, f).\create \; v &= \ell.\pput \; (f \; v) \; v
\end{align*}

\begin{claim}
If $\ell : S \xLeftrightarrow{\delta_S, \partial (S, V), \partial V}
V$ and $f: S \longrightarrow V$, then $\default(\ell, f) : S
\xLeftrightarrow{\delta_S, \partial (S, V), \partial V} V$.
\end{claim}
\begin{proof}
Follows from the fact that $\ell : S \xLeftrightarrow{\delta_S, \partial (S, V),
\partial V} V$.
\end{proof}
\subsection{Product}
Let $\ell_1 : S_1 \Leftrightarrow V_1$ and $\ell_2 : S_2 \Leftrightarrow V_2$ be
lenses. Define the lens $\ell_1 \times \ell_2 : (S_1 \times S_2) \Leftrightarrow
(V_1 \times V_2)$ by
\begin{align*}
(\ell_1 \times \ell_2).\get \; (s_1, s_2) &= (\ell_1.\get \; s_1, \ell_2.\get
\; s_1)\\
(\ell_1 \times \ell_2).\pput \; (s_1, s_2) \; (v_1, v_2) &= (\ell_1.\pput \; s_1
\; v_1, \ell_2.\pput \; s_2 \; v_2)\\
(\ell_1 \times \ell_2).\create \; (v_1, v_2) &= (\ell_1.\create \; v_1,
\ell_2.\create \; v_2)
\end{align*}
\begin{claim}\label{productislens}
If $\ell_1 :S_1 \xLeftrightarrow{\delta_{S_1}, \partial (S_1, V_1),
\delta_{V_1}} V_1$ and $\ell_2 : S_2 \xLeftrightarrow{\delta_{S_2}, \partial
(S_2, V_2), \delta_{V_2}} V_2$, then \\
$\ell_1 \times \ell_2  : S_1 \times S_2 
\xLeftrightarrow{\delta_{S_1} \times \delta_{S_2}, \partial (S_1, V_1) \times
\partial (S_2, V_2), \delta_{V_1} \times \delta_{V_2}} V_1 \times V_2$.
\end{claim}
\begin{proof}
We need to show that
$$(\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \; (\delta((\ell_1 \times
\ell_2).\pput \; (s_1, s_2) \; (v_1, v_2), (s_1, s_2))) = \delta((v_1, v_2),
(\ell_1 \times \ell_2).\get \; (s_1, s_2))$$ which follows from the fact that
\begin{align*}
&(\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \; (\delta((\ell_1 \times
\ell_2).\pput \; (s_1, s_2) \; (v_1, v_2), (s_1, s_2)))\\
&= (\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \;
(\delta((\ell_1.\pput \; s_1 \; v_1, \ell_2.\pput \; s_2 \; v_2), (s_1, s_2))\\
&= (\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \;
(\delta(\ell_1.\pput \; s_1 \; v_1, s_1), \delta(\ell_2.\pput \; s_2 \; v_2,
s_2))\\
&= ((\partial (S_1, V_1)).\get \; (\delta(\ell_1.\pput \;
s_1 \; v_1, s_1)), (\partial (S_2, V_2)).\get \; (\delta(\ell_2.\pput \; s_2 \;
v_2, s_2)))\\
&= (\delta(v_1, \ell_1.\get \; s_1), \delta(v_2, \ell_2.\get \; s_2))\\
&= \delta((v_1, v_2), (\ell_1.\get \; s_1, \ell_2.\get \; s_2))\\
&= \delta((v_1, v_2), (\ell_1 \times \ell_2).\get \; (s_1, s_2))
\end{align*}
\end{proof}
\subsection{Sum}
Let $\ell_1 : S_1 \Leftrightarrow V_1$ and $\ell_2 : S_2 \Leftrightarrow V_2$ be
lenses. Define the lens
$\ell_1 \sqcup \ell_2 :
(S_1 \sqcup S_2) \Leftrightarrow (V_1 \sqcup V_2)$ by
\begin{align*}
(\ell_1 \sqcup \ell_2).\get \; s &=
\begin{cases}
\ell_1.\get \; s & \text{if } s \in S_1\\
\ell_2.\get \; s & \text{if } s \in S_2\\
\end{cases}\\
(\ell_1 \sqcup \ell_2).\pput \; s \; v&=
\begin{cases}
\ell_1.\pput \; s \; v& \text{if } s \in S_1, \; v \in V_1\\
\ell_2.\pput \; s \; v& \text{if } s \in S_2, \; v \in V_2\\
\ell_1.\create \; v & \text{if } s \in S_2, \; v \in V_1\\
\ell_2.\create \; v & \text{if } s \in S_1, \; v \in V_2\\
\end{cases}\\
(\ell_1 \sqcup \ell_2).\create \; v &=
\begin{cases}
\ell_1.\create \; v & \text{if } v \in V_1\\
\ell_2.\create \; v & \text{if } v \in V_2\\
\end{cases}
\end{align*}

\begin{claim}
If $\ell_1 : S_1  \xLeftrightarrow{\delta_{S_1}, \partial (S_1, V_1),
\delta_{V_1}} V_1$ and $\ell_2 : S_2 \xLeftrightarrow{\delta_{S_2}, \partial
(S_2, V_2), \delta_{V_2}} V_2$, then \\
$\ell_1 \sqcup \ell_2 : S_1 \sqcup S_2 \xLeftrightarrow{\delta_{S_1} \sqcup
\delta_{S_2}, \partial (S_1, V_1) \sqcup \partial (S_2, V_2) \sqcup (\ell_1
\times \ell_2), \delta_{V_1} \sqcup \delta_{V_2}} V_1 \sqcup V_2$
\end{claim}
\begin{proof}
We need to show that
$$(\partial (S_1, V_1) \sqcup \partial (S_2, V_2)).\get \;
(\delta((\ell_1 \sqcup \ell_2).\pput \; s \; v, s)) = \delta(v, (\ell_1 \sqcup
\ell_2).\get \; s)$$

The result holds inductively if $s \in S_1$ and $v \in V_1$ or $s \in S_2$ and
$v \in V_2$. Otherwise, assume without loss of generality
that $s \in S_1$ and $v \in V_2$. Then
\begin{align*}
(\partial (S_1, V_1) \sqcup \partial (S_2, V_2)).\get \;
(\delta((\ell_1 \sqcup \ell_2).\pput \; s \; v, s)) &=
(\partial (S_1, V_1) \sqcup \partial (S_2, V_2)).\get \; (\delta(\ell_2.\create
\; v, s))\\
&= (\ell_1 \times \ell_2).\get \; (s, \ell_2.\create \; v)\\
&= (\ell_1.\get \; s, \ell_2.\get \; (\ell_2.\create \; v))\\
&= (\ell_1.\get \; s, v)\\
&= \delta(v, \ell_1.\get \; s)\\
&= \delta(v, (\ell_1 \sqcup \ell_2).\get \; s)
\end{align*}
\end{proof}
\subsection{Positional Iteration}
Let $\ell : S \Leftrightarrow V$ be a lens and define the lens $\ell^*:
S^* \Leftrightarrow V^*$ by
\begin{align*}
\ell^* .\get \; (s_1, \ldots, s_n) &= (\ell.\get \; s_1) \cdot \ldots
\cdot (\ell.\get \; s_n)\\
\ell^*.\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n) &= s'_1 \cdot
\ldots \cdot s'_n\\
\text{ where } s'_i &= \begin{cases}
\ell.\pput \; s_i \; v_i & \text{if } 1 \leq i \leq \min\{m, n\}\\
\ell.\create \; v_i & \text{if } \min\{m, n\} < i \leq n
\end{cases}\\
\ell^*.\create \; (v_1, \ldots, v_n) &= (\ell.\create \; v_1) \cdot \ldots
\cdot (\ell.\create \; v_n)
\end{align*}

\begin{claim}\label{positionalisvalid}
If $\ell: S \xLeftrightarrow{\delta_S, \partial (S, V), \delta_V}
V$, then $\ell^*: S^* \xLeftrightarrow{\delta_{S}^*, (\partial (S, V) \sqcup
\ell)^*, \delta_{V}^*} V^*$.
\end{claim}
\begin{proof}
We need to show that
\begin{multline}
(\partial (S, V) \sqcup \ell)^*.\get \;
(\delta_S^*(\ell^*.\pput \; (s_1, \ldots, s_n) \; (v_1, \ldots,
v_m), (s_1, \ldots, s_n))) \\
= \delta_V^*((v_1, \ldots, v_m), \ell^*.\get \; (s_1, \ldots,
s_n))
\end{multline}
Let $1 \leq i \leq \min\{m, n\}$. Observe that $v_i = \ell.\get \; s_i$ if and
only if $\ell.\pput \; s_i \; v_i = s_i$, so $\delta_V^*$ omits
$\delta_V(v_i, \ell.\get \; s_i)$ in position $i$ if and only if
$\delta_S^*$ also omits $\delta_S(\ell.\pput \; s_i \; v_i, s_i)$ in
position $i$; if no omission occurs, then $\delta(S, V).\get \;
(\delta(\ell.\pput \; s_i \; v_i, s_i)) = \delta(v_i, \ell.\get \; s_i)$ since
$(\ell, \partial(S,V))$ is a delta lens. For $\min\{m,n\} < i \leq m$, then
$\delta_V^*$ adds $v_i$ in position $i$, and $\delta_S^*$ adds
$\ell.\create \; v_i$ in position $i$, which gives the desired result since
$$(\partial (S, V) \sqcup \ell).\get \; (\ell.\create \; v_i) = \ell.\get \;
(\ell.\create \; v_i)) = v_i$$
\end{proof}
\subsection{Matching Iteration}
\iffalse
When defining a $\match(\mathcal{M}, \delta)$ difference function, we often wish for
$\mathcal{M}$ to be an optimal set of matchings for some alignment strategy.

For instance, define the set of matchings $\mathcal{M}_{POS}$ by
$$\mathcal{M}_{POS} = \{\varphi_{(a_1, \ldots, a_m), (b_1, \ldots, b_n)}:[m]
\longrightarrow [n] \; | \; m \leq n, \varphi_{(a_1, \ldots, a_m), (b_1,
\ldots, b_n)}(i) = i\}$$ Then $\mathcal{M}_{POS}$ is an optimal set of
matchings for a positional alignment strategy.

For the non-crossing strategy, we want to our matchings to be increasing rather
than positional. That is, given a valid set of matchings $\mathcal{M}$, let
$\mathcal{M}_{a, b} = \varphi_{a, b} \in \mathcal{M}$, and for $a, a' \in P$
with $a=(a_1, \ldots, a_m)$ and $a'=(a'_1, \ldots, a'_n)$, define
$\mathcal{M}^{NC}_{a, b}$ by
$$\mathcal{M}^{NC}_{a, a'} = \argmin_{\substack{\mathcal{M}_{a, a'} \\
\mathcal{M}_{a, a'} \text{ is increasing}}} |\match(\mathcal{M}_{a, a'},
\delta)(a, a')|$$
There may be multiple matchings which for which the size of the delta is
minimal, in which case we pick the lexicographically smallest matching.
Then $\mathcal{M}^{NC}$ is an optimal set of matchings for a non-crossing
alignment strategy.

Finally, for a min-edit distance strategy, we simply remove the condition that
the matchings should be increasing. That is, define
$\mathcal{M}^{MED}_{a, b}$ by
$$\mathcal{M}^{MED}_{a, a'} = \argmin_{\mathcal{M}_{a, a'}}
|\match(\mathcal{M}_{a, a'}, \delta)(a, a')|$$
Then $\mathcal{M}_{MED}$ is an optimal set of matchings for a min-edit distance
alignment strategy.
\fi
Assume that $\ell: S \Leftrightarrow V$. Let $\mathcal{M}_V = \{\varphi_{x, y}
\; | \; x, y \in V^*\}$ be a set of alignments between $x$ and $y$ for each $x,
y \in V^*$. For each $s, s' \in S^*$, let $\varphi_{s, s'} = \varphi_{\ell.\get
\; s, \ell.\get \; s'}$. Define $\mathcal{M}_S$ by $\mathcal{M}_S =
\{\varphi_{s, s'} \; | \; s, s' \in S^*\}$. Define the lens
$\match(\mathcal{M}_V, \ell) : S^* \Leftrightarrow V^*$ by
\begin{align*}
\match(\mathcal{M}_V, \ell) .\get \; (s_1, \ldots, s_n) &= (\ell.\get \; s_1)
\cdot \ldots \cdot (\ell.\get \; s_n)\\
\match(\mathcal{M}_V, \ell) .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n)
&= (s'_1 , \ldots , s'_n)\\
\text{ where } s'_i &= \begin{cases}
\ell.\pput \; s_{\psi(i)} \; v_i & \text{if } i \in Dom(\psi)\\
\ell.\create \; v_i & \text{if } i \not \in Dom(\psi)
\end{cases}\\
\text{ and } \psi &= \varphi_{(v_1, \ldots, v_n), ((\ell.\get \; s_1), \ldots,
(\ell.\get \; s_m))}\\
\match(\mathcal{M}_V, \ell).\create \; (v_1, \ldots, v_n) &= (\ell.\create \;
v_1) \cdot \ldots \cdot (\ell.\create \; v_n)
\end{align*}
Assume that 
\begin{claim}
If $\ell: S \xLeftrightarrow{\delta_{S}, \partial (S, V), \delta_{V}}
V$, then $\match(\mathcal{M}_V, \ell) : S^*
\xLeftrightarrow{\match(\mathcal{M}_S, \delta_{S}), \partial (S, V)^*,
\match(\mathcal{M}_V, \delta_V)} V^*$.
\end{claim}
\begin{proof}
Observe that the matching used to compute the delta on the source side is
the same as the matching used to compute the matching on the view side, hence we
may take the update lens to be position and proceed with the rest of the proof
as in \cref{positionalisvalid}.
\end{proof}
\subsection{Composition}
Let $\ell_1 : S \Leftrightarrow U$ and $\ell_2 : U \Leftrightarrow V$ be lenses.
Define the lens $\ell_2 \circ \ell_1 : S \Leftrightarrow V$ by
\begin{align*}
(\ell_2 \circ \ell_1).\get \; s &= \ell_2.\get \; (\ell_1.\get \; s)\\
(\ell_2 \circ \ell_1).\pput \; s \; v &= \ell_1.\pput \; s \; (\ell_2.\pput \;
(\ell_1.\get \; s) \; v) \\
(\ell_2 \circ \ell_1).\create \; v &= \ell_1.\create \; (\ell_2.\create \; v)
\end{align*}

\begin{claim}
If $\ell_1 : S_1  \xLeftrightarrow{\delta_{S_1}, \partial (S_1, V_1),
\delta_{V_1}} V_1$ and $\ell_2 : S_2 \xLeftrightarrow{\delta_{S_2}, \partial
(S_2, V_2), \delta_{V_2}} V_2$, then \\
$\ell_2 \circ \ell_1 :  S \xLeftrightarrow{\delta_S, \partial
(S_2, V_2) \circ \partial (S_1, V_1), \delta_V} V$.
\end{claim}
\begin{proof}
We need to show that
$$
(\partial (S_2, V_2) \circ \partial (S_1, V_1)).\get \; (\delta(\ell_2 \circ
\ell_1).\pput \; s \; v, s) = \delta(v, (\ell_2 \circ \ell_1).\get \; s)$$

This follows from the fact that
\begin{align*}
&(\partial (S_2, V_2) \circ \partial (S_1, V_1)).\get \; (\delta_S((\ell_2 \circ
\ell_1).\pput \; s \; v, s))\\
&=(\partial (S_2, V_2) \circ \partial (S_1, V_1)).\get \; (\delta_S(\ell_1.\pput
\; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; v), s))\\
&= (\partial (S_2, V_2)).\get \; ((\partial (S_1, V_1)).\get \;
(\delta_S(\ell_1.\pput \; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; v), s)))\\
&= (\partial (S_2, V_2)).\get \; (\delta_U(\ell_2.\pput \; (\ell_1.\get \; s)
\; v, \ell_1.\get \; s))\\
&= \delta_V(v, \ell_2.\get \; (\ell_1.\get \; s))\\
&= \delta_V(v, (\ell_2 \circ \ell_1).\get \; s)
\end{align*}
\end{proof}
\iffalse
\section{Symmetric Delta Lenses}
Assume that $(X, \partial X, \delta_X)$ and $(Y, \partial Y, \delta_Y)$
are difference structures. A symmetric delta lens is a pair $(\ell, \partial
\ell)$ where $\ell : X \Leftrightarrow Y$ is a symmetric lens, and
$\partial \ell : \partial X \Leftrightarrow \partial Y$ is a symmetric lens
such that
\begin{enumerate}
  \item
  if $\ell.putr(x, c) = (y, c)$ and $\ell.putl(y', c) = (x', c')$,
  then $\partial \ell.putl \; (\delta(y, y'), \delta(c, c')) = (\delta(x,
  x'), \delta(c, c'))$
  \item
  if $\ell.putl(y, c) = (x, c)$ and $\ell.putr(x', c) = (y', c')$, then
  $\partial \ell.putr \; (\delta(x, x'), \delta(c, c')) = (\delta(y, y'),
  \delta(c, c'))$
\end{enumerate}
\begin{fact}[Identity Lens]
Given any symmetric lens $\partial X : X \Leftrightarrow X$, $(id_X, \partial
X)$ is a symmetric delta lens
\end{fact}
\begin{fact}[Bijective Lenses]
Assume that $f : X \longrightarrow Y$ and $g : \partial X \longrightarrow
\partial Y$ are bijections. Assume that $g(\delta(x, x')) = \delta(f(x),
f(x'))$. Then $(bij_f, bij_g)$ is a symmetric delta lens.
\end{fact}
\begin{fact}[Lens Composition]
If $(\ell_1, \partial \ell_1) : (X, \partial X, \delta_X) \Leftrightarrow (Y,
\partial Y, \delta_Y)$ and $(\ell_2, \partial \ell_2): (Y, \partial Y,
\delta_Y) \Leftrightarrow (Z, \partial Z, \delta_Z)$ are symmetric delta
lenses, then $(\ell_1;\ell_2, \partial \ell_1;\partial \ell_2)$ is a symmetric
delta lens.
\end{fact}
\begin{fact}[Dual of a Lens]
If $(\ell, \partial \ell)$ is symmetric delta lens, then $(\ell^{op}, \partial
\ell^{op})$ is a symmetric delta lens.
\end{fact}
\begin{fact}[Terminal Lens]
$(term_x, term_{\delta(x, x)})$ is a symmetric delta lens
\end{fact}
\begin{corollary}[Disconnect Lens]
If $x \in X$ and $y \in Y$, then $(term_x;term_y^{op},
term_{\delta(x, x)};term_{\delta(y, y)}^{op})$ is a symmetric delta lens.
\end{corollary}
\begin{fact}[Tensor Product Lens]
If $(\ell_1, \partial \ell_1)$ and $(\ell_2, \partial \ell_2)$ are symmetric
delta lenses, then $(\ell_1 \otimes \ell_2, \partial \ell_1 \otimes \partial
\ell_2)$ is a symmetric delta lens.
\end{fact}
\begin{corollary}[Projection Lens]
Projection lenses are symmetric delta lenses.
\end{corollary}
\begin{fact}[Forgetful Tensor Sum Lens]
If $(\ell_1, \partial \ell_1)$ and $(\ell_2, \partial \ell_2)$ are symmetric
delta lenses, then $(\ell_1 \oplus^f \ell_2, \partial \ell_1 \oplus^f \partial
\ell_2 \oplus (\ell_1 \otimes \ell_2))$ is a symmetric delta lens.
\end{fact}
\begin{fact}[Forgetful List Mapping Lens]
If $(\ell, \partial \ell)$ is a symmetric delta lens, then $(map^f(\ell)
\ell_2, map^f(\partial \ell))$ is a symmetric delta lens.
\end{fact}
\begin{remark}[Retentive Lenses]
The retentive tensor sum and list mapping lenses are in general not symmetric
delta lenses with the obvious definitions of difference functions.
\end{remark}
\fi
\section{Relational Lenses}
We define a {\em relation} $R$ of type $((T_1, \ldots, T_n), P, F)$ to be a
multiset $R$ where $T_i, P$ are sets satisfying $R \subseteq P \subseteq T_1
\times \ldots \times T_n$, $F \subseteq [n] \times [n]$, and if $i, j \in F$,
then for all $r, r' \in R$, if $r_i = r'_i$ then $r_j = r'_j$. We interpret the
$T_i$ to be a set of {\em attributes}, $R$ to be a set of {\em records} with
$dom(r) = (T_1, \ldots, T_n)$ for all $r \in R$, $P$ to be a {\em predicate}
such that the records in $R$ satisfy $P$, and $F$ to be a set of {\em
functional dependencies} on the attributes of the records in $R$.

Let $G(F)$ be the graph induced by $F$ on $T_1, \ldots, T_n$ i.e. there is an
edge from $T_i$ to $T_j$ if and only if $(i, j) \in F$. We say that $F$ is
in {\em tree form} if $G(F)$ is a DAG and every vertex has indegree at most 1;
following \cite{bohannon2006relational},we will always assume that $F$ is in
tree form from here on. Observe that every record $r$ is completely determined
by nodes of indegree 0 in $G(F)$: if $Roots(G(F)) = \{T_{i_1}, \ldots,
T_{i_k}\}$ is the set of nodes of indegree 0 and outdegree greater than 0 in
$G(F)$ and $Isolated(G(F)) = \{T_{j_1}, \ldots, T_{j_\ell}\}$ is the set of
nodes of indegree 0 and outdegree 0 in $G(F)$, then for any $r = (r_1, \ldots,
r_n)\in R$, then the map $((r_{i_1}, \ldots, r_{i_k}), (r_{j_1}, \ldots,
r_{j_\ell}))$ is a bijection since the $r_{i_1}, \ldots, r_{i_k}$ determine all
nodes of indegree greater than 0, which leaves just the nodes $r_{j_1}, \ldots,
r_{j_\ell}$.

In order to define lenses in relations, we first need to describe data
structures representing the relation; lenses will operate on the data
structures rather than on relations. Let $Outputs(F) = G(F) \setminus
(Roots(G(F)) \cup Isolated(G(F)))$. We choose to represent a relation with the
following data:
\begin{enumerate}
  \item
  A partial map $data : (T_{i_1} \times \ldots \times T_{i_k})
  \longrightarrow 2^{(T_{j_1} \times \ldots \times T_{j_{\ell}})}$ such that
  there exists a record $r \in R$ with $r[T_{i_1}, \ldots, T_{i_k}] =
  (t_{i_1}, \ldots, t_{i_k})$ if and only if
  $$data(t_{i_1}, \ldots, t_{i_k}) =  \{\{r[T_{j_1}, \ldots, T_{j_{\ell}}] \in
  R \; | \; r_{i_m} = t_{i_m} \text{ for }1 \leq m \leq k\}\}$$
  \item
  A map $dependencies : outputs(F) \longrightarrow \cup_{i,j} 2^{T_i \times
  T_j}$ such that $(p, q) \in dependencies(j)$ if and only if there exists a
  record $r \in R$ such that $r[T_i, T_j] = (p, q)$, where $i$ is such that
  $T_i \longrightarrow T_j \in F$ ($i$ is unique as $F$ is assumed to be in tree
  form).
\end{enumerate}
\subsection{Select}
We say that a predicate $P$ ignores a set of attributes $X$ if for all $r, r'
\in R$, if $r[dom(R) \setminus X] = r'[dom(R) \setminus X]$ implies that $r
\in P$ if and only if $r' \in P$.

The {\em select} lens on $R$ evaluates to a relation which is the same as $R$
but which satisfies a predicate $P$. The inferrence rule for deriving the select
lens is the following:
\begin{prooftree}
\AxiomC{$R : ((T_1, \ldots, T_2), Q, F)$}
\AxiomC{$S : ((T_1, \ldots, T_2), P \cap Q, F)$}
\AxiomC{$F$ is in tree form}
\AxiomC{$P$ ignores $Outputs(F)$}
\QuaternaryInfC{$(select \; from \; R\; where \; P \; as \; S) : R
\Leftrightarrow S$}
\end{prooftree}
In the get direction, the select lens simply restricts $R$ to $P$. Now consider
$put \; s \; r$. The select lens proceeds as follows:
\begin{enumerate}
  \item
  First, it computes $(data_s, dependencies_s)$ and $(data_r, dependencies_r)$.
  \item
  Next, it updates $dependencies_r$ using $dependencies_s$ to get
  $dependencies_r'$.
  \item
  Next it filters $data_r$ to contain only records not in $P$ to get $data_r'$.
  \item
  Next, it computes $(data_r' \cup data_s, dependencies_r')$.
  \item
  Finally, it deletes the records from $data_r' \cup data_s$ that would cause
  PUT-GET to fail.
\end{enumerate}
\section{Matching Lenses}
\begin{definition}
A {\em basic lens} $\ell : S \xLeftrightarrow{C} V$ is a 4-tuple of functions
$\ell.\get : S \longrightarrow V$, $\ell.\res : S \longrightarrow C$,
$\ell.\pput : V \longrightarrow C \longrightarrow S$ and $\ell.\create : V
\longrightarrow S$ satisfying the following properties:
\begin{align*}
\ell.\get \; (\ell.\pput \; v \; c) &= v \tag{PUTGET}\\
\ell.\pput \; (\ell.\get \; s) \; (\ell.\res \; s) &= s \tag{GETPUT}
\end{align*}
\end{definition}
Idea: First remove matched parts to leave unmatched skeleton. Next remove
unmatched parts to leave matched skeleton and then flatten.
%% Acknowledgments
\begin{acks}                            %% acks environment is optional
%% contents suppressed with 'anonymous'
%% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
%% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
%% acknowledge financial support and will be used by metadata
%% extraction tools.
This material is based upon work supported by the
\grantsponsor{GS100000001}{National Science
Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
No.~\grantnum{GS100000001}{nnnnnnn} and Grant
No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
conclusions or recommendations expressed in this material are those
of the author and do not necessarily reflect the views of the
National Science Foundation.
\end{acks}
\fi
\bibliographystyle{plain}
\bibliography{local}
%% Appendix
%%\appendix
%%\section{Appendix}

\end{document}

### 2018-06-15 papers/DifferenceLenses/main.tex
Made some changes

@@
 As a result, the transformations derived using these languages can lead to unexpected results.
-In this paper we present a formalization of bx transformations that enables a programmer to express non-trivial guarantees on update propagation.
-We will then use this formalization of bx transformations to describe a compositional language of {\em difference lenses}: each difference lens gives non-trivial guarantees on how it propagates non-trivial updates.
-We will then embed various lens programming languages into the language of difference lenses, and describe non-trivial guarantees that these languages provide beyond consistency restoration.
+In this paper we present a formalization for bx transformations that requires a bx to provide guarantees on how it propagates updates.
+We will then describe a specific bx language which we call {\em difference lenses}; this language is based on the {\em basic lenses} defined in the bx programming language {\em Boomerang}, but each difference lens comes with explicit guarantees on update propagation.
+We will then express various lens programming languages as difference lenses, and hence describe non-trivial guarantees on update propagation that these languages provide as well.
 \end{abstract}
@@
 $$\ell.\pput \; [``Ada \; Lovelace", \; ``Alan \; Turing"] \; [``Turing", \; ``Lovelace"] = [``Ada \; Turing", \; ``Alan \; Lovelace"]$$
-Consider another example where a source consists of a list of names and a number while the target consists of a list of names.
-Assume that $\ell$ is a bx which in the \get direction simply discards the number:
-$$\ell.\get \; ([``Ada \; Lovelace", \; ``Alan \; Turing"], \; 2)  = [``Lovelace", \; ``Turing"]$$
-A user might expect that $\ell$ should always restore the number in the \pput direction so that the following result occurs:
-$$\ell.\pput \; ([``Ada \; Lovelace", \; ``Alan \; Turing"], \; 2) \; [``Lovelace"] = ([``Ada \; Lovelace"], \; 2)$$
-However, perhaps the intended semantics of the source is that the number which is the second component of a source denotes the length of the list in the first component of the source so that an application of \pput leads to the following result:
-$$\ell.\pput \; ([``Ada \; Lovelace", \; ``Alan \; Turing"], \; 2) \; [``Lovelace"] = ([``Ada \; Lovelace"], \; 1)$$
-In both of these examples, there are multiple ways to restore consistency; that is, there are multiple ways to define $\ell.\pput \; s \; v$ in such a way that $\ell.\get \; (\ell.\pput \; s \; v) = v$.
-Moreover, the source that results from a \pput application may be vastly different from what a user expects, or may break an invariant that is not explicitly maintained by the source.
+Consider another example where both a source and a target consists of a list where every element on the list is a name and a year.
+Assume that $\ell$ is a bx which in the \get direction returns all elements of the source list that have year greater than or equal to 1900:
+$$\ell.\get \; [``Ada \; Lovelace, \; 1815", \; ``Alan \; Turing, \; 1912"] = [``Alan \; Turing, \; 1912"]$$
+Suppose that the entry $``Ada \; Lovelace, \; 1915"$ is then added to the target list. Should $\ell.\pput$ replace the old entry containing the name $``Ada \; Lovelace"$,
+\begin{align*}
+&\ell.\pput \; [``Ada \; Lovelace, \; 1815", \; ``Alan \; Turing, \; 1912"] \; [``Alan \; Turing, \; 1912", \; ``Ada \; Lovelace, \; 1915"] \\
+&= [``Ada \; Lovelace, \; 1915", \; ``Alan \; Turing, \; 1912"]
+\end{align*}
+or should the \pput application retain both versions of $``Ada \; Lovelace"$?
+\begin{align*}
+&\ell.\pput \; [``Ada \; Lovelace, \; 1815", \; ``Alan \; Turing, \; 1912"] \; [``Alan \; Turing, \; 1912", \; ``Ada \; Lovelace, \; 1915"] \\
+&= [``Ada \; Lovelace, \; 1815", \; ``Alan \; Turing, \; 1912", \; ``Ada \; Lovelace, \; 1915"]
+\end{align*}
+In both of these examples, there are multiple ways to restore consistency---that is, there are multiple ways to define $\ell.\pput \; s \; v$ in such a way that $\ell.\get \; (\ell.\pput \; s \; v) = v$---and the results from the different \pput applications may be vastly different from what a user expects.
 
@@
 
-We take the approach that a bx should not have to explicitly propagate updates when going from the source to the target and vice versa.
-Instead, the programmer should be able provide guarantees on how a bx propagates updates.
-For instance, in the second example, let $\delta_S$ be a function which takes two source elements $(\ell_1, n_1)$ and $(\ell_2, n_2)$ and returns the number $n_1 - n_2$, and let $\delta_T$ be a function which takes two target lists $\ell_1$ and $\ell_2$ and returns the number $length(\ell_1) - \length(\ell_2)$.
-Then the programmer should be able to express the fact that the bx restores the length of the list in the backward direction, which is expressed using the following equation:
-$$\delta_S((\ell_1, n_1), \ell.\pput \; (\ell_1, n_1) \; \ell_2) = \delta_V(\ell.\get \; (\ell_1, n_1), \ell_2) $$
-In this paper, we introduce a formalization of bx transformations that enables a programmer to express guarantees such as the one above and show how our formalism can be used to characterize update propagation behaviour of bidirectional transformations.
-
-The major technical contributions that we will make in this paper are the following:
+We take the approach that a bidirectional programming language should have explicit guarantees on how programs written in that language propagate updates; ideally these guarantees should be strictly stronger than hippocraticness.
+This way a programmer who derives bx transformations in this language has a clearer understanding of the behaviour of the programs that she writes.
+
+In this paper we will show how this can be done, by describing a formalization of bx transformations that explicitly accounts for update propagation, and presenting a series of examples showing how this principle can be applied to existing bx programming languages.
+More specifically, the major technical contributions that we will make in this paper are the following:
 \begin{enumerate}
 \item
-  We present a formalization of bx transformations that enables a programmer to express non-trivial update propagation behaviour.
-  \item
-  We use our formalization to develop a compositional language of {\em difference lenses} based on the {\em basic lenses} implemented in the lens programming language Boomerang \cite{foster2009bidirectional}, and describe non-trivial update propagation guarantees that these lenses provide.
+We present a formalization for bx transformations that requires a bx to provide guarantees on how it propagates updates. 
+\item
+We describe a specific bx language which we call {\em difference lenses}; this language is based on the {\em basic lenses} defined in the bx programming language {\em Boomerang}\cite{foster2009bidirectional}, but each difference lens comes with explicit guarantees on update propagation.
+As a result of this work, we will describe non-trivial update propagation properties that Boomerang's basic lenses provide.
   \item

### 2018-06-22 papers/DifferenceLenses/main.tex
Added technical material

@@
 \newcommand{\match}{\ensuremath{\kw{match}}\xspace}
-
-\DeclareMathOperator*{\argmin}{arg\,min} % thin space, limits underneath in
-\DeclareMathOperator*{\argmax}{argmax} % no space, limits underneath in
+\newcommand{\filter}{\ensuremath{\kw{filter}}\xspace}
+\newcommand{\strfilter}{\ensuremath{\kw{str\_filter}}\xspace}
+\newcommand{\strunfilter}{\ensuremath{\kw{str\_unfilter}}\xspace}
+\newcommand{\mergelens}{\ensuremath{\kw{merge}}\xspace}
+
+\DeclareMathOperator*{\argmin}{argmin}
+\DeclareMathOperator*{\argmax}{argmax}
 
@@
 Consider another example where both a source and a target consists of a list where every element on the list is a name and a year.
-Assume that $\ell$ is a bx which in the \get direction returns all elements of the source list that have year greater than or equal to 1900:
-$$\ell.\get \; [``Ada \; Lovelace, \; 1815", \; ``Alan \; Turing, \; 1912"] = [``Alan \; Turing, \; 1912"]$$
-Suppose that the entry $``Ada \; Lovelace, \; 1915"$ is then added to the target list. Should $\ell.\pput$ replace the old entry containing the name $``Ada \; Lovelace"$,
-\begin{align*}
-&\ell.\pput \; [``Ada \; Lovelace, \; 1815", \; ``Alan \; Turing, \; 1912"] \; [``Alan \; Turing, \; 1912", \; ``Ada \; Lovelace, \; 1915"] \\
-&= [``Ada \; Lovelace, \; 1915", \; ``Alan \; Turing, \; 1912"]
+Assume that $\ell$ is a bx which in the \get direction returns all elements of the source list that have year less than 1900:
+\begin{align*}
+&\ell.\get \; [``Ada \; Lovelace, \; 1915", \; ``Alan \; Turing, \; 1912", \; ``George \; Boole, \; 1815"] = [``George \; Boole, \; 1815"]
+\end{align*}
+Suppose that the entry $``Ada \; Lovelace, \; 1815"$ is then added to the target list. Should $\ell.\pput$ replace the old entry containing the name $``Ada \; Lovelace"$,
+\begin{align*}
+&\ell.\pput \; [``Ada \; Lovelace, \; 1915", \; ``Alan \; Turing, \; 1912", \; ``George \; Boole, \; 1815"]\\
+&[``Alan \; Turing, \; 1912", \; ``Ada \; Lovelace, \; 1815"] \\
+&= [``Ada \; Lovelace, \; 1815", \; ``Alan \; Turing, \; 1912", \; ``George \; Boole, \; 1815"]
 \end{align*}
@@
 \begin{align*}
-&\ell.\pput \; [``Ada \; Lovelace, \; 1815", \; ``Alan \; Turing, \; 1912"] \; [``Alan \; Turing, \; 1912", \; ``Ada \; Lovelace, \; 1915"] \\
-&= [``Ada \; Lovelace, \; 1815", \; ``Alan \; Turing, \; 1912", \; ``Ada \; Lovelace, \; 1915"]
+&\ell.\pput \; [``Ada \; Lovelace, \; 1915", \; ``Alan \; Turing, \; 1912", \; ``George \; Boole, \; 1815"] \\
+& [``Alan \; Turing, \; 1912", \; ``Ada \; Lovelace, \; 1815"]\\
+&= [``Ada \; Lovelace, \; 1915", \; ``Alan \; Turing, \; 1912", \; ``George \; Boole, \; 1815", \; ``Ada \; Lovelace, \; 1815"]
 \end{align*}
@@
 
-We take the approach that a bidirectional programming language should have explicit guarantees on how programs written in that language propagate updates; ideally these guarantees should be strictly stronger than hippocraticness.
-This way a programmer who derives bx transformations in this language has a clearer understanding of the behaviour of the programs that she writes.
-
-In this paper we will show how this can be done, by describing a formalization of bx transformations that explicitly accounts for update propagation, and presenting a series of examples showing how this principle can be applied to existing bx programming languages.
-More specifically, the major technical contributions that we will make in this paper are the following:
+Each one of these solutions has significant shortcomings.
+While matching lenses successfully address the issue of alignment, they do not address the general issue of correct update propagation.
+Edit lenses and delta lenses all assume that the lens propagates a specific set of updates explicitly, but this is unrealistic since since in general users do not update data using a small predetermined langauge of edits.
+Instead users edit the data in any way, for example using text editors or IDEs, and present the updatad data to the bx wholesale.
+Finally, least change lenses are in general computationally intractable \cite{buneman2002propagation} and also lack a notion of composition---that is, the composition of two deterministic least change lenses is not always a least change lens\cite{macedo2013composing}.
+
+We take the approach that instead of having a bx manually propagate a fixed set of edits, a bidirectional programming language should have explicit guarantees on how programs written in that language propagate updates; ideally these guarantees should be strictly stronger than hippocraticness.
+Consider again the first example that we presented, which is an instance of the alignment problem.
+In that example, the list
+$[``Lovelace", \; ``Turing"]$ was updated to the list $[``Turing", \; ``Lovelace"]$.
+The programmer may wish to consider these two lists equivalent or not: if she does, then she should use a min-edit distance ``metric'' and demand that the source side respects this metric.
+On the other hand, if she wishes to use a positional metric to compare these two lists, then she should similarly demand that the source side respects this metric.
+In each case, the programmer does not care how the bx restores consistency; she cares only that the bx restores consistency and respects the appropriate metric.
+
+In this paper, we present a general approach for describing update propapagtion guarantees for a bx without requiring that the bx manually handle updates.
+In particular, we will enhance a standard definition of bx transformations that only takes into account a consistency relation with a ``difference'' relation that describes a relationship between source updates and target updates.
+With this new formalization, a well behaved bx will not only be hippocratic, but will also be required to preserve the difference relation that is part of the bx.
+
+We will then use this formalization to describe a particular language of bx transformations which we call {\em difference lenses}.
+Difference lenses are based on the {\em basic lenses} defined in the lens programming language {\em Boomerang} \cite{foster2009bidirectional}, but come with the guarantee that the lens respects ``differences'' on the source and view side of a lens.
+Intuitively, this means that if one starts with a source, maps it into a view, edits the view and then maps the updated view back into the old source, then the difference between the new and old source is at most the difference between the old and new view.
+Difference lenses will have combinators that deal with local alignment; in particular, difference lenses have positional, non-crossing, and min-edit distance iteration combinators that preserve positional, non-crossing, and min-edit distance differences respectively.
+We will then build on difference lenses to describe a language of {\em matching difference lenses} that deal with global alignment.
+
+\sam{TODO: Talk about relational lenses.}
+
+The major technical contributions that we will make in this paper are the following:
 \begin{enumerate}
@@
 \item
-We describe a specific bx language which we call {\em difference lenses}; this language is based on the {\em basic lenses} defined in the bx programming language {\em Boomerang}\cite{foster2009bidirectional}, but each difference lens comes with explicit guarantees on update propagation.
+We describe a specific bx language which we call {\em difference lenses}; this language is based on the {\em basic lenses} defined in the lens programming language {\em Boomerang}\cite{foster2009bidirectional}, but each difference lens comes with explicit guarantees on update propagation.
 As a result of this work, we will describe non-trivial update propagation properties that Boomerang's basic lenses provide.
@@
 \end{enumerate}
-\iffalse
- we describe a compositional language of lenses called {\em difference lenses}, where each difference $\ell$ satisfies the following equation,
-\begin{equation}\label{difflenslaw}
-\delta_S(s, \ell.\pput \; s \; v) = \delta_V(\ell.\get \; s, v)
-\end{equation}
-for appropriate real-valued $\delta_S : S \times S \longrightarrow \mathbb{R}^{\geq 0}$ and $\delta_S : S \times S \longrightarrow \mathbb{R}^{\geq 0}$.
-
-\cref{difflenslaw}, which already appears in Gibbons et.
-al.
-\cite{gibbons2017principles} as the principle of {\em Weak Least Surprise}, is simple and intuitive but presents a challenge when trying to develop a compositional lens langauge that satisfies it.
-For example, assume that a source consists of a name-and-date pairs while a view consists of a name.
-Assume that a lens $\ell$ simply discards dates in the \get direction:
-\begin{lstlisting}[mathescape=true]
-$\ell.\get$ "Ada Lovelace, 12-10-1815" = "Ada Lovelace"
-\end{lstlisting}
-Consider the iteration lens $\ell^*$, where the source is now a list of name-and-date pairs while a view consists of a list of names.
-Whenever the length of a view list is larger than the length of a source list, then a \pput application of $\ell^*$ must initialize a source from a view from scratch:
-\begin{lstlisting}[mathescape=true]
-$\ell.\pput$ [] ["Ada Lovelace"] = ["Ada Lovelace, 00-00-0000"]
-\end{lstlisting}
-Now if we are to define $\delta_S^*$ and $\delta_V^*$ inductively on source and view lists respectively in such a way that \cref{difflenslaw} holds of $\ell^*$, then the following equation would have to hold:
-\begin{lstlisting}[mathescape=true]
-$\delta_S^*$(["Ada Lovelace, 00-00-0000"], []) = $\delta_V^*$(["Ada Lovelace"], [])
-\end{lstlisting}
-In a case such as this when a source is {\em created} from a view as opposed to arising from a view being \pput into an existing source, the source delta will be defined so as to defer to the view delta by mapping the created source back into the view; in particular, for the lens $\ell^*$ in the example, then we have
-\begin{lstlisting}[mathescape=true]
-$\delta_S^*$(["Ada Lovelace, 00-00-0000"], [])
-= $\delta_V^*$([$\ell.\get$ "Ada Lovelace, 00-00-0000"], [])
-= $\delta_V^*$(["Ada Lovelace"], [])
-\end{lstlisting}
-
-By now there exist many programming languages that enable one to define {\em
-bidirectional transformations}, programs that can be run in two
-directions. Each of these languages provides guarantees on how the two
-transformations interact e.g. that they are bijective, that a round-trip on
-one or both directions does not cause non-trivial updates etc. Many of these
-guarantees have to do with restoring or preserving consistency. To formalize
-this statement, we describe a widely used formulation of bidirectional
-transformations due to Stevens \cite{stevens2010bidirectional} but originally
-articulated in a more general form by Meertens \cite{meertens1998designing}:
-\begin{definition}
-A bidirectional transformation (bx) $R: M \Leftrightarrow N$ comprises of a
-consistency relation $R \subseteq M \times N$ and restorers $\overrightarrow{R}
-: M \times N \rightarrow N$ and $\overleftarrow{R} : M \times N \longrightarrow
-N$.
-\end{definition}
-\begin{definition}
-A bx $R : M \Leftrightarrow N$ is {\em correct} if the consistency restorers do
-restore consistency: that is, for all $m \in M, n \in N$, we have $R(m,
-\overrightarrow{R}(m, n))$, and dually. $R$ is {\em hippocratic} if the
-consistency restorers make no change to an already-consistent pair of models:
-that is, for all $m \in M$, $n \in N$, if $R(m, n)$ then $\overrightarrow{R}(m,
-n) = n$ and dually. $R$ is {\em well behaved} if it is correct and hippocratic.
-\end{definition}
-
-Many bidirectional programming languages thus enable one to define
-well behaved bx's: consistency restoration is guaranteed by correctness of the
-bx while consistency preservation is guaranteed by the hippocraticness of the
-bx. This paper is primarily concerned with well behaved bx's where
-$\overrightarrow{R}$, the \get function of the bx, ignores it's $N$-argument and
-where the consistency relation $R$ is defined by $R(m, \overrightarrow{R}(m))$
-for all $m \in M$--- these transformations are the well known and widely
-understood {\em lenses} of Foster et al \cite{foster2007combinators}--- though
-the ideas we develop could be used in other bidirectional settings as well.
-
-When the consistency relation of a bx $R$ is very fine, i.e. when each $m \in M$
-is consistent with only a few $n \in N$ and vice versa, then a well behaved lens
-gives fairly strong guarantees on $\overleftarrow{R}$ and $\overrightarrow{R}$.
-However, there are many cases that arise in practice where there are many
-different ways to restore consistency; when this happens, well-behavedness may
-be too weak to rule out ``bad'' behaviour or to clarify the behaviour of a bx in
-subtle cases. Various solutions have been proposed to address this issue, for
-example {\em matching lenses} that deal with the issue of alignment
-\cite{barbosa2010matching}, {\em edit lenses} \cite{hofmann2012edit} and {\em
-delta lenses} \cite{diskin2011asymmetric,diskin2011state,pacheco2012delta}
-that explicitly propagate edits through the lens, and {\em least change
-lenses} that require that the source value resulting from a \pput application
-is not only acceptable, but also one of closest to the original $s$ among the
-sources that share the same view $v$ \cite{macedo2013composing}.
-
-This paper takes a different approach to specifying update propagation by
-identifying a condition called {\em $\partial D$ preservation} which we describe
-below:
-\begin{definition}
-Let $R : M \Leftrightarrow N$ be a bidirectional transformation and $\partial D
-\subseteq (M \times M) \times (N \times N)$. Then $R$ {\em preserves $\partial
-D$} if for all $m \in M$, $n, n' \in N$ , if $R(m, n)$ then $\partial D((m,
-\overleftarrow{R}(m, n')), (\overrightarrow{R}(m, n), n'))$. $R$ {\em
-strongly preserves $\partial D$} if the dual condition also holds.
-\end{definition}
-\noindent The idea behind $\partial D$ preservation is that if $m$ and $n$ are
-already consistent and $\overrightarrow{R}(m,n)$ is updated to produce $n'$,
-then the difference between $\overrightarrow{R}(m,n)$ and $n'$ is related
-to the difference between $m$ and $\overleftarrow{R}(m,n')$. $R$
-strongly preserves $\partial D$ if it preserves $\partial D$ in both directions.
-
-For lenses, $\partial D$ preservation says that difference between $\ell.\get \;
-s$ and $v$ is related to the difference between $s$ and $\ell.\pput \; s \;
-v$. One way to formalize this idea is to model the difference between two sources $s, s'
-\in S$ and two views $v, v' \in V$ as functions $\delta_S : S \times S
-\longrightarrow \partial S$ and $\delta_V : V \times V \longrightarrow \partial
-V$ for some sets $\partial S$ and $\partial V$, and to capture the
-relationship between $\delta_S$ and $\delta_V$ by a lens. In
-particular, given a lens $\partial(S, V) : \partial S \Leftrightarrow \partial
-V$ we will ask if the following equation holds of $\ell$:
-\begin{equation}\label{difflenslaw}
-\partial(S, V).\get \; (\delta_S(s, \ell.\pput \; s \; v)) = \delta_V(\ell.\get
-\; s, v)
-\end{equation}
-If $\ell : S \Leftrightarrow V$ satisfies \cref{difflenslaw} we say that 
-$\ell$ is a {\em difference  lens} with respect to $(\delta_S, 
-\partial(S, V), \delta_V)$, which we denote with the typing relation 
-$\ell : S \xLeftrightarrow{\delta_S, \partial(S, V), \delta_V} V$.
-
-For example, consider two versions of an iteration lens $\ell_1 : S^*
-\Leftrightarrow V^*$ and $\ell_2 : S^* \Leftrightarrow V^*$ defined inductively
-from a lens $\ell : S \Leftrightarrow V$. Both of these lens have the same get
-behaviour: given a list $s_1, \ldots, s_n$ of sources, they both produce
-$\ell.\get \; s_1, \ldots, \ell.\get \; s_n$. The \pput behaviour of the two
-lenses is different though: $\ell_1$ simply aligns views to sources
-positionally, while $\ell_2$ computes the optimal alignment on the view side and
-uses this alignment to put views back into sources.
-
-Let \lstinline|name = [A-Z] . [a-z]*|, let the source $S$ equal 
-\lstinline|name . " " . name| and the view $V$ equal \lstinline|name|.
-Let $\ell : S \Leftrightarrow V$ be the lens
-\lstinline|del name . del " " . copy name| which given a pair of names produces
-the second name in the \get direction, and restores the first name in the \pput
-direction:
-$$\ell.\get \; ``Ada \; Turing" = ``Turing" \text{ and }\; \ell.\pput \; ``Ada
-\; Turing" \; ``Lovelace" = ``Ada \; Lovelace"$$
-Define the
-function $\delta_V : V \times V \longrightarrow \mathbb{N}^{\geq 0}$ by 
-$$\delta_V(n, n') = 
-\begin{cases}
-0 & \text{if }n=n'\\
-1 & \text{otherwise}
-\end{cases}$$
-Define $\delta_S : S \times S \longrightarrow \mathbb{N}^{\geq 0}$ by
-$$\delta_S(m \cdot `` \quad " \cdot n, m' \cdot `` \quad " \cdot n') =
-\begin{cases}
-0 & \text{if } m = m', n = n'\\
-1 & \text{if } m = m', n \neq n'\\
-1 & \text{if } m \neq m', n = n'\\
-2 & \text{if } m \neq m', n \neq n'
-\end{cases}$$
-Let the lens $\partial(S, V) : \mathbb{N}^{\geq 0} \Leftrightarrow
-\mathbb{N}^{\geq 0}$ be given by $\partial(S, V) = \ccopy(\mathbb{N}^{\geq
-0})$. Then $\ell : S \xLeftrightarrow{\delta_S, \partial(S, V), \delta_V} V$.
-
-Now let $xs, ys$ be two lists of single names, with $length(xs) \geq length(ys)$
-Define $\delta_V^* : V^* \times V^* \longrightarrow \mathbb{Z}^* \times V^*$ on $xs$ and
-$ys$ by aligning elements of $xs$ maximally with elements of $ys$ and then using $\delta_V$
-on these pairs to produce a list $ds$ of numbers. In this process, some elements of
-$ys$ will not be aligned if $length(xs) < length(ys)$, in which case we simply collect
-these in another list of elements of $V$. For instance,
-$$\delta_V^*([``Lovelace", ``Kleene", ``Turing"], [``Turing", ``Lovelace"]) = ([0,0], [``Kleene"])$$
-since we align \lstinline|"Lovelace"| to \lstinline|"Lovelace"| and 
-\lstinline|"Turing"| to \lstinline|"Turing"| for a difference of 0 in each case, which leaves
-\lstinline|"Kleene"| unaligned.
-
-On the source side, let $ss, ts$ be two lists of two names with $length(ss) \geq length(ts)$.
-Define $\delta_{S, \ell}^* : S^* \times S^* \longrightarrow \mathbb{Z}^* \times S^*$ by taking the 
-alignment used to compute $\delta_V^*(\ell^*.\get \; ss, \ell^*.\get \; ts)$ and then using that
-alignment in the same manner that $\delta_V^*$ does. For instance
-\begin{align*}
-&\delta_{S, \ell}^*([``Ada \; Lovelace", ``Stephen \; Kleene", ``Alan \; Turing"], 
-[``Alan \; Turing", ``Ada \; Lovelace"])\\
-&= ([0,0], [``Stephen \; Kleene"])
-\end{align*}
-Recall that the \pput behaviour of $\ell_1$ simply aligns views to sources
-positionally, while $\ell_2$ computes the optimal alignment on the view side and
-uses this alignment to put views back into sources. From this we can prove that
-$$\ell_2 : S^* \xLeftrightarrow{\delta_{S, \ell}^*, \partial(S, V)^* \times \ell^*, \delta_V^*} V^*$$
-On the other hand, the judgement $\ell_1 : S^* \xLeftrightarrow{\delta_{S, \ell}^*, 
-\partial(S, V)^* \times \ell^*, \delta_V^*} V^*$ does not holds since $\ell_1$ does not satisfy \cref{difflenslaw} :
-$$(\partial(S, V)^* \times \ell^*).\get \; (\delta_{S, \ell}^*(s, \ell.\pput \; s \; v)) = \delta_V^*(\ell.\get
-\; s, v)
-$$
-This is witnessed by the fact that if $s = [``Ada \; Lovelace", ``Stephen \; Kleene", ``Alan \; Turing"]$ and \\
-$v = [``Kleene", ``Turing", ``Lovelace"]$, then 
-\begin{align*}
-&\ell_1.\pput \; [``Ada \; Lovelace", ``Stephen \; Kleene", ``Alan \; Turing"] \; [``Kleene", ``Turing", ``Lovelace"]\\
-&= [``Ada \; Kleene", ``Stephen \; Turing", ``Alan \; Lovelace"]
-\end{align*}
-Also, $\delta_{S, \ell}^*(s, \ell.\pput \; s \; v) = ([1,1,1], [])$, $\delta_V^*(\ell.\get
-\; s, v) = ([0,0,0], [])$, but
-\begin{align*}
-(\partial(S, V)^* \times \ell^*).\get \; (\delta_{S, \ell}^*(s, \ell.\pput \; s \; v)) &=
-(\partial(S, V)^* \times \ell^*).\get \; ([1,1,1], [])\\
-&= ([1,1,1], [])\\
-&\neq ([0,0,0], [])\\
-&= \delta_V^*(\ell.\get \; s, v)
-\end{align*}
-Here the positional lens $\ell_1$ is not a difference lens with respect to $(\delta_{S, \ell}^*, 
-\partial(S, V)^* \times \ell^*, \delta_V^*)$ since it does not align sources to views in the same way that
-$\delta_V^*$ does. On the other hand, $\ell_2$ does, hence is a difference lens with respect to $(\delta_{S, \ell}^*, 
-\partial(S, V)^* \times \ell^*, \delta_V^*)$.
-
-In the rest of this paper, we give concrete examples of how to use $\partial D$ preservation to characterize 
-the guarantees on update propagation provided by various lens programming languages. In particular,
-the major technical contributions that we will make in this paper are the following:
-\begin{enumerate}
-  \item
-  We formalize a condition called $\partial D$ preservation which provides a way
-  of constraining the updates that can be made by a bidirectional
-  transformation. We give examples of relations $\partial D$ as well as
-  examples and counterexamples of lenses that are $\partial D$ preserving.
-  \item
-  We describe an instance of $\partial D$ preservation that is satisfied by {\em difference lenses}.
-We give a set of combinators for {\em difference lenses} and describe the update propagation guarantees that the lenses derived using these combinators give.
-  \item
-  Finally, we show how to represent matching lenses and relational lenses
-  \cite{bohannon2006relational} as difference lenses, and hence
-  describe non-trivial guarantees beyond correctness and hippocraticness that
-  these lens languages provide as well.
-\end{enumerate}
-\fi
 \section{Laying Down BX Laws}
-\begin{definition}
+\begin{definition}[Stevens \cite{stevens2010bidirectional}]
 A bidirectional transformation (bx) $R : M \Leftrightarrow N$ comprises of 
@@
 a consistency relation $R \subseteq M \times N$,
-\item
-a relation $S \subseteq (M \times M) \times (N \times N)$ such that for all $m, m' \in M$, $n, n' \in N$,
-\begin{enumerate}
-\item
-if $(m, m') \stackrel{S}{\sim} (n, n)$ then $m = m'$,
-\item
-if $(m, m) \stackrel{S}{\sim} (n, n')$ then $n = n'$, and
-\end{enumerate}
 \item
@@
 \end{definition}
-Intuitively, the relation $S$ relates differences on the $M$ side to differences on the $N$ side.
-For instance, $S$ could be the relation $(m, m') \stackrel{S}{\sim} (n,n')$ if and only if $\delta_M(m,m') = \delta_N(n,n')$ for some metrics $\delta_M$ and $\delta_N$ on $M$ and $N$ respectively.
-The relation is required to send only trivial deltas to trivial deltas.
-\begin{definition}
-A bx $R : M \Leftrightarrow N$ is {\em $S$-preserving} for all $m \in M$, $n, n' \in N$, then if $R(m, n)$ then $(m, \overleftarrow{R}(m,n')) \stackrel{S}{\sim} (n, n')$
-\end{definition}
-Intuitively $S$-preservation says that if we start in a situation where $m$ is consistent with $n$, update $n$ to get $n'$, then map $n'$ back into $m$ via $\overleftarrow{R}(m,n')$, then the difference between $n$ and $n'$ is related to the difference between $m$ and $\overleftarrow{R}(m,n')$.
+
 \begin{definition}
 A bx $R : M \Leftrightarrow N$ is hippocratic if for all $m \in M$, $n \in N$, if $R(m, n)$ then $\overrightarrow{R}(m, n) = n$ and $\overleftarrow{R}(m,n) = m$.
+$R$ is {\em correct} if the consistency restorers do restore consistency: that is, for all $m \in M, n \in N$, we have $R(m, \overrightarrow{R}(m, n))$, and dually.
+$R$ is {\em well behaved} if it is correct and hippocratic.
+\end{definition}
+As we saw in the examples in the introduction, consistency restoration often times is too loose of a constrain to prevent a well-behaved bx from causing unexpected behaviour.
+What is missing from this model is how the bx handles non-trivial updates.
+The following definition factors in this missing piece, by relating updates on one side of a bx to updates on the other side of the bx.
+\begin{definition}
+A difference bidirectional transformation (diff-bx) $(R,S) : M \Leftrightarrow N$ comprises of 
+\begin{enumerate}
+\item
+a consistency relation $R \subseteq M \times N$,
+\item
+a difference relation $S \subseteq (M \times M) \times (N \times N)$ such that for all $m, m' \in M$, $n, n' \in N$,
+\begin{enumerate}
+\item
+if $(m, m') \stackrel{S}{\sim} (n, n)$ then $m = m'$,
+\item
+if $(m, m) \stackrel{S}{\sim} (n, n')$ then $n = n'$, and
+\end{enumerate}
+\item
+restorers $\overrightarrow{R} : M \times N \rightarrow N$ and $\overleftarrow{R} : M \times N \longrightarrow
+N$.
+\end{enumerate}
+\end{definition}
+Intuitively, the relation $S$ relates differences on the $M$ side to differences on the $N$ side. The relation is required to send only trivial differences to trivial differences.
+\begin{definition}
+A diff-bx $(R,S) : M \Leftrightarrow N$ is {\em $S$-preserving} if for all $m \in M$, $n, n' \in N$, then if $R(m, n)$ then $(m, \overleftarrow{R}(m,n')) \stackrel{S}{\sim} (n, n')$
 \end{definition}
 \begin{fact}
-For any $R : N \Leftrightarrow N$ and any $S \subseteq (M \times M) \times (N \times N)$, if $R$ is $S$-preserving, then $R$ is hippocratic.
+If $(R,S) : M \Leftrightarrow N$ is $S$-preserving, then $R$ is hippocratic.
 \end{fact}
 \begin{definition}
-A bx $R : M \Leftrightarrow N$ is {\em correct} if the consistency restorers do
-restore consistency: that is, for all $m \in M, n \in N$, we have $R(m,
-\overrightarrow{R}(m, n))$, and dually.
-\end{definition}
-\iffalse
-A {\em lens} is a program that can be run in two directions. In the
-forward direction, the \get component of a lens produces a {\em view}
-from a {\em source}, while in the backward direction, the \pput component folds
-an updated view back into the original source. Formally, Foster
-\cite{foster2007combinators} defined a lens $\ell : S
-\Leftrightarrow V$ from a source $S$ to a view $V$ to be a pair of functions
-$\ell.\get : S \longrightarrow V$ and $\ell.\pput : S \longrightarrow V
-\longrightarrow S$ satisfying the
-following properties:
+A diff-bx $(R,S) : M \Leftrightarrow N$ is {\em well behaved} if $(R,S)$ is $S$-preserving and $R$ is correct.
+\end{definition}
+In the next section we describe a language for deriving well-behaved difference bidirectional transformations based on the {\em basic lenses} implemented in the lens programming language {\em Boomerang}.
+\section{Difference Lenses}
+\begin{definition}
+A lens $\ell : S \Leftrightarrow V$ is a well behaved bx where the forward transformation $\overrightarrow{\ell} : S \times V \longrightarrow V$ does not depend on its $V$-argument, and where the consistency relation $\ell$ is given by $(s, v) \in R$ if and only if $\overrightarrow{\ell}(s) = v$
+
+Equivalently, a lens $\ell : S \Leftrightarrow V$ is a pair of functions $\ell.\get : S \longrightarrow V$ and $\ell.\pput : S \longrightarrow V \longrightarrow V$ such that the following equations hold:
 \begin{align*}
@@
 \end{align*}
-The PUTGET law ensures that updates to the view are translated
-``exactly''---i.e. that, given a view, the \pput function produces a source
-that \get maps back to the very same view, while the GETPUT law ensures a
-``stability'' property for the source---i.e. it requires that the \pput
-function returns the original source unmodified whenever the update to the view
-is a no-op.
-
-Unfortunately, while these laws give some guarantees on how the \get and \pput
-components of a lens interact, they do not specify the behaviour of the \pput
-component when the update to the view is non-trivial. Consequently, a
-non-trivial update to a view could lead to an unexpected update to the source.
-Various solutions have been proposed to address this issue, for example {\em
-matching lenses} to deal with the issue of alignment \cite{barbosa2010matching},
-{\em edit lenses} \cite{hofmann2012edit} and {\em delta lenses}
-\cite{diskin2011asymmetric,diskin2011state,pacheco2012delta} which explicitly
-propagate edits through the lens, and {\em least change lenses} which
-add a new LEASTPUTGET law which requires that the source value resulting from a
-\pput application is not only acceptable, but also one of closest to the
-original $s$ among the sources that share the same view $v$
-\cite{macedo2013composing}.
-
-This paper describes a general method for specifying what updates are allowable
-for a given lens without requiring that the lens explicitly propagate updates.
-Our approach is guided by the simple idea the idea that if one starts with a
-source $s$, calls the \get function on $s$ to acquire the view $\ell.\get \;
-s$, updates $\ell.\get \; s$ to get a new view $v$, then the ``difference''
-between $v$ and $\ell.\get \; s$ and $v$ should be related to the difference
-between $\ell.\pput \; s \; v$ and $v$.
-
-To formalize this idea, we first recall a widely used definition of
-bidirectional transformations
-\cite{meertens1998designing,stevens2010bidirectional} definition of a
-bidirectional transformations:
+\end{definition}
+In order to strengthen hippocraticness (i.e.
+the GETPUT) law, we need to identify a difference relation $S$ that relates source updates to view updates.
+Our choice is relate the pair $(\ell.\pput \; s \; v, s)$ to the pair $(v, \ell.\get \; s)$ if and only if $\delta_S(\ell.\pput \; s \; v, s) \geq \delta_V(v, \ell.\get \; s)$ for some real-valued functions $\delta_S : S \times S \longrightarrow \mathbb{R}^{\geq 0}$ and $\delta_V : V \times V \longrightarrow \mathbb{R}^{\geq 0}$.
+
+The functions $\delta_S$ and $\delta_V$ will depend only on $S$ and $V$, and will be designed to measure how similar two sources (or views) are to each other.
+More specifically, given a regular expression $R$ and strings $s, s' \in \mathcal{L}(R)$, we will define a number $\delta_R(s, s')$ inductively on $R$, with $0 \leq \delta_R(s, s') \leq 1$, and a high value of $\delta_R(s, s')$ indicates that $s$ and $s'$ are similar---in particular, if $s = s'$, then $\delta_R(s, s') = 1$, and if $s \neq s'$ and either $s = \varepsilon$ or $s' = \varepsilon$, then $\delta_R(s, s') = 0$.
+We will also expand our definition of the regular expression $R^*$, having three different variants $R^*_{POS}$, $R^*_{NC}$ and $R^*_{MED}$, where the differences are computed using positional, non-crossing and min-edit distance alignment strategies respectively, so that for all $s, s' \in \mathcal{L}(R^*)$, $\delta_{R^*_{MED}}(s, s') \geq \delta_{R^*_{NC}}(s, s') \geq \delta_{R^*_{MED}}(s, s')$.
+\subsection{Difference Functions on Regular Languages}
 \begin{definition}
-A bidirectional transformation (bx) $R: M \Leftrightarrow N$ comprises of a
-consistency relation $R \subseteq M \times N$ and restorers $\overrightarrow{R}
-: M \times N \rightarrow N$ and $\overleftarrow{R} : M \times N \longrightarrow
-N$.
+Given a finite alphabet $\Sigma$, a {\em difference regular expression} is any expression generated by the following grammar,
+$$R := s \; | \; R_1 \cdot R_2 \; | \; R_1 \cup R_2 \; | \; R^*_{POS} \; | \; R^*_{NC} \; | \; R^*_{MED}$$
+where $s$ ranges over $\Sigma^*$ and $R$ ranges over regular expressions. The languages of $R^*_{POS}$, $R^*_{NC}$ and $R^*_{MED}$ are all the same, namely $\mathcal{L}(R^*)$.
 \end{definition}
 \begin{definition}
-A bx $R : M \leftrightarrow N$ is {\em correct} if the consistency restorers do
-restore consistency: that is, for all $m \in M, n \in N$, we have $R(m,
-\overrightarrow{R}(m, n))$, and dually. $R$ is {\em hippocratic} if the
-consistency restores make no change to and already-consistent pair of models:
-that is, for all $m \in M$, $n \in N$, if $R(m, n)$ then $\overrightarrow{R}(m,
-n) = n$ and dually. $R$ is {\em well behaved} if it is correct and hippocratic.
-\end{definition}
-
-Observe that $\ell : S \Leftrightarrow V$ is a lens, let $R$ be the
-consistency relation given by $R(s, \ell.\get \; s)$ for all $s \in
-S$, $\overrightarrow{\ell}(s, v) = \ell.\get \; s$, and $\overleftarrow{\ell} =
-\ell.\pput$. By the PUTGET law, $\ell.\get \; (\ell.\pput \; s \; v) = v$,
-hence $\ell$ is correct, and by the GETPUT laws, $\ell.\pput \; s \; (\ell.\get
-\; s) = s$ and $\ell.\get \; (\ell.\create \; v) = v$, hence $\ell$ is
-hippocractic. Therefore $\ell$ is a well-behaved bx.
-
-If $R$ is a well-behaved bx then $R$ restores consistency when a source or view
-is updated (correctness) and preserves consistency when the source and view are
-already consistent (hippocraticness). In order to express
-
-We will then apply this method to define a typing judgement for lenses which
-constrains how a lens propagates updates. We will describe a compositional
-language of {\em difference lenses} which satisfy this new typing judgement,
-and give examples of ordinary lenses that are not difference lenses. Finally,
-we will show how to represent matching lenses and relational lenses
-\cite{bohannon2006relational} as difference lenses, and hence describe the
-guarantees that these lens languages provide.
-
-For example suppose that the source is an XML document representing the
-names, dates and nationalities of a collection of classical music
-composers such as,
-
-\begin{lstlisting}
-<composers>
-<composer>
-<name>Jean Sibelius</name>
-<dates>1865-1956</dates>
-<nationality>Finnish</nationality>
-</composer>
-<composer>
-<name>Aaron Copland</name>
-<dates>1910-1990</dates>
-<nationality>American</nationality>
-</composer>
-<composer>
-<name>Benjamin Briten</name>
-<dates>1913-1976</dates>
-<nationality>English</nationality>
-</composer>
-</composers>
-\end{lstlisting}
-
-while the view is a list if ASCII text representing the names and dates of each
-composer:
-\begin{lstlisting}
-Jean Sibelius, 1865-1956
-Aaron Copland, 1910-1990
-Benjamin Briten, 1913-1976
-\end{lstlisting}
-
-After computing the initial view, we might want to edit it in some way---e.g.,
-correcting the error in Sibelius's death date and the misspelling in Britten's
-name
-\begin{lstlisting}
-Jean Sibelius, 1865-|\colorbox{orange}{1957}|
-Aaron Copland, 1910-1990
-Benjamin |\colorbox{orange}{Britten}|, 1913-1976
-\end{lstlisting}
-and push the changes back into the original XML format:
-\begin{lstlisting}
-<composers>
-<composer>
-<name>Jean Sibelius</name>
-<dates>1865-|\colorbox{orange}{1957}|</dates>
-<nationality>Finnish</nationality>
-</composer>
-<composer>
-<name>Aaron Copland</name>
-<dates>1910-1990</dates>
-<nationality>American</nationality>
-</composer>
-<composer>
-<name>Benjamin |\colorbox{orange}{Britten}|</name>
-<dates>1913-1976</dates>
-<nationality>English</nationality>
-</composer>
-</composers>
-\end{lstlisting}
-
-Lenses are designed to perform these kinds of bidirectional transformations. If
-$\ell$ is a suitably defined lens for the transformation that we
-have just described, then the \get component of the lens performs the
-XML to ASCII transformation that produces an ASCII view from an XML source,
-while the \pput component of the lens performs the ASCII to XML transformation
-that folds back an updated ASCII view back into the original XML source.
-
-For example, suppose that a source $S$ consists of a pair consisting of a
-boolean value and an integer, and that the view $V$ for $S$ consists of a
-single integer. Further, suppose that we wish to define a \pput function for a
-lens $\ell$ whose \get component is the function which given a pair $(b, n)$
-simply produces the boolean value $b$. Let $f : \mathbb{B} \times \mathbb{Z}
-\longrightarrow \mathbb{Z}$ be any function from $\mathbb{B} \times \mathbb{Z}
-\longrightarrow \mathbb{Z}$. Then the lens $\ell_{f}$ defined by
-\begin{align*}
-\ell_{f}.\get \; (b, n) &= b\\
-\ell_{f}.\pput \; (b, n) \; b' &= \begin{cases}
-(b', n) & \text{if }b = b'\\
-(b', f(b', n)) & \text{otherwise}\\
+Given an unambiguous difference regular expression $R$, define $\delta_R : \mathcal{L}(R) \times \mathcal{L}(R) \longrightarrow \mathbb{R}^{\geq 0}$ inductively as follows:
+\begin{align*}
+\delta_s(t, t') &= 1\\
+\delta_{R_1 \cdot R_2}(s_1 \cdot s_2, s'_1 \cdot s'_2) &= \frac{1}{2}(\delta_{R_1}(s_1, s'_1) + \delta_{R_2}(s_2, s'_2))\\
+\delta_{R_1 \cup R_2}(s, s') &= 
+\begin{cases}
+\delta_{R_1}(s, s') & \text{if }s, s' \in \mathcal{L}(R_1) \\
+\delta_{R_2}(s, s') & \text{if }s, s' \in \mathcal{L}(R_2)\\
+0 & \text{otherwise }
 \end{cases}\\
-\ell.\create \; b &= (b,f(b, 0))
-\end{align*}
-is a valid lens.
-
-A user who applies a lens $\ell$ of type $\ell : S \Leftrightarrow V$ may expect
-that $\ell.\pput \; (true, 1) \; false = (false, 0)$, reasoning that the
-integer produced by the \pput function should be the number representation of
-the boolean value $false$. Or perhaps she might expect that $\ell.\pput \;
-(true, 1) \; false = (false, 1)$, reasoning that the two pieces of data in a
-pair are independent and that the reasonable \pput behaviour in this case is to
-restore the orginial integer. In either case, the user may be shocked to find
-that in fact $\ell.\pput (true, 0) \; false = (false, 48)$! Given the \get
-behaviour which just produces the boolean value from a pair, since $\ell_f$ is a
-valid lens for any $f$, then the \pput function of a valid lens of type
-$\mathbb{B} \times \mathbb{Z} \Leftrightarrow \mathbb{Z}$ can restore the
-integer in a pair in any way so long as the old boolean value and the updated
-boolean value are different.
-
-In order to give a more constrained specification for the desired
-lens we define the difference $\delta_S((a, b), (c, d))$ between $(a, b)$ and
-$(c, d)$ by $$\delta_S((a, b), (c,
-d)) = \begin{cases}
-0 & \text{if }a = c, b = d\\
-1 & \text{if }a=c, b \neq d\\
-1 & \text{if }a \neq c, b = d\\
-2 & \text{if }a \neq c, b \neq d
-\end{cases}$$
-On the view side, we define the difference $\delta_V(a, b)$
-between $a$ and $b$ by, $$\delta_V(a, b) = \begin{cases}
-0 & \text{if } a = b\\
-1 & \text{if } a \neq b
-\end{cases}$$
-
-Next we define a lens $\partial(S, V) : \partial S \Leftrightarrow \partial V$
-which describes how differences should be propagated. For this example, we
-define $\partial(S,V)$ to simply be the \ccopy lens from $\partial S$ to
-$\partial V$, keeping in mind that $\partial S = \partial V = \mathbb{Z}$.
-\cref{difflenslaw} enables us to rule out $\ell_{bad}$, since
-\begin{align*}
-(\partial (S, V)).\get \; (\delta_S(\ell_{bad}.\pput \; (1,2) \; 3, (1,2))) &=
-(\partial (S, V)).\get \; (\delta_S((0,3), (1,2)))\\
-&= (\partial (S, V)).\get \; 2\\
-&= 2\\
-&\neq 1\\
-&= \delta_V(3, 2)\\
-&= \delta_V(3, \ell_{bad}.\get \; (1, 2))
-\end{align*}
-Consequently $\ell_{bad}$ is not a valid delta lens. On the
-other hand, the lens $\ell_{good} : S \Leftrightarrow V$ defined by
-\begin{align*}
-\ell_{good}.\get \; (a, b) &= b\\
-\ell_{good}.\pput \; (a, b) \; c &= (a, c)\\
-\ell_{good}.\create \; b &= (0, b)
-\end{align*}
-is indeed a valid delta lens since the its \pput component always restores the
-first factor.
-
-\section{Strengthening Lens Laws}
+\delta_{R^*_{POS}}(s_1 \cdot \ldots \cdot s_m, t_1 \cdot \ldots \cdot t_n) &= 
+\begin{cases}
+0 & \text{if } m=0 \text{ or } n=0\\
+\max\limits_{\{\varphi : [m] \longrightarrow [n] \; | \; \varphi(i) = i\}}\frac{1}{n}\sum\limits_{i=1}^{m}\ \delta_R(s_i, t_{\varphi(i)})& \text{if } m \leq n\\
+\max\limits_{\{\varphi : [n] \longrightarrow [m] \; | \; \varphi(i) = i\}}\frac{1}{m}\sum\limits_{i=1}^{m}\ \delta_R(t_i, s_{\varphi(i)})& \text{otherwise}
+\end{cases}\\
+\delta_{R^*_{NC}}(s_1 \cdot \ldots \cdot s_m, t_1 \cdot \ldots \cdot t_n) &= 
+\begin{cases}
+0 & \text{if } m=0 \text{ or } n=0\\
+\max\limits_{\{\varphi : [m] \longrightarrow [n] \; | \; \varphi \text{ is strictly increasing}\}}\frac{1}{n}\sum\limits_{i=1}^{m}\ \delta_R(s_i, t_{\varphi(i)})& \text{if } m \leq n\\
+\max\limits_{\{\varphi : [n] \longrightarrow [m] \; | \; \varphi \text{ is strictly increasing}\}}\frac{1}{m}\sum\limits_{i=1}^{m}\ \delta_R(t_i, s_{\varphi(i)})& \text{otherwise}
+\end{cases}\\
+\delta_{R^*_{MED}}(s_1 \cdot \ldots \cdot s_m, t_1 \cdot \ldots \cdot t_n) &= 
+\begin{cases}
+0 & \text{if } m=0 \text{ or } n=0\\
+\max\limits_{\{\varphi : [m] \longrightarrow [n] \; | \; \varphi \text{ is injective}\}}\frac{1}{n}\sum\limits_{i=1}^{m}\ \delta_R(s_i, t_{\varphi(i)})& \text{if } m \leq n\\
+\max\limits_{\{\varphi : [n] \longrightarrow [m] \; | \; \varphi \text{ is injective}\}}\frac{1}{m}\sum\limits_{i=1}^{m}\ \delta_R(t_i, s_{\varphi(i)})& \text{otherwise}
+\end{cases}
+\end{align*}
+\end{definition}
+\begin{fact}
+For any regular expression $R$ and $s, s' \in \mathcal{L}(R)$
+\begin{enumerate}
+\item
+$0 \leq \delta_R(s, s') \leq 1$.
+\item
+$\delta_R(s, s) = 1$.
+\item
+If $s \neq s'$ and either $s = \varepsilon$ or $s' = \varepsilon$, then $\delta_R(s, s') = 0$
+\item
+For all $s, s' \in \mathcal{L}(R^*)$, $\delta_{R^*_{MED}}(s, s') \geq \delta_{R^*_{NC}}(s, s') \geq \delta_{R^*_{MED}}(s, s')$.
+\item
+$\delta_R(s, s') = \delta_R(s', s)$.
+\end{enumerate}
+
+\end{fact}
+\subsection{Lens Combinators}
 \begin{definition}
-A bidirectional transformation (bx) $R: M \Leftrightarrow N$ comprises of a
-consistency relation $R \subseteq M \times N$ and restorers $\overrightarrow{R}
-: M \times N \rightarrow N$ and $\overleftarrow{R} : M \times N \longrightarrow
-N$.
-\end{definition}
-\begin{definition}
-A bx is {\em correct} if the consistency restorers do restore consistency:
-that is, for all $m \in M, n \in N$, we have $R(m, \overrightarrow{R}(m, n))$,
-and dually.
-\end{definition}
-\begin{definition}
-A bx is {\em hippocratic} if the consistency restores make no change
-to and already-consistent pair of models: that is, for all $m \in M$, $n \in N$,
-if $R(m, n)$ then $\overrightarrow{R}(m, n) = n$ and dually.
-\end{definition}
-\begin{definition}
-A bx is {\em well behaved} if it is correct and hippocratic.
-\end{definition}
-For lenses, if $\ell : S \Leftrightarrow V$ is a lens, we take the
-consistency relation $R$ to be given by $R(s, \ell.\get \; s)$ for all $s \in
-S$, $\overrightarrow{\ell}(s, v) = \ell.\get \; s$, and $\overleftarrow{\ell} =
-\ell.\pput$. By the PUTGET law, $\ell.\get \; (\ell.\pput \; s \; v) = v$,
-hence $\ell$ is correct, and by the GETPUT laws, $\ell.\pput \; s \; (\ell.\get
-\; s) = s$ and $\ell.\get \; (\ell.\create \; v) = v$, hence $\ell$ is
-hippocractic. Therefore $\ell$ is a well-behaved bx.
-
-Using the notation of Stevens \cite{stevens2010bidirectional},
-
-The major shortcoming of this model of bx transformations is that it does not
-take into account how changes should be propagated by transformation; a
-well-behaved bx need only restore consistency when models go out of sync or
-maintain consisteny if models are already in sync. This becomes an issue when
-there are many different ways of restoring consistency
-
-\section{Motivation and Related Work}
-Foster et. al. \cite{foster2007combinators,foster2009bidirectional}
-defined a lens $\ell : S \Leftrightarrow V$ from a source $S$ to a view $V$ to
-be a triple of functions $\ell.\get : S \longrightarrow V$, $\ell.\pput : S
-\longrightarrow S \longrightarrow V$ and $\ell.\create : V \longrightarrow S$
-satisfying the following properties:
-\begin{align*}
-\ell.\get \; (\ell.\pput \; s \; v) &= v \tag{PUTGET}\\
-\ell.\get \; (\ell.\create \; v) &= v \tag{CREATEGET}\\
-\ell.\pput \; s \; (\ell.\get \; s) &= s \tag{GETPUT}
-\end{align*}
-While these two laws give some guarantees on how the \get, \pput and \create
-components of a lens interact, they do not specify the behaviour of the \pput
-component when the update to the view is non-trivial, because the GETPUT law
-only says that the \pput component must return the original source unmodified
-whenever the update to the view is a no-op. Consequently, a non-trivial update
-to a view could lead to an undesired update to the source.
-
-Therefore, several lens formalism have been proposed that explicitly have a
-notion of an {\em update}. The first of these was Barbosa et. al.'s {\em
-matching lenses}\cite{barbosa2010matching}. Matching lenses specifically deal
-with the issue of {\em alignment}, since Foster et. al.'s original iteration
-lens updates records positionally:
-\begin{align*}
-\ell^* .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n) &= t_1 \cdot
-\ldots \cdot t_n\\
-\text{ where } t_i &= \begin{cases}
-\ell.\pput \; s_i \; v_i & \text{if } 1 \leq i \leq \min\{m, n\}\\
-\ell.\create \; v_i & \text{if } \min\{m, n\} < i \leq n
-\end{cases}\\
-\end{align*}
-In their work, Barbosa et. al. enrich the types of lenses with ``chunks''
-identifying reorderable pieces of the source and view that should be re-aligned
-after an update, and formulate behavioral laws that capture essential
-constraints on the handling of chunks. They develop a core language
-of matching lenses for strings, together with a set of ``alignment
-combinators'' that implement a variety of alignment strategies. The obvious
-shortcoming of matching lenses is that they only deal with alignment; they do
-not solve the general issue of update propagation.
-
-The next line of work is Diskin et. al.'s {\em delta
-lenses}\cite{diskin2011asymmetric,diskin2011state}. Diskin et. al.'s delta
-lenses work as follows: let $\ell : S \Leftrightarrow V$ be a lens from
-$S$ to $V$. Assume that $s$ is a source, and that $v = \ell.\get s$. A user
-performs an edits $d$ on $v$ to get $v'$: $v \xrightarrow{d} v'$. The \pput
-function now takes the view edit $d$ and the original source $s$ and produces a
-new source $s'$: $s \xrightarrow{\pput(d, s)} s'$. One thing that Diskin et. al.
-do not do is that they do not give any concrete examples of delta lenses,
-no composition, no notion of equivalence, and no combinators for constructing
-delta lenses.
-
-The next line of work is Hofmann et. al.'s {\em edit
-lenses}\cite{hofmann2012edit}. Edit lenses are similar to delta lenses, but edit
-lenses are more explicit. Specifically, the source and view edits are
-represented as elements of a {\em monoid}, and the lens acts on sources and
-views via monoid actions. That is if $v$ is a view and $d_v$ is an edit that
-can be made on $v$ to produce $v'$, then $d_v \cdot v = v'$. The intended usage
-of an edit lens is as follows. There are two users, one holding an element of
-$S$ the other one an element of $V$ both referred to as {\em replicas}.
-Initially, they hold $init_S \in S$ and $init_V \in V$ respectively. The users
-then perform actions and propagate them across the lens. An action consists of
-producing an edit $d_s$ (or $d_v$), applying it to one's current replica $s$
-(resp. $v$), putting the edit through the lens to obtain an edit $d_v$ (resp.
-$d_s$), and asking the user on the other side to apply $d_v$ (resp. $d_s$) to
-their replica. One thing that Hofmann et. al. do that Diskin et. al. do not do
-is that they do give concrete examples of edit lenses, define a suitable notion
-of composition on them, and define combinators for constructing edit lenses.
-
-The final line of work in this area is Pacheco et. al.'s
-version of delta lenses\cite{pacheco2012delta}. Pacheco et. al. are the
-first to give concrete examples of delta lenses, define a suitable notion
-of composition on them, and define combinators for constructing delta
-lenses. Pacheco et al. define a delta lens $\ell$ (d-lens for short), denote by
-$\ell: S \trianglerighteq_{\blacktriangle} V$ to be a bidirectional
-transformations that comprises four total functions:
-\begin{align*}
-\get &: S \longrightarrow V\\
-\pput &: \forall(v, s) : V \times S. (v \triangle \get \;s) \longrightarrow
-SA\\
-\get_{\blacktriangle} &: \forall\{s':S, s:S\}. (s' \triangle s)
-\longrightarrow (\get \; s')\triangle (\get \; s)\\
-\pput_{\blacktriangle} &:\forall\{(v, s):V \times S\},d:(v \triangle \get
-\;s).(\pput \; (v, s) \; d) \triangle s
-\end{align*}
-that satisfy the following propertes:
-\begin{align*}
-\get \; (\pput \; (v, s) \; d) = v \tag{PUTGET}\\
-\pput \; (\get \; s, s) \; id = s \tag{GETPUT}\\
-\get_{\blacktriangle} \; (\pput_{\blacktriangle} \; d) = d
-\tag{$\text{PUTGET}_{\blacktriangle}$}\\
-\pput_{\blacktriangle} \; id = id \tag{$\text{PUTGET}_{\blacktriangle}$}
-\end{align*}
-The \get function is a just like a \get function in Foster's et. al.'s original
-work. The \pput function is different though. Let $s \in S$ and $v \in V$, and
-let $d$ be an update from $v$ to $\ell.\get \; s$; for example, $d$ is a
-sequence of insertions and deletions that transforms $v$ to $\ell.\get \; s$.
-Then \pput takes in $s$ and $v$ as input, and in addition receives $d$ to
-compute $\pput \; (v, s) \; d$.
-
-The $\get_{\blacktriangle}$ and $\pput_{\blacktriangle}$ functions are functions
-that transform deltas to deltas. The last two axioms for delta lenses say that
-if $d$ is a delta from $v$ to $\ell.\get \; s$ and $\Delta$ is a delta from
-$\pput \; (v, s) \; d$ to $s$, then $\get_{\blacktriangle} \; \Delta = d$, and
-that if $d$ is the identity delta, then $\Delta$ is also the identity delta.
-
-In addition to describing a delta lens language, Pacheco et. al. also
-address the following shortcoming with edit lenses: edit lenses include
-combinators for constructing lenses over {\em container types}.
-In the simplest case, container types can be thought of as algebraic datatypes
-such as lists and trees; each value of a container type has a certain {\em
-shape}, and each shape has a set of {\em positions} associated with it. For
-instance, the shape of a list of size $n$ is just $n$, while the positions of
-such a list are the numbers $0, 1, \ldots, n-1$.
-
-The shortcoming of the container edit lenses addressed by Pacheco et. al.
-elements is that these lenses always preserve the number of positions; for
-instance if $\ell$ is a container lens that maps lists to trees, and if $xs$ is
-a list of size $n$, then the number of positions of $\ell.\get \; xs$ is also
-$n$. Pacheco et. al. define a new set of combinators for container lenses that
-can say map a list of size $5$ to a tree with 4 nodes.
-
-In this article we show that it is not necessary for the \pput component of a lens
-to explicitly propagate updates: instead, we can specify what the put function should
-evaluate to, by imposing a more precise type judgement on lenses.
-
-The idea of the stronger typing judgement is that we embed sources and views inside
-of {\em difference structures} and then we define a lens that describes how the
-differences behave as they cross in-between the source and the view. A
-lens that is validly typed is then a lens that preserves this overall structure.
-
-Here are the actual definitions:
-\begin{definition}
-A weighted set is a pair $(X, |\cdot|)$ where $X$ is a set and $|\cdot| : X
-\longrightarrow \mathbb{R}^{\geq 0}$ is a map from $X$ to the non-negative
-reals.
-\end{definition}
-\begin{definition}
-A difference structure is a triple $(S, \partial S, \delta_S)$ where $S$ and
-$\partial S$ are weighted sets, and $\delta_S:
-S \times S \longrightarrow \partial S$ is a function satisfying
-\begin{align*}
-\delta(s, s') &= \delta(s', s)\tag{A1}\\
-|\delta(s, s)| &= 0\tag{A2}\\
-|\delta(s, s')| &\leq |s| + |s'|\tag{A3}
-\end{align*}
-\end{definition}
-\begin{definition}
-Let $(S, \partial S, \delta_S)$ and $(V, \partial V, \delta_V)$ be
-difference structures. Let $\partial(S, V) : \partial S \Leftrightarrow \partial
-V$ be a lens from $\partial S$ to $\partial V$. Then a lens $\ell : S
-\Leftrightarrow V$ also has type $\ell : (S, \partial S,
-\delta_S) \xLeftrightarrow{\partial(S, V)} (V, \partial V, \delta_V)$ if for
-all $s \in S, v \in V$,
+A lens $\ell : S \Leftrightarrow V$ is a {\em difference lens} if it satisfies the following equation:
 \begin{equation}\label{difflenslaw}
-(\partial (S, V)).\get\; (\delta_S(\ell.\pput \; s \; v, s)) = \delta_V(v,
-\ell.\get \; s)
+\delta_S(\ell.\pput \; s \;v, s) \geq \delta_V(v, \ell.\get \; )
 \end{equation}
 \end{definition}
-The extra difference structure on the sources and the views, as well as the lens
-$\partial(S, V)$ force the lens $\ell$ to satisfy \cref{difflenslaw}; this law is
-therefore a specification that constrains what $\ell$ can do.
-
-For example, suppose that a source $S$ consists of a pair of integers, and that the
-view $V$ for $S$ consists of a single integer. Further, suppose that we wish
-for the \get component of any lens from $S$ to $V$ to simply project away the
-first factor:that is for any $\ell : \mathbb{Z} \times \mathbb{Z}
-\Leftrightarrow \mathbb{Z}$, $\ell.\get \; (a, b) = b$ for all integers $a, b$.
-Suppose further that we wish to force the \pput function of any lens $\ell$
-defined on the data to restore the factor projected away by the \get function:
-that is, we wish to enforce that for all integers $a, b, c$, $\pi_1(\ell.\pput
-\; (a, b) \; c) = a$, where $\pi_1$ is the projection on the first factor.
-
-Now consider the lens $\ell_{bad}$ defined by
-\begin{align*}
-\ell_{bad}.\get \; (a, b) &= b\\
-\ell_{bad}.\pput \; (a, b) \; c &= \begin{cases}
-(a, c) & \text{if }b = c\\
-(0, c) & \text{otherwise}\\
-\end{cases}\\
-\ell.\create \; b &= (0,b)
-\end{align*}
-Then $\ell_{bad}$ is a valid lens even though $\ell_{bad}$ does not
-always restore the first factor. For example, observe that $\ell.\pput (1, 2) \;
-3 = (0,3)$. In order to give a more constrained specification for the desired
-lens we define the difference $\delta_S((a, b), (c, d))$ between $(a, b)$ and
-$(c, d)$ by $$\delta_S((a, b), (c,
-d)) = \begin{cases}
-0 & \text{if }a = c, b = d\\
-1 & \text{if }a=c, b \neq d\\
-1 & \text{if }a \neq c, b = d\\
-2 & \text{if }a \neq c, b \neq d
-\end{cases}$$
-On the view side, we define the difference $\delta_V(a, b)$
-between $a$ and $b$ by, $$\delta_V(a, b) = \begin{cases}
-0 & \text{if } a = b\\
-1 & \text{if } a \neq b
-\end{cases}$$
-
-Next we define a lens $\partial(S, V) : \partial S \Leftrightarrow \partial V$
-which describes how differences should be propagated. For this example, we
-define $\partial(S,V)$ to simply be the \ccopy lens from $\partial S$ to
-$\partial V$, keeping in mind that $\partial S = \partial V = \mathbb{Z}$.
-\cref{difflenslaw} enables us to rule out $\ell_{bad}$, since
-\begin{align*}
-(\partial (S, V)).\get \; (\delta_S(\ell_{bad}.\pput \; (1,2) \; 3, (1,2))) &=
-(\partial (S, V)).\get \; (\delta_S((0,3), (1,2)))\\
-&= (\partial (S, V)).\get \; 2\\
-&= 2\\
-&\neq 1\\
-&= \delta_V(3, 2)\\
-&= \delta_V(3, \ell_{bad}.\get \; (1, 2))
-\end{align*}
-Consequently $\ell_{bad}$ is not a valid delta lens. On the
-other hand, the lens $\ell_{good} : S \Leftrightarrow V$ defined by
-\begin{align*}
-\ell_{good}.\get \; (a, b) &= b\\
-\ell_{good}.\pput \; (a, b) \; c &= (a, c)\\
-\ell_{good}.\create \; b &= (0, b)
-\end{align*}
-is indeed a valid delta lens since the its \pput component always restores the
-first factor.
-
-For a more realistic example using a difference lens combinator, let $S
-= name \cdot dates$ and $V = name$. Let $\ell :
-S \Leftrightarrow V = (copy \; name) \cdot (del \; dates)$, and suppose that we
-wish to define a lens of type $S^* \Leftrightarrow V^*$, where the difference
-function used on the view side is the min-edit distance difference function.
-Assuming that we inductively use the difference function
-$$ \delta(d_1, d_2) =
-\begin{cases}
-0 & \text{if }d_1 = d_2 \\
-1 & \text{otherwise}
-\end{cases}
-$$
-for both the name and date fields, then with the min-edit distance difference
-function, the difference between the string $v$ given by
-\begin{lstlisting}
-John Doe
-Jane Smith
-\end{lstlisting}
-and the string $v'$ given by
-\begin{lstlisting}
-Jane Smith
-John Doe
-\end{lstlisting}
-is the empty list. Let $s$ =
-\begin{lstlisting}
-John Doe,1900-2000
-Jane Smith,1901-2001
-\end{lstlisting}Then $\ell.\get \; s = v$. Now assume that $v$ is edited to $v'$
-given by
-\begin{lstlisting}
-Mark Moe
-John Doe
-Jane Smith
-\end{lstlisting}
-where a new record \lstinline|Mark Moe| has been added to $v$.
-With the min-edit distance on the view side, $\delta(v', \ell.\get \; s)$ is
-the singleton list [``Mark Moe''].
-
-On the source side, the min-edit distance matching lens $\ell^*_{MIN-EDIT}$,
-$\ell^*_{MIN-EDIT}.\pput \; s \; v'$ is the string
-\begin{lstlisting}
-Mark Moe,0000-0000
-John Doe,1900-2000
-Jane Smith,1901-2001
-\end{lstlisting}
-while with the positional lens $\ell^*$, then $\ell^*.\pput \; s \; v'$ is the
-string
-\begin{lstlisting}
-Mark Moe,1900-2000
-John Doe,1901-2001
-Jane Smith,0000-0000
-\end{lstlisting}
-With the min-edit distance difference function,
-$\delta(\ell^*_{MIN-EDIT}.\pput \; s \; v', s)$ is the singleton list
-$[\text{``Mark Moe, 0000-0000''}]$, while with the
-positional lens, $\delta(\ell^*.\pput \; s \; v', s)$ is the list
-given by
-\begin{lstlisting}
-Mark Moe,1900-2000
-(0,1)
-(0,1)
-\end{lstlisting}
-The min-edit distance satisfies \cref{difflenslaw} since
-\begin{align*}
-(\partial (S, V))^*.\get\; (\delta_S(\ell^*_{MIN-EDIT}.\pput \; s \; v', s)) &=
-(\partial (S, V))^*.\get\; ([\text{``Mark Moe, 1900-2000''}])\\
-&= [\text{``Mark Moe''}]\\
-&=
-\delta_V(v', \ell.\get \; s)
-\end{align*}
-On the other hand, the positional lens does not satisfy \cref{difflenslaw} since
-\begin{align*}
-(\partial (S, V))^*.\get\; (\delta_S(\ell^*.\pput \; s \; v', s)) &=
-(\partial (S, V))^*.\get\; ([\text{``Mark Moe, 1900-2000''}, (0,1), (0,1)])\\
-&=[\text{``Mark Moe''}, (0,*), (0,*)]\\
-& \neq \delta_V(v', \ell.\get \; s)
-\end{align*}
-Let $\mathcal{M}_V$ be the set of min-edit distance alignments on the view side,
-and $\mathcal{M}_S$ the corresponding alignments on the source side.
-Then, the typing judgement
-$$\ell^*_{MIN-EDIT}:(S, (\partial S \sqcup S)^*, \match(\mathcal{M}_S, \delta_s))
-\xLeftrightarrow{\partial(S, V)^*} (V, (\partial V \sqcup V)^*, \match(\mathcal{M}_V, \delta_v))$$
-is valid, whereas the typing judgement
-$$\ell^*:(S, (\partial S \sqcup S)^*, \match(\mathcal{M}_S, \delta_s))
-\xLeftrightarrow{\partial(S, V)^*} (V, (\partial V \sqcup V)^*, \match(\mathcal{M}_V, \delta_v))$$
-is not.
-\iffalse
-For example, if $\ell : S \Leftrightarrow V$ is a lens, Foster's initial
-definition of the iteration lens $\ell^* : S^* \Leftrightarrow V^*$ was
-\begin{align*}
-\ell^* .\get \; (s_1, \ldots, s_n) &= (\ell.\get \; s_1) \cdot \ldots
-\cdot (\ell.\get \; s_n)\\
-\ell^* .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n) &= t_1 \cdot
-\ldots \cdot t_n\\
-\text{ where } t_i &= \begin{cases}
-\ell.\pput \; s_i \; v_i & \text{if } 1 \leq i \leq \min\{m, n\}\\
-\ell.\create \; v_i & \text{if } \min\{m, n\} < i \leq n
-\end{cases}\\
-\ell^*.\create \; (v_1, \ldots, v_n) &= (\ell.\create \; v_1) \cdot \ldots
-\cdot (\ell.\create \; v_n)
-\end{align*}
-This lens always propagates updates positionally, which is unintended behaviour
-in many cases. For example suppose that $\ell$ is a lens that transforms
-source data in the form of an XML document representing the names, dates and
-nationalities of a collection of classical music composers such as,
-
-\begin{lstlisting}
-<composers>
-<composer>
-<name>Jean Sibelius</name>
-<dates>1865-1957</dates>
-<nationality>Finnish</nationality>
-</composer>
-<composer>
-<name>Aaron Copland</name>
-<dates>1910-1990</dates>
-<nationality>American</nationality>
-</composer>
-<composer>
-<name>Benjamin Britten</name>
-<dates>1913-1976</dates>
-<nationality>English</nationality>
-</composer>
-</composers>
-\end{lstlisting}
-into view data stoored in ASCII text representing the names and dates of each
-composer:
-\begin{lstlisting}
-Jean Sibelius, 1865-1957
-Aaron Copland, 1910-1990
-Benjamin Britten, 1913-1976
-\end{lstlisting}
-Now suppose that the elements in the view are reordered:
-\begin{lstlisting}
-Aaron Copland, 1910-1990
-Benjamin Britten, 1913-1976
-Jean Sibelius, 1865-1957
-\end{lstlisting}
-Then with the positional iteration lens, the updated source is
-\begin{lstlisting}
-<composers>
-<composer>
-<name>Aaron Copeland</name>
-<dates>1910-1990</dates>
-<nationality>Finnish</nationality>
-</composer>
-<composer>
-<name>Benjamin Britten</name>
-<dates>1913-1976</dates>
-<nationality>American</nationality>
-</composer>
-<composer>
-<name>Jean Sibelius</name>
-<dates>1865-1957</dates>
-<nationality>English</nationality>
-</composer>
-</composers>
-\end{lstlisting}
-so each of the composers now has the wrong nationality in the source.
-
-\section{Difference functions}
-\begin{definition}
-A weighted set is a pair $(X, |\cdot|)$ where $X$ is a set and $|\cdot| : X
-\longrightarrow \mathbb{R}^{\geq 0}$ is a map from $X$ to the non-negative
-reals.
-\end{definition}
-\begin{definition}
-A difference structure is a triple $(S, \partial S, \delta_S)$ where $S$ and
-$\partial S$ are weighted sets, and $\delta_S:
-S \times S \longrightarrow \partial S$ is a function satisfying
-\begin{align*}
-\delta(s, s') &= \delta(s', s)\tag{A1}\\
-|\delta(s, s)| &= 0\tag{A2}\\
-|\delta(s, s')| &\leq |s| + |s'|\tag{A3}
-\end{align*}
-We refer to the function $\delta$ as the difference function of the structure.
-\end{definition}
-\iffalse
-
-\begin{example}
-Let $X$ be any set. Define $x \leq y$ if and only if $x=y$, and define $|X|=1$
-for all $x \in X$. Let $Q = \{0,1\}$, and define
-$$\delta(a, b) = \begin{cases} 1 & \text{if } a \neq b\\
-0 & \text{otherwise}
-\end{cases}$$
-Then $\delta$ is a difference function.
-\end{example}
-
-Various different edit distances on strings give a way of defining delta
-functions on sets of strings:
-\begin{example}
-The {\em Levenshtein distance} between two words is the minimum number of
-single-character edits (insertions, deletions or substitutions) required to
-change one word into the other. For example, the Levenshtein distance between
-"kitten" and "sitting" is 3, since the following three edits change one into
-the other, and there is no way to do it with fewer than three edits:
-
-\begin{enumerate}
-  \item
-  kitten -> sitten (substitution of "s" for "k")
-  \item
-  sitten -> sittin (substitution of "i" for "e")
-  \item
-  sittin -> sitting (insertion of "g" at the end).
-\end{enumerate}
-\end{example}
-\begin{example}
-In contrast, the Longest Common Subsequence(LCS) distance only allows for
-insertions and deletions:
-\begin{enumerate}
-  \item
-  delete k at 0
-  \item
-  insert s at 0
-  \item
-  delete e at 4
-  \item
-  insert i at 4
-  \item
-  insert g at 6
-\end{enumerate}
-for a total cost/distance of 5 operations.
-\end{example}
-Both Levenshtein and LCS distances are integer-valued distance functions since
-they are metrics.
-\begin{example}
-If all the strings in a set are of the same length, then the {\em Hamming
-distance} is a difference function since the Hamming distance is also a metric.
-\end{example}
-\begin{example}
-Alignment algorithms also give us a way of defining integer-valued distance
-functions on sets of strings by taking the difference of two strings to be the
-total number of characters that are not equal according to the alignment.
-
-When deciding which alignment to choose between two strings, we often wish to
-choose an alignment that is optimal for alignment strategy. With
-the positional alignment strategy, given two strings $s$ and $t$, then $s_1$ is
-matched to $t_1$, $s_2$ is matched to $t_2$ and so on.
-Also, $\mathcal{M}_{POS}$ is an optimal set of alignments for a positional
-alignment strategy.
-
-For example, using the positional alignment strategy on the strings "kitten" and
-"sitting", then "k" is matched to "s", "i" to "i" and so on. The characters
-which are included in the delta from "kitten" are "k", "e" while the
-in the characters included from "sitting" are "s", "i" and "g", hence the delta
-is equal to 5.
-
-For the non-crossing strategy, we choose among alignments which are increasing
-rather than positional, while for the min-edit distance strategy, we consider
-all possible alignments.
-\end{example}
-
-\begin{example}\label{naturalnumbers}
-Let $P$ be the set of non-negative integers with weight function $|a| = a$. Then
-$\delta$ is a difference function.
-Define $\delta(a, b) = |a-b|$. Then $\delta$ is a difference function.
-\end{example}
-\begin{example}
-Let $X$ be a normed vector space and define $\delta(x, y) = |x-y|$. Then
-$\delta$ is a difference function (in fact, $\delta$ is also a metric).
-\end{example}
-\begin{example}\label{divisibility}
-Let $P$ be the set of positive integers with weight function $|\cdot|$ defined
-by $|p_1^{a_1} \cdot \ldots \cdot p_n^{a_n}| = a_1 + \ldots + a_n$. Define
-$\delta(a, b) = ab/(gcd(a, b))^2$.
-Then $\delta$ is a difference function.
-\end{example}
-\begin{example}\label{boolean}
-Let $X$ be a set and $P$ the set of finite subsets of $X$ with weight function
-given by the size of the set. Define $\delta(A, B) = (A \setminus B) \cup (B
-\setminus A)$. Then $\delta$ is a difference function.
-\end{example}
-\begin{example}\label{general}
-Generalizing \cref{divisibility} and
-\cref{boolean}, let $P$ be a graded complete semi-modular lattice satisfying
-$|\bigwedge P|=0$. Define $\delta(a, b) = \bigwedge \{x \; | \; (a \wedge b)
-\vee x = a\} \vee \bigwedge \{x \; | \; (a \wedge b) \vee x = b\}$. Then
-$\delta$ is a difference function.
-\end{example}
-\begin{proof}
-Let $a, b \in P$. Let $X = \{x \; | \; (a \wedge b) \vee x = a\}$ and $Y = \{y
-\; | \; (a \wedge b) \vee y = b\}$
-\begin{enumerate}
-  \item[(A1)]
-  Certainly $\delta(a, b) = \delta(b, a)$.
-  \item[(A2)]
-  Observe that $\bigwedge X = \bigwedge Y = \bot$, hence $\delta(a,
-  a) = \bot$, and $|\delta(a, a)| = |\bot| = 0$.
-  \item[(A3)]
-  Since $P$ is semi-modular, then $|x \wedge y| + |x \vee y| \leq |x| + |y|$
-  for all $x, y \in P$, from which it follows that $$|\delta(a, b)| =
-  \left|\bigwedge X \vee \bigwedge Y\right| \leq \left|\bigwedge X \vee
-  \bigwedge Y\right| + \left|\bigwedge X \wedge \bigwedge Y\right| \leq
-  \left|\bigwedge X\right| + \left|\bigwedge Y\right|$$
-\end{enumerate}
-\end{proof}
-\fi
-
-\subsection{Constructions on difference functions}
-\subsubsection{Products}
-Given difference functions $\delta_i: P_i \times P_i \longrightarrow Q_i$ for
-$i \in \{1,2\}$, define $\delta_1 \times \delta_2 :(P_1 \times P_2) \times
-(P_1 \times P_2) \longrightarrow (Q_1 \times Q_2)$ by
-$$(\delta_1 \times \delta_2)((x, y), (x', y')) = (\delta_1(x, x'), \delta_2(y,
-y'))$$
+\cref{difflenslaw} can be expressed by a relation $S_{\ell} \subseteq (S \times S) \times (V \times V)$ by $(\ell.\pput \; s \;v, s)$ is related to $(v, \ell.\get \; s)$ if and only if \cref{difflenslaw} holds.
+
+In the following part of the paper we describe a language for defining difference lenses on regular languages.
+\subsubsection{Copy}
+Let $R$ be any unambiguous difference regular expression. Define the lens $\ccopy(R) : R \Leftrightarrow R$ by
+\begin{align}
+\ccopy(R).\get \; s &= s\\
+\ccopy(R).\pput \; s \; v &= v\\
+\ccopy(R).\create \; v &= v
+\end{align}
 \begin{claim}
-$\delta_1 \times \delta_2$ is a difference function.
-\end{claim}
-\begin{proof}
-Let $x, x' \in P_1$ and $y, y' \in P_2$.
-\begin{enumerate}
-  \item[(A1)]
-  Observe that
-  \begin{align*}
-(\delta_1 \times \delta_2)((x, y), (x', y')) &= (\delta_1(x, x'),
-\delta_2(y, y'))\\
-&= (\delta_1(x', x), \delta_2(y', y))\\
-&= (\delta_1 \times \delta_2)((x', y'), (x, y))
-\end{align*}
-\item[(A2)] Observe that
-$$|(\delta_1 \times \delta_2)((x, y), (x, y))| = |(\delta_1(x, x),
-\delta_2(y, y))| = |\delta_1(x, x)| + |\delta_2(y,y)| = 0 + 0 = 0$$
-\item[(A3)]
-Observe that
-\begin{align*}
-|(\delta_1 \times \delta_2)((x, y), (x', y'))| &= |(\delta_1(x, x'),
-\delta_2(y, y'))|\\
-&= |(\delta_1(x, x')| + |(\delta_2(y, y')|\\
-&\leq (|x| + |x'|) + (|y| + |y'|)\\
-&= (|x| + |y|) + (|x'| + |y'|)\\
-&= |(x, y)| + |(x', y')|
-\end{align*}
-\end{enumerate}
-\end{proof}
-\subsubsection{Coproducts}
-Given difference functions $\delta_i: P_i \times P_i \longrightarrow Q_i$ for
-$i \in \{1,2\}$, define $\delta_1 \sqcup \delta_2 :P_1 \sqcup P_2
-\longrightarrow Q_1 \sqcup Q_2 \sqcup (P_1 \times P_2)$ by
-$$(\delta_1 \sqcup \delta_2)(x, y) =
-\begin{cases}
-\delta_1(x, y) & \text{if }x, y \in P_1\\
-\delta_2(x, y) & \text{if }x, y \in P_2\\
-(x, y) & \text{if }x \in P_1, y \in P_2\\
-(y, x) & \text{if }x \in P_2, y \in P_1
-\end{cases}$$
-\begin{claim}
-$\delta_1 \sqcup \delta_2$ is a difference function
-\end{claim}
-\begin{proof}
-Let $x, x' \in P_1$ and $y, y' \in P_2$.
-\begin{enumerate}
-  \item[(A1)]
-  If $x, y \in P_i$, then $(\delta_1 \sqcup \delta_2)(x, y) = \delta_i(x, y) =
-  \delta_i(y, x)$. If $x \in P_1$ and $y \in P_2$, then $(\delta_1 \sqcup
-  \delta_2)(x, y) = (x, y) = (\delta_1 \sqcup \delta_2)(y, x)$, and similarly
-  if $x \in P_2$ and $y \in P_1$, then $(\delta_1 \sqcup \delta_2)(x, y) =
-  (y,x) = (\delta_1 \sqcup \delta_2)(y, x)$
-  \item[(A2)] If $x \in P_i$, then $|(\delta_1 \sqcup \delta_2)(x, x)| =
-  |\delta_i(x, x)| = 0$
-  \item[(A3)]
-  If $x, y \in P_i$, then $|(\delta_1 \sqcup \delta_2)(x, y)| = |\delta_i(x, y)|
-  \leq |x| + |y|$. If $x \in P_1$ and $y \in P_2$, then $|(\delta_1 \sqcup
-  \delta_2)(x, y)| = |(x, y)| = |x| + |y|$, and if $x \in P_2$ and $y \in
-  P_1$, then $|(\delta_1 \sqcup \delta_2)(x, y)| = (|(y, x)|) = |y| + |x|$.
-\end{enumerate}
-\end{proof}
-\subsubsection{List difference functions}
-Let $P$ be a weighted set. Let $x=(x_1, \ldots, x_m), y=(y_1, \ldots, y_n)$ be
-in $P^*$. Given a difference function $\delta: P \times P \longrightarrow Q$, we
-define the list difference function $\delta^*:P^* \longrightarrow (Q \sqcup
-P)^*$ on $x$ and $y$ to be the list given by $\delta(x_i, y_i)$ for $1 \leq i
-\leq m$, except that we omit $\delta(x_i, y_i)$ in position $i$ if $x_i = y_i$.
-For $m < i \leq n$ we add $y_i$ at position $i$. Finally we collapse the list
-to get rid of the empty spaces left after the omissions.
-
-For instance suppose that $S$ is the set of all letters in the alphabet and
-$\partial S = \{\epsilon\} \sqcup \{\{x, y\} \; | \; x, y \in S, x \neq y\}$
-with $$\delta(x, y) =
-\begin{cases}
-\epsilon & \text{if } x = y\\
-\{x,y\} & \text{if } x \neq y
-\end{cases}$$
-Then $\delta_S^*((a, c, b, d), (a, b, c, d, e)) = (\{b, c\}, \{b, c\},
-e)$ where we match and remove the first $a$ and the fourth $d$ in. Since $b \neq
-c$ we have $\{b,c\}$ in the second and third positions, and since $e$ is
-unmatched, we include $e$ in the fifth position. Finally we collapse the list to
-get rid of the two empty spaces created.
-
-\begin{claim}
-$\delta^*$ is a difference function
-\end{claim}
-\begin{proof}
-Let $x = (x_1, \ldots, x_m), y = (y_1, \ldots, y_n) \in P^*$.
-\begin{enumerate}
-  \item[(A1)]
-  By construction $\delta^*(x, y) = \delta^*(y, x)$.
-  \item[(A2)] Observe that $|\delta^*(x, x)| = \sum_{i=1}^m|\delta(x_i,
-  x_i)| =  \sum_{i=1}^m 0 = 0$.
-  \item[(A3)]
-  Assume WLOG that $m \leq n$. Then
-  \begin{align*}
-|\delta^*(x, y)| &=
-\sum_{i=1}^m|\delta(x_i, y_i)| + \sum_{i=m+1}^n |y_i|\\
-&\leq \sum_{i=1}^m|x_i| + |y_i| + \sum_{i=m+1}^n |y_i|\\
-&= \sum_{i=1}^m|x_i| + \sum_{i=1}^n |y_i|\\
-&= |x| + |y|
-\end{align*}
-\end{enumerate}
-\end{proof}
-\subsubsection{Matching difference functions}
-Now let $\mathcal{M} = \{\varphi_{(x_1, \ldots, x_m), (y_1, \ldots, y_n)} \; |
-\; x_i, y_j \in P, m \leq n\}$ be a set of injective functions $\varphi_{(x_1,
-\ldots, x_m), (y_1, \ldots, y_n)} : [m] \longrightarrow [n]$. Given a difference
-function $\delta:
-P \times P \longrightarrow Q$, we define the difference function
-$\match(\mathcal{M}, \delta):P^* \longrightarrow (Q \sqcup P)^*$ on $x$
-and $y$ to be the list given by $\delta(x_i, y_{\varphi(i)})$ in position
-$\varphi(i)$ for $1 \leq i \leq m$, except that we omit $\delta(x_i,
-y_{\varphi(i)})$ in position $\varphi(i)$ if $x_i = y_{\varphi(i)}$. For $i \in
-[n] \setminus \varphi([m])$ we add $y_i$ at position $i$. Finally we collapse
-the list to get rid of the empty spaces left after the omissions.
-
-Going back to our previous example, let $x = \{a, c, b, d\}$ and $y = \{a, b, c,
-d, e\}$, and assume that $\varphi_{x, y} : [4] \longrightarrow [5]$ is given by
-$$\varphi(1) = 1, \; \varphi(2) = 3, \; \varphi(3) = 2, \text{ and } \varphi(4)
-= 4$$ Then $\match(\mathcal{M}, \delta)((a, c, b, d), (a, b, c, d, e)) =
-(e)$.
-\begin{claim}
-$\match(\mathcal{M}, \delta)$ is a difference function.
-\end{claim}
-\begin{proof}
-Let $x = \{x_1, \ldots, x_m\}, y = \{y_1, \ldots, y_n\} \in \match(P)$.
-\begin{enumerate}
-  \item[(A1)]
-  By construction $\match(\mathcal{M}, \delta)(x, y) =
-  \match(\mathcal{M}, \delta)(y, x)$.
-  \item[(A2)] Observe that $|(\match(\mathcal{M}, \delta))(x, x)| =
-  \sum_{i=1}^m|\delta(x_i, x_{\varphi(i)}| = |\delta(x_i, x_i)| =   \sum_{i=1}^m
-  0 = 0$.
-  \item[(A3)]
-  Assume WLOG that $m \leq n$. Then
-  \begin{align*}
-|\match(\mathcal{M}, \delta)(x, y)| &=
-\sum_{i=1}^m|\delta(x_i, y_{\varphi(i)})| + \sum_{i \in [n] \setminus
-\varphi([m])} |y_i|\\
-&\leq \sum_{i=1}^m|x_i| + |y_{\varphi(i)}| + \sum_{i \in [n] \setminus
-\varphi([m])} |y_i|\\
-&= \sum_{i=1}^m|x_i| + \sum_{i=1}^n |y_i|\\
-&= |x| + |y|
-\end{align*}
-\end{enumerate}
-\end{proof}
-\fi
-\section{Lens Combinators}
-\subsection{Copy}
-Let $P$ be any set. Define the $\ccopy(P) : P \Leftrightarrow P$ by
-\begin{align}
-\ccopy(P).\get \; s &= s\\
-\ccopy(P).\pput \; s \; v &= v\\
-\ccopy(P).\create \; v &= v
-\end{align}
-Then $\ccopy(P)$ is a lens. Now let $\partial S$ be a set, $\delta :
-S \times S \longrightarrow \partial S$ a function.
-\begin{claim}
-$\ccopy(S) : S \xLeftrightarrow{\delta , \ccopy(\partial S), \delta} S$.
-\end{claim}
-\begin{proof}
-We need to show that $(\ccopy(\partial S)).\get \; (\delta(\ccopy(S).\pput
-\; s \; v, s)) = \delta(v, \ccopy(S).\get \; s)$ which follows from the fact
-that $$ (\ccopy(\partial S)).\get \; (\delta(\ccopy(S).\pput \; s
-\; v, s)) = \delta(\ccopy(S).\pput \; s \; v, s) = \delta(v, \ccopy(S).\get
-\; s) $$
-\end{proof}
-\subsection{Constant}
-Let $S$ be any set with $s'\in S$. Define the \const lens $\const(S) : S
-\Leftrightarrow \{\star\}$ by
-\begin{align*}
-\const.\get \; s &= \star\\
+If $R$ is an unambiguous difference regular expression, then $\ccopy(R) : R \Leftrightarrow R$ is a difference lens.
+\end{claim}
+\begin{proof}
+We need to show that $\delta_R(\ccopy(S).\pput
+\; s \; v, s)) \geq \delta_R(v, \ccopy(S).\get \; s)$ which follows from the fact that $\delta_R(\ccopy(S).\pput \; s
+\; v, s)) = \delta_R(v, s) = \delta_R(v, \ccopy(S).\get \; s) $.
+\end{proof}
+\subsubsection{Constant}
+Let $R$ be any unambiguous difference regular expression and $c \in \Sigma^*$. Assume that $s' \in \mathcal{L}(R)$. Define the \const lens $\const(R, c) : R
+\Leftrightarrow c$ by
+\begin{align*}
+\const.\get \; s &= c\\
 \const.\pput \; s \; v &= s\\
@@
 \end{align*}
-Then $\const(S)$ is a lens. 
 \begin{claim}
-Let $\partial S$ be any set, and $\delta : S \times S \longrightarrow \partial
-S$ any function satisfying $\delta(s, s) = \delta(s', s')$ for all $s, s' \in
-S$. Define $\beta : \{\star\} \times \{\star\} \longrightarrow \partial S$ by
-$\beta(\star, \star) = \delta(s, s)$ for any $s \in S$. Let $\ell : \partial S
-\Leftrightarrow \partial S$ be any lens satisfying $\ell.\get \; \delta(s, s) =
-\delta(s, s)$ for all $s \in S$. Then $\const(S): S \xLeftrightarrow{\delta,
-\ell, \beta} \{\star\}$.
-\end{claim}
-\begin{proof}
-We need to show that $\ell.\get \; (\delta(\const(S).\pput \; s \; \star, s)) =
-\beta (\star, \const(S).\get \; s)$
-which follows from the fact that
-$$ \ell.\get \; (\delta(\const(S).\pput \; s \; \star, s)) = \ell.\get
-\; \delta(s, s) = \delta(s, s) = \beta(\star, \star) = \beta(\star,
-\const(S).\get \; s)$$
-\end{proof}
-
-\subsection{Default}
-Let $\ell : S \Leftrightarrow V$ be a lens. Let $f :
+If $R$ is an unabmbiguous difference regular expression, then $\const(R, c)$ is a difference lens.
+\end{claim}
+\begin{proof}
+We need to show that $\delta_R(\const(R, c).\pput
+\; s \; v, s)) \geq \delta_c(v, \const(R, c).\get \; s)$ which follows from the fact that $\delta_R(\const(R, c).\pput
+\; s \; v, s)) = \delta_R(s, s) = 1 = \delta_c(c, c) = \delta_R(v, \const(R, c).\get \; s)$.
+\end{proof}
+
+\subsubsection{Default}
+Let $\ell : S \Leftrightarrow V$ be a difference lens. Let $f :
 V \longrightarrow S$ be any function. Define the lens $\default(\ell, f) :
@@
 \begin{claim}
-If $\ell : S \xLeftrightarrow{\delta_S, \partial (S, V), \partial V}
-V$ and $f: S \longrightarrow V$, then $\default(\ell, f) : S
-\xLeftrightarrow{\delta_S, \partial (S, V), \partial V} V$.
-\end{claim}
-\begin{proof}
-Follows from the fact that $\ell : S \xLeftrightarrow{\delta_S, \partial (S, V),
-\partial V} V$.
-\end{proof}
-\subsection{Product}
+If $\ell : S \Leftrightarrow V$ is a difference lens and $f: S \longrightarrow V$, then $\default(\ell, f) : S \Leftrightarrow V$ is also a difference lens.
+\end{claim}
+\begin{proof}
+Follows from the fact that $\ell$ is a difference lens.
+\end{proof}
+\subsubsection{Product}
+Let $\ell_1 : S_1 \Leftrightarrow V_1$ and $\ell_2 : S_2 \Leftrightarrow V_2$ be difference
+lenses. If $S_1 \cdot^{!} S_2$ and $V_1 \cdot^{!} V_2$, then define the lens $\ell_1 \cdot \ell_2 : S_1 \cdot S_2 \Leftrightarrow
+V_1 \cdot V_2$ by
+\begin{align*}
+(\ell_1 \cdot \ell_2).\get \; (s_1 \cdot s_2) &= (\ell_1.\get \; s_1) \cdot (\ell_2.\get \; s_2)\\
+(\ell_1 \cdot \ell_2).\pput \; (s_1 \cdot s_2) \; (v_1 \cdot v_2) &= (\ell_1.\pput \; s_1
+\; v_1) \cdot (\ell_2.\pput \; s_2 \; v_2)\\
+(\ell_1 \cdot \ell_2).\create \; (v_1 \cdot v_2) &= (\ell_1.\create \; v_1) \cdot (\ell_2.\create \; v_2)
+\end{align*}
+\begin{claim}\label{productislens}
+If $\ell_1 :S_1 \Leftrightarrow V_1$ and $\ell_2 : S_2 \Leftrightarrow V_2$ are difference lenses with $S_1 \cdot^{!} S_2$ and $V_1 \cdot^{!} V_2$, then
+$\ell_1 \cdot \ell_2  : S_1 \cdot S_2 \Leftrightarrow V_1 \cdot V_2$ is also a difference lens.
+\end{claim}
+\begin{proof}
+We need to show that
+$$\delta((\ell_1 \cdot
+\ell_2).\pput \; (s_1 \cdot s_2) \; (v_1 \cdot v_2), (s_1, s_2))) = \delta((v_1 \cdot v_2),
+(\ell_1 \cdot \ell_2).\get \; (s_1 \cdot s_2))$$ which follows from the fact that
+\begin{align*}
+\delta((\ell_1 \cdot
+\ell_2).\pput \; (s_1 \cdot s_2) \; (v_1 \cdot v_2), (s_1 \cdot s_2))) &= \delta((\ell_1.\pput \; s_1 \; v_1) \cdot
+(\ell_2.\pput \; s_2 \; v_2), (s_1 \cdot s_2)))  \\
+&= \frac{1}{2}(\delta(\ell_1.\pput \; s_1 \; v_1, s_1) + \delta(\ell_1.\pput \; s_2 \; v_2, s_2))\\
+&\geq \frac{1}{2}(\delta(v_1, \ell_1.\get \; s_1) + \delta(v_2, \ell.\get \; s_2))\\
+&= \delta((v_1 \cdot v_2), (\ell_1.\get \; s_1) \cdot (\ell_2.\get \; s_2))\\
+&= \delta((v_1 \cdot v_2),
+(\ell_1 \cdot \ell_2).\get \; (s_1 \cdot s_2))
+\end{align*}
+\end{proof}
+\subsubsection{Swap}
+Let $\ell_1 : S_1 \Leftrightarrow V_1$ and $\ell_2 : S_2 \Leftrightarrow V_2$ be difference
+lenses. If $S_1 \cdot^{!} S_2$ and $V_2 \cdot^{!} V_1$, then define the lens $\ell_1 \sim \ell_2 : S_1 \cdot S_2 \Leftrightarrow
+V_2 \cdot V_1$ by
+\begin{align*}
+(\ell_1 \sim \ell_2).\get \; (s_1 \cdot s_2) &= (\ell_2.\get \; s_2) \cdot (\ell_1.\get \; s_1)\\
+(\ell_1 \sim \ell_2).\pput \; (s_1 \cdot s_2) \; (v_2 \cdot v_1) &= (\ell_1.\pput \; s_1
+\; v_1) \cdot (\ell_2.\pput \; s_2 \; v_2)\\
+(\ell_1 \sim \ell_2).\create \; (v_2 \cdot v_1) &= (\ell_1.\create \; v_1) \cdot (\ell_2.\create \; v_2)
+\end{align*}
+\begin{claim}\label{productislens}
+If $\ell_1 :S_1 \Leftrightarrow V_1$ and $\ell_2 : S_2 \Leftrightarrow V_2$ are difference lenses with $S_1 \cdot^{!} S_2$ and $V_2 \cdot^{!} V_1$, then
+$\ell_1 \sim \ell_2  : S_1 \cdot S_2 \Leftrightarrow V_2 \cdot V_1$ is also a difference lens.
+\end{claim}
+\begin{proof}
+The proof is similar to the concatenation case.
+\end{proof}
+\subsubsection{Sum}
 Let $\ell_1 : S_1 \Leftrightarrow V_1$ and $\ell_2 : S_2 \Leftrightarrow V_2$ be
-lenses. Define the lens $\ell_1 \times \ell_2 : (S_1 \times S_2) \Leftrightarrow
-(V_1 \times V_2)$ by
-\begin{align*}
-(\ell_1 \times \ell_2).\get \; (s_1, s_2) &= (\ell_1.\get \; s_1, \ell_2.\get
-\; s_1)\\
-(\ell_1 \times \ell_2).\pput \; (s_1, s_2) \; (v_1, v_2) &= (\ell_1.\pput \; s_1
-\; v_1, \ell_2.\pput \; s_2 \; v_2)\\
-(\ell_1 \times \ell_2).\create \; (v_1, v_2) &= (\ell_1.\create \; v_1,
-\ell_2.\create \; v_2)
-\end{align*}
-\begin{claim}\label{productislens}
-If $\ell_1 :S_1 \xLeftrightarrow{\delta_{S_1}, \partial (S_1, V_1),
-\delta_{V_1}} V_1$ and $\ell_2 : S_2 \xLeftrightarrow{\delta_{S_2}, \partial
-(S_2, V_2), \delta_{V_2}} V_2$, then \\
-$\ell_1 \times \ell_2  : S_1 \times S_2 
-\xLeftrightarrow{\delta_{S_1} \times \delta_{S_2}, \partial (S_1, V_1) \times
-\partial (S_2, V_2), \delta_{V_1} \times \delta_{V_2}} V_1 \times V_2$.
-\end{claim}
-\begin{proof}
-We need to show that
-$$(\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \; (\delta((\ell_1 \times
-\ell_2).\pput \; (s_1, s_2) \; (v_1, v_2), (s_1, s_2))) = \delta((v_1, v_2),
-(\ell_1 \times \ell_2).\get \; (s_1, s_2))$$ which follows from the fact that
-\begin{align*}
-&(\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \; (\delta((\ell_1 \times
-\ell_2).\pput \; (s_1, s_2) \; (v_1, v_2), (s_1, s_2)))\\
-&= (\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \;
-(\delta((\ell_1.\pput \; s_1 \; v_1, \ell_2.\pput \; s_2 \; v_2), (s_1, s_2))\\
-&= (\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \;
-(\delta(\ell_1.\pput \; s_1 \; v_1, s_1), \delta(\ell_2.\pput \; s_2 \; v_2,
-s_2))\\
-&= ((\partial (S_1, V_1)).\get \; (\delta(\ell_1.\pput \;
-s_1 \; v_1, s_1)), (\partial (S_2, V_2)).\get \; (\delta(\ell_2.\pput \; s_2 \;
-v_2, s_2)))\\
-&= (\delta(v_1, \ell_1.\get \; s_1), \delta(v_2, \ell_2.\get \; s_2))\\
-&= \delta((v_1, v_2), (\ell_1.\get \; s_1, \ell_2.\get \; s_2))\\
-&= \delta((v_1, v_2), (\ell_1 \times \ell_2).\get \; (s_1, s_2))
-\end{align*}
-\end{proof}
-\subsection{Sum}
-Let $\ell_1 : S_1 \Leftrightarrow V_1$ and $\ell_2 : S_2 \Leftrightarrow V_2$ be
-lenses. Define the lens
-$\ell_1 \sqcup \ell_2 :
-(S_1 \sqcup S_2) \Leftrightarrow (V_1 \sqcup V_2)$ by
-\begin{align*}
-(\ell_1 \sqcup \ell_2).\get \; s &=
+difference lenses with $\mathcal{L}(S_1) \cap \mathcal{L}(S_2) = \varnothing$ and $\mathcal{L}(V_1) \cap \mathcal{L}(V_2) = \varnothing$. Define the lens
+$\ell_1 \cup \ell_2 :
+S_1 \cup S_2 \Leftrightarrow V_1 \cup V_2$ by
+\begin{align*}
+(\ell_1 \cup \ell_2).\get \; s &=
 \begin{cases}
@@
 \end{cases}\\
-(\ell_1 \sqcup \ell_2).\pput \; s \; v&=
+(\ell_1 \cup \ell_2).\pput \; s \; v&=
 \begin{cases}
@@
 \end{cases}\\
-(\ell_1 \sqcup \ell_2).\create \; v &=
+(\ell_1 \cup \ell_2).\create \; v &=
 \begin{cases}
@@
 \begin{claim}
-If $\ell_1 : S_1  \xLeftrightarrow{\delta_{S_1}, \partial (S_1, V_1),
-\delta_{V_1}} V_1$ and $\ell_2 : S_2 \xLeftrightarrow{\delta_{S_2}, \partial
-(S_2, V_2), \delta_{V_2}} V_2$, then \\
-$\ell_1 \sqcup \ell_2 : S_1 \sqcup S_2 \xLeftrightarrow{\delta_{S_1} \sqcup
-\delta_{S_2}, \partial (S_1, V_1) \sqcup \partial (S_2, V_2) \sqcup (\ell_1
-\times \ell_2), \delta_{V_1} \sqcup \delta_{V_2}} V_1 \sqcup V_2$
-\end{claim}
-\begin{proof}
+If $\ell_1 : S_1 \Leftrightarrow V_1$ and $\ell_2 : S_2 \Leftrightarrow V_2$ are difference lenses, with $\mathcal{L}(S_1) \cap \mathcal{L}(S_2) = \varnothing$ and $\mathcal{L}(V_1) \cap \mathcal{L}(V_2) = \varnothing$ then $\ell_1 \cup \ell_2 :
+S_1 \cup S_2 \Leftrightarrow V_1 \cup V_2$ is also a difference lens.
+\end{claim}
+\begin{proof}
+If the source and the view are in $S_i$ and $V_i$ respectively for $i \in \{1,2\}$, then the result holds from the fact that $\ell_1$ and $\ell_2$ are difference lenses, otherwise the difference on both the source and view side are equal to 0, from which the result follows.
+\end{proof}
+\subsubsection{Iteration}
+Let $\ell : S \Leftrightarrow V$ is a difference lens with $S^{!*}$ and $V^{!*}$.
+Let $s_1 \cdot \ldots \cdot s_m$ be a source, and $v_1 \cdot \ldots \cdot v_n$ a view.
+If $n \leq m$ and $\varphi : [n] \longrightarrow [m]$, define $ v \xhookrightarrow{\ell, \varphi} s = s'_1 \cdot \ldots \cdot s'_n$ by
+$$s'_i = \begin{cases}
+\ell.\pput \; s_{\varphi(i)} \; v_i & \text{if } i \in Dom(\varphi)\\
+\ell.\create \; v_i & \text{if } i \not \in Dom(\varphi)
+\end{cases}$$
+If $m < n$ and $\varphi : [m] \longrightarrow [n]$, define $ v \xhookrightarrow{\ell, \varphi} s = s'_1 \cdot \ldots \cdot s'_n$ by 
+$$s'_i = \begin{cases}
+\ell.\pput \; s_{\varphi^{-1}(i)} \; v_i & \text{if } i \in \varphi([m])\\
+\ell.\create \; v_i & \text{if } i \not \in \varphi([m])
+\end{cases}$$
+
+Using the notation $len(v)$ and $len(s)$ to mean $len(v) = n$ if $v = v_1 \cdot \ldots \cdot v_n$ and $len(s) = m$ if $s = s_1 \cdot \ldots \cdot s_m$, define the lenses $\ell^*_{POS} : S^*_{POS} \Leftrightarrow V^*_{POS}$, $\ell^*_{NC} : S^*_{NC} \Leftrightarrow V^*_{NC}$ and $\ell^*_{MED} : S^*_{MED} \Leftrightarrow V^*_{MED}$ by:
+\begin{align*}
+\ell^*_{POS}.\get \; (s_1 \cdot  \ldots \cdot s_n) &= (\ell.\get \; s_1)
+\cdot \ldots \cdot (\ell.\get \; s_n)\\
+\ell^*_{POS}.\pput \; s \; v
+&= v \xhookrightarrow{\psi} s\\
+\text{where } \psi &= \begin{cases}
+\argmax \limits_{\{\varphi : [len(v)] \longrightarrow [len(s)] \; | \; \varphi(i) = i\}} \delta_{S^*_{NC}}(v \xhookrightarrow{\ell, \varphi} s, s) & \text{if } len(v) \leq len(s)\\
+\argmax \limits_{\{\varphi : [len(s)] \longrightarrow [len(v)] \; | \; \varphi(i) = i\}} \delta_{S^*_{NC}}(v \xhookrightarrow{\ell, \varphi} s, s) & \text{if } len(s) < len(v)
+\end{cases}\\
+\ell^*_{POS}.\create \; (v_1, \ldots, v_n) &= (\ell.\create \;
+v_1) \cdot \ldots \cdot (\ell.\create \; v_n)
+\end{align*}
+\begin{align*}
+\ell^*_{NC}.\get \; (s_1 \cdot  \ldots \cdot s_n) &= (\ell.\get \; s_1)
+\cdot \ldots \cdot (\ell.\get \; s_n)\\
+\ell^*_{NC}.\pput \; s \; v
+&= v \xhookrightarrow{\psi} s\\
+\text{where } \psi &= \begin{cases}
+\argmax \limits_{\{\varphi : [len(v)] \longrightarrow [len(s)] \; | \; \varphi \text{ is str. increasing}\}} \delta_{S^*_{NC}}(v \xhookrightarrow{\ell, \varphi} s, s) & \text{if } len(v) \leq len(s)\\
+\argmax \limits_{\{\varphi : [len(s)] \longrightarrow [len(v)] \; | \; \varphi \text{ is str. increasing}\}} \delta_{S^*_{NC}}(v \xhookrightarrow{\ell, \varphi} s, s) & \text{if } len(s) < len(v)
+\end{cases}\\
+\ell^*_{NC}.\create \; (v_1, \ldots, v_n) &= (\ell.\create \;
+v_1) \cdot \ldots \cdot (\ell.\create \; v_n)
+\end{align*}
+\begin{align*}
+\ell^*_{MED}.\get \; (s_1 \cdot  \ldots \cdot s_n) &= (\ell.\get \; s_1)
+\cdot \ldots \cdot (\ell.\get \; s_n)\\
+\ell^*_{MED}.\pput \; s \; v
+&= v \xhookrightarrow{\ell, \psi} s\\
+\text{where } \psi &= \begin{cases}
+\argmax \limits_{\{\varphi : [len(v)] \longrightarrow [len(s)] \; | \; \varphi \text{ is injective}\}} \delta_{S^*_{NC}}(v \xhookrightarrow{\ell, \varphi} s, s) & \text{if } len(v) \leq len(s)\\
+\argmax \limits_{\{\varphi : [len(s)] \longrightarrow [len(v)] \; | \; \varphi \text{ is injective}\}} \delta_{S^*_{NC}}(v \xhookrightarrow{\ell, \varphi} s, s) & \text{if } len(s) < len(v)
+\end{cases}\\
+\ell^*_{MED}.\create \; (v_1, \ldots, v_n) &= (\ell.\create \;
+v_1) \cdot \ldots \cdot (\ell.\create \; v_n)
+\end{align*}
+For the non-crossing and min-edit distance lens, there may be multiple alignments $\psi$ that maximize the respective differences, in which case the lens chooses the lexicographically smallest of those alignments.
+
+\begin{claim}
+If $\ell : S \Leftrightarrow V$ is a difference lens with $S^{!*}$ and $V^{!*}$, then $\ell^*_{POS}$, $\ell^*_{NC}$ and $\ell^*_{MED}$ are also difference lenses.
+\end{claim}
+\begin{proof}
+We will show that $\ell^*_{MED}$ is a difference lens; the proof for each of the other lenses is similar.
+
 We need to show that
-$$(\partial (S_1, V_1) \sqcup \partial (S_2, V_2)).\get \;
-(\delta((\ell_1 \sqcup \ell_2).\pput \; s \; v, s)) = \delta(v, (\ell_1 \sqcup
-\ell_2).\get \; s)$$
-
-The result holds inductively if $s \in S_1$ and $v \in V_1$ or $s \in S_2$ and
-$v \in V_2$. Otherwise, assume without loss of generality
-that $s \in S_1$ and $v \in V_2$. Then
-\begin{align*}
-(\partial (S_1, V_1) \sqcup \partial (S_2, V_2)).\get \;
-(\delta((\ell_1 \sqcup \ell_2).\pput \; s \; v, s)) &=
-(\partial (S_1, V_1) \sqcup \partial (S_2, V_2)).\get \; (\delta(\ell_2.\create
-\; v, s))\\
-&= (\ell_1 \times \ell_2).\get \; (s, \ell_2.\create \; v)\\
-&= (\ell_1.\get \; s, \ell_2.\get \; (\ell_2.\create \; v))\\
-&= (\ell_1.\get \; s, v)\\
-&= \delta(v, \ell_1.\get \; s)\\
-&= \delta(v, (\ell_1 \sqcup \ell_2).\get \; s)
-\end{align*}
-\end{proof}
-\subsection{Positional Iteration}
-Let $\ell : S \Leftrightarrow V$ be a lens and define the lens $\ell^*:
-S^* \Leftrightarrow V^*$ by
-\begin{align*}
-\ell^* .\get \; (s_1, \ldots, s_n) &= (\ell.\get \; s_1) \cdot \ldots
-\cdot (\ell.\get \; s_n)\\
-\ell^*.\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n) &= s'_1 \cdot
-\ldots \cdot s'_n\\
-\text{ where } s'_i &= \begin{cases}
-\ell.\pput \; s_i \; v_i & \text{if } 1 \leq i \leq \min\{m, n\}\\
-\ell.\create \; v_i & \text{if } \min\{m, n\} < i \leq n
-\end{cases}\\
-\ell^*.\create \; (v_1, \ldots, v_n) &= (\ell.\create \; v_1) \cdot \ldots
-\cdot (\ell.\create \; v_n)
-\end{align*}
-
-\begin{claim}\label{positionalisvalid}
-If $\ell: S \xLeftrightarrow{\delta_S, \partial (S, V), \delta_V}
-V$, then $\ell^*: S^* \xLeftrightarrow{\delta_{S}^*, (\partial (S, V) \sqcup
-\ell)^*, \delta_{V}^*} V^*$.
-\end{claim}
-\begin{proof}
-We need to show that
-\begin{multline}
-(\partial (S, V) \sqcup \ell)^*.\get \;
-(\delta_S^*(\ell^*.\pput \; (s_1, \ldots, s_n) \; (v_1, \ldots,
-v_m), (s_1, \ldots, s_n))) \\
-= \delta_V^*((v_1, \ldots, v_m), \ell^*.\get \; (s_1, \ldots,
-s_n))
-\end{multline}
-Let $1 \leq i \leq \min\{m, n\}$. Observe that $v_i = \ell.\get \; s_i$ if and
-only if $\ell.\pput \; s_i \; v_i = s_i$, so $\delta_V^*$ omits
-$\delta_V(v_i, \ell.\get \; s_i)$ in position $i$ if and only if
-$\delta_S^*$ also omits $\delta_S(\ell.\pput \; s_i \; v_i, s_i)$ in
-position $i$; if no omission occurs, then $\delta(S, V).\get \;
-(\delta(\ell.\pput \; s_i \; v_i, s_i)) = \delta(v_i, \ell.\get \; s_i)$ since
-$(\ell, \partial(S,V))$ is a delta lens. For $\min\{m,n\} < i \leq m$, then
-$\delta_V^*$ adds $v_i$ in position $i$, and $\delta_S^*$ adds
-$\ell.\create \; v_i$ in position $i$, which gives the desired result since
-$$(\partial (S, V) \sqcup \ell).\get \; (\ell.\create \; v_i) = \ell.\get \;
-(\ell.\create \; v_i)) = v_i$$
-\end{proof}
-\subsection{Matching Iteration}
-\iffalse
-When defining a $\match(\mathcal{M}, \delta)$ difference function, we often wish for
-$\mathcal{M}$ to be an optimal set of matchings for some alignment strategy.
-
-For instance, define the set of matchings $\mathcal{M}_{POS}$ by
-$$\mathcal{M}_{POS} = \{\varphi_{(a_1, \ldots, a_m), (b_1, \ldots, b_n)}:[m]
-\longrightarrow [n] \; | \; m \leq n, \varphi_{(a_1, \ldots, a_m), (b_1,
-\ldots, b_n)}(i) = i\}$$ Then $\mathcal{M}_{POS}$ is an optimal set of
-matchings for a positional alignment strategy.
-
-For the non-crossing strategy, we want to our matchings to be increasing rather
-than positional. That is, given a valid set of matchings $\mathcal{M}$, let
-$\mathcal{M}_{a, b} = \varphi_{a, b} \in \mathcal{M}$, and for $a, a' \in P$
-with $a=(a_1, \ldots, a_m)$ and $a'=(a'_1, \ldots, a'_n)$, define
-$\mathcal{M}^{NC}_{a, b}$ by
-$$\mathcal{M}^{NC}_{a, a'} = \argmin_{\substack{\mathcal{M}_{a, a'} \\
-\mathcal{M}_{a, a'} \text{ is increasing}}} |\match(\mathcal{M}_{a, a'},
-\delta)(a, a')|$$
-There may be multiple matchings which for which the size of the delta is
-minimal, in which case we pick the lexicographically smallest matching.
-Then $\mathcal{M}^{NC}$ is an optimal set of matchings for a non-crossing
-alignment strategy.
-
-Finally, for a min-edit distance strategy, we simply remove the condition that
-the matchings should be increasing. That is, define
-$\mathcal{M}^{MED}_{a, b}$ by
-$$\mathcal{M}^{MED}_{a, a'} = \argmin_{\mathcal{M}_{a, a'}}
-|\match(\mathcal{M}_{a, a'}, \delta)(a, a')|$$
-Then $\mathcal{M}_{MED}$ is an optimal set of matchings for a min-edit distance
-alignment strategy.
-\fi
-Assume that $\ell: S \Leftrightarrow V$. Let $\mathcal{M}_V = \{\varphi_{x, y}
-\; | \; x, y \in V^*\}$ be a set of alignments between $x$ and $y$ for each $x,
-y \in V^*$. For each $s, s' \in S^*$, let $\varphi_{s, s'} = \varphi_{\ell.\get
-\; s, \ell.\get \; s'}$. Define $\mathcal{M}_S$ by $\mathcal{M}_S =
-\{\varphi_{s, s'} \; | \; s, s' \in S^*\}$. Define the lens
-$\match(\mathcal{M}_V, \ell) : S^* \Leftrightarrow V^*$ by
-\begin{align*}
-\match(\mathcal{M}_V, \ell) .\get \; (s_1, \ldots, s_n) &= (\ell.\get \; s_1)
-\cdot \ldots \cdot (\ell.\get \; s_n)\\
-\match(\mathcal{M}_V, \ell) .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n)
-&= (s'_1 , \ldots , s'_n)\\
-\text{ where } s'_i &= \begin{cases}
-\ell.\pput \; s_{\psi(i)} \; v_i & \text{if } i \in Dom(\psi)\\
-\ell.\create \; v_i & \text{if } i \not \in Dom(\psi)
-\end{cases}\\
-\text{ and } \psi &= \varphi_{(v_1, \ldots, v_n), ((\ell.\get \; s_1), \ldots,
-(\ell.\get \; s_m))}\\
-\match(\mathcal{M}_V, \ell).\create \; (v_1, \ldots, v_n) &= (\ell.\create \;
-v_1) \cdot \ldots \cdot (\ell.\create \; v_n)
-\end{align*}
-Assume that 
+$$\delta_{S^*_{MED}}(\ell^*_{MED}.\pput \; (s_1 \cdot \ldots \cdot s_m) \; (v_1 \cdot \ldots \cdot v_n), s_1 \cdot \ldots \cdot s_m) \geq \delta_{V^*_{MED}}(v_1 \cdot \ldots \cdot v_n, (\ell.\get \; s_1) \cdot \ldots (\ell.\get \; s_m))$$
+Assume wthout loss of generality that $n \leq m$.
+An optimal min-edit distance alignment $\varphi : [n] \longrightarrow [m]$ is used to compute the difference on the view side.
+Since $\ell$ is a difference lens, the following equation holds,
+$$\delta_S(\ell.\pput \; s_{\varphi(i)} \; v_i, s_{\varphi(i)}) \geq \delta_V(v_i, \ell.\get \; s_{\varphi(i)})$$
+from which it follows that
+$$\delta_{S^*_{MED}}(v_1 \cdot \ldots v_n \xhookrightarrow{\ell, \varphi} s_1 \cdot \ldots \cdot s_m, s_1 \cdot \ldots \cdot s_m) \geq \delta_{V^*_{NC}}(v_1 \cdot \ldots \cdot v_n, (\ell.\get \; s_1) \cdot \ldots (\ell.\get \; s_m))$$
+Since the \pput function of $\ell^*_{MED}$ uses an alignment $\psi$ that maximizes the min-edit distance over all injective alignments mapping $v$ into $s$, then
+$$\delta_{S^*_{MED}}(v_1 \cdot \ldots v_n \xhookrightarrow{\ell, \psi} s_1 \cdot \ldots \cdot s_m, s_1 \cdot \ldots \cdot s_m) \geq \delta_{S^*_{MED}}(v_1 \cdot \ldots v_n \xhookrightarrow{\varphi} s_1 \cdot \ldots \cdot s_m, s_1 \cdot \ldots \cdot s_m)$$
+
+It thus follows that,
+$$\delta_{S^*_{MED}}(v_1 \cdot \ldots v_n \xhookrightarrow{\ell, \psi} s_1 \cdot \ldots \cdot s_m, s_1 \cdot \ldots \cdot s_m) \geq \delta_{V^*_{NC}}(v_1 \cdot \ldots \cdot v_n, (\ell.\get \; s_1) \cdot \ldots (\ell.\get \; s_m))$$
+hence,
+$$\delta_{S^*_{MED}}(\ell^*_{MED}.\pput \; (s_1 \cdot \ldots \cdot s_m) \; (v_1 \cdot \ldots \cdot v_n), s_1 \cdot \ldots \cdot s_m) \geq \delta_{V^*_{MED}}(v_1 \cdot \ldots \cdot v_n, (\ell.\get \; s_1) \cdot \ldots (\ell.\get \; s_m))$$
+which is what we wanted to show.
+\end{proof}
+\subsubsection{Subtyping}
+Consider a positional iteration lens $\ell^*_{POS} : S^*_{POS} \Leftrightarrow V^*_{POS}$.
+Since $\delta_{S^*_{NC}}(s, s') \geq \delta_{S^*_{POS}}(s, s')$ for all $s, s' \in \mathcal{L}(R)$, it follows $\ell^*_{POS} : S^*_{NC} \Leftrightarrow V^*_{POS}$ is also a valid typing judgement.
+This observation suggests a {\em subtyping} relation on unambiguous difference regular expressions.
+
+We define the subtying relation $\leq$ on difference regular expressions as follows:
+\begin{enumerate}
+\item
+If $R$ is unambiguous, then $R \leq R$.
+\item
+If $R^{!*}$, then $R^*_{POS} \leq R^*_{NC}$, $R^*_{POS} \leq R^*_{MED}$ and $R^*_{POS} \leq R^*_{NC}$.
+\item $R \leq S$ and $S \leq T$ then $R \leq T$.
+\item If $R_1 \leq S_1$ and $R_2 \leq S_2$ with $R_1 \cdot ^! R_2$ and $S_1 \cdot^! S_2$ then $R_1 \cdot R_2 \leq S_1 \cdot S_2$.
+\item If $R_1 \leq S_1$ and $R_2 \leq S_2$ with $\mathcal{L}(R_1) \cap \mathcal{L}(R_2) = \varnothing$ and $\mathcal{L}(S_1) \cap \mathcal{L}(S_2) = \varnothing$ then $R_1 \cup R_2 \leq S_1 \cup S_2$.
+\item
+If $R \leq S$ with $R^{!*}$ and $S^{!*}$ then $R^*_{POS} \leq S^*_{POS}$, $R^*_{NC} \leq S^*_{NC}$ and $R^*_{MED} \leq S^*_{MED}$.
+\end{enumerate}
 \begin{claim}
-If $\ell: S \xLeftrightarrow{\delta_{S}, \partial (S, V), \delta_{V}}
-V$, then $\match(\mathcal{M}_V, \ell) : S^*
-\xLeftrightarrow{\match(\mathcal{M}_S, \delta_{S}), \partial (S, V)^*,
-\match(\mathcal{M}_V, \delta_V)} V^*$.
-\end{claim}
-\begin{proof}
-Observe that the matching used to compute the delta on the source side is
-the same as the matching used to compute the matching on the view side, hence we
-may take the update lens to be position and proceed with the rest of the proof
-as in \cref{positionalisvalid}.
-\end{proof}
-\subsection{Composition}
-Let $\ell_1 : S \Leftrightarrow U$ and $\ell_2 : U \Leftrightarrow V$ be lenses.
-Define the lens $\ell_2 \circ \ell_1 : S \Leftrightarrow V$ by
-\begin{align*}
-(\ell_2 \circ \ell_1).\get \; s &= \ell_2.\get \; (\ell_1.\get \; s)\\
-(\ell_2 \circ \ell_1).\pput \; s \; v &= \ell_1.\pput \; s \; (\ell_2.\pput \;
+\begin{enumerate}
+\item
+If $R \leq S$ then $\mathcal{L}(R) = \mathcal{L}(S)$.
+\item
+If $R \leq S$, then $\delta_R(s, s') \geq \delta_S(s, s')$ for all $s, s' \in \mathcal{L}(R) = \mathcal{L}(S)$.
+\end{enumerate}
+\begin{corollary}
+If $\ell : S \Leftrightarrow V$ is a difference lens with $S' \leq S$ and $V \leq V'$, then $\ell : S' \Leftrightarrow V'$ is also a difference lens.
+\end{corollary}
+\end{claim}
+\subsubsection{Composition}
+Let $\ell_1 : S \Leftrightarrow U_1$ and $\ell_2 : U_2 \Leftrightarrow V$ be difference lenses with $U_1 \leq U_2$. Define the lens $\ell_1 \; ; \; \ell_2 : S \Leftrightarrow V$ by
+\begin{align*}
+(\ell_1 \; ; \; \ell_2).\get \; s &= \ell_2.\get \; (\ell_1.\get \; s)\\
+(\ell_1 \; ; \; \ell_2).\pput \; s \; v &= \ell_1.\pput \; s \; (\ell_2.\pput \;
 (\ell_1.\get \; s) \; v) \\
-(\ell_2 \circ \ell_1).\create \; v &= \ell_1.\create \; (\ell_2.\create \; v)
+(\ell_1 \; ; \; \ell_2).\create \; v &= \ell_1.\create \; (\ell_2.\create \; v)
 \end{align*}
@@
 \begin{claim}
-If $\ell_1 : S_1  \xLeftrightarrow{\delta_{S_1}, \partial (S_1, V_1),
-\delta_{V_1}} V_1$ and $\ell_2 : S_2 \xLeftrightarrow{\delta_{S_2}, \partial
-(S_2, V_2), \delta_{V_2}} V_2$, then \\
-$\ell_2 \circ \ell_1 :  S \xLeftrightarrow{\delta_S, \partial
-(S_2, V_2) \circ \partial (S_1, V_1), \delta_V} V$.
+If $\ell_1 : S  \Leftrightarrow U_1$ and $\ell_2 : U_2 \Leftrightarrow V$ are difference lenses with $U_1 \leq U_2$ then
+$\ell_1 \; ; \; \ell_2 :  S \Leftrightarrow V$ is also a difference lens.
 \end{claim}
@@
 $$
-(\partial (S_2, V_2) \circ \partial (S_1, V_1)).\get \; (\delta(\ell_2 \circ
-\ell_1).\pput \; s \; v, s) = \delta(v, (\ell_2 \circ \ell_1).\get \; s)$$
+\delta_S(\ell_1 \; ; \; \ell_2).\pput \; s \; v, s) \geq \delta_V(v, (\ell_1 \; ; \; \ell_2).\get \; s)$$
 
@@
 \begin{align*}
-&(\partial (S_2, V_2) \circ \partial (S_1, V_1)).\get \; (\delta_S((\ell_2 \circ
-\ell_1).\pput \; s \; v, s))\\
-&=(\partial (S_2, V_2) \circ \partial (S_1, V_1)).\get \; (\delta_S(\ell_1.\pput
-\; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; v), s))\\
-&= (\partial (S_2, V_2)).\get \; ((\partial (S_1, V_1)).\get \;
-(\delta_S(\ell_1.\pput \; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; v), s)))\\
-&= (\partial (S_2, V_2)).\get \; (\delta_U(\ell_2.\pput \; (\ell_1.\get \; s)
-\; v, \ell_1.\get \; s))\\
-&= \delta_V(v, \ell_2.\get \; (\ell_1.\get \; s))\\
-&= \delta_V(v, (\ell_2 \circ \ell_1).\get \; s)
-\end{align*}
+\delta_S(\ell_1 \; ; \; \ell_2).\pput \; s \; v, s) &=
+\delta_S(\ell_1.\pput \; s \; (\ell_2.\pput \;
+(\ell_1.\get \; s) \; v), s)\\
+&\geq \delta_{U_1}(\ell_2.\pput \;
+(\ell_1.\get \; s) \; v, \ell_1.\get \; s)\\
+&\geq \delta_{U_2}(\ell_2.\pput \;
+(\ell_1.\get \; s) \; v, \ell_1.\get \; s)\\
+&\geq \delta_V(v, \ell_2.\get \; (\ell_1.\get \; s))\\
+&= \delta_V(v, (\ell_1 \; ; \; \ell_2).\get \; s)
+\end{align*}
+\end{proof}
+\subsubsection{Filter}
+Let $R$ and $S$ be difference regular expressions.
+Assume that $\mathcal{L}(R) \cap \mathcal{L}(S) = \varnothing$, and that $(R \cup S)^{!*}$.
+Define the lens $\filter_{MED} \; R \; S : (R \sqcup S)^*_{MED} \Leftrightarrow R^*_{MED}$ by
+\begin{align*}
+(\filter \; R \; S).\get \; s &= \strfilter \; R \; s\\
+(\filter \; R \; S).\pput \; s \; v &= \strunfilter \; S \; s \;v\\
+(\filter \; R \; S).\create \; v &= v
+\end{align*}
+where \strfilter and \strunfilter are defined as follows:
+\begin{lstlisting}[mathescape=true]
+let rec $\strfilter \; R \; xs$ = 
+match $xs$ with
+$|$ $\varepsilon$ -> $\varepsilon$
+$|$ $x \cdot xs'$ when $x \in R$ -> $x \cdot (\strfilter \; R \; xs')$
+$|$ $x \cdot xs'$ when $x \not \in R$ -> $\strfilter \; R \; xs'$
+
+let rec $\strunfilter \; S \; xs \; ys$ = 
+match $xs, ys$ with
+$|$ $\varepsilon$,_ -> $ys$
+$|$ $x \cdot xs'$,_ when $x \in S$ -> $x \cdot (\strunfilter \; S \; xs' \; ys)$
+$|$ $x \cdot xs',\varepsilon$ when $x \not \in S$ -> $\strunfilter \; R \; xs' \; \varepsilon$
+$|$ $x \cdot xs',y \cdot ys'$ when $x \not \in S$ -> $y \cdot (\strunfilter \; R \; xs' \; ys')$
+\end{lstlisting}
+Given a source $s \in (R \cup S)^*$, the lens $\filter \; R \; S$ removes all elements that match $S$ in $s$ in the \get direction and restores them in the \pput direction.
+\begin{claim}
+Let $R$ and $S$ be difference regular expressions with $\mathcal{L}(R) \cap \mathcal{L}(S) = \varnothing$ and $(R \cup S)^{!*}$.
+Then $\filter_{POS} \; R \; S : (R \sqcup S)^*_{MED} \Leftrightarrow R^*_{MED}$ is a difference lens.
+\end{claim}
+\begin{proof}
+We need to show that
+$$
+\delta_{(R \cup S)^*_{MED}} (\strunfilter \; S \; s \; v, s) \geq \delta_{R^*_{MED}}(v, \strfilter \; R \; s)$$
+Observe that if $s = s_1 \cdot \ldots cdot s_m$ with $s_{i_1}, \ldots, s_{i_p} \in R$ and $s_{j_1}, \ldots, s_{j_p} \in S$, then $\strfilter \; R \; s = s_{i_1} \cdot \ldots \cdot s_{i_n}$.
+Assume that $v = v_1 \cdot \ldots v_{n'}$ and assume without loss of generality that $n \leq n'$.
+Then since $\delta_{R^*_{MED}}(v, \strfilter \; R \; s) \leq 1$, it follows that
+$$
+\delta_{(R \cup S)^*_{MED}} (\strunfilter \; S \; s \; v, s) = \frac{m + \delta_{R^*_{MED}}(v, \strfilter \; R \; s) * n'}{m+n'} \geq \delta_{R^*_{MED}}(v, \strfilter \; R \; s)
+$$
+which is what we wanted to show.
+\end{proof}
+\subsubsection{Merge}
+If $R$ is a difference regular expression with $R \cdot^! R$, then define the \mergelens lens $\mergelens \; R : R \cdot R \Leftrightarrow R$ as follows:
+\begin{align*}
+(\mergelens \; R).\get \; (s \cdot s') &= s\\
+(\mergelens \; R).\pput \; (s \cdot s') \; t &= 
+\begin{cases}
+t \cdot t & \text{if }s = s'\\
+t \cdot s' & \text{otherwise}
+\end{cases}\\
+(\mergelens \; R).\create s &= s \cdot s
+\end{align*}
+\begin{claim}
+If $R$ is a difference regular expression with $R \cdot^! R$, then $\mergelens \; R : R \cdot R \Leftrightarrow R$ is a difference lens.
+\end{claim}
+\begin{proof}
+We need to show that
+$$\delta_{R \cdot R}((\mergelens \; R).\pput \; (s \cdot s') \; t, (s \cdot s')) \geq \delta_R(t, s)$$
+If $s = s'$, then we need to show that
+$$\delta_{R \cdot R}(t \cdot t, s \cdot s) \geq \delta_R(t, s)$$
+which follows from the fact that
+If $s \neq s'$, then we need to show that
+$$\delta_{R \cdot R}(t \cdot t, s \cdot s) = \frac{1}{2}\left(\delta_R(t, s) + \delta_R(t, s)\right) = \delta_R(t, s)$$
+$$\delta_{R \cdot R}(t \cdot s', s \cdot s') = \frac{1}{2}\left(\delta_R(t, s) + \delta_R(s', s')\right) = \frac{1}{2}\left(\delta_R(t, s) + 1\right) \geq \delta_R(t, s)$$
+which follows from the fact that $\delta_R(t, s) \leq 1$.
 \end{proof}
 \iffalse
-\section{Symmetric Delta Lenses}
-Assume that $(X, \partial X, \delta_X)$ and $(Y, \partial Y, \delta_Y)$
-are difference structures. A symmetric delta lens is a pair $(\ell, \partial
-\ell)$ where $\ell : X \Leftrightarrow Y$ is a symmetric lens, and
-$\partial \ell : \partial X \Leftrightarrow \partial Y$ is a symmetric lens
-such that
-\begin{enumerate}
-  \item
-  if $\ell.putr(x, c) = (y, c)$ and $\ell.putl(y', c) = (x', c')$,
-  then $\partial \ell.putl \; (\delta(y, y'), \delta(c, c')) = (\delta(x,
-  x'), \delta(c, c'))$
-  \item
-  if $\ell.putl(y, c) = (x, c)$ and $\ell.putr(x', c) = (y', c')$, then
-  $\partial \ell.putr \; (\delta(x, x'), \delta(c, c')) = (\delta(y, y'),
-  \delta(c, c'))$
-\end{enumerate}
-\begin{fact}[Identity Lens]
-Given any symmetric lens $\partial X : X \Leftrightarrow X$, $(id_X, \partial
-X)$ is a symmetric delta lens
-\end{fact}
-\begin{fact}[Bijective Lenses]
-Assume that $f : X \longrightarrow Y$ and $g : \partial X \longrightarrow
-\partial Y$ are bijections. Assume that $g(\delta(x, x')) = \delta(f(x),
-f(x'))$. Then $(bij_f, bij_g)$ is a symmetric delta lens.
-\end{fact}
-\begin{fact}[Lens Composition]
-If $(\ell_1, \partial \ell_1) : (X, \partial X, \delta_X) \Leftrightarrow (Y,
-\partial Y, \delta_Y)$ and $(\ell_2, \partial \ell_2): (Y, \partial Y,
-\delta_Y) \Leftrightarrow (Z, \partial Z, \delta_Z)$ are symmetric delta
-lenses, then $(\ell_1;\ell_2, \partial \ell_1;\partial \ell_2)$ is a symmetric
-delta lens.
-\end{fact}
-\begin{fact}[Dual of a Lens]
-If $(\ell, \partial \ell)$ is symmetric delta lens, then $(\ell^{op}, \partial
-\ell^{op})$ is a symmetric delta lens.
-\end{fact}
-\begin{fact}[Terminal Lens]
-$(term_x, term_{\delta(x, x)})$ is a symmetric delta lens
-\end{fact}
-\begin{corollary}[Disconnect Lens]
-If $x \in X$ and $y \in Y$, then $(term_x;term_y^{op},
-term_{\delta(x, x)};term_{\delta(y, y)}^{op})$ is a symmetric delta lens.
-\end{corollary}
-\begin{fact}[Tensor Product Lens]
-If $(\ell_1, \partial \ell_1)$ and $(\ell_2, \partial \ell_2)$ are symmetric
-delta lenses, then $(\ell_1 \otimes \ell_2, \partial \ell_1 \otimes \partial
-\ell_2)$ is a symmetric delta lens.
-\end{fact}
-\begin{corollary}[Projection Lens]
-Projection lenses are symmetric delta lenses.
-\end{corollary}
-\begin{fact}[Forgetful Tensor Sum Lens]
-If $(\ell_1, \partial \ell_1)$ and $(\ell_2, \partial \ell_2)$ are symmetric
-delta lenses, then $(\ell_1 \oplus^f \ell_2, \partial \ell_1 \oplus^f \partial
-\ell_2 \oplus (\ell_1 \otimes \ell_2))$ is a symmetric delta lens.
-\end{fact}
-\begin{fact}[Forgetful List Mapping Lens]
-If $(\ell, \partial \ell)$ is a symmetric delta lens, then $(map^f(\ell)
-\ell_2, map^f(\partial \ell))$ is a symmetric delta lens.
-\end{fact}
-\begin{remark}[Retentive Lenses]
-The retentive tensor sum and list mapping lenses are in general not symmetric
-delta lenses with the obvious definitions of difference functions.
-\end{remark}
-\fi
+\section{Matching Lenses}
+\begin{definition}
+A {\em basic lens} $\ell : S \xLeftrightarrow{C} V$ is a 4-tuple of functions
+$\ell.\get : S \longrightarrow V$, $\ell.\res : S \longrightarrow C$,
+$\ell.\pput : V \longrightarrow C \longrightarrow S$ and $\ell.\create : V
+\longrightarrow S$ satisfying the following properties:
+\begin{align*}
+\ell.\get \; (\ell.\pput \; v \; c) &= v \tag{PUTGET}\\
+\ell.\pput \; (\ell.\get \; s) \; (\ell.\res \; s) &= s \tag{GETPUT}
+\end{align*}
+\end{definition}
+Idea: First remove matched parts to leave unmatched skeleton. Next remove
+unmatched parts to leave matched skeleton and then flatten.
 \section{Relational Lenses}
@@
 \end{enumerate}
-\section{Matching Lenses}
-\begin{definition}
-A {\em basic lens} $\ell : S \xLeftrightarrow{C} V$ is a 4-tuple of functions
-$\ell.\get : S \longrightarrow V$, $\ell.\res : S \longrightarrow C$,
-$\ell.\pput : V \longrightarrow C \longrightarrow S$ and $\ell.\create : V
-\longrightarrow S$ satisfying the following properties:
-\begin{align*}
-\ell.\get \; (\ell.\pput \; v \; c) &= v \tag{PUTGET}\\
-\ell.\pput \; (\ell.\get \; s) \; (\ell.\res \; s) &= s \tag{GETPUT}
-\end{align*}
-\end{definition}
-Idea: First remove matched parts to leave unmatched skeleton. Next remove
-unmatched parts to leave matched skeleton and then flatten.
+\fi
 %% Acknowledgments
@@
 \end{acks}
-\fi
+
 \bibliographystyle{plain}
@@
 \end{document}
-\newif\ifdraft\drafttrue  % set true to show comments
-\newif\ifplentyoftime\plentyoftimefalse  % :-)
-
-% % For double-blind review submission, w/o CCS and ACM Reference (max
-% submission space)
-\documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-% % For double-blind review submission, w/ CCS and ACM Reference
-% \documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-% % For single-blind review submission, w/o CCS and ACM Reference (max
-% submission space)
-% \documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-% % For single-blind review submission, w/ CCS and ACM Reference
-% \documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true} % For
-% final camera-ready submission, w/ required CCS and ACM Reference
-% \documentclass[acmsmall]{acmart}\settopmatter{}
-
-% % Journal information % Supplied to authors by publisher for camera-ready
-% submission; % use defaults for review submission.
-\acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{POPL} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2018}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-% % Copyright information % Supplied to authors (based on authors' rights
-% management selection; % see authors.acm.org) by publisher for camera-ready
-% submission; % use 'none' for review submission.
-\setcopyright{none}
-% \setcopyright{acmcopyright} \setcopyright{acmlicensed}
-% \setcopyright{rightsretained} \ccopyrightyear{2018}           %% If different
-% from \acmYear
-
-% % Bibliography style %\bibliographystyle{ACM-Reference-Format} % Citation
-% style % Note: author/year citations are required for papers published as an %
-% issue of PACMPL.
-% %\citestyle{acmauthoryear}   %% For author/year citations
-
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%% % Note:
-% Authors migrating a paper from PACMPL format to traditional % SIGPLAN
-% proceedings format must update the '\documentclass' and % topmatter commands
-% above; see 'acmart-sigplanproc-template.tex'.
-% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\usepackage{amsmath, amssymb, amsthm, enumerate, array, extarrows, mathrsfs,
-mathtools, stmaryrd, listings, xspace, bussproofs}
-\theoremstyle{definition}
-\usepackage[capitalize]{cleveref}
-\newtheorem{definition}{Definition}
-\newtheorem{example}{Example}
-\newtheorem{conjecture}{Conjecture}
-\newtheorem{question}{Question}
-\newtheorem{fact}{Fact}
-\newtheorem{claim}{Claim}
-\newtheorem{remark}{Remark}
-
-% Macros Colors
-\definecolor{dkblue}{rgb}{0,0.1,0.5}
-\definecolor{dkgreen}{rgb}{0,0.6,0}
-\definecolor{dkred}{rgb}{0.6,0,0}
-\definecolor{dkpurple}{rgb}{0.7,0,0.4}
-\definecolor{olive}{rgb}{0.4, 0.4, 0.0}
-\definecolor{teal}{rgb}{0.0,0.5,0.5}
-\definecolor{orange}{rgb}{0.9,0.6,0.2}
-\definecolor{lightyellow}{RGB}{255, 255, 179}
-\definecolor{lightgreen}{RGB}{170, 255, 220}
-\definecolor{teal}{RGB}{141,211,199}
-\definecolor{darkbrown}{RGB}{121,37,0}
-
-\newcommand{\FINISH}[3]{\ifdraft\textcolor{#1}{[#2: #3]}\fi}
-\newcommand{\bcp}[1]{\FINISH{dkred}{B}{#1}}
-\newcommand{\BCP}[1]{\FINISH{dkred}{B}{\bf #1}}
-\newcommand{\afm}[1]{\FINISH{dkgreen}{A}{#1}}
-\newcommand{\dpw}[1]{\FINISH{dkblue}{D}{#1}} % Toronto Maple Leafs Blue :-)
-\newcommand{\saz}[1]{\FINISH{orange}{SZ}{#1}}
-\newcommand{\ksf}[1]{\FINISH{teal}{K}{#1}}
-\newcommand{\sam}[1]{\FINISH{dkpurple}{SM}{#1}}
-
-\newcommand{\kw}[1]{\ensuremath{\mathsf{#1}}\xspace}
-\newcommand{\get}{\ensuremath{\kw{get}}\xspace}
-\newcommand{\pput}{\ensuremath{\kw{put}}\xspace}
-\newcommand{\create}{\ensuremath{\kw{create}}\xspace}
-\newcommand{\res}{\ensuremath{\kw{res}}\xspace}
-\newcommand{\ccopy}{\ensuremath{\kw{copy}}\xspace}
-\newcommand{\const}{\ensuremath{\kw{const}}\xspace}
-\newcommand{\default}{\ensuremath{\kw{default}}\xspace}
-\newcommand{\length}{\ensuremath{\kw{length}}\xspace}
-\newcommand{\match}{\ensuremath{\kw{match}}\xspace}
-
-\DeclareMathOperator*{\argmin}{arg\,min} % thin space, limits underneath in
-\DeclareMathOperator*{\argmax}{argmax} % no space, limits underneath in
-
-\lstset{ basicstyle=\ttfamily, escapeinside=|| }
-\begin{document}
-
-% % Title information
-\title{Lenses with a Difference}         %% [Short Title] is optional;
-% % when present, will be used in % header instead of Full Title.
-% %\titlenote{with title note}             %% \titlenote is optional; % can be
-% repeated if necessary; % contents suppressed with 'anonymous'
-% %\subtitle{Subtitle}                     %% \subtitle is optional
-% %\subtitlenote{with subtitle note}       %% \subtitlenote is optional; % can
-% be repeated if necessary; % contents suppressed with 'anonymous'
-
-% % Author information % Contents and number of authors suppressed with
-% 'anonymous'.
-% % Each author should be introduced by \author, followed by % \authornote
-% (optional), \orcid (optional), \affiliation, and % \email.
-% % An author may have multiple affiliations and/or emails; repeat the %
-% appropriate command.
-% % Many elements are not rendered, but should be provided for metadata %
-% extraction tools.
-
-% % Author with single affiliation.
-% \author{First1 Last1} \authornote{with author1 note}          %% \authornote
-% is optional; % can be repeated if necessary \orcid{nnnn-nnnn-nnnn-nnnn} %%
-% \orcid is optional \affiliation{ \positition{Position1}
-% \department{Department1} %% \department is recommended
-% \institution{Institution1} %% \institution is required \streetaddress{Street1
-% Address1} \city{City1} \state{State1} \posittcode{Post-Code1}
-% \country{Country1}                    %% \country is recommended }
-% \email{first1.last1@inst1.edu}          %% \email is recommended
-
-\author{Solomon Maina}
-\position{PhD Student}
-\department{Computer Science}              %% \department is recommended
-\institution{University of Pennsylvania}            %% \institution is required
-\country{USA}                    %% \country is recommended }
-\email{smaina@seas.upenn.edu}          %% \email is recommended
-
-% % Abstract % Note: \begin{abstract}...\end{abstract} environment must come %
-% before \maketitle command
-\begin{abstract}
-Many of the bidirectional (bx) programming languages that exist today focus on
-restoring consistency of models. Consequently, these languages often provide two
-guarantees: (1) that a transformation restores consistency when an
-update is made to one of the models, and (2) that models that are already
-consistent are not updated by the transformation. Unfortunately, in situations
-where there are many different ways of restoring consistency, these languages
-do not constrain how a lens chooses to do so. As a result, the bx
-transformations derived using these languages can lead to unexpected results. In
-this paper we formalize a condition called {\em $\partial D$ preservation} that
-describes whether a bx transformation propagates updates correctly with respect
-to the relation $\partial D$, and then use this condition to characterize the
-guarantees that various lens programming languages make with regard to update
-propagation.
-\end{abstract}
-
-%% Keywords
-%% comma separated list
-\keywords{bidirectional programming, lenses}  %% \keywords
-% are mandatory in final camera-ready submission
-
-%% \maketitle
-%% Note: \maketitle command must come after title commands, author
-%% commands, abstract environment, Computing Classification System
-%% environment and commands, and keywords command.
-\maketitle
-
-\section{Introduction}
-By now there exist many programming languages that enable one to define {\em
-bidirectional transformations}, programs that can be run in two
-directions. Each of these languages provides guarantees on how the two
-transformations interact e.g. that they are bijective, that a round-trip on
-one or both directions does not cause non-trivial updates etc. Many of these
-guarantees have to do with restoring or preserving consistency. To formalize
-this statement, we describe a widely used formulation of bidirectional
-transformations due to Stevens \cite{stevens2010bidirectional} but originally
-articulated in a more general form by Meertens \cite{meertens1998designing}:
-\begin{definition}
-A bidirectional transformation (bx) $R: M \Leftrightarrow N$ comprises of a
-consistency relation $R \subseteq M \times N$ and restorers $\overrightarrow{R}
-: M \times N \rightarrow N$ and $\overleftarrow{R} : M \times N \longrightarrow
-N$.
-\end{definition}
-\begin{definition}
-A bx $R : M \Leftrightarrow N$ is {\em correct} if the consistency restorers do
-restore consistency: that is, for all $m \in M, n \in N$, we have $R(m,
-\overrightarrow{R}(m, n))$, and dually. $R$ is {\em hippocratic} if the
-consistency restorers make no change to an already-consistent pair of models:
-that is, for all $m \in M$, $n \in N$, if $R(m, n)$ then $\overrightarrow{R}(m,
-n) = n$ and dually. $R$ is {\em well behaved} if it is correct and hippocratic.
-\end{definition}
-
-Many bidirectional programming languages thus enable one to define
-well behaved bx's: consistency restoration is guaranteed by correctness of the
-bx while consistency preservation is guaranteed by the hippocraticness of the
-bx. This paper is primarily concerned with well behaved bx's where
-$\overrightarrow{R}$, the \get function of the bx, ignores it's $N$-argument and
-where the consistency relation $R$ is defined by $R(m, \overrightarrow{R}(m))$
-for all $m \in M$--- these transformations are the well known and widely
-understood {\em lenses} of Foster et al \cite{foster2007combinators}--- though
-the ideas we develop could be used in other bidirectional settings as well.
-
-When the consistency relation of a bx $R$ is very fine, i.e. when each $m \in M$
-is consistent with only a few $n \in N$ and vice versa, then a well behaved lens
-gives fairly strong guarantees on $\overleftarrow{R}$ and $\overrightarrow{R}$.
-However, there are many cases that arise in practice where there are many
-different ways to restore consistency; when this happens, well-behavedness may
-be too weak to rule out ``bad'' behaviour or to clarify the behaviour of a bx in
-subtle cases. Various solutions have been proposed to address this issue, for
-example {\em matching lenses} that deal with the issue of alignment
-\cite{barbosa2010matching}, {\em edit lenses} \cite{hofmann2012edit} and {\em
-delta lenses} \cite{diskin2011asymmetric,diskin2011state,pacheco2012delta}
-that explicitly propagate edits through the lens, and {\em least change
-lenses} that require that the source value resulting from a \pput application
-is not only acceptable, but also one of closest to the original $s$ among the
-sources that share the same view $v$ \cite{macedo2013composing}.
-
-This paper takes a different approach to specifying update propagation by
-identifying a condition called {\em $\partial D$ preservation} which we describe
-below:
-\begin{definition}
-Let $R : M \Leftrightarrow N$ be a bidirectional transformation and $\partial D
-\subseteq (M \times M) \times (N \times N)$. Then $R$ {\em preserves $\partial
-D$} if for all $m \in M$, $n, n' \in N$ , if $R(m, n)$ then $\partial D((m,
-\overleftarrow{R}(m, n')), (\overrightarrow{R}(m, n), n'))$. $R$ {\em
-strongly preserves $\partial D$} if the dual condition also holds.
-\end{definition}
-\noindent The idea behind $\partial D$ preservation is that if $m$ and $n$ are
-already consistent and $\overrightarrow{R}(m,n)$ is updated to produce $n'$,
-then the difference between $\overrightarrow{R}(m,n)$ and $n'$ is related
-to the difference between $m$ and $\overleftarrow{R}(m,n')$. $R$
-strongly preserves $\partial D$ if it preserves $\partial D$ in both directions.
-
-For lenses, $\partial D$ preservation says that difference between $\ell.\get \;
-s$ and $v$ is related to the difference between $s$ and $\ell.\pput \; s \;
-v$. One way to formalize this idea is to model the difference between two sources $s, s'
-\in S$ and two views $v, v' \in V$ as functions $\delta_S : S \times S
-\longrightarrow \partial S$ and $\delta_V : V \times V \longrightarrow \partial
-V$ for some sets $\partial S$ and $\partial V$, and to capture the
-relationship between $\delta_S$ and $\delta_V$ by a lens. In
-particular, given a lens $\partial(S, V) : \partial S \Leftrightarrow \partial
-V$ we will ask if the following equation holds of $\ell$:
-\begin{equation}\label{difflenslaw}
-\partial(S, V).\get \; (\delta_S(s, \ell.\pput \; s \; v)) = \delta_V(\ell.\get
-\; s, v)
-\end{equation}
-If $\ell : S \Leftrightarrow V$ satisfies \cref{difflenslaw} we say that 
-$\ell$ is a {\em difference  lens} with respect to $(\delta_S, 
-\partial(S, V), \delta_V)$, which we denote with the typing relation 
-$\ell : S \xLeftrightarrow{\delta_S, \partial(S, V), \delta_V} V$.
-
-For example, consider two versions of an iteration lens $\ell_1 : S^*
-\Leftrightarrow V^*$ and $\ell_2 : S^* \Leftrightarrow V^*$ defined inductively
-from a lens $\ell : S \Leftrightarrow V$. Both of these lens have the same get
-behaviour: given a list $s_1, \ldots, s_n$ of sources, they both produce
-$\ell.\get \; s_1, \ldots, \ell.\get \; s_n$. The \pput behaviour of the two
-lenses is different though: $\ell_1$ simply aligns views to sources
-positionally, while $\ell_2$ computes the optimal alignment on the view side and
-uses this alignment to put views back into sources.
-
-Let \lstinline|name = [A-Z] . [a-z]*|, let the source $S$ equal 
-\lstinline|name . " " . name| and the view $V$ equal \lstinline|name|.
-Let $\ell : S \Leftrightarrow V$ be the lens
-\lstinline|del name . del " " . copy name| which given a pair of names produces
-the second name in the \get direction, and restores the first name in the \pput
-direction:
-$$\ell.\get \; ``Ada \; Turing" = ``Turing" \text{ and }\; \ell.\pput \; ``Ada
-\; Turing" \; ``Lovelace" = ``Ada \; Lovelace"$$
-Define the
-function $\delta_V : V \times V \longrightarrow \mathbb{N}^{\geq 0}$ by 
-$$\delta_V(n, n') = 
-\begin{cases}
-0 & \text{if }n=n'\\
-1 & \text{otherwise}
-\end{cases}$$
-Define $\delta_S : S \times S \longrightarrow \mathbb{N}^{\geq 0}$ by
-$$\delta_S(m \cdot `` \quad " \cdot n, m' \cdot `` \quad " \cdot n') =
-\begin{cases}
-0 & \text{if } m = m', n = n'\\
-1 & \text{if } m = m', n \neq n'\\
-1 & \text{if } m \neq m', n = n'\\
-2 & \text{if } m \neq m', n \neq n'
-\end{cases}$$
-Let the lens $\partial(S, V) : \mathbb{N}^{\geq 0} \Leftrightarrow
-\mathbb{N}^{\geq 0}$ be given by $\partial(S, V) = \ccopy(\mathbb{N}^{\geq
-0})$. Then $\ell : S \xLeftrightarrow{\delta_S, \partial(S, V), \delta_V} V$.
-
-Now let $xs, ys$ be two lists of single names, with $length(xs) \geq length(ys)$
-Define $\delta_V^* : V^* \times V^* \longrightarrow \mathbb{Z}^* \times V^*$ on $xs$ and
-$ys$ by aligning elements of $xs$ maximally with elements of $ys$ and then using $\delta_V$
-on these pairs to produce a list $ds$ of numbers. In this process, some elements of
-$ys$ will not be aligned if $length(xs) < length(ys)$, in which case we simply collect
-these in another list of elements of $V$. For instance,
-$$\delta_V^*([``Lovelace", ``Kleene", ``Turing"], [``Turing", ``Lovelace"]) = ([0,0], [``Kleene"])$$
-since we align \lstinline|"Lovelace"| to \lstinline|"Lovelace"| and 
-\lstinline|"Turing"| to \lstinline|"Turing"| for a difference of 0 in each case, which leaves
-\lstinline|"Kleene"| unaligned.
-
-On the source side, let $ss, ts$ be two lists of two names with $length(ss) \geq length(ts)$.
-Define $\delta_{S, \ell}^* : S^* \times S^* \longrightarrow \mathbb{Z}^* \times S^*$ by taking the 
-alignment used to compute $\delta_V^*(\ell^*.\get \; ss, \ell^*.\get \; ts)$ and then using that
-alignment in the same manner that $\delta_V^*$ does. For instance
-\begin{align*}
-&\delta_{S, \ell}^*([``Ada \; Lovelace", ``Stephen \; Kleene", ``Alan \; Turing"], 
-[``Alan \; Turing", ``Ada \; Lovelace"])\\
-&= ([0,0], [``Stephen \; Kleene"])
-\end{align*}
-Recall that the \pput behaviour of $\ell_1$ simply aligns views to sources
-positionally, while $\ell_2$ computes the optimal alignment on the view side and
-uses this alignment to put views back into sources. From this we can prove that
-$$\ell_2 : S^* \xLeftrightarrow{\delta_{S, \ell}^*, \partial(S, V)^* \times \ell^*, \delta_V^*} V^*$$
-On the other hand, the judgement $\ell_1 : S^* \xLeftrightarrow{\delta_{S, \ell}^*, 
-\partial(S, V)^* \times \ell^*, \delta_V^*} V^*$ does not holds since $\ell_1$ does not satisfy \cref{difflenslaw} :
-$$(\partial(S, V)^* \times \ell^*).\get \; (\delta_{S, \ell}^*(s, \ell.\pput \; s \; v)) = \delta_V^*(\ell.\get
-\; s, v)
-$$
-This is witnessed by the fact that if $s = [``Ada \; Lovelace", ``Stephen \; Kleene", ``Alan \; Turing"]$ and \\
-$v = [``Kleene", ``Turing", ``Lovelace"]$, then 
-\begin{align*}
-&\ell_1.\pput \; [``Ada \; Lovelace", ``Stephen \; Kleene", ``Alan \; Turing"] \; [``Kleene", ``Turing", ``Lovelace"]\\
-&= [``Ada \; Kleene", ``Stephen \; Turing", ``Alan \; Lovelace"]
-\end{align*}
-Also, $\delta_{S, \ell}^*(s, \ell.\pput \; s \; v) = ([1,1,1], [])$, $\delta_V^*(\ell.\get
-\; s, v) = ([0,0,0], [])$, but
-\begin{align*}
-(\partial(S, V)^* \times \ell^*).\get \; (\delta_{S, \ell}^*(s, \ell.\pput \; s \; v)) &=
-(\partial(S, V)^* \times \ell^*).\get \; ([1,1,1], [])\\
-&= ([1,1,1], [])\\
-&\neq ([0,0,0], [])\\
-&= \delta_V^*(\ell.\get \; s, v)
-\end{align*}
-Here the positional lens $\ell_1$ is not a difference lens with respect to $(\delta_{S, \ell}^*, 
-\partial(S, V)^* \times \ell^*, \delta_V^*)$ since it does not align sources to views in the same way that
-$\delta_V^*$ does. On the other hand, $\ell_2$ does, hence is a difference lens with respect to $(\delta_{S, \ell}^*, 
-\partial(S, V)^* \times \ell^*, \delta_V^*)$.
-
-In the rest of this paper, we give concrete examples of how to use $\partial D$ preservation to characterize 
-the guarantees on update propagation provided by various lens programming languages. In particular,
-the major technical contributions that we will make in this paper are the following:
-\begin{enumerate}
-  \item
-  We formalize a condition called $\partial D$ preservation which provides a way
-  of constraining the updates that can be made by a bidirectional
-  transformation. We give examples of relations $\partial D$ as well as
-  examples and counterexamples of lenses that are $\partial D$ preserving.
-  \item
-  We describe an instance of $\partial D$ preservation that is satisfied by {\em difference lenses}.
-We give a set of combinators for {\em difference lenses} and describe the update propagation guarantees that the lenses derived using these combinators give.
-  \item
-  Finally, we show how to represent matching lenses and relational lenses
-  \cite{bohannon2006relational} as difference lenses, and hence
-  describe non-trivial guarantees beyond correctness and hippocraticness that
-  these lens languages provide as well.
-\end{enumerate}
-\iffalse
-A {\em lens} is a program that can be run in two directions. In the
-forward direction, the \get component of a lens produces a {\em view}
-from a {\em source}, while in the backward direction, the \pput component folds
-an updated view back into the original source. Formally, Foster
-\cite{foster2007combinators} defined a lens $\ell : S
-\Leftrightarrow V$ from a source $S$ to a view $V$ to be a pair of functions
-$\ell.\get : S \longrightarrow V$ and $\ell.\pput : S \longrightarrow V
-\longrightarrow S$ satisfying the
-following properties:
-\begin{align*}
-\ell.\get \; (\ell.\pput \; s \; v) &= v \tag{PUTGET}\\
-\ell.\pput \; s \; (\ell.\get \; s) &= s \tag{GETPUT}
-\end{align*}
-The PUTGET law ensures that updates to the view are translated
-``exactly''---i.e. that, given a view, the \pput function produces a source
-that \get maps back to the very same view, while the GETPUT law ensures a
-``stability'' property for the source---i.e. it requires that the \pput
-function returns the original source unmodified whenever the update to the view
-is a no-op.
-
-Unfortunately, while these laws give some guarantees on how the \get and \pput
-components of a lens interact, they do not specify the behaviour of the \pput
-component when the update to the view is non-trivial. Consequently, a
-non-trivial update to a view could lead to an unexpected update to the source.
-Various solutions have been proposed to address this issue, for example {\em
-matching lenses} to deal with the issue of alignment \cite{barbosa2010matching},
-{\em edit lenses} \cite{hofmann2012edit} and {\em delta lenses}
-\cite{diskin2011asymmetric,diskin2011state,pacheco2012delta} which explicitly
-propagate edits through the lens, and {\em least change lenses} which
-add a new LEASTPUTGET law which requires that the source value resulting from a
-\pput application is not only acceptable, but also one of closest to the
-original $s$ among the sources that share the same view $v$
-\cite{macedo2013composing}.
-
-This paper describes a general method for specifying what updates are allowable
-for a given lens without requiring that the lens explicitly propagate updates.
-Our approach is guided by the simple idea the idea that if one starts with a
-source $s$, calls the \get function on $s$ to acquire the view $\ell.\get \;
-s$, updates $\ell.\get \; s$ to get a new view $v$, then the ``difference''
-between $v$ and $\ell.\get \; s$ and $v$ should be related to the difference
-between $\ell.\pput \; s \; v$ and $v$.
-
-To formalize this idea, we first recall a widely used definition of
-bidirectional transformations
-\cite{meertens1998designing,stevens2010bidirectional} definition of a
-bidirectional transformations:
-\begin{definition}
-A bidirectional transformation (bx) $R: M \Leftrightarrow N$ comprises of a
-consistency relation $R \subseteq M \times N$ and restorers $\overrightarrow{R}
-: M \times N \rightarrow N$ and $\overleftarrow{R} : M \times N \longrightarrow
-N$.
-\end{definition}
-\begin{definition}
-A bx $R : M \leftrightarrow N$ is {\em correct} if the consistency restorers do
-restore consistency: that is, for all $m \in M, n \in N$, we have $R(m,
-\overrightarrow{R}(m, n))$, and dually. $R$ is {\em hippocratic} if the
-consistency restores make no change to and already-consistent pair of models:
-that is, for all $m \in M$, $n \in N$, if $R(m, n)$ then $\overrightarrow{R}(m,
-n) = n$ and dually. $R$ is {\em well behaved} if it is correct and hippocratic.
-\end{definition}
-
-Observe that $\ell : S \Leftrightarrow V$ is a lens, let $R$ be the
-consistency relation given by $R(s, \ell.\get \; s)$ for all $s \in
-S$, $\overrightarrow{\ell}(s, v) = \ell.\get \; s$, and $\overleftarrow{\ell} =
-\ell.\pput$. By the PUTGET law, $\ell.\get \; (\ell.\pput \; s \; v) = v$,
-hence $\ell$ is correct, and by the GETPUT laws, $\ell.\pput \; s \; (\ell.\get
-\; s) = s$ and $\ell.\get \; (\ell.\create \; v) = v$, hence $\ell$ is
-hippocractic. Therefore $\ell$ is a well-behaved bx.
-
-If $R$ is a well-behaved bx then $R$ restores consistency when a source or view
-is updated (correctness) and preserves consistency when the source and view are
-already consistent (hippocraticness). In order to express
-
-We will then apply this method to define a typing judgement for lenses which
-constrains how a lens propagates updates. We will describe a compositional
-language of {\em difference lenses} which satisfy this new typing judgement,
-and give examples of ordinary lenses that are not difference lenses. Finally,
-we will show how to represent matching lenses and relational lenses
-\cite{bohannon2006relational} as difference lenses, and hence describe the
-guarantees that these lens languages provide.
-
-For example suppose that the source is an XML document representing the
-names, dates and nationalities of a collection of classical music
-composers such as,
-
-\begin{lstlisting}
-<composers>
-<composer>
-<name>Jean Sibelius</name>
-<dates>1865-1956</dates>
-<nationality>Finnish</nationality>
-</composer>
-<composer>
-<name>Aaron Copland</name>
-<dates>1910-1990</dates>
-<nationality>American</nationality>
-</composer>
-<composer>
-<name>Benjamin Briten</name>
-<dates>1913-1976</dates>
-<nationality>English</nationality>
-</composer>
-</composers>
-\end{lstlisting}
-
-while the view is a list if ASCII text representing the names and dates of each
-composer:
-\begin{lstlisting}
-Jean Sibelius, 1865-1956
-Aaron Copland, 1910-1990
-Benjamin Briten, 1913-1976
-\end{lstlisting}
-
-After computing the initial view, we might want to edit it in some way---e.g.,
-correcting the error in Sibelius's death date and the misspelling in Britten's
-name
-\begin{lstlisting}
-Jean Sibelius, 1865-|\colorbox{orange}{1957}|
-Aaron Copland, 1910-1990
-Benjamin |\colorbox{orange}{Britten}|, 1913-1976
-\end{lstlisting}
-and push the changes back into the original XML format:
-\begin{lstlisting}
-<composers>
-<composer>
-<name>Jean Sibelius</name>
-<dates>1865-|\colorbox{orange}{1957}|</dates>
-<nationality>Finnish</nationality>
-</composer>
-<composer>
-<name>Aaron Copland</name>
-<dates>1910-1990</dates>
-<nationality>American</nationality>
-</composer>
-<composer>
-<name>Benjamin |\colorbox{orange}{Britten}|</name>
-<dates>1913-1976</dates>
-<nationality>English</nationality>
-</composer>
-</composers>
-\end{lstlisting}
-
-Lenses are designed to perform these kinds of bidirectional transformations. If
-$\ell$ is a suitably defined lens for the transformation that we
-have just described, then the \get component of the lens performs the
-XML to ASCII transformation that produces an ASCII view from an XML source,
-while the \pput component of the lens performs the ASCII to XML transformation
-that folds back an updated ASCII view back into the original XML source.
-
-For example, suppose that a source $S$ consists of a pair consisting of a
-boolean value and an integer, and that the view $V$ for $S$ consists of a
-single integer. Further, suppose that we wish to define a \pput function for a
-lens $\ell$ whose \get component is the function which given a pair $(b, n)$
-simply produces the boolean value $b$. Let $f : \mathbb{B} \times \mathbb{Z}
-\longrightarrow \mathbb{Z}$ be any function from $\mathbb{B} \times \mathbb{Z}
-\longrightarrow \mathbb{Z}$. Then the lens $\ell_{f}$ defined by
-\begin{align*}
-\ell_{f}.\get \; (b, n) &= b\\
-\ell_{f}.\pput \; (b, n) \; b' &= \begin{cases}
-(b', n) & \text{if }b = b'\\
-(b', f(b', n)) & \text{otherwise}\\
-\end{cases}\\
-\ell.\create \; b &= (b,f(b, 0))
-\end{align*}
-is a valid lens.
-
-A user who applies a lens $\ell$ of type $\ell : S \Leftrightarrow V$ may expect
-that $\ell.\pput \; (true, 1) \; false = (false, 0)$, reasoning that the
-integer produced by the \pput function should be the number representation of
-the boolean value $false$. Or perhaps she might expect that $\ell.\pput \;
-(true, 1) \; false = (false, 1)$, reasoning that the two pieces of data in a
-pair are independent and that the reasonable \pput behaviour in this case is to
-restore the orginial integer. In either case, the user may be shocked to find
-that in fact $\ell.\pput (true, 0) \; false = (false, 48)$! Given the \get
-behaviour which just produces the boolean value from a pair, since $\ell_f$ is a
-valid lens for any $f$, then the \pput function of a valid lens of type
-$\mathbb{B} \times \mathbb{Z} \Leftrightarrow \mathbb{Z}$ can restore the
-integer in a pair in any way so long as the old boolean value and the updated
-boolean value are different.
-
-In order to give a more constrained specification for the desired
-lens we define the difference $\delta_S((a, b), (c, d))$ between $(a, b)$ and
-$(c, d)$ by $$\delta_S((a, b), (c,
-d)) = \begin{cases}
-0 & \text{if }a = c, b = d\\
-1 & \text{if }a=c, b \neq d\\
-1 & \text{if }a \neq c, b = d\\
-2 & \text{if }a \neq c, b \neq d
-\end{cases}$$
-On the view side, we define the difference $\delta_V(a, b)$
-between $a$ and $b$ by, $$\delta_V(a, b) = \begin{cases}
-0 & \text{if } a = b\\
-1 & \text{if } a \neq b
-\end{cases}$$
-
-Next we define a lens $\partial(S, V) : \partial S \Leftrightarrow \partial V$
-which describes how differences should be propagated. For this example, we
-define $\partial(S,V)$ to simply be the \ccopy lens from $\partial S$ to
-$\partial V$, keeping in mind that $\partial S = \partial V = \mathbb{Z}$.
-\cref{difflenslaw} enables us to rule out $\ell_{bad}$, since
-\begin{align*}
-(\partial (S, V)).\get \; (\delta_S(\ell_{bad}.\pput \; (1,2) \; 3, (1,2))) &=
-(\partial (S, V)).\get \; (\delta_S((0,3), (1,2)))\\
-&= (\partial (S, V)).\get \; 2\\
-&= 2\\
-&\neq 1\\
-&= \delta_V(3, 2)\\
-&= \delta_V(3, \ell_{bad}.\get \; (1, 2))
-\end{align*}
-Consequently $\ell_{bad}$ is not a valid delta lens. On the
-other hand, the lens $\ell_{good} : S \Leftrightarrow V$ defined by
-\begin{align*}
-\ell_{good}.\get \; (a, b) &= b\\
-\ell_{good}.\pput \; (a, b) \; c &= (a, c)\\
-\ell_{good}.\create \; b &= (0, b)
-\end{align*}
-is indeed a valid delta lens since the its \pput component always restores the
-first factor.
-
-\section{Strengthening Lens Laws}
-\begin{definition}
-A bidirectional transformation (bx) $R: M \Leftrightarrow N$ comprises of a
-consistency relation $R \subseteq M \times N$ and restorers $\overrightarrow{R}
-: M \times N \rightarrow N$ and $\overleftarrow{R} : M \times N \longrightarrow
-N$.
-\end{definition}
-\begin{definition}
-A bx is {\em correct} if the consistency restorers do restore consistency:
-that is, for all $m \in M, n \in N$, we have $R(m, \overrightarrow{R}(m, n))$,
-and dually.
-\end{definition}
-\begin{definition}
-A bx is {\em hippocratic} if the consistency restores make no change
-to and already-consistent pair of models: that is, for all $m \in M$, $n \in N$,
-if $R(m, n)$ then $\overrightarrow{R}(m, n) = n$ and dually.
-\end{definition}
-\begin{definition}
-A bx is {\em well behaved} if it is correct and hippocratic.
-\end{definition}
-For lenses, if $\ell : S \Leftrightarrow V$ is a lens, we take the
-consistency relation $R$ to be given by $R(s, \ell.\get \; s)$ for all $s \in
-S$, $\overrightarrow{\ell}(s, v) = \ell.\get \; s$, and $\overleftarrow{\ell} =
-\ell.\pput$. By the PUTGET law, $\ell.\get \; (\ell.\pput \; s \; v) = v$,
-hence $\ell$ is correct, and by the GETPUT laws, $\ell.\pput \; s \; (\ell.\get
-\; s) = s$ and $\ell.\get \; (\ell.\create \; v) = v$, hence $\ell$ is
-hippocractic. Therefore $\ell$ is a well-behaved bx.
-
-Using the notation of Stevens \cite{stevens2010bidirectional},
-
-The major shortcoming of this model of bx transformations is that it does not
-take into account how changes should be propagated by transformation; a
-well-behaved bx need only restore consistency when models go out of sync or
-maintain consisteny if models are already in sync. This becomes an issue when
-there are many different ways of restoring consistency
-
-\section{Motivation and Related Work}
-Foster et. al. \cite{foster2007combinators,foster2009bidirectional}
-defined a lens $\ell : S \Leftrightarrow V$ from a source $S$ to a view $V$ to
-be a triple of functions $\ell.\get : S \longrightarrow V$, $\ell.\pput : S
-\longrightarrow S \longrightarrow V$ and $\ell.\create : V \longrightarrow S$
-satisfying the following properties:
-\begin{align*}
-\ell.\get \; (\ell.\pput \; s \; v) &= v \tag{PUTGET}\\
-\ell.\get \; (\ell.\create \; v) &= v \tag{CREATEGET}\\
-\ell.\pput \; s \; (\ell.\get \; s) &= s \tag{GETPUT}
-\end{align*}
-While these two laws give some guarantees on how the \get, \pput and \create
-components of a lens interact, they do not specify the behaviour of the \pput
-component when the update to the view is non-trivial, because the GETPUT law
-only says that the \pput component must return the original source unmodified
-whenever the update to the view is a no-op. Consequently, a non-trivial update
-to a view could lead to an undesired update to the source.
-
-Therefore, several lens formalism have been proposed that explicitly have a
-notion of an {\em update}. The first of these was Barbosa et. al.'s {\em
-matching lenses}\cite{barbosa2010matching}. Matching lenses specifically deal
-with the issue of {\em alignment}, since Foster et. al.'s original iteration
-lens updates records positionally:
-\begin{align*}
-\ell^* .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n) &= t_1 \cdot
-\ldots \cdot t_n\\
-\text{ where } t_i &= \begin{cases}
-\ell.\pput \; s_i \; v_i & \text{if } 1 \leq i \leq \min\{m, n\}\\
-\ell.\create \; v_i & \text{if } \min\{m, n\} < i \leq n
-\end{cases}\\
-\end{align*}
-In their work, Barbosa et. al. enrich the types of lenses with ``chunks''
-identifying reorderable pieces of the source and view that should be re-aligned
-after an update, and formulate behavioral laws that capture essential
-constraints on the handling of chunks. They develop a core language
-of matching lenses for strings, together with a set of ``alignment
-combinators'' that implement a variety of alignment strategies. The obvious
-shortcoming of matching lenses is that they only deal with alignment; they do
-not solve the general issue of update propagation.
-
-The next line of work is Diskin et. al.'s {\em delta
-lenses}\cite{diskin2011asymmetric,diskin2011state}. Diskin et. al.'s delta
-lenses work as follows: let $\ell : S \Leftrightarrow V$ be a lens from
-$S$ to $V$. Assume that $s$ is a source, and that $v = \ell.\get s$. A user
-performs an edits $d$ on $v$ to get $v'$: $v \xrightarrow{d} v'$. The \pput
-function now takes the view edit $d$ and the original source $s$ and produces a
-new source $s'$: $s \xrightarrow{\pput(d, s)} s'$. One thing that Diskin et. al.
-do not do is that they do not give any concrete examples of delta lenses,
-no composition, no notion of equivalence, and no combinators for constructing
-delta lenses.
-
-The next line of work is Hofmann et. al.'s {\em edit
-lenses}\cite{hofmann2012edit}. Edit lenses are similar to delta lenses, but edit
-lenses are more explicit. Specifically, the source and view edits are
-represented as elements of a {\em monoid}, and the lens acts on sources and
-views via monoid actions. That is if $v$ is a view and $d_v$ is an edit that
-can be made on $v$ to produce $v'$, then $d_v \cdot v = v'$. The intended usage
-of an edit lens is as follows. There are two users, one holding an element of
-$S$ the other one an element of $V$ both referred to as {\em replicas}.
-Initially, they hold $init_S \in S$ and $init_V \in V$ respectively. The users
-then perform actions and propagate them across the lens. An action consists of
-producing an edit $d_s$ (or $d_v$), applying it to one's current replica $s$
-(resp. $v$), putting the edit through the lens to obtain an edit $d_v$ (resp.
-$d_s$), and asking the user on the other side to apply $d_v$ (resp. $d_s$) to
-their replica. One thing that Hofmann et. al. do that Diskin et. al. do not do
-is that they do give concrete examples of edit lenses, define a suitable notion
-of composition on them, and define combinators for constructing edit lenses.
-
-The final line of work in this area is Pacheco et. al.'s
-version of delta lenses\cite{pacheco2012delta}. Pacheco et. al. are the
-first to give concrete examples of delta lenses, define a suitable notion
-of composition on them, and define combinators for constructing delta
-lenses. Pacheco et al. define a delta lens $\ell$ (d-lens for short), denote by
-$\ell: S \trianglerighteq_{\blacktriangle} V$ to be a bidirectional
-transformations that comprises four total functions:
-\begin{align*}
-\get &: S \longrightarrow V\\
-\pput &: \forall(v, s) : V \times S. (v \triangle \get \;s) \longrightarrow
-SA\\
-\get_{\blacktriangle} &: \forall\{s':S, s:S\}. (s' \triangle s)
-\longrightarrow (\get \; s')\triangle (\get \; s)\\
-\pput_{\blacktriangle} &:\forall\{(v, s):V \times S\},d:(v \triangle \get
-\;s).(\pput \; (v, s) \; d) \triangle s
-\end{align*}
-that satisfy the following propertes:
-\begin{align*}
-\get \; (\pput \; (v, s) \; d) = v \tag{PUTGET}\\
-\pput \; (\get \; s, s) \; id = s \tag{GETPUT}\\
-\get_{\blacktriangle} \; (\pput_{\blacktriangle} \; d) = d
-\tag{$\text{PUTGET}_{\blacktriangle}$}\\
-\pput_{\blacktriangle} \; id = id \tag{$\text{PUTGET}_{\blacktriangle}$}
-\end{align*}
-The \get function is a just like a \get function in Foster's et. al.'s original
-work. The \pput function is different though. Let $s \in S$ and $v \in V$, and
-let $d$ be an update from $v$ to $\ell.\get \; s$; for example, $d$ is a
-sequence of insertions and deletions that transforms $v$ to $\ell.\get \; s$.
-Then \pput takes in $s$ and $v$ as input, and in addition receives $d$ to
-compute $\pput \; (v, s) \; d$.
-
-The $\get_{\blacktriangle}$ and $\pput_{\blacktriangle}$ functions are functions
-that transform deltas to deltas. The last two axioms for delta lenses say that
-if $d$ is a delta from $v$ to $\ell.\get \; s$ and $\Delta$ is a delta from
-$\pput \; (v, s) \; d$ to $s$, then $\get_{\blacktriangle} \; \Delta = d$, and
-that if $d$ is the identity delta, then $\Delta$ is also the identity delta.
-
-In addition to describing a delta lens language, Pacheco et. al. also
-address the following shortcoming with edit lenses: edit lenses include
-combinators for constructing lenses over {\em container types}.
-In the simplest case, container types can be thought of as algebraic datatypes
-such as lists and trees; each value of a container type has a certain {\em
-shape}, and each shape has a set of {\em positions} associated with it. For
-instance, the shape of a list of size $n$ is just $n$, while the positions of
-such a list are the numbers $0, 1, \ldots, n-1$.
-
-The shortcoming of the container edit lenses addressed by Pacheco et. al.
-elements is that these lenses always preserve the number of positions; for
-instance if $\ell$ is a container lens that maps lists to trees, and if $xs$ is
-a list of size $n$, then the number of positions of $\ell.\get \; xs$ is also
-$n$. Pacheco et. al. define a new set of combinators for container lenses that
-can say map a list of size $5$ to a tree with 4 nodes.
-
-In this article we show that it is not necessary for the \pput component of a lens
-to explicitly propagate updates: instead, we can specify what the put function should
-evaluate to, by imposing a more precise type judgement on lenses.
-
-The idea of the stronger typing judgement is that we embed sources and views inside
-of {\em difference structures} and then we define a lens that describes how the
-differences behave as they cross in-between the source and the view. A
-lens that is validly typed is then a lens that preserves this overall structure.
-
-Here are the actual definitions:
-\begin{definition}
-A weighted set is a pair $(X, |\cdot|)$ where $X$ is a set and $|\cdot| : X
-\longrightarrow \mathbb{R}^{\geq 0}$ is a map from $X$ to the non-negative
-reals.
-\end{definition}
-\begin{definition}
-A difference structure is a triple $(S, \partial S, \delta_S)$ where $S$ and
-$\partial S$ are weighted sets, and $\delta_S:
-S \times S \longrightarrow \partial S$ is a function satisfying
-\begin{align*}
-\delta(s, s') &= \delta(s', s)\tag{A1}\\
-|\delta(s, s)| &= 0\tag{A2}\\
-|\delta(s, s')| &\leq |s| + |s'|\tag{A3}
-\end{align*}
-\end{definition}
-\begin{definition}
-Let $(S, \partial S, \delta_S)$ and $(V, \partial V, \delta_V)$ be
-difference structures. Let $\partial(S, V) : \partial S \Leftrightarrow \partial
-V$ be a lens from $\partial S$ to $\partial V$. Then a lens $\ell : S
-\Leftrightarrow V$ also has type $\ell : (S, \partial S,
-\delta_S) \xLeftrightarrow{\partial(S, V)} (V, \partial V, \delta_V)$ if for
-all $s \in S, v \in V$,
-\begin{equation}\label{difflenslaw}
-(\partial (S, V)).\get\; (\delta_S(\ell.\pput \; s \; v, s)) = \delta_V(v,
-\ell.\get \; s)
-\end{equation}
-\end{definition}
-The extra difference structure on the sources and the views, as well as the lens
-$\partial(S, V)$ force the lens $\ell$ to satisfy \cref{difflenslaw}; this law is
-therefore a specification that constrains what $\ell$ can do.
-
-For example, suppose that a source $S$ consists of a pair of integers, and that the
-view $V$ for $S$ consists of a single integer. Further, suppose that we wish
-for the \get component of any lens from $S$ to $V$ to simply project away the
-first factor:that is for any $\ell : \mathbb{Z} \times \mathbb{Z}
-\Leftrightarrow \mathbb{Z}$, $\ell.\get \; (a, b) = b$ for all integers $a, b$.
-Suppose further that we wish to force the \pput function of any lens $\ell$
-defined on the data to restore the factor projected away by the \get function:
-that is, we wish to enforce that for all integers $a, b, c$, $\pi_1(\ell.\pput
-\; (a, b) \; c) = a$, where $\pi_1$ is the projection on the first factor.
-
-Now consider the lens $\ell_{bad}$ defined by
-\begin{align*}
-\ell_{bad}.\get \; (a, b) &= b\\
-\ell_{bad}.\pput \; (a, b) \; c &= \begin{cases}
-(a, c) & \text{if }b = c\\
-(0, c) & \text{otherwise}\\
-\end{cases}\\
-\ell.\create \; b &= (0,b)
-\end{align*}
-Then $\ell_{bad}$ is a valid lens even though $\ell_{bad}$ does not
-always restore the first factor. For example, observe that $\ell.\pput (1, 2) \;
-3 = (0,3)$. In order to give a more constrained specification for the desired
-lens we define the difference $\delta_S((a, b), (c, d))$ between $(a, b)$ and
-$(c, d)$ by $$\delta_S((a, b), (c,
-d)) = \begin{cases}
-0 & \text{if }a = c, b = d\\
-1 & \text{if }a=c, b \neq d\\
-1 & \text{if }a \neq c, b = d\\
-2 & \text{if }a \neq c, b \neq d
-\end{cases}$$
-On the view side, we define the difference $\delta_V(a, b)$
-between $a$ and $b$ by, $$\delta_V(a, b) = \begin{cases}
-0 & \text{if } a = b\\
-1 & \text{if } a \neq b
-\end{cases}$$
-
-Next we define a lens $\partial(S, V) : \partial S \Leftrightarrow \partial V$
-which describes how differences should be propagated. For this example, we
-define $\partial(S,V)$ to simply be the \ccopy lens from $\partial S$ to
-$\partial V$, keeping in mind that $\partial S = \partial V = \mathbb{Z}$.
-\cref{difflenslaw} enables us to rule out $\ell_{bad}$, since
-\begin{align*}
-(\partial (S, V)).\get \; (\delta_S(\ell_{bad}.\pput \; (1,2) \; 3, (1,2))) &=
-(\partial (S, V)).\get \; (\delta_S((0,3), (1,2)))\\
-&= (\partial (S, V)).\get \; 2\\
-&= 2\\
-&\neq 1\\
-&= \delta_V(3, 2)\\
-&= \delta_V(3, \ell_{bad}.\get \; (1, 2))
-\end{align*}
-Consequently $\ell_{bad}$ is not a valid delta lens. On the
-other hand, the lens $\ell_{good} : S \Leftrightarrow V$ defined by
-\begin{align*}
-\ell_{good}.\get \; (a, b) &= b\\
-\ell_{good}.\pput \; (a, b) \; c &= (a, c)\\
-\ell_{good}.\create \; b &= (0, b)
-\end{align*}
-is indeed a valid delta lens since the its \pput component always restores the
-first factor.
-
-For a more realistic example using a difference lens combinator, let $S
-= name \cdot dates$ and $V = name$. Let $\ell :
-S \Leftrightarrow V = (copy \; name) \cdot (del \; dates)$, and suppose that we
-wish to define a lens of type $S^* \Leftrightarrow V^*$, where the difference
-function used on the view side is the min-edit distance difference function.
-Assuming that we inductively use the difference function
-$$ \delta(d_1, d_2) =
-\begin{cases}
-0 & \text{if }d_1 = d_2 \\
-1 & \text{otherwise}
-\end{cases}
-$$
-for both the name and date fields, then with the min-edit distance difference
-function, the difference between the string $v$ given by
-\begin{lstlisting}
-John Doe
-Jane Smith
-\end{lstlisting}
-and the string $v'$ given by
-\begin{lstlisting}
-Jane Smith
-John Doe
-\end{lstlisting}
-is the empty list. Let $s$ =
-\begin{lstlisting}
-John Doe,1900-2000
-Jane Smith,1901-2001
-\end{lstlisting}Then $\ell.\get \; s = v$. Now assume that $v$ is edited to $v'$
-given by
-\begin{lstlisting}
-Mark Moe
-John Doe
-Jane Smith
-\end{lstlisting}
-where a new record \lstinline|Mark Moe| has been added to $v$.
-With the min-edit distance on the view side, $\delta(v', \ell.\get \; s)$ is
-the singleton list [``Mark Moe''].
-
-On the source side, the min-edit distance matching lens $\ell^*_{MIN-EDIT}$,
-$\ell^*_{MIN-EDIT}.\pput \; s \; v'$ is the string
-\begin{lstlisting}
-Mark Moe,0000-0000
-John Doe,1900-2000
-Jane Smith,1901-2001
-\end{lstlisting}
-while with the positional lens $\ell^*$, then $\ell^*.\pput \; s \; v'$ is the
-string
-\begin{lstlisting}
-Mark Moe,1900-2000
-John Doe,1901-2001
-Jane Smith,0000-0000
-\end{lstlisting}
-With the min-edit distance difference function,
-$\delta(\ell^*_{MIN-EDIT}.\pput \; s \; v', s)$ is the singleton list
-$[\text{``Mark Moe, 0000-0000''}]$, while with the
-positional lens, $\delta(\ell^*.\pput \; s \; v', s)$ is the list
-given by
-\begin{lstlisting}
-Mark Moe,1900-2000
-(0,1)
-(0,1)
-\end{lstlisting}
-The min-edit distance satisfies \cref{difflenslaw} since
-\begin{align*}
-(\partial (S, V))^*.\get\; (\delta_S(\ell^*_{MIN-EDIT}.\pput \; s \; v', s)) &=
-(\partial (S, V))^*.\get\; ([\text{``Mark Moe, 1900-2000''}])\\
-&= [\text{``Mark Moe''}]\\
-&=
-\delta_V(v', \ell.\get \; s)
-\end{align*}
-On the other hand, the positional lens does not satisfy \cref{difflenslaw} since
-\begin{align*}
-(\partial (S, V))^*.\get\; (\delta_S(\ell^*.\pput \; s \; v', s)) &=
-(\partial (S, V))^*.\get\; ([\text{``Mark Moe, 1900-2000''}, (0,1), (0,1)])\\
-&=[\text{``Mark Moe''}, (0,*), (0,*)]\\
-& \neq \delta_V(v', \ell.\get \; s)
-\end{align*}
-Let $\mathcal{M}_V$ be the set of min-edit distance alignments on the view side,
-and $\mathcal{M}_S$ the corresponding alignments on the source side.
-Then, the typing judgement
-$$\ell^*_{MIN-EDIT}:(S, (\partial S \sqcup S)^*, \match(\mathcal{M}_S, \delta_s))
-\xLeftrightarrow{\partial(S, V)^*} (V, (\partial V \sqcup V)^*, \match(\mathcal{M}_V, \delta_v))$$
-is valid, whereas the typing judgement
-$$\ell^*:(S, (\partial S \sqcup S)^*, \match(\mathcal{M}_S, \delta_s))
-\xLeftrightarrow{\partial(S, V)^*} (V, (\partial V \sqcup V)^*, \match(\mathcal{M}_V, \delta_v))$$
-is not.
-\iffalse
-For example, if $\ell : S \Leftrightarrow V$ is a lens, Foster's initial
-definition of the iteration lens $\ell^* : S^* \Leftrightarrow V^*$ was
-\begin{align*}
-\ell^* .\get \; (s_1, \ldots, s_n) &= (\ell.\get \; s_1) \cdot \ldots
-\cdot (\ell.\get \; s_n)\\
-\ell^* .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n) &= t_1 \cdot
-\ldots \cdot t_n\\
-\text{ where } t_i &= \begin{cases}
-\ell.\pput \; s_i \; v_i & \text{if } 1 \leq i \leq \min\{m, n\}\\
-\ell.\create \; v_i & \text{if } \min\{m, n\} < i \leq n
-\end{cases}\\
-\ell^*.\create \; (v_1, \ldots, v_n) &= (\ell.\create \; v_1) \cdot \ldots
-\cdot (\ell.\create \; v_n)
-\end{align*}
-This lens always propagates updates positionally, which is unintended behaviour
-in many cases. For example suppose that $\ell$ is a lens that transforms
-source data in the form of an XML document representing the names, dates and
-nationalities of a collection of classical music composers such as,
-
-\begin{lstlisting}
-<composers>
-<composer>
-<name>Jean Sibelius</name>
-<dates>1865-1957</dates>
-<nationality>Finnish</nationality>
-</composer>
-<composer>
-<name>Aaron Copland</name>
-<dates>1910-1990</dates>
-<nationality>American</nationality>
-</composer>
-<composer>
-<name>Benjamin Britten</name>
-<dates>1913-1976</dates>
-<nationality>English</nationality>
-</composer>
-</composers>
-\end{lstlisting}
-into view data stoored in ASCII text representing the names and dates of each
-composer:
-\begin{lstlisting}
-Jean Sibelius, 1865-1957
-Aaron Copland, 1910-1990
-Benjamin Britten, 1913-1976
-\end{lstlisting}
-Now suppose that the elements in the view are reordered:
-\begin{lstlisting}
-Aaron Copland, 1910-1990
-Benjamin Britten, 1913-1976
-Jean Sibelius, 1865-1957
-\end{lstlisting}
-Then with the positional iteration lens, the updated source is
-\begin{lstlisting}
-<composers>
-<composer>
-<name>Aaron Copeland</name>
-<dates>1910-1990</dates>
-<nationality>Finnish</nationality>
-</composer>
-<composer>
-<name>Benjamin Britten</name>
-<dates>1913-1976</dates>
-<nationality>American</nationality>
-</composer>
-<composer>
-<name>Jean Sibelius</name>
-<dates>1865-1957</dates>
-<nationality>English</nationality>
-</composer>
-</composers>
-\end{lstlisting}
-so each of the composers now has the wrong nationality in the source.
-
-\section{Difference functions}
-\begin{definition}
-A weighted set is a pair $(X, |\cdot|)$ where $X$ is a set and $|\cdot| : X
-\longrightarrow \mathbb{R}^{\geq 0}$ is a map from $X$ to the non-negative
-reals.
-\end{definition}
-\begin{definition}
-A difference structure is a triple $(S, \partial S, \delta_S)$ where $S$ and
-$\partial S$ are weighted sets, and $\delta_S:
-S \times S \longrightarrow \partial S$ is a function satisfying
-\begin{align*}
-\delta(s, s') &= \delta(s', s)\tag{A1}\\
-|\delta(s, s)| &= 0\tag{A2}\\
-|\delta(s, s')| &\leq |s| + |s'|\tag{A3}
-\end{align*}
-We refer to the function $\delta$ as the difference function of the structure.
-\end{definition}
-\iffalse
-
-\begin{example}
-Let $X$ be any set. Define $x \leq y$ if and only if $x=y$, and define $|X|=1$
-for all $x \in X$. Let $Q = \{0,1\}$, and define
-$$\delta(a, b) = \begin{cases} 1 & \text{if } a \neq b\\
-0 & \text{otherwise}
-\end{cases}$$
-Then $\delta$ is a difference function.
-\end{example}
-
-Various different edit distances on strings give a way of defining delta
-functions on sets of strings:
-\begin{example}
-The {\em Levenshtein distance} between two words is the minimum number of
-single-character edits (insertions, deletions or substitutions) required to
-change one word into the other. For example, the Levenshtein distance between
-"kitten" and "sitting" is 3, since the following three edits change one into
-the other, and there is no way to do it with fewer than three edits:
-
-\begin{enumerate}
-  \item
-  kitten -> sitten (substitution of "s" for "k")
-  \item
-  sitten -> sittin (substitution of "i" for "e")
-  \item
-  sittin -> sitting (insertion of "g" at the end).
-\end{enumerate}
-\end{example}
-\begin{example}
-In contrast, the Longest Common Subsequence(LCS) distance only allows for
-insertions and deletions:
-\begin{enumerate}
-  \item
-  delete k at 0
-  \item
-  insert s at 0
-  \item
-  delete e at 4
-  \item
-  insert i at 4
-  \item
-  insert g at 6
-\end{enumerate}
-for a total cost/distance of 5 operations.
-\end{example}
-Both Levenshtein and LCS distances are integer-valued distance functions since
-they are metrics.
-\begin{example}
-If all the strings in a set are of the same length, then the {\em Hamming
-distance} is a difference function since the Hamming distance is also a metric.
-\end{example}
-\begin{example}
-Alignment algorithms also give us a way of defining integer-valued distance
-functions on sets of strings by taking the difference of two strings to be the
-total number of characters that are not equal according to the alignment.
-
-When deciding which alignment to choose between two strings, we often wish to
-choose an alignment that is optimal for alignment strategy. With
-the positional alignment strategy, given two strings $s$ and $t$, then $s_1$ is
-matched to $t_1$, $s_2$ is matched to $t_2$ and so on.
-Also, $\mathcal{M}_{POS}$ is an optimal set of alignments for a positional
-alignment strategy.
-
-For example, using the positional alignment strategy on the strings "kitten" and
-"sitting", then "k" is matched to "s", "i" to "i" and so on. The characters
-which are included in the delta from "kitten" are "k", "e" while the
-in the characters included from "sitting" are "s", "i" and "g", hence the delta
-is equal to 5.
-
-For the non-crossing strategy, we choose among alignments which are increasing
-rather than positional, while for the min-edit distance strategy, we consider
-all possible alignments.
-\end{example}
-
-\begin{example}\label{naturalnumbers}
-Let $P$ be the set of non-negative integers with weight function $|a| = a$. Then
-$\delta$ is a difference function.
-Define $\delta(a, b) = |a-b|$. Then $\delta$ is a difference function.
-\end{example}
-\begin{example}
-Let $X$ be a normed vector space and define $\delta(x, y) = |x-y|$. Then
-$\delta$ is a difference function (in fact, $\delta$ is also a metric).
-\end{example}
-\begin{example}\label{divisibility}
-Let $P$ be the set of positive integers with weight function $|\cdot|$ defined
-by $|p_1^{a_1} \cdot \ldots \cdot p_n^{a_n}| = a_1 + \ldots + a_n$. Define
-$\delta(a, b) = ab/(gcd(a, b))^2$.
-Then $\delta$ is a difference function.
-\end{example}
-\begin{example}\label{boolean}
-Let $X$ be a set and $P$ the set of finite subsets of $X$ with weight function
-given by the size of the set. Define $\delta(A, B) = (A \setminus B) \cup (B
-\setminus A)$. Then $\delta$ is a difference function.
-\end{example}
-\begin{example}\label{general}
-Generalizing \cref{divisibility} and
-\cref{boolean}, let $P$ be a graded complete semi-modular lattice satisfying
-$|\bigwedge P|=0$. Define $\delta(a, b) = \bigwedge \{x \; | \; (a \wedge b)
-\vee x = a\} \vee \bigwedge \{x \; | \; (a \wedge b) \vee x = b\}$. Then
-$\delta$ is a difference function.
-\end{example}
-\begin{proof}
-Let $a, b \in P$. Let $X = \{x \; | \; (a \wedge b) \vee x = a\}$ and $Y = \{y
-\; | \; (a \wedge b) \vee y = b\}$
-\begin{enumerate}
-  \item[(A1)]
-  Certainly $\delta(a, b) = \delta(b, a)$.
-  \item[(A2)]
-  Observe that $\bigwedge X = \bigwedge Y = \bot$, hence $\delta(a,
-  a) = \bot$, and $|\delta(a, a)| = |\bot| = 0$.
-  \item[(A3)]
-  Since $P$ is semi-modular, then $|x \wedge y| + |x \vee y| \leq |x| + |y|$
-  for all $x, y \in P$, from which it follows that $$|\delta(a, b)| =
-  \left|\bigwedge X \vee \bigwedge Y\right| \leq \left|\bigwedge X \vee
-  \bigwedge Y\right| + \left|\bigwedge X \wedge \bigwedge Y\right| \leq
-  \left|\bigwedge X\right| + \left|\bigwedge Y\right|$$
-\end{enumerate}
-\end{proof}
-\fi
-
-\subsection{Constructions on difference functions}
-\subsubsection{Products}
-Given difference functions $\delta_i: P_i \times P_i \longrightarrow Q_i$ for
-$i \in \{1,2\}$, define $\delta_1 \times \delta_2 :(P_1 \times P_2) \times
-(P_1 \times P_2) \longrightarrow (Q_1 \times Q_2)$ by
-$$(\delta_1 \times \delta_2)((x, y), (x', y')) = (\delta_1(x, x'), \delta_2(y,
-y'))$$
-\begin{claim}
-$\delta_1 \times \delta_2$ is a difference function.
-\end{claim}
-\begin{proof}
-Let $x, x' \in P_1$ and $y, y' \in P_2$.
-\begin{enumerate}
-  \item[(A1)]
-  Observe that
-  \begin{align*}
-(\delta_1 \times \delta_2)((x, y), (x', y')) &= (\delta_1(x, x'),
-\delta_2(y, y'))\\
-&= (\delta_1(x', x), \delta_2(y', y))\\
-&= (\delta_1 \times \delta_2)((x', y'), (x, y))
-\end{align*}
-\item[(A2)] Observe that
-$$|(\delta_1 \times \delta_2)((x, y), (x, y))| = |(\delta_1(x, x),
-\delta_2(y, y))| = |\delta_1(x, x)| + |\delta_2(y,y)| = 0 + 0 = 0$$
-\item[(A3)]
-Observe that
-\begin{align*}
-|(\delta_1 \times \delta_2)((x, y), (x', y'))| &= |(\delta_1(x, x'),
-\delta_2(y, y'))|\\
-&= |(\delta_1(x, x')| + |(\delta_2(y, y')|\\
-&\leq (|x| + |x'|) + (|y| + |y'|)\\
-&= (|x| + |y|) + (|x'| + |y'|)\\
-&= |(x, y)| + |(x', y')|
-\end{align*}
-\end{enumerate}
-\end{proof}
-\subsubsection{Coproducts}
-Given difference functions $\delta_i: P_i \times P_i \longrightarrow Q_i$ for
-$i \in \{1,2\}$, define $\delta_1 \sqcup \delta_2 :P_1 \sqcup P_2
-\longrightarrow Q_1 \sqcup Q_2 \sqcup (P_1 \times P_2)$ by
-$$(\delta_1 \sqcup \delta_2)(x, y) =
-\begin{cases}
-\delta_1(x, y) & \text{if }x, y \in P_1\\
-\delta_2(x, y) & \text{if }x, y \in P_2\\
-(x, y) & \text{if }x \in P_1, y \in P_2\\
-(y, x) & \text{if }x \in P_2, y \in P_1
-\end{cases}$$
-\begin{claim}
-$\delta_1 \sqcup \delta_2$ is a difference function
-\end{claim}
-\begin{proof}
-Let $x, x' \in P_1$ and $y, y' \in P_2$.
-\begin{enumerate}
-  \item[(A1)]
-  If $x, y \in P_i$, then $(\delta_1 \sqcup \delta_2)(x, y) = \delta_i(x, y) =
-  \delta_i(y, x)$. If $x \in P_1$ and $y \in P_2$, then $(\delta_1 \sqcup
-  \delta_2)(x, y) = (x, y) = (\delta_1 \sqcup \delta_2)(y, x)$, and similarly
-  if $x \in P_2$ and $y \in P_1$, then $(\delta_1 \sqcup \delta_2)(x, y) =
-  (y,x) = (\delta_1 \sqcup \delta_2)(y, x)$
-  \item[(A2)] If $x \in P_i$, then $|(\delta_1 \sqcup \delta_2)(x, x)| =
-  |\delta_i(x, x)| = 0$
-  \item[(A3)]
-  If $x, y \in P_i$, then $|(\delta_1 \sqcup \delta_2)(x, y)| = |\delta_i(x, y)|
-  \leq |x| + |y|$. If $x \in P_1$ and $y \in P_2$, then $|(\delta_1 \sqcup
-  \delta_2)(x, y)| = |(x, y)| = |x| + |y|$, and if $x \in P_2$ and $y \in
-  P_1$, then $|(\delta_1 \sqcup \delta_2)(x, y)| = (|(y, x)|) = |y| + |x|$.
-\end{enumerate}
-\end{proof}
-\subsubsection{List difference functions}
-Let $P$ be a weighted set. Let $x=(x_1, \ldots, x_m), y=(y_1, \ldots, y_n)$ be
-in $P^*$. Given a difference function $\delta: P \times P \longrightarrow Q$, we
-define the list difference function $\delta^*:P^* \longrightarrow (Q \sqcup
-P)^*$ on $x$ and $y$ to be the list given by $\delta(x_i, y_i)$ for $1 \leq i
-\leq m$, except that we omit $\delta(x_i, y_i)$ in position $i$ if $x_i = y_i$.
-For $m < i \leq n$ we add $y_i$ at position $i$. Finally we collapse the list
-to get rid of the empty spaces left after the omissions.
-
-For instance suppose that $S$ is the set of all letters in the alphabet and
-$\partial S = \{\epsilon\} \sqcup \{\{x, y\} \; | \; x, y \in S, x \neq y\}$
-with $$\delta(x, y) =
-\begin{cases}
-\epsilon & \text{if } x = y\\
-\{x,y\} & \text{if } x \neq y
-\end{cases}$$
-Then $\delta_S^*((a, c, b, d), (a, b, c, d, e)) = (\{b, c\}, \{b, c\},
-e)$ where we match and remove the first $a$ and the fourth $d$ in. Since $b \neq
-c$ we have $\{b,c\}$ in the second and third positions, and since $e$ is
-unmatched, we include $e$ in the fifth position. Finally we collapse the list to
-get rid of the two empty spaces created.
-
-\begin{claim}
-$\delta^*$ is a difference function
-\end{claim}
-\begin{proof}
-Let $x = (x_1, \ldots, x_m), y = (y_1, \ldots, y_n) \in P^*$.
-\begin{enumerate}
-  \item[(A1)]
-  By construction $\delta^*(x, y) = \delta^*(y, x)$.
-  \item[(A2)] Observe that $|\delta^*(x, x)| = \sum_{i=1}^m|\delta(x_i,
-  x_i)| =  \sum_{i=1}^m 0 = 0$.
-  \item[(A3)]
-  Assume WLOG that $m \leq n$. Then
-  \begin{align*}
-|\delta^*(x, y)| &=
-\sum_{i=1}^m|\delta(x_i, y_i)| + \sum_{i=m+1}^n |y_i|\\
-&\leq \sum_{i=1}^m|x_i| + |y_i| + \sum_{i=m+1}^n |y_i|\\
-&= \sum_{i=1}^m|x_i| + \sum_{i=1}^n |y_i|\\
-&= |x| + |y|
-\end{align*}
-\end{enumerate}
-\end{proof}
-\subsubsection{Matching difference functions}
-Now let $\mathcal{M} = \{\varphi_{(x_1, \ldots, x_m), (y_1, \ldots, y_n)} \; |
-\; x_i, y_j \in P, m \leq n\}$ be a set of injective functions $\varphi_{(x_1,
-\ldots, x_m), (y_1, \ldots, y_n)} : [m] \longrightarrow [n]$. Given a difference
-function $\delta:
-P \times P \longrightarrow Q$, we define the difference function
-$\match(\mathcal{M}, \delta):P^* \longrightarrow (Q \sqcup P)^*$ on $x$
-and $y$ to be the list given by $\delta(x_i, y_{\varphi(i)})$ in position
-$\varphi(i)$ for $1 \leq i \leq m$, except that we omit $\delta(x_i,
-y_{\varphi(i)})$ in position $\varphi(i)$ if $x_i = y_{\varphi(i)}$. For $i \in
-[n] \setminus \varphi([m])$ we add $y_i$ at position $i$. Finally we collapse
-the list to get rid of the empty spaces left after the omissions.
-
-Going back to our previous example, let $x = \{a, c, b, d\}$ and $y = \{a, b, c,
-d, e\}$, and assume that $\varphi_{x, y} : [4] \longrightarrow [5]$ is given by
-$$\varphi(1) = 1, \; \varphi(2) = 3, \; \varphi(3) = 2, \text{ and } \varphi(4)
-= 4$$ Then $\match(\mathcal{M}, \delta)((a, c, b, d), (a, b, c, d, e)) =
-(e)$.
-\begin{claim}
-$\match(\mathcal{M}, \delta)$ is a difference function.
-\end{claim}
-\begin{proof}
-Let $x = \{x_1, \ldots, x_m\}, y = \{y_1, \ldots, y_n\} \in \match(P)$.
-\begin{enumerate}
-  \item[(A1)]
-  By construction $\match(\mathcal{M}, \delta)(x, y) =
-  \match(\mathcal{M}, \delta)(y, x)$.
-  \item[(A2)] Observe that $|(\match(\mathcal{M}, \delta))(x, x)| =
-  \sum_{i=1}^m|\delta(x_i, x_{\varphi(i)}| = |\delta(x_i, x_i)| =   \sum_{i=1}^m
-  0 = 0$.
-  \item[(A3)]
-  Assume WLOG that $m \leq n$. Then
-  \begin{align*}
-|\match(\mathcal{M}, \delta)(x, y)| &=
-\sum_{i=1}^m|\delta(x_i, y_{\varphi(i)})| + \sum_{i \in [n] \setminus
-\varphi([m])} |y_i|\\
-&\leq \sum_{i=1}^m|x_i| + |y_{\varphi(i)}| + \sum_{i \in [n] \setminus
-\varphi([m])} |y_i|\\
-&= \sum_{i=1}^m|x_i| + \sum_{i=1}^n |y_i|\\
-&= |x| + |y|
-\end{align*}
-\end{enumerate}
-\end{proof}
-\fi
-\section{Lens Combinators}
-\subsection{Copy}
-Let $P$ be any set. Define the $\ccopy(P) : P \Leftrightarrow P$ by
-\begin{align}
-\ccopy(P).\get \; s &= s\\
-\ccopy(P).\pput \; s \; v &= v\\
-\ccopy(P).\create \; v &= v
-\end{align}
-Then $\ccopy(P)$ is a lens. Now let $\partial S$ be a set, $\delta :
-S \times S \longrightarrow \partial S$ a function.
-\begin{claim}
-$\ccopy(S) : S \xLeftrightarrow{\delta , \ccopy(\partial S), \delta} S$.
-\end{claim}
-\begin{proof}
-We need to show that $(\ccopy(\partial S)).\get \; (\delta(\ccopy(S).\pput
-\; s \; v, s)) = \delta(v, \ccopy(S).\get \; s)$ which follows from the fact
-that $$ (\ccopy(\partial S)).\get \; (\delta(\ccopy(S).\pput \; s
-\; v, s)) = \delta(\ccopy(S).\pput \; s \; v, s) = \delta(v, \ccopy(S).\get
-\; s) $$
-\end{proof}
-\subsection{Constant}
-Let $S$ be any set with $s'\in S$. Define the \const lens $\const(S) : S
-\Leftrightarrow \{\star\}$ by
-\begin{align*}
-\const.\get \; s &= \star\\
-\const.\pput \; s \; v &= s\\
-\const.\create \; v &= s'
-\end{align*}
-Then $\const(S)$ is a lens. 
-\begin{claim}
-Let $\partial S$ be any set, and $\delta : S \times S \longrightarrow \partial
-S$ any function satisfying $\delta(s, s) = \delta(s', s')$ for all $s, s' \in
-S$. Define $\beta : \{\star\} \times \{\star\} \longrightarrow \partial S$ by
-$\beta(\star, \star) = \delta(s, s)$ for any $s \in S$. Let $\ell : \partial S
-\Leftrightarrow \partial S$ be any lens satisfying $\ell.\get \; \delta(s, s) =
-\delta(s, s)$ for all $s \in S$. Then $\const(S): S \xLeftrightarrow{\delta,
-\ell, \beta} \{\star\}$.
-\end{claim}
-\begin{proof}
-We need to show that $\ell.\get \; (\delta(\const(S).\pput \; s \; \star, s)) =
-\beta (\star, \const(S).\get \; s)$
-which follows from the fact that
-$$ \ell.\get \; (\delta(\const(S).\pput \; s \; \star, s)) = \ell.\get
-\; \delta(s, s) = \delta(s, s) = \beta(\star, \star) = \beta(\star,
-\const(S).\get \; s)$$
-\end{proof}
-
-\subsection{Default}
-Let $\ell : S \Leftrightarrow V$ be a lens. Let $f :
-V \longrightarrow S$ be any function. Define the lens $\default(\ell, f) :
-S \Leftrightarrow V$ by
-\begin{align*}
-\default(\ell, f).\get \; s &= \ell.\get \; s\\
-\default(\ell, f).\pput \; s \; v &= \ell.\pput \; s \; v\\
-\default(\ell, f).\create \; v &= \ell.\pput \; (f \; v) \; v
-\end{align*}
-
-\begin{claim}
-If $\ell : S \xLeftrightarrow{\delta_S, \partial (S, V), \partial V}
-V$ and $f: S \longrightarrow V$, then $\default(\ell, f) : S
-\xLeftrightarrow{\delta_S, \partial (S, V), \partial V} V$.
-\end{claim}
-\begin{proof}
-Follows from the fact that $\ell : S \xLeftrightarrow{\delta_S, \partial (S, V),
-\partial V} V$.
-\end{proof}
-\subsection{Product}
-Let $\ell_1 : S_1 \Leftrightarrow V_1$ and $\ell_2 : S_2 \Leftrightarrow V_2$ be
-lenses. Define the lens $\ell_1 \times \ell_2 : (S_1 \times S_2) \Leftrightarrow
-(V_1 \times V_2)$ by
-\begin{align*}
-(\ell_1 \times \ell_2).\get \; (s_1, s_2) &= (\ell_1.\get \; s_1, \ell_2.\get
-\; s_1)\\
-(\ell_1 \times \ell_2).\pput \; (s_1, s_2) \; (v_1, v_2) &= (\ell_1.\pput \; s_1
-\; v_1, \ell_2.\pput \; s_2 \; v_2)\\
-(\ell_1 \times \ell_2).\create \; (v_1, v_2) &= (\ell_1.\create \; v_1,
-\ell_2.\create \; v_2)
-\end{align*}
-\begin{claim}\label{productislens}
-If $\ell_1 :S_1 \xLeftrightarrow{\delta_{S_1}, \partial (S_1, V_1),
-\delta_{V_1}} V_1$ and $\ell_2 : S_2 \xLeftrightarrow{\delta_{S_2}, \partial
-(S_2, V_2), \delta_{V_2}} V_2$, then \\
-$\ell_1 \times \ell_2  : S_1 \times S_2 
-\xLeftrightarrow{\delta_{S_1} \times \delta_{S_2}, \partial (S_1, V_1) \times
-\partial (S_2, V_2), \delta_{V_1} \times \delta_{V_2}} V_1 \times V_2$.
-\end{claim}
-\begin{proof}
-We need to show that
-$$(\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \; (\delta((\ell_1 \times
-\ell_2).\pput \; (s_1, s_2) \; (v_1, v_2), (s_1, s_2))) = \delta((v_1, v_2),
-(\ell_1 \times \ell_2).\get \; (s_1, s_2))$$ which follows from the fact that
-\begin{align*}
-&(\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \; (\delta((\ell_1 \times
-\ell_2).\pput \; (s_1, s_2) \; (v_1, v_2), (s_1, s_2)))\\
-&= (\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \;
-(\delta((\ell_1.\pput \; s_1 \; v_1, \ell_2.\pput \; s_2 \; v_2), (s_1, s_2))\\
-&= (\partial (S_1, V_1) \times \partial (S_2, V_2)).\get \;
-(\delta(\ell_1.\pput \; s_1 \; v_1, s_1), \delta(\ell_2.\pput \; s_2 \; v_2,
-s_2))\\
-&= ((\partial (S_1, V_1)).\get \; (\delta(\ell_1.\pput \;
-s_1 \; v_1, s_1)), (\partial (S_2, V_2)).\get \; (\delta(\ell_2.\pput \; s_2 \;
-v_2, s_2)))\\
-&= (\delta(v_1, \ell_1.\get \; s_1), \delta(v_2, \ell_2.\get \; s_2))\\
-&= \delta((v_1, v_2), (\ell_1.\get \; s_1, \ell_2.\get \; s_2))\\
-&= \delta((v_1, v_2), (\ell_1 \times \ell_2).\get \; (s_1, s_2))
-\end{align*}
-\end{proof}
-\subsection{Sum}
-Let $\ell_1 : S_1 \Leftrightarrow V_1$ and $\ell_2 : S_2 \Leftrightarrow V_2$ be
-lenses. Define the lens
-$\ell_1 \sqcup \ell_2 :
-(S_1 \sqcup S_2) \Leftrightarrow (V_1 \sqcup V_2)$ by
-\begin{align*}
-(\ell_1 \sqcup \ell_2).\get \; s &=
-\begin{cases}
-\ell_1.\get \; s & \text{if } s \in S_1\\
-\ell_2.\get \; s & \text{if } s \in S_2\\
-\end{cases}\\
-(\ell_1 \sqcup \ell_2).\pput \; s \; v&=
-\begin{cases}
-\ell_1.\pput \; s \; v& \text{if } s \in S_1, \; v \in V_1\\
-\ell_2.\pput \; s \; v& \text{if } s \in S_2, \; v \in V_2\\
-\ell_1.\create \; v & \text{if } s \in S_2, \; v \in V_1\\
-\ell_2.\create \; v & \text{if } s \in S_1, \; v \in V_2\\
-\end{cases}\\
-(\ell_1 \sqcup \ell_2).\create \; v &=
-\begin{cases}
-\ell_1.\create \; v & \text{if } v \in V_1\\
-\ell_2.\create \; v & \text{if } v \in V_2\\
-\end{cases}
-\end{align*}
-
-\begin{claim}
-If $\ell_1 : S_1  \xLeftrightarrow{\delta_{S_1}, \partial (S_1, V_1),
-\delta_{V_1}} V_1$ and $\ell_2 : S_2 \xLeftrightarrow{\delta_{S_2}, \partial
-(S_2, V_2), \delta_{V_2}} V_2$, then \\
-$\ell_1 \sqcup \ell_2 : S_1 \sqcup S_2 \xLeftrightarrow{\delta_{S_1} \sqcup
-\delta_{S_2}, \partial (S_1, V_1) \sqcup \partial (S_2, V_2) \sqcup (\ell_1
-\times \ell_2), \delta_{V_1} \sqcup \delta_{V_2}} V_1 \sqcup V_2$
-\end{claim}
-\begin{proof}
-We need to show that
-$$(\partial (S_1, V_1) \sqcup \partial (S_2, V_2)).\get \;
-(\delta((\ell_1 \sqcup \ell_2).\pput \; s \; v, s)) = \delta(v, (\ell_1 \sqcup
-\ell_2).\get \; s)$$
-
-The result holds inductively if $s \in S_1$ and $v \in V_1$ or $s \in S_2$ and
-$v \in V_2$. Otherwise, assume without loss of generality
-that $s \in S_1$ and $v \in V_2$. Then
-\begin{align*}
-(\partial (S_1, V_1) \sqcup \partial (S_2, V_2)).\get \;
-(\delta((\ell_1 \sqcup \ell_2).\pput \; s \; v, s)) &=
-(\partial (S_1, V_1) \sqcup \partial (S_2, V_2)).\get \; (\delta(\ell_2.\create
-\; v, s))\\
-&= (\ell_1 \times \ell_2).\get \; (s, \ell_2.\create \; v)\\
-&= (\ell_1.\get \; s, \ell_2.\get \; (\ell_2.\create \; v))\\
-&= (\ell_1.\get \; s, v)\\
-&= \delta(v, \ell_1.\get \; s)\\
-&= \delta(v, (\ell_1 \sqcup \ell_2).\get \; s)
-\end{align*}
-\end{proof}
-\subsection{Positional Iteration}
-Let $\ell : S \Leftrightarrow V$ be a lens and define the lens $\ell^*:
-S^* \Leftrightarrow V^*$ by
-\begin{align*}
-\ell^* .\get \; (s_1, \ldots, s_n) &= (\ell.\get \; s_1) \cdot \ldots
-\cdot (\ell.\get \; s_n)\\
-\ell^*.\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n) &= s'_1 \cdot
-\ldots \cdot s'_n\\
-\text{ where } s'_i &= \begin{cases}
-\ell.\pput \; s_i \; v_i & \text{if } 1 \leq i \leq \min\{m, n\}\\
-\ell.\create \; v_i & \text{if } \min\{m, n\} < i \leq n
-\end{cases}\\
-\ell^*.\create \; (v_1, \ldots, v_n) &= (\ell.\create \; v_1) \cdot \ldots
-\cdot (\ell.\create \; v_n)
-\end{align*}
-
-\begin{claim}\label{positionalisvalid}
-If $\ell: S \xLeftrightarrow{\delta_S, \partial (S, V), \delta_V}
-V$, then $\ell^*: S^* \xLeftrightarrow{\delta_{S}^*, (\partial (S, V) \sqcup
-\ell)^*, \delta_{V}^*} V^*$.
-\end{claim}
-\begin{proof}
-We need to show that
-\begin{multline}
-(\partial (S, V) \sqcup \ell)^*.\get \;
-(\delta_S^*(\ell^*.\pput \; (s_1, \ldots, s_n) \; (v_1, \ldots,
-v_m), (s_1, \ldots, s_n))) \\
-= \delta_V^*((v_1, \ldots, v_m), \ell^*.\get \; (s_1, \ldots,
-s_n))
-\end{multline}
-Let $1 \leq i \leq \min\{m, n\}$. Observe that $v_i = \ell.\get \; s_i$ if and
-only if $\ell.\pput \; s_i \; v_i = s_i$, so $\delta_V^*$ omits
-$\delta_V(v_i, \ell.\get \; s_i)$ in position $i$ if and only if
-$\delta_S^*$ also omits $\delta_S(\ell.\pput \; s_i \; v_i, s_i)$ in
-position $i$; if no omission occurs, then $\delta(S, V).\get \;
-(\delta(\ell.\pput \; s_i \; v_i, s_i)) = \delta(v_i, \ell.\get \; s_i)$ since
-$(\ell, \partial(S,V))$ is a delta lens. For $\min\{m,n\} < i \leq m$, then
-$\delta_V^*$ adds $v_i$ in position $i$, and $\delta_S^*$ adds
-$\ell.\create \; v_i$ in position $i$, which gives the desired result since
-$$(\partial (S, V) \sqcup \ell).\get \; (\ell.\create \; v_i) = \ell.\get \;
-(\ell.\create \; v_i)) = v_i$$
-\end{proof}
-\subsection{Matching Iteration}
-\iffalse
-When defining a $\match(\mathcal{M}, \delta)$ difference function, we often wish for
-$\mathcal{M}$ to be an optimal set of matchings for some alignment strategy.
-
-For instance, define the set of matchings $\mathcal{M}_{POS}$ by
-$$\mathcal{M}_{POS} = \{\varphi_{(a_1, \ldots, a_m), (b_1, \ldots, b_n)}:[m]
-\longrightarrow [n] \; | \; m \leq n, \varphi_{(a_1, \ldots, a_m), (b_1,
-\ldots, b_n)}(i) = i\}$$ Then $\mathcal{M}_{POS}$ is an optimal set of
-matchings for a positional alignment strategy.
-
-For the non-crossing strategy, we want to our matchings to be increasing rather
-than positional. That is, given a valid set of matchings $\mathcal{M}$, let
-$\mathcal{M}_{a, b} = \varphi_{a, b} \in \mathcal{M}$, and for $a, a' \in P$
-with $a=(a_1, \ldots, a_m)$ and $a'=(a'_1, \ldots, a'_n)$, define
-$\mathcal{M}^{NC}_{a, b}$ by
-$$\mathcal{M}^{NC}_{a, a'} = \argmin_{\substack{\mathcal{M}_{a, a'} \\
-\mathcal{M}_{a, a'} \text{ is increasing}}} |\match(\mathcal{M}_{a, a'},
-\delta)(a, a')|$$
-There may be multiple matchings which for which the size of the delta is
-minimal, in which case we pick the lexicographically smallest matching.
-Then $\mathcal{M}^{NC}$ is an optimal set of matchings for a non-crossing
-alignment strategy.
-
-Finally, for a min-edit distance strategy, we simply remove the condition that
-the matchings should be increasing. That is, define
-$\mathcal{M}^{MED}_{a, b}$ by
-$$\mathcal{M}^{MED}_{a, a'} = \argmin_{\mathcal{M}_{a, a'}}
-|\match(\mathcal{M}_{a, a'}, \delta)(a, a')|$$
-Then $\mathcal{M}_{MED}$ is an optimal set of matchings for a min-edit distance
-alignment strategy.
-\fi
-Assume that $\ell: S \Leftrightarrow V$. Let $\mathcal{M}_V = \{\varphi_{x, y}
-\; | \; x, y \in V^*\}$ be a set of alignments between $x$ and $y$ for each $x,
-y \in V^*$. For each $s, s' \in S^*$, let $\varphi_{s, s'} = \varphi_{\ell.\get
-\; s, \ell.\get \; s'}$. Define $\mathcal{M}_S$ by $\mathcal{M}_S =
-\{\varphi_{s, s'} \; | \; s, s' \in S^*\}$. Define the lens
-$\match(\mathcal{M}_V, \ell) : S^* \Leftrightarrow V^*$ by
-\begin{align*}
-\match(\mathcal{M}_V, \ell) .\get \; (s_1, \ldots, s_n) &= (\ell.\get \; s_1)
-\cdot \ldots \cdot (\ell.\get \; s_n)\\
-\match(\mathcal{M}_V, \ell) .\pput \; (s_1, \ldots, s_m) \; (v_1, \ldots, v_n)
-&= (s'_1 , \ldots , s'_n)\\
-\text{ where } s'_i &= \begin{cases}
-\ell.\pput \; s_{\psi(i)} \; v_i & \text{if } i \in Dom(\psi)\\
-\ell.\create \; v_i & \text{if } i \not \in Dom(\psi)
-\end{cases}\\
-\text{ and } \psi &= \varphi_{(v_1, \ldots, v_n), ((\ell.\get \; s_1), \ldots,
-(\ell.\get \; s_m))}\\
-\match(\mathcal{M}_V, \ell).\create \; (v_1, \ldots, v_n) &= (\ell.\create \;
-v_1) \cdot \ldots \cdot (\ell.\create \; v_n)
-\end{align*}
-Assume that 
-\begin{claim}
-If $\ell: S \xLeftrightarrow{\delta_{S}, \partial (S, V), \delta_{V}}
-V$, then $\match(\mathcal{M}_V, \ell) : S^*
-\xLeftrightarrow{\match(\mathcal{M}_S, \delta_{S}), \partial (S, V)^*,
-\match(\mathcal{M}_V, \delta_V)} V^*$.
-\end{claim}
-\begin{proof}
-Observe that the matching used to compute the delta on the source side is
-the same as the matching used to compute the matching on the view side, hence we
-may take the update lens to be position and proceed with the rest of the proof
-as in \cref{positionalisvalid}.
-\end{proof}
-\subsection{Composition}
-Let $\ell_1 : S \Leftrightarrow U$ and $\ell_2 : U \Leftrightarrow V$ be lenses.
-Define the lens $\ell_2 \circ \ell_1 : S \Leftrightarrow V$ by
-\begin{align*}
-(\ell_2 \circ \ell_1).\get \; s &= \ell_2.\get \; (\ell_1.\get \; s)\\
-(\ell_2 \circ \ell_1).\pput \; s \; v &= \ell_1.\pput \; s \; (\ell_2.\pput \;
-(\ell_1.\get \; s) \; v) \\
-(\ell_2 \circ \ell_1).\create \; v &= \ell_1.\create \; (\ell_2.\create \; v)
-\end{align*}
-
-\begin{claim}
-If $\ell_1 : S_1  \xLeftrightarrow{\delta_{S_1}, \partial (S_1, V_1),
-\delta_{V_1}} V_1$ and $\ell_2 : S_2 \xLeftrightarrow{\delta_{S_2}, \partial
-(S_2, V_2), \delta_{V_2}} V_2$, then \\
-$\ell_2 \circ \ell_1 :  S \xLeftrightarrow{\delta_S, \partial
-(S_2, V_2) \circ \partial (S_1, V_1), \delta_V} V$.
-\end{claim}
-\begin{proof}
-We need to show that
-$$
-(\partial (S_2, V_2) \circ \partial (S_1, V_1)).\get \; (\delta(\ell_2 \circ
-\ell_1).\pput \; s \; v, s) = \delta(v, (\ell_2 \circ \ell_1).\get \; s)$$
-
-This follows from the fact that
-\begin{align*}
-&(\partial (S_2, V_2) \circ \partial (S_1, V_1)).\get \; (\delta_S((\ell_2 \circ
-\ell_1).\pput \; s \; v, s))\\
-&=(\partial (S_2, V_2) \circ \partial (S_1, V_1)).\get \; (\delta_S(\ell_1.\pput
-\; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; v), s))\\
-&= (\partial (S_2, V_2)).\get \; ((\partial (S_1, V_1)).\get \;
-(\delta_S(\ell_1.\pput \; s \; (\ell_2.\pput \; (\ell_1.\get \; s) \; v), s)))\\
-&= (\partial (S_2, V_2)).\get \; (\delta_U(\ell_2.\pput \; (\ell_1.\get \; s)
-\; v, \ell_1.\get \; s))\\
-&= \delta_V(v, \ell_2.\get \; (\ell_1.\get \; s))\\
-&= \delta_V(v, (\ell_2 \circ \ell_1).\get \; s)
-\end{align*}
-\end{proof}
-\iffalse
-\section{Symmetric Delta Lenses}
-Assume that $(X, \partial X, \delta_X)$ and $(Y, \partial Y, \delta_Y)$
-are difference structures. A symmetric delta lens is a pair $(\ell, \partial
-\ell)$ where $\ell : X \Leftrightarrow Y$ is a symmetric lens, and
-$\partial \ell : \partial X \Leftrightarrow \partial Y$ is a symmetric lens
-such that
-\begin{enumerate}
-  \item
-  if $\ell.putr(x, c) = (y, c)$ and $\ell.putl(y', c) = (x', c')$,
-  then $\partial \ell.putl \; (\delta(y, y'), \delta(c, c')) = (\delta(x,
-  x'), \delta(c, c'))$
-  \item
-  if $\ell.putl(y, c) = (x, c)$ and $\ell.putr(x', c) = (y', c')$, then
-  $\partial \ell.putr \; (\delta(x, x'), \delta(c, c')) = (\delta(y, y'),
-  \delta(c, c'))$
-\end{enumerate}
-\begin{fact}[Identity Lens]
-Given any symmetric lens $\partial X : X \Leftrightarrow X$, $(id_X, \partial
-X)$ is a symmetric delta lens
-\end{fact}
-\begin{fact}[Bijective Lenses]
-Assume that $f : X \longrightarrow Y$ and $g : \partial X \longrightarrow
-\partial Y$ are bijections. Assume that $g(\delta(x, x')) = \delta(f(x),
-f(x'))$. Then $(bij_f, bij_g)$ is a symmetric delta lens.
-\end{fact}
-\begin{fact}[Lens Composition]
-If $(\ell_1, \partial \ell_1) : (X, \partial X, \delta_X) \Leftrightarrow (Y,
-\partial Y, \delta_Y)$ and $(\ell_2, \partial \ell_2): (Y, \partial Y,
-\delta_Y) \Leftrightarrow (Z, \partial Z, \delta_Z)$ are symmetric delta
-lenses, then $(\ell_1;\ell_2, \partial \ell_1;\partial \ell_2)$ is a symmetric
-delta lens.
-\end{fact}
-\begin{fact}[Dual of a Lens]
-If $(\ell, \partial \ell)$ is symmetric delta lens, then $(\ell^{op}, \partial
-\ell^{op})$ is a symmetric delta lens.
-\end{fact}
-\begin{fact}[Terminal Lens]
-$(term_x, term_{\delta(x, x)})$ is a symmetric delta lens
-\end{fact}
-\begin{corollary}[Disconnect Lens]
-If $x \in X$ and $y \in Y$, then $(term_x;term_y^{op},
-term_{\delta(x, x)};term_{\delta(y, y)}^{op})$ is a symmetric delta lens.
-\end{corollary}
-\begin{fact}[Tensor Product Lens]
-If $(\ell_1, \partial \ell_1)$ and $(\ell_2, \partial \ell_2)$ are symmetric
-delta lenses, then $(\ell_1 \otimes \ell_2, \partial \ell_1 \otimes \partial
-\ell_2)$ is a symmetric delta lens.
-\end{fact}
-\begin{corollary}[Projection Lens]
-Projection lenses are symmetric delta lenses.
-\end{corollary}
-\begin{fact}[Forgetful Tensor Sum Lens]
-If $(\ell_1, \partial \ell_1)$ and $(\ell_2, \partial \ell_2)$ are symmetric
-delta lenses, then $(\ell_1 \oplus^f \ell_2, \partial \ell_1 \oplus^f \partial
-\ell_2 \oplus (\ell_1 \otimes \ell_2))$ is a symmetric delta lens.
-\end{fact}
-\begin{fact}[Forgetful List Mapping Lens]
-If $(\ell, \partial \ell)$ is a symmetric delta lens, then $(map^f(\ell)
-\ell_2, map^f(\partial \ell))$ is a symmetric delta lens.
-\end{fact}
-\begin{remark}[Retentive Lenses]
-The retentive tensor sum and list mapping lenses are in general not symmetric
-delta lenses with the obvious definitions of difference functions.
-\end{remark}
-\fi
-\section{Relational Lenses}
-We define a {\em relation} $R$ of type $((T_1, \ldots, T_n), P, F)$ to be a
-multiset $R$ where $T_i, P$ are sets satisfying $R \subseteq P \subseteq T_1
-\times \ldots \times T_n$, $F \subseteq [n] \times [n]$, and if $i, j \in F$,
-then for all $r, r' \in R$, if $r_i = r'_i$ then $r_j = r'_j$. We interpret the
-$T_i$ to be a set of {\em attributes}, $R$ to be a set of {\em records} with
-$dom(r) = (T_1, \ldots, T_n)$ for all $r \in R$, $P$ to be a {\em predicate}
-such that the records in $R$ satisfy $P$, and $F$ to be a set of {\em
-functional dependencies} on the attributes of the records in $R$.
-
-Let $G(F)$ be the graph induced by $F$ on $T_1, \ldots, T_n$ i.e. there is an
-edge from $T_i$ to $T_j$ if and only if $(i, j) \in F$. We say that $F$ is
-in {\em tree form} if $G(F)$ is a DAG and every vertex has indegree at most 1;
-following \cite{bohannon2006relational},we will always assume that $F$ is in
-tree form from here on. Observe that every record $r$ is completely determined
-by nodes of indegree 0 in $G(F)$: if $Roots(G(F)) = \{T_{i_1}, \ldots,
-T_{i_k}\}$ is the set of nodes of indegree 0 and outdegree greater than 0 in
-$G(F)$ and $Isolated(G(F)) = \{T_{j_1}, \ldots, T_{j_\ell}\}$ is the set of
-nodes of indegree 0 and outdegree 0 in $G(F)$, then for any $r = (r_1, \ldots,
-r_n)\in R$, then the map $((r_{i_1}, \ldots, r_{i_k}), (r_{j_1}, \ldots,
-r_{j_\ell}))$ is a bijection since the $r_{i_1}, \ldots, r_{i_k}$ determine all
-nodes of indegree greater than 0, which leaves just the nodes $r_{j_1}, \ldots,
-r_{j_\ell}$.
-
-In order to define lenses in relations, we first need to describe data
-structures representing the relation; lenses will operate on the data
-structures rather than on relations. Let $Outputs(F) = G(F) \setminus
-(Roots(G(F)) \cup Isolated(G(F)))$. We choose to represent a relation with the
-following data:
-\begin{enumerate}
-  \item
-  A partial map $data : (T_{i_1} \times \ldots \times T_{i_k})
-  \longrightarrow 2^{(T_{j_1} \times \ldots \times T_{j_{\ell}})}$ such that
-  there exists a record $r \in R$ with $r[T_{i_1}, \ldots, T_{i_k}] =
-  (t_{i_1}, \ldots, t_{i_k})$ if and only if
-  $$data(t_{i_1}, \ldots, t_{i_k}) =  \{\{r[T_{j_1}, \ldots, T_{j_{\ell}}] \in
-  R \; | \; r_{i_m} = t_{i_m} \text{ for }1 \leq m \leq k\}\}$$
-  \item
-  A map $dependencies : outputs(F) \longrightarrow \cup_{i,j} 2^{T_i \times
-  T_j}$ such that $(p, q) \in dependencies(j)$ if and only if there exists a
-  record $r \in R$ such that $r[T_i, T_j] = (p, q)$, where $i$ is such that
-  $T_i \longrightarrow T_j \in F$ ($i$ is unique as $F$ is assumed to be in tree
-  form).
-\end{enumerate}
-\subsection{Select}
-We say that a predicate $P$ ignores a set of attributes $X$ if for all $r, r'
-\in R$, if $r[dom(R) \setminus X] = r'[dom(R) \setminus X]$ implies that $r
-\in P$ if and only if $r' \in P$.
-
-The {\em select} lens on $R$ evaluates to a relation which is the same as $R$
-but which satisfies a predicate $P$. The inferrence rule for deriving the select
-lens is the following:
-\begin{prooftree}
-\AxiomC{$R : ((T_1, \ldots, T_2), Q, F)$}
-\AxiomC{$S : ((T_1, \ldots, T_2), P \cap Q, F)$}
-\AxiomC{$F$ is in tree form}
-\AxiomC{$P$ ignores $Outputs(F)$}
-\QuaternaryInfC{$(select \; from \; R\; where \; P \; as \; S) : R
-\Leftrightarrow S$}
-\end{prooftree}
-In the get direction, the select lens simply restricts $R$ to $P$. Now consider
-$put \; s \; r$. The select lens proceeds as follows:
-\begin{enumerate}
-  \item
-  First, it computes $(data_s, dependencies_s)$ and $(data_r, dependencies_r)$.
-  \item
-  Next, it updates $dependencies_r$ using $dependencies_s$ to get
-  $dependencies_r'$.
-  \item
-  Next it filters $data_r$ to contain only records not in $P$ to get $data_r'$.
-  \item
-  Next, it computes $(data_r' \cup data_s, dependencies_r')$.
-  \item
-  Finally, it deletes the records from $data_r' \cup data_s$ that would cause
-  PUT-GET to fail.
-\end{enumerate}
-\section{Matching Lenses}
-\begin{definition}
-A {\em basic lens} $\ell : S \xLeftrightarrow{C} V$ is a 4-tuple of functions
-$\ell.\get : S \longrightarrow V$, $\ell.\res : S \longrightarrow C$,
-$\ell.\pput : V \longrightarrow C \longrightarrow S$ and $\ell.\create : V
-\longrightarrow S$ satisfying the following properties:
-\begin{align*}
-\ell.\get \; (\ell.\pput \; v \; c) &= v \tag{PUTGET}\\
-\ell.\pput \; (\ell.\get \; s) \; (\ell.\res \; s) &= s \tag{GETPUT}
-\end{align*}
-\end{definition}
-Idea: First remove matched parts to leave unmatched skeleton. Next remove
-unmatched parts to leave matched skeleton and then flatten.
-%% Acknowledgments
-\begin{acks}                            %% acks environment is optional
-%% contents suppressed with 'anonymous'
-%% Commands \grantsponsor{<sponsorID>}{<name>}{<url>} and
-%% \grantnum[<url>]{<sponsorID>}{<number>} should be used to
-%% acknowledge financial support and will be used by metadata
-%% extraction tools.
-This material is based upon work supported by the
-\grantsponsor{GS100000001}{National Science
-Foundation}{http://dx.doi.org/10.13039/100000001} under Grant
-No.~\grantnum{GS100000001}{nnnnnnn} and Grant
-No.~\grantnum{GS100000001}{mmmmmmm}.  Any opinions, findings, and
-conclusions or recommendations expressed in this material are those
-of the author and do not necessarily reflect the views of the
-National Science Foundation.
-\end{acks}
-\fi
-\bibliographystyle{plain}
-\bibliography{local}
-%% Appendix
-%%\appendix
-%%\section{Appendix}
-
-\end{document}

### 2018-06-24 papers/DifferenceLenses/main.tex
Made mandatory changes

@@
 \begin{cases}
-0 & \text{if } m=0 \text{ or } n=0\\
-\max\limits_{\{\varphi : [m] \longrightarrow [n] \; | \; \varphi(i) = i\}}\frac{1}{n}\sum\limits_{i=1}^{m}\ \delta_R(s_i, t_{\varphi(i)})& \text{if } m \leq n\\
+1 & \text{if } m = 0, n = 0\\ 
+0 & \text{if } m = 0, n \neq 0\\
+0 & \text{if } m \neq 0, n = 0\\\max\limits_{\{\varphi : [m] \longrightarrow [n] \; | \; \varphi(i) = i\}}\frac{1}{n}\sum\limits_{i=1}^{m}\ \delta_R(s_i, t_{\varphi(i)})& \text{if } m \leq n\\
 \max\limits_{\{\varphi : [n] \longrightarrow [m] \; | \; \varphi(i) = i\}}\frac{1}{m}\sum\limits_{i=1}^{m}\ \delta_R(t_i, s_{\varphi(i)})& \text{otherwise}
@@
 \begin{cases}
-0 & \text{if } m=0 \text{ or } n=0\\
+1 & \text{if } m = 0, n = 0\\ 
+0 & \text{if } m = 0, n \neq 0\\
+0 & \text{if } m \neq 0, n = 0\\
 \max\limits_{\{\varphi : [m] \longrightarrow [n] \; | \; \varphi \text{ is strictly increasing}\}}\frac{1}{n}\sum\limits_{i=1}^{m}\ \delta_R(s_i, t_{\varphi(i)})& \text{if } m \leq n\\
@@
 \begin{cases}
-0 & \text{if } m=0 \text{ or } n=0\\
-\max\limits_{\{\varphi : [m] \longrightarrow [n] \; | \; \varphi \text{ is injective}\}}\frac{1}{n}\sum\limits_{i=1}^{m}\ \delta_R(s_i, t_{\varphi(i)})& \text{if } m \leq n\\
+1 & \text{if } m = 0, n = 0\\ 
+0 & \text{if } m = 0, n \neq 0\\
+0 & \text{if } m \neq 0, n = 0\\\max\limits_{\{\varphi : [m] \longrightarrow [n] \; | \; \varphi \text{ is injective}\}}\frac{1}{n}\sum\limits_{i=1}^{m}\ \delta_R(s_i, t_{\varphi(i)})& \text{if } m \leq n\\
 \max\limits_{\{\varphi : [n] \longrightarrow [m] \; | \; \varphi \text{ is injective}\}}\frac{1}{m}\sum\limits_{i=1}^{m}\ \delta_R(t_i, s_{\varphi(i)})& \text{otherwise}
@@
 \end{proof}
+\subsection{Matching Lenses}
+Difference lenses provide three iteration combinators for dealing with the issue of {\em local alignment}.
+However, if the ``chunks'' that a programmer wishes to align do not occur underneath a star regular expression, then the difference lenses combinators will not align them.
+Consider for example the lens:
+\begin{lstlisting}
+let k : lens = (copy [A-Z]) . ([a-z]<->"")
+let l : lens = k . (copy "," . k)*
+\end{lstlisting}
+If we try to put the view \lstinline|"Z,Y,W,X"| into the source \lstinline|"Xx,Yy,Zz"|, then using a min-edit distance iteration lens, we will get the result \lstinline|"Zx,Yy,Wz,Xa"|, because the view element $``Z"$, being out outside the star, will be put into $``Xx"$, and the iteration lens will align the chunks remaining in the star.
+What is needed in this example is a lens which performs a ``global alignment'' strategy.
+
+Barbosa's {\em matching lenses} \cite{barbosa2010matching} do just this, by allowing a programmer to mark every chunk that they wish to align anywhere in a lens term.
+The guarantee that matching lenses give is that alignment used to match view chunks is the same alignment that is used to match source chunks.
+
 \iffalse
-\section{Matching Lenses}
-\begin{definition}
-A {\em basic lens} $\ell : S \xLeftrightarrow{C} V$ is a 4-tuple of functions
-$\ell.\get : S \longrightarrow V$, $\ell.\res : S \longrightarrow C$,
-$\ell.\pput : V \longrightarrow C \longrightarrow S$ and $\ell.\create : V
-\longrightarrow S$ satisfying the following properties:
-\begin{align*}
-\ell.\get \; (\ell.\pput \; v \; c) &= v \tag{PUTGET}\\
-\ell.\pput \; (\ell.\get \; s) \; (\ell.\res \; s) &= s \tag{GETPUT}
-\end{align*}
-\end{definition}
-Idea: First remove matched parts to leave unmatched skeleton. Next remove
-unmatched parts to leave matched skeleton and then flatten.
 \section{Relational Lenses}

### 2018-06-24 papers/QuotientLenses/main.tex
Made mandatory changes

@@
 
-Strong unambiguity is necessary in the premises of the regular QRE combinators
-because it ensures that the canonizing function of a QRE is well-defined.
-For example, consider the QRE \lstinline{"a"$^*$ . (collapse "a"$^*$ $\to$ "a")}. The
+These unambiguity constraints can be a little fiddly in practice.
+As a simple example of this, consider writing a regular expression for comma-separated lists of strings.
+Our first impulse might be to write it as,
+\begin{lstlisting}
+CSL = anychar+ . ("," . anychar+)*
+\end{lstlisting}
+(where . is concatenation), but this regular expression is ambiguous, as \lstinline|anychar| is a big union of a of single characters including comma.
+
+While the unambiguity constraints consequently appear to compromise compositionality of QREs, they can usually be circumvented by making a small changes to the offending regular expression: for instance, in the preceding example, then we need to write,
+\begin{lstlisting}
+CSL = anycharexceptcomma+ . ("," . anycharexceptcomma+)*
+\end{lstlisting}
+(assuming \lstinline|anycharexceptcomma| denotes a big union of every single character string except comma).
+
+Unambiguity constraints are necessary because they ensure that the canonizing function of a QRE is well-defined.
+For example, consider the QRE \lstinline|"a"$^*$ . (collapse "a"$^*$ $\to$ "a")|. The
 behavior of this QRE is not well-defined since the string
 \cd{"aaa"} can be canonized to any of \cd{"a"}, \cd{"aa"}, or \cd{"aaa"}
-depending on how \cd{"aaa"} is parsed. We also require that the regular
-combinators applied to the kernels of QREs be unambiguous since the underlying
+depending on how \cd{"aaa"} is parsed. The unambiguity constraints are also applied to the kernels of QREs since the underlying
 bijective string lens of a QRE lens operates on kernels, and bijective string lenses
@@
 from FlashFill~\cite{flashfill}, and benchmarks derived from
-Augeas~\cite{augeas2}.  In these 39 benchmarks, 10 of
+Augeas~\cite{augeas2} (we also experimented using our QRE implementation to synthesize quotient lenses between XML, RDF and and JSON formats using data from data.gov database; the data consisted of census statistics, demographic statistics, wage comparosion data, and crime index data).
+In these 39 benchmarks, 10 of
 the data formats had to be modified to work with the bijectivity
@@
 non-bijective problems in the benchmark suite.  On
-average, \canonizeAndSpecSize{} used~44\% more AST nodes
-than \QRESize{}, requiring an average of~237 more AST nodes. On 
-average, \LensAndSpecSize{} used~193\% more AST nodes than \QRESize{}, requiring an
-average of~1267 more AST nodes. These figures suggest that introducing QREs saves
+average (using a geometric mean), \canonizeAndSpecSize{} used~38.5\% more AST nodes
+than \QRESize{}, requiring an average of~214 more AST nodes. On 
+average, \LensAndSpecSize{} used~180\% more AST nodes than \QRESize{}, requiring an
+average of~998 more AST nodes. These figures suggest that introducing QREs saves
 programmers significant effort compared to both Optician and basic

### 2018-06-24 papers/QuotientLenses/main.tex
Fixed some typos

@@
 relations indicating the essential information (QREs).
-A tantalyzing possibility would be to use the bijective string lens
+A tantalizing possibility would be to use the bijective string lens
 synthesis procedure as a subroutine for a quotient lens synthesis procedure. 
@@
 \noindent
-Intuitivley, $W(Q)$ is the regular expression representing the
+Intuitively, $W(Q)$ is the regular expression representing the
 external format, while $K(Q)$ is the regular expression representing
@@
 mechanism. Figure~\ref{fig:times}(a) shows that \QOpt{} was able to synthesize
-all of the Optician benchmarks at a speed competative with the old version.
+all of the Optician benchmarks at a speed competitive with the old version.
 There is a small amount of additional overhead introduced by QREs in calculating
@@
 $R$; here a flat value of $R$ is a value of $R$ with all left/right tags,
-parantheses and list brackets removed. The bidirectional part of BiFluX involves
+parentheses and list brackets removed. The bidirectional part of BiFluX involves
 transforming source regular expressions to view regular expressions so that
@@
 procedures are complete in the sense that they can disambiguate any ambiguous
-type, but they do show that when these procedures succed the normalized types
+type, but they do show that when these procedures succeed the normalized types
 are unambiguous.
@@
 More concretely, their approach gives the developer an explicit option to
-formulat her understanding of two object models being different. The technical
+formulate her understanding of two object models being different. The technical
 realization is as follows: The developer specifies ``classifying term'', a
@@
 QREs and canonizers aim at identifying or characterizing the elements that
-should be treated equally by a lens, the focus of the work by Hilken et. al. is
+should be treated equally by a lens, the focus of the work by Hilken et al. is
 on classes that represent specific patterns of particular relevance to the
@@
 is significantly more complex.  In particular, because Optician types do not
-have full canonical forms, Anders et. al. used a pseudo-canonical form that
+have full canonical forms, Anders et al. used a pseudo-canonical form that
 captures part of the equivalence relation over types were used. To preserve

### 2018-06-26 papers/DifferenceLenses/main.tex
Edited some text and added some technical material

@@
 As a result, the transformations derived using these languages can lead to unexpected results.
-In this paper we present a formalization for bx transformations that requires a bx to provide guarantees on how it propagates updates.
-We will then describe a specific bx language which we call {\em difference lenses}; this language is based on the {\em basic lenses} defined in the bx programming language {\em Boomerang}, but each difference lens comes with explicit guarantees on update propagation.
-We will then express various lens programming languages as difference lenses, and hence describe non-trivial guarantees on update propagation that these languages provide as well.
+In this paper we present a formalization for bx transformations that requires a bx to provide guarantees on how it propagates updates and use this formalization to describe a bx language of {\em difference lenses} based on the basic lenses defined in the lens programming language Boomerang.
+Each difference lens will be shown to respect a ``difference'' measure that estimates how similar two pieces of data are on both sides of a bx; the fact that difference lenses have this property demonstrates that they  have stronger guarantees than consistency restoration.
+We will then turn our attention to Boomerang's matching lenses, where we will define an appropriate difference measure for data acted on by matching lenses, and show that matching lenses respect this refined measure.
 \end{abstract}
@@
 Each one of these solutions has significant shortcomings.
-While matching lenses successfully address the issue of alignment, they do not address the general issue of correct update propagation.
-Edit lenses and delta lenses all assume that the lens propagates a specific set of updates explicitly, but this is unrealistic since since in general users do not update data using a small predetermined langauge of edits.
-Instead users edit the data in any way, for example using text editors or IDEs, and present the updatad data to the bx wholesale.
+Firstly, while matching lenses successfully address the issue of alignment, they do not address the general issue of correct update propagation.
+Secondly, edit lenses and delta lenses all assume that the lens propagates a specific set of updates explicitly, but this is unrealistic since since in general users do not update data using a small predetermined langauge of edits; instead users edit the data in any way, for example using text editors or IDEs, and present the updatad data to the bx wholesale.
 Finally, least change lenses are in general computationally intractable \cite{buneman2002propagation} and also lack a notion of composition---that is, the composition of two deterministic least change lenses is not always a least change lens\cite{macedo2013composing}.
@@
 $[``Lovelace", \; ``Turing"]$ was updated to the list $[``Turing", \; ``Lovelace"]$.
-The programmer may wish to consider these two lists equivalent or not: if she does, then she should use a min-edit distance ``metric'' and demand that the source side respects this metric.
-On the other hand, if she wishes to use a positional metric to compare these two lists, then she should similarly demand that the source side respects this metric.
+The programmer may wish to consider these two lists equivalent or not: if she does, then she should use a min-edit distance ``metric'' and demand that the source side respects this metric by applying the ``flip'' permutation applying the bx in reverse.
+On the other hand, if she wishes to use a positional metric to compare these two lists, then she should similarly demand that the source side respects this metric and use the identity permuation when applying the bx in reverse.
 In each case, the programmer does not care how the bx restores consistency; she cares only that the bx restores consistency and respects the appropriate metric.
@@
 In this paper, we present a general approach for describing update propapagtion guarantees for a bx without requiring that the bx manually handle updates.
-In particular, we will enhance a standard definition of bx transformations that only takes into account a consistency relation with a ``difference'' relation that describes a relationship between source updates and target updates.
+In particular, we will enhance a standard definition of bx transformations that only takes into account a consistency relation with a ``difference'' relation that describes the relationship between source updates and target updates.
 With this new formalization, a well behaved bx will not only be hippocratic, but will also be required to preserve the difference relation that is part of the bx.
 
-We will then use this formalization to describe a particular language of bx transformations which we call {\em difference lenses}.
-Difference lenses are based on the {\em basic lenses} defined in the lens programming language {\em Boomerang} \cite{foster2009bidirectional}, but come with the guarantee that the lens respects ``differences'' on the source and view side of a lens.
-Intuitively, this means that if one starts with a source, maps it into a view, edits the view and then maps the updated view back into the old source, then the difference between the new and old source is at most the difference between the old and new view.
-Difference lenses will have combinators that deal with local alignment; in particular, difference lenses have positional, non-crossing, and min-edit distance iteration combinators that preserve positional, non-crossing, and min-edit distance differences respectively.
-We will then build on difference lenses to describe a language of {\em matching difference lenses} that deal with global alignment.
+In this paper we present a formalization for bx transformations that requires a bx to provide guarantees on how it propagates updates and use this formalization to describe a bx language of {\em difference lenses} based on the basic lenses defined in the lens programming language Boomerang.
+Each difference lens will be shown to respect a ``difference'' measure that measures how similar two pieces of data are on both sides of a bx.
+Intuitively, this means that if one starts with some source data, maps it into a target format using a difference lens, edits the target data, and then maps the updated data back into the old source, then the difference between the new and old source is at most the difference between the old and new targets.
+Difference lenses will have combinators that deal with {\em local alignment}; in particular, difference lenses will have positional, non-crossing, and min-edit distance iteration combinators that preserve positional, non-crossing, and min-edit distance differences respectively.
 
-\sam{TODO: Talk about relational lenses.}
+We will then turn our attention to Boomerang's matching lenses, where we will define an appropriate difference measure for data acted on by matching lenses, and show that matching lenses respect this new measure.
+Matching lenses perform {\em global alignment} of data; for example, they can align data inside of an starred regular expression to data outside of a starred regular expression.
+The measure that we use for matching lenses will thus seperate the difference between aligned chunks and non-aligned chunks and compute the average of the two to get a final result.
 
@@
 \item
-We describe a specific bx language which we call {\em difference lenses}; this language is based on the {\em basic lenses} defined in the lens programming language {\em Boomerang}\cite{foster2009bidirectional}, but each difference lens comes with explicit guarantees on update propagation.
-As a result of this work, we will describe non-trivial update propagation properties that Boomerang's basic lenses provide.
+We describe a bx language of {\em difference lenses} based on the basic lenses defined in the lens programming language Boomerang, and show that each difference lens respects a ``difference'' measure that estimates how similar two pieces of data are on both sides of a bx.
   \item
-  We show how to represent matching lenses and  relational lenses \cite{bohannon2006relational} as difference lenses, and hence describe non-trivial guarantees that these lens languages provide as well.
+We define an appropriate difference measure for data acted on by {\em matching lenses} \cite{barbosa2010matching}, and show that matching lenses respect this refined measure.
 \end{enumerate}
@@
 $\delta_R(s, s) = 1$.
-\item
-If $s \neq s'$ and either $s = \varepsilon$ or $s' = \varepsilon$, then $\delta_R(s, s') = 0$
 \item
@@
 Assume that $\mathcal{L}(R) \cap \mathcal{L}(S) = \varnothing$, and that $(R \cup S)^{!*}$.
-Define the lens $\filter_{MED} \; R \; S : (R \sqcup S)^*_{MED} \Leftrightarrow R^*_{MED}$ by
+Define the lens $\filter_{MED} \; R \; S : (R \cup S)^*_{MED} \Leftrightarrow R^*_{MED}$ by
 \begin{align*}
@@
 Let $R$ and $S$ be difference regular expressions with $\mathcal{L}(R) \cap \mathcal{L}(S) = \varnothing$ and $(R \cup S)^{!*}$.
-Then $\filter_{POS} \; R \; S : (R \sqcup S)^*_{MED} \Leftrightarrow R^*_{MED}$ is a difference lens.
+Then $\filter_{POS} \; R \; S : (R \cup S)^*_{MED} \Leftrightarrow R^*_{MED}$ is a difference lens.
 \end{claim}
@@
 which follows from the fact that
+$$\delta_{R \cdot R}(t \cdot t, s \cdot s) = \frac{1}{2}\left(\delta_R(t, s) + \delta_R(t, s)\right) = \delta_R(t, s)$$
 If $s \neq s'$, then we need to show that
-$$\delta_{R \cdot R}(t \cdot t, s \cdot s) = \frac{1}{2}\left(\delta_R(t, s) + \delta_R(t, s)\right) = \delta_R(t, s)$$
 $$\delta_{R \cdot R}(t \cdot s', s \cdot s') = \frac{1}{2}\left(\delta_R(t, s) + \delta_R(s', s')\right) = \frac{1}{2}\left(\delta_R(t, s) + 1\right) \geq \delta_R(t, s)$$
@@
 Barbosa's {\em matching lenses} \cite{barbosa2010matching} do just this, by allowing a programmer to mark every chunk that they wish to align anywhere in a lens term.
-The guarantee that matching lenses give is that alignment used to match view chunks is the same alignment that is used to match source chunks.
+For example, to align all chunks handled using the lens \lstinline|k| in the following example, then we would wrap each occurrence of \lstinline|k| with the ``chunk'' annotation \lstinline|<>| and specify which alignment to use i.e.
+positional, non-crossing or min-edit distance, which leads to the following lens:
+\begin{lstlisting}
+let k : lens = (copy [A-Z]) . ([a-z]<->"")
+let l : lens = <best:k> . (copy "," . <best:k>)*
+\end{lstlisting}
+The \pput behaviour of this new lens is,
+\begin{lstlisting}
+l.put "Z,Y,W,X" into "Xx,Yy,Zz" = "Zz,Yy,Wa,XX"
+\end{lstlisting}
+Because matching lenses perform a ``global'' alignment of chunks i.e.
+an alignment which may match chunks in different positions of a regular expression, then the difference $\delta_R$ is not an appropriate measure to compare the difference between two sources or views.
+Instead, given a {\em chunk-annotated} regular expression, we will pull out all strings which match the annotated part of the regular expression and compute the difference of those separately from the parts of the regular expression that are not chunk annotated, and then take the average of the two.
+\begin{definition}
+Let $\Sigma$ be a finite alphabet. A {\em regular expression} is any term generated by the following grammar,
+$$R := \varnothing \; | \; s \; | \; R \cdot R' \; | \; R \cup R' \; | \; R^*$$
+where $R$ ranges over regular expressions and $s$ ranges over $\Sigma^*$.
+\end{definition}
+\begin{definition}
+A {\em chunk-annotated regular expression} is any term generated by the following grammar,
+$$A := R \; | \; <R,t> \; | \; A \cdot A' \; | \; A \cup A' \; | \; A^*$$
+where $A$ ranges over chunk-annotated regular expressions $R$ ranges over regular expressions, and $t$ ranges over $\{POS,NC,MED\}$
+\end{definition}
+Let $A$ be an unambiguous chunk-annotated regular expression built from the alphabet $\Sigma$. Let $* \not \in \Sigma$ and $s \in \mathcal{L}(A)$. Define $S(A, s)$ and $L(A, s)$ inductively by
+\begin{align*}
+S(R, s) &= s\\
+S(<R,t>, <s>) &= *\\
+S(A \cdot A', s \cdot s') &= S(A, s) \cdot S(A', s')\\
+S(A \cup A', s) &=
+\begin{cases}
+S(A, s) & \text{if } s \in \mathcal{L}(A)\\
+S(A', s) & \text{if } s \in \mathcal{L}(A')
+\end{cases}\\
+S(A^*, s_1 \cdot \ldots \cdot s_n) &= S(A, s_1) \cdot \ldots \cdot T(A, s_n)\\
+\\
+L(R, s) &= []\\
+L(<R,t>, <s>) &= [s]\\
+L(A \cdot A', s \cdot s') &=  L(A,s) +\!\!+ L(A',s')\\
+L(A \cup A', s) &=
+\begin{cases}
+L(A, s) & \text{if } s \in \mathcal{L}(A)\\
+L(A', s) & \text{if } s \in \mathcal{L}(A')
+\end{cases}\\
+L(A^*, s_1 \cdot \ldots \cdot s_n) &= L(A, s_1) +\!\!+ \ldots +\!\!+ L(A, s_n)
+\end{align*}
+The string $S(A, s)$ is the ``skeleton'' of $s$ that matches the annotated regular expression $A$, while the list $L(A, s)$ is the list of strings inside $s$ that have been marked as chunks for alignment.
 
+Define the function $pick : List(A \; option) \longrightarrow A \; option$ by
+$$
+pick \; [a_1, \ldots, a_n] =
+\begin{cases}
+Some \; a & \text{if } i > 0 \text{ and }a_i \neq None \text{ implies } a_i = a\\
+None &\text{otherwise}
+\end{cases}
+$$
+Given a chunk-annotated regular expression $A$, define $M(A)$ and $N(A)$ as follows:
+\begin{align*}
+M(R) &= R\\
+M(<R,t>) &= *\\
+M(A \cdot A') &= M(A) \cdot M(A')\\
+M(A \cup A') &= M(A) \cup M(A')\\
+M(A^*) &= M(A)^*\\
+\\
+N(R) &= None\\
+N(<R, t>) &= Some \; (R, t)\\
+N(A \cdot A') &= pick \; [N(A),N(A')]\\
+N(A \cup A') &= pick \; [N(A),N(A')]\\
+N(A^*) &= pick [M(A)]
+\end{align*}
+$M(A)$ is the skeletal part of a chunk-annotated regular expresion, whereas $N(A)$ is the regular expression inside $A$ that actually matches chunks to be aligned.
+
+For any unambiguous regular expression $R$, let $List(R)$ be the set of all lists with elements in $\mathcal{L}(R)$, and observe that we can define three difference functions $\delta^{List(R)}_{POS}$ $\delta^{List(R)}_{NC}$ and $\delta^{List(R)}_{MED}$ on $List(R)$ in the same way that we define $\delta_{R^*_{POS}}$, $\delta_{R^*_{NC}}$ and $\delta_{R^*_{MED}}$.
+For example, define
+$$
+\delta^{List(R)}_{MED}([s_1, \ldots, s_m], [t_1, \ldots, t_n]) = 
+\begin{cases}
+1 & \text{if } m = 0, n = 0\\ 
+0 & \text{if } m = 0, n \neq 0\\
+0 & \text{if } m \neq 0, n = 0\\\max\limits_{\{\varphi : [m] \longrightarrow [n] \; | \; \varphi \text{ is injective}\}}\frac{1}{n}\sum\limits_{i=1}^{m}\ \delta_R(s_i, t_{\varphi(i)})& \text{if } m \leq n\\
+\max\limits_{\{\varphi : [n] \longrightarrow [m] \; | \; \varphi \text{ is injective}\}}\frac{1}{m}\sum\limits_{i=1}^{m}\ \delta_R(t_i, s_{\varphi(i)})& \text{otherwise}
+\end{cases}
+$$
+
+\begin{definition}
+If $A$ is a chunk-annotated unambiguous regular expression with $s,s' \in \mathcal{L}(A)$ and $N(A) = (R, t)$, define $\delta_A(s, s')$ by
+$$
+\delta_A(s, s) = 
+\frac{1}{2} \left(\delta_{M(V)}(S(A,s), S(A,s')) + \delta^{List(R)}_{t}(L(A, s), L(A, s'))\right)
+$$
+\end{definition}
+Having defined a way of measuring the difference between two chunk-annotated strings, we now give a brief summary of the design of a simplified version of matching lenses, and explain how a matching lens $\ell$ is used to define a basic lens $\lfloor \ell \rfloor$ that performs alignment of chunks in a chunk-annotated string.
+We will then prove that $\lfloor \ell \rfloor$ is a difference lens with respect to the measure $\delta_A$ that we defined for chunk-annotated regular expressions.
+\begin{definition}
+Given a basic lens $k : S_k \Leftrightarrow V_k$, sets $S$, $V$ that have chunks in $S_k$ and $V_k$ respectively, and a set $C$ of ``complements'', a matching lens $\ell : S \xLeftrightarrow{C, k} V$ comprises four functions:
+\begin{align*}
+\ell.\get &: S \longrightarrow V\\
+\ell.\res &: S \longrightarrow (C \times (\mathbb{N} \rightharpoonup S_k))\\
+\ell.\pput &: V \longrightarrow (C \times (\mathbb{N} \rightharpoonup S_k)) \longrightarrow S\\
+\ell.\create &: V \longrightarrow (\mathbb{N} \rightharpoonup S_k) \longrightarrow S
+\end{align*}
+that satisfy the following laws,
+\begin{align*}
+\ell.\get \; (\ell.\pput \; v \; 
+(c,r)) &= v \tag{PUTGET}\\
+\ell.\pput \; (\ell.\get \; s) \; (\ell.\res \; s) &= s \tag{GETPUT}\\
+\text{if } n \in (locations(v) \cap dom(r)) &\text{ then } (\ell.\pput \; v \; (c, r))[n] = k.\pput \; v[n] \; (r(n)) \tag{CHUNKPUT}\\
+\text{if } n \not \in (locations(v) \cap dom(r)) &\text{ then } (\ell.\pput \; v \; (c, r))[n] = k.\create \; v[n] \tag{NOCHUNKPUT}
+\end{align*}
+where $locations(s)$ is the number of chunks in $s$. (Matching lenses are required to satisfy other laws as well, but these have been ommitted for brevity).
+\end{definition}
+The \pput function of a matching lens takes a complement $c$ and partial map $r$ from natural numbers to chunks of type $S_k$.
+Intuitively, when the lens encounters a the $n^{th}$ chunk inside of a string, it uses the basic lens $k$ by putting the $n^{th}$ chunk of $v$ into the $n^{th}$ chunk in $r$, and if $r$ does not contain any chunk at position $n$ then $r$ uses the \create function of $k$ to.
+Otherwise, the lens uses the complement $c$ in the ``skeletal'' parts of the string i.e.
+the parts of the string that are not marked as chunks.
+
+Now let $\ell : S \xLeftrightarrow{C, k}$ be a matching lens with $k : S_k \Leftrightarrow V_k$.
+Assume that $N(V) = Some \; (R, t)$, and define the basic lens $\lfloor \ell \rfloor : S \Leftrightarrow V$ by
+\begin{align*}
+\lfloor \ell \rfloor.\get \; s &= \ell.\get \; s\\
+\lfloor \ell \rfloor.\pput \; s \; v&= \ell.\pput \; v \; (c, r \circ g)\\
+\text{where } (c, r) &= \ell.\res \;s \\
+\text{and } g &\text{ aligns } L(V, chunks(v)) \text{ to } L(V, chunks(\ell.\get \; s)),\\
+&\text{ maximizing } \delta^{List(R)}_{t}(L(V, chunks(v)), L(V, chunks(\ell.\get \; s)))\\
+\lfloor \ell \rfloor.\create \; v &= \ell.\create \; v \; \varnothing
+\end{align*}
+\begin{claim}
+$\delta_S(\lfloor \ell \rfloor.\pput \; s \;v, s) \geq \delta_V(v, \lfloor \ell \rfloor.\get \; s)$
+\end{claim}
+\begin{proof}
+Let $N(S) = Some \; (R', t)$.
+Since $$\delta_S(\lfloor \ell \rfloor.\pput \; s \;v, s) = \frac{1}{2} \left(\delta_{M(S)}(S(S,\lfloor \ell \rfloor.\pput \; s \;v), S(S,s)) + \delta^{List(R')}_{t}(L(S, \lfloor \ell \rfloor.\pput \; s \;v), L(S, s))\right)$$
+and
+$$\delta_V(v, \lfloor \ell \rfloor.\get \; s) = \frac{1}{2} \left(\delta_{M(V)}(S(V,v), S(V,\lfloor \ell \rfloor.\get \; s)) + \delta^{List(R)}_{t}(L(V, v), L(V, \lfloor \ell \rfloor.\get \; s))\right)$$
+then it is enough to show that
+$$\delta_{M(S)}(S(S,\lfloor \ell \rfloor.\pput \; s \;v), S(S,s)) \geq \delta_{M(V)}(S(V,v), S(V,\lfloor \ell \rfloor.\get \; s))$$
+and that
+$$\delta^{List(R')}_{t}(L(S, \lfloor \ell \rfloor.\pput \; s \;v), L(S, s)) \geq \delta^{List(R)}_{t}(L(V, v), L(V, \lfloor \ell \rfloor.\get \; s))$$
+Both of these follow from the fact that $k$ and the skeleton of $\ell$ are both basic lenses.
+\end{proof}
 \iffalse

### 2018-06-27 papers/SymmetricLenses/main.tex
a first attempt at a new intro. I will continue to review it tonight and then release it

@@
 \begin{abstract}
-\saz{This abstract too long and doesn't get to the point quickly enough. We
-  should tighten it up considerably.}\bcp{+1.  But not just tighten ---
-  the whole story doesn't feel very compelling.}
-  Users commonly need to synchronize the data shared between formats, so edits
-  to the data in one format gets propagated to the data in the other format.
-  Manually writing the functions for these data synchronizers is tedious and
-  error prone. Instead of manually writing these synchronizers, bidirectional
-  languages can be used to express \emph{lenses}, groups of synchronization
-  functions with guarantees about round-trip behavior. While these bidirectional
-  languages help make programming these synthesizers less error-prone, they
-  remain difficult to program in. Synthesizing bidirectional languages allows
-  users to use lenses without manually programming them. Current
-  synthesizers work only on the simplest types of lenses, when the two formats
-  describe exactly the same data; this is often not the case.
-
-  Oftentimes, each format may contain information the other lacks. Such
-  synchronization tasks are described by one of the richest classes of lenses,
-  symmetric lenses. However, symmetric lenses in their current formulation
-  require external\bcp{why not ``internal''?} state in the form of a \emph{complement}. This statefulness
-  makes symmetric lenses both difficult to use and difficult to
-  synthesize. \bcp{This paragraph seems too dense.}
-
-  In this work, we reformulate\bcp{More a restriction than a reformulation.}
-  symmetric lenses as \emph{simple symmetric 
-    lenses}. While this formulation is less expressive than symmetric lenses, it
-  expresses many\bcp{weak} symmetric lenses while requiring \emph{no external state}. We
-  provide combinators for simple symmetric lenses on string data, and integrate
-  these combinators into Boomerang\bcp{undefined} while providing full backwards compatibility.
-
-\bcp{This bit feels more detailed / technical than needed for an abstract.}
-  We develop a type-directed synthesis algorithm for generating lenses in this
-  language from regular expressions describing the data formats. A significant
-  technical challenge in this work arises from the fact that our target language
-  permits synchronization between formats that don't have exactly the same data
-  -- data in one format need not be present in the other format. This gives the
-  synthesis algorithm much more freedom, as data can be completely unaligned
-  (edits in one format are never propagated), completely aligned (the formats
-  are in bijective correspondence), or anywhere in between. Under the assumption
-  that users typically want to align as much data as possible, we provide an
-  information theoretic metric for how much data is left unaligned based on the
-  theory of \emph{stochastic regular expressions}. Furthermore, we use this
-  metric to develop a greedy synthesis algorithm that tries to minimize
-  unaligned data. We experimentally demonstrate that our algorithm can
-  efficiently synthesize simple symmetric lenses on a variety of benchmark
-  suites consisting of synchronizing file formats from configuration files,
-  application-specific data storage files, and data cleaning tasks.
+%% \saz{This abstract too long and doesn't get to the point quickly enough. We
+%%   should tighten it up considerably.}\bcp{+1.  But not just tighten ---
+%%   the whole story doesn't feel very compelling.}
+%%   Users commonly need to synchronize the data shared between formats, so edits
+%%   to the data in one format gets propagated to the data in the other format.
+%%   Manually writing the functions for these data synchronizers is tedious and
+%%   error prone. Instead of manually writing these synchronizers, bidirectional
+%%   languages can be used to express \emph{lenses}, groups of synchronization
+%%   functions with guarantees about round-trip behavior. While these bidirectional
+%%   languages help make programming these synthesizers less error-prone, they
+%%   remain difficult to program in. Synthesizing bidirectional languages allows
+%%   users to use lenses without manually programming them. Current
+%%   synthesizers work only on the simplest types of lenses, when the two formats
+%%   describe exactly the same data; this is often not the case.
+
+%%   Oftentimes, each format may contain information the other lacks. Such
+%%   synchronization tasks are described by one of the richest classes of lenses,
+%%   symmetric lenses. However, symmetric lenses in their current formulation
+%%   require external\bcp{why not ``internal''?} state in the form of a \emph{complement}. This statefulness
+%%   makes symmetric lenses both difficult to use and difficult to
+%%   synthesize. \bcp{This paragraph seems too dense.}
+
+%%   In this work, we reformulate\bcp{More a restriction than a reformulation.}
+%%   symmetric lenses as \emph{simple symmetric 
+%%     lenses}. While this formulation is less expressive than symmetric lenses, it
+%%   expresses many\bcp{weak} symmetric lenses while requiring \emph{no external state}. We
+%%   provide combinators for simple symmetric lenses on string data, and integrate
+%%   these combinators into Boomerang\bcp{undefined} while providing full backwards compatibility.
+
+%% \bcp{This bit feels more detailed / technical than needed for an abstract.}
+%%   We develop a type-directed synthesis algorithm for generating lenses in this
+%%   language from regular expressions describing the data formats. A significant
+%%   technical challenge in this work arises from the fact that our target language
+%%   permits synchronization between formats that don't have exactly the same data
+%%   -- data in one format need not be present in the other format. This gives the
+%%   synthesis algorithm much more freedom, as data can be completely unaligned
+%%   (edits in one format are never propagated), completely aligned (the formats
+%%   are in bijective correspondence), or anywhere in between. Under the assumption
+%%   that users typically want to align as much data as possible, we provide an
+%%   information theoretic metric for how much data is left unaligned based on the
+%%   theory of \emph{stochastic regular expressions}. Furthermore, we use this
+%%   metric to develop a greedy synthesis algorithm that tries to minimize
+%%   unaligned data. We experimentally demonstrate that our algorithm can
+%%   efficiently synthesize simple symmetric lenses on a variety of benchmark
+%%   suites consisting of synchronizing file formats from configuration files,
+%%   application-specific data storage files, and data cleaning tasks.
 
@@
 \section{Introduction}
+
+\noindent
+\dpw{XXXXXX New attempt at intro starts here}
+
+In a big-data world, one will often find the same, or similar, data stored in a
+variety of different formats, with each format processed by a different set of tools.
+For instance, electronic calendars come in iCalendar, vCalendar and h-event
+formats~\cite{calendar-formats}; US
+taxation data can be transmitted via Tax XML (used by the US government for e-filing)
+or TXF (used by TurboTax)~\cite{tax-formats}; and sewing machines can read embroidery files transmitted
+as PES formats, DST formats or several others~\cite{embroidery-formats}.
+
+One way to interconvert between these varied data formats is to use a
+\emph{lens}~\cite{?}---a bi-directional
+program that can transform data represented in one format $S$ to data represented
+in a related format $T$,
+and vice versa.  There are a wide variety of different kinds of lenses, but well-defined lens
+languages typically come with guarantees that help ensure data is
+not lost or corrupted as it is transformed back and forth between formats.
+
+Lens languages, like any domain-specific language, also cut down the space of
+possible programs one can write, and the smaller the space of programs,
+the easier it is to find the program one is looking for.  
+%This feature can be a negative, if the desired transformation
+%falls outside of the scope provided by the language, but can also be positive, as complex
+%properties (like invertability) can be enforced automatically.
+%Another advantage of reducing the
+%set of programs is that doing so can facilitate program synthesis, which is essentially
+%a search through the space of possible computations that can be expressed as a lens.
+Recent research~\cite{miltner+:bijective-synthesis,maina+:quotient-synthesis}
+has exploited this phenomenon by demonstrating it is possible to synthesize relatively
+large, complex lenses from specifications.
+More specifically, given types that describe source and target formats, these synthesis
+engines will quickly find bijective lenses~\cite{miltner+:bijective-synthesis} or
+bijections modulo isomorphism~\cite{maina+quotient-synthesis} (\EG, the isomorphism that
+treats whitespace as irrelevant) for medium-sized files, such as those housing linux
+application configurations.  Oftentimes, these algorithms find such lenses in less than
+a second, and without the aid of any examples, just with types describing the given
+formats.
+
+One of reasons these algorithms are so effective, even relative to successful synthesis projects
+in other domains, is because the specifications cut down the space of programs very far
+indeed.  In particular, the restriction to bijectivity is both a blessing and a curse.
+It is a blessing, because given an input format, unless the same kind of data appears in
+both formats in multiple places, there are typically hardly any bijective lenses between
+the data formats:  If the synthesis algorithm can find a bijection then it is very likely
+to be the one the user wants.  On the other hand, the bijectivity restriction is a curse
+because in reality, bijections are very few and very far between.  It is much more common
+for data formats to have some data in common, but also to contain some differences.  For
+instance, two related tables might share three of five columns, but not the other two.
+Synchronizing such tables involves synchronizing three of the columns while managing the
+other two columns separately.  Indeed, in our study of Augeas~\cite{augeus}, a RedHat system that
+uses lenses to edit Linux files, we found that XXXXX of YYYYY lenses were not bijective.
+\dpw{fill in with accurate stats.}
+
+A natural way to expand the scope of such tools is to consider a more
+powerful language of lenses.  In particular, \emph{symmetric
+lenses}~\cite{symmetric-lenses} encompass a broad collection of bidirectional
+transformations and include bijections as well as transformations that
+allow each side of the transaction to retain information not present
+in the other side.  If $\ell$ is a symmetric lense between formats $S$ and $T$, we can
+use it to transform data from $S$ to $T$ or vice versa equally well.  We would typically
+use $\ell$ in the forwards direction (from $S$ to $T$) on a string $s$ when a user updates
+$s$---using $\ell$ in this direction will propagate those updates to $T$.
+However, because data in format $T$ may contain some information not present in format $T$,
+the $S \rightarrow T$ transformation will take
+both a string $s \in S$ and a string $t \in T$ as arguments.
+Inuitively, the lens should extract the data ``shared'' between the two formats from $s$
+as it generates an output $t'$.  If $s$ is missing information, the transformation should
+extract the missing components from $t$.
+Initially, we hypothesized that extending the algorithms
+of Miltner \ETAL~\cite{miltner+:bijective-synthesis} to this richer class of
+transformations would be relatively straightforward:  Simply extend the set
+of combinators and execute a search with similar heuristics.  However, in our early,
+naive attempts to do so, we ran
+into two different kinds of difficulties. \dpw{I got into the weeds a little bit there.
+The goal was to set up the idea that there are ``two parts'' to a symmetric
+lens transformation: the information from $s$ and the information from $t$.  In
+a future paragraph in the intro we are going to explain that the central heuristic
+involves minimizing the information from $t$.}
+
+The first difficulty was pragmatic:  symmetric lenses actually
+usually operate over three structures, the source $S$, the target $T$ and a
+complement $C$~\cite{symmetric-lenses}.
+The latter contains all the information not present in one of $S$ or $T$.  Logistically,
+the complement object must be stored away somewhere and infrastructure must be developed
+to manage it.  
+Perhaps more relevant though, is that compliments complicate specifications, even when
+those specifications are just examples.  The issues is that one cannot
+just give examples of source and target; one must also give guidance on how to manage
+the ``hidden state.''  In order to avoid such complexities, we defined a new
+variant of symmetric lenses, which we call \emph{simple symmetric lenses}.  Our new
+lenses do not require compliments, only need source and target data.
+In order to characterize the expressive power of simple symmetric lenses
+theoretically, we prove simple symmetric lenses
+include all (and only) those symmetric lenses that satisfy
+an intuitive property called \emph{forgetfulness}.  In other words, they are the
+complete set of symmetric lenses that do not require a ``memory'' to retain data
+from past instances of $S$ or $T$ when making a round trip.\dpw{fix that}
+From a pragmatic perspective, we justify our design choices by extending
+the Boomerang implementation with symmetric lenses (they had never before been implemented),
+applying them successfully to a range of real-world applications,
+and demonstrating that they are compatible, not just with basic bijective combinators,
+but with a rich set of advanced features including quotient lenses and matching lenses.
+
+The second difficulty appears to be more fundamental\dpw{better word?}:
+While the number of bijective lenses
+between two related formats is often tiny, the number of symmetric lenses is often
+large.  In practice, this meant that when a search algorithm simply selects the smallest
+bijective lens it finds that relates two formats, that lens has a good chance of being the
+one a user wants.  In contrast, and to our surprise, our first experiments showed that
+when a search algorithm selects the smallest symmetric lens, it typically fails
+to find the lens the user wants!  The flexibity these lenses admit when it comes to
+eliding information generates an explosion of possibilities.  We needed to discover
+new principles for identifying the ``likely'' transformations and to
+develop new algorithms to search
+the space of program transformations using these principles.
+\dpw{Were the bijective lenses chosen using a ``smaller'' criteria?  I
+forget... if not edit the prior paragraph.}\dpw{Ideally, we could back up the claim
+that there are many symmetric lenses and not many bijective lenses by graphing the
+number of such lenses somehow.  eg, in a past synthesis paper with Steve (POPL 2016)
+we graphed
+the untyped terms of type $\mathsf{int}\rightarrow\mathsf{int}$.  I wonder if we can
+do something similar here.}
+
+To overcome this second difficulty, we turned to information theory
+to try to find a way to characterize the likely and unlikely lenses.  Intuitively, a ``likely'' lens
+is one that propagates a lot of information from source to target and also from target
+back to source.  Conversely, given the source, the target is ``unlikely'' to require very much
+additional information in order to reconstruct it---if it were, then one is unlikely to
+be attempting to synthesize a lens between the two formats!  In order to implement
+this intuition, we use \emph{stochastic regular expressions} (SREs).
+Stochastic regular expressions define a set of strings, as do ordinary regular expressions,
+but they also provide the expected distribution of strings within that set.  These
+distributions allow us to compute the amount of information required to reconstruct
+an element of the set.  For example, the language of the SRE $a |_{0.5} b$ contains two elements
+(the strings $a$ and $b$) and the probability 0.5 annotating the union indicates that both
+elements are equally likely.  Consquently, we observe that a transformation produces
+an element of that language,
+we will gain one bit of information.  On the other hand, consider the language of the SRE $c$.
+It contains just one element (the string $c$).  If we observe that our algorithm
+generates $c$, we gain zero bits of information.
+Hence, a lens is ``likely,'' in a technical sense, if relative to the target SRE $T$,
+very little information from $t$ is required to generate it. \dpw{check that I'm saying
+that correctly (and repair).  I think it is a bit off.}
+
+With SREs in hand, we develop a new algorithm for searching for likely lenses.
+This algorithm is constructed from a pair of collaborating search procedures.
+Given source and target SREs $S$ and $T$, the first algorithm searches for two \emph{related} SREs
+$S'$ and $T'$ that are compatible (a heuristic that estimates the likelihood
+our second algorithm will succeed).  $S'$ and $T'$ are related in the sense
+that they define the same regular sets as $S$ and $T$ (respectively) and the distribution
+of strings is identical.  Indeed, we guarantee these two properties by implementing
+a search procedure that obeys the star semi-ring axioms of regular expression equivalence,
+and which we have proven does not alter the probability distributions.
+Given the two compatible SREs,
+the second algorithm uses a greedy, type-directed algorithm find a likely lens between them.
+If algorithm estimates the lens it has discovered is more likely than
+any future lens it can possibly find, the search will be terminated.  Otherwise,
+this lens is stored away (if it is the best found so far) and the search continues with
+algorithm 1 again.
+
+We implemented this algorithm and evaluated its effectiveness on a range of benchmarks
+from XXXX to YYYYY. We found that ... \dpw{give a number of two that characterize
+our empirical findings.}
+
+In summary, our central contributions are as follows:
+\begin{itemize}
+\item We define the space of simple symmetric lenses, which lies between asymmetric lenses and full symmetric lenses and has the useful property that they don't have any internal state.
+We characterize simple as a subset of full symmetric lenses in terms of a semantic ``forgetfulness'' property.
+We analyze the gap between simple and full symmetric lenses and argue that simple ones suffice for many applications.
+\item We show how to synthesize simple symmetric lenses (including all asymmetric lenses as a special case), using a novel application of stochastic regular expressions to guide the search process.
+In particular, we show that the RE equivalences we are also valid when their semantics is enriched with distributions.
+\item We extend Boomerang implementation with simple symmetric lenses, showing that they can easily be combined with a rich variety of other lens extensions including quotients and matching lenses.
+\end{itemize}
+
+%Such synthesis efforts
+%make it easier to use lenses as programmers do not need to be familiar with the syntax
+%of a new language, and even advanced users can benefit from the productivity boost.
+
+\noindent
+\dpw{XXXXXX New attempt at intro ends here}
+
 % state of the world
 \bcp{Not very compelling:}Similar data can come in varieties of formats: scheduled jobs can be stored as
-cron jobs for Linux or as Scheduled Tasks for Windows; web service resources can
-be provided to clients in legacy or modern formats; calendar events can be
-stored in CSV files or ICS files. Oftentimes, files should be synchronized
-across these formats: when I edit a task to run more often on my Linux machine,
-I want this task to run more often on my Windows machine; when I update data
-through a legacy API, I want those changes to be reflected when I retrieve it
-through a modern API; when I create a new calendar event in my ICS file, I want
+cron jobs for Linux or as Scheduled Tasks for Windows; web service
+resources can be provided to clients in legacy or modern formats;
+calendar events can be stored in CSV files or ICS files. Oftentimes,
+files should be synchronized across these formats: when I edit a task
+to run more often on my Linux machine, I want this task to run more
+often on my Windows machine; when I update data through a legacy API,
+I want those changes to be reflected when I retrieve it through a
+modern API; when I create a new calendar event in my ICS file, I want
 my CSV calendar files to be updated to reflect this change.

### 2018-06-28 papers/SymmetricLenses/main.tex
intro stuff

@@
 
-In a big-data world, one will often find the same, or similar, data stored in a
-variety of different formats, with each format processed by a different set of tools.
-For instance, electronic calendars come in iCalendar, vCalendar and h-event
-formats~\cite{calendar-formats}; US
-taxation data can be transmitted via Tax XML (used by the US government for e-filing)
-or TXF (used by TurboTax)~\cite{tax-formats}; and sewing machines can read embroidery files transmitted
-as PES formats, DST formats or several others~\cite{embroidery-formats}.
+In a big-data world, one will often find similar data stored in a
+variety of different formats. For instance, electronic calendars come in iCalendar, vCalendar and h-event
+formats~\cite{calendar-formats}; US taxation data can be transmitted via Tax XML (used by the US government
+for e-filing) or TXF (used by TurboTax)~\cite{tax-formats}; and sewing machines can read embroidery files
+transmitted as PES formats, DST formats or several others~\cite{embroidery-formats}. \dpw{Editorial comment:
+Phil Wadler forced his very irritating long-form citations into the standard format. ug. how stupid is that.}
 
@@
 \emph{lens}~\cite{?}---a bi-directional
-program that can transform data represented in one format $S$ to data represented
-in a related format $T$,
+program that can transform data represented in a format $S$ to a format $T$,
 and vice versa.  There are a wide variety of different kinds of lenses, but well-defined lens
 languages typically come with guarantees that help ensure data is
-not lost or corrupted as it is transformed back and forth between formats.
+not lost or corrupted as it is transformed back and forth between different representations.
 
@@
 %a search through the space of possible computations that can be expressed as a lens.
-Recent research~\cite{miltner+:bijective-synthesis,maina+:quotient-synthesis}
-has exploited this phenomenon by demonstrating it is possible to synthesize relatively
+Recent research~\cite{optician}
+has exploited this fact by demonstrating it is possible to synthesize relatively
 large, complex lenses from specifications.
-More specifically, given types that describe source and target formats, these synthesis
-engines will quickly find bijective lenses~\cite{miltner+:bijective-synthesis} or
-bijections modulo isomorphism~\cite{maina+quotient-synthesis} (\EG, the isomorphism that
-treats whitespace as irrelevant) for medium-sized files, such as those housing linux
+In particular, given types that describe source and target formats, there are synthesis
+algorithms that will find bijective lenses~\cite{optician}
+for medium-sized files, such as those housing linux
 application configurations.  Oftentimes, these algorithms find such lenses in less than
-a second, and without the aid of any examples, just with types describing the given
-formats.
+a second, and without the aid of any examples at all, with just types describing the source and
+target formats.  \dpw{I would cite the ICFP paper (cite string: maina+:quotient-synthesis) as to appear but I'm not sure how to do that
+without possibly violating the spirit of double-blind reviewing. ie, if it hasn't appeared yet,
+how could we know the content if we aren't the authors?  Mentioning it in the 3rd person does not
+seem to suffice ...}
 
 One of reasons these algorithms are so effective, even relative to successful synthesis projects
-in other domains, is because the specifications cut down the space of programs very far
-indeed.  In particular, the restriction to bijectivity is both a blessing and a curse.
-It is a blessing, because given an input format, unless the same kind of data appears in
-both formats in multiple places, there are typically hardly any bijective lenses between
-the data formats:  If the synthesis algorithm can find a bijection then it is very likely
-to be the one the user wants.  On the other hand, the bijectivity restriction is a curse
-because in reality, bijections are very few and very far between.  It is much more common
-for data formats to have some data in common, but also to contain some differences.  For
-instance, two related tables might share three of five columns, but not the other two.
-Synchronizing such tables involves synchronizing three of the columns while managing the
-other two columns separately.  Indeed, in our study of Augeas~\cite{augeus}, a RedHat system that
+in other domains, is because unless the same kind of data appears in a format many times,
+there simply are not very many different bijections between related data formats.  Hence,
+if the synthesis algorithm can find any bijection at all, then it is very likely
+to be the one the user wants.  However, there is also a price for narrowing the synthesis
+domain so dramatically:  Many real data formats are almost, but not quite, bijections.
+More often two related data formats have some data in common, but also contain some differences.
+For instance, two related tables might share four of five columns, but not the last.  Alternately, one ad hoc
+configuration
+file might host some format-specific meta data, such as a date, a time or reference number, whereas the other
+does not.  Indeed, in our study of Augeas~\cite{augeus}, a RedHat system that
 uses lenses to edit Linux files, we found that XXXXX of YYYYY lenses were not bijective.
-\dpw{fill in with accurate stats.}
+\dpw{fill in with accurate stats and/or more concrete (and real) compelling examples; I made up the above.
+mention that the POPL paper benchmarks all had to be edited?}
+Consequently, past tools that search for strict bijections will rarely be applicable in practice.
 
@@
 allow each side of the transaction to retain information not present
-in the other side.  If $\ell$ is a symmetric lense between formats $S$ and $T$, we can
-use it to transform data from $S$ to $T$ or vice versa equally well.  We would typically
-use $\ell$ in the forwards direction (from $S$ to $T$) on a string $s$ when a user updates
-$s$---using $\ell$ in this direction will propagate those updates to $T$.
+in the other side.  If $\ell$ is a symmetric lens between formats $S$ and $T$, one can
+use it to transform data from $S$ to $T$, or vice versa equally well.  A user would typically
+run $\ell$ in the forwards direction (from $S$ to $T$) on a string $s$ when a user updates
+$s$---executing $\ell$ in this direction will propagate those updates to $T$.
 However, because data in format $T$ may contain some information not present in format $T$,
@@
 extract the missing components from $t$.
+
 Initially, we hypothesized that extending the algorithms
-of Miltner \ETAL~\cite{miltner+:bijective-synthesis} to this richer class of
-transformations would be relatively straightforward:  Simply extend the set
-of combinators and execute a search with similar heuristics.  However, in our early,
+of Miltner \ETAL~\cite{miltner+:bijective-synthesis} to operate over symmetric lenses
+as opposed to bijective lenses 
+would be relatively straightforward:  Simply replace one set of combinators with another
+and execute a search with similar heuristics.  However, in our early,
 naive attempts to do so, we ran
@@
 complement $C$~\cite{symmetric-lenses}.
-The latter contains all the information not present in one of $S$ or $T$.  Logistically,
+The latter object contains all the information not present in one of $S$ or $T$.  Logistically,
 the complement object must be stored away somewhere and infrastructure must be developed
 to manage it.  
-Perhaps more relevant though, is that compliments complicate specifications, even when
+Perhaps more relevant though, is that complements complicate specifications, even when
 those specifications are just examples.  The issues is that one cannot
@@
 variant of symmetric lenses, which we call \emph{simple symmetric lenses}.  Our new
-lenses do not require compliments, only need source and target data.
+lenses do not require complements; they only need source and target data as inputs.
+
 In order to characterize the expressive power of simple symmetric lenses
@@
 complete set of symmetric lenses that do not require a ``memory'' to retain data
-from past instances of $S$ or $T$ when making a round trip.\dpw{fix that}
+from past instances of $S$ or $T$ when making a round trip.\dpw{fix that to make the intuitive
+explaination sufficiently accurate}
 From a pragmatic perspective, we justify our design choices by extending
@@
 between two related formats is often tiny, the number of symmetric lenses is often
-large.  In practice, this meant that when a search algorithm simply selects the smallest
-bijective lens it finds that relates two formats, that lens has a good chance of being the
-one a user wants.  In contrast, and to our surprise, our first experiments showed that
+large.  In practice, the difference in cardinality means that when a search algorithm selects the smallest
+bijective lens it finds, that lens has a good chance of being the
+one a user wants.  In contrast, and to our surprise, our experiments show \dpw{can we have an experiment that
+shows this?} that
 when a search algorithm selects the smallest symmetric lens, it typically fails
-to find the lens the user wants!  The flexibity these lenses admit when it comes to
-eliding information generates an explosion of possibilities.  We needed to discover
-new principles for identifying the ``likely'' transformations and to
-develop new algorithms to search
-the space of program transformations using these principles.
+to find the lens the user wants.  The flexibity these lenses admit when it comes to
+eliding information generates an explosion of possible transformations.
+To disambiguate between the many possible lenses, we need
+new principles for identifying the ``likely'' ones, and we need
+new algorithms to search the space of program transformations using these principles.
 \dpw{Were the bijective lenses chosen using a ``smaller'' criteria?  I
@@
 be attempting to synthesize a lens between the two formats!  In order to implement
-this intuition, we use \emph{stochastic regular expressions} (SREs).
+this intuition, we turn to \emph{stochastic regular expressions} (SREs).
 Stochastic regular expressions define a set of strings, as do ordinary regular expressions,
-but they also provide the expected distribution of strings within that set.  These
-distributions allow us to compute the amount of information required to reconstruct
+but they also provide the expected distribution of strings within that set.  Such
+distributions allow us to compute the amount of information gained when presented
 an element of the set.  For example, the language of the SRE $a |_{0.5} b$ contains two elements
-(the strings $a$ and $b$) and the probability 0.5 annotating the union indicates that both
-elements are equally likely.  Consquently, we observe that a transformation produces
-an element of that language,
-we will gain one bit of information.  On the other hand, consider the language of the SRE $c$.
+(the strings $a$ and $b$).  The probability 0.5 annotating the union indicates that both
+elements are equally likely to appear.  Consequently, if we observe 
+an element of that language (either $a$ or $b$), we will gain one bit of information.  Said another way,
+one bit, together with the SRE, suffices to recover the string in question.
+On the other hand, consider the language of the SRE $c$.
 It contains just one element (the string $c$).  If we observe that our algorithm
-generates $c$, we gain zero bits of information.
+generates $c$, we gain zero bits of information --- the string was already determined.
 Hence, a lens is ``likely,'' in a technical sense, if relative to the target SRE $T$,
-very little information from $t$ is required to generate it. \dpw{check that I'm saying
-that correctly (and repair).  I think it is a bit off.}
-
-With SREs in hand, we develop a new algorithm for searching for likely lenses.
+very little information from $t$ is required to generate it.
+
+With SREs in hand, we develop a new algorithm for synthesizing likely lenses.
 This algorithm is constructed from a pair of collaborating search procedures.
@@
 a search procedure that obeys the star semi-ring axioms of regular expression equivalence,
-and which we have proven does not alter the probability distributions.
+and which we have proven does not alter the probability distributions of the languages.
+
 Given the two compatible SREs,
@@
 \item We define the space of simple symmetric lenses, which lies between asymmetric lenses and full symmetric lenses and has the useful property that they don't have any internal state.
-We characterize simple as a subset of full symmetric lenses in terms of a semantic ``forgetfulness'' property.
-We analyze the gap between simple and full symmetric lenses and argue that simple ones suffice for many applications.
-\item We show how to synthesize simple symmetric lenses (including all asymmetric lenses as a special case), using a novel application of stochastic regular expressions to guide the search process.
-In particular, we show that the RE equivalences we are also valid when their semantics is enriched with distributions.
-\item We extend Boomerang implementation with simple symmetric lenses, showing that they can easily be combined with a rich variety of other lens extensions including quotients and matching lenses.
+We characterize
+the simple lenses semantically as the subset of full symmetric lenses for which a ``forgetfulness''
+property holds. We also analyze the gap between simple and full symmetric lenses empirically and argue that
+simple ones suffice for many applications.
+\item We show how to synthesize simple symmetric lenses (including all asymmetric lenses as a special case),
+using a novel application of stochastic regular expressions to guide the search process.  In particular, we
+show that the RE star semi-ring equivalences remain valid when their semantics are enriched by distributions.
+\item We extend the Boomerang implementation with simple symmetric lenses, showing that they can easily be
+combined with a rich variety of other lens extensions including quotients and matching lenses.  We evaluate
+our implementation over a set of benchmarks drawn from XXXXX and demonstrate that YYYY ... numbers ....
 \end{itemize}

### 2018-06-29 papers/SymmetricLenses/main.tex
notes on new intro

@@
 transmitted as PES formats, DST formats or several others~\cite{embroidery-formats}. \dpw{Editorial comment:
-Phil Wadler forced his very irritating long-form citations into the standard format. ug. how stupid is that.}
+Phil Wadler forced his very irritating long-form citations into the
+standard format. ug. how stupid is that.}
+\ksf{I like the embroidery example and I also hate the long form citations.}
 
@@
 for medium-sized files, such as those housing linux
-application configurations.  Oftentimes, these algorithms find such lenses in less than
+application configurations.  \ksf{why medium sized?  You can have very
+large files with simple format descriptions that are bijective and for
+which our previous systems would have worked just fine.} Oftentimes, these algorithms find such lenses in less than
 a second, and without the aid of any examples at all, with just types describing the source and
@@
 how could we know the content if we aren't the authors?  Mentioning it in the 3rd person does not
-seem to suffice ...}
+seem to suffice ...} \ksf{The ICFP 2018 website lists the paper; I
+think it is okay to cite it.  The reviewer can assume we contacted the
+authors for a copy.  We could ask Stephanie for a ruling if we want to
+be sure.  One of the principles of double blind as adopted by SIGPLAN
+is not to reduce the quality of the paper; omitting this citation
+woudl have that affect.}
 
@@
 variant of symmetric lenses, which we call \emph{simple symmetric lenses}.  Our new
-lenses do not require complements; they only need source and target data as inputs.
+lenses do not require complements; they only need source and target
+data as inputs. \ksf{This seems magic.  Say something about how it is
+possible/under what circumstances it is possible to not need a
+complement?
+At the price of some expressiveness, our new lenses eliminate the need
+for complements; they require only source and target data as inputs.
+}
 
@@
 Given source and target SREs $S$ and $T$, the first algorithm searches for two \emph{related} SREs
-$S'$ and $T'$ that are compatible (a heuristic that estimates the likelihood
+$S'$ and $T'$ that are \emph{compatible} (a heuristic that estimates the likelihood
 our second algorithm will succeed).  $S'$ and $T'$ are related in the sense

### 2018-06-29 papers/SymmetricLenses/main.tex
Comments and a bit of polishing

@@
 is not to reduce the quality of the paper; omitting this citation
-woudl have that affect.}
+woudl have that effect.}\bcp{+1}
 
@@
 From a pragmatic perspective, we justify our design choices by extending
-the Boomerang implementation with symmetric lenses (they had never before been implemented),
+the Boomerang implementation~\cite{Boomerang} with symmetric lenses (which
+had never before been implemented\bcp{Confusing: we implement only simple
+  ones, but it was the earlier full form that had never been implemented}),
 applying them successfully to a range of real-world applications,
-and demonstrating that they are compatible, not just with basic bijective combinators,
+and demonstrating that they are compatible, not just with basic bijective
+combinators,\bcp{confusing: they are not exactly ``compatible'' with the
+  bijective combinators---they enrich them}
 but with a rich set of advanced features including quotient lenses and matching lenses.
 
-The second difficulty appears to be more fundamental\dpw{better word?}:
+The second difficulty with synthesizing symmetric lenses is more fundamental\dpw{better word?}:
 While the number of bijective lenses
@@
 back to source.  Conversely, given the source, the target is ``unlikely'' to require very much
-additional information in order to reconstruct it---if it were, then one is unlikely to
-be attempting to synthesize a lens between the two formats!  In order to implement
+additional information in order to reconstruct it---if it were\bcp{were
+  what?  likely, or unlikely?}, then one is unlikely to
+be attempting to synthesize a lens between the two formats\bcp{why? even if
+  two big formats only share a little bit of common information, mightn't
+  one want a lens between them?}!  In order to implement
 this intuition, we turn to \emph{stochastic regular expressions} (SREs).
@@
 an element of that language (either $a$ or $b$), we will gain one bit of information.  Said another way,
-one bit, together with the SRE, suffices to recover the string in question.
+one bit, together with the SRE, suffices to recover the string in
+question.\bcp{Is this implying that, if the annotation were
+  different, we'd need something other than one bit to recover the string??}
 On the other hand, consider the language of the SRE $c$.
@@
 generates $c$, we gain zero bits of information --- the string was already determined.
-Hence, a lens is ``likely,'' in a technical sense, if relative to the target SRE $T$,
+Hence\bcp{?}, a lens is ``likely,'' in a technical sense, if relative to the target SRE $T$,
 very little information from $t$ is required to generate it.
 
-With SREs in hand, we develop a new algorithm for synthesizing likely lenses.
+With simple symmetric lenses and SREs in hand, we develop a new algorithm for synthesizing likely lenses.
 This algorithm is constructed from a pair of collaborating search procedures.
@@
 a search procedure that obeys the star semi-ring axioms of regular expression equivalence,
-and which we have proven does not alter the probability distributions of the languages.
+and which we show does not alter the probability distributions of the languages.
 
@@
 the second algorithm uses a greedy, type-directed algorithm find a likely lens between them.
-If algorithm estimates the lens it has discovered is more likely than
-any future lens it can possibly find, the search will be terminated.  Otherwise,
+If algorithm estimates that the lens it has discovered is more likely than
+any future lens it can possibly find, the search terminates.  Otherwise,
 this lens is stored away (if it is the best found so far) and the search continues with
@@
 
-We implemented this algorithm and evaluated its effectiveness on a range of benchmarks
-from XXXX to YYYYY. We found that ... \dpw{give a number of two that characterize
+We implemented this\bcp{which?} algorithm and evaluated its effectiveness on a range of benchmarks
+from XXXX to YYYYY. We found that ... \dpw{give a number or two that characterize
 our empirical findings.}
 
-In summary, our central contributions are as follows:
+In summary, our main contributions are:\bcp{annotate with section numbers}
 \begin{itemize}

### 2018-06-29 papers/SymmetricLenses/main.tex
minor wording changes

--- 2018-06-29 papers/SymmetricLenses/main.tex
@@
 Recent research~\cite{optician}
-has exploited this fact by demonstrating it is possible to synthesize relatively
+has exploited this fact by demonstrating that it is possible to synthesize relatively
 large, complex lenses from specifications.
@@
 is not to reduce the quality of the paper; omitting this citation
-woudl have that affect.}
+woudl have that affect.}\saz{Stephanie says we should cite it in the third
+person, even if that's a bit awkward, but otherwise citing it is OK.}
 
@@
 
-Initially, we hypothesized that extending the algorithms
+Initially, we hypothesized that extending the synthesis algorithms
 of Miltner \ETAL~\cite{miltner+:bijective-synthesis} to operate over symmetric lenses
@@
 In order to characterize the expressive power of simple symmetric lenses
-theoretically, we prove simple symmetric lenses
+theoretically, we prove that simple symmetric lenses
 include all (and only) those symmetric lenses that satisfy

### 2018-06-29 papers/SymmetricLenses/main.tex

--- 2018-06-29 papers/SymmetricLenses/main.tex
@@
 is not to reduce the quality of the paper; omitting this citation
-woudl have that effect.}\bcp{+1}
+would have that affect.}\bcp{+1}\saz{Stephanie says we should cite it in the third
+person, even if that's a bit awkward, but otherwise citing it is OK.}
 

### 2018-06-29 papers/SymmetricLenses/main.tex
add random stuff in my lenssynth dir

--- 2018-06-29 papers/SymmetricLenses/main.tex
@@
 \begin{abstract}
-%% \saz{This abstract too long and doesn't get to the point quickly enough. We
-%%   should tighten it up considerably.}\bcp{+1.  But not just tighten ---
-%%   the whole story doesn't feel very compelling.}
-%%   Users commonly need to synchronize the data shared between formats, so edits
-%%   to the data in one format gets propagated to the data in the other format.
-%%   Manually writing the functions for these data synchronizers is tedious and
-%%   error prone. Instead of manually writing these synchronizers, bidirectional
-%%   languages can be used to express \emph{lenses}, groups of synchronization
-%%   functions with guarantees about round-trip behavior. While these bidirectional
-%%   languages help make programming these synthesizers less error-prone, they
-%%   remain difficult to program in. Synthesizing bidirectional languages allows
-%%   users to use lenses without manually programming them. Current
-%%   synthesizers work only on the simplest types of lenses, when the two formats
-%%   describe exactly the same data; this is often not the case.
-
-%%   Oftentimes, each format may contain information the other lacks. Such
-%%   synchronization tasks are described by one of the richest classes of lenses,
-%%   symmetric lenses. However, symmetric lenses in their current formulation
-%%   require external\bcp{why not ``internal''?} state in the form of a \emph{complement}. This statefulness
-%%   makes symmetric lenses both difficult to use and difficult to
-%%   synthesize. \bcp{This paragraph seems too dense.}
-
-%%   In this work, we reformulate\bcp{More a restriction than a reformulation.}
-%%   symmetric lenses as \emph{simple symmetric 
-%%     lenses}. While this formulation is less expressive than symmetric lenses, it
-%%   expresses many\bcp{weak} symmetric lenses while requiring \emph{no external state}. We
-%%   provide combinators for simple symmetric lenses on string data, and integrate
-%%   these combinators into Boomerang\bcp{undefined} while providing full backwards compatibility.
-
-%% \bcp{This bit feels more detailed / technical than needed for an abstract.}
-%%   We develop a type-directed synthesis algorithm for generating lenses in this
-%%   language from regular expressions describing the data formats. A significant
-%%   technical challenge in this work arises from the fact that our target language
-%%   permits synchronization between formats that don't have exactly the same data
-%%   -- data in one format need not be present in the other format. This gives the
-%%   synthesis algorithm much more freedom, as data can be completely unaligned
-%%   (edits in one format are never propagated), completely aligned (the formats
-%%   are in bijective correspondence), or anywhere in between. Under the assumption
-%%   that users typically want to align as much data as possible, we provide an
-%%   information theoretic metric for how much data is left unaligned based on the
-%%   theory of \emph{stochastic regular expressions}. Furthermore, we use this
-%%   metric to develop a greedy synthesis algorithm that tries to minimize
-%%   unaligned data. We experimentally demonstrate that our algorithm can
-%%   efficiently synthesize simple symmetric lenses on a variety of benchmark
-%%   suites consisting of synchronizing file formats from configuration files,
-%%   application-specific data storage files, and data cleaning tasks.
+\saz{This abstract too long and doesn't get to the point quickly enough. We
+  should tighten it up considerably.}\bcp{+1.  But not just tighten ---
+  the whole story doesn't feel very compelling.}
+  Users commonly need to synchronize the data shared between formats, so edits
+  to the data in one format gets propagated to the data in the other format.
+  Manually writing the functions for these data synchronizers is tedious and
+  error prone. Instead of manually writing these synchronizers, bidirectional
+  languages can be used to express \emph{lenses}, groups of synchronization
+  functions with guarantees about round-trip behavior. While these bidirectional
+  languages help make programming these synthesizers less error-prone, they
+  remain difficult to program in. Synthesizing bidirectional languages allows
+  users to use lenses without manually programming them. Current
+  synthesizers work only on the simplest types of lenses, when the two formats
+  describe exactly the same data; this is often not the case.
+
+  Oftentimes, each format may contain information the other lacks. Such
+  synchronization tasks are described by one of the richest classes of lenses,
+  symmetric lenses. However, symmetric lenses in their current formulation
+  require external\bcp{why not ``internal''?} state in the form of a \emph{complement}. This statefulness
+  makes symmetric lenses both difficult to use and difficult to
+  synthesize. \bcp{This paragraph seems too dense.}
+
+  In this work, we reformulate\bcp{More a restriction than a reformulation.}
+  symmetric lenses as \emph{simple symmetric 
+    lenses}. While this formulation is less expressive than symmetric lenses, it
+  expresses many\bcp{weak} symmetric lenses while requiring \emph{no external state}. We
+  provide combinators for simple symmetric lenses on string data, and integrate
+  these combinators into Boomerang\bcp{undefined} while providing full backwards compatibility.
+
+\bcp{This bit feels more detailed / technical than needed for an abstract.}
+  We develop a type-directed synthesis algorithm for generating lenses in this
+  language from regular expressions describing the data formats. A significant
+  technical challenge in this work arises from the fact that our target language
+  permits synchronization between formats that don't have exactly the same data
+  -- data in one format need not be present in the other format. This gives the
+  synthesis algorithm much more freedom, as data can be completely unaligned
+  (edits in one format are never propagated), completely aligned (the formats
+  are in bijective correspondence), or anywhere in between. Under the assumption
+  that users typically want to align as much data as possible, we provide an
+  information theoretic metric for how much data is left unaligned based on the
+  theory of \emph{stochastic regular expressions}. Furthermore, we use this
+  metric to develop a greedy synthesis algorithm that tries to minimize
+  unaligned data. We experimentally demonstrate that our algorithm can
+  efficiently synthesize simple symmetric lenses on a variety of benchmark
+  suites consisting of synchronizing file formats from configuration files,
+  application-specific data storage files, and data cleaning tasks.
 
@@
 \section{Introduction}
-
-\noindent
-\dpw{XXXXXX New attempt at intro starts here}
-
-In a big-data world, one will often find similar data stored in a
-variety of different formats. For instance, electronic calendars come in iCalendar, vCalendar and h-event
-formats~\cite{calendar-formats}; US taxation data can be transmitted via Tax XML (used by the US government
-for e-filing) or TXF (used by TurboTax)~\cite{tax-formats}; and sewing machines can read embroidery files
-transmitted as PES formats, DST formats or several others~\cite{embroidery-formats}. \dpw{Editorial comment:
-Phil Wadler forced his very irritating long-form citations into the
-standard format. ug. how stupid is that.}
-\ksf{I like the embroidery example and I also hate the long form citations.}
-
-One way to interconvert between these varied data formats is to use a
-\emph{lens}~\cite{?}---a bi-directional
-program that can transform data represented in a format $S$ to a format $T$,
-and vice versa.  There are a wide variety of different kinds of lenses, but well-defined lens
-languages typically come with guarantees that help ensure data is
-not lost or corrupted as it is transformed back and forth between different representations.
-
-Lens languages, like any domain-specific language, also cut down the space of
-possible programs one can write, and the smaller the space of programs,
-the easier it is to find the program one is looking for.  
-%This feature can be a negative, if the desired transformation
-%falls outside of the scope provided by the language, but can also be positive, as complex
-%properties (like invertability) can be enforced automatically.
-%Another advantage of reducing the
-%set of programs is that doing so can facilitate program synthesis, which is essentially
-%a search through the space of possible computations that can be expressed as a lens.
-Recent research~\cite{optician}
-has exploited this fact by demonstrating that it is possible to synthesize relatively
-large, complex lenses from specifications.
-In particular, given types that describe source and target formats, there are synthesis
-algorithms that will find bijective lenses~\cite{optician}
-for medium-sized files, such as those housing linux
-application configurations.  \ksf{why medium sized?  You can have very
-large files with simple format descriptions that are bijective and for
-which our previous systems would have worked just fine.} Oftentimes, these algorithms find such lenses in less than
-a second, and without the aid of any examples at all, with just types describing the source and
-target formats.  \dpw{I would cite the ICFP paper (cite string: maina+:quotient-synthesis) as to appear but I'm not sure how to do that
-without possibly violating the spirit of double-blind reviewing. ie, if it hasn't appeared yet,
-how could we know the content if we aren't the authors?  Mentioning it in the 3rd person does not
-seem to suffice ...} \ksf{The ICFP 2018 website lists the paper; I
-think it is okay to cite it.  The reviewer can assume we contacted the
-authors for a copy.  We could ask Stephanie for a ruling if we want to
-be sure.  One of the principles of double blind as adopted by SIGPLAN
-is not to reduce the quality of the paper; omitting this citation
-woudl have that affect.}\saz{Stephanie says we should cite it in the third
-person, even if that's a bit awkward, but otherwise citing it is OK.}
-
-One of reasons these algorithms are so effective, even relative to successful synthesis projects
-in other domains, is because unless the same kind of data appears in a format many times,
-there simply are not very many different bijections between related data formats.  Hence,
-if the synthesis algorithm can find any bijection at all, then it is very likely
-to be the one the user wants.  However, there is also a price for narrowing the synthesis
-domain so dramatically:  Many real data formats are almost, but not quite, bijections.
-More often two related data formats have some data in common, but also contain some differences.
-For instance, two related tables might share four of five columns, but not the last.  Alternately, one ad hoc
-configuration
-file might host some format-specific meta data, such as a date, a time or reference number, whereas the other
-does not.  Indeed, in our study of Augeas~\cite{augeus}, a RedHat system that
-uses lenses to edit Linux files, we found that XXXXX of YYYYY lenses were not bijective.
-\dpw{fill in with accurate stats and/or more concrete (and real) compelling examples; I made up the above.
-mention that the POPL paper benchmarks all had to be edited?}
-Consequently, past tools that search for strict bijections will rarely be applicable in practice.
-
-A natural way to expand the scope of such tools is to consider a more
-powerful language of lenses.  In particular, \emph{symmetric
-lenses}~\cite{symmetric-lenses} encompass a broad collection of bidirectional
-transformations and include bijections as well as transformations that
-allow each side of the transaction to retain information not present
-in the other side.  If $\ell$ is a symmetric lens between formats $S$ and $T$, one can
-use it to transform data from $S$ to $T$, or vice versa equally well.  A user would typically
-run $\ell$ in the forwards direction (from $S$ to $T$) on a string $s$ when a user updates
-$s$---executing $\ell$ in this direction will propagate those updates to $T$.
-However, because data in format $T$ may contain some information not present in format $T$,
-the $S \rightarrow T$ transformation will take
-both a string $s \in S$ and a string $t \in T$ as arguments.
-Inuitively, the lens should extract the data ``shared'' between the two formats from $s$
-as it generates an output $t'$.  If $s$ is missing information, the transformation should
-extract the missing components from $t$.
-
-Initially, we hypothesized that extending the synthesis algorithms
-of Miltner \ETAL~\cite{miltner+:bijective-synthesis} to operate over symmetric lenses
-as opposed to bijective lenses 
-would be relatively straightforward:  Simply replace one set of combinators with another
-and execute a search with similar heuristics.  However, in our early,
-naive attempts to do so, we ran
-into two different kinds of difficulties. \dpw{I got into the weeds a little bit there.
-The goal was to set up the idea that there are ``two parts'' to a symmetric
-lens transformation: the information from $s$ and the information from $t$.  In
-a future paragraph in the intro we are going to explain that the central heuristic
-involves minimizing the information from $t$.}
-
-The first difficulty was pragmatic:  symmetric lenses actually
-usually operate over three structures, the source $S$, the target $T$ and a
-complement $C$~\cite{symmetric-lenses}.
-The latter object contains all the information not present in one of $S$ or $T$.  Logistically,
-the complement object must be stored away somewhere and infrastructure must be developed
-to manage it.  
-Perhaps more relevant though, is that complements complicate specifications, even when
-those specifications are just examples.  The issues is that one cannot
-just give examples of source and target; one must also give guidance on how to manage
-the ``hidden state.''  In order to avoid such complexities, we defined a new
-variant of symmetric lenses, which we call \emph{simple symmetric lenses}.  Our new
-lenses do not require complements; they only need source and target
-data as inputs. \ksf{This seems magic.  Say something about how it is
-possible/under what circumstances it is possible to not need a
-complement?
-At the price of some expressiveness, our new lenses eliminate the need
-for complements; they require only source and target data as inputs.
-}
-
-In order to characterize the expressive power of simple symmetric lenses
-theoretically, we prove that simple symmetric lenses
-include all (and only) those symmetric lenses that satisfy
-an intuitive property called \emph{forgetfulness}.  In other words, they are the
-complete set of symmetric lenses that do not require a ``memory'' to retain data
-from past instances of $S$ or $T$ when making a round trip.\dpw{fix that to make the intuitive
-explaination sufficiently accurate}
-From a pragmatic perspective, we justify our design choices by extending
-the Boomerang implementation with symmetric lenses (they had never before been implemented),
-applying them successfully to a range of real-world applications,
-and demonstrating that they are compatible, not just with basic bijective combinators,
-but with a rich set of advanced features including quotient lenses and matching lenses.
-
-The second difficulty appears to be more fundamental\dpw{better word?}:
-While the number of bijective lenses
-between two related formats is often tiny, the number of symmetric lenses is often
-large.  In practice, the difference in cardinality means that when a search algorithm selects the smallest
-bijective lens it finds, that lens has a good chance of being the
-one a user wants.  In contrast, and to our surprise, our experiments show \dpw{can we have an experiment that
-shows this?} that
-when a search algorithm selects the smallest symmetric lens, it typically fails
-to find the lens the user wants.  The flexibity these lenses admit when it comes to
-eliding information generates an explosion of possible transformations.
-To disambiguate between the many possible lenses, we need
-new principles for identifying the ``likely'' ones, and we need
-new algorithms to search the space of program transformations using these principles.
-\dpw{Were the bijective lenses chosen using a ``smaller'' criteria?  I
-forget... if not edit the prior paragraph.}\dpw{Ideally, we could back up the claim
-that there are many symmetric lenses and not many bijective lenses by graphing the
-number of such lenses somehow.  eg, in a past synthesis paper with Steve (POPL 2016)
-we graphed
-the untyped terms of type $\mathsf{int}\rightarrow\mathsf{int}$.  I wonder if we can
-do something similar here.}
-
-To overcome this second difficulty, we turned to information theory
-to try to find a way to characterize the likely and unlikely lenses.  Intuitively, a ``likely'' lens
-is one that propagates a lot of information from source to target and also from target
-back to source.  Conversely, given the source, the target is ``unlikely'' to require very much
-additional information in order to reconstruct it---if it were, then one is unlikely to
-be attempting to synthesize a lens between the two formats!  In order to implement
-this intuition, we turn to \emph{stochastic regular expressions} (SREs).
-Stochastic regular expressions define a set of strings, as do ordinary regular expressions,
-but they also provide the expected distribution of strings within that set.  Such
-distributions allow us to compute the amount of information gained when presented
-an element of the set.  For example, the language of the SRE $a |_{0.5} b$ contains two elements
-(the strings $a$ and $b$).  The probability 0.5 annotating the union indicates that both
-elements are equally likely to appear.  Consequently, if we observe 
-an element of that language (either $a$ or $b$), we will gain one bit of information.  Said another way,
-one bit, together with the SRE, suffices to recover the string in question.
-On the other hand, consider the language of the SRE $c$.
-It contains just one element (the string $c$).  If we observe that our algorithm
-generates $c$, we gain zero bits of information --- the string was already determined.
-Hence, a lens is ``likely,'' in a technical sense, if relative to the target SRE $T$,
-very little information from $t$ is required to generate it.
-
-With SREs in hand, we develop a new algorithm for synthesizing likely lenses.
-This algorithm is constructed from a pair of collaborating search procedures.
-Given source and target SREs $S$ and $T$, the first algorithm searches for two \emph{related} SREs
-$S'$ and $T'$ that are \emph{compatible} (a heuristic that estimates the likelihood
-our second algorithm will succeed).  $S'$ and $T'$ are related in the sense
-that they define the same regular sets as $S$ and $T$ (respectively) and the distribution
-of strings is identical.  Indeed, we guarantee these two properties by implementing
-a search procedure that obeys the star semi-ring axioms of regular expression equivalence,
-and which we have proven does not alter the probability distributions of the languages.
-
-Given the two compatible SREs,
-the second algorithm uses a greedy, type-directed algorithm find a likely lens between them.
-If algorithm estimates the lens it has discovered is more likely than
-any future lens it can possibly find, the search will be terminated.  Otherwise,
-this lens is stored away (if it is the best found so far) and the search continues with
-algorithm 1 again.
-
-We implemented this algorithm and evaluated its effectiveness on a range of benchmarks
-from XXXX to YYYYY. We found that ... \dpw{give a number of two that characterize
-our empirical findings.}
-
-In summary, our central contributions are as follows:
-\begin{itemize}
-\item We define the space of simple symmetric lenses, which lies between asymmetric lenses and full symmetric lenses and has the useful property that they don't have any internal state.
-We characterize
-the simple lenses semantically as the subset of full symmetric lenses for which a ``forgetfulness''
-property holds. We also analyze the gap between simple and full symmetric lenses empirically and argue that
-simple ones suffice for many applications.
-\item We show how to synthesize simple symmetric lenses (including all asymmetric lenses as a special case),
-using a novel application of stochastic regular expressions to guide the search process.  In particular, we
-show that the RE star semi-ring equivalences remain valid when their semantics are enriched by distributions.
-\item We extend the Boomerang implementation with simple symmetric lenses, showing that they can easily be
-combined with a rich variety of other lens extensions including quotients and matching lenses.  We evaluate
-our implementation over a set of benchmarks drawn from XXXXX and demonstrate that YYYY ... numbers ....
-\end{itemize}
-
-%Such synthesis efforts
-%make it easier to use lenses as programmers do not need to be familiar with the syntax
-%of a new language, and even advanced users can benefit from the productivity boost.
-
-\noindent
-\dpw{XXXXXX New attempt at intro ends here}
-
 % state of the world
 \bcp{Not very compelling:}Similar data can come in varieties of formats: scheduled jobs can be stored as
-cron jobs for Linux or as Scheduled Tasks for Windows; web service
-resources can be provided to clients in legacy or modern formats;
-calendar events can be stored in CSV files or ICS files. Oftentimes,
-files should be synchronized across these formats: when I edit a task
-to run more often on my Linux machine, I want this task to run more
-often on my Windows machine; when I update data through a legacy API,
-I want those changes to be reflected when I retrieve it through a
-modern API; when I create a new calendar event in my ICS file, I want
+cron jobs for Linux or as Scheduled Tasks for Windows; web service resources can
+be provided to clients in legacy or modern formats; calendar events can be
+stored in CSV files or ICS files. Oftentimes, files should be synchronized
+across these formats: when I edit a task to run more often on my Linux machine,
+I want this task to run more often on my Windows machine; when I update data
+through a legacy API, I want those changes to be reflected when I retrieve it
+through a modern API; when I create a new calendar event in my ICS file, I want
 my CSV calendar files to be updated to reflect this change.
@@
 
+% begin stoch-rx
+% end stoch-rx
+
 % begin ssl
@@
 \]
-Type equivalence enables a lens of type $S \Leftrightarrow T$ to be used as a lens of type $S' \Leftrightarrow T'$ if $S$ equivalent to $S'$ and $T$ is equivalent to $T'$.
-Type equivalence is useful both for addressing type annotations and for
-letting compositions go through.\sam{I do not understand this statement.}
+Type equivalence enables a lens of type $S \Leftrightarrow T$ to be used as a
+lens of type $S' \Leftrightarrow T'$ if $S$ equivalent to $S'$ and $T$ is
+equivalent to $T'$. Type equivalence is useful both for addressing type
+annotations, and for making well-typed compositions.
 %\begin{center}
@@
 annotations would be.  The probabilities are defined as follows:
-\sam{Why in the definition of star do we multiply by $p^n * (1-p)$? I think this 
-should be explained}
 \begin{center}
@@
 \end{center}
+In the stochastic regular expression
+$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation \Probability
+represents the likelihood the string is in $\Regex_1$. In the stochastic regular
+expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability
+how much more likely having a string involving more than $n$ iterations is than
+having a string involving $n$ iterations.
 
@@
 Stochastic DNF regular expressions are hierarchically built, where stochastic
-DNF regular expressions are lists of stochastic sequences. Stochastic sequences
+DNF regular expressions (\DNFRegex, \DNFRegexAlt) are lists of stochastic sequences. Stochastic sequences
 themselves are lists of interleaved strings and stochastic atoms.  Stochastic

### 2018-06-29 papers/BijectiveLenses/ConferencePaper/transform-data.py

@@ -51,25 +51,20 @@ def generate_examples_required_graph(input_csv):
     zero_count_ind = 0
     one_to_five_count_ind = 1
     six_to_ten_count_ind = 2
-    eleven_to_fifteen_count_ind = 3
-    sixteen_to_twenty_count_ind = 4
-    over_twenty_count_ind = 5
+    over_twenty_count_ind = 3
 
     zero_count_text = "0"
     one_to_five_count_text = "1-5"
     six_to_ten_count_text = "6-10"
     eleven_to_fifteen_count_text = "11-15"
     sixteen_to_twenty_count_text = "16-20"
-    over_twenty_count_text = ">20"
+    over_twenty_count_text = ">10"
     ind_to_text = [zero_count_text,
                    one_to_five_count_text,
                    six_to_ten_count_text,
-                   eleven_to_fifteen_count_text,
-                   sixteen_to_twenty_count_text,
                    over_twenty_count_text]
 
-    experimental_values = [0,0,0,0,0,0,]
-    determinizing_values = [0,0,0,0,0,0,]
+    experimental_values = [0,0,0,0,]
 
     def add_to_correct_group(count_values, n):
         if n < 0.0:
@@ -80,40 +75,35 @@ def generate_examples_required_graph(input_csv):
             count_values[one_to_five_count_ind] = count_values[one_to_five_count_ind]+1
         elif n <= 10.0:
             count_values[six_to_ten_count_ind] = count_values[six_to_ten_count_ind]+1
-        elif n <= 15.0:
-            count_values[eleven_to_fifteen_count_ind] = count_values[eleven_to_fifteen_count_ind]+1
-        elif n <= 20.0:
-            count_values[sixteen_to_twenty_count_ind] = count_values[sixteen_to_twenty_count_ind]+1
         else:
             count_values[over_twenty_count_ind] = count_values[over_twenty_count_ind]+1
 
     experimental_vals = project_column_from_csv(input_csv, "ExamplesRequired")
     for example_num in experimental_vals:
         add_to_correct_group(experimental_values, float(example_num))
-    determinizing_vals = project_column_from_csv(input_csv, "MaxExampleCount")
-    for example_num in determinizing_vals:
-        add_to_correct_group(determinizing_values, float(example_num))
     
 
[…]

### 2018-06-29 papers/DifferenceLenses/main.tex

@@
 Each one of these solutions has significant shortcomings.
-Firstly, while matching lenses successfully address the issue of alignment, they do not address the general issue of correct update propagation.
-Secondly, edit lenses and delta lenses all assume that the lens propagates a specific set of updates explicitly, but this is unrealistic since since in general users do not update data using a small predetermined langauge of edits; instead users edit the data in any way, for example using text editors or IDEs, and present the updatad data to the bx wholesale.
+While matching lenses successfully address the issue of alignment, they do not address the general issue of correct update propagation.
+Edit lenses and delta lenses all assume that the lens propagates a specific set of updates explicitly, which is unrealistic since since in general users do not update data using a small predetermined langauge of edits; instead they edit the data in any way, for example using text editors or IDEs, and present the updatad data to the bx wholesale.
 Finally, least change lenses are in general computationally intractable \cite{buneman2002propagation} and also lack a notion of composition---that is, the composition of two deterministic least change lenses is not always a least change lens\cite{macedo2013composing}.
@@
 0 & \text{if } m = 0, n \neq 0\\
-0 & \text{if } m \neq 0, n = 0\\\max\limits_{\{\varphi : [m] \longrightarrow [n] \; | \; \varphi(i) = i\}}\frac{1}{n}\sum\limits_{i=1}^{m}\ \delta_R(s_i, t_{\varphi(i)})& \text{if } m \leq n\\
-\max\limits_{\{\varphi : [n] \longrightarrow [m] \; | \; \varphi(i) = i\}}\frac{1}{m}\sum\limits_{i=1}^{m}\ \delta_R(t_i, s_{\varphi(i)})& \text{otherwise}
+0 & \text{if } m \neq 0, n = 0\\\max\limits_{\{\varphi : [m] \rightharpoonup [n] \; | \; \varphi(i) = i\}}\sum\limits_{i \in Dom(\varphi)}\ \frac{\delta_R(s_i, t_{\varphi(i)})}{\max(m,n)}& \text{otherwise}
 \end{cases}\\
@@
 0 & \text{if } m \neq 0, n = 0\\
-\max\limits_{\{\varphi : [m] \longrightarrow [n] \; | \; \varphi \text{ is strictly increasing}\}}\frac{1}{n}\sum\limits_{i=1}^{m}\ \delta_R(s_i, t_{\varphi(i)})& \text{if } m \leq n\\
-\max\limits_{\{\varphi : [n] \longrightarrow [m] \; | \; \varphi \text{ is strictly increasing}\}}\frac{1}{m}\sum\limits_{i=1}^{m}\ \delta_R(t_i, s_{\varphi(i)})& \text{otherwise}
+\max\limits_{\{\varphi : [m] \rightharpoonup [n] \; | \; \varphi \text{ is str. increasing}\}}\sum\limits_{i \in Dom(\varphi)}\ \frac{\delta_R(s_i, t_{\varphi(i)})}{\max(m,n)}& \text{otherwise}
 \end{cases}\\
@@
 0 & \text{if } m = 0, n \neq 0\\
-0 & \text{if } m \neq 0, n = 0\\\max\limits_{\{\varphi : [m] \longrightarrow [n] \; | \; \varphi \text{ is injective}\}}\frac{1}{n}\sum\limits_{i=1}^{m}\ \delta_R(s_i, t_{\varphi(i)})& \text{if } m \leq n\\
-\max\limits_{\{\varphi : [n] \longrightarrow [m] \; | \; \varphi \text{ is injective}\}}\frac{1}{m}\sum\limits_{i=1}^{m}\ \delta_R(t_i, s_{\varphi(i)})& \text{otherwise}
+0 & \text{if } m \neq 0, n = 0\\\max\limits_{\{\varphi : [m] \rightharpoonup [n] \; | \; \varphi \text{ is injective}\}}\sum\limits_{i \in Dom(\varphi)}\ \frac{\delta_R(s_i, t_{\varphi(i)})}{\max(m,n)}& \text{otherwise }
 \end{cases}
@@
 \item
+$\delta_R(s, s') = \delta_R(s', s)$.
+\item
 For all $s, s' \in \mathcal{L}(R^*)$, $\delta_{R^*_{MED}}(s, s') \geq \delta_{R^*_{NC}}(s, s') \geq \delta_{R^*_{MED}}(s, s')$.
-\item
-$\delta_R(s, s') = \delta_R(s', s)$.
 \end{enumerate}
@@
 Let $s_1 \cdot \ldots \cdot s_m$ be a source, and $v_1 \cdot \ldots \cdot v_n$ a view.
-If $n \leq m$ and $\varphi : [n] \longrightarrow [m]$, define $ v \xhookrightarrow{\ell, \varphi} s = s'_1 \cdot \ldots \cdot s'_n$ by
+Let $\varphi : [n] \rightharpoonup [m]$ and define $ v \xhookrightarrow{\ell, \varphi} s = s'_1 \cdot \ldots \cdot s'_n$ by
 $$s'_i = \begin{cases}
@@
 \ell.\create \; v_i & \text{if } i \not \in Dom(\varphi)
-\end{cases}$$
-If $m < n$ and $\varphi : [m] \longrightarrow [n]$, define $ v \xhookrightarrow{\ell, \varphi} s = s'_1 \cdot \ldots \cdot s'_n$ by 
-$$s'_i = \begin{cases}
-\ell.\pput \; s_{\varphi^{-1}(i)} \; v_i & \text{if } i \in \varphi([m])\\
-\ell.\create \; v_i & \text{if } i \not \in \varphi([m])
 \end{cases}$$
@@
 &= v \xhookrightarrow{\psi} s\\
-\text{where } \psi &= \begin{cases}
-\argmax \limits_{\{\varphi : [len(v)] \longrightarrow [len(s)] \; | \; \varphi(i) = i\}} \delta_{S^*_{NC}}(v \xhookrightarrow{\ell, \varphi} s, s) & \text{if } len(v) \leq len(s)\\
-\argmax \limits_{\{\varphi : [len(s)] \longrightarrow [len(v)] \; | \; \varphi(i) = i\}} \delta_{S^*_{NC}}(v \xhookrightarrow{\ell, \varphi} s, s) & \text{if } len(s) < len(v)
-\end{cases}\\
+\text{where } \psi &= 
+\argmax \limits_{\{\varphi : [len(v)] \rightharpoonup [len(s)] \; | \; \varphi(i) = i\}} \delta_{S^*_{POS}}(v \xhookrightarrow{\ell, \varphi} s, s)\\
 \ell^*_{POS}.\create \; (v_1, \ldots, v_n) &= (\ell.\create \;
@@
 &= v \xhookrightarrow{\psi} s\\
-\text{where } \psi &= \begin{cases}
-\argmax \limits_{\{\varphi : [len(v)] \longrightarrow [len(s)] \; | \; \varphi \text{ is str. increasing}\}} \delta_{S^*_{NC}}(v \xhookrightarrow{\ell, \varphi} s, s) & \text{if } len(v) \leq len(s)\\
-\argmax \limits_{\{\varphi : [len(s)] \longrightarrow [len(v)] \; | \; \varphi \text{ is str. increasing}\}} \delta_{S^*_{NC}}(v \xhookrightarrow{\ell, \varphi} s, s) & \text{if } len(s) < len(v)
-\end{cases}\\
+\text{where } \psi &= 
+\argmax \limits_{\{\varphi : [len(v)] \rightharpoonup [len(s)] \; | \; \varphi \text{ is str. increasing}\}} \delta_{S^*_{NC}}(v \xhookrightarrow{\ell, \varphi} s, s) \\
 \ell^*_{NC}.\create \; (v_1, \ldots, v_n) &= (\ell.\create \;
@@
 &= v \xhookrightarrow{\ell, \psi} s\\
-\text{where } \psi &= \begin{cases}
-\argmax \limits_{\{\varphi : [len(v)] \longrightarrow [len(s)] \; | \; \varphi \text{ is injective}\}} \delta_{S^*_{NC}}(v \xhookrightarrow{\ell, \varphi} s, s) & \text{if } len(v) \leq len(s)\\
-\argmax \limits_{\{\varphi : [len(s)] \longrightarrow [len(v)] \; | \; \varphi \text{ is injective}\}} \delta_{S^*_{NC}}(v \xhookrightarrow{\ell, \varphi} s, s) & \text{if } len(s) < len(v)
-\end{cases}\\
+\text{where } \psi &= 
+\argmax \limits_{\{\varphi : [len(v)] \rightharpoonup [len(s)] \; | \; \varphi \text{ is injective}\}} \delta_{S^*_{MED}}(v \xhookrightarrow{\ell, \varphi} s, s)\\
 \ell^*_{MED}.\create \; (v_1, \ldots, v_n) &= (\ell.\create \;
@@
 \end{align*}
-For the non-crossing and min-edit distance lens, there may be multiple alignments $\psi$ that maximize the respective differences, in which case the lens chooses the lexicographically smallest of those alignments.
-
 \begin{claim}
@@
 $$\delta_{S^*_{MED}}(\ell^*_{MED}.\pput \; (s_1 \cdot \ldots \cdot s_m) \; (v_1 \cdot \ldots \cdot v_n), s_1 \cdot \ldots \cdot s_m) \geq \delta_{V^*_{MED}}(v_1 \cdot \ldots \cdot v_n, (\ell.\get \; s_1) \cdot \ldots (\ell.\get \; s_m))$$
-Assume wthout loss of generality that $n \leq m$.
-An optimal min-edit distance alignment $\varphi : [n] \longrightarrow [m]$ is used to compute the difference on the view side.
-Since $\ell$ is a difference lens, the following equation holds,
+An optimal min-edit distance alignment $\varphi : [n] \rightharpoonup [m]$ is used to compute the difference on the view side. Since $\ell$ is a difference lens, the following equation holds,
 $$\delta_S(\ell.\pput \; s_{\varphi(i)} \; v_i, s_{\varphi(i)}) \geq \delta_V(v_i, \ell.\get \; s_{\varphi(i)})$$
 from which it follows that
-$$\delta_{S^*_{MED}}(v_1 \cdot \ldots v_n \xhookrightarrow{\ell, \varphi} s_1 \cdot \ldots \cdot s_m, s_1 \cdot \ldots \cdot s_m) \geq \delta_{V^*_{NC}}(v_1 \cdot \ldots \cdot v_n, (\ell.\get \; s_1) \cdot \ldots (\ell.\get \; s_m))$$
+$$\delta_{S^*_{MED}}(v_1 \cdot \ldots v_n \xhookrightarrow{\ell, \varphi} s_1 \cdot \ldots \cdot s_m, s_1 \cdot \ldots \cdot s_m) \geq \delta_{V^*_{MED}}(v_1 \cdot \ldots \cdot v_n, (\ell.\get \; s_1) \cdot \ldots (\ell.\get \; s_m))$$
 Since the \pput function of $\ell^*_{MED}$ uses an alignment $\psi$ that maximizes the min-edit distance over all injective alignments mapping $v$ into $s$, then
@@
 It thus follows that,
-$$\delta_{S^*_{MED}}(v_1 \cdot \ldots v_n \xhookrightarrow{\ell, \psi} s_1 \cdot \ldots \cdot s_m, s_1 \cdot \ldots \cdot s_m) \geq \delta_{V^*_{NC}}(v_1 \cdot \ldots \cdot v_n, (\ell.\get \; s_1) \cdot \ldots (\ell.\get \; s_m))$$
+$$\delta_{S^*_{MED}}(v_1 \cdot \ldots v_n \xhookrightarrow{\ell, \psi} s_1 \cdot \ldots \cdot s_m, s_1 \cdot \ldots \cdot s_m) \geq \delta_{V^*_{MED}}(v_1 \cdot \ldots \cdot v_n, (\ell.\get \; s_1) \cdot \ldots (\ell.\get \; s_m))$$
 hence,
@@
 \delta_{(R \cup S)^*_{MED}} (\strunfilter \; S \; s \; v, s) \geq \delta_{R^*_{MED}}(v, \strfilter \; R \; s)$$
-Observe that if $s = s_1 \cdot \ldots cdot s_m$ with $s_{i_1}, \ldots, s_{i_p} \in R$ and $s_{j_1}, \ldots, s_{j_p} \in S$, then $\strfilter \; R \; s = s_{i_1} \cdot \ldots \cdot s_{i_n}$.
-Assume that $v = v_1 \cdot \ldots v_{n'}$ and assume without loss of generality that $n \leq n'$.
+Observe that if $s = s_1 \cdot \ldots \cdot s_m$ with $s_{i_1}, \ldots, s_{i_p} \in R$ and $s_{j_1}, \ldots, s_{j_p} \in S$, then $\strfilter \; R \; s = s_{i_1} \cdot \ldots \cdot s_{i_n}$.
+Assume that $v = v_1 \cdot \ldots v_{n'}$.
 Then since $\delta_{R^*_{MED}}(v, \strfilter \; R \; s) \leq 1$, it follows that
 $$
-\delta_{(R \cup S)^*_{MED}} (\strunfilter \; S \; s \; v, s) = \frac{m + \delta_{R^*_{MED}}(v, \strfilter \; R \; s) * n'}{m+n'} \geq \delta_{R^*_{MED}}(v, \strfilter \; R \; s)
+\delta_{(R \cup S)^*_{MED}} (\strunfilter \; S \; s \; v, s) = \frac{m + \delta_{R^*_{MED}}(v, \strfilter \; R \; s) * \max(n,n')}{m+\max(n,n')} \geq \delta_{R^*_{MED}}(v, \strfilter \; R \; s)
 $$
@@
 0 & \text{if } m = 0, n \neq 0\\
-0 & \text{if } m \neq 0, n = 0\\\max\limits_{\{\varphi : [m] \longrightarrow [n] \; | \; \varphi \text{ is injective}\}}\frac{1}{n}\sum\limits_{i=1}^{m}\ \delta_R(s_i, t_{\varphi(i)})& \text{if } m \leq n\\
-\max\limits_{\{\varphi : [n] \longrightarrow [m] \; | \; \varphi \text{ is injective}\}}\frac{1}{m}\sum\limits_{i=1}^{m}\ \delta_R(t_i, s_{\varphi(i)})& \text{otherwise}
+0 & \text{if } m \neq 0, n = 0\\\max\limits_{\{\varphi : [m] \rightharpoonup [n] \; | \; \varphi \text{ is injective}\}}\sum\limits_{i=1}^{m}\ \frac{\delta_R(s_i, t_{\varphi(i)})}{\max(m,n)}& \text{otherwise}
 \end{cases}

### 2018-06-29 papers/QuotientLenses/main.tex

@@
 
-Recent work~\cite{optician} developed a tool called Optician to synthesize a
+Recent work~\cite{miltner2017synthesizing} developed a tool called Optician to synthesize a
 subset of the {\em bijective lenses} that can be derived in the lens
@@
 that guide the synthesis algorithm. This presents a challenge for synthesizing
-quotient lenses since a specification of its source and target formats
+quotient lenses since a specification of a lens's source and target formats
 needs to account for the equivalence relations defined on the
@@
 % via a bijective lens between the respective canonical formats of the QREs. This
-% idea is simple and natural, but begs the following question: do we give up
+% idea is simple and natural, but provokes the following question: do we give up
 % expressiveness if we restrict ourselves to this form? We will show that this
@@
 Recent work described the Optician
-algorithm/tool~\cite{optician}, which can synthesize bijective lenses
+algorithm/tool~\cite{miltner2017synthesizing}, which can synthesize bijective lenses
 such as \cd{bib_to_end}, obviating the sometimes tedious tasks
@@
 and $q.\lput : \mathcal{L}(S) \longrightarrow \mathcal{L}(R)$ such that for all
-$r \in \mathcal{L}(R)$ and $s \in \mathcal{L}(S)$, we have $q.\lput(q_1.\get(r))
-\equiv_R r$ and $q.\get(q_1.\lput(s)) \equiv_R s$. Moreover, if $r \equiv_R
+$r \in \mathcal{L}(R)$ and $s \in \mathcal{L}(S)$, we have $q.\lput(q.\get(r))
+\equiv_R r$ and $q.\get(q.\lput(s)) \equiv_S s$. Moreover, if $r \equiv_R
 r'$, then $q.\get(r) = q.\get(r')$, and if $s \equiv_S s'$, then $q.\lput(s) =
@@
 that the {\em get} and {\em put} functions lifted to $\mathcal{L}(R)/{\equiv_R}$
-and $\mathcal{L}(S)/{\equiv_S}$ are mutual inverses. These laws are similar
+and $\mathcal{L}(S)/{\equiv_S}$ are mutual inverses. These laws are similar to the
 bijective lens laws \cref{bijectivelenslaws}, except that the equality
@@
 source language of $\ell$ is $\mathcal{L}(R)$ and the target language of $\ell$
-is $\mathcal{L}(S)$. IThe final rule in the figure deserves special attention.
+is $\mathcal{L}(S)$. The final rule in the figure deserves special attention.
 It states that a lens $\ell : R \Leftrightarrow S$ can also be considered to be
@@
 
-The Optician framework~\cite{optician} showed how to do such lens synthesis in
+The Optician framework~\cite{miltner2017synthesizing} showed how to do such lens synthesis in
 the case for bijective lenses. Here we show how to reduce QRE lens synthesis to
@@
 details on this algorithm can be found in the original paper on synthesizing
-bijective lenses~\cite{optician}.
+bijective lenses~\cite{miltner2017synthesizing}.
 
@@
   the external source and target formats and (2) to estimate the
-  burdern of writing the requisite canonizers
+  burden of writing the requisite canonizers
   by hand.  We count the nodes in the examples because they would be
@@
 
-A newer language which supports quotenting is BiFluX~\cite{pacheco2014biflux}, a
+A newer language which supports quotenting is FliPpr \cite{matsuda2013flippr}.
+FliPpr is a program transformation system that uses program inversion to produce a CFG parser from a pretty-printer.
+FliPpr includes biased choice combinators that choose canonical representatives for sets of strings.
+For example, one can define variants of (white)spaces with the choice operator as,
+\begin{center}
+\begin{lstlisting}
+nil = text "" <+ space
+space = (text " " <+ text "\n") <> nil
+\end{lstlisting}
+\end{center}
+where \lstinline|nil| and \lstinline|space| pretty-print \lstinline|""| and \lstinline|" "| respectively, but represent zero-or-more and one-or-more whitespaces in parsing.
+FliPpr also allows annotations for optional parentheses in programs.
+These features enable a user to easily define a pretty-printer that also recognizes ugly strings that get mapped to a canonical representative modulo whitespace and parentheses quotienting.
+
+Another programming language that supports quotienting is BiFluX~\cite{pacheco2014biflux}, a
 bidirectional functional update language for XML. BiFluX is inspired by
@@
 In BiFluX, a regular expression $R$ is said to be unambiguous if there is only
-one way to parse a {\em flat} value of type $R$ to a stuctured value of type
+one way to parse a {\em flat} value of type $R$ to a structured value of type
 $R$; here a flat value of $R$ is a value of $R$ with all left/right tags,
@@
 For example, imagine that a book database has source type \\
-$S$ = \lstinline{books[book[title[string], author[string]+]*]}
-with view type $V$ = \lstinline{title[string]}. When the source type is
+$S$ = \lstinline|books[book[title[string], author[string]+]*]|
+with view type $V$ = \lstinline|title[string]|. When the source type is
 transformed to the view type for matching, only titles are kept, the element
 labels are destroyed and all authors are replaced for the empty sequence, thus
-leaving the intermediate type $S'$ = \lstinline{(title[string],(),()*)*}. In
+leaving the intermediate type $S'$ = \lstinline|(title[string],(),()*)*|. In
 BiFluX, $V$ is considered a subtype of $S'$ because the flat values of $V$ are
 also flat values of $S'$. Since the intermediate type is $S'$ ambiguous, a view
-sequence \lstinline{[title["mybook"]]} would be upcast into an intermediate
-value \lstinline{[(title["mybook"],((), [ ]))]} that would in turn lead to an
-updated source \lstinline{books[[book[(title["mybook"],(aut, [ ]))]]]}, which is
+sequence \lstinline|[title["mybook"]]| would be upcast into an intermediate
+value \lstinline|[(title["mybook"],((), [ ]))]| that would in turn lead to an
+updated source \lstinline|books[[book[(title["mybook"],(aut, [ ]))]]]|, which is
 clearly unsatisfactory as it would discard the entire book database except the
-first author \lstinline{aut} of the first book.
+first author \lstinline|aut| of the first book.
 \fi
@@
 low-level details including selecting field names for elements and parsing data
-from client responses. In their work, Rajkumar et. al. represent formlets as
-functions that take a source of names---concretely, an integer $n$ that should
-be used as the next name---as an argument and produce a triple consisting of an
+from client responses. In their work, Rajkumar et. al. represent a formlens as a
+function that takes an optional initial value of a form and a list of integers as a name source and produces a triple consisting of an
 HTML document, a {\em collect} function, and a modified namesource:
-\begin{center}
-\begin{lstlisting}
-type Formlet a = Int $\to$ (Html, Env $\to$ a, Int)
-\end{lstlisting}
-\end{center}
-The names of any generated form fields are drawn from the namesource, and the
-{\em collect} function looks up precisely these names. For example the
-\kw{inputInt} combinator below builds a form that accepts an integer,
-\begin{center}
-\begin{lstlisting}
-inputInt :: Formlet Int
-inputInt i = let n = show i in
-             ($\kw{inputTag}$ {name = n, value = ""},$\lambda e \longrightarrow \kw{read}$ ($\kw{fromJust}$ ($\kw{lookup}$ n e)),i+1)
-\end{lstlisting}
-\end{center}
-where \kw{inputTag} constructs an input element, \kw{lookup}
-retrieves a value from an association list \kw{fromJust} extracts the
-encapsulated value from a \kw{Maybe} value, and \kw{read} parses an \kw{Int}
-from a \kw{String}.
-
-Rajkumar et. al. observe that formlets are a useful abstraction, but they only
-address one half of the problem---they make it easy to construct a function
-that {\em produces} a value of type \codefont{a}, but they do not provide a way
-to describe a function that {\em consumes} a value of type \codefont{a} and
-embeds it in a form. A different way to think about this problem is to observe
-that forms are often used to present an {\em updateable view} of the data
-source. Therefore, Rajkumar et. al. leverage the power of bidirectional
-transformations to define formlets that behave like updatable views, yielding
-the abstraction of formlenses.
-
-The Formlens type is given by
 \begin{center}
@@
 \end{center}
-with the differences between this definition and the standard definition for
-formlets being an extra \codefont{Maybe\; a} parameter that provides an
-optional initial value of the form, the namesource being changed to a list of
-integers, and the {\em collect} function being allowed to return an optional
-value.
-
-Rajkumar et. al. consider only formlenses that are well-formed in that a
+The names of any generated form fields are drawn from the name source, and the
+{\em collect} function looks up precisely these names
+
+Rajkumar et. al. consider only formlenses that are well-formed in the sense that a
 well-formed formlens should only draw names from the namesource provided as an
@@
 synthesis assumes that the synthesizer is provided with a collection of examples.
-Optician and Optician$_Q$ differ in that they require the programmer to supply
+Optician and QRE-Enhanced Optician differ in that they require the programmer to supply
 both examples {\em and} format descriptions in the form of regular expressions
@@
 is significantly more complex.  In particular, because Optician types do not
-have full canonical forms, Anders et al. used a pseudo-canonical form that
-captures part of the equivalence relation over types were used. To preserve
+have full canonical forms, Miltner et al. \cite{miltner2017synthesizing} used a pseudo-canonical form that
+captures part of the equivalence relation over types. To preserve
 completeness, they pushed some of the remaining parts of the type equivalence

### 2018-06-29 papers/QuotientLenses/main.tex
Fixed some more typos

@@
 string that matches a regular expression $R$ to some canonical
-representative in another regular expression $R'$. The equivalence
+representative in another regular expression $R'$ where $\mathcal{L}(R') 
+\subseteq \mathcal{L}(R)$. The equivalence
 relation defined by the normalize combinator is hence the equivalence
@@
 \\[2ex]
-\multicolumn{3}{l}{\canonize(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
-\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)})(w_{\sigma(1)} \cdot
+\multicolumn{3}{l}{\canonize(\perm{Q_1, \ldots, Q_n}{Q})(w_{\sigma(1)} \cdot
 s_1 \cdot \ldots \cdot s_{n-1} \cdot w_{\sigma(n)})}\\[.5ex]
-\multicolumn{3}{l}{\ =\ \ \canonize(Q_1)(r_1)(w_1) \cdot \canonize(Q)(s_1)
-\cdot \ldots \cdot \canonize(Q)(s_{n-1})\cdot\canonize(Q_n)(r_n)(w_n)}
+\multicolumn{3}{l}{\ =\ \ \canonize(Q_1)(w_1) \cdot \canonize(Q)(s_1)
+\cdot \ldots \cdot \canonize(Q)(s_{n-1})\cdot\canonize(Q_n)(w_n)}
 \end{array}
@@
 \begin{array}{l}
-\canonize(\perm{Q_1, \ldots, Q_n}{Q})(r_{\sigma(1)}
-\cdot s_1 \cdot \ldots \cdot s_{n-1} \cdot r_{\sigma(n)})(w_{\sigma(1)} \cdot s_1
+\canonize(\perm{Q_1, \ldots, Q_n}{Q})(w_{\sigma(1)} \cdot s_1
   \cdot \ldots \cdot s_{n-1} \cdot w_{\sigma(n)})\\[.5ex]
-\ =\ \ \canonize(Q_1)(r_1)(w_1) \cdot \canonize(Q)(s_1) \cdot \ldots \cdot
-\canonize(Q)(s_{n-1})\cdot\canonize(Q_n)(r_n)(w_n)
+\ =\ \ \canonize(Q_1)(w_1) \cdot \canonize(Q)(s_1) \cdot \ldots \cdot
+\canonize(Q)(s_{n-1})\cdot\canonize(Q_n)(w_n)
 \end{array}
@@
 \cdot r_m = r'_1 \cdot \ldots \cdot r'_n$, then $m = n$ and $r_i = r'_i$. We say
-that a regular expression $R$ is \textit{strongly unambiguous}\cite{Sippu1988}
+that a regular expression $R$ is \textit{strongly unambiguous} \cite{Sippu1988}
 if and only if (1) $R = \varnothing$, or (2) $R = S_1 \cdot S_2$ with $S_1,
@@
 \ell_2 \rrbracket\}\\
-\llbracket \ell_1 \cdot \ell_2 \rrbracket &= \{(s \cdot t, s' \cdot t) \sep
+\llbracket \ell_1 \cdot \ell_2 \rrbracket &= \{(s \cdot t, s' \cdot t') \sep
 (s, s') \in \llbracket \ell_1 \rrbracket \text{ and } (t, t') \in \llbracket
@@
 \llbracket \ell_1 \; ; \; \ell_2 \rrbracket &= \{(r, t) \sep \text{there exists }s
-\text{ such that } (r, s) \in \llbracket \ell_1 \rrbracket \text{ and } (s, t)
+\text{ with } (r, s) \in \llbracket \ell_1 \rrbracket \text{ and } (s, t)
 \in \llbracket \ell_2 \rrbracket\}
@@
 $$ q := \lift(\ell) \sep  q_1 \cdot q_2 \sep \swap(q_1, q_2) \sep (q_1 \sep q_2)
-\sep q^* \sep q_1 \; ; \; q_2 \sep \lquot(q, \ell) \sep \rquot(\ell, q)$$
-where $\ell$ ranges over bijective lenses.
+\sep q^* \sep q_1 \; ; \; q_2 \sep \lquot(q, Q) \sep \rquot(Q, q)$$
+where $Q$ ranges over QREs.
 
@@
 quotienting actually occurs: the $\lquot$ and $\rquot$ combinators. The
-$\lquot(Q, q)$ combinator takes a quotient lens $q$ and quotients the source
+$\lquot(Q, q)$ combinator takes a quotient lens $q$ and a QRE $Q$ and quotients the source
 data using $Q$, assuming that the source data forms a complete set of
-representatives for the equivalence relation $\eqrel{Q}$. The $\lquot(q, Q)$
+representatives for the equivalence relation $\eqrel{Q}$. The $\rquot(q, Q)$
 combinator does the same but on the target data.

### 2018-06-29 papers/SymmetricLenses/main.tex
add numbers and additional info

@@
 
-One of reasons these algorithms are so effective, even relative to successful synthesis projects
-in other domains, is because unless the same kind of data appears in a format many times,
-there simply are not very many different bijections between related data formats.  Hence,
-if the synthesis algorithm can find any bijection at all, then it is very likely
-to be the one the user wants.  However, there is also a price for narrowing the synthesis
-domain so dramatically:  Many real data formats are almost, but not quite, bijections.
-More often two related data formats have some data in common, but also contain some differences.
-For instance, two related tables might share four of five columns, but not the last.  Alternately, one ad hoc
-configuration
-file might host some format-specific meta data, such as a date, a time or reference number, whereas the other
-does not.  Indeed, in our study of Augeas~\cite{augeus}, a RedHat system that
-uses lenses to edit Linux files, we found that XXXXX of YYYYY lenses were not bijective.
-\dpw{fill in with accurate stats and/or more concrete (and real) compelling examples; I made up the above.
-mention that the POPL paper benchmarks all had to be edited?}
-Consequently, past tools that search for strict bijections will rarely be applicable in practice.
+One of reasons these algorithms are so effective, even relative to successful
+synthesis projects in other domains, is because unless the same kind of data
+appears in a format many times, there simply are not very many different
+bijections between related data formats. Hence, if the synthesis algorithm can
+find any bijection at all, then it is very likely to be the one the user wants.
+However, there is also a price for narrowing the synthesis domain so
+dramatically: Many real data formats are almost, but not quite, bijections. More
+often two related data formats have some data in common, but also contain some
+differences. For instance, two related tables might share four of five columns,
+but not the last. Alternately, one ad hoc configuration file might host some
+format-specific meta data, such as a date, a time or reference number, whereas
+the other does not. Indeed, all of our benchmarks taken from Flash
+Fill~\cite{flashfill} and all of our benchmarks that synchronize file formats
+contain data present in one format that isn't present in the other.
+Consequently, past tools that search for strict bijections will rarely be
+applicable in practice.
 
@@
 
-We implemented this\bcp{which?} algorithm and evaluated its effectiveness on a range of benchmarks
-from XXXX to YYYYY. We found that ... \dpw{give a number or two that characterize
-our empirical findings.}
+We implemented this\bcp{which?} algorithm and evaluated its effectiveness on a
+range of benchmarks, including data cleaning tasks taken from Flash Fill,
+information extraction tasks taken from Augeas, and ad-hoc format
+synchronization tasks taken from a variety of locations\footnote{The majority of
+  these format synchronization tasks were found from the following index:
+  \url{http://fileformats.archiveteam.org/wiki/Electronic_File_Formats}}.
 
@@
 show that the RE star semi-ring equivalences remain valid when their semantics are enriched by distributions.
-\item We extend the Boomerang implementation with simple symmetric lenses, showing that they can easily be
-combined with a rich variety of other lens extensions including quotients and matching lenses.  We evaluate
-our implementation over a set of benchmarks drawn from XXXXX and demonstrate that YYYY ... numbers ....
+\item We extend the Boomerang implementation with simple symmetric lenses,
+  showing that they can easily be combined with a rich variety of other lens
+  extensions including quotients and matching lenses. We evaluate our
+  implementation over a set of benchmarks drawn from data cleaning, information
+  extraction, and file synchronization tasks and demonstrate that we can
+  synthesize all of them in under half a minute.
 \end{itemize}

### 2018-06-30 papers/SymmetricLenses/main.tex
woo get in changes

@@
 ]
-0,20,40 cp -r /Library/ "$\$$(date)"
-
-* df -h | sendmail admin@domain.com
+Jane Doe: 35000
+John Public: 37500
 \end{lstlisting}
@@
 ]
-<Tasks>
- <Task>
-  <Command>c</Command>
-  <Time>0</Time><Time>20</Time> <Time>40</Time>
- </Task>
- <Task>
-  <Command>c</Command>
-  <AllTimes/>
- </Task>
-<Tasks>
+lastname,firstname,company
+Doe,Jane,Healthcare Inc.
+Public,John,Insurance Co.
 \end{lstlisting}
@@
 \begin{lstlisting}
-let minute_list = minute . ("," . minute)$^*$ in
-let times = "*" | minute_list in
-let line = times . " " . command in
-let crontab = "" | line . ("\n\n" line)*
+let salary = number | "UNSET"
+let emp_salary = name . " " . name . ": " salary
+let emp_salaries = "" | emp_salary . ("\n" emp_salary)*
 \end{lstlisting}
@@
 \begin{lstlisting}
-let minute_xml = "<Time>" . minute . "</Time>" in
-let minutes_xml = minute_xml . minute_xml$^*$ in
-let xml_time = "<AllTimes/>" | minutes_xml in
-let command_xml = 
-  "<Command>" . command . "</Command>"
-in
-let xml_task = command_xml . xml_time in
-let scheduled_job = "<Tasks>" . xml_task$^*$ . "</Tasks>"
+let company = (name . ("Co." | "Inc." | "Ltd.")) | "UNSET" in
+let emp_row = name . "," . name "," company in
+let header = "firstname,lastname,company" in
+let emp_insurances = header . ("\n" . emp_row)*
 \end{lstlisting}
@@
       \hline
-      crontab & scheduled.job \\
+      company.mgmt & company.hr \\
       \hline
@@
   }
-  \caption{An example crontab and Scheduled Tasks file, and regular expressions
-    describing the formats. For simplicity, these are modified versions of the
-    real formats.}
+  \caption{Example files used by management and HR, and regular expressions
+    describing the formats of these files.}
   \label{fig:minimized-representations}
@@
 
-We motivate our problem with an extended example. Consider the previously
-described goal of keeping Linux cron jobs (crontab) and Windows Scheduled Tasks
-(scheduled.job) synchronized. These are each formats specifying when to run
-recurring tasks. For the sake of exposition, we use simplified versions crontab
-and scheduled.job formats, with examples and corresponding regular expressions
-shown in Figure~\ref{fig:minimized-representations}. Each file contains a list
-of tasks set to run at specific times every hour. In both formats, users
-can either provide specific minutes at which they want the program to run, or they
-can specify that the program should run at every minute.
-
-We wish for the jobs to be synchronized across our Windows and Linux devices.
-However, the commands that need to be run are typically different for each
-device. Paths are different in Windows and Linux, and, while there is typically a
-Windows equvialent for a Linux command, that equivalent command is not always
-named the same thing. So, while we want to be able to have the same number of
-commands, with the commands running at the same times, we don't want to
-align the actual commands that are run.
+\afm{i commented out the previous version, because this one was quickly written
+  up to give you something to work off. Feel free to delete it if it isnt
+  helpful. } We motivate our problem with an extended example. In our
+theoretical company, human resources and upper management both keep information
+about employees. HR needs to know the names of the employees as well as their
+insurance providers. Management needs to know the names of employees and their
+salary. Two example files, and regular expressions describing the formats, are
+shown in Figure~\ref{fig:minimized-representations}. These two files should be
+synchronized: if management adds a new employee, ``Richard Roe: 32500'', a new
+row should be added to HR's file, ``Roe,Richard,UNSET''. Furthermore, if HR
+changes ``John Public'' to ``Jon Public'', that name change should be reflected
+in both files. However, not all the data should be synchronized. The HR file
+should not get altered due to salary changes, and the management file should not
+get altered due to insurance company changes. While names should be matched
+across the two formats, insurance company and salary information should be left
+unmatched. Such a synchronization can be achieved with simple symmetric lenses.
+
+%Consider the previously
+%described goal of keeping Linux cron jobs (crontab) and Windows Scheduled Tasks
+%(scheduled.job) synchronized. These are each formats specifying when to run
+%recurring tasks. For the sake of exposition, we use simplified versions crontab
+%and scheduled.job formats, with examples and corresponding regular expressions
+%shown in Figure~\ref{fig:minimized-representations}. Each file contains a list
+%of tasks set to run at specific times every hour. In both formats, users
+%can either provide specific minutes at which they want the program to run, or they
+%can specify that the program should run at every minute.
+%
+%We wish for the jobs to be synchronized across our Windows and Linux devices.
+%However, the commands that need to be run are typically different for each
+%device. Paths are different in Windows and Linux, and, while there is typically a
+%Windows equvialent for a Linux command, that equivalent command is not always
+%named the same thing. So, while we want to be able to have the same number of
+%commands, with the commands running at the same times, we don't want to
+%align the actual commands that are run.
 
@@
 and two \emph{put} functions. Create functions are used to initialize data in
-one format when provided with data in the other. If a crontab is initialized, a
-synchronized Scheduled Tasks file should also be created, and vice versa. Put
-functions are used to propagate edits from one format to the other. If a crontab
-is updated, we should be able to combine it with a non-updated Scheduled Tasks
-file, and output a new Scheduled Tasks file, synchronized to the updated
-crontab. To have a simple symmetric lens is to have these required
-synchronization functions. Formally, a simple symmetric lens between $X$ and $Y$
-comprises 4 functions:
+one format when provided with data in the other. If a company.mgmt file is
+initialized, a synchronized company.hr file should also be created, and vice
+versa. Put functions are used to propagate edits from one format to the other.
+If a company.hr file is updated, we should be able to combine it with a
+non-updated company.mgmt file, and output a new company.mgmt file, synchronized
+to the updated company.hr. To have a simple symmetric lens is to have these
+required synchronization functions. Formally, a simple symmetric lens between
+$X$ and $Y$ comprises 4 functions:
 \begin{enumerate}
@@
 
-For synchronization, when the original crontab or Scheduled Tasks file is
+For synchronization, when the original company.hr or company.mgmt file is
 created, a synchronized file of the other form is created from it with
@@
 future edits, changes are made with the puts, a \PutR would propagate the shared
-information from the crontab, and knit it together with the information specific
-to the Scheduled Tasks file.
+information from the company.hr file, and knit it together with the information
+specific to the company.mgmt file.
 
@@
 $\LanguageOf{\RegexAlt}$. At the end, we want our symmetric lenses to by typed
-between \lstinline{crontab} and \lstinline{scheduled_job}.
+between \lstinline{emp_salaries} and \lstinline{emp_insurance}.  The finished
+lens is shown in Figure~\ref{fig:example_lens}.
 
@@
 identity function on strings in that regular expression. The identity lens fully
-aligns the data formats: when one data format is updated, the other format gets
-the exact same value. In our example, when a job is told to run at a specific
-time in crontab, it is also expected to run at that same time in the Scheduled
-Jobs list. The identity lens propagates that information.
+matches the data formats: when one data format is updated, the other format gets
+the exact same value. In our example, when a name is updated, it is also
+expected to updated in the other format. The identity lens propagates that
+information.
 %
 \begin{lstlisting}
-id(minute) : minute $\Leftrightarrow$ minute
+id(name) : name $\Leftrightarrow$ name
 \end{lstlisting}
@@
 data in the first format and data in the second format, designated default
-strings are used for the other format. In our extended example, the Linux
-commands should not be synchronized with the Windows commands, so a disconnect
-lens would be used. We want our default behavior to be failures, under the
-assumption this failure will raise a warning, in case somebody forgets to find
-the analogous command of the other operating system.
+strings are used for the other format. In our extended example, salaries and
+insurance companies should not be matched, so a disconnect lens would be used.
+We want our default behavior to specify that the value has not been set.
 %
 \begin{lstlisting}
-disc(command,command_xml,"false","<Command>SET ERRORLEVEL=1</Command>") 
-  : command $\Leftrightarrow$ command_xml
+disc(salary,company,"UNSET","UNSET") : salary $\Leftrightarrow$ company
 \end{lstlisting}
@@
 \begin{lstlisting}
-or(minutes_map,alltimes_map) : times $\Leftrightarrow$ xml_time
+or(id(""),multi_emp_map) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_row) . ("\n" emp_row)*
 \end{lstlisting}
 
-Data can also be concatenated, and so lens mappings can be concatenated.
-However, synchronizing concatenated data can happen in a number of ways, as data
-can appear in the same order, and data can be reordered. For example, in the
-crontab, time data appears before command data, whereas in the Scheduled Jobs
-file, command data comes before time data. If \lstinline{command_map} provides a
-lens between commands in each format, and \lstinline{time_map} provides a lens
-between time data in each format, those lenses can be composed with a \SwapLens
-lens.
+Data can also be concatenated, and so lens mappings can be concatenated.  This
+is achieved with the concat map.
 %
 \begin{lstlisting}
-swap(command_map,time_map) : line $\Leftrightarrow$ xml_task
+concat(ins(header),multi_emp_map) : emp_salaries $\Leftrightarrow$ emp_insurance
 \end{lstlisting}
@@
 \lstinline{concat($\Lens_1$,$\Lens_2$)} or as the infix style
-\lstinline{$\Lens_1$ . $\Lens_2$} for data that is not reordered.
+\lstinline{$\Lens_1$ . $\Lens_2$}.
 
@@
 argument and synchronizes each element of the iteration with that lens. For
-example, if \lstinline{full_task_map} synchronizes crontab lines prepended by
-two newlines and xml for a single task, iterating that lens builds a lens that
-synchronizes a list of tasks.
+example, if \lstinline{single_emp_map} synchronizes a single employee's data,
+iterating that lens builds a lens that synchronizes a list of tasks.
 %
 \begin{lstlisting}
-iterate(full_task_map) : ("\n\n" . line)$^*$ $\Leftrightarrow$ xml_task$^*$
+iterate(id("\n") . single_emp_map) : ("\n" . emp_salary)$^*$ $\Leftrightarrow$ ("\n" . emp_row)$^*$
 \end{lstlisting}
@@
 \begin{lstlisting}
-let full_minute_map = ins("<Time>") . id(minute) . ins("</Time>") in
-let comma_minute_map = del(",") . full_minute_map in
-let minutes_map = full_minute_map . iterate(comma_minute_map) in
-let alltimes_map = del("*") . ins("<AllTimes/>") in
-let times_map = or(minutes_map,alltimes_map) in
-
-let command_core = 
-  disc(command,command_xml,"false","<Command>SET ERRORLEVEL=1</Command>")
-in
-let command_map = 
-  ins("<Command>") . command_core . ins("</Command>")) . del(" ") 
-in
-
-let task_map = swap(command_map,time_map) in
-let full_task_map = 
-  ins("<Task>") . del("\n\n") . task_map . ins("</Task>") 
-in
-let repeated_task_map = full_task_map . iterate(full_task_map)
-let nonempty_task_map = ins("<Tasks>") . repeated_task_map . ins("</Tasks>")
-
-or(ins("<Tasks></Tasks>"),nonempty_task_map) : crontab $\Leftrightarrow$ scheduled_job
+let name_map = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
+let single_emp_map = name_map . disc(salary,company,"UNSET","UNSET") in
+let multi_emp_map = ins("\n") . single_emp_map . iterate(id("\n") . single_emp_map) in
+let emp_map = or(id(""),multi_emp_map) in
+ins(header) . multi_emp_map : emp_salaries $\Leftrightarrow$ emp_insurance
 \end{lstlisting}
   \caption{A lens, written in our language for building simple symmetric lenses
-    on string data, that synchronizes crontab files and Scheduled Task files.
+    on string data, that synchronizes management and hr employee files.
     Note that we can annotate our lenses with types, to confirm the lens
     operates on the languages of the regular expressions. In this work, we
-    synthesize the lens itself from such regular expressions. \saz{Say something
-      about the possibility of synthesizing this lens?}}
+    synthesize the lens itself from such regular expressions.}
   \label{fig:example_lens}
@@
 
-Using these combinators, we can build up a lens that synchronizes crontabs and
-Scheduled Task files, shown in Figure~\ref{fig:example_lens}. This lens uses one
-typing rule not previously mentioned: type equivalence. While
-\lstinline{scheduled_job} is a equivalent to a regular expression with an
-outermost disjunction, it is not syntactically equivalent to one. Without type
-equivalence, the provided lens wouldn't actually be well-typed, but type
-equivalence states that if a lens is typed between a pair of regular
-expressions, it is also typed between an equivalent pair.
+Using these combinators, we can build up a lens that synchronizes HR and
+management employee files. This lens uses one typing rule not previously
+mentioned: type equivalence. While \lstinline{scheduled_job} is a equivalent to
+a regular expression with an outermost disjunction, it is not syntactically
+equivalent to one. Without type equivalence, the provided lens wouldn't actually
+be well-typed, but type equivalence states that if a lens is typed between a
+pair of regular expressions, it is also typed between an equivalent pair.
 
@@
 pair, and satisfies the input/output examples. In this example, we wish to
-synthesize a lens from \lstinline{crontab} and \lstinline{scheduled_job}, using
-the strings in Figure~\ref{fig:minimized-representations} as an input-output
-example for \CreateR.
+synthesize a lens from \lstinline{emp_salaries} and \lstinline{emp_insurance},
+using the strings in Figure~\ref{fig:minimized-representations} as an
+input-output example for \CreateR.
 
@@
 between two regular expressions. For example, one possible lens between
-\lstinline{crontab} and \lstinline{scheduled_job} is the one in
+\lstinline{emp_salaries} and \lstinline{emp_insurance} is the one in
 Figure~\ref{fig:example_lens}, but the lens
@@
 \begin{lstlisting}
-disc(crontab,scheduled_job,rep(crontab),rep(scheduled_job))
+disc(emp_salaries,emp_insurance,rep(emp_salaries),rep(emp_insurance))
 \end{lstlisting}
@@
 
-Now that we have a notion of best, we need a way to
-search for that best lens.
+Now that we have a notion of best, we need a way to search for that best lens.
 Our overall approach involves two communicating synthesizers: one (\RXSearch)
@@
 Figure~\ref{fig:example_lens}. This lens would \emph{not} be in the class
-characterized by \lstinline{crontab} and \lstinline{scheduled_job}, as that
-typing derivation requires unrolling \lstinline{xml_task$^*$} within
-\lstinline{scheduled_job} into \lstinline{"" | (xml_task . xml_task$^*$)}.
+characterized by \lstinline{emp_salaries} and \lstinline{emp_insurance}, as that
+typing derivation requires unrolling \lstinline{("\n" . emp_salary)$^*$} within
+\lstinline{emp_salaries} into \lstinline{"" | (("\n" . emp_salary) . ("\n" . emp_salary)$^*$)}.
 
@@
 lens. Moreover, we may find a great lens characterized by
-\lstinline{crontab} and \lstinline{scheduled_job} that is not the
+\lstinline{emp_salaries} and \lstinline{emp_insurances} that is not the
 desired lens shown in Figure~\ref{fig:example_lens}.  The
@@
     $\CreateR{} \App \String'$ & = & $\StringAlt$\\
-    $\CreateL{} \App \StringAlt'$ & = & $\String$\\
-    $\PutR{} \App \String' \App \StringAlt'$ & = & $\StringAlt'$\\
+    $\CreateL{} \App \StringAlt'''''$ & = & $\String$\\
+    $\PutR{} \App \String'' \App \StringAlt'$ & = & $\StringAlt'$\\
     $\PutL{} \App \StringAlt' \App \String'$ & = & $\String'$

### 2018-06-30 papers/SymmetricLenses/main.tex
add example of entropy calcs

@@
 \usepackage{listings}
+\usepackage{letltxmacro}
 \usepackage{sansmath}
@@
 which users can refine with additional examples.
+
+\afm{calculations for cost of various lenses between \lstinline{emp_salary} and
+  \lstinline{emp_row}} Consider the lens \lstinline{single_emp_map} between
+\lstinline{emp_salary} and \lstinline{emp_row}. Given this lens, the expected
+entropy of recovering the left format from the right
+($\EntropyOf{\lstinline{emp_salary} \Given \lstinline{emp_row},
+  \lstinline{single_emp_map} }$) is the sum of the expected entropy of
+recovering each of its subcomponents, $\EntropyOf{\lstinline{emp_salary} \Given
+  \lstinline{emp_row}, \lstinline{single_emp_map} } = \EntropyOf{\lstinline{name
+    . " " . name} \Given \lstinline{name . "," . name}, \lstinline{name_map}
+}$\\ + $\EntropyOf{\lstinline{": " . salary} \Given
+  \lstinline{"," . company}, \lstinline{disc(salary,company,"UNSET","UNSET")}
+}$.
+
+The entropy of the identity lens is zero. As the two formats are in bijective
+correspondence, we know exactly which string is present in the left-hand format.
+The entropy of the disconnect lens is just the entropy of
+\lstinline{": " . salary}.  Knowing which string is present in the right-hand
+format tells us nothing about what string is present in the left-hand format. So
+$\EntropyOf{\lstinline{emp_salary} \Given \lstinline{emp_row},
+  \lstinline{single_emp_map} } = \EntropyOf{\lstinline{": " . salary} }$
+
+Analogously, $\EntropyOf{\lstinline{emp_row} \Given \lstinline{emp_salary},
+  \lstinline{single_emp_map} }$ is just $\EntropyOf{\lstinline{"," . company}
+}$, so the total cost of the lens is $\EntropyOf{\lstinline{": " . salary}} +
+\EntropyOf{\lstinline{"," . company}}$
+
+Consider the trivial lens
+\lstinline{disc(emp_salary,emp_row,rep(emp_salary),rep(emp_row))}. As this is
+the disconnect lens, the cost of each side is the entropy of the data format, so
+the total cost is $\EntropyOf{\lstinline{emp_salary}} +
+\EntropyOf{\lstinline{emp_row}}}$. \lstinline{emp_salary} contains strictly more
+imformation than \EntropyOf{\lstinline{": " . salary}}, and \lstinline{emp_row}
+contains strictly more information than \EntropyOf{\lstinline{": " . salary}},
+so by our metric \lstinline{single_emp_map} is preferable to the disconnect
+lens.
 % end overview

### 2018-06-30 papers/SymmetricLenses/main.tex
edits in progress to section 2

@@
 ]
-Jane Doe: 35000
+Jane Doe: 38000
 John Public: 37500
@@
 ]
-lastname,firstname,company
-Doe,Jane,Healthcare Inc.
-Public,John,Insurance Co.
+Name,Company
+Jane Doe,Healthcare Inc.
+John Public,Insurance Co.
 \end{lstlisting}
@@
 \begin{lstlisting}
-let salary = number | "UNSET"
+let salary = number | "N/A"
 let emp_salary = name . " " . name . ": " salary
@@
 \begin{lstlisting}
-let company = (name . ("Co." | "Inc." | "Ltd.")) | "UNSET" in
-let emp_row = name . "," . name "," company in
-let header = "firstname,lastname,company" in
-let emp_insurances = header . ("\n" . emp_row)*
+let company = (name . ("Co." | "Inc." | "Ltd.")) | "UNK" in
+let emp_ins = name . " " . name "," company in
+let header = "Name,Company" in
+let emp_insurance = header . ("\n" . emp_ins)*
 \end{lstlisting}
@@
       \hline
-      company.mgmt & company.hr \\
+      \texttt{company.mgmt} data file & \texttt{company.hr} data file \\
       \hline
@@
       \hline
-      \listingC & \listingD \\
+      \cellcolor{yellow}\texttt{company.mgmt} regular expression & 
+      \cellcolor{green}\texttt{company.hr}  regular expression\\
+      \hline
+      \cellcolor{yellow}\listingC & \cellcolor{green}\listingD \\
       \hline
@@
   }
-  \caption{Example files used by management and HR, and regular expressions
-    describing the formats of these files.}
+  \caption{Example data files and corresponding regular expressions
+  used by management and HR, respectively, to represent employee salaries and health
+  insurance companies.  
+   \ksf{fix colors so regular expressions are just slightly darker
+  shade than data file.}}  
   \label{fig:minimized-representations}
@@
 
-\afm{i commented out the previous version, because this one was quickly written
-  up to give you something to work off. Feel free to delete it if it isnt
-  helpful. } We motivate our problem with an extended example. In our
-theoretical company, human resources and upper management both keep information
-about employees. HR needs to know the names of the employees as well as their
-insurance providers. Management needs to know the names of employees and their
-salary. Two example files, and regular expressions describing the formats, are
-shown in Figure~\ref{fig:minimized-representations}. These two files should be
-synchronized: if management adds a new employee, ``Richard Roe: 32500'', a new
-row should be added to HR's file, ``Roe,Richard,UNSET''. Furthermore, if HR
+In this section, we give an overview of simple symmetric lenses and
+our corresponding synthesis algorithm using an example drawn from a
+hypothetical U.S. company.  In this company, human resources and upper
+management store information about employees in separate files:
+Management records the names of employees and their salaries while
+Human Resources (HR) stores the names of employees and their health
+insurance providers.  Figure~\ref{fig:minimized-representations} show
+two example files and regular expressions describing the formats.
+The company wants to keep these two files synchronized:
+if management adds a new employee, ``Chris Roe: 32500'', 
+the corresponding entry ``Chris Roe, UNK'' should be added to HR's file.
+
+Furthermore, if HR
 changes ``John Public'' to ``Jon Public'', that name change should be reflected

### 2018-06-30 papers/SymmetricLenses/main.tex
fixed typo
in progress edits to section 2

@@
 our corresponding synthesis algorithm using an example drawn from a
-hypothetical U.S. company.  In this company, human resources and upper
-management store information about employees in separate files:
-Management records the names of employees and their salaries while
-Human Resources (HR) stores the names of employees and their health
-insurance providers.  Figure~\ref{fig:minimized-representations} show
-two example files and regular expressions describing the formats.
-The company wants to keep these two files synchronized:
-if management adds a new employee, ``Chris Roe: 32500'', 
-the corresponding entry ``Chris Roe, UNK'' should be added to HR's file.
-
+hypothetical U.S. company.  In this company, management and
+human resources (HR) store information about employees in separate files:
+management records the names of employees and their salaries while
+HR stores the names of employees and their health
+insurance providers.  Figure~\ref{fig:minimized-representations} gives
+examples of the two types of files and regular expressions describing
+the two formats.
+The company uses simple symmetric lenses to keep these files synchronized.
+If management adds a new employee, ``Chris Roe: 32500'',  a lens adds 
+the corresponding entry ``Chris Roe, UNK'' to HR's file.  The
+sentinel value ``UNK'' represents the fact that the employee's
+insurance company is currently unknown.  A similar update happens if
+HR adds a new employee before management, in which case the sentinel
+value ``N/A'' reflects the unknown salary information. 
 Furthermore, if HR
-changes ``John Public'' to ``Jon Public'', that name change should be reflected
-in both files. However, not all the data should be synchronized. The HR file
-should not get altered due to salary changes, and the management file should not
-get altered due to insurance company changes. While names should be matched
-across the two formats, insurance company and salary information should be left
-unmatched. Such a synchronization can be achieved with simple symmetric lenses.
+corrects an error in an employee's name, say changing ``John Public''
+to ``Jon Public'', the lens mirrors that name change in management's
+file. Not all the data is mirrored, however. The management file 
+is not updated in response to insurance changes, and the HR file
+is oblivious to salary changes.  Simple symmetric lenses are
+appropriate for keeping these two files synchronized because they have
+a mix of shared and unshared information.
 
@@
 
-Lenses synchronize files by providing 4 functions, two \emph{create} functions,
+Simple symmetric lenses synchronize files using four functions,
+two \emph{create} functions, 
 and two \emph{put} functions. Create functions are used to initialize data in
@@
 the total cost is $\EntropyOf{\lstinline{emp_salary}} +
-\EntropyOf{\lstinline{emp_row}}}$. \lstinline{emp_salary} contains strictly more
+\EntropyOf{\lstinline{emp_row}}$. \lstinline{emp_salary} contains strictly more
 imformation than \EntropyOf{\lstinline{": " . salary}}, and \lstinline{emp_row}

### 2018-06-30 papers/SymmetricLenses/main.tex
in progress edits to section 2

@@
 
-Simple symmetric lenses synchronize files using four functions,
-two \emph{create} functions, 
-and two \emph{put} functions. Create functions are used to initialize data in
-one format when provided with data in the other. If a company.mgmt file is
-initialized, a synchronized company.hr file should also be created, and vice
-versa. Put functions are used to propagate edits from one format to the other.
-If a company.hr file is updated, we should be able to combine it with a
-non-updated company.mgmt file, and output a new company.mgmt file, synchronized
-to the updated company.hr. To have a simple symmetric lens is to have these
-required synchronization functions. Formally, a simple symmetric lens between
-$X$ and $Y$ comprises 4 functions:
+\subsection{Simple Symmetric Lenses}
+A simple symmetric lens between formats described by types $\Regex$
+and $\RegexAlt$ 
+is defined by four functions: 
 \begin{enumerate}
-\item $\CreateR \OfType X \rightarrow Y$
-\item $\CreateL \OfType Y \rightarrow X$
-\item $\PutR \OfType X \rightarrow Y \rightarrow Y$
-\item $\PutL \OfType Y \rightarrow X \rightarrow X$
+\item $\CreateR \OfType \Regex \rightarrow \RegexAlt$
+\item $\CreateL \OfType \RegexAlt \rightarrow \Regex$
+\item $\PutR \OfType \Regex \rightarrow \RegexAlt \rightarrow \RegexAlt$
+\item $\PutL \OfType \RegexAlt \rightarrow \Regex \rightarrow \Regex$
 \end{enumerate}
-Furthermore, if a file is not edited, but is then resynchronized using a put
-function, no data should be changed. To be a simple symmetric lens, one must
-satisfy the following ``round-tripping'' laws, guaranteeing identity edits won't
-change the file data.
+\noindent
+that collectively satisfy four ``round-tripping'' laws: 
 \begin{gather}
   \tag{\CreatePutRL}
-  \PutLOf{(\CreateROf{x})}{x} = x\\
+  \PutLOf{(\CreateROf{s})}{s} = s\\
   \tag{\CreatePutLR}
-  \PutROf{(\CreateLOf{y})}{y} = y\\
+  \PutROf{(\CreateLOf{t})}{t} = t\\
   \tag{\PutRL}
-  \PutLOf{(\PutROf{x}{y})}{x} = x\\
+  \PutLOf{(\PutROf{s}{t})}{s} = s\\
   \tag{\PutLR}
-  \PutROf{(\PutLOf{y}{x})}{y} = y
+  \PutROf{(\PutLOf{t}{s})}{t} = t
 \end{gather}
-
-For synchronization, when the original company.hr or company.mgmt file is
-created, a synchronized file of the other form is created from it with
-$\CreateR$ or $\CreateL$ (depending on which file was initially created). On
-future edits, changes are made with the puts, a \PutR would propagate the shared
-information from the company.hr file, and knit it together with the information
-specific to the company.mgmt file.
-
-\subsection{Programming Symmetric Lenses} We now build up a symmetric lens by
-example. Our symmetric lens expressions are typed between regular expression
-pairs, if \lstinline{$\Lens$ : $\Regex \Leftrightarrow \RegexAlt$}, then $\Lens$
-encodes a simple symmetric lens between $\LanguageOf{\Regex}$ and
-$\LanguageOf{\RegexAlt}$. At the end, we want our symmetric lenses to by typed
-between \lstinline{emp_salaries} and \lstinline{emp_insurance}.  The finished
-lens is shown in Figure~\ref{fig:example_lens}.
-
-The identity lens takes, as argument, a regular expression, and performs the
+\noindent
+The two \emph{create} functions provide default values for times when
+data appears on one side but not the other (eg, the ``N/A'' salary
+entry added to the management file when HR inserts a new employee).
+The two \emph{put} functions propagate edits from one format to the
+other by combining the change from one file with the previous value
+from the other file. The roundtripping laws guarantee that identify edits to one file do
+not change the data in the other file.
+
+%%For synchronization, when the original company.hr or company.mgmt
+%%file is created, a synchronized file of the other form is created
+%%from it with $\CreateR$ or $\CreateL$ (depending on which file was
+%%initially created). On future edits, changes are made with the puts,
+%%a \PutR would propagate the shared information from the company.hr
+%%file, and knit it together with the information specific to the company.mgmt file.
+
+We use regular expressions to denote the types of data formats, and we
+impose a typing discipline on our lenses.  Specifically, if 
+$\Regex$ and $\RegexAlt$ are regular expressions, then the notation
+\lstinline{$\Lens$ : $\Regex \Leftrightarrow \RegexAlt$} indicates that $\Lens$
+is a simple symmetric lens between formats described by $\LanguageOf{\Regex}$ and
+$\LanguageOf{\RegexAlt}$. 
+When clear from context, we use the term \textit{lens} to denote a
+simple symmetric lens.
+
+\subsection{Simple Symmetric Lens Combinators} 
+\ksf{This section needs more polishing}
+Generally, we do not expect programmers to define simple symmetric lenses by
+writing the four functions by hand and showing they satisfy the
+roundtripping laws.  Instead, we provide simple symmetric lens
+combinators to build up the requisite lenses from simpler ones.
+We explain these combinators by using them to define a lens between
+the two employee data formats in Figure~\ref{fig:minimized-representations}. 
+The finished lens, which appears in Figure~\ref{fig:example_lens}, has
+the type 
+\lstinline{emp_salaries} $\Leftrightarrow$ \lstinline{emp_insurance}, where
+\lstinline{emp_salaries} and \lstinline{emp_insurance} are regular
+expressions defined in Figure~\ref{fig:minimized-representations}.
+
+The simplest combinator is the identity lens, which takes as an argument a
+regular expression, and performs the 
 identity function on strings in that regular expression. The identity lens fully
@@
 the exact same value. In our example, when a name is updated, it is also
-expected to updated in the other format. The identity lens propagates that
+expected to be updated in the other format. The identity lens propagates that
 information.
@@
 
-Where the identity lens aligns data, the disconnect lens does the opposite: it
-leaves data unaligned. The disconnect lens takes 4 arguments: two regular
-expressions and two strings. On creates, because there is no association between
-data in the first format and data in the second format, designated default
-strings are used for the other format. In our extended example, salaries and
-insurance companies should not be matched, so a disconnect lens would be used.
-We want our default behavior to specify that the value has not been set.
+In contrast to the identity lens, the disconnect lens leaves data
+unaligned. The disconnect lens takes four arguments: two regular 
+expressions and two strings. The regular expressions specify the
+formats of the data on the two sides while the strings provide default
+values.  In our extended example, salaries and
+insurance companies unrelated, so we use a disconnect lens.  The
+defaults are strings that indicate the information is missing.
 %
 \begin{lstlisting}
-disc(salary,company,"UNSET","UNSET") : salary $\Leftrightarrow$ company
+disc(salary,company,"N/A","UNK") : salary $\Leftrightarrow$ company
 \end{lstlisting}
@@
 Finally, there are a number of lenses that compositionally build up new lenses
-from old ones. For example, data can appear in one form or another: a program
-can run during a specific set of minutes, or that program can run the every
-minute. If \lstinline{minutes_map} and \lstinline{alltimes_map} are lenses that
-synchronize those settings in each format, then those lenses can be composed
-with an \OrLens lens.
+from old ones. 
+For example, data can be concatenated, and so we provide
+the \lstinline{concat} combinator that sequences lenses, usually
+written using an infix notation \lstinline{$\Lens_1$ . $\Lens_2$}:
 %
 \begin{lstlisting}
-or(id(""),multi_emp_map) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_row) . ("\n" emp_row)*
+id(name) . ins(",") . id(name) . del(": ") . ins(",")
 \end{lstlisting}
-
-Data can also be concatenated, and so lens mappings can be concatenated.  This
-is achieved with the concat map.
+% concat(ins(header),employees_lens) : emp_salaries $\Leftrightarrow$ emp_insurance
+%
+
+Data can also appear in multiple forms. For example,
+a list of employees with insurance information can either be empty or
+be a sequence of employee/insurance records: \ksf{This example is
+vacuous as you could eliminate the or in favor of the *.  Do somethign
+else to introduce or?}
 %
 \begin{lstlisting}
-concat(ins(header),multi_emp_map) : emp_salaries $\Leftrightarrow$ emp_insurance
+or(id(""),employees_lens) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_ins) . ("\n" emp_ins)*
 \end{lstlisting}
-%
-Two lenses, $\Lens_1$ and $\Lens_2$, can be concatenated as
-\lstinline{concat($\Lens_1$,$\Lens_2$)} or as the infix style
-\lstinline{$\Lens_1$ . $\Lens_2$}.
 
@@
 argument and synchronizes each element of the iteration with that lens. For
-example, if \lstinline{single_emp_map} synchronizes a single employee's data,
-iterating that lens builds a lens that synchronizes a list of tasks.
+example, if \lstinline{employee_lens} synchronizes a single employee's data,
+iterating that lens builds a lens that synchronizes a list of employees.
 %
 \begin{lstlisting}
-iterate(id("\n") . single_emp_map) : ("\n" . emp_salary)$^*$ $\Leftrightarrow$ ("\n" . emp_row)$^*$
+iterate(id("\n") . employee_lens) : ("\n" . emp_salary)$^*$ $\Leftrightarrow$ ("\n" . emp_ins)$^*$
 \end{lstlisting}
@@
 \begin{lstlisting}
-let name_map = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
-let single_emp_map = name_map . disc(salary,company,"UNSET","UNSET") in
-let multi_emp_map = ins("\n") . single_emp_map . iterate(id("\n") . single_emp_map) in
-let emp_map = or(id(""),multi_emp_map) in
-ins(header) . multi_emp_map : emp_salaries $\Leftrightarrow$ emp_insurance
+let name_lens = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
+let employee_lens = name_lens . disc(salary,company,"N/A","UNK") in
+let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
+let emp_lens = or(id(""),employees_lens) in
+ins(header) . employees_lens : emp_salaries $\Leftrightarrow$ emp_insurance
 \end{lstlisting}
   \caption{A lens, written in our language for building simple symmetric lenses
-    on string data, that synchronizes management and hr employee files.
+    on string data, that synchronizes management and HR employee files.
     Note that we can annotate our lenses with types, to confirm the lens
@@
 
-Using these combinators, we can build up a lens that synchronizes HR and
-management employee files. This lens uses one typing rule not previously
+Using these combinators, we can build up a lens that synchronizes
+management and HR employee files.
+
+\ksf{This paragraph needs to be updated, if nothing else, to reflect
+the new example. Also, we haven't mentioned any typing rules so the
+first sentence doesn't make sense.  Perhaps the whole topic could be
+defered to later where we can explain why such a rule is necessary?}
+This lens uses one typing rule not previously
 mentioned: type equivalence. While \lstinline{scheduled_job} is a equivalent to
@@
 
-Note that, while the programmed lens should be quite simple, it was relatively
-difficult to write. We had to provide a lot of scaffolding to map between the
-desired formats, even though the idea was quite simple: align the times but
-leave the commands unaligned. Furthermore, this is inherent to manually writing
-these synchronization functions as a programmer must describe which components
-of the strings are aligned with which, and which components are scaffolding.
+Note that, while the synchronization task is quite simple, the
+requisite lens is relatively complex. Furthermore, this complexity is
+inherent in manually writing simple symmetric lenses because the
+lenses must precisely describe the portions of the data that should be
+connected and those that should be disconnected.  For disconnected
+pieces, the lens must further specifiy appropriate defaults.
 

### 2018-06-30 papers/SymmetricLenses/main.tex
edits in progress

@@
       \hline
-      \cellcolor{yellow}\texttt{company.mgmt} regular expression & 
-      \cellcolor{green}\texttt{company.hr}  regular expression\\
+      \cellcolor{vlightyellow}\texttt{company.mgmt} regular expression & 
+      \cellcolor{vlightgreen}\texttt{company.hr}  regular expression\\
       \hline
-      \cellcolor{yellow}\listingC & \cellcolor{green}\listingD \\
+      \cellcolor{vlightyellow}\listingC & \cellcolor{vlightgreen}\listingD \\
       \hline
@@
   used by management and HR, respectively, to represent employee salaries and health
-  insurance companies.  
-   \ksf{fix colors so regular expressions are just slightly darker
-  shade than data file.}}  
+  insurance companies.  }  
   \label{fig:minimized-representations}
@@
 \lstinline{$\Lens$ : $\Regex \Leftrightarrow \RegexAlt$} indicates that $\Lens$
-is a simple symmetric lens between formats described by $\LanguageOf{\Regex}$ and
+is a simple symmetric lens between $\LanguageOf{\Regex}$ and
 $\LanguageOf{\RegexAlt}$. 
@@
 find a symmetric lens expression that is typed between the regular expression
-pair, and satisfies the input/output examples. In this example, we wish to
+pair, and satisfies the input/output examples. We call such a lens
+a \emph{satisfying lens}.
+In this example, we wish to
 synthesize a lens from \lstinline{emp_salaries} and \lstinline{emp_insurance},
-using the strings in Figure~\ref{fig:minimized-representations} as an
-input-output example for \CreateR.
-
-However, our simple symmetric lens combinators permit many well-typed programs
-between two regular expressions. For example, one possible lens between
-\lstinline{emp_salaries} and \lstinline{emp_insurance} is the one in
-Figure~\ref{fig:example_lens}, but the lens
+using the example data in Figure~\ref{fig:minimized-representations} as an
+input-output example from which to draw the required defaut values for 
+\Disconnect lenses.
+
+A challenge is that the simple symmetric lens combinators permit many
+well-typed lenses
+between a given pair of regular expressions. For example, 
+Figure~\ref{fig:example_lens} gives one possible lens between regular
+expressions  \lstinline{emp_salaries} and \lstinline{emp_insurance},
+but 
 %
@@
 %
-is also a valid lens (where \lstinline{rep($\Regex$)} is a string in the
+is another valid lens (where \lstinline{rep($\Regex$)} is a string in the
 language of $\Regex$). Furthermore, adding the synchronized files from
@@
 chosen input-output example on \PutR or \PutL, this problem continues presenting
-itself for the sublenses, and is highlighted even more on larger formats. While
-we could merely ask programmers to provide large, tight specifications for their
+itself for the sublenses, and is highlighted even more on larger
+formats. 
+\ksf{I don't understand the above.  We shouldn't explain the issue in
+terms of \PutR and \PutL as we are focused on the combinators rather
+than the raw functions.}
+While
+we could ask programmers to provide detailed and precise specifications for their
 synthesis tasks, we would prefer the algorithm to choose the right lens
@@
 
-But what is the best satisfying lens? Here, we take the view that the best satisfying lens is the
-one that aligns as much data as is permitted by the type system and the
-examples. More formally, we wish to minimize the expected number of bits
-required to recover data in one format from a synchronized string in the other.
-The expected number of bits is a well known concept in information theory: the
-\emph{entropy} of that data source.
-
-The entropy of a data source is based on the probability distribution of that
-data source. Given a set $S$, and a probability distribution over that set $P
+But what is the best satisfying lens? We propose that the
+best satisfying lens is the satisyfing lens that connects as much data
+as possible between the two formats. More formally, we
+propose that the best satisfying lens is the satisfying lens that 
+minimizes the expected number of bits required to recover data in
+one format from synchronized data in the other.  
+In more detail, we say that two strings $s$ and $t$
+are \emph{synchronized according to lens $l$} if 
+$l.\PutR\ s\ t = t$ and $l.\PutL\ t\ s = s$ where the notation $l.\PutR$
+extracts the \PutR function from simple symmetric lens $l$; similarly
+for \PutL.
+We say that we can \emph{recover} $s$ from $t$ using bits $b$ and lens
+$l$ if we can reconstruct $s$ from $t$, $b$, and $l$.  For example,
+given the \IdentityLens{} lens, we can recover $s$ from $t$ using
+no bits because in this case, $s$ is just $t$.  In contrast, given
+the \Disconnect{} lens, we need enough bits to fully encode $s$ in
+order to recover it from $t$ because $t$ contains no useful
+information.  This calculation shows that if both \IdentityLens{}
+and \Disconnect{} are satisfying lenses, then \IdentityLens{} is
+better.
+
+The expected number of bits required to recover some data is a
+well-known concept in information theory: it is the \emph{entropy} of
+the data. Calculating the 
+entropy of some data requires a probability distribution over the
+space of possible values the data can take on.  Specifically,
+given a set $S$ and a probability distribution over that set $P
 \OfType S \rightarrow \Reals$, the entropy of the $(S,P)$ pair is $-\Sigma_{s\in
-  S}P(s)\Log_2P(s)$. While we have ways of expressing sets of strings (with
-regular expressions), we do not have probability distributions. Here, we
-express these probability distributions jointly with the language through
-\emph{stochastic regular expressions}, regular expressions annotated with
-probabilities. Our algorithm guesses a reasonable probability distribution from
+  S}P(s)\Log_2P(s)$.
+
+In our setting, we already have a way of expressing sets of data:
+regular expressions.  What we need is a way to express probability
+distributions over that data.
+To that end, we adopt \emph{stochastic regular expressions}\ksf{is
+there a citation?}, 
+which are regular expressions where each operator is annotated with a
+probability.  Stochastic regular expressions
+allow us to jointly specify the allowable data and give a probability
+distribution over it.  From this information, we can calculate the
+entropy of the stochastic regular expression.
+
+Given a lens $\Lens$ between two stochastic regular expressions $\Regex$ and
+$\RegexAlt$, we can calculate the expected number of bits required
+to recover a string $s$ in $\LanguageOf{\Regex}$ from a synchronized
+string $t$ in $\LanguageOf{\RegexAlt}$. 
+This expected value is just the conditional entropy of the $\Regex$ given
+$\RegexAlt$ and $\Lens$. 
+To ensure symmetry, we seek to find the lens that
+minimizes the sum of that entropy with the conditional entropy of $\RegexAlt$
+given $\Regex$ and $\Lens$.
+
+\ksf{This piece might better go somewhere deeper in the paper.}
+Our algorithm guesses a reasonable probability distribution from
 the syntax, though if users manually provided stochastic regular expressions, or
@@
 we could instead use that.
-
-With sets and probability distributions, we can find the expected number of bits
-required to recover one string from a synchronized string. This expected value
-is the conditional entropy of the stochastic regular expression ($\Regex$) given
-another stochastic regular expression ($\RegexAlt$), and a lens that shows which
-strings are synchronized with which ($\Lens$). We aim to find the lens that
-minimizes the sum of that entropy with the conditional entropy of $\RegexAlt$
-given $\Regex$ and $\Lens$.
 

### 2018-07-01 papers/SymmetricLenses/main.tex
edits in progress to section 2

@@
 
-Given a lens $\Lens$ between two stochastic regular expressions $\Regex$ and
-$\RegexAlt$, we can calculate the expected number of bits required
+This infrastructure gives us what we need to find the best
+satisfying lens from a collection of satisfying lenses.  All such
+lenses must synchronize data between the same two 
+stochastic regular expressions $\Regex$ and $\RegexAlt$.
+For any such lens $\Lens$, we can calculate the expected number of
+bits required 
 to recover a string $s$ in $\LanguageOf{\Regex}$ from a synchronized
 string $t$ in $\LanguageOf{\RegexAlt}$. 
-This expected value is just the conditional entropy of the $\Regex$ given
-$\RegexAlt$ and $\Lens$. 
-To ensure symmetry, we seek to find the lens that
-minimizes the sum of that entropy with the conditional entropy of $\RegexAlt$
-given $\Regex$ and $\Lens$.
-
-\ksf{This piece might better go somewhere deeper in the paper.}
-Our algorithm guesses a reasonable probability distribution from
-the syntax, though if users manually provided stochastic regular expressions, or
-a large suite of data in the formats through which we could learn the annotations,
-we could instead use that.
-
-Now that we have a notion of best, we need a way to search for that best lens.
-Our overall approach involves two communicating synthesizers: one (\RXSearch)
-that proposes regular expression pairs, semantically but not syntactically
-equivalent to the originals; one (\GreedySynth) that tries to find the best lens
-within a class of lenses, characterized by such regular expression pairs. Given
-two regular expressions, \GreedySynth can quickly search through the class of
-lenses that can be typed between them, where all instances of type equivalence
-involve no star equivalences. For example, consider the lens shown in
-Figure~\ref{fig:example_lens}. This lens would \emph{not} be in the class
+This expected number of bits is just the conditional entropy of
+$\Regex$ given $\RegexAlt$ and $\Lens$. 
+To ensure symmetry, the entropy we assign to $\Lens$ is the sum of 
+the conditional entropy of $\Regex$ given $\RegexAlt$ and $\Lens$ with 
+the conditional entropy of $\RegexAlt$ given $\Regex$ and $\Lens$.
+
+To obtain stochastic regular expressions from the plain regular
+expressions that we expect users to supply, we use a heurstic 
+that attempts to assign
+probability annotations that give each string in the language equal
+probablity.  We can also take as input a stochastic regular
+expression rather than a plain one, which allows the user to
+specify a data-specific probability distribution.  Alternatively,
+the user could provide a large suite of example data from which we
+could infer the necessary annotations to model the supplied data.
+
+\ksf{Should we be talking about ``best'' or just ``good'' or ``low''?
+There are typically a collection of ``best'' lenses, and we choose the
+one that requires the simplest permutation.}
+Now that we have a notion of best, we need a way to search for that
+best lens.  Previous work on synthesis~\ksf{add citations} has shown
+that we can use types to dramatically restrict the search space and
+improve the effectiveness of synthesis.  In our setting, types are
+pairs of stochastic regular expressions, which come with a rich
+algebraic structure that makes each regular expression semantically
+equal to infinitely many others.  To tame that search space and
+leverage the benefits of type-directed synthesis, we split our
+algorithm into two communicating synthesizers following earlier
+work~\cite{quotient-lens-synthesis}.  The first
+synthesizer, \RXSearch, navigates the space of semantically equivalent
+regular expressions by iteratively applying rewrite rules that
+preserve both the semantics and the probability
+distributions of those stochastic regular expressions. \RXSearch ranks
+pairs of stochastic regular expressions 
+by the number of rewrite rule applications required to obtain the pair
+from the stochastic regular expressions given as input. It passes the
+pairs off to the second synthesizer, \GreedySynth, in rank order.
+
+At a high level, \GreedySynth looks for lenses between a given pair of
+stochastic regular expressions $\Regex$ and $\RegexAlt$ that have the
+best conditional entropy by performing a type-directed search. 
+In more detail, \GreedySynth first uses type equalities based on semiring 
+equivalences (ignoring star semiring equivalences) to normalize the
+stochastic regular expressions provided by \RXSearch into 
+\emph{stochastic n-ary DNF regular expressions}.  
+Then, \GreedySynth uses the syntax of these n-ary DNF regular
+expressions to find highly normalized lenses, in a
+form we call \emph{simple symmetric n-ary DNF lenses}.  These lenses include
+neither a composition operator nor a type equivalence rule.  These
+restrictions mean that each simple symmetric n-ary DNF lens is typed
+by a single pair of stochastic n-ary DNF regular expressions and they
+make \GreedySynth's search space finite.  The variations that \GreedySynth
+considers include choosing between \IdentityLens{} and  \Disconnect
+lenses and deciding how to match multiple occurrences of the same kind
+of data, \EG, the two occurrences of \lstinline{name} in the
+management and HR data formats (we want to map first names to first
+names and last names to last names). 
+Finally, \GreedySynth converts this lens 
+back into a simple symmetric lens by converting the n-ary forms into
+the binary forms provided in the surface language.
+
+As an example of the partnership between \RXSearch and \GreedySynth,
+consider the lens shown in Figure~\ref{fig:example_lens}. 
+This lens is \emph{not} in the set of lenses 
 characterized by \lstinline{emp_salaries} and \lstinline{emp_insurance}, as that
@@
 \lstinline{emp_salaries} into \lstinline{"" | (("\n" . emp_salary) . ("\n" . emp_salary)$^*$)}.
-
-How does \RXSearch propose stochastic regular expression pairs? We define
-rewrite rules for stochastic regular expressions, and prove those rewrites
-preserve language and probability distributions. By applying these rewrites, we
-can generate new regular expression pairs, syntactically distinct but
-semantically equivalent to the originals. \RXSearch proposes the regular
-expression pairs which are most similar to the originals first (in that they
-require fewer rewrite rules to generate), and iteratively propose pairs that are
-increasingly dissimilar to the original pair (in that they require more rewrite
-rules to generate).
-
-Critically for this work, \GreedySynth efficiently searches for one of the best
-lenses in a class of lenses, characterized by a stochastic regular expression
-pair. \GreedySynth works by searching through a smaller search space. First,
-\GreedySynth converts the provided stochastic regular expression pair into a
-\emph{stochastic DNF regular expression} pair. Then, the algorithm uses the
-syntax of the DNF regular expression pair to greedily find a highly normalized
-version of a simple symmetric lens, a \emph{symmetric DNF lens}. Finally, this
-symmetric DNF lens is converted into a simple symmetric lens. DNF Lenses include
-neither a composition operator nor a type equivalence rule. Unlike simple
-symmetric lenses, for which a single lens term is typed between many possible
-regular expression pairs, each DNF lens is typed only between a single regular
-expression pair. The relationship between DNF lens syntax and DNF regular
-expression syntax is very tight, permitting an efficient
-synthesis algorithm for DNF lenses from DNF regular expressions.
 

### 2018-07-01 papers/SymmetricLenses/main.tex
Finished pass over section 2

@@
   }
-  \caption{Example data files and corresponding regular expressions
-  used by management and HR, respectively, to represent employee salaries and health
-  insurance companies.  }  
+  \caption{Hypothetical example data files and corresponding regular expressions
+  used by management and HR at a US company to represent employee salaries and health
+  insurance providers, respectively. \ksf{Why
+  is \lstinline{emp_salaries} written with an or instead of just using
+  the star operator? Can we simplify?} }  
   \label{fig:minimized-representations}
@@
 is oblivious to salary changes.  Simple symmetric lenses are
-appropriate for keeping these two files synchronized because they have
+appropriate for keeping these two files synchronized because the files have
 a mix of shared and unshared information.
@@
 and $\RegexAlt$ 
-is defined by four functions: 
-\begin{enumerate}
-\item $\CreateR \OfType \Regex \rightarrow \RegexAlt$
-\item $\CreateL \OfType \RegexAlt \rightarrow \Regex$
-\item $\PutR \OfType \Regex \rightarrow \RegexAlt \rightarrow \RegexAlt$
-\item $\PutL \OfType \RegexAlt \rightarrow \Regex \rightarrow \Regex$
-\end{enumerate}
+is defined by four functions
+
+%\begin{enumerate}
+%\item $\CreateR \OfType \Regex \rightarrow \RegexAlt$
+%\item $\CreateL \OfType \RegexAlt \rightarrow \Regex$
+%\item $\PutR \OfType \Regex \rightarrow \RegexAlt \rightarrow \RegexAlt$
+%\item $\PutL \OfType \RegexAlt \rightarrow \Regex \rightarrow \Regex$
+%\end{enumerate}
+
+\begin{gather}
+ \tag{1}
+ \mbox{$\CreateR \OfType \Regex \rightarrow \RegexAlt$}\\
+ \tag{2}
+ \mbox{$\CreateL \OfType \RegexAlt \rightarrow \Regex$}\\
+ \tag{3}
+ \mbox{$\PutR \OfType \Regex \rightarrow \RegexAlt \rightarrow \RegexAlt$}\\
+ \tag{4}
+ \mbox{$\PutL \OfType \RegexAlt \rightarrow \Regex \rightarrow \Regex$}
+\end{gather}
+
 \noindent
@@
 \noindent
-The two \emph{create} functions provide default values for times when
-data appears on one side but not the other (eg, the ``N/A'' salary
+The two create functions provide default values for times when
+data appears on one side but not the other (\EG, the ``N/A'' salary
 entry added to the management file when HR inserts a new employee).
-The two \emph{put} functions propagate edits from one format to the
+The two put functions propagate edits from one format to the
 other by combining the change from one file with the previous value
@@
 \subsection{Simple Symmetric Lens Combinators} 
-\ksf{This section needs more polishing}
 Generally, we do not expect programmers to define simple symmetric lenses by
@@
 combinators to build up the requisite lenses from simpler ones.
-We explain these combinators by using them to define a lens between
+We briefly explain a selection of these combinators by using them to
+define a lens between 
 the two employee data formats in Figure~\ref{fig:minimized-representations}. 
@@
 
-The simplest combinator is the identity lens, which takes as an argument a
-regular expression, and performs the 
-identity function on strings in that regular expression. The identity lens fully
-matches the data formats: when one data format is updated, the other format gets
-the exact same value. In our example, when a name is updated, it is also
-expected to be updated in the other format. The identity lens propagates that
-information.
+The simplest combinator is the identity lens \IdentityLens, which takes as an
+argument a regular expression $\Regex$ and performs an identity mapping on data
+matching $\Regex$. The identity lens fully connects the
+corresponding data in the two formats: when one format is
+updated, the other format gets the exact same value. In our example,
+when a \lstinline{name} is updated in one file, the identify lens ensures it is
+also updated in the other file.
+
 %
@@
 
-In contrast to the identity lens, the disconnect lens leaves data
-unaligned. The disconnect lens takes four arguments: two regular 
+In contrast to the identity lens, the disconnect lens \Disconnect leaves data
+unconnected. The disconnect lens takes four arguments: two regular 
 expressions and two strings. The regular expressions specify the
-formats of the data on the two sides while the strings provide default
-values.  In our extended example, salaries and
-insurance companies unrelated, so we use a disconnect lens.  The
-defaults are strings that indicate the information is missing.
+formats of the corresponding data on the two sides while the strings provide default
+values.  In our extended example, 
+the \lstinline{salary} and \lstinline{company} fields are independent,
+so we use the disconnect lens to relate them.
+The defaults are strings that indicate the information is missing.
 %
 \begin{lstlisting}
-disc(salary,company,"N/A","UNK") : salary $\Leftrightarrow$ company
+disc(salary, company, "N/A", "UNK") : salary $\Leftrightarrow$ company
 \end{lstlisting}
 %
-When building a disconnect lens between two strings, one of which is merely the
-empty string, the lenses \lstinline{ins} and \lstinline{del} can instead be
-used, where \lstinline{ins($t$) = disc("",$t$,"",$t$)} and
-\lstinline{del($s$) = disc($s$,"",$s$,"")}.
-
-Finally, there are a number of lenses that compositionally build up new lenses
-from old ones. 
-For example, data can be concatenated, and so we provide
-the \lstinline{concat} combinator that sequences lenses, usually
-written using an infix notation \lstinline{$\Lens_1$ . $\Lens_2$}:
+The insert \lstinline{ins} and delete \lstinline{del} lenses are
+syntactic sugar for uses of the disconnect lens in which a string
+constant is omitted entirely from the source and target formats,
+respectively. \ksf{Are these in the right order?}
+\begin{lstlisting}
+ ins($t$) = disc("", $t$, "", $t$)
+ del($s$) = disc($s$, "", $s$, "")
+\end{lstlisting}
+
+Finally, there are a number of standard lenses that compositionally
+construct more complex lenses from simpler ones.  These combinators include 
+\lstinline{concat} for concatenation, 
+\lstinline{or} for variation, and
+\IterateLens for sequences.
+We usually write \lstinline{concat} using an infix
+notation \lstinline{$\Lens_1$ . $\Lens_2$}
+as in:
 %
@@
 %
-
-Data can also appear in multiple forms. For example,
-a list of employees with insurance information can either be empty or
-be a sequence of employee/insurance records: \ksf{This example is
-vacuous as you could eliminate the or in favor of the *.  Do somethign
-else to introduce or?}
+which appears in the definition of the \lstinline{name_lens} in our
+running example. 
+%\ksf{I don't think this example is necessary.}
+%Data can also appear in multiple forms. For example,
+%a list of employees with insurance information can either be empty or
+%be a sequence of employee/insurance records: \ksf{This example is
+%vacuous as you could eliminate the or in favor of the *.  Do somethign
+%else to introduce or?}
 %
-\begin{lstlisting}
-or(id(""),employees_lens) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_ins) . ("\n" emp_ins)*
-\end{lstlisting}
-
-Lastly, the \IterateLens lens synchronizes repeated data, aligning the number of
-times the data is repeated. The \IterateLens combinator takes a lens as an
-argument and synchronizes each element of the iteration with that lens. For
-example, if \lstinline{employee_lens} synchronizes a single employee's data,
-iterating that lens builds a lens that synchronizes a list of employees.
+%\begin{lstlisting}
+%or(id(""),employees_lens) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_ins) . ("\n" emp_ins)*
+%\end{lstlisting}
+The \IterateLens lens also appears in our running example, where it
+synchronizes data for a list of employees when given a
+lens \lstinline{employee_lens} that synchronizes  data for a single employee:
 %
@@
 let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
-let emp_lens = or(id(""),employees_lens) in
 ins(header) . employees_lens : emp_salaries $\Leftrightarrow$ emp_insurance
 \end{lstlisting}
+%\ksf{I don't think this is necesssary.}
+%let emp_lens = or(id(""),employees_lens) in
   \caption{A lens, written in our language for building simple symmetric lenses
@@
 
-Using these combinators, we can build up a lens that synchronizes
-management and HR employee files.
-
-\ksf{This paragraph needs to be updated, if nothing else, to reflect
-the new example. Also, we haven't mentioned any typing rules so the
-first sentence doesn't make sense.  Perhaps the whole topic could be
-defered to later where we can explain why such a rule is necessary?}
-This lens uses one typing rule not previously
-mentioned: type equivalence. While \lstinline{scheduled_job} is a equivalent to
-a regular expression with an outermost disjunction, it is not syntactically
-equivalent to one. Without type equivalence, the provided lens wouldn't actually
-be well-typed, but type equivalence states that if a lens is typed between a
-pair of regular expressions, it is also typed between an equivalent pair.
-
-Note that, while the synchronization task is quite simple, the
+% Using these combinators, we can build up a lens that synchronizes
+% management and HR employee files.
+
+We provide typing rules for each of the simple symmetric combinators,
+most of which are syntax directed.  There is one additional
+typing rule, which allows us to use equivalence relations on regular
+expressions to convert the type of any lens to a semantically equal
+but syntactically different pair of regular expressions.  The flexibility this
+type equivalence rule provides is useful when a different
+representation of a regular expression would facilitate synthesis.
+
+%%\ksf{Do we need to use
+%%type equivalence to type the employee example?  If so, perhaps we
+%%should explain her}
+%%
+%%This lens uses one typing rule not previously
+%%mentioned: type equivalence. While \lstinline{scheduled_job} is a equivalent to
+%%a regular expression with an outermost disjunction, it is not syntactically
+%%equivalent to one. Without type equivalence, the provided lens wouldn't actually
+%%be well-typed, but type equivalence states that if a lens is typed between a
+%%pair of regular expressions, it is also typed between an equivalent pair.
+
+Note that, while the synchronization task for our running example is quite simple, the
 requisite lens is relatively complex. Furthermore, this complexity is
@@
 
-To alleviate these difficulties, we wish to synthesize symmetric lenses---given
+To alleviate the difficulty of manually writing simple symmetric
+lenses, we wish to synthesize them instead. Given 
 a pair of regular expressions and a set of input-output examples, we want to
-find a symmetric lens expression that is typed between the regular expression
-pair, and satisfies the input/output examples. We call such a lens
+find a simple symmetric lens that is typed by the regular expression
+pair and satisfies the input/output examples. We call such a lens
 a \emph{satisfying lens}.
-In this example, we wish to
-synthesize a lens from \lstinline{emp_salaries} and \lstinline{emp_insurance},
-using the example data in Figure~\ref{fig:minimized-representations} as an
-input-output example from which to draw the required defaut values for 
-\Disconnect lenses.
+In our running example, we wish to
+synthesize a lens between \lstinline{emp_salaries} and \lstinline{emp_insurance},
+using as an input-output example the data in
+Figure~\ref{fig:minimized-representations}.  The algorithm uses such
+examples to find suitable defaults for \Disconnect lenses and to
+figure out how to match data that could be synchronised in multiple
+ways.
 
@@
 is another valid lens (where \lstinline{rep($\Regex$)} is a string in the
-language of $\Regex$). Furthermore, adding the synchronized files from
-Figure~\ref{fig:minimized-representations} as an input-output example helps
-little. The default strings given to \Disconnect merely need to be the strings
-in the input-output examples, and the lens will still satisfy the specification.
-While the disconnect lens is prohibited by two examples, or by a carefully
-chosen input-output example on \PutR or \PutL, this problem continues presenting
-itself for the sublenses, and is highlighted even more on larger
-formats. 
-\ksf{I don't understand the above.  We shouldn't explain the issue in
-terms of \PutR and \PutL as we are focused on the combinators rather
-than the raw functions.}
+language of $\Regex$). Furthermore, adding the input-output example in
+Figure~\ref{fig:minimized-representations} 
+helps, but only a little, because it requires many examples to rule
+out all the possible occurrences of nested \Disconnect lenses,
+particularly in complex formats. 
+%\ksf{I don't understand the following.  We shouldn't explain the issue in
+%terms of \PutR and \PutL as we are focused on the combinators rather
+%than the raw functions.}
+%The default strings given to \Disconnect merely need to be the strings
+%in the input-output examples, and the lens will still satisfy the
+%specification.  While the disconnect lens is prohibited by two
+%examples, or by a carefully chosen input-output example on \PutR
+%or \PutL, this problem continues presenting itself for the sublenses,
+%and is highlighted even more on larger formats.
 While
@@
 automatically. Instead of merely finding \emph{any} satisfying lens, we wish to
-synthesize the \emph{best} satisfying lens.
-
-But what is the best satisfying lens? We propose that the
-best satisfying lens is the satisyfing lens that connects as much data
-as possible between the two formats. More formally, we
-propose that the best satisfying lens is the satisfying lens that 
-minimizes the expected number of bits required to recover data in
-one format from synchronized data in the other.  
-In more detail, we say that two strings $s$ and $t$
-are \emph{synchronized according to lens $l$} if 
-$l.\PutR\ s\ t = t$ and $l.\PutL\ t\ s = s$ where the notation $l.\PutR$
-extracts the \PutR function from simple symmetric lens $l$; similarly
-for \PutL.
-We say that we can \emph{recover} $s$ from $t$ using bits $b$ and lens
-$l$ if we can reconstruct $s$ from $t$, $b$, and $l$.  For example,
-given the \IdentityLens{} lens, we can recover $s$ from $t$ using
-no bits because in this case, $s$ is just $t$.  In contrast, given
+synthesize a satisfying lens that is likely to please the user.
+
+But what is a likely satisfying lens? We propose that a satisfying
+lens is more likely if it connects more data between the two
+formats, \IE, by using an identity rather than a disconnect lens.  The
+likeliest satisfying lenses would connect as much data as
+possible. More formally, we propose that the likelihood of a
+satisfying lens is the expected number of bits required to recover
+data in one format from synchronized data in the other.  Higher
+likelihoods correspond to fewer numbers of bits. In more
+detail, we say that two strings $s$ and $t$ are \emph{synchronized
+according to lens $l$} if $l.\PutR\ s\ t = t$ and $l.\PutL\ t\ s = s$.\footnote{
+The notation $l.\PutR$ extracts the \PutR function from simple
+symmetric lens $l$; similarly for \PutL.}  We say that we
+can \emph{recover} $s$ from $t$ using bits $b$ and lens $l$ if we can
+reconstruct $s$ from $t$, $b$, and $l$.  For example, given
+the \IdentityLens{} lens, we can recover $s$ from $t$ using no bits
+because in this case, $s$ is just $t$.  In contrast, given
 the \Disconnect{} lens, we need enough bits to fully encode $s$ in
@@
 information.  This calculation shows that if both \IdentityLens{}
-and \Disconnect{} are satisfying lenses, then \IdentityLens{} is
-better.
-
-The expected number of bits required to recover some data is a
+and \Disconnect{} are satisfying lenses, then \IdentityLens{} is more
+likely.
+
+The expected number of bits required to recover data is a
 well-known concept in information theory: it is the \emph{entropy} of
 the data. Calculating the 
-entropy of some data requires a probability distribution over the
-space of possible values the data can take on.  Specifically,
+entropy of data requires a probability distribution over the
+space of possible values for the data.  Specifically,
 given a set $S$ and a probability distribution over that set $P
@@
 
-This infrastructure gives us what we need to find the best
-satisfying lens from a collection of satisfying lenses.  All such
+This infrastructure gives us what we need to find likely
+satisfying lenses from a collection of satisfying lenses.  All such
 lenses must synchronize data between the same two 
@@
 $\Regex$ given $\RegexAlt$ and $\Lens$. 
-To ensure symmetry, the entropy we assign to $\Lens$ is the sum of 
-the conditional entropy of $\Regex$ given $\RegexAlt$ and $\Lens$ with 
+To ensure symmetry, the likelihood we assign to $\Lens$ is the sum of 
+the conditional entropy of $\Regex$ given $\RegexAlt$ and $\Lens$ and
 the conditional entropy of $\RegexAlt$ given $\Regex$ and $\Lens$.
@@
 
-\ksf{Should we be talking about ``best'' or just ``good'' or ``low''?
-There are typically a collection of ``best'' lenses, and we choose the
-one that requires the simplest permutation.}
-Now that we have a notion of best, we need a way to search for that
-best lens.  Previous work on synthesis~\ksf{add citations} has shown
+Now that we have a way to calculate likelihood, we need a way to search for
+it.
+Previous work on synthesis~\ksf{add citations} has shown
 that we can use types to dramatically restrict the search space and
@@
 preserve both the semantics and the probability
-distributions of those stochastic regular expressions. \RXSearch ranks
-pairs of stochastic regular expressions 
+distributions of those stochastic regular expressions. \RXSearch
+ranks pairs of stochastic regular expressions 
 by the number of rewrite rule applications required to obtain the pair
 from the stochastic regular expressions given as input. It passes the
-pairs off to the second synthesizer, \GreedySynth, in rank order.
+pairs off to the second synthesizer, \GreedySynth, in rank order with
+the smallest first.
 
@@
 stochastic regular expressions $\Regex$ and $\RegexAlt$ that have the
-best conditional entropy by performing a type-directed search. 
+highest likelihood by performing a type-directed search. 
 In more detail, \GreedySynth first uses type equalities based on semiring 
-equivalences (ignoring star semiring equivalences) to normalize the
+equivalences (ignoring the star semiring equivalences that are managed
+by \RXSearch) to normalize the
 stochastic regular expressions provided by \RXSearch into 
@@
 restrictions mean that each simple symmetric n-ary DNF lens is typed
-by a single pair of stochastic n-ary DNF regular expressions and they
-make \GreedySynth's search space finite.  The variations that \GreedySynth
+by a single pair of stochastic n-ary DNF regular expressions and 
+\GreedySynth's search space for a given pair of regular expressions is
+finite.  The variations that \GreedySynth 
 considers include choosing between \IdentityLens{} and  \Disconnect
@@
 consider the lens shown in Figure~\ref{fig:example_lens}. 
-This lens is \emph{not} in the set of lenses 
-characterized by \lstinline{emp_salaries} and \lstinline{emp_insurance}, as that
-typing derivation requires unrolling \lstinline{("\n" . emp_salary)$^*$} within
-\lstinline{emp_salaries} into \lstinline{"" | (("\n" . emp_salary) . ("\n" . emp_salary)$^*$)}.
-
-Our strategy involving communicating synthesizers permits both a way of
-enumerating classes and efficiently searching through them, but it poses a
-problem: when do we stop? While we may have found a great lens between two regular
-expressions, a different regular expression pair may provide an even better
-lens. Moreover, we may find a great lens characterized by
-\lstinline{emp_salaries} and \lstinline{emp_insurances} that is not the
-desired lens shown in Figure~\ref{fig:example_lens}.  The
-tension is that we want to find a good lens, but we also want to give quick feedback to
-the user. After finding the lens in Figure~\ref{fig:example_lens}, we dont want
-the algorithm to spend much longer searching for better lenses.
-
-\saz{The following is hard to follow} Eventually, we cut the proposal-synthesis
-loop short. Before proposing new types \RXSearch evaluates whether it should
-continue proposing regular expression pairs, or whether it is content with lens
-of lowest cost seen thus far. If there are a large number of regular expression
-pairs that are equally dissimilar to the originals or if the current lens is
-very well synchronized, \RXSearch stops. By doing this \RXSearch continues
-searching for better lenses, but does not tackle huge search spaces unless it
-needs to. In this way, the algorithm can provide lenses quickly to the users,
-which users can refine with additional examples.
-
-\afm{calculations for cost of various lenses between \lstinline{emp_salary} and
-  \lstinline{emp_row}} Consider the lens \lstinline{single_emp_map} between
-\lstinline{emp_salary} and \lstinline{emp_row}. Given this lens, the expected
-entropy of recovering the left format from the right
-($\EntropyOf{\lstinline{emp_salary} \Given \lstinline{emp_row},
-  \lstinline{single_emp_map} }$) is the sum of the expected entropy of
-recovering each of its subcomponents, $\EntropyOf{\lstinline{emp_salary} \Given
-  \lstinline{emp_row}, \lstinline{single_emp_map} } = \EntropyOf{\lstinline{name
-    . " " . name} \Given \lstinline{name . "," . name}, \lstinline{name_map}
-}$\\ + $\EntropyOf{\lstinline{": " . salary} \Given
-  \lstinline{"," . company}, \lstinline{disc(salary,company,"UNSET","UNSET")}
-}$.
-
-The entropy of the identity lens is zero. As the two formats are in bijective
-correspondence, we know exactly which string is present in the left-hand format.
-The entropy of the disconnect lens is just the entropy of
-\lstinline{": " . salary}.  Knowing which string is present in the right-hand
-format tells us nothing about what string is present in the left-hand format. So
-$\EntropyOf{\lstinline{emp_salary} \Given \lstinline{emp_row},
-  \lstinline{single_emp_map} } = \EntropyOf{\lstinline{": " . salary} }$
-
-Analogously, $\EntropyOf{\lstinline{emp_row} \Given \lstinline{emp_salary},
-  \lstinline{single_emp_map} }$ is just $\EntropyOf{\lstinline{"," . company}
-}$, so the total cost of the lens is $\EntropyOf{\lstinline{": " . salary}} +
-\EntropyOf{\lstinline{"," . company}}$
-
-Consider the trivial lens
-\lstinline{disc(emp_salary,emp_row,rep(emp_salary),rep(emp_row))}. As this is
-the disconnect lens, the cost of each side is the entropy of the data format, so
-the total cost is $\EntropyOf{\lstinline{emp_salary}} +
-\EntropyOf{\lstinline{emp_row}}$. \lstinline{emp_salary} contains strictly more
-imformation than \EntropyOf{\lstinline{": " . salary}}, and \lstinline{emp_row}
-contains strictly more information than \EntropyOf{\lstinline{": " . salary}},
-so by our metric \lstinline{single_emp_map} is preferable to the disconnect
-lens.
+Inferring this lens requires \RXSearch to use a star semiring axiom to
+rewrite the regular expression \lstinline{("\n" . emp_salary)$^*$} within
+\lstinline{emp_salaries} into 
+\lstinline{"" | (("\n" . emp_salary). ("\n" . emp_salary)$^*$)} 
+so that \GreedySynth can find the appropriate lens.
+
+Our strategy of communicating synthesizers gives us a way to
+enumerate pairs of stochastic regular expressions of increasing
+rank and to efficiently search through them, but it poses a
+problem: when do we stop? We may have found a promising lens
+between a pair of stochastic regular expressions, but a different 
+pair we haven't yet discovered may give rise to an even better lens. 
+Or, we may have found a promising lens between a given pair of
+stochastic regular expressions, but the discovered lens is not 
+acceptable to the users.
+
+Our algorithm must resolve a three-way tension between the quality of
+the inferred lens, the number of examples the user must provide to
+eliminate unwanted lenses, and the amount of time it takes to return a
+result.  For example, if the algorithm has already found the lens in
+Figure~\ref{fig:example_lens}, we don't want to spend a lot of time
+searching for an even better lens.
+
+To resolve this tension, \RXSearch uses heuristics to judge whether to
+return the current best satisfying lens to the user or to pass the
+next set of unrollings to \GreedySynth.
+The heuristics favor stopping if the current best satisfying
+lens has relatively high likelihood, indicating the lens is very promising.
+The heuristics also favor stopping if \RXSearch 
+has delievered to \GreedySynth all pairs of stochastic regular
+expressions at a given rank and there is a large number of pairs at the next
+rank because searching through all such pairs will take a long time.
+With this approach, the algorithm can quickly respond with a satisfying
+lens with relatively high likelihood.  If users are unhappy with the result, they can refine the
+search by supplying additional examples, which serve both to rule out
+the previously proposed lens(es) and to reduce the size of the search
+space by cutting down on the number of satisfying lenses.
+
+\ksf{Commented out entropy calculation; I don't think we need that
+level of detail; leaving the text in case we change our mind :-)}
+%Not sure we need this level of detail after all.
+%\afm{calculations for cost of various lenses between \lstinline{emp_salary} and
+%  \lstinline{emp_row}} Consider the lens \lstinline{single_emp_map} between
+%\lstinline{emp_salary} and \lstinline{emp_row}. Given this lens, the expected
+%entropy of recovering the left format from the right
+%($\EntropyOf{\lstinline{emp_salary} \Given \lstinline{emp_row},
+%  \lstinline{single_emp_map} }$) is the sum of the expected entropy of
+%recovering each of its subcomponents, $\EntropyOf{\lstinline{emp_salary} \Given
+%  \lstinline{emp_row}, \lstinline{single_emp_map} } = \EntropyOf{\lstinline{name
+%    . " " . name} \Given \lstinline{name . "," . name}, \lstinline{name_map}
+%}$\\ + $\EntropyOf{\lstinline{": " . salary} \Given
+%  \lstinline{"," . company}, \lstinline{disc(salary,company,"UNSET","UNSET")}
+%}$.
+%
+%The entropy of the identity lens is zero. As the two formats are in bijective
+%correspondence, we know exactly which string is present in the left-hand format.
+%The entropy of the disconnect lens is just the entropy of
+%\lstinline{": " . salary}.  Knowing which string is present in the right-hand
+%format tells us nothing about what string is present in the left-hand format. So
+%$\EntropyOf{\lstinline{emp_salary} \Given \lstinline{emp_row},
+%  \lstinline{single_emp_map} } = \EntropyOf{\lstinline{": " . salary} }$
+%
+%Analogously, $\EntropyOf{\lstinline{emp_row} \Given \lstinline{emp_salary},
+%  \lstinline{single_emp_map} }$ is just $\EntropyOf{\lstinline{"," . company}
+%}$, so the total cost of the lens is $\EntropyOf{\lstinline{": " . salary}} +
+%\EntropyOf{\lstinline{"," . company}}$
+
+%Consider the trivial lens
+%\lstinline{disc(emp_salary,emp_row,rep(emp_salary),rep(emp_row))}. As this is
+%the disconnect lens, the cost of each side is the entropy of the data format, so
+%the total cost is $\EntropyOf{\lstinline{emp_salary}} +
+%\EntropyOf{\lstinline{emp_row}}$. \lstinline{emp_salary} contains strictly more
+%imformation than \EntropyOf{\lstinline{": " . salary}}, and \lstinline{emp_row}
+%contains strictly more information than \EntropyOf{\lstinline{": " . salary}},
+%so by our metric \lstinline{single_emp_map} is preferable to the disconnect
+%lens.
 % end overview

### 2018-07-01 papers/SymmetricLenses/main.tex
minor changes

--- 2018-06-30 papers/SymmetricLenses/main.tex
@@
 \usepackage{amssymb}
-\usepackage[inline]{enumitem}
+\usepackage{enumitem}
 \usepackage{multirow,bigdelim}
@@
 ]
-Jane Doe: 38000
+Jane Doe: 35000
 John Public: 37500
@@
 ]
-Name,Company
-Jane Doe,Healthcare Inc.
-John Public,Insurance Co.
+lastname,firstname,company
+Doe,Jane,Healthcare Inc.
+Public,John,Insurance Co.
 \end{lstlisting}
@@
 \begin{lstlisting}
-let salary = number | "N/A"
+let salary = number | "UNSET"
 let emp_salary = name . " " . name . ": " salary
@@
 \begin{lstlisting}
-let company = (name . ("Co." | "Inc." | "Ltd.")) | "UNK" in
-let emp_ins = name . " " . name "," company in
-let header = "Name,Company" in
-let emp_insurance = header . ("\n" . emp_ins)*
+let company = (name . ("Co." | "Inc." | "Ltd.")) | "UNSET" in
+let emp_row = name . "," . name "," company in
+let header = "firstname,lastname,company" in
+let emp_insurances = header . ("\n" . emp_row)*
 \end{lstlisting}
@@
       \hline
-      \texttt{company.mgmt} data file & \texttt{company.hr} data file \\
+      company.mgmt & company.hr \\
       \hline
@@
       \hline
-      \cellcolor{vlightyellow}\texttt{company.mgmt} regular expression & 
-      \cellcolor{vlightgreen}\texttt{company.hr}  regular expression\\
-      \hline
-      \cellcolor{vlightyellow}\listingC & \cellcolor{vlightgreen}\listingD \\
+      \listingC & \listingD \\
       \hline
@@
   }
-  \caption{Example data files and corresponding regular expressions
-  used by management and HR, respectively, to represent employee salaries and health
-  insurance companies.  }  
+  \caption{Example files used by management and HR, and regular expressions
+    describing the formats of these files.}
   \label{fig:minimized-representations}
@@
 
-In this section, we give an overview of simple symmetric lenses and
-our corresponding synthesis algorithm using an example drawn from a
-hypothetical U.S. company.  In this company, management and
-human resources (HR) store information about employees in separate files:
-management records the names of employees and their salaries while
-HR stores the names of employees and their health
-insurance providers.  Figure~\ref{fig:minimized-representations} gives
-examples of the two types of files and regular expressions describing
-the two formats.
-The company uses simple symmetric lenses to keep these files synchronized.
-If management adds a new employee, ``Chris Roe: 32500'',  a lens adds 
-the corresponding entry ``Chris Roe, UNK'' to HR's file.  The
-sentinel value ``UNK'' represents the fact that the employee's
-insurance company is currently unknown.  A similar update happens if
-HR adds a new employee before management, in which case the sentinel
-value ``N/A'' reflects the unknown salary information. 
-Furthermore, if HR
-corrects an error in an employee's name, say changing ``John Public''
-to ``Jon Public'', the lens mirrors that name change in management's
-file. Not all the data is mirrored, however. The management file 
-is not updated in response to insurance changes, and the HR file
-is oblivious to salary changes.  Simple symmetric lenses are
-appropriate for keeping these two files synchronized because they have
-a mix of shared and unshared information.
+\afm{i commented out the previous version, because this one was quickly written
+  up to give you something to work off. Feel free to delete it if it isnt
+  helpful. } We motivate our problem with an extended example. In our
+theoretical company, human resources and upper management both keep information
+about employees. HR needs to know the names of the employees as well as their
+insurance providers. Management needs to know the names of employees and their
+salary. Two example files, and regular expressions describing the formats, are
+shown in Figure~\ref{fig:minimized-representations}. These two files should be
+synchronized: if management adds a new employee, ``Richard Roe: 32500'', a new
+row should be added to HR's file, ``Roe,Richard,UNSET''. Furthermore, if HR
+changes ``John Public'' to ``Jon Public'', that name change should be reflected
+in both files. However, not all the data should be synchronized. The HR file
+should not get altered due to salary changes, and the management file should not
+get altered due to insurance company changes. While names should be matched
+across the two formats, insurance company and salary information should be left
+unmatched. Such a synchronization can be achieved with simple symmetric lenses.
 
@@
 
-\subsection{Simple Symmetric Lenses}
-A simple symmetric lens between formats described by types $\Regex$
-and $\RegexAlt$ 
-is defined by four functions: 
+Lenses synchronize files by providing 4 functions, two \emph{create} functions,
+and two \emph{put} functions. Create functions are used to initialize data in
+one format when provided with data in the other. If a company.mgmt file is
+initialized, a synchronized company.hr file should also be created, and vice
+versa. Put functions are used to propagate edits from one format to the other.
+If a company.hr file is updated, we should be able to combine it with a
+non-updated company.mgmt file, and output a new company.mgmt file, synchronized
+to the updated company.hr. To have a simple symmetric lens is to have these
+required synchronization functions. Formally, a simple symmetric lens between
+$X$ and $Y$ comprises 4 functions:
 \begin{enumerate}
-\item $\CreateR \OfType \Regex \rightarrow \RegexAlt$
-\item $\CreateL \OfType \RegexAlt \rightarrow \Regex$
-\item $\PutR \OfType \Regex \rightarrow \RegexAlt \rightarrow \RegexAlt$
-\item $\PutL \OfType \RegexAlt \rightarrow \Regex \rightarrow \Regex$
+\item $\CreateR \OfType X \rightarrow Y$
+\item $\CreateL \OfType Y \rightarrow X$
+\item $\PutR \OfType X \rightarrow Y \rightarrow Y$
+\item $\PutL \OfType Y \rightarrow X \rightarrow X$
 \end{enumerate}
-\noindent
-that collectively satisfy four ``round-tripping'' laws: 
+Furthermore, if a file is not edited, but is then resynchronized using a put
+function, no data should be changed. To be a simple symmetric lens, one must
+satisfy the following ``round-tripping'' laws, guaranteeing identity edits won't
+change the file data.
 \begin{gather}
   \tag{\CreatePutRL}
-  \PutLOf{(\CreateROf{s})}{s} = s\\
+  \PutLOf{(\CreateROf{x})}{x} = x\\
   \tag{\CreatePutLR}
-  \PutROf{(\CreateLOf{t})}{t} = t\\
+  \PutROf{(\CreateLOf{y})}{y} = y\\
   \tag{\PutRL}
-  \PutLOf{(\PutROf{s}{t})}{s} = s\\
+  \PutLOf{(\PutROf{x}{y})}{x} = x\\
   \tag{\PutLR}
-  \PutROf{(\PutLOf{t}{s})}{t} = t
+  \PutROf{(\PutLOf{y}{x})}{y} = y
 \end{gather}
-\noindent
-The two \emph{create} functions provide default values for times when
-data appears on one side but not the other (eg, the ``N/A'' salary
-entry added to the management file when HR inserts a new employee).
-The two \emph{put} functions propagate edits from one format to the
-other by combining the change from one file with the previous value
-from the other file. The roundtripping laws guarantee that identify edits to one file do
-not change the data in the other file.
-
-%%For synchronization, when the original company.hr or company.mgmt
-%%file is created, a synchronized file of the other form is created
-%%from it with $\CreateR$ or $\CreateL$ (depending on which file was
-%%initially created). On future edits, changes are made with the puts,
-%%a \PutR would propagate the shared information from the company.hr
-%%file, and knit it together with the information specific to the company.mgmt file.
-
-We use regular expressions to denote the types of data formats, and we
-impose a typing discipline on our lenses.  Specifically, if 
-$\Regex$ and $\RegexAlt$ are regular expressions, then the notation
-\lstinline{$\Lens$ : $\Regex \Leftrightarrow \RegexAlt$} indicates that $\Lens$
-is a simple symmetric lens between $\LanguageOf{\Regex}$ and
-$\LanguageOf{\RegexAlt}$. 
-When clear from context, we use the term \textit{lens} to denote a
-simple symmetric lens.
-
-\subsection{Simple Symmetric Lens Combinators} 
-\ksf{This section needs more polishing}
-Generally, we do not expect programmers to define simple symmetric lenses by
-writing the four functions by hand and showing they satisfy the
-roundtripping laws.  Instead, we provide simple symmetric lens
-combinators to build up the requisite lenses from simpler ones.
-We explain these combinators by using them to define a lens between
-the two employee data formats in Figure~\ref{fig:minimized-representations}. 
-The finished lens, which appears in Figure~\ref{fig:example_lens}, has
-the type 
-\lstinline{emp_salaries} $\Leftrightarrow$ \lstinline{emp_insurance}, where
-\lstinline{emp_salaries} and \lstinline{emp_insurance} are regular
-expressions defined in Figure~\ref{fig:minimized-representations}.
-
-The simplest combinator is the identity lens, which takes as an argument a
-regular expression, and performs the 
+
+For synchronization, when the original company.hr or company.mgmt file is
+created, a synchronized file of the other form is created from it with
+$\CreateR$ or $\CreateL$ (depending on which file was initially created). On
+future edits, changes are made with the puts, a \PutR would propagate the shared
+information from the company.hr file, and knit it together with the information
+specific to the company.mgmt file.
+
+\subsection{Programming Symmetric Lenses} We now build up a symmetric lens by
+example. Our symmetric lens expressions are typed between regular expression
+pairs, if \lstinline{$\Lens$ : $\Regex \Leftrightarrow \RegexAlt$}, then $\Lens$
+encodes a simple symmetric lens between $\LanguageOf{\Regex}$ and
+$\LanguageOf{\RegexAlt}$. At the end, we want our symmetric lenses to by typed
+between \lstinline{emp_salaries} and \lstinline{emp_insurance}.  The finished
+lens is shown in Figure~\ref{fig:example_lens}.
+
+The identity lens takes, as argument, a regular expression, and performs the
 identity function on strings in that regular expression. The identity lens fully
@@
 the exact same value. In our example, when a name is updated, it is also
-expected to be updated in the other format. The identity lens propagates that
+expected to updated in the other format. The identity lens propagates that
 information.
@@
 
-In contrast to the identity lens, the disconnect lens leaves data
-unaligned. The disconnect lens takes four arguments: two regular 
-expressions and two strings. The regular expressions specify the
-formats of the data on the two sides while the strings provide default
-values.  In our extended example, salaries and
-insurance companies unrelated, so we use a disconnect lens.  The
-defaults are strings that indicate the information is missing.
+Where the identity lens aligns data, the disconnect lens does the opposite: it
+leaves data unaligned. The disconnect lens takes 4 arguments: two regular
+expressions and two strings. On creates, because there is no association between
+data in the first format and data in the second format, designated default
+strings are used for the other format. In our extended example, salaries and
+insurance companies should not be matched, so a disconnect lens would be used.
+We want our default behavior to specify that the value has not been set.
 %
 \begin{lstlisting}
-disc(salary,company,"N/A","UNK") : salary $\Leftrightarrow$ company
+disc(salary,company,"UNSET","UNSET") : salary $\Leftrightarrow$ company
 \end{lstlisting}
@@
 Finally, there are a number of lenses that compositionally build up new lenses
-from old ones. 
-For example, data can be concatenated, and so we provide
-the \lstinline{concat} combinator that sequences lenses, usually
-written using an infix notation \lstinline{$\Lens_1$ . $\Lens_2$}:
+from old ones. For example, data can appear in one form or another: a program
+can run during a specific set of minutes, or that program can run the every
+minute. If \lstinline{minutes_map} and \lstinline{alltimes_map} are lenses that
+synchronize those settings in each format, then those lenses can be composed
+with an \OrLens lens.
 %
 \begin{lstlisting}
-id(name) . ins(",") . id(name) . del(": ") . ins(",")
+or(id(""),multi_emp_map) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_row) . ("\n" emp_row)*
 \end{lstlisting}
-% concat(ins(header),employees_lens) : emp_salaries $\Leftrightarrow$ emp_insurance
-%
-
-Data can also appear in multiple forms. For example,
-a list of employees with insurance information can either be empty or
-be a sequence of employee/insurance records: \ksf{This example is
-vacuous as you could eliminate the or in favor of the *.  Do somethign
-else to introduce or?}
+
+Data can also be concatenated, and so lens mappings can be concatenated.  This
+is achieved with the concat map.
 %
 \begin{lstlisting}
-or(id(""),employees_lens) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_ins) . ("\n" emp_ins)*
+concat(ins(header),multi_emp_map) : emp_salaries $\Leftrightarrow$ emp_insurance
 \end{lstlisting}
+%
+Two lenses, $\Lens_1$ and $\Lens_2$, can be concatenated as
+\lstinline{concat($\Lens_1$,$\Lens_2$)} or as the infix style
+\lstinline{$\Lens_1$ . $\Lens_2$}.
 
@@
 argument and synchronizes each element of the iteration with that lens. For
-example, if \lstinline{employee_lens} synchronizes a single employee's data,
-iterating that lens builds a lens that synchronizes a list of employees.
+example, if \lstinline{single_emp_map} synchronizes a single employee's data,
+iterating that lens builds a lens that synchronizes a list of tasks.
 %
 \begin{lstlisting}
-iterate(id("\n") . employee_lens) : ("\n" . emp_salary)$^*$ $\Leftrightarrow$ ("\n" . emp_ins)$^*$
+iterate(id("\n") . single_emp_map) : ("\n" . emp_salary)$^*$ $\Leftrightarrow$ ("\n" . emp_row)$^*$
 \end{lstlisting}
@@
 \begin{lstlisting}
-let name_lens = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
-let employee_lens = name_lens . disc(salary,company,"N/A","UNK") in
-let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
-let emp_lens = or(id(""),employees_lens) in
-ins(header) . employees_lens : emp_salaries $\Leftrightarrow$ emp_insurance
+let name_map = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
+let single_emp_map = name_map . disc(salary,company,"UNSET","UNSET") in
+let multi_emp_map = ins("\n") . single_emp_map . iterate(id("\n") . single_emp_map) in
+let emp_map = or(id(""),multi_emp_map) in
+ins(header) . multi_emp_map : emp_salaries $\Leftrightarrow$ emp_insurance
 \end{lstlisting}
   \caption{A lens, written in our language for building simple symmetric lenses
-    on string data, that synchronizes management and HR employee files.
+    on string data, that synchronizes management and hr employee files.
     Note that we can annotate our lenses with types, to confirm the lens
@@
 
-Using these combinators, we can build up a lens that synchronizes
-management and HR employee files.
-
-\ksf{This paragraph needs to be updated, if nothing else, to reflect
-the new example. Also, we haven't mentioned any typing rules so the
-first sentence doesn't make sense.  Perhaps the whole topic could be
-defered to later where we can explain why such a rule is necessary?}
-This lens uses one typing rule not previously
+Using these combinators, we can build up a lens that synchronizes HR and
+management employee files. This lens uses one typing rule not previously
 mentioned: type equivalence. While \lstinline{scheduled_job} is a equivalent to
@@
 
-Note that, while the synchronization task is quite simple, the
-requisite lens is relatively complex. Furthermore, this complexity is
-inherent in manually writing simple symmetric lenses because the
-lenses must precisely describe the portions of the data that should be
-connected and those that should be disconnected.  For disconnected
-pieces, the lens must further specifiy appropriate defaults.
+Note that, while the programmed lens should be quite simple, it was relatively
+difficult to write. We had to provide a lot of scaffolding to map between the
+desired formats, even though the idea was quite simple: align the times but
+leave the commands unaligned. Furthermore, this is inherent to manually writing
+these synchronization functions as a programmer must describe which components
+of the strings are aligned with which, and which components are scaffolding.
 
@@
 find a symmetric lens expression that is typed between the regular expression
-pair, and satisfies the input/output examples. We call such a lens
-a \emph{satisfying lens}.
-In this example, we wish to
+pair, and satisfies the input/output examples. In this example, we wish to
 synthesize a lens from \lstinline{emp_salaries} and \lstinline{emp_insurance},
-using the example data in Figure~\ref{fig:minimized-representations} as an
-input-output example from which to draw the required defaut values for 
-\Disconnect lenses.
-
-A challenge is that the simple symmetric lens combinators permit many
-well-typed lenses
-between a given pair of regular expressions. For example, 
-Figure~\ref{fig:example_lens} gives one possible lens between regular
-expressions  \lstinline{emp_salaries} and \lstinline{emp_insurance},
-but 
+using the strings in Figure~\ref{fig:minimized-representations} as an
+input-output example for \CreateR.
+
+However, our simple symmetric lens combinators permit many well-typed programs
+between two regular expressions. For example, one possible lens between
+\lstinline{emp_salaries} and \lstinline{emp_insurance} is the one in
+Figure~\ref{fig:example_lens}, but the lens
 %
@@
 %
-is another valid lens (where \lstinline{rep($\Regex$)} is a string in the
+is also a valid lens (where \lstinline{rep($\Regex$)} is a string in the
 language of $\Regex$). Furthermore, adding the synchronized files from
@@
 chosen input-output example on \PutR or \PutL, this problem continues presenting
-itself for the sublenses, and is highlighted even more on larger
-formats. 
-\ksf{I don't understand the above.  We shouldn't explain the issue in
-terms of \PutR and \PutL as we are focused on the combinators rather
-than the raw functions.}
-While
-we could ask programmers to provide detailed and precise specifications for their
+itself for the sublenses, and is highlighted even more on larger formats. While
+we could merely ask programmers to provide large, tight specifications for their
 synthesis tasks, we would prefer the algorithm to choose the right lens
@@
 
-But what is the best satisfying lens? We propose that the
-best satisfying lens is the satisyfing lens that connects as much data
-as possible between the two formats. More formally, we
-propose that the best satisfying lens is the satisfying lens that 
-minimizes the expected number of bits required to recover data in
-one format from synchronized data in the other.  
-In more detail, we say that two strings $s$ and $t$
-are \emph{synchronized according to lens $l$} if 
-$l.\PutR\ s\ t = t$ and $l.\PutL\ t\ s = s$ where the notation $l.\PutR$
-extracts the \PutR function from simple symmetric lens $l$; similarly
-for \PutL.
-We say that we can \emph{recover} $s$ from $t$ using bits $b$ and lens
-$l$ if we can reconstruct $s$ from $t$, $b$, and $l$.  For example,
-given the \IdentityLens{} lens, we can recover $s$ from $t$ using
-no bits because in this case, $s$ is just $t$.  In contrast, given
-the \Disconnect{} lens, we need enough bits to fully encode $s$ in
-order to recover it from $t$ because $t$ contains no useful
-information.  This calculation shows that if both \IdentityLens{}
-and \Disconnect{} are satisfying lenses, then \IdentityLens{} is
-better.
-
-The expected number of bits required to recover some data is a
-well-known concept in information theory: it is the \emph{entropy} of
-the data. Calculating the 
-entropy of some data requires a probability distribution over the
-space of possible values the data can take on.  Specifically,
-given a set $S$ and a probability distribution over that set $P
+But what is the best satisfying lens? Here, we take the view that the best satisfying lens is the
+one that aligns as much data as is permitted by the type system and the
+examples. More formally, we wish to minimize the expected number of bits
+required to recover data in one format from a synchronized string in the other.
+The expected number of bits is a well known concept in information theory: the
+\emph{entropy} of that data source.
+
+The entropy of a data source is based on the probability distribution of that
+data source. Given a set $S$, and a probability distribution over that set $P
 \OfType S \rightarrow \Reals$, the entropy of the $(S,P)$ pair is $-\Sigma_{s\in
-  S}P(s)\Log_2P(s)$.
-
-In our setting, we already have a way of expressing sets of data:
-regular expressions.  What we need is a way to express probability
-distributions over that data.
-To that end, we adopt \emph{stochastic regular expressions}\ksf{is
-there a citation?}, 
-which are regular expressions where each operator is annotated with a
-probability.  Stochastic regular expressions
-allow us to jointly specify the allowable data and give a probability
-distribution over it.  From this information, we can calculate the
-entropy of the stochastic regular expression.
-
-Given a lens $\Lens$ between two stochastic regular expressions $\Regex$ and
-$\RegexAlt$, we can calculate the expected number of bits required
-to recover a string $s$ in $\LanguageOf{\Regex}$ from a synchronized
-string $t$ in $\LanguageOf{\RegexAlt}$. 
-This expected value is just the conditional entropy of the $\Regex$ given
-$\RegexAlt$ and $\Lens$. 
-To ensure symmetry, we seek to find the lens that
-minimizes the sum of that entropy with the conditional entropy of $\RegexAlt$
-given $\Regex$ and $\Lens$.
-
-\ksf{This piece might better go somewhere deeper in the paper.}
-Our algorithm guesses a reasonable probability distribution from
+  S}P(s)\Log_2P(s)$. While we have ways of expressing sets of strings (with
+regular expressions), we do not have probability distributions. Here, we
+express these probability distributions jointly with the language through
+\emph{stochastic regular expressions}, regular expressions annotated with
+probabilities. Our algorithm guesses a reasonable probability distribution from
 the syntax, though if users manually provided stochastic regular expressions, or
@@
 we could instead use that.
+
+With sets and probability distributions, we can find the expected number of bits
+required to recover one string from a synchronized string. This expected value
+is the conditional entropy of the stochastic regular expression ($\Regex$) given
+another stochastic regular expression ($\RegexAlt$), and a lens that shows which
+strings are synchronized with which ($\Lens$). We aim to find the lens that
+minimizes the sum of that entropy with the conditional entropy of $\RegexAlt$
+given $\Regex$ and $\Lens$.
 
@@
 the disconnect lens, the cost of each side is the entropy of the data format, so
-the total cost is $\EntropyOf{\lstinline{emp_salary}} +
+the total cost is $\EntropyOf{\lstinline{emp_salary} } +
 \EntropyOf{\lstinline{emp_row}}$. \lstinline{emp_salary} contains strictly more
-imformation than \EntropyOf{\lstinline{": " . salary}}, and \lstinline{emp_row}
-contains strictly more information than \EntropyOf{\lstinline{": " . salary}},
-so by our metric \lstinline{single_emp_map} is preferable to the disconnect
-lens.
+imformation than $\EntropyOf{\lstinline{": " . salary}}$, and
+\lstinline{emp_row} contains strictly more information than
+\EntropyOf{\lstinline{": " . salary}}, so by our metric
+\lstinline{single_emp_map} is preferable to the disconnect lens.
 % end overview
@@
 \label{sec:implementation}
+% end implementation
+
+% begin evaluation
+\section{Evaluation}
+\label{sec:evaluation}
 We have implemented our symmetric DNF lenses and the synthesis of them as an
@@
 
-We have extended Optician with our symmetric synthesis algorithm.
-% end implementation
-
-% begin evaluation
-\section{Evaluation}
-\label{sec:evaluation}
+We have extended Optician with our symmetric synthesis algorithm. Users can
+write synthesis tasks alongside lens combinators, use those synthesis results in
+the manual writing of new lenses, and use previously defined lenses in the
+synthesis of new lenses.
+
+In this work, we aim to answer two primary research questions:
+\begin{itemize}
+\item Is our synthesis procedure efficient enough to be used in everyday
+  development?  Can it be used in place of the existing bijective lens synthesis
+  algorithm integrated with Boomerang?
+\item Is our information theoretic cost metric useful in finding good lenses.
+\end{itemize}
+
+\subsection{Benchmark Suite Construction}
+Our benchmarks came from four different sources.
+\begin{enumerate}
+\item We adapted \afm{8} data cleaning benchmarks from Flash Fill. In adopting
+  these benchmarks, we go beyond the data cleaning tasks, and synthesize
+  synchronization functions between the ad-hoc data and their cleaned forms.
+  Optician also adopted benchmarks from Flash Fill, but required alterations to
+  be in bijective form. With our symmetric enhancement, we Optician's four
+  benchmarks, and 4 new Flash Fill benchmarks with no alterations.
+\item We adapted \afm{29} benchmarks from Augeas.  These benchmarks are the same
+  as those present in Optician.  While these benchmarks had to be slightly
+  adopted for bijectivity, previous work addressed this lack of bijectivity by
+  synthesizing quotient bijective lenses~\cite{?}.
+\item We used \afm{8} benchmarks from synchronization tasks on synchronizing a
+  variety of real-life formats. These formats range from the synchronizing REST
+  and JSON web resource descriptions to synchronizing BibTeX and EndNote
+  citation descriptions.
+\end{enumerate}
+
+\subsection{Efficiency}
+
+To answer our first research question, we analyze the runtime of our algorithm,
+and compare it to the existing algorithm for synthesizing bijective lenses.
+
+We synthesized lenses in 4 modes:
+
+\begin{tabulary}{\linewidth}{rL}
+  \SSOpt{}: & We synthesize lenses using our symmetric synthesis algorithm with all optimizations enabled.\\
+  \SSNCOpt{}: & We synthesize lenses using our symmetric synthesis algorithm,
+                but compositional synthesis is not enabled.\\
+  \BSOpt{}: &  We synthesize lenses using the existing bijective lens synthesis
+              algorithm with all optimizations enabled\\
+  \BSNCOpt{}: & We synthesize lenses using the existing bijective lens synthesis
+                algorithm, but compositional synthesis is not enabled.\\
+\end{tabulary}\\
+
+We include running these algorithms without compositional synthesis to stress
+test the algorithms. While compositional synthesis makes the procedure easier,
+and demonstrates the potential to be used in arbitrarily hard problems, it is
+dependent on the user identifying important sublenses between subcomponents of
+the final formats. It is therefore interesting to see how many problems each
+algorithm can solve, when trying to synthesize ``in one go''.
+
+For each benchmark in our benchmark suite, we ran every mode on it with a
+timeout of 1 minute, and average the result over 5 runs. Figure~\ref{fig:times}
+summarizes the results of these tests.
 % end evaluation
@@
   \tag{\PutRL}
-  \frac{\PutRSymOf{(x,c)} = (y,c')}{\PutLSymOf{(y,c')} = (x,c')}
+  \frac{\PutRSymOf{(x,c)} = (y,c'')}{\PutLSymOf{(y,c')} = (x,c')}
 \end{equation}

### 2018-07-01 papers/SymmetricLenses/main.tex

--- 2018-07-01 papers/SymmetricLenses/main.tex
@@
 
+%Consider the trivial lens
+%\lstinline{disc(emp_salary,emp_row,rep(emp_salary),rep(emp_row))}. As this is
+%the disconnect lens, the cost of each side is the entropy of the data format, so
+%the total cost is $\EntropyOf{\lstinline{emp_salary}} +
+%\EntropyOf{\lstinline{emp_row}}}$. \lstinline{emp_salary} contains strictly more
+%imformation than \EntropyOf{\lstinline{": " . salary}}, and \lstinline{emp_row}
+%contains strictly more information than \EntropyOf{\lstinline{": " . salary}},
+%so by our metric \lstinline{single_emp_map} is preferable to the disconnect
+%lens.
+
 \ksf{Commented out entropy calculation; I don't think we need that
@@
 %lens.
+Consider the trivial lens
+\lstinline{disc(emp_salary,emp_row,rep(emp_salary),rep(emp_row))}. As this is
+the disconnect lens, the cost of each side is the entropy of the data format, so
+the total cost is $\EntropyOf{\lstinline{emp_salary} } +
+\EntropyOf{\lstinline{emp_row}}$. \lstinline{emp_salary} contains strictly more
+imformation than $\EntropyOf{\lstinline{": " . salary}}$, and
+\lstinline{emp_row} contains strictly more information than
+\EntropyOf{\lstinline{": " . salary}}, so by our metric
+\lstinline{single_emp_map} is preferable to the disconnect lens.
 % end overview

### 2018-07-01 papers/SymmetricLenses/main.tex
eval efficiency

@@
 
+\begin{figure}
+  \includegraphics{generated-graphs/times}
+  \caption{ Number of benchmarks that can be solved by a given algorithm in a
+    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
+    \BSOpt{} is the full bijective lens synthesis algorithm. \SSNCOpt{} is the
+    symmetric synthesis algorithm without using a library of existing lenses.
+    \BSNCOpt{} is the bijective synthesis algorithm without using a library of
+    existing lenses. Our symmetric synthesis algorithm is able to complete all
+    benchmarks in our benchmark suite in under 30 seconds, and is able to
+    complete most of the benchmarks without using compositional synthesis. }
+  \label{fig:times}
+\end{figure}
+
 For each benchmark in our benchmark suite, we ran every mode on it with a
-timeout of 1 minute, and average the result over 5 runs. Figure~\ref{fig:times}
-summarizes the results of these tests.
+timeout of 30 seconds, and average the result over 5 runs.
+Figure~\ref{fig:times} summarizes the results of these tests. We find that our
+algorithm is able to synthesize all of the benchmarks in our benchmark suite in
+under 30 seconds.
+
+The slowest completed benchmark for both the bijective and symmetric synthesis
+algorithms is \lstinline{xml_to_augeas.boom}, a benchmark that converts
+arbitrary XML up to depth 3 into a serialized version of the structured
+dictionary representation used in Augeas.  This benchmark takes 18.9 seconds for
+the symmetric synthesis algorithm to complete, and 9.3 seconds the bijective
+synthesis algorithm to complete.
 % end evaluation

### 2018-07-01 papers/SymmetricLenses/transform-data.py

@@ -1,0 +1,417 @@
+#!/usr/local/bin/python
+
+from __future__ import print_function
+from easyprocess import EasyProcess
+
+import os
+import csv
+from os.path import splitext, join
+import subprocess
+import sys
+import time
+import matplotlib
+import matplotlib as mpl
+mpl.use('pgf')
+import numpy as np
+import matplotlib.pyplot as plt
+
+plt.rc('font', size=10)
+plt.rc('legend', fontsize=10)
+plt.rcParams['text.usetex'] = True
+plt.rcParams['text.latex.preamble'] = '\usepackage{libertine},\usepackage[libertine]{newtxmath},\usepackage[T1]{fontenc}'
+
+generated_graphs_base = "generated-graphs/"
+transformed_data_base = "transformed-data/"
+
+def ensure_dir(f):
+    d = os.path.dirname(f)
+    if not os.path.exists(d):
+        os.makedirs(d)
+
+
+def print_usage(args):
+    print("Usage: {0} <file1>".format(args[0]))
+
+def retrieve_csv(filename):
+    csv_rows = []
+    with open(filename, 'rb') as csvfile:
+        reader = csv.DictReader(csvfile)
+        for row in reader:
+            csv_rows.append(row)
+    return csv_rows
+
+def project_column_from_csv(csv_obj, col_name):
+    return [r[col_name] for r in csv_obj]
+
+def write_to_filename(filename, s):
+    with open(filename, "wb") as f:
+        f.write(s)
+
+def generate_examples_required_graph(input_csv):
+    zero_count_ind = 0
+    one_to_five_count_ind = 1
+    six_to_ten_count_ind = 2
+    over_twenty_count_ind = 3
+
+    zero_count_text = "0"
+    one_to_five_count_text = "1-5"
+    six_to_ten_count_text = "6-10"
+    eleven_to_fifteen_count_text = "11-15"
+    sixteen_to_twenty_count_text = "16-20"
+    over_twenty_count_text = ">10"
+    ind_to_text = [zero_count_text,
+                   one_to_five_count_text,
+                   six_to_ten_count_text,
+                   over_twenty_count_text]
+
+    experimental_values = [0,0,0,0,]
+
+    def add_to_correct_group(count_values, n):
+        if n < 0.0:
+            raise Exception("SOMETHING WENT WRONG")
+        if n == 0.0:
+            count_values[zero_count_ind] = count_values[zero_count_ind]+1
+        elif n <= 5.0:
[…]

### 2018-07-02 papers/SymmetricLenses/main.tex
add eval and conclusion and do a big revision

@@
 %%   application-specific data storage files, and data cleaning tasks.
+TODO: abstract
 
@@
 \noindent
-\dpw{XXXXXX New attempt at intro starts here}
-
 In a big-data world, one will often find similar data stored in a
@@
 %a search through the space of possible computations that can be expressed as a lens.
-Recent research~\cite{optician}
-has exploited this fact by demonstrating that it is possible to synthesize relatively
-large, complex lenses from specifications.
-In particular, given types that describe source and target formats, there are synthesis
-algorithms that will find bijective lenses~\cite{optician}
-for medium-sized files, such as those housing linux
-application configurations.  \ksf{why medium sized?  You can have very
-large files with simple format descriptions that are bijective and for
-which our previous systems would have worked just fine.} Oftentimes, these algorithms find such lenses in less than
-a second, and without the aid of any examples at all, with just types describing the source and
-target formats.  \dpw{I would cite the ICFP paper (cite string: maina+:quotient-synthesis) as to appear but I'm not sure how to do that
-without possibly violating the spirit of double-blind reviewing. ie, if it hasn't appeared yet,
-how could we know the content if we aren't the authors?  Mentioning it in the 3rd person does not
-seem to suffice ...} \ksf{The ICFP 2018 website lists the paper; I
-think it is okay to cite it.  The reviewer can assume we contacted the
-authors for a copy.  We could ask Stephanie for a ruling if we want to
-be sure.  One of the principles of double blind as adopted by SIGPLAN
-is not to reduce the quality of the paper; omitting this citation
-would have that affect.}\bcp{+1}\saz{Stephanie says we should cite it in the third
-person, even if that's a bit awkward, but otherwise citing it is OK.}
+Recent research~\cite{optician} has exploited this fact by demonstrating that it
+is possible to synthesize relatively large, complex lenses from specifications.
+In particular, given types that describe source and target formats, there are
+synthesis algorithms that will find bijective lenses~\cite{optician} for files
+in ad-hoc formats, such as those housing linux application configurations.
+Oftentimes, these algorithms find such lenses in less than a second, and without
+the aid of any examples at all, with just types describing the source and target
+formats. \dpw{I would cite the ICFP paper (cite string:
+  maina+:quotient-synthesis) as to appear but I'm not sure how to do that
+  without possibly violating the spirit of double-blind reviewing. ie, if it
+  hasn't appeared yet, how could we know the content if we aren't the authors?
+  Mentioning it in the 3rd person does not seem to suffice ...} \ksf{The ICFP
+  2018 website lists the paper; I think it is okay to cite it. The reviewer can
+  assume we contacted the authors for a copy. We could ask Stephanie for a
+  ruling if we want to be sure. One of the principles of double blind as adopted
+  by SIGPLAN is not to reduce the quality of the paper; omitting this citation
+  would have that affect.}\bcp{+1}\saz{Stephanie says we should cite it in the
+  third person, even if that's a bit awkward, but otherwise citing it is OK.}
 
@@
 the other does not. Indeed, all of our benchmarks taken from Flash
-Fill~\cite{flashfill} and all of our benchmarks that synchronize file formats
-contain data present in one format that isn't present in the other.
-Consequently, past tools that search for strict bijections will rarely be
-applicable in practice.
+Fill~\cite{flashfill} and all of our benchmarks that synchronize between two
+ad-hoc file formats. Consequently, past tools that search for strict bijections
+will rarely be applicable in practice.
 
@@
 
-The first difficulty was pragmatic:  symmetric lenses actually
-usually operate over three structures, the source $S$, the target $T$ and a
-complement $C$~\cite{symmetric-lenses}.
-The latter object contains all the information not present in one of $S$ or $T$.  Logistically,
-the complement object must be stored away somewhere and infrastructure must be developed
-to manage it.  
-Perhaps more relevant though, is that complements complicate specifications, even when
-those specifications are just examples.  The issues is that one cannot
-just give examples of source and target; one must also give guidance on how to manage
-the ``hidden state.''  In order to avoid such complexities, we defined a new
-variant of symmetric lenses, which we call \emph{simple symmetric lenses}.  Our new
-lenses do not require complements; they only need source and target
-data as inputs. \ksf{This seems magic.  Say something about how it is
-possible/under what circumstances it is possible to not need a
-complement?
-At the price of some expressiveness, our new lenses eliminate the need
-for complements; they require only source and target data as inputs.
-}
+The first difficulty was pragmatic: symmetric lenses actually usually operate
+over three structures, the source $S$, the target $T$ and a complement
+$C$~\cite{symmetric-lenses}. The latter object contains all the information not
+present in one of $S$ or $T$. Logistically, the complement object must be stored
+away somewhere and infrastructure must be developed to manage it. Perhaps more
+relevant though, is that complements complicate specifications, even when those
+specifications are just examples. The issues is that one cannot just give
+examples of source and target; one must also give guidance on how to manage the
+``hidden state.'' In order to avoid such complexities, we defined a new variant
+of symmetric lenses, which we call \emph{simple symmetric lenses}. At the price
+of some expressiveness, our new lenses do not require complements; they only
+need source and target data as inputs.
 
 In order to characterize the expressive power of simple symmetric lenses
-theoretically, we prove that simple symmetric lenses
-include all (and only) those symmetric lenses that satisfy
-an intuitive property called \emph{forgetfulness}.  In other words, they are the
-complete set of symmetric lenses that do not require a ``memory'' to retain data
-from past instances of $S$ or $T$ when making a round trip.\dpw{fix that to make the intuitive
-explaination sufficiently accurate}
-From a pragmatic perspective, we justify our design choices by extending
-the Boomerang implementation~\cite{Boomerang} with symmetric lenses (which
-had never before been implemented\bcp{Confusing: we implement only simple
-  ones, but it was the earlier full form that had never been implemented}),
-applying them successfully to a range of real-world applications,
-and demonstrating that they are compatible, not just with basic bijective
-combinators,\bcp{confusing: they are not exactly ``compatible'' with the
-  bijective combinators---they enrich them}
-but with a rich set of advanced features including quotient lenses and matching lenses.
-
-The second difficulty with synthesizing symmetric lenses is more fundamental\dpw{better word?}:
-While the number of bijective lenses
-between two related formats is often tiny, the number of symmetric lenses is often
-large.  In practice, the difference in cardinality means that when a search algorithm selects the smallest
-bijective lens it finds, that lens has a good chance of being the
-one a user wants.  In contrast, and to our surprise, our experiments show \dpw{can we have an experiment that
-shows this?} that
-when a search algorithm selects the smallest symmetric lens, it typically fails
-to find the lens the user wants.  The flexibity these lenses admit when it comes to
-eliding information generates an explosion of possible transformations.
-To disambiguate between the many possible lenses, we need
-new principles for identifying the ``likely'' ones, and we need
-new algorithms to search the space of program transformations using these principles.
-\dpw{Were the bijective lenses chosen using a ``smaller'' criteria?  I
-forget... if not edit the prior paragraph.}\dpw{Ideally, we could back up the claim
-that there are many symmetric lenses and not many bijective lenses by graphing the
-number of such lenses somehow.  eg, in a past synthesis paper with Steve (POPL 2016)
-we graphed
-the untyped terms of type $\mathsf{int}\rightarrow\mathsf{int}$.  I wonder if we can
-do something similar here.}
-
-To overcome this second difficulty, we turned to information theory
-to try to find a way to characterize the likely and unlikely lenses.  Intuitively, a ``likely'' lens
-is one that propagates a lot of information from source to target and also from target
-back to source.  Conversely, given the source, the target is ``unlikely'' to require very much
-additional information in order to reconstruct it---if it were\bcp{were
-  what?  likely, or unlikely?}, then one is unlikely to
-be attempting to synthesize a lens between the two formats\bcp{why? even if
-  two big formats only share a little bit of common information, mightn't
-  one want a lens between them?}!  In order to implement
-this intuition, we turn to \emph{stochastic regular expressions} (SREs).
-Stochastic regular expressions define a set of strings, as do ordinary regular expressions,
-but they also provide the expected distribution of strings within that set.  Such
-distributions allow us to compute the amount of information gained when presented
-an element of the set.  For example, the language of the SRE $a |_{0.5} b$ contains two elements
-(the strings $a$ and $b$).  The probability 0.5 annotating the union indicates that both
-elements are equally likely to appear.  Consequently, if we observe 
-an element of that language (either $a$ or $b$), we will gain one bit of information.  Said another way,
-one bit, together with the SRE, suffices to recover the string in
-question.\bcp{Is this implying that, if the annotation were
-  different, we'd need something other than one bit to recover the string??}
-On the other hand, consider the language of the SRE $c$.
-It contains just one element (the string $c$).  If we observe that our algorithm
-generates $c$, we gain zero bits of information --- the string was already determined.
-Hence\bcp{?}, a lens is ``likely,'' in a technical sense, if relative to the target SRE $T$,
-very little information from $t$ is required to generate it.
-
-With simple symmetric lenses and SREs in hand, we develop a new algorithm for synthesizing likely lenses.
-This algorithm is constructed from a pair of collaborating search procedures.
-Given source and target SREs $S$ and $T$, the first algorithm searches for two \emph{related} SREs
-$S'$ and $T'$ that are \emph{compatible} (a heuristic that estimates the likelihood
-our second algorithm will succeed).  $S'$ and $T'$ are related in the sense
-that they define the same regular sets as $S$ and $T$ (respectively) and the distribution
-of strings is identical.  Indeed, we guarantee these two properties by implementing
-a search procedure that obeys the star semi-ring axioms of regular expression equivalence,
-and which we show does not alter the probability distributions of the languages.
-
-Given the two compatible SREs,
-the second algorithm uses a greedy, type-directed algorithm find a likely lens between them.
-If algorithm estimates that the lens it has discovered is more likely than
-any future lens it can possibly find, the search terminates.  Otherwise,
-this lens is stored away (if it is the best found so far) and the search continues with
-algorithm 1 again.
+theoretically, we prove that simple symmetric lenses include all (and only)
+those symmetric lenses that satisfy an intuitive property called
+\emph{forgetfulness}. In other words, they are the complete set of symmetric
+lenses that do not require external ``memory'' to recover data from past
+instances of $S$ or $T$ when making a round trip -- they only need the most
+recent instance. From a pragmatic perspective, we justify our design choices by
+extending the Boomerang implementation~\cite{Boomerang} with simple symmetric
+lenses applying them successfully to a range of real-world applications, and
+demonstrating that they can enrich, not just with basic bijective combinators,
+but with a rich set of advanced features including quotient lenses and matching
+lenses.
+
+The second difficulty with synthesizing symmetric lenses is more
+fundamental\dpw{better word?}: While the number of bijective lenses between two
+related formats is often tiny, the number of symmetric lenses is often large. In
+practice, the difference in cardinality means that when a search algorithm
+selects the first bijective lens it finds, that lens has a good chance of being
+the one a user wants. In contrast, and to our surprise, our experiments show
+that if our search algorithm selects the first symmetric lens it finds, it
+typically fails to find the lens the user wants. The flexibity these lenses
+admit when it comes to eliding information generates an explosion of possible
+transformations. To disambiguate between the many possible lenses, we need new
+principles for identifying the ``likely'' ones, and we need new algorithms to
+search the space of program transformations using these principles. \dpw{Were
+  the bijective lenses chosen using a ``smaller'' criteria? I forget... if not
+  edit the prior paragraph.}\afm{no, chosen using a ``first'' criteria.
+  Edited.}\dpw{Ideally, we could back up the claim that there are many symmetric
+  lenses and not many bijective lenses by graphing the number of such lenses
+  somehow. eg, in a past synthesis paper with Steve (POPL 2016) we graphed the
+  untyped terms of type $\mathsf{int}\rightarrow\mathsf{int}$. I wonder if we
+  can do something similar here.}\afm{What about lenses between ``a''* and
+  ''b``*}
+
+To overcome this second difficulty, we turned to information theory to try to
+find a way to characterize the likely and unlikely lenses. Intuitively, a
+``likely'' lens is one that propagates a lot of information from source to
+target and also from target back to source. Conversely, given the source, the
+target is ``unlikely'' to require very much additional information in order to
+reconstruct it---if it were\bcp{were what? likely, or unlikely?}, then one is
+unlikely to be attempting to synthesize a lens between the two formats\bcp{why?
+  even if two big formats only share a little bit of common information,
+  mightn't one want a lens between them?}! In order to implement this intuition,
+we turn to \emph{stochastic regular expressions} (SREs). Stochastic regular
+expressions define a set of strings, as do ordinary regular expressions, but
+they also provide the expected distribution of strings within that set. Such
+distributions allow us to compute the amount of information gained when
+presented an element of the set. For example, the language of the SRE
+$\lstinline{"a"} |_{0.5}
+\lstinline{"b"}$ contains two elements (the strings
+\lstinline{"a"} and
+\lstinline{"b"}). The probability 0.5 annotating the union indicates that both
+elements are equally likely to appear. Consequently, if we observe an element of
+that language (either \lstinline{"a"} or \lstinline{"b"}), we will gain one bit
+of information. Said another way, one bit, together with the SRE, suffices to
+recover the string in question.\bcp{Is this implying that, if the annotation
+  were different, we'd need something other than one bit to recover the
+  string??}
+\afm{From an information theoretic perspective, yes, because that perspective
+  views it as the amount of information if we were communicating a stream of
+  these choices.} On the other hand, consider the language of the SRE \lstinline{c}. It contains
+just one element (the string $c$). If we observe that our algorithm generates
+$\lstinline{"c"}$, we gain zero bits of information --- the string was already determined.
+Hence\bcp{?}, a lens is ``likely,'' in a technical sense, if relative to the
+target SRE $T$, very little information from $t$ is required to generate it.
+
+\begin{figure}
+  \includegraphics[width=.5\textwidth]{high-level-algorithm.pdf}
+  \caption{Schematic Diagram for our symmetric lens synthesis algorithm. Regular
+    expressions \Regex and \RegexAlt, and a set of examples \Examples, are
+    provided as input. \RXSearch searches through stochastic regular expression
+    pairs, equivalent to the originals, and proposes them to \GreedySynth.
+    \GreedySynth finds a lens, typed between the generated equivalent pairs.
+    When \RXSearch determines it has an optimal lens, it returns that lens.}
+  \label{fig:high-level-algorithm}
+\end{figure}
+
+With simple symmetric lenses and SREs in hand, we develop a new algorithm for
+synthesizing likely lenses, based on the previous algorithm for synthesizing
+bijective lenses. This algorithm is constructed from \RXSearch and \GreedySynth,
+a pair of collaborating search procedures, shown in
+Figure~\ref{fig:high-level-algorithm}. Given source and target SREs $S$ and $T$,
+the \RXSearch searches for two \emph{related} SREs $S'$ and $T'$ that are
+\emph{compatible} (a heuristic that estimates the likelihood our second
+algorithm will succeed). $S'$ and $T'$ are related in the sense that they define
+the same regular sets as $S$ and $T$ (respectively) and the distribution of
+strings is identical. Indeed, we guarantee these two properties by implementing
+a search procedure that obeys the star semi-ring axioms of regular expression
+equivalence, and which we show does not alter the probability distributions of
+the languages.
+
+Given the two compatible SREs, \GreedySynth uses a greedy, type-directed
+algorithm find a likely lens between them. If algorithm estimates that the lens
+it has discovered is more likely than any future lens it can possibly find, the
+search terminates. Otherwise, this lens is stored away (if it is the best found
+so far) and the search continues with \RXSearch again.
 
@@
 \begin{itemize}
-\item We define the space of simple symmetric lenses, which lies between asymmetric lenses and full symmetric lenses and has the useful property that they don't have any internal state.
-We characterize
-the simple lenses semantically as the subset of full symmetric lenses for which a ``forgetfulness''
-property holds. We also analyze the gap between simple and full symmetric lenses empirically and argue that
-simple ones suffice for many applications.
+\item We define the space of simple symmetric lenses, which lies between
+  asymmetric lenses and full symmetric lenses and has the useful property that
+  they don't have any internal state (\S\ref{sec:overview}). We characterize the
+  simple lenses semantically as the subset of full symmetric lenses for which a
+  ``forgetfulness'' property holds (\S\ref{sec:related}). We also analyze the
+  gap between simple and full symmetric lenses empirically and argue that simple
+  ones suffice for many applications (\S\ref{sec:evaluation}).
 \item We show how to synthesize simple symmetric lenses (including all asymmetric lenses as a special case),
 using a novel application of stochastic regular expressions to guide the search process.  In particular, we
-show that the RE star semi-ring equivalences remain valid when their semantics are enriched by distributions.
+show that the RE star semi-ring equivalences remain valid when their semantics
+are enriched by distributions (\S\ref{sec:synthesis}).
 \item We extend the Boomerang implementation with simple symmetric lenses,
@@
   extraction, and file synchronization tasks and demonstrate that we can
-  synthesize all of them in under half a minute.
+  synthesize all of them in under half a minute (\S\ref{sec:evaluation}).
 \end{itemize}
@@
 \noindent
-\dpw{XXXXXX New attempt at intro ends here}
 
 % state of the world
-\bcp{Not very compelling:}Similar data can come in varieties of formats: scheduled jobs can be stored as
-cron jobs for Linux or as Scheduled Tasks for Windows; web service
-resources can be provided to clients in legacy or modern formats;
-calendar events can be stored in CSV files or ICS files. Oftentimes,
-files should be synchronized across these formats: when I edit a task
-to run more often on my Linux machine, I want this task to run more
-often on my Windows machine; when I update data through a legacy API,
-I want those changes to be reflected when I retrieve it through a
-modern API; when I create a new calendar event in my ICS file, I want
-my CSV calendar files to be updated to reflect this change.
-
-Unfortunately, writing such synchronizers is quite difficult; we must write
-parsers both formats, specify how the parsed data is converted into the other
-format, and specify how to propagate updates from one format to the other.
-Bidirectional languages ease the difficulty of writing such synchronizers,
-because a single, well-typed program expresses a \emph{lens}---a group of
-conversion functions guaranteed to satisfy round-tripping laws. However,
-bidirectional languages are themselves difficult to program in, often requiring
-users to permute data in fiddly ways while satisfying the type system's complex
-unambiguity constraints.
-
-A better approach is to \emph{synthesize} such conversion functions. However,
-prior work on synthesizing lenses assumes a bijection between the source and
-target, which is a severe limitation in practice. Oftentimes, fields present in
-one format are not present in the other: the commands scheduled to run on Linux
-machines are typically different than the commands scheduled to run on Windows
-machines; updated APIs can expose new information, or omit legacy information;
-ICS files contain a URL field that Google Calendar CSV files do not expose.
-
-\emph{Symmetric lenses}~\cite{symmetric-lenses} formalize this richer class of
-non-bijective conversion functions by using additional state, called the
-\emph{complement}, to help with synchronization. While increasing
-expressiveness, the existence of the complement makes symmetric lenses difficult
-to use in real-world synchronization tasks, because lens programs must maintain
-the synchronized files as well as the complements. The complement further
-complicates synthesis, because input-output examples no longer suffice; instead
-edit sequences must be provided.
-
-In this paper, we contrain symmetric lenses to \emph{simple symmetric lenses}.
-These constrained lenses do not use a complement, and only use the files that we
-aim to synchronize as state. We identify a property, \emph{forgetfulness}, that
-identifies the subset of symmetric lenses that can be written as simple
-symmetric lenses. We provide combinators for building simple symmetric lenses on
-string data, and integrate these combinators with Boomerang, a language for
-bidirectional programming on strings. Both Boomerang and our extension type lens
-terms between pairs of regular expressions. The typing judgment $\Lens : \Regex
-\Leftrightarrow \RegexAlt$ means the lens term $\Lens$ encodes synchronizing
-functions between the format described by $\Regex$ and the format described by
-$\RegexAlt$.
-
-We provide a tool for synthesizing simple symmetric lenses when given a suite of
-input-output examples and a pair of regular expressions describing the data
-formats. The synthesis algorithm uses type-directed synthesis
-on the two regular expressions to synthesize a lens that is well-typed between
-the provided regular expressions and satisfies the provided input-output
-examples.
-
-\saz{This paragraph is hard to follow --- maybe introduce $\ell : S
-  \Leftrightarrow T$ and use the names $S$ and $T$?}\bcp{The names make it a
-bit better, but it is all still rather abstract.  Indeed, the whole intro is
-quite abstract---could it not be driven by an example?} A primary difficulty is
-that there are a relatively many well-typed simple symmetric lenses. If we are
-trying to find a lens between $\Regex$ and $\RegexAlt$, there is now a choice:
-do $\Regex$ and $\RegexAlt$ represent the same data, and should be aligned
-(changes to data in the language of $\Regex$ are propagated to synchronized
-strings in the language of $\RegexAlt$), or should it be left unaligned (changes
-to data in $\Regex$ update synchronized strings in $\RegexAlt$). Or instead
-should certain components of $\Regex$ be aligned with certain components of
-$\RegexAlt$, with the other components left unaligned.
-
-Operating under the assumption that users wish to align as much data as
-possible, our algorithm tries to align as much data as the types and examples
-allow. This task gets harder when there are more complex decisions involved. For
-example, if there are two fields in one format, and one in the other, which field
-in the first format should be aligned to the sole field of the second format? Or
-would it be better for the field on the right to be aligneded to a mish-mash
-of subcomponents from each field?\bcp{Wouldn't these questions be answered
-  by examples?} To answer these questions, we develop a cost
-metric for reasoning about which lens aligns the most information.
-
-In particular, the \emph{cost} of a lens is the expected number of bits required to
-recover a string in one format when a synchronized string in the other format is
-provided. In this model, bijective lenses have zero cost, and as more
-information is left unsychronized, the cost is increased.
-
-Our model is based on \emph{stochastic regular expressions}~\cite{?}, which
-represent both a language and a probability distribution over that language
-simultaneously.  We develop a way to syntactically infer the expected
-information content (or \emph{entropy}) of an unambiguous regular
-expression. This entropy is used to calculate the expected information content
-to recover a string, when given a string it is synchronized with.\bcp{I
-know we're going to get to more technical detail later, but at this point I
-am confused: a string matching a RE can in general be arbitrarily long, so
-what kind of sense does it make to talk about its ``expected'' information content?}
-
-\begin{figure}
-  \includegraphics[width=.5\textwidth]{high-level-algorithm.pdf}
-  \caption{Schematic Diagram for \SOptician. Regular expressions \Regex and
-    \RegexAlt, and a set of examples \Examples, are provided as input. \RXSearch
-    searches through stochastic regular expression pairs, equivalent to the
-    originals, and proposes them to \GreedySynth. \GreedySynth finds a lens,
-    typed between the generated equivalent pairs. When \RXSearch determines it
-    has an optimal lens, it returns that lens.}
-  \label{fig:high-level-algorithm}
-\end{figure}
-
-\bcp{It gets rather technical and incomprehensible for the next three
-  paragraphs...}
-
-We develop an algorithm that synthesizes a low-cost lens, shown in
-Figure~\ref{fig:high-level-algorithm}. This algorithm consists of two
-communicating synthesizers, one that proposes candidate stochastic regular
-expression pairs that it thinks will align well, and one that greedily searches
-for a lens between those two regular expressions that minimizes cost and does
-not traverse star-related equivalences\bcp{???}. When the first algorithm is content with
-the lens it has found, it returns that lens.
-
-\bcp{Make clear that these algorithms are based on earlier / published ones.}Our first algorithm, \RXSearch, proposes candidate stochastic regular
-expressions. These candidate stochastic regular expressions are generated by
-traversing \emph{star semiring} equivalences on stochastic regular expressions.
-These proposed stochastic regular expression pairs have equivalent languages and
-probability distributions as the original ones: we prove that the star-semiring
-rewrites this synthesizer applies are sound rewrites on stochastic regular
-expressions.
-
-Our second algorithm, \GreedySynth, utilizes alternative languages for
-stochastic regular expressions and lenses, \emph{stochastic DNF regular
-  expressions} and \emph{symmetric DNF lenses}. While these languages are hard
-to read and understand, they are highly normalized and permit an efficient
-type-directed synthesis algorithm. We develop our cost metric on DNF lenses, and
-provide a greedy algorithm that tries to find a symmetric DNF lens of minimal
-cost. The synthesized symmetric DNF lens is then converted to a lens in the
-original simple symmetric lens language.
-
-We evaluate our algorithm on a benchmark suite consisting of synchronizing file
-formats from configuration files, application-specific data storage files, and
-data cleaning tasks. We find that our algorithm is able to synthesize each of
-these lenses in under 20 seconds.
-
-In summary, our contributions are:
-\begin{enumerate}
-\item We develop \emph{simple symmetric lenses} a new formulation for
-  bidirectional programs that can project data from both sides~(\S\ref{sec:overview}),
-  and describe a language that expresses such lenses on string
-  data~(\S\ref{sec:ssl}).  \bcp{A new language, or a variant of
-    Hofmann/Pierce/Wagner?} 
-\item We develop a means to traverse stochastic regular expression equivalences,
-  while preserving probability distributions and
-  languages~(\S\ref{subsec:stoch-rx}).
-\item We develop stochastic DNF regular expressions and DNF
-  lenses~(\S\ref{subsec:greedy-synth}). We develop a sound algorithm for
-  converting stochastic regular expressions to stochastic DNF regular
-  expressions and a information-theoretic cost metric on DNF lenses.
-\item We extended Boomerang with our simple symmetric lens combinators, and
-  integrated our algorithm with the Optician synthesis system within Boomerang,
-  allowing synthesis tasks as Boomerang
-  expressions~(\S\ref{sec:implementation}).
-\item We evaluate our algorithm on a benchmark suite consisting of synchronizing
-  file formats from configuration files, application-specific data storage
-  files, and data cleaning tasks (\S\ref{sec:evaluation}).  \bcp{... and
-    conclude what?}
-\end{enumerate}
 % end introduction
@@
 
-\ksf{Commented out entropy calculation; I don't think we need that
-level of detail; leaving the text in case we change our mind :-)}
 %Not sure we need this level of detail after all.
@@
 
+\section{Stochastic Regular Expressions}
+\label{sec:srx}
+Important to our algorithm is finding the expected number of bits needed to
+recover a string in one data source from a synchronized string in the other data
+source. To do this, we must first develop a probability model for our language.
+We do this with \emph{stochastic regular expressions}, regular expressions
+annotated with probability information. With this, we can jointly express a
+language and a probability distribution over that language.
+
+\begin{center}
+  \begin{tabular}{rcl}
+    \Regex{},\RegexAlt{}
+    & \GEq{} & $\String$ \\
+    & \GBar{} & $\emptyset$ \\
+    & \GBar{} & $\PRegexStar{\Regex}{p}$ \\
+    & \GBar{} & $\RegexConcat{\Regex_1}{\Regex_2}$ \\
+    & \GBar{} & $\PRegexOr{\Regex_1}{\Regex_2}{p}$
+  \end{tabular}
+\end{center}
+
+The language is defined like the regular expression without probability
+annotations would be.  The probabilities are defined as follows:
+\begin{center}
+  \begin{tabular}{rcl}
+    $\ProbabilityOf{\String}{\String''}$
+    & =
+    & $\begin{cases*}1 & if $\String = \String''$\\ 0 & otherwise\end{cases*}$ \\
+
+    $\ProbabilityOf{\emptyset}{\String}$
+    & =
+    & $0$ \\
+
+    $\ProbabilityOf{\RegexConcat{\Regex_1}{\Regex_2}}{s}$
+
+    & =
+    & $\Sigma_{\String = \String_1\String_2}\ProbabilityOf{\Regex_1}{\String_1}*\ProbabilityOf{\Regex_2}{\String_2}$ \\
+
+    $\ProbabilityOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability}}{\String}$
+    & =
+    & $\Probability * \ProbabilityOf{\Regex_1}{\String} +
+      (1-\Probability) * \ProbabilityOf{\Regex_2}{\String}$\\
+
+    $\ProbabilityOf{\PRegexStar{\Regex}{\Probability}}{\String}$
+    & =
+    & $\Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\Regex}{\String_i}$\\
+  \end{tabular}
+\end{center}
+In the stochastic regular expression
+$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation \Probability
+represents the likelihood the string is in $\Regex_1$. In the stochastic regular
+expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability
+how much more likely having a string involving more than $n$ iterations is than
+having a string involving $n$ iterations.
+
+\subsection{Stochastic Regular Expression Equivalences}
+$\RXSearch$ needs to be able to find equivalent stochastic regular expressions.
+However, as of yet, there is no syntactic means to reason about stochastic
+regular expression equivalence. We extend the \emph{star-semiring} equivalences
+to stochastic regular expressions.
+
+\begin{center}
+  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
+    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
+    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
+    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
+    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
+    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{(1-\Probability_1)*\Probability_2})}{\Probability_1*\Probability_2} & \OrAssociativityRule{}  \\
+    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
+    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
+    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
+    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
+    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
+    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability} & \UnrollstarLeftRule{} \\
+    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{\Probability} & \UnrollstarRightRule{} 
+  \end{tabular}
+\end{center}
+
+\begin{theorem}
+  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
+  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
+\end{theorem}
+
+\subsection{Stochastic Regular Expression Entropy}
+The entropy of a stochastic regular expression can be computed directly from its
+syntax, when the stochastic regular expression is unambiguous and contains only
+nonempty subcomponents.
+\begin{center}
+  \begin{tabular}{rcl}
+    $\EntropyOf{s}$
+    & =
+    & 1\\
+
+    $\EntropyOf{\PRegexStar{\Regex}{\Probability}}$
+    & =
+    & $\frac{\Probability}{1-\Probability}(\EntropyOf{\Regex} - \Log_2\Probability)
+      - \Log_2(1-\Probability)$\\
+
+    $\EntropyOf{\PRegexConcat{\Regex}{\RegexAlt}}$
+    & =
+    & \EntropyOf{\Regex} + \EntropyOf{\RegexAlt}\\
+
+    $\EntropyOf{\PRegexOr{\Regex}{\RegexAlt}{\Probability}}$
+    & =
+    & $\Probability*(\EntropyOf{\Regex} - \Log_2(\Probability)) + (1-\Probability)*(\EntropyOf{\RegexAlt} - \Log_2(1-\Probability))$\\
+  \end{tabular}
+\end{center}
+We have proven that this syntactic computation corresponds to the entropy of the
+stochastic regular expression.
+
+\begin{theorem}
+  \label{thm:correct_entropy}
+  If $\Regex$ is unambiguous and contains no empty subcomponents,
+  $\EntropyOf{\Regex}$ is the entropy of $\Regex$.
+\end{theorem}
+
+% \caption{Regular Expression Equivalences}
+% \label{fig:regex-equivalence-rules}
+
 % begin ssl
@@
 \label{sec:ssl}
-\sam{Perhaps we shoud introduce stochastic regular expressions first before introducing the lens combinators?
-This is because the type equivalence combinator mentions regexp equivalence, but this has not been introduced by this point in the paper.}
 We now give a formal description of all the combinators in our symmetric lens
-language. These lens terms are typed between pairs of regular expressions: the
-typing judgment $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$ means that
-$\Lens$ is a well typed lens between $\LanguageOf{\Regex}$ and
-$\LanguageOf{\RegexAlt}$.
+language. These lens terms are typed between pairs of stochastic regular
+expressions: the typing judgment $\Lens \OfType \Regex \Leftrightarrow
+\RegexAlt$ means that $\Lens$ is a well typed lens between $\LanguageOf{\Regex}$
+and $\LanguageOf{\RegexAlt}$. These lens terms can also be typed between regular
+expressions, but we present in terms of stochastic regular expressions to assist
+in describing the entropy of the lenses.
 
@@
     $\CreateL{} \App \StringAlt'''''$ & = & $\String$\\
-    $\PutR{} \App \String'' \App \StringAlt'$ & = & $\StringAlt'$\\
+    $\PutR{} \App \String''' \App \StringAlt'$ & = & $\StringAlt'$\\
     $\PutL{} \App \StringAlt' \App \String'$ & = & $\String'$
@@
     \Lens \OfType \Regex \Leftrightarrow \RegexAlt\\
-    \Regex \equiv \Regex'\\
-    \RegexAlt \equiv \RegexAlt'
+    \Regex \SSREquiv \Regex'\\
+    \RegexAlt \SSREquiv \RegexAlt'
   }
@@
 annotations, and for making well-typed compositions.
+
+\subsection{Lens Costs}
+Our cost metric is based on the expected amount of information required to
+recover a string in one data format from the other. We use the function
+$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate the expected amount
+of information required to recover a string in $\Regex$ from a string in
+$\RegexAlt$, synchronized by $\Lens$. We use the function
+$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to calculate the expected amount
+of information required to recover a string in $\RegexAlt$ from a string in
+$\Regex$, synchronized by $\Lens$.
+\begin{center}
+  \begin{tabular}{rcl}
+    $\LEntropyOf{\Regex \Given \IdentityLensOf{\Regex}, \Regex}$
+    & =
+    & 0\\
+
+    $\LEntropyOf{\Regex \Given \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, \RegexAlt}$
+    & =
+    & \EntropyOf{\Regex}\\
+
+    $\LEntropyOf{\PRegexStar{\Regex}{\Probability} \Given \IterateLensOf{\Lens}^, \PRegexStar{\RegexAlt}{\ProbabilityAlt}}$
+    & =
+    & $\frac{\ProbabilityAlt}{1-\ProbabilityAlt}(\LEntropyOf{\Regex \Given \Lens, \RegexAlt})$\\
+
+    $\LEntropyOf{\Regex_1 \Concat \Regex_2 \Given \ConcatLensOf{\Lens_1}{\Lens_2}, \RegexAlt_1 \Concat \RegexAlt_2}$
+    & =
+    & $\LEntropyOf{\Regex_1 \Given \Lens_1^\leftarrow, \RegexAlt_1} + \LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2}$\\
+
+    $\LEntropyOf{\Regex_1 \Concat \Regex_2 \Given \SwapLensOf{\Lens_1}{\Lens_2}, \RegexAlt_2 \Concat \RegexAlt_1}$
+    & =
+    & $\LEntropyOf{\Regex_1 \Given \Lens_1^\leftarrow, \RegexAlt_1} + \LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2}$\\
+
+    $\LEntropyOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability} \Given \OrLensOf{\Lens_1}{\Lens_2}, \PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}}$
+    & =
+    & $\ProbabilityAlt*(\LEntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt_1}) + (1-\ProbabilityAlt)*(\LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2})$\\
+
+    $\LEntropyOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability} \Given \OrLensOf{\Lens_1}{\Lens_2}, \PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}}$
+    & =
+    & $\ProbabilityAlt*(\LEntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt_1}) + (1-\ProbabilityAlt)*(\LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2})$\\
+
+    $\LEntropyOf{\Regex \Given \MergeROf{\Lens_1}{\Lens_2}, \PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}}$
+    & =
+    & $\ProbabilityAlt*(\LEntropyOf{\Regex \Given \Lens_1, \RegexAlt_1}) + (1-\ProbabilityAlt)*(\LEntropyOf{\Regex \Given \Lens_2, \RegexAlt_2})$\\
+
+    $\LEntropyOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability} \Given \MergeLOf{\Lens_1}{\Lens_2}, \RegexAlt}$
+    & =
+    & $\Probability*(\LEntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt} - \Log_2(\Probability)) + $\\
+    &
+    & $(1-\Probability)*(\LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt} - \Log_2(1 - \Probability))$\\
+
+    $\LEntropyOf{\RegexAlt \Given \InvertOf{\Lens}, \Regex}$
+    & =
+    & $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$\\
+  \end{tabular}
+\end{center}
+$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ is defined symmetrically.  These
+functions calculate the expected number of bits to recover one data format from
+a synchronized string in the other format.
+The cost of a lens, $c(\Lens \OfType \Regex \Leftrightarrow \RegexAlt) =
+\LEntropyOf{\Regex \Given \Lens, \RegexAlt} + \REntropyOf{\RegexAlt \Given
+  \Lens, \Regex}$ is the
+sum of recovering the left format from the right, and the right from the left.
+We have proven theorems demonstrating the calculated entropy corresponds to the
+actual conditional entropy to recover the data.
+
+\begin{theorem}
+  Let $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$, where $\Lens$ does not
+  include composition, $\Regex$ and $\RegexAlt$ and unambiguous, and neither
+  $\Regex$ nor $\RegexAlt$ contain any empty subcomponents.
+  \begin{enumerate}
+  \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ calculates the expected
+    information content of $\SetOf{s \SuchThat s \in \LanguageOf{\Regex}}$,
+    given $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt} \BooleanAnd
+      \Lens.\PutLOf{t}{s} = s}$
+  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ calculates the expected
+    information content of $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$,
+    given $\SetOf{s \SuchThat s \in \LanguageOf{\Regex} \BooleanAnd
+      \Lens.\PutROf{s}{t} = t}$
+  \end{enumerate}
+\end{theorem}
+
+Note that this method doesn't work for lenses with composition. When composing
+lenses sequentially, the lenses on the sides could project or add information
+when synchronizing with the central type. While the composition may be a
+bijection (and should have zero cost), it can be difficult to see this looking
+merely at the syntax.
+
+However, there are many useful lenses involving composition. We synthesize
+lenses in a normal form as DNF lenses. DNF lenses do not include composition, so
+when we greedily search for a minimal lens, we calculate the cost through DNF
+lenses.
+
 %\begin{center}
@@
 \section{Synthesis}
+\label{sec:synthesis}
 Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm. The
@@
 of lenses by proposing regular expression pairs that characterize those classes.
-$\GreedySynth$ greedily finds a low cost lens within each proposed lens class,
-if there is one that satisfies the specification. These lenses are then provided
-to $\RXSearch$. Eventually, $\RXSearch$ is content with one of the lenses, and
-returns that lens to the user.
+These lens classes are proposed based on the distance from the original
+stochastic regular expressions: classes characterized by stochastic regular
+expressions closer to the originals are proposed earlier. $\GreedySynth$
+greedily finds a low cost lens within each proposed lens class, if there is one
+that satisfies the specification. These lenses are then provided to $\RXSearch$.
+Eventually, $\RXSearch$ terminates and returns the best lens found thus far.
 
@@
 
-\subsection{Stochastic Regular Expressions}
-\label{subsec:stoch-rx}
-Important to our algorithm is finding the expected number of bits needed to
-recover a string in one data source from a synchronized string in the other data
-source. To do this, we must first develop a probability model for our language.
-We do this with \emph{stochastic regular expressions}, regular expressions
-annotated with probability information. With this, we can jointly express a
-language and a probability distribution over that language.
-
-\begin{center}
-  \begin{tabular}{rcl}
-    \Regex{},\RegexAlt{}
-    & \GEq{} & $\String$ \\
-    & \GBar{} & $\emptyset$ \\
-    & \GBar{} & $\PRegexStar{\Regex}{p}$ \\
-    & \GBar{} & $\RegexConcat{\Regex_1}{\Regex_2}$ \\
-    & \GBar{} & $\PRegexOr{\Regex_1}{\Regex_2}{p}$
-  \end{tabular}
-\end{center}
-
-The language is defined like the regular expression without probability
-annotations would be.  The probabilities are defined as follows:
-\begin{center}
-  \begin{tabular}{rcl}
-    $\ProbabilityOf{\String}{\String''}$
-    & =
-    & $\begin{cases*}1 & if $\String = \String''$\\ 0 & otherwise\end{cases*}$ \\
-
-    $\ProbabilityOf{\emptyset}{\String}$
-    & =
-    & $0$ \\
-
-    $\ProbabilityOf{\RegexConcat{\Regex_1}{\Regex_2}}{s}$
-
-    & =
-    & $\Sigma_{\String = \String_1\String_2}\ProbabilityOf{\Regex_1}{\String_1}*\ProbabilityOf{\Regex_2}{\String_2}$ \\
-
-    $\ProbabilityOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability}}{\String}$
-    & =
-    & $\Probability * \ProbabilityOf{\Regex_1}{\String} +
-      (1-\Probability) * \ProbabilityOf{\Regex_2}{\String}$\\
-
-    $\ProbabilityOf{\PRegexStar{\Regex}{\Probability}}{\String}$
-    & =
-    & $\Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\Regex}{\String_i}$\\
-  \end{tabular}
-\end{center}
-In the stochastic regular expression
-$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation \Probability
-represents the likelihood the string is in $\Regex_1$. In the stochastic regular
-expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability
-how much more likely having a string involving more than $n$ iterations is than
-having a string involving $n$ iterations.
-
-$\RXSearch$ needs to be able to find equivalent stochastic regular expressions.
-However, as of yet, there is no syntactic means to reason about stochastic
-regular expression equivalence. We extend the \emph{star-semiring} equivalences
-to stochastic regular expressions.
-
-\begin{center}
-  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
-    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
-    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
-    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
-    \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
-    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{(1-\Probability_1)*\Probability_2})}{\Probability_1*\Probability_2} & \OrAssociativityRule{}  \\
-    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
-    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
-    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
-    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
-    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
-    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability} & \UnrollstarLeftRule{} \\
-    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{\Probability} & \UnrollstarRightRule{} 
-  \end{tabular}
-\end{center}
-
-\begin{theorem}
-  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
-  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
-\end{theorem}
-
-Finally, we wish to infer a reasonable probability distribution for stochastic
-regular expressions.  While these probability distributions could be either
-input by users or inferred from example data, we found that our inferred
-probability distributions work in practice. We defer our description of
-\ToStochastic until \S\ref{sec:dnf}, as it is easier to describe our desired
-probability distribution in the context of DNF regular expressions.
-% \caption{Regular Expression Equivalences}
-% \label{fig:regex-equivalence-rules}
-
 \subsection{\RXSearch}
@@
 
-\RXSearch terminates when the number of regular expressions of the smallest
-unproposed size exceeds the cost of the current best lens. At this point,
-\RXSearch stops proposing regular expression pairs, and instead returns to the
-user the best lens found thus far. Note this means that if \RXSearch finds a
-bijective lens (a lens of cost 0) it will immediately return.
+\RXSearch terminates when the sum of the number of regular expressions of the
+smallest unproposed size with the distance of those regular expressions from the
+originals exceeds the cost of the current best lens. At this point, \RXSearch
+stops proposing regular expression pairs, and instead returns to the user the
+best lens found thus far. Note this means that if \RXSearch finds a bijective
+lens (a lens of cost 0) it will immediately return.
 %
@@
 Stochastic DNF regular expressions are hierarchically built, where stochastic
-DNF regular expressions (\DNFRegex, \DNFRegexAlt) are lists of stochastic sequences. Stochastic sequences
+DNF regular expressions (\DNFRegex, \DNFRegexAlt) are lists of stochastic
+sequences. Stochastic sequences (\Sequence, \SequenceAlt)
 themselves are lists of interleaved strings and stochastic atoms.  Stochastic
-atoms are iterated stochastic DNF regular expressions.
-\sam{I think we should explicitly say what sets $A, B$, $SQ, TQ$ and $DS, DT$ range over.}
+atoms (\Atom,\AtomAlt) are iterated stochastic DNF regular expressions.
+
 \begin{center}
@@
 expressions whose DNF form gives equal probability to all sequences, and gives
-iterations a .8 chance to continue iterating.\sam{I think we should mention that
-.8 is chosen because it works well in practice.}
+iterations a .8 chance to continue iterating, as that gives a reasonable
+distribution of strings when generating them from the regular expressions.
 
 \paragraph*{Entropy}
-\sam{We should define what entropy means and why it is significant}
-In the situation where a DNF regular expression is
-unambiguous, we have defined an efficient, syntactic means to find the entropy
-of the DNF regular expression.
+We have developed a syntactic means of finding the entropy of a stochastic DNF
+regular expression, like we have for stochastic regular expressions.  This
+enables us to efficiently find the entropy without first converting to a
+stochastic regular expression.
+
 \begin{center}
@@
 
-As before, an atom lens \AtomLens{} is merely the iteration of a symmetric DNF
-Lens.
-
-\paragraph*{Symmetric DNF Lenses: Typing and Semantics}
-\sam{I found that I understood the semantics easier than the typing
-rules. Perhaps we should switch the order that we do them in?}
-The typing judgment is a 3-ary relation over a single DNF lens, and two DNF
-regular expressions. If $\SDNFLens \OfType \DNFRegex \Leftrightarrow
-\DNFRegexAlt$, then the $\SDNFLens.\CreateR$, $\SDNFLens.\CreateL$,
-$\SDNFLens.\PutR$, and $\SDNFLens.\PutL$ functions form a symmetric lens.
-
-The typing judgement has 3 components. The first is the sublens components,
-confirming that the sequence lenses the DNF lens is comprised of are all
-well-typed. The second guarantees that if each sequence on the left has a
-sequence lens that can be used for \CreateR{}s and \PutR{}s. The last guarantees
-the same for sequences on the right, with \CreateL{} and \PutL{}.
+An atom lens \AtomLens{} is merely the iteration of a symmetric DNF Lens.
+
+While DNF lenses are useful because they permit an efficient greedy synthesis
+procedure, the particulars of their typing and semantics are not the primary
+intellectual contributions to our work. The details of the typing and semantics
+are included in the supplementary material present in the appendix.
+
+\paragraph*{Information-Theoretic Cost Metric}
+Now that we've developed symmetric DNF lenses, it's important to develop a means
+to guage how bijective the lens is.
+
+Instead of searching for \emph{any} term that satisfies the specification, we
+search for the \emph{best} term that satisfies the specification. This begs the
+question: what metric should we use to determine which term is best?
+Intuitively, we aim to synthesize the lens which is ``closest'' to being a
+bijection. Somewhat more formally, we aim to synthesize a lens which minimizes
+the expected number of bits which must be used to recover the source from the
+target, and the target from the source.
+
+Much like we defined a syntactic means to find the expected number of bits to
+recover string from a synchronized string in another format for simple symmetric
+string lenses, we do the same for symmetric DNF lenses.  Unlike simple symmetric
+string lenses, DNF lenses do not have composition, so the entropy can be defined
+over all terms.  $\LEntropyOf{\DNFRegex \Given \SDNFLens, \DNFRegexAlt}$ is the
+expected number of bits required to recover a string in $\DNFRegex$ from a
+synchronized string in $\DNFRegexAlt$.
 
 \[
-  \inferrule*
-  {
-    \SSQLens_1 \OfType \Sequence_{i_1} \Leftrightarrow \SequenceAlt_{j_1}\\
-    \ldots\\
-    \SSQLens_p \OfType \Sequence_{i_p} \Leftrightarrow \SequenceAlt_{j_p}\\\\
-    i_{c_1} = 1\\
-    \ldots\\
-    i_{c_q} = q\\\\
-    j_{d_1} = 1\\
-    \ldots\\
-    j_{d_r} = r
-  }
-  {
+  \begin{array}{rl}
+    & \LEntropyOf{\DNFOf{\Sequence_1 \DNFSep \ldots \DNFSep \Sequence_n}\\
+    & \hspace*{1.21em}\Given
       (\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
@@
       ,\ListOf{c_1;\ldots;c_q}
-    ,\ListOf{d_1;\ldots;d_r})
-    \OfType\\\\
-    \DNFOf{(\Sequence_1,\Probability_1) \DNFSep \ldots \DNFSep (\Sequence_q,\Probability_q)}
-    \Leftrightarrow
-    \DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1) \DNFSep \ldots \DNFSep (\SequenceAlt_r,\ProbabilityAlt_q)}
-  }
+      ,\ListOf{d_1;\ldots;d_r})\\
+    & \hspace*{1.21em},~\DNFOf{\SequenceAlt_1 \DNFSep \ldots \DNFSep \SequenceAlt_m}}\\
+    = & \frac{\Sigma_{j=1}^m(H_j)}{m}\\
+    & \text{where } H_j = \frac{\Sigma_{k\SuchThat j_k = j}
+      \LEntropyOf{\Sequence_{i_k}\Given\SequenceLens_k,\SequenceAlt_j}}{\SizeOf{\SetOf{k\SuchThat
+      j_k = j}}} + log_2\SizeOf{\SetOf{k\SuchThat j_k = j}}
+  \end{array}
 \]
 
-The \CreateR{} function looks for the sequence the provided string matches. If
-the string matches sequence $\Sequence_x$, then the lens will look in the create
-list at position $x$ to find which sequence lens to use. Then, the sequence lens
-transforms the given string using that sequence lens.
-
-The \PutR{} function finds what pairs sequences the input source and view
-strings match.  If there that pair of sequences have a sequence lens between
-them, then the DNF lens merely performs that sequence lens on the provided
-strings.  If there isn't a pair of sequence lenses between them, then \CreateR{}
-is performed on the source, with the view forgotten.
-
-The \CreateL{} and \PutL{} functions are defined symmetrically.
-
-\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
-  $\CreateR{} \App s$ & = & $\SSQLens_{c_x}.\CreateR{} \App s$ if $s \in \Sequence_x$\\
-  $\CreateL{} \App v$ & = & $\SSQLens_{d_y}.\CreateL{} \App v$ if $v \in \SequenceAlt_y$\\
-  $\PutR{} \App s \App v$ & = &
-                               $\begin{cases*}
-                                 \SSQLens_x.\PutR{} \App s \App v & if $s \in \Sequence_{i_x}$ and $v \in \SequenceAlt_{j_x}$\\
-                                 \CreateR{} \App s & if $\nexists x.$ $s \in \Sequence_{i_x}$ and $v \in \SequenceAlt_{j_x}$
-                               \end{cases*}$\\
-  $\PutL{} \App v \App s$ & = &
-                               $\begin{cases*}
-                                 \SSQLens_y.\PutL{} \App v \App s & if $v \in \SequenceAlt_{j_y}$ and $s \in \Sequence_{i_y}$\\
-                                 \CreateL{} \App v & if $\nexists y.$ $v \in \SequenceAlt_{j_y}$ and $s \in \Sequence_{i_y}$
-                               \end{cases*}$
-\end{tabular}
-
-\paragraph*{Symmetric Sequence Lenses: Typing and Semantics}
-The typing judgment is a 3-ary relation over a single Sequence lens, and two
-sequences. If $\SSQLens \OfType \Sequence \Leftrightarrow \SequenceAlt$, then
-the $\SSQLens.\CreateR$, $\SSQLens.\CreateL$, $\SSQLens.\PutR$, and
-$\SSQLens.\PutL$ functions form a symmetric lens.
-
-The typing judgement has 4 components. The first is the sublens components,
-confirming that the atom lenses the sequence lens is comprised of all are
-well-typed. The second guarantees that if each string that will be used for
-\CreateR{}s are members of the correct atoms.  The third guarantees
-the same for strings and atoms on the right, with \CreateL.  The last guarantees
-that each atom is mapped by at most one atom lens.
-
-\[
-  \inferrule*
-  {
-    \SAtomLens_1 \OfType \Atom_{i_1} \Leftrightarrow \AtomAlt_{j_1}\\
-    \ldots\\
-    \SAtomLens_p \OfType \Atom_{i_p} \Leftrightarrow \AtomAlt_{j_p}\\\\
-    \String_1 \in \Atom_1\\
-    \ldots\\
-    \String_q \in \Atom_q\\\\
-    \StringAlt_1 \in \AtomAlt_1\\
-    \ldots\\
-    \StringAlt_r \in \AtomAlt_r\\\\
-    i_x = i_y \BooleanImplies x = y\\
-    j_x = j_y \BooleanImplies x = y
-  }
-  {
-    (\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
-      \ldots\SeqLSep
-      (i_p,j_p,\SAtomLens_p)}
-    ,\ListOf{\String_1;\ldots;\String_q}
-    ,\ListOf{\StringAlt_1;\ldots;\StringAlt_r})\\
-    \OfType
-    \SequenceOf{\String_0' \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_q
-      \SeqSep \String_q'}
-    \Leftrightarrow
-    \SequenceOf{\StringAlt_0' \SeqSep \AtomAlt_1 \SeqSep \ldots \SeqSep \AtomAlt_r \SeqSep \StringAlt_r'}
-  }
-\]
-
-For each component of the string matching an atom, the \CreateR function looks
-for the atom lens that maps on the atom. If there is such an atom lens,
-\SAtomLens, then that the sequence lens puts the provided string into the
-default string for the target atom. If there is no such atom lens, then the
-sequence lens merely uses the default string.
-
-For each component of the string matching an atom, the \PutR function looks
-for the atom lens that maps on the atom. If there is such an atom lens,
-\SAtomLens, then that the atom lens puts the provided string of the source atom into the
-string of the target atom. If there is no such atom lens, then the
-sequence lens merely recovers the target's string.
-
-The \CreateL{} and \PutL{} functions are defined symmetrically.
-
-\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
-  $\CreateR{} \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q'$
-  & = 
-  & $\StringAlt_0' \Concat \StringAlt_1'' \Concat \ldots \Concat
-    \StringAlt_r'' \Concat \StringAlt_r'$\\
-  & & where $\StringAlt_y'' =
-    \begin{cases*}
-      \SAtomLens_k.\PutROf{\String_{i_k}''}{\StringAlt_y} & if $j_k = y$\\
-      \StringAlt_y & if $\nexists k. j_k = y$\\
-    \end{cases*}$\\
-  $\CreateL{} \App \StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q''
-  \Concat \StringAlt_q'$
-  & = 
-  & $\String_0' \Concat \String_1'' \Concat \ldots \Concat
-    \String_r'' \Concat \String_r'$\\
-  & & where $\String_x'' =
-    \begin{cases*}
-      \SAtomLens_k.\PutROf{\StringAlt_{j_k}''}{\String_x} & if $i_k = x$\\
-      \String_x & if $\nexists k. i_k = x$\\
-    \end{cases*}$\\
-  $\PutR{} \App (\String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q') \App (\StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q'' \Concat \StringAlt_q')$
-  & =
-  & $\StringAlt_0'\Concat \StringAlt_1''' \Concat \ldots \Concat \StringAlt_q''' \Concat \StringAlt_q'$\\
-  & & where
-      $t_y''' =
-      \begin{cases*}
-        \SAtomLens_k.\PutROf{\String_{i_k}''}{\StringAlt_{y}''} & if $j_k = y$\\
-        \StringAlt_y'' & if $\nexists k. j_k = y$\\
-      \end{cases*}$\\
-  $\PutL{} \App (\StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q'') \Concat (\StringAlt_q' \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q')$
-  & =
-  & $\String_0'\Concat \String_1''' \Concat \ldots \Concat \String_r''' \Concat \String_r'$\\
-  & & where
-      $s_x''' =
-      \begin{cases*}
-        \SAtomLens_k.\PutLOf{\StringAlt_{j_k}''}{\String_{x}''} & if $i_k = x$\\
-        \String_x'' & if $\nexists k. i_k = x$\\
-      \end{cases*}$\\
-\end{tabular}
-
-\paragraph*{Symmetric Atom Lenses: Typing and Semantics}
-The typing judgment is a 3-ary relation over a single atom lens, and two
-atoms. If $\SAtomLens \OfType \Atom \Leftrightarrow \AtomAlt$, then
-the $\SAtomLens.\CreateR$, $\SAtomLens.\CreateL$, $\SAtomLens.\PutR$, and
-$\SAtomLens.\PutL$ functions form a symmetric lens.
-
-The typing judgement just confirms that the DNF lens that comprises the
-sequence lens is also well typed.
-
-\[
-  \inferrule*
-  {
-    \SDNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt
-  }
-  {
-    \StarOf{\SDNFLens}
-    \OfType \PRegexStar{\DNFRegex}{\Probability}
-    \Leftrightarrow \PRegexStar{\DNFRegexAlt}{\ProbabilityAlt}
-  }
-\]
-
-For each component of the string matching an atom, the \CreateR function looks
-for the atom lens that maps on the atom. If there is such an atom lens,
-\SAtomLens, then that the sequence lens puts the provided string into the
-default string for the target atom. If there is no such atom lens, then the
-sequence lens merely uses the default string.
-
-For each component of the string matching an atom, the \PutR function looks
-for the atom lens that maps on the atom. If there is such an atom lens,
-\SAtomLens, then that the atom lens puts the provided string of the source atom into the
-string of the target atom. If there is no such atom lens, then the
-sequence lens merely recovers the target's string.
-
-The \CreateL{} and \PutL{} functions are defined symmetrically.
-
-\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
-  $\CreateR{} \App \String_0 \Concat \ldots \Concat \String_n$
-  & = 
-  & $\StringAlt_1 \Concat \ldots \Concat \StringAlt_n$
-  where $\StringAlt_i = \SDNFLens.\CreateR \App \String_i$\\
-  $\CreateL{} \App \StringAlt_0 \Concat \ldots \Concat \StringAlt_m$
-  & = 
-  & $\String_1 \Concat \ldots \Concat \String_m$
-  where $\String_i = \SDNFLens.\CreateL \App \StringAlt_i$\\
-  $\PutR{} \App (\String_0 \Concat \ldots \Concat \String_n) \App
-  (\StringAlt_1 \Concat \ldots \Concat \StringAlt_m)$
-  & = 
-  & $\StringAlt_1' \Concat \ldots \Concat \StringAlt_n'$
-    where $\StringAlt_i' =
-    \begin{cases*}
-      \SDNFLens.\PutR \App \String_i \App \StringAlt_i & if $i \leq m$\\
-      \SDNFLens.\CreateR \App \String_i & otherwise
-    \end{cases*}$\\
-  $\PutL{} \App (\StringAlt_1 \Concat \ldots \Concat \StringAlt_m) \App
-  (\String_0 \Concat \ldots \Concat \String_n)$
-  & = 
-  & $\String_1' \Concat \ldots \Concat \String_m'$
-    where $\String_i' =
-    \begin{cases*}
-      \SDNFLens.\PutL \App \StringAlt_i \App \String_i & if $i \leq n$\\
-      \SDNFLens.\CreateL \App \StringAlt_i & otherwise
-    \end{cases*}$\\
-\end{tabular}
-
-\paragraph*{Information-Theoretic Cost Metric}
-Now that we've developed symmetric DNF lenses, it's important to develop a means
-to guage how bijective the lens is.
-
-Instead of searching for \emph{any} term that satisfies the specification, we
-search for the \emph{best} term that satisfies the specification. This begs the
-question: what metric should we use to determine which term is best? Intuitively,
-we aim to synthesize the lens which is \emph{closest} to being a bijection.
-Somewhat more formally, we aim to synthesize a lens which minimizes the expected
-number of bits which must be used to recover the source from the target, and the
-target from the source. The expected number of bits that are needed to encode
-informatation from a data source is a well known concept in information theory,
-and is known as the \emph{entropy} of the data source.
-
-The entropy of $\DNFRegexAlt$ given $\DNFRegex$ and $\SDNFLens$,
-$\EntropyOf{\DNFRegexAlt \Given \DNFRegex,\SDNFLens}$ is defined symmetrically.
-With these $\operatorname*{argmin}_\theta$ terms defined, the ``best'' DNF lens
-between $\DNFRegex$ and $\DNFRegexAlt$ is $\ArgminOver{\SetOf{\SDNFLens \SuchThat
-    \SDNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt}}
-\EntropyOf{\DNFRegex \Given \DNFRegexAlt,\SDNFLens} + \EntropyOf{\DNFRegexAlt
-  \Given \DNFRegex,\SDNFLens}$.
-\sam{We should definitely explain the formulae for entropy computing entropy. Also I did not understand $argmin_{\theta}$ terms were being talked about in the 
-previous sentence.}
-\begin{centering}
+This entropy calculation for symmetric DNF lenses requires a similar entropy
+calculation for symmetric sequence lenses and symmetric atom lenses.
+
 \[
   \begin{array}{rl}
-      & \EntropyOf{\DNFOf{\Sequence_1 \DNFSep \ldots \DNFSep \Sequence_n}\\
-      & \hspace*{1em}\Given
-        \DNFOf{\SequenceAlt_1 \DNFSep \ldots \DNFSep \SequenceAlt_m}\\
-      & \hspace*{1.21em},~(\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
-        \ldots\DNFLSep
-        (i_p,j_p,\SSQLens_p)}
-        ,\ListOf{c_1;\ldots;c_q}
-        ,\ListOf{d_1;\ldots;d_r})}\\
-      =\\
-      & \frac{\Sigma_{j=1}^m(\frac{\Sigma_{k\SuchThat j_k = j}
-        \EntropyOf{\Sequence_{i_k}\Given\SequenceAlt_j,\SequenceLens_k}}{\SizeOf{\SetOf{k\SuchThat
-        j_k = j}}} + log_2\SizeOf{\SetOf{k\SuchThat j_k = j}})}{m}\\\\
-      \\
-
-      & \EntropyOf{
+      & \LEntropyOf{
         \SequenceOf{\String_0,\Atom_1,\ldots,\Atom_n,\String_n}\\
       & \hspace*{1em}\Given
-        \SequenceOf{\StringAlt_0,\AtomAlt_1,\ldots,\AtomAlt_m,\StringAlt_m}\\
-      & \hspace*{1.21em},~
         (\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
@@
         ,\ListOf{(k_1,\String_1);\ldots;(k_q,\String_q)}
-        ,\ListOf{(l_1,\StringAlt_1);\ldots;(l_r,\StringAlt_r)})}\\
-      =\\
-      & \Sigma_{x=1}^p\EntropyOf{\Atom_{i_x} \Given \AtomAlt_{i_x},\AtomLens_x} +
-        \Sigma_{x=1}^q\EntropyOf{\Atom_{k_x}}\\\\
-      \\
-
-      & \EntropyOf{\StarOf{\DNFRegex} \Given \StarOf{\DNFRegexAlt},\StarOf{\SDNFLens}}\\
-      =\\
-      & 4*\EntropyOf{\DNFRegex\Given\DNFRegexAlt,\SDNFLens}
+        ,\ListOf{(l_1,\StringAlt_1);\ldots;(l_r,\StringAlt_r)})\\
+      & \hspace*{1.21em},~
+        \SequenceOf{\StringAlt_0,\AtomAlt_1,\ldots,\AtomAlt_m,\StringAlt_m}}\\
+      =
+      & \Sigma_{x=1}^p\LEntropyOf{\Atom_{i_x} \Given \AtomAlt_{i_x},\AtomLens_x} +
+        \Sigma_{x=1}^q\LEntropyOf{\Atom_{k_x}}
   \end{array}
 \]
-\end{centering}
+
+The entropy calculation for symmetric sequence lenses requires a similar entropy
+calculation for symmetric atom lenses, which in turn relies on the entropy
+calculation for symmetric DNF lenses.
+
+\[\begin{array}{rl}
+      & \LEntropyOf{\PRegexStar{\DNFRegex}{\Probability} \Given \StarOf{\DNFRegexAlt},\StarOf{\SDNFLens}}\\
+      =\\
+      & 4*\LEntropyOf{\DNFRegex\Given\DNFRegexAlt,\SDNFLens}
+  \end{array}
+\]
+The expected number of bits required to recover a string in $\DNFRegexAlt$ from
+a synchronized string in $\DNFRegex$, $\REntropyOf{\DNFRegexAlt \Given
+  \SDNFLens, \DNFRegex}$ is defined symmetrically.
 
@@
 sequence lenses (\GreedySeqSynth).
-\sam{Could it be better to describe the algorithm in English first before giving
-pseudo code? I think that putting the algorithm first might cause readers to read the 
-pseudo code and get confused whereas if the algorithm is first described in English then they have a better chance of understanding it.}
+
+\paragraph*{Symmetric DNF Lenses} Algorithm~\ref{alg:greedysynth} presents
+\GreedySynth, a means to synthesize symmetric DNF lenses. The inputs to
+\GreedySynth are a pair of stochastic DNF regular expressions and a suite of
+input-output examples. First, \PCF{CannotMap} checks if there could possibly be
+a lens satisfying the examples. If there is not, \GreedySynth returns \None
+early. After this, \GreedySynth finds the best lenses between each sequence pair
+of the left and right DNF regular expressions, given the examples that match
+them. An internal state for the algorithm is initialized from these. This state
+consists of two things: an intermediate partial DNF Lens and a priority queue
+consisting of potential sequence lenses. On initialization, \GreedySynth checks
+which sequence lenses are \emph{required} to be included, from the examples.
+Those are added into the partial lens, and the rest are thrown into the priority
+queue. For example, an example if an input output example is provided, the
+sequences that match those provided strings must be mapped to each other with a
+sequence lens. \GreedySynth then loops until there are no more sequence lenses
+in the priority queue. Within this loop, sequence lenses are popped. If a popped
+sequence lens maps between two sequences which aren't yet included in a sequence
+lens, then this sequence lens is added to the partial DNF lens. Furthermore, we
+update the priorities of our sequence lenses as the algorithm proceeds: if two
+sequence lenses have the same source, the second one to be popped gets a higher
+cost than it originally had, to address the information that needs to be stored
+for including that source of non-bijectivity. Finally, after all sequences have
+been popped, the partial DNF lens is converted into a symmetric DNF lens. This
+is only possible if all sequences are involved in some sequence lens: if they
+are not, \PCF{GreedyState.ToDNFLens} instead returns \None.
+
 \begin{algorithm}
@@
 
-\paragraph*{Symmetric DNF Lenses} Algorithm~\ref{alg:greedysynth} presents
-\GreedySynth, a means to synthesize symmetric DNF lenses. The inputs to
-\GreedySynth are a pair of stochastic DNF regular expressions and a suite of
-input-output examples. First, \PCF{CannotMap} checks if there is possibly a lens
-between the given DNF regular expressions that is consistent with the examples.
-\sam{I think we should reword this to say something like CANNOTMAP checks to see if the specification cannot be satisfied.
-To me it sounds like checking if a lens possibly exists is already going into synthesis.}If there is not, \GreedySynth returns \None early.
-After this, \GreedySynth
-finds the best lenses between each sequence pair of the left and right DNF
-regular expressions. An internal state for the algorithm is initialized from
-these. This state consists of two things: an intermediate partial DNF Lens and a
-priority queue consisting of potential sequence lenses. On initialization,
-\GreedySynth checks which sequence lenses are \emph{required} to be included,
-from the examples. Those are added into the partial lens, and the rest are
-thrown into the priority queue.
-\sam{Perhaps a small example on initialization would help? I was wondering what
-the partial DNF lens and potential sequence lenses could be initialized as}
-\GreedySynth then loops until there are no more sequence lenses in the priority
-queue. Within this loop, sequence lenses are popped. If a popped sequence lens
-maps between two sequences which aren't yet included in a sequence lens, then
-this sequence lens is added to the partial DNF lens. Furthermore, we update the
-priorities of our sequence lenses as the algorithm proceeds: if two sequence
-lenses have the same source, the second one to be popped gets a higher cost than
-it originally had, to address the information that needs to be stored for
-including that source of non-bijectivity. Finally, after all sequences have been
-popped, the partial DNF lens is converted into a symmetric DNF lens. This
-is only possible if all sequences are involved in some sequence lens: if they
-are not, \PCF{GreedyState.ToDNFLens} instead returns \None.
+\paragraph*{Symmetric Sequence lenses} Algorithm~\ref{alg:greedyseqsynth}
+presents \GreedySeqSynth, a means to synthesizes symmetric sequence lenses. This
+algorithm uses \PCF{AtomSynth}, a simple function that synthesizes atom lenses
+by iterating the DNF lens between its subcomponents. The inputs to
+\GreedySeqSynth are a pair of stochastic atoms and a suite of input-output
+examples. First, \PCF{CannotMap} checks if there is a possibility of a lens
+between the given sequences regular expressions that is consistent with the
+examples. If there is not, \GreedySeqSynth returns \None early. After this,
+\GreedySeqSynth finds the best lenses between each atom pair of the left and
+right sequences. An internal state for the algorithm is initialized from these.
+This state consists of two things: an intermediate sequence lens and a priority
+queue consisting of potential sequence lenses. On initialization,
+\GreedySeqSynth throws all the atom lenses into the priority queue, and the
+intermediate sequence lens is one that projects all the atoms.
+
+\GreedySeqSynth then loops until there are no more atom lenses in the priority
+queue. Within this loop, atom lenses are popped. The atom lens is then added if
+it is either necessary, or useful. Within a sequence lens, each atom can either
+be projected or be synchronized with exactly one atom in the other sequence, and
+examples can show that certain atoms \emph{need} to be mapped. In such
+situations, that atom lens is necessary. A useful atom lens is where adding that
+lens reduces the overall cost of the partial lens. If both atoms are not
+involved in an atom lens, the lens will always be useful, but sometimes adding
+that lens requires removing another lens. If the cost added by removing that
+other lens is more than is removed by adding the proposed lens, then this atom
+lens is not included in the sequence lens. Finally, after all atoms have been
+popped, the sequence lens is validated and returned. The sequence lens is
+validated if all atoms that need to be mapped, are: if they are not,
+\PCF{GreedySeqState.ToDNFLens} instead returns \None.
 
@@
 
-\paragraph*{Symmetric Sequence lenses} Algorithm~\ref{alg:greedyseqsynth}
-presents \GreedySeqSynth, a means to synthesizes symmetric sequence lenses. This
-algorithm uses \PCF{AtomSynth}, a simple function that synthesizes atom lenses
-by iterating the DNF lens between its subcomponents. The inputs to
-\GreedySeqSynth are a pair of stochastic atoms and a suite of input-output
-examples. First, \PCF{CannotMap} checks if there is a possibility of a lens
-between the given sequences regular expressions that is consistent with the
-examples. If there is not, \GreedySeqSynth returns \None early. After this,
-\GreedySeqSynth finds the best lenses between each atom pair of the left and
-right sequences. An internal state for the algorithm is initialized from these.
-This state consists of two things: an intermediate sequence lens and a priority
-queue consisting of potential sequence lenses. On initialization,
-\GreedySeqSynth throws all the atom lenses into the priority queue, and the
-intermediate sequence lens is one that projects all the atoms.
-
-\GreedySeqSynth then loops until there are no more atom lenses in the priority
-queue. Within this loop, atom lenses are popped. The atom lens is then added if
-it is either necessary, or useful. Within a sequence lens, each atom can either
-be projected or be synchronized with exactly one atom in the other sequence, and
-examples can show that certain atoms \emph{need} to be mapped. In such
-situations, that atom lens is necessary. A useful atom lens is where adding that
-lens reduces the overall cost of the partial lens. If both atoms are not
-involved in an atom lens, the lens will always be useful, but sometimes adding
-that lens requires removing another lens. If the cost added by removing that
-other lens is more than is removed by adding the proposed lens, then this atom
-lens is not included in the sequence lens. Finally, after all atoms have been
-popped, the sequence lens is validated and returned. The sequence lens is
-validated if all atoms that need to be mapped, are: if they are not,
-\PCF{GreedySeqState.ToDNFLens} instead returns \None.
-
 \subsection{Optimizations}
@@
 ``closed'' annotations to regular expressions, an expansion inference algorithm,
-hash-consing and memoization, and compositional synthesis to our algorithm.
+hash-consing and memoization, and compositional synthesis to our algorithm. We
+also enable an additional parameter in our specifications that customizes how
+permissive \RXSearch{} is in accepting proposed lenses.
 
@@
 we enhance our rewrites by changing an annotation from closed to open. This
-forces changes to \RXSearch and \GreedySynth.  \RXSearch is changed in that now
-it can perform new rewrites.  \GreedySynth is changed in that now there is an
-additional type of atom that can be mapped between.  Two regular expressions can
-map to each other only through the identity lens: two regular expressions can
-map to each other by the identity lens, if they are the same and all the
-examples say they are the same, otherwise each regular expression is left
-disconnected.
-\sam{I did not understand this section. In particluar, I'm still not sure I
-understand a closed regexp vs an open regexp. Perhaps an example would help?}
+forces changes to \RXSearch and \GreedySynth. \RXSearch now can rewrite closed
+regular expressions to open ones. \GreedySynth is provided with the means to map
+closed regular expressions to each other. Two closed regular expressions can map
+to each other only through the identity lens: two regular expressions can map to
+each other by the identity lens, if they are the same closed regular expression
+and they parse the same string components in all examples.
+
 \paragraph*{Expansion Inference} The prior optimization helped \GreedySynth, but
@@
 examples in a desired way.
+
+\paragraph*{\RXSearch Permissiveness Customization}
+While termination heuristic works for many situations: there are situations
+where a large number of expansions must be performed to find the correct lens.
+While users could prevent incorrect lenses with a large number of examples, we
+find it simpler to let the users specify that \RXSearch should be more
+restrictive in which lenses it accepts. If the user's desired transformation has
+a particularly high cost, the algorithm may spend a long time searching in vain
+for lenses with lower cost.  By tuning \RXSearch to me more permissive in the
+lenses it accepts, this search can be short-circuited.
 % end synthesis
-
-% begin implementation
-\section{Implementation}
-\label{sec:implementation}
-% end implementation
 
@@
 In this work, we aim to answer two primary research questions:
-\begin{itemize}
+\begin{enumerate}
 \item Is our synthesis procedure efficient enough to be used in everyday
@@
   algorithm integrated with Boomerang?
-\item Is our information theoretic cost metric useful in finding good lenses.
-\end{itemize}
+\item Are our heuristics useful for reducing the specifications users have to
+  provide.
+\end{enumerate}
 
@@
 the final formats. It is therefore interesting to see how many problems each
-algorithm can solve, when trying to synthesize ``in one go''.
+algorithm can solve, when trying to synthesize ``in one go.''
 
@@
 The slowest completed benchmark for both the bijective and symmetric synthesis
-algorithms is \lstinline{xml_to_augeas.boom}, a benchmark that converts
+algorithms is \texttt{xml\_to\_augeas.boom}, a benchmark that converts
 arbitrary XML up to depth 3 into a serialized version of the structured
@@
 synthesis algorithm to complete.
+
+\subsection{Heuristic Usefulness}
+We evaluate the usefulness of (1) our information theoretic metric, and (2) our
+termination heuristic.  To this end, we run our program in 3 different modes:\\
+\begin{tabulary}{\linewidth}{rL}
+  \NoTPOpt{}: & We synthesize lenses using our symmetric synthesis algorithm,
+                but users cannot tune when \RXSearch returns using a termination parameter.\\
+  \FCOpt{}: &  The first time \GreedySynth finds a lens, it returns that lens.\\
+  \NMOpt{}: & \GreedySynth does not have any notion of cost, and searches for
+              any lens that satisfies the specification. \\
+\end{tabulary}
+
+Figure~\ref{fig:heuristic} summarizes the result of this experiment.  We find
+that \afm{TODO}.
 % end evaluation
@@
 There is related work in a number of directions. There is related work in
-symmetric lenses, related work on synthesizing synchronization functions,
-related work on string transformation synthesis, and related work on
-type-directed synthesis.
+symmetric lenses, string transformation synthesis, and type-directed synthesis.
+We also refer the reader to a recent, general survey on synthesis~\ref{?}.
 
@@
   \tag{\PutRL}
-  \frac{\PutRSymOf{(x,c)} = (y,c'')}{\PutLSymOf{(y,c')} = (x,c')}
+  \frac{\PutRSymOf{(x,c)} = (y,c'')}{\PutLSymOf{(y,c'')} = (x,c')}
 \end{equation}
@@
 
-\subsection{Synthesizing Synchronization Functions}
-
 \subsection{String Transformation Synthesis}
+Much of the research on string transformation synthesis is on synthesizing
+string transformations from input/output
+specifications~\cite{flashfill,flashextract,?}. While this provides an easier
+interaction model with the user than this work, it requires the algorithm to
+learn the data formats from the examples. This makes the task much more
+difficult for the synthesis algorithm, particularly in addressing iterations and
+disjunctions. Prior work has shown that learning from examples and data format
+descriptions enables synthesis of much more complex transformations than are
+permitted by examples alone. Recent work has gone into making these algorithms
+more robust by delaying the selection of program until seeing what further
+inputs the program is run on~\cite{?}. By doing this, the synthesis algorithm
+can learn additional information about the formats from the use of the program.
+
+In many of these synthesis algorithms the correct transformation is chosen by
+principals like ``minimize constants''~\cite{?,?,?} and ``most specific
+generalization''~\cite{?,?,?}. We are the first work to our knowledge to use an
+information-theoretic measure for finding the correct transformation. However,
+much of our metrics are compatible with existing metrics: avoiding \Disconnect
+lenses minimizes constants and creates more specfic transformations (\Disconnect
+lenses permit more synchronized data).
+
+Optician~\ref{?} synchronizes string transformation functions, but prior
+versions were very restrictive in which transformations were permitted.
+Originally, Optician only synchronized transformations between formats in
+bijective correspondence, but few formats satisfy that constraint. Recent
+enhancements have permitted synthesis of more expressive lenses, quotient
+bijective lenses. These enhancements were made by strengthening the
+specifications from regular expressions to QREs, regular expressions annotated
+with information on normal forms of the language. This enhancement permits
+synthesizing transformations between formats whose normal forms are in bijective
+correspondence. While we have not integrated our system to permit QRE
+specifications, we see no reason the two systems cannot be integrated.
+
+Other work synthesizing invertible transformations was done by inverting a
+previously defined transformation.  This has been done by exploiting
+parametricity to invert transformations~\cite{?}, and by inverting symbolic
+finite automata~\cite{?}.
 
 \subsection{Type-Directed Synthesis}
+
+Type-directed synthesis uses type information to tighten specifications and
+reduce the number of possible programs. Typically, type-directed synthesis
+techniques use the types to orient the search space.  Our work interestingly
+deals with types with many semantic equivalences on them. InSynth~\cite{?}
+deals with the same problem when trying to synthesize transformations involving
+library functions. InSynth addresses this in a similar manner to this work: by
+normalizing the types we don't need to search through the equivalences on those
+types.
 
@@
 \label{sec:conc}
+We have developed a synthesis algorithm for synthesizing synchronization
+functions between data formats that may not be in bijective correspondence. We
+do this by synthesizing symmetric lenses. However, existing symmetric lens
+formulations use complex complements to store external state. We identify a
+restriction on symmetric lenses, simple symmetric lenses. We develop combinators
+for simple symmetric lenses on string data. We show how to synthesize these
+lenses from regular expression specifications. Prior work could return
+\emph{any} satisfying lens, as the type system was restrictive enough that any
+satisfying lens would typically be the correct satisfying lens. Because of the
+increased flexibility of the symmetric lens language, we identify an
+information-theoretic heuristic for finding a likely satisfying lens. We
+developed a greedy synthesis algorithm that aims to find the lens of lowest
+cost. We demonstrate empiraclly that our algorithm can find lenses quickly, and
+our heuristic makes finding the correct lens require fewer specifications.
+
 % end conclusion
@@
 \onecolumn
+
+\paragraph*{Symmetric DNF Lenses: Typing and Semantics}
+The typing judgment is a 3-ary relation over a single DNF lens, and two DNF
+regular expressions. If $\SDNFLens \OfType \DNFRegex \Leftrightarrow
+\DNFRegexAlt$, then the $\SDNFLens.\CreateR$, $\SDNFLens.\CreateL$,
+$\SDNFLens.\PutR$, and $\SDNFLens.\PutL$ functions form a symmetric lens.
+
+The typing judgement has 3 components. The first is the sublens components,
+confirming that the sequence lenses the DNF lens is comprised of are all
+well-typed. The second guarantees that if each sequence on the left has a
+sequence lens that can be used for \CreateR{}s and \PutR{}s. The last guarantees
+the same for sequences on the right, with \CreateL{} and \PutL{}.
+
+\[
+  \inferrule*
+  {
+    \SSQLens_1 \OfType \Sequence_{i_1} \Leftrightarrow \SequenceAlt_{j_1}\\
+    \ldots\\
+    \SSQLens_p \OfType \Sequence_{i_p} \Leftrightarrow \SequenceAlt_{j_p}\\\\
+    i_{c_1} = 1\\
+    \ldots\\
+    i_{c_q} = q\\\\
+    j_{d_1} = 1\\
+    \ldots\\
+    j_{d_r} = r
+  }
+  {
+    (\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
+      \ldots\DNFLSep
+      (i_p,j_p,\SSQLens_p)}
+    ,\ListOf{c_1;\ldots;c_q}
+    ,\ListOf{d_1;\ldots;d_r})
+    \OfType\\\\
+    \DNFOf{(\Sequence_1,\Probability_1) \DNFSep \ldots \DNFSep (\Sequence_q,\Probability_q)}
+    \Leftrightarrow
+    \DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1) \DNFSep \ldots \DNFSep (\SequenceAlt_r,\ProbabilityAlt_q)}
+  }
+\]
+
+The \CreateR{} function looks for the sequence the provided string matches. If
+the string matches sequence $\Sequence_x$, then the lens will look in the create
+list at position $x$ to find which sequence lens to use. Then, the sequence lens
+transforms the given string using that sequence lens.
+
+The \PutR{} function finds what pairs sequences the input source and view
+strings match.  If there that pair of sequences have a sequence lens between
+them, then the DNF lens merely performs that sequence lens on the provided
+strings.  If there isn't a pair of sequence lenses between them, then \CreateR{}
+is performed on the source, with the view forgotten.
+
+The \CreateL{} and \PutL{} functions are defined symmetrically.
+
+\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+  $\CreateR{} \App s$ & = & $\SSQLens_{c_x}.\CreateR{} \App s$ if $s \in \Sequence_x$\\
+  $\CreateL{} \App v$ & = & $\SSQLens_{d_y}.\CreateL{} \App v$ if $v \in \SequenceAlt_y$\\
+  $\PutR{} \App s \App v$ & = &
+                               $\begin{cases*}
+                                 \SSQLens_x.\PutR{} \App s \App v & if $s \in \Sequence_{i_x}$ and $v \in \SequenceAlt_{j_x}$\\
+                                 \CreateR{} \App s & if $\nexists x.$ $s \in \Sequence_{i_x}$ and $v \in \SequenceAlt_{j_x}$
+                               \end{cases*}$\\
+  $\PutL{} \App v \App s$ & = &
+                               $\begin{cases*}
+                                 \SSQLens_y.\PutL{} \App v \App s & if $v \in \SequenceAlt_{j_y}$ and $s \in \Sequence_{i_y}$\\
+                                 \CreateL{} \App v & if $\nexists y.$ $v \in \SequenceAlt_{j_y}$ and $s \in \Sequence_{i_y}$
+                               \end{cases*}$
+\end{tabular}
+
+\paragraph*{Symmetric Sequence Lenses: Typing and Semantics}
+The typing judgment is a 3-ary relation over a single Sequence lens, and two
+sequences. If $\SSQLens \OfType \Sequence \Leftrightarrow \SequenceAlt$, then
+the $\SSQLens.\CreateR$, $\SSQLens.\CreateL$, $\SSQLens.\PutR$, and
+$\SSQLens.\PutL$ functions form a symmetric lens.
+
+The typing judgement has 4 components. The first is the sublens components,
+confirming that the atom lenses the sequence lens is comprised of all are
+well-typed. The second guarantees that if each string that will be used for
+\CreateR{}s are members of the correct atoms.  The third guarantees
+the same for strings and atoms on the right, with \CreateL.  The last guarantees
+that each atom is mapped by at most one atom lens.
+
+\[
+  \inferrule*
+  {
+    \SAtomLens_1 \OfType \Atom_{i_1} \Leftrightarrow \AtomAlt_{j_1}\\
+    \ldots\\
+    \SAtomLens_p \OfType \Atom_{i_p} \Leftrightarrow \AtomAlt_{j_p}\\\\
+    \String_1 \in \Atom_1\\
+    \ldots\\
+    \String_q \in \Atom_q\\\\
+    \StringAlt_1 \in \AtomAlt_1\\
+    \ldots\\
+    \StringAlt_r \in \AtomAlt_r\\\\
+    i_x = i_y \BooleanImplies x = y\\
+    j_x = j_y \BooleanImplies x = y
+  }
+  {
+    (\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
+      \ldots\SeqLSep
+      (i_p,j_p,\SAtomLens_p)}
+    ,\ListOf{\String_1;\ldots;\String_q}
+    ,\ListOf{\StringAlt_1;\ldots;\StringAlt_r})\\
+    \OfType
+    \SequenceOf{\String_0'''' \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_q
+      \SeqSep \String_q'}
+    \Leftrightarrow
+    \SequenceOf{\StringAlt_0' \SeqSep \AtomAlt_1 \SeqSep \ldots \SeqSep \AtomAlt_r \SeqSep \StringAlt_r'}
+  }
+\]
+
+For each component of the string matching an atom, the \CreateR function looks
+for the atom lens that maps on the atom. If there is such an atom lens,
+\SAtomLens, then that the sequence lens puts the provided string into the
+default string for the target atom. If there is no such atom lens, then the
+sequence lens merely uses the default string.
+
+For each component of the string matching an atom, the \PutR function looks
+for the atom lens that maps on the atom. If there is such an atom lens,
+\SAtomLens, then that the atom lens puts the provided string of the source atom into the
+string of the target atom. If there is no such atom lens, then the
+sequence lens merely recovers the target's string.
+
+The \CreateL{} and \PutL{} functions are defined symmetrically.
+
+\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+  $\CreateR{} \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q'$
+  & = 
+  & $\StringAlt_0' \Concat \StringAlt_1'' \Concat \ldots \Concat
+    \StringAlt_r'' \Concat \StringAlt_r'$\\
+  & & where $\StringAlt_y'' =
+    \begin{cases*}
+      \SAtomLens_k.\PutROf{\String_{i_k}''}{\StringAlt_y} & if $j_k = y$\\
+      \StringAlt_y & if $\nexists k. j_k = y$\\
+    \end{cases*}$\\
+  $\CreateL{} \App \StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q''
+  \Concat \StringAlt_q'$
+  & = 
+  & $\String_0' \Concat \String_1'' \Concat \ldots \Concat
+    \String_r'' \Concat \String_r'$\\
+  & & where $\String_x'' =
+    \begin{cases*}
+      \SAtomLens_k.\PutROf{\StringAlt_{j_k}''}{\String_x} & if $i_k = x$\\
+      \String_x & if $\nexists k. i_k = x$\\
+    \end{cases*}$\\
+  $\PutR{} \App (\String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q') \App (\StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q'' \Concat \StringAlt_q')$
+  & =
+  & $\StringAlt_0'\Concat \StringAlt_1''' \Concat \ldots \Concat \StringAlt_q''' \Concat \StringAlt_q'$\\
+  & & where
+      $t_y''' =
+      \begin{cases*}
+        \SAtomLens_k.\PutROf{\String_{i_k}''}{\StringAlt_{y}''} & if $j_k = y$\\
+        \StringAlt_y'' & if $\nexists k. j_k = y$\\
+      \end{cases*}$\\
+  $\PutL{} \App (\StringAlt_0'\Concat \StringAlt_1'' \Concat \ldots \Concat \StringAlt_q'') \Concat (\StringAlt_q' \App \String_0'\Concat \String_1'' \Concat \ldots \Concat \String_q'' \Concat \String_q')$
+  & =
+  & $\String_0'\Concat \String_1''' \Concat \ldots \Concat \String_r''' \Concat \String_r'$\\
+  & & where
+      $s_x''' =
+      \begin{cases*}
+        \SAtomLens_k.\PutLOf{\StringAlt_{j_k}''}{\String_{x}''} & if $i_k = x$\\
+        \String_x'' & if $\nexists k. i_k = x$\\
+      \end{cases*}$\\
+\end{tabular}
+
+\paragraph*{Symmetric Atom Lenses: Typing and Semantics}
+The typing judgment is a 3-ary relation over a single atom lens, and two
+atoms. If $\SAtomLens \OfType \Atom \Leftrightarrow \AtomAlt$, then
+the $\SAtomLens.\CreateR$, $\SAtomLens.\CreateL$, $\SAtomLens.\PutR$, and
+$\SAtomLens.\PutL$ functions form a symmetric lens.
+
+The typing judgement just confirms that the DNF lens that comprises the
+sequence lens is also well typed.
+
+\[
+  \inferrule*
+  {
+    \SDNFLens \OfType \DNFRegex \Leftrightarrow \DNFRegexAlt
+  }
+  {
+    \StarOf{\SDNFLens}
+    \OfType \PRegexStar{\DNFRegex}{\Probability}
+    \Leftrightarrow \PRegexStar{\DNFRegexAlt}{\ProbabilityAlt}
+  }
+\]
+
+For each component of the string matching an atom, the \CreateR function looks
+for the atom lens that maps on the atom. If there is such an atom lens,
+\SAtomLens, then that the sequence lens puts the provided string into the
+default string for the target atom. If there is no such atom lens, then the
+sequence lens merely uses the default string.
+
+For each component of the string matching an atom, the \PutR function looks
+for the atom lens that maps on the atom. If there is such an atom lens,
+\SAtomLens, then that the atom lens puts the provided string of the source atom into the
+string of the target atom. If there is no such atom lens, then the
+sequence lens merely recovers the target's string.
+
+The \CreateL{} and \PutL{} functions are defined symmetrically.
+
+\begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+  $\CreateR{} \App \String_0 \Concat \ldots \Concat \String_n$
+  & = 
+  & $\StringAlt_1 \Concat \ldots \Concat \StringAlt_n$
+  where $\StringAlt_i = \SDNFLens.\CreateR \App \String_i$\\
+  $\CreateL{} \App \StringAlt_0 \Concat \ldots \Concat \StringAlt_m$
+  & = 
+  & $\String_1 \Concat \ldots \Concat \String_m$
+  where $\String_i = \SDNFLens.\CreateL \App \StringAlt_i$\\
+  $\PutR{} \App (\String_0 \Concat \ldots \Concat \String_n) \App
+  (\StringAlt_1 \Concat \ldots \Concat \StringAlt_m)$
+  & = 
+  & $\StringAlt_1' \Concat \ldots \Concat \StringAlt_n'$
+    where $\StringAlt_i' =
+    \begin{cases*}
+      \SDNFLens.\PutR \App \String_i \App \StringAlt_i & if $i \leq m$\\
+      \SDNFLens.\CreateR \App \String_i & otherwise
+    \end{cases*}$\\
+  $\PutL{} \App (\StringAlt_1 \Concat \ldots \Concat \StringAlt_m) \App
+  (\String_0 \Concat \ldots \Concat \String_n)$
+  & = 
+  & $\String_1' \Concat \ldots \Concat \String_m'$
+    where $\String_i' =
+    \begin{cases*}
+      \SDNFLens.\PutL \App \StringAlt_i \App \String_i & if $i \leq n$\\
+      \SDNFLens.\CreateL \App \StringAlt_i & otherwise
+    \end{cases*}$\\
+\end{tabular}
 
@@
 
+% old intro
+%\bcp{Not very compelling:}Similar data can come in varieties of formats: scheduled jobs can be stored as
+%cron jobs for Linux or as Scheduled Tasks for Windows; web service
+%resources can be provided to clients in legacy or modern formats;
+%calendar events can be stored in CSV files or ICS files. Oftentimes,
+%files should be synchronized across these formats: when I edit a task
+%to run more often on my Linux machine, I want this task to run more
+%often on my Windows machine; when I update data through a legacy API,
+%I want those changes to be reflected when I retrieve it through a
+%modern API; when I create a new calendar event in my ICS file, I want
+%my CSV calendar files to be updated to reflect this change.
+%
+%Unfortunately, writing such synchronizers is quite difficult; we must write
+%parsers both formats, specify how the parsed data is converted into the other
+%format, and specify how to propagate updates from one format to the other.
+%Bidirectional languages ease the difficulty of writing such synchronizers,
+%because a single, well-typed program expresses a \emph{lens}---a group of
+%conversion functions guaranteed to satisfy round-tripping laws. However,
+%bidirectional languages are themselves difficult to program in, often requiring
+%users to permute data in fiddly ways while satisfying the type system's complex
+%unambiguity constraints.
+%
+%A better approach is to \emph{synthesize} such conversion functions. However,
+%prior work on synthesizing lenses assumes a bijection between the source and
+%target, which is a severe limitation in practice. Oftentimes, fields present in
+%one format are not present in the other: the commands scheduled to run on Linux
+%machines are typically different than the commands scheduled to run on Windows
+%machines; updated APIs can expose new information, or omit legacy information;
+%ICS files contain a URL field that Google Calendar CSV files do not expose.
+%
+%\emph{Symmetric lenses}~\cite{symmetric-lenses} formalize this richer class of
+%non-bijective conversion functions by using additional state, called the
+%\emph{complement}, to help with synchronization. While increasing
+%expressiveness, the existence of the complement makes symmetric lenses difficult
+%to use in real-world synchronization tasks, because lens programs must maintain
+%the synchronized files as well as the complements. The complement further
+%complicates synthesis, because input-output examples no longer suffice; instead
+%edit sequences must be provided.
+%
+%In this paper, we contrain symmetric lenses to \emph{simple symmetric lenses}.
+%These constrained lenses do not use a complement, and only use the files that we
+%aim to synchronize as state. We identify a property, \emph{forgetfulness}, that
+%identifies the subset of symmetric lenses that can be written as simple
+%symmetric lenses. We provide combinators for building simple symmetric lenses on
+%string data, and integrate these combinators with Boomerang, a language for
+%bidirectional programming on strings. Both Boomerang and our extension type lens
+%terms between pairs of regular expressions. The typing judgment $\Lens : \Regex
+%\Leftrightarrow \RegexAlt$ means the lens term $\Lens$ encodes synchronizing
+%functions between the format described by $\Regex$ and the format described by
+%$\RegexAlt$.
+%
+%We provide a tool for synthesizing simple symmetric lenses when given a suite of
+%input-output examples and a pair of regular expressions describing the data
+%formats. The synthesis algorithm uses type-directed synthesis
+%on the two regular expressions to synthesize a lens that is well-typed between
+%the provided regular expressions and satisfies the provided input-output
+%examples.
+%
+%\saz{This paragraph is hard to follow --- maybe introduce $\ell : S
+%  \Leftrightarrow T$ and use the names $S$ and $T$?}\bcp{The names make it a
+%bit better, but it is all still rather abstract.  Indeed, the whole intro is
+%quite abstract---could it not be driven by an example?} A primary difficulty is
+%that there are a relatively many well-typed simple symmetric lenses. If we are
+%trying to find a lens between $\Regex$ and $\RegexAlt$, there is now a choice:
+%do $\Regex$ and $\RegexAlt$ represent the same data, and should be aligned
+%(changes to data in the language of $\Regex$ are propagated to synchronized
+%strings in the language of $\RegexAlt$), or should it be left unaligned (changes
+%to data in $\Regex$ update synchronized strings in $\RegexAlt$). Or instead
+%should certain components of $\Regex$ be aligned with certain components of
+%$\RegexAlt$, with the other components left unaligned.
+%
+%Operating under the assumption that users wish to align as much data as
+%possible, our algorithm tries to align as much data as the types and examples
+%allow. This task gets harder when there are more complex decisions involved. For
+%example, if there are two fields in one format, and one in the other, which field
+%in the first format should be aligned to the sole field of the second format? Or
+%would it be better for the field on the right to be aligneded to a mish-mash
+%of subcomponents from each field?\bcp{Wouldn't these questions be answered
+%  by examples?} To answer these questions, we develop a cost
+%metric for reasoning about which lens aligns the most information.
+%
+%In particular, the \emph{cost} of a lens is the expected number of bits required to
+%recover a string in one format when a synchronized string in the other format is
+%provided. In this model, bijective lenses have zero cost, and as more
+%information is left unsychronized, the cost is increased.
+%
+%Our model is based on \emph{stochastic regular expressions}~\cite{?}, which
+%represent both a language and a probability distribution over that language
+%simultaneously.  We develop a way to syntactically infer the expected
+%information content (or \emph{entropy}) of an unambiguous regular
+%expression. This entropy is used to calculate the expected information content
+%to recover a string, when given a string it is synchronized with.\bcp{I
+%know we're going to get to more technical detail later, but at this point I
+%am confused: a string matching a RE can in general be arbitrarily long, so
+%what kind of sense does it make to talk about its ``expected'' information content?}
+%
+%\begin{figure}
+%  \includegraphics[width=.5\textwidth]{high-level-algorithm.pdf}
+%  \caption{Schematic Diagram for \SOptician. Regular expressions \Regex and
+%    \RegexAlt, and a set of examples \Examples, are provided as input. \RXSearch
+%    searches through stochastic regular expression pairs, equivalent to the
+%    originals, and proposes them to \GreedySynth. \GreedySynth finds a lens,
+%    typed between the generated equivalent pairs. When \RXSearch determines it
+%    has an optimal lens, it returns that lens.}
+%  \label{fig:high-level-algorithm}
+%\end{figure}
+%
+%\bcp{It gets rather technical and incomprehensible for the next three
+%  paragraphs...} 
+%
+%We develop an algorithm that synthesizes a low-cost lens, shown in
+%Figure~\ref{fig:high-level-algorithm}. This algorithm consists of two
+%communicating synthesizers, one that proposes candidate stochastic regular
+%expression pairs that it thinks will align well, and one that greedily searches
+%for a lens between those two regular expressions that minimizes cost and does
+%not traverse star-related equivalences\bcp{???}. When the first algorithm is content with
+%the lens it has found, it returns that lens. 
+%
+%\bcp{Make clear that these algorithms are based on earlier / published ones.}Our first algorithm, \RXSearch, proposes candidate stochastic regular
+%expressions. These candidate stochastic regular expressions are generated by
+%traversing \emph{star semiring} equivalences on stochastic regular expressions.
+%These proposed stochastic regular expression pairs have equivalent languages and
+%probability distributions as the original ones: we prove that the star-semiring
+%rewrites this synthesizer applies are sound rewrites on stochastic regular
+%expressions.
+%
+%Our second algorithm, \GreedySynth, utilizes alternative languages for
+%stochastic regular expressions and lenses, \emph{stochastic DNF regular
+%  expressions} and \emph{symmetric DNF lenses}. While these languages are hard
+%to read and understand, they are highly normalized and permit an efficient
+%type-directed synthesis algorithm. We develop our cost metric on DNF lenses, and
+%provide a greedy algorithm that tries to find a symmetric DNF lens of minimal
+%cost. The synthesized symmetric DNF lens is then converted to a lens in the
+%original simple symmetric lens language.
+%
+%
+%We evaluate our algorithm on a benchmark suite consisting of synchronizing file
+%formats from configuration files, application-specific data storage files, and
+%data cleaning tasks. We find that our algorithm is able to synthesize each of
+%these lenses in under 20 seconds.
+%
+%In summary, our contributions are:
+%\begin{enumerate}
+%\item We develop \emph{simple symmetric lenses} a new formulation for
+%  bidirectional programs that can project data from both sides~(\S\ref{sec:overview}),
+%  and describe a language that expresses such lenses on string
+%  data~(\S\ref{sec:ssl}).  \bcp{A new language, or a variant of
+%    Hofmann/Pierce/Wagner?} 
+%\item We develop a means to traverse stochastic regular expression equivalences,
+%  while preserving probability distributions and
+%  languages~(\S\ref{subsec:stoch-rx}).
+%\item We develop stochastic DNF regular expressions and DNF
+%  lenses~(\S\ref{subsec:greedy-synth}). We develop a sound algorithm for
+%  converting stochastic regular expressions to stochastic DNF regular
+%  expressions and a information-theoretic cost metric on DNF lenses.
+%\item We extended Boomerang with our simple symmetric lens combinators, and
+%  integrated our algorithm with the Optician synthesis system within Boomerang,
+%  allowing synthesis tasks as Boomerang
+%  expressions~(\S\ref{sec:implementation}).
+%\item We evaluate our algorithm on a benchmark suite consisting of synchronizing
+%  file formats from configuration files, application-specific data storage
+%  files, and data cleaning tasks (\S\ref{sec:evaluation}).  \bcp{... and
+%    conclude what?}
+%\end{enumerate}
+
 %%% Local Variables:

### 2018-07-02 papers/QuotientLenses/main.tex
Converted to camera ready state

--- 2018-06-29 papers/QuotientLenses/main.tex
@@
 
-\documentclass[acmsmall,review,anonymous]{acmart}
-\settopmatter{printfolios=true,printccs=false,printacmref=false}
-% % For double-blind review submission, w/ CCS and ACM Reference
-% \documentclass[acmsmall,review,anonymous]{acmart}\settopmatter{printfolios=true}
-% % For single-blind review submission, w/o CCS and ACM Reference (max
-% submission space)
-% \documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
-% % For single-blind review submission, w/ CCS and ACM Reference
-% \documentclass[acmsmall,review]{acmart}\settopmatter{printfolios=true} % For
-% final camera-ready submission, w/ required CCS and ACM Reference
-% \documentclass[acmsmall]{acmart}\settopmatter{}
-
-% % Journal information % Supplied to authors by publisher for camera-ready
-% submission; % use defaults for review submission.
+\documentclass[acmsmall,screen]{acmart}
+
+\pdfpagewidth=8.5in 
+\pdfpageheight=11in
+
+\settopmatter{}
+
+\setcopyright{none}
+\acmPrice{}
+\acmDOI{10.1145/3234603}
+\acmYear{2018}
+\copyrightyear{2018}
 \acmJournal{PACMPL}
-\acmVolume{1}
-\acmNumber{ICFP} % CONF = POPL or ICFP or OOPSLA
-\acmArticle{1}
-\acmYear{2018}
-\acmMonth{1}
-\acmDOI{} % \acmDOI{10.1145/nnnnnnn.nnnnnnn}
-\startPage{1}
-
-% % Copyright information % Supplied to authors (based on authors' rights
-% management selection; % see authors.acm.org) by publisher for camera-ready
-% submission; % use 'none' for review submission.
-\setcopyright{none}
-\usepackage{algorithm, amsmath, amssymb, verbatim, enumerate, graphicx,
-  centernot, tikz, array, mathtools, bussproofs, stmaryrd, enumitem,
-  stackengine, subcaption}
-\captionsetup{compatibility=false}
-\usepackage{relsize}
-\usepackage{listings}
-\usepackage{proof}
-\usepackage{xspace}
+\acmVolume{2}
+\acmNumber{ICFP}
+\acmArticle{14}
+\acmMonth{9}
+
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10011007.10011006.10011050.10011017</concept_id>
+<concept_desc>Software and its engineering~Domain specific languages</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10011007.10011006.10011066.10011070</concept_id>
+<concept_desc>Software and its engineering~Application specific development environments</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Software and its engineering~Domain specific languages}
+\ccsdesc[300]{Software and its engineering~Application specific development environments}
+
+\keywords{Bidirectional Programming, Program Synthesis, Type-Directed Synthesis,
+Type Systems}
+
+\bibliographystyle{ACM-Reference-Format}
+\citestyle{acmauthoryear}
+
+\author{Solomon Maina}
+\affiliation{
+  \institution{University of Pennsylvnia}
+  \country{USA}
+}
+\email{smaina@seas.upenn.edu}
+
+\author{Anders Miltner}
+\affiliation{
+  \institution{Princeton University}
+  \country{USA}
+}
+\email{amiltner@cs.princeton.edu}
+
+\author{Kathleen Fisher}
+\affiliation{
+  \institution{Tufts University}
+  \country{USA}
+}
+\email{kfisher@eecs.tufts.edu}
+
+\author{Benjamin C. Pierce}
+\affiliation{
+  \institution{University of Pennsylvania}
+  \country{USA}
+}
+\email{bcpierce@cis.upenn.edu}
+
+\author{David Walker}
+\affiliation{
+  \institution{Princeton University}
+  \country{USA}
+}
+\email{dpw@cs.princeton.edu}
+
+\author{Steve Zdancewic}
+\affiliation{
+  \institution{University of Pennsylvania}
+  \country{USA}
+}
+\email{stevez@cis.upenn.edu}
+
+\usepackage{algorithm, amsmath, amssymb, verbatim, enumerate, graphicx, centernot, tikz, array, mathtools, bussproofs, stmaryrd, enumitem, stackengine, subcaption,flushend,relsize, listings, proof, xspace, booktabs}
 \usepackage[noend]{algpseudocode}
 \usepackage[capitalize]{cleveref}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
+\usepackage{microtype}
+
+\captionsetup{compatibility=false}
 
@@
 
+\clubpenalty = 10000
+\widowpenalty = 10000
+\displaywidowpenalty = 10000
 % %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
@@
 
-\keywords{quotient regular expressions, QRE lens, synthesis}
 \maketitle
-
+\renewcommand{\shortauthors}{S. Maina, A. Miltner, K. Fisher, B. C. Pierce, D. Walker, S. Zdancewic}
 \section{Introduction}
@@
 from FlashFill~\cite{flashfill}, and benchmarks derived from
-Augeas~\cite{augeas2} (we also experimented using our QRE implementation to synthesize quotient lenses between XML, RDF and and JSON formats using data from data.gov database; the data consisted of census statistics, demographic statistics, wage comparosion data, and crime index data).
+Augeas~\cite{augeas2} (we also experimented using our QRE implementation to synthesize quotient lenses between XML, RDF and and JSON formats using data from the data.gov database; the data consisted of census statistics, demographic statistics, wage comparosion data, and crime index data).
 In these 39 benchmarks, 10 of
@@
 
-\bibliographystyle{plain}
-\bibliography{local}
-
 \appendix
@@
 \Leftrightarrow K(Q_2)$ such that:
-\begin{align*}
-\llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket\circ
-\canonize(Q_1)\\
-\llbracket q \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
+$$
+\llbracket q \rrbracket.\get = \llbracket \ell \rrbracket\circ
+\canonize(Q_1) \text{ and } 
+\llbracket q \rrbracket.\lput = \llbracket \ell \rrbracket^{-1} \circ
 \canonize(Q_2)
-\end{align*}
+$$
 We now prove this theorem.
@@
 \Leftrightarrow Q_3$, $Q_1$ is well formed and $K(Q_1) = W(Q_2)$. Then:
-\begin{align*}
-\llbracket \lquot(Q_1, q) \rrbracket.\get  &= \llbracket q
-\rrbracket.\get \circ \canonize(Q_1)\\
-\llbracket \lquot(Q_1, q) \rrbracket.\lput &= \llbracket q \rrbracket.\lput
-\end{align*}
+$$
+\llbracket \lquot(Q_1, q) \rrbracket.\get  = \llbracket q
+\rrbracket.\get \circ \canonize(Q_1) \text{ and }
+\llbracket \lquot(Q_1, q) \rrbracket.\lput = \llbracket q \rrbracket.\lput
+$$
 By the induction hypothesis, there exists a bijective lens $\ell :
 K(Q_2) \Leftrightarrow K(Q_3)$ such that:
-\begin{align*}
-\llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket \circ
-\canonize(Q_2)\\
-\llbracket q \rrbracket.\lput &= \llbracket \ell \rrbracket^{-1} \circ
+$$
+\llbracket q \rrbracket.\get = \llbracket \ell \rrbracket \circ
+\canonize(Q_2) \text{ and }
+\llbracket q \rrbracket.\lput = \llbracket \ell \rrbracket^{-1} \circ
 \canonize(Q_3)
-\end{align*}
+$$
 Consequently
@@
 $q_2 : Q_1 \Leftrightarrow Q_2$. Then:
-\begin{align*}
-\llbracket q_1 \; ; \; q_2 \rrbracket.\get &= \llbracket q_2
-\rrbracket.\get\circ \llbracket q_1 \rrbracket.\get, \text{ and }\\
-\llbracket q_1 \; ; \; q_2 \rrbracket.\lput &= \llbracket q_1 \rrbracket.\lput
+$$
+\llbracket q_1 \; ; \; q_2 \rrbracket.\get = \llbracket q_2
+\rrbracket.\get\circ \llbracket q_1 \rrbracket.\get \text{ and }
+\llbracket q_1 \; ; \; q_2 \rrbracket.\lput = \llbracket q_1 \rrbracket.\lput
 \circ \llbracket q_2 \rrbracket.\lput
-\end{align*}
+$$
 By the induction hypothesis, there exist bijective lenses,
@@
 that,
-\begin{align*}
-\llbracket q_1 \rrbracket.\get &= \llbracket \ell_1 \rrbracket \circ
-\canonize(c)\\
-\llbracket q_1 \rrbracket.\lput &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
-\canonize(Q_1)
-\end{align*}
-and
-\begin{align*}
-\llbracket q_2 \rrbracket.\get &= \llbracket \ell_2 \rrbracket \circ
+\[\begin{array}{l l}
+\llbracket q_1 \rrbracket.\get = \llbracket \ell_1 \rrbracket \circ
+\canonize(c) & \llbracket q_2 \rrbracket.\get = \llbracket \ell_2 \rrbracket \circ
 \canonize(Q_1)\\
-\llbracket q_2 \rrbracket.\lput &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
+\llbracket q_1 \rrbracket.\lput = {\llbracket \ell_1 \rrbracket}^{-1} \circ
+\canonize(Q_1) & \llbracket q_2 \rrbracket.\lput = {\llbracket \ell_2 \rrbracket}^{-1} \circ
 \canonize(Q_2)
-\end{align*}
+\end{array}
+\]
 Consequently:
@@
 \end{align*}
-A similar argument shows that:
-$$\llbracket q_1 \; ; \; q_2 \rrbracket.\lput =
+A similar argument shows that $\llbracket q_1 \; ; \; q_2 \rrbracket.\lput =
 \llbracket \ell_1 \; ; \; \ell_2 \rrbracket^{-1} \circ
-\canonize(c)$$
+\canonize(c)$
 \item
@@
 $W(Q_1)^{*!}$ and $W(Q_2)^{*!}$ and $K(Q_1)^{*!}$ and $K(Q_2)^{*!}$. Then:
-\begin{align*}
-\llbracket q^* \rrbracket.\get &= (\llbracket q \rrbracket.\get)^*, \text{
-and }\\
-\llbracket q^* \rrbracket.\lput &= (\llbracket q \rrbracket.\lput)^*
-\end{align*}
+$$
+\llbracket q^* \rrbracket.\get = (\llbracket q \rrbracket.\get)^* \text{ and }
+\llbracket q^* \rrbracket.\lput = (\llbracket q \rrbracket.\lput)^*
+$$
 By the induction hypothesis there exists a bijective lens $\ell : K(Q_1)
 \Leftrightarrow K(Q_2)$ such that:
-\begin{align*}
-\llbracket q \rrbracket.\get &= \llbracket \ell \rrbracket \circ
-\canonize(Q_1)\\
-\llbracket q \rrbracket.\lput &= {\llbracket \ell \rrbracket}^{-1} \circ
+$$
+\llbracket q \rrbracket.\get = \llbracket \ell \rrbracket \circ
+\canonize(Q_1) \text{ and }
+\llbracket q \rrbracket.\lput = {\llbracket \ell \rrbracket}^{-1} \circ
 \canonize(Q_2)
-\end{align*}
+$$
 Consequentlty:
@@
 K(Q_3) \cdot^! K(Q_4)$. Then:
-\begin{align*}
-\llbracket q_1 \cdot q_2 \rrbracket.\get &= \llbracket q_1 \rrbracket.\get \cdot
-\llbracket q_2 \rrbracket.\get, \text{ and }\\
-\llbracket q_1 \cdot q_2 \rrbracket.\lput &= \llbracket q_1 \rrbracket.\lput
+$$
+\llbracket q_1 \cdot q_2 \rrbracket.\get = \llbracket q_1 \rrbracket.\get \cdot
+\llbracket q_2 \rrbracket.\get \text{ and }
+\llbracket q_1 \cdot q_2 \rrbracket.\lput = \llbracket q_1 \rrbracket.\lput
 \cdot \llbracket q_2 \rrbracket.\lput
-\end{align*}
+$$
 By the induction hypothesis, there exist bijective lenses $\ell_1 : K(Q_1)
 \Leftrightarrow K(Q_3)$ and $\ell_2 : K(Q_2) \Leftrightarrow K(Q_4)$ such that,
-\begin{align*}
-\llbracket q \rrbracket.\get &= \llbracket \ell_1 \rrbracket \circ
-\canonize(Q_1)\\
-\llbracket q \rrbracket.\lput &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
-\canonize(Q_3)
-\end{align*}
-and
-\begin{align*}
-\llbracket q_2 \rrbracket.\get &= \llbracket \ell_2 \rrbracket \circ
+\[\begin{array}{l l}
+\llbracket q \rrbracket.\get = \llbracket \ell_1 \rrbracket \circ
+\canonize(Q_1) & \llbracket q_2 \rrbracket.\get = \llbracket \ell_2 \rrbracket \circ
 \canonize(Q_2)\\
-\llbracket q_2 \rrbracket.\lput &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
+\llbracket q \rrbracket.\lput = {\llbracket \ell_1 \rrbracket}^{-1} \circ
+\canonize(Q_3) & \llbracket q_2 \rrbracket.\lput = {\llbracket \ell_2 \rrbracket}^{-1} \circ
 \canonize(Q_4)
-\end{align*}
+\end{array}
+\]
 Consequently:
@@
 \end{align*}
-Similarly:
-$$
+Similarly, $
 \llbracket q_1 \cdot q_2\rrbracket.\lput = \llbracket \ell_1 \cdot  \ell_2
-\rrbracket^{-1} \circ \canonize(Q_3 \cdot Q_4) $$
+\rrbracket^{-1} \circ \canonize(Q_3 \cdot Q_4)$
 \item
@@
 \Leftrightarrow K(Q_3)$ and $\ell_2 : K(Q_2) \Leftrightarrow K(Q_4)$ such that,
-\begin{align*}
-\llbracket q_1 \rrbracket.\get &= \llbracket \ell_1 \rrbracket \circ
-\canonize(Q_1)\\
-\llbracket q_1 \rrbracket.\lput &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
-\canonize(Q_3)
-\end{align*}
-and
-\begin{align*}
-\llbracket q_2 \rrbracket.\get &= \llbracket \ell_2 \rrbracket \circ
+\[\begin{array}{l l}
+\llbracket q_1 \rrbracket.\get = \llbracket \ell_1 \rrbracket \circ
+\canonize(Q_1) & \llbracket q_2 \rrbracket.\get = \llbracket \ell_2 \rrbracket \circ
 \canonize(Q_2)\\
-\llbracket q_2 \rrbracket.\lput &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
+\llbracket q_1 \rrbracket.\lput = {\llbracket \ell_1 \rrbracket}^{-1} \circ
+\canonize(Q_3) & \llbracket q_2 \rrbracket.\lput = {\llbracket \ell_2 \rrbracket}^{-1} \circ
 \canonize(Q_4)
-\end{align*}
+\end{array}
+\]
 Consequently:
@@
 \end{align*}
-Similarly:
-$$
+Similarly, $
 \llbracket \swap(q_1, q_2) \rrbracket.\lput = (\llbracket \swap(\ell_1, \ell_2)
-\rrbracket)^{-1} \circ (\canonize(Q_4) \cdot \canonize(Q_3))$$
+\rrbracket)^{-1} \circ (\canonize(Q_4) \cdot \canonize(Q_3))$
 \item
@@
 \Leftrightarrow K(Q_3)$ and $\ell_2 : K(Q_2) \Leftrightarrow K(Q_4)$ such that,
-\begin{align*}
-\llbracket q_1 \rrbracket.\get &= \llbracket \ell_1 \rrbracket \circ
-\canonize(Q_1)\\
-\llbracket q_1 \rrbracket.\lput &= {\llbracket \ell_1 \rrbracket}^{-1} \circ
-\canonize(Q_3)
-\end{align*}
-and
-\begin{align*}
-\llbracket q_2 \rrbracket.\get &= \llbracket \ell_2 \rrbracket \circ
+\[\begin{array}{l l }
+\llbracket q_1 \rrbracket.\get = \llbracket \ell_1 \rrbracket \circ
+\canonize(Q_1) & \llbracket q_2 \rrbracket.\get = \llbracket \ell_2 \rrbracket \circ
 \canonize(Q_2)\\
-\llbracket q_2 \rrbracket.\lput &= {\llbracket \ell_2 \rrbracket}^{-1} \circ
+\llbracket q_1 \rrbracket.\lput = {\llbracket \ell_1 \rrbracket}^{-1} \circ
+\canonize(Q_3) & \llbracket q_2 \rrbracket.\lput = {\llbracket \ell_2 \rrbracket}^{-1} \circ
 \canonize(Q_4)
-\end{align*}
+\end{array}
+\]
 Consequently:
@@
 so $\llbracket q_1 \sep q_2 \rrbracket.\get = \llbracket \ell_1 \sep
-\ell_2 \rrbracket \circ \canonize(Q_1 \sep Q_2)$. A similar argument shows
-that $\llbracket q_1 \sep q_2 \rrbracket.\lput = \llbracket \ell_1 \sep
+\ell_2 \rrbracket \circ \canonize(Q_1 \sep Q_2)$. Similarly, $\llbracket q_1 \sep q_2 \rrbracket.\lput = \llbracket \ell_1 \sep
 \ell_2 \rrbracket^{-1} \circ \canonize(Q_3 \sep Q_4)$.\\
@@
 
+\bibliography{local}
+
 \end{document}

### 2018-07-02 papers/SymmetricLenses/main.tex

@@
 %a search through the space of possible computations that can be expressed as a lens.
-Recent research~\cite{optician} has exploited this fact by demonstrating that it
+Recent research~\cite{optician, maina+:quotient-synthesis} has exploited this fact by demonstrating that it
 is possible to synthesize relatively large, complex lenses from specifications.
 In particular, given types that describe source and target formats, there are
-synthesis algorithms that will find bijective lenses~\cite{optician} for files
+synthesis algorithms\bcp{more than one?  I know there are two papers, but
+  it's really the same synthesis algorithm} that will find bijective
+lenses
+% ~\cite{optician}
+for files 
 in ad-hoc formats, such as those housing linux application configurations.
@@
 the aid of any examples at all, with just types describing the source and target
-formats. \dpw{I would cite the ICFP paper (cite string:
-  maina+:quotient-synthesis) as to appear but I'm not sure how to do that
-  without possibly violating the spirit of double-blind reviewing. ie, if it
-  hasn't appeared yet, how could we know the content if we aren't the authors?
-  Mentioning it in the 3rd person does not seem to suffice ...} \ksf{The ICFP
-  2018 website lists the paper; I think it is okay to cite it. The reviewer can
-  assume we contacted the authors for a copy. We could ask Stephanie for a
-  ruling if we want to be sure. One of the principles of double blind as adopted
-  by SIGPLAN is not to reduce the quality of the paper; omitting this citation
-  would have that affect.}\bcp{+1}\saz{Stephanie says we should cite it in the
-  third person, even if that's a bit awkward, but otherwise citing it is OK.}
+formats. 
+% \dpw{I would cite the ICFP paper (cite string:
+%   maina+:quotient-synthesis) as to appear but I'm not sure how to do that
+%   without possibly violating the spirit of double-blind reviewing. ie, if it
+%   hasn't appeared yet, how could we know the content if we aren't the authors?
+%   Mentioning it in the 3rd person does not seem to suffice ...} \ksf{The ICFP
+%   2018 website lists the paper; I think it is okay to cite it. The reviewer can
+%   assume we contacted the authors for a copy. We could ask Stephanie for a
+%   ruling if we want to be sure. One of the principles of double blind as adopted
+%   by SIGPLAN is not to reduce the quality of the paper; omitting this citation
+%   would have that affect.}\bcp{+1}\saz{Stephanie says we should cite it in the
+%   third person, even if that's a bit awkward, but otherwise citing it is OK.}
 
@@
   can do something similar here.}\afm{What about lenses between ``a''* and
-  ''b``*}
+  ''b``*}\bcp{Aren't there an infinite number of those, even if we restrict
+  to bijective ones?}
 
@@
   views it as the amount of information if we were communicating a stream of
-  these choices.} On the other hand, consider the language of the SRE \lstinline{c}. It contains
+  these choices.}\bcp{I see---very interesting.  What about mentioning this
+  point, to get people ready for some information-theoretic thinking
+  later...?} 
+On the other hand, consider the language of the SRE \lstinline{c}. It contains 
 just one element (the string $c$). If we observe that our algorithm generates
@@
 
-In summary, our main contributions are:\bcp{annotate with section numbers}
+In summary, our main contributions are:
 \begin{itemize}
@@
   asymmetric lenses and full symmetric lenses and has the useful property that
-  they don't have any internal state (\S\ref{sec:overview}). We characterize the
-  simple lenses semantically as the subset of full symmetric lenses for which a
-  ``forgetfulness'' property holds (\S\ref{sec:related}). We also analyze the
-  gap between simple and full symmetric lenses empirically and argue that simple
-  ones suffice for many applications (\S\ref{sec:evaluation}).
+  its members don't have any internal state (\S\ref{sec:overview}). 
 \item We show how to synthesize simple symmetric lenses (including all asymmetric lenses as a special case),
@@
   synthesize all of them in under half a minute (\S\ref{sec:evaluation}).
+%
+  We also analyze the gap between simple and full symmetric lenses
+  empirically and argue that simple ones suffice for many applications.
+\item We show that the class of simple lenses can be characterized
+  semantically as the subset of full symmetric lenses for which a simple
+  ``forgetfulness'' property holds (\S\ref{sec:related}).  \bcp{I suggest
+    splitting this out into its own section, since there is other related
+    work that we'll want to discuss in a Related Work section.}
 \end{itemize}
@@
 HR adds a new employee before management, in which case the sentinel
-value ``N/A'' reflects the unknown salary information. 
+value ``N/A'' reflects the unknown salary information\bcp{can we just use
+  ``UNKNOWN'' on both sides?  Or is there some reason to make it more
+  complicated?} 
 Furthermore, if HR
@@
 and $\RegexAlt$ 
-is defined by four functions
+is defined by four functions\bcp{I think a picture would be helpful.  I can probably
+  make one pretty easily by tweaking the one from the original symmetric paper.}
 
@@
 
-In contrast to the identity lens, the disconnect lens \Disconnect leaves data
-unconnected. The disconnect lens takes four arguments: two regular 
+In contrast to the identity lens, the disconnect lens, \Disconnect, leaves data
+unconnected.\bcp{Can we call it ``disconnect''?} The disconnect lens takes four arguments: two regular 
 expressions and two strings. The regular expressions specify the
@@
 \end{center}
-We have proven that this syntactic computation corresponds to the entropy of the
+We have proven\bcp{This is a POPL paper: Proofs of important theorems should
+be included.} that this syntactic computation corresponds to the entropy of the
 stochastic regular expression.
@@
 We have proven theorems demonstrating the calculated entropy corresponds to the
-actual conditional entropy to recover the data.
+actual conditional entropy to recover the data.\bcp{Again, they should be
+  included.}
 
@@
 \]
-We have proven this conversion respects languages and probability distributions.
+We have proven this conversion respects languages and probability
+distributions.\bcp{Should be included.}
 

### 2018-07-03 papers/SymmetricLenses/main.tex
Some more comments

@@
 
+\bcp{This is a POPL paper: Proofs of important theorems should
+be included.  (This one might be easy enough that we can just label it as a
+``Fact'' and omit the proof, or better yet we can say that the proof is straightforward.}
+
 \subsection{Stochastic Regular Expression Entropy}
-The entropy of a stochastic regular expression can be computed directly from its
-syntax, when the stochastic regular expression is unambiguous and contains only
-nonempty subcomponents.
+The {\em entropy} of a stochastic regular expression can be computed
+directly from its syntax, when the stochastic regular expression is
+unambiguous and contains only nonempty subcomponents.\bcp{What is an empty
+  subcomponent?} 
 \begin{center}
@@
 \end{center}
-We have proven\bcp{This is a POPL paper: Proofs of important theorems should
-be included.} that this syntactic computation corresponds to the entropy of the
-stochastic regular expression.
+We have proven\bcp{Again, some kind of proof is needed.} that this syntactic
+computation corresponds to the entropy of the stochastic regular expression.
 
@@
 $\PRegexStar{\DNFRegex}{.8}$. In particular, $\ToDNFRegex$ generates regular
-expressions whose DNF form gives equal probability to all sequences, and gives
+expressions whose DNF form gives equal probability to all sequences\bcp{Why?}, and gives
 iterations a .8 chance to continue iterating, as that gives a reasonable
-distribution of strings when generating them from the regular expressions.
+distribution of strings when generating them from the regular
+expressions\bcp{What evaluation did we do to
+  determine that it is the right value?}.
 

### 2018-07-03 papers/SymmetricLenses/main.tex
Some more comments

@@
 \usepackage{xspace}
+\usepackage{bbm}
+
+\newtheorem*{theorem*}{Theorem}
 
@@
 \EntropyOf{\lstinline{emp_row}}$. \lstinline{emp_salary} contains strictly more
-imformation than $\EntropyOf{\lstinline{": " . salary}}$, and
+information than $\EntropyOf{\lstinline{": " . salary}}$, and
 \lstinline{emp_row} contains strictly more information than
@@
     & =
-    & $\Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\Regex}{\String_i}$\\
+    & $\Sigma_n \Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\Regex}{\String_i}$\\
   \end{tabular}
@@
     \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
-    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{(1-\Probability_1)*\Probability_2})}{\Probability_1*\Probability_2} & \OrAssociativityRule{}  \\
+    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{\frac{(1-\Probability_1)*\Probability_2}{1 - \Probability_1 * \Probability_2}})}{\Probability_1*\Probability_2} & \OrAssociativityRule{}  \\
     \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
@@
     \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
-    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability} & \UnrollstarLeftRule{} \\
-    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{\Probability} & \UnrollstarRightRule{} 
+    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{1-\Probability} & \UnrollstarLeftRule{} \\
+    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{1-\Probability} & \UnrollstarRightRule{} 
   \end{tabular}
@@
 recover a string in one data format from the other. We use the function
-$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate the expected amount
-of information required to recover a string in $\Regex$ from a string in
-$\RegexAlt$, synchronized by $\Lens$. We use the function
 $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to calculate the expected amount
 of information required to recover a string in $\RegexAlt$ from a string in
-$\Regex$, synchronized by $\Lens$.
+$\Regex$, synchronized by $\Lens$. We use the function $\LEntropyOf{\Regex
+  \Given \Lens, \RegexAlt}$ to calculate the expected amount of information
+required to recover a string in $\Regex$ from a string in $\RegexAlt$,
+synchronized by $\Lens$.
 \begin{center}
   \begin{tabular}{rcl}
-    $\LEntropyOf{\Regex \Given \IdentityLensOf{\Regex}, \Regex}$
+    $\REntropyOf{\RegexAlt \Given \IdentityLensOf{\RegexAlt}, \RegexAlt}$
     & =
@@
 
-    $\LEntropyOf{\Regex \Given \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, \RegexAlt}$
-    & =
-    & \EntropyOf{\Regex}\\
-
-    $\LEntropyOf{\PRegexStar{\Regex}{\Probability} \Given \IterateLensOf{\Lens}^, \PRegexStar{\RegexAlt}{\ProbabilityAlt}}$
-    & =
-    & $\frac{\ProbabilityAlt}{1-\ProbabilityAlt}(\LEntropyOf{\Regex \Given \Lens, \RegexAlt})$\\
-
-    $\LEntropyOf{\Regex_1 \Concat \Regex_2 \Given \ConcatLensOf{\Lens_1}{\Lens_2}, \RegexAlt_1 \Concat \RegexAlt_2}$
-    & =
-    & $\LEntropyOf{\Regex_1 \Given \Lens_1^\leftarrow, \RegexAlt_1} + \LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2}$\\
-
-    $\LEntropyOf{\Regex_1 \Concat \Regex_2 \Given \SwapLensOf{\Lens_1}{\Lens_2}, \RegexAlt_2 \Concat \RegexAlt_1}$
-    & =
-    & $\LEntropyOf{\Regex_1 \Given \Lens_1^\leftarrow, \RegexAlt_1} + \LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2}$\\
-
-    $\LEntropyOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability} \Given \OrLensOf{\Lens_1}{\Lens_2}, \PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}}$
-    & =
-    & $\ProbabilityAlt*(\LEntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt_1}) + (1-\ProbabilityAlt)*(\LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2})$\\
-
-    $\LEntropyOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability} \Given \OrLensOf{\Lens_1}{\Lens_2}, \PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}}$
-    & =
-    & $\ProbabilityAlt*(\LEntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt_1}) + (1-\ProbabilityAlt)*(\LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2})$\\
-
-    $\LEntropyOf{\Regex \Given \MergeROf{\Lens_1}{\Lens_2}, \PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}}$
-    & =
-    & $\ProbabilityAlt*(\LEntropyOf{\Regex \Given \Lens_1, \RegexAlt_1}) + (1-\ProbabilityAlt)*(\LEntropyOf{\Regex \Given \Lens_2, \RegexAlt_2})$\\
-
-    $\LEntropyOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability} \Given \MergeLOf{\Lens_1}{\Lens_2}, \RegexAlt}$
-    & =
-    & $\Probability*(\LEntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt} - \Log_2(\Probability)) + $\\
+    $\REntropyOf{\RegexAlt \Given \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, \Regex}$
+    & =
+    & \EntropyOf{\RegexAlt}\\
+
+    $\REntropyOf{\PRegexStar{\RegexAlt}{\ProbabilityAlt} \Given \IterateLensOf{\Lens}, \PRegexStar{\Regex}{\Probability}}$
+    & =
+    & $\frac{\Probability}{1-\Probability}(\REntropyOf{\RegexAlt \Given \Lens, \Regex})$\\
+
+    $\REntropyOf{\RegexAlt_1 \Concat \RegexAlt_2 \Given \ConcatLensOf{\Lens_1}{\Lens_2}, \Regex_1 \Concat \Regex_2}$
+    & =
+    & $\REntropyOf{\RegexAlt_1 \Given \Lens_1, \Regex_1} + \REntropyOf{\RegexAlt_2 \Given \Lens_2, \Regex_2}$\\
+
+    $\REntropyOf{\RegexAlt_2 \Concat \RegexAlt_1 \Given \SwapLensOf{\Lens_1}{\Lens_2}, \Regex_1 \Concat \Regex_2}$
+    & =
+    & $\REntropyOf{\RegexAlt_2 \Given \Lens_1, \Regex_2} + \REntropyOf{\RegexAlt_1 \Given \Lens_1, \Regex_1}$\\
+
+    $\REntropyOf{\PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt} \Given \OrLensOf{\Lens_1}{\Lens_2}, \PRegexOr{\Regex_1}{\Regex_2}{\Probability}}$
+    & =
+    & $\Probability*(\REntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt_1} ) + (1-\Probability)*(\REntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2})$\\
+
+    $\REntropyOf{\PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt} \Given \MergeLOf{\Lens_1}{\Lens_2}, \Regex}$
+    & =
+    & $\ProbabilityAlt*(\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1}) + (1-\ProbabilityAlt)*(\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2})$\\
+
+    $\REntropyOf{\RegexAlt \Given \MergeROf{\Lens_1}{\Lens_2}, \PRegexOr{\Regex_1}{\Regex_2}{\Probability}}$
+    & =
+    & $\Probability*(\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1} - \Log_2(\Probability)) + $\\
     &
-    & $(1-\Probability)*(\LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt} - \Log_2(1 - \Probability))$\\
-
-    $\LEntropyOf{\RegexAlt \Given \InvertOf{\Lens}, \Regex}$
-    & =
-    & $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$\\
+    & $(1-\Probability)*(\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2} - \Log_2(1 - \Probability))$\\
+
+    $\REntropyOf{\Regex \Given \InvertOf{\Lens}, \RegexAlt}$
+    & =
+    & $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$\\
   \end{tabular}
 \end{center}
-$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ is defined symmetrically.  These
+$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ is defined symmetrically.  These
 functions calculate the expected number of bits to recover one data format from
@@
 We have proven theorems demonstrating the calculated entropy corresponds to the
-actual conditional entropy to recover the data.\bcp{Again, they should be
+actual conditional entropy to recover the data.\bcp{Again, the proofs should be
   included.}
@@
   \begin{enumerate}
+  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ calculates the expected
+    information content of $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$,
+    given $\SetOf{s \SuchThat s \in \LanguageOf{\Regex} \BooleanAnd
+      \Lens.\PutROf{s}{t} = t}$
   \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ calculates the expected
@@
       \Lens.\PutLOf{t}{s} = s}$
-  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ calculates the expected
-    information content of $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$,
-    given $\SetOf{s \SuchThat s \in \LanguageOf{\Regex} \BooleanAnd
-      \Lens.\PutROf{s}{t} = t}$
   \end{enumerate}
@@
     & =
-    & $\Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\DNFRegex}{\String_i}$\\
+    & $\Sigma_n \Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\DNFRegex}{\String_i}$\\
 
@@
 
-    $\EntropyOf{\SequenceOf{\String_0 \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_n \SeqSep \String_n}}{\String'}$
+    $\EntropyOf{\SequenceOf{\String_0 \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_n \SeqSep \String_n}}$
     & =
@@
 string lenses, DNF lenses do not have composition, so the entropy can be defined
-over all terms.  $\LEntropyOf{\DNFRegex \Given \SDNFLens, \DNFRegexAlt}$ is the
+over all terms.  $\REntropyOf{\DNFRegex \Given \SDNFLens, \DNFRegexAlt}$ is the
 expected number of bits required to recover a string in $\DNFRegex$ from a
@@
   \begin{array}{rl}
-    & \LEntropyOf{\DNFOf{\Sequence_1 \DNFSep \ldots \DNFSep \Sequence_n}\\
+    & \REntropyOf{\DNFOf{\Sequence_1 \DNFSep \ldots \DNFSep \Sequence_n}\\
     & \hspace*{1.21em}\Given
@@
     & \text{where } H_j = \frac{\Sigma_{k\SuchThat j_k = j}
-      \LEntropyOf{\Sequence_{i_k}\Given\SequenceLens_k,\SequenceAlt_j}}{\SizeOf{\SetOf{k\SuchThat
+      \REntropyOf{\Sequence_{i_k}\Given\SequenceLens_k,\SequenceAlt_j}}{\SizeOf{\SetOf{k\SuchThat
       j_k = j}}} + log_2\SizeOf{\SetOf{k\SuchThat j_k = j}}
@@
   \begin{array}{rl}
-      & \LEntropyOf{
+      & \REntropyOf{
         \SequenceOf{\String_0,\Atom_1,\ldots,\Atom_n,\String_n}\\
@@
       =
-      & \Sigma_{x=1}^p\LEntropyOf{\Atom_{i_x} \Given \AtomAlt_{i_x},\AtomLens_x} +
-        \Sigma_{x=1}^q\LEntropyOf{\Atom_{k_x}}
+      & \Sigma_{x=1}^p\REntropyOf{\Atom_{i_x} \Given \AtomAlt_{i_x},\AtomLens_x} +
+        \Sigma_{x=1}^q\REntropyOf{\Atom_{k_x}}
   \end{array}
@@
 \[\begin{array}{rl}
-      & \LEntropyOf{\PRegexStar{\DNFRegex}{\Probability} \Given \StarOf{\DNFRegexAlt},\StarOf{\SDNFLens}}\\
+      & \REntropyOf{\PRegexStar{\DNFRegex}{\Probability} \Given \StarOf{\DNFRegexAlt},\StarOf{\SDNFLens}}\\
       =\\
-      & 4*\LEntropyOf{\DNFRegex\Given\DNFRegexAlt,\SDNFLens}
+      & 4*\REntropyOf{\DNFRegex\Given\DNFRegexAlt,\SDNFLens}
   \end{array}
@@
 mapped to each other either if they are the same regular expression and have
-examples agreeing with the examples, or if there is a lens which is a
+examples agreeing with the examples\bcp{What does ``have
+examples agreeing with the examples'' mean??}, or if there is a lens which is a
 composition of predefined lenses between them, where the lens alters the
@@
 
-\paragraph*{\RXSearch Permissiveness Customization}
-While termination heuristic works for many situations: there are situations
+\paragraph*{\RXSearch Permissiveness Customization}\bcp{I don't follow this:}
+While the termination heuristic works for many situations, there are situations
 where a large number of expansions must be performed to find the correct lens.
@@
 
-We have extended Optician with our symmetric synthesis algorithm. Users can
+We have extended Optician\bcp{this is the first mention of the word
+  Optician!} with our symmetric synthesis algorithm. Users can 
 write synthesis tasks alongside lens combinators, use those synthesis results in
@@
 
-In this work, we aim to answer two primary research questions:
+In this evaluation, we aim to answer two primary questions:
 \begin{enumerate}
 \item Is our synthesis procedure efficient enough to be used in everyday
-  development?  Can it be used in place of the existing bijective lens synthesis
-  algorithm integrated with Boomerang?
-\item Are our heuristics useful for reducing the specifications users have to
-  provide.
+  development?  In particular, can it be used in place of the existing
+  bijective lens synthesis algorithm integrated with Boomerang (which
+  \citet{miltner+:bijective-synthesis} have argued is reasonably fast in
+  practice)? 
+  \item Are the heuristics we have described useful for reducing the 
+  ``specification overhead'' for users?  \bcp{Maybe it would be better to
+    ask {\em which} heuristics are the most useful.  Also, are we talking
+    about reducing the number of examples needed?  It's not clear.  (And:
+    Don't the heuristics also make it faster??)}
 \end{enumerate}
 
-\subsection{Benchmark Suite Construction}
-Our benchmarks came from four different sources.
+\subsection{Benchmark Suite}
+Our benchmarks are drawn from four different sources.
 \begin{enumerate}
 \item We adapted \afm{8} data cleaning benchmarks from Flash Fill. In adopting
-  these benchmarks, we go beyond the data cleaning tasks, and synthesize
-  synchronization functions between the ad-hoc data and their cleaned forms.
+  these benchmarks, we go beyond the data cleaning tasks\bcp{confusing: you
+    say these are data cleaning tasks, then immediately say we go beyond
+    data cleaning tasks.  (I know the next bit of the sentence makes it
+    (somewhat) clearer, but the reader is already lost right here.)}, and
+  synthesize 
+  synchronization functions\bcp{what is a ``synchronization function''}
+  between the ad-hoc data and their cleaned forms. 
   Optician also adopted benchmarks from Flash Fill, but required alterations to
-  be in bijective form. With our symmetric enhancement, we Optician's four
-  benchmarks, and 4 new Flash Fill benchmarks with no alterations.
+  be in bijective form\bcp{Mysterious}. With our symmetric enhancement, we
+  \bcp{words missing?} Optician's four 
+  benchmarks and 4 new Flash Fill benchmarks with no alterations.
 \item We adapted \afm{29} benchmarks from Augeas.  These benchmarks are the same
@@
 
+\bcp{Why is this a good set of benchmarks for the purposes described earlier
+in the section?}
+
 \subsection{Efficiency}
 
-To answer our first research question, we analyze the runtime of our algorithm,
-and compare it to the existing algorithm for synthesizing bijective lenses.
+To answer our first research question, we measure the runtime of our algorithm
+and compare it to the existing algorithm for synthesizing bijective
+lenses.\bcp{Note that this comparison doesn't, directly answer the question
+  ``Is it fast enough.''}
 
@@
     benchmarks in our benchmark suite in under 30 seconds, and is able to
-    complete most of the benchmarks without using compositional synthesis. }
+    complete most of the benchmarks without using compositional
+    synthesis. \bcp{The interesting part of the graph is all squished
+      against the Y axis.  Can we make the X axis logarithmic?}}
   \label{fig:times}
@@
 
-\subsection{Heuristic Usefulness}
+\subsection{The Effect of Our Heuristic}
 We evaluate the usefulness of (1) our information theoretic metric, and (2) our
@@
   \FCOpt{}: &  The first time \GreedySynth finds a lens, it returns that lens.\\
-  \NMOpt{}: & \GreedySynth does not have any notion of cost, and searches for
-              any lens that satisfies the specification. \\
+  \NMCCOpt{}: & The lens cluster that contains is identified through \SSOpt{}. \GreedySynth is then called on that pair of SREs. \\
+  \NMOpt{}: & Neither \GreedySynth nor \RXSearch have any notion of cost; they
+              both merely search for
+              any lens that satisfies the specification. 
 \end{tabulary}
 
-Figure~\ref{fig:heuristic} summarizes the result of this experiment.  We find
-that \afm{TODO}.
+\begin{figure}
+  \includegraphics{generated-graphs/metrics_importance}
+  \caption{Number of benchmarks that synthesize the correct lens by a given
+    algorithm. \NoTPOpt{} is the full symmetric synthesis algorithm, but
+    termination parameters are ignored. \FCOpt{} provides a notion of cost to
+    \GreedySynth, but once a satisfying lens is greedily found, that lens is
+    returned. \NMCCOpt{} identifies the lens cluster in which the desired 
+    lens lies.  \GreedySynth{} is then run on that cluster with no notion of
+    cost. \NMOpt{} removes all notion of cost, and merely finds the first
+    satisfying lens.\bcp{Am I reading this wrong, or does even the best
+      algorithm fail to find the correct lens something like 20\% of the
+      time?!?!?!}} 
+  \label{fig:metric}
+\end{figure}
+
+Figure~\ref{fig:metric} summarizes the result of this experiment. We find that
+our metrics are very important for finding the correct lens. Our
+information-theoretic metric is critical for finding the correct lens. Indeed,
+even when provided with the correct lens cluster, without a notion of cost, we
+would need constrain our search space more with additional examples to find the
+correct lens.
+
+Our termination condition is also quite important. When running in \FCOpt{},
+only 5 lenses are found. The first cluster in which a lens is found is rarely
+the correct cluster. However, our termination condition is not perfect.
+Sometimes users must tell the algorithm to ``keep going'' or ``stop early.''
+Without termination parameters, our algorithm finds the correct lens for 36 of
+our 45 benchmarks, 9 of them need help in the form of a termination parameter
+for finding the correct lens.
 % end evaluation
@@
   \tag{\PutRL}
-  \frac{\PutRSymOf{(x,c)} = (y,c'')}{\PutLSymOf{(y,c'')} = (x,c')}
+  \frac{\PutRSymOf{(x,c)} = (y,c'')}{\PutLSymOf{(y,c'''')} = (x,c')}
 \end{equation}
@@
 \fi
-
+\section{Forgetful Symmetric Lenses}
+\begin{theorem*}
+  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
+  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
+\end{theorem*}
+The following lemma is
+\begin{proof}
+\begin{enumerate}
+
+\item
+($S \; |_1 \; \varnothing \equiv^S S$)
+
+For all $s \in \mathcal{L}(S \; | \; \varnothing)$, then $P_{S \; |_1 \; \varnothing} = 1 * P_S(s) = P_S(s)$.
+\item
+($S \cdot \varnothing \equiv^S \varnothing$)
+
+For all $s \in \mathcal{L}(S \cdot \varnothing)$, the $P_{S \cdot \varnothing}(s) = \sum_{s_1 \cdot s_2}P_S(s_1) * P_{\varnothing}(s_2)$, but this sum is empty, hence $P_{S \cdot \varnothing}(s) = 0 = P_{\varnothing}(s)$.
+\item
+($\varnothing \cdot S \equiv^S \varnothing$)
+
+Similar to the previous case.
+\item
+($(S \cdot S') \cdot S'' \equiv^S S \cdot (S' \cdot S'')$). Let $s \in \mathcal{L}(S \cdot S' \cdot S'')$
+
+Then
+\begin{align*}
+P_{(S \cdot S') \cdot S''}(s) &= \sum_{s=s_4 \cdot s_3}P_{S \cdot S'}(s_4) * P_{S''}(s_3)\\
+&= \sum_{s=s_4 \cdot s_3}\left(\sum_{s_4=s_1 \cdot s_2} P_{S}(s_1) * P_{S'}(s_2)\right)* P_{S''}(s_3)\\
+&= \sum_{s=s_1\cdot s_2 \cdot s_3} P_{S}(s_1) * P_{S'}(s_2) * P_{S''}(s_3)\\
+P_{S \cdot (S' \cdot S'')}(s) &= \sum_{s=s_1\cdot s_4} P_{S}(s) * \left(\sum_{s_4=s_2\cdot s_3}P_{S'}(s_2) * P_{S''}(s_3)\right)
+\end{align*}
+\item
+($(S \; |_{p_1} \; S') \; |_{p_2} \; S'' \equiv^S S \; |_{p_1 * p_2} \; (S' \; |_{\frac{(1-p_1)*p_2}{1-p_1*p_2}} \; S'')$)
+
+Let $s \in \mathcal{L}(S \; |\; S' \; | \; S'')$. Then
+\begin{align*}
+P_{(S \; |_{p_1} \; S') \; |_{p_2} \; S''}(s) &= p_2 * P_{S \; |_{p_1} \; S'}(s) + (1-p_2)P_{S''}(s)\\
+&= (p_2 * (p_1 * P_S(s) + (1-p_1) * P_{S'}(s))) + (1-p_2)P_{S''}(s)\\
+&= (p_2 * (p_1 * P_S(s) + P_{S'}(s)-p_1* P_{S'}(s))) + (1-p_2)P_{S''}(s)\\
+&= p_2*p_1 * P_S(s) + p_2*P_{S'}(s)-p_2*p_1* P_{S'}(s) + (1-p_2)P_{S''}(s)\\
+&= p_2*p_1 * P_S(s) + (1-p_1) * p_2 * P_{S'}(s)+ (1-p_2)P_{S''}(s)\\
+S \; |_{p_1 * p_2} \; (S' \; |_{\frac{(1-p_1)*p_2}{1-p_1*p_2}} \; S'') &= (p_1p_2)  P_S(s) + (1-p_1p_2)  \left(\frac{(1-p_1)p_2}{1-p_1p_2}  P_{S'}(s) + \left(1-\frac{(1-p_1)p_2}{1-p_1p_2}\right)P_{S''}(s)\right)
+\end{align*}
+\item
+($S \; |_p \; T \equiv^S T \; |_{1-p} \; S$)
+
+For all $s \in \mathcal{L}(S \; | \; T)$, we have
+$$P_{S \; |_p \; T}(s) = p*P_S(s) + (1-p)*P_T(s) = (1-p) * P_T(s) + (1-(1-p))*P_s(s) = P_{T \; |_{1-p} \; S}(s)$$
+\item
+($S \cdot (S' \; |_p \; S'' ) \equiv^S (S \cdot S') \; |_p \; (S \cdot S'')$)
+
+For all $s \in \mathcal{L}(S \cdot (S' \sep S''))$ we have
+\begin{align*}
+P_{S \cdot (S' \; |_p \; S'' )} &= \sum_{s_1 \cdot s_2 = s}P_S(s_1) * P_{S' \; |_p \; S''}(s)\\
+&= \sum_{s_1 \cdot s_2 = s}P_S(s_1) * (p * P_{S'}(s_2) + (1-p) * P_{S''}(s_2))\\
+&= \sum_{s_1 \cdot s_2 = s}p * P_S(s_1) * P_{S'}(s_2) + \sum_{s_1 \cdot s_2 = s}(1-p) * P_S(s_1) * P_{S''}(s_2)\\
+P_{(S \cdot S') \; |_p \; (S \cdot S'')}(s) &= p*P_{S \cdot S'}(s) + (1-p)*P_{S \cdot S''}(s)
+\end{align*}
+\item
+$((S' \; |_p \; S'') \cdot S \equiv^S (S' \cdot S) \; |_p \; (S'' \cdot S))$
+
+For all $s \in \mathcal{L}((S' \; | \; S'') \cdot S)$,  we have
+\begin{align*}
+P_{(S' \; |_p \; S'') \cdot S}(s) &= \sum_{s_1 \cdot s_2 = s}P_{S' \; |_p \; S''}(s_1) * P_S(s_2)\\
+&= \sum_{s_1 \cdot s_2 = s}(p * P_{S'}(s_1) + (1-p) * P_{S''}(s_1))* P_S(s_2)\\
+&= \sum_{s_1 \cdot s_2=s}p*P_{S'}(s_1) * P_S(s_2) + \sum_{s_1 \cdot s_2 = s}(1-p) * P_{S''}(s_1) * P_S(s_2)\\
+P_{(S' \cdot S) \; |_p \; (S'' \cdot S)}&= p * P_{S' \cdot S}(s) + (1-p) * P_{S'' \cdot S}(s)
+\end{align*}
+\item
+($\varepsilon \cdot S \equiv^S S$)
+
+For all $s \in \mathcal{L}(S)$, we have
+$$P_{\epsilon \cdot S}(s) = \sum_{s_1 \cdot s_2 = s}P_{\epsilon}(s_1) * P_S(s_2) = P_{\epsilon}(\epsilon) * P_S(s_2) = P_S(s)$$
+since $s_2 = s$ in the computation above.
+\item
+($S \cdot \epsilon \equiv^S S$)
+
+Similar to the previous case.
+\item
+($S^{*p} \equiv^S \epsilon \; |_{(1-p)} \; (S^{*p} \cdot S)$)
+
+Let $s \in \mathcal{L}(S^*)$, in which case we have
+\begin{align*}
+P_{\epsilon \; |_{1-p} \; (S^{*p} \cdot S)}(s) &= (1-p) * P_{\epsilon}(s) + p * P_{S^{*p} \cdot S}(s)\\
+&= (1-p) * P_{\epsilon}(s) + p * \sum_{a \cdot b = s}\left(\sum_n \sum_{s_1 \ldots \cdot s_n=a} p^n (1-p) \prod_{i=1}^n P_S(s_i)\right) * P_S(b)\\
+&= (1-p) * P_{\epsilon}(s) + \sum_{a \cdot b = s}\left(\sum_n \sum_{s_1 \cdot \ldots \cdot s_n=s} p^{n+1} (1-p) \prod_{i=1}^n P_S(s_i) * P_S(b)\right)\\
+&= (1-p) * P_{\epsilon}(s) + \sum_n \sum_{s_1 \cdot \ldots \cdot s_{n+1}=s} p^{n+1} (1-p) \prod_{i=1}^{n+1} P_S(s_i)\\
+&= (1-p) * P_{\epsilon}(s) + \left(\sum_n \sum_{s_1 \cdot \ldots \cdot s_n = s} p^n * (1-p) * \prod_{i=1}^n P_S(s_i)\right) - (1-p) * (\mathbbm{1}_{s = \varepsilon})
+\end{align*}
+Since
+$$
+P_{S^{*p}}(s) = \sum_n \sum_{s_1 \cdot \ldots \cdot s_n = s} p^n * (1-p) * \prod_{i=1}^n P_S(s_i)
+$$
+then if $s = \epsilon$, the terms $(1-p) * P_{\epsilon}(s)$ and $(1-p) * \mathbbm{1}_{s=\varepsilon}$ cancel each other out.
+Otherwise, if $s \neq \epsilon$, then $(1-p) * P_{\epsilon}(s) = 0 = (1-p) * (\mathbbm{1}_{s = \varepsilon})$ from which the result follows.
+\item
+($S^{*p} \equiv^S \epsilon \; |_{(1-p)} \; (S \cdot S^{*p})$)
+
+Similar to the previous case.
+\end{enumerate}
+\end{proof}
+
+\begin{theorem*}
+  $\ProbabilityOf{\Regex}{\String} = \ProbabilityOf{\ToDNFRegex
+    \Regex}{\String}$ and $\LanguageOf{\Regex} =
+  \LanguageOf{\ToDNFRegexOf{\Regex}}$
+\end{theorem*}
+\begin{proof}
+By \sam{COPY PASTE PROOF FROM BIJECTIVE LENS PAPER} $\LanguageOf{\Regex} =
+  \LanguageOf{\ToDNFRegexOf{\Regex}}$, so we will prove the result $\ProbabilityOf{\Regex}{\String} = \ProbabilityOf{\ToDNFRegex
+    \Regex}{\String}$.
+\begin{enumerate}
+\item
+$(S = s)$
+
+By definition, $\Downarrow s = \langle ([s],1) \rangle$ and $P_{\langle ([s], 1) \rangle}(s') = 
+\begin{cases}
+1 & \text{if } s = s'\\
+0 & \text{otherwise}
+\end{cases}$ hence $P_{\langle ([s],1) \rangle}(s') = P_s(s')$.
+\item
+$(S = \varnothing)$
+
+By definition, $\Downarrow \varnothing = \langle \rangle$ and $P_{\langle \rangle}(s) = 0 = P_{\varnothing}(s)$
+\item
+$(S = S^{*p})$
+
+By definition, $\Downarrow (S^{*p}) = \langle [\epsilon \cdot (\Downarrow S)^{*p} \cdot \epsilon] \rangle $. By the induction hypothesis, $P_{(\Downarrow S)} = P_{S}$, hence
+\begin{align*}
+P_{[\epsilon \cdot (\Downarrow S)^{*p}] \cdot \epsilon}(s) &= P_{(\Downarrow S)^{*p}}(s)\\
+&= \sum_n \sum_{s_1 \cdot \ldots \cdot s_n=s} p^n * (1-p) * \prod_{i=1}^n P_{(\Downarrow S)}(s_i)\\
+&= \sum_n \sum_{s_1 \cdot \ldots \cdot s_n=s} p^n * (1-p) * \prod_{i=1}^n P_{S}(s_i)\\
+&= P_{S^{*p}}(s)
+\end{align*}
+\item
+$(S = S_1 \cdot S_2)$
+
+By definition, $\Downarrow (S_1 \cdot S_2) = (\Downarrow S_1) \odot (\Downarrow S_2)$.
+By the induction hypothesis, $P_{S_1} = P_{(\Downarrow S_1)}$, and $P_{S_2} = P_{(\Downarrow S_2)}$.
+Assume that $\Downarrow S_1 = \langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_m, p_n)\rangle$ and $\Downarrow S_2 = \langle (TQ_1,q_1) \; | \; \ldots \; | \; (TQ_n, q_m) \rangle$.
+Then
+\begin{align*}
+P_{S_1}(s) &= \sum_{i=1}^n p_i P_{SQ_i}(s)\\
+P_{S_1}(s) &= \sum_{j=1}^m q_i P_{TQ_j}(s)\\
+\end{align*}
+hence
+\begin{align*}
+P_{S_1 \cdot S_2}(s) &= \sum_{s_1 \cdot s_2 = s} P_{S_1}(s_1)* P_{S_2}(s_2)\\
+&= \sum_{s_1 \cdot s_2 = s} \left(\sum_{i=1}^n p_i P_{SQ_i}(s_1)\right)* \left(\sum_{j=1}^m q_i P_{TQ_j}(s_2)\right)\\
+&= \sum_{s_1 \cdot s_2 = s} \sum_{i=1}^n \sum_{j=1}^m (p_i * q_j) * (P_{SQ_i}(s_1) *  P_{TQ_j}(s_2))\\
+&= \sum_{i=1}^n \sum_{j=1}^m (p_i * q_j)\sum_{s_1 \cdot s_2 = s} (P_{SQ_i}(s_1) *  P_{TQ_j}(s_2))
+\end{align*}
+By definition,
+$\ConcatDNFOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}{\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}=$\\
+      $\DNFLeft (\ConcatSequenceOf{\Sequence_1}{\SequenceAlt_1},\Probability_1*\ProbabilityAlt_1)\DNFSep \ldots
+      \DNFSep
+      (\ConcatSequenceOf{\Sequence_1}{\SequenceAlt_m},\Probability_1*\ProbabilityAlt_m)\DNFSep
+      \ldots$\\
+      $\DNFSep
+      (\ConcatSequenceOf{\Sequence_n}{\SequenceAlt_1},\Probability_n*\ProbabilityAlt_1)\DNFSep
+      \ldots \DNFSep
+      (\ConcatSequenceOf{\Sequence_n}{\SequenceAlt_m},\Probability_n * \ProbabilityAlt_m) \DNFRight$
+
+where
+$$\ConcatSequenceOf{[\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n]}{[\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\StringAlt_m]}=
+  [\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n\Concat\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\StringAlt_m]$$
+Hence
+$$P_{\langle (SQ_1, p_1) \; | \; \ldots \; | \;(SQ_n,p_n)\rangle \odot \langle (TQ_1, q_1) \; | \; \ldots \; | \;(TQ_m,q_m)\rangle}(s) = \sum_{i=1}^n \sum_{j=1}^m(p_i * q_j) * P_{SQ_i \odot_{SQ} TQ_j}(s)$$
+Since $P_{[s_0 \cdot A_1 \cdot \ldots \cdot A_n \cdot s_n]}(s) = \sum_{s_0 \cdot s'_1 \cdot \ldots \cdot s'_n \cdot s_n = s} \prod_{i=1}^n P_{A_i}(s'i)$, then
+\begin{align*}
+P_{[s_0 \cdot A_1 \cdot \ldots \cdot A_n \cdot s_n] \odot_{SQ} [t_0 \cdot B_1 \cdot \ldots \cdot B_m \cdot t_m]}(s) &= \sum_{s_0 \cdot s'_1 \cdot \ldots \cdot s'_n \cdot s_n \cdot t_0 \cdot t'_1 \cdot \ldots t'_m \cdot t_m =s } \prod_{i=1}^n P_{A_i}(s'_i) \prod_{j=1}^m P_{B_j}(t'_j)\\
+&= \sum_{a \cdot b = s} \sum_{s_0 \cdot s'_1 \cdot \ldots \cdot s'_n \cdot s_n = a} \sum_{t_0 \cdot t'_1 \cdot \ldots \cdot t'_m \cdot t_m = b} \prod_{i=1}^n P_{A_i}(s'_i) \prod_{j=1}^m P_{B_j}(t'_j)\\
+& \sum_{a \cdot b = s} \left(\sum_{s_0 \cdot s'_1 \cdot \ldots \cdot s'_n \cdot s_n = a} \prod_{i=1}^n P_{A_i}(s'_i)\right) \left(\sum_{t_0 \cdot t'_1 \cdot \ldots \cdot t'_m \cdot t_m = b} \prod_{j=1}^m P_{B_j}(t'_j)\right)\\
+&= \sum_{a \cdot b = s}P_{[s_0 \cdot A_1 \cdot \ldots \cdot A_n \cdot s_n]}(a) * P_{[t_0 \cdot B_1 \cdot \ldots \cdot B_m \cdot t_m]}(b)
+\end{align*}
+In other words, $P_{SQ \odot_{SQ} TQ}(s) = \sum_{a \cdot b = s}P_{SQ}(a) * P_{TQ}(b)$. Hence, 
+\begin{align*}
+P_{S_1 \cdot S_2}(s) &= \sum_{i=1}^n \sum_{j=1}^m (p_i * q_j)\sum_{s_1 \cdot s_2 = s} (P_{SQ_i}(s_1) *  P_{TQ_j}(s_2))\\
+&= \sum_{i=1}^n \sum_{j=1}^m (p_i * q_j)\sum_{s_1 \cdot s_2 = s} P_{SQ_i \odot_{SQ} TQ_j}(s)\\
+&= P_{\langle (SQ_1, p_1) \; | \; \ldots \; | \;(SQ_n,p_n)\rangle \odot \langle (TQ_1, q_1) \; | \; \ldots \; | \;(TQ_m,q_m)\rangle}(s)\\
+&= P_{(\Downarrow S_1) \odot (\Downarrow S_2)}(s)
+\end{align*}
+which is what we wanted to show.
+\item
+$(S = S_1 \; |_p \; S_2)$
+
+By definition $\Downarrow (S_1 \; |_p \; S_2) = (\Downarrow S_1) \oplus_p (\Downarrow S_2)$.
+By the induction hypothesis $P_{S_1} = P_{\Downarrow S_1}$ and $P_{S_2} = P_{\Downarrow S_2}$.
+Assume that $\Downarrow S_1 = \langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_m, p_n)\rangle$ and $\Downarrow S_2 = \langle (TQ_1,q_1) \; | \; \ldots \; | \; (TQ_n, q_m) \rangle$.
+By definition,
+
+  $\OrDNFOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}{\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}{\Probability} =$\\
+  $\DNFOf{(\Sequence_1,\Probability_1*\Probability)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n*\Probability)\DNFSep(\SequenceAlt_1,\ProbabilityAlt_1*(1-\Probability))\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m*(1-\Probability))}$
+  hence
+  \begin{align*}
+  P_{(\Downarrow S_1) \; |_p \; (\Downarrow S_2)}(s) &= P_{\langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_n, p_n)\rangle \oplus_p \langle (TQ1, q_1) \; | \; \ldots \; | \; (TQ_m, q_m)\rangle}(s) \\
+  &= P_{\langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_n, p_n) \; | \; (TQ1, q_1) \; | \; \ldots \; | \; (TQ_m, q_m)\rangle}(s)\\
+  &= \sum_{i=1}^m (p * p_i) P_{SQ_i}(s) + \sum_{j=1}^m ((1-p) * q_j) P_{TQ_j}(s)\\
+  &= p * \sum_{i=1}^m p_i P_{SQ_i}(s) + (1-p)* \sum_{j=1}^m q_j * P_{TQ_j}(s)\\
+&= p * P_{\langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_n, p_n)\rangle}(s) + (1-p) * P_{\langle (TQ1, q_1) \; | \; \ldots \; | \; (TQ_m, q_m)\rangle}(s)\\
+&= p * P_{\Downarrow S_1}(s) + (1-p) * P_{\Downarrow S_2}(s) 
+  \end{align*}
+
+  Since $P_{S_1} = P_{\Downarrow S_1}$ and $P_{S_1} = P_{\Downarrow S_1}$, then 
+  \begin{align*}
+  P_{S_1}(s) &= \sum_{i=1}^n p_i * P_{SQ_i}(s)\\
+  P_{S_2}(s) &= \sum_{j=1}^m q_m * P_{TQ_j}(s)
+  \end{align*}
+  hence
+  \begin{align*}
+  P_{S_1 \; |_p \; S_2}(s) &= p * (P_{S_1}(s)) + (1-p) * (P_{S_2}(s))\\
+  &= p \left(\sum_{i=1}^n p_i * P_{SQ_i}(s)\right) + (1-p)\left(\sum_{j=1}^m q_m * P_{TQ_j}(s)\right)\\
+  &= p * P_{\Downarrow S_1}(s) + (1-p) * P_{\Downarrow S_2}(s)\\
+  &= P_{(\Downarrow S_1) \; |_p \; (\Downarrow S_2)}(s)
+  \end{align*}
+  which is what we wanted to show. This completes the proof.
+\end{enumerate}
+\end{proof}
+\begin{theorem*}
+  \label{thm:correct_entropy}
+  If $\Regex$ is unambiguous and contains no empty subcomponents,
+  $\EntropyOf{\Regex}$ is the entropy of $\mathcal{L}(\Regex)$.
+\end{theorem*}
+\begin{proof}
+Given a discrete set $X$ and a probability distribution $P$ on $X$, we define the entropy $H(X)$ of $X$ by
+$$H(X) = -\sum_{x \in X}P(x)\log_2{P(x)}$$
+We prove the theorem by induction on $S$.
+\begin{enumerate}
+\item
+($S = s$)
+
+Then $\mathbb{H}(s) = 0 = H(\mathcal{L}(s)) = H(\{s\})$
+\item
+$(S = S^{*p})$
+
+Since $S^{*p}$ is unambiguous, then $S$ is also anambiguous, and for each $s \in \mathcal{L}(S^{*p})$, there exist unique $t_1, \ldots, t_n \in \mathcal{L}(S)$ such that $s = t^s_1 \cdot \ldots \cdot t^s_{n_s}$.
+By the induction hypothesis, $H(\mathcal{L}(S)) = \mathbb{H}(S)$, thus
+\begin{align*}
+H(\mathcal{L}(S^{*p})) &= - \sum_{s \in \mathcal{L}(S^{*p})} P_{S^{*p}}(s) \log_2 P_{S^{*p}}(s)\\
+&= -(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)\log_2 \left(p^n * (1 - p) * \prod_{i=1}^n P_S(s_i)\right)\\
+&= -(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)\left(n \log_2 p  + \log_2 (1 - p) + \log_2 \left(\prod_{i=1}^n P_S(s_i)\right)\right)\\
+&= -(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)\left(n \log_2 p  + \log_2 (1 - p) + \log_2 \left(\prod_{i=1}^n P_S(s_i)\right)\right)\\
+&= -(1-p) * \log_2 p \sum_n n * p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)
+-(1-p) * \log_2 (1 - p) \sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i) \\
+&-(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)\log_2 \left(\prod_{i=1}^n P_S(s_i)\right)\\
+&= -\log_2 p * \frac{p}{(1-p)} -\log_2 (1 - p) 
+-(1-p)\sum_n p^n H(\mathcal{L}(S^n))\\
+&= -\log_2 p * \frac{p}{(1-p)} -\log_2 (1 - p) 
+-(1-p)\sum_n p^n n H(\mathcal{L}(S))\\
+&= -\log_2 p * \frac{p}{(1-p)} -\log_2 (1 - p) 
+-\frac{p}{(1-p)} * H(\mathcal{L}(S))\\
+&= \frac{p}{(1-p)}(H(\mathcal{L}(S)) - \log_2 p) -\log_2 (1 - p)\\
+&= \frac{p}{(1-p)}(\mathbb{H}(S) - \log_2 p) -\log_2 (1 - p)
+\end{align*}
+\item
+($S = S_1 \cdot S_2$)
+
+Since $S$ is unambigous, then $S_1$ and $S_2$ are also unambiguous, and for each $s \in \mathcal{L}(S_1 \cdot S_2)$ there exists a unique $s_1 \in \mathcal{L}(S_1)$ and a unique $s_2 \in \mathcal{L}(S_2)$ such that $s = s_1 \cdot s_2$.
+By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$, hence
+\begin{align*}
+H(\mathcal{L}(S_1 \cdot S_2))(s) &= -\sum_{s_1 \in S_1} \sum_{s_2 \in S_2}P_{S_1}(s_1)*P_{s_2}(s_2) \log_2(P_{S_1}(s_1)*P_{S_2}(s_2))\\
+&= -\sum_{s_1 \in S_1} P_{S_1}(s_1)\sum_{s_2 \in S_2}P_{s_2}(s_2) (\log_2(P_{S_1}(s_1) + \log_2 P_{S_2}(s_2)))\\
+&= -\sum_{s_1 \in S_1} P_{S_1}(s_1) \sum_{s_2 \in S_2}P_{S_2}(s_2)\log_2(P_{S_1}(s_1)) + \sum_{s_1 \in S_1} P_{S_1}(s_1)\sum_{s_2 \in S_2} P_{S_2}(s_2)\log_2 P_{S_2}(s_2)\\
+&= -\sum_{s_1 \in S_1} P_{S_1}(s_1) \log_2(P_{S_1}(s_1)) + \sum_{s_2 \in S_2} P_{S_2}(s_2)\log_2 P_{S_2}(s_2)\\
+&= H(\mathcal{L}(S_1)) + H(\mathcal{L}(S_2))\\
+&= \mathbb{H}(S_1) + \mathbb{H}(S_2)
+\end{align*}
+\item
+($S = S_1 \; |_p \; S_2$)
+
+Since $S$ is unambiguous, then $S_1$ and $S_2$ are also unambiguous, and every $s \in \mathcal{L}(S_1 \; |_p \; S_2)$ is either in $\mathcal{L}(S_1)$ or in $\mathcal{L}(S_2)$.
+By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$.
+Then
+\begin{align*}
+H(\mathcal{L}(S_1 \; |_p \; S_2)) &= -\sum_{s \in \mathcal{L}(S_1)}(p * P_{S_1}(s))\log_2 (p * P_{S_1}(s)) - \sum_{s \in \mathcal{L}(S_2)}((1-p) * P_{S_2}(s))\log_2 ((1-p) * P_{S_2}(s))\\
+&=- p\sum_{s \in \mathcal{L}(S_1)}P_{S_1}(s)(\log_2 p + \log_2 P_{S_1}(s)) - (1-p)\sum_{s \in \mathcal{L}(S_2)}P_{S_2}(s)(\log_2 (1-p) + \log_2 P_{S_2}(s))\\
+&= p * (\log_2 p +  H(\mathcal{L}(S_1))) + (1-p) * ( \log_2 (1-p) + H(\mathcal{L}(S_2)))\\
+&= p * (\log_2 p +  \mathbb{H}(S_1)) + (1-p) * ( \log_2 (1-p) + \mathbb{H}(S_2))\\
+\end{align*}
+\end{enumerate}
+\end{proof}
+\begin{theorem*}
+  Let $\Lens$ be an asymmetric lens. $\Lens$ is also a simple symmetric lens,
+  where
+  \begin{center}
+    \begin{tabular}{rcl}
+      $\Lens.\CreateROf{x}$ & $=$ & $\Lens.get \App x$\\
+      $\Lens.\CreateLOf{y}$ & $=$ & $\Lens.create \App y$\\
+      $\Lens.\PutROf{x}{y}$ & $=$ & $\Lens.get \App x$\\
+      $\Lens.\PutLOf{y}{x}$ & $=$ & $\Lens.put \App y \App x$
+    \end{tabular}
+  \end{center}
+\end{theorem*}
+\begin{proof}
+Let $\ell$ be an asymmetric lens. Then $\ell$ satisfies the following laws:
+\begin{align*}
+\ell.\get \; (\ell.\pput \; v \; s) &= v \tag{PUTGET}\\
+\ell.\pput \; (\ell.\get \; s) \; s&= s \tag{GETPUT}\\
+\ell.\get \; (\ell.\pput \; v \; s) &= v \tag{CREATGET}
+\end{align*}
+Define the simple symmetric lens $\lceil \ell \rceil$ by 
+\begin{align*}
+\lceil \ell \rceil.\CreateROf{s} &= \ell.\get \; s\\
+\lceil \ell \rceil.\CreateLOf{v} &= \ell.\create \; v\\
+\lceil \ell \rceil.\PutROf{s}{v} &= \ell.\get \; s\\
+\lceil \ell \rceil.\PutLOf{v}{s} &= \ell.\pput \; v \; s\\
+\end{align*}
+We now show that $\lceil \ell \rceil$ satisfies the simple symmetric lens laws:
+\begin{align*}
+\lceil \ell \rceil.\PutLOf{(\lceil \ell \rceil.\CreateROf{s})}{s} &= \ell.\pput \; (\ell.\get \; s) \; s = s\\
+\lceil \ell \rceil.\PutROf{(\lceil \ell \rceil.\CreateLOf{v})}{v} &= \ell.\get \; (\ell.\create \; v) = v\\
+\lceil \ell \rceil.\PutLOf{(\lceil \ell \rceil.\PutROf{s}{v})}{s} &= \ell.\pput \; (\ell.\get) \; s = s\\
+\lceil \ell \rceil.\PutROf{(\lceil \ell \rceil.\PutLOf{v}{s})}{v} &= \ell.\get \; (\ell.\pput \; v \; s) = v
+\end{align*}
+\end{proof}
+
+\begin{theorem*}
+  Let $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$, where $\Lens$ does not
+  include composition, $\Regex$ and $\RegexAlt$ and unambiguous, and neither
+  $\Regex$ nor $\RegexAlt$ contain any empty subcomponents.
+  \begin{enumerate}
+  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ calculates the expected
+    information content of $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$,
+    given $\SetOf{s \SuchThat s \in \LanguageOf{\Regex} \BooleanAnd
+      \Lens.\PutROf{s}{t} = t}$
+  \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ calculates the expected
+    information content of $\SetOf{s \SuchThat s \in \LanguageOf{\Regex}}$,
+    given $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt} \BooleanAnd
+      \Lens.\PutLOf{t}{s} = s}$
+  \end{enumerate}
+\end{theorem*}
+\begin{proof}
+Given $s \in S$, let $V_{\ell}(s) = \{t \in \mathcal{L}(T) \; | \; \ell.\PutROf{s}{t} = t\}$. We want to show that
+$$\sum_{s \in \mathcal{L}(S)}P_{\mathcal{L}(S)}(s)H(V_{\ell}(s)) = \mathbb{H}^{\rightarrow}(S \; | \; \ell, T)$$
+We proceed by induction over $\ell$. 
+\begin{enumerate}
+\item
+($\ell = \IdentityLensOf{S}$)
+
+Observe that $V_{\IdentityLensOf{S}}(s) = \{t \in \mathcal{L}(S) \; | \; \IdentityLensOf{S}.\ell.\PutROf{s}{t} = t\} = \{t \in \mathcal{L}(S) \; | \; t = s\} = \{s\}$.
+
+Consequently, $H(V_{\IdentityLensOf{S}}(s)) = H(\{s\}) = 0$, from which it follows that 
+$$\sum_{s \in \mathcal{L}(S)} P_{\mathcal{L}(S)}(s) * H(V_{\IdentityLensOf{S}}(s)) = 0 = \mathbb{H}^{\rightarrow}(S \; | \; \IdentityLensOf{S}, T)$$
+\item
+($\ell = \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}$)
+
+Observe that 
+$$V_{\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}}(s) = \{t \in \mathcal{L}(T) \; | \; \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}.\PutROf{s}{t} = t\} = \{t \in \mathcal{L}(T) \; | \; t = t\} = \mathcal{L}(T)$$
+Consequently
+$$\sum_{s \in \mathcal{L}(S)}P_{S}(s)H(V_{\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}}(s)) = \sum_{s \in \mathcal{L}(S)}P_{S}(s)H(\mathcal{L}(T)) = H(\mathcal{L}(T)) = \mathbb{H}(T) = \mathbb{H}^{\rightarrow}(S \; | \; \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, T)$$
+\item
+($\ell = \IterateLensOf{\Lens} : S^{*p} \Leftrightarrow T^{*q}$)
+
+Observe that 
+\begin{align*}
+V_{\IterateLensOf{\Lens}}(s_1 \cdot \ldots \cdot s_n) &= \{t_1 \cdot \ldots \cdot t_m \; | \; \IterateLensOf{\Lens}.\PutROf{(s_1 \cdot \ldots \cdot s_n) }{(t_1 \cdot \ldots \cdot t_m)} = t_1 \cdot \ldots \cdot t_m\}\\
+&= \{t_1 \cdot \ldots \cdot t_n \; | \; \ell.\PutROf{s_i}{t_i} = t_i \text{ for }1 \leq i \leq n\}\\
+&= V_{\ell}(s_1) \cdot \ldots \cdot V_{\ell}(s_n)
+\end{align*}
+Consequently
+\begin{align*}
+\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V_{\IterateLensOf{\Lens}}(s))
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n)\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i) H(V_{\ell}(s_1) \cdot \ldots \cdot V_{\ell}(s_n))\\
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n)\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i) \left(\sum_{j=1}^n H(V_{\ell}(s_j)) \right)\\
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \sum_{s_n} P_S(s_n) \left(\sum_{j=1}^n H(V_{\ell}(s_j)) \right)
+\end{align*}
+We focus on the term $\sum_{s_n} P_S(s_n) \sum_{j=1}^n H(V_{\ell}(s_j))$:
+\begin{align*}
+\sum_{s_n} P_S(s_n) \sum_{j=1}^n H(V_{\ell}(s_j)) &= \sum_{s_n} P_S(s_n)H(V_{\ell}(s_n)) + P_S(s_n) \sum_{j=1}^{n-1} H(V_{\ell}(s_j))\\
+&= \mathbb{H}^{\rightarrow}(S \; | \; \ell, T) + \sum_{j=1}^{n-1}H(V_{\ell}(s_j))
+\end{align*}
+Consequently,
+\begin{align*}
+\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V_{\IterateLensOf{\Lens}}(s))
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left( \mathbb{H}^{\rightarrow}(S \; | \; \ell, T) + \sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)\\
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1})\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \mathbb{H}^{\rightarrow}(S \; | \; \ell, T)\\
+&+(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)\\
+&= \mathbb{H}^{\rightarrow}(S \; | \; \ell, T) * (1-p)\sum_n p^n \\
+&+(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)
+\end{align*}
+Unrolling the term 
+$$\sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)$$ 
+$(n-1)$ more times gives
+$$\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V_{\IterateLensOf{\Lens}}(s))
+= \mathbb{H}^{\rightarrow}(S \; | \; \ell, T) * (1-p)\sum_n n * p^n  = \frac{p}{1-p} * \mathbb{H}^{\rightarrow}(S \; | \; \ell, T)$$
+which is what we wanted to show.
+\item
+($S = \ConcatLensOf{\Lens_1}{\Lens_2}$)
+
+Observe that 
+\begin{align*}
+V_{ \ConcatLensOf{\Lens_1}{\Lens_2}}(s_1 \cdot s_2) &= \{t_1 \cdot t_2 \in \mathcal{L}(T_1 \cdot T_2) \; | \;  \ConcatLensOf{\Lens_1}{\Lens_2}.\PutROf{(s_1 \cdot s_2)}{(t_1 \cdot t_2)} = t_1 \cdot t_2\}\\
+&= \{t_1 \cdot t_2 \in \mathcal{L}(T_1 \cdot T_2) \; | \;  \ell_1.\PutROf{s_1}{t_1 } = t_1  \text{ and } \ell_2.\PutROf{t_2}{t_2} = t_2\}\\
+&= V_{\ell_1}(s_1) \cdot V_{\ell_2}(s_2)
+\end{align*}
+Consequently,
+\begin{align*}
+&\sum_{s \in \mathcal{L}(S_1 \cdot S_2)}P_{S_1 \cdot S_2}(s) H(V_{\ConcatLensOf{\Lens_1}{\Lens_2}}(s))\\
+&= \sum_{s_1 \in \mathcal{L}(s_1)} P_{S_1}(s_1)\sum_{s_2 \in \mathcal{L}(s_2)}P_{S_2}(s_2)(H(V_{\ell_1}(s_1)) + H(V_{\ell_2}(s_2)))\\
+&= \sum_{s_1 \in \mathcal{L}(s_1)} P_{S_1}(s_1)\sum_{s_2 \in \mathcal{L}(s_2)}P_{S_2}(s_2)H(V_{\ell_1}(s_1)) + \sum_{s_1 \in \mathcal{L}(s_1)} P_{S_1}(s_1)\sum_{s_2 \in \mathcal{L}(s_2)}P_{S_2}(s_2)H(V_{\ell_2}(s_2))\\
+&= \mathbb{H}^{\rightarrow}(S_1 \; | \; \ell_1, T_1) + \mathbb{H}^{\rightarrow}(S_2 \; | \; \ell_2, T_2)
+\end{align*}
+\item
+($S = \SwapLensOf{\Lens_1}{\Lens_2}$)
+
+Similar to the previous case.
+\item
+($S = \OrLensOf{\Lens_1}{\Lens_2}$)
+
+Observe that 
+$$
+V_{ \OrLensOf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \OrLensOf{\Lens_1}{\Lens_2}.\ell.\PutROf{s}{t} = t\}
+$$
+If $s \in \mathcal{L}(S_1)$, then
+$$
+V_{ \OrLensOf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \ell_1.\PutROf{s}{t} = t\} = V_{\ell_1}(s)
+$$
+while if $s \in \mathcal{L}(S_2)$, then
+$$
+V_{ \OrLensOf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \ell_2.\PutROf{s}{t} = t\} = V_{\ell_2}(s)
+$$
+Consequently,
+\begin{align*}
+&\sum_{s \in \mathcal{L}(S_1 \; |_p \; S_2)}P_{S_1 \; |_p \; S_2}(s) H(V_{\OrLensOf{\Lens_1}{\Lens_2}}(s))\\
+&= p * \sum_{s \in \mathcal{L}(S_1)}P_{S_1}(s) H(V_{\ell_1}(s)) + (1-p) * \sum_{s \in \mathcal{L}(S_2)}P_{S_2}(s) H(V_{\ell_2}(s))\\
+&= p * (\mathbb{H}^{\rightarrow}(S_1 \; | \; \ell_1,T_1)) + (1-p) * (\mathbb{H}^{\rightarrow}(S_2 \; | \; \ell_2,T_2))
+\end{align*}
+\item
+($S = \MergeROf{\Lens_1}{\Lens_2}$)
+
+Observe that 
+$$
+V_{ \MergeROf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T) \; | \;  \MergeROf{\Lens_1}{\Lens_2}.\PutROf{s}{t} = t\}
+$$
+If $s \in \mathcal{L}(S_1)$, then
+$$
+V_{ \MergeROf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T) \; | \;  \ell_1.\PutROf{s}{t} = t\} = V_{\ell_1}(s)
+$$
+while if $s \in \mathcal{L}(S_2)$, then
+$$
+V_{ \MergeROf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T) \; | \;  \ell_2.\PutROf{s}{t} = t\} = V_{\ell_2}(s)
+$$
+Consequently,
+\begin{align*}
+&\sum_{s \in \mathcal{L}(S_1 \; |_p \; S_2)}P_{S_1 \; |_p \; S_2}(s) H(V_{\MergeROf{\Lens_1}{\Lens_2}}(s))\\
+&= p * \sum_{s \in \mathcal{L}(S_1)}P_{S_1}(s) H(V_{\ell_1}(s)) + (1-p) * \sum_{s \in \mathcal{L}(S_2)}P_{S_2}(s) H(V_{\ell_2}(s))\\
+&= p * (\mathbb{H}^{\rightarrow}(S_1 \; | \; \ell_1,T_1) ) + (1-p) * (\mathbb{H}^{\rightarrow}(S_2 \; | \; \ell_2,T_2) )
+\end{align*}
+\item
+($S = \MergeLOf{\Lens_1}{\Lens_2}$)
+
+Observe that 
+\begin{align*}
+V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s) &= \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \MergeLOf{\Lens_1}{\Lens_2}.\PutROf{s}{t} = t\}\\
+&= \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \ell_1.\PutROf{s}{t} = t \text{ or } \ell_2.\PutROf{s}{t} = t\}\\
+&= V_{\ell_1}(s) \cup V_{\ell_2}(s)
+\end{align*}
+First we consider $H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))$:
+\begin{align*}
+-H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)) &= \sum_{t \in V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}(t) \log_2 P_{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}(t)\\
+&= \sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \\
+&+ \sum_{t \in \mathcal{L}(T_2) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+\end{align*}
+We focus on the term 
+$$\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}$$
+Observe that
+\begin{align*}
+&\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
+&= \frac{q}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t) \log_2 q * P_{T_1}(t)\\
+&- \frac{q * \log_2 P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t)
+\end{align*}
+Again we focus on the term
+$$\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t) \log_2 q * P_{T_1}(t)$$
+Observe that
+\begin{align*}
+&\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t) \log_2 q * P_{T_1}(t) = \log_2 q * P_{T_1}(V_{\ell_1}(s))\\
+& + P_{T_1}(V_{\ell_1}(s))\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{P_{T_1}(t)}{P_{T_1}(V_{\ell_1}(s))} \log_2 \left( P_{T_1}(V_{\ell_1}(s)) *  \frac{P_{T_1}(t)}{P_{T_1}(V_{\ell_1}(s))} \right)\\
+&= \log_2 q * P_{T_1}(V_{\ell_1}(s)) + \log_2 P_{T_1}(V_{\ell_1}(s)) * P_{T_1}(V_{\ell_1}(s)) + P_{T_1}(V_{\ell_1}(s))* -H(V_{\ell_1}(s))\\
+&= P_{T_1}(V_{\ell_1}(s))[\log_2 q + \log_2 P_{T_1}(V_{\ell_1}(s)) - H(V_{\ell_1}(s))]
+\end{align*}
+Hence
+\begin{align*}
+&\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \\
+&= \frac{q}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t) \log_2 q * P_{T_1}(t)\\
+&- \frac{q * \log_2 P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t)\\
+&= \frac{q * P_{T_1}(V_{\ell_1}(s))[\log_2 q + \log_2 P_{T_1}(V_{\ell_1}(s)) - H(V_{\ell_1}(s))]}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
+&- \frac{q * \log_2 P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)) * P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
+&= \frac{q * P_{T_1}(V_{\ell_1}(s)) \left[\log_2 q + \log_2 \left[\frac{P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right]  -H(V_{\ell_1}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+\end{align*}
+Similarly,
+\begin{align*}
+&\sum_{t \in \mathcal{L}(T_2) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
+&= \frac{(1-q) * P_{T_2}(V_{\ell_2}(s)) \left[\log_2 (1-q) + \log_2 \left[\frac{P_{T_2}(V_{\ell_2}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] - H(V_{\ell_2}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+\end{align*}
+\begin{align*}
+&-H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))\\
+&= \frac{q * P_{T_1}(V_{\ell_1}(s)) \left[\log_2 q + \log_2 \left[\frac{P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] - H(V_{\ell_1}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
+&+ \frac{(1-q) * P_{T_2}(V_{\ell_2}(s)) \left[\log_2 (1-q) + \log_2 \left[\frac{P_{T_2}(V_{\ell_2}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] - H(V_{\ell_2}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+\end{align*}
+\end{enumerate}
+
+\end{proof}
 \end{document}

### 2018-07-03 papers/SymmetricLenses/main.tex
Added proof of Theorem 1

--- 2018-07-02 papers/SymmetricLenses/main.tex
@@
 \fi
-
+\section{Forgetful Symmetric Lenses}
+\begin{theorem*}
+  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
+  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
+\end{theorem*}
+The following lemma is
+\begin{proof}
+\begin{enumerate}
+
+\item
+($S \; |_1 \; \varnothing \equiv^S S$)
+
+For all $s \in \mathcal{L}(S \; | \; \varnothing)$, then $P_{S \; |_1 \; \varnothing} = 1 * P_S(s) = P_S(s)$.
+\item
+($S \cdot \varnothing \equiv^S \varnothing$)
+
+For all $s \in \mathcal{L}(S \cdot \varnothing)$, the $P_{S \cdot \varnothing}(s) = \sum_{s_1 \cdot s_2}P_S(s_1) * P_{\varnothing}(s_2)$, but this sum is empty, hence $P_{S \cdot \varnothing}(s) = 0 = P_{\varnothing}(s)$.
+\item
+($\varnothing \cdot S \equiv^S \varnothing$)
+
+Similar to the previous case.
+\item
+($(S \cdot S') \cdot S'' \equiv^S S \cdot (S' \cdot S'')$). Let $s \in \mathcal{L}(S \cdot S' \cdot S'')$
+
+Then
+\begin{align*}
+P_{(S \cdot S') \cdot S''}(s) &= \sum_{s=s_4 \cdot s_3}P_{S \cdot S'}(s_4) * P_{S''}(s_3)\\
+&= \sum_{s=s_4 \cdot s_3}\left(\sum_{s_4=s_1 \cdot s_2} P_{S}(s_1) * P_{S'}(s_2)\right)* P_{S''}(s_3)\\
+&= \sum_{s=s_1\cdot s_2 \cdot s_3} P_{S}(s_1) * P_{S'}(s_2) * P_{S''}(s_3)\\
+P_{S \cdot (S' \cdot S'')}(s) &= \sum_{s=s_1\cdot s_4} P_{S}(s) * \left(\sum_{s_4=s_2\cdot s_3}P_{S'}(s_2) * P_{S''}(s_3)\right)
+\end{align*}
+\item
+($(S \; |_{p_1} \; S') \; |_{p_2} \; S'' \equiv^S S \; |_{p_1 * p_2} \; (S' \; |_{\frac{(1-p_1)*p_2}{1-p_1*p_2}} \; S'')$)
+
+Let $s \in \mathcal{L}(S \; |\; S' \; | \; S'')$. Then
+\begin{align*}
+P_{(S \; |_{p_1} \; S') \; |_{p_2} \; S''}(s) &= p_2 * P_{S \; |_{p_1} \; S'}(s) + (1-p_2)P_{S''}(s)\\
+&= (p_2 * (p_1 * P_S(s) + (1-p_1) * P_{S'}(s))) + (1-p_2)P_{S''}(s)\\
+&= (p_2 * (p_1 * P_S(s) + P_{S'}(s)-p_1* P_{S'}(s))) + (1-p_2)P_{S''}(s)\\
+&= p_2*p_1 * P_S(s) + p_2*P_{S'}(s)-p_2*p_1* P_{S'}(s) + (1-p_2)P_{S''}(s)\\
+&= p_2*p_1 * P_S(s) + (1-p_1) * p_2 * P_{S'}(s)+ (1-p_2)P_{S''}(s)\\
+S \; |_{p_1 * p_2} \; (S' \; |_{\frac{(1-p_1)*p_2}{1-p_1*p_2}} \; S'') &= (p_1p_2)  P_S(s) + (1-p_1p_2)  \left(\frac{(1-p_1)p_2}{1-p_1p_2}  P_{S'}(s) + \left(1-\frac{(1-p_1)p_2}{1-p_1p_2}\right)P_{S''}(s)\right)
+\end{align*}
+\item
+($S \; |_p \; T \equiv^S T \; |_{1-p} \; S$)
+
+For all $s \in \mathcal{L}(S \; | \; T)$, we have
+$$P_{S \; |_p \; T}(s) = p*P_S(s) + (1-p)*P_T(s) = (1-p) * P_T(s) + (1-(1-p))*P_s(s) = P_{T \; |_{1-p} \; S}(s)$$
+\item
+($S \cdot (S' \; |_p \; S'' ) \equiv^S (S \cdot S') \; |_p \; (S \cdot S'')$)
+
+For all $s \in \mathcal{L}(S \cdot (S' \sep S''))$ we have
+\begin{align*}
+P_{S \cdot (S' \; |_p \; S'' )} &= \sum_{s_1 \cdot s_2 = s}P_S(s_1) * P_{S' \; |_p \; S''}(s)\\
+&= \sum_{s_1 \cdot s_2 = s}P_S(s_1) * (p * P_{S'}(s_2) + (1-p) * P_{S''}(s_2))\\
+&= \sum_{s_1 \cdot s_2 = s}p * P_S(s_1) * P_{S'}(s_2) + \sum_{s_1 \cdot s_2 = s}(1-p) * P_S(s_1) * P_{S''}(s_2)\\
+P_{(S \cdot S') \; |_p \; (S \cdot S'')}(s) &= p*P_{S \cdot S'}(s) + (1-p)*P_{S \cdot S''}(s)
+\end{align*}
+\item
+$((S' \; |_p \; S'') \cdot S \equiv^S (S' \cdot S) \; |_p \; (S'' \cdot S))$
+
+For all $s \in \mathcal{L}((S' \; | \; S'') \cdot S)$,  we have
+\begin{align*}
+P_{(S' \; |_p \; S'') \cdot S}(s) &= \sum_{s_1 \cdot s_2 = s}P_{S' \; |_p \; S''}(s_1) * P_S(s_2)\\
+&= \sum_{s_1 \cdot s_2 = s}(p * P_{S'}(s_1) + (1-p) * P_{S''}(s_1))* P_S(s_2)\\
+&= \sum_{s_1 \cdot s_2=s}p*P_{S'}(s_1) * P_S(s_2) + \sum_{s_1 \cdot s_2 = s}(1-p) * P_{S''}(s_1) * P_S(s_2)\\
+P_{(S' \cdot S) \; |_p \; (S'' \cdot S)}&= p * P_{S' \cdot S}(s) + (1-p) * P_{S'' \cdot S}(s)
+\end{align*}
+\item
+($\varepsilon \cdot S \equiv^S S$)
+
+For all $s \in \mathcal{L}(S)$, we have
+$$P_{\epsilon \cdot S}(s) = \sum_{s_1 \cdot s_2 = s}P_{\epsilon}(s_1) * P_S(s_2) = P_{\epsilon}(\epsilon) * P_S(s_2) = P_S(s)$$
+since $s_2 = s$ in the computation above.
+\item
+($S \cdot \epsilon \equiv^S S$)
+
+Similar to the previous case.
+\item
+($S^{*p} \equiv^S \epsilon \; |_{(1-p)} \; (S^{*p} \cdot S)$)
+
+Let $s \in \mathcal{L}(S^*)$, in which case we have
+\begin{align*}
+P_{\epsilon \; |_{1-p} \; (S^{*p} \cdot S)}(s) &= (1-p) * P_{\epsilon}(s) + p * P_{S^{*p} \cdot S}(s)\\
+&= (1-p) * P_{\epsilon}(s) + p * \sum_{a \cdot b = s}\left(\sum_n \sum_{s_1 \ldots \cdot s_n=a} p^n (1-p) \prod_{i=1}^n P_S(s_i)\right) * P_S(b)\\
+&= (1-p) * P_{\epsilon}(s) + \sum_{a \cdot b = s}\left(\sum_n \sum_{s_1 \cdot \ldots \cdot s_n=s} p^{n+1} (1-p) \prod_{i=1}^n P_S(s_i) * P_S(b)\right)\\
+&= (1-p) * P_{\epsilon}(s) + \sum_n \sum_{s_1 \cdot \ldots \cdot s_{n+1}=s} p^{n+1} (1-p) \prod_{i=1}^{n+1} P_S(s_i)\\
+&= (1-p) * P_{\epsilon}(s) + \left(\sum_n \sum_{s_1 \cdot \ldots \cdot s_n = s} p^n * (1-p) * \prod_{i=1}^n P_S(s_i)\right) - (1-p) * (\mathbbm{1}_{s = \varepsilon})
+\end{align*}
+Since
+$$
+P_{S^{*p}}(s) = \sum_n \sum_{s_1 \cdot \ldots \cdot s_n = s} p^n * (1-p) * \prod_{i=1}^n P_S(s_i)
+$$
+then if $s = \epsilon$, the terms $(1-p) * P_{\epsilon}(s)$ and $(1-p) * \mathbbm{1}_{s=\varepsilon}$ cancel each other out.
+Otherwise, if $s \neq \epsilon$, then $(1-p) * P_{\epsilon}(s) = 0 = (1-p) * (\mathbbm{1}_{s = \varepsilon})$ from which the result follows.
+\item
+($S^{*p} \equiv^S \epsilon \; |_{(1-p)} \; (S \cdot S^{*p})$)
+
+Similar to the previous case.
+\end{enumerate}
+\end{proof}
 \end{document}

### 2018-07-03 papers/SymmetricLenses/main.tex
More comments

--- 2018-07-03 papers/SymmetricLenses/main.tex
@@
 \end{enumerate}
+\end{proof}
+
+\begin{theorem*}
+  $\ProbabilityOf{\Regex}{\String} = \ProbabilityOf{\ToDNFRegex
+    \Regex}{\String}$ and $\LanguageOf{\Regex} =
+  \LanguageOf{\ToDNFRegexOf{\Regex}}$
+\end{theorem*}
+\begin{proof}
+By \sam{COPY PASTE PROOF FROM BIJECTIVE LENS PAPER} $\LanguageOf{\Regex} =
+  \LanguageOf{\ToDNFRegexOf{\Regex}}$, so we will prove the result $\ProbabilityOf{\Regex}{\String} = \ProbabilityOf{\ToDNFRegex
+    \Regex}{\String}$.
+\begin{enumerate}
+\item
+$(S = s)$
+
+By definition, $\Downarrow s = \langle ([s],1) \rangle$ and $P_{\langle ([s], 1) \rangle}(s') = 
+\begin{cases}
+1 & \text{if } s = s'\\
+0 & \text{otherwise}
+\end{cases}$ hence $P_{\langle ([s],1) \rangle}(s') = P_s(s')$.
+\item
+$(S = \varnothing)$
+
+By definition, $\Downarrow \varnothing = \langle \rangle$ and $P_{\langle \rangle}(s) = 0 = P_{\varnothing}(s)$
+\item
+$(S = S^{*p})$
+
+By definition, $\Downarrow (S^{*p}) = \langle [\epsilon \cdot (\Downarrow S)^{*p} \cdot \epsilon] \rangle $. By the induction hypothesis, $P_{(\Downarrow S)} = P_{S}$, hence
+\begin{align*}
+P_{[\epsilon \cdot (\Downarrow S)^{*p}] \cdot \epsilon}(s) &= P_{(\Downarrow S)^{*p}}(s)\\
+&= \sum_n \sum_{s_1 \cdot \ldots \cdot s_n=s} p^n * (1-p) * \prod_{i=1}^n P_{(\Downarrow S)}(s_i)\\
+&= \sum_n \sum_{s_1 \cdot \ldots \cdot s_n=s} p^n * (1-p) * \prod_{i=1}^n P_{S}(s_i)\\
+&= P_{S^{*p}}(s)
+\end{align*}
+\item
+$(S = S_1 \cdot S_2)$
+
+By definition, $\Downarrow (S_1 \cdot S_2) = (\Downarrow S_1) \odot (\Downarrow S_2)$.
+By the induction hypothesis, $P_{S_1} = P_{(\Downarrow S_1)}$, and $P_{S_2} = P_{(\Downarrow S_2)}$.
+Assume that $\Downarrow S_1 = \langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_m, p_n)\rangle$ and $\Downarrow S_2 = \langle (TQ_1,q_1) \; | \; \ldots \; | \; (TQ_n, q_m) \rangle$.
+Then
+\begin{align*}
+P_{S_1}(s) &= \sum_{i=1}^n p_i P_{SQ_i}(s)\\
+P_{S_1}(s) &= \sum_{j=1}^m q_i P_{TQ_j}(s)\\
+\end{align*}
+hence
+\begin{align*}
+P_{S_1 \cdot S_2}(s) &= \sum_{s_1 \cdot s_2 = s} P_{S_1}(s_1)* P_{S_2}(s_2)\\
+&= \sum_{s_1 \cdot s_2 = s} \left(\sum_{i=1}^n p_i P_{SQ_i}(s_1)\right)* \left(\sum_{j=1}^m q_i P_{TQ_j}(s_2)\right)\\
+&= \sum_{s_1 \cdot s_2 = s} \sum_{i=1}^n \sum_{j=1}^m (p_i * q_j) * (P_{SQ_i}(s_1) *  P_{TQ_j}(s_2))\\
+&= \sum_{i=1}^n \sum_{j=1}^m (p_i * q_j)\sum_{s_1 \cdot s_2 = s} (P_{SQ_i}(s_1) *  P_{TQ_j}(s_2))
+\end{align*}
+By definition,
+$\ConcatDNFOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}{\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}=$\\
+      $\DNFLeft (\ConcatSequenceOf{\Sequence_1}{\SequenceAlt_1},\Probability_1*\ProbabilityAlt_1)\DNFSep \ldots
+      \DNFSep
+      (\ConcatSequenceOf{\Sequence_1}{\SequenceAlt_m},\Probability_1*\ProbabilityAlt_m)\DNFSep
+      \ldots$\\
+      $\DNFSep
+      (\ConcatSequenceOf{\Sequence_n}{\SequenceAlt_1},\Probability_n*\ProbabilityAlt_1)\DNFSep
+      \ldots \DNFSep
+      (\ConcatSequenceOf{\Sequence_n}{\SequenceAlt_m},\Probability_n * \ProbabilityAlt_m) \DNFRight$
+
+where
+$$\ConcatSequenceOf{[\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n]}{[\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\StringAlt_m]}=
+  [\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n\Concat\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\StringAlt_m]$$
+Hence
+$$P_{\langle (SQ_1, p_1) \; | \; \ldots \; | \;(SQ_n,p_n)\rangle \odot \langle (TQ_1, q_1) \; | \; \ldots \; | \;(TQ_m,q_m)\rangle}(s) = \sum_{i=1}^n \sum_{j=1}^m(p_i * q_j) * P_{SQ_i \odot_{SQ} TQ_j}(s)$$
+Since $P_{[s_0 \cdot A_1 \cdot \ldots \cdot A_n \cdot s_n]}(s) = \sum_{s_0 \cdot s'_1 \cdot \ldots \cdot s'_n \cdot s_n = s} \prod_{i=1}^n P_{A_i}(s'i)$, then
+\begin{align*}
+P_{[s_0 \cdot A_1 \cdot \ldots \cdot A_n \cdot s_n] \odot_{SQ} [t_0 \cdot B_1 \cdot \ldots \cdot B_m \cdot t_m]}(s) &= \sum_{s_0 \cdot s'_1 \cdot \ldots \cdot s'_n \cdot s_n \cdot t_0 \cdot t'_1 \cdot \ldots t'_m \cdot t_m =s } \prod_{i=1}^n P_{A_i}(s'_i) \prod_{j=1}^m P_{B_j}(t'_j)\\
+&= \sum_{a \cdot b = s} \sum_{s_0 \cdot s'_1 \cdot \ldots \cdot s'_n \cdot s_n = a} \sum_{t_0 \cdot t'_1 \cdot \ldots \cdot t'_m \cdot t_m = b} \prod_{i=1}^n P_{A_i}(s'_i) \prod_{j=1}^m P_{B_j}(t'_j)\\
+& \sum_{a \cdot b = s} \left(\sum_{s_0 \cdot s'_1 \cdot \ldots \cdot s'_n \cdot s_n = a} \prod_{i=1}^n P_{A_i}(s'_i)\right) \left(\sum_{t_0 \cdot t'_1 \cdot \ldots \cdot t'_m \cdot t_m = b} \prod_{j=1}^m P_{B_j}(t'_j)\right)\\
+&= \sum_{a \cdot b = s}P_{[s_0 \cdot A_1 \cdot \ldots \cdot A_n \cdot s_n]}(a) * P_{[t_0 \cdot B_1 \cdot \ldots \cdot B_m \cdot t_m]}(b)
+\end{align*}
+In other words, $P_{SQ \odot_{SQ} TQ}(s) = \sum_{a \cdot b = s}P_{SQ}(a) * P_{TQ}(b)$. Hence, 
+\begin{align*}
+P_{S_1 \cdot S_2}(s) &= \sum_{i=1}^n \sum_{j=1}^m (p_i * q_j)\sum_{s_1 \cdot s_2 = s} (P_{SQ_i}(s_1) *  P_{TQ_j}(s_2))\\
+&= \sum_{i=1}^n \sum_{j=1}^m (p_i * q_j)\sum_{s_1 \cdot s_2 = s} P_{SQ_i \odot_{SQ} TQ_j}(s)\\
+&= P_{\langle (SQ_1, p_1) \; | \; \ldots \; | \;(SQ_n,p_n)\rangle \odot \langle (TQ_1, q_1) \; | \; \ldots \; | \;(TQ_m,q_m)\rangle}(s)\\
+&= P_{(\Downarrow S_1) \odot (\Downarrow S_2)}(s)
+\end{align*}
+which is what we wanted to show.
+\item
+$(S = S_1 \; |_p \; S_2)$
+
+By definition $\Downarrow (S_1 \; |_p \; S_2) = (\Downarrow S_1) \oplus_p (\Downarrow S_2)$.
+By the induction hypothesis $P_{S_1} = P_{\Downarrow S_1}$ and $P_{S_2} = P_{\Downarrow S_2}$.
+Assume that $\Downarrow S_1 = \langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_m, p_n)\rangle$ and $\Downarrow S_2 = \langle (TQ_1,q_1) \; | \; \ldots \; | \; (TQ_n, q_m) \rangle$.
+By definition,
+
+  $\OrDNFOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}{\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}{\Probability} =$\\
+  $\DNFOf{(\Sequence_1,\Probability_1*\Probability)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n*\Probability)\DNFSep(\SequenceAlt_1,\ProbabilityAlt_1*(1-\Probability))\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m*(1-\Probability))}$
+  hence
+  \begin{align*}
+  P_{(\Downarrow S_1) \; |_p \; (\Downarrow S_2)}(s) &= P_{\langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_n, p_n)\rangle \oplus_p \langle (TQ1, q_1) \; | \; \ldots \; | \; (TQ_m, q_m)\rangle}(s) \\
+  &= P_{\langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_n, p_n) \; | \; (TQ1, q_1) \; | \; \ldots \; | \; (TQ_m, q_m)\rangle}(s)\\
+  &= \sum_{i=1}^m (p * p_i) P_{SQ_i}(s) + \sum_{j=1}^m ((1-p) * q_j) P_{TQ_j}(s)\\
+  &= p * \sum_{i=1}^m p_i P_{SQ_i}(s) + (1-p)* \sum_{j=1}^m q_j * P_{TQ_j}(s)\\
+&= p * P_{\langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_n, p_n)\rangle}(s) + (1-p) * P_{\langle (TQ1, q_1) \; | \; \ldots \; | \; (TQ_m, q_m)\rangle}(s)\\
+&= p * P_{\Downarrow S_1}(s) + (1-p) * P_{\Downarrow S_2}(s) 
+  \end{align*}
+
+  Since $P_{S_1} = P_{\Downarrow S_1}$ and $P_{S_1} = P_{\Downarrow S_1}$, then 
+  \begin{align*}
+  P_{S_1}(s) &= \sum_{i=1}^n p_i * P_{SQ_i}(s)\\
+  P_{S_2}(s) &= \sum_{j=1}^m q_m * P_{TQ_j}(s)
+  \end{align*}
+  hence
+  \begin{align*}
+  P_{S_1 \; |_p \; S_2}(s) &= p * (P_{S_1}(s)) + (1-p) * (P_{S_2}(s))\\
+  &= p \left(\sum_{i=1}^n p_i * P_{SQ_i}(s)\right) + (1-p)\left(\sum_{j=1}^m q_m * P_{TQ_j}(s)\right)\\
+  &= p * P_{\Downarrow S_1}(s) + (1-p) * P_{\Downarrow S_2}(s)\\
+  &= P_{(\Downarrow S_1) \; |_p \; (\Downarrow S_2)}(s)
+  \end{align*}
+  which is what we wanted to show. This completes the proof.
+\end{enumerate}
+\end{proof}
+\begin{theorem*}
+  \label{thm:correct_entropy}
+  If $\Regex$ is unambiguous and contains no empty subcomponents,
+  $\EntropyOf{\Regex}$ is the entropy of $\mathcal{L}(\Regex)$.
+\end{theorem*}
+\begin{proof}
+Given a discrete set $X$ and a probability distribution $P$ on $X$, we define the entropy $H(X)$ of $X$ by
+$$H(X) = -\sum_{x \in X}P(x)\log_2{P(x)}$$
+We prove the theorem by induction on $S$.
+\begin{enumerate}
+\item
+($S = s$)
+
+Then $\mathbb{H}(s) = 0 = H(\mathcal{L}(s)) = H(\{s\})$
+\item
+$(S = S^{*p})$
+
+Since $S^{*p}$ is unambiguous, then $S$ is also anambiguous, and for each $s \in \mathcal{L}(S^{*p})$, there exist unique $t_1, \ldots, t_n \in \mathcal{L}(S)$ such that $s = t^s_1 \cdot \ldots \cdot t^s_{n_s}$.
+By the induction hypothesis, $H(\mathcal{L}(S)) = \mathbb{H}(S)$, thus
+\begin{align*}
+H(\mathcal{L}(S^{*p})) &= - \sum_{s \in \mathcal{L}(S^{*p})} P_{S^{*p}}(s) \log_2 P_{S^{*p}}(s)\\
+&= -(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)\log_2 \left(p^n * (1 - p) * \prod_{i=1}^n P_S(s_i)\right)\\
+&= -(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)\left(n \log_2 p  + \log_2 (1 - p) + \log_2 \left(\prod_{i=1}^n P_S(s_i)\right)\right)\\
+&= -(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)\left(n \log_2 p  + \log_2 (1 - p) + \log_2 \left(\prod_{i=1}^n P_S(s_i)\right)\right)\\
+&= -(1-p) * \log_2 p \sum_n n * p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)
+-(1-p) * \log_2 (1 - p) \sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i) \\
+&-(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)\log_2 \left(\prod_{i=1}^n P_S(s_i)\right)\\
+&= -\log_2 p * \frac{p}{(1-p)} -\log_2 (1 - p) 
+-(1-p)\sum_n p^n H(\mathcal{L}(S^n))\\
+&= -\log_2 p * \frac{p}{(1-p)} -\log_2 (1 - p) 
+-(1-p)\sum_n p^n n H(\mathcal{L}(S))\\
+&= -\log_2 p * \frac{p}{(1-p)} -\log_2 (1 - p) 
+-\frac{p}{(1-p)} * H(\mathcal{L}(S))\\
+&= \frac{p}{(1-p)}(H(\mathcal{L}(S)) - \log_2 p) -\log_2 (1 - p)\\
+&= \frac{p}{(1-p)}(\mathbb{H}(S) - \log_2 p) -\log_2 (1 - p)
+\end{align*}
+\item
+($S = S_1 \cdot S_2$)
+
+Since $S$ is unambigous, then $S_1$ and $S_2$ are also unambiguous, and for each $s \in \mathcal{L}(S_1 \cdot S_2)$ there exists a unique $s_1 \in \mathcal{L}(S_1)$ and a unique $s_2 \in \mathcal{L}(S_2)$ such that $s = s_1 \cdot s_2$.
+By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$, hence
+\begin{align*}
+H(\mathcal{L}(S_1 \cdot S_2))(s) &= -\sum_{s_1 \in S_1} \sum_{s_2 \in S_2}P_{S_1}(s_1)*P_{s_2}(s_2) \log_2(P_{S_1}(s_1)*P_{S_2}(s_2))\\
+&= -\sum_{s_1 \in S_1} P_{S_1}(s_1)\sum_{s_2 \in S_2}P_{s_2}(s_2) (\log_2(P_{S_1}(s_1) + \log_2 P_{S_2}(s_2)))\\
+&= -\sum_{s_1 \in S_1} P_{S_1}(s_1) \sum_{s_2 \in S_2}P_{S_2}(s_2)\log_2(P_{S_1}(s_1)) + \sum_{s_1 \in S_1} P_{S_1}(s_1)\sum_{s_2 \in S_2} P_{S_2}(s_2)\log_2 P_{S_2}(s_2)\\
+&= -\sum_{s_1 \in S_1} P_{S_1}(s_1) \log_2(P_{S_1}(s_1)) + \sum_{s_2 \in S_2} P_{S_2}(s_2)\log_2 P_{S_2}(s_2)\\
+&= H(\mathcal{L}(S_1)) + H(\mathcal{L}(S_2))\\
+&= \mathbb{H}(S_1) + \mathbb{H}(S_2)
+\end{align*}
+\item
+($S = S_1 \; |_p \; S_2$)
+
+Since $S$ is unambiguous, then $S_1$ and $S_2$ are also unambiguous, and every $s \in \mathcal{L}(S_1 \; |_p \; S_2)$ is either in $\mathcal{L}(S_1)$ or in $\mathcal{L}(S_2)$.
+By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$.
+Then
+\begin{align*}
+H(\mathcal{L}(S_1 \; |_p \; S_2)) &= -\sum_{s \in \mathcal{L}(S_1)}(p * P_{S_1}(s))\log_2 (p * P_{S_1}(s)) - \sum_{s \in \mathcal{L}(S_2)}((1-p) * P_{S_2}(s))\log_2 ((1-p) * P_{S_2}(s))\\
+&=- p\sum_{s \in \mathcal{L}(S_1)}P_{S_1}(s)(\log_2 p + \log_2 P_{S_1}(s)) - (1-p)\sum_{s \in \mathcal{L}(S_2)}P_{S_2}(s)(\log_2 (1-p) + \log_2 P_{S_2}(s))\\
+&= p * (\log_2 p +  H(\mathcal{L}(S_1))) + (1-p) * ( \log_2 (1-p) + H(\mathcal{L}(S_2)))\\
+&= p * (\log_2 p +  \mathbb{H}(S_1)) + (1-p) * ( \log_2 (1-p) + \mathbb{H}(S_2))\\
+\end{align*}
+\end{enumerate}
+\end{proof}
+\begin{theorem*}
+  Let $\Lens$ be an asymmetric lens. $\Lens$ is also a simple symmetric lens,
+  where
+  \begin{center}
+    \begin{tabular}{rcl}
+      $\Lens.\CreateROf{x}$ & $=$ & $\Lens.get \App x$\\
+      $\Lens.\CreateLOf{y}$ & $=$ & $\Lens.create \App y$\\
+      $\Lens.\PutROf{x}{y}$ & $=$ & $\Lens.get \App x$\\
+      $\Lens.\PutLOf{y}{x}$ & $=$ & $\Lens.put \App y \App x$
+    \end{tabular}
+  \end{center}
+\end{theorem*}
+\begin{proof}
+Let $\ell$ be an asymmetric lens. Then $\ell$ satisfies the following laws:
+\begin{align*}
+\ell.\get \; (\ell.\pput \; v \; s) &= v \tag{PUTGET}\\
+\ell.\pput \; (\ell.\get \; s) \; s&= s \tag{GETPUT}\\
+\ell.\get \; (\ell.\pput \; v \; s) &= v \tag{CREATGET}
+\end{align*}
+Define the simple symmetric lens $\lceil \ell \rceil$ by 
+\begin{align*}
+\lceil \ell \rceil.\CreateROf{s} &= \ell.\get \; s\\
+\lceil \ell \rceil.\CreateLOf{v} &= \ell.\create \; v\\
+\lceil \ell \rceil.\PutROf{s}{v} &= \ell.\get \; s\\
+\lceil \ell \rceil.\PutLOf{v}{s} &= \ell.\pput \; v \; s\\
+\end{align*}
+We now show that $\lceil \ell \rceil$ satisfies the simple symmetric lens laws:
+\begin{align*}
+\lceil \ell \rceil.\PutLOf{(\lceil \ell \rceil.\CreateROf{s})}{s} &= \ell.\pput \; (\ell.\get \; s) \; s = s\\
+\lceil \ell \rceil.\PutROf{(\lceil \ell \rceil.\CreateLOf{v})}{v} &= \ell.\get \; (\ell.\create \; v) = v\\
+\lceil \ell \rceil.\PutLOf{(\lceil \ell \rceil.\PutROf{s}{v})}{s} &= \ell.\pput \; (\ell.\get) \; s = s\\
+\lceil \ell \rceil.\PutROf{(\lceil \ell \rceil.\PutLOf{v}{s})}{v} &= \ell.\get \; (\ell.\pput \; v \; s) = v
+\end{align*}
+\end{proof}
+
+\begin{theorem*}
+  Let $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$, where $\Lens$ does not
+  include composition, $\Regex$ and $\RegexAlt$ and unambiguous, and neither
+  $\Regex$ nor $\RegexAlt$ contain any empty subcomponents.
+  \begin{enumerate}
+  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ calculates the expected
+    information content of $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$,
+    given $\SetOf{s \SuchThat s \in \LanguageOf{\Regex} \BooleanAnd
+      \Lens.\PutROf{s}{t} = t}$
+  \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ calculates the expected
+    information content of $\SetOf{s \SuchThat s \in \LanguageOf{\Regex}}$,
+    given $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt} \BooleanAnd
+      \Lens.\PutLOf{t}{s} = s}$
+  \end{enumerate}
+\end{theorem*}
+\begin{proof}
+Given $s \in S$, let $V_{\ell}(s) = \{t \in \mathcal{L}(T) \; | \; \ell.\PutROf{s}{t} = t\}$. We want to show that
+$$\sum_{s \in \mathcal{L}(S)}P_{\mathcal{L}(S)}(s)H(V_{\ell}(s)) = \mathbb{H}^{\rightarrow}(S \; | \; \ell, T)$$
+We proceed by induction over $\ell$. 
+\begin{enumerate}
+\item
+($\ell = \IdentityLensOf{S}$)
+
+Observe that $V_{\IdentityLensOf{S}}(s) = \{t \in \mathcal{L}(S) \; | \; \IdentityLensOf{S}.\ell.\PutROf{s}{t} = t\} = \{t \in \mathcal{L}(S) \; | \; t = s\} = \{s\}$.
+
+Consequently, $H(V_{\IdentityLensOf{S}}(s)) = H(\{s\}) = 0$, from which it follows that 
+$$\sum_{s \in \mathcal{L}(S)} P_{\mathcal{L}(S)}(s) * H(V_{\IdentityLensOf{S}}(s)) = 0 = \mathbb{H}^{\rightarrow}(S \; | \; \IdentityLensOf{S}, T)$$
+\item
+($\ell = \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}$)
+
+Observe that 
+$$V_{\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}}(s) = \{t \in \mathcal{L}(T) \; | \; \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}.\PutROf{s}{t} = t\} = \{t \in \mathcal{L}(T) \; | \; t = t\} = \mathcal{L}(T)$$
+Consequently
+$$\sum_{s \in \mathcal{L}(S)}P_{S}(s)H(V_{\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}}(s)) = \sum_{s \in \mathcal{L}(S)}P_{S}(s)H(\mathcal{L}(T)) = H(\mathcal{L}(T)) = \mathbb{H}(T) = \mathbb{H}^{\rightarrow}(S \; | \; \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, T)$$
+\item
+($\ell = \IterateLensOf{\Lens} : S^{*p} \Leftrightarrow T^{*q}$)
+
+Observe that 
+\begin{align*}
+V_{\IterateLensOf{\Lens}}(s_1 \cdot \ldots \cdot s_n) &= \{t_1 \cdot \ldots \cdot t_m \; | \; \IterateLensOf{\Lens}.\PutROf{(s_1 \cdot \ldots \cdot s_n) }{(t_1 \cdot \ldots \cdot t_m)} = t_1 \cdot \ldots \cdot t_m\}\\
+&= \{t_1 \cdot \ldots \cdot t_n \; | \; \ell.\PutROf{s_i}{t_i} = t_i \text{ for }1 \leq i \leq n\}\\
+&= V_{\ell}(s_1) \cdot \ldots \cdot V_{\ell}(s_n)
+\end{align*}
+Consequently
+\begin{align*}
+\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V_{\IterateLensOf{\Lens}}(s))
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n)\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i) H(V_{\ell}(s_1) \cdot \ldots \cdot V_{\ell}(s_n))\\
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n)\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i) \left(\sum_{j=1}^n H(V_{\ell}(s_j)) \right)\\
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \sum_{s_n} P_S(s_n) \left(\sum_{j=1}^n H(V_{\ell}(s_j)) \right)
+\end{align*}
+We focus on the term $\sum_{s_n} P_S(s_n) \sum_{j=1}^n H(V_{\ell}(s_j))$:
+\begin{align*}
+\sum_{s_n} P_S(s_n) \sum_{j=1}^n H(V_{\ell}(s_j)) &= \sum_{s_n} P_S(s_n)H(V_{\ell}(s_n)) + P_S(s_n) \sum_{j=1}^{n-1} H(V_{\ell}(s_j))\\
+&= \mathbb{H}^{\rightarrow}(S \; | \; \ell, T) + \sum_{j=1}^{n-1}H(V_{\ell}(s_j))
+\end{align*}
+Consequently,
+\begin{align*}
+\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V_{\IterateLensOf{\Lens}}(s))
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left( \mathbb{H}^{\rightarrow}(S \; | \; \ell, T) + \sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)\\
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1})\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \mathbb{H}^{\rightarrow}(S \; | \; \ell, T)\\
+&+(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)\\
+&= \mathbb{H}^{\rightarrow}(S \; | \; \ell, T) * (1-p)\sum_n p^n \\
+&+(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)
+\end{align*}
+Unrolling the term 
+$$\sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)$$ 
+$(n-1)$ more times gives
+$$\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V_{\IterateLensOf{\Lens}}(s))
+= \mathbb{H}^{\rightarrow}(S \; | \; \ell, T) * (1-p)\sum_n n * p^n  = \frac{p}{1-p} * \mathbb{H}^{\rightarrow}(S \; | \; \ell, T)$$
+which is what we wanted to show.
+\item
+($S = \ConcatLensOf{\Lens_1}{\Lens_2}$)
+
+Observe that 
+\begin{align*}
+V_{ \ConcatLensOf{\Lens_1}{\Lens_2}}(s_1 \cdot s_2) &= \{t_1 \cdot t_2 \in \mathcal{L}(T_1 \cdot T_2) \; | \;  \ConcatLensOf{\Lens_1}{\Lens_2}.\PutROf{(s_1 \cdot s_2)}{(t_1 \cdot t_2)} = t_1 \cdot t_2\}\\
+&= \{t_1 \cdot t_2 \in \mathcal{L}(T_1 \cdot T_2) \; | \;  \ell_1.\PutROf{s_1}{t_1 } = t_1  \text{ and } \ell_2.\PutROf{t_2}{t_2} = t_2\}\\
+&= V_{\ell_1}(s_1) \cdot V_{\ell_2}(s_2)
+\end{align*}
+Consequently,
+\begin{align*}
+&\sum_{s \in \mathcal{L}(S_1 \cdot S_2)}P_{S_1 \cdot S_2}(s) H(V_{\ConcatLensOf{\Lens_1}{\Lens_2}}(s))\\
+&= \sum_{s_1 \in \mathcal{L}(s_1)} P_{S_1}(s_1)\sum_{s_2 \in \mathcal{L}(s_2)}P_{S_2}(s_2)(H(V_{\ell_1}(s_1)) + H(V_{\ell_2}(s_2)))\\
+&= \sum_{s_1 \in \mathcal{L}(s_1)} P_{S_1}(s_1)\sum_{s_2 \in \mathcal{L}(s_2)}P_{S_2}(s_2)H(V_{\ell_1}(s_1)) + \sum_{s_1 \in \mathcal{L}(s_1)} P_{S_1}(s_1)\sum_{s_2 \in \mathcal{L}(s_2)}P_{S_2}(s_2)H(V_{\ell_2}(s_2))\\
+&= \mathbb{H}^{\rightarrow}(S_1 \; | \; \ell_1, T_1) + \mathbb{H}^{\rightarrow}(S_2 \; | \; \ell_2, T_2)
+\end{align*}
+\item
+($S = \SwapLensOf{\Lens_1}{\Lens_2}$)
+
+Similar to the previous case.
+\item
+($S = \OrLensOf{\Lens_1}{\Lens_2}$)
+
+Observe that 
+$$
+V_{ \OrLensOf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \OrLensOf{\Lens_1}{\Lens_2}.\ell.\PutROf{s}{t} = t\}
+$$
+If $s \in \mathcal{L}(S_1)$, then
+$$
+V_{ \OrLensOf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \ell_1.\PutROf{s}{t} = t\} = V_{\ell_1}(s)
+$$
+while if $s \in \mathcal{L}(S_2)$, then
+$$
+V_{ \OrLensOf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \ell_2.\PutROf{s}{t} = t\} = V_{\ell_2}(s)
+$$
+Consequently,
+\begin{align*}
+&\sum_{s \in \mathcal{L}(S_1 \; |_p \; S_2)}P_{S_1 \; |_p \; S_2}(s) H(V_{\OrLensOf{\Lens_1}{\Lens_2}}(s))\\
+&= p * \sum_{s \in \mathcal{L}(S_1)}P_{S_1}(s) H(V_{\ell_1}(s)) + (1-p) * \sum_{s \in \mathcal{L}(S_2)}P_{S_2}(s) H(V_{\ell_2}(s))\\
+&= p * (\mathbb{H}^{\rightarrow}(S_1 \; | \; \ell_1,T_1)) + (1-p) * (\mathbb{H}^{\rightarrow}(S_2 \; | \; \ell_2,T_2))
+\end{align*}
+\item
+($S = \MergeROf{\Lens_1}{\Lens_2}$)
+
+Observe that 
+$$
+V_{ \MergeROf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T) \; | \;  \MergeROf{\Lens_1}{\Lens_2}.\PutROf{s}{t} = t\}
+$$
+If $s \in \mathcal{L}(S_1)$, then
+$$
+V_{ \MergeROf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T) \; | \;  \ell_1.\PutROf{s}{t} = t\} = V_{\ell_1}(s)
+$$
+while if $s \in \mathcal{L}(S_2)$, then
+$$
+V_{ \MergeROf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T) \; | \;  \ell_2.\PutROf{s}{t} = t\} = V_{\ell_2}(s)
+$$
+Consequently,
+\begin{align*}
+&\sum_{s \in \mathcal{L}(S_1 \; |_p \; S_2)}P_{S_1 \; |_p \; S_2}(s) H(V_{\MergeROf{\Lens_1}{\Lens_2}}(s))\\
+&= p * \sum_{s \in \mathcal{L}(S_1)}P_{S_1}(s) H(V_{\ell_1}(s)) + (1-p) * \sum_{s \in \mathcal{L}(S_2)}P_{S_2}(s) H(V_{\ell_2}(s))\\
+&= p * (\mathbb{H}^{\rightarrow}(S_1 \; | \; \ell_1,T_1) ) + (1-p) * (\mathbb{H}^{\rightarrow}(S_2 \; | \; \ell_2,T_2) )
+\end{align*}
+\item
+($S = \MergeLOf{\Lens_1}{\Lens_2}$)
+
+Observe that 
+\begin{align*}
+V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s) &= \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \MergeLOf{\Lens_1}{\Lens_2}.\PutROf{s}{t} = t\}\\
+&= \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \ell_1.\PutROf{s}{t} = t \text{ or } \ell_2.\PutROf{s}{t} = t\}\\
+&= V_{\ell_1}(s) \cup V_{\ell_2}(s)
+\end{align*}
+First we consider $H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))$:
+\begin{align*}
+-H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)) &= \sum_{t \in V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}(t) \log_2 P_{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}(t)\\
+&= \sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \\
+&+ \sum_{t \in \mathcal{L}(T_2) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+\end{align*}
+We focus on the term 
+$$\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}$$
+Observe that
+\begin{align*}
+&\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
+&= \frac{q}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t) \log_2 q * P_{T_1}(t)\\
+&- \frac{q * \log_2 P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t)
+\end{align*}
+Again we focus on the term
+$$\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t) \log_2 q * P_{T_1}(t)$$
+Observe that
+\begin{align*}
+&\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t) \log_2 q * P_{T_1}(t) = \log_2 q * P_{T_1}(V_{\ell_1}(s))\\
+& + P_{T_1}(V_{\ell_1}(s))\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{P_{T_1}(t)}{P_{T_1}(V_{\ell_1}(s))} \log_2 \left( P_{T_1}(V_{\ell_1}(s)) *  \frac{P_{T_1}(t)}{P_{T_1}(V_{\ell_1}(s))} \right)\\
+&= \log_2 q * P_{T_1}(V_{\ell_1}(s)) + \log_2 P_{T_1}(V_{\ell_1}(s)) * P_{T_1}(V_{\ell_1}(s)) + P_{T_1}(V_{\ell_1}(s))* -H(V_{\ell_1}(s))\\
+&= P_{T_1}(V_{\ell_1}(s))[\log_2 q + \log_2 P_{T_1}(V_{\ell_1}(s)) - H(V_{\ell_1}(s))]
+\end{align*}
+Hence
+\begin{align*}
+&\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \\
+&= \frac{q}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t) \log_2 q * P_{T_1}(t)\\
+&- \frac{q * \log_2 P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t)\\
+&= \frac{q * P_{T_1}(V_{\ell_1}(s))[\log_2 q + \log_2 P_{T_1}(V_{\ell_1}(s)) - H(V_{\ell_1}(s))]}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
+&- \frac{q * \log_2 P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)) * P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
+&= \frac{q * P_{T_1}(V_{\ell_1}(s)) \left[\log_2 q + \log_2 \left[\frac{P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right]  -H(V_{\ell_1}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+\end{align*}
+Similarly,
+\begin{align*}
+&\sum_{t \in \mathcal{L}(T_2) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
+&= \frac{(1-q) * P_{T_2}(V_{\ell_2}(s)) \left[\log_2 (1-q) + \log_2 \left[\frac{P_{T_2}(V_{\ell_2}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] - H(V_{\ell_2}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+\end{align*}
+\begin{align*}
+&-H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))\\
+&= \frac{q * P_{T_1}(V_{\ell_1}(s)) \left[\log_2 q + \log_2 \left[\frac{P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] - H(V_{\ell_1}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
+&+ \frac{(1-q) * P_{T_2}(V_{\ell_2}(s)) \left[\log_2 (1-q) + \log_2 \left[\frac{P_{T_2}(V_{\ell_2}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] - H(V_{\ell_2}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+\end{align*}
+\end{enumerate}
+
 \end{proof}

### 2018-07-03 papers/SymmetricLenses/main.tex
merge cherry-pick

@@
 \subsection{Benchmark Suite}
-Our benchmarks are drawn from four different sources.
+Our benchmarks are drawn from four different sources.\bcp{What makes this a good
+  set of benchmarks for present purposes?}
 \begin{enumerate}
@@
   between the ad-hoc data and their cleaned forms. 
-  Optician also adopted benchmarks from Flash Fill, but required alterations to
+  \bcp{Clarify:}Optician also adopted benchmarks from Flash Fill, but required alterations to
   be in bijective form\bcp{Mysterious}. With our symmetric enhancement, we
@@
 \item We adapted \afm{29} benchmarks from Augeas.  These benchmarks are the same
-  as those present in Optician.  While these benchmarks had to be slightly
-  adopted for bijectivity, previous work addressed this lack of bijectivity by
+  as those present in Optician.  \bcp{I (think I) understand what this
+    means, but readers will not.}While these benchmarks had to be slightly
+  adapted for bijectivity, previous work addressed this lack of bijectivity by
   synthesizing quotient bijective lenses~\cite{?}.
@@
   \includegraphics{generated-graphs/times}
-  \caption{ Number of benchmarks that can be solved by a given algorithm in a
+  \caption{Number of benchmarks that can be solved by a given algorithm in a
     given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
@@
 \subsection{Symmetric Lenses}
+\bcp{I suggest adding an adjective---e.g., ``classic symmetric lenses''---to
+avoid confusion between the old ones and our new ones.}
 \paragraph*{Formalization} We have reformulated the problem addressed by
@@
 initial value of $C$, and specifies default behavior when there isn't existing
-data to stitch with.  In our example, the complement would consist of two lists
+data to stitch with.  In our example\bcp{which example?}, the complement
+would consist of two lists 
 of commands, where one list contains the Linux commands and one list contains
@@
 By describing our lenses in this complement style, we can express an incredibly
-wide class of lenses, as these complements permit storing any amount of
+wide class\bcp{where is the evidence that this class of lenses is {\em
+    usefully} wider than simple symmetric lenses?} of lenses, as these complements permit storing any amount of
 information in them.  However, they pose problems when implementing them in
@@
 synchronization forces us to maintain not just two files that we aim to
-sycnrhonize, but three. This is unintuitive, and creates many problems in
+sycnrhonize, but three. \bcp{I don't find it unintuitive, and I think we
+  should be careful not to oversell the difficulty of storing the
+  complements.  It requires more engineering, but it's not rocket
+  science.}This is unintuitive, and creates many problems in 
 practice. For example, consider the situation where I am synchronizing a crontab
@@
 and hook up that complement to be used for that synchronization task. This is a
-huge divergence from other synchronization utilities, where all that needs to be
+huge divergence from other synchronization utilities\bcp{No.  For example,
+  Unison maintains a database on each replica recording, for each file, what
+its hash was after the last time the two replicas were synchronized.}, where all
+that needs to be 
 done is maintain the files. Using those for synchronization, edits can be sent
@@
 
+\bcp{I guess we could also show that equivalence of simple symmetric lenses
+  is just the identity relation?}
+
 Despite being less expressive than symmetric lenses, simple symmetric lenses are
@@
 
-\begin{theorem}
+\begin{theorem}\bcp{To prove this, we need to state the type and laws for
+  asymmetric lenses.}
   Let $\Lens$ be an asymmetric lens. $\Lens$ is also a simple symmetric lens,

### 2018-07-03 papers/SymmetricLenses/main.tex
Even more comments

--- 2018-07-03 papers/SymmetricLenses/main.tex
@@
+>
 \newif\ifdraft\drafttrue  % set true to show comments
@@
 \subsection{Benchmark Suite Construction}
-Our benchmarks came from four different sources.
+Our benchmarks came from four different sources.\bcp{What makes this a good
+  set of benchmarks for present purposes?}
 \begin{enumerate}
@@
   synchronization functions between the ad-hoc data and their cleaned forms.
-  Optician also adopted benchmarks from Flash Fill, but required alterations to
+  \bcp{Clarify:}Optician also adopted benchmarks from Flash Fill, but required alterations to
   be in bijective form. With our symmetric enhancement, we Optician's four
@@
 
-To answer our first research question, we analyze the runtime of our algorithm,
-and compare it to the existing algorithm for synthesizing bijective lenses.
+To answer our first research question, we analyze\bcp{measure?} the runtime of our algorithm,
+and compare it to the existing algorithm for synthesizing bijective
+lenses.\bcp{Why is this the right thing to compare against?}
 
@@
     benchmarks in our benchmark suite in under 30 seconds, and is able to
-    complete most of the benchmarks without using compositional synthesis. }
+    complete most of the benchmarks without using compositional
+    synthesis. \bcp{The interesting part of the figure seems to be all
+      squished up against the Y axis.  Could we use a log-scale for the X
+      axis?}} 
   \label{fig:times}

### 2018-07-04 papers/SymmetricLenses/transform-data.py
add metric

@@ -43,59 +43,35 @@ def retrieve_csv(filename):
 def project_column_from_csv(csv_obj, col_name):
     return [r[col_name] for r in csv_obj]
 
+def count_real_vals(csv_obj, col_name):
+    col = project_column_from_csv(csv_obj, col_name)
+    return len([val for val in col if val != "[-1]"])
+
 def write_to_filename(filename, s):
     with open(filename, "wb") as f:
         f.write(s)
 
 def generate_examples_required_graph(input_csv):
-    zero_count_ind = 0
-    one_to_five_count_ind = 1
-    six_to_ten_count_ind = 2
-    over_twenty_count_ind = 3
-
-    zero_count_text = "0"
-    one_to_five_count_text = "1-5"
-    six_to_ten_count_text = "6-10"
-    eleven_to_fifteen_count_text = "11-15"
-    sixteen_to_twenty_count_text = "16-20"
-    over_twenty_count_text = ">10"
-    ind_to_text = [zero_count_text,
-                   one_to_five_count_text,
-                   six_to_ten_count_text,
-                   over_twenty_count_text]
-
-    experimental_values = [0,0,0,0,]
-
-    def add_to_correct_group(count_values, n):
-        if n < 0.0:
-            raise Exception("SOMETHING WENT WRONG")
-        if n == 0.0:
-            count_values[zero_count_ind] = count_values[zero_count_ind]+1
-        elif n <= 5.0:
-            count_values[one_to_five_count_ind] = count_values[one_to_five_count_ind]+1
-        elif n <= 10.0:
-            count_values[six_to_ten_count_ind] = count_values[six_to_ten_count_ind]+1
-        else:
-            count_values[over_twenty_count_ind] = count_values[over_twenty_count_ind]+1
-
-    experimental_vals = project_column_from_csv(input_csv, "ExamplesRequired")
-    for example_num in experimental_vals:
-        add_to_correct_group(experimental_values, float(example_num))
-    
+    nm_count = count_real_vals(input_csv,"NM")
+    nmcc_count = count_real_vals(input_csv,"NMCC")
+    fc_count = count_real_vals(input_csv,"FC")
+    notp_count = count_real_vals(input_csv,"NoTP")
 
     ind = np.arange(4)
     width = 0.35
 
     fig, ax = plt.subplots()
 
[…]

### 2018-07-04 papers/SymmetricLenses/main.tex
add graphs

--- 2018-07-03 papers/SymmetricLenses/main.tex
@@
 \EntropyOf{\lstinline{emp_row}}$. \lstinline{emp_salary} contains strictly more
-information than $\EntropyOf{\lstinline{": " . salary}}$, and
+imformation than $\EntropyOf{\lstinline{": " . salary}}$, and
 \lstinline{emp_row} contains strictly more information than
@@
 We have proven theorems demonstrating the calculated entropy corresponds to the
-actual conditional entropy to recover the data.\bcp{Again, the proofs should be
+actual conditional entropy to recover the data.\bcp{Again, they should be
   included.}
@@
 mapped to each other either if they are the same regular expression and have
-examples agreeing with the examples\bcp{What does ``have
-examples agreeing with the examples'' mean??}, or if there is a lens which is a
+examples agreeing with the examples, or if there is a lens which is a
 composition of predefined lenses between them, where the lens alters the
@@
 
-\paragraph*{\RXSearch Permissiveness Customization}\bcp{I don't follow this:}
-While the termination heuristic works for many situations, there are situations
+\paragraph*{\RXSearch Permissiveness Customization}
+While termination heuristic works for many situations: there are situations
 where a large number of expansions must be performed to find the correct lens.
@@
 
-We have extended Optician\bcp{this is the first mention of the word
-  Optician!} with our symmetric synthesis algorithm. Users can 
+We have extended Optician with our symmetric synthesis algorithm. Users can
 write synthesis tasks alongside lens combinators, use those synthesis results in
@@
 
-In this evaluation, we aim to answer two primary questions:
+In this work, we aim to answer two primary research questions:
 \begin{enumerate}
 \item Is our synthesis procedure efficient enough to be used in everyday
-  development?  In particular, can it be used in place of the existing
-  bijective lens synthesis algorithm integrated with Boomerang (which
-  \citet{miltner+:bijective-synthesis} have argued is reasonably fast in
-  practice)? 
-  \item Are the heuristics we have described useful for reducing the 
-  ``specification overhead'' for users?  \bcp{Maybe it would be better to
-    ask {\em which} heuristics are the most useful.  Also, are we talking
-    about reducing the number of examples needed?  It's not clear.  (And:
-    Don't the heuristics also make it faster??)}
+  development?  Can it be used in place of the existing bijective lens synthesis
+  algorithm integrated with Boomerang?
+\item Are our heuristics useful for reducing the specifications users have to
+  provide.
 \end{enumerate}
 
-\subsection{Benchmark Suite}
-Our benchmarks are drawn from four different sources.
+\subsection{Benchmark Suite Construction}
+Our benchmarks came from four different sources.
 \begin{enumerate}
 \item We adapted \afm{8} data cleaning benchmarks from Flash Fill. In adopting
-  these benchmarks, we go beyond the data cleaning tasks\bcp{confusing: you
-    say these are data cleaning tasks, then immediately say we go beyond
-    data cleaning tasks.  (I know the next bit of the sentence makes it
-    (somewhat) clearer, but the reader is already lost right here.)}, and
-  synthesize 
-  synchronization functions\bcp{what is a ``synchronization function''}
-  between the ad-hoc data and their cleaned forms. 
+  these benchmarks, we go beyond the data cleaning tasks, and synthesize
+  synchronization functions between the ad-hoc data and their cleaned forms.
   Optician also adopted benchmarks from Flash Fill, but required alterations to
-  be in bijective form\bcp{Mysterious}. With our symmetric enhancement, we
-  \bcp{words missing?} Optician's four 
-  benchmarks and 4 new Flash Fill benchmarks with no alterations.
+  be in bijective form. With our symmetric enhancement, we Optician's four
+  benchmarks, and 4 new Flash Fill benchmarks with no alterations.
 \item We adapted \afm{29} benchmarks from Augeas.  These benchmarks are the same
@@
 
-\bcp{Why is this a good set of benchmarks for the purposes described earlier
-in the section?}
-
 \subsection{Efficiency}
 
-To answer our first research question, we measure the runtime of our algorithm
-and compare it to the existing algorithm for synthesizing bijective
-lenses.\bcp{Note that this comparison doesn't, directly answer the question
-  ``Is it fast enough.''}
+To answer our first research question, we analyze the runtime of our algorithm,
+and compare it to the existing algorithm for synthesizing bijective lenses.
 
@@
     benchmarks in our benchmark suite in under 30 seconds, and is able to
-    complete most of the benchmarks without using compositional
-    synthesis. \bcp{The interesting part of the graph is all squished
-      against the Y axis.  Can we make the X axis logarithmic?}}
+    complete most of the benchmarks without using compositional synthesis. }
   \label{fig:times}
@@
 
-\subsection{The Effect of Our Heuristic}
+\subsection{Heuristic Usefulness}
 We evaluate the usefulness of (1) our information theoretic metric, and (2) our
@@
     cost. \NMOpt{} removes all notion of cost, and merely finds the first
-    satisfying lens.\bcp{Am I reading this wrong, or does even the best
-      algorithm fail to find the correct lens something like 20\% of the
-      time?!?!?!}} 
+    satisfying lens.}
   \label{fig:metric}

### 2018-07-04 papers/SymmetricLenses/main.tex
add all the graphs yay

--- 2018-07-03 papers/SymmetricLenses/main.tex
@@
->
 \newif\ifdraft\drafttrue  % set true to show comments
@@
   \FCOpt{}: &  The first time \GreedySynth finds a lens, it returns that lens.\\
-  \NMOpt{}: & \GreedySynth does not have any notion of cost, and searches for
-              any lens that satisfies the specification. \\
+  \NMCCOpt{}: & The lens cluster that contains is identified through \SSOpt{}. \GreedySynth is then called on that pair of SREs. \\
+  \NMOpt{}: & Neither \GreedySynth nor \RXSearch have any notion of cost; they
+              both merely search for
+              any lens that satisfies the specification. 
 \end{tabulary}
 
-Figure~\ref{fig:heuristic} summarizes the result of this experiment.  We find
-that \afm{TODO}.
+\begin{figure}
+  \includegraphics{generated-graphs/metrics_importance}
+  \caption{Number of benchmarks that synthesize the correct lens by a given
+    algorithm. \NoTPOpt{} is the full symmetric synthesis algorithm, but
+    termination parameters are ignored. \FCOpt{} provides a notion of cost to
+    \GreedySynth, but once a satisfying lens is greedily found, that lens is
+    returned. \NMCCOpt{} identifies the lens cluster in which the desired 
+    lens lies.  \GreedySynth{} is then run on that cluster with no notion of
+    cost. \NMOpt{} removes all notion of cost, and merely finds the first
+    satisfying lens.}}
+  \label{fig:metric}
+\end{figure}
+
+Figure~\ref{fig:metric} summarizes the result of this experiment. We find that
+our metrics are very important for finding the correct lens. Our
+information-theoretic metric is critical for finding the correct lens. Indeed,
+even when provided with the correct lens cluster, without a notion of cost, we
+would need constrain our search space more with additional examples to find the
+correct lens.
+
+Our termination condition is also quite important. When running in \FCOpt{},
+only 5 lenses are found. The first cluster in which a lens is found is rarely
+the correct cluster. However, our termination condition is not perfect.
+Sometimes users must tell the algorithm to ``keep going'' or ``stop early.''
+Without termination parameters, our algorithm finds the correct lens for 36 of
+our 45 benchmarks, 9 of them need help in the form of a termination parameter
+for finding the correct lens.
 % end evaluation

### 2018-07-04 papers/SymmetricLenses/transform-data.py

@@ -56,6 +56,10 @@ def generate_examples_required_graph(input_csv):
     nmcc_count = count_real_vals(input_csv,"NMCC")
     fc_count = count_real_vals(input_csv,"FC")
     notp_count = count_real_vals(input_csv,"NoTP")
+    print(nm_count)
+    print(nmcc_count)
+    print(fc_count)
+    print(notp_count)
 
     ind = np.arange(4)
     width = 0.35
@@ -64,7 +68,7 @@ def generate_examples_required_graph(input_csv):
 
     rects1 = ax.bar(ind, [nm_count,nmcc_count,fc_count,notp_count], width, color='#ffffb3', align='center')
 
-    ax.set_ylabel('Benchmarks Completed')
+    ax.set_ylabel('Benchmarks\nCompleted')
     ax.set_xlabel('Run Mode')
     ax.set_xticks(ind)
     ax.set_xticklabels(["\\textbf{NM}","\\textbf{NMCC}","\\textbf{FC}","\\textbf{NoTP}"])

### 2018-07-04 papers/SymmetricLenses/main.tex
Theorem 4 proof ongoing

--- 2018-07-03 papers/SymmetricLenses/main.tex
@@
 
-\bcp{This is a POPL paper: Proofs of important theorems should
-be included.  (This one might be easy enough that we can just label it as a
-``Fact'' and omit the proof, or better yet we can say that the proof is straightforward.}
-
 \subsection{Stochastic Regular Expression Entropy}
-The {\em entropy} of a stochastic regular expression can be computed
-directly from its syntax, when the stochastic regular expression is
-unambiguous and contains only nonempty subcomponents.\bcp{What is an empty
-  subcomponent?} 
+The entropy of a stochastic regular expression can be computed directly from its
+syntax, when the stochastic regular expression is unambiguous and contains only
+nonempty subcomponents.
 \begin{center}
@@
 \end{center}
-We have proven\bcp{Again, some kind of proof is needed.} that this syntactic
-computation corresponds to the entropy of the stochastic regular expression.
+We have proven\bcp{This is a POPL paper: Proofs of important theorems should
+be included.} that this syntactic computation corresponds to the entropy of the
+stochastic regular expression.
 
@@
 $\PRegexStar{\DNFRegex}{.8}$. In particular, $\ToDNFRegex$ generates regular
-expressions whose DNF form gives equal probability to all sequences\bcp{Why?}, and gives
+expressions whose DNF form gives equal probability to all sequences, and gives
 iterations a .8 chance to continue iterating, as that gives a reasonable
-distribution of strings when generating them from the regular
-expressions\bcp{What evaluation did we do to
-  determine that it is the right value?}.
+distribution of strings when generating them from the regular expressions.
 
@@
 \subsection{Benchmark Suite Construction}
-Our benchmarks came from four different sources.\bcp{What makes this a good
-  set of benchmarks for present purposes?}
+Our benchmarks came from four different sources.
 \begin{enumerate}
@@
   synchronization functions between the ad-hoc data and their cleaned forms.
-  \bcp{Clarify:}Optician also adopted benchmarks from Flash Fill, but required alterations to
+  Optician also adopted benchmarks from Flash Fill, but required alterations to
   be in bijective form. With our symmetric enhancement, we Optician's four
@@
 \item We adapted \afm{29} benchmarks from Augeas.  These benchmarks are the same
-  as those present in Optician.  \bcp{I (think I) understand what this
-    means, but readers will not.}While these benchmarks had to be slightly
-  adapted for bijectivity, previous work addressed this lack of bijectivity by
+  as those present in Optician.  While these benchmarks had to be slightly
+  adopted for bijectivity, previous work addressed this lack of bijectivity by
   synthesizing quotient bijective lenses~\cite{?}.
@@
 
-To answer our first research question, we analyze\bcp{measure?} the runtime of our algorithm,
-and compare it to the existing algorithm for synthesizing bijective
-lenses.\bcp{Why is this the right thing to compare against?}
+To answer our first research question, we analyze the runtime of our algorithm,
+and compare it to the existing algorithm for synthesizing bijective lenses.
 
@@
   \includegraphics{generated-graphs/times}
-  \caption{Number of benchmarks that can be solved by a given algorithm in a
+  \caption{ Number of benchmarks that can be solved by a given algorithm in a
     given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
@@
     benchmarks in our benchmark suite in under 30 seconds, and is able to
-    complete most of the benchmarks without using compositional
-    synthesis. \bcp{The interesting part of the figure seems to be all
-      squished up against the Y axis.  Could we use a log-scale for the X
-      axis?}} 
+    complete most of the benchmarks without using compositional synthesis. }
   \label{fig:times}
@@
 \subsection{Symmetric Lenses}
-\bcp{I suggest adding an adjective---e.g., ``classic symmetric lenses''---to
-avoid confusion between the old ones and our new ones.}
 \paragraph*{Formalization} We have reformulated the problem addressed by
@@
 initial value of $C$, and specifies default behavior when there isn't existing
-data to stitch with.  In our example\bcp{which example?}, the complement
-would consist of two lists 
+data to stitch with.  In our example, the complement would consist of two lists
 of commands, where one list contains the Linux commands and one list contains
@@
 By describing our lenses in this complement style, we can express an incredibly
-wide class\bcp{where is the evidence that this class of lenses is {\em
-    usefully} wider than simple symmetric lenses?} of lenses, as these complements permit storing any amount of
+wide class of lenses, as these complements permit storing any amount of
 information in them.  However, they pose problems when implementing them in
@@
 synchronization forces us to maintain not just two files that we aim to
-sycnrhonize, but three. \bcp{I don't find it unintuitive, and I think we
-  should be careful not to oversell the difficulty of storing the
-  complements.  It requires more engineering, but it's not rocket
-  science.}This is unintuitive, and creates many problems in 
+sycnrhonize, but three. This is unintuitive, and creates many problems in
 practice. For example, consider the situation where I am synchronizing a crontab
@@
 and hook up that complement to be used for that synchronization task. This is a
-huge divergence from other synchronization utilities\bcp{No.  For example,
-  Unison maintains a database on each replica recording, for each file, what
-its hash was after the last time the two replicas were synchronized.}, where all
-that needs to be 
+huge divergence from other synchronization utilities, where all that needs to be
 done is maintain the files. Using those for synchronization, edits can be sent
@@
 
-\bcp{I guess we could also show that equivalence of simple symmetric lenses
-  is just the identity relation?}
-
 Despite being less expressive than symmetric lenses, simple symmetric lenses are
@@
 
-\begin{theorem}\bcp{To prove this, we need to state the type and laws for
-  asymmetric lenses.}
+\begin{theorem}
   Let $\Lens$ be an asymmetric lens. $\Lens$ is also a simple symmetric lens,
@@
 Similar to the previous case.
+\end{enumerate}
+\end{proof}
+\begin{theorem*}
+  $\ProbabilityOf{\Regex}{\String} = \ProbabilityOf{\ToDNFRegex
+    \Regex}{\String}$ and $\LanguageOf{\Regex} =
+  \LanguageOf{\ToDNFRegexOf{\Regex}}$
+\end{theorem*}
+\begin{proof}
+\begin{enumerate}
+\item
+$(S = s)$
+
+By definition, $\Downarrow s = \langle ([s],1) \rangle$. The language of $\langle ([s],1) \rangle$ is just $s$ and $P_{\langle ([s], 1) \rangle}(s') = 
+\begin{cases}
+1 & \text{if } s = s'\\
+0 & \text{otherwise}
+\end{cases}$ hence $P_{\langle ([s],1) \rangle}(s') = P_s(s')$.
+\item
+$(S = \varnothing)$
+
+By definition, $\Downarrow \varnothing = \langle \rangle$. The language of $\langle \rangle$ is the empty set, and $P_{\langle \rangle}(s) = 0 = P_{\varnothing}(s)$
+\item
+$(S = S^{*p})$
+
+By definition, $\Downarrow (S^{*p}) = \langle [\epsilon \cdot (\Downarrow S)^{*p} \cdot \epsilon] \rangle $.
+By the induction hypothesis, $P_{(\Downarrow S)} = P_{S}$ and $\mathcal{L}(\Downarrow S) = \mathcal{S}$, hence $\mathcal{L}(\Downarrow S) = \mathcal{L}(S)$.
+Also,
+\begin{align*}
+P_{[\epsilon \cdot (\Downarrow S)^{*p}] \cdot \epsilon}(s) &= P_{(\Downarrow S)^{*p}}(s)\\
+&= \sum_n \sum_{s_1 \cdot \ldots \cdot s_n=s} p^n * (1-p) * \prod_{i=1}^n P_{(\Downarrow S)}(s_i)\\
+&= \sum_n \sum_{s_1 \cdot \ldots \cdot s_n=s} p^n * (1-p) * \prod_{i=1}^n P_{S}(s_i)\\
+&= P_{S^{*p}}(s)
+\end{align*}
+\item
+$(S = S_1 \cdot S_2)$
+
+By definition, $\Downarrow (S_1 \cdot S_2) = (\Downarrow S_1) \odot (\Downarrow S_2)$.
+By the induction hypothesis, $\mathcal{L}(S_1) = \mathcal{L}(\Downarrow S_1)$ and $P_{S_1} = P_{(\Downarrow S_1)}$, and $\mathcal{L}(S_2) = \mathcal{L}(\Downarrow S_2)$ and $P_{S_2} = P_{(\Downarrow S_2)}$
 \end{enumerate}

### 2018-07-04 papers/SymmetricLenses/main.tex
Added concat case in Theorem 4

--- 2018-07-04 papers/SymmetricLenses/main.tex
@@
 \end{proof}
+
 \begin{theorem*}
@@
 \begin{proof}
+By \sam{COPY PASTE PROOF FROM BIJECTIVE LENS PAPER} $\LanguageOf{\Regex} =
+  \LanguageOf{\ToDNFRegexOf{\Regex}}$, so we will prove the result $\ProbabilityOf{\Regex}{\String} = \ProbabilityOf{\ToDNFRegex
+    \Regex}{\String}$.
 \begin{enumerate}
@@
 
-By definition, $\Downarrow s = \langle ([s],1) \rangle$. The language of $\langle ([s],1) \rangle$ is just $s$ and $P_{\langle ([s], 1) \rangle}(s') = 
+By definition, $\Downarrow s = \langle ([s],1) \rangle$ and $P_{\langle ([s], 1) \rangle}(s') = 
 \begin{cases}
@@
 
-By definition, $\Downarrow \varnothing = \langle \rangle$. The language of $\langle \rangle$ is the empty set, and $P_{\langle \rangle}(s) = 0 = P_{\varnothing}(s)$
+By definition, $\Downarrow \varnothing = \langle \rangle$ and $P_{\langle \rangle}(s) = 0 = P_{\varnothing}(s)$
 \item
@@
 
-By definition, $\Downarrow (S^{*p}) = \langle [\epsilon \cdot (\Downarrow S)^{*p} \cdot \epsilon] \rangle $.
-By the induction hypothesis, $P_{(\Downarrow S)} = P_{S}$ and $\mathcal{L}(\Downarrow S) = \mathcal{S}$, hence $\mathcal{L}(\Downarrow S) = \mathcal{L}(S)$.
-Also,
+By definition, $\Downarrow (S^{*p}) = \langle [\epsilon \cdot (\Downarrow S)^{*p} \cdot \epsilon] \rangle $. By the induction hypothesis, $P_{(\Downarrow S)} = P_{S}$, hence
 \begin{align*}
@@
 By definition, $\Downarrow (S_1 \cdot S_2) = (\Downarrow S_1) \odot (\Downarrow S_2)$.
-By the induction hypothesis, $\mathcal{L}(S_1) = \mathcal{L}(\Downarrow S_1)$ and $P_{S_1} = P_{(\Downarrow S_1)}$, and $\mathcal{L}(S_2) = \mathcal{L}(\Downarrow S_2)$ and $P_{S_2} = P_{(\Downarrow S_2)}$
+By the induction hypothesis, $P_{S_1} = P_{(\Downarrow S_1)}$, and $P_{S_2} = P_{(\Downarrow S_2)}$.
+Assume that $\Downarrow S_1 = \langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_m, p_n)\rangle$ and $\Downarrow S_2 = \langle (TQ_1,q_1) \; | \; \ldots \; | \; (TQ_n, q_m) \rangle$.
+Then
+\begin{align*}
+P_{S_1}(s) &= \sum_{i=1}^n p_i P_{SQ_i}(s)\\
+P_{S_1}(s) &= \sum_{j=1}^m q_i P_{TQ_j}(s)\\
+\end{align*}
+hence
+\begin{align*}
+P_{S_1 \cdot S_2}(s) &= \sum_{s_1 \cdot s_2 = s} P_{S_1}(s_1)* P_{S_2}(s_2)\\
+&= \sum_{s_1 \cdot s_2 = s} \left(\sum_{i=1}^n p_i P_{SQ_i}(s_1)\right)* \left(\sum_{j=1}^m q_i P_{TQ_j}(s_2)\right)\\
+&= \sum_{s_1 \cdot s_2 = s} \sum_{i=1}^n \sum_{j=1}^m (p_i * q_j) * (P_{SQ_i}(s_1) *  P_{TQ_j}(s_2))\\
+&= \sum_{i=1}^n \sum_{j=1}^m (p_i * q_j)\sum_{s_1 \cdot s_2 = s} (P_{SQ_i}(s_1) *  P_{TQ_j}(s_2))
+\end{align*}
+By definition,
+$\ConcatDNFOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}{\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}=$\\
+      $\DNFLeft (\ConcatSequenceOf{\Sequence_1}{\SequenceAlt_1},\Probability_1*\ProbabilityAlt_1)\DNFSep \ldots
+      \DNFSep
+      (\ConcatSequenceOf{\Sequence_1}{\SequenceAlt_m},\Probability_1*\ProbabilityAlt_m)\DNFSep
+      \ldots$\\
+      $\DNFSep
+      (\ConcatSequenceOf{\Sequence_n}{\SequenceAlt_1},\Probability_n*\ProbabilityAlt_1)\DNFSep
+      \ldots \DNFSep
+      (\ConcatSequenceOf{\Sequence_n}{\SequenceAlt_m},\Probability_n * \ProbabilityAlt_m) \DNFRight$
+
+where
+$$\ConcatSequenceOf{[\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n]}{[\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\StringAlt_m]}=
+  [\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n\Concat\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\StringAlt_m]$$
+Hence
+$$P_{\langle (SQ_1, p_1) \; | \; \ldots \; | \;(SQ_n,p_n)\rangle \odot \langle (TQ_1, q_1) \; | \; \ldots \; | \;(TQ_m,q_m)\rangle}(s) = \sum_{i=1}^n \sum_{j=1}^m(p_i * q_j) * P_{SQ_i \odot_{SQ} TQ_j}(s)$$
+Since $P_{[s_0 \cdot A_1 \cdot \ldots \cdot A_n \cdot s_n]}(s) = \sum_{s_0 \cdot s'_1 \cdot \ldots \cdot s'_n \cdot s_n = s} \prod_{i=1}^n P_{A_i}(s'i)$, then
+\begin{align*}
+P_{[s_0 \cdot A_1 \cdot \ldots \cdot A_n \cdot s_n] \odot_{SQ} [t_0 \cdot B_1 \cdot \ldots \cdot B_m \cdot t_m]}(s) &= \sum_{s_0 \cdot s'_1 \cdot \ldots \cdot s'_n \cdot s_n \cdot t_0 \cdot t'_1 \cdot \ldots t'_m \cdot t_m =s } \prod_{i=1}^n P_{A_i}(s'_i) \prod_{j=1}^m P_{B_j}(t'_j)\\
+&= \sum_{a \cdot b = s} \sum_{s_0 \cdot s'_1 \cdot \ldots \cdot s'_n \cdot s_n = a} \sum_{t_0 \cdot t'_1 \cdot \ldots \cdot t'_m \cdot t_m = b} \prod_{i=1}^n P_{A_i}(s'_i) \prod_{j=1}^m P_{B_j}(t'_j)\\
+& \sum_{a \cdot b = s} \left(\sum_{s_0 \cdot s'_1 \cdot \ldots \cdot s'_n \cdot s_n = a} \prod_{i=1}^n P_{A_i}(s'_i)\right) \left(\sum_{t_0 \cdot t'_1 \cdot \ldots \cdot t'_m \cdot t_m = b} \prod_{j=1}^m P_{B_j}(t'_j)\right)\\
+&= \sum_{a \cdot b = s}P_{[s_0 \cdot A_1 \cdot \ldots \cdot A_n \cdot s_n]}(a) * P_{[t_0 \cdot B_1 \cdot \ldots \cdot B_m \cdot t_m]}(b)
+\end{align*}
+In other words, $P_{SQ \odot_{SQ} TQ}(s) = \sum_{a \cdot b = s}P_{SQ}(a) * P_{TQ}(b)$. Hence, 
+\begin{align*}
+P_{S_1 \cdot S_2}(s) &= \sum_{i=1}^n \sum_{j=1}^m (p_i * q_j)\sum_{s_1 \cdot s_2 = s} (P_{SQ_i}(s_1) *  P_{TQ_j}(s_2))\\
+&= \sum_{i=1}^n \sum_{j=1}^m (p_i * q_j)\sum_{s_1 \cdot s_2 = s} P_{SQ_i \odot_{SQ} TQ_j}(s)\\
+&= P_{\langle (SQ_1, p_1) \; | \; \ldots \; | \;(SQ_n,p_n)\rangle \odot \langle (TQ_1, q_1) \; | \; \ldots \; | \;(TQ_m,q_m)\rangle}(s)\\
+&= P_{(\Downarrow S_1) \odot (\Downarrow S_2)}(s)
+\end{align*}
+which is what we wanted to show.
+\item
 \end{enumerate}

### 2018-07-04 papers/SymmetricLenses/main.tex
Finished proof of Theorem 4

@@
 \item
+$(S = S_1 \; |_p \; S_2)$
+
+By definition $\Downarrow (S_1 \; |_p \; S_2) = (\Downarrow S_1) \oplus_p (\Downarrow S_2)$.
+By the induction hypothesis $P_{S_1} = P_{\Downarrow S_1}$ and $P_{S_2} = P_{\Downarrow S_2}$.
+Assume that $\Downarrow S_1 = \langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_m, p_n)\rangle$ and $\Downarrow S_2 = \langle (TQ_1,q_1) \; | \; \ldots \; | \; (TQ_n, q_m) \rangle$.
+By definition,
+
+  $\OrDNFOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}{\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}{\Probability} =$\\
+  $\DNFOf{(\Sequence_1,\Probability_1*\Probability)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n*\Probability)\DNFSep(\SequenceAlt_1,\ProbabilityAlt_1*(1-\Probability))\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m*(1-\Probability))}$
+  hence
+  \begin{align*}
+  P_{(\Downarrow S_1) \; |_p \; (\Downarrow S_2)}(s) &= P_{\langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_n, p_n)\rangle \oplus_p \langle (TQ1, q_1) \; | \; \ldots \; | \; (TQ_m, q_m)\rangle}(s) \\
+  &= P_{\langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_n, p_n) \; | \; (TQ1, q_1) \; | \; \ldots \; | \; (TQ_m, q_m)\rangle}(s)\\
+  &= \sum_{i=1}^m (p * p_i) P_{SQ_i}(s) + \sum_{j=1}^m ((1-p) * q_j) P_{TQ_j}(s)\\
+  &= p * \sum_{i=1}^m p_i P_{SQ_i}(s) + (1-p)* \sum_{j=1}^m q_j * P_{TQ_j}(s)\\
+&= p * P_{\langle (SQ_1,p_1) \; | \; \ldots \; | \; (SQ_n, p_n)\rangle}(s) + (1-p) * P_{\langle (TQ1, q_1) \; | \; \ldots \; | \; (TQ_m, q_m)\rangle}(s)\\
+&= p * P_{\Downarrow S_1}(s) + (1-p) * P_{\Downarrow S_2}(s) 
+  \end{align*}
+
+  Since $P_{S_1} = P_{\Downarrow S_1}$ and $P_{S_1} = P_{\Downarrow S_1}$, then 
+  \begin{align*}
+  P_{S_1}(s) &= \sum_{i=1}^n p_i * P_{SQ_i}(s)\\
+  P_{S_2}(s) &= \sum_{j=1}^m q_m * P_{TQ_j}(s)
+  \end{align*}
+  hence
+  \begin{align*}
+  P_{S_1 \; |_p \; S_2}(s) &= p * (P_{S_1}(s)) + (1-p) * (P_{S_2}(s))\\
+  &= p \left(\sum_{i=1}^n p_i * P_{SQ_i}(s)\right) + (1-p)\left(\sum_{j=1}^m q_m * P_{TQ_j}(s)\right)\\
+  &= p * P_{\Downarrow S_1}(s) + (1-p) * P_{\Downarrow S_2}(s)\\
+  &= P_{(\Downarrow S_1) \; |_p \; (\Downarrow S_2)}(s)
+  \end{align*}
+  which is what we wanted to show. This completes the proof.
 \end{enumerate}

### 2018-07-04 papers/SymmetricLenses/main.tex
Finished proof of Theorem 2

@@
   which is what we wanted to show. This completes the proof.
+\end{enumerate}
+\end{proof}
+\begin{theorem*}
+  \label{thm:correct_entropy}
+  If $\Regex$ is unambiguous and contains no empty subcomponents,
+  $\EntropyOf{\Regex}$ is the entropy of $\mathcal{L}(\Regex)$.
+\end{theorem*}
+\begin{proof}
+Given a discrete set $X$ and a probability distribution $P$ on $X$, we define the entropy $H(X)$ of $X$ by
+$$H(X) = -\sum_{x \in X}P(x)\log_2{P(x)}$$
+We prove the theorem by induction on $S$.
+\begin{enumerate}
+\item
+($S = s$)
+
+Then $\mathbb{H}(s) = 0 = H(\mathcal{L}(s)) = H(\{s\})$
+\item
+$(S = S^{*p})$
+
+Since $S^{*p}$ is unambiguous, then $S$ is also anambiguous, and for each $s \in \mathcal{L}(S^{*p})$, there exist unique $t_1, \ldots, t_n \in \mathcal{L}(S)$ such that $s = t^s_1 \cdot \ldots \cdot t^s_{n_s}$.
+By the induction hypothesis, $H(\mathcal{L}(S)) = \mathbb{H}(S)$, thus
+\begin{align*}
+H(\mathcal{L}(S^{*p})) &= - \sum_{s \in \mathcal{L}(S^{*p})} P_{S^{*p}}(s) \log_2 P_{S^{*p}}(s)\\
+&= -(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)\log_2 \left(p^n * (1 - p) * \prod_{i=1}^n P_S(s_i)\right)\\
+&= -(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)\left(n \log_2 p  + \log_2 (1 - p) + \log_2 \left(\prod_{i=1}^n P_S(s_i)\right)\right)\\
+&= -(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)\left(n \log_2 p  + \log_2 (1 - p) + \log_2 \left(\prod_{i=1}^n P_S(s_i)\right)\right)\\
+&= -(1-p) * \log_2 p \sum_n n * p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)
+-(1-p) * \log_2 (1 - p) \sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i) \\
+&-(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i)\log_2 \left(\prod_{i=1}^n P_S(s_i)\right)\\
+&= -\log_2 p * \frac{p}{(1-p)} -\log_2 (1 - p) 
+-(1-p)\sum_n p^n H(\mathcal{L}(S^n))\\
+&= -\log_2 p * \frac{p}{(1-p)} -\log_2 (1 - p) 
+-(1-p)\sum_n p^n n H(\mathcal{L}(S))\\
+&= -\log_2 p * \frac{p}{(1-p)} -\log_2 (1 - p) 
+-\frac{p}{(1-p)} * H(\mathcal{L}(S))\\
+&= \frac{p}{(1-p)}(H(\mathcal{L}(S)) - \log_2 p) -\log_2 (1 - p)\\
+&= \frac{p}{(1-p)}(\mathbb{H}(S) - \log_2 p) -\log_2 (1 - p)
+\end{align*}
+\item
+($S = S_1 \cdot S_2$)
+
+Since $S$ is unambigous, then $S_1$ and $S_2$ are also unambiguous, and for each $s \in \mathcal{L}(S_1 \cdot S_2)$ there exists a unique $s_1 \in \mathcal{L}(S_1)$ and a unique $s_2 \in \mathcal{L}(S_2)$ such that $s = s_1 \cdot s_2$.
+By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$, hence
+\begin{align*}
+H(\mathcal{L}(S_1 \cdot S_2))(s) &= -\sum_{s_1 \in S_1} \sum_{s_2 \in S_2}P_{S_1}(s_1)*P_{s_2}(s_2) \log_2(P_{S_1}(s_1)*P_{S_2}(s_2))\\
+&= -\sum_{s_1 \in S_1} P_{S_1}(s_1)\sum_{s_2 \in S_2}P_{s_2}(s_2) (\log_2(P_{S_1}(s_1) + \log_2 P_{S_2}(s_2)))\\
+&= -\sum_{s_1 \in S_1} P_{S_1}(s_1) \sum_{s_2 \in S_2}P_{S_2}(s_2)\log_2(P_{S_1}(s_1)) + \sum_{s_1 \in S_1} P_{S_1}(s_1)\sum_{s_2 \in S_2} P_{S_2}(s_2)\log_2 P_{S_2}(s_2)\\
+&= -\sum_{s_1 \in S_1} P_{S_1}(s_1) \log_2(P_{S_1}(s_1)) + \sum_{s_2 \in S_2} P_{S_2}(s_2)\log_2 P_{S_2}(s_2)\\
+&= H(\mathcal{L}(S_1)) + H(\mathcal{L}(S_2))\\
+&= \mathbb{H}(S_1) + \mathbb{H}(S_2)
+\end{align*}
+\item
+($S = S_1 \; |_p \; S_2$)
+
+Since $S$ is unambiguous, then $S_1$ and $S_2$ are also unambiguous, and every $s \in \mathcal{L}(S_1 \; |_p \; S_2)$ is either in $\mathcal{L}(S_1)$ or in $\mathcal{L}(S_2)$.
+By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$.
+Then
+\begin{align*}
+H(\mathcal{L}(S_1 \; |_p \; S_2)) &= -\sum_{s \in \mathcal{L}(S_1)}(p * P_{S_1}(s))\log_2 (p * P_{S_1}(s)) - \sum_{s \in \mathcal{L}(S_2)}((1-p) * P_{S_2}(s))\log_2 ((1-p) * P_{S_2}(s))\\
+&=- p\sum_{s \in \mathcal{L}(S_1)}P_{S_1}(s)(\log_2 p + \log_2 P_{S_1}(s)) - (1-p)\sum_{s \in \mathcal{L}(S_2)}P_{S_2}(s)(\log_2 (1-p) + \log_2 P_{S_2}(s))\\
+&= p * (\log_2 p +  H(\mathcal{L}(S_1))) + (1-p) * ( \log_2 (1-p) + H(\mathcal{L}(S_2)))\\
+&= p * (\log_2 p +  \mathbb{H}(S_1)) + (1-p) * ( \log_2 (1-p) + \mathbb{H}(S_2))\\
+\end{align*}
 \end{enumerate}

### 2018-07-04 papers/SymmetricLenses/main.tex
Finished proof of Theorem 7
Finished like a third of Theorem 3

@@
 \end{enumerate}
+\end{proof}
+\begin{theorem*}
+  Let $\Lens$ be an asymmetric lens. $\Lens$ is also a simple symmetric lens,
+  where
+  \begin{center}
+    \begin{tabular}{rcl}
+      $\Lens.\CreateROf{x}$ & $=$ & $\Lens.get \App x$\\
+      $\Lens.\CreateLOf{y}$ & $=$ & $\Lens.create \App y$\\
+      $\Lens.\PutROf{x}{y}$ & $=$ & $\Lens.get \App x$\\
+      $\Lens.\PutLOf{y}{x}$ & $=$ & $\Lens.put \App y \App x$
+    \end{tabular}
+  \end{center}
+\end{theorem*}
+\begin{proof}
+Let $\ell$ be an asymmetric lens. Then $\ell$ satisfies the following laws:
+\begin{align*}
+\ell.\get \; (\ell.\pput \; v \; s) &= v \tag{PUTGET}\\
+\ell.\pput \; (\ell.\get \; s) \; s&= s \tag{GETPUT}\\
+\ell.\get \; (\ell.\pput \; v \; s) &= v \tag{CREATGET}
+\end{align*}
+Define the simple symmetric lens $\lceil \ell \rceil$ by 
+\begin{align*}
+\lceil \ell \rceil.\CreateROf{s} &= \ell.\get \; s\\
+\lceil \ell \rceil.\CreateLOf{v} &= \ell.\create \; v\\
+\lceil \ell \rceil.\PutROf{s}{v} &= \ell.\get \; s\\
+\lceil \ell \rceil.\PutLOf{v}{s} &= \ell.\pput \; v \; s\\
+\end{align*}
+We now show that $\lceil \ell \rceil$ satisfies the simple symmetric lens laws:
+\begin{align*}
+\lceil \ell \rceil.\PutLOf{(\lceil \ell \rceil.\CreateROf{s})}{s} &= \ell.\pput \; (\ell.\get \; s) \; s = s\\
+\lceil \ell \rceil.\PutROf{(\lceil \ell \rceil.\CreateLOf{v})}{v} &= \ell.\get \; (\ell.\create \; v) = v\\
+\lceil \ell \rceil.\PutLOf{(\lceil \ell \rceil.\PutROf{s}{v})}{s} &= \ell.\pput \; (\ell.\get) \; s = s\\
+\lceil \ell \rceil.\PutROf{(\lceil \ell \rceil.\PutLOf{v}{s})}{v} &= \ell.\get \; (\ell.\pput \; v \; s) = v
+\end{align*}
+\end{proof}
+
+\begin{theorem*}
+  Let $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$, where $\Lens$ does not
+  include composition, $\Regex$ and $\RegexAlt$ and unambiguous, and neither
+  $\Regex$ nor $\RegexAlt$ contain any empty subcomponents.
+  \begin{enumerate}
+  \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ calculates the expected
+    information content of $\SetOf{s \SuchThat s \in \LanguageOf{\Regex}}$,
+    given $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt} \BooleanAnd
+      \Lens.\PutLOf{t}{s} = s}$
+  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ calculates the expected
+    information content of $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$,
+    given $\SetOf{s \SuchThat s \in \LanguageOf{\Regex} \BooleanAnd
+      \Lens.\PutROf{s}{t} = t}$
+  \end{enumerate}
+\end{theorem*}
+\begin{proof}
+Given $s \in S$, let $V_{\ell}(s) = \{t \in \mathcal{L}(T) \; | \; \ell.\PutLOf{t}{s} = s\}$. We want to show that
+$$\sum_{s \in \mathcal{L}(S)}P_{\mathcal{L}(S)}H(V_{\ell}(s)) = \mathbb{H}^{\leftarrow}(S \; | \; \ell, T)$$
+We proceed by induction over $\ell$. 
+\begin{center}
+  \begin{tabular}{l@{\ }l@{\ }l@{\ }l}
+            & & \GBar{} & $\ConcatLensOf{\Lens_1}{\Lens_2}$ \\
+            & & \GBar{} & $\SwapLensOf{\Lens_1}{\Lens_2}$ \\
+            & & \GBar{} & $\OrLensOf{\Lens_1}{\Lens_2}$ \\
+            & & \GBar{} & $\ComposeLensOf{\Lens_1}{\Lens_2}$ \\
+            & & \GBar{} & $\MergeROf{\Lens_1}{\Lens_2}$ \\
+            & & \GBar{} & $\MergeLOf{\Lens_1}{\Lens_2}$ \\
+            & & \GBar{} & $\InvertOf{\Lens}$ \\
+  \end{tabular}
+\end{center}
+\begin{enumerate}
+\item
+($\ell = \IdentityLensOf{S}$)
+
+Observe that $V_{\IdentityLensOf{S}}(s) = \{t \in \mathcal{L}(S) \; | \; \IdentityLensOf{S}.\PutLOf{t}{s} = s\} = \{t \in \mathcal{L}(S) \; | \; t = s\} = \{s\}$.
+
+Consequently, $H(V_{\IdentityLensOf{S}}(s)) = H(\{s\}) = 0$, from which it follows that 
+$$\sum_{s \in \mathcal{L}(S)} P_{\mathcal{L}(S)}(s) * H(V_{\IdentityLensOf{S}}(s)) = 0 = \mathbb{H}^{\leftarrow}(S \; | \; \IdentityLensOf{S}, T)$$
+\item
+($\ell = \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}$)
+
+Observe that 
+$$V_{\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}}(s) = \{t \in \mathcal{L}(T) \; | \; \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}.\PutLOf{t}{s} = s\} = \{t \in \mathcal{L}(T) \; | \; s = s\} = \mathcal{L}(T)$$
+Consequently
+$$\sum_{s \in \mathcal{L}(S)}P_{S}(s)H(V_{\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}}(s)) = \sum_{s \in \mathcal{L}(S)}P_{\mathcal{S}}(s)H(\mathcal{L}(T)) = H(\mathcal{L}(T)) = \mathbb{H}(T) = \mathbb{H}^{\leftarrow}(S \; | \; \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, T)$$
+\item
+($\ell = \IterateLensOf{\Lens} : S^{*p} \Leftrightarrow T^{*q}$)
+
+Observe that 
+\begin{align*}
+V_{\IterateLensOf{\Lens}}(s_1 \cdot \ldots \cdot s_n) &= \{t_1 \cdot \ldots \cdot t_m \; | \; \IterateLensOf{\Lens}.\PutLOf{(t_1 \cdot \ldots \cdot t_m)}{(s_1 \cdot \ldots \cdot s_n)} = s_1 \cdot \ldots \cdot s_n\}\\
+&= \{t_1 \cdot \ldots \cdot t_n \; | \; \ell.\PutLOf{t_i}{s_i} = s_i \text{ for }1 \leq i \leq n\}\\
+&= V_{\ell}(s_1) \cdot \ldots \cdot V_{\ell}(s_n)
+\end{align*}
+Consequently
+\begin{align*}
+\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V_{\IterateLensOf{\Lens}}(s))
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n)\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i) H(V_{\ell}(s_1) \cdot \ldots \cdot V_{\ell}(s_n))\\
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n)\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i) \left(\sum_{j=1}^n H(V_{\ell}(s_j)) \right)\\
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \sum_{s_n} P_S(s_n) \left(\sum_{j=1}^n H(V_{\ell}(s_j)) \right)
+\end{align*}
+We focus on the term $\sum_{s_n} P_S(s_n) \sum_{j=1}^n H(V_{\ell}(s_j))$:
+\begin{align*}
+\sum_{s_n} P_S(s_n) \sum_{j=1}^n H(V_{\ell}(s_j)) &= \sum_{s_n} P_S(s_n)H(V_{\ell}(s_n)) + P_S(s_n) \sum_{j=1}^{n-1} H(V_{\ell}(s_j))\\
+&= \mathbb{H}^{\leftarrow}(S \; | \; \ell, T) + \sum_{j=1}^{n-1}H(V_{\ell}(s_j))
+\end{align*}
+Consequently,
+\begin{align*}
+\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V_{\IterateLensOf{\Lens}}(s))
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left( \mathbb{H}^{\leftarrow}(S \; | \; \ell, T) + \sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)\\
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1})\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \mathbb{H}^{\leftarrow}(S \; | \; \ell, T)\\
+&+(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)\\
+&= \mathbb{H}^{\leftarrow}(S \; | \; \ell, T) * (1-p)\sum_n p^n \\
+&+(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)
+\end{align*}
+Unrolling the term 
+$$\sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)$$ 
+$(n-1)$ more times gives
+$$\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V_{\IterateLensOf{\Lens}}(s))
+= \mathbb{H}^{\leftarrow}(S \; | \; \ell, T) * (1-p)\sum_n n * p^n  = \frac{p}{1-p} * \mathbb{H}^{\leftarrow}(S \; | \; \ell, T)$$
+which is what we wanted to show.
+\end{enumerate}
+
 \end{proof}

### 2018-07-05 papers/SymmetricLenses/main.tex
Stuck at merge left

--- 2018-07-04 papers/SymmetricLenses/main.tex
@@
 recover a string in one data format from the other. We use the function
-$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate the expected amount
+$\REntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate the expected amount
 of information required to recover a string in $\Regex$ from a string in
@@
   \begin{tabular}{rcl}
-    $\LEntropyOf{\Regex \Given \IdentityLensOf{\Regex}, \Regex}$
+    $\REntropyOf{\Regex \Given \IdentityLensOf{\Regex}, \Regex}$
     & =
@@
 
-    $\LEntropyOf{\Regex \Given \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, \RegexAlt}$
-    & =
-    & \EntropyOf{\Regex}\\
-
-    $\LEntropyOf{\PRegexStar{\Regex}{\Probability} \Given \IterateLensOf{\Lens}^, \PRegexStar{\RegexAlt}{\ProbabilityAlt}}$
-    & =
-    & $\frac{\ProbabilityAlt}{1-\ProbabilityAlt}(\LEntropyOf{\Regex \Given \Lens, \RegexAlt})$\\
-
-    $\LEntropyOf{\Regex_1 \Concat \Regex_2 \Given \ConcatLensOf{\Lens_1}{\Lens_2}, \RegexAlt_1 \Concat \RegexAlt_2}$
-    & =
-    & $\LEntropyOf{\Regex_1 \Given \Lens_1^\leftarrow, \RegexAlt_1} + \LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2}$\\
-
-    $\LEntropyOf{\Regex_1 \Concat \Regex_2 \Given \SwapLensOf{\Lens_1}{\Lens_2}, \RegexAlt_2 \Concat \RegexAlt_1}$
-    & =
-    & $\LEntropyOf{\Regex_1 \Given \Lens_1^\leftarrow, \RegexAlt_1} + \LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2}$\\
-
-    $\LEntropyOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability} \Given \OrLensOf{\Lens_1}{\Lens_2}, \PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}}$
-    & =
-    & $\ProbabilityAlt*(\LEntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt_1}) + (1-\ProbabilityAlt)*(\LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2})$\\
-
-    $\LEntropyOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability} \Given \OrLensOf{\Lens_1}{\Lens_2}, \PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}}$
-    & =
-    & $\ProbabilityAlt*(\LEntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt_1}) + (1-\ProbabilityAlt)*(\LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2})$\\
-
-    $\LEntropyOf{\Regex \Given \MergeROf{\Lens_1}{\Lens_2}, \PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}}$
-    & =
-    & $\ProbabilityAlt*(\LEntropyOf{\Regex \Given \Lens_1, \RegexAlt_1}) + (1-\ProbabilityAlt)*(\LEntropyOf{\Regex \Given \Lens_2, \RegexAlt_2})$\\
-
-    $\LEntropyOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability} \Given \MergeLOf{\Lens_1}{\Lens_2}, \RegexAlt}$
-    & =
-    & $\Probability*(\LEntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt} - \Log_2(\Probability)) + $\\
+    $\REntropyOf{\Regex \Given \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, \RegexAlt}$
+    & =
+    & \EntropyOf{\RegexAlt}\\
+
+    $\REntropyOf{\PRegexStar{\Regex}{\Probability} \Given \IterateLensOf{\Lens}^, \PRegexStar{\RegexAlt}{\ProbabilityAlt}}$
+    & =
+    & $\frac{\Probability}{1-\Probability}(\REntropyOf{\Regex \Given \Lens, \RegexAlt})$\\
+
+    $\REntropyOf{\Regex_1 \Concat \Regex_2 \Given \ConcatLensOf{\Lens_1}{\Lens_2}, \RegexAlt_1 \Concat \RegexAlt_2}$
+    & =
+    & $\REntropyOf{\Regex_1 \Given \Lens_1^\leftarrow, \RegexAlt_1} + \REntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2}$\\
+
+    $\REntropyOf{\Regex_1 \Concat \Regex_2 \Given \SwapLensOf{\Lens_1}{\Lens_2}, \RegexAlt_2 \Concat \RegexAlt_1}$
+    & =
+    & $\REntropyOf{\Regex_1 \Given \Lens_1^\leftarrow, \RegexAlt_1} + \REntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2}$\\
+
+    $\REntropyOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability} \Given \OrLensOf{\Lens_1}{\Lens_2}, \PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}}$
+    & =
+    & $\Probability*(\REntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt_1} ) + (1-\Probability)*(\REntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2})$\\
+
+    $\REntropyOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability} \Given \MergeROf{\Lens_1}{\Lens_2}, \RegexAlt}$
+    & =
+    & $\Probability*(\REntropyOf{\Regex \Given \Lens_1, \RegexAlt_1}) + (1-\Probability)*(\REntropyOf{\Regex \Given \Lens_2, \RegexAlt_2})$\\
+
+    $\REntropyOf{\Regex \Given \MergeLOf{\Lens_1}{\Lens_2}, \PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}}$
+    & =
+    & $\ProbabilityAlt*(\REntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt} - \Log_2(\ProbabilityAlt)) + $\\
     &
-    & $(1-\Probability)*(\LEntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt} - \Log_2(1 - \Probability))$\\
-
-    $\LEntropyOf{\RegexAlt \Given \InvertOf{\Lens}, \Regex}$
+    & $(1-\ProbabilityAlt)*(\REntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt} - \Log_2(1 - \ProbabilityAlt))$\\
+
+    $\REntropyOf{\RegexAlt \Given \InvertOf{\Lens}, \Regex}$
     & =
@@
 \end{center}
-$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ is defined symmetrically.  These
+$\LEntropyOf{\RegexAlt \Given \Lens, \Regex}$ is defined symmetrically.  These
 functions calculate the expected number of bits to recover one data format from
@@
   \begin{enumerate}
-  \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ calculates the expected
+  \item $\REntropyOf{\Regex \Given \Lens, \RegexAlt}$ calculates the expected
     information content of $\SetOf{s \SuchThat s \in \LanguageOf{\Regex}}$,
@@
 \end{enumerate}
+\end{proof}
+\begin{theorem*}
+  Let $\Lens$ be an asymmetric lens. $\Lens$ is also a simple symmetric lens,
+  where
+  \begin{center}
+    \begin{tabular}{rcl}
+      $\Lens.\CreateROf{x}$ & $=$ & $\Lens.get \App x$\\
+      $\Lens.\CreateLOf{y}$ & $=$ & $\Lens.create \App y$\\
+      $\Lens.\PutROf{x}{y}$ & $=$ & $\Lens.get \App x$\\
+      $\Lens.\PutLOf{y}{x}$ & $=$ & $\Lens.put \App y \App x$
+    \end{tabular}
+  \end{center}
+\end{theorem*}
+\begin{proof}
+Let $\ell$ be an asymmetric lens. Then $\ell$ satisfies the following laws:
+\begin{align*}
+\ell.\get \; (\ell.\pput \; v \; s) &= v \tag{PUTGET}\\
+\ell.\pput \; (\ell.\get \; s) \; s&= s \tag{GETPUT}\\
+\ell.\get \; (\ell.\pput \; v \; s) &= v \tag{CREATGET}
+\end{align*}
+Define the simple symmetric lens $\lceil \ell \rceil$ by 
+\begin{align*}
+\lceil \ell \rceil.\CreateROf{s} &= \ell.\get \; s\\
+\lceil \ell \rceil.\CreateLOf{v} &= \ell.\create \; v\\
+\lceil \ell \rceil.\PutROf{s}{v} &= \ell.\get \; s\\
+\lceil \ell \rceil.\PutLOf{v}{s} &= \ell.\pput \; v \; s\\
+\end{align*}
+We now show that $\lceil \ell \rceil$ satisfies the simple symmetric lens laws:
+\begin{align*}
+\lceil \ell \rceil.\PutLOf{(\lceil \ell \rceil.\CreateROf{s})}{s} &= \ell.\pput \; (\ell.\get \; s) \; s = s\\
+\lceil \ell \rceil.\PutROf{(\lceil \ell \rceil.\CreateLOf{v})}{v} &= \ell.\get \; (\ell.\create \; v) = v\\
+\lceil \ell \rceil.\PutLOf{(\lceil \ell \rceil.\PutROf{s}{v})}{s} &= \ell.\pput \; (\ell.\get) \; s = s\\
+\lceil \ell \rceil.\PutROf{(\lceil \ell \rceil.\PutLOf{v}{s})}{v} &= \ell.\get \; (\ell.\pput \; v \; s) = v
+\end{align*}
+\end{proof}
+
+\begin{theorem*}
+  Let $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$, where $\Lens$ does not
+  include composition, $\Regex$ and $\RegexAlt$ and unambiguous, and neither
+  $\Regex$ nor $\RegexAlt$ contain any empty subcomponents.
+  \begin{enumerate}
+  \item $\REntropyOf{\Regex \Given \Lens, \RegexAlt}$ calculates the expected
+    information content of $\SetOf{s \SuchThat s \in \LanguageOf{\Regex}}$,
+    given $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt} \BooleanAnd
+      \Lens.\PutLOf{t}{s} = s}$
+  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ calculates the expected
+    information content of $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$,
+    given $\SetOf{s \SuchThat s \in \LanguageOf{\Regex} \BooleanAnd
+      \Lens.\PutROf{s}{t} = t}$
+  \end{enumerate}
+\end{theorem*}
+\begin{proof}
+Given $s \in S$, let $V_{\ell}(s) = \{t \in \mathcal{L}(T) \; | \; \ell.\PutROf{s}{t} = t\}$. We want to show that
+$$\sum_{s \in \mathcal{L}(S)}P_{\mathcal{L}(S)}(s)H(V_{\ell}(s)) = \mathbb{H}^{\rightarrow}(S \; | \; \ell, T)$$
+We proceed by induction over $\ell$. 
+\begin{enumerate}
+\item
+($\ell = \IdentityLensOf{S}$)
+
+Observe that $V_{\IdentityLensOf{S}}(s) = \{t \in \mathcal{L}(S) \; | \; \IdentityLensOf{S}.\ell.\PutROf{s}{t} = t\} = \{t \in \mathcal{L}(S) \; | \; t = s\} = \{s\}$.
+
+Consequently, $H(V_{\IdentityLensOf{S}}(s)) = H(\{s\}) = 0$, from which it follows that 
+$$\sum_{s \in \mathcal{L}(S)} P_{\mathcal{L}(S)}(s) * H(V_{\IdentityLensOf{S}}(s)) = 0 = \mathbb{H}^{\rightarrow}(S \; | \; \IdentityLensOf{S}, T)$$
+\item
+($\ell = \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}$)
+
+Observe that 
+$$V_{\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}}(s) = \{t \in \mathcal{L}(T) \; | \; \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}.\PutROf{s}{t} = t\} = \{t \in \mathcal{L}(T) \; | \; t = t\} = \mathcal{L}(T)$$
+Consequently
+$$\sum_{s \in \mathcal{L}(S)}P_{S}(s)H(V_{\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}}(s)) = \sum_{s \in \mathcal{L}(S)}P_{S}(s)H(\mathcal{L}(T)) = H(\mathcal{L}(T)) = \mathbb{H}(T) = \mathbb{H}^{\rightarrow}(S \; | \; \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, T)$$
+\item
+($\ell = \IterateLensOf{\Lens} : S^{*p} \Leftrightarrow T^{*q}$)
+
+Observe that 
+\begin{align*}
+V_{\IterateLensOf{\Lens}}(s_1 \cdot \ldots \cdot s_n) &= \{t_1 \cdot \ldots \cdot t_m \; | \; \IterateLensOf{\Lens}.\PutROf{(s_1 \cdot \ldots \cdot s_n) }{(t_1 \cdot \ldots \cdot t_m)} = t_1 \cdot \ldots \cdot t_m\}\\
+&= \{t_1 \cdot \ldots \cdot t_n \; | \; \ell.\PutROf{s_i}{t_i} = t_i \text{ for }1 \leq i \leq n\}\\
+&= V_{\ell}(s_1) \cdot \ldots \cdot V_{\ell}(s_n)
+\end{align*}
+Consequently
+\begin{align*}
+\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V_{\IterateLensOf{\Lens}}(s))
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n)\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i) H(V_{\ell}(s_1) \cdot \ldots \cdot V_{\ell}(s_n))\\
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n)\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i) \left(\sum_{j=1}^n H(V_{\ell}(s_j)) \right)\\
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \sum_{s_n} P_S(s_n) \left(\sum_{j=1}^n H(V_{\ell}(s_j)) \right)
+\end{align*}
+We focus on the term $\sum_{s_n} P_S(s_n) \sum_{j=1}^n H(V_{\ell}(s_j))$:
+\begin{align*}
+\sum_{s_n} P_S(s_n) \sum_{j=1}^n H(V_{\ell}(s_j)) &= \sum_{s_n} P_S(s_n)H(V_{\ell}(s_n)) + P_S(s_n) \sum_{j=1}^{n-1} H(V_{\ell}(s_j))\\
+&= \mathbb{H}^{\rightarrow}(S \; | \; \ell, T) + \sum_{j=1}^{n-1}H(V_{\ell}(s_j))
+\end{align*}
+Consequently,
+\begin{align*}
+\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V_{\IterateLensOf{\Lens}}(s))
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left( \mathbb{H}^{\rightarrow}(S \; | \; \ell, T) + \sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)\\
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1})\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \mathbb{H}^{\rightarrow}(S \; | \; \ell, T)\\
+&+(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)\\
+&= \mathbb{H}^{\rightarrow}(S \; | \; \ell, T) * (1-p)\sum_n p^n \\
+&+(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)
+\end{align*}
+Unrolling the term 
+$$\sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)$$ 
+$(n-1)$ more times gives
+$$\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V_{\IterateLensOf{\Lens}}(s))
+= \mathbb{H}^{\rightarrow}(S \; | \; \ell, T) * (1-p)\sum_n n * p^n  = \frac{p}{1-p} * \mathbb{H}^{\rightarrow}(S \; | \; \ell, T)$$
+which is what we wanted to show.
+\item
+($S = \ConcatLensOf{\Lens_1}{\Lens_2}$)
+
+Observe that 
+\begin{align*}
+V_{ \ConcatLensOf{\Lens_1}{\Lens_2}}(s_1 \cdot s_2) &= \{t_1 \cdot t_2 \in \mathcal{L}(T_1 \cdot T_2) \; | \;  \ConcatLensOf{\Lens_1}{\Lens_2}.\PutROf{(s_1 \cdot s_2)}{(t_1 \cdot t_2)} = t_1 \cdot t_2\}\\
+&= \{t_1 \cdot t_2 \in \mathcal{L}(T_1 \cdot T_2) \; | \;  \ell_1.\PutROf{s_1}{t_1 } = t_1  \text{ and } \ell_2.\PutROf{t_2}{t_2} = t_2\}\\
+&= V_{\ell_1}(s_1) \cdot V_{\ell_2}(s_2)
+\end{align*}
+Consequently,
+\begin{align*}
+&\sum_{s \in \mathcal{L}(S_1 \cdot S_2)}P_{S_1 \cdot S_2}(s) H(V_{\ConcatLensOf{\Lens_1}{\Lens_2}}(s))\\
+&= \sum_{s_1 \in \mathcal{L}(s_1)} P_{S_1}(s_1)\sum_{s_2 \in \mathcal{L}(s_2)}P_{S_2}(s_2)(H(V_{\ell_1}(s_1)) + H(V_{\ell_2}(s_2)))\\
+&= \sum_{s_1 \in \mathcal{L}(s_1)} P_{S_1}(s_1)\sum_{s_2 \in \mathcal{L}(s_2)}P_{S_2}(s_2)H(V_{\ell_1}(s_1)) + \sum_{s_1 \in \mathcal{L}(s_1)} P_{S_1}(s_1)\sum_{s_2 \in \mathcal{L}(s_2)}P_{S_2}(s_2)H(V_{\ell_2}(s_2))\\
+&= \mathbb{H}^{\rightarrow}(S_1 \; | \; \ell_1, T_1) + \mathbb{H}^{\rightarrow}(S_2 \; | \; \ell_2, T_2)
+\end{align*}
+\item
+($S = \SwapLensOf{\Lens_1}{\Lens_2}$)
+
+Similar to the previous case.
+\item
+($S = \OrLensOf{\Lens_1}{\Lens_2}$)
+
+Observe that 
+$$
+V_{ \OrLensOf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \OrLensOf{\Lens_1}{\Lens_2}.\ell.\PutROf{s}{t} = t\}
+$$
+If $s \in \mathcal{L}(S_1)$, then
+$$
+V_{ \OrLensOf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \ell_1.\PutROf{s}{t} = t\} = V_{\ell_1}(s)
+$$
+while if $s \in \mathcal{L}(S_2)$, then
+$$
+V_{ \OrLensOf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \ell_2.\PutROf{s}{t} = t\} = V_{\ell_2}(s)
+$$
+Consequently,
+\begin{align*}
+&\sum_{s \in \mathcal{L}(S_1 \; |_p \; S_2)}P_{S_1 \; |_p \; S_2}(s) H(V_{\OrLensOf{\Lens_1}{\Lens_2}}(s))\\
+&= p * \sum_{s \in \mathcal{L}(S_1)}P_{S_1}(s) H(V_{\ell_1}(s)) + (1-p) * \sum_{s \in \mathcal{L}(S_2)}P_{S_2}(s) H(V_{\ell_2}(s))\\
+&= p * (\mathbb{H}^{\rightarrow}(S_1 \; | \; \ell_1,T_1)) + (1-p) * (\mathbb{H}^{\rightarrow}(S_2 \; | \; \ell_2,T_2))
+\end{align*}
+\item
+($S = \MergeROf{\Lens_1}{\Lens_2}$)
+
+Observe that 
+$$
+V_{ \MergeROf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T) \; | \;  \MergeROf{\Lens_1}{\Lens_2}.\ell.\PutROf{s}{t} = t\}
+$$
+If $s \in \mathcal{L}(S_1)$, then
+$$
+V_{ \MergeROf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T) \; | \;  \ell_1.\PutROf{s}{t} = t\} = V_{\ell_1}(s)
+$$
+while if $s \in \mathcal{L}(S_2)$, then
+$$
+V_{ \MergeROf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T) \; | \;  \ell_2.\PutROf{s}{t} = t\} = V_{\ell_2}(s)
+$$
+Consequently,
+\begin{align*}
+&\sum_{s \in \mathcal{L}(S_1 \; |_p \; S_2)}P_{S_1 \; |_p \; S_2}(s) H(V_{\MergeROf{\Lens_1}{\Lens_2}}(s))\\
+&= p * \sum_{s \in \mathcal{L}(S_1)}P_{S_1}(s) H(V_{\ell_1}(s)) + (1-p) * \sum_{s \in \mathcal{L}(S_2)}P_{S_2}(s) H(V_{\ell_2}(s))\\
+&= p * (\mathbb{H}^{\rightarrow}(S_1 \; | \; \ell_1,T_1) ) + (1-p) * (\mathbb{H}^{\rightarrow}(S_2 \; | \; \ell_2,T_2) )
+\end{align*}
+\item
+($S = \MergeLOf{\Lens_1}{\Lens_2}$)
+
+Observe that 
+\begin{align*}
+V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s) &= \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \MergeLOf{\Lens_1}{\Lens_2}.\PutROf{s}{t} = t\}\\
+&= \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \ell_1.\PutROf{s}{t} = t \text{ or } \ell_2.\PutROf{s}{t} = t\}\\
+&= V_{\ell_1}(s) \cup V_{\ell_2}(s)
+\end{align*}
+First we consider $H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))$:
+\begin{align*}
+H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)) &= \sum_{t \in V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}(t) \log_2 P_{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}(t)\\
+&= \sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \\
+&+ \sum_{t \in \mathcal{L}(T_2) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+\end{align*}
+We focus on the term 
+$$\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}$$
+Observe that
+\begin{align*}
+&\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
+&= \frac{q}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t) \log_2 q * P_{T_1}(t)\\
+&- \frac{q * \log_2 P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t)
+\end{align*}
+Again we focus on the term
+$$\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t) \log_2 q * P_{T_1}(t)$$
+Observe that
+\begin{align*}
+\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t) \log_2 q * P_{T_1}(t) &= \log_2 q * P_{T_1}(V_{\ell_1}(s)) + H(V_{\ell_1}(s))
+\end{align*}
+Hence
+\begin{align*}
+&\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \\
+&= \frac{q}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t) \log_2 q * P_{T_1}(t)\\
+&- \frac{q * \log_2 P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t)\\
+&= \frac{q \left[P_{T_1}(V_{\ell_1}(s)) \left[\log_2 \frac{q}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \right] + H(V_{\ell_1}(s))\right]}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+\end{align*}
+Similarly,
+\begin{align*}
+&\sum_{t \in \mathcal{L}(T_2) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
+&= \frac{(1-q)\left[P_{T_2}(V_{\ell_2}(s)) \left[\log_2 \frac{(1-q)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] + H(V_{\ell_2}(s))\right]}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+\end{align*}
+\end{enumerate}
+
 \end{proof}

### 2018-07-05 papers/SymmetricLenses/main.tex
Found error in Theorem 3 but still stuck

@@
 $$
-V_{ \MergeROf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T) \; | \;  \MergeROf{\Lens_1}{\Lens_2}.\ell.\PutROf{s}{t} = t\}
+V_{ \MergeROf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T) \; | \;  \MergeROf{\Lens_1}{\Lens_2}.\PutROf{s}{t} = t\}
 $$
@@
 \begin{align*}
-H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)) &= \sum_{t \in V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}(t) \log_2 P_{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}(t)\\
+-H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)) &= \sum_{t \in V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}(t) \log_2 P_{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}(t)\\
 &= \sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \\
@@
 \begin{align*}
-\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t) \log_2 q * P_{T_1}(t) &= \log_2 q * P_{T_1}(V_{\ell_1}(s)) + H(V_{\ell_1}(s))
+&\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t) \log_2 q * P_{T_1}(t) = \log_2 q * P_{T_1}(V_{\ell_1}(s))\\
+& + P_{T_1}(V_{\ell_1}(s))\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{P_{T_1}(t)}{P_{T_1}(V_{\ell_1}(s))} \log_2 \left( P_{T_1}(V_{\ell_1}(s)) *  \frac{P_{T_1}(t)}{P_{T_1}(V_{\ell_1}(s))} \right)\\
+&= \log_2 q * P_{T_1}(V_{\ell_1}(s)) + \log_2 P_{T_1}(V_{\ell_1}(s)) * P_{T_1}(V_{\ell_1}(s)) + P_{T_1}(V_{\ell_1}(s))*H(V_{\ell_1}(s))\\
+&= P_{T_1}(V_{\ell_1}(s))[\log_2 q + \log_2 P_{T_1}(V_{\ell_1}(s)) + H(V_{\ell_1}(s))]
 \end{align*}
@@
 &- \frac{q * \log_2 P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t)\\
-&= \frac{q \left[P_{T_1}(V_{\ell_1}(s)) \left[\log_2 \frac{q}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \right] + H(V_{\ell_1}(s))\right]}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+&= \frac{q * P_{T_1}(V_{\ell_1}(s))[\log_2 q + \log_2 P_{T_1}(V_{\ell_1}(s)) + H(V_{\ell_1}(s))]}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
+&- \frac{q * \log_2 P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)) * P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
+&= \frac{q * P_{T_1}(V_{\ell_1}(s)) \left[\log_2 q + \log_2 \left[\frac{P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] + H(V_{\ell_1}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
 \end{align*}
@@
 &\sum_{t \in \mathcal{L}(T_2) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
-&= \frac{(1-q)\left[P_{T_2}(V_{\ell_2}(s)) \left[\log_2 \frac{(1-q)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] + H(V_{\ell_2}(s))\right]}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+&= \frac{(1-q) * P_{T_2}(V_{\ell_2}(s)) \left[\log_2 (1-q) + \log_2 \left[\frac{P_{T_2}(V_{\ell_2}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] + H(V_{\ell_2}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+\end{align*}
+\begin{align*}
+&-H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))\\
+&= \frac{q * P_{T_1}(V_{\ell_1}(s)) \left[\log_2 q + \log_2 \left[\frac{P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] + H(V_{\ell_1}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
+&+ \frac{(1-q) * P_{T_2}(V_{\ell_2}(s)) \left[\log_2 (1-q) + \log_2 \left[\frac{P_{T_2}(V_{\ell_2}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] + H(V_{\ell_2}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
 \end{align*}

### 2018-07-05 papers/SymmetricLenses/main.tex
Fixed a sign
add in typo fix

@@
       \Lens.\PutLOf{t}{s} = s}$
-  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ calculates the expected
+  \item $\LEntropyOf{\RegexAlt \Given \Lens, \Regex}$ calculates the expected
     information content of $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$,
@@
 & + P_{T_1}(V_{\ell_1}(s))\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{P_{T_1}(t)}{P_{T_1}(V_{\ell_1}(s))} \log_2 \left( P_{T_1}(V_{\ell_1}(s)) *  \frac{P_{T_1}(t)}{P_{T_1}(V_{\ell_1}(s))} \right)\\
-&= \log_2 q * P_{T_1}(V_{\ell_1}(s)) + \log_2 P_{T_1}(V_{\ell_1}(s)) * P_{T_1}(V_{\ell_1}(s)) + P_{T_1}(V_{\ell_1}(s))*H(V_{\ell_1}(s))\\
-&= P_{T_1}(V_{\ell_1}(s))[\log_2 q + \log_2 P_{T_1}(V_{\ell_1}(s)) + H(V_{\ell_1}(s))]
+&= \log_2 q * P_{T_1}(V_{\ell_1}(s)) + \log_2 P_{T_1}(V_{\ell_1}(s)) * P_{T_1}(V_{\ell_1}(s)) + P_{T_1}(V_{\ell_1}(s))* -H(V_{\ell_1}(s))\\
+&= P_{T_1}(V_{\ell_1}(s))[\log_2 q + \log_2 P_{T_1}(V_{\ell_1}(s)) - H(V_{\ell_1}(s))]
 \end{align*}
@@
 &- \frac{q * \log_2 P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t)\\
-&= \frac{q * P_{T_1}(V_{\ell_1}(s))[\log_2 q + \log_2 P_{T_1}(V_{\ell_1}(s)) + H(V_{\ell_1}(s))]}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
+&= \frac{q * P_{T_1}(V_{\ell_1}(s))[\log_2 q + \log_2 P_{T_1}(V_{\ell_1}(s)) - H(V_{\ell_1}(s))]}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
 &- \frac{q * \log_2 P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)) * P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
-&= \frac{q * P_{T_1}(V_{\ell_1}(s)) \left[\log_2 q + \log_2 \left[\frac{P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] + H(V_{\ell_1}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+&= \frac{q * P_{T_1}(V_{\ell_1}(s)) \left[\log_2 q + \log_2 \left[\frac{P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right]  -H(V_{\ell_1}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
 \end{align*}
@@
 &\sum_{t \in \mathcal{L}(T_2) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
-&= \frac{(1-q) * P_{T_2}(V_{\ell_2}(s)) \left[\log_2 (1-q) + \log_2 \left[\frac{P_{T_2}(V_{\ell_2}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] + H(V_{\ell_2}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+&= \frac{(1-q) * P_{T_2}(V_{\ell_2}(s)) \left[\log_2 (1-q) + \log_2 \left[\frac{P_{T_2}(V_{\ell_2}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] - H(V_{\ell_2}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
 \end{align*}
@@
 &-H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))\\
-&= \frac{q * P_{T_1}(V_{\ell_1}(s)) \left[\log_2 q + \log_2 \left[\frac{P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] + H(V_{\ell_1}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
-&+ \frac{(1-q) * P_{T_2}(V_{\ell_2}(s)) \left[\log_2 (1-q) + \log_2 \left[\frac{P_{T_2}(V_{\ell_2}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] + H(V_{\ell_2}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+&= \frac{q * P_{T_1}(V_{\ell_1}(s)) \left[\log_2 q + \log_2 \left[\frac{P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] - H(V_{\ell_1}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
+&+ \frac{(1-q) * P_{T_2}(V_{\ell_2}(s)) \left[\log_2 (1-q) + \log_2 \left[\frac{P_{T_2}(V_{\ell_2}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] - H(V_{\ell_2}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
 \end{align*}

### 2018-07-05 papers/SymmetricLenses/main.tex
fix theorem statement

@@
 recover a string in one data format from the other. We use the function
-$\REntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate the expected amount
-of information required to recover a string in $\Regex$ from a string in
-$\RegexAlt$, synchronized by $\Lens$. We use the function
 $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to calculate the expected amount
 of information required to recover a string in $\RegexAlt$ from a string in
-$\Regex$, synchronized by $\Lens$.
+$\Regex$, synchronized by $\Lens$. We use the function $\LEntropyOf{\Regex
+  \Given \Lens, \RegexAlt}$ to calculate the expected amount of information
+required to recover a string in $\Regex$ from a string in $\RegexAlt$,
+synchronized by $\Lens$.
 \begin{center}
   \begin{tabular}{rcl}
-    $\REntropyOf{\Regex \Given \IdentityLensOf{\Regex}, \Regex}$
+    $\REntropyOf{\RegexAlt \Given \IdentityLensOf{\RegexAlt}, \RegexAlt}$
     & =
@@
 
-    $\REntropyOf{\Regex \Given \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, \RegexAlt}$
+    $\REntropyOf{\RegexAlt \Given \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, \Regex}$
     & =
@@
 
-    $\REntropyOf{\PRegexStar{\Regex}{\Probability} \Given \IterateLensOf{\Lens}^, \PRegexStar{\RegexAlt}{\ProbabilityAlt}}$
-    & =
-    & $\frac{\Probability}{1-\Probability}(\REntropyOf{\Regex \Given \Lens, \RegexAlt})$\\
-
-    $\REntropyOf{\Regex_1 \Concat \Regex_2 \Given \ConcatLensOf{\Lens_1}{\Lens_2}, \RegexAlt_1 \Concat \RegexAlt_2}$
-    & =
-    & $\REntropyOf{\Regex_1 \Given \Lens_1^\leftarrow, \RegexAlt_1} + \REntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2}$\\
-
-    $\REntropyOf{\Regex_1 \Concat \Regex_2 \Given \SwapLensOf{\Lens_1}{\Lens_2}, \RegexAlt_2 \Concat \RegexAlt_1}$
-    & =
-    & $\REntropyOf{\Regex_1 \Given \Lens_1^\leftarrow, \RegexAlt_1} + \REntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2}$\\
-
-    $\REntropyOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability} \Given \OrLensOf{\Lens_1}{\Lens_2}, \PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}}$
+    $\REntropyOf{\PRegexStar{\RegexAlt}{\ProbabilityAlt} \Given \IterateLensOf{\Lens}, \PRegexStar{\Regex}{\Probability}}$
+    & =
+    & $\frac{\Probability}{1-\Probability}(\REntropyOf{\RegexAlt \Given \Lens, \Regex})$\\
+
+    $\REntropyOf{\RegexAlt_1 \Concat \RegexAlt_2 \Given \ConcatLensOf{\Lens_1}{\Lens_2}, \Regex_1 \Concat \Regex_2}$
+    & =
+    & $\REntropyOf{\RegexAlt_1 \Given \Lens_1, \Regex_1} + \REntropyOf{\RegexAlt_2 \Given \Lens_2, \Regex_2}$\\
+
+    $\REntropyOf{\RegexAlt_2 \Concat \RegexAlt_1 \Given \SwapLensOf{\Lens_1}{\Lens_2}, \Regex_1 \Concat \Regex_2}$
+    & =
+    & $\REntropyOf{\RegexAlt_2 \Given \Lens_1, \Regex_2} + \REntropyOf{\RegexAlt_1 \Given \Lens_1, \Regex_1}$\\
+
+    $\REntropyOf{\PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt} \Given \OrLensOf{\Lens_1}{\Lens_2}, \PRegexOr{\Regex_1}{\Regex_2}{\Probability}}$
     & =
@@
 
-    $\REntropyOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability} \Given \MergeROf{\Lens_1}{\Lens_2}, \RegexAlt}$
-    & =
-    & $\Probability*(\REntropyOf{\Regex \Given \Lens_1, \RegexAlt_1}) + (1-\Probability)*(\REntropyOf{\Regex \Given \Lens_2, \RegexAlt_2})$\\
-
-    $\REntropyOf{\Regex \Given \MergeLOf{\Lens_1}{\Lens_2}, \PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}}$
-    & =
-    & $\ProbabilityAlt*(\REntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt} - \Log_2(\ProbabilityAlt)) + $\\
+    $\REntropyOf{\PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt} \Given \MergeLOf{\Lens_1}{\Lens_2}, \Regex}$
+    & =
+    & $\ProbabilityAlt*(\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1}) + (1-\ProbabilityAlt)*(\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2})$\\
+
+    $\REntropyOf{\RegexAlt \Given \MergeROf{\Lens_1}{\Lens_2}, \PRegexOr{\Regex_1}{\Regex_2}{\Probability}}$
+    & =
+    & $\Probability*(\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1} - \Log_2(\Probability)) + $\\
     &
-    & $(1-\ProbabilityAlt)*(\REntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt} - \Log_2(1 - \ProbabilityAlt))$\\
-
-    $\REntropyOf{\RegexAlt \Given \InvertOf{\Lens}, \Regex}$
-    & =
-    & $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$\\
+    & $(1-\Probability)*(\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2} - \Log_2(1 - \Probability))$\\
+
+    $\REntropyOf{\Regex \Given \InvertOf{\Lens}, \RegexAlt}$
+    & =
+    & $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$\\
   \end{tabular}
 \end{center}
-$\LEntropyOf{\RegexAlt \Given \Lens, \Regex}$ is defined symmetrically.  These
+$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ is defined symmetrically.  These
 functions calculate the expected number of bits to recover one data format from
@@
   \begin{enumerate}
-  \item $\REntropyOf{\Regex \Given \Lens, \RegexAlt}$ calculates the expected
+  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ calculates the expected
+    information content of $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$,
+    given $\SetOf{s \SuchThat s \in \LanguageOf{\Regex} \BooleanAnd
+      \Lens.\PutROf{s}{t} = t}$
+  \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ calculates the expected
     information content of $\SetOf{s \SuchThat s \in \LanguageOf{\Regex}}$,
@@
       \Lens.\PutLOf{t}{s} = s}$
-  \item $\LEntropyOf{\RegexAlt \Given \Lens, \Regex}$ calculates the expected
-    information content of $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$,
-    given $\SetOf{s \SuchThat s \in \LanguageOf{\Regex} \BooleanAnd
-      \Lens.\PutROf{s}{t} = t}$
   \end{enumerate}

### 2018-07-06 papers/SymmetricLenses/main.tex
add main

--- 2018-07-05 papers/SymmetricLenses/main.tex
@@
   \FCOpt{}: &  The first time \GreedySynth finds a lens, it returns that lens.\\
-  \NMOpt{}: & \GreedySynth does not have any notion of cost, and searches for
-              any lens that satisfies the specification. \\
+  \NMCCOpt{}: & The lens cluster that contains is identified through \SSOpt{}. \GreedySynth is then called on that pair of SREs. \\
+  \NMOpt{}: & Neither \GreedySynth nor \RXSearch have any notion of cost; they
+              both merely search for
+              any lens that satisfies the specification. 
 \end{tabulary}
 
-Figure~\ref{fig:heuristic} summarizes the result of this experiment.  We find
-that \afm{TODO}.
+\begin{figure}
+  \includegraphics{generated-graphs/metrics_importance}
+  \caption{Number of benchmarks that synthesize the correct lens by a given
+    algorithm. \NoTPOpt{} is the full symmetric synthesis algorithm, but
+    termination parameters are ignored. \FCOpt{} provides a notion of cost to
+    \GreedySynth, but once a satisfying lens is greedily found, that lens is
+    returned. \NMCCOpt{} identifies the lens cluster in which the desired 
+    lens lies.  \GreedySynth{} is then run on that cluster with no notion of
+    cost. \NMOpt{} removes all notion of cost, and merely finds the first
+    satisfying lens.}
+  \label{fig:metric}
+\end{figure}
+
+Figure~\ref{fig:metric} summarizes the result of this experiment. We find that
+our metrics are very important for finding the correct lens. Our
+information-theoretic metric is critical for finding the correct lens. Indeed,
+even when provided with the correct lens cluster, without a notion of cost, we
+would need constrain our search space more with additional examples to find the
+correct lens.
+
+Our termination condition is also quite important. When running in \FCOpt{},
+only 5 lenses are found. The first cluster in which a lens is found is rarely
+the correct cluster. However, our termination condition is not perfect.
+Sometimes users must tell the algorithm to ``keep going'' or ``stop early.''
+Without termination parameters, our algorithm finds the correct lens for 36 of
+our 45 benchmarks, 9 of them need help in the form of a termination parameter
+for finding the correct lens.
 % end evaluation

### 2018-07-06 papers/QuotientLenses/main.tex
Added acknowledgement
A bit of polishing

@@
 (where . is concatenation), but this regular expression is ambiguous, as \lstinline|anychar| is a big union of a of single characters including comma.
-
-While the unambiguity constraints consequently appear to compromise compositionality of QREs, they can usually be circumvented by making a small changes to the offending regular expression: for instance, in the preceding example, then we need to write,
+Instead, in this case, we need to write
 \begin{lstlisting}
@@
 from FlashFill~\cite{flashfill}, and benchmarks derived from
-Augeas~\cite{augeas2} (we also experimented using our QRE implementation to synthesize quotient lenses between XML, RDF and and JSON formats using data from the data.gov database; the data consisted of census statistics, demographic statistics, wage comparosion data, and crime index data).
+Augeas~\cite{augeas2} (we also experimented using our QRE implementation to synthesize quotient lenses between XML, RDF and and JSON formats using data from the data.gov database; the data consisted of census statistics, demographic statistics, wage comparison data, and crime index data).
 In these 39 benchmarks, 10 of
@@
 \end{center}
-where \lstinline|nil| and \lstinline|space| pretty-print \lstinline|""| and \lstinline|" "| respectively, but represent zero-or-more and one-or-more whitespaces in parsing.
-FliPpr also allows annotations for optional parentheses in programs.
-These features enable a user to easily define a pretty-printer that also recognizes ugly strings that get mapped to a canonical representative modulo whitespace and parentheses quotienting.
+where \lstinline|nil| and \lstinline|space| pretty-print \lstinline|""| and
+\lstinline|" "| respectively, but represent zero-or-more and one-or-more
+whitespaces in parsing. FliPpr also allows annotations for optional
+parentheses in programs. These features enable a user to easily define a
+pretty-printer and a parser that also recognizes ``ugly'' strings; after a
+round trip through the parser and printer, such 
+strings will be mapped to a canonical representative modulo whitespace and parentheses quotienting.
 
@@
 Theorem~\ref{normal form}.
+
+\begin{acks}
+  We thank the anonymous reviewers for their useful feedback and discussions and Nate Foster for extensive assistance in integrating
+  QRE-enhanced \QOpt{} into Boomerang.  This research has been supported in part by DARPA
+  award FA8750-17-2-0028 and ONR 568751 (SynCrypt).
+\end{acks}
 

### 2018-07-07 papers/SymmetricLenses/main.tex
add new theorem

--- 2018-07-06 papers/SymmetricLenses/main.tex
@@
 recover a string in one data format from the other. We use the function
-$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to calculate the expected amount
-of information required to recover a string in $\RegexAlt$ from a string in
+$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to bound the expected amount of
+information required to recover a string in $\RegexAlt$ from a string in
 $\Regex$, synchronized by $\Lens$. We use the function $\LEntropyOf{\Regex
-  \Given \Lens, \RegexAlt}$ to calculate the expected amount of information
-required to recover a string in $\Regex$ from a string in $\RegexAlt$,
-synchronized by $\Lens$.
+  \Given \Lens, \RegexAlt}$ to bound the expected amount of information required
+to recover a string in $\Regex$ from a string in $\RegexAlt$, synchronized by
+$\Lens$.
 \begin{center}
@@
     & =
-    & 0\\
+    & [0,0]\\
 
@@
     & =
-    & \EntropyOf{\RegexAlt}\\
+    & [\EntropyOf{\RegexAlt},\EntropyOf{\RegexAlt}]\\
 
@@
 
+    $\REntropyOf{\RegexAlt \Given \MergeROf{\Lens_1}{\Lens_2}, \PRegexOr{\Regex_1}{\Regex_2}{\Probability}}$
+    & =
+    & $\Probability*(\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1}) + (1-\Probability)*(\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2})$\\
+
     $\REntropyOf{\PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt} \Given \MergeLOf{\Lens_1}{\Lens_2}, \Regex}$
     & =
-    & $\ProbabilityAlt*(\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1}) + (1-\ProbabilityAlt)*(\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2})$\\
-
-    $\REntropyOf{\RegexAlt \Given \MergeROf{\Lens_1}{\Lens_2}, \PRegexOr{\Regex_1}{\Regex_2}{\Probability}}$
-    & =
-    & $\Probability*(\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1} - \Log_2(\Probability)) + $\\
+    & $\ProbabilityAlt*(\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1}) + $\\
     &
-    & $(1-\Probability)*(\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2} - \Log_2(1 - \Probability))$\\
+    & $(1-\ProbabilityAlt)*(\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2} - \Log_2(1 - \ProbabilityAlt))+$\\
+    &
+    & $(0, -\Log_2(\ProbabilityAlt)*\ProbabilityAlt -\Log_2(1 - \ProbabilityAlt)*(1-\ProbabilityAlt)]$\\
 
@@
 $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ is defined symmetrically. These
-functions calculate the expected number of bits to recover one data format from
-a synchronized string in the other format.
+functions bound the expected number of bits to recover one data format from a
+synchronized string in the other format. Notice, we would be able to calculate
+the exact entropy, were it not for \MergeL. If $\MergeLOf{\Lens_1}{\Lens_2}
+\OfType \Regex \Leftrightarrow
+\PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}$, given a string in $s$, we
+need to determine if the synchronized string is in $\Regex$ or $\RegexAlt$.
+However, this information content is dependent on the how likely the
+synchronized string is to be in $\Regex$ or $\RegexAlt$. In the worst case, the
+expected information content to encode this choice is
+$-\Log_2(\ProbabilityAlt)*\ProbabilityAlt -\Log_2(1 -
+\ProbabilityAlt)*(1-\ProbabilityAlt)$, but the expected information content
+could be arbitrarily small.
+
 The cost of a lens, $c(\Lens \OfType \Regex \Leftrightarrow \RegexAlt) =
-\LEntropyOf{\Regex \Given \Lens, \RegexAlt} + \REntropyOf{\RegexAlt \Given
-  \Lens, \Regex}$ is the
-sum of recovering the left format from the right, and the right from the left.
-We have proven theorems demonstrating the calculated entropy corresponds to the
-actual conditional entropy to recover the data.\bcp{Again, they should be
+\MaxOf{\LEntropyOf{\Regex \Given \Lens, \RegexAlt}} +
+\MaxOf{\REntropyOf{\RegexAlt \Given \Lens, \Regex}}$ is the maximum sum of
+recovering the left format from the right, and the right from the left. We have
+proven theorems demonstrating the calculated entropy corresponds to the actual
+conditional entropy to recover the data.\bcp{Again, the proofs should be
   included.}
@@
   \begin{enumerate}
-  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ calculates the expected
+  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ bounds the expected
     information content of $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$,
@@
       \Lens.\PutROf{s}{t} = t}$
-  \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ calculates the expected
+  \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ bounds the expected
     information content of $\SetOf{s \SuchThat s \in \LanguageOf{\Regex}}$,

### 2018-07-07 papers/QuotientLenses/main.tex

--- 2018-07-02 papers/QuotientLenses/main.tex
@@
   QRE lenses to the problem of synthesizing bijective string lenses, which was
-  previously studied in past work~\cite{optician} (Section~\ref{synth}).
+  previously studied in past work~\cite{miltner2017synthesizing} (Section~\ref{synth}).
   \item We extend Boomerang with QREs and QRE lens synthesis and demonstrate
@@
 Theorem~\ref{normal form}.
+
+\begin{acks}
+  We thank our anonymous reviewers for their useful feedback and discussions and Nate Foster for his extensive assistance in integrating
+  QRE-enhanced \QOpt{} into Boomerang.  This research has been supported in part by DARPA
+  award FA8750-17-2-0028 and ONR 568751 (SynCrypt).
+\end{acks}
 

### 2018-07-07 papers/SymmetricLenses/main.tex
Still stuck... urgh

@@
 \end{align*}
-First we consider $H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))$:
+To compute $H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))$, we use the formula 
+$$H(P_1, P_2) = H(m(P_1), m(P_2)) + m(P_1)H\left(\frac{P_1}{m(P_1)}\right) + m(P_2)H\left(\frac{P_2}{m(P_2)}\right)$$
+if the probability space $P$ is the sum of measure spaces $P_1$ and $P_2$, and $m(P_i)$ is the measure of $P_i$:
+\begin{align*}
+&H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)) = H \left(\frac{q * P_{T_1}V_{\ell_1}(s)}{P_{T_1 \; |_q \; T_2}{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}}, \frac{(1-q) * P_{T_2}V_{\ell_2}(s)}{P_{T_1 \; |_q \; T_2}{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}}\right)\\
+&= \frac{q * P_{T_1}V_{\ell_1}(s)}{P_{T_1 \; |_q \; T_2}{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}} H(V_{\ell_1}(s)) + \frac{(1-q) * P_{T_2}V_{\ell_2}(s)}{P_{T_1 \; |_q \; T_2}{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}} H(V_{\ell_2}(s))
+\end{align*}
+
+\iffalse
 \begin{align*}
@@
 \end{align*}
+Consequently,
 \begin{align*}
 &-H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))\\
-&= \frac{q * P_{T_1}(V_{\ell_1}(s)) \left[\log_2 q + \log_2 \left[\frac{P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] - H(V_{\ell_1}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
-&+ \frac{(1-q) * P_{T_2}(V_{\ell_2}(s)) \left[\log_2 (1-q) + \log_2 \left[\frac{P_{T_2}(V_{\ell_2}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] - H(V_{\ell_2}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+&= \frac{q * P_{T_1}(V_{\ell_1}(s)) \left[\log_2 \left[\frac{q * P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] - H(V_{\ell_1}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
+&+ \frac{(1-q) * P_{T_2}(V_{\ell_2}(s)) \left[\log_2 \left[\frac{(1-q) * P_{T_2}(V_{\ell_2}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] - H(V_{\ell_2}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
 \end{align*}
+\fi
 \end{enumerate}

### 2018-07-07 papers/SymmetricLenses/main.tex
blah

--- 2018-07-07 papers/SymmetricLenses/main.tex
@@
 commentstyle=\itshape\rmfamily,
-emph={rep,iterate,synth,collapse,perm,squash,normalize,using,ins,del,lens,let,get,put,rquot,lquot,id,swap,concat,or,disc,merge_left,merge_right,const},
+emph={rep,iterate,synth,collapse,perm,squash,normalize,using,ins,del,lens,let,get,put,rquot,lquot,id,swap,concat,or,disconnect,merge_left,merge_right,const},
 emphstyle=\upshape\sffamily\color{dkpurple}, 
@@
 \noindent
-In a big-data world, one will often find similar data stored in a
-variety of different formats. For instance, electronic calendars come in iCalendar, vCalendar and h-event
-formats~\cite{calendar-formats}; US taxation data can be transmitted via Tax XML (used by the US government
-for e-filing) or TXF (used by TurboTax)~\cite{tax-formats}; and sewing machines can read embroidery files
-transmitted as PES formats, DST formats or several others~\cite{embroidery-formats}. \dpw{Editorial comment:
-Phil Wadler forced his very irritating long-form citations into the
-standard format. ug. how stupid is that.}
-\ksf{I like the embroidery example and I also hate the long form citations.}
+In a big-data world, one will often find similar data stored in a variety of
+different formats. For instance, electronic calendars come in iCalendar,
+vCalendar and h-event formats~\cite{calendar-formats}; US taxation data can be
+transmitted via Tax XML (used by the US government for e-filing) or TXF (used by
+TurboTax)~\cite{tax-formats}; and sewing machines can read embroidery files
+transmitted as PES formats, DST formats or several
+others~\cite{embroidery-formats}.
 
@@
 %a search through the space of possible computations that can be expressed as a lens.
-Recent research~\cite{optician, maina+:quotient-synthesis} has exploited this fact by demonstrating that it
-is possible to synthesize relatively large, complex lenses from specifications.
-In particular, given types that describe source and target formats, there are
-synthesis algorithms\bcp{more than one?  I know there are two papers, but
-  it's really the same synthesis algorithm} that will find bijective
-lenses
+Recent research~\cite{optician, maina+:quotient-synthesis} has exploited this
+fact by demonstrating that it is possible to synthesize relatively large,
+complex lenses from specifications. In particular, given types that describe
+source and target formats, the bijective synthesis algorithm presented in this
+research will find bijective lenses
 % ~\cite{optician}
-for files 
-in ad-hoc formats, such as those housing linux application configurations.
-Oftentimes, these algorithms find such lenses in less than a second, and without
-the aid of any examples at all, with just types describing the source and target
-formats. 
+for files in ad-hoc formats, such as those housing linux application
+configurations. Oftentimes, these algorithms find such lenses in less than a
+second, and without the aid of any examples at all, with just types describing
+the source and target formats.
 % \dpw{I would cite the ICFP paper (cite string:
@@
 principles for identifying the ``likely'' ones, and we need new algorithms to
-search the space of program transformations using these principles. \dpw{Were
-  the bijective lenses chosen using a ``smaller'' criteria? I forget... if not
-  edit the prior paragraph.}\afm{no, chosen using a ``first'' criteria.
-  Edited.}\dpw{Ideally, we could back up the claim that there are many symmetric
-  lenses and not many bijective lenses by graphing the number of such lenses
-  somehow. eg, in a past synthesis paper with Steve (POPL 2016) we graphed the
-  untyped terms of type $\mathsf{int}\rightarrow\mathsf{int}$. I wonder if we
-  can do something similar here.}\afm{What about lenses between ``a''* and
-  ''b``*}\bcp{Aren't there an infinite number of those, even if we restrict
-  to bijective ones?}
+search the space of program transformations using these principles.
 
@@
 target is ``unlikely'' to require very much additional information in order to
-reconstruct it---if it were\bcp{were what? likely, or unlikely?}, then one is
-unlikely to be attempting to synthesize a lens between the two formats\bcp{why?
-  even if two big formats only share a little bit of common information,
-  mightn't one want a lens between them?}! In order to implement this intuition,
-we turn to \emph{stochastic regular expressions} (SREs). Stochastic regular
-expressions define a set of strings, as do ordinary regular expressions, but
-they also provide the expected distribution of strings within that set. Such
+reconstruct it, under the assumption that users typically want to synchronize
+formats with a large amount of shared information. In order to implement this
+intuition, we turn to \emph{stochastic regular expressions} (SREs). Stochastic
+regular expressions define a set of strings, as do ordinary regular expressions,
+but they also provide the expected distribution of strings within that set. Such
 distributions allow us to compute the amount of information gained when
@@
 $\lstinline{"a"} |_{0.5}
-\lstinline{"b"}$ contains two elements (the strings
-\lstinline{"a"} and
-\lstinline{"b"}). The probability 0.5 annotating the union indicates that both
-elements are equally likely to appear. Consequently, if we observe an element of
-that language (either \lstinline{"a"} or \lstinline{"b"}), we will gain one bit
-of information. Said another way, one bit, together with the SRE, suffices to
-recover the string in question.\bcp{Is this implying that, if the annotation
-  were different, we'd need something other than one bit to recover the
-  string??}
-\afm{From an information theoretic perspective, yes, because that perspective
-  views it as the amount of information if we were communicating a stream of
-  these choices.}\bcp{I see---very interesting.  What about mentioning this
-  point, to get people ready for some information-theoretic thinking
-  later...?} 
-On the other hand, consider the language of the SRE \lstinline{c}. It contains 
-just one element (the string $c$). If we observe that our algorithm generates
-$\lstinline{"c"}$, we gain zero bits of information --- the string was already determined.
-Hence\bcp{?}, a lens is ``likely,'' in a technical sense, if relative to the
-target SRE $T$, very little information from $t$ is required to generate it.
+\lstinline{"b"}$ contains two elements (the strings \lstinline{"a"} and \lstinline{"b"}).
+The probability 0.5 annotating the union indicates that both elements are
+equally likely to appear. Consequently, if we observe an element of that
+language (either \lstinline{"a"} or
+\lstinline{"b"}), we will gain one bit of information. Said another way, one bit, together with
+the SRE, suffices to recover the string in question. Interestingly, if the
+probability annotation were anything but 0.5, the expected information content
+(or entropy) would actually be less than 1 (given a stream of such choices, on
+average we could encode the the contents of stream in fewer bits than the the
+number of choices present). On the other hand, consider the language of the SRE
+\lstinline{c}. It contains just one element (the string $c$). If we observe that
+our algorithm generates $\lstinline{"c"}$, we gain zero bits of information ---
+the string was already determined. Using this foundation, a lens $\Lens \OfType
+\Regex \Leftrightarrow \RegexAlt$ is ``likely,''
+in a technical sense, if given a string $s$ in the source $S$, very little information
+is needed to determine which synchronized string $t$ is present in the target
+$T$.
 
@@
 
-We implemented this\bcp{which?} algorithm and evaluated its effectiveness on a
-range of benchmarks, including data cleaning tasks taken from Flash Fill,
-information extraction tasks taken from Augeas, and ad-hoc format
-synchronization tasks taken from a variety of locations\footnote{The majority of
-  these format synchronization tasks were found from the following index:
+We implemented the \GreedySynth and \RXSearch algorithms and evaluated their
+effectiveness of synthesizing synchronization functions on a range of
+benchmarks, including data cleaning tasks taken from Flash Fill, information
+extraction tasks taken from Augeas, and ad-hoc format synchronization tasks
+taken from a variety of locations\footnote{The majority of these format
+  synchronization tasks were found from the following index:
   \url{http://fileformats.archiveteam.org/wiki/Electronic_File_Formats}}.
@@
   empirically and argue that simple ones suffice for many applications.
-\item We show that the class of simple lenses can be characterized
-  semantically as the subset of full symmetric lenses for which a simple
-  ``forgetfulness'' property holds (\S\ref{sec:related}).  \bcp{I suggest
-    splitting this out into its own section, since there is other related
-    work that we'll want to discuss in a Related Work section.}
+\item We show that the class of simple lenses can be characterized semantically
+  as the subset of full symmetric lenses for which a simple ``forgetfulness''
+  property holds (\S\ref{sec:relationship}).
 \end{itemize}
@@
 \begin{lstlisting}
-let salary = number | "N/A"
+let salary = number | "unk"
 let emp_salary = name . " " . name . ": " salary
@@
   \resizebox{\columnwidth}{!}{%
-    \begin{tabular}{|>{\columncolor{lightyellow}}c|>{\columncolor{lightgreen}}c|}
+    \begin{tabular}{|>{\columncolor{vlightyellow}}c|>{\columncolor{vlightgreen}}c|}
       \hline
@@
       \hline
-      \cellcolor{vlightyellow}\texttt{company.mgmt} regular expression & 
-      \cellcolor{vlightgreen}\texttt{company.hr}  regular expression\\
+      \cellcolor{lightyellow}\texttt{company.mgmt} regular expression & 
+      \cellcolor{lightgreen}\texttt{company.hr}  regular expression\\
       \hline
-      \cellcolor{vlightyellow}\listingC & \cellcolor{vlightgreen}\listingD \\
+      \cellcolor{lightyellow}\listingC & \cellcolor{lightgreen}\listingD \\
       \hline
@@
   is \lstinline{emp_salaries} written with an or instead of just using
-  the star operator? Can we simplify?} }  
+  the star operator? Can we simplify?}\afm{It does require the or operator, if
+  we don't want it to have trailing or preceding newlines.  Howevever, I
+  intentionally had this as not having those to force a discussion about
+  rewrites and searching through rewrites.} }  
   \label{fig:minimized-representations}
@@
 
-In this section, we give an overview of simple symmetric lenses and
-our corresponding synthesis algorithm using an example drawn from a
-hypothetical U.S. company.  In this company, management and
-human resources (HR) store information about employees in separate files:
-management records the names of employees and their salaries while
-HR stores the names of employees and their health
-insurance providers.  Figure~\ref{fig:minimized-representations} gives
-examples of the two types of files and regular expressions describing
-the two formats.
-The company uses simple symmetric lenses to keep these files synchronized.
-If management adds a new employee, ``Chris Roe: 32500'',  a lens adds 
-the corresponding entry ``Chris Roe, UNK'' to HR's file.  The
-sentinel value ``UNK'' represents the fact that the employee's
-insurance company is currently unknown.  A similar update happens if
-HR adds a new employee before management, in which case the sentinel
-value ``N/A'' reflects the unknown salary information\bcp{can we just use
-  ``UNKNOWN'' on both sides?  Or is there some reason to make it more
-  complicated?} 
-Furthermore, if HR
-corrects an error in an employee's name, say changing ``John Public''
-to ``Jon Public'', the lens mirrors that name change in management's
-file. Not all the data is mirrored, however. The management file 
-is not updated in response to insurance changes, and the HR file
-is oblivious to salary changes.  Simple symmetric lenses are
-appropriate for keeping these two files synchronized because the files have
-a mix of shared and unshared information.
+In this section, we give an overview of simple symmetric lenses and our
+corresponding synthesis algorithm using an example drawn from a hypothetical
+U.S. company. In this company, management and human resources (HR) store
+information about employees in separate files: management records the names of
+employees and their salaries while HR stores the names of employees and their
+health insurance providers. Figure~\ref{fig:minimized-representations} gives
+examples of the two types of files and regular expressions describing the two
+formats. The company uses simple symmetric lenses to keep these files
+synchronized. If management adds a new employee, ``Chris Roe: 32500'', a lens
+adds the corresponding entry ``Chris Roe, UNK'' to HR's file. The sentinel value
+``UNK'' represents the fact that the employee's insurance company is currently
+unknown. A similar update happens if HR adds a new employee before management,
+in which case the sentinel value ``unk'' reflects the unknown salary information
+Furthermore, if HR corrects an error in an employee's name, say changing ``John
+Public'' to ``Jon Public'', the lens mirrors that name change in management's
+file. Not all the data is mirrored, however. The management file is not updated
+in response to insurance changes, and the HR file is oblivious to salary
+changes. Simple symmetric lenses are appropriate for keeping these two files
+synchronized because the files have a mix of shared and unshared information.
 
@@
   make one pretty easily by tweaking the one from the original symmetric paper.}
+\afm{That would be awesome.}
 
@@
 The two create functions provide default values for times when
-data appears on one side but not the other (\EG, the ``N/A'' salary
+data appears on one side but not the other (\EG, the ``unk'' salary
 entry added to the management file when HR inserts a new employee).
@@
 when a \lstinline{name} is updated in one file, the identify lens ensures it is
-also updated in the other file.
+also updated in the other file. In this manner, both the \CreateR and \CreateL
+functions are the identity function ($\CreateR \App s = s$), and the put
+functions merely return the first argument ($\PutROf{s_1}{s_2} = s_2$).
 
@@
 In contrast to the identity lens, the disconnect lens, \Disconnect, leaves data
-unconnected.\bcp{Can we call it ``disconnect''?} The disconnect lens takes four arguments: two regular 
-expressions and two strings. The regular expressions specify the
-formats of the corresponding data on the two sides while the strings provide default
-values.  In our extended example, 
-the \lstinline{salary} and \lstinline{company} fields are independent,
-so we use the disconnect lens to relate them.
-The defaults are strings that indicate the information is missing.
+unconnected. The disconnect lens takes four arguments: two regular expressions
+and two strings. The regular expressions specify the formats of the
+corresponding data on the two sides while the strings provide default values.
+On creates, the input values are thrown away, and default values are returned ($\CreateROf{\String} =
+\lstinline{"UNK"}$), and on puts, the second argument is used, and the first is
+thrown away ($\PutROf{\String}{\StringAlt} = \StringAlt$).  In
+our extended example, the \lstinline{salary} and
+\lstinline{company} fields are independent, so we use the disconnect lens to relate them. The defaults are strings that indicate the information is missing.
 %
 \begin{lstlisting}
-disc(salary, company, "N/A", "UNK") : salary $\Leftrightarrow$ company
+disconnect(salary, company, "unk", "UNK") : salary $\Leftrightarrow$ company
 \end{lstlisting}
@@
 constant is omitted entirely from the source and target formats,
-respectively. \ksf{Are these in the right order?}
+respectively.
 \begin{lstlisting}
- ins($t$) = disc("", $t$, "", $t$)
- del($s$) = disc($s$, "", $s$, "")
+ ins($t$) = disconnect("", $t$, "", $t$)
+ del($s$) = disconnect($s$, "", $s$, "")
 \end{lstlisting}
 
-Finally, there are a number of standard lenses that compositionally
-construct more complex lenses from simpler ones.  These combinators include 
-\lstinline{concat} for concatenation, 
-\lstinline{or} for variation, and
-\IterateLens for sequences.
-We usually write \lstinline{concat} using an infix
-notation \lstinline{$\Lens_1$ . $\Lens_2$}
-as in:
+Finally, there are a number of standard lenses that compositionally construct
+more complex lenses from simpler ones. These combinators include concatenation
+($\ConcatLensOf{\Lens_1}{\Lens_2}$ or $\Lens_1 . \Lens_2$), variation
+($\OrLensOf{\Lens_1}{\Lens_2}$), and iteration ($\IterateLensOf{\Lens}$).  For example;
 %
@@
 %
-which appears in the definition of the \lstinline{name_lens} in our
-running example. 
+appears in the definition of the \lstinline{name_lens} in our running example.
 %\ksf{I don't think this example is necessary.}
@@
 let name_lens = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
-let employee_lens = name_lens . disc(salary,company,"N/A","UNK") in
+let employee_lens = name_lens . disconnect(salary,company,"unk","UNK") in
 let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
@@
 
-We provide typing rules for each of the simple symmetric combinators,
-most of which are syntax directed.  There is one additional
-typing rule, which allows us to use equivalence relations on regular
-expressions to convert the type of any lens to a semantically equal
-but syntactically different pair of regular expressions.  The flexibility this
-type equivalence rule provides is useful when a different
-representation of a regular expression would facilitate synthesis.
+We provide typing rules for each of the simple symmetric combinators, most of
+which are syntax directed. However, regular expressions have a number of
+equivalences on them, and sometimes we must convert one regular expression type
+to another, equivalent regular expression type. Our typing rules include type
+equivalence, so if a lens is typed between two regular expressions, it is also
+typed between two equivalent ones. The flexibility this type equivalence rule
+provides is useful when a different representation of a regular expression would
+facilitate synthesis.
 
@@
 
-Note that, while the synchronization task for our running example is quite simple, the
-requisite lens is relatively complex. Furthermore, this complexity is
-inherent in manually writing simple symmetric lenses because the
-lenses must precisely describe the portions of the data that should be
-connected and those that should be disconnected.  For disconnected
-pieces, the lens must further specifiy appropriate defaults.
-
 \subsection{Synthesizing Symmetric Lenses}
 
-To alleviate the difficulty of manually writing simple symmetric
-lenses, we wish to synthesize them instead. Given 
-a pair of regular expressions and a set of input-output examples, we want to
-find a simple symmetric lens that is typed by the regular expression
-pair and satisfies the input/output examples. We call such a lens
-a \emph{satisfying lens}.
-In our running example, we wish to
-synthesize a lens between \lstinline{emp_salaries} and \lstinline{emp_insurance},
-using as an input-output example the data in
-Figure~\ref{fig:minimized-representations}.  The algorithm uses such
-examples to find suitable defaults for \Disconnect lenses and to
-figure out how to match data that could be synchronised in multiple
-ways.
+While our example is relatively simple for demonstration, writing such lenses
+for real-world formats is quite difficult, as these formats can be large and
+transform data in complex and unintuitive ways. To alleviate the difficulty of
+manually writing simple symmetric lenses, we wish to synthesize them instead.
+Given a pair of regular expressions and a set of input-output examples, we want
+to find a simple symmetric lens that is typed by the regular expression pair and
+satisfies the input/output examples. We call such a lens a \emph{satisfying
+  lens}. In our running example, we wish to synthesize a lens between
+\lstinline{emp_salaries} and \lstinline{emp_insurance}, using as an input-output
+example the data in Figure~\ref{fig:minimized-representations}. The algorithm
+uses such examples to find suitable defaults for \Disconnect lenses and to
+figure out how to match data that could be synchronised in multiple ways.
 
@@
 \begin{lstlisting}
-disc(emp_salaries,emp_insurance,rep(emp_salaries),rep(emp_insurance))
+disconnect(emp_salaries,emp_insurance,rep(emp_salaries),rep(emp_insurance))
 \end{lstlisting}
@@
 %and is highlighted even more on larger formats.
-While
-we could ask programmers to provide detailed and precise specifications for their
-synthesis tasks, we would prefer the algorithm to choose the right lens
-automatically. Instead of merely finding \emph{any} satisfying lens, we wish to
-synthesize a satisfying lens that is likely to please the user.
+While we could ask programmers to provide detailed and precise specifications
+for their synthesis tasks, we would prefer an algorithm that can intelligently
+choose a lens despite the underspecification. Instead of merely finding
+\emph{any} satisfying lens, we wish to synthesize a satisfying lens that is
+likely to please the user.
 
@@
 distributions over that data.
-To that end, we adopt \emph{stochastic regular expressions}\ksf{is
-there a citation?}, 
+To that end, we adopt \emph{stochastic regular expressions}\cite{?}, 
 which are regular expressions where each operator is annotated with a
@@
 
-To obtain stochastic regular expressions from the plain regular
-expressions that we expect users to supply, we use a heurstic 
-that attempts to assign
-probability annotations that give each string in the language equal
-probablity.  We can also take as input a stochastic regular
-expression rather than a plain one, which allows the user to
-specify a data-specific probability distribution.  Alternatively,
-the user could provide a large suite of example data from which we
-could infer the necessary annotations to model the supplied data.
+To obtain stochastic regular expressions from the plain regular expressions that
+we expect users to supply, we use a heurstic that attempts to assign probability
+annotations that give each string in the language equal probablity. We can also
+take as input a stochastic regular expression rather than a plain one, which
+allows the user to specify a data-specific probability distribution.
+Alternatively, the user could provide a large suite of example data from which
+we could infer the necessary annotations to model the supplied data. We found
+our heuristically chosen distribution works well in practice.
 
@@
 it.
-Previous work on synthesis~\ksf{add citations} has shown
+Previous work on synthesis~\cite{?,?,?} has shown
 that we can use types to dramatically restrict the search space and
@@
 % begin related-work
-\section{Related Work}
-\label{sec:related}
-There is related work in a number of directions. There is related work in
-symmetric lenses, string transformation synthesis, and type-directed synthesis.
-We also refer the reader to a recent, general survey on synthesis~\ref{?}.
-
-\subsection{Symmetric Lenses}
+\section{Relationship To Classic Symmetric Lenses}
+\label{sec:relationship}
 \bcp{I suggest adding an adjective---e.g., ``classic symmetric lenses''---to
@@
 \end{theorem}
+
+\section{Related Work}
+\label{sec:related}
+There is related work in a number of directions. There is related work in string
+transformation synthesis and type-directed synthesis. We also refer the reader
+to a recent, general survey on synthesis~\ref{?}.
 

### 2018-07-07 papers/SymmetricLenses/main.tex
add some edits

@@
 \label{sec:srx}
-Important to our algorithm is finding the expected number of bits needed to
-recover a string in one data source from a synchronized string in the other data
-source. To do this, we must first develop a probability model for our language.
-We do this with \emph{stochastic regular expressions}, regular expressions
-annotated with probability information. With this, we can jointly express a
-language and a probability distribution over that language.
+To characterize the likely lenses, we must compute the expected number of bits
+needed to recover a string in one data source from a synchronized string in the
+other data source. To do this, we must first develop a probabilistic langauge
+model for our language. We do this with \emph{stochastic regular expressions},
+regular expressions annotated with probability information. With this, we can
+jointly express a language and a probability distribution over that language.
 
@@
     & \GEq{} & $\String$ \\
-    & \GBar{} & $\emptyset$ \\
     & \GBar{} & $\PRegexStar{\Regex}{p}$ \\
@@
 
-The language is defined like the regular expression without probability
-annotations would be.  The probabilities are defined as follows:
+In the stochastic regular expression
+$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation \Probability
+represents the likelihood the string is in $\Regex_1$. In the stochastic regular
+expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability how
+much more likely having a string involving more than $n$ iterations is than
+having a string involving $n$ iterations. The semantics of SREs is defined by
+the function $P$ below.
+
 \begin{center}
@@
 
-    $\ProbabilityOf{\emptyset}{\String}$
-    & =
-    & $0$ \\
-
     $\ProbabilityOf{\RegexConcat{\Regex_1}{\Regex_2}}{s}$
@@
 \end{center}
-In the stochastic regular expression
-$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation \Probability
-represents the likelihood the string is in $\Regex_1$. In the stochastic regular
-expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability
-how much more likely having a string involving more than $n$ iterations is than
-having a string involving $n$ iterations.
 
@@
   \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
-    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
-    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
-    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
     \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
@@
 \end{theorem}
-
-\bcp{This is a POPL paper: Proofs of important theorems should
-be included.  (This one might be easy enough that we can just label it as a
-``Fact'' and omit the proof, or better yet we can say that the proof is straightforward.}
+\begin{proof}
+  The proofs of this theorem, and subsequent theorems in this paper, appears in
+  the auxiliary materials accompanying this submission.
+\end{proof}
 
 \subsection{Stochastic Regular Expression Entropy}
-The {\em entropy} of a stochastic regular expression can be computed
-directly from its syntax, when the stochastic regular expression is
-unambiguous and contains only nonempty subcomponents.\bcp{What is an empty
-  subcomponent?} 
+The entropy of a data source $S$ is $-\Sigma_{s\in S}P(s)\Log_2P(s)$, the
+expected number of bits required to describe a random element in $S$. The
+entropy of a stochastic regular expression can be computed directly from its
+syntax, when the stochastic regular expression is unambiguous.
 \begin{center}
@@
 \end{center}
-We have proven\bcp{Again, some kind of proof is needed.} that this syntactic
-computation corresponds to the entropy of the stochastic regular expression.
+We have proven that this syntactic computation corresponds to the entropy of the
+stochastic regular expression.
 

### 2018-07-07 papers/SymmetricLenses/main.tex
make disconnect change
make some changes

@@
 
-In contrast to the identity lens, the disconnect lens, \Disconnect, leaves data
+In contrast to the identity lens, the \Disconnect lens, leaves data
 unconnected. The disconnect lens takes four arguments: two regular expressions
@@
     & \GEq{} & $\String$ \\
+    & \GBar{} & $\emptyset$ \\
     & \GBar{} & $\PRegexStar{\Regex}{p}$ \\
@@
 
+    $\ProbabilityOf{\emptyset}{\String}$
+    & =
+    & $0$ \\
+
     $\ProbabilityOf{\RegexConcat{\Regex_1}{\Regex_2}}{s}$
@@
   \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
+    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
+    $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
+    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
     \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
@@
 \subsection{Stochastic Regular Expression Entropy}
-The entropy of a data source $S$ is $-\Sigma_{s\in S}P(s)\Log_2P(s)$, the
-expected number of bits required to describe a random element in $S$. The
-entropy of a stochastic regular expression can be computed directly from its
-syntax, when the stochastic regular expression is unambiguous.
+The entropy of a data source $S$ is the expected number of bits required to
+describe a random element in $S$ ($-\Sigma_{s\in S}P(s)\Log_2P(s)$). The entropy
+of a stochastic regular expression can be computed directly from its syntax,
+when the stochastic regular expression is unambiguous, and contains no nonempty
+components.
 \begin{center}
@@
   \label{thm:correct_entropy}
-  If $\Regex$ is unambiguous and contains no empty subcomponents,
+  If $\Regex$ is unambiguous and does not contain $\emptyset$ as a subterm,
   $\EntropyOf{\Regex}$ is the entropy of $\Regex$.
 \end{theorem}
+
+Without unambiguity constraints it is not easy to calculate the entropy. For
+example, consider the SRE $\PRegexOr{\lstinline{"a"}}{\lstinline{"a"}}{.5}$.
+Using the above computation, this SRE has an entropy of $1$, but the entropy
+should be $0$. Similar issues occur when trying to find the entropy when
+$\emptyset$ is a subterm, so we do not define the entropy on $\emptyset$.
+
+In our work, we do not find this to be a problem, we require unambiguous regular
+expressions as input to our synthesis procedure, and we can easily preprocess
+emptysets out of the stochastic regular expressions that are themselves nonempty
+by traversing the correct star-semiring equivalences
+(\PRegexOr{\emptyset}{\lstinline{"s"}}{.5} \SSREquiv \lstinline{"s"}).
 

### 2018-07-07 papers/SymmetricLenses/main.tex
nits

@@
 our algorithm generates $\lstinline{"c"}$, we gain zero bits of information ---
-the string was already determined. Using this foundation, a lens $\Lens \OfType
-\Regex \Leftrightarrow \RegexAlt$ is ``likely,''
+the string was already determined. Using this foundation, a lens $\Lens$ that
+synchronizes data between $\Regex$ and $\RegexAlt$ is ``likely,''
 in a technical sense, if given a string $s$ in the source $S$, very little information
@@
 \subsection{Simple Symmetric Lenses}
-A simple symmetric lens between formats described by types $\Regex$
+A simple symmetric lens between formats described by types $\BRegex$
 and $\RegexAlt$ 

### 2018-07-08 papers/SymmetricLenses/main.tex
add lotta stuff

@@
 \subsection{Simple Symmetric Lenses}
-A simple symmetric lens between formats described by types $\BRegex$
-and $\RegexAlt$ 
-is defined by four functions\bcp{I think a picture would be helpful.  I can probably
-  make one pretty easily by tweaking the one from the original symmetric paper.}
-\afm{That would be awesome.}
+A simple symmetric lens between formats described by regular expression types
+$\BRegex$ and $\BRegexAlt$\footnote{We use nonterminals \BRegex and \BRegexAlt
+  for regular expressions, using \Regex and \RegexAlt for stochastic regular
+  expressions.} is defined by four functions.\bcp{I think a picture would be
+  helpful. I can probably make one pretty easily by tweaking the one from the
+  original symmetric paper.} \afm{That would be awesome.}
 
@@
  \tag{1}
- \mbox{$\CreateR \OfType \Regex \rightarrow \RegexAlt$}\\
+ \mbox{$\CreateR \OfType \BRegex \rightarrow \BRegexAlt$}\\
  \tag{2}
- \mbox{$\CreateL \OfType \RegexAlt \rightarrow \Regex$}\\
+ \mbox{$\CreateL \OfType \BRegexAlt \rightarrow \BRegex$}\\
  \tag{3}
- \mbox{$\PutR \OfType \Regex \rightarrow \RegexAlt \rightarrow \RegexAlt$}\\
+ \mbox{$\PutR \OfType \BRegex \rightarrow \BRegexAlt \rightarrow \BRegexAlt$}\\
  \tag{4}
- \mbox{$\PutL \OfType \RegexAlt \rightarrow \Regex \rightarrow \Regex$}
+ \mbox{$\PutL \OfType \BRegexAlt \rightarrow \BRegex \rightarrow \BRegex$}
 \end{gather}
@@
 impose a typing discipline on our lenses.  Specifically, if 
-$\Regex$ and $\RegexAlt$ are regular expressions, then the notation
-\lstinline{$\Lens$ : $\Regex \Leftrightarrow \RegexAlt$} indicates that $\Lens$
-is a simple symmetric lens between $\LanguageOf{\Regex}$ and
-$\LanguageOf{\RegexAlt}$. 
+$\BRegex$ and $\BRegexAlt$ are regular expressions, then the notation
+\lstinline{$\Lens$ : $\BRegex \Leftrightarrow \BRegexAlt$} indicates that $\Lens$
+is a simple symmetric lens between $\LanguageOf{\BRegex}$ and
+$\LanguageOf{\BRegexAlt}$. 
 When clear from context, we use the term \textit{lens} to denote a
@@
 The simplest combinator is the identity lens \IdentityLens, which takes as an
-argument a regular expression $\Regex$ and performs an identity mapping on data
-matching $\Regex$. The identity lens fully connects the
+argument a regular expression $\BRegex$ and performs an identity mapping on data
+matching $\BRegex$. The identity lens fully connects the
 corresponding data in the two formats: when one format is
@@
 %
-is another valid lens (where \lstinline{rep($\Regex$)} is a string in the
-language of $\Regex$). Furthermore, adding the input-output example in
+is another valid lens (where \lstinline{rep($\BRegex$)} is a string in the
+language of $\BRegex$). Furthermore, adding the input-output example in
 Figure~\ref{fig:minimized-representations} 
@@
 distributions over that data.
-To that end, we adopt \emph{stochastic regular expressions}\cite{?}, 
+To that end, we adopt \emph{stochastic regular expressions}\cite{?} (SREs), 
 which are regular expressions where each operator is annotated with a
@@
 distribution over it.  From this information, we can calculate the
-entropy of the stochastic regular expression.
-
-This infrastructure gives us what we need to find likely
-satisfying lenses from a collection of satisfying lenses.  All such
-lenses must synchronize data between the same two 
-stochastic regular expressions $\Regex$ and $\RegexAlt$.
-For any such lens $\Lens$, we can calculate the expected number of
-bits required 
-to recover a string $s$ in $\LanguageOf{\Regex}$ from a synchronized
-string $t$ in $\LanguageOf{\RegexAlt}$. 
-This expected number of bits is just the conditional entropy of
-$\Regex$ given $\RegexAlt$ and $\Lens$. 
-To ensure symmetry, the likelihood we assign to $\Lens$ is the sum of 
-the conditional entropy of $\Regex$ given $\RegexAlt$ and $\Lens$ and
-the conditional entropy of $\RegexAlt$ given $\Regex$ and $\Lens$.
-
-To obtain stochastic regular expressions from the plain regular expressions that
+entropy of the SRE.
+
+This infrastructure gives us what we need to find likely satisfying lenses from
+a collection of satisfying lenses. All such lenses must synchronize data between
+the same two SREs $\Regex$ and $\RegexAlt$. For any such lens $\Lens$, we can
+calculate the expected number of bits required to recover a string $s$ in
+$\LanguageOf{\Regex}$ from a synchronized string $t$ in
+$\LanguageOf{\RegexAlt}$. This expected number of bits is just the conditional
+entropy of $\Regex$ given $\RegexAlt$ and $\Lens$. To ensure symmetry, the
+likelihood we assign to $\Lens$ is the sum of the conditional entropy of
+$\Regex$ given $\RegexAlt$ and $\Lens$ and the conditional entropy of
+$\RegexAlt$ given $\Regex$ and $\Lens$.
+
+To obtain SREs from the plain regular expressions that
 we expect users to supply, we use a heurstic that attempts to assign probability
@@
 regular expressions annotated with probability information. With this, we can
-jointly express a language and a probability distribution over that language.
+jointly express a language and a probability distribution over that language. In
+the following grammar, $\String$ ranges over strings, and $\Probability$ ranges
+over real numbers between 0 and 1, exclusive.
 
@@
 In the stochastic regular expression
-$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation \Probability
+$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation $\Probability$
 represents the likelihood the string is in $\Regex_1$. In the stochastic regular
@@
 language. These lens terms are typed between pairs of stochastic regular
-expressions: the typing judgment $\Lens \OfType \Regex \Leftrightarrow
-\RegexAlt$ means that $\Lens$ is a well typed lens between $\LanguageOf{\Regex}$
-and $\LanguageOf{\RegexAlt}$. These lens terms can also be typed between regular
-expressions, but we present in terms of stochastic regular expressions to assist
-in describing the entropy of the lenses.
+expressions: the typing judgment $\Lens \OfType \BRegex \Leftrightarrow
+\BRegexAlt$ means that $\Lens$ is a well typed lens between
+$\LanguageOf{\BRegex}$ and $\LanguageOf{\BRegexAlt}$. While we are performing
+synthesis on stochastic regular expressions, lenses are typed between regular
+expressions.  We can recover a regular expression from a stochastic regular
+expression by deleting probability annotations.  In a slight abuse of notation,
+if $\Regex$ is a stochastic regular expression, $\BRegex$ is the regular
+expression obtained by removing probability annotations.
 
@@
     % REGEX
-    \Lens{} & \GEq{} & & $\IdentityLensOf{\Regex}$\\
-            & & \GBar{} & $\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}$ \\
+    \Lens{} & \GEq{} & & $\IdentityLensOf{\BRegex}$\\
+            & & \GBar{} & $\DisconnectOf{\BRegex}{\BRegexAlt}{\String}{\StringAlt}$ \\
             & & \GBar{} & $\IterateLensOf{\Lens}$ \\
@@
   {
-    \IdentityLensOf{\Regex} \OfType \Regex \Leftrightarrow \Regex
+    \IdentityLensOf{\BRegex} \OfType \BRegex \Leftrightarrow \BRegex
   }
@@
   {
-    \String \in \LanguageOf{\Regex}\\
-    \StringAlt \in \LanguageOf{\RegexAlt}
-  }
-  {
-    \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}
-    \OfType \Regex \Leftrightarrow \RegexAlt
+    \String \in \LanguageOf{\BRegex}\\
+    \StringAlt \in \LanguageOf{\BRegexAlt}
+  }
+  {
+    \DisconnectOf{\BRegex}{\BRegexAlt}{\String}{\StringAlt}
+    \OfType \BRegex \Leftrightarrow \BRegexAlt
   }
@@
   {
-    \Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
-    \Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt_2
-  }
-  {
-    \ConcatLensOf{\Lens_1}{\Lens_2} \OfType \Regex_1 \Concat \Regex_2
+    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt_1\\
+    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt_2
+  }
+  {
+    \ConcatLensOf{\Lens_1}{\Lens_2} \OfType \BRegex_1 \Concat \BRegex_2
     \Leftrightarrow
-    \RegexAlt_1 \Concat \RegexAlt_2
+    \BRegexAlt_1 \Concat \BRegexAlt_2
   }
@@
   {
-    \Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
-    \Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt_2
-  }
-  {
-    \SwapLensOf{\Lens_1}{\Lens_2} \OfType \Regex_1 \Concat \Regex_2
+    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt_1\\
+    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt_2
+  }
+  {
+    \SwapLensOf{\Lens_1}{\Lens_2} \OfType \BRegex_1 \Concat \BRegex_2
     \Leftrightarrow
-    \RegexAlt_2 \Concat \RegexAlt_1
+    \BRegexAlt_2 \Concat \BRegexAlt_1
   }
@@
   {
-    \Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
-    \Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt_2
+    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt_1\\
+    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt_2
   }
@@
     \OrLensOf{\Lens_1}{\Lens_2} \OfType
-    \RegexOr{\Regex_1}{\Regex_2}
+    \RegexOr{\BRegex_1}{\BRegex_2}
     \Leftrightarrow
-    \RegexOr{\RegexAlt_1}{\RegexAlt_2}
+    \RegexOr{\BRegexAlt_1}{\BRegexAlt_2}
   }
@@
     & $\begin{cases*}
-      \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\Regex_1}$\\
-      \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\Regex_2}$
+      \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_1}$\\
+      \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_2}$
       \end{cases*}$\\
@@
     & $\begin{cases*}
-      \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\RegexAlt_1}$\\
-      \Lens_2.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\RegexAlt_2}$
+      \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1}$\\
+      \Lens_2.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2}$
       \end{cases*}$\\
@@
     & $\begin{cases*}
-        \Lens_1.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\Regex_1} \BooleanAnd \StringAlt\in\LanguageOf{\RegexAlt_1}$\\
-        \Lens_2.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\Regex_2} \BooleanAnd \StringAlt\in\LanguageOf{\RegexAlt_2}$\\
-        \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\Regex_1} \BooleanAnd \StringAlt\in\LanguageOf{\RegexAlt_2}$\\
-        \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\Regex_2} \BooleanAnd \StringAlt\in\LanguageOf{\RegexAlt_1}$
+        \Lens_1.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\BRegex_1} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_1}$\\
+        \Lens_2.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\BRegex_2} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_2}$\\
+        \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_1} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_2}$\\
+        \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_2} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_1}$
       \end{cases*}$\\
@@
     & $\begin{cases*}
-        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_1} \BooleanAnd \String\in\LanguageOf{\Regex_1}$\\
-        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_2} \BooleanAnd \String\in\LanguageOf{\Regex_2}$\\
-        \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\RegexAlt_1} \BooleanAnd \String\in\LanguageOf{\Regex_2}$\\
-        \Lens_2.\CreateLOf{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_2} \BooleanAnd \String\in\LanguageOf{\Regex_1}$
+        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1} \BooleanAnd \String\in\LanguageOf{\BRegex_1}$\\
+        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2} \BooleanAnd \String\in\LanguageOf{\BRegex_2}$\\
+        \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1} \BooleanAnd \String\in\LanguageOf{\BRegex_2}$\\
+        \Lens_2.\CreateLOf{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2} \BooleanAnd \String\in\LanguageOf{\BRegex_1}$
       \end{cases*}$\\
@@
   {
-    \Lens \OfType \Regex \Leftrightarrow \RegexAlt
+    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt
   }
@@
     \IterateLensOf{\Lens} \OfType
-    \StarOf{\Regex}
+    \StarOf{\BRegex}
     \Leftrightarrow
-    \StarOf{\RegexAlt}
+    \StarOf{\BRegexAlt}
   }
@@
   {
-    \Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt\\
-    \Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt
+    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt\\
+    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt
   }
@@
     \MergeROf{\Lens_1}{\Lens_2} \OfType
-    \RegexOr{\Regex_1}{\Regex_2}
+    \RegexOr{\BRegex_1}{\BRegex_2}
     \Leftrightarrow
-    \RegexAlt
+    \BRegexAlt
   }
@@
     & $\begin{cases*}
-      \Lens_1.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\Regex_1}$\\
-      \Lens_2.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\Regex_2}$
+      \Lens_1.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\BRegex_1}$\\
+      \Lens_2.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\BRegex_2}$
     \end{cases*}$\\
@@
     & $\begin{cases*}
-        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\String\in\LanguageOf{\Regex_1}$\\
-        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\String\in\LanguageOf{\Regex_2}$
+        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\String\in\LanguageOf{\BRegex_1}$\\
+        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\String\in\LanguageOf{\BRegex_2}$
       \end{cases*}$\\
@@
   {
-    \Lens_1 \OfType \Regex \Leftrightarrow \RegexAlt_1\\
-    \Lens_2 \OfType \Regex \Leftrightarrow \RegexAlt_2
+    \Lens_1 \OfType \BRegex \Leftrightarrow \BRegexAlt_1\\
+    \Lens_2 \OfType \BRegex \Leftrightarrow \BRegexAlt_2
   }
@@
     \MergeLOf{\Lens_1}{\Lens_2} \OfType
-    \Regex
+    \BRegex
     \Leftrightarrow
-    \RegexOr{\RegexAlt_1}{\RegexAlt_2}
+    \RegexOr{\BRegexAlt_1}{\BRegexAlt_2}
   }
@@
     & $\begin{cases*}
-      \Lens_1.\PutROf{\String}{\StringAlt} & if $\StringAlt\in\LanguageOf{\RegexAlt_1}$\\
-      \Lens_2.\PutROf{\String}{\StringAlt} & if $\StringAlt\in\LanguageOf{\RegexAlt_2}$
+      \Lens_1.\PutROf{\String}{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1}$\\
+      \Lens_2.\PutROf{\String}{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2}$
     \end{cases*}$\\
@@
     & $\begin{cases*}
-        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_1}$\\
-        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\RegexAlt_2}$
+        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1}$\\
+        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2}$
       \end{cases*}$\\
@@
   {
-    \Lens_1 \OfType \Regex \Leftrightarrow \RegexAlt\\
-    \Lens_2 \OfType \RegexAlt \Leftrightarrow \RegexAltAlt\\
+    \Lens_1 \OfType \BRegex \Leftrightarrow \BRegexAlt\\
+    \Lens_2 \OfType \BRegexAlt \Leftrightarrow \BRegexAltAlt\\
   }
@@
     \ComposeLensOf{\Lens_1}{\Lens_2} \OfType
-    \Regex \Leftrightarrow \RegexAltAlt
+    \BRegex \Leftrightarrow \BRegexAltAlt
   }
@@
   {
-    \Lens \OfType \Regex \Leftrightarrow \RegexAlt
-  }
-  {
-    \InvertOf{\Lens} \OfType \RegexAlt \Leftrightarrow \Regex
+    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt
+  }
+  {
+    \InvertOf{\Lens} \OfType \BRegexAlt \Leftrightarrow \BRegex
   }
@@
   {
-    \Lens \OfType \Regex \Leftrightarrow \RegexAlt\\
-    \Regex \SSREquiv \Regex'\\
-    \RegexAlt \SSREquiv \RegexAlt'
-  }
-  {
-    \Lens \OfType \Regex' \Leftrightarrow \RegexAlt'
+    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt\\
+    \BRegex \SSREquiv \BRegex'\\
+    \BRegexAlt \SSREquiv \BRegexAlt'
+  }
+  {
+    \Lens \OfType \BRegex' \Leftrightarrow \BRegexAlt'
   }
@@
 
-The cost of a lens, $c(\Lens \OfType \Regex \Leftrightarrow \RegexAlt) =
+The cost of a lens between two SREs $c(\Lens, \Regex, \RegexAlt) =
 \MaxOf{\LEntropyOf{\Regex \Given \Lens, \RegexAlt}} +
@@
   \begin{enumerate}
-  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ bounds the expected
-    information content of $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$,
-    given $\SetOf{s \SuchThat s \in \LanguageOf{\Regex} \BooleanAnd
-      \Lens.\PutROf{s}{t} = t}$
-  \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ bounds the expected
-    information content of $\SetOf{s \SuchThat s \in \LanguageOf{\Regex}}$,
-    given $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt} \BooleanAnd
-      \Lens.\PutLOf{t}{s} = s}$
+  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ bounds the entropy of
+    $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$, given $\SetOf{s
+      \SuchThat s \in \LanguageOf{\Regex} \BooleanAnd \Lens.\PutROf{s}{t} = t}$
+  \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ bounds the entropy of
+    $\SetOf{s \SuchThat s \in \LanguageOf{\Regex}}$, given $\SetOf{t \SuchThat t
+      \in \LanguageOf{\RegexAlt} \BooleanAnd \Lens.\PutLOf{t}{s} = s}$
   \end{enumerate}
@@
 
-Note that this method doesn't work for lenses with composition. When composing
-lenses sequentially, the lenses on the sides could project or add information
-when synchronizing with the central type. While the composition may be a
-bijection (and should have zero cost), it can be difficult to see this looking
-merely at the syntax.
-
-However, there are many useful lenses involving composition. We synthesize
-lenses in a normal form as DNF lenses. DNF lenses do not include composition, so
-when we greedily search for a minimal lens, we calculate the cost through DNF
-lenses.
+Note we have not defined entropy for lens compositions. Consider the lens
+$\Lens_1 ; \Lens_2$, where both $\Lens_1$ and $\Lens_2$ are non-bijective.
+However, their composition could actually be bijective, for example if the
+intermediary format contains information not present in either the source nor
+the target. While this composition may be a bijection (and should have zero
+cost), it is difficult to see this looking merely at the syntax.
+
+Notice that our definition of $\Entropy$ contains no case for sequential
+composition $\Lens_1 ; \Lens_2$ and our theorem excludes lenses that contain
+such compositions. Unfortunately, defining the entropy of lenses involving
+composition is challenging because $\Lens_1$ might, for instance, add some
+information that is subsequently projected away in $\Lens_2$. Such operations
+can cancel, leaving us with a zero-entropy bijection. However, detecting such
+cancellations directly is complicated. Fortunately, we are able to avoid such
+complications by synthesizing \emph{DNF lenses} — simple symmetric lenses that
+inhabit a disjunctive normal form. DNF lenses have a number of properties that
+facilitate synthesis: they cut down the program search space dramatically and
+they do not contain instances of composition.
 
@@
 Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm. The
-provided regular expressions are first converted into stochastic regular
+input regular expressions are first converted into stochastic regular
 expressions with \ToStochastic. Next, $\RXSearch$ is initialized with that pair
@@
 \subsection{Stochastic DNF Regular Expressions}
-Stochastic DNF regular expressions are used as a highly normalized form of
-stochastic regular expressions. Due to this highly normalized nature, where type
+Stochastic DNF regular expressions are a highly normalized form of stochastic
+regular expressions. Due to this highly normalized nature, where type
 equivalence is often needed for simple symmetric lenses, such rules aren't
@@
 but syntactically distinct regular expressions have syntactically equal DNF
-forms.
-
-Stochastic DNF regular expressions are hierarchically built, where stochastic
-DNF regular expressions (\DNFRegex, \DNFRegexAlt) are lists of stochastic
-sequences. Stochastic sequences (\Sequence, \SequenceAlt)
+forms.  Like how stochastic regular expressions are regular expressions with
+probability annotations, stochastic DNF regular expressions are DNF regular
+expressions with probability annotations.  Like with SREs and regular
+expressions, if $\DNFRegex$ is a stochastic DNF regular expression,
+$\overline{\DNFRegex}$ is the DNF regular expression generated by erasing
+probability annotations.
+
+Syntactically, stochastic DNF regular expressions (\DNFRegex, \DNFRegexAlt) are
+lists of stochastic sequences. Stochastic sequences (\Sequence, \SequenceAlt)
 themselves are lists of interleaved strings and stochastic atoms. Stochastic
@@
 We have proven this conversion respects languages and probability
-distributions.\bcp{Should be included.}
+distributions.
 
@@
 $\PRegexStar{\DNFRegex}{.8}$. In particular, $\ToDNFRegex$ generates regular
-expressions whose DNF form gives equal probability to all sequences\bcp{Why?}, and gives
-iterations a .8 chance to continue iterating, as that gives a reasonable
-distribution of strings when generating them from the regular
-expressions\bcp{What evaluation did we do to
-  determine that it is the right value?}.
+expressions whose DNF form gives equal probability to all sequences, and gives
+iterations a .8 chance to continue iterating. In our experience generating
+random strings from regular expressions, these probabilities provide good
+distributions of strings -- stars are iterated 4 times on average, and no
+individual choice in a series disjunctions is chosen disproportionately often.
 
@@
 regular expression, like we have for stochastic regular expressions. This
-enables us to efficiently find the entropy without first converting to a
-stochastic regular expression.
+enables us to efficiently find the entropy without first converting a DNF SRE to
+a stochastic regular expression.
 
@@
 
-This means we also have an algorithm for finding the entropy of an unambiguous
-stochastic regular expression, as we can merely convert the regular expression
-into DNF form, and find the entropy of it.
-
 \subsection{Symmetric DNF Lenses}
 Symmetric DNF lenses are an intermediate synthesis target for \GreedySynth.
-Because of the lack of a type equivalence rule or a composition rule, a
-symmetric DNF lens is syntactically quite similar to the stochastic DNF regular
-expression pair it is typed between. This similarity is part of what permits
-such an efficient type-directed synthesis algorithm of them.
+Because of the lack of a type equivalence rule or a composition lens, the set of
+possible lenses with type $\overline{\DNFRegex} \Leftrightarrow
+\overline{\DNFRegexAlt}$ is quite small. Consequently, an algorithm for
+synthesizing such lenses can be driven primarily by the type of the lenses.
 

### 2018-07-08 papers/SymmetricLenses/main.tex
made some eval changes

@@
 proven theorems demonstrating the calculated entropy corresponds to the actual
-conditional entropy to recover the data.\bcp{Again, the proofs should be
-  included.}
+conditional entropy to recover the data.
 
@@
 relatively efficiently, it suffers from an exponential blowup when converting
-regular expressions to DNF form.\sam{Perhaps explain why this blowup arises?} This can be minimized by not converting
+regular expressions to DNF form. This can be minimized by not converting
 \emph{all} of the regular expression into DNF form, but only converting
@@
 synthesis. We alter \GreedySynth such that two regular expression atoms can be
-mapped to each other either if they are the same regular expression and have
-examples agreeing with the examples\bcp{What does ``have
-examples agreeing with the examples'' mean??}, or if there is a lens which is a
-composition of predefined lenses between them, where the lens alters the
-examples in a desired way.
+mapped to each other either if they are the same regular expression and agree
+with the examples, or if there is a lens which is a composition of predefined
+lenses between them, where the lens agrees with the examples.
 
@@
 
-We have extended Optician\bcp{this is the first mention of the word
-  Optician!} with our symmetric synthesis algorithm. Users can 
-write synthesis tasks alongside lens combinators, use those synthesis results in
-the manual writing of new lenses, and use previously defined lenses in the
-synthesis of new lenses.
-
-In this evaluation, we aim to answer two primary questions:
+We have extended the Optician synthesis plugin to Boomerang~\cite{?} with our
+symmetric synthesis algorithm. In doing so, users can write synthesis tasks
+alongside lens combinators, use those synthesis results in the manual writing of
+new lenses, and use previously defined lenses in the synthesis of new lenses.
+
+In this evaluation, we aim to answer three primary questions:
 \begin{enumerate}
 \item Is our synthesis procedure efficient enough to be used in everyday
-  development?  In particular, can it be used in place of the existing
-  bijective lens synthesis algorithm integrated with Boomerang (which
-  \citet{miltner+:bijective-synthesis} have argued is reasonably fast in
-  practice)? 
-  \item Are the heuristics we have described useful for reducing the 
-  ``specification overhead'' for users?  \bcp{Maybe it would be better to
-    ask {\em which} heuristics are the most useful.  Also, are we talking
-    about reducing the number of examples needed?  It's not clear.  (And:
-    Don't the heuristics also make it faster??)}
+  development?
+
+\item How much slower do we perform on bijective lens synthesis benchmarks than
+  the existing Optician system customized for bijective lenses.
+
+\item Which heuristics we have described useful for reducing the ``specification
+  overhead'' for users? In particular, which heuristics are useful in
+  synthesizing the desired lens, instead of merely a lens that satisfies the
+  specification.
 \end{enumerate}
@@
 \subsection{Benchmark Suite}
-Our benchmarks are drawn from four different sources.\bcp{What makes this a good
-  set of benchmarks for present purposes?}
+Our benchmarks are drawn from three different sources.
 \begin{enumerate}
-\item We adapted \afm{8} data cleaning benchmarks from Flash Fill. In adopting
-  these benchmarks, we go beyond the data cleaning tasks\bcp{confusing: you
-    say these are data cleaning tasks, then immediately say we go beyond
-    data cleaning tasks.  (I know the next bit of the sentence makes it
-    (somewhat) clearer, but the reader is already lost right here.)}, and
-  synthesize 
-  synchronization functions\bcp{what is a ``synchronization function''}
-  between the ad-hoc data and their cleaned forms. 
-  \bcp{Clarify:}Optician also adopted benchmarks from Flash Fill, but required alterations to
-  be in bijective form\bcp{Mysterious}. With our symmetric enhancement, we
-  \bcp{words missing?} Optician's four 
-  benchmarks and 4 new Flash Fill benchmarks with no alterations.
-\item We adapted \afm{29} benchmarks from Augeas.  These benchmarks are the same
-  as those present in Optician.  \bcp{I (think I) understand what this
-    means, but readers will not.}While these benchmarks had to be slightly
-  adapted for bijectivity, previous work addressed this lack of bijectivity by
-  synthesizing quotient bijective lenses~\cite{?}.
+\item We adapted \afm{8} data cleaning benchmarks from Flash Fill. In these
+  benchmarks, we not only provide functions that clean the data, but we also
+  provide other synchronization functions: if the cleaned data is updated, those
+  changes are reflected in the original format.
+\item We adapted \afm{29} benchmarks from Augeas. These benchmarks turn the
+  ad-hoc data present in Linux configuration files into a structured, dictionary
+  representation.
 \item We used \afm{8} benchmarks from synchronization tasks on synchronizing a
@@
 
-\bcp{Why is this a good set of benchmarks for the purposes described earlier
-in the section?}
-
-\subsection{Efficiency}
+A realistic benchmarks suite is cretical in answering the first research
+question. To determine whether our algorithm is fast enough to be used in
+everyday development, we need to evaluate whether we can solve real-world
+problems efficiently. All three of our benchmark sources test real life
+problems: the data cleaning tasks presented in Flash Fill are either derived
+from online help forums or taken from the Excel product team; Augeas is a
+utility that transforms Linux configuration files into a manageable dictionary
+form; the many of the remaining benchmarks are from synchronizing real-life data
+formats, like BibTeX and EndNote.
+
+\subsection{Useful in Everyday Development}
 
 To answer our first research question, we measure the runtime of our algorithm
-and compare it to the existing algorithm for synthesizing bijective
-lenses.\bcp{Note that this comparison doesn't, directly answer the question
-  ``Is it fast enough.''}
-
-We synthesized lenses in 4 modes:
+on our benchmark suite.
+
+We synthesized lenses in 2 modes:
 
@@
                 but compositional synthesis is not enabled.\\
-  \BSOpt{}: &  We synthesize lenses using the existing bijective lens synthesis
-              algorithm with all optimizations enabled\\
-  \BSNCOpt{}: & We synthesize lenses using the existing bijective lens synthesis
-                algorithm, but compositional synthesis is not enabled.\\
 \end{tabulary}\\
 
-We include running these algorithms without compositional synthesis to stress
-test the algorithms. While compositional synthesis makes the procedure easier,
-and demonstrates the potential to be used in arbitrarily hard problems, it is
-dependent on the user identifying important sublenses between subcomponents of
-the final formats. It is therefore interesting to see how many problems each
-algorithm can solve, when trying to synthesize ``in one go.''
+We include running the symmetric synthesis algorithm without compositional
+synthesis to stress test the algorithms. While compositional synthesis makes the
+procedure easier, and demonstrates the potential to be used in arbitrarily hard
+problems, it is dependent on the user identifying important sublenses between
+subcomponents of the final formats. It is therefore interesting to see how many
+problems each algorithm can solve, when trying to synthesize ``in one go.''
 
@@
     benchmarks in our benchmark suite in under 30 seconds, and is able to
-    complete most of the benchmarks without using compositional
-    synthesis. \bcp{The interesting part of the graph is all squished
-      against the Y axis.  Can we make the X axis logarithmic?}}
+    complete most of the benchmarks without using compositional synthesis.
+    \bcp{The interesting part of the graph is all squished against the Y axis.
+      Can we make the X axis logarithmic?}}
   \label{fig:times}
@@
 
-The slowest completed benchmark for both the bijective and symmetric synthesis
-algorithms is \texttt{xml\_to\_augeas.boom}, a benchmark that converts
-arbitrary XML up to depth 3 into a serialized version of the structured
-dictionary representation used in Augeas.  This benchmark takes 18.9 seconds for
-the symmetric synthesis algorithm to complete, and 9.3 seconds the bijective
-synthesis algorithm to complete.
-
-\subsection{The Effect of Our Heuristic}
+\subsection{Slowdown Compared to Bijective Synthesis}
+
+To compare to the existing bijective synthesis algorithm, we run our symmetric
+synthesis algorithm on the original Optician benchmark suite. \afm{8} of these
+benchmarks had to be slightly altered with tighter specifications or termination
+parameters, as we had to alter the original specifications to generate the
+correct lens.
+
+To perform this comparison, we synthesized lenses in 2 modes:
+
+\begin{tabulary}{\linewidth}{rL}
+  \SSOpt{}: & We synthesize lenses using our symmetric synthesis algorithm with all optimizations enabled.\\
+  \BSOpt{}: & We synthesize lenses using the existing bijective synthesis
+              algorithm with all optimizations enabled.\\
+\end{tabulary}\\
+
+For each benchmark in our benchmark suite, we ran every mode on it with a
+timeout of 30 seconds, and average the result over 5 runs.
+Figure~\ref{fig:comparison-times} summarizes the results of these tests. On
+average, \SSOpt{} took 2.1 times (\afm{3.5} seconds) longer to complete than
+\BSOpt{}.
+
+The slowest completed benchmark for both synthesis algorithms is
+\texttt{xml\_to\_augeas.boom}, a benchmark that converts arbitrary XML up to
+depth 3 into a serialized version of the structured dictionary representation
+used in Augeas. This benchmark takes 18.9 seconds for the symmetric synthesis
+algorithm to complete, and 9.3 seconds the bijective synthesis algorithm to
+complete.
+
+\subsection{The Effects of Our Heuristics}
 We evaluate the usefulness of (1) our information theoretic metric, and (2) our

### 2018-07-08 papers/SymmetricLenses/main.tex
add this work

@@
   asymmetric lenses and full symmetric lenses and has the useful property that
-  its members don't have any internal state (\S\ref{sec:overview}). 
-\item We show how to synthesize simple symmetric lenses (including all asymmetric lenses as a special case),
-using a novel application of stochastic regular expressions to guide the search process.  In particular, we
-show that the RE star semi-ring equivalences remain valid when their semantics
-are enriched by distributions (\S\ref{sec:synthesis}).
+  its members don't have any internal state (\S\ref{sec:overview}). We describe
+  a language for simple symmetric string lenses and the cost metric for that
+  language (\S\ref{sec:ssl}).
+\item We show how to synthesize simple symmetric lenses (including all
+  asymmetric lenses as a special case), using a novel application of stochastic
+  regular expressions to guide the search process (\S\ref{sec:synthesis}). In
+  particular, we show that the RE star semi-ring equivalences remain valid when
+  their semantics are enriched by distributions (\S\ref{sec:sre}).
 \item We extend the Boomerang implementation with simple symmetric lenses,
@@
 \end{itemize}
+We finish the paper with a discussion of related work (\S\ref{sec:related}) and
+our final thoughts (\S\ref{sec:conc}).
 
@@
 In this section, we give an overview of simple symmetric lenses and our
-corresponding synthesis algorithm using an example drawn from a hypothetical
+synthesis algorithm for them, using a simplified example drawn from a hypothetical
 U.S. company. In this company, management and human resources (HR) store
-information about employees in separate files: management records the names of
+information about employees in separate text files: management records the
+names of 
 employees and their salaries while HR stores the names of employees and their
 health insurance providers. Figure~\ref{fig:minimized-representations} gives
-examples of the two types of files and regular expressions describing the two
-formats. The company uses simple symmetric lenses to keep these files
+examples of the two file formats and regular expressions describing them.
+
+The company uses a simple symmetric lens to keep these files
 synchronized. If management adds a new employee, ``Chris Roe: 32500'', a lens
@@
 \subsection{Simple Symmetric Lenses}
-A simple symmetric lens between formats described by regular expression types
-$\BRegex$ and $\BRegexAlt$\footnote{We use nonterminals \BRegex and \BRegexAlt
-  for regular expressions, using \Regex and \RegexAlt for stochastic regular
-  expressions.} is defined by four functions.\bcp{I think a picture would be
-  helpful. I can probably make one pretty easily by tweaking the one from the
-  original symmetric paper.} \afm{That would be awesome.}
+A simple symmetric lens between formats described by types $X$ and $Y$ is
+defined by four functions.\bcp{I think a picture would be helpful. I can
+  probably make one pretty easily by tweaking the one from the original
+  symmetric paper.} \afm{That would be awesome.}
 
@@
  \tag{1}
- \mbox{$\CreateR \OfType \BRegex \rightarrow \BRegexAlt$}\\
+ \mbox{$\CreateR \OfType X \rightarrow Y$}\\
  \tag{2}
- \mbox{$\CreateL \OfType \BRegexAlt \rightarrow \BRegex$}\\
+ \mbox{$\CreateL \OfType Y \rightarrow X$}\\
  \tag{3}
- \mbox{$\PutR \OfType \BRegex \rightarrow \BRegexAlt \rightarrow \BRegexAlt$}\\
+ \mbox{$\PutR \OfType X \rightarrow Y \rightarrow Y$}\\
  \tag{4}
- \mbox{$\PutL \OfType \BRegexAlt \rightarrow \BRegex \rightarrow \BRegex$}
+ \mbox{$\PutL \OfType Y \rightarrow X \rightarrow X$}
 \end{gather}
@@
   \tag{\CreatePutRL}
-  \PutLOf{(\CreateROf{s})}{s} = s\\
+  \PutLOf{(\CreateROf{x})}{x} = x\\
   \tag{\CreatePutLR}
-  \PutROf{(\CreateLOf{t})}{t} = t\\
+  \PutROf{(\CreateLOf{y})}{y} = y\\
   \tag{\PutRL}
-  \PutLOf{(\PutROf{s}{t})}{s} = s\\
+  \PutLOf{(\PutROf{x}{y})}{x} = x\\
   \tag{\PutLR}
-  \PutROf{(\PutLOf{t}{s})}{t} = t
+  \PutROf{(\PutLOf{y}{x})}{y} = y
 \end{gather}
 \noindent
-The two create functions provide default values for times when
-data appears on one side but not the other (\EG, the ``unk'' salary
-entry added to the management file when HR inserts a new employee).
-The two put functions propagate edits from one format to the
-other by combining the change from one file with the previous value
-from the other file. The roundtripping laws guarantee that identify edits to one file do
-not change the data in the other file.
+The two {\sc Create} functions propagate shared information and fill in default
+values when introducing new data. (\EG, the ``unk'' salary entry is added
+alongside the name to the management file when HR inserts a new employee). The
+two {\sc Put} functions propagate edits from one format to the other by
+combining the change from one file with the previous value from the other file.
+The roundtripping laws guarantee that if two formats are already synchronized,
+{\sc Put}s merely return the second argument, unchanged.
+
+This definition differs from previous formulations of simple symmetric lenses,
+as previous formulations required a \emph{complement}. Instead of comprising
+four functions, existing formulations merely comprised two functions, which
+combined complements with edited formats to create updated complements and
+synchronized data. We provide a detailed comparison to classic symmetric lenses
+in \S\ref{sec:relationship}.
 
@@
 
-We use regular expressions to denote the types of data formats, and we
-impose a typing discipline on our lenses.  Specifically, if 
-$\BRegex$ and $\BRegexAlt$ are regular expressions, then the notation
-\lstinline{$\Lens$ : $\BRegex \Leftrightarrow \BRegexAlt$} indicates that $\Lens$
-is a simple symmetric lens between $\LanguageOf{\BRegex}$ and
-$\LanguageOf{\BRegexAlt}$. 
-When clear from context, we use the term \textit{lens} to denote a
-simple symmetric lens.
-
 \subsection{Simple Symmetric Lens Combinators}
+
 Generally, we do not expect programmers to define simple symmetric lenses by
-writing the four functions by hand and showing they satisfy the
-roundtripping laws.  Instead, we provide simple symmetric lens
-combinators to build up the requisite lenses from simpler ones.
-We briefly explain a selection of these combinators by using them to
-define a lens between 
-the two employee data formats in Figure~\ref{fig:minimized-representations}. 
-The finished lens, which appears in Figure~\ref{fig:example_lens}, has
-the type 
-\lstinline{emp_salaries} $\Leftrightarrow$ \lstinline{emp_insurance}, where
-\lstinline{emp_salaries} and \lstinline{emp_insurance} are regular
-expressions defined in Figure~\ref{fig:minimized-representations}.
+writing the four functions by hand and showing they satisfy the roundtripping
+laws. Instead, we provide simple symmetric lens combinators to build up the
+requisite lenses from simpler ones, and focus our combinators on providing
+lenses between string data formats.
+
+We use regular expressions to denote the types of data formats, and we impose a
+typing discipline on our lenses. Specifically, if $\BRegex$ and $\BRegexAlt$ are
+regular expressions\footnote{We use \BRegex and \BRegexAlt to denote regular
+  expressions. We later use the plain nonterminals \Regex and \RegexAlt for
+  stochastic regular expressions.}, then the notation
+\lstinline{$\Lens$ : $\BRegex \Leftrightarrow \BRegexAlt$} indicates that $\Lens$ is a simple symmetric lens between $\LanguageOf{\BRegex}$ and $\LanguageOf{\BRegexAlt}$.
+When clear from context, we use the term \textit{lens} to denote a simple symmetric lens.
+
+We briefly explain a selection of these
+combinators by using them to define a lens between the two employee data formats
+in Figure~\ref{fig:minimized-representations}. The finished lens, which appears
+in Figure~\ref{fig:example_lens}, has the type
+\lstinline{emp_salaries} $\Leftrightarrow$
+\lstinline{emp_insurance}, where
+\lstinline{emp_salaries} and
+\lstinline{emp_insurance} are regular expressions defined in Figure~\ref{fig:minimized-representations}.
 
@@
 
-In contrast to the identity lens, the \Disconnect lens, leaves data
-unconnected. The disconnect lens takes four arguments: two regular expressions
-and two strings. The regular expressions specify the formats of the
-corresponding data on the two sides while the strings provide default values.
-On creates, the input values are thrown away, and default values are returned ($\CreateROf{\String} =
+In contrast to the identity lens, the \Disconnect lens does not propagate edits
+from one format to the other. The disconnect lens takes four arguments: two
+regular expressions and two strings. The regular expressions specify the formats
+of the corresponding data on the two sides while the strings provide default
+values. On creates, the input values are thrown away, and default values are
+returned ($\CreateROf{\String} =
 \lstinline{"UNK"}$), and on puts, the second argument is used, and the first is
-thrown away ($\PutROf{\String}{\StringAlt} = \StringAlt$).  In
-our extended example, the \lstinline{salary} and
-\lstinline{company} fields are independent, so we use the disconnect lens to relate them. The defaults are strings that indicate the information is missing.
+thrown away ($\PutROf{\String}{\StringAlt} = \StringAlt$). In our extended
+example, the \lstinline{salary} and \lstinline{company} fields are independent,
+so we use the disconnect lens to relate them\bcp{Potentially confusing to
+  readers (and, even to me, hacky). Why not relate each of them to the constant
+  empty string?}\afm{I'm fine with the empty string.  I've seen sentinel values
+  like UNK used before in practice, but if it makes the example cleaner and more
+  comprehensible, sure.  Anybody else have thoughts on the matter?}.
+The defaults are strings that indicate the information is missing.
 %
@@
 equivalences on them, and sometimes we must convert one regular expression type
-to another, equivalent regular expression type. Our typing rules include type
-equivalence, so if a lens is typed between two regular expressions, it is also
-typed between two equivalent ones. The flexibility this type equivalence rule
-provides is useful when a different representation of a regular expression would
-facilitate synthesis.
+to another, equivalent, regular expression type. Our typing rules include type
+equivalence, so if a lens has type $\Regex \Leftrightarrow \RegexAlt$, then it
+is also has type $\Regex' \Leftrightarrow \RegexAlt'$, where $\Regex$ is
+equivalent to $\Regex'$, and $\RegexAlt$ is equivalent to $\RegexAlt'$. The
+flexibility this type equivalence rule provides is useful when a different
+representation of a regular expression would facilitate synthesis.
 
@@
 for real-world formats is quite difficult, as these formats can be large and
-transform data in complex and unintuitive ways. To alleviate the difficulty of
+represent data in complex and unintuitive ways. To alleviate the difficulty of
 manually writing simple symmetric lenses, we wish to synthesize them instead.
@@
 %
-is another valid lens (where \lstinline{rep($\BRegex$)} is a string in the
+is another valid lens (where \lstinline{rep($\BRegex$)} denotes some string in the
 language of $\BRegex$). Furthermore, adding the input-output example in
 Figure~\ref{fig:minimized-representations} 
-helps, but only a little, because it requires many examples to rule
-out all the possible occurrences of nested \Disconnect lenses,
+helps, but only a little: it may, in general, require {\em many} examples to rule
+out all possible occurrences of nested \Disconnect lenses,
 particularly in complex formats. 
@@
 
-But what is a likely satisfying lens? We propose that a satisfying
+But what is a ``likely'' satisfying lens? We propose the following
+heuristic: a satisfying
 lens is more likely if it connects more data between the two
-formats, \IE, by using an identity rather than a disconnect lens.  The
+formats, \EG, by using an identity rather than a disconnect lens.  The
 likeliest satisfying lenses would connect as much data as
-possible. More formally, we propose that the likelihood of a
-satisfying lens is the expected number of bits required to recover
+possible. More formally, we define the likelihood of a
+satisfying lens as the expected number of bits required to recover
 data in one format from synchronized data in the other.  Higher
-likelihoods correspond to fewer numbers of bits. In more
+likelihoods correspond to lower numbers of bits. In more
 detail, we say that two strings $s$ and $t$ are \emph{synchronized
@@
 the \IdentityLens{} lens, we can recover $s$ from $t$ using no bits
-because in this case, $s$ is just $t$.  In contrast, given
+because, in this case, $s$ is just $t$.  In contrast, given
 the \Disconnect{} lens, we need enough bits to fully encode $s$ in
@@
 
-The expected number of bits required to recover data is a
-well-known concept in information theory: it is the \emph{entropy} of
-the data. Calculating the 
-entropy of data requires a probability distribution over the
-space of possible values for the data.  Specifically,
-given a set $S$ and a probability distribution over that set $P
-\OfType S \rightarrow \Reals$, the entropy of the $(S,P)$ pair is $-\Sigma_{s\in
-  S}P(s)\Log_2P(s)$.
+The expected number of bits required to recover data is a well-known concept in
+information theory: it is the \emph{entropy} of the data. Calculating the
+entropy of data requires a probability distribution over the space of possible
+values for the data. Specifically, given a set $S$ and a probability
+distribution over that set $P \OfType S \rightarrow \Reals$, the entropy of $S$
+with probability distribution $P$ is $-\Sigma_{s\in S}P(s)\Log_2P(s)$.
 
@@
 distributions over that data.
-To that end, we adopt \emph{stochastic regular expressions}\cite{?} (SREs), 
+To that end, we adopt \emph{stochastic regular expressions}~\cite{?} (SREs), 
 which are regular expressions where each operator is annotated with a
-probability.  Stochastic regular expressions
-allow us to jointly specify the allowable data and give a probability
+probability.  A stochastic regular expression
+specifies both a set of strings and a probability
 distribution over it.  From this information, we can calculate the
@@
 likelihood we assign to $\Lens$ is the sum of the conditional entropy of
-$\Regex$ given $\RegexAlt$ and $\Lens$ and the conditional entropy of
-$\RegexAlt$ given $\Regex$ and $\Lens$.
+$\Regex$ given $\RegexAlt$ and $\Lens$, \bcp{Am I supposed to know what this
+  means?} and the conditional entropy of $\RegexAlt$ given $\Regex$ and
+$\Lens$.\bcp{Why is this the right (or a sensible) thing to do, besides the fact
+  that it ensures symmetry?}
 
@@
 Now that we have a way to calculate likelihood, we need a way to search for
-it.
-Previous work on synthesis~\cite{?,?,?} has shown
-that we can use types to dramatically restrict the search space and
-improve the effectiveness of synthesis.  In our setting, types are
-pairs of stochastic regular expressions, which come with a rich
-algebraic structure that makes each regular expression semantically
-equal to infinitely many others.  To tame that search space and
-leverage the benefits of type-directed synthesis, we split our
-algorithm into two communicating synthesizers following earlier
-work~\cite{quotient-lens-synthesis}.  The first
-synthesizer, \RXSearch, navigates the space of semantically equivalent
-regular expressions by iteratively applying rewrite rules that
-preserve both the semantics and the probability
-distributions of those stochastic regular expressions. \RXSearch
-ranks pairs of stochastic regular expressions 
-by the number of rewrite rule applications required to obtain the pair
-from the stochastic regular expressions given as input. It passes the
-pairs off to the second synthesizer, \GreedySynth, in rank order with
+likely lenses. Previous work on synthesis~\cite{?,?,?} has shown that we can use
+types to dramatically restrict the search space and improve the effectiveness of
+synthesis. In our setting, types are pairs of stochastic regular expressions,
+which come with a rich algebraic structure that makes each regular expression
+semantically equal to infinitely many others. To tame that search space and
+leverage the benefits of type-directed synthesis, we split our algorithm into
+two communicating synthesizers following earlier
+work~\cite{quotient-lens-synthesis}. The first synthesizer, \RXSearch, navigates
+the space of semantically equivalent regular expressions by iteratively applying
+rewrite rules that preserve both the semantics and the probability distributions
+of those stochastic regular expressions. \RXSearch ranks pairs of stochastic
+regular expressions by the number of rewrite rule applications required to
+obtain the pair from the stochastic regular expressions given as input. It
+passes the pairs off to the second synthesizer, \GreedySynth, in rank order with
 the smallest first.
@@
 At a high level, \GreedySynth looks for lenses between a given pair of
-stochastic regular expressions $\Regex$ and $\RegexAlt$ that have the
-highest likelihood by performing a type-directed search. 
-In more detail, \GreedySynth first uses type equalities based on semiring 
-equivalences (ignoring the star semiring equivalences that are managed
-by \RXSearch) to normalize the
-stochastic regular expressions provided by \RXSearch into 
-\emph{stochastic n-ary DNF regular expressions}.  
-Then, \GreedySynth uses the syntax of these n-ary DNF regular
-expressions to find highly normalized lenses, in a
-form we call \emph{simple symmetric n-ary DNF lenses}.  These lenses include
-neither a composition operator nor a type equivalence rule.  These
-restrictions mean that each simple symmetric n-ary DNF lens is typed
-by a single pair of stochastic n-ary DNF regular expressions and 
-\GreedySynth's search space for a given pair of regular expressions is
-finite.  The variations that \GreedySynth 
-considers include choosing between \IdentityLens{} and  \Disconnect
-lenses and deciding how to match multiple occurrences of the same kind
-of data, \EG, the two occurrences of \lstinline{name} in the
-management and HR data formats (we want to map first names to first
-names and last names to last names). 
-Finally, \GreedySynth converts this lens 
-back into a simple symmetric lens by converting the n-ary forms into
-the binary forms provided in the surface language.
+stochastic regular expressions $\Regex$ and $\RegexAlt$ that have the highest
+likelihood by performing a type-directed search. In more detail, \GreedySynth
+first normalizes the stochastic regular expressions provided by \RXSearch into
+\emph{stochastic DNF regular expressions}. These stochastic DNF regular
+expressions are, intuitively, stochastic regular expressions with disjunctions
+distributed over concatenations, where associativity information has been
+removed. Then, \GreedySynth uses the syntax of these n-ary DNF regular
+expressions to find highly normalized lenses, in a form we call \emph{simple
+  symmetric n-ary DNF lenses}. These lenses include neither a composition
+operator nor a type equivalence rule. These restrictions mean that each simple
+symmetric n-ary DNF lens is typed by a single pair of stochastic n-ary DNF
+regular expressions and \GreedySynth's search space for a given pair of regular
+expressions is finite. The variations that \GreedySynth considers include
+choosing between \IdentityLens{} and \Disconnect lenses and deciding how to
+match multiple occurrences of the same kind of data, \EG, the two occurrences of
+\lstinline{name} in the management and HR data formats (we want to map first
+names to first names and last names to last names). Finally, \GreedySynth
+converts this lens back into a simple symmetric lens by converting the n-ary
+forms into the binary forms provided in the surface language.
 
@@
 \section{Stochastic Regular Expressions}
-\label{sec:srx}
+\label{sec:sre}
 To characterize the likely lenses, we must compute the expected number of bits
@@
 much more likely having a string involving more than $n$ iterations is than
-having a string involving $n$ iterations. The semantics of SREs is defined by
-the function $P$ below.
+having a string involving exactly $n$ iterations. The semantics of SREs is
+defined by the function $P$ below.
 
@@
 $\RXSearch$ needs to be able to find equivalent stochastic regular expressions.
-However, as of yet, there is no syntactic means to reason about stochastic
-regular expression equivalence. We extend the \emph{star-semiring} equivalences
-to stochastic regular expressions.
+However, existing work does not reason about stochastic regular expression
+equivalence. We extend the \emph{star-semiring}~\cite{?} equivalences to
+stochastic regular expressions.\bcp{Why these?}
 
@@
   The proofs of this theorem, and subsequent theorems in this paper, appears in
-  the auxiliary materials accompanying this submission.
+  the auxiliary materials accompanying this submission.\bcp{Not in the appendix?}
 \end{proof}
@@
 of a stochastic regular expression can be computed directly from its syntax,
-when the stochastic regular expression is unambiguous, and contains no nonempty
-components.
+when the stochastic regular expression is unambiguous and contains no nonempty\bcp{empty?}
+subcomponents.
 \begin{center}
@@
 Using the above computation, this SRE has an entropy of $1$, but the entropy
-should be $0$. Similar issues occur when trying to find the entropy when
+should be $0$.\bcp{This doesn't show that it's not easy---it just shows that
+one particular way of computing it is wrong.} Similar issues occur\bcp{Can we at least hint what they are?} when trying to find the entropy when
 $\emptyset$ is a subterm, so we do not define the entropy on $\emptyset$.
 
-In our work, we do not find this to be a problem, we require unambiguous regular
+In our work, we do not find this to be a problem\bcp{Why might one expect
+  it to be a problem, rather than just a technicality?}: we require
+unambiguous regular 
 expressions as input to our synthesis procedure, and we can easily preprocess
@@
 by traversing the correct star-semiring equivalences
-(\PRegexOr{\emptyset}{\lstinline{"s"}}{.5} \SSREquiv \lstinline{"s"}).
+(\EG, \PRegexOr{\emptyset}{\lstinline{"s"}}{.5} \SSREquiv \lstinline{"s"}).
 
@@
 
-We have already described over a number of these lenses in \S\ref{sec:overview}.
+We have already described a number of these lenses in \S\ref{sec:overview}.
 The compose lens $\Lens_1 ; \Lens_2$ composes two lenses sequentially, it
@@
 We now provide the typing derivations and semantics for these lenses. For
-simplicity, we define the typing derivations and semantics only over unambiguous
+simplicity\bcp{Is there any known issue with relaxing this restriction?
+  I've often found ti convenient to use id lenses with ambiguous REs...}, we define the typing derivations and semantics only over unambiguous
 regular expressions, even though some lenses can be defined over ambiguous
@@
 The \MergeR lens is interesting because it merges data where one data can be in
-two formats, and one data has only one format. In previous work, this was
+two formats, and one data has only one format. In previous work\bcp{where?},
+this was
 combined into \OrLens{}, where \OrLens{} could have ambiguous types, but we
@@
 recover a string in one data format from the other. We use the function
-$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to bound the expected amount of
+$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to bound\bcp{``bound'' or
+  ``calculate''?  I.e., is what we calculate correct, or is it only known to
+be an upper bound?  If (as it seems) the answer is the latter, then why?}
+the expected amount of 
 information required to recover a string in $\RegexAlt$ from a string in
@@
   \end{algorithmic}
+\bcp{Why does the RETURN statement not get a number?}
 \end{algorithm}
@@
 best lens found thus far. Note this means that if \RXSearch finds a bijective
-lens (a lens of cost 0) it will immediately return.
+lens (a lens of cost 0\bcp{have we proved that the bijective lenses are
+  exactly the ones with zero cost?  That would be nice to state clearly someplace...}) it will immediately return.
 %
@@
 \ToDNFRegex defined, it is easier to explain the function that converts regular
-expressions into stochastic regular expressions, \ToStochastic. If $\Regex$ is a
+expressions into stochastic regular expressions, \ToStochastic\bcp{from the
+  description above, I thought this step was optional (because we allowed
+  the user to write SREs directly)}. If $\Regex$ is a
 stochastic DNF regular expression generated by \ToStochastic, then
@@
 $\PRegexStar{\DNFRegex}{.8}$. In particular, $\ToDNFRegex$ generates regular
-expressions whose DNF form gives equal probability to all sequences, and gives
+expressions whose DNF form gives equal probability to all sequences\bcp{Why?}\bcp{Also, what does this mean?  If a given RE denotes an
+infinite language, then there is no way to give equal probability to every
+string in that language, is there???}, and gives
 iterations a .8 chance to continue iterating. In our experience generating
-random strings from regular expressions, these probabilities provide good
+random strings from regular expressions\bcp{what does generating random
+  strings have to do with synthesis? i.e., why is this a good thing to measure?}, these probabilities provide good
 distributions of strings -- stars are iterated 4 times on average, and no
@@
 as well as some additional information. Symmetric sequence lenses consist of a
-list of symmetric atom lenses, as well as some additional information, and
+list of symmetric atom lenses\bcp{why are they called atom lenses?}, as well
+as some additional information, and
 symmetric atom lenses are merely iterations of symmetric DNF lenses. In the
@@
 \label{subsec:greedy-synth}
+\bcp{The following is all pretty incomprehensibly dense.  A little running example
+might help.}
 

### 2018-07-08 papers/SymmetricLenses/main.tex
add this work

@@
   \begin{enumerate}
-  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ calculates the expected
+  \item $\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ is an upper bound for the expected
     information content of $\SetOf{t \SuchThat t \in \LanguageOf{\RegexAlt}}$,
@@
       \Lens.\PutROf{s}{t} = t}$
-  \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ calculates the expected
+  \item $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ is an upper bound for the expected
     information content of $\SetOf{s \SuchThat s \in \LanguageOf{\Regex}}$,
@@
 \begin{proof}
-Given $s \in S$, let $V_{\ell}(s) = \{t \in \mathcal{L}(T) \; | \; \ell.\PutROf{s}{t} = t\}$. We want to show that
-$$\sum_{s \in \mathcal{L}(S)}P_{\mathcal{L}(S)}(s)H(V_{\ell}(s)) = \mathbb{H}^{\rightarrow}(S \; | \; \ell, T)$$
-We proceed by induction over $\ell$. 
+Let $H(T \; | \; \ell, S)$ be the expected information content of $\{t \; | \; t \in \mathcal{L}(T)\}$ given $\{s \; | \; s \in \mathcal{L}(S) \wedge \ell.\PutROf{s}{t} = t\}$.
+Given $s \in S$, let $V^{\rightarrow}_{\ell}(s) = \{t \in \mathcal{L}(T) \; | \; \ell.\PutROf{s}{t} = t\}$.
+We want to show that
+$$H(T \; | \; \ell, S) = \sum_{s \in \mathcal{L}(S)}P_{\mathcal{L}(S)}(s)H(V^{\rightarrow}_{\ell}(s)) \leq \mathbb{H}^{\rightarrow}(T \; | \; \ell, S)$$
+(The proofs for $\mathbb{H}^{\leftarrow}(S \; | \; \ell, T)$ will follow the same pattern). We proceed by induction over $\ell$. 
 \begin{enumerate}
@@
 
-Observe that $V_{\IdentityLensOf{S}}(s) = \{t \in \mathcal{L}(S) \; | \; \IdentityLensOf{S}.\ell.\PutROf{s}{t} = t\} = \{t \in \mathcal{L}(S) \; | \; t = s\} = \{s\}$.
-
-Consequently, $H(V_{\IdentityLensOf{S}}(s)) = H(\{s\}) = 0$, from which it follows that 
-$$\sum_{s \in \mathcal{L}(S)} P_{\mathcal{L}(S)}(s) * H(V_{\IdentityLensOf{S}}(s)) = 0 = \mathbb{H}^{\rightarrow}(S \; | \; \IdentityLensOf{S}, T)$$
+Observe that $V^{\rightarrow}_{\IdentityLensOf{S}}(s) = \{t \in \mathcal{L}(S) \; | \; \IdentityLensOf{S}.\ell.\PutROf{s}{t} = t\} = \{t \in \mathcal{L}(S) \; | \; t = s\} = \{s\}$.
+
+Consequently, $H(V^{\rightarrow}_{\IdentityLensOf{S}}(s)) = H(\{s\}) = 0$, from which it follows that 
+$$\sum_{s \in \mathcal{L}(S)} P_{\mathcal{L}(S)}(s) * H(V^{\rightarrow}_{\IdentityLensOf{S}}(s)) = 0 = \mathbb{H}^{\rightarrow}(S \; | \; \IdentityLensOf{S}, T)$$
 \item
@@
 Observe that 
-$$V_{\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}}(s) = \{t \in \mathcal{L}(T) \; | \; \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}.\PutROf{s}{t} = t\} = \{t \in \mathcal{L}(T) \; | \; t = t\} = \mathcal{L}(T)$$
+$$V^{\rightarrow}_{\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}}(s) = \{t \in \mathcal{L}(T) \; | \; \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}.\PutROf{s}{t} = t\} = \{t \in \mathcal{L}(T) \; | \; t = t\} = \mathcal{L}(T)$$
 Consequently
-$$\sum_{s \in \mathcal{L}(S)}P_{S}(s)H(V_{\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}}(s)) = \sum_{s \in \mathcal{L}(S)}P_{S}(s)H(\mathcal{L}(T)) = H(\mathcal{L}(T)) = \mathbb{H}(T) = \mathbb{H}^{\rightarrow}(S \; | \; \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, T)$$
+\begin{align*}
+\sum_{s \in \mathcal{L}(S)}P_{S}(s)H(V^{\rightarrow}_{\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}}(s)) &= \sum_{s \in \mathcal{L}(S)}P_{S}(s)H(\mathcal{L}(T))\\
+&= H(\mathcal{L}(T))\\
+&= \mathbb{H}(T)\\
+&= \mathbb{H}^{\rightarrow}(T \; | \; \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, S)
+\end{align*}
 \item
@@
 \begin{align*}
-V_{\IterateLensOf{\Lens}}(s_1 \cdot \ldots \cdot s_n) &= \{t_1 \cdot \ldots \cdot t_m \; | \; \IterateLensOf{\Lens}.\PutROf{(s_1 \cdot \ldots \cdot s_n) }{(t_1 \cdot \ldots \cdot t_m)} = t_1 \cdot \ldots \cdot t_m\}\\
+V^{\rightarrow}_{\IterateLensOf{\Lens}}(s_1 \cdot \ldots \cdot s_n) &= \{t_1 \cdot \ldots \cdot t_m \; | \; \IterateLensOf{\Lens}.\PutROf{(s_1 \cdot \ldots \cdot s_n) }{(t_1 \cdot \ldots \cdot t_m)} = t_1 \cdot \ldots \cdot t_m\}\\
 &= \{t_1 \cdot \ldots \cdot t_n \; | \; \ell.\PutROf{s_i}{t_i} = t_i \text{ for }1 \leq i \leq n\}\\
-&= V_{\ell}(s_1) \cdot \ldots \cdot V_{\ell}(s_n)
+&= V^{\rightarrow}_{\ell}(s_1) \cdot \ldots \cdot V^{\rightarrow}_{\ell}(s_n)
 \end{align*}
@@
 \begin{align*}
-\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V_{\IterateLensOf{\Lens}}(s))
-&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n)\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i) H(V_{\ell}(s_1) \cdot \ldots \cdot V_{\ell}(s_n))\\
-&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n)\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i) \left(\sum_{j=1}^n H(V_{\ell}(s_j)) \right)\\
-&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \sum_{s_n} P_S(s_n) \left(\sum_{j=1}^n H(V_{\ell}(s_j)) \right)
+\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V^{\rightarrow}_{\IterateLensOf{\Lens}}(s))
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n)\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i) H(V^{\rightarrow}_{\ell}(s_1) \cdot \ldots \cdot V^{\rightarrow}_{\ell}(s_n))\\
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_n)\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^n P_S(s_i) \left(\sum_{j=1}^n H(V^{\rightarrow}_{\ell}(s_j)) \right)\\
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \sum_{s_n} P_S(s_n) \left(\sum_{j=1}^n H(V^{\rightarrow}_{\ell}(s_j)) \right)
 \end{align*}
-We focus on the term $\sum_{s_n} P_S(s_n) \sum_{j=1}^n H(V_{\ell}(s_j))$:
+We focus on the term $\sum_{s_n} P_S(s_n) \sum_{j=1}^n H(V^{\rightarrow}_{\ell}(s_j))$:
 \begin{align*}
-\sum_{s_n} P_S(s_n) \sum_{j=1}^n H(V_{\ell}(s_j)) &= \sum_{s_n} P_S(s_n)H(V_{\ell}(s_n)) + P_S(s_n) \sum_{j=1}^{n-1} H(V_{\ell}(s_j))\\
-&= \mathbb{H}^{\rightarrow}(S \; | \; \ell, T) + \sum_{j=1}^{n-1}H(V_{\ell}(s_j))
+\sum_{s_n} P_S(s_n) \sum_{j=1}^n H(V^{\rightarrow}_{\ell}(s_j)) &= \sum_{s_n} P_S(s_n)H(V^{\rightarrow}_{\ell}(s_n)) + P_S(s_n) \sum_{j=1}^{n-1} H(V^{\rightarrow}_{\ell}(s_j))\\
+&= H(T \; | \; \ell, S) + \sum_{j=1}^{n-1}H(V^{\rightarrow}_{\ell}(s_j))
 \end{align*}
@@
 \begin{align*}
-\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V_{\IterateLensOf{\Lens}}(s))
-&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left( \mathbb{H}^{\rightarrow}(S \; | \; \ell, T) + \sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)\\
-&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1})\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \mathbb{H}^{\rightarrow}(S \; | \; \ell, T)\\
-&+(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)\\
-&= \mathbb{H}^{\rightarrow}(S \; | \; \ell, T) * (1-p)\sum_n p^n \\
-&+(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)
+\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V^{\rightarrow}_{\IterateLensOf{\Lens}}(s))
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left( H(S \; | \; \ell, T) + \sum_{j=1}^{n-1}H(V^{\rightarrow}_{\ell}(s_j))\right)\\
+&= (1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1})\\s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) H(S \; | \; \ell, T)\\
+&+(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V^{\rightarrow}_{\ell}(s_j))\right)\\
+&= H(S \; | \; \ell, T) * (1-p)\sum_n p^n \\
+&+(1-p)\sum_n p^n \sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V^{\rightarrow}_{\ell}(s_j))\right)
 \end{align*}
 Unrolling the term 
-$$\sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V_{\ell}(s_j))\right)$$ 
+$$\sum_{\substack{(s_1, \ldots, s_{n-1}) \\ s_i \in \mathcal{L}(S)}} \prod_{i=1}^{n-1}P_S(s_i) \left(\sum_{j=1}^{n-1}H(V^{\rightarrow}_{\ell}(s_j))\right)$$ 
 $(n-1)$ more times gives
-$$\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V_{\IterateLensOf{\Lens}}(s))
-= \mathbb{H}^{\rightarrow}(S \; | \; \ell, T) * (1-p)\sum_n n * p^n  = \frac{p}{1-p} * \mathbb{H}^{\rightarrow}(S \; | \; \ell, T)$$
+\begin{align*}
+H(T^{*q} \; | \; V^{\rightarrow}_{\IterateLensOf{\Lens}}(s), S^{*p}) &=
+\sum_{s \in \mathcal{L}(S^{*p})}P_{S^{*p}}(s) H(V^{\rightarrow}_{\IterateLensOf{\Lens}}(s))\\
+&= H(T \; | \; \ell, S) * (1-p)\sum_n n * p^n \\
+&= \frac{p}{1-p} * H(T \; | \; \ell, S)\\
+&\leq \frac{p}{1-p} * \mathbb{H}^{\rightarrow}(T \; | \; \ell, S)\\
+&= \mathbb{H}(T^{*q} \; | \; V^{\rightarrow}_{\IterateLensOf{\Lens}}(s), S^{*p})
+\end{align*}
+
 which is what we wanted to show.
@@
 \begin{align*}
-V_{ \ConcatLensOf{\Lens_1}{\Lens_2}}(s_1 \cdot s_2) &= \{t_1 \cdot t_2 \in \mathcal{L}(T_1 \cdot T_2) \; | \;  \ConcatLensOf{\Lens_1}{\Lens_2}.\PutROf{(s_1 \cdot s_2)}{(t_1 \cdot t_2)} = t_1 \cdot t_2\}\\
+V^{\rightarrow}_{ \ConcatLensOf{\Lens_1}{\Lens_2}}(s_1 \cdot s_2) &= \{t_1 \cdot t_2 \in \mathcal{L}(T_1 \cdot T_2) \; | \;  \ConcatLensOf{\Lens_1}{\Lens_2}.\PutROf{(s_1 \cdot s_2)}{(t_1 \cdot t_2)} = t_1 \cdot t_2\}\\
 &= \{t_1 \cdot t_2 \in \mathcal{L}(T_1 \cdot T_2) \; | \;  \ell_1.\PutROf{s_1}{t_1 } = t_1  \text{ and } \ell_2.\PutROf{t_2}{t_2} = t_2\}\\
-&= V_{\ell_1}(s_1) \cdot V_{\ell_2}(s_2)
+&= V^{\rightarrow}_{\ell_1}(s_1) \cdot V^{\rightarrow}_{\ell_2}(s_2)
 \end{align*}
@@
 \begin{align*}
-&\sum_{s \in \mathcal{L}(S_1 \cdot S_2)}P_{S_1 \cdot S_2}(s) H(V_{\ConcatLensOf{\Lens_1}{\Lens_2}}(s))\\
-&= \sum_{s_1 \in \mathcal{L}(s_1)} P_{S_1}(s_1)\sum_{s_2 \in \mathcal{L}(s_2)}P_{S_2}(s_2)(H(V_{\ell_1}(s_1)) + H(V_{\ell_2}(s_2)))\\
-&= \sum_{s_1 \in \mathcal{L}(s_1)} P_{S_1}(s_1)\sum_{s_2 \in \mathcal{L}(s_2)}P_{S_2}(s_2)H(V_{\ell_1}(s_1)) + \sum_{s_1 \in \mathcal{L}(s_1)} P_{S_1}(s_1)\sum_{s_2 \in \mathcal{L}(s_2)}P_{S_2}(s_2)H(V_{\ell_2}(s_2))\\
-&= \mathbb{H}^{\rightarrow}(S_1 \; | \; \ell_1, T_1) + \mathbb{H}^{\rightarrow}(S_2 \; | \; \ell_2, T_2)
+&H(T_1 \cdot T_2 \; | \; \ConcatLensOf{\Lens_1}{\Lens_2}, S_1 \cdot S_2)\\
+& =\sum_{s \in \mathcal{L}(S_1 \cdot S_2)}P_{S_1 \cdot S_2}(s) H(V^{\rightarrow}_{\ConcatLensOf{\Lens_1}{\Lens_2}}(s))\\
+&= \sum_{s_1 \in \mathcal{L}(s_1)} P_{S_1}(s_1)\sum_{s_2 \in \mathcal{L}(s_2)}P_{S_2}(s_2)(H(V^{\rightarrow}_{\ell_1}(s_1)) + H(V^{\rightarrow}_{\ell_2}(s_2)))\\
+&= \sum_{s_1 \in \mathcal{L}(s_1)} P_{S_1}(s_1)\sum_{s_2 \in \mathcal{L}(s_2)}P_{S_2}(s_2)H(V^{\rightarrow}_{\ell_1}(s_1)) + \sum_{s_1 \in \mathcal{L}(s_1)} P_{S_1}(s_1)\sum_{s_2 \in \mathcal{L}(s_2)}P_{S_2}(s_2)H(V^{\rightarrow}_{\ell_2}(s_2))\\
+&= H(T_1 \; | \; \ell_1, S_1) + H(T_2 \; | \; \ell_2, S_2)\\
+&\leq \mathbb{H}^{\rightarrow}(T_1 \; | \; \ell_1, S_1) + \mathbb{H}^{\rightarrow}(T_2 \; | \; \ell_2, S_2)\\
+&= \mathbb{H}(T_1 \cdot T_2 \; | \; \ConcatLensOf{\Lens_1}{\Lens_2}, S_1 \cdot S_2)
 \end{align*}
@@
 $$
-V_{ \OrLensOf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \OrLensOf{\Lens_1}{\Lens_2}.\ell.\PutROf{s}{t} = t\}
+V^{\rightarrow}_{ \OrLensOf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \OrLensOf{\Lens_1}{\Lens_2}.\ell.\PutROf{s}{t} = t\}
 $$
@@
 $$
-V_{ \OrLensOf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \ell_1.\PutROf{s}{t} = t\} = V_{\ell_1}(s)
+V^{\rightarrow}_{ \OrLensOf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \ell_1.\PutROf{s}{t} = t\} = V^{\rightarrow}_{\ell_1}(s)
 $$
@@
 $$
-V_{ \OrLensOf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \ell_2.\PutROf{s}{t} = t\} = V_{\ell_2}(s)
+V^{\rightarrow}_{ \OrLensOf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \ell_2.\PutROf{s}{t} = t\} = V^{\rightarrow}_{\ell_2}(s)
 $$
@@
 \begin{align*}
-&\sum_{s \in \mathcal{L}(S_1 \; |_p \; S_2)}P_{S_1 \; |_p \; S_2}(s) H(V_{\OrLensOf{\Lens_1}{\Lens_2}}(s))\\
-&= p * \sum_{s \in \mathcal{L}(S_1)}P_{S_1}(s) H(V_{\ell_1}(s)) + (1-p) * \sum_{s \in \mathcal{L}(S_2)}P_{S_2}(s) H(V_{\ell_2}(s))\\
-&= p * (\mathbb{H}^{\rightarrow}(S_1 \; | \; \ell_1,T_1)) + (1-p) * (\mathbb{H}^{\rightarrow}(S_2 \; | \; \ell_2,T_2))
+H(T_1 \; |_q \; T_2 \; | \; \OrLensOf{\Lens_1}{\Lens_2}, S_1 \; |_p \; S_2) &= \sum_{s \in \mathcal{L}(S_1 \; |_p \; S_2)}P_{S_1 \; |_p \; S_2}(s) H(V^{\rightarrow}_{\OrLensOf{\Lens_1}{\Lens_2}}(s))\\
+&= p * \sum_{s \in \mathcal{L}(S_1)}P_{S_1}(s) H(V^{\rightarrow}_{\ell_1}(s)) + (1-p) * \sum_{s \in \mathcal{L}(S_2)}P_{S_2}(s) H(V^{\rightarrow}_{\ell_2}(s))\\
+&= p * (H(T_1 \; | \; \ell_1,S_1)) + (1-p) * (H(T_2 \; | \; \ell_2,S_2))\\
+&\leq p * (\mathbb{H}^{\rightarrow}(T_1 \; | \; \ell_1,S_1)) + (1-p) * (\mathbb{H}^{\rightarrow}(T_2 \; | \; \ell_2,S_2))\\
+&= \mathbb{H}(T_1 \; |_q \; T_2 \; | \; \OrLensOf{\Lens_1}{\Lens_2}, S_1 \; |_p \; S_2) 
 \end{align*}
@@
 $$
-V_{ \MergeROf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T) \; | \;  \MergeROf{\Lens_1}{\Lens_2}.\PutROf{s}{t} = t\}
+V^{\rightarrow}_{ \MergeROf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T) \; | \;  \MergeROf{\Lens_1}{\Lens_2}.\PutROf{s}{t} = t\}
 $$
@@
 $$
-V_{ \MergeROf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T) \; | \;  \ell_1.\PutROf{s}{t} = t\} = V_{\ell_1}(s)
+V^{\rightarrow}_{ \MergeROf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T) \; | \;  \ell_1.\PutROf{s}{t} = t\} = V^{\rightarrow}_{\ell_1}(s)
 $$
@@
 $$
-V_{ \MergeROf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T) \; | \;  \ell_2.\PutROf{s}{t} = t\} = V_{\ell_2}(s)
+V^{\rightarrow}_{ \MergeROf{\Lens_1}{\Lens_2}}(s) = \{t \in \mathcal{L}(T) \; | \;  \ell_2.\PutROf{s}{t} = t\} = V^{\rightarrow}_{\ell_2}(s)
 $$
@@
 \begin{align*}
-&\sum_{s \in \mathcal{L}(S_1 \; |_p \; S_2)}P_{S_1 \; |_p \; S_2}(s) H(V_{\MergeROf{\Lens_1}{\Lens_2}}(s))\\
-&= p * \sum_{s \in \mathcal{L}(S_1)}P_{S_1}(s) H(V_{\ell_1}(s)) + (1-p) * \sum_{s \in \mathcal{L}(S_2)}P_{S_2}(s) H(V_{\ell_2}(s))\\
-&= p * (\mathbb{H}^{\rightarrow}(S_1 \; | \; \ell_1,T_1) ) + (1-p) * (\mathbb{H}^{\rightarrow}(S_2 \; | \; \ell_2,T_2) )
+H(T \; | \; \MergeROf{\Lens_1}{\Lens_2}, S_1 \; |_p \; S_2) &= \sum_{s \in \mathcal{L}(S_1 \; |_p \; S_2)}P_{S_1 \; |_p \; S_2}(s) H(V^{\rightarrow}_{\MergeROf{\Lens_1}{\Lens_2}}(s))\\
+&= p * \sum_{s \in \mathcal{L}(S_1)}P_{S_1}(s) H(V^{\rightarrow}_{\ell_1}(s)) + (1-p) * \sum_{s \in \mathcal{L}(S_2)}P_{S_2}(s) H(V^{\rightarrow}_{\ell_2}(s))\\
+&= p * (H(T_1 \; | \; \ell_1, S_1) ) + (1-p) * (H(T_2 \; | \; \ell_2, S_2) )\\
+&\leq p * (\mathbb{H}^{\rightarrow}(T_1 \; | \; \ell_1, S_1) ) + (1-p) * (\mathbb{H}^{\rightarrow}(T_2 \; | \; \ell_2, S_2) )\\
+&= \mathbb{H}(T \; | \; \MergeROf{\Lens_1}{\Lens_2}, S_1 \; |_p \; S_2) 
 \end{align*}
@@
 \begin{align*}
-V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s) &= \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \MergeLOf{\Lens_1}{\Lens_2}.\PutROf{s}{t} = t\}\\
+V^{\rightarrow}_{ \MergeLOf{\Lens_1}{\Lens_2}}(s) &= \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \MergeLOf{\Lens_1}{\Lens_2}.\PutROf{s}{t} = t\}\\
 &= \{t \in \mathcal{L}(T_1 \; |_q \; T_2) \; | \;  \ell_1.\PutROf{s}{t} = t \text{ or } \ell_2.\PutROf{s}{t} = t\}\\
-&= V_{\ell_1}(s) \cup V_{\ell_2}(s)
+&= V^{\rightarrow}_{\ell_1}(s) \cup V^{\rightarrow}_{\ell_2}(s)
 \end{align*}
-To compute $H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))$, we use the formula 
+To compute $H(V^{\rightarrow}_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))$, we use the formula 
 $$H(P_1, P_2) = H(m(P_1), m(P_2)) + m(P_1)H\left(\frac{P_1}{m(P_1)}\right) + m(P_2)H\left(\frac{P_2}{m(P_2)}\right)$$
@@
 \begin{align*}
-&H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)) = H \left(\frac{q * P_{T_1}V_{\ell_1}(s)}{P_{T_1 \; |_q \; T_2}{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}}, \frac{(1-q) * P_{T_2}V_{\ell_2}(s)}{P_{T_1 \; |_q \; T_2}{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}}\right)\\
-&= \frac{q * P_{T_1}V_{\ell_1}(s)}{P_{T_1 \; |_q \; T_2}{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}} H(V_{\ell_1}(s)) + \frac{(1-q) * P_{T_2}V_{\ell_2}(s)}{P_{T_1 \; |_q \; T_2}{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}} H(V_{\ell_2}(s))
-\end{align*}
-
-\iffalse
-\begin{align*}
--H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)) &= \sum_{t \in V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}(t) \log_2 P_{V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}(t)\\
-&= \sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \\
-&+ \sum_{t \in \mathcal{L}(T_2) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
-\end{align*}
-We focus on the term 
-$$\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}$$
-Observe that
-\begin{align*}
-&\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
-&= \frac{q}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t) \log_2 q * P_{T_1}(t)\\
-&- \frac{q * \log_2 P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t)
-\end{align*}
-Again we focus on the term
-$$\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t) \log_2 q * P_{T_1}(t)$$
-Observe that
-\begin{align*}
-&\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t) \log_2 q * P_{T_1}(t) = \log_2 q * P_{T_1}(V_{\ell_1}(s))\\
-& + P_{T_1}(V_{\ell_1}(s))\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{P_{T_1}(t)}{P_{T_1}(V_{\ell_1}(s))} \log_2 \left( P_{T_1}(V_{\ell_1}(s)) *  \frac{P_{T_1}(t)}{P_{T_1}(V_{\ell_1}(s))} \right)\\
-&= \log_2 q * P_{T_1}(V_{\ell_1}(s)) + \log_2 P_{T_1}(V_{\ell_1}(s)) * P_{T_1}(V_{\ell_1}(s)) + P_{T_1}(V_{\ell_1}(s))* -H(V_{\ell_1}(s))\\
-&= P_{T_1}(V_{\ell_1}(s))[\log_2 q + \log_2 P_{T_1}(V_{\ell_1}(s)) - H(V_{\ell_1}(s))]
-\end{align*}
-Hence
-\begin{align*}
-&\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{q * P_{T_1}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \\
-&= \frac{q}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t) \log_2 q * P_{T_1}(t)\\
-&- \frac{q * \log_2 P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\sum_{t \in \mathcal{L}(T_1) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}P_{T_1}(t)\\
-&= \frac{q * P_{T_1}(V_{\ell_1}(s))[\log_2 q + \log_2 P_{T_1}(V_{\ell_1}(s)) - H(V_{\ell_1}(s))]}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
-&- \frac{q * \log_2 P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)) * P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
-&= \frac{q * P_{T_1}(V_{\ell_1}(s)) \left[\log_2 q + \log_2 \left[\frac{P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right]  -H(V_{\ell_1}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
-\end{align*}
-Similarly,
-\begin{align*}
-&\sum_{t \in \mathcal{L}(T_2) \cap V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)}\frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))} \log_2 \frac{(1-q) * P_{T_2}(t)}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
-&= \frac{(1-q) * P_{T_2}(V_{\ell_2}(s)) \left[\log_2 (1-q) + \log_2 \left[\frac{P_{T_2}(V_{\ell_2}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] - H(V_{\ell_2}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+&H(V^{\rightarrow}_{ \MergeLOf{\Lens_1}{\Lens_2}}(s)) =\\
+&H \left(\frac{q * P_{T_1}V^{\rightarrow}_{\ell_1}(s)}{q * P_{T_1}V^{\rightarrow}_{\ell_1}(s) + (1-q) * P_{T_2}V^{\rightarrow}_{\ell_2}(s)}, \frac{(1-q) * P_{T_2}V^{\rightarrow}_{\ell_2}(s)}{q * P_{T_1}V^{\rightarrow}_{\ell_1}(s) + (1-q) * P_{T_2}V^{\rightarrow}_{\ell_2}(s)}\right)\\
+&+ \frac{q * P_{T_1}V^{\rightarrow}_{\ell_1}(s)}{q * P_{T_1}V^{\rightarrow}_{\ell_1}(s) + (1-q) * P_{T_2}V^{\rightarrow}_{\ell_2}(s)} H(V^{\rightarrow}_{\ell_1}(s)) + \frac{(1-q) * P_{T_2}V^{\rightarrow}_{\ell_2}(s)}{q * P_{T_1}V^{\rightarrow}_{\ell_1}(s) + (1-q) * P_{T_2}V^{\rightarrow}_{\ell_2}(s)} H(V^{\rightarrow}_{\ell_2}(s))
 \end{align*}
@@
 \begin{align*}
-&-H(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))\\
-&= \frac{q * P_{T_1}(V_{\ell_1}(s)) \left[\log_2 \left[\frac{q * P_{T_1}(V_{\ell_1}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] - H(V_{\ell_1}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\\
-&+ \frac{(1-q) * P_{T_2}(V_{\ell_2}(s)) \left[\log_2 \left[\frac{(1-q) * P_{T_2}(V_{\ell_2}(s))}{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}\right] - H(V_{\ell_2}(s)) \right] }{P(V_{ \MergeLOf{\Lens_1}{\Lens_2}}(s))}
+&H(T_1 \; |_q \; T_2 \; | \; \MergeLOf{\Lens_1}{\Lens_2}, S) = \\
+&\sum_{s \in \mathcal{L}(S)}H \left(\frac{q * P_{T_1}V^{\rightarrow}_{\ell_1}(s)}{q * P_{T_1}V^{\rightarrow}_{\ell_1}(s) + (1-q) * P_{T_2}V^{\rightarrow}_{\ell_2}(s)}, \frac{(1-q) * P_{T_2}V^{\rightarrow}_{\ell_2}(s)}{q * P_{T_1}V^{\rightarrow}_{\ell_1}(s) + (1-q) * P_{T_2}V^{\rightarrow}_{\ell_2}(s)}\right)P_S(s)\\
+&+ \sum_{s \in \mathcal{L}(S)}\left(\frac{q * P_{T_1}V^{\rightarrow}_{\ell_1}(s)}{q * P_{T_1}V^{\rightarrow}_{\ell_1}(s) + (1-q) * P_{T_2}V^{\rightarrow}_{\ell_2}(s)}\right)P_S(s)H(V^{\rightarrow}_{\ell_1}(s))\\
+&+ \sum_{s \in \mathcal{L}(S)}\left(\frac{(1-q) * P_{T_2}V^{\rightarrow}_{\ell_2}(s)}{q * P_{T_1}V^{\rightarrow}_{\ell_1}(s) + (1-q) * P_{T_2}V^{\rightarrow}_{\ell_2}(s)}\right)P_S(s)H(V^{\rightarrow}_{\ell_2}(s))
 \end{align*}
-\fi
+Since
+\begin{align*}
+H \left(\frac{q * P_{T_1}V^{\rightarrow}_{\ell_1}(s)}{q * P_{T_1}V^{\rightarrow}_{\ell_1}(s) + (1-q) * P_{T_2}V^{\rightarrow}_{\ell_2}(s)}, \frac{(1-q) * P_{T_2}V^{\rightarrow}_{\ell_2}(s)}{q * P_{T_1}V^{\rightarrow}_{\ell_1}(s) + (1-q) * P_{T_2}V^{\rightarrow}_{\ell_2}(s)}\right) &\leq 1\\
+\left(\frac{q * P_{T_1}V^{\rightarrow}_{\ell_1}(s)}{q * P_{T_1}V^{\rightarrow}_{\ell_1}(s) + (1-q) * P_{T_2}V^{\rightarrow}_{\ell_2}(s)}\right) &\leq 1\\
+\left(\frac{(1-q) * P_{T_2}V^{\rightarrow}_{\ell_2}(s)}{q * P_{T_1}V^{\rightarrow}_{\ell_1}(s) + (1-q) * P_{T_2}V^{\rightarrow}_{\ell_2}(s)}\right) &\leq 1
+\end{align*}
+Then
+\begin{align*}
+H(T_1 \; |_q \; T_2 \; | \; \MergeLOf{\Lens_1}{\Lens_2}, S) & \leq \sum_{s \in \mathcal{L}(s)}P_S(s) + \sum_{s \in \mathcal{L}(s)}P_S(s)H(V^{\rightarrow}_{\ell_1}(s)) + \sum_{s \in \mathcal{L}(s)}P_S(s)H(V^{\rightarrow}_{\ell_2})(s)\\
+&= H(T_1 \; | \; \ell_1, S) + H(T_2 \; | \; \ell_2, S)\\
+&\leq 1 + \mathbb{H}^{\rightarrow}(T_1 \; | \; \ell_1, S) + \mathbb{H}^{\rightarrow}(T_2 \; | \; \ell_2, S)
+\end{align*}
+\item
+($S = \REntropyOf{\Regex \Given \InvertOf{\Lens}, \RegexAlt}$)
+
+Observe that $$V^{\rightarrow}_{\InvertOf{\Lens}}(t) = \{s \in \mathcal{L}(S) \; | \; \InvertOf{\Lens}.\PutROf{s}{t} = t\} = \{s \in \mathcal{L}(S) \; | \; \ell.\PutLOf{t}{s} = s\} = V^{\leftarrow}_{\ell}(t)$$
+Consequently
+$$
+H(S \; | \; \InvertOf{\Lens}, T) = \sum_{t \in \mathcal{L}(T)}P_T(t) H(V^{\rightarrow}_{\InvertOf{\Lens}}(t))
+= \sum_{t \in \mathcal{L}(T)}P_T(t) H(V^{\leftarrow}_{\ell}(t))
+\leq \mathbb{H}^{\leftarrow}(S\; | \; \ell, T)
+$$
 \end{enumerate}

### 2018-07-08 papers/SymmetricLenses/main.tex
add impl revisions

--- 2018-07-08 papers/SymmetricLenses/main.tex
@@
 
-To obtain SREs from the plain regular expressions that
-we expect users to supply, we use a heurstic that attempts to assign probability
-annotations that give each string in the language equal probablity. We can also
-take as input a stochastic regular expression rather than a plain one, which
-allows the user to specify a data-specific probability distribution.
-Alternatively, the user could provide a large suite of example data from which
-we could infer the necessary annotations to model the supplied data. We found
-our heuristically chosen distribution works well in practice.
+To obtain SREs from the plain regular expressions that we expect users to
+supply, we use a heurstic that attempts to assign probability annotations that
+give each string in the language equal probablity. Our algorithm could also take
+as input a stochastic regular expression rather than a plain one, which allows
+the user to specify a data-specific probability distribution, either manually
+written or inferred from a large dataset. We found our heuristically chosen
+distribution works well in practice.
 
@@
   The proofs of this theorem, and subsequent theorems in this paper, appears in
-  the auxiliary materials accompanying this submission.\bcp{Not in the appendix?}
+  the appendix present in the full version of the paper accompanying this
+  submission.
 \end{proof}
@@
 of a stochastic regular expression can be computed directly from its syntax,
-when the stochastic regular expression is unambiguous and contains no nonempty\bcp{empty?}
+when the stochastic regular expression is unambiguous and contains no empty
 subcomponents.
@@
 The \MergeR lens is interesting because it merges data where one data can be in
-two formats, and one data has only one format. In previous work\bcp{where?},
-this was
-combined into \OrLens{}, where \OrLens{} could have ambiguous types, but we
+two formats, and one data has only one format. In previous work~\cite{?}, this
+was combined into \OrLens{}, where \OrLens{} could have ambiguous types, but we
 find it more clear to have explicit merge operators: it is easier to see what
@@
     \EndWhile
-    \ReturnVal{\Call{RXSearch.Best}{\RXSearchState}}
+    \State \ReturnVal{\Call{RXSearch.Best}{\RXSearchState}}
     \EndFunction
   \end{algorithmic}
-\bcp{Why does the RETURN statement not get a number?}
 \end{algorithm}
@@
 \RXSearch proposes rewrites incrementally, where it proposes regular expression
-pairs with the smallest number of rewrites first, and when there are no more
-proposed regular expressions of that size, \RXSearch proposes regular
-expressions of larger size.
+pairs with the smallest number of rewrites first. When there are no more
+proposed regular expressions derived from $n$ rewrites, \RXSearch proposes
+regular expressions derived from $n+1$ rewrites.
 
@@
 best lens found thus far. Note this means that if \RXSearch finds a bijective
-lens (a lens of cost 0\bcp{have we proved that the bijective lenses are
-  exactly the ones with zero cost?  That would be nice to state clearly someplace...}) it will immediately return.
+lens, which has zero cost it will immediately return.
 %
@@
 \ToDNFRegex defined, it is easier to explain the function that converts regular
-expressions into stochastic regular expressions, \ToStochastic\bcp{from the
-  description above, I thought this step was optional (because we allowed
-  the user to write SREs directly)}. If $\Regex$ is a
-stochastic DNF regular expression generated by \ToStochastic, then
-$\ToDNFRegexOf{\Regex} = \DNFOf{(\Sequence_1,\frac{1}{n}) \DNFSep \ldots \DNFSep
+expressions into stochastic regular expressions, \ToStochastic. If users
+specified stochastic regular expressions themselves, this step could be skipped,
+though our integration into Boomerang does not permit users to provide
+stochastic regular expressions. If $\Regex$ is a stochastic DNF regular
+expression generated by \ToStochastic, then $\ToDNFRegexOf{\Regex} =
+\DNFOf{(\Sequence_1,\frac{1}{n}) \DNFSep \ldots \DNFSep
   (\Sequence_n,\frac{1}{n})}$ for some sequences $\Sequence_1\ldots\Sequence_n$,
@@
 $\PRegexStar{\DNFRegex}{.8}$. In particular, $\ToDNFRegex$ generates regular
-expressions whose DNF form gives equal probability to all sequences\bcp{Why?}\bcp{Also, what does this mean?  If a given RE denotes an
-infinite language, then there is no way to give equal probability to every
-string in that language, is there???}, and gives
-iterations a .8 chance to continue iterating. In our experience generating
-random strings from regular expressions\bcp{what does generating random
-  strings have to do with synthesis? i.e., why is this a good thing to measure?}, these probabilities provide good
-distributions of strings -- stars are iterated 4 times on average, and no
-individual choice in a series disjunctions is chosen disproportionately often.
+expressions whose DNF form gives equal probability to all sequence subcomponents
+of the DNF regular expression., and gives iterations a .8 chance to continue
+iterating. In our experience generating random strings from regular expressions,
+these probabilities provide good distributions of strings -- stars are iterated
+4 times on average, and no individual choice in a series disjunctions is chosen
+disproportionately often. As we have found these distributions generate
+reasonable strings, we believe they describe the shape of data well.
 
@@
 Like stochastic DNF regular expressions, symmetric DNF Lenses are hierarchically
-structured. Symmetric DNF lenses consist of a list of symmetric sequence lenses,
-as well as some additional information. Symmetric sequence lenses consist of a
-list of symmetric atom lenses\bcp{why are they called atom lenses?}, as well
-as some additional information, and
-symmetric atom lenses are merely iterations of symmetric DNF lenses. In the
+structured. Symmetric DNF lenses, which provide lens functions between DNF
+regular expressions, consist of a list of symmetric sequence lenses, as well as
+some additional information. Symmetric sequence lenses, which provide lens
+functions betwen sequences, consist of a list of symmetric atom lenses, as well
+as some additional information. Symmetric atom lenses, which provide lens
+functions between atoms, are merely iterations of symmetric DNF lenses. In the
 following syntax, the nonterminals $i$, $j$, $p$, $q$, $r$, $c$, and $d$ all
@@
 \[\begin{array}{rl}
-      & \REntropyOf{\PRegexStar{\DNFRegex}{\Probability} \Given \StarOf{\DNFRegexAlt},\StarOf{\SDNFLens}}\\
-      =\\
-      & 4*\REntropyOf{\DNFRegex\Given\DNFRegexAlt,\SDNFLens}
+      & \REntropyOf{\PRegexStar{\DNFRegexAlt}{\ProbabilityAlt} \Given \PRegexStar{\DNFRegex}{\Probability},\StarOf{\SDNFLens}}\\
+      =
+      & \frac{\Probability}{1-\Probability}\REntropyOf{\DNFRegexAlt\Given\DNFRegex,\SDNFLens}
   \end{array}
@@
 \label{subsec:greedy-synth}
-\bcp{The following is all pretty incomprehensibly dense.  A little running example
-might help.}
-
 The greedy algorithm is actually two greedy algorithms, one that greedily finds
@@
 sequence lenses (\GreedySeqSynth).
+We describe this synthesis algorithm in the context of trying to find a lens
+between \lstinline{"" | name.name*} and \lstinline{"" | name}, with the example
+\lstinline{creater "John Doe" = "John"}.
 
@@
 \GreedySynth are a pair of stochastic DNF regular expressions and a suite of
-input-output examples. First, \PCF{CannotMap} checks if there could possibly be
-a lens satisfying the examples. If there is not, \GreedySynth returns \None
-early. After this, \GreedySynth finds the best lenses between each sequence pair
-of the left and right DNF regular expressions, given the examples that match
-them. An internal state for the algorithm is initialized from these. This state
-consists of two things: an intermediate partial DNF Lens and a priority queue
-consisting of potential sequence lenses. On initialization, \GreedySynth checks
-which sequence lenses are \emph{required} to be included, from the examples.
-Those are added into the partial lens, and the rest are thrown into the priority
-queue. For example, an example if an input output example is provided, the
-sequences that match those provided strings must be mapped to each other with a
-sequence lens. \GreedySynth then loops until there are no more sequence lenses
-in the priority queue. Within this loop, sequence lenses are popped. If a popped
-sequence lens maps between two sequences which aren't yet included in a sequence
-lens, then this sequence lens is added to the partial DNF lens. Furthermore, we
-update the priorities of our sequence lenses as the algorithm proceeds: if two
-sequence lenses have the same source, the second one to be popped gets a higher
-cost than it originally had, to address the information that needs to be stored
-for including that source of non-bijectivity. Finally, after all sequences have
-been popped, the partial DNF lens is converted into a symmetric DNF lens. This
-is only possible if all sequences are involved in some sequence lens: if they
-are not, \PCF{GreedyState.ToDNFLens} instead returns \None.
+input-output examples.
+
+First, \PCF{CannotMap} checks if there could possibly be a lens satisfying the
+examples. If there is not, \GreedySynth returns \None early. After this,
+\GreedySynth finds the best lenses between each sequence pair of the left and
+right DNF regular expressions, given the examples that match them. An internal
+state for the algorithm is initialized from these. This state consists of two
+things: an intermediate partial DNF Lens and a priority queue consisting of
+potential sequence lenses.
+
+\GreedySynth then loops until there are no more sequence lenses in the priority
+queue. Within this loop, sequence lenses are popped. If a popped sequence lens
+maps between two sequences which aren't yet included in a sequence lens, then
+this sequence lens is added to the partial DNF lens. Furthermore, we update the
+priorities of our sequence lenses as the algorithm proceeds: if two sequence
+lenses have the same source, the second one to be popped gets a higher cost than
+it originally had, to address the information that needs to be stored for
+including that source of non-bijectivity.
+
+In our example, \GreedySynth may first pop the sequence pair \lstinline{""} and
+\lstinline{""}, as there is a bijective sequence lens between them. After this,
+the sequence lens between \lstinline{name.name*} and \lstinline{name} would be
+popped. While examples prevent lenses between the other sequence pairs, if there
+were other lenses, they would be popped, but not added to the ongoing state, as
+they would not be a useful addition to the state.
+
+Finally, after all sequences have been
+popped, the partial DNF lens is converted into a symmetric DNF lens. This is
+only possible if all sequences are involved in some sequence lens: if they are
+not, \PCF{GreedyState.ToDNFLens} instead returns \None.
 
@@
     \EndWhile
-    \ReturnVal{\Call{GreedyState.ToDNFLens}{\GreedyState}}
+    \State \ReturnVal{\Call{GreedyState.ToDNFLens}{\GreedyState}}
     \EndFunction
@@
 \paragraph*{Symmetric Sequence lenses} Algorithm~\ref{alg:greedyseqsynth}
-presents \GreedySeqSynth, a means to synthesizes symmetric sequence lenses. This
-algorithm uses \PCF{AtomSynth}, a simple function that synthesizes atom lenses
-by iterating the DNF lens between its subcomponents. The inputs to
+presents \GreedySeqSynth, a means to synthesizes symmetric sequence lenses.
+
+This algorithm uses \PCF{AtomSynth}, a simple function that synthesizes atom
+lenses by iterating the DNF lens between its subcomponents. The inputs to
 \GreedySeqSynth are a pair of stochastic atoms and a suite of input-output
@@
 
+For example, if trying to find a sequence lens between \lstinline{name.name*}
+and \lstinline{name}, lenses are found between \lstinline{name} and
+\lstinline{name}, and \lstinline{name} and \lstinline{name*}, and added to the
+queue.
+
 \GreedySeqSynth then loops until there are no more atom lenses in the priority
@@
 other lens is more than is removed by adding the proposed lens, then this atom
-lens is not included in the sequence lens. Finally, after all atoms have been
+lens is not included in the sequence lens. For example, the bijective lens
+between \lstinline{name} and \lstinline{name} would first be found, and added to
+the state.  Next, the disconnect lens between \lstinline{name} and
+\lstinline{name*} would be popped.  Because this lens is not useful, it is not
+added to the state.
+
+Finally, after all atoms have been
 popped, the sequence lens is validated and returned. The sequence lens is
@@
     \EndWhile
-    \ReturnVal{\Call{GreedySeqState.ToDNFLens}{\GreedyState}}
+    \State \ReturnVal{\Call{GreedySeqState.ToDNFLens}{\GreedyState}}
     \EndFunction
@@
 
-\paragraph*{\RXSearch Permissiveness Customization}\bcp{I don't follow this:}
-While the termination heuristic works for many situations, there are situations
-where a large number of expansions must be performed to find the correct lens.
-While users could prevent incorrect lenses with a large number of examples, we
+\paragraph*{\RXSearch Permissiveness Customization}
+In this work, we are constrained by three competing constraints. (1) We want to
+return lenses quickly to users, (2) we want to return good lenses, and (3) we
+don't want to require users to write a large number of examples in the
+specifications. Our termination heuristic works well for most situations, but
+there are situations where a large number of expansions must be performed to
+find the correct lens.
+
+Users could prevent incorrect lenses with a large number of examples, but it can
+be quite difficult to write a large number of examples for complex formats. We
 find it simpler to let the users specify that \RXSearch should be more
-restrictive in which lenses it accepts. If the user's desired transformation has
+restrictive in which lenses it accepts, so users do not have to write a lot of
+examples.
+
+This constraint goes the other way too. If the user's desired transformation has
 a particularly high cost, the algorithm may spend a long time searching in vain
@@
 lenses it accepts, this search can be short-circuited.
+
+While our defaults termination condition quickly finds the correct lens for
+\afm{35} of our \afm{45} benchmarks, our termination condition is not
+one-size-fits all. Particularly complex lenses, or lenses that project a
+particularly large amount of data, can be synthesized more easily and more
+quickly by customizing how permissive \RXSearch is in returning lenses.
 % end synthesis

### 2018-07-08 papers/SymmetricLenses/main.tex

@@
 extension to Boomerang. We added symmetric lens combinators to Boomerang, and
-implemented Boomerang's previous asymmetric lens combinators within the
-symmetric lens combinators.
+implemented Boomerang's asymmetric lens combinators \sam{all of Boomerang's combinators?} using the symmetric lens combinators.
 
@@
 \item How much slower do we perform on bijective lens synthesis benchmarks than
-  the existing Optician system customized for bijective lenses.
-
-\item Which heuristics we have described useful for reducing the ``specification
+  the existing Optician system customized for bijective lenses?
+
+\item Which heuristics are useful for reducing the ``specification
   overhead'' for users? In particular, which heuristics are useful in
   synthesizing the desired lens, instead of merely a lens that satisfies the
-  specification.
+  specification?
 \end{enumerate}
@@
   ad-hoc data present in Linux configuration files into a structured, dictionary
-  representation.
-\item We used \afm{8} benchmarks from synchronization tasks on synchronizing a
-  variety of real-life formats. These formats range from the synchronizing REST
+  representation. \sam{What does data cleaning mean?}
+\item We used \afm{8} benchmarks from synchronization tasks between a
+  variety of real-life formats. These tasks range from synchronizing REST
   and JSON web resource descriptions to synchronizing BibTeX and EndNote
@@
 
-A realistic benchmarks suite is cretical in answering the first research
+A realistic benchmarks suite is critical in answering the first research
 question. To determine whether our algorithm is fast enough to be used in

### 2018-07-08 papers/SymmetricLenses/main.tex
stuff

@@
 \label{sec:synthesis}
-Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm. The
+Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm at a
+high level of abstraction. The
 input regular expressions are first converted into stochastic regular
-expressions with \ToStochastic. Next, $\RXSearch$ is initialized with that pair
-of stochastic regular expressions. $\RXSearch$ then proposes candidate classes
-of lenses by proposing regular expression pairs that characterize those classes.
-These lens classes are proposed based on the distance from the original
-stochastic regular expressions: classes characterized by stochastic regular
-expressions closer to the originals are proposed earlier. $\GreedySynth$
-greedily finds a low cost lens within each proposed lens class, if there is one
-that satisfies the specification. These lenses are then provided to $\RXSearch$.
-Eventually, $\RXSearch$ terminates and returns the best lens found thus far.
-
+expressions with \ToStochastic. Once that conversion has taken place,
+the $\RXSearch$ algorithm executes.
+This algorithm itself is defined via
+five functions:
+(1) $\PCF{\RXSearch.Create}$,
+(2) $\PCF{\RXSearch.Continue}$,
+(3) $\PCF{\RXSearch.Propose}$,
+(4) $\PCF{\RXSearch.Enqueue}$, and
+(5) $\PCF{\RXSearch.Best}$.
+First, $\PCF{\RXSearch.Create}$ creates a priority queue ($rxs$)
+containing a single element:  the initial pair of stochastic regular
+expressions.  Next, 
+$\RXSearch$ enters a loop that searches for likely lenses.  The loop
+terminates when the algorithm believes it is unlikely
+to find a better lens than the best one it has found so far (a
+termination condition defined by $\PCF{\RXSearch.Continue}$).
+Within each iteration of the loop, it:
+\begin{itemize}
+\item generates the next class ($S$, $T$) of lenses to
+search ($\PCF{\RXSearch.Propose}$),
+\item executes $\GreedySynth$ to find a best lens in that class if
+one exists ($lco$),
+using the examples $exs$ to filter out potential lenses that do not satisfy
+the specification, and
+\item adds this best lens to the set of possibilities
+under consideration $\PCF{\RXSearch.Add}$
+\end{itemize}
+When the loop terminates, the search returns the globally best lens found
+($\PCF{\RXSearch.Best}$).  Each element of this algorithm will be explained in
+further depth in the following subsections.
+
+\dpw{Note: I added overbars to regexps in Algorithm 1 below ... double check me
+for correctness.}
 \begin{algorithm}
@@
   \begin{algorithmic}[1]
-    \Function{SynthSymLens}{$\Regex,\RegexAlt,\Examples$}
-    \State $\Regex \gets \Call{\ToStochastic}{\Regex}$
-    \State $\RegexAlt \gets \Call{\ToStochastic}{\RegexAlt}$
-    \State $\RXSearchState \gets \Call{\RXSearch.Init}{\Regex,\RegexAlt}$
+    \Function{SynthSymLens}{$\BRegex,\BRegexAlt,\Examples$}
+    \State $\Regex \gets \Call{\ToStochastic}{\BRegex}$
+    \State $\RegexAlt \gets \Call{\ToStochastic}{\BRegexAlt}$
+    \State $\RXSearchState \gets \Call{\RXSearch.Create}{\Regex,\RegexAlt}$
     \While{$\Call{\RXSearch.Continue}{\RXSearchState}$}
@@
     \State $\LCO \gets \Call{\GreedySynth}{\Examples,\Regex,\RegexAlt}$
-    \State $\RXSearchState \gets \Call{\RXSearch.Consider}{\RXSearchState,\LCO}$
+    \State $\RXSearchState \gets \Call{\RXSearch.Add}{\RXSearchState,\LCO}$
     \EndWhile
@@
 
-\subsection{\RXSearch}
+\subsection{\PCF{\RXSearch.Propose}}
 The first synthesizer is one that searches through regular expression pairs.
@@
 \subsection{Stochastic DNF Regular Expressions}
-Stochastic DNF regular expressions are a highly normalized form of stochastic
-regular expressions. Due to this highly normalized nature, where type
-equivalence is often needed for simple symmetric lenses, such rules aren't
-needed when typed between DNF regular expressions, as many semantically equal
-but syntactically distinct regular expressions have syntactically equal DNF
-forms.  Like how stochastic regular expressions are regular expressions with
-probability annotations, stochastic DNF regular expressions are DNF regular
-expressions with probability annotations.  Like with SREs and regular
+Stochastic DNF regular expressions (\SDNFREabbrev) are a pseudo-normal form
+for stochastic regular expressions.  Of all the star-semiring equations
+required to relate two \SDNFREabbrevs, only those relating to Kleene star
+(\EG, \emph{Unrollstar$_L$} and \emph{Unrollstar$_R$}) are needed to relate
+two \SDNFREabbrevs, other axioms, like associativity, commutativity
+or distributivity, are unneeded.  This important property makes it possible to
+construct an efficient algorithm ($\GreedySynth$), directed
+by the syntax of two \SDNFREabbrevs, to synthesize any lenses in the
+same ``class'' (\IE, any lenses that have the same type when the
+Kleene star axioms are unused).  This property, and the resulting
+efficiency of $\GreedySynth$ is key to the effectiveness of our
+core algorithms.
+
+Just as SREs are REs with probability annotations,
+\SDNFREabbrevs are \DNFREabbrevs with probability annotations.
+As with SREs and regular
 expressions, if $\DNFRegex$ is a stochastic DNF regular expression,

### 2018-07-08 papers/SymmetricLenses/main.tex
minor changes
edit

@@
     & =
-    & $\ProbabilityAlt*(\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1}) + $\\
-    &
-    & $(1-\ProbabilityAlt)*(\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2} - \Log_2(1 - \ProbabilityAlt))+$\\
-    &
-    & $(0, -\Log_2(\ProbabilityAlt)*\ProbabilityAlt -\Log_2(1 - \ProbabilityAlt)*(1-\ProbabilityAlt)]$\\
+    & $(0,1+\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1}+\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2})$\\
 
@@
 expressions into stochastic regular expressions, \ToStochastic. If users
-specified stochastic regular expressions themselves, this step could be skipped,
-though our integration into Boomerang does not permit users to provide
+specified stochastic regular expressions themselves (perhaps inferred from
+example data), this step could be skipped,
+though our implementation does not (yet) permit users to provide such
 stochastic regular expressions. If $\Regex$ is a stochastic DNF regular

### 2018-07-09 papers/SymmetricLenses/main.tex
fix algs

--- 2018-07-08 papers/SymmetricLenses/main.tex
@@
     & =
-    & $\ProbabilityAlt*(\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1}) + $\\
-    &
-    & $(1-\ProbabilityAlt)*(\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2} - \Log_2(1 - \ProbabilityAlt))+$\\
-    &
-    & $(0, -\Log_2(\ProbabilityAlt)*\ProbabilityAlt -\Log_2(1 - \ProbabilityAlt)*(1-\ProbabilityAlt)]$\\
+    & $(0,\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1}+\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2}+1)$\\
 
@@
 
-The cost of a lens between two SREs $c(\Lens, \Regex, \RegexAlt) =
+The cost of a lens between two SREs $cost(\Lens, \Regex, \RegexAlt) =
 \MaxOf{\LEntropyOf{\Regex \Given \Lens, \RegexAlt}} +
@@
 \label{sec:synthesis}
-Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm at a
-high level of abstraction. The
-input regular expressions are first converted into stochastic regular
-expressions with \ToStochastic. Once that conversion has taken place,
-the $\RXSearch$ algorithm executes.
-This algorithm itself is defined via
-five functions:
-(1) $\PCF{\RXSearch.Create}$,
-(2) $\PCF{\RXSearch.Continue}$,
-(3) $\PCF{\RXSearch.Propose}$,
-(4) $\PCF{\RXSearch.Enqueue}$, and
-(5) $\PCF{\RXSearch.Best}$.
-First, $\PCF{\RXSearch.Create}$ creates a priority queue ($rxs$)
-containing a single element:  the initial pair of stochastic regular
-expressions.  Next, 
-$\RXSearch$ enters a loop that searches for likely lenses.  The loop
-terminates when the algorithm believes it is unlikely
-to find a better lens than the best one it has found so far (a
-termination condition defined by $\PCF{\RXSearch.Continue}$).
-Within each iteration of the loop, it:
+Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm at a high
+level of abstraction. The input regular expressions are first converted into
+stochastic regular expressions with \ToStochastic. Once that conversion has
+taken place, the $\RXSearch$ algorithm executes. This algorithm itself is
+defined via five functions: (1) $\PCF{\PQ.Create}$, (2) $\PCF{\PQ.Pop}$, (3)
+$\PCF{Expand}$, (4) $\PCF{\PQ.Push}$, and (5) $\PCF{\Continue}$. First,
+$\PCF{\PQ.Create}$ creates a priority queue ($pq$) containing a single element:
+the initial pair of stochastic regular expressions. Next, $\PCF{SynthSymLens}$
+enters a loop that searches for likely lenses. The loop terminates when the
+algorithm believes it is unlikely to find a better lens than the best one it has
+found so far (a termination condition defined by $\PCF{\Continue}$). Within each
+iteration of the loop, it:
 \begin{itemize}
 \item generates the next class ($S$, $T$) of lenses to
-search ($\PCF{\RXSearch.Propose}$),
+search ($\PCF{\PQ.Pop}$),
 \item executes $\GreedySynth$ to find a best lens in that class if
-one exists ($lco$),
+one exists ($\LC$),
 using the examples $exs$ to filter out potential lenses that do not satisfy
-the specification, and
-\item adds this best lens to the set of possibilities
-under consideration $\PCF{\RXSearch.Add}$
+the specification,
+\item replaces $best$ with $\LC$, if $\LC$ has lower cost, and
+\item adds all of the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)
+  to the priority queue with $\PCF{\PQ.Push}$
 \end{itemize}
@@
 
-\dpw{Note: I added overbars to regexps in Algorithm 1 below ... double check me
-for correctness.}
 \begin{algorithm}
@@
     \State $\RegexAlt \gets \Call{\ToStochastic}{\BRegexAlt}$
-    \State $\RXSearchState \gets \Call{\RXSearch.Create}{\Regex,\RegexAlt}$
-    \While{$\Call{\RXSearch.Continue}{\RXSearchState}$}
-    \State $(\Regex,\RegexAlt,\RXSearchState) \gets \Call{\RXSearch.Propose}{\RXSearchState}$
-    \State $\LCO \gets \Call{\GreedySynth}{\Examples,\Regex,\RegexAlt}$
-    \State $\RXSearchState \gets \Call{\RXSearch.Add}{\RXSearchState,\LCO}$
+    \State $\RXSearchState \gets \Call{\PQ.Create}{\Regex,\RegexAlt}$
+    \State $\Best \gets \None$
+    \While{$\Call{\Continue}{\RXSearchState,\Best}$}
+    \State $(\Regex,\RegexAlt) \gets \Call{\PQ.Pop}{\RXSearchState}$
+    \State $\LC \gets \Call{\GreedySynth}{\Examples,\Regex,\RegexAlt}$
+    \If{$\Call{Cost}{\LC} < \Call{Cost}{\Best}$}
+    \State $\Best \gets \LC$
+    \EndIf
+    \State $\Call{\PQ.Push}{\RXSearchState,\Expand(\Regex,\RegexAlt)}$
     \EndWhile
-    \State \ReturnVal{\Call{RXSearch.Best}{\RXSearchState}}
+    \State \ReturnVal{best}
     \EndFunction
@@
 
-\subsection{\PCF{\RXSearch.Propose}}
+\subsection{Search Through Types}
 The first synthesizer is one that searches through regular expression pairs.
-This synthesizer, \RXSearch, proposes regular expressions, semantically but not
-syntactically equivalent to the originals.  In particular, this synthesizer
-performs the following rewrites:
+This synthesizer, comprising \Expand, \PQ, and \RXSearch, proposes regular
+expressions, semantically but not syntactically equivalent to the originals. In
+particular, this synthesizer performs the following rewrites:
 \begin{center}
@@
 
-\RXSearch proposes rewrites incrementally, where it proposes regular expression
-pairs with the smallest number of rewrites first. When there are no more
-proposed regular expressions derived from $n$ rewrites, \RXSearch proposes
-regular expressions derived from $n+1$ rewrites.
-
-\RXSearch terminates when the sum of the number of regular expressions of the
+Our priority queue proposes stochastic regular expressions generated from fewer
+rewrites first. Only when there are no more proposed regular expressions derived
+from $n$ rewrites will \PCF{\PQ.Pop} propose regular expressions derived from
+$n+1$ rewrites.
+
+The procedure \PCF{\Continue} terminates when the sum of the number of regular expressions of the
 smallest unproposed size with the distance of those regular expressions from the
@@
 concatenation of its subcomponents, and the language of a stochastic atom is the
-iteration of its subcomponent.
+iteration of its subcomponent. For
+$\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}$
+to be a valid stochastic DNF regular expression, the probabilities must sum to
+one ($\sum_{i=0}^n\Probability_i = 1$).
 
@@
       \EmptyString},1)}$
-  \caption{DNF Regular Expression Functions}
+  \caption{Stochastic DNF Regular Expression Functions}
   \label{fig:dnf-regex-functions}
@@
     = & \frac{\Sigma_{j=1}^m(H_j)}{m}\\
-    & \text{where } H_j = \frac{\Sigma_{k\SuchThat j_k = j}
-      \REntropyOf{\Sequence_{i_k}\Given\SequenceLens_k,\SequenceAlt_j}}{\SizeOf{\SetOf{k\SuchThat
-      j_k = j}}} + log_2\SizeOf{\SetOf{k\SuchThat j_k = j}}
+    & \text{where } H_j = (0,\Sigma_{\SetOf{k\SuchThat j_k = j}}
+      \REntropyOf{\Sequence_{i_k}\Given\SequenceLens_k,\SequenceAlt_j})
   \end{array}

### 2018-07-09 papers/SymmetricLenses/main.tex
fix build break

@@
 symmetric DNF lenses (\GreedySynth) and one that greedily finds symmetric
-sequence lenses (\GreedySeqSynth).
-We describe this synthesis algorithm in the context of trying to find a lens
-between \lstinline{"" | name.name*} and \lstinline{"" | name}, with the example
-\lstinline{creater "John Doe" = "John"}.
+sequence lenses (\GreedySeqSynth). We describe this synthesis algorithm in the
+context of trying to find a lens between \lstinline{"" | name.name*} and
+\lstinline{"" | name}.
 
@@
 \GreedySynth finds the best lenses between each sequence pair of the left and
-right DNF regular expressions, given the examples that match them. An internal
-state for the algorithm is initialized from these. This state consists of two
-things: an intermediate partial DNF Lens and a priority queue consisting of
-potential sequence lenses.
+right DNF regular expressions, given the examples that match them. A priority
+queue containing these sequence lenses is initialized from these
+($\PCF{PQ.Create}$). The symmetric lens is built up iteratively from these
+sequence lenses, the state of this partial lens is tracked in the lens builder,
+$lb$.
 
@@
     \Call{\CartesianMap}{\GreedySeqSynth(\Examples),\ListOf{\Sequence_1;\ldots;\Sequence_n},\ListOf{\SequenceAlt_1;\ldots;\SequenceAlt_m}}$
-    \State $\GreedyState \gets \Call{GreedyState.Init}{\SLS}$
-    \While{$\Call{GreedyState.Continue}{\GreedyState}$}
-    \State $(\GreedyState,\SSQLens) \gets
-    \Call{GreedyState.PopBest}{\GreedyState}$
-    \If{$\Call{GreedyState.UsefulAdd}{\GreedyState,\SSQLens}$}
-    \State $\GreedyState \gets \Call{GreedyState.AddSeq}{\GreedyState,\SSQLens}$
+    \State $\GreedyState \gets \Call{PQ.Create}{\SLS}$
+    \State $lb \gets \Call{LensBuilder.Empty}{}$
+    \While{$\Call{PQ.IsNonempty}{\GreedyState}$}
+    \State $\SSQLens \gets
+    \Call{PQ.PopBest}{\GreedyState}$
+    \If{$\Call{LensBuilder.UsefulAdd}{lb,\SSQLens}$}
+    \State $lb \gets \Call{GreedyState.AddSeq}{lb,\SSQLens}$
     \EndIf
     \EndWhile
-    \State \ReturnVal{\Call{GreedyState.ToDNFLens}{\GreedyState}}
+    \State \ReturnVal{\Call{LensBuilder.ToDNFLens}{\GreedyState}}
     \EndFunction
@@
     \Else
-    \State $\ReturnVal{\Call{Map}{\PCF{Iterate},\GreedySynth(\Examples,\DNFRegex,\DNFRegexAlt)}}$
+    \State $\ReturnVal{\IterateLensOf{\GreedySynth(\Examples,\DNFRegex,\DNFRegexAlt)}}$
     \EndIf

### 2018-07-09 papers/SymmetricLenses/main.tex
Added need lemma from bijective lens paper

@@
 \usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
+\usepackage[capitalise]{cleveref}
 \usepackage{amsmath}
@@
 \end{proof}
-
+\begin{mylemma}[Equivalence of \ConcatSequence{} and \Concat{}]
+  If $\LanguageOf{\Regex}=\LanguageOf{\Sequence}$,
+  and $\LanguageOf{\RegexAlt}=\LanguageOf{\SequenceAlt}$,
+  then $\LanguageOf{\RegexConcat{\Regex}{\RegexAlt}}=\LanguageOf{\ConcatSequenceOf{\Sequence}{\SequenceAlt}}$.
+\end{mylemma}
+\begin{proof}
+  Let $\Sequence=\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots
+    \SeqSep\Atom_n\SeqSep\String_n}$, and
+  let $\SequenceAlt=[\StringAlt_0\SeqSep\AtomAlt_1\SeqSep\ldots
+  \SeqSep\AtomAlt_m\SeqSep\StringAlt_m]$. Then
+  \begin{align*}  
+    \LanguageOf{\ConcatSequenceOf{\Sequence}{\SequenceAlt}} & = 
+                                                              \LanguageOf{\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots
+                                                              \SeqSep\Atom_n\SeqSep\String_n\Concat\StringAlt_0\SeqSep{}
+                                                              \AtomAlt_1\SeqSep\ldots\SeqSep\AtomAlt_m\SeqSep\StringAlt_m}} \\
+                                                            & = 
+                                                              \{\String_0\Concat\String_1'\Concat\ldots\Concat\String_n'\Concat\String_n
+                                                              \Concat\StringAlt_0\Concat\StringAlt_1'\Concat\ldots
+                                                              \Concat\StringAlt_m'\Concat\StringAlt_m\SuchThat{} \String_i'\in\LanguageOf{\Atom_i} \BooleanAnd{}
+                                                              \StringAlt_i'\in\LanguageOf{\AtomAlt_i}\}\\
+                                                            & = 
+                                                              \{\String\Concat\StringAlt{} \SuchThat{} \String\in\LanguageOf{\Sequence}
+                                                              \BooleanAnd{} \StringAlt\in\LanguageOf{\SequenceAlt}\}\\
+                                                            & =
+                                                              \{\String\Concat\StringAlt{} \SuchThat{} \String\in\LanguageOf{\Regex}
+                                                              \BooleanAnd{} \StringAlt\in\LanguageOf{\RegexAlt}\}\\
+                                                            & =
+                                                              \LanguageOf{\RegexConcat{\Regex}{\RegexAlt}}
+\end{align*}
+\end{proof}
+
+\begin{mylemma}[Equivalence of \ConcatDNF{} and \Concat{}]
+  \label{lem:cdnfeq}
+  If $\LanguageOf{\Regex}=\LanguageOf{\DNFRegex}$,
+  and $\LanguageOf{\RegexAlt}=\LanguageOf{\DNFRegexAlt}$,
+  then $\LanguageOf{\RegexConcat{\Regex}{\RegexAlt}}=
+  \LanguageOf{\ConcatDNFOf{\DNFRegex}{\DNFRegexAlt}}$.
+\end{mylemma}
+\begin{proof}
+  Let $\DNFRegex=\DNFOf{\Sequence_0\DNFSep\ldots\DNFSep\Sequence_n}$, and
+  let $\DNFRegexAlt=\DNFOf{\SequenceAlt_0\DNFSep\ldots\DNFSep\SequenceAlt_m}$
+  \begin{align*}
+    \LanguageOf{\ConcatDNFOf{\DNFRegex}{\DNFRegexAlt}} & = 
+                                                         \LanguageOf{\DNFOf{\ConcatSequenceOf{\Sequence_i}{\SequenceAlt_j}
+                                                         \text{ for $i\in\RangeIncInc{1}{n}$, $j\in\RangeIncInc{1}{m}$}}} \\
+                                                       & = 
+                                                         \{\String\SuchThat \String\in\ConcatSequenceOf{\Sequence_i}{\SequenceAlt_j} \text{ where $i\in\RangeIncInc{1}{n}$, $j\in\RangeIncInc{1}{m}$}\}\\
+                                                       & = 
+                                                         \{\String\Concat\StringAlt{} \SuchThat{} \String\in\LanguageOf{\Sequence_i}
+                                                         \BooleanAnd{} \StringAlt\in\LanguageOf{\SequenceAlt_j}\} \text{ where $i\in\RangeIncInc{1}{n}$, $j\in\RangeIncInc{1}{m}$}\}\\
+                                                       & =
+                                                         \{\String\Concat\StringAlt{} \SuchThat{} \String\in\LanguageOf{\DNFRegex}
+                                                         \BooleanAnd{} \StringAlt\in\LanguageOf{\DNFRegexAlt}\}\\
+                                                       & =
+                                                         \{\String\Concat\StringAlt{} \SuchThat{} \String\in\LanguageOf{\Regex}
+                                                         \BooleanAnd{} \StringAlt\in\LanguageOf{\RegexAlt}\}\\
+                                                       & =
+                                                         \LanguageOf{\RegexConcat{\Regex}{\RegexAlt}}
+\end{align*}
+\end{proof}
+
+\begin{mylemma}[Equivalence of $\Atom$ and $\AtomToDNFOf{\Atom}$]
+  \label{lem:atomtodnfeq}
+  $\LanguageOf{\Atom} = \LanguageOf{\AtomToDNFOf{\Atom}}$
+\end{mylemma}
+\begin{proof}
+\begin{align*}
+  \LanguageOf{\AtomToDNFOf{\Atom}} &=
+  \LanguageOf{\DNFOf{\SequenceOf{\EmptyString \SeqSep \Atom \SeqSep \EmptyString}}}\\
+  & =
+  \SetOf{\String \SuchThat \String \in
+    \LanguageOf{\SequenceOf{\EmptyString \SeqSep \Atom \SeqSep \EmptyString}}} \\
+  & =
+  \SetOf{\EmptyString\Concat\String\Concat\EmptyString \SuchThat \String \in
+    \LanguageOf{\Atom}} \\
+   & = \SetOf{\String \SuchThat \String \in
+    \LanguageOf{\Atom}} \\
+    &= \LanguageOf{\Atom}
+    \end{align*}
+\end{proof}
+
+\begin{mylemma}[Equivalence of \OrDNF{} and \Or{}]
+  \label{lem:odnfeq}
+  If $\LanguageOf{\Regex}=\LanguageOf{\DNFRegex}$,
+  and $\LanguageOf{\RegexAlt}=\LanguageOf{\DNFRegexAlt}$,
+  then $\LanguageOf{\RegexOr{\Regex}{\RegexAlt}}=
+  \LanguageOf{\OrDNFOf{\DNFRegex}{\DNFRegexAlt}}$.
+\end{mylemma}
+\begin{proof}
+  Let $\DNFRegex=\DNFOf{\Sequence_0\DNFSep\ldots\DNFSep\Sequence_n}$, and
+  let $\DNFRegexAlt=\DNFOf{\SequenceAlt_0\DNFSep\ldots\DNFSep\SequenceAlt_m}$
+
+  \begin{align*}
+    \mathcal{L}(DS \oplus DT)& = 
+                                                     \LanguageOf{\DNFOf{\Sequence_0\DNFSep\ldots\DNFSep\Sequence_n\DNFSep
+                                                     \SequenceAlt_1\DNFSep\ldots\DNFSep\SequenceAlt_m}}\\
+                                                   & = 
+                                                     \{\String\SuchThat{} \String\in\Sequence_i\vee\String\in\SequenceAlt_j \text{ where $i\in\RangeIncInc{1}{n}$, $j\in\RangeIncInc{1}{m}$}\}\\
+                                                   & = 
+                                                     \{\String{} \SuchThat{} \String\in\LanguageOf{\DNFRegex}
+                                                     \BooleanOr{} \String\in\LanguageOf{\DNFRegexAlt}\}\\
+                                                   & =
+                                                     \{\String \SuchThat{} \String\in\LanguageOf{\Regex}
+                                                     \BooleanOr{} \String\in\LanguageOf{\RegexAlt}\}\\
+                                                   & =
+                                                     \LanguageOf{\RegexOr{\Regex}{\RegexAlt}}
+  \end{align*}
+\end{proof}
+
+\begin{theorem}\label{ConversionPreservesSemantics}
+  For all regular expressions \Regex{},
+  $\LanguageOf{\ToDNFRegexOf{\Regex}}=\LanguageOf{\Regex{}}$.
+\end{theorem}
+\begin{proof}
+  By structural induction.
+
+  Let $\Regex=\String$.
+  $\LanguageOf{\ToDNFRegex(\String)}=\LanguageOf{\DNFOf{\SequenceOf{\String}}}=
+  \{\String\}=\LanguageOf{\String}$
+
+  Let $\Regex=\emptyset$.
+  $\LanguageOf{\ToDNFRegex(\emptyset)}=\LanguageOf{\DNFOf{}} =
+  \{\} = \LanguageOf{\emptyset}$.
+
+  Let $\Regex=\StarOf{\Regex'}$.
+  By induction assumption, $\LanguageOf{\ToDNFRegex(\Regex')}=
+  \LanguageOf{\Regex'}$.
+  \begin{align*}
+    \LanguageOf{\ToDNFRegex(\StarOf{\DNFRegex'})} & =
+                                                    \LanguageOf{\DNFOf{\SequenceOf{\StarOf{\ToDNFRegex(\Regex')}}}}\\
+                                                  & =
+                                                    \{\String\SuchThat\String\in
+                                                    \LanguageOf{\SequenceOf{\StarOf{\ToDNFRegex(\Regex')}}}\}\\
+                                                  & = 
+                                                    \{\String\SuchThat{} \String\in\LanguageOf{\StarOf{\ToDNFRegex(\Regex')}}\}\\
+                                                  & =
+                                                    \{\String_1\Concat\ldots\Concat\String_n\SuchThat{}
+                                                    n\in\Nats \BooleanAnd\String_i\in\LanguageOf{\ToDNFRegex(\Regex')}\}\\
+                                                  & =
+                                                    \{\String_1\Concat\ldots\Concat\String_n\SuchThat{}
+                                                    n\in\Nats\BooleanAnd\String_i\in\LanguageOf{\Regex'}\}\\
+                                                  & = \LanguageOf{\StarOf{\Regex'}}
+  \end{align*}
+
+  Let $\Regex=\RegexConcat{\Regex_1}{\Regex_2}$.
+  By induction assumption,
+  $\LanguageOf{\ToDNFRegex(\Regex_1)}=\LanguageOf{\Regex_1}$, and
+  $\LanguageOf{\ToDNFRegex(\Regex_2)}=\LanguageOf{\Regex_2}$.
+  $\ToDNFRegex(\RegexConcat{\Regex_1}{\Regex_2})=
+  \ConcatDNFOf{\ToDNFRegex(\Regex_1)}{\ToDNFRegex(\Regex_2)}$.
+  By Lemma~\ref{lem:cdnfeq},
+  $\RegexConcat{\Regex_1}{\Regex_2}=
+  \ConcatDNFOf{\ToDNFRegex(\Regex_1)}{\ToDNFRegex(\Regex_2)}$.
+
+  Let $\Regex=\RegexOr{\Regex_1}{\Regex_2}$.
+  By induction assumption,
+  $\LanguageOf{\ToDNFRegex(\Regex_1)}=\LanguageOf{\Regex_1}$, and
+  $\LanguageOf{\ToDNFRegex(\Regex_2)}=\LanguageOf{\Regex_2}$.
+
+  $\Downarrow(S \; | \; T) = (\Downarrow S) \oplus (\Downarrow T)$. By Lemma~\ref{lem:odnfeq},
+  $\mathcal{L}(S \; | \; T) = \mathcal{L}((\Downarrow S) \oplus (\Downarrow T))$.
+
+\end{proof}
 \begin{theorem*}
@@
 \begin{proof}
-By \sam{COPY PASTE PROOF FROM BIJECTIVE LENS PAPER} $\LanguageOf{\Regex} =
+By \cref{ConversionPreservesSemantics}, $\LanguageOf{\Regex} =
   \LanguageOf{\ToDNFRegexOf{\Regex}}$, so we will prove the result $\ProbabilityOf{\Regex}{\String} = \ProbabilityOf{\ToDNFRegex

### 2018-07-09 papers/SymmetricLenses/main.tex
add better algorithms descriptions and algs

--- 2018-07-09 papers/SymmetricLenses/main.tex
@@
 right DNF regular expressions, given the examples that match them. A priority
-queue containing these sequence lenses is initialized from these
-($\PCF{PQ.Create}$). The symmetric lens is built up iteratively from these
+queue containing these sequence lenses is then initialized with
+$\PCF{PQ.Create}$. The symmetric lens is built up iteratively from these
 sequence lenses, the state of this partial lens is tracked in the lens builder,
@@
 \GreedySynth then loops until there are no more sequence lenses in the priority
-queue. Within this loop, sequence lenses are popped. If a popped sequence lens
-maps between two sequences which aren't yet included in a sequence lens, then
-this sequence lens is added to the partial DNF lens. Furthermore, we update the
-priorities of our sequence lenses as the algorithm proceeds: if two sequence
-lenses have the same source, the second one to be popped gets a higher cost than
-it originally had, to address the information that needs to be stored for
-including that source of non-bijectivity.
+queue. Within this loop, sequence lenses are popped. If this sequence lens is
+``useful'', then that sequence lens is included in the final DNF lens. If the
+source or target of a popped sequence lens is not the source or target of
+another sequence lens, then this sequence lens is considered useful. If the
+examples require two sequences have a lens between them, then any sequence lens
+between them is considered useful. We update the priorities of our sequence
+lenses as the algorithm proceeds: if two sequence lenses have the same source,
+the second one to be popped gets a higher cost than it originally had, to
+address the information that needs to be stored for including that source of
+non-bijectivity.
 
 In our example, \GreedySynth may first pop the sequence pair \lstinline{""} and
-\lstinline{""}, as there is a bijective sequence lens between them. After this,
-the sequence lens between \lstinline{name.name*} and \lstinline{name} would be
-popped. While examples prevent lenses between the other sequence pairs, if there
-were other lenses, they would be popped, but not added to the ongoing state, as
-they would not be a useful addition to the state.
-
-Finally, after all sequences have been
-popped, the partial DNF lens is converted into a symmetric DNF lens. This is
-only possible if all sequences are involved in some sequence lens: if they are
-not, \PCF{GreedyState.ToDNFLens} instead returns \None.
+\lstinline{""}, as there is a bijective sequence lens between them. As neither
+\lstinline{""} is involved in a sequence lens, this lens is considered useful.
+After this, the sequence lens between \lstinline{name.name*} and
+\lstinline{name} would be popped: while that lens is not bijective: it is still
+better than the alternatives. As all sequences are now involved in sequence lenses,
+and there are no examples to make other lenses useful, no further sequence
+lenses would be added to the lens builder.
+
+Finally, after all sequences have been popped, the partial DNF lens $lb$ is
+converted into a symmetric DNF lens. This is only possible if all sequences are
+involved in some sequence lens: if they are not, \PCF{GreedyState.ToDNFLens}
+instead returns \None.
 
@@
     \State $\SSQLens \gets
-    \Call{PQ.PopBest}{\GreedyState}$
-    \If{$\Call{LensBuilder.UsefulAdd}{lb,\SSQLens}$}
+    \Call{PQ.Pop}{\GreedyState}$
+    \If{$\Call{LensBuilder.UsefulAdd}{lb,\SSQLens,\Examples}$}
     \State $lb \gets \Call{GreedyState.AddSeq}{lb,\SSQLens}$
@@
 \GreedySeqSynth finds the best lenses between each atom pair of the left and
-right sequences. An internal state for the algorithm is initialized from these.
-This state consists of two things: an intermediate sequence lens and a priority
-queue consisting of potential sequence lenses. On initialization,
-\GreedySeqSynth throws all the atom lenses into the priority queue, and the
-intermediate sequence lens is one that projects all the atoms.
-
-For example, if trying to find a sequence lens between \lstinline{name.name*}
-and \lstinline{name}, lenses are found between \lstinline{name} and
-\lstinline{name}, and \lstinline{name} and \lstinline{name*}, and added to the
-queue.
+right sequences. A priority queue containing these lenses is then initialized
+with \PCF{PQ.Create}. The symmetric sequence lens is built up iteratively from
+these atom lenses, the state of the partial lens is tracked in the sequence lens
+builder, $slb$.
 
@@
 queue. Within this loop, atom lenses are popped. The atom lens is then added if
-it is either necessary, or useful. Within a sequence lens, each atom can either
-be projected or be synchronized with exactly one atom in the other sequence, and
-examples can show that certain atoms \emph{need} to be mapped. In such
-situations, that atom lens is necessary. A useful atom lens is where adding that
-lens reduces the overall cost of the partial lens. If both atoms are not
-involved in an atom lens, the lens will always be useful, but sometimes adding
-that lens requires removing another lens. If the cost added by removing that
-other lens is more than is removed by adding the proposed lens, then this atom
-lens is not included in the sequence lens. For example, the bijective lens
-between \lstinline{name} and \lstinline{name} would first be found, and added to
-the state.  Next, the disconnect lens between \lstinline{name} and
-\lstinline{name*} would be popped.  Because this lens is not useful, it is not
-added to the state.
-
-Finally, after all atoms have been
-popped, the sequence lens is validated and returned. The sequence lens is
-validated if all atoms that need to be mapped, are: if they are not,
-\PCF{GreedySeqState.ToDNFLens} instead returns \None.
+it is useful. An atom lens is useful if adding it will lower the cost of the
+generated sequence lens, or if examples show that one of the atoms must be
+connected to another lens. The priorities of these lenses change over time, as
+each atom can only be involved in one lens. Adding a lens may necessitate
+removing another, so not all popped atom lenses will lower the cost of the
+generated sequence lens.
+
+Finally, after all atoms have been popped, the sequence lens is validated and
+returned. The algorithm confirms that all atoms that need to be connected with
+an atom lens, are. If they are not, \PCF{SLensBuilder.ToDNFLens} instead returns
+\None.
 
@@
     \Call{\CartesianMap}{\AtomSynth(\Examples),\ListOf{\Atom_1;\ldots;\Atom_n},\ListOf{\AtomAlt_1;\ldots;\AtomAlt_m}}$
-    \State $\GreedyState \gets \Call{GreedySeqState.Init}{\ALS}$
-    \While{$\Call{GreedySeqState.Continue}{\GreedyState}$}
-    \State $(\GreedyState,\SAtomLens) \gets
-    \Call{GreedySeqState.PopBest}{\GreedyState}$
-    \If{$\Call{GreedySeqState.UsefulAdd}{\GreedyState,\SAtomLens}
-      \BooleanOr \Call{NecessaryAdd}{\Examples,\SAtomLens}$}
-    \State $\GreedyState \gets \Call{GreedySeqState.AddAtom}{\GreedyState,\SAtomLens}$
+    \State $\GreedyState \gets \Call{PQ.Init}{\ALS}$
+    \State $slb \gets \Call{SLensBuilder.Empty}{}$
+    \While{$\Call{PQ.IsNonempty}{\GreedyState}$}
+    \State $\SAtomLens \gets
+    \Call{PQ.Pop}{\GreedyState}$
+    \If{$\Call{SLensBuilder.UsefulAdd}{slb,\SAtomLens,\Examples}$}
+    \State $\GreedyState \gets \Call{SLensBuilder.AddAtom}{\GreedyState,\SAtomLens}$
     \EndIf
     \EndWhile
-    \State \ReturnVal{\Call{GreedySeqState.ToDNFLens}{\GreedyState}}
+    \State \ReturnVal{\Call{SLensBuilder.ToDNFLens}{\GreedyState}}
     \EndFunction

### 2018-07-09 papers/SymmetricLenses/main.tex
minor fix

@@
 needed to recover a string in one data source from a synchronized string in the
-other data source. To do this, we must first develop a probabilistic langauge
+other data source. To do this, we must first develop a probabilistic language
 model for our language. We do this with \emph{stochastic regular expressions},
@@
 
-An atom lens \AtomLens{} is merely the iteration of a symmetric DNF Lens.
-
 While DNF lenses are useful because they permit an efficient greedy synthesis
@@
 \paragraph*{Information-Theoretic Cost Metric}
-Now that we've developed symmetric DNF lenses, it's important to develop a means
-to guage how bijective the lens is.
-
 Instead of searching for \emph{any} term that satisfies the specification, we

### 2018-07-09 papers/SymmetricLenses/main.tex
edits to alg

@@
 level of abstraction. The input regular expressions are first converted into
-stochastic regular expressions with \ToStochastic. Once that conversion has
-taken place, the $\RXSearch$ algorithm executes. This algorithm itself is
-defined via five functions: (1) $\PCF{\PQ.Create}$, (2) $\PCF{\PQ.Pop}$, (3)
-$\PCF{Expand}$, (4) $\PCF{\PQ.Push}$, and (5) $\PCF{\Continue}$. First,
-$\PCF{\PQ.Create}$ creates a priority queue ($pq$) containing a single element:
-the initial pair of stochastic regular expressions. Next, $\PCF{SynthSymLens}$
+stochastic regular expressions with \ToStochastic.  
+%Once that conversion has
+%taken place, the $\RXSearch$ algorithm executes. This algorithm itself is
+%defined via five functions: (1) $\PCF{\PQ.Create}$, (2) $\PCF{\PQ.Pop}$, (3)
+%$\PCF{Expand}$, (4) $\PCF{\PQ.Push}$, and (5) $\PCF{\Continue}$. First,
+This pair of SREs are used to initialize a priority queue ($pq$).
+Next, $\PCF{SynthSymLens}$
 enters a loop that searches for likely lenses. The loop terminates when the
@@
 \begin{itemize}
-\item generates the next class ($S$, $T$) of lenses to
-search ($\PCF{\PQ.Pop}$),
+\item pops the next class ($S$, $T$) of lenses to
+search of the priority queue ($\PCF{\PQ.Pop}$),
 \item executes $\GreedySynth$ to find a best lens in that class if
@@
 the specification,
-\item replaces $best$ with $\LC$, if $\LC$ has lower cost, and
+\item replaces $best$ with $\LC$, if $\LC$ has lower cost according
+to our information theoretic metric, and
 \item adds all of the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)

### 2018-07-09 papers/SymmetricLenses/main.tex
Added proof of Theorem 5

@@
 \end{proof}
-\begin{theorem*}
-  \label{thm:correct_entropy}
+\begin{theorem}
+  \label{CorrectEntropy}
   If $\Regex$ is unambiguous and contains no empty subcomponents,
   $\EntropyOf{\Regex}$ is the entropy of $\mathcal{L}(\Regex)$.
-\end{theorem*}
+\end{theorem}
 \begin{proof}
@@
 
+\end{proof}
+\begin{theorem}
+If $DS$ is unambiguous, then $\EntropyOf{\DNFRegex}$ is the entropy of $P_{\DNFRegex}$.
+\end{theorem}
+\begin{proof}
+By mutual induction over $DS$, $SQ$ and $A$ (the computations of the entropies follow the same pattern as the respective computations in \cref{CorrectEntropy}).
+\begin{enumerate}
+\item
+($A = DS^{*p}$)
+
+By the induction hypothesis $H(DS) = \mathbb{H}(DS)$, from which it follows that
+$$H(DS^{*p}) = \frac{p}{1-p}(H(DS) - \log_2 p) - \log_2 (1-p) = \mathbb{H}(DS^{*p})$$
+\item
+($SQ = [s_0 \cdot A_1 \cdot \ldots \cdot A_n \cdot s_n]$)
+
+By the induction hypothesis, hence by unambiguity $H(A_i) = \mathbb{H}(A_i)$, thus
+$$H([s_0 \cdot A_1 \cdot \ldots \cdot A_n \cdot s_n]) = \sum_{i=1}^n H(A_i) = \sum_{i=1}^n \mathbb{H}(A_i) = \mathbb{H}([s_0 \cdot A_1 \cdot \ldots \cdot A_n \cdot s_n])$$
+\item
+$(DS = \langle (SQ_1, p_1) \; | \; \ldots \; | \; (SQ_n, p_n)\rangle)$
+
+By the induction hypothesis, $H(SQ_i, p_i) = \mathbb{H}(SQ_i, p_i)$. Using the formula
+$$H(P) = H(m(P_1), \ldots m(P_n)) + \sum_{i=1}^n m(P_i)H(P_i)$$
+if $P$ is a measure space satisfying $P = P_1 + \ldots + P_n$, then
+\begin{align*}
+H(\langle (SQ_1, p_1) \; | \; \ldots \; | \; (SQ_n, p_n)\rangle)
+&= H(p_1, \ldots, p_n) + \sum_{i=1}^n p_i H(SQ_i, p_i)\\
+&= \sum_{i=1}^n p_i \mathbb{H}(SQ_i, p_i) + p_i * \log_2 p_i\\
+&= \mathbb{H}(\langle (SQ_1, p_1) \; | \; \ldots \; | \; (SQ_n, p_n)\rangle)
+\end{align*}
+\end{enumerate}
 \end{proof}

### 2018-07-09 papers/SymmetricLenses/main.tex
some more changes

--- 2018-07-09 papers/SymmetricLenses/main.tex
@@
     expressions \Regex and \RegexAlt, and a set of examples \Examples, are
-    provided as input. \RXSearch searches through stochastic regular expression
+    provided as input. \Expand searches through stochastic regular expression
     pairs, equivalent to the originals, and proposes them to \GreedySynth.
     \GreedySynth finds a lens, typed between the generated equivalent pairs.
-    When \RXSearch determines it has an optimal lens, it returns that lens.}
+    When the algorithm determines it has an optimal lens, it returns that lens.}
   \label{fig:high-level-algorithm}
@@
 synthesizing likely lenses, based on the previous algorithm for synthesizing
-bijective lenses. This algorithm is constructed from \RXSearch and \GreedySynth,
+bijective lenses. This algorithm is constructed from \Expand and \GreedySynth,
 a pair of collaborating search procedures, shown in
 Figure~\ref{fig:high-level-algorithm}. Given source and target SREs $S$ and $T$,
-the \RXSearch searches for two \emph{related} SREs $S'$ and $T'$ that are
+the \Expand searches for two \emph{related} SREs $S'$ and $T'$ that are
 \emph{compatible} (a heuristic that estimates the likelihood our second
@@
 search terminates. Otherwise, this lens is stored away (if it is the best found
-so far) and the search continues with \RXSearch again.
-
-We implemented the \GreedySynth and \RXSearch algorithms and evaluated their
+so far) and the search continues with \Expand again.
+
+We implemented the \GreedySynth and \Expand algorithms and evaluated their
 effectiveness of synthesizing synchronization functions on a range of
@@
 the same two SREs $\Regex$ and $\RegexAlt$. For any such lens $\Lens$, we can
-calculate the expected number of bits required to recover a string $s$ in
-$\LanguageOf{\Regex}$ from a synchronized string $t$ in
-$\LanguageOf{\RegexAlt}$. This expected number of bits is just the conditional
-entropy of $\Regex$ given $\RegexAlt$ and $\Lens$. To ensure symmetry, the
-likelihood we assign to $\Lens$ is the sum of the conditional entropy of
-$\Regex$ given $\RegexAlt$ and $\Lens$, \bcp{Am I supposed to know what this
-  means?} and the conditional entropy of $\RegexAlt$ given $\Regex$ and
-$\Lens$.\bcp{Why is this the right (or a sensible) thing to do, besides the fact
-  that it ensures symmetry?}
+calculate the expected number of bits required to recover a string $t$ in
+$\LanguageOf{\RegexAlt}$ from a synchronized string $s$ in
+$\LanguageOf{\Regex}$. This expectation is the \emph{conditional entropy of
+  $\Regex$ given $\RegexAlt$ and $\Lens$}, formally $\sum_{s \in
+  \Regex}\ProbabilityOf{\Regex}{s}*\EntropyOf{\SetOf{t \SuchThat
+    \Lens.\PutROf{s}{t} = t}}$. The likelihood we assign to $\Lens$ is the sum
+of the conditional entropy of $\Regex$ given $\RegexAlt$ and $\Lens$ and the
+conditional entropy of $\RegexAlt$ given $\Regex$ and $\Lens$. This metric
+assigns higher cost to lenses where knowing the string present in \Regex
+provides little information about what string is present in \RegexAlt. With this
+metric, bijections have zero cost -- given a string $s \in \Regex$, we know
+exactly what string is present in $\RegexAlt$, as exactly one string is
+synchronized to it.
 
@@
 two communicating synthesizers following earlier
-work~\cite{quotient-lens-synthesis}. The first synthesizer, \RXSearch, navigates
-the space of semantically equivalent regular expressions by iteratively applying
-rewrite rules that preserve both the semantics and the probability distributions
-of those stochastic regular expressions. \RXSearch ranks pairs of stochastic
+work~\cite{quotient-lens-synthesis}. The first synthesizer, \Expand, navigates
+the space of semantically equivalent regular expressions by applying rewrite
+rules that preserve both the semantics and the probability distributions of
+those stochastic regular expressions. Our algorithm ranks pairs of stochastic
 regular expressions by the number of rewrite rule applications required to
@@
 likelihood by performing a type-directed search. In more detail, \GreedySynth
-first normalizes the stochastic regular expressions provided by \RXSearch into
+first normalizes the stochastic regular expressions provided by \Expand into
 \emph{stochastic DNF regular expressions}. These stochastic DNF regular
@@
 
-As an example of the partnership between \RXSearch and \GreedySynth,
+As an example of the partnership between \Expand and \GreedySynth,
 consider the lens shown in Figure~\ref{fig:example_lens}. 
-Inferring this lens requires \RXSearch to use a star semiring axiom to
+Inferring this lens requires \Expand to use a star semiring axiom to
 rewrite the regular expression \lstinline{("\n" . emp_salary)$^*$} within
@@
 
-To resolve this tension, \RXSearch uses heuristics to judge whether to
+To resolve this tension, our algorithm uses heuristics to judge whether to
 return the current best satisfying lens to the user or to pass the
@@
 lens has relatively high likelihood, indicating the lens is very promising.
-The heuristics also favor stopping if \RXSearch 
+The heuristics also favor stopping if \Expand
 has delievered to \GreedySynth all pairs of stochastic regular
@@
 \subsection{Stochastic Regular Expression Equivalences}
-$\RXSearch$ needs to be able to find equivalent stochastic regular expressions.
-However, existing work does not reason about stochastic regular expression
-equivalence. We extend the \emph{star-semiring}~\cite{?} equivalences to
-stochastic regular expressions.\bcp{Why these?}
+$\Expand$ needs to be able to find equivalent stochastic regular expressions.
+However, existing work does not provide a means to reason about stochastic
+regular expression equivalence. We extend the \emph{star-semiring}~\cite{?}
+equivalences to stochastic regular expressions.
 
@@
 
+We prove these star-semiring equivalences sound, as those are the equivalences
+we either traverse (with \Expand), or normalize across (as part of
+\GreedySynth).
+
 \subsection{Stochastic Regular Expression Entropy}
@@
 
-Without unambiguity constraints it is not easy to calculate the entropy. For
-example, consider the SRE $\PRegexOr{\lstinline{"a"}}{\lstinline{"a"}}{.5}$.
-Using the above computation, this SRE has an entropy of $1$, but the entropy
-should be $0$.\bcp{This doesn't show that it's not easy---it just shows that
-one particular way of computing it is wrong.} Similar issues occur\bcp{Can we at least hint what they are?} when trying to find the entropy when
-$\emptyset$ is a subterm, so we do not define the entropy on $\emptyset$.
-
-In our work, we do not find this to be a problem\bcp{Why might one expect
-  it to be a problem, rather than just a technicality?}: we require
-unambiguous regular 
-expressions as input to our synthesis procedure, and we can easily preprocess
-emptysets out of the stochastic regular expressions that are themselves nonempty
-by traversing the correct star-semiring equivalences
-(\EG, \PRegexOr{\emptyset}{\lstinline{"s"}}{.5} \SSREquiv \lstinline{"s"}).
+Without unambiguity constraints it is not easy to calculate the entropy. This
+method is certainly incorrect without unambiguity constraints: consider the SRE
+$\PRegexOr{\lstinline{"a"}}{\lstinline{"a"}}{.5}$. Using the above computation,s
+this SRE has an entropy of $1$, but the entropy should be $0$. These
+difficulties become particularlly complex when the languages are ambiguous in
+subtle ways, the algorithm must be correct when only some of the two languages
+overlap. Similar issues occur when trying to find the entropy when $\emptyset$
+is a subterm (what should the entropy of $\PRegexStar{\emptyset}{.5}$ be), so we
+do not define the entropy on $\emptyset$.
+
+In our work, these conditions are merely a technicality, we require unambiguous
+regular expressions as input to our synthesis procedure, and we can easily
+preprocess emptysets out of the stochastic regular expressions that are
+themselves nonempty by traversing the correct star-semiring equivalences (\EG,
+\PRegexOr{\emptyset}{\lstinline{"s"}}{.5} \SSREquiv \lstinline{"s"}).
 
@@
 We now provide the typing derivations and semantics for these lenses. For
-simplicity\bcp{Is there any known issue with relaxing this restriction?
-  I've often found ti convenient to use id lenses with ambiguous REs...}, we define the typing derivations and semantics only over unambiguous
-regular expressions, even though some lenses can be defined over ambiguous
-regular expressions (for example, the identity lens doesn't need any ambiguity
-constraints).
+simplicity of presentation,\bcp{Is there any known issue with relaxing this
+  restriction? I've often found ti convenient to use id lenses with ambiguous
+  REs...}\afm{no known issues, I just didn't want to go into specifics about
+  ambiguity in an already-full paper}, we define the typing derivations and
+semantics only over unambiguous regular expressions, even though some lenses can
+be defined over ambiguous regular expressions (for example, the identity lens
+doesn't need any ambiguity constraints).
 
@@
 recover a string in one data format from the other. We use the function
-$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to bound\bcp{``bound'' or
-  ``calculate''?  I.e., is what we calculate correct, or is it only known to
-be an upper bound?  If (as it seems) the answer is the latter, then why?}
-the expected amount of 
-information required to recover a string in $\RegexAlt$ from a string in
-$\Regex$, synchronized by $\Lens$. We use the function $\LEntropyOf{\Regex
-  \Given \Lens, \RegexAlt}$ to bound the expected amount of information required
-to recover a string in $\Regex$ from a string in $\RegexAlt$, synchronized by
-$\Lens$.
+$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to calculate bounds of the
+expected amount of information required to recover a string in $\RegexAlt$ from
+a string in $\Regex$, synchronized by $\Lens$. We use the function
+$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate bounds of the
+expected amount of information required to recover a string in $\Regex$ from a
+string in $\RegexAlt$, synchronized by $\Lens$.
 \begin{center}
@@
 functions bound the expected number of bits to recover one data format from a
-synchronized string in the other format. Notice, we would be able to calculate
-the exact entropy, were it not for \MergeL. If $\MergeLOf{\Lens_1}{\Lens_2}
-\OfType \Regex \Leftrightarrow
+synchronized string in the other format. Notice, we would be able to exactly
+calculate the conditional entropy, were it not for \MergeL. If
+$\MergeLOf{\Lens_1}{\Lens_2} \OfType \Regex \Leftrightarrow
 \PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}$, given a string in $s$, we
@@
 However, this information content is dependent on the how likely the
-synchronized string is to be in $\Regex$ or $\RegexAlt$. In the worst case, the
-expected information content to encode this choice is
-$-\Log_2(\ProbabilityAlt)*\ProbabilityAlt -\Log_2(1 -
-\ProbabilityAlt)*(1-\ProbabilityAlt)$, but the expected information content
-could be arbitrarily small.
+synchronized string is to be in $\Regex$ or $\RegexAlt$. Furthermore, we
+typically calculate the conditional entropy exactly, as merges are relatively
+uncommon on practice; only \afm{2} of the lenses sythesized in our benchmarks
+include merges.
 
@@
 stochastic regular expressions with \ToStochastic. Once that conversion has
-taken place, the $\RXSearch$ algorithm executes. This algorithm itself is
-defined via five functions: (1) $\PCF{\PQ.Create}$, (2) $\PCF{\PQ.Pop}$, (3)
-$\PCF{Expand}$, (4) $\PCF{\PQ.Push}$, and (5) $\PCF{\Continue}$. First,
-$\PCF{\PQ.Create}$ creates a priority queue ($pq$) containing a single element:
-the initial pair of stochastic regular expressions. Next, $\PCF{SynthSymLens}$
-enters a loop that searches for likely lenses. The loop terminates when the
-algorithm believes it is unlikely to find a better lens than the best one it has
-found so far (a termination condition defined by $\PCF{\Continue}$). Within each
-iteration of the loop, it:
+taken place, the we begin searching for the most likely lens. This search
+itself is defined via five functions: (1) $\PCF{\PQ.Create}$, (2)
+$\PCF{\PQ.Pop}$, (3) $\PCF{Expand}$, (4) $\PCF{\PQ.Push}$, and (5)
+$\PCF{\Continue}$. First, $\PCF{\PQ.Create}$ creates a priority queue ($pq$)
+containing a single element: the initial pair of stochastic regular expressions.
+Next, $\PCF{SynthSymLens}$ enters a loop that searches for likely lenses. The
+loop terminates when the algorithm believes it is unlikely to find a better lens
+than the best one it has found so far (a termination condition defined by
+$\PCF{\Continue}$). Within each iteration of the loop, it:
 \begin{itemize}
@@
 When the loop terminates, the search returns the globally best lens found
-($\PCF{\RXSearch.Best}$).  Each element of this algorithm will be explained in
+($best$).  Each element of this algorithm will be explained in
 further depth in the following subsections.

### 2018-07-09 papers/SymmetricLenses/main.tex

@@
 level of abstraction. The input regular expressions are first converted into
-stochastic regular expressions with \ToStochastic. Once that conversion has
-taken place, the we begin searching for the most likely lens. This search
-itself is defined via five functions: (1) $\PCF{\PQ.Create}$, (2)
-$\PCF{\PQ.Pop}$, (3) $\PCF{Expand}$, (4) $\PCF{\PQ.Push}$, and (5)
-$\PCF{\Continue}$. First, $\PCF{\PQ.Create}$ creates a priority queue ($pq$)
-containing a single element: the initial pair of stochastic regular expressions.
-Next, $\PCF{SynthSymLens}$ enters a loop that searches for likely lenses. The
-loop terminates when the algorithm believes it is unlikely to find a better lens
-than the best one it has found so far (a termination condition defined by
-$\PCF{\Continue}$). Within each iteration of the loop, it:
+stochastic regular expressions with \ToStochastic.  
+%Once that conversion has
+%taken place, the $\RXSearch$ algorithm executes. This algorithm itself is
+%defined via five functions: (1) $\PCF{\PQ.Create}$, (2) $\PCF{\PQ.Pop}$, (3)
+%$\PCF{Expand}$, (4) $\PCF{\PQ.Push}$, and (5) $\PCF{\Continue}$. First,
+This pair of SREs are used to initialize a priority queue ($pq$).
+Next, $\PCF{SynthSymLens}$
+enters a loop that searches for likely lenses. The loop terminates when the
+algorithm believes it is unlikely to find a better lens than the best one it has
+found so far (a termination condition defined by $\PCF{\Continue}$). Within each
+iteration of the loop, it:
 \begin{itemize}
-\item generates the next class ($S$, $T$) of lenses to
-search ($\PCF{\PQ.Pop}$),
+\item pops the next class ($S$, $T$) of lenses to
+search of the priority queue ($\PCF{\PQ.Pop}$),
 \item executes $\GreedySynth$ to find a best lens in that class if

### 2018-07-09 papers/SymmetricLenses/main.tex
add examples for explanation

@@
 \paragraph*{Syntax}
-Like stochastic DNF regular expressions, symmetric DNF Lenses are hierarchically
-structured. Symmetric DNF lenses, which provide lens functions between DNF
-regular expressions, consist of a list of symmetric sequence lenses, as well as
-some additional information. Symmetric sequence lenses, which provide lens
-functions betwen sequences, consist of a list of symmetric atom lenses, as well
-as some additional information. Symmetric atom lenses, which provide lens
-functions between atoms, are merely iterations of symmetric DNF lenses. In the
-following syntax, the nonterminals $i$, $j$, $p$, $q$, $r$, $c$, and $d$ all
-represent natural numbers. The nonterminals $\String$ and $\StringAlt$ represent
-strings.
+Roughly speaking, symmetric DNF lenses are an n-ary union of symmetric sequence
+lenses, which are, in turn, an n-ary sequence of atomic lenses. Atomic lenses
+are iterations of symmetric DNF lenses. In the following syntax, the
+nonterminals $i$, $j$, $p$, $q$, $r$, $c$, and $d$ all represent natural
+numbers. The nonterminals $\String$ and $\StringAlt$ represent strings.
 \begin{center}
@@
                           ,\ListOf{(\String_1,\Atom_1);\ldots;(\String_q,\Atom_q)}
-                          ,\ListOf{\StringAlt_1;\ldots;\StringAlt_r})$ \\
+                          ,\ListOf{(\StringAlt_1,\AtomAlt_1);\ldots;(\StringAlt_r,\AtomAlt_r)})$ \\
     \SDNFLens{} & \GEq{} & $(\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
@@
 
-In greater detail, a symmetric DNF Lens \SDNFLens{} consists of three
+As an example, consider the symmetric DNF lens:
+$(\SDNFLensOf{(1,1,\SSQLens_1)\DNFLSep (1,2,\SSQLens_2)}
+,\ListOf{2},\ListOf{1;2})$, synchronizing data between $\DNFOf{\Sequence_1}$ and
+$\DNFOf{\SequenceAlt_1 \DNFSep \SequenceAlt_2}$. This lens merges both
+$\SequenceAlt_1$ and $\SequenceAlt_2$ into $\Sequence_1$, as they both are
+involved in a lens with $\Sequence_1$. When performing a \CreateR, $\SSQLens_2$
+is used. When performing a \CreateL on a string in $\Sequence_1$, $\SSQLens_1$
+is used, and when performing a \CreateL on a string in $\Sequence_2$,
+$\SSQLens_2$ is used. More generally, a symmetric DNF Lens \SDNFLens{} consists
+of three components. First is the mapping component. The mapping component is a
+list of triples $(i,j,\SSQLens)$. In this, $i$ and $j$ refer to sequence $i$ and
+sequence $j$ in the left and right DNF regular expressions, respectively. The
+third component of the triple, \SSQLens{}, is a sequence lens that maps between
+the two sequences specified by $i$ and $j$. The second component is a list of
+integers $c$. The integer, $c_i$ at position $i$ states that when performing a
+$\CreateR$ on a string matching sequence $\Sequence_i$, it will use the sequence
+lens that maps it to sequence $\SequenceAlt_{c_i}$ on the right. The third
+component is a list of integers $d$, and follows the same pattern as the second
+component, but maps from right to left.
+
+As an example, consider the symmetric sequence lens:
+$(\SSQLensOf{(1,2,\SAtomLens_1)} ,\ListOf{}
+,\ListOf{(\StringAlt_1',\AtomAlt_1)})$, synchronizing data between
+$\SequenceOf{\String_0 \SeqSep \Atom_1 \SeqSep \String_1}$ and
+$\SequenceOf{\StringAlt_0 \SeqSep \AtomAlt_1 \SeqSep \StringAlt_1 \SeqSep
+  \AtomAlt_2 \SeqSep \StringAlt_2}$. This lens maps $\Sequence_1$ to
+$\SequenceAlt_2$, and leaves $\AtomAlt_1$ disconnected. On a \CreateR, the
+string created is a concatenation of $\StringAlt_1'$ with the default from
+$\SAtomLens_1$. More generally, a sequence lens \SSQLens{} consists of three
 components. First is the mapping component. The mapping component is a list of
-triples $(i,j,\SSQLens)$. In this, $i$ and $j$ refer to sequence $i$ and
-sequence $j$ in the left and right DNF regular expressions, respectively. The
-third component of the triple, \SSQLens{}, is a sequence lens that maps between the two
-sequences specified by $i$ and $j$. The second component is a list of integers
-$c$. The integer, $c_i$ at position $i$ states that when performing a $\CreateR$
-on a string matching sequence $\Sequence_i$, it will use the sequence lens that
-maps it to sequence $\SequenceAlt_{c_i}$ on the right. The third component is a
-list of integers $d$, and follows the same pattern as the second component, but
-maps from right to left.
-
-A Sequence lens \SSQLens{} consists of three components. First is the mapping
-component. The mapping component is a list of triples $(i,j,\SAtomLens)$. In
-this, $i$ and $j$ refer to atom $i$ and atom $j$ on the left and right
-sequences, respectively. The third component of the triple, \SAtomLens{}, is an atom lens that
-maps between the two atoms specified by $i$ and $j$. The second component is a
-list of strings, $\String$. The string, $s_i$ at position $i$ states that when
-performing a $\CreateL$ on a string matching atom $\Atom_i$, it will do one of
-two things. If $\Atom_i$ is mapped to in the mapping component, then performing
-the atom's $\CreateL$ will actually $\PutL$ into that string. If $\Atom_i$ is
-not mapped to, then $\String_i$ will be created as a representative of
-$\Atom_i$.
+triples $(i,j,\SAtomLens)$. In this, $i$ and $j$ refer to atom $i$ and atom $j$
+on the left and right sequences, respectively. The third component of the
+triple, \SAtomLens{}, is an atom lens that maps between the two atoms specified
+by $i$ and $j$. The second component is a list of strings, $\String$. The
+string, $s_i$ at position $i$ states that when performing a $\CreateL$ on a
+string matching atom $\Atom_i$, it will do one of two things. If $\Atom_i$ is
+mapped to in the mapping component, then performing the atom's $\CreateL$ will
+actually $\PutL$ into that string. If $\Atom_i$ is not mapped to, then
+$\String_i$ will be created as a representative of $\Atom_i$.
 
@@
 procedure, the particulars of their typing and semantics are not the primary
-intellectual contributions to our work. The details of the typing and semantics
+intellectual contributions of our work. The details of the typing and semantics
 are included in the supplementary material present in the appendix.
@@
 \paragraph*{Information-Theoretic Cost Metric}
-Instead of searching for \emph{any} term that satisfies the specification, we
-search for the \emph{best} term that satisfies the specification. This begs the
-question: what metric should we use to determine which term is best?
-Intuitively, we aim to synthesize the lens which is ``closest'' to being a
-bijection. Somewhat more formally, we aim to synthesize a lens which minimizes
-the expected number of bits which must be used to recover the source from the
-target, and the target from the source.
-
 Much like we defined a syntactic means to find the expected number of bits to
@@
 extension to Boomerang. We added symmetric lens combinators to Boomerang, and
-implemented Boomerang's asymmetric lens combinators \sam{all of Boomerang's combinators?} using the symmetric lens combinators.
+implemented all Boomerang's asymmetric lens combinators using a combination of
+the symmetric lens combinators, as well as other symmetric lens combinators for
+the asymmetric theory extensions present in Boomerang, like matching lenses and
+quotient lenses.
 
@@
   ad-hoc data present in Linux configuration files into a structured, dictionary
-  representation. \sam{What does data cleaning mean?}
+  representation.
 \item We used \afm{8} benchmarks from synchronization tasks between a

### 2018-07-09 papers/SymmetricLenses/main.tex
Couple quick reactions

@@
   like UNK used before in practice, but if it makes the example cleaner and more
-  comprehensible, sure.  Anybody else have thoughts on the matter?}.
+  comprehensible, sure.  Anybody else have thoughts on the matter?}\bcp{No,
+  I meant why not concatenate two disconnect lenses, one relating salary to
+  the empty string and one relating the empty string to company.}.
 The defaults are strings that indicate the information is missing.
@@
   REs...}\afm{no known issues, I just didn't want to go into specifics about
-  ambiguity in an already-full paper}, we define the typing derivations and
+  ambiguity in an already-full paper}\bcp{OK, we can just comment that we
+  believe relaxing this restriction is unproblematic.}, we define the typing derivations and
 semantics only over unambiguous regular expressions, even though some lenses can

### 2018-07-09 papers/SymmetricLenses/main.tex
Responded to some comments in Section 7

@@
 \label{sec:relationship}
-\bcp{I suggest adding an adjective---e.g., ``classic symmetric lenses''---to
-avoid confusion between the old ones and our new ones.}
 \paragraph*{Formalization} We have reformulated the problem addressed by
@@
 initial value of $C$, and specifies default behavior when there isn't existing
-data to stitch with.  In our example\bcp{which example?}, the complement
-would consist of two lists 
-of commands, where one list contains the Linux commands and one list contains
-the Windows commands.
+data to stitch with.
+In the example in Figure~\ref{fig:minimized-representations} the complement would consist of a list of pairs with each pair consisting of an integral salaray value and the name of a company.
 
@@
 
-\paragraph*{Issues with Symmetric Lenses} Using symmetric lenses for
+\paragraph*{Issues with Symmetric Lenses}
+\iffalse
+Using symmetric lenses for
 synchronization forces us to maintain not just two files that we aim to
@@
 additional state needed.
-
-For synthesis, because each lens has a custom complement, we can no longer
+\fi
+For one, using classical symmetric lenses for synchronization forces us to maintain three files instead of two, which requires more engineering.
+The more significant issues with classical symmetric lenses arise in synthesis.
+Firstly, because each lens has a custom complement, we can no longer
 specify the put functions though mere input/output examples.  Instead, we have
@@
 
+\paragraph*{Relationship with Asymmetric Lenses}
+Formally, an {\em asymmetric lens} $ \ell : S \Leftrightarrow V$ is a triple of functions $\ell.\get : S \longrightarrow V$, $\ell.\pput : V \longrightarrow S \longrightarrow S$ and $\ell.\create : V \longrightarrow S$ satisfying the following laws \cite{Focal2005-long2}:
+\begin{align*}
+\ell.\get \; (\ell.\pput \; s \; v) &= v \tag{PUTGET}\\
+\ell.\pput \; s \; (\ell.\get \; s) &= s \tag{GETPUT}\\
+\ell.\get \; (\ell.\create \; v) &= v \tag{CREATEGET}
+\end{align*}
 Despite being less expressive than symmetric lenses, simple symmetric lenses are

### 2018-07-09 papers/SymmetricLenses/main.tex
A few more changes to Section 7

@@
 % begin related-work
-\section{Relationship To Classic Symmetric Lenses}
+\section{Relationship To Classical Symmetric Lenses}
 \label{sec:relationship}
-\paragraph*{Formalization} We have reformulated the problem addressed by
-symmetric lenses with simple symmetric lenses. Classically, these types of
-synchronization tasks are represented as symmetric lenses. A symmetric lens
+\paragraph*{Formalization} We have reformulated the problem addressed by classical
+symmetric lenses with simple symmetric lenses. A classical symmetric lens
 $\Lens$ between $X$ and $Y$ consists of 4 components: a complement $C$, a
@@
 
-Furthermore, two symmetric lenses are equivalent if they always output the same
+Furthermore, two classical symmetric lenses are equivalent if they always output the same
 formats on any sequence of edits. Formally, given a lens $\Lens \in X
@@
 
-\paragraph*{Issues with Symmetric Lenses}
+\paragraph*{Issues with Classical Symmetric Lenses}
 \iffalse
@@
 \paragraph*{Relation with Simple Symmetric Lenses} To formalize how simple
-symmetric lenses lenses work as synchronization tools, we define an $apply$
-function on them, much like we define an $apply$ on symmetric lenses. The
-function $apply$, given a symmetric lens and an optional pair of elements of the
+symmetric lenses work as synchronization tools, we define an $apply$
+function on them, much like we define an $apply$ on classical symmetric lenses. The
+function $apply$, given a simple symmetric lens and an optional pair of elements of the
 source and target, is a function from sequences of put objects to sequences of
-put objects. If $apply(\Lens,\None,es,es')$, then that means, starting with no
+put objects. If $apply(\Lens,\None,es) = es'$, then that means, starting with no
 prior data, after edit $es_i$, the lens \Lens synchronizes the other data to
-$es_i'$. If $apply(\Lens,\SomeOf{(x,y)},es,es')$, then that means, starting with
+$es_i'$. If $apply(\Lens,\SomeOf{(x,y)},es) = es'$, then that means, starting with
 data $x$ and $y$ on the left and right, respectively, after edit $es_i$, the
@@
 \begin{theorem}
-  Let $\Lens$ be a symmetric lens. The lens $\Lens$ is equivalent to a forgetful
+  Let $\Lens$ be a classical symmetric lens. The lens $\Lens$ is equivalent to a forgetful
   lens if, and only if, there exists a simple symmetric lens $\Lens'$ where
@@
   is just the identity relation?}
+\sam{No; for example the identity lens of a concatenation of two regular expressions is the same as concatenating the two identity lenses on the respective regular expressions}
 

### 2018-07-09 papers/SymmetricLenses/main.tex
add termination condition descritpion improvement

--- 2018-07-09 papers/SymmetricLenses/main.tex
@@
   like UNK used before in practice, but if it makes the example cleaner and more
-  comprehensible, sure.  Anybody else have thoughts on the matter?}\bcp{No,
-  I meant why not concatenate two disconnect lenses, one relating salary to
-  the empty string and one relating the empty string to company.}.
+  comprehensible, sure.  Anybody else have thoughts on the matter?}.
 The defaults are strings that indicate the information is missing.
@@
   REs...}\afm{no known issues, I just didn't want to go into specifics about
-  ambiguity in an already-full paper}\bcp{OK, we can just comment that we
-  believe relaxing this restriction is unproblematic.}, we define the typing derivations and
+  ambiguity in an already-full paper}, we define the typing derivations and
 semantics only over unambiguous regular expressions, even though some lenses can
@@
 
-The procedure \PCF{\Continue} terminates when the sum of the number of regular expressions of the
-smallest unproposed size with the distance of those regular expressions from the
-originals exceeds the cost of the current best lens. At this point, \RXSearch
-stops proposing regular expression pairs, and instead returns to the user the
-best lens found thus far. Note this means that if \RXSearch finds a bijective
-lens, which has zero cost it will immediately return.
+The procedure \PCF{\Continue} returns false based on the how long the search has
+been going, and how hard it expects the next class of problems to be. In
+particular, if $\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at
+distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
+current best lens has cost $c$:
+\begin{gather*}
+  \Call{Continue}{pq} := c >= d + \Log_2(n)
+\end{gather*}
+
+When $\Call{Continue}{pq}$ returns $\False$, \RXSearch stops proposing regular
+expression pairs, and instead returns to the user the best lens found thus far.
+Note this means that if \RXSearch finds a bijective lens, which has zero cost it
+will immediately return.
 %
@@
 hash-consing and memoization, and compositional synthesis to our algorithm. We
-also enable an additional parameter in our specifications that customizes how
-permissive \RXSearch{} is in accepting proposed lenses.
+also enable an additional parameter in our specifications that customizes the
+termination condition for our procedure.
 
@@
 
-\paragraph*{\RXSearch Permissiveness Customization}
+\paragraph*{Termination Customization}
 In this work, we are constrained by three competing constraints. (1) We want to
@@
 
+For these types of situations, users can input a custom number, $t$, to
+customize the strictness of the termination condition. In particular, if
+$\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at distance $d$, the
+number of pairs in $pq$ at distance $d$ is $n$, and the current best lens has
+cost $c$:
+\begin{gather*}
+  \Call{Continue}{pq} := c >= \Min(d + \Log_2(n) - t, 0)
+\end{gather*}
+With a positive $t$, the termination condition is more strict, and the algorithm
+will spend longer searching for more likely lenses. With a negative $t$ the
+termination condition is less strict, and the algorithm will spend less time
+searching for more likely lenses.  Our default $t$ is set to $0$.
+
 While our defaults termination condition quickly finds the correct lens for
@@
 one-size-fits all. Particularly complex lenses, or lenses that project a
-particularly large amount of data, can be synthesized more easily and more
-quickly by customizing how permissive \RXSearch is in returning lenses.
+particularly large amount of data, can be synthesized more easily and quickly by
+customizing how permissive \RXSearch is in returning lenses.
 % end synthesis

### 2018-07-09 papers/SymmetricLenses/main.tex
Half an abstract -- will write more on it tomorrow if no one beats me to it

@@
 %%   application-specific data storage files, and data cleaning tasks.
-TODO: abstract
+
+(The idea of lenses.  Their usefulness and popularity.)
+
+Recent work by~\citet{miltner+:bijective-synthesis} has also demonstrated
+how techniques from {\em type-directed program synthesis} can be used to
+efficiently synthesize the simplest class of lenses: so-called {\em
+  bijective lenses}, where no information is discarded when transforming
+data in either direction.
+
+We show here how to extend this algorithm to a much richer class of lenses,
+which we call {\em simple symmetric lenses}; this class includes all
+bijective and standard ``asymmetric'' lenses~\cite{OriginalFosterPOPLpaper},
+as well as a rich subset of the full-blown ``symmetric lenses'' studied by
+\citet{symmetric-lenses}.
+
+(why our algorithm is harder; how we use information theory to help)
 

### 2018-07-09 papers/SymmetricLenses/main.tex
related

@@
 % begin related-work
-\section{Relationship To Classical Symmetric Lenses}
+\section{Related Work}
+
+In this paper, we have designed and implemented a new, pragmatic
+formulation of symmetric lenses as well as new program synthesis
+techniques.  In this section, we consider the relationship with
+classical symmetric lenses in depth and also comment on related
+program synthesis techniques.
+
+\subsection{Relationship with Classical Symmetric Lenses}
 \label{sec:relationship}
-\paragraph*{Formalization} We have reformulated the problem addressed by classical
-symmetric lenses with simple symmetric lenses. A classical symmetric lens
+%\paragraph*{Formalization}
+%We have reformulated the problem addressed by classical
+%symmetric lenses with simple symmetric lenses.
+A classical symmetric lens
 $\Lens$ between $X$ and $Y$ consists of 4 components: a complement $C$, a
-designated element $init \in C$, and two functions for propagating changes in
-each format to the other:
+designated element $init \in C$, and two functions,
+$\PutRSym$ and $\PutLSym$, that propagate changes in one format to the other.
 \begin{enumerate}
@@
 \end{enumerate}
-In this formulation, the components of each format that aren't shared with the
-other format are stored in the complement. When one format is edited, the \PutR
-or \PutL function stitches together that edited data with the data for the other
-side, stored in the complement.  The designated element, $init$, is used as the
-initial value of $C$, and specifies default behavior when there isn't existing
-data to stitch with.
-In the example in Figure~\ref{fig:minimized-representations} the complement would consist of a list of pairs with each pair consisting of an integral salaray value and the name of a company.
-
-These components only make sense for synchronization when they satisfy
-round-tripping laws.  This guarantees that if an identity edit is propagated, no
-data gets changed.
+In this formulation, data unique to each side are stored in the complement.
+When one format is edited, the \PutR
+or \PutL function stitches together the edited data with data
+stored in the complement.  The  $init$ element is used as the
+initial value of $C$ and specifies default behavior when data is missing.
+For instance, to implement the
+scenario in Figure~\ref{fig:minimized-representations}, the complement
+would consist of a list of pairs of salary and company name.
+
+Classical symmetric lenses satisfy the following equational laws.
 \begin{equation}
@@
 
-Furthermore, two classical symmetric lenses are equivalent if they always output the same
-formats on any sequence of edits. Formally, given a lens $\Lens \in X
-\leftrightarrow Y$, an \emph{edit} for $\Lens$ is a member of $X + Y$. Consider
+Two classical symmetric lenses are equivalent if they output the same
+formats given any sequence of edits. Formally, given a lens
+$\Lens \in X \leftrightarrow Y$,
+an \emph{edit} for $\Lens$ is a member of $X + Y$. Consider
 the function $apply$, which, given a lens and an element of that lens's
@@
 
-\section{Related Work}
-\label{sec:related}
-There is related work in a number of directions. There is related work in string
-transformation synthesis and type-directed synthesis. We also refer the reader
-to a recent, general survey on synthesis~\ref{?}.
-
-\subsection{String Transformation Synthesis}
-Much of the research on string transformation synthesis is on synthesizing
-string transformations from input/output
-specifications~\cite{flashfill,flashextract,?}. While this provides an easier
-interaction model with the user than this work, it requires the algorithm to
-learn the data formats from the examples. This makes the task much more
-difficult for the synthesis algorithm, particularly in addressing iterations and
-disjunctions. Prior work has shown that learning from examples and data format
-descriptions enables synthesis of much more complex transformations than are
-permitted by examples alone. Recent work has gone into making these algorithms
-more robust by delaying the selection of program until seeing what further
-inputs the program is run on~\cite{?}. By doing this, the synthesis algorithm
-can learn additional information about the formats from the use of the program.
-
-In many of these synthesis algorithms the correct transformation is chosen by
-principals like ``minimize constants''~\cite{?,?,?} and ``most specific
-generalization''~\cite{?,?,?}. We are the first work to our knowledge to use an
-information-theoretic measure for finding the correct transformation. However,
-much of our metrics are compatible with existing metrics: avoiding \Disconnect
-lenses minimizes constants and creates more specfic transformations (\Disconnect
-lenses permit more synchronized data).
-
-Optician~\ref{?} synchronizes string transformation functions, but prior
-versions were very restrictive in which transformations were permitted.
-Originally, Optician only synchronized transformations between formats in
-bijective correspondence, but few formats satisfy that constraint. Recent
-enhancements have permitted synthesis of more expressive lenses, quotient
-bijective lenses. These enhancements were made by strengthening the
-specifications from regular expressions to QREs, regular expressions annotated
-with information on normal forms of the language. This enhancement permits
-synthesizing transformations between formats whose normal forms are in bijective
-correspondence. While we have not integrated our system to permit QRE
-specifications, we see no reason the two systems cannot be integrated.
-
-Other work synthesizing invertible transformations was done by inverting a
-previously defined transformation.  This has been done by exploiting
-parametricity to invert transformations~\cite{?}, and by inverting symbolic
-finite automata~\cite{?}.
-
-\subsection{Type-Directed Synthesis}
-
-Type-directed synthesis uses type information to tighten specifications and
-reduce the number of possible programs. Typically, type-directed synthesis
-techniques use the types to orient the search space.  Our work interestingly
-deals with types with many semantic equivalences on them. InSynth~\cite{?}
-deals with the same problem when trying to synthesize transformations involving
-library functions. InSynth addresses this in a similar manner to this work: by
-normalizing the types we don't need to search through the equivalences on those
-types.
+\subsection{Data Transformation Synthesis}
+Over the past decade, the programming languages community has explored
+the synthesis of programs from a wide variety of angles.  One of
+the key ideas is typically to narrow the program search space by focusing
+on a specific domain, as we have done here.
+%, and as has been advocated
+%more generally under the rubric of syntax-guided synthesis~\cite{sygus}.
+
+Automation of string transformations, in particular, has been the focus
+of much prior attention.  For example, Gulwani and others work on FlashFill
+generates one-way spreadsheet transformations from input/output
+examples~\cite{flashfill,flashextract}.  On the one hand,
+FlashFill is easier to use because one need not specify the type of
+the data being transformed.  On the other hand, this type information
+makes it possible to transform more complex formats---Flash Fill
+does not synthesize programs with nested loops, for instance,
+and hence is incapable of synthesizing the majority of our
+benchmarks, even in one direction~\cite{flashfill}.
+%% Flashfill, designed
+%% for transformation of the cells of spreadsheets, does not scale to
+%% more complex formats
+%% While this provides an easier
+%% interaction model with the user than this work, it requires the algorithm to
+%% learn the data formats from the examples. This makes the task much more
+%% difficult for the synthesis algorithm, particularly in addressing iterations and
+%% disjunctions. Prior work has shown that learning from examples and data format
+%% descriptions enables synthesis of much more complex transformations than are
+%% permitted by examples alone. Recent work has gone into making these algorithms
+%% more robust by delaying the selection of program until seeing what further
+%% inputs the program is run on~\cite{?}. By doing this, the synthesis algorithm
+%% can learn additional information about the formats from the use of the program.
+
+All pragmatic synthesis algorithms use heuristics of one kind or another.  One of the
+goals of the current paper is to try to ground those heuristics in a broader theory,
+information theory,
+with the hope that this theory may inform future design decisions and help us understand
+heuristics crafted in other tools and possibly in other domains.  For instance, we
+speculate that some of the heuristics used in
+FlashFill (to take one well-documented example) may be connected to some of the principles
+laid out here.  For instance, FlashFill prioritizes the substring constructor over the
+constant string constructor when ranking possible programs. Such a choice may be understood
+in information-theoretic terms as the translation from target back to source would require
+more information.  Said another way, our information theoretic metrics measure injectivity
+(in both directions, in our case, to measure bijectivity) but applied in one direction, as
+is the case for FlashFill, would avoid the highly injective constant functions.  Beyond
+constants, another FlashFill heuristic prefers ``wider'' character classes (or token classes)
+over ``narrower'' ones.  Selecting a wider set of strings to copy from left-to-right, viewed
+in the reverse direction, from right-to-left, results in a function that is closer to being
+injective, as measured by information theory.  These are just a few of the heuristics used
+in Flashfill; more research would be required (and public disclosure of the details necessary)
+to understand the connections fully.
+
+%% In many of these synthesis algorithms the correct transformation is chosen by
+%% principals like ``minimize constants''~\cite{?,?,?} and ``most specific
+%% generalization''~\cite{?,?,?}. We are the first work to our knowledge to use an
+%% information-theoretic measure for finding the correct transformation. However,
+%% much of our metrics are compatible with existing metrics: avoiding \Disconnect
+%% lenses minimizes constants and creates more specfic transformations (\Disconnect
+%% lenses permit more synchronized data).
+
+As discussed in the introduction,
+the Optician tool~\cite{optician,maina+:quotient-synthesis} was a building block for
+our work.  Optician synthesized bijective transformations~\cite{optician} and
+bijections modulo quotients~\cite{maina+:quotient-synthesis}, but could
+not synthesize more complex bidirectional transformations where one format contains
+important information not present in the other---a common situation in the real
+world.  From a technical perspective, the first key novelty in our work involves the definition, theory,
+analysis, and implementation of a new class of simple symmetric lenses, designed
+for synthesis.  The second key technical innovation involves the use of stochastic
+regular expressions and information theory to guide the search for program transformations.
+As mentioned earlier, we believe such information theoretic techniques may have broad utility
+in helping us understand how to formulate a search for a data transformation function.
+%% The central contributions
+%% only synchronized transformations between formats in
+%% bijective correspondence, but few formats satisfy that constraint. Recent
+%% enhancements have permitted synthesis of more expressive lenses, quotient
+%% bijective lenses. These enhancements were made by strengthening the
+%% specifications from regular expressions to QREs, regular expressions annotated
+%% with information on normal forms of the language. This enhancement permits
+%% synthesizing transformations between formats whose normal forms are in bijective
+%% correspondence. While we have not integrated our system to permit QRE
+%% specifications, we see no reason the two systems cannot be integrated.
+
+While our tool uses types and examples to specify invertable transformations, other
+tools have been shown to synthesize a backwards transformation from ordinary code designed
+to implement the forwards transformation.  For instance, Hu and
+D'Antoni~\cite{program-inversion-symbolic-transducers} show how to invert transformations
+using symbolic finite automata, and
+Voigtl\"{a}nder~\cite{bidirectionalization-for-free} demonstrates how to
+construct reverse transformations from forwards transformations
+by exploiting parametricity properties.  These kinds of tools are very useful, but in different
+circumstances from ours (namely, when one already has the code to implement one direction of
+the transformation).
+
+%% Other work synthesizing invertible transformations was done by inverting a
+%% previously defined transformation.  This has been done by exploiting
+%% parametricity to invert transformations~\cite{}, and by inverting symbolic
+%% finite automata~\cite{?}.
+
+%% \subsection{Type-Directed Synthesis}
+
+%% Type-directed synthesis uses type information to tighten specifications and
+%% reduce the number of possible programs. Typically, type-directed synthesis
+%% techniques use the types to orient the search space.  Our work interestingly
+%% deals with types with many semantic equivalences on them. InSynth~\cite{?}
+%% deals with the same problem when trying to synthesize transformations involving
+%% library functions. InSynth addresses this in a similar manner to this work: by
+%% normalizing the types we don't need to search through the equivalences on those
+%% types.
 

### 2018-07-09 papers/SymmetricLenses/main.tex
stuff

@@
 the transformation).
+
+More broadly, information theory and probabilistic languages are common tools
+in natural language understanding,  machine translation, information retrieval,
+data extraction and grammatical inference (see Pereira~\cite{pereira:info-theory},
+for an introduction).  Indeed, our work was inspired, in part, by the PADS format
+inference tool~\cite{pads:synthesis}, which was in turn inspired by earlier work on
+probabilistic gramatical inference
+and information extraction~\cite{kushmerick-thesis,arasu:extracting}.  More specifically, PADS
+used an information-theoretic metric to help judge the quality of inferred grammars
+for ad hoc data formats.  However, PADS did not synthesize data transformers, and
+we are not aware of the use of information-theoretic principles in type-directed or syntax-guided
+synthesis of DSL programs.
 

### 2018-07-09 papers/SymmetricLenses/main.tex
stuff

@@
 formulation of symmetric lenses as well as new program synthesis
-techniques.  In this section, we consider the relationship with
-classical symmetric lenses in depth and also comment on related
+techniques.  In this section, we analyze the relationship
+between our simple symmetric
+lenses and two previous lens languages:  classical symmetric lenses
+and asymmetric lenses.  We also comment on related
 program synthesis techniques.
 
-\subsection{Relationship with Classical Symmetric Lenses}
+\subsection{Relationship with Classical Lens Languages}
 \label{sec:relationship}
@@
 %symmetric lenses with simple symmetric lenses.
+
+\paragraph*{Symmetric Lenses}
 A classical symmetric lens
@@
 \]
-
+\noindent
 Two lenses, $\Lens_1$ and $\Lens_2$, are equivalent if
@@
 
-By describing our lenses in this complement style, we can express an incredibly
-wide class\bcp{where is the evidence that this class of lenses is {\em
-    usefully} wider than simple symmetric lenses?} of lenses, as these complements permit storing any amount of
-information in them.  However, they pose problems when implementing them in
-real-world systems and in synthesizing them.
-
-\paragraph*{Issues with Classical Symmetric Lenses}
+%% By describing our lenses in this complement style, we can express an incredibly
+%% wide class\bcp{where is the evidence that this class of lenses is {\em
+%%     usefully} wider than simple symmetric lenses?} of lenses, as these complements permit storing any amount of
+%% information in them.  However, they pose problems when implementing them in
+%% real-world systems and in synthesizing them.
+
+%\paragraph*{Issues with Classical Symmetric Lenses}
 \iffalse
@@
 \fi
-For one, using classical symmetric lenses for synchronization forces us to maintain three files instead of two, which requires more engineering.
-The more significant issues with classical symmetric lenses arise in synthesis.
-Firstly, because each lens has a custom complement, we can no longer
-specify the put functions though mere input/output examples.  Instead, we have
-to specify the behavior through the observable behavior of edit sequences.
-Furthermore, there are lenses that require arbitrarily long edit sequences to
-differentiate between: consider the lens that keeps two formats equal for the
-first $k$ non-identity edits, then keeps them completely unsynchronized for all
-future edits.  This requires an edit sequence of $k$ edits to differentiate
-between this lens and the identity lens.  Simple symmetric lenses avoid this
-problem by not allowing such lenses: we can't keep track of the number of
-non-identity edits performed unless there are explicit fields in the data
-formats keeping track of them.  While we are reducing the expressivity of our
-lenses with this restriction, we feel the benefits outweigh the costs.
+While classical symmetric are more expressive than simple symmetric
+lenses, it is unclear whether this additional expressiveness is useful in
+practice.  Moreover, classic symmetric lenses come with 
+some disadvantages.
+For one, an implementation of classical symmetric lenses
+would need to squirrel away the complement in
+an auxiliary file somwhere, which likely requires more engineering.
+However, more significant issues arise in synthesis.
+First, because each lens has a custom complement, one can no longer
+specify the put functions through input/output examples alone.  One
+alternative would be to enrich specifications with edit sequentials;
+another would be to specify the structure of complements explicitly
+(though the latter would be somewhat akin to specifying the internal
+state of a program).  In either case, the complexity of the specifications
+increases.
+%% addition, users would have to specify some auxilary 
+%% to specify the behavior through the observable behavior of edit sequences.
+%% Second, there are lenses that require arbitrarily long edit sequences to
+%% differentiate between: consider the lens that keeps two formats equal for the
+%% first $k$ non-identity edits, then keeps them completely unsynchronized for all
+%% future edits.  This requires an edit sequence of $k$ edits to differentiate
+%% between this lens and the identity lens.  Simple symmetric lenses avoid this
+%% problem by disallowing such lenses: we can't keep track of the number of
+%% non-identity edits performed unless there are explicit fields in the data
+%% formats keeping track of them.  While we are reducing the expressivity of our
+%% lenses with this restriction, we feel the benefits outweigh the costs.
 
@@
 symmetric lenses work as synchronization tools, we define an $apply$
-function on them, much like we define an $apply$ on classical symmetric lenses. The
-function $apply$, given a simple symmetric lens and an optional pair of elements of the
-source and target, is a function from sequences of put objects to sequences of
-put objects. If $apply(\Lens,\None,es) = es'$, then that means, starting with no
-prior data, after edit $es_i$, the lens \Lens synchronizes the other data to
-$es_i'$. If $apply(\Lens,\SomeOf{(x,y)},es) = es'$, then that means, starting with
+function on them, much like one defines an $apply$ on classical symmetric lenses:
+Given a simple symmetric lens and an optional pair of 
+source and target strings,
+$apply$ is a function from sequences of put objects to sequences of
+put objects. If $apply(\Lens,\None,es) = es'$, then starting with no
+prior data, after edit $es_i$, the lens \Lens synchronizes the other data\dpw{Is the ``other data'' $es$? or is it something else? If it is $es$ then we can say ($es$) in the sentence.} to
+$es_i'$. \dpw{I wasn't sure what a ``put object'' was.  Is it a function? (If so, perhaps
+call it a function and give its type?)} If $apply(\Lens,\SomeOf{(x,y)},es) = es'$, then that means, starting with
 data $x$ and $y$ on the left and right, respectively, after edit $es_i$, the
@@
 
-We've reformulated symmetric lenses in a style similar to existing asymmetric
+We have reformulated symmetric lenses in a style similar to existing asymmetric
 lens formulations. However, some symmetric lenses certainly are inexpressible in
@@
 with a restriction on symmetric lenses. We define \emph{forgetful symmetric
-  lenses} to be symmetric lenses which satisfy the following additional laws:
+  lenses} to be symmetric lenses that satisfy the following additional laws.
 \begin{equation}
@@
 
-Intuitively, this says that the complements are uniquely determined by the most
-recently input $x$ and $y$. This corresponds closely with simple symmetric
+Intuitively, these equations state that complements are uniquely determined by the most
+recently input $x$ and $y$. Hence, such lenses correspond closely \dpw{replace ``closely'' with
+``exactly''?} with simple symmetric
 lenses, where all state is maintained by the current state of the $x$ and $y$
@@
 
-\paragraph*{Relationship with Asymmetric Lenses}
+\paragraph*{Asymmetric Lenses}
 Formally, an {\em asymmetric lens} $ \ell : S \Leftrightarrow V$ is a triple of functions $\ell.\get : S \longrightarrow V$, $\ell.\pput : V \longrightarrow S \longrightarrow S$ and $\ell.\create : V \longrightarrow S$ satisfying the following laws \cite{Focal2005-long2}:
@@
 the key ideas is typically to narrow the program search space by focusing
-on a specific domain, as we have done here.
+on a specific domain, and imposing constraints on syntax~\cite{sygus} or
+typing~\cite{augustsson-2004,gvero-pldi-2013,osera+:pldi15,feser-pldi-2015,scherer-icfp-2015,frankle+:popl16,}, or both.
 %, and as has been advocated

### 2018-07-09 papers/SymmetricLenses/main.tex
cite cleaned up a bit

@@
 
-Recent work by~\citet{miltner+:bijective-synthesis} has also demonstrated
+Recent work by~\citet{optician} has also demonstrated
 how techniques from {\em type-directed program synthesis} can be used to
@@
 Initially, we hypothesized that extending the synthesis algorithms
-of Miltner \ETAL~\cite{miltner+:bijective-synthesis} to operate over symmetric lenses
+of Miltner \ETAL~\cite{optician} to operate over symmetric lenses
 as opposed to bijective lenses 
@@
 Now that we have a way to calculate likelihood, we need a way to search for
-likely lenses. Previous work on synthesis~\cite{?,?,?} has shown that we can use
+likely lenses. Previous work on synthesis~\cite{augustsson-2004,osera+:pldi15,feser-pldi-2015,frankle+:popl16} has shown that we can use
 types to dramatically restrict the search space and improve the effectiveness of
@@
 two communicating synthesizers following earlier
-work~\cite{quotient-lens-synthesis}. The first synthesizer, \Expand, navigates
+work~\cite{maina+:quotient-synthesis}. The first synthesizer, \Expand, navigates
 the space of semantically equivalent regular expressions by applying rewrite
@@
 \section{Related Work}
+\label{sec:related}
 
@@
 on a specific domain, and imposing constraints on syntax~\cite{sygus} or
-typing~\cite{augustsson-2004,gvero-pldi-2013,osera+:pldi15,feser-pldi-2015,scherer-icfp-2015,frankle+:popl16,}, or both.
+typing~\cite{augustsson-2004,gvero-pldi-2013,osera+:pldi15,feser-pldi-2015,scherer-icfp-2015,frankle+:popl16}, or both.
 %, and as has been advocated
@@
 generates one-way spreadsheet transformations from input/output
-examples~\cite{flashfill,flashextract}.  On the one hand,
+examples~\cite{flashfill,le-pldi-2014}.  On the one hand,
 FlashFill is easier to use because one need not specify the type of

### 2018-07-09 papers/SymmetricLenses/main.tex
add some changes to main

@@
 returned ($\CreateROf{\String} =
-\lstinline{"UNK"}$), and on puts, the second argument is used, and the first is
+\lstinline{"unk"}$), and on puts, the second argument is used, and the first is
 thrown away ($\PutROf{\String}{\StringAlt} = \StringAlt$). In our extended
-example, the \lstinline{salary} and \lstinline{company} fields are independent,
-so we use the disconnect lens to relate them\bcp{Potentially confusing to
-  readers (and, even to me, hacky). Why not relate each of them to the constant
-  empty string?}\afm{I'm fine with the empty string.  I've seen sentinel values
-  like UNK used before in practice, but if it makes the example cleaner and more
-  comprehensible, sure.  Anybody else have thoughts on the matter?}\bcp{No,
-  I meant why not concatenate two disconnect lenses, one relating salary to
-  the empty string and one relating the empty string to company.}.
+example, the \lstinline{salary} field is only present in management files,
+so we use the disconnect lens to ensure salary edits do not cause updates to the
+HR file.
 The defaults are strings that indicate the information is missing.
@@
 \begin{lstlisting}
-disconnect(salary, company, "unk", "UNK") : salary $\Leftrightarrow$ company
+disconnect(salary, "", "unk", "") : salary $\Leftrightarrow$ company
 \end{lstlisting}
@@
 let name_lens = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
-let employee_lens = name_lens . disconnect(salary,company,"unk","UNK") in
+let employee_lens = name_lens . disconnect(salary,"","unk","") 
+                                                    . disconnect("",company,"","UNK") in
 let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
@@
 We now provide the typing derivations and semantics for these lenses. For
-simplicity of presentation,\bcp{Is there any known issue with relaxing this
-  restriction? I've often found ti convenient to use id lenses with ambiguous
-  REs...}\afm{no known issues, I just didn't want to go into specifics about
-  ambiguity in an already-full paper}\bcp{OK, we can just comment that we
-  believe relaxing this restriction is unproblematic.}, we define the typing derivations and
-semantics only over unambiguous regular expressions, even though some lenses can
-be defined over ambiguous regular expressions (for example, the identity lens
-doesn't need any ambiguity constraints).
+simplicity of presentation, we define the typing derivations and semantics only
+over unambiguous regular expressions, even though some lenses can be defined
+over ambiguous regular expressions (for example, the identity lens doesn't need
+any ambiguity constraints). However, we believe relaxing this restriction is
+unproblematic.
 
@@
 \paragraph*{Relation with Simple Symmetric Lenses} To formalize how simple
-symmetric lenses work as synchronization tools, we define an $apply$
-function on them, much like one defines an $apply$ on classical symmetric lenses:
-Given a simple symmetric lens and an optional pair of 
-source and target strings,
-$apply$ is a function from sequences of put objects to sequences of
-put objects. If $apply(\Lens,\None,es) = es'$, then starting with no
-prior data, after edit $es_i$, the lens \Lens synchronizes the other data\dpw{Is the ``other data'' $es$? or is it something else? If it is $es$ then we can say ($es$) in the sentence.} to
-$es_i'$. \dpw{I wasn't sure what a ``put object'' was.  Is it a function? (If so, perhaps
-call it a function and give its type?)} If $apply(\Lens,\SomeOf{(x,y)},es) = es'$, then that means, starting with
-data $x$ and $y$ on the left and right, respectively, after edit $es_i$, the
-lens \Lens synchronizes the other data to $es_i'$.
+symmetric lenses work as synchronization tools, we define an $apply$ function on
+them, much like one defines an $apply$ on classical symmetric lenses: Given a
+simple symmetric lens and an optional pair of source and target strings, $apply$
+is a function from sequences of edits to sequences of edits. If
+$apply(\Lens,\None,es) = es'$, then starting with no prior data, after edit
+$es_i$, the lens \Lens synchronizes the updates the data in the non-edited
+format (the right format if $es_i = \InLOf{x}$, and the left format if $es_i =
+\InROf{y}$) to $es_i'$. If $apply(\Lens,\SomeOf{(x,y)},es) = es'$, then that
+means, starting with data $x$ and $y$ on the left and right, respectively, after
+edit $es_i$, the lens \Lens synchronizes the other data to $es_i'$.
 
@@
   is just the identity relation?}
-\sam{No; for example the identity lens of a concatenation of two regular expressions is the same as concatenating the two identity lenses on the respective regular expressions}
+\sam{No; for example the identity lens of a concatenation of two regular
+  expressions is the same as concatenating the two identity lenses on the
+  respective regular expressions}
+\afm{I think Benjamin likely means identity relation on the semantic meaning of
+  the lenses, which is true.  Would this be a useful thing to prove? It's a 2
+  liner.}
 

### 2018-07-09 papers/SymmetricLenses/main.tex
Fexed some spacing issues in Section 4

@@
 \documentclass[acmsmall,screen,anonymous]{acmart}
-\settopmatter{}
+\settopmatter{printacmref=false}
 
@@
 \newtheorem*{theorem*}{Theorem}
+\newenvironment{centermath}
+ {\begin{center}$\displaystyle}
+ {$\end{center}}
 
@@
   \renewcommand\footnotetextcopyrightpermission[1]{} % removes footnote with conference information in first column
-
 %\setcopyright{rightsretained}
@@
 
-\begin{CCSXML}
-<ccs2012>
-<concept>
-<concept_id>10011007.10011006.10011050.10011017</concept_id>
-<concept_desc>Software and its engineering~Domain specific languages</concept_desc>
-<concept_significance>500</concept_significance>
-</concept>
-<concept>
-<concept_id>10011007.10011006.10011066.10011070</concept_id>
-<concept_desc>Software and its engineering~Application specific development environments</concept_desc>
-<concept_significance>300</concept_significance>
-</concept>
-</ccs2012>
-\end{CCSXML}
-
-\ccsdesc[500]{Software and its engineering~Domain specific languages}
-\ccsdesc[300]{Software and its engineering~Application specific development environments}
+%\begin{CCSXML}
+%<ccs2012>
+%<concept>
+%<concept_id>10011007.10011006.10011050.10011017</concept_id>
+%<concept_desc>Software and its engineering~Domain specific languages</concept_desc>
+%<concept_significance>500</concept_significance>
+%</concept>
+%<concept>
+%<concept_id>10011007.10011006.10011066.10011070</concept_id>
+%<concept_desc>Software and its engineering~Application specific development environments</concept_desc>
+%<concept_significance>300</concept_significance>
+%</concept>
+%</ccs2012>
+%\end{CCSXML}
+%
+%\ccsdesc[500]{Software and its engineering~Domain specific languages}
+%\ccsdesc[300]{Software and its engineering~Application specific development environments}
 
@@
 
-\keywords{Bidirectional Programming, Program Synthesis, Type-Directed Synthesis,
-Type Systems}
+%\keywords{Bidirectional Programming, Program Synthesis, Type-Directed Synthesis,
+%Type Systems}
 
@@
 
-\begin{center}
-  \begin{tabular}{rcl}
-    \Regex{},\RegexAlt{}
-    & \GEq{} & $\String$ \\
-    & \GBar{} & $\emptyset$ \\
-    & \GBar{} & $\PRegexStar{\Regex}{p}$ \\
-    & \GBar{} & $\RegexConcat{\Regex_1}{\Regex_2}$ \\
-    & \GBar{} & $\PRegexOr{\Regex_1}{\Regex_2}{p}$
-  \end{tabular}
-\end{center}
+$$\Regex{},\RegexAlt{} \GEq{}  \String  \; | \; \emptyset \; | \; \PRegexStar{\Regex}{p} \; | \; \RegexConcat{\Regex_1}{\Regex_2} \; | \; \PRegexOr{\Regex_1}{\Regex_2}{p}$$
 
@@
 
-% begin ssl
 \section{Simple Symmetric String Lenses}
@@
 
-\begin{center}
-  \begin{tabular}{l@{\ }l@{\ }l@{\ }l}
-    % REGEX
-    \Lens{} & \GEq{} & & $\IdentityLensOf{\BRegex}$\\
-            & & \GBar{} & $\DisconnectOf{\BRegex}{\BRegexAlt}{\String}{\StringAlt}$ \\
-            & & \GBar{} & $\IterateLensOf{\Lens}$ \\
-            & & \GBar{} & $\ConcatLensOf{\Lens_1}{\Lens_2}$ \\
-            & & \GBar{} & $\SwapLensOf{\Lens_1}{\Lens_2}$ \\
-            & & \GBar{} & $\OrLensOf{\Lens_1}{\Lens_2}$ \\
-            & & \GBar{} & $\ComposeLensOf{\Lens_1}{\Lens_2}$ \\
-            & & \GBar{} & $\MergeROf{\Lens_1}{\Lens_2}$ \\
-            & & \GBar{} & $\MergeLOf{\Lens_1}{\Lens_2}$ \\
-            & & \GBar{} & $\InvertOf{\Lens}$ \\
-  \end{tabular}
-\end{center}
+\begin{align*}
+\ell ::= &\IdentityLensOf{\BRegex} \; | \; \DisconnectOf{\BRegex}{\BRegexAlt}{\String}{\StringAlt} \; | \; \IterateLensOf{\Lens} \; | \; \ConcatLensOf{\Lens_1}{\Lens_2} \; | \; \SwapLensOf{\Lens_1}{\Lens_2} \; | \; \OrLensOf{\Lens_1}{\Lens_2} \; | \; \ComposeLensOf{\Lens_1}{\Lens_2} \; |\\
+&\MergeROf{\Lens_1}{\Lens_2} \; | \; \MergeLOf{\Lens_1}{\Lens_2} \; | \; \InvertOf{\Lens}
+\end{align*}
 
@@
 We now provide the typing derivations and semantics for these lenses. For
-simplicity of presentation, we define the typing derivations and semantics only
-over unambiguous regular expressions, even though some lenses can be defined
-over ambiguous regular expressions (for example, the identity lens doesn't need
-any ambiguity constraints). However, we believe relaxing this restriction is
-unproblematic.
-
-\[
-  \centering
+simplicity of presentation,\bcp{Is there any known issue with relaxing this
+  restriction? I've often found ti convenient to use id lenses with ambiguous
+  REs...}\afm{no known issues, I just didn't want to go into specifics about
+  ambiguity in an already-full paper}\bcp{OK, we can just comment that we
+  believe relaxing this restriction is unproblematic.}, we define the typing derivations and
+semantics only over unambiguous regular expressions, even though some lenses can
+be defined over ambiguous regular expressions (for example, the identity lens
+doesn't need any ambiguity constraints).
+
+\begin{centermath}
+\begin{array}[b]{l@{\qquad \; \qquad}l}
+\begin{array}[b]{c}
+\quad\\
   \inferrule*
@@
   }
-\]
-\begin{center}
-  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
-    $\CreateR{} \App s$ & = & $s$\\
-    $\CreateL{} \App s$ & = & $s$\\
-    $\PutR{} \App s_1 \App s_2$ & = & $s_1$\\
-    $\PutL{} \App s_1 \App s_2$ & = & $s_1$
-  \end{tabular}
-\end{center}
-
+  \\
+  \quad
+  \end{array}
+&
+\begin{array}[b]{r@{\ }c@{\ }l}
+    \CreateR{} \App s & = & s\\
+    \CreateL{} \App s & = & s\\
+    \PutR{} \App s_1 \App s_2 & = & s_1\\
+    \PutL{} \App s_1 \App s_2 & = & s_1
+  \end{array}
+\end{array}
+\end{centermath}
 Note that the identity lens ignores the second argument in the put functions.
@@
 
-\[
-  \centering
+\begin{centermath}
+\begin{array}[b]{l@{\qquad \; \qquad}l}
   \inferrule*
@@
   }
-\]
-\begin{center}
-  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+&
+  \begin{array}{@{}r@{\ }c@{\ }l@{}}
     $\CreateR{} \App \String'$ & = & $\StringAlt$\\
@@
     $\PutL{} \App \StringAlt' \App \String'$ & = & $\String'$
-  \end{tabular}
-\end{center}
+  \end{array}
+\end{array}
+\end{centermath}
+
 Just as the identity lens ignores the second argument in puts, disconnect lenses
@@
 
-\[
-  \centering
+\begin{centermath}
+\begin{tabular}[b]{l@{\qquad}l}
+$
   \inferrule*
@@
   }
-\]
-\begin{center}
+  $
+&
   \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
@@
   \end{tabular}
-\end{center}
+\end{tabular}
+\end{centermath}
 The \OrLens lens deals with data that can come in one form or another. If the
@@
 
-\[
+\begin{centermath}
+\begin{tabular}[b]{l@{\qquad}l}
+$
   \centering
@@
   }
-\]
-\begin{center}
+$
+&
   \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
@@
   \end{tabular}
-\end{center}
+\end{tabular}
+\end{centermath}
+
 The \MergeR lens is interesting because it merges data where one data can be in
@@
   \end{tabular}
-\end{center}
+\end{tabular}
+\end{centermath}
 The \MergeL lens is symmetric to \MergeR.
 
-\[
-  \centering
+\begin{centermath}
+\begin{tabular}[b]{l@{\qquad}l}
+  $
   \inferrule*
@@
   \end{tabular}
-\end{center}
+\end{tabular}
+\end{centermath}
 Composing is interesting in the put functions. Because puts require intermediary

### 2018-07-09 papers/SymmetricLenses/main.tex
Fixed some more spacing issues

--- 2018-07-09 papers/SymmetricLenses/main.tex
@@
 \PassOptionsToPackage{usenames,dvipsnames,svgnames,table}{xcolor}
-\documentclass[acmsmall,screen,anonymous]{acmart}
-\settopmatter{printacmref=false}
+\documentclass[acmsmall,review,screen,anonymous]{acmart}
+\settopmatter{printacmref=false,printccs=false}
 
@@
 lost.
-
-\[
-  \centering
-  \inferrule*
-  {
-    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt
-  }
-  {
-    \IterateLensOf{\Lens} \OfType
-    \StarOf{\BRegex}
-    \Leftrightarrow
-    \StarOf{\BRegexAlt}
-  }
-\]
 %\begin{center}
@@
 %\end{center}
-The \IterateLens lens deals with iterated data.
-
 \begin{centermath}
@@
 
-\[
-  \centering
+\begin{centermath}
+\begin{tabular}[b]{l@{\qquad}l}
+$
 \inferrule*
@@
   {
+    \IterateLensOf{\Lens} \OfType
+    \StarOf{\BRegex}
+    \Leftrightarrow
+    \StarOf{\BRegexAlt}
+  }
+  $
+  &
+  $
+  \inferrule*
+  {
+    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt
+  }
+  {
     \InvertOf{\Lens} \OfType \BRegexAlt \Leftrightarrow \BRegex
   }
-\]
+  $
+\end{tabular}
+\end{centermath}
 %\begin{center}
@@
 %\end{center}
-Inverting reverses the direction of a lens: creating on the right becomes creating on the left and vice versa, and putting on the right becomes putting on the left and vice versa.
+The \IterateLens lens deals with iterated data, while inverting reverses the direction of a lens: creating on the right becomes creating on the left and vice versa, and putting on the right becomes putting on the left and vice versa.
 The invert combonator is particularly useful when chaining many compositions together, as it can be used to align the central types.
-
 \[
@@
 \]
+
 Type equivalence enables a lens of type $S \Leftrightarrow T$ to be used as a
@@
 \begin{gather*}
-  \Call{Continue}{pq} := c >= d + \Log_2(n)
+  \Call{Continue}{pq} := c > d + \Log_2(n)
 \end{gather*}
@@
 \begin{gather*}
-  \Call{Continue}{pq} := c >= \Min(d + \Log_2(n) - t, 0)
+  \Call{Continue}{pq} := c > \Min(d + \Log_2(n) - t, 0)
 \end{gather*}

### 2018-07-09 papers/SymmetricLenses/main.tex
One more spacing fix

@@
   \includegraphics[width=.5\textwidth]{high-level-algorithm.pdf}
-  \caption{Schematic Diagram for our symmetric lens synthesis algorithm. Regular
+  \caption{Schematic diagram for our symmetric lens synthesis algorithm. Regular
     expressions \Regex and \RegexAlt, and a set of examples \Examples, are
@@
     \GreedySynth finds a lens, typed between the generated equivalent pairs.
-    When the algorithm determines it has an optimal lens, it returns that lens.}
+    When the algorithm determines it has an optimal lens, it returns
+  that lens. \ksf{Figure needs to be changed to call first search
+  Expand; Are the S and T regular expressions (which should have bars)
+  or stochastic regular expressions (which should be called such)?}}
   \label{fig:high-level-algorithm}
@@
   used by management and HR at a US company to represent employee salaries and health
-  insurance providers, respectively. \ksf{Why
-  is \lstinline{emp_salaries} written with an or instead of just using
-  the star operator? Can we simplify?}\afm{It does require the or operator, if
-  we don't want it to have trailing or preceding newlines.  Howevever, I
-  intentionally had this as not having those to force a discussion about
-  rewrites and searching through rewrites.} }  
+  insurance providers, respectively. }  
   \label{fig:minimized-representations}
@@
 
-\[
-  \centering
+\begin{centermath}
+\begin{tabular}[b]{l@{\qquad \; \qquad}l}
+$
   \inferrule*
@@
   }
-\]
+$
+&
+$
+  \inferrule*
+  {
+    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt_1\\
+    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt_2
+  }
+  {
+    \SwapLensOf{\Lens_1}{\Lens_2} \OfType \BRegex_1 \Concat \BRegex_2
+    \Leftrightarrow
+    \BRegexAlt_2 \Concat \BRegexAlt_1
+  }
+$
+\end{tabular}
+\end{centermath}
 %\begin{center}
@@
 Concat is similar to concatenation in existing string lens languages like
-Boomerang.  For such terms, we do not provide the semantics, and merely refer
-readers to existing work.
-
-\[
-  \centering
-  \inferrule*
-  {
-    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt_1\\
-    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt_2
-  }
-  {
-    \SwapLensOf{\Lens_1}{\Lens_2} \OfType \BRegex_1 \Concat \BRegex_2
-    \Leftrightarrow
-    \BRegexAlt_2 \Concat \BRegexAlt_1
-  }
-\]
+Boomerang.  For such terms, we do not provide the semantics, and merely refer readers to existing work. The swap combinator is similar to concat, though the second regular expression
+is swapped.
 %\begin{center}
@@
 %\end{center}
-The swap combinator is similar to concat, though the second regular expression
-is swapped.
-
 \begin{centermath}

### 2018-07-09 papers/SymmetricLenses/main.tex
add side stuff

--- 2018-07-09 papers/SymmetricLenses/main.tex
@@
 lost.
+
+\[
+  \centering
+  \inferrule*
+  {
+    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt
+  }
+  {
+    \IterateLensOf{\Lens} \OfType
+    \StarOf{\BRegex}
+    \Leftrightarrow
+    \StarOf{\BRegexAlt}
+  }
+\]
 %\begin{center}
@@
 %\end{center}
+The \IterateLens lens deals with iterated data.
+
 \begin{centermath}
@@
 
-\begin{centermath}
-\begin{tabular}[b]{l@{\qquad}l}
-$
+\[
+  \centering
   \inferrule*
@@
   {
-    \IterateLensOf{\Lens} \OfType
-    \StarOf{\BRegex}
-    \Leftrightarrow
-    \StarOf{\BRegexAlt}
-  }
-  $
-  &
-  $
-  \inferrule*
-  {
-    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt
-  }
-  {
     \InvertOf{\Lens} \OfType \BRegexAlt \Leftrightarrow \BRegex
   }
-  $
-\end{tabular}
-\end{centermath}
+\]
 %\begin{center}
@@
 %\end{center}
-The \IterateLens lens deals with iterated data, while inverting reverses the direction of a lens: creating on the right becomes creating on the left and vice versa, and putting on the right becomes putting on the left and vice versa.
+Inverting reverses the direction of a lens: creating on the right becomes creating on the left and vice versa, and putting on the right becomes putting on the left and vice versa.
 The invert combonator is particularly useful when chaining many compositions together, as it can be used to align the central types.
+
 \[
@@
 \]
-
 Type equivalence enables a lens of type $S \Leftrightarrow T$ to be used as a

### 2018-07-10 papers/SymmetricLenses/main.tex
fixed minor typos
in progress edits to intro

--- 2018-07-09 papers/SymmetricLenses/main.tex
@@
 \noindent
-In a big-data world, one will often find similar data stored in a variety of
-different formats. For instance, electronic calendars come in iCalendar,
+In a big-data world, similar data is often stored in a variety of
+formats. For instance, electronic calendars come in iCalendar,
 vCalendar and h-event formats~\cite{calendar-formats}; US taxation data can be
@@
 One way to interconvert between these varied data formats is to use a
-\emph{lens}~\cite{?}---a bi-directional
+\emph{lens}~\cite{Focal2005}---a bi-directional
 program that can transform data represented in a format $S$ to a format $T$,
-and vice versa.  There are a wide variety of different kinds of lenses, but well-defined lens
-languages typically come with guarantees that help ensure data is
+and vice versa.  There are many different kinds of lenses, but well-defined lens
+languages typically provide guarantees that help ensure data is
 not lost or corrupted as it is transformed back and forth between different representations.
 
-Lens languages, like any domain-specific language, also cut down the space of
-possible programs one can write, and the smaller the space of programs,
-the easier it is to find the program one is looking for.  
+While lenses facilitate writing principled data
+transformations, \textit{synthesizing} lenses 
+can make the task even easier.  Lens languages are a ripe target for
+synthesis because lens languages, like many domain-specific languages,
+reduce the space of possible programs, making search-based synthesis easier. 
 %This feature can be a negative, if the desired transformation
@@
 Recent research~\cite{optician, maina+:quotient-synthesis} has exploited this
-fact by demonstrating that it is possible to synthesize relatively large,
+observation by demonstrating that it is possible to synthesize relatively large,
 complex lenses from specifications. In particular, given types that describe
-source and target formats, the bijective synthesis algorithm presented in this
-research will find bijective lenses
+source and target formats and optional examples of corresponding data, the
+bijective synthesis algorithm presented in these 
+papers will find bijective lenses to convert between data
 % ~\cite{optician}
-for files in ad-hoc formats, such as those housing linux application
-configurations. Oftentimes, these algorithms find such lenses in less than a
-second, and without the aid of any examples at all, with just types describing
-the source and target formats.
+in ad hoc formats, such as linux 
+configuration files that represent the same data in different ways. 
+Often, these algorithms can find an appropriate lens
+in less than a second without requiring any examples.
+%, with just types describing the source and target formats.
 % \dpw{I would cite the ICFP paper (cite string:
@@
 %   third person, even if that's a bit awkward, but otherwise citing it is OK.}
-
-One of reasons these algorithms are so effective, even relative to successful
-synthesis projects in other domains, is because unless the same kind of data
-appears in a format many times, there simply are not very many different
-bijections between related data formats. Hence, if the synthesis algorithm can
-find any bijection at all, then it is very likely to be the one the user wants.
-However, there is also a price for narrowing the synthesis domain so
-dramatically: Many real data formats are almost, but not quite, bijections. More
-often two related data formats have some data in common, but also contain some
-differences. For instance, two related tables might share four of five columns,
-but not the last. Alternately, one ad hoc configuration file might host some
-format-specific meta data, such as a date, a time or reference number, whereas
-the other does not. Indeed, all of our benchmarks taken from Flash
-Fill~\cite{flashfill} and all of our benchmarks that synchronize between two
-ad-hoc file formats. Consequently, past tools that search for strict bijections
-will rarely be applicable in practice.
+One reason that bijective lens synthesis is so effective, even relative to successful
+synthesis projects in other domains, is that there are simply not very
+many different bijections between related data formats (unless the
+same kind of data appears in the two formats many times). 
+Hence, if the synthesis algorithm can
+find any bijection at all, then it is very likely to be the one the
+user wants.
+
+However, there is a price for narrowing the synthesis domain so
+dramatically: Many real data formats are almost, but not quite,
+bijections. More often two related data formats have some data in
+common, but also contain some differences. For instance, two related
+tables might share four of five columns, but each have a different
+fifth column. Alternately, one ad hoc configuration file might include
+some file-specific meta data, such as a date, a time or reference
+number, whereas the other does not. Indeed, all of our benchmarks
+taken from Flash Fill~\cite{flashfill} and all of our benchmarks that
+synchronize between two ad hoc file formats described in
+Section \ref{sec:evaluation} \ksf{is this right?} have this 
+characteristic. Consequently, past tools that search for strict
+bijections will rarely be applicable in practice.
 

### 2018-07-10 papers/SymmetricLenses/main.tex
add graphs and analysis

--- 2018-07-09 papers/SymmetricLenses/main.tex
@@
 We now provide the typing derivations and semantics for these lenses. For
-simplicity of presentation,\bcp{Is there any known issue with relaxing this
-  restriction? I've often found ti convenient to use id lenses with ambiguous
-  REs...}\afm{no known issues, I just didn't want to go into specifics about
-  ambiguity in an already-full paper}\bcp{OK, we can just comment that we
-  believe relaxing this restriction is unproblematic.}, we define the typing derivations and
-semantics only over unambiguous regular expressions, even though some lenses can
-be defined over ambiguous regular expressions (for example, the identity lens
-doesn't need any ambiguity constraints).
+simplicity of presentation, we define the typing derivations and semantics only
+over unambiguous regular expressions, even though some lenses can be defined
+over ambiguous regular expressions (for example, the identity lens doesn't need
+any ambiguity constraints). We believe relaxing this restriction is
+unproblematic.
 
@@
     benchmarks in our benchmark suite in under 30 seconds, and is able to
-    complete most of the benchmarks without using compositional synthesis.
-    \bcp{The interesting part of the graph is all squished against the Y axis.
-      Can we make the X axis logarithmic?}}
+    complete a majority of the benchmarks without using compositional
+    synthesis.}
   \label{fig:times}
@@
 algorithm is able to synthesize all of the benchmarks in our benchmark suite in
-under 30 seconds.
+under 30 seconds. Without compositional synthesis, the synthesis algorithm is
+able to solve 26 problem instances.
 
@@
 timeout of 30 seconds, and average the result over 5 runs.
-Figure~\ref{fig:comparison-times} summarizes the results of these tests. On
+Figure~\ref{fig:times_bijective} summarizes the results of these tests. On
 average, \SSOpt{} took 2.1 times (\afm{3.5} seconds) longer to complete than
 \BSOpt{}.
+
+\begin{figure}
+  \includegraphics{generated-graphs/times_bijective}
+  \caption{Number of benchmarks that can be solved by a given algorithm in a
+    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
+    \BSOpt{} is the full bijective lens synthesis algorithm. \SSNCOpt{} is the
+    symmetric synthesis algorithm without using a library of existing lenses.
+    \BSNCOpt{} is the bijective synthesis algorithm without using a library of
+    existing lenses. Our symmetric synthesis algorithm is able to complete all
+    benchmarks in our benchmark suite in under 30 seconds, and is able to
+    complete most of the benchmarks without using compositional synthesis.}
+  \label{fig:times_bijective}
+\end{figure}
 
@@
 \begin{tabulary}{\linewidth}{rL}
-  \NoTPOpt{}: & We synthesize lenses using our symmetric synthesis algorithm,
-                but users cannot tune when \RXSearch returns using a termination parameter.\\
-  \FCOpt{}: &  The first time \GreedySynth finds a lens, it returns that lens.\\
-  \NMCCOpt{}: & The lens cluster that contains is identified through \SSOpt{}. \GreedySynth is then called on that pair of SREs. \\
-  \NMOpt{}: & Neither \GreedySynth nor \RXSearch have any notion of cost; they
-              both merely search for
-              any lens that satisfies the specification. 
+  \AnyOpt{}: & We synthesize any lens that satisfies the specification, with no
+               notion of cost.\\
+  \FLOpt{}: &  The first time \GreedySynth finds a lens, the synthesis procedure
+              returns that lens.\\
+  \CCOpt{}: & We replace our information theoretic cost metric with one where
+              the cost of the lens is the number of disonnects plus the number
+              of merges.\\
+  \ZTOpt{}: & \ZTOpt{} runs the algorithm with the termination parameter
+    universally set to 0 (default). \\
+  \TFTOpt{}: & \TFTOpt{} runs the algorithm with the
+    termination parameter universally set to 25. \\
+  \NTFTOpt{}: & \NTFTOpt{} runs the algorithm with the
+    termination parameter universally set to -25. \\
 \end{tabulary}
@@
   \caption{Number of benchmarks that synthesize the correct lens by a given
-    algorithm. \NoTPOpt{} is the full symmetric synthesis algorithm, but
-    termination parameters are ignored. \FCOpt{} provides a notion of cost to
-    \GreedySynth, but once a satisfying lens is greedily found, that lens is
-    returned. \NMCCOpt{} identifies the lens cluster in which the desired 
-    lens lies.  \GreedySynth{} is then run on that cluster with no notion of
-    cost. \NMOpt{} removes all notion of cost, and merely finds the first
-    satisfying lens.\bcp{Am I reading this wrong, or does even the best
-      algorithm fail to find the correct lens something like 20\% of the
-      time?!?!?!}} 
+    algorithm. \AnyOpt{} provides no notion of cost, and merely returns the
+    first lens it finds that satisfies the specification. \FLOpt{} provides a
+    notion of cost to \GreedySynth, but once a satisfying lens is greedily
+    found, that lens is returned. \CCOpt{} synthesizes lenses, where the cost of
+    a lens is the number of disconnects plus the number of merges. \ZTOpt{} runs
+    the algorithm with the termination parameter universally set to 0 (default).
+    \TFTOpt{} runs the algorithm with the termination parameter universally set
+    to 25. \NTFTOpt{} runs the algorithm with the termination parameter
+    universally set to -25.}
   \label{fig:metric}
@@
 information-theoretic metric is critical for finding the correct lens. Indeed,
-even when provided with the correct lens cluster, without a notion of cost, we
-would need constrain our search space more with additional examples to find the
-correct lens.
-
-Our termination condition is also quite important. When running in \FCOpt{},
-only 5 lenses are found. The first cluster in which a lens is found is rarely
-the correct cluster. However, our termination condition is not perfect.
+only \afm{10} of our benchmarks succeeded when running in \CCOpt{} mode.
+
+Our termination condition is also quite important. When running in \FLOpt{},
+only \afm{5} lenses are found. The first cluster in which a lens is found is
+rarely the correct cluster. However, our termination condition is not perfect.
 Sometimes users must tell the algorithm to ``keep going'' or ``stop early.''
@@
 our 45 benchmarks, 9 of them need help in the form of a termination parameter
-for finding the correct lens.
+for finding the correct lens. That being said, our default parameter finds a
+good middle-ground. Both \TFTOpt{} and \NTFTOpt{} synthesized the correct lens
+less often than \ZTOpt.  \TFTOpt{} necessitates too good a lens, so the search
+procedure keeps spinning, and \NTFTOpt{} terminates early, not finding the
+correct lens.
 % end evaluation

### 2018-07-10 papers/SymmetricLenses/transform-data.py

@@ -53,28 +53,32 @@ def write_to_filename(filename, s):
 
 def generate_examples_required_graph(input_csv):
     nm_count = count_real_vals(input_csv,"NM")
-    nmcc_count = count_real_vals(input_csv,"NMCC")
     fc_count = count_real_vals(input_csv,"FC")
     notp_count = count_real_vals(input_csv,"NoTP")
+    tc25_count = count_real_vals(input_csv,"TC25")
+    tcn25_count = count_real_vals(input_csv,"TCN25")
+    constcostcc_count = count_real_vals(input_csv,"ConstCostCC")
     print(nm_count)
-    print(nmcc_count)
     print(fc_count)
+    print(tcn25_count)
     print(notp_count)
+    print(tc25_count)
+    print(constcostcc_count)
 
-    ind = np.arange(4)
+    ind = np.arange(6)
     width = 0.35
 
     fig, ax = plt.subplots()
 
-    rects1 = ax.bar(ind, [nm_count,nmcc_count,fc_count,notp_count], width, color='#ffffb3', align='center')
+    rects1 = ax.bar(ind, [nm_count,fc_count,constcostcc_count,notp_count,tc25_count,tcn25_count], width, color='#ffffb3', align='center')
 
     ax.set_ylabel('Benchmarks\nCompleted')
     ax.set_xlabel('Run Mode')
     ax.set_xticks(ind)
-    ax.set_xticklabels(["\\textbf{NM}","\\textbf{NMCC}","\\textbf{FC}","\\textbf{NoTP}"])
+    ax.set_xticklabels(["\\textbf{Any}","\\textbf{FL}","\\textbf{DC}","\\textbf{t=0}","\\textbf{t=25}","\\textbf{t=-25}"])
 
-    fig = plt.figure(2,tight_layout=True)
-    ax.step([-.5,3.5],[45.1,45.1],label="Benchmark Count",linestyle=":",
+    fig = plt.figure(3,tight_layout=True)
+    ax.step([-.5,5.5],[45.1,45.1],label="Benchmark Count",linestyle=":",
             linewidth=1, dashes=(1,1))
 
     plt.tick_params(
@@ -83,7 +87,7 @@ def generate_examples_required_graph(input_csv):
                         bottom='off',      # ticks along the bottom edge are off
                             top='off') # labels along the bottom edge are off
     fig.set_figheight(1)
-    fig.set_figwidth(3)
+    fig.set_figwidth(4)
 
[…]

### 2018-07-10 papers/SymmetricLenses/main.tex
a pass through section 5

--- 2018-07-10 papers/SymmetricLenses/main.tex
@@
 %$\PCF{Expand}$, (4) $\PCF{\PQ.Push}$, and (5) $\PCF{\Continue}$. First,
-This pair of SREs are used to initialize a priority queue ($pq$).
+This pair of SREs is used to initialize a priority queue ($pq$).
 Next, $\PCF{SynthSymLens}$
@@
 \item pops the next class ($S$, $T$) of lenses to
-search of the priority queue ($\PCF{\PQ.Pop}$),
+search off of the priority queue ($\PCF{\PQ.Pop}$),
 \item executes $\GreedySynth$ to find a best lens in that class if
@@
 \item adds all of the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)
-  to the priority queue with $\PCF{\PQ.Push}$
+  to the priority queue with $\PCF{\PQ.Push}$ \saz{what determines the priority order?}
 \end{itemize}
@@
 The first synthesizer is one that searches through regular expression pairs.
-This synthesizer, comprising \Expand, \PQ, and \RXSearch, proposes regular
-expressions, semantically but not syntactically equivalent to the originals. In
+This synthesizer, comprising \Expand, \PQ, and \RXSearch \saz{The name \RXSearch
+  does not appear in the figure.  I get the impression that it is the name of
+  the while loop, but perhaps we can make that clearer in the algorithm code?
+  Maybe by putting a brace around the while loop and giving it a name? Here we
+  could say something like ``... and a loop that continues searching until a
+  satisfactory lens is found (we call this loop \RXSearch)}.  It proposes regular
+expressions that are semantically, but not syntactically, equivalent to the originals. In
 particular, this synthesizer performs the following rewrites:
@@
                                             \{\String_1\Concat\ldots\Concat\String_n
-                                            \SuchThat \forall i \String_i\in\LanguageOf{\DNFRegex}\}\\
+                                            \SuchThat \forall i,  \String_i\in\LanguageOf{\DNFRegex}\}\\
     \LanguageOf{\SequenceOf{\String_0\SeqSep\Atom_1\SeqSep\ldots\SeqSep\Atom_n\SeqSep\String_n}}\ =\  & 
@@
 \]
-We have proven this conversion respects languages and probability
+We have proved that this conversion respects languages and probability
 distributions.
@@
 iterating. In our experience generating random strings from regular expressions,
-these probabilities provide good distributions of strings -- stars are iterated
+these probabilities provide good distributions of strings---stars are iterated
 4 times on average, and no individual choice in a series disjunctions is chosen
@@
 
+\saz{The following two paragraphs are pretty similar (was there a bad merge at
+  some point?) and very low level.  I also
+  could not really follow either of the examples at all.  If we
+  are looking for things to cut, I would propose dropping most of the content of
+  the next two paragraphs, merging the remainder, and skipping pretty directly to the paragraph
+  that starts with ``While DNF lenses...''}
 As an example, consider the symmetric DNF lens:
@@
 lens that maps it to sequence $\SequenceAlt_{c_i}$ on the right. The third
-component is a list of integers $d$, and follows the same pattern as the second
+component is a list of integers $d$ that follows the same pattern as the second
 component, but maps from right to left.
@@
 \label{subsec:greedy-synth}
-The greedy algorithm is actually two greedy algorithms, one that greedily finds
+\saz{We should call out the structural similarity
+  between the DNF grammar and the organization of these algorithms.}
+The greedy algorithm is actually two\saz{three! don't forget AtomSynth} greedy algorithms, one that greedily finds
 symmetric DNF lenses (\GreedySynth) and one that greedily finds symmetric
@@
 \GreedySynth, a means to synthesize symmetric DNF lenses. The inputs to
-\GreedySynth are a pair of stochastic DNF regular expressions and a suite of
-input-output examples.
-
-First, \PCF{CannotMap} checks if there could possibly be a lens satisfying the
-examples. If there is not, \GreedySynth returns \None early. After this,
-\GreedySynth finds the best lenses between each sequence pair of the left and
-right DNF regular expressions, given the examples that match them. A priority
-queue containing these sequence lenses is then initialized with
-$\PCF{PQ.Create}$. The symmetric lens is built up iteratively from these
-sequence lenses, the state of this partial lens is tracked in the lens builder,
-$lb$.
-
-\GreedySynth then loops until there are no more sequence lenses in the priority
-queue. Within this loop, sequence lenses are popped. If this sequence lens is
-``useful'', then that sequence lens is included in the final DNF lens. If the
-source or target of a popped sequence lens is not the source or target of
-another sequence lens, then this sequence lens is considered useful. If the
-examples require two sequences have a lens between them, then any sequence lens
-between them is considered useful. We update the priorities of our sequence
-lenses as the algorithm proceeds: if two sequence lenses have the same source,
-the second one to be popped gets a higher cost than it originally had, to
-address the information that needs to be stored for including that source of
-non-bijectivity.
-
-In our example, \GreedySynth may first pop the sequence pair \lstinline{""} and
-\lstinline{""}, as there is a bijective sequence lens between them. As neither
+\GreedySynth are a suite of input-output examples and a pair of stochastic DNF regular expressions.
+
+First, \PCF{CannotMap} checks whether there is no possible lens satisfying the
+examples. If so, \GreedySynth returns \None early. Otherwise, \GreedySynth finds
+the best lenses (given the examples that match them) between all sequence pairs
+$(SQ_i, TQ_j)$ drawn from the left and right DNF regular expressions. (The
+function \CartesianMap maps its argument across the cross product of the input
+lists).  A priority queue containing these sequence lenses is then initialized
+with $\PCF{PQ.Create}$.\saz{Ordered how?} The symmetric lens is then built up
+iteratively from these sequence lenses, where the state of the partially constructed lens is
+tracked in the lens builder, $lb$.
+
+\GreedySynth loops until there are no more sequence lenses in the priority
+queue. Within this loop, a sequence lense is popped from the queue and, if it is
+``useful'', then it is included in the final DNF lens. The lens is considered to
+be useful when its source (or target) is not already the source (or target) of
+an already chosen sequence lens. If the examples require that two sequences have a
+lens between them, then any sequence lens between them is considered useful. We
+update the priorities of the sequence lenses as the algorithm proceeds: if two
+sequence lenses have the same source, the second one to be popped gets a higher
+cost than it originally had, to account for the information that needs to be stored
+for including that source of non-bijectivity.
+
+In our example, \GreedySynth might first pop the sequence pair \lstinline{""} and
+\lstinline{""}, because there is a bijective sequence lens between them. As neither
 \lstinline{""} is involved in a sequence lens, this lens is considered useful.
-After this, the sequence lens between \lstinline{name.name*} and
+Next, the sequence lens between \lstinline{name.name*} and
 \lstinline{name} would be popped: while that lens is not bijective: it is still
@@
 \paragraph*{Symmetric Sequence lenses} Algorithm~\ref{alg:greedyseqsynth}
-presents \GreedySeqSynth, a means to synthesizes symmetric sequence lenses.
-
-This algorithm uses \PCF{AtomSynth}, a simple function that synthesizes atom
-lenses by iterating the DNF lens between its subcomponents. The inputs to
-\GreedySeqSynth are a pair of stochastic atoms and a suite of input-output
-examples. First, \PCF{CannotMap} checks if there is a possibility of a lens
-between the given sequences regular expressions that is consistent with the
-examples. If there is not, \GreedySeqSynth returns \None early. After this,
+presents \GreedySeqSynth, which synthesizes symmetric sequence lenses using an
+algorithm whose structure is similar to \GreedySynth{}'s.  It calls \PCF{AtomSynth}, a
+function that synthesizes atom lenses by iterating a DNF lens between
+its subcomponents.
+
+The inputs to \GreedySeqSynth are a suite of input-output
+examples and a pair of stochastic atoms. As in \GreedySynth, \GreedySeqSynth
+returns \None early if there is no possible lens. Afterward,
 \GreedySeqSynth finds the best lenses between each atom pair of the left and
-right sequences. A priority queue containing these lenses is then initialized
+right sequences, and organizes them into a priority queue 
 with \PCF{PQ.Create}. The symmetric sequence lens is built up iteratively from
-these atom lenses, the state of the partial lens is tracked in the sequence lens
+these atom lenses, where the state of the partially built lens is tracked in the sequence lens
 builder, $slb$.
@@
 \GreedySeqSynth then loops until there are no more atom lenses in the priority
-queue. Within this loop, atom lenses are popped. The atom lens is then added if
-it is useful. An atom lens is useful if adding it will lower the cost of the
+queue. Within this loop, atom lenses are popped from the queue and the
+``useful'' ones are kept.  An atom lens is useful if adding it will lower the cost of the
 generated sequence lens, or if examples show that one of the atoms must be
-connected to another lens. The priorities of these lenses change over time, as
+connected to another lens.\saz{I'm not sure what ``must be connected to another
+  lens means''}  The priorities of these lenses change over time, because
 each atom can only be involved in one lens. Adding a lens may necessitate
 removing another, so not all popped atom lenses will lower the cost of the
-generated sequence lens.
+generated sequence lens.\saz{I don't understand the last two sentences.}
 
@@
 
-We included a number of optimizations in our algorithm. We added ``open'' and
-``closed'' annotations to regular expressions, an expansion inference algorithm,
-hash-consing and memoization, and compositional synthesis to our algorithm. We
-also enable an additional parameter in our specifications that customizes the
-termination condition for our procedure.
-
-\paragraph*{Open and Closed Regular Expressions} While GreedySynth acts
+Our implementation includes a number of optimizations to this algorithm.  We added ``open'' and
+``closed'' annotations to regular expressions \saz{what do these mean? I guess
+  ``open'' means ``subject to  DNF conversion'' and ``closed'' means not?}, an expansion inference algorithm,
+hash-consing and memoization, along with compositional synthesis. The implementation
+also allows an additional parameter in specifications that customizes the
+termination condition for the synthesis procedure.
+
+\paragraph*{Open and Closed Regular Expressions} While \GreedySynth acts
 relatively efficiently, it suffers from an exponential blowup when converting
@@
 \emph{all} of the regular expression into DNF form, but only converting
-\emph{some} of it to DNF form. We add in an additional atom, where an atom can
-be an either an iteration of a DNF regular expression or it can be a regular
-expression. When converted into a DNF regular expression, if $\Regex$ is a
-stochastic regular expression, annotated as closed, then $\ToDNFRegexOf{\Regex}
-= \AtomToDNFOf{\Regex}$.
-
-At the start of synthesis, we annotate our regular expressions as closed. Now,
-we enhance our rewrites by changing an annotation from closed to open. This
-forces changes to \RXSearch and \GreedySynth. \RXSearch now can rewrite closed
-regular expressions to open ones. \GreedySynth is provided with the means to map
-closed regular expressions to each other. Two closed regular expressions can map
-to each other only through the identity lens: two regular expressions can map to
-each other by the identity lens, if they are the same closed regular expression
-and they parse the same string components in all examples.
+\emph{some} of it to DNF form, and doing so incrementally.  This change is easy
+to make by adding non-DNF stochastic regular expressions as a new type of
+``closed'' atom.  When converted into a DNF regular expression, if $\Regex$ is
+such a closed stochastic regular expression, then
+$\ToDNFRegexOf{\Regex} = \AtomToDNFOf{\Regex}$.
+
+At the start of synthesis, all regular expressions are closed. \RXSearch is
+modified so that it is now permitted to rewrite closed regular expressions to
+open ones (thereby triggering DNF normalization). \GreedySynth is allowed to to
+map closed regular expressions to each other, but only using the identity
+lens---they must parse the same string components in all examples.
 
 \paragraph*{Expansion Inference} The prior optimization helped \GreedySynth, but
-it made \RXSearch have a much harder job.  To alleviate this difficulty, we
-added in expansion inference.  Expansion inference identifies when a regular
-expression does not have a chance to map through any means other than
-disconnected.  Expansion inference opens those regular expressions.
+it made \RXSearch have the harder job of determining when to open closed
+expressions.  To alleviate this difficulty, we added in expansion inference,
+which identifies when a regular expression cannot possibly map by any means
+other than the disconnect lens.  Expansion inference opens those regular
+expressions to allow for further exploration.
 
@@
 \paragraph*{Compositional Synthesis} Compositional synthesis allows us to use
-previously defined lenses - whether they are previously defined by users or by
+previously defined lenses---whether they are previously defined by users or by
 synthesis. We alter \GreedySynth such that two regular expression atoms can be

### 2018-07-10 papers/SymmetricLenses/main.tex
some comments / changes to section 4

@@
 \label{sec:ssl}
+\saz{This paragraph used to say that lenses are typed between stochastic REs
+  and it also says that lenses are typed between REs.  I think its should just
+  be REs? Double check that this is correct}
+
 We now give a formal description of all the combinators in our symmetric lens
-language. These lens terms are typed between pairs of stochastic regular
-expressions: the typing judgment $\Lens \OfType \BRegex \Leftrightarrow
-\BRegexAlt$ means that $\Lens$ is a well typed lens between
-$\LanguageOf{\BRegex}$ and $\LanguageOf{\BRegexAlt}$. While we are performing
-synthesis on stochastic regular expressions, lenses are typed between regular
-expressions.  We can recover a regular expression from a stochastic regular
-expression by deleting probability annotations.  In a slight abuse of notation,
-if $\Regex$ is a stochastic regular expression, $\BRegex$ is the regular
-expression obtained by removing probability annotations.
+language.  While our algorithms perform synthesis on stochastic regular
+expressions, lenses themselves are typed between regular expressions.  We can
+recover a regular expression from a stochastic regular expression by deleting
+probability annotations.  In a slight abuse of notation\saz{Why is this an
+  abuse?}, if $\Regex$ is a stochastic regular expression, $\BRegex$ is the
+regular expression obtained by removing probability annotations. The typing
+judgment $\Lens \OfType \BRegex \Leftrightarrow \BRegexAlt$ means that $\Lens$
+is a well typed lens between $\LanguageOf{\BRegex}$ and
+$\LanguageOf{\BRegexAlt}$.
 
@@
 functions bound the expected number of bits to recover one data format from a
-synchronized string in the other format. Notice, we would be able to exactly
+synchronized string in the other format. Note that we would be able to exactly
 calculate the conditional entropy, were it not for \MergeL. If
@@
 \PRegexOr{\RegexAlt_1}{\RegexAlt_2}{\ProbabilityAlt}$, given a string in $s$, we
-need to determine if the synchronized string is in $\Regex$ or $\RegexAlt$.
+need to determine if the synchronized string is in $\RegexAlt_1$ or $\RegexAlt_2$.
 However, this information content is dependent on the how likely the
-synchronized string is to be in $\Regex$ or $\RegexAlt$. Furthermore, we
+synchronized string is to be in $\RegexAlt_1$ or $\RegexAlt_2$. Nevertheless, we
 typically calculate the conditional entropy exactly, as merges are relatively
@@
   Let $\Lens \OfType \Regex \Leftrightarrow \RegexAlt$, where $\Lens$ does not
-  include composition, $\Regex$ and $\RegexAlt$ and unambiguous, and neither
+  include composition, $\Regex$ and $\RegexAlt$ are unambiguous, and neither
   $\Regex$ nor $\RegexAlt$ contain any empty subcomponents.
@@
 
-Note we have not defined entropy for lens compositions. Consider the lens
-$\Lens_1 ; \Lens_2$, where both $\Lens_1$ and $\Lens_2$ are non-bijective.
-However, their composition could actually be bijective, for example if the
-intermediary format contains information not present in either the source nor
-the target. While this composition may be a bijection (and should have zero
-cost), it is difficult to see this looking merely at the syntax.
-
-Notice that our definition of $\Entropy$ contains no case for sequential
+% SAZ: This looks like a bad merge, since the paragraph below is better
+% Note we have not defined entropy for lens compositions. Consider the lens
+% $\Lens_1 ; \Lens_2$, where both $\Lens_1$ and $\Lens_2$ are non-bijective.
+% However, their composition could actually be bijective, for example if the
+% intermediary format contains information not present in either the source nor
+% the target. While this composition may be a bijection (and should have zero
+% cost), it is difficult to see this looking merely at the syntax.
+
+Note that our definition of $\Entropy$ contains no case for sequential
 composition $\Lens_1 ; \Lens_2$ and our theorem excludes lenses that contain
@@
 information that is subsequently projected away in $\Lens_2$. Such operations
-can cancel, leaving us with a zero-entropy bijection. However, detecting such
-cancellations directly is complicated. Fortunately, we are able to avoid such
-complications by synthesizing \emph{DNF lenses} — simple symmetric lenses that
-inhabit a disjunctive normal form. DNF lenses have a number of properties that
-facilitate synthesis: they cut down the program search space dramatically and
-they do not contain instances of composition.
+can cancel, leaving a zero-entropy bijection composed from two non-zero entropy
+transformations. However, detecting such cancellations directly is complicated and
+this property is difficult to determine merely from syntax. Fortunately, we are able to
+avoid such complication altogether by synthesizing \emph{DNF lenses}---simple symmetric
+lenses that inhabit a disjunctive normal form. DNF lenses have a number of
+properties that facilitate synthesis: they cut down the program search space
+dramatically and they do not contain instances of composition.
 

### 2018-07-10 papers/SymmetricLenses/main.tex
a few tweaks to section 2

@@
 updated, the other format gets the exact same value. In our example,
-when a \lstinline{name} is updated in one file, the identify lens ensures it is
+when a \lstinline{name} is updated in one file, the identity lens ensures it is
 also updated in the other file. In this manner, both the \CreateR and \CreateL
@@
 to another, equivalent, regular expression type. Our typing rules include type
-equivalence, so if a lens has type $\Regex \Leftrightarrow \RegexAlt$, then it
-is also has type $\Regex' \Leftrightarrow \RegexAlt'$, where $\Regex$ is
-equivalent to $\Regex'$, and $\RegexAlt$ is equivalent to $\RegexAlt'$. The
+equivalence, so if a lens has type $\BRegex \Leftrightarrow \BRegexAlt$, then it
+is also has type $\BRegex' \Leftrightarrow \BRegexAlt'$, where $\BRegex$ is
+equivalent to $\BRegex'$, and $\BRegexAlt$ is equivalent to $\BRegexAlt'$. The
 flexibility this type equivalence rule provides is useful when a different
@@
 The expected number of bits required to recover data is a well-known concept in
-information theory: it is the \emph{entropy} of the data. Calculating the
+information theory: it is the \emph{entropy} of the data. \saz{Cite Shannon?} Calculating the
 entropy of data requires a probability distribution over the space of possible
@@
 provides little information about what string is present in \RegexAlt. With this
-metric, bijections have zero cost -- given a string $s \in \Regex$, we know
+metric, bijections have zero cost---given a string $s \in \Regex$, we know
 exactly what string is present in $\RegexAlt$, as exactly one string is

### 2018-07-10 papers/SymmetricLenses/main.tex
in progress edits to intro

@@
 bijective synthesis algorithm presented in these 
-papers will find bijective lenses to convert between data
+papers will find bijective lenses to convert between data stored
 % ~\cite{optician}
-in ad hoc formats, such as linux 
-configuration files that represent the same data in different ways. 
+in different formats,
+such as linux configuration files from different systems that represent the same data in different ways.  
 Often, these algorithms can find an appropriate lens
@@
 One reason that bijective lens synthesis is so effective, even relative to successful
-synthesis projects in other domains, is that there are simply not very
-many different bijections between related data formats (unless the
+synthesis projects in other domains, is that there are not very
+many bijections between related data formats (unless the
 same kind of data appears in the two formats many times). 
@@
 A natural way to expand the scope of such tools is to consider a more
-powerful language of lenses.  In particular, \emph{symmetric
+exrpressive language of lenses.  For example, \emph{symmetric
 lenses}~\cite{symmetric-lenses} encompass a broad collection of bidirectional
-transformations and include bijections as well as transformations that
-allow each side of the transaction to retain information not present
-in the other side.  If $\ell$ is a symmetric lens between formats $S$ and $T$, one can
-use it to transform data from $S$ to $T$, or vice versa equally well.  A user would typically
-run $\ell$ in the forwards direction (from $S$ to $T$) on a string $s$ when a user updates
-$s$---executing $\ell$ in this direction will propagate those updates to $T$.
-However, because data in format $T$ may contain some information not present in format $T$,
-the $S \rightarrow T$ transformation will take
-both a string $s \in S$ and a string $t \in T$ as arguments.
-Inuitively, the lens should extract the data ``shared'' between the two formats from $s$
-as it generates an output $t'$.  If $s$ is missing information, the transformation should
-extract the missing components from $t$.
+transformations that include bijections as well as more flexibile transformations that
+allow each side to retain information not present
+in the other.  If $\ell$ is a symmetric lens between formats $S$ and
+$T$, then $\ell$  can transform data from $S$ to $T$ or vice versa
+equally well.  A user would typically 
+run $\ell$ in the forwards direction (from $S$ to $T$) when a user updates
+a string $s \in S$ to propagate those updates to $T$.
+However, because data in format $S$ may contain some information not present in format $T$,
+the $S \rightarrow T$ transformation takes
+\emph{both} the edited string $s \in S$ and a string $t \in T$ as arguments.
+Inuitively, the lens extracts the data ``shared'' between the two
+formats from $s$
+and uses $t$ to supply any missing information.
+% required by $T$ that is missing
+%as it generates an output $t'$.  If $s$ is missing information, the transformation should
+%extract the missing components from $t$.
 
 Initially, we hypothesized that extending the synthesis algorithms
-of Miltner \ETAL~\cite{optician} to operate over symmetric lenses
-as opposed to bijective lenses 
-would be relatively straightforward:  Simply replace one set of combinators with another
-and execute a search with similar heuristics.  However, in our early,
-naive attempts to do so, we ran
-into two different kinds of difficulties. \dpw{I got into the weeds a little bit there.
+of Miltner \ETAL~\cite{optician} to operate over symmetric rather than
+bijective lenses
+would be relatively straightforward:  Simply replace the bijective
+combinators with symmetric ones 
+and search using similar heuristics.  However, 
+this naive approach runs into two difficulties. \dpw{I got into the
+weeds a little bit there. 
 The goal was to set up the idea that there are ``two parts'' to a symmetric

### 2018-07-10 papers/SymmetricLenses/main.tex

@@
 
-(The idea of lenses.  Their usefulness and popularity.)
-
-Recent work by~\citet{optician} has also demonstrated
-how techniques from {\em type-directed program synthesis} can be used to
-efficiently synthesize the simplest class of lenses: so-called {\em
-  bijective lenses}, where no information is discarded when transforming
-data in either direction.
-
-We show here how to extend this algorithm to a much richer class of lenses,
-which we call {\em simple symmetric lenses}; this class includes all
-bijective and standard ``asymmetric'' lenses~\cite{OriginalFosterPOPLpaper},
-as well as a rich subset of the full-blown ``symmetric lenses'' studied by
-\citet{symmetric-lenses}.
-
-(why our algorithm is harder; how we use information theory to help)
-
+A {\em lens} is a program that can be run both ``front to back'' and ``back
+to front,'' allowing data to be transformed in two directions.  Since their
+introduction by~\citet{Focal2005-long2}, lenses have been extensively
+studied and applied.  Recent work~\cite{optician} has also demonstrated how
+techniques from {\em type-directed program synthesis} can be used to
+efficiently synthesize a very simple class of lenses---so-called {\em
+  bijective lenses} over string data---given a pair of types (regular
+expressions) and a small number of examples.
+
+We show how to extend this synthesis algorithm to a much richer class of
+lenses, which we call {\em simple symmetric lenses}, including both
+bijective lenses and the more widely used ``asymmetric''
+lenses~\cite{Focal2005-long2}, as well as a rich subset of the full-blown
+``symmetric lenses'' proposed by \citet{symmetric-lenses}.  Simple symmetric
+lenses are of independent theoretical interest, being the largest class of
+symmetric lenses that do not rely on persistent internal state.
+
+Synthesizing simple symmetric lenses is substantially more challenging that
+synthesizing bijective lenses: since some of the information on each side
+can be ``disconnected'' from the other side, there will in general be {\em
+  many} lenses that agree with a given example.  To guide the search
+process, we use {\em stochastic regular expressions} and ideas from
+information theory to estimate the amount of information propagated by a
+candidate lens.
+
+We describe an implementation of simple symmetric lenses and our synthesis
+procedure as extensions to the Boomerang language~\cite{Boomerang} and
+evaluate its performance on a range of benchmark examples.
 \end{abstract}

### 2018-07-10 papers/SymmetricLenses/main.tex
add data to abstract
Merged David's comments and suggestions

@@
 We describe an implementation of simple symmetric lenses and our synthesis
-procedure as extensions to the Boomerang language~\cite{Boomerang} and
-evaluate its performance on a range of benchmark examples.
+procedure as extensions to the Boomerang language~\cite{Boomerang} and evaluate
+its performance on a range of benchmark examples. We find this implementation is
+able to synthesize all the lenses of this benchmark suite in under 30 seconds.
 \end{abstract}
@@
 and search using similar heuristics.  However, 
-this naive approach runs into two difficulties. \dpw{I got into the
-weeds a little bit there. 
-The goal was to set up the idea that there are ``two parts'' to a symmetric
-lens transformation: the information from $s$ and the information from $t$.  In
-a future paragraph in the intro we are going to explain that the central heuristic
-involves minimizing the information from $t$.}
-
-The first difficulty was pragmatic: symmetric lenses actually usually operate
-over three structures, the source $S$, the target $T$ and a complement
-$C$~\cite{symmetric-lenses}. The latter object contains all the information not
-present in one of $S$ or $T$. Logistically, the complement object must be stored
-away somewhere and infrastructure must be developed to manage it. Perhaps more
-relevant though, is that complements complicate specifications, even when those
-specifications are just examples. The issues is that one cannot just give
-examples of source and target; one must also give guidance on how to manage the
-``hidden state.'' In order to avoid such complexities, we defined a new variant
-of symmetric lenses, which we call \emph{simple symmetric lenses}. At the price
-of some expressiveness, our new lenses do not require complements; they only
-need source and target data as inputs.
-
-In order to characterize the expressive power of simple symmetric lenses
-theoretically, we prove that simple symmetric lenses include all (and only)
+this naive approach runs into two difficulties.
+
+The first difficulty is pragmatic. Symmetric lenses usually operate
+over three structures: the source $S$, the target $T$ and a complement
+$C$ that contains the information not
+present in either $S$ or $T$~\cite{symmetric-lenses}. Logistically, the complement must be stored
+somewhere and managed. More significantly, complements complicate synthesis
+specifications.
+Users can no longer just give examples of source and target
+pairings; instead they must also indicate how to manage the 
+``hidden state.'' To avoid such complexities, we define a new variant
+of symmetric lenses: \emph{simple symmetric lenses}. At the price
+of some expressiveness, these new lenses do not require complements.
+%only the usual source and target.
+%
+To characterize their expressive power,
+we prove that simple symmetric lenses include all (and only)
 those symmetric lenses that satisfy an intuitive property called
-\emph{forgetfulness}. In other words, they are the complete set of symmetric
+\emph{forgetfulness}, which we define in Section~\ref{sec:relationship}.
+Intuitively, simple symmetric lenses are the complete set of symmetric
 lenses that do not require external ``memory'' to recover data from past
-instances of $S$ or $T$ when making a round trip -- they only need the most
-recent instance. From a pragmatic perspective, we justify our design choices by
-extending the Boomerang implementation~\cite{Boomerang} with simple symmetric
-lenses applying them successfully to a range of real-world applications, and
-demonstrating that they can enrich, not just with basic bijective combinators,
-but with a rich set of advanced features including quotient lenses and matching
-lenses.
+instances of $S$ or $T$ when making a round trip. They only need the most
+recent instance. 
+We justify this loss of expressiveness pragmatically. 
+We extend Boomerang~\cite{Boomerang} with simple symmetric lenses and
+successfully use them on a range of real-world applications, showing
+in the process that simple symmetric lenses integrate smoothly with 
+other advanced lenses in Boomerang, including quotient
+lenses and matching lenses.
 
 The second difficulty with synthesizing symmetric lenses is more
-fundamental\dpw{better word?}: While the number of bijective lenses between two
+fundamental: While the number of bijective lenses between two
 related formats is often tiny, the number of symmetric lenses is often large. In
@@
 \subsection{Optimizations}
+\label{subsec:optimizations}
 
@@
 the asymmetric theory extensions present in Boomerang, like matching lenses and
-quotient lenses.
-
-We have extended the Optician synthesis plugin to Boomerang~\cite{?} with our
-symmetric synthesis algorithm. In doing so, users can write synthesis tasks
+quotient lenses. We also integrated our synthesis engine into Boomerang. It allows users to write synthesis tasks
 alongside lens combinators, use those synthesis results in the manual writing of
@@
 
-\item How much slower do we perform on bijective lens synthesis benchmarks than
+\item How much slower is our tool on bijective lens synthesis benchmarks than
   the existing Optician system customized for bijective lenses?
@@
 \begin{enumerate}
-\item We adapted \afm{8} data cleaning benchmarks from Flash Fill. In these
-  benchmarks, we not only provide functions that clean the data, but we also
-  provide other synchronization functions: if the cleaned data is updated, those
-  changes are reflected in the original format.
+\item We adapted \afm{8} data cleaning benchmarks from Flash Fill, though, of course, our tool produces bidirectional transformations rather than one-way transformers like Flash Fill.
 \item We adapted \afm{29} benchmarks from Augeas. These benchmarks turn the
@@
 
-To answer our first research question, we measure the runtime of our algorithm
-on our benchmark suite.
-
-We synthesized lenses in 2 modes:
+To begin, we measure the run time of our algorithm in 2 modes:
 
 \begin{tabulary}{\linewidth}{rL}
-  \SSOpt{}: & We synthesize lenses using our symmetric synthesis algorithm with all optimizations enabled.\\
-  \SSNCOpt{}: & We synthesize lenses using our symmetric synthesis algorithm,
-                but compositional synthesis is not enabled.\\
+  \SSOpt{}: & The symmetric synthesis algorithm with all optimizations enabled.\\
+  \SSNCOpt{}: & The symmetric synthesis algorithm, but compositional synthesis not enabled.\\
 \end{tabulary}\\
 
-We include running the symmetric synthesis algorithm without compositional
-synthesis to stress test the algorithms. While compositional synthesis makes the
-procedure easier, and demonstrates the potential to be used in arbitrarily hard
-problems, it is dependent on the user identifying important sublenses between
-subcomponents of the final formats. It is therefore interesting to see how many
-problems each algorithm can solve, when trying to synthesize ``in one go.''
+Recall that compositional synthesis exploits user annotations to break a benchmark into a series of smaller synthesis steps.
+In our experience, compositional synthesis (SS mode) allows our system to scale to arbitrarily large and complex formats.
+However, SSNC mode, which uses no such annotations and synthesizes a complete lens all at once provides a useful experimental stress test for the system.
+Recall from Section
 
@@
 able to solve 26 problem instances.
+Recall from Section~\ref{subsec:optimizations} that users may tune the termination metric.
+Success on this set of benchmarks requires we tune that termination codition (e.g.
+for some benchmarks, allowing the system flexibility to spend more time considering more of the search space).
+In Section~\ref{subsec:effect-of-heuristics}, we will consider the effect of fixing the termination metric.
 
@@
 benchmarks had to be slightly altered with tighter specifications or termination
-parameters, as we had to alter the original specifications to generate the
+parameters\dpw{What does that mean? You had to alter it to get the system to work?}, as we had to alter the original specifications to generate the
 correct lens.
@@
 \begin{tabulary}{\linewidth}{rL}
-  \SSOpt{}: & We synthesize lenses using our symmetric synthesis algorithm with all optimizations enabled.\\
-  \BSOpt{}: & We synthesize lenses using the existing bijective synthesis
+  \SSOpt{}: & The symmetric synthesis algorithm with all optimizations enabled.\\
+  \BSOpt{}: & The existing bijective synthesis
               algorithm with all optimizations enabled.\\
@@
 average, \SSOpt{} took 2.1 times (\afm{3.5} seconds) longer to complete than
-\BSOpt{}.
-
-\begin{figure}
-  \includegraphics{generated-graphs/times_bijective}
-  \caption{Number of benchmarks that can be solved by a given algorithm in a
-    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
-    \BSOpt{} is the full bijective lens synthesis algorithm. \SSNCOpt{} is the
-    symmetric synthesis algorithm without using a library of existing lenses.
-    \BSNCOpt{} is the bijective synthesis algorithm without using a library of
-    existing lenses. Our symmetric synthesis algorithm is able to complete all
-    benchmarks in our benchmark suite in under 30 seconds, and is able to
-    complete most of the benchmarks without using compositional synthesis.}
-  \label{fig:times_bijective}
-\end{figure}
-
-The slowest completed benchmark for both synthesis algorithms is
+\BSOpt{}. The slowest completed benchmark for both synthesis algorithms is
 \texttt{xml\_to\_augeas.boom}, a benchmark that converts arbitrary XML up to
@@
 
+\begin{figure}
+  \includegraphics{generated-graphs/times_bijective}
+  \caption{Number of benchmarks that can be solved by a given algorithm in a
+    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
+    \BSOpt{} is the full bijective lens synthesis algorithm.}
+  \label{fig:times_bijective}
+\end{figure}
+
 \subsection{The Effects of Our Heuristics}
+\label{subsec:effect-of-heuristics}
 We evaluate the usefulness of (1) our information theoretic metric, and (2) our
@@
 \begin{tabulary}{\linewidth}{rL}
-  \AnyOpt{}: & We synthesize any lens that satisfies the specification, with no
+  \AnyOpt{}: & We synthesize any lens \dpw{Did you mean the following?
+But is the order truly arbitrary?
+Is size or anything else a driver?:] We synthesize lenses in an arbitrary order (i.e.
+without use of the information theoretic preference metric)} that satisfies the specification, with no
                notion of cost.\\
   \FLOpt{}: &  The first time \GreedySynth finds a lens, the synthesis procedure
-              returns that lens.\\
+              returns that lens.\dpw{What controls the ``greediness''? Say a little something else to more clearly differentiate from Any.}\\
   \CCOpt{}: & We replace our information theoretic cost metric with one where
@@
 
-Figure~\ref{fig:metric} summarizes the result of this experiment. We find that
+We experimented with the \CCOpt{} mode to determine whether the complexity of the information theoretic measure was really needed.
+Related work on string transformations has often used simpler measures such as ``avoid constants'' which align with, but are simpler than our measures.
+The \CCOpt{} mode is an example of such a simple measure --- it operates by counting disconnects, which put a complete stop to information transfer, like constants, and merges, which eliminate the information in a union.
+Figure~\ref{fig:metric} summarizes the result of this experiment.
+We find that
 our metrics are very important for finding the correct lens. Our

### 2018-07-10 papers/SymmetricLenses/main.tex

@@
 A simple symmetric lens between formats described by types $X$ and $Y$ is
-defined by four functions.\bcp{I think a picture would be helpful. I can
-  probably make one pretty easily by tweaking the one from the original
-  symmetric paper.} \afm{That would be awesome.}
+defined by four functions.
 

### 2018-07-10 papers/SymmetricLenses/main.tex

@@
 A natural way to expand the scope of such tools is to consider a more
-exrpressive language of lenses.  For example, \emph{symmetric
+expressive language of lenses.  For example, \emph{symmetric
 lenses}~\cite{symmetric-lenses} encompass a broad collection of bidirectional
@@
 fundamental: While the number of bijective lenses between two
-related formats is often tiny, the number of symmetric lenses is often large. In
+related formats is often tiny, the number of simple symmetric lenses is much larger. In
 practice, the difference in cardinality means that when a search algorithm
-selects the first bijective lens it finds, that lens has a good chance of being
-the one a user wants. In contrast, and to our surprise, our experiments show
-that if our search algorithm selects the first symmetric lens it finds, it
-typically fails to find the lens the user wants. The flexibity these lenses
-admit when it comes to eliding information generates an explosion of possible
-transformations. To disambiguate between the many possible lenses, we need new
-principles for identifying the ``likely'' ones, and we need new algorithms to
-search the space of program transformations using these principles.
-
-To overcome this second difficulty, we turned to information theory to try to
-find a way to characterize the likely and unlikely lenses. Intuitively, a
-``likely'' lens is one that propagates a lot of information from source to
-target and also from target back to source. Conversely, given the source, the
-target is ``unlikely'' to require very much additional information in order to
-reconstruct it, under the assumption that users typically want to synchronize
-formats with a large amount of shared information. In order to implement this
+returns the first bijective lens it finds, that lens has a good chance of being
+the desired lens. In contrast, and to our initial surprise, experiments showed
+that if a naive search algorithm selects the first simple symmetric lens it
+finds, it is typically \emph{not} the desired lens.  The flexibility
+that such lenses provide in eliding information 
+generates an explosion of possible transformations. To disambiguate
+between the many possible lenses, we need new 
+principles for identifying the ``likely'' ones, and we need new
+algorithms guided by those principles to
+search the space of program transformations.
+
+To overcome this second difficulty, we turn to information theory 
+to characterize the likely and unlikely lenses. 
+We consider a ``likely'' lens to be one that propagates a lot of
+information from source to 
+target and vice versa. Conversely, an ``unlikely'' lens requires a
+large amount of additional information to recover the target given a
+source and vice versa.  
+This categorization follows from the intuition
+that users will typically want to synchronise formats with lots
+of shared information, and so lenses that preserve 
+more information are preferable.
+
+To implement this
 intuition, we turn to \emph{stochastic regular expressions} (SREs). Stochastic
@@
 distributions allow us to compute the amount of information gained when
-presented an element of the set. For example, the language of the SRE
-$\lstinline{"a"} |_{0.5}
-\lstinline{"b"}$ contains two elements (the strings \lstinline{"a"} and \lstinline{"b"}).
-The probability 0.5 annotating the union indicates that both elements are
-equally likely to appear. Consequently, if we observe an element of that
-language (either \lstinline{"a"} or
-\lstinline{"b"}), we will gain one bit of information. Said another way, one bit, together with
-the SRE, suffices to recover the string in question. Interestingly, if the
-probability annotation were anything but 0.5, the expected information content
-(or entropy) would actually be less than 1 (given a stream of such choices, on
-average we could encode the the contents of stream in fewer bits than the the
-number of choices present). On the other hand, consider the language of the SRE
-\lstinline{c}. It contains just one element (the string $c$). If we observe that
-our algorithm generates $\lstinline{"c"}$, we gain zero bits of information ---
-the string was already determined. Using this foundation, a lens $\Lens$ that
-synchronizes data between $\Regex$ and $\RegexAlt$ is ``likely,''
-in a technical sense, if given a string $s$ in the source $S$, very little information
-is needed to determine which synchronized string $t$ is present in the target
-$T$.
+presented an element of the set.
 
@@
 entropy of the SRE.
+
+\ksf{Consider integrating more detailed discussion of SREs that was
+formerly in the intro.  Commented out here for now.}
+%%To implement this intuition, we turn to \emph{stochastic regular
+%%expressions} (SREs). Stochastic 
+%%regular expressions define a set of strings, as do ordinary regular expressions,
+%%but they also provide the expected distribution of strings within that set. Such
+%%distributions allow us to compute the amount of information gained when
+%%presented an element of the set. For example, the language of the SRE
+%%$\lstinline{"a"} |_{0.5}
+%%\lstinline{"b"}$ contains two elements (the strings \lstinline{"a"} and \lstinline{"b"}).
+%%The probability 0.5 annotating the union indicates that both elements are
+%%equally likely to appear. Consequently, if we observe an element of that
+%%language (either \lstinline{"a"} or
+%%\lstinline{"b"}), we will gain one bit of information. Said another way, one bit, together with
+%%the SRE, suffices to recover the string in question. Interestingly, if the
+%%probability annotation were anything but 0.5, the expected information content
+%%(or entropy) would actually be less than 1 (given a stream of such choices, on
+%%average we could encode the the contents of stream in fewer bits than the the
+%%number of choices present). On the other hand, consider the language of the SRE
+%%\lstinline{c}. It contains just one element (the string $c$). If we observe that
+%%our algorithm generates $\lstinline{"c"}$, we gain zero bits of information ---
+%%the string was already determined. Using this foundation, a lens $\Lens$ that
+%%synchronizes data between $\Regex$ and $\RegexAlt$ is ``likely,''
+%%in a technical sense, if given a string $s$ in the source $S$, very little information
+%%is needed to determine which synchronized string $t$ is present in the target
+%%$T$.
 

### 2018-07-10 papers/SymmetricLenses/main.tex
in progress edits to intro

--- 2018-07-10 papers/SymmetricLenses/main.tex
@@
 \subsection{Optimizations}
-\label{subsec:optimizations}
 
@@
 the asymmetric theory extensions present in Boomerang, like matching lenses and
-quotient lenses. We also integrated our synthesis engine into Boomerang. It allows users to write synthesis tasks
+quotient lenses.
+
+We have extended the Optician synthesis plugin to Boomerang~\cite{?} with our
+symmetric synthesis algorithm. In doing so, users can write synthesis tasks
 alongside lens combinators, use those synthesis results in the manual writing of
@@
 
-\item How much slower is our tool on bijective lens synthesis benchmarks than
+\item How much slower do we perform on bijective lens synthesis benchmarks than
   the existing Optician system customized for bijective lenses?
@@
 \begin{enumerate}
-\item We adapted \afm{8} data cleaning benchmarks from Flash Fill, though, of course, our tool produces bidirectional transformations rather than one-way transformers like Flash Fill.
+\item We adapted \afm{8} data cleaning benchmarks from Flash Fill. In these
+  benchmarks, we not only provide functions that clean the data, but we also
+  provide other synchronization functions: if the cleaned data is updated, those
+  changes are reflected in the original format.
 \item We adapted \afm{29} benchmarks from Augeas. These benchmarks turn the
@@
 
-To begin, we measure the run time of our algorithm in 2 modes:
+To answer our first research question, we measure the runtime of our algorithm
+on our benchmark suite.
+
+We synthesized lenses in 2 modes:
 
 \begin{tabulary}{\linewidth}{rL}
-  \SSOpt{}: & The symmetric synthesis algorithm with all optimizations enabled.\\
-  \SSNCOpt{}: & The symmetric synthesis algorithm, but compositional synthesis not enabled.\\
+  \SSOpt{}: & We synthesize lenses using our symmetric synthesis algorithm with all optimizations enabled.\\
+  \SSNCOpt{}: & We synthesize lenses using our symmetric synthesis algorithm,
+                but compositional synthesis is not enabled.\\
 \end{tabulary}\\
 
-Recall that compositional synthesis exploits user annotations to break a benchmark into a series of smaller synthesis steps.
-In our experience, compositional synthesis (SS mode) allows our system to scale to arbitrarily large and complex formats.
-However, SSNC mode, which uses no such annotations and synthesizes a complete lens all at once provides a useful experimental stress test for the system.
-Recall from Section
+We include running the symmetric synthesis algorithm without compositional
+synthesis to stress test the algorithms. While compositional synthesis makes the
+procedure easier, and demonstrates the potential to be used in arbitrarily hard
+problems, it is dependent on the user identifying important sublenses between
+subcomponents of the final formats. It is therefore interesting to see how many
+problems each algorithm can solve, when trying to synthesize ``in one go.''
 
@@
 able to solve 26 problem instances.
-Recall from Section~\ref{subsec:optimizations} that users may tune the termination metric.
-Success on this set of benchmarks requires we tune that termination codition (e.g.
-for some benchmarks, allowing the system flexibility to spend more time considering more of the search space).
-In Section~\ref{subsec:effect-of-heuristics}, we will consider the effect of fixing the termination metric.
 
@@
 benchmarks had to be slightly altered with tighter specifications or termination
-parameters\dpw{What does that mean? You had to alter it to get the system to work?}, as we had to alter the original specifications to generate the
+parameters, as we had to alter the original specifications to generate the
 correct lens.
@@
 \begin{tabulary}{\linewidth}{rL}
-  \SSOpt{}: & The symmetric synthesis algorithm with all optimizations enabled.\\
-  \BSOpt{}: & The existing bijective synthesis
+  \SSOpt{}: & We synthesize lenses using our symmetric synthesis algorithm with all optimizations enabled.\\
+  \BSOpt{}: & We synthesize lenses using the existing bijective synthesis
               algorithm with all optimizations enabled.\\
@@
 average, \SSOpt{} took 2.1 times (\afm{3.5} seconds) longer to complete than
-\BSOpt{}. The slowest completed benchmark for both synthesis algorithms is
+\BSOpt{}.
+
+\begin{figure}
+  \includegraphics{generated-graphs/times_bijective}
+  \caption{Number of benchmarks that can be solved by a given algorithm in a
+    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
+    \BSOpt{} is the full bijective lens synthesis algorithm. \SSNCOpt{} is the
+    symmetric synthesis algorithm without using a library of existing lenses.
+    \BSNCOpt{} is the bijective synthesis algorithm without using a library of
+    existing lenses. Our symmetric synthesis algorithm is able to complete all
+    benchmarks in our benchmark suite in under 30 seconds, and is able to
+    complete most of the benchmarks without using compositional synthesis.}
+  \label{fig:times_bijective}
+\end{figure}
+
+The slowest completed benchmark for both synthesis algorithms is
 \texttt{xml\_to\_augeas.boom}, a benchmark that converts arbitrary XML up to
@@
 
-\begin{figure}
-  \includegraphics{generated-graphs/times_bijective}
-  \caption{Number of benchmarks that can be solved by a given algorithm in a
-    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
-    \BSOpt{} is the full bijective lens synthesis algorithm.}
-  \label{fig:times_bijective}
-\end{figure}
-
 \subsection{The Effects of Our Heuristics}
-\label{subsec:effect-of-heuristics}
 We evaluate the usefulness of (1) our information theoretic metric, and (2) our
@@
 \begin{tabulary}{\linewidth}{rL}
-  \AnyOpt{}: & We synthesize any lens \dpw{Did you mean the following?
-But is the order truly arbitrary?
-Is size or anything else a driver?:] We synthesize lenses in an arbitrary order (i.e.
-without use of the information theoretic preference metric)} that satisfies the specification, with no
+  \AnyOpt{}: & We synthesize any lens that satisfies the specification, with no
                notion of cost.\\
   \FLOpt{}: &  The first time \GreedySynth finds a lens, the synthesis procedure
-              returns that lens.\dpw{What controls the ``greediness''? Say a little something else to more clearly differentiate from Any.}\\
+              returns that lens.\\
   \CCOpt{}: & We replace our information theoretic cost metric with one where
@@
 
-We experimented with the \CCOpt{} mode to determine whether the complexity of the information theoretic measure was really needed.
-Related work on string transformations has often used simpler measures such as ``avoid constants'' which align with, but are simpler than our measures.
-The \CCOpt{} mode is an example of such a simple measure --- it operates by counting disconnects, which put a complete stop to information transfer, like constants, and merges, which eliminate the information in a union.
-Figure~\ref{fig:metric} summarizes the result of this experiment.
-We find that
+Figure~\ref{fig:metric} summarizes the result of this experiment. We find that
 our metrics are very important for finding the correct lens. Our

### 2018-07-10 papers/SymmetricLenses/main.tex
add high level alg and main tex

@@
 and search using similar heuristics.  However, 
-this naive approach runs into two difficulties.
-
-The first difficulty is pragmatic. Symmetric lenses usually operate
-over three structures: the source $S$, the target $T$ and a complement
-$C$ that contains the information not
-present in either $S$ or $T$~\cite{symmetric-lenses}. Logistically, the complement must be stored
-somewhere and managed. More significantly, complements complicate synthesis
-specifications.
-Users can no longer just give examples of source and target
-pairings; instead they must also indicate how to manage the 
-``hidden state.'' To avoid such complexities, we define a new variant
-of symmetric lenses: \emph{simple symmetric lenses}. At the price
-of some expressiveness, these new lenses do not require complements.
-%only the usual source and target.
-%
-To characterize their expressive power,
-we prove that simple symmetric lenses include all (and only)
+this naive approach runs into two difficulties. \dpw{I got into the
+weeds a little bit there. 
+The goal was to set up the idea that there are ``two parts'' to a symmetric
+lens transformation: the information from $s$ and the information from $t$.  In
+a future paragraph in the intro we are going to explain that the central heuristic
+involves minimizing the information from $t$.}
+
+The first difficulty was pragmatic: symmetric lenses actually usually operate
+over three structures, the source $S$, the target $T$ and a complement
+$C$~\cite{symmetric-lenses}. The latter object contains all the information not
+present in one of $S$ or $T$. Logistically, the complement object must be stored
+away somewhere and infrastructure must be developed to manage it. Perhaps more
+relevant though, is that complements complicate specifications, even when those
+specifications are just examples. The issues is that one cannot just give
+examples of source and target; one must also give guidance on how to manage the
+``hidden state.'' In order to avoid such complexities, we defined a new variant
+of symmetric lenses, which we call \emph{simple symmetric lenses}. At the price
+of some expressiveness, our new lenses do not require complements; they only
+need source and target data as inputs.
+
+In order to characterize the expressive power of simple symmetric lenses
+theoretically, we prove that simple symmetric lenses include all (and only)
 those symmetric lenses that satisfy an intuitive property called
-\emph{forgetfulness}, which we define in Section~\ref{sec:relationship}.
-Intuitively, simple symmetric lenses are the complete set of symmetric
+\emph{forgetfulness}. In other words, they are the complete set of symmetric
 lenses that do not require external ``memory'' to recover data from past
-instances of $S$ or $T$ when making a round trip. They only need the most
-recent instance. 
-We justify this loss of expressiveness pragmatically. 
-We extend Boomerang~\cite{Boomerang} with simple symmetric lenses and
-successfully use them on a range of real-world applications, showing
-in the process that simple symmetric lenses integrate smoothly with 
-other advanced lenses in Boomerang, including quotient
-lenses and matching lenses.
+instances of $S$ or $T$ when making a round trip -- they only need the most
+recent instance. From a pragmatic perspective, we justify our design choices by
+extending the Boomerang implementation~\cite{Boomerang} with simple symmetric
+lenses applying them successfully to a range of real-world applications, and
+demonstrating that they can enrich, not just with basic bijective combinators,
+but with a rich set of advanced features including quotient lenses and matching
+lenses.
 
 The second difficulty with synthesizing symmetric lenses is more
-fundamental: While the number of bijective lenses between two
+fundamental\dpw{better word?}: While the number of bijective lenses between two
 related formats is often tiny, the number of symmetric lenses is often large. In

### 2018-07-10 papers/SymmetricLenses/main.tex
add response to steve comments

--- 2018-07-10 papers/SymmetricLenses/main.tex
@@
 The expected number of bits required to recover data is a well-known concept in
-information theory: it is the \emph{entropy} of the data. \saz{Cite Shannon?} Calculating the
-entropy of data requires a probability distribution over the space of possible
-values for the data. Specifically, given a set $S$ and a probability
-distribution over that set $P \OfType S \rightarrow \Reals$, the entropy of $S$
-with probability distribution $P$ is $-\Sigma_{s\in S}P(s)\Log_2P(s)$.
+information theory: it is the \emph{entropy} of the data~\cite{Shannon1948}.
+Calculating the entropy of data requires a probability distribution over the
+space of possible values for the data. Specifically, given a set $S$ and a
+probability distribution over that set $P \OfType S \rightarrow \Reals$, the
+entropy of $S$ with probability distribution $P$ is $-\Sigma_{s\in
+  S}P(s)\Log_2P(s)$.
 
@@
 \label{sec:ssl}
-\saz{This paragraph used to say that lenses are typed between stochastic REs
-  and it also says that lenses are typed between REs.  I think its should just
-  be REs? Double check that this is correct}
-
 We now give a formal description of all the combinators in our symmetric lens
@@
 probability annotations. In a slight abuse of notation\saz{Why is this an
-  abuse?}, if $\Regex$ is a stochastic regular expression, $\BRegex$ is the
-regular expression obtained by removing probability annotations. The typing
-judgment $\Lens \OfType \BRegex \Leftrightarrow \BRegexAlt$ means that $\Lens$
-is a well typed lens between $\LanguageOf{\BRegex}$ and
-$\LanguageOf{\BRegexAlt}$.
+  abuse?}\afm{We also use \BRegex{},\BRegexAlt{} as nonterminals}, if $\Regex$
+is a stochastic regular expression, $\BRegex$ is the regular expression obtained
+by removing probability annotations. The typing judgment $\Lens \OfType \BRegex
+\Leftrightarrow \BRegexAlt$ means that $\Lens$ is a well typed lens between
+$\LanguageOf{\BRegex}$ and $\LanguageOf{\BRegexAlt}$.
 
@@
 \item adds all of the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)
-  to the priority queue with $\PCF{\PQ.Push}$ \saz{what determines the priority order?}
+  to the priority queue with $\PCF{\PQ.Push}$. The priority of a SRE pair is the
+  number of rewrites needed to derive the pair from the originals.
 \end{itemize}
@@
 The first synthesizer is one that searches through regular expression pairs.
-This synthesizer, comprising \Expand, \PQ, and \RXSearch \saz{The name \RXSearch
-  does not appear in the figure.  I get the impression that it is the name of
-  the while loop, but perhaps we can make that clearer in the algorithm code?
-  Maybe by putting a brace around the while loop and giving it a name? Here we
-  could say something like ``... and a loop that continues searching until a
-  satisfactory lens is found (we call this loop \RXSearch)}.  It proposes regular
-expressions that are semantically, but not syntactically, equivalent to the originals. In
-particular, this synthesizer performs the following rewrites:
+This synthesizer comprises \Expand, \PQ, and \Continue. It proposes regular
+expressions that are semantically, but not syntactically, equivalent to the
+originals. In particular, this synthesizer performs the following rewrites:
 \begin{center}
@@
 \label{subsec:greedy-synth}
-\saz{We should call out the structural similarity
-  between the DNF grammar and the organization of these algorithms.}
-The greedy algorithm is actually two\saz{three! don't forget AtomSynth} greedy algorithms, one that greedily finds
-symmetric DNF lenses (\GreedySynth) and one that greedily finds symmetric
-sequence lenses (\GreedySeqSynth). We describe this synthesis algorithm in the
-context of trying to find a lens between \lstinline{"" | name.name*} and
-\lstinline{"" | name}.
+The greedy algorithm is actually three algorithms, one that greedily finds
+symmetric DNF lenses (\GreedySynth), one that greedily finds symmetric sequence
+lenses (\GreedySeqSynth), and one that finds symmetric atom lenses
+(\AtomSynth{}). These three algorithms are hierarchically structured:
+\GreedySynth relies on \GreedySeqSynth, \GreedySeqSynth relies on \AtomSynth,
+and \AtomSynth rlies on \GreedySynth. The structure of the algorithms mirrors
+the structure of symmetric DNF lenses and \SDNFREabbrevs. We describe this
+synthesis algorithm in the context of trying to find a lens between
+\lstinline{"" | name.name*} and \lstinline{"" | name}.
 
@@
 function \CartesianMap maps its argument across the cross product of the input
-lists).  A priority queue containing these sequence lenses is then initialized
-with $\PCF{PQ.Create}$.\saz{Ordered how?} The symmetric lens is then built up
-iteratively from these sequence lenses, where the state of the partially constructed lens is
-tracked in the lens builder, $lb$.
+lists). A priority queue containing these sequence lenses, ordered by cost, is
+then initialized with $\PCF{PQ.Create}$. The symmetric lens is then built up
+iteratively from these sequence lenses, where the state of the partially
+constructed lens is tracked in the lens builder, $lb$.
 
@@
 converted into a symmetric DNF lens. This is only possible if all sequences are
-involved in some sequence lens: if they are not, \PCF{GreedyState.ToDNFLens}
+involved in some sequence lens: if they are not, \PCF{LensBuilder.ToDNFLens}
 instead returns \None.
@@
     \If{$\Call{LensBuilder.UsefulAdd}{lb,\SSQLens,\Examples}$}
-    \State $lb \gets \Call{GreedyState.AddSeq}{lb,\SSQLens}$
+    \State $lb \gets \Call{LensBuilder.AddSeq}{lb,\SSQLens}$
     \EndIf
@@
 
-The inputs to \GreedySeqSynth are a suite of input-output
-examples and a pair of stochastic atoms. As in \GreedySynth, \GreedySeqSynth
-returns \None early if there is no possible lens. Afterward,
-\GreedySeqSynth finds the best lenses between each atom pair of the left and
-right sequences, and organizes them into a priority queue 
-with \PCF{PQ.Create}. The symmetric sequence lens is built up iteratively from
-these atom lenses, where the state of the partially built lens is tracked in the sequence lens
-builder, $slb$.
+The inputs to \GreedySeqSynth are a suite of input-output examples and a pair of
+stochastic atoms. As in \GreedySynth, \GreedySeqSynth returns \None early if
+there is no possible lens. Afterward, \GreedySeqSynth finds the best lenses
+between each atom pair of the left and right sequences, and organizes them into
+a priority queue ordered by cost with \PCF{PQ.Create}. The symmetric sequence
+lens is built up iteratively from these atom lenses, where the state of the
+partially built lens is tracked in the sequence lens builder, $slb$.
 
@@
 queue. Within this loop, atom lenses are popped from the queue and the
-``useful'' ones are kept.  An atom lens is useful if adding it will lower the cost of the
-generated sequence lens, or if examples show that one of the atoms must be
-connected to another lens.\saz{I'm not sure what ``must be connected to another
-  lens means''}  The priorities of these lenses change over time, because
-each atom can only be involved in one lens. Adding a lens may necessitate
-removing another, so not all popped atom lenses will lower the cost of the
-generated sequence lens.\saz{I don't understand the last two sentences.}
+``useful'' ones are kept. An atom lens is useful if adding it will lower the
+cost of the generated sequence lens, or if examples show that one of the atoms
+must not be disconnected. Each atom can only be part of one lens at a time, so
+the algorithm must at times disconnect previously connected atoms to connect an
+atom that must not be disconnected.
 
 Finally, after all atoms have been popped, the sequence lens is validated and
-returned. The algorithm confirms that all atoms that need to be connected with
-an atom lens, are. If they are not, \PCF{SLensBuilder.ToDNFLens} instead returns
-\None.
+returned. The algorithm confirms that all atoms that must not be disconnected
+are involved in an atom lens. If they are not, \PCF{SLensBuilder.ToDNFLens}
+instead returns \None.
 
@@
 
-Our implementation includes a number of optimizations to this algorithm.  We added ``open'' and
-``closed'' annotations to regular expressions \saz{what do these mean? I guess
-  ``open'' means ``subject to  DNF conversion'' and ``closed'' means not?}, an expansion inference algorithm,
-hash-consing and memoization, along with compositional synthesis. The implementation
-also allows an additional parameter in specifications that customizes the
-termination condition for the synthesis procedure.
+Our implementation includes a number of optimizations to this algorithm. We
+added ``closed'' annotations to regular expressions \saz{what do these mean? I
+  guess ``open'' means ``subject to DNF conversion'' and ``closed'' means
+  not?}\afm{I describe in greater detail below, should I add in additional info
+  here too?}, an expansion inference algorithm, hash-consing and memoization,
+along with compositional synthesis. The implementation also allows an additional
+parameter in specifications that customizes the termination condition for the
+synthesis procedure.
 
@@
 ``closed'' atom. When converted into a DNF regular expression, if $\Regex$ is
-such a closed stochastic regular expression, then
-$\ToDNFRegexOf{\Regex} = \AtomToDNFOf{\Regex}$.
+annotated as ``closed'', then $\ToDNFRegexOf{\Regex} = \AtomToDNFOf{\Regex}$.
+Regular expressions that are not annotated as closed, are considered ``open.''
 
@@
   \tag{\PutLR}
-  \frac{\PutLSymOf{(y,c)} = (x,c')}{\PutRSymOf{(x,c')} = (y,c')}
+  \frac{\PutLSymOf{(y,c)} = (x,c')}{\PutRSymOf{(x,c'')} = (y,c')}
 \end{equation}

### 2018-07-10 papers/SymmetricLenses/main.tex
eval changes

@@
 synthesis algorithm on the original Optician benchmark suite. \afm{8} of these
-benchmarks had to be slightly altered with tighter specifications or termination
-parameters\dpw{What does that mean? You had to alter it to get the system to work?}, as we had to alter the original specifications to generate the
-correct lens.
+benchmarks had to be slightly altered to work with our new system. We altered
+these benchmarks by either providing additional examples termination parameters.
+Without these alterations, the symmetric synthesis synthesized a lens that fit
+the specification but not the intended lens.
 
@@
 \begin{tabulary}{\linewidth}{rL}
-  \AnyOpt{}: & We synthesize any lens \dpw{Did you mean the following?
-But is the order truly arbitrary?
-Is size or anything else a driver?:] We synthesize lenses in an arbitrary order (i.e.
-without use of the information theoretic preference metric)} that satisfies the specification, with no
-               notion of cost.\\
+  \AnyOpt{}: & We synthesize lenses without use of the information
+               theoretic preference metric.  Once \GreedySynth finds a lens, the
+  synthesis procedure returns that lens.\\
   \FLOpt{}: &  The first time \GreedySynth finds a lens, the synthesis procedure
-              returns that lens.\dpw{What controls the ``greediness''? Say a little something else to more clearly differentiate from Any.}\\
+              returns that lens. However, \GreedySynth uses the information
+              theoretic preference metric to choose a lens.\\
   \CCOpt{}: & We replace our information theoretic cost metric with one where

### 2018-07-10 papers/SymmetricLenses/main.tex
fill in citations

@@
 distributions over that data.
-To that end, we adopt \emph{stochastic regular expressions}~\cite{?} (SREs), 
+To that end, we adopt \emph{stochastic regular expressions}~\cite{stoch-def,stoch-rnn} (SREs), 
 which are regular expressions where each operator is annotated with a
@@
 However, existing work does not provide a means to reason about stochastic
-regular expression equivalence. We extend the \emph{star-semiring}~\cite{?}
+regular expression equivalence. We extend the \emph{star-semiring}~\cite{Droste2009}
 equivalences to stochastic regular expressions.
@@
 The \MergeR lens is interesting because it merges data where one data can be in
-two formats, and one data has only one format. In previous work~\cite{?}, this
-was combined into \OrLens{}, where \OrLens{} could have ambiguous types, but we
-find it more clear to have explicit merge operators: it is easier to see what
-lens the synthesis algorithm is creating.
+two formats, and one data has only one format. In previous
+work~\cite{Boomerang}, this was combined into \OrLens{}, where
+\OrLens{} could have ambiguous types, but we find it more clear to have explicit
+merge operators: it is easier to see what lens the synthesis algorithm is
+creating.
 

### 2018-07-10 papers/SymmetricLenses/main.tex
in progress edits to section 1

--- 2018-07-10 papers/SymmetricLenses/main.tex
@@
 more information are preferable.
-
 To implement this
-intuition, we turn to \emph{stochastic regular expressions} (SREs). Stochastic
-regular expressions define a set of strings, as do ordinary regular expressions,
-but they also provide the expected distribution of strings within that set. Such
-distributions allow us to compute the amount of information gained when
-presented an element of the set.
+intuition, we turn to \emph{stochastic regular expressions}
+(SREs)~\cite{?}, which simultaneously define a set of strings \`a la
+regular expressions and a probability distribution over those
+strings.  Using this information, we can calculate the shared
+information content of two formats and our measure of
+likeliness.
 
@@
   \includegraphics[width=.5\textwidth]{high-level-algorithm.pdf}
-  \caption{Schematic diagram for our symmetric lens synthesis algorithm. Regular
-    expressions \Regex and \RegexAlt, and a set of examples \Examples, are
-    provided as input. \Expand searches through stochastic regular expression
-    pairs, equivalent to the originals, and proposes them to \GreedySynth.
-    \GreedySynth finds a lens, typed between the generated equivalent pairs.
-    When the algorithm determines it has an optimal lens, it returns
-  that lens. \ksf{Figure needs to be changed to call first search
-  Expand; Are the S and T regular expressions (which should have bars)
-  or stochastic regular expressions (which should be called such)?}}
+  \caption{Schematic diagram for the simple symmetric lens synthesis
+  algorithm. The user provides 
+   regular expressions \BRegex and \BRegexAlt and a set of examples \Examples{}
+    as input. \Expand first converts \BRegex and \BRegexAlt to
+    stochastic regular expressions \Regex and \RegexAlt with default
+    probabilities. It then finds pairs of stochastic regular expressions
+    equivalent to \Regex and \RegexAlt and iteratively proposes them to \GreedySynth.
+    \GreedySynth finds a lens typed between the supplied SREs.
+    When the algorithm determines it has found a likely lens, it
+  returns it.  \ksf{can we change the subscript on the returned lens
+  to ``likely'' instead of ``best''? Also, is it the cost that is
+  returned or the likeliness }
+   }
   \label{fig:high-level-algorithm}
@@
 
-With simple symmetric lenses and SREs in hand, we develop a new algorithm for
-synthesizing likely lenses, based on the previous algorithm for synthesizing
-bijective lenses. This algorithm is constructed from \Expand and \GreedySynth,
-a pair of collaborating search procedures, shown in
-Figure~\ref{fig:high-level-algorithm}. Given source and target SREs $S$ and $T$,
-the \Expand searches for two \emph{related} SREs $S'$ and $T'$ that are
-\emph{compatible} (a heuristic that estimates the likelihood our second
-algorithm will succeed). $S'$ and $T'$ are related in the sense that they define
-the same regular sets as $S$ and $T$ (respectively) and the distribution of
-strings is identical. Indeed, we guarantee these two properties by implementing
-a search procedure that obeys the star semi-ring axioms of regular expression
-equivalence, and which we show does not alter the probability distributions of
-the languages.
-
-Given the two compatible SREs, \GreedySynth uses a greedy, type-directed
-algorithm find a likely lens between them. If algorithm estimates that the lens
-it has discovered is more likely than any future lens it can possibly find, the
-search terminates. Otherwise, this lens is stored away (if it is the best found
-so far) and the search continues with \Expand again.
+With simple symmetric lenses and our SRE-based likeliness measure in
+hand, we develop a new algorithm for synthesizing likely lenses.
+At its core, the algorithm performs a type-directed search between the 
+descriptions of the source and target formats, measuring success using
+the likeliness measure.
+However, regular expressions enjoy a rich algebraic structure, meaning
+there are infinitely many regular expressions equivalent to the
+original pair, and the lens returned by a type-directed search depends
+upon the chosen representation. 
+To tame this complexity, we divide the synthesis algorithm into two communicating
+search procedures (Figure~\ref{fig:high-level-algorithm}), following
+earlier work~\cite{maina+:quotient-synthesis}.
+The first algorithm, \Expand, uses rewriting rules to iteratively
+propose new pairs of stochastic regular expressions equivalent to the
+original pair.
+The second algorithm, \GreedySynth, takes a pair of such SREs as input
+and uses a greedy, type-directed algoithm to find a simple symmetric
+lens between them, returning the lens and its likelihood score
+to \Expand.  The algorithm terminates when the likilihood score of the
+best lens found so far compared with 
+the complexity of newly generated SREs 
+suggests that further searching is unlikely to be fruitful.
+
+%%
+%%searches for two \emph{related} SREs $S'$ and $T'$ that are
+%%\emph{compatible} (a heuristic that estimates the likelihood our second
+%%algorithm will succeed). $S'$ and $T'$ are related in the sense that they define
+%%the same regular sets as $\BRegex$ and $\BRegexAlt$ (respectively) and the distribution of
+%%strings is identical. Indeed, we guarantee these two properties by implementing
+%%a search procedure that obeys the star semi-ring axioms of regular expression
+%%equivalence, and which we show does not alter the probability distributions of
+%%the languages.
+
+%%Given the two compatible SREs, \GreedySynth uses a greedy, type-directed
+%%algorithm find a likely lens between them. If algorithm estimates that the lens
+%%it has discovered is more likely than any future lens it can possibly find, the
+%%search terminates. Otherwise, this lens is stored away (if it is the best found
+%%so far) and the search continues with \Expand again.
 

### 2018-07-10 papers/SymmetricLenses/main.tex
add changes

--- 2018-07-10 papers/SymmetricLenses/main.tex
@@
 \paragraph*{Symmetric Lenses}
-A classical symmetric lens
-$\Lens$ between $X$ and $Y$ consists of 4 components: a complement $C$, a
-designated element $init \in C$, and two functions,
-$\PutRSym$ and $\PutLSym$, that propagate changes in one format to the other.
-\begin{enumerate}
-\item $\PutRSym \OfType X \times C \rightarrow Y \times C$
-\item $\PutLSym \OfType Y \times C \rightarrow X \times C$
-\end{enumerate}
-In this formulation, data unique to each side are stored in the complement.
-When one format is edited, the \PutR
-or \PutL function stitches together the edited data with data
-stored in the complement.  The  $init$ element is used as the
-initial value of $C$ and specifies default behavior when data is missing.
-For instance, to implement the
-scenario in Figure~\ref{fig:minimized-representations}, the complement
-would consist of a list of pairs of salary and company name.
-
+A classical symmetric lens $\Lens$ between $X$ and $Y$ consists of 4 components:
+a complement $C$, a designated element $init \in C$, and two functions,
+$\PutRSym \OfType X \times C \rightarrow Y \times C$ and $\PutLSym \OfType Y
+\times C \rightarrow X \times C$, that propagate changes in one format to the
+other.
+
+In this formulation, data unique to each side are stored in the complement. When
+one format is edited, the \PutR or \PutL function stitches together the edited
+data with data stored in the complement. The $init$ element is the initial value
+of $C$ and specifies default behavior when data is missing. For instance, to
+implement the scenario in Figure~\ref{fig:minimized-representations}, the
+complement would consist of a list of pairs of salary and company name.
 Classical symmetric lenses satisfy the following equational laws.

### 2018-07-10 papers/SymmetricLenses/main.tex
add edits to remove horrible dnf lenses

@@
 Symmetric DNF lenses are an intermediate synthesis target for \GreedySynth.
-Because of the lack of a type equivalence rule or a composition lens, the set of
-possible lenses with type $\overline{\DNFRegex} \Leftrightarrow
-\overline{\DNFRegexAlt}$ is quite small. Consequently, an algorithm for
-synthesizing such lenses can be driven primarily by the type of the lenses.
-
-\paragraph*{Syntax}
-Roughly speaking, symmetric DNF lenses are an n-ary union of symmetric sequence
-lenses, which are, in turn, an n-ary sequence of atomic lenses. Atomic lenses
-are iterations of symmetric DNF lenses. In the following syntax, the
-nonterminals $i$, $j$, $p$, $q$, $r$, $c$, and $d$ all represent natural
-numbers. The nonterminals $\String$ and $\StringAlt$ represent strings.
-\begin{center}
-  \begin{tabular}{@{}r@{\ }c@{}l@{}}
-    % REGEX
-    \SAtomLens{} & \GEq{} & $\StarOf{\SDNFLens}$ \\
-    \SSQLens{} & \GEq{} & $(\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
-                          \ldots\SeqLSep
-                          (i_p,j_p,\SAtomLens_p)}
-                          ,\ListOf{(\String_1,\Atom_1);\ldots;(\String_q,\Atom_q)}
-                          ,\ListOf{(\StringAlt_1,\AtomAlt_1);\ldots;(\StringAlt_r,\AtomAlt_r)})$ \\
-    \SDNFLens{} & \GEq{} & $(\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
-                           \ldots\DNFLSep
-                           (i_p,j_p,\SSQLens_p)}
-                           ,\ListOf{c_1;\ldots;c_q}
-                           ,\ListOf{d_1;\ldots;d_r})$ \\
-  \end{tabular}
-\end{center}
-
-\saz{The following two paragraphs are pretty similar (was there a bad merge at
-  some point?) and very low level.  I also
-  could not really follow either of the examples at all.  If we
-  are looking for things to cut, I would propose dropping most of the content of
-  the next two paragraphs, merging the remainder, and skipping pretty directly to the paragraph
-  that starts with ``While DNF lenses...''}
-As an example, consider the symmetric DNF lens:
-$(\SDNFLensOf{(1,1,\SSQLens_1)\DNFLSep (1,2,\SSQLens_2)}
-,\ListOf{2},\ListOf{1;2})$, synchronizing data between $\DNFOf{\Sequence_1}$ and
-$\DNFOf{\SequenceAlt_1 \DNFSep \SequenceAlt_2}$. This lens merges both
-$\SequenceAlt_1$ and $\SequenceAlt_2$ into $\Sequence_1$, as they both are
-involved in a lens with $\Sequence_1$. When performing a \CreateR, $\SSQLens_2$
-is used. When performing a \CreateL on a string in $\Sequence_1$, $\SSQLens_1$
-is used, and when performing a \CreateL on a string in $\Sequence_2$,
-$\SSQLens_2$ is used. More generally, a symmetric DNF Lens \SDNFLens{} consists
-of three components. First is the mapping component. The mapping component is a
-list of triples $(i,j,\SSQLens)$. In this, $i$ and $j$ refer to sequence $i$ and
-sequence $j$ in the left and right DNF regular expressions, respectively. The
-third component of the triple, \SSQLens{}, is a sequence lens that maps between
-the two sequences specified by $i$ and $j$. The second component is a list of
-integers $c$. The integer, $c_i$ at position $i$ states that when performing a
-$\CreateR$ on a string matching sequence $\Sequence_i$, it will use the sequence
-lens that maps it to sequence $\SequenceAlt_{c_i}$ on the right. The third
-component is a list of integers $d$ that follows the same pattern as the second
-component, but maps from right to left.
-
-As an example, consider the symmetric sequence lens:
-$(\SSQLensOf{(1,2,\SAtomLens_1)} ,\ListOf{}
-,\ListOf{(\StringAlt_1',\AtomAlt_1)})$, synchronizing data between
-$\SequenceOf{\String_0 \SeqSep \Atom_1 \SeqSep \String_1}$ and
-$\SequenceOf{\StringAlt_0 \SeqSep \AtomAlt_1 \SeqSep \StringAlt_1 \SeqSep
-  \AtomAlt_2 \SeqSep \StringAlt_2}$. This lens maps $\Sequence_1$ to
-$\SequenceAlt_2$, and leaves $\AtomAlt_1$ disconnected. On a \CreateR, the
-string created is a concatenation of $\StringAlt_1'$ with the default from
-$\SAtomLens_1$. More generally, a sequence lens \SSQLens{} consists of three
-components. First is the mapping component. The mapping component is a list of
-triples $(i,j,\SAtomLens)$. In this, $i$ and $j$ refer to atom $i$ and atom $j$
-on the left and right sequences, respectively. The third component of the
-triple, \SAtomLens{}, is an atom lens that maps between the two atoms specified
-by $i$ and $j$. The second component is a list of strings, $\String$. The
-string, $s_i$ at position $i$ states that when performing a $\CreateL$ on a
-string matching atom $\Atom_i$, it will do one of two things. If $\Atom_i$ is
-mapped to in the mapping component, then performing the atom's $\CreateL$ will
-actually $\PutL$ into that string. If $\Atom_i$ is not mapped to, then
-$\String_i$ will be created as a representative of $\Atom_i$.
-
-While DNF lenses are useful because they permit an efficient greedy synthesis
-procedure, the particulars of their typing and semantics are not the primary
-intellectual contributions of our work. The details of the typing and semantics
-are included in the supplementary material present in the appendix.
-
-\paragraph*{Information-Theoretic Cost Metric}
-Much like we defined a syntactic means to find the expected number of bits to
-recover string from a synchronized string in another format for simple symmetric
-string lenses, we do the same for symmetric DNF lenses. Unlike simple symmetric
-string lenses, DNF lenses do not have composition, so the entropy can be defined
-over all terms. $\REntropyOf{\DNFRegex \Given \SDNFLens, \DNFRegexAlt}$ is the
-expected number of bits required to recover a string in $\DNFRegex$ from a
-synchronized string in $\DNFRegexAlt$.
-
-\[
-  \begin{array}{rl}
-    & \REntropyOf{\DNFOf{\Sequence_1 \DNFSep \ldots \DNFSep \Sequence_n}\\
-    & \hspace*{1.21em}\Given
-      (\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
-      \ldots\DNFLSep
-      (i_p,j_p,\SSQLens_p)}
-      ,\ListOf{c_1;\ldots;c_q}
-      ,\ListOf{d_1;\ldots;d_r})\\
-    & \hspace*{1.21em},~\DNFOf{\SequenceAlt_1 \DNFSep \ldots \DNFSep \SequenceAlt_m}}\\
-    = & \frac{\Sigma_{j=1}^m(H_j)}{m}\\
-    & \text{where } H_j = (0,\Sigma_{\SetOf{k\SuchThat j_k = j}}
-      \REntropyOf{\Sequence_{i_k}\Given\SequenceLens_k,\SequenceAlt_j})
-  \end{array}
-\]
-
-This entropy calculation for symmetric DNF lenses requires a similar entropy
-calculation for symmetric sequence lenses and symmetric atom lenses.
-
-\[
-  \begin{array}{rl}
-      & \REntropyOf{
-        \SequenceOf{\String_0,\Atom_1,\ldots,\Atom_n,\String_n}\\
-      & \hspace*{1em}\Given
-        (\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
-        \ldots\SeqLSep
-        (i_p,j_p,\SAtomLens_p)}
-        ,\ListOf{(k_1,\String_1);\ldots;(k_q,\String_q)}
-        ,\ListOf{(l_1,\StringAlt_1);\ldots;(l_r,\StringAlt_r)})\\
-      & \hspace*{1.21em},~
-        \SequenceOf{\StringAlt_0,\AtomAlt_1,\ldots,\AtomAlt_m,\StringAlt_m}}\\
-      =
-      & \Sigma_{x=1}^p\REntropyOf{\Atom_{i_x} \Given \AtomAlt_{i_x},\AtomLens_x} +
-        \Sigma_{x=1}^q\REntropyOf{\Atom_{k_x}}
-  \end{array}
-\]
-
-The entropy calculation for symmetric sequence lenses requires a similar entropy
-calculation for symmetric atom lenses, which in turn relies on the entropy
-calculation for symmetric DNF lenses.
-
-\[\begin{array}{rl}
-      & \REntropyOf{\PRegexStar{\DNFRegexAlt}{\ProbabilityAlt} \Given \PRegexStar{\DNFRegex}{\Probability},\StarOf{\SDNFLens}}\\
-      =
-      & \frac{\Probability}{1-\Probability}\REntropyOf{\DNFRegexAlt\Given\DNFRegex,\SDNFLens}
-  \end{array}
-\]
-The expected number of bits required to recover a string in $\DNFRegexAlt$ from
-a synchronized string in $\DNFRegex$, $\REntropyOf{\DNFRegexAlt \Given
-  \SDNFLens, \DNFRegex}$ is defined symmetrically.
+There are many fewer symmetric DNF lenses than symmetric regular lenses. In
+fact, if one does not use the star-unrolling axioms, there are only finitely
+many DNF lenses of a given type (though there are still many more symmetric DNF
+lenses than there are DNF bijective lenses).
+
+Intuitively, a symmetric DNF lens $sdl$ is a union of symmetric sequence lenses,
+$ssql_1 \ldots ssql_p$, a symmetric DNF sequence lens is a concatenation of
+symmetric atom lenses, $sal_1 \ldots sal_n$, and a symmetric atom lens $sal$ is
+an iteration of a symmetric DNF lens.
+
+Just as we analyzed the information content of ordinary regular expressions, we
+can analyze the information content of DNF regular expressions. As before, we
+use $\REntropyOf{\DNFRegexAlt \Given \SDNFLens, \DNFRegex}$ to calculate bounds
+of the expected amount of information required to recover a string in
+$\DNFRegexAlt$ from a string in $\DNFRegex$, synchronized by $\SDNFLens$. We use
+the function $\LEntropyOf{\DNFRegex \Given \SDNFLens, \DNFRegexAlt}$ to
+calculate bounds of the expected amount of information required to recover a
+string in $\DNFRegex$ from a string in $\DNFRegexAlt$, synchronized by
+$\SDNFLens$.
+
+The details of these definitions are syntactically tedious, but not
+intellectually difficult. We elide them here but include them in the appendix.
 
@@
 Classical symmetric lenses satisfy the following equational laws.
-\begin{equation}
-  \tag{\PutRL}
-  \frac{\PutRSymOf{(x,c)} = (y,c'')}{\PutLSymOf{(y,c'''')} = (x,c')}
-\end{equation}
-\begin{equation}
-  \tag{\PutLR}
-  \frac{\PutLSymOf{(y,c)} = (x,c')}{\PutRSymOf{(x,c'')} = (y,c')}
-\end{equation}
+\begin{centermath}
+\begin{array}[b]{l@{\qquad \; \qquad}l}
+  \begin{mathprooftree}
+    \AxiomC{$\PutRSymOf{(x,c)} = (y,c')$}
+    \UnaryInfC{$\PutLSymOf{(y,c')} = (x,c')$}
+  \end{mathprooftree}
+&
+  \begin{mathprooftree}
+    \AxiomC{$\PutLSymOf{(y,c)} = (x,c')$}
+    \UnaryInfC{$\PutRSymOf{(x,c')} = (y,c')$}
+  \end{mathprooftree}
+\end{array}
+\end{centermath}
 
@@
 
-\paragraph*{Symmetric DNF Lenses: Typing and Semantics}
+\paragraph*{Symmetric DNF Lenses: Syntax, Cost, Typing, and Semantics}
+
+\paragraph*{Syntax}
+Roughly speaking, symmetric DNF lenses are an n-ary union of symmetric sequence
+lenses, which are, in turn, an n-ary sequence of atomic lenses. Atomic lenses
+are iterations of symmetric DNF lenses. In the following syntax, the
+nonterminals $i$, $j$, $p$, $q$, $r$, $c$, and $d$ all represent natural
+numbers. The nonterminals $\String$ and $\StringAlt$ represent strings.
+\begin{center}
+  \begin{tabular}{@{}r@{\ }c@{}l@{}}
+    % REGEX
+    \SAtomLens{} & \GEq{} & $\StarOf{\SDNFLens}$ \\
+    \SSQLens{} & \GEq{} & $(\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
+                          \ldots\SeqLSep
+                          (i_p,j_p,\SAtomLens_p)}
+                          ,\ListOf{(\String_1,\Atom_1);\ldots;(\String_q,\Atom_q)}
+                          ,\ListOf{(\StringAlt_1,\AtomAlt_1);\ldots;(\StringAlt_r,\AtomAlt_r)})$ \\
+    \SDNFLens{} & \GEq{} & $(\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
+                           \ldots\DNFLSep
+                           (i_p,j_p,\SSQLens_p)}
+                           ,\ListOf{c_1;\ldots;c_q}
+                           ,\ListOf{d_1;\ldots;d_r})$ \\
+  \end{tabular}
+\end{center}
+
+\paragraph*{Information-Theoretic Cost Metric}
+Much like we defined a syntactic means to find the expected number of bits to
+recover string from a synchronized string in another format for simple symmetric
+string lenses, we do the same for symmetric DNF lenses. Unlike simple symmetric
+string lenses, DNF lenses do not have composition, so the entropy can be defined
+over all terms. $\LEntropyOf{\DNFRegex \Given \SDNFLens, \DNFRegexAlt}$ is the
+expected number of bits required to recover a string in $\DNFRegex$ from a
+synchronized string in $\DNFRegexAlt$.
+
+\[
+  \begin{array}{rl}
+    & \LEntropyOf{\DNFOf{\Sequence_1 \DNFSep \ldots \DNFSep \Sequence_n}\\
+    & \hspace*{1.21em}\Given
+      (\SDNFLensOf{(i_1,j_1,\SSQLens_1)\DNFLSep
+      \ldots\DNFLSep
+      (i_p,j_p,\SSQLens_p)}
+      ,\ListOf{c_1;\ldots;c_q}
+      ,\ListOf{d_1;\ldots;d_r})\\
+    & \hspace*{1.21em},~\DNFOf{\SequenceAlt_1 \DNFSep \ldots \DNFSep \SequenceAlt_m}}\\
+    = & \frac{\Sigma_{j=1}^m(H_j)}{m}\\
+    & \text{where } H_j = (0,\Sigma_{\SetOf{k\SuchThat j_k = j}}
+      \LEntropyOf{\Sequence_{i_k}\Given\SequenceLens_k,\SequenceAlt_j})
+  \end{array}
+\]
+
+This entropy calculation for symmetric DNF lenses requires a similar entropy
+calculation for symmetric sequence lenses and symmetric atom lenses.
+
+\[
+  \begin{array}{rl}
+      & \REntropyOf{
+        \SequenceOf{\String_0,\Atom_1,\ldots,\Atom_n,\String_n}\\
+      & \hspace*{1em}\Given
+        (\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
+        \ldots\SeqLSep
+        (i_p,j_p,\SAtomLens_p)}
+        ,\ListOf{(k_1,\String_1);\ldots;(k_q,\String_q)}
+        ,\ListOf{(l_1,\StringAlt_1);\ldots;(l_r,\StringAlt_r)})\\
+      & \hspace*{1.21em},~
+        \SequenceOf{\StringAlt_0,\AtomAlt_1,\ldots,\AtomAlt_m,\StringAlt_m}}\\
+      =
+      & \Sigma_{x=1}^p\REntropyOf{\Atom_{i_x} \Given \AtomAlt_{i_x},\AtomLens_x} +
+        \Sigma_{x=1}^q\REntropyOf{\Atom_{k_x}}
+  \end{array}
+\]
+
+The entropy calculation for symmetric sequence lenses requires a similar entropy
+calculation for symmetric atom lenses, which in turn relies on the entropy
+calculation for symmetric DNF lenses.
+
+\[\begin{array}{rl}
+      & \REntropyOf{\PRegexStar{\DNFRegexAlt}{\ProbabilityAlt} \Given \PRegexStar{\DNFRegex}{\Probability},\StarOf{\SDNFLens}}\\
+      =
+      & \frac{\Probability}{1-\Probability}\REntropyOf{\DNFRegexAlt\Given\DNFRegex,\SDNFLens}
+  \end{array}
+\]
+The expected number of bits required to recover a string in $\DNFRegexAlt$ from
+a synchronized string in $\DNFRegex$, $\REntropyOf{\DNFRegexAlt \Given
+  \SDNFLens, \DNFRegex}$ is defined symmetrically.
+
 The typing judgment is a 3-ary relation over a single DNF lens, and two DNF
@@
 \[
-  \inferrule*
-  {
+  \inferrule* {
     \SAtomLens_1 \OfType \Atom_{i_1} \Leftrightarrow \AtomAlt_{j_1}\\
@@
     \SAtomLens_p \OfType \Atom_{i_p} \Leftrightarrow \AtomAlt_{j_p}\\\\
-    \String_1 \in \Atom_1\\
+    \String_{c_1} \in \Atom_{c_1}\\
     \ldots\\
-    \String_q \in \Atom_q\\\\
-    \StringAlt_1 \in \AtomAlt_1\\
+    \String_{c_{q'}} \in \Atom_{c_{q'}}\\\\
+    \StringAlt_{d_1} \in \AtomAlt_{d_1}\\
     \ldots\\
-    \StringAlt_r \in \AtomAlt_r\\\\
+    \StringAlt_{d_{r'}} \in \AtomAlt_{d_{r'}}\\\\
     i_x = i_y \BooleanImplies x = y\\
-    j_x = j_y \BooleanImplies x = y
-  }
-  {
-    (\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep
-      \ldots\SeqLSep
+    j_x = j_y \BooleanImplies x = y } {
+    (\SSQLensOf{(i_1,j_1,\SAtomLens_1)\SeqLSep \ldots\SeqLSep
       (i_p,j_p,\SAtomLens_p)}
-    ,\ListOf{\String_1;\ldots;\String_q}
-    ,\ListOf{\StringAlt_1;\ldots;\StringAlt_r})\\
-    \OfType
-    \SequenceOf{\String_0'''' \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_q
-      \SeqSep \String_q'}
-    \Leftrightarrow
-    \SequenceOf{\StringAlt_0' \SeqSep \AtomAlt_1 \SeqSep \ldots \SeqSep \AtomAlt_r \SeqSep \StringAlt_r'}
-  }
+    ,\ListOf{(\String_{c_1},\Atom_{c_1});\ldots;(\String_{c_{q'}},\Atom_{c_{q'}})}
+    ,\ListOf{(\StringAlt_{d_1},\Atom_{d_1});\ldots;(\StringAlt_{d_r},\Atom_{d_r})})\\
+    \OfType \SequenceOf{\String_0'''' \SeqSep \Atom_1 \SeqSep \ldots \SeqSep
+      \Atom_{q'} \SeqSep \String_{q}'} \Leftrightarrow \SequenceOf{\StringAlt_0'
+      \SeqSep \AtomAlt_1 \SeqSep \ldots \SeqSep \AtomAlt_r \SeqSep
+      \StringAlt_{r}'} }
 \]

### 2018-07-10 papers/SymmetricLenses/main.tex
in progress edits to section 1

@@
 intuition, we turn to \emph{stochastic regular expressions}
-(SREs)~\cite{?}, which simultaneously define a set of strings \`a la
+(SREs)~\cite{stoch-rnn,stoch-def}, which simultaneously define a set of strings \`a la
 regular expressions and a probability distribution over those
@@
     When the algorithm determines it has found a likely lens, it
-  returns it.  \ksf{can we change the subscript on the returned lens
-  to ``likely'' instead of ``best''? Also, is it the cost that is
-  returned or the likeliness }
+  returns it.  \ksf{Can we change cost to likeliness?}
    }
@@
 We implemented the \GreedySynth and \Expand algorithms and evaluated their
-effectiveness of synthesizing synchronization functions on a range of
-benchmarks, including data cleaning tasks taken from Flash Fill, information
-extraction tasks taken from Augeas, and ad-hoc format synchronization tasks
-taken from a variety of locations\footnote{The majority of these format
+effectiveness at synthesizing lenses on a range of
+benchmarks, including data cleaning tasks taken from Flash Fill~\cite{gulwani-popl-2014}, information
+extraction tasks taken from Augeas~\cite{Augeas}, and ad-hoc format synchronization tasks
+taken from a variety of other locations\footnote{The majority of these format
   synchronization tasks were found from the following index:
@@
 \begin{itemize}
-\item We define the space of simple symmetric lenses, which lies between
-  asymmetric lenses and full symmetric lenses and has the useful property that
-  its members don't have any internal state (\S\ref{sec:overview}). We describe
-  a language for simple symmetric string lenses and the cost metric for that
-  language (\S\ref{sec:ssl}).
+\item We define simple symmetric lenses, which 
+  have the useful property that they do not maintain any hidden state
+  (\S\ref{sec:overview}).  
+  We describe a language for defining them and describe how to
+  calculate the likelihood of a  particular lens (\S\ref{sec:ssl}). 
+  We show simple symmetric lenses
+  fall between asymmetric lenses~\cite{Focal2005-long2} and full
+  symmetric lenses~\cite{symmetric-lenses} in expressiveness (\S\ref{sec:related}). 
 \item We show how to synthesize simple symmetric lenses (including all
@@
   regular expressions to guide the search process (\S\ref{sec:synthesis}). In
-  particular, we show that the RE star semi-ring equivalences remain valid when
+  particular, we show that the star semi-ring equivalences on regular
+  expressions remain valid
   their semantics are enriched by distributions (\S\ref{sec:sre}).

### 2018-07-10 papers/SymmetricLenses/main.tex
finished pass over intro

@@
 Hence, if the synthesis algorithm can
-find any bijection at all, then it is very likely to be the one the
+find any bijection at all, then it is very likely to find the one the
 user wants.
@@
 common, but also contain some differences. For instance, two related
-tables might share four of five columns, but each have a different
+tables might share four of five columns, but each has a different
 fifth column. Alternately, one ad hoc configuration file might include
-some file-specific meta data, such as a date, a time or reference
+some file-specific meta data, such as a date, a time or a reference
 number, whereas the other does not. Indeed, all of our benchmarks
@@
 those symmetric lenses that satisfy an intuitive property called
-\emph{forgetfulness}, which we define in Section~\ref{sec:relationship}.
+\emph{forgetfulness}.
+%, which we define in Section~\ref{sec:relationship}.
 Intuitively, simple symmetric lenses are the complete set of symmetric
@@
 We implemented the \GreedySynth and \Expand algorithms and evaluated their
-effectiveness at synthesizing lenses on a range of
-benchmarks, including data cleaning tasks taken from Flash Fill~\cite{gulwani-popl-2014}, information
-extraction tasks taken from Augeas~\cite{Augeas}, and ad-hoc format synchronization tasks
+effectiveness at synthesizing lenses on a 
+benchmark suite of 45 problems, including data cleaning tasks taken
+from Flash Fill~\cite{gulwani-popl-2014}, information 
+extraction tasks \ksf{Dave wants to change information extraction to
+soemthing else.} taken from Augeas~\cite{Augeas}, and ad-hoc format synchronization tasks
 taken from a variety of other locations\footnote{The majority of these format
@@
 \begin{itemize}
-\item We define simple symmetric lenses, which 
-  have the useful property that they do not maintain any hidden state
+\item We define simple symmetric lenses, which have
+  the useful property that they do not maintain any hidden state
   (\S\ref{sec:overview}).  
-  We describe a language for defining them and describe how to
-  calculate the likelihood of a  particular lens (\S\ref{sec:ssl}). 
-  We show simple symmetric lenses
+  We give a language for defining such lenses and describe how to
+  calculate their likelihoods (\S\ref{sec:ssl}). 
+\item  We show that simple symmetric lenses
   fall between asymmetric lenses~\cite{Focal2005-long2} and full
-  symmetric lenses~\cite{symmetric-lenses} in expressiveness (\S\ref{sec:related}). 
+  symmetric lenses~\cite{symmetric-lenses} in expressiveness
+  (\S\ref{sec:related}).  We further show that the class of simple
+  lenses can be characterized semantically as the subset of full
+  symmetric lenses for which a simple ``forgetfulness'' property holds
+  (\S\ref{sec:relationship}).
 \item We show how to synthesize simple symmetric lenses (including all
   asymmetric lenses as a special case), using a novel application of stochastic
-  regular expressions to guide the search process (\S\ref{sec:synthesis}). In
-  particular, we show that the star semi-ring equivalences on regular
-  expressions remain valid
-  their semantics are enriched by distributions (\S\ref{sec:sre}).
-\item We extend the Boomerang implementation with simple symmetric lenses,
-  showing that they can easily be combined with a rich variety of other lens
-  extensions including quotients and matching lenses. We evaluate our
-  implementation over a set of benchmarks drawn from data cleaning, information
-  extraction, and file synchronization tasks and demonstrate that we can
-  synthesize all of them in under half a minute (\S\ref{sec:evaluation}).
-%
-  We also analyze the gap between simple and full symmetric lenses
-  empirically and argue that simple ones suffice for many applications.
-\item We show that the class of simple lenses can be characterized semantically
-  as the subset of full symmetric lenses for which a simple ``forgetfulness''
-  property holds (\S\ref{sec:relationship}).
+  regular expressions to guide the search process (\S\ref{sec:synthesis}). 
+  To support that effort, we show that the star semi-ring equivalences on regular
+  expressions remain valid when applied to stochastic regular
+  expressions, preserving the probability distributions as well as the languages
+  (\S\ref{sec:sre}).
+\item We extend Boomerang with simple symmetric lenses,
+  showing that such lenses can integrate smoothly with 
+  other lens extensions such as quotient lenses and matching
+  lenses.  We evaluate our
+  implementation on a set of 45 lens synthesis benchmarks drawn from
+  data cleaning, \ksf{Dave wants to change information extraction to
+soemthing else.} information 
+  extraction, and file synchronization tasks.  We demonstrate that we can
+  synthesize simple symmetric lenses for all of the benchmarks in
+  under 30 seconds (\S\ref{sec:evaluation}). 
 \end{itemize}
 We finish the paper with a discussion of related work (\S\ref{sec:related}) and
-our final thoughts (\S\ref{sec:conc}).
+our concluding thoughts (\S\ref{sec:conc}).
 

### 2018-07-10 papers/SymmetricLenses/main.tex
finished pass over abstract

@@
 Synthesizing simple symmetric lenses is substantially more challenging that
-synthesizing bijective lenses: since some of the information on each side
+synthesizing bijective lenses: Since some of the information on each side
 can be ``disconnected'' from the other side, there will in general be {\em
@@
 We describe an implementation of simple symmetric lenses and our synthesis
-procedure as extensions to the Boomerang language~\cite{Boomerang} and evaluate
-its performance on a range of benchmark examples. We find this implementation is
-able to synthesize all the lenses of this benchmark suite in under 30 seconds.
+procedure as extensions to the Boomerang language~\cite{Boomerang}. We
+evaluate
+its performance on 45 benchmark examples drawn from FlashFill, Augeus,
+and an online collection of electronic file formats. We find this implementation is
+able to synthesize all the lenses of the benchmark suite in under 30 seconds.
 \end{abstract}

### 2018-07-10 papers/SymmetricLenses/main.tex
add some minimizations

--- 2018-07-10 papers/SymmetricLenses/main.tex
@@
 intuition, we turn to \emph{stochastic regular expressions}
-(SREs)~\cite{stoch-rnn,stoch-def}, which simultaneously define a set of strings \`a la
+(SREs)~\cite{?}, which simultaneously define a set of strings \`a la
 regular expressions and a probability distribution over those
@@
     When the algorithm determines it has found a likely lens, it
-  returns it.  \ksf{Can we change cost to likeliness?}
+  returns it.  \ksf{can we change the subscript on the returned lens
+  to ``likely'' instead of ``best''? Also, is it the cost that is
+  returned or the likeliness }
    }
@@
 We implemented the \GreedySynth and \Expand algorithms and evaluated their
-effectiveness at synthesizing lenses on a range of
-benchmarks, including data cleaning tasks taken from Flash Fill~\cite{gulwani-popl-2014}, information
-extraction tasks taken from Augeas~\cite{Augeas}, and ad-hoc format synchronization tasks
-taken from a variety of other locations\footnote{The majority of these format
+effectiveness of synthesizing synchronization functions on a range of
+benchmarks, including data cleaning tasks taken from Flash Fill, information
+extraction tasks taken from Augeas, and ad-hoc format synchronization tasks
+taken from a variety of locations\footnote{The majority of these format
   synchronization tasks were found from the following index:
@@
 \begin{itemize}
-\item We define simple symmetric lenses, which 
-  have the useful property that they do not maintain any hidden state
-  (\S\ref{sec:overview}).  
-  We describe a language for defining them and describe how to
-  calculate the likelihood of a  particular lens (\S\ref{sec:ssl}). 
-  We show simple symmetric lenses
-  fall between asymmetric lenses~\cite{Focal2005-long2} and full
-  symmetric lenses~\cite{symmetric-lenses} in expressiveness (\S\ref{sec:related}). 
+\item We define the space of simple symmetric lenses, which lies between
+  asymmetric lenses and full symmetric lenses and has the useful property that
+  its members don't have any internal state (\S\ref{sec:overview}). We describe
+  a language for simple symmetric string lenses and the cost metric for that
+  language (\S\ref{sec:ssl}).
 \item We show how to synthesize simple symmetric lenses (including all
@@
   regular expressions to guide the search process (\S\ref{sec:synthesis}). In
-  particular, we show that the star semi-ring equivalences on regular
-  expressions remain valid
+  particular, we show that the RE star semi-ring equivalences remain valid when
   their semantics are enriched by distributions (\S\ref{sec:sre}).
@@
 
-We now provide the typing derivations and semantics for these lenses. For
-simplicity of presentation, we define the typing derivations and semantics only
-over unambiguous regular expressions, even though some lenses can be defined
-over ambiguous regular expressions (for example, the identity lens doesn't need
-any ambiguity constraints). We believe relaxing this restriction is
+We now provide the typing derivations and semantics for these lenses. We do not
+include the semantics for combinators equivalent to those presented in prior
+work~\cite{symmetric-lenses,boomerang}. Furthermore, we only present \CreateR
+and \PutR when \CreateL and \PutL are symmetric. Full details are present in the
+appendix. For simplicity of presentation, we define the typing derivations and
+semantics only over unambiguous regular expressions, even though some lenses can
+be defined over ambiguous regular expressions (for example, the identity lens
+doesn't need any ambiguity constraints). We believe relaxing this restriction is
 unproblematic.
@@
     \CreateR{} \App s & = & s\\
-    \CreateL{} \App s & = & s\\
     \PutR{} \App s_1 \App s_2 & = & s_1\\
-    \PutL{} \App s_1 \App s_2 & = & s_1
   \end{array}
@@
     $\CreateR{} \App \String'$ & = & $\StringAlt$\\
-    $\CreateL{} \App \StringAlt'''''$ & = & $\String$\\
     $\PutR{} \App \String''' \App \StringAlt'$ & = & $\StringAlt'$\\
-    $\PutL{} \App \StringAlt' \App \String'$ & = & $\String'$
   \end{array}
@@
 
-    $\CreateL{} \App \StringAlt$
-    & =
-    & $\begin{cases*}
-      \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1}$\\
-      \Lens_2.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2}$
-      \end{cases*}$\\
-
     $\PutR{} \App \String \App \StringAlt$
@@
         \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_2} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_1}$
-      \end{cases*}$\\
-
-    $\PutL{} \App \StringAlt \App \String$
-    & =
-    & $\begin{cases*}
-        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1} \BooleanAnd \String\in\LanguageOf{\BRegex_1}$\\
-        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2} \BooleanAnd \String\in\LanguageOf{\BRegex_2}$\\
-        \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1} \BooleanAnd \String\in\LanguageOf{\BRegex_2}$\\
-        \Lens_2.\CreateLOf{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2} \BooleanAnd \String\in\LanguageOf{\BRegex_1}$
       \end{cases*}$\\
@@
 \begin{centermath}
-\begin{tabular}[b]{l@{\qquad}l}
-$
   \inferrule*
@@
   }
-$
-&
-  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
-    $\CreateR{} \App \String$
-    & =
-    & $\Lens_1.\CreateROf{\String}$\\
-
-    $\CreateL{} \App \StringAlt$
-    & =
-    & $\begin{cases*}
-      \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1}$\\
-      \Lens_2.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2}$
-      \end{cases*}$\\
-
-    $\PutR{} \App \String \App \StringAlt$
-    & =
-    & $\begin{cases*}
-      \Lens_1.\PutROf{\String}{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1}$\\
-      \Lens_2.\PutROf{\String}{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2}$
-    \end{cases*}$\\
-
-    $\PutL{} \App \StringAlt \App \String$
-    & =
-    & $\begin{cases*}
-        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1}$\\
-        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2}$
-      \end{cases*}$\\
-  \end{tabular}
-\end{tabular}
 \end{centermath}
@@
 \onecolumn
+
+\paragraph*{Symmetric Lenses full Detail}
+\begin{centermath}
+\begin{array}[b]{l@{\qquad \; \qquad}l}
+\begin{array}[b]{c}
+\quad\\
+  \inferrule*
+  {
+  }
+  {
+    \IdentityLensOf{\BRegex} \OfType \BRegex \Leftrightarrow \BRegex
+  }
+  \\
+  \quad
+  \end{array}
+&
+\begin{array}[b]{r@{\ }c@{\ }l}
+    \CreateR{} \App s & = & s\\
+    \CreateL{} \App s & = & s\\
+    \PutR{} \App s_1 \App s_2 & = & s_1\\
+    \PutL{} \App s_1 \App s_2 & = & s_1\\
+  \end{array}
+\end{array}
+\end{centermath}
+Note that the identity lens ignores the second argument in the put functions.
+Because the two formats are fully synchronized, no knowledge of the prior data
+is needed.
+
+\begin{centermath}
+\begin{array}[b]{l@{\qquad \; \qquad}l}
+  \inferrule*
+  {
+    \String \in \LanguageOf{\BRegex}\\
+    \StringAlt \in \LanguageOf{\BRegexAlt}
+  }
+  {
+    \DisconnectOf{\BRegex}{\BRegexAlt}{\String}{\StringAlt}
+    \OfType \BRegex \Leftrightarrow \BRegexAlt
+  }
+&
+  \begin{array}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App \String'$ & = & $\StringAlt$\\
+    $\CreateL{} \App \StringAlt'''''$ & = & $\String$\\
+    $\PutR{} \App \String''' \App \StringAlt'$ & = & $\StringAlt'$\\
+    $\PutL{} \App \StringAlt' \App \String'$ & = & $\String'$
+  \end{array}
+\end{array}
+\end{centermath}
+
+Just as the identity lens ignores the second argument in puts, disconnect lenses
+ignore the first in both puts and creates.  The data is unsynchronized in these
+two formats, information from one format doesn't impact the other.
+
+\begin{centermath}
+\begin{tabular}[b]{l@{\qquad \; \qquad}l}
+$
+  \inferrule*
+  {
+    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt_1\\
+    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt_2
+  }
+  {
+    \ConcatLensOf{\Lens_1}{\Lens_2} \OfType \BRegex_1 \Concat \BRegex_2
+    \Leftrightarrow
+    \BRegexAlt_1 \Concat \BRegexAlt_2
+  }
+$
+&
+$
+  \inferrule*
+  {
+    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt_1\\
+    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt_2
+  }
+  {
+    \SwapLensOf{\Lens_1}{\Lens_2} \OfType \BRegex_1 \Concat \BRegex_2
+    \Leftrightarrow
+    \BRegexAlt_2 \Concat \BRegexAlt_1
+  }
+$
+\end{tabular}
+\end{centermath}
+\begin{center}
+  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App \String_1 \String_2$ & = & $(\Lens_1.\CreateROf{\String_1})(\Lens_2.\CreateROf{\String_2})$\\
+    $\CreateL{} \App \StringAlt_1 \StringAlt_2$ & = & $(\Lens_1.\CreateLOf{\StringAlt_1})(\Lens_2.\CreateLOf{\StringAlt_2})$\\
+    $\PutR{} \App (\String_1\String_2) \App (\StringAlt_1\StringAlt_2)$ & = & $(\Lens_1.\PutROf{\String_1}{\StringAlt_1})(\Lens_2.\PutROf{\String_2}{\StringAlt_2})$\\
+    $\PutL{} \App (\StringAlt_1\StringAlt_2) \App (\String_1\String_2)$ & = & $(\Lens_1.\PutLOf{\StringAlt_1}{\String_1})(\Lens_2.\PutLOf{\StringAlt_2}{\String_2})$\\
+  \end{tabular}
+\end{center}
+
+Concat is similar to concatenation in existing string lens languages like
+Boomerang.  For such terms, we do not provide the semantics, and merely refer readers to existing work. The swap combinator is similar to concat, though the second regular expression
+is swapped.
+\begin{center}
+  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App \String_1\String_2$ & = & $(\Lens_2.\CreateROf{\String_2})(\Lens_1.\CreateROf{\String_1})$\\
+    $\CreateL{} \App \StringAlt_2\StringAlt_1$ & = & $(\Lens_1.\CreateLOf{\StringAlt_1})(\Lens_2.\CreateLOf{\StringAlt_2})$\\
+    $\PutR{} \App (\String_1\String_2) \App (\StringAlt_2\StringAlt_1)$ & = & $(\Lens_2.\PutROf{\String_2}{\StringAlt_2})(\Lens_1.\PutROf{\String_1}{\StringAlt_1})$\\
+    $\PutL{} \App (\StringAlt_2\StringAlt_1) \App (\String_1\String_2)$ & = & $(\Lens_1.\PutLOf{\StringAlt_1}{\String_1})(\Lens_2.\PutLOf{\StringAlt_2}{\String_2})$\\
+  \end{tabular}
+\end{center}
+\begin{centermath}
+\begin{tabular}[b]{l@{\qquad}l}
+$
+  \inferrule*
+  {
+    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt_1\\
+    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt_2
+  }
+  {
+    \OrLensOf{\Lens_1}{\Lens_2} \OfType
+    \RegexOr{\BRegex_1}{\BRegex_2}
+    \Leftrightarrow
+    \RegexOr{\BRegexAlt_1}{\BRegexAlt_2}
+  }
+  $
+&
+  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App \String$
+    & =
+    & $\begin{cases*}
+      \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_1}$\\
+      \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_2}$
+      \end{cases*}$\\
+
+    $\CreateL{} \App \StringAlt$
+    & =
+    & $\begin{cases*}
+      \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1}$\\
+      \Lens_2.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2}$
+      \end{cases*}$\\
+
+    $\PutR{} \App \String \App \StringAlt$
+    & =
+    & $\begin{cases*}
+        \Lens_1.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\BRegex_1} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_1}$\\
+        \Lens_2.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\BRegex_2} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_2}$\\
+        \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_1} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_2}$\\
+        \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_2} \BooleanAnd \StringAlt\in\LanguageOf{\BRegexAlt_1}$
+      \end{cases*}$\\
+
+    $\PutL{} \App \StringAlt \App \String$
+    & =
+    & $\begin{cases*}
+        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1} \BooleanAnd \String\in\LanguageOf{\BRegex_1}$\\
+        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2} \BooleanAnd \String\in\LanguageOf{\BRegex_2}$\\
+        \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1} \BooleanAnd \String\in\LanguageOf{\BRegex_2}$\\
+        \Lens_2.\CreateLOf{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2} \BooleanAnd \String\in\LanguageOf{\BRegex_1}$
+      \end{cases*}$\\
+  \end{tabular}
+\end{tabular}
+\end{centermath}
+The \OrLens lens deals with data that can come in one form or another. If the
+data gets changed from one format to the other, information in the old format is
+lost.
+\begin{center}
+  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App \String_1\ldots\String_n$
+    & =
+    & $(\Lens.\CreateROf{\String_1})\ldots(\Lens.\CreateROf{\String_n})$\\
+
+    $\CreateL{} \App \StringAlt_1\ldots\StringAlt_n$
+    & =
+    & $(\Lens.\CreateLOf{\StringAlt_1})\ldots(\Lens.\CreateLOf{\StringAlt_n})$\\
+
+    $\PutR{} \App (\String_1\ldots\String_n) \App (\StringAlt_1\ldots\StringAlt_m)$
+    & =
+    & $\StringAlt_1'\ldots\StringAlt_n'$ where $\StringAlt_i' =
+      \begin{cases*}
+        \Lens.\PutROf{\String_i}{\StringAlt_i} & if $i \leq m$\\
+        \Lens.\CreateROf{\String_i} & otherwise
+      \end{cases*}$\\
+    $\PutL{} \App (\StringAlt_1\ldots\StringAlt_m) \App (\String_1\ldots\String_n)$
+    & =
+    & $\String_1'\ldots\String_n'$ where $\String_i' =
+      \begin{cases*}
+        \Lens.\PutROf{\StringAlt_i}{\String_i} & if $i \leq n$\\
+        \Lens.\CreateROf{\StringAlt_i} & otherwise
+      \end{cases*}$
+  \end{tabular}
+\end{center}
+\begin{centermath}
+\begin{tabular}[b]{l@{\qquad}l}
+$
+  \centering
+  \inferrule*
+  {
+    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt\\
+    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt
+  }
+  {
+    \MergeROf{\Lens_1}{\Lens_2} \OfType
+    \RegexOr{\BRegex_1}{\BRegex_2}
+    \Leftrightarrow
+    \BRegexAlt
+  }
+$
+&
+  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App \String$
+    & =
+    & $\begin{cases*}
+      \Lens_1.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_1}$\\
+      \Lens_2.\CreateROf{\String} & if $\String\in\LanguageOf{\BRegex_2}$
+      \end{cases*}$\\
+
+    $\CreateL{} \App \StringAlt$
+    & =
+    & $\Lens_1.\CreateLOf{\StringAlt}$\\
+
+    $\PutR{} \App \String \App \StringAlt$
+    & =
+    & $\begin{cases*}
+      \Lens_1.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\BRegex_1}$\\
+      \Lens_2.\PutROf{\String}{\StringAlt} & if $\String\in\LanguageOf{\BRegex_2}$
+    \end{cases*}$\\
+
+    $\PutL{} \App \StringAlt \App \String$
+    & =
+    & $\begin{cases*}
+        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\String\in\LanguageOf{\BRegex_1}$\\
+        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\String\in\LanguageOf{\BRegex_2}$
+      \end{cases*}$\\
+  \end{tabular}
+\end{tabular}
+\end{centermath}
+
+The \MergeR lens is interesting because it merges data where one data can be in
+two formats, and one data has only one format. In previous
+work~\cite{Boomerang}, this was combined into \OrLens{}, where
+\OrLens{} could have ambiguous types, but we find it more clear to have explicit
+merge operators: it is easier to see what lens the synthesis algorithm is
+creating.
+
+\begin{centermath}
+\begin{tabular}[b]{l@{\qquad}l}
+$
+  \inferrule*
+  {
+    \Lens_1 \OfType \BRegex \Leftrightarrow \BRegexAlt_1\\
+    \Lens_2 \OfType \BRegex \Leftrightarrow \BRegexAlt_2
+  }
+  {
+    \MergeLOf{\Lens_1}{\Lens_2} \OfType
+    \BRegex
+    \Leftrightarrow
+    \RegexOr{\BRegexAlt_1}{\BRegexAlt_2}
+  }
+$
+&
+  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App \String$
+    & =
+    & $\Lens_1.\CreateROf{\String}$\\
+
+    $\CreateL{} \App \StringAlt$
+    & =
+    & $\begin{cases*}
+      \Lens_1.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1}$\\
+      \Lens_2.\CreateLOf{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2}$
+      \end{cases*}$\\
+
+    $\PutR{} \App \String \App \StringAlt$
+    & =
+    & $\begin{cases*}
+      \Lens_1.\PutROf{\String}{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1}$\\
+      \Lens_2.\PutROf{\String}{\StringAlt} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2}$
+    \end{cases*}$\\
+
+    $\PutL{} \App \StringAlt \App \String$
+    & =
+    & $\begin{cases*}
+        \Lens_1.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_1}$\\
+        \Lens_2.\PutLOf{\StringAlt}{\String} & if $\StringAlt\in\LanguageOf{\BRegexAlt_2}$
+      \end{cases*}$\\
+  \end{tabular}
+\end{tabular}
+\end{centermath}
+The \MergeL lens is symmetric to \MergeR.
+
+\begin{centermath}
+\begin{tabular}[b]{l@{\qquad}l}
+  $
+  \inferrule*
+  {
+    \Lens_1 \OfType \BRegex \Leftrightarrow \BRegexAlt\\
+    \Lens_2 \OfType \BRegexAlt \Leftrightarrow \BRegexAltAlt\\
+  }
+  {
+    \ComposeLensOf{\Lens_1}{\Lens_2} \OfType
+    \BRegex \Leftrightarrow \BRegexAltAlt
+  }
+$
+&
+  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App \String$ & = & $\Lens_2.\CreateROf{(\Lens_1.\CreateROf{\String})}$\\
+    $\CreateL{} \App \StringAlt$ & = & $\Lens_1.\CreateLOf{(\Lens_2.\CreateLOf{\StringAlt})}$\\
+    $\PutR{} \App \String \App \StringAltAlt$ & = & $\Lens_2.\PutROf{(\Lens_1.\PutROf{\String}{(\Lens_2.\CreateLOf{\StringAltAlt})})}{\StringAltAlt}$\\
+    $\PutL{} \App \StringAltAlt \App \String$ & = & $\Lens_1.\PutLOf{(\Lens_2.\PutLOf{\StringAltAlt}{(\Lens_2.\CreateROf{\String})})}{\String}$
+  \end{tabular}
+\end{tabular}
+\end{centermath}
+Composing is interesting in the put functions. Because puts require intermediary
+data, we recreate that intermediary data with creates.
+
+\begin{centermath}
+\begin{tabular}[b]{l@{\qquad}l}
+$
+\inferrule*
+  {
+    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt
+  }
+  {
+    \IterateLensOf{\Lens} \OfType
+    \StarOf{\BRegex}
+    \Leftrightarrow
+    \StarOf{\BRegexAlt}
+  }
+  $
+  &
+  $
+  \inferrule*
+  {
+    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt
+  }
+  {
+    \InvertOf{\Lens} \OfType \BRegexAlt \Leftrightarrow \BRegex
+  }
+  $
+\end{tabular}
+\end{centermath}
+\begin{center}
+  \begin{tabular}{@{}r@{\ }c@{\ }l@{}}
+    $\CreateR{} \App \StringAlt$ & = & $\Lens.\CreateLOf{\StringAlt}$\\
+    $\CreateL{} \App \String$ & = & $\Lens.\CreateROf{\String}$\\
+    $\PutR{} \App \StringAlt \App \String$ & = & $\Lens.\PutLOf{\StringAlt}{\String}$\\
+    $\PutL{} \App \String \App \StringAlt$ & = & $\Lens.\PutROf{\String}{\StringAlt}$
+  \end{tabular}
+\end{center}
+The \IterateLens lens deals with iterated data, while inverting reverses the direction of a lens: creating on the right becomes creating on the left and vice versa, and putting on the right becomes putting on the left and vice versa.
+The invert combonator is particularly useful when chaining many compositions together, as it can be used to align the central types.
+\[
+  \centering
+  \inferrule*
+  {
+    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt\\
+    \BRegex \SSREquiv \BRegex'\\
+    \BRegexAlt \SSREquiv \BRegexAlt'
+  }
+  {
+    \Lens \OfType \BRegex' \Leftrightarrow \BRegexAlt'
+  }
+\]
+
+Type equivalence enables a lens of type $S \Leftrightarrow T$ to be used as a
+lens of type $S' \Leftrightarrow T'$ if $S$ equivalent to $S'$ and $T$ is
+equivalent to $T'$. Type equivalence is useful both for addressing type
+annotations, and for making well-typed compositions.
 

### 2018-07-11 papers/SymmetricLenses/main.tex
minor edits to intro and ad-hoc --> ad hoc

@@
 some file-specific meta data, such as a date, a time or a reference
-number, whereas the other does not. Indeed, all of our benchmarks
-taken from Flash Fill~\cite{flashfill} and all of our benchmarks that
-synchronize between two ad hoc file formats described in
-Section \ref{sec:evaluation} \ksf{is this right?} have this 
-characteristic. Consequently, past tools that search for strict
-bijections will rarely be applicable in practice.
+number, whereas the other does not. Indeed of the 45 benchmark problems
+described in Section \ref{sec:evaluation}, all of the benchmarks
+taken from Flash Fill~\cite{flashfill} and all of the benchmarks that
+synchronize between two ad hoc file formats have this 
+characteristic, \afm{num} benchmarks in total. Consequently, tools that search
+only for strict bijections will rarely be applicable in practice.
 
@@
 extraction tasks \ksf{Dave wants to change information extraction to
-soemthing else.} taken from Augeas~\cite{Augeas}, and ad-hoc format synchronization tasks
+soemthing else.} taken from Augeas~\cite{Augeas}, and ad hoc format synchronization tasks
 taken from a variety of other locations\footnote{The majority of these format
@@
 \item We adapted \afm{29} benchmarks from Augeas. These benchmarks turn the
-  ad-hoc data present in Linux configuration files into a structured, dictionary
+  ad hoc data present in Linux configuration files into a structured, dictionary
   representation.

### 2018-07-11 papers/SymmetricLenses/main.tex
Spacing fix applied
some edits to section 2

@@
 A simple symmetric lens between formats described by types $X$ and $Y$ is
-defined by four functions.
+defined by four functions that collectively satisfy four ``round-tripping'' laws:
 
@@
 %\end{enumerate}
-
+\vskip 1.5ex
+\begin{minipage}[c]{.3\textwidth}
 \begin{gather}
@@
 \end{gather} 
-
-\noindent
-that collectively satisfy four ``round-tripping'' laws: 
-\begin{gather}
+\end{minipage} 
+\begin{minipage}[c]{.6\textwidth}
+\begin{align}
   \tag{\CreatePutRL}
@@
   \PutROf{(\PutLOf{y}{x})}{y} = y
-\end{gather}
+\end{align}
+\end{minipage} 
+\vskip 1ex
 \noindent
-The two {\sc Create} functions propagate shared information and fill in default
+The two \KW{create} functions propagate shared information and fill in default
 values when introducing new data. (\EG, the ``unk'' salary entry is added
 alongside the name to the management file when HR inserts a new employee). The
-two {\sc Put} functions propagate edits from one format to the other by
+two \KW{put} functions propagate edits from one format to the other by
 combining the change from one file with the previous value from the other file.
-The roundtripping laws guarantee that if two formats are already synchronized,
-{\sc Put}s merely return the second argument, unchanged.
-
-This definition differs from previous formulations of simple symmetric lenses,
-as previous formulations required a \emph{complement}. Instead of comprising
-four functions, existing formulations merely comprised two functions, which
-combined complements with edited formats to create updated complements and
-synchronized data. We provide a detailed comparison to classic symmetric lenses
+The round-tripping laws guarantee that if two formats are already synchronized,
+\KW{put}s merely return the second argument, unchanged.
+
+This definition differs from formulations of classic symmetric lenses,
+which require a \emph{complement} but only two functions.  These two
+functions combine complements with edited data to create updated
+complements and synchronized data, one function for each direction.
+We provide a detailed comparison to classic symmetric lenses
 in \S\ref{sec:relationship}.
@@
 Generally, we do not expect programmers to define simple symmetric lenses by
-writing the four functions by hand and showing they satisfy the roundtripping
+writing the four functions by hand and showing they satisfy the round-tripping
 laws. Instead, we provide simple symmetric lens combinators to build up the
-requisite lenses from simpler ones, and focus our combinators on providing
+requisite lenses from simpler ones. We focus on combinators for defining
 lenses between string data formats.
@@
 Figure~\ref{fig:minimized-representations} 
-helps, but only a little: it may, in general, require {\em many} examples to rule
+helps, but only a little: It may, in general, require {\em many} examples to rule
 out all possible occurrences of nested \Disconnect lenses,
@@
 edit $es_i$, the lens $\Lens$ synchronizes the other data to $es'_i$.
-\[
+\begin{centermath}
+\begin{tabular}[b]{l@{\qquad \; \qquad}l}
+$
   \inferrule
@@
   }
-\]
-\[
+$
+&
+$
   \inferrule
@@
   }
-\]
+$
+\end{tabular}
+\end{centermath}
 \[
@@
 
-\[
+\begin{centermath}
+\begin{tabular}[b]{l@{\qquad \; \qquad}l}
+$
   \inferrule

### 2018-07-11 papers/SymmetricLenses/main.tex
fixed capitalization error in boomerang citation that was generaing bibtex errors
removed information extraction from benchmark description

--- 2018-07-11 papers/SymmetricLenses/main.tex
@@
 benchmark suite of 45 problems, including data cleaning tasks taken
-from Flash Fill~\cite{gulwani-popl-2014}, information 
-extraction tasks \ksf{Dave wants to change information extraction to
-soemthing else.} taken from Augeas~\cite{Augeas}, and ad hoc format synchronization tasks
+from Flash Fill~\cite{gulwani-popl-2014}, view maintenance
+tasks taken from Augeas~\cite{Augeas}, and ad hoc format synchronization tasks
 taken from a variety of other locations\footnote{The majority of these format
@@
   implementation on a set of 45 lens synthesis benchmarks drawn from
-  data cleaning, \ksf{Dave wants to change information extraction to
-soemthing else.} information 
-  extraction, and file synchronization tasks.  We demonstrate that we can
+  data cleaning, view maintenance, 
+  and file synchronization tasks.  We demonstrate that we can
   synthesize simple symmetric lenses for all of the benchmarks in
@@
 But what is a ``likely'' satisfying lens? We propose the following
-heuristic: a satisfying
+heuristic: A satisfying
 lens is more likely if it connects more data between the two
@@
 
-\ksf{Consider integrating more detailed discussion of SREs that was
-formerly in the intro.  Commented out here for now.}
+%% \ksf{Consider integrating more detailed discussion of SREs that was
+%%formerly in the intro.  Commented out here for now.}
 %%To implement this intuition, we turn to \emph{stochastic regular
@@
 conditional entropy of $\RegexAlt$ given $\Regex$ and $\Lens$. This metric
-assigns higher cost to lenses where knowing the string present in \Regex
-provides little information about what string is present in \RegexAlt. With this
-metric, bijections have zero cost---given a string $s \in \Regex$, we know
-exactly what string is present in $\RegexAlt$, as exactly one string is
-synchronized to it.
+assigns higher entropy (or cost) to lenses where knowing the string
+on one side 
+provides little information about the string on the other side. 
+It assigns zero entropy to bijections because given a string $s \in \Regex$,
+the bijection exactly determines
+the corresponding string in 
+$\RegexAlt$.
 
 To obtain SREs from the plain regular expressions that we expect users to
-supply, we use a heurstic that attempts to assign probability annotations that
+supply, we use a heuristic that attempts to assign probability annotations that
 give each string in the language equal probablity. Our algorithm could also take
-as input a stochastic regular expression rather than a plain one, which allows
+as input a stochastic regular expression rather than a plain one,
+which would allow
 the user to specify a data-specific probability distribution, either manually
-written or inferred from a large dataset. We found our heuristically chosen
-distribution works well in practice.
+calculated or inferred from a large dataset. We found our heuristic
+works well in practice, so we have not explored these options.
 
@@
 include the semantics for combinators equivalent to those presented in prior
-work~\cite{symmetric-lenses,boomerang}. Furthermore, we only present \CreateR
+work~\cite{symmetric-lenses,Boomerang}. Furthermore, we only present \CreateR
 and \PutR when \CreateL and \PutL are symmetric. Full details are present in the

### 2018-07-11 papers/SymmetricLenses/main.tex
add ref to fig 4

--- 2018-07-11 papers/SymmetricLenses/main.tex
@@
 some file-specific meta data, such as a date, a time or a reference
-number, whereas the other does not. Indeed of the 45 benchmark problems
-described in Section \ref{sec:evaluation}, all of the benchmarks
-taken from Flash Fill~\cite{flashfill} and all of the benchmarks that
-synchronize between two ad hoc file formats have this 
-characteristic, \afm{num} benchmarks in total. Consequently, tools that search
-only for strict bijections will rarely be applicable in practice.
+number, whereas the other does not. Indeed, all of our benchmarks
+taken from Flash Fill~\cite{flashfill} and all of our benchmarks that
+synchronize between two ad hoc file formats described in
+Section \ref{sec:evaluation} \ksf{is this right?} have this 
+characteristic. Consequently, past tools that search for strict
+bijections will rarely be applicable in practice.
 
@@
 extraction tasks \ksf{Dave wants to change information extraction to
-soemthing else.} taken from Augeas~\cite{Augeas}, and ad hoc format synchronization tasks
+soemthing else.} taken from Augeas~\cite{Augeas}, and ad-hoc format synchronization tasks
 taken from a variety of other locations\footnote{The majority of these format
@@
 \]
-We have proved that this conversion respects languages and probability
-distributions.
+This conversion utilizes operators defined in
+Figure~\ref{fig:dnf-regex-functions}. We have proved that this conversion
+respects languages and probability distributions.
 
@@
 \item We adapted \afm{29} benchmarks from Augeas. These benchmarks turn the
-  ad hoc data present in Linux configuration files into a structured, dictionary
+  ad-hoc data present in Linux configuration files into a structured, dictionary
   representation.
@@
 edit $es_i$, the lens $\Lens$ synchronizes the other data to $es'_i$.
-\begin{centermath}
-\begin{tabular}[b]{l@{\qquad \; \qquad}l}
-$
+\[
   \inferrule
@@
   }
-$
-&
-$
+\]
+\[
   \inferrule
@@
   }
-$
-\end{tabular}
-\end{centermath}
+\]
 \[
@@
 
-\begin{centermath}
-\begin{tabular}[b]{l@{\qquad \; \qquad}l}
-$
+\[
   \inferrule

### 2018-07-11 papers/SymmetricLenses/main.tex
edits to section 1 and section 2

--- 2018-07-11 papers/SymmetricLenses/main.tex
@@
 
+%%As a further example, consider the trivial lens
+%%\lstinline{disc(emp_salary,emp_row,rep(emp_salary),rep(emp_row))}. As this is
+%%the disconnect lens, the cost of each side is the entropy of the data format, so
+%%the total cost is $\EntropyOf{\lstinline{emp_salary} } +
+%%\EntropyOf{\lstinline{emp_row}}$. The lens \lstinline{emp_salary} contains strictly more
+%%information than $\EntropyOf{\lstinline{": " . salary}}$, and
+%%\lstinline{emp_row} contains strictly more information than
+%%\EntropyOf{\lstinline{": " . salary}}, so by our metric
+%%\lstinline{single_emp_map} is preferable to the disconnect lens.
+
 To obtain SREs from the plain regular expressions that we expect users to
@@
 %lens.
-Consider the trivial lens
-\lstinline{disc(emp_salary,emp_row,rep(emp_salary),rep(emp_row))}. As this is
-the disconnect lens, the cost of each side is the entropy of the data format, so
-the total cost is $\EntropyOf{\lstinline{emp_salary} } +
-\EntropyOf{\lstinline{emp_row}}$. \lstinline{emp_salary} contains strictly more
-information than $\EntropyOf{\lstinline{": " . salary}}$, and
-\lstinline{emp_row} contains strictly more information than
-\EntropyOf{\lstinline{": " . salary}}, so by our metric
-\lstinline{single_emp_map} is preferable to the disconnect lens.
 % end overview

### 2018-07-11 papers/SymmetricLenses/main.tex
minor edit
deleted extra paragraph in section 2
big edit on related

--- 2018-07-11 papers/SymmetricLenses/main.tex
@@
 
-Two classical symmetric lenses are equivalent if they output the same
-formats given any sequence of edits. Formally, given a lens
-$\Lens \in X \leftrightarrow Y$,
-an \emph{edit} for $\Lens$ is a member of $X + Y$. Consider
-the function $apply$, which, given a lens and an element of that lens's
-complement, is a function from sequences of edits to sequences of edits. If
-$apply(\Lens,c,es) = es'$, then that means, starting with complement $c$, after
-edit $es_i$, the lens $\Lens$ synchronizes the other data to $es'_i$.
+Two classical symmetric lenses are equivalent if they output the same formats
+given any sequence of edits. Formally, given a lens $\Lens \in X \leftrightarrow
+Y$, an \emph{edit} for $\Lens$ is a member of $X + Y$. Consider the function
+$apply$, which, given a lens and an element of that lens's complement, is a
+function from sequences of edits to sequences of edits. If $apply(\Lens,c,es) =
+es'$, then given complement $c$ and edits $es_i$, the lens $\Lens$ generates
+$es'_i$.
 \begin{centermath}
@@
 Two lenses, $\Lens_1$ and $\Lens_2$, are equivalent if
-$apply(\Lens_1,\Lens_1.init,es) = apply(\Lens_2,\Lens_2.init,es)$ for all edit
-sequences $es$.
+$apply(\Lens_1,\Lens_1.init,es) = apply(\Lens_2,\Lens_2.init,es)$ for all $es$.
 
@@
 %\paragraph*{Issues with Classical Symmetric Lenses}
-\iffalse
-Using symmetric lenses for
-synchronization forces us to maintain not just two files that we aim to
-sycnrhonize, but three. \bcp{I don't find it unintuitive, and I think we
-  should be careful not to oversell the difficulty of storing the
-  complements.  It requires more engineering, but it's not rocket
-  science.}This is unintuitive, and creates many problems in 
-practice. For example, consider the situation where I am synchronizing a crontab
-and job manager file. I am synchronizing across computers, so I must either
-maintain a copy of the complement on each computer, and keep them synchronized,
-or figure out a computer to hold the complement, and make sure all edits go
-through that complement. Furthermore, if I want to change the synchronization to
-a different computer, I don't need to merely make sure that my files are moved
-to the computer, I also have to make sure that I correctly move the complement,
-and hook up that complement to be used for that synchronization task. This is a
-huge divergence from other synchronization utilities\bcp{No.  For example,
-  Unison maintains a database on each replica recording, for each file, what
-its hash was after the last time the two replicas were synchronized.}, where all
-that needs to be 
-done is maintain the files. Using those for synchronization, edits can be sent
-over the wire, and the appropriate function can stitch together those edits, no
-additional state needed.
-\fi
-While classical symmetric are more expressive than simple symmetric
-lenses, it is unclear whether this additional expressiveness is useful in
-practice.  Moreover, classic symmetric lenses come with 
-some disadvantages.
-For one, an implementation of classical symmetric lenses
-would need to squirrel away the complement in
-an auxiliary file somwhere, which likely requires more engineering.
-However, more significant issues arise in synthesis.
-First, because each lens has a custom complement, one can no longer
-specify the put functions through input/output examples alone.  One
-alternative would be to enrich specifications with edit sequentials;
-another would be to specify the structure of complements explicitly
-(though the latter would be somewhat akin to specifying the internal
-state of a program).  In either case, the complexity of the specifications
-increases.
+\iffalse Using symmetric lenses for synchronization forces us to maintain not
+just two files that we aim to sycnrhonize, but three. \bcp{I don't find it
+  unintuitive, and I think we should be careful not to oversell the difficulty
+  of storing the complements. It requires more engineering, but it's not rocket
+  science.}This is unintuitive, and creates many problems in practice. For
+example, consider the situation where I am synchronizing a crontab and job
+manager file. I am synchronizing across computers, so I must either maintain a
+copy of the complement on each computer, and keep them synchronized, or figure
+out a computer to hold the complement, and make sure all edits go through that
+complement. Furthermore, if I want to change the synchronization to a different
+computer, I don't need to merely make sure that my files are moved to the
+computer, I also have to make sure that I correctly move the complement, and
+hook up that complement to be used for that synchronization task. This is a huge
+divergence from other synchronization utilities\bcp{No. For example, Unison
+  maintains a database on each replica recording, for each file, what its hash
+  was after the last time the two replicas were synchronized.}, where all that
+needs to be done is maintain the files. Using those for synchronization, edits
+can be sent over the wire, and the appropriate function can stitch together
+those edits, no additional state needed. \fi While classical symmetric lenses
+are technically more expressive than simple symmetric lenses, it is unclear
+whether this additional expressiveness is useful in practice. In our judgement,
+the complications they introduce into synthesis did not justify the potential
+additional expressiveness: because each lens has a custom
+complement, one can no longer specify the put functions through input/output
+examples alone. One alternative would be to enrich specifications with edit
+sequentials; another would be to specify the structure of complements explicitly
+(though the latter would be somewhat akin to specifying the internal state of a
+program). In either case, the complexity of the specifications increases.
 %% addition, users would have to specify some auxilary 
@@
 
-\paragraph*{Relation with Simple Symmetric Lenses} To formalize how simple
-symmetric lenses work as synchronization tools, we define an $apply$ function on
-them, much like one defines an $apply$ on classical symmetric lenses: Given a
-simple symmetric lens and an optional pair of source and target strings, $apply$
-is a function from sequences of edits to sequences of edits. If
+\paragraph*{Relation with Simple Symmetric Lenses} To compare classic and simple
+symmetric lenses, we define an $apply$ function on simple lenses as well. If
 $apply(\Lens,\None,es) = es'$, then starting with no prior data, after edit
-$es_i$, the lens \Lens synchronizes the updates the data in the non-edited
-format (the right format if $es_i = \InLOf{x}$, and the left format if $es_i =
-\InROf{y}$) to $es_i'$. If $apply(\Lens,\SomeOf{(x,y)},es) = es'$, then that
-means, starting with data $x$ and $y$ on the left and right, respectively, after
-edit $es_i$, the lens \Lens synchronizes the other data to $es_i'$.
+$es_i$, the lens \Lens generates $es_i'$ (the right format if $es_i =
+\InLOf{x}$, and the left format if $es_i = \InROf{y}$). If
+$apply(\Lens,\SomeOf{(x,y)},es) = es'$, then starting with data $x$
+and $y$ on the left and right, respectively, after edit $es_i$, the lens \Lens
+generates $es_i'$.
 
@@
 
-We have reformulated symmetric lenses in a style similar to existing asymmetric
-lens formulations. However, some symmetric lenses certainly are inexpressible in
-the simple symmetric lens formulation. We formalize which are expressible
-with a restriction on symmetric lenses. We define \emph{forgetful symmetric
+Next, we define \emph{forgetful symmetric
   lenses} to be symmetric lenses that satisfy the following additional laws.
@@
 
-Intuitively, these equations state that complements are uniquely determined by the most
-recently input $x$ and $y$. Hence, such lenses correspond closely \dpw{replace ``closely'' with
-``exactly''?} with simple symmetric
-lenses, where all state is maintained by the current state of the $x$ and $y$
-data. Forgetful symmetric lenses express exactly the same $apply$ function as
-simple symmetric lenses.
+Intuitively, these equations state that complements are uniquely determined by
+the most recently input $x$ and $y$. Hence, such lenses correspond exactly with
+simple symmetric lenses, where all state is maintained by the current state of
+the $x$ and $y$ data. Forgetful symmetric lenses express exactly the same
+$apply$ function as simple symmetric lenses.
 
@@
 \end{theorem}
-
-\bcp{I guess we could also show that equivalence of simple symmetric lenses
-  is just the identity relation?}
-\sam{No; for example the identity lens of a concatenation of two regular
-  expressions is the same as concatenating the two identity lenses on the
-  respective regular expressions}
-\afm{I think Benjamin likely means identity relation on the semantic meaning of
-  the lenses, which is true.  Would this be a useful thing to prove? It's a 2
-  liner.}
 
@@
 FlashFill (to take one well-documented example) may be connected to some of the principles
-laid out here.  For instance, FlashFill prioritizes the substring constructor over the
-constant string constructor when ranking possible programs. Such a choice may be understood
-in information-theoretic terms as the translation from target back to source would require
-more information.  Said another way, our information theoretic metrics measure injectivity
-(in both directions, in our case, to measure bijectivity) but applied in one direction, as
-is the case for FlashFill, would avoid the highly injective constant functions.  Beyond
-constants, another FlashFill heuristic prefers ``wider'' character classes (or token classes)
-over ``narrower'' ones.  Selecting a wider set of strings to copy from left-to-right, viewed
-in the reverse direction, from right-to-left, results in a function that is closer to being
-injective, as measured by information theory.  These are just a few of the heuristics used
-in Flashfill; more research would be required (and public disclosure of the details necessary)
-to understand the connections fully.
+laid out here.  For instance, Flash Fill prioritizes the substring constructor over the
+constant string constructor when ranking possible programs. Such a choice is
+consistent with our information-theoretic viewpoint as the constant function
+throws away a great deal of information about the source string being
+transformed. Likewise, Flash Fill prefers ``wider'' character classes over
+``narrower'' ones. Again such a choice is implied by information theory -- the
+wider class preserves more information during translation.  More broadly, we
+hope our information theoretic analysis provides a principle that can be
+used to understand the heuristic choices made in related work.
 
@@
 
-More broadly, information theory and probabilistic languages are common tools
-in natural language understanding,  machine translation, information retrieval,
-data extraction and grammatical inference (see Pereira~\cite{pereira:info-theory},
+More broadly, information theory and probabilistic languages are common tools in
+natural language understanding, machine translation, information retrieval, data
+extraction and grammatical inference (see Pereira~\cite{pereira:info-theory},
 for an introduction). Indeed, our work was inspired, in part, by the PADS format
-inference tool~\cite{pads:synthesis}, which was in turn inspired by earlier work on
-probabilistic gramatical inference
-and information extraction~\cite{kushmerick-thesis,arasu:extracting}.  More specifically, PADS
-used an information-theoretic metric to help judge the quality of inferred grammars
-for ad hoc data formats.  However, PADS did not synthesize data transformers, and
-we are not aware of the use of information-theoretic principles in type-directed or syntax-guided
-synthesis of DSL programs.
+inference tool~\cite{pads:synthesis}, which was in turn inspired by earlier work
+on probabilistic gramatical inference and information
+extraction~\cite{kushmerick-thesis,arasu:extracting}. PADS did not synthesize
+data transformers, and we are not aware of the use of information-theoretic
+principles in type-directed or syntax-guided synthesis of DSL programs.
 
@@
 We have developed a synthesis algorithm for synthesizing synchronization
-functions between data formats that may not be in bijective correspondence. We
-do this by synthesizing symmetric lenses. However, existing symmetric lens
-formulations use complex complements to store external state. We identify a
-restriction on symmetric lenses, simple symmetric lenses. We develop combinators
-for simple symmetric lenses on string data. We show how to synthesize these
-lenses from regular expression specifications. Prior work could return
-\emph{any} satisfying lens, as the type system was restrictive enough that any
-satisfying lens would typically be the correct satisfying lens. Because of the
-increased flexibility of the symmetric lens language, we identify an
-information-theoretic heuristic for finding a likely satisfying lens. We
-developed a greedy synthesis algorithm that aims to find the lens of lowest
-cost. We demonstrate empiraclly that our algorithm can find lenses quickly, and
-our heuristic makes finding the correct lens require fewer specifications.
+functions between data formats that may not be in bijective correspondence. More
+specifically, we identify a subset of symmetric lenses, simple symmetric lenses,
+develop new combinators for them, and show how to synthesize them from regular
+expression specifications. In order to guide the search for ``likely'' lenses,
+we introduce new search principles based on information theory. Based on these
+principles, we designed, implemented, and evaluated a new tool for lens
+synthesis.
 
@@
   \begin{array}{@{}r@{\ }c@{\ }l@{}}
-    $\CreateR{} \App \String'$ & = & $\StringAlt$\\
+    $\CreateR{} \App \String''''$ & = & $\StringAlt$\\
     $\CreateL{} \App \StringAlt'''''$ & = & $\String$\\

### 2018-07-11 papers/SymmetricLenses/main.tex

@@
     When the algorithm determines it has found a likely lens, it
-  returns it.  \ksf{Can we change cost to likeliness?}
+  returns it.  \ksf{Can we change cost to likelihood and l\_likely to l\_best?}
    }
@@
 \KW{put}s merely return the second argument, unchanged.
-
 This definition differs from formulations of classic symmetric lenses,

### 2018-07-11 papers/SymmetricLenses/main.tex
add main

@@
 \usepackage{amsmath}
+\usepackage{makecell}%To keep spacing of text in tables
 \usepackage{nccmath}
@@
  {$\end{center}}
+\setcellgapes{4pt}%parameter for the spacing
 
@@
 
-\subsection{Stochastic Regular Expression Equivalences}
-$\Expand$ needs to be able to find equivalent stochastic regular expressions.
-However, existing work does not provide a means to reason about stochastic
-regular expression equivalence. We extend the \emph{star-semiring}~\cite{Droste2009}
-equivalences to stochastic regular expressions.
-
-\begin{center}
+\begin{figure}
   \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
@@
   \end{tabular}
-\end{center}
+  \caption{Stochastic Regular Expression Star-Semiring Equivalence}
+  \label{fig:ssr-sre}
+\end{figure}
+
+\subsection{Stochastic Regular Expression Equivalences}
+$\Expand$ needs to be able to find equivalent stochastic regular expressions.
+However, existing work does not provide a means to reason about stochastic
+regular expression equivalence. We extend the
+\emph{star-semiring}~\cite{Droste2009} equivalences to stochastic regular
+expressions (Figure~\ref{fig:ssr-sre}).
 
@@
 \begin{centermath}
-\begin{array}[b]{l@{\qquad \; \qquad}l}
-\begin{array}[b]{c}
-\quad\\
+  \begin{tabular}{m{1cm}@{\hspace*{6em}}m{1cm}@{\hspace*{3em}}}
+    $
   \inferrule*
@@
     \IdentityLensOf{\BRegex} \OfType \BRegex \Leftrightarrow \BRegex
-  }
-  \\
-  \quad
-  \end{array}
+  }$
 &
-\begin{array}[b]{r@{\ }c@{\ }l}
+$\begin{array}[b]{r@{\ }c@{\ }l}
     \CreateR{} \App s & = & s\\
-    \PutR{} \App s_1 \App s_2 & = & s_1\\
-  \end{array}
-\end{array}
+    \PutR{} \App s_1 \App s_2 & = & s_1
+  \end{array}$
+\end{tabular}
 \end{centermath}
@@
 \begin{centermath}
-\begin{array}[b]{l@{\qquad \; \qquad}l}
+  \begin{tabular}{m{5cm}m{2cm}}
+    $
     \inferrule*
@@
     \OfType \BRegex \Leftrightarrow \BRegexAlt
-  }
+    }$
     &
+      $
   \begin{array}{@{}r@{\ }c@{\ }l@{}}
     $\CreateR{} \App \String'$ & = & $\StringAlt$\\
-    $\PutR{} \App \String''' \App \StringAlt'$ & = & $\StringAlt'$\\
-  \end{array}
-\end{array}
+    $\PutR{} \App \String' \App \StringAlt'$ & = & $\StringAlt'$
+  \end{array}$
+\end{tabular}
 \end{centermath}
@@
 \begin{centermath}
-\begin{tabular}[b]{l@{\qquad}l}
+  \begin{tabular}{m{5cm}m{7cm}}
 $
@@
 \begin{centermath}
-\begin{tabular}[b]{l@{\qquad}l}
+  \begin{tabular}{m{5cm}m{6cm}}
 $
@@
 \begin{centermath}
-\begin{tabular}[b]{l@{\qquad}l}
+  \begin{tabular}{m{5cm}m{6cm}}
   $

### 2018-07-11 papers/SymmetricLenses/main.tex
a pass through section 5

--- 2018-07-11 papers/SymmetricLenses/main.tex
@@
 to our information theoretic metric, and
-\item adds all of the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)
-  to the priority queue with $\PCF{\PQ.Push}$. The priority of a SRE pair is the
+\item adds the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)
+  to the priority queue.  The priority of a SRE pair is the
   number of rewrites needed to derive the pair from the originals.
@@
 When the loop terminates, the search returns the globally best lens found
-($best$).  Each element of this algorithm will be explained in
+($best$).  Each subroutine of this algorithm will be explained in
 further depth in the following subsections.
 
-\begin{algorithm}
+\begin{algorithm}[t]
   \caption{\SynthSymLens}
@@
 expressions that are semantically, but not syntactically, equivalent to the
-originals. In particular, this synthesizer performs the following rewrites:
+originals by performing the following rewrites:
 \begin{center}
@@
 
-Our priority queue proposes stochastic regular expressions generated from fewer
+The priority queue yields stochastic regular expressions generated using fewer
 rewrites first. Only when there are no more proposed regular expressions derived
@@
 
-When $\Call{Continue}{pq}$ returns $\False$, \RXSearch stops proposing regular
+When $\Call{Continue}{pq}$ returns $\False$, \Expand stops proposing regular
 expression pairs, and instead returns to the user the best lens found thus far.
-Note this means that if \RXSearch finds a bijective lens, which has zero cost it
+Note this means that if the algorithm finds a bijective lens, which has zero cost, it
 will immediately return.
@@
 (\EG, \emph{Unrollstar$_L$} and \emph{Unrollstar$_R$}) are needed to relate
-two \SDNFREabbrevs, other axioms, like associativity, commutativity
+two \SDNFREabbrevs; other axioms, like associativity, commutativity
 or distributivity, are unneeded.  This important property makes it possible to
@@
 
-Intuitively, stochastic DNF regular expressions are essentially stochastic
+Intuitively, stochastic DNF regular expressions are stochastic
 regular expressions with all concatenations fully distributed over all
@@
 As these DNF regular expressions are \emph{stochastic}, they are annotated with
-probabilities to express a probability distribution, in addition to a language. 
+probabilities to express a probability distribution, in addition to a
+language. \saz{I find the use of asterisk for the $\times$ or $\cdot$ or maybe
+  just juxtaposition slightly jarring.}
 \begin{center}
@@
 
-We have a means of turning stochastic regular expressions into stochastic DNF
-regular expressions. The conversion algorithm itself, written
+The algorithm for converting a stochastic regular expressions into its
+corresponding stochastic DNF regular expressions, written
 $\ToDNFRegexOf{\Regex}$, is defined below.
@@
 \]
-This conversion utilizes operators defined in
+This conversion relies on operators defined in
 Figure~\ref{fig:dnf-regex-functions}. We have proved that this conversion
@@
 \paragraph*{\ToStochastic} With stochastic DNF regular expressions and
-\ToDNFRegex defined, it is easier to explain the function that converts regular
-expressions into stochastic regular expressions, \ToStochastic. If users
-specified stochastic regular expressions themselves (perhaps inferred from
+\ToDNFRegex defined, it is easier to explain \ToStochastic, the function that converts regular
+expressions into stochastic regular expressions. If users
+provided stochastic regular expressions themselves (perhaps inferred from
 example data), this step could be skipped,
@@
 \label{subsec:greedy-synth}
-The greedy algorithm is actually three algorithms, one that greedily finds
+The synthesis procedure comprises three algorithms: one that greedily finds
 symmetric DNF lenses (\GreedySynth), one that greedily finds symmetric sequence
@@
 the structure of symmetric DNF lenses and \SDNFREabbrevs. We describe this
-synthesis algorithm in the context of trying to find a lens between
+synthesis algorithm using the running example of searching for a lens between
 \lstinline{"" | name.name*} and \lstinline{"" | name}.
@@
 \paragraph*{Symmetric DNF Lenses} Algorithm~\ref{alg:greedysynth} presents
-\GreedySynth, a means to synthesize symmetric DNF lenses. The inputs to
-\GreedySynth are a suite of input-output examples and a pair of stochastic DNF regular expressions.
-
-First, \PCF{CannotMap} checks whether there is no possible lens satisfying the
-examples. If so, \GreedySynth returns \None early. Otherwise, \GreedySynth finds
+\GreedySynth, which synthesizes symmetric DNF lenses. Its inputs 
+are a suite of input-output examples and a pair of stochastic DNF regular expressions.
+%
+First, \PCF{CannotMap} determines whether there is no lens satisfying the
+examples, and if so, \GreedySynth returns \None immediately. Otherwise, \GreedySynth finds
 the best lenses (given the examples that match them) between all sequence pairs
@@
 queue. Within this loop, a sequence lense is popped from the queue and, if it is
-``useful'', then it is included in the final DNF lens. The lens is considered to
-be useful when its source (or target) is not already the source (or target) of
+``useful,'' it is included in the final DNF lens. The lens is considered to
+be useful when its source (or target) is \textit{not} already the source (or target) of
 an already chosen sequence lens. If the examples require that two sequences have a
@@
 
-\begin{algorithm}
+\begin{algorithm}[t]
   \caption{\GreedySynth}
@@
 presents \GreedySeqSynth, which synthesizes symmetric sequence lenses using an
-algorithm whose structure is similar to \GreedySynth{}'s.  It calls \PCF{AtomSynth}, a
-function that synthesizes atom lenses by iterating a DNF lens between
+algorithm whose structure is similar to \GreedySynth{}'s.  It calls \PCF{AtomSynth}, 
+which synthesizes atom lenses by iterating a DNF lens between
 its subcomponents.
@@
 \GreedySeqSynth then loops until there are no more atom lenses in the priority
-queue. Within this loop, atom lenses are popped from the queue and the
-``useful'' ones are kept. An atom lens is useful if adding it will lower the
-cost of the generated sequence lens, or if examples show that one of the atoms
-must not be disconnected. Each atom can only be part of one lens at a time, so
-the algorithm must at times disconnect previously connected atoms to connect an
+queue. Within the loop, a popped atom lens is considered 
+``useful'' if adding it to the sequence will lower the
+cost of the generated sequence lens, or if examples show that one of its atoms
+must not be disconnected. Each atom can be part of only one lens at a time, so
+the algorithm must sometimes a previously chosen atom lens in order to connect an
 atom that must not be disconnected.
-
-Finally, after all atoms have been popped, the sequence lens is validated and
-returned. The algorithm confirms that all atoms that must not be disconnected
-are involved in an atom lens. If they are not, \PCF{SLensBuilder.ToDNFLens}
-instead returns \None.
-
-\begin{algorithm}
+After all atoms have been popped, the sequence lens is validated and
+returned. The algorithm succeeds when all atoms that must not be disconnected
+are involved in an atom lens, but \PCF{SLensBuilder.ToDNFLens}
+returns \None otherwise.
+
+\begin{algorithm}[t]
   \caption{\GreedySeqSynth}
@@
 
-Our implementation includes a number of optimizations to this algorithm. We
-added ``closed'' annotations to regular expressions \saz{what do these mean? I
-  guess ``open'' means ``subject to DNF conversion'' and ``closed'' means
-  not?}\afm{I describe in greater detail below, should I add in additional info
-  here too?}, an expansion inference algorithm, hash-consing and memoization,
-along with compositional synthesis. The implementation also allows an additional
-parameter in specifications that customizes the termination condition for the
-synthesis procedure.
+Our implementation includes a number of optimizations not described above:
+annotations that guide DNF conversion; an expansion inference
+algorithm; hash-consing and memoization; and compositional synthesis. The
+implementation also provides for an additional specification parameter that
+customizes the termination condition for the synthesis procedure.
 
@@
 relatively efficiently, it suffers from an exponential blowup when converting
-regular expressions to DNF form. This can be minimized by not converting
-\emph{all} of the regular expression into DNF form, but only converting
-\emph{some} of it to DNF form, and doing so incrementally. This change is easy
-to make by adding non-DNF stochastic regular expressions as a new type of
+regular expressions to DNF form. This can be mitigated by not converting
+\emph{all} of the regular expression to DNF form, but only converting
+\emph{some} of it, and doing so incrementally. This change is easy
+to support by adding non-DNF stochastic regular expressions as a new type of
 ``closed'' atom. When converted into a DNF regular expression, if $\Regex$ is
@@
 
-At the start of synthesis, all regular expressions are closed. \RXSearch is
+At the start of synthesis, all regular expressions are closed. \Expand is
 modified so that it is now permitted to rewrite closed regular expressions to
 open ones (thereby triggering DNF normalization). \GreedySynth is allowed to to
-map closed regular expressions to each other, but only using the identity
+map closed regular expressions to each other, but only by using the identity
 lens---they must parse the same string components in all examples.
 
-\paragraph*{Expansion Inference} The prior optimization helped \GreedySynth, but
-it made \RXSearch have the harder job of determining when to open closed
-expressions.  To alleviate this difficulty, we added in expansion inference,
-which identifies when a regular expression cannot possibly map by any means
-other than the disconnect lens.  Expansion inference opens those regular
-expressions to allow for further exploration.
-
-\paragraph*{Hash-consing and Memoization} We further help \GreedySynth by
+\paragraph*{Expansion Inference} Distinguishing open and closed regular
+expressions helps \GreedySynth, but doing so means that \Expand has the harder job of
+determining when to open closed expressions.  To that end,
+expansion inference identifies when a regular expression cannot
+possibly map by any means other than the ``trivial'' disconnect lens. Expansion inference
+opens those regular expressions to allow for further exploration.
+
+\paragraph*{Hash-consing and Memoization} We further improve \GreedySynth by
 memoizing its results.  However, our stochastic DNF regular expressions are
@@
 mapped to each other either if they are the same regular expression and agree
-with the examples, or if there is a lens which is a composition of predefined
-lenses between them, where the lens agrees with the examples.
+with the examples, or if there is an existing lens (which is a composition of predefined
+lenses) that already agrees with the examples.
 
 \paragraph*{Termination Customization}
-In this work, we are constrained by three competing constraints. (1) We want to
-return lenses quickly to users, (2) we want to return good lenses, and (3) we
-don't want to require users to write a large number of examples in the
+In this work, there are three competing constraints. (1) We want to return
+lenses quickly to users, (2) we want to return good lenses, and (3) we don't
+want to require users to write a large number of examples in their
 specifications. Our termination heuristic works well for most situations, but
-there are situations where a large number of expansions must be performed to
-find the correct lens.
-
-Users could prevent incorrect lenses with a large number of examples, but it can
-be quite difficult to write a large number of examples for complex formats. We
-find it simpler to let the users specify that \RXSearch should be more
-restrictive in which lenses it accepts, so users do not have to write a lot of
-examples.
-
-This constraint goes the other way too. If the user's desired transformation has
-a particularly high cost, the algorithm may spend a long time searching in vain
-for lenses with lower cost. By tuning \RXSearch to me more permissive in the
-lenses it accepts, this search can be short-circuited.
-
-For these types of situations, users can input a custom number, $t$, to
-customize the strictness of the termination condition. In particular, if
-$\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at distance $d$, the
-number of pairs in $pq$ at distance $d$ is $n$, and the current best lens has
-cost $c$:
+there are cases where a large number of expansions must be performed to find the
+correct lens.  Users could work around that problem by providing a large number
+of examples, but it can be difficult to write a large number of informative
+examples for complex formats. We find it simpler to instead let the users
+specify that \SynthSymLens should be more restrictive in which lenses it
+accepts.  Conversely, if the user's desired transformation has a particularly
+high cost, the algorithm may spend a long time searching in vain for lenses with
+lower cost. By tuning \SynthSymLens to be more permissive, such a search can be
+short-circuited.
+
+For these situations, users can provide a number, $t$, which affects
+the strictness of the termination condition, adjusting the \Continue formula
+as follows. When $\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at
+distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
+current best lens has cost $c$:
 \begin{gather*}
@@
 \end{gather*}
-With a positive $t$, the termination condition is more strict, and the algorithm
+With a positive $t$, the termination condition is stricter, and the algorithm
 will spend longer searching for more likely lenses. With a negative $t$ the
 termination condition is less strict, and the algorithm will spend less time
-searching for more likely lenses.  Our default $t$ is set to $0$.
+searching for more likely lenses.  Our default $t$ is $0$.
 
@@
 \afm{35} of our \afm{45} benchmarks, our termination condition is not
-one-size-fits all. Particularly complex lenses, or lenses that project a
-particularly large amount of data, can be synthesized more easily and quickly by
-customizing how permissive \RXSearch is in returning lenses.
+one-size-fits all. Complex lenses, or lenses that project a
+large amount of data, can be synthesized more easily and quickly by
+adjusting $t$.
 % end synthesis
@@
 
-Two classical symmetric lenses are equivalent if they output the same formats
-given any sequence of edits. Formally, given a lens $\Lens \in X \leftrightarrow
-Y$, an \emph{edit} for $\Lens$ is a member of $X + Y$. Consider the function
-$apply$, which, given a lens and an element of that lens's complement, is a
-function from sequences of edits to sequences of edits. If $apply(\Lens,c,es) =
-es'$, then given complement $c$ and edits $es_i$, the lens $\Lens$ generates
-$es'_i$.
+Two classical symmetric lenses are equivalent if they output the same
+formats given any sequence of edits. Formally, given a lens
+$\Lens \in X \leftrightarrow Y$,
+an \emph{edit} for $\Lens$ is a member of $X + Y$. Consider
+the function $apply$, which, given a lens and an element of that lens's
+complement, is a function from sequences of edits to sequences of edits. If
+$apply(\Lens,c,es) = es'$, then that means, starting with complement $c$, after
+edit $es_i$, the lens $\Lens$ synchronizes the other data to $es'_i$.
 \begin{centermath}
@@
 Two lenses, $\Lens_1$ and $\Lens_2$, are equivalent if
-$apply(\Lens_1,\Lens_1.init,es) = apply(\Lens_2,\Lens_2.init,es)$ for all $es$.
+$apply(\Lens_1,\Lens_1.init,es) = apply(\Lens_2,\Lens_2.init,es)$ for all edit
+sequences $es$.
 
@@
 %\paragraph*{Issues with Classical Symmetric Lenses}
-\iffalse Using symmetric lenses for synchronization forces us to maintain not
-just two files that we aim to sycnrhonize, but three. \bcp{I don't find it
-  unintuitive, and I think we should be careful not to oversell the difficulty
-  of storing the complements. It requires more engineering, but it's not rocket
-  science.}This is unintuitive, and creates many problems in practice. For
-example, consider the situation where I am synchronizing a crontab and job
-manager file. I am synchronizing across computers, so I must either maintain a
-copy of the complement on each computer, and keep them synchronized, or figure
-out a computer to hold the complement, and make sure all edits go through that
-complement. Furthermore, if I want to change the synchronization to a different
-computer, I don't need to merely make sure that my files are moved to the
-computer, I also have to make sure that I correctly move the complement, and
-hook up that complement to be used for that synchronization task. This is a huge
-divergence from other synchronization utilities\bcp{No. For example, Unison
-  maintains a database on each replica recording, for each file, what its hash
-  was after the last time the two replicas were synchronized.}, where all that
-needs to be done is maintain the files. Using those for synchronization, edits
-can be sent over the wire, and the appropriate function can stitch together
-those edits, no additional state needed. \fi While classical symmetric lenses
-are technically more expressive than simple symmetric lenses, it is unclear
-whether this additional expressiveness is useful in practice. In our judgement,
-the complications they introduce into synthesis did not justify the potential
-additional expressiveness: because each lens has a custom
-complement, one can no longer specify the put functions through input/output
-examples alone. One alternative would be to enrich specifications with edit
-sequentials; another would be to specify the structure of complements explicitly
-(though the latter would be somewhat akin to specifying the internal state of a
-program). In either case, the complexity of the specifications increases.
+\iffalse
+Using symmetric lenses for
+synchronization forces us to maintain not just two files that we aim to
+sycnrhonize, but three. \bcp{I don't find it unintuitive, and I think we
+  should be careful not to oversell the difficulty of storing the
+  complements.  It requires more engineering, but it's not rocket
+  science.}This is unintuitive, and creates many problems in 
+practice. For example, consider the situation where I am synchronizing a crontab
+and job manager file. I am synchronizing across computers, so I must either
+maintain a copy of the complement on each computer, and keep them synchronized,
+or figure out a computer to hold the complement, and make sure all edits go
+through that complement. Furthermore, if I want to change the synchronization to
+a different computer, I don't need to merely make sure that my files are moved
+to the computer, I also have to make sure that I correctly move the complement,
+and hook up that complement to be used for that synchronization task. This is a
+huge divergence from other synchronization utilities\bcp{No.  For example,
+  Unison maintains a database on each replica recording, for each file, what
+its hash was after the last time the two replicas were synchronized.}, where all
+that needs to be 
+done is maintain the files. Using those for synchronization, edits can be sent
+over the wire, and the appropriate function can stitch together those edits, no
+additional state needed.
+\fi
+While classical symmetric are more expressive than simple symmetric
+lenses, it is unclear whether this additional expressiveness is useful in
+practice.  Moreover, classic symmetric lenses come with 
+some disadvantages.
+For one, an implementation of classical symmetric lenses
+would need to squirrel away the complement in
+an auxiliary file somwhere, which likely requires more engineering.
+However, more significant issues arise in synthesis.
+First, because each lens has a custom complement, one can no longer
+specify the put functions through input/output examples alone.  One
+alternative would be to enrich specifications with edit sequentials;
+another would be to specify the structure of complements explicitly
+(though the latter would be somewhat akin to specifying the internal
+state of a program).  In either case, the complexity of the specifications
+increases.
 %% addition, users would have to specify some auxilary 
@@
 
-\paragraph*{Relation with Simple Symmetric Lenses} To compare classic and simple
-symmetric lenses, we define an $apply$ function on simple lenses as well. If
+\paragraph*{Relation with Simple Symmetric Lenses} To formalize how simple
+symmetric lenses work as synchronization tools, we define an $apply$ function on
+them, much like one defines an $apply$ on classical symmetric lenses: Given a
+simple symmetric lens and an optional pair of source and target strings, $apply$
+is a function from sequences of edits to sequences of edits. If
 $apply(\Lens,\None,es) = es'$, then starting with no prior data, after edit
-$es_i$, the lens \Lens generates $es_i'$ (the right format if $es_i =
-\InLOf{x}$, and the left format if $es_i = \InROf{y}$). If
-$apply(\Lens,\SomeOf{(x,y)},es) = es'$, then starting with data $x$
-and $y$ on the left and right, respectively, after edit $es_i$, the lens \Lens
-generates $es_i'$.
+$es_i$, the lens \Lens synchronizes the updates the data in the non-edited
+format (the right format if $es_i = \InLOf{x}$, and the left format if $es_i =
+\InROf{y}$) to $es_i'$. If $apply(\Lens,\SomeOf{(x,y)},es) = es'$, then that
+means, starting with data $x$ and $y$ on the left and right, respectively, after
+edit $es_i$, the lens \Lens synchronizes the other data to $es_i'$.
 
@@
 
-Next, we define \emph{forgetful symmetric
+We have reformulated symmetric lenses in a style similar to existing asymmetric
+lens formulations. However, some symmetric lenses certainly are inexpressible in
+the simple symmetric lens formulation. We formalize which are expressible
+with a restriction on symmetric lenses. We define \emph{forgetful symmetric
   lenses} to be symmetric lenses that satisfy the following additional laws.
@@
 
-Intuitively, these equations state that complements are uniquely determined by
-the most recently input $x$ and $y$. Hence, such lenses correspond exactly with
-simple symmetric lenses, where all state is maintained by the current state of
-the $x$ and $y$ data. Forgetful symmetric lenses express exactly the same
-$apply$ function as simple symmetric lenses.
+Intuitively, these equations state that complements are uniquely determined by the most
+recently input $x$ and $y$. Hence, such lenses correspond closely \dpw{replace ``closely'' with
+``exactly''?} with simple symmetric
+lenses, where all state is maintained by the current state of the $x$ and $y$
+data. Forgetful symmetric lenses express exactly the same $apply$ function as
+simple symmetric lenses.
 
@@
 \end{theorem}
+
+\bcp{I guess we could also show that equivalence of simple symmetric lenses
+  is just the identity relation?}
+\sam{No; for example the identity lens of a concatenation of two regular
+  expressions is the same as concatenating the two identity lenses on the
+  respective regular expressions}
+\afm{I think Benjamin likely means identity relation on the semantic meaning of
+  the lenses, which is true.  Would this be a useful thing to prove? It's a 2
+  liner.}
 
@@
 FlashFill (to take one well-documented example) may be connected to some of the principles
-laid out here.  For instance, Flash Fill prioritizes the substring constructor over the
-constant string constructor when ranking possible programs. Such a choice is
-consistent with our information-theoretic viewpoint as the constant function
-throws away a great deal of information about the source string being
-transformed. Likewise, Flash Fill prefers ``wider'' character classes over
-``narrower'' ones. Again such a choice is implied by information theory -- the
-wider class preserves more information during translation.  More broadly, we
-hope our information theoretic analysis provides a principle that can be
-used to understand the heuristic choices made in related work.
+laid out here.  For instance, FlashFill prioritizes the substring constructor over the
+constant string constructor when ranking possible programs. Such a choice may be understood
+in information-theoretic terms as the translation from target back to source would require
+more information.  Said another way, our information theoretic metrics measure injectivity
+(in both directions, in our case, to measure bijectivity) but applied in one direction, as
+is the case for FlashFill, would avoid the highly injective constant functions.  Beyond
+constants, another FlashFill heuristic prefers ``wider'' character classes (or token classes)
+over ``narrower'' ones.  Selecting a wider set of strings to copy from left-to-right, viewed
+in the reverse direction, from right-to-left, results in a function that is closer to being
+injective, as measured by information theory.  These are just a few of the heuristics used
+in Flashfill; more research would be required (and public disclosure of the details necessary)
+to understand the connections fully.
 
@@
 
-More broadly, information theory and probabilistic languages are common tools in
-natural language understanding, machine translation, information retrieval, data
-extraction and grammatical inference (see Pereira~\cite{pereira:info-theory},
+More broadly, information theory and probabilistic languages are common tools
+in natural language understanding,  machine translation, information retrieval,
+data extraction and grammatical inference (see Pereira~\cite{pereira:info-theory},
 for an introduction).  Indeed, our work was inspired, in part, by the PADS format
-inference tool~\cite{pads:synthesis}, which was in turn inspired by earlier work
-on probabilistic gramatical inference and information
-extraction~\cite{kushmerick-thesis,arasu:extracting}. PADS did not synthesize
-data transformers, and we are not aware of the use of information-theoretic
-principles in type-directed or syntax-guided synthesis of DSL programs.
+inference tool~\cite{pads:synthesis}, which was in turn inspired by earlier work on
+probabilistic gramatical inference
+and information extraction~\cite{kushmerick-thesis,arasu:extracting}.  More specifically, PADS
+used an information-theoretic metric to help judge the quality of inferred grammars
+for ad hoc data formats.  However, PADS did not synthesize data transformers, and
+we are not aware of the use of information-theoretic principles in type-directed or syntax-guided
+synthesis of DSL programs.
 
@@
 We have developed a synthesis algorithm for synthesizing synchronization
-functions between data formats that may not be in bijective correspondence. More
-specifically, we identify a subset of symmetric lenses, simple symmetric lenses,
-develop new combinators for them, and show how to synthesize them from regular
-expression specifications. In order to guide the search for ``likely'' lenses,
-we introduce new search principles based on information theory. Based on these
-principles, we designed, implemented, and evaluated a new tool for lens
-synthesis.
+functions between data formats that may not be in bijective correspondence. We
+do this by synthesizing symmetric lenses. However, existing symmetric lens
+formulations use complex complements to store external state. We identify a
+restriction on symmetric lenses, simple symmetric lenses. We develop combinators
+for simple symmetric lenses on string data. We show how to synthesize these
+lenses from regular expression specifications. Prior work could return
+\emph{any} satisfying lens, as the type system was restrictive enough that any
+satisfying lens would typically be the correct satisfying lens. Because of the
+increased flexibility of the symmetric lens language, we identify an
+information-theoretic heuristic for finding a likely satisfying lens. We
+developed a greedy synthesis algorithm that aims to find the lens of lowest
+cost. We demonstrate empiraclly that our algorithm can find lenses quickly, and
+our heuristic makes finding the correct lens require fewer specifications.
 
@@
   \begin{array}{@{}r@{\ }c@{\ }l@{}}
-    $\CreateR{} \App \String''''$ & = & $\StringAlt$\\
+    $\CreateR{} \App \String'$ & = & $\StringAlt$\\
     $\CreateL{} \App \StringAlt'''''$ & = & $\String$\\

### 2018-07-11 papers/SymmetricLenses/main.tex
pass through section 2

@@
 %let emp_lens = or(id(""),employees_lens) in
-  \caption{A lens, written in our language for building simple symmetric lenses
+  \caption{A lens, written in our language for defining simple symmetric lenses
     on string data, that synchronizes management and HR employee files.
-    Note that we can annotate our lenses with types, to confirm the lens
-    operates on the languages of the regular expressions. In this work, we
-    synthesize the lens itself from such regular expressions.}
+    Note that we can annotate lenses with types, ensuring that the lens
+    operates on strings that match the regular expressions. In this work, we
+    \textit{synthesize} the lens itself from such regular expressions.}
   \label{fig:example_lens}
@@
 
-The expected number of bits required to recover data is a well-known concept in
-information theory: it is the \emph{entropy} of the data~\cite{Shannon1948}.
-Calculating the entropy of data requires a probability distribution over the
+The expected number of bits required to recover data is the well-known information-theoretic concept 
+of the \emph{entropy} of the data~\cite{Shannon1948}.
+Calculating entropy requires a probability distribution over the
 space of possible values for the data. Specifically, given a set $S$ and a
-probability distribution over that set $P \OfType S \rightarrow \Reals$, the
-entropy of $S$ with probability distribution $P$ is $-\Sigma_{s\in
-  S}P(s)\Log_2P(s)$.
+probability distribution $P \OfType S \rightarrow \Reals$ over $S$, the
+entropy is $-\Sigma_{s\in  S}P(s)\Log_2P(s)$.
 
@@
 To that end, we adopt \emph{stochastic regular expressions}~\cite{stoch-def,stoch-rnn} (SREs), 
-which are regular expressions where each operator is annotated with a
+which are regular expressions in which each operator is annotated with a
 probability.  A stochastic regular expression
 specifies both a set of strings and a probability
-distribution over it.  From this information, we can calculate the
+distribution over that set, enabling us to calculate the
 entropy of the SRE.
@@
 \lstinline{name} in the management and HR data formats (we want to map first
-names to first names and last names to last names). Finally, \GreedySynth
-converts this lens back into a simple symmetric lens by converting the n-ary
+names to first names and last names to last names). \GreedySynth
+converts yields a simple symmetric lens by converting the n-ary
 forms into the binary forms provided in the surface language.

### 2018-07-11 papers/SymmetricLenses/main.tex
minor edits to section 6

@@
 extension to Boomerang. We added symmetric lens combinators to Boomerang, and
-implemented all Boomerang's asymmetric lens combinators using a combination of
+implemented all of Boomerang's asymmetric lens combinators using a combination of
 the symmetric lens combinators, as well as other symmetric lens combinators for
@@
 
-Recall that compositional synthesis exploits user annotations to break a benchmark into a series of smaller synthesis steps.
-In our experience, compositional synthesis (SS mode) allows our system to scale to arbitrarily large and complex formats.
-However, SSNC mode, which uses no such annotations and synthesizes a complete lens all at once provides a useful experimental stress test for the system.
-Recall from Section
+Recall that compositional synthesis exploits user annotations to break a
+benchmark into a series of smaller synthesis steps. In our experience,
+compositional synthesis (SS mode) allows our system to scale to arbitrarily
+large and complex formats. However, SSNC mode, which uses no such annotations
+and synthesizes a complete lens all at once provides a useful experimental
+stress test for the system.
 
@@
     given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
-    \BSOpt{} is the full bijective lens synthesis algorithm. \SSNCOpt{} is the
+    \saz{BS isn't in the figure}\BSOpt{} is the full bijective lens synthesis algorithm. \SSNCOpt{} is the
     symmetric synthesis algorithm without using a library of existing lenses.
-    \BSNCOpt{} is the bijective synthesis algorithm without using a library of
+    \saz{BSNC isn't in the figure}\BSNCOpt{} is the bijective synthesis algorithm without using a library of
     existing lenses. Our symmetric synthesis algorithm is able to complete all
@@
 For each benchmark in our benchmark suite, we ran every mode on it with a
-timeout of 30 seconds, and average the result over 5 runs.
+timeout of 30 seconds, and averaged the result over 5 runs.
 Figure~\ref{fig:times} summarizes the results of these tests. We find that our
@@
 under 30 seconds. Without compositional synthesis, the synthesis algorithm is
-able to solve 26 problem instances.
-Recall from Section~\ref{subsec:optimizations} that users may tune the termination metric.
-Success on this set of benchmarks requires we tune that termination codition (e.g.
-for some benchmarks, allowing the system flexibility to spend more time considering more of the search space).
-In Section~\ref{subsec:effect-of-heuristics}, we will consider the effect of fixing the termination metric.
+able to solve 26 problem instances. Recall from
+Section~\ref{subsec:optimizations} that users may tune the termination
+metric. Success on this set of benchmarks requires we tune that termination
+codition (e.g. for some benchmarks, allowing the system flexibility to spend
+more time considering more of the search space). In
+Section~\ref{subsec:effect-of-heuristics}, we will consider the effect of fixing
+the termination metric. \saz{If there's space: Maybe say something about how hard it was to tune?}
 
@@
 benchmarks had to be slightly altered to work with our new system. We altered
-these benchmarks by either providing additional examples termination parameters.
-Without these alterations, the symmetric synthesis synthesized a lens that fit
-the specification but not the intended lens.
+these benchmarks by either providing additional examples or tuning termination parameters.
+Without these alterations, the symmetric synthesis yielded a lens that fit
+the specification but that was not the desired lens.
 
@@
 For each benchmark in our benchmark suite, we ran every mode on it with a
-timeout of 30 seconds, and average the result over 5 runs.
+timeout of 30 seconds, and averaged the result over 5 runs.
 Figure~\ref{fig:times_bijective} summarizes the results of these tests. On
@@
 We evaluate the usefulness of (1) our information theoretic metric, and (2) our
-termination heuristic.  To this end, we run our program in 3 different modes:\\
+termination heuristic.  To this end, we run our program in several different modes:\\
 \begin{tabulary}{\linewidth}{rL}
@@
 Related work on string transformations has often used simpler measures such as ``avoid constants'' which align with, but are simpler than our measures.
-The \CCOpt{} mode is an example of such a simple measure --- it operates by counting disconnects, which put a complete stop to information transfer, like constants, and merges, which eliminate the information in a union.
+The \CCOpt{} mode is an example of such a simple measure---it operates by counting disconnects, which put a complete stop to information transfer, like constants, and merges, which eliminate the information in a union.
 Figure~\ref{fig:metric} summarizes the result of this experiment.

### 2018-07-11 papers/SymmetricLenses/main.tex
add new run data and fill in concrete numbers

--- 2018-07-11 papers/SymmetricLenses/main.tex
@@
 evaluate
-its performance on 45 benchmark examples drawn from FlashFill, Augeus,
+its performance on 48 benchmark examples drawn from FlashFill, Augeus,
 and an online collection of electronic file formats. We find this implementation is
@@
 some file-specific meta data, such as a date, a time or a reference
-number, whereas the other does not. Indeed of the 45 benchmark problems
+number, whereas the other does not. Indeed of the 48 benchmark problems
 described in Section \ref{sec:evaluation}, all of the benchmarks
@@
 synchronize between two ad hoc file formats have this 
-characteristic, \afm{num} benchmarks in total. Consequently, tools that search
+characteristic, 48 benchmarks in total. Consequently, tools that search
 only for strict bijections will rarely be applicable in practice.
@@
   \includegraphics[width=.5\textwidth]{high-level-algorithm.pdf}
-  \caption{Schematic diagram for the simple symmetric lens synthesis
-  algorithm. The user provides 
-   regular expressions \BRegex and \BRegexAlt and a set of examples \Examples{}
-    as input. \Expand first converts \BRegex and \BRegexAlt to
-    stochastic regular expressions \Regex and \RegexAlt with default
+  \caption{Schematic diagram for the simple symmetric lens synthesis algorithm.
+    The user provides regular expressions \BRegex and \BRegexAlt and a set of
+    examples \Examples{} as input. \Expand first converts \BRegex and \BRegexAlt
+    to stochastic regular expressions \Regex and \RegexAlt with default
     probabilities. It then finds pairs of stochastic regular expressions
-    equivalent to \Regex and \RegexAlt and iteratively proposes them to \GreedySynth.
-    \GreedySynth finds a lens typed between the supplied SREs.
-    When the algorithm determines it has found a likely lens, it
-  returns it.  \ksf{Can we change cost to likelihood and l\_likely to l\_best?}
-   }
+    equivalent to \Regex and \RegexAlt and iteratively proposes them to
+    \GreedySynth. \GreedySynth finds a lens typed between the supplied SREs.
+    When the algorithm determines it has found a likely lens, it returns it.}
   \label{fig:high-level-algorithm}
@@
 We implemented the \GreedySynth and \Expand algorithms and evaluated their
-effectiveness at synthesizing lenses on a 
-benchmark suite of 45 problems, including data cleaning tasks taken
-from Flash Fill~\cite{gulwani-popl-2014}, view maintenance
-tasks taken from Augeas~\cite{Augeas}, and ad hoc format synchronization tasks
-taken from a variety of other locations\footnote{The majority of these format
-  synchronization tasks were found from the following index:
-  \url{http://fileformats.archiveteam.org/wiki/Electronic_File_Formats}}.
+effectiveness at synthesizing lenses on a benchmark suite of 48 problems,
+including data cleaning tasks taken from Flash Fill~\cite{gulwani-popl-2014},
+view maintenance tasks taken from Augeas~\cite{Augeas}, and ad hoc format
+synchronization tasks taken from a variety of other locations\footnote{The
+  majority of these format synchronization tasks were derived from the
+  bidirectional programming literature.}.
 
@@
   lenses.  We evaluate our
-  implementation on a set of 45 lens synthesis benchmarks drawn from
+  implementation on a set of 48 lens synthesis benchmarks drawn from
   data cleaning, view maintenance, 
@@
 typically calculate the conditional entropy exactly, as merges are relatively
-uncommon on practice; only \afm{2} of the lenses sythesized in our benchmarks
+uncommon on practice; only 2 of the lenses sythesized in our benchmarks
 include merges.
@@
 to our information theoretic metric, and
-\item adds the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)
-  to the priority queue.  The priority of a SRE pair is the
+\item adds all of the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)
+  to the priority queue with $\PCF{\PQ.Push}$. The priority of a SRE pair is the
   number of rewrites needed to derive the pair from the originals.
@@
 When the loop terminates, the search returns the globally best lens found
-($best$).  Each subroutine of this algorithm will be explained in
+($best$).  Each element of this algorithm will be explained in
 further depth in the following subsections.
 
-\begin{algorithm}[t]
+\begin{algorithm}
   \caption{\SynthSymLens}
@@
 expressions that are semantically, but not syntactically, equivalent to the
-originals by performing the following rewrites:
+originals. In particular, this synthesizer performs the following rewrites:
 \begin{center}
@@
 
-The priority queue yields stochastic regular expressions generated using fewer
+Our priority queue proposes stochastic regular expressions generated from fewer
 rewrites first. Only when there are no more proposed regular expressions derived
@@
 
-When $\Call{Continue}{pq}$ returns $\False$, \Expand stops proposing regular
+When $\Call{Continue}{pq}$ returns $\False$, \RXSearch stops proposing regular
 expression pairs, and instead returns to the user the best lens found thus far.
-Note this means that if the algorithm finds a bijective lens, which has zero cost, it
+Note this means that if \RXSearch finds a bijective lens, which has zero cost it
 will immediately return.
@@
 (\EG, \emph{Unrollstar$_L$} and \emph{Unrollstar$_R$}) are needed to relate
-two \SDNFREabbrevs; other axioms, like associativity, commutativity
+two \SDNFREabbrevs, other axioms, like associativity, commutativity
 or distributivity, are unneeded.  This important property makes it possible to
@@
 
-Intuitively, stochastic DNF regular expressions are stochastic
+Intuitively, stochastic DNF regular expressions are essentially stochastic
 regular expressions with all concatenations fully distributed over all
@@
 As these DNF regular expressions are \emph{stochastic}, they are annotated with
-probabilities to express a probability distribution, in addition to a
-language. \saz{I find the use of asterisk for the $\times$ or $\cdot$ or maybe
-  just juxtaposition slightly jarring.}
+probabilities to express a probability distribution, in addition to a language. 
 \begin{center}
@@
 
-The algorithm for converting a stochastic regular expressions into its
-corresponding stochastic DNF regular expressions, written
+We have a means of turning stochastic regular expressions into stochastic DNF
+regular expressions. The conversion algorithm itself, written
 $\ToDNFRegexOf{\Regex}$, is defined below.
@@
 \]
-This conversion relies on operators defined in
+This conversion utilizes operators defined in
 Figure~\ref{fig:dnf-regex-functions}. We have proved that this conversion
@@
 \paragraph*{\ToStochastic} With stochastic DNF regular expressions and
-\ToDNFRegex defined, it is easier to explain \ToStochastic, the function that converts regular
-expressions into stochastic regular expressions. If users
-provided stochastic regular expressions themselves (perhaps inferred from
+\ToDNFRegex defined, it is easier to explain the function that converts regular
+expressions into stochastic regular expressions, \ToStochastic. If users
+specified stochastic regular expressions themselves (perhaps inferred from
 example data), this step could be skipped,
@@
 \label{subsec:greedy-synth}
-The synthesis procedure comprises three algorithms: one that greedily finds
+The greedy algorithm is actually three algorithms, one that greedily finds
 symmetric DNF lenses (\GreedySynth), one that greedily finds symmetric sequence
@@
 the structure of symmetric DNF lenses and \SDNFREabbrevs. We describe this
-synthesis algorithm using the running example of searching for a lens between
+synthesis algorithm in the context of trying to find a lens between
 \lstinline{"" | name.name*} and \lstinline{"" | name}.
@@
 \paragraph*{Symmetric DNF Lenses} Algorithm~\ref{alg:greedysynth} presents
-\GreedySynth, which synthesizes symmetric DNF lenses. Its inputs 
-are a suite of input-output examples and a pair of stochastic DNF regular expressions.
-%
-First, \PCF{CannotMap} determines whether there is no lens satisfying the
-examples, and if so, \GreedySynth returns \None immediately. Otherwise, \GreedySynth finds
+\GreedySynth, a means to synthesize symmetric DNF lenses. The inputs to
+\GreedySynth are a suite of input-output examples and a pair of stochastic DNF regular expressions.
+
+First, \PCF{CannotMap} checks whether there is no possible lens satisfying the
+examples. If so, \GreedySynth returns \None early. Otherwise, \GreedySynth finds
 the best lenses (given the examples that match them) between all sequence pairs
@@
 queue. Within this loop, a sequence lense is popped from the queue and, if it is
-``useful,'' it is included in the final DNF lens. The lens is considered to
-be useful when its source (or target) is \textit{not} already the source (or target) of
+``useful'', then it is included in the final DNF lens. The lens is considered to
+be useful when its source (or target) is not already the source (or target) of
 an already chosen sequence lens. If the examples require that two sequences have a
@@
 
-\begin{algorithm}[t]
+\begin{algorithm}
   \caption{\GreedySynth}
@@
 presents \GreedySeqSynth, which synthesizes symmetric sequence lenses using an
-algorithm whose structure is similar to \GreedySynth{}'s.  It calls \PCF{AtomSynth}, 
-which synthesizes atom lenses by iterating a DNF lens between
+algorithm whose structure is similar to \GreedySynth{}'s.  It calls \PCF{AtomSynth}, a
+function that synthesizes atom lenses by iterating a DNF lens between
 its subcomponents.
@@
 \GreedySeqSynth then loops until there are no more atom lenses in the priority
-queue. Within the loop, a popped atom lens is considered 
-``useful'' if adding it to the sequence will lower the
-cost of the generated sequence lens, or if examples show that one of its atoms
-must not be disconnected. Each atom can be part of only one lens at a time, so
-the algorithm must sometimes a previously chosen atom lens in order to connect an
+queue. Within this loop, atom lenses are popped from the queue and the
+``useful'' ones are kept. An atom lens is useful if adding it will lower the
+cost of the generated sequence lens, or if examples show that one of the atoms
+must not be disconnected. Each atom can only be part of one lens at a time, so
+the algorithm must at times disconnect previously connected atoms to connect an
 atom that must not be disconnected.
-After all atoms have been popped, the sequence lens is validated and
-returned. The algorithm succeeds when all atoms that must not be disconnected
-are involved in an atom lens, but \PCF{SLensBuilder.ToDNFLens}
-returns \None otherwise.
-
-\begin{algorithm}[t]
+
+Finally, after all atoms have been popped, the sequence lens is validated and
+returned. The algorithm confirms that all atoms that must not be disconnected
+are involved in an atom lens. If they are not, \PCF{SLensBuilder.ToDNFLens}
+instead returns \None.
+
+\begin{algorithm}
   \caption{\GreedySeqSynth}
@@
 
-Our implementation includes a number of optimizations not described above:
-annotations that guide DNF conversion; an expansion inference
-algorithm; hash-consing and memoization; and compositional synthesis. The
-implementation also provides for an additional specification parameter that
-customizes the termination condition for the synthesis procedure.
+Our implementation includes a number of optimizations to this algorithm. We
+added ``closed'' annotations to regular expressions \saz{what do these mean? I
+  guess ``open'' means ``subject to DNF conversion'' and ``closed'' means
+  not?}\afm{I describe in greater detail below, should I add in additional info
+  here too?}, an expansion inference algorithm, hash-consing and memoization,
+along with compositional synthesis. The implementation also allows an additional
+parameter in specifications that customizes the termination condition for the
+synthesis procedure.
 
@@
 relatively efficiently, it suffers from an exponential blowup when converting
-regular expressions to DNF form. This can be mitigated by not converting
-\emph{all} of the regular expression to DNF form, but only converting
-\emph{some} of it, and doing so incrementally. This change is easy
-to support by adding non-DNF stochastic regular expressions as a new type of
+regular expressions to DNF form. This can be minimized by not converting
+\emph{all} of the regular expression into DNF form, but only converting
+\emph{some} of it to DNF form, and doing so incrementally. This change is easy
+to make by adding non-DNF stochastic regular expressions as a new type of
 ``closed'' atom. When converted into a DNF regular expression, if $\Regex$ is
@@
 
-At the start of synthesis, all regular expressions are closed. \Expand is
+At the start of synthesis, all regular expressions are closed. \RXSearch is
 modified so that it is now permitted to rewrite closed regular expressions to
 open ones (thereby triggering DNF normalization). \GreedySynth is allowed to to
-map closed regular expressions to each other, but only by using the identity
+map closed regular expressions to each other, but only using the identity
 lens---they must parse the same string components in all examples.
 
-\paragraph*{Expansion Inference} Distinguishing open and closed regular
-expressions helps \GreedySynth, but doing so means that \Expand has the harder job of
-determining when to open closed expressions.  To that end,
-expansion inference identifies when a regular expression cannot
-possibly map by any means other than the ``trivial'' disconnect lens. Expansion inference
-opens those regular expressions to allow for further exploration.
-
-\paragraph*{Hash-consing and Memoization} We further improve \GreedySynth by
+\paragraph*{Expansion Inference} The prior optimization helped \GreedySynth, but
+it made \RXSearch have the harder job of determining when to open closed
+expressions.  To alleviate this difficulty, we added in expansion inference,
+which identifies when a regular expression cannot possibly map by any means
+other than the disconnect lens.  Expansion inference opens those regular
+expressions to allow for further exploration.
+
+\paragraph*{Hash-consing and Memoization} We further help \GreedySynth by
 memoizing its results.  However, our stochastic DNF regular expressions are
@@
 mapped to each other either if they are the same regular expression and agree
-with the examples, or if there is an existing lens (which is a composition of predefined
-lenses) that already agrees with the examples.
+with the examples, or if there is a lens which is a composition of predefined
+lenses between them, where the lens agrees with the examples.
 
 \paragraph*{Termination Customization}
-In this work, there are three competing constraints. (1) We want to return
-lenses quickly to users, (2) we want to return good lenses, and (3) we don't
-want to require users to write a large number of examples in their
+In this work, we are constrained by three competing constraints. (1) We want to
+return lenses quickly to users, (2) we want to return good lenses, and (3) we
+don't want to require users to write a large number of examples in the
 specifications. Our termination heuristic works well for most situations, but
-there are cases where a large number of expansions must be performed to find the
-correct lens.  Users could work around that problem by providing a large number
-of examples, but it can be difficult to write a large number of informative
-examples for complex formats. We find it simpler to instead let the users
-specify that \SynthSymLens should be more restrictive in which lenses it
-accepts.  Conversely, if the user's desired transformation has a particularly
-high cost, the algorithm may spend a long time searching in vain for lenses with
-lower cost. By tuning \SynthSymLens to be more permissive, such a search can be
-short-circuited.
-
-For these situations, users can provide a number, $t$, which affects
-the strictness of the termination condition, adjusting the \Continue formula
-as follows. When $\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at
-distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
-current best lens has cost $c$:
+there are situations where a large number of expansions must be performed to
+find the correct lens.
+
+Users could prevent incorrect lenses with a large number of examples, but it can
+be quite difficult to write a large number of examples for complex formats. We
+find it simpler to let the users specify that \RXSearch should be more
+restrictive in which lenses it accepts, so users do not have to write a lot of
+examples.
+
+This constraint goes the other way too. If the user's desired transformation has
+a particularly high cost, the algorithm may spend a long time searching in vain
+for lenses with lower cost. By tuning \RXSearch to me more permissive in the
+lenses it accepts, this search can be short-circuited.
+
+For these types of situations, users can input a custom number, $t$, to
+customize the strictness of the termination condition. In particular, if
+$\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at distance $d$, the
+number of pairs in $pq$ at distance $d$ is $n$, and the current best lens has
+cost $c$:
 \begin{gather*}
@@
 \end{gather*}
-With a positive $t$, the termination condition is stricter, and the algorithm
+With a positive $t$, the termination condition is more strict, and the algorithm
 will spend longer searching for more likely lenses. With a negative $t$ the
 termination condition is less strict, and the algorithm will spend less time
-searching for more likely lenses.  Our default $t$ is $0$.
+searching for more likely lenses.  Our default $t$ is set to $0$.
 
 While our defaults termination condition quickly finds the correct lens for
-\afm{35} of our \afm{45} benchmarks, our termination condition is not
-one-size-fits all. Complex lenses, or lenses that project a
-large amount of data, can be synthesized more easily and quickly by
-adjusting $t$.
+39 of our 48 benchmarks, our termination condition is not
+one-size-fits all. Particularly complex lenses, or lenses that project a
+particularly large amount of data, can be synthesized more easily and quickly by
+customizing how permissive \RXSearch is in returning lenses.
 % end synthesis
@@
 \begin{enumerate}
-\item We adapted \afm{8} data cleaning benchmarks from Flash Fill, though, of course, our tool produces bidirectional transformations rather than one-way transformers like Flash Fill.
-\item We adapted \afm{29} benchmarks from Augeas. These benchmarks turn the
+\item We adapted 8 data cleaning benchmarks from Flash Fill, though, of course, our tool produces bidirectional transformations rather than one-way transformers like Flash Fill.
+\item We adapted 29 benchmarks from Augeas. These benchmarks turn the
   ad hoc data present in Linux configuration files into a structured, dictionary
   representation.
-\item We used \afm{8} benchmarks from synchronization tasks between a
-  variety of real-life formats. These tasks range from synchronizing REST
-  and JSON web resource descriptions to synchronizing BibTeX and EndNote
-  citation descriptions.
+\item We used 11 benchmarks from synchronization tasks between a variety of
+  formats taken from the bidirectional programming literature. These tasks range
+  from synchronizing REST and JSON web resource descriptions to synchronizing
+  BibTeX and EndNote citation descriptions.
 \end{enumerate}
@@
 To compare to the existing bijective synthesis algorithm, we run our symmetric
-synthesis algorithm on the original Optician benchmark suite. \afm{8} of these
+synthesis algorithm on the original Optician benchmark suite. 4 of these
 benchmarks had to be slightly altered to work with our new system. We altered
@@
 For each benchmark in our benchmark suite, we ran every mode on it with a
-timeout of 30 seconds, and average the result over 5 runs.
+timeout of 60 seconds, and average the result over 5 runs.
 Figure~\ref{fig:times_bijective} summarizes the results of these tests. On
-average, \SSOpt{} took 2.1 times (\afm{3.5} seconds) longer to complete than
+average, \SSOpt{} took 1.3 times (0.5 seconds) longer to complete than
 \BSOpt{}. The slowest completed benchmark for both synthesis algorithms is
@@
 information-theoretic metric is critical for finding the correct lens. Indeed,
-only \afm{10} of our benchmarks succeeded when running in \CCOpt{} mode.
+only 10 of our benchmarks succeeded when running in \CCOpt{} mode.
 
 Our termination condition is also quite important. When running in \FLOpt{},
-only \afm{5} lenses are found. The first cluster in which a lens is found is
+only 5 lenses are found. The first cluster in which a lens is found is
 rarely the correct cluster. However, our termination condition is not perfect.
@@
 Without termination parameters, our algorithm finds the correct lens for 36 of
-our 45 benchmarks, 9 of them need help in the form of a termination parameter
+our 48 benchmarks, 9 of them need help in the form of a termination parameter
 for finding the correct lens. That being said, our default parameter finds a

### 2018-07-11 papers/SymmetricLenses/main.tex

--- 2018-07-11 papers/SymmetricLenses/main.tex
@@
 While our defaults termination condition quickly finds the correct lens for
-\afm{35} of our \afm{45} benchmarks, our termination condition is not
+39 of our 48 benchmarks, our termination condition is not
 one-size-fits all. Complex lenses, or lenses that project a
@@
 For each benchmark in our benchmark suite, we ran every mode on it with a
-timeout of 30 seconds, and averaged the result over 5 runs.
+timeout of 60 seconds, and averaged the result over 5 runs.
 Figure~\ref{fig:times_bijective} summarizes the results of these tests. On
-average, \SSOpt{} took 2.1 times (\afm{3.5} seconds) longer to complete than
+average, \SSOpt{} took 1.3 times (0.5 seconds) longer to complete than
 \BSOpt{}. The slowest completed benchmark for both synthesis algorithms is

### 2018-07-11 papers/SymmetricLenses/main.tex
minor nit fixes
edits to 5

@@
 \paragraph*{Open and Closed Regular Expressions} While \GreedySynth acts
-relatively efficiently, it suffers from an exponential blowup when converting
-regular expressions to DNF form. This can be mitigated by not converting
-\emph{all} of the regular expression to DNF form, but only converting
-\emph{some} of it, and doing so incrementally. This change is easy
-to support by adding non-DNF stochastic regular expressions as a new type of
-``closed'' atom. When converted into a DNF regular expression, if $\Regex$ is
-annotated as ``closed'', then $\ToDNFRegexOf{\Regex} = \AtomToDNFOf{\Regex}$.
-Regular expressions that are not annotated as closed, are considered ``open.''
-
-At the start of synthesis, all regular expressions are closed. \Expand is
-modified so that it is now permitted to rewrite closed regular expressions to
-open ones (thereby triggering DNF normalization). \GreedySynth is allowed to to
-map closed regular expressions to each other, but only by using the identity
-lens---they must parse the same string components in all examples.
-
-\paragraph*{Expansion Inference} Distinguishing open and closed regular
-expressions helps \GreedySynth, but doing so means that \Expand has the harder job of
-determining when to open closed expressions.  To that end,
-expansion inference identifies when a regular expression cannot
-possibly map by any means other than the ``trivial'' disconnect lens. Expansion inference
-opens those regular expressions to allow for further exploration.
+relatively efficiently, it can suffer from an exponential blowup when converting
+regular expressions to DNF form. This problem can be mitigated by avoiding the conversion
+of some REs to DNF form and by performing the conversion lazily when necessarily.
+More specifically, \Expand labels some uncoverted REs as ``closed,'' which means the
+type-directed \GreedySynth algorithm treats them as DNF atoms and does not dig into
+them recursively.  In other words,
+given a pair of closed REs, \GreedySynth can either construct
+the identity lens between them (it will do this if they are the same RE), or it can
+construct a disconnect lens between them.
+Regular expressions that are not annotated as closed are considered ``open.''
+%is allowed to to
+%map closed regular expressions to each other, but only by using the identity
+%lens---they must parse the same string components in all examples.
+%This change is easy
+%to support by adding non-DNF stochastic regular expressions as a new type of
+%``closed'' atom. When converted into a DNF regular expression, if $\Regex$ is
+%annotated as ``closed'', then $\ToDNFRegexOf{\Regex} = \AtomToDNFOf{\Regex}$.
+
+\paragraph*{Expansion Inference} Distinguishing between open and closed regular
+expressions improves the efficiency of \GreedySynth, but forces \Expand to decide
+which closed expressions to open.
+At the start of synthesis, all regular expressions are closed. \Expand 
+rewrites selected closed regular expressions to
+open ones (thereby triggering DNF normalization)
+when the disconnect lens is the only lens that will ever apply.
 
 \paragraph*{Hash-consing and Memoization} We further improve \GreedySynth by
-memoizing its results.  However, our stochastic DNF regular expressions are
-quite large and so hashing them is computationally intensive.  We use less
-memory and allow for quick comparisons and hashing of these DNF regular
-expressions by hash-consing them.
+hash-consing stochastic DNF regular expressions.  This optimization enables
+fast equality checks and saves memory.\dpw{we could probably cut this paragraph.}
+%memoizing its results.  However, our stochastic DNF regular expressions are
+%quite large and so hashing them is computationally intensive.  We use less
+%memory and allow for quick comparisons and hashing of these DNF regular
+%expressions by
 
 \paragraph*{Compositional Synthesis} Compositional synthesis allows us to use
-previously defined lenses---whether they are previously defined by users or by
-synthesis. We alter \GreedySynth such that two regular expression atoms can be
-mapped to each other either if they are the same regular expression and agree
-with the examples, or if there is an existing lens (which is a composition of predefined
-lenses) that already agrees with the examples.
+previously defined lenses, whether they are previously defined by users or by
+synthesis. As the synthesizer processes a Boomerang file, it accumulates
+lens definitions and their types.  It tackles synthesis problems one after another
+and there may be many such problems in a given program.
+During synthesis, if an existing lens has the right type and agrees with the examples.
+\GreedySynth will use it.  Consequently, if a large synthesis problem cannot be solved all
+at once, a user can first generate subproblems for parts of a data format, and the
+larger problem can subsequently use solutions to those subproblems.  In our experience,
+this is a very powerful tool that allows synthesis to tackle problems of arbitrary
+complexity.
 
 \paragraph*{Termination Customization}
-In this work, there are three competing constraints. (1) We want to return
-lenses quickly to users, (2) we want to return good lenses, and (3) we don't
-want to require users to write a large number of examples in their
-specifications. Our termination heuristic works well for most situations, but
+In this work, there are three competing constraints. The tool should (1) return
+lenses quickly to users, (2) return the highest quality lenses, and (3)
+not require users to write a large number of examples.
+Our termination heuristic works well for most situations, but
 there are cases where a large number of expansions must be performed to find the
-correct lens.  Users could work around that problem by providing a large number
-of examples, but it can be difficult to write a large number of informative
-examples for complex formats. We find it simpler to instead let the users
-specify that \SynthSymLens should be more restrictive in which lenses it
-accepts.  Conversely, if the user's desired transformation has a particularly
+correct lens.  Users can work around such a problem by providing a large number
+of examples, but constructing examples is a chore, especially when formats are
+complicated.  We have often found it simpler to allow users to change the
+synthesis termination condition to demand \SynthSymLens be more judgemental
+when it comes to deciding it has found a ``good enough'' lens.  
+Conversely, if the user's desired transformation has a particularly
 high cost, the algorithm may spend a long time searching in vain for lenses with
@@
 
-For these situations, users can provide a number, $t$, which affects
+For these situations, users can provide a number, $t$, that affects
 the strictness of the termination condition, adjusting the \Continue formula
@@
 With a positive $t$, the termination condition is stricter, and the algorithm
-will spend longer searching for more likely lenses. With a negative $t$ the
+will spend longer searching for more likely lenses. With a negative $t$, the
 termination condition is less strict, and the algorithm will spend less time
@@
 
-While our defaults termination condition quickly finds the correct lens for
+As discussed in the following section,
+while our default termination condition quickly finds the correct lens for
 39 of our 48 benchmarks, our termination condition is not
@@
 large amount of data, can be synthesized more easily and quickly by
-adjusting $t$.
+adjusting $t$. \dpw{should this paragraph be cut and its content left for the
+next section?}
 % end synthesis
@@
 \item We used 11 benchmarks from synchronization tasks between a variety of
-  formats taken from the bidirectional programming literature. These tasks range
-  from synchronizing REST and JSON web resource descriptions to synchronizing
-  BibTeX and EndNote citation descriptions.
+  formats derived from real-world examples and the bidirectional programming
+  literature. These tasks range from synchronizing REST and JSON web resource
+  descriptions to synchronizing BibTeX and EndNote citation descriptions.
 \end{enumerate}
@@
 
-\subsection{Useful in Everyday Development}
+\subsection{Usefulness in Everyday Development}
 

### 2018-07-11 papers/SymmetricLenses/main.tex
add some changes to optimizations

@@
 Regular expressions that are not annotated as closed are considered ``open.''
+
+Distinguishing between open and closed regular expressions improves the
+efficiency of \GreedySynth, but forces \Expand to decide which closed
+expressions to open. At the start of synthesis, all regular expressions are
+closed, and \Expand rewrites selected closed regular expressions to open ones
+(thereby triggering DNF normalization).
 %is allowed to to
@@
 
-\paragraph*{Expansion Inference} Distinguishing between open and closed regular
-expressions improves the efficiency of \GreedySynth, but forces \Expand to decide
-which closed expressions to open.
-At the start of synthesis, all regular expressions are closed. \Expand 
-rewrites selected closed regular expressions to
-open ones (thereby triggering DNF normalization)
-when the disconnect lens is the only lens that will ever apply.
+\paragraph*{Expansion Inference} These additional rewrites make the search
+through possible regular expressions much harder. Our algorithm automatically
+will identify when certain closed regular expressions can \emph{only} be
+involved in a disconnect lens.  Such regular expressions will automatically be
+opened.
 

### 2018-07-11 papers/SymmetricLenses/main.tex
respond to steve comments

@@
     given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
-    \saz{BS isn't in the figure}\BSOpt{} is the full bijective lens synthesis algorithm. \SSNCOpt{} is the
-    symmetric synthesis algorithm without using a library of existing lenses.
-    \saz{BSNC isn't in the figure}\BSNCOpt{} is the bijective synthesis algorithm without using a library of
+    \SSNCOpt{} is the symmetric synthesis algorithm without using a library of
     existing lenses. Our symmetric synthesis algorithm is able to complete all
@@
 able to solve 26 problem instances. Recall from
-Section~\ref{subsec:optimizations} that users may tune the termination
-metric. Success on this set of benchmarks requires we tune that termination
-codition (e.g. for some benchmarks, allowing the system flexibility to spend
-more time considering more of the search space). In
+Section~\ref{subsec:optimizations} that users may tune the termination metric.
+Success on this set of benchmarks requires we tune that termination codition
+(e.g. for some benchmarks, allowing the system flexibility to spend more time
+considering more of the search space). In
 Section~\ref{subsec:effect-of-heuristics}, we will consider the effect of fixing
-the termination metric. \saz{If there's space: Maybe say something about how hard it was to tune?}
+the termination metric. In practice, we found customizing the termination metric
+to be quite easy: if manual inspection of the lens showed there were too few
+bijections, we would increase the termination parameter. If synthesis took too
+long, we would decrease it.
 

### 2018-07-11 papers/SymmetricLenses/main.tex
no more asterisks for times
add nit changes

@@
 described in Section \ref{sec:evaluation}, all of the benchmarks
-taken from Flash Fill~\cite{flashfill} and all of the benchmarks that
+taken from Flash Fill~\cite{flashfill} and many of the benchmarks that
 synchronize between two ad hoc file formats have this 
-characteristic, 48 benchmarks in total. Consequently, tools that search
-only for strict bijections will rarely be applicable in practice.
+characteristic, 13 benchmarks in total. Consequently, tools that search
+only for strict bijections are often not applicable in practice.
 
@@
 and search using similar heuristics.  However, 
-this naive approach runs into two difficulties.
+this na\"ive approach runs into two difficulties.
 
@@
 the desired lens. In contrast, and to our initial surprise, experiments showed
-that if a naive search algorithm selects the first simple symmetric lens it
+that if a na\"ive search algorithm selects the first simple symmetric lens it
 finds, it is typically \emph{not} the desired lens.  The flexibility
@@
 let emp_salary = name . " " . name . ": " salary
-let emp_salaries = "" | emp_salary . ("\n" emp_salary)*
+let emp_salaries = "" | emp_salary . ("\n" emp_salary)$^*$
 \end{lstlisting}
@@
 let header = "Name,Company" in
-let emp_insurance = header . ("\n" . emp_ins)*
+let emp_insurance = header . ("\n" . emp_ins)$^*$
 \end{lstlisting}
@@
   $\Regex$ given $\RegexAlt$ and $\Lens$}, formally $\sum_{s \in
-  \Regex}\ProbabilityOf{\Regex}{s}*\EntropyOf{\SetOf{t \SuchThat
+  \Regex}\ProbabilityOf{\Regex}{s}\EntropyOf{\SetOf{t \SuchThat
     \Lens.\PutROf{s}{t} = t}}$. The likelihood we assign to $\Lens$ is the sum
@@
     & =
-    & $\Sigma_{\String = \String_1\String_2}\ProbabilityOf{\Regex_1}{\String_1}*\ProbabilityOf{\Regex_2}{\String_2}$ \\
+    & $\Sigma_{\String = \String_1\String_2}\ProbabilityOf{\Regex_1}{\String_1}\ProbabilityOf{\Regex_2}{\String_2}$ \\
 
@@
     & =
-    & $\Probability * \ProbabilityOf{\Regex_1}{\String} +
-      (1-\Probability) * \ProbabilityOf{\Regex_2}{\String}$\\
+    & $\Probability\ProbabilityOf{\Regex_1}{\String} +
+      (1-\Probability)\ProbabilityOf{\Regex_2}{\String}$\\
 
@@
     & =
-    & $\Sigma_n \Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\Regex}{\String_i}$\\
+    & $\Sigma_n \Sigma_{\String = \String_1 \ldots \String_n}\Probability^n(1-\Probability)\Pi_{i=1}^n\ProbabilityOf{\Regex}{\String_i}$\\
   \end{tabular}
@@
     \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
-    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{\frac{(1-\Probability_1)*\Probability_2}{1 - \Probability_1 * \Probability_2}})}{\Probability_1*\Probability_2} & \OrAssociativityRule{}  \\
+    \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{\frac{(1-\Probability_1)\Probability_2}{1 - \Probability_1\Probability_2}})}{\Probability_1\Probability_2} & \OrAssociativityRule{}  \\
     \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
@@
     & =
-    & $\Probability*(\EntropyOf{\Regex} - \Log_2(\Probability)) + (1-\Probability)*(\EntropyOf{\RegexAlt} - \Log_2(1-\Probability))$\\
+    & $\Probability(\EntropyOf{\Regex} - \Log_2(\Probability)) + (1-\Probability)(\EntropyOf{\RegexAlt} - \Log_2(1-\Probability))$\\
   \end{tabular}
@@
     & =
-    & $\frac{\Probability}{1-\Probability}(\REntropyOf{\RegexAlt \Given \Lens, \Regex})$\\
+    & $\frac{\Probability}{1-\Probability}\REntropyOf{\RegexAlt \Given \Lens, \Regex}$\\
 
@@
     & =
-    & $\Probability*(\REntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt_1} ) + (1-\Probability)*(\REntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2})$\\
+    & $\Probability\REntropyOf{\Regex_1 \Given \Lens_1, \RegexAlt_1} + (1-\Probability)\REntropyOf{\Regex_2 \Given \Lens_2, \RegexAlt_2}$\\
 
@@
     & =
-    & $\Probability*(\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1}) + (1-\Probability)*(\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2})$\\
+    & $\Probability\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1} + (1-\Probability)\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2}$\\
 
@@
 probabilities to express a probability distribution, in addition to a
-language. \saz{I find the use of asterisk for the $\times$ or $\cdot$ or maybe
-  just juxtaposition slightly jarring.}
+language.
 \begin{center}
@@
     & =
-    & $\Sigma_n \Sigma_{\String = \String_1 \ldots \String_n}\Probability^n*(1-\Probability)*\Pi_{i=1}^n\ProbabilityOf{\DNFRegex}{\String_i}$\\
+    & $\Sigma_n \Sigma_{\String = \String_1 \ldots \String_n}\Probability^n(1-\Probability)\Pi_{i=1}^n\ProbabilityOf{\DNFRegex}{\String_i}$\\
 
@@
     & =
-    & $\Sigma_{i=1}^n\Probability_i * \ProbabilityOf{\Sequence_i}{\String}$\\
+    & $\Sigma_{i=1}^n\Probability_i\ProbabilityOf{\Sequence_i}{\String}$\\
   \end{tabular}
@@
   $\ConcatDNFOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}{\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}=$\\
-      $\DNFLeft (\ConcatSequenceOf{\Sequence_1}{\SequenceAlt_1},\Probability_1*\ProbabilityAlt_1)\DNFSep \ldots
+      $\DNFLeft (\ConcatSequenceOf{\Sequence_1}{\SequenceAlt_1},\Probability_1\ProbabilityAlt_1)\DNFSep \ldots
       \DNFSep
-      (\ConcatSequenceOf{\Sequence_1}{\SequenceAlt_m},\Probability_1*\ProbabilityAlt_m)\DNFSep
+      (\ConcatSequenceOf{\Sequence_1}{\SequenceAlt_m},\Probability_1\ProbabilityAlt_m)\DNFSep
       \ldots$\\
       $\DNFSep
-      (\ConcatSequenceOf{\Sequence_n}{\SequenceAlt_1},\Probability_n*\ProbabilityAlt_1)\DNFSep
+      (\ConcatSequenceOf{\Sequence_n}{\SequenceAlt_1},\Probability_n\ProbabilityAlt_1)\DNFSep
       \ldots \DNFSep
-      (\ConcatSequenceOf{\Sequence_n}{\SequenceAlt_m},\Probability_n * \ProbabilityAlt_m) \DNFRight$
+      (\ConcatSequenceOf{\Sequence_n}{\SequenceAlt_m},\Probability_n\ProbabilityAlt_m) \DNFRight$
 
@@
   $\OrDNFOf{\DNFOf{(\Sequence_1,\Probability_1)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n)}}{\DNFOf{(\SequenceAlt_1,\ProbabilityAlt_1)\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m)}}{\Probability} =$\\
-  $\DNFOf{(\Sequence_1,\Probability_1*\Probability)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n*\Probability)\DNFSep(\SequenceAlt_1,\ProbabilityAlt_1*(1-\Probability))\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m*(1-\Probability))}$
+  $\DNFOf{(\Sequence_1,\Probability_1\Probability)\DNFSep\ldots\DNFSep(\Sequence_n,\Probability_n\Probability)\DNFSep(\SequenceAlt_1,\ProbabilityAlt_1(1-\Probability))\DNFSep\ldots\DNFSep(\SequenceAlt_m,\ProbabilityAlt_m(1-\Probability))}$
 
@@
 synthesis algorithm using the running example of searching for a lens between
-\lstinline{"" | name.name*} and \lstinline{"" | name}.
+\lstinline{"" | name.name$^*$} and \lstinline{"" | name}.
 
@@
 \lstinline{""} is involved in a sequence lens, this lens is considered useful.
-Next, the sequence lens between \lstinline{name.name*} and
+Next, the sequence lens between \lstinline{name.name$^*$} and
 \lstinline{name} would be popped: while that lens is not bijective: it is still

### 2018-07-11 papers/SymmetricLenses/main.tex
Comments and polishing

@@
 
-A {\em lens} is a program that can be run both ``front to back'' and ``back
+{\em Lenses} are programs that can be run both ``front to back'' and ``back
 to front,'' allowing data to be transformed in two directions.  Since their
@@
 evaluate
-its performance on 48 benchmark examples drawn from FlashFill, Augeus,
-and an online collection of electronic file formats. We find this implementation is
-able to synthesize all the lenses of the benchmark suite in under 30 seconds.
+its performance on 48 benchmark examples drawn from FlashFill~\bcp{cite}, Augeus~\bcp{cite},
+and an online collection of electronic file formats. Our
+implementation can synthesize all these lenses in under
+30 seconds. 
 \end{abstract}
@@
 \noindent
-In a big-data world, similar data is often stored in a variety of
-formats. For instance, electronic calendars come in iCalendar,
-vCalendar and h-event formats~\cite{calendar-formats}; US taxation data can be
-transmitted via Tax XML (used by the US government for e-filing) or TXF (used by
-TurboTax)~\cite{tax-formats}; and sewing machines can read embroidery files
-transmitted as PES formats, DST formats or several
-others~\cite{embroidery-formats}.
-
-One way to interconvert between these varied data formats is to use a
-\emph{lens}~\cite{Focal2005}---a bi-directional
-program that can transform data represented in a format $S$ to a format $T$,
-and vice versa.  There are many different kinds of lenses, but well-defined lens
-languages typically provide guarantees that help ensure data is
-not lost or corrupted as it is transformed back and forth between different representations.
-
-While lenses facilitate writing principled data
-transformations, \textit{synthesizing} lenses 
-can make the task even easier.  Lens languages are a ripe target for
-synthesis because lens languages, like many domain-specific languages,
+In the big-data world, similar information is often stored or transmitted in
+different formats. For instance, electronic calendars come in iCalendar,
+vCalendar, and h-event formats~\cite{calendar-formats}; US taxation data can
+be transmitted via Tax XML (used by the US government for e-filing) or TXF
+(used by TurboTax)~\cite{tax-formats}; and sewing machines can use
+embroidery files in formats with names like PES, DST, and several
+others~\cite{embroidery-formats}.  \bcp{It would be nice to be able to give
+  a quick example of a truly symmetric case---do any of these examples work
+  for that?}
+
+One convenient way to interconvert between these varied data formats is to
+use a \emph{lens}~\cite{Focal2005}---a bidirectional program that can
+transform data represented in a format $S$ to a format $T$, and vice versa.
+offering ``round tripping'' guarantees that help ensure data is not lost or
+corrupted as it is transformed back and forth between different
+representations.
+
+While domain-specific languages for writing lenses can facilitate building
+principled data transformations, \textit{synthesizing} lenses 
+can make the task even easier.  Indeed, lens languages offer excellent
+targets for synthesis, since their specialized, sub-Turing-complete syntax
+and expressive type systems drastically
 reduce the space of possible programs, making search-based synthesis easier. 
-%This feature can be a negative, if the desired transformation
-%falls outside of the scope provided by the language, but can also be positive, as complex
-%properties (like invertability) can be enforced automatically.
-%Another advantage of reducing the
-%set of programs is that doing so can facilitate program synthesis, which is essentially
-%a search through the space of possible computations that can be expressed as a lens.
-Recent research~\cite{optician, maina+:quotient-synthesis} has exploited this
-observation by demonstrating that it is possible to synthesize relatively large,
-complex lenses from specifications. In particular, given types that describe
-source and target formats and optional examples of corresponding data, the
-bijective synthesis algorithm presented in these 
-papers will find bijective lenses to convert between data stored
-% ~\cite{optician}
+%
+Recent work~\cite{optician} has exploited this
+observation, demonstrating that it is feasible to synthesize quite complex
+examples, albeit only for a simple form of lenses---so-called bijective
+ones. Specifically, given a pair of types describing the intended
+source and target formats plus a small number of examples of corresponding
+data, the bijective synthesis algorithm developed in these two
+papers will find a bijective lens that converts between data stored
 in different formats,
@@
 search procedures (Figure~\ref{fig:high-level-algorithm}), following
-earlier work~\cite{maina+:quotient-synthesis}.
+earlier work~\cite{maina+:quotient-synthesis}\bcp{Is this the right citation?}.
 The first algorithm, \Expand, uses rewriting rules to iteratively
@@
 
-In summary, our main contributions are:
+In summary, our contributions are:\bcp{The section references do not proceed
+  in order, so this doesn't work very well as an outline of the paper.
+  Should we try to reorder the contributions?  Write a separate outline?
+  Live with it as-is?}
 \begin{itemize}
-\item We define simple symmetric lenses, which have
+\item We define {\em simple symmetric lenses}, with the
   the useful property that they do not maintain any hidden state
   (\S\ref{sec:overview}).  
-  We give a language for defining such lenses and describe how to
+  We present a collection of combinators for defining these lenses and
+  describe how to 
   calculate their likelihoods (\S\ref{sec:ssl}). 
@@
   lenses can be characterized semantically as the subset of full
-  symmetric lenses for which a simple ``forgetfulness'' property holds
-  (\S\ref{sec:relationship}).
+  symmetric lenses for which a simple ``forgetfulness'' property holds.
 \item We show how to synthesize simple symmetric lenses (including all
@@
   regular expressions to guide the search process (\S\ref{sec:synthesis}). 
-  To support that effort, we show that the star semi-ring equivalences on regular
+  To support that effort, we show that the star semi-ring equivalences on
+  regular 
   expressions remain valid when applied to stochastic regular
@@
   (\S\ref{sec:sre}).
-\item We extend Boomerang with simple symmetric lenses,
-  showing that such lenses can integrate smoothly with 
-  other lens extensions such as quotient lenses and matching
-  lenses.  We evaluate our
+\item We extend the Boomerang implementation with simple symmetric lenses,
+  demonstrating that they integrate smoothly with 
+  other lens extensions such as quotient lenses~\bcp{citation} and matching
+  lenses~\bcp{citation}.  We evaluate our
   implementation on a set of 48 lens synthesis benchmarks drawn from
@@
 \end{itemize}
-We finish the paper with a discussion of related work (\S\ref{sec:related}) and
-our concluding thoughts (\S\ref{sec:conc}).
+We close with a discussion of related work (\S\ref{sec:related}) and
+concluding thoughts (\S\ref{sec:conc}).
 
@@
 
-\subsection{Synthesizing Symmetric Lenses}
+\subsection{Synthesizing Simple Symmetric Lenses}
 
@@
 two communicating synthesizers following earlier
-work~\cite{maina+:quotient-synthesis}. The first synthesizer, \Expand, navigates
+work~\cite{maina+:quotient-synthesis}\bcp{Some of this text seems to be
+  identical to text in the intro.  That will annoy careful readers.}. The
+first synthesizer, \Expand, navigates 
 the space of semantically equivalent regular expressions by applying rewrite
@@
 over real numbers between 0 and 1, exclusive.
-
 $$\Regex{},\RegexAlt{} \GEq{}  \String  \; | \; \emptyset \; | \; \PRegexStar{\Regex}{p} \; | \; \RegexConcat{\Regex_1}{\Regex_2} \; | \; \PRegexOr{\Regex_1}{\Regex_2}{p}$$
@@
   \end{tabular}
-  \caption{Stochastic Regular Expression Star-Semiring Equivalence}
+  \caption{Stochastic Regular Expression Star-Semiring Equivalence\bcp{Can
+      we make the skips between the separate sections of the definition a
+      bit larger, and increase the margin on the left?  It looks (even)
+      denser than it is, at the moment.}}
   \label{fig:ssr-sre}
@@
 \item replaces $best$ with $\LC$, if $\LC$ has lower cost according
-to our information theoretic metric, and
+to our information-theoretic metric, and
 \item adds the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)
@@
 corresponding stochastic DNF regular expressions, written
-$\ToDNFRegexOf{\Regex}$, is defined below.
+$\ToDNFRegexOf{\Regex}$, is defined below.\bcp{Could put this in two columns
+to save space.}
 \[
@@
     \Regex}{\String}$ and $\LanguageOf{\Regex} =
-  \LanguageOf{\ToDNFRegexOf{\Regex}}$
+  \LanguageOf{\ToDNFRegexOf{\Regex}}$.
 \end{theorem}
@@
 \label{subsec:effect-of-heuristics}
-We evaluate the usefulness of (1) our information theoretic metric, and (2) our
+We evaluate the usefulness of (1) our information-theoretic metric, and (2) our
 termination heuristic.  To this end, we run our program in several different modes:\\
 \begin{tabulary}{\linewidth}{rL}
-  \AnyOpt{}: & We synthesize lenses without use of the information
-               theoretic preference metric.  Once \GreedySynth finds a lens, the
+  \AnyOpt{}: & We synthesize lenses without use of the information-theoretic
+  preference metric.  Once \GreedySynth finds a lens, the 
   synthesis procedure returns that lens.\\
   \FLOpt{}: &  The first time \GreedySynth finds a lens, the synthesis procedure
-              returns that lens. However, \GreedySynth uses the information
-              theoretic preference metric to choose a lens.\\
-  \CCOpt{}: & We replace our information theoretic cost metric with one where
+              returns that lens. However, \GreedySynth uses the
+              information-theoretic preference metric to choose a lens.\\ 
+  \CCOpt{}: & We replace our information-theoretic cost metric with one where
               the cost of the lens is the number of disonnects plus the number
@@
 
-We experimented with the \CCOpt{} mode to determine whether the complexity of the information theoretic measure was really needed.
+We experimented with the \CCOpt{} mode to determine whether the complexity of the information-theoretic measure was really needed.
 Related work on string transformations has often used simpler measures such as ``avoid constants'' which align with, but are simpler than our measures.
@@
   Let $\Lens$ be an asymmetric lens. $\Lens$ is also a simple symmetric lens,
-  where
+  where:\bcp{This could be set in two columns.}
   \begin{center}
@@
 wider class preserves more information during translation.  More broadly, we
-hope our information theoretic analysis provides a principle that can be
-used to understand the heuristic choices made in related work.
+hope our information-theoretic analysis provides a basis for
+understanding the heuristic choices made in related work.
 
@@
 regular expressions and information theory to guide the search for program transformations.
-As mentioned earlier, we believe such information theoretic techniques may have broad utility
+As mentioned earlier, we believe such information-theoretic techniques may have broad utility
 in helping us understand how to formulate a search for a data transformation function.

### 2018-07-11 papers/SymmetricLenses/main.tex
add it

--- 2018-07-11 papers/SymmetricLenses/main.tex
@@
 %$\PCF{Expand}$, (4) $\PCF{\PQ.Push}$, and (5) $\PCF{\Continue}$. First,
-This pair of SREs is used to initialize a priority queue ($pq$).
-Next, $\PCF{SynthSymLens}$
-enters a loop that searches for likely lenses. The loop terminates when the
-algorithm believes it is unlikely to find a better lens than the best one it has
-found so far (a termination condition defined by $\PCF{\Continue}$). Within each
-iteration of the loop, it:
+This pair of SREs is used to initialize a priority queue ($pq$). The priority of
+a SRE pair is the number of rewrites needed to derive the pair from the
+originals. Next, $\PCF{SynthSymLens}$ enters a loop that searches for likely
+lenses. The loop terminates when the algorithm believes it is unlikely to find a
+better lens than the best one it has found so far (a termination condition
+defined by $\PCF{\Continue}$). Within each iteration of the loop, it:
 \begin{itemize}
@@
 \item adds the SREs derived from rewriting $S$ and $T$ ($\Expand(S,T)$)
-  to the priority queue.  The priority of a SRE pair is the
-  number of rewrites needed to derive the pair from the originals.
+  to the priority queue.
 \end{itemize}
@@
 further depth in the following subsections.
+
+\emph{Stochastic DNF regular expressions} are critical to the success of our
+algorithm. Stochastic DNF regular expressions (\SDNFREabbrev{}s) are a
+pseudo-normal form for stochastic regular expressions. Of all the star-semiring
+equations required to relate two \SDNFREabbrevs, only those relating to Kleene
+star (\EG, \emph{Unrollstar$_L$} and \emph{Unrollstar$_R$}) are needed to relate
+two \SDNFREabbrevs; other axioms, like associativity, commutativity or
+distributivity, are unneeded. This important property makes it possible to
+construct an efficient algorithm ($\GreedySynth$), directed by the syntax of two
+\SDNFREabbrevs, to synthesize any lenses in the same ``class'' (\IE, any lenses
+that have the same type when the Kleene star axioms are unused). This property,
+and the resulting efficiency of $\GreedySynth$ is key to the effectiveness of
+our core algorithms.
 
@@
 
-\subsection{Search Through Types}
-The first synthesizer is one that searches through regular expression pairs.
-This synthesizer comprises \Expand, \PQ, and \Continue. It proposes regular
-expressions that are semantically, but not syntactically, equivalent to the
-originals by performing the following rewrites:
+\subsection{Searching for $(\Regex,\RegexAlt)$ Candidate Classes}
+The first phase of the synthesis algorithm looks for pairs of SREs
+$(\Regex,\RegexAlt)$ to drive the \GreedySynth algorithm.  These pairs are
+generated using the star unrolling axioms:
 \begin{center}
@@
 \end{center}
-As well as the congruence rules that allow these rewrites to be applied on
+as well as the congruence rules that allow these rewrites to be applied on
 subexpressions.
-
 The priority queue yields stochastic regular expressions generated using fewer
@@
 distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
-current best lens has cost $c$:
+current best lens has cost $c$, then:
 \begin{gather*}
@@
 
-When $\Call{Continue}{pq}$ returns $\False$, \Expand stops proposing regular
-expression pairs, and instead returns to the user the best lens found thus far.
-Note this means that if the algorithm finds a bijective lens, which has zero cost, it
-will immediately return.
+When $\Call{Continue}{pq}$ returns $\False$, the algorithm stops proposing
+regular expression pairs, and instead returns to the user the best lens found
+thus far. If the algorithm finds a bijective lens, which
+has zero cost, it will immediately return.
 %
 \subsection{Stochastic DNF Regular Expressions}
-Stochastic DNF regular expressions (\SDNFREabbrev) are a pseudo-normal form
-for stochastic regular expressions.  Of all the star-semiring equations
-required to relate two \SDNFREabbrevs, only those relating to Kleene star
-(\EG, \emph{Unrollstar$_L$} and \emph{Unrollstar$_R$}) are needed to relate
-two \SDNFREabbrevs; other axioms, like associativity, commutativity
-or distributivity, are unneeded.  This important property makes it possible to
-construct an efficient algorithm ($\GreedySynth$), directed
-by the syntax of two \SDNFREabbrevs, to synthesize any lenses in the
-same ``class'' (\IE, any lenses that have the same type when the
-Kleene star axioms are unused).  This property, and the resulting
-efficiency of $\GreedySynth$ is key to the effectiveness of our
-core algorithms.
-
 Just as SREs are REs with probability annotations,
@@
 the function $\LEntropyOf{\DNFRegex \Given \SDNFLens, \DNFRegexAlt}$ to
-calculate bounds of the expected amount of information required to recover a
+calculate bounds on the expected amount of information required to recover a
 string in $\DNFRegex$ from a string in $\DNFRegexAlt$, synchronized by
@@
 and \AtomSynth rlies on \GreedySynth. The structure of the algorithms mirrors
-the structure of symmetric DNF lenses and \SDNFREabbrevs. We describe this
-synthesis algorithm using the running example of searching for a lens between
-\lstinline{"" | name.name$^*$} and \lstinline{"" | name}.
+the structure of symmetric DNF lenses and \SDNFREabbrevs.
 
@@
 \GreedySynth loops until there are no more sequence lenses in the priority
-queue. Within this loop, a sequence lense is popped from the queue and, if it is
+queue. Within this loop, a sequence lens is popped from the queue and, if it is
 ``useful,'' it is included in the final DNF lens. The lens is considered to
@@
 
-In our example, \GreedySynth might first pop the sequence pair \lstinline{""} and
-\lstinline{""}, because there is a bijective sequence lens between them. As neither
-\lstinline{""} is involved in a sequence lens, this lens is considered useful.
-Next, the sequence lens between \lstinline{name.name$^*$} and
-\lstinline{name} would be popped: while that lens is not bijective: it is still
-better than the alternatives. As all sequences are now involved in sequence lenses,
-and there are no examples to make other lenses useful, no further sequence
-lenses would be added to the lens builder.
+As an example, consider searching for a lens between \lstinline{"" | name.name$^*$} and \lstinline{"" | name}. \GreedySynth might first pop the
+sequence pair \lstinline{""} and \lstinline{""}, because there is a bijective
+sequence lens between them. As neither \lstinline{""} is involved in a sequence
+lens, this lens is considered useful. Next, the sequence lens between
+\lstinline{name.name$^*$} and \lstinline{name} would be popped: while that lens
+is not bijective: it is still better than the alternatives. As all sequences are
+now involved in sequence lenses, and there are no examples to make other lenses
+useful, no further sequence lenses would be added to the lens builder.
 
@@
 
+\ifappendices
 \appendix
-
-\ifappendices
 
@@
 
-\fi
 \section{Forgetful Symmetric Lenses}
@@
 \end{proof}
+
+\fi
 \end{document}
@@
 %    conclude what?}
-%\end{enumerate}
+%  \end{enumerate}
 

### 2018-07-11 papers/SymmetricLenses/main.tex
clarification on first sentence of eval

@@
 extension to Boomerang. We added symmetric lens combinators to Boomerang, and
-implemented all of Boomerang's asymmetric lens combinators using a combination of
-the symmetric lens combinators, as well as other symmetric lens combinators for
-the asymmetric theory extensions present in Boomerang, like matching lenses and
-quotient lenses. We also integrated our synthesis engine into Boomerang. It allows users to write synthesis tasks
+implemented all of Boomerang's asymmetric lens combinators 
+using a combination of the symmetric lens combinators presented in this
+paper and symmetric versions of the asymmetric theory extensions
+already present in Boomerang, like matching lenses and quotient lenses.
+We also integrated our synthesis engine into Boomerang. It allows users to write synthesis tasks
 alongside lens combinators, use those synthesis results in the manual writing of

### 2018-07-11 papers/SymmetricLenses/main.tex
local changes

--- 2018-07-11 papers/SymmetricLenses/main.tex
@@
 different formats. For instance, electronic calendars come in iCalendar,
-vCalendar, and h-event formats~\cite{calendar-formats}; US taxation data can
-be transmitted via Tax XML (used by the US government for e-filing) or TXF
-(used by TurboTax)~\cite{tax-formats}; and sewing machines can use
-embroidery files in formats with names like PES, DST, and several
-others~\cite{embroidery-formats}.  \bcp{It would be nice to be able to give
-  a quick example of a truly symmetric case---do any of these examples work
-  for that?}
+vCalendar, and h-event formats~\cite{calendar-formats}; US taxation data can be
+transmitted via Tax XML (used by the US government for e-filing) or TXF (used by
+TurboTax)~\cite{tax-formats}; and sewing machines can use embroidery files in
+formats with names like PES, DST, and several others~\cite{embroidery-formats}.
+\bcp{It would be nice to be able to give a quick example of a truly symmetric
+  case---do any of these examples work for that?}\afm{h-event and vCalendar are
+  truly symmetric. Even the fields related to location alone are truly
+  symmetric (though it is subtle to see this).}
 
@@
 other advanced lenses in Boomerang, including quotient
-lenses and matching lenses.
+lenses~\cite{quotientlenses} and matching lenses~\cite{matchinglenses}.
 
@@
 search procedures (Figure~\ref{fig:high-level-algorithm}), following
-earlier work~\cite{maina+:quotient-synthesis}\bcp{Is this the right citation?}.
+earlier work~\cite{optician}.
 The first algorithm, \Expand, uses rewriting rules to iteratively
@@
   Should we try to reorder the contributions?  Write a separate outline?
-  Live with it as-is?}
+  Live with it as-is?}\afm{Personally, I don't see it as a problem.  It's nice
+  when the contributions mimic the paper, I am not jarred when they don't.}
 \begin{itemize}
@@
   demonstrating that they integrate smoothly with 
-  other lens extensions such as quotient lenses~\bcp{citation} and matching
-  lenses~\bcp{citation}.  We evaluate our
+  other lens extensions such as quotient lenses~\cite{quotientlenses} and matching
+  lenses~\cite{matchinglenses}.  We evaluate our
   implementation on a set of 48 lens synthesis benchmarks drawn from
@@
 Now that we have a way to calculate likelihood, we need a way to search for
-likely lenses. Previous work on synthesis~\cite{augustsson-2004,osera+:pldi15,feser-pldi-2015,frankle+:popl16} has shown that we can use
-types to dramatically restrict the search space and improve the effectiveness of
-synthesis. In our setting, types are pairs of stochastic regular expressions,
-which come with a rich algebraic structure that makes each regular expression
-semantically equal to infinitely many others. To tame that search space and
-leverage the benefits of type-directed synthesis, we split our algorithm into
-two communicating synthesizers following earlier
-work~\cite{maina+:quotient-synthesis}\bcp{Some of this text seems to be
-  identical to text in the intro.  That will annoy careful readers.}. The
-first synthesizer, \Expand, navigates 
-the space of semantically equivalent regular expressions by applying rewrite
-rules that preserve both the semantics and the probability distributions of
-those stochastic regular expressions. Our algorithm ranks pairs of stochastic
+likely lenses. Previous work on
+synthesis~\cite{augustsson-2004,osera+:pldi15,feser-pldi-2015,frankle+:popl16}
+has shown that we can use types to dramatically restrict the search space and
+improve the effectiveness of synthesis. In our setting, types are pairs of
+stochastic regular expressions where each regular expression is semantically
+equal to infinitely many others. We follow existing work on bidirectional
+program synthesis, and split our algorithm into two communicating
+synthesizers~\cite{maina+:quotient-synthesis}. The first synthesizer, \Expand,
+navigates the space of semantically equivalent regular expressions by applying
+rewrite rules that preserve both the semantics and the probability distributions
+of those stochastic regular expressions. Our algorithm ranks pairs of stochastic
 regular expressions by the number of rewrite rule applications required to
@@
 \begin{figure}
-  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
+  \setlength\extrarowheight{3pt}
+  \begin{tabular}{@{\hspace*{5em}}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
     \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
@@
   \end{tabular}
-  \caption{Stochastic Regular Expression Star-Semiring Equivalence\bcp{Can
-      we make the skips between the separate sections of the definition a
-      bit larger, and increase the margin on the left?  It looks (even)
-      denser than it is, at the moment.}}
+  \caption{Stochastic Regular Expression Star-Semiring Equivalence}
   \label{fig:ssr-sre}
@@
 recover a regular expression from a stochastic regular expression by deleting
-probability annotations. In a slight abuse of notation\saz{Why is this an
-  abuse?}\afm{We also use \BRegex{},\BRegexAlt{} as nonterminals}, if $\Regex$
-is a stochastic regular expression, $\BRegex$ is the regular expression obtained
-by removing probability annotations. The typing judgment $\Lens \OfType \BRegex
+probability annotations. In a slight abuse of notation, if $\Regex$ is a
+stochastic regular expression, $\BRegex$ is the regular expression obtained by
+removing probability annotations. The typing judgment $\Lens \OfType \BRegex
 \Leftrightarrow \BRegexAlt$ means that $\Lens$ is a well typed lens between
@@
 hash-consing stochastic DNF regular expressions.  This optimization enables
-fast equality checks and saves memory.\dpw{we could probably cut this paragraph.}
+fast equality checks and saves memory.
 %memoizing its results.  However, our stochastic DNF regular expressions are
@@
 searching for more likely lenses.  Our default $t$ is $0$.
-
-As discussed in the following section,
-while our default termination condition quickly finds the correct lens for
-39 of our 48 benchmarks, our termination condition is not
-one-size-fits all. Complex lenses, or lenses that project a
-large amount of data, can be synthesized more easily and quickly by
-adjusting $t$. \dpw{should this paragraph be cut and its content left for the
-next section?}
 % end synthesis
@@
 %\paragraph*{Issues with Classical Symmetric Lenses}
-\iffalse Using symmetric lenses for synchronization forces us to maintain not
-just two files that we aim to sycnrhonize, but three. \bcp{I don't find it
-  unintuitive, and I think we should be careful not to oversell the difficulty
-  of storing the complements. It requires more engineering, but it's not rocket
-  science.}This is unintuitive, and creates many problems in practice. For
-example, consider the situation where I am synchronizing a crontab and job
-manager file. I am synchronizing across computers, so I must either maintain a
-copy of the complement on each computer, and keep them synchronized, or figure
-out a computer to hold the complement, and make sure all edits go through that
-complement. Furthermore, if I want to change the synchronization to a different
-computer, I don't need to merely make sure that my files are moved to the
-computer, I also have to make sure that I correctly move the complement, and
-hook up that complement to be used for that synchronization task. This is a huge
-divergence from other synchronization utilities\bcp{No. For example, Unison
-  maintains a database on each replica recording, for each file, what its hash
-  was after the last time the two replicas were synchronized.}, where all that
-needs to be done is maintain the files. Using those for synchronization, edits
-can be sent over the wire, and the appropriate function can stitch together
-those edits, no additional state needed. \fi While classical symmetric lenses
+%\iffalse Using symmetric lenses for synchronization forces us to maintain not
+%just two files that we aim to sycnrhonize, but three. \bcp{I don't find it
+%  unintuitive, and I think we should be careful not to oversell the difficulty
+%  of storing the complements. It requires more engineering, but it's not rocket
+%  science.}This is unintuitive, and creates many problems in practice. For
+%example, consider the situation where I am synchronizing a crontab and job
+%manager file. I am synchronizing across computers, so I must either maintain a
+%copy of the complement on each computer, and keep them synchronized, or figure
+%out a computer to hold the complement, and make sure all edits go through that
+%complement. Furthermore, if I want to change the synchronization to a different
+%computer, I don't need to merely make sure that my files are moved to the
+%computer, I also have to make sure that I correctly move the complement, and
+%hook up that complement to be used for that synchronization task. This is a huge
+%divergence from other synchronization utilities\bcp{No. For example, Unison
+%  maintains a database on each replica recording, for each file, what its hash
+%  was after the last time the two replicas were synchronized.}, where all that
+%needs to be done is maintain the files. Using those for synchronization, edits
+%can be sent over the wire, and the appropriate function can stitch together
+%those edits, no additional state needed. \fi
+
+While classical symmetric lenses
 are technically more expressive than simple symmetric lenses, it is unclear

### 2018-07-11 papers/SymmetricLenses/transform-data.py
add graphs and transformer

@@ -78,7 +78,7 @@ def generate_examples_required_graph(input_csv):
     ax.set_xticklabels(["\\textbf{Any}","\\textbf{FL}","\\textbf{DC}","\\textbf{t=0}","\\textbf{t=25}","\\textbf{t=-25}"])
 
     fig = plt.figure(3,tight_layout=True)
-    ax.step([-.5,5.5],[45.1,45.1],label="Benchmark Count",linestyle=":",
+    ax.step([-.5,5.5],[48.1,48.1],label="Benchmark Count",linestyle=":",
             linewidth=1, dashes=(1,1))
 
     plt.tick_params(
@@ -363,7 +363,7 @@ def generate_time_vs_tasks_graph(input_csv):
     normal_size = 2
     full_size = 3
 
-    ax.step([0,60],[45.1,45.1],label="Benchmark Count",linestyle=":",
+    ax.step([0,60],[48.1,48.1],label="Benchmark Count",linestyle=":",
             linewidth=1, dashes=(1,1))
     create_step_plot("SS","\\textbf{SS}",'-',normal_size)
     create_step_plot("SSNC","\\textbf{SSNC}",':',normal_size)

### 2018-07-12 papers/SymmetricLenses/main.tex
minor fixes

@@
 We describe an implementation of simple symmetric lenses and our synthesis
-procedure as extensions to the Boomerang language~\cite{Boomerang}. We
-evaluate
-its performance on 48 benchmark examples drawn from FlashFill~\bcp{cite}, Augeus~\bcp{cite},
-and an online collection of electronic file formats. Our
-implementation can synthesize all these lenses in under
-30 seconds. 
+procedure as extensions to the Boomerang language~\cite{Boomerang}. We evaluate
+its performance on 48 benchmark examples drawn from Flash Fill~\cite{flashfill},
+Augeus~\cite{augeas2}, the bidirectional programming literature, and electronic
+file format synchronization tasks. Our implementation can synthesize all these
+lenses in under 30 seconds.
 \end{abstract}
@@
 synchronization tasks taken from a variety of other locations\footnote{The
-  majority of these format synchronization tasks were derived from the
+  majority of these format synchronization tasks were taken from the
   bidirectional programming literature.}.
@@
 corresponding stochastic DNF regular expressions, written
-$\ToDNFRegexOf{\Regex}$, is defined below.\bcp{Could put this in two columns
-to save space.}
+$\ToDNFRegexOf{\Regex}$, is defined below.
 \[
-  \begin{array}{rcl}
-    \ToDNFRegexOf{\String} & = & \DNFOf{(\SequenceOf{\String},1)}\\
-    \ToDNFRegexOf{\emptyset} & = & \DNFOf{}\\
+  \begin{array}{rcl@{\hspace*{3em}}rcl}
+    \ToDNFRegexOf{\String} 
+    & = 
+    & \DNFOf{(\SequenceOf{\String},1)}
+
+    & \ToDNFRegexOf{(\RegexConcat{\Regex_1}{\Regex_2})} 
+    & = 
+    & \ToDNFRegexOf{\Regex_1} \ConcatDNF \ToDNFRegexOf{\Regex_2} \\
+
+    \ToDNFRegexOf{\emptyset} 
+    & = 
+    & \DNFOf{}
+
+    & \ToDNFRegexOf{(\PRegexOr{\Regex_1}{\Regex_2}{\Probability})} 
+    & = 
+    & \ToDNFRegexOf{\Regex_1} \OrDNF_{\Probability} \ToDNFRegexOf{\Regex_2} \\
+
     \ToDNFRegexOf{(\PRegexStar{\Regex}{\Probability})} & = & \AtomToDNFOf{\PRegexStar{(\ToDNFRegexOf{\Regex})}{\Probability}}\\
-    \ToDNFRegexOf{(\RegexConcat{\Regex_1}{\Regex_2})} & = & \ToDNFRegexOf{\Regex_1} \ConcatDNF \ToDNFRegexOf{\Regex_2}\\
-    \ToDNFRegexOf{(\PRegexOr{\Regex_1}{\Regex_2}{\Probability})} & = & \ToDNFRegexOf{\Regex_1} \OrDNF_{\Probability} \ToDNFRegexOf{\Regex_2}\\
+
   \end{array}
@@
 We evaluate the usefulness of (1) our information-theoretic metric, and (2) our
-termination heuristic.  To this end, we run our program in several different modes:\\
+termination heuristic.  To this end, we run our program in several different
+modes:\\
 \begin{tabulary}{\linewidth}{rL}
@@
   Let $\Lens$ be an asymmetric lens. $\Lens$ is also a simple symmetric lens,
-  where:\bcp{This could be set in two columns.}
+  where:
   \begin{center}
-    \begin{tabular}{rcl}
-      $\Lens.\CreateROf{x}$ & $=$ & $\Lens.get \App x$\\
-      $\Lens.\CreateLOf{y}$ & $=$ & $\Lens.create \App y$\\
-      $\Lens.\PutROf{x}{y}$ & $=$ & $\Lens.get \App x$\\
+    \begin{tabular}{rcl@{\hspace*{3em}}rcl}
+      $\Lens.\CreateLOf{y}$ & $=$ & $\Lens.create \App y$
+      & $\Lens.\CreateROf{x}$ & $=$ & $\Lens.get \App x$\\
       $\Lens.\PutLOf{y}{x}$ & $=$ & $\Lens.put \App y \App x$
+      & $\Lens.\PutROf{x}{y}$ & $=$ & $\Lens.get \App x$
     \end{tabular}

### 2018-07-12 papers/SymmetricLenses/main.tex
change alg to 60 sec
add more info

@@
 use a \emph{lens}~\cite{Focal2005}---a bidirectional program that can
-transform data represented in a format $S$ to a format $T$, and vice versa.
+transform data represented in a format $S$ to a format $T$, and vice versa,
 offering ``round tripping'' guarantees that help ensure data is not lost or
@@
 %
-Recent work~\cite{optician} has exploited this
-observation, demonstrating that it is feasible to synthesize quite complex
-examples, albeit only for a simple form of lenses---so-called bijective
-ones. Specifically, given a pair of types describing the intended
-source and target formats plus a small number of examples of corresponding
-data, the bijective synthesis algorithm developed in these two
-papers will find a bijective lens that converts between data stored
-in different formats,
-such as linux configuration files from different systems that represent the same data in different ways.  
-Often, these algorithms can find an appropriate lens
-in less than a second without requiring any examples.
+Recent work~\cite{optician} has exploited this observation, demonstrating that
+it is feasible to synthesize quite complex examples, albeit only for a simple
+form of lenses---so-called bijective lenses. Specifically, given a pair of types
+describing the intended source and target formats plus a small number of
+examples of corresponding data, the bijective synthesis algorithm will find a
+bijective lens that converts between data stored the provided formats. Often,
+these algorithms can find an appropriate lens in less than a second without
+requiring any examples.
 %, with just types describing the source and target formats.
@@
 synchronize between two ad hoc file formats have this 
-characteristic, 13 benchmarks in total. Consequently, tools that search
-only for strict bijections are often not applicable in practice.
+characteristic, 13 benchmarks in total. Consequently, searching
+only for strict bijections is often not applicable in practice.
 
@@
 Initially, we hypothesized that extending the synthesis algorithms
-of Miltner \ETAL~\cite{optician} to operate over symmetric rather than
+of \citet{optician} to operate over symmetric rather than
 bijective lenses
@@
   Live with it as-is?}\afm{Personally, I don't see it as a problem.  It's nice
-  when the contributions mimic the paper, I am not jarred when they don't.}
+  when the contributions mirror the paper, I am not jarred when they don't.}
 \begin{itemize}
@@
 For each benchmark in our benchmark suite, we ran every mode on it with a
-timeout of 30 seconds, and averaged the result over 5 runs.
+timeout of 60 seconds, and averaged the result over 5 runs.
 Figure~\ref{fig:times} summarizes the results of these tests. We find that our
@@
 tools have been shown to synthesize a backwards transformation from ordinary code designed
-to implement the forwards transformation.  For instance, Hu and
-D'Antoni~\cite{program-inversion-symbolic-transducers} show how to invert transformations
+to implement the forwards transformation.  For instance, \citet{program-inversion-symbolic-transducers} show how to invert transformations
 using symbolic finite automata, and
-Voigtl\"{a}nder~\cite{bidirectionalization-for-free} demonstrates how to
+\citet{bidirectionalization-for-free} demonstrates how to
 construct reverse transformations from forwards transformations
@@
 natural language understanding, machine translation, information retrieval, data
-extraction and grammatical inference (see Pereira~\cite{pereira:info-theory},
+extraction and grammatical inference (see \citet{pereira:info-theory},
 for an introduction). Indeed, our work was inspired, in part, by the PADS format

### 2018-07-12 papers/SymmetricLenses/main.tex
merge main

@@
 to front,'' allowing data to be transformed in two directions.  Since their
-introduction by~\citet{Focal2005-long2}, lenses have been extensively
+introduction by~\citet{Focal2005-long}, lenses have been extensively
 studied and applied.  Recent work~\cite{optician} has also demonstrated how
@@
 bijective lenses and the more widely used ``asymmetric''
-lenses~\cite{Focal2005-long2}, as well as a rich subset of the full-blown
+lenses~\cite{Focal2005-long}, as well as a rich subset of the full-blown
 ``symmetric lenses'' proposed by \citet{symmetric-lenses}.  Simple symmetric

### 2018-07-12 papers/SymmetricLenses/main.tex
merge kathleens changes in

@@
 \label{sec:evaluation}
-We have implemented our symmetric DNF lenses and the synthesis of them as an
-extension to Boomerang. We added symmetric lens combinators to Boomerang, and
-implemented all of Boomerang's asymmetric lens combinators 
-using a combination of the symmetric lens combinators presented in this
-paper and symmetric versions of the asymmetric theory extensions
+We have implemented symmetric DNF lenses as an extension to
+Boomerang. Specifically, we added symmetric lens combinators to
+Boomerang, and implemented all of Boomerang's asymmetric lens
+combinators  using a combination of the symmetric lens combinators presented in this
+paper and symmetric versions of the asymmetric extensions
 already present in Boomerang, like matching lenses and quotient lenses.
-We also integrated our synthesis engine into Boomerang. It allows users to write synthesis tasks
-alongside lens combinators, use those synthesis results in the manual writing of
-new lenses, and use previously defined lenses in the synthesis of new lenses.
+We also integrated our synthesis engine into Boomerang. This
+integration allows users to write synthesis tasks 
+alongside lens combinators, incorporate synthesis results into 
+manually-written lenses, and reference previously defined lenses in
+the synthesis of new lenses.
 
@@
 \begin{enumerate}
-\item Is our synthesis procedure efficient enough to be used in everyday
+\item Does the synthesis algorithm find the correct lens?
+\item Is the synthesis procedure efficient enough to be used in everyday
   development?
-
 \item How much slower is our tool on bijective lens synthesis benchmarks than
@@
 
-\item Which heuristics are useful for reducing the ``specification
+\item Which heuristics are useful
+for reducing the ``specification 
   overhead'' for users? In particular, which heuristics are useful in
-  synthesizing the desired lens, instead of merely a lens that satisfies the
+  synthesizing the desired lens instead of merely a lens that satisfies the
   specification?
@@
 \begin{enumerate}
-\item We adapted 8 data cleaning benchmarks from Flash Fill, though, of course, our tool produces bidirectional transformations rather than one-way transformers like Flash Fill.
+\item We adapted 8 data cleaning benchmarks from Flash Fill. Note
+that our tool produces bidirectional transformations rather
+than one-way transformers like Flash Fill. 
 \item We adapted 29 benchmarks from Augeas. These benchmarks turn the
-  ad hoc data present in Linux configuration files into a structured, dictionary
+  ad hoc data present in Linux configuration files into a structured dictionary
   representation.
-\item We used 11 benchmarks from synchronization tasks between a variety of
-  formats derived from real-world examples and the bidirectional programming
+\item We created 11 additional benchmarks 
+  derived from real-world examples and/or the bidirectional programming
   literature. These tasks range from synchronizing REST and JSON web resource
@@
 \end{enumerate}
-
-A realistic benchmarks suite is critical in answering the first research
-question. To determine whether our algorithm is fast enough to be used in
-everyday development, we need to evaluate whether we can solve real-world
-problems efficiently. All three of our benchmark sources test real life
-problems: the data cleaning tasks presented in Flash Fill are either derived
-from online help forums or taken from the Excel product team; Augeas is a
-utility that transforms Linux configuration files into a manageable dictionary
-form; the many of the remaining benchmarks are from synchronizing real-life data
-formats, like BibTeX and EndNote.
+Each benchmark comprises a source and target regular expression, a
+collection of examples, and an optional termination condition
+parameter. 
+%%A realistic benchmarks suite is critical in answering the first research
+%%question. To determine whether our algorithm is fast enough to be used in
+%%everyday development, we need to evaluate whether we can solve real-world
+%%problems efficiently.
+
+All three sources for our benchmarks provide realistic problems
+suitable for answering the research questions:
+Flash Fill  data cleaning tasks are either derived from online help
+forums or taken from the Excel product team;  
+Augeas is a utility that transforms Linux configuration files into a manageable dictionary
+form; and
+many of the remaining benchmarks arise from synchronizing real-world
+data formats such as BibTeX and EndNote.
+
+\subsection{Lens correctness}
+To determine whether the synthesized lenses were the desired ones,
+we ran the symmetric synthesis algorithm on the 48 benchmarks and then
+manually inspected the generated lenses.  In all 48 cases, the
+synthesized benchmark was correct.
 
@@
 
-To begin, we measure the run time of our algorithm in 2 modes:
+We measured the run time of our algorithm in two modes:
 
@@
     \SSNCOpt{} is the symmetric synthesis algorithm without using a library of
-    existing lenses. Our symmetric synthesis algorithm is able to complete all
-    benchmarks in our benchmark suite in under 30 seconds, and is able to
-    complete a majority of the benchmarks without using compositional
+    existing lenses. The symmetric synthesis algorithm is able to complete all
+    benchmarks in under 30 seconds elapsed total time. In addition, it is able to
+    complete a majority of the benchmarks in that timeframe without
+  using compositional 
     synthesis.}
@@
 
-For each benchmark in our benchmark suite, we ran every mode on it with a
+For each benchmark in our benchmark suite, we ran it in both modes with a
 timeout of 60 seconds, and averaged the result over 5 runs.
@@
 Section~\ref{subsec:optimizations} that users may tune the termination metric.
-Success on this set of benchmarks requires we tune that termination codition
+Success on this set of benchmarks requires we tune that termination condition
 (e.g. for some benchmarks, allowing the system flexibility to spend more time
@@
 To compare to the existing bijective synthesis algorithm, we run our symmetric
-synthesis algorithm on the original Optician benchmark suite. 4 of these
-benchmarks had to be slightly altered to work with our new system. We altered
-these benchmarks by either providing additional examples or tuning termination parameters.
-Without these alterations, the symmetric synthesis yielded a lens that fit
+synthesis algorithm on the original Optician benchmarks, comprised of
+39 bijective synthesis tasks. We had to slightly alter four of these
+benchmarks, either by providing additional examples or by tuning termination parameters.
+Without these alterations, symmetric synthesis yielded a lens that fit
 the specification but that was not the desired lens.
 
-To perform this comparison, we synthesized lenses in 2 modes:
+To perform this comparison, we synthesized lenses in two modes:
 
@@
 
-For each benchmark in our benchmark suite, we ran every mode on it with a
-timeout of 60 seconds, and averaged the result over 5 runs.
+For each benchmark, we ran it in both modes with a
+timeout of 60 seconds and averaged the result over 5 runs.
 Figure~\ref{fig:times_bijective} summarizes the results of these tests. On
@@
 We evaluate the usefulness of (1) our information-theoretic metric, and (2) our
-termination heuristic.  To this end, we run our program in several different
-modes:\\
+termination heuristic.  To this end, we run our program in several different modes:\\[1ex]
 \begin{tabulary}{\linewidth}{rL}
-  \AnyOpt{}: & We synthesize lenses without use of the information-theoretic
-  preference metric.  Once \GreedySynth finds a lens, the 
-  synthesis procedure returns that lens.\\
-  \FLOpt{}: &  The first time \GreedySynth finds a lens, the synthesis procedure
-              returns that lens. However, \GreedySynth uses the
-              information-theoretic preference metric to choose a lens.\\ 
+  \AnyOpt{}: & We ignore the information-theoretic preference metric entirely,
+  returning the first lens \GreedySynth finds without using this information. \\ 
+  \FLOpt{}: &  \GreedySynth uses the information-theoretic preference
+  metric to find a lens, but only the original pair of stochastic
+  regular expressions are considered as end-points.  \\
   \CCOpt{}: & We replace our information-theoretic cost metric with one where
@@
 
-We experimented with the \CCOpt{} mode to determine whether the complexity of the information-theoretic measure was really needed.
-Related work on string transformations has often used simpler measures such as ``avoid constants'' which align with, but are simpler than our measures.
-The \CCOpt{} mode is an example of such a simple measure---it operates by counting disconnects, which put a complete stop to information transfer, like constants, and merges, which eliminate the information in a union.
-Figure~\ref{fig:metric} summarizes the result of this experiment.
-We find that
-our metrics are very important for finding the correct lens. Our
-information-theoretic metric is critical for finding the correct lens. Indeed,
-only 10 of our benchmarks succeeded when running in \CCOpt{} mode.
-
-Our termination condition is also quite important. When running in \FLOpt{},
-only 5 lenses are found. The first cluster in which a lens is found is
-rarely the correct cluster. However, our termination condition is not perfect.
+We experimented with the \CCOpt{} mode to determine whether the
+complexity of the information-theoretic measure was really
+needed. Related work on string transformations has often used simpler
+measures such as ``avoid constants'' which align with, but are simpler
+than our measures. The \CCOpt{} mode is an example of such a simple
+measure---it operates by counting disconnects, which put a complete
+stop to information transfer, and merges, which
+eliminate the information in a union.
+
+Figure~\ref{fig:metric}
+summarizes the result of these experiments. The data reveal that 
+the information-theoretic metric is critical for finding the correct
+lens: Only 10 of the benchmarks succeeded when running in \CCOpt{} mode.
+The termination condition is also quite important. When running
+in \FLOpt{} mode, the algorithm only discovers 5 lenses, which shows
+that the first cluster that contains a satisfying lens is rarely the
+correct cluster.  However, our termination condition is not perfect.
 Sometimes users must tell the algorithm to ``keep going'' or ``stop early.''
 Without termination parameters, our algorithm finds the correct lens for 36 of
-our 48 benchmarks, 9 of them need help in the form of a termination parameter
+our 48 benchmarks.  Nine of them need help in the form of a termination parameter
 for finding the correct lens. That being said, our default parameter finds a
 good middle-ground. Both \TFTOpt{} and \NTFTOpt{} synthesized the correct lens
-less often than \ZTOpt.  \TFTOpt{} necessitates too good a lens, so the search
-procedure keeps spinning, and \NTFTOpt{} terminates early, not finding the
-correct lens.
+less often than \ZTOpt.  \TFTOpt{} demands too good a lens, so the search
+procedure keeps spinning, while \NTFTOpt{} terminates early, not
+always trying hard enough to find the correct lens.
 % end evaluation

### 2018-07-12 papers/SymmetricLenses/main.tex
add fl change

@@
   \FLOpt{}: &  \GreedySynth uses the information-theoretic preference
-  metric to find a lens, but only the original pair of stochastic
-  regular expressions are considered as end-points.  \\
+              metric to find a lens, and we return immediately upon finding a
+              satisfying lens, instead of returning based on \Continue.  \\
   \CCOpt{}: & We replace our information-theoretic cost metric with one where

### 2018-07-12 papers/SymmetricLenses/main.tex
fixed typo
mentioned tuning and interaction in the abstract

@@
 Augeus~\cite{augeas2}, the bidirectional programming literature, and electronic
-file format synchronization tasks. Our implementation can synthesize all these
+file format synchronization tasks. After modest tuning, which is encouraged by
+the interactive nature of the tool, our implementation can synthesize all these
 lenses in under 30 seconds.
@@
 
-In this evaluation, we aim to answer three primary questions:
+In this evaluation, we aim to answer four primary questions:
 \begin{enumerate}

### 2018-07-12 papers/SymmetricLenses/main.tex
make change

--- 2018-07-12 papers/SymmetricLenses/main.tex
@@
 One convenient way to interconvert between these varied data formats is to
-use a \emph{lens}~\cite{Focal2005}---a bidirectional program that can
+use a \emph{lens}~\cite{Focal2005-long}---a bidirectional program that can
 transform data represented in a format $S$ to a format $T$, and vice versa,
-offering ``round tripping'' guarantees that help ensure data is not lost or
+providing ``round tripping'' guarantees that help ensure data is not lost or
 corrupted as it is transformed back and forth between different
@@
 
-However, there is a price for narrowing the synthesis domain so
-dramatically: Many real data formats are almost, but not quite,
-bijections. More often two related data formats have some data in
-common, but also contain some differences. For instance, two related
-tables might share four of five columns, but each has a different
-fifth column. Alternately, one ad hoc configuration file might include
-some file-specific meta data, such as a date, a time or a reference
-number, whereas the other does not. Indeed of the 48 benchmark problems
-described in Section \ref{sec:evaluation}, all of the benchmarks
-taken from Flash Fill~\cite{flashfill} and many of the benchmarks that
-synchronize between two ad hoc file formats have this 
-characteristic, 13 benchmarks in total. Consequently, searching
-only for strict bijections is often not applicable in practice.
+However, there is a price for narrowing the synthesis domain so dramatically:
+many synchronization tasks cannot by synthesized! Many real data formats are
+almost, but not quite, bijections. More often two related data formats have some
+data in common, but also contain some differences. For instance, two related
+tables might share four of five columns, but each has a different fifth column.
+Alternately, one ad hoc configuration file might include some file-specific meta
+data, such as a date, a time or a reference number, whereas the other does not.
+Indeed of the 48 benchmark problems described in Section \ref{sec:evaluation},
+all of the benchmarks taken from Flash Fill~\cite{flashfill} and many of the
+benchmarks that synchronize between two ad hoc file formats have this
+characteristic, 13 benchmarks in total. Consequently, searching only for strict
+bijections is often not applicable in practice.
 
@@
 \item  We show that simple symmetric lenses
-  fall between asymmetric lenses~\cite{Focal2005-long2} and full
+  fall between asymmetric lenses~\cite{Focal2005-long} and full
   symmetric lenses~\cite{symmetric-lenses} in expressiveness
@@
 \paragraph*{Asymmetric Lenses}
-Formally, an {\em asymmetric lens} $ \ell : S \Leftrightarrow V$ is a triple of functions $\ell.\get : S \longrightarrow V$, $\ell.\pput : V \longrightarrow S \longrightarrow S$ and $\ell.\create : V \longrightarrow S$ satisfying the following laws \cite{Focal2005-long2}:
+Formally, an {\em asymmetric lens} $ \ell : S \Leftrightarrow V$ is a triple of functions $\ell.\get : S \longrightarrow V$, $\ell.\pput : V \longrightarrow S \longrightarrow S$ and $\ell.\create : V \longrightarrow S$ satisfying the following laws \cite{Focal2005-long}:
 \begin{align*}

### 2018-07-12 papers/SymmetricLenses/main.tex
sentence in the intro

@@
   majority of these format synchronization tasks were taken from the
-  bidirectional programming literature.}.
+  bidirectional programming literature.}.  Our tool is intended to be responsive
+enough for interactive use by the programmer, who can control the search process
+via examples and a ``tuning'' parameter that governs how much effort is spent
+trying to improve on existing solutions.
 

### 2018-07-12 papers/SymmetricLenses/main.tex
updated the contributions

@@
 \item We extend the Boomerang implementation with simple symmetric lenses,
-  demonstrating that they integrate smoothly with 
-  other lens extensions such as quotient lenses~\cite{quotientlenses} and matching
-  lenses~\cite{matchinglenses}.  We evaluate our
-  implementation on a set of 48 lens synthesis benchmarks drawn from
-  data cleaning, view maintenance, 
-  and file synchronization tasks.  We demonstrate that we can
-  synthesize simple symmetric lenses for all of the benchmarks in
+  demonstrating that they integrate smoothly with other lens extensions such as
+  quotient lenses~\cite{quotientlenses} and matching
+  lenses~\cite{matchinglenses}.  We evaluate our implementation and the effects
+  of various optimizations and the tuning parameter on a set of 48 lens
+  synthesis benchmarks drawn from data cleaning, view maintenance, and file
+  synchronization tasks.  We demonstrate that, with only modest interactive
+  tuning, the system can synthesize simple symmetric lenses for all of the benchmarks in
   under 30 seconds (\S\ref{sec:evaluation}).

### 2018-07-12 papers/SymmetricLenses/main.tex
pass through evaluation

@@
 To resolve this tension, our algorithm uses heuristics to judge whether to
-return the current best satisfying lens to the user or to pass the
-next set of unrollings to \GreedySynth.
-The heuristics favor stopping if the current best satisfying
-lens has relatively high likelihood, indicating the lens is very promising.
-The heuristics also favor stopping if \Expand
-has delievered to \GreedySynth all pairs of stochastic regular
-expressions at a given rank and there is a large number of pairs at the next
-rank because searching through all such pairs will take a long time.
-With this approach, the algorithm can quickly respond with a satisfying
-lens with relatively high likelihood.  If users are unhappy with the result, they can refine the
-search by supplying additional examples, which serve both to rule out
-the previously proposed lens(es) and to reduce the size of the search
-space by cutting down on the number of satisfying lenses.
+return the current best satisfying lens to the user or to pass the next set of
+unrollings to \GreedySynth.  The heuristics favor stopping if the current best
+satisfying lens has relatively high likelihood, indicating the lens is very
+promising.  The heuristics also favor stopping if \Expand has delivered to
+\GreedySynth all pairs of stochastic regular expressions at a given rank and
+there is a large number of pairs at the next rank because searching through all
+such pairs will take a long time.  With this approach, the algorithm can quickly
+respond with a satisfying lens with relatively high likelihood.  If users are
+unhappy with the result, they can either refine the search by supplying
+additional examples, which serve both to rule out the previously proposed
+lens(es) and to reduce the size of the search space by cutting down on the
+number of satisfying lenses, or supply a tuning parameter (discussed in 
+\S\ref{subsec:optimizations}) that adjusts how much effort is put into the search.
 
@@
 
-\subsection{Optimizations}
+\subsection{Optimizations and Termination Tuning}
 \label{subsec:optimizations}
@@
 implementation also provides for an additional specification parameter that
-customizes the termination condition for the synthesis procedure.
+customizes the termination condition for the synthesis procedure.  The
+optimizations make the system performant enough for interactive use, and the
+tuning parameter gives the user a ``knob'' by which the algorithm can be
+directed to try harder to find a good lens.
 
@@
 combinators  using a combination of the symmetric lens combinators presented in this
-paper and symmetric versions of the asymmetric extensions
-already present in Boomerang, like matching lenses and quotient lenses.
+paper and symmetric versions of the asymmetric extensions (like matching lenses and quotient lenses)
+already present in Boomerang.
 We also integrated our synthesis engine into Boomerang. This
@@
 \begin{enumerate}
-\item Does the synthesis algorithm find the correct lens?
+\item Can the algorithm (with suitable examples and termination
+  tuning) find the correct lens?
 \item Is the synthesis procedure efficient enough to be used in everyday
@@
   the existing Optician system customized for bijective lenses?
-
-\item Which heuristics are useful
-for reducing the ``specification 
-  overhead'' for users? In particular, which heuristics are useful in
-  synthesizing the desired lens instead of merely a lens that satisfies the
-  specification?
+\item How much termination tuning is necessary and how does the heuristic affect the
+  results?
 \end{enumerate}
@@
 \end{enumerate}
-Each benchmark comprises a source and target regular expression, a
-collection of examples, and an optional termination condition
-parameter. 
+
+All three sources for our benchmarks provide realistic problems suitable for
+answering the research questions: Flash Fill data cleaning tasks are either
+derived from online help forums or taken from the Excel product team; Augeas is
+a utility that transforms Linux configuration files into a manageable dictionary
+form; and many of the remaining benchmarks arise from synchronizing real-world
+data formats such as BibTeX and EndNote.  Each task consists of a source and
+target regular expression.\saz{Do we want to say anything about the sizes of these?}
+
+\subsection{Synthesizing Correct Lenses}
+
+Because our tool is intended to be interactive---the user typically starts with
+zero or one examples, runs the tools, determines whether the output is the
+desired one and, if not, either adds an example or adjusts the termination
+metric and then tries again---it is hard to cleanly evaluate the net user
+experience.  The primary concern is whether it is usable to generate
+synthesize correct lenses.
+
+To determine whether the system can synthesize the desired lenses, we ran the
+symmetric synthesis algorithm interactively on all 48 tasks,
+working with the system to create sufficient examples and adjust the termination
+tuning.  In all cases, the desired lens was obtained, which we verified by
+manually inspecting the generated lens programs.
+%
+The majority of the tasks required only a single example and none required
+more than three examples in order to synthesize the desired lens.\footnote{In
+  one benchmark, we supplied a fourth example that was later discovered to 
+  be unnecessary.}
+
+Adjusting the tuning parameter was needed in only 9 of the 48 tasks. In practice, we
+found that customizing the termination metric to be quite easy: if manual
+inspection of the lens showed there were too few bijections, we would increase
+the termination parameter. If synthesis took too long, we would decrease it.
+Section~\ref{subsec:effect-of-heuristics}, studies the effect of fixing the
+termination metric.
+
 %%A realistic benchmarks suite is critical in answering the first research
@@
 
-All three sources for our benchmarks provide realistic problems
-suitable for answering the research questions:
-Flash Fill  data cleaning tasks are either derived from online help
-forums or taken from the Excel product team;  
-Augeas is a utility that transforms Linux configuration files into a manageable dictionary
-form; and
-many of the remaining benchmarks arise from synchronizing real-world
-data formats such as BibTeX and EndNote.
-
-\subsection{Lens correctness}
-To determine whether the synthesized lenses were the desired ones,
-we ran the symmetric synthesis algorithm on the 48 benchmarks and then
-manually inspected the generated lenses.  In all 48 cases, the
-synthesized benchmark was correct.
-
-\subsection{Usefulness in Everyday Development}
-
-We measured the run time of our algorithm in two modes:
+\subsection{Effectiveness of Compositional Synthesis}
+
+Having determined appropriate examples and timing parameters for the 48
+benchmarks, we evaluate the performance of the system by 
+measuring the running time of our algorithm in two modes:
 
@@
   \SSOpt{}: & The symmetric synthesis algorithm with all optimizations enabled.\\
-  \SSNCOpt{}: & The symmetric synthesis algorithm, but compositional synthesis not enabled.\\
+  \SSNCOpt{}: & The symmetric synthesis algorithm, with no compositional synthesis enabled.\\
 \end{tabulary}\\
@@
 compositional synthesis (SS mode) allows our system to scale to arbitrarily
-large and complex formats. However, SSNC mode, which uses no such annotations
-and synthesizes a complete lens all at once provides a useful experimental
-stress test for the system.
+large and complex formats; measuring it provides an estimate of the
+responsiveness of the system when used as intended.\saz{Anders check this!} However, SSNC mode, which
+uses no such annotations and synthesizes a complete lens all at once provides a
+useful experimental stress test for the system.
 
@@
     existing lenses. The symmetric synthesis algorithm is able to complete all
-    benchmarks in under 30 seconds elapsed total time. In addition, it is able to
-    complete a majority of the benchmarks in that timeframe without
-  using compositional 
-    synthesis.}
+    benchmarks in under 30 seconds elapsed total time. Without compositional
+    synthesis it is able to complete 26. Each benchmark specification includes
+    source and target regular expressions, between one and three sufficient
+    examples, and, if necessary, the timing parameter determined to be
+    sufficient for correctness during a prior interactive session.}
   \label{fig:times}
@@
 
-For each benchmark in our benchmark suite, we ran it in both modes with a
-timeout of 60 seconds, and averaged the result over 5 runs.
-Figure~\ref{fig:times} summarizes the results of these tests. We find that our
-algorithm is able to synthesize all of the benchmarks in our benchmark suite in
-under 30 seconds. Without compositional synthesis, the synthesis algorithm is
-able to solve 26 problem instances. Recall from
-Section~\ref{subsec:optimizations} that users may tune the termination metric.
-Success on this set of benchmarks requires we tune that termination condition
-(e.g. for some benchmarks, allowing the system flexibility to spend more time
-considering more of the search space). In
-Section~\ref{subsec:effect-of-heuristics}, we will consider the effect of fixing
-the termination metric. In practice, we found customizing the termination metric
-to be quite easy: if manual inspection of the lens showed there were too few
-bijections, we would increase the termination parameter. If synthesis took too
-long, we would decrease it.
+For each benchmark in the suite and each mode, we ran the system with a timeout
+of 60 seconds, averaging the result over 5 runs.  Figure~\ref{fig:times}
+summarizes the results of these tests. We find that our algorithm is able to
+synthesize all of the benchmarks in under 30 seconds. Without compositional
+synthesis, the synthesis algorithm is able to solve only 26 problem
+instances.
+
+% Section~\ref{subsec:optimizations} that users may tune the termination metric.
+% Success on this set of benchmarks requires we tune that termination condition
+% (e.g. for some benchmarks, allowing the system flexibility to spend more time
+% considering more of the search space). In
 
@@
 synthesis algorithm on the original Optician benchmarks, comprised of
-39 bijective synthesis tasks. We had to slightly alter four of these
+39 bijective synthesis tasks.\footnote{We had to slightly alter four of these
 benchmarks, either by providing additional examples or by tuning termination parameters.
 Without these alterations, symmetric synthesis yielded a lens that fit
-the specification but that was not the desired lens.
+the specification but that was undesired.}
 
@@
 Sometimes users must tell the algorithm to ``keep going'' or ``stop early.''
-Without termination parameters, our algorithm finds the correct lens for 36 of
-our 48 benchmarks.  Nine of them need help in the form of a termination parameter
-for finding the correct lens. That being said, our default parameter finds a
+Without termination parameters, our algorithm finds the correct lens for 39 of
+our 48 benchmarks; nine required tuning 
+to find the correct lens. Nevertheless, our default parameter sits in a
 good middle-ground. Both \TFTOpt{} and \NTFTOpt{} synthesized the correct lens

### 2018-07-12 papers/SymmetricLenses/main.tex
push some minor changes

@@
-\newif\ifdraft\drafttrue  % set true to show comments
+\newif\ifdraft\draftfalse  % set true to show comments
 %\newif\ifdraft\draftfalse  % set true to show comments
@@
   the existing Optician system customized for bijective lenses?
-\item How much termination tuning is necessary and how does the heuristic affect the
-  results?
+\item How effective is the information-theoretic search heuristic and how does
+  termination tuning affect the results?
 \end{enumerate}
@@
 working with the system to create sufficient examples and adjust the termination
-tuning.  In all cases, the desired lens was obtained, which we verified by
+parameter.  In all cases, the desired lens was obtained, which we verified by
 manually inspecting the generated lens programs.
@@
 
-Adjusting the tuning parameter was needed in only 9 of the 48 tasks. In practice, we
+Adjusting the termination parameter was needed in only 9 of the 48 tasks. In practice, we
 found that customizing the termination metric to be quite easy: if manual

### 2018-07-12 papers/SymmetricLenses/main.tex
ah

@@
 
-Having determined appropriate examples and timing parameters for the 48
-benchmarks, we evaluate the performance of the system by 
-measuring the running time of our algorithm in two modes:
+Having determined appropriate examples and termination parameters for the 48
+benchmarks, we evaluate the performance of the system by measuring the running
+time of our algorithm in two modes:
 

### 2018-07-12 papers/SymmetricLenses/main.tex
space tweaks

--- 2018-07-12 papers/SymmetricLenses/main.tex
@@
-\newif\ifdraft\drafttrue  % set true to show comments
-%\newif\ifdraft\draftfalse  % set true to show comments
+%\newif\ifdraft\drafttrue  % set true to show comments
+\newif\ifdraft\draftfalse  % set true to show comments
 \newif\ifanon\anonfalse    % set true to suppress names, etc.
@@
   \includegraphics[width=.5\textwidth]{high-level-algorithm.pdf}
+  \vspace{-2ex}
   \caption{Schematic diagram for the simple symmetric lens synthesis algorithm.
@@
 \lstinline{name.name$^*$} and \lstinline{name} would be popped: while that lens
-is not bijective: it is still better than the alternatives. As all sequences are
+is not bijective it is still better than the alternatives. As all sequences are
 now involved in sequence lenses, and there are no examples to make other lenses
-useful, no further sequence lenses would be added to the lens builder.
+useful, no more sequence lenses would be added to the lens builder.
 
@@
   the existing Optician system customized for bijective lenses?
-\item How much termination tuning is necessary and how does the heuristic affect the
-  results?
+\item How effective is the information-theoretic search heuristic and how does termination tuning affect the results?
 \end{enumerate}
@@
   \includegraphics{generated-graphs/times}
+  \vspace{-4ex}
   \caption{Number of benchmarks that can be solved by a given algorithm in a
@@
 
-\begin{figure}
+\begin{figure}[t]
   \includegraphics{generated-graphs/times_bijective}
+  \vspace{-4ex}
   \caption{Number of benchmarks that can be solved by a given algorithm in a
@@
   \FLOpt{}: &  \GreedySynth uses the information-theoretic preference
-              metric to find a lens, and we return immediately upon finding a
+              metric to search, but returns immediately upon finding a
               satisfying lens, instead of returning based on \Continue.  \\
@@
               of merges.\\
+\end{tabulary}
+
+\begin{tabulary}{\linewidth}{rL}
   \ZTOpt{}: & \ZTOpt{} runs the algorithm with the termination parameter
@@
 \end{tabulary}
-
-\begin{figure}
-  \includegraphics{generated-graphs/metrics_importance}
-  \caption{Number of benchmarks that synthesize the correct lens by a given
-    algorithm. \AnyOpt{} provides no notion of cost, and merely returns the
-    first lens it finds that satisfies the specification. \FLOpt{} provides a
-    notion of cost to \GreedySynth, but once a satisfying lens is greedily
-    found, that lens is returned. \CCOpt{} synthesizes lenses, where the cost of
-    a lens is the number of disconnects plus the number of merges. \ZTOpt{} runs
-    the algorithm with the termination parameter universally set to 0 (default).
-    \TFTOpt{} runs the algorithm with the termination parameter universally set
-    to 25. \NTFTOpt{} runs the algorithm with the termination parameter
-    universally set to -25.}
-  \label{fig:metric}
-\end{figure}
 
@@
 \label{sec:related}
+
+\begin{figure}[t]
+  \includegraphics{generated-graphs/metrics_importance}
+  \vspace{-4ex}
+  \caption{Number of benchmarks that synthesize the correct lens by a given
+    algorithm. \AnyOpt{} provides no notion of cost, and merely returns the
+    first lens it finds that satisfies the specification. \FLOpt{} provides a
+    notion of cost to \GreedySynth, but once a satisfying lens is greedily
+    found, that lens is returned. \CCOpt{} synthesizes lenses, where the cost of
+    a lens is the number of disconnects plus the number of merges. \ZTOpt{} runs
+    the algorithm with the termination parameter universally set to 0 (default).
+    \TFTOpt{} runs the algorithm with the termination parameter universally set
+    to 25. \NTFTOpt{} runs the algorithm with the termination parameter
+    universally set to -25.}
+  \label{fig:metric}
+\end{figure}
 

### 2018-07-12 papers/SymmetricLenses/transform-data.py
change graph heigh

@@ -334,7 +334,7 @@ def generate_time_vs_tasks_graph_vs_bijective(input_csv):
     plt.yticks(np.arange(0, 40.1, 10))
 
     fig = plt.figure(2,tight_layout=True)
-    fig.set_figheight(2)
+    fig.set_figheight(1.75)
     fig.set_figwidth(4)
 
     fig.savefig(generated_graphs_base + "times_bijective.eps", bbox_inches='tight')
@@ -379,7 +379,7 @@ def generate_time_vs_tasks_graph(input_csv):
     plt.yticks(np.arange(0, 50.1, 10))
 
     fig = plt.figure(1,tight_layout=True)
-    fig.set_figheight(2)
+    fig.set_figheight(1.75)
     fig.set_figwidth(4)
 
     fig.savefig(generated_graphs_base + "times.eps", bbox_inches='tight')

### 2018-07-12 papers/SymmetricLenses/main.tex
more space tweaks

@@
 
-\GreedySeqSynth then loops until there are no more atom lenses in the priority
-queue. Within the loop, a popped atom lens is considered 
+\GreedySeqSynth loops until there are no more atom lenses in the priority
+queue. In the loop, a popped atom lens is considered 
 ``useful'' if adding it to the sequence will lower the
@@
 must not be disconnected. Each atom can be part of only one lens at a time, so
-the algorithm must sometimes a previously chosen atom lens in order to connect an
-atom that must not be disconnected.
-After all atoms have been popped, the sequence lens is validated and
-returned. The algorithm succeeds when all atoms that must not be disconnected
-are involved in an atom lens, but \PCF{SLensBuilder.ToDNFLens}
+the algorithm must sometimes remove a previously chosen atom lens in order to connect 
+one that must not be disconnected.
+% After all atoms have been popped, the sequence lens is validated and
+% returned.
+The algorithm succeeds when all atoms that must not be disconnected
+are involved in an atom lens; \PCF{SLensBuilder.ToDNFLens}
 returns \None otherwise.
@@
 \paragraph*{Expansion Inference} These additional rewrites make the search
-through possible regular expressions much harder. Our algorithm automatically
-will identify when certain closed regular expressions can \emph{only} be
+through possible regular expressions harder. Our algorithm 
+identifies when certain closed regular expressions can \emph{only} be
 involved in a disconnect lens.  Such regular expressions will automatically be

### 2018-07-12 papers/SymmetricLenses/main.tex
fits in 25 pages

@@
 \begin{figure}
-  \setlength\extrarowheight{3pt}
   \begin{tabular}{@{\hspace*{5em}}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
@@
 
-All three sources for our benchmarks provide realistic problems suitable for
+All three sources provide realistic problems suitable for
 answering the research questions: Flash Fill data cleaning tasks are either
@@
 
-To determine whether the system can synthesize the desired lenses, we ran the
-symmetric synthesis algorithm interactively on all 48 tasks,
-working with the system to create sufficient examples and adjust the termination
-parameter.  In all cases, the desired lens was obtained, which we verified by
-manually inspecting the generated lens programs.
+To determine whether the system can synthesize the desired lenses, we ran it
+interactively on all 48 tasks, working with the system to create sufficient
+examples and adjust the termination parameter.  In all cases, the desired lens
+was obtained, which we verified by manually inspecting the generated lens
+programs.
 %

### 2018-07-12 papers/SymmetricLenses/main.tex
add more text and cleaning

@@
 \newif\iffull\fullfalse   % set true for long version
-\newif\ifappendices\appendicestrue
+\newif\ifappendices\appendicesfalse
 
@@
 found that customizing the termination metric to be quite easy: if manual
-inspection of the lens showed there were too few bijections, we would increase
+inspection of the lens showed there were too few \IdentityLens{}s, and too many
+\Disconnect{}s or merges, we would increase
 the termination parameter. If synthesis took too long, we would decrease it.
-Section~\ref{subsec:effect-of-heuristics}, studies the effect of fixing the
+Section~\ref{subsec:effect-of-heuristics} studies the effect of fixing the
 termination metric.
@@
 
-Recall that compositional synthesis exploits user annotations to break a
-benchmark into a series of smaller synthesis steps. In our experience,
-compositional synthesis (SS mode) allows our system to scale to arbitrarily
-large and complex formats; measuring it provides an estimate of the
-responsiveness of the system when used as intended.\saz{Anders check this!} However, SSNC mode, which
-uses no such annotations and synthesizes a complete lens all at once provides a
-useful experimental stress test for the system.
+Recall that compositional synthesis allows users to break a benchmark into a
+series of smaller synthesis tasks, whose solutions are utilized in more complex
+synthesis procedures. Compositional synthesis (\SSOpt{} mode)
+allows our system to scale to arbitrarily large and complex formats; measuring
+it shows of the responsiveness of the system when used as
+intended. \SSNCOpt{} mode, which synthesizes a complete lens all at once,
+provides a useful experimental stress test for the system.
 
@@
 synthesize all of the benchmarks in under 30 seconds. Without compositional
-synthesis, the synthesis algorithm is able to solve only 26 problem
-instances.
+synthesis, the synthesis algorithm is able to solve 26 problem instances.
 

### 2018-07-12 papers/SymmetricLenses/main.tex
Checkpoint -- finished polishing intro

@@
-%\newif\ifdraft\drafttrue  % set true to show comments
+% \newif\ifdraft\drafttrue  % set true to show comments
 \newif\ifdraft\draftfalse  % set true to show comments
@@
 
+% BCP: To make sure we don't leave comments on!
+\bcp{XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX }
+
 {\em Lenses} are programs that can be run both ``front to back'' and ``back
-to front,'' allowing data to be transformed in two directions.  Since their
+to front,'' allowing data and updates to be transformed in two directions.
+Since their 
 introduction by~\citet{Focal2005-long}, lenses have been extensively
@@
 
-We show how to extend this synthesis algorithm to a much richer class of
+We show how to extend this synthesis algorithm to a much wider class of
 lenses, which we call {\em simple symmetric lenses}, including both
@@
 \noindent
-In the big-data world, similar information is often stored or transmitted in
-different formats. For instance, electronic calendars come in iCalendar,
-vCalendar, and h-event formats~\cite{calendar-formats}; US taxation data can be
-transmitted via Tax XML (used by the US government for e-filing) or TXF (used by
-TurboTax)~\cite{tax-formats}; and sewing machines can use embroidery files in
-formats with names like PES, DST, and several others~\cite{embroidery-formats}.
-\bcp{It would be nice to be able to give a quick example of a truly symmetric
-  case---do any of these examples work for that?}\afm{h-event and vCalendar are
-  truly symmetric. Even the fields related to location alone are truly
-  symmetric (though it is subtle to see this).}
-
-One convenient way to interconvert between these varied data formats is to
-use a \emph{lens}~\cite{Focal2005-long}---a bidirectional program that can
-transform data represented in a format $S$ to a format $T$, and vice versa,
-providing ``round tripping'' guarantees that help ensure data is not lost or
-corrupted as it is transformed back and forth between different
-representations.
+In today's big-data world, similar information is sometimes
+stored in different places and in different formats. For instance, electronic
+calendars come in iCalendar, vCalendar, and h-event
+formats~\cite{calendar-formats}; US taxation data can be transmitted via Tax
+XML (used by the US government for e-filing) or TXF (used by
+TurboTax)~\cite{tax-formats}; and sewing machines can use embroidery files
+in formats with names like PES, DST, and several
+others~\cite{embroidery-formats}.
+%
+\bcp{I'm not satisfied with the embroidery example: It doesn't obviously
+  correspond to a bidirectional situation (the sewing machine does not need
+  to send back the embroidery file to anybody).  The tax example could be
+  criticized on the same grounds.}
+%
+\bcp{Also, it would be nice to be able to give a quick example of a truly
+  symmetric case---do any of these examples work for that?}\afm{h-event and
+  vCalendar are truly symmetric. Even the fields related to location alone
+  are truly symmetric (though it is subtle to see this).}
+
+One convenient way to maintain consistency between these varied data formats
+is to use a \emph{lens}~\cite{Focal2005-long}---a bidirectional program that
+can transform updates to data represented in a format $S$ to another format
+$T$ and vice versa, providing ``round tripping'' guarantees that help ensure
+information is not lost or corrupted as it is transformed back and forth
+between different representations.
 
@@
 can make the task even easier.  Indeed, lens languages offer excellent
-targets for synthesis, since their specialized, sub-Turing-complete syntax
+targets for synthesis: their specialized, sub-Turing-complete syntax
 and expressive type systems drastically
@@
 Recent work~\cite{optician} has exploited this observation, demonstrating that
-it is feasible to synthesize quite complex examples, albeit only for a simple
-form of lenses---so-called bijective lenses. Specifically, given a pair of types
-describing the intended source and target formats plus a small number of
+it is feasible to synthesize quite complex examples, albeit only for a
+restricted class of lenses---so-called bijective lenses. Specifically, given 
+source and target formats plus a small number of
 examples of corresponding data, the bijective synthesis algorithm will find a
-bijective lens that converts between data stored the provided formats. Often,
-these algorithms can find an appropriate lens in less than a second without
-requiring any examples.
+bijective lens matching the examples, if one exists. Often,
+it can find an appropriate lens in less than a second with
+no examples at all.
 %, with just types describing the source and target formats.
@@
 %   third person, even if that's a bit awkward, but otherwise citing it is OK.}
-One reason that bijective lens synthesis is so effective, even relative to successful
-synthesis projects in other domains, is that there are not very
-many bijections between related data formats (unless the
-same kind of data appears in the two formats many times). 
+One reason that bijective lens synthesis is so effective, even relative to
+successful synthesis tools in other domains, is that there are typically not
+very many bijections between related data formats. 
 Hence, if the synthesis algorithm can
-find any bijection at all, then it is very likely to find the one the
-user wants.
-
-However, there is a price for narrowing the synthesis domain so dramatically:
-many synchronization tasks cannot by synthesized! Many real data formats are
-almost, but not quite, bijections. More often two related data formats have some
-data in common, but also contain some differences. For instance, two related
-tables might share four of five columns, but each has a different fifth column.
-Alternately, one ad hoc configuration file might include some file-specific meta
-data, such as a date, a time or a reference number, whereas the other does not.
+find any bijection at all, it is likely to be the right one.
+
+However, many real-world data formats share just part of their information
+content with other representations.  For instance, two related
+database tables might share four out of five columns, but each might have a
+different fifth column. 
+Or, one ad-hoc system configuration file might include some format-specific meta
+data, such as a date or a reference number, whereas the same configuration
+file on another operating system does not. 
 Indeed of the 48 benchmark problems described in Section \ref{sec:evaluation},
@@
 benchmarks that synchronize between two ad hoc file formats have this
-characteristic, 13 benchmarks in total. Consequently, searching only for strict
-bijections is often not applicable in practice.
-
-A natural way to expand the scope of such tools is to consider a more
-expressive language of lenses.  For example, \emph{symmetric
-lenses}~\cite{symmetric-lenses} encompass a broad collection of bidirectional
-transformations that include bijections as well as more flexibile transformations that
-allow each side to retain information not present
-in the other.  If $\ell$ is a symmetric lens between formats $S$ and
-$T$, then $\ell$  can transform data from $S$ to $T$ or vice versa
-equally well.  A user would typically 
-run $\ell$ in the forwards direction (from $S$ to $T$) when a user updates
-a string $s \in S$ to propagate those updates to $T$.
-However, because data in format $S$ may contain some information not present in format $T$,
-the $S \rightarrow T$ transformation takes
-\emph{both} the edited string $s \in S$ and a string $t \in T$ as arguments.
-Inuitively, the lens extracts the data ``shared'' between the two
-formats from $s$
-and uses $t$ to supply any missing information.
+characteristic, 13 benchmarks in total. Consequently, synthesizing only strict
+bijections is often not useful in practice.
+
+Could we extend synthesis tools to more kinds of lenses?  In particular,
+what about \emph{symmetric lenses}~\cite{symmetric-lenses}, a class that
+includes bijective lenses, ``asymmetric'' lenses where the transformation
+from $S$ to $T$ can throw away information that must then be restored when
+returning from $T$ to $S$, and even more flexibile transformations that
+allow each side to retain information not present in the other?
+
+\bcp{We use $S$ and $T$ here but $X$ and $Y$ below.}
+
+% A symmetric lens between $S$ and $T$ represents a pair of functions, $\PutR$
+% and $\PutL$.  The $\PutR$ component has type
+% $S \times T \rightarrow T$---that is, it takes a ``new'' $S$ and an ``old''
+% $T$ and builds a ``new'' $T$ using its first argument for the shared
+% information and its second argument for information that is only available
+% on the $T$ side.  Conversely, the $\PutL$ component has type
+% $T \times S \rightarrow S$.
+
+% If $\ell$ is a symmetric lens between formats $S$ and
+% $T$, then $\ell$  can transform data from $S$ to $T$ or vice versa
+% equally well.  A user would typically 
+% run $\ell$ in the forwards direction (from $S$ to $T$) when a user updates
+% a string $s \in S$ to propagate those updates to $T$.
+% However, because data in format $S$ may contain some information not present in format $T$,
+% the $S \rightarrow T$ transformation takes
+% \emph{both} the edited string $s \in S$ and a string $t \in T$ as arguments.
+% Inuitively, the lens extracts the data ``shared'' between the two
+% formats from $s$
+% and uses $t$ to supply any missing information.
+%
 % required by $T$ that is missing
-%as it generates an output $t'$.  If $s$ is missing information, the transformation should
+%as it generates an output $t'$.  If $s$ is missing information, the
+%transformation should 
 %extract the missing components from $t$.
 
-Initially, we hypothesized that extending the synthesis algorithms
-of \citet{optician} to operate over symmetric rather than
-bijective lenses
-would be relatively straightforward:  Simply replace the bijective
-combinators with symmetric ones 
-and search using similar heuristics.  However, 
-this na\"ive approach runs into two difficulties.
-
-The first difficulty is pragmatic. Symmetric lenses usually operate
+One might hope that extending the algorithms from \citet{optician} to
+synthesize symmetric rather than bijective lenses would be relatively
+straightforward: Simply replace the bijective combinators with symmetric
+ones and search using similar heuristics.  However, this na\"ive approach
+encounters two difficulties.
+
+The first is pragmatic. ``Classical'' symmetric lenses as presented
+by ~\citet{symmetric-lenses} operate
 over three structures: the source $S$, the target $T$ and a complement
 $C$ that contains the information not
-present in either $S$ or $T$~\cite{symmetric-lenses}. Logistically, the complement must be stored
-somewhere and managed. More significantly, complements complicate synthesis
-specifications.
-Users can no longer just give examples of source and target
-pairings; instead they must also indicate how to manage the 
-``hidden state.'' To avoid such complexities, we define a new variant
-of symmetric lenses: \emph{simple symmetric lenses}. At the price
-of some expressiveness, these new lenses do not require complements.
+present in either $S$ or $T$. Logistically, these complements must be stored
+and managed somehow.  More fundamentially, complements complicate
+synthesis {\em specifications}---instead of just giving single examples of
+source and target 
+pairings, users would have to give longer ``interaction sequences'' to 
+show how a lens should behave. To avoid these complexities, we define a
+restricted variant 
+of symmetric lenses, which we call \emph{simple symmetric lenses}. 
 %only the usual source and target.
 %
-To characterize their expressive power,
-we prove that simple symmetric lenses include all (and only)
-those symmetric lenses that satisfy an intuitive property called
-\emph{forgetfulness}.
 %, which we define in Section~\ref{sec:relationship}.
@@
 recent instance.
-We justify this loss of expressiveness pragmatically. 
-We extend Boomerang~\cite{Boomerang} with simple symmetric lenses and
-successfully use them on a range of real-world applications, showing
-in the process that simple symmetric lenses integrate smoothly with 
-other advanced lenses in Boomerang, including quotient
-lenses~\cite{quotientlenses} and matching lenses~\cite{matchinglenses}.
-
-The second difficulty with synthesizing symmetric lenses is more
-fundamental: While the number of bijective lenses between two
-related formats is often tiny, the number of simple symmetric lenses is much larger. In
-practice, the difference in cardinality means that when a search algorithm
-returns the first bijective lens it finds, that lens has a good chance of being
-the desired lens. In contrast, and to our initial surprise, experiments showed
-that if a na\"ive search algorithm selects the first simple symmetric lens it
-finds, it is typically \emph{not} the desired lens.  The flexibility
-that such lenses provide in eliding information 
-generates an explosion of possible transformations. To disambiguate
-between the many possible lenses, we need new 
-principles for identifying the ``likely'' ones, and we need new
-algorithms guided by those principles to
-search the space of program transformations.
-
-To overcome this second difficulty, we turn to information theory 
-to characterize the likely and unlikely lenses. 
-We consider a ``likely'' lens to be one that propagates a lot of
+
+To characterize their expressive power formally, we show that simple
+symmetric lenses are exactly the symmetric lenses that satisfy an intuitive
+property called \emph{forgetfulness}.  To check that they are expressive
+enough in practice, we add simple symmetric lenses to the Boomerang
+language~\cite{Boomerang} and apply them to a range of real-world
+applications, showing in the process that simple symmetric lenses integrate
+smoothly with other advanced lens features provided by Boomerang, including
+quotient lenses~\cite{quotientlenses,maina+:quotient-synthesis} and matching
+lenses~\cite{matchinglenses}.
+
+However, we are still left with a second difficulty with synthesizing
+symmetric lenses: While the number of bijective lenses between two related
+formats is typically tiny, the number of simple symmetric lenses is
+typically huge.  Thus, if a na\"ive search algorithm selects the first
+simple symmetric lens it finds, this will generally \emph{not} be the one we
+want.  We need a new principle for identifying ``more likely'' lenses and a
+more sophisticated synthesis algorithm that uses this principles to search
+the space more intelligently.
+
+For these, we turn to information theory. 
+We consider a ``likely'' lens to be one that propagates ``a lot'' of
 information from source to 
-target and vice versa. Conversely, an ``unlikely'' lens requires a
+target and vice versa. Conversely, an unlikely lens requires a
 large amount of additional information to recover the target given a
-source and vice versa.  
+source.  
 This categorization follows from the intuition
 that users will typically want to synchronise formats with lots
-of shared information, and so lenses that preserve 
-more information are preferable.
-To implement this
-intuition, we turn to \emph{stochastic regular expressions}
-(SREs)~\cite{stoch-rnn,stoch-def}, which simultaneously define a set of strings \`a la
-regular expressions and a probability distribution over those
-strings.  Using this information, we can calculate the shared
-information content of two formats and our measure of
-likeliness.
+of shared information, and so lenses that propagate
+more information are preferable to ones that propagate less.
+%
+This intuition is formalized using \emph{stochastic regular expressions}
+(SREs)~\cite{stoch-rnn,stoch-def}, which simultaneously define a set of
+strings and a probability distribution over those strings.  Using this
+probability distribution, we can calculate the shared information content of
+two formats\bcp{It's not obvious at this point what the {\em shared}
+  information content has to do with anything.} and our measure of likelihood.
 
@@
     \GreedySynth. \GreedySynth finds a lens typed between the supplied SREs.
-    When the algorithm determines it has found a likely lens, it returns it.}
+    When the algorithm finds a likely lens, it returns it.\bcp{Later: We
+      could save significant space if we put the inputs and outputs on the
+      sides instead of top and bottom}}
   \label{fig:high-level-algorithm}
@@
 
-With simple symmetric lenses and our SRE-based likeliness measure in
+With simple symmetric lenses and our SRE-based likelihood measure in
 hand, we develop a new algorithm for synthesizing likely lenses.
@@
 descriptions of the source and target formats, measuring success using
-the likeliness measure.
-However, regular expressions enjoy a rich algebraic structure, meaning
-there are infinitely many regular expressions equivalent to the
-original pair, and the lens returned by a type-directed search depends
-upon the chosen representation. 
-To tame this complexity, we divide the synthesis algorithm into two communicating
-search procedures (Figure~\ref{fig:high-level-algorithm}), following
-earlier work~\cite{optician}.
-The first algorithm, \Expand, uses rewriting rules to iteratively
+the likelihood measure.
+The interesting complications arise from the need to deal with
+regular-expression equivalences.  There are infinitely many regular
+expressions equivalent to a given one, and the lens returned by a
+type-directed search will depend which of the infinitely many possible
+representations are chosen for its source and target formats.  Moreover,
+there may be {\em no} lens connecting the 
+source and target formats selected by the user: in general, the synthesis
+algorithm has to search to find equivalent ones for which a suitable lens
+exists.  
+To tame this complexity, we divide the synthesis algorithm into two
+communicating search procedures (Figure~\ref{fig:high-level-algorithm}),
+following~\citet{optician}. 
+The first, \Expand, uses rewriting rules to iteratively
 propose new pairs of stochastic regular expressions equivalent to the
 original pair.
-The second algorithm, \GreedySynth, takes a pair of such SREs as input
-and uses a greedy, type-directed algoithm to find a simple symmetric
+The second, \GreedySynth, takes a pair of SREs as input
+and uses a greedy, type-directed algorithm to find a simple symmetric
 lens between them, returning the lens and its likelihood score
-to \Expand.  The algorithm terminates when the likilihood score of the
-best lens found so far compared with 
+to \Expand.  The whole synthesis algorithm terminates when comparing the
+likilihood score of the best lens found so far with 
 the complexity of newly generated SREs 
@@
 
-We implemented the \GreedySynth and \Expand algorithms and evaluated their
-effectiveness at synthesizing lenses on a benchmark suite of 48 problems,
-including data cleaning tasks taken from Flash Fill~\cite{gulwani-popl-2014},
-view maintenance tasks taken from Augeas~\cite{Augeas}, and ad hoc format
-synchronization tasks taken from a variety of other locations\footnote{The
-  majority of these format synchronization tasks were taken from the
-  bidirectional programming literature.}.  Our tool is intended to be responsive
-enough for interactive use by the programmer, who can control the search process
-via examples and a ``tuning'' parameter that governs how much effort is spent
-trying to improve on existing solutions.
+We added this synthesis procedure to the Boomerang implementation and
+evaluated its effectiveness on a benchmark suite of 48 problems, including
+data-cleaning tasks taken from Flash Fill~\cite{gulwani-popl-2014},
+view-maintenance tasks taken from Augeas~\cite{augeas2}, and ad-hoc
+synchronization tasks from a variety of other sources (mostly from the
+bidirectional programming literature).  The ``programmer'' can control the
+search process via examples and a {\em tuning parameter} that governs how
+much effort is spent trying to improve on existing solutions.
 
@@
   Should we try to reorder the contributions?  Write a separate outline?
-  Live with it as-is?}\afm{Personally, I don't see it as a problem.  It's nice
-  when the contributions mirror the paper, I am not jarred when they don't.}
+  Live with it as-is?}\afm{Personally, I don't see it as a problem.  It's
+  nice when the contributions mirror the paper, I am not jarred when they
+  don't.}\bcp{But this way, the reader has no birds-eye view of the paper at
+  this point...}
 \begin{itemize}
-\item We define {\em simple symmetric lenses}, with the
-  the useful property that they do not maintain any hidden state
-  (\S\ref{sec:overview}).  
-  We present a collection of combinators for defining these lenses and
-  describe how to 
-  calculate their likelihoods (\S\ref{sec:ssl}). 
-\item  We show that simple symmetric lenses
-  fall between asymmetric lenses~\cite{Focal2005-long} and full
-  symmetric lenses~\cite{symmetric-lenses} in expressiveness
-  (\S\ref{sec:related}).  We further show that the class of simple
-  lenses can be characterized semantically as the subset of full
-  symmetric lenses for which a simple ``forgetfulness'' property holds.
-\item We show how to synthesize simple symmetric lenses (including all
-  asymmetric lenses as a special case), using a novel application of stochastic
-  regular expressions to guide the search process (\S\ref{sec:synthesis}). 
-  To support that effort, we show that the star semi-ring equivalences on
-  regular 
-  expressions remain valid when applied to stochastic regular
-  expressions, preserving the probability distributions as well as the languages
-  (\S\ref{sec:sre}).
+%
+\item We define {\em simple symmetric lenses}, a subclass of symmetric
+lenses that do not maintain any hidden state (\S\ref{sec:overview}).  We
+present a collection of combinators for building simple symmetric lenses and
+describe how to calculate their likelihoods (\S\ref{sec:ssl}).
+%
+\item We show that simple symmetric lenses fall between asymmetric
+lenses~\cite{Focal2005-long} and full symmetric
+lenses~\cite{symmetric-lenses} in expressiveness.  We further show that the
+class of simple lenses can be characterized semantically as the subset of
+full symmetric lenses for which a simple ``forgetfulness'' property holds
+(\S\ref{sec:related}).
+%
+\item We give an algorithm for synthesizing simple symmetric lenses, using a
+novel application of stochastic regular expressions to guide the search
+process (\S\ref{sec:synthesis})\bcp{``and prove it correct''???}.  Along the
+way, we show that {\em star semi-ring equivalences} on regular expressions
+remain valid when applied to stochastic regular expressions, preserving
+probability distributions as well as languages (\S\ref{sec:sre}).
+%
 \item We extend the Boomerang implementation with simple symmetric lenses,
-  demonstrating that they integrate smoothly with other lens extensions such as
-  quotient lenses~\cite{quotientlenses} and matching
-  lenses~\cite{matchinglenses}.  We evaluate our implementation and the effects
-  of various optimizations and the tuning parameter on a set of 48 lens
-  synthesis benchmarks drawn from data cleaning, view maintenance, and file
-  synchronization tasks.  We demonstrate that, with only modest interactive
-  tuning, the system can synthesize simple symmetric lenses for all of the benchmarks in
-  under 30 seconds (\S\ref{sec:evaluation}).
+demonstrating that they integrate smoothly with other lens extensions such
+as quotient lenses~\cite{quotientlenses} and matching
+lenses~\cite{matchinglenses}.  We evaluate our implementation and the
+effects of various optimizations and the tuning parameter on a set of 48
+lens synthesis benchmarks drawn from data cleaning, view maintenance, and
+file synchronization tasks.  With only modest interactive tuning, the system
+can synthesize simple symmetric lenses for all of the benchmarks in under 30
+seconds (\S\ref{sec:evaluation}).
 \end{itemize}
-We close with a discussion of related work (\S\ref{sec:related}) and
+We close with related work (\S\ref{sec:related}) and
 concluding thoughts (\S\ref{sec:conc}).
@@
 are technically more expressive than simple symmetric lenses, it is unclear
-whether this additional expressiveness is useful in practice. In our judgement,
+whether this additional expressiveness is useful in practice.\bcp{That's a
+  total cop-out! We have much more to say about this point, and we {\em
+    should} have even more to say than we do.} In our judgement,
 the complications they introduce into synthesis did not justify the potential

### 2018-07-12 papers/SymmetricLenses/main.tex
Submitted version

@@
 \newif\ifanon\anonfalse    % set true to suppress names, etc.
-\newif\iffull\fullfalse   % set true for long version
 \newif\ifappendices\appendicesfalse
@@
 %%strings is identical. Indeed, we guarantee these two properties by implementing
-%%a search procedure that obeys the star semi-ring axioms of regular expression
+%%a search procedure that obeys the star-semiring axioms of regular expression
 %%equivalence, and which we show does not alter the probability distributions of
@@
   nice when the contributions mirror the paper, I am not jarred when they
-  don't.}\bcp{But this way, the reader has no birds-eye view of the paper at
-  this point...}
+  don't.}\bcp{But then the reader has no birds-eye view of the paper at
+  this point.  I tried reorganizing, which I think also helps clarify the
+  contributions.} 
 \begin{itemize}
@@
 \item We define {\em simple symmetric lenses}, a subclass of symmetric
-lenses that do not maintain any hidden state (\S\ref{sec:overview}).  We
+lenses that do not maintain any hidden state (\S\ref{sec:overview}).  
+%
+We
 present a collection of combinators for building simple symmetric lenses and
 describe how to calculate their likelihoods (\S\ref{sec:ssl}).
-%
-\item We show that simple symmetric lenses fall between asymmetric
-lenses~\cite{Focal2005-long} and full symmetric
-lenses~\cite{symmetric-lenses} in expressiveness.  We further show that the
-class of simple lenses can be characterized semantically as the subset of
-full symmetric lenses for which a simple ``forgetfulness'' property holds
-(\S\ref{sec:related}).
-%
-\item We give an algorithm for synthesizing simple symmetric lenses, using a
+\item We give
+an algorithm for synthesizing lenses built from these combinators, using a 
 novel application of stochastic regular expressions to guide the search
-process (\S\ref{sec:synthesis})\bcp{``and prove it correct''???}.  Along the
-way, we show that {\em star semi-ring equivalences} on regular expressions
-remain valid when applied to stochastic regular expressions, preserving
-probability distributions as well as languages (\S\ref{sec:sre}).
+process (\S\ref{sec:synthesis})\bcp{``and prove it correct''???}.  
 %
@@
 seconds (\S\ref{sec:evaluation}).
+%
+\item We show that simple symmetric lenses fall between asymmetric
+lenses~\cite{Focal2005-long} and full symmetric
+lenses~\cite{symmetric-lenses} in expressiveness.  Indeed, the
+class of simple lenses can be characterized semantically as the subset of
+full symmetric lenses for which a simple ``forgetfulness'' property holds
+(\S\ref{sec:related}).
 \end{itemize}
+Along the
+way, we show that {\em star-semiring equivalences} on regular expressions
+remain valid when applied to stochastic regular expressions, preserving
+probability distributions as well as languages (\S\ref{sec:sre}).
 We close with related work (\S\ref{sec:related}) and
 concluding thoughts (\S\ref{sec:conc}).
+Elided proofs can be found in the appendices\bcp{The appendices are pretty
+  hideous and hard to navigate!!}
+\ifappendices\else included with the supplemental material\fi.
 
@@
       \hline
-      \texttt{company.mgmt} data file & \texttt{company.hr} data file \\
+      Management's data & HR's data \\
       \hline
@@
       \hline
-      \cellcolor{lightyellow}\texttt{company.mgmt} regular expression & 
-      \cellcolor{lightgreen}\texttt{company.hr}  regular expression\\
+      \cellcolor{lightyellow}Management's type & 
+      \cellcolor{lightgreen} HR's type\\
       \hline
@@
   used by management and HR at a US company to represent employee salaries and health
-  insurance providers, respectively. }  
+  insurance providers, respectively. \bcp{The colors are a bit loud,
+    especially since there is so little color elsewhere.}
+  %
+  \bcp{Suggest replacing ``Name,Company'' with ``First Last, Company'' (too
+    many ``name''s flying around)}
+  %
+  \bcp{More substantially, thinking about this example makes me wonder about
+    a whole different way of specifying simple symmetric lenses: rather than
+    just giving corresponding source and target pairs, could we give triples
+    of the form ``new source / old target / new target'' and ``new target /
+    old source / new source''?  I know these would be a bit more work to
+    write, but they would also give the synthesis procedure way more
+    information to work with---it seems to me it should be able to do much
+    less guessing about where things are disconnected.  Would this allow us
+    to drop all the information theory?  Or would we still need it?  (In
+    retrospect, this seems like a fairly obvious thing to try, so it seems
+    important to try it and report on how it performs relative to what we
+    propose here...)}  }
   \label{fig:minimized-representations}
@@
 
-In this section, we give an overview of simple symmetric lenses and our
-synthesis algorithm for them, using a simplified example drawn from a hypothetical
+We begin with an overview of simple symmetric lenses and our
+synthesis algorithm, using a simplified example drawn from a hypothetical
 U.S. company. In this company, management and human resources (HR) store
-information about employees in separate text files: management records the
+information about employees in separate text files: management stores the
 names of 
@@
 The company uses a simple symmetric lens to keep these files
-synchronized. If management adds a new employee, ``Chris Roe: 32500'', a lens
-adds the corresponding entry ``Chris Roe, UNK'' to HR's file. The sentinel value
+synchronized. When management adds a new employee, say ``Chris Roe: 32500'',
+this lens 
+adds the corresponding entry ``Chris Roe, UNK'' to HR's file. The default value
 ``UNK'' represents the fact that the employee's insurance company is currently
-unknown. A similar update happens if HR adds a new employee before management,
-in which case the sentinel value ``unk'' reflects the unknown salary information
+unknown. A similar update happens if HR adds a new employee before
+management knows about them,
+in which case the sentinel value ``unk'' reflects unknown salary information
+on management's side.
 Furthermore, if HR corrects an error in an employee's name, say changing ``John
-Public'' to ``Jon Public'', the lens mirrors that name change in management's
-file. Not all the data is mirrored, however. The management file is not updated
-in response to insurance changes, and the HR file is oblivious to salary
+Public'' to ``Jon Public'', the lens mirrors this change into management's
+file. Not all the data is mirrored, however: management's file is not updated
+in response to insurance changes, and HR's is oblivious to salary
 changes. Simple symmetric lenses are appropriate for keeping these two files
-synchronized because the files have a mix of shared and unshared information.
+synchronized because they contain a mix of shared and unshared information.
 
@@
 \subsection{Simple Symmetric Lenses}
-A simple symmetric lens between formats described by types $X$ and $Y$ is
-defined by four functions that collectively satisfy four ``round-tripping'' laws:
+Semantically, a simple symmetric lens between sets $X$ and
+$Y$ comprises four functions that collectively satisfy four
+``round-tripping'' laws:
 
@@
 \vskip 1ex
+\bcp{I want less space between the put/create laws and their names.}
 \noindent
-The two \KW{create} functions propagate shared information and fill in default
-values when introducing new data. (\EG, the ``unk'' salary entry is added
-alongside the name to the management file when HR inserts a new employee). The
-two \KW{put} functions propagate edits from one format to the other by
-combining the change from one file with the previous value from the other file.
-The round-tripping laws guarantee that if two formats are already synchronized,
-\KW{put}s merely return the second argument, unchanged.
-This definition differs from formulations of classic symmetric lenses,
-which require a \emph{complement} but only two functions.  These two
-functions combine complements with edited data to create updated
-complements and synchronized data, one function for each direction.
-We provide a detailed comparison to classic symmetric lenses
-in \S\ref{sec:relationship}.
+The two \KW{create} functions are used to propagate shared information and
+fill in default values when introducing new data (\EG, the ``unk'' salary
+entry is added alongside the name to the management file when HR inserts a
+new employee). The two \KW{put} functions propagate edits from one format to
+the other by combining a new value from one with an old value from the
+other.  The round-tripping laws guarantee that pushing an ``un-edited''
+value from one side (the result of a put or create) back through a lens in
+the other direction will get back to exactly where we began.
+
+This definition differs from classic symmetric lenses in that it does not
+involve a \emph{complement}.  We give a detailed comparison in
+\S\ref{sec:relationship}.
 
@@
 Generally, we do not expect programmers to define simple symmetric lenses by
-writing the four functions by hand and showing they satisfy the round-tripping
-laws. Instead, we provide simple symmetric lens combinators to build up the
-requisite lenses from simpler ones. We focus on combinators for defining
-lenses between string data formats.
-
-We use regular expressions to denote the types of data formats, and we impose a
-typing discipline on our lenses. Specifically, if $\BRegex$ and $\BRegexAlt$ are
-regular expressions\footnote{We use \BRegex and \BRegexAlt to denote regular
-  expressions. We later use the plain nonterminals \Regex and \RegexAlt for
-  stochastic regular expressions.}, then the notation
-\lstinline{$\Lens$ : $\BRegex \Leftrightarrow \BRegexAlt$} indicates that $\Lens$ is a simple symmetric lens between $\LanguageOf{\BRegex}$ and $\LanguageOf{\BRegexAlt}$.
-When clear from context, we use the term \textit{lens} to denote a simple symmetric lens.
-
-We briefly explain a selection of these
-combinators by using them to define a lens between the two employee data formats
+writing four functions by hand and showing they satisfy the round-tripping
+laws. Instead, we provide a set of combinators---a domain-specific
+language---for building complex lenses from simpler ones.  Such languages
+have been given for building lenses over many different kinds of data,
+including database relations~\cite{BohannonPierceVaughan} and algebraic data
+types~\cite{symmetric-lenses}\bcp{and what else? graphs, etc.?}.  We focus
+here on combinators for defining lenses between string data formats
+described by regular expressions.
+
+If $\BRegex$ and $\BRegexAlt$ are regular expressions, then
+\lstinline{$\Lens$ : $\BRegex \Leftrightarrow \BRegexAlt$} indicates that
+$\Lens$ is a simple symmetric lens between $\LanguageOf{\BRegex}$ and
+$\LanguageOf{\BRegexAlt}$. (Later, we will use undecorated variables for
+stochastic regular expressions; we mark plain REs with overbars.)  When
+clear from context, we shoren ``simple symmetric lens'' to just
+``lens.''
+
+We illustrate some of these
+combinators by defining a lens between the two employee data formats
 in Figure~\ref{fig:minimized-representations}. The finished lens, which appears
-in Figure~\ref{fig:example_lens}, has the type
+in Figure~\ref{fig:example_lens}, has type
 \lstinline{emp_salaries} $\Leftrightarrow$
-\lstinline{emp_insurance}, where
-\lstinline{emp_salaries} and
-\lstinline{emp_insurance} are regular expressions defined in Figure~\ref{fig:minimized-representations}.
+\lstinline{emp_insurance}.
 
@@
 argument a regular expression $\BRegex$ and performs an identity mapping on data
-matching $\BRegex$. The identity lens fully connects the
-corresponding data in the two formats: when one format is
-updated, the other format gets the exact same value. In our example,
-when a \lstinline{name} is updated in one file, the identity lens ensures it is
-also updated in the other file. In this manner, both the \CreateR and \CreateL
-functions are the identity function ($\CreateR \App s = s$), and the put
-functions merely return the first argument ($\PutROf{s_1}{s_2} = s_2$).
-
+matching $\BRegex$. 
 %
@@
 %
-
-In contrast to the identity lens, the \Disconnect lens does not propagate edits
-from one format to the other. The disconnect lens takes four arguments: two
+The identity lens fully connects the corresponding data in the two formats:
+when one is updated, the other gets the exact same value. Both the \CreateR
+and \CreateL functions are the identity function ($\CreateR \App s = s$),
+and the put functions merely return the first argument
+($\PutROf{s_1}{s_2} = s_2$).
+%
+In our example, when a \lstinline{name} is updated in one file, the identity
+lens ensures it is also updated in the other file.
+
+In contrast, the \Disconnect lens does not propagate any edits
+from one format to the other.  It takes four arguments: two
 regular expressions and two strings. The regular expressions specify the formats
 of the corresponding data on the two sides while the strings provide default
-values. On creates, the input values are thrown away, and default values are
+values. 
+%
+\begin{lstlisting}
+disconnect(salary, "", "unk", "") : salary $\Leftrightarrow$ company
+\end{lstlisting}
+%
+On creates, the input values are thrown away, and default values are
 returned ($\CreateROf{\String} =
-\lstinline{"unk"}$), and on puts, the second argument is used, and the first is
+\lstinline{"unk"}$), and on puts, the second argument is used and the first is
 thrown away ($\PutROf{\String}{\StringAlt} = \StringAlt$). In our extended
@@
 The defaults are strings that indicate the information is missing.
-%
-\begin{lstlisting}
-disconnect(salary, "", "unk", "") : salary $\Leftrightarrow$ company
-\end{lstlisting}
-%
+
 The insert \lstinline{ins} and delete \lstinline{del} lenses are
@@
 
-Finally, there are a number of standard lenses that compositionally construct
-more complex lenses from simpler ones. These combinators include concatenation
+Finally, there are a number of combinators that compositionally construct
+more complex lenses from simpler ones. These include concatenation
 ($\ConcatLensOf{\Lens_1}{\Lens_2}$ or $\Lens_1 . \Lens_2$), variation
-($\OrLensOf{\Lens_1}{\Lens_2}$), and iteration ($\IterateLensOf{\Lens}$).  For example;
+($\OrLensOf{\Lens_1}{\Lens_2}$), and iteration ($\IterateLensOf{\Lens}$).
+For example, the composite lens \lstinline{name_lens} in the running
+example
 %
 \begin{lstlisting}
-id(name) . ins(",") . id(name) . del(": ") . ins(",")
+id(name) . id(" ") . id(name) . del(": ") . ins(",")
 \end{lstlisting}
+%
+will transform ``\lstinline{Jane Doe: }'' to ``\lstinline{Jane Doe,}'' in
+the forward direction.
+\bcp{We don't actually give a nontrivial example of how disconnect works.}
+
 % concat(ins(header),employees_lens) : emp_salaries $\Leftrightarrow$ emp_insurance
 %
-appears in the definition of the \lstinline{name_lens} in our running example.
 %\ksf{I don't think this example is necessary.}
@@
 \begin{lstlisting}
-let name_lens = id(name) . ins(",") . id(name) . del(": ") . ins(",") in
+let name_lens = id(name) . id(" ") . id(name) . del(": ") . ins(",") in
 let employee_lens = name_lens . disconnect(salary,"","unk","") 
@@
 %let emp_lens = or(id(""),employees_lens) in
-  \caption{A lens, written in our language for defining simple symmetric lenses
-    on string data, that synchronizes management and HR employee files.
-    Note that we can annotate lenses with types, ensuring that the lens
-    operates on strings that match the regular expressions. In this work, we
-    \textit{synthesize} the lens itself from such regular expressions.}
+  \caption{A lens that synchronizes management and HR employee files\bcp{I'm
+    pretty sure there was an error in name\_lens: the thing between the two
+    names should be id(" ") (was ins('','')).  Crossing my fingers that I'm
+    making a correct fix...}}
   \label{fig:example_lens}
@@
 
-%% \ksf{Consider integrating more detailed discussion of SREs that was
-%%formerly in the intro.  Commented out here for now.}
-%%To implement this intuition, we turn to \emph{stochastic regular
-%%expressions} (SREs). Stochastic 
-%%regular expressions define a set of strings, as do ordinary regular expressions,
-%%but they also provide the expected distribution of strings within that set. Such
-%%distributions allow us to compute the amount of information gained when
-%%presented an element of the set. For example, the language of the SRE
-%%$\lstinline{"a"} |_{0.5}
-%%\lstinline{"b"}$ contains two elements (the strings \lstinline{"a"} and \lstinline{"b"}).
-%%The probability 0.5 annotating the union indicates that both elements are
-%%equally likely to appear. Consequently, if we observe an element of that
-%%language (either \lstinline{"a"} or
-%%\lstinline{"b"}), we will gain one bit of information. Said another way, one bit, together with
-%%the SRE, suffices to recover the string in question. Interestingly, if the
-%%probability annotation were anything but 0.5, the expected information content
-%%(or entropy) would actually be less than 1 (given a stream of such choices, on
-%%average we could encode the the contents of stream in fewer bits than the the
-%%number of choices present). On the other hand, consider the language of the SRE
-%%\lstinline{c}. It contains just one element (the string $c$). If we observe that
-%%our algorithm generates $\lstinline{"c"}$, we gain zero bits of information ---
-%%the string was already determined. Using this foundation, a lens $\Lens$ that
-%%synchronizes data between $\Regex$ and $\RegexAlt$ is ``likely,''
-%%in a technical sense, if given a string $s$ in the source $S$, very little information
-%%is needed to determine which synchronized string $t$ is present in the target
-%%$T$.
-
 This infrastructure gives us what we need to find likely satisfying lenses from
@@
 consider the lens shown in Figure~\ref{fig:example_lens}. 
-Inferring this lens requires \Expand to use a star semiring axiom to
+Inferring this lens requires \Expand to use a star-semiring axiom to
 rewrite the regular expression \lstinline{("\n" . emp_salary)$^*$} within
@@
 
-Without unambiguity constraints it is not easy to calculate the entropy. This
-method is certainly incorrect without unambiguity constraints: consider the SRE
+Without unambiguity constraints it is not easy to calculate entropy (and the
+method we have given is certainly incorrect).  For example, consider the SRE
 $\PRegexOr{\lstinline{"a"}}{\lstinline{"a"}}{.5}$. Using the above computation,s
@@
 overlap. Similar issues occur when trying to find the entropy when $\emptyset$
-is a subterm (what should the entropy of $\PRegexStar{\emptyset}{.5}$ be), so we
-do not define the entropy on $\emptyset$.
-
-In our work, these conditions are merely a technicality, we require unambiguous
-regular expressions as input to our synthesis procedure, and we can easily
-preprocess emptysets out of the stochastic regular expressions that are
-themselves nonempty by traversing the correct star-semiring equivalences (\EG,
+is a subterm (what should the entropy of $\PRegexStar{\emptyset}{.5}$ be?),
+so we do not define entropy on $\emptyset$.
+
+Fortunately, we already require unambiguous regular expressions as input to
+our synthesis procedure for other reasons, and we can easily preprocess
+emptysets out of the stochastic regular expressions that are themselves
+nonempty by traversing the appropriate star-semiring equivalences (\EG,
 \PRegexOr{\emptyset}{\lstinline{"s"}}{.5} \SSREquiv \lstinline{"s"}).
@@
 
-We now provide the typing derivations and semantics for these lenses. We do not
-include the semantics for combinators equivalent to those presented in prior
-work~\cite{symmetric-lenses,Boomerang}. Furthermore, we only present \CreateR
-and \PutR when \CreateL and \PutL are symmetric. Full details are present in the
-appendix. For simplicity of presentation, we define the typing derivations and
-semantics only over unambiguous regular expressions, even though some lenses can
-be defined over ambiguous regular expressions (for example, the identity lens
-doesn't need any ambiguity constraints). We believe relaxing this restriction is
-unproblematic.
+We now provide the typing derivations and semantics for these lenses. We do
+not include the semantics for combinators equivalent to those presented in
+prior work~\cite{symmetric-lenses,Boomerang}. Furthermore, we only present
+\CreateR and \PutR when \CreateL and \PutL are symmetric; full details can
+be found in an appendix\ifappendices\else{} (in the supplemental
+materials)\fi. For simplicity of presentation, we use unambiguous regular
+expressions everywhere, even though some lenses can actually be defined over
+ambiguous regular expressions (for example, the identity lens doesn't need
+any ambiguity constraints\bcp{Is this the only example?}); we believe
+relaxing this restriction is unproblematic.
 
@@
 distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
-current best lens has cost $c$, then:
-\begin{gather*}
-  \Call{Continue}{pq} := c > d + \Log_2(n)
-\end{gather*}
+current best lens has cost $c$, then
+  $\Call{Continue}{pq}$ checks whether $c > d + \Log_2(n)$.
 
@@
 
-For these situations, users can provide a number, $t$, that affects
-the strictness of the termination condition, adjusting the \Continue formula
-as follows. When $\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at
+For these situations, users can provide a number $t$ that affects
+the strictness of the termination condition by adjusting the \Continue formula
+as follows. When $\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, $\Regex$ and
+$\RegexAlt$ are at
 distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
-current best lens has cost $c$:
+current best lens has cost $c$,
 \begin{gather*}
-  \Call{Continue}{pq} := c > \Min(d + \Log_2(n) - t, 0)
+  \Call{Continue}{pq} := c > \Min(d + \Log_2(n) - t, 0).
 \end{gather*}
@@
 termination condition is less strict, and the algorithm will spend less time
-searching for more likely lenses.  Our default $t$ is $0$.
+searching.  The default $t$ is $0$.
 % end synthesis

### 2018-09-18 papers/SymmetricLenses/response_popl19.md
add response draft

Thank you for your comments!

We begin with brief responses to the most significant concerns, followed by a
(long and optional) appendix addressing individual questions.

                                -------------

     - How do we determine what a "correct" lens is? Can we make this process a 
       bit more automatic and objective?

Determining what a correct lens is quite a difficult problem. What a "correct"
lens is inherently subjective -- determining which components of one format
should be synchronized to another requires knowledge of the formats. If there
were a example large suite of synchronized files of these formats, we could use
this example suite as a ground truth. Unfortunately, such a suite does not
exist, and we then must take it upon ourselves to recognize which components of
the file should be synchronized to which.

In validating that the synthesized lenses were correct, we applied the types of
approaches we use in validating our code is correct in everyday development.
1.  We manually inspected the code. This is certainly a nontrivial task, as
    these generated lenses are quite large. However, this task is easier than it
    may initially sound. Much of lenses maintain a rigid structure:
    Concatenations in the lenses typically follow concatenations in the regular
    expressions, disjunctions in the lenses typically follow disjunctions in the
    regular expressions. The primary difficulties arise when seeing how the more
    complex combinators interact. Are the swaps in the right place? Are the
    correct components disconnected? Are lenses merged in the correct way? While
    validation certainly still requires work, instead of closely inspecting
    hundreds of lines of code, we instead must closely inspect a few dozen, and
    give a more cursory glance to the others.
2.  We built unit tests to validate the more complex lenses. Much like everyday
    development, manual inspection of code is rarely sufficient for determining
    correctness. In addition to reading through the code, we added in unit tests
    to validate that the generated code performs correctly on representative
    examples.
3.  Lastly, in addition to unit tests, we ran the lenses on a number of strings,
    taken from the unit tests and the examples, but with minor alterations.
    These quick tests we did not add as unit tests, but served as additional
    validation for ourselves as we were constructing the benchmark suite.

We recognize we did not go into significant enough detail on this in the
submitted version, and will gladly add in such detail on validation in the final
version.

                                -------------

     - Is our synthesis procedure overfit to our benchmarks? Why does assigning
       a fixed probability work?

Our 39 bijective benchmarks came from the original "Synthesizing Bijective
Lenses" benchmark suite without alteration to the regular expressions. Our
additional examples, in particular the data cleaning tasks, were taken from the
original FlashFill paper. Unfortunately, there is no public release of the
FlashFill benchmarks, so our examples had to be taken from the paper, instead of
from a larger benchmark suite.

Assigning a fixed probability is certainly less desirable than inferring it from
data. With a learned distribution, we expect our algorithm to perform better: it
can recognize that certain disjunctions both have high probability, and would
strongly incentivize aligning those critical disjuncts to the critical disjuncts
in the other format. With a learned distribution, we expect that tasks that
currently require 2-3 examples would instead require 1-2. However, this is not
to say that our tool will perform badly as-is on distributions that are far from
our computed ones. On such distributions, our tool will merely not get the
benefits provided by such distributions. It would get an advantage in having an
idea beforehand which components should map to which with a learned
distribution, but merely loses on that advantage without it.

(* This paragraph is to respond to a complaint only of R3, maybe should move
down to below the fold *)

This is not to say that SREs and our semantic metric are not critical for
success. While we use a learned distribution, using such a semantic metric
provides a means of comparison across proposed pairs. With a syntactic metric,
we aren't making an "apples-to-apples" comparison, one lens might have a higher
score, but that score would be based on a different distribution of data - it
actually would be less likely with a different data distribution. Keeping the
probabilities of the two formats is critical for comparing between proposed
types.

===========================================================================
===========================================================================
Detailed responses to individual questions in the reviews follow...

===========================================================================
Review #43A

Paper summary
-------------
This paper investigates the automatic synthesis of symmetric lenses, i.e.,
programs that allow data to be synchronized (updated in both directions)
between two data sources, such as two databases. Specifically, this paper
identifies a well-behaving subclass of symmetric lenses, called simple
symmetric lenses, for which the authors are able to develop an efficient
synthesis algorithm. The main difference between simple and ordinary
symmetric lenses is that the former do not contain a notion of complement,
significantly reducing the data the user has to provide for synthesis.

In detail, this paper:

- Introduces a refinement of the notion of symmetric lens, called simple
  symmetric lenses, that are well-suited for synthesis purposes. The
  paper organises their definition into a smalltyped language that provides
  a set of combinators for building complex lenses from simpler ones.

- Proposes a (mostly, up to provided examples and tuning of parameters)
  automatic synthesis algorithm, which 1) first expands the given regular
  expressions into stochastic regular expressions, and 2) then tries to
  iteratively find a "likely" simple symmetric lens between stochastic
  regular expressions equivalent to the original ones. The algorithm
  terminates when certain preference and termination conditions are
  satisfied.

- Extends the standard notion of star-equivalence to stochastic regular
  expressions so as to compare and (soundly) rewrite them.

- Combines stochastic regular expressions with ideas from information
  theory, using the entropy of data sources to compute costs for lenses,
  so as to guide synthesis towards "more likely" candidate lenses, i.e.,
  those lenses that propagate a lot of information.

- Defines a notion of DNF stochastic regular expressions and a 
  corresponding notion of DNF symmetric lenses that are used 
  as an intermediate representation in lens synthesis.

- Incorporates the proposed synthesis algorithm into the Boomerang
  language and evaluates it on 48 benchmarks that cover a variety of
  real world data formats, ranging from converting Linux configuration
  files from ad hoc textual presentation to structured dictionaries to
  synchronising BibTex and EndNote citation descriptions.

  While the naive implementation of this synthesis algorithm was only
  able to find 26 of the required lenses, then a mildly optimised
  variant of it (supporting compositional synthesis) was able to
  synthesise good lenses for all 48 benchmarks, each under 30 sec.

  The paper also demonstrates that when restricted to bijective
  lenses, the proposed synthesis algorithm is roughly on bar with
  an existing dedicated synthesis algorithm for such lenses.

As I think the paper does a good job of combining theoretical and
practical research, e.g., by identifying a useful and large subclass
of symmetric lenses, and demonstrating that lenses in this class can
be efficiently (mostly automatically) synthesised for many real world
data formats using just a few examples. Also, as far as I can tell,
the developed mathematics is correct (modulo some typos listed below).
Thus I recommend this paper to be accepted at POPL 2018. However, as
I am not an expert on program synthesis, I will leave scrutinising
the related work and this work's relationship to it to experts.

Strengths
---------
The paper does a good job of combining theoretical and practical
research, e.g., by identifying a useful and large subclass of
symmetric lenses; by using information-theoretic ideas to define
a notion of lens cost that can be used effectively for candidate
selection in synthesis; and by demonstrating that the proposed
algorithm works well on various real world data formats.

Weaknesses
----------
I have no significant complaints about this paper apart from Section 3
not making it clear that stochastic regular expressions are not a
contribution of the authors (albeit it is mentioned in passing on p2).
My other comments are mostly about layout and (what are hopefully) typos.

Comments for author
-------------------
Comments:

1. It is usually a good style not to include citations in abstracts.

2. You should mention Optician when referring to the work of Miltner
   et al. in the Introduction. Otherwise its mention in the beginning
   of Section 6 kind of comes out of thin air.

3. p5 l236,

   Shouldn't the second argument to `disconnect` be `company`?

4. p6, l288,

   If I understood correctly, this parituclar lens `disconnect(...)`
   picks the default values at random, but is still valid because it
   satisfies the required lens type. If so, please make it clearer.

5. p7, l321,

   At this point the $\mathbb{H}$ notation for entropy has not yet
   been introduced, perhaps introduce it above for `S` and `P`.

6. p9, l438,

   The combination of `S` and the entropy formula in the parentheses
   looks a bit like some sort of function application, please make
   it clearer that it isn't.

7. p10, l449

   It might be worth reminding the reader what unambiguous regular
   expressions are.

8. p16, l736,

   It could be worthwhile pointing out explicitly that the structure 
   of symmetric DNF lenses mirrors exactly that of SDNF REs.

9. p16, l777,

   Could you give concrete examples of situations where all sequences
   are involved in sequence lenses, but there are examples that could
   make other lenses more useful?

10. p18, l842

    Could you explain in some more detail what criteria does `EXPAND` use
    to decide to open some closed expressions but not others?

11. p19, l887

    Surely the definition of `CONTINUE(pq)` should be using `max`? Otherwise
    I don't see how the definition makes sense, because using surely `min` 
    will just pick 0 for t>=0, because the distance and the logarithm are
    positive values.

12. p21, l1022

    Which preference metric are the t=0, t=25, t=-25 benchmarks using?

Typos:

- p3, l119, "will depend which" -> "will depend on which"

- p5, l220, "we shoren" -> "we shorten"

- p5, l230, "putR s1 s2 = s2" -> "putR s1 s2 = s1"

- p8, l360, "GREEDYSYNTH converts yields" -> "GREEDYSYNTH yields"

- p9, l414, "annotation p how" -> "annotation p represents how"

- p10, l478, "typing derivations" -> "typing rules"

- p11, l478, "these lenses" -> "these combinators"

- p12, l565, "H->(T|l1,S1) + H->(T|l2,S2)" -> "H->(T1|l1,S) + H->(T2|l2,S)"

- p12, l573, "uncommon on practice" -> "uncommon in practice"

- p12, l575, "maximum sum of recovering" -> "sum of maximums of recovering"

- p12, l583, "definition of H contains" -> "definition of H-> contains"

- p14, l657, "SDNF REs are SDNF REs" -> "SDNF REs are DNF REs"

- p15, l715, "expression., " -> "expression, "

- p15, l721, "syntactic means of" -> "syntactic means for"

- p15, l723, "DNF SRE" -> "SDNF RE"

- p17, l820, "pair of stochastic atoms" -> "pair of lists of stochastic atoms"

- p19, l926, "runs the tools, determines" -> "runs the tool, determines"

- p21, l1100, "simple lenses" -> "simple symmetric lenses"

Questions for authors’ response
---------------------------------
First, I would really like if the authors could provide answers to
questions 9, 10, 11, and 12 above.

Second, from what I remember from the literature on (ordinary) symmetric 
lenses, they do not lend themselves very well to analysis in terms of 
spans of asymmetric lenses. With simple symmetric lenses, do you think 
the situation is better now that you have omitted the complement part?

Thirdly, do the authors expect to be able to extend the proposed synthesis
algorithm to handle any of the many variants of updates-based lenses, 
such as edit lenses, delta lenses, and update lenses?

===========================================================================

Review #43B
Paper summary
-------------
The paper proposes simple symmetric lenses --- symmetric lenses that
are forgetful, i.e. the complement can be determined by the most
recent “source” and “view” --- and believe they are practically
useful. Different from traditional approaches, after defining basic
combinators (of simple symmetric lenses), the paper votes for
synthesising lenses instead of manually writing them. To do so, the
paper proposes a novel SRE and information-theoretic based approach,
which makes it stand out from the previous synthesising work on
(quotient) bijective lenses. The evaluation is based on the previous
Optician benchmarks including 48 tasks. With all optimisations turned
on, it is able to finish each work in 30 seconds. The paper also
discusses the relationship with asymmetric lenses and symmetric
lenses, and proves that simple symmetric lenses lie in between them.

Strengths
---------
- The SRE and information-theoretic based synthesising approach is
  novel and may have much influence on follow-up research.

- The evaluation shows that the authors' approach has very fast speed
  on the tasks adapted from other real world application. (One
  problem of program synthesis is its low speed.)

- The appendix provides details about the specification of simple
  symmetric lens combinators and proofs of theorem in the paper.

Weaknesses
----------
- The main weakness is in the evaluation part. There are many
  unclear/controversial parts regarding the evaluation. For example,
  the lack of description of the benchmark environment; the way they
  test whether the synthesised lens is correct; the description of the
  tasks to be synthesised. See detailed comments below. (This defect
  does not prevent it from being a good paper.)

- The first subsection of Related Work did not give any reference to
  the related work except for Foster et al. 2007, which makes the
  paper unfriendly and hard to verify. The appendix is also in a
  messy.

Comments for author
-------------------
L11: The reference Martin Hofmann et al. (2015) does not help since it
is not yet published and cannot be obtained from Internet. Martin
Hofmann et al. (2011) may be the choice for this time.

L230: `putR s1 s2 = s2`  --> `putR s1 s2 = s1`

L238: `createR s = "unk"` --> `createL _ = "unk"`  
These two errors will be misery for non-experts.  
BTW, the authors sometimes use `s` and `v`, sometimes use `s1` and `s2`. It will be better to use a uniform notation.

L401 and L411: `ϵ` is not in the grammar of SRE; it appears without
any explanation.

L442: H(s) should be 0. Otherwise the entropy of the SRE in L452 is 2.

L641 and L642: previously they are `S*p -> ϵ |1-p (S・S*p)` and `S*p
-> ϵ |1-p (S*p・S)`, instead of `S*p -> ϵ |p (S・S*p)` and `S*p -> ϵ
|p (S*p・S)`.

L484 and L485:
I cannot see the reason to relax the restriction regarding unambiguous
REs. What will happen if primitive lens combinators for unambiguous
REs are composed with the ones for ambiguous REs? Since complex lenses
are composed from (ambiguous and unambiguous) primitive lenses, in
practice the restriction should be always there.

L557 and below: Why some of the entropy is a single value while some
of them seems ranging over an interval? If the interval is used to
handle `merge_left`, then the entropy of all combinators should be an
interval. Currently only `id`, `disconnect`, and `merge_left` yield
intervals.

L584 ~ L606: From some other perspective, if the overall effect is
just a bijection, there is no reason to further decompose the
bijection into non-bijective ones in practice.

L718 and L719: Different kinds of data definitely have different
shapes. I cannot understand why assigning a fixed probability will fit
in all situations. Could you explain it a bit?

L747: I skimmed this subsection and believe that the algorithm is
correct (since there is an implementation). There are too many small
caps in the algorithm figures.

L866 - L868: The user provides two REs describing the source and view
as input, which should already give some useful structure. For
example, the top-level nonterminal may be described by several other
non-terminals. This structure indeed suggests that a big problem can
be divided into several subproblems.  Why do the user still need to
manually "generate subproblems"? Is it avoidable?

L900: Before introducing the benchmark suite, the authors had better
provide the specification, regarding CPU, memory, etc., of their
benchmark environment since an important measurement is the speed of
the synthesis process.

L909 and L956: About the benchmark suite, how many of the benchmarks
in fact only need bijective lenses? This definitely should be
considered since (1) the benchmark suite is adapted from the "original
Optician benchmarks, comprised of 39 bijective synthesis tasks", and
(2) "Adjusting the termination parameter was needed in only 9 of the
48 tasks". I suppose it is not, but still wonder, if for all of the 48
- 39 = 9 non bijective tasks, we need to adjust their termination
parameters?

L922: The authors may add a few words on how representative and
complex the selected tasks is. For example, I got the data from the
paper *Deep API Programmer: Learning to Program with APIs* that there
is 238 real-world FlashFill benchmarks; then why the 8 data cleaning
tasks here are more representative among the 238 tasks? The reviewers
(and readers in the future) may want to know to which extent the job
of writing lenses can be automatically done.

L926 and L952: This is the most confusing point for me. The meaning of
"correct lenses" here seems very subjective as the authors let the
user determine whether the synthesised lens is desired. Can we make
this process a bit automatic and objective?

For instance, usually a natural and reasonable approach is to divide
examples into a "training" set and a "validation" set; we use the
training set to synthesise desired lenses and use the validation set
to test if the synthesised lenses work well (although this does not
prove its correctness). Even if the user is finally involved in
determining whether the lens is a desired one, this process may reduce
the burden of the user.

(I also wonder how can the user verifies whether a complex lens of
hundreds or thousands of lines code is the desired one by directly
inspecting the lens program?)

L945 and L993 (Fig.6 and Fig.7): Since the time is only within 30
seconds, why not make Figures whose x-axes are linear? With the
"non-linear figures", I cannot tell that SS took 1.3 times longer than
BS. In addition, since we only care about the average time of tasks
(and the maximum time of a single task), there is no reason to use a
line chart, whose shape is drastically affected by the order of
tasks. A simple table is better, if the details of tasks (and their
order) are not given.

L1055: Again, if the default (termination) parameter can only find
bijective lenses, it is not that good.

L1066: When discussing symmetric lenses, the authors need to give a
reference to Symmetric Lenses by Martin Hofmann et al. (2011) since
the definitions and notations are from there.

L1079: The authors silently move to introduce edit lenses, although
different from Edit Lenses by Martin Hofmann et al. (2012), a short
introduction and a reference are needed. Moreover, the authors may
need to explicitly mention that the edits in this section are
"overwrites"; otherwise it is difficult to understand why the edits
have the same type as the source and view (ignoring the sum types made
by inl and inr.), as usually they are some "deltas".

L1080: The authors did not explain the notation `l ∈ X ⟷ Y` as it
hides the complement `C`.

L1105 and below: Similarly, the authors need to explain the inference
rules and the involved symbols more. For instance, what is `xyo`?
There is even no note telling the reader that some explanation can be
found in the appendix.

L1111 and L1112: I cannot figure out where does `x` below the
"horizontal line" come from

L1170: I do believe information-theoretic analysis is of extremely
useful; if possible, adding one more example will make the argument
stronger.

##Things that hinder reading

L262: this is the first time that "forward direction"
appears. Besides, for symmetric lenses, it is unclear what is the
forward direction.

L364: `... within emp_salaries ...`. According to the context, this
`emp_salaries` is supposed to to be found in Figure 3 and there
happens to be a lens type called `emp_salaries`. But in fact the
`emp_salaries` should refer to the one in Figure 2.

L659: The notation (DS,DT) might need some explanation. It is not a
tuple but just introduce two preferable names DS and DT.

L1101 and L 1103: the authors use "edit `esi`" here but use "edits
`esi`" at L1082. Is `esi` a single edit or a sequence of edits? Having
read it many times, I eventually understand that it should be the
single edit that we focus on; so that `esi = inl x`.

L1509 and L1515: the two lines exceed page limit. Luckily the exceeded
parts can be guessed.

## minors:

L14: that -> than

L54: Indeed of -> Indeed among

L220: shoren -> shorten

L372: between -> among

L414 and L415: some words are missing in the sentence so that it is
not easy to understand.

L426: a means to reason -> a means of reasoning

L431: present -> presented ?

L438: the authors had better to put the definition of entropy
somewhere earlier in this sentence. Otherwise it syntactically looks
like the definition of data source S.

L451: a redundant ",s"

L453: "in subtle ways, the algorithm" -> "in subtle ways and the algorithm"

L459: I do not understand the example. Which law gives the result?

L542: combonator -> combinator

L547: equivalent -> is equivalent

L553: does `H->(l,s | T)` looks better than the original `H->(T | l,s)`? There is no reason to use the same argument order for `H->` and `H<-` since we can think them as different functions.

L620: I cannot parse the sentence

L656: SDNF REs are SDNF REs with probability annotations -> SDNF REs
are DNF REs with probability annotations.

BTW, the authors had introduced the abbreviation SRE in the
introduction. But in this subsection the authors use too many
different abbreviations:

```
SDNF REs
stochastic DNF REs
stochastic DNF regular expressions

SREs
stochastic REs
stochastic regular expressions

REs
regular expressions
```
Please consider use the same notation throughout.

L658: `DS bar` is introduced but seems to be never used.

L715: expression., -> expression,

L734: than there are DNF bijective lenses -> than DNF bijective lenses

L845: uncoverted -> unconverted

L928: whether it is usable to generate synthesize correct lenses ->
whether it is usable to synthesize correct lenses

L929: I prefer to remove the "the" in the first sentence.

L971: measuring it shows of the responsiveness -> measuring it shows
the responsiveness

L1080 ~ L1082: the description of the function apply is its curried
version while the usage below is the uncurried version.

L1082: edits `esi` -> edit `esi`

L1099: I guess the title should be "Relation with Symmetric Lenses"

L1099: classic -> classical

L1136: `l.put s v` -> `l.put v s`

L1138 `l.put s (l.get s)` -> `l.put (l.get s) s`

L1512: I found `(y,c2') = l.putR (x2', c1')`, do you mean `(y,c2') =
l.putR (x2', c2')`

L1526 and L1530: sometimes x and y are used and sometimes s and v.

L1702: this section is not FORGETFUL SYMMETRIC LENSES (which is the
previous section)

L2088: perhaps the theorem about lenses should be put to the previous
section. This section is about SRE and information theory.

L2098: this is CREATEGET, not PUTGET again

L2112: `l.get` -> `l.get s`

Questions for authors’ response
---------------------------------
Please clarify the evaluation part.

===========================================================================
Review #43C

Paper summary
-------------
- This paper proposes a technique for synthesizing a form of lenses from input-output examples and types. 
- Lenses are transformations from one data type to another. They are typically used to transform between data formats, e.g., CSV to JSON.
- Previous work on synthesizing lenses covered the bijective case, where the synthesized lens is not lossy.
- This paper builds on top of prior work to synthesize lossy lenses, which they call simple symmetric lenses. These lenses allow data transformations that lose data, e.g., extracting one column from a table.
- Compared to bijective lenses, the search space is less constrained. The observation of previous work was that the bijectivity constraint is so restrictive that any lens synthesized is likely to be the right one. By removing this constraint, the search space is much larger.
- To work around that restriction, the authors propose to use probabilistic types, which define a probability distribution over inhabitants of the type. Then, they propose the heuristic where they look for lenses that minimize entropy in the transformation — roughly the number of bits needed to recover one format from the other.
- The synthesis algorithm largely follows that from POPL18 on bijective lenses — types are defined via regular expressions, regular expressions are unrolled (expanded) into DNF form, and then a greedy search looks for a lens for that pair of types.
- The experiments demonstrate the usefulness of the search heuristics used—particularly using entropy to guide the search.

Strengths
---------
+ Interesting application of synthesis
+ Elegant set of combinators defining simple symmetric lenses
+ Intriguing information-theoretic search heuristic

Weaknesses
----------
— Small algorithmic delta over POPL18

— Unclear whether stochastic REs are needed

Comments for author
-------------------
I enjoyed reading this paper. I think the technique is elegant and well-reasoned. The usability aspect is questionable: It seems that if the user has defined complex regular expressions defining the pair of languages, then they can probably go ahead and just implement the lens without resorting to examples.

While the paper begins with a nice high-level illustration and examples, the technical sections lack illustrative examples and motivation that it’s hard to appreciate some points. For example, Section 4.1 introduces the lens cost computation, but no example of such computation is given. More important: It is unclear to the general reader why the Expand procedure is needed — i.e., why do we need to rewrite types for GreedySynth.

The ToStochastic procedure felt quite ad hoc to me. The paper makes a point that if the user knew the distribution (e.g., if they automatically infer the parameters from data), then they can supply it. But will that really help the search? Essentially, I wish there were examples (even synthetic ones) where knowing the distribution of data can clearly influence the synthesis procedure. The paper resorts to a fixed distribution that gives the same probability to each disjunct. This raises two concerns:

1.  that this overfits to the hand-picked benchmarks and 
2.  that the entropy calculation and stochastic REs are an overkill.

Why not use a syntactic heuristic. That’s what DC is trying to do, but perhaps a slightly more sophisticated syntactic heuristic could achieve the same effect. So, while the use of entropy is cool, appealing, and new in synthesis (to my knowledge), the paper does not convince me that it is needed here.

Section 6 or 7 should discuss examples that are outside the expressivity of the lens language and/or take too much time. It was hard for me to infer what the limitations of the approach are.

Questions for authors’ response
---------------------------------
My main concern is whether the use of stochastic REs is really useful (see above)

Are stochastic REs new? I didn't see any citations when they were introduced.

### 2018-09-18 papers/SymmetricLenses/response_popl19.md
add fine-grained repsonses

@@
 exist, and we then must take it upon ourselves to recognize which components of
-the file should be synchronized to which.
+the file should be synchronized to which, and validate that the lenses
+corresponded to our internal notion of correctness.
 
 In validating that the synthesized lenses were correct, we applied the types of
-approaches we use in validating our code is correct in everyday development.
+approaches we use in validating that our code is correct in everyday
+development.
 1.  We manually inspected the code. This is certainly a nontrivial task, as
@@
 
-Paper summary
                                 -------------
-This paper investigates the automatic synthesis of symmetric lenses, i.e.,
-programs that allow data to be synchronized (updated in both directions)
-between two data sources, such as two databases. Specifically, this paper
-identifies a well-behaving subclass of symmetric lenses, called simple
-symmetric lenses, for which the authors are able to develop an efficient
-synthesis algorithm. The main difference between simple and ordinary
-symmetric lenses is that the former do not contain a notion of complement,
-significantly reducing the data the user has to provide for synthesis.
 
-In detail, this paper:
-
-- Introduces a refinement of the notion of symmetric lens, called simple
-  symmetric lenses, that are well-suited for synthesis purposes. The
-  paper organises their definition into a smalltyped language that provides
-  a set of combinators for building complex lenses from simpler ones.
-
-- Proposes a (mostly, up to provided examples and tuning of parameters)
-  automatic synthesis algorithm, which 1) first expands the given regular
-  expressions into stochastic regular expressions, and 2) then tries to
-  iteratively find a "likely" simple symmetric lens between stochastic
-  regular expressions equivalent to the original ones. The algorithm
-  terminates when certain preference and termination conditions are
-  satisfied.
-
-- Extends the standard notion of star-equivalence to stochastic regular
-  expressions so as to compare and (soundly) rewrite them.
-
-- Combines stochastic regular expressions with ideas from information
-  theory, using the entropy of data sources to compute costs for lenses,
-  so as to guide synthesis towards "more likely" candidate lenses, i.e.,
-  those lenses that propagate a lot of information.
-
-- Defines a notion of DNF stochastic regular expressions and a 
-  corresponding notion of DNF symmetric lenses that are used 
-  as an intermediate representation in lens synthesis.
-
-- Incorporates the proposed synthesis algorithm into the Boomerang
-  language and evaluates it on 48 benchmarks that cover a variety of
-  real world data formats, ranging from converting Linux configuration
-  files from ad hoc textual presentation to structured dictionaries to
-  synchronising BibTex and EndNote citation descriptions.
-
-  While the naive implementation of this synthesis algorithm was only
-  able to find 26 of the required lenses, then a mildly optimised
-  variant of it (supporting compositional synthesis) was able to
-  synthesise good lenses for all 48 benchmarks, each under 30 sec.
-
-  The paper also demonstrates that when restricted to bijective
-  lenses, the proposed synthesis algorithm is roughly on bar with
-  an existing dedicated synthesis algorithm for such lenses.
-
-As I think the paper does a good job of combining theoretical and
-practical research, e.g., by identifying a useful and large subclass
-of symmetric lenses, and demonstrating that lenses in this class can
-be efficiently (mostly automatically) synthesised for many real world
-data formats using just a few examples. Also, as far as I can tell,
-the developed mathematics is correct (modulo some typos listed below).
-Thus I recommend this paper to be accepted at POPL 2018. However, as
-I am not an expert on program synthesis, I will leave scrutinising
-the related work and this work's relationship to it to experts.
-
-Strengths
----------
-The paper does a good job of combining theoretical and practical
-research, e.g., by identifying a useful and large subclass of
-symmetric lenses; by using information-theoretic ideas to define
-a notion of lens cost that can be used effectively for candidate
-selection in synthesis; and by demonstrating that the proposed
-algorithm works well on various real world data formats.
-
-Weaknesses
-----------
-I have no significant complaints about this paper apart from Section 3
+     - I have no significant complaints about this paper apart from Section 3 
        not making it clear that stochastic regular expressions are not a
        contribution of the authors (albeit it is mentioned in passing on p2).
-My other comments are mostly about layout and (what are hopefully) typos.
 
-Comments for author
--------------------
-Comments:
+We did not intend to forget such citations or claim that we are introducing
+stochastic regular expressions. We will be sure, in future versions, to make
+this distinction more clear.
 
-1. It is usually a good style not to include citations in abstracts.
+                                -------------
 
-2. You should mention Optician when referring to the work of Miltner
-   et al. in the Introduction. Otherwise its mention in the beginning
-   of Section 6 kind of comes out of thin air.
-
-3. p5 l236,
+     - 3. p5 l236,
 
@@
 
-4. p6, l288,
+We should have the second type be "" (to be consistent with what we use in the
+final lens)
+
+                                -------------
+
+     - 4. p6, l288,
 
@@
 
-5. p7, l321,
+This is correct
 
-   At this point the $\mathbb{H}$ notation for entropy has not yet
-   been introduced, perhaps introduce it above for `S` and `P`.
+                                -------------
 
-6. p9, l438,
-
-   The combination of `S` and the entropy formula in the parentheses
-   looks a bit like some sort of function application, please make
-   it clearer that it isn't.
-
-7. p10, l449
-
-   It might be worth reminding the reader what unambiguous regular
-   expressions are.
-
-8. p16, l736,
-
-   It could be worthwhile pointing out explicitly that the structure 
-   of symmetric DNF lenses mirrors exactly that of SDNF REs.
-
-9. p16, l777,
+     - 9. p16, l777,
 
@@
 
-10. p18, l842
+Yes, one of our benchmarks is such an example. In particular, consider trying to
+synchronize Windows "Scheduled Tasks" and Linux "cron jobs." In such a task, we
+wish to synchronize the times that jobs are run, but not synchronize the
+commands to run the jobs themselves. In such a scenario, we could create a
+bijective lens, but this is not correct behavior. Instead, examples can
+demonstrate that these components should remain unsynchronized.
+
+                                -------------
+
+     - 10. p18, l842
 
@@
 
-11. p19, l887
+Full details are present in the prior work on Synthesizing Bijective Lenses
+(page 16). From a high level perspective, we process both regular expression
+trees, looking for situations where a regular expression is as a subcomponent of
+one format, but not the other, and expand such REs (this is where hashconsing
+provides its greatest benefits). Furthermore, we look for instances where a
+regular expressions may be present in one format, but is only contained within a
+closed regular expressions in the other, and expand the closed RE containing the
+missing RE. We then make this apply in slightly more places by also taking into
+account how deeply nested within stars certain REs occur.
+
+                                -------------
+
+     - 11. p19, l887
 
@@
 
-12. p21, l1022
+Correct, this should be max
+
+                                -------------
+
+     - 12. p21, l1022
 
@@
 
-Typos:
+These benchmarks are using the full "SS" mode. The only difference between these
+benchmarks and the full mode is that termination parameters are ignored, and
+are instead required to use the 0, 25, and -25 parameters.
 
-- p3, l119, "will depend which" -> "will depend on which"
+                                -------------
 
-- p5, l220, "we shoren" -> "we shorten"
-
-- p5, l230, "putR s1 s2 = s2" -> "putR s1 s2 = s1"
-
-- p8, l360, "GREEDYSYNTH converts yields" -> "GREEDYSYNTH yields"
-
-- p9, l414, "annotation p how" -> "annotation p represents how"
-
-- p10, l478, "typing derivations" -> "typing rules"
-
-- p11, l478, "these lenses" -> "these combinators"
-
-- p12, l565, "H->(T|l1,S1) + H->(T|l2,S2)" -> "H->(T1|l1,S) + H->(T2|l2,S)"
-
-- p12, l573, "uncommon on practice" -> "uncommon in practice"
-
-- p12, l575, "maximum sum of recovering" -> "sum of maximums of recovering"
-
-- p12, l583, "definition of H contains" -> "definition of H-> contains"
-
-- p14, l657, "SDNF REs are SDNF REs" -> "SDNF REs are DNF REs"
-
-- p15, l715, "expression., " -> "expression, "
-
-- p15, l721, "syntactic means of" -> "syntactic means for"
-
-- p15, l723, "DNF SRE" -> "SDNF RE"
-
-- p17, l820, "pair of stochastic atoms" -> "pair of lists of stochastic atoms"
-
-- p19, l926, "runs the tools, determines" -> "runs the tool, determines"
-
-- p21, l1100, "simple lenses" -> "simple symmetric lenses"
-
-Questions for authors’ response
----------------------------------
-First, I would really like if the authors could provide answers to
-questions 9, 10, 11, and 12 above.
-
-Second, from what I remember from the literature on (ordinary) symmetric 
+      - Second, from what I remember from the literature on (ordinary) symmetric
         lenses, they do not lend themselves very well to analysis in terms of 
@@
 
-Thirdly, do the authors expect to be able to extend the proposed synthesis
-algorithm to handle any of the many variants of updates-based lenses, 
-such as edit lenses, delta lenses, and update lenses?
+TODO: Learn Category Theory
+
+                                -------------
+
+       - Thirdly, do the authors expect to be able to extend the proposed 
+         synthesis algorithm to handle any of the many variants of updates-based
+         lenses, such as edit lenses, delta lenses, and update lenses?
+
+Yes, I do expect that the algorithm can be extended to handle such variants,
+though likely it would require some additional work.
+
+Delta lenses would require additional work in understanding "shape" constructs.
+Currently our language only permits basic iteration, but shapes allow for
+different types of mergings. This would require changes to AtomSynth, in that we
+would we would need to discover the correct shapes of the data, and align them
+in the correct ways.
+
+Both edit and update lenses would require new notions of specifications. Both
+lens types permit updates that are not merely pushing data through, but instead
+are applying specific elements of an edit monoid. We expect that our current IO
+approach to specifications would not work as we wished for such transformations.
+(Also we would remove complements from edit lenses).
 
 ===========================================================================
+Review #43B
 
-Review #43B
-Paper summary
                                 -------------
-The paper proposes simple symmetric lenses --- symmetric lenses that
-are forgetful, i.e. the complement can be determined by the most
-recent “source” and “view” --- and believe they are practically
-useful. Different from traditional approaches, after defining basic
-combinators (of simple symmetric lenses), the paper votes for
-synthesising lenses instead of manually writing them. To do so, the
-paper proposes a novel SRE and information-theoretic based approach,
-which makes it stand out from the previous synthesising work on
-(quotient) bijective lenses. The evaluation is based on the previous
-Optician benchmarks including 48 tasks. With all optimisations turned
-on, it is able to finish each work in 30 seconds. The paper also
-discusses the relationship with asymmetric lenses and symmetric
-lenses, and proves that simple symmetric lenses lie in between them.
 
-Strengths
----------
-- The SRE and information-theoretic based synthesising approach is
-  novel and may have much influence on follow-up research.
+     - The main weakness is in the evaluation part. There are many 
+       unclear/controversial parts regarding the evaluation. For example, the 
+       lack of description of the benchmark environment; the way they test 
+       whether the synthesised lens is correct; the description of the tasks to
+       be synthesised. See detailed comments below. (This defect does not
+       prevent it from being a good paper.)
 
-- The evaluation shows that the authors' approach has very fast speed
-  on the tasks adapted from other real world application. (One
-  problem of program synthesis is its low speed.)
+All benchmarks were performed on a 2.5 GHz Intel Core i7 processor with 16 GB of
+1600 MHz DDR3 running macOS High Sierra. We provide a more detailed answer our
+benchmark suite and method of finding correctness in the main portion of the
+paper.
 
-- The appendix provides details about the specification of simple
-  symmetric lens combinators and proofs of theorem in the paper.
-
-Weaknesses
-----------
-- The main weakness is in the evaluation part. There are many
-  unclear/controversial parts regarding the evaluation. For example,
-  the lack of description of the benchmark environment; the way they
-  test whether the synthesised lens is correct; the description of the
-  tasks to be synthesised. See detailed comments below. (This defect
-  does not prevent it from being a good paper.)
+                                -------------
 
@@
 
-Comments for author
--------------------
-L11: The reference Martin Hofmann et al. (2015) does not help since it
+We apologize for the lack of citations and messy appendix. For the first
+subsection of Related Work, we got carried away in making formal claims, and did
+not include information comparisons to other lens languages. We will make sure
+to make such comparisons and citations in the final version of the paper. We
+will similarly improve the appendix.
+
+                                -------------
+
+     - L11: The reference Martin Hofmann et al. (2015) does not help since it
        is not yet published and cannot be obtained from Internet. Martin
@@
 
-L230: `putR s1 s2 = s2`  --> `putR s1 s2 = s1`
+Thank you, you are correct, we will change this citation.
 
-L238: `createR s = "unk"` --> `createL _ = "unk"`  
-These two errors will be misery for non-experts.  
-BTW, the authors sometimes use `s` and `v`, sometimes use `s1` and `s2`. It will be better to use a uniform notation.
+                                -------------
 
-L401 and L411: `ϵ` is not in the grammar of SRE; it appears without
-any explanation.
+     - L442: H(s) should be 0. Otherwise the entropy of the SRE in L452 is 2.
 
-L442: H(s) should be 0. Otherwise the entropy of the SRE in L452 is 2.
+You are correct.
 
-L641 and L642: previously they are `S*p -> ϵ |1-p (S・S*p)` and `S*p
--> ϵ |1-p (S*p・S)`, instead of `S*p -> ϵ |p (S・S*p)` and `S*p -> ϵ
-|p (S*p・S)`.
+                                -------------
 
-L484 and L485:
+     - L484 and L485:
        I cannot see the reason to relax the restriction regarding unambiguous
@@
 
-L557 and below: Why some of the entropy is a single value while some
+Within our synthesis algorithm, you are, in general, correct. However, Boomerang
+exists beyond Optician, and permits lenses like id("a"|"a"). In our extension to
+Boomerang, we do not remove such freedom, and continue to permit such languages.
+
+                                -------------
+
+     - L557 and below: Why some of the entropy is a single value while some
        of them seems ranging over an interval? If the interval is used to
@@
 
-L584 ~ L606: From some other perspective, if the overall effect is
-just a bijection, there is no reason to further decompose the
-bijection into non-bijective ones in practice.
+This was us trying to make our equations more readable, but we did not
+sufficiently explain what our syntax meant. The calls to H(...) return pairs,
+and when multiplied by constants, they act as follows:
 
-L718 and L719: Different kinds of data definitely have different
+a x (b,c) = (a x b,a x c)
+(a,b) + (c,d) = (a + b,c + d)
+
+We will make sure to explain this more clearly in a final version of the paper.
+
+                                -------------
+
+     - L718 and L719: Different kinds of data definitely have different
        shapes. I cannot understand why assigning a fixed probability will fit
@@
 
-L747: I skimmed this subsection and believe that the algorithm is
-correct (since there is an implementation). There are too many small
-caps in the algorithm figures.
+See the main response above the fold.
 
-L866 - L868: The user provides two REs describing the source and view
+                                -------------
+
+     - L866 - L868: The user provides two REs describing the source and view
        as input, which should already give some useful structure. For
@@
 
-L900: Before introducing the benchmark suite, the authors had better
+We do not believe such subproblems are straightforwardly avoidable. While we can
+attempt to match up subcomponents, this matching oftentimes will not work
+because of, for example, issues where various equivalences must be applied (for
+example: ("a"|"b")|("c"|"d") <-> "a"|("b"|("c"|"d"))) do not break down. If we
+had attempted to initially try breaking down the problem syntactically, we would
+have wasted time. This is particularly an issue because bad specifications could
+take a very long time, as there may be no good lenses between the attempted
+proposed subcomponents, so there would potentially be a great deal of wasted
+time in these attempts.
+
+                                -------------
+
+     - L900: Before introducing the benchmark suite, the authors had better
        provide the specification, regarding CPU, memory, etc., of their
@@
 
-L909 and L956: About the benchmark suite, how many of the benchmarks
+All benchmarks were performed on a 2.5 GHz Intel Core i7 processor with 16 GB of
+1600 MHz DDR3 running macOS High Sierra.
+
+                                -------------
+
+     - L909 and L956: About the benchmark suite, how many of the benchmarks
        in fact only need bijective lenses? This definitely should be
@@
 
-L922: The authors may add a few words on how representative and
+35 of the benchmarks only need bijective lenses (4 of the original Optician
+benchmarks had to be altered to be bijective because of projected data, in the
+symmetric benchmark suite we removed such alterations.) Of the 9 benchmarks that
+required the termination parameter, 4 of them required only bijective lenses.
+
+                                -------------
+
+     - L922: The authors may add a few words on how representative and
        complex the selected tasks is. For example, I got the data from the
@@
 
-L926 and L952: This is the most confusing point for me. The meaning of
+We chose our benchmarks from the FlashFill paper, because the 238 tasks are,
+unfortunately, not publicly available. In general, our benchmark suite goes
+beyond the level of difficulty required for data cleaning. Many real-life file
+format descriptors involve large amounts of disjunctions and nested iterations,
+which FlashFill is either bad at (for disjunctions) or cannot do (nested
+iterations).
+
+A more thorough comparison to FlashFill is present in the previous Synthesizing
+Bijective Lenses paper.
+
+                                -------------
+
+     - L926 and L952: This is the most confusing point for me. The meaning of
        "correct lenses" here seems very subjective as the authors let the
@@
 
-L945 and L993 (Fig.6 and Fig.7): Since the time is only within 30
+We respond to this in detail in the main response.
+
+                                -------------
+
+     - L945 and L993 (Fig.6 and Fig.7): Since the time is only within 30
        seconds, why not make Figures whose x-axes are linear? With the
@@
 
-L1055: Again, if the default (termination) parameter can only find
-bijective lenses, it is not that good.
+We chose a logarithmic x axis to show the differences on the fast benchmarks, as
+they otherwise would be imperceptable. However, if the impact on comparing the
+slower benchmarks is too significant, we can change to a linear x axis.
 
-L1066: When discussing symmetric lenses, the authors need to give a
+Our graph does not include any specific order of tasks. At any given time, a
+reader can see how many of the benchmarks are able to complete within that time
+(for example, 41 benchmarks require complete in under a second in the full
+synthesis algorithm). We can reword the description of these graphs to make this
+more clear.
+
+                                -------------
+
+     - L1066: When discussing symmetric lenses, the authors need to give a
        reference to Symmetric Lenses by Martin Hofmann et al. (2011) since
@@
 
-L1080: The authors did not explain the notation `l ∈ X ⟷ Y` as it
-hides the complement `C`.
-
-L1105 and below: Similarly, the authors need to explain the inference
-rules and the involved symbols more. For instance, what is `xyo`?
-There is even no note telling the reader that some explanation can be
+We will update the lens section of the related work to be more thorough on the
+related work that we do not make formal statements about, and include proper
+citations.
 found in the appendix.
 
-L1111 and L1112: I cannot figure out where does `x` below the
+                                -------------
+
+     - L1111 and L1112: I cannot figure out where does `x` below the
        "horizontal line" come from
 
-L1170: I do believe information-theoretic analysis is of extremely
-useful; if possible, adding one more example will make the argument
-stronger.
-
-##Things that hinder reading
-
-L262: this is the first time that "forward direction"
-appears. Besides, for symmetric lenses, it is unclear what is the
-forward direction.
-
-L364: `... within emp_salaries ...`. According to the context, this
-`emp_salaries` is supposed to to be found in Figure 3 and there
-happens to be a lens type called `emp_salaries`. But in fact the
-`emp_salaries` should refer to the one in Figure 2.
-
-L659: The notation (DS,DT) might need some explanation. It is not a
-tuple but just introduce two preferable names DS and DT.
-
-L1101 and L 1103: the authors use "edit `esi`" here but use "edits
-`esi`" at L1082. Is `esi` a single edit or a sequence of edits? Having
-read it many times, I eventually understand that it should be the
-single edit that we focus on; so that `esi = inl x`.
-
-L1509 and L1515: the two lines exceed page limit. Luckily the exceeded
-parts can be guessed.
-
-## minors:
-
-L14: that -> than
-
-L54: Indeed of -> Indeed among
-
-L220: shoren -> shorten
-
-L372: between -> among
-
-L414 and L415: some words are missing in the sentence so that it is
-not easy to understand.
-
-L426: a means to reason -> a means of reasoning
-
-L431: present -> presented ?
-
-L438: the authors had better to put the definition of entropy
-somewhere earlier in this sentence. Otherwise it syntactically looks
-like the definition of data source S.
-
-L451: a redundant ",s"
-
-L453: "in subtle ways, the algorithm" -> "in subtle ways and the algorithm"
-
-L459: I do not understand the example. Which law gives the result?
-
-L542: combonator -> combinator
-
-L547: equivalent -> is equivalent
-
-L553: does `H->(l,s | T)` looks better than the original `H->(T | l,s)`? There is no reason to use the same argument order for `H->` and `H<-` since we can think them as different functions.
-
-L620: I cannot parse the sentence
-
-L656: SDNF REs are SDNF REs with probability annotations -> SDNF REs
-are DNF REs with probability annotations.
-
-BTW, the authors had introduced the abbreviation SRE in the
-introduction. But in this subsection the authors use too many
-different abbreviations:
-
-```
-SDNF REs
-stochastic DNF REs
-stochastic DNF regular expressions
-
-SREs
-stochastic REs
-stochastic regular expressions
-
-REs
-regular expressions
-```
-Please consider use the same notation throughout.
-
-L658: `DS bar` is introduced but seems to be never used.
-
-L715: expression., -> expression,
-
-L734: than there are DNF bijective lenses -> than DNF bijective lenses
-
-L845: uncoverted -> unconverted
-
-L928: whether it is usable to generate synthesize correct lenses ->
-whether it is usable to synthesize correct lenses
-
-L929: I prefer to remove the "the" in the first sentence.
-
-L971: measuring it shows of the responsiveness -> measuring it shows
-the responsiveness
-
-L1080 ~ L1082: the description of the function apply is its curried
-version while the usage below is the uncurried version.
-
-L1082: edits `esi` -> edit `esi`
-
-L1099: I guess the title should be "Relation with Symmetric Lenses"
-
-L1099: classic -> classical
-
-L1136: `l.put s v` -> `l.put v s`
-
-L1138 `l.put s (l.get s)` -> `l.put (l.get s) s`
-
-L1512: I found `(y,c2') = l.putR (x2', c1')`, do you mean `(y,c2') =
-l.putR (x2', c2')`
-
-L1526 and L1530: sometimes x and y are used and sometimes s and v.
-
-L1702: this section is not FORGETFUL SYMMETRIC LENSES (which is the
-previous section)
-
-L2088: perhaps the theorem about lenses should be put to the previous
-section. This section is about SRE and information theory.
-
-L2098: this is CREATEGET, not PUTGET again
-
-L2112: `l.get` -> `l.get s`
-
-Questions for authors’ response
----------------------------------
-Please clarify the evaluation part.
+This is true for any arbitrary "x"
 
@@
 
-Paper summary
                                 -------------
-- This paper proposes a technique for synthesizing a form of lenses from input-output examples and types. 
-- Lenses are transformations from one data type to another. They are typically used to transform between data formats, e.g., CSV to JSON.
-- Previous work on synthesizing lenses covered the bijective case, where the synthesized lens is not lossy.
-- This paper builds on top of prior work to synthesize lossy lenses, which they call simple symmetric lenses.
-These lenses allow data transformations that lose data, e.g., extracting one column from a table.
-- Compared to bijective lenses, the search space is less constrained.
-The observation of previous work was that the bijectivity constraint is so restrictive that any lens synthesized is likely to be the right one.
-By removing this constraint, the search space is much larger.
-- To work around that restriction, the authors propose to use probabilistic types, which define a probability distribution over inhabitants of the type.
-Then, they propose the heuristic where they look for lenses that minimize entropy in the transformation — roughly the number of bits needed to recover one format from the other.
-- The synthesis algorithm largely follows that from POPL18 on bijective lenses — types are defined via regular expressions, regular expressions are unrolled (expanded) into DNF form, and then a greedy search looks for a lens for that pair of types.
-- The experiments demonstrate the usefulness of the search heuristics used—particularly using entropy to guide the search.
 
-Strengths
----------
-+ Interesting application of synthesis
-+ Elegant set of combinators defining simple symmetric lenses
-+ Intriguing information-theoretic search heuristic
+    - The usability aspect is questionable: It seems that if the user has 
+      defined complex regular expressions defining the pair of languages, then
+      they can probably go ahead and just implement the lens without resorting 
+      to examples.
 
-Weaknesses
-----------
-— Small algorithmic delta over POPL18
+As a demonstration about the difficulty of writing lenses, consider the
+following bidirectional program
 
-— Unclear whether stochastic REs are needed
+let single_author_convert =
+  ins " au - "
+    . lens_swap
+        (NAME . del ",")
+        (lens_swap WSP NAME)*
+in
 
-Comments for author
--------------------
-I enjoyed reading this paper.
-I think the technique is elegant and well-reasoned.
-The usability aspect is questionable: It seems that if the user has defined complex regular expressions defining the pair of languages, then they can probably go ahead and just implement the lens without resorting to examples.
+Even the relatively simple lens that permutes the 3 elements needed for
+single_author_convert is fairly complex.  The complexity becomes even more
+apparent when working with large lenses comprised of many sub-lenses (though
+for brevity we give here the more minimal example above).  Lenses provide
+great power with their invertibility guarantees, but they come at the cost
+of thinking about many fiddly details when writing the terms.  Worrying
+about these fiddly details while ALSO thinking about unambiguity
+restrictions is a very difficult task!
 
-While the paper begins with a nice high-level illustration and examples, the technical sections lack illustrative examples and motivation that it’s hard to appreciate some points.
-For example, Section 4.1 introduces the lens cost computation, but no example of such computation is given.
-More important: It is unclear to the general reader why the Expand procedure is needed — i.e., why do we need to rewrite types for GreedySynth.
+Engineering best practices dictate that we annotate the lenses with their types.
+While Boomerang does not require these annotations - being able to infer the
+types from the terms - the types of the term serve as documentation for future
+programmers to understand what formats the lens maps between. Furthermore, the
+types provide resilience in the face of future lens modifications, as the types
+ensure that the lens maps exactly between strings of the provided formats.
+Invalid inputs are not accepted nor are the outputs ever invalid. Consequently,
+in a well-maintained codebase, we are not requiring additional work. Instead, we
+are saving programmers from having to writing the lens by leveraging work they
+should already be performing.
 
-The ToStochastic procedure felt quite ad hoc to me.
-The paper makes a point that if the user knew the distribution (e.g., if they automatically infer the parameters from data), then they can supply it.
-But will that really help the search?
-Essentially, I wish there were examples (even synthetic ones) where knowing the distribution of data can clearly influence the synthesis procedure.
-The paper resorts to a fixed distribution that gives the same probability to each disjunct.
-This raises two concerns:
+                                -------------
+
+    - More important: It is unclear to the general reader why the Expand
+      procedure is needed — i.e., why do we need to rewrite types for
+      GreedySynth.
+
+We will make this more clear in future versions of the paper. Expand is needed
+because GreedySynth alone cannot traverse regular expression equivalences. If
+given the two regular expressions "" | "a"+, and "a"\*, GreedySynth alone would
+merge "" to "a"\* (with a disconect), and merge "a"+ (also with a disconect).
+However, contuing search through equivalent REs helps find the bijective lens.
+
+                                -------------
+
+    - The ToStochastic procedure felt quite ad hoc to me. The paper makes a 
+      point that if the user knew the distribution (e.g., if they automatically
+      infer the parameters from data), then they can supply it. But will that
+      really help the search? Essentially, I wish there were examples (even 
+      synthetic ones) where knowing the distribution of data can clearly 
+      influence the synthesis procedure.
+
+Consider the situation where the input is two regular expressions:
+"a"+ | "b"+ <=> "c" | "d"
+
+There are two lenses that are both require projecting the same iteration that
+satisfies this spec, "a"+ can go to "c", or "a"+ can go to "d" (and "b" goes to
+the other).
+
+However, if both "a"+ and "c" are high probability, then they would become
+aligned first, as it is more important they get low cost lenses, as they occur
+more often. This is desired, as in synchronized formats, we expect synchronized
+disjuncts to have similar probabilities.
+
+                                -------------
+
+    - The paper resorts to a fixed 
+      distribution that gives the same probability to each disjunct. This raises
+      two concerns:
 
@@
 
-Why not use a syntactic heuristic.
-That’s what DC is trying to do, but perhaps a slightly more sophisticated syntactic heuristic could achieve the same effect.
-So, while the use of entropy is cool, appealing, and new in synthesis (to my knowledge), the paper does not convince me that it is needed here.
+      Why not use a syntactic heuristic. That’s what DC is trying to do, but 
+      perhaps a slightly more sophisticated syntactic heuristic could achieve 
+      the same effect. So, while the use of entropy is cool, appealing, and new 
+      in synthesis (to my knowledge), the paper does not convince me that it is 
+      needed here.
 
-Section 6 or 7 should discuss examples that are outside the expressivity of the lens language and/or take too much time. It was hard for me to infer what the limitations of the approach are.
+We respond to these issues in the main reponse.
 
-Questions for authors’ response
----------------------------------
-My main concern is whether the use of stochastic REs is really useful (see above)
+                                -------------
 
-Are stochastic REs new? I didn't see any citations when they were introduced.
+    - Section 6 or 7 should discuss examples that are outside the expressivity 
+      of the lens language and/or take too much time. It was hard for me to 
+      infer what the limitations of the approach are.
+
+We agree that section 7 requires more, informal comparisons to other lenses, for
+us to discuss comparisons between our framework and other types of lenses. In
+general, for different types of lenses to work, we require additional work,
+either in the synthesis algorithm (for different types of combinators like
+matching lenses or delta lenses), or in both the synthesis algorithm and
+specifications (like for edit lenses and update lenses).
+
+                                -------------
+
+    - My main concern is whether the use of stochastic REs is really useful (see
+      above)
+
+Stochastic REs, and more generally, a semantic cost metric is critical for
+comparing between responses by GreedySynth.
+
+                                -------------
+
+    - Are stochastic REs new? I didn't see any citations when they were 
+      introduced.
+
+Stochastic REs are not new. We cite them when we first mention them on page 2,
+but agree we should provide additional citations when they are formally
+introduced. We do not mean to imply we are introducing stochastic regular
+expressions, though we are introducing syntactic means to find their entropy.

### 2018-09-18 papers/SymmetricLenses/response_popl19.md
make no merge conflicts

@@
 exist, and we then must take it upon ourselves to recognize which components of
-the file should be synchronized to which, and validate that the lenses
-corresponded to our internal notion of correctness.
+the file should be synchronized to which.
 
 In validating that the synthesized lenses were correct, we applied the types of
-approaches we use in validating that our code is correct in everyday
-development.
+approaches we use in validating our code is correct in everyday development.
 1.  We manually inspected the code. This is certainly a nontrivial task, as

### 2018-09-19 papers/SymmetricLenses/response_popl19.md
Polishing and comments on the response

@@
+BCP: The first section is too long -- 500 words max.
+
 Thank you for your comments!
@@
 We begin with brief responses to the most significant concerns, followed by a
-(long and optional) appendix addressing individual questions.
+(long, optional) appendix addressing individual questions point by point.
 
@@
 down to below the fold *) 
+BCP: Yes, maybe so.  Though I wish we could leave at least a brief remark
+here, since other reviewers may be wondering now.
 
@@
 
-We did not intend to forget such citations or claim that we are introducing
-stochastic regular expressions. We will be sure, in future versions, to make
-this distinction more clear.
+Yes, we will make this clearer on p. 3 too.
 
@@
 
+BCP: ??
+
                                 -------------
@@
 
-This is correct
+This is correct -- we will clarify.
 
@@
 
-Yes, one of our benchmarks is such an example. In particular, consider trying to
+BCP: I'm not sure I understand the question!
+
+One of our benchmarks is such an example. Consider trying to
 synchronize Windows "Scheduled Tasks" and Linux "cron jobs." In such a task, we
-wish to synchronize the times that jobs are run, but not synchronize the
-commands to run the jobs themselves. In such a scenario, we could create a
+might wish to synchronize the times that jobs are run, but not synchronize the
+commands to run the jobs themselves. We could create a
 bijective lens, but this is not correct behavior. Instead, examples can
@@
 
-Full details are present in the prior work on Synthesizing Bijective Lenses
-(page 16). From a high level perspective, we process both regular expression
-trees, looking for situations where a regular expression is as a subcomponent of
-one format, but not the other, and expand such REs (this is where hashconsing
-provides its greatest benefits). Furthermore, we look for instances where a
-regular expressions may be present in one format, but is only contained within a
-closed regular expressions in the other, and expand the closed RE containing the
-missing RE. We then make this apply in slightly more places by also taking into
-account how deeply nested within stars certain REs occur.
+At a high level, we process both regular expression trees, looking for
+situations where a regular expression is as a subcomponent of one format but
+not the other, and expand such REs (this is where hash-consing provides its
+greatest benefits). Furthermore, we look for instances where a regular
+expression may be present in one format, but is only contained within a
+closed regular expression in the other, and we expand the closed RE
+containing the missing RE in this case. We then make this apply in slightly
+more places by also taking into account how deeply nested within stars
+certain REs occur.
+
+BCP: What is "closed"?
+
+Full details can be found in the prior work on Synthesizing Bijective Lenses
+(page 16).
 
@@
 
-Correct, this should be max
+Right, this should be max!
 
@@
 
-Yes, I do expect that the algorithm can be extended to handle such variants,
-though likely it would require some additional work.
+Yes, we do have some hope that the algorithm can be extended to handle such
+variants.
 
-Delta lenses would require additional work in understanding "shape" constructs.
-Currently our language only permits basic iteration, but shapes allow for
-different types of mergings. This would require changes to AtomSynth, in that we
-would we would need to discover the correct shapes of the data, and align them
-in the correct ways.
+Delta lenses would require additional work in understanding "shape"
+constructs.  Currently our language only permits basic iteration, whereas
+shapes allow for different types of mergings. Allowing this would require
+changes to the AtomSynth algorithm, in that we would we would need to
+discover the correct shapes of the data, and align them in the correct ways.
 
-Both edit and update lenses would require new notions of specifications. Both
-lens types permit updates that are not merely pushing data through, but instead
-are applying specific elements of an edit monoid. We expect that our current IO
-approach to specifications would not work as we wished for such transformations.
-(Also we would remove complements from edit lenses).
+Both edit and update lenses would require new forms of specifications, since
+both permit updates that are not merely pushing data through, but instead
+are applying specific elements of an edit monoid.  (Also we would remove
+complements from edit lenses).
 
@@
 Review #43B
-
-                                -------------
 
@@
 
-All benchmarks were performed on a 2.5 GHz Intel Core i7 processor with 16 GB of
-1600 MHz DDR3 running macOS High Sierra. We provide a more detailed answer our
-benchmark suite and method of finding correctness in the main portion of the
-paper.
+We responded to the more substantive points here (the benchmark suite and
+judging correctness) "above the fold" at the top of this response.
+
+All benchmarks were performed on a 2.5 GHz Intel Core i7 processor with 16
+GB of 1600 MHz DDR3 running macOS High Sierra.
 
@@
 
-We apologize for the lack of citations and messy appendix. For the first
-subsection of Related Work, we got carried away in making formal claims, and did
-not include information comparisons to other lens languages. We will make sure
-to make such comparisons and citations in the final version of the paper. We
-will similarly improve the appendix.
+We apologize for the lack of citations and the messy appendix. For the first
+subsection of Related Work, we got carried away in making formal claims, and
+did not include information comparisons to other lens languages. We will
+make sure to make such comparisons and citations in the next version of the
+paper. We will similarly improve the appendix.
+
+BCP: information comparisons?
 
@@
 Thank you, you are correct, we will change this citation.
+
+BCP: We should push the final version of this paper out to Arxiv.
 
@@
 
-Within our synthesis algorithm, you are, in general, correct. However, Boomerang
-exists beyond Optician, and permits lenses like id("a"|"a"). In our extension to
-Boomerang, we do not remove such freedom, and continue to permit such languages.
+Within our synthesis algorithm, you are, in general, correct. However,
+Boomerang exists beyond Optician, and permits lenses like id("a"|"a"). In
+our extension to Boomerang, we do not remove this freedom.
 
@@
 
-We chose our benchmarks from the FlashFill paper, because the 238 tasks are,
+We chose our benchmarks from the FlashFill paper because the 238 tasks are,
 unfortunately, not publicly available. In general, our benchmark suite goes
-beyond the level of difficulty required for data cleaning. Many real-life file
-format descriptors involve large amounts of disjunctions and nested iterations,
-which FlashFill is either bad at (for disjunctions) or cannot do (nested
-iterations).
+beyond the level of difficulty required for data cleaning. Many real-life
+file format descriptors involve large numbers of disjunctions and nested
+iterations, which FlashFill is either bad at (for disjunctions) or cannot
+handle at all (nested iterations).
 
@@
 
-We respond to this in detail in the main response.
+We responded to this in detail above the fold.
 
@@
 
-                                -------------
-
     - The usability aspect is questionable: It seems that if the user has 
@@
 
-Engineering best practices dictate that we annotate the lenses with their types.
-While Boomerang does not require these annotations - being able to infer the
-types from the terms - the types of the term serve as documentation for future
-programmers to understand what formats the lens maps between. Furthermore, the
-types provide resilience in the face of future lens modifications, as the types
-ensure that the lens maps exactly between strings of the provided formats.
-Invalid inputs are not accepted nor are the outputs ever invalid. Consequently,
-in a well-maintained codebase, we are not requiring additional work. Instead, we
-are saving programmers from having to writing the lens by leveraging work they
-should already be performing.
+Engineering best practices dictate that we annotate the lenses with their
+types.  While Boomerang does not require these annotations - being able to
+infer the types from the terms - the types of the term serve as
+documentation for future programmers to understand what formats the lens
+maps between. Furthermore, the types provide resilience in the face of
+future lens modifications, as the types ensure that the lens maps exactly
+between strings of the provided formats.  Invalid inputs are not accepted
+nor are the outputs ever invalid. Consequently, in a well-maintained
+codebase, we are not requiring additional work. Instead, we are saving
+programmers from having to writing the lens by leveraging work they should
+already be performing.
 
@@
 
-We will make this more clear in future versions of the paper. Expand is needed
-because GreedySynth alone cannot traverse regular expression equivalences. If
-given the two regular expressions "" | "a"+, and "a"\*, GreedySynth alone would
-merge "" to "a"\* (with a disconect), and merge "a"+ (also with a disconect).
-However, contuing search through equivalent REs helps find the bijective lens.
+We will make this more clear in future versions of the paper. Expand is
+needed because GreedySynth alone cannot traverse regular expression
+equivalences. If given the two regular expressions "" | "a"+, and "a"\*,
+GreedySynth alone would merge "" to "a"\* (with a disconect), and merge "a"+
+(also with a disconect).  However, contuing search through equivalent REs
+helps find the bijective lens.
 
@@
 
-Stochastic REs are not new. We cite them when we first mention them on page 2,
-but agree we should provide additional citations when they are formally
-introduced. We do not mean to imply we are introducing stochastic regular
-expressions, though we are introducing syntactic means to find their entropy.
+Stochastic REs are not new. We cite them when we first mention them on page
+2, but agree we should provide additional citations when they are formally
+introduced. We do believe our syntactic technique for calculating their
+entropy is new.
+
+BCP: Rephased the last sentence -- is it right?

### 2018-09-19 papers/SymmetricLenses/response_popl19.md
edits

--- 2018-09-18 papers/SymmetricLenses/response_popl19.md
@@
 
-Determining what a correct lens is quite a difficult problem. What a "correct"
-lens is inherently subjective -- determining which components of one format
-should be synchronized to another requires knowledge of the formats. If there
-were a example large suite of synchronized files of these formats, we could use
-this example suite as a ground truth. Unfortunately, such a suite does not
-exist, and we then must take it upon ourselves to recognize which components of
-the file should be synchronized to which.
+Beyond the format specifications and examples, we do not have
+specifications that define the conversion functions.  Hence, we do not
+have an automated way to measure the correctness of our lenses (beyond
+checking that each synthesized lens does indeed transform the given examples
+correctly).  In this way, our work is similar to other published
+synthesis-from-examples research, such as Gulwani's work on FlashFill and related projects.
 
-In validating that the synthesized lenses were correct, we applied the types of
-approaches we use in validating our code is correct in everyday development.
-1.  We manually inspected the code. This is certainly a nontrivial task, as
-    these generated lenses are quite large. However, this task is easier than it
-    may initially sound. Much of lenses maintain a rigid structure:
-    Concatenations in the lenses typically follow concatenations in the regular
-    expressions, disjunctions in the lenses typically follow disjunctions in the
-    regular expressions. The primary difficulties arise when seeing how the more
-    complex combinators interact. Are the swaps in the right place? Are the
-    correct components disconnected? Are lenses merged in the correct way? While
-    validation certainly still requires work, instead of closely inspecting
-    hundreds of lines of code, we instead must closely inspect a few dozen, and
-    give a more cursory glance to the others.
-2.  We built unit tests to validate the more complex lenses. Much like everyday
-    development, manual inspection of code is rarely sufficient for determining
-    correctness. In addition to reading through the code, we added in unit tests
-    to validate that the generated code performs correctly on representative
-    examples.
-3.  Lastly, in addition to unit tests, we ran the lenses on a number of strings,
-    taken from the unit tests and the examples, but with minor alterations.
-    These quick tests we did not add as unit tests, but served as additional
+To check the correctness of our tool on our benchmarks, we did what
+programmers usually do:
+
+1.  We manually inspected the code. This was certainly a nontrivial task, as
+    these generated lenses are quite large, but perhaps easier than it
+    may initially sound. Some of the lenses' structure follows from the
+	structure of the regular expressions that describe source and
+    target format --- this is relatively easy to validate. However, it
+    can be more difficult to check the interactions of the complex
+    combinators: Are the swaps in the right place? Are the
+    correct components disconnected? Are lenses merged in the correct
+    way?  These complex combinators often occupied a few dozen lines
+    of code rather than a few hundred, making our task easier.   (Certainly, checking
+    the code is easier than creating it de novo.)
+2.  We built a few unit tests to validate the more complex lenses. 
+3.  We ran the lenses on a number of strings, taken from the unit
+    tests and the examples, but with minor alterations.
+    We did not add these tests as additional unit tests, but they served as additional
     validation for ourselves as we were constructing the benchmark suite.

### 2018-09-19 papers/SymmetricLenses/response_popl19.md
edits

@@
 
-     - Is our synthesis procedure overfit to our benchmarks? Why does assigning
-       a fixed probability work?
+      - Why does assigning a fixed probability work?
 
-Our 39 bijective benchmarks came from the original "Synthesizing Bijective
-Lenses" benchmark suite without alteration to the regular expressions. Our
-additional examples, in particular the data cleaning tasks, were taken from the
-original FlashFill paper. Unfortunately, there is no public release of the
-FlashFill benchmarks, so our examples had to be taken from the paper, instead of
-from a larger benchmark suite.
+We assign uniform distributions to all of the components to avoid
+preferring one conjunct or disjunct over another.  We do this at the
+moment because we have little additional information and our
+experiments demonstrate that it is quite effective at the moment
+(more effective than simple heuristics such as preferring constants.
+On the other hand, we believe that if we learned a distribution from
+example data, we might be able to achieve even better results.  As an
+example, consider trying to learn a lens with this type:
 
-Assigning a fixed probability is certainly less desirable than inferring it from
-data. With a learned distribution, we expect our algorithm to perform better: it
-can recognize that certain disjunctions both have high probability, and would
-strongly incentivize aligning those critical disjuncts to the critical disjuncts
-in the other format. With a learned distribution, we expect that tasks that
-currently require 2-3 examples would instead require 1-2. However, this is not
-to say that our tool will perform badly as-is on distributions that are far from
-our computed ones. On such distributions, our tool will merely not get the
-benefits provided by such distributions. It would get an advantage in having an
-idea beforehand which components should map to which with a learned
-distribution, but merely loses on that advantage without it.
+"a"+ | "b"+ <=> "c" | "d"
 
-(* This paragraph is to respond to a complaint only of R3, maybe should move
-down to below the fold *) 
-BCP: Yes, maybe so.  Though I wish we could leave at least a brief remark
-here, since other reviewers may be wondering now.
+There are (at least) two lenses of similar quality:
 
-This is not to say that SREs and our semantic metric are not critical for
-success. While we use a learned distribution, using such a semantic metric
-provides a means of comparison across proposed pairs. With a syntactic metric,
-we aren't making an "apples-to-apples" comparison, one lens might have a higher
-score, but that score would be based on a different distribution of data - it
-actually would be less likely with a different data distribution. Keeping the
-probabilities of the two formats is critical for comparing between proposed
-types.
+"a"+ can go to "c", or "a"+ can go to "d"
+
+(and "b" goes to the other).
+
+With a fixed weighting, both lenses would be equivalent.  However, by
+analyzing example data, we might find that both
+"a"+ and "c" are high probability. Such a distribution would suggest
+that "a"+ be transformed into "c" (and information-theoretically, such
+a lens would have a better score).    If such weights were inputs to the
+system, our synthesis algorithm would generate a lens that transformed
+"a"+ into "c" rather than "d".
 

### 2018-09-19 papers/SymmetricLenses/response_popl19.md
fixed minor typos

@@
 At a high level, we process both regular expression trees, looking for
-situations where a regular expression is as a subcomponent of one format but
+situations where a regular expression is a subcomponent of one format but
 not the other, and expand such REs (this is where hash-consing provides its
@@
 
-We responded to the more substantive points here (the benchmark suite and
+We responded to the key points in this comment (the benchmark suite and
 judging correctness) "above the fold" at the top of this response.
@@
 citations.
-found in the appendix.
 
@@
         (lens_swap WSP NAME)*
-in
+in ...
 

### 2018-09-19 papers/SymmetricLenses/response_popl19.md
do some clean

--- 2018-09-19 papers/SymmetricLenses/response_popl19.md
@@
 
-Beyond the format specifications and examples, we do not have
-specifications that define the conversion functions.  Hence, we do not
-have an automated way to measure the correctness of our lenses (beyond
-checking that each synthesized lens does indeed transform the given examples
-correctly).  In this way, our work is similar to other published
-synthesis-from-examples research, such as Gulwani's work on FlashFill and related projects.
+Beyond the format specifications and examples, we do not have specifications
+that define the conversion functions. Hence, we do not have an automated way to
+measure the correctness of our lenses (beyond checking that each synthesized
+lens does indeed transform the given examples correctly). In this way, our work
+is similar to other published synthesis-from-examples research, such as
+Gulwani's work on FlashFill and related projects.
 
@@
 1.  We manually inspected the code. This was certainly a nontrivial task, as
-    these generated lenses are quite large, but perhaps easier than it
-    may initially sound. Some of the lenses' structure follows from the
-	structure of the regular expressions that describe source and
-    target format --- this is relatively easy to validate. However, it
-    can be more difficult to check the interactions of the complex
-    combinators: Are the swaps in the right place? Are the
-    correct components disconnected? Are lenses merged in the correct
-    way?  These complex combinators often occupied a few dozen lines
-    of code rather than a few hundred, making our task easier.   (Certainly, checking
-    the code is easier than creating it de novo.)
+    these generated lenses are quite large, but perhaps easier than it may
+    initially sound. Some of the lenses' structure follows from the structure of
+    the regular expressions that describe source and target format --- this is
+    relatively easy to validate. However, it can be more difficult to check the
+    interactions of the complex combinators: Are the swaps in the right place?
+    Are the correct components disconnected? Are lenses merged in the correct
+    way? These complex combinators often occupied a few dozen lines of code
+    rather than a few hundred, making our task easier. (Certainly, checking the
+    code is easier than creating it de novo.)
 2.  We built a few unit tests to validate the more complex lenses. 
-3.  We ran the lenses on a number of strings, taken from the unit
-    tests and the examples, but with minor alterations.
-    We did not add these tests as additional unit tests, but they served as additional
-    validation for ourselves as we were constructing the benchmark suite.
+3.  We ran the lenses on a number of strings, taken from the unit tests and the
+    examples, but with minor alterations. We did not add these tests as
+    additional unit tests, but they served as additional validation for
+    ourselves as we were constructing the benchmark suite.
 
@@
 
-      - Why does assigning a fixed probability work?
+      - There were a number of related questions about our information theoretic
+        measure: Is it useful? Why not use a syntactic metric? Why does a
+        fixed distribution work?  Will knowing the data distribution help?
 
-We assign uniform distributions to all of the components to avoid
-preferring one conjunct or disjunct over another.  We do this at the
-moment because we have little additional information and our
-experiments demonstrate that it is quite effective at the moment
-(more effective than simple heuristics such as preferring constants.
-On the other hand, we believe that if we learned a distribution from
-example data, we might be able to achieve even better results.  As an
-example, consider trying to learn a lens with this type:
+Having a guiding principle behind these heuristics is, in itself, useful. Ad-hoc
+syntactic metrics that are present in other systems (and we claim: are typically
+building towards our underlying principle). Finding a simple underlying
+principle (be as bijective as possible, which we formalize via information
+theory) is better than a metric that simply applies only to our domain. We claim
+that information theory is a general-purpose way to group all these metrics
+where people give higher cost to their "less bijective" combinators (avoiding
+constants, for example).
 
-"a"+ | "b"+ <=> "c" | "d"
+A fixed distribution works because, the core thing we are aiming for is "more
+bijective." A fixed distribution is merely a way to not unfairly weight one
+conjunct or disjunct over another in the presence of not knowing other
+information. Even with our inferred distribution, we are still aiming towards
+"more bijective".
 
-There are (at least) two lenses of similar quality:
-
-"a"+ can go to "c", or "a"+ can go to "d"
-
-(and "b" goes to the other).
-
-With a fixed weighting, both lenses would be equivalent.  However, by
-analyzing example data, we might find that both
-"a"+ and "c" are high probability. Such a distribution would suggest
-that "a"+ be transformed into "c" (and information-theoretically, such
-a lens would have a better score).    If such weights were inputs to the
-system, our synthesis algorithm would generate a lens that transformed
-"a"+ into "c" rather than "d".
+Knowing the distribution will help in certain scenarios, and hinder in others.
+With a learned distribution, our information theoretic measure is closer to the
+ground-truth, permitting our learned lens to, in fact, be "more bijective."
+However, we recognize that, while we believe "more bijective" is a solid guiding
+principle, there are certain situations where the most bijective lens is not the
+desired one: this is why we must further use examples for specifications. We
+could certainly imagine situations where a learned distribution would find the
+wrong lens (without additional examples), and synthesizing with the fixed
+distribution would find the correct one, and vice-versa. However, we claim that,
+as we typically want "more bijective," and a learned distribution gets closer to
+a ground-truth on that, a learned distribution likely would perform better.
 
@@
 BCP: ?? 
+AFM: In hindsight, I think I will just delete this question - it's more
+of a "finding-errata" question than an actual one
 
@@
 
+Within a DNF lens itself, all sequences are *required* to be involved with
+another sequence lens: otherwise when that side of the DNF SRE is provided -
+what disjunct on the other side should that lens be mapped to? Of course, there
+are other ways the disjunct can be mapped: but these must happen at a different
+level of the synthesis algorithm. For example - the entirety of that DNF regular
+expression (which itself could be deeply nested within other DNF regular
+expression components) could be projected. Furthermore, a specific sequence
+could be mapped to another, but all the components of that sequence lens
+actually may be disconnects. However, all sequences *must* be involved in a
+sequence lens in a well-typed DNF lens.
+
 BCP: I'm not sure I understand the question! 
-
-One of our benchmarks is such an example. Consider trying to
-synchronize Windows "Scheduled Tasks" and Linux "cron jobs." In such a task, we
-might wish to synchronize the times that jobs are run, but not synchronize the
-commands to run the jobs themselves. We could create a
-bijective lens, but this is not correct behavior. Instead, examples can
-demonstrate that these components should remain unsynchronized.
+AFM: I also now realize I didn't. After rereading the section, I think this
+response is more appropriate.
 
@@
 BCP: What is "closed"?
+AFM: The internals of the Regular Expression are not exposed to GreedySynth,
+they are the analogue of User Defined regular expressions in the previous paper,
+but with this closed/open approach, we aren't dependent on where users place
+variables (also making our integration into Boomerang significantly less hacky).
 
@@
 We apologize for the lack of citations and the messy appendix. For the first
-subsection of Related Work, we got carried away in making formal claims, and
-did not include information comparisons to other lens languages. We will
-make sure to make such comparisons and citations in the next version of the
-paper. We will similarly improve the appendix.
-
-BCP: information comparisons?
+subsection of Related Work, we got carried away in making formal claims, and did
+not include informal comparisons to other lens languages. We will make sure to
+make such comparisons and citations in the next version of the paper. We will
+similarly improve the appendix.
 
@@
 Thank you, you are correct, we will change this citation.
-
-BCP: We should push the final version of this paper out to Arxiv.
 
@@
 
-Consider the situation where the input is two regular expressions:
-"a"+ | "b"+ <=> "c" | "d"
+We respond to this primarily above.  For a specific example:
 
-There are two lenses that are both require projecting the same iteration that
-satisfies this spec, "a"+ can go to "c", or "a"+ can go to "d" (and "b" goes to
-the other).
+Consider finding a lens of type:
+("a" |(.5) "b") |(.5) "c" <=> "x" |(.1) "y"
 
-However, if both "a"+ and "c" are high probability, then they would become
-aligned first, as it is more important they get low cost lenses, as they occur
-more often. This is desired, as in synchronized formats, we expect synchronized
-disjuncts to have similar probabilities.
+In this situation, the lowest cost lens would merge "a" and "b" into "x", and
+keep "c" mapped to "y".
+
+Consider finding a lens of type:
+("a" |(.5) "b") |(.5) "c" <=> "x" |(.9) "y"
+
+In this situation, the lowest cost lens would merge "a" and "b" into "y", and
+keep "c" mapped to "x".
+
+We are happy to provide an example such as this one in the paper.
 
@@
 Stochastic REs, and more generally, a semantic cost metric is critical for
-comparing between responses by GreedySynth.
+comparing between responses by GreedySynth. Without this semantic cost, the same
+function would likely have different costs if found through different calls to
+GreedySynth. Syntactic metrics, we feel, aim to approximate how good the
+underlying result is: in this work we are able to find a semantic metric that we
+can be found using syntactic information.
 
@@
 entropy is new.
-
-BCP: Rephased the last sentence -- is it right?

### 2018-09-20 papers/SymmetricLenses/response_popl19.md
some wordsmithing

@@
     code is easier than creating it de novo.)
-2.  We built a few unit tests to validate the more complex lenses. 
+2.  We built unit tests to validate the more complex lenses. 
 3.  We ran the lenses on a number of strings, taken from the unit tests and the
@@
     ourselves as we were constructing the benchmark suite.
+
+KSF: Does 3 add anything??  We could delete it and save some words.
 
@@
 
-Having a guiding principle behind these heuristics is, in itself, useful. Ad-hoc
-syntactic metrics that are present in other systems (and we claim: are typically
-building towards our underlying principle). Finding a simple underlying
-principle (be as bijective as possible, which we formalize via information
-theory) is better than a metric that simply applies only to our domain. We claim
-that information theory is a general-purpose way to group all these metrics
-where people give higher cost to their "less bijective" combinators (avoiding
-constants, for example).
+Having a principle guiding our heuristics is, in itself, useful as it
+helps explain the efficacy of the heuristic.  A general-purpose
+principle that cuts across multiple domains, (be as bijective as
+possible, formalized via standard information theory) provides more
+general insights than heuristics driven by a single domain. We claim
+that information theory is a general-purpose way to group the many
+metrics in which people assign higher costs to "less bijective"
+combinators (avoiding constants, for example).  We believe that the ad
+hoc syntactic metrics that are present in other systems are typically
+building towards something like the information theoretic measure we
+propose.
 

### 2018-09-20 papers/SymmetricLenses/response_popl19.md
changes to bottom repsonses

@@
 KSF: Does 3 add anything?? We could delete it and save some words.
+AFM: I just wanted to give them full information about how we made sure the
+lenses were correct, and this was part of my process, because that's their
+primary fear. This comes from R2, who is the main person I want to make happy.
 
@@
 
-We should have the second type be "" (to be consistent with what we use in the
-final lens)
-
-BCP: ?? 
-AFM: In hindsight, I think I will just delete this question - it's more
-of a "finding-errata" question than an actual one
+Correct. However, in signsight, we should change the example such that the
+second type is "" (to be consistent with what we use in the full lens)
 
@@
 BCP: I'm not sure I understand the question! 
-AFM: I also now realize I didn't. After rereading the section, I think this
+AFM: I also now realize I also didn't. After rereading the section, I think this
 response is more appropriate.
@@
 
-           Surely the definition of `CONTINUE(pq)` should be using `max`? Otherwise
-           I don't see how the definition makes sense, because using surely `min` 
-           will just pick 0 for t>=0, because the distance and the logarithm are
-           positive values.
+           Surely the definition of `CONTINUE(pq)` should be using `max`? 
+           Otherwise I don't see how the definition makes sense, because using 
+           surely `min` will just pick 0 for t>=0, because the distance and the
+           logarithm are positive values.
 
@@
 
-TODO: Learn Category Theory
+We are not experts in this area, but from our brief research we think that it
+may be slightly better. From "Spans of Delta Lenses," p1, equivalence relations
+are needed in both the span-presentation and set-based presentation of symmetric
+lenses to address "'hidden' details such as their *complements*." Simple
+symmetric lenses intentionally have no such details. However, there seem to be
+other complexities related to lens laws, requiring restrictions like very
+well-behavedness, as (from "Spans of Lenses", p1) "in the category whose
+morphisms are lenses, what one might expect to be the pullback of lenses need
+not satisfy the universal property of a pullback." We do not believe our
+restriction helps obviate these difficulties.
 
@@
 sufficiently explain what our syntax meant. The calls to H(...) return pairs,
-and when multiplied by constants, they act as follows:
+and when multiplied by constants, or added together, they act as follows:
 
@@
 
-See the main response above the fold.
+See the main response "above the fold" at the top of this response.
 
@@
 
-We responded to this in detail above the fold.
+We respond to this in detail above the fold.
 
@@
 
-This is true for any arbitrary "x"
+This is true for any arbitrary x (as that x gets "overwritten" by the user
+provided x')
 
@@
 
-We respond to this primarily above.  For a specific example:
+We respond to this primarily in the main response "above the fold" at the top of
+this response. For a specific example:
 
@@
 general, for different types of lenses to work, we require additional work,
-either in the synthesis algorithm (for different types of combinators like
-matching lenses or delta lenses), or in both the synthesis algorithm and
-specifications (like for edit lenses and update lenses).
+either in the synthesis algorithm (for synthesizing different types of
+combinators like matching lenses or delta lenses), or in both the synthesis
+algorithm and specifications (like for edit lenses and update lenses).
 

### 2018-09-20 papers/SymmetricLenses/response_popl19.md
changes to top

@@
     relatively easy to validate. However, it can be more difficult to check the
-    interactions of the complex combinators: Are the swaps in the right place?
-    Are the correct components disconnected? Are lenses merged in the correct
-    way? These complex combinators often occupied a few dozen lines of code
+    interactions of the complex combinators, like swaps, disconnects, and
+    merges. These complex combinators often occupied a few dozen lines of code
     rather than a few hundred, making our task easier. (Certainly, checking the
@@
 
-A fixed distribution works because, the core thing we are aiming for is "more
+A fixed distribution works because the core thing we are aiming for is "more
 bijective." A fixed distribution is merely a way to not unfairly weight one
-conjunct or disjunct over another in the presence of not knowing other
-information. Even with our inferred distribution, we are still aiming towards
-"more bijective".
+conjunct or disjunct over another in the absence of additional information.
+However, even with our fixed distribution, we are still aiming towards "more
+bijective".
 
@@
 With a learned distribution, our information theoretic measure is closer to the
-ground-truth, permitting our learned lens to, in fact, be "more bijective."
-However, we recognize that, while we believe "more bijective" is a solid guiding
-principle, there are certain situations where the most bijective lens is not the
-desired one: this is why we must further use examples for specifications. We
-could certainly imagine situations where a learned distribution would find the
-wrong lens (without additional examples), and synthesizing with the fixed
-distribution would find the correct one, and vice-versa. However, we claim that,
-as we typically want "more bijective," and a learned distribution gets closer to
-a ground-truth on that, a learned distribution likely would perform better.
+ground-truth, permitting our learned lens to be "more bijective." However, while
+we believe "more bijective" is a solid guiding principle, there are certain
+situations where the most bijective lens is not the desired one: this is why we
+must further use examples for specifications. We could certainly imagine
+situations where a learned distribution would not find the desired lens (without
+additional examples), and synthesizing with the fixed distribution would find
+the desired one, and vice-versa. However, we claim that, as we typically want
+"more bijective," and a learned distribution is closer to a ground-truth, a
+learned distribution likely would perform better.
 

### 2018-09-20 papers/SymmetricLenses/response_popl19.md
edits

@@
     code is easier than creating it de novo.)
-2.  We built unit tests to validate the more complex lenses. 
-3.  We ran the lenses on a number of strings, taken from the unit tests and the
-    examples, but with minor alterations. We did not add these tests as
-    additional unit tests, but they served as additional validation for
-    ourselves as we were constructing the benchmark suite.
-
-KSF: Does 3 add anything?? We could delete it and save some words.
-AFM: I just wanted to give them full information about how we made sure the
-lenses were correct, and this was part of my process, because that's their
-primary fear. This comes from R2, who is the main person I want to make happy.
+2.  We ran a set of tests to validate the more complex lenses.
 
@@
       - There were a number of related questions about our information theoretic
-        measure: Is it useful? Why not use a syntactic metric? Why does a
+        measure: Is it useful? Why not use a (complex) syntactic metric? Why does a
         fixed distribution work?  Will knowing the data distribution help?
 
-Having a principle guiding our heuristics is, in itself, useful as it
-helps explain the efficacy of the heuristic.  A general-purpose
-principle that cuts across multiple domains, (be as bijective as
-possible, formalized via standard information theory) provides more
-general insights than heuristics driven by a single domain. We claim
-that information theory is a general-purpose way to group the many
-metrics in which people assign higher costs to "less bijective"
-combinators (avoiding constants, for example).  We believe that the ad
-hoc syntactic metrics that are present in other systems are typically
-building towards something like the information theoretic measure we
-propose.
+It is difficult (impossible) to prove that our information-theoretic
+measure is better than all possible "syntactic" metrics, but our
+experiments show that it does do well.  In particular, it outperforms
+simple syntactic heuristics.  More important though, is the fact that
+our heuristic is based on a relatively simple, clear idea that is
+relatively easy to communicate and implement.  This idea can
+potentially be reused in a variety of related settings (and then
+perhaps augmented with ad hoc, domain-specific heuristics in those
+settings).  In contrast, complex ad hoc syntactic metrics provide
+little value beyond the single artifact in which they are used.
 
-A fixed distribution works because the core thing we are aiming for is "more
-bijective." A fixed distribution is merely a way to not unfairly weight one
-conjunct or disjunct over another in the absence of additional information.
-However, even with our fixed distribution, we are still aiming towards "more
-bijective".
+In addition, we believe that our information-theoretic ideas have
+value because they help us understand some of the ad hoc metrics used in
+the past: those in which people assign higher costs to "less
+bijective" combinators in simpler ways (avoiding constants, for
+example).  We believe that the ad hoc syntactic metrics that are
+present in other systems are often building towards something like
+the information theoretic measure we propose.
+
+A fixed distribution works because the core goal is to generate "more
+bijective" transformations.  A fixed distribution avoids unfairly
+weighting one conjunct or disjunct over another in the absence of
+additional information.  However, even with our fixed distribution, we
+are still aiming towards "more bijective".
+
+Knowing the distribution will help in certain scenarios, as it could
+help to disambiguate between several possible "equally bijective"
+lenses.  More broadly, the weightings can provide evidence that
+data from one (heavily-weighted) component on the left flows to
+another (heavily-weighted) component on the right.  For example:
+
+ANDERS:
+
+ANDERS POSSIBLY CUT THIS PARAGRAPH.  (I don't think it is detailed
+enough/concrete enough to explain exactly how a weighted distribution
+helps/hurts.):
 

### 2018-09-20 papers/SymmetricLenses/response_popl19.md
edits

@@
 A fixed distribution works because the core goal is to generate "more
-bijective" transformations.  A fixed distribution avoids unfairly
-weighting one conjunct or disjunct over another in the absence of
-additional information.  However, even with our fixed distribution, we
-are still aiming towards "more bijective".
+bijective" transformations --- the intuition is that information
+should be preserved, when possible, as opposed to thrown away, when
+data is translated between formats.  A fixed, uniform distribution avoids
+unfairly weighting one conjunct or disjunct over another and unfairly
+throwing certain information away.  In the absence of additional
+information, this kind of fairness seems to be the best we can do.
 
-Knowing the distribution will help in certain scenarios, as it could
-help to disambiguate between several possible "equally bijective"
-lenses.  More broadly, the weightings can provide evidence that
-data from one (heavily-weighted) component on the left flows to
-another (heavily-weighted) component on the right.  For example:
+On the other hand, knowing the distribution can help the system make a
+choice about what to throw away --- given a choice, it prefers to
+throw away less data.  Assume (x?.9) means "x appears
+with probability .9 and is "" otherwise.  Consider a conversion between
+these formats:
 
-ANDERS:
+(x?.999).(y?.001) <==> z
 
-ANDERS POSSIBLY CUT THIS PARAGRAPH.  (I don't think it is detailed
-enough/concrete enough to explain exactly how a weighted distribution
-helps/hurts.):
-
-Knowing the distribution will help in certain scenarios, and hinder in others.
-With a learned distribution, our information theoretic measure is closer to the
-ground-truth, permitting our learned lens to be "more bijective." However, while
-we believe "more bijective" is a solid guiding principle, there are certain
-situations where the most bijective lens is not the desired one: this is why we
-must further use examples for specifications. We could certainly imagine
-situations where a learned distribution would not find the desired lens (without
-additional examples), and synthesizing with the fixed distribution would find
-the desired one, and vice-versa. However, we claim that, as we typically want
-"more bijective," and a learned distribution is closer to a ground-truth, a
-learned distribution likely would perform better.
+Here, x appears more frequently than y on the left.  Hence, the system will prefer
+a lens that projects y and maps x to z over a lens that projects x and
+maps y to z as the latter tends to throw away much more information
+and is "less bijective" in an information theoretic sense.
 

### 2018-09-20 papers/SymmetricLenses/response_popl19.md
edits
Polishing response

@@
 
-We begin with brief responses to the most significant concerns, followed by a
-(long, optional) appendix addressing individual questions point by point.
+We begin with brief responses to the most significant concerns, followed by
+an (optional) appendix addressing individual questions point by point.
 
@@
 
-Beyond the format specifications and examples, we do not have specifications
-that define the conversion functions. Hence, we do not have an automated way to
-measure the correctness of our lenses (beyond checking that each synthesized
-lens does indeed transform the given examples correctly). In this way, our work
-is similar to other published synthesis-from-examples research, such as
-Gulwani's work on FlashFill and related projects.
+We agree we did not go into enough detail on this, and we can add more in a
+final version.
 
-To check the correctness of our tool on our benchmarks, we did what
+We do not have an automated way to measure the correctness of our lenses
+(beyond checking that each synthesized lens does indeed transform the given
+examples correctly). In this respect, our work is similar to other published
+synthesis-from-examples research, such as Gulwani's work on FlashFill and
+related projects.
+
+To assess the behavior of our tool on our benchmarks, we did what
 programmers usually do:
 
-1.  We manually inspected the code. This was certainly a nontrivial task, as
-    these generated lenses are quite large, but perhaps easier than it may
-    initially sound. Some of the lenses' structure follows from the structure of
-    the regular expressions that describe source and target format --- this is
-    relatively easy to validate. However, it can be more difficult to check the
-    interactions of the complex combinators, like swaps, disconnects, and
-    merges. These complex combinators often occupied a few dozen lines of code
-    rather than a few hundred, making our task easier. (Certainly, checking the
-    code is easier than creating it de novo.)
+1.  We manually inspected the code.
+
+    This was a nontrivial task, as generated lenses can be quite large, but
+    it was easier than it might initially sound. Some of the lenses'
+    structure should follow from the structure of the regular expressions
+    describing source and target formats; this is relatively easy to
+    validate. The more difficult part was checking the interactions of
+    complex combinators like swaps, disconnects, and merges. Still these
+    bits typically spanned only a few dozen lines of code, making our task
+    easier. Certainly, checking the code is easier than creating it de novo.
+
 2.  We ran a set of tests to validate the more complex lenses.
 
-We recognize we did not go into significant enough detail on this in the
-submitted version, and will gladly add in such detail on validation in the final
-version.
+BCP: #2 needs a few more words
 
@@
 
-      - There were a number of related questions about our information theoretic
-        measure: Is it useful? Why not use a (complex) syntactic metric? Why does a
-        fixed distribution work?  Will knowing the data distribution help?
+      - A number of related questions about our information-
+        theoretic measure: Is it useful? Why not use a 
+        (perhaps complex) syntactic metric? 
+        Why does a fixed distribution work?  
+        Will knowing the data distribution help?
 
-It is difficult (impossible) to prove that our information-theoretic
-measure is better than all possible "syntactic" metrics, but our
-experiments show that it does do well.  In particular, it outperforms
-simple syntactic heuristics.  More important though, is the fact that
-our heuristic is based on a relatively simple, clear idea that is
-relatively easy to communicate and implement.  This idea can
-potentially be reused in a variety of related settings (and then
-perhaps augmented with ad hoc, domain-specific heuristics in those
-settings).  In contrast, complex ad hoc syntactic metrics provide
-little value beyond the single artifact in which they are used.
+BCP: Difficult, or impossible???
 
-In addition, we believe that our information-theoretic ideas have
-value because they help us understand some of the ad hoc metrics used in
-the past: those in which people assign higher costs to "less
-bijective" combinators in simpler ways (avoiding constants, for
-example).  We believe that the ad hoc syntactic metrics that are
-present in other systems are often building towards something like
-the information theoretic measure we propose.
+It is difficult (impossible) to prove that our information-theoretic measure
+is better than all possible "syntactic" metrics, but our experiments show
+that it performs well.  In particular, it outperforms simple syntactic
+heuristics.  More important is the fact that our heuristic is based on a
+simple, clear idea that is easy to communicate and implement.  This idea can
+potentially be reused in a variety of related settings (and perhaps
+_augmented_ with domain-specific heuristics).
+
+BCP: I'd delete this: "By contrast, ad hoc syntactic metrics provide little
+value beyond the single artifact in which they are used."
+
+Moreover, our information-theoretic presentation has helped us understand
+some of the ad hoc metrics used in the past -- [BCP: Whose?] those in which
+people assign higher costs to "less bijective" combinators in simpler ways
+(avoiding constants, for example).  These ad hoc syntactic metrics seem to
+be approximating something like the information-theoretic measure we
+work with.
 
 A fixed distribution works because the core goal is to generate "more
-bijective" transformations --- the intuition is that information
-should be preserved, when possible, as opposed to thrown away, when
-data is translated between formats.  A fixed, uniform distribution avoids
-unfairly weighting one conjunct or disjunct over another and unfairly
-throwing certain information away.  In the absence of additional
-information, this kind of fairness seems to be the best we can do.
+bijective" transformations; the intuition is that information should be
+preserved, when possible, when data is translated between formats.  A fixed,
+uniform distribution avoids unfairly weighting one conjunct or disjunct over
+another and unfairly throwing information away.  In the absence of
+additional information, this kind of fairness seems to be the best we can
+do.  [BCP: This is not very satisfying.  Seems like the question is exactly
+"What kind of additional information might be useful and how could it be
+communicated?"  Perhaps we could comment that this could be a direction for
+future work.]
+
+BCP: Oh, maybe the next paragraph is the discussion I was just asking for.
+But I'm confused by the way it's introduced: Is this something we do now, or
+something we could do?
 
 On the other hand, knowing the distribution can help the system make a
-choice about what to throw away --- given a choice, it prefers to
-throw away less data.  Assume (x?.9) means "x appears
-with probability .9 and is "" otherwise.  Consider a conversion between
-these formats:
+choice about what to throw away: given a choice, it prefers to throw away
+less data.  For example, let (x?.9) mean "x appears with probability .9 and
+is "" otherwise and consider a conversion between these formats:
 
@@
 
-Here, x appears more frequently than y on the left.  Hence, the system will prefer
-a lens that projects y and maps x to z over a lens that projects x and
-maps y to z as the latter tends to throw away much more information
-and is "less bijective" in an information theoretic sense.
+Here, x appears more frequently than y on the left.  Hence, the system will
+prefer a lens that projects y and maps x to z over a lens that projects x
+and maps y to z, as the latter tends to throw away much more information and
+is "less bijective" in an information-theoretic sense.
 
@@
 ===========================================================================
+End of main response.
+
 Detailed responses to individual questions in the reviews follow...
@@
 
-Correct. However, in signsight, we should change the example such that the
+Correct. However, in hindsight, we should change the example such that the
 second type is "" (to be consistent with what we use in the full lens)
@@
 Within a DNF lens itself, all sequences are *required* to be involved with
-another sequence lens: otherwise when that side of the DNF SRE is provided -
-what disjunct on the other side should that lens be mapped to? Of course, there
-are other ways the disjunct can be mapped: but these must happen at a different
-level of the synthesis algorithm. For example - the entirety of that DNF regular
-expression (which itself could be deeply nested within other DNF regular
-expression components) could be projected. Furthermore, a specific sequence
-could be mapped to another, but all the components of that sequence lens
-actually may be disconnects. However, all sequences *must* be involved in a
-sequence lens in a well-typed DNF lens.
-
-BCP: I'm not sure I understand the question! 
-AFM: I also now realize I also didn't. After rereading the section, I think this
-response is more appropriate.
+another sequence lens: otherwise, when that side of the DNF SRE is provided,
+what disjunct on the other side should that lens be mapped to? Of course,
+there are other ways the disjunct can be mapped, but these must happen at a
+different level of the synthesis algorithm. For example, the entirety of
+that DNF regular expression (which itself could be deeply nested within
+other DNF regular expression components) could be projected. Furthermore, a
+specific sequence could be mapped to another, but all the components of that
+sequence lens actually may be disconnects. However, all sequences *must* be
+involved in a sequence lens in a well-typed DNF lens.
 
@@
 not the other, and expand such REs (this is where hash-consing provides its
-greatest benefits). Furthermore, we look for instances where a regular
-expression may be present in one format, but is only contained within a
-closed regular expression in the other, and we expand the closed RE
-containing the missing RE in this case. We then make this apply in slightly
-more places by also taking into account how deeply nested within stars
-certain REs occur.
+greatest benefits). Also, we look for instances where a regular expression
+may be present in one format, but is only contained within a "closed"
+regular expression in the other, and we expand the closed RE containing the
+missing RE in this case. We make this apply in a few more places by taking
+into account how deeply nested within stars certain REs occur.
 
@@
 variables (also making our integration into Boomerang significantly less hacky).
+
+BCP: OK.  Can a short version of that make it into a parenthesis after
+"closed" above?
 
@@
 
-We are not experts in this area, but from our brief research we think that it
-may be slightly better. From "Spans of Delta Lenses," p1, equivalence relations
-are needed in both the span-presentation and set-based presentation of symmetric
-lenses to address "'hidden' details such as their *complements*." Simple
-symmetric lenses intentionally have no such details. However, there seem to be
-other complexities related to lens laws, requiring restrictions like very
-well-behavedness, as (from "Spans of Lenses", p1) "in the category whose
-morphisms are lenses, what one might expect to be the pullback of lenses need
-not satisfy the universal property of a pullback." We do not believe our
-restriction helps obviate these difficulties.
+It may be slightly better, but the details are tricky. From "Spans of Delta
+Lenses," p1, equivalence relations are needed in both the span-presentation
+and set-based presentation of symmetric lenses to address "'hidden' details
+such as their *complements*." Simple symmetric lenses intentionally have no
+such details. However, there seem to be other complexities related to lens
+laws, requiring restrictions like very well-behavedness, as (from "Spans of
+Lenses", p1) "in the category whose morphisms are lenses, what one might
+expect to be the pullback of lenses need not satisfy the universal property
+of a pullback." We do not believe our restriction helps obviate these
+difficulties.
 
@@
 
-We do not believe such subproblems are straightforwardly avoidable. While we can
-attempt to match up subcomponents, this matching oftentimes will not work
-because of, for example, issues where various equivalences must be applied (for
-example: ("a"|"b")|("c"|"d") <-> "a"|("b"|("c"|"d"))) do not break down. If we
-had attempted to initially try breaking down the problem syntactically, we would
-have wasted time. This is particularly an issue because bad specifications could
-take a very long time, as there may be no good lenses between the attempted
-proposed subcomponents, so there would potentially be a great deal of wasted
-time in these attempts.
+We do not believe such subproblems are straightforwardly avoidable. While we
+can attempt to match up subcomponents, this matching oftentimes will not
+work because of, for example, issues where various equivalences must be
+applied (for example: ("a"|"b")|("c"|"d") <-> "a"|("b"|("c"|"d"))) do not
+break down. If we had attempted to initially try breaking down the problem
+syntactically, we would have wasted time. This is particularly an issue
+because bad specifications could take a very long time, as there may be no
+good lenses between the attempted proposed subcomponents, so there would
+potentially be a great deal of wasted time in these attempts.
 
@@
 
-All benchmarks were performed on a 2.5 GHz Intel Core i7 processor with 16 GB of
-1600 MHz DDR3 running macOS High Sierra.
+All benchmarks were performed on a 2.5 GHz Intel Core i7 processor with 16
+GB of 1600 MHz DDR3 running macOS High Sierra.
 
@@
 35 of the benchmarks only need bijective lenses (4 of the original Optician
-benchmarks had to be altered to be bijective because of projected data, in the
-symmetric benchmark suite we removed such alterations.) Of the 9 benchmarks that
-required the termination parameter, 4 of them required only bijective lenses.
+benchmarks had to be altered to be bijective because of projected data, in
+the symmetric benchmark suite we removed such alterations.) Of the 9
+benchmarks that required the termination parameter, 4 of them required only
+bijective lenses.
 
@@
 
-A more thorough comparison to FlashFill is present in the previous Synthesizing
-Bijective Lenses paper.
+A more thorough comparison to FlashFill can be found in the previous
+Synthesizing Bijective Lenses paper.
 
@@
 
-We respond to this in detail above the fold.
+See our detailed response "above the fold."
 
@@
 
-We chose a logarithmic x axis to show the differences on the fast benchmarks, as
-they otherwise would be imperceptable. However, if the impact on comparing the
-slower benchmarks is too significant, we can change to a linear x axis.
+We chose a logarithmic x axis to show the differences on the fast
+benchmarks, as they otherwise would be imperceptable. However, if the impact
+on comparing the slower benchmarks is too significant, we can change to a
+linear x axis.
 
 Our graph does not include any specific order of tasks. At any given time, a
-reader can see how many of the benchmarks are able to complete within that time
-(for example, 41 benchmarks require complete in under a second in the full
-synthesis algorithm). We can reword the description of these graphs to make this
-more clear.
+reader can see how many of the benchmarks are able to complete within that
+time (for example, 41 benchmarks require complete in under a second in the
+full synthesis algorithm). We can reword the description of these graphs to
+make this more clear.
 
@@
 
-We will update the lens section of the related work to be more thorough on the
-related work that we do not make formal statements about, and include proper
-citations.
+We will update the lens section of the related work to be more thorough on
+the related work that we do not make formal statements about, and include
+proper citations.
 
@@
 single_author_convert is fairly complex.  The complexity becomes even more
-apparent when working with large lenses comprised of many sub-lenses (though
-for brevity we give here the more minimal example above).  Lenses provide
-great power with their invertibility guarantees, but they come at the cost
-of thinking about many fiddly details when writing the terms.  Worrying
+apparent when working with large lenses built from many sub-lenses.  Lenses
+provide great power with their invertibility guarantees, but they come at
+the cost of thinking about fiddly details when writing the terms.  Worrying
 about these fiddly details while ALSO thinking about unambiguity
-restrictions is a very difficult task!
+restrictions is a difficult task!
 
-Engineering best practices dictate that we annotate the lenses with their
-types.  While Boomerang does not require these annotations - being able to
-infer the types from the terms - the types of the term serve as
-documentation for future programmers to understand what formats the lens
-maps between. Furthermore, the types provide resilience in the face of
-future lens modifications, as the types ensure that the lens maps exactly
-between strings of the provided formats.  Invalid inputs are not accepted
-nor are the outputs ever invalid. Consequently, in a well-maintained
-codebase, we are not requiring additional work. Instead, we are saving
-programmers from having to writing the lens by leveraging work they should
-already be performing.
+Engineering best practices suggest that it's a good idea to annotate lenses
+with their types.  While Boomerang does not require these annotations (it
+can infer the types from the terms), types serve as documentation for future
+programmers to understand what formats the lens maps between. Furthermore,
+the types provide resilience in the face of future lens modifications,
+ensuring that the lens maps exactly between strings of the provided formats.
 
@@
 
-We will make this more clear in future versions of the paper. Expand is
-needed because GreedySynth alone cannot traverse regular expression
-equivalences. If given the two regular expressions "" | "a"+, and "a"\*,
-GreedySynth alone would merge "" to "a"\* (with a disconect), and merge "a"+
-(also with a disconect).  However, contuing search through equivalent REs
-helps find the bijective lens.
+Expand is needed because GreedySynth alone cannot traverse regular
+expression equivalences. If given the two regular expressions "" | "a"+, and
+"a"\*, GreedySynth alone would merge "" to "a"\* (with a disconect), and
+merge "a"+ (also with a disconect).  However, contuing search through
+equivalent REs helps find the bijective lens.
 
@@
 
-We respond to this primarily in the main response "above the fold" at the top of
-this response. For a specific example:
+We respond to this primarily in the main response "above the fold" at the
+top of this response. For a specific example:
 
 Consider finding a lens of type:
+
     ("a" |(.5) "b") |(.5) "c" <=> "x" |(.1) "y"
@@
 Consider finding a lens of type:
+
     ("a" |(.5) "b") |(.5) "c" <=> "x" |(.9) "y"
@@
 
-We are happy to provide an example such as this one in the paper.
+We can provide an example like this in the paper.
 
@@
 
-We respond to these issues in the main reponse.
+See the main reponse.
 
@@
 
-We agree that section 7 requires more, informal comparisons to other lenses, for
-us to discuss comparisons between our framework and other types of lenses. In
-general, for different types of lenses to work, we require additional work,
-either in the synthesis algorithm (for synthesizing different types of
-combinators like matching lenses or delta lenses), or in both the synthesis
-algorithm and specifications (like for edit lenses and update lenses).
+We agree that section 7 requires more, informal comparisons to other lenses,
+for us to discuss comparisons between our framework and other types of
+lenses. In general, for different types of lenses to work, we require
+additional work, either in the synthesis algorithm (for synthesizing
+different types of combinators like matching lenses or delta lenses), or in
+both the synthesis algorithm and specifications (like for edit lenses and
+update lenses).
 
@@
 
-Stochastic REs, and more generally, a semantic cost metric is critical for
-comparing between responses by GreedySynth. Without this semantic cost, the same
-function would likely have different costs if found through different calls to
-GreedySynth. Syntactic metrics, we feel, aim to approximate how good the
-underlying result is: in this work we are able to find a semantic metric that we
-can be found using syntactic information.
+Stochastic REs, and more generally, a semantic cost metric are critical for
+comparing between responses by GreedySynth. Without this semantic cost, the
+same function would likely have different costs if found through different
+calls to GreedySynth. Syntactic metrics, we feel, aim to approximate how
+good the underlying result is: in this work we are able to find a semantic
+metric that we can be found using syntactic information.
 
@@
 introduced. We do believe our syntactic technique for calculating their
-entropy is new.
+entropy is novel.

### 2018-09-20 papers/SymmetricLenses/response_popl19.md
changes

--- 2018-09-20 papers/SymmetricLenses/response_popl19.md
@@
       - There were a number of related questions about our information theoretic
-        measure: Is it useful? Why not use a (complex) syntactic metric? Why does a
-        fixed distribution work?  Will knowing the data distribution help?
+        measure: Is it useful? Why not use a (complex) syntactic metric? Why 
+        does a fixed distribution work?  Will knowing the data distribution 
+        help?
 
@@
 
-Here, x appears more frequently than y on the left.  Hence, the system will prefer
-a lens that projects y and maps x to z over a lens that projects x and
-maps y to z as the latter tends to throw away much more information
-and is "less bijective" in an information theoretic sense.
+Here, x appears more frequently than y on the left. Hence, the system will
+prefer a lens that projects y and maps x to z over a lens that projects x and
+maps y to z as the latter tends to throw away much more information and is "less
+bijective" in an information theoretic sense.
 
@@
 
-BCP: I'm not sure I understand the question! 
-AFM: I also now realize I also didn't. After rereading the section, I think this
-response is more appropriate.
-
                                 -------------
@@
 
-BCP: What is "closed"?
-AFM: The internals of the Regular Expression are not exposed to GreedySynth,
-they are the analogue of User Defined regular expressions in the previous paper,
-but with this closed/open approach, we aren't dependent on where users place
-variables (also making our integration into Boomerang significantly less hacky).
-
 Full details can be found in the prior work on Synthesizing Bijective Lenses
-(page 16).
+(page 16), though in that context, "user-defined data types" are used instead of
+"closed regular expressions."
 
@@
 
-We responded to the key points in this comment (the benchmark suite and
-judging correctness) "above the fold" at the top of this response.
+We responded to issues on judging correctness "above the fold" at the top of
+this response.
+
+We chose our benchmarks from the FlashFill paper because the 238 tasks are,
+unfortunately, not publicly available. In general, our benchmark suite goes
+beyond the level of difficulty required for data cleaning. Many real-life
+file format descriptors involve large numbers of disjunctions and nested
+iterations, which FlashFill is either bad at (for disjunctions) or cannot
+handle at all (nested iterations).
+
+A more thorough comparison to FlashFill is present in the previous Synthesizing
+Bijective Lenses paper.
 
@@
 We respond to this primarily in the main response "above the fold" at the top of
-this response. For a specific example:
+this response.
 
-Consider finding a lens of type:
-("a" |(.5) "b") |(.5) "c" <=> "x" |(.1) "y"
-
-In this situation, the lowest cost lens would merge "a" and "b" into "x", and
-keep "c" mapped to "y".
-
-Consider finding a lens of type:
-("a" |(.5) "b") |(.5) "c" <=> "x" |(.9) "y"
-
-In this situation, the lowest cost lens would merge "a" and "b" into "y", and
-keep "c" mapped to "x".
-
-We are happy to provide an example such as this one in the paper.
+We are happy to provide an example such as the one above in the paper.
 

### 2018-09-20 papers/SymmetricLenses/response_popl19.md

--- 2018-09-20 papers/SymmetricLenses/response_popl19.md
@@
 BCP: Difficult, or impossible???
-
 It is difficult (impossible) to prove that our information-theoretic measure
@@
 
-BCP: What is "closed"?
-AFM: The internals of the Regular Expression are not exposed to GreedySynth,
-they are the analogue of User Defined regular expressions in the previous paper,
-but with this closed/open approach, we aren't dependent on where users place
-variables (also making our integration into Boomerang significantly less hacky).
-
 BCP: OK.  Can a short version of that make it into a parenthesis after
@@
 Full details can be found in the prior work on Synthesizing Bijective Lenses
-(page 16).
+(page 16), though in that context, "user-defined data types" are used instead of
+"closed regular expressions."
 

### 2018-09-20 papers/SymmetricLenses/response_popl19.md
updated and reacted

@@
 
-BCP: Difficult, or impossible???
-It is difficult (impossible) to prove that our information-theoretic measure
+It is impossible to prove that our information-theoretic measure
 is better than all possible "syntactic" metrics, but our experiments show

### 2018-09-20 papers/SymmetricLenses/response_popl19.md
minor changes

@@
 
-    This was a nontrivial task, as generated lenses can be quite large, but
-    it was easier than it might initially sound. Some of the lenses'
-    structure should follow from the structure of the regular expressions
-    describing source and target formats; this is relatively easy to
-    validate. The more difficult part was checking the interactions of
-    complex combinators like swaps, disconnects, and merges. Still these
-    bits typically spanned only a few dozen lines of code, making our task
-    easier. Certainly, checking the code is easier than creating it de novo.
+    This was a nontrivial task, as generated lenses can be quite large, but it
+    was easier than it might initially sound. Some of the lenses' structure
+    should follow from the structure of the regular expressions describing
+    source and target formats; this is relatively easy to validate. The more
+    difficult part was checking the interactions of complex combinators like
+    swaps, disconnects, and merges. Still these bits typically spanned only a
+    few dozen lines of code, making our task easier. Certainly, checking the
+    code is easier than creating it de novo.
 
-2.  We ran a set of tests to validate the more complex lenses.
+2.  We ran a set of tests to validate the more complex lenses. These tests
+    confirmed the output of creates and puts were as we expected them to be.
 
-BCP: #2 needs a few more words
+After manually validating that SS found the desired lenses, we validated the
+lenses output in other modes by comparing those lenses to the ground truth of
+SS.
 
@@
 
-BCP: I'd delete this: "By contrast, ad hoc syntactic metrics provide little
-value beyond the single artifact in which they are used."
+Moreover, our information-theoretic presentation has helped us understand some
+of the ad hoc metrics used in the past -- like those used in FlashFill and
+Refazer -- those in which people assign higher costs to "less bijective"
+combinators in simpler ways (avoiding constants, for example). These ad hoc
+syntactic metrics seem to be approximating something like the
+information-theoretic measure we work with.
 
-Moreover, our information-theoretic presentation has helped us understand
-some of the ad hoc metrics used in the past -- [BCP: Whose?] those in which
-people assign higher costs to "less bijective" combinators in simpler ways
-(avoiding constants, for example).  These ad hoc syntactic metrics seem to
-be approximating something like the information-theoretic measure we
-work with.
+A fixed distribution works because the core goal is to generate "more bijective"
+transformations; the intuition is that information should be preserved, when
+possible, when data is translated between formats. A fixed, uniform distribution
+avoids unfairly weighting one conjunct or disjunct over another and unfairly
+throwing information away. In the absence of a distribution, this kind of
+fairness seems to be the best we can do.
 
-A fixed distribution works because the core goal is to generate "more
-bijective" transformations; the intuition is that information should be
-preserved, when possible, when data is translated between formats.  A fixed,
-uniform distribution avoids unfairly weighting one conjunct or disjunct over
-another and unfairly throwing information away.  In the absence of
-additional information, this kind of fairness seems to be the best we can
-do.  [BCP: This is not very satisfying.  Seems like the question is exactly
-"What kind of additional information might be useful and how could it be
-communicated?"  Perhaps we could comment that this could be a direction for
-future work.]
-
-BCP: Oh, maybe the next paragraph is the discussion I was just asking for.
-But I'm confused by the way it's introduced: Is this something we do now, or
-something we could do?
-
-On the other hand, knowing the distribution can help the system make a
-choice about what to throw away: given a choice, it prefers to throw away
-less data.  For example, let (x?.9) mean "x appears with probability .9 and
-is "" otherwise and consider a conversion between these formats:
+However, if given the distribution, the system can make a more informed choice
+about what to throw away: given a choice, it prefers to throw away less data.
+For example, let (x?.9) mean "x appears with probability .9 and is "" otherwise
+and consider a conversion between these formats:
 
@@
 At a high level, we process both regular expression trees, looking for
-situations where a regular expression is a subcomponent of one format but
+situations where a closed regular expression is a subcomponent of one format but
 not the other, and expand such REs (this is where hash-consing provides its
-greatest benefits). Also, we look for instances where a regular expression
-may be present in one format, but is only contained within a "closed"
-regular expression in the other, and we expand the closed RE containing the
-missing RE in this case. We make this apply in a few more places by taking
-into account how deeply nested within stars certain REs occur.
-
-BCP: OK.  Can a short version of that make it into a parenthesis after
-"closed" above?
+greatest benefits). As the internals of closed REs are not exposed to
+GreedySynth, this is necessary to synchronize that closed RE with anything but
+disconnect. Also, we look for instances where a regular expression may be
+present in one format, but is only contained within a closed regular expression
+in the other, and we expand the closed RE containing the missing RE in this
+case. We make this apply in a few more places by taking into account how deeply
+nested within stars certain REs occur.
 
@@
 
-We respond to this primarily in the main response "above the fold" at the
-top of this response. For a specific example:
+We respond to this in the main response "above the fold".
 
-Consider finding a lens of type:
-
-    ("a" |(.5) "b") |(.5) "c" <=> "x" |(.1) "y"
-
-In this situation, the lowest cost lens would merge "a" and "b" into "x", and
-keep "c" mapped to "y".
-
-Consider finding a lens of type:
-
-    ("a" |(.5) "b") |(.5) "c" <=> "x" |(.9) "y"
-
-In this situation, the lowest cost lens would merge "a" and "b" into "y", and
-keep "c" mapped to "x".
-
-We can provide an example like this in the paper.
+We can provide an example, like the one in the main response, in the paper.
 

### 2018-09-20 papers/SymmetricLenses/response_popl19.md
clean

@@
 After manually validating that SS found the desired lenses, we validated the
-lenses output in other modes by comparing those lenses to the ground truth of
-SS.
+lenses synthesized in other modes by comparing those lenses to the ground truth
+of SS.
 
@@
 avoids unfairly weighting one conjunct or disjunct over another and unfairly
-throwing information away. In the absence of a distribution, this kind of
-fairness seems to be the best we can do.
+throwing information away. If not given a distribution this kind of fairness
+seems to be the best we can do.
 

### 2018-09-21 papers/SymmetricLenses/response_popl19.md
One more comment
minor edits

@@
-BCP: The first section is too long -- 500 words max.
-
 Thank you for your comments!
@@
 We begin with brief responses to the most significant concerns, followed by
-an (optional) appendix addressing individual questions point by point.
+an (optional) appendix addressing individual questions.
 
@@
 
-     - How do we determine what a "correct" lens is? Can we make this process a 
-       bit more automatic and objective?
+     - How do we determine whether a lens is "correct"? Can we make this process  
+       more automatic and objective?
 
-We agree we did not go into enough detail on this, and we can add more in a
-final version.
+We agree we should discuss this point in more detail; we will do so in
+the final version.
 
 We do not have an automated way to measure the correctness of our lenses
-(beyond checking that each synthesized lens does indeed transform the given
-examples correctly). In this respect, our work is similar to other published
+beyond checking that each synthesized lens does indeed transform the given
+examples correctly. In this respect, our work is similar to other published
 synthesis-from-examples research, such as Gulwani's work on FlashFill and
@@
 
-To assess the behavior of our tool on our benchmarks, we did what
+To assess the correctness of our tool on the benchmarks, we did what
 programmers usually do:
@@
 
-    This was a nontrivial task, as generated lenses can be quite large, but it
+    This task was nontrivial, as generated lenses can be quite large, but it
     was easier than it might initially sound. Some of the lenses' structure
@@
     difficult part was checking the interactions of complex combinators like
-    swaps, disconnects, and merges. Still these bits typically spanned only a
+    swaps, disconnects, and merges. Still these bits typically span only a
     few dozen lines of code, making our task easier. Certainly, checking the
@@
 
-2.  We ran a set of tests to validate the more complex lenses. These tests
-    confirmed the output of creates and puts were as we expected them to be.
+2.  We ran a series of unit tests to validate the more complex lenses. These tests
+    confirmed that the output of creates and puts were as expected.
 
 After manually validating that SS found the desired lenses, we validated the
-lenses synthesized in other modes by comparing those lenses to the ground truth
-of SS.
+lenses synthesized in other modes by comparing those lenses to the ground
+truth of SS.
 

### 2018-09-21 papers/SymmetricLenses/response_popl19.md
minor word change

--- 2018-09-20 papers/SymmetricLenses/response_popl19.md
@@
-BCP: The first section is too long -- 500 words max.
-
 Thank you for your comments!
 
-We begin with brief responses to the most significant concerns, followed by
-an (optional) appendix addressing individual questions point by point.
+We begin with brief responses to the most significant concerns, followed by an
+(optional) appendix addressing individual questions point-by-point.
 
@@
 
-     - How do we determine what a "correct" lens is? Can we make this process a 
-       bit more automatic and objective?
+     - How do we determine what a "correct" lens is? Can we make this process 
+       more automatic and objective?
 
-We agree we did not go into enough detail on this, and we can add more in a
-final version.
+We agree we did not go into enough detail on this; we will add more in a final
+version.
 
-We do not have an automated way to measure the correctness of our lenses
-(beyond checking that each synthesized lens does indeed transform the given
-examples correctly). In this respect, our work is similar to other published
-synthesis-from-examples research, such as Gulwani's work on FlashFill and
-related projects.
+We do not have an automated way to measure the correctness of our lenses (beyond
+validating synthesized lenses conform to their specifications). In this respect,
+our work is similar to other published synthesis-from-examples research, like
+FlashFill and related projects.
 
-To assess the behavior of our tool on our benchmarks, we did what
-programmers usually do:
+To validate our tool synthesized the intended lenses, we did what programmers
+usually do:
 
@@
     This was a nontrivial task, as generated lenses can be quite large, but it
-    was easier than it might initially sound. Some of the lenses' structure
-    should follow from the structure of the regular expressions describing
-    source and target formats; this is relatively easy to validate. The more
-    difficult part was checking the interactions of complex combinators like
-    swaps, disconnects, and merges. Still these bits typically spanned only a
-    few dozen lines of code, making our task easier. Certainly, checking the
+    was easier than it might sound. Some of the lenses' structure should follow
+    the structure of their regular expression types; this is relatively easy to
+    validate. The more difficult part was checking the interactions of complex
+    combinators like swaps, disconnects, and merges. However, these bits
+    typically spanned only a few dozen lines of code. Certainly, checking the
     code is easier than creating it de novo.
 
-2.  We ran a set of tests to validate the more complex lenses. These tests
-    confirmed the output of creates and puts were as we expected them to be.
+2.  We ran a set of tests to validate the more complex lenses. These tested that
+    the synthesized lenses performed creates and puts as we expected them to.
 
 After manually validating that SS found the desired lenses, we validated the
-lenses synthesized in other modes by comparing those lenses to the ground truth
-of SS.
+lenses synthesized in other modes by comparing those lenses to the ground
+truth of SS.
 
@@
 
-It is impossible to prove that our information-theoretic measure
-is better than all possible "syntactic" metrics, but our experiments show
-that it performs well.  In particular, it outperforms simple syntactic
-heuristics.  More important is the fact that our heuristic is based on a
+It is impossible to prove that our information-theoretic measure is better than
+all possible "syntactic" metrics; our experiments show that it performs
+well, outperforming simple ones. More importantly, our heuristic is based on a
 simple, clear idea that is easy to communicate and implement. This idea can
-potentially be reused in a variety of related settings (and perhaps
-_augmented_ with domain-specific heuristics).
+potentially be reused in a variety of related settings (and perhaps _augmented_
+with domain-specific heuristics).
 
-Moreover, our information-theoretic presentation has helped us understand some
-of the ad hoc metrics used in the past -- like those used in FlashFill and
-Refazer -- those in which people assign higher costs to "less bijective"
-combinators in simpler ways (avoiding constants, for example). These ad hoc
-syntactic metrics seem to be approximating something like the
-information-theoretic measure we work with.
+Moreover, our information-theoretic presentation helped us understand some of
+the ad hoc metrics used in the past -- like in FlashFill and Refazer -- in which
+"less bijective" combinators are assigned higher costs in simpler ways (like
+avoiding constants). These ad hoc syntactic metrics seem to be approximating
+something like the information-theoretic measure we work with.
 
 A fixed distribution works because the core goal is to generate "more bijective"
-transformations; the intuition is that information should be preserved, when
-possible, when data is translated between formats. A fixed, uniform distribution
-avoids unfairly weighting one conjunct or disjunct over another and unfairly
-throwing information away. If not given a distribution this kind of fairness
-seems to be the best we can do.
+transformations; information should be preserved, when possible, when
+translating data between formats. A fixed, uniform distribution avoids unfairly
+weighting one conjunct or disjunct over another, unfairly throwing information
+away. When not given a distribution this kind of fairness seems to be the best
+we can do.
 
-However, if given the distribution, the system can make a more informed choice
-about what to throw away: given a choice, it prefers to throw away less data.
-For example, let (x?.9) mean "x appears with probability .9 and is "" otherwise
-and consider a conversion between these formats:
+If given the distribution, the system can make a more informed choice about what
+to throw away. Let (x?.9) mean "x appears with probability .9 and is ""
+otherwise." Consider a conversion between these formats:
 
@@
 
-Here, x appears more frequently than y on the left.  Hence, the system will
-prefer a lens that projects y and maps x to z over a lens that projects x
-and maps y to z, as the latter tends to throw away much more information and
-is "less bijective" in an information-theoretic sense.
+Here, x appears more frequently than y on the left. The system will prefer a
+lens that projects y and maps x to z over a lens that projects x and maps y to
+z; the latter tends to throw away more information and is "less bijective" in an
+information-theoretic sense.
 
@@
 Correct. However, in hindsight, we should change the example such that the
-second type is "" (to be consistent with what we use in the full lens)
+second type is "" (to be consistent with what we use in the full lens in Figure
+3)
 
@@
 
+35 of the 48 symmetric benchmarks synthesized the same lenses as in the
+Synthesizing Bijective Lenses paper. 4 of the benchmarks in Synthesizing
+Bijective Lenses had to be altered to be bijective to account for projecting
+information: those 4 were altered to their original forms, and we built our
+algorithm on those 39 benchmarks. The remaining 9 benchmarks were added without
+any changes to the synthesis algorithm (although we did make bugfixes while
+adding those benchmarks). The 39 benchmarks from Optician hence were certainly
+not hand-picked, and our synthesis algorithm was certainly not overtuned to the
+remaining 9 benchmarks.
+
                                 -------------
@@
 
-We agree that section 7 requires more, informal comparisons to other lenses,
-for us to discuss comparisons between our framework and other types of
-lenses. In general, for different types of lenses to work, we require
-additional work, either in the synthesis algorithm (for synthesizing
-different types of combinators like matching lenses or delta lenses), or in
-both the synthesis algorithm and specifications (like for edit lenses and
-update lenses).
+We agree that section 7 requires more, informal comparisons to other lenses, for
+us to discuss comparisons between our framework and other types of lenses. In
+general, for different types of lenses to work, we require additional work,
+either in the synthesis algorithm (for synthesizing different types of
+combinators like matching lenses or delta lenses), or in both the synthesis
+algorithm and specifications (like for edit lenses and update lenses).
 

### 2018-09-21 papers/SymmetricLenses/response_popl19.md

@@
 
-We begin with brief responses to the most significant concerns, followed by an
-(optional) appendix addressing individual questions point-by-point.
+We begin with brief responses to the most significant concerns, followed by
+an (optional) appendix addressing individual questions.
 
@@
 
-     - How do we determine what a "correct" lens is? Can we make this process 
+     - How do we determine whether a lens is "correct"? Can we make this process
        more automatic and objective?
 
-We agree we did not go into enough detail on this; we will add more in a final
+We agree we should discuss this point in more detail; we will do so in the final
 version.
 
-We do not have an automated way to measure the correctness of our lenses (beyond
-validating synthesized lenses conform to their specifications). In this respect,
+We do not have an automated way to measure the correctness of our lenses beyond
+validating synthesized lenses conform to their specifications. In this respect,
 our work is similar to other published synthesis-from-examples research, like
@@
 
-    This was a nontrivial task, as generated lenses can be quite large, but it
+    This task was nontrivial, as generated lenses can be quite large, but it
     was easier than it might sound. Some of the lenses' structure should follow
@@
     combinators like swaps, disconnects, and merges. However, these bits
-    typically spanned only a few dozen lines of code. Certainly, checking the
+    typically span only a few dozen lines of code. Certainly, checking the
     code is easier than creating it de novo.
 
-2.  We ran a set of tests to validate the more complex lenses. These tested that
-    the synthesized lenses performed creates and puts as we expected them to.
+2.  We ran a series of unit tests to validate the more complex lenses. These
+    tests confirmed that the output of creates and puts were as expected.
 

### 2018-09-21 papers/SymmetricLenses/response_popl19.md
add on average

@@
 lens that projects y and maps x to z over a lens that projects x and maps y to
-z; the latter tends to throw away more information and is "less bijective" in an
-information-theoretic sense.
+z; the latter tends to throw away more information on average and is "less
+bijective" in an information-theoretic sense.
 

### 2018-09-21 papers/SymmetricLenses/response_popl19.md
dave edits

@@
 We chose our benchmarks from the FlashFill paper because the 238 tasks are,
-unfortunately, not publicly available. In general, our benchmark suite goes
-beyond the level of difficulty required for data cleaning. Many real-life
+unfortunately, not publicly available.  Many real-life
 file format descriptors involve large numbers of disjunctions and nested
@@
 
-A more thorough comparison to FlashFill is present in the previous Synthesizing
-Bijective Lenses paper.
+In our previous work on synthesizing bijective lenses, we compared
+FlashFill to our bijective synthesis algorithm in more detail, finding
+that FlashFill was able to synthesize just 5 of the 39 examples that
+we had at that time due to the presence of nested iterations.  In any
+subsequent version of this paper, we would be happy to compare our
+work with FlashFill in greater detail.
 
@@
 
-This was us trying to make our equations more readable, but we did not
-sufficiently explain what our syntax meant. The calls to H(...) return pairs,
+We were attempting to make our equations more readable, but we did not
+sufficiently explain what our syntax meant. The calls to H(...) returns pairs,
 and when multiplied by constants, or added together, they act as follows:
@@
 
-We chose our benchmarks from the FlashFill paper because the 238 tasks are,
-unfortunately, not publicly available. In general, our benchmark suite goes
-beyond the level of difficulty required for data cleaning. Many real-life
-file format descriptors involve large numbers of disjunctions and nested
-iterations, which FlashFill is either bad at (for disjunctions) or cannot
-handle at all (nested iterations).
+We chose our benchmarks from the FlashFill paper because the 238 tasks
+are, unfortunately, not publicly available.  Many real-life file
+format descriptors involve large numbers of disjunctions and nested
+iterations, which FlashFill is either bad at (for disjunctions) or
+cannot handle at all (nested iterations).
 
-A more thorough comparison to FlashFill can be found in the previous
-Synthesizing Bijective Lenses paper.
+In our previous work, we compared FlashFill to our bijective synthesis
+algorithm in more detail, finding that FlashFill was able to
+synthesize just 5 of the 39 examples that we had at that time due to
+the presence of nested iterations.  In any subsequent version of this
+paper, we would be happy to compare our work with FlashFill in greater
+detail.
 
@@
 
-Engineering best practices suggest that it's a good idea to annotate lenses
+Moreover, engineering best practices suggest that it's a good idea to annotate lenses
 with their types.  While Boomerang does not require these annotations (it
 can infer the types from the terms), types serve as documentation for future
-programmers to understand what formats the lens maps between. Furthermore,
+programmers to understand the formats that lenses process. Furthermore,
 the types provide resilience in the face of future lens modifications,
-ensuring that the lens maps exactly between strings of the provided formats.
+ensuring that the lens maps exactly between strings of the provided
+formats.  Hence, if lens authors are going to write down the types
+anyway, why not allow our system to synthesize the fiddly terms from
+those types for the programmer?  It seems clear work will be saved.
 
@@
 
-Expand is needed because GreedySynth alone cannot traverse regular
+Expand is needed because GreedySynth alone does not account for regular
 expression equivalences. If given the two regular expressions "" | "a"+, and
-"a"\*, GreedySynth alone would merge "" to "a"\* (with a disconect), and
-merge "a"+ (also with a disconect).  However, contuing search through
-equivalent REs helps find the bijective lens.
+"a"\*, GreedySynth alone will not synthesize the right lens:  it will
+merge "" to "a"\* (with a disconect), and merge "a"+ (also with a
+disconect).  The Expand algorithm searches for semantically equivalent
+REs that are "compatible" in the sense that they allow GreedySynth to
+find additional lenses between them.  In particular, in the example
+given here, using Expand allows us to find a bijective lens.
 
@@
 
-We respond to this in the main response "above the fold".
+Please see above, in the initial 500 words of this reponse.
 
-We can provide an example, like the one in the main response, in the paper.
+We can provide an example, like the one in the main response above, in the paper.
 
@@
 
-See the main reponse.
+Please see above, in the initial 500 words of this reponse.
 
@@
 any changes to the synthesis algorithm (although we did make bugfixes while
-adding those benchmarks). The 39 benchmarks from Optician hence were certainly
-not hand-picked, and our synthesis algorithm was certainly not overtuned to the
+adding those benchmarks). The 39 benchmarks from Optician hence were
+not hand-picked, and our synthesis algorithm was not overtuned to the
 remaining 9 benchmarks.
@@
 
-Stochastic REs, and more generally, a semantic cost metric are critical for
-comparing between responses by GreedySynth. Without this semantic cost, the
-same function would likely have different costs if found through different
-calls to GreedySynth. Syntactic metrics, we feel, aim to approximate how
-good the underlying result is: in this work we are able to find a semantic
-metric that we can be found using syntactic information.
+Please see above, in the initial 500 words of this response.
 

### 2018-09-21 papers/SymmetricLenses/response_popl19.md
add better response

@@
 
-See our detailed response "above the fold."
+See our detailed response "above the fold," for a description of our process.
+
+Without an existing lens to compare to and without an existing set of
+input/output examples, it is not obvious to us how to make the process of
+checking correctness more automatic and objective.
+
+We are interested in finding ways to make validating lenses easier for the user,
+for example by finding examples of puts and creates that differentiate between
+highly ranked lenses. However, how exactly this process would work is
+not obvious, and we leave this for future work.
 

### 2018-09-21 papers/SymmetricLenses/response_popl19.md
Kathleens edits

--- 2018-09-21 papers/SymmetricLenses/response_popl19.md
@@
 
-We do not have an automated way to measure the correctness of our lenses beyond
-validating synthesized lenses conform to their specifications. In this respect,
-our work is similar to other published synthesis-from-examples research, like
-FlashFill and related projects.
+We do not have an automated way to measure the correctness of our
+lenses beyond confirming that they conform to their specifications. In
+this respect, our work is similar to other published
+synthesis-from-examples research, like FlashFill and related projects.
 
-To validate our tool synthesized the intended lenses, we did what programmers
-usually do:
+To validate the synthesized lenses, we did what programmers usually do:
 
@@
 
-After manually validating that SS found the desired lenses, we validated the
-lenses synthesized in other modes by comparing those lenses to the ground
-truth of SS.
+After manually validating that the SS mode found the desired lenses,
+we validated other modes by matching their output to that of SS.
 
@@
 
-A fixed distribution works because the core goal is to generate "more bijective"
-transformations; information should be preserved, when possible, when
-translating data between formats. A fixed, uniform distribution avoids unfairly
-weighting one conjunct or disjunct over another, unfairly throwing information
-away. When not given a distribution this kind of fairness seems to be the best
-we can do.
+A fixed distribution works because the core goal is to generate "more
+bijective" transformations; information should be preserved, wherever
+possible, when translating data between formats. A fixed, uniform
+distribution avoids weighting one conjunct or disjunct over
+another. When not given a distribution, this kind of "fairness" seems
+to be the best we can do.
 
@@
 
-Here, x appears more frequently than y on the left. The system will prefer a
-lens that projects y and maps x to z over a lens that projects x and maps y to
-z; the latter tends to throw away more information on average and is "less
-bijective" in an information-theoretic sense.
+Here, x appears more frequently than y on the left. The system will
+prefer a lens that projects y and maps x to z over a lens that
+projects x and maps y to z because the latter throws away more
+information on average and is "less bijective" in an
+information-theoretic sense.
 
@@
 
-Yes, we will make this clearer on p. 3 too.
+Yes, we will make this clear on p. 3 too.
 
@@
 
-Correct. However, in hindsight, we should change the example such that the
-second type is "" (to be consistent with what we use in the full lens in Figure
-3)
+Correct. However, in hindsight, we should change the example so that the
+second type is "" to be consistent with the full lens in Figure 3.
 
@@
 
-This is correct -- we will clarify.
+You are correct. We will clarify.
 
@@
 
-Within a DNF lens itself, all sequences are *required* to be involved with
-another sequence lens: otherwise, when that side of the DNF SRE is provided,
-what disjunct on the other side should that lens be mapped to? Of course,
-there are other ways the disjunct can be mapped, but these must happen at a
-different level of the synthesis algorithm. For example, the entirety of
-that DNF regular expression (which itself could be deeply nested within
-other DNF regular expression components) could be projected. Furthermore, a
-specific sequence could be mapped to another, but all the components of that
-sequence lens actually may be disconnects. However, all sequences *must* be
-involved in a sequence lens in a well-typed DNF lens.
+Within a DNF lens, all sequences are *required* to pair with another
+sequence lens. Otherwise, when given one side of the DNF SRE, what
+disjunct on the other side should it be mapped to?  There are other
+ways the disjunct can be mapped, but they must happen at a different
+level of the synthesis algorithm. For example, the entirety of that
+DNF regular expression (which itself could be deeply nested within
+other DNF regular expression components) could be
+projected. Furthermore, a specific sequence could be mapped to
+another, but all the components of that sequence lens may be
+disconnects. However, all sequences *must* be involved in a sequence
+lens in a well-typed DNF lens.
 
@@
 
-At a high level, we process both regular expression trees, looking for
-situations where a closed regular expression is a subcomponent of one format but
-not the other, and expand such REs (this is where hash-consing provides its
-greatest benefits). As the internals of closed REs are not exposed to
-GreedySynth, this is necessary to synchronize that closed RE with anything but
-disconnect. Also, we look for instances where a regular expression may be
-present in one format, but is only contained within a closed regular expression
-in the other, and we expand the closed RE containing the missing RE in this
-case. We make this apply in a few more places by taking into account how deeply
+EXPAND processes both regular expression trees, looking for situations
+where a closed regular expression is a subcomponent of one format but
+not the other.  We expand such REs (hash-consing provides great
+benefits here). As the internals of closed REs are not exposed to
+GreedySynth, this expansion is necessary to synchronize the closed RE
+with anything but disconnect. EXPAND also looks for instances where a
+regular expression is present in one format but is only contained
+within a closed regular expression in the other.  In this case, we
+expand the closed RE containing the missing RE.  We apply this
+rewriting in a few more places by taking into account how deeply
 nested within stars certain REs occur.
 
-Full details can be found in the prior work on Synthesizing Bijective Lenses
-(page 16), though in that context, "user-defined data types" are used instead of
-"closed regular expressions."
+Full details can be found in the prior work on Synthesizing Bijective
+Lenses (page 16), though in that context, the term "user-defined data
+types" is used instead of "closed regular expressions."
 
@@
 
-Right, this should be max!
+Right!
 
@@
 
-These benchmarks are using the full "SS" mode. The only difference between these
-benchmarks and the full mode is that termination parameters are ignored, and
-are instead required to use the 0, 25, and -25 parameters.
+These benchmarks use the full "SS" mode. The only difference between these
+benchmarks and the full mode is that they ignore termination
+parameters and use the 0, 25, and -25 parameters instead.
+
+KSF: Anders, what is the difference between full SS mode and full mode?
 
@@
 
-Yes, we do have some hope that the algorithm can be extended to handle such
-variants.
+Yes.
 
 Delta lenses would require additional work in understanding "shape"
-constructs.  Currently our language only permits basic iteration, whereas
-shapes allow for different types of mergings. Allowing this would require
-changes to the AtomSynth algorithm, in that we would we would need to
-discover the correct shapes of the data, and align them in the correct ways.
+constructs.  Currently our language only permits basic iteration,
+whereas shapes allow for different types of mergings. This extension
+would require changes to the AtomSynth algorithm in that we would need
+to discover the correct shapes of the data and align them correctly.
 
-Both edit and update lenses would require new forms of specifications, since
-both permit updates that are not merely pushing data through, but instead
-are applying specific elements of an edit monoid.  (Also we would remove
-complements from edit lenses).
+Both edit and update lenses would require new forms of specifications,
+since both permit updates that are not merely pushing data through,
+but instead are applying elements of an edit monoid.  (Also we would
+remove complements from edit lenses).
 
@@
 
-We responded to issues on judging correctness "above the fold" at the top of
-this response.
+We responded to issues on judging correctness at the top of this response.
 
-We chose our benchmarks from the FlashFill paper because the 238 tasks are,
-unfortunately, not publicly available.  Many real-life
-file format descriptors involve large numbers of disjunctions and nested
-iterations, which FlashFill is either bad at (for disjunctions) or cannot
-handle at all (nested iterations).
+We chose our benchmarks from the FlashFill paper because the fuller
+set of 238 tasks are not publicly available.
 
-In our previous work on synthesizing bijective lenses, we compared
-FlashFill to our bijective synthesis algorithm in more detail, finding
-that FlashFill was able to synthesize just 5 of the 39 examples that
-we had at that time due to the presence of nested iterations.  In any
-subsequent version of this paper, we would be happy to compare our
-work with FlashFill in greater detail.
+Many real-life file format descriptors involve large numbers of
+disjunctions and nested iterations. FlashFill struggles with
+disjunctions and does not handle nested iterations. In our previous
+work on synthesizing bijective lenses, we compared FlashFill to our
+bijective synthesis algorithm in more detail, finding that FlashFill
+was able to synthesize just 5 of the 39 examples that we had at that
+time because of nested iterations.  In future version of this paper,
+we would be happy to compare our work with FlashFill in greater
+detail.
 
@@
 
-We apologize for the lack of citations and the messy appendix. For the first
-subsection of Related Work, we got carried away in making formal claims, and did
-not include informal comparisons to other lens languages. We will make sure to
-make such comparisons and citations in the next version of the paper. We will
-similarly improve the appendix.
+We apologize for the lack of citations and the messy appendix. For the
+first subsection of Related Work, we focused on making formal claims,
+and did not include informal comparisons to other lens languages. We
+will make sure to make such comparisons and include a fuller set of
+citations in the next version of the paper. We will similarly improve
+the appendix.
 
@@
 
-Thank you, you are correct, we will change this citation.
+Thank you!  We will change this citation.
 
@@
 
-You are correct.
+You are correct; we will fix it.
 
@@
 
-Within our synthesis algorithm, you are, in general, correct. However,
-Boomerang exists beyond Optician, and permits lenses like id("a"|"a"). In
+Within our synthesis algorithm, you are correct. However, Boomerang
+includes more than Optician and permits lenses like id("a"|"a"). In
 our extension to Boomerang, we do not remove this freedom.
@@
 We were attempting to make our equations more readable, but we did not
-sufficiently explain what our syntax meant. The calls to H(...) returns pairs,
+sufficiently explain our syntax. The calls to H(...) return pairs,
 and when multiplied by constants, or added together, they act as follows:
 
-  a x (b,c) = (a x b,a x c)
+  a * (b,c) = (a * b,a * c)
   (a,b) + (c,d) = (a + b,c + d)
 
-We will make sure to explain this more clearly in a final version of the paper.
+We will explain this more clearly in a final version of the paper.
 
@@
 
-See the main response "above the fold" at the top of this response.
+See the shared section at the top of this response.
 
@@
 
-We do not believe such subproblems are straightforwardly avoidable. While we
-can attempt to match up subcomponents, this matching oftentimes will not
-work because of, for example, issues where various equivalences must be
-applied (for example: ("a"|"b")|("c"|"d") <-> "a"|("b"|("c"|"d"))) do not
-break down. If we had attempted to initially try breaking down the problem
-syntactically, we would have wasted time. This is particularly an issue
-because bad specifications could take a very long time, as there may be no
-good lenses between the attempted proposed subcomponents, so there would
-potentially be a great deal of wasted time in these attempts.
+We do not believe such subproblems are straightforwardly
+avoidable. While we can attempt to match up subcomponents, this
+matching will often not work, for example in cases where various
+equivalences must be applied (such as: ("a"|"b")|("c"|"d") <->
+"a"|("b"|("c"|"d"))). If we had attempted to break down the problem
+syntactically, we could have wasted a lot of time trying to find a
+good lens that doesn't exist.
 
@@
 All benchmarks were performed on a 2.5 GHz Intel Core i7 processor with 16
-GB of 1600 MHz DDR3 running macOS High Sierra.
+GB of 1600 MHz DDR3 running macOS High Sierra. We will add this
+information.
 
@@
 
-We chose our benchmarks from the FlashFill paper because the 238 tasks
-are, unfortunately, not publicly available.  Many real-life file
-format descriptors involve large numbers of disjunctions and nested
-iterations, which FlashFill is either bad at (for disjunctions) or
-cannot handle at all (nested iterations).
+We chose our benchmarks from the FlashFill paper because the fuller
+set of 238 tasks are not publicly available.
 
-In our previous work, we compared FlashFill to our bijective synthesis
-algorithm in more detail, finding that FlashFill was able to
-synthesize just 5 of the 39 examples that we had at that time due to
-the presence of nested iterations.  In any subsequent version of this
-paper, we would be happy to compare our work with FlashFill in greater
+Many real-life file format descriptors involve large numbers of
+disjunctions and nested iterations. FlashFill struggles with
+disjunctions and does not handle nested iterations. In our previous
+work on synthesizing bijective lenses, we compared FlashFill to our
+bijective synthesis algorithm in more detail, finding that FlashFill
+was able to synthesize just 5 of the 39 examples that we had at that
+time because of nested iterations.  In future version of this paper,
+we would be happy to compare our work with FlashFill in greater
 detail.
@@
 
-See our detailed response "above the fold."
+See our response "above the fold" for a description of our process.
+
+Without an existing lens to compare to and without an existing set of
+input/output examples, it is not obvious how to make the process of
+checking correctness more automatic and objective.
+
+We are interested in finding ways to make validating lenses easier for the user,
+for example by finding examples of puts and creates that differentiate between
+highly ranked lenses. However, we leave this task for future work.
 
@@
 
-Our graph does not include any specific order of tasks. At any given time, a
-reader can see how many of the benchmarks are able to complete within that
-time (for example, 41 benchmarks require complete in under a second in the
-full synthesis algorithm). We can reword the description of these graphs to
-make this more clear.
+Our graph does not include any specific order of tasks. At any given
+time, a reader can see how many of the benchmarks are able to complete
+within that time (for example, 41 benchmarks complete in under a
+second in the full synthesis algorithm). We can reword the description
+of these graphs to make this more clear.
 
@@
 
-We will update the lens section of the related work to be more thorough on
-the related work that we do not make formal statements about, and include
-proper citations.
+We will update the lens portion of the related work section to include
+a more thorough deription of related work that we do not make formal
+statements about, and we will include the relevant citations.
 
@@
 
-As a demonstration about the difficulty of writing lenses, consider the
+As a demonstration of the difficulty of writing lenses, consider the
 following bidirectional program
@@
 
-Even the relatively simple lens that permutes the 3 elements needed for
-single_author_convert is fairly complex.  The complexity becomes even more
-apparent when working with large lenses built from many sub-lenses.  Lenses
-provide great power with their invertibility guarantees, but they come at
-the cost of thinking about fiddly details when writing the terms.  Worrying
-about these fiddly details while ALSO thinking about unambiguity
-restrictions is a difficult task!
+Even this relatively simple lens that permutes the 3 elements needed
+for single_author_convert is fairly complex.  The complexity becomes
+even more apparent when working with large lenses built from many
+sub-lenses.  Lenses provide great power with their invertibility
+guarantees, but they come at the cost of thinking about fiddly details
+when writing the terms.  Worrying about these details while ALSO
+thinking about unambiguity restrictions is a difficult task!
 
@@
 
-Please see above, in the initial 500 words of this reponse.
+See the shared section at the top of this response.
 
@@
 
-Please see above, in the initial 500 words of this reponse.
+See the main response at the top of this response.
 
@@
 
-Please see above, in the initial 500 words of this response.
+See the shared section at the top of this response.
 

### 2018-09-21 papers/SymmetricLenses/response_popl19.md
response clean

@@
 
-These benchmarks use the full "SS" mode. The only difference between these
-benchmarks and the full mode is that they ignore termination
-parameters and use the 0, 25, and -25 parameters instead.
-
-KSF: Anders, what is the difference between full SS mode and full mode?
+These benchmarks use the full "SS" mode. The only difference between these t=0,
+t=25, and t=-25 benchmarks and the full "SS" mode is that they ignore
+termination parameters and use the 0, 25, and -25 parameters instead.
 

### 2019-01-25 papers/SymmetricLenses/main.tex
add changes to paper

@@
 Synthesizing simple symmetric lenses is substantially more challenging that
-synthesizing bijective lenses: Since some of the information on each side
-can be ``disconnected'' from the other side, there will in general be {\em
-  many} lenses that agree with a given example.  To guide the search
-process, we use {\em stochastic regular expressions} and ideas from
-information theory to estimate the amount of information propagated by a
-candidate lens.
+synthesizing bijective lenses: Since some of the information on each side can be
+``disconnected'' from the other side, there will in general be {\em many} lenses
+that agree with a given example. To guide the search process, we use {\em
+  stochastic regular expressions} and ideas from information theory to estimate
+the amount of information propagated by a candidate lens, in conjunction with
+annotations to state whether certain bits of information are irrelevant, or must
+be maintained.
 
@@
 Augeus~\cite{augeas2}, the bidirectional programming literature, and electronic
-file format synchronization tasks. After modest tuning, which is encouraged by
-the interactive nature of the tool, our implementation can synthesize all these
+file format synchronization tasks. Our implementation can synthesize all these
 lenses in under 30 seconds.
@@
 
-We added this synthesis procedure to the Boomerang implementation and
-evaluated its effectiveness on a benchmark suite of 48 problems, including
-data-cleaning tasks taken from Flash Fill~\cite{gulwani-popl-2014},
-view-maintenance tasks taken from Augeas~\cite{augeas2}, and ad-hoc
-synchronization tasks from a variety of other sources (mostly from the
-bidirectional programming literature).  The ``programmer'' can control the
-search process via examples and a {\em tuning parameter} that governs how
-much effort is spent trying to improve on existing solutions.
+We added this synthesis procedure to the Boomerang implementation and evaluated
+its effectiveness on a benchmark suite of 48 problems, including data-cleaning
+tasks taken from Flash Fill~\cite{gulwani-popl-2014}, view-maintenance tasks
+taken from Augeas~\cite{augeas2}, and ad-hoc synchronization tasks from a
+variety of other sources (mostly from the bidirectional programming literature).
+The ``programmer'' can control the search process via examples and annotations
+that informs the synthesis algorithm as to whether certain information is
+irrelevant, or must maintained.
 
@@
 \item We extend the Boomerang implementation with simple symmetric lenses,
-demonstrating that they integrate smoothly with other lens extensions such
-as quotient lenses~\cite{quotientlenses} and matching
-lenses~\cite{matchinglenses}.  We evaluate our implementation and the
-effects of various optimizations and the tuning parameter on a set of 48
-lens synthesis benchmarks drawn from data cleaning, view maintenance, and
-file synchronization tasks.  With only modest interactive tuning, the system
-can synthesize simple symmetric lenses for all of the benchmarks in under 30
-seconds (\S\ref{sec:evaluation}).
+  demonstrating that they integrate smoothly with other lens extensions such as
+  quotient lenses~\cite{quotientlenses} and matching
+  lenses~\cite{matchinglenses}. We evaluate our implementation, the effects of
+  various optimizations, and our inclusion of relevance annotations on a set of
+  48 lens synthesis benchmarks drawn from data cleaning, view maintenance, and
+  file synchronization tasks. The system can synthesize simple symmetric lenses
+  for all of the benchmarks in under 30 seconds (\S\ref{sec:evaluation}).
 %
@@
 lens(es) and to reduce the size of the search space by cutting down on the
-number of satisfying lenses, or supply a tuning parameter (discussed in 
-\S\ref{subsec:optimizations}) that adjusts how much effort is put into the search.
+number of satisfying lenses, or they can supply annotations (discussed in
+\S\ref{subsec:optimizations}) informing describing whether certain information
+should be retained through format conversion.
 
@@
 
-\subsection{Optimizations and Termination Tuning}
+\subsection{Optimizations and Relevance Annotations}
 \label{subsec:optimizations}
@@
 Our implementation includes a number of optimizations not described above:
-annotations that guide DNF conversion; an expansion inference
-algorithm; hash-consing and memoization; and compositional synthesis. The
-implementation also provides for an additional specification parameter that
-customizes the termination condition for the synthesis procedure.  The
-optimizations make the system performant enough for interactive use, and the
-tuning parameter gives the user a ``knob'' by which the algorithm can be
-directed to try harder to find a good lens.
+annotations that guide DNF conversion; an expansion inference algorithm;
+hash-consing and memoization; and compositional synthesis. The implementation
+also provides for additional specifications in the form of annotations that help
+describe which information is irrelevant to the lens. The optimizations make the
+system performant enough for interactive use, and the annotations help ensure
+the desired information is projected.
 
@@
 
-\paragraph*{Termination Customization}
+\paragraph*{Relevance Annotations}
 In this work, there are three competing constraints. The tool should (1) return
@@
 Our termination heuristic works well for most situations, but
-there are cases where a large number of expansions must be performed to find the
-correct lens.  Users can work around such a problem by providing a large number
-of examples, but constructing examples is a chore, especially when formats are
-complicated.  We have often found it simpler to allow users to change the
-synthesis termination condition to demand \SynthSymLens be more judgemental
-when it comes to deciding it has found a ``good enough'' lens.  
-Conversely, if the user's desired transformation has a particularly
-high cost, the algorithm may spend a long time searching in vain for lenses with
-lower cost. By tuning \SynthSymLens to be more permissive, such a search can be
-short-circuited.
-
-For these situations, users can provide a number $t$ that affects
-the strictness of the termination condition by adjusting the \Continue formula
-as follows. When $\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, $\Regex$ and
-$\RegexAlt$ are at
-distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
-current best lens has cost $c$,
-\begin{gather*}
-  \Call{Continue}{pq} := c > \Min(d + \Log_2(n) - t, 0).
-\end{gather*}
-With a positive $t$, the termination condition is stricter, and the algorithm
-will spend longer searching for more likely lenses. With a negative $t$, the
-termination condition is less strict, and the algorithm will spend less time
-searching.  The default $t$ is $0$.
+there are cases where it fails.
+
+In particular, our system works hard to find lenses that project \emph{as little
+  information as possible}. If the desired lens requires projecting a large
+amount of information, the termination condition will be satisfied ``too late.''
+While \GreedySynth may have found what the user knows to be the desired lens,
+the full algorithm will continue trying to find lenses that project less
+information.
+
+Conversely, the termination heuristic is satisfied algorithm cuts short: if it
+had continued searching for lenses it would have found the desired one, but the
+termination heuristic was satisfied ``too early.'' We wish for our algorithm to
+provide feedback in a timely manner, but sometimes this prohibits finding the
+desired lens.
+
+To address these issues, we allow annotations on the regular expressions
+dictating relevance. The first, $\SkipOf{\Regex}$, says the information
+content of $\Regex$ is irrelevant to the desired lens, and can the data can be
+safely projected. The second, $\SRequireOf{\Regex}$, says the information of
+$\Regex$ is critical to the desired lens, and none of the information can be
+lost when converting between the formats.
+
+If synthesizing a lens seems to be taking a long time, it is spending time
+checking it cannot find a lens that projects less information. By ``skipping''
+portions of the SRE that \emph{should} be projected, the algorithm will not
+continue searching for a better lens when it has already found the correct one.
+In practice, we implement \Skip by setting the information content of
+$\Skip(\Regex)$ to zero.
+
+If the algorithm returns a lens with too disconnects and merges, one could
+provide additional examples to constrain the lens more. However, for large
+formats, writing such examples is time consuming and error-prone. Instead, if
+one recognizes that information from $\Regex$ is incorrectly lost when converted
+between formats, they can prevent that information from being lost by annotating
+it $\SRequireOf{\Regex}$. One can guarantee a lens between \Regex and \RegexAlt
+is bijective by annotating them as $\SRequireOf{\Regex}$ and
+$\SRequireOf{\RegexAlt}$. In practice, we implement \SRequire by setting the
+cost of any lens that loses ``required'' information to infinity.
 % end synthesis
@@
 \begin{enumerate}
-\item Can the algorithm (with suitable examples and termination
-  tuning) find the correct lens?
+\item Can the algorithm (with suitable examples and annotations) find the correct lens?
 \item Is the synthesis procedure efficient enough to be used in everyday
@@
   the existing Optician system customized for bijective lenses?
-\item How effective is the information-theoretic search heuristic and how does termination tuning affect the results?
+\item How effective is the information-theoretic search heuristic and how do our
+  annotations affect the results?
 \end{enumerate}
@@
 zero or one examples, runs the tools, determines whether the output is the
-desired one and, if not, either adds an example or adjusts the termination
-metric and then tries again---it is hard to cleanly evaluate the net user
-experience.  The primary concern is whether it is usable to generate
-synthesize correct lenses.
+desired one and, if not, either adds an example or relevance annotations and
+then tries again---it is hard to cleanly evaluate the net user experience. The
+primary concern is whether it is usable to generate synthesize correct lenses.
 
@@
 interactively on all 48 tasks, working with the system to create sufficient
-examples and adjust the termination parameter.  In all cases, the desired lens
-was obtained, which we verified by manually inspecting the generated lens
-programs.
+examples and provide useful relevance annotations.  In all cases, the desired lens
+was obtained.
 %
@@
 
-Adjusting the termination parameter was needed in only 9 of the 48 tasks. In practice, we
-found that customizing the termination metric to be quite easy: if manual
-inspection of the lens showed there were too few \IdentityLens{}s, and too many
-\Disconnect{}s or merges, we would increase
-the termination parameter. If synthesis took too long, we would decrease it.
-Section~\ref{subsec:effect-of-heuristics} studies the effect of fixing the
-termination metric.
+Providing relevance annotations was needed in only 8 of the 48 tasks. In
+practice, we found that adding such annotations quite easy: if manual inspection
+of the lens showed there were too few \IdentityLens{}s, and too many
+\Disconnect{}s or merges, we would add \SRequire annotations. If synthesis took
+too long, we would add \Skip annotations.
+Section~\ref{subsec:effect-of-heuristics} studies the effects of removing such
+annotations.
+
+We verified that \SSOpt{} generated the correct lenses the way programmers often
+validate their programs: we manually inspected the code and ran unit tests on
+the synthesized code. To determine whether the synthesis procedure generated the
+correct lens when running in modes other than \SSOpt{}, we compared generated lens
+to the lens synthesized by \SSOpt{}.
 
@@
 
-Having determined appropriate examples and termination parameters for the 48
+Having determined appropriate examples and annotations for the 48
 benchmarks, we evaluate the performance of the system by measuring the running
@@
     synthesis it is able to complete 26. Each benchmark specification includes
-    source and target regular expressions, between one and three sufficient
-    examples, and, if necessary, the timing parameter determined to be
-    sufficient for correctness during a prior interactive session.}
+    source and target (potentially annotated) regular expressions, and between
+    one and three sufficient examples.}
   \label{fig:times}
@@
 39 bijective synthesis tasks.\footnote{We had to slightly alter four of these
-benchmarks, either by providing additional examples or by tuning termination parameters.
+benchmarks, either by providing additional examples or by adding in \SRequire annotations.
 Without these alterations, symmetric synthesis yielded a lens that fit
@@
               of merges.\\
-\end{tabulary}
-
-\begin{tabulary}{\linewidth}{rL}
-  \ZTOpt{}: & \ZTOpt{} runs the algorithm with the termination parameter
-              universally set to 0 (default). \\
-  \TFTOpt{}: & \TFTOpt{} runs the algorithm with the
-               termination parameter universally set to 25. \\
-  \NTFTOpt{}: & \NTFTOpt{} runs the algorithm with the
-                termination parameter universally set to -25. \\
+  \NSOpt{}: & We ignore all \Skip annotations in the SRE specifications. \\
+  \NROpt{}: & We ignore all \SRequire annotations in the SRE specifications. \\
 \end{tabulary}
@@
 
-Figure~\ref{fig:metric}
-summarizes the result of these experiments. The data reveal that 
-the information-theoretic metric is critical for finding the correct
-lens: Only 10 of the benchmarks succeeded when running in \CCOpt{} mode.
-The termination condition is also quite important. When running
-in \FLOpt{} mode, the algorithm only discovers 5 lenses, which shows
-that the first cluster that contains a satisfying lens is rarely the
-correct cluster.  However, our termination condition is not perfect.
-Sometimes users must tell the algorithm to ``keep going'' or ``stop early.''
-Without termination parameters, our algorithm finds the correct lens for 39 of
-our 48 benchmarks; nine required tuning 
-to find the correct lens. Nevertheless, our default parameter sits in a
-good middle-ground. Both \TFTOpt{} and \NTFTOpt{} synthesized the correct lens
-less often than \ZTOpt.  \TFTOpt{} demands too good a lens, so the search
-procedure keeps spinning, while \NTFTOpt{} terminates early, not
-always trying hard enough to find the correct lens.
+Figure~\ref{fig:metric} summarizes the result of these experiments. The data
+reveal that the information-theoretic metric is critical for finding the correct
+lens: Only 10 of the benchmarks succeeded when running in \CCOpt{} mode. The
+termination condition is also quite important. When running in \FLOpt{} mode,
+the algorithm only discovers 5 lenses, which shows that the first cluster that
+contains a satisfying lens is rarely the correct cluster. However, our algorithm
+is not perfect and fails when either it is very difficult to find the desired
+lens (necessitating \SRequire) or when a large amount of data is projected
+(necessitating \Skip). Without any annotations, our algorithm finds the correct
+lens for 40 of our 48 benchmarks; eight required relevance annotations to find
+the correct lens.
 % end evaluation
@@
     found, that lens is returned. \CCOpt{} synthesizes lenses, where the cost of
-    a lens is the number of disconnects plus the number of merges. \ZTOpt{} runs
-    the algorithm with the termination parameter universally set to 0 (default).
-    \TFTOpt{} runs the algorithm with the termination parameter universally set
-    to 25. \NTFTOpt{} runs the algorithm with the termination parameter
-    universally set to -25.}
+    a lens is the number of disconnects plus the number of merges. \NSOpt
+    ignores all \Skip annotations while running the algorithm. \NROpt ignores
+    all \SRequire annotations while running the algorithm.}
   \label{fig:metric}

### 2019-01-27 papers/SymmetricLenses/main.tex
minor fixes

@@
 
-Synthesizing simple symmetric lenses is substantially more challenging that
+Synthesizing simple symmetric lenses is substantially more challenging than
 synthesizing bijective lenses: Since some of the information on each side can be
@@
 the amount of information propagated by a candidate lens, in conjunction with
-annotations to state whether certain bits of information are irrelevant, or must
-be maintained.
+annotations to state whether certain bits of information are irrelevant to the
+lens, or must be maintained.
 
@@
 The ``programmer'' can control the search process via examples and annotations
-that informs the synthesis algorithm as to whether certain information is
-irrelevant, or must maintained.
+that informs the synthesis algorithm as to whether certain bits of information
+are irrelevant, or must maintained.
 

### 2019-01-28 papers/SymmetricLenses/main.tex
Some polishing on the abstract

@@
 {\em Lenses} are programs that can be run both ``front to back'' and ``back
-to front,'' allowing data and updates to be transformed in two directions.
+to front,'' allowing updates to both input and output data to be transferred
+in both directions.
 Since their 
 introduction by~\citet{Focal2005-long}, lenses have been extensively
-studied and applied.  Recent work~\cite{optician} has also demonstrated how
+studied, extended, and applied.  Recent work~\cite{optician} has also demonstrated how
 techniques from {\em type-directed program synthesis} can be used to
-efficiently synthesize a very simple class of lenses---so-called {\em
+efficiently synthesize a simple class of lenses---so-called {\em
   bijective lenses} over string data---given a pair of types (regular
@@
 We show how to extend this synthesis algorithm to a much wider class of
-lenses, which we call {\em simple symmetric lenses}, including both
-bijective lenses and the more widely used ``asymmetric''
-lenses~\cite{Focal2005-long}, as well as a rich subset of the full-blown
-``symmetric lenses'' proposed by \citet{symmetric-lenses}.  Simple symmetric
-lenses are of independent theoretical interest, being the largest class of
-symmetric lenses that do not rely on persistent internal state.
+lenses, which we call {\em simple symmetric lenses}, including not only all
+of the bijective lenses and the more widely used ``asymmetric''
+lenses~\cite{Focal2005-long}, but also a rich subset of the full-blown
+``symmetric lenses'' proposed by \citet{symmetric-lenses}.
 
 Synthesizing simple symmetric lenses is substantially more challenging than
-synthesizing bijective lenses: Since some of the information on each side can be
-``disconnected'' from the other side, there will in general be {\em many} lenses
-that agree with a given example. To guide the search process, we use {\em
-  stochastic regular expressions} and ideas from information theory to estimate
-the amount of information propagated by a candidate lens, in conjunction with
-annotations to state whether certain bits of information are irrelevant to the
-lens, or must be maintained.
+synthesizing bijective lenses: Since some of the information on each side
+can be ``disconnected'' from the other side, there will in general be {\em
+  many} lenses that agree with a given example. To guide the search process,
+we use {\em stochastic regular expressions} and ideas from information
+theory to estimate the amount of information propagated by a candidate lens,
+in conjunction with user annotations identifying parts of the source and
+target data structures as either {\em irrelevant} or {\em essential}.
+
+Simple symmetric lenses are also of independent theoretical interest, being
+the largest class of symmetric lenses that do not rely on persistent
+internal state.
 
@@
 \noindent
-In today's big-data world, similar information is sometimes
+In today's big-data world, similar information is often
 stored in different places and in different formats. For instance, electronic
@@
 
-One convenient way to maintain consistency between these varied data formats
+One convenient way to maintain consistency between such varied data formats
 is to use a \emph{lens}~\cite{Focal2005-long}---a bidirectional program that
@@
   hideous and hard to navigate!!}
-\ifappendices\else included with the supplemental material\fi.
+\ifappendices\else included with the supplemental material\bcp{Right?}\fi.
 

### 2019-01-28 papers/SymmetricLenses/main.tex
Nit

@@
 can make the task even easier.  Indeed, lens languages offer excellent
-targets for synthesis: their specialized, sub-Turing-complete syntax
+targets for search-based synthesis, since their specialized, sub-Turing-complete syntax
 and expressive type systems drastically
-reduce the space of possible programs, making search-based synthesis easier. 
+limit the space of possible programs. 
 %

### 2019-01-28 papers/SymmetricLenses/main.tex
Minor polishing

@@
-% \newif\ifdraft\drafttrue  % set true to show comments
-\newif\ifdraft\draftfalse  % set true to show comments
+\newif\ifdraft\drafttrue  % set true to show comments
+% \newif\ifdraft\draftfalse  % set true to show comments
 \newif\ifanon\anonfalse    % set true to suppress names, etc.
@@
 
-% BCP: To make sure we don't leave comments on!
-\bcp{XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX }
-
 {\em Lenses} are programs that can be run both ``front to back'' and ``back
@@
 in both directions.
-Since their 
-introduction by~\citet{Focal2005-long}, lenses have been extensively
+Since their introduction by~\citet{Focal2005-long}, lenses have been extensively
 studied, extended, and applied.  Recent work~\cite{optician} has also demonstrated how
@@
   bijective lenses} over string data---given a pair of types (regular
-expressions) and a small number of examples.
+expressions) and a small number of examples.  \bcp{Why is the next line not indented?}
 
@@
 
-While domain-specific languages for writing lenses can facilitate building
-principled data transformations, \textit{synthesizing} lenses 
+Domain-specific languages for writing lenses can facilitate building
+principled data transformations,
+\textit{Synthesizing} lenses 
 can make the task even easier.  Indeed, lens languages offer excellent

### 2019-01-28 papers/SymmetricLenses/main.tex
fix nits

--- 2019-01-28 papers/SymmetricLenses/main.tex
@@
 %
-Recent work~\cite{optician} has exploited this observation, demonstrating that
+Recently, Optician~\cite{optician} has exploited this observation, demonstrating that
 it is feasible to synthesize quite complex examples, albeit only for a
@@
 file on another operating system does not. 
-Indeed of the 48 benchmark problems described in Section \ref{sec:evaluation},
+Indeed among the 48 benchmark problems described in Section \ref{sec:evaluation},
 all of the benchmarks taken from Flash Fill~\cite{flashfill} and many of the
@@
 expressions equivalent to a given one, and the lens returned by a
-type-directed search will depend which of the infinitely many possible
+type-directed search will depend on which of the infinitely many possible
 representations are chosen for its source and target formats.  Moreover,
@@
 
-This definition differs from classic symmetric lenses in that it does not
+This definition differs from classical symmetric lenses in that it does not
 involve a \emph{complement}.  We give a detailed comparison in
@@
 stochastic regular expressions; we mark plain REs with overbars.)  When
-clear from context, we shoren ``simple symmetric lens'' to just
+clear from context, we shorten ``simple symmetric lens'' to just
 ``lens.''
@@
 and the put functions merely return the first argument
-($\PutROf{s_1}{s_2} = s_2$).
+($\PutROf{\String}{\StringAlt} = \String$).
 %
@@
 \begin{lstlisting}
-disconnect(salary, "", "unk", "") : salary $\Leftrightarrow$ company
+disconnect(salary, "", "unk", "") : salary $\Leftrightarrow$ ""
 \end{lstlisting}
@@
 On creates, the input values are thrown away, and default values are
-returned ($\CreateROf{\String} =
+returned ($\CreateLOf{\StringAlt} =
 \lstinline{"unk"}$), and on puts, the second argument is used and the first is
@@
 will transform ``\lstinline{Jane Doe: }'' to ``\lstinline{Jane Doe,}'' in
-the forward direction.
+the forward direction (converting from the left format to the right).
 \bcp{We don't actually give a nontrivial example of how disconnect works.}
@@
 %
-is another valid lens (where \lstinline{rep($\BRegex$)} denotes some string in the
+is another well-typed lens (where \lstinline{rep($\BRegex$)} denotes some string in the
 language of $\BRegex$). Furthermore, adding the input-output example in
@@
 
-The expected number of bits required to recover data is the well-known information-theoretic concept 
-of the \emph{entropy} of the data~\cite{Shannon1948}.
-Calculating entropy requires a probability distribution over the
-space of possible values for the data. Specifically, given a set $S$ and a
-probability distribution $P \OfType S \rightarrow \Reals$ over $S$, the
-entropy is $-\Sigma_{s\in  S}P(s)\Log_2P(s)$.
+The expected number of bits required to recover data is the well-known
+information-theoretic concept of the \emph{entropy} of the
+data~\cite{Shannon1948}. Calculating entropy requires a probability distribution
+over the space of possible values for the data. Specifically, given a set $S$
+and a probability distribution $P \OfType S \rightarrow \Reals$ over $S$, the
+entropy $\EntropyOf{S,P}$ is $-\Sigma_{s\in S}P(s)\Log_2P(s)$.
 
@@
 names to first names and last names to last names). \GreedySynth
-converts yields a simple symmetric lens by converting the n-ary
+yields a simple symmetric lens by converting the n-ary
 forms into the binary forms provided in the surface language.
@@
 
-Our algorithm must resolve a three-way tension between the quality of
+Our algorithm must resolve a three-way tension among the quality of
 the inferred lens, the number of examples the user must provide to
@@
 other data source. To do this, we must first develop a probabilistic language
-model for our language. We do this with \emph{stochastic regular expressions},
+model for our language. We do this with \emph{stochastic regular expressions}~\cite{stoch-rnn,stoch-def},
 regular expressions annotated with probability information. With this, we can
@@
 $\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation $\Probability$
-represents the likelihood the string is in $\Regex_1$. In the stochastic regular
-expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability how
+represents the likelihood that the string is in $\Regex_1$. In the stochastic regular
+expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability
+represents how
 much more likely having a string involving more than $n$ iterations is than
@@
 $\Expand$ needs to be able to find equivalent stochastic regular expressions.
-However, existing work does not provide a means to reason about stochastic
+However, existing work does not provide a means of reasoning about stochastic
 regular expression equivalence. We extend the
@@
   The proofs of this theorem, and subsequent theorems in this paper, appears in
-  the appendix present in the full version of the paper accompanying this
+  the appendix presented in the full version of the paper accompanying this
   submission.
@@
 The entropy of a data source $S$ is the expected number of bits required to
-describe a random element in $S$ ($-\Sigma_{s\in S}P(s)\Log_2P(s)$). The entropy
-of a stochastic regular expression can be computed directly from its syntax,
-when the stochastic regular expression is unambiguous and contains no empty
-subcomponents.
+describe a random element in $S$ (formally $-\Sigma_{s\in S}P(s)\Log_2P(s)$). The
+entropy of a stochastic regular expression can be computed directly from its
+syntax, when the stochastic regular expression is unambiguous and contains no
+empty subcomponents.
 \begin{center}
@@
     & =
-    & 1\\
+    & 0\\
 
@@
 We have proven that this syntactic computation corresponds to the entropy of the
-stochastic regular expression.
+stochastic regular expression when the stochastic regular expression is
+unambiguous (parses strings uniquely).
 
@@
 method we have given is certainly incorrect).  For example, consider the SRE
-$\PRegexOr{\lstinline{"a"}}{\lstinline{"a"}}{.5}$. Using the above computation,s
+$\PRegexOr{\lstinline{"a"}}{\lstinline{"a"}}{.5}$. Using the above computation,
 this SRE has an entropy of $1$, but the entropy should be $0$. These
 difficulties become particularlly complex when the languages are ambiguous in
-subtle ways, the algorithm must be correct when only some of the two languages
+subtle ways and the algorithm must be correct when only some of the two languages
 overlap. Similar issues occur when trying to find the entropy when $\emptyset$
@@
 
-We now provide the typing derivations and semantics for these lenses. We do
+We now provide the typing rules and semantics for these combinators. We do
 not include the semantics for combinators equivalent to those presented in
@@
     \CreateR{} \App s & = & s\\
-    \PutR{} \App s_1 \App s_2 & = & s_1
+    \PutR{} \App s \App t & = & s
   \end{array}$
@@
 The \IterateLens lens deals with iterated data, while inverting reverses the direction of a lens: creating on the right becomes creating on the left and vice versa, and putting on the right becomes putting on the left and vice versa.
-The invert combonator is particularly useful when chaining many compositions together, as it can be used to align the central types.
+The invert combinator is particularly useful when chaining many compositions together, as it can be used to align the central types.
 \[
@@
 Type equivalence enables a lens of type $S \Leftrightarrow T$ to be used as a
-lens of type $S' \Leftrightarrow T'$ if $S$ equivalent to $S'$ and $T$ is
+lens of type $S' \Leftrightarrow T'$ if $S$ is equivalent to $S'$ and $T$ is
 equivalent to $T'$. Type equivalence is useful both for addressing type
@@
 expected amount of information required to recover a string in $\Regex$ from a
-string in $\RegexAlt$, synchronized by $\Lens$.
+string in $\RegexAlt$, synchronized by $\Lens$. We use the notation $a(b,c)$ to
+mean $(a*b,a*c)$, and the notation $(a,b)+(c,d)$ to mean $(a+c,b+d)$.
 \begin{center}
@@
     & =
-    & $(0,\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1}+\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2}+1)$\\
+    & $(0,\REntropyOf{\RegexAlt_1 \Given \Lens_1, \Regex}+\REntropyOf{\RegexAlt_2 \Given \Lens_2, \Regex}+1)$\\
 
@@
 typically calculate the conditional entropy exactly, as merges are relatively
-uncommon on practice; only 2 of the lenses sythesized in our benchmarks
+uncommon in practice; only 2 of the lenses sythesized in our benchmarks
 include merges.
@@
 \MaxOf{\LEntropyOf{\Regex \Given \Lens, \RegexAlt}} +
-\MaxOf{\REntropyOf{\RegexAlt \Given \Lens, \Regex}}$ is the maximum sum of
-recovering the left format from the right, and the right from the left. We have
+\MaxOf{\REntropyOf{\RegexAlt \Given \Lens, \Regex}}$ is the sum of the maximum
+expected number of bits to
+recover the left format from the right, and the right from the left. We have
 proven theorems demonstrating the calculated entropy corresponds to the actual
@@
 
-Note that our definition of $\Entropy$ contains no case for sequential
+Note that our definition of $\Entropy^\rightarrow$ contains no case for sequential
 composition $\Lens_1 ; \Lens_2$ and our theorem excludes lenses that contain
@@
     & \Rewrite
-    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability}\\
+    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{1-\Probability}\\
 
@@
     & \Rewrite
-    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{\Regex{}}{\Probability}}{\Regex{}})}{\Probability}
+    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{\Regex{}}{\Probability}}{\Regex{}})}{1-\Probability}
   \end{tabular}
@@
 expressions whose DNF form gives equal probability to all sequence subcomponents
-of the DNF regular expression., and gives iterations a .8 chance to continue
+of the DNF regular expression, and gives iterations a .8 chance to continue
 iterating. In our experience generating random strings from regular expressions,
@@
 \paragraph*{Entropy}
-We have developed a syntactic means of finding the entropy of a stochastic DNF
+We have developed a syntactic means for finding the entropy of a stochastic DNF
 regular expression, like we have for stochastic regular expressions. This
-enables us to efficiently find the entropy without first converting a DNF SRE to
+enables us to efficiently find the entropy without first converting a \SDNFREabbrev to
 a stochastic regular expression.
@@
 many DNF lenses of a given type (though there are still many more symmetric DNF
-lenses than there are DNF bijective lenses).
+lenses than DNF bijective lenses).
 
@@
 symmetric atom lenses, $sal_1 \ldots sal_n$, and a symmetric atom lens $sal$ is
-an iteration of a symmetric DNF lens.
+an iteration of a symmetric DNF lens. Note that the structure of symmetric DNF
+lenses exactly mirrors that of SDNF REs.
 
@@
 The inputs to \GreedySeqSynth are a suite of input-output examples and a pair of
-stochastic atoms. As in \GreedySynth, \GreedySeqSynth returns \None early if
+lists of stochastic atoms. As in \GreedySynth, \GreedySeqSynth returns \None early if
 there is no possible lens. Afterward, \GreedySeqSynth finds the best lenses
@@
 \paragraph*{Expansion Inference} These additional rewrites make the search
-through possible regular expressions harder. Our algorithm 
-identifies when certain closed regular expressions can \emph{only} be
-involved in a disconnect lens.  Such regular expressions will automatically be
-opened.
+through possible regular expressions harder. Our algorithm identifies when
+certain closed regular expressions can \emph{only} be involved in a disconnect
+lens (unless opened). Such regular expressions will automatically be opened.
 
@@
 combinators  using a combination of the symmetric lens combinators presented in this
-paper and symmetric versions of the asymmetric extensions (like matching lenses and quotient lenses)
+paper and symmetric versions of the asymmetric extensions (like matching lenses~\cite{matchinglenses} and quotient lenses\cite{quotientlenses})
 already present in Boomerang.
@@
 the synthesis of new lenses.
+
+All benchmarks were performed on a 2.5 GHz Intel Core i7 processor with 16
+GB of 1600 MHz DDR3 running macOS High Sierra.
 
@@
 Because our tool is intended to be interactive---the user typically starts with
-zero or one examples, runs the tools, determines whether the output is the
+zero or one examples, runs the tool, determines whether the output is the
 desired one and, if not, either adds an example or relevance annotations and
 then tries again---it is hard to cleanly evaluate the net user experience. The
-primary concern is whether it is usable to generate synthesize correct lenses.
-
-To determine whether the system can synthesize the desired lenses, we ran it
+primary concern is whether it is usable to synthesize correct lenses.
+
+To determine whether the system can synthesize desired lenses, we ran it
 interactively on all 48 tasks, working with the system to create sufficient
@@
 allows our system to scale to arbitrarily large and complex formats; measuring
-it shows of the responsiveness of the system when used as
+it shows the responsiveness of the system when used as
 intended. \SSNCOpt{} mode, which synthesizes a complete lens all at once,
@@
 \paragraph*{Symmetric Lenses}
-A classical symmetric lens $\Lens$ between $X$ and $Y$ consists of 4 components:
+A classical symmetric lens\cite{symmetric-lenses} $\Lens$ between $X$ and $Y$ consists of 4 components:
 a complement $C$, a designated element $init \in C$, and two functions,
@@
 Two classical symmetric lenses are equivalent if they output the same formats
-given any sequence of edits. Formally, given a lens $\Lens \in X \leftrightarrow
-Y$, an \emph{edit} for $\Lens$ is a member of $X + Y$. Consider the function
+given any sequence of edits. Formally, given a lens $\Lens$ between $X$ and $Y$, an \emph{edit} for $\Lens$ is a member of $X + Y$. Consider the function
 $apply$, which, given a lens and an element of that lens's complement, is a
 function from sequences of edits to sequences of edits. If $apply(\Lens,c,es) =
-es'$, then given complement $c$ and edits $es_i$, the lens $\Lens$ generates
+es'$, then given complement $c$ and edit $es_i$, the lens $\Lens$ generates
 $es'_i$.
@@
 
-\paragraph*{Relation with Simple Symmetric Lenses} To compare classic and simple
-symmetric lenses, we define an $apply$ function on simple lenses as well. If
+\paragraph*{Relation with Simple Symmetric Lenses} To compare classical and simple
+symmetric lenses, we define an $apply$ function on simple symmetric lenses as well. If
 $apply(\Lens,\None,es) = es'$, then starting with no prior data, after edit
@@
 
+\paragraph*{Other Lens Formulations}
+Bidirectional programming has an extensive literature, with many extensions onto
+basic lens formulations, like quotient lenses\cite{quotientlenses} and matching
+lenses\cite{matchinglenses}. Interested readers can consult
+\citet{DBLP:conf/icmt/CzarneckiFHLST09} for a slightly dated survey of lenses
+and other lens-like structures.
+
 \subsection{Data Transformation Synthesis}
@@
     \CreateL{} \App s & = & s\\
-    \PutR{} \App s_1 \App s_2 & = & s_1\\
-    \PutL{} \App s_1 \App s_2 & = & s_1\\
+    \PutR{} \App s \App t & = & s\\
+    \PutL{} \App s \App t & = & s\\
   \end{array}

### 2019-01-28 papers/SymmetricLenses/main.tex

@@
 
-While domain-specific languages for writing lenses can facilitate building
-principled data transformations, \textit{synthesizing} lenses 
+Domain-specific languages for writing lenses can facilitate building
+principled data transformations,
+\textit{Synthesizing} lenses 
 can make the task even easier.  Indeed, lens languages offer excellent
-targets for synthesis: their specialized, sub-Turing-complete syntax
+targets for search-based synthesis, since their specialized, sub-Turing-complete syntax
 and expressive type systems drastically
-reduce the space of possible programs, making search-based synthesis easier. 
+limit the space of possible programs. 
 %

### 2019-01-28 papers/SymmetricLenses/main.tex
Polishing

@@
 \noindent
-In today's big-data world, similar information is often
+In today's data-rich world, similar information is often
 stored in different places and in different formats. For instance, electronic
@@
 between different representations.
-
 Domain-specific languages for writing lenses can facilitate building
-principled data transformations,
+principled data transformations.
+
 \textit{Synthesizing} lenses 
-can make the task even easier.  Indeed, lens languages offer excellent
+can make the task even easier.  Indeed, lens languages are excellent
 targets for search-based synthesis, since their specialized, sub-Turing-complete syntax
@@
 %
-Recently, Optician~\cite{optician} has exploited this observation, demonstrating that
+In particular, the Optician~\cite{optician} system has demonstrated that
 it is feasible to synthesize quite complex examples, albeit only for a
@@
 source and target formats plus a small number of
-examples of corresponding data, the bijective synthesis algorithm will find a
-bijective lens matching the examples, if one exists. Often,
-it can find an appropriate lens in less than a second with
+examples of corresponding data, Optician's synthesis algorithm is guaranteed
+to find a bijective lens matching the examples, if one exists. Indeed, 
+it can often find an appropriate lens in less than a second with
 no examples at all.
@@
 %   third person, even if that's a bit awkward, but otherwise citing it is OK.}
-One reason that bijective lens synthesis is so effective, even relative to
+One reason that bijective lens synthesis can be so effective, even relative to
 successful synthesis tools in other domains, is that there are typically not
-very many bijections between related data formats. 
-Hence, if the synthesis algorithm can
-find any bijection at all, it is likely to be the right one.
-
-However, many real-world data formats share just part of their information
-content with other representations.  For instance, two related
-database tables might share four out of five columns, but each might have a
-different fifth column. 
-Or, one ad-hoc system configuration file might include some format-specific meta
-data, such as a date or a reference number, whereas the same configuration
-file on another operating system does not. 
-Indeed among the 48 benchmark problems described in Section \ref{sec:evaluation},
-all of the benchmarks taken from Flash Fill~\cite{flashfill} and many of the
-benchmarks that synchronize between two ad hoc file formats have this
-characteristic, 13 benchmarks in total. Consequently, synthesizing only strict
-bijections is often not useful in practice.
-
-Could we extend synthesis tools to more kinds of lenses?  In particular,
-what about \emph{symmetric lenses}~\cite{symmetric-lenses}, a class that
-includes bijective lenses, ``asymmetric'' lenses where the transformation
-from $S$ to $T$ can throw away information that must then be restored when
-returning from $T$ to $S$, and even more flexibile transformations that
-allow each side to retain information not present in the other?
+very many bijections between a given pair of data formats. 
+If the synthesis algorithm can
+find any bijection at all, it is fairly likely to be the intended one.
+
+However, the set of real-world use cases for bijective lenses is also
+somewhat limited.  A more common situation is where two real-world data
+formats share just {\em some} of their information content.  For instance,
+two related database tables might share four out of five columns, but each
+might have a different fifth column.  Or, one ad-hoc system configuration
+file might include some format-specific meta data, such as a date or a
+reference number, whereas the same configuration file on another operating
+system does not.  Indeed among the 48 benchmark problems described in
+Section \ref{sec:evaluation} below, all of the benchmarks taken from Flash
+Fill~\cite{flashfill} and many of the benchmarks that synchronize between
+two ad hoc file formats have this characteristic, 13 benchmarks in
+total.
+
+Can Optician's basic synthesis procedure be extended to a richer class of
+lenses? In particular, what about \emph{symmetric
+  lenses}~\cite{symmetric-lenses}---a much larger class that includes
+bijective lenses, ``asymmetric'' lenses (where the transformation from a
+source structure $S$ to a target $T$ can throw away information that must
+then be restored when transferring from $T$ to $S$), and even more flexibile
+transformations that allow each side to retain information not present in
+the other?
 

### 2019-01-29 papers/SymmetricLenses/main.tex
Polishing on intro and beginning of 2

@@
 %
-In particular, the Optician~\cite{optician} system has demonstrated that
+In particular, the Optician~\cite{optician} algorithm\bcp{there's a
+  potential for confusion between ``Optician'' and ``Boomerang''---are they
+  distinct systems?  Is one part of the other?  Is one a system and one just
+  an algorithm?  It's not very clear.} has demonstrated that
 it is feasible to synthesize quite complex examples, albeit only for a
@@
 enough in practice, we add simple symmetric lenses to the Boomerang
-language~\cite{Boomerang} and apply them to a range of real-world
+language~\cite{Boomerang}\bcp{this seems to assume that readers already know what
+  Boomerang is} and apply them to a range of real-world
 applications, showing in the process that simple symmetric lenses integrate
@@
 
-However, we are still left with a second difficulty with synthesizing
-symmetric lenses: While the number of bijective lenses between two related
-formats is typically tiny, the number of simple symmetric lenses is
-typically huge.  Thus, if a na\"ive search algorithm selects the first
-simple symmetric lens it finds, this will generally \emph{not} be the one we
-want.  We need a new principle for identifying ``more likely'' lenses and a
-more sophisticated synthesis algorithm that uses this principles to search
-the space more intelligently.
-
-For these, we turn to information theory. 
+This leaves us with just the second difficulty with synthesizing symmetric
+lenses: While the number of bijective lenses between two given formats is
+typically tiny, the number of simple symmetric lenses is typically enormous.
+If a na\"ive search algorithm just selects the first simple symmetric lens
+it finds, this will generally \emph{not} be the one the user wanted.  We
+need a new principle for identifying ``more likely'' lenses and a more
+sophisticated synthesis algorithm that uses this principle to search the
+space more intelligently. For these, we turn to information theory.
+
 We consider a ``likely'' lens to be one that propagates ``a lot'' of
 information from source to 
-target and vice versa. Conversely, an unlikely lens requires a
+target and vice versa. Conversely, an unlikely lens is one that requires a
 large amount of additional information to recover the target given a
-source.  
+source.\bcp{I'm nervous about the way we're talking about source and target
+  here.  For one thing, even the words ``source'' and ``target'' connote a
+  directionality that is misleading here.  And even ignoring that issue, a
+  lot of the discussion (like here) sounds as though we are talking about
+  unidirectional transformations.}  
 This categorization follows from the intuition
 that users will typically want to synchronise formats with lots
-of shared information, and so lenses that propagate
-more information are preferable to ones that propagate less.
+of shared information\bcp{really?  It's hard to be certain without lots of
+  practical experience, but a priori it seems equally reasonable for users
+  to want to pull little bits of shared information out of huge structures
+  and keep them in sync.  But maybe we can weaken the claim and just say
+  that we have an intuition that this will at least be {\em one} large class
+  of application scenarios, and this is the one we focus on}, and so lenses
+that propagate more information are preferable to ones that propagate less.
 %
@@
 probability distribution, we can calculate the shared information content of
-two formats\bcp{It's not obvious at this point what the {\em shared}
-  information content has to do with anything.} and our measure of likelihood.
+two formats and our measure of likelihood.
 
@@
 expressions equivalent to a given one, and the lens returned by a
-type-directed search will depend on which of the infinitely many possible
+type-directed search will in general depend on which of the possible
 representations are chosen for its source and target formats.  Moreover,
 there may be {\em no} lens connecting the 
-source and target formats selected by the user: in general, the synthesis
+source and target formats selected\bcp{until we get past the colon, it is
+  not clear that we are talking here about formats {\em not} modulo
+  equivalence.} by the user: in general, the synthesis 
 algorithm has to search to find equivalent ones for which a suitable lens
@@
 lens between them, returning the lens and its likelihood score
-to \Expand.  The whole synthesis algorithm terminates when comparing the
+to \Expand.  The whole synthesis algorithm terminates when a comparison of the
 likilihood score of the best lens found so far with 
-the complexity of newly generated SREs 
+the complexity\bcp{what does it mean to compare likelihood with complexity?}
+of newly generated SREs  
 suggests that further searching is unlikely to be fruitful.
@@
 
-We added this synthesis procedure to the Boomerang implementation and evaluated
+We have added this synthesis procedure to the Boomerang system and evaluated
 its effectiveness on a benchmark suite of 48 problems, including data-cleaning
@@
 variety of other sources (mostly from the bidirectional programming literature).
-The ``programmer'' can control the search process via examples and annotations
+The ``programmer''\bcp{user?} can control the search process via examples and annotations
 that informs the synthesis algorithm as to whether certain bits of information
@@
 
-In summary, our contributions are:\bcp{The section references do not proceed
-  in order, so this doesn't work very well as an outline of the paper.
-  Should we try to reorder the contributions?  Write a separate outline?
-  Live with it as-is?}\afm{Personally, I don't see it as a problem.  It's
-  nice when the contributions mirror the paper, I am not jarred when they
-  don't.}\bcp{But then the reader has no birds-eye view of the paper at
-  this point.  I tried reorganizing, which I think also helps clarify the
-  contributions.} 
+In summary, our contributions are:
 \begin{itemize}
@@
 novel application of stochastic regular expressions to guide the search
-process (\S\ref{sec:synthesis})\bcp{``and prove it correct''???}.  
+process (\S\ref{sec:synthesis})\bcp{``and prove it correct''!  (right?)}.  
 %
@@
 way, we show that {\em star-semiring equivalences} on regular expressions
-remain valid when applied to stochastic regular expressions, preserving
+remain valid when extended to stochastic regular expressions, preserving
 probability distributions as well as languages (\S\ref{sec:sre}).
@@
 fill in default values when introducing new data (\EG, the ``unk'' salary
-entry is added alongside the name to the management file when HR inserts a
+entry is added\bcp{is added by what?  A create function, I guess, but the
+  connection is not clear} alongside the name to the management file when HR inserts a
 new employee). The two \KW{put} functions propagate edits from one format to
@@
 value from one side (the result of a put or create) back through a lens in
-the other direction will get back to exactly where we began.
+the other direction will get back to exactly where we began.  \bcp{That
+  paragraph seems likely to be quite opaque to any reader that is not
+  already familiar with (symmetric) lenses.  How about a concrete example?}
 
@@
 
-Generally, we do not expect programmers to define simple symmetric lenses by
+We do not expect programmers to define simple symmetric lenses by
 writing four functions by hand and showing they satisfy the round-tripping

### 2019-01-29 papers/SymmetricLenses/main.tex
Checkpoint

@@
     important to try it and report on how it performs relative to what we
-    propose here...)}  }
+    propose here...)}\afm{Oh sorry, yes we can write such examples.
+I personally found that typically the same information could be provided through a series of 2 creates (aka: if we want a disconnect: creater a = b; creater a' = b:  if we want an identity: creater a = a; creater a' = a').
+However, we also permit putr and putl examples of the proposed form in the tool
+Much like other aspects of the work, any situation with examples ultimately only constrains the subcomponents of the lens corresponding to that specific parse tree (aka: if I give an example about A in A|B, it only constrains "A").
+I don't think it removes the need for the information theory, as all the other non-constrained subcomponents of the lens can still be "anything."
+}  }
   \label{fig:minimized-representations}
@@
 Semantically, a simple symmetric lens between sets $X$ and
-$Y$ comprises four functions that collectively satisfy four
+$Y$ comprises four functions subject to four
 ``round-tripping'' laws:
@@
 \vskip 1.5ex
+\begin{center}
 \begin{minipage}[c]{.3\textwidth}
@@
 \end{minipage} 
-\begin{minipage}[c]{.6\textwidth}
+\hspace{.4in}
+\begin{minipage}[c]{.5\textwidth}
 \begin{align}
@@
 \end{minipage} 
+\end{center}
 \vskip 1ex
-\bcp{I want less space between the put/create laws and their names.}
+\bcp{Are we going to use the numbers next to the laws??  Seems like not, so
+  why have them at all?}
 \noindent
@@
 
-We do not expect programmers to define simple symmetric lenses by
-writing four functions by hand and showing they satisfy the round-tripping
-laws.  Instead, we provide a set of combinators---a domain-specific
-language---for building complex lenses from simpler ones.  Such languages
+In principle, programmers can define simple symmetric lenses just by
+writing four arbitrary functions by hand and showing they satisfy the round-tripping
+laws.  But a better idea is to streamline the process by providing a set of
+combinators---a domain-specific 
+language---for building law-abiding complex lenses from simpler ones.  Such
+bidirectional languages
 have been given for building lenses over many different kinds of data,
@@
 $\Lens$ is a simple symmetric lens between $\LanguageOf{\BRegex}$ and
-$\LanguageOf{\BRegexAlt}$. (Later, we will use undecorated variables for
+$\LanguageOf{\BRegexAlt}$. (We will use undecorated variables later for
 stochastic regular expressions; we mark plain REs with overbars.)  When
@@
 
-In contrast, the \Disconnect lens does not propagate any edits
+In contrast, the \Disconnect lens does not propagate any edits at all
 from one format to the other.  It takes four arguments: two
 regular expressions and two strings. The regular expressions specify the formats
-of the corresponding data on the two sides while the strings provide default
-values. 
+on the two sides, while the strings provide default values. 
 %
@@
 
-The insert \lstinline{ins} and delete \lstinline{del} lenses are
+The insert lens \lstinline{ins} and the delete lens \lstinline{del} are
 syntactic sugar for uses of the disconnect lens in which a string
-constant is omitted entirely from the source and target formats,
-respectively.
+constant is omitted entirely from the source or target format.
 \begin{lstlisting}
@@
 \end{lstlisting}
-
-Finally, there are a number of combinators that compositionally construct
+The \lstinline{ins} lens inserts a constant string when going from
+left to right, while \lstinline{del} inserts a string when going from right
+to left.
+
+Finally, there are a number of combinators that construct
 more complex lenses from simpler ones. These include concatenation
-($\ConcatLensOf{\Lens_1}{\Lens_2}$ or $\Lens_1 . \Lens_2$), variation
+($\ConcatLensOf{\Lens_1}{\Lens_2}$, often written $\Lens_1 . \Lens_2$), variation
 ($\OrLensOf{\Lens_1}{\Lens_2}$), and iteration ($\IterateLensOf{\Lens}$).
@@
 will transform ``\lstinline{Jane Doe: }'' to ``\lstinline{Jane Doe,}'' in
-the forward direction (converting from the left format to the right).
+the left-to-right direction.
 \bcp{We don't actually give a nontrivial example of how disconnect works.}
@@
     pretty sure there was an error in name\_lens: the thing between the two
-    names should be id(" ") (was ins('','')).  Crossing my fingers that I'm
+    names should be id(" ") (it was ins('','')).  Crossing my fingers that I'm
     making a correct fix...}}
@@
 
-We provide typing rules for each of the simple symmetric combinators, most of
-which are syntax directed. However, regular expressions have a number of
+We provide typing rules for the simple symmetric combinators, most of
+them syntax directed. However, regular expressions have a number of
 equivalences on them, and sometimes we must convert one regular expression type
-to another, equivalent, regular expression type. Our typing rules include type
+to another, equivalent, regular expression type. \bcp{The next sentence is
+  dense.  Just show the rule?}Our typing rules include type
 equivalence, so if a lens has type $\BRegex \Leftrightarrow \BRegexAlt$, then it
 is also has type $\BRegex' \Leftrightarrow \BRegexAlt'$, where $\BRegex$ is
-equivalent to $\BRegex'$, and $\BRegexAlt$ is equivalent to $\BRegexAlt'$. The
+equivalent to $\BRegex'$, and $\BRegexAlt$ is equivalent to
+$\BRegexAlt'$. \bcp{We're not talking about synthesis yet, so this next
+  sentence seems off-track.  Drop it?}The
 flexibility this type equivalence rule provides is useful when a different
@@
 
-While our example is relatively simple for demonstration, writing such lenses
-for real-world formats is quite difficult, as these formats can be large and
-represent data in complex and unintuitive ways. To alleviate the difficulty of
-manually writing simple symmetric lenses, we wish to synthesize them instead.
-Given a pair of regular expressions and a set of input-output examples, we want
+Lens expressions for small examples like the one above are pretty easy to
+write.  But for larger and more complex formats, it can become quite difficult to
+write lenses by hand.  Instead, we wish to synthesize them from examples.
+
+That is, given a pair of regular expressions and a set of input-output examples, we want
 to find a simple symmetric lens that is typed by the regular expression pair and
-satisfies the input/output examples. We call such a lens a \emph{satisfying
+satisfies all the input/output examples. We call such a lens a \emph{satisfying
   lens}. In our running example, we wish to synthesize a lens between
 \lstinline{emp_salaries} and \lstinline{emp_insurance}, using as an input-output
-example the data in Figure~\ref{fig:minimized-representations}. The algorithm
+example the data in Figure~\ref{fig:minimized-representations}.
+\bcp{Suggest deleting this sentence:}The algorithm
 uses such examples to find suitable defaults for \Disconnect lenses and to
@@
 
-A challenge is that the simple symmetric lens combinators permit many
+One challenge is that the simple symmetric lens combinators permit many
 well-typed lenses
@@
 \begin{lstlisting}
-disconnect(emp_salaries,emp_insurance,rep(emp_salaries),rep(emp_insurance))
+disconnect(emp_salaries, emp_insurance, rep(emp_salaries), rep(emp_insurance))
 \end{lstlisting}
@@
 is another well-typed lens (where \lstinline{rep($\BRegex$)} denotes some string in the
-language of $\BRegex$). Furthermore, adding the input-output example in
+language of $\BRegex$).  Adding the input-output example in
 Figure~\ref{fig:minimized-representations} 
-helps, but only a little: It may, in general, require {\em many} examples to rule
+helps, but only a little: In general, {\em many} examples may be required to rule
 out all possible occurrences of nested \Disconnect lenses,
@@
 
-But what is a ``likely'' satisfying lens? We propose the following
+How can we identify such a ``likely'' lens? We propose the following
 heuristic: A satisfying
-lens is more likely if it connects more data between the two
+lens is ``more likely'' if it ``connects'' more data between the two
 formats, \EG, by using an identity rather than a disconnect lens.  The
@@
 data in one format from synchronized data in the other.  Higher
-likelihoods correspond to lower numbers of bits. In more
-detail, we say that two strings $s$ and $t$ are \emph{synchronized
+likelihoods correspond to fewer bits. Yet more formally, we say that two
+strings $s$ and $t$ are \emph{synchronized 
 according to lens $l$} if $l.\PutR\ s\ t = t$ and $l.\PutL\ t\ s = s$.\footnote{
 The notation $l.\PutR$ extracts the \PutR function from simple
-symmetric lens $l$; similarly for \PutL.}  We say that we
+symmetric lens $l$; similarly for \PutL.\bcp{Let's introduce this notation
+  earlier when talking about lenses; shoving it in a footnote here is clunky.}}  We say that we
 can \emph{recover} $s$ from $t$ using bits $b$ and lens $l$ if we can
@@
 order to recover it from $t$ because $t$ contains no useful
-information.  This calculation shows that if both \IdentityLens{}
-and \Disconnect{} are satisfying lenses, then \IdentityLens{} is more
-likely.
-
-The expected number of bits required to recover data is the well-known
-information-theoretic concept of the \emph{entropy} of the
-data~\cite{Shannon1948}. Calculating entropy requires a probability distribution
-over the space of possible values for the data. Specifically, given a set $S$
-and a probability distribution $P \OfType S \rightarrow \Reals$ over $S$, the
-entropy $\EntropyOf{S,P}$ is $-\Sigma_{s\in S}P(s)\Log_2P(s)$.
+information.  Thus, if both \IdentityLens{}
+and \Disconnect{} are satisfying lenses for a particular pair of formats,
+then \IdentityLens{} will be more likely.
+
+The expected number of bits required to recover a piece of data corresponds
+to the well-known information-theoretic concept of
+\emph{entropy}~\cite{Shannon1948}. Calculating entropy requires a
+probability distribution over the space of possible values for the
+data. Specifically, given a set $S$ and a probability distribution
+$P \OfType S \rightarrow \Reals$ over $S$, the entropy $\EntropyOf{S,P}$ is
+$-\Sigma_{s\in S}P(s)\Log_2P(s)$.  \bcp{That definition requires some more words of
+  explanation!  And probably an example or two.  It is fundamental to the
+  rest of the paper: if a reader doesn't follow it, they will be lost.  (I
+  am a little lost myself... is entropy a property of an individual piece of
+  data (as the wording above suggests), or is it a property of the set to
+  which the data belongs?)}
 

### 2019-01-29 papers/SymmetricLenses/main.tex
Checkpoint

@@
 
-This infrastructure gives us what we need to find likely satisfying lenses from
+\bcp{This paragraph has a funny flow: The first sentence implies that
+  entropy is all we need.  The second sentence then says that, actually, we
+  need something else!}Entropy is the tool we need to
+choose likely satisfying lenses from 
 a collection of satisfying lenses. All such lenses must synchronize data between
-the same two SREs $\Regex$ and $\RegexAlt$. For any such lens $\Lens$, we can
+the same two SREs $\Regex$ and $\RegexAlt$. For any lens $\Lens$, we can
 calculate the expected number of bits required to recover a string $t$ in
-$\LanguageOf{\RegexAlt}$ from a synchronized string $s$ in
+$\LanguageOf{\RegexAlt}$ from a synchronized string\bcp{Have we defined
+  what a ``synchronized string'' is?} $s$ in
 $\LanguageOf{\Regex}$.  This expectation is the \emph{conditional entropy of
   $\Regex$ given $\RegexAlt$ and $\Lens$}, formally $\sum_{s \in
-  \Regex}\ProbabilityOf{\Regex}{s}\EntropyOf{\SetOf{t \SuchThat
-    \Lens.\PutROf{s}{t} = t}}$. The likelihood we assign to $\Lens$ is the sum
+  \Regex}\ProbabilityOf{\Regex}{s}\cdot\EntropyOf{\SetOf{t \SuchThat
+    \Lens.\PutROf{s}{t} = t}}$.\bcp{It seems H is applied here to one argument.  Are
+  we supposed to know what that means.} The likelihood we assign to $\Lens$ is the sum
 of the conditional entropy of $\Regex$ given $\RegexAlt$ and $\Lens$ and the

### 2019-01-29 papers/SymmetricLenses/main.tex
Checkpoint and done for today

@@
 
-To obtain SREs from the plain regular expressions that we expect users to
-supply, we use a heuristic that attempts to assign probability annotations that
-give each string in the language equal probablity. Our algorithm could also take
-as input a stochastic regular expression rather than a plain one,
-which would allow
-the user to specify a data-specific probability distribution, either manually
-calculated or inferred from a large dataset. We found our heuristic
-works well in practice, so we have not explored these options.
+To obtain SREs from the plain regular expressions that users write, we use a
+heuristic that attempts to assign probability annotations that give each
+string in the language equal probablity. Our synthesis algorithm could
+equally well take as input a\bcp{pair of} stochastic regular expression rather than a
+plain one, permitting users to specify a data-specific probability
+distribution (either manually calculated or inferred from some dataset of
+interest). However, we have found that our heuristic works well in practice,
+so we have not explored these options.
 
 Now that we have a way to calculate likelihood, we need a way to search for
-likely lenses. Previous work on
-synthesis~\cite{augustsson-2004,osera+:pldi15,feser-pldi-2015,frankle+:popl16}
-has shown that we can use types to dramatically restrict the search space and
-improve the effectiveness of synthesis. In our setting, types are pairs of
-stochastic regular expressions where each regular expression is semantically
-equal to infinitely many others. We follow existing work on bidirectional
-program synthesis, and split our algorithm into two communicating
-synthesizers~\cite{maina+:quotient-synthesis}. The first synthesizer, \Expand,
-navigates the space of semantically equivalent regular expressions by applying
-rewrite rules that preserve both the semantics and the probability distributions
-of those stochastic regular expressions. Our algorithm ranks pairs of stochastic
-regular expressions by the number of rewrite rule applications required to
-obtain the pair from the stochastic regular expressions given as input. It
-passes the pairs off to the second synthesizer, \GreedySynth, in rank order with
-the smallest first.
-
-At a high level, \GreedySynth looks for lenses between a given pair of
-stochastic regular expressions $\Regex$ and $\RegexAlt$ that have the highest
-likelihood by performing a type-directed search. In more detail, \GreedySynth
+likely lenses. Previous work~\cite{augustsson-2004, osera+:pldi15,
+  feser-pldi-2015, frankle+:popl16} has shown that we can use types to
+dramatically restrict the search space and improve the effectiveness of
+synthesis. In our setting, types are pairs of stochastic regular
+expressions\bcp{the rest of the sentence feels out of place (and it's
+  grammatically confusing, suggesting that there might be an RE that is NOT
+  semantically equivalent to infinitely many others)}
+where each regular expression is semantically equivalent to infinitely many
+others. Following existing work on bidirectional program
+synthesis~\cite{maina+:quotient-synthesis}, we 
+split our algorithm into two communicating
+synthesizers. The first synthesizer\bcp{why
+call this algorithm a synthesizer?},
+\Expand, navigates the space of semantically equivalent regular expressions
+by applying rewrite rules that preserve both the semantics and the
+probability distributions of those stochastic regular expressions. This
+algorithm ranks pairs of stochastic regular expressions by the number of
+rewrite rule applications required to obtain each pair from the one given as
+input. It passes the pairs off to the second 
+synthesizer, \GreedySynth, in rank order with the smallest first.
+
+Abstractly, \GreedySynth looks for highest-likelihood lenses between a given pair of
+stochastic regular expressions $\Regex$ and $\RegexAlt$ by performing a
+type-directed search. In more detail, \GreedySynth 
 first normalizes the stochastic regular expressions provided by \Expand into
-\emph{stochastic DNF regular expressions}. These stochastic DNF regular
-expressions are, intuitively, stochastic regular expressions with disjunctions
-distributed over concatenations, where associativity information has been
-removed. Then, \GreedySynth uses the syntax of these n-ary DNF regular
-expressions to find highly normalized lenses, in a form we call \emph{simple
+\emph{stochastic DNF regular expressions}---a constrained representation of
+stochastic regular expressions with disjunctions  
+distributed over concatenations and with associativity information\bcp{what
+  is ``associativity information''??} 
+removed.  \GreedySynth then uses the syntax of these n-ary DNF regular
+expressions to find highly normalized\bcp{is ``highly normalized'' different
+from ``normalized''?} lenses, in a form we call \emph{simple
   symmetric n-ary DNF lenses}. These lenses include neither a composition
@@
 symmetric n-ary DNF lens is typed by a single pair of stochastic n-ary DNF
-regular expressions and \GreedySynth's search space for a given pair of regular
+regular expressions, hence \GreedySynth's search space for a given pair of regular
 expressions is finite. The variations that \GreedySynth considers include
@@
 \lstinline{name} in the management and HR data formats (we want to map first
-names to first names and last names to last names). \GreedySynth
+names to first names and last names to last names\bcp{the relationship of
+  the parenthesized comment to what comes before is not clear}). Finally, \GreedySynth
 yields a simple symmetric lens by converting the n-ary
@@
 
+%%%%%%%%%%%%%%% BCP stopped here
+
 As an example of the partnership between \Expand and \GreedySynth,
 consider the lens shown in Figure~\ref{fig:example_lens}. 
-Inferring this lens requires \Expand to use a star-semiring axiom to
+Inferring this lens requires \Expand to use a star-semiring axiom\bcp{are we
+assuming readers know what that is?} to
 rewrite the regular expression \lstinline{("\n" . emp_salary)$^*$} within

### 2019-01-30 papers/SymmetricLenses/main.tex
More polishing and comments

@@
 
-%%%%%%%%%%%%%%% BCP stopped here
-
-As an example of the partnership between \Expand and \GreedySynth,
+To illustrate the partnership between \Expand and \GreedySynth,
 consider the lens shown in Figure~\ref{fig:example_lens}. 
@@
 \lstinline{"" | (("\n" . emp_salary). ("\n" . emp_salary)$^*$)} 
-so that \GreedySynth can find the appropriate lens.
-
-Our strategy of communicating synthesizers gives us a way to
+so that \GreedySynth can find the appropriate lens.  \bcp{This example would
+be more helpful if it were spelled out in full.  What is the path from the
+start of the search to the point where we find this solution?}
+
+This architecture of communicating synthesizers gives us a way to
 enumerate pairs of stochastic regular expressions of increasing
 rank and to efficiently search through them, but it poses a
-problem: when do we stop? We may have found a promising lens
+problem: when do we stop?\bcp{This question would arise independent of the
+  architecture of the search algorithm!} We may have found a promising lens
 between a pair of stochastic regular expressions, but a different 
@@
 stochastic regular expressions, but the discovered lens is not 
-acceptable to the users.
-
-Our algorithm must resolve a three-way tension among the quality of
+acceptable to the users. \bcp{How would the algorithm know this?}
+
+The search algorithm must resolve a three-way tension among the quality of
 the inferred lens, the number of examples the user must provide to
@@
 Figure~\ref{fig:example_lens}, we don't want to spend a lot of time
-searching for an even better lens.
-
-To resolve this tension, our algorithm uses heuristics to judge whether to
+searching for an even better lens.\bcp{How does this example illustrate a
+  three-way tension?}
+
+To resolve this tension, the algorithm uses heuristics to judge whether to
 return the current best satisfying lens to the user or to pass the next set of
 unrollings to \GreedySynth. The heuristics favor stopping if the current best
-satisfying lens has relatively high likelihood, indicating the lens is very
+satisfying lens has relatively high\bcp{high compared to what??} likelihood,
+indicating the lens is very 
 promising. The heuristics also favor stopping if \Expand has delivered to
 \GreedySynth all pairs of stochastic regular expressions at a given rank and
-there is a large number of pairs at the next rank because searching through all
+there is a large number of pairs at the next rank, because searching through all
 such pairs will take a long time. With this approach, the algorithm can quickly
-respond with a satisfying lens with relatively high likelihood. If users are
+respond with a satisfying lens with relatively high likelihood. If the user is
 unhappy with the result, they can either refine the search by supplying
@@
 lens(es) and to reduce the size of the search space by cutting down on the
-number of satisfying lenses, or they can supply annotations (discussed in
-\S\ref{subsec:optimizations}) informing describing whether certain information
-should be retained through format conversion.
+number of satisfying lenses, or they can supply annotations on the source
+and target SREs indicating that certain information
+either must be retained or must be discarded by the lens (see
+\S\ref{subsec:optimizations}\bcp{this reference to the later section with
+  details is nice, but why don't we have similar references throughout this
+  overview??}).
 
@@
 \label{sec:sre}
-To characterize the likely lenses, we must compute the expected number of bits
-needed to recover a string in one data source from a synchronized string in the
-other data source. To do this, we must first develop a probabilistic language
-model for our language. We do this with \emph{stochastic regular expressions}~\cite{stoch-rnn,stoch-def},
-regular expressions annotated with probability information. With this, we can
-jointly express a language and a probability distribution over that language. In
-the following grammar, $\String$ ranges over strings, and $\Probability$ ranges
-over real numbers between 0 and 1, exclusive.
-$$\Regex{},\RegexAlt{} \GEq{}  \String  \; | \; \emptyset \; | \; \PRegexStar{\Regex}{p} \; | \; \RegexConcat{\Regex_1}{\Regex_2} \; | \; \PRegexOr{\Regex_1}{\Regex_2}{p}$$
-
-In the stochastic regular expression
+To characterize likely lenses, we must compute the expected number of bits
+needed to recover a string in one data source from a synchronized\bcp{again,
+needs a definition} string in the
+other data source. To do this, we must first develop a probabilistic 
+model for our language. We do this with \emph{stochastic regular
+  expressions} (SREs)---regular expressions annotated with probability
+information~\cite{stoch-rnn,stoch-def}---which 
+jointly express a language and a probability distribution over that language. 
+$$\Regex{},\RegexAlt{} \quad\GEq{}\quad  \String  \quad|\quad \emptyset \quad|\quad \PRegexStar{\Regex}{p} \quad|\quad \RegexConcat{\Regex_1}{\Regex_2} \quad|\quad \PRegexOr{\Regex_1}{\Regex_2}{p}$$
+(Lowercase $\String$ ranges over constant strings and $\Probability$ ranges
+over real numbers between 0 and 1, exclusive.)
+
+The semantics of an SRE $S$ is a probability distribution $P_S$ defined as
+follows.
+\begin{center}
+  \begin{tabular}{rcl}
+    $\ProbabilityOf{\String}{\String''}$
+    & =
+    & $\begin{cases*}1 & if $\String = \String''$\\ 0 & otherwise\end{cases*}$ \\
+
+    $\ProbabilityOf{\emptyset}{\String}$
+    & =
+    & $0$ \\
+
+    $\ProbabilityOf{\RegexConcat{\Regex_1}{\Regex_2}}{s}$
+
+    & =
+    & $\Sigma_{\String = \String_1\String_2}\ProbabilityOf{\Regex_1}{\String_1}\ProbabilityOf{\Regex_2}{\String_2}$ \\
+
+    $\ProbabilityOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability}}{\String}$
+    & =
+    & $\Probability\ProbabilityOf{\Regex_1}{\String} +
+      (1-\Probability)\ProbabilityOf{\Regex_2}{\String}$\\
+
+    $\ProbabilityOf{\PRegexStar{\Regex}{\Probability}}{\String}$
+    & =
+    & $\Sigma_n \Sigma_{\String = \String_1 \ldots \String_n}\Probability^n(1-\Probability)\Pi_{i=1}^n\ProbabilityOf{\Regex}{\String_i}$\\
+  \end{tabular}
+\end{center}
+In the stochastic regular expression\bcp{abbreviate to SRE here and in many
+  places elsewhere}
 $\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation $\Probability$
-represents the likelihood that the string is in $\Regex_1$. In the stochastic regular
+represents the likelihood that the string is in $\Regex_1$\bcp{That seems
+  both confusing (what is ``the string''??) and probably inaccurate (what if
+  $S_1$ and $S_2$ are the same?)}. In the stochastic regular
 expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability
@@
 much more likely having a string involving more than $n$ iterations is than
-having a string involving exactly $n$ iterations. The semantics of SREs is
-defined by the function $P$ below.
-
-\begin{center}
-  \begin{tabular}{rcl}
-    $\ProbabilityOf{\String}{\String''}$
-    & =
-    & $\begin{cases*}1 & if $\String = \String''$\\ 0 & otherwise\end{cases*}$ \\
-
-    $\ProbabilityOf{\emptyset}{\String}$
-    & =
-    & $0$ \\
-
-    $\ProbabilityOf{\RegexConcat{\Regex_1}{\Regex_2}}{s}$
-
-    & =
-    & $\Sigma_{\String = \String_1\String_2}\ProbabilityOf{\Regex_1}{\String_1}\ProbabilityOf{\Regex_2}{\String_2}$ \\
-
-    $\ProbabilityOf{\PRegexOr{\Regex_1}{\Regex_2}{\Probability}}{\String}$
-    & =
-    & $\Probability\ProbabilityOf{\Regex_1}{\String} +
-      (1-\Probability)\ProbabilityOf{\Regex_2}{\String}$\\
-
-    $\ProbabilityOf{\PRegexStar{\Regex}{\Probability}}{\String}$
-    & =
-    & $\Sigma_n \Sigma_{\String = \String_1 \ldots \String_n}\Probability^n(1-\Probability)\Pi_{i=1}^n\ProbabilityOf{\Regex}{\String_i}$\\
-  \end{tabular}
-\end{center}
+having a string involving exactly $n$ iterations. \bcp{Last sentence needs
+  an example.  Probably the first sentence also.}
 
@@
 \subsection{Stochastic Regular Expression Equivalences}
-$\Expand$ needs to be able to find equivalent stochastic regular expressions.
+The $\Expand$ algorithm needs to be able to enumerate SREs that are
+``equivalent'' to a given one.
 However, existing work does not provide a means of reasoning about stochastic
-regular expression equivalence. We extend the
-\emph{star-semiring}~\cite{Droste2009} equivalences to stochastic regular
+regular expression equivalence\bcp{Does not provide a means of {\em
+    reasoning} about it, or does not even define it?}. We extend the
+\emph{star-semiring}~\cite{Droste2009} equivalences\bcp{again, we seem to be
+assuming that all readers will already know what they are...} to stochastic regular
 expressions (Figure~\ref{fig:ssr-sre}).

### 2019-01-31 papers/SymmetricLenses/main.tex
More polishing -- done for now

@@
 \newif\ifanon\anonfalse    % set true to suppress names, etc.
-\newif\ifappendices\appendicesfalse
+\newif\ifappendices\appendicestrue
 
@@
 regular expression equivalence\bcp{Does not provide a means of {\em
-    reasoning} about it, or does not even define it?}. We extend the
+    reasoning} about it, or does not even define
+  it?}. Figure~\ref{fig:ssr-sre} shows how we to extend the
 \emph{star-semiring}~\cite{Droste2009} equivalences\bcp{again, we seem to be
 assuming that all readers will already know what they are...} to stochastic regular
-expressions (Figure~\ref{fig:ssr-sre}).
+expressions.\bcp{Notation: I don't see any reason for the $s$ superscript in
+$\SSREquiv$}
 
@@
 \begin{proof}
-  The proofs of this theorem, and subsequent theorems in this paper, appears in
-  the appendix presented in the full version of the paper accompanying this
-  submission.
+  See the appendix%
+\ifappendices\else{} included with the supplemental
+material\bcp{Right?}\fi.  \bcp{It took me quite a while to find it.  We
+  should give an actual pointer to a specific theorem in the appendix.}
 \end{proof}
 
-We prove these star-semiring equivalences sound, as those are the equivalences
-we either traverse (with \Expand), or normalize across (as part of
-\GreedySynth).
+This theorem will come in handy as we traverse (with \Expand) or normalize
+across (as part of \GreedySynth) star-semiring equivalences.
 
@@
 The entropy of a data source $S$ is the expected number of bits required to
-describe a random element in $S$ (formally $-\Sigma_{s\in S}P(s)\Log_2P(s)$). The
+describe a random element in $S$ (formally $-\Sigma_{s\in S}\,P(s)\cdot \Log_2P(s)$). The
 entropy of a stochastic regular expression can be computed directly from its
-syntax, when the stochastic regular expression is unambiguous and contains no
+syntax, when it is unambiguous\bcp{undefined term} and contains no
 empty subcomponents.
@@
 \end{center}
-We have proven that this syntactic computation corresponds to the entropy of the
+\bcp{This sentence says exactly what the previous one does.  Drop.}We have proven that this syntactic computation corresponds to the entropy of the
 stochastic regular expression when the stochastic regular expression is
 unambiguous (parses strings uniquely).
+
+\bcp{Again, this is a critical definition for understanding the paper, and
+  it is pretty gnarly.  Examples, please!}
 
@@
   If $\Regex$ is unambiguous and does not contain $\emptyset$ as a subterm,
-  $\EntropyOf{\Regex}$ is the entropy of $\Regex$.
+  $\EntropyOf{\Regex}$ is the entropy of $\Regex$.\bcp{Should it not be $-\EntropyOf{\Regex}$??}
 \end{theorem}
 
-Without unambiguity constraints it is not easy to calculate entropy (and the
-method we have given is certainly incorrect).  For example, consider the SRE
-$\PRegexOr{\lstinline{"a"}}{\lstinline{"a"}}{.5}$. Using the above computation,
-this SRE has an entropy of $1$, but the entropy should be $0$. These
-difficulties become particularlly complex when the languages are ambiguous in
-subtle ways and the algorithm must be correct when only some of the two languages
-overlap. Similar issues occur when trying to find the entropy when $\emptyset$
-is a subterm (what should the entropy of $\PRegexStar{\emptyset}{.5}$ be?),
-so we do not define entropy on $\emptyset$.
+Without unambiguity constraints, the method above for calculating entropy does
+not work, and it is not easy to see how to generalize it.  For example,
+consider the SRE $\PRegexOr{\lstinline{"a"}}{\lstinline{"a"}}{.5}$. Using
+the above computation, this SRE has an entropy of $1$, but the entropy
+should be $0$. These difficulties become particularly complex when the
+languages are ambiguous in subtle ways and the algorithm must be correct
+when only some of the two languages overlap. Similar issues occur when
+trying to find the entropy when $\emptyset$ is a subterm (what should the
+entropy of $\PRegexStar{\emptyset}{.5}$ be?), so we do not define entropy
+on~$\emptyset$.
 
@@
 our synthesis procedure for other reasons, and we can easily preprocess
-emptysets out of the stochastic regular expressions that are themselves
-nonempty by traversing the appropriate star-semiring equivalences (\EG,
-\PRegexOr{\emptyset}{\lstinline{"s"}}{.5} \SSREquiv \lstinline{"s"}).
+empty subexpressions out of SREs that are themselves nonempty using the
+star-semiring equivalences (\EG, $\PRegexOr{\emptyset}{\lstinline{"s"}}{.5}
+\SSREquiv \lstinline{"s"}$).
 
@@
 % \label{fig:regex-equivalence-rules}
+
+%%%%%%%%%%%%%%  BCP: stopped reading carefully here...
 
@@
 recover a regular expression from a stochastic regular expression by deleting
-probability annotations. In a slight abuse of notation, if $\Regex$ is a
+probability annotations. In a slight abuse of notation\bcp{why is this an
+  abuse of notation??}, if $\Regex$ is a
 stochastic regular expression, $\BRegex$ is the regular expression obtained by
@@
 \Leftrightarrow \BRegexAlt$ means that $\Lens$ is a well typed lens between
-$\LanguageOf{\BRegex}$ and $\LanguageOf{\BRegexAlt}$.
+$\LanguageOf{\BRegex}$ and $\LanguageOf{\BRegexAlt}$.  \bcp{Ugly typesetting:}
 
@@
 
-We have already described a number of these lenses in \S\ref{sec:overview}.
-The compose lens $\Lens_1 ; \Lens_2$ composes two lenses sequentially, it
+We have already described a number of these combinators in \S\ref{sec:overview}.
+Briefly (full definitions are below), the compose lens $\Lens_1 ; \Lens_2$
+composes two lenses sequentially, it 
 transforms from one data format to the other by first transforming to an
@@
 
-We now provide the typing rules and semantics for these combinators. We do
-not include the semantics for combinators equivalent to those presented in
-prior work~\cite{symmetric-lenses,Boomerang}. Furthermore, we only present
-\CreateR and \PutR when \CreateL and \PutL are symmetric; full details can
-be found in an appendix\ifappendices\else{} (in the supplemental
-materials)\fi. For simplicity of presentation, we use unambiguous regular
-expressions everywhere, even though some lenses can actually be defined over
-ambiguous regular expressions (for example, the identity lens doesn't need
-any ambiguity constraints\bcp{Is this the only example?}); we believe
-relaxing this restriction is unproblematic.
+We now give typing rules and semantics for these combinators. We do not
+include the semantics for combinators equivalent to those presented in prior
+work~\cite{symmetric-lenses,Boomerang}\bcp{Huh?}. Furthermore, we present
+just \CreateR and \PutR in cases where \CreateL and \PutL are symmetric;
+full details can be found in an appendix\bcp{where,
+  exactly??}\ifappendices\else{} (in the supplemental materials)\fi. For
+simplicity of presentation, we use unambiguous regular expressions
+everywhere, even though some lenses can actually be defined over ambiguous
+regular expressions (for example, the identity lens doesn't need any
+ambiguity constraints\bcp{Is this the only example?}); we believe relaxing
+this restriction is unproblematic.
 
@@
 data gets changed from one format to the other, information in the old format is
-lost.
+lost.  \bcp{We should highlight here the fact that this is a significant
+  loss of power for some examples, compared to full symmetric lenses.}
 %\begin{center}
@@
 expected amount of information required to recover a string in $\RegexAlt$ from
-a string in $\Regex$, synchronized by $\Lens$. We use the function
+a string in $\Regex$, synchronized by $\Lens$. Similarly, we use the function
 $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate bounds of the
 expected amount of information required to recover a string in $\Regex$ from a
-string in $\RegexAlt$, synchronized by $\Lens$. We use the notation $a(b,c)$ to
-mean $(a*b,a*c)$, and the notation $(a,b)+(c,d)$ to mean $(a+c,b+d)$.
+string in $\RegexAlt$, synchronized by $\Lens$. We write $a(b,c)$ to
+mean $(a*b,a*c)$, and $(a,b)+(c,d)$ to mean $(a+c,b+d)$.
 \begin{center}
@@
 synchronized string in the other format. Note that we would be able to exactly
-calculate the conditional entropy, were it not for \MergeL. If
+calculate the conditional entropy, were it not for \MergeL\bcp{and \MergeR??}. If
 $\MergeLOf{\Lens_1}{\Lens_2} \OfType \Regex \Leftrightarrow
@@
 composition $\Lens_1 ; \Lens_2$ and our theorem excludes lenses that contain
-such compositions. Unfortunately, defining the entropy of lenses involving
+such compositions.  Defining the entropy of lenses involving
 composition is challenging because $\Lens_1$ might, for instance, add some
@@
 this property is difficult to determine merely from syntax. Fortunately, we are able to
-avoid such complication altogether by synthesizing \emph{DNF lenses}---simple symmetric
+sidestep such considerations by synthesizing \emph{DNF lenses}---simple symmetric
 lenses that inhabit a disjunctive normal form. DNF lenses have a number of
 properties that facilitate synthesis: they cut down the program search space
-dramatically and they do not contain instances of composition.
+dramatically and they do not contain instances of composition.\bcp{The last
+  sentence basically repeats what the rest of the paragraph already says.}
 
@@
 The details of these definitions are syntactically tedious, but not
-intellectually difficult. We elide them here but include them in the appendix.
+intellectually difficult. We elide them here but include them in the appendix.\bcp{Where?}
 
@@
 
-\paragraph*{Relevance Annotations}
+\paragraph*{Relevance Annotations}\bcp{Wonder whether this should be its own
+top-level section...?}
 In this work, there are three competing constraints. The tool should (1) return
@@
 
-Conversely, the termination heuristic is satisfied algorithm cuts short: if it
+Conversely, the termination heuristic is satisfied algorithm cuts
+short\bcp{huh??}: if it 
 had continued searching for lenses it would have found the desired one, but the
@@
 
-If synthesizing a lens seems to be taking a long time, it is spending time
+If synthesizing a lens seems to be taking a long time, it is\bcp{always?} spending time
 checking it cannot find a lens that projects less information. By ``skipping''
@@
 
-If the algorithm returns a lens with too disconnects and merges, one could
+If the algorithm returns a lens with too disconnects\bcp{??} and merges, one could
 provide additional examples to constrain the lens more. However, for large
@@
 cost of any lens that loses ``required'' information to infinity.
+
+\bcp{This discussion seems both too brief---maybe some concrete examples
+  would help?---and too informal: E..g., What do skip and require have to do
+  with probability distributions and entropy?}
+
 % end synthesis
@@
 Boomerang. Specifically, we added symmetric lens combinators to
-Boomerang, and implemented all of Boomerang's asymmetric lens
+Boomerang\bcp{that's exactly what the previous sentence said!}, and
+implemented all of Boomerang's asymmetric lens 
 combinators  using a combination of the symmetric lens combinators presented in this
@@
 
-\paragraph*{Relation with Simple Symmetric Lenses} To compare classical and simple
+\paragraph*{Symmetric Lenses vs.{} Simple Symmetric Lenses} To compare classical and simple
 symmetric lenses, we define an $apply$ function on simple symmetric lenses as well. If
@@
 
-\paragraph*{Other Lens Formulations}
+\paragraph*{Other Lens Formulations}\bcp{This paragraph doesn't add much}
 Bidirectional programming has an extensive literature, with many extensions onto
@@
 Since $S$ is unambigous, then $S_1$ and $S_2$ are also unambiguous, and for each $s \in \mathcal{L}(S_1 \cdot S_2)$ there exists a unique $s_1 \in \mathcal{L}(S_1)$ and a unique $s_2 \in \mathcal{L}(S_2)$ such that $s = s_1 \cdot s_2$.
-By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$, hence
+By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$, hence\bcp{Don't understand what's going on with the $-$
+  signs in the last few lines here.}
 \begin{align*}
@@
 
-Since $S$ is unambiguous, then $S_1$ and $S_2$ are also unambiguous, and every $s \in \mathcal{L}(S_1 \; |_p \; S_2)$ is either in $\mathcal{L}(S_1)$ or in $\mathcal{L}(S_2)$.
-By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$.
-Then
+\bcp{``Since...then...'' is not english.}
+Since $S$ is unambiguous, then $S_1$ and $S_2$ are also unambiguous, and
+every $s \in \mathcal{L}(S_1 \; |_p \; S_2)$ is either in $\mathcal{L}(S_1)$
+or in $\mathcal{L}(S_2)$. By the induction hypothesis, $H(\mathcal{L}(S_1))
+= \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) =
+\mathbb{H}(S_2)$. Then\bcp{Again, don't understand what's going on with the $-$
+  signs.}
 \begin{align*}

### 2019-02-05 papers/SymmetricLenses/main.tex
update to feedback

@@
 \newif\ifanon\anonfalse    % set true to suppress names, etc.
-\newif\ifappendices\appendicestrue
+\newif\ifappendices\appendicesfalse
 
@@
   bijective lenses} over string data---given a pair of types (regular
-expressions) and a small number of examples.  \bcp{Why is the next line not indented?}
+expressions) and a small number of examples.  \bcp{Why is the next line not
+  indented?} \afm{I have truly no idea, deferring until I get a stackoverflow answer.}
 
@@
   criticized on the same grounds.}
-%
-\bcp{Also, it would be nice to be able to give a quick example of a truly
-  symmetric case---do any of these examples work for that?}\afm{h-event and
-  vCalendar are truly symmetric. Even the fields related to location alone
-  are truly symmetric (though it is subtle to see this).}
+\afm{I still think that there are situations where the embroidery file makes
+  sense (I'm sharing my design with a friend, who then tweaks it and sends it
+  back. There may also need to file-specific information, where one contains
+  parameters like how fast to embroider and some doesn't, to be honest I don't
+  know much about embroidery though. Thinking about other examples, however.)}
 
@@
 Domain-specific languages for writing lenses can facilitate building
-principled data transformations.
-
-\textit{Synthesizing} lenses 
-can make the task even easier.  Indeed, lens languages are excellent
-targets for search-based synthesis, since their specialized, sub-Turing-complete syntax
-and expressive type systems drastically
+principled data transformations.  For example, Boomerang~cite{boomerang} is a
+language for expressing bidirectional string transformations.
+
+\textit{Synthesizing} lenses can make the task even easier. Indeed, lens
+languages, like Boomerang, are excellent targets for search-based synthesis, since their
+specialized, sub-Turing-complete syntax and expressive type systems drastically
 limit the space of possible programs.
 %
-In particular, the Optician~\cite{optician} algorithm\bcp{there's a
-  potential for confusion between ``Optician'' and ``Boomerang''---are they
-  distinct systems?  Is one part of the other?  Is one a system and one just
-  an algorithm?  It's not very clear.} has demonstrated that
-it is feasible to synthesize quite complex examples, albeit only for a
-restricted class of lenses---so-called bijective lenses. Specifically, given 
-source and target formats plus a small number of
-examples of corresponding data, Optician's synthesis algorithm is guaranteed
-to find a bijective lens matching the examples, if one exists. Indeed, 
-it can often find an appropriate lens in less than a second with
-no examples at all.
+In particular, the recent Optician~\cite{optician} enhancement to Boomerang has
+demonstrated that it is feasible to synthesize quite complex examples, albeit
+only for a restricted class of lenses---so-called bijective lenses.
+Specifically, given source and target formats plus a small number of examples of
+corresponding data, Optician's synthesis algorithm is guaranteed to find a
+bijective lens matching the examples, if one exists. Indeed, it can often find
+an appropriate lens in less than a second with no examples at all.
 %, with just types describing the source and target formats.
@@
 bijective lenses, ``asymmetric'' lenses (where the transformation from a
-source structure $S$ to a target $T$ can throw away information that must
-then be restored when transferring from $T$ to $S$), and even more flexibile
+source structure to a target can throw away information that must
+then be restored when transferring from target to source), and even more flexibile
 transformations that allow each side to retain information not present in
 the other?
-
-\bcp{We use $S$ and $T$ here but $X$ and $Y$ below.}
 
@@
 by ~\citet{symmetric-lenses} operate
-over three structures: the source $S$, the target $T$ and a complement
+over three structures: the ``left'' structure $X$, the ``right'' $Y$ and a complement
 $C$ that contains the information not
-present in either $S$ or $T$. Logistically, these complements must be stored
+present in either $X$ or $Y$. Logistically, these complements must be stored
 and managed somehow.  More fundamentially, complements complicate
@@
 lenses that do not require external ``memory'' to recover data from past
-instances of $S$ or $T$ when making a round trip. They only need the most
+instances of $X$ or $Y$ when making a round trip. They only need the most
 recent instance.
@@
 enough in practice, we add simple symmetric lenses to the Boomerang
-language~\cite{Boomerang}\bcp{this seems to assume that readers already know what
-  Boomerang is} and apply them to a range of real-world
+language~\cite{Boomerang} and apply them to a range of real-world
 applications, showing in the process that simple symmetric lenses integrate
@@
 
-We consider a ``likely'' lens to be one that propagates ``a lot'' of
-information from source to 
-target and vice versa. Conversely, an unlikely lens is one that requires a
-large amount of additional information to recover the target given a
-source.\bcp{I'm nervous about the way we're talking about source and target
-  here.  For one thing, even the words ``source'' and ``target'' connote a
-  directionality that is misleading here.  And even ignoring that issue, a
-  lot of the discussion (like here) sounds as though we are talking about
-  unidirectional transformations.}  
-This categorization follows from the intuition
-that users will typically want to synchronise formats with lots
-of shared information\bcp{really?  It's hard to be certain without lots of
-  practical experience, but a priori it seems equally reasonable for users
-  to want to pull little bits of shared information out of huge structures
-  and keep them in sync.  But maybe we can weaken the claim and just say
-  that we have an intuition that this will at least be {\em one} large class
-  of application scenarios, and this is the one we focus on}, and so lenses
-that propagate more information are preferable to ones that propagate less.
+We consider a ``likely'' lens to be one that propagates ``a lot'' of information
+from the left data format to the right and vice versa. Conversely, an unlikely
+lens is one that requires a large amount of additional information to recover
+either of the formats given the other. For many application scenarios, users
+will want to synchronise formats with lots of shared information (and lenses
+with little shared information are typically easier to write). Our tool is built
+primarily for these complex scnearios, and so lenses that propagate more
+information are preferable in our context to ones that propagate less.
 %
@@
 
-With simple symmetric lenses and our SRE-based likelihood measure in
-hand, we develop a new algorithm for synthesizing likely lenses.
-At its core, the algorithm performs a type-directed search between the 
-descriptions of the source and target formats, measuring success using
-the likelihood measure.
-The interesting complications arise from the need to deal with
-regular-expression equivalences.  There are infinitely many regular
-expressions equivalent to a given one, and the lens returned by a
-type-directed search will in general depend on which of the possible
-representations are chosen for its source and target formats.  Moreover,
-there may be {\em no} lens connecting the 
-source and target formats selected\bcp{until we get past the colon, it is
-  not clear that we are talking here about formats {\em not} modulo
-  equivalence.} by the user: in general, the synthesis 
-algorithm has to search to find equivalent ones for which a suitable lens
-exists.  
-To tame this complexity, we divide the synthesis algorithm into two
-communicating search procedures (Figure~\ref{fig:high-level-algorithm}),
-following~\citet{optician}. 
-The first, \Expand, uses rewriting rules to iteratively
-propose new pairs of stochastic regular expressions equivalent to the
-original pair.
-The second, \GreedySynth, takes a pair of SREs as input
-and uses a greedy, type-directed algorithm to find a simple symmetric
-lens between them, returning the lens and its likelihood score
-to \Expand.  The whole synthesis algorithm terminates when a comparison of the
-likilihood score of the best lens found so far with 
-the complexity\bcp{what does it mean to compare likelihood with complexity?}
-of newly generated SREs  
-suggests that further searching is unlikely to be fruitful.
+With simple symmetric lenses and our SRE-based likelihood measure in hand, we
+develop a new algorithm for synthesizing likely lenses. At its core, the
+algorithm performs a type-directed search between the descriptions of the source
+and target formats, measuring success using the likelihood measure. The
+interesting complications arise from the need to deal with regular-expression
+equivalences. There are infinitely many regular expressions equivalent to a
+given one, and the lens returned by a type-directed search will in general
+depend on which of the possible representations are chosen for its source and
+target formats. Moreover, if Boomerang's type-equivalence rule is disallowed,
+there may be {\em no} well-typed lens connecting the source and target formats
+selected by the user: in general, the synthesis algorithm has to search through
+equivalent regular expression types for which a suitable lens exists. To tame
+this complexity, we divide the synthesis algorithm into two communicating search
+procedures (Figure~\ref{fig:high-level-algorithm}), following~\citet{optician}.
+The first, \Expand, uses rewriting rules to iteratively propose new pairs of
+stochastic regular expressions equivalent to the original pair. The second,
+\GreedySynth, takes a pair of SREs as input and uses a greedy, type-directed
+algorithm to find a simple symmetric lens between them, returning the lens and
+its likelihood score to \Expand. The whole synthesis algorithm heuristically
+terminates when the most likely lens is no longer unlikely enough to justify
+the time cost of searching through newly generated SREs.
 
@@
 variety of other sources (mostly from the bidirectional programming literature).
-The ``programmer''\bcp{user?} can control the search process via examples and annotations
+The user can control the search process via examples and annotations
 that informs the synthesis algorithm as to whether certain bits of information
@@
 %
-We
-present a collection of combinators for building simple symmetric lenses and
+We present a collection of combinators for building simple symmetric lenses and
 describe how to calculate their likelihoods (\S\ref{sec:ssl}).
-\item We give
-an algorithm for synthesizing lenses built from these combinators, using a 
-novel application of stochastic regular expressions to guide the search
-process (\S\ref{sec:synthesis})\bcp{``and prove it correct''!  (right?)}.  
+\item We give an algorithm for synthesizing lenses built from these combinators,
+  using a novel application of stochastic regular expressions to guide the
+  search process (\S\ref{sec:synthesis}) and prove that our metrics have deep
+  semantic meaning.
 %
@@
   hideous and hard to navigate!!}
-\ifappendices\else included with the supplemental material\bcp{Right?}\fi.
+\ifappendices\else included with the supplemental material\fi.
 
@@
 \begin{gather}
- \tag{1}
- \mbox{$\CreateR \OfType X \rightarrow Y$}\\
- \tag{2}
- \mbox{$\CreateL \OfType Y \rightarrow X$}\\
- \tag{3}
- \mbox{$\PutR \OfType X \rightarrow Y \rightarrow Y$}\\
- \tag{4}
- \mbox{$\PutL \OfType Y \rightarrow X \rightarrow X$}
+ \mbox{$\CreateR \OfType X \rightarrow Y$} \nonumber \\
+ \mbox{$\CreateL \OfType Y \rightarrow X$} \nonumber \\
+ \mbox{$\PutR \OfType X \rightarrow Y \rightarrow Y$} \nonumber \\
+ \mbox{$\PutL \OfType Y \rightarrow X \rightarrow X$} \nonumber
 \end{gather} 
@@
 \vskip 1ex
-\bcp{Are we going to use the numbers next to the laws??  Seems like not, so
-  why have them at all?}
 \noindent
-The two \KW{create} functions are used to propagate shared information and
-fill in default values when introducing new data (\EG, the ``unk'' salary
-entry is added\bcp{is added by what?  A create function, I guess, but the
-  connection is not clear} alongside the name to the management file when HR inserts a
-new employee). The two \KW{put} functions propagate edits from one format to
-the other by combining a new value from one with an old value from the
-other.  The round-tripping laws guarantee that pushing an ``un-edited''
-value from one side (the result of a put or create) back through a lens in
-the other direction will get back to exactly where we began.  \bcp{That
-  paragraph seems likely to be quite opaque to any reader that is not
-  already familiar with (symmetric) lenses.  How about a concrete example?}
-
-This definition differs from classical symmetric lenses in that it does not
-involve a \emph{complement}.  We give a detailed comparison in
+The two \KW{create} functions are used to propagate shared information and fill
+in default values when introducing new data (\EG, on \KW{create} the ``unk''
+salary entry is added alongside the name to the management file when HR inserts
+a new employee). The two \KW{put} functions propagate edits from one format to
+the other by combining a new value from one with an old value from the other.
+The notation $\Lens.\PutR$ extracts the \PutR function from simple symmetric
+lens $\Lens$; similarly for \PutL. These four functions can be used to keep the
+common information between two file types in synch. For example, if a new file
+of the left-hand format is created, the \CreateR function will a new file in the
+right-hand format. If the right-hand format is edited, the \PutL function can
+update the left-hand format with the changed information.
+
+The round-tripping laws guarantee that pushing an ``un-edited'' value from one
+side (the result of a put or create) back through a lens in the other direction
+will get back to exactly where we began. For example, consider,
+\lstinline{emp_lens}, a lens between
+\lstinline{emp_salaries} and
+\lstinline{emp_insurance}. Applying
+\lstinline{emp_lens}.\CreateR to a salary file creates a insurance file with the
+same employee names in the file, with \lstinline{"UNK"} for each insurance
+company. The round-tripping laws guarantee that using \lstinline{emp_lens}.\PutL
+to put the generated insurance file into the original salary file will return
+the original salary file, unmodified. If \KW{put}ting a string $s$ into a string
+$t$ results in the original string $s$, then $s$ and $t$ are considered
+\emph{synchronized}.
+
+This simple symmetric lens definition differs from classical symmetric lenses in
+that it does not involve a \emph{complement}. We give a detailed comparison in
 \S\ref{sec:relationship}.
@@
 
-In principle, programmers can define simple symmetric lenses just by
-writing four arbitrary functions by hand and showing they satisfy the round-tripping
+In principle, programmers can define simple symmetric lenses just by writing
+four arbitrary functions by hand and showing they satisfy the round-tripping
 laws. But a better idea is to streamline the process by providing a set of
-combinators---a domain-specific 
-language---for building law-abiding complex lenses from simpler ones.  Such
-bidirectional languages
-have been given for building lenses over many different kinds of data,
-including database relations~\cite{BohannonPierceVaughan} and algebraic data
-types~\cite{symmetric-lenses}\bcp{and what else? graphs, etc.?}.  We focus
-here on combinators for defining lenses between string data formats
+combinators---a domain-specific language---for building law-abiding complex
+lenses from simpler ones. Such bidirectional languages have been given for
+building lenses over many different kinds of data, including database
+relations~\cite{BohannonPierceVaughan} algebraic data
+types~\cite{symmetric-lenses}, graphs~\cite{tgg}, and more.
+
+We focus here on combinators for defining lenses between string data formats
 described by regular expressions.
@@
 HR file.
-The defaults are strings that indicate the information is missing.
+For example, with the above lens, $\PutL \App \lstinline{""} \App 60000$ will return 60000,
+and \PutR will always return \lstinline{""}.
 
@@
 the left-to-right direction.
-\bcp{We don't actually give a nontrivial example of how disconnect works.}
 
@@
 %let emp_lens = or(id(""),employees_lens) in
-  \caption{A lens that synchronizes management and HR employee files\bcp{I'm
-    pretty sure there was an error in name\_lens: the thing between the two
-    names should be id(" ") (it was ins('','')).  Crossing my fingers that I'm
-    making a correct fix...}}
+  \caption{A lens that synchronizes management and HR employee files}
   \label{fig:example_lens}
@@
 equivalences on them, and sometimes we must convert one regular expression type
-to another, equivalent, regular expression type. \bcp{The next sentence is
-  dense.  Just show the rule?}Our typing rules include type
-equivalence, so if a lens has type $\BRegex \Leftrightarrow \BRegexAlt$, then it
-is also has type $\BRegex' \Leftrightarrow \BRegexAlt'$, where $\BRegex$ is
-equivalent to $\BRegex'$, and $\BRegexAlt$ is equivalent to
-$\BRegexAlt'$. \bcp{We're not talking about synthesis yet, so this next
-  sentence seems off-track.  Drop it?}The
-flexibility this type equivalence rule provides is useful when a different
-representation of a regular expression would facilitate synthesis.
+to another, equivalent, regular expression type.
+%
+\[
+  \centering
+  \inferrule*
+  {
+    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt\\
+    \BRegex \SSREquiv \BRegex'\\
+    \BRegexAlt \SSREquiv \BRegexAlt'
+  }
+  {
+    \Lens \OfType \BRegex' \Leftrightarrow \BRegexAlt'
+  }
+\]
+%
+This rule is needed for lenses like the final one in
+Figure~\ref{fig:example_lens} to be well-typed.
 
@@
 example the data in Figure~\ref{fig:minimized-representations}.
-\bcp{Suggest deleting this sentence:}The algorithm
-uses such examples to find suitable defaults for \Disconnect lenses and to
-figure out how to match data that could be synchronised in multiple ways.
 
@@
 
-How can we identify such a ``likely'' lens? We propose the following
-heuristic: A satisfying
-lens is ``more likely'' if it ``connects'' more data between the two
-formats, \EG, by using an identity rather than a disconnect lens.  The
-likeliest satisfying lenses would connect as much data as
-possible. More formally, we define the likelihood of a
-satisfying lens as the expected number of bits required to recover
-data in one format from synchronized data in the other.  Higher
-likelihoods correspond to fewer bits. Yet more formally, we say that two
-strings $s$ and $t$ are \emph{synchronized 
-according to lens $l$} if $l.\PutR\ s\ t = t$ and $l.\PutL\ t\ s = s$.\footnote{
-The notation $l.\PutR$ extracts the \PutR function from simple
-symmetric lens $l$; similarly for \PutL.\bcp{Let's introduce this notation
-  earlier when talking about lenses; shoving it in a footnote here is clunky.}}  We say that we
-can \emph{recover} $s$ from $t$ using bits $b$ and lens $l$ if we can
-reconstruct $s$ from $t$, $b$, and $l$.  For example, given
-the \IdentityLens{} lens, we can recover $s$ from $t$ using no bits
-because, in this case, $s$ is just $t$.  In contrast, given
-the \Disconnect{} lens, we need enough bits to fully encode $s$ in
-order to recover it from $t$ because $t$ contains no useful
-information.  Thus, if both \IdentityLens{}
-and \Disconnect{} are satisfying lenses for a particular pair of formats,
-then \IdentityLens{} will be more likely.
-
-The expected number of bits required to recover a piece of data corresponds
-to the well-known information-theoretic concept of
-\emph{entropy}~\cite{Shannon1948}. Calculating entropy requires a
-probability distribution over the space of possible values for the
-data. Specifically, given a set $S$ and a probability distribution
-$P \OfType S \rightarrow \Reals$ over $S$, the entropy $\EntropyOf{S,P}$ is
-$-\Sigma_{s\in S}P(s)\Log_2P(s)$.  \bcp{That definition requires some more words of
-  explanation!  And probably an example or two.  It is fundamental to the
-  rest of the paper: if a reader doesn't follow it, they will be lost.  (I
-  am a little lost myself... is entropy a property of an individual piece of
-  data (as the wording above suggests), or is it a property of the set to
-  which the data belongs?)}
+How can we identify such a ``likely'' lens? We propose the following heuristic:
+A satisfying lens is ``more likely'' if it ``connects'' more data between the
+two formats, \EG, by using an identity rather than a disconnect lens. The
+likeliest satisfying lenses would connect as much data as possible. More
+formally, we define the likelihood of a satisfying lens as the expected number
+of bits required to recover data in one format from synchronized data in the
+other. Higher likelihoods correspond to fewer bits. Yet more formally, we say
+that two strings $s$ and $t$ are \emph{synchronized according to lens $l$} if
+$l.\PutR\ s\ t = t$ and $l.\PutL\ t\ s = s$. We say that we can \emph{recover}
+$s$ from $t$ using bits $b$ and lens $l$ if we can reconstruct $s$ from $t$,
+$b$, and $l$. For example, given the \IdentityLens{} lens, we can recover $s$
+from $t$ using no bits because, in this case, $s$ is just $t$. In contrast,
+given the \Disconnect{} lens, we need enough bits to fully encode $s$ in order
+to recover it from $t$ because $t$ contains no useful information. Thus, if both
+\IdentityLens{} and \Disconnect{} are satisfying lenses for a particular pair of
+formats, then \IdentityLens{} will be more likely.
+
+The expected number of bits required to recover a piece of data corresponds to
+the well-known information-theoretic concept of
+\emph{entropy}~\cite{Shannon1948}. Calculating entropy requires a probability
+distribution over the space of possible values for the data. Specifically, given
+a set $S$ and a probability distribution $P \OfType S \rightarrow \Reals$ over
+$S$, the entropy $\EntropyOf{S,P}$ is $-\Sigma_{s\in S}P(s)\Log_2P(s)$. In
+information theory, each element of $S$ has \emph{information content}, the
+number of bits required to specify that element. The information content of an
+element $s$ is $-\Log_2P(s)$. For example, let $S = \SetOf{a,b_1,b_2}$, where
+$P(a)=.5$ and $P(b_1) = .25 = P(b_2)$. The information content of $a$ is $1$,
+and the information content of each $b_i$ is 2. The entropy of $S$ calculates
+the expected number of bits required to specify a random element pulled from
+$S$, in other words, the probability of each element times its information
+content. In our example, the entropy of $S$ is $1.5$. When $P$ is clear, we will
+often use the shorthand of $\EntropyOf{S}$ for $\EntropyOf{S,P}$.
 
@@
 which are regular expressions in which each operator is annotated with a
-probability.  A stochastic regular expression
+probability (see \S\ref{sec:sre}).  A stochastic regular expression
 specifies both a set of strings and a probability
@@
 
-\bcp{This paragraph has a funny flow: The first sentence implies that
-  entropy is all we need.  The second sentence then says that, actually, we
-  need something else!}Entropy is the tool we need to
-choose likely satisfying lenses from 
-a collection of satisfying lenses. All such lenses must synchronize data between
-the same two SREs $\Regex$ and $\RegexAlt$. For any lens $\Lens$, we can
-calculate the expected number of bits required to recover a string $t$ in
-$\LanguageOf{\RegexAlt}$ from a synchronized string\bcp{Have we defined
-  what a ``synchronized string'' is?} $s$ in
-$\LanguageOf{\Regex}$.  This expectation is the \emph{conditional entropy of
-  $\Regex$ given $\RegexAlt$ and $\Lens$}, formally $\sum_{s \in
-  \Regex}\ProbabilityOf{\Regex}{s}\cdot\EntropyOf{\SetOf{t \SuchThat
-    \Lens.\PutROf{s}{t} = t}}$.\bcp{It seems H is applied here to one argument.  Are
-  we supposed to know what that means.} The likelihood we assign to $\Lens$ is the sum
+In the synthesis setting, there are typically an infinite number of possible
+lenses that satisfy the specification. There are many choices on which
+components of the data should be synchronized vs which should be left
+unsynchronized. We use entropy to guage the relative likelihood of such lenses
+in our synthesis algorithm (see \S\ref{subsec:lenscosts}). For any lens $\Lens$, we can calculate the expected
+number of bits required to recover a string $t$ in $\LanguageOf{\RegexAlt}$ from
+a synchronized string $s$ in $\LanguageOf{\Regex}$. This expectation is the
+\emph{conditional entropy of $\Regex$ given $\RegexAlt$ and $\Lens$}, formally
+$\sum_{s \in \Regex}\ProbabilityOf{\Regex}{s}\cdot\EntropyOf{\SetOf{t \SuchThat
+    \Lens.\PutROf{s}{t} = t}}$. The likelihood we assign to $\Lens$ is the sum
 of the conditional entropy of $\Regex$ given $\RegexAlt$ and $\Lens$ and the
 conditional entropy of $\RegexAlt$ given $\Regex$ and $\Lens$. This metric
-assigns higher entropy (or cost) to lenses where knowing the string
-on one side 
-provides little information about the string on the other side. 
-It assigns zero entropy to bijections because given a string $s \in \Regex$,
-the bijection exactly determines
-the corresponding string in 
-$\RegexAlt$.
+assigns higher entropy (or cost) to lenses where knowing the string on one side
+provides little information about the string on the other side. It assigns zero
+entropy to bijections because given a string $s \in \Regex$, the bijection
+exactly determines the corresponding string in $\RegexAlt$.
 
@@
 string in the language equal probablity. Our synthesis algorithm could
-equally well take as input a\bcp{pair of} stochastic regular expression rather than a
+equally well take as input a pair of stochastic regular expression rather than a
 plain one, permitting users to specify a data-specific probability
@@
 synthesis. In our setting, types are pairs of stochastic regular
-expressions\bcp{the rest of the sentence feels out of place (and it's
-  grammatically confusing, suggesting that there might be an RE that is NOT
-  semantically equivalent to infinitely many others)}
-where each regular expression is semantically equivalent to infinitely many
-others. Following existing work on bidirectional program
+expressions. The fact that each regular expression is semantically equivalent to
+infinitely many others complicates the problem.
+Following existing work on bidirectional program
 synthesis~\cite{maina+:quotient-synthesis}, we 
 split our algorithm into two communicating
-synthesizers. The first synthesizer\bcp{why
-call this algorithm a synthesizer?},
+searches. The first search,
 \Expand, navigates the space of semantically equivalent regular expressions
@@
 input. It passes the pairs off to the second 
-synthesizer, \GreedySynth, in rank order with the smallest first.
-
-Abstractly, \GreedySynth looks for highest-likelihood lenses between a given pair of
-stochastic regular expressions $\Regex$ and $\RegexAlt$ by performing a
-type-directed search. In more detail, \GreedySynth 
-first normalizes the stochastic regular expressions provided by \Expand into
-\emph{stochastic DNF regular expressions}---a constrained representation of
-stochastic regular expressions with disjunctions  
-distributed over concatenations and with associativity information\bcp{what
-  is ``associativity information''??} 
-removed.  \GreedySynth then uses the syntax of these n-ary DNF regular
-expressions to find highly normalized\bcp{is ``highly normalized'' different
-from ``normalized''?} lenses, in a form we call \emph{simple
-  symmetric n-ary DNF lenses}. These lenses include neither a composition
-operator nor a type equivalence rule. These restrictions mean that each simple
-symmetric n-ary DNF lens is typed by a single pair of stochastic n-ary DNF
-regular expressions, hence \GreedySynth's search space for a given pair of regular
-expressions is finite. The variations that \GreedySynth considers include
-choosing between \IdentityLens{} and \Disconnect lenses and deciding how to
-match multiple occurrences of the same kind of data, \EG, the two occurrences of
-\lstinline{name} in the management and HR data formats (we want to map first
-names to first names and last names to last names\bcp{the relationship of
-  the parenthesized comment to what comes before is not clear}). Finally, \GreedySynth
-yields a simple symmetric lens by converting the n-ary
-forms into the binary forms provided in the surface language.
-
-To illustrate the partnership between \Expand and \GreedySynth,
-consider the lens shown in Figure~\ref{fig:example_lens}. 
-Inferring this lens requires \Expand to use a star-semiring axiom\bcp{are we
-assuming readers know what that is?} to
+search procedure, \GreedySynth, in rank order with the smallest first.
+
+Abstractly, \GreedySynth looks for highest-likelihood lenses between a given
+pair of stochastic regular expressions $\Regex$ and $\RegexAlt$ by performing a
+type-directed search. In more detail, \GreedySynth first normalizes the
+stochastic regular expressions provided by \Expand into \emph{stochastic DNF
+  regular expressions}---a constrained representation of stochastic regular
+expressions with disjunctions distributed over concatenations and with
+concatenations and disjunctions normalized to operating over lists (see \S\ref{subsec:sdnfre}).
+\GreedySynth then uses the syntax of these n-ary DNF regular expressions to find
+normalized lenses, in a form we call \emph{simple symmetric n-ary DNF lenses}.
+These lenses include neither a composition operator nor a type equivalence rule.
+These restrictions mean that each simple symmetric n-ary DNF lens is typed by a
+single pair of stochastic n-ary DNF regular expressions, hence \GreedySynth's
+search space for a given pair of regular expressions is finite. The variations
+that \GreedySynth considers include choosing between \IdentityLens{} and
+\Disconnect lenses and deciding how to match multiple occurrences of the same
+kind of data, \EG, the two occurrences of \lstinline{name} in the management and
+HR data formats (should the algorithm match the first instance of
+\lstinline{name} on the left with the first instance of \lstinline{name} on the
+right, or to the second). Finally, \GreedySynth yields a simple symmetric lens
+by converting the n-ary forms into the binary forms provided in the surface
+language.
+
+To illustrate the partnership between \Expand and \GreedySynth, consider
+synthesizing the lens shown in Figure~\ref{fig:example_lens}. When \GreedySynth
+initially tries to find a lens between \lstinline{emp_salaries} and
+\lstinline{emp_insurance}, it hits some snags: there are complex disjunctions
+and concatenations in \lstinline{emp_salaries} outside the iteration, but no
+such complexity in \lstinline{emp_insurance}. Because of this disparity between
+the formats, \GreedySynth finds an unlikely lens (involving undesired
+\KW{merge}s and \KW{disconnect}s), and the search continues. However, after
+\Expand applies a star-semiring axiom (presented in Figure~\ref{fig:ssr-sre}) to
 rewrite the regular expression \lstinline{("\n" . emp_salary)$^*$} within
 \lstinline{emp_salaries} into
-\lstinline{"" | (("\n" . emp_salary). ("\n" . emp_salary)$^*$)} 
-so that \GreedySynth can find the appropriate lens.  \bcp{This example would
-be more helpful if it were spelled out in full.  What is the path from the
-start of the search to the point where we find this solution?}
-
-This architecture of communicating synthesizers gives us a way to
-enumerate pairs of stochastic regular expressions of increasing
-rank and to efficiently search through them, but it poses a
-problem: when do we stop?\bcp{This question would arise independent of the
-  architecture of the search algorithm!} We may have found a promising lens
-between a pair of stochastic regular expressions, but a different 
-pair we haven't yet discovered may give rise to an even better lens. 
-Or, we may have found a promising lens between a given pair of
-stochastic regular expressions, but the discovered lens is not 
-acceptable to the users. \bcp{How would the algorithm know this?}
-
-The search algorithm must resolve a three-way tension among the quality of
-the inferred lens, the number of examples the user must provide to
-eliminate unwanted lenses, and the amount of time it takes to return a
+\lstinline{"" | (("\n" . emp_salary) . ("\n" . emp_salary)$^*$)}, the
+subcomponents line up better, and the more likely correct lens is instead found.
+
+This architecture of communicating synthesizers gives us a way to enumerate
+pairs of stochastic regular expressions of increasing rank and to efficiently
+search through them, but it poses a problem: when should \Expand stop proposing
+new SRE pairs? We may have found a promising lens between a pair of stochastic
+regular expressions, but a different pair we haven't yet discovered may give
+rise to an even better lens.
+
+The search algorithm must resolve a tension among the quality of
+the inferred lens and the amount of time it takes to return a
 result.  For example, if the algorithm has already found the lens in
 Figure~\ref{fig:example_lens}, we don't want to spend a lot of time
-searching for an even better lens.\bcp{How does this example illustrate a
-  three-way tension?}
+searching for an even better lens.
 
@@
 unrollings to \GreedySynth. The heuristics favor stopping if the current best
-satisfying lens has relatively high\bcp{high compared to what??} likelihood,
-indicating the lens is very 
-promising. The heuristics also favor stopping if \Expand has delivered to
-\GreedySynth all pairs of stochastic regular expressions at a given rank and
-there is a large number of pairs at the next rank, because searching through all
-such pairs will take a long time. With this approach, the algorithm can quickly
-respond with a satisfying lens with relatively high likelihood. If the user is
-unhappy with the result, they can either refine the search by supplying
-additional examples, which serve both to rule out the previously proposed
-lens(es) and to reduce the size of the search space by cutting down on the
-number of satisfying lenses, or they can supply annotations on the source
-and target SREs indicating that certain information
+satisfying lens is very likely, indicating the lens is very promising (for
+example, if the satisfying lens loses no information, the algorithm should
+terminate for no other lens will be more likely). The heuristics also favor
+stopping if \Expand has delivered to \GreedySynth all pairs of stochastic
+regular expressions at a given rank and there is a large number of pairs at the
+next rank, because searching through all such pairs will take a long time. With
+this approach, the algorithm can quickly respond with a satisfying lens with
+relatively high likelihood. If the user is unhappy with the result, they can
+either refine the search by supplying additional examples, which serve both to
+rule out the previously proposed lens(es) and to reduce the size of the search
+space by cutting down on the number of satisfying lenses, or they can supply
+annotations on the source and target SREs indicating that certain information
 either must be retained or must be discarded by the lens (see
-\S\ref{subsec:optimizations}\bcp{this reference to the later section with
-  details is nice, but why don't we have similar references throughout this
-  overview??}).
+\S\ref{subsec:optimizations}).
 
@@
 To characterize likely lenses, we must compute the expected number of bits
-needed to recover a string in one data source from a synchronized\bcp{again,
-needs a definition} string in the
+needed to recover a string in one data source from a synchronized string in the
 other data source. To do this, we must first develop a probabilistic 
@@
 \end{center}
-In the stochastic regular expression\bcp{abbreviate to SRE here and in many
-  places elsewhere}
-$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation $\Probability$
-represents the likelihood that the string is in $\Regex_1$\bcp{That seems
-  both confusing (what is ``the string''??) and probably inaccurate (what if
-  $S_1$ and $S_2$ are the same?)}. In the stochastic regular
-expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability
-represents how
+In the SRE $\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation
+$\Probability$ represent that, if generating arbitrary strings from
+$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, strings should be pulled from
+$\Regex_1$ with probability $\Probability$ (\EG pulling a string from
+$\PRegexOr{\lstinline{"a"}}{\lstinline{"b"}}{.2}$ would retrieve \lstinline{"a"}
+20\% of the time). In the SRE
+$\PRegexStar{\Regex}{\Probability}$, the annotation \Probability represents how
 much more likely having a string involving more than $n$ iterations is than
-having a string involving exactly $n$ iterations. \bcp{Last sentence needs
-  an example.  Probably the first sentence also.}
+having a string involving exactly $n$ iterations (\EG pulling a random string
+from $\PRegexStar{\lstinline{"c"}}{.8}$ would give \lstinline{""} 20\% of the
+time, \lstinline{"c"} 16\% of the time, and \lstinline{"cc"} 12.8\% of the time).
 
@@
 The $\Expand$ algorithm needs to be able to enumerate SREs that are
-``equivalent'' to a given one.
-However, existing work does not provide a means of reasoning about stochastic
-regular expression equivalence\bcp{Does not provide a means of {\em
-    reasoning} about it, or does not even define
-  it?}. Figure~\ref{fig:ssr-sre} shows how we to extend the
-\emph{star-semiring}~\cite{Droste2009} equivalences\bcp{again, we seem to be
-assuming that all readers will already know what they are...} to stochastic regular
-expressions.\bcp{Notation: I don't see any reason for the $s$ superscript in
-$\SSREquiv$}
+``equivalent'' to a given one. However, existing work does not define any notion
+of stochastic regular expression equivalence. Figure~\ref{fig:ssr-sre} shows how
+we to extend the \emph{star-semiring}~\cite{Droste2009} equivalences (a
+\emph{finer} notion of equivalence than semantic equivalence) to
+SREs.
 
@@
 \ifappendices\else{} included with the supplemental
-material\bcp{Right?}\fi.  \bcp{It took me quite a while to find it.  We
+material\fi.  \bcp{It took me quite a while to find it.  We
   should give an actual pointer to a specific theorem in the appendix.}
@@
 The entropy of a data source $S$ is the expected number of bits required to
-describe a random element in $S$ (formally $-\Sigma_{s\in S}\,P(s)\cdot \Log_2P(s)$). The
-entropy of a stochastic regular expression can be computed directly from its
-syntax, when it is unambiguous\bcp{undefined term} and contains no
-empty subcomponents.
+describe a random element in $S$ (formally $-\Sigma_{s\in S}\,P(s)\cdot
+\Log_2P(s)$). The entropy of a SRE can be computed directly from its syntax,
+when each string is uniquely parsable (\IE{} the SRE is unambiguous) and contains
+no empty subcomponents.
 \begin{center}
@@
 \end{center}
-\bcp{This sentence says exactly what the previous one does.  Drop.}We have proven that this syntactic computation corresponds to the entropy of the
-stochastic regular expression when the stochastic regular expression is
-unambiguous (parses strings uniquely).
-
-\bcp{Again, this is a critical definition for understanding the paper, and
-  it is pretty gnarly.  Examples, please!}
+
+For example, the entropy of $\PRegexOr{\lstinline{"a"}}{(\lstinline{"b"})}{.5}$
+is 1. This is expected because on average 1 bit is required to encode whether a
+randomly pulled string is \lstinline{"a"} or \lstinline{"b"}. As an additional
+example, a fixed string has no information content, and so has no entropy.
 
@@
   If $\Regex$ is unambiguous and does not contain $\emptyset$ as a subterm,
-  $\EntropyOf{\Regex}$ is the entropy of $\Regex$.\bcp{Should it not be $-\EntropyOf{\Regex}$??}
+  $\EntropyOf{\Regex}$ is the entropy of $\Regex$.
 \end{theorem}
@@
 recover a regular expression from a stochastic regular expression by deleting
-probability annotations. In a slight abuse of notation\bcp{why is this an
-  abuse of notation??}, if $\Regex$ is a
+probability annotations. In a slight abuse of notation (as we are also using
+these symbols as nonterminals), if $\Regex$ is a
 stochastic regular expression, $\BRegex$ is the regular expression obtained by
@@
 \Leftrightarrow \BRegexAlt$ means that $\Lens$ is a well typed lens between
-$\LanguageOf{\BRegex}$ and $\LanguageOf{\BRegexAlt}$.  \bcp{Ugly typesetting:}
+the language of \BRegex and the language of \BRegexAlt.
 
@@
 
-We now give typing rules and semantics for these combinators. We do not
+We now give typing rules and semantics for (some of) these combinators. We do not
 include the semantics for combinators equivalent to those presented in prior
-work~\cite{symmetric-lenses,Boomerang}\bcp{Huh?}. Furthermore, we present
+work~\cite{Boomerang}. Furthermore, we present
 just \CreateR and \PutR in cases where \CreateL and \PutL are symmetric;
@@
 simplicity of presentation, we use unambiguous regular expressions
-everywhere, even though some lenses can actually be defined over ambiguous
-regular expressions (for example, the identity lens doesn't need any
-ambiguity constraints\bcp{Is this the only example?}); we believe relaxing
+everywhere, even though the identity and disconnect lenses can actually be defined over ambiguous
+regular expressions; we believe relaxing
 this restriction is unproblematic.
@@
 data gets changed from one format to the other, information in the old format is
-lost.  \bcp{We should highlight here the fact that this is a significant
-  loss of power for some examples, compared to full symmetric lenses.}
+lost. This differs from the \OrLens lens of classical symmetric
+lenses, as those can retain such information in the complement (see
+\S\ref{sec:related}).
 %\begin{center}
@@
 \subsection{Lens Costs}
+\label{subsec:lenscosts}
 Our cost metric is based on the expected amount of information required to
@@
 synchronized string in the other format. Note that we would be able to exactly
-calculate the conditional entropy, were it not for \MergeL\bcp{and \MergeR??}. If
+calculate the conditional entropy, were it not for \MergeL and \MergeR. If
 $\MergeLOf{\Lens_1}{\Lens_2} \OfType \Regex \Leftrightarrow
@@
 sidestep such considerations by synthesizing \emph{DNF lenses}---simple symmetric
-lenses that inhabit a disjunctive normal form. DNF lenses have a number of
-properties that facilitate synthesis: they cut down the program search space
-dramatically and they do not contain instances of composition.\bcp{The last
-  sentence basically repeats what the rest of the paragraph already says.}
+lenses that inhabit a disjunctive normal form.
 
@@
 \subsection{Stochastic DNF Regular Expressions}
+\label{subsec:sdnfre}
 Just as SREs are REs with probability annotations,
@@
 \paragraph*{Relevance Annotations}\bcp{Wonder whether this should be its own
-top-level section...?}
+top-level section...?}\afm{I'm not sure, any other thoughts? To me it is
+inherently an optimization, though it's a more important and interesting one
+than others}
 In this work, there are three competing constraints. The tool should (1) return
@@
 
-Conversely, the termination heuristic is satisfied algorithm cuts
-short\bcp{huh??}: if it 
+Conversely, sometimes the termination heuristic is overly permissive and algorithm cuts
+short: if it 
 had continued searching for lenses it would have found the desired one, but the
@@
 
-If synthesizing a lens seems to be taking a long time, it is\bcp{always?} spending time
-checking it cannot find a lens that projects less information. By ``skipping''
-portions of the SRE that \emph{should} be projected, the algorithm will not
-continue searching for a better lens when it has already found the correct one.
-In practice, we implement \Skip by setting the information content of
-$\Skip(\Regex)$ to zero.
-
-If the algorithm returns a lens with too disconnects\bcp{??} and merges, one could
+If synthesizing a lens seems to be taking a long time, it is typically spending
+time checking it cannot find a lens that projects less information. By
+``skipping'' portions of the SRE that \emph{should} be projected, the algorithm
+will not continue searching for a better lens when it has already found the
+correct one. In practice, we implement \Skip by setting the information content
+of $\Skip(\Regex)$ to zero.
+%
+\begin{gather*}
+  \EntropyOf{\Skip(\Regex)} = 0
+\end{gather*}
+%
+If synthesizing a lens between \lstinline{emp_salaries} and
+\lstinline{emp_insurance} from Figure~\ref{fig:minimized-representations} is
+taking overly long, the synthesis procedure is uncomfortable disconnecting
+\lstinline{company} and \lstinline{salary} information. Redefining
+\lstinline{company} and \lstinline{salary} as $\SkipOf{\lstinline{company}}$ and
+$\SkipOf{\lstinline{salary}}$ would set their information content to zero,
+alleviating such issues.
+
+If the algorithm returns a lens with too many disconnects and merges, one could
 provide additional examples to constrain the lens more. However, for large
@@
 cost of any lens that loses ``required'' information to infinity.
-
-\bcp{This discussion seems both too brief---maybe some concrete examples
-  would help?---and too informal: E..g., What do skip and require have to do
-  with probability distributions and entropy?}
-
+%
+\begin{center}
+  \begin{tabular}{rcl}
+    $\REntropyOf{\SRequireOf{\RegexAlt} \Given \Lens, \Regex}$
+    & =
+    & $\begin{cases}
+      \infty & \text{if }\REntropyOf{\SRequireOf{\RegexAlt} \Given \Lens, \Regex}
+      \neq 0\\
+    0 & \text{otherwise}
+  \end{cases}$\\
+
+    $\REntropyOf{\RegexAlt \Given \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, \Regex}$
+    & =
+    & $\begin{cases}
+      \infty & \vspace*{-.3em}\text{if $\RegexAlt$ contains $\SRequireOf{\RegexAlt'}$ as a subexpression,}\\
+      & \text{where $\EntropyOf{\RegexAlt'} \neq 0$}\\
+    [\EntropyOf{\RegexAlt},\EntropyOf{\RegexAlt}] & \text{otherwise}
+  \end{cases}$\\
+  \end{tabular}
+\end{center}
+
+%
+If synthesizing a lens between \lstinline{emp_salaries} and
+\lstinline{emp_insurance} from Figure~\ref{fig:minimized-representations}
+quickly returns a lens with too many disconnects and merges (perhaps
+disconnecting instances of \lstinline{name} in \lstinline{emp_salary} and
+\lstinline{emp_ins}), the termination heuristic is being overly permissive. By
+replacing \lstinline{name} with $\SRequireOf{\lstinline{name}}$ in
+\lstinline{emp_salary} and \lstinline{emp_ins}, name information can no longer
+be lost, and future synthesis calls would continue until a more likely lens is
+found.
 % end synthesis
@@
 \label{sec:evaluation}
-We have implemented symmetric DNF lenses as an extension to
-Boomerang. Specifically, we added symmetric lens combinators to
-Boomerang\bcp{that's exactly what the previous sentence said!}, and
-implemented all of Boomerang's asymmetric lens 
-combinators  using a combination of the symmetric lens combinators presented in this
-paper and symmetric versions of the asymmetric extensions (like matching lenses~\cite{matchinglenses} and quotient lenses\cite{quotientlenses})
-already present in Boomerang.
-We also integrated our synthesis engine into Boomerang. This
-integration allows users to write synthesis tasks 
-alongside lens combinators, incorporate synthesis results into 
-manually-written lenses, and reference previously defined lenses in
-the synthesis of new lenses.
+We have implemented symmetric DNF lenses as an extension to Boomerang.
+We have
+implemented all of Boomerang's asymmetric lens combinators using a
+combination of the symmetric lens combinators presented in this paper and
+symmetric versions of the asymmetric extensions (like matching
+lenses~\cite{matchinglenses} and quotient lenses\cite{quotientlenses}) already
+present in Boomerang. We also integrated our synthesis engine into Boomerang.
+This integration allows users to write synthesis tasks alongside lens
+combinators, incorporate synthesis results into manually-written lenses, and
+reference previously defined lenses in the synthesis of new lenses.
 
@@
 
-While classical symmetric lenses
-are technically more expressive than simple symmetric lenses, it is unclear
-whether this additional expressiveness is useful in practice.\bcp{That's a
-  total cop-out! We have much more to say about this point, and we {\em
-    should} have even more to say than we do.} In our judgement,
-the complications they introduce into synthesis did not justify the potential
-additional expressiveness: because each lens has a custom
-complement, one can no longer specify the put functions through input/output
-examples alone. One alternative would be to enrich specifications with edit
-sequentials; another would be to specify the structure of complements explicitly
-(though the latter would be somewhat akin to specifying the internal state of a
-program). In either case, the complexity of the specifications increases.
+While classical symmetric lenses are technically more expressive than simple
+symmetric lenses, we still believe simple symmetric lenses are useful in
+practice. Indeed, all asymmetric lenses (we later show) are expressible as
+simple symmetric lenses. The primary loss is the loss of ``memory'' within the
+complement. In classical symmetric lenses, \OrLens lenses retain information, so
+users don't lose information when swapping from one disjunct to the other, the
+information is stored in the complement. Simple symmetric lenses have no such
+complement, so they mimic the forgetful \OrLens lens of classical symmetric
+lenses.
+
+In our judgement, the complications they introduce into synthesis did
+not justify the potential additional expressiveness: because each lens has a
+custom complement, one can no longer specify the put functions through
+input/output examples alone. One alternative would be to enrich specifications
+with edit sequentials; another would be to specify the structure of complements
+explicitly (though the latter would be somewhat akin to specifying the internal
+state of a program). In either case, the complexity of the specifications
+increases.
 %% addition, users would have to specify some auxilary 
@@
 
-\paragraph*{Other Lens Formulations}\bcp{This paragraph doesn't add much}
+\paragraph*{Other Lens Formulations}\bcp{This paragraph doesn't add
+  much}\afm{the POPL reviewers complained we didn't have this type of
+  paragraph, I personally agree that it doesn't add much and feels unnecessary}
 Bidirectional programming has an extensive literature, with many extensions onto

### 2019-02-06 papers/BXProposal/main.tex
add bx proposal

\documentclass[a4paper]{article}
\usepackage{graphicx}
\usepackage{listings}
\usepackage{onecolpceurws}
\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}

\title{Expanding the Power of Lens Synthesis With Annotations}

% I've never done a talk before, should this be everybody or just me
\author{
Anders Miltner \\ Princeton University\\
Princeton, NJ 08540 \\ amiltner@cs.princeton.edu
\and
Solomon Maina \\ University of Pennsylvania\\
Philadelphia, PA 19104 \\ smaina@seas.upenn.edu
\and
Kathleen Fisher \\ Tufts University\\
Medford, MA 02155 \\ kfisher@eecs.tufts.edu
\and
Benjamin C. Pierce \\ University of Pennsylvania\\
Philadelphia, PA 19104 \\ bcpierce@cis.upenn.edu
\and
David Walker \\ Princeton University\\
Princeton, NJ 08540 \\ dpw@cs.princeton.edu
\and
Steve Zdancewic \\ University of Pennsylvania\\
Philadelphia, PA 19104 \\ stevez@cis.upenn.edu
}

\institution{}

\begin{document}
\maketitle

\definecolor{dkblue}{rgb}{0,0.1,0.5}
\definecolor{dkgreen}{rgb}{0,0.6,0}
\definecolor{dkred}{rgb}{0.6,0,0}
\definecolor{dkpurple}{rgb}{0.4,0,0.6}
\definecolor{olive}{rgb}{0.4, 0.4, 0.0}
\definecolor{teal}{rgb}{0.0,0.5,0.5}
\definecolor{orange}{rgb}{0.9,0.6,0.2}
\definecolor{lightyellow}{RGB}{255, 255, 179}
\definecolor{vlightyellow}{RGB}{255, 255, 200}
\definecolor{lightgreen}{RGB}{170, 255, 220}
\definecolor{vlightgreen}{RGB}{190, 255, 230}
\definecolor{teal}{RGB}{141,211,199}
\definecolor{darkbrown}{RGB}{121,37,0}

\lstset{ language=Caml, basicstyle=\upshape\sffamily,
keywordstyle=\upshape\sffamily\color{dkpurple}, keepspaces=true,
framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=true,
commentstyle=\itshape\rmfamily,
emph={rep,iterate,synth,collapse,perm,squash,normalize,using,ins,del,lens,let,get,put,rquot,lquot,id,swap,concat,or,disconnect,merge_left,merge_right,const,skip},
emphstyle=\upshape\sffamily\color{dkpurple}, 
columns=fullflexible,
mathescape, 
xleftmargin=1.5em,
% BCP: I find this distracting:
stringstyle=\sffamily\color{dkblue},
}

\begin{abstract}
  Recent work has shown how to synthesize a class of string bijections from
  regular expression specifications. While these bijective string
  transformations can be useful for synchronizing data formats, it may still be
  useful to keep interesting formats synchronized when they are not in bijective
  correspondence. In this talk, we aim to show how we can expand the class of
  synthesized transformations beyond bijections, as long as users are willing to
  provide annotations that help identify a ``bijective core.''
\end{abstract}
\vskip 32pt

\section{Synthesizing Bijective Transformations}

Bidirectional languages are often difficult to program in. The can use
unfamiliar paradigms, and their type systems can sometimes be complex, making
fixing type errors oftentimes nontrivial. This is evident in the language
Boomerang~\cite{boomerang}, a DSL for writing lenses~\cite{Focal2005-long2}
on strings. In Boomerang, users can write bidirectional programs between data
formats: programs are lenses and types are regular expressions representing the
formats the lenses operate on. Instead of requiring users to manually write
bidirectional transformations, we find it preferable to synthesize bidirectional
transformations from their types.

Optician is a Boomerang extension that synthesizes a subset of Boomerang lenses,
bijective lenses. When provided with a pair of regular expressions, Optician
will synthesize one of Boomerang's bijective lenses that converts between the
provided regular expressions.

\section{Building Complex Transformations With Bijective Cores}

Bijections are an important class of bidirectional transformations, but many
useful transformations are not completely bijective. Boomerang permits defining
non-bijective transformations, such as lenses and quotient lenses, though
Optician cannot synthesize them. Lenses are bidirectional transformations that
may lose information when going from one side to the other, while obeying
certain roundtripping properties. Quotient lenses~\cite{quotientlenses} obey
similar properties to lenses, but also permit treating certain pieces of
information as irrelevant -- normalizing them to a canonical representative.

While we wish to make creating these lenses easier, Optician's synthesis algorithm
only works on bijective lenses. However, lenses and quotient lenses seem to have
a ``bijective core'', a component that merely changes the shape of data.
This bijective component can then be surrounded by projection and quotient
components to form a full-fledged quotient lens.

At its base, this surrounding can be done solely with Optician and Boomerang,
but doing so requires repeated code. For example, consider the example of
converting an index of musical composers (with names, birth and death dates, and
nationalities) to a simple index of musical composers (with just names and
nationalities), but with irrelevant whitespace. In this example, ``.''
represents concatenation and ``\lstinline{*}'' represents the Kleene star.

\begin{lstlisting}
  let full_composers : regex = (name . dates . nationality)*
  let composers_nodates : regex = (nationality . " "* . name)*
\end{lstlisting}

In this example, the bijective core is the bijection that transmits nationality and
names (while swapping them), the projection component removes \lstinline{dates},
and the quotient component normalizes the whitespace.

We can build a projection component for \lstinline{full_composers} using
existing Boomerang combinators.
%
\begin{lstlisting}
  let full_projection = iterate (name . (del dates) . nationality)
\end{lstlisting}
%
Here, \lstinline{del} deletes data when going in one direction, and restores it
when going backwards.

We can also build a normalization component for \lstinline{dates} using existing
Boomerang combinators.
%
\begin{lstlisting}
  let ws_norm = normalize (del " "*)
  let nodates_norm = iterate (nationality . (rquot ws_norm) . name)
\end{lstlisting}
%
Here, \lstinline{normalize} turns a lens into a ``canonizer'' (in effect, a
normalization function). In this example, \lstinline{normalize} canonizes an
arbitrarily long list of whitespace to the empty string. Then, the canonizer is
hooked into a lens with \lstinline{rquot}, essentially normalizing strings when
the lens is being operated from right-to-left.

With these normalization and projection components defined, the bijective core
can be synthesized (using \lstinline{synth $R$ <=> $S$}), and the full lens can be
provided by composing all the subcomponents (using ``;'').
\begin{lstlisting}
  let bij_core = synth (targetType full_projection) <=> (sourceType nodates_norm)
  let composers : full_composers <=> composers_nodates = full_projections ; bij_core ; nodates_norm
\end{lstlisting}

Note how in both the projection and quotient component, the bijective aspects
are repeated with no real alteration -- they are merely required scaffolding
for expressing the projections and normalizations.

\section{Projection and Quotient Annotations}

Instead of manually writing explicit canonization and projection functions, we
consider an alternative approach: by annotating our regular expressions with
canonization and projection annotations, we can write our data formats and find
the bijective core in a single step. This approach has already been employed in
synthesizing bidirectional transformations that are bijections (up to
quotients)~\cite{maina+:quotient-synthesis}, but here we extend it with projections as well.

Consider the left-hand format of \lstinline{full_composers}. We expand our
regular expression types with the new combinator \lstinline{skip}. Annotating a
regular expression to be skipped does exactly what \lstinline{full_projection}
does in the previous code -- it sends the minified format to the core synthesis
engine, then composes the generated lens with a lens that deletes the regular
expression from the data format.
%
\begin{lstlisting}
  let full_composers' : annotated_regex = (name . (skip dates) . nationality)*
\end{lstlisting}
%

Consider the right-hand format of \lstinline{composers_nodates}. We expand our
regular expression types with a number of new combinators (to address different
types of quotients), one of which is
\lstinline{squash}~\cite{maina+:quotient-synthesis}. Annotating a regular
expression to be squashed does exactly what \lstinline{nodates_norm} does in the
previous code -- it sends the normalized format to the core synthesis engine,
then composes the generated lens with a lens that normalizes the whitespace to a
representative (in the example, the empty string).
%
\begin{lstlisting}
  let composers_nodates' : annotated_regex = (nationality . (squash " "*) . name)*
\end{lstlisting}
%

Now, with annotated regular expressions, we can synthesize the composers lens much
more compactly.
%
\begin{lstlisting}
  let composers'  = synth full_composers' <=> composers_nodates'
\end{lstlisting}
%
Synthesizing with these annotated regular expressions performs the same steps
that we previously did manually: it breaks the lens into concatenation and
normalization subcomponents, sends the bijective core to the existing synthesis
algorithm, and composes the generated subcomponents with the synthesized
bijective core.

\bibliographystyle{alpha} 
\bibliography{local}
%inline the .bbl file directly for mailing to authors.

\end{document}

### 2019-02-07 papers/BXProposal/main.tex
add main

@@
 
+While we have shown how these annotations can work for quotient lenses, we could
+even imagine getting them to work for symmetric lenses. If we viewed symmetric
+lenses as also having a bijective core (with projections on both sides instead
+of just one), providing \lstinline{skip} annotations on the right could even
+make synthesizing symmetric lenses easier. Of course, this would require
+symmetric lenses in Boomerang for the generated lenses to be well-typed.
+
 \bibliographystyle{alpha} 

### 2019-02-11 papers/BXProposal/main.tex
Edits to BX proposal

@@
 
-\title{Expanding the Power of Lens Synthesis With Annotations}
+\title{Expanding the Power of Lens Synthesis \\ (talk proposal)}
 
@@
 \author{
-Anders Miltner \\ Princeton University\\
-Princeton, NJ 08540 \\ amiltner@cs.princeton.edu
+Anders Miltner \\ Princeton University \\ \ 
+% Princeton, NJ 08540 \\ amiltner@cs.princeton.edu
 \and
-Solomon Maina \\ University of Pennsylvania\\
-Philadelphia, PA 19104 \\ smaina@seas.upenn.edu
+Solomon Maina \\ University of Pennsylvania
+% Philadelphia, PA 19104 \\ smaina@seas.upenn.edu
 \and
-Kathleen Fisher \\ Tufts University\\
-Medford, MA 02155 \\ kfisher@eecs.tufts.edu
+Kathleen Fisher \\ Tufts University
+% Medford, MA 02155 \\ kfisher@eecs.tufts.edu
 \and
-Benjamin C. Pierce \\ University of Pennsylvania\\
-Philadelphia, PA 19104 \\ bcpierce@cis.upenn.edu
+Benjamin C. Pierce \\ University of Pennsylvania
+% Philadelphia, PA 19104 \\ bcpierce@cis.upenn.edu
 \and
-David Walker \\ Princeton University\\
-Princeton, NJ 08540 \\ dpw@cs.princeton.edu
+David Walker \\ Princeton University
+% Princeton, NJ 08540 \\ dpw@cs.princeton.edu
 \and
-Steve Zdancewic \\ University of Pennsylvania\\
-Philadelphia, PA 19104 \\ stevez@cis.upenn.edu
+Steve Zdancewic \\ University of Pennsylvania
+% Philadelphia, PA 19104 \\ stevez@cis.upenn.edu
 }
@@
 Recent work has shown how to synthesize a class of string bijections from
-  regular expression specifications. While these bijective string
-  transformations can be useful for synchronizing data formats, it may still be
-  useful to keep interesting formats synchronized when they are not in bijective
-  correspondence. In this talk, we aim to show how we can expand the class of
-  synthesized transformations beyond bijections, as long as users are willing to
-  provide annotations that help identify a ``bijective core.''
+regular expression specifications. These bijective string transformations
+can be useful for synchronizing simple data formats, but many interesting
+formats require non-bijective lenses. In this talk, we show how to expand
+the class of synthesized transformations beyond bijections to a large subset
+of full symmetric lenses (including all the standard asymmetric lenses)
+using lightweight user annotations that help identify a ``bijective core.''
 \end{abstract}
-\vskip 32pt
 
@@
 
-Bidirectional languages are often difficult to program in. The can use
-unfamiliar paradigms, and their type systems can sometimes be complex, making
-fixing type errors oftentimes nontrivial. This is evident in the language
-Boomerang~\cite{boomerang}, a DSL for writing lenses~\cite{Focal2005-long2}
-on strings. In Boomerang, users can write bidirectional programs between data
-formats: programs are lenses and types are regular expressions representing the
-formats the lenses operate on. Instead of requiring users to manually write
-bidirectional transformations, we find it preferable to synthesize bidirectional
-transformations from their types.
+Bidirectional languages can be difficult to program in, proposing unfamiliar
+programming paradigms, novel syntax, and subtle typing constraints.
+Fortunately, their restricted expressive power, together with the
+fine-grained type systems needed to guarantee that well-types programs
+satisfy strong round-tripping properites, make them a perfect target for
+program synthesis.
 
-Optician is a Boomerang extension that synthesizes a subset of Boomerang lenses,
-bijective lenses. When provided with a pair of regular expressions, Optician
-will synthesize one of Boomerang's bijective lenses that converts between the
-provided regular expressions.
+The {\em Optician} tool is an extension to Boomerang~\cite{boomerang}---a
+DSL for writing lenses~\cite{Focal2005-long2} on strings---that synthesizes
+{\em bijective} Boomerang lenses when provided with a pair of types (regular
+expressions) and a small number of examples (often zero).
 

### 2019-02-12 papers/BXProposal/main.tex
minor changes before pulling bcp

--- 2019-02-07 papers/BXProposal/main.tex
@@
 \begin{abstract}
-  Recent work has shown how to synthesize a class of string bijections from
+  \emph{AFM: I will remove this from the submission, but am including it in the
+    document for people to review. It is due today (February 12)} Recent work has shown how to synthesize a class of string bijections from
   regular expression specifications. While these bijective string
@@
   correspondence. In this talk, we aim to show how we can expand the class of
-  synthesized transformations beyond bijections, as long as users are willing to
-  provide annotations that help identify a ``bijective core.''
+  synthesized transformations beyond bijections, for example into classical
+  lenses, quotient lenses, and even symmetric lenses, as long as users are
+  willing to provide annotations that help identify a ``bijective core.''
 \end{abstract}
@@
 
-Bidirectional languages are often difficult to program in. The can use
-unfamiliar paradigms, and their type systems can sometimes be complex, making
-fixing type errors oftentimes nontrivial. This is evident in the language
-Boomerang~\cite{boomerang}, a DSL for writing lenses~\cite{Focal2005-long2}
-on strings. In Boomerang, users can write bidirectional programs between data
-formats: programs are lenses and types are regular expressions representing the
-formats the lenses operate on. Instead of requiring users to manually write
-bidirectional transformations, we find it preferable to synthesize bidirectional
-transformations from their types.
+Learning any new domain-specific language takes time and energy. Any support we
+can provide to help users ``get over the hump'' of learning a new DSL is going
+to increase the rate of adoption for such a DSL. Indeed, due to the need to
+think in ``two directions'' at once, bidirectional languages are more difficult
+to learn than usual. This difficulty is evident in the language
+Boomerang~\cite{boomerang}, a DSL for writing bidirectional programs on strings.
+In Boomerang, users can write bidirectional programs between data formats.
+Boomerang programs encode lenses a class of bidirectional objects. Boomerang
+expressions are bidirectional programs on strings, and Boomerang types are
+regular expressions representing the formats the lenses operate on. Instead of
+requiring users to manually write bidirectional transformations, we find it
+preferable to synthesize these transformations from their types.
 
@@
 Bijections are an important class of bidirectional transformations, but many
-useful transformations are not completely bijective. Boomerang permits defining
-non-bijective transformations, such as lenses and quotient lenses, though
-Optician cannot synthesize them. Lenses are bidirectional transformations that
-may lose information when going from one side to the other, while obeying
-certain roundtripping properties. Quotient lenses~\cite{quotientlenses} obey
-similar properties to lenses, but also permit treating certain pieces of
+useful transformations are not completely bijective. Boomerang expressions are
+not necessarily bijections -- they encode lenses~\cite{Focal2005-long2}, and
+various lens extensions like quotient lenses~\cite{quotientlenses}. Lenses are
+bidirectional transformations that may lose information when going from one side
+to the other, while obeying certain roundtripping properties. Quotient lenses
+obey similar properties to lenses, but also permit treating certain pieces of
 information as irrelevant -- normalizing them to a canonical representative.
@@
 
-At its base, this surrounding can be done solely with Optician and Boomerang,
-but doing so requires repeated code. For example, consider the example of
-converting an index of musical composers (with names, birth and death dates, and
-nationalities) to a simple index of musical composers (with just names and
-nationalities), but with irrelevant whitespace. In this example, ``.''
+At its base, surrounding bijections in this way can be done solely with Optician
+and Boomerang, but doing so requires repeated code. For example, consider the
+example of converting an index of musical composers (with names, birth and death
+dates, and nationalities) to a simple index of musical composers (with just
+names and nationalities), but with irrelevant whitespace. In this example, ``.''
 represents concatenation and ``\lstinline{*}'' represents the Kleene star.
@@
 regular expression to be skipped does exactly what \lstinline{full_projection}
-does in the previous code -- it sends the minified format to the core synthesis
-engine, then composes the generated lens with a lens that deletes the regular
-expression from the data format.
+does in the previous code -- it sends the format, after removing the projected
+portions, to the core synthesis engine, then composes the generated lens with a
+lens that deletes the regular expression from the data format.
 %
@@
 
+Furthermore, we think this approach could scale to other lens extensions, like
+matching lenses and dictionary lenses, and other types of bidirectional
+constructs beyond merely lenses. All that is needed for this approach is a
+user-provided a reduction to a ``bijective core,'' and a means to synthesize
+that core. This means that, if a bijective synthesis algorithm is discovered for
+(for example) trees or graphs, then annotations can be used to synthesize more
+complex bidirectional programs for those structures as well.
+
 \bibliographystyle{alpha} 

### 2019-02-12 papers/BXProposal/main.tex

--- 2019-02-11 papers/BXProposal/main.tex
@@
 \begin{abstract}
+emph{AFM: I will remove this from the submission, but am including it in the
+    document for people to review. It is due today (February 12)}
 Recent work has shown how to synthesize a class of string bijections from

### 2019-02-12 papers/SymmetricLenses/main.tex

@@
 inherently an optimization, though it's a more important and interesting one
-than others}
+than others}\bcp{My impression was that, before we invented these
+annotations, things were really not working very well.  So they may be
+``just an optimization,'' but they are a super-important one.  Moreover,
+they strongly impact the way the user interacts with the system.  I think
+they should figure prominently in the way we explain what we've done.}
 In this work, there are three competing constraints. The tool should (1) return

### 2019-02-12 papers/BXProposal/main.tex
wordsmithing and updated abstract

@@
 \begin{abstract}
-emph{AFM: I will remove this from the submission, but am including it in the
-    document for people to review. It is due today (February 12)}
-Recent work has shown how to synthesize a class of string bijections from
-regular expression specifications. These bijective string transformations
-can be useful for synchronizing simple data formats, but many interesting
-formats require non-bijective lenses. In this talk, we show how to expand
-the class of synthesized transformations beyond bijections to a large subset
-of full symmetric lenses (including all the standard asymmetric lenses)
-using lightweight user annotations that help identify a ``bijective core.''
+  {\color{dkblue} \emph{AFM: I will remove this from the submission, but am including it in the
+    document for people to review. It is due today (February 12)}} Recent work
+  has shown how to synthesize a class of string bijections from regular
+  expression specifications. These bijective string transformations can be
+  useful for synchronizing simple data formats, but many interesting formats
+  require non-bijective lenses. In this talk, we show how to expand the class of
+  synthesized transformations beyond bijections to quotient lenses and a large
+  subset of full symmetric lenses (including all the standard asymmetric lenses)
+  through lightweight user annotations that help identify a ``bijective core.''
 \end{abstract}
@@
 Bijections are an important class of bidirectional transformations, but many
-useful transformations are not completely bijective. Boomerang expressions are
-not necessarily bijections -- they encode lenses~\cite{Focal2005-long2}, and
+useful transformations are not completely bijective. In particular, Boomerang
+expressions are not necessarily bijections -- they encode lenses and
 various lens extensions like quotient lenses~\cite{quotientlenses}. Lenses are
@@
 
-While we wish to make creating these lenses easier, Optician's synthesis algorithm
-only works on bijective lenses. However, lenses and quotient lenses seem to have
-a ``bijective core'', a component that merely changes the shape of data.
-This bijective component can then be surrounded by projection and quotient
-components to form a full-fledged quotient lens.
+While we wish to make creating these lenses easier, Optician's synthesis
+algorithm only works on bijective lenses. Fortunately, many lenses and quotient
+lenses seem to have a ``bijective core'', a component that merely changes the
+shape of data, but does not remove any data. This bijective component can then
+be surrounded by projection and quotient components to form a full-fledged
+quotient lens.
 
-At its base, surrounding bijections in this way can be done solely with Optician
-and Boomerang, but doing so requires repeated code. For example, consider the
-example of converting an index of musical composers (with names, birth and death
+Surrounding bijections in this way can be already be done with Optician
+and Boomerang, but doing so requires repeated code. Consider the
+converting an index of musical composers (with names, birth and death
 dates, and nationalities) to a simple index of musical composers (with just
-names and nationalities), but with irrelevant whitespace. In this example, ``.''
+names and nationalities and irrelevant whitespace). In this example, ``.''
 represents concatenation and ``\lstinline{*}'' represents the Kleene star.
@@
 In this example, the bijective core is the bijection that transmits nationality and
-names (while swapping them), the projection component removes \lstinline{dates},
-and the quotient component normalizes the whitespace.
+names (while swapping them), the projection component removes \lstinline{dates}
+when operating from left-to-right and recovers it when operating from right-to-left,
+and the quotient component normalizes the whitespace to the empty string.
 
@@
 arbitrarily long list of whitespace to the empty string. Then, the canonizer is
-hooked into a lens with \lstinline{rquot}, essentially normalizing strings when
-the lens is being operated from right-to-left.
+hooked into a lens with \lstinline{rquot}, normalizing strings when the lens is
+applied right-to-left, and performing the identity when applied from left-to-right.
 
@@
 user-provided a reduction to a ``bijective core,'' and a means to synthesize
-that core. This means that, if a bijective synthesis algorithm is discovered for
+that core. If a bijective synthesis algorithm is discovered for
 (for example) trees or graphs, then annotations can be used to synthesize more

### 2019-02-12 papers/SymmetricLenses/main.tex
add main and macros fixes appendix and other stuff

--- 2019-02-05 papers/SymmetricLenses/main.tex
@@
   bijective lenses} over string data---given a pair of types (regular
-expressions) and a small number of examples.  \bcp{Why is the next line not
-  indented?} \afm{I have truly no idea, deferring until I get a stackoverflow answer.}
-
-We show how to extend this synthesis algorithm to a much wider class of
+expressions) and a small number of examples.
+
+\hspace*{1em}We show how to extend this synthesis algorithm to a much wider class of
 lenses, which we call {\em simple symmetric lenses}, including not only all
@@
 concluding thoughts (\S\ref{sec:conc}).
-Elided proofs can be found in the appendices\bcp{The appendices are pretty
-  hideous and hard to navigate!!}
+Elided proofs can be found in the appendices
 \ifappendices\else included with the supplemental material\fi.
@@
 ]
-Name,Company
+FirstLast,Company
 Jane Doe,Healthcare Inc.
@@
 let emp_ins = name . " " . name "," company in
-let header = "Name,Company" in
+let header = "FirstLast,Company" in
 let emp_insurance = header . ("\n" . emp_ins)$^*$
@@
   \resizebox{\columnwidth}{!}{%
-    \begin{tabular}{|>{\columncolor{vlightyellow}}c|>{\columncolor{vlightgreen}}c|}
+    \begin{tabular}{|>{\columncolor{vlightyellow}}c|>{\columncolor{vlightblue}}c|}
       \hline
@@
       \cellcolor{lightyellow}Management's type & 
-      \cellcolor{lightgreen} HR's type\\
+      \cellcolor{lightblue} HR's type\\
       \hline
-      \cellcolor{lightyellow}\listingC & \cellcolor{lightgreen}\listingD \\
+      \cellcolor{lightyellow}\listingC & \cellcolor{lightblue}\listingD \\
       \hline
@@
   used by management and HR at a US company to represent employee salaries and health
-  insurance providers, respectively. \bcp{The colors are a bit loud,
-    especially since there is so little color elsewhere.}
-  %
-  \bcp{Suggest replacing ``Name,Company'' with ``First Last, Company'' (too
-    many ``name''s flying around)}
-  %
-  \bcp{More substantially, thinking about this example makes me wonder about
-    a whole different way of specifying simple symmetric lenses: rather than
-    just giving corresponding source and target pairs, could we give triples
-    of the form ``new source / old target / new target'' and ``new target /
-    old source / new source''?  I know these would be a bit more work to
-    write, but they would also give the synthesis procedure way more
-    information to work with---it seems to me it should be able to do much
-    less guessing about where things are disconnected.  Would this allow us
-    to drop all the information theory?  Or would we still need it?  (In
-    retrospect, this seems like a fairly obvious thing to try, so it seems
-    important to try it and report on how it performs relative to what we
-    propose here...)}\afm{Oh sorry, yes we can write such examples.
-I personally found that typically the same information could be provided through a series of 2 creates (aka: if we want a disconnect: creater a = b; creater a' = b:  if we want an identity: creater a = a; creater a' = a').
-However, we also permit putr and putl examples of the proposed form in the tool
-Much like other aspects of the work, any situation with examples ultimately only constrains the subcomponents of the lens corresponding to that specific parse tree (aka: if I give an example about A in A|B, it only constrains "A").
-I don't think it removes the need for the information theory, as all the other non-constrained subcomponents of the lens can still be "anything."
-}  }
+  insurance providers, respectively.
+  }
   \label{fig:minimized-representations}
@@
 \begin{proof}
-  See the appendix%
-\ifappendices\else{} included with the supplemental
-material\fi.  \bcp{It took me quite a while to find it.  We
-  should give an actual pointer to a specific theorem in the appendix.}
+  See the appendix %
+\ifappendices  Theorem~\ref{thm:semantics-correctness} \else{} included with the supplemental
+material\fi.
 \end{proof}
@@
 just \CreateR and \PutR in cases where \CreateL and \PutL are symmetric;
-full details can be found in an appendix\bcp{where,
-  exactly??}\ifappendices\else{} (in the supplemental materials)\fi. For
+full details can be found in an appendix\ifappendices \S\ref{sec:appendixlenses}  \else{} (in the supplemental materials)\fi. For
 simplicity of presentation, we use unambiguous regular expressions
@@
 The details of these definitions are syntactically tedious, but not
-intellectually difficult. We elide them here but include them in the appendix.\bcp{Where?}
+intellectually difficult. We elide them here but include them in the
+appendix \ifappendices \S\ref{sec:appendixdnf}.  \else{} (in the supplemental
+materials).\fi
 
@@
 
-\subsection{Optimizations and Relevance Annotations}
+\subsection{Optimizations}
 \label{subsec:optimizations}
@@
 
-\paragraph*{Relevance Annotations}\bcp{Wonder whether this should be its own
-top-level section...?}\afm{I'm not sure, any other thoughts? To me it is
-inherently an optimization, though it's a more important and interesting one
-than others}
+\subsection{Relevance Annotations}
+
 In this work, there are three competing constraints. The tool should (1) return
@@
 lenses and two previous lens languages:  classical symmetric lenses
-and asymmetric lenses.  We also comment on related
+and asymmetric lenses. Proofs about these relationships are included in
+the appendix
+\ifappendices  \S\ref{sec:appendixforget}. \else{} included with the supplemental
+material.\fi
+
+We also comment on related
 program synthesis techniques.
@@
 
-\paragraph*{Symmetric Lenses full Detail}
+\section{Symmetric Lenses full Detail}
+\label{sec:appendixlenses}
+
+Here we provide fullly detailed information on symmetric lenses -- full typing
+rules and full semantics.
+
 \begin{centermath}
@@
 
-\paragraph*{Symmetric DNF Lenses: Syntax, Cost, Typing, and Semantics}
+\section{Symmetric DNF Lenses: Syntax, Cost, Typing, and Semantics}
+\label{sec:appendixdnf}
+
+Here we give full details on symmetric DNF lenses, including syntax, cost,
+typing, and semantics.
 
@@
 \section{Forgetful Symmetric Lenses}
+\label{sec:appendixforget}
 
@@
   \label{prop:forget-rl}
-  Let $\Lens$ be a forgetful symmetric lens.  If $(x_1',c_1') = \Lens.\PutL \App
-  (y,\Snd \App (\Lens.\PutR \App (x,c_1)))$, and
+  Let $\Lens$ be a forgetful symmetric lens.  If $(x_1',c_1') = \Lens.\PutL \App$
+  $(y,\Snd \App (\Lens.\PutR \App (x,c_1)))$, and
   $(x_2',c_2') = \Lens.\PutL \App
@@
   \label{prop:forget-lr}
-  Let $\Lens$ be a forgetful symmetric lens.  If $(y_1',c_1') = \Lens.\PutR \App
-  (x,\Snd \App (\Lens.\PutL \App (y,c_1)))$, and
+  Let $\Lens$ be a forgetful symmetric lens.  If $(y_1',c_1') = \Lens.\PutR \App$
+  $(x,\Snd \App (\Lens.\PutL \App (y,c_1)))$, and
   $(x_2',c_2') = \Lens.\PutL \App
@@
 
-\section{Forgetful Symmetric Lenses}
-\begin{theorem*}
+\section{SREs and Information Theory}
+\begin{theorem}
+  \label{thm:semantics-correctness}
   If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
   \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
-\end{theorem*}
+\end{theorem}
 The following lemma is
@@
 By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$, hence\bcp{Don't understand what's going on with the $-$
-  signs in the last few lines here.}
+  signs in the last few lines here.}\afm{in the definition of entropy there is a negative. this is to
+  address that we're taking the log of a probability, which inherently will
+  return a negative number}
 \begin{align*}
@@
 
-\bcp{``Since...then...'' is not english.}
-Since $S$ is unambiguous, then $S_1$ and $S_2$ are also unambiguous, and
+Since $S$ is unambiguous, it must be that $S_1$ and $S_2$ are also unambiguous, and
 every $s \in \mathcal{L}(S_1 \; |_p \; S_2)$ is either in $\mathcal{L}(S_1)$

### 2019-02-16 papers/SymmetricLenses/main.tex
dunno

--- 2019-02-05 papers/SymmetricLenses/main.tex
@@
-\newif\ifdraft\drafttrue  % set true to show comments
-% \newif\ifdraft\draftfalse  % set true to show comments
+% \newif\ifdraft\drafttrue  % set true to show comments
+\newif\ifdraft\draftfalse  % set true to show comments
 \newif\ifanon\anonfalse    % set true to suppress names, etc.
@@
 \PassOptionsToPackage{usenames,dvipsnames,svgnames,table}{xcolor}
-\documentclass[acmsmall,review,screen,anonymous]{acmart}
+%\documentclass[acmsmall,review,screen,anonymous]{acmart}
+\documentclass[acmsmall,review,screen]{acmart}
 \settopmatter{printacmref=false,printccs=false}
@@
 
+% BCP: To make sure we don't leave comments on!
+\bcp{XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX XXXX }
+
 {\em Lenses} are programs that can be run both ``front to back'' and ``back
-to front,'' allowing updates to both input and output data to be transferred
-in both directions.
-Since their introduction by~\citet{Focal2005-long}, lenses have been extensively
-studied, extended, and applied.  Recent work~\cite{optician} has also demonstrated how
+to front,'' allowing data and updates to be transformed in two directions.
+Since their 
+introduction by~\citet{Focal2005-long}, lenses have been extensively
+studied and applied.  Recent work~\cite{optician} has also demonstrated how
 techniques from {\em type-directed program synthesis} can be used to
-efficiently synthesize a simple class of lenses---so-called {\em
+efficiently synthesize a very simple class of lenses---so-called {\em
   bijective lenses} over string data---given a pair of types (regular
-expressions) and a small number of examples.  \bcp{Why is the next line not
-  indented?} \afm{I have truly no idea, deferring until I get a stackoverflow answer.}
+expressions) and a small number of examples.
 
 We show how to extend this synthesis algorithm to a much wider class of
-lenses, which we call {\em simple symmetric lenses}, including not only all
-of the bijective lenses and the more widely used ``asymmetric''
-lenses~\cite{Focal2005-long}, but also a rich subset of the full-blown
-``symmetric lenses'' proposed by \citet{symmetric-lenses}.
-
-Synthesizing simple symmetric lenses is substantially more challenging than
+lenses, which we call {\em simple symmetric lenses}, including both
+bijective lenses and the more widely used ``asymmetric''
+lenses~\cite{Focal2005-long}, as well as a rich subset of the full-blown
+``symmetric lenses'' proposed by \citet{symmetric-lenses}.  Simple symmetric
+lenses are of independent theoretical interest, being the largest class of
+symmetric lenses that do not rely on persistent internal state.
+
+Synthesizing simple symmetric lenses is substantially more challenging that
 synthesizing bijective lenses: Since some of the information on each side
 can be ``disconnected'' from the other side, there will in general be {\em
-  many} lenses that agree with a given example. To guide the search process,
-we use {\em stochastic regular expressions} and ideas from information
-theory to estimate the amount of information propagated by a candidate lens,
-in conjunction with user annotations identifying parts of the source and
-target data structures as either {\em irrelevant} or {\em essential}.
-
-Simple symmetric lenses are also of independent theoretical interest, being
-the largest class of symmetric lenses that do not rely on persistent
-internal state.
+  many} lenses that agree with a given example.  To guide the search
+process, we use {\em stochastic regular expressions} and ideas from
+information theory to estimate the amount of information propagated by a
+candidate lens.
 
@@
 
-\ifanon
+%\ifanon
 %\authorinfo{}
@@
 %           {}
-\maketitle
+%\maketitle
 % \vspace*{-6cm}
-\else
+%\else
 \author{Anders Miltner}
@@
 \maketitle
-\fi
+%\fi
 
@@
 \noindent
-In today's data-rich world, similar information is often
+In today's big-data world, similar information is sometimes
 stored in different places and in different formats. For instance, electronic
@@
   criticized on the same grounds.}
-\afm{I still think that there are situations where the embroidery file makes
-  sense (I'm sharing my design with a friend, who then tweaks it and sends it
-  back. There may also need to file-specific information, where one contains
-  parameters like how fast to embroider and some doesn't, to be honest I don't
-  know much about embroidery though. Thinking about other examples, however.)}
-
-One convenient way to maintain consistency between such varied data formats
+%
+\bcp{Also, it would be nice to be able to give a quick example of a truly
+  symmetric case---do any of these examples work for that?}\afm{h-event and
+  vCalendar are truly symmetric. Even the fields related to location alone
+  are truly symmetric (though it is subtle to see this).}
+
+One convenient way to maintain consistency between these varied data formats
 is to use a \emph{lens}~\cite{Focal2005-long}---a bidirectional program that
@@
 between different representations.
-Domain-specific languages for writing lenses can facilitate building
-principled data transformations.  For example, Boomerang~cite{boomerang} is a
-language for expressing bidirectional string transformations.
-
-\textit{Synthesizing} lenses can make the task even easier. Indeed, lens
-languages, like Boomerang, are excellent targets for search-based synthesis, since their
-specialized, sub-Turing-complete syntax and expressive type systems drastically
-limit the space of possible programs.
-%
-In particular, the recent Optician~\cite{optician} enhancement to Boomerang has
-demonstrated that it is feasible to synthesize quite complex examples, albeit
-only for a restricted class of lenses---so-called bijective lenses.
-Specifically, given source and target formats plus a small number of examples of
-corresponding data, Optician's synthesis algorithm is guaranteed to find a
-bijective lens matching the examples, if one exists. Indeed, it can often find
-an appropriate lens in less than a second with no examples at all.
+
+While domain-specific languages for writing lenses can facilitate building
+principled data transformations, \textit{synthesizing} lenses 
+can make the task even easier.  Indeed, lens languages offer excellent
+targets for synthesis: their specialized, sub-Turing-complete syntax
+and expressive type systems drastically
+reduce the space of possible programs, making search-based synthesis easier. 
+%
+Recent work~\cite{optician} has exploited this observation, demonstrating that
+it is feasible to synthesize quite complex examples, albeit only for a
+restricted class of lenses---so-called bijective lenses. Specifically, given 
+source and target formats plus a small number of
+examples of corresponding data, the bijective synthesis algorithm will find a
+bijective lens matching the examples, if one exists. Often,
+it can find an appropriate lens in less than a second with
+no examples at all.
 %, with just types describing the source and target formats.
@@
 %   third person, even if that's a bit awkward, but otherwise citing it is OK.}
-One reason that bijective lens synthesis can be so effective, even relative to
+One reason that bijective lens synthesis is so effective, even relative to
 successful synthesis tools in other domains, is that there are typically not
-very many bijections between a given pair of data formats. 
-If the synthesis algorithm can
-find any bijection at all, it is fairly likely to be the intended one.
-
-However, the set of real-world use cases for bijective lenses is also
-somewhat limited.  A more common situation is where two real-world data
-formats share just {\em some} of their information content.  For instance,
-two related database tables might share four out of five columns, but each
-might have a different fifth column.  Or, one ad-hoc system configuration
-file might include some format-specific meta data, such as a date or a
-reference number, whereas the same configuration file on another operating
-system does not.  Indeed among the 48 benchmark problems described in
-Section \ref{sec:evaluation} below, all of the benchmarks taken from Flash
-Fill~\cite{flashfill} and many of the benchmarks that synchronize between
-two ad hoc file formats have this characteristic, 13 benchmarks in
-total.
-
-Can Optician's basic synthesis procedure be extended to a richer class of
-lenses? In particular, what about \emph{symmetric
-  lenses}~\cite{symmetric-lenses}---a much larger class that includes
-bijective lenses, ``asymmetric'' lenses (where the transformation from a
-source structure to a target can throw away information that must
-then be restored when transferring from target to source), and even more flexibile
-transformations that allow each side to retain information not present in
-the other?
+very many bijections between related data formats. 
+Hence, if the synthesis algorithm can
+find any bijection at all, it is likely to be the right one.
+
+However, many real-world data formats share just part of their information
+content with other representations.  For instance, two related
+database tables might share four out of five columns, but each might have a
+different fifth column. 
+Or, one ad-hoc system configuration file might include some format-specific meta
+data, such as a date or a reference number, whereas the same configuration
+file on another operating system does not. 
+Indeed of the 48 benchmark problems described in Section \ref{sec:evaluation},
+all of the benchmarks taken from Flash Fill~\cite{flashfill} and many of the
+benchmarks that synchronize between two ad hoc file formats have this
+characteristic, 13 benchmarks in total. Consequently, synthesizing only strict
+bijections is often not useful in practice.
+
+Could we extend synthesis tools to more kinds of lenses?  In particular,
+what about \emph{symmetric lenses}~\cite{symmetric-lenses}, a class that
+includes bijective lenses, ``asymmetric'' lenses where the transformation
+from $S$ to $T$ can throw away information that must then be restored when
+returning from $T$ to $S$, and even more flexibile transformations that
+allow each side to retain information not present in the other?
+
+\bcp{We use $S$ and $T$ here but $X$ and $Y$ below.}
 
@@
 by ~\citet{symmetric-lenses} operate
-over three structures: the ``left'' structure $X$, the ``right'' $Y$ and a complement
+over three structures: the source $S$, the target $T$ and a complement
 $C$ that contains the information not
-present in either $X$ or $Y$. Logistically, these complements must be stored
+present in either $S$ or $T$. Logistically, these complements must be stored
 and managed somehow.  More fundamentially, complements complicate
@@
 lenses that do not require external ``memory'' to recover data from past
-instances of $X$ or $Y$ when making a round trip. They only need the most
+instances of $S$ or $T$ when making a round trip. They only need the most
 recent instance.
@@
 
-This leaves us with just the second difficulty with synthesizing symmetric
-lenses: While the number of bijective lenses between two given formats is
-typically tiny, the number of simple symmetric lenses is typically enormous.
-If a na\"ive search algorithm just selects the first simple symmetric lens
-it finds, this will generally \emph{not} be the one the user wanted.  We
-need a new principle for identifying ``more likely'' lenses and a more
-sophisticated synthesis algorithm that uses this principle to search the
-space more intelligently. For these, we turn to information theory.
-
-We consider a ``likely'' lens to be one that propagates ``a lot'' of information
-from the left data format to the right and vice versa. Conversely, an unlikely
-lens is one that requires a large amount of additional information to recover
-either of the formats given the other. For many application scenarios, users
-will want to synchronise formats with lots of shared information (and lenses
-with little shared information are typically easier to write). Our tool is built
-primarily for these complex scnearios, and so lenses that propagate more
-information are preferable in our context to ones that propagate less.
+However, we are still left with a second difficulty with synthesizing
+symmetric lenses: While the number of bijective lenses between two related
+formats is typically tiny, the number of simple symmetric lenses is
+typically huge.  Thus, if a na\"ive search algorithm selects the first
+simple symmetric lens it finds, this will generally \emph{not} be the one we
+want.  We need a new principle for identifying ``more likely'' lenses and a
+more sophisticated synthesis algorithm that uses this principles to search
+the space more intelligently.
+
+For these, we turn to information theory. 
+We consider a ``likely'' lens to be one that propagates ``a lot'' of
+information from source to 
+target and vice versa. Conversely, an unlikely lens requires a
+large amount of additional information to recover the target given a
+source.  
+This categorization follows from the intuition
+that users will typically want to synchronise formats with lots
+of shared information, and so lenses that propagate
+more information are preferable to ones that propagate less.
 %
@@
 probability distribution, we can calculate the shared information content of
-two formats and our measure of likelihood.
+two formats\bcp{It's not obvious at this point what the {\em shared}
+  information content has to do with anything.} and our measure of likelihood.
 
@@
 
-With simple symmetric lenses and our SRE-based likelihood measure in hand, we
-develop a new algorithm for synthesizing likely lenses. At its core, the
-algorithm performs a type-directed search between the descriptions of the source
-and target formats, measuring success using the likelihood measure. The
-interesting complications arise from the need to deal with regular-expression
-equivalences. There are infinitely many regular expressions equivalent to a
-given one, and the lens returned by a type-directed search will in general
-depend on which of the possible representations are chosen for its source and
-target formats. Moreover, if Boomerang's type-equivalence rule is disallowed,
-there may be {\em no} well-typed lens connecting the source and target formats
-selected by the user: in general, the synthesis algorithm has to search through
-equivalent regular expression types for which a suitable lens exists. To tame
-this complexity, we divide the synthesis algorithm into two communicating search
-procedures (Figure~\ref{fig:high-level-algorithm}), following~\citet{optician}.
-The first, \Expand, uses rewriting rules to iteratively propose new pairs of
-stochastic regular expressions equivalent to the original pair. The second,
-\GreedySynth, takes a pair of SREs as input and uses a greedy, type-directed
-algorithm to find a simple symmetric lens between them, returning the lens and
-its likelihood score to \Expand. The whole synthesis algorithm heuristically
-terminates when the most likely lens is no longer unlikely enough to justify
-the time cost of searching through newly generated SREs.
+With simple symmetric lenses and our SRE-based likelihood measure in
+hand, we develop a new algorithm for synthesizing likely lenses.
+At its core, the algorithm performs a type-directed search between the 
+descriptions of the source and target formats, measuring success using
+the likelihood measure.
+The interesting complications arise from the need to deal with
+regular-expression equivalences.  There are infinitely many regular
+expressions equivalent to a given one, and the lens returned by a
+type-directed search will depend which of the infinitely many possible
+representations are chosen for its source and target formats.  Moreover,
+there may be {\em no} lens connecting the 
+source and target formats selected by the user: in general, the synthesis
+algorithm has to search to find equivalent ones for which a suitable lens
+exists.  
+To tame this complexity, we divide the synthesis algorithm into two
+communicating search procedures (Figure~\ref{fig:high-level-algorithm}),
+following~\citet{optician}. 
+The first, \Expand, uses rewriting rules to iteratively
+propose new pairs of stochastic regular expressions equivalent to the
+original pair.
+The second, \GreedySynth, takes a pair of SREs as input
+and uses a greedy, type-directed algorithm to find a simple symmetric
+lens between them, returning the lens and its likelihood score
+to \Expand.  The whole synthesis algorithm terminates when comparing the
+likilihood score of the best lens found so far with 
+the complexity of newly generated SREs 
+suggests that further searching is unlikely to be fruitful.
 
@@
 
-We have added this synthesis procedure to the Boomerang system and evaluated
-its effectiveness on a benchmark suite of 48 problems, including data-cleaning
-tasks taken from Flash Fill~\cite{gulwani-popl-2014}, view-maintenance tasks
-taken from Augeas~\cite{augeas2}, and ad-hoc synchronization tasks from a
-variety of other sources (mostly from the bidirectional programming literature).
-The user can control the search process via examples and annotations
-that informs the synthesis algorithm as to whether certain bits of information
-are irrelevant, or must maintained.
-
-In summary, our contributions are:
+We added this synthesis procedure to the Boomerang implementation and
+evaluated its effectiveness on a benchmark suite of 48 problems, including
+data-cleaning tasks taken from Flash Fill~\cite{gulwani-popl-2014},
+view-maintenance tasks taken from Augeas~\cite{augeas2}, and ad-hoc
+synchronization tasks from a variety of other sources (mostly from the
+bidirectional programming literature).  The ``programmer'' can control the
+search process via examples and a {\em tuning parameter} that governs how
+much effort is spent trying to improve on existing solutions.
+
+In summary, our contributions are:\bcp{The section references do not proceed
+  in order, so this doesn't work very well as an outline of the paper.
+  Should we try to reorder the contributions?  Write a separate outline?
+  Live with it as-is?}\afm{Personally, I don't see it as a problem.  It's
+  nice when the contributions mirror the paper, I am not jarred when they
+  don't.}\bcp{But then the reader has no birds-eye view of the paper at
+  this point.  I tried reorganizing, which I think also helps clarify the
+  contributions.} 
 \begin{itemize}
@@
 %
-We present a collection of combinators for building simple symmetric lenses and
+We
+present a collection of combinators for building simple symmetric lenses and
 describe how to calculate their likelihoods (\S\ref{sec:ssl}).
-\item We give an algorithm for synthesizing lenses built from these combinators,
-  using a novel application of stochastic regular expressions to guide the
-  search process (\S\ref{sec:synthesis}) and prove that our metrics have deep
-  semantic meaning.
+\item We give
+an algorithm for synthesizing lenses built from these combinators, using a 
+novel application of stochastic regular expressions to guide the search
+process (\S\ref{sec:synthesis})\bcp{``and prove it correct''???}.  
 %
 \item We extend the Boomerang implementation with simple symmetric lenses,
-  demonstrating that they integrate smoothly with other lens extensions such as
-  quotient lenses~\cite{quotientlenses} and matching
-  lenses~\cite{matchinglenses}. We evaluate our implementation, the effects of
-  various optimizations, and our inclusion of relevance annotations on a set of
-  48 lens synthesis benchmarks drawn from data cleaning, view maintenance, and
-  file synchronization tasks. The system can synthesize simple symmetric lenses
-  for all of the benchmarks in under 30 seconds (\S\ref{sec:evaluation}).
+demonstrating that they integrate smoothly with other lens extensions such
+as quotient lenses~\cite{quotientlenses} and matching
+lenses~\cite{matchinglenses}.  We evaluate our implementation and the
+effects of various optimizations and the tuning parameter on a set of 48
+lens synthesis benchmarks drawn from data cleaning, view maintenance, and
+file synchronization tasks.  With only modest interactive tuning, the system
+can synthesize simple symmetric lenses for all of the benchmarks in under 30
+seconds (\S\ref{sec:evaluation}).
 %
@@
 way, we show that {\em star-semiring equivalences} on regular expressions
-remain valid when extended to stochastic regular expressions, preserving
+remain valid when applied to stochastic regular expressions, preserving
 probability distributions as well as languages (\S\ref{sec:sre}).
@@
     important to try it and report on how it performs relative to what we
-    propose here...)}\afm{Oh sorry, yes we can write such examples.
-I personally found that typically the same information could be provided through a series of 2 creates (aka: if we want a disconnect: creater a = b; creater a' = b:  if we want an identity: creater a = a; creater a' = a').
-However, we also permit putr and putl examples of the proposed form in the tool
-Much like other aspects of the work, any situation with examples ultimately only constrains the subcomponents of the lens corresponding to that specific parse tree (aka: if I give an example about A in A|B, it only constrains "A").
-I don't think it removes the need for the information theory, as all the other non-constrained subcomponents of the lens can still be "anything."
-}  }
+    propose here...)}  }
   \label{fig:minimized-representations}
@@
 Semantically, a simple symmetric lens between sets $X$ and
-$Y$ comprises four functions subject to four
+$Y$ comprises four functions that collectively satisfy four
 ``round-tripping'' laws:
@@
 \vskip 1.5ex
-\begin{center}
 \begin{minipage}[c]{.3\textwidth}
 \begin{gather}
- \mbox{$\CreateR \OfType X \rightarrow Y$} \nonumber \\
- \mbox{$\CreateL \OfType Y \rightarrow X$} \nonumber \\
- \mbox{$\PutR \OfType X \rightarrow Y \rightarrow Y$} \nonumber \\
- \mbox{$\PutL \OfType Y \rightarrow X \rightarrow X$} \nonumber
+ \tag{1}
+ \mbox{$\CreateR \OfType X \rightarrow Y$}\\
+ \tag{2}
+ \mbox{$\CreateL \OfType Y \rightarrow X$}\\
+ \tag{3}
+ \mbox{$\PutR \OfType X \rightarrow Y \rightarrow Y$}\\
+ \tag{4}
+ \mbox{$\PutL \OfType Y \rightarrow X \rightarrow X$}
 \end{gather} 
 \end{minipage} 
-\hspace{.4in}
-\begin{minipage}[c]{.5\textwidth}
+\begin{minipage}[c]{.6\textwidth}
 \begin{align}
@@
 \end{minipage} 
-\end{center}
 \vskip 1ex
+\bcp{I want less space between the put/create laws and their names.}
 \noindent
-The two \KW{create} functions are used to propagate shared information and fill
-in default values when introducing new data (\EG, on \KW{create} the ``unk''
-salary entry is added alongside the name to the management file when HR inserts
-a new employee). The two \KW{put} functions propagate edits from one format to
-the other by combining a new value from one with an old value from the other.
-The notation $\Lens.\PutR$ extracts the \PutR function from simple symmetric
-lens $\Lens$; similarly for \PutL. These four functions can be used to keep the
-common information between two file types in synch. For example, if a new file
-of the left-hand format is created, the \CreateR function will a new file in the
-right-hand format. If the right-hand format is edited, the \PutL function can
-update the left-hand format with the changed information.
-
-The round-tripping laws guarantee that pushing an ``un-edited'' value from one
-side (the result of a put or create) back through a lens in the other direction
-will get back to exactly where we began. For example, consider,
-\lstinline{emp_lens}, a lens between
-\lstinline{emp_salaries} and
-\lstinline{emp_insurance}. Applying
-\lstinline{emp_lens}.\CreateR to a salary file creates a insurance file with the
-same employee names in the file, with \lstinline{"UNK"} for each insurance
-company. The round-tripping laws guarantee that using \lstinline{emp_lens}.\PutL
-to put the generated insurance file into the original salary file will return
-the original salary file, unmodified. If \KW{put}ting a string $s$ into a string
-$t$ results in the original string $s$, then $s$ and $t$ are considered
-\emph{synchronized}.
-
-This simple symmetric lens definition differs from classical symmetric lenses in
-that it does not involve a \emph{complement}. We give a detailed comparison in
+The two \KW{create} functions are used to propagate shared information and
+fill in default values when introducing new data (\EG, the ``unk'' salary
+entry is added alongside the name to the management file when HR inserts a
+new employee). The two \KW{put} functions propagate edits from one format to
+the other by combining a new value from one with an old value from the
+other.  The round-tripping laws guarantee that pushing an ``un-edited''
+value from one side (the result of a put or create) back through a lens in
+the other direction will get back to exactly where we began.
+
+This definition differs from classic symmetric lenses in that it does not
+involve a \emph{complement}.  We give a detailed comparison in
 \S\ref{sec:relationship}.
@@
 
-In principle, programmers can define simple symmetric lenses just by writing
-four arbitrary functions by hand and showing they satisfy the round-tripping
-laws. But a better idea is to streamline the process by providing a set of
-combinators---a domain-specific language---for building law-abiding complex
-lenses from simpler ones. Such bidirectional languages have been given for
-building lenses over many different kinds of data, including database
-relations~\cite{BohannonPierceVaughan} algebraic data
-types~\cite{symmetric-lenses}, graphs~\cite{tgg}, and more.
-
-We focus here on combinators for defining lenses between string data formats
+Generally, we do not expect programmers to define simple symmetric lenses by
+writing four functions by hand and showing they satisfy the round-tripping
+laws. Instead, we provide a set of combinators---a domain-specific
+language---for building complex lenses from simpler ones.  Such languages
+have been given for building lenses over many different kinds of data,
+including database relations~\cite{BohannonPierceVaughan} and algebraic data
+types~\cite{symmetric-lenses}\bcp{and what else? graphs, etc.?}.  We focus
+here on combinators for defining lenses between string data formats
 described by regular expressions.
@@
 $\Lens$ is a simple symmetric lens between $\LanguageOf{\BRegex}$ and
-$\LanguageOf{\BRegexAlt}$. (We will use undecorated variables later for
+$\LanguageOf{\BRegexAlt}$. (Later, we will use undecorated variables for
 stochastic regular expressions; we mark plain REs with overbars.)  When
-clear from context, we shorten ``simple symmetric lens'' to just
+clear from context, we shoren ``simple symmetric lens'' to just
 ``lens.''
@@
 and the put functions merely return the first argument
-($\PutROf{\String}{\StringAlt} = \String$).
+($\PutROf{s_1}{s_2} = s_2$).
 %
@@
 
-In contrast, the \Disconnect lens does not propagate any edits at all
+In contrast, the \Disconnect lens does not propagate any edits
 from one format to the other.  It takes four arguments: two
 regular expressions and two strings. The regular expressions specify the formats
-on the two sides, while the strings provide default values. 
+of the corresponding data on the two sides while the strings provide default
+values. 
 %
 \begin{lstlisting}
-disconnect(salary, "", "unk", "") : salary $\Leftrightarrow$ ""
+disconnect(salary, "", "unk", "") : salary $\Leftrightarrow$ company
 \end{lstlisting}
@@
 On creates, the input values are thrown away, and default values are
-returned ($\CreateLOf{\StringAlt} =
+returned ($\CreateROf{\String} =
 \lstinline{"unk"}$), and on puts, the second argument is used and the first is
@@
 HR file.
-For example, with the above lens, $\PutL \App \lstinline{""} \App 60000$ will return 60000,
-and \PutR will always return \lstinline{""}.
-
-The insert lens \lstinline{ins} and the delete lens \lstinline{del} are
+The defaults are strings that indicate the information is missing.
+
+The insert \lstinline{ins} and delete \lstinline{del} lenses are
 syntactic sugar for uses of the disconnect lens in which a string
-constant is omitted entirely from the source or target format.
+constant is omitted entirely from the source and target formats,
+respectively.
 \begin{lstlisting}
@@
 \end{lstlisting}
-The \lstinline{ins} lens inserts a constant string when going from
-left to right, while \lstinline{del} inserts a string when going from right
-to left.
-
-Finally, there are a number of combinators that construct
+
+Finally, there are a number of combinators that compositionally construct
 more complex lenses from simpler ones. These include concatenation
-($\ConcatLensOf{\Lens_1}{\Lens_2}$, often written $\Lens_1 . \Lens_2$), variation
+($\ConcatLensOf{\Lens_1}{\Lens_2}$ or $\Lens_1 . \Lens_2$), variation
 ($\OrLensOf{\Lens_1}{\Lens_2}$), and iteration ($\IterateLensOf{\Lens}$).
@@
 will transform ``\lstinline{Jane Doe: }'' to ``\lstinline{Jane Doe,}'' in
-the left-to-right direction.
+the forward direction.
+\bcp{We don't actually give a nontrivial example of how disconnect works.}
 
@@
 %let emp_lens = or(id(""),employees_lens) in
-  \caption{A lens that synchronizes management and HR employee files}
+  \caption{A lens that synchronizes management and HR employee files\bcp{I'm
+    pretty sure there was an error in name\_lens: the thing between the two
+    names should be id(" ") (was ins('','')).  Crossing my fingers that I'm
+    making a correct fix...}}
   \label{fig:example_lens}
@@
 
-We provide typing rules for the simple symmetric combinators, most of
-them syntax directed. However, regular expressions have a number of
+We provide typing rules for each of the simple symmetric combinators, most of
+which are syntax directed. However, regular expressions have a number of
 equivalences on them, and sometimes we must convert one regular expression type
-to another, equivalent, regular expression type.
-%
-\[
-  \centering
-  \inferrule*
-  {
-    \Lens \OfType \BRegex \Leftrightarrow \BRegexAlt\\
-    \BRegex \SSREquiv \BRegex'\\
-    \BRegexAlt \SSREquiv \BRegexAlt'
-  }
-  {
-    \Lens \OfType \BRegex' \Leftrightarrow \BRegexAlt'
-  }
-\]
-%
-This rule is needed for lenses like the final one in
-Figure~\ref{fig:example_lens} to be well-typed.
+to another, equivalent, regular expression type. Our typing rules include type
+equivalence, so if a lens has type $\BRegex \Leftrightarrow \BRegexAlt$, then it
+is also has type $\BRegex' \Leftrightarrow \BRegexAlt'$, where $\BRegex$ is
+equivalent to $\BRegex'$, and $\BRegexAlt$ is equivalent to $\BRegexAlt'$. The
+flexibility this type equivalence rule provides is useful when a different
+representation of a regular expression would facilitate synthesis.
 
@@
 
-Lens expressions for small examples like the one above are pretty easy to
-write.  But for larger and more complex formats, it can become quite difficult to
-write lenses by hand.  Instead, we wish to synthesize them from examples.
-
-That is, given a pair of regular expressions and a set of input-output examples, we want
+While our example is relatively simple for demonstration, writing such lenses
+for real-world formats is quite difficult, as these formats can be large and
+represent data in complex and unintuitive ways. To alleviate the difficulty of
+manually writing simple symmetric lenses, we wish to synthesize them instead.
+Given a pair of regular expressions and a set of input-output examples, we want
 to find a simple symmetric lens that is typed by the regular expression pair and
-satisfies all the input/output examples. We call such a lens a \emph{satisfying
+satisfies the input/output examples. We call such a lens a \emph{satisfying
   lens}. In our running example, we wish to synthesize a lens between
 \lstinline{emp_salaries} and \lstinline{emp_insurance}, using as an input-output
-example the data in Figure~\ref{fig:minimized-representations}.
-
-One challenge is that the simple symmetric lens combinators permit many
+example the data in Figure~\ref{fig:minimized-representations}. The algorithm
+uses such examples to find suitable defaults for \Disconnect lenses and to
+figure out how to match data that could be synchronised in multiple ways.
+
+A challenge is that the simple symmetric lens combinators permit many
 well-typed lenses
@@
 \begin{lstlisting}
-disconnect(emp_salaries, emp_insurance, rep(emp_salaries), rep(emp_insurance))
+disconnect(emp_salaries,emp_insurance,rep(emp_salaries),rep(emp_insurance))
 \end{lstlisting}
 %
-is another well-typed lens (where \lstinline{rep($\BRegex$)} denotes some string in the
-language of $\BRegex$).  Adding the input-output example in
+is another valid lens (where \lstinline{rep($\BRegex$)} denotes some string in the
+language of $\BRegex$). Furthermore, adding the input-output example in
 Figure~\ref{fig:minimized-representations} 
-helps, but only a little: In general, {\em many} examples may be required to rule
+helps, but only a little: It may, in general, require {\em many} examples to rule
 out all possible occurrences of nested \Disconnect lenses,
@@
 
-How can we identify such a ``likely'' lens? We propose the following heuristic:
-A satisfying lens is ``more likely'' if it ``connects'' more data between the
-two formats, \EG, by using an identity rather than a disconnect lens. The
-likeliest satisfying lenses would connect as much data as possible. More
-formally, we define the likelihood of a satisfying lens as the expected number
-of bits required to recover data in one format from synchronized data in the
-other. Higher likelihoods correspond to fewer bits. Yet more formally, we say
-that two strings $s$ and $t$ are \emph{synchronized according to lens $l$} if
-$l.\PutR\ s\ t = t$ and $l.\PutL\ t\ s = s$. We say that we can \emph{recover}
-$s$ from $t$ using bits $b$ and lens $l$ if we can reconstruct $s$ from $t$,
-$b$, and $l$. For example, given the \IdentityLens{} lens, we can recover $s$
-from $t$ using no bits because, in this case, $s$ is just $t$. In contrast,
-given the \Disconnect{} lens, we need enough bits to fully encode $s$ in order
-to recover it from $t$ because $t$ contains no useful information. Thus, if both
-\IdentityLens{} and \Disconnect{} are satisfying lenses for a particular pair of
-formats, then \IdentityLens{} will be more likely.
-
-The expected number of bits required to recover a piece of data corresponds to
-the well-known information-theoretic concept of
-\emph{entropy}~\cite{Shannon1948}. Calculating entropy requires a probability
-distribution over the space of possible values for the data. Specifically, given
-a set $S$ and a probability distribution $P \OfType S \rightarrow \Reals$ over
-$S$, the entropy $\EntropyOf{S,P}$ is $-\Sigma_{s\in S}P(s)\Log_2P(s)$. In
-information theory, each element of $S$ has \emph{information content}, the
-number of bits required to specify that element. The information content of an
-element $s$ is $-\Log_2P(s)$. For example, let $S = \SetOf{a,b_1,b_2}$, where
-$P(a)=.5$ and $P(b_1) = .25 = P(b_2)$. The information content of $a$ is $1$,
-and the information content of each $b_i$ is 2. The entropy of $S$ calculates
-the expected number of bits required to specify a random element pulled from
-$S$, in other words, the probability of each element times its information
-content. In our example, the entropy of $S$ is $1.5$. When $P$ is clear, we will
-often use the shorthand of $\EntropyOf{S}$ for $\EntropyOf{S,P}$.
+But what is a ``likely'' satisfying lens? We propose the following
+heuristic: A satisfying
+lens is more likely if it connects more data between the two
+formats, \EG, by using an identity rather than a disconnect lens.  The
+likeliest satisfying lenses would connect as much data as
+possible. More formally, we define the likelihood of a
+satisfying lens as the expected number of bits required to recover
+data in one format from synchronized data in the other.  Higher
+likelihoods correspond to lower numbers of bits. In more
+detail, we say that two strings $s$ and $t$ are \emph{synchronized
+according to lens $l$} if $l.\PutR\ s\ t = t$ and $l.\PutL\ t\ s = s$.\footnote{
+The notation $l.\PutR$ extracts the \PutR function from simple
+symmetric lens $l$; similarly for \PutL.}  We say that we
+can \emph{recover} $s$ from $t$ using bits $b$ and lens $l$ if we can
+reconstruct $s$ from $t$, $b$, and $l$.  For example, given
+the \IdentityLens{} lens, we can recover $s$ from $t$ using no bits
+because, in this case, $s$ is just $t$.  In contrast, given
+the \Disconnect{} lens, we need enough bits to fully encode $s$ in
+order to recover it from $t$ because $t$ contains no useful
+information.  This calculation shows that if both \IdentityLens{}
+and \Disconnect{} are satisfying lenses, then \IdentityLens{} is more
+likely.
+
+The expected number of bits required to recover data is the well-known information-theoretic concept 
+of the \emph{entropy} of the data~\cite{Shannon1948}.
+Calculating entropy requires a probability distribution over the
+space of possible values for the data. Specifically, given a set $S$ and a
+probability distribution $P \OfType S \rightarrow \Reals$ over $S$, the
+entropy is $-\Sigma_{s\in  S}P(s)\Log_2P(s)$.
 
@@
 which are regular expressions in which each operator is annotated with a
-probability (see \S\ref{sec:sre}).  A stochastic regular expression
+probability.  A stochastic regular expression
 specifies both a set of strings and a probability
@@
 
-In the synthesis setting, there are typically an infinite number of possible
-lenses that satisfy the specification. There are many choices on which
-components of the data should be synchronized vs which should be left
-unsynchronized. We use entropy to guage the relative likelihood of such lenses
-in our synthesis algorithm (see \S\ref{subsec:lenscosts}). For any lens $\Lens$, we can calculate the expected
-number of bits required to recover a string $t$ in $\LanguageOf{\RegexAlt}$ from
-a synchronized string $s$ in $\LanguageOf{\Regex}$. This expectation is the
-\emph{conditional entropy of $\Regex$ given $\RegexAlt$ and $\Lens$}, formally
-$\sum_{s \in \Regex}\ProbabilityOf{\Regex}{s}\cdot\EntropyOf{\SetOf{t \SuchThat
+This infrastructure gives us what we need to find likely satisfying lenses from
+a collection of satisfying lenses. All such lenses must synchronize data between
+the same two SREs $\Regex$ and $\RegexAlt$. For any such lens $\Lens$, we can
+calculate the expected number of bits required to recover a string $t$ in
+$\LanguageOf{\RegexAlt}$ from a synchronized string $s$ in
+$\LanguageOf{\Regex}$. This expectation is the \emph{conditional entropy of
+  $\Regex$ given $\RegexAlt$ and $\Lens$}, formally $\sum_{s \in
+  \Regex}\ProbabilityOf{\Regex}{s}\EntropyOf{\SetOf{t \SuchThat
     \Lens.\PutROf{s}{t} = t}}$. The likelihood we assign to $\Lens$ is the sum
@@
 conditional entropy of $\RegexAlt$ given $\Regex$ and $\Lens$. This metric
-assigns higher entropy (or cost) to lenses where knowing the string on one side
-provides little information about the string on the other side. It assigns zero
-entropy to bijections because given a string $s \in \Regex$, the bijection
-exactly determines the corresponding string in $\RegexAlt$.
+assigns higher entropy (or cost) to lenses where knowing the string
+on one side 
+provides little information about the string on the other side. 
+It assigns zero entropy to bijections because given a string $s \in \Regex$,
+the bijection exactly determines
+the corresponding string in 
+$\RegexAlt$.
 
@@
 
-To obtain SREs from the plain regular expressions that users write, we use a
-heuristic that attempts to assign probability annotations that give each
-string in the language equal probablity. Our synthesis algorithm could
-equally well take as input a pair of stochastic regular expression rather than a
-plain one, permitting users to specify a data-specific probability
-distribution (either manually calculated or inferred from some dataset of
-interest). However, we have found that our heuristic works well in practice,
-so we have not explored these options.
+To obtain SREs from the plain regular expressions that we expect users to
+supply, we use a heuristic that attempts to assign probability annotations that
+give each string in the language equal probablity. Our algorithm could also take
+as input a stochastic regular expression rather than a plain one,
+which would allow
+the user to specify a data-specific probability distribution, either manually
+calculated or inferred from a large dataset. We found our heuristic
+works well in practice, so we have not explored these options.
 
 Now that we have a way to calculate likelihood, we need a way to search for
-likely lenses. Previous work~\cite{augustsson-2004, osera+:pldi15,
-  feser-pldi-2015, frankle+:popl16} has shown that we can use types to
-dramatically restrict the search space and improve the effectiveness of
-synthesis. In our setting, types are pairs of stochastic regular
-expressions. The fact that each regular expression is semantically equivalent to
-infinitely many others complicates the problem.
-Following existing work on bidirectional program
-synthesis~\cite{maina+:quotient-synthesis}, we 
-split our algorithm into two communicating
-searches. The first search,
-\Expand, navigates the space of semantically equivalent regular expressions
-by applying rewrite rules that preserve both the semantics and the
-probability distributions of those stochastic regular expressions. This
-algorithm ranks pairs of stochastic regular expressions by the number of
-rewrite rule applications required to obtain each pair from the one given as
-input. It passes the pairs off to the second 
-search procedure, \GreedySynth, in rank order with the smallest first.
-
-Abstractly, \GreedySynth looks for highest-likelihood lenses between a given
-pair of stochastic regular expressions $\Regex$ and $\RegexAlt$ by performing a
-type-directed search. In more detail, \GreedySynth first normalizes the
-stochastic regular expressions provided by \Expand into \emph{stochastic DNF
-  regular expressions}---a constrained representation of stochastic regular
-expressions with disjunctions distributed over concatenations and with
-concatenations and disjunctions normalized to operating over lists (see \S\ref{subsec:sdnfre}).
-\GreedySynth then uses the syntax of these n-ary DNF regular expressions to find
-normalized lenses, in a form we call \emph{simple symmetric n-ary DNF lenses}.
-These lenses include neither a composition operator nor a type equivalence rule.
-These restrictions mean that each simple symmetric n-ary DNF lens is typed by a
-single pair of stochastic n-ary DNF regular expressions, hence \GreedySynth's
-search space for a given pair of regular expressions is finite. The variations
-that \GreedySynth considers include choosing between \IdentityLens{} and
-\Disconnect lenses and deciding how to match multiple occurrences of the same
-kind of data, \EG, the two occurrences of \lstinline{name} in the management and
-HR data formats (should the algorithm match the first instance of
-\lstinline{name} on the left with the first instance of \lstinline{name} on the
-right, or to the second). Finally, \GreedySynth yields a simple symmetric lens
-by converting the n-ary forms into the binary forms provided in the surface
-language.
-
-To illustrate the partnership between \Expand and \GreedySynth, consider
-synthesizing the lens shown in Figure~\ref{fig:example_lens}. When \GreedySynth
-initially tries to find a lens between \lstinline{emp_salaries} and
-\lstinline{emp_insurance}, it hits some snags: there are complex disjunctions
-and concatenations in \lstinline{emp_salaries} outside the iteration, but no
-such complexity in \lstinline{emp_insurance}. Because of this disparity between
-the formats, \GreedySynth finds an unlikely lens (involving undesired
-\KW{merge}s and \KW{disconnect}s), and the search continues. However, after
-\Expand applies a star-semiring axiom (presented in Figure~\ref{fig:ssr-sre}) to
+likely lenses. Previous work on
+synthesis~\cite{augustsson-2004,osera+:pldi15,feser-pldi-2015,frankle+:popl16}
+has shown that we can use types to dramatically restrict the search space and
+improve the effectiveness of synthesis. In our setting, types are pairs of
+stochastic regular expressions where each regular expression is semantically
+equal to infinitely many others. We follow existing work on bidirectional
+program synthesis, and split our algorithm into two communicating
+synthesizers~\cite{maina+:quotient-synthesis}. The first synthesizer, \Expand,
+navigates the space of semantically equivalent regular expressions by applying
+rewrite rules that preserve both the semantics and the probability distributions
+of those stochastic regular expressions. Our algorithm ranks pairs of stochastic
+regular expressions by the number of rewrite rule applications required to
+obtain the pair from the stochastic regular expressions given as input. It
+passes the pairs off to the second synthesizer, \GreedySynth, in rank order with
+the smallest first.
+
+At a high level, \GreedySynth looks for lenses between a given pair of
+stochastic regular expressions $\Regex$ and $\RegexAlt$ that have the highest
+likelihood by performing a type-directed search. In more detail, \GreedySynth
+first normalizes the stochastic regular expressions provided by \Expand into
+\emph{stochastic DNF regular expressions}. These stochastic DNF regular
+expressions are, intuitively, stochastic regular expressions with disjunctions
+distributed over concatenations, where associativity information has been
+removed. Then, \GreedySynth uses the syntax of these n-ary DNF regular
+expressions to find highly normalized lenses, in a form we call \emph{simple
+  symmetric n-ary DNF lenses}. These lenses include neither a composition
+operator nor a type equivalence rule. These restrictions mean that each simple
+symmetric n-ary DNF lens is typed by a single pair of stochastic n-ary DNF
+regular expressions and \GreedySynth's search space for a given pair of regular
+expressions is finite. The variations that \GreedySynth considers include
+choosing between \IdentityLens{} and \Disconnect lenses and deciding how to
+match multiple occurrences of the same kind of data, \EG, the two occurrences of
+\lstinline{name} in the management and HR data formats (we want to map first
+names to first names and last names to last names). \GreedySynth
+converts yields a simple symmetric lens by converting the n-ary
+forms into the binary forms provided in the surface language.
+
+As an example of the partnership between \Expand and \GreedySynth,
+consider the lens shown in Figure~\ref{fig:example_lens}. 
+Inferring this lens requires \Expand to use a star-semiring axiom to
 rewrite the regular expression \lstinline{("\n" . emp_salary)$^*$} within
 \lstinline{emp_salaries} into 
-\lstinline{"" | (("\n" . emp_salary) . ("\n" . emp_salary)$^*$)}, the
-subcomponents line up better, and the more likely correct lens is instead found.
-
-This architecture of communicating synthesizers gives us a way to enumerate
-pairs of stochastic regular expressions of increasing rank and to efficiently
-search through them, but it poses a problem: when should \Expand stop proposing
-new SRE pairs? We may have found a promising lens between a pair of stochastic
-regular expressions, but a different pair we haven't yet discovered may give
-rise to an even better lens.
-
-The search algorithm must resolve a tension among the quality of
-the inferred lens and the amount of time it takes to return a
+\lstinline{"" | (("\n" . emp_salary). ("\n" . emp_salary)$^*$)} 
+so that \GreedySynth can find the appropriate lens.
+
+Our strategy of communicating synthesizers gives us a way to
+enumerate pairs of stochastic regular expressions of increasing
+rank and to efficiently search through them, but it poses a
+problem: when do we stop? We may have found a promising lens
+between a pair of stochastic regular expressions, but a different 
+pair we haven't yet discovered may give rise to an even better lens. 
+Or, we may have found a promising lens between a given pair of
+stochastic regular expressions, but the discovered lens is not 
+acceptable to the users.
+
+Our algorithm must resolve a three-way tension between the quality of
+the inferred lens, the number of examples the user must provide to
+eliminate unwanted lenses, and the amount of time it takes to return a
 result.  For example, if the algorithm has already found the lens in
@@
 
-To resolve this tension, the algorithm uses heuristics to judge whether to
+To resolve this tension, our algorithm uses heuristics to judge whether to
 return the current best satisfying lens to the user or to pass the next set of
 unrollings to \GreedySynth.  The heuristics favor stopping if the current best
-satisfying lens is very likely, indicating the lens is very promising (for
-example, if the satisfying lens loses no information, the algorithm should
-terminate for no other lens will be more likely). The heuristics also favor
-stopping if \Expand has delivered to \GreedySynth all pairs of stochastic
-regular expressions at a given rank and there is a large number of pairs at the
-next rank, because searching through all such pairs will take a long time. With
-this approach, the algorithm can quickly respond with a satisfying lens with
-relatively high likelihood. If the user is unhappy with the result, they can
-either refine the search by supplying additional examples, which serve both to
-rule out the previously proposed lens(es) and to reduce the size of the search
-space by cutting down on the number of satisfying lenses, or they can supply
-annotations on the source and target SREs indicating that certain information
-either must be retained or must be discarded by the lens (see
-\S\ref{subsec:optimizations}).
+satisfying lens has relatively high likelihood, indicating the lens is very
+promising.  The heuristics also favor stopping if \Expand has delivered to
+\GreedySynth all pairs of stochastic regular expressions at a given rank and
+there is a large number of pairs at the next rank because searching through all
+such pairs will take a long time.  With this approach, the algorithm can quickly
+respond with a satisfying lens with relatively high likelihood.  If users are
+unhappy with the result, they can either refine the search by supplying
+additional examples, which serve both to rule out the previously proposed
+lens(es) and to reduce the size of the search space by cutting down on the
+number of satisfying lenses, or supply a tuning parameter (discussed in 
+\S\ref{subsec:optimizations}) that adjusts how much effort is put into the search.
 
@@
 \label{sec:sre}
-To characterize likely lenses, we must compute the expected number of bits
+To characterize the likely lenses, we must compute the expected number of bits
 needed to recover a string in one data source from a synchronized string in the
-other data source. To do this, we must first develop a probabilistic 
-model for our language. We do this with \emph{stochastic regular
-  expressions} (SREs)---regular expressions annotated with probability
-information~\cite{stoch-rnn,stoch-def}---which 
-jointly express a language and a probability distribution over that language. 
-$$\Regex{},\RegexAlt{} \quad\GEq{}\quad  \String  \quad|\quad \emptyset \quad|\quad \PRegexStar{\Regex}{p} \quad|\quad \RegexConcat{\Regex_1}{\Regex_2} \quad|\quad \PRegexOr{\Regex_1}{\Regex_2}{p}$$
-(Lowercase $\String$ ranges over constant strings and $\Probability$ ranges
-over real numbers between 0 and 1, exclusive.)
-
-The semantics of an SRE $S$ is a probability distribution $P_S$ defined as
-follows.
+other data source. To do this, we must first develop a probabilistic language
+model for our language. We do this with \emph{stochastic regular expressions},
+regular expressions annotated with probability information. With this, we can
+jointly express a language and a probability distribution over that language. In
+the following grammar, $\String$ ranges over strings, and $\Probability$ ranges
+over real numbers between 0 and 1, exclusive.
+$$\Regex{},\RegexAlt{} \GEq{}  \String  \; | \; \emptyset \; | \; \PRegexStar{\Regex}{p} \; | \; \RegexConcat{\Regex_1}{\Regex_2} \; | \; \PRegexOr{\Regex_1}{\Regex_2}{p}$$
+
+In the stochastic regular expression
+$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation $\Probability$
+represents the likelihood the string is in $\Regex_1$. In the stochastic regular
+expression $\PRegexStar{\Regex}{\Probability}$, the annotation \Probability how
+much more likely having a string involving more than $n$ iterations is than
+having a string involving exactly $n$ iterations. The semantics of SREs is
+defined by the function $P$ below.
+
 \begin{center}
@@
 \end{center}
-In the SRE $\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation
-$\Probability$ represent that, if generating arbitrary strings from
-$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, strings should be pulled from
-$\Regex_1$ with probability $\Probability$ (\EG pulling a string from
-$\PRegexOr{\lstinline{"a"}}{\lstinline{"b"}}{.2}$ would retrieve \lstinline{"a"}
-20\% of the time). In the SRE
-$\PRegexStar{\Regex}{\Probability}$, the annotation \Probability represents how
-much more likely having a string involving more than $n$ iterations is than
-having a string involving exactly $n$ iterations (\EG pulling a random string
-from $\PRegexStar{\lstinline{"c"}}{.8}$ would give \lstinline{""} 20\% of the
-time, \lstinline{"c"} 16\% of the time, and \lstinline{"cc"} 12.8\% of the time).
 
@@
 \subsection{Stochastic Regular Expression Equivalences}
-The $\Expand$ algorithm needs to be able to enumerate SREs that are
-``equivalent'' to a given one. However, existing work does not define any notion
-of stochastic regular expression equivalence. Figure~\ref{fig:ssr-sre} shows how
-we to extend the \emph{star-semiring}~\cite{Droste2009} equivalences (a
-\emph{finer} notion of equivalence than semantic equivalence) to
-SREs.
+$\Expand$ needs to be able to find equivalent stochastic regular expressions.
+However, existing work does not provide a means to reason about stochastic
+regular expression equivalence. We extend the
+\emph{star-semiring}~\cite{Droste2009} equivalences to stochastic regular
+expressions (Figure~\ref{fig:ssr-sre}).
 
@@
 \begin{proof}
-  See the appendix%
-\ifappendices\else{} included with the supplemental
-material\fi.  \bcp{It took me quite a while to find it.  We
-  should give an actual pointer to a specific theorem in the appendix.}
+  The proofs of this theorem, and subsequent theorems in this paper, appears in
+  the appendix present in the full version of the paper accompanying this
+  submission.
 \end{proof}
 
-This theorem will come in handy as we traverse (with \Expand) or normalize
-across (as part of \GreedySynth) star-semiring equivalences.
+We prove these star-semiring equivalences sound, as those are the equivalences
+we either traverse (with \Expand), or normalize across (as part of
+\GreedySynth).
 
@@
 The entropy of a data source $S$ is the expected number of bits required to
-describe a random element in $S$ (formally $-\Sigma_{s\in S}\,P(s)\cdot
-\Log_2P(s)$). The entropy of a SRE can be computed directly from its syntax,
-when each string is uniquely parsable (\IE{} the SRE is unambiguous) and contains
-no empty subcomponents.
+describe a random element in $S$ ($-\Sigma_{s\in S}P(s)\Log_2P(s)$). The entropy
+of a stochastic regular expression can be computed directly from its syntax,
+when the stochastic regular expression is unambiguous and contains no empty
+subcomponents.
 \begin{center}
@@
     & =
-    & 0\\
+    & 1\\
 
@@
 \end{center}
-
-For example, the entropy of $\PRegexOr{\lstinline{"a"}}{(\lstinline{"b"})}{.5}$
-is 1. This is expected because on average 1 bit is required to encode whether a
-randomly pulled string is \lstinline{"a"} or \lstinline{"b"}. As an additional
-example, a fixed string has no information content, and so has no entropy.
+We have proven that this syntactic computation corresponds to the entropy of the
+stochastic regular expression.
 
@@
 
-Without unambiguity constraints, the method above for calculating entropy does
-not work, and it is not easy to see how to generalize it.  For example,
-consider the SRE $\PRegexOr{\lstinline{"a"}}{\lstinline{"a"}}{.5}$. Using
-the above computation, this SRE has an entropy of $1$, but the entropy
-should be $0$. These difficulties become particularly complex when the
-languages are ambiguous in subtle ways and the algorithm must be correct
-when only some of the two languages overlap. Similar issues occur when
-trying to find the entropy when $\emptyset$ is a subterm (what should the
-entropy of $\PRegexStar{\emptyset}{.5}$ be?), so we do not define entropy
-on~$\emptyset$.
+Without unambiguity constraints it is not easy to calculate entropy (and the
+method we have given is certainly incorrect).  For example, consider the SRE
+$\PRegexOr{\lstinline{"a"}}{\lstinline{"a"}}{.5}$. Using the above computation,s
+this SRE has an entropy of $1$, but the entropy should be $0$. These
+difficulties become particularlly complex when the languages are ambiguous in
+subtle ways, the algorithm must be correct when only some of the two languages
+overlap. Similar issues occur when trying to find the entropy when $\emptyset$
+is a subterm (what should the entropy of $\PRegexStar{\emptyset}{.5}$ be?),
+so we do not define entropy on $\emptyset$.
 
@@
 our synthesis procedure for other reasons, and we can easily preprocess
-empty subexpressions out of SREs that are themselves nonempty using the
-star-semiring equivalences (\EG, $\PRegexOr{\emptyset}{\lstinline{"s"}}{.5}
-\SSREquiv \lstinline{"s"}$).
+emptysets out of the stochastic regular expressions that are themselves
+nonempty by traversing the appropriate star-semiring equivalences (\EG,
+\PRegexOr{\emptyset}{\lstinline{"s"}}{.5} \SSREquiv \lstinline{"s"}).
 
@@
 % \label{fig:regex-equivalence-rules}
-
-%%%%%%%%%%%%%%  BCP: stopped reading carefully here...
 
@@
 recover a regular expression from a stochastic regular expression by deleting
-probability annotations. In a slight abuse of notation (as we are also using
-these symbols as nonterminals), if $\Regex$ is a
+probability annotations. In a slight abuse of notation, if $\Regex$ is a
 stochastic regular expression, $\BRegex$ is the regular expression obtained by
@@
 \Leftrightarrow \BRegexAlt$ means that $\Lens$ is a well typed lens between
-the language of \BRegex and the language of \BRegexAlt.
+$\LanguageOf{\BRegex}$ and $\LanguageOf{\BRegexAlt}$.
 
@@
 
-We have already described a number of these combinators in \S\ref{sec:overview}.
-Briefly (full definitions are below), the compose lens $\Lens_1 ; \Lens_2$
-composes two lenses sequentially, it 
+We have already described a number of these lenses in \S\ref{sec:overview}.
+The compose lens $\Lens_1 ; \Lens_2$ composes two lenses sequentially, it
 transforms from one data format to the other by first transforming to an
@@
 
-We now give typing rules and semantics for (some of) these combinators. We do not
-include the semantics for combinators equivalent to those presented in prior
-work~\cite{Boomerang}. Furthermore, we present
-just \CreateR and \PutR in cases where \CreateL and \PutL are symmetric;
-full details can be found in an appendix\bcp{where,
-  exactly??}\ifappendices\else{} (in the supplemental materials)\fi. For
-simplicity of presentation, we use unambiguous regular expressions
-everywhere, even though the identity and disconnect lenses can actually be defined over ambiguous
-regular expressions; we believe relaxing
-this restriction is unproblematic.
+We now provide the typing derivations and semantics for these lenses. We do
+not include the semantics for combinators equivalent to those presented in
+prior work~\cite{symmetric-lenses,Boomerang}. Furthermore, we only present
+\CreateR and \PutR when \CreateL and \PutL are symmetric; full details can
+be found in an appendix\ifappendices\else{} (in the supplemental
+materials)\fi. For simplicity of presentation, we use unambiguous regular
+expressions everywhere, even though some lenses can actually be defined over
+ambiguous regular expressions (for example, the identity lens doesn't need
+any ambiguity constraints\bcp{Is this the only example?}); we believe
+relaxing this restriction is unproblematic.
 
@@
     \CreateR{} \App s & = & s\\
-    \PutR{} \App s \App t & = & s
+    \PutR{} \App s_1 \App s_2 & = & s_1
   \end{array}$
@@
 data gets changed from one format to the other, information in the old format is
-lost. This differs from the \OrLens lens of classical symmetric
-lenses, as those can retain such information in the complement (see
-\S\ref{sec:related}).
+lost.
 %\begin{center}
@@
 The \IterateLens lens deals with iterated data, while inverting reverses the direction of a lens: creating on the right becomes creating on the left and vice versa, and putting on the right becomes putting on the left and vice versa.
-The invert combinator is particularly useful when chaining many compositions together, as it can be used to align the central types.
+The invert combonator is particularly useful when chaining many compositions together, as it can be used to align the central types.
 \[
@@
 Type equivalence enables a lens of type $S \Leftrightarrow T$ to be used as a
-lens of type $S' \Leftrightarrow T'$ if $S$ is equivalent to $S'$ and $T$ is
+lens of type $S' \Leftrightarrow T'$ if $S$ equivalent to $S'$ and $T$ is
 equivalent to $T'$. Type equivalence is useful both for addressing type
@@
 \subsection{Lens Costs}
-\label{subsec:lenscosts}
 Our cost metric is based on the expected amount of information required to
@@
 expected amount of information required to recover a string in $\RegexAlt$ from
-a string in $\Regex$, synchronized by $\Lens$. Similarly, we use the function
+a string in $\Regex$, synchronized by $\Lens$. We use the function
 $\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate bounds of the
 expected amount of information required to recover a string in $\Regex$ from a
-string in $\RegexAlt$, synchronized by $\Lens$. We write $a(b,c)$ to
-mean $(a*b,a*c)$, and $(a,b)+(c,d)$ to mean $(a+c,b+d)$.
+string in $\RegexAlt$, synchronized by $\Lens$.
 \begin{center}
@@
     & =
-    & $(0,\REntropyOf{\RegexAlt_1 \Given \Lens_1, \Regex}+\REntropyOf{\RegexAlt_2 \Given \Lens_2, \Regex}+1)$\\
+    & $(0,\REntropyOf{\RegexAlt \Given \Lens_1, \Regex_1}+\REntropyOf{\RegexAlt \Given \Lens_2, \Regex_2}+1)$\\
 
@@
 synchronized string in the other format. Note that we would be able to exactly
-calculate the conditional entropy, were it not for \MergeL and \MergeR. If
+calculate the conditional entropy, were it not for \MergeL. If
 $\MergeLOf{\Lens_1}{\Lens_2} \OfType \Regex \Leftrightarrow
@@
 typically calculate the conditional entropy exactly, as merges are relatively
-uncommon in practice; only 2 of the lenses sythesized in our benchmarks
+uncommon on practice; only 2 of the lenses sythesized in our benchmarks
 include merges.
@@
 \MaxOf{\LEntropyOf{\Regex \Given \Lens, \RegexAlt}} +
-\MaxOf{\REntropyOf{\RegexAlt \Given \Lens, \Regex}}$ is the sum of the maximum
-expected number of bits to
-recover the left format from the right, and the right from the left. We have
+\MaxOf{\REntropyOf{\RegexAlt \Given \Lens, \Regex}}$ is the maximum sum of
+recovering the left format from the right, and the right from the left. We have
 proven theorems demonstrating the calculated entropy corresponds to the actual
@@
 
-Note that our definition of $\Entropy^\rightarrow$ contains no case for sequential
+Note that our definition of $\Entropy$ contains no case for sequential
 composition $\Lens_1 ; \Lens_2$ and our theorem excludes lenses that contain
-such compositions.  Defining the entropy of lenses involving
+such compositions. Unfortunately, defining the entropy of lenses involving
 composition is challenging because $\Lens_1$ might, for instance, add some
@@
 this property is difficult to determine merely from syntax. Fortunately, we are able to
-sidestep such considerations by synthesizing \emph{DNF lenses}---simple symmetric
-lenses that inhabit a disjunctive normal form.
+avoid such complication altogether by synthesizing \emph{DNF lenses}---simple symmetric
+lenses that inhabit a disjunctive normal form. DNF lenses have a number of
+properties that facilitate synthesis: they cut down the program search space
+dramatically and they do not contain instances of composition.
 
@@
     & \Rewrite
-    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{1-\Probability}\\
+    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{\Probability}\\
 
@@
     & \Rewrite
-    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{\Regex{}}{\Probability}}{\Regex{}})}{1-\Probability}
+    & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{\Regex{}}{\Probability}}{\Regex{}})}{\Probability}
   \end{tabular}
@@
 \subsection{Stochastic DNF Regular Expressions}
-\label{subsec:sdnfre}
 Just as SREs are REs with probability annotations,
@@
 expressions whose DNF form gives equal probability to all sequence subcomponents
-of the DNF regular expression, and gives iterations a .8 chance to continue
+of the DNF regular expression., and gives iterations a .8 chance to continue
 iterating. In our experience generating random strings from regular expressions,
@@
 \paragraph*{Entropy}
-We have developed a syntactic means for finding the entropy of a stochastic DNF
+We have developed a syntactic means of finding the entropy of a stochastic DNF
 regular expression, like we have for stochastic regular expressions. This
-enables us to efficiently find the entropy without first converting a \SDNFREabbrev to
+enables us to efficiently find the entropy without first converting a DNF SRE to
 a stochastic regular expression.
@@
 many DNF lenses of a given type (though there are still many more symmetric DNF
-lenses than DNF bijective lenses).
+lenses than there are DNF bijective lenses).
 
@@
 symmetric atom lenses, $sal_1 \ldots sal_n$, and a symmetric atom lens $sal$ is
-an iteration of a symmetric DNF lens. Note that the structure of symmetric DNF
-lenses exactly mirrors that of SDNF REs.
+an iteration of a symmetric DNF lens.
 
@@
 The details of these definitions are syntactically tedious, but not
-intellectually difficult. We elide them here but include them in the appendix.\bcp{Where?}
+intellectually difficult. We elide them here but include them in the appendix.
 
@@
 The inputs to \GreedySeqSynth are a suite of input-output examples and a pair of
-lists of stochastic atoms. As in \GreedySynth, \GreedySeqSynth returns \None early if
+stochastic atoms. As in \GreedySynth, \GreedySeqSynth returns \None early if
 there is no possible lens. Afterward, \GreedySeqSynth finds the best lenses
@@
 
-\subsection{Optimizations and Relevance Annotations}
+\subsection{Optimizations and Termination Tuning}
 \label{subsec:optimizations}
@@
 Our implementation includes a number of optimizations not described above:
-annotations that guide DNF conversion; an expansion inference algorithm;
-hash-consing and memoization; and compositional synthesis. The implementation
-also provides for additional specifications in the form of annotations that help
-describe which information is irrelevant to the lens. The optimizations make the
-system performant enough for interactive use, and the annotations help ensure
-the desired information is projected.
+annotations that guide DNF conversion; an expansion inference
+algorithm; hash-consing and memoization; and compositional synthesis. The
+implementation also provides for an additional specification parameter that
+customizes the termination condition for the synthesis procedure.  The
+optimizations make the system performant enough for interactive use, and the
+tuning parameter gives the user a ``knob'' by which the algorithm can be
+directed to try harder to find a good lens.
 
@@
 \paragraph*{Expansion Inference} These additional rewrites make the search
-through possible regular expressions harder. Our algorithm identifies when
-certain closed regular expressions can \emph{only} be involved in a disconnect
-lens (unless opened). Such regular expressions will automatically be opened.
+through possible regular expressions harder. Our algorithm 
+identifies when certain closed regular expressions can \emph{only} be
+involved in a disconnect lens.  Such regular expressions will automatically be
+opened.
 
@@
 
-\paragraph*{Relevance Annotations}\bcp{Wonder whether this should be its own
-top-level section...?}\afm{I'm not sure, any other thoughts? To me it is
-inherently an optimization, though it's a more important and interesting one
-than others}
+\paragraph*{Termination Customization}
 In this work, there are three competing constraints. The tool should (1) return
@@
 Our termination heuristic works well for most situations, but
-there are cases where it fails.
-
-In particular, our system works hard to find lenses that project \emph{as little
-  information as possible}. If the desired lens requires projecting a large
-amount of information, the termination condition will be satisfied ``too late.''
-While \GreedySynth may have found what the user knows to be the desired lens,
-the full algorithm will continue trying to find lenses that project less
-information.
-
-Conversely, sometimes the termination heuristic is overly permissive and algorithm cuts
-short: if it 
-had continued searching for lenses it would have found the desired one, but the
-termination heuristic was satisfied ``too early.'' We wish for our algorithm to
-provide feedback in a timely manner, but sometimes this prohibits finding the
-desired lens.
-
-To address these issues, we allow annotations on the regular expressions
-dictating relevance. The first, $\SkipOf{\Regex}$, says the information
-content of $\Regex$ is irrelevant to the desired lens, and can the data can be
-safely projected. The second, $\SRequireOf{\Regex}$, says the information of
-$\Regex$ is critical to the desired lens, and none of the information can be
-lost when converting between the formats.
-
-If synthesizing a lens seems to be taking a long time, it is typically spending
-time checking it cannot find a lens that projects less information. By
-``skipping'' portions of the SRE that \emph{should} be projected, the algorithm
-will not continue searching for a better lens when it has already found the
-correct one. In practice, we implement \Skip by setting the information content
-of $\Skip(\Regex)$ to zero.
-%
+there are cases where a large number of expansions must be performed to find the
+correct lens.  Users can work around such a problem by providing a large number
+of examples, but constructing examples is a chore, especially when formats are
+complicated.  We have often found it simpler to allow users to change the
+synthesis termination condition to demand \SynthSymLens be more judgemental
+when it comes to deciding it has found a ``good enough'' lens.  
+Conversely, if the user's desired transformation has a particularly
+high cost, the algorithm may spend a long time searching in vain for lenses with
+lower cost. By tuning \SynthSymLens to be more permissive, such a search can be
+short-circuited.
+
+For these situations, users can provide a number $t$ that affects
+the strictness of the termination condition by adjusting the \Continue formula
+as follows. When $\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, $\Regex$ and
+$\RegexAlt$ are at
+distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
+current best lens has cost $c$,
 \begin{gather*}
-  \EntropyOf{\Skip(\Regex)} = 0
+  \Call{Continue}{pq} := c > \Min(d + \Log_2(n) - t, 0).
 \end{gather*}
-%
-If synthesizing a lens between \lstinline{emp_salaries} and
-\lstinline{emp_insurance} from Figure~\ref{fig:minimized-representations} is
-taking overly long, the synthesis procedure is uncomfortable disconnecting
-\lstinline{company} and \lstinline{salary} information. Redefining
-\lstinline{company} and \lstinline{salary} as $\SkipOf{\lstinline{company}}$ and
-$\SkipOf{\lstinline{salary}}$ would set their information content to zero,
-alleviating such issues.
-
-If the algorithm returns a lens with too many disconnects and merges, one could
-provide additional examples to constrain the lens more. However, for large
-formats, writing such examples is time consuming and error-prone. Instead, if
-one recognizes that information from $\Regex$ is incorrectly lost when converted
-between formats, they can prevent that information from being lost by annotating
-it $\SRequireOf{\Regex}$. One can guarantee a lens between \Regex and \RegexAlt
-is bijective by annotating them as $\SRequireOf{\Regex}$ and
-$\SRequireOf{\RegexAlt}$. In practice, we implement \SRequire by setting the
-cost of any lens that loses ``required'' information to infinity.
-%
-\begin{center}
-  \begin{tabular}{rcl}
-    $\REntropyOf{\SRequireOf{\RegexAlt} \Given \Lens, \Regex}$
-    & =
-    & $\begin{cases}
-      \infty & \text{if }\REntropyOf{\SRequireOf{\RegexAlt} \Given \Lens, \Regex}
-      \neq 0\\
-    0 & \text{otherwise}
-  \end{cases}$\\
-
-    $\REntropyOf{\RegexAlt \Given \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, \Regex}$
-    & =
-    & $\begin{cases}
-      \infty & \vspace*{-.3em}\text{if $\RegexAlt$ contains $\SRequireOf{\RegexAlt'}$ as a subexpression,}\\
-      & \text{where $\EntropyOf{\RegexAlt'} \neq 0$}\\
-    [\EntropyOf{\RegexAlt},\EntropyOf{\RegexAlt}] & \text{otherwise}
-  \end{cases}$\\
-  \end{tabular}
-\end{center}
-
-%
-If synthesizing a lens between \lstinline{emp_salaries} and
-\lstinline{emp_insurance} from Figure~\ref{fig:minimized-representations}
-quickly returns a lens with too many disconnects and merges (perhaps
-disconnecting instances of \lstinline{name} in \lstinline{emp_salary} and
-\lstinline{emp_ins}), the termination heuristic is being overly permissive. By
-replacing \lstinline{name} with $\SRequireOf{\lstinline{name}}$ in
-\lstinline{emp_salary} and \lstinline{emp_ins}, name information can no longer
-be lost, and future synthesis calls would continue until a more likely lens is
-found.
+With a positive $t$, the termination condition is stricter, and the algorithm
+will spend longer searching for more likely lenses. With a negative $t$, the
+termination condition is less strict, and the algorithm will spend less time
+searching.  The default $t$ is $0$.
 % end synthesis
@@
 \label{sec:evaluation}
-We have implemented symmetric DNF lenses as an extension to Boomerang.
-We have
-implemented all of Boomerang's asymmetric lens combinators using a
-combination of the symmetric lens combinators presented in this paper and
-symmetric versions of the asymmetric extensions (like matching
-lenses~\cite{matchinglenses} and quotient lenses\cite{quotientlenses}) already
-present in Boomerang. We also integrated our synthesis engine into Boomerang.
-This integration allows users to write synthesis tasks alongside lens
-combinators, incorporate synthesis results into manually-written lenses, and
-reference previously defined lenses in the synthesis of new lenses.
-
-All benchmarks were performed on a 2.5 GHz Intel Core i7 processor with 16
-GB of 1600 MHz DDR3 running macOS High Sierra.
+We have implemented symmetric DNF lenses as an extension to
+Boomerang. Specifically, we added symmetric lens combinators to
+Boomerang, and implemented all of Boomerang's asymmetric lens
+combinators  using a combination of the symmetric lens combinators presented in this
+paper and symmetric versions of the asymmetric extensions (like matching lenses and quotient lenses)
+already present in Boomerang.
+We also integrated our synthesis engine into Boomerang. This
+integration allows users to write synthesis tasks 
+alongside lens combinators, incorporate synthesis results into 
+manually-written lenses, and reference previously defined lenses in
+the synthesis of new lenses.
 
@@
 \begin{enumerate}
-\item Can the algorithm (with suitable examples and annotations) find the correct lens?
+\item Can the algorithm (with suitable examples and termination
+  tuning) find the correct lens?
 \item Is the synthesis procedure efficient enough to be used in everyday
@@
   the existing Optician system customized for bijective lenses?
-\item How effective is the information-theoretic search heuristic and how do our
-  annotations affect the results?
+\item How effective is the information-theoretic search heuristic and how does termination tuning affect the results?
 \end{enumerate}
@@
 Because our tool is intended to be interactive---the user typically starts with
-zero or one examples, runs the tool, determines whether the output is the
-desired one and, if not, either adds an example or relevance annotations and
-then tries again---it is hard to cleanly evaluate the net user experience. The
-primary concern is whether it is usable to synthesize correct lenses.
-
-To determine whether the system can synthesize desired lenses, we ran it
+zero or one examples, runs the tools, determines whether the output is the
+desired one and, if not, either adds an example or adjusts the termination
+metric and then tries again---it is hard to cleanly evaluate the net user
+experience.  The primary concern is whether it is usable to generate
+synthesize correct lenses.
+
+To determine whether the system can synthesize the desired lenses, we ran it
 interactively on all 48 tasks, working with the system to create sufficient
-examples and provide useful relevance annotations.  In all cases, the desired lens
-was obtained.
+examples and adjust the termination parameter.  In all cases, the desired lens
+was obtained, which we verified by manually inspecting the generated lens
+programs.
 %
@@
 
-Providing relevance annotations was needed in only 8 of the 48 tasks. In
-practice, we found that adding such annotations quite easy: if manual inspection
-of the lens showed there were too few \IdentityLens{}s, and too many
-\Disconnect{}s or merges, we would add \SRequire annotations. If synthesis took
-too long, we would add \Skip annotations.
-Section~\ref{subsec:effect-of-heuristics} studies the effects of removing such
-annotations.
-
-We verified that \SSOpt{} generated the correct lenses the way programmers often
-validate their programs: we manually inspected the code and ran unit tests on
-the synthesized code. To determine whether the synthesis procedure generated the
-correct lens when running in modes other than \SSOpt{}, we compared generated lens
-to the lens synthesized by \SSOpt{}.
+Adjusting the termination parameter was needed in only 9 of the 48 tasks. In practice, we
+found that customizing the termination metric to be quite easy: if manual
+inspection of the lens showed there were too few \IdentityLens{}s, and too many
+\Disconnect{}s or merges, we would increase
+the termination parameter. If synthesis took too long, we would decrease it.
+Section~\ref{subsec:effect-of-heuristics} studies the effect of fixing the
+termination metric.
 
@@
 
-Having determined appropriate examples and annotations for the 48
+Having determined appropriate examples and termination parameters for the 48
 benchmarks, we evaluate the performance of the system by measuring the running
@@
 allows our system to scale to arbitrarily large and complex formats; measuring
-it shows the responsiveness of the system when used as
+it shows of the responsiveness of the system when used as
 intended. \SSNCOpt{} mode, which synthesizes a complete lens all at once,
@@
     synthesis it is able to complete 26. Each benchmark specification includes
-    source and target (potentially annotated) regular expressions, and between
-    one and three sufficient examples.}
+    source and target regular expressions, between one and three sufficient
+    examples, and, if necessary, the timing parameter determined to be
+    sufficient for correctness during a prior interactive session.}
   \label{fig:times}
@@
 39 bijective synthesis tasks.\footnote{We had to slightly alter four of these
-benchmarks, either by providing additional examples or by adding in \SRequire annotations.
+benchmarks, either by providing additional examples or by tuning termination parameters.
 Without these alterations, symmetric synthesis yielded a lens that fit
@@
               of merges.\\
-  \NSOpt{}: & We ignore all \Skip annotations in the SRE specifications. \\
-  \NROpt{}: & We ignore all \SRequire annotations in the SRE specifications. \\
+\end{tabulary}
+
+\begin{tabulary}{\linewidth}{rL}
+  \ZTOpt{}: & \ZTOpt{} runs the algorithm with the termination parameter
+              universally set to 0 (default). \\
+  \TFTOpt{}: & \TFTOpt{} runs the algorithm with the
+               termination parameter universally set to 25. \\
+  \NTFTOpt{}: & \NTFTOpt{} runs the algorithm with the
+                termination parameter universally set to -25. \\
 \end{tabulary}
@@
 
-Figure~\ref{fig:metric} summarizes the result of these experiments. The data
-reveal that the information-theoretic metric is critical for finding the correct
-lens: Only 10 of the benchmarks succeeded when running in \CCOpt{} mode. The
-termination condition is also quite important. When running in \FLOpt{} mode,
-the algorithm only discovers 5 lenses, which shows that the first cluster that
-contains a satisfying lens is rarely the correct cluster. However, our algorithm
-is not perfect and fails when either it is very difficult to find the desired
-lens (necessitating \SRequire) or when a large amount of data is projected
-(necessitating \Skip). Without any annotations, our algorithm finds the correct
-lens for 40 of our 48 benchmarks; eight required relevance annotations to find
-the correct lens.
+Figure~\ref{fig:metric}
+summarizes the result of these experiments. The data reveal that 
+the information-theoretic metric is critical for finding the correct
+lens: Only 10 of the benchmarks succeeded when running in \CCOpt{} mode.
+The termination condition is also quite important. When running
+in \FLOpt{} mode, the algorithm only discovers 5 lenses, which shows
+that the first cluster that contains a satisfying lens is rarely the
+correct cluster.  However, our termination condition is not perfect.
+Sometimes users must tell the algorithm to ``keep going'' or ``stop early.''
+Without termination parameters, our algorithm finds the correct lens for 39 of
+our 48 benchmarks; nine required tuning 
+to find the correct lens. Nevertheless, our default parameter sits in a
+good middle-ground. Both \TFTOpt{} and \NTFTOpt{} synthesized the correct lens
+less often than \ZTOpt.  \TFTOpt{} demands too good a lens, so the search
+procedure keeps spinning, while \NTFTOpt{} terminates early, not
+always trying hard enough to find the correct lens.
 % end evaluation
@@
     found, that lens is returned. \CCOpt{} synthesizes lenses, where the cost of
-    a lens is the number of disconnects plus the number of merges. \NSOpt
-    ignores all \Skip annotations while running the algorithm. \NROpt ignores
-    all \SRequire annotations while running the algorithm.}
+    a lens is the number of disconnects plus the number of merges. \ZTOpt{} runs
+    the algorithm with the termination parameter universally set to 0 (default).
+    \TFTOpt{} runs the algorithm with the termination parameter universally set
+    to 25. \NTFTOpt{} runs the algorithm with the termination parameter
+    universally set to -25.}
   \label{fig:metric}
@@
 \paragraph*{Symmetric Lenses}
-A classical symmetric lens\cite{symmetric-lenses} $\Lens$ between $X$ and $Y$ consists of 4 components:
+A classical symmetric lens $\Lens$ between $X$ and $Y$ consists of 4 components:
 a complement $C$, a designated element $init \in C$, and two functions,
@@
 Two classical symmetric lenses are equivalent if they output the same formats
-given any sequence of edits. Formally, given a lens $\Lens$ between $X$ and $Y$, an \emph{edit} for $\Lens$ is a member of $X + Y$. Consider the function
+given any sequence of edits. Formally, given a lens $\Lens \in X \leftrightarrow
+Y$, an \emph{edit} for $\Lens$ is a member of $X + Y$. Consider the function
 $apply$, which, given a lens and an element of that lens's complement, is a
 function from sequences of edits to sequences of edits. If $apply(\Lens,c,es) =
-es'$, then given complement $c$ and edit $es_i$, the lens $\Lens$ generates
+es'$, then given complement $c$ and edits $es_i$, the lens $\Lens$ generates
 $es'_i$.
@@
 
-While classical symmetric lenses are technically more expressive than simple
-symmetric lenses, we still believe simple symmetric lenses are useful in
-practice. Indeed, all asymmetric lenses (we later show) are expressible as
-simple symmetric lenses. The primary loss is the loss of ``memory'' within the
-complement. In classical symmetric lenses, \OrLens lenses retain information, so
-users don't lose information when swapping from one disjunct to the other, the
-information is stored in the complement. Simple symmetric lenses have no such
-complement, so they mimic the forgetful \OrLens lens of classical symmetric
-lenses.
-
-In our judgement, the complications they introduce into synthesis did
-not justify the potential additional expressiveness: because each lens has a
-custom complement, one can no longer specify the put functions through
-input/output examples alone. One alternative would be to enrich specifications
-with edit sequentials; another would be to specify the structure of complements
-explicitly (though the latter would be somewhat akin to specifying the internal
-state of a program). In either case, the complexity of the specifications
-increases.
+While classical symmetric lenses
+are technically more expressive than simple symmetric lenses, it is unclear
+whether this additional expressiveness is useful in practice.\bcp{That's a
+  total cop-out! We have much more to say about this point, and we {\em
+    should} have even more to say than we do.} In our judgement,
+the complications they introduce into synthesis did not justify the potential
+additional expressiveness: because each lens has a custom
+complement, one can no longer specify the put functions through input/output
+examples alone. One alternative would be to enrich specifications with edit
+sequentials; another would be to specify the structure of complements explicitly
+(though the latter would be somewhat akin to specifying the internal state of a
+program). In either case, the complexity of the specifications increases.
 %% addition, users would have to specify some auxilary 
@@
 
-\paragraph*{Symmetric Lenses vs.{} Simple Symmetric Lenses} To compare classical and simple
-symmetric lenses, we define an $apply$ function on simple symmetric lenses as well. If
+\paragraph*{Relation with Simple Symmetric Lenses} To compare classic and simple
+symmetric lenses, we define an $apply$ function on simple lenses as well. If
 $apply(\Lens,\None,es) = es'$, then starting with no prior data, after edit
@@
 
-\paragraph*{Other Lens Formulations}\bcp{This paragraph doesn't add
-  much}\afm{the POPL reviewers complained we didn't have this type of
-  paragraph, I personally agree that it doesn't add much and feels unnecessary}
-Bidirectional programming has an extensive literature, with many extensions onto
-basic lens formulations, like quotient lenses\cite{quotientlenses} and matching
-lenses\cite{matchinglenses}. Interested readers can consult
-\citet{DBLP:conf/icmt/CzarneckiFHLST09} for a slightly dated survey of lenses
-and other lens-like structures.
-
 \subsection{Data Transformation Synthesis}
@@
     \CreateL{} \App s & = & s\\
-    \PutR{} \App s \App t & = & s\\
-    \PutL{} \App s \App t & = & s\\
+    \PutR{} \App s_1 \App s_2 & = & s_1\\
+    \PutL{} \App s_1 \App s_2 & = & s_1\\
   \end{array}
@@
 Since $S$ is unambigous, then $S_1$ and $S_2$ are also unambiguous, and for each $s \in \mathcal{L}(S_1 \cdot S_2)$ there exists a unique $s_1 \in \mathcal{L}(S_1)$ and a unique $s_2 \in \mathcal{L}(S_2)$ such that $s = s_1 \cdot s_2$.
-By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$, hence\bcp{Don't understand what's going on with the $-$
-  signs in the last few lines here.}
+By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$, hence
 \begin{align*}
@@
 
-\bcp{``Since...then...'' is not english.}
-Since $S$ is unambiguous, then $S_1$ and $S_2$ are also unambiguous, and
-every $s \in \mathcal{L}(S_1 \; |_p \; S_2)$ is either in $\mathcal{L}(S_1)$
-or in $\mathcal{L}(S_2)$. By the induction hypothesis, $H(\mathcal{L}(S_1))
-= \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) =
-\mathbb{H}(S_2)$. Then\bcp{Again, don't understand what's going on with the $-$
-  signs.}
+Since $S$ is unambiguous, then $S_1$ and $S_2$ are also unambiguous, and every $s \in \mathcal{L}(S_1 \; |_p \; S_2)$ is either in $\mathcal{L}(S_1)$ or in $\mathcal{L}(S_2)$.
+By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$.
+Then
 \begin{align*}

### 2019-02-16 report/final-report.tex
start

\documentclass[12pt]{article}

\newif\ifdraft\drafttrue  % set true to show comments
% \newif\ifdraft\draftfalse  % set true to show comments
\newif\ifanon\anonfalse    % set true to suppress names, etc.
\newif\ifappendices\appendicesfalse

%\PassOptionsToPackage{usenames,dvipsnames,svgnames,table}{xcolor}

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage{amsmath}
\usepackage[capitalise]{cleveref}
\usepackage{makecell}%To keep spacing of text in tables
\usepackage{nccmath}
\usepackage{mathtools}
\usepackage{bussproofs}
\usepackage{varwidth}
\usepackage{amsthm}
\usepackage{csvsimple}
\usepackage{thmtools,thm-restate}
\usepackage{changepage}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage{enumitem}
\usepackage{multirow,bigdelim}
\usepackage{multicol}
\usepackage{siunitx}
\usepackage{listings}
\usepackage{letltxmacro}
\usepackage{sansmath}
\usepackage{url}
\usepackage{flushend}
\usepackage{microtype}
\usepackage[utf8]{inputenc}
\usepackage{mathpartir}
\usepackage{empheq}
\usepackage{array}
\usepackage{pgfplots}
\usepackage{stmaryrd}
\usepackage{courier}
\usepackage{qtree}
\usepackage[normalem]{ulem}
\usepackage{relsize}
\usepackage{tikz}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{tabularx}
\usepackage{stackengine}
\usepackage{caption}
\usepackage{wrapfig}
\usepackage{remreset}
\usepackage{tabulary}
\usepackage{xspace}
\usepackage{bbm}

\newtheorem*{theorem*}{Theorem}
\newenvironment{centermath}
 {\begin{center}$\displaystyle}
 {$\end{center}}
\setcellgapes{4pt}%parameter for the spacing

\lstset{ language=Caml, basicstyle=\upshape\sffamily,
keywordstyle=\upshape\sffamily\color{dkpurple}, keepspaces=true,
framexleftmargin=1ex, framexrightmargin=1ex, showstringspaces=true,
commentstyle=\itshape\rmfamily,
emph={rep,iterate,synth,collapse,perm,squash,normalize,using,ins,del,lens,let,get,put,rquot,lquot,id,swap,concat,or,disconnect,merge_left,merge_right,const},
emphstyle=\upshape\sffamily\color{dkpurple}, 
columns=fullflexible,
mathescape, 
xleftmargin=1.5em,
% BCP: I find this distracting:
stringstyle=\sffamily\color{dkblue},
}
\makeatletter
     \let\lst@oldvisiblespace\lst@visiblespace
     \def\lst@visiblespace{\,\lst@oldvisiblespace\,}
\makeatother

\setlength{\belowdisplayskip}{0pt} \setlength{\belowdisplayshortskip}{0pt}
\setlength{\abovedisplayskip}{0pt} \setlength{\abovedisplayshortskip}{0pt}

\usetikzlibrary{
  er,
  matrix,
  shapes,
  arrows,
  positioning,
  fit,
  calc,
  pgfplots.groupplots,
  arrows.meta
}
\tikzset{>={Latex}}

%%%% Hyperlinks – must come late!
%\usepackage[pdftex,%
%            pdfpagelabels,%
%            linkcolor=blue,%
%            citecolor=blue,%
%            filecolor=blue,%
%            urlcolor=blue]
%           {hyperref}

\input{macros}

\begin{document}

\pagestyle{empty}

\begin{center}

  \large \textbf{Synthesizing Data Wranglers\\\vspace{1cm} FA8750-17-2-0028}

  \vspace{1in}

  \normalsize
  %\begin{tabular}{rl}
  Kathleen Fisher (Tufts University) \\
  Benjamin Pierce  (University of Pennsylvania) \\
  David Walker (Princeton University) \\
  Steve Zdancewic (University of Pennsylvania)
 %\end{tabular}
  \vspace{1in}

  Final Report

\end{center}

\newpage
\pagestyle{plain}

\renewcommand{\thepage}{\roman{page}}% Roman numerals for page count
\setcounter{page}{1}% Start page number with 2

\tableofcontents
\newpage

\listoffigures
\newpage

\renewcommand{\thepage}{\arabic{page}}% Arabic numerals for page counter
\setcounter{page}{1}% Start page number with 2

\section{Summary}

\section{Introduction}

Building information-processing systems and maintaining them over a long
period of time is a tedious, labor-intensive process.  One key challenge is
that such systems must often interact with a large number of
\emph{ad hoc data sources}---partially structured data sources represented
in non-standard formats.  Ad hoc data sources include various
different kinds of system log files as well as scientific data sources
generated by experiments.  These ad hoc data sources are often
produced by other automated systems, and each requires custom tools.
Over time, the data sources tend to evolve---fields are added,
removed, or co-opted, variants are added, etc.  Today, to manage these
changes, engineers must manually re-code parsers and/or insert
adaptors while maintaining the desired semantics.  Such manual work is
not only time-consuming, but exceedingly error-prone.  Moreover,
errors in environment-facing interfaces can not only lead to
corruption of important data, but also to significant security
vulnerabilities.  In order to design and implement survivable,
long-lived, complex software systems that are robust to changes 
in their operational environment---the vision of
DARPA's BRASS program---it is necessary to develop new, easier-to-use
and more robust programming systems for managing these ad hoc data
sources as they evolve.

To help alleviate this problem, we studied algorithms 
for automatically synthesizing adaptors between related data
sources, given (i) the \emph{type} of each
source, as well as (ii) a small collection of
representative \emph{examples} of the desired translation.  The
adaptors we synthesize are \emph{bidirectional}, meaning
that we synthesize transformations that may be applied both
backwards and forwards (from source A to B as well as B back to A).
Such bidirectional transformations may help faciliate evolution and
maintainence of long-lived systems by making it possible to upgrade
one component of a (possibly distributed) system, while it continues
to interact correctly with other components and with its environment.
The bidirectional transformations will be guaranteed to preserve
strong invertability laws, thereby reducing the likelihood of
inadvertant data corruption.  In addition, parsing components
will be synthesized by a compiler rather than being manually coded,
thereby reducing the likelihood of the buffer overruns that lead to
many security vulnerabilities.

The core result of our research is a new algorithm for
the synthesis of \emph{bijective string lenses}.  Bijective string lenses (\emph{i.e.,}
bidirectional transformations) define a limited set of
transformations between strings.  The domain and range of such
transformations are determined by regular expressions ({\emph{i.e.,}
  regular expressions serve as the types of these transformations).
 In addition, as their name suggests, these transformations are
 \emph{bijections}.
 In other words, the information content of source A is preserved (though usually
rearranged) when data is translated to target B, and vice versa when B
is translated back to A.  Such transformations can rearrange fields of
a record or insert new kinds of syntactic separators (\emph{e.g.,} replacing a
comma with a vertical bar, or the name of one HTML tag with another)
but they cannot implement more general transformations that elide
irrelevant details, such as the amount of whitespace that separates
two tokens.

Despite their limitations, bijections and bijective synthesis form a
useful foundation on top which more general transformations between
data sets may be built.  Hence, in the second half of our of project,
we extended the regular expression-based specifications of data types
with \emph{isomorphisms} between strings, also known as
\emph{quotients}.  To synthesize transformations between sets of
strings modulo isomorphisms, we generate canonizing functions followed
by bijections (using the original bijective synthesis algorithm).  The
so-called \emph{quotient lenses}~\cite{Foster:quotient-lenses}we generate are capable of handling
``irrelevant'' differences between structures such as white space or
permutations of items in a row of data set and hence expand the set of
transformations our system can define significantly.  Finally, we explored the
synthesis of \emph{symmetric lenses}~\cite{hofmann+:lenses} , again building upon the
bijective platform that we started with.  Symmetric lenses are able to
ignore arbitrary chunks of data in A when generating target data B,
and vice-versa.  For example, a serial number specific to one data set
may be ignored when we translate to a second.  Symmetric lenses
further expand the set of allowed transformations.

We measured the effectiveness of our algorithms on a set of
benchmarks drawn from the Augeas system~\cite{augeas}---a system for
transforming and editing Linux configuration files.  Past synthesis
tools, such as FlashFill~\cite{flashfill} were unable to translate
most Augeas file formats.  However, our synthesis toolkit was able to
transform all 40 of the Augeas file formats that we analyzed.

We have written three academic papers~\cite{?}
describing and evaluating our algorithms and produced open-source our code.

\section{Methods, Assumptions, and Procedures}

\section{Results and Discussion}

Reffing a \Cref{fig:eg}.

\begin{figure}
  a picture is here
  \caption{An example figure}
  \label{fig:eg}
\end{figure}

\section{Conclusions}

During the course of this project, we have designed, analyzed and implemented algorithms that demonstrate it is possible to synthesize three classes of bidirectional transformations:  (1) pure bijective transformations, (2) bijections modulo equivalences classes and (3) bijections modulo projections.  The synthesis algorithms we have designed take inputs that include a format specification (which includes specification of equivalence classes) and a collection of examples.  As the class of transformations becomes richer, the number of potential programs grows dramatically.  As a result, the corresponding inference algorithm slows and its ability to guess the transformation desired by the user decreases, leading to reduced accuracy.  We found that it was possible to overcome such challenges through new heuristics that use information theory to help guide the search for program transformations.  In addition, we found the use of compositional synthesis, which is the process of breaking down a complex synthesis task into smaller, more mangeable subtasks critical when attempting to scale our prototype system up to be able to handle larger transformation task.

\bibliographystyle{apalike}
\bibliography{local.bib,bcp.bib}

\section{List of Symbols, Abbreviations, and Acronyms}
\begin{tabular}{ll}
  DARPA & Defense Advanced Research Project Agency \\
\end{tabular}

\end{document}

%%% Local Variables:
%%% mode: latex
%%% TeX-master: t
%%% End:

### 2019-02-18 papers/SymmetricLenses/main.tex
updated with dave's suggestions

--- 2019-02-12 papers/SymmetricLenses/main.tex
@@
 in both directions.
-Since their introduction by~\citet{Focal2005-long}, lenses have been extensively
-studied, extended, and applied.  Recent work~\cite{optician} has also demonstrated how
+Since their introduction by Foster \ETAL, lenses have been extensively
+studied, extended, and applied.  Recent work has also demonstrated how
 techniques from {\em type-directed program synthesis} can be used to
@@
 of the bijective lenses and the more widely used ``asymmetric''
-lenses~\cite{Focal2005-long}, but also a rich subset of the full-blown
-``symmetric lenses'' proposed by \citet{symmetric-lenses}.
+lenses, but also a rich subset of the full-blown
+``symmetric lenses'' proposed by Hofmann \ETAL.
 
@@
 synthesizing bijective lenses: Since some of the information on each side
-can be ``disconnected'' from the other side, there will in general be {\em
+can be ``disconnected'' from the other side, there will, in general, be {\em
   many} lenses that agree with a given example. To guide the search process,
@@
 We describe an implementation of simple symmetric lenses and our synthesis
-procedure as extensions to the Boomerang language~\cite{Boomerang}. We evaluate
-its performance on 48 benchmark examples drawn from Flash Fill~\cite{flashfill},
-Augeus~\cite{augeas2}, the bidirectional programming literature, and electronic
+procedure as extensions to the Boomerang language. We evaluate
+its performance on 48 benchmark examples drawn from Flash Fill,
+Augeus, the bidirectional programming literature, and electronic
 file format synchronization tasks. Our implementation can synthesize all these
@@
 between different representations.
-Domain-specific languages for writing lenses can facilitate building
-principled data transformations.  For example, Boomerang~cite{boomerang} is a
+Domain-specific languages for writing lenses can facilitate generation of
+principled data transformations.  For example, Boomerang~\cite{boomerang} is a
 language for expressing bidirectional string transformations.
@@
 might have a different fifth column.  Or, one ad-hoc system configuration
-file might include some format-specific meta data, such as a date or a
+file might include some format-specific metadata, such as a date or a
 reference number, whereas the same configuration file on another operating
@@
 $C$ that contains the information not
-present in either $X$ or $Y$. Logistically, these complements must be stored
+present in either $X$ or $Y$. These complements must be stored
 and managed somehow.  More fundamentially, complements complicate
@@
 strings and a probability distribution over those strings.  Using this
-probability distribution, we can calculate the shared information content of
-two formats and our measure of likelihood.
+probability distribution, we can calculate the likelihood of a given lens.
 
@@
 selected by the user: in general, the synthesis algorithm has to search through
-equivalent regular expression types for which a suitable lens exists. To tame
+equivalent regular expression types to find the most likely lens. To tame
 this complexity, we divide the synthesis algorithm into two communicating search
@@
 its likelihood score to \Expand. The whole synthesis algorithm heuristically
-terminates when the most likely lens is no longer unlikely enough to justify
-the time cost of searching through newly generated SREs.
+terminates when a sufficiently likely lens is found.
 
@@
 
-We have added this synthesis procedure to the Boomerang system and evaluated
-its effectiveness on a benchmark suite of 48 problems, including data-cleaning
+We added this synthesis procedure to the Boomerang system and explored its
+effectiveness on a range of applications. Users set up a Boomerang synthesis
+task by providing the source and target types and optionally supplying one or
+more examples. In addition, users can override the default mechanism for
+calculating the information content of a SRE by
+asserting that certain strings are \emph{essential} or \emph{irrelevant},
+thereby forcing certain data to either be retained or discarded during the
+transformations. We evaluated our system on a benchmark suite of 48 problems, including data-cleaning
 tasks taken from Flash Fill~\cite{gulwani-popl-2014}, view-maintenance tasks
@@
 variety of other sources (mostly from the bidirectional programming literature).
-The user can control the search process via examples and annotations
-that informs the synthesis algorithm as to whether certain bits of information
-are irrelevant, or must maintained.
 
@@
   quotient lenses~\cite{quotientlenses} and matching
-  lenses~\cite{matchinglenses}. We evaluate our implementation, the effects of
+  lenses~\cite{matchinglenses}.
+
+\item We also extend Boomerang with symmetric lens synthesis specifications,
+  which include types, examples, and relevance annotations. We evaluate our implementation, the effects of
   various optimizations, and our inclusion of relevance annotations on a set of

### 2019-02-19 report/final-report.tex
some work on the methods section

@@
 \usepackage{bbm}
+\usepackage{fullpage}
+\usepackage[explicit]{titlesec}
+
+% \renewcommand*\thesection{\arabic{section}.0}
+% \renewcommand*\thesubsection{\arabic{section}.\arabic{subsection}}
 
@@
 
-\section{Summary}
+\section{SUMMARY}
 
-\section{Introduction}
+\section{INTRODUCTION}
 
@@
 We have written three academic papers~\cite{?}
-describing and evaluating our algorithms and produced open-source our code.
+describing and evaluating our algorithms and produced open-source our code that
+implements those algorithms.
 
-\section{Methods, Assumptions, and Procedures}
+\section{METHODS, ASSUMPTIONS, AND PROCEDURES}
 
-\section{Results and Discussion}
+Our approach combines two strategies.  We use \textit{domain-specific programming
+  language}, specifically, the Boomerang~\cite{boomerang, Matching10} language,
+which is designed for writing bidirectional string transformations.  We also exploit
+\textit{type- and example-directed program synthesis}, which uses enumerative
+techniques to search a space of candidate solutions.  The search is constrained
+by the type information about the program, as well as by example instances
+provided by the user.
+
+\subsection{Lenses and lens domain-specific programming languages}
+
+A lens comprises two functions,
+\emph{get} and \emph{put}.  The \emph{get} function translates
+source data into the target format.  If the target data is updated, the
+\emph{put} function translates this edited data back into the
+source format.  
+A benefit of lens-based languages is that they use a single term
+to express both 
+\emph{get} and \emph{put}.
+Furthermore, well-typed lenses give rise to 
+\emph{get} and \emph{put} functions 
+guaranteed to satisfy desirable invertibility properties.
+
+Lens-based languages are present in variety of tools and have found mainstream
+industrial use.
+Boomerang~\cite{boomerang, Matching10} lenses provide
+guarantees on transformations between {\em ad hoc} string document formats.
+Augeas~\cite{augeas}, a popular tool that reads Linux system configuration
+files, uses the \emph{get} part of a lens to transform configuration
+files into a canonical tree representation that users can edit
+either manually or 
+programmatically.  It uses the lens's \emph{put} to merge the edited
+results back into the original string format.  Other lens-based languages and
+tools include 
+%
+GRoundTram~\cite{Hidaka2011GRoundTramAI},
+%
+BiFluX~\cite{DBLP:conf/ppdp/PachecoZH14}, 
+%
+BiYacc~\cite{DBLP:conf/staf/ZhuK0SH15},
+%
+Brul~\cite{DBLP:conf/etaps/ZanLKH16},
+%
+BiGUL \cite{DBLP:conf/pepm/KoZH16}, 
+%
+bidirectional variants of 
+relational algebra~\cite{BohannonPierceVaughan},
+spreadsheet formulas~\cite{DBLP:conf/vl/MacedoPSC14},
+graph query languages~\cite{DBLP:conf/icfp/HidakaHIKMN10},
+and
+XML transformation languages~\cite{DBLP:conf/pepm/LiuHT07}.
+
+\subsection{Bijective lens synthesis}
+
+As inputs, our bijective synthesis procedure takes regular expressions specifying the
+source, $S$, and
+target, $T$, formats, plus a collection of concrete examples of the desired 
+transformation.  Format specifications are supplied as ordinary regular
+expressions.
+Because regular expressions are so widely understood, we anticipate such
+inputs will be substantially easier for everyday programmers to work with
+than the unfamiliar syntax of lenses.
+Moreover, these format descriptions communicate a
+great deal of information to the synthesis system.  Thus, requiring user input
+of regular expressions makes synthesis robust, 
+helps the system scale to large and complex data sources, and 
+constrains the search space sufficiently that the user typically needs
+to give very few, if any, examples.
+
+The goal of the synthesis algorithm is to find a lens
+$\ell : S \Leftrightarrow T$ that corresponds to a bijection from the language
+of a source regular expression $S$ to a target language of a target regular
+expression $T$.   Such a lens should satisfy the \textit{bijective lens laws}:
+
+\begin{equation}\label{bijectivelenslaws} \ell.\get \;
+  (\ell.\lput \; t) = t \text{, and } \ell.\lput \; (\ell.\get \; s) = s
+\end{equation}
+
+\begin{figure}
+  \centering
+  \small 
+  \begin{tikzpicture}[auto,node distance=1.5cm]
+    \node[text width=1.5cm,minimum height=.6cm,align=center,draw,rectangle] (todnfregex) {\ToDNFRegex{}};
+
+    \node[align=right, anchor=east] (regex1) [left = .6cm of todnfregex.north west]{\Regex{}};
+    \node[align=right, anchor=east] (regex2) [left = .6cm of todnfregex.south west]{\RegexAlt{}};
+    \node[align=right, anchor=east] (exs) [below = .2cm of regex2]{ \Examples{} };
+
+    \node[align=center] (dnfregex1) [right = .4cm of todnfregex.north east]{\DNFRegex{}};
+    \node[align=center] (dnfregex2) [right = .4cm of todnfregex.south east]{\DNFRegexAlt{}};
+
+    \node[text width=2.3cm,minimum height=.6cm,align=center,draw,rectangle] [right = 1.45cm of todnfregex.east] (synthdnflens) {\SynthDNFLens{}};
+    \node[align=center] [above = .7cm of synthdnflens] (optician) {\Optician{}};
+
+    \node[align=center] [right = .4cm of synthdnflens] (dnflens) {\DNFLens{}};
+
+    \node[text width=1.5cm,minimum height=.6cm,align=center,draw,rectangle] [right = .4cm of dnflens] (tolens) {\ToLens{}};
+
+    \node[align=center] [right = .6cm of tolens] (lens) {\Lens{}};
+
+    \path[->] (regex1.east) edge (todnfregex.north west);
+    \path[->] (regex2.east) edge (todnfregex.south west);
+
+    \path[->] (todnfregex.north east) edge (dnfregex1.west);
+    \path[->] (todnfregex.south east) edge (dnfregex2.west);
+
+    \path[->] (dnfregex1.east) edge (synthdnflens.north west);
+    \path[->] (dnfregex2.east) edge (synthdnflens.south west);
+
+    \path[->] (synthdnflens) edge (dnflens);
+
+    \path[->] (dnflens) edge (tolens);
+
+    \path[->] (tolens) edge (lens);
+    \draw[->] ($(exs.east)+(-3pt,0)$) -| node(exsedge) {} (synthdnflens);
+    \node[fit={($(todnfregex.west)+(-4pt,0)$) ($(tolens.east)+(4pt,0)$) (exsedge) (dnfregex1) (optician) (dnfregex2)},draw] (surrounding) {};
+    % Now place a relation (ID=rel1)
+    %\node[text width=2cm,align=center,draw, rectangle] (sketch-gen) [right = .75cm of spec] {\TypeProp{}};
+    %\node (below-gen) [below=.5cm of sketch-gen] {};
+    %\node[text width=2cm,align=center,draw, rectangle] (sketch-compl)
+    %     [right = .25cm of sketch-gen] {\RigidSynth{}};
+    %\node (below-compl) [below=.5cm of sketch-compl] {};
+    %\node[align=center] (lens) [right = .75cm of sketch-compl] {Lens}; 
+    %% Draw an edge between rel1 and node1; rel1 and node2
+    %\path[->] (spec) edge node (start-alg) {} (sketch-gen);
+    %\path[->] (sketch-gen) edge node(middle) {} (sketch-compl);
+    %\path[->] (sketch-compl) edge node[near start](success) {\Success{}} (lens);
+    %\draw[<-] (sketch-gen.south) -- +(0,-.5) -| node[above left](failure){\Failure{}} (sketch-compl.south);
+
+    %\node (synth-name) [above=.5cm of middle] {\Optician{}};
+    %
+    %\node[fit=(sketch-gen) (sketch-compl) (start-alg) (synth-name) (failure)
+    %(success) ,draw] (surrounding) {};
+
+  \end{tikzpicture}
+  \caption{Schematic Diagram for \Optician{}.  Regular expressions, \Regex{} and
+    \RegexAlt{}, and examples, \Examples{}, are given as input.
+    First, the function \ToDNFRegex{} converts \Regex{} and \RegexAlt{} into
+    their respective DNF forms, \DNFRegex{} and \DNFRegexAlt{}.
+    Next, \SynthDNFLens{} synthesizes a DNF lens, \DNFLens{}, from \Regex{},
+    \RegexAlt{}, and \Examples{}.
+    Finally, \ToLens{} converts \DNFLens{} into \Lens{}, a lens in Boomerang
+    that is equivalent to \DNFLens{}.}
+  \label{fig:schematic-diagram-synthesis}
+\end{figure}
+
+Figure~\Cref{fig:schematic-diagram-synthesis} shows a high-level,
+schematic diagram for \Optician{}.
+First, \Optician{} uses the function \ToDNFRegex{} to convert the input
+regular expressions into DNF regular expressions.  Next,
+\SynthDNFLens{} performs type-directed synthesis on these DNF regular
+expressions and the input examples to synthesize a DNF lens.  Finally, this DNF lens is
+converted back into a regular lens with the function \ToLens{}, and returned to the user.
+
+\subsection{Quotient lens synthesis}
+
+Quotient lenses are lenses in which the lens laws are
+loosened so that they hold modulo an equivalence relation on the source and
+target data respectively; as above, we are concerned with {\em bijective
+quotient lenses} which are lenses for which Equation \ref{bijectivelenslaws}
+holds modulo equivalence relations $\equiv_S$ and $\equiv_T$ defined on the source
+and target data respectively:
+
+\begin{equation}\label{quotientlenslaws}
+\ell.\get \; (\ell.\lput \; t) \equiv_T t \text{, and } \ell.\lput \; (\ell.\get
+\; s) \equiv_S s
+\end{equation}
+
+The inputs to the quotient lens synthesis problem, as in the bijective case,
+include source and target regular expressions $S$ and $T$ and a set of example
+instances.  The equivalence relations $\equiv_S$ and $\equiv_T$ are also given
+as inputs to the synthesis algorithm.  As we describe below, one result of our
+work is a language of \textit{quotient regular expressions} that concisely
+specifies $S$ and $\equiv_S$ simultaneously.
+
+\section{RESULTS AND DISCUSSION}
 
@@
 
-\section{Conclusions}
+\section{CONCLUSIONS}
 

### 2019-02-19 papers/BXProposal/main.tex

@@
 
+\vspace*{-3em}
+
 \begin{abstract}
-  {\color{dkblue} \emph{AFM: I will remove this from the submission, but am including it in the
-    document for people to review. It is due today (February 12)}} Recent work
+  Recent work
   has shown how to synthesize a class of string bijections from regular
@@
 
-Instead of manually writing explicit canonization and projection functions, we
+Instead of writing explicit canonization and projection functions, we
 consider an alternative approach: by annotating our regular expressions with
@@
 synthesizing bidirectional transformations that are bijections (up to
-quotients)~\cite{maina+:quotient-synthesis}, but here we extend it with projections as well.
+quotients)~\cite{maina+:quotient-synthesis}, here we extend it with projections as well.
 
@@
 %
-Synthesizing with these annotated regular expressions performs the same steps
-that we previously did manually: it breaks the lens into concatenation and
-normalization subcomponents, sends the bijective core to the existing synthesis
-algorithm, and composes the generated subcomponents with the synthesized
-bijective core.
 

### 2019-02-19 papers/SymmetricLenses/main.tex
bx proposal changes and make an algorithm more horizontal
add macros and main

@@
 % \newif\ifdraft\draftfalse  % set true to show comments
-\newif\ifanon\anonfalse    % set true to suppress names, etc.
+\newif\ifanon\anontrue    % set true to suppress names, etc.
 \newif\ifappendices\appendicesfalse
@@
 \PassOptionsToPackage{usenames,dvipsnames,svgnames,table}{xcolor}
+%\documentclass[acmsmall,review,screen,anonymous]{acmart}
 \documentclass[acmsmall,review,screen,anonymous]{acmart}
@@
 stringstyle=\sffamily\color{dkblue},
+aboveskip=.3em,
+belowskip=.3em,
 }
@@
 \begin{figure}
-  \includegraphics[width=.5\textwidth]{high-level-algorithm.pdf}
+  \includegraphics[width=.63\textwidth]{high-level-algorithm.pdf}
   \vspace{-2ex}
@@
     \GreedySynth. \GreedySynth finds a lens typed between the supplied SREs.
-    When the algorithm finds a likely lens, it returns it.\bcp{Later: We
-      could save significant space if we put the inputs and outputs on the
-      sides instead of top and bottom}}
+    When the algorithm finds a likely lens, it returns it.}
   \label{fig:high-level-algorithm}
@@
 %\end{enumerate}
-\vskip 1.5ex
 \begin{center}
@@
 \end{center}
-\vskip 1ex
 \noindent
+\vskip -.5em
 The two \KW{create} functions are used to propagate shared information and fill
@@
 \begin{figure}
-  \begin{tabular}{@{\hspace*{5em}}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{}}
-    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} \\
+  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r@{\hspace*{3em}}r@{\hspace{1em}}c@{\hspace{1em}}l@{}r}
+    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} &
     $\RegexConcat{\Regex}{\emptyset}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionRightRule{} \\
-    $\RegexConcat{\emptyset}{\Regex}$ & $\SSREquiv$ & $\emptyset$ & \EmptyProjectionLeftRule{} \\
+    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} &
+    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
+    \PRegexOr{\Regex}{\emptyset}{1} & $\SSREquiv$ & \Regex{} & \OrIdentityRule{} &
+    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
+    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} &
+    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
+    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{1-\Probability} & \UnrollstarLeftRule{} &
+    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{1-\Probability} & \UnrollstarRightRule{} 
+  \end{tabular}
+  \begin{tabular}{@{}r@{\hspace{1em}}c@{\hspace{1em}}l@{\hspace{1em}}r@{}}
     \RegexConcat{(\RegexConcat{\Regex{}}{\Regex'})}{\Regex''} & $\SSREquiv$ & \RegexConcat{\Regex{}}{(\RegexConcat{\Regex'}{\Regex''})} & \ConcatAssocRule{}  \\
     \PRegexOr{(\PRegexOr{\Regex}{\Regex'}{\Probability_1})}{\Regex''}{\Probability_2} & $\SSREquiv$ & \PRegexOr{\Regex}{(\PRegexOr{\Regex'}{\Regex''}{\frac{(1-\Probability_1)\Probability_2}{1 - \Probability_1\Probability_2}})}{\Probability_1\Probability_2} & \OrAssociativityRule{}  \\
-    \PRegexOr{\Regex{}}{\RegexAlt{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\RegexAlt{}}{\Regex{}}{1-\Probability} & \OrCommutativityRule{}\\
-    \RegexConcat{\Regex{}}{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}}{\Regex{}'})}{(\RegexConcat{\Regex{}}{\Regex{}''})}{\Probability} & \DistributivityLeftRule{} \\
-    \RegexConcat{(\PRegexOr{\Regex{}'}{\Regex{}''}{\Probability})}{\Regex{}} & $\SSREquiv$ & \PRegexOr{(\RegexConcat{\Regex{}'}{\Regex{}})}{(\RegexConcat{\Regex{}''}{\Regex{}})}{\Probability} & \DistributivityRightRule{} \\
-    \RegexConcat{\EmptyString{}}{\Regex{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityLeftRule{} \\
-    \RegexConcat{\Regex{}}{\EmptyString{}} & $\SSREquiv$ & \Regex{} & \ConcatIdentityRightRule{} \\
-    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\Regex{}}{\PRegexStar{\Regex{}}{\Probability}})}{1-\Probability} & \UnrollstarLeftRule{} \\
-    \PRegexStar{\Regex{}}{\Probability} & $\SSREquiv$ & \PRegexOr{\EmptyString{}}{(\RegexConcat{\PRegexStar{{\Regex{}}}{\Probability}}{\Regex{}})}{1-\Probability} & \UnrollstarRightRule{} 
   \end{tabular}
@@
 equations required to relate two \SDNFREabbrevs, only those relating to Kleene
-star (\EG, \emph{Unrollstar$_L$} and \emph{Unrollstar$_R$}) are needed to relate
+star (\EG, \UnrollstarLeftRule{} and \UnrollstarLeftRule{}) are needed to relate
 two \SDNFREabbrevs; other axioms, like associativity, commutativity or

### 2019-02-19 papers/QuotientLenses/main.tex
Made changes for publishing

@@
 
-\setcopyright{none}
+\setcopyright{rightsretained}
 \acmPrice{}
-\acmDOI{10.1145/3234603}
+\acmDOI{10.1145/3236775}
 \acmYear{2018}
@@
 \acmNumber{ICFP}
-\acmArticle{14}
+\acmArticle{80}
 \acmMonth{9}
@@
 \usepackage{microtype}
+\usepackage[all]{nowidow}
 
@@
 inessential aspects of string data formats; each QRE specifies,
-simulteneously, a regular language and an equivalence relation on it.
+simultaneously, a regular language and an equivalence relation on it.
 Second, we introduce {\em QRE lenses}, i.e., lenses mapping between QREs.
@@
 Theorem~\ref{normal form}.
-
-\begin{acks}
-  We thank our anonymous reviewers for their useful feedback and discussions and Nate Foster for his extensive assistance in integrating
-  QRE-enhanced \QOpt{} into Boomerang.  This research has been supported in part by DARPA
-  award FA8750-17-2-0028 and ONR 568751 (SynCrypt).
-\end{acks}
 
@@
 With soundness and completeness, we trivially get the correctness theorem.
+\begin{acks}
+  We thank our anonymous reviewers for their useful feedback and discussions and Nate Foster for his extensive assistance in integrating
+  QRE-enhanced \QOpt{} into Boomerang.  This research has been supported in part by DARPA
+  award FA8750-17-2-0028 and ONR 568751 (SynCrypt).
+\end{acks}
 

### 2019-02-19 papers/SymmetricLenses/main.tex
update

@@
 enough in practice, we add simple symmetric lenses to the Boomerang
-language~\cite{Boomerang} and apply them to a range of real-world
+language~\cite{boomerang} and apply them to a range of real-world
 applications, showing in the process that simple symmetric lenses integrate
@@
 include the semantics for combinators equivalent to those presented in prior
-work~\cite{Boomerang}. Furthermore, we present
+work~\cite{boomerang}. Furthermore, we present
 just \CreateR and \PutR in cases where \CreateL and \PutL are symmetric;
@@
 two formats, and one data has only one format. In previous
-work~\cite{Boomerang}, this was combined into \OrLens{}, where
+work~\cite{boomerang}, this was combined into \OrLens{}, where
 \OrLens{} could have ambiguous types, but we find it more clear to have explicit

### 2019-02-19 papers/SymmetricLenses/main.tex
fixes

@@
 \subsection{Simple Symmetric Lenses}
-Semantically, a simple symmetric lens between sets $X$ and
-$Y$ comprises four functions subject to four
-``round-tripping'' laws:
+Semantically, a simple symmetric lens (or lens when clear from context) between
+sets $X$ and $Y$ comprises four functions subject to four ``round-tripping''
+laws.
 
@@
 the other by combining a new value from one with an old value from the other.
-The notation $\Lens.\PutR$ extracts the \PutR function from simple symmetric
+The notation $\Lens.\PutR$ extracts the \PutR function from the
 lens $\Lens$; similarly for \PutL. These four functions can be used to keep the
@@
 side (the result of a put or create) back through a lens in the other direction
-will get back to exactly where we began. For example, consider,
-\lstinline{emp_lens}, a lens between
-\lstinline{emp_salaries} and
-\lstinline{emp_insurance}. Applying
-\lstinline{emp_lens}.\CreateR to a salary file creates a insurance file with the
-same employee names in the file, with \lstinline{"UNK"} for each insurance
-company. The round-tripping laws guarantee that using \lstinline{emp_lens}.\PutL
-to put the generated insurance file into the original salary file will return
-the original salary file, unmodified. If \KW{put}ting a string $s$ into a string
-$t$ results in the original string $s$, then $s$ and $t$ are considered
-\emph{synchronized}.
+will get back to exactly where we began. For example, consider a simple
+symmetric lens
+synchronizing the employee data formats in
+Figure~\ref{fig:minimized-representations}. Applying the \CreateR function to a
+salary file creates a insurance file with the same employee names in the file,
+with
+\lstinline{"UNK"} for each insurance company. The round-tripping
+laws guarantees that applying the \PutL function to the generated insurance file
+and the original salary file will return the original salary file, unmodified.
+If \KW{put}ting a string $s$ into a string $t$ results in the original string $s$, then $s$ and $t$ are considered \emph{synchronized}.
 
@@
 laws. But a better idea is to streamline the process by providing a set of
-combinators---a domain-specific language---for building law-abiding complex
+combinators---a domain-specific language---for building complex, law-abiding
 lenses from simpler ones. Such bidirectional languages have been given for
@@
 described by regular expressions.
-
 If $\BRegex$ and $\BRegexAlt$ are regular expressions, then
@@
 $\LanguageOf{\BRegexAlt}$. (We will use undecorated variables later for
-stochastic regular expressions; we mark plain REs with overbars.)  When
-clear from context, we shorten ``simple symmetric lens'' to just
-``lens.''
-
+stochastic regular expressions; we mark plain REs with overbars.)
 We illustrate some of these
-combinators by defining a lens between the two employee data formats
-in Figure~\ref{fig:minimized-representations}. The finished lens, which appears
-in Figure~\ref{fig:example_lens}, has type
-\lstinline{emp_salaries} $\Leftrightarrow$
-\lstinline{emp_insurance}.
+combinators by defining lenses on subcomponents of the employee data formats.
 
@@
 %
-The identity lens fully connects the corresponding data in the two formats:
-when one is updated, the other gets the exact same value. Both the \CreateR
-and \CreateL functions are the identity function ($\CreateR \App s = s$),
-and the put functions merely return the first argument
+The identity lens moves data back and forth from source to target without
+changing it. Both the \CreateR and \CreateL functions are the identity function
+($\CreateR \App s = s$), and the put functions merely return the first argument
 ($\PutROf{\String}{\StringAlt} = \String$).
-%
-In our example, when a \lstinline{name} is updated in one file, the identity
-lens ensures it is also updated in the other file.
-
-In contrast, the \Disconnect lens does not propagate any edits at all
-from one format to the other.  It takes four arguments: two
-regular expressions and two strings. The regular expressions specify the formats
+
+In contrast, $\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}$ does not propagate any data at all
+from one format to the other.  The \Disconnect lens takes four arguments: two
+regular expressions ($\Regex, \RegexAlt$) and two strings ($\String, \StringAlt$). The regular expressions specify the formats
 on the two sides, while the strings provide default values. 
@@
 \lstinline{"unk"}$), and on puts, the second argument is used and the first is
-thrown away ($\PutROf{\String}{\StringAlt} = \StringAlt$). In our extended
+thrown away ($\PutROf{\String}{\StringAlt} = \StringAlt$). For 
 example, the \lstinline{salary} field is only present in management files,
-so we use the disconnect lens to ensure salary edits do not cause updates to the
+so the disconnect lens can ensure salary edits do not cause updates to the
 HR file.
@@
 ($\OrLensOf{\Lens_1}{\Lens_2}$), and iteration ($\IterateLensOf{\Lens}$).
-For example, the composite lens \lstinline{name_lens} in the running
-example
+For example, we could use the following lens
 %
@@
 %
-will transform ``\lstinline{Jane Doe: }'' to ``\lstinline{Jane Doe,}'' in
+to transform ``\lstinline{Jane Doe: }'' to ``\lstinline{Jane Doe,}'' in
 the left-to-right direction.
@@
 %\end{lstlisting}
-The \IterateLens lens also appears in our running example, where it
-synchronizes data for a list of employees when given a
-lens \lstinline{employee_lens} that synchronizes  data for a single employee:
+The \IterateLens lens is useful for synchronizing a series of items or rows in a
+table. For example given a
+lens \lstinline{employee_lens} that synchronizes data for a single employee the
+following lens
 %
@@
 %
+transforms a list of employees in employees in the Management format to a list
+of employees in the HR format in the left-to-right direction. These combinators
+are combined in figure~\ref{fig:example_lens} to construct a complete lens
+between the employee formats.
 
@@
 
-We provide typing rules for the simple symmetric combinators, most of
-them syntax directed. However, regular expressions have a number of
-equivalences on them, and sometimes we must convert one regular expression type
-to another, equivalent, regular expression type.
+Most of the typing rules for lenses are syntax driven, where only a handful of
+rules result in a given type.  For example, the type of a
+\ConcatLens lens is the concatenation of the regular expressions of the sub-lenses.
+\[
+  \inferrule*
+  {
+    \Lens_1 \OfType \BRegex_1 \Leftrightarrow \BRegexAlt_1\\
+    \Lens_2 \OfType \BRegex_2 \Leftrightarrow \BRegexAlt_2
+  }
+  {
+    \ConcatLensOf{\Lens_1}{\Lens_2} \OfType \BRegex_1 \Concat \BRegex_2
+    \Leftrightarrow
+    \BRegexAlt_1 \Concat \BRegexAlt_2
+  }
+\]
+With one notable exception, lens typing rules maintain structural similarity
+between regular expression types and lens terms; type equivalence does not. Type
+equivalence permits a well-typed lens to be well-typed among any equivalent
+regular expression pair.
 %
@@
 
-\subsection{Synthesizing Simple Symmetric Lenses}
-
-Lens expressions for small examples like the one above are pretty easy to
-write.  But for larger and more complex formats, it can become quite difficult to
+\subsection{Ranking Simple Symmetric Lenses}
+
+Even for small examples, lenses like the one above take a little getting used
+to. For larger and more complex formats, it can become quite difficult to
 write lenses by hand.  Instead, we wish to synthesize them from examples. 
-
 That is, given a pair of regular expressions and a set of input-output examples, we want
@@
 \begin{lstlisting}
-disconnect(emp_salaries, emp_insurance, rep(emp_salaries), rep(emp_insurance))
+disconnect(emp_salaries, emp_insurance, ..., ...)
 \end{lstlisting}
 %
-is another well-typed lens (where \lstinline{rep($\BRegex$)} denotes some string in the
-language of $\BRegex$).  Adding the input-output example in
-Figure~\ref{fig:minimized-representations} 
-helps, but only a little: In general, {\em many} examples may be required to rule
-out all possible occurrences of nested \Disconnect lenses,
+is another well-typed lens that satisfies the example in
+Figure~\ref{fig:minimized-representations} (where \lstinline{...} represents
+components of the example itself). In general, {\em many} examples may be
+required to rule out all possible occurrences of nested \Disconnect lenses,
 particularly in complex formats.
@@
 How can we identify such a ``likely'' lens? We propose the following heuristic:
-A satisfying lens is ``more likely'' if it ``connects'' more data between the
-two formats, \EG, by using an identity rather than a disconnect lens. The
-likeliest satisfying lenses would connect as much data as possible. More
-formally, we define the likelihood of a satisfying lens as the expected number
-of bits required to recover data in one format from synchronized data in the
-other. Higher likelihoods correspond to fewer bits. Yet more formally, we say
-that two strings $s$ and $t$ are \emph{synchronized according to lens $l$} if
-$l.\PutR\ s\ t = t$ and $l.\PutL\ t\ s = s$. We say that we can \emph{recover}
-$s$ from $t$ using bits $b$ and lens $l$ if we can reconstruct $s$ from $t$,
-$b$, and $l$. For example, given the \IdentityLens{} lens, we can recover $s$
-from $t$ using no bits because, in this case, $s$ is just $t$. In contrast,
-given the \Disconnect{} lens, we need enough bits to fully encode $s$ in order
-to recover it from $t$ because $t$ contains no useful information. Thus, if both
+A satisfying lens is ``more likely'' if it u;ses more data from one format to
+construct the other. One manifestation of this principle involves choosing the
+identity lens (which uses all the data) rather than the disconnect lens (which
+uses none of the data). We define the likelihood of a satisfying lens as the
+expected number of bits required to recover data in one format from synchronized
+data in the other. Higher likelihoods correspond to fewer bits. We say that two
+strings $s$ and $t$ are \emph{synchronized according to lens $l$} if $l.\PutR\
+s\ t = t$ and $l.\PutL\ t\ s = s$. We say that we can \emph{recover} $s$ from
+$t$ using bits $b$ and lens $l$ if we can reconstruct $s$ from $t$, $b$, and
+$l$. For example, given the \IdentityLens{} lens, we can recover $s$ from $t$
+using no bits because, in this case, $s$ is just $t$. In contrast, given the
+\Disconnect{} lens, we need enough bits to fully encode $s$ in order to recover
+it from $t$ because $t$ contains no useful information. Thus, if both
 \IdentityLens{} and \Disconnect{} are satisfying lenses for a particular pair of
@@
 $S$, the entropy $\EntropyOf{S,P}$ is $-\Sigma_{s\in S}P(s)\Log_2P(s)$. In
-information theory, each element of $S$ has \emph{information content}, the
-number of bits required to specify that element. The information content of an
-element $s$ is $-\Log_2P(s)$. For example, let $S = \SetOf{a,b_1,b_2}$, where
-$P(a)=.5$ and $P(b_1) = .25 = P(b_2)$. The information content of $a$ is $1$,
-and the information content of each $b_i$ is 2. The entropy of $S$ calculates
-the expected number of bits required to specify a random element pulled from
-$S$, in other words, the probability of each element times its information
-content. In our example, the entropy of $S$ is $1.5$. When $P$ is clear, we will
-often use the shorthand of $\EntropyOf{S}$ for $\EntropyOf{S,P}$.
-
-In our setting, we already have a way of expressing sets of data:
-regular expressions.  What we need is a way to express probability
-distributions over that data.
-To that end, we adopt \emph{stochastic regular expressions}~\cite{stoch-def,stoch-rnn} (SREs), 
-which are regular expressions in which each operator is annotated with a
-probability (see \S\ref{sec:sre}).  A stochastic regular expression
-specifies both a set of strings and a probability
-distribution over that set, enabling us to calculate the
-entropy of the SRE.
-
-In the synthesis setting, there are typically an infinite number of possible
-lenses that satisfy the specification. There are many choices on which
-components of the data should be synchronized vs which should be left
-unsynchronized. We use entropy to guage the relative likelihood of such lenses
-in our synthesis algorithm (see \S\ref{subsec:lenscosts}). For any lens $\Lens$, we can calculate the expected
-number of bits required to recover a string $t$ in $\LanguageOf{\RegexAlt}$ from
-a synchronized string $s$ in $\LanguageOf{\Regex}$. This expectation is the
-\emph{conditional entropy of $\Regex$ given $\RegexAlt$ and $\Lens$}, formally
-$\sum_{s \in \Regex}\ProbabilityOf{\Regex}{s}\cdot\EntropyOf{\SetOf{t \SuchThat
+information theory, each element of $s \in S$ has \emph{information content},
+the number of bits required to specify that element in a perfect encoding scheme
+($-\Log_2P(s)$). The entropy of $S$ calculates the expected number of bits
+required to specify a random element pulled from $S$, in other words, the
+probability of each element times its information content. Entropy maintains
+that if a data source contains many possible elements, where none have
+significantly higher probability than others, it will have high information
+content. Data sources with few elements that don't have vanishingly small
+probabilities have small entropy. When $P$ is clear, we will often use the
+shorthand of $\EntropyOf{S}$ for $\EntropyOf{S,P}$.
+
+In our setting, we already have a way of expressing sets of data: regular
+expressions. To calculate entropy, what we need is a way to express probability
+distributions over that data. To that end, we adopt \emph{stochastic regular
+  expressions}~\cite{stoch-def,stoch-rnn} (SREs), which are regular expressions
+in which each operator is annotated with a probability (see \S\ref{sec:sre}). A
+stochastic regular expression specifies both a set of strings and a probability
+distribution over that set.
+
+We use entropy to guage the relative likelihood of such lenses in our synthesis
+algorithm (see \S\ref{subsec:lenscosts}). For any lens $\Lens$, we can calculate
+the expected number of bits required to recover a string $t$ in
+$\LanguageOf{\RegexAlt}$ from a synchronized string $s$ in
+$\LanguageOf{\Regex}$. This expectation is the \emph{conditional entropy of
+  $\Regex$ given $\RegexAlt$ and $\Lens$}, formally $\sum_{s \in
+  \Regex}\ProbabilityOf{\Regex}{s}\cdot\EntropyOf{\SetOf{t \SuchThat
     \Lens.\PutROf{s}{t} = t}}$. The likelihood we assign to $\Lens$ is the sum
@@
 
-To obtain SREs from the plain regular expressions that users write, we use a
-heuristic that attempts to assign probability annotations that give each
-string in the language equal probablity. Our synthesis algorithm could
-equally well take as input a pair of stochastic regular expression rather than a
-plain one, permitting users to specify a data-specific probability
-distribution (either manually calculated or inferred from some dataset of
-interest). However, we have found that our heuristic works well in practice,
-so we have not explored these options.
-
+To obtain SREs from the plain regular expressions that users write, by default
+we use a heuristic that attempts to assign probability annotations that give
+each string in the language equal probablity. In other words, this heuristic
+tries to make each bit of data no more or less likely than any other.
+Our synthesis algorithm could equally well take as input a pair of stochastic
+regular expression rather than a plain one, permitting users to specify a
+data-specific probability distribution (either manually calculated or inferred
+from some dataset of interest).
+
+However, sometimes users already know that certain data should or should not be
+used to construct the other format. We introduce relevance annotations to SREs
+that enable users to specify whether a piece of data should be used to construct
+the other format (with \SRequire) or not (with \Skip). In our example,
+\lstinline{salary} and \lstinline{company} could safely be skipped (as they are
+disconnected), where \lstinline{name} in \lstinline{emp_salary} and
+\lstinline{emp_ins} be annotated as required (as they are converted with the
+identity lens).
+
+\subsection{Searching for Likely Lenses}
 Now that we have a way to calculate likelihood, we need a way to search for
@@
 pair of stochastic regular expressions $\Regex$ and $\RegexAlt$ by performing a
-type-directed search. In more detail, \GreedySynth first normalizes the
-stochastic regular expressions provided by \Expand into \emph{stochastic DNF
-  regular expressions}---a constrained representation of stochastic regular
-expressions with disjunctions distributed over concatenations and with
-concatenations and disjunctions normalized to operating over lists (see \S\ref{subsec:sdnfre}).
+type-directed search. In more detail, \GreedySynth first converts the stochastic
+regular expressions provided by \Expand into \emph{stochastic DNF regular
+  expressions}---a constrained representation of stochastic regular expressions
+with disjunctions distributed over concatenations and with concatenations and
+disjunctions normalized to operating over lists (see \S\ref{subsec:sdnfre}).
 \GreedySynth then uses the syntax of these n-ary DNF regular expressions to find
@@
 These lenses include neither a composition operator nor a type equivalence rule.
-These restrictions mean that each simple symmetric n-ary DNF lens is typed by a
-single pair of stochastic n-ary DNF regular expressions, hence \GreedySynth's
-search space for a given pair of regular expressions is finite. The variations
-that \GreedySynth considers include choosing between \IdentityLens{} and
-\Disconnect lenses and deciding how to match multiple occurrences of the same
-kind of data, \EG, the two occurrences of \lstinline{name} in the management and
-HR data formats (should the algorithm match the first instance of
-\lstinline{name} on the left with the first instance of \lstinline{name} on the
-right, or to the second). Finally, \GreedySynth yields a simple symmetric lens
-by converting the n-ary forms into the binary forms provided in the surface
+These restrictions mean that there are only a few simple symmetric n-ary DNF
+lenses that are well-typed between each pair of stochastic n-ary DNF regular
+expressions, hence \GreedySynth's search space for a given pair of regular
+expressions is finite. Finally, \GreedySynth yields a simple symmetric lens by
+converting the n-ary forms into the binary forms provided in the surface
 language.
-
-To illustrate the partnership between \Expand and \GreedySynth, consider
-synthesizing the lens shown in Figure~\ref{fig:example_lens}. When \GreedySynth
-initially tries to find a lens between \lstinline{emp_salaries} and
-\lstinline{emp_insurance}, it hits some snags: there are complex disjunctions
-and concatenations in \lstinline{emp_salaries} outside the iteration, but no
-such complexity in \lstinline{emp_insurance}. Because of this disparity between
-the formats, \GreedySynth finds an unlikely lens (involving undesired
-\KW{merge}s and \KW{disconnect}s), and the search continues. However, after
-\Expand applies a star-semiring axiom (presented in Figure~\ref{fig:ssr-sre}) to
-rewrite the regular expression \lstinline{("\n" . emp_salary)$^*$} within
-\lstinline{emp_salaries} into
-\lstinline{"" | (("\n" . emp_salary) . ("\n" . emp_salary)$^*$)}, the
-subcomponents line up better, and the more likely correct lens is instead found.
 
@@
 regular expressions, but a different pair we haven't yet discovered may give
-rise to an even better lens.
-
-The search algorithm must resolve a tension among the quality of
-the inferred lens and the amount of time it takes to return a
+rise to an even better lens. The search algorithm must resolve a tension between
+the quality of the inferred lens and the amount of time it takes to return a
 result. For example, if the algorithm has already found the lens in
-Figure~\ref{fig:example_lens}, we don't want to spend a lot of time
-searching for an even better lens.
-
-To resolve this tension, the algorithm uses heuristics to judge whether to
-return the current best satisfying lens to the user or to pass the next set of
-unrollings to \GreedySynth. The heuristics favor stopping if the current best
-satisfying lens is very likely, indicating the lens is very promising (for
-example, if the satisfying lens loses no information, the algorithm should
-terminate for no other lens will be more likely). The heuristics also favor
-stopping if \Expand has delivered to \GreedySynth all pairs of stochastic
-regular expressions at a given rank and there is a large number of pairs at the
-next rank, because searching through all such pairs will take a long time. With
-this approach, the algorithm can quickly respond with a satisfying lens with
-relatively high likelihood. If the user is unhappy with the result, they can
-either refine the search by supplying additional examples, which serve both to
-rule out the previously proposed lens(es) and to reduce the size of the search
-space by cutting down on the number of satisfying lenses, or they can supply
-annotations on the source and target SREs indicating that certain information
-either must be retained or must be discarded by the lens (see
-\S\ref{subsec:optimizations}).
+Figure~\ref{fig:example_lens}, we don't want to spend a lot of time searching
+for an even better lens. To resolve this tension, the algorithm uses heuristics
+to judge whether to return the current best satisfying lens to the user or to
+pass the next set of unrollings to \GreedySynth. The heuristics favor stopping
+if the current best satisfying lens is very likely, indicating the lens is very
+promising (for example, if the satisfying lens loses no information, the
+algorithm should terminate for no other lens will be more likely). The
+heuristics also favor stopping if \Expand has delivered to \GreedySynth all
+pairs of stochastic regular expressions at a given rank and there is a large
+number of pairs at the next rank, because searching through all such pairs will
+take a long time. With this approach, the algorithm can quickly respond with a
+satisfying lens with relatively high likelihood. If the user is unhappy with the
+result, they can either refine the search by supplying additional examples,
+which serve both to rule out the previously proposed lens(es) and to reduce the
+size of the search space by cutting down on the number of satisfying lenses, or
+they can supply annotations on the source and target SREs indicating that
+certain information either must be retained or must be discarded by the lens
+(see \S\ref{subsec:optimizations}).
 

### 2019-02-22 report/final-report.tex
added some experimental results

@@
 
-Reffing a \Cref{fig:eg}.
+% From bijective lens paper
+
+% \newcommand{\SOptician}{Optician\textsubscript{S}}
+% \newcommand{\SynthSymLens}{\PCF{SynthSymLens}\xspace}
+% \newcommand{\RXSearchState}{\ensuremath{\mathit{pq}}}
+% \newcommand{\ToStochastic}{\PCF{ToStochastic}\xspace}
+% \newcommand{\LC}{\ensuremath{\mathit{lc}}}
+% \newcommand{\Best}{\ensuremath{\mathit{best}}}
+% \newcommand{\RXSearch}{\PCF{RXSearch}\xspace}
+% \newcommand{\Continue}{\PCF{Continue}\xspace}
+% \newcommand{\PQ}{\PCF{PQ}\xspace}
+% \newcommand{\SynthDNFLens}{\PCF{SynthDNFLens}}
+% \newcommand{\ToLens}{\ensuremath{\Uparrow}}
+% \newcommand{\ToLensOf}[1]{\ensuremath{\ToLens{}\mkern-4mu #1}}
+% \newcommand{\ToDNFRegexText}{\PCF{ToDNFRegex}}
+% \newcommand{\Beautify}{\PCF{Beautify}}
+% \newcommand{\RigidSynth}{\PCF{RigidSynth}}
+% \newcommand{\GreedySynth}{\PCF{GreedySynth}\xspace}
+% \newcommand{\RigidSynthInternal}{\PCF{RigidSynthInternal}}
+% \newcommand{\RigidSynthSequence}{\PCF{RigidSynthSeq}}
+% \newcommand{\RigidSynthAtom}{\PCF{RigidSynthAtom}}
+% \newcommand{\GetDNFNormalizer}{\PCF{GetDNFNormalizer}}
+% \newcommand{\CreatePQueue}{\PCF{CreatePQueue}}
+% \newcommand{\GetTransitiveSet}{\PCF{GetTransitiveSet}}
+% \newcommand{\GetCurrentSet}{\PCF{GetCurrentSet}}
+% \newcommand{\Pop}{\PCF{Pop}}
+% \newcommand{\ExpandOnce}{\PCF{ExpandOnce}}
+% \newcommand{\ExpandRequired}{\PCF{ExpandRequired}}
+% \newcommand{\FixProblemElts}{\PCF{FixProblemElts}}
+% \newcommand{\Expand}{\PCF{Expand}\xspace}
+% \newcommand{\ForceExpand}{\PCF{ForceExpand}}
+% \newcommand{\Reveal}{\PCF{Reveal}}
+% \newcommand{\Map}{\PCF{Map}}
+% \newcommand{\EnqueueMany}{\PCF{EnqueueMany}}
+% \newcommand{\ReturnVal}[1]{\ensuremath{\Return\,#1}}
+% \newcommand{\CurrentSet}{\ensuremath{\mathit{CS}}}
+% \newcommand{\TransitiveSet}{\ensuremath{\mathit{TS}}}
+
+% \newcommand{\SSOpt}{\ensuremath{\mathbf{SS}}}
+% \newcommand{\SSNCOpt}{\ensuremath{\mathbf{SSNC}}}
+% \newcommand{\BSOpt}{\ensuremath{\mathbf{BS}}}
+% \newcommand{\BSNCOpt}{\ensuremath{\mathbf{BSNC}}}
+% \newcommand{\AnyOpt}{\ensuremath{\mathbf{Any}}}
+% \newcommand{\FLOpt}{\ensuremath{\mathbf{FL}}}
+% \newcommand{\CCOpt}{\ensuremath{\mathbf{DC}}}
+% \newcommand{\NSOpt}{\textbf{NS}}
+% \newcommand{\NROpt}{\textbf{NR}}
+
+We have implemented our synthesis algorithms in
+of OCaml and integrated them into Boomerang system,
+where they are available as open source code~\cite{GitHub}.
+Users of Boomerang can write lenses by hand, specify them as
+synthesis tasks, or do a combination of both.
+
+We have analyzed our algorithms theoretically, proving various
+soundness and completeness results.  We have also analyzed our
+algorithms empirically, studying their impact on a range of real and
+synthetic benchmarks.  Our full results have been published in a
+series of conference papers and technical
+reports~\cite{bijective,quotient,symmetric}.  In the rest of this
+section, we summarize some of the key results.
+
+\subsection{Bijective Synthesis Algorithms}
+
+We first study the effectiveness of the pure bijective synthesis
+algorithm by evaluating its performance on a set of 39
+benchmark programs.
+
+All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
+of 1600 MHz DDR3 running macOS Sierra.
+
+\paragraph*{Benchmark Suite Construction}
+We constructed our benchmarks primarily by adapting examples from
+Augeas~\cite{augeas} and 
+Flash Fill~\cite{gulwani-popl-2014}.
+
+Augeas is a configuration editing system for Linux that uses lens
+combinators similar to those in Boomerang. However, it transforms
+strings on the left to structured trees on the right rather than
+transforming strings to strings.
+We adapted these Augeas lenses to our setting by converting the
+right-hand sides to strings that correspond to serialized versions
+of the tree formats.  
+We derived 29 of the benchmark tests by
+adapting the first 27 lenses in alphabetical order, as well as the lenses
+\CF{aug/xml-firstlevel} and \CF{aug/xml} that were referenced
+by the `A' lenses.
+Furthermore, the 12 last synthesis problems derived
+from Augeas were tested after \Optician{} was
+finalized, demonstrating that the optimizations were not
+overtuned to perform well on the testing data.
+
+Flash Fill is a system that allows users to specify common string
+transformations by example~\cite{gulwani-popl-2014}.  
+We derived three benchmarks from the first few examples in the
+paper and one from the running example on
+extracting phone numbers.
+
+Both Augeas and Flash Fill permit non-bijective transformations.
+To test our system on these benchmarks, we had to convert them into
+bijective transformations by hand.  More specifically, in many of the
+benchmarks, we normalized
+the whitespace in the documents so the amount of whitespace was equal
+in the source and target.  In a few of the benchmarks, either the
+source or the target was missing information present in the other
+format.  We modified those benchmarks by adding the missing
+information back into the file.
+
+Finally, we added custom examples to highlight weaknesses of
+our algorithm (\CF{cap-prob} and \CF{2-cap-prob}) 
+and to test situations for which we thought the tool would be
+particularly useful (\CF{workitem-probs}, \CF{date-probs}, \CF{bib-prob},
+and \CF{addr-probs}).   These examples convert between work item formats, date
+formats, bibliography formats, and address formats, respectively.
 
 \begin{figure}
-  a picture is here
-  \caption{An example figure}
-  \label{fig:eg}
+  \centering
+  \begin{subfigure}[b]{.49\textwidth}
+    \centering
+    \includegraphics{figs/specsizes}
+    \caption{}
+    \label{subfig:lenssize}
+  \end{subfigure}
+  \begin{subfigure}[b]{.49\textwidth}
+    \includegraphics{figs/examplesused}
+    \caption{}
+    \label{subfig:examplesused}
+  \end{subfigure}
+  \caption{Sizes of Specifications.
+    In (a), we show how many benchmarks are defined in our suite using a
+    given number of AST nodes or fewer.
+    In (b), we show how many benchmarks are defined in our suite using a
+    given number of examples or fewer.}
+  \label{fig:definition-sizes}
 \end{figure}
+
+Figure~\ref{fig:definition-sizes} shows the complexity of our type-based
+specifications as well as our example counts.
+An average benchmark has a type-based specification that can be represented using
+310 AST node, and requires 1.1 input/output examples.
+Our benchmarks vary from simple problems, like changing the
+representation of dates
+(with a specification size of 85, and a generated lens size of 79), to 
+complex tasks, like transforming configuration files for server monitoring
+software into dictionary form (with a specification size of 670 and
+a generated lens size of 651).  On average, the size of the generated lens is
+89\% the size of its type specifications.
+
+\paragraph*{Importance of Examples}
+
+\begin{figure}
+  \centering
+  \includegraphics{figs/examples.eps}
+  \caption{Average number of random examples required to synthesize benchmark
+    programs.  {\bf Experimental Average} is the average number of randomly
+    generated examples needed to correctly synthesize the lens.  {\bf
+      Determinize Permutations} is the theoretical number of examples required
+    to determinize the choice all the permutations in \RigidSynth{}.
+    In practice, far fewer examples are
+    needed to synthesize the correct lens than would be predicted by the number
+    required to determinize permutations.}
+  \label{fig:exs-reqd}
+\end{figure}
+
+To evaluate how many user-supplied examples the algorithm requires in
+practice, we \textit{randomly} generated appropriate source/target
+pairs, mimicking what a na\"{i}ve user might do.  We did not write the
+examples by hand out of concern that our knowledge of the synthesis
+algorithm might bias the selection. Figure~\ref{fig:exs-reqd} shows
+the number of randomly generated examples it takes to synthesize the
+correct lens averaged over ten runs.  The synthesis algorithm almost never needs
+any examples: only 5 benchmarks need a nonzero number of examples to
+synthesize the correct lens and only one, \CF{cust/workitem-probs} required over
+10 randomly generated examples.
+A clever user may be able to reduce the
+number of examples further by selecting examples carefully; we
+synthesized \CF{cust/workitem-probs} with only 8 examples.
+
+These numbers are low because there are relatively few well-typed
+bijective lenses between any two source and target regular expressions. 
+As one would expect, the benchmarks where there are multiple ways to
+map source data to the target (and vice versa) require the most examples.
+For example, the benchmark \CF{cust/workitem-probs} requires a large number of
+examples because it
+must differentiate between data in different text fields in both the
+source and target and map between them appropriately.  As these text fields are
+heavily permuted
+(the legacy format ordered fields by a numeric ID, where
+the modern format ordered fields alphabetically) and fields can be
+omitted, a number of examples are needed to correctly identify the mapping
+between fields.
+
+The average number of examples to
+infer the correct lens does not tell the whole story.  The system will
+stop as soon as it finds a well typed lens that satisfies the supplied examples.
+This inferred lens may or may not 
+correctly handle unseen examples that correspond to
+unexercised portions of the source and target regular expressions.
+Figure~\ref{fig:exs-reqd} lists
+the number of examples that are required to determinize the generation of
+permutations in \RigidSynth{}.
+Intuitively, this number represents the maximum number of
+examples that a user must supply to guide the synthesis engine if it
+always guesses the wrong permutation when multiple permutations can be used to
+satisfy the specification.
+
+The average number of examples is so much lower than the maximum
+number of required examples because of correspondences in how we wrote
+the regular expressions for the source and target data formats. 
+Specifically, when we had corresponding disjunctions in both the
+source and the target, we ordered them the same way.  The algorithm
+uses the supplied ordering to guide its search, and so the system
+requires fewer examples.   We did not write the examples in this style
+to facilitate synthesis, but rather because maintaining similar
+subparts in similar orderings makes the types much easier to 
+read. We expect that most users would do the same.
+
+\paragraph*{Comparison Against Other Tools}
+%
+We are the first tool to synthesize bidirectional transformations between data
+formats, so there is no tool to which we can make an apple-to-apples comparison.
+Instead, we compare against tools for generating unidirectional
+transformations instead. 
+Figure~\ref{fig:synthesis-times} includes a comparison against two other
+well-known tools that synthesize
+text transformation and extraction functions from examples -- Flash Fill and FlashExtract.  For this
+evaluation, we used the version of these tools distributed through the
+PROSE project~\cite{prose}.
+
+\begin{figure}
+  \includegraphics{figs/times}
+  \caption{
+    Number of benchmarks that can be solved by a given algorithm in a given
+    amount of time.
+    \Optician{} is our bijective synthesis algorithm including a set
+    of optimizations we have implemented.
+    \FlashExtractMode{} is the existing FlashExtract system.  \FlashFillMode{} is
+    the existing Flash Fill system.  \NaiveMode{} is na\"{i}ve type-directed 
+    synthesis on the bijective lens combinators.  Our synthesis algorithm performs
+    better than the na\"{i}ve approach and other string transformation systems,
+    and our optimizations speed up the algorithm enough that all tasks become
+    solvable.
+  }
+  \label{fig:synthesis-times}
+\end{figure}
+
+To generate specifications for Flash Fill, we generated input/output
+specifications by generating random elements of the source language, and
+running the lens on those elements to generate elements of the target language.
+These were then fed to Flash Fill.
+
+To generate specifications for FlashExtract, we extracted portions of strings
+mapped in the generated lens either through an identity
+transformation or through a previously synthesized lens, whereas strings that were
+mapped through use of \ConstLens{} were considered boilerplate and so not
+extracted.
+
+As these tools were designed for a broader audience, they put less of a burden
+on the user.  These tools only use input/output examples (for Flash
+Fill), or marked text regions (for FlashExtract), as opposed
+to \Optician{}'s use of regular expressions to constrain the format of
+the input and output.  By using regular expressions,
+\Optician{} is able to synthesize significantly more programs
+than either existing tool.
+
+Flash Fill and FlashExtract have two tasks: to determine how
+the data is transformed, they must also infer the structure of the data, a
+difficult job for complex formats.
+In particular, neither Flash Fill nor FlashExtract was able to synthesize
+transformations or extractions present under two iterations, a type of format
+that is notoriously hard to infer.
+These types of dual iterations are pervasive in Linux configuration
+files, making Flash Fill and FlashExtract ill suited for many of the synthesis
+tasks present in our test suite.
+
+Furthermore, as unidirectional transformations, Flash Fill and FlashExtract have
+a more expressive calculus.  To guarantee bidirectionality, our syntax must be
+highly restrictive, providing a smaller search space to traverse.
 

### 2019-02-22 report/final-report.tex
experiments

@@
 Flash Fill~\cite{gulwani-popl-2014}.
-
+%
 Augeas is a configuration editing system for Linux that uses lens
@@
 overtuned to perform well on the testing data.
-
+%
 Flash Fill is a system that allows users to specify common string
@@
 
+\subsection{Quotient Synthesis Algorithms}
+
+% DPW:  From Quotient lenses paper
+
+\newcommand{\wf}[1]{\ensuremath{#1\;\mathsf{wf}}}
+
+% FOR Regular Expression names
+\newcommand{\re}[1]{\ensuremath{\mathtt{#1}}}
+\newcommand{\codefont}[1]{\ensuremath{\mathsf{#1}}}
+\newcommand{\kw}[1]{\textcolor{dkblue}{\ensuremath{\mathsf{#1}}}}
+\newcommand{\collapse}[2]{\ensuremath{\kw{collapse} \; #1 \mapsto #2}}
+\newcommand{\squash}[3]{\ensuremath{\kw{squash} \; #1 \rightarrow #2\; \kw{using} \; #3}}
+\newcommand{\perm}[2]{\ensuremath{\kw{perm}(#1)\; \kw{with}\; #2}}
+\newcommand{\normalize}[3]{\ensuremath{\kw{normalize}(#1, #2, #3)}}
+\newcommand{\eqrel}[1]{\ensuremath{\equiv_{#1}}}
+
+\newcommand{\canonize}{\ensuremath{\kw{canonize}}}
+
+\newcommand{\Name}{Optometrist\xspace}
+
+\newcommand{\QRESize}{\textbf{QS}}
+\newcommand{\canonizeAndSpecSize}{\textbf{BS}}
+\newcommand{\LensAndSpecSize}{\textbf{NS}}
+
+%\newcommand{\QOpt}{Optician_Q}
+\newcommand{\QOpt}{QRE-enhanced Optician}
+\newcommand{\OpticianRuntime}{\textbf{Optician}}
+\newcommand{\QREOptician}{\textbf{Optician\textsubscript{Q}}}
+\newcommand{\SystemOnOptician}{\textbf{QO}}
+\newcommand{\SystemOnBenchmarks}{\textbf{QQ}}
+\newcommand{\cd}[1]{\lstinline[backgroundcolor=\color{white}]$#1$}
+
+We have implemented QREs and the quotient lens synthesis algorithm described
+above as an extension to the Boomerang
+interpreter~\cite{boomerang,quotientlenses}. We have extended the existing
+Optician tool to synthesize QRE lenses from QREs. We will use 
+``\QOpt'' to denote our extended version of Optician, and just plain ``Optician'' to denote
+the pre-QRE version of Optician. The synthesis algorithm produces Boomerang lens
+values, so Boomerang gives synthesized lenses the same first-class status as hand-written ones. To
+evaluate the effectiveness of QREs, QRE lenses, and QRE lens synthesis, we
+conducted experiments to answer the following questions:
+%
+\begin{itemize}
+  \item {\bf Ease of use.} Does synthesizing QRE lenses from QREs permit an
+  easier development process than writing lenses by hand?
+  Does synthesizing QRE lenses from QREs permit an easier development
+  process than manually writing canonizers and then synthesizing lenses between
+  their canonized forms?  
+  \item {\bf Performance.} Is the synthesis algorithm/implementation fast enough to be
+  used as part of a standard development process?
+\end{itemize}
+%
+All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
+of 1600 MHz DDR3 running macOS High Sierra.
+
+\paragraph*{Benchmark Suite Construction}
+
+To evaluate our QRE implementation, we adapted 39 lens synthesis tasks
+from the benchmark suite of the original Optician system. These
+benchmarks are a combination of custom benchmarks, benchmarks derived
+from FlashFill~\cite{flashfill}, and benchmarks derived from
+Augeas~\cite{augeas2} (we also experimented using our QRE implementation to synthesize quotient lenses between XML, RDF and and JSON formats using data from the data.gov database; the data consisted of census statistics, demographic statistics, wage comparosion data, and crime index data).
+In these 39 benchmarks, 10 of
+the data formats had to be modified to work with the bijectivity
+constraints that Optician required. For instance, when one
+representation permits whitespace where the other does not, we
+modified the original version of the benchmark to allow more whitespace,
+thereby restoring bijectivity (but altering the data format). With the
+new QRE support, we were able to remove these alterations.  This
+experience alone suggests that QREs make the lens development process
+more flexible.
+
+\paragraph*{Ease of Use}
+
+\begin{figure}[t]
+\includegraphics{qfigs/asts.eps}
+\caption{AST node measurements for each of the three approaches
+on each of the 10 non-bijective benchmark problems.
+Benchmarks are sorted in order of increasing complexity as measured by
+the number of AST nodes in the source and target format descriptions. 
+QRE Synthesis requires far fewer AST nodes than the other
+two approaches.}
+%%\caption{Count of benchmark programs definable using a given AST count. We find
+%%that it takes far fewer AST nodes to define benchmark lenses using QRE
+%%synthesis than with Optician without QRE synthesis or without synthesis.}
+\label{fig:asts}
+\end{figure}
+
+To evaluate the impact of QRE lens synthsis on programmer effort,
+we focus our attention on the 10 problems in the benchmark suite that
+are not bijective and hence require non-trivial canonizers. 
+(Optician already handles the other problems with minimal programmer
+effort.)
+
+We are interested in comparing three different approaches, which vary
+in the amount of synthesis used. 
+In the first approach, which we call \QRESize{} for QRE Synthesis, the programmer uses
+QRE lens synthesis.  She must write QRE specifications of the source and
+target formats and she may give examples.
+In the second approach, which we call \canonizeAndSpecSize{} for
+Bijective Synthesis, the
+programmer uses bijective lens synthesis \`a la Optician.
+She must write canonizers by hand, along with 
+regular expressions to describe the external
+representations of the source and target formats. (The internal
+formats can be inferred from the canonizers.) She may also
+provide examples to help in the synthesis of the bijective lens.
+In the third approach, which we call \LensAndSpecSize{} for No Synthesis, the
+programmer writes the lens between the source and target formats
+entirely by hand, including the descriptions of the source and target
+formats.
+
+For each problem in the benchmark suite, we calculate the following
+measures as proxies for the level of programmer effort when using each
+the three approaches:
+
+%
+\begin{itemize}
+  \item[\QRESize{}:] 
+  The number of AST nodes in the QRE specifications for the source and
+  target formats, including examples. 
+  \item[\canonizeAndSpecSize{}:] 
+  The sum of (1) the number of AST nodes in $W(q)$ for each QRE $q$ in the source and target
+  formats, (2) the number of AST nodes in $\canonize(q)$ for each QRE $q$ with a
+  non-trivial canonizer, and (3) the number of AST nodes in the
+  examples.  We use (1) to estimate the burden of describing
+  the external source and target formats and (2) to estimate the
+  burden of writing the requisite canonizers
+  by hand.  We count the nodes in the examples because they would be
+  fed to the bijective synthesizer.  
+  These counts are an approximation, as both $W(q)$ and $\canonize(q)$ are
+  automatically generated from the corresponding QRE $q$, and it is
+  possible that a human-written version might be smaller.
+  \item[\LensAndSpecSize{}:] The sum of (1) the number of AST nodes in
+  $W(q)$ for each QRE $q$ in the source and target formats and (2) the
+  number of AST nodes in the synthesized QRE lens.  We use (1) to
+  estimate the burdern of describing the source and target formats
+  and (2) to estimate the burdern of writing the appropriate lens by
+  hand. These counts are also approximations, as
+  $W(q)$ and the synthesized lens may be larger than one written by hand.
+\end{itemize}
+
+Figure~\ref{fig:asts} shows each of these measures for the 10
+non-bijective problems in the benchmark suite.  On
+average (using a geometric mean), \canonizeAndSpecSize{} used~38.5\% more AST nodes
+than \QRESize{}, requiring an average of~214 more AST nodes. On 
+average, \LensAndSpecSize{} used~180\% more AST nodes than \QRESize{}, requiring an
+average of~998 more AST nodes. These figures suggest that introducing QREs saves
+programmers significant effort compared to both Optician and basic
+Boomerang.
+
+\paragraph*{Maintaining Competitive Performance}
+
+\begin{figure}[t]
+\centering
+\begin{subfigure}[b]{.49\textwidth}
+\centering
+\includegraphics{qfigs/times_opt}
+\caption{}
+\label{subfig:lenssize}
+\end{subfigure}
+\begin{subfigure}[b]{.49\textwidth}
+\includegraphics{qfigs/times_new.eps}
+\caption{}
+\label{subfig:examplesused}
+\end{subfigure}
+\caption{Runtimes measurements. In (a), we run Optician and \QOpt{}
+  on the Optician benchmarks.  We find that there is only a negligible
+  performance overhead incurred by using QREs. 
+  In (b), we run \QOpt{} on the 10 Optician benchmarks
+  previously edited to make them bijective, after removing those edits
+  and then extending the synthesis specification to include QREs.
+  (In other words, we restored them to their original state, added QREs,
+  and then ran \QOpt{}). 
+  We find that \QOpt{} is able to
+  synthesize all quotient lenses in under 10 seconds, and typically finishes in
+  under 5 seconds.}
+\label{fig:times}
+\end{figure}
+
+To assess the performance of QRE synthesis, we are interested in two different
+questions. First, how does the performance of \QOpt{} compare to the performance
+of Optician on benchmarks that do not require QREs? The answer to this question
+tells us how much overhead we have introduced by adopting the more general
+mechanism. Figure~\ref{fig:times}(a) shows that \QOpt{} was able to synthesize
+all of the Optician benchmarks at a speed competitive with the old version.
+There is a small amount of additional overhead introduced by QREs in calculating
+the $W$ and $K$ functions, resulting in a slight decrease in performance.
+
+Second, how much time does it take for \QOpt{} to synthesize a QRE
+lens when running on a non-bijective benchmark problem?  
+Figure~\ref{fig:times}(b) shows the amount of time required to infer a
+lens for each of the 10 benchmark programs with nontrivial quotients.  
+We find that \QOpt{} is able to synthesize all quotient lenses in
+under 10~seconds, and typically finishes in under 5~seconds.
+
 \section{CONCLUSIONS}

### 2019-02-22 report/final-report.tex
stuff

@@
 
-\section{SUMMARY}
+\section{Summary}
 
-\section{INTRODUCTION}
+\section{Introduction}
 
@@
 Despite their limitations, bijections and bijective synthesis form a
-useful foundation on top which more general transformations between
+useful foundation on top of which more general transformations between
 data sets may be built.  Hence, in the second half of our of project,
@@
 by bijections (using the original bijective synthesis algorithm).  The
-so-called \emph{quotient lenses}~\cite{Foster:quotient-lenses}we generate are capable of handling
+so-called \emph{quotient lenses}~\cite{quotientlenses} we generate are capable of handling
 ``irrelevant'' differences between structures such as white space or
@@
 transformations our system can define significantly.  Finally, we explored the
-synthesis of \emph{symmetric lenses}~\cite{hofmann+:lenses} , again building upon the
+synthesis of \emph{symmetric lenses}~\cite{symmetric-lenses} , again building upon the
 bijective platform that we started with.  Symmetric lenses are able to
@@
 
-We measured the effectiveness of our algorithms on a set of
-benchmarks drawn from the Augeas system~\cite{augeas}---a system for
-transforming and editing Linux configuration files.  Past synthesis
+We measured the effectiveness of our algorithms on a set of 39
+benchmarks drawn from the Augeas system~\cite{augeas},  system for
+transforming and editing Linux configuration files, and
+FlashFill~\cite{flashfill}, an extension to Microsoft Excel for
+synthesizing columns of spreadsheets.  Past synthesis
 tools, such as FlashFill~\cite{flashfill} were unable to translate
 most Augeas file formats.  However, our synthesis toolkit was able to
-transform all 40 of the Augeas file formats that we analyzed.
+generate transforms for all 39 of the Augeas file formats that we analyzed.
 
-We have written three academic papers~\cite{?}
-describing and evaluating our algorithms and produced open-source our code that
-implements those algorithms.
+We have written
+papers~\cite{bijective-synthesis,quotient-synthesis,symmetric-synthesis}
+describing and evaluating our algorithms and produced open-source 
+code~\cite{GitHub} that implements those algorithms.  In this report,
+we summarize some of our key methods, results and findings, focusing
+on the core algorithms for bijective synthesis and quotient synthesis.
 
-\section{METHODS, ASSUMPTIONS, AND PROCEDURES}
+\section{Methods, Assumptions and Procedures}
 
@@
 %
+bidirectional variants of 
+relational algebra~\cite{BohannonPierceVaughan},
+%
 BiGUL \cite{DBLP:conf/pepm/KoZH16}, 
 %
-bidirectional variants of 
-relational algebra~\cite{BohannonPierceVaughan},
 spreadsheet formulas~\cite{DBLP:conf/vl/MacedoPSC14},
@@
 instances.  The equivalence relations $\equiv_S$ and $\equiv_T$ are also given
-as inputs to the synthesis algorithm.  As we describe below, one result of our
-work is a language of \textit{quotient regular expressions} that concisely
-specifies $S$ and $\equiv_S$ simultaneously.
+as inputs to the synthesis algorithm.  Definition of such equivalences
+are given via a new language of \textit{quotient regular expressions}
+that we have defined.  Quotient regular expressions
+describe both  $S$ (or $T$) and $\equiv_S$ (or $\equiv_T$)
+simulataneously in one compact notation.
 
-\section{RESULTS AND DISCUSSION}
+\section{Results and Discussion}
 
@@
 series of conference papers and technical
-reports~\cite{bijective,quotient,symmetric}.  In the rest of this
+reports~\cite{bijective-synthesis,quotient-synthesis,symmetric-synthesis}.  In the rest of this
 section, we summarize some of the key results.
@@
 benchmark programs.  
-
+%
 All evaluations were performed on a 2.5 GHz Intel Core i7 processor with 16 GB
@@
 
-We have implemented QREs and the quotient lens synthesis algorithm described
-above as an extension to the Boomerang
-interpreter~\cite{boomerang,quotientlenses}. We have extended the existing
-Optician tool to synthesize QRE lenses from QREs. We will use 
+We have extended the bijective synthesis tool to synthesize quotient lenses. We will use 
 ``\QOpt'' to denote our extended version of Optician, and just plain ``Optician'' to denote
-the pre-QRE version of Optician. The synthesis algorithm produces Boomerang lens
-values, so Boomerang gives synthesized lenses the same first-class status as hand-written ones. To
-evaluate the effectiveness of QREs, QRE lenses, and QRE lens synthesis, we
-conducted experiments to answer the following questions:
-%
-\begin{itemize}
-  \item {\bf Ease of use.} Does synthesizing QRE lenses from QREs permit an
-  easier development process than writing lenses by hand?
-  Does synthesizing QRE lenses from QREs permit an easier development
-  process than manually writing canonizers and then synthesizing lenses between
-  their canonized forms?  
-  \item {\bf Performance.} Is the synthesis algorithm/implementation fast enough to be
-  used as part of a standard development process?
-\end{itemize}
+the pre-quotient version of Optician. The synthesis algorithm produces Boomerang lens
+values, so Boomerang gives synthesized lenses the same first-class status as hand-written ones. 
 %
@@
 
-To evaluate our QRE implementation, we adapted 39 lens synthesis tasks
-from the benchmark suite of the original Optician system. These
-benchmarks are a combination of custom benchmarks, benchmarks derived
-from FlashFill~\cite{flashfill}, and benchmarks derived from
-Augeas~\cite{augeas2} (we also experimented using our QRE implementation to synthesize quotient lenses between XML, RDF and and JSON formats using data from the data.gov database; the data consisted of census statistics, demographic statistics, wage comparosion data, and crime index data).
-In these 39 benchmarks, 10 of
-the data formats had to be modified to work with the bijectivity
-constraints that Optician required. For instance, when one
-representation permits whitespace where the other does not, we
-modified the original version of the benchmark to allow more whitespace,
-thereby restoring bijectivity (but altering the data format). With the
-new QRE support, we were able to remove these alterations.  This
-experience alone suggests that QREs make the lens development process
-more flexible.
+We analyzed the same 39 lens synthesis tasks from the original
+Optician system.
+We also experimented using our tool to synthesize quotient lenses
+between XML, RDF and and JSON formats using data from the data.gov
+database; the data consisted of census statistics, demographic
+statistics, wage comparosion data, and crime index data).
+Recall that to use bijective synthesis,  we had to modify 10 of 39
+benchmarks make them bijective (for instance, by normalizing whitespace)
+Because quotient lenses are more expressive, we were able avoid such
+modifications. This
+experience alone points to the benefits that quotient lens synthesis
+brings to the table.
 
-\paragraph*{Ease of Use}
+\paragraph*{Evaluating programmer effort}
 
@@
 in the amount of synthesis used. 
-In the first approach, which we call \QRESize{} for QRE Synthesis, the programmer uses
-QRE lens synthesis.  She must write QRE specifications of the source and
+In the first approach, which we call \QRESize{}, the programmer uses
+quotient lens synthesis.  She must write QRE specifications of the source and
 target formats and she may give examples.
@@
 
-\paragraph*{Maintaining Competitive Performance}
+\paragraph*{Quotient Lens Performance}
 
@@
 
-To assess the performance of QRE synthesis, we are interested in two different
+To assess the performance of quotient synthesis, we are interested in two different
 questions. First, how does the performance of \QOpt{} compare to the performance
-of Optician on benchmarks that do not require QREs? The answer to this question
+of Optician on benchmarks that do not require quotients? The answer to this question
 tells us how much overhead we have introduced by adopting the more general
@@
 There is a small amount of additional overhead introduced by QREs in calculating
-the $W$ and $K$ functions, resulting in a slight decrease in performance.
+equivalences, resulting in a slight decrease in performance.
 
-Second, how much time does it take for \QOpt{} to synthesize a QRE
+Second, how much time does it take for \QOpt{} to synthesize a quotient
 lens when running on a non-bijective benchmark problem?  
@@
 
-\section{CONCLUSIONS}
+\section{Conclusions}
 

### 2019-02-25 papers/SymmetricLenses/main.tex
minor typo
add minor fixes

@@
 \noindent
-In today's data-rich world, similar information is often
-stored in different places and in different formats. For instance, electronic
-calendars come in iCalendar, vCalendar, and h-event
-formats~\cite{calendar-formats}; US taxation data can be transmitted via Tax
-XML (used by the US government for e-filing) or TXF (used by
-TurboTax)~\cite{tax-formats}; and sewing machines can use embroidery files
-in formats with names like PES, DST, and several
-others~\cite{embroidery-formats}.
-%
-\bcp{I'm not satisfied with the embroidery example: It doesn't obviously
-  correspond to a bidirectional situation (the sewing machine does not need
-  to send back the embroidery file to anybody).  The tax example could be
-  criticized on the same grounds.}
-\afm{I still think that there are situations where the embroidery file makes
-  sense (I'm sharing my design with a friend, who then tweaks it and sends it
-  back. There may also need to file-specific information, where one contains
-  parameters like how fast to embroider and some doesn't, to be honest I don't
-  know much about embroidery though. Thinking about other examples, however.)}
+In today's data-rich world, similar information is often stored in different
+places and in different formats. For instance, electronic calendars come in
+iCalendar, vCalendar, and h-event formats~\cite{calendar-formats}; US taxation
+data can be transmitted via Tax XML (used by the US government for e-filing) or
+TXF (used by TurboTax)~\cite{tax-formats}; configuration files can be stored in
+ad-hoc or structured file formats~\cite{augeas} (for example, scheduled commands
+can be stored in a crontab or as Windows's Scheduled Tasks xml).
 
@@
 How can we identify such a ``likely'' lens? We propose the following heuristic:
-A satisfying lens is ``more likely'' if it u;ses more data from one format to
+A satisfying lens is ``more likely'' if it uses more data from one format to
 construct the other. One manifestation of this principle involves choosing the
@@
 
-All three sources provide realistic problems suitable for
-answering the research questions: Flash Fill data cleaning tasks are either
-derived from online help forums or taken from the Excel product team; Augeas is
-a utility that transforms Linux configuration files into a manageable dictionary
-form; and many of the remaining benchmarks arise from synchronizing real-world
-data formats such as BibTeX and EndNote.  Each task consists of a source and
-target regular expression.\saz{Do we want to say anything about the sizes of these?}
+All three sources provide realistic problems suitable for answering the research
+questions: Flash Fill data cleaning tasks are either derived from online help
+forums or taken from the Excel product team; Augeas is a utility that transforms
+Linux configuration files into a manageable dictionary form; and many of the
+remaining benchmarks arise from synchronizing real-world data formats such as
+BibTeX and EndNote. Each task consists of a source and target regular
+expression. These tasks can be quite quite large, as we are synthesizing complex
+transformations between real file formats -- some of the lenses and
+specifications are over 500 AST nodes.
 

### 2019-02-26 report/final-report.tex
fixed Figure 1 text
drafted summary section

@@
 \section{Summary}
+
+Maintaining software systems that manage evolving representations of
+data is tedious and error prone.  In this study, we have developed
+algorithms for automatically synthesizing \emph{lenses}, which are
+software adaptors that convert between two different representations
+of the same information.  Lenses are a special form of program that
+guarantee so-called \emph{round-tripping properties} that ensure the
+transformations are well-behaved.  Our synthesis algorithms take as
+input (i) the \emph{type} of each source as a regular expression, as
+well as (ii) a small collection of representative \emph{examples} of
+the desired translation.  They produceas output \emph{bijective string
+  lenses}.  Bijective string lenses are a special class of lenses that
+only exist when the two data sources are related via a bijection.
+Despite this limitation, bijections and bijective synthesis form a
+useful foundation on top of which more general transformations between
+data sets may be built.  We explore two of these extensions: quotient
+lenses and symmetric lenses.  Quotient lenses apply when two formats
+are ``morally'' but not technically bijective.  Examples include two
+formats that differ in their use of white space or in the order of
+sequences of elements.  Symmetric lenses allow one side or the other to have
+information not present in the other.  When converting between the
+formats, defaults take the place of missing information.
+
+We measured the effectiveness of our
+algorithms on a set of 39 benchmarks drawn from Augeas~\cite{augeas},
+which is a system for transforming and editing Linux configuration
+files, and from FlashFill~\cite{flashfill}, which is an extension to
+Microsoft Excel for synthesizing columns of spreadsheets.
+On these benchmarks, we find it is possible to synthesize lenses of
+comparable quality to hand-written lenses from format descriptions and
+a very small number of examples, leveraging 
+the large amount of information embedded in the format descriptions. 
+We further found that as the formats diverge and the class of lenses
+between them becomes richer, the synthesis problem becomes more
+difficult and additional heuristics need to be considered.
+
+We have written
+papers~\cite{bijective-synthesis,quotient-synthesis,symmetric-synthesis}
+describing and evaluating our algorithms. We produced open-source 
+code~\cite{GitHub} that implements those algorithms.  In this report,
+we summarize our key methods, results, and findings, focusing
+on the core algorithms for bijective synthesis and quotient synthesis.
 
@@
 
-    \node[text width=2.3cm,minimum height=.6cm,align=center,draw,rectangle] [right = 1.45cm of todnfregex.east] (synthdnflens) {\SynthDNFLens{}};
+    \node[text width=3.1cm,minimum height=.6cm,align=center,draw,rectangle] [right = 1.45cm of todnfregex.east] (synthdnflens) {\SynthDNFLens{}};
     \node[align=center] [above = .7cm of synthdnflens] (optician) {\Optician{}};

### 2019-02-26 report/final-report.tex
minor edit to conclusion

@@
 
-During the course of this project, we have designed, analyzed and implemented algorithms that demonstrate it is possible to synthesize three classes of bidirectional transformations:  (1) pure bijective transformations, (2) bijections modulo equivalences classes and (3) bijections modulo projections.
-The synthesis algorithms we have designed take inputs that include a format specification (which includes specification of equivalence classes) and a collection of examples.
-As the class of transformations becomes richer, the number of potential programs grows dramatically.
-As a result, the corresponding inference algorithm slows and its ability to guess the transformation desired by the user decreases, leading to reduced accuracy.
-We found that it was possible to overcome such challenges through new heuristics that use information theory to help guide the search for program transformations.
-In addition, we found the use of compositional synthesis, which is the process of breaking down a complex synthesis task into smaller, more mangeable subtasks critical when attempting to scale our prototype system up to be able to handle larger transformation task.
+During the course of this project, we have designed, analyzed and
+implemented algorithms that demonstrate it is possible to synthesize
+three classes of bidirectional transformations: (1) pure bijective
+transformations, (2) bijections modulo equivalences classes (quotient
+lenses) and (3)
+bijections modulo projections (symmetric lenses).  The synthesis algorithms we have
+designed take inputs that include a format specification (which
+includes specification of equivalence classes) and a collection of
+examples.  As the class of transformations becomes richer, the number
+of potential programs grows dramatically.  As a result, the
+corresponding inference algorithm slows and its ability to guess the
+transformation desired by the user decreases, leading to reduced
+accuracy.  We found that it was possible to overcome such challenges
+through new heuristics that use information theory to help guide the
+search for program transformations.  In addition, we found the use of
+compositional synthesis, which is the process of breaking down a
+complex synthesis task into smaller, more mangeable subtasks critical
+when attempting to scale our prototype system up to be able to handle
+larger transformation task.
 

### 2019-02-26 papers/SymmetricLenses/main.tex
edits to abstract

@@
 lenses, but also a rich subset of the full-blown
-``symmetric lenses'' proposed by Hofmann \ETAL.
-
+``symmetric lenses'' proposed by Hofmann \ETAL  
+Intuitively, simple symmetric lenses allow some information to be present
+on one side but not the other and vice versa.
+Simple symmetric lenses are of independent theoretical interest, being
+the largest class of symmetric lenses that do not rely on persistent
+internal state.  
 Synthesizing simple symmetric lenses is substantially more challenging than
@@
 we use {\em stochastic regular expressions} and ideas from information
-theory to estimate the amount of information propagated by a candidate lens,
-in conjunction with user annotations identifying parts of the source and
+theory to estimate the amount of information propagated by a candidate
+lens, generally prefering lenses that propagate more information, 
+as well as user annotations identifying parts of the source and
 target data structures as either {\em irrelevant} or {\em essential}.
-
-Simple symmetric lenses are also of independent theoretical interest, being
-the largest class of symmetric lenses that do not rely on persistent
-internal state.
 

### 2019-02-27 papers/SymmetricLenses/main.tex
fix issues

--- 2019-02-25 papers/SymmetricLenses/main.tex
@@
 commentstyle=\itshape\rmfamily,
-emph={rep,iterate,synth,collapse,perm,squash,normalize,using,ins,del,lens,let,get,put,rquot,lquot,id,swap,concat,or,disconnect,merge_left,merge_right,const},
+emph={rep,iterate,synth,collapse,perm,squash,normalize,using,ins,del,lens,let,get,put,rquot,lquot,id,swap,concat,or,disconnect,merge_left,merge_right,const,skip,require},
 emphstyle=\upshape\sffamily\color{dkpurple}, 
@@
 \begin{lstlisting}
-let company = (name . ("Co." | "Inc." | "Ltd.")) | "UNK" in
+let company = (co_name . ("Co." | "Inc." | "Ltd.")) | "UNK" in
 let emp_ins = name . " " . name "," company in
@@
 
-Without unambiguity constraints, the method above for calculating entropy does
-not work, and it is not easy to see how to generalize it.  For example,
-consider the SRE $\PRegexOr{\lstinline{"a"}}{\lstinline{"a"}}{.5}$. Using
-the above computation, this SRE has an entropy of $1$, but the entropy
-should be $0$. These difficulties become particularly complex when the
-languages are ambiguous in subtle ways and the algorithm must be correct
-when only some of the two languages overlap. Similar issues occur when
-trying to find the entropy when $\emptyset$ is a subterm (what should the
-entropy of $\PRegexStar{\emptyset}{.5}$ be?), so we do not define entropy
-on~$\emptyset$.
+To understand the difficulties caused by ambiguity, consider the SRE
+$\PRegexOr{\lstinline{"a"}}{\lstinline{"a"}}{.5}$. The formula above defines the
+entropy to be 1, but the true entropy is 0 (\IE, no bits are needed to know the
+generated string will be \lstinline{"a"}). Similar issues occur when trying to
+find the entropy when $\emptyset$ is a subterm (what should the entropy of
+$\PRegexStar{\emptyset}{.5}$ be?), so we do not define entropy on~$\emptyset$.
 
@@
 \label{sec:ssl}
-We now give a formal description of all the combinators in our symmetric lens
-language. While our algorithms perform synthesis on stochastic regular
-expressions, lenses themselves are typed between regular expressions. We can
-recover a regular expression from a stochastic regular expression by deleting
-probability annotations. In a slight abuse of notation (as we are also using
-these symbols as nonterminals), if $\Regex$ is a
-stochastic regular expression, $\BRegex$ is the regular expression obtained by
-removing probability annotations. The typing judgment $\Lens \OfType \BRegex
-\Leftrightarrow \BRegexAlt$ means that $\Lens$ is a well typed lens between
-the language of \BRegex and the language of \BRegexAlt.
+Simple symmetric string lenses are bidirectional programs that convert string
+data between formats. These lenses operate over regular expressions, not SREs.
+If $\Regex$ is a stochastic regular expression, $\BRegex$ is the regular
+expression obtained by removing probability annotations.
 
@@
 We have already described a number of these combinators in \S\ref{sec:overview}.
-Briefly (full definitions are below), the compose lens $\Lens_1 ; \Lens_2$
+Briefly (full definitions are below), the lens $\Lens_1 ; \Lens_2$
 composes two lenses sequentially, it 
@@
 intermediate format, shared as the target of $\Lens_1$ and the source of
-$\Lens_2$. The \MergeR lens combines two sublenses which share the same target.
-Which lens gets used depends on what data is in the left-hand format, if it is
+$\Lens_2$. The \MergeR lens combines two sublenses that operate on disjoint
+source formats and share the same target.
+The lens that gets used depends on what data is in the left-hand format, if the data is
 in the domain of $\Lens_1$ then $\Lens_1$ gets used, and similarly for
@@
 
-We now give typing rules and semantics for (some of) these combinators. We do not
-include the semantics for combinators equivalent to those presented in prior
-work~\cite{boomerang}. Furthermore, we present
-just \CreateR and \PutR in cases where \CreateL and \PutL are symmetric;
-full details can be found in an appendix\ifappendices \S\ref{sec:appendixlenses}  \else{} (in the supplemental materials)\fi. For
-simplicity of presentation, we use unambiguous regular expressions
-everywhere, even though the identity and disconnect lenses can actually be defined over ambiguous
-regular expressions; we believe relaxing
-this restriction is unproblematic.
+We now give typing rules and semantics for (some of) these combinators. We write
+$\Lens \OfType \BRegex \Leftrightarrow \BRegexAlt$ to indicate that $\Lens$ is a
+well typed lens between the language of \BRegex and the language of \BRegexAlt.
+We present just \CreateR and \PutR in cases where \CreateL and \PutL are
+symmetric; full details can be found in an appendix\ifappendices
+\S\ref{sec:appendixlenses} \else{} (in the supplemental materials)\fi. For
+simplicity of presentation, we use unambiguous regular expressions everywhere,
+even though the identity and disconnect lenses can actually be defined over
+ambiguous regular expressions; relaxing this restriction is not problematic.
 
@@
 Just as the identity lens ignores the second argument in puts, disconnect lenses
-ignore the first in both puts and creates.  The data is unsynchronized in these
+ignore the first argument in both puts and creates.  The data is unsynchronized in these
 two formats, information from one format doesn't impact the other.
@@
 recover a string in one data format from the other. We use the function
-$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to calculate bounds of the
+$\REntropyOf{\RegexAlt \Given \Lens, \Regex}$ to calculate bounds on the
 expected amount of information required to recover a string in $\RegexAlt$ from
 a string in $\Regex$, synchronized by $\Lens$. Similarly, we use the function
-$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate bounds of the
+$\LEntropyOf{\Regex \Given \Lens, \RegexAlt}$ to calculate bounds on the
 expected amount of information required to recover a string in $\Regex$ from a
@@
 sidestep such considerations by synthesizing \emph{DNF lenses}---simple symmetric
-lenses that inhabit a disjunctive normal form.
+lenses that inhabit a disjunctive normal form that does not include composition.
 
@@
 Algorithm~\ref{alg:synth-sym-lens} presents our synthesis algorithm at a high
-level of abstraction. The input regular expressions are first converted into
-stochastic regular expressions with \ToStochastic.  
-%Once that conversion has
-%taken place, the $\RXSearch$ algorithm executes. This algorithm itself is
-%defined via five functions: (1) $\PCF{\PQ.Create}$, (2) $\PCF{\PQ.Pop}$, (3)
-%$\PCF{Expand}$, (4) $\PCF{\PQ.Push}$, and (5) $\PCF{\Continue}$. First,
-This pair of SREs is used to initialize a priority queue ($pq$). The priority of
-a SRE pair is the number of rewrites needed to derive the pair from the
-originals. Next, $\PCF{SynthSymLens}$ enters a loop that searches for likely
-lenses. The loop terminates when the algorithm believes it is unlikely to find a
-better lens than the best one it has found so far (a termination condition
-defined by $\PCF{\Continue}$). Within each iteration of the loop, it:
+level of abstraction. This algorithm searches for likely lenses in priority
+order one ``class'' at a time using a \GreedySynth subroutine. Each class is the
+set of lenses that can by typed by a pair of regular expressions, modulo a set
+of simple axioms such as associativity, commutativity, and distributivity. The
+\Expand subroutine generates new classes using the star-unrolling axioms
+(\UnrollstarLeftRule and \UnrollstarRightRule).
+
+To summarize, the input regular
+expressions are first converted into stochastic regular expressions with
+\ToStochastic. This pair of SREs is used to initialize a priority queue ($pq$).
+The priority of a SRE pair is the number of rewrites needed to derive the pair
+from the originals. Next, $\PCF{SynthSymLens}$ enters a loop that searches for
+likely lenses. The loop terminates when the algorithm believes it is unlikely to
+find a better lens than the best one it has found so far (a termination
+condition defined by $\PCF{\Continue}$). Within each iteration of the loop, it:
 \begin{itemize}
@@
 further depth in the following subsections.
-
-\emph{Stochastic DNF regular expressions} are critical to the success of our
-algorithm. Stochastic DNF regular expressions (\SDNFREabbrev{}s) are a
-pseudo-normal form for stochastic regular expressions. Of all the star-semiring
-equations required to relate two \SDNFREabbrevs, only those relating to Kleene
-star (\EG, \UnrollstarLeftRule{} and \UnrollstarLeftRule{}) are needed to relate
-two \SDNFREabbrevs; other axioms, like associativity, commutativity or
-distributivity, are unneeded. This important property makes it possible to
-construct an efficient algorithm ($\GreedySynth$), directed by the syntax of two
-\SDNFREabbrevs, to synthesize any lenses in the same ``class'' (\IE, any lenses
-that have the same type when the Kleene star axioms are unused). This property,
-and the resulting efficiency of $\GreedySynth$ is key to the effectiveness of
-our core algorithms.
 
@@
 distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
-current best lens has cost $c$, then
-  $\Call{Continue}{pq}$ checks whether $c > d + \Log_2(n)$.
+current best lens has cost $c$, then $\Call{Continue}{pq}$ checks whether $c > d
++ \Log_2(n)$. This termination condition is based around two primary principles:
+do not search overly deep ($d$), and do not tackle a frontier that would take
+too long to process ($\Log_2(n)$). The log of the frontier is included because
+the frontier grows much faster than lens costs grow. Lens cost grows with the
+log of the expected number of choices in a given lens (due to its information
+theoretic basis), so the frontier calculation does too.
 
@@
 \label{subsec:sdnfre}
-Just as SREs are REs with probability annotations,
-\SDNFREabbrevs are \DNFREabbrevs with probability annotations.
-As with SREs and regular
-expressions, if $\DNFRegex$ is a stochastic DNF regular expression,
-$\overline{\DNFRegex}$ is the DNF regular expression generated by erasing
-probability annotations.
+The \GreedySynth subroutine converts input SREs into a temporary pseudo-normal
+form, \emph{Stochastic DNF regular expressions} (\SDNFREabbrev{}s).
+\SDNFREabbrev{}s normalize across many of the star-semiring equivalences -- if
+two regular expressions are equal modulo differences in associativity,
+commutativity, or distributivity, their corresponding SREs are syntactically
+equal.
 
@@
 corresponding stochastic DNF regular expressions, written
-$\ToDNFRegexOf{\Regex}$, is defined below.
+$\ToDNFRegexOf{\Regex}$, is defined below. This conversion relies on operators
+defined in Figure~\ref{fig:dnf-regex-functions}.
 \[
@@
 \]
-This conversion relies on operators defined in
-Figure~\ref{fig:dnf-regex-functions}. We have proved that this conversion
-respects languages and probability distributions.
+After this conversion has taken place, the sequences are ordered (thereby
+normalizing for commutativity differences).
+We have proven that this conversion respects languages and probability
+distributions.
 
@@
 
+\paragraph*{Entropy}
+We have developed a syntactic means for finding the entropy of a stochastic DNF
+regular expression, like we have for stochastic regular expressions. This
+enables us to efficiently find the entropy without first converting a \SDNFREabbrev to
+a stochastic regular expression.
+
+\begin{center}
+  \begin{tabular}{rcl}
+    $\EntropyOf{\PRegexStar{\DNFRegex}{\Probability}}$
+    & =
+    & $\frac{\Probability}{1-\Probability}(\EntropyOf{\DNFRegex} - \Log_2\Probability)
+      - \Log_2(1-\Probability)
+      $\\
+
+    $\EntropyOf{\SequenceOf{\String_0 \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_n \SeqSep \String_n}}$
+    & =
+    & $\Sigma_{i = 1}^n\EntropyOf{\Atom_i}$ \\
+
+    $\EntropyOf{\DNFOf{(\Sequence_1,\Probability_1) \DNFSep \ldots \DNFSep (\Sequence_n,\Probability_n)}}$
+    & =
+    & $\Sigma_{i=1}^n\Probability_i(\EntropyOf{\Sequence_i}+\Log_2\Probability_i)$\\
+  \end{tabular}
+\end{center}
+
+\begin{theorem}
+  $\EntropyOf{\DNFRegex}$ is the entropy of $P_{\DNFRegex}$.
+\end{theorem}
+
 \paragraph*{\ToStochastic} With stochastic DNF regular expressions and
-\ToDNFRegex defined, it is easier to explain \ToStochastic, the function that converts regular
-expressions into stochastic regular expressions. If users
-provided stochastic regular expressions themselves (perhaps inferred from
-example data), this step could be skipped,
-though our implementation does not (yet) permit users to provide such
-stochastic regular expressions. If $\Regex$ is a stochastic DNF regular
-expression generated by \ToStochastic, then $\ToDNFRegexOf{\Regex} =
-\DNFOf{(\Sequence_1,\frac{1}{n}) \DNFSep \ldots \DNFSep
+\ToDNFRegex defined, it is easier to explain \ToStochastic, the function that
+converts regular expressions into stochastic regular expressions. If $\Regex$ is
+a stochastic regular expression generated by \ToStochastic, then when put into
+\SDNFREabbrev form,
+$\ToDNFRegexOf{\Regex} = \DNFOf{(\Sequence_1,\frac{1}{n}) \DNFSep \ldots \DNFSep
   (\Sequence_n,\frac{1}{n})}$ for some sequences $\Sequence_1\ldots\Sequence_n$,
@@
 iterating. In our experience generating random strings from regular expressions,
-these probabilities provide good distributions of strings---stars are iterated
-4 times on average, and no individual choice in a series disjunctions is chosen
-disproportionately often. As we have found these distributions generate
-reasonable strings, we believe they describe the shape of data well.
-
-\paragraph*{Entropy}
-We have developed a syntactic means for finding the entropy of a stochastic DNF
-regular expression, like we have for stochastic regular expressions. This
-enables us to efficiently find the entropy without first converting a \SDNFREabbrev to
-a stochastic regular expression.
-
+these probabilities provide good distributions of strings---stars are iterated 4
+times on average, and no individual choice in a series of disjunctions is chosen
+disproportionately often.
+
+\subsection{Relevance Annotations}
+
+Even though our generated probability distribution works well in most
+situations, it is not perfect. Consider synthesizing a lens between the formats
+shown in Figure~\ref{fig:minimized-representations}. While salary information is
+present in \lstinline{emp_salaries}, but not in \lstinline{emp_insurance}, the
+algorithm might spend a long time (fruitlessly) trying to construct lenses that
+transform the salary to information present in \lstinline{emp_insurance}.
+
+One solution would be to cut off the search early. However, then one runs into
+the opposite problem: in other scenarios salary information may be present in
+the other format, but it may well take quite a bit of work to find a
+transformation that connects the salary in \lstinline{emp_salaries} to the
+salary information in the second format. In such a scenario, early termination
+may leave salary information discarded where it should be retained.
+
+We can solve both problems by allowing users to augment the specifications with
+\emph{relevance annotations}. Sometimes, users have external knowledge that certain
+information appears exclusively in one format or the other or conversely appears
+in both formats. By communicating this knowledge to the synthesis algorithm
+through relevance annotations, users can force the synthesis of lenses that
+discard or retain certain information.
+The first annotation, $\SkipOf{\Regex}$, says the information of
+$\Regex$ appears only in $\Regex$, and can safely be
+projected. The second annotation, $\SRequireOf{\Regex}$, says the information of
+$\Regex$ appears in the other format and cannot be discarded.
+
+For example, the user can easily recognize that salary information of
+\lstinline{emp_salaries} is not present in \lstinline{emp_insurance}. By
+annotating the \lstinline{salary} field as \lstinline{skip(salary)}, users can
+add this knowledge to the specification to optimize the search. In practice,
+we define \Skip by setting the information content of $\Skip(\Regex)$ to zero.
+\begin{gather*}
+  \EntropyOf{\Skip(\Regex)} = 0
+\end{gather*}
+
+Similarly, users can recognize that employee names are present in both files. By
+annotating instances of \lstinline{name} as \lstinline{require(name)}, users can
+add this knowledge to the specification to optimize the search and enforce that
+the returned lens that retains \lstinline{name} information. In practice, we
+define \SRequire by setting the cost of any lens that loses ``required''
+information to infinity.
+%
 \begin{center}
   \begin{tabular}{rcl}
-    $\EntropyOf{\PRegexStar{\DNFRegex}{\Probability}}$
-    & =
-    & $\frac{\Probability}{1-\Probability}(\EntropyOf{\DNFRegex} - \Log_2\Probability)
-      - \Log_2(1-\Probability)
-      $\\
-
-    $\EntropyOf{\SequenceOf{\String_0 \SeqSep \Atom_1 \SeqSep \ldots \SeqSep \Atom_n \SeqSep \String_n}}$
-    & =
-    & $\Sigma_{i = 1}^n\EntropyOf{\Atom_i}$ \\
-
-    $\EntropyOf{\DNFOf{(\Sequence_1,\Probability_1) \DNFSep \ldots \DNFSep (\Sequence_n,\Probability_n)}}$
-    & =
-    & $\Sigma_{i=1}^n\Probability_i(\EntropyOf{\Sequence_i}+\Log_2\Probability_i)$\\
+    $\REntropyOf{\SRequireOf{\RegexAlt} \Given \Lens, \Regex}$
+    & =
+    & $\begin{cases}
+      \infty & \text{if }\REntropyOf{\SRequireOf{\RegexAlt} \Given \Lens, \Regex}
+      \neq 0\\
+    0 & \text{otherwise}
+  \end{cases}$\\
+
+    $\REntropyOf{\RegexAlt \Given \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, \Regex}$
+    & =
+    & $\begin{cases}
+      \infty & \vspace*{-.3em}\text{if $\RegexAlt$ contains $\SRequireOf{\RegexAlt'}$ as a}\\
+      & \text{subexpression, where $\EntropyOf{\RegexAlt'} \neq 0$}\\
+    [\EntropyOf{\RegexAlt},\EntropyOf{\RegexAlt}] & \text{otherwise}
+  \end{cases}$\\
   \end{tabular}
 \end{center}
-
-\begin{theorem}
-  $\EntropyOf{\DNFRegex}$ is the entropy of $P_{\DNFRegex}$.
-\end{theorem}
 
@@
 
-Intuitively, a symmetric DNF lens $sdl$ is a union of symmetric sequence lenses,
-$ssql_1 \ldots ssql_p$, a symmetric DNF sequence lens is a concatenation of
-symmetric atom lenses, $sal_1 \ldots sal_n$, and a symmetric atom lens $sal$ is
-an iteration of a symmetric DNF lens. Note that the structure of symmetric DNF
-lenses exactly mirrors that of SDNF REs.
+The structure of symmetric DNF lenses exactly mirrors that of SDNF REs. A
+symmetric DNF lens $sdl$ is a union of symmetric sequence lenses, $ssql_1 \ldots
+ssql_p$, a symmetric DNF sequence lens is a concatenation of symmetric atom
+lenses, $sal_1 \ldots sal_n$, and a symmetric atom lens $sal$ is an iteration of
+a symmetric DNF lens.
 
@@
 use $\REntropyOf{\DNFRegexAlt \Given \SDNFLens, \DNFRegex}$ to calculate bounds
-of the expected amount of information required to recover a string in
+on the expected amount of information required to recover a string in
 $\DNFRegexAlt$ from a string in $\DNFRegex$, synchronized by $\SDNFLens$. We use
@@
 complexity.
-
-\subsection{Relevance Annotations}
-
-In this work, there are three competing constraints. The tool should (1) return
-lenses quickly to users, (2) return the highest quality lenses, and (3)
-not require users to write a large number of examples.
-Our termination heuristic works well for most situations, but
-there are cases where it fails.
-
-In particular, our system works hard to find lenses that project \emph{as little
-  information as possible}. If the desired lens requires projecting a large
-amount of information, the termination condition will be satisfied ``too late.''
-While \GreedySynth may have found what the user knows to be the desired lens,
-the full algorithm will continue trying to find lenses that project less
-information.
-
-Conversely, sometimes the termination heuristic is overly permissive and algorithm cuts
-short: if it 
-had continued searching for lenses it would have found the desired one, but the
-termination heuristic was satisfied ``too early.'' We wish for our algorithm to
-provide feedback in a timely manner, but sometimes this prohibits finding the
-desired lens.
-
-To address these issues, we allow annotations on the regular expressions
-dictating relevance. The first, $\SkipOf{\Regex}$, says the information
-content of $\Regex$ is irrelevant to the desired lens, and can the data can be
-safely projected. The second, $\SRequireOf{\Regex}$, says the information of
-$\Regex$ is critical to the desired lens, and none of the information can be
-lost when converting between the formats.
-
-If synthesizing a lens seems to be taking a long time, it is typically spending
-time checking it cannot find a lens that projects less information. By
-``skipping'' portions of the SRE that \emph{should} be projected, the algorithm
-will not continue searching for a better lens when it has already found the
-correct one. In practice, we implement \Skip by setting the information content
-of $\Skip(\Regex)$ to zero.
-%
-\begin{gather*}
-  \EntropyOf{\Skip(\Regex)} = 0
-\end{gather*}
-%
-If synthesizing a lens between \lstinline{emp_salaries} and
-\lstinline{emp_insurance} from Figure~\ref{fig:minimized-representations} is
-taking overly long, the synthesis procedure is uncomfortable disconnecting
-\lstinline{company} and \lstinline{salary} information. Redefining
-\lstinline{company} and \lstinline{salary} as $\SkipOf{\lstinline{company}}$ and
-$\SkipOf{\lstinline{salary}}$ would set their information content to zero,
-alleviating such issues.
-
-If the algorithm returns a lens with too many disconnects and merges, one could
-provide additional examples to constrain the lens more. However, for large
-formats, writing such examples is time consuming and error-prone. Instead, if
-one recognizes that information from $\Regex$ is incorrectly lost when converted
-between formats, they can prevent that information from being lost by annotating
-it $\SRequireOf{\Regex}$. One can guarantee a lens between \Regex and \RegexAlt
-is bijective by annotating them as $\SRequireOf{\Regex}$ and
-$\SRequireOf{\RegexAlt}$. In practice, we implement \SRequire by setting the
-cost of any lens that loses ``required'' information to infinity.
-%
-\begin{center}
-  \begin{tabular}{rcl}
-    $\REntropyOf{\SRequireOf{\RegexAlt} \Given \Lens, \Regex}$
-    & =
-    & $\begin{cases}
-      \infty & \text{if }\REntropyOf{\SRequireOf{\RegexAlt} \Given \Lens, \Regex}
-      \neq 0\\
-    0 & \text{otherwise}
-  \end{cases}$\\
-
-    $\REntropyOf{\RegexAlt \Given \DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}, \Regex}$
-    & =
-    & $\begin{cases}
-      \infty & \vspace*{-.3em}\text{if $\RegexAlt$ contains $\SRequireOf{\RegexAlt'}$ as a subexpression,}\\
-      & \text{where $\EntropyOf{\RegexAlt'} \neq 0$}\\
-    [\EntropyOf{\RegexAlt},\EntropyOf{\RegexAlt}] & \text{otherwise}
-  \end{cases}$\\
-  \end{tabular}
-\end{center}
-
-%
-If synthesizing a lens between \lstinline{emp_salaries} and
-\lstinline{emp_insurance} from Figure~\ref{fig:minimized-representations}
-quickly returns a lens with too many disconnects and merges (perhaps
-disconnecting instances of \lstinline{name} in \lstinline{emp_salary} and
-\lstinline{emp_ins}), the termination heuristic is being overly permissive. By
-replacing \lstinline{name} with $\SRequireOf{\lstinline{name}}$ in
-\lstinline{emp_salary} and \lstinline{emp_ins}, name information can no longer
-be lost, and future synthesis calls would continue until a more likely lens is
-found.
 % end synthesis

### 2019-02-27 papers/SymmetricLenses/main.tex
update relevance annotations to be smoother text

@@
 \emph{relevance annotations}. Sometimes, users have external knowledge that certain
-information appears exclusively in one format or the other or conversely appears
-in both formats. By communicating this knowledge to the synthesis algorithm
+information appears exclusively in one format or the other, or they may know the
+information is present both formats. By communicating this knowledge to the synthesis algorithm
 through relevance annotations, users can force the synthesis of lenses that
@@
 annotations that guide DNF conversion; an expansion inference algorithm;
-hash-consing and memoization; and compositional synthesis. The implementation
-also provides for additional specifications in the form of annotations that help
-describe which information is irrelevant to the lens. The optimizations make the
-system performant enough for interactive use, and the annotations help ensure
-the desired information is projected.
+hash-consing and memoization; and compositional synthesis. The optimizations
+make the system performant enough for interactive use, and the annotations help
+ensure the desired information is projected.
 

### 2019-02-27 papers/SymmetricLenses/main.tex
make 25 pages yay

@@
 %
-In particular, the recent Optician~\cite{optician} enhancement to Boomerang has
+The recent Optician~\cite{optician} enhancement to Boomerang has
 demonstrated that it is feasible to synthesize quite complex examples, albeit
 only for a restricted class of lenses---so-called bijective lenses.
-Specifically, given source and target formats plus a small number of examples of
+Given source and target formats plus a small number of examples of
 corresponding data, Optician's synthesis algorithm is guaranteed to find a
@@
 If the synthesis algorithm can
-find any bijection at all, it is fairly likely to be the intended one.
+find any bijection, it is fairly likely to be the intended one.
 
@@
 To characterize their expressive power formally, we show that simple
-symmetric lenses are exactly the symmetric lenses that satisfy an intuitive
-property called \emph{forgetfulness}.  To check that they are expressive
-enough in practice, we add simple symmetric lenses to the Boomerang
-language~\cite{boomerang} and apply them to a range of real-world
-applications, showing in the process that simple symmetric lenses integrate
+symmetric lenses are exactly symmetric lenses that satisfy an intuitive
+property called \emph{forgetfulness}.  We show they are expressive
+enough in practice by adding simple symmetric lenses to the Boomerang
+language~\cite{boomerang} and applying them to a range of real-world
+applications, showing that simple symmetric lenses integrate
 smoothly with other advanced lens features provided by Boomerang, including
@@
 
-We consider a ``likely'' lens to be one that propagates ``a lot'' of information
-from the left data format to the right and vice versa. Conversely, an unlikely
-lens is one that requires a large amount of additional information to recover
+We consider ``likely'' lenses to be ones that propagate ``a lot'' of information
+from the left data format to the right and vice versa. Conversely, unlikely
+lenses require a large amount of additional information to recover
 either of the formats given the other. For many application scenarios, users
@@
 algorithm performs a type-directed search between the descriptions of the source
-and target formats, measuring success using the likelihood measure. The
-interesting complications arise from the need to deal with regular-expression
+and target formats, measuring success using the likelihood measure.
+Interesting complications arise from the need to deal with regular-expression
 equivalences. There are infinitely many regular expressions equivalent to a
@@
 procedures (Figure~\ref{fig:high-level-algorithm}), following~\citet{optician}.
-The first, \Expand, uses rewriting rules to iteratively propose new pairs of
+The first, \Expand, uses rewriting rules to propose new pairs of
 stochastic regular expressions equivalent to the original pair. The second,
-\GreedySynth, takes a pair of SREs as input and uses a greedy, type-directed
-algorithm to find a simple symmetric lens between them, returning the lens and
+\GreedySynth, uses a greedy, type-directed
+algorithm to find a simple symmetric lens between input SRE pairs, returning the lens and
 its likelihood score to \Expand. The whole synthesis algorithm heuristically
@@
 transformations. We evaluated our system on a benchmark suite of 48 problems, including data-cleaning
-tasks taken from Flash Fill~\cite{gulwani-popl-2014}, view-maintenance tasks
+tasks taken from Flash Fill~\cite{flashfill}, view-maintenance tasks
 taken from Augeas~\cite{augeas2}, and ad-hoc synchronization tasks from a
@@
 
-\item We also extend Boomerang with symmetric lens synthesis specifications,
-  which include types, examples, and relevance annotations. We evaluate our implementation, the effects of
-  various optimizations, and our inclusion of relevance annotations on a set of
+\item We also extend Boomerang with symmetric lens synthesis specifications:
+  including types, examples, and relevance annotations. We evaluate our implementation, the effects of
+  optimizations, and our inclusion of relevance annotations on a set of
   48 lens synthesis benchmarks drawn from data cleaning, view maintenance, and
@@
 Along the
-way, we show that {\em star-semiring equivalences} on regular expressions
+way, we show the {\em star-semiring equivalences} on regular expressions
 remain valid when extended to stochastic regular expressions, preserving
@@
 expected number of bits required to recover data in one format from synchronized
-data in the other. Higher likelihoods correspond to fewer bits. We say that two
+data in the other. Higher likelihoods correspond to fewer bits. Two
 strings $s$ and $t$ are \emph{synchronized according to lens $l$} if $l.\PutR\
-s\ t = t$ and $l.\PutL\ t\ s = s$. We say that we can \emph{recover} $s$ from
+s\ t = t$ and $l.\PutL\ t\ s = s$. We can \emph{recover} $s$ from
 $t$ using bits $b$ and lens $l$ if we can reconstruct $s$ from $t$, $b$, and
@@
 ($-\Log_2P(s)$). The entropy of $S$ calculates the expected number of bits
-required to specify a random element pulled from $S$, in other words, the
+required to specify an element drawn from $S$, in other words, the
 probability of each element times its information content. Entropy maintains
@@
 needed to recover a string in one data source from a synchronized string in the
-other data source. To do this, we must first develop a probabilistic 
-model for our language. We do this with \emph{stochastic regular
+other data source. To do this, we first develop a probabilistic 
+model for our language using \emph{stochastic regular
   expressions} (SREs)---regular expressions annotated with probability
-information~\cite{stoch-rnn,stoch-def}---which 
+information~\cite{stoch-rnn,stoch-def}---that
 jointly express a language and a probability distribution over that language. 
@@
 over real numbers between 0 and 1, exclusive.)
-
 The semantics of an SRE $S$ is a probability distribution $P_S$ defined as
@@
 \end{center}
-In the SRE $\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, the annotation
-$\Probability$ represent that, if generating arbitrary strings from
-$\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$, strings should be pulled from
-$\Regex_1$ with probability $\Probability$ (\EG pulling a string from
-$\PRegexOr{\lstinline{"a"}}{\lstinline{"b"}}{.2}$ would retrieve \lstinline{"a"}
-20\% of the time). In the SRE
-$\PRegexStar{\Regex}{\Probability}$, the annotation \Probability represents how
-much more likely having a string involving more than $n$ iterations is than
-having a string involving exactly $n$ iterations (\EG pulling a random string
-from $\PRegexStar{\lstinline{"c"}}{.8}$ would give \lstinline{""} 20\% of the
-time, \lstinline{"c"} 16\% of the time, and \lstinline{"cc"} 12.8\% of the time).
+One may think of SREs as string generators. Under this interpretation, the
+constant SRE $\String$ always generates the string $\String$ and never any other
+string. The SRE $\PRegexOr{\Regex_1}{\Regex_2}{\Probability}$ generates a string
+from $\Regex_1$ with probability $\Probability$ and generates a string from
+$\Regex_2$ with probability $1-\Probability$. The SRE
+$\PRegexStar{\lstinline{"c"}}{\Probability}$ generates the empty string with
+probability $1-\Probability$, and generates any other string in the language with
+probability $\Probability$. More generally,
+$\PRegexStar{\Regex}{\Probability}$ generates strings in $\Regex^n$ with
+probability $\Probability^n(1-\Probability)$.
 
@@
 \subsection{Stochastic Regular Expression Equivalences}
-The $\Expand$ algorithm needs to be able to enumerate SREs that are
+The $\Expand$ algorithm enumerates SREs that are
 ``equivalent'' to a given one. However, existing work does not define any notion
 of stochastic regular expression equivalence. Figure~\ref{fig:ssr-sre} shows how
-we to extend the \emph{star-semiring}~\cite{Droste2009} equivalences (a
+we to extend the \emph{star-semiring}~\cite{starsemiring} equivalences (a
 \emph{finer} notion of equivalence than semantic equivalence) to
@@
 The entropy of a data source $S$ is the expected number of bits required to
-describe a random element in $S$ (formally $-\Sigma_{s\in S}\,P(s)\cdot
+describe an element drawn from $S$ (formally $-\Sigma_{s\in S}\,P(s)\cdot
 \Log_2P(s)$). The entropy of a SRE can be computed directly from its syntax,
@@
 
-For example, the entropy of $\PRegexOr{\lstinline{"a"}}{(\lstinline{"b"})}{.5}$
-is 1. This is expected because on average 1 bit is required to encode whether a
-randomly pulled string is \lstinline{"a"} or \lstinline{"b"}. As an additional
-example, a fixed string has no information content, and so has no entropy.
+For example, the entropy of $\Regex =
+\PRegexOr{\lstinline{"a"}}{(\lstinline{"b"})}{.5}$ is 1. The best encoding of a
+stream of elements from $\Regex$ will use, on average, 1 bit per element to
+determine whether that element is \lstinline{"a"} or \lstinline{"b"}. As an
+additional example, a fixed string has no information content, and so has no
+entropy.
 
@@
 
-The procedure \PCF{\Continue} returns false based on the how long the search has
-been going, and how hard it expects the next class of problems to be. In
-particular, if $\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at
+The procedure \PCF{\Continue} terminates the loop based on the how long the
+search has been going, and how hard it expects the next class of problems to be.
+In particular, if $\PCF{PQ.Peek}(pq) = (\Regex,\RegexAlt)$, that RE pair is at
 distance $d$, the number of pairs in $pq$ at distance $d$ is $n$, and the
-current best lens has cost $c$, then $\Call{Continue}{pq}$ checks whether $c > d
-+ \Log_2(n)$. This termination condition is based around two primary principles:
-do not search overly deep ($d$), and do not tackle a frontier that would take
-too long to process ($\Log_2(n)$). The log of the frontier is included because
-the frontier grows much faster than lens costs grow. Lens cost grows with the
-log of the expected number of choices in a given lens (due to its information
-theoretic basis), so the frontier calculation does too.
-
-When $\Call{Continue}{pq}$ returns $\False$, the algorithm stops proposing
+current best lens has cost $c$, then $\Call{Continue}{pq}$ continues the loop
+while $c < d + \Log_2(n)$. This termination condition is based around two
+primary principles: do not search overly deep ($d$), and do not tackle a
+frontier that would take too long to process ($\Log_2(n)$). The log of the
+frontier is included because the frontier grows much faster than lens costs
+grow. Lens cost grows with the log of the expected number of choices in a given
+lens (due to its information theoretic basis), so the frontier calculation does
+too.
+
+When $\Call{Continue}{pq}$ terminates the loop, the algorithm stops proposing
 regular expression pairs, and instead returns to the user the best lens found
@@
 
-The algorithm for converting a stochastic regular expressions into its
-corresponding stochastic DNF regular expressions, written
+The algorithm for converting a stochastic regular expressions $\Regex$ into its
+corresponding \SDNFREabbrev form, written
 $\ToDNFRegexOf{\Regex}$, is defined below. This conversion relies on operators
@@
 \]
-After this conversion has taken place, the sequences are ordered (thereby
-normalizing for commutativity differences).
-We have proven that this conversion respects languages and probability
-distributions.
+After this syntactic conversion has taken place, the sequences are ordered
+(normalizing commutativity differences). This conversion respects languages and
+probability distributions.
 
@@
 a stochastic regular expression generated by \ToStochastic, then when put into
-\SDNFREabbrev form,
-$\ToDNFRegexOf{\Regex} = \DNFOf{(\Sequence_1,\frac{1}{n}) \DNFSep \ldots \DNFSep
-  (\Sequence_n,\frac{1}{n})}$ for some sequences $\Sequence_1\ldots\Sequence_n$,
-and every stochastic atom generated by \ToStochastic is
-$\PRegexStar{\DNFRegex}{.8}$. In particular, $\ToDNFRegex$ generates regular
-expressions whose DNF form gives equal probability to all sequence subcomponents
-of the DNF regular expression, and gives iterations a .8 chance to continue
-iterating. In our experience generating random strings from regular expressions,
-these probabilities provide good distributions of strings---stars are iterated 4
-times on average, and no individual choice in a series of disjunctions is chosen
-disproportionately often.
+\SDNFREabbrev form, $\ToDNFRegexOf{\Regex} = \DNFOf{(\Sequence_1,\frac{1}{n})
+  \DNFSep \ldots \DNFSep (\Sequence_n,\frac{1}{n})}$ for some sequences
+$\Sequence_1\ldots\Sequence_n$, and every stochastic atom generated by
+\ToStochastic is $\PRegexStar{\DNFRegex}{.8}$. In particular, $\ToDNFRegex$
+generates regular expressions whose DNF form gives equal probability to all
+sequence subcomponents of the \SDNFREabbrevs, and gives a .8 chance for
+stars to continue iterating. In our experience generating random strings from regular
+expressions, these probabilities provide good distributions of strings---stars
+are iterated 4 times on average, and no individual choice in a series of
+disjunctions is chosen disproportionately often.
 
@@
 situations, it is not perfect. Consider synthesizing a lens between the formats
-shown in Figure~\ref{fig:minimized-representations}. While salary information is
+shown in Figure~\ref{fig:minimized-representations}. Because salary information is
 present in \lstinline{emp_salaries}, but not in \lstinline{emp_insurance}, the
 algorithm might spend a long time (fruitlessly) trying to construct lenses that
-transform the salary to information present in \lstinline{emp_insurance}.
+transform the salary to information present in \lstinline{emp_insurance} even
+though that is impossible. In a similar but more elaborate example, such wasted
+processing effort may cause synthesis to fail to terminate in any reasonable
+amount of time.
 
 One solution would be to cut off the search early. However, then one runs into
-the opposite problem: in other scenarios salary information may be present in
-the other format, but it may well take quite a bit of work to find a
+the opposite problem: In other scenarios, salary information may be present in
+the other format, but it may take quite a bit of work to find a
 transformation that connects the salary in \lstinline{emp_salaries} to the
-salary information in the second format. In such a scenario, early termination
-may leave salary information discarded where it should be retained.
+salary in the second format. Hence, early termination
+may cut off synthesis before the right lens is found.
 
@@
 
-For example, the user can easily recognize that salary information of
-\lstinline{emp_salaries} is not present in \lstinline{emp_insurance}. By
-annotating the \lstinline{salary} field as \lstinline{skip(salary)}, users can
-add this knowledge to the specification to optimize the search. In practice,
-we define \Skip by setting the information content of $\Skip(\Regex)$ to zero.
+For example, users can easily recognize that salary information is not present
+in \lstinline{emp_insurance}. By annotating the \lstinline{salary} field as
+\lstinline{skip(salary)}, users can add this knowledge to the specification to
+optimize the search. In practice, we define the information content of
+$\Skip(\Regex)$ to be zero.
 \begin{gather*}
@@
 annotating instances of \lstinline{name} as \lstinline{require(name)}, users can
-add this knowledge to the specification to optimize the search and enforce that
-the returned lens that retains \lstinline{name} information. In practice, we
-define \SRequire by setting the cost of any lens that loses ``required''
-information to infinity.
+add this knowledge to the specification to optimize the search and force the
+the generated lens to retain \lstinline{name} information. In practice, we
+define the cost of any lens that loses ``required'' information to be infinite.
 %
@@
 
-The structure of symmetric DNF lenses exactly mirrors that of SDNF REs. A
-symmetric DNF lens $sdl$ is a union of symmetric sequence lenses, $ssql_1 \ldots
-ssql_p$, a symmetric DNF sequence lens is a concatenation of symmetric atom
-lenses, $sal_1 \ldots sal_n$, and a symmetric atom lens $sal$ is an iteration of
-a symmetric DNF lens.
+The structure of symmetric DNF lenses mirrors that of SDNF REs. A symmetric DNF
+lens ($sdl$) is a union of symmetric sequence lenses, a symmetric sequence
+lens ($ssql$) is a concatenation of symmetric atom lenses, and a
+symmetric atom lens ($sal$) is an iteration of a symmetric DNF lens.
 
@@
 the structure of symmetric DNF lenses and \SDNFREabbrevs.
-
-\paragraph*{Symmetric DNF Lenses} Algorithm~\ref{alg:greedysynth} presents
-\GreedySynth, which synthesizes symmetric DNF lenses. Its inputs 
-are a suite of input-output examples and a pair of stochastic DNF regular expressions.
-%
-First, \PCF{CannotMap} determines whether there is no lens satisfying the
-examples, and if so, \GreedySynth returns \None immediately. Otherwise, \GreedySynth finds
-the best lenses (given the examples that match them) between all sequence pairs
-$(SQ_i, TQ_j)$ drawn from the left and right DNF regular expressions. (The
-function \CartesianMap maps its argument across the cross product of the input
-lists). A priority queue containing these sequence lenses, ordered by cost, is
-then initialized with $\PCF{PQ.Create}$. The symmetric lens is then built up
-iteratively from these sequence lenses, where the state of the partially
-constructed lens is tracked in the lens builder, $lb$.
-
-\GreedySynth loops until there are no more sequence lenses in the priority
-queue. Within this loop, a sequence lens is popped from the queue and, if it is
-``useful,'' it is included in the final DNF lens. The lens is considered to
-be useful when its source (or target) is \textit{not} already the source (or target) of
-an already chosen sequence lens. If the examples require that two sequences have a
-lens between them, then any sequence lens between them is considered useful. We
-update the priorities of the sequence lenses as the algorithm proceeds: if two
-sequence lenses have the same source, the second one to be popped gets a higher
-cost than it originally had, to account for the information that needs to be stored
-for including that source of non-bijectivity.
-
-As an example, consider searching for a lens between \lstinline{"" | name.name$^*$} and \lstinline{"" | name}. \GreedySynth might first pop the
-sequence pair \lstinline{""} and \lstinline{""}, because there is a bijective
-sequence lens between them. As neither \lstinline{""} is involved in a sequence
-lens, this lens is considered useful. Next, the sequence lens between
-\lstinline{name.name$^*$} and \lstinline{name} would be popped: while that lens
-is not bijective it is still better than the alternatives. As all sequences are
-now involved in sequence lenses, and there are no examples to make other lenses
-useful, no more sequence lenses would be added to the lens builder.
-
-Finally, after all sequences have been popped, the partial DNF lens $lb$ is
-converted into a symmetric DNF lens. This is only possible if all sequences are
-involved in some sequence lens: if they are not, \PCF{LensBuilder.ToDNFLens}
-instead returns \None.
 
@@
 
+\paragraph*{Symmetric DNF Lenses} Algorithm~\ref{alg:greedysynth} presents
+\GreedySynth, which synthesizes symmetric DNF lenses. Its inputs 
+are a suite of input-output examples and a pair of stochastic DNF regular expressions.
+%
+First, \PCF{CannotMap} determines whether there is no lens satisfying the
+examples, and if so, \GreedySynth returns \None immediately. Otherwise, \GreedySynth finds
+the best lenses (given the examples that match them) between all sequence pairs
+$(SQ_i, TQ_j)$ drawn from the left and right DNF regular expressions. (The
+function \CartesianMap maps its argument across the cross product of the input
+lists). A priority queue containing these sequence lenses, ordered by cost, is
+then initialized with $\PCF{PQ.Create}$. The symmetric lens is then built up
+iteratively from these sequence lenses, where the state of the partially
+constructed lens is tracked in the lens builder, $lb$.
+
+\GreedySynth loops until there are no more sequence lenses in the priority
+queue. Within this loop, a sequence lens is popped from the queue and, if it is
+``useful,'' is included in the final DNF lens. The lens is considered to be
+useful when its source (or target) is \textit{not} already the source (or
+target) of an included sequence lens. If examples require that two
+sequences have a lens between them, such lenses are considered useful. The
+priorities of the sequence lenses update as the algorithm proceeds: if two
+sequence lenses have the same source, the second one to be popped gets a higher
+cost than it originally had; information must now be stored for including that
+source of non-bijectivity.
+
+As an example, consider searching for a lens between \lstinline{"" | name.name$^*$} and \lstinline{"" | name}. \GreedySynth might first pop the
+sequence pair \lstinline{""} and \lstinline{""}, because there is a bijective
+sequence lens between them. As neither \lstinline{""} is involved in a sequence
+lens, this lens is considered useful. Next, the sequence lens between
+\lstinline{name.name$^*$} and \lstinline{name} would be popped: while that lens
+is not bijective it is still better than the alternatives. As all sequences are
+now involved in sequence lenses, and there are no examples to make other lenses
+useful, no more sequence lenses would be added to the lens builder.
+
+Finally, after all sequences have been popped, the partial DNF lens $lb$ is
+converted into a symmetric DNF lens. This is only possible if all sequences are
+involved in some sequence lens: if they are not, \PCF{LensBuilder.ToDNFLens}
+instead returns \None.
+
 \paragraph*{Symmetric Sequence lenses} Algorithm~\ref{alg:greedyseqsynth}
@@
 annotations that guide DNF conversion; an expansion inference algorithm;
-hash-consing and memoization; and compositional synthesis. The optimizations
+and compositional synthesis. The optimizations
 make the system performant enough for interactive use, and the annotations help
@@
 
-\paragraph*{Hash-consing and Memoization} We further improve \GreedySynth by
-hash-consing stochastic DNF regular expressions.  This optimization enables
-fast equality checks and saves memory.
-%memoizing its results.  However, our stochastic DNF regular expressions are
-%quite large and so hashing them is computationally intensive.  We use less
-%memory and allow for quick comparisons and hashing of these DNF regular
-%expressions by
-
 \paragraph*{Compositional Synthesis} Compositional synthesis allows us to use
@@
 \label{sec:evaluation}
-We have implemented symmetric DNF lenses as an extension to Boomerang.
-We have
-implemented all of Boomerang's asymmetric lens combinators using a
-combination of the symmetric lens combinators presented in this paper and
-symmetric versions of the asymmetric extensions (like matching
-lenses~\cite{matchinglenses} and quotient lenses\cite{quotientlenses}) already
-present in Boomerang. We also integrated our synthesis engine into Boomerang.
-This integration allows users to write synthesis tasks alongside lens
-combinators, incorporate synthesis results into manually-written lenses, and
-reference previously defined lenses in the synthesis of new lenses.
-
-All benchmarks were performed on a 2.5 GHz Intel Core i7 processor with 16
-GB of 1600 MHz DDR3 running macOS High Sierra.
+We implemented simple symmetric lenses as an extension to
+Boomerang~\cite{boomerang}. In doing so, we reimplemented Boomerang's
+asymmetric lens combinators using a combination of the symmetric
+combinators presented in this paper and symmetric versions of asymmetric
+extensions (like matching lenses~\cite{matchinglenses} and quotient
+lenses~\cite{quotientlenses}) already present in Boomerang. We also integrated
+our synthesis engine into Boomerang, allowing users to write synthesis tasks
+alongside lens combinators, incorporate synthesis results into manually-written
+lenses, and reference previously defined lenses during synthesis. All experiments
+were performed on a 2.5 GHz Intel Core i7 processor with 16 GB of 1600 MHz DDR3
+running macOS Mojave.
 
@@
 \item How much slower is our tool on bijective lens synthesis benchmarks than
-  the existing Optician system customized for bijective lenses?
+  prior work customized for bijective lenses~\cite{optician}?
 \item How effective is the information-theoretic search heuristic and how do our
@@
 \begin{enumerate}
-\item We adapted 8 data cleaning benchmarks from Flash Fill. Note
-that our tool produces bidirectional transformations rather
-than one-way transformers like Flash Fill. 
-\item We adapted 29 benchmarks from Augeas. These benchmarks turn the
-  ad hoc data present in Linux configuration files into a structured dictionary
-  representation.
+\item We adapted 8 data cleaning benchmarks from Flash Fill~\cite{flashfill}. Flash Fill data
+  cleaning tasks are either derived from online help forums or taken from the
+  Excel product team. Note that our tool produces bidirectional transformations
+  rather than one-way transformers like Flash Fill. We ensure one direction of
+  our bidirectional transformers performs the same unidirectional transformation
+  as Flash Fill -- we ignore the other direction.
+\item We adapted 29 benchmarks from Augeas~\cite{augeas}. Augeas is a utility
+  that bidirectionally converts between Linux configuration files and an
+  in-memory dictionaries. In our benchmarks, we synthesize lenses between Linux
+  configuration files and serialized versions of Augeas dictionaries.
 \item We created 11 additional benchmarks 
@@
 
-All three sources provide realistic problems suitable for answering the research
-questions: Flash Fill data cleaning tasks are either derived from online help
-forums or taken from the Excel product team; Augeas is a utility that transforms
-Linux configuration files into a manageable dictionary form; and many of the
-remaining benchmarks arise from synchronizing real-world data formats such as
-BibTeX and EndNote. Each task consists of a source and target regular
-expression. These tasks can be quite quite large, as we are synthesizing complex
+Each task consists of a source and target regular expression and a set of
+examples. These tasks can be quite quite large, as we are synthesizing complex
 transformations between real file formats -- some of the lenses and
-specifications are over 500 AST nodes.
+specifications are over 500 AST nodes \afm{run the experiments and collect the
+  data at some point}.
 
 \subsection{Synthesizing Correct Lenses}
-
-Because our tool is intended to be interactive---the user typically starts with
-zero or one examples, runs the tool, determines whether the output is the
-desired one and, if not, either adds an example or relevance annotations and
-then tries again---it is hard to cleanly evaluate the net user experience. The
-primary concern is whether it is usable to synthesize correct lenses.
 
@@
 
-We verified that \SSOpt{} generated the correct lenses the way programmers often
+We verified that our default running mode (\SSOpt{}) generated the correct lenses the way programmers often
 validate their programs: we manually inspected the code and ran unit tests on
@@
 
-\subsection{The Effects of Our Heuristics}
+\subsection{The Effects of Heuristics and Relevance Annotations}
 \label{subsec:effect-of-heuristics}
 We evaluate the usefulness of (1) our information-theoretic metric, and (2) our
-termination heuristic.  To this end, we run our program in several different modes:\\[1ex]
+relevance annotations.  To this end, we run our program in several different modes:\\[1ex]
 \begin{tabulary}{\linewidth}{rL}
@@
 needed. Related work on string transformations has often used simpler
-measures such as ``avoid constants'' which align with, but are simpler
+measures such as ``avoid constants'' that align with, but are simpler
 than our measures. The \CCOpt{} mode is an example of such a simple
@@
 eliminate the information in a union.
-
-Figure~\ref{fig:metric} summarizes the result of these experiments. The data
-reveal that the information-theoretic metric is critical for finding the correct
-lens: Only 10 of the benchmarks succeeded when running in \CCOpt{} mode. The
-termination condition is also quite important. When running in \FLOpt{} mode,
-the algorithm only discovers 5 lenses, which shows that the first cluster that
-contains a satisfying lens is rarely the correct cluster. However, our algorithm
-is not perfect and fails when either it is very difficult to find the desired
-lens (necessitating \SRequire) or when a large amount of data is projected
-(necessitating \Skip). Without any annotations, our algorithm finds the correct
-lens for 40 of our 48 benchmarks; eight required relevance annotations to find
-the correct lens.
-% end evaluation
-
-% begin related-work
-\section{Related Work}
-\label{sec:related}
 
@@
 
+Figure~\ref{fig:metric} summarizes the result of these experiments. The data
+reveal that the information-theoretic metric is critical for finding the correct
+lens: Only 10 of the benchmarks succeeded when running in \CCOpt{} mode. The
+termination condition is also quite important. When running in \FLOpt{} mode,
+the algorithm only discovers 5 lenses, which shows that the first cluster that
+contains a satisfying lens is rarely the correct cluster. However, our algorithm
+is not perfect and fails when either it is very difficult to find the desired
+lens (necessitating \SRequire) or when a large amount of data is projected
+(necessitating \Skip). Without any annotations, our algorithm finds the correct
+lens for 40 of our 48 benchmarks; eight required relevance annotations to find
+the correct lens.
+% end evaluation
+
+% begin related-work
+\section{Related Work}
+\label{sec:related}
+
 In this paper, we have designed and implemented a new, pragmatic
@@
 the appendix
-\ifappendices  \S\ref{sec:appendixforget}. \else{} included with the supplemental
-material.\fi
-
-We also comment on related
+\ifappendices  \S\ref{sec:appendixforget}. \else{}included with the supplemental
+material.  \fi We also comment on related
 program synthesis techniques.
@@
 Intuitively, these equations state that complements are uniquely determined by
-the most recently input $x$ and $y$. Hence, such lenses correspond exactly with
-simple symmetric lenses, where all state is maintained by the current state of
-the $x$ and $y$ data. Forgetful symmetric lenses express exactly the same
-$apply$ function as simple symmetric lenses.
+the most recently input $x$ and $y$. Such lenses correspond exactly with simple
+symmetric lenses, where all state is maintained by the $x$ and $y$ data.
 
@@
 \end{align*}
-Despite being less expressive than symmetric lenses, simple symmetric lenses are
-strictly more expressive than classical asymmetric lenses.
+Simple symmetric lenses are strictly more expressive than classical asymmetric
+lenses.
 
@@
 Bidirectional programming has an extensive literature, with many extensions onto
-basic lens formulations, like quotient lenses\cite{quotientlenses} and matching
-lenses\cite{matchinglenses}. Interested readers can consult
-\citet{DBLP:conf/icmt/CzarneckiFHLST09} for a slightly dated survey of lenses
-and other lens-like structures.
+basic lenses, like quotient lenses~\cite{quotientlenses} and matching
+lenses~\cite{matchinglenses}. Readers can consult
+\citet{DBLP:conf/icmt/CzarneckiFHLST09} for a survey of lenses
+and lens-like structures.
 
 \subsection{Data Transformation Synthesis}
-Over the past decade, the programming languages community has explored
-the synthesis of programs from a wide variety of angles.  One of
-the key ideas is typically to narrow the program search space by focusing
-on a specific domain, and imposing constraints on syntax~\cite{sygus} or
-typing~\cite{augustsson-2004,gvero-pldi-2013,osera+:pldi15,feser-pldi-2015,scherer-icfp-2015,frankle+:popl16}, or both.
+Over the past decade, the programming languages community has explored the
+synthesis of programs from a wide variety of angles. One of the key ideas is
+typically to narrow the program search space by focusing on a specific domain,
+and imposing constraints on syntax~\cite{sygus},
+typing~\cite{augustsson-2004,gvero-pldi-2013,osera+:pldi15,feser-pldi-2015,scherer-icfp-2015,frankle+:popl16},
+or both.
 %, and as has been advocated
@@
 
-All pragmatic synthesis algorithms use heuristics of one kind or another.  One of the
-goals of the current paper is to try to ground those heuristics in a broader theory,
-information theory,
-with the hope that this theory may inform future design decisions and help us understand
-heuristics crafted in other tools and possibly in other domains.  For instance, we
-speculate that some of the heuristics used in
-FlashFill (to take one well-documented example) may be connected to some of the principles
-laid out here.  For instance, Flash Fill prioritizes the substring constructor over the
-constant string constructor when ranking possible programs. Such a choice is
-consistent with our information-theoretic viewpoint as the constant function
-throws away a great deal of information about the source string being
-transformed. Likewise, Flash Fill prefers ``wider'' character classes over
-``narrower'' ones. Again such a choice is implied by information theory -- the
-wider class preserves more information during translation.  More broadly, we
-hope our information-theoretic analysis provides a basis for
-understanding the heuristic choices made in related work.
+All pragmatic synthesis algorithms use heuristics of one kind or another. One of
+the goals of the current paper is to try to ground those heuristics in a broader
+theory, information theory, with the hope that this theory may inform future
+design decisions and help us understand heuristics crafted in other tools and
+possibly in other domains. For instance, we speculate that some of the
+heuristics used in FlashFill (to take one well-documented example) may be
+connected to some of the principles laid out here. For instance, Flash Fill
+prioritizes the substring constructor over the constant string constructor when
+ranking possible programs. Such a choice is consistent with our
+information-theoretic viewpoint as the constant function throws away a great
+deal of information about the source string being transformed. Likewise, Flash
+Fill prefers ``wider'' character classes over ``narrower'' ones. Again such a
+choice is implied by information theory -- the wider class preserves more
+information during translation. More broadly, we hope our information-theoretic
+analysis provides a basis for understanding the heuristic choices made in
+related work.
 

### 2019-02-28 papers/SymmetricLenses/main.tex
add fixes and keep 25 pages

@@
-\newif\ifdraft\drafttrue  % set true to show comments
+\newif\ifdraft\draftfalse  % set true to show comments
 % \newif\ifdraft\draftfalse  % set true to show comments
@@
 Each task consists of a source and target regular expression and a set of
-examples. These tasks can be quite quite large, as we are synthesizing complex
-transformations between real file formats -- some of the lenses and
-specifications are over 500 AST nodes \afm{run the experiments and collect the
-  data at some point}.
+examples. We are synthesizing transformations between real formats, so these
+tasks are can be quite large -- on average, the size of each RE
+in the specification is 1637 and the size of the synthesized lens is 3254.
 

### 2019-02-28 papers/SymmetricLenses/main.tex
turned off drafts
Comments and polishing

@@
-\newif\ifdraft\draftfalse  % set true to show comments
+\newif\ifdraft\drafttrue  % set true to show comments
 % \newif\ifdraft\draftfalse  % set true to show comments
@@
 {\em Lenses} are programs that can be run both ``front to back'' and ``back
-to front,'' allowing updates to both input and output data to be transferred
-in both directions.
+to front,'' allowing updates to either their source or their target data to be
+transferred in both directions. 
 Since their introduction by Foster \ETAL, lenses have been extensively
@@
 
-\hspace*{1em}We show how to extend this synthesis algorithm to a much wider class of
-lenses, which we call {\em simple symmetric lenses}, including not only all
-of the bijective lenses and the more widely used ``asymmetric''
-lenses, but also a rich subset of the full-blown
+\hspace*{1em}We extend this synthesis algorithm to a much broader class of
+lenses, called {\em simple symmetric lenses}, including all
+bijective lenses, all of the popular category of ``asymmetric''
+lenses, and a rich subset of the more powerful
 ``symmetric lenses'' proposed by Hofmann \ETAL  
@@
 on one side but not the other and vice versa.
-Simple symmetric lenses are of independent theoretical interest, being
+They are of independent theoretical interest, being
 the largest class of symmetric lenses that do not rely on persistent
 internal state.
+
 Synthesizing simple symmetric lenses is substantially more challenging than
@@
 lens, generally prefering lenses that propagate more information, 
-as well as user annotations identifying parts of the source and
+as well as user annotations marking parts of the source and
 target data structures as either {\em irrelevant} or {\em essential}.
@@
 file format synchronization tasks. Our implementation can synthesize all these
-lenses in under 30 seconds.
+lenses in under 30 seconds.  \bcp{All of them together take 30 seconds?  Or
+  the maximum time it takes for one is 30 seconds?}
 \end{abstract}
@@
 can transform updates to data represented in a format $S$ to another format
-$T$ and vice versa, providing ``round tripping'' guarantees that help ensure
+$T$ and vice versa, providing round-tripping guarantees that help ensure
 information is not lost or corrupted as it is transformed back and forth
@@
 \textit{Synthesizing} lenses can make the task even easier. Indeed, lens
-languages, like Boomerang, are excellent targets for search-based synthesis, since their
-specialized, sub-Turing-complete syntax and expressive type systems drastically
-limit the space of possible programs.
-%
-The recent Optician~\cite{optician} enhancement to Boomerang has
-demonstrated that it is feasible to synthesize quite complex examples, albeit
+languages like Boomerang are excellent targets for search-based synthesis,
+since their 
+specialized, sub-Turing-complete syntax and their expressive type systems
+drastically limit the space of possible programs.
+%
+The recent Optician synthesizer for Boomerang~\cite{optician} has
+demonstrated the feasibility of synthesizing quite complex examples, albeit
 only for a restricted class of lenses---so-called bijective lenses.
@@
 very many bijections between a given pair of data formats. 
-If the synthesis algorithm can
-find any bijection, it is fairly likely to be the intended one.
-
-However, the set of real-world use cases for bijective lenses is also
-somewhat limited.  A more common situation is where two real-world data
+If the synthesis algorithm finds any bijection, it is fairly likely to be
+the intended one.
+
+However, the set of real-world use cases for bijective lenses, where two
+data formats contain different arrangements of precisely the same
+information, is  
+limited.  A more common situation is where two data
 formats share just {\em some} of their information content.  For instance,
 two related database tables might share four out of five columns, but each
-might have a different fifth column.  Or, one ad-hoc system configuration
+might have a different fifth column.  Or one ad-hoc system-configuration
 file might include some format-specific metadata, such as a date or a
-reference number, whereas the same configuration file on another operating
+reference number, while the same configuration file on another operating
 system does not.  Indeed among the 48 benchmark problems described in
@@
 two ad hoc file formats have this characteristic, 13 benchmarks in
-total.
+total.  \bcp{I don't understand this.  We say that the non-bijective
+  situation is ``more common,'' but we support this by saying that 13 out of
+48 benchmarks (that we chose!) are non-bijective.  Or am I misreading??}
 

### 2019-02-28 papers/SymmetricLenses/main.tex
respond to comments and update intro and abstract

@@
 We describe an implementation of simple symmetric lenses and our synthesis
-procedure as extensions to the Boomerang language. We evaluate
-its performance on 48 benchmark examples drawn from Flash Fill,
-Augeus, the bidirectional programming literature, and electronic
-file format synchronization tasks. Our implementation can synthesize all these
-lenses in under 30 seconds.  \bcp{All of them together take 30 seconds?  Or
-  the maximum time it takes for one is 30 seconds?}
+procedure as extensions to the Boomerang language. We evaluate its performance
+on 48 benchmark examples drawn from Flash Fill, Augeus, the bidirectional
+programming literature, and electronic file format synchronization tasks. Our
+implementation can synthesize each of these lenses in under 30 seconds.
 \end{abstract}
@@
 information, is  
-limited.  A more common situation is where two data
+limited.  Oftentimes, two data
 formats share just {\em some} of their information content.  For instance,
@@
   situation is ``more common,'' but we support this by saying that 13 out of
-48 benchmarks (that we chose!) are non-bijective.  Or am I misreading??}
+48 benchmarks (that we chose!) are non-bijective.  Or am I misreading??} \afm{I
+think this is due, in part, to where we got the benchmarks (Augeas typically removes
+little information). I do think that
+it is more common, but I suppose I have no real data on that. I've updated
+the wording to be ``safer''.}
 

### 2019-02-28 papers/SymmetricLenses/main.tex
More comments

@@
 Can Optician's basic synthesis procedure be extended to a richer class of
-lenses? In particular, what about \emph{symmetric
+lenses? Could we even imagine synthesizing all \emph{symmetric
   lenses}~\cite{symmetric-lenses}---a much larger class that includes
@@
 
-The first is pragmatic. ``Classical'' symmetric lenses as presented
+The first of these is pragmatic. ``Classical'' symmetric lenses as presented
 by ~\citet{symmetric-lenses} operate
-over three structures: the ``left'' structure $X$, the ``right'' $Y$ and a complement
+over three structures: a ``left'' structure $X$, a ``right'' structure
+$Y$ and a ``complement'' 
 $C$ that contains the information not
@@
 restricted variant 
-of symmetric lenses, which we call \emph{simple symmetric lenses}. 
+of symmetric lenses,  called \emph{simple symmetric lenses}. 
 %only the usual source and target.
@@
 %, which we define in Section~\ref{sec:relationship}.
-Intuitively, simple symmetric lenses are the complete set of symmetric
+Intuitively, simple symmetric lenses are symmetric
 lenses that do not require external ``memory'' to recover data from past
@@
 
-To characterize their expressive power formally, we show that simple
-symmetric lenses are exactly symmetric lenses that satisfy an intuitive
-property called \emph{forgetfulness}.  We show they are expressive
-enough in practice by adding simple symmetric lenses to the Boomerang
-language~\cite{boomerang} and applying them to a range of real-world
-applications, showing that simple symmetric lenses integrate
-smoothly with other advanced lens features provided by Boomerang, including
-quotient lenses~\cite{quotientlenses,maina+:quotient-synthesis} and matching
-lenses~\cite{matchinglenses}.
-
-This leaves us with just the second difficulty with synthesizing symmetric
-lenses: While the number of bijective lenses between two given formats is
+Formally, we characterize the expressiveness of simple symmetric lenses by
+proving that they are exactly the symmetric lenses
+that satisfy an intuitive property called \emph{forgetfulness}.  We also show
+they are expressive enough for many practical uses by adding simple
+symmetric lenses to 
+the Boomerang language~\cite{boomerang} and applying them to a range of
+real-world applications.  This exercise also demonstrates that simple
+symmetric lenses integrate smoothly with other advanced lens features
+provided by Boomerang, including {\em quotient
+  lenses}~\cite{quotientlenses,maina+:quotient-synthesis} and {\em matching
+  lenses}~\cite{matchinglenses}.\bcp{How does it demonstrate this?  Did we
+  actually {\em use} symmetric quotient or matching lenses for anything?  Or
+are we just observing that these features can coexist in the same
+implementation even if using them together might not make sense?}
+
+This leaves us with the second difficulty in synthesizing symmetric
+lenses: Whereas the number of bijective lenses between two given formats is
 typically tiny, the number of simple symmetric lenses is typically enormous.
@@
 sophisticated synthesis algorithm that uses this principle to search the
-space more intelligently. For these, we turn to information theory.
-
-We consider ``likely'' lenses to be ones that propagate ``a lot'' of information
-from the left data format to the right and vice versa. Conversely, unlikely
-lenses require a large amount of additional information to recover
-either of the formats given the other. For many application scenarios, users
+space more intelligently.
+
+For these, we turn to information theory. We consider ``likely'' lenses to
+be ones that propagate ``a lot'' of information 
+from the left data format to the right and vice versa. Conversely, ``unlikely''
+lenses are ones that require a large amount of additional information to recover
+one of the formats given the other. \bcp{Next two sentences are not very
+  persuasive.  And they may be a bit out of date anyway, not that we have
+  these annotations.  We can just say that we take ``propagate as much as
+  possible'' as the default and use explicit annotations on the source and
+  target types to mark information that the user knows should not be propagated.}For many application scenarios, users
 will want to synchronise formats with lots of shared information (and lenses
 with little shared information are typically easier to write). Our tool is built
-primarily for these complex scnearios, and so lenses that propagate more
+primarily for these complex scenarios, and so lenses that propagate more
 information are preferable in our context to ones that propagate less.
@@
 
-With simple symmetric lenses and our SRE-based likelihood measure in hand, we
-develop a new algorithm for synthesizing likely lenses. At its core, the
+With simple symmetric lenses and this SRE-based likelihood measure in hand, we
+propose a new algorithm for synthesizing likely lenses. At its core, the
 algorithm performs a type-directed search between the descriptions of the source
 and target formats, measuring success using the likelihood measure.
+
 Interesting complications arise from the need to deal with regular-expression

### 2019-02-28 papers/SymmetricLenses/main.tex
More comments

@@
 two ad hoc file formats have this characteristic, 13 benchmarks in
-total.  \bcp{I don't understand this.  We say that the non-bijective
-  situation is ``more common,'' but we support this by saying that 13 out of
-48 benchmarks (that we chose!) are non-bijective.  Or am I misreading??} \afm{I
-think this is due, in part, to where we got the benchmarks (Augeas typically removes
-little information). I do think that
-it is more common, but I suppose I have no real data on that. I've updated
-the wording to be ``safer''.}
+total.
 

### 2019-02-28 papers/SymmetricLenses/main.tex
react to feedback and make safer statements

--- 2019-02-28 papers/SymmetricLenses/main.tex
@@
 real-world applications.  This exercise also demonstrates that simple
-symmetric lenses integrate smoothly with other advanced lens features
+symmetric lenses can coexist with, and be extended by, other advanced lens features
 provided by Boomerang, including {\em quotient
@@
 implementation even if using them together might not make sense?}
+\afm{Good point, updated.}
 
@@
 lenses are ones that require a large amount of additional information to recover
-one of the formats given the other. \bcp{Next two sentences are not very
-  persuasive.  And they may be a bit out of date anyway, not that we have
-  these annotations.  We can just say that we take ``propagate as much as
-  possible'' as the default and use explicit annotations on the source and
-  target types to mark information that the user knows should not be propagated.}For many application scenarios, users
-will want to synchronise formats with lots of shared information (and lenses
-with little shared information are typically easier to write). Our tool is built
-primarily for these complex scenarios, and so lenses that propagate more
-information are preferable in our context to ones that propagate less.
-%
-This intuition is formalized using \emph{stochastic regular expressions}
+one of the formats given the other. By default, our synthesis algorithm prefers
+the lenses that propagate more information.
+This preference is formalized using \emph{stochastic regular expressions}
 (SREs)~\cite{stoch-rnn,stoch-def}, which simultaneously define a set of

### 2019-02-28 papers/SymmetricLenses/main.tex
remove comments

@@
   lenses}~\cite{quotientlenses,maina+:quotient-synthesis} and {\em matching
-  lenses}~\cite{matchinglenses}.\bcp{How does it demonstrate this?  Did we
-  actually {\em use} symmetric quotient or matching lenses for anything?  Or
-are we just observing that these features can coexist in the same
-implementation even if using them together might not make sense?}
-\afm{Good point, updated.}
+  lenses}~\cite{matchinglenses}.
 

### 2019-02-28 papers/SymmetricLenses/main.tex
More comments

@@
 Fill~\cite{flashfill} and many of the benchmarks that synchronize between
-two ad hoc file formats have this characteristic, 13 benchmarks in
-total.
+two ad hoc file formats have this characteristic (13 benchmarks in
+total).  \bcp{The safer wording is better, but the reader is still left
+  wondering ``If they are demonstrating a system for synthesizing
+  non-bijective lenses, why did they choose a benchmark suite where 35/48
+  examples are bijective??''}
 
@@
 
-One might hope that extending the algorithms from \citet{optician} to
+One might first hope that extending the algorithms from \citet{optician} to
 synthesize symmetric rather than bijective lenses would be relatively
@@
 
-The first of these is pragmatic. ``Classical'' symmetric lenses as presented
-by ~\citet{symmetric-lenses} operate
+The first of these is pragmatic.  Symmetric lenses as presented
+by~\citet{symmetric-lenses} operate
 over three structures: a ``left'' structure $X$, a ``right'' structure
@@
 depend on which of the possible representations are chosen for its source and
-target formats. Moreover, if Boomerang's type-equivalence rule is disallowed,
-there may be {\em no} well-typed lens connecting the source and target formats
+target formats. \bcp{There's a disconnect here: The reader was probably not
+  thinking about Boomerang's type-equivalence rule (indeed, very likely, did
+  not even know there was such a thing).}Moreover, if Boomerang's type-equivalence rule is disallowed,
+there may be {\em no} well-typed lens connecting a particular pair of source
+and target formats 
 selected by the user: in general, the synthesis algorithm has to search through
@@
 thereby forcing certain data to either be retained or discarded during the
-transformations. We evaluated our system on a benchmark suite of 48 problems, including data-cleaning
-tasks taken from Flash Fill~\cite{flashfill}, view-maintenance tasks
-taken from Augeas~\cite{augeas2}, and ad-hoc synchronization tasks from a
-variety of other sources (mostly from the bidirectional programming literature).
+transformations.
+
+We evaluated our system on a benchmark suite of 48 problems, including
+data-cleaning tasks taken from Flash Fill~\cite{flashfill}, view-maintenance
+tasks taken from Augeas~\cite{augeas2}, and ad-hoc synchronization tasks
+from a variety of other sources (mostly from the bidirectional programming
+literature).  \bcp{``... and we found...''?}  \bcp{Also, this paragraph is
+  redundant with the corresponding contribution a bit below.  I think it
+  would improve the flow if we made this one a bit longer (and stronger) and
+  the contribution paragraph shorter, just summarizing what this one says.}
 

### 2019-02-28 papers/SymmetricLenses/main.tex
Finished with section 1

@@
 
-In summary, our contributions are:
+In summary, our contributions are as follows:
 \begin{itemize}
@@
 \item We define {\em simple symmetric lenses}, a subclass of symmetric
-lenses that do not maintain any hidden state (\S\ref{sec:overview}).  
+lenses with no hidden state,
 %
-We present a collection of combinators for building simple symmetric lenses and
-describe how to calculate their likelihoods (\S\ref{sec:ssl}).
+we propose a syntax for them---a collection
+of combinators for building simple symmetric lenses---and we
+describe how to calculate their likelihoods
+(\S\ref{sec:overview}--\S\ref{sec:ssl}).  
 \item We give an algorithm for synthesizing lenses built from these combinators,
@@
   search process (\S\ref{sec:synthesis}) and prove that our metrics have deep
-  semantic meaning.
+  semantic meaning. \bcp{``Semantic meaning'' is redundant.  And ``deep
+    semantic meaning'' just sounds silly.  Can we say it more precisely?}
 %
-\item We extend the Boomerang implementation with simple symmetric lenses,
-  demonstrating that they integrate smoothly with other lens extensions such as
-  quotient lenses~\cite{quotientlenses} and matching
-  lenses~\cite{matchinglenses}.
-
-\item We also extend Boomerang with symmetric lens synthesis specifications:
-  including types, examples, and relevance annotations. We evaluate our implementation, the effects of
+\item We extend the Boomerang implementation with simple symmetric lenses
+and lens synthesis specifications, 
+  including types, examples, and relevance annotations. We evaluate our
+  implementation, the effects of
   optimizations, and our inclusion of relevance annotations on a set of
@@
 %
-\item We show that simple symmetric lenses fall between asymmetric
+\item We prove that simple symmetric lenses fall between asymmetric
 lenses~\cite{Focal2005-long} and full symmetric
@@
 Along the
-way, we show the {\em star-semiring equivalences} on regular expressions
+way, we show that {\em star-semiring equivalences}\bcp{citation} on regular
+expressions 
 remain valid when extended to stochastic regular expressions, preserving
@@
 \end{theorem}
-\begin{proof}
-  See the appendix %
-\ifappendices  Theorem~\ref{thm:semantics-correctness} \else{} included with the supplemental
-material\fi.
-\end{proof}
-
-This theorem will come in handy as we traverse (with \Expand) or normalize
-across (as part of \GreedySynth) star-semiring equivalences.
+
+(For the proof, see the appendix \ifappendices
+Theorem~\ref{thm:semantics-correctness} \else{} included with the
+supplemental material\fi.)  This theorem will come in handy as we traverse
+(with \Expand) or normalize across (as part of \GreedySynth) star-semiring
+equivalences.
 
@@
 \subsection{Lens Costs}
-\label{subsec:lenscosts}
+\label{subsec:lenscosts} \bcp{In the intro, we talked about ``likelihood.''
+Now we seem to be switching the terminology to ``cost.''}
 Our cost metric is based on the expected amount of information required to

### 2019-02-28 papers/SymmetricLenses/main.tex
More comments and polish

@@
 \subsection{Simple Symmetric Lenses}
-Semantically, a simple symmetric lens (or lens when clear from context) between
-sets $X$ and $Y$ comprises four functions subject to four ``round-tripping''
+Semantically, a simple symmetric lens (we will just say ``lens'' when clear
+from context) between 
+sets $X$ and $Y$ comprises four functions subject to four round-tripping
 laws.
-
+\vspace*{.7ex}
+\begin{trivlist}
+\item 
 %\begin{enumerate}
@@
 \end{center}
-\noindent
-\vskip -.5em
-The two \KW{create} functions are used to propagate shared information and fill
+\end{trivlist}
+The two \KW{create} functions are used to \proposedelete{propagate shared
+  information and} fill 
 in default values when introducing new data (\EG, on \KW{create} the ``unk''
@@
 the other by combining a new value from one with an old value from the other.
-The notation $\Lens.\PutR$ extracts the \PutR function from the
-lens $\Lens$; similarly for \PutL. These four functions can be used to keep the
-common information between two file types in synch. For example, if a new file
-of the left-hand format is created, the \CreateR function will a new file in the
-right-hand format. If the right-hand format is edited, the \PutL function can
-update the left-hand format with the changed information.
+The record projection notation $\Lens.\PutR$ extracts the \PutR function
+from the 
+lens $\Lens$. These four functions can be used to keep the
+common information between two file types in sync. For example, if a new file
+of the left-hand format is created, the \CreateR function will build a new
+file in the 
+right-hand format. If the right-hand file is then edited, the \PutL function can
+update the left-hand file with the changed information.
 
@@
 side (the result of a put or create) back through a lens in the other direction
-will get back to exactly where we began. For example, consider a simple
-symmetric lens
+will get back to exactly where we began. For example, consider a lens
 synchronizing the employee data formats in
@@
 
-This simple symmetric lens definition differs from classical symmetric lenses in
-that it does not involve a \emph{complement}. We give a detailed comparison in
+Simple symmetric lenses differ from ''classical'' symmetric lenses in that
+they do not involve a \emph{complement}. We give a detailed comparison in
 \S\ref{sec:relationship}.
@@
 
-In principle, programmers can define simple symmetric lenses just by writing
+In principle, programmers can define lenses just by writing
 four arbitrary functions by hand and showing they satisfy the round-tripping
-laws. But a better idea is to streamline the process by providing a set of
-combinators---a domain-specific language---for building complex, law-abiding
+laws, but this can be tedious and error prone.  A better idea is to provide
+a set of combinators---a domain-specific language---for building complex,
+law-abiding 
 lenses from simpler ones. Such bidirectional languages have been given for
-building lenses over many different kinds of data, including database
+many different kinds of data, including database
 relations~\cite{BohannonPierceVaughan} algebraic data
 types~\cite{symmetric-lenses}, graphs~\cite{tgg}, and more.
-
 We focus here on combinators for defining lenses between string data formats
 described by regular expressions.
+
 If $\BRegex$ and $\BRegexAlt$ are regular expressions, then
@@
 $\LanguageOf{\BRegexAlt}$. (We will use undecorated variables later for
-stochastic regular expressions; we mark plain REs with overbars.)
+stochastic regular expressions, so we mark plain REs with overbars.)
 We illustrate some of these
@@
 The simplest combinator is the identity lens \IdentityLens, which takes as an
-argument a regular expression $\BRegex$ and performs an identity mapping on data
-matching $\BRegex$. 
+argument a regular expression $\BRegex$ and propagates data unchanged
+in both directions.
 %
@@
 HR file.
-For example, with the above lens, $\PutL \App \lstinline{""} \App 60000$ will return 60000,
+With the above lens, $\PutL \App \lstinline{""} \App 60000$ will return 60000,
 and \PutR will always return \lstinline{""}.
@@
 ($\OrLensOf{\Lens_1}{\Lens_2}$), and iteration ($\IterateLensOf{\Lens}$).
-For example, we could use the following lens
+For example, we could use the lens
 %
@@
 table. For example given a
-lens \lstinline{employee_lens} that synchronizes data for a single employee the
-following lens
+lens \lstinline{employee_lens} that synchronizes data for a single employee, the
+lens\bcp{In all these examples, I'd like to have one more space on each side
+of the ``:''}
 %
@@
 transforms a list of employees in employees in the Management format to a list
-of employees in the HR format in the left-to-right direction. These combinators
+of employees in the HR format and vice versa. These combinators
 are combined in figure~\ref{fig:example_lens} to construct a complete lens
@@
 
-Most of the typing rules for lenses are syntax driven, where only a handful of
-rules result in a given type.  For example, the type of a
+\bcp{I wonder whether the rest of this subsection can be cut... It is both
+  hard to understand and too brief to really say much, and I wonder whether
+  we need it for anything between here and where we give it a real
+  explanation.}
+
+Most of the typing rules for lenses are syntax directed, where only a handful of
+rules result in a given type\bcp{syntax directed usually means that the
+  conclusion of at most
+  {\em one} rule will match a given goal; if that's not the case here, then
+  we need to speak more carefully}.  For example, the type of a
 \ConcatLens lens is the concatenation of the regular expressions of the sub-lenses.
@@
 \]
-With one notable exception, lens typing rules maintain structural similarity
-between regular expression types and lens terms; type equivalence does not. Type
-equivalence permits a well-typed lens to be well-typed among any equivalent
-regular expression pair.
+The typing  rules for maintain structural similarity
+between regular expression types and lens terms, with one exception: The
+type equivalence rule permits a well-typed lens to be well-typed among any
+equivalent regular expression pair.
 %

### 2019-02-28 papers/SymmetricLenses/main.tex
More comments and polish

@@
 
-Even for small examples, lenses like the one above take a little getting used
-to. For larger and more complex formats, it can become quite difficult to
-write lenses by hand.  Instead, we wish to synthesize them from examples. 
-That is, given a pair of regular expressions and a set of input-output examples, we want
+For larger and more complex formats, it can become quite difficult to
+write lenses by hand, as we did above.  Instead, we would like to synthesize
+them from types and examples.  
+That is, given a pair of regular expressions and a set of input-output
+examples, we want 
 to find a simple symmetric lens that is typed by the regular expression pair and
-satisfies all the input/output examples. We call such a lens a \emph{satisfying
-  lens}. In our running example, we wish to synthesize a lens between
-\lstinline{emp_salaries} and \lstinline{emp_insurance}, using as an input-output
-example the data in Figure~\ref{fig:minimized-representations}.
+satisfies all the input/output examples. We this a \emph{satisfying
+  lens}. For our running example, suppose we wish to synthesize a lens between
+\lstinline{emp_salaries} and \lstinline{emp_insurance}, using as sole
+input-output example the data in Figure~\ref{fig:minimized-representations}.
 
@@
 Figure~\ref{fig:minimized-representations} (where \lstinline{...} represents
-components of the example itself). In general, {\em many} examples may be
-required to rule out all possible occurrences of nested \Disconnect lenses,
+components of the example itself\bcp{huh?}). In general, {\em many} examples
+may be 
+required to rule out all possible occurrences of nested\bcp{??} \Disconnect lenses,
 particularly in complex formats.
@@
 %and is highlighted even more on larger formats.
-While we could ask programmers to provide detailed and precise specifications
+\proposedelete{While we could ask programmers to provide detailed and precise specifications
 for their synthesis tasks, we would prefer an algorithm that can intelligently
-choose a lens despite the underspecification. Instead of merely finding
+choose a lens despite the underspecification.} Instead of merely finding
 \emph{any} satisfying lens, we wish to synthesize a satisfying lens that is
@@
 A satisfying lens is ``more likely'' if it uses more data from one format to
-construct the other. One manifestation of this principle involves choosing the
-identity lens (which uses all the data) rather than the disconnect lens (which
-uses none of the data). We define the likelihood of a satisfying lens as the
-expected number of bits required to recover data in one format from synchronized
-data in the other. Higher likelihoods correspond to fewer bits. Two
-strings $s$ and $t$ are \emph{synchronized according to lens $l$} if $l.\PutR\
+construct the other.  For example, the
+identity lens (which uses all the data) is more likely than the disconnect
+lens (which uses none). Formally, we define the {\em likelihood} of a
+satisfying lens as the 
+expected number of bits required to recover data in one format from 
+data in the other; higher likelihoods correspond to fewer bits. Two
+strings $s$ and $t$ are \emph{synchronized according to lens
+  $l$}\bcp{Earlier we used $\Lens$, now we are switching to $l$...} if $l.\PutR\
 s\ t = t$ and $l.\PutL\ t\ s = s$. We can \emph{recover} $s$ from 
@@
 \Disconnect{} lens, we need enough bits to fully encode $s$ in order to recover
-it from $t$ because $t$ contains no useful information. Thus, if both
+it from $t$ because all the information in $t$ gets thrown away. Thus, if both
 \IdentityLens{} and \Disconnect{} are satisfying lenses for a particular pair of
-formats, then \IdentityLens{} will be more likely.
+formats, \IdentityLens{} will be more likely.
 

### 2019-02-28 papers/SymmetricLenses/main.tex
Finished with polishing 1 and 2. Lots of comments. Switching to other papers now, but I will be available for discussion.

@@
 $S$, the entropy $\EntropyOf{S,P}$ is $-\Sigma_{s\in S}P(s)\Log_2P(s)$. In
-information theory, each element of $s \in S$ has \emph{information content},
-the number of bits required to specify that element in a perfect encoding scheme
-($-\Log_2P(s)$). The entropy of $S$ calculates the expected number of bits
-required to specify an element drawn from $S$, in other words, the
-probability of each element times its information content. Entropy maintains
-that if a data source contains many possible elements, where none have
+information theory, the \emph{information content} of each element of $s \in
+S$ is 
+the number of bits required to specify $s$ in a perfect encoding scheme
+($-\Log_2P(s)$). The entropy of $S$ is the expected number of bits
+required to specify an element drawn from $S$---the
+probability of each element times its information content. Entropy captures
+the intuition that, if a data source contains many possible elements and
+none have 
 significantly higher probability than others, it will have high information
-content. Data sources with few elements that don't have vanishingly small
-probabilities have small entropy. When $P$ is clear, we will often use the
-shorthand of $\EntropyOf{S}$ for $\EntropyOf{S,P}$.
-
-In our setting, we already have a way of expressing sets of data: regular
+content. Data sources where most elements have small
+probabilities have small entropy. When $P$ is clear from context, we will
+often use the shorthand of $\EntropyOf{S}$ for $\EntropyOf{S,P}$.
+
+In the present setting, we already have a way of expressing sets of data:
+regular 
 expressions. To calculate entropy, what we need is a way to express probability
-distributions over that data. To that end, we adopt \emph{stochastic regular
+distributions over those sets. To that end, we adopt \emph{stochastic regular
   expressions}~\cite{stoch-def,stoch-rnn} (SREs), which are regular expressions
 in which each operator is annotated with a probability (see \S\ref{sec:sre}). A
-stochastic regular expression specifies both a set of strings and a probability
-distribution over that set.
-
-We use entropy to guage the relative likelihood of such lenses in our synthesis
+stochastic regular expression thus specifies both a set of strings at the
+same time as a probability distribution over that set.
+
+We use entropy to gauge the relative likelihood of lenses in our synthesis
 algorithm (see \S\ref{subsec:lenscosts}). For any lens $\Lens$, we can calculate
@@
 conditional entropy of $\RegexAlt$ given $\Regex$ and $\Lens$. This metric
-assigns higher entropy (or cost) to lenses where knowing the string on one side
+assigns higher entropy (or cost\bcp{= lower likelihood?  I'm still confused
+  by these different ways of speaking about what seems to be the same
+  thing.}) to lenses where knowing the string on one side 
 provides little information about the string on the other side. It assigns zero
@@
 
-To obtain SREs from the plain regular expressions that users write, by default
-we use a heuristic that attempts to assign probability annotations that give
+To obtain SREs from the plain regular expressions that users write, 
+we use a default heuristic that attempts to assign probability annotations
+giving
 each string in the language equal probablity. In other words, this heuristic
-tries to make each bit of data no more or less likely than any other.
-Our synthesis algorithm could equally well take as input a pair of stochastic
-regular expression rather than a plain one, permitting users to specify a
+tries to make each bit of data no more or less likely than any
+other.\bcp{huh?} 
+Alternatively, our synthesis algorithm could take as input a pair of stochastic
+regular expressions rather than plain ones, permitting users to specify a
 data-specific probability distribution (either manually calculated or inferred
-from some dataset of interest).
-
-However, sometimes users already know that certain data should or should not be
+from some dataset of interest).\bcp{So why didn't we do it this way?}
+
+However\bcp{``However'' means that something from before is being
+  contradicted.  What is it in this case?}, sometimes users already know that certain data should or should not be
 used to construct the other format. We introduce relevance annotations to SREs
@@
 \lstinline{emp_ins} be annotated as required (as they are converted with the
-identity lens).
+identity lens).  \bcp{That paragraph seems way too brief and high level.
+  These annotations were added in response to significant complaints from an
+earlier PC, but the paper reads as though they are an afterthought of little
+consequence; they deserve more attention and explanation.}
 
@@
 likely lenses. Previous work~\cite{augustsson-2004, osera+:pldi15,
-  feser-pldi-2015, frankle+:popl16} has shown that we can use types to
+  feser-pldi-2015, frankle+:popl16} has shown that types can be used to
 dramatically restrict the search space and improve the effectiveness of
 synthesis. In our setting, types are pairs of stochastic regular
-expressions. The fact that each regular expression is semantically equivalent to
-infinitely many others complicates the problem.
+expressions, and the fact that each regular expression is semantically
+equivalent to infinitely many others complicates the problem.
+
 Following existing work on bidirectional program
@@
 split our algorithm into two communicating
-searches. The first search,
+searche procedures. The first,
 \Expand, navigates the space of semantically equivalent regular expressions
-by applying rewrite rules that preserve both the semantics and the
-probability distributions of those stochastic regular expressions. This
+by applying rewrite rules that preserve both semantics and
+probability distributions. This
 algorithm ranks pairs of stochastic regular expressions by the number of
@@
 
-Abstractly, \GreedySynth looks for highest-likelihood lenses between a given
+\GreedySynth looks for highest-likelihood lenses between a given
 pair of stochastic regular expressions $\Regex$ and $\RegexAlt$ by performing a
-type-directed search. In more detail, \GreedySynth first converts the stochastic
+type-directed search. It first converts the stochastic
 regular expressions provided by \Expand into \emph{stochastic DNF regular
@@
 with disjunctions distributed over concatenations and with concatenations and
-disjunctions normalized to operating over lists (see \S\ref{subsec:sdnfre}).
-\GreedySynth then uses the syntax of these n-ary DNF regular expressions to find
-normalized lenses, in a form we call \emph{simple symmetric n-ary DNF lenses}.
-These lenses include neither a composition operator nor a type equivalence rule.
+disjunctions normalized to operating over lists.
+Then it uses the syntax of these n-ary DNF regular expressions to find
+normalized lenses in a form we call \emph{simple symmetric n-ary DNF lenses}.
+These involve neither a composition operator nor a type equivalence rule.
 These restrictions mean that there are only a few simple symmetric n-ary DNF
-lenses that are well-typed between each pair of stochastic n-ary DNF regular
-expressions, hence \GreedySynth's search space for a given pair of regular
+lenses that are well typed between a given pair of stochastic n-ary DNF regular
+expressions, so \GreedySynth's search space for a given pair of regular
 expressions is finite. Finally, \GreedySynth yields a simple symmetric lens by
-converting the n-ary forms into the binary forms provided in the surface
+converting the n-ary syntax back to the binary forms provided in the surface
 language.
@@
 search through them, but it poses a problem: when should \Expand stop proposing
-new SRE pairs? We may have found a promising lens between a pair of stochastic
+new SRE pairs? We might have found a promising lens between a pair of stochastic
 regular expressions, but a different pair we haven't yet discovered may give
@@
 promising (for example, if the satisfying lens loses no information, the
-algorithm should terminate for no other lens will be more likely). The
+algorithm should terminate for no other lens will be more likely\bcp{this
+  case is 100\% obvious, which makes it not so interesting as an example.
+  The reader is left wondering what happens with pretty likely but not
+  maximally likely lenses?}). The
 heuristics also favor stopping if \Expand has delivered to \GreedySynth all
@@
 result, they can either refine the search by supplying additional examples,
-which serve both to rule out the previously proposed lens(es) and to reduce the
+which serve both to rule out previously proposed lenses and to reduce the
 size of the search space by cutting down on the number of satisfying lenses, or
@@
 certain information either must be retained or must be discarded by the lens
-(see \S\ref{subsec:optimizations}).
+(see \S\ref{subsec:optimizations}). \bcp{Again, I feel like we're
+  downplaying this aspect too much.}
 

### 2019-02-28 papers/SymmetricLenses/main.tex
Couple more comments

@@
   \vspace{-4ex}
-  \caption{Number of benchmarks that can be solved by a given algorithm in a
+  \caption{\bcp{What are the little tick marks on the top of the figure?
+      And why do we need a line for ``benchmark count''?}Number of
+    benchmarks that can be solved by a given algorithm in a 
     given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
@@
 \label{sec:conc}
-We have developed a synthesis algorithm for synthesizing synchronization
-functions between data formats that may not be in bijective correspondence. More
-specifically, we identify a subset of symmetric lenses, simple symmetric lenses,
+We described a synthesis algorithm for synthesizing synchronization
+functions between data formats that may not be in bijective correspondence.
+We identified a subset of symmetric lenses, simple symmetric lenses, 
 develop new combinators for them, and show how to synthesize them from regular
@@
 synthesis.
+
+\bcp{Future work??????????????}
 

### 2019-02-28 papers/SymmetricLenses/main.tex
minor changes

--- 2019-02-28 papers/SymmetricLenses/main.tex
@@
 formats share just {\em some} of their information content.  For instance,
-two related database tables might share four out of five columns, but each
-might have a different fifth column.  Or one ad-hoc system-configuration
+one ad-hoc system-configuration
 file might include some format-specific metadata, such as a date or a
@@
 Fill~\cite{flashfill} and many of the benchmarks that synchronize between
-two ad hoc file formats have this characteristic (13 benchmarks in
-total).  \bcp{The safer wording is better, but the reader is still left
-  wondering ``If they are demonstrating a system for synthesizing
-  non-bijective lenses, why did they choose a benchmark suite where 35/48
-  examples are bijective??''}
+two ad hoc file formats have this characteristic.
 
@@
 then be restored when transferring from target to source), and even more flexibile
-transformations that allow each side to retain information not present in
-the other?
+transformations that allow each side to throw away information?
 
@@
 If a na\"ive search algorithm just selects the first simple symmetric lens
-it finds, this will generally \emph{not} be the one the user wanted.  We
+it finds, the returned lens will generally \emph{not} be the one the user wanted.  We
 need a new principle for identifying ``more likely'' lenses and a more
@@
 propose a new algorithm for synthesizing likely lenses. At its core, the
-algorithm performs a type-directed search between the descriptions of the source
-and target formats, measuring success using the likelihood measure.
+algorithm performs a type-directed search between descriptions of the data
+formats, measuring success using the likelihood measure.
 
@@
 depend on which of the possible representations are chosen for its source and
-target formats. \bcp{There's a disconnect here: The reader was probably not
-  thinking about Boomerang's type-equivalence rule (indeed, very likely, did
-  not even know there was such a thing).}Moreover, if Boomerang's type-equivalence rule is disallowed,
-there may be {\em no} well-typed lens connecting a particular pair of source
-and target formats 
-selected by the user: in general, the synthesis algorithm has to search through
+target formats. Moreover, certain lenses may not be well-typed unless the
+format representations are replaced by equivalent ones:
+in general, the synthesis algorithm has to search through
 equivalent regular expression types to find the most likely lens. To tame
@@
 effectiveness on a range of applications. Users set up a Boomerang synthesis
-task by providing the source and target types and optionally supplying one or
-more examples. In addition, users can override the default mechanism for
+task by providing two regular expressions and optionally supplying input-output
+examples. Users can also override the default mechanism for
 calculating the information content of a SRE by
 asserting that certain strings are \emph{essential} or \emph{irrelevant},
-thereby forcing certain data to either be retained or discarded during the
+forcing certain data to either be retained or discarded during the
 transformations.
 
-We evaluated our system on a benchmark suite of 48 problems, including
-data-cleaning tasks taken from Flash Fill~\cite{flashfill}, view-maintenance
-tasks taken from Augeas~\cite{augeas2}, and ad-hoc synchronization tasks
-from a variety of other sources (mostly from the bidirectional programming
-literature).  \bcp{``... and we found...''?}  \bcp{Also, this paragraph is
-  redundant with the corresponding contribution a bit below.  I think it
-  would improve the flow if we made this one a bit longer (and stronger) and
-  the contribution paragraph shorter, just summarizing what this one says.}
-
-In summary, our contributions are as follows:
+We evaluate our implementation, the effects of optimizations, and our inclusion
+of relevance annotations on a set of 48 lens synthesis benchmarks drawn from
+data cleaning, view maintenance, and file synchronization tasks. We find the
+system can synthesize simple symmetric lenses for all of the benchmarks in under
+30 seconds (\S\ref{sec:evaluation}).
+
+In summary, our contributions are:
 \begin{itemize}
@@
 \item We define {\em simple symmetric lenses}, a subclass of symmetric
-lenses with no hidden state,
-%
-we propose a syntax for them---a collection
-of combinators for building simple symmetric lenses---and we
-describe how to calculate their likelihoods
-(\S\ref{sec:overview}--\S\ref{sec:ssl}).  
+lenses that do not maintain any hidden state (\S\ref{sec:overview}).  
+%
+We present a collection of combinators for building simple symmetric lenses and
+describe how to calculate their likelihoods (\S\ref{sec:ssl}).
 \item We give an algorithm for synthesizing lenses built from these combinators,
@@
   search process (\S\ref{sec:synthesis}) and prove that our metrics have deep
-  semantic meaning. \bcp{``Semantic meaning'' is redundant.  And ``deep
-    semantic meaning'' just sounds silly.  Can we say it more precisely?}
-%
-\item We extend the Boomerang implementation with simple symmetric lenses
-and lens synthesis specifications, 
-  including types, examples, and relevance annotations. We evaluate our
-  implementation, the effects of
-  optimizations, and our inclusion of relevance annotations on a set of
-  48 lens synthesis benchmarks drawn from data cleaning, view maintenance, and
-  file synchronization tasks. The system can synthesize simple symmetric lenses
-  for all of the benchmarks in under 30 seconds (\S\ref{sec:evaluation}).
-%
-\item We prove that simple symmetric lenses fall between asymmetric
+  semantic meaning.
+%
+\item We extend the Boomerang implementation with simple symmetric lenses,
+  demonstrating that they integrate smoothly with other lens extensions such as
+  quotient lenses~\cite{quotientlenses} and matching
+  lenses~\cite{matchinglenses}.
+
+\item We also extend Boomerang with symmetric lens synthesis specifications:
+  including types, examples, and relevance annotations. We evaluate our system
+  on a set of 48 benchmarks, and find it synthesizes each of them in under
+  30 seconds.
+%
+\item We show that simple symmetric lenses fall between asymmetric
 lenses~\cite{Focal2005-long} and full symmetric
@@
 Along the
-way, we show that {\em star-semiring equivalences}\bcp{citation} on regular
-expressions 
+way, we show the {\em star-semiring equivalences} on regular expressions
 remain valid when extended to stochastic regular expressions, preserving
@@
 \subsection{Simple Symmetric Lenses}
-Semantically, a simple symmetric lens (we will just say ``lens'' when clear
-from context) between 
-sets $X$ and $Y$ comprises four functions subject to four round-tripping
+Semantically, a simple symmetric lens (or lens when clear from context) between
+sets $X$ and $Y$ comprises four functions subject to four ``round-tripping''
 laws.
-\vspace*{.7ex}
-\begin{trivlist}
-\item 
+
 %\begin{enumerate}
@@
 \end{center}
-\end{trivlist}
-The two \KW{create} functions are used to \proposedelete{propagate shared
-  information and} fill 
+\noindent
+\vskip -.5em
+The two \KW{create} functions are used to propagate shared information and fill
 in default values when introducing new data (\EG, on \KW{create} the ``unk''
@@
 the other by combining a new value from one with an old value from the other.
-The record projection notation $\Lens.\PutR$ extracts the \PutR function
-from the 
-lens $\Lens$. These four functions can be used to keep the
-common information between two file types in sync. For example, if a new file
-of the left-hand format is created, the \CreateR function will build a new
-file in the 
-right-hand format. If the right-hand file is then edited, the \PutL function can
-update the left-hand file with the changed information.
+The notation $\Lens.\PutR$ extracts the \PutR function from the
+lens $\Lens$; similarly for \PutL. These four functions can be used to keep the
+common information between two file types in synch. For example, if a new file
+of the left-hand format is created, the \CreateR function will a new file in the
+right-hand format. If the right-hand format is edited, the \PutL function can
+update the left-hand format with the changed information.
 
@@
 side (the result of a put or create) back through a lens in the other direction
-will get back to exactly where we began. For example, consider a lens
+will get back to exactly where we began. For example, consider a simple
+symmetric lens
 synchronizing the employee data formats in
@@
 
-Simple symmetric lenses differ from ''classical'' symmetric lenses in that
-they do not involve a \emph{complement}. We give a detailed comparison in
+This simple symmetric lens definition differs from classical symmetric lenses in
+that it does not involve a \emph{complement}. We give a detailed comparison in
 \S\ref{sec:relationship}.
@@
 
-In principle, programmers can define lenses just by writing
+In principle, programmers can define simple symmetric lenses just by writing
 four arbitrary functions by hand and showing they satisfy the round-tripping
-laws, but this can be tedious and error prone.  A better idea is to provide
-a set of combinators---a domain-specific language---for building complex,
-law-abiding 
+laws. But a better idea is to streamline the process by providing a set of
+combinators---a domain-specific language---for building complex, law-abiding
 lenses from simpler ones. Such bidirectional languages have been given for
-many different kinds of data, including database
+building lenses over many different kinds of data, including database
 relations~\cite{BohannonPierceVaughan} algebraic data
 types~\cite{symmetric-lenses}, graphs~\cite{tgg}, and more.
+
 We focus here on combinators for defining lenses between string data formats
 described by regular expressions.
-
 If $\BRegex$ and $\BRegexAlt$ are regular expressions, then
@@
 $\LanguageOf{\BRegexAlt}$. (We will use undecorated variables later for
-stochastic regular expressions, so we mark plain REs with overbars.)
+stochastic regular expressions; we mark plain REs with overbars.)
 We illustrate some of these
@@
 The simplest combinator is the identity lens \IdentityLens, which takes as an
-argument a regular expression $\BRegex$ and propagates data unchanged
-in both directions.
+argument a regular expression $\BRegex$ and performs an identity mapping on data
+matching $\BRegex$. 
 %
@@
 HR file.
-With the above lens, $\PutL \App \lstinline{""} \App 60000$ will return 60000,
+For example, with the above lens, $\PutL \App \lstinline{""} \App 60000$ will return 60000,
 and \PutR will always return \lstinline{""}.
@@
 ($\OrLensOf{\Lens_1}{\Lens_2}$), and iteration ($\IterateLensOf{\Lens}$).
-For example, we could use the lens
+For example, we could use the following lens
 %
@@
 table. For example given a
-lens \lstinline{employee_lens} that synchronizes data for a single employee, the
-lens\bcp{In all these examples, I'd like to have one more space on each side
-of the ``:''}
+lens \lstinline{employee_lens} that synchronizes data for a single employee the
+following lens
 %
@@
 transforms a list of employees in employees in the Management format to a list
-of employees in the HR format and vice versa. These combinators
+of employees in the HR format in the left-to-right direction. These combinators
 are combined in figure~\ref{fig:example_lens} to construct a complete lens
@@
 
-\bcp{I wonder whether the rest of this subsection can be cut... It is both
-  hard to understand and too brief to really say much, and I wonder whether
-  we need it for anything between here and where we give it a real
-  explanation.}
-
-Most of the typing rules for lenses are syntax directed, where only a handful of
-rules result in a given type\bcp{syntax directed usually means that the
-  conclusion of at most
-  {\em one} rule will match a given goal; if that's not the case here, then
-  we need to speak more carefully}.  For example, the type of a
+Most of the typing rules for lenses are syntax driven, where only a handful of
+rules result in a given type.  For example, the type of a
 \ConcatLens lens is the concatenation of the regular expressions of the sub-lenses.
@@
 \]
-The typing  rules for maintain structural similarity
-between regular expression types and lens terms, with one exception: The
-type equivalence rule permits a well-typed lens to be well-typed among any
-equivalent regular expression pair.
+With one notable exception, lens typing rules maintain structural similarity
+between regular expression types and lens terms; type equivalence does not. Type
+equivalence permits a well-typed lens to be well-typed among any equivalent
+regular expression pair.
 %
@@
 
-For larger and more complex formats, it can become quite difficult to
-write lenses by hand, as we did above.  Instead, we would like to synthesize
-them from types and examples.  
-That is, given a pair of regular expressions and a set of input-output
-examples, we want 
+Even for small examples, lenses like the one above take a little getting used
+to. For larger and more complex formats, it can become quite difficult to
+write lenses by hand.  Instead, we wish to synthesize them from examples. 
+That is, given a pair of regular expressions and a set of input-output examples, we want
 to find a simple symmetric lens that is typed by the regular expression pair and
-satisfies all the input/output examples. We this a \emph{satisfying
-  lens}. For our running example, suppose we wish to synthesize a lens between
-\lstinline{emp_salaries} and \lstinline{emp_insurance}, using as sole
-input-output example the data in Figure~\ref{fig:minimized-representations}.
+satisfies all the input/output examples. We call such a lens a \emph{satisfying
+  lens}. In our running example, we wish to synthesize a lens between
+\lstinline{emp_salaries} and \lstinline{emp_insurance}, using as an input-output
+example the data in Figure~\ref{fig:minimized-representations}.
 
@@
 Figure~\ref{fig:minimized-representations} (where \lstinline{...} represents
-components of the example itself\bcp{huh?}). In general, {\em many} examples
-may be 
-required to rule out all possible occurrences of nested\bcp{??} \Disconnect lenses,
+components of the example itself). In general, {\em many} examples may be
+required to rule out all possible occurrences of nested \Disconnect lenses,
 particularly in complex formats.
@@
 %and is highlighted even more on larger formats.
-\proposedelete{While we could ask programmers to provide detailed and precise specifications
+While we could ask programmers to provide detailed and precise specifications
 for their synthesis tasks, we would prefer an algorithm that can intelligently
-choose a lens despite the underspecification.} Instead of merely finding
+choose a lens despite the underspecification. Instead of merely finding
 \emph{any} satisfying lens, we wish to synthesize a satisfying lens that is
@@
 A satisfying lens is ``more likely'' if it uses more data from one format to
-construct the other.  For example, the
-identity lens (which uses all the data) is more likely than the disconnect
-lens (which uses none). Formally, we define the {\em likelihood} of a
-satisfying lens as the 
-expected number of bits required to recover data in one format from 
-data in the other; higher likelihoods correspond to fewer bits. Two
-strings $s$ and $t$ are \emph{synchronized according to lens
-  $l$}\bcp{Earlier we used $\Lens$, now we are switching to $l$...} if $l.\PutR\
+construct the other. One manifestation of this principle involves choosing the
+identity lens (which uses all the data) rather than the disconnect lens (which
+uses none of the data). We define the likelihood of a satisfying lens as the
+expected number of bits required to recover data in one format from synchronized
+data in the other. Higher likelihoods correspond to fewer bits. Two
+strings $s$ and $t$ are \emph{synchronized according to lens $l$} if $l.\PutR\
 s\ t = t$ and $l.\PutL\ t\ s = s$. We can \emph{recover} $s$ from
@@
 \Disconnect{} lens, we need enough bits to fully encode $s$ in order to recover
-it from $t$ because all the information in $t$ gets thrown away. Thus, if both
+it from $t$ because $t$ contains no useful information. Thus, if both
 \IdentityLens{} and \Disconnect{} are satisfying lenses for a particular pair of
-formats, \IdentityLens{} will be more likely.
+formats, then \IdentityLens{} will be more likely.
 
@@
 \end{theorem}
-
-(For the proof, see the appendix \ifappendices
-Theorem~\ref{thm:semantics-correctness} \else{} included with the
-supplemental material\fi.)  This theorem will come in handy as we traverse
-(with \Expand) or normalize across (as part of \GreedySynth) star-semiring
-equivalences.
+\begin{proof}
+  See the appendix %
+\ifappendices  Theorem~\ref{thm:semantics-correctness} \else{} included with the supplemental
+material\fi.
+\end{proof}
+
+This theorem will come in handy as we traverse (with \Expand) or normalize
+across (as part of \GreedySynth) star-semiring equivalences.
 
@@
 \subsection{Lens Costs}
-\label{subsec:lenscosts} \bcp{In the intro, we talked about ``likelihood.''
-Now we seem to be switching the terminology to ``cost.''}
+\label{subsec:lenscosts}
 Our cost metric is based on the expected amount of information required to
@@
 
-\paragraph*{Compositional Synthesis} Compositional synthesis allows us to use
-previously defined lenses, whether they are previously defined by users or by
-synthesis. As the synthesizer processes a Boomerang file, it accumulates
-lens definitions and their types.  It tackles synthesis problems one after another
+\paragraph*{Compositional Synthesis} Compositional synthesis allows \GreedySynth to use
+previously defined (by users or synthesis) lenses. As the synthesizer processes a Boomerang file, it accumulates
+lens definitions and types.  It tackles synthesis problems one after another
 and there may be many such problems in a given program.
-During synthesis, if an existing lens has the right type and agrees with the examples.
-\GreedySynth will use it.  Consequently, if a large synthesis problem cannot be solved all
-at once, a user can first generate subproblems for parts of a data format, and the
-larger problem can subsequently use solutions to those subproblems.  In our experience,
+During synthesis, if an existing lens has the right type and agrees with the examples,
+\GreedySynth will use it.  If a large synthesis problem cannot be solved all
+at once, a user can generate subproblems for parts of a data format, and the
+larger problem can subsequently use solutions to those subproblems. For example,
+when given the synthesis task:
+%
+\begin{lstlisting}
+let l_1 = synth R <=> S
+let l_2 = synth R' <=> S'
+\end{lstlisting}
+%
+the synthesis algorithm use the previously generated \lstinline{l_1} in the
+synthesis of \lstinline{l_2}.
+In our experience,
 this is a very powerful tool that allows synthesis to tackle problems of arbitrary
@@
   our bidirectional transformers performs the same unidirectional transformation
-  as Flash Fill -- we ignore the other direction.
+  as Flash Fill -- the other direction is determined from the round-tripping
+  laws. None of these benchmarks were bijective.
 \item We adapted 29 benchmarks from Augeas~\cite{augeas}. Augeas is a utility
@@
   in-memory dictionaries. In our benchmarks, we synthesize lenses between Linux
-  configuration files and serialized versions of Augeas dictionaries.
+  configuration files and serialized versions of Augeas dictionaries. Because
+  these benchmarks merely transformed the information into a structured form,
+  synthesized lenses were all bijective.
 \item We created 11 additional benchmarks 
@@
   literature. These tasks range from synchronizing REST and JSON web resource
-  descriptions to synchronizing BibTeX and EndNote citation descriptions.
+  descriptions to synchronizing BibTeX and EndNote citation descriptions. Five
+  of these benchmarks were not bijective lenses.
 \end{enumerate}
@@
 The majority of the tasks required only a single example and none required
-more than three examples in order to synthesize the desired lens.\footnote{In
+more than three examples to synthesize the desired lens.\footnote{In
   one benchmark, we supplied a fourth example that was later discovered to 
@@
 \begin{tabulary}{\linewidth}{rL}
-  \SSOpt{}: & The symmetric synthesis algorithm with all optimizations enabled.\\
-  \SSNCOpt{}: & The symmetric synthesis algorithm, with no compositional synthesis enabled.\\
+  \SSOpt{}: & Run the symmetric synthesis algorithm with all optimizations enabled.\\
+  \SSNCOpt{}: & Run the symmetric synthesis algorithm, with no compositional synthesis enabled.\\
 \end{tabulary}\\
@@
 synthesize all of the benchmarks in under 30 seconds. Without compositional
-synthesis, the synthesis algorithm is able to solve 26 problem instances.
+synthesis, the synthesis algorithm is able to solve 26 out of 48 problem instances.
 
@@
 \label{subsec:effect-of-heuristics}
-We evaluate the usefulness of (1) our information-theoretic metric, and (2) our
+We evaluate the usefulness of (1) our information-theoretic metric, (2) our
+termination heuristic and (3) our
 relevance annotations.  To this end, we run our program in several different modes:\\[1ex]
 \begin{tabulary}{\linewidth}{rL}
-  \AnyOpt{}: & We ignore the information-theoretic preference metric entirely,
-               returning the first lens \GreedySynth finds without using this information. \\ 
-  \FLOpt{}: &  \GreedySynth uses the information-theoretic preference
-              metric to search, but returns immediately upon finding a
-              satisfying lens, instead of returning based on \Continue.  \\
-  \CCOpt{}: & We replace our information-theoretic cost metric with one where
+  \AnyOpt{}: & Ignore the information-theoretic preference metric (\IE, all valid
+               lenses have cost 0). \\ 
+  \FLOpt{}: &  Return the first highest ranked lens \GreedySynth returns (\IE,
+              ignore the termination heuristic). \\
+  \CCOpt{}: & Replace our information-theoretic cost metric with one where
               the cost of the lens is the number of disonnects plus the number
               of merges.\\
-  \NSOpt{}: & We ignore all \Skip annotations in the SRE specifications. \\
-  \NROpt{}: & We ignore all \SRequire annotations in the SRE specifications. \\
+  \NSOpt{}: & Ignore all \Skip annotations in the SRE specifications. \\
+  \NROpt{}: & Ignore all \SRequire annotations in the SRE specifications. \\
 \end{tabulary}
@@
 termination condition is also quite important. When running in \FLOpt{} mode,
-the algorithm only discovers 5 lenses, which shows that the first cluster that
-contains a satisfying lens is rarely the correct cluster. However, our algorithm
+the algorithm only discovers 5 lenses, which shows that the first class that
+contains a satisfying lens is rarely the correct class. However, our algorithm
 is not perfect and fails when either it is very difficult to find the desired
@@
 and asymmetric lenses. Proofs about these relationships are included in
-the appendix
-\ifappendices  \S\ref{sec:appendixforget}. \else{}included with the supplemental
+the appendix \ifappendices  \S\ref{sec:appendixforget}. \else{}included with the supplemental
 material.  \fi We also comment on related
@@
 
-While classical symmetric lenses are technically more expressive than simple
-symmetric lenses, we still believe simple symmetric lenses are useful in
-practice. Indeed, all asymmetric lenses (we later show) are expressible as
+Simple symmetric lenses are a strict subset of classical symmetric lenses, but
+quite a useful one. For instance, all asymmetric lenses are expressible as
 simple symmetric lenses. The primary loss is the loss of ``memory'' within the
-complement. In classical symmetric lenses, \OrLens lenses retain information, so
-users don't lose information when swapping from one disjunct to the other, the
-information is stored in the complement. Simple symmetric lenses have no such
-complement, so they mimic the forgetful \OrLens lens of classical symmetric
-lenses.
-
-In our judgement, the complications they introduce into synthesis did
-not justify the potential additional expressiveness: because each lens has a
+complement. In classical symmetric lenses, disjunctive (\OrLens) lenses retain information
+about both possible formats. If a user edits a format from one disjuncted format
+to the other, the information contained in that first disjunct is retained
+within the complement. Simple symmetric lenses have no such complement, so they
+mimic the forgetful disjunctive lens of classical symmetric lenses.
+
+Though classical symmetric lenses are more expressive, they introduce drawbacks
+for synthesis: because each lens has a
 custom complement, one can no longer specify the put functions through
 input/output examples alone. One alternative would be to enrich specifications
-with edit sequentials; another would be to specify the structure of complements
+with edit sequences; another would be to specify the structure of complements
 explicitly (though the latter would be somewhat akin to specifying the internal
@@
 Intuitively, these equations state that complements are uniquely determined by
-the most recently input $x$ and $y$. Such lenses correspond exactly with simple
+the most recent input $x$ and $y$. Such lenses correspond exactly with simple
 symmetric lenses, where all state is maintained by the $x$ and $y$ data.
@@
 
-\paragraph*{Other Lens Formulations}\bcp{This paragraph doesn't add
-  much}\afm{the POPL reviewers complained we didn't have this type of
-  paragraph, I personally agree that it doesn't add much and feels unnecessary}
+\paragraph*{Other Lens Formulations}
 Bidirectional programming has an extensive literature, with many extensions onto
@@
 the data being transformed.  On the other hand, this type information
-makes it possible to transform more complex formats---Flash Fill
-does not synthesize programs with nested loops, for instance,
+makes it possible to transform more complex formats. For example, Flash Fill
+does not synthesize programs with nested loops~\cite{flashfill},
 and hence is incapable of synthesizing the majority of our
-benchmarks, even in one direction~\cite{flashfill}.
+benchmarks, even in one direction.
 %% Flashfill, designed
@@
 Fill prefers ``wider'' character classes over ``narrower'' ones. Again such a
-choice is implied by information theory -- the wider class preserves more
+choice is implied by information theory---the wider class preserves more
 information during translation. More broadly, we hope our information-theoretic
@@
 expression specifications. In order to guide the search for ``likely'' lenses,
-we introduce new search principles based on information theory. Based on these
-principles, we designed, implemented, and evaluated a new tool for lens
-synthesis.
+we introduce new search principles based on information theory and allow users
+to control this search via relevance annotations. To evaluate the effectiveness
+of these ideas, we designed and implemented a new tool for symmetric lens
+synthesis and integrated it into the Boomerang lens programming framework. Our
+experiments on 48 benchmarks demonstrate that we can synthesize complex lenses for
+real-life formats in under 30 seconds.
 

### 2019-02-28 papers/SymmetricLenses/main.tex

--- 2019-02-28 papers/SymmetricLenses/main.tex
@@
 effectiveness on a range of applications. Users set up a Boomerang synthesis
-task by providing the source and target types and optionally supplying one or
-more examples. In addition, users can override the default mechanism for
+task by providing two regular expressions and optionally supplying input-output
+examples. Users can also override the default mechanism for
 calculating the information content of a SRE by
 asserting that certain strings are \emph{essential} or \emph{irrelevant},
-thereby forcing certain data to either be retained or discarded during the
+forcing certain data to either be retained or discarded during the
 transformations.
 
-We evaluated our system on a benchmark suite of 48 problems, including
-data-cleaning tasks taken from Flash Fill~\cite{flashfill}, view-maintenance
-tasks taken from Augeas~\cite{augeas2}, and ad-hoc synchronization tasks
-from a variety of other sources (mostly from the bidirectional programming
-literature).  \bcp{``... and we found...''?}  \bcp{Also, this paragraph is
-  redundant with the corresponding contribution a bit below.  I think it
-  would improve the flow if we made this one a bit longer (and stronger) and
-  the contribution paragraph shorter, just summarizing what this one says.}
+We evaluate our implementation, the effects of optimizations, and our inclusion
+of relevance annotations on a set of 48 lens synthesis benchmarks drawn from
+data cleaning, view maintenance, and file synchronization tasks. We find the
+system can synthesize simple symmetric lenses for all of the benchmarks in under
+30 seconds (\S\ref{sec:evaluation}).
 

### 2019-03-01 papers/SymmetricLenses/main.tex
add updated info

@@
   using a novel application of stochastic regular expressions to guide the
-  search process (\S\ref{sec:synthesis}) and prove that our metrics have deep
-  semantic meaning. \bcp{``Semantic meaning'' is redundant.  And ``deep
-    semantic meaning'' just sounds silly.  Can we say it more precisely?}
+  search process (\S\ref{sec:synthesis}) and prove that our metrics correspond
+  to the information loss of a given lens.
 %
@@
 Along the
-way, we show that {\em star-semiring equivalences}\bcp{citation} on regular
+way, we show that {\em star-semiring equivalences}~\cite{starsemiring} on regular
 expressions 
@@
 \end{trivlist}
-The two \KW{create} functions are used to \proposedelete{propagate shared
-  information and} fill 
+The two \KW{create} functions are used to fill 
 in default values when introducing new data (\EG, on \KW{create} the ``unk''
@@
 lens \lstinline{employee_lens} that synchronizes data for a single employee, the
-lens\bcp{In all these examples, I'd like to have one more space on each side
-of the ``:''}
+lens
 %
@@
   we need it for anything between here and where we give it a real
-  explanation.}
-
-Most of the typing rules for lenses are syntax directed, where only a handful of
-rules result in a given type\bcp{syntax directed usually means that the
-  conclusion of at most
-  {\em one} rule will match a given goal; if that's not the case here, then
-  we need to speak more carefully}.  For example, the type of a
+  explanation.} \afm{I believe it is necessary for 2.4. I tried giving it a more
+terse treatment, but that caused its own set of problems.}
+
+Lenses are typed by pairs of regular expressions and only a handful of
+rules result in a given type.  For example, the type of a
 \ConcatLens lens is the concatenation of the regular expressions of the sub-lenses.
@@
 Figure~\ref{fig:minimized-representations} (where \lstinline{...} represents
-components of the example itself\bcp{huh?}). In general, {\em many} examples
+the example strings). In general, {\em many} examples
 may be 
-required to rule out all possible occurrences of nested\bcp{??} \Disconnect lenses,
+required to rule out all possible occurrences of \Disconnect lenses,
 particularly in complex formats.
@@
 %and is highlighted even more on larger formats.
-\proposedelete{While we could ask programmers to provide detailed and precise specifications
-for their synthesis tasks, we would prefer an algorithm that can intelligently
-choose a lens despite the underspecification.} Instead of merely finding
+Instead of merely finding
 \emph{any} satisfying lens, we wish to synthesize a satisfying lens that is
@@
 strings $s$ and $t$ are \emph{synchronized according to lens
-  $l$}\bcp{Earlier we used $\Lens$, now we are switching to $l$...} if $l.\PutR\
-s\ t = t$ and $l.\PutL\ t\ s = s$. We can \emph{recover} $s$ from 
-$t$ using bits $b$ and lens $l$ if we can reconstruct $s$ from $t$, $b$, and
-$l$. For example, given the \IdentityLens{} lens, we can recover $s$ from $t$
+  $\Lens$} if $\Lens.\PutR\
+s\ t = t$ and $\Lens.\PutL\ t\ s = s$. We can \emph{recover} $s$ from 
+$t$ using bits $b$ and lens $\Lens$ if we can reconstruct $s$ from $t$, $b$, and
+$\Lens$. For example, given the \IdentityLens{} lens, we can recover $s$ from $t$
 using no bits because, in this case, $s$ is just $t$. In contrast, given the
@@
 conditional entropy of $\RegexAlt$ given $\Regex$ and $\Lens$. This metric
-assigns higher entropy (or cost\bcp{= lower likelihood?  I'm still confused
-  by these different ways of speaking about what seems to be the same
-  thing.}) to lenses where knowing the string on one side 
+assigns higher entropy (or likelihood) to lenses where knowing the string on one side 
 provides little information about the string on the other side. It assigns zero
@@
 giving
-each string in the language equal probablity. In other words, this heuristic
-tries to make each bit of data no more or less likely than any
-other.\bcp{huh?} 
-Alternatively, our synthesis algorithm could take as input a pair of stochastic
-regular expressions rather than plain ones, permitting users to specify a
-data-specific probability distribution (either manually calculated or inferred
-from some dataset of interest).\bcp{So why didn't we do it this way?}
-
-However\bcp{``However'' means that something from before is being
-  contradicted.  What is it in this case?}, sometimes users already know that certain data should or should not be
+each string in the language equal probablity (not prioritizing one piece of
+information over another).
+
+Sometimes users already know that certain data should or should not be
 used to construct the other format. We introduce relevance annotations to SREs
@@
 \lstinline{emp_ins} be annotated as required (as they are converted with the
-identity lens).  \bcp{That paragraph seems way too brief and high level.
-  These annotations were added in response to significant complaints from an
-earlier PC, but the paper reads as though they are an afterthought of little
-consequence; they deserve more attention and explanation.}
+identity lens). Doing so both constrains the problem (as \lstinline{name}s must
+be synchronized) and makes it easier (as the algorithm does not waste time
+looking for \lstinline{salary} information in the insurance format). In this
+way, users can tweak the
+lens likelihoods with their external knowledge.
 
@@
 promising (for example, if the satisfying lens loses no information, the
-algorithm should terminate for no other lens will be more likely\bcp{this
-  case is 100\% obvious, which makes it not so interesting as an example.
-  The reader is left wondering what happens with pretty likely but not
-  maximally likely lenses?}). The
+algorithm should terminate for no other lens will be more likely). The
 heuristics also favor stopping if \Expand has delivered to \GreedySynth all
@@
 number of pairs at the next rank, because searching through all such pairs will
-take a long time. With this approach, the algorithm can quickly respond with a
+take a long time (for example, if the satisfying lens loses only a little
+information,
+but searching for a better one will take a long time, the discovered lens is
+returned).
+With this approach, the algorithm can quickly return a
 satisfying lens with relatively high likelihood. If the user is unhappy with the
@@
 certain information either must be retained or must be discarded by the lens
-(see \S\ref{subsec:optimizations}). \bcp{Again, I feel like we're
-  downplaying this aspect too much.}
+(see \S\ref{subsec:relevanceannotations}).
 
@@
 
-\subsection{Lens Costs}
-\label{subsec:lenscosts} \bcp{In the intro, we talked about ``likelihood.''
-Now we seem to be switching the terminology to ``cost.''}
-Our cost metric is based on the expected amount of information required to
+\subsection{Lens Likelihoods}
+\label{subsec:lenscosts}
+Our likelihood metric is based on the expected amount of information required to
 recover a string in one data format from the other. We use the function
@@
 
-The cost of a lens between two SREs $cost(\Lens, \Regex, \RegexAlt) =
-\MaxOf{\LEntropyOf{\Regex \Given \Lens, \RegexAlt}} +
-\MaxOf{\REntropyOf{\RegexAlt \Given \Lens, \Regex}}$ is the sum of the maximum
-expected number of bits to
-recover the left format from the right, and the right from the left. We have
-proven theorems demonstrating the calculated entropy corresponds to the actual
-conditional entropy to recover the data.
+The likelihood of a lens is the negative of its \emph{cost}. The cost of a lens
+between two SREs $cost(\Lens, \Regex, \RegexAlt) = \MaxOf{\LEntropyOf{\Regex
+    \Given \Lens, \RegexAlt}} + \MaxOf{\REntropyOf{\RegexAlt \Given \Lens,
+    \Regex}}$ is the sum of the maximum expected number of bits to recover the
+left format from the right, and the right from the left. We have proven theorems
+demonstrating the calculated entropy corresponds to the actual conditional
+entropy to recover the data.
 
@@
 \item executes $\GreedySynth$ to find a best lens in that class if
-one exists ($\LC$),
+one exists ($\Lens$),
 using the examples $exs$ to filter out potential lenses that do not satisfy
 the specification,
-\item replaces $best$ with $\LC$, if $\LC$ has lower cost according
+\item replaces $best$ with $\Lens$, if $\Lens$ is more likely according
 to our information-theoretic metric, and
@@
     \State $(\Regex,\RegexAlt) \gets \Call{\PQ.Pop}{\RXSearchState}$
-    \State $\LC \gets \Call{\GreedySynth}{\Examples,\Regex,\RegexAlt}$
-    \If{$\Call{Cost}{\LC} < \Call{Cost}{\Best}$}
-    \State $\Best \gets \LC$
+    \State $\Lens \gets \Call{\GreedySynth}{\Examples,\Regex,\RegexAlt}$
+    \If{$\Call{Cost}{\Lens} < \Call{Cost}{\Best}$}
+    \State $\Best \gets \Lens$
     \EndIf
@@
 \subsection{Relevance Annotations}
+\label{subsec:relevanceannotations}
 
@@
 the generated lens to retain \lstinline{name} information. In practice, we
-define the cost of any lens that loses ``required'' information to be infinite.
+make any lens that loses ``required'' information infinitely unlikely.
 %
@@
 function \CartesianMap maps its argument across the cross product of the input
-lists). A priority queue containing these sequence lenses, ordered by cost, is
+lists). A priority queue containing these sequence lenses, ordered by likelihood, is
 then initialized with $\PCF{PQ.Create}$. The symmetric lens is then built up
@@
 between each atom pair of the left and right sequences, and organizes them into
-a priority queue ordered by cost with \PCF{PQ.Create}. The symmetric sequence
+a priority queue ordered by likelihood with \PCF{PQ.Create}. The symmetric sequence
 lens is built up iteratively from these atom lenses, where the state of the
@@
   \vspace{-4ex}
-  \caption{\bcp{What are the little tick marks on the top of the figure?
-      And why do we need a line for ``benchmark count''?}Number of
+  \caption{Number of
     benchmarks that can be solved by a given algorithm in a 
@@
 \begin{tabulary}{\linewidth}{rL}
-  \SSOpt{}: & The symmetric synthesis algorithm with all optimizations enabled.\\
   \BSOpt{}: & The existing bijective synthesis
               algorithm with all optimizations enabled.\\
+  \SSOpt{}: & The symmetric synthesis algorithm with all optimizations enabled.\\
 \end{tabulary}\\
@@
 
-\bcp{Future work??????????????}
+\bcp{Future work??????????????}\afm{I don't think there's room for it}
 
@@
 Since $S$ is unambigous, then $S_1$ and $S_2$ are also unambiguous, and for each $s \in \mathcal{L}(S_1 \cdot S_2)$ there exists a unique $s_1 \in \mathcal{L}(S_1)$ and a unique $s_2 \in \mathcal{L}(S_2)$ such that $s = s_1 \cdot s_2$.
-By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$, hence\bcp{Don't understand what's going on with the $-$
-  signs in the last few lines here.}\afm{in the definition of entropy there is a negative. this is to
-  address that we're taking the log of a probability, which inherently will
-  return a negative number}
+By the induction hypothesis, $H(\mathcal{L}(S_1)) = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) = \mathbb{H}(S_2)$, hence
 \begin{align*}
@@
 = \mathbb{H}(S_1)$ and $H(\mathcal{L}(S_2)) =
-\mathbb{H}(S_2)$. Then\bcp{Again, don't understand what's going on with the $-$
-  signs.}
+\mathbb{H}(S_2)$. Then
 \begin{align*}

### 2019-03-01 papers/SymmetricLenses/transform-data.py

@@ -76,8 +76,8 @@ def generate_examples_required_graph(input_csv):
     ax.set_xticklabels(["\\textbf{Any}","\\textbf{FL}","\\textbf{DC}","\\textbf{NS}","\\textbf{NR}"])
 
     fig = plt.figure(3,tight_layout=True)
-    ax.step([-.5,4.5],[48.1,48.1],label="Benchmark Count",linestyle=":",
-            linewidth=1, dashes=(1,1))
+    #ax.step([-.5,4.5],[48.1,48.1],label="Benchmark Count",linestyle=":",
+    #        linewidth=1, dashes=(1,1))
 
     plt.tick_params(
                 axis='x',          # changes apply to the x-axis
@@ -309,22 +309,23 @@ def generate_time_vs_tasks_graph_vs_bijective(input_csv):
             x_completed_counts.append(acc)
         x_completed_counts = [0] + x_completed_counts[:len(x_completed_counts)-1]
         if (style != '-'):
-            ax.step(x_vals,x_completed_counts,label=outputname,linestyle=style,linewidth=width, dashes=(5,1))
+            ax.step(x_vals,x_completed_counts,label=outputname,linestyle=style,linewidth=width, dashes=(5,1), color='g')
         else:
             ax.step(x_vals,x_completed_counts,label=outputname,linestyle=style,linewidth=width)
 
     normal_size = 2
     full_size = 3
 
-    ax.step([0,60],[39.1,39.1],label="Benchmark Count",linestyle=":",
-            linewidth=1, dashes=(1,1))
-    create_step_plot("SS","\\textbf{SS}",'-',normal_size)
+    #ax.step([0,60],[39.1,39.1],label="Benchmark Count",linestyle=":",
+    #        linewidth=1, dashes=(1,1))
     create_step_plot("BS","\\textbf{BS}",':',normal_size)
+    create_step_plot("SS","\\textbf{SS}",'-',normal_size)
 
     ax.set_ylabel('Benchmarks\nCompleted')
     ax.set_xlabel('Time (s)')
 
-    l = ax.legend(bbox_to_anchor=(1.6,1),borderaxespad=0)
+    #l = ax.legend(bbox_to_anchor=(1.6,1),borderaxespad=0)
+    l = ax.legend(bbox_to_anchor=(1.25,1),borderaxespad=0)
     plt.setp(l.texts) 
 
[…]

### 2019-03-01 papers/SymmetricLenses/main.tex
remove discussions
spellcheck and other checks

@@
 theory to estimate the amount of information propagated by a candidate
-lens, generally prefering lenses that propagate more information, 
+lens, generally preferring lenses that propagate more information, 
 as well as user annotations marking parts of the source and
@@
 procedure as extensions to the Boomerang language. We evaluate its performance
-on 48 benchmark examples drawn from Flash Fill, Augeus, the bidirectional
+on 48 benchmark examples drawn from Flash Fill, Augeas, the bidirectional
 programming literature, and electronic file format synchronization tasks. Our
@@
 reference number, while the same configuration file on another operating
-system does not.  Indeed among the 48 benchmark problems described in
-Section \ref{sec:evaluation} below, all of the benchmarks taken from Flash
+system does not.  Indeed among the benchmark problems described in
+Section \ref{sec:evaluation}, all of the benchmarks taken from Flash
 Fill~\cite{flashfill} and many of the benchmarks that synchronize between
@@
 source structure to a target can throw away information that must
-then be restored when transferring from target to source), and even more flexibile
+then be restored when transferring from target to source), and even more flexible
 transformations that allow each side to throw away information?
@@
 present in either $X$ or $Y$. These complements must be stored
-and managed somehow.  More fundamentially, complements complicate
+and managed somehow.  More fundamentally, complements complicate
 synthesis {\em specifications}---instead of just giving single examples of
@@
 
-The round-tripping laws guarantee that pushing an ``un-edited'' value from one
+The round-tripping laws guarantee that pushing an ``unedited'' value from one
 side (the result of a put or create) back through a lens in the other direction
@@
 and the original salary file will return the original salary file, unmodified.
-If \KW{put}ting a string $s$ into a string $t$ results in the original string $s$, then $s$ and $t$ are considered \emph{synchronized}.
 
@@
 
-\bcp{I wonder whether the rest of this subsection can be cut... It is both
-  hard to understand and too brief to really say much, and I wonder whether
-  we need it for anything between here and where we give it a real
-  explanation.} \afm{I believe it is necessary for 2.4. I tried giving it a more
-terse treatment, but that caused its own set of problems.}
-
 Lenses are typed by pairs of regular expressions and only a handful of
-rules result in a given type.  For example, the type of a
+rules result in a given type.
+Many of these rules provide tight relationships between terms and types, for example, the type of a
 \ConcatLens lens is the concatenation of the regular expressions of the sub-lenses.
@@
 to find a simple symmetric lens that is typed by the regular expression pair and
-satisfies all the input/output examples. We this a \emph{satisfying
+satisfies all the input/output examples. We call this a \emph{satisfying
   lens}. For our running example, suppose we wish to synthesize a lens between
@@
 \begin{lstlisting}
-disconnect(emp_salaries, emp_insurance, ..., ...)
+disconnect(emp_salaries, emp_insurance, $s$, $t$)
 \end{lstlisting}
@@
 is another well-typed lens that satisfies the example in
-Figure~\ref{fig:minimized-representations} (where \lstinline{...} represents
-the example strings). In general, {\em many} examples
+Figure~\ref{fig:minimized-representations} (where $s$ and $t$ represent
+the provided example strings). In general, {\em many} examples
 may be 
@@
 none have 
-significantly higher probability than others, it will have high information
-content. Data sources where most elements have small
-probabilities have small entropy. When $P$ is clear from context, we will
+significantly higher probability than others, it will have high entropy. Data
+sources with just a few high probability elements have low
+entropy. When $P$ is clear from context, we will
 often use the shorthand of $\EntropyOf{S}$ for $\EntropyOf{S,P}$.
@@
 giving
-each string in the language equal probablity (not prioritizing one piece of
+each string in the language equal probability (not prioritizing one piece of
 information over another).
@@
 disconnected), where \lstinline{name} in \lstinline{emp_salary} and
-\lstinline{emp_ins} be annotated as required (as they are converted with the
+\lstinline{emp_ins} could be annotated as required (as they are converted with the
 identity lens). Doing so both constrains the problem (as \lstinline{name}s must
@@
 split our algorithm into two communicating
-searche procedures. The first,
+search procedures. The first,
 \Expand, navigates the space of semantically equivalent regular expressions
@@
 These involve neither a composition operator nor a type equivalence rule.
-These restrictions mean that there are only a few simple symmetric n-ary DNF
+These restrictions mean that there are comparatively few simple symmetric n-ary DNF
 lenses that are well typed between a given pair of stochastic n-ary DNF regular
@@
 to judge whether to return the current best satisfying lens to the user or to
-pass the next set of unrollings to \GreedySynth. The heuristics favor stopping
+pass the next set of equivalent SREs to \GreedySynth. The heuristics favor stopping
 if the current best satisfying lens is very likely, indicating the lens is very
@@
 number of pairs at the next rank, because searching through all such pairs will
-take a long time (for example, if the satisfying lens loses only a little
-information,
-but searching for a better one will take a long time, the discovered lens is
+take a long time (\IE, if a satisfying lens loses only a little
+information
+and searching for a better one will take a long time, the discovered lens is
 returned).
@@
 For example, the entropy of $\Regex =
-\PRegexOr{\lstinline{"a"}}{(\lstinline{"b"})}{.5}$ is 1. The best encoding of a
+\PRegexOr{\lstinline{"a"}}{\lstinline{"b"}}{.5}$ is 1. The best encoding of a
 stream of elements from $\Regex$ will use, on average, 1 bit per element to
@@
 typically calculate the conditional entropy exactly, as merges are relatively
-uncommon in practice; only 2 of the lenses sythesized in our benchmarks
+uncommon in practice; only 2 of the lenses synthesized in our benchmarks
 include merges.
@@
 sidestep such considerations by synthesizing \emph{DNF lenses}---simple symmetric
-lenses that inhabit a disjunctive normal form that does not include composition.
+lenses that inhabit a disjunctive normal form that does not include
+composition.
 
@@
 \Expand subroutine generates new classes using the star-unrolling axioms
-(\UnrollstarLeftRule and \UnrollstarRightRule).
+(\UnrollstarLeftRule{} and \UnrollstarRightRule).
 
@@
 one exists ($\Lens$),
-using the examples $exs$ to filter out potential lenses that do not satisfy
+using the examples $\Examples$ to filter out potential lenses that do not satisfy
 the specification,
@@
 add this knowledge to the specification to optimize the search and force the
-the generated lens to retain \lstinline{name} information. In practice, we
+generated lens to retain \lstinline{name} information. In practice, we
 make any lens that loses ``required'' information infinitely unlikely.
@@
 \GreedySynth relies on \GreedySeqSynth, \GreedySeqSynth relies on \AtomSynth,
-and \AtomSynth rlies on \GreedySynth. The structure of the algorithms mirrors
+and \AtomSynth relies on \GreedySynth. The structure of the algorithms mirrors
 the structure of symmetric DNF lenses and \SDNFREabbrevs.
@@
     \State $\GreedyState \gets \Call{PQ.Create}{\SLS}$
-    \State $lb \gets \Call{LensBuilder.Empty}{}$
+    \State $\LensBuilder \gets \Call{LensBuilder.Empty}{}$
     \While{$\Call{PQ.IsNonempty}{\GreedyState}$}
@@
     \Call{PQ.Pop}{\GreedyState}$
-    \If{$\Call{LensBuilder.UsefulAdd}{lb,\SSQLens,\Examples}$}
-    \State $lb \gets \Call{LensBuilder.AddSeq}{lb,\SSQLens}$
+    \If{$\Call{LensBuilder.UsefulAdd}{\LensBuilder,\SSQLens,\Examples}$}
+    \State $\LensBuilder \gets \Call{LensBuilder.AddSeq}{\LensBuilder,\SSQLens}$
     \EndIf
@@
 iteratively from these sequence lenses, where the state of the partially
-constructed lens is tracked in the lens builder, $lb$.
+constructed lens is tracked in the lens builder, \LensBuilder.
 
@@
 
-Finally, after all sequences have been popped, the partial DNF lens $lb$ is
+Finally, after all sequences have been popped, the partial DNF lens \LensBuilder is
 converted into a symmetric DNF lens. This is only possible if all sequences are
@@
 lens is built up iteratively from these atom lenses, where the state of the
-partially built lens is tracked in the sequence lens builder, $slb$.
+partially built lens is tracked in the sequence lens builder, \SeqLensBuilder.
 
@@
     \State $\GreedyState \gets \Call{PQ.Init}{\ALS}$
-    \State $slb \gets \Call{SLensBuilder.Empty}{}$
+    \State $\SeqLensBuilder \gets \Call{SLensBuilder.Empty}{}$
     \While{$\Call{PQ.IsNonempty}{\GreedyState}$}
@@
     \Call{PQ.Pop}{\GreedyState}$
-    \If{$\Call{SLensBuilder.UsefulAdd}{slb,\SAtomLens,\Examples}$}
+    \If{$\Call{SLensBuilder.UsefulAdd}{\SeqLensBuilder,\SAtomLens,\Examples}$}
     \State $\GreedyState \gets \Call{SLensBuilder.AddAtom}{\GreedyState,\SAtomLens}$
@@
 and compositional synthesis. The optimizations
-make the system performant enough for interactive use, and the annotations help
-ensure the desired information is projected.
+make the system performant enough for interactive use.
 
@@
 relatively efficiently, it can suffer from an exponential blowup when converting
-regular expressions to DNF form. This problem can be mitigated by avoiding the conversion
-of some REs to DNF form and by performing the conversion lazily when necessarily.
-More specifically, \Expand labels some uncoverted REs as ``closed,'' which means the
+SREs to DNF form. This problem can be mitigated by avoiding the conversion
+of some SREs to DNF form and by performing the conversion lazily when necessarily.
+More specifically, \Expand labels some unconverted SREs as ``closed,'' which means the
 type-directed \GreedySynth algorithm treats them as DNF atoms and does not dig into
 them recursively.  In other words,
-given a pair of closed REs, \GreedySynth can either construct
-the identity lens between them (it will do this if they are the same RE), or it can
+given a pair of closed SREs, \GreedySynth can either construct
+the identity lens between them (it will do this if they are the same SRE), or it can
 construct a disconnect lens between them.
@@
   our bidirectional transformers performs the same unidirectional transformation
-  as Flash Fill -- the other direction is determined from the round-tripping
+  as Flash Fill---the other direction is determined from the round-tripping
   laws. None of these benchmarks were bijective.
@@
   \CCOpt{}: & Replace our information-theoretic cost metric with one where
-              the cost of the lens is the number of disonnects plus the number
+              the cost of the lens is the number of disconnects plus the number
               of merges.\\
@@
 
-While our tool uses types and examples to specify invertable transformations, other
+While our tool uses types and examples to specify invertible transformations, other
 tools have been shown to synthesize a backwards transformation from ordinary code designed
@@
 inference tool~\cite{pads:synthesis}, which was in turn inspired by earlier work
-on probabilistic gramatical inference and information
+on probabilistic grammatical inference and information
 extraction~\cite{kushmerick-thesis,arasu:extracting}. PADS did not synthesize
@@
 real-life formats in under 30 seconds.
-
-\bcp{Future work??????????????}\afm{I don't think there's room for it}
 
@@
 %%% End:
+
+%  LocalWords:  SREs SRE emp

### 2019-03-03 report/final-report.tex
added symmetric synthesis to methods section

@@
 
-Figure~\Cref{fig:schematic-diagram-synthesis} shows a high-level,
+\Cref{fig:schematic-diagram-synthesis} shows a high-level,
 schematic diagram for \Optician{}.
@@
 simulataneously in one compact notation.
+
+\subsection{Simple symmetric lens synthesis}
+Although it is very useful to be able to synthesize bijective or
+quotient lenses between data formats, such lenses do not fully solve the
+problem. 
+Bijective and quotient lenses require the two data formats to have
+either precisely or morally the same information content. 
+Many related data formats in practice have overlapping
+information: some information may be present in one format but not the
+other and vice versa.
+
+Symmetric lenses~\cite{HofmannPierceWagner10:POPL} address this
+limitation.  General symmetric lenses introduce the notion of 
+\textit{complement} that stores the information necessary to fully
+reconstruct one format from the other.  Because such complements would
+complicate the synthesis specificaiton problem, we instead define a
+more restricted language of \textit{simple symmetric lenses}, which
+are the largest class of symmetric lenses that do not rely on
+persistent internal state to synchronise between two related data
+formats, a property called \textit{forgetfulness}.  (Instead of a
+complement, simple symmetric lenses rely on defaults to replace
+missing information.)
+
+A challenge in adopting the type-directed synthesis approach to simple
+symmetric lenses is the number of such lenses.  Whereas the number of
+bijective lenses between two formats is typically tiny, the
+number of simple symmetric lenses is typically enormous. 
+If a na\"ive  search algorithm just selects the first simple symmetric
+lens it finds, the returned lens will generally not be the one the
+user wanted. Symmetric synthesis requires a new principle for
+identifying “more likely” lenses and a more sophisticated synthesis
+algorithm that uses this principle to search the space more
+intelligently.
+
+For these, we turn to information theory. We consider ``likely'' lenses to
+be ones that propagate ``a lot'' of information 
+from the left data format to the right and vice versa. Conversely, ``unlikely''
+lenses are ones that require a large amount of additional information to recover
+one of the formats given the other. By default, our synthesis algorithm prefers
+lenses that propagate more information.
+This preference is formalized using \emph{stochastic regular expressions}
+(SREs)~\cite{stoch-rnn,stoch-def}, which simultaneously define a set of
+strings and a probability distribution over those strings.  Using this
+probability distribution, we can calculate the likelihood of a given
+lens.
+We also allow users to override the default mechanism for
+calculating the information content of a SRE by
+asserting that certain strings are \emph{essential} or \emph{irrelevant},
+forcing certain data to either be retained or discarded during the
+transformations.
+
+\begin{figure}
+ \centering
+  \includegraphics[width=.63\textwidth]{high-level-algorithm.pdf}
+  \vspace{-2ex}
+  \caption{Schematic diagram for the simple symmetric lens synthesis algorithm.
+    The user provides regular expressions \BRegex and \BRegexAlt and a set of
+    examples \Examples{} as input. \Expand first converts \BRegex and \BRegexAlt
+    to stochastic regular expressions \Regex and \RegexAlt with default
+    probabilities. It then finds pairs of stochastic regular expressions
+    equivalent to \Regex and \RegexAlt and iteratively proposes them to
+    \GreedySynth. \GreedySynth finds a lens typed between the supplied SREs.
+    When the algorithm finds a likely lens, it returns it.}
+  \label{fig:high-level-algorithm}
+\end{figure}
+
+With simple symmetric lenses and this SRE-based likelihood measure in hand, we
+propose a new algorithm for synthesizing likely lenses. At its core, the
+algorithm performs a type-directed search between descriptions of the data
+formats, measuring success using the likelihood measure.
+
+Interesting complications arise from the need to deal with regular-expression
+equivalences. There are infinitely many regular expressions equivalent to a
+given one, and the lens returned by a type-directed search will in general
+depend on which of the possible representations are chosen for its source and
+target formats. Moreover, certain lenses may not be well-typed unless the
+format representations are replaced by equivalent ones:
+in general, the synthesis algorithm has to search through
+equivalent regular expression types to find the most likely lens. To tame
+this complexity, we divide the synthesis algorithm into two communicating search
+procedures (Figure~\ref{fig:high-level-algorithm}), following~\cite{optician}.
+The first, \Expand, uses rewriting rules to propose new pairs of
+stochastic regular expressions equivalent to the original pair. The second,
+\GreedySynth, uses a greedy, type-directed
+algorithm to find a simple symmetric lens between input SRE pairs, returning the lens and
+its likelihood score to \Expand. The whole synthesis algorithm heuristically
+terminates when a sufficiently likely lens is found.
 

### 2019-03-03 report/final-report.tex
results section for simple symmetric synthesis

@@
 
+\subsection{Simple Symmetric Synthesis Algorithms}
+We implemented simple symmetric lenses as an extension to
+Boomerang~\cite{boomerang}.  
+We also integrated our synthesis engine into Boomerang, allowing users
+to write synthesis tasks 
+alongside lens combinators, incorporate synthesis results into manually-written
+lenses, and reference previously defined lenses during synthesis. All experiments
+were performed on a 2.5 GHz Intel Core i7 processor with 16 GB of 1600 MHz DDR3
+running macOS Mojave.
+
+\paragraph*{Benchmark Suite Construction}
+Our benchmarks are drawn from three different sources.
+\begin{enumerate}
+\item We adapted 8 data cleaning benchmarks from Flash Fill~\cite{flashfill}. Flash Fill data
+  cleaning tasks are either derived from online help forums or taken from the
+  Excel product team. Note that our tool produces bidirectional transformations
+  rather than one-way transformers like Flash Fill. We ensure one direction of
+  our bidirectional transformers performs the same unidirectional transformation
+  as Flash Fill---the other direction is determined from the round-tripping
+  laws. None of these benchmarks were bijective.
+\item We adapted 29 benchmarks from Augeas~\cite{augeas}. Augeas is a utility
+  that bidirectionally converts between Linux configuration files and an
+  in-memory dictionaries. In our benchmarks, we synthesize lenses between Linux
+  configuration files and serialized versions of Augeas dictionaries. Because
+  these benchmarks merely transformed the information into a structured form,
+  synthesized lenses were all bijective.
+\item We created 11 additional benchmarks 
+  derived from real-world examples and/or the bidirectional programming
+  literature. These tasks range from synchronizing REST and JSON web resource
+  descriptions to synchronizing BibTeX and EndNote citation descriptions. Five
+  of these benchmarks were not bijective lenses.
+\end{enumerate}
+
+\paragraph*{Synthesizing Correct Simple Symmetric Lenses}
+
+To determine whether the system can synthesize desired lenses, we ran
+it interactively on all 48 benchmark tasks, working with the system to
+create sufficient examples and provide useful relevance annotations
+(\IE, \emph{essential} and \emph{irrelevant} markings, written as
+\SRequire and \Skip, respectively).
+In all cases, the desired lens was obtained.
+%
+The majority of the tasks required only a single example and none required
+more than three examples to synthesize the desired lens.
+%
+Providing relevance annotations 
+was needed in only 8 of the 48 tasks. In
+practice, we found that adding such annotations was quite easy: if manual inspection
+of the lens showed there were too few \IdentityLens{}s, and too many
+\Disconnect{}s\footnote{
+A \Disconnect{} is a lens combinator that marks a place where
+information appears in one format but not the 
+other: there is a \emph{disconnect} between the two formats. 
+}
+  or merges, we would add \SRequire annotations. If synthesis took
+too long, we would add \Skip annotations.
+%
+We verified that the default running mode of our synthesis tool
+(\SSOpt{}) generated the 
+correct lenses the way programmers often validate their programs: we
+manually inspected the code and ran unit tests on the synthesized
+code. To determine whether the synthesis procedure generated the
+correct lens when running in modes other than \SSOpt{}, we compared
+generated lens to the lens synthesized by \SSOpt{}.
+
+\paragraph*{Effectiveness of Optimizations in Simple Symmetric Synthesis}
+
+Having determined appropriate examples and annotations for the 48
+benchmarks, we evaluate the performance of the system by measuring the running
+time of our algorithm in two modes:
+\vskip .5ex
+\begin{tabulary}{\linewidth}{rL}
+  \SSOpt{}: & Run the symmetric synthesis algorithm with all optimizations enabled.\\
+  \SSNCOpt{}: & Run the symmetric synthesis algorithm with no compositional synthesis enabled.\\
+\end{tabulary}\\
+\noindent
+Compositional synthesis allows users to break a benchmark into a
+series of smaller synthesis tasks, whose solutions are utilized in more complex
+synthesis procedures. Compositional synthesis (\SSOpt{} mode)
+allows our system to scale to arbitrarily large and complex formats; measuring
+it shows the responsiveness of the system when used as
+intended. \SSNCOpt{} mode, which synthesizes a complete lens all at once,
+provides a useful experimental stress test for the system.
+
+\begin{figure}
+  \centering
+  \includegraphics{times}
+  \vspace{-2ex}
+  \caption{Number of
+    benchmarks that can be solved by a given algorithm in a 
+    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
+    \SSNCOpt{} is the symmetric synthesis algorithm without using a library of
+    existing lenses. The symmetric synthesis algorithm is able to complete all
+    benchmarks in under 30 seconds elapsed total time. Without compositional
+    synthesis it is able to complete 26. Each benchmark specification includes
+    source and target (potentially annotated) regular expressions, and between
+    one and three sufficient examples.}
+  \label{fig:times}
+\end{figure}
+
+For each benchmark in the suite and each mode, we ran the system with a timeout
+of 60 seconds, averaging the result over 5 runs. Figure~\ref{fig:times}
+summarizes the results of these tests. We find that our algorithm is able to
+synthesize all of the benchmarks in under 30 seconds. Without compositional
+synthesis, the synthesis algorithm is able to solve 26 out of 48 problem instances.
+
+% Section~\ref{subsec:optimizations} that users may tune the termination metric.
+% Success on this set of benchmarks requires we tune that termination condition
+% (e.g. for some benchmarks, allowing the system flexibility to spend more time
+% considering more of the search space). In
+
+\paragraph*{Slowdown Compared to Bijective Synthesis}
+
+To compare to the existing bijective synthesis algorithm, we run our symmetric
+synthesis algorithm on the original Optician benchmarks, comprised of
+39 bijective synthesis tasks.\footnote{We had to slightly alter four of these
+benchmarks, either by providing additional examples or by adding \SRequire annotations.
+Without these alterations, symmetric synthesis yielded a lens that fit
+the specification but that was undesired.}
+To perform this comparison, we synthesized lenses in two modes:
+\vskip 2ex
+\begin{tabulary}{\linewidth}{rL}
+  \BSOpt{}: & The existing bijective synthesis
+              algorithm with all optimizations enabled.\\
+  \SSOpt{}: & The symmetric synthesis algorithm with all optimizations enabled.\\
+\end{tabulary}\\[2ex]
+\noindent
+For each benchmark, we ran it in both modes with a
+timeout of 60 seconds and averaged the result over 5 runs.
+Figure~\ref{fig:times_bijective} summarizes the results of these tests. On
+average, \SSOpt{} took 1.3 times (0.5 seconds) longer to complete than
+\BSOpt{}. The slowest completed benchmark for both synthesis algorithms is
+\texttt{xml\_to\_augeas.boom}, a benchmark that converts arbitrary XML up to
+depth 3 into a serialized version of the structured dictionary representation
+used in Augeas. This benchmark takes 18.9 seconds for the symmetric synthesis
+algorithm to complete, and 9.3 seconds the bijective synthesis algorithm to
+complete.
+
+\begin{figure}[t]
+  \includegraphics{times_bijective}
+  \vspace{-2ex}
+  \caption{Number of benchmarks that can be solved by a given algorithm in a
+    given amount of time. \SSOpt{} is the full symmetric synthesis algorithm.
+    \BSOpt{} is the full bijective lens synthesis algorithm.}
+  \label{fig:times_bijective}
+\end{figure}
+
+\paragraph*{The Effects of Heuristics and Relevance Annotations}
+We evaluate the usefulness of (1) our information-theoretic metric, (2) our
+termination heuristic and (3) our
+relevance annotations.  To this end, we run our program in several different modes:\\[2ex]
+\begin{tabulary}{\linewidth}{rL}
+  \AnyOpt{}: & Ignore the information-theoretic metric (\IE, all valid
+               lenses have cost 0). \\ 
+  \FLOpt{}: &  Return the first highest ranked lens \GreedySynth returns (\IE,
+              ignore the termination heuristic). \\
+  \CCOpt{}: & Replace our information-theoretic cost metric with one where
+              the cost of the lens is the number of disconnects plus the number
+              of merges.\\
+  \NSOpt{}: & Ignore all \Skip annotations in the SRE specifications. \\
+  \NROpt{}: & Ignore all \SRequire annotations in the SRE specifications. \\[2ex]
+\end{tabulary}
+
+\noindent
+We experimented with the \CCOpt{} mode to determine whether the
+complexity of the information-theoretic measure was really
+needed. Related work on string transformations has often used simpler
+measures such as ``avoid constants'' that align with, but are simpler
+than our measures. The \CCOpt{} mode is an example of such a simple
+measure---it operates by counting disconnects, which put a complete
+stop to information transfer, and merges, which
+eliminate the information in a union.
+
+\begin{figure}[t]
+  \includegraphics{metrics_importance}
+  \vspace{-2ex}
+  \caption{Number of benchmarks that synthesize the correct lens by a given
+    algorithm. \AnyOpt{} provides no notion of cost, and merely returns the
+    first lens it finds that satisfies the specification. \FLOpt{} provides a
+    notion of cost to \GreedySynth, but once a satisfying lens is greedily
+    found, that lens is returned. \CCOpt{} synthesizes lenses, where the cost of
+    a lens is the number of disconnects plus the number of merges. \NSOpt{}
+    ignores all \Skip annotations while running the algorithm. \NROpt{} ignores
+    all \SRequire annotations while running the algorithm.}
+  \label{fig:metric}
+\end{figure}
+
+Figure~\ref{fig:metric} summarizes the result of these experiments. The data
+reveal that the information-theoretic metric is critical for finding the correct
+lens: Only 10 of the benchmarks succeeded when running in \CCOpt{} mode. The
+termination condition is also quite important. When running in \FLOpt{} mode,
+the algorithm only discovers 5 lenses, which shows that the first class that
+contains a satisfying lens is rarely the correct class. However, our algorithm
+is not perfect and fails when either it is very difficult to find the desired
+lens (necessitating \SRequire) or when a large amount of data is projected
+(necessitating \Skip). Without any annotations, our algorithm finds the correct
+lens for 40 of our 48 benchmarks; eight required relevance annotations to find
+the correct lens.
+% end evaluation
+
 \section{Conclusions}

### 2019-03-03 report/final-report.tex
added to list of acronyms and symbols

@@
 we summarize our key methods, results, and findings, focusing
-on the core algorithms for bijective synthesis and quotient synthesis.
+on the core algorithms for bijective synthesis, quotient synthesis and
+simple symmetric synthesis.
 
@@
 transformations our system can define significantly.  Finally, we explored the
-synthesis of \emph{symmetric lenses}~\cite{symmetric-lenses} , again building upon the
+synthesis of \emph{symmetric lenses}~\cite{symmetric-lenses}, again building upon the
 bijective platform that we started with.  Symmetric lenses are able to
 ignore arbitrary chunks of data in A when generating target data B,
-and vice-versa.  For example, a serial number specific to one data set
+and vice versa.  For example, a serial number specific to one data set
 may be ignored when we translate to a second.  Symmetric lenses
@@
 \begin{figure}[t]
+\centering
 \includegraphics{qfigs/asts.eps}
@@
 \section{List of Symbols, Abbreviations, and Acronyms}
-\begin{tabular}{ll}
+\begin{tabular}{p{1in}p{5in}}
+  \ToDNFRegex{}         & An operation that converts regular   expressions into DNF form\\
+  \ToLens{}             & An operation that converts a DNF lens to a
+  lens \\
+  $\equiv_S$, $\equiv_T$ & Equivalence relations on data matching
+                           regular expressions $S$ and $T$
+                           respectively \\
+  \AnyOpt{}             & Mode for simple symmetric synthesis that ignores the information-theoretic metric \\ 
+  AST     & Abstract syntax tree\\
   DARPA   & Defense Advanced Research Project Agency \\
+  \CCOpt{}              & Mode for simple symmetric synthesis that
+                          replace our information-theoretic cost metric with one where 
+                          the cost of the lens is the number of disconnects plus the number
+                          of merges.\\
+  \DNFLens{}            & Lens in disjunctive normal form\\
+  DNF                   & Disjunctive normal form\\
+  \DNFRegex{}, \DNFRegexAlt{}  & Regular expression in disjunctive normal form\\
+  \Examples{}           & Examples \\
+  \FLOpt{}              & Mode for simple symmetric synthesis that
+                          returns the first highest ranked lens found
+                          by  \GreedySynth \\
+  \Lens{}               & A lens \\
+  \NROpt{}              & Mode for simple symmetric synthesis that
+                          ignores all \SRequire annotations in the SRE specifications. \\
+  \NSOpt{}              & Mode for simple symmetric syntehsis that
+                          ignores all \Skip annotations in the SRE specifications. \\ 
+  QRE                   & Quotient regular expression\\
+  SRE                   & Stochastic regular expression\\
+  \BRegex and \BRegexAlt & Regular expressions provided by the user\\
+  \Regex{}, \RegexAlt{} & Regular expressions\\
+  $S \Leftrightarrow T$ & Type of a lens that maps between data
+                          described by regular expressions $S$ and $T$\\
+  \SSOpt{}              & Full symmetric synthesis algorithm\\
+  \SSNCOpt{}            & Symmetric synthesis algorithm without using a library of
+    existing lenses.\\
+  \BSOpt{}              & Bijective lens synthesis algorithm\\
+
 \end{tabular}

### 2019-03-04 report/final-report.tex
Checkpoint

@@
 
-  \large \textbf{Synthesizing Data Wranglers\\\vspace{1cm} FA8750-17-2-0028}
-
   \vspace{1in}
 
-  \normalsize
+  \Large \textbf{Synthesizing Data Wranglers\\\vspace{1cm} FA8750-17-2-0028}
+
+  \vspace{1in}
+
+  \large
   %\begin{tabular}{rl}
@@
 software adaptors that convert between two different representations
-of the same information.  Lenses are a special form of program that
+of the same information~\cite{Focal2005}.  Lenses are a special form of
+programs that 
 guarantee so-called \emph{round-tripping properties} that ensure the
 transformations are well-behaved.  Our synthesis algorithms take as
-input (i) the \emph{type} of each source as a regular expression, as
-well as (ii) a small collection of representative \emph{examples} of
-the desired translation.  They produceas output \emph{bijective string
-  lenses}.  Bijective string lenses are a special class of lenses that
-only exist when the two data sources are related via a bijection.
-Despite this limitation, bijections and bijective synthesis form a
-useful foundation on top of which more general transformations between
-data sets may be built.  We explore two of these extensions: quotient
-lenses and symmetric lenses.  Quotient lenses apply when two formats
-are ``morally'' but not technically bijective.  Examples include two
+input (i) the \emph{type} of each source as a regular expression, and
+(ii) a small collection of representative \emph{examples} of
+the desired translation.  They produce as output different varieties of \emph{string
+  lenses}.  Have investigated three specific varieties in detail: {\em
+  bijective lenses}, {\em quotient
+lenses} and {\em simple symmetric lenses}.  Bijective lenses address the
+simplest case, when the two formats we want to convert between have exactly
+the same information content, modulo rearrangement and reformatting.  
+Quotient lenses apply when two
+formats are ``morally'' but not technically bijective.  Examples include two
 formats that differ in their use of white space or in the order of
-sequences of elements.  Symmetric lenses allow one side or the other to have
-information not present in the other.  When converting between the
+sequences of elements.  Simple symmetric lenses allow one side or the other
+to have 
+information not present in the other; when converting between the
 formats, defaults take the place of missing information.
@@
 
-We have written
-papers~\cite{bijective-synthesis,quotient-synthesis,symmetric-synthesis}
-describing and evaluating our algorithms. We produced open-source 
+The products of this investigation include several published
+papers~\cite{bijective-synthesis, quotient-synthesis, symmetric-synthesis}
+describing and evaluating our algorithms, as well as open-source 
 code~\cite{GitHub} that implements those algorithms.  In this report,
 we summarize our key methods, results, and findings, focusing
-on the core algorithms for bijective synthesis, quotient synthesis and
+on the core algorithms for bijective synthesis, quotient synthesis, and
 simple symmetric synthesis.

### 2019-03-04 report/final-report.tex
Checkpoint new section 3

@@
 
+% Colors
+\definecolor{dkblue}{rgb}{0,0.1,0.5}
+\definecolor{dkgreen}{rgb}{0,0.6,0}
+\definecolor{dkred}{rgb}{0.6,0,0}
+\definecolor{dkpurple}{rgb}{0.2,0,0.4}
+\definecolor{olive}{rgb}{0.4, 0.4, 0.0}
+\definecolor{teal}{rgb}{0.0,0.5,0.5}
+\definecolor{orange}{rgb}{0.9,0.6,0.2}
+\definecolor{lightyellow}{RGB}{255, 255, 240}
+\definecolor{vlightyellow}{RGB}{255, 255, 245}
+\definecolor{lightgreen}{RGB}{235, 255, 240}
+\definecolor{vlightgreen}{RGB}{247, 255, 250}
+\definecolor{lightblue}{RGB}{245, 255, 255}
+\definecolor{vlightblue}{RGB}{250, 255, 255}
+\definecolor{teal}{RGB}{141,211,199}
+\definecolor{darkbrown}{RGB}{121,37,0}
+
 \newtheorem*{theorem*}{Theorem}
@@
 
-\section{Methods, Assumptions and Procedures}
-
-Our approach combines two strategies.  We use \textit{domain-specific programming
-  language}, specifically, the Boomerang~\cite{boomerang, Matching10} language,
-which is designed for writing bidirectional string transformations.  We also exploit
-\textit{type- and example-directed program synthesis}, which uses enumerative
-techniques to search a space of candidate solutions.  The search is constrained
-by the type information about the program, as well as by example instances
-provided by the user.
-
-\subsection{Lenses and lens domain-specific programming languages}
+\section{Lenses}\bcp{Is a separate section the best place for this?  Is
+  this the kind of example we had in mind?  Is it OK that I cannibalized the
+later subsection on lenses?  (We could alternatively move all this there.)}
+
+\begin{figure}
+\small
+  % https://tex.stackexchange.com/questions/168741/showing-two-listings-in-a-table-side-by-side
+  \setbox0=\hbox{%
+    \begin{minipage}{3.5in}
+\begin{lstlisting}[
+stringstyle=\upshape\sffamily,
+xleftmargin=0pt,
+]
+Jane Doe: 38000
+John Public: 37500
+\end{lstlisting}
+    \end{minipage}
+  }
+  \savestack{\listingA}{\box0}
+
+  \setbox0=\hbox{%
+    \begin{minipage}{3.5in}
+\begin{lstlisting}[
+xleftmargin=0pt,
+stringstyle=\upshape\sffamily
+]
+FirstLast,Company
+Jane Doe,Healthcare Inc.
+John Public,Insurance Co.
+\end{lstlisting}
+    \end{minipage}
+  }
+  \savestack{\listingB}{\box0}
+
+  \setbox0=\hbox{%
+    \begin{minipage}{3.5in}
+\begin{lstlisting}[xleftmargin=0pt]
+let salary = number | "unk"
+let emp_salary =   name . " "
+                 . name . ": " 
+                 . salary
+let emp_salaries = "" 
+                 |   emp_salary
+                   . ("\n" emp_salary)$^*$
+\end{lstlisting}
+    \end{minipage}
+  }
+  \savestack{\listingC}{\box0}
+
+  \setbox0=\hbox{%
+    \begin{minipage}{3.5in}
+\begin{lstlisting}[xleftmargin=0pt]
+let company = (co_name . ("Co." | "Inc." | "Ltd.")) 
+            | "UNK" 
+let emp_ins = name . " " . name "," company 
+let header = "FirstLast,Company" 
+let emp_insurance = header . ("\n" . emp_ins)$^*$
+\end{lstlisting}
+    \end{minipage}
+  }
+  \savestack{\listingD}{\box0}
+
+  \centering
+  \resizebox{\columnwidth}{!}{%
+    \begin{tabular}{|>{\columncolor{vlightyellow}}c|>{\columncolor{vlightblue}}c|}
+      \hline
+      Management's data & HR's data \\
+      \hline
+      \listingA & \listingB \\
+      \hline
+      \cellcolor{lightyellow}Management's type & 
+      \cellcolor{lightblue} HR's type\\
+      \hline
+      \cellcolor{lightyellow}\listingC & \cellcolor{lightblue}\listingD \\
+      \hline
+    \end{tabular}
+  }
+  \caption{Hypothetical example data files and corresponding regular expressions
+  used by management and HR at a US company to represent employee salaries and health
+  insurance providers, respectively.
+  }
+  \label{fig:minimized-representations}
+\end{figure}
+
+Before surveying our methods and results on synthesis, we quickly sketch an
+example of what a lens looks like, using a simplified situation drawn from a
+hypothetical U.S. company. In this company, management and human resources
+(HR) store information about employees in separate text files: management
+stores the names of employees and their salaries while HR stores the names
+of employees and their health insurance
+providers. Figure~\ref{fig:minimized-representations} gives examples of the
+two file formats and regular expressions describing them.
+
+The company uses a simple symmetric lens to keep these files
+synchronized. When management adds a new employee, say ``Chris Roe: 32500'',
+this lens 
+adds the corresponding entry ``Chris Roe, UNK'' to HR's file. The default value
+``UNK'' represents the fact that the employee's insurance company is currently
+unknown. A similar update happens if HR adds a new employee before
+management knows about them,
+in which case the sentinel value ``unk'' reflects unknown salary information
+on management's side.
+Furthermore, if HR corrects an error in an employee's name, say changing ``John
+Public'' to ``Jon Public'', the lens mirrors this change into management's
+file. Not all the data is mirrored, however: management's file is not updated
+in response to insurance changes, and HR's is oblivious to salary
+changes. Simple symmetric lenses are appropriate for keeping these two files
+synchronized because they contain a mix of shared and unshared information.
+
+%Consider the previously
+%described goal of keeping Linux cron jobs (crontab) and Windows Scheduled Tasks
+%(scheduled.job) synchronized. These are each formats specifying when to run
+%recurring tasks. For the sake of exposition, we use simplified versions crontab
+%and scheduled.job formats, with examples and corresponding regular expressions
+%shown in Figure~\ref{fig:minimized-representations}. Each file contains a list
+%of tasks set to run at specific times every hour. In both formats, users
+%can either provide specific minutes at which they want the program to run, or they
+%can specify that the program should run at every minute.
+%
+%We wish for the jobs to be synchronized across our Windows and Linux devices.
+%However, the commands that need to be run are typically different for each
+%device. Paths are different in Windows and Linux, and, while there is typically a
+%Windows equvialent for a Linux command, that equivalent command is not always
+%named the same thing. So, while we want to be able to have the same number of
+%commands, with the commands running at the same times, we don't want to
+%align the actual commands that are run.
+
+In principle, programmers can define lenses just by writing four arbitrary
+functions by hand and showing they satisfy the round-tripping laws, but this
+can be tedious and error prone.  A better idea is to provide a set of
+combinators---a domain-specific language---for building complex, law-abiding
+lenses from simpler ones.
 
@@
 XML transformation languages~\cite{DBLP:conf/pepm/LiuHT07}.
+
+If $\BRegex$ and $\BRegexAlt$ are regular expressions, then
+\lstinline{$\Lens$ : $\BRegex \Leftrightarrow \BRegexAlt$} indicates that
+$\Lens$ is a simple symmetric lens between $\LanguageOf{\BRegex}$ and
+$\LanguageOf{\BRegexAlt}$. (We will use undecorated variables later for
+stochastic regular expressions, so we mark plain REs with overbars.)
+We illustrate some of these
+combinators by defining lenses on subcomponents of the employee data formats.
+
+The simplest combinator is the identity lens \IdentityLens, which takes as an
+argument a regular expression $\BRegex$ and propagates data unchanged
+in both directions.
+%
+\begin{lstlisting}
+id(name)  :  name $\Leftrightarrow$ name
+\end{lstlisting}
+%
+The identity lens moves data back and forth from source to target without
+changing it. Both the \CreateR and \CreateL functions are the identity function
+($\CreateR \App s = s$), and the put functions merely return the first argument
+($\PutROf{\String}{\StringAlt} = \String$).
+
+In contrast, $\DisconnectOf{\Regex}{\RegexAlt}{\String}{\StringAlt}$ does not propagate any data at all
+from one format to the other.  The \Disconnect lens takes four arguments: two
+regular expressions ($\Regex, \RegexAlt$) and two strings ($\String, \StringAlt$). The regular expressions specify the formats
+on the two sides, while the strings provide default values. 
+%
+\begin{lstlisting}
+disconnect(salary, "", "unk", "")  :  salary $\Leftrightarrow$ ""
+\end{lstlisting}
+%
+On creates, the input values are thrown away, and default values are
+returned ($\CreateLOf{\StringAlt} =
+\lstinline{"unk"}$), and on puts, the second argument is used and the first is
+thrown away ($\PutROf{\String}{\StringAlt} = \StringAlt$). For 
+example, the \lstinline{salary} field is only present in management files,
+so the disconnect lens can ensure salary edits do not cause updates to the
+HR file.
+With the above lens, $\PutL \App \lstinline{""} \App 60000$ will return 60000,
+and \PutR will always return \lstinline{""}.
+
+The insert lens \lstinline{ins} and the delete lens \lstinline{del} are
+syntactic sugar for uses of the disconnect lens in which a string
+constant is omitted entirely from the source or target format.
+\begin{lstlisting}
+ ins($t$) = disconnect("", $t$, "", $t$)
+ del($s$) = disconnect($s$, "", $s$, "")
+\end{lstlisting}
+The \lstinline{ins} lens inserts a constant string when going from
+left to right, while \lstinline{del} inserts a string when going from right
+to left.
+
+Finally, there are a number of combinators that construct
+more complex lenses from simpler ones. These include concatenation
+($\ConcatLensOf{\Lens_1}{\Lens_2}$, often written $\Lens_1 . \Lens_2$), variation
+($\OrLensOf{\Lens_1}{\Lens_2}$), and iteration ($\IterateLensOf{\Lens}$).
+For example, we could use the lens
+%
+\begin{lstlisting}
+id(name) . id(" ") . id(name) . del(": ") . ins(",")
+\end{lstlisting}
+%
+to transform ``\lstinline{Jane Doe: }'' to ``\lstinline{Jane Doe,}'' in
+the left-to-right direction.
+
+% concat(ins(header),employees_lens) : emp_salaries $\Leftrightarrow$ emp_insurance
+%
+%\ksf{I don't think this example is necessary.}
+%Data can also appear in multiple forms. For example,
+%a list of employees with insurance information can either be empty or
+%be a sequence of employee/insurance records: \ksf{This example is
+%vacuous as you could eliminate the or in favor of the *.  Do somethign
+%else to introduce or?}
+%
+%\begin{lstlisting}
+%or(id(""),employees_lens) : emp_salaries $\Leftrightarrow$ "" | ("\n" emp_ins) . ("\n" emp_ins)*
+%\end{lstlisting}
+The \IterateLens lens is useful for synchronizing a series of items or rows in a
+table. For example given a
+lens \lstinline{employee_lens} that synchronizes data for a single employee, the
+lens
+%
+\begin{lstlisting}
+iterate(id("\n") . employee_lens)  :  ("\n" . emp_salary)$^*$ $\Leftrightarrow$ ("\n" . emp_ins)$^*$
+\end{lstlisting}
+%
+transforms a list of employees in employees in the Management format to a list
+of employees in the HR format and vice versa. These combinators
+are combined in figure~\ref{fig:example_lens} to construct a complete lens
+between the employee formats.
+
+\begin{figure}
+\begin{lstlisting}
+let name_lens = id(name) . id(" ") . id(name) . del(": ") . ins(",") in
+let employee_lens = name_lens . disconnect(salary,"","unk","") 
+                                                    . disconnect("",company,"","UNK") in
+let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
+ins(header) . employees_lens  :  emp_salaries $\Leftrightarrow$ emp_insurance
+\end{lstlisting}
+%\ksf{I don't think this is necesssary.}
+%let emp_lens = or(id(""),employees_lens) in
+  \caption{A lens that synchronizes management and HR employee files}
+  \label{fig:example_lens}
+\end{figure}
+
+\section{Methods, Assumptions and Procedures}
+
+Our approach to lens synthesis combines two strategies.  First, as
+illustrated in the previous section, we use \textit{domain-specific
+  programming language}, specifically the Boomerang  language~\cite{boomerang,
+  Matching10}, which is designed for writing bidirectional string
+transformations.  Second, we exploit \textit{type- and example-directed
+  program synthesis}, which uses enumerative techniques to search a space of
+candidate solutions.  The search is constrained by the type information
+about the program, as well as by example instances provided by the user.
+
+% \subsection{Lenses and lens domain-specific programming languages}
 
@@
 
-During the course of this project, we have designed, analyzed and
-implemented algorithms that demonstrate it is possible to synthesize
-three classes of bidirectional transformations: (1) pure bijective
-transformations, (2) bijections modulo equivalences classes (quotient
-lenses) and (3)
-bijections modulo projections (symmetric lenses).  The synthesis algorithms we have
-designed take inputs that include a format specification (which
-includes specification of equivalence classes) and a collection of
-examples.  As the class of transformations becomes richer, the number
-of potential programs grows dramatically.  As a result, the
-corresponding inference algorithm slows and its ability to guess the
-transformation desired by the user decreases, leading to reduced
-accuracy.  We found that it was possible to overcome such challenges
-through new heuristics that use information theory to help guide the
-search for program transformations.  In addition, we found the use of
-compositional synthesis, which is the process of breaking down a
-complex synthesis task into smaller, more mangeable subtasks critical
-when attempting to scale our prototype system up to be able to handle
-larger transformation task.
+In this project, we designed, analyzed and implemented algorithms that
+synthesize three classes of bidirectional transformations: (1) pure
+bijective transformations, (2) bijections modulo equivalences classes
+(quotient lenses) and (3) bijections modulo projections (simple symmetric
+lenses).  Our synthesis algorithms take as inputs a format specification
+(which includes specification of equivalence classes) and a collection of
+examples.  As the class of transformations becomes richer, the number of
+potential programs grows dramatically.  As a result, the corresponding
+inference algorithm slows and its ability to guess the transformation
+desired by the user decreases, leading to reduced accuracy.  We overcame
+these challenges through new heuristics that use information theory to help
+guide the search for program transformations.  In addition, to scale our
+prototype to larger transformation tasks, we found it necessary to work
+compositionally, by breaking down complex synthesis tasks into smaller, more
+mangeable subtasks.
 

### 2019-03-04 report/final-report.tex
intro
typo
stuff

@@
 the desired translation.  They produce as output different varieties of \emph{string
-  lenses}.  Have investigated three specific varieties in detail: {\em
+  lenses}.
+
+We have investigated three specific varieties of string lenses in detail: {\em
   bijective lenses}, {\em quotient
@@
 \end{subfigure}
-\caption{Runtimes measurements. In (a), we run Optician and \QOpt{}
+\caption{Runtime measurements. In (a), we run Optician and \QOpt{}
   on the Optician benchmarks.  We find that there is only a negligible
@@
   AST     & Abstract syntax tree\\  
+  \BSOpt{}              & Bijective lens synthesis algorithm\\
   DARPA   & Defense Advanced Research Project Agency \\
@@
                           by  \GreedySynth \\
+    JSON & JavaScript Object Notation---a standard file format used to
+         encode data objects\\
   \Lens{}               & A lens \\
@@
   QRE                   & Quotient regular expression\\
+    RDF & Resource Description Framework---used to model web resources \\
   SRE                   & Stochastic regular expression\\
@@
     existing lenses.\\
-  \BSOpt{}              & Bijective lens synthesis algorithm\\
-
+  XML & Extensible Markup Language---used in web services \\
 \end{tabular}

### 2019-03-04 report/final-report.tex
moving lens section

@@
 
-\section{Lenses}\bcp{Is a separate section the best place for this?  Is
-  this the kind of example we had in mind?  Is it OK that I cannibalized the
-later subsection on lenses?  (We could alternatively move all this there.)}
+\section{Methods, Assumptions and Procedures}
+
+Our approach to lens synthesis combines two strategies.  First, 
+we use a \textit{domain-specific
+  programming language}, Boomerang~\cite{boomerang,Matching10},
+which is designed for writing bidirectional string
+transformations.  Second, we exploit \textit{type- and example-directed
+  program synthesis}, which uses enumerative techniques to search a space of
+candidate solutions.  The search is constrained by the type information
+about the program, as well as by example instances provided by the user.
+
+% \subsection{Lenses and lens domain-specific programming languages}
+
+\subsection{Lenses}
 
@@
 Before surveying our methods and results on synthesis, we quickly sketch an
-example of what a lens looks like, using a simplified situation drawn from a
+example of what a Boomerang lens looks like, using a simplified situation drawn from a
 hypothetical U.S. company. In this company, management and human resources
@@
 \end{figure}
-
-\section{Methods, Assumptions and Procedures}
-
-Our approach to lens synthesis combines two strategies.  First, as
-illustrated in the previous section, we use \textit{domain-specific
-  programming language}, specifically the Boomerang  language~\cite{boomerang,
-  Matching10}, which is designed for writing bidirectional string
-transformations.  Second, we exploit \textit{type- and example-directed
-  program synthesis}, which uses enumerative techniques to search a space of
-candidate solutions.  The search is constrained by the type information
-about the program, as well as by example instances provided by the user.
-
-% \subsection{Lenses and lens domain-specific programming languages}
 

### 2019-03-04 report/final-report.tex
Checkpoint

@@
 
-We have implemented our synthesis algorithms in
-of OCaml and integrated them into Boomerang system,
-where they are available as open source code~\cite{GitHub}.
-Users of Boomerang can write lenses by hand, specify them as
-synthesis tasks, or do a combination of both.
-
-We have analyzed our algorithms theoretically, proving various
-soundness and completeness results.  We have also analyzed our
-algorithms empirically, studying their impact on a range of real and
-synthetic benchmarks.  Our full results have been published in a
-series of conference papers and technical
-reports~\cite{bijective-synthesis,quotient-synthesis,symmetric-synthesis}.  In the rest of this
-section, we summarize some of the key results.
+We implemented our synthesis algorithms in the OCaml programming language
+and integrated them into Boomerang system, where they are available as open
+source code~\cite{GitHub}\bcp{Two more pointers needed}.  Users of Boomerang
+can write lenses by hand, specify them as synthesis tasks, or do a
+combination of both.
+
+We analyzed our algorithms theoretically, proving various soundness and
+completeness results.  We also analyzed these algorithms empirically,
+studying their impact on a range of real and synthetic benchmarks.  Our
+results have been published in a series of conference papers and technical
+reports~\cite{bijective-synthesis,quotient-synthesis,symmetric-synthesis}.
+In the rest of this section, we summarize key results.
 
@@
 
-We first study the effectiveness of the pure bijective synthesis
+We first show the effectiveness of the pure bijective synthesis
 algorithm by evaluating its performance on a set of 39
@@
 Augeas is a configuration editing system for Linux that uses lens
-combinators similar to those in Boomerang. However, it transforms
-strings on the left to structured trees on the right rather than
-transforming strings to strings.
-We adapted these Augeas lenses to our setting by converting the
-right-hand sides to strings that correspond to serialized versions
-of the tree formats.  
+combinators similar to those in Boomerang.  
+% , with some small
+% differences. However, it transforms 
+% strings on the left to structured trees on the right rather than
+% transforming strings to strings.
+% We adapted these Augeas lenses to our setting by converting the
+% right-hand sides to strings that correspond to serialized versions
+% of the tree formats
 We derived 29 of the benchmark tests by
-adapting the first 27 lenses in alphabetical order, as well as the lenses
+adapting the first 27 lenses from the Augeas library in alphabetical order,
+as well as the lenses 
 \CF{aug/xml-firstlevel} and \CF{aug/xml} that were referenced
@@
 Furthermore, the 12 last synthesis problems derived
-from Augeas were tested after \Optician{} was
+from Augeas were tested after our synthesizer was was
 finalized, demonstrating that the optimizations were not
-overtuned to perform well on the testing data.
+over-tuned to perform well on the testing data.
 %
@@
 transformations by example~\cite{gulwani-popl-2014}.  
-We derived three benchmarks from the first few examples in the
+We derived three benchmarks from the first few examples in this
 paper and one from the running example on
@@
 To test our system on these benchmarks, we had to convert them into
-bijective transformations by hand.  More specifically, in many of the
+bijective transformations (by hand).  Specifically, in many of the
 benchmarks, we normalized
@@
 
-Finally, we added custom examples to highlight weaknesses of
-our algorithm (\CF{cap-prob} and \CF{2-cap-prob}) 
+Finally, we added a few custom examples designed to highlight known
+weaknesses of our algorithm
+%(\CF{cap-prob} and \CF{2-cap-prob}) 
 and to test situations for which we thought the tool would be
-particularly useful (\CF{workitem-probs}, \CF{date-probs}, \CF{bib-prob},
-and \CF{addr-probs}).   These examples convert between work item formats, date
+particularly useful.
+% (\CF{workitem-probs}, \CF{date-probs}, \CF{bib-prob},
+% and \CF{addr-probs}).
+These examples convert between work item formats, date
 formats, bibliography formats, and address formats, respectively.

### 2019-03-04 report/final-report.tex
acronyms

--- 2019-03-04 report/final-report.tex
@@
   \Examples{}           & Examples \\
+  \Expand       & A component of the synthesis algorithm that proposes candidate
+                  types equivalent to the input
   \FLOpt{}              & Mode for simple symmetric synthesis that
@@
                           by  \GreedySynth \\
+  \GreedySynth & Synthesis procedure that searches greedily for candidate solutions
+  HR & Human Resources \\
+  HTML & Hyptertext Markup Language \\
   JSON & JavaScript Object Notation---a standard file format used to
@@
   QRE                   & Quotient regular expression\\
+  \textbf{QS}  & Number of nodes in the abstract syntax of QRE specifications \\
   RDF & Resource Description Framework---used to model web resources \\
+  REST & Representational State Transfer\\
   SRE                   & Stochastic regular expression\\

### 2019-03-04 report/final-report.tex
fix

@@
   \Expand       & A component of the synthesis algorithm that proposes candidate
-                  types equivalent to the input
+                  types equivalent to the input \\
   \FLOpt{}              & Mode for simple symmetric synthesis that
@@
                           by  \GreedySynth \\
-  \GreedySynth & Synthesis procedure that searches greedily for candidate solutions
+  \GreedySynth & Synthesis procedure that searches greedily for candidate
+                 solutions \\
   HR & Human Resources \\

### 2019-03-04 report/final-report.tex
added refs to optician, a few more acronyms

@@
 describing and evaluating our algorithms, as well as open-source 
-code~\cite{GitHub} that implements those algorithms.  In this report,
+code~\cite{GitHub} that implements those algorithms in (several variants of) a
+prototype tool called \Optician{}.  In this report,
 we summarize our key methods, results, and findings, focusing
@@
 describing and evaluating our algorithms and produced open-source
-code~\cite{GitHub} that implements those algorithms.  In this report,
-we summarize some of our key methods, results and findings, focusing
-on the core algorithms for bijective synthesis and quotient synthesis.
+code~\cite{GitHub} that implements those algorithms in (several variants of) a
+tool called \Optician{}.  In this report, we summarize some of our key methods,
+results and findings, focusing on the core algorithms for bijective synthesis
+and quotient synthesis.
 
@@
   \ToDNFRegex{}         & An operation that converts regular   expressions into DNF form\\
-  \ToLens{}             & An operation that converts a DNF lens to a
-  lens \\
+  \ToLens{}             & An operation that converts a DNF lens to a  lens \\
   $\equiv_S$, $\equiv_T$ & Equivalence relations on data matching
@@
   \Lens{}               & A lens \\
+  $\mathcal{L}(\BRegex{})$  & The language of strings for the regular expression
+                             \BRegex{} \\
   \NROpt{}              & Mode for simple symmetric synthesis that

### 2019-03-04 report/final-report.tex
more acronyms

@@
 \usepackage[explicit]{titlesec}
+\usepackage{longtable}
 
@@
 \section{List of Symbols, Abbreviations, and Acronyms}
-\begin{tabular}{p{1in}p{5in}}
+\begin{longtable}{p{1in}p{5in}}
   \ToDNFRegex{}         & An operation that converts regular   expressions into DNF form\\
@@
                           by  \GreedySynth \\
-  \GreedySynth & Synthesis procedure that searches greedily for candidate
+  \footnotesize{\GreedySynth} & Synthesis procedure that searches greedily for candidate
                  solutions \\
@@
                           ignores all \Skip annotations in the SRE specifications. \\ 
+  \Optician{} & Our prototype implementation of the lens synthesis algorithms \\
   QRE                   & Quotient regular expression\\
@@
     existing lenses.\\
+  {\footnotesize \SynthDNFLens{}} & A type-directed lens synthesis algorithm \\
   XML & Extensible Markup Language---used in web services \\
-\end{tabular}
+\end{longtable}
 

### 2019-03-04 report/final-report.tex
Checkpoint

@@
 practice, we \textit{randomly} generated appropriate source/target
-pairs, mimicking what a na\"{i}ve user might do.  We did not write the
+pairs, mimicking what a na\"{i}ve user might do.  (We did not write the
 examples by hand out of concern that our knowledge of the synthesis
-algorithm might bias the selection. Figure~\ref{fig:exs-reqd} shows
+algorithm might bias the selection.) Figure~\ref{fig:exs-reqd} shows
 the number of randomly generated examples it takes to synthesize the
-correct lens averaged over ten runs.  The synthesis algorithm almost never needs
-any examples: only 5 benchmarks need a nonzero number of examples to
-synthesize the correct lens and only one, \CF{cust/workitem-probs} required over
+correct lens, averaged over ten runs.  Strikingly, the synthesis algorithm
+almost never needs 
+any examples at all: only 5 benchmarks need a nonzero number of examples to
+synthesize the correct lens and only one required more than
 10 randomly generated examples.
-A clever user may be able to reduce the
-number of examples further by selecting examples carefully; we
-synthesized \CF{cust/workitem-probs} with only 8 examples.
+(A clever user may be able to reduce the
+number of examples further by selecting examples carefully; we were able to
+synthesize this last test case with only 8 examples.)
 
@@
 map source data to the target (and vice versa) require the most examples.
-For example, the benchmark \CF{cust/workitem-probs} requires a large number of
+For example, one benchmark requires a large number of
 examples because it
@@
 %
-We are the first tool to synthesize bidirectional transformations between data
+Ours is the first tool to synthesize bidirectional transformations between data
 formats, so there is no tool to which we can make an apple-to-apples comparison.
-Instead, we compare against tools for generating unidirectional
-transformations instead. 
+Instead, we compared against tools for generating unidirectional
+transformations. 
 Figure~\ref{fig:synthesis-times} includes a comparison against two other
@@
 Fill), or marked text regions (for FlashExtract), as opposed
-to \Optician{}'s use of regular expressions to constrain the format of
+to our use of regular expressions to constrain the format of
 the input and output.  By using regular expressions,
-\Optician{} is able to synthesize significantly more programs
+we can synthesize significantly more programs
 than either existing tool.
@@
 In particular, neither Flash Fill nor FlashExtract was able to synthesize
-transformations or extractions present under two iterations, a type of format
+transformations or extractions where the regular expressions describing the
+formats involved two nested iteration (Kleene-star) operators---a type of format
 that is notoriously hard to infer.
-These types of dual iterations are pervasive in Linux configuration
-files, making Flash Fill and FlashExtract ill suited for many of the synthesis
-tasks present in our test suite.
-
-Furthermore, as unidirectional transformations, Flash Fill and FlashExtract have
-a more expressive calculus.  To guarantee bidirectionality, our syntax must be
-highly restrictive, providing a smaller search space to traverse.
+However, this kind of nested iteration is pervasive in Linux configuration
+file formats, making Flash Fill and FlashExtract ill suited for many of the
+synthesis tasks in our test suite.
+
+Conversely, since they deal with only unidirectional transformations, Flash
+Fill and FlashExtract can offer a more expressive set of operators.  To
+guarantee bidirectionality, our syntax must be highly restrictive, providing
+a smaller search space to traverse.
 
@@
 
-We have extended the bijective synthesis tool to synthesize quotient lenses. We will use 
-``\QOpt'' to denote our extended version of Optician, and just plain ``Optician'' to denote
-the pre-quotient version of Optician. The synthesis algorithm produces Boomerang lens
-values, so Boomerang gives synthesized lenses the same first-class status as hand-written ones. 
+Our next thread of work was extending the bijective synthesis tool to
+synthesize quotient lenses. In this subsection, we will use ``\QOpt'' to
+denote our extended version of Optician, and just plain ``Optician'' to
+denote the pre-quotient version of Optician. The synthesis algorithm
+produces Boomerang lens values, so Boomerang gives synthesized lenses the
+same first-class status as hand-written ones.
 %
@@
 
-We analyzed the same 39 lens synthesis tasks from the original
+We analyzed the same 39 lens synthesis tasks as we did for the original
 Optician system.
@@
 statistics, wage comparosion data, and crime index data).
-Recall that to use bijective synthesis,  we had to modify 10 of 39
-benchmarks make them bijective (for instance, by normalizing whitespace)
+Recall that, to use bijective synthesis, we had to modify some of the 
+benchmarks (10 of 39) to make them bijective---for instance, by normalizing whitespace.
 Because quotient lenses are more expressive, we were able avoid such
@@
 
-\paragraph*{Evaluating programmer effort}
+\paragraph*{Evaluating Programmer Effort}
 
@@
 
-For each problem in the benchmark suite, we calculate the following
-measures as proxies for the level of programmer effort when using each
-the three approaches:
+For each problem in the benchmark suite, we calculate the following measures
+as proxies for the level of programmer effort when using each the three
+approaches: \bcp{This seems very detailed, and it includes some technical
+  terms (like ``canonize(q)'') that readers of this summary may not
+  know.  But maybe it's OK?}
 

### 2019-03-04 report/final-report.tex
opticianq
Checkpoint

@@
 
-\subsection{Simple Symmetric Synthesis Algorithms}
-We implemented simple symmetric lenses as an extension to
+\subsection{Simple Symmetric Lens Synthesis}
+Finally, we implemented simple symmetric lenses as an extension to
 Boomerang~\cite{boomerang}.  
-We also integrated our synthesis engine into Boomerang, allowing users
+This allows users
 to write synthesis tasks 
@@
 \paragraph*{Benchmark Suite Construction}
-Our benchmarks are drawn from three different sources.
+Our benchmarks for this experiment are drawn from three different sources.
 \begin{enumerate}
@@
   laws. None of these benchmarks were bijective.
-\item We adapted 29 benchmarks from Augeas~\cite{augeas}. Augeas is a utility
-  that bidirectionally converts between Linux configuration files and an
-  in-memory dictionaries. In our benchmarks, we synthesize lenses between Linux
-  configuration files and serialized versions of Augeas dictionaries. Because
-  these benchmarks merely transformed the information into a structured form,
+\item We adapted 29 benchmarks from Augeas~\cite{augeas}. Because
+  these benchmarks merely transform the information into a structured form,
   synthesized lenses were all bijective.
@@
 \Disconnect{}s\footnote{
-A \Disconnect{} is a lens combinator that marks a place where
+Recall that \Disconnect{} is a lens combinator that marks a place where
 information appears in one format but not the 
@@
 of 60 seconds, averaging the result over 5 runs. Figure~\ref{fig:times}
-summarizes the results of these tests. We find that our algorithm is able to
+summarizes the results of these tests. We found that our algorithm is able to
 synthesize all of the benchmarks in under 30 seconds. Without compositional
@@
 
-To compare to the existing bijective synthesis algorithm, we run our symmetric
+To compare to the existing bijective synthesis algorithm, we ran our symmetric
 synthesis algorithm on the original Optician benchmarks, comprised of
@@
 Without these alterations, symmetric synthesis yielded a lens that fit
-the specification but that was undesired.}
+the specification but that was not what the programmer wanted.}
 To perform this comparison, we synthesized lenses in two modes:
@@
 (necessitating \Skip). Without any annotations, our algorithm finds the correct
-lens for 40 of our 48 benchmarks; eight required relevance annotations to find
-the correct lens.
+lens for 40 of our 48 benchmarks; the other eight required relevance
+annotations to find the correct lens.
 % end evaluation

### 2019-03-04 report/final-report.tex
section 5

--- 2019-03-04 report/final-report.tex
@@
 examples.  As the class of transformations becomes richer, the number of
-potential programs grows dramatically.  As a result, the corresponding
+potential programs grows dramatically.  As a result, the na\"{i}ve
 inference algorithm slows and its ability to guess the transformation
 desired by the user decreases, leading to reduced accuracy.  We overcame
-these challenges through new heuristics that use information theory to help
+these challenges by applying heuristics that use information theory to help
 guide the search for program transformations.  In addition, to scale our
-prototype to larger transformation tasks, we found it necessary to work
+prototypes to larger transformation tasks, we found it necessary to work
 compositionally, by breaking down complex synthesis tasks into smaller, more
-mangeable subtasks.
+mangeable subtasks.  The resulting system, embodied by (variants of) our
+prototype \Optician{} implementation, out perform existing state-of-the-art
+tools for synthesis tasks in this domain.
 
@@
   \Optician{} & Our prototype implementation of the lens synthesis algorithms \\
+  \Optician{}$_Q$ & Optician enhanced with support for QRE synthesis \\
   QRE                   & Quotient regular expression\\

### 2019-03-04 report/final-report.tex
bib stuff

--- 2019-03-04 report/final-report.tex
@@
 describing and evaluating our algorithms, as well as open-source 
-code~\cite{GitHub} that implements those algorithms in (several variants of) a
+code~\cite{BijectiveOptician,QuotientOptician,SymmetricOptician} that implements those algorithms in (several variants of) a
 prototype tool called \Optician{}.  In this report,
@@
 describing and evaluating our algorithms and produced open-source
-code~\cite{GitHub} that implements those algorithms in (several variants of) a
+code~\cite{BijectiveOptician,QuotientOptician,SymmetricOptician} that implements those algorithms in (several variants of) a
 tool called \Optician{}.  In this report, we summarize some of our key methods,
@@
 We implemented our synthesis algorithms in the OCaml programming language
-and integrated them into Boomerang system, where they are available as open
-source code~\cite{GitHub}\bcp{Two more pointers needed}.  Users of Boomerang
-can write lenses by hand, specify them as synthesis tasks, or do a
-combination of both.
+where they are available as open
+source code~\cite{BijectiveOptician,QuotientOptician,SymmetricOptician}.
 

### 2019-03-04 report/final-report.tex
fig caption

--- 2019-03-04 report/final-report.tex
@@
   \caption{Average number of random examples required to synthesize benchmark
-    programs.  {\bf Experimental Average} is the average number of randomly
+    programs.  {\bf Examples Required} is the average number of randomly
     generated examples needed to correctly synthesize the lens.  {\bf
@@
 \subsection{Simple Symmetric Lens Synthesis}
-Finally, we implemented simple symmetric lenses as an extension to
+In addition to the results above, we implemented simple symmetric lenses as an extension to
 Boomerang~\cite{boomerang}.  

### 2019-03-04 report/final-report.tex
caption

@@
   \caption{Average number of random examples required to synthesize benchmark
-    programs.  {\bf Examples Required} is the average number of randomly
-    generated examples needed to correctly synthesize the lens.  {\bf
-      Determinize Permutations} is the theoretical number of examples required
-    to determinize the choice all the permutations in \RigidSynth{}.
-    In practice, far fewer examples are
-    needed to synthesize the correct lens than would be predicted by the number
-    required to determinize permutations.}
+    programs. %  {\bf Examples Required} is the average number of randomly
+    % generated examples needed to correctly synthesize the lens.
+    % {\bf
+    %   Determinize Permutations} is the theoretical number of examples required
+    % to determinize the choice all the permutations in \RigidSynth{}.
+    % In practice, far fewer examples are
+    % needed to synthesize the correct lens than would be predicted by the number
+    % required to determinize permutations.
+  }
   \label{fig:exs-reqd}

### 2019-04-16 report/final-report.tex

@@
 unexercised portions of the source and target regular expressions.
-Figure~\ref{fig:exs-reqd} lists
+We also calculated
 the number of examples that are required to determinize the generation of
@@
 
-The average number of examples is so much lower than the maximum
-number of required examples because of correspondences in how we wrote
-the regular expressions for the source and target data formats. 
-Specifically, when we had corresponding disjunctions in both the
-source and the target, we ordered them the same way.  The algorithm
-uses the supplied ordering to guide its search, and so the system
-requires fewer examples.   We did not write the examples in this style
-to facilitate synthesis, but rather because maintaining similar
-subparts in similar orderings makes the types much easier to 
-read. We expect that most users would do the same.
+The average number of examples needed is much lower than the maximum number of
+required examples because of correspondences in how we wrote the regular
+expressions for the source and target data formats.  Specifically, when we had
+corresponding disjunctions in both the source and the target, we ordered them
+the same way.  The algorithm uses the supplied ordering to guide its search, and
+so the system requires fewer examples.  We did not write the examples in this
+style to facilitate synthesis, but rather because maintaining similar subparts
+in similar orderings makes the types much easier to read. We expect that most
+users would do the same.
 
@@
 as proxies for the level of programmer effort when using each the three
-approaches: \bcp{This seems very detailed, and it includes some technical
-  terms (like ``canonize(q)'') that readers of this summary may not
-  know.  But maybe it's OK?}
+approaches:
 
@@
   The sum of (1) the number of AST nodes in $W(q)$ for each QRE $q$ in the source and target
-  formats, (2) the number of AST nodes in $\canonize(q)$ for each QRE $q$ with a
-  non-trivial canonizer, and (3) the number of AST nodes in the
+  formats, (2) the number of AST nodes in $\canonize(q)$ for each QRE $q$
+  requiring a  non-trivial canonizer (``canonizers'' as transformations that
+  project a string to its equivalence class representative; they are written as
+  programs in Boomerang), and (3) the number of AST nodes in the
   examples.  We use (1) to estimate the burden of describing

### 2019-04-16 papers/SymmetricLenses/rebuttal_icfp19.txt
rebuttal

Thank you for your feedback. We respond the major questions and concerns, then
provide an (optional) point-by-point response to individual concerns. We are
happy to include any of the information presented in this rebuttal in the final
version of the paper.

    - Why is writing two regexes easier for the user than writing a lens? Do
      input formats often come equipped with regexes already? Can the
      synthesizer use such pre-existing regexes out of the box, or do they have
      to be "massaged" by the user?

While we require users to write regular expressions, programming regular
expressions is easier than programming lenses:

* Regular expressions are widely understood.  Few programmers know about
  programming with lenses.

* A regular expression describes one data format, while a lens deals with two
  at the same time.

* In the (common) case where the two formats have Kleene stars in different
  places, manual lens programming requires mentally transforming both to a
  common "aligned" form.  Our synthesis algorithm does this alignment
  automatically.  This alignment can be quite challenging for programmers.

As an explicit example, consider the following function:

        let last_head = NAME . "," . (WSP . NAME)*
        let last_tail = (NAME . WSP)* NAME

        let name_conversion : last_head <=> last_tail =
            lens_swap
                (NAME . del ",")
                (lens_swap WSP NAME)*

While the name_conversion expression may have more AST nodes than the two
regular expressions, we find it much harder to write.

Engineering best practices dictate that we annotate the lenses with their
types.  While Boomerang does not require these annotations - being able to
infer the types from the terms - the types of the term serve as
documentation for future programmers to understand what formats the lens
maps between.  Furthermore, the types provide resilience in the face of future
lens modifications, as the types ensure that the lens maps exactly
between strings of the provided formats. Invalid inputs are not accepted
nor are the outputs ever invalid.  Consequently, we are not requiring
additional work. Instead, we are saving programmers from having to 
writing the lens by leveraging work they should already be performing.

In our experiments, we manually programmed the regular expressions. While
writing these regular expressions, we did not massage them to be amenable to
synthesis. In general, the synthesizer is able to handle regular expressions
expressed in multiple different ways, due to the traversal of regular expression
equivalences. However, if the regular expressions require a transformation that
was not a star-semiring transformation, such a massaging would be required.

    - What is the relationship between our information-theoretic ranking, and
      prior work like Unsupervised Learning by Program Synthesis

(afm: this can potentially go after the fold, as it was only mentioned in
passing, not as a critical question of the reviewer)

Unsupervised Learning by Program Synthesis (ULPS) expands upon previous research
in unsupervised machine learning. Unsupervised learning attempts to "induce good
latent representations of a data set" [Ellis et al. 2015]. Given a set of data
points, or "outputs," ULPS attempts to find good representations of these
outputs as a combination of finding a single function, and a sequence of
"inputs" that correspond to the outputs. The latent representation is a
combination of the corresponding input, and the transformation from inputs to
outputs.

Our work is similar to this work in that we both use information-theoretic
principles to find the desired programs, but the use of information theory in
each is different. In ULPS, minimizing the information theoretic cost metric is
the only goal. In this metric, the desire is to encode the output data in as few
bits as possible, so the example data is tightly coupled with the problem. In
our work, we use information theory to formalize the notion of "more bijective,"
providing a firm theoretical foundation for previous ad-hoc ranking functions.
In our case, the examples serve merely as a specification, but the inclusion or
exclusion of such examples does not impact "how bijective" certain functions
are. We do not consider this a weakness, but an inherent aspect to our
formalization. Furthermore, we would not consider it a good idea to assume that
we should try to minimize the data recovery costs on the examples more than on
other pieces of data -- non-bijective input/output pairs are often useful.

    - The authors do not consider lens compositions in their synthesis. Are
      there any justifications not to consider them? In existing results that
      synthesize bijective (quotient) lenses, one of the reasons not to consider
      lens composition in their synthesis is that ruling out them does not
      change the expressive power? Can you say a similar argument for this case?

We do not synthesize lens compositions for the same reasons as previous work --
such compositions are quite difficult to generate. You are correct that we do
not prove that such a restriction does not limit the expressive power. However,
we do not consider the symmetric DNF lens language a core contribution
(symmetric DNF lenses are not even defined outside of the appendix) and do not
believe such a proof would add much to the paper. However, we do provide
experimental evidence that the composition operator is not needed in practice;
we are able to synthesize all our benchmarks.

    - There is no explanation of the selection of the benchmarks. What are their
      characteristics and why those exercise interesting aspects of the
      algorithm.

We describe the selection of the benchmark suite in section 6.1. 29 of the
benchmarks are taken from Augeas, a bidirectional tool for editing Linux
configuration files. These examples were quite large, and involved some very
complex transformations, though these benchmarks were typically bijective. These
demonstrated how our tool performed at large, complex transformations. The 8
benchmarks from Flash Fill provided a range of transformations, none of which
were not bijective. These demonstrated how our tool performed on tasks that lost
large amounts of data, and sometimes required extended transformations through
Expand. The remaining 11 benchmarks tested how our tool performed at the type of
transformations we are familiar with, from transformations that would be useful
in day-to-day software engineering or classic transformations from the
bidirectional programming literature.

    - There is no attempt to explain what are the sources of worse performance
      for SS.

The primary impact on speed is the use of GreedySynth, instead of the bijective
RigidSynth. For GreedySynth, lenses are found between regular expressions that,
ultimately, do not show up in the final lens. RigidSynth does not incur this
performance penalty -- it only synthesizes lenses between regular expressions
for which there exists a bijective lens between them.

=================================================================================
Reviewer A
=================================================================================

    - Moreover, on top of the regexes, the user needs to provide examples and
      relevance annotations, and the only way to do so seems to be through an
      iterative process (by inspecting the synthesized lens and adding more
      examples / annotations when needed). Such iterative approach would be
      challenging for a user who is not sufficiently familiar with lenses to
      write one themselves.

We described the iterative process assuming a lazy user who didn't want to
provide more specifications than were necessary. However, such an iterative
process is not required, and if the user is confident that certain information
is "required" or may safely "skipped," they can safely add this information --
it will not slow down synthesis. Similarly, if the user already has a benchmark
suite, they may safely include such examples.

    - To add to the previous point: the evaluation could do a better job
      quantifying these limitations. For example, it would be good to see for
      each benchmark the size of annotations required, as well as how many
      "existing lenses" were used in compositional synthesis and where they came
      from (some standard library of lenses? or did the authors manually
      decompose each problem into subproblems? if so, then how many subproblems
      were there?)

There were a total of 12 uses of "require" in the benchmark suite. There were a
total of 4 uses of "skip" in the benchmark suite. In total 73 existing lenses
were used in compositional synthesis (about 1.5 per benchmark on average).
Typically, we manually decomposed these lenses, though some used the to_upper
function. We are happy to include a figure detailing the subtasks specified
during compositional synthesis in the final version of the paper (similar to the
one on p24 of Synthesizing Bijective Lenses).

    - Finally, the evaluation could also do a better job confirming that there's
      indeed a benefit to synthesizing a lens, rather than using a tool like
      FlashFill to synthesize the two functions independently. The related work
      section mentions that FlashFill cannot handle nested loops. However,
      FlashFill has evolved a lot since its original publication in 2011, so it
      would be interesting to see an empirical comparison.

We included an extended comparison to FlashFill in previous work, and did not
include such a comparison for lack of space. We are happy to make comparison to
FlashFill cited more explicitly.

    - The introduction mentions that stochastic regular expressions guide the
      search process. I would like to see it better explained in the main text
      of the paper: do they actually guide the search or merely rank the
      results?

Due to the Greedy nature of GreedySynth, the stochastic regular expressions
guide the search (but the stochastic annotations only guide the search insofar
as the lens costs guide the search).

    - L 284: please explain why this rule is needed for the lens in Fig. 3 to be
      well-typed.

Without the type equivalence rule, the lens would be well-typed between
emp_salaries and header . ("" | ("\n" . emp_ins) . ("\n" . emp_ins)∗). To be
well-typed between emp_salaries and emp_insurance, the right-hand type needs to
be transformed into an equivalent form.

    - Section 5.6: expansion inference seems to be a crucial part of the
      algorithm but is described very briefly. I would like to see more detail.

Expansion inference is the same as is performed in bijective lens inference, and
we did not include it for space. We will gladly make such a citation more
explicit.

    - Can you explain why you need the lower bound [for intervals]?

While we do not need it in the algorithm, we include it to make formal claims
about bounding the cost. We describe the need for intervals in 4.1.

### 2019-04-17 papers/SymmetricLenses/rebuttal_icfp19.txt
add rebuttal fine-grained responses

@@
 While we do not need it in the algorithm, we include it to make formal claims
-about bounding the cost. We describe the need for intervals in 4.1.
+about bounding the cost. We describe why we need intervals for the formal claims
+in 4.1.
+
+=================================================================================
+Reviewer B
+=================================================================================
+
+    - Theorem 1 is about the soundness of the equational theory shown in Fig. 4.
+      Can you say the completeness?
+
+The equational theory shown in Fig. 4. is incomplete, for it does not include
+some of the non-star-semiring equivalences (for example: the regular expression
+equivalence of (A | B)* = (A*B)*A* does not currently have an analogue in this
+equational theory). Furthermore, it is currently unclear whether the natural
+extensions to Conway's equational theory is sufficient for completeness, or
+whether Kozen and Solomaa's deductive systems are easily amenable to SREs.
+
+    - p10:L466 for other reasons
+
+The other reasons are to guarantee that any lens that is synthesized by our
+synthesis algorithm is guaranteed to be well-typed.
+
+    - p10:L489 relaxing this restrict is not problemantic
+
+For all these combinators, we can merely straightforwardly follow the style of
+unambiguity requirements on the types that was used in prior lens work. For
+example, the original Boomerang paper explicitly includes such unambiguity
+checks within the typing judgments, these unambiguity checks can be used in the
+same way for the symmetric lens variants of these combinators.
+
+    - p12:L563-572 I would like to read more explanations on why
+      \mathbb{H}^{\to} is defined as this.
+
+In general, these derivations follow the information theory, so as to make
+Theorem 3 true. Most of these strictly follow the rules of information theory,
+the only ones that are more complex are merges. Calculating the expected
+information to recover a string from a lens using merges does not involve a
+simple inductive calculation, but requires reasoning about the probilities of
+individual strings in the unioned format, conditioned on the probabilities of
+the source format.
+
+    - p13:L559 Cost(\ell) < Cost(best). What is the meaning of "<"? Since
+      Cost(\ell) can return a range instead of a number, the definition of "<"
+      is not so obvious.
+
+We define the cost as the sum of the maximum of \mathbb{H}^{\to} and the maximum
+of \mathbb{H}^{\leftarrow}.
+
+    - p14:L650 distance d.  What is "distance"?
+
+The distance is the number of star-semiring equivalences that must be applied to
+the pair of regular expressions to convert to the pair.
+
+    - p17:L806 CannotMap determines whether there is no lens satisfying the
+      examples.
+
+This is done by looking at the example parses. For example, given two regular
+expressions "a" | "b" and "c", and the example data putl "c" "b" = "a", we can
+determine that there cannot be a lens between "c" and "b".
+
+    - p17:L806 CannotMap determines whether there is no lens satisfying the
+      examples. How do you determine this?
+
+This is done by looking at the example parses. For example, given two regular
+expressions "a" | "b" and "c", and the example data putl "c" "b" = "a", we can
+determine that there cannot be a lens between "c" and "b".
+
+=================================================================================
+Reviewer C
+=================================================================================
+
+    - 5 runs per benchmark are not sufficient to draw statistically significant
+      conclusions. There is no report of errors or confidence intervals for the
+      mean time calculations.
+
+We only ran 5 times because there was so little variance between the times, as
+our procedure is deterministic. Rerunning, we found our average standard
+deviation while running the full symmetric algorithm was [STDDEV]. We are happy
+to both include standard deviations and rerun additional times.
+
+    - The graphs drop a lot of useful information. In particular they do not
+      show for which benchmark SS does better than SSNC or BS. This information
+      together with an analysis of the benchmarks can help understand the
+      performance of the proposed algorithm. The number of benchmarks that
+      terminate in less than 30s offers almost no insight.
+
+In general, SS always performs better than SSNC and SS always performs worse
+than BS. We ran for 30s under the assumption that users would be unwilling to
+wait for responses much longer than that. We are happy to rerun with a longer
+timeout.
+
+    - Similar to the above, the graph in figure 8 and the accompanying
+      discussion are not sufficient to understand what features of the algorithm
+      help or do not help with what kind of inputs.
+
+Requires help with continuing searching when the desired lenses require a large
+number of Expand applications. Skips help when the desired lenses require
+projecting large amounts of information. Using the DC metric only works for
+lenses where the desired lens appears quite early. FL is the same as DC, but is
+even more dependent on the desired lens appearing early (it *must* appear in the
+first cluster containing a satisfying program).

### 2019-04-17 papers/SymmetricLenses/rebuttal_icfp19.txt
Some notes

@@
-Thank you for your feedback. We respond the major questions and concerns, then
-provide an (optional) point-by-point response to individual concerns. We are
-happy to include any of the information presented in this rebuttal in the final
-version of the paper.
+Thank you for the thoughtful feedback. We first respond briefly to major
+questions and concerns, then provide an (optional, "below the fold")
+point-by-point response to individual concerns. We are happy to include any
+of the information presented in this rebuttal in the final version of the
+paper.
 
@@
 
-While we require users to write regular expressions, programming regular
-expressions is easier than programming lenses:
+BCP: This seems too long for "above the fold"
 
-* Regular expressions are widely understood.  Few programmers know about
-  programming with lenses.
+Writing regular expressions is significantly easier than programming lenses:
 
-* A regular expression describes one data format, while a lens deals with two
-  at the same time.
+* Regular expressions are widely understood, while few programmers know
+  about programming with lenses.
 
-* In the (common) case where the two formats have Kleene stars in different
-  places, manual lens programming requires mentally transforming both to a
-  common "aligned" form.  Our synthesis algorithm does this alignment
-  automatically.  This alignment can be quite challenging for programmers.
+* Each of the two regular expressions describes just one data format, while
+  a lens has to deal with both at the same time.
 
-As an explicit example, consider the following function:
+* In the (common, in our experience) case where the natural way of writing
+  the two formats puts Kleene stars in different places, manual lens
+  programming requires mentally transforming both to a common "aligned"
+  form, which can be quite challenging for programmers.  Our synthesis
+  algorithm does this alignment automatically.
+
+BCP: Maybe the example is too detailed -- move it to "under the fold"?
+As an explicit example of the last point, consider the following function:
 
@@
 regular expressions, we find it much harder to write.
+BCP: Is there a "not" missing?
 
+BCP: The point in this paragraph seems weaker than the others.  We can still
+  make it, but we should acknowledge that it is less important.
 Engineering best practices dictate that we annotate the lenses with their
@@
 was not a star-semiring transformation, such a massaging would be required.
+BCP: What would be an example of the last point?
 
@@
 passing, not as a critical question of the reviewer)
+BCP: Maybe a short version here?
 
@@
 we are able to synthesize all our benchmarks.
+
+BCP: I had not actually realized that we didn't prove that composition is
+admissible!  This is a significant technical shortcoming, no?  Is the proof
+hard?  Or is it maybe noit true?
 

### 2019-04-17 papers/SymmetricLenses/rebuttal_icfp19.txt
include stddev
revisions

@@
       to be "massaged" by the user?
-
-BCP: This seems too long for "above the fold"
 
@@
 
-BCP: Maybe the example is too detailed -- move it to "under the fold"?
-As an explicit example of the last point, consider the following function:
-
-        let last_head = NAME . "," . (WSP . NAME)*
-        let last_tail = (NAME . WSP)* NAME
-
-        let name_conversion : last_head <=> last_tail =
-            lens_swap
-                (NAME . del ",")
-                (lens_swap WSP NAME)*
-
-While the name_conversion expression may have more AST nodes than the two
-regular expressions, we find it much harder to write.
-BCP: Is there a "not" missing?
-
-BCP: The point in this paragraph seems weaker than the others.  We can still
-  make it, but we should acknowledge that it is less important.
-Engineering best practices dictate that we annotate the lenses with their
-types.  While Boomerang does not require these annotations - being able to
-infer the types from the terms - the types of the term serve as
-documentation for future programmers to understand what formats the lens
-maps between.  Furthermore, the types provide resilience in the face of future
-lens modifications, as the types ensure that the lens maps exactly
-between strings of the provided formats. Invalid inputs are not accepted
-nor are the outputs ever invalid.  Consequently, we are not requiring
-additional work. Instead, we are saving programmers from having to 
-writing the lens by leveraging work they should already be performing.
-
 In our experiments, we manually programmed the regular expressions. While
@@
 was not a star-semiring transformation, such a massaging would be required.
+
 BCP: What would be an example of the last point?
+AFM: Consider creating a lens of the following type: (a+b)* <=> (a*b)*a*.
+Finding the identity lens between these two regular expressions would require
+using sumstar, a regular expression equivalence we do not traverse.
 
@@
 BCP: Maybe a short version here?
+AFM: I couldn't find a natural way to split it into main component/subcomponent
+so I just made it more terse
 
 Unsupervised Learning by Program Synthesis (ULPS) expands upon previous research
-in unsupervised machine learning. Unsupervised learning attempts to "induce good
-latent representations of a data set" [Ellis et al. 2015]. Given a set of data
-points, or "outputs," ULPS attempts to find good representations of these
-outputs as a combination of finding a single function, and a sequence of
-"inputs" that correspond to the outputs. The latent representation is a
-combination of the corresponding input, and the transformation from inputs to
-outputs.
+in unsupervised machine learning. Given a set of data points, or "outputs," ULPS
+attempts to find good representations of these outputs as a combination of
+finding a single function, and a sequence of "inputs" that correspond to the
+outputs. The latent representation of the example data is a combination of the
+corresponding input, and the transformation from inputs to outputs.
 
-Our work is similar to this work in that we both use information-theoretic
-principles to find the desired programs, but the use of information theory in
-each is different. In ULPS, minimizing the information theoretic cost metric is
-the only goal. In this metric, the desire is to encode the output data in as few
-bits as possible, so the example data is tightly coupled with the problem. In
+Our work is similar to ULPS in that we both use information-theoretic principles
+to find the desired programs, but the use of information theory in each is quite
+different. In ULPS, the synthesizer aims to encode the example data in as few
+bits as possible, so the example data is tightly coupled with the problem (this
+is very similar to the approach in "From Dirt To Shovels", Fisher et al.). In
 our work, we use information theory to formalize the notion of "more bijective,"
@@
 exclusion of such examples does not impact "how bijective" certain functions
-are. We do not consider this a weakness, but an inherent aspect to our
-formalization. Furthermore, we would not consider it a good idea to assume that
-we should try to minimize the data recovery costs on the examples more than on
-other pieces of data -- non-bijective input/output pairs are often useful.
+are. This is not a weakness, merely a consequence of solving a different
+problem. We shouldn't prioritize minimizing the data recovery costs on the
+examples more than on other pieces of data -- non-bijective input/output pairs
+are often useful.
 
@@
 
+The omission of this proof is pragmatic -- the proof would be similar to that
+required for bijective lenses, which was nearly 100 pages. Due to the
+enhancements to DNF Lenses, Symmetric DNF lenses would need to be reproven
+bisimilar and confluent with respect to the regular expression rewrite rules.
+
 BCP: I had not actually realized that we didn't prove that composition is
@@
 hard?  Or is it maybe noit true?
+AFM: The proof is quite hard, comparable to the near 100 page one done in
+bijective lenses, not simpler like the QRE one. I don't personally consider
+it a major shortcoming -- as I don't think it really is part of our core
+contributions. I included a brief description of why it is hard.
 
@@
 =================================================================================
+
+    - Why is writing two regexes easier for the user than writing a lens? Do
+      input formats often come equipped with regexes already? Can the
+      synthesizer use such pre-existing regexes out of the box, or do they have
+      to be "massaged" by the user? (Extended Response)
+
+Below is an explicit example of of the difficulty of writing lenses vs writing
+regular expressions.
+
+        let last_head = NAME . "," . (WSP . NAME)*
+        let last_tail = (NAME . WSP)* NAME
+
+        let name_conversion : last_head <=> last_tail =
+            lens_swap
+                (NAME . del ",")
+                (lens_swap WSP NAME)*
+
+While the name_conversion expression has about as many AST nodes than the two
+regular expressions, we find it much harder to write.
+
+BCP: The point in this paragraph seems weaker than the others.  We can still
+  make it, but we should acknowledge that it is less important.
+AFM: I moved it down from the main response, is that sufficient?
+
+Engineering best practices dictate that we annotate the lenses with their
+types.  While Boomerang does not require these annotations - being able to
+infer the types from the terms - the types of the term serve as
+documentation for future programmers to understand what formats the lens
+maps between.  Furthermore, the types provide resilience in the face of future
+lens modifications, as the types ensure that the lens maps exactly
+between strings of the provided formats. Invalid inputs are not accepted
+nor are the outputs ever invalid.  Consequently, we are not requiring
+additional work. Instead, we are saving programmers from having to 
+writing the lens by leveraging work they should already be performing.
 
@@
 FlashFill cited more explicitly.
+
+    - What is the relationship between our information-theoretic ranking, and
+      prior work like Unsupervised Learning by Program Synthesis. (Extended
+      Response)
 
@@
 our procedure is deterministic. Rerunning, we found our average standard
-deviation while running the full symmetric algorithm was [STDDEV]. We are happy
+deviation while running the full symmetric algorithm was .0162. We are happy
 to both include standard deviations and rerun additional times.

### 2019-04-17 report/final-report.tex
edits

@@
 \documentclass[12pt]{article}
+%\usepackage[style=ieee]{biblatex}
+%\addbibresource{local.bib,bcp.bib}
+%\documentclass{IEEEtran}
+\usepackage{etoolbox}
+\patchcmd{\thebibliography}{\section*{\refname}}{}{}{}
+
+\providetoggle{blx@skipbiblist}
 
@@
 
+\newcommand{\FS}{\vspace{10pt}}
+
 \newtheorem*{theorem*}{Theorem}
@@
   Kathleen Fisher (Tufts University) \\
+  Solomon Maina (University of Pennsylvania) \\
+  Anders Miltner (Princeton University) \\
   Benjamin Pierce  (University of Pennsylvania) \\
@@
 
+\addcontentsline{toc}{section}{{Contents}}
 \tableofcontents
@@
 
+%\cleardoublepage
+% \phantomsection
+\addcontentsline{toc}{section}{{List of Figures}}
 \listoffigures
+
 \newpage
@@
 
+\newpage
 \section{Introduction}
@@
 in their operational environment---the vision of
-DARPA's BRASS program---it is necessary to develop new, easier-to-use
+the Defense Advanced Research Project Agency's (DARPA) Building
+Resource Adaptive Software Systems (BRASS) program---it is necessary to develop new, easier-to-use
 and more robust programming systems for managing these ad hoc data
@@
 a record or insert new kinds of syntactic separators (\emph{e.g.,} replacing a
-comma with a vertical bar, or the name of one HTML tag with another)
+comma with a vertical bar, or the name of one Hypertext Markup
+Language (HTML) tag with another)
 but they cannot implement more general transformations that elide
@@
 
+\newpage
 \section{Methods, Assumptions and Procedures}
@@
 
-Lens-based languages are present in variety of tools and have found mainstream
+Lens-based languages are present in a variety of tools and have found mainstream
 industrial use.
@@
 and
-XML transformation languages~\cite{DBLP:conf/pepm/LiuHT07}.
-
-If $\BRegex$ and $\BRegexAlt$ are regular expressions, then
+Extensible Markup Language (XML) transformation languages~\cite{DBLP:conf/pepm/LiuHT07}.
+
+If $\BRegex$ and $\BRegexAlt$ are regular expressions then
 \lstinline{$\Lens$ : $\BRegex \Leftrightarrow \BRegexAlt$} indicates that
@@
 %
-transforms a list of employees in employees in the Management format to a list
+transforms a list of employees in the Management format to a list
 of employees in the HR format and vice versa. These combinators
-are combined in figure~\ref{fig:example_lens} to construct a complete lens
+are combined in Figure~\ref{fig:example_lens} to construct a complete lens
 between the employee formats.
@@
   \label{fig:example_lens}
+  \FS
 \end{figure}
@@
   \label{fig:schematic-diagram-synthesis}
+  \FS
 \end{figure}
@@
 
+\newpage
 \section{Results and Discussion}
@@
 Furthermore, the 12 last synthesis problems derived
-from Augeas were tested after our synthesizer was was
+from Augeas were tested after our synthesizer was
 finalized, demonstrating that the optimizations were not
@@
 An average benchmark has a type-based specification that can be represented using
-310 AST node, and requires 1.1 input/output examples.
+310 Abstract Syntax Tree (AST) nodes, and requires 1.1 input/output examples.
 Our benchmarks vary from simple problems, like changing the
@@
 We also experimented using our tool to synthesize quotient lenses
-between XML, RDF and and JSON formats using data from the data.gov
+between XML, Resource Description Framework (RDF) and
+Javascript Object Notation (JSON) formats using data from the data.gov
 database; the data consisted of census statistics, demographic
-statistics, wage comparosion data, and crime index data).
+statistics, wage comparison data, and crime index data).
 Recall that, to use bijective synthesis, we had to modify some of the 
@@
 target formats and she may give examples.
-In the second approach, which we call \canonizeAndSpecSize{} for
-Bijective Synthesis, the
-programmer uses bijective lens synthesis \`a la Optician.
+In the second approach, ``Bijective Synthesis'' (\canonizeAndSpecSize{}), the
+programmer uses the Optician bijective synthesis algorithm.
 She must write canonizers by hand, along with 
@@
 provide examples to help in the synthesis of the bijective lens.
-In the third approach, which we call \LensAndSpecSize{} for No Synthesis, the
+In the third approach, ``No Synthesis'' (\LensAndSpecSize{}), the
 programmer writes the lens between the source and target formats
@@
 For each problem in the benchmark suite, we calculate the following measures
-as proxies for the level of programmer effort when using each the three
+as proxies for the level of programmer effort when using each of the three
 approaches:
@@
   programs in Boomerang), and (3) the number of AST nodes in the
-  examples.  We use (1) to estimate the burden of describing
+  examples.  We use (1) to estimate the cost of describing
   the external source and target formats and (2) to estimate the
-  burden of writing the requisite canonizers
+  cost of writing the requisite canonizers
   by hand.  We count the nodes in the examples because they would be
@@
 
-\subsection{Simple Symmetric Lens Synthesis}
+\subsection{Simple Symmetric Synthesis Algorithms}
 In addition to the results above, we implemented simple symmetric lenses as an extension to
@@
     one and three sufficient examples.}
-  \label{fig:times}
+  \label{fig:symm-times}
 \end{figure}
@@
 For each benchmark in the suite and each mode, we ran the system with a timeout
-of 60 seconds, averaging the result over 5 runs. Figure~\ref{fig:times}
+of 60 seconds, averaging the result over 5 runs. Figure~\ref{fig:symm-times}
 summarizes the results of these tests. We found that our algorithm is able to
@@
 Without these alterations, symmetric synthesis yielded a lens that fit
-the specification but that was not what the programmer wanted.}
+the specification but was not bijective and hence was different from
+the one generated by Optician.}
 To perform this comparison, we synthesized lenses in two modes:
@@
 
+\newpage
 \section{Conclusions}
@@
 
+\newpage
+
+\section{References}
+
+%\printbibliography[heading=none]
 \bibliographystyle{apalike}
+%\bibliographystyle{IEEEtran}
 \bibliography{local.bib,bcp.bib}
 
+\newpage
 \section{List of Symbols, Abbreviations, and Acronyms}
@@
   AST     & Abstract syntax tree\\
+  BRASS & Building Resource Adaptive Software Systems \\
   \BSOpt{}              & Bijective lens synthesis algorithm\\
@@
   HR & Human Resources \\
-  HTML & Hyptertext Markup Language \\
+  HTML & Hypertext Markup Language \\
   JSON & JavaScript Object Notation---a standard file format used to
@@
   \textbf{QS}  & Number of nodes in the abstract syntax of QRE specifications \\
-  RDF & Resource Description Framework---used to model web resources \\
+  RDF & Resource Description Framework---used to model web resources
+  \\
+  RE & Regular Expression \\
   REST & Representational State Transfer\\

### 2019-04-17 papers/SymmetricLenses/rebuttal_icfp19.txt

@@
 
-    - What is the relationship between our information-theoretic ranking, and
-      prior work like Unsupervised Learning by Program Synthesis. (Extended
-      Response)
-
     - The introduction mentions that stochastic regular expressions guide the
@@
 Due to the Greedy nature of GreedySynth, the stochastic regular expressions
-guide the search (but the stochastic annotations only guide the search insofar
-as the lens costs guide the search).
+guide the search (insofar as the lens costs guide the search).
 
@@
 well-typed between emp_salaries and emp_insurance, the right-hand type needs to
-be transformed into an equivalent form.
+be transformed into an equivalent form, without the explicit star unrolling.
 

### 2019-04-18 papers/SymmetricLenses/rebuttal_icfp19.txt
Update rebuttal_icfp19.txt

@@
 
+KSF: In the above, we need to explain how we picked the examples in a way that 
+didn't cherry pick ones that would work for us.  The above text doesn't do that.
+
     - There is no attempt to explain what are the sources of worse performance
@@
 for which there exists a bijective lens between them.
+
+KSF: I'm not following the above argument, which makes me think the reviewers won't either. 
+Can you revise to make the text more self-contained?
 

### 2019-04-18 papers/SymmetricLenses/rebuttal_icfp19.txt
edits

--- 2019-04-17 papers/SymmetricLenses/rebuttal_icfp19.txt
@@
 writing these regular expressions, we did not massage them to be amenable to
-synthesis. In general, the synthesizer is able to handle regular expressions
-expressed in multiple different ways, due to the traversal of regular expression
-equivalences. However, if the regular expressions require a transformation that
-was not a star-semiring transformation, such a massaging would be required.
+synthesis. The synthesizer implements the star-semiring regular
+expression equivalence axioms so that users do not have to worry about
+the way they write expressions themselves.
 
@@
 
-Unsupervised Learning by Program Synthesis (ULPS) expands upon previous research
-in unsupervised machine learning. Given a set of data points, or "outputs," ULPS
-attempts to find good representations of these outputs as a combination of
-finding a single function, and a sequence of "inputs" that correspond to the
-outputs. The latent representation of the example data is a combination of the
-corresponding input, and the transformation from inputs to outputs.
+In a nutshell, Unsupervised Learning by Program Synthesis (ULPS)
+learns a description of a single data set.  Ideally, such an
+descriptions are compact and information theory is used as a measure
+of the compactness of the description learned.  In this sense, ULPS
+shares more similarities with Fisher et al's paper on "From Dirt To
+Shovels" (which also seeks to learn a description of a single data
+source) than it does with the current paper.
 
-Our work is similar to ULPS in that we both use information-theoretic principles
-to find the desired programs, but the use of information theory in each is quite
-different. In ULPS, the synthesizer aims to encode the example data in as few
-bits as possible, so the example data is tightly coupled with the problem (this
-is very similar to the approach in "From Dirt To Shovels", Fisher et al.). In
-our work, we use information theory to formalize the notion of "more bijective,"
-providing a firm theoretical foundation for previous ad-hoc ranking functions.
-In our case, the examples serve merely as a specification, but the inclusion or
-exclusion of such examples does not impact "how bijective" certain functions
-are. This is not a weakness, merely a consequence of solving a different
-problem. We shouldn't prioritize minimizing the data recovery costs on the
-examples more than on other pieces of data -- non-bijective input/output pairs
-are often useful.
+In contrast, in our current paper, we are attempting to learn a
+translation from data set to another.  However, there are many different
+candidate translations.  To select amongst the candidate translations,
+we choose the translation that preserves as much information from
+source to target (and vice versa) as possible.
+
+In summary, both systems use information theory but the objectives of
+the two systems are different and the properties measured by
+information theory are also different.
+
+(Please also see below for an extended response.)
 
@@
 
-We do not synthesize lens compositions for the same reasons as previous work --
-such compositions are quite difficult to generate. You are correct that we do
-not prove that such a restriction does not limit the expressive power. However,
-we do not consider the symmetric DNF lens language a core contribution
-(symmetric DNF lenses are not even defined outside of the appendix) and do not
-believe such a proof would add much to the paper. However, we do provide
-experimental evidence that the composition operator is not needed in practice;
-we are able to synthesize all our benchmarks.
-
-The omission of this proof is pragmatic -- the proof would be similar to that
-required for bijective lenses, which was nearly 100 pages. Due to the
-enhancements to DNF Lenses, Symmetric DNF lenses would need to be reproven
-bisimilar and confluent with respect to the regular expression rewrite rules.
-
-BCP: I had not actually realized that we didn't prove that composition is
-admissible!  This is a significant technical shortcoming, no?  Is the proof
-hard?  Or is it maybe noit true?
-AFM: The proof is quite hard, comparable to the near 100 page one done in
-bijective lenses, not simpler like the QRE one. I don't personally consider
-it a major shortcoming -- as I don't think it really is part of our core
-contributions. I included a brief description of why it is hard.
+We do not synthesize lens compositions for the same reasons as
+previous work -- such compositions are quite difficult to generate
+(doing so requires synthesizing an intermediate regular expression
+from out of thin air).  We conjecture that adding composition lenses
+would not increase the expressive power of the language but we have
+not proven this fact.  The analogous proof for bijective lenses
+required nearly 100 pages of work and we did not believe the result
+was important enough to invest this amount of work.  We chose to focus
+our energy on understanding the information theoretic ranking methods,
+and the relationship between simple symmetric lenses and classical
+symmetric lenses, which we view as the core contributions of this paper.
 
@@
 
-We describe the selection of the benchmark suite in section 6.1. 29 of the
-benchmarks are taken from Augeas, a bidirectional tool for editing Linux
-configuration files. These examples were quite large, and involved some very
-complex transformations, though these benchmarks were typically bijective. These
-demonstrated how our tool performed at large, complex transformations. The 8
-benchmarks from Flash Fill provided a range of transformations, none of which
-were not bijective. These demonstrated how our tool performed on tasks that lost
-large amounts of data, and sometimes required extended transformations through
-Expand. The remaining 11 benchmarks tested how our tool performed at the type of
-transformations we are familiar with, from transformations that would be useful
-in day-to-day software engineering or classic transformations from the
-bidirectional programming literature.
+We describe the selection of the benchmark suite in section 6.1.
+
+29 of the benchmarks are taken from Augeas, a bidirectional tool for editing Linux
+configuration files.  We selected these 29 benchmarks by selecting
+Augeas files one-by-one in alphabetical order (we did not cherry-pick
+programs).  We stopped when we didn't feel that adding more benchmarks
+would be informative.
+
+These examples were quite large, and involved some very complex
+transformations, though these benchmarks were typically
+bijective. These benchmarks demonstrated how our tool performed on
+large, complex transformations. The 8 benchmarks from Flash Fill
+provided a range of transformations, none of which were not
+bijective. Again, the FlashFill benchmarks were the first 8 benchmarks
+in the paper.  These benchmarks demonstrated how our tool performed on
+transformation tasks that
+discard large amounts of data.  On occasion, these benchmarks require
+the use of many regular expression equivalence axioms.  The remaining
+11 benchmarks were chosen in an ad hoc manner --- we simply looked
+around for useful day-to-day example data for which it would be useful to have lenses.
 
@@
       for SS.
+
+Bijective synthesis is faster than symmetric synthesis because ...
 

### 2019-04-18 papers/SymmetricLenses/rebuttal_icfp19.txt
rebuttal

--- 2019-04-18 papers/SymmetricLenses/rebuttal_icfp19.txt
@@
 in the paper.  These benchmarks demonstrated how our tool performed on
-transformation tasks that
-discard large amounts of data.  On occasion, these benchmarks require
-the use of many regular expression equivalence axioms.  The remaining
-11 benchmarks were chosen in an ad hoc manner --- we simply looked
-around for useful day-to-day example data for which it would be useful to have lenses.
+transformation tasks that discard large amounts of data.  On occasion,
+these benchmarks require the use of many regular expression
+equivalence axioms.  The remaining 11 benchmarks were chosen in an ad
+hoc manner --- we simply looked around for useful day-to-day example
+data for which it would be useful to have lenses.
 
@@
 
-Bijective synthesis is faster than symmetric synthesis because ...
+Both the bijective synthesis and the symmetric synthesis use a pair of
+collaborating synthesizers that involves (1) a search for a compatible
+pair of regular expressions and (2) search for a lens given those
+regular expressions.
 
-The primary impact on speed is the use of GreedySynth, instead of the bijective
-RigidSynth. For GreedySynth, lenses are found between regular expressions that,
-ultimately, do not show up in the final lens. RigidSynth does not incur this
-performance penalty -- it only synthesizes lenses between regular expressions
-for which there exists a bijective lens between them.
+Bijective synthesis is faster than symmetric synthesis because part
+(2) is much faster.  More specifically, a bijection must translate
+*all* data on the left into data on the right.  However, a symmetric
+synthesis problem has a choice of *which* data on the left to translate
+into data on the right.  This gives rise to a large set of possible
+choices, and symmetric synthesis must consider all of them.
 

### 2019-04-19 papers/SymmetricLenses/rebuttal_icfp19.txt
Update rebuttal_icfp19.txt

@@
 In a nutshell, Unsupervised Learning by Program Synthesis (ULPS)
-learns a description of a single data set.  Ideally, such an
+learns a description of a single data set.  Ideally, such 
 descriptions are compact and information theory is used as a measure
@@
 large, complex transformations. The 8 benchmarks from Flash Fill
-provided a range of transformations, none of which were not
+provided a range of transformations, all of which were 
 bijective. Again, the FlashFill benchmarks were the first 8 benchmarks
@@
 
-For all these combinators, we can merely straightforwardly follow the style of
+For all these combinators, we can follow the style of
 unambiguity requirements on the types that was used in prior lens work. For
 example, the original Boomerang paper explicitly includes such unambiguity
-checks within the typing judgments, these unambiguity checks can be used in the
+checks within the typing judgments; these unambiguity checks can be used in the
 same way for the symmetric lens variants of these combinators.
@@
 In general, these derivations follow the information theory, so as to make
-Theorem 3 true. Most of these strictly follow the rules of information theory,
-the only ones that are more complex are merges. Calculating the expected
+Theorem 3 true. Most strictly follow the rules of information theory.
+The only ones that are more complex are merges. Calculating the expected
 information to recover a string from a lens using merges does not involve a
-simple inductive calculation, but requires reasoning about the probilities of
+simple inductive calculation, but requires reasoning about the probabilities of
 individual strings in the unioned format, conditioned on the probabilities of
@@
       examples.
-
-This is done by looking at the example parses. For example, given two regular
-expressions "a" | "b" and "c", and the example data putl "c" "b" = "a", we can
-determine that there cannot be a lens between "c" and "b".
-
-    - p17:L806 CannotMap determines whether there is no lens satisfying the
-      examples. How do you determine this?
 

### 2019-04-19 papers/SymmetricLenses/rebuttal_icfp19.txt
rebuttal

--- 2019-04-18 papers/SymmetricLenses/rebuttal_icfp19.txt
@@
 
-BCP: What would be an example of the last point?
-AFM: Consider creating a lens of the following type: (a+b)* <=> (a*b)*a*.
-Finding the identity lens between these two regular expressions would require
-using sumstar, a regular expression equivalence we do not traverse.
-
     - What is the relationship between our information-theoretic ranking, and
       prior work like Unsupervised Learning by Program Synthesis
-
-(afm: this can potentially go after the fold, as it was only mentioned in
-passing, not as a critical question of the reviewer)
-BCP: Maybe a short version here?
-AFM: I couldn't find a natural way to split it into main component/subcomponent
-so I just made it more terse
 
@@
 information theory are also different.
-
-(Please also see below for an extended response.)
 
@@
 
-29 of the benchmarks are taken from Augeas, a bidirectional tool for editing Linux
-configuration files.  We selected these 29 benchmarks by selecting
-Augeas files one-by-one in alphabetical order (we did not cherry-pick
-programs).  We stopped when we didn't feel that adding more benchmarks
-would be informative.
+29 of the benchmarks are taken from Augeas, a bidirectional tool for
+editing Linux configuration files.  We selected these benchmarks by
+selecting Augeas files one-by-one in alphabetical order (we did
+not cherry-pick programs).  We stopped when we didn't feel that
+adding more benchmarks would be informative. These examples were
+quite large, and involved some very complex transformations, though
+these benchmarks were typically bijective. These benchmarks
+demonstrated how our tool performed on large, complex transformations.
 
-These examples were quite large, and involved some very complex
-transformations, though these benchmarks were typically
-bijective. These benchmarks demonstrated how our tool performed on
-large, complex transformations. The 8 benchmarks from Flash Fill
-provided a range of transformations, none of which were not
-bijective. Again, the FlashFill benchmarks were the first 8 benchmarks
-in the paper.  These benchmarks demonstrated how our tool performed on
-transformation tasks that discard large amounts of data.  On occasion,
-these benchmarks require the use of many regular expression
-equivalence axioms.  The remaining 11 benchmarks were chosen in an ad
-hoc manner --- we simply looked around for useful day-to-day example
-data for which it would be useful to have lenses.
+The 8 benchmarks from Flash Fill provided a range of transformations,
+none of which were not bijective. Again, the FlashFill benchmarks were
+the first 8 benchmarks in the paper.  These benchmarks demonstrated
+how our tool performed on transformation tasks that discard large
+amounts of data.  On occasion, these benchmarks require the use of
+many regular expression equivalence axioms.
 
-    - There is no attempt to explain what are the sources of worse performance
-      for SS.
+The remaining 11 benchmarks were chosen in an ad hoc manner --- we
+simply looked around for day-to-day data formats for which it would
+be useful to have lenses.
 
-Both the bijective synthesis and the symmetric synthesis use a pair of
-collaborating synthesizers that involves (1) a search for a compatible
-pair of regular expressions and (2) search for a lens given those
-regular expressions.
+    - Why is symmetric synthesis slower than bijective synthesis.
+
+Both the bijective synthesis and the symmetric synthesis engines use a
+pair of collaborating synthesizers that involves (1) a search for a
+compatible pair of regular expressions and (2) search for a lens given
+those regular expressions.
 
@@
 regular expressions, we find it much harder to write.
-
-BCP: The point in this paragraph seems weaker than the others.  We can still
-  make it, but we should acknowledge that it is less important.
-AFM: I moved it down from the main response, is that sufficient?
 
@@
 
-We included an extended comparison to FlashFill in previous work, and did not
-include such a comparison for lack of space. We are happy to make comparison to
-FlashFill cited more explicitly.
+We included an extended comparison to a current version of FlashFill in previous
+work, and did not include such a comparison in this work for lack of space.
 
@@
 equational theory). Furthermore, it is currently unclear whether the natural
-extensions to Conway's equational theory is sufficient for completeness, or
+extensions to Conway's equational theory is sufficient for completeness, and
 whether Kozen and Solomaa's deductive systems are easily amenable to SREs.
@@
 
-In general, these derivations follow the information theory, so as to make
-Theorem 3 true. Most of these strictly follow the rules of information theory,
-the only ones that are more complex are merges. Calculating the expected
-information to recover a string from a lens using merges does not involve a
-simple inductive calculation, but requires reasoning about the probilities of
-individual strings in the unioned format, conditioned on the probabilities of
-the source format.
+In general, these derivations are defined so as to make Theorem 3 true. Most of
+these simply follow the decompositions provided by information theory, the
+only ones that are more complex are merges. Calculating the expected information
+to recover a string from a lens using merges does not involve a simple inductive
+calculation, but requires reasoning about the probilities of individual strings
+in the unioned format, conditioned on the probabilities of the source format.
 

### 2019-04-19 papers/SymmetricLenses/rebuttal_icfp19.txt

@@
 quite large, and involved some very complex transformations, though
-these benchmarks were typically bijective. These benchmarks
-demonstrated how our tool performed on large, complex transformations.
+these benchmarks were bijective. These benchmarks demonstrated 
+how our tool performed on large, complex transformations.
 
 The 8 benchmarks from Flash Fill provided a range of transformations,
-none of which were not bijective. Again, the FlashFill benchmarks were
+none of which were bijective. Again, the FlashFill benchmarks were
 the first 8 benchmarks in the paper.  These benchmarks demonstrated
@@
 We included an extended comparison to a current version of FlashFill in previous
-work, and did not include such a comparison in this work for lack of space.
+work, and did not include such a comparison in this work for lack of space. We
+are happy to make comparison to FlashFill cited more explicitly.
 
@@
 
-In general, these derivations are defined so as to make Theorem 3 true. Most of
-these simply follow the decompositions provided by information theory, the
-only ones that are more complex are merges. Calculating the expected information
-to recover a string from a lens using merges does not involve a simple inductive
-calculation, but requires reasoning about the probilities of individual strings
-in the unioned format, conditioned on the probabilities of the source format.
+In general, these derivations follow the information theory, so as to make
+Theorem 3 true. Most strictly follow the rules of information theory.
+The only ones that are more complex are merges. Calculating the expected
+information to recover a string from a lens using merges does not involve a
+simple inductive calculation, but requires reasoning about the probabilities of
+individual strings in the unioned format, conditioned on the probabilities of
+the source format. Because of this complexity, we chose to bound the information
+content instead of calculate it exactly.
 

### 2019-04-19 papers/SymmetricLenses/rebuttal_icfp19.txt
add rebuttal yay
Some polishing -- more later

@@
 questions and concerns, then provide an (optional, "below the fold")
-point-by-point response to individual concerns. We are happy to include any
-of the information presented in this rebuttal in the final version of the
-paper.
+point-by-point response to individual concerns.
 
@@
       input formats often come equipped with regexes already? Can the
-      synthesizer use such pre-existing regexes out of the box, or do they have
-      to be "massaged" by the user?
+      synthesizer use such pre-existing regexes out of the box, or do they
+      have to be "massaged" by the user?
 
@@
 
-* Regular expressions are widely understood, while few programmers know
-  about programming with lenses.
+* Regular expressions are widely understood, whereas few programmers are
+  fluent at building lenses.
 
-* Each of the two regular expressions describes just one data format, while
-  a lens has to deal with both at the same time.
+* Each of the two regular expressions describes a single data format, while
+  a lens has to deal with two formats at the same time.
 
@@
 
-In our experiments, we manually programmed the regular expressions. While
-writing these regular expressions, we did not massage them to be amenable to
-synthesis. The synthesizer implements the star-semiring regular
-expression equivalence axioms so that users do not have to worry about
-the way they write expressions themselves.
+In our experiments, we manually programmed the regular expressions. However,
+when writing these regular expressions, we did not massage them to be
+amenable to synthesis. The synthesizer implements the star-semiring regular
+expression equivalence axioms, so that users do not have to worry about the
+way they write expressions themselves.
 
-    - What is the relationship between our information-theoretic ranking, and
-      prior work like Unsupervised Learning by Program Synthesis
+    - What is the relationship between our information-theoretic ranking,
+      and prior work like Unsupervised Learning by Program Synthesis
 
@@
     - The authors do not consider lens compositions in their synthesis. Are
-      there any justifications not to consider them? In existing results that
-      synthesize bijective (quotient) lenses, one of the reasons not to consider
-      lens composition in their synthesis is that ruling out them does not
-      change the expressive power? Can you say a similar argument for this case?
+      there any justifications not to consider them? In existing results
+      that synthesize bijective (quotient) lenses, one of the reasons not to
+      consider lens composition in their synthesis is that ruling out them
+      does not change the expressive power? Can you say a similar argument
+      for this case?
 
@@
 
-    - There is no explanation of the selection of the benchmarks. What are their
-      characteristics and why those exercise interesting aspects of the
-      algorithm.
+    - There is no explanation of the selection of the benchmarks. What are
+      their characteristics and why those exercise interesting aspects of
+      the algorithm.
 
@@
 (2) is much faster.  More specifically, a bijection must translate
-*all* data on the left into data on the right.  However, a symmetric
+*all* data on the left into data on the right.  By contrast, a symmetric
 synthesis problem has a choice of *which* data on the left to translate
@@
 work, and did not include such a comparison in this work for lack of space. We
-are happy to make comparison to FlashFill cited more explicitly.
+are happy to make this comparison to FlashFill cited more explicitly.
 

### 2019-05-08 papers/SymmetricLenses/Conference/rebuttal_icfp19.txt
move conference paper to conference directory

--- 2019-04-19 papers/SymmetricLenses/rebuttal_icfp19.txt
@@
 point-by-point response to individual concerns.
+
+-----
 
@@
 
+-----
+
     - What is the relationship between our information-theoretic ranking,
@@
 
+-----
+
     - The authors do not consider lens compositions in their synthesis. Are
@@
 
-We do not synthesize lens compositions for the same reasons as
-previous work -- such compositions are quite difficult to generate
-(doing so requires synthesizing an intermediate regular expression
-from out of thin air).  We conjecture that adding composition lenses
-would not increase the expressive power of the language but we have
-not proven this fact.  The analogous proof for bijective lenses
-required nearly 100 pages of work and we did not believe the result
-was important enough to invest this amount of work.  We chose to focus
-our energy on understanding the information theoretic ranking methods,
-and the relationship between simple symmetric lenses and classical
-symmetric lenses, which we view as the core contributions of this paper.
+We do not synthesize lens compositions for the same reasons as in previous
+work -- such compositions are quite difficult to generate (doing so requires
+synthesizing an intermediate regular expression from out of thin air).
+
+We conjecture that adding composition lenses would not increase the
+expressive power of the language but we have not proven this.  (Honestly,
+the analogous proof for bijective lenses required nearly 100 pages and we
+did not believe the result was important enough to invest this amount of
+work, since the whole algorithm relies on heuristics to terminate in a
+reasonable amount of time anyway.)  We chose to focus our energy on
+understanding the information-theoretic ranking methods, and the
+relationship between simple symmetric lenses and classical symmetric lenses,
+which we view as the core contributions of this paper.
+
+-----
 
@@
 
-29 of the benchmarks are taken from Augeas, a bidirectional tool for
-editing Linux configuration files.  We selected these benchmarks by
-selecting Augeas files one-by-one in alphabetical order (we did
-not cherry-pick programs).  We stopped when we didn't feel that
-adding more benchmarks would be informative. These examples were
-quite large, and involved some very complex transformations, though
-these benchmarks were bijective. These benchmarks demonstrated 
-how our tool performed on large, complex transformations.
+29 of the benchmarks are taken from Augeas, a bidirectional tool for editing
+Linux configuration files.  We chose these benchmarks by selecting Augeas
+files one-by-one in alphabetical order (we did not cherry-pick programs).
+We stopped when we didn't feel that adding more benchmarks would be
+informative. These examples were fairly large and involved some quite
+complex transformations, though only bijective ones. The purpose of these
+benchmarks was to demonstrate how our tool performed on large, complex
+transformations.
 
-The 8 benchmarks from Flash Fill provided a range of transformations,
-none of which were bijective. Again, the FlashFill benchmarks were
-the first 8 benchmarks in the paper.  These benchmarks demonstrated
-how our tool performed on transformation tasks that discard large
-amounts of data.  On occasion, these benchmarks require the use of
-many regular expression equivalence axioms.
+The 8 benchmarks from Flash Fill (the first 8 in the paper) illustrate a
+range of sample transformations, none of them bijective.  These benchmarks
+demonstrate how our tool performs on transformation tasks that _discard_
+large amounts of data.  Some of them require the use of many regular
+expression equivalence axioms.
 
-The remaining 11 benchmarks were chosen in an ad hoc manner --- we
-simply looked around for day-to-day data formats for which it would
-be useful to have lenses.
+The remaining 11 benchmarks were chosen in an ad hoc manner: we simply
+looked around for day-to-day data formats for which it would be useful to
+have lenses.
+
+-----
 
@@
 
-Both the bijective synthesis and the symmetric synthesis engines use a
-pair of collaborating synthesizers that involves (1) a search for a
-compatible pair of regular expressions and (2) search for a lens given
-those regular expressions.
+Both the bijective synthesis and the symmetric synthesis engines use a pair
+of collaborating synthesizers that involves (1) a search for a compatible
+pair of regular expressions and (2) search for a lens given those regular
+expressions.
 
-Bijective synthesis is faster than symmetric synthesis because part
-(2) is much faster.  More specifically, a bijection must translate
-*all* data on the left into data on the right.  By contrast, a symmetric
-synthesis problem has a choice of *which* data on the left to translate
-into data on the right.  This gives rise to a large set of possible
-choices, and symmetric synthesis must consider all of them.
+Bijective synthesis is faster than symmetric synthesis because part (2) is
+much faster.  Specifically, a bijection must translate *all* data on the
+left into data on the right, and this fact constrains the search.  By
+contrast, a symmetric synthesis problem has a choice of *which* data on the
+left to translate into data on the right.  This gives rise to a large set of
+possible choices, and symmetric synthesis must consider all of them.
 
+(Detailed responses to the reviews follow.  The rest of the file is optional.)
 =================================================================================
@@
 
-While the name_conversion expression has about as many AST nodes than the two
-regular expressions, we find it much harder to write.
+While the name_conversion lens has about as many AST nodes than the two
+regular expressions, we find it much trickier to write.
 
-Engineering best practices dictate that we annotate the lenses with their
-types.  While Boomerang does not require these annotations - being able to
-infer the types from the terms - the types of the term serve as
+Arguably, these regular expressions are useful to write anyway -- after all,
+good engineering practice often recommends annotating programs with their
+types!  While Boomerang does not require these annotations - being able to
+infer the types from the terms - the types of the term do serve as
 documentation for future programmers to understand what formats the lens
-maps between.  Furthermore, the types provide resilience in the face of future
-lens modifications, as the types ensure that the lens maps exactly
-between strings of the provided formats. Invalid inputs are not accepted
-nor are the outputs ever invalid.  Consequently, we are not requiring
-additional work. Instead, we are saving programmers from having to 
-writing the lens by leveraging work they should already be performing.
+maps between, and this may be especially useful if the lens itself is
+synthesized.  Furthermore, type annotationss provide resilience in the face
+of future changes, as the types ensure that the lens maps exactly between
+strings of the provided formats.
+
+-----
 
@@
 process is not required, and if the user is confident that certain information
-is "required" or may safely "skipped," they can safely add this information --
+is "required" or may safely "skipped," they can safely add this information; 
 it will not slow down synthesis. Similarly, if the user already has a benchmark
-suite, they may safely include such examples.
+suite, they may safely include all of it right at the beginning.
+
+-----
 
@@
 
-There were a total of 12 uses of "require" in the benchmark suite. There were a
-total of 4 uses of "skip" in the benchmark suite. In total 73 existing lenses
-were used in compositional synthesis (about 1.5 per benchmark on average).
-Typically, we manually decomposed these lenses, though some used the to_upper
-function. We are happy to include a figure detailing the subtasks specified
-during compositional synthesis in the final version of the paper (similar to the
-one on p24 of Synthesizing Bijective Lenses).
+There were a total of 12 uses of "require" in the benchmark suite. There
+were a total of 4 uses of "skip" in the benchmark suite. In total 73
+existing lenses were used in compositional synthesis (about 1.5 per
+benchmark on average).  Typically, we manually decomposed these lenses,
+though some used the to_upper function. (We like the idea of including a
+figure detailing the subtasks specified during compositional synthesis in
+the final version of the paper, similar to the one on p24 of Synthesizing
+Bijective Lenses.)
+
+-----
 
@@
 
-We included an extended comparison to a current version of FlashFill in previous
-work, and did not include such a comparison in this work for lack of space. We
-are happy to make this comparison to FlashFill cited more explicitly.
+We included an extended comparison to a current version of FlashFill in
+previous work, but did not include such a comparison in this work for lack
+of space. We are happy to cite this comparison to FlashFill more explicitly.
+
+-----
 
@@
 
+-----
+
     - L 284: please explain why this rule is needed for the lens in Fig. 3 to be
@@
 
+-----
+
     - Section 5.6: expansion inference seems to be a crucial part of the
@@
 
+-----
+
     - Can you explain why you need the lower bound [for intervals]?
 
-While we do not need it in the algorithm, we include it to make formal claims
-about bounding the cost. We describe why we need intervals for the formal claims
-in 4.1.
+While we do not need it in the algorithm, we include it to make formal
+claims about bounding the cost. We describe why we need intervals for the
+formal claims in 4.1.
 
@@
 
-The equational theory shown in Fig. 4. is incomplete, for it does not include
-some of the non-star-semiring equivalences (for example: the regular expression
-equivalence of (A | B)* = (A*B)*A* does not currently have an analogue in this
-equational theory). Furthermore, it is currently unclear whether the natural
-extensions to Conway's equational theory is sufficient for completeness, and
-whether Kozen and Solomaa's deductive systems are easily amenable to SREs.
+The equational theory shown in Fig. 4. is incomplete: it does not include
+some of the non-star-semiring equivalences (for example: the regular
+expression equivalence of (A | B)* = (A*B)*A* does not currently have an
+analogue in this equational theory). Furthermore, it is currently unclear to
+us whether the natural extension to Conway's equational theory is sufficient
+for completeness, or whether Kozen and Solomaa's deductive systems are
+easily amenable to SREs.
+
+-----
 
@@
 
+-----
+
     - p10:L489 relaxing this restrict is not problemantic
 
-For all these combinators, we can follow the style of
-unambiguity requirements on the types that was used in prior lens work. For
-example, the original Boomerang paper explicitly includes such unambiguity
-checks within the typing judgments; these unambiguity checks can be used in the
+For all these combinators, we can follow the style of "unambiguity
+requirements" on the types that was used in prior lens work. For example,
+the original Boomerang paper explicitly includes such unambiguity checks
+within its typing judgments; these unambiguity checks can be used in the
 same way for the symmetric lens variants of these combinators.
+
+-----
 
@@
 In general, these derivations follow the information theory, so as to make
-Theorem 3 true. Most strictly follow the rules of information theory.
-The only ones that are more complex are merges. Calculating the expected
+Theorem 3 true. Most strictly follow the rules of information theory.  The
+only ones that are more complex are merges. Calculating the expected
 information to recover a string from a lens using merges does not involve a
-simple inductive calculation, but requires reasoning about the probabilities of
-individual strings in the unioned format, conditioned on the probabilities of
-the source format. Because of this complexity, we chose to bound the information
-content instead of calculate it exactly.
+simple inductive calculation, but requires reasoning about the probabilities
+of individual strings in the unioned format, conditioned on the
+probabilities of the source format. Because of this complexity, we chose to
+bound the information content instead of calculate it exactly.
+
+-----
 
@@
 
+-----
+
     - p14:L650 distance d.  What is "distance"?
@@
 the pair of regular expressions to convert to the pair.
+
+-----
 
@@
 
-We only ran 5 times because there was so little variance between the times, as
+We only ran five times because there was so little variance between the times, as
 our procedure is deterministic. Rerunning, we found our average standard
@@
 to both include standard deviations and rerun additional times.
+
+-----
 
@@
 
+-----
+
     - Similar to the above, the graph in figure 8 and the accompanying
@@
 
-Requires help with continuing searching when the desired lenses require a large
-number of Expand applications. Skips help when the desired lenses require
-projecting large amounts of information. Using the DC metric only works for
-lenses where the desired lens appears quite early. FL is the same as DC, but is
-even more dependent on the desired lens appearing early (it *must* appear in the
-first cluster containing a satisfying program).
+"Require"s help with continuing searching when the desired lenses require a
+large number of Expand applications. "Skip"s help when the desired lenses
+involve projecting large amounts of information. Using the DC metric only
+works for lenses where the desired lens appears quite early. FL is the same
+as DC, but is even more dependent on the desired lens appearing early (it
+*must* appear in the first cluster containing a satisfying program).

### 2019-05-15 papers/SymmetricLenses/Conference/main.tex
move conference paper to conference directory
add updates

--- 2019-03-01 papers/SymmetricLenses/main.tex
@@
 % \newif\ifdraft\draftfalse  % set true to show comments
-\newif\ifanon\anontrue    % set true to suppress names, etc.
+\newif\ifanon\anonfalse    % set true to suppress names, etc.
 \newif\ifappendices\appendicesfalse
@@
 %\documentclass[acmsmall,review,screen,anonymous]{acmart}
-\documentclass[acmsmall,review,screen,anonymous]{acmart}
-\settopmatter{printacmref=false,printccs=false}
+\documentclass[acmsmall,screen]{acmart}
+
+\pdfpagewidth=8.5in 
+\pdfpageheight=11in
+
+\settopmatter{}
+
+\setcopyright{rightsretained}
+\acmPrice{}
+\acmDOI{xxx.xxx/xxxx}
+\acmYear{2019}
+\copyrightyear{2019}
+\acmJournal{PACMPL}
+\acmVolume{2}
+\acmNumber{ICFP}
+\acmArticle{80}
+\acmMonth{8}
 
@@
 
+\begin{CCSXML}
+<ccs2012>
+<concept>
+<concept_id>10011007.10011006.10011050.10011017</concept_id>
+<concept_desc>Software and its engineering~Domain specific languages</concept_desc>
+<concept_significance>500</concept_significance>
+</concept>
+<concept>
+<concept_id>10011007.10011006.10011066.10011070</concept_id>
+<concept_desc>Software and its engineering~Application specific development environments</concept_desc>
+<concept_significance>300</concept_significance>
+</concept>
+</ccs2012>
+\end{CCSXML}
+
+\ccsdesc[500]{Software and its engineering~Domain specific languages}
+\ccsdesc[300]{Software and its engineering~Application specific development environments}
+
+\keywords{Bidirectional Programming, Program Synthesis, Type-Directed Synthesis,
+Type Systems}
+
 %\begin{CCSXML}
@@
 places and in different formats. For instance, electronic calendars come in
-iCalendar, vCalendar, and h-event formats~\cite{calendar-formats}; US taxation
-data can be transmitted via Tax XML (used by the US government for e-filing) or
+iCalendar, vCalendar, and h-event formats~\cite{calendar-formats}; U.S. taxation
+data can be transmitted via Tax XML (used by the U.S. government for e-filing) or
 TXF (used by TurboTax)~\cite{tax-formats}; configuration files can be stored in
@@
 corresponding data, Optician's synthesis algorithm is guaranteed to find a
-bijective lens matching the examples, if one exists. Indeed, it can often find
-an appropriate lens in less than a second with no examples at all.
+bijective lens matching the examples, if one exists, often requiring no
+examples at all. While specifying regular expressions can be cumbersome, doing so
+is much easier than programming lenses -- regular expressions are widely
+understood whereas few programmers are fluent in lenses, and writing lenses
+require users reason about how to align two formats at once.
 %, with just types describing the source and target formats.
@@
 concluding thoughts (\S\ref{sec:conc}).
-Elided proofs can be found in the appendices
-\ifappendices\else included with the supplemental material\fi.
+Elided proofs can be found
+\ifappendices in the appendices\else in the full version of the paper~\cite{sslfull}\fi.
 
@@
 \begin{lstlisting}
-let company = (co_name . ("Co." | "Inc." | "Ltd.")) | "UNK" in
-let emp_ins = name . " " . name "," company in
-let header = "FirstLast,Company" in
+let company = (co_name . ("Co." | "Inc." | "Ltd.")) | "UNK"
+let emp_ins = name . " " . name "," company
+let header = "FirstLast,Company"
 let emp_insurance = header . ("\n" . emp_ins)$^*$
@@
   \caption{Hypothetical example data files and corresponding regular expressions
-  used by management and HR at a US company to represent employee salaries and health
+  used by management and HR at a company to represent employee salaries and health
   insurance providers, respectively.
@@
 synthesis algorithm, using a simplified example drawn from a hypothetical
-U.S. company. In this company, management and human resources (HR) store
+company. In this company, management and human resources (HR) store
 information about employees in separate text files: management stores the
@@
 Figure~\ref{fig:minimized-representations}. Applying the \CreateR function to a
-salary file creates a insurance file with the same employee names in the file,
+salary file creates an insurance file with the same employee names in the file,
 with
@@
 \begin{lstlisting}
-let name_lens = id(name) . id(" ") . id(name) . del(": ") . ins(",") in
+let name_lens = id(name) . id(" ") . id(name) . del(": ") . ins(",")
 let employee_lens = name_lens . disconnect(salary,"","unk","") 
-                                                    . disconnect("",company,"","UNK") in
-let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens) in
-ins(header) . employees_lens  :  emp_salaries $\Leftrightarrow$ emp_insurance
+                                                    . disconnect("",company,"","UNK")
+let employees_lens = ins("\n") . employee_lens . iterate(id("\n") . employee_lens)
+let full_lens  :  emp_salaries $\Leftrightarrow$ emp_insurance = ins(header) . employees_lens
 \end{lstlisting}
@@
 \]
-The typing  rules for maintain structural similarity
+The typing  rules maintain structural similarity
 between regular expression types and lens terms, with one exception: The
@@
 conditional entropy of $\RegexAlt$ given $\Regex$ and $\Lens$. This metric
-assigns higher entropy (or likelihood) to lenses where knowing the string on one side 
+assigns higher entropy (or lower likelihood) to lenses where knowing the string on one side 
 provides little information about the string on the other side. It assigns zero
@@
 $\Regex_2$ with probability $1-\Probability$. The SRE
-$\PRegexStar{\lstinline{"c"}}{\Probability}$ generates the empty string with
-probability $1-\Probability$, and generates any other string in the language with
-probability $\Probability$. More generally,
 $\PRegexStar{\Regex}{\Probability}$ generates strings in $\Regex^n$ with
-probability $\Probability^n(1-\Probability)$.
+probability $\Probability^n(1-\Probability)$. For example, if the empty string
+is not in $\Regex$, $\PRegexStar{\Regex}{\Probability}$ will generate the empty
+string with probability $1-\Probability$.
 
@@
 of stochastic regular expression equivalence. Figure~\ref{fig:ssr-sre} shows how
-we to extend the \emph{star-semiring}~\cite{starsemiring} equivalences (a
+we extend the \emph{star-semiring}~\cite{starsemiring} equivalences (a
 \emph{finer} notion of equivalence than semantic equivalence) to
@@
 \begin{theorem}
-  If $\Regex \SSREquiv \RegexAlt$ then $\ProbabilityOf{\Regex}{\String} =
-  \ProbabilityOf{\RegexAlt}{\String}$, for all strings $\String \in \LanguageOf{\Regex}$.
+  If $\Regex \SSREquiv \RegexAlt$, then
+  $\ProbabilityDistOf{\Regex} = \ProbabilityDistOf{\RegexAlt}$.
 \end{theorem}
 
-(For the proof, see the appendix \ifappendices
-Theorem~\ref{thm:semantics-correctness} \else{} included with the
-supplemental material\fi.)  This theorem will come in handy as we traverse
+\ifappendices(For the proof, see the appendix 
+Theorem~\ref{thm:semantics-correctness}.) \else\fi  This theorem will come in handy as we traverse
 (with \Expand) or normalize across (as part of \GreedySynth) star-semiring
@@
 
-Fortunately, we already require unambiguous regular expressions as input to
-our synthesis procedure for other reasons, and we can easily preprocess
-empty subexpressions out of SREs that are themselves nonempty using the
-star-semiring equivalences (\EG, $\PRegexOr{\emptyset}{\lstinline{"s"}}{.5}
-\SSREquiv \lstinline{"s"}$).
+Fortunately, we already require unambiguous regular expressions as input to our
+synthesis procedure to guarantee the synthesized lenses are well-typed, and we
+can easily preprocess empty subexpressions out of SREs that are themselves
+nonempty using the star-semiring equivalences (\EG,
+$\PRegexOr{\emptyset}{\lstinline{"s"}}{.5} \SSREquiv \lstinline{"s"}$).
 
@@
 We present just \CreateR and \PutR in cases where \CreateL and \PutL are
-symmetric; full details can be found in an appendix\ifappendices
-\S\ref{sec:appendixlenses} \else{} (in the supplemental materials)\fi. For
+symmetric; full details can be found in\ifappendices{} appendix
+\S\ref{sec:appendixlenses} \else{} the full version of the paper\fi. For
 simplicity of presentation, we use unambiguous regular expressions everywhere,
@@
 
-The \MergeR lens is interesting because it merges data where one data can be in
+The \MergeR lens is interesting because it merges data where one piece of data can be in
 two formats, and one data has only one format. In previous
@@
 expected amount of information required to recover a string in $\Regex$ from a
-string in $\RegexAlt$, synchronized by $\Lens$. We write $a(b,c)$ to
-mean $(a*b,a*c)$, and $(a,b)+(c,d)$ to mean $(a+c,b+d)$.
+string in $\RegexAlt$, synchronized by $\Lens$. We write $a[b,c]$ to
+mean $[ab,ac]$, and $[a,b]+[c,d]$ to mean $[a+c,b+d]$.
 \begin{center}
@@
     & =
-    & $(0,\REntropyOf{\RegexAlt_1 \Given \Lens_1, \Regex}+\REntropyOf{\RegexAlt_2 \Given \Lens_2, \Regex}+1)$\\
+    & $[0,\REntropyOf{\RegexAlt_1 \Given \Lens_1, \Regex}+\REntropyOf{\RegexAlt_2 \Given \Lens_2, \Regex}+1]$\\
 
@@
 need to determine if the synchronized string is in $\RegexAlt_1$ or $\RegexAlt_2$.
-However, this information content is dependent on the how likely the
+However, this information content is dependent on how likely the
 synchronized string is to be in $\RegexAlt_1$ or $\RegexAlt_2$. Nevertheless, we
@@
 \SDNFREabbrev{}s normalize across many of the star-semiring equivalences -- if
-two regular expressions are equal modulo differences in associativity,
+two regular expressions are equivalent modulo differences in associativity,
 commutativity, or distributivity, their corresponding SREs are syntactically
@@
 The details of these definitions are syntactically tedious, but not
-intellectually difficult. We elide them here but include them in the
-appendix \ifappendices \S\ref{sec:appendixdnf}.  \else{} (in the supplemental
-materials).\fi
+intellectually difficult. We elide them here but include them in
+\ifappendices appendix \S\ref{sec:appendixdnf}.  \else{}the full version of the
+paper.  \fi In the original Optician paper~\cite{optician}, DNF lenses were
+proven equivalent in expressiveness to standard lenses. While we conjecture
+symmetric DNF lenses are equivalent in expressivity to our standard symmetric
+lenses, we have not proven this equivalence.
 
@@
 
-As an example, consider searching for a lens between \lstinline{"" | name.name$^*$} and \lstinline{"" | name}. \GreedySynth might first pop the
-sequence pair \lstinline{""} and \lstinline{""}, because there is a bijective
-sequence lens between them. As neither \lstinline{""} is involved in a sequence
+As an example, consider searching for a lens between \lstinline{"" | name.name$^*$}
+and \lstinline{"" | name}. \GreedySynth might first the
+sequence lens between the
+sequences \lstinline{""} and \lstinline{""}, because it is a bijective
+sequence lens between. As neither \lstinline{""} is involved in a sequence
 lens, this lens is considered useful. Next, the sequence lens between
@@
     \Call{\CartesianMap}{\AtomSynth(\Examples),\ListOf{\Atom_1;\ldots;\Atom_n},\ListOf{\AtomAlt_1;\ldots;\AtomAlt_m}}$
-    \State $\GreedyState \gets \Call{PQ.Init}{\ALS}$
+    \State $\GreedyState \gets \Call{PQ.Create}{\ALS}$
     \State $\SeqLensBuilder \gets \Call{SLensBuilder.Empty}{}$
@@
 certain closed regular expressions can \emph{only} be involved in a disconnect
-lens (unless opened). Such regular expressions will automatically be opened.
+lens (unless opened). Such regular expressions will automatically be opened. The
+full details of expansion inference are explained in \citet{optician}.
 
@@
   cleaning tasks are either derived from online help forums or taken from the
-  Excel product team. Note that our tool produces bidirectional transformations
+  Excel product team. The 8 we chose are merely the first 8 found in the Flash
+  Fill paper. Note that our tool produces bidirectional transformations
   rather than one-way transformers like Flash Fill. We ensure one direction of
@@
   in-memory dictionaries. In our benchmarks, we synthesize lenses between Linux
-  configuration files and serialized versions of Augeas dictionaries. Because
-  these benchmarks merely transformed the information into a structured form,
-  synthesized lenses were all bijective.
+  configuration files and serialized versions of Augeas dictionaries. We
+  selected the first 28 benchmarks from Augeas in alphabetical order, and
+  included one additional benchmark we considered particularly difficult,
+  \texttt{xml\_to\_augeas.boom}. Because these benchmarks merely transformed the
+  information into a structured form, synthesized lenses were all bijective.
 \item We created 11 additional benchmarks 
@@
 Each task consists of a source and target regular expression and a set of
-examples. We are synthesizing transformations between real formats, so these
-tasks are can be quite large -- on average, the size of each RE
-in the specification is 1637 and the size of the synthesized lens is 3254.
+examples. We manually programmed the regular expressions. However, when writing
+these regular expressions, we did not massage them to be amenable to synthesis.
+The synthesizer implements the star-semiring regular expression equivalence
+axioms, so that users do not have to worry about the way they write expressions
+themselves. We are synthesizing transformations between real formats, so these
+tasks are can be quite large -- on average, the size of each RE in the
+specification is 1637 and the size of the synthesized lens is 3254.
 
@@
     benchmarks in under 30 seconds elapsed total time. Without compositional
-    synthesis it is able to complete 26. Each benchmark specification includes
+    synthesis it is able to complete 31. Each benchmark specification includes
     source and target (potentially annotated) regular expressions, and between
@@
 synthesize all of the benchmarks in under 30 seconds. Without compositional
-synthesis, the synthesis algorithm is able to solve 26 out of 48 problem instances.
+synthesis, the synthesis algorithm is able to solve 31 out of 48 problem instances.
 
@@
 complete.
+
+Both the bijective synthesis and the symmetric synthesis engines use a pair of
+collaborating synthesizers that involves (1) a search for a compatible pair of
+regular expressions and (2) search for a lens given those regular expressions.
+Bijective synthesis is faster than symmetric synthesis because part (2) is much
+faster. Specifically, a bijection must translate all data on the left into data
+on the right, and this fact constrains the search. By contrast, a symmetric
+synthesis problem has a choice of which data on the left to translate into data
+on the right. This gives rise to a large set of possible choices, and symmetric
+synthesis must consider all of them.
 
@@
 We experimented with the \CCOpt{} mode to determine whether the
-complexity of the information-theoretic measure was really
+complexity of the information-theoretic measure is really
 needed. Related work on string transformations has often used simpler
@@
 and asymmetric lenses. Proofs about these relationships are included in
-the appendix \ifappendices  \S\ref{sec:appendixforget}. \else{}included with the supplemental
-material.  \fi We also comment on related
+\ifappendices appendix \S\ref{sec:appendixforget}. \else{}the full version of
+the paper.  \fi We also comment on related
 program synthesis techniques.
@@
     \UnaryInfC{$\Lens.putl(y,c_2) = (\_,c_2')$}
-    \AxiomC{$putr(x,c_1') = (\_,c_1'')$}
+    \AxiomC{$\Lens.putr(x,c_1') = (\_,c_1'')$}
     \def\extraVskip{.5pt}
@@
 Automation of string transformations, in particular, has been the focus
-of much prior attention.  For example, Gulwani and others work on FlashFill
+of much prior attention.  For example, Gulwani's and others' work on FlashFill
 generates one-way spreadsheet transformations from input/output
@@
 and hence is incapable of synthesizing the majority of our
-benchmarks, even in one direction.
+benchmarks, even in one direction. A comparison of Optician (on
+bijective benchmarks) to these synthesis tools is included in~\citet{optician}.
 %% Flashfill, designed
@@
 natural language understanding, machine translation, information retrieval, data
-extraction and grammatical inference (see \citet{pereira:info-theory},
-for an introduction). Indeed, our work was inspired, in part, by the PADS format
+extraction and grammatical inference (see \citet{pereira:info-theory}, for an
+introduction). Indeed, our work was inspired, in part, by the PADS format
 inference tool~\cite{pads:synthesis}, which was in turn inspired by earlier work
@@
 data transformers, and we are not aware of the use of information-theoretic
-principles in type-directed or syntax-guided synthesis of DSL programs.
+principles in type-directed or syntax-guided synthesis of DSL programs. More
+recently, the principles used in \citet{pads:synthesis} have been applied in
+unsupervised learning~\cite{ULPS}. These works learn compact descriptions of a
+single data set. Ideally, such descriptions are compact and information theory
+is used as a measure of the compactness of the description learned. In contrast,
+we are attempting to learn a translation from data set to another. However,
+there are many different candidate translations. To select amongst the candidate
+translations, we choose the translation that preserves as much information from
+source to target (and vice versa) as possible.
 
@@
 real-life formats in under 30 seconds.
-
 % end conclusion
@@
 \begin{acks}
+  We thank our anonymous reviewers for their useful feedback and discussions and
+  Nate Foster and Michael Greenberg for their help integrating Optician into
+  Boomerang. This research has been supported in part by DARPA award
+  FA8750-17-2-0028 and ONR 568751 (SynCrypt).
 \end{acks}

### 2019-05-29 papers/SymmetricLenses/ArtifactEvaluationPaper/main.tex
more updates

--- 2017-11-10 papers/BijectiveLenses/ArtifactEvaluationPaper/main.tex
@@
 
-\title{Synthesizing Bijective Lenses: Artifact Evaluation}
+\title{Synthesizing Symmetric Lenses: Artifact Evaluation}
 
@@
 \email{kfisher@eecs.tufts.edu}
+
+\author{Solomon Maina}
+\affiliation{University of Pennsylvania, USA}
+\email{smaina@seas.upenn.edu}
 
@@
 \item Install a Virtual Machine Manager (we tested with VirtualBox)
-\item Download our virtual machine image (BijectiveLensSynthesis.ova in\\
+\item Download our virtual machine image (SymmetricLensSynthesis.ova in\\
   {\color{blue}
-    \url{https://drive.google.com/drive/folders/0B5OYxhNvX675cTN4ZzVEMmJ0Mlk})}
+    \url{https://drive.google.com/drive/folders/1en7-K6Kxp7MvF3Z1GGfpCDBIKuF-84YM})}
 \item Load our virtual machine from your virtual machine manager
@@
 
-We suggest the committe dedicate over 8GB RAM to this environment.
-While we get similar results with less RAM, some of the existing tools we
-compare our artifact to use up to 8GB RAM.  Less RAM may cause these tools to
-perform worse than presented in our data.
+We suggest the committe dedicate at least 8GB RAM to this environment.
 
-The user for the virtual machine is bls, with password bls.
+The user for the virtual machine is ae, with password 2019.
 
@@
 If you want to install on a custom environment, certain programs and libraries
-must be installed.  We provide the commands for installation in Ubuntu 16.04.3
+must be installed.  We provide the commands for installation in Ubuntu 16.04.6
 in parenthesis next to the program or library name.
@@
   opam update)
-\item Switch opam to OCaml 4.03.0 (opam switch 4.03.0; eval `opam config env`)
 \item Install necessary system packages (opam depext conf-m4.1)
+\item Install OCaml's JBuilder (opam install jbuilder)
 \item Install OCaml's Core (opam install core)
@@
 \item Install OCaml's ppx\_deriving (opam install ppx\_deriving)
+\item Install OCaml's ppx\_hash (opam install ppx\_hash)
 
@@
 \item Install Python's tk (sudo apt install python-tk)
+\item Install Python's numpy (sudo apt install numpy)
 
-\item Install DotNet runtime environment, and C\# and Visual Basic libraries and
-  compilers (sudo apt install mono-complete; sudo apt install mono-vbnc)
-
-\item Install the codebase\\
-  (git clone https://github.com/Miltnoid/BijectiveLensSynthArtifactEvaluation.git)
+\item Download the codebase\\
+  (git clone https://github.com/Miltnoid/SymmetricLensSynthArtifactEval.git)
 \end{enumerate}
@@
 To validate: navigate to the directory the codebase is installed in
-(/home/bls/LensSynth in the virtual machine).
-Then run the following commands:
-\begin{enumerate}
-\item make regenerate-specs
-\item make regenerate-data
-\end{enumerate}
+(/home/ae/SymmetricLensSynthArtifactEval in the virtual machine).
+Then run the following command: ``make regenerate-data''
 
-After these commands are run, the figures will be present at the following
+After this command is run, the figures will be present at the following
 locations: 
 \begin{itemize}
-\item Figure 8 will be present at \$/generated-graphs/times.eps
-\item Figure 9 will be present at \$/generated-graphs/uninferred.eps
-\item Figure 10 will be present at \$/generated-graphs/compositional.eps
-\item Figure 11 will be present at \$/generated-graphs/examples.eps
+\item Figure 6 will be present at \$/generated-graphs/times.eps
+\item Figure 7 will be present at \$/generated-graphs/times\_bijective.eps
+\item Figure 8 will be present at \$/generated-graphs/metrics\_importance.eps
 \end{itemize}
 
-The command ''make regenerate-specs`` will take about an hour to complete.  It
+The command ``make regenerate-data'' will take a few hours to complete. It
 must run to completion for the data to be saved.
@@
 
-The command, ``make regenerate-data'' can be stopped when partially completed.
-After a benchmark has had all tests run on it, that data is saved, so future
-runs will not rerun tests on that benchmark.  Each benchmark takes between 0 and
-3 hours to complete, with information being printed to stdout about what
-test and benchmark is being run.  The total time to complete all tests takes
-between 2 and 3 days.
-
 \subsection{Directory Information}
@@
 If the artifact committee want to do more in-depth testing, we provide
-information on the individual project directories.  The means to run individual
-programs are described in README.txt files in the directories.
+information on the individual project directories. Boomerang can be built by
+running ``make'' in the \$/boomerang directory, and run with ./boomerang.exe.
 
-\paragraph*{\$/program}  Our tool is present in the program directory.  The
-source code is provided in \$/program/src .  The directory \$/program/tests
-contains a test suite, consisting of microbenchmarks and the tested benchmarks.
-The directory \$/program/example-inputs contains the tested benchmarks.
-The directory \$/program/generated\_data is the output location for the csv created from
-program runs.  The directories \$/program/generated\_extraction\_specs and
-\$/program/generated\_io\_specs are the output locations for the specifications
-given to Flash Fill and FlashExtract.
+\paragraph*{\$/boomerang} Boomerang enhanced with symmetric lenses is present in
+the program directory. The optician source code is provided in
+\$/boomerang/optician . The directory \$/program/examples contains example
+boomerang programs, and the directory \$/program/examples/synth\_examples
+contains examples using synthesis.
 
-\paragraph*{\$/comparisons}  When created by ``make regenerate-specs'', the
-directories \$/comparisons/extraction\_specs and \$/comparisons/io\_specs
-contain the specifications given to Flash Fill and FlashExtract.  The directory
-\$/comparisons/prose/Transformation.Text contains the generated data for its
-runs under generated\_data, and the source code for our tests in
-SampleProgram.cs.  The directory
-\$/comparisons/prose/Extraction.Text contains the generated data for its
-runs under generated\_data, and the source code for our tests in
-LearningSamples.cs.
-
-\paragraph*{\$/generated-data}  After data is created by the individual tests,
-it is then combined and placed into \$/generated-data/data.csv.
+\paragraph*{\$/generated-data} After data is created by the individual tests, it
+is then combined and placed into \$/generated-data/data.csv.
 

### 2019-05-30 papers/SymmetricLenses/Conference/main.tex
update main and cover letter

@@
 }
-\email{smaina@cis.upenn.edu}
+\email{smaina@seas.upenn.edu}
 
@@
 \maketitle
+\renewcommand{\shortauthors}{A. Miltner, S. Maina, K. Fisher, B. C. Pierce, D. Walker, S. Zdancewic}
 %\fi
@@
 understood whereas few programmers are fluent in lenses, and writing lenses
-require users reason about how to align two formats at once.
+requires users to reason about how to align two formats at once.
 %, with just types describing the source and target formats.
@@
 $\PRegexStar{\Regex}{\Probability}$ generates strings in $\Regex^n$ with
-probability $\Probability^n(1-\Probability)$. For example, if the empty string
-is not in $\Regex$, $\PRegexStar{\Regex}{\Probability}$ will generate the empty
-string with probability $1-\Probability$.
+probability $\Probability^n(1-\Probability)$. For example,
+$\PRegexStar{\Regex}{\Probability}$ will generate a string in $\Regex$ with
+probability $\Probability(1-\Probability)$ and a string in
+$\RegexConcat{\Regex}{\Regex}$ with probability $\Probability^2(1-\Probability)$.
 
@@
 As an example, consider searching for a lens between \lstinline{"" | name.name$^*$}
-and \lstinline{"" | name}. \GreedySynth might first the
+and \lstinline{"" | name}. \GreedySynth might first pop the
 sequence lens between the
@@
 tasks are can be quite large -- on average, the size of each RE in the
-specification is 1637 and the size of the synthesized lens is 3254.
+specification is 1637 and the size of the synthesized lens is 3254. For each of
+these tasks, we selected tasks until we did not feel additional examples would
+be elucidating.
 
@@
 synthesize all of the benchmarks in under 30 seconds. Without compositional
-synthesis, the synthesis algorithm is able to solve 31 out of 48 problem instances.
+synthesis, the synthesis algorithm is able to solve 31 out of 48 problem
+instances. In total, 73 existing lenses were used in compositional synthesis
+(about 1.5 per benchmark on average).
 
@@
 
-Both the bijective synthesis and the symmetric synthesis engines use a pair of
-collaborating synthesizers that involves (1) a search for a compatible pair of
-regular expressions and (2) search for a lens given those regular expressions.
-Bijective synthesis is faster than symmetric synthesis because part (2) is much
-faster. Specifically, a bijection must translate all data on the left into data
-on the right, and this fact constrains the search. By contrast, a symmetric
-synthesis problem has a choice of which data on the left to translate into data
-on the right. This gives rise to a large set of possible choices, and symmetric
-synthesis must consider all of them.
+Both the bijective synthesis and the symmetric synthesis engines use a pair of collaborating
+synthesizers that (1) search for a compatible pair of regular expressions and (2) search
+for a lens given those regular expressions. Bijective synthesis is faster than symmetric synthesis
+because part (2) is much faster. Specifically, a bijection must translate all data on the left into data
+on the right, and this fact constrains the search. By contrast, a symmetric synthesis problem has a
+choice of which data on the left to translate into data on the right. This choice gives rise to a large set of
+other choices, and symmetric synthesis must consider all of them.
 
@@
 lens for 40 of our 48 benchmarks; eight required relevance annotations to find
-the correct lens.
+the correct lens. In total, there are 12 uses of \SRequire, and 4 uses of \Skip
+in our benchmark suite.
 % end evaluation
@@
 recently, the principles used in \citet{pads:synthesis} have been applied in
-unsupervised learning~\cite{ULPS}. These works learn compact descriptions of a
+unsupervised learning~\cite{ULPS}. This work learns compact descriptions of a
 single data set. Ideally, such descriptions are compact and information theory
 is used as a measure of the compactness of the description learned. In contrast,
-we are attempting to learn a translation from data set to another. However,
+we are attempting to learn a translation from one data set to another. However,
 there are many different candidate translations. To select amongst the candidate

### 2020-04-07 papers/BijectiveLenses/IWC/main.tex
first draft

% easychair.tex,v 3.5 2017/03/15

\documentclass{easychair}
%\documentclass[EPiC]{easychair}
%\documentclass[EPiCempty]{easychair}
%\documentclass[debug]{easychair}
%\documentclass[verbose]{easychair}
%\documentclass[notimes]{easychair}
%\documentclass[withtimes]{easychair}
%\documentclass[a4paper]{easychair}
%\documentclass[letterpaper]{easychair}

\usepackage{doc}

%\newif\ifdraft\drafttrue      % Uncomment to turn on comments
\newif\ifdraft\draftfalse   % Uncomment to turn off comments

% use this if you have a long article and want to create an index
% \usepackage{makeidx}

% In order to save space or manage large tables or figures in a
% landcape-like text, you can use the rotating and pdflscape
% packages. Uncomment the desired from the below.
%
% \usepackage{rotating}
% \usepackage{pdflscape}

% Some of our commands for this guide.
%
\newcommand{\easychair}{\textsf{easychair}}
\newcommand{\miktex}{MiK{\TeX}}
\newcommand{\texniccenter}{{\TeX}nicCenter}
\newcommand{\makefile}{\texttt{Makefile}}
\newcommand{\latexeditor}{LEd}

%\makeindex

%% Front Matter
%%
% Regular title as in the article class.
%
%
\title{Confluence in Lens Synthesis}

% Authors are joined by \and. Their affiliations are given by \inst, which indexes
% into the list defined using \institute
%
\author{
Anders Miltner\inst{1}
\and
Kathleen Fisher\inst{2}
\and
Benjamin C. Pierce\inst{3}
\and
David Walker\inst{4}
\and
Steve Zdancewic\inst{5}
}

% Institutes for affiliations are also joined by \and,
\institute{
  Princeton University\\
  \email{amiltner@cs.princeton.edu}
\and
   Tufts University\\
   \email{kfisher@eecs.tufts.edu}\\
\and
   University of Pennsylvania\\
   \email{bcpierce@cis.upenn.edu}\\
\and
   Princeton University\\
  \email{dpw@cs.princeton.edu}
\and
  University of Pennsylvania\\
  \email{stevez@cis.upenn.edu}\\
}

%  \authorrunning{} has to be set for the shorter version of the authors' names;
% otherwise a warning will be rendered in the running heads. When processed by
% EasyChair, this command is mandatory: a document without \authorrunning
% will be rejected by EasyChair

\authorrunning{Miltner, Fisher, Pierce, Walker and Zdancewic}

% \titlerunning{} has to be set to either the main title or its shorter
% version for the running heads. When processed by
% EasyChair, this command is mandatory: a document without \titlerunning
% will be rejected by EasyChair
\titlerunning{Confluence in Lens Synthesis}

\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{bussproofs}
\usepackage{amsthm}
\usepackage{csvsimple}
\usepackage{thmtools,thm-restate}
\usepackage{changepage}
\usepackage{booktabs}
\usepackage{amssymb}
\usepackage[inline]{enumitem}
\usepackage{multirow,bigdelim}
\usepackage{multicol}
\usepackage{siunitx}
\usepackage{listings}
\usepackage{sansmath}
\usepackage{url}
\usepackage{flushend}
\usepackage{microtype}
\usepackage[utf8]{inputenc}
\usepackage{mathpartir}
\usepackage{empheq}
\usepackage{array}
\usepackage{pgfplots}
\usepackage{stmaryrd}
\usepackage{courier}
\usepackage{qtree}
\usepackage[normalem]{ulem}
\usepackage{relsize}
\usepackage{tikz}
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{textcomp}
\usepackage{tabularx}
\usepackage{stackengine}
\usepackage{caption}
\usepackage{wrapfig}
\usepackage{remreset}
\usepackage{tabulary}

\usetikzlibrary{
  er,
  matrix,
  shapes,
  arrows,
  positioning,
  fit,
  calc,
  pgfplots.groupplots,
  arrows.meta
}
\tikzset{>={Latex}}

\input{macros}

\begin{document}

\maketitle

\begin{abstract}
  Bidirectional programs are programs where a single expression has multiple
  interpretations, encoding a number of functions with a single piece of syntax.
  Lenses are bidirectional programs with semantic guarantees, and lens languages
  have been developed for synchronizing string formats, building updatable views
  in databases, and more. These languages often are typed, where a well-typed
  expression will provide the semantic guarantees. Recent research has shown how
  string lenses can be synthesized from their types, pairs of regular
  expressions. However, providing completeness guarantees is quite tricky on
  these languages, due in large part to the many equivalences on regular
  expressions.

  In proving a completeness guarantee, a confluence-like property,
  $R$-confluence, must be shown. This property can be thought of as a
  generalization of standard confluence. Where standard confluence is
  parameterized by a set $S$ and a binary relation $\rightarrow$, indexed
  confluence is parameterized by an additional binary relation $R \subseteq S
  \times S$. In this model, confluence is the specific case where $R$ is just
  equality. We show how the proof that the transitive closure of a rewrite
  system is $R$-confluent differs from the proof strategy for standard
  confluence.
\end{abstract}

\section{Introduction}

Bidirectional transformations are pervasive in modern software systems, occuring
as database views and view updaters, parsers and pretty-printers, data
synchronization tools, and more. Instead of manually building the functions that
comprise a bidirectional transformation, programmers can build them both ``at
once'' using a bidirectional programming language. Bidirectional programming
languages have been developed for creating view updaters~\cite{?}, Linux
configuration file editors~\cite{?}, direct manipulation programming
systems~\cite{?}, and more~\cite{?,?,?}. Lenses are a particularly well-behaved
class of bidirectional programs, where the underlying transformations are
guaranteed to satisfy a number of ``round-tripping'' laws. Lens-based
bidirectional programming languages often provide round-tripping guarantees
through a set of typing rules; well-typed lens expressions are guaranteed to
satisfy the round-tripping laws.

To make bidirectional programming easier, Optician~\cite{?,?,?} synthesizes
bidirectional programs from these types, specifically in Boomerang, a language
for synchronizing string data. Optician takes as input two regular expressions
(which serve as the type of a Boomerang lens) and a set of examples specifying
input-output behavior, and synthesizes a well-typed lens between the languages
of those regular expressions.

The judgement $\Lens : \Regex \Leftrightarrow \RegexAlt$ means that $\Lens$
bidirectionally converts between $\LanguageOf{\Regex}$ and
$\LanguageOf{\RegexAlt}$. For synthesis, given $\Regex$ and $\RegexAlt$, we must
find a lens $\Lens$ such that $\Lens : \Regex \Leftrightarrow \RegexAlt$. There
are a number of typing rules for lenses, but the rules for type-equivalence and
composition (shown below) do not permit an efficient search algorithm. Note that
in type equivalence we use a finer notion of equivalence than language
equivalence, we instead use star-semiring equivalence. We were unable to make
our proof strategy work for language equivalence, so instead guarantee
completeness only on the subset of lenses well-typed with star-semiring
equivalence.
\begin{mathpar}
  \inferrule[\RewriteRegexLensRule{}]
  {
    \Lens \OfType \Regex \Leftrightarrow \RegexAlt\\
    \Regex \SSREquiv \Regex'\\
    \RegexAlt \SSREquiv \RegexAlt'
  }
  {
    \Lens \OfType \Regex' \Leftrightarrow \RegexAlt'
  }

  \inferrule[\ComposeLensRule{}]
  {
    \Lens_1 \OfType \Regex_1 \Leftrightarrow \Regex_2\\
    \Lens_2 \OfType \Regex_2 \Leftrightarrow \Regex_3\\
  }
  {
    \ComposeLensOf{\Lens_1}{\Lens_2} \OfType \Regex_1 \Leftrightarrow \Regex_3
  }
\end{mathpar}

Trying to synthesize a lens from the types by applying typing rules ``in
reverse'' gets confounded by these two rules. They require no significant
restriction on the output types, and so can be applied at any time. Furthermore,
applying them in reverse requires pulling regular expressions ``out of thin
air.'' For \RewriteRegexLensRule, these regular expressions need only be
equivalent to the input regular expressions, and for \ComposeLensRule, there are
no restrictions at all!

Our solution is to synthesize in an alternative language of disjunctive normal
form lenses (DNF lenses), typed by disjunctive number form regular expressions
(DNF REs). Both DNF lenses and DNF REs can be thought of as lenses and regular
expressions, put into a pseudonormal form.

\begin{figure}
  \centering
  \begin{tikzpicture}[auto,node distance=1.5cm]
    \node[text width=1.5cm,minimum height=.6cm,align=center,draw,rectangle] (todnfregex) {\ToDNFRegex{}};

    \node[align=right, anchor=east] (regex1) [left = .6cm of todnfregex.north west]{\Regex{}};
    \node[align=right, anchor=east] (regex2) [left = .6cm of todnfregex.south west]{\RegexAlt{}};
    \node[align=right, anchor=east] (exs) [below = .2cm of regex2]{ \Examples{} };

    \node[align=center] (dnfregex1) [right = .4cm of todnfregex.north east]{\DNFRegex{}};
    \node[align=center] (dnfregex2) [right = .4cm of todnfregex.south east]{\DNFRegexAlt{}};

    \node[text width=2.9cm,minimum height=.6cm,align=center,draw,rectangle] [right = 1.45cm of todnfregex.east] (synthdnflens) {\SynthDNFLens{}};
    \node[align=center] [above = .7cm of synthdnflens] (optician) {\Optician{}};

    \node[align=center] [right = .4cm of synthdnflens] (dnflens) {\DNFLens{}};

    \node[text width=1.5cm,minimum height=.6cm,align=center,draw,rectangle] [right = .4cm of dnflens] (tolens) {\ToLens{}};

    \node[align=center] [right = .6cm of tolens] (lens) {\Lens{}};

    \path[->] (regex1.east) edge (todnfregex.north west);
    \path[->] (regex2.east) edge (todnfregex.south west);

    \path[->] (todnfregex.north east) edge (dnfregex1.west);
    \path[->] (todnfregex.south east) edge (dnfregex2.west);

    \path[->] (dnfregex1.east) edge (synthdnflens.north west);
    \path[->] (dnfregex2.east) edge (synthdnflens.south west);

    \path[->] (synthdnflens) edge (dnflens);

    \path[->] (dnflens) edge (tolens);

    \path[->] (tolens) edge (lens);
    \draw[->] ($(exs.east)+(-3pt,0)$) -| node(exsedge) {} (synthdnflens);
    \node[fit={($(todnfregex.west)+(-4pt,0)$) ($(tolens.east)+(4pt,0)$) (exsedge) (dnfregex1) (optician) (dnfregex2)},draw] (surrounding) {};
    % Now place a relation (ID=rel1)
    %\node[text width=2cm,align=center,draw, rectangle] (sketch-gen) [right = .75cm of spec] {\TypeProp{}};
    %\node (below-gen) [below=.5cm of sketch-gen] {};
    %\node[text width=2cm,align=center,draw, rectangle] (sketch-compl)
    %     [right = .25cm of sketch-gen] {\RigidSynth{}};
    %\node (below-compl) [below=.5cm of sketch-compl] {};
    %\node[align=center] (lens) [right = .75cm of sketch-compl] {Lens}; 
    %% Draw an edge between rel1 and node1; rel1 and node2
    %\path[->] (spec) edge node (start-alg) {} (sketch-gen);
    %\path[->] (sketch-gen) edge node(middle) {} (sketch-compl);
    %\path[->] (sketch-compl) edge node[near start](success) {\Success{}} (lens);
    %\draw[<-] (sketch-gen.south) -- +(0,-.5) -| node[above left](failure){\Failure{}} (sketch-compl.south);

    %\node (synth-name) [above=.5cm of middle] {\Optician{}};
    %
    %\node[fit=(sketch-gen) (sketch-compl) (start-alg) (synth-name) (failure) (success) ,draw] (surrounding) {};
  \end{tikzpicture}
  \caption{Schematic Diagram for \Optician{}.  Regular expressions, \Regex{} and
    \RegexAlt{}, and examples, \Examples{}, are given as input.
    First, the function \ToDNFRegex{} converts \Regex{} and \RegexAlt{} into
    their respective DNF forms, \DNFRegex{} and \DNFRegexAlt{}.
    Next, \SynthDNFLens{} synthesizes a DNF lens, \DNFLens{}, from \Regex{},
    \RegexAlt{}, and \Examples{}.
    Finally, \ToLens{} converts \DNFLens{} into \Lens{}, a lens in Boomerang
    that is equivalent to \DNFLens{}.}
  \label{fig:schematic-diagram-synthesis}
\end{figure}

Two typing judgements are involved in typing a DNF lens, $\DNFLens : \DNFRegex
\Leftrightarrow \DNFRegexAlt$ and $\DNFLens \OfRewritelessType \DNFRegex
\Leftrightarrow \DNFRegexAlt$. In both instances, the judgement means $\DNFLens$
is a DNF lens that converts between $\LanguageOf{\DNFRegex}$ and
$\LanguageOf{\DNFRegexAlt}$. If $\DNFLens \OfRewritelessType \DNFRegex
\Leftrightarrow \DNFRegexAlt$, then the typing derivation of $\DNFLens$ never
rewrites the DNF REs. The only typing rule for $\DNFLens : \DNFRegex
\Leftrightarrow \DNFRegexAlt$ is:
\begin{mathpar}
  \inferrule[\RewriteDNFRegexLensRule{}]
  {
    \DNFRegex' \StarOf{\Rewrite} \DNFRegex\\
    \DNFRegexAlt' \StarOf{\Rewrite} \DNFRegexAlt \\
    \DNFLens \OfRewritelessType \MapsBetweenTypeOf{\DNFRegex}{\DNFRegexAlt}
  }
  {
    \DNFLens \OfType \MapsBetweenTypeOf{\DNFRegex'}{\DNFRegexAlt'}
  }
\end{mathpar}
To type a DNF lens, one first applies a number of rewrites to the DNF REs, then
builds up a DNF lens using rewriteless typing rules. Because of this, equivalent
types only need to be generated once, and composition never need be applied.
This permits an efficient algorithm that first searchest through type
equivalences with rewrites, and second builds up a rewriteless DNF lens purely
through the syntax of the DNF RE. This then permits the full synthesis
algorithm, shown in Figure~\ref{fig:schematic-diagram-synthesis}. The input
regular expressions are converted into DNF form with $\ToDNFRegex$. Next, a
well-typed DNF lens is synthesized, and then converted back into a standard lens
language with $\ToLens$.

Proving this search procedure complete reduces to proving DNF lenses complete
with respect to our standard lens language. In particular, we wish the prove the
following theorem:
\begin{theorem}
  If $\Lens : \Regex \Leftrightarrow \RegexAlt$, then there exists a DNF lens,
  $\DNFLens$, such that $\DNFLens \OfType \ToDNFRegexOf{\Regex} \Leftrightarrow
  \ToDNFRegexOf{\RegexAlt}$, such that the semantics of $\Lens$ and $\DNFLens$
  are equivalent.
\end{theorem}

We prove this by induction, and particular difficulty lies in the lens
equivalence rule. We begin this case below:

\begin{mathpar}
  \inferrule
  {
    \Lens \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
    \Regex \SSREquiv \Regex_2\\
    \RegexAlt \SSREquiv \RegexAlt_2
  }
  {
    \Lens \OfType \Regex_2 \Leftrightarrow \RegexAlt_2
  }
\end{mathpar}

By induction assumption, there exists $\DNFLens \OfType \ToDNFRegexOf{\Regex_1}
\Leftrightarrow \ToDNFRegexOf{\RegexAlt_1}$, where the semantics of $\DNFLens$
are equivalent to those of $\Lens$. By inversion
\begin{mathpar}
  \inferrule
  {
    \ToDNFRegexOf{\Regex_1} \StarOf{\Rewrite} \DNFRegex_1\\
    \ToDNFRegexOf{\RegexAlt_1} \StarOf{\Rewrite} \DNFRegexAlt_1\\
    \DNFLens \OfRewritelessType \DNFRegex_1 \Leftrightarrow \DNFRegexAlt_1
  }
  {
    \DNFLens \OfType \ToDNFRegexOf{\Regex_1} \Leftrightarrow \ToDNFRegexOf{\RegexAlt_1}
  }
\end{mathpar}

To complete this case, we need to find a DNF lens $\DNFLens' \OfType
\ToDNFRegexOf{\Regex_2} \Leftrightarrow \ToDNFRegexOf{\RegexAlt_2}$ with
equivalent semantics to $\Lens$.

We first show that there exists $\DNFRegex_2$ and $\DNFRegexAlt_2$ such that
$\ToDNFRegexOf{\Regex_2} \StarOf{\rightarrow} \DNFRegex_2$ and
$\ToDNFRegexOf{\DNFRegex_1} \StarOf{\rightarrow} \DNFRegex_2$ and
$\ToDNFRegexOf{\RegexAlt_2} \StarOf{\rightarrow} \DNFRegexAlt_2$ and
$\ToDNFRegexOf{\DNFRegexAlt_1} \StarOf{\rightarrow} \DNFRegexAlt_2$. To do this,
we first prove that $\StarOf{\rightarrow}$ is confluent, and that if $\Regex
\SSREquiv \RegexAlt$, then $\ToDNFRegexOf{\Regex}$ is related to
$\ToDNFRegexOf{\RegexAlt}$ through the transitive, reflexive closure of
$\rightarrow$. We then apply a corollary of confluence to show the existance of
such a $\DNFRegex_2$ and $\DNFRegexAlt_2$

\begin{figure}
  \centering
    \includegraphics[scale=.4]{equiv-case.pdf}
    \caption{Diagram constructing a well-typed DNF lens between
      \ToDNFRegexOf{\Regex_2} and \ToDNFRegexOf{\RegexAlt_2}.}
    \label{fig:equiv-case}
\end{figure}

After this, we have $\DNFLens \OfRewritelessType \DNFRegex_1 \Leftrightarrow
\DNFRegexAlt_1$ and $\DNFRegex_1 \StarOf{\rightarrow} \DNFRegex_2$ and
$\DNFRegexAlt_1 \StarOf{\rightarrow} \DNFRegexAlt_2$. If we can prove a
confluence-like property that would show the existance of some $\DNFLens'
\OfRewritelessType \DNFRegex_3 \Leftrightarrow \DNFRegexAlt_3$, where
$\DNFRegex_2 \StarOf{\rightarrow} \DNFRegex_3$ and $\DNFRegexAlt_2
\StarOf{\rightarrow} \DNFRegexAlt_3$, we would be done. This property is
$R$-confluence for a properly chosen $R$. This case is diagrammed in Figure~\ref{fig:equiv-case}

\section{$R$-Confluence Formulation}
Let $S$ be an underlying set, and $\rightarrow$ and $R$ be binary relations. The
indexed rewrite system $(S,\rightarrow)$ is $R$-confluent, if for all $s_1,s_2 \in
S$, if $R(s_1,s_2)$, $s_1 \rightarrow s_1'$, and $s_2 \rightarrow s_2'$, then
there exist $s_1''$ and $s_2''$ such that $R(s_1'',s_2'')$, $s_1' \rightarrow
s_1''$, and $s_2 \rightarrow s_2''$.

In the context of lens synthesis, $S$ is the set of DNF REs,
the rewrites are $\StarOf{\rightarrow}$, and given a DNF lens $\DNFLens$,
$R_\DNFLens(\DNFRegex,\DNFRegexAlt)$ is true if there exists a DNF lens $\DNFLens'$ such
that $\DNFLens' \OfRewritelessType \DNFRegex \Leftrightarrow \DNFRegexAlt$ and
$\DNFLens$ is equivalent to $\DNFLens'$.

If we followed the standard way to prove a rewrite system confluent, we would
find a set of rewrite rules $\ParallelRewrite$ such that
$\StarOf{\ParallelRewrite} = \StarOf{\Rewrite}$ and $(S,\ParallelRewrite)$ is
$R_\DNFLens$-confluent. Then, we need to prove that
$(S,\StarOf{\ParallelRewrite})$ is $R_\DNFLens$-confluent from knowledge that
$(S,\ParallelRewrite)$ is $R_\DNFLens$-confluent. However, we will show this
approach does not work.

\section{Proving (S,\StarOf{\rightarrow}) $R$-Confluent}

\begin{figure}
  \centering
    \includegraphics[scale=.4]{proof-attempt.pdf}
    \caption{If trying to prove $R$-confluence for the transitive closure of a
      rewrite system in the standard way, we get stuck at this point in the
      inductive case.}
    \label{fig:proof-attempt}
\end{figure}

The standard way of proving $(S,\StarOf{\rightarrow})$ to be $=$-confluent is to
first prove $(S,\rightarrow)$ to be $=$-confluent, then to perform two
inductions, first on the left arrows then on the right arrows. This approach
does not work for arbitrary $R$-confluence. Consider the inductive case, where
$R(s_1,t_1)$ and $s_1 \StarOf{\rightarrow} s_2$ and $s_2 \rightarrow s_3$ and
$t_1 \StarOf{\rightarrow} t_2$, shown in Figure~\ref{fig:proof-attempt}. By
applying the inducion hypothesis, we know there exists $s_4, t_4$ such that $s_2
\rightarrow s_4$ and $t_2 \StarOf{\rightarrow} t_4$, and $R(s_4,t_4)$. We know
that $s_2 \rightarrow s_3$ and $s_2 \rightarrow s_4$, but we only know that
$(S,\rightarrow)$ is $R$-confluent, and so we are now stuck.

To get around this issue, we require a different set of properties.
\begin{enumerate}
\item $(S,\rightarrow)$ must be $=$-confluent.
\item $(S,\rightarrow)$ must be \emph{$R$-bisimilar}. In other words,
  if $R(s_1,t_1)$, and $s_1 \rightarrow s_2$, then there exists $t_2$ such that
  $t_1 \rightarrow t_2$ and $R(s_2,t_2)$; and if $R(s_1,t_1)$, and $t_1
  \rightarrow t_2$, then there exists $s_2$ such that $s_1 \rightarrow s_2$ and
  $R(s_2,t_2)$.
\end{enumerate}

Now we can proceed with our proof.

\begin{theorem}
  \label{lem:right-side}
  Let $(S,\rightarrow)$ be $=$-confluent, and $(S,\rightarrow)$ be
  $R$-bisimilar. If $R(s_1,t_1)$ and $s_1 \StarOf{\rightarrow} s_2$ and $t_1
  \StarOf{\rightarrow} t_2$, then there exists $s_3,t_3$ such that $s_2
  \StarOf{\rightarrow} s_3$ and $t_2 \rightarrow t_3$ and $R(s_3,t_3)$.
\end{theorem}
\begin{proof}
  By induction on $s_1 \StarOf{\rightarrow} s_2$
  \begin{case}[Base Case]
    Let $R(s_1,t_1)$ and $t_1 \StarOf{\rightarrow} t_2$. As
    $(S,\StarOf{\rightarrow})$ is $R$-bisimilar, there exists $s_2$ such that
    $s_1 \StarOf{\rightarrow} s_2$ and $R(s_2,t_2)$, as desired.
  \end{case}

  \begin{figure}
    \centering
    \includegraphics[scale=.4]{solved-proof.pdf}
    \caption{Diagram showing how we prove the inductive case of $R$-confluence
      of the transitive closure of a rewrite system. $R(s_4,t_4)$ comes from the
      inductive hypothesis. The existance of $s_5$ is guaranteed through
      $=$-confluence. $R(s_5,t_5)$ comes from $R$-bisimilarity.}
    \label{fig:solved-proof}
  \end{figure}

  \begin{case}[Inductive Case]
    Let $R(s_1,t_1)$ and $s_1 \StarOf{\rightarrow} s_2$ and $s_2 \rightarrow
    s_3$ and $t_1 \StarOf{\rightarrow} t_2$. By the induction hypothesis, there
    exists $s_4,t_4$ such that $R(s_4,t_4)$ and $s_3 \StarOf{\rightarrow} s_4$
    and $t_3 \rightarrow t_4$.

    As $(S,\rightarrow)$ is $=$-confluent, we know from prior work
    $(S,\StarOf{\rightarrow})$ is $=$-confluent. Because of this, there exists
    $s_5$ such that $s_3 \StarOf{\rightarrow} s_5$ and $s_4 \StarOf{\rightarrow}
    s_5$. As $(S,\StarOf{\rightarrow})$ is $R$-bisimilar, there exists $t_5$
    such that $t_4 \StarOf{\rightarrow} t_5$ and $R(s_5,t_5)$, as desired. This
    case is diagrammed in Figure~\ref{fig:solved-proof}.
  \end{case}
\end{proof}

%------------------------------------------------------------------------------

\label{sect:bib}
\bibliographystyle{plain}
%\bibliographystyle{alpha}
%\bibliographystyle{unsrt}
%\bibliographystyle{abbrv}
\bibliography{local}

%------------------------------------------------------------------------------

%------------------------------------------------------------------------------
% Index
%\printindex

%------------------------------------------------------------------------------
\end{document}

### 2020-04-07 papers/BijectiveLenses/IWC/main.tex
add more info

@@
 once'' using a bidirectional programming language. Bidirectional programming
-languages have been developed for creating view updaters~\cite{?}, Linux
-configuration file editors~\cite{?}, direct manipulation programming
-systems~\cite{?}, and more~\cite{?,?,?}. Lenses are a particularly well-behaved
-class of bidirectional programs, where the underlying transformations are
-guaranteed to satisfy a number of ``round-tripping'' laws. Lens-based
-bidirectional programming languages often provide round-tripping guarantees
-through a set of typing rules; well-typed lens expressions are guaranteed to
-satisfy the round-tripping laws.
+languages have been developed for creating view
+updaters~\cite{BohannonPierceVaughan}, Linux configuration file
+editors~\cite{augeas2}, direct manipulation programming
+systems~\cite{bidirectionaleval}, and
+more~\cite{DBLP:conf/pepm/KoZH16,DBLP:conf/icfp/HidakaHIKMN10,DBLP:conf/staf/ZhuK0SH15}. Lenses are a
+particularly well-behaved class of bidirectional programs, where the underlying
+transformations are guaranteed to satisfy a number of ``round-tripping'' laws.
+Lens-based bidirectional programming languages often provide round-tripping
+guarantees through a set of typing rules; well-typed lens expressions are
+guaranteed to satisfy the round-tripping laws.
 
-To make bidirectional programming easier, Optician~\cite{?,?,?} synthesizes
+To make bidirectional programming easier, Optician~\cite{optician} synthesizes
 bidirectional programs from these types, specifically in Boomerang, a language
@@
 %\bibliographystyle{abbrv}
-\bibliography{local}
+\bibliography{local,bcp}
 

### 2020-04-09 papers/BijectiveLenses/IWC/main.tex
minor edits

@@
 in type equivalence we use a finer notion of equivalence than language
-equivalence, we instead use star-semiring equivalence. We were unable to make
+equivalence, we instead use star-semiring equivalence, denoted $\SSREquiv$. We were unable to make
 our proof strategy work for language equivalence, so instead guarantee
@@
 Our solution is to synthesize in an alternative language of disjunctive normal
-form lenses (DNF lenses), typed by disjunctive number form regular expressions
+form lenses (DNF lenses), typed by disjunctive normal form regular expressions
 (DNF REs). Both DNF lenses and DNF REs can be thought of as lenses and regular
@@
 To type a DNF lens, one first applies a number of rewrites to the DNF REs, then
-builds up a DNF lens using rewriteless typing rules. Because of this, equivalent
+builds up a DNF lens using rewriteless typing rules. Because of this
+phase distinction, equivalent
 types only need to be generated once, and composition never need be applied.
-This permits an efficient algorithm that first searchest through type
-equivalences with rewrites, and second builds up a rewriteless DNF lens purely
-through the syntax of the DNF RE. This then permits the full synthesis
-algorithm, shown in Figure~\ref{fig:schematic-diagram-synthesis}. The input
-regular expressions are converted into DNF form with $\ToDNFRegex$. Next, a
-well-typed DNF lens is synthesized, and then converted back into a standard lens
-language with $\ToLens$.
+This structure corresponds to an efficient algorithm that first searches through type
+equivalences with rewrites and then builds up a rewriteless DNF
+lens as directed by the syntax of the DNF RE. 
+Figure~\ref{fig:schematic-diagram-synthesis} shows the full synthesis
+algorithm. The input
+regular expressions are converted into DNF form using the
+$\ToDNFRegex$ function. Next, a
+well-typed DNF lens is synthesized and then converted back into a standard lens
+language with the $\ToLens$ function.
 
@@
   $\DNFLens$, such that $\DNFLens \OfType \ToDNFRegexOf{\Regex} \Leftrightarrow
-  \ToDNFRegexOf{\RegexAlt}$, such that the semantics of $\Lens$ and $\DNFLens$
+  \ToDNFRegexOf{\RegexAlt}$ and the semantics of $\Lens$ and $\DNFLens$
   are equivalent.
@@
 
-We prove this by induction, and particular difficulty lies in the lens
+We prove this propety by induction. The particular difficulty lies in the lens
 equivalence rule. We begin this case below:
@@
 $t_1 \StarOf{\rightarrow} t_2$, shown in Figure~\ref{fig:proof-attempt}. By
-applying the inducion hypothesis, we know there exists $s_4, t_4$ such that $s_2
+applying the induction hypothesis, we know there exists $s_4, t_4$ such that $s_2
 \rightarrow s_4$ and $t_2 \StarOf{\rightarrow} t_4$, and $R(s_4,t_4)$. We know

### 2020-04-17 papers/BijectiveLenses/IWC/main.tex
update abstract

@@
 \begin{abstract}
-  Bidirectional programs are programs where a single expression has multiple
-  interpretations, encoding a number of functions with a single piece of syntax.
-  Lenses are bidirectional programs with semantic guarantees, and lens languages
-  have been developed for synchronizing string formats, building updatable views
-  in databases, and more. These languages often are typed, where a well-typed
-  expression will provide the semantic guarantees. Recent research has shown how
-  string lenses can be synthesized from their types, pairs of regular
-  expressions. However, providing completeness guarantees is quite tricky on
-  these languages, due in large part to the many equivalences on regular
-  expressions.
+  A lens is a program that can be executed both forwards and backwards, from
+  input to output and from output back to input again. Domain-specific languages
+  for defining lenses have been developed to help users synchronize text files,
+  and construct different ``views'' of databases, among other applications.
+  Recent research has shown how string lenses can be synthesized from their
+  types, which are pairs of regular expressions. However, guaranteeing that we
+  can synthesize all possible lenses is quite tricky on these languages, due in
+  large part to the many equivalences on regular expressions.
 
-  In proving a completeness guarantee, a confluence-like property,
-  $R$-confluence, must be shown. This property can be thought of as a
-  generalization of standard confluence. Where standard confluence is
-  parameterized by a set $S$ and a binary relation $\rightarrow$, indexed
-  confluence is parameterized by an additional binary relation $R \subseteq S
-  \times S$. In this model, confluence is the specific case where $R$ is just
-  equality. We show how the proof that the transitive closure of a rewrite
-  system is $R$-confluent differs from the proof strategy for standard
-  confluence.
+  Our proof that we can synthesize all string lenses involves proving a
+  confluence-like property, indexed by a an additional binary relation $R$. In
+  this model, standard confluence is the specific case where $R$ is equality. We
+  show how our proof that the transitive closure of a rewrite system is
+  $R$-confluent differs from the proof strategy for standard confluence.
 \end{abstract}
@@
 
-To make bidirectional programming easier, Optician~\cite{optician} synthesizes
-bidirectional programs from these types, specifically in Boomerang, a language
-for synchronizing string data. Optician takes as input two regular expressions
-(which serve as the type of a Boomerang lens) and a set of examples specifying
+Optician~\cite{optician}, an extension of Boomerang, makes bidirectional
+programming easier by supporting synthesis of bidirectional string
+transformations. More specifically, it takes as input two regular expressions
+($\Regex$ and $\RegexAlt$, which serve as the type of a Boomerang lens) and a set of examples specifying
 input-output behavior, and synthesizes a well-typed lens between the languages
@@
 
-The judgement $\Lens : \Regex \Leftrightarrow \RegexAlt$ means that $\Lens$
-bidirectionally converts between $\LanguageOf{\Regex}$ and
-$\LanguageOf{\RegexAlt}$. For synthesis, given $\Regex$ and $\RegexAlt$, we must
+To indicate a lens $\Lens$ is well typed, and converts between the languages of
+$\Regex$ and $\RegexAlt$, we write $\Lens : \Regex \Leftrightarrow \RegexAlt$ For synthesis, given $\Regex$ and $\RegexAlt$, we must
 find a lens $\Lens$ such that $\Lens : \Regex \Leftrightarrow \RegexAlt$. There

### 2020-04-21 papers/BijectiveLenses/IWC/main.tex
add main and citations

@@
   Our proof that we can synthesize all string lenses involves proving a
-  confluence-like property, indexed by a an additional binary relation $R$. In
+  confluence-like property, indexed by a an additional binary relation $R$. We
+  call this property $R$-confluence. In
   this model, standard confluence is the specific case where $R$ is equality. We
@@
 To indicate a lens $\Lens$ is well typed, and converts between the languages of
-$\Regex$ and $\RegexAlt$, we write $\Lens : \Regex \Leftrightarrow \RegexAlt$ For synthesis, given $\Regex$ and $\RegexAlt$, we must
-find a lens $\Lens$ such that $\Lens : \Regex \Leftrightarrow \RegexAlt$. There
-are a number of typing rules for lenses, but the rules for type-equivalence and
-composition (shown below) do not permit an efficient search algorithm. Note that
-in type equivalence we use a finer notion of equivalence than language
-equivalence, we instead use star-semiring equivalence, denoted $\SSREquiv$. We were unable to make
-our proof strategy work for language equivalence, so instead guarantee
-completeness only on the subset of lenses well-typed with star-semiring
-equivalence.
+$\Regex$ and $\RegexAlt$, we write $\Lens : \Regex \Leftrightarrow \RegexAlt$
+For synthesis, given $\Regex$ and $\RegexAlt$, we must find a lens $\Lens$ such
+that $\Lens : \Regex \Leftrightarrow \RegexAlt$. In the context of lens
+synthesis, there are three sorts of lens typing rules: syntax-directed rules,
+composition, and type equivalence.
+
+For syntax-directed rules, the syntax for the types closely mirrors the syntax
+of the expressions. As an example, consider the following rule for
+disjunction:
+\begin{mathpar}
+  \inferrule[\OrLensRule{}]
+  {
+    \Lens_1 \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
+    \Lens_2 \OfType \Regex_2 \Leftrightarrow \RegexAlt_2\\\\
+    \UnambigOrOf{\Regex_1}{\Regex_2}\\
+    \UnambigOrOf{\RegexAlt_1}{\RegexAlt_2}
+  }
+  {
+    \OrLensOf{\Lens_1}{\Lens_2} \OfType
+    \Regex_1 \Or \Regex_2
+    \Leftrightarrow \RegexAlt_1 \Or \RegexAlt_2
+  }
+\end{mathpar}
+Consider finding a lens of type $\Regex_1 \Or \Regex_2 \Leftrightarrow
+\RegexAlt_1 \Or \RegexAlt_2$. With only syntax-directed rules, the only lens
+that can be well-typed would be an or lens.
+
+Composition sequentially composes two lenses.
+\begin{mathpar}
+  \inferrule[\ComposeLensRule{}]
+  {
+    \Lens_1 \OfType \Regex_1 \Leftrightarrow \Regex_2\\
+    \Lens_2 \OfType \Regex_2 \Leftrightarrow \Regex_3\\
+  }
+  {
+    \ComposeLensOf{\Lens_1}{\Lens_2} \OfType \Regex_1 \Leftrightarrow \Regex_3
+  }
+\end{mathpar}
+Composition is difficult in the context of lens synthesis. If trying to
+synthesize a composition lens, one has to pull the central regular expression
+``out of thin air.''
+
+The last typing rule is type-equivalence. If two regular expressions are
+star-semiring equivalent to the type of a lens, those equivalent regular
+expressions also serve as the type of the lens.
 \begin{mathpar}
@@
   }
+\end{mathpar}
+This rule is difficult in the context of synthesis, as it forces a search
+through equivalent regular expressions.  This rule can also be applied at any
+point in the derivation, which makes the search even harder.
 
-  \inferrule[\ComposeLensRule{}]
-  {
-    \Lens_1 \OfType \Regex_1 \Leftrightarrow \Regex_2\\
-    \Lens_2 \OfType \Regex_2 \Leftrightarrow \Regex_3\\
-  }
-  {
-    \ComposeLensOf{\Lens_1}{\Lens_2} \OfType \Regex_1 \Leftrightarrow \Regex_3
-  }
-\end{mathpar}
+%\begin{figure}
+%  \centering
+%  \begin{tikzpicture}[auto,node distance=1.5cm]
+%    \node[text width=1.5cm,minimum height=.6cm,align=center,draw,rectangle] (todnfregex) {\ToDNFRegex{}};
+%    
+%    \node[align=right, anchor=east] (regex1) [left = .6cm of todnfregex.north west]{\Regex{}};
+%    \node[align=right, anchor=east] (regex2) [left = .6cm of todnfregex.south west]{\RegexAlt{}};
+%    \node[align=right, anchor=east] (exs) [below = .2cm of regex2]{ \Examples{} };
+%    
+%    \node[align=center] (dnfregex1) [right = .4cm of todnfregex.north east]{\DNFRegex{}};
+%    \node[align=center] (dnfregex2) [right = .4cm of todnfregex.south east]{\DNFRegexAlt{}};
+%
+%    \node[text width=2.9cm,minimum height=.6cm,align=center,draw,rectangle] [right = 1.45cm of todnfregex.east] (synthdnflens) {\SynthDNFLens{}};
+%    \node[align=center] [above = .7cm of synthdnflens] (optician) {\Optician{}};
+%    
+%    \node[align=center] [right = .4cm of synthdnflens] (dnflens) {\DNFLens{}};
+%    
+%    \node[text width=1.5cm,minimum height=.6cm,align=center,draw,rectangle] [right = .4cm of dnflens] (tolens) {\ToLens{}};
+%    
+%    \node[align=center] [right = .6cm of tolens] (lens) {\Lens{}};
+%    
+%    
+%    \path[->] (regex1.east) edge (todnfregex.north west);
+%    \path[->] (regex2.east) edge (todnfregex.south west);
+%    
+%    \path[->] (todnfregex.north east) edge (dnfregex1.west);
+%    \path[->] (todnfregex.south east) edge (dnfregex2.west);
+%    
+%    \path[->] (dnfregex1.east) edge (synthdnflens.north west);
+%    \path[->] (dnfregex2.east) edge (synthdnflens.south west);
+%    
+%    \path[->] (synthdnflens) edge (dnflens);
+%    
+%    \path[->] (dnflens) edge (tolens);
+%    
+%    \path[->] (tolens) edge (lens);
+%    \draw[->] ($(exs.east)+(-3pt,0)$) -| node(exsedge) {} (synthdnflens);
+%    \node[fit={($(todnfregex.west)+(-4pt,0)$) ($(tolens.east)+(4pt,0)$) (exsedge) (dnfregex1) (optician) (dnfregex2)},draw] (surrounding) {};
+%    % Now place a relation (ID=rel1)
+%    %\node[text width=2cm,align=center,draw, rectangle] (sketch-gen) [right = .75cm of spec] {\TypeProp{}};
+%    %\node (below-gen) [below=.5cm of sketch-gen] {};
+%    %\node[text width=2cm,align=center,draw, rectangle] (sketch-compl)
+%    %     [right = .25cm of sketch-gen] {\RigidSynth{}};
+%    %\node (below-compl) [below=.5cm of sketch-compl] {};
+%    %\node[align=center] (lens) [right = .75cm of sketch-compl] {Lens}; 
+%    %% Draw an edge between rel1 and node1; rel1 and node2
+%    %\path[->] (spec) edge node (start-alg) {} (sketch-gen);
+%    %\path[->] (sketch-gen) edge node(middle) {} (sketch-compl);
+%    %\path[->] (sketch-compl) edge node[near start](success) {\Success{}} (lens);
+%    %\draw[<-] (sketch-gen.south) -- +(0,-.5) -| node[above left](failure){\Failure{}} (sketch-compl.south);
+%
+%    %\node (synth-name) [above=.5cm of middle] {\Optician{}};
+%    %
+%    %\node[fit=(sketch-gen) (sketch-compl) (start-alg) (synth-name) (failure) (success) ,draw] (surrounding) {};
+%  \end{tikzpicture}
+%  \caption{Schematic Diagram for \Optician{}.  Regular expressions, \Regex{} and
+%    \RegexAlt{}, and examples, \Examples{}, are given as input.
+%    First, the function \ToDNFRegex{} converts \Regex{} and \RegexAlt{} into
+%    their respective DNF forms, \DNFRegex{} and \DNFRegexAlt{}.
+%    Next, \SynthDNFLens{} synthesizes a DNF lens, \DNFLens{}, from \Regex{},
+%    \RegexAlt{}, and \Examples{}.
+%    Finally, \ToLens{} converts \DNFLens{} into \Lens{}, a lens in Boomerang
+%    that is equivalent to \DNFLens{}.}
+%  \label{fig:schematic-diagram-synthesis}
+%\end{figure}
 
-Trying to synthesize a lens from the types by applying typing rules ``in
-reverse'' gets confounded by these two rules. They require no significant
-restriction on the output types, and so can be applied at any time. Furthermore,
-applying them in reverse requires pulling regular expressions ``out of thin
-air.'' For \RewriteRegexLensRule, these regular expressions need only be
-equivalent to the input regular expressions, and for \ComposeLensRule, there are
-no restrictions at all!
+To address the difficulties with composition and type-equivalence rules, we
+synthesize lenses in an alternative language of disjunctive normal form (DNF)
+lenses. DNF lenses are in pseudonormal form, containing no composition operator,
+and so their synthesis never needs to pull regular expressions ``out of thin
+air.'' The types of DNF lenses are pairs of regular expressions in a
+pseudonormal form, DNF regular expressions. Because DNF regular expressions are
+in a pseudonormal form, fewer equivalent regular expressions need to be searched
+through.
 
-Our solution is to synthesize in an alternative language of disjunctive normal
-form lenses (DNF lenses), typed by disjunctive normal form regular expressions
-(DNF REs). Both DNF lenses and DNF REs can be thought of as lenses and regular
-expressions, put into a pseudonormal form.
-
-\begin{figure}
-  \centering
-  \begin{tikzpicture}[auto,node distance=1.5cm]
-    \node[text width=1.5cm,minimum height=.6cm,align=center,draw,rectangle] (todnfregex) {\ToDNFRegex{}};
-
-    \node[align=right, anchor=east] (regex1) [left = .6cm of todnfregex.north west]{\Regex{}};
-    \node[align=right, anchor=east] (regex2) [left = .6cm of todnfregex.south west]{\RegexAlt{}};
-    \node[align=right, anchor=east] (exs) [below = .2cm of regex2]{ \Examples{} };
-
-    \node[align=center] (dnfregex1) [right = .4cm of todnfregex.north east]{\DNFRegex{}};
-    \node[align=center] (dnfregex2) [right = .4cm of todnfregex.south east]{\DNFRegexAlt{}};
-
-    \node[text width=2.9cm,minimum height=.6cm,align=center,draw,rectangle] [right = 1.45cm of todnfregex.east] (synthdnflens) {\SynthDNFLens{}};
-    \node[align=center] [above = .7cm of synthdnflens] (optician) {\Optician{}};
-
-    \node[align=center] [right = .4cm of synthdnflens] (dnflens) {\DNFLens{}};
-
-    \node[text width=1.5cm,minimum height=.6cm,align=center,draw,rectangle] [right = .4cm of dnflens] (tolens) {\ToLens{}};
-
-    \node[align=center] [right = .6cm of tolens] (lens) {\Lens{}};
-
-    \path[->] (regex1.east) edge (todnfregex.north west);
-    \path[->] (regex2.east) edge (todnfregex.south west);
-
-    \path[->] (todnfregex.north east) edge (dnfregex1.west);
-    \path[->] (todnfregex.south east) edge (dnfregex2.west);
-
-    \path[->] (dnfregex1.east) edge (synthdnflens.north west);
-    \path[->] (dnfregex2.east) edge (synthdnflens.south west);
-
-    \path[->] (synthdnflens) edge (dnflens);
-
-    \path[->] (dnflens) edge (tolens);
-
-    \path[->] (tolens) edge (lens);
-    \draw[->] ($(exs.east)+(-3pt,0)$) -| node(exsedge) {} (synthdnflens);
-    \node[fit={($(todnfregex.west)+(-4pt,0)$) ($(tolens.east)+(4pt,0)$) (exsedge) (dnfregex1) (optician) (dnfregex2)},draw] (surrounding) {};
-    % Now place a relation (ID=rel1)
-    %\node[text width=2cm,align=center,draw, rectangle] (sketch-gen) [right = .75cm of spec] {\TypeProp{}};
-    %\node (below-gen) [below=.5cm of sketch-gen] {};
-    %\node[text width=2cm,align=center,draw, rectangle] (sketch-compl)
-    %     [right = .25cm of sketch-gen] {\RigidSynth{}};
-    %\node (below-compl) [below=.5cm of sketch-compl] {};
-    %\node[align=center] (lens) [right = .75cm of sketch-compl] {Lens}; 
-    %% Draw an edge between rel1 and node1; rel1 and node2
-    %\path[->] (spec) edge node (start-alg) {} (sketch-gen);
-    %\path[->] (sketch-gen) edge node(middle) {} (sketch-compl);
-    %\path[->] (sketch-compl) edge node[near start](success) {\Success{}} (lens);
-    %\draw[<-] (sketch-gen.south) -- +(0,-.5) -| node[above left](failure){\Failure{}} (sketch-compl.south);
-
-    %\node (synth-name) [above=.5cm of middle] {\Optician{}};
-    %
-    %\node[fit=(sketch-gen) (sketch-compl) (start-alg) (synth-name) (failure) (success) ,draw] (surrounding) {};
-  \end{tikzpicture}
-  \caption{Schematic Diagram for \Optician{}.  Regular expressions, \Regex{} and
-    \RegexAlt{}, and examples, \Examples{}, are given as input.
-    First, the function \ToDNFRegex{} converts \Regex{} and \RegexAlt{} into
-    their respective DNF forms, \DNFRegex{} and \DNFRegexAlt{}.
-    Next, \SynthDNFLens{} synthesizes a DNF lens, \DNFLens{}, from \Regex{},
-    \RegexAlt{}, and \Examples{}.
-    Finally, \ToLens{} converts \DNFLens{} into \Lens{}, a lens in Boomerang
-    that is equivalent to \DNFLens{}.}
-  \label{fig:schematic-diagram-synthesis}
-\end{figure}
-
-Two typing judgements are involved in typing a DNF lens, $\DNFLens : \DNFRegex
-\Leftrightarrow \DNFRegexAlt$ and $\DNFLens \OfRewritelessType \DNFRegex
-\Leftrightarrow \DNFRegexAlt$. In both instances, the judgement means $\DNFLens$
-is a DNF lens that converts between $\LanguageOf{\DNFRegex}$ and
-$\LanguageOf{\DNFRegexAlt}$. If $\DNFLens \OfRewritelessType \DNFRegex
-\Leftrightarrow \DNFRegexAlt$, then the typing derivation of $\DNFLens$ never
-rewrites the DNF REs. The only typing rule for $\DNFLens : \DNFRegex
-\Leftrightarrow \DNFRegexAlt$ is:
+In our search algorithm, we only search through equivalent regular expressions
+once, before processing any syntax directed-rules. We formalize this in the
+typing of DNF regular expressions by only permitting the application of
+type-equivalence once, after all syntactic rules have been applied. This is
+enforced by having two typing judgements: one for the ``rewriteless'' type of
+the lens (meaning no type-equivalence rules were applied) and one of the
+``full'' type of the lens. If $\DNFLens \OfRewritelessType \DNFRegex
+\Leftrightarrow \DNFRegexAlt$, then $\DNFLens$ is a DNF lens of rewriteless type
+$\DNFRegex \Leftrightarrow \DNFRegexAlt$. If $\DNFLens \OfType \DNFRegex
+\Leftrightarrow \DNFRegexAlt$, then $\DNFLens$ is a DNF lens of full type
+$\DNFRegex \Leftrightarrow \DNFRegexAlt$. The following rule is used to get the
+full type of a DNF lens from the rewriteless type.
 \begin{mathpar}
@@
 \end{mathpar}
-To type a DNF lens, one first applies a number of rewrites to the DNF REs, then
-builds up a DNF lens using rewriteless typing rules. Because of this
-phase distinction, equivalent
-types only need to be generated once, and composition never need be applied.
-This structure corresponds to an efficient algorithm that first searches through type
-equivalences with rewrites and then builds up a rewriteless DNF
-lens as directed by the syntax of the DNF RE. 
-Figure~\ref{fig:schematic-diagram-synthesis} shows the full synthesis
-algorithm. The input
-regular expressions are converted into DNF form using the
-$\ToDNFRegex$ function. Next, a
-well-typed DNF lens is synthesized and then converted back into a standard lens
-language with the $\ToLens$ function.
+Note that instead of using directionless equivalences,
+\RewriteDNFRegexLensRule{} uses directioned rewrites. The relationship between
+the two is formalized by the following theorem:
+\begin{theorem}
+  \label{thm:thm1}
+  If $\Regex \SSREquiv \RegexAlt$, then $\ToDNFRegexOf{\Regex}
+  \equiv_{\rightarrow} \ToDNFRegexOf{\RegexAlt}$, where $\ToDNFRegexOf{\Regex}$
+  and $\ToDNFRegexOf{\RegexAlt}$ are $\Regex$ and $\RegexAlt$ in DNF form
+  (respectively), and $\equiv_{\rightarrow}$ is the transitive, reflexive, and
+  symmetric closure of $\rightarrow$.
+\end{theorem}
 
-Proving this search procedure complete reduces to proving DNF lenses complete
-with respect to our standard lens language. In particular, we wish the prove the
-following theorem:
+Proving that our search procedure can generate any lens reduces to proving DNF
+lenses complete with respect to our standard lens language. In particular, we
+wish the prove the following theorem:
 \begin{theorem}
@@
 
-We prove this propety by induction. The particular difficulty lies in the lens
-equivalence rule. We begin this case below:
+We prove this propety by induction on the structure of the typing derivation.
+The particular difficulty lies in the lens equivalence rule. We begin this case
+below:
 
@@
 \Leftrightarrow \ToDNFRegexOf{\RegexAlt_1}$, where the semantics of $\DNFLens$
-are equivalent to those of $\Lens$. By inversion
+are equivalent to those of $\Lens$. By inversion on the derivation of $\DNFLens
+\OfType \ToDNFRegexOf{\Regex_1} \Leftrightarrow \ToDNFRegexOf{\RegexAlt_1}$,
+there exists $\DNFRegex_1$ and $\DNFRegexAlt_1$ such that:
 \begin{mathpar}
@@
 $\ToDNFRegexOf{\DNFRegexAlt_1} \StarOf{\rightarrow} \DNFRegexAlt_2$. To do this,
-we first prove that $\StarOf{\rightarrow}$ is confluent, and that if $\Regex
-\SSREquiv \RegexAlt$, then $\ToDNFRegexOf{\Regex}$ is related to
-$\ToDNFRegexOf{\RegexAlt}$ through the transitive, reflexive closure of
-$\rightarrow$. We then apply a corollary of confluence to show the existance of
-such a $\DNFRegex_2$ and $\DNFRegexAlt_2$
+we first prove that $\StarOf{\rightarrow}$ is confluent. By
+Theorem~\ref{thm:thm1}, we know that $\ToDNFRegexOf{\Regex_1} \equiv_\rightarrow
+\ToDNFRegexOf{\Regex_2}$. Such $\DNFRegex_2$ and $\DNFRegexAlt_2$ are guaranteed
+to exist by a corollary to confluence.
 
@@
 \StarOf{\rightarrow} \DNFRegexAlt_3$, we would be done. This property is
-$R$-confluence for a properly chosen $R$. This case is diagrammed in Figure~\ref{fig:equiv-case}
+$R$-confluence for a properly chosen $R$. This case is diagrammed in
+Figure~\ref{fig:equiv-case}.
 
@@
 
+\section{Related Work}
+
+The full proof of completeness is contained in the appendix of the full version
+of the original optician paper~\cite{extended-version}. The original proof of
+$R$-confluence for the transitive closure of $\rightarrow$ previously required
+additional assumptions. These unnecessary assumptions have been identified and
+removed in this version of the proof. Future work used the proof of completeness
+over our lens language to show that quotient bijective lenses are also
+synthesizeable~\cite{maina+:quotient-synthesis}. Lastly, while synthesizeability
+was not proven for symmetric lenses~\cite{soptician}, such a proof would likely
+have proven $R$-confluence in a similar manner. Our proof used approaches
+inspired by the original Church-Rosser paper~\cite{Church1936SomePO}.
+
+This work continues a trend in making programming easier through
+synthesis~\cite{gulwani2017program}. While synthesis is one approach to make
+bidirectional programming easier, it is not the only approach. Recent work has
+gone into building lenses without requiring the point-free combinator
+style~\cite{10.1007/978-3-319-89884-1_2}. Other work has found
+applicative~\cite{10.1145/2858949.2784750} and
+monadic~\cite{10.1007/978-3-030-17184-1_6} approaches to compositionally
+building lenses.
+
 %------------------------------------------------------------------------------

### 2020-06-11 papers/BijectiveLenses/IWC/main.tex
update iwc

@@
 
-  Our proof that we can synthesize all string lenses involves proving a
-  confluence-like property, indexed by a an additional binary relation $R$. We
-  call this property $R$-confluence. In
-  this model, standard confluence is the specific case where $R$ is equality. We
-  show how our proof that the transitive closure of a rewrite system is
-  $R$-confluent differs from the proof strategy for standard confluence.
+  The proof that all string lenses are synthesizeable involves proving a
+  confluence-like property, parameterized by a an additional binary relation
+  $R$. We call this property $R$-confluence. In this model, standard confluence
+  is the specific case where $R$ is equality. In this paper, we show how
+  existing techniques for demonstrating confluence do not work in the domain of
+  $R$-confluence, and find that if the rewrite system is $=$-confluent and
+  satisfies a commutativity property with $R$, then the system is $R$-confluent.
 \end{abstract}
@@
 transformations. More specifically, it takes as input two regular expressions
-($\Regex$ and $\RegexAlt$, which serve as the type of a Boomerang lens) and a set of examples specifying
-input-output behavior, and synthesizes a well-typed lens between the languages
-of those regular expressions.
+($\Regex$ and $\RegexAlt$, which serve as the type of a Boomerang lens) and a
+set of examples specifying input-output behavior, and synthesizes a well-typed
+lens between the languages of those regular expressions. For brevity, we will
+not provide formal definitions for some aspects of Optician; the interested
+reader can find such definitions in the original Optician paper~\cite{optician}.
+Furthermore, examples detailing the use of synthesized lenses in practice (which
+we also elide for space) can be found in that paper and follow-up
+work~\cite{maina+:quotient-synthesis,soptician}.
 
 To indicate a lens $\Lens$ is well typed, and converts between the languages of
-$\Regex$ and $\RegexAlt$, we write $\Lens : \Regex \Leftrightarrow \RegexAlt$
+$\Regex$ and $\RegexAlt$, we write $\Lens : \Regex \Leftrightarrow \RegexAlt$.
 For synthesis, given $\Regex$ and $\RegexAlt$, we must find a lens $\Lens$ such
@@
   and $\ToDNFRegexOf{\RegexAlt}$ are $\Regex$ and $\RegexAlt$ in DNF form
-  (respectively), and $\equiv_{\rightarrow}$ is the transitive, reflexive, and
+  (respectively), and $\equiv_{\rightarrow}$ is the reflexive, transitive, and
   symmetric closure of $\rightarrow$.
@@
 lenses complete with respect to our standard lens language. In particular, we
-wish the prove the following theorem:
+wish the prove:
 \begin{theorem}
@@
 We prove this propety by induction on the structure of the typing derivation.
-The particular difficulty lies in the lens equivalence rule. We begin this case
+Particular difficulty lies in the lens equivalence rule. We begin this case
 below:
@@
     \Lens \OfType \Regex_1 \Leftrightarrow \RegexAlt_1\\
-    \Regex \SSREquiv \Regex_2\\
-    \RegexAlt \SSREquiv \RegexAlt_2
+    \Regex_1 \SSREquiv \Regex_2\\
+    \RegexAlt_1 \SSREquiv \RegexAlt_2
   }
@@
 
-We first show that there exists $\DNFRegex_2$ and $\DNFRegexAlt_2$ such that
+We first show that there exist $\DNFRegex_2$ and $\DNFRegexAlt_2$ such that
 $\ToDNFRegexOf{\Regex_2} \StarOf{\rightarrow} \DNFRegex_2$ and
@@
 $\ToDNFRegexOf{\DNFRegexAlt_1} \StarOf{\rightarrow} \DNFRegexAlt_2$. To do this,
-we first prove that $\StarOf{\rightarrow}$ is confluent. By
+we first prove that $\rightarrow$ is confluent. By
 Theorem~\ref{thm:thm1}, we know that $\ToDNFRegexOf{\Regex_1} \equiv_\rightarrow
-\ToDNFRegexOf{\Regex_2}$. Such $\DNFRegex_2$ and $\DNFRegexAlt_2$ are guaranteed
-to exist by a corollary to confluence.
+\ToDNFRegexOf{\Regex_2}$, so confluence implies the existance of such a $\DNFRegex_2$ and $\DNFRegexAlt_2$.
 
@@
     \caption{Diagram constructing a well-typed DNF lens between
-      \ToDNFRegexOf{\Regex_2} and \ToDNFRegexOf{\RegexAlt_2}.}
+      \ToDNFRegexOf{\Regex_2} and \ToDNFRegexOf{\RegexAlt_2}. Single lined
+      arrows indicate rewrites ($\rightarrow$), and triple lines indicate
+      equivalence ($\equiv_{\rightarrow}$).}
     \label{fig:equiv-case}
@@
 \StarOf{\rightarrow} \DNFRegexAlt_3$, we would be done. This property is
-$R$-confluence for a properly chosen $R$. This case is diagrammed in
-Figure~\ref{fig:equiv-case}.
+$R$-confluence for a properly chosen $R$ (which we describe in \S\ref{sec:rcf}).
+This case is diagrammed in Figure~\ref{fig:equiv-case}.
 
 \section{$R$-Confluence Formulation}
-Let $S$ be an underlying set, and $\rightarrow$ and $R$ be binary relations. The
-indexed rewrite system $(S,\rightarrow)$ is $R$-confluent, if for all $s_1,s_2 \in
-S$, if $R(s_1,s_2)$, $s_1 \rightarrow s_1'$, and $s_2 \rightarrow s_2'$, then
-there exist $s_1''$ and $s_2''$ such that $R(s_1'',s_2'')$, $s_1' \rightarrow
-s_1''$, and $s_2 \rightarrow s_2''$.
+\label{sec:rcf}
+Let $S$ be an underlying set, and $\rightarrow$ and $R$ be
+binary relations. The rewrite system $(S,\rightarrow)$ is $R$-confluent, if for
+all $s_1,s_2 \in S$, if $R(s_1,s_2)$, $s_1 \StarOf{\rightarrow} s_1'$, and $s_2
+\StarOf{\rightarrow} s_2'$, then there exist $s_1''$ and $s_2''$ such that
+$R(s_1'',s_2'')$, $s_1' \StarOf{\rightarrow} s_1''$, and $s_2
+\StarOf{\rightarrow} s_2''$.
 
@@
 
-If we followed the standard way to prove a rewrite system confluent, we would
-find a set of rewrite rules $\ParallelRewrite$ such that
-$\StarOf{\ParallelRewrite} = \StarOf{\Rewrite}$ and $(S,\ParallelRewrite)$ is
-$R_\DNFLens$-confluent. Then, we need to prove that
-$(S,\StarOf{\ParallelRewrite})$ is $R_\DNFLens$-confluent from knowledge that
-$(S,\ParallelRewrite)$ is $R_\DNFLens$-confluent. However, we will show this
-approach does not work.
+Now that $R$-confluence has been formally defined, we can ask ourselves: ``What
+is a good approach to proving $R$-confluence?'' One approach is to prove that
+our rewrite system is locally confluent, which is equivalent to $=$-confluence
+in a terminating system~\cite{huetconf}. Unfortunately, our rewrites are not
+terminating, so this approach does not work.
+
+An approach pioneered by Tait and Martin-L\"{o}f~\cite{barendrecht} still works
+in non-terminating systems. This approach uses the \emph{diamond property}: A
+rewrite system $(S,\rightarrow)$ satisfies the diamond property if $s_1
+\rightarrow s_2$ and $s_1 \rightarrow s_3$ implies that there exists $s_4$ such
+that $s_2 \rightarrow s_4$, and $s_2 \rightarrow s_4$. If a rewrite system
+satisfies the diamond property, then it is also confluent. Unfortunately, this
+approach does not work, as the parameterized version of the diamond property
+does not imply $R$-confluence.
 
@@
   \centering
-    \includegraphics[scale=.4]{proof-attempt.pdf}
-    \caption{If trying to prove $R$-confluence for the transitive closure of a
-      rewrite system in the standard way, we get stuck at this point in the
-      inductive case.}
-    \label{fig:proof-attempt}
+    \includegraphics[scale=.4]{cex.pdf}
+    \caption{Rewrite system that satisfies the $R$-diamond property, but is not
+      $R$-confluent.}
+    \label{fig:cex}
 \end{figure}
 
-The standard way of proving $(S,\StarOf{\rightarrow})$ to be $=$-confluent is to
-first prove $(S,\rightarrow)$ to be $=$-confluent, then to perform two
-inductions, first on the left arrows then on the right arrows. This approach
-does not work for arbitrary $R$-confluence. Consider the inductive case, where
-$R(s_1,t_1)$ and $s_1 \StarOf{\rightarrow} s_2$ and $s_2 \rightarrow s_3$ and
-$t_1 \StarOf{\rightarrow} t_2$, shown in Figure~\ref{fig:proof-attempt}. By
-applying the induction hypothesis, we know there exists $s_4, t_4$ such that $s_2
-\rightarrow s_4$ and $t_2 \StarOf{\rightarrow} t_4$, and $R(s_4,t_4)$. We know
-that $s_2 \rightarrow s_3$ and $s_2 \rightarrow s_4$, but we only know that
-$(S,\rightarrow)$ is $R$-confluent, and so we are now stuck.
+In the Tait and Martin-L\"{o}f approach to proving $(S,\rightarrow)$ confluent,
+one must first prove $(S,\rightarrow)$ satisfies the diamond property. Consider
+a parameterized property analogous to the diamond property, the
+\emph{$R$-diamond property}: A rewrite system $(S,\rightarrow)$ satisfies the
+$R$-diamond property if $s_1 \rightarrow s_2$ and $t_1 \rightarrow t_2$ and
+$R(s_1,t_1)$ implies that there exists $s_3,t_3$ such that $s_2 \rightarrow s_3$
+and $t_2 \rightarrow t_3$ and $R(s_3,t_3)$.
+
+However, satisfying the $R$-diamond property is not sufficient for
+$R$-confluence. Consider the simple rewrite system shown in
+Figure~\ref{fig:cex}. In this rewrite system, there are 3 elements, $s_1,s_2$,
+and $t$. In this setup, $R(s_1,t)$ and $s_1 \rightarrow s_2$. $R$-confluence
+requires some $s_3,t'$ such that $s_2 \StarOf{\rightarrow} s_3$ and $t
+\StarOf{\rightarrow} t'$ and $R(s_3,t')$, but no such values exist.
 
@@
 \item $(S,\rightarrow)$ must be $=$-confluent.
-\item $(S,\rightarrow)$ must be \emph{$R$-bisimilar}. In other words,
-  if $R(s_1,t_1)$, and $s_1 \rightarrow s_2$, then there exists $t_2$ such that
-  $t_1 \rightarrow t_2$ and $R(s_2,t_2)$; and if $R(s_1,t_1)$, and $t_1
-  \rightarrow t_2$, then there exists $s_2$ such that $s_1 \rightarrow s_2$ and
-  $R(s_2,t_2)$.
+\item $R$ must be a bisimilulation relation for $(S,\StarOf{\rightarrow})$. In
+  other words if $R(s_1,t_1)$, and $s_1 \StarOf{\rightarrow} s_2$, then there
+  exists $t_2$ such that $t_1 \StarOf{\rightarrow} t_2$ and $R(s_2,t_2)$; and if
+  $R(s_1,t_1)$, and $t_1 \StarOf{\rightarrow} t_2$, then there exists $s_2$ such
+  that $s_1 \StarOf{\rightarrow} s_2$ and $R(s_2,t_2)$.
 \end{enumerate}
-
-Now we can proceed with our proof.
 
@@
   \label{lem:right-side}
-  Let $(S,\rightarrow)$ be $=$-confluent, and $(S,\rightarrow)$ be
-  $R$-bisimilar. If $R(s_1,t_1)$ and $s_1 \StarOf{\rightarrow} s_2$ and $t_1
-  \StarOf{\rightarrow} t_2$, then there exists $s_3,t_3$ such that $s_2
-  \StarOf{\rightarrow} s_3$ and $t_2 \rightarrow t_3$ and $R(s_3,t_3)$.
+  Let $(S,\rightarrow)$ be $=$-confluent, and $R$ be a bisimilulation relation
+  for $(S,\StarOf{\rightarrow})$. If $R(s_1,t_1)$ and $s_1 \StarOf{\rightarrow}
+  s_2$ and $t_1 \StarOf{\rightarrow} t_2$, then there exists $s_3,t_3$ such that
+  $s_2 \StarOf{\rightarrow} s_3$ and $t_2 \rightarrow t_3$ and $R(s_3,t_3)$.
 \end{theorem}
 \begin{proof}
-  By induction on $s_1 \StarOf{\rightarrow} s_2$
+  By induction length of the reduction of $s_1 \StarOf{\rightarrow} s_2$
   \begin{case}[Base Case]
     Let $R(s_1,t_1)$ and $t_1 \StarOf{\rightarrow} t_2$. As
-    $(S,\StarOf{\rightarrow})$ is $R$-bisimilar, there exists $s_2$ such that
+    $R$ is a bisimulation relation for $(S,\StarOf{\rightarrow})$, there exists $s_2$ such that
     $s_1 \StarOf{\rightarrow} s_2$ and $R(s_2,t_2)$, as desired.
@@
       inductive hypothesis. The existance of $s_5$ is guaranteed through
-      $=$-confluence. $R(s_5,t_5)$ comes from $R$-bisimilarity.}
+      $=$-confluence. $R(s_5,t_5)$ is guaranteed because $R$ is a bisimulation
+      relation.}
     \label{fig:solved-proof}
@@
     s_3$ and $t_1 \StarOf{\rightarrow} t_2$. By the induction hypothesis, there
-    exists $s_4,t_4$ such that $R(s_4,t_4)$ and $s_3 \StarOf{\rightarrow} s_4$
-    and $t_3 \rightarrow t_4$.
+    exists $s_4,t_4$ such that $R(s_4,t_4)$ and $s_2 \StarOf{\rightarrow} s_4$
+    and $t_2 \rightarrow t_4$.
 
-    As $(S,\rightarrow)$ is $=$-confluent, we know from prior work
-    $(S,\StarOf{\rightarrow})$ is $=$-confluent. Because of this, there exists
-    $s_5$ such that $s_3 \StarOf{\rightarrow} s_5$ and $s_4 \StarOf{\rightarrow}
-    s_5$. As $(S,\StarOf{\rightarrow})$ is $R$-bisimilar, there exists $t_5$
-    such that $t_4 \StarOf{\rightarrow} t_5$ and $R(s_5,t_5)$, as desired. This
-    case is diagrammed in Figure~\ref{fig:solved-proof}.
+    Because $(S,\rightarrow)$ is $=$-confluent, there
+    exists $s_5$ such that $s_3 \StarOf{\rightarrow} s_5$ and $s_4
+    \StarOf{\rightarrow} s_5$. As $(S,\StarOf{\rightarrow})$ is a bisimilulation
+    relation on $(S,\StarOf{\rightarrow})$,
+    there exists $t_5$ such that $t_4 \StarOf{\rightarrow} t_5$ and
+    $R(s_5,t_5)$, as desired. This case is diagrammed in
+    Figure~\ref{fig:solved-proof}.
   \end{case}
@@
 
+The concept of $R$-confluence is related to the notion of \emph{confluence
+  modulo $\sim$}~\cite{huetconf}. The definition of confluent modulo $\sim{}$ is
+almost the same as $\sim$-confluence, the only difference is that confluence
+modulo $\sim{}$ requires $\sim{}$ to an equivalence relation. Conditions that suffice to prove a rewrite system confluent modulo
+$\sim$ are not generally sufficient to prove $R$-confluence (and vice versa).
+Furthermore, our bisimulation relations are closely related to local coherence
+modulo $\sim{}$.
+
+Bisimulation relations come from concurrency theory~\cite{bisimcoind}, but a
+related notion, commuting rewrites~\cite{commut}, appear in the confluence
+literature. We require a single rewrite of $R$ to commute with an arbitrary
+number of rewrites of $\rightarrow$, which commuting rewrites do not express.
+
 The full proof of completeness is contained in the appendix of the full version
 of the original optician paper~\cite{extended-version}. The original proof of
-$R$-confluence for the transitive closure of $\rightarrow$ previously required
+$R$-confluence for the transitive closure of $\rightarrow$ required
 additional assumptions. These unnecessary assumptions have been identified and
-removed in this version of the proof. Future work used the proof of completeness
+removed in this paper. Future work used the proof of completeness
 over our lens language to show that quotient bijective lenses are also
@@
 was not proven for symmetric lenses~\cite{soptician}, such a proof would likely
-have proven $R$-confluence in a similar manner. Our proof used approaches
-inspired by the original Church-Rosser paper~\cite{Church1936SomePO}.
+have proven $R$-confluence in a similar manner.
 
@@
 synthesis~\cite{gulwani2017program}. While synthesis is one approach to make
-bidirectional programming easier, it is not the only approach. Recent work has
-gone into building lenses without requiring the point-free combinator
+bidirectional programming easier, it is not the only approach. Work has
+gone into building lenses without requiring a point-free combinator
 style~\cite{10.1007/978-3-319-89884-1_2}. Other work has found
